### 2024-05-14 index.qmd
20240513 - chapter numbering for PDF

# Preface {.unnumbered .unlisted}

This is a Quarto book.

To learn more about Quarto books visit <https://quarto.org/docs/books>.

```{r}
1 + 1
```

\mainmatter

### 2024-05-14 index.qmd
20240513 - preface

@@
 
-This is a Quarto book.
+This is a book in progress—it is incomplete.
+I will continue to add to and update it as I am able.
 
-To learn more about Quarto books visit <https://quarto.org/docs/books>.
+## Open Access {#sec-openAccess}
 
-```{r}
-1 + 1
+This is an open-access book.
+This means that it is freely available for anyone to access.
+
+## License {#sec-license}
+
+![Creative Commons License](Images/CC-BY.png){width=50%}
+
+The online version of this book is licensed under the [Creative Commons Attribution License](https://creativecommons.org/licenses/by/4.0/).
+In short, you can use my work as long as you cite it.
+
+## Citation {#sec-citation}
+
+The APA-style citation for the book is:
+
+Petersen, I. T. (2024). *Fantasy football analytics*. Version 0.0.1. University of Iowa Libraries. <https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook>. [INSERT DOI LINK]
+
+<!--
+[![DOI](Images/zenodo_doiBadge.png)](https://doi.org/10.5281/zenodo.6466589)
+-->
+
+The BibTeX citation for the book is:
+
+```
+@book{petersenFantasyFootballAnalytics,
+  title = {Fantasy football analytics},
+  author = {Petersen, Isaac T.},
+  year = {2024},
+  publisher = {{University of Iowa Libraries}},
+  note = {Version 0.0.1},
+  doi = {INSERT},
+  isbn = {INSERT},
+  url = {https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook}
+}
 ```
 
+## Accessibility {#sec-accessibility}
+
+I strive to follow [principles of accessibility](https://bookdown.org/yihui/rmarkdown-cookbook/html-accessibility.html) (archived at https://perma.cc/8XJ9-Q6QJ) to make the book content accessible to people with visual impairments and physical disabilities.
+If there are additional ways I can make the content more accessible, please let me know.
+
+## How to Contribute {#sec-contribute}
+
+This is an open-access textbook.
+My goal is to share data analysis strategies for free!
+Anyone is welcome to contribute to the project.
+If you would like to contribute, feel free to [open an issue](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues) or [create a pull request](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/pulls) on GitHub.\index{GitHub}
+The GitHub repository for the book is located here:\index{GitHub} <https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook>.
+If you have data or analysis examples that are you willing to share and include in the book, feel free to contact me.
+
+## Acknowledgments {#sec-acknowledgments}
+
 \mainmatter

### 2024-05-14 intro.qmd
20240513 - intro

# Introduction

How can we use information to make predictions about uncertain events?
This book is about empiricism (basing theories on observed data) and judgment, prediction, and decision making in the context of uncertainty.
The book provides an introduction to modern analytical techniques used to make informed predictions, test theories, and draw conclusions from a given dataset.

## What is Fantasy Football? (#sec-whatIsFantasyFootball)

Fantasy football is an online game where participants assemble (i.e., “draft”) imaginary teams composed of real-life National Football League (NFL) players.
In this game, participants compete against their opponents (friends/coworkers/classmates), accumulating points based on players' actual statistical performances in games.
The goal is to outscore one’s opponent each week to win matches and ultimately claim victory in the league.

## Why Focus on Fantasy Football? {#sec-whyFantasyFootball}

Many students find statistics intimidating in part because of how it is typically taught—with examples like dice rolls and coin flips that are not particularly exciting to students.
My contention is that applied examples are a more effective lens to teach many concepts in psychology and data analysis.
It can be more engaging and relatable to learn statistics in the applied context of sports, a domain that is more intuitive to many.
Many people play fantasy sports.
This book involves applying statistics to a particular domain (football).
People actually want to learn statistical principles and methods when they can apply them to interesting questions (e.g., sports).
In my opinion [and supported by evidence; @Motz2013], this is a much more effective way of engaging people and teaching statistics than in the context of abstract coin flips and dice rolls.
Fantasy football relies heavily on prediction—trying to predict which players will perform best and selecting them accordingly.
In this way, fantasy football provides a plethora of decision making opportunities in the face of uncertainty, and a wealth of data for analyzing these decisions.
However, unlike many other applied domains in psychology, fantasy football (1) allows a person to see the accuracy of their predictions on a timely basis and (2) provides a safe environment for friendly competition.
Thus, it provides a unique domain to evaluate—and improve—the accuracy of various prediction models.

## Educational Value {#sec-educationalValue}

Skills in data analysis are highly valuable.
This book includes practical and conceptual tools that build a foundation for critical thinking.
The book aims to help readers evaluate theory in the light of evidence (and vice versa) and to refine decision making in the context of uncertainty.
Readers will learn about the ways that psychological science (and related disciplines) poses questions, formulates hypotheses, designs studies to test those questions, and interprets the findings, collectively with the aim to answer questions, improve decision making, and solve problems.

Of course, this is not a traditional psychology textbook.
However, the class incorporates important psychological concepts, such as cognitive biases in judgment and prediction, etc.
In the modern world of big data, research and society need people who know how to make sense of the information around us.
Psychology is in a prime position to teach applied statistics to a wide variety of students, most of whom will not have careers as psychologists.
Psychology can teach the importance of statistics given humans’ cognitive biases.
It can also teach about how these biases can influence how people interpret statistics.
This course will teach readers the applications of statistics (prediction) and research methods (empiricism) to answer questions they find interesting, while applying scientific and psychological rigor.

## Learning Objectives {#sec-objectives}

This book aims to help readers accomplish the following learning objectives:

- Apply empirical inference and appreciate the value it provides over speculative supposition.
- Ask educated questions when confronted with decisions in the face of uncertainty.
- Understand human decision making, including common heuristics and cognitive biases and how to mitigate them analytically.
- Engage in critical thinking about causality, including devising plausible alternative explanations for observed effects.
- Understand causal inference including confounding, causal pathways, and counterfactuals.
- Think empirically about human behavior and performance.
- Describe the strengths and weaknesses of humans versus computers in prediction scenarios.
- Manipulate and summarize datasets.
- Critically evaluate the strengths and limitations of different statistical models and methodologies used in predicting uncertain events, enhancing their understanding of statistical inference and model selection.
- Use various analytical techniques for predicting the outcome of uncertain events, and for uncovering latent causes of patterns in observed data.
- Interpret findings from various statistical approaches and evaluate the accuracy of predictions.
- Engage in iterative problem-solving processes, refining analytical approaches based on feedback and outcomes, and adapting strategies accordingly.
- Communicate statistical findings and analyses in both written and oral formats, demonstrating proficiency in presenting complex information to diverse audiences.
- Make sense of big data.
- Use practical analytical skills that can be applied in future research and job settings.

## Disclosures {#sec-disclosures}

I am the Owner of Fantasy Football Analytics, LLC, which operates <https://fantasyfootballanalytics.net>.

## Disclaimer {#sec-disclaimer}

"*This material probably won't win you fantasy football championships.
You could take what we learn and apply it to fantasy football and you might become 5 percent more likely to win.
Or...
Consider the broader relevance of this.
You could learn data analysis and figure out ways to apply it to other systems.
And you could be making a six-figure salary within the next five years.*"
– Benjamin Motz, Ph.D.

### 2024-05-14 index.qmd
20240513 - typo
20240514 - cut

@@
 
-![Creative Commons License](Images/CC-BY.png){width=50%}
+![Creative Commons License](images/CC-BY.png){width=50%}
 
@@
 
-<!--
-[![DOI](Images/zenodo_doiBadge.png)](https://doi.org/10.5281/zenodo.6466589)
--->
-
 The BibTeX citation for the book is:
-
-```
-@book{petersenFantasyFootballAnalytics,
-  title = {Fantasy football analytics},
-  author = {Petersen, Isaac T.},
-  year = {2024},
-  publisher = {{University of Iowa Libraries}},
-  note = {Version 0.0.1},
-  doi = {INSERT},
-  isbn = {INSERT},
-  url = {https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook}
-}
-```
 
@@
 
-I strive to follow [principles of accessibility](https://bookdown.org/yihui/rmarkdown-cookbook/html-accessibility.html) (archived at https://perma.cc/8XJ9-Q6QJ) to make the book content accessible to people with visual impairments and physical disabilities.
+I strive to follow [principles of accessibility](https://bookdown.org/yihui/rmarkdown-cookbook/html-accessibility.html) (archived at <https://perma.cc/8XJ9-Q6QJ>) to make the book content accessible to people with visual impairments and physical disabilities.
 If there are additional ways I can make the content more accessible, please let me know.

### 2024-05-14 index.qmd
20240514 - update

@@
 I will continue to add to and update it as I am able.
+
+The following code chunk throws an error if I remove it:
+
+```{r}
+1 + 1
+```
 
@@
 
+<!--
+[![DOI](images/zenodo_doiBadge.png)](https://doi.org/10.5281/zenodo.6466589)
+-->
+
 The BibTeX citation for the book is:
+
+```
+@book{petersenFantasyFootballAnalytics,
+  title = {Fantasy football analytics},
+  author = {Petersen, Isaac T.},
+  year = {2024},
+  publisher = {{University of Iowa Libraries}},
+  note = {Version 0.0.1},
+  doi = {INSERT},
+  isbn = {INSERT},
+  url = {https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook}
+}
+```
 

### 2024-05-14 getting-started.qmd
20240514 - R initial setup

# Getting Started with R for Data Analysis

The book uses `R` for statistical analyses (<http://www.r-project.org>).
R is a free software environment; you can download it at no charge here: <https://cran.r-project.org>.

## Initial Setup {#sec-initialSetup}

To get started, follow the following steps:

1. Install `R`: <https://cran.r-project.org>
1. Install `RStudio Desktop`: <https://posit.co/download/rstudio-desktop>
1. After installing `RStudio`, open `RStudio` and run the following code in the console to install several key `R` packages:
```{r}
#| eval: false

install.packages(c("petersenlab","tidyverse","psych"))
```

If you are in Dr. Petersen's class, also perform the following steps:

1. Set up a free account on [GitHub.com](https://github.com).
1. Download GitHub Desktop: <https://desktop.github.com/>

## Installing Packages {#sec-installingPackages}

You can install R packages using the following syntax:

```{r}
#| eval: false

install.packages("INSERT_PACKAGE_NAME_HERE")
```

### 2024-05-16 getting-started.qmd
20240514 - cleanup
20240515 - callout
20240516 - section headings for cross-reference

@@
-# Getting Started with R for Data Analysis
+# Getting Started with R for Data Analysis {#sec-gettingStarted}
 
@@
 
+::: {#nte-gettingStarted .callout-note title="If you are in Dr. Petersen's class"}
 If you are in Dr. Petersen's class, also perform the following steps:
@@
 1. Set up a free account on [GitHub.com](https://github.com).
-1. Download GitHub Desktop: <https://desktop.github.com/>
+1. Download GitHub Desktop: <https://desktop.github.com>
+:::
 

### 2024-05-16 intro.qmd
20240514 - update
20240514 - todo
20240514 - cleanup
20240514 - embed video
20240516 - section headings for cross-reference

@@
-# Introduction
+# Introduction {#sec-intro}
 
@@
 
-## What is Fantasy Football? (#sec-whatIsFantasyFootball)
+## What is Fantasy Football? {#sec-whatIsFantasyFootball}
 
-Fantasy football is an online game where participants assemble (i.e., “draft”) imaginary teams composed of real-life National Football League (NFL) players.
-In this game, participants compete against their opponents (friends/coworkers/classmates), accumulating points based on players' actual statistical performances in games.
-The goal is to outscore one’s opponent each week to win matches and ultimately claim victory in the league.
+Fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.
+In this game, participants compete against their opponents (e.g., friends/coworkers/classmates), accumulating points based on players' actual statistical performances in games.
+The goal is to outscore one's opponent each week to win matches and ultimately claim victory in the league.
 
@@
 Psychology is in a prime position to teach applied statistics to a wide variety of students, most of whom will not have careers as psychologists.
-Psychology can teach the importance of statistics given humans’ cognitive biases.
+Psychology can teach the importance of statistics given humans' cognitive biases.
 It can also teach about how these biases can influence how people interpret statistics.
@@
 – Benjamin Motz, Ph.D.
+
+::: {.content-visible when-format="html"}
+
+![Benjamin Motz on statistics and fantasy football course](images/motzVideo.mp4){width=100%}
+
+:::

### 2024-05-16 fantasy-football.qmd
20240516 - fantasy football chapter

# Intro to Football and Fantasy {#sec-introFantasy}

This chapter provides a brief primer on football and fantasy football.
If you are already familiar with fantasy football, feel free to skip this chapter.

## Football {#sec-football}

## Fantasy Football {#sec-fantasyFootball}

As noted in the [Introduction](intro.qmd) Fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.
A fantasy football "league" is composed of various participants (each of whom has an imaginary team), and they play against each other.
Leagues are commonly composed of 8 or 10 teams, but leagues can have more or fewer participants/teams.
In the game of fantasy football, participants accumulate points on a weekly basis based on players' actual statistical performances in NFL games.
A participant's goal is to outscore their opponent each week to win matches and ultimately claim victory in the league.

Fantasy football relies heavily on prediction—trying to predict which players will perform best and selecting them accordingly.

### 2024-05-16 intro.qmd
20240516 - convert "class"/"course" to "book"

@@
 Of course, this is not a traditional psychology textbook.
-However, the class incorporates important psychological concepts, such as cognitive biases in judgment and prediction, etc.
+However, the book incorporates important psychological concepts, such as cognitive biases in judgment and prediction, etc.
 In the modern world of big data, research and society need people who know how to make sense of the information around us.
@@
 It can also teach about how these biases can influence how people interpret statistics.
-This course will teach readers the applications of statistics (prediction) and research methods (empiricism) to answer questions they find interesting, while applying scientific and psychological rigor.
+This book will teach readers the applications of statistics (prediction) and research methods (empiricism) to answer questions they find interesting, while applying scientific and psychological rigor.
 

### 2024-05-16 fantasy-football.qmd
20240516 - typo

@@
 
-As noted in the [Introduction](intro.qmd) Fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.
+As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.
 A fantasy football "league" is composed of various participants (each of whom has an imaginary team), and they play against each other.

### 2024-05-18 fantasy-football.qmd
20240518 - fantasy description

@@
 As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.
-A fantasy football "league" is composed of various participants (each of whom has an imaginary team), and they play against each other.
-Leagues are commonly composed of 8 or 10 teams, but leagues can have more or fewer participants/teams.
+The participants are in charge of managing and making strategic decisions for their imaginary team to have the best possible team that will score the most points.
+Thus, the participants are called "managers".
+Managers make decisions such as selecting which players to draft, selecting which players to play (i.e., "start") on a weekly basis, identifying players to pick up from the remaining pool of available players (i.e., waiver wire), and making trades with other teams.
+
+A fantasy football "league" is composed of various imaginary (i.e., "fantasy") teams—and their associated manager.
+In the fantasy league, the managers' fantasy teams play against each other.
+A fantasy league is commonly composed of 8 or 10 fantasy teams, but leagues can have more or fewer participants/teams.
 In the game of fantasy football, participants accumulate points on a weekly basis based on players' actual statistical performances in NFL games.

### 2024-05-18 actuarial.qmd
20240518 - placeholders

# Human Judgment Versus Actuarial Approaches to Prediction {#sec-judgmentVsActuarial}

### 2024-05-18 decision-making.qmd
20240518 - placeholders

# Decision Making in the Context of Uncertainty {#sec-decisionMaking}

### 2024-05-18 evaluating-prediction-accuracy.qmd
20240518 - placeholders

# Evaluation of Prediction/Forecasting Accuracy {#sec-predictionAccuracy}

### 2024-05-18 modern-portfolio-theory.qmd
20240518 - placeholders

# Modern Portfolio Theory {#sec-modernPortfolioTheory}

### 2024-05-18 mythbusters.qmd
20240518 - placeholders

# Mythbusters: Putting Fantasy Football Beliefs/Anecdotes to the Test {#sec-mythbusters}

### 2024-05-18 sports-cognitive-psychology.qmd
20240518 - placeholders

# Sports and Cognitive Psychology {#sec-sportsCognitivePsychology}

### 2024-05-18 fantasy-football.qmd
20240518 - fantasy intro and scoring

@@
 
-As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.
+### Overview of Fantasy Football {#sec-fantasyOverview}
+
+As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]
 The participants are in charge of managing and making strategic decisions for their imaginary team to have the best possible team that will score the most points.
@@
 Managers make decisions such as selecting which players to draft, selecting which players to play (i.e., "start") on a weekly basis, identifying players to pick up from the remaining pool of available players (i.e., waiver wire), and making trades with other teams.
+Fantasy football relies heavily on prediction—trying to predict which players will perform best and selecting them accordingly.
+
+### The Fantasy League {#sec-league}
 
@@
 In the fantasy league, the managers' fantasy teams play against each other.
-A fantasy league is commonly composed of 8 or 10 fantasy teams, but leagues can have more or fewer participants/teams.
-In the game of fantasy football, participants accumulate points on a weekly basis based on players' actual statistical performances in NFL games.
-A participant's goal is to outscore their opponent each week to win matches and ultimately claim victory in the league.
+A fantasy league is commonly composed of 8 or 10 fantasy teams, but leagues can have more or fewer teams.
 
-Fantasy football relies heavily on prediction—trying to predict which players will perform best and selecting them accordingly.
+### The Roster of a Fantasy Team {#sec-roster}
+
+On a given roster, a manager has a "starting lineup" and a "bench".
+Each week, the manager decides which players on their roster to put in the starting lineup, and which to keep on the bench.
+In many leagues, a starting lineup is composed of one or more spots for each of the following:
+
+<u>Offensive players:</u>
+
+- Quarterback (QB)
+- Running Back (RB)
+- Wide Receiver (WR)
+- Tight End (TE)
+- Flex Position—a flexible position that can involve a player from various positions: e.g., a RB or WR or TE
+
+<u>Kickers:</u>
+
+- Kicker (K)
+
+<u>Defense/Special Teams:</u>
+
+- Team Defense (DST/D/DEF) or Individual Defensive Player (IDP)
+
+### Scoring {#sec-scoring}
+
+In the game of fantasy football, managers accumulate points on a weekly basis based on players' actual statistical performances in NFL games.
+Managers receive points for only those players who are on their starting lineup (not players on their bench).
+A manager's goal is to outscore their opponent each week to win matches and ultimately claim victory in the league.
+Scoring settings can differ from league to league.
+
+Here are common scoring settings for fantasy leagues:
+
+<u>Offensive players:</u>
+
+| Statistical category                              | Points               |
+|---------------------------------------------------|----------------------|
+| Rushing or receiving TD                           | 6                    |
+| Returning a kick or punt for a TD                 | 6                    |
+| Returning or recovering a fumble for a TD         | 6                    |
+| Passing TD                                        | 4                    |
+| Passing INT                                       | −2                   |
+| Fumble lost                                       | −2                   |
+| Rushing, passing, or receiving 2 point conversion | 2                    |
+| Rushing or receiving yards                        | 1 point per 10 yards |
+| Passing yards                                     | 1 point per 25 yards |
+
+Note: "TD" = touchdown; "INT" = interception
+
+Other common (but not necessarily standard) statistical categories include:
+
+- receptions (called "points per reception" [PPR] leagues)
+- return yards
+- passing attempts
+- rushing attempts
+
+<u>Kickers:</u>
+
+| Statistical category                              | Points |
+|---------------------------------------------------|--------|
+| FG made: 50+ yards                                | 5      |
+| FG made: 40–49 yards                              | 4      |
+| FG made: 39 yards or less                         | 3      |
+| Rushing, passing, or receiving 2 point conversion | 2      |
+| Point after attempt made                          | 1      |
+| Point after attempt missed                        | −1     |
+| Missed FG: 0–39 yards                             | −2     |
+| Missed FG: 40–49 yards                            | −1     |
+
+Note: "FG" = field goal
+
+<u>Team Defense/Special Teams:</u>
+
+| Statistical category          | Points |
+|-------------------------------|--------|
+| Defensive or special teams TD | 3      |
+| Interception                  | 2      |
+| Fumble recovery               | 2      |
+| Blocked punt, PAT, or FG      | 2      |
+| Safety                        | 2      |
+| Sack                          | 1      |
+
+Note: "PAT" = point after touchdown; "FG" = field goal
+
+<u>Individual Defensive Players:</u>
+
+| Statistical category  | Points |
+|-----------------------|--------|
+| Tackle solo           | 1      |
+| Tackle assist         | 0.5    |
+| Tackle for loss       | 1      |
+| Sack                  | 2      |
+| Interception          | 4      |
+| Fumble force          | 2      |
+| Fumble recovery       | 2      |
+| TD                    | 6      |
+| Safety                | 2      |
+| Pass defended         | 1      |
+| Block kick            | 2      |
+| Extra point returned  | 2      |
+
+Note: "TD" = touchdown
+
+Other common (but not necessarily standard) statistical categories include:
+
+- turnover return yards
+
+Common scoring abbreviations:
+
+- "TD" = touchdown
+- "INT" = interception
+- "yds" = yards
+- "ATT" = attempts
+- "2 pt conversion" = two-point conversion
+- "FG"
+- "PAT" = point after touchdown (i.e., extra point/point after attempt)

### 2024-05-18 fantasy-football.qmd
20240518 - headings
20240518 - typo

@@
 
-As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]
+As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available for [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]
 The participants are in charge of managing and making strategic decisions for their imaginary team to have the best possible team that will score the most points.
@@
 
+#### Scoring Overview {#sec-scoringOverview}
+
 In the game of fantasy football, managers accumulate points on a weekly basis based on players' actual statistical performances in NFL games.
@@
 
-Here are common scoring settings for fantasy leagues:
+Below are common scoring settings for fantasy leagues.
 
-<u>Offensive players:</u>
+#### Offensive Players {#sec-scoringOffense}
 
@@
 
-<u>Kickers:</u>
+#### Kickers {#sec-scoringKickers}
 
@@
 
-<u>Team Defense/Special Teams:</u>
+#### Team Defense/Special Teams {#sec-scoringDST}
 
@@
 
-<u>Individual Defensive Players:</u>
+#### Individual Defensive Players {#sec-scoringIDP}
 
@@
 
-Common scoring abbreviations:
+#### Common Scoring Abbreviations {#sec-scoringAbbreviations}
 

### 2024-05-18 fantasy-football.qmd
20240518 - edits

@@
 In the fantasy league, the managers' fantasy teams play against each other.
-A fantasy league is commonly composed of 8 or 10 fantasy teams, but leagues can have more or fewer teams.
+A fantasy league is commonly composed of 8, 10, or 12 fantasy teams, but leagues can have more or fewer teams.
 
@@
 Each week, the manager decides which players on their roster to put in the starting lineup, and which to keep on the bench.
-In many leagues, a starting lineup is composed of one or more spots for each of the following:
+In many leagues, a starting lineup is composed of offensive players, a kicker, and defense/special teams:
 
@@
 
-- Quarterback (QB)
-- Running Back (RB)
-- Wide Receiver (WR)
-- Tight End (TE)
-- Flex Position—a flexible position that can involve a player from various positions: e.g., a RB or WR or TE
+| Position           | Typical Number of Players in Starting Lineup |
+|--------------------|----------------------------------------------|
+| Quarterback (QB)   | 1                                            |
+| Running Back (RB)  | 2                                            |
+| Wide Receiver (WR) | 2                                            |
+| Tight End (TE)     | 1                                            |
+| Flex Position      | 1                                            |
+
+A "flex position" is a flexible position that can involve a player from various positions: e.g., a RB or WR or TE.
 
@@
 
-- Kicker (K)
+- one Kicker (K)
 
@@
 
-- Team Defense (DST/D/DEF) or Individual Defensive Player (IDP)
+- one Team Defense (DST/D/DEF) or multiple Individual Defensive Players (IDP)
 

### 2024-05-18 fantasy-football.qmd
20240518 - video intro to fantasy football
20240518 - updates

@@
 Fantasy football relies heavily on prediction—trying to predict which players will perform best and selecting them accordingly.
+
+::: {.content-visible when-format="html"}
+
+![Video that provides introduction to fantasy football](images/fantasyFootball.mp4){width=100%}
+
+:::
 
@@
 
-- receptions (called "points per reception" [PPR] leagues)
+- receptions (called "point per reception" [PPR] leagues)
 - return yards

### 2024-05-18 fantasy-football.qmd
20240518 - update to save only to html and revealjs
20240518 - video introduction to American football
20240518 - video titles

@@
 
+::: {.content-visible when-format="html:js"}
+
+![Video that provides introduction to American football.](images/football.mp4){width=100%}
+
+:::
+
 ### Overview of Fantasy Football {#sec-fantasyOverview}
@@
 
-::: {.content-visible when-format="html"}
+::: {.content-visible when-format="html:js"}
 
-![Video that provides introduction to fantasy football](images/fantasyFootball.mp4){width=100%}
+![Video that provides introduction to fantasy football.](images/fantasyFootball.mp4){width=100%}
 

### 2024-05-18 intro.qmd
20240518 - edits
20240518 - update to save only to html and revealjs
20240518 - video titles
20240518 - add video introductions

@@
 
-::: {.content-visible when-format="html"}
+::: {.content-visible when-format="html:js"}
 
-![Benjamin Motz on statistics and fantasy football course](images/motzVideo.mp4){width=100%}
+Here is a video of a Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football:
+
+![Video of Prof. Benjamin Motz discussing statistics and fantasy football course.](images/motzVideo.mp4){width=100%}
 

### 2024-05-18 fantasy-football.qmd
20240518 - move content
20240518 - add video introductions
20240518 - table captions

@@
 
-## Fantasy Football {#sec-fantasyFootball}
+::: {.content-visible when-format="html:js"}
 
-::: {.content-visible when-format="html:js"}
+Here is a video that provides a brief introduction to American football:
 
@@
 :::
+
+## Fantasy Football {#sec-fantasyFootball}
 
@@
 ::: {.content-visible when-format="html:js"}
+
+Here is a video that provides a brief introduction to fantasy football:
 
@@
 | Flex Position      | 1                                            |
+
+: Offensive Players in the Starting Lineup {#tbl-lineupOffense tbl-colwidths="[25,75]"}
 
@@
 
+: Common Scoring Settings for Offensive Players {#tbl-scoringOffense tbl-colwidths="[75,25]"}
+
 Note: "TD" = touchdown; "INT" = interception
@@
 
+: Common Scoring Settings for Kickers {#tbl-scoringKickers tbl-colwidths="[75,25]"}
+
 Note: "FG" = field goal
@@
 | Sack                          | 1      |
+
+: Common Scoring Settings for Team Defense/Special Teams {#tbl-scoringDST tbl-colwidths="[75,25]"}
 
@@
 
+: Common Scoring Settings for Individual Defensive Players {#tbl-scoringIDP tbl-colwidths="[75,25]"}
+
 Note: "TD" = touchdown

### 2024-05-18 fantasy-football.qmd
20240518 - table column widths

--- 2024-05-18 fantasy-football.qmd
@@
 
+: Common Scoring Settings for Offensive Players {#tbl-scoringOffense tbl-colwidths="[50,50]"}
+
 Note: "TD" = touchdown; "INT" = interception
@@
 
+: Common Scoring Settings for Kickers {#tbl-scoringKickers tbl-colwidths="[50,50]"}
+
 Note: "FG" = field goal
@@
 | Sack                          | 1      |
+
+: Common Scoring Settings for Team Defense/Special Teams {#tbl-scoringDST tbl-colwidths="[50,50]"}
 
@@
 
+: Common Scoring Settings for Individual Defensive Players {#tbl-scoringIDP tbl-colwidths="[50,50]"}
+
 Note: "TD" = touchdown

### 2024-05-18 fantasy-football.qmd
20240518 - define abbreviations

@@
 
-Note: "PAT" = point after touchdown; "FG" = field goal
+Note: "TD" = touchdown; "PAT" = point after touchdown; "FG" = field goal
 
@@
 - "2 pt conversion" = two-point conversion
-- "FG"
+- "FG" = field goal
 - "PAT" = point after touchdown (i.e., extra point/point after attempt)

### 2024-05-19 fantasy-football.qmd
20240519 - description of football

@@
 
-This chapter provides a brief primer on football and fantasy football.
+This chapter provides a brief primer on (American) football and fantasy football.
 If you are already familiar with fantasy football, feel free to skip this chapter.
@@
 ## Football {#sec-football}
+
+The goal in football is for a team to score more points than their opponent.
+
+### The Roster {#sec-footballRoster}
+
+Each team has 11 players on the field at a time.
+The particular players who are on the field will depend on the situation, but usually includes one of the three subsets of players:
+
+1. Offense
+1. Defense
+1. Special Teams
+
+#### Offense {#sec-footballRosterOffense}
+
+The offense is on the field when the team has the ball.
+
+Players on offense include:
+
+- quarterback (QB)
+- running back (RB), also called halfback or tailback
+- fullback (FB)
+- wide receiver (WR)
+- tight end (TE)
+- offensive line (OL)
+    - center (C)
+    - offensive guard (OG)
+    - offensive tackle (OT)
+
+Quarterbacks are the most important player on the offense.
+They help lead the team down the field.
+Quarterbacks receive the ball from the Center at the beginning of the play, and they can either hand the ball off (typically to a running back or fullback), pass the ball (typically to a wide receiver or tight end), or run the ball.
+
+Running backs take a hand-off from the quarterback to execute a running play.
+They may also catch short passes from the quarterback or help protect (i.e., block for) the quarterback from the defensive players who are trying to tackle the quarterback.
+
+Wide receivers catch passes from the quarterback to execute a passing play.
+On running plays, they provide protection for the player running the ball (e.g., the running back) so they can get as far as possible without being tackled.
+
+Tight ends block for running and passing plays, and they catch passes from the quarterback.
+
+The offensive line blocks for running and passing plays.
+On passing plays, they provide protection for the quarterback so the quarterback has time to pass the ball without being tackled.
+On running plays, they provide protection for the player running the ball (e.g., the running back) so they can get as far as possible without being tackled.
+
+#### Defense {#sec-footballRosterDefense}
+
+The defense is on the field when the team does not have the ball (i.e., when the opposing team has the ball).
+
+Players on defense include:
+
+- defensive line (DL)
+    - defensive end (DE)
+    - defensive tackle (DT)
+- linebacker (LB)
+    - middle linebacker (MLB)
+    - outside linebacker (OLB)
+- defensive back (DB)
+    - cornerback (CB)
+    - safety (S)
+        - free safety (FS)
+        - strong safety (SS)
+
+#### Special Teams {#sec-footballRosterSpecialTeams}
+
+The special teams involves specialist players who are on the field during all kicking plays including kickoffs, field goals, and punts.
+
+Players on special teams include:
+
+- kicker (K)
+- punter (P)
+- holder
+- long snapper
+- punt returner
+- kick returner
+- and other players intended to block for or to tackle the ball carrier
+
+### The Field {#sec-footballField}
+
+The football field is rectangular and is 120 yards long and 53 1/3 yards wide.
+At each end of the 120-yard field is a team's end zone.
+Each end zone is 10 yards long.
+Thus, the distance from one end zone to the other end zone is 100 yards.
+Behind each end zone is a field goal post.
+
+### The Gameplay {#sec-footballGameplay}
+
+At the beginning of the game, there is a coin flip to determine which teams receives the ball first and which team takes which side of the field.
+During the kickoff, the kicking team kicks the ball to the receiving team, who has the option to return the kick.
+The offense starts their possession at the 25 yard line—if there is no return (i.e., a touchback)—or wherever the kick returner is tackled or goes out of bounds.
+
+The team with the ball (i.e., the offense) has four opportunities ("downs") to advance the ball 10 yards.
+A team can advance the ball either by running it or passing and catching it.
+At the end of a rushing play, the ball advances to wherever the ball carrier is tackled or goes out of bounds (i.e., wherever the player is "down").
+At the end of a passing play, if the thrown ball is caught (i.e., a completed pass), the ball advances to wherever the ball carrier is tackled or goes out of bounds.
+If the thrown ball is not caught (i.e., an incomplete pass), the ball does not advance.
+Wherever the ball is advanced to dictates where the next play begins.
+The yard position on the field where the next play takes place from is known as the line of scrimmage.
+Neither team can cross the line the line of scrimmage until the next play begins.
+To begin the play, the ball is placed on the line of scrimmage and the Center gives (or "snaps") the ball to the Quarterback.
+
+If the team advances the ball 10 or more yards within four downs, the team is awarded a new set of downs and is given four more downs to advance the ball 10 more yards.
+If the team advances the ball all the way to the other team's end zone, they score a touchdown.
+If the team fails to advance the ball 10 or more yards within four downs, the team loses the ball, and the other team takes possession at that spot on the field.
+There are risks of giving the other team the ball with a short distance to score.
+Thus, on fourth down, instead of trying to advance the ball for a first down, a team may choose to kick a field goal—to get points—or to punt.
+
+A field goal involves a kicker kicking the ball with an intent to kick the ball through the field goal posts ("uprights").
+To achieve points, the field goal must be above the cross bar and be in between the uprights (extended vertically) when it crosses the cross bar.
+
+Punting involves a punter kicking the ball to the other team with an intent to give their opponent worse field position, thus making it harder for the other team to score.
+The punting team tries to pin the opponent as close as possible to the opponent's end zone (i.e., as far as possible from the own team's end zone), so they have a longer distance to go to score a touchdown.
+
+There are multiple ways that ball possession can switch from the offense to the other team.
+After scoring a touchdown, field goal, or safety, there is a kickoff, in which the scoring team kicks the ball to the opponent.
+Another way that the ball swithces possession to the other team is if the team commits a turnover.
+The defense can force a turnover by an interception, fumble recovery, or turnover on downs.
+A turnover due to an interception occurs when the defense catches the the quarterback's pass.
+A turnover due to a fumble recovery occurs when an offensive player, who had possession of the ball, loses the ball before being down or scoring a touchdown and the ball is recovered by the opponent.
+A turnover on downs occurs when the team attempts to achieve the remainder of the needed 10 yards to go on fourth down but fails.
+
+Other football-related situations include tackles for loss and sacks.
+A tackle for loss occurs when a (non-quarterback) is tackled behind the line of scrimmage.
+A sack occurs when a quarterback is tackled behind the line of scrimmage.
+A pass defended occurs when a defensive back knocks down the ball in the air so that the indended receiver cannot catch the ball.
+
+### The Scoring {#sec-footballScoring}
+
+The goal of the team with the ball (i.e., the offense) is to score points.
+It can do this by either advancing the ball into the other team's end zone (6 points) or by kicking a field goal (3 points).
+Advancing the ball in the other team's end zone is called a touchdown.
+After a touchdown, the offense chooses to attempt either a point-after-touchdown (PAT) or a two-point conversion.
+A PAT is a short kick attempt from the 15-yard line that, if it goes through the goal posts ("uprights"), is worth 1 point.
+A two-point conversion is a single-scoring opportunity from the 3-yard line (i.e., 3 yards away from the end zone).
+If the offense scores (i.e., advances the ball into the end zone) from the 3-yard line, the team is awarded 2 points.
+
+A team can kick a field goal from any distance as long as the kick goes through the goal posts.
+The current record for the longest field goal is 66 yards (by Justin Tucker in 2021).
+
+A safety occurs when the offense is tackled with the ball in their own end zone.
+When a safety occurs, the opposing team (i.e., defense) is awarded two points.
+
+### Glossary of Terms {#sec-footballGlossary}
+
+- running play ("run") or rushing play (or "rush")
+- passing play (or "pass")
+- passing attempt
+- rushing attempt
+- passing completion
+- passing incompletion
+- passing yards
+- rushing yards
+- receving yards
+- reception
+- touchdown
+- passing touchdown
+- rushing touchdown
+- receiving touchdown
+- two-point conversion
+- block
+- kickoff
+- field goal
+- point after touchdown (PAT)
+- extra point returned
+- punt
+- fumble lost
+- fumble forced
+- fumble recovery
+- interception
+- tackle
+- tackle solo
+- tackle assist
+- tackle for loss
+- sack
+- pass defended
+- safety
 
@@
 
-A "flex position" is a flexible position that can involve a player from various positions: e.g., a RB or WR or TE.
+A "flex position" is a flexible position that can involve a player from various positions: e.g., a running back, wide receiver, or tight end.
+In fantasy football, fullbacks are classified as running backs because they both run the ball.
 
@@
 | Fumble lost                                       | −2                   |
-| Rushing, passing, or receiving 2 point conversion | 2                    |
+| Rushing, passing, or receiving 2-point conversion | 2                    |
 | Rushing or receiving yards                        | 1 point per 10 yards |
@@
 | FG made: 39 yards or less                         | 3      |
-| Rushing, passing, or receiving 2 point conversion | 2      |
-| Point after attempt made                          | 1      |
-| Point after attempt missed                        | −1     |
+| Rushing, passing, or receiving 2-point conversion | 2      |
+| Point after touchdown attempt made                | 1      |
+| Point after touchdown attempt missed              | −1     |
 | Missed FG: 0–39 yards                             | −2     |
@@
 | Interception          | 4      |
-| Fumble force          | 2      |
+| Fumble forced         | 2      |
 | Fumble recovery       | 2      |
@@
 | Pass defended         | 1      |
-| Block kick            | 2      |
+| Blocked kick          | 2      |
 | Extra point returned  | 2      |
@@
 - "ATT" = attempts
-- "2 pt conversion" = two-point conversion
+- "2-pt conversion" = two-point conversion
 - "FG" = field goal

### 2024-05-19 fantasy-football.qmd
20240519 - capitalization

@@
 The goal in football is for a team to score more points than their opponent.
+
+::: {.content-visible when-format="html:js"}
+
+Here is a video that provides a brief introduction to American football:
+
+![Video that provides introduction to American football.](images/football.mp4){width=100%}
+
+:::
 
@@
 
-- quarterback (QB)
-- running back (RB), also called halfback or tailback
-- fullback (FB)
-- wide receiver (WR)
-- tight end (TE)
-- offensive line (OL)
-    - center (C)
-    - offensive guard (OG)
-    - offensive tackle (OT)
+- Quarterback (QB)
+- Running Back (RB), also called halfback or tailback
+- Fullback (FB)
+- Wide Receiver (WR)
+- Tight End (TE)
+- Offensive Line (OL)
+    - Center (C)
+    - Offensive Guard (OG)
+    - Offensive Tackle (OT)
 
@@
 They help lead the team down the field.
-Quarterbacks receive the ball from the Center at the beginning of the play, and they can either hand the ball off (typically to a running back or fullback), pass the ball (typically to a wide receiver or tight end), or run the ball.
+Quarterbacks receive the ball from the Center at the beginning of the play, and they can either hand the ball off (typically to a Running Back or Fullback), pass the ball (typically to a Wide Receiver or Tight End), or run the ball.
 
-Running backs take a hand-off from the quarterback to execute a running play.
-They may also catch short passes from the quarterback or help protect (i.e., block for) the quarterback from the defensive players who are trying to tackle the quarterback.
+Running Backs take a hand-off from the Quarterback to execute a running play.
+They may also catch short passes from the Quarterback or help protect (i.e., block for) the Quarterback from the defensive players who are trying to tackle the Quarterback.
 
-Wide receivers catch passes from the quarterback to execute a passing play.
+Wide Receivers catch passes from the Quarterback to execute a passing play.
 On running plays, they provide protection for the player running the ball (e.g., the running back) so they can get as far as possible without being tackled.
 
-Tight ends block for running and passing plays, and they catch passes from the quarterback.
+Tight Ends block for running and passing plays, and they catch passes from the Quarterback.
 
-The offensive line blocks for running and passing plays.
-On passing plays, they provide protection for the quarterback so the quarterback has time to pass the ball without being tackled.
+The players on the Offensive Line block for running and passing plays.
+On passing plays, they provide protection for the Quarterback so the Quarterback has time to pass the ball without being tackled.
 On running plays, they provide protection for the player running the ball (e.g., the running back) so they can get as far as possible without being tackled.
@@
 
-- defensive line (DL)
-    - defensive end (DE)
-    - defensive tackle (DT)
-- linebacker (LB)
-    - middle linebacker (MLB)
-    - outside linebacker (OLB)
-- defensive back (DB)
-    - cornerback (CB)
-    - safety (S)
-        - free safety (FS)
-        - strong safety (SS)
+- Defensive Line (DL)
+    - Defensive End (DE)
+    - Defensive Tackle (DT)
+- Linebacker (LB)
+    - Middle Linebacker (MLB)
+    - Outside Linebacker (OLB)
+- Defensive Back (DB)
+    - Cornerback (CB)
+    - Safety (S)
+        - Free Safety (FS)
+        - Strong Safety (SS)
 
@@
 
-- kicker (K)
-- punter (P)
-- holder
-- long snapper
-- punt returner
-- kick returner
+- Kicker (K)
+- Punter (P)
+- Holder
+- Long Snapper
+- Punt Returner
+- Kick Returner
 - and other players intended to block for or to tackle the ball carrier
@@
 Other football-related situations include tackles for loss and sacks.
-A tackle for loss occurs when a (non-quarterback) is tackled behind the line of scrimmage.
-A sack occurs when a quarterback is tackled behind the line of scrimmage.
-A pass defended occurs when a defensive back knocks down the ball in the air so that the indended receiver cannot catch the ball.
+A tackle for loss occurs when a ball carrier is tackled behind the line of scrimmage.
+A sack occurs when a Quarterback is tackled with the ball behind the line of scrimmage.
+A pass defended occurs when a defensive player knocks down the ball in the air so that the indended receiver cannot catch the ball.
 
@@
 
-::: {.content-visible when-format="html:js"}
-
-Here is a video that provides a brief introduction to American football:
-
-![Video that provides introduction to American football.](images/football.mp4){width=100%}
-
-:::
-
 ## Fantasy Football {#sec-fantasyFootball}
@@
 
-A "flex position" is a flexible position that can involve a player from various positions: e.g., a running back, wide receiver, or tight end.
-In fantasy football, fullbacks are classified as running backs because they both run the ball.
+A "flex position" is a flexible position that can involve a player from various positions: e.g., a Running Back, Wide Receiver, or Tight End.
+In fantasy football, Fullbacks are classified as Running Backs because they both run the ball.
 

### 2024-05-19 fantasy-football.qmd
20240519 - more additions and edits

@@
 - Quarterback (QB)
-- Running Back (RB), also called halfback or tailback
+- Running Back (RB)
+    - Halfback or Tailback
     - Fullback (FB)
@@
 Quarterbacks receive the ball from the Center at the beginning of the play, and they can either hand the ball off (typically to a Running Back or Fullback), pass the ball (typically to a Wide Receiver or Tight End), or run the ball.
+Quarterbacks tend to have a strong arm for throwing the ball far and accurately.
 
@@
 They may also catch short passes from the Quarterback or help protect (i.e., block for) the Quarterback from the defensive players who are trying to tackle the Quarterback.
+Halfbacks or Tailbacks tend to be quick and agile.
+Fullbacks tend to be strong and powerful.
 
 Wide Receivers catch passes from the Quarterback to execute a passing play.
-On running plays, they provide protection for the player running the ball (e.g., the running back) so they can get as far as possible without being tackled.
+On running plays, they provide protection for the player running the ball (e.g., the Running Back) so the ball carrier can get as far as possible without being tackled.
+Wide receivers tend to be tall, fast, have good hands (can catch the ball well), and can jump high.
 
 Tight Ends block for running and passing plays, and they catch passes from the Quarterback.
+Tight ends tend to be strong and have good hands.
 
@@
 On passing plays, they provide protection for the Quarterback so the Quarterback has time to pass the ball without being tackled.
-On running plays, they provide protection for the player running the ball (e.g., the running back) so they can get as far as possible without being tackled.
+On running plays, they provide protection for the player running the ball (e.g., the Running Back) so the ball carrier can get as far as possible without being tackled.
+Players on the Offensive Line tend to be large so they can provide adequate protection for the Quarterback and Running Back.
 
@@
 - Linebacker (LB)
-    - Middle Linebacker (MLB)
+    - Middle (or Inside) Linebacker (MLB)
     - Outside Linebacker (OLB)
-- Defensive Back (DB)
+- "The Secondary" / Defensive Back (DB)
     - Cornerback (CB)
@@
         - Strong Safety (SS)
+
+The players on the defense attempt to tackle the offensive players for as short of gains as possible and attempt to prevent completed passes.
+
+On passing plays, the players on the Defensive Line try to apply pressure to the Quarterback and try to sack them.
+On rushing plays, the players on the Defensive Line try to tackle the ball carrier.
+Players on the Defensive Line tend to be large yet quick so they can apply pressure to the Quarterback.
+
+Linebackers are versatile in that, on a given play, they may attempt to a) "blitz" to sack the Quarterback, b) stop the Running Back, or c) prevent a completed pass.
+Linebackers tend to be strong yet agile.
+
+Defensive Backs are specialist pass defenders.
+The main role of Cornerbacks is to cover the Wide Receivers.
+Safeties serve as the last line of defense for longer passes.
+Defensive Backs tend to be quick and agile.
 
@@
 - and other players intended to block for or to tackle the ball carrier
+
+On a field goal attempt, the Long Snapper snaps the ball to the Holder, who holds the ball for the Kicker.
+The Kicker attempts field goals and, during kickoffs, kicks the ball to the opposing team.
+During kickoffs, the Kick Returner catches the kicked ball and returns it for as many yards as possible.
+During a punt play, the Long Snapper snaps the ball to the Punter who kicks (i.e., punts) the ball to the opposing team.
+The Punt Returner catches the punted ball and returns it for as many yards as possible.
 
@@
 A "flex position" is a flexible position that can involve a player from various positions: e.g., a Running Back, Wide Receiver, or Tight End.
-In fantasy football, Fullbacks are classified as Running Backs because they both run the ball.
 

### 2024-05-19 fantasy-football.qmd
20240519 - dual threats
20240519 - example formation
20240519 - football field diagram

@@
 
-![Video that provides introduction to American football.](images/football.mp4){width=100%}
+![Video that Provides An Introduction to American Football.](images/football.mp4){width=100%}
 
@@
 ### The Roster {#sec-footballRoster}
+
+#### Overview {#sec-footballRosterOverview}
 
@@
 1. Special Teams
+
+An example formation is depicted in Figure @fig-footballFormation.
+
+![An Example Football Formation for the Offense and Defense.](images/footballFormation.png){#fig-footballFormation}
 
@@
 Quarterbacks tend to have a strong arm for throwing the ball far and accurately.
+Some quarterbacks are fast and are considered "dual threats" to pass or run.
 
@@
 Behind each end zone is a field goal post.
+A diagram of a football field is depicted in Figure @fig-footballField.
+
+::: {#fig-footballField}
+![](images/footballField.png)
+
+A Diagram of a Football Field.
+The yard markers depict the distance from the nearest end zone.
+The orange shaded area is called the "red zone", where chances of scoring points are highest.
+The original figure was modified by adding field goal posts.
+(Figure retrieved from https://commons.wikimedia.org/wiki/File:American_football_field.svg)
+:::
 
@@
 Wherever the ball is advanced to dictates where the next play begins.
-The yard position on the field where the next play takes place from is known as the line of scrimmage.
+The yard position on the field where the next play takes place from is known as the "line of scrimmage".
 Neither team can cross the line the line of scrimmage until the next play begins.
@@
 
-If the team advances the ball 10 or more yards within four downs, the team is awarded a new set of downs and is given four more downs to advance the ball 10 more yards.
+If the team advances the ball 10 or more yards within four downs, the team receives a "first down" and is awarded a new set of downs—four more downs to advance the ball 10 more yards.
 If the team advances the ball all the way to the other team's end zone, they score a touchdown.
@@
 
-![Video that provides introduction to fantasy football.](images/fantasyFootball.mp4){width=100%}
+![Video that Provides an Introduction to Fantasy Football.](images/fantasyFootball.mp4){width=100%}
 

### 2024-05-19 fantasy-football.qmd
20240519 - updated football field
20240519 - link in figure caption

@@
 ::: {#fig-footballField}
-![](images/footballField.png)
+![](images/footballField.png){width=50%}
 
@@
 The orange shaded area is called the "red zone", where chances of scoring points are highest.
-The original figure was modified by adding field goal posts.
-(Figure retrieved from https://commons.wikimedia.org/wiki/File:American_football_field.svg)
+The original figure was modified to depict field goal posts.
+(Figure retrieved from <https://commons.wikimedia.org/wiki/File:American_football_field.svg>)
 :::

### 2024-05-19 fantasy-football.qmd
20240519 - updated formation
20240519 - concision

@@
 
-![An Example Football Formation for the Offense and Defense.](images/footballFormation.png){#fig-footballFormation}
+::: {#fig-footballFormation}
+![](images/footballFormation.png)
+
+An Example Football Formation for the Offense and Defense. The solid line indicates the line of scrimmage. The arrow indicates the direction the offense tries to advance the ball.
+:::
 
@@
 - Tight End (TE)
-- Offensive Line (OL)
+- Offensive Linemen (OL)
     - Center (C)
@@
 
-The players on the Offensive Line block for running and passing plays.
+Offensive Linemen block for running and passing plays.
 On passing plays, they provide protection for the Quarterback so the Quarterback has time to pass the ball without being tackled.
 On running plays, they provide protection for the player running the ball (e.g., the Running Back) so the ball carrier can get as far as possible without being tackled.
-Players on the Offensive Line tend to be large so they can provide adequate protection for the Quarterback and Running Back.
+Offensive Linemen tend to be large so they can provide adequate protection for the Quarterback and Running Back.
 
@@
 
-- Defensive Line (DL)
+- Defensive Linemen (DL)
     - Defensive End (DE)
@@
 
-On passing plays, the players on the Defensive Line try to apply pressure to the Quarterback and try to sack them.
-On rushing plays, the players on the Defensive Line try to tackle the ball carrier.
-Players on the Defensive Line tend to be large yet quick so they can apply pressure to the Quarterback.
+On passing plays, Defensive Linemen try to apply pressure to the Quarterback and try to sack them.
+On rushing plays, Defensive Linemen try to tackle the ball carrier.
+Defensive Linemen tend to be large yet quick so they can apply pressure to the Quarterback.
 

### 2024-05-19 fantasy-football.qmd
20240519 - edits

@@
 
+### The Objective {#sec-football}
+
 The goal in football is for a team to score more points than their opponent.
+A game lasts 60 minutes, and it is separated into four 15-minute quarters.
+The team with the most points when the time runs out wins.
 
@@
 At the end of a passing play, if the thrown ball is caught (i.e., a completed pass), the ball advances to wherever the ball carrier is tackled or goes out of bounds.
-If the thrown ball is not caught (i.e., an incomplete pass), the ball does not advance.
+If the thrown ball is not caught in bounds before the ball hits the ground (i.e., an incomplete pass), the ball does not advance.
 Wherever the ball is advanced to dictates where the next play begins.
@@
 A field goal involves a kicker kicking the ball with an intent to kick the ball through the field goal posts ("uprights").
-To achieve points, the field goal must be above the cross bar and be in between the uprights (extended vertically) when it crosses the cross bar.
+To score points by making a field goal, the kicked ball must go between the uprights (extended vertically) and over the cross bar.
 

### 2024-05-19 index.qmd
20240514 - error, message, warning
20240514 - remove code chunk that was throwing an error
20240516 - section headings for cross-reference
20240519 - acknowledgments
20240519 - FFA

@@
-# Preface {.unnumbered .unlisted}
+# Preface {#sec-preface .unnumbered .unlisted}
 
@@
 I will continue to add to and update it as I am able.
-
-The following code chunk throws an error if I remove it:
-
-```{r}
-1 + 1
-```
 
@@
 
+I thank Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.
+I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, and Adam Peterson, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
+
 \mainmatter

### 2024-05-20 intro.qmd
20240519 - football field diagram
20240520 - about the book

@@
 # Introduction {#sec-intro}
+
+## About this Book {#sec-aboutBook}
 
@@
 The book provides an introduction to modern analytical techniques used to make informed predictions, test theories, and draw conclusions from a given dataset.
+
+This book was originally written for a undergraduate-level course entitled, "Fantasy Football: Predictive Analytics and Empiricism".
+The chapters provide an overview of topics that each could have its own class and textbook, such as [causal inference](causal-inference.qmd), [factor analysis](factor-analysis.qmd), [cluster analysis](cluster-analysis.qmd), [principal component analysis](pca.qmd), [machine learning](machine-learning.qmd), [cognitive biases](cognitive-bias.qmd), [modern portfolio theory](modern-portfolio-theory.qmd), [data visualization](data-visualization.qmd), [simulation](simulation.qmd), etc.
+The book gives readers an overview of the breadth of the approaches to prediction and empiricism.
+As a consequence, the book does not cover any one technique or approach in great depth.
 
@@
 
-![Video of Prof. Benjamin Motz discussing statistics and fantasy football course.](images/motzVideo.mp4){width=100%}
+![Video of Prof. Benjamin Motz Discussing Statistics and His Fantasy Football Course.](images/motzVideo.mp4){width=100%}
 

### 2024-05-20 cognitive-bias.qmd
Built site for gh-pages

# Heuristics and Cognitive Biases in Prediction {#sec-cognitiveBias}

### 2024-05-20 pca.qmd
Built site for gh-pages

# Data Reduction: Principal Component Analysis {#sec-pca}

### 2024-05-20 simulation.qmd
Built site for gh-pages

# Simulation: the Monte Carlo Method {#sec-simulation}

### 2024-05-20 index.qmd
20240520 - Patrick Carroll

@@
 
-I thank Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.
+I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.
 I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, and Adam Peterson, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
+I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.
 

### 2024-05-20 intro.qmd
20240520 - Patrick Carroll

@@
 
-Many students find statistics intimidating in part because of how it is typically taught—with examples like dice rolls and coin flips that are not particularly exciting to students.
+I was fortunate to have an excellent instructor who taught me the value of learning statistics to answer interesting and important questions.
+That is, I do not find statistics intrinsically interesting; rather, I find them interesting because of what they allow me to do.
+Many students find statistics intimidating in part because of how it is typically taught—with examples like dice rolls and coin flips that are (seemingly irrelevant and) boring to students.
 My contention is that applied examples are a more effective lens to teach many concepts in psychology and data analysis.

### 2024-05-22 getting-started.qmd
20240521 - add data from nflfastR

@@
 1. After installing `RStudio`, open `RStudio` and run the following code in the console to install several key `R` packages:
+
 ```{r}
@@
 ```
+
+For instance, you can use the following code to install the `nflfastR` package:
+
+```{r}
+#| eval: false
+
+install.packages("nflfastR")
+```
+
+## Load Packages {#sec-loadPackages}
+
+```{r}
+library("nflfastR")
+library("tidyverse")
+```
+
+## Download Football Data {#sec-downloadFootballData}
+
+### Play-By-Play Data {#sec-downloadPlayByPlay}
+
+To download play-by-play data from prior weeks and seasons, we can use the `load_pbp()` function of the `nflfastR` package:
+
+```{r}
+#| eval: false
+
+nfl_pbp <- nflfastR::load_pbp(seasons = TRUE)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+library("here")
+
+save(
+  nfl_pbp,
+  file = here("data", "nfl_pbp.RData")
+)
+```
+
+```{r}
+#| include: false
+
+library("here")
+
+load(file = here("data", "nfl_pbp.RData"))
+```
+
+### Player Info {#sec-downloadPlayerInfo}
+
+## Calculations {#sec-calculations}
+
+### Historical Actual Player Statistics {#sec-calculateActualStats}
+
+TODO: compute player stats for defenses (`calculate_player_stats_def()`) and kickers (`calculate_player_stats_kicking()`).
+
+Note: The following code takes a while to run.
+
+First, subset the data by year:
+
+```{r}
+#| eval: false
+
+nfl_pbp_1999 <- nfl_pbp %>% filter(season == 1999)
+nfl_pbp_2000 <- nfl_pbp %>% filter(season == 2000)
+nfl_pbp_2001 <- nfl_pbp %>% filter(season == 2001)
+nfl_pbp_2002 <- nfl_pbp %>% filter(season == 2002)
+nfl_pbp_2003 <- nfl_pbp %>% filter(season == 2003)
+nfl_pbp_2004 <- nfl_pbp %>% filter(season == 2004)
+nfl_pbp_2005 <- nfl_pbp %>% filter(season == 2005)
+nfl_pbp_2006 <- nfl_pbp %>% filter(season == 2006)
+nfl_pbp_2007 <- nfl_pbp %>% filter(season == 2007)
+nfl_pbp_2008 <- nfl_pbp %>% filter(season == 2008)
+nfl_pbp_2009 <- nfl_pbp %>% filter(season == 2009)
+nfl_pbp_2010 <- nfl_pbp %>% filter(season == 2010)
+nfl_pbp_2011 <- nfl_pbp %>% filter(season == 2011)
+nfl_pbp_2012 <- nfl_pbp %>% filter(season == 2012)
+nfl_pbp_2013 <- nfl_pbp %>% filter(season == 2013)
+nfl_pbp_2014 <- nfl_pbp %>% filter(season == 2014)
+nfl_pbp_2015 <- nfl_pbp %>% filter(season == 2015)
+nfl_pbp_2016 <- nfl_pbp %>% filter(season == 2016)
+nfl_pbp_2017 <- nfl_pbp %>% filter(season == 2017)
+nfl_pbp_2018 <- nfl_pbp %>% filter(season == 2018)
+nfl_pbp_2019 <- nfl_pbp %>% filter(season == 2019)
+nfl_pbp_2020 <- nfl_pbp %>% filter(season == 2020)
+nfl_pbp_2021 <- nfl_pbp %>% filter(season == 2021)
+nfl_pbp_2022 <- nfl_pbp %>% filter(season == 2022)
+nfl_pbp_2023 <- nfl_pbp %>% filter(season == 2023)
+```
+
+Second, compute the weekly stats:
+
+```{r}
+#| eval: false
+
+nfl_actualStats_weekly <- nflfastR::calculate_player_stats(
+  nfl_pbp,
+  weekly = TRUE)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_actualStats_weekly,
+  file = here("data", "nfl_actualStats_weekly.RData")
+)
+```
+
+Third, compute the career stats:
+
+```{r}
+#| eval: false
+
+nfl_actualStats_career <- nflfastR::calculate_player_stats(
+  nfl_pbp,
+  weekly = FALSE)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_actualStats_career,
+  file = here("data", "nfl_actualStats_career.RData")
+)
+```
+
+Fourth, compute seasonal stats by year:
+
+```{r}
+#| eval: false
+
+nfl_actualStats_seasonal_1999 <- nflfastR::calculate_player_stats(
+  nfl_pbp_1999,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2000 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2000,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2001 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2001,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2002 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2002,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2003 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2003,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2004 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2004,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2005 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2005,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2006 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2006,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2007 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2007,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2008 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2008,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2009 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2009,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2010 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2010,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2011 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2011,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2012 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2012,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2013 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2013,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2014 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2014,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2015 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2015,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2016 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2016,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2017 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2017,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2018 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2018,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2019 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2019,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2020 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2020,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2021 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2021,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2022 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2022,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal_2023 <- nflfastR::calculate_player_stats(
+  nfl_pbp_2023,
+  weekly = FALSE)
+
+nfl_actualStats_seasonal <- bind_rows(
+  nfl_actualStats_seasonal_1999,
+  nfl_actualStats_seasonal_2000,
+  nfl_actualStats_seasonal_2001,
+  nfl_actualStats_seasonal_2002,
+  nfl_actualStats_seasonal_2003,
+  nfl_actualStats_seasonal_2004,
+  nfl_actualStats_seasonal_2005,
+  nfl_actualStats_seasonal_2006,
+  nfl_actualStats_seasonal_2007,
+  nfl_actualStats_seasonal_2008,
+  nfl_actualStats_seasonal_2009,
+  nfl_actualStats_seasonal_2010,
+  nfl_actualStats_seasonal_2011,
+  nfl_actualStats_seasonal_2012,
+  nfl_actualStats_seasonal_2013,
+  nfl_actualStats_seasonal_2014,
+  nfl_actualStats_seasonal_2015,
+  nfl_actualStats_seasonal_2016,
+  nfl_actualStats_seasonal_2017,
+  nfl_actualStats_seasonal_2018,
+  nfl_actualStats_seasonal_2019,
+  nfl_actualStats_seasonal_2020,
+  nfl_actualStats_seasonal_2021,
+  nfl_actualStats_seasonal_2022,
+  nfl_actualStats_seasonal_2023,
+  .id = "id"
+)
+
+nfl_actualStats_seasonal$season <- NA
+
+nfl_actualStats_seasonal$id <- as.integer(nfl_actualStats_seasonal$id)
+
+for(i in 1:max(nfl_actualStats_seasonal$id)){
+  nfl_actualStats_seasonal$season[which(nfl_actualStats_seasonal$id == i)] <- 1998 + i
+}
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_actualStats_seasonal,
+  file = here("data", "nfl_actualStats_seasonal.RData")
+)
+```
+
+### Historical Actual Fantasy Points {#sec-calculateActualPoints}

### 2024-05-22 getting-started.qmd
20240522 - try loading git lfs data file
Revert "20240522 - try loading git lfs data file"
This reverts commit 0e8f950bb10689b76cc914bd41f072854e3f9aeb.

@@
 ```{r}
-#| eval: false
-
 nfl_pbp <- nflfastR::load_pbp(seasons = TRUE)
@@
 )
-```
-
-```{r}
-#| include: false
-
-library("here")
-
-load(file = here("data", "nfl_pbp.RData"))
 ```
@@
 
+```{r}
+#| include: false
+
+library("here")
+
+load(file = here("data", "nfl_actualStats_weekly.RData"))
+```
+
 Third, compute the career stats:
@@
 )
+```
+
+```{r}
+#| include: false
+
+library("here")
+
+load(file = here("data", "nfl_actualStats_career.RData"))
 ```
@@
 
+```{r}
+#| include: false
+
+library("here")
+
+load(file = here("data", "nfl_actualStats_seasonal.RData"))
+```
+
 ### Historical Actual Fantasy Points {#sec-calculateActualPoints}

### 2024-05-22 getting-started.qmd
20240522 - clean up formatting
20240522 - update computations

--- 2024-05-16 getting-started.qmd
@@
 ```
+
+For instance, you can use the following code to install the `nflreadr` package:
+
+```{r}
+#| eval: false
+
+install.packages("nflreadr")
+```
+
+## Load Packages {#sec-loadPackages}
+
+```{r}
+library("nflreadr")
+library("nflfastR")
+library("progressr")
+library("tidyverse")
+```
+
+## Download Football Data {#sec-downloadFootballData}
+
+### Player Info {#sec-downloadPlayerInfo}
+
+### Play-By-Play Data {#sec-downloadPlayByPlay}
+
+To download play-by-play data from prior weeks and seasons, we can use the `load_pbp()` function of the `nflreadr` package.
+We add a progress bar using the `with_progress()` function from the `progressr` package because it takes a while to run.
+
+::: {#nte-downloadPlayByPlay .callout-note title="Downloading play-by-play data"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+
+nfl_pbp <- progressr::with_progress(
+  nflreadr::load_pbp(seasons = TRUE))
+```
+
+```{r}
+#| include: false
+
+nfl_pbp <- nflreadr::load_pbp(seasons = TRUE)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+library("here")
+
+save(
+  nfl_pbp,
+  file = here("data", "nfl_pbp.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_pbp.RData"))
+```
+
+### Historical Weekly Actual Player Statistics {#sec-downloadActualStats}
+
+We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package:
+
+```{r}
+#| eval: false
+
+nfl_actualStats_offense_weekly <- progressr::with_progress(
+  nflreadr::load_player_stats(
+    seasons = TRUE,
+    stat_type = "offense"))
+
+nfl_actualStats_defense_weekly <- progressr::with_progress(
+  nflreadr::load_player_stats(
+    seasons = TRUE,
+    stat_type = "defense"))
+
+nfl_actualStats_kicking_weekly <- progressr::with_progress(
+  nflreadr::load_player_stats(
+    seasons = TRUE,
+    stat_type = "kicking"))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_actualStats_offense_weekly, nfl_actualStats_defense_weekly, nfl_actualStats_kicking_weekly,
+  file = here("data", "nfl_actualStats_weekly.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_actualStats_weekly.RData"))
+```
+
+## Calculations {#sec-calculations}
+
+### Historical Actual Player Statistics {#sec-calculateActualStats}
+
+In addition to week-by-week actual player statistics, we can also compute historical actual player statistics as a function of different timeframes, including season-by-season and career statistics.
+
+#### Career Statistics {#sec-calculateActualCareerStats}
+
+First, we can compute the players' career statistics using the `calculate_player_stats()`, `calculate_player_stats_def()`, and `calculate_player_stats_kicking()` functions from the `nflfastR` package for offensive players, defensive players, and kickers, respectively.
+
+::: {#nte-calculateActualCareerStats .callout-note title="Calculating players' career statistics"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+
+nfl_actualStats_offense_career <- nflfastR::calculate_player_stats(
+  nfl_pbp,
+  weekly = FALSE)
+
+nfl_actualStats_defense_career <- nflfastR::calculate_player_stats_def(
+  nfl_pbp,
+  weekly = FALSE)
+
+nfl_actualStats_kicking_career <- nflfastR::calculate_player_stats_kicking(
+  nfl_pbp,
+  weekly = FALSE)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_actualStats_offense_career, nfl_actualStats_defense_career, nfl_actualStats_kicking_career,
+  file = here("data", "nfl_actualStats_career.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_actualStats_career.RData"))
+```
+
+#### Season-by-Season Statistics {#sec-calculateActualSeasonalStats}
+
+Second, we can compute the players' season-by-season statistics.
+
+```{r}
+seasons <- unique(nfl_pbp$season)
+
+nfl_pbp_seasonalList <- list()
+nfl_actualStats_offense_seasonalList <- list()
+nfl_actualStats_defense_seasonalList <- list()
+nfl_actualStats_kicking_seasonalList <- list()
+```
+
+::: {#nte-calculateActualSeasonalStats .callout-note title="Calculating players' season-by-season statistics"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+
+pb <- txtProgressBar(
+  min = 0,
+  max = length(seasons),
+  style = 3)
+
+for(i in 1:length(seasons)){
+  # Subset play-by-play data by season
+  nfl_pbp_seasonalList[[i]] <- nfl_pbp %>% 
+    filter(season == seasons[i])
+
+  # Compute actual statistics by season
+  nfl_actualStats_offense_seasonalList[[i]] <- 
+    nflfastR::calculate_player_stats(
+      nfl_pbp_seasonalList[[i]],
+      weekly = FALSE)
+
+  nfl_actualStats_defense_seasonalList[[i]] <- 
+    nflfastR::calculate_player_stats_def(
+      nfl_pbp_seasonalList[[i]],
+      weekly = FALSE)
+
+  nfl_actualStats_kicking_seasonalList[[i]] <- 
+    nflfastR::calculate_player_stats_kicking(
+      nfl_pbp_seasonalList[[i]],
+      weekly = FALSE)
+
+  nfl_actualStats_offense_seasonalList[[i]]$season <- seasons[i]
+  nfl_actualStats_defense_seasonalList[[i]]$season <- seasons[i]
+  nfl_actualStats_kicking_seasonalList[[i]]$season <- seasons[i]
+
+  print(
+    paste("Completed computing projections for season: ", seasons[i], sep = ""))
+
+  # Update the progress bar
+  setTxtProgressBar(pb, i)
+}
+
+# Close the progress bar
+close(pb)
+
+nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonalList %>% 
+  bind_rows()
+nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonalList %>% 
+  bind_rows()
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonalList %>% 
+  bind_rows()
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_actualStats_offense_seasonal, nfl_actualStats_defense_seasonal, nfl_actualStats_kicking_seasonal,
+  file = here("data", "nfl_actualStats_seasonal.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_actualStats_seasonal.RData"))
+```
+
+#### Week-by-Week Statistics {#sec-calculateActualWeeklyStats}
+
+We already load players' week-by-week statistics [above](#downloadActualStats).
+Nevertheless, we could compute players' weekly statistics from the play-by-play data using the following syntax:
+
+```{r}
+#| eval: false
+
+nfl_actualStats_weekly <- nflfastR::calculate_player_stats(
+  nfl_pbp,
+  weekly = TRUE)
+```
+
+### Historical Actual Fantasy Points {#sec-calculateActualPoints}

### 2024-05-22 getting-started.qmd
20240522 - fix error
20240522 - fix cross-reference
20240522 - weekly calculations

@@
 
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_pbp.RData"))
-```
-
 ### Historical Weekly Actual Player Statistics {#sec-downloadActualStats}
@@
 #| include: false
+
+library("here")
 
@@
 
-We already load players' week-by-week statistics [above](#downloadActualStats).
+We already load players' week-by-week statistics [above](#sec-downloadActualStats).
 Nevertheless, we could compute players' weekly statistics from the play-by-play data using the following syntax:
@@
 
-nfl_actualStats_weekly <- nflfastR::calculate_player_stats(
+nfl_actualStats_offense_weekly <- nflfastR::calculate_player_stats(
+  nfl_pbp,
+  weekly = TRUE)
+
+nfl_actualStats_defense_weekly <- nflfastR::calculate_player_stats_def(
+  nfl_pbp,
+  weekly = TRUE)
+
+nfl_actualStats_kicking_weekly <- nflfastR::calculate_player_stats_kicking(
   nfl_pbp,

### 2024-05-23 getting-started.qmd
20240522 - rosters
20240522 - fix error
20240522 - duplicate section identifier
20240523 - add data and data dictionaries

@@
 
+### Players {#sec-downloadPlayers}
+
+```{r}
+#| eval: false
+
+nfl_players <- progressr::with_progress(
+  nflreadr::load_players())
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_players,
+  file = here("data", "nfl_players.RData")
+)
+```
+
+```{r}
+#| include: false
+
+library("here")
+
+load(file = here("data", "nfl_players.RData"))
+```
+
+### Teams {#sec-downloadTeams}
+
+```{r}
+#| eval: false
+
+nfl_teams <- progressr::with_progress(
+  nflreadr::load_teams(current = TRUE))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_teams,
+  file = here("data", "nfl_teams.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_teams.RData"))
+```
+
 ### Player Info {#sec-downloadPlayerInfo}
+
+### Rosters {#sec-downloadRosters}
+
+A Data Dictionary for rosters is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_rosters.html>
+
+```{r}
+#| eval: false
+
+nfl_rosters <- progressr::with_progress(
+  nflreadr::load_rosters(seasons = TRUE))
+
+nfl_rosters_weekly <- progressr::with_progress(
+  nflreadr::load_rosters_weekly(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_rosters,
+  file = here("data", "nfl_rosters.RData")
+)
+
+save(
+  nfl_rosters_weekly,
+  file = here("data", "nfl_rosters_weekly.RData")
+)
+```
+
+```{r}
+#| include: false
+
+library("here")
+
+load(file = here("data", "nfl_rosters.RData"))
+load(file = here("data", "nfl_rosters_weekly.RData"))
+```
 
@@
 We add a progress bar using the `with_progress()` function from the `progressr` package because it takes a while to run.
+A Data Dictionary for the play-by-play data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>
 
@@
 
-library("here")
-
 save(
@@
 
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
+```
+
+### Participation {#sec-downloadParticipation}
+
+A Data Dictionary for the participation data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_participation.html>
+
+```{r}
+nfl_participation <- progressr::with_progress(
+  nflreadr::load_participation(
+    seasons = TRUE,
+    include_pbp = TRUE))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_participation,
+  file = here("data", "nfl_participation.RData")
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
+```
+
 ### Historical Weekly Actual Player Statistics {#sec-downloadActualStats}
 
-We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package:
+We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package.
+A Data Dictionary for statistics for offensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
+A Data Dictionary for statistics for defensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.
 
@@
 
-library("here")
-
 load(file = here("data", "nfl_actualStats_weekly.RData"))
 ```
+
+## Data Dictionary {#sec-dataDictionary}
+
+Data Dictionaries are metadata that describe the meaning of the variables in a datset.
+You can find Data Dictionaries for the various NFL datasets at the following link: <https://nflreadr.nflverse.com/articles/index.html>.
 

### 2024-05-24 getting-started.qmd
20240524 - add data files

@@
 
+### Game Schedules {#sec-downloadSchedules}
+
+```{r}
+#| eval: false
+
+nfl_schedules <- progressr::with_progress(
+  nflreadr::load_schedules(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_schedules,
+  file = here("data", "nfl_schedules.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_schedules.RData"))
+```
+
+### Draft Picks {#sec-draftPicks}
+
+```{r}
+#| eval: false
+
+nfl_draftPicks <- progressr::with_progress(
+  nflreadr::load_draft_picks(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_draftPicks,
+  file = here("data", "nfl_draftPicks.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_draftPicks.RData"))
+```
+
 ### Play-By-Play Data {#sec-downloadPlayByPlay}
@@
 load(file = here("data", "nfl_actualStats_weekly.RData"))
+```
+
+### Injuries {#sec-injuries}
+
+```{r}
+#| eval: false
+
+nfl_injuries <- progressr::with_progress(
+  nflreadr::load_injuries(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_injuries,
+  file = here("data", "nfl_injuries.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_injuries.RData"))
+```
+
+### Snap Counts {#sec-snapCounts}
+
+```{r}
+#| eval: false
+
+nfl_snapCounts <- progressr::with_progress(
+  nflreadr::load_snap_counts(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_snapCounts,
+  file = here("data", "nfl_snapCounts.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_snapCounts.RData"))
+```
+
+### ESPN QBR {#sec-espnQBR}
+
+```{r}
+#| eval: false
+
+nfl_espnQBR_seasonal <- progressr::with_progress(
+  nflreadr::load_espn_qbr(
+    seasons = TRUE,
+    summary_type = c("season")))
+
+nfl_espnQBR_weekly <- progressr::with_progress(
+  nflreadr::load_espn_qbr(
+    seasons = TRUE,
+    summary_type = c("weekly")))
+
+nfl_espnQBR_weekly$game_week <- as.character(nfl_espnQBR_weekly$game_week)
+
+nfl_espnQBR <- bind_rows(
+  nfl_espnQBR_seasonal,
+  nfl_espnQBR_weekly
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_espnQBR,
+  file = here("data", "nfl_espnQBR.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_espnQBR.RData"))
+```
+
+### Advanced Stats from PFR {#sec-advancedStatsPFR}
+
+```{r}
+#| eval: false
+
+nfl_advancedStatsPFR_pass_seasonal <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("pass"),
+    summary_level = c("season")))
+
+nfl_advancedStatsPFR_pass_weekly <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("pass"),
+    summary_level = c("week")))
+
+nfl_advancedStatsPFR_rush_seasonal <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("rush"),
+    summary_level = c("season")))
+
+nfl_advancedStatsPFR_rush_weekly <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("rush"),
+    summary_level = c("week")))
+
+nfl_advancedStatsPFR_rec_seasonal <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("rec"),
+    summary_level = c("season")))
+
+nfl_advancedStatsPFR_rec_weekly <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("rec"),
+    summary_level = c("week")))
+
+nfl_advancedStatsPFR_def_seasonal <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("def"),
+    summary_level = c("season")))
+
+nfl_advancedStatsPFR_def_weekly <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("def"),
+    summary_level = c("week")))
+
+nfl_advancedStatsPFR <- bind_rows(
+  nfl_advancedStatsPFR_pass_seasonal,
+  nfl_advancedStatsPFR_pass_weekly,
+  nfl_advancedStatsPFR_rush_seasonal,
+  nfl_advancedStatsPFR_rush_weekly,
+  nfl_advancedStatsPFR_rec_seasonal,
+  nfl_advancedStatsPFR_rec_weekly,
+  nfl_advancedStatsPFR_def_seasonal,
+  nfl_advancedStatsPFR_def_weekly,
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_advancedStatsPFR,
+  file = here("data", "nfl_advancedStatsPFR.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_advancedStatsPFR.RData"))
+```
+
+### Player Contracts {#sec-playerContracts}
+
+```{r}
+#| eval: false
+
+nfl_playerContracts <- progressr::with_progress(
+  nflreadr::load_contracts())
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_playerContracts,
+  file = here("data", "nfl_playerContracts.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_playerContracts.RData"))
+```
+
+### FTN Charting Data {#sec-ftnCharting}
+
+```{r}
+#| eval: false
+
+nfl_ftnCharting <- progressr::with_progress(
+  nflreadr::load_ftn_charting(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_ftnCharting,
+  file = here("data", "nfl_ftnCharting.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_ftnCharting.RData"))
+```
+
+### Fantasy Player IDs {#sec-fantasyPlayerIDs}
+
+```{r}
+#| eval: false
+
+nfl_playerIDs <- progressr::with_progress(
+  nflreadr::load_ff_playerids())
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_playerIDs,
+  file = here("data", "nfl_playerIDs.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_playerIDs.RData"))
+```
+
+### FantasyPros Rankings {#sec-fantasyProsRankings}
+
+```{r}
+#| eval: false
+
+#nfl_rankings <- progressr::with_progress( # currently throws error
+#  nflreadr::load_ff_rankings(type = "all"))
+
+nfl_rankings_draft <- progressr::with_progress(
+  nflreadr::load_ff_rankings(type = "draft"))
+
+nfl_rankings_weekly <- progressr::with_progress(
+  nflreadr::load_ff_rankings(type = "week"))
+
+nfl_rankings <- bind_rows(
+  nfl_rankings_draft,
+  nfl_rankings_weekly
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_rankings,
+  file = here("data", "nfl_rankings.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_rankings.RData"))
+```
+
+### Expected Fantasy Points {#sec-expectedFantasyPoints}
+
+```{r}
+#| eval: false
+
+nfl_expectedFantasyPoints_weekly <- progressr::with_progress(
+  nflreadr::load_ff_opportunity(
+    seasons = TRUE,
+    stat_type = "weekly",
+    model_version = "latest"
+  ))
+
+nfl_expectedFantasyPoints_pass <- progressr::with_progress(
+  nflreadr::load_ff_opportunity(
+    seasons = TRUE,
+    stat_type = "pbp_pass",
+    model_version = "latest"
+  ))
+
+nfl_expectedFantasyPoints_rush <- progressr::with_progress(
+  nflreadr::load_ff_opportunity(
+    seasons = TRUE,
+    stat_type = "pbp_rush",
+    model_version = "latest"
+  ))
+
+nfl_expectedFantasyPoints_weekly$season <- as.integer(nfl_expectedFantasyPoints_weekly$season)
+
+nfl_expectedFantasyPoints_offense <- bind_rows(
+  nfl_expectedFantasyPoints_pass,
+  nfl_expectedFantasyPoints_rush
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_expectedFantasyPoints_weekly,
+  file = here("data", "nfl_expectedFantasyPoints_weekly.RData")
+)
+
+save(
+  nfl_expectedFantasyPoints_offense,
+  file = here("data", "nfl_expectedFantasyPoints_offense.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))
+
+load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
 ```

### 2024-05-24 getting-started.qmd
20240524 - add data

@@
 
+### The Combine {#sec-combine}
+
+```{r}
+#| eval: false
+
+nfl_combine <- progressr::with_progress(
+  nflreadr::load_combine(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_combine,
+  file = here("data", "nfl_combine.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_combine.RData"))
+```
+
 ### Draft Picks {#sec-draftPicks}
@@
 load(file = here("data", "nfl_draftPicks.RData"))
+```
+
+### Weekly Depth Charts {#sec-depthCharts}
+
+```{r}
+#| eval: false
+
+nfl_depthCharts <- progressr::with_progress(
+  nflreadr::load_depth_charts(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_depthCharts,
+  file = here("data", "nfl_depthCharts.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_depthCharts.RData"))
 ```
@@
 load(file = here("data", "nfl_espnQBR.RData"))
+```
+
+### NFL Next Gen Stats {#sec-nextGenStats}
+
+```{r}
+#| eval: false
+
+nfl_nextGenStats_pass_weekly <- progressr::with_progress(
+  nflreadr::load_nextgen_stats(
+    seasons = TRUE,
+    stat_type = c("passing")))
+
+nfl_nextGenStats_rush_weekly <- progressr::with_progress(
+  nflreadr::load_nextgen_stats(
+    seasons = TRUE,
+    stat_type = c("rushing")))
+
+nfl_nextGenStats_rec_weekly <- progressr::with_progress(
+  nflreadr::load_nextgen_stats(
+    seasons = TRUE,
+    stat_type = c("receiving")))
+
+nfl_nextGenStats_weekly <- bind_rows(
+  nfl_nextGenStats_pass_weekly,
+  nfl_nextGenStats_rush_weekly,
+  nfl_nextGenStats_rec_weekly
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_nextGenStats_weekly,
+  file = here("data", "nfl_nextGenStats_weekly.RData")
+)
+```
+
+```{r}
+#| include: false
+
+load(file = here("data", "nfl_nextGenStats_weekly.RData"))
 ```

### 2024-05-24 getting-started.qmd
20240524 - data dictionaries

@@
 
+A Data Dictionary for game schedules data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_schedules.html>
+
 ```{r}
@@
 ### The Combine {#sec-combine}
+
+A Data Dictionary for data from the combine is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_combine.html>
 
@@
 
+A Data Dictionary for draft picks data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_draft_picks.html>
+
 ```{r}
@@
 
-### Weekly Depth Charts {#sec-depthCharts}
+### Depth Charts {#sec-depthCharts}
+
+A Data Dictionary for data from weekly depth charts is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_depth_charts.html>
 
@@
 
+A Data Dictionary for injury data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_injuries.html>
+
 ```{r}
@@
 
+A Data Dictionary for snap counts data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_snap_counts.html>
+
 ```{r}
@@
 ### ESPN QBR {#sec-espnQBR}
+
+A Data Dictionary for ESPN QBR data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_espn_qbr.html>
 
@@
 ### NFL Next Gen Stats {#sec-nextGenStats}
+
+A Data Dictionary for NFL Next Gen Stats data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_nextgen_stats.html>
 
@@
 ### Advanced Stats from PFR {#sec-advancedStatsPFR}
+
+A Data Dictionary for PFR passing data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html>
 
@@
 
+A Data Dictionary for player contracts data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_contracts.html>
+
 ```{r}
@@
 ### FTN Charting Data {#sec-ftnCharting}
+
+A Data Dictionary for FTN Charting data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ftn_charting.html>
 
@@
 
+A Data Dictionary for fantasy player ID data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_playerids.html>
+
 ```{r}
@@
 ### FantasyPros Rankings {#sec-fantasyProsRankings}
+
+A Data Dictionary for FantasyPros ranking data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_rankings.html>
 
@@
 ### Expected Fantasy Points {#sec-expectedFantasyPoints}
+
+A Data Dictionary for expected fantasy points data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_opportunity.html>
 

### 2024-05-25 getting-started.qmd
20240525 - calculate player age

@@
 library("progressr")
+library("lubridate")
 library("tidyverse")
@@
 ## Download Football Data {#sec-downloadFootballData}
+
+```{r}
+#| eval: false
+#| include: false
+
+library("here")
+
+load(file = here("data", "nfl_players.RData"))
+load(file = here("data", "nfl_teams.RData"))
+load(file = here("data", "nfl_rosters.RData"))
+load(file = here("data", "nfl_rosters_weekly.RData"))
+load(file = here("data", "nfl_schedules.RData"))
+load(file = here("data", "nfl_combine.RData"))
+load(file = here("data", "nfl_draftPicks.RData"))
+load(file = here("data", "nfl_depthCharts.RData"))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
+load(file = here("data", "nfl_actualStats_weekly.RData"))
+load(file = here("data", "nfl_injuries.RData"))
+load(file = here("data", "nfl_snapCounts.RData"))
+load(file = here("data", "nfl_espnQBR.RData"))
+load(file = here("data", "nfl_nextGenStats_weekly.RData"))
+load(file = here("data", "nfl_advancedStatsPFR.RData"))
+load(file = here("data", "nfl_playerContracts.RData"))
+load(file = here("data", "nfl_ftnCharting.RData"))
+load(file = here("data", "nfl_playerIDs.RData"))
+load(file = here("data", "nfl_rankings.RData"))
+load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))
+load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
+```
 
@@
 #| include: false
-
-library("here")
 
@@
 load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))
-
 load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
@@
 ### Historical Actual Fantasy Points {#sec-calculateActualPoints}
+
+### Player Age {#sec-calculateActualPoints}
+
+```{r}
+# Reshape from wide to long format
+nfl_actualStats_offense_weekly_long <- nfl_actualStats_offense_weekly %>% 
+  pivot_longer(
+    cols = c(recent_team, opponent_team),
+    names_to = "role",
+    values_to = "team")
+
+# Reshape from wide to long format
+nfl_schedules_long <- nfl_schedules %>% 
+  pivot_longer(
+    cols = c(home_team, away_team),
+    names_to = "role",
+    values_to = "team")
+
+# Perform separate inner join operations for the home_team and away_team
+nfl_actualStats_offense_weekly_home <- inner_join(
+  nfl_actualStats_offense_weekly_long,
+  nfl_schedules,
+  by = c("season","week","team" = "home_team")) %>% 
+  mutate(home_away = "home_team")
+
+nfl_actualStats_offense_weekly_away <- inner_join(
+  nfl_actualStats_offense_weekly_long,
+  nfl_schedules,
+  by = c("season","week","team" = "away_team")) %>% 
+  mutate(home_away = "away_team")
+
+# Combine the results of the join operations
+nfl_actualStats_offense_weekly_schedules_long <- bind_rows(
+  nfl_actualStats_offense_weekly_home,
+  nfl_actualStats_offense_weekly_away)
+
+# Reshape from long to wide
+nfl_actualStats_offense_weekly_schedules <- nfl_actualStats_offense_weekly_schedules_long %>%
+  distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
+  pivot_wider(
+    names_from = home_away,
+    values_from = team)
+
+# Merge player birthdate and the game date
+player_birthdate_gameday <- left_join(
+  nfl_actualStats_offense_weekly_schedules,
+  unique(nfl_players[,c("gsis_id","birth_date")]),
+  by = c("player_id" = "gsis_id")
+)
+
+player_birthdate_gameday$birth_date <- ymd(player_birthdate_gameday$birth_date)
+player_birthdate_gameday$gameday <- ymd(player_birthdate_gameday$gameday)
+
+# Calculate player's age for a given week as the difference between their birthdate and the game date
+player_birthdate_gameday$age <- interval(
+  start = player_birthdate_gameday$birth_date,
+  end = player_birthdate_gameday$gameday
+) %>% 
+  time_length(unit = "years")
+
+# Merge with player info
+player_age <-  left_join(
+  player_birthdate_gameday,
+  nfl_players %>% select(-birth_date),
+  by = c("season", "player_id" = "gsis_id"))
+
+# Add game_id to weekly stats to facilitate merging
+nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly %>% 
+  left_join(
+    player_age[,c("season","week","player_id","game_id")],
+    by = c("season","week","player_id"))
+
+# Merge with player weekly stats
+player_age_stats <- left_join(
+  player_age,
+  nfl_actualStats_offense_weekly,
+  by = c(c("season","week","player_id","game_id")))
+
+player_age_stats$years_of_experience <- as.integer(player_age_stats$years_of_experience)
+```

### 2024-05-25 getting-started.qmd
20240525 - player age

@@
 
+library("here")
+
 load(file = here("data", "nfl_players.RData"))
@@
 
-# Reshape from wide to long format
-nfl_schedules_long <- nfl_schedules %>% 
-  pivot_longer(
-    cols = c(home_team, away_team),
-    names_to = "role",
-    values_to = "team")
-
 # Perform separate inner join operations for the home_team and away_team
@@
 
+nfl_actualStats_defense_weekly_home <- inner_join(
+  nfl_actualStats_defense_weekly,
+  nfl_schedules,
+  by = c("season","week","team" = "home_team")) %>% 
+  mutate(home_away = "home_team")
+
+nfl_actualStats_defense_weekly_away <- inner_join(
+  nfl_actualStats_defense_weekly,
+  nfl_schedules,
+  by = c("season","week","team" = "away_team")) %>% 
+  mutate(home_away = "away_team")
+
+nfl_actualStats_kicking_weekly_home <- inner_join(
+  nfl_actualStats_kicking_weekly,
+  nfl_schedules,
+  by = c("season","week","team" = "home_team")) %>% 
+  mutate(home_away = "home_team")
+
+nfl_actualStats_kicking_weekly_away <- inner_join(
+  nfl_actualStats_kicking_weekly,
+  nfl_schedules,
+  by = c("season","week","team" = "away_team")) %>% 
+  mutate(home_away = "away_team")
+
 # Combine the results of the join operations
@@
 
+nfl_actualStats_defense_weekly_schedules_long <- bind_rows(
+  nfl_actualStats_defense_weekly_home,
+  nfl_actualStats_defense_weekly_away)
+
+nfl_actualStats_kicking_weekly_schedules_long <- bind_rows(
+  nfl_actualStats_kicking_weekly_home,
+  nfl_actualStats_kicking_weekly_away)
+
 # Reshape from long to wide
-nfl_actualStats_offense_weekly_schedules <- nfl_actualStats_offense_weekly_schedules_long %>%
+player_game_gameday_offense <- nfl_actualStats_offense_weekly_schedules_long %>%
   distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
@@
 
+player_game_gameday_defense <- nfl_actualStats_defense_weekly_schedules_long %>%
+  distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
+  pivot_wider(
+    names_from = home_away,
+    values_from = team)
+
+player_game_gameday_kicking <- nfl_actualStats_kicking_weekly_schedules_long %>%
+  distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
+  pivot_wider(
+    names_from = home_away,
+    values_from = team)
+
 # Merge player birthdate and the game date
-player_birthdate_gameday <- left_join(
-  nfl_actualStats_offense_weekly_schedules,
+player_game_birthdate_gameday_offense <- left_join(
+  player_game_gameday_offense,
   unique(nfl_players[,c("gsis_id","birth_date")]),
@@
 
-player_birthdate_gameday$birth_date <- ymd(player_birthdate_gameday$birth_date)
-player_birthdate_gameday$gameday <- ymd(player_birthdate_gameday$gameday)
+player_game_birthdate_gameday_defense <- left_join(
+  player_game_gameday_defense,
+  unique(nfl_players[,c("gsis_id","birth_date")]),
+  by = c("player_id" = "gsis_id")
+)
+
+player_game_birthdate_gameday_kicking <- left_join(
+  player_game_gameday_kicking,
+  unique(nfl_players[,c("gsis_id","birth_date")]),
+  by = c("player_id" = "gsis_id")
+)
+
+player_game_birthdate_gameday_offense$birth_date <- ymd(player_game_birthdate_gameday_offense$birth_date)
+player_game_birthdate_gameday_offense$gameday <- ymd(player_game_birthdate_gameday_offense$gameday)
+
+player_game_birthdate_gameday_defense$birth_date <- ymd(player_game_birthdate_gameday_defense$birth_date)
+player_game_birthdate_gameday_defense$gameday <- ymd(player_game_birthdate_gameday_defense$gameday)
+
+player_game_birthdate_gameday_kicking$birth_date <- ymd(player_game_birthdate_gameday_kicking$birth_date)
+player_game_birthdate_gameday_kicking$gameday <- ymd(player_game_birthdate_gameday_kicking$gameday)
 
 # Calculate player's age for a given week as the difference between their birthdate and the game date
-player_birthdate_gameday$age <- interval(
-  start = player_birthdate_gameday$birth_date,
-  end = player_birthdate_gameday$gameday
+player_game_birthdate_gameday_offense$age <- interval(
+  start = player_game_birthdate_gameday_offense$birth_date,
+  end = player_game_birthdate_gameday_offense$gameday
 ) %>% 
@@
 
+player_game_birthdate_gameday_defense$age <- interval(
+  start = player_game_birthdate_gameday_defense$birth_date,
+  end = player_game_birthdate_gameday_defense$gameday
+) %>% 
+  time_length(unit = "years")
+
+player_game_birthdate_gameday_kicking$age <- interval(
+  start = player_game_birthdate_gameday_kicking$birth_date,
+  end = player_game_birthdate_gameday_kicking$gameday
+) %>% 
+  time_length(unit = "years")
+
 # Merge with player info
-player_age <-  left_join(
-  player_birthdate_gameday,
+player_age_offense <-  left_join(
+  player_game_birthdate_gameday_offense,
   nfl_players %>% select(-birth_date),
@@
 
+player_age_defense <-  left_join(
+  player_game_birthdate_gameday_defense,
+  nfl_players %>% select(-birth_date),
+  by = c("season", "player_id" = "gsis_id"))
+
+player_age_kicking <-  left_join(
+  player_game_birthdate_gameday_kicking,
+  nfl_players %>% select(-birth_date),
+  by = c("season", "player_id" = "gsis_id"))
+
 # Add game_id to weekly stats to facilitate merging
-nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly %>% 
+nfl_actualStats_game_offense_weekly <- nfl_actualStats_offense_weekly %>% 
   left_join(
-    player_age[,c("season","week","player_id","game_id")],
+    player_age_offense[,c("season","week","player_id","game_id")],
     by = c("season","week","player_id"))
 
+nfl_actualStats_game_defense_weekly <- nfl_actualStats_defense_weekly %>% 
+  left_join(
+    player_age_offense[,c("season","week","player_id","game_id")],
+    by = c("season","week","player_id"))
+
+nfl_actualStats_game_kicking_weekly <- nfl_actualStats_kicking_weekly %>% 
+  left_join(
+    player_age_offense[,c("season","week","player_id","game_id")],
+    by = c("season","week","player_id"))
+
 # Merge with player weekly stats
-player_age_stats <- left_join(
-  player_age,
-  nfl_actualStats_offense_weekly,
+player_age_stats_offense <- left_join(
+  player_age_offense,
+  nfl_actualStats_game_offense_weekly,
   by = c(c("season","week","player_id","game_id")))
 
-player_age_stats$years_of_experience <- as.integer(player_age_stats$years_of_experience)
-```
+player_age_stats_defense <- left_join(
+  player_age_defense,
+  nfl_actualStats_game_defense_weekly,
+  by = c(c("season","week","player_id","game_id")))
+
+player_age_stats_kicking <- left_join(
+  player_age_kicking,
+  nfl_actualStats_game_kicking_weekly,
+  by = c(c("season","week","player_id","game_id")))
+
+player_age_stats_offense$years_of_experience <- as.integer(player_age_stats_offense$years_of_experience)
+player_age_stats_defense$years_of_experience <- as.integer(player_age_stats_defense$years_of_experience)
+player_age_stats_kicking$years_of_experience <- as.integer(player_age_stats_kicking$years_of_experience)
+```

### 2024-05-25 getting-started.qmd
20240525 - remove duplicate identifier
20240525 - fix merge error

@@
 
-### Player Age {#sec-calculateActualPoints}
+### Player Age {#sec-calculatePlayerAge}
 
@@
   player_game_birthdate_gameday_offense,
-  nfl_players %>% select(-birth_date),
-  by = c("season", "player_id" = "gsis_id"))
+  nfl_players %>% select(-birth_date, -season),
+  by = c("player_id" = "gsis_id"))
 
@@
   player_game_birthdate_gameday_defense,
-  nfl_players %>% select(-birth_date),
-  by = c("season", "player_id" = "gsis_id"))
+  nfl_players %>% select(-birth_date, -season),
+  by = c("player_id" = "gsis_id"))
 
@@
   player_game_birthdate_gameday_kicking,
-  nfl_players %>% select(-birth_date),
-  by = c("season", "player_id" = "gsis_id"))
+  nfl_players %>% select(-birth_date, -season),
+  by = c("player_id" = "gsis_id"))
 

### 2024-05-25 bugs.qmd
20240525 - bugs to report

# Bugs {#sec-bugs}

Bugs to report (`nflreadr`):

- `players` database has a season column but it is only populated for some players, and it is only populated with 2020
- `player_stats` database of weekly stats does not have a `game_id` column

### 2024-05-26 getting-started.qmd
20240525 - plot of EPA per play

@@
 library("nflfastR")
+library("nflplotR")
 library("progressr")
@@
 ```
+
+## Plotting {#sec-plotting}
+
+### Defensive and Offensive EPA per Play {#sec-plotDefOffEPA}
+
+Expected points added (EPA) per play by the team with possession.
+
+```{r}
+pbp_regularSeason <- nfl_pbp %>% 
+  dplyr::filter(
+    season == 2023,
+    season_type == "REG") %>%
+  dplyr::filter(!is.na(posteam) & (rush == 1 | pass == 1))
+
+epa_offense <- pbp_regularSeason %>%
+  dplyr::group_by(team = posteam) %>%
+  dplyr::summarise(off_epa = mean(epa, na.rm = TRUE))
+
+epa_defense <- pbp_regularSeason %>%
+  dplyr::group_by(team = defteam) %>%
+  dplyr::summarise(def_epa = mean(epa, na.rm = TRUE))
+
+epa_combined <- epa_offense %>%
+  dplyr::inner_join(epa_defense, by = "team")
+
+ggplot2::ggplot(
+  data = epa_combined,
+  ggplot2::aes(
+    x = off_epa,
+    y = def_epa)) +
+  nflplotR::geom_mean_lines(
+    ggplot2::aes(
+      x0 = off_epa ,
+      y0 = def_epa)) +
+  nflplotR::geom_nfl_logos(
+    ggplot2::aes(
+      team_abbr = team),
+      width = 0.065,
+      alpha = 0.7) +
+  ggplot2::labs(
+    x = "Offense EPA/play",
+    y = "Defense EPA/play",
+    title = "2023 NFL Offensive and Defensive EPA per Play"
+  ) +
+  ggplot2::theme_classic() +
+  ggplot2::scale_y_reverse()
+```

### 2024-05-26 getting-started.qmd
20240525 - figure label and caption

@@
 ```{r}
+#| label: fig-plotDefOffEPA
+#| fig-cap: "2023 NFL Offensive and Defensive EPA per Play"
+
 pbp_regularSeason <- nfl_pbp %>% 

### 2024-05-26 getting-started.qmd
20240525 - rushing EPA by age; ypc by age

@@
 
+# Downloaded Data
 load(file = here("data", "nfl_players.RData"))
@@
 load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
+
+# Calculated Data
+load(file = here("data", "nfl_actualStats_career.RData"))
+load(file = here("data", "nfl_actualStats_seasonal.RData"))
 ```
@@
 player_age_stats_offense <- left_join(
-  player_age_offense,
+  player_age_offense %>% select(-position, -position_group),
   nfl_actualStats_game_offense_weekly,
@@
 player_age_stats_defense <- left_join(
-  player_age_defense,
+  player_age_defense %>% select(-position, -position_group),
   nfl_actualStats_game_defense_weekly,
@@
 player_age_stats_kicking <- left_join(
-  player_age_kicking,
+  player_age_kicking %>% select(-position, -position_group),
   nfl_actualStats_game_kicking_weekly,
@@
 player_age_stats_kicking$years_of_experience <- as.integer(player_age_stats_kicking$years_of_experience)
+
+# Merge player info with seasonal stats
+player_seasonal_offense <- left_join(
+  nfl_actualStats_offense_seasonal,
+  nfl_players %>% select(-position, -position_group, -season),
+  by = c("player_id" = "gsis_id")
+)
+
+player_seasonal_defense <- left_join(
+  nfl_actualStats_defense_seasonal,
+  nfl_players %>% select(-position, -position_group, -season),
+  by = c("player_id" = "gsis_id")
+)
+
+player_seasonal_kicking <- left_join(
+  nfl_actualStats_kicking_seasonal,
+  nfl_players %>% select(-position, -position_group, -season),
+  by = c("player_id" = "gsis_id")
+)
+
+# Calculate age
+season_startdate <- nfl_schedules %>% 
+  group_by(season) %>% 
+  summarise(startdate = min(gameday, na.rm = TRUE))
+
+player_seasonal_offense <- player_seasonal_offense %>% 
+  left_join(
+    season_startdate,
+    by = "season"
+  )
+
+player_seasonal_defense <- player_seasonal_defense %>% 
+  left_join(
+    season_startdate,
+    by = "season"
+  )
+
+player_seasonal_kicking <- player_seasonal_kicking %>% 
+  left_join(
+    season_startdate,
+    by = "season"
+  )
+
+player_seasonal_offense$age <- interval(
+  start = player_seasonal_offense$birth_date,
+  end = player_seasonal_offense$startdate
+) %>% 
+  time_length(unit = "years")
+
+player_seasonal_defense$age <- interval(
+  start = player_seasonal_defense$birth_date,
+  end = player_seasonal_defense$startdate
+) %>% 
+  time_length(unit = "years")
+
+player_seasonal_kicking$age <- interval(
+  start = player_seasonal_kicking$birth_date,
+  end = player_seasonal_kicking$startdate
+) %>% 
+  time_length(unit = "years")
 ```
@@
 ## Plotting {#sec-plotting}
+
+### Rushing Yards per Carry By Player Age {#sec-plotRushEPAbyAge}
+
+```{r}
+# Prepare Data
+rushing_attempts <- nfl_pbp %>% 
+  dplyr::filter(
+    season_type == "REG") %>% 
+    filter(
+      rush == 1,
+      rush_attempt == 1,
+      qb_scramble == 0,
+      qb_dropback == 0,
+      !is.na(rushing_yards))
+
+rb_yardsPerCarry <- rushing_attempts %>% 
+  group_by(rusher_id, season) %>% 
+  summarise(
+    ypc = mean(rushing_yards, na.rm = TRUE),
+    rush_attempts = n(),
+    .groups = "drop") %>% 
+  ungroup() %>% 
+  left_join(
+    nfl_players %>% select(-season),
+    by = c("rusher_id" = "gsis_id")
+  ) %>% 
+  filter(
+    position_group == "RB",
+    rush_attempts >= 50) %>% 
+  left_join(
+    season_startdate,
+    by = "season"
+  )
+
+rb_yardsPerCarry$age <- interval(
+  start = rb_yardsPerCarry$birth_date,
+  end = rb_yardsPerCarry$startdate
+) %>% 
+  time_length(unit = "years")
+
+# Create Plot
+ggplot2::ggplot(
+  data = rb_yardsPerCarry,
+  ggplot2::aes(
+    x = age,
+    y = ypc)) +
+  ggplot2::geom_point() +
+  ggplot2::geom_smooth() +
+  ggplot2::labs(
+    x = "Rushing Back Age (years)",
+    y = "Rushing Yards per Carry/season",
+    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
+    subtitle = "(minimum 50 rushing attempts)"
+  ) +
+  ggplot2::theme_classic()
+```
+
+```{r}
+# Subset Data
+rb_seasonal <- player_seasonal_offense %>% 
+  filter(position_group == "RB")
+
+# Create Plot
+ggplot2::ggplot(
+  data = rb_seasonal,
+  ggplot2::aes(
+    x = age,
+    y = rushing_epa)) +
+  ggplot2::geom_point() +
+  ggplot2::geom_smooth() +
+  ggplot2::labs(
+    x = "Rushing Back Age (years)",
+    y = "Rushing EPA/season",
+    title = "2023 NFL Rushing EPA per Season by Player Age"
+  ) +
+  ggplot2::theme_classic()
+```
 

### 2024-05-26 getting-started.qmd
20240526 - variable names and logical operators

@@
 
+## Variable Names {#sec-variableNames}
+
+To see the names of variables in a data frame, use the following syntax:
+
+```{r}
+names(nfl_players)
+```
+
+## Logical Operators {#sec-logicalOperators}
+
+### Is Equal To: `==` {#sec-logicalOperatorEqual}
+
+```{r}
+nfl_players$position_group == "RB"
+```
+
+### Is Not Equal To: `!=` {#sec-logicalOperatorNotEqual}
+
+```{r}
+nfl_players$position_group != "RB"
+```
+
+### Is Greater Than: `>` {#sec-logicalOperatorGreaterThan}
+
+```{r}
+nfl_players$weight > 300
+```
+
+### Is Less Than: `<` {#sec-logicalOperatorLessThan}
+
+```{r}
+nfl_players$weight < 300
+```
+
+### Is Greater Than or Equal To: `>=` {#sec-logicalOperatorGreaterThanEqual}
+
+```{r}
+nfl_players$weight >= 300
+```
+
+### Is Less Than or Equal To: `<=` {#sec-logicalOperatorLessThanEqual}
+
+```{r}
+nfl_players$weight <= 300
+```
+
+### Is In a Value of Another Vector: `%in%` {#sec-logicalOperatorInVector}
+
+```{r}
+nfl_players$position_group %in% c("QB","RB","WR")
+```
+
+### Is Not In a Value of Another Vector: `!(%in%)` {#sec-logicalOperatorNotInVector}
+
+```{r}
+!(nfl_players$position_group %in% c("QB","RB","WR"))
+```
+
+### Is Missing: `is.na()` {#sec-logicalOperatorMissing}
+
+```{r}
+is.na(nfl_players$college_name)
+```
+
+### Is Not Missing: `!is.na()` {#sec-logicalOperatorNotMissing}
+
+```{r}
+!is.na(nfl_players$college_name)
+```
+
+### And: `&`{#sec-logicalOperatorAnd}
+
+```{r}
+nfl_players$position_group == "RB" & nfl_players$weight > 230
+```
+
+### Or: `|` {#sec-logicalOperatorOr}
+
+```{r}
+nfl_players$position_group == "RB" | nfl_players$weight > 230
+```
+
 ## Calculations {#sec-calculations}

### 2024-05-26 getting-started.qmd
20240526 - subset

@@
 
+## Subset {#sec-subset}
+
+To subset a data frame, use brackets to specify the subset of rows and columns to keep, where the value/vector before the comma specifies the rows to keep, and the value/vector after the comma specifies the columns to keep:
+
+```{r, eval = FALSE}
+dataframe[rowsToKeep, columnsToKeep]
+```
+
+You can subset by using any of the following:
+
+- numeric indices of the rows/columns to keep (or drop)
+- names of the rows/columns to keep (or drop)
+- values of `TRUE` and `FALSE` corresponding to which rows/columns to keep
+
+### One Variable {#sec-subsetOneVariable}
+
+To subset one variable, use the following syntax:
+
+```{r}
+nfl_players$display_name
+```
+
+or:
+
+```{r}
+nfl_players[,"display_name"]
+```
+
+### Particular Rows of One Variable {#sec-subsetParticularRowsOneVariable}
+
+To subset one variable, use the following syntax:
+
+```{r}
+nfl_players$display_name[which(nfl_players$position_group == "RB")]
+```
+
+or:
+
+```{r}
+nfl_players[which(nfl_players$position_group == "RB"), "display_name"]
+```
+
+### Particular Columns (Variables) {#sec-subsetParticularColumns}
+
+To subset particular columns/variables, use the following syntax:
+
+#### Base `R` {#sec-subsetParticularColumnsBaseR}
+
+```{r}
+subsetVars <- c("status","display_name","college_name")
+
+nfl_players[,c(1,2,9)]
+nfl_players[,c("status","display_name","college_name")]
+nfl_players[,subsetVars]
+```
+
+Or, to drop columns:
+
+```{r}
+dropVars <- c("status","college_name")
+
+nfl_players[,-c(1,9)]
+nfl_players[,!(names(nfl_players) %in% c("status","college_name"))]
+nfl_players[,!(names(nfl_players) %in% dropVars)]
+```
+
+#### Tidyverse {#sec-subsetParticularColumnsTidyverse}
+
+```{r}
+nfl_players %>%
+  select(status, display_name, college_name)
+
+nfl_players %>%
+  select(status:college_name)
+
+nfl_players %>%
+  select(all_of(subsetVars))
+```
+
+Or, to drop columns:
+
+```{r}
+nfl_players %>%
+  select(-status, -college_name)
+
+nfl_players %>%
+  select(-c(status:college_name))
+
+nfl_players %>%
+  select(-all_of(dropVars))
+```
+
+### Particular Rows {#sec-subsetParticularRows}
+
+To subset particular rows, use the following syntax:
+
+#### Base `R` {#sec-subsetParticularRowsBaseR}
+
+```{r}
+subsetRows <- c(1,3,5)
+
+nfl_players[c(1,3,5),]
+nfl_players[subsetRows,]
+nfl_players[which(nfl_players$position_group == "RB"),]
+```
+
+#### Tidyverse {#sec-subsetParticularRowsTidyverse}
+
+```{r}
+nfl_players %>%
+  filter(position_group == "RB")
+
+nfl_players %>%
+  filter(position_group == "RB", weight <= 250)
+
+nfl_players %>%
+  filter(position_group == "RB" | weight >= 250)
+```
+
+### Particular Rows and Columns {#sec-subsetParticularRowsAndColumns}
+
+To subset particular rows and columns, use the following syntax:
+
+#### Base `R` {#sec-subsetParticularRowsAndColumnsBaseR}
+
+```{r}
+nfl_players[c(1,3,5), c(1,2,3)]
+nfl_players[subsetRows, subsetVars]
+nfl_players[which(nfl_players$position_group == "RB"), subsetVars]
+```
+
+#### Tidyverse {#sec-subsetParticularRowsAndColumnsTidyverse}
+
+```{r}
+nfl_players %>%
+  filter(position_group == "RB") %>%
+  select(all_of(subsetVars))
+```
+
 ## Calculations {#sec-calculations}

### 2024-05-26 getting-started.qmd
20240526 - additional data management

@@
 ```{r}
+#| eval: false
+
 nfl_players$position_group == "RB"
@@
 ```{r}
+#| eval: false
+
 nfl_players$position_group != "RB"
@@
 ```{r}
+#| eval: false
+
 nfl_players$weight > 300
@@
 ```{r}
+#| eval: false
+
 nfl_players$weight < 300
@@
 ```{r}
+#| eval: false
+
 nfl_players$weight >= 300
@@
 ```{r}
+#| eval: false
+
 nfl_players$weight <= 300
@@
 ```{r}
+#| eval: false
+
 nfl_players$position_group %in% c("QB","RB","WR")
@@
 ```{r}
+#| eval: false
+
 !(nfl_players$position_group %in% c("QB","RB","WR"))
@@
 ```{r}
+#| eval: false
+
 is.na(nfl_players$college_name)
@@
 ```{r}
+#| eval: false
+
 !is.na(nfl_players$college_name)
@@
 ```{r}
+#| eval: false
+
 nfl_players$position_group == "RB" & nfl_players$weight > 230
@@
 ```{r}
+#| eval: false
+
 nfl_players$position_group == "RB" | nfl_players$weight > 230
@@
 
-```{r, eval = FALSE}
+```{r}
+#| eval: false
+
 dataframe[rowsToKeep, columnsToKeep]
@@
 ```{r}
+#| eval: false
+
 nfl_players$display_name
@@
 ```{r}
+#| eval: false
+
 nfl_players$display_name[which(nfl_players$position_group == "RB")]
@@
   select(all_of(subsetVars))
+```
+
+## View Data {#sec-viewData}
+
+### All Data {#sec-viewDataAll}
+
+To view data, use the following syntax:
+
+```{r}
+#| eval: false
+
+View(nfl_players)
+```
+
+### First 6 Rows/Elements {#sec-viewDataHead}
+
+To view only the first six rows (if a data frame) or elements (if a vector), use the following syntax:
+
+```{r}
+head(nfl_players)
+head(nfl_players$display_name)
+```
+
+## Data Characteristics {#sec-dataCharacteristics}
+
+### Data Structure {#sec-dataCharacteristicsStructure}
+
+```{r}
+str(nfl_players)
+```
+
+### Data Dimensions {#sec-dataCharacteristicsDimensions}
+
+Number of rows and columns:
+
+```{r}
+dim(nfl_players)
+```
+
+### Number of Elements {#sec-dataCharacteristicsLength}
+
+```{r}
+length(nfl_players$display_name)
+```
+
+### Number of Missing Elements {#sec-dataCharacteristicsMissingElements}
+
+```{r}
+length(nfl_players$college_name[which(is.na(nfl_players$college_name))])
+```
+
+### Number of Non-Missing Elements {#sec-dataCharacteristicsNonMissingElements}
+
+```{r}
+length(nfl_players$college_name[which(!is.na(nfl_players$college_name))])
+length(na.omit(nfl_players$college_name))
+```
+
+## Create New Variables {#sec-createNewVars}
+
+To create a new variable, use the following syntax:
+
+```{r}
+nfl_players$newVar <- NA
+```
+
+Here is an example of creating a new variable:
+
+```{r}
+nfl_players$newVar <- 1:nrow(nfl_players)
+```
+
+```{r}
+#| include: false
+
+nfl_players$newVar <- NULL
+```
+
+## Create a Data Frame {#sec-createDF}
+
+Here is an example of creating a data frame:
+
+```{r}
+mydata <- data.frame(
+  ID = c(1:5, 1047:1051),
+  cat = sample(
+    0:1,
+    10,
+    replace = TRUE)
+)
+
+mydata
+```
+
+## Recode Variables {#sec-recodeVars}
+
+Here is an example of recoding a variable:
+
+```{r}
+mydata$oldVar1 <- NA
+mydata$oldVar1[which(mydata$cat == 0)] <- "dog"
+mydata$oldVar1[which(mydata$cat == 1)] <- "cat"
+
+mydata$oldVar2 <- NA
+mydata$oldVar2[which(mydata$cat == 0)] <- "no"
+mydata$oldVar2[which(mydata$cat == 1)] <- "yes"
+```
+
+Recode multiple variables:
+
+```{r}
+mydata %>%
+  mutate(across(c(
+    oldVar1:oldVar2),
+    ~ case_match(
+      .,
+      c("dog","no") ~ 0,
+      c("cat","yes") ~ 1)))
+```
+
+## Rename Variables {#renameVars}
+
+```{r}
+mydata <- mydata %>% 
+  rename(
+    newVar1 = oldVar1,
+    newVar2 = oldVar2)
+```
+
+Using a vector of variable names:
+
+```{r}
+#| eval: false
+
+varNamesFrom <- c("oldVar1","oldVar2")
+varNamesTo <- c("newVar1","newVar2")
+
+mydata <- mydata %>% 
+  rename_with(~ varNamesTo, all_of(varNamesFrom))
+```
+
+## Convert the Types of Variables {#convertVarTypes}
+
+One variable:
+
+```{r}
+mydata$factorVar <- factor(mydata$ID)
+mydata$numericVar <- as.numeric(mydata$cat)
+mydata$integerVar <- as.integer(mydata$cat)
+mydata$characterVar <- as.character(mydata$newVar1)
+```
+
+Multiple variables:
+
+```{r}
+mydata %>%
+  mutate(across(c(
+    ID,
+    cat),
+    as.numeric))
+
+mydata %>%
+  mutate(across(
+    ID:cat,
+    as.numeric))
+
+mydata %>%
+  mutate(across(where(is.factor), as.character))
 ```

### 2024-05-26 getting-started.qmd
20240526 - join types

@@
 
+## Merging/Joins {#sec-merging}
+
+### Overview {#sec-mergingOverview}
+
+Merging (also called joining) merges two data objects using a shared set of variables called "keys."
+The keys are the variable(s) that uniquely identify each row (i.e., they account for the levels of nesting).
+In some data objects, the key might be the participant's ID (e.g., `participantID`).
+However, some data objects have multiple keys.
+For instance, in long form data objects, each participant may have multiple rows corresponding to multiple timepoints.
+In this case, the keys are `participantID` and `timepoint`.
+If a participant has multiple rows corresponding to timepoints and measures, the keys are `participantID`, `timepoint`, and `measure`.
+In general, each row should have a value on each of the keys; there should be no missingness in the keys.
+
+To merge two objects, the keys must be present in both objects.
+The keys are used to merge the variables in object 1 (`x`) with the variables in object 2 (`y`).
+Different merge types select different rows to merge.
+
+Note: if the two objects include variables with the same name (apart from the keys), `R` will not know how you want each to appear in the merged object.
+So, it will add a suffix (e.g., `.x`, `.y`) to each common variable to indicate which object (i.e., object `x` or object `y`) the variable came from, where object `x` is the first object—i.e., the object to which object `y` (the second object) is merged.
+In general, apart from the keys, you should not include variables with the same name in two objects to be merged.
+To prevent this, either remove or rename the shared variable in one of the objects, or include the shared variable as a key.
+However, as described above, you should include it as a key ***only*** if it uniquely identifies each row in terms of levels of nesting.
+
+### Data Before Merging {#sec-mergingData}
+
+Here are the data in the `nfl_actualStats_offense_weekly` object:
+
+```{r}
+nfl_actualStats_offense_weekly
+
+dim(nfl_actualStats_offense_weekly)
+
+dim(distinct(nfl_actualStats_offense_weekly, player_id, season, week, .keep_all = TRUE))
+
+str(nfl_actualStats_offense_weekly[,c("player_id", "season", "week")])
+```
+
+The data are structured in player-season-week form.
+That is, every row in the dataset is uniquely identified by the variables, `player_id`, `season`, and `week`.
+
+Here are the data in the `nfl_expectedFantasyPoints_weekly` object:
+
+```{r}
+nfl_players
+
+dim(nfl_players)
+
+dim(distinct(nfl_players, gsis_id, .keep_all = TRUE))
+```
+
+### Types of Joins {#sec-mergeTypes}
+
+#### Visual Overview of Join Types {#sec-mergeTypesVisual}
+
+Below is a visual that depicts various types of merges/joins.
+Object `x` is the circle labeled as `x`.
+Object `y` is the circle labeled as `y`.
+The area of overlap in the Venn diagram indicates the rows on the keys that are shared between the two objects (e.g., the same `player_id`, `season`, and `week`).
+The non-overlapping area indicates the rows on the keys that are unique to each object.
+The shaded blue area indicates which rows (on the keys) are kept in the merged object from each of the two objects, when using each of the merge types.
+For instance, a left outer join keeps the shared rows and the rows that are unique to object `x`, but it drops the rows that are unique to object `y`.
+
+![Types of merges/joins](images/joinTypes.png)
+
+#### Full Outer Join {#sec-fullJoin}
+
+A full outer join includes all rows in $x$ **or** $y$.
+It returns columns from $x$ and $y$.
+Here is how to merge two data frames using a full outer join (i.e., "full join"):
+
+```{r}
+fullJoinData <- full_join(
+  nfl_actualStats_offense_weekly,
+  nfl_players,
+  by = c("player_id" = "gsis_id"))
+
+fullJoinData
+dim(fullJoinData)
+```
+
+#### Left Outer Join {#sec-leftJoin}
+
+A left outer join includes all rows in $x$.
+It returns columns from $x$ and $y$.
+Here is how to merge two data frames using a left outer join ("left join"):
+
+```{r}
+leftJoinData <- left_join(
+  nfl_actualStats_offense_weekly,
+  nfl_players,
+  by = c("player_id" = "gsis_id"))
+
+leftJoinData
+dim(leftJoinData)
+```
+
+#### Right Outer Join {#sec-rightJoin}
+
+A right outer join includes all rows in $y$.
+It returns columns from $x$ and $y$.
+Here is how to merge two data frames using a right outer join ("right join"):
+
+```{r}
+rightJoinData <- right_join(
+  nfl_actualStats_offense_weekly,
+  nfl_players,
+  by = c("player_id" = "gsis_id"))
+
+rightJoinData
+dim(rightJoinData)
+```
+
+#### Inner Join {#sec-innerJoin}
+
+An inner join includes all rows that are in **both** $x$ **and** $y$.
+An inner join will return one row of $x$ for each matching row of $y$, and can duplicate values of records on either side (left or right) if $x$ and $y$ have more than one matching record.
+It returns columns from $x$ and $y$.
+Here is how to merge two data frames using an inner join:
+
+```{r}
+innerJoinData <- inner_join(
+  nfl_actualStats_offense_weekly,
+  nfl_players,
+  by = c("player_id" = "gsis_id"))
+
+innerJoinData
+dim(innerJoinData)
+```
+
+#### Semi Join {#sec-semiJoin}
+
+A semi join is a filter.
+A left semi join returns all rows from $x$ **with** a match in $y$.
+That is, it filters out records from $x$ that are not in $y$.
+Unlike an [inner join](#innerJoin), a left semi join will never duplicate rows of $x$, and it includes columns from only $x$ (not from $y$). 
+Here is how to merge two data frames using a left semi join:
+
+```{r}
+semiJoinData <- semi_join(
+  nfl_actualStats_offense_weekly,
+  nfl_players,
+  by = c("player_id" = "gsis_id"))
+
+semiJoinData
+dim(semiJoinData)
+```
+
+#### Anti Join {#sec-antiJoin}
+
+An anti join is a filter.
+A left anti join returns all rows from $x$ **without** a match in $y$.
+That is, it filters out records from $x$ that are in $y$.
+It returns columns from only $x$ (not from $y$).
+Here is how to merge two data frames using a left anti join:
+
+```{r}
+antiJoinData <- anti_join(
+  nfl_actualStats_offense_weekly,
+  nfl_players,
+  by = c("player_id" = "gsis_id"))
+
+antiJoinData
+dim(antiJoinData)
+```
+
+#### Cross Join {#sec-crossJoin}
+
+A cross join combines each row in $x$ with each row in $y$.
+
+```{r}
+crossJoinData <- cross_join(
+  unique(nfl_actualStats_offense_weekly[,c("player_id","player_display_name")]),
+  data.frame(season = 2020:2024))
+
+crossJoinData
+dim(crossJoinData)
+```
+
 ## Calculations {#sec-calculations}

### 2024-05-26 getting-started.qmd
20240526 - clean up
20240526 - long to wide; wide to long

@@
 dim(distinct(nfl_actualStats_offense_weekly, player_id, season, week, .keep_all = TRUE))
-
-str(nfl_actualStats_offense_weekly[,c("player_id", "season", "week")])
 ```
@@
 dim(crossJoinData)
+```
+
+## Transform Data from Long to Wide {#sec-longToWide}
+
+Original data:
+
+```{r}
+dataLong <- nfl_actualStats_offense_weekly %>% 
+  select(player_id, player_display_name, season, week, fantasy_points)
+
+dataLong
+```
+
+Data widened by two variable (`season` and `week`), using `tidyverse`:
+
+```{r}
+dataLong %>% 
+  pivot_wider(
+    names_from = week,
+    names_glue = "{.value}_week{week}",
+    values_from = fantasy_points)
+
+dataLong %>% 
+  pivot_wider(
+    names_from = c(season, week),
+    names_glue = "{.value}_{season}_week{week}",
+    values_from = fantasy_points)
+```
+
+## Transform Data from Wide to Long {#sec-wideToLong}
+
+Original data:
+
+```{r}
+dataWide <- nfl_actualStats_offense_weekly %>% 
+  select(player_id, player_display_name, season, week, recent_team, opponent_team)
+
+dataWide
+```
+
+Data in long form, transformed from wide form using `tidyverse`:
+
+```{r}
+dataWide %>% 
+  pivot_longer(
+    cols = c(recent_team, opponent_team),
+    names_to = "role",
+    values_to = "team")
 ```

### 2024-05-26 getting-started.qmd
20240526 - 4th down and risk aversion

@@
 library("nflfastR")
+library("nfl4th")
 library("nflplotR")
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
+```
+
+### 4th Down Data {#sec-download4thDown}
+
+```{r}
+nfl_4thdown <- nfl4th::load_4th_pbp(seasons = 2014:2023)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_4thdown,
+  file = here("data", "nfl_4thdown.RData")
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
 ```

### 2024-05-26 sports-cognitive-psychology.qmd
20240526 - 4th down and risk aversion

@@
 # Sports and Cognitive Psychology {#sec-sportsCognitivePsychology}
+
+## Load Packages {#sec-loadPackages}
+
+```{r}
+library("nflreadr")
+library("tidyverse")
+```
+
+## Download Football Data {#sec-downloadFootballData}
+
+```{r}
+#| eval: false
+#| include: false
+
+library("here")
+
+# Downloaded Data
+load(file = here("data", "nfl_players.RData"))
+load(file = here("data", "nfl_teams.RData"))
+load(file = here("data", "nfl_rosters.RData"))
+load(file = here("data", "nfl_rosters_weekly.RData"))
+load(file = here("data", "nfl_schedules.RData"))
+load(file = here("data", "nfl_combine.RData"))
+load(file = here("data", "nfl_draftPicks.RData"))
+load(file = here("data", "nfl_depthCharts.RData"))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
+load(file = here("data", "nfl_actualStats_weekly.RData"))
+load(file = here("data", "nfl_injuries.RData"))
+load(file = here("data", "nfl_snapCounts.RData"))
+load(file = here("data", "nfl_espnQBR.RData"))
+load(file = here("data", "nfl_nextGenStats_weekly.RData"))
+load(file = here("data", "nfl_advancedStatsPFR.RData"))
+load(file = here("data", "nfl_playerContracts.RData"))
+load(file = here("data", "nfl_ftnCharting.RData"))
+load(file = here("data", "nfl_playerIDs.RData"))
+load(file = here("data", "nfl_rankings.RData"))
+load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))
+load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
+
+# Calculated Data
+load(file = here("data", "nfl_actualStats_career.RData"))
+load(file = here("data", "nfl_actualStats_seasonal.RData"))
+```
+
+## Coaching and Risk Aversion {#sec-coachingRiskAversion}
+
+```{r}
+nfl_pbp4thDown <- nfl_pbp %>% 
+  filter(down == 4) %>% 
+  filter(!(play_type %in% c("no_play","qb_kneel")))
+
+nfl_pbp4thDown$goForIt <- NA
+nfl_pbp4thDown$goForIt[which(nfl_pbp4thDown$play_type %in% c("field_goal","punt"))] <- 0
+nfl_pbp4thDown$goForIt[which(nfl_pbp4thDown$play_type %in% c("pass","run"))] <- 1
+
+nfl_pbp4thDownPlotData <- nfl_pbp4thDown %>% 
+  filter(!is.na(goForIt)) %>% 
+  group_by(season) %>% 
+  summarise(
+    goForItPct = mean(goForIt, na.rm = TRUE),
+    n = n(),
+    sd = sd(goForIt),
+    se = sd / n
+  )
+
+ggplot2::ggplot(
+  data = nfl_pbp4thDownPlotData,
+  ggplot2::aes(
+    x = season,
+    y = goForItPct)) +
+  geom_point() +
+  geom_line() +
+  geom_ribbon(
+    aes(
+      y = goForItPct,
+      ymin = goForItPct - qnorm(0.975)*se,
+      ymax = goForItPct + qnorm(0.975)*se),
+    alpha = 0.2) +
+  scale_y_continuous(
+    limits = c(0, NA)
+  ) +
+  ggplot2::labs(
+    x = "Season",
+    y = "Proportion of 4th Down Plays that are Attempts (to Get the First Down)",
+    title = "4th Down Attempts (Proportion) by Season",
+  ) +
+  ggplot2::theme_classic()
+```

### 2024-05-26 getting-started.qmd
20240526 - comment out 4th down data
20240526 - shrink file size

@@
 ```{r}
+#| eval: false
+
 nfl_4thdown <- nfl4th::load_4th_pbp(seasons = 2014:2023)
@@
 ```{r}
+#| eval: false
+
 subsetVars <- c("status","display_name","college_name")
@@
 ```{r}
+#| eval: false
+
 dropVars <- c("status","college_name")
@@
 ```{r}
+#| eval: false
+
 nfl_players %>%
@@
 ```{r}
+#| eval: false
+
 subsetRows <- c(1,3,5)
@@
 ```{r}
+#| eval: false
+
 nfl_players[c(1,3,5), c(1,2,3)]
@@
 ```{r}
-nfl_actualStats_offense_weekly
-
 dim(nfl_actualStats_offense_weekly)
@@
 ```{r}
-nfl_players
-
 dim(nfl_players)
@@
 ```{r}
+#| eval: false
+
 fullJoinData <- full_join(
@@
 
-fullJoinData
 dim(fullJoinData)
@@
 ```{r}
+#| eval: false
+
 leftJoinData <- left_join(
@@
 
-leftJoinData
 dim(leftJoinData)
@@
 ```{r}
+#| eval: false
+
 rightJoinData <- right_join(
@@
 
-rightJoinData
 dim(rightJoinData)
@@
 ```{r}
+#| eval: false
+
 innerJoinData <- inner_join(
@@
 
-innerJoinData
 dim(innerJoinData)
@@
 ```{r}
+#| eval: false
+
 semiJoinData <- semi_join(
@@
 
-semiJoinData
 dim(semiJoinData)
@@
 ```{r}
+#| eval: false
+
 antiJoinData <- anti_join(
@@
 
-antiJoinData
 dim(antiJoinData)
@@
 ```{r}
+#| eval: false
+
 crossJoinData <- cross_join(
@@
 ```{r}
+#| eval: false
+
 dataLong <- nfl_actualStats_offense_weekly %>% 
@@
 
-dataLong
+dim(dataLong)
 ```
@@
 ```{r}
-dataLong %>% 
-  pivot_wider(
-    names_from = week,
-    names_glue = "{.value}_week{week}",
-    values_from = fantasy_points)
-
-dataLong %>% 
+#| eval: false
+
+dataWide <- dataLong %>% 
   pivot_wider(
@@
     values_from = fantasy_points)
+
+dim(dataWide)
 ```
@@
 ```{r}
+#| eval: false
+
 dataWide <- nfl_actualStats_offense_weekly %>% 
@@
 
-dataWide
+dim(dataWide)
 ```
@@
 ```{r}
-dataWide %>% 
+#| eval: false
+
+dataLong <- dataWide %>% 
   pivot_longer(
@@
     values_to = "team")
+
+dim(dataLong)
 ```

### 2024-05-26 sports-cognitive-psychology.qmd
20240526 - coaching risk aversion

@@
 ```
+
+Adapted from Ben Baldwin: https://www.nfl4th.com/articles/4th-down-research.html
+
+```{r}
+# labels on the plot
+text_df <- tibble(
+  label = c("NFL coaches<br>in <span style='color:#00BFC4'>**2023**</span>", "NFL coaches<br>in <span style='color:#F8766D'>**2014**</span>"),
+  x = c(6, 8.2),
+  y = c(80, 37),
+  angle = c(10, 10),
+  color = c("black", "black")
+)
+
+nfl_4thdown %>%
+  filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2014, 2023)) %>%
+  ggplot(aes(go_boost, go, color = as.factor(season))) + 
+  geom_richtext(data = text_df,   
+                aes(x, y, label = label, angle = angle), 
+                color = "black", fill = NA, label.color = NA, size = 5) + 
+  geom_vline(xintercept = 0) +
+  stat_smooth(method = "gam", method.args = list(gamma = 1), formula = y ~ s(x, bs = "cs", k = 10), show.legend = F, se = F, linewidth = 4) +
+  # this is just to get the plot to draw the full 0 to 100 range
+  geom_hline(yintercept = 100, alpha = 0) +
+  geom_hline(yintercept = 0, alpha = 0) +
+  ggplot2::theme_classic() +
+  labs(x = "Gain in win probability by going for it",
+       y = "Go-for-it percentage",
+       subtitle = "4th down decisions in 2023 versus 2014, win prob. > 20%",
+       title = glue::glue("How <span style='color:red'>math</span> is changing football")) +
+  theme(
+    legend.position = "none",
+    plot.title = element_markdown(size = 22, hjust = 0.5),
+    plot.subtitle = element_markdown(size = 14, hjust = 0.5),
+    axis.title.x = element_text(size = 14, face="bold"),
+    axis.title.y = element_text(size = 14, face="bold")
+  ) +
+  scale_y_continuous(breaks = scales::pretty_breaks(n = 4), limits = c(0, 100), expand = c(0,0)) +
+  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(-10, 10), expand = c(0,0)) +
+  annotate("text", x= -1.2, y= 70, label = "Should\nkick", color="black", size = 5) +
+  annotate("text", x= 1.2, y= 70, label = "Should\ngo for it", color="black", size = 5) +
+  geom_segment(
+    aes(x = -.1, y = 80, xend = -2, yend = 80),
+    arrow = arrow(length = unit(0.05, "npc")),
+    color = "black", size = 2
+    ) +
+  geom_segment(
+    aes(x = .1, y = 80, xend = 2, yend = 80),
+    arrow = arrow(length = unit(0.05, "npc")),
+    color = "black", size = 2
+  )
+```

### 2024-05-26 mythbusters.qmd
20240526 - contract year

@@
 # Mythbusters: Putting Fantasy Football Beliefs/Anecdotes to the Test {#sec-mythbusters}
+
+## Do Players Score More Points in their Contract Year? {#sec-contractYear}
+
+```{r}
+nfl_playerContracts_subset %>% 
+  filter(player == "Aaron Rodgers") %>% 
+  select(player:years, contractYear, next_contract_start, played_in_contract_year)
+
+nfl_playerContracts_subset %>% 
+  filter(player %in% c("Jared Allen", "Aaron Rodgers")) %>% 
+  select(player:years, contractYear, next_contract_start, played_in_contract_year)
+
+# Subset to remove players without a year signed
+nfl_playerContracts_subset <- nfl_playerContracts %>% 
+  filter(!is.na(year_signed) & year_signed != 0)
+
+# Determine the contract year for a given contract
+nfl_playerContracts_subset$contractYear <- nfl_playerContracts_subset$year_signed + nfl_playerContracts_subset$years - 1
+
+# Arrange contracts by player and year_signed
+nfl_playerContracts_subset <- nfl_playerContracts_subset %>%
+  group_by(player, position) %>% 
+  arrange(player, position, -year_signed) %>% 
+  ungroup()
+
+# Determine if the player played in the original contract year
+nfl_playerContracts_subset <- nfl_playerContracts_subset %>%
+  group_by(player, position) %>%
+  mutate(
+    next_contract_start = lag(year_signed)) %>%
+  ungroup() %>%
+  mutate(
+    played_in_contract_year = ifelse(
+      is.na(next_contract_start) | contractYear < next_contract_start,
+      TRUE,
+      FALSE))
+
+# Merge with seasonal fantasy points data
+```

### 2024-05-27 getting-started.qmd
20240526 - placeholder
20240526 - load ffanalytics package

@@
 ```{r}
+library("ffanalytics")
 library("nflreadr")
@@
 
+#### Weekly {#sec-calculateActualPointsWeekly}
+
+#### Seasonal {#sec-calculateActualPointsWeekly}
+
+#### Career {#sec-calculateActualPointsWeekly}
+
 ### Player Age {#sec-calculatePlayerAge}

### 2024-05-27 fantasy-football.qmd
20240521 - remove git lfs
20240527 - audience and number of players

@@
 
-### The Objective {#sec-football}
+Football is the most widely watched sport in the United States.^[https://news.gallup.com/poll/610046/football-retains-dominant-position-favorite-sport.aspx (archived at https://perma.cc/X2UG-RAAK); https://www.statista.com/statistics/1430289/most-watched-sports-leagues-usa/ (archived at https://perma.cc/JNU6-S96A)]
+
+### The Objective {#sec-footballObjective}
 
@@
 
+Fantasy football is one of the most widely played games in the history of games.
+It is estimated that around 29–60 million people play fantasy football.^[https://thefsga.org/industry-demographics/ (archived at https://perma.cc/9PB8-ZDJJ); https://www.statista.com/topics/10895/fantasy-sports-in-the-us/ (archived at https://perma.cc/8YSN-UUNT)]
 As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available for [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]

### 2024-05-27 fantasy-football.qmd
20240527 - links

--- 2024-05-19 fantasy-football.qmd
@@
 
-### The Objective {#sec-football}
+Football is the most widely watched sport in the United States.^[<https://news.gallup.com/poll/610046/football-retains-dominant-position-favorite-sport.aspx> (archived at <https://perma.cc/X2UG-RAAK>); <https://www.statista.com/statistics/1430289/most-watched-sports-leagues-usa/> (archived at <https://perma.cc/JNU6-S96A>)]
+
+### The Objective {#sec-footballObjective}
 
@@
 
+Fantasy football is one of the most widely played games in the history of games.
+It is estimated that around 29–60 million people play fantasy football.^[<https://thefsga.org/industry-demographics/> (archived at <https://perma.cc/9PB8-ZDJJ>); <https://www.statista.com/topics/10895/fantasy-sports-in-the-us/> (archived at <https://perma.cc/8YSN-UUNT>)]
 As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available for [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]

### 2024-05-27 fantasy-football.qmd
20240527 - update numbers

@@
 Fantasy football is one of the most widely played games in the history of games.
-It is estimated that around 29–60 million people play fantasy football.^[<https://thefsga.org/industry-demographics/> (archived at <https://perma.cc/9PB8-ZDJJ>); <https://www.statista.com/topics/10895/fantasy-sports-in-the-us/> (archived at <https://perma.cc/8YSN-UUNT>)]
+It is estimated that around 29–62 million people play fantasy football.^[<https://thefsga.org/industry-demographics/> (archived at <https://perma.cc/9PB8-ZDJJ>); <https://www.statista.com/topics/10895/fantasy-sports-in-the-us/> (archived at <https://perma.cc/8YSN-UUNT>)]
 As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available for [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]

### 2024-05-27 fantasy-football.qmd
20240527 - numbers for fantasy sports vs fantasy football

@@
 Fantasy football is one of the most widely played games in the history of games.
-It is estimated that around 29–62 million people play fantasy football.^[<https://thefsga.org/industry-demographics/> (archived at <https://perma.cc/9PB8-ZDJJ>); <https://www.statista.com/topics/10895/fantasy-sports-in-the-us/> (archived at <https://perma.cc/8YSN-UUNT>)]
+It is estimated that around 62 million people play fantasy sports^[<https://thefsga.org/industry-demographics/> (archived at <https://perma.cc/9PB8-ZDJJ>)], of whom around 29 million play fantasy football.^[<https://www.statista.com/topics/10895/fantasy-sports-in-the-us/> (archived at <https://perma.cc/8YSN-UUNT>)]
 As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available for [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]

### 2024-05-28 fantasy-football.qmd
20240527 - glossary

@@
 - Running Back (RB)
-    - Halfback or Tailback
+    - Halfback (HB) or Tailback (TB)
     - Fullback (FB)
@@
 - Tight End (TE)
-- Offensive Linemen (OL)
+- Offensive Linemen (OL), part of the "Offensive Line"
     - Center (C)
@@
 
-Quarterbacks are the most important player on the offense.
+The quarterback is the most important player on the offense.
 They help lead the team down the field.
-Quarterbacks receive the ball from the Center at the beginning of the play, and they can either hand the ball off (typically to a Running Back or Fullback), pass the ball (typically to a Wide Receiver or Tight End), or run the ball.
+The quarterback receives the ball from the Center at the beginning of the play, and they can either hand the ball off (typically to a Running Back or Fullback), pass the ball (typically to a Wide Receiver or Tight End), or run the ball.
 Quarterbacks tend to have a strong arm for throwing the ball far and accurately.
@@
 
-Running Backs take a hand-off from the Quarterback to execute a running play.
+Running Backs take a hand-off from the Quarterback to execute a running play (i.e., a rush).
 They may also catch short passes from the Quarterback or help protect (i.e., block for) the Quarterback from the defensive players who are trying to tackle the Quarterback.
-Halfbacks or Tailbacks tend to be quick and agile.
+Halfbacks and Tailbacks tend to be quick and agile.
 Fullbacks tend to be strong and powerful.
@@
 
-- Defensive Linemen (DL)
+- Defensive Linemen (DL), part of the "Defensive Line"
     - Defensive End (DE)
@@
     - Outside Linebacker (OLB)
-- "The Secondary" / Defensive Back (DB)
+- Defensive Back (DB), part of the "Secondary"
     - Cornerback (CB)
@@
 
-On passing plays, Defensive Linemen try to apply pressure to the Quarterback and try to sack them.
-On rushing plays, Defensive Linemen try to tackle the ball carrier.
+On passing plays, Defensive Linemen try to apply pressure to the Quarterback and try to tackle the Quarterback behind the line of scrimmage before the Quarterback can throw the ball (i.e., a sack).
+On rushing plays, Defensive Linemen try to tackle the ball carrier to prevent the ball carrier from advancing the ball (i.e., gaining yards).
 Defensive Linemen tend to be large yet quick so they can apply pressure to the Quarterback.
@@
 
-The football field is rectangular and is 120 yards long and 53 1/3 yards wide.
+The football field is rectangular and is 120 yards long and 53 1/3 yards wide (109.73 m x 48.77 m).^[One yard is equal to three feet. A yard is just smaller than a meter (0.9144 meters).]
 At each end of the 120-yard field is a team's end zone.
-Each end zone is 10 yards long.
-Thus, the distance from one end zone to the other end zone is 100 yards.
+Each end zone is 10 yards long (9.14 m).
+Thus, the distance from one end zone to the other end zone is 100 yards (91.44 m).
 Behind each end zone is a field goal post.
@@
 
-The team with the ball (i.e., the offense) has four opportunities ("downs") to advance the ball 10 yards.
-A team can advance the ball either by running it or passing and catching it.
+The team with the ball (i.e., the offense) has four opportunities ("downs") to advance the ball (i.e., gain) 10 yards.
+A team can advance the ball either by running it or by throwing (i.e., passing) and catching it.
 At the end of a rushing play, the ball advances to wherever the ball carrier is tackled or goes out of bounds (i.e., wherever the player is "down").
@@
 After scoring a touchdown, field goal, or safety, there is a kickoff, in which the scoring team kicks the ball to the opponent.
-Another way that the ball swithces possession to the other team is if the team commits a turnover.
+Another way that the ball switches possession to the other team is if the team commits a turnover.
 The defense can force a turnover by an interception, fumble recovery, or turnover on downs.
-A turnover due to an interception occurs when the defense catches the the quarterback's pass.
+A turnover due to an interception occurs when a defensive player catches the quarterback's pass.
 A turnover due to a fumble recovery occurs when an offensive player, who had possession of the ball, loses the ball before being down or scoring a touchdown and the ball is recovered by the opponent.
-A turnover on downs occurs when the team attempts to achieve the remainder of the needed 10 yards to go on fourth down but fails.
+A turnover on downs occurs when the team attempts on fourth down to achieve the remainder of the needed 10 yards to go but fails.
 
@@
 After a touchdown, the offense chooses to attempt either a point-after-touchdown (PAT) or a two-point conversion.
-A PAT is a short kick attempt from the 15-yard line that, if it goes through the goal posts ("uprights"), is worth 1 point.
+A PAT is a short kick attempt from the 15-yard line (i.e., 15 yards away from the end zone) that, if it goes through the goal posts ("uprights") and over the cross bar, is worth 1 point.
 A two-point conversion is a single-scoring opportunity from the 3-yard line (i.e., 3 yards away from the end zone).
@@
 A safety occurs when the offense is tackled with the ball in their own end zone.
-When a safety occurs, the opposing team (i.e., defense) is awarded two points.
+When a safety occurs, the opposing team (i.e., defense) is awarded two points and the ball.
 
@@
 
-- running play ("run") or rushing play (or "rush")
-- passing play (or "pass")
-- passing attempt
-- rushing attempt
-- passing completion
-- passing incompletion
-- passing yards
-- rushing yards
-- receving yards
-- reception
-- touchdown
-- passing touchdown
-- rushing touchdown
-- receiving touchdown
-- two-point conversion
-- block
-- kickoff
-- field goal
-- point after touchdown (PAT)
-- extra point returned
-- punt
-- fumble lost
-- fumble forced
-- fumble recovery
-- interception
-- tackle
-- tackle solo
-- tackle assist
-- tackle for loss
-- sack
-- pass defended
-- safety
+- running play ("run") or rushing play (or "rush")—the attempt by an offensive player, typically the Running Back or Quarterback, to advance the ball "on the ground" by running it—not by passing it forward
+- passing play (or "pass")—the attempt by an offensive player, typically the Quarterback, to advance the ball by throwing it forward to an offensive player
+- passing attempt—the attempt to advance the ball by passing it (i.e., a thrown pass)
+- rushing attempt—the attempt to advance the ball by running it
+- passing completion—a thrown pass that is succesfully caught by an offensive player
+- passing incompletion—a thrown pass that is not caught by an offensive player
+- passing yards—the distance (in yards) the player advanced the ball by throwing it
+- rushing yards—the distance (in yards) the player advanced the ball by running it
+- receving yards—the distance (in yards) the player advanced the ball by catching thrown passes and then running with it further upfield
+- kick/punt return yards—the distance (in yards) the player advanced the ball by returning kicks or punts
+- turnover return yards—the distance (in yards) the player advanced the ball by returning turnovers
+- reception—a pass that is caught by the offensive player
+- touchdown—advancing the ball into the opponent's end zone either by a) throwing a completed pass that ends up in the end zone, b) running it into the end zone, c) catching it in the end zone, or d) catching it and then running it into the end zone
+- passing touchdown—advancing the ball into the opponent's end zone either by throwing a completed pass that ends up in the end zone
+- rushing touchdown—advancing the ball into the opponent's end zone either by running it into the end zone
+- receiving touchdown—advancing the ball into the opponent's end zone either by catching it in the end zone or by catching it and then running it into the end zone
+- kick/punt return touchdown—advancing the ball into the opponent's end zone when returning a kick or punt
+- turnover return touchdown—advancing the ball into the opponent's end zone when returning a turnover (i.e., interception or fumble)
+- two-point conversion—a single-scoring opportunity from the 3-yard line (i.e., 3 yards away from the end zone) that is an option given to a team that scores a touchdown;
+if the offense scores (i.e., advances the ball into the end zone) from the 3-yard line, the team is awarded 2 points
+- block—when the defense/special teams blocks a kick or field goal by hitting the ball just after it is kicked to prevent the ball from going far
+- kickoff—the kicking team kicks the ball to the receiving team, who has the option to return the kick
+- field goal—a kicker kicks the ball with an intent to kick the ball through the field goal posts ("uprights").
+To score points by making a field goal, the kicked ball must go between the uprights (extended vertically) and over the cross bar.
+If the field goal attempt is successful, the team gains 3 points.
+- point after touchdown (PAT)—a short kick attempt from the 15-yard line (i.e., 15 yards away from the end zone) that, if it goes through the goal posts ("uprights") and over the cross bar, is worth 1 point
+- extra point returned—if the defense/special teams returns the ball into the opponent's end zone during a point after touchdown (PAT) attempt, it is worth 2 points
+- punt—a punter kicks the ball to the other team with an intent to give their opponent worse field position, thus making it harder for the other team to score
+- fumble lost—when an offensive player, who had possession of the ball, loses the ball before being down or scoring a touchdown and the ball is recovered by the opponent
+- fumble forced—when a defensive player knocks the ball out of the hands of an offensive player, who had possession of the ball
+- fumble recovery—when a defensive player recovers a fumble by the opponent
+- interception—when a defensive player catches a pass from an offensive player
+- tackle—when a player brings down the ball carrier
+- tackle solo—when a player is the main tackler (i.e., the primary player to bring down the ball carrier)
+- tackle assist—when a player is one of two or more players who, together, bring down the ball carrier
+- tackle for loss—when an offensive player is tackled with the ball behind the line of scrimmage
+- sack—when a Quarterback is tackled with the ball behind the line of scrimmage
+- pass defended—when a defensive player knocks down the ball in the air so that the indended receiver cannot catch the ball
+- safety—when the offense is tackled with the ball in their own end zone
 

### 2024-05-29 getting-started.qmd
20240526 - placeholder
20240527 - formatting
20240528 - better data example

@@
-# Getting Started with R for Data Analysis {#sec-gettingStarted}
+# Getting Started with `R` for Data Analysis {#sec-gettingStarted}
 
@@
 
-You can install R packages using the following syntax:
+You can install `R` packages using the following syntax:
 
@@
 
+## Create a Data Frame {#sec-createDF}
+
+Here is an example of creating a data frame:
+
+```{r}
+players <- data.frame(
+  ID = 1:10,
+  name = c(
+    "Ken Cussion",
+    "Ben Sacked",
+    "Justin Time",
+    "Lionel Scrimmage",
+    "Drew Blood",
+    "Hugo Long",
+    "Chase Emdown",
+    "Name8",
+    "Name9",
+    "Name10"),
+  position = c("QB","QB","RB","RB","WR","WR","WR","WR","TE","TE"),
+  age = c(40, 30, 20, 18, 23, 27, 32, 26, 23, NA)
+  )
+
+fantasyPoints <- data.frame(
+  ID = c(2, 6, 11, 12),
+  fantasyPoints = c(250, 170, 65, 15)
+)
+```
+
 ## Variable Names {#sec-variableNames}
@@
 names(nfl_players)
+
+names(players)
+
+names(fantasyPoints)
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$position_group == "RB"
+players$position == "RB"
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$position_group != "RB"
+players$position != "RB"
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$weight > 300
+players$age > 30
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$weight < 300
+players$age < 30
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$weight >= 300
+players$age >= 30
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$weight <= 300
+players$age <= 30
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$position_group %in% c("QB","RB","WR")
+players$position %in% c("RB","WR")
 ```
@@
 ```{r}
-#| eval: false
-
-!(nfl_players$position_group %in% c("QB","RB","WR"))
+!(players$position %in% c("RB","WR"))
 ```
@@
 ```{r}
-#| eval: false
-
-is.na(nfl_players$college_name)
+is.na(players$age)
 ```
@@
 ```{r}
-#| eval: false
-
-!is.na(nfl_players$college_name)
+!is.na(players$age)
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$position_group == "RB" & nfl_players$weight > 230
+players$position == "WR" & players$age > 26
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$position_group == "RB" | nfl_players$weight > 230
+players$position == "WR" | players$age > 23
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$display_name
+players$name
 ```
@@
 ```{r}
-nfl_players[,"display_name"]
+players[,"name"]
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players$display_name[which(nfl_players$position_group == "RB")]
+players$name[which(players$position == "RB")]
 ```
@@
 ```{r}
-nfl_players[which(nfl_players$position_group == "RB"), "display_name"]
+players[which(players$position == "RB"), "name"]
 ```
@@
 ```{r}
-#| eval: false
-
-subsetVars <- c("status","display_name","college_name")
-
-nfl_players[,c(1,2,9)]
-nfl_players[,c("status","display_name","college_name")]
-nfl_players[,subsetVars]
+subsetVars <- c("name","age")
+
+players[,c(2,3)]
+players[,c("name","age")]
+players[,subsetVars]
 ```
@@
 ```{r}
-#| eval: false
-
-dropVars <- c("status","college_name")
-
-nfl_players[,-c(1,9)]
-nfl_players[,!(names(nfl_players) %in% c("status","college_name"))]
-nfl_players[,!(names(nfl_players) %in% dropVars)]
+dropVars <- c("name","age")
+
+players[,-c(2,3)]
+players[,!(names(players) %in% c("name","age"))]
+players[,!(names(players) %in% dropVars)]
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players %>%
-  select(status, display_name, college_name)
-
-nfl_players %>%
-  select(status:college_name)
-
-nfl_players %>%
+players %>%
+  select(name, age)
+
+players %>%
+  select(name:age)
+
+players %>%
   select(all_of(subsetVars))
@@
 ```{r}
-#| eval: false
-
-nfl_players %>%
-  select(-status, -college_name)
-
-nfl_players %>%
-  select(-c(status:college_name))
-
-nfl_players %>%
+players %>%
+  select(-name, -age)
+
+players %>%
+  select(-c(position:age))
+
+players %>%
   select(-all_of(dropVars))
@@
 ```{r}
-#| eval: false
-
-subsetRows <- c(1,3,5)
-
-nfl_players[c(1,3,5),]
-nfl_players[subsetRows,]
-nfl_players[which(nfl_players$position_group == "RB"),]
+subsetRows <- c(3,4)
+
+players[c(3,4),]
+players[subsetRows,]
+players[which(players$position == "RB"),]
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players %>%
-  filter(position_group == "RB")
-
-nfl_players %>%
-  filter(position_group == "RB", weight <= 250)
-
-nfl_players %>%
-  filter(position_group == "RB" | weight >= 250)
+players %>%
+  filter(position == "WR")
+
+players %>%
+  filter(position == "WR", age <= 26)
+
+players %>%
+  filter(position == "WR" | age >= 26)
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players[c(1,3,5), c(1,2,3)]
-nfl_players[subsetRows, subsetVars]
-nfl_players[which(nfl_players$position_group == "RB"), subsetVars]
+players[c(3,4), c(2,4)]
+players[subsetRows, subsetVars]
+players[which(players$position == "RB"), subsetVars]
 ```
@@
 ```{r}
-#| eval: false
-
-nfl_players %>%
-  filter(position_group == "RB") %>%
+players %>%
+  filter(position == "RB") %>%
   select(all_of(subsetVars))
@@
 
-View(nfl_players)
+View(players)
 ```
@@
 ```{r}
-nfl_players$newVar <- NA
+players$newVar <- NA
 ```
@@
 ```{r}
-nfl_players$newVar <- 1:nrow(nfl_players)
-```
-
-```{r}
-#| include: false
-
-nfl_players$newVar <- NULL
-```
-
-## Create a Data Frame {#sec-createDF}
-
-Here is an example of creating a data frame:
-
-```{r}
-mydata <- data.frame(
-  ID = c(1:5, 1047:1051),
-  cat = sample(
-    0:1,
-    10,
-    replace = TRUE)
-)
-
-mydata
+players$newVar <- 1:nrow(players)
+```
+
+```{r}
+#| include: false
+
+players$newVar <- NULL
 ```
@@
 ```{r}
-mydata$oldVar1 <- NA
-mydata$oldVar1[which(mydata$cat == 0)] <- "dog"
-mydata$oldVar1[which(mydata$cat == 1)] <- "cat"
-
-mydata$oldVar2 <- NA
-mydata$oldVar2[which(mydata$cat == 0)] <- "no"
-mydata$oldVar2[which(mydata$cat == 1)] <- "yes"
+players$oldVar1 <- NA
+players$oldVar1[which(players$position == "QB")] <- "quarterback"
+players$oldVar1[which(players$position == "RB")] <- "running back"
+players$oldVar1[which(players$position == "WR")] <- "wide receiver"
+players$oldVar1[which(players$position == "TE")] <- "tight end"
+
+players$oldVar2 <- NA
+players$oldVar2[which(players$age < 30)] <- "young"
+players$oldVar2[which(players$age >= 30)] <- "old"
 ```
@@
 ```{r}
-mydata %>%
+players %>%
   mutate(across(c(
@@
       .,
-      c("dog","no") ~ 0,
-      c("cat","yes") ~ 1)))
+      c("quarterback","old","running back") ~ 0,
+      c("wide receiver","tight end","young") ~ 1)))
 ```
@@
 ```{r}
-mydata <- mydata %>% 
+players <- players %>% 
   rename(
@@
 
-mydata <- mydata %>% 
+players <- players %>% 
   rename_with(~ varNamesTo, all_of(varNamesFrom))
@@
 ```{r}
-mydata$factorVar <- factor(mydata$ID)
-mydata$numericVar <- as.numeric(mydata$cat)
-mydata$integerVar <- as.integer(mydata$cat)
-mydata$characterVar <- as.character(mydata$newVar1)
+players$factorVar <- factor(players$ID)
+players$numericVar <- as.numeric(players$age)
+players$integerVar <- as.integer(players$newVar)
+players$characterVar <- as.character(players$newVar1)
 ```
@@
 ```{r}
-mydata %>%
+players %>%
   mutate(across(c(
     ID,
-    cat),
+    age),
     as.numeric))
 
-mydata %>%
+players %>%
   mutate(across(
-    ID:cat,
-    as.numeric))
-
-mydata %>%
+    age:newVar1,
+    as.character))
+
+players %>%
   mutate(across(where(is.factor), as.character))
@@
 
-Here are the data in the `nfl_actualStats_offense_weekly` object:
-
-```{r}
-dim(nfl_actualStats_offense_weekly)
-
-dim(distinct(nfl_actualStats_offense_weekly, player_id, season, week, .keep_all = TRUE))
-```
-
-The data are structured in player-season-week form.
-That is, every row in the dataset is uniquely identified by the variables, `player_id`, `season`, and `week`.
-
-Here are the data in the `nfl_expectedFantasyPoints_weekly` object:
-
-```{r}
-dim(nfl_players)
-
-dim(distinct(nfl_players, gsis_id, .keep_all = TRUE))
+Here are the data in the `players` object:
+
+```{r}
+players
+
+dim(players)
+```
+
+The data are structured in ID form.
+That is, every row in the dataset is uniquely identified by the variable, `ID`.
+
+Here are the data in the `fantasyPoints` object:
+
+```{r}
+fantasyPoints
+
+dim(fantasyPoints)
 ```
@@
 ```{r}
-#| eval: false
-
 fullJoinData <- full_join(
-  nfl_actualStats_offense_weekly,
-  nfl_players,
-  by = c("player_id" = "gsis_id"))
+  players,
+  fantasyPoints,
+  by = "ID")
+
+fullJoinData
 
@@
 ```{r}
-#| eval: false
-
 leftJoinData <- left_join(
-  nfl_actualStats_offense_weekly,
-  nfl_players,
-  by = c("player_id" = "gsis_id"))
+  players,
+  fantasyPoints,
+  by = "ID")
+
+leftJoinData
 
@@
 ```{r}
-#| eval: false
-
 rightJoinData <- right_join(
-  nfl_actualStats_offense_weekly,
-  nfl_players,
-  by = c("player_id" = "gsis_id"))
+  players,
+  fantasyPoints,
+  by = "ID")
+
+rightJoinData
 
@@
 ```{r}
-#| eval: false
-
 innerJoinData <- inner_join(
-  nfl_actualStats_offense_weekly,
-  nfl_players,
-  by = c("player_id" = "gsis_id"))
+  players,
+  fantasyPoints,
+  by = "ID")
+
+innerJoinData
 
@@
 ```{r}
-#| eval: false
-
 semiJoinData <- semi_join(
-  nfl_actualStats_offense_weekly,
-  nfl_players,
-  by = c("player_id" = "gsis_id"))
+  players,
+  fantasyPoints,
+  by = "ID")
+
+semiJoinData
 
@@
 ```{r}
-#| eval: false
-
 antiJoinData <- anti_join(
-  nfl_actualStats_offense_weekly,
-  nfl_players,
-  by = c("player_id" = "gsis_id"))
+  players,
+  fantasyPoints,
+  by = "ID")
+
+antiJoinData
 
@@
 ```{r}
-#| eval: false
-
 crossJoinData <- cross_join(
-  unique(nfl_actualStats_offense_weekly[,c("player_id","player_display_name")]),
-  data.frame(season = 2020:2024))
+  players,
+  fantasyPoints)
 
 crossJoinData
+
 dim(crossJoinData)
@@
 
+Here are the data in the `nfl_actualStats_offense_weekly` object.
+The data are structured in player-season-week form.
+That is, every row in the dataset is uniquely identified by the variables, `player_id`, `season`, and `week`.
+
 Original data:
@@
 ```{r}
-#| eval: false
-
 dataLong <- nfl_actualStats_offense_weekly %>% 
@@
 
-Data widened by two variable (`season` and `week`), using `tidyverse`:
-
-```{r}
-#| eval: false
-
+Below, we widen the data widened by two variables (`season` and `week`), using `tidyverse`, so that the data are now in player form (where each row is uniquely identified by the `player_id` variable):
+
+```{r}
 dataWide <- dataLong %>% 
@@
 ```{r}
-#| eval: false
-
 dataWide <- nfl_actualStats_offense_weekly %>% 
@@
 ```{r}
-#| eval: false
-
 dataLong <- dataWide %>% 
@@
 ### Historical Actual Fantasy Points {#sec-calculateActualPoints}
+
+Specify scoring settings:
 

### 2024-05-29 getting-started.qmd
20240528 - fix error and warning

@@
 players$numericVar <- as.numeric(players$age)
-players$integerVar <- as.integer(players$newVar)
-players$characterVar <- as.character(players$newVar1)
+players$integerVar <- as.integer(players$newVar1)
+players$characterVar <- as.character(players$newVar2)
 ```
@@
 
-#### Seasonal {#sec-calculateActualPointsWeekly}
-
-#### Career {#sec-calculateActualPointsWeekly}
+#### Seasonal {#sec-calculateActualPointsSeasonal}
+
+#### Career {#sec-calculateActualPointsCareer}
 

### 2024-05-29 getting-started.qmd
20240528 - edit for consistency

@@
 
-players[,c(2,3)]
+players[,c(2,4)]
 players[,c("name","age")]
@@
 
-players[,-c(2,3)]
+players[,-c(2,4)]
 players[,!(names(players) %in% c("name","age"))]

### 2024-05-31 getting-started.qmd
20240529 - add variables names
20240530 - names

@@
     "Ben Sacked",
-    "Justin Time",
+    "Ron Ingback",
     "Lionel Scrimmage",
@@
     "Chase Emdown",
-    "Name8",
+    "Justin Time",
     "Name9",
@@
 dim(dataLong)
+
+names(dataLong)
 ```
@@
 dim(dataWide)
+
+names(dataWide)
 ```

### 2024-06-01 getting-started.qmd
20240601 - names

@@
     "Ron Ingback",
+    "Rhonda Ball",
+    "Hugo Long",
     "Lionel Scrimmage",
     "Drew Blood",
-    "Hugo Long",
     "Chase Emdown",
     "Justin Time",
-    "Name9",
     "Name10"),

### 2024-06-01 getting-started.qmd
20240601 - player names

--- 2024-05-31 getting-started.qmd
@@
 players <- data.frame(
-  ID = 1:10,
+  ID = 1:11,
   name = c(
@@
     "Ben Sacked",
+    "Chuck Downfield",
     "Ron Ingback",
+    "Rhonda Ball",
+    "Hugo Long",
     "Lionel Scrimmage",
     "Drew Blood",
-    "Hugo Long",
     "Chase Emdown",
     "Justin Time",
-    "Name9",
-    "Name10"),
-  position = c("QB","QB","RB","RB","WR","WR","WR","WR","TE","TE"),
-  age = c(40, 30, 20, 18, 23, 27, 32, 26, 23, NA)
+    "Spike D'Ball"),
+  position = c("QB","QB","QB","RB","RB","WR","WR","WR","WR","TE","TE"),
+  age = c(40, 30, 24, 20, 18, 23, 27, 32, 26, 23, NA)
   )
@@
 fantasyPoints <- data.frame(
-  ID = c(2, 6, 11, 12),
+  ID = c(2, 7, 12, 13),
   fantasyPoints = c(250, 170, 65, 15)

### 2024-06-01 getting-started.qmd
20240601 - names

--- 2024-05-31 getting-started.qmd
@@
 players <- data.frame(
-  ID = 1:10,
+  ID = 1:12,
   name = c(
@@
     "Ben Sacked",
+    "Chuck Downfield",
     "Ron Ingback",
+    "Rhonda Ball",
+    "Hugo Long",
     "Lionel Scrimmage",
     "Drew Blood",
-    "Hugo Long",
     "Chase Emdown",
     "Justin Time",
-    "Name9",
-    "Name10"),
-  position = c("QB","QB","RB","RB","WR","WR","WR","WR","TE","TE"),
-  age = c(40, 30, 20, 18, 23, 27, 32, 26, 23, NA)
+    "Spike D'Ball",
+    "Isac Ulooz"),
+  position = c("QB","QB","QB","RB","RB","WR","WR","WR","WR","TE","TE","LB"),
+  age = c(40, 30, 24, 20, 18, 23, 27, 32, 26, 23, NA, 37)
   )
@@
 fantasyPoints <- data.frame(
-  ID = c(2, 6, 11, 12),
+  ID = c(2, 7, 13, 14),
   fantasyPoints = c(250, 170, 65, 15)

### 2024-06-01 getting-started.qmd
20240601 - updates

@@
 players %>%
-  select(-c(position:age))
+  select(-c(name:age))
 
@@
 ```{r}
-subsetRows <- c(3,4)
-
-players[c(3,4),]
+subsetRows <- c(4,5)
+
+players[c(4,5),]
 players[subsetRows,]
@@
 ```{r}
-players[c(3,4), c(2,4)]
+players[c(4,5), c(2,4)]
 players[subsetRows, subsetVars]
@@
 The keys are the variable(s) that uniquely identify each row (i.e., they account for the levels of nesting).
-In some data objects, the key might be the participant's ID (e.g., `participantID`).
+In some data objects, the key might be the player's identification number (e.g., `player_id`).
 However, some data objects have multiple keys.
-For instance, in long form data objects, each participant may have multiple rows corresponding to multiple timepoints.
-In this case, the keys are `participantID` and `timepoint`.
-If a participant has multiple rows corresponding to timepoints and measures, the keys are `participantID`, `timepoint`, and `measure`.
+For instance, in long form data objects, each participant may have multiple rows corresponding to multiple seasons.
+In this case, the keys may be `player_id` and `season`.
+If a participant has multiple rows corresponding to seasons and games/weeks, the keys are `player_id`, `season`, and `week`.
 In general, each row should have a value on each of the keys; there should be no missingness in the keys.
 
-To merge two objects, the keys must be present in both objects.
+To merge two objects, the key(s) that will be used to match the records must be present in both objects.
 The keys are used to merge the variables in object 1 (`x`) with the variables in object 2 (`y`).

### 2024-06-01 player-evaluation.qmd
20240601 - player evaluation

# Player Evaluation {#sec-playerEvaluation}

There are various domains of criteria we can consider when evaluating a football player's fantasy prospects.
Potential domains to consider include:

- athletic profile
- historical performance
- age and career stage

# Athletic Profile

- height
- weight
- hand size
- agility
- strength
- speed
- jumping ability
- reaction time

# Historical Performance

- performance in college
    - draft position
- performance in the NFL
- efficiency
- consistency

# Health

- current injury status
- injury history

# Age and Career Stage

- age
- touches

# Situational Factors

Situational factors are influenced by many factors, including skill of the player, skill of their teammates, coaching style, strategy of the opposing team, game scripts, etc.

- team quality
- role on team
- teammates
- opportunity and usage
    - snap count
    - touches/targets
    - red zone usage

# Matchups

- strength of schedule
- weekly matchup

# Cognitive and Motivational Factors

- work ethic and mental toughness
- whether they are in a contract year

### 2024-06-01 player-evaluation.qmd
20240601 - incentives
20240601 - stock picking

@@
 # Player Evaluation {#sec-playerEvaluation}
+
+Evaluating players for fantasy football could be thought of as similar to evaluating companies when buying stocks.
+You want to evaluate and compare various assets so that you get the assets with the best value.
 
@@
 - work ethic and mental toughness
+- incentives
+    - contract performance incentives
     - whether they are in a contract year

### 2024-06-02 player-evaluation.qmd
20240601 - athletic profile and historical performance

--- 2024-06-01 player-evaluation.qmd
@@
 # Player Evaluation {#sec-playerEvaluation}
+
+Evaluating players for fantasy football could be thought of as similar to the process of evaluating companies when picking stocks to buy.
+You want to evaluate and compare various assets so that you get the assets with the best value.
 
@@
 - historical performance
+- health
 - age and career stage
+- situational factors
+- matchups
+- cognitive and motivational factors
 
@@
 
+- body shape
     - height
@@
     - hand size
+- body function
     - agility
@@
     - speed
+    - acceleration/explosiveness
     - jumping ability
-- reaction time
+
+In terms of body shape, we might consider a player's height, weight, and hand size.
+Height allows players to see over opponents and to reach balls higher in the air.
+Thus, greater height is particularly valuable for Quarterbacks and Wide Receivers.
+Heavier players are tougher to budge and to tackle.
+Greater weight is particularly valuable for Linemen, Fullbacks, and Tight Ends, but it can also be valuable—to a deree—for Quarterbacks, Running Backs, and Wide Receivers.
+Hand size is particularly valuable for people catching the ball; thus, a larger hand size is particularly valuable for Wide Receivers and Tight Ends.
+
+In terms of body function, we can consider a player's agility, strength, speed, acceleration/explosiveness, and jumping ability.
+For Wide Receivers, speed, explosiveness, and jumping ability are particularly valuable.
+For Running Backs, agility, strength, speed, and explosiveness are particularly valuable.
 
 # Historical Performance
+
+> "The best predictor of future behavior is past behavior." – Unknown
+
+> "Past performance does not guarantee future results." – A common disclaimer about investments.
 
@@
 - consistency
+
+It is important to consider a player's past performance.
+However, the extent to which historical performance may predict future performance may depend on many factors such as (a) the similarity of the prior situation to the current situation, (b) how long ago the prior situation was, and (c) the extent to which the player (or situation) has changed in the interim.
+For rookies, the player does not have prior seasons of performance in the National Football League (NFL) to draw upon.
+Thus, when evaluating rookies, it can be helpful to consider their performance in college or in their prior leagues.
+However, there are large differences between the situation in college and the situation in the NFL, so prior success in college may not portend future success in the NFL.
+An indicator that intends to be prognostic of future performance, and that accounts for past performance, is a player's draft position—that is, how early (or late) was a player selected in the NFL Draft.
+The earlier a player was selected in the NFL Draft, the greater likelihood that the player will perform well.
+
+For players who have played in the NFL, past performance becomes more relevant because, presumably, the prior situation is more similar (than was their situation in college) to their current situation.
+Nevertheless, lots of things change from game to game and season to season: injuries, coaches, coaching strategies, teammates, etc.
+So just because a player performed well or poorly in a given game or season does not necessarily mean that they will perform similarly in subsequent games/seasons.
+Nevertheless, historical performance is one of the best indicators we have.
+
+In addition to how many fantasy points a player scores in terms of historical performance, we also care about efficiency.
+How efficient were they given the number of opportunities they had.
+If they were relatively more efficient, they will likely score more points than many of their peers when given more opportunities.
+If they were relativelly inefficient, their capacity to score fantasy points may be more dependent on touches/opportunities.
+Efficiency might be operationalized by indicators such as yards per passing attempt, yards per rushing attempt, yards per target, yards per reception, etc.
 
@@
 
+- reaction time
+- intelligence
 - work ethic and mental toughness
+- incentives
+    - contract performance incentives
     - whether they are in a contract year

### 2024-06-02 player-evaluation.qmd
20240601 - empirical question

@@
 - cognitive and motivational factors
+
+The discussion that follows is based on my and others' *impressions* of which characteristics may be valuable to consider when evaluating players.
+However, the extent to which any factor is actually relevant for predicting future performance is an empirical question and should be evaluated empirically.
 

### 2024-06-02 player-evaluation.qmd
20240602 - player evaluation

@@
 
-# Athletic Profile
+# Athletic Profile {sec-evalAthleticProfile}
+
+Factors related to a player's athletic profile include factors such as:
 
@@
 
-# Historical Performance
+Many aspects of a player's athletic profile are available from the National Football League (NFL) Combine, which is especially relevant for evaluating rookies.
+We demonstrate how to import data from the NFL Combine in Section @sec-combine.
+
+# Historical Performance {sec-evalHistoricalPerformance}
 
@@
 > "Past performance does not guarantee future results." – A common disclaimer about investments.
+
+Factors relating to historical performance to consider could include:
 
@@
 However, the extent to which historical performance may predict future performance may depend on many factors such as (a) the similarity of the prior situation to the current situation, (b) how long ago the prior situation was, and (c) the extent to which the player (or situation) has changed in the interim.
-For rookies, the player does not have prior seasons of performance in the National Football League (NFL) to draw upon.
+For rookies, the player does not have prior seasons of performance in the NFL to draw upon.
 Thus, when evaluating rookies, it can be helpful to consider their performance in college or in their prior leagues.
@@
 
+We demonstrate how to import historical player statistics in Section @sec-downloadActualStats.
+We demonstrate how to calculate historical player statistics in Section @sec-calculateActualStats.
+We demonstrate how to calculate historical fantasy points in Section @sec-calculateActualPoints.
+
 In addition to how many fantasy points a player scores in terms of historical performance, we also care about efficiency.
-How efficient were they given the number of opportunities they had.
+How efficient were they given the number of opportunities they had?
 If they were relatively more efficient, they will likely score more points than many of their peers when given more opportunities.
@@
 
-# Health
+# Health {sec-evalHealth}
+
+Health-related factors to consider include:
 
@@
 
-# Age and Career Stage
+It is also important to consider a player's past and current health status.
+In terms of a player's current health status, it is important to consider whether they are injured or are playing at less than 100% of their typical health.
+In terms of a player's prior health status, one can consider their injury history, including the frequency and severity of injuries and their prognosis.
+
+We demonstrate how to import injury reports in Section @sec-injuries.
+
+# Age and Career Stage {sec-evalAgeCareerStage}
+
+Age and career stage-related factors include:
 
 - age
+- experience
 - touches
 
-# Situational Factors
+A player's age is relevant because of important age-related changes in a player's speed, ability to recover from injury, etc.
+A player's experience is relevant because players develop knowledge and skills with greater experience.
+A player's prior touches/usage is also relevant, because it speaks to how many hits a player may have taken.
+For players who take more hits, it may be more likely that their bodies "break down" sooner.
 
-Situational factors are influenced by many factors, including skill of the player, skill of their teammates, coaching style, strategy of the opposing team, game scripts, etc.
+# Situational Factors {sec-evalSituation}
+
+Situational factors one could consider include:
 
@@
 
-# Matchups
+Football is a team sport.
+A player is embedded within a broader team context; it is important to consider the strength of their team context insofar as it may support— or detract from—a player's performance.
+For instance, for a Quarterback, it is important to consider how strong the pass blocking is from the Offensive Line.
+Will they have enough time to throw the ball, or will they be constantly under pressure to be sacked?
+It is also important to consider the strength of the pass catchers—the Wide Receivers and Tight Ends.
+For a Running Back, it is important to consider how strong the run blocking is from the Offensive Line.
+For a Wide Receiver, it is important to consider how strong the pass blocking is, and how strong the Quarterback is.
+
+It is also important to consider a player's role on the team.
+Is the player a starter or a backup?
+Related to this, it is important to consider the strength of one's teammates.
+For a given Running Back, if a teammate is better at running the ball, this may take away from how much the player sees the field.
+For a given Wide Receiver, if a teammate is better at catching the ball, this may take some targets away from the player.
+However, the team's top defensive back is often matched up against the team's top Wide Receiver.
+So, if the team's top Wide Receiver is matched up against a particularly strong Defensive Back, the second- and third-best Wide Receivers may more targets than usual.
+
+It is also important to consider a player's opportunity and usage, which are influenced by many factors, including the skill of the player, the skill of their teammates, the role of the player on the team, the coaching style, the strategy of the opposing team, game scripts, etc.
+In terms of the player's opportunity and usage, how many snaps do they get?
+How many touches and/or targets do they receive?
+Being on the field for more snaps and receiving more touches and/or targets means that the player has more opportunities to score fantasy points.
+Are they targeted in the red zone?
+Red zone targets are more likely to lead to touchdown scoring opportunities, which are particularly valuable in fantasy football.
+
+# Matchups {sec-evalMatchups}
+
+Matchup-related factors to consider include:
 
@@
 
-# Cognitive and Motivational Factors
+Another aspect to consider is how challenging their matchup(s) and strength of schedule is.
+For a Quarterback, it is valuable to consider how strong the oppenent's passing defense is.
+For a Running Back, how strong is the running defense?
+For a Wide Receiver, how strong is the passing defense and the Defensive Back that is likely to be assigned to guard them?
+
+# Cognitive and Motivational Factors {sec-evalCogMotivational}
+
+Other factors to consider include cognitive and motivational factors
+Cognitive and motivational factors one could consider include:
 
 - reaction time
-- intelligence
+- knowledge and intelligence
 - work ethic and mental toughness
@@
     - whether they are in a contract year
+
+A player's knowledge, intelligence, and reaction time can help them gain an upper-hand even when they may not be the fastest or strongest.
+A player's work ethic and mental toughness may help them be resilient and persevere in the face of challenges.
+Contact-related incentives may lead a player to put forth greater effort.
+For instance, a contract may have a performance incentive that provides a player greater compensation if they achieve a particular performance milestone (e.g., receiving yards).
+Another potential incentive is if a player is in what is called their "contract year" (i.e., the last year of their current contract).
+If a player is in the last year of their current contract, they have an incentive to perform well so they can get re-signed to a new contract.

### 2024-06-02 player-evaluation.qmd
20240602 - integration of evaluation information

@@
 
-The discussion that follows is based on my and others' *impressions* of which characteristics may be valuable to consider when evaluating players.
+The discussion that follows is based on my and others' *impressions* of some of the characteristics that may be valuable to consider when evaluating players.
 However, the extent to which any factor is actually relevant for predicting future performance is an empirical question and should be evaluated empirically.
 
-# Athletic Profile {sec-evalAthleticProfile}
+## Athletic Profile {#sec-evalAthleticProfile}
 
@@
 
-# Historical Performance {sec-evalHistoricalPerformance}
+## Historical Performance {#sec-evalHistoricalPerformance}
 
@@
 
-# Health {sec-evalHealth}
+## Health {#sec-evalHealth}
 
@@
 
-# Age and Career Stage {sec-evalAgeCareerStage}
+## Age and Career Stage {#sec-evalAgeCareerStage}
 
@@
 
-# Situational Factors {sec-evalSituation}
+## Situational Factors {#sec-evalSituation}
 
@@
 
-# Matchups {sec-evalMatchups}
+## Matchups {#sec-evalMatchups}
 
@@
 
-# Cognitive and Motivational Factors {sec-evalCogMotivational}
+## Cognitive and Motivational Factors {#sec-evalCogMotivational}
 
@@
 If a player is in the last year of their current contract, they have an incentive to perform well so they can get re-signed to a new contract.
+
+## Putting it Altogether {#sec-evalIntegration}
+
+After performing an evaluation of the relevant domain(s) for a given player, then one must integrate the evaluation information across domains to make a judgment about a player's overall value.
+When thinking about a player's value, it can be worth thinking of a player's upside (or ceiling) and a player's downside (or flow).
+Player that are more consistent may show higher downside but a lower upside.
+Younger, less experienced players may show a higher upside but a lower downside.
+
+The extent to which you prioritize a higher upside versus a higher downside may depend on many factors.
+For instance, when drafting players, you may prioritize drafting players with the highest downside (i.e., the safest players), whereas you may draft sleepers (i.e., players with higher upside) for your bench.
+When choosing which players to start in a given week, if you are predicted to beat a team handily, it may make sense to start the players with the highest downside.
+By contrast, if you are predicted to lose to a team by a good margin, it may make sense to start the players with the highest upside.

### 2024-06-02 player-evaluation.qmd
20240602 - so-called intangibles

@@
 
-Other factors to consider include cognitive and motivational factors
+Other factors to consider include cognitive and motivational factors.
+Some coaches refer to these as the "X Factor" or "the intangibles."
+However, just as any other construct in psychology, we can devise ways to operationalize them.
+
 Cognitive and motivational factors one could consider include:

### 2024-06-03 draft.qmd
20240602 - the fantasy draft

# The Fantasy Draft {#sec-draft}

## Types of Fantasy Drafts {#sec-draftTypes}

There are several types of drafts in fantasy football.
The most common types of drafts are snake drafts and auction drafts.

### Snake Draft {#sec-draftSnake}

In a snake draft, the participants (i.e., managers) are assigned a draft order.
In the first round, the managers draft in that order.
In the second round, the managers draft in reverse order.
It continues to "snake" in this way, round after round, so that the person who has the first pick in a given round has the last pick in the next round, and whoever has the last pick in a given round has the first pick in the next round.

### Auction Draft [#sec-draftAuction]

In an auction draft, the managers are assigned a nomination order and there is a salary cap (e.g., $200).
The first manager chooses which player to nominate.
Then, the managers bid on that player like in an auction.
In order to bid, the manager must raise the price by at least $1.
If two managers want to obtain the same player, they may continue to raise the amount until one manager backs out and is no longer to bid by raising the price.
The highest bidder wins (i.e., drafts) that player.
Then, the second manager nominates a player, and the managers bid on that player.
This process repeats until all teams have drafted their allotment of players.

In the auction draft, unlike the snake draft, all players are available to all teams.
For instance, in the snake draft, the first 9 players drafted are unavailable to the 10th pick of the first round.
So, if you have the 10th pick and want the top-ranked player, this player would not be available to you in the snake draft.
However, in the auction draft, every player is available to every manager, so long as the manager is able and willing to bid enough.

## Draft Strategy {#sec-draftStrategy}

There is no one "right" draft strategy.
Sometimes it works best to "zig" when everyone else is "zagging".
For instance, if you notice that everyone else is drafting Wide Receivers, this may mean that other managers are over-valuing Wide Receivers, and this could be a nice opportunity to draft a Running Back for good value.

An important concept in the draft is "dropoff."
*Dropoff* at a given position, is the difference—in terms of projected fantasy points—between (a) the best available player remaining at that position and (b) the second-best available player remaining at that position.
If there is a bigger dropoff at a given position, there may be greater value in drafting the top player from that position.

### Snake Draft {#sec-draftStrategySnake}

An important concept in the snake draft is a player's value over a typical replacement player (shortened to "value over replacement player").
A player's *value over a replacement player* (VORP) is the difference between (a) a player's projected fantasy points and (b) the fantasy points that you would be expected to get from a typical bench player at that position.

### Auction Draft [#sec-draftStrategyAuction]

### 2024-06-03 draft.qmd
20240602 - dropoff example

@@
 If there is a bigger dropoff at a given position, there may be greater value in drafting the top player from that position.
+For instance, consider the following scenario:
+"Quarterback A" is projected to score 325 points, and "Quarterback B" is projected to score 320 points.
+"Tight End A" is projected to score 230 points, and "Tight End B" is projected to score 150 points.
+In this example, there is a much greater dropoff for Tight Ends than there is for Quarterbacks.
+Thus, even though "Quarterback A" is projected to score more points than "Tight End A", "Tight End A" may be more valuable because there is still a good Quarterback available if someone else drafts "Quarterback A".
 

### 2024-06-04 draft.qmd
20240603 - fix headings for cross-references
20240603 - clarification
20240604 - more player info

@@
 
-### Auction Draft [#sec-draftAuction]
+### Auction Draft {#sec-draftAuction}
 
@@
 
-An important concept in the draft is "dropoff."
+In general, you will first want to generate the rankings you will use to select which players to prioritize.
+You may generate your rankings based one or more of the following:
+
+    - your [player evaluation](player-evaluation.qmd)
+    - expert or aggregated rankings
+    - players' Average Draft Position (ADP) in other league drafts (for snake drafts), or players' Average Auction Value (AAV) in other league drafts (for auction drafts)
+    - expert or aggregated projections
+
+An important concept in the draft is "dropoff", which is described in Section @sec-fantasyValue.
 *Dropoff* at a given position, is the difference—in terms of projected fantasy points—between (a) the best available player remaining at that position and (b) the second-best available player remaining at that position.
@@
 
-An important concept in the snake draft is a player's value over a typical replacement player (shortened to "value over replacement player").
+An important concept in the snake draft is a player's value over a typical replacement player at that position (shortened to "value over replacement player"), which is described in Section @sec-fantasyValue.
 A player's *value over a replacement player* (VORP) is the difference between (a) a player's projected fantasy points and (b) the fantasy points that you would be expected to get from a typical bench player at that position.
 
-### Auction Draft [#sec-draftStrategyAuction]
+### Auction Draft {#sec-draftStrategyAuction}

### 2024-06-04 player-evaluation.qmd
20240604 - more player info

@@
 - cognitive and motivational factors
+- fantasy value
 
@@
 
-In addition to how many fantasy points a player scores in terms of historical performance, we also care about efficiency.
+In addition to how many fantasy points a player scores in terms of historical performance, we also care about efficiency and consistency.
 How efficient were they given the number of opportunities they had?
@@
 Efficiency might be operationalized by indicators such as yards per passing attempt, yards per rushing attempt, yards per target, yards per reception, etc.
+
+In terms of consistency, how consistent was the player they from game to game and from season to season?
+For instance, we could examine the standard deviations of players' fantasy points across games in a given season.
+However, the standard deviation tends to be upwardly biased as the mean increases.
+So, we can account for the player's mean fantasy points per game by dividing their game-to-game standard deviation of fantasy points by their mean fantasy points across games.
+This is known as the coefficient of variation (CV):
+
+$$
+CV = \frac{\sigma}{\mu}
+$$
+
+Players with a lower standard deviation and a lower coefficient of variation (of fantasy points across games) are more consistent.
+
+[ADD EXAMPLE]
 
@@
 
+## Fantasy Value {#sec-fantasyValue}
+
+There are several sources that one can draw upon to evaluate a player's fantasy value:
+
+- expert or aggregated rankings
+- players' Average Draft Position (ADP) in other league drafts
+- players' Average Auction Value (AAV) in other league drafts
+- expert or aggregated projections
+
+[ADD LINKS TO SOURCES OF RANKINGS/PROJECTIONS/ADP/AAV, ETC]
+
+It is important to keep in mind that rankings, average draft position (ADP), and average auction values (AAV) are specific to roster and scoring settings of a particular league.
+For instance, [DESCRIBE]
+Thus, if using rankings, ADP, or AAV, it is important to find ones from leagues that mirror—as closely as possible—your league settings.
+
+Projected statistics (e.g., projected passing touchdowns) are agnostic to league settings and can thus be used to generate league-specific fantasy projections and rankings.
+Thus, projected statisitics may be more useful than rankings because they can be used to generate rankings for your particular league settings.
+For instance, if you know how many touchdowns, yards, and interceptions a Quarterback is a projected to throw (in addition to any other relevant categories for the player, e.g., rushing yards and touchdowns), you can calculate how many fantasy points the Quarterback is expected to gain in your league (or in any league).
+Thus, you can calculate ranking from projections, but you cannot reverse engineer projections from rankings.
+
+There are many potential indices that can be worthwhile to consider, including a player's:
+
+- dropoff
+- value over replacement player (VORP)
+- uncertainty
+
+A player's *dropoff*, is the difference between (a) the player's projected points and (b) the projected points of the next-best player at that position.
+We can also consider a player's value over a typical replacement player at that position (shortened to "value over replacement player").
+A player's *value over a replacement player* (VORP) is the difference between (a) a player's projected fantasy points and (b) the fantasy points that you would be expected to get from a typical bench player at that position.
+
+A player's *uncertainty* is how much variability there is in projections or rankings for a given player across sources.
+For instance, consider a scenario where three experts provide ratings about two players, Player A and Player B.
+Player A is projected to score 300, 310, and 290 points by experts 1, 2, and 3, respectively.
+Player B is projected to score 400, 300, and 200 points by experts 1, 2, and 3, respectively.
+In this case, both players are (on average) projected to score the same number of points (300).
+
+```{r}
+exampleData <- data.frame(
+  player = c(rep("A", 3), rep("B", 3)),
+  expert = c(1:3, 1:3),
+  projectedPoints = c(300, 310, 290, 400, 300, 200)
+)
+
+playerA_mean <- mean(exampleData$projectedPoints[which(exampleData$player == "A")])
+playerB_mean <- mean(exampleData$projectedPoints[which(exampleData$player == "B")])
+
+playerA_mean
+playerB_mean
+```
+
+However, the players differ considerably in their uncertainty (i.e., the source-to-source variability in their projections), as operationalized with the standard deviation and coefficient variation of projected points across sources for a given player.
+
+```{r}
+playerA_sd <- sd(exampleData$projectedPoints[which(exampleData$player == "A")])
+playerb_sd <- sd(exampleData$projectedPoints[which(exampleData$player == "B")])
+
+playerA_cv <- playerA_mean / playerA_sd
+playerB_cv <- playerB_mean / playerb_sd
+
+playerA_mean
+playerB_mean
+
+playerA_sd
+playerb_sd
+
+playerA_cv
+playerB_cv
+```
+
+Uncertainty is not necessarily a bad.
+It just means we have less confidence about how they may be expected to perform.
+Thus, players with greater uncertainty are risky and tend to have a higher upside (or ceiling) and a lower downside (or floor).
+
 ## Putting it Altogether {#sec-evalIntegration}
@@
 After performing an evaluation of the relevant domain(s) for a given player, then one must integrate the evaluation information across domains to make a judgment about a player's overall value.
-When thinking about a player's value, it can be worth thinking of a player's upside (or ceiling) and a player's downside (or flow).
+When thinking about a player's value, it can be worth thinking of a player's upside and a player's downside.
 Player that are more consistent may show higher downside but a lower upside.

### 2024-06-05 player-evaluation.qmd
20240604 - clean up
20240604 - edit
20240604 - relative athletic score (RAS)

@@
     - hand size
+    - wing span (arm length)
 - body function
@@
 
-In terms of body shape, we might consider a player's height, weight, and hand size.
+In terms of body shape, we might consider a player's height, weight, hand size, and wing span (arm length).
 Height allows players to see over opponents and to reach balls higher in the air.
@@
 Greater weight is particularly valuable for Linemen, Fullbacks, and Tight Ends, but it can also be valuable—to a deree—for Quarterbacks, Running Backs, and Wide Receivers.
-Hand size is particularly valuable for people catching the ball; thus, a larger hand size is particularly valuable for Wide Receivers and Tight Ends.
+Hand size and wing span is particularly valuable for people catching the ball; thus, a larger hand size and longer wing span are particularly valuable for Wide Receivers and Tight Ends.
 
@@
 We demonstrate how to import data from the NFL Combine in Section @sec-combine.
+There are also calculators that integrate information about body shape and information from the NFL Combine to determine a player's relative athletic score (RAS) for their position: <https://ras.football/ras-calculator/>
 
@@
 
-playerA_mean
-playerB_mean
-
 playerA_sd
@@
 
-Uncertainty is not necessarily a bad.
+Uncertainty is not necessarily a bad thing for a player.
 It just means we have less confidence about how they may be expected to perform.

### 2024-06-05 draft.qmd
20240604 - bullets
20240604 - edit

@@
 
-    - your [player evaluation](player-evaluation.qmd)
+- your [evaluation of players](player-evaluation.qmd)
 - expert or aggregated rankings

### 2024-06-05 player-evaluation.qmd
20240604 - links

@@
 
-[ADD LINKS TO SOURCES OF RANKINGS/PROJECTIONS/ADP/AAV, ETC]
+Fantasy rankings are provided by many sources.
+[FantasyPros](https://www.fantasypros.com/nfl/rankings/consensus-cheatsheets.php) aggregates fantasy rankings across sources.
+[Fantasy Football Analytics](https://fantasyfootballanalytics.net) creates fantasy rankings from projections that are aggregated across sources (see the webapp here: <https://apps.fantasyfootballanalytics.net>).
+
+Average Draft Position (ADP) data are provided by [FantasyPros](https://www.fantasypros.com/nfl/adp/overall.php).
+Average Auction Value (AAV) data are also provided by [FantasyPros](https://www.fantasypros.com/nfl/auction-values/calculator.php).
+
+Projections that are aggregated across sources are provided by [Fantasy Football Analytics](https://fantasyfootballanalytics.net) (see the webapp here: <https://apps.fantasyfootballanalytics.net>) and by [FantasyPros](https://www.fantasypros.com/nfl/auction-values/calculator.php).
 

### 2024-06-05 player-evaluation.qmd
20240605 - edits

@@
 
-Uncertainty is not necessarily a bad thing for a player.
-It just means we have less confidence about how they may be expected to perform.
+Uncertainty is not necessarily a bad characteristic of a player's projected points.
+It just means we have less confidence about how the player may be expected to perform.
 Thus, players with greater uncertainty are risky and tend to have a higher upside (or ceiling) and a lower downside (or floor).

### 2024-06-05 draft.qmd
20240604 - links
20240605 - VORP

@@
 
+Section @sec-fantasyValue describes where to obtain aggregated rankings, aggregated projections, ADP, and AAV data.
+
 An important concept in the draft is "dropoff", which is described in Section @sec-fantasyValue.
@@
 
-An important concept in the snake draft is a player's value over a typical replacement player at that position (shortened to "value over replacement player"), which is described in Section @sec-fantasyValue.
-A player's *value over a replacement player* (VORP) is the difference between (a) a player's projected fantasy points and (b) the fantasy points that you would be expected to get from a typical bench player at that position.
+An important concept in the snake draft is a player's value over a typical replacement player at that position (shortened to "value over replacement player"; VORP), which is described in Section @sec-fantasyValue.
+In general, your goal is to draft the team whose weekly starting lineup has the greatest VORP.
 

### 2024-06-05 player-evaluation.qmd
20240605 - VORP

@@
 
-A player's *dropoff*, is the difference between (a) the player's projected points and (b) the projected points of the next-best player at that position.
-We can also consider a player's value over a typical replacement player at that position (shortened to "value over replacement player").
+A player's *dropoff* is the difference between (a) the player's projected points and (b) the projected points of the next-best player at that position.
+
+Because players from some positions (e.g., Quarterbacks) tend to score more points than players from other positions (e.g., Wide Receivers), it would be inadvisable to compare players across different positions based on projected points.
+In order to more fairly compare players across positions, we can consider a player's value over a typical replacement player at that position (shortened to "value over replacement player").
 A player's *value over a replacement player* (VORP) is the difference between (a) a player's projected fantasy points and (b) the fantasy points that you would be expected to get from a typical bench player at that position.
+Thus, VORP provides an index of how much added value a player provides.
 

### 2024-06-05 getting-started.qmd
20240605 - figure captions

@@
 ```{r}
+#| label: fig-rushYPCperSeasonByAge
+#| fig-cap: "2023 NFL Rushing Yards Per Carry per Season by Player Age"
+
 # Prepare Data
@@
 ```{r}
+#| label: fig-rushEPAperSeasonByAge
+#| fig-cap: "2023 NFL Rushing Expected Points Added (EPA) per Season by Player Age"
+
 # Subset Data
@@
     y = "Rushing EPA/season",
-    title = "2023 NFL Rushing EPA per Season by Player Age"
+    title = "2023 NFL Rushing Expected Points Added (EPA) per Season by Player Age"
   ) +

### 2024-06-05 player-evaluation.qmd
20240605 - density plot

@@
 
+Here is a depiction of a density plot of projected points for a player with a low, medium, and high uncertainty:
+
+```{r}
+#| label: fig-playerUncertainty
+#| fig-cap: "Density Plot of Projected Points for Three Players"
+
+playerA <- rnorm(1000000, mean = 150, sd = 5)
+playerB <- rnorm(1000000, mean = 150, sd = 15)
+playerC <- rnorm(1000000, mean = 150, sd = 30)
+
+mydata <- data.frame(playerA, playerB, playerC)
+
+mydata_long <- mydata %>% 
+  pivot_longer(
+    cols = everything(),
+    names_to = "player",
+    values_to = "points"
+  ) %>% 
+  mutate(
+    name = case_match(
+      player,
+      "playerA" ~ "Player A",
+      "playerB" ~ "Player B",
+      "playerC" ~ "Player C",
+    )
+  )
+
+ggplot2::ggplot(
+  data = mydata_long,
+  ggplot2::aes(
+    x = points,
+    fill = name
+  )
+) +
+  ggplot2::geom_density(alpha = .3) + 
+  ggplot2::labs(
+    x = "Players' Projected Points",
+    title = "Density Plot of Projected Points for Three Players"
+  ) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(legend.title = element_blank())
+```
+
 Uncertainty is not necessarily a bad characteristic of a player's projected points.

### 2024-06-05 base-rates.qmd
20240605 - getting started/load packages

# Base Rates {#sec-baseRates}

## Getting Started

### Load Packages

```{r}

```

### 2024-06-06 player-evaluation.qmd
20240605 - getting started/load packages
20240605 - code folding
20240606 - coefficient of variation

@@
 # Player Evaluation {#sec-playerEvaluation}
+
+## Getting Started {#sec-playerEvaluationGettingStarted}
+
+### Load Packages {#sec-playerEvaluationLoadPackages}
+
+```{r}
+library("tidyverse")
+```
+
+## Overview
 
@@
 Players with a lower standard deviation and a lower coefficient of variation (of fantasy points across games) are more consistent.
+In the example below, Player 2 might be preferable to Player 1 because Player 2 is more consistent; Player 1 is more "boom-or-bust."
+Despite showing a similar mean of fantasy points across weeks, Player 2 shows a smaller week-to-week standard deviation and coefficient of variation.
 
-[ADD EXAMPLE]
+```{r}
+set.seed(1)
+
+playerScoresByWeek <- data.frame(
+  player1_scores = rnorm(17, mean = 20, sd = 7),
+  player2_scores = rnorm(17, mean = 20, sd = 4),
+  player3_scores = rnorm(17, mean = 10, sd = 4),
+  player4_scores = rnorm(17, mean = 10, sd = 1)
+)
+
+consistencyData <- data.frame(t(playerScoresByWeek))
+
+weekNames <- paste("week", 1:17, sep = "")
+
+names(consistencyData) <- weekNames
+row.names(consistencyData) <- NULL
+
+consistencyData$mean <- rowMeans(consistencyData[,weekNames])
+consistencyData$sd <- apply(consistencyData, 1, sd)
+consistencyData$cv <- consistencyData$sd / consistencyData$mean
+
+consistencyData$player <- c(1, 2, 3, 4)
+
+consistencyData <- consistencyData %>% 
+  select(player, mean, sd, cv, week1:week17)
+
+round(consistencyData, 2)
+```
 
@@
 #| fig-cap: "Density Plot of Projected Points for Three Players"
+#| code-fold: true
 

### 2024-06-06 player-evaluation.qmd
20240606 - define symbols

@@
 However, the standard deviation tends to be upwardly biased as the mean increases.
-So, we can account for the player's mean fantasy points per game by dividing their game-to-game standard deviation of fantasy points by their mean fantasy points across games.
+So, we can account for the player's mean fantasy points per game by dividing their game-to-game standard deviation of fantasy points ($\sigma$) by their mean fantasy points across games ($\mu$).
 This is known as the coefficient of variation (CV):

### 2024-06-08 draft.qmd
20240605 - getting started/load packages
20240608 - cross-references and content

@@
 # The Fantasy Draft {#sec-draft}
+
+## Getting Started {#sec-draftGettingStarted}
+
+### Load Packages {#sec-draftLoadPackages}
+
+```{r}
+
+```
 
@@
 
+### Overview {#sec-draftStrategyOverview}
+
 There is no one "right" draft strategy.
@@
 - your [evaluation of players](player-evaluation.qmd)
-- expert or aggregated rankings
-- players' Average Draft Position (ADP) in other league drafts (for snake drafts), or players' Average Auction Value (AAV) in other league drafts (for auction drafts)
-- expert or aggregated projections
+- [expert or aggregated rankings](#sec-fantasyValueRankings)
+- [layperson rankings](#sec-fantasyValueADPAAV)
+    - players' [Average Draft Position](#sec-fantasyValueADPAAV) (ADP) in other league drafts (for [snake drafts](#sec-draftStrategySnake))
+    - players' [Average Auction Value](#sec-fantasyValueADPAAV) (AAV) in other league drafts (for [auction drafts](sec-draftStrategyAuction))
+- [expert or aggregated projections](#sec-fantasyValueProjections)
+- [indices derived from rankings and projections](#sec-fantasyValueIndices)
 
-Section @sec-fantasyValue describes where to obtain aggregated rankings, aggregated projections, ADP, and AAV data.
+Section @sec-fantasyValueSources describes where to obtain [aggregated rankings](#sec-fantasyValueRankings), [aggregated projections](#sec-fantasyValueProjections), [ADP](#sec-fantasyValueADPAAV), and [AAV](#sec-fantasyValueADPAAV) data.
 
-An important concept in the draft is "dropoff", which is described in Section @sec-fantasyValue.
-*Dropoff* at a given position, is the difference—in terms of projected fantasy points—between (a) the best available player remaining at that position and (b) the second-best available player remaining at that position.
-If there is a bigger dropoff at a given position, there may be greater value in drafting the top player from that position.
+An important concept in the draft is "[dropoff](#sec-fantasyValueDropoff)", which is described in Section @sec-fantasyValueDropoff.
+[*Dropoff*](#sec-fantasyValueDropoff) at a given position, is the difference—in terms of projected fantasy points—between (a) the best available player remaining at that position and (b) the second-best available player remaining at that position.
+If there is a bigger [dropoff](#sec-fantasyValueDropoff) at a given position, there may be greater value in drafting the top player from that position.
 For instance, consider the following scenario:
@@
 "Tight End A" is projected to score 230 points, and "Tight End B" is projected to score 150 points.
-In this example, there is a much greater dropoff for Tight Ends than there is for Quarterbacks.
+In this example, there is a much greater [dropoff](#sec-fantasyValueDropoff) for Tight Ends than there is for Quarterbacks.
 Thus, even though "Quarterback A" is projected to score more points than "Tight End A", "Tight End A" may be more valuable because there is still a good Quarterback available if someone else drafts "Quarterback A".
+
+Another important concept is a player's [value over a typical replacement player](#sec-fantasyValueVORP) at that position (shortened to "value over replacement player"; VORP), which is described in Section @sec-fantasyValueVORP.
+
+Another important concept is a player's [uncertainty](#sec-fantasyValueUncertainty), which is described in Section @sec-fantasyValueUncertainty.
+
+In both [snake](#sec-draftStrategySnake) and [auction](#sec-draftStrategyAuction) draft formats, your goal is to draft the team whose weekly starting lineup scores the most points and thus the collection of players with the greatest [VORP](#sec-fantasyValueVORP).
+However the draft strategies to achieve that differ between [snake](#sec-draftStrategySnake) versus [auction](#sec-draftStrategyAuction) drafts.
 
@@
 
-An important concept in the snake draft is a player's value over a typical replacement player at that position (shortened to "value over replacement player"; VORP), which is described in Section @sec-fantasyValue.
-In general, your goal is to draft the team whose weekly starting lineup has the greatest VORP.
+In general, your goal is to draft the team whose weekly starting lineup has the greatest [VORP](#sec-fantasyValueVORP).
+Consequently, you are often looking to pick the player with the highest [VORP](#sec-fantasyValueVORP) at a given selection, while keeping in mind (a) the [dropoff](#sec-fantasyValueDropoff) of players at other positions and (b) which players may be available at subsequent picks so that you do not sacrifice too much later value with a given selection.
+For instance, if a particular Quarterback has a slightly higher [VORP](#sec-fantasyValueVORP) than a particular Running Back, but the Quarterback is likely to be available at the manager's next pick but the Running Back is likely to be unavailable at their next pick, it might make more sense to draft the Running Back.
+
+For your starting lineup, it may make sense—especially with your earliest selections—when comparing two players with equivalent [VORP](#sec-fantasyValueVORP), to prioritize players with higher [consistency](#sec-evalHistoricalConsistency) and lower [uncertainty](#sec-fantasyValueUncertainty), because they may be considered "safer" with a higher floor.
+However, when drafting your bench players, it make make more sense to prioritize high risk, high reward players with greater [uncertainty](#sec-fantasyValueUncertainty), because they may have a higher ceiling.
+The Spurs in the National Basketball Association (NBA) were well-reputed for excelling in this draft strategy.
+They frequently used their second-round picks to draft high-risk, high-reward players; sometimes, the secound round pick was a bust, but other times their second round picks including Manu Ginobili and Tony Parker greatly outperformed expectations and contributed to the team's strong success.
 

### 2024-06-08 player-evaluation.qmd
20240608 - cross-references and content

@@
 
+### Overview {#sec-evalHistoricalPerformanceOverview}
+
 > "The best predictor of future behavior is past behavior." – Unknown
@@
 
-In addition to how many fantasy points a player scores in terms of historical performance, we also care about efficiency and consistency.
+### Efficiency {#sec-evalHistoricalEfficiency}
+
+In addition to how many fantasy points a player scores in terms of historical performance, we also care about efficiency and [consistency](#sec-evalHistoricalConsistency).
 How efficient were they given the number of opportunities they had?
@@
 Efficiency might be operationalized by indicators such as yards per passing attempt, yards per rushing attempt, yards per target, yards per reception, etc.
+
+### Consistency {#sec-evalHistoricalConsistency}
 
@@
 
+### Sources From Which to Evaluate Fantasy Value {#sec-fantasyValueSources}
+
 There are several sources that one can draw upon to evaluate a player's fantasy value:
@@
 - expert or aggregated rankings
-- players' Average Draft Position (ADP) in other league drafts
-- players' Average Auction Value (AAV) in other league drafts
+- layperson rankings
+    - players' Average Draft Position (ADP) in other league [snake drafts](#sec-draftStrategySnake)
+    - players' Average Auction Value (AAV) in other league [auction drafts](#sec-draftStrategyAuction)
 - expert or aggregated projections
 
-Fantasy rankings are provided by many sources.
+#### Expert Fantasy Rankings {#sec-fantasyValueRankings}
+
+Fantasy rankings (by so-called "experts") are provided by many sources.
+To reduce some of the bias due to a given source, some services aggregate projections across sources, consistent with a "wisdom of the crowd" approach.
 [FantasyPros](https://www.fantasypros.com/nfl/rankings/consensus-cheatsheets.php) aggregates fantasy rankings across sources.
@@
 
-Average Draft Position (ADP) data are provided by [FantasyPros](https://www.fantasypros.com/nfl/adp/overall.php).
-Average Auction Value (AAV) data are also provided by [FantasyPros](https://www.fantasypros.com/nfl/auction-values/calculator.php).
+#### Layperson Fantasy Rankings: ADP and AAV {#sec-fantasyValueADPAAV}
 
+Average Draft Position (ADP) and Average Auction Value (AAV), are based on league drafts, mostly composed of everyday people.
+ADP is based on [snake drafts](#sec-draftStrategySnake), whereas AAV is based on [auction drafts](#sec-draftStrategyAuction).
+Thus, ADP and AAV are consistent with a "wisdom of the crowd" approach, and I refer to them as forms of rankings by laypeople.
+ADP data are provided by [FantasyPros](https://www.fantasypros.com/nfl/adp/overall.php).
+AAV data are also provided by [FantasyPros](https://www.fantasypros.com/nfl/auction-values/calculator.php).
+
+#### Projections {#sec-fantasyValueProjections}
+
+Projections are provided by various sources.
+Projections (and rankings, for that matter) are a bit of a black box.
+It is often unclear how they were derived by a particular source.
+That is, it is unclear how much of the projection was based on statistical analysis versus conjecture.
+
+To reduce some of the bias due to a given source, some services aggregate projections across sources, consistent with a "wisdom of the crowd" approach.
 Projections that are aggregated across sources are provided by [Fantasy Football Analytics](https://fantasyfootballanalytics.net) (see the webapp here: <https://apps.fantasyfootballanalytics.net>) and by [FantasyPros](https://www.fantasypros.com/nfl/auction-values/calculator.php).
 
-It is important to keep in mind that rankings, average draft position (ADP), and average auction values (AAV) are specific to roster and scoring settings of a particular league.
-For instance, [DESCRIBE]
+#### Benefits of Using Projections Rather than Rankings {#sec-fantasyValueProjectionsVsRankings}
+
+It is important to keep in mind that rankings, ADP, and AAV are specific to roster and scoring settings of a particular league.
+For instance, in point-per-reception (PPR) leagues, players who catch lots of passes (Wide Receivers, Tight Ends, and some Running Backs) are valued more highly.
+As another example, Quarterbacks are valued more highly in 2-Quarterback leagues.
 Thus, if using rankings, ADP, or AAV, it is important to find ones from leagues that mirror—as closely as possible—your league settings.
@@
 
+### Indices to Evaluate Fantasy Value {#sec-fantasyValueIndices}
+
+Based on the sources above (rankings, ADP, AAV, and projections), we can derive multiple indices to evaluate fantasy value.
 There are many potential indices that can be worthwhile to consider, including a player's:
@@
 
+#### Dropoff {#sec-fantasyValueDropoff}
+
 A player's *dropoff* is the difference between (a) the player's projected points and (b) the projected points of the next-best player at that position.
+
+#### Value Over Replacement Player {#sec-fantasyValueVORP}
 
@@
 Thus, VORP provides an index of how much added value a player provides.
+
+#### Uncertainty {#sec-fantasyValueUncertainty}
 

### 2024-06-08 draft.qmd
20240608 - reorganize

@@
 In both [snake](#sec-draftStrategySnake) and [auction](#sec-draftStrategyAuction) draft formats, your goal is to draft the team whose weekly starting lineup scores the most points and thus the collection of players with the greatest [VORP](#sec-fantasyValueVORP).
-However the draft strategies to achieve that differ between [snake](#sec-draftStrategySnake) versus [auction](#sec-draftStrategyAuction) drafts.
+For your starting lineup, it may make sense—especially with your earliest selections—when comparing two players with equivalent [VORP](#sec-fantasyValueVORP), to prioritize players with higher [consistency](#sec-evalHistoricalConsistency) and lower [uncertainty](#sec-fantasyValueUncertainty), because they may be considered "safer" with a higher floor.
+However, when drafting your bench players, it make make more sense to prioritize high risk, high reward players with greater [uncertainty](#sec-fantasyValueUncertainty), because they may have a higher ceiling.
+The Spurs in the National Basketball Association (NBA) were well-reputed for excelling in this draft strategy.
+They frequently used their second-round picks to draft high-risk, high-reward players; sometimes, the secound round pick was a bust, but other times their second round picks including Manu Ginobili and Tony Parker greatly outperformed expectations and contributed to the team's strong success.
+
+However, the draft strategies to achieve the "optimal lineup" differ between [snake](#sec-draftStrategySnake) versus [auction](#sec-draftStrategyAuction) drafts.
 
@@
 
-For your starting lineup, it may make sense—especially with your earliest selections—when comparing two players with equivalent [VORP](#sec-fantasyValueVORP), to prioritize players with higher [consistency](#sec-evalHistoricalConsistency) and lower [uncertainty](#sec-fantasyValueUncertainty), because they may be considered "safer" with a higher floor.
-However, when drafting your bench players, it make make more sense to prioritize high risk, high reward players with greater [uncertainty](#sec-fantasyValueUncertainty), because they may have a higher ceiling.
-The Spurs in the National Basketball Association (NBA) were well-reputed for excelling in this draft strategy.
-They frequently used their second-round picks to draft high-risk, high-reward players; sometimes, the secound round pick was a bust, but other times their second round picks including Manu Ginobili and Tony Parker greatly outperformed expectations and contributed to the team's strong success.
-
 ### Auction Draft {#sec-draftStrategyAuction}

### 2024-06-08 draft.qmd
20240608 - auction draft

@@
 For your starting lineup, it may make sense—especially with your earliest selections—when comparing two players with equivalent [VORP](#sec-fantasyValueVORP), to prioritize players with higher [consistency](#sec-evalHistoricalConsistency) and lower [uncertainty](#sec-fantasyValueUncertainty), because they may be considered "safer" with a higher floor.
-However, when drafting your bench players, it make make more sense to prioritize high risk, high reward players with greater [uncertainty](#sec-fantasyValueUncertainty), because they may have a higher ceiling.
-The Spurs in the National Basketball Association (NBA) were well-reputed for excelling in this draft strategy.
-They frequently used their second-round picks to draft high-risk, high-reward players; sometimes, the secound round pick was a bust, but other times their second round picks including Manu Ginobili and Tony Parker greatly outperformed expectations and contributed to the team's strong success.
+However, when drafting players for your bench, it make make more sense to prioritize high risk, high reward players with greater [uncertainty](#sec-fantasyValueUncertainty), because they may have a higher ceiling (i.e., this is a *lower* risk kind of selection with later picks).
+The Spurs in the National Basketball Association (NBA) were well-reputed for [excelling in this draft strategy](https://harvardsportsanalysis.org/2013/11/beating-the-nba-draft-does-any-team-outperform-expectations/) (archived at <https://perma.cc/X7NW-WZC6>).
+They frequently used their second-round picks to draft high-risk, high-reward players.
+Sometimes, the secound round pick was a bust, but they have little to lose with a failed second round pick; and other times their second round picks including DeJuan Blair, Goran Dragic, Luis Scola, and Manu Ginobili greatly outperformed expectations, and several of them contributed to the team's strong success.
 
@@
 ### Auction Draft {#sec-draftStrategyAuction}
+
+According to [an analysis](https://harvardsportsanalysis.wordpress.com/wp-content/uploads/2012/04/fantasyfootballdraftanalysis1.pdf) by the Harvard Sports Analysis Collective (archived at <https://perma.cc/P7RX-92UU>), the majority of the manager's salary cap should be spent on the starting lineup, and you should spend less on bench players.
+This is known as the "stars and scrubs" draft strategy.
+Based on the analysis, the author recommended applying a 10% premium to the top players and a 10% discount to the lower-tiered players.
+The idea behind the approach is that a player on your bench does not contribute to the team's points and, thus, most players drafted to your bench do not contribute much to the team's points throughout the season.
+That said, bench players can be important in the case of a starter's injury or under-performance.
+So, it is recommended to draft starters with lower [uncertainty](#sec-fantasyValueUncertainty) who are safer.
+In contrast to your starting lineup, you may look to draft players on your bench who have greater [uncertainty](#sec-fantasyValueUncertainty) for their high reward potential in a low-risk selection given the lower price.

### 2024-06-08 draft.qmd
20240608 - draft comparison

@@
 
-In the auction draft, unlike the snake draft, all players are available to all teams.
-For instance, in the snake draft, the first 9 players drafted are unavailable to the 10th pick of the first round.
+### Comparison {#sec-draftComparison}
+
+Snake drafts are more common than auction drafts.
+Snake drafts tend to be quicker than auction drafts.
+However, auction drafts are more fair than snake drafts.
+In an auction draft, unlike a snake draft, all players are available to all teams.
+For instance, in a snake draft, the first 9 players drafted are unavailable to the 10th pick of the first round.
 So, if you have the 10th pick and want the top-ranked player, this player would not be available to you in the snake draft.

### 2024-06-08 draft.qmd
20240608 - high risk, high reward

@@
 For your starting lineup, it may make sense—especially with your earliest selections—when comparing two players with equivalent [VORP](#sec-fantasyValueVORP), to prioritize players with higher [consistency](#sec-evalHistoricalConsistency) and lower [uncertainty](#sec-fantasyValueUncertainty), because they may be considered "safer" with a higher floor.
-However, when drafting players for your bench, it make make more sense to prioritize high risk, high reward players with greater [uncertainty](#sec-fantasyValueUncertainty), because they may have a higher ceiling (i.e., this is a *lower* risk kind of selection with later picks).
+However, when drafting players for your bench, it make make more sense to prioritize high-risk, high reward players with greater [uncertainty](#sec-fantasyValueUncertainty), because they may have a higher ceiling.
+Players with a higher ceiling have a potential to be "sleepers"—players who are valued low (i.e., with a high [ADP](#sec-fantasyValueADPAAV) or low [AAV](#sec-fantasyValueADPAAV)) and who outperform their valuation.
+Note that, although players with greater [uncertainty](#sec-fantasyValueUncertainty) are high-risk, high-reward players, selecting this kind of a player for your bench (i.e., in a late round or for a small cost) is a *lower* risk selection, because you have less to lose with later/lower-cost picks.
+That is, even though the *player* is higher risk, selecting a higher risk player for your bench is a lower risk *decision*.
+
 The Spurs in the National Basketball Association (NBA) were well-reputed for [excelling in this draft strategy](https://harvardsportsanalysis.org/2013/11/beating-the-nba-draft-does-any-team-outperform-expectations/) (archived at <https://perma.cc/X7NW-WZC6>).
 They frequently used their second-round picks to draft high-risk, high-reward players.
-Sometimes, the secound round pick was a bust, but they have little to lose with a failed second round pick; and other times their second round picks including DeJuan Blair, Goran Dragic, Luis Scola, and Manu Ginobili greatly outperformed expectations, and several of them contributed to the team's strong success.
+Sometimes, the secound round pick was a bust, but they have little to lose with a failed second round pick.
+Other times, their second round picks—including DeJuan Blair, Goran Dragic, Luis Scola, and Manu Ginobili—greatly outperformed expectations.
+Thanks, in part, to this draft strategy, the team showed strong extended success for nearly three decades from 1989 through the late-2010s.
 

### 2024-06-08 index.qmd
20240608 - ways to contribute

@@
 I will continue to add to and update it as I am able.
+
+## How to Contribute {#sec-contribute}
+
+This is an open-access textbook.
+My goal is to share data analysis strategies for free!
+Anyone is welcome to contribute to the project.
+If you would like to contribute, please consider one of the following:
+
+  - [open an issue](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues) or [create a pull request](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/pulls) on the [book's GitHub repository](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook).\index{GitHub}
+  - [buy me a coffee](https://www.paypal.com/donate/?business=Q4FMMAMFQYTX8&no_recurring=0&item_name=Support+me+in+developing+this+resource+for+fantasy+football+analytics...+Even+a+cup+of+coffee+%28%244.99%29+helps+me+stay+awake%21&currency_code=USD)—Support me in developing this resource for fantasy football analytics...
+Even a cup of coffee ($4.99) helps me stay awake!
+
+```{=html}
+<form action="https://www.paypal.com/donate" method="post" target="_top">
+<input type="hidden" name="business" value="Q4FMMAMFQYTX8" />
+<input type="hidden" name="no_recurring" value="0" />
+<input type="hidden" name="item_name" value="Support me in developing this resource for fantasy football analytics... Even a cup of coffee ($4.99) helps me stay awake!" />
+<input type="hidden" name="currency_code" value="USD" />
+<input type="image" src="https://www.paypalobjects.com/en_US/i/btn/btn_donate_LG.gif" border="0" name="submit" title="PayPal - The safer, easier way to pay online!" alt="Donate with PayPal button" />
+<img alt="" border="0" src="https://www.paypal.com/en_US/i/scr/pixel.gif" width="1" height="1" />
+</form>
+```
+
+The GitHub repository for the book is located here:\index{GitHub} <https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook>.
+If you have data or analysis examples that are you willing to share and include in the book, feel free to contact me.
 
@@
 
-## How to Contribute {#sec-contribute}
-
-This is an open-access textbook.
-My goal is to share data analysis strategies for free!
-Anyone is welcome to contribute to the project.
-If you would like to contribute, feel free to [open an issue](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues) or [create a pull request](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/pulls) on GitHub.\index{GitHub}
-The GitHub repository for the book is located here:\index{GitHub} <https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook>.
-If you have data or analysis examples that are you willing to share and include in the book, feel free to contact me.
-
 ## Acknowledgments {#sec-acknowledgments}

### 2024-06-08 basic-statistics.qmd
20240605 - getting started/load packages
20240608 - getting started
20240608 - statistical decision tree

--- 2024-06-05 base-rates.qmd
@@
-# Base Rates {#sec-baseRates}
+# Basic Statistics {#sec-basicStats}
 
-## Getting Started
+## Getting Started {#sec-basicStatsGettingStarted}
 
-### Load Packages
+### Load Packages {#sec-basicStatsLoadPackages}
 
@@
 ```
+
+## Types of Data
+
+## Statistical Decision Tree
+
+An example statistical decision tree is depicted in Figure @fig-statisticalDecisionTree.
+
+::: {#fig-statisticalDecisionTree}
+![](images/statisticalDecisionTree.png)
+
+A Statistical Decision Tree For Choosing an Appropriate Statistical Procedure.
+Obtained from: <https://commons.wikimedia.org/wiki/File:InferentialStatisticalDecisionMakingTrees.pdf>.
+The original source is: Corston, R., & Colman, A.
+M.
+(2000).
+*A crash course in SPSS for Windows*.
+Wiley-Blackwell.
+:::

### 2024-06-09 index.qmd
20240608 - conditional formatting
20240608 - about the author

@@
+::: {.content-hidden unless-format="pdf"}
+
 # Preface {#sec-preface .unnumbered .unlisted}
+
+:::
+
+::: {.content-hidden when-format="pdf"}
+
+# Preface {#sec-preface}
+
+:::
 
@@
 
+## About the Author {#sec-author}
+
+I am an Assistant Professor in the Department of Psychological and Brain Sciences at the University of Iowa.
+I am a licensed psychologist with expertise in child clinical psychology.
+Why am I writing about fantasy football and data analysis?
+Because fantasy football involves the intersection of two things I love: sports and statistics.
+
+Through my training, I have learned the value of statistics for answering important questions that I find interesting.
+In graduate training, I came to the realization that statistics are relevant not only for psychology and science, but also for domains that I enjoy as hobbies, including sports and fantasy sports.
+I have played in a longstanding fantasy football league for over 20 years (since my junior year of high school) with old friends from high school.
+I wanted to apply what I was learning about statistics to help others improve their performance in fantasy football and to help people—including those who might not otherwise be interested—to learn statistics.
+So I began blogging online about the value of applying statistics to improve decision making in fantasy football.
+Apparently, many people were interested in learning statistics when they could apply them to a domain that they find interesting like fantasy football.
+My blog eventually became [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), a website that uses advanced statistics to help people win their fantasy football leagues.
+
+In terms of my `R` and statistics background, I have published many [peer-reviewed publications](https://developmental-psychopathology.lab.uiowa.edu/publications) that employ advanced statistical methods, have published a [book on psychological assessment](https://www.routledge.com/9781032413068) that includes applied examples in `R`, and have published [an `R` package](https://cran.r-project.org/web/packages/petersenlab/index.html) on the Comprehensive R Archive Network (CRAN).
+
 ## Accessibility {#sec-accessibility}

### 2024-06-09 index.qmd
20240608 - clean up
20240608 - petersenlab package

@@
-::: {.content-hidden unless-format="pdf"}
-
 # Preface {#sec-preface .unnumbered .unlisted}
-
-:::
-
-::: {.content-hidden when-format="pdf"}
-
-# Preface {#sec-preface}
-
-:::
 
@@
 
-In terms of my `R` and statistics background, I have published many [peer-reviewed publications](https://developmental-psychopathology.lab.uiowa.edu/publications) that employ advanced statistical methods, have published a [book on psychological assessment](https://www.routledge.com/9781032413068) that includes applied examples in `R`, and have published [an `R` package](https://cran.r-project.org/web/packages/petersenlab/index.html) on the Comprehensive R Archive Network (CRAN).
+In terms of my `R` and statistics background, I have published many [peer-reviewed publications](https://developmental-psychopathology.lab.uiowa.edu/publications) that employ advanced statistical methods, have published a [book on psychological assessment](https://www.routledge.com/9781032413068) that includes applied examples in `R`, and have published [the `petersenlab` `R` package](https://cran.r-project.org/web/packages/petersenlab/index.html) on the Comprehensive R Archive Network (CRAN).
 

### 2024-06-09 draft.qmd
20240608 - update names
20240608 - clean up

@@
 
-Section @sec-fantasyValueSources describes where to obtain [aggregated rankings](#sec-fantasyValueRankings), [aggregated projections](#sec-fantasyValueProjections), [ADP](#sec-fantasyValueADPAAV), and [AAV](#sec-fantasyValueADPAAV) data.
+@sec-fantasyValueSources describes where to obtain [aggregated rankings](#sec-fantasyValueRankings), [aggregated projections](#sec-fantasyValueProjections), [ADP](#sec-fantasyValueADPAAV), and [AAV](#sec-fantasyValueADPAAV) data.
 
-An important concept in the draft is "[dropoff](#sec-fantasyValueDropoff)", which is described in Section @sec-fantasyValueDropoff.
+An important concept in the draft is "[dropoff](#sec-fantasyValueDropoff)", which is described in @sec-fantasyValueDropoff.
 [*Dropoff*](#sec-fantasyValueDropoff) at a given position, is the difference—in terms of projected fantasy points—between (a) the best available player remaining at that position and (b) the second-best available player remaining at that position.
@@
 
-Another important concept is a player's [value over a typical replacement player](#sec-fantasyValueVORP) at that position (shortened to "value over replacement player"; VORP), which is described in Section @sec-fantasyValueVORP.
+Another important concept is a player's [value over a typical replacement player](#sec-fantasyValueVORP) at that position (shortened to "value over replacement player"; VORP), which is described in @sec-fantasyValueVORP.
 
-Another important concept is a player's [uncertainty](#sec-fantasyValueUncertainty), which is described in Section @sec-fantasyValueUncertainty.
+Another important concept is a player's [uncertainty](#sec-fantasyValueUncertainty), which is described in @sec-fantasyValueUncertainty.
 
@@
 Sometimes, the secound round pick was a bust, but they have little to lose with a failed second round pick.
-Other times, their second round picks—including DeJuan Blair, Goran Dragic, Luis Scola, and Manu Ginobili—greatly outperformed expectations.
+Other times, their second round picks—including Willie Anderson, DeJuan Blair, Goran Dragic, Luis Scola, and Manu Ginóbili—greatly outperformed expectations.
 Thanks, in part, to this draft strategy, the team showed strong extended success for nearly three decades from 1989 through the late-2010s.

### 2024-06-09 player-evaluation.qmd
20240608 - clean up

@@
 Many aspects of a player's athletic profile are available from the National Football League (NFL) Combine, which is especially relevant for evaluating rookies.
-We demonstrate how to import data from the NFL Combine in Section @sec-combine.
+We demonstrate how to import data from the NFL Combine in @sec-combine.
 There are also calculators that integrate information about body shape and information from the NFL Combine to determine a player's relative athletic score (RAS) for their position: <https://ras.football/ras-calculator/>
@@
 
-We demonstrate how to import historical player statistics in Section @sec-downloadActualStats.
-We demonstrate how to calculate historical player statistics in Section @sec-calculateActualStats.
-We demonstrate how to calculate historical fantasy points in Section @sec-calculateActualPoints.
+We demonstrate how to import historical player statistics in @sec-downloadActualStats.
+We demonstrate how to calculate historical player statistics in @sec-calculateActualStats.
+We demonstrate how to calculate historical fantasy points in @sec-calculateActualPoints.
 
@@
 
-We demonstrate how to import injury reports in Section @sec-injuries.
+We demonstrate how to import injury reports in @sec-injuries.
 

### 2024-06-09 basic-statistics.qmd
20240608 - cleanup
20240609 - levels of measurement

@@
 
-## Types of Data
+## Levels of Measurement {#sec-levelsOfMeasurement}
+
+It is important to know the levels of measurement of your data, because the level(s) of measurement of your data constrain the types of comparisons and analyses that you can meaningfully perform.
+There are four levels of measurement that any variable can have:
+
+- nominal
+- ordinal
+- interval
+- ratio
+
+Each is described below:
+
+### Nominal {#sec-nominal}
+
+A variable is considered nominal if it is composed of qualitative classifications.
+You cannot meaningfully evaluate whether one number in the variable is larger than another variable in the variable because higher numbers do not reflect higher levels of the concept.
+Examples of nominal variables include:
+
+- sex (e.g., 1 = male; 2 = female)
+- race (e.g., 1 = American Indian; 2 = Asian; 3 = Black; 4 = Pacific Islander; 5 = White)
+- ethnicity (e.g., 0 = Non-Hispanic/Latino; 1 = Hispanic/Latino)
+- jersey number
+- zip code
+
+A football player's jersey number is an example of a nominal variable.
+A jersey number of 7 is not higher on whatever concept of interest compared to a jersey number of 6.
+
+To examine the central tendency of a nominal variable, you can determine the mode, but you cannot calculate a mean or median.
+
+### Ordinal {#sec-ordinal}
+
+A variable is considered ordinal if the classifications are ordered.
+However, ordinal variables do not have equally spaced intervals.
+Examples of ordinal intervals include:
+
+- likert response scales (e.g., 1 = strongly disagree; 2 = disagree; 3 = neutral; 4 = agree; 5 = strongly agree)
+- educational attainment (e.g., 1 = no formal education; 2 = elementary school; 3 = middle school; 4 = high school; 5 = college; 6 = graduate degree)
+- academic grades on A–F scale (e.g., 1 = A; 2 = B; 3 = C; 4 = D; 5 = F)
+- player rank (1 = 1st; 2 = 2nd; 3 = 3rd, etc.)
+
+A football player's fantasy rank is an example of an ordinal variable.
+A player with a fantasy rank of 1 has a higher rank than a player with a rank of 2, but it is not known how far apart each player is—i.e., the intervals do not all reflect the same distance.
+For instance, the distance between the top-ranked player and the 2nd-best player might be 30 points, whereas the distance between the 2nd-best player and the 3rd-best player might be 2 points.
+
+To examine the central tendency of ordinal data, the median and mode are most appropriate; however, the mean may be used (unlike for nominal data).
+
+### Interval {#sec-interval}
+
+A variable is considered interval if the classifications are ordered (similar to ordinal data) and have equally spaced intervals (unlike ordinal data).
+However, interval variables do not have a meaningful zero that reflects absence.
+Examples of interval data include:
+
+- temperature on the Fahrenheit or Celsius scale
+- time of day
+
+For instance, the temperature difference between 80 and 90 degrees Fahrenheit is the same as the temperature difference between 90 and 100 degrees Fahrenheit.
+However, 0 degrees Fahrenheit does not reflect absence of temperature/heat.
+
+Interval data can be meaningfully added or subtracted.
+For instance, if a game starts at 4 pm and ends at 7 pm, you know the game lasted 3 hours ($7 - 4 = 3$).
+However, interval data cannot be meaningfully multiplied or divided.
+For instance, 100 degrees Fahrenheit is not twice as hot as 50 degrees Fahrenheit.
+
+To examine the central tendency of interval data, you can compute the mean, median, or mode.
+
+### Ratio {#sec-ratio}
+
+A variable is considered ratio if the classifications are ordered (similar to ordinal data), have equally spaced intervals (like interval data), and have an absolute zero point that reflects absence of the concept.
+Examples of ratio data include:
+
+- height
+- weight
+- age
+- distance
+- speed
+- volume
+- time elapsed
+- income
+- years of formal education
+- points in football
+
+Ratio data can be meaningfully, added, subtracted, multiplied, or divided.
+A player who weighs 350 pounds weighs twice as much as someone who weighs 175 pounds.
+
+To examine the central tendency of ratio data, you can compute the mean, median, or mode.
+
+## Descriptive Statistics
+
+## Inferential Statistics
 
@@
 
-An example statistical decision tree is depicted in Figure @fig-statisticalDecisionTree.
+An example statistical decision tree is depicted in @fig-statisticalDecisionTree.
 

### 2024-06-10 index.qmd
20240609 - ffanalytics package

@@
 
-In terms of my `R` and statistics background, I have published many [peer-reviewed publications](https://developmental-psychopathology.lab.uiowa.edu/publications) that employ advanced statistical methods, have published a [book on psychological assessment](https://www.routledge.com/9781032413068) that includes applied examples in `R`, and have published [the `petersenlab` `R` package](https://cran.r-project.org/web/packages/petersenlab/index.html) on the Comprehensive R Archive Network (CRAN).
+In terms of my `R` and statistics background, I have published many [peer-reviewed publications](https://developmental-psychopathology.lab.uiowa.edu/publications) that employ advanced statistical methods, have published a [book on psychological assessment](https://www.routledge.com/9781032413068) that includes applied examples in `R`, and have published the [`petersenlab` `R` package](https://cran.r-project.org/web/packages/petersenlab/index.html) on the Comprehensive R Archive Network (CRAN).
+I am also a co-author of the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) that provides free utilities for downloading fantasy football projections and additional fantasy-relevant data, and for calculating projected points given your league settings.
 

### 2024-06-10 basic-statistics.qmd
20240609 - clean up

@@
 A variable is considered nominal if it is composed of qualitative classifications.
-You cannot meaningfully evaluate whether one number in the variable is larger than another variable in the variable because higher numbers do not reflect higher levels of the concept.
+You cannot meaningfully evaluate whether one number in the variable is larger than another number in the variable because higher numbers do not reflect higher levels of the concept.
 Examples of nominal variables include:
@@
 - ethnicity (e.g., 0 = Non-Hispanic/Latino; 1 = Hispanic/Latino)
+- zip code
 - jersey number
-- zip code
 
@@
 
-Ratio data can be meaningfully, added, subtracted, multiplied, or divided.
+Ratio data can be meaningfully added, subtracted, multiplied, or divided.
 A player who weighs 350 pounds weighs twice as much as someone who weighs 175 pounds.

### 2024-06-12 basic-statistics.qmd
20240610 - clarification
20240612 - interactive statistical decision tree

@@
 
+For instance, points in football has order, equally spaced intervals, and an absolute zero—a team cannot score less than zero points, and zero points reflects absence of points (though it could be argued to be interval data because zero points does not reflect absence of skill.)
+
 Ratio data can be meaningfully added, subtracted, multiplied, or divided.
@@
 :::
+
+For an online, interactive statistical decision tree to help you decide which statistical analysis to use, see here: <https://www.statsflowchart.co.uk>

### 2024-06-13 player-evaluation.qmd
20240613 - code folding

@@
 ```{r}
+#| code-fold: true
+
 set.seed(1)
@@
 However, just as any other construct in psychology, we can devise ways to operationalize them.
+Insofar as they are observable, they are measurable.
 
@@
 ```{r}
+#| code-fold: true
+
 exampleData <- data.frame(
@@
 
+playerA_sd <- sd(exampleData$projectedPoints[which(exampleData$player == "A")])
+playerB_sd <- sd(exampleData$projectedPoints[which(exampleData$player == "B")])
+
+playerA_cv <- playerA_mean / playerA_sd
+playerB_cv <- playerB_mean / playerb_sd
+```
+
+```{r}
 playerA_mean
@@
 ```{r}
-playerA_sd <- sd(exampleData$projectedPoints[which(exampleData$player == "A")])
-playerb_sd <- sd(exampleData$projectedPoints[which(exampleData$player == "B")])
-
-playerA_cv <- playerA_mean / playerA_sd
-playerB_cv <- playerB_mean / playerb_sd
-
 playerA_sd
-playerb_sd
+playerB_sd
 

### 2024-06-15 basic-statistics.qmd
20240612 - statistical decision tree
20240614 - research designs

@@
 ```
+
+## Research Designs {#sec-researchDesigns}
+
+There are three broad types of research designs:
+
+- experiment
+- correlational/observational study
+- case study
+
+### Experiment {#sec-experiment}
+
+In an *experiment*, there are one or more things that we manipulate to see how the manipulation influences the process of interest.
+The variable that we manipulate is the *independent variable*.
+By contast, the *dependent variable* is the variable that we evaluate to determine whether it was influenced by the manipulation (i.e., by the independent variable).
+
+For instance, we may be interested to evaluate whether players perform better (e.g., run faster) if they drink a sports drink compared when they drink only water.
+Our hypothesis might be that players will be expected to perform better when they drink a sports drink (compared to when they drink only water).
+To this this research question and hypothesis, we might conduct an experiment by randomly assigning some players during practice to receive a sports drink and some players to receive only water.
+In this case, our independent variable is whether the player receives a sports drink.
+Our dependent variable might be their 40-yard dash time during practice.
+
+### Correlational/Observational Study {#sec-correlationalStudy}
+
+In a correlational (aka observational) study, we do not manipulate a variable to see how the manipulation influences another variable.
+Instead, we examine how two variables, a predictor and an outcome variable, are associated.
+The hypothesized cause is called the predictor variable.
+The hypothesized effect is called the outcome variable.
+In this way, the predictor variable is similar to the independent variable, and the outcome variable is similar to the dependent variable.
+However, unlike the independent and dependent variables in an experiment, the predictor and outcome variables in a correlational study are not manipulated.
+
+For instance, to use a correlational study to test the possibility that players who drink sports drinks perform better than players who drink only water, we could examine whether the players who drink sports drinks during a game score more fantasy points than players who drink only water during the game.
+In this case, our predictor variable is whether the players drinks sports drinks during a game.
+Our outcome variable is the number of fantasy points the player scored.
+
+### Case Study {#sec-caseStudy}
+
+In a case study, we assess a small sample of individuals (commonly only one person or a few people), often with rich qualitative information.
+Themes may be coded from the qualitative information, which may help inform inferences about whether some process may have played a role in influencing the outcome of interest.
+
+For instance, to use a case study to evalute the possibilty that players who drink sports drinks perform better than players who drink only water, we could conduct an in-depth interview with a player.
+In the interview, we might ask the player how they performed in games with versus without a sports drink and have them discuss whether they believe the sports drink improved their performance (and if so, how).
+Then, based on the player's responses, we might code the responses to extract themes and to make a qualitative judgement of whether or not the player likely performed better during games in which they had a sports drinl
+
+### Other Features of the Research Design: Number of Timepoints
+
+In addition to whether the research design is an [experiment](#sec-experiment), [correlational/observational study](#sec-correlationalStudy), or a [case study](#sec-caseStudy), a research design can also have one or multiple timepoints.
+The differing number of timepoints allow studies to be characterized as one of the following:
+
+- cross-sectional
+- longitudinal
+
+#### Cross-Sectional {#sec-crossSectional}
+
+A study with one timepoint is a *cross-sectional study*.
+
+#### Longitudinal Design {#sec-longitudinal}
+
+A study with more than one timepoint is a *longitudinal study*.
+When the **same measures** are assessed at each of multiple timepoints, we refer to this as a "repeated measures" design.
+
+## Research Design Validity {#sec-researchDesignValidity}
 
@@
 A Statistical Decision Tree For Choosing an Appropriate Statistical Procedure.
-Obtained from: <https://commons.wikimedia.org/wiki/File:InferentialStatisticalDecisionMakingTrees.pdf>.
+Adapted from: <https://commons.wikimedia.org/wiki/File:InferentialStatisticalDecisionMakingTrees.pdf>.
 The original source is: Corston, R., & Colman, A.
@@
 Wiley-Blackwell.
+Changes were made to the original, including the addition of several statistical tests.
 :::

### 2024-06-15 getting-started.qmd
20240608 - packages to install
20240614 - wide vs long form

@@
 
-install.packages(c("petersenlab","tidyverse","psych"))
+install.packages(
+  c("petersenlab","ffanalytics","nflreadr","nflfastR","nfl4th","nflplotR",
+  "gsisdecoder","progressr","lubridate","tidyverse","psych"))
 ```
@@
 
+Depending on the analysis, it may be important to restructure the data to be in long or wide form.
+When the data are in wide form, each player has only one row.
+When the data are in long form, each player has multiple rows—e.g., a row for each game.
+The data structure is called wide or long form because a dataset in wide form has more columns and fewer rows (i.e., it appears wider and shorter), whereas a dataset in long form has more rows and fewer columns (i.e., it appears narrower and taller).
+
 Here are the data in the `nfl_actualStats_offense_weekly` object.
-The data are structured in player-season-week form.
+The data are structured in "player-season-week form".
 That is, every row in the dataset is uniquely identified by the variables, `player_id`, `season`, and `week`.
+This is an example of long form, because each player has multiple rows.
 
@@
 
-Below, we widen the data widened by two variables (`season` and `week`), using `tidyverse`, so that the data are now in player form (where each row is uniquely identified by the `player_id` variable):
+Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "player form" (where each row is uniquely identified by the `player_id` variable):
 
@@
 ## Transform Data from Wide to Long {#sec-wideToLong}
+
+Conversely, we can also restructure data from wide to long.
 

### 2024-06-15 basic-statistics.qmd
20240615 - research methods

--- 2024-06-10 basic-statistics.qmd
@@
 ```
-
-## Levels of Measurement {#sec-levelsOfMeasurement}
-
-It is important to know the levels of measurement of your data, because the level(s) of measurement of your data constrain the types of comparisons and analyses that you can meaningfully perform.
-There are four levels of measurement that any variable can have:
-
-- nominal
-- ordinal
-- interval
-- ratio
-
-Each is described below:
-
-### Nominal {#sec-nominal}
-
-A variable is considered nominal if it is composed of qualitative classifications.
-You cannot meaningfully evaluate whether one number in the variable is larger than another number in the variable because higher numbers do not reflect higher levels of the concept.
-Examples of nominal variables include:
-
-- sex (e.g., 1 = male; 2 = female)
-- race (e.g., 1 = American Indian; 2 = Asian; 3 = Black; 4 = Pacific Islander; 5 = White)
-- ethnicity (e.g., 0 = Non-Hispanic/Latino; 1 = Hispanic/Latino)
-- zip code
-- jersey number
-
-A football player's jersey number is an example of a nominal variable.
-A jersey number of 7 is not higher on whatever concept of interest compared to a jersey number of 6.
-
-To examine the central tendency of a nominal variable, you can determine the mode, but you cannot calculate a mean or median.
-
-### Ordinal {#sec-ordinal}
-
-A variable is considered ordinal if the classifications are ordered.
-However, ordinal variables do not have equally spaced intervals.
-Examples of ordinal intervals include:
-
-- likert response scales (e.g., 1 = strongly disagree; 2 = disagree; 3 = neutral; 4 = agree; 5 = strongly agree)
-- educational attainment (e.g., 1 = no formal education; 2 = elementary school; 3 = middle school; 4 = high school; 5 = college; 6 = graduate degree)
-- academic grades on A–F scale (e.g., 1 = A; 2 = B; 3 = C; 4 = D; 5 = F)
-- player rank (1 = 1st; 2 = 2nd; 3 = 3rd, etc.)
-
-A football player's fantasy rank is an example of an ordinal variable.
-A player with a fantasy rank of 1 has a higher rank than a player with a rank of 2, but it is not known how far apart each player is—i.e., the intervals do not all reflect the same distance.
-For instance, the distance between the top-ranked player and the 2nd-best player might be 30 points, whereas the distance between the 2nd-best player and the 3rd-best player might be 2 points.
-
-To examine the central tendency of ordinal data, the median and mode are most appropriate; however, the mean may be used (unlike for nominal data).
-
-### Interval {#sec-interval}
-
-A variable is considered interval if the classifications are ordered (similar to ordinal data) and have equally spaced intervals (unlike ordinal data).
-However, interval variables do not have a meaningful zero that reflects absence.
-Examples of interval data include:
-
-- temperature on the Fahrenheit or Celsius scale
-- time of day
-
-For instance, the temperature difference between 80 and 90 degrees Fahrenheit is the same as the temperature difference between 90 and 100 degrees Fahrenheit.
-However, 0 degrees Fahrenheit does not reflect absence of temperature/heat.
-
-Interval data can be meaningfully added or subtracted.
-For instance, if a game starts at 4 pm and ends at 7 pm, you know the game lasted 3 hours ($7 - 4 = 3$).
-However, interval data cannot be meaningfully multiplied or divided.
-For instance, 100 degrees Fahrenheit is not twice as hot as 50 degrees Fahrenheit.
-
-To examine the central tendency of interval data, you can compute the mean, median, or mode.
-
-### Ratio {#sec-ratio}
-
-A variable is considered ratio if the classifications are ordered (similar to ordinal data), have equally spaced intervals (like interval data), and have an absolute zero point that reflects absence of the concept.
-Examples of ratio data include:
-
-- height
-- weight
-- age
-- distance
-- speed
-- volume
-- time elapsed
-- income
-- years of formal education
-- points in football
-
-Ratio data can be meaningfully added, subtracted, multiplied, or divided.
-A player who weighs 350 pounds weighs twice as much as someone who weighs 175 pounds.
-
-To examine the central tendency of ratio data, you can compute the mean, median, or mode.
 
@@
 Wiley-Blackwell.
+Changes were made to the original, including the addition of several statistical tests.
 :::
+
+For an online, interactive statistical decision tree to help you decide which statistical analysis to use, see here: <https://www.statsflowchart.co.uk>

### 2024-06-15 research-methods.qmd
20240615 - research methods

# Research Methods {#sec-researchMethods}

## Getting Started {#sec-researchMethodsGettingStarted}

### Load Packages {#sec-researchMethodsLoadPackages}

```{r}

```

## Research Designs {#sec-researchDesigns}

There are three broad types of research designs:

- experiment
- correlational/observational study
- case study

### Experiment {#sec-experiment}

In an *experiment*, there are one or more things that we manipulate to see how the manipulation influences the process of interest.
The variable that we manipulate is the *independent variable*.
By contast, the *dependent variable* is the variable that we evaluate to determine whether it was influenced by the manipulation (i.e., by the independent variable).

For instance, we may be interested to evaluate whether players perform better (e.g., run faster) if they drink a sports drink compared when they drink only water.
Our hypothesis might be that players will be expected to perform better when they drink a sports drink (compared to when they drink only water).
To this this research question and hypothesis, we might conduct an experiment by randomly assigning some players during practice to receive a sports drink and some players to receive only water.
In this case, our independent variable is whether the player receives a sports drink.
Our dependent variable might be their 40-yard dash time during practice.

### Correlational/Observational Study {#sec-correlationalStudy}

In a correlational (aka observational) study, we do not manipulate a variable to see how the manipulation influences another variable.
Instead, we examine how two variables, a predictor and an outcome variable, are associated.
The hypothesized cause is called the predictor variable.
The hypothesized effect is called the outcome variable.
In this way, the predictor variable is similar to the independent variable, and the outcome variable is similar to the dependent variable.
However, unlike the independent and dependent variables in an experiment, the predictor and outcome variables in a correlational study are not manipulated.

For instance, to use a correlational study to test the possibility that players who drink sports drinks perform better than players who drink only water, we could examine whether the players who drink sports drinks during a game score more fantasy points than players who drink only water during the game.
In this case, our predictor variable is whether the players drinks sports drinks during a game.
Our outcome variable is the number of fantasy points the player scored.

### Case Study {#sec-caseStudy}

In a case study, we assess a small sample of individuals (commonly only one person or a few people), often with rich qualitative information.
Themes may be coded from the qualitative information, which may help inform inferences about whether some process may have played a role in influencing the outcome of interest.

For instance, to use a case study to evalute the possibilty that players who drink sports drinks perform better than players who drink only water, we could conduct an in-depth interview with a player.
In the interview, we might ask the player how they performed in games with versus without a sports drink and have them discuss whether they believe the sports drink improved their performance (and if so, how).
Then, based on the player's responses, we might code the responses to extract themes and to make a qualitative judgement of whether or not the player likely performed better during games in which they had a sports drinl

### Other Features of the Research Design {#sec-researchDesignOtherFeatures}

#### Number of Timepoints {#sec-researchDesignTimepoints}

In addition to whether the research design is an [experiment](#sec-experiment), [correlational/observational study](#sec-correlationalStudy), or a [case study](#sec-caseStudy), a research design can also have one or multiple timepoints.
The differing number of timepoints allow studies to be characterized as one of the following:

- cross-sectional
- longitudinal

##### Cross-Sectional {#sec-crossSectional}

A study with one timepoint is a *cross-sectional study*.

##### Longitudinal Design {#sec-longitudinal}

A study with more than one timepoint is a *longitudinal study*.
When the **same measures** are assessed at each of multiple timepoints, we refer to this as a "repeated measures" design.

#### Within- or Between-Subject {#sec-withinBetweenSubject}

##### Within-Subject Design {#sec-withinSubject}

##### Between-Subject Design {#sec-betweenSubject}

## Research Design Validity {#sec-researchDesignValidity}

### Internal Validity {#sec-internalValidity}

### External Validity {#sec-externalValidity}

### Conclusion Validity {#sec-externalValidity}

## Levels of Measurement {#sec-levelsOfMeasurement}

It is important to know the levels of measurement of your data, because the level(s) of measurement of your data constrain the types of comparisons and analyses that you can meaningfully perform.
There are four levels of measurement that any variable can have:

- nominal
- ordinal
- interval
- ratio

Each is described below:

### Nominal {#sec-nominal}

A variable is considered nominal if it is composed of qualitative classifications.
You cannot meaningfully evaluate whether one number in the variable is larger than another number in the variable because higher numbers do not reflect higher levels of the concept.
Examples of nominal variables include:

- sex (e.g., 1 = male; 2 = female)
- race (e.g., 1 = American Indian; 2 = Asian; 3 = Black; 4 = Pacific Islander; 5 = White)
- ethnicity (e.g., 0 = Non-Hispanic/Latino; 1 = Hispanic/Latino)
- zip code
- jersey number

A football player's jersey number is an example of a nominal variable.
A jersey number of 7 is not higher on whatever concept of interest compared to a jersey number of 6.

To examine the central tendency of a nominal variable, you can determine the mode, but you cannot calculate a mean or median.

### Ordinal {#sec-ordinal}

A variable is considered ordinal if the classifications are ordered.
However, ordinal variables do not have equally spaced intervals.
Examples of ordinal intervals include:

- likert response scales (e.g., 1 = strongly disagree; 2 = disagree; 3 = neutral; 4 = agree; 5 = strongly agree)
- educational attainment (e.g., 1 = no formal education; 2 = elementary school; 3 = middle school; 4 = high school; 5 = college; 6 = graduate degree)
- academic grades on A–F scale (e.g., 1 = A; 2 = B; 3 = C; 4 = D; 5 = F)
- player rank (1 = 1st; 2 = 2nd; 3 = 3rd, etc.)

A football player's fantasy rank is an example of an ordinal variable.
A player with a fantasy rank of 1 has a higher rank than a player with a rank of 2, but it is not known how far apart each player is—i.e., the intervals do not all reflect the same distance.
For instance, the distance between the top-ranked player and the 2nd-best player might be 30 points, whereas the distance between the 2nd-best player and the 3rd-best player might be 2 points.

To examine the central tendency of ordinal data, the median and mode are most appropriate; however, the mean may be used (unlike for nominal data).

### Interval {#sec-interval}

A variable is considered interval if the classifications are ordered (similar to ordinal data) and have equally spaced intervals (unlike ordinal data).
However, interval variables do not have a meaningful zero that reflects absence.
Examples of interval data include:

- temperature on the Fahrenheit or Celsius scale
- time of day

For instance, the temperature difference between 80 and 90 degrees Fahrenheit is the same as the temperature difference between 90 and 100 degrees Fahrenheit.
However, 0 degrees Fahrenheit does not reflect absence of temperature/heat.

Interval data can be meaningfully added or subtracted.
For instance, if a game starts at 4 pm and ends at 7 pm, you know the game lasted 3 hours ($7 - 4 = 3$).
However, interval data cannot be meaningfully multiplied or divided.
For instance, 100 degrees Fahrenheit is not twice as hot as 50 degrees Fahrenheit.

To examine the central tendency of interval data, you can compute the mean, median, or mode.

### Ratio {#sec-ratio}

A variable is considered ratio if the classifications are ordered (similar to ordinal data), have equally spaced intervals (like interval data), and have an absolute zero point that reflects absence of the concept.
Examples of ratio data include:

- height
- weight
- age
- distance
- speed
- volume
- time elapsed
- income
- years of formal education
- points in football

For instance, points in football has order, equally spaced intervals, and an absolute zero—a team cannot score less than zero points, and zero points reflects absence of points (though it could be argued to be interval data because zero points does not reflect absence of skill.)

Ratio data can be meaningfully added, subtracted, multiplied, or divided.
A player who weighs 350 pounds weighs twice as much as someone who weighs 175 pounds.

To examine the central tendency of ratio data, you can compute the mean, median, or mode.

## Measurement Validity {#sec-measurementValidity}

### Reliability {#sec-reliability}

### Validity {#sec-validity}

### 2024-06-15 research-methods.qmd
20240615 - sample vs population; research design validity

@@
 ```
+
+## Sample vs Population {#sec-sampleVsPopulation}
+
+In research, it is important to distinguish between the sample and the target population.
+The target *population* is who you want your study's findings to generalize to.
+For instance, if we want our findings to lead to inferences we can draw regarding all current NFL players, then NFL players are our target population.
+However, despite our best efforts to recruit all NFL players into our study, we may not succeed in doing that.
+The participants (i.e., people or players) who we successfully recruit to be in our study represent our *sample*.
+
+It is rare for the sample to include all people who are in the target population.
+It can be costly to recruit large samples, and many potential participants may decline to participate for a variety of reasons (insufficient time, lack of interest in the study, distrust of scientists, etc.).
+Thus, our goals are (a) to recruit as many people from the population as possible and (b) for the sample to be as *representative* of the population as possible.
+
+For increasing the representativeness of the sample (with respect to the population), we might conduct a *random sample*, in which each person in the population (i.e., each NFL player) has equal likelihood of being selected.
+For instance, we might randomly select 250 players to recruit to the study.
+True random samples, though strong in aspiration, are difficult and costly to achieve.
+In reality, many researchers conduct convenience sampling.
+A convenience sample is recruited because it is convenient (i.e., less costly and time-consuming).
+
+For instance, many studies examine college students—in part, because they are easy to recruit.
+If our target population is NFL players but we are unable to recruit NFL players into our study, we could easily recruit a large sample of college students.
+Although the convenience sample may afford a very large sample, the college student sample may not be representative of the target population (NFL players).
+Thus, the findings in our study may not *generalize* to NFL players—that is, what we learn in college students may not apply in the same way among NFL players.
+For instance, if we learn that consumption of sports drinks (compared to drinking only water) improves running speed among college students, that may not be the case among NFL players.
 
@@
 
-A study with one timepoint is a *cross-sectional study*.
+A *cross-sectional study* is a study with one timepoint.
+
+For instance, in a cross-sectional study evaluating whether having a sports drink improves player performance, we might assess players' drinking behavior and performance during only game 1.
 
@@
 
-A study with more than one timepoint is a *longitudinal study*.
-When the **same measures** are assessed at each of multiple timepoints, we refer to this as a "repeated measures" design.
+A *longitudinal study* is a study with more than one timepoint.
+When the <u>same measures<u> are assessed at each of multiple timepoints, we refer to this as a "repeated measures" design.
+
+In a longitudinal study evaluating whether having a sports drink improves player performance, we might assess players' drinking behavior and performance during each game of the season, and possibly across multiple seasons.
 
@@
 
+A research design can also be within-subject, between-subject, or both.
+
 ##### Within-Subject Design {#sec-withinSubject}
+
+A *within-subject design* is one in which each participant (i.e., person or player) receives multiple levels of the independent variable (or predictor).
+
+For instance, in an experiment evaluating whether having a sports drink improves player performance, we might assign players to drink the sports drink in the first half of the game and to drink only water in the second half of the game.
+Or we could assign some of the players to drink sports drink in the first half and water in the second half, and assign the other players to drink water in the first half and sports drink in the second half.
+
+In a correlational study evaluating whether having a sports drink improves player performance, we might evaluate how within-person changes in sports drink consumption are associated with within-person changes in performance.
+That is, we could evaluate, when a given player has a sports drink (or more sports drinks), do they perform better than <u>when the same individual</u> has only water (or fewer sports drinks)?
 
@@
 
+A *between-subject design* is one in which each participant (i.e., person or player) receives only one level of the independent variable.
+
+For instance, in an experiment evaluating whether having a sports drink improves player performance, we might assign some players to drink the sports drink but the other players to drink only water.
+
+In a correlational study evaluating whether having a sports drink improves player performance, we might evaluate whether people who drink sports drinks tend to perform better than players who drink only water.
+Or, we could evaluate whether players who drink more sports drinks perform better than players who drink fewer sports drinks (i.e., whether the number of sports drinks consumed during a game is correlated with player performance).
+
 ## Research Design Validity {#sec-researchDesignValidity}
+
+Research design validity involves the accuracy of inferences from a study.
+There are three types of research design validity:
+
+- internal validity
+- external validity
+- conclusion validity
 
@@
 
+Internal validity is the extent to which we can be confident that the associations identified in the study are causal.
+
 ### External Validity {#sec-externalValidity}
 
+External validity is the extent to which we can be confident that findings from the study play out similarly in the real world—that is, the findings generalize to the target population.
+
 ### Conclusion Validity {#sec-externalValidity}
+
+Conclusion validity is the extent to which a study's conclusions are reasonable about the association among variables based on the data.
+That is, were the correct statistical analyses performed, and are the interpretations of the findings from those analyses correct?
 

### 2024-06-15 research-methods.qmd
20240615 - mediation/moderation

@@
 
-In an *experiment*, there are one or more things that we manipulate to see how the manipulation influences the process of interest.
+In an *experiment*, there are one or more things (i.e., variables) that we manipulate to see how the manipulation influences the process of interest.
 The variable that we manipulate is the *independent variable*.
-By contast, the *dependent variable* is the variable that we evaluate to determine whether it was influenced by the manipulation (i.e., by the independent variable).
+By contast, the *dependent variable* is the variable that we evaluate to determine whether it was influenced by the manipulation (i.e., by the independent variable).Besides the independent and dependent variables, the researcher attempts to hold everything else constant through processes including standardization and random assignment.
+*Standardization* involves using the same procedures to assess each participant, so that scores can be fairly compared across participants (and groups).
+Random assignment involves randomly assigning participants to conditions of the independent variable, so the people in each condition are comparable and do not differ systematically.
 
@@
 Themes may be coded from the qualitative information, which may help inform inferences about whether some process may have played a role in influencing the outcome of interest.
+The inferences are then drawn in a subjective, qualitative way.
+Testimonials and anecdotes are examples are case studies.
 
@@
 In the interview, we might ask the player how they performed in games with versus without a sports drink and have them discuss whether they believe the sports drink improved their performance (and if so, how).
-Then, based on the player's responses, we might code the responses to extract themes and to make a qualitative judgement of whether or not the player likely performed better during games in which they had a sports drinl
+Then, based on the player's responses, we might code the responses to extract themes and to make a qualitative judgement of whether or not the player likely performed better during games in which they had a sports drink.
 
@@
 
+Cross-sectional studies are more common than longitudinal studies because cross-sectional studies are less costly and time-consuming.
+They can provide a helpful starting point to test findings more rigorously in subsequent longitudinal studies.
+
 ##### Longitudinal Design {#sec-longitudinal}
@@
 
+Longitudinal studies are less common than cross-sectional studies because longitudinal studies are more costly and time-consuming.
+Nevertheless, longitudinal studies can allow us test our hypotheses more rigorously, because they can allow us to test whether changes in the predictor/indepdnent variable leads to changes in the outcome/dependent variable.
+Thus, compared to cross-sectional studies, longitudinal studies can provide greater confidence in causal inferences.
+
 #### Within- or Between-Subject {#sec-withinBetweenSubject}
@@
 A research design can also be within-subject, between-subject, or both.
+A study can involve both within-subject and between-subject comparisons if one predictor/independent variable is within-subject and another predictor/independent variable is between-subject.
 
@@
 
+### Tradeoffs Between Internal and External Validity {#sec-internalExternalValidityTradeoffs}
+
+There is a tradeoff between [internal](#sec-internalValidity) and [external](#sec-externalValidity) validity—a single research design cannot have both high [internal](#sec-internalValidity) and high [external validity](#sec-externalValidity).
+Each study and design has weaknesses.
+Some research designs are better suited for making causal inferences, whereas other designs tend to be better suited for making inferences that generalize to the real world.
+The research design that is best suited to making causal inferences is an [experiment](#sec-experiment) because it is the design in which the researcher has the greatest control over the variables.
+Thus, [experiments](#sec-experiment) tend to have higher [internal validity](#sec-internalValidity) than other research designs.
+However, by manipulating one variable and holding everything else constant, the research takes place in a very standardized fashion that can become like studying a process in a vacuum.
+So, even if a process is theoretically causal in a vacuum, it may act differently in the real world when it interacts with other processes.
+
+[Correlational designs](#sec-correlationalStudy) have greater capacity for [external validity](#sec-externalValidity) than [experimental designs](#sec-experiment) because the participants can be observed in their natural environments to evaluate how variables are related in the real world.
+However, the greater [external validity](#sec-externalValidity) comes at a cost of lower [internal validity](#sec-internalValidity).
+[Correlational designs](#sec-correlationalStudy) are not well-positioned to make causal inferences.
+As the maxim goes, "correlation does not mean causation"—just because two variables are associated does not necessarily mean that they are causally related.
+There could be third variables (i.e., confounds) that act as a common cause of both the predictor and outcome variables, thus explaining why they are correlated.
+
+For instance, it may not be that sport drink consumption per se influences player performance; rather, it may be that players who are more intelligent or have more financial resources tend to drink more sports drinks and also tend to perform better.
+In this case, intelligence or financial resources may be a confound that influences both sports drink consumption and player performance, but sports drink consumptions—though correlated with player performance—does not influence player performance.
+Or, the variables could be associated due to the opposite direction of effect (i.e., the outcome variable may influence the predictor variable).
+For instance, greater performance may lead players to drink more sports drink (rather than the reverse).
+[Correlational studies](#sec-correlationalStudy) can account for potential confounds using *covariates* or for the reverse direction of effect using longitudinal designs, but the researcher has less control over the variables than in an [experiment](#sec-experiment).
+
+As the [internal validity](#sec-internalValidity) of a study's design increases, its [external validity](#sec-externalValidity) tends to decrease.
+The greater control we have over variables (and, therefore, have greater confidence about causal inferences), the lower the likelihood that the findings reflect what happens in the real world because it is studying things in a metaphorical vacuum.
+Because no single research design can have both high [internal](#sec-internalValidity) and [external](#sec-externalValidity) validity, scientific inquiry needs a combination of many different research designs so we can be more confident in our inferences—[experimental designs](#sec-experiment) for making causal inferences and [correlational designs](#sec-correlationalStudy) for making inferences that are more likely to reflect the real world.
+
+[Case studies](#sec-caseStudy), because they have smaller sample sizes and inferences drawn in a subjective, qualitative way, tend to have lower [external validity](#sec-externalValidity) than both [experimental](#sec-experiment) and [correlational](#sec-correlationalStudy) studies.
+[Case studies](#sec-caseStudy) also tend to have lower [internal validity](#sec-internalValidity) because they have less control over variables, and thus fail to remove the possibility of illusory correlations, potential confounds, or the reverse direction of effect.
+Thus, [case studies](#sec-caseStudy) are among the weakest forms of evidence.
+Nevertheless, case studies can still be useful for generating hypotheses that can then be tested empirically with a larger sample in [experimental](#sec-experiment) or [correlational](#sec-correlationalStudy) studies.
+
 ### Conclusion Validity {#sec-externalValidity}
@@
 That is, were the correct statistical analyses performed, and are the interpretations of the findings from those analyses correct?
+
+## Mediation vs Moderation {#sec-mediationModeration}
+
+Both types of effects involve (at least) three variables:
+
+1. An independent/predictor variable, which will be labeled as `X`.
+1. A dependent/outcome variable, which will be labeled as `Y`.
+1. The mediator or moderator variable, which will be labeled as `M`.
+
+![Mediation Versus Moderation Mnemonic](images/mediationModerationMnemonic.png)
+
+### Mediation {#mediation}
+
+#### Overview {#mediationOverview}
+
+**Mediation** is a <u>causal chain of events</u>, where one variable (a mediator variable) at least partially <u>explains (or accounts for) the association between two other variables</u> (the predictor variable and the outcome variable).
+In mediation, a predictor (`X`) leads to a mediator (`M`), which leads to an outcome (`Y`).
+Mediation answers the question of, "<u>**Why (or how)** does `X` influence `Y`?</u> A <u>mediator</u> (`M`) is a variable that helps explain the assocation between two other variables, and it answers the question of why/how `X` influences `Y`.
+That is, the mediator is the variable that helps explain how/why `X` is related to `Y`.
+In other words, you can think of the mediator as the <u>mechanism</u> that helps explain why `X` has an impact on `Y`.
+The association between `X` and `Y` <u>gets smaller</u> when accounting for `M`.
+Visually this can be written as:
+
+![Mediation](images/mediation.png)
+
+where `X` is causing `M`, which in turn is causing `Y`.
+In other words, `X` leads to `M`, and `M` leads to `Y`.
+
+For instance, if we determine that consuming sports drinks improves player performance, we may want to know how/why.
+That is, what is the mechanism that leads consumption of sports drinks <u>to improve</u> player performance?
+We might hypothesize that consumption of sports drink helps increase a player's hydration, which in turn will improve the player's performance.
+In this case, increased hydration <u>mediates</u> (i.e., helps explain or account for) the effect of the sports drink consumption on improved player performance.
+
+Question: <u>Why/how</u> does sports drink consumpion lead players to perform better?\
+Answer: increased hydration
+
+As a picture, we can draw this assocation as:
+
+[INSERT]
+
+#### Types of Mediation {#mediationTypes}
+
+##### Full Mediation {#fullMediation}
+
+When one mechanism fully accounts for the effect of the predictor variable on the outcome variable, this is known as **full mediation**, as depicted below:
+
+![Full Mediation](images/mediation.png)
+
+##### Partial Mediation {#partialMediation}
+
+When a single process partially—but does not fully—accounts for the effect of the predictor variable on the outcome variable; this is known as **partial mediation** and is depicted below:
+
+![Partial Mediation](images/mediationPartial.png)
+
+##### Multiple Mediators {#multipleMediators}
+
+In addition, there can be multiple mediators/mechanisms that account for the effect of a predictor variable on an outcome variable, as depicted below:
+
+![Multiple Mediators](images/mediationMultiple.png)
+
+### Moderation (i.e., Interaction) {#moderation}
+
+#### Overview {#moderationOverview}
+
+**Moderation** (sometimes called an "interaction"), on the other hand, occurs when there is a variable or condition (`M`; called a "moderator") that <u>changes the assocation</u> between `X` and `Y`.
+That is, the effect of the predictor variable on the outcome variable differs at different levels of the moderator variable.
+In these cases, `X` and `M` <u>work together</u> to have an effect on `Y`; here `X` <u>does not have a direct effect</u> on `M`.
+Moderation answers the question of, "<u>**For whom** does `X` influence `Y`?</u>"
+If `X` influences `Y` more strongly for some people or in some circumstances, we would say that there is an interaction such that the <u>effect of `X` on `Y` depends on `M`</u>, as depicted below:
+
+![Moderation](images/moderation.png)
+
+For example, if the effect of consuming sports drinks on player performance differs for Quarterbacks and Wide Receivers, the interaction could be depicted below:
+
+[INSERT]
+
+An interaction can be identified visually by <u>non-parallel lines</u> at different levels of the moderator.
 

### 2024-06-15 research-methods.qmd
20240615 - mediation/moderation figures
20240615 - typo
20240615 - correlation and causation

@@
 
+#### Correlation Does Not Imply Causation {#sec-correlationCausation}
+
+As the maxim goes, "correlation does not imply causation"—just because two variables are associated does not necessarily mean that they are causally related.
+
+Just because $X$ is associated with $Y$ does not mean that $X$ causes $Y$.
+Consider that you find an association between variables $X$ and $Y$.
+
+There are several reasons why you might observe an association between $X$ and $Y$:
+
+- $X$ causes $Y$
+- $Y$ causes $X$
+- a third variable (i.e., confound), $Z$, influences both $X$ and $Y$
+- the association between $X$ and $Y$ is spurious
+
+For instance, one possibility is that the association we observed reflects our hypothesis that $X$ causes $Y$, as depicted in @fig-correlationAndCausation1.
+That is, consumption of more sports drink may improve players' performance.
+
+::: {#fig-correlationAndCausation1}
+![](images/correlationAndCausation_1.png)
+
+Hypothesized Causal Effect Based on an Observed Association Between $X$ and $Y$, Such That $X$ Causes $Y$.
+:::
+
+However, a second possibility is that the association reflects the opposite direction of effect, where $Y$ actually causes $X$, as depicted in @fig-correlationAndCausation2.
+For instance, greater performance may lead players to drink more sports drink (rather than the reverse).
+
+::: {#fig-correlationAndCausation2}
+![](images/correlationAndCausation_2.png)
+
+Reverse (Opposite) Direction of Effect From the Hypothesized Effect, Where $Y$ Causes $X$.
+:::
+
+A third possibility is that the association could reflect the influence of a third variable.
+If a third variable is a common cause of each and accounts for their association, it is a *confound*.
+An observed association between $X$ and $Y$ could reflect a confound—i.e., a cause ($Z$) that influences both $X$ and $Y$, which explains why $X$ and $Y$ are correlated even though they are not causally related.
+A third variable confound that is a common cause of both $X$ and $Y$ is depicted in @fig-correlationAndCausation3.
+For instance, it may not be that sport drink consumption per se influences player performance; rather, it may be that players who are more intelligent or have more financial resources tend to drink more sports drinks and also tend to perform better.
+In this case, intelligence or financial resources may be a confound that influences both sports drink consumption and player performance, but sports drink consumptions—though correlated with player performance—does not influence player performance.
+
+::: {#fig-correlationAndCausation3}
+![](images/correlationAndCausation_3.png)
+
+Confounded Association Between $X$ and $Y$ due to a Common Cause, $Z$.
+:::
+
+Lastly, the association might be spurious.
+It might just reflect random variation (i.e., chance), and that when tested on an independent sample, what appeared as an association in the original dataset may not hold when testing the association in a new dataset.
+
 ### Case Study {#sec-caseStudy}
@@
 A *longitudinal study* is a study with more than one timepoint.
-When the <u>same measures<u> are assessed at each of multiple timepoints, we refer to this as a "repeated measures" design.
+When the <u>same measures</u> are assessed at each of multiple timepoints, we refer to this as a "repeated measures" design.
 
@@
 [Correlational designs](#sec-correlationalStudy) are not well-positioned to make causal inferences.
-As the maxim goes, "correlation does not mean causation"—just because two variables are associated does not necessarily mean that they are causally related.
-There could be third variables (i.e., confounds) that act as a common cause of both the predictor and outcome variables, thus explaining why they are correlated.
-
-For instance, it may not be that sport drink consumption per se influences player performance; rather, it may be that players who are more intelligent or have more financial resources tend to drink more sports drinks and also tend to perform better.
-In this case, intelligence or financial resources may be a confound that influences both sports drink consumption and player performance, but sports drink consumptions—though correlated with player performance—does not influence player performance.
-Or, the variables could be associated due to the opposite direction of effect (i.e., the outcome variable may influence the predictor variable).
-For instance, greater performance may lead players to drink more sports drink (rather than the reverse).
 [Correlational studies](#sec-correlationalStudy) can account for potential confounds using *covariates* or for the reverse direction of effect using longitudinal designs, but the researcher has less control over the variables than in an [experiment](#sec-experiment).
@@
 
-![Mediation Versus Moderation Mnemonic](images/mediationModerationMnemonic.png)
+A mnemonic to help remember the difference between [mediation](#mediation) and [moderation](#moderation) is in @fig-mediationModerationMnemonic.
+
+::: {#fig-mediationModerationMnemonic}
+![](images/mediationModerationMnemonic.png)
+
+Mediation Versus Moderation Mnemonic.
+:::
 
@@
 The association between `X` and `Y` <u>gets smaller</u> when accounting for `M`.
-Visually this can be written as:
+Visually this can be written as in @#fig-mediation:
 
-![Mediation](images/mediation.png)
+::: {#fig-mediation}
+![](images/mediation.png)
+
+Mediation.
+:::
 
@@
 
-As a picture, we can draw this assocation as:
+As a picture, we can draw this assocation as in @fig-mediationExample:
 
-[INSERT]
+::: {#fig-mediationExample}
+![](images/mediationExample.png)
+
+Mediation Example.
+:::
 
@@
 
-When one mechanism fully accounts for the effect of the predictor variable on the outcome variable, this is known as **full mediation**, as depicted below:
+When one mechanism fully accounts for the effect of the predictor variable on the outcome variable, this is known as **full mediation**, as depicted in @fullMediation:
 
-![Full Mediation](images/mediation.png)
+::: {#fig-fullMediation}
+![](images/mediation.png)
+
+Full Mediation.
+:::
 
@@
 
-When a single process partially—but does not fully—accounts for the effect of the predictor variable on the outcome variable; this is known as **partial mediation** and is depicted below:
+When a single process partially—but does not fully—accounts for the effect of the predictor variable on the outcome variable; this is known as **partial mediation** and is depicted in @fig-fullMediation:
 
-![Partial Mediation](images/mediationPartial.png)
+::: {#fig-fullMediation}
+![](images/mediationPartial.png)
+
+Partial Mediation.
+:::
 
@@
 
-In addition, there can be multiple mediators/mechanisms that account for the effect of a predictor variable on an outcome variable, as depicted below:
+In addition, there can be multiple mediators/mechanisms that account for the effect of a predictor variable on an outcome variable, as depicted in @fig-multipleMediators:
 
-![Multiple Mediators](images/mediationMultiple.png)
+::: {#fig-multipleMediators}
+![](images/mediationMultiple.png)
+
+Multiple Mediators.
+:::
 
@@
 Moderation answers the question of, "<u>**For whom** does `X` influence `Y`?</u>"
-If `X` influences `Y` more strongly for some people or in some circumstances, we would say that there is an interaction such that the <u>effect of `X` on `Y` depends on `M`</u>, as depicted below:
+If `X` influences `Y` more strongly for some people or in some circumstances, we would say that there is an interaction such that the <u>effect of `X` on `Y` depends on `M`</u>, as depicted in @fig-moderation:
 
-![Moderation](images/moderation.png)
+::: {#fig-moderation}
+![](images/moderation.png)
 
-For example, if the effect of consuming sports drinks on player performance differs for Quarterbacks and Wide Receivers, the interaction could be depicted below:
+Moderation.
+:::
 
-[INSERT]
+For example, if the effect of consuming sports drinks on player performance differs for Quarterbacks and Wide Receivers, the interaction could be depicted in @moderationExamplePathDiagram and @fig-moderationExampleInteractionGraph:
+
+::: {#fig-moderationExamplePathDiagram}
+![](images/moderationExample1.png)
+
+Moderation Example: Path Diagram.
+:::
+
+::: {#fig-moderationExampleInteractionGraph}
+![](images/moderationExample2.png)
+
+Moderation Example: Interaction Graph.
+:::
 

### 2024-06-15 research-methods.qmd
20240615 - typos

@@
 
-### Conclusion Validity {#sec-externalValidity}
+### Conclusion Validity {#sec-conclusionValidity}
 
@@
 
-When one mechanism fully accounts for the effect of the predictor variable on the outcome variable, this is known as **full mediation**, as depicted in @fullMediation:
+When one mechanism fully accounts for the effect of the predictor variable on the outcome variable, this is known as **full mediation**, as depicted in @fig-fullMediation:
 
@@
 
-When a single process partially—but does not fully—accounts for the effect of the predictor variable on the outcome variable; this is known as **partial mediation** and is depicted in @fig-fullMediation:
+When a single process partially—but does not fully—accounts for the effect of the predictor variable on the outcome variable; this is known as **partial mediation** and is depicted in @fig-partialMediation:
 
-::: {#fig-fullMediation}
+::: {#fig-partialMediation}
 ![](images/mediationPartial.png)
@@
 
-For example, if the effect of consuming sports drinks on player performance differs for Quarterbacks and Wide Receivers, the interaction could be depicted in @moderationExamplePathDiagram and @fig-moderationExampleInteractionGraph:
+For example, if the effect of consuming sports drinks on player performance differs for Quarterbacks and Wide Receivers, the interaction could be depicted in @fig-moderationExamplePathDiagram and @fig-moderationExampleInteractionGraph:
 

### 2024-06-15 research-methods.qmd
20240615 - typo
20240615 - moderation clarification

@@
 The association between `X` and `Y` <u>gets smaller</u> when accounting for `M`.
-Visually this can be written as in @#fig-mediation:
+Visually this can be written as in @fig-mediation:
 
@@
 An interaction can be identified visually by <u>non-parallel lines</u> at different levels of the moderator.
+In this example, the player's position moderates the effect consuming sports drinks on player performance.
+In particular, there is a <u>strong positive association</u> between consuming sports drinks and player performance for Wide Receivers (as evidenced by the upward slope of the best-fit regression line), whereas there is <u>no association</u> between consuming sports drinks and player performance for Quarterbacks (as evidenced by the flat line).
 

### 2024-06-16 research-methods.qmd
20240615 - relabel
20240615 - reliability, validity

@@
 
-## Measurement Validity {#sec-measurementValidity}
+## Psychometrics {#sec-psychometrics}
 
-### Reliability {#sec-reliability}
+### Measurement Reliability {#sec-reliability}
 
-### Validity {#sec-validity}
+The *reliability* of a measure's scores deals with the *consistency* of measurement.
+This book focuses on the following types of reliability:
+
+- [test–retest reliability](#sec-testRetestReliability)
+- [inter-rater reliability](#sec-interraterReliability)
+- [intra-rater reliability](#sec-intraraterReliability)
+- [internal consistency](#sec-internalConsistency)
+- [parallel-forms reliability](#sec-parallelFormsReliability)
+
+#### Test–Retest Reliability {#sec-testRetestReliability}
+
+Test–retest reliability evaluates the consistency of scores across time.
+For a construct that is expected to be stable across time (e.g., hand size in adults), we would expect our measurements to be consistent across time.
+The consistency of scores across time can be examined in terms of relative or absolute test–retest reliability.
+Relative test–retest reliability—i.e., the consistency of individual differences across time—is commonly evaluated using the coefficient of stability (i.e., the Pearson correlation coefficent).
+Absolute test–retest reliability—i.e., the absolute consistency of people's scores across time—is commonly evaluated using the coefficient of repeatability.
+
+#### Inter-Rater Reliability {#sec-interraterReliability}
+
+Inter-rater reliability evaluates the consistency of scores across raters.
+For instance, if we have a strong measure for assessing college players' aptitude to succeed in the NFL, the measure should yield a similar score for a given player regardless of which (trained) rater (e.g., coach) uses it to rate the player.
+The consistency of scores across raters is commonly evaluated using the intraclass correlation coefficient (for continuous variables) and Cohen's kappa ($\kappa$; for categorical variables).
+
+#### Intra-Rater Reliability {#sec-intraraterReliability}
+
+Intra-rater reliability evaluates the consistency of scores within a given rater.
+If we have a strong measure for assessing college players' aptitude to succeed in the NFL, the measure should yield a similar score for a given player from the same (trained) rater (e.g., coach) each time they rate the same player.
+The consistency of scores within raters can be evaluated using similar approaches as those evaluating [inter-rater reliability](#sec-interraterReliability).
+
+#### Internal Consistency {#sec-internalConsistency}
+
+Internal consistency evaluates the consistency of scores across items within a measure.
+If we develop a strong questionnaire measure to assess a college players' aptitude to succeed in the NFL, the scores should be relatively consistent across items.
+The consistency of scores across items within a measure is commonly evaluated using Cronbach's alpha ($\alpha$) or McDonald's omega ($\omega$).
+
+#### Parallel-Forms Reliability {#sec-parallelFormsReliability}
+
+Parallel-forms reliability evaluates the consistency of scores across different but equivalent forms of a measure.
+If we develop two equivalent versions of the Wonderlic Contemporary Cognitive Ability Test (Form A and Form B) so that players sitting next to each other do not receive the same items, we would expect a player's score on Form A would be similar to their score on Form B.
+Parallel-forms reliability is is commonly evaluated using the coefficient of equivalence (i.e., the Pearson correlation coefficent).
+
+### Measurement Validity {#sec-validity}
+
+The *validity* of a measure's scores deals with the *accuracy* of measurement.
+This book focuses on the following types of validity:
+
+- [face validity](#sec-faceValidity)
+- [content validity](#sec-contentValidity)
+- [criterion-related validity](#sec-criterionValidity)
+    - [concurrent (criterion-related) validity](#sec-concurrentValidity)
+    - [predictive (criterion-related) validity](#sec-predictiveValidity)
+- [construct validity](#sec-constructValidity)
+- [convergent validity](#sec-convergentValidity)
+- [discriminant validity](#sec-discriminantValidity)
+- [incremental validity](#sec-incrementalValidity)
+
+#### Face Validity {#sec-faceValidity}
+
+Face validity evaluates the extent to which a measure "looks like" (on its face) it assesses the construct of interest.
+For instance, if a measure is developed to assess aptitude of Wide Receivers for the position, it would be considered to have face validity if everyday (lay) people believe that it assesses aptitude for being a successful Wide Receiver.
+
+#### Content Validity {#sec-contentValidity}
+
+Content validity evaluates the extent to which the measure assesses the full breadth of the content, as determined by context experts.
+For the measure to have content validity, it should not have gaps (missing content facets) or intrusions (facets of other constructs).
+For instance, a strong measure for assessing a player's aptitude to succeed in the NFL might need to include a player's speed, strength, size, lateral quickness, etc.
+If the measure is missing their speed, this would be a content gap.
+If the measure assesses a construct-irrelevant facet (e.g., their attractiveness), this would be a content intrusion.
+
+#### Criterion-Related Validity {#sec-criterionValidity}
+
+Criterion-related validity evaluates the extent to which the measure's scores are related to meaningful variables of interest.
+Criterion-related validity is commonly evaluated using a Pearson correlation or some form of regression.
+
+There are two types of criterion-related validity:
+
+- concurrent (criterion-related) validity
+- predictive (criterion-related) validity
+
+##### Concurrent (Criterion-Related) Validity {#sec-concurrentValidity}
+
+Concurrent criterion-related validity evaluates the extent to which the measure's scores are related to meaningful variables of interest assessed at the same point in time.
+That is, concurrent validity could evaluate whether current player statistics (e.g., passing yards) are associated with their fantasy points.
+
+##### Predictive (Criterion-Related) Validity {#sec-predictiveValidity}
+
+Predictive criterion-related validity evaluates the extent to which the measure's scores are related to meaningful variables of interest that are assessed at a later point in time.
+For example, predictive validity could evaluate whether scores on the measure we developed to assess a player's aptitude to succeed in the NFL predicts later performance in the NFL.
+
+#### Construct Validity {#sec-constructValidity}
+
+Construct validity evaluates the extent to which the measure's scores accurately assess the construct of interest.
+If we develop a measure with intent to assess aptitude for being a successful Running Back, and it appears to more accurately assess aptitude for being a successful Wide Receiver, then our measure has poor construct validity for assessing aptitude for being a successful Running Back.
+Construct validity subsumes [convergent](#sec-convergentValidity) and [discriminant](##sec-discriminantValidity) validity, in addition to all of the other forms of measurement validity.
+
+#### Convergent Validity {#sec-convergentValidity}
+
+Convergent validity evaluates the extent to which the measure's scores are related to other measures of the same construct.
+For instance, if we develop a new measure to assess intelligence, its scores should be related to scores from other measures designed to assess intelligence (e.g., Wonderlic Contemporary Cognitive Ability Test).
+
+#### Discriminant Validity {#sec-discriminantValidity}
+
+Discriminant validity evaluates the extent to which the measure's scores are unrelated to measures of the different constructs.
+For instance, if we develop a new measure to assess intelligence, its scores should be less strongly associated with measures of other constructs (e.g., measures of happiness).
+
+#### Incremental Validity {#sec-incrementalValidity}
+
+Incremental validity evalutes the extent to which the measure's scores provide an increase in predictive accuracy compared to other information that is easily and cheaply available.
+That is, in order to be useful, a strong measure should tell us something that we did not already know.
+For instance, if we develop a strong measure of intelligence, it should result in increased predictive accuracy (for success in the NFL) compared to when just relying on the Wonderlic Contemporary Cognitive Ability Test.
+
+### Reliability vs Validity {#sec-reliabilityVsValidity}
+
+Reliability and validity are different but related.
+Reliability refers to the *consistency* of scores, whereas accuracy refers to the *accuracy* of scores.
+Validity depends on reliability.
+Reliability is necessary—but insufficient for—validity.
+That is, consistency is necessary—but insufficient for—accuracy.
+As depicted in @fig-reliabilityVsValidity, a measure can be no more valid than it is reliable.
+
+::: {#fig-reliabilityVsValidity}
+![](images/reliabilityVsValidity.png)
+
+Reliability Versus Validity.
+:::

### 2024-06-16 index.qmd
20240615 - CITES

@@
 
-In terms of my `R` and statistics background, I have published many [peer-reviewed publications](https://developmental-psychopathology.lab.uiowa.edu/publications) that employ advanced statistical methods, have published a [book on psychological assessment](https://www.routledge.com/9781032413068) that includes applied examples in `R`, and have published the [`petersenlab` `R` package](https://cran.r-project.org/web/packages/petersenlab/index.html) on the Comprehensive R Archive Network (CRAN).
+In terms of my `R` and statistics background, I have published many [peer-reviewed publications](https://developmental-psychopathology.lab.uiowa.edu/publications) that employ advanced statistical methods, have published a [book on psychological assessment](https://www.routledge.com/9781032413068) [@Petersen2024a; @PetersenPrinciplesPsychAssessment] that includes applied examples in `R`, and have published the [`petersenlab` `R` package](https://cran.r-project.org/web/packages/petersenlab/index.html) [@R-petersenlab] on the Comprehensive R Archive Network (CRAN).
 I am also a co-author of the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) that provides free utilities for downloading fantasy football projections and additional fantasy-relevant data, and for calculating projected points given your league settings.

### 2024-06-16 research-methods.qmd
20240615 - CITES

@@
 
+Below, I provide brief discussions of various aspects of measurement reliability and validity.
+For more information on these and other aspects of psychometrics, see @Petersen2024a and @PetersenPrinciplesPsychAssessment.
+
 ### Measurement Reliability {#sec-reliability}
@@
 - [parallel-forms reliability](#sec-parallelFormsReliability)
+
+For more information on these and other aspects of reliability, see <https://isaactpetersen.github.io/Principles-Psychological-Assessment/reliability.html> [@Petersen2024a; @PetersenPrinciplesPsychAssessment].
 
@@
 - [incremental validity](#sec-incrementalValidity)
+
+For more information on these and other aspects of validity, see <https://isaactpetersen.github.io/Principles-Psychological-Assessment/validity.html> [@Petersen2024a; @PetersenPrinciplesPsychAssessment].
 

### 2024-06-16 research-methods.qmd
20240615 - clarification

@@
 Inter-rater reliability evaluates the consistency of scores across raters.
-For instance, if we have a strong measure for assessing college players' aptitude to succeed in the NFL, the measure should yield a similar score for a given player regardless of which (trained) rater (e.g., coach) uses it to rate the player.
+For instance, if we have a strong measure for assessing college players' aptitude to succeed in the NFL, the measure should yield a similar score for a given player regardless of which (trained) rater (e.g., coach or talent scout) uses it to rate the player.
 The consistency of scores across raters is commonly evaluated using the intraclass correlation coefficient (for continuous variables) and Cohen's kappa ($\kappa$; for categorical variables).
@@
 Intra-rater reliability evaluates the consistency of scores within a given rater.
-If we have a strong measure for assessing college players' aptitude to succeed in the NFL, the measure should yield a similar score for a given player from the same (trained) rater (e.g., coach) each time they rate the same player.
+If we have a strong measure for assessing college players' aptitude to succeed in the NFL, the measure should yield a similar score for a given player from the same (trained) rater (e.g., coach or talent scout) each time they rate the same player (assuming the player's aptitude has not changed).
 The consistency of scores within raters can be evaluated using similar approaches as those evaluating [inter-rater reliability](#sec-interraterReliability).
@@
 As depicted in @fig-reliabilityVsValidity, a measure can be no more valid than it is reliable.
+A measure can be consistent but inaccurate; however, a measure cannot be accurate but inconsistent.
 

### 2024-06-16 basic-statistics.qmd
20240615 - sample vs population; research design validity
20240615 - placeholders
20240616 - basic descriptive and inferential stats

@@
 
-## Descriptive Statistics
+## Descriptive Statistics {#sec-descriptiveStatistics}
 
-## Inferential Statistics
+Descriptive statistics are used to describe data.
+For instance, they may be used to describe the center, spread, or shape of the data.
 
-## Statistical Decision Tree
+There are various indices of each.
+Indices to describe the center of a variable's data include:
+
+- mean
+- median
+- mode
+
+Indices to describe the center of a variable's data include:
+
+- standard deviation
+- variance
+
+Indices to describe the shape of a variable's data include:
+
+- skewness
+- kurtosis
+
+## Inferential Statistics {#sec-inferentialStatistics}
+
+Inferential statistics are used to draw inferences regarding whether there is (a) a difference in level on variable across groups or (b) an association between variables.
+For instance, inferential statistics may be used to evaluate whether Quarterbacks tend to have longer careers compared to Running Backs.
+Or, they could be used to evaluate whether number of carries is associated with injury likelihood.
+To apply inferential statistics, we make use of the null hypothesis ($H_0$) and the alternative hypothesis ($H_1$).
+
+### Null Hypothesis ($H_0$)
+
+When testing whether there are differences in level across groups on a variable of interest, the null hypothesis ($H_0$) is that there is <u>no difference</u> in level across groups.
+For instance, when testing whether Quarterbacks tend to have longer careers compared to Running Backs, the null hypothesis ($H_0$) is that Quarterbacks do not systematically differ from Running Backs in the length of their career.
+
+When testing whether there is an association between variables, the null hypothesis ($H_0$) is that there is <u>no association</u> between the variables.
+For instance, when testing whether number of carries is associated with injury likelihood, the null hypothesis ($H_0$) is that there is no association between number of carries and injury likelihood.
+
+### Alternative Hypothesis ($H_1$)
+
+The alternative hypothesis ($H_1$) is the researcher's hypothesis that they want to evaluate.
+An alternative hypothesis ($H_1$) might directional or non-directional.
+
+Directional hypotheses specify a particular direction, such as which group will have larger scores or which direction (positive or negative) two variables will be associated.
+Examples of directional hypotheses include:
+
+- Quarterbacks have <u>longer</u> careers compared to Running Backs
+- Number of carries is <u>positively</u> associated with injury likelihood
+
+Non-directional hypotheses do not specify a particular direction.
+For instance, non-directional hypotheses may state that two groups differ but do not specify which group will have larger scores.
+Or, non-directional hypotheses may state that two variables are associated but do not state what the sign is of the association—i.e., positive or negative.
+Examples of non-directional hypotheses include:
+
+- Quarterbacks <u>differ</u> in the length of their careers compared to Running Backs
+- Number of carries is <u>associated</u> with injury likelihood
+
+### Statistical Significance {#sec-statisticalSignificance}
+
+In science, statistical significance is evaluated with the *p*-value.
+The *p*-value represents a conditional probability—it examines the probability of one event given another event.
+In particular, the *p*-value evaluates the likelihood that you would detect a result as extreme as (or more extreme than) the one observed (in terms of the magnitude of the difference or of the association) given that the null hypothesis ($H_0$) is true.
+
+This can be expressed as:
+
+Left-tailed directional test (i.e., Quarterbacks have <u>shorter</u> careers than Running Backs; or number of carries is <u>negatively</u> associated with injury likelihood):
+$$
+p\text{-value} = Pr(T \le t | H_0)
+$$
+
+Right-tailed directional test (i.e., Quarterbacks have <u>longer</u> careers than Running Backs; or number of carries is <u>positively</u> associated with injury likelihood):
+
+$$
+p\text{-value} = Pr(T \ge t | H_0)
+$$
+
+Two-tailed non-directional test (i.e., Quarterbacks <u>differ</u> in the length of their careers compared to Running Backs; or number of carries is <u>associated</u> with injury likelihood):
+
+$$
+p\text{-value} = 2 \times Pr(T \ge |t| | H_0)
+$$
+
+where $T$ is the test statistic of interest (e.g., $t$-, $r-$, $F$-, or $B$-coefficient, depending on the test), and $t$ is the observed test statistic.
+
+That is, for a test of group differences, the *p*-value evaluates the likelihood that you would observe a difference as large or larger than the one you observed between the groups if there were no systematic difference between the groups.
+For instance, when evaluating whether Quarterbacks have <u>longer</u> careers than Running Backs, and you observed a mean difference of 0.03 years, the *p*-value evaluates the likelihood that you would observe a difference as larger or larger than 0.03 years between the groups if Quarterbacks do not differ from Running Backs in terms of the length of their career.
+
+That is, for a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed between the groups if there were no association between the variables.
+For instance, when evaluating whether number of carries is <u>positively</u> associated with injury likelihood, and you observed a correlation coefficient of $r = .25$ between number of carries and injury likelihood, the *p*-value evaluates the likelihood that you would observe a correlation as strong or stronger than $r = .25$ between the variables if number of carries is not associated with injury likelihood.
+
+### Practical Significance {#sec-practicalSignificance}
+
+## Statistical Decision Tree {#sec-statisticalDecisionTree}
 

### 2024-06-16 basic-statistics.qmd
20240616 - statistical and practical significance

@@
 
-where $T$ is the test statistic of interest (e.g., $t$-, $r-$, $F$-, or $B$-coefficient, depending on the test), and $t$ is the observed test statistic.
+where $T$ is the test statistic of interest (e.g., $t$-, $r-$, $F$-, or $B$-coefficient, depending on the test), $t$ is the observed test statistic, and $Pr$ is the probability of the event.
 
@@
 
+In science, we typically consider an effect to be *statistically significant* if the *p*-value is less than .05.
+That is, there is a small chance (5%; or 1 in 20 times) that we would observe an effect as extreme as (or more extreme than) the effect observed, if the null hypothesis were true.
+Statistical significance involves the *consistency* of an effect/association/difference; it suggests that the association/difference is reliably non-zero.
+However, just because something is statistically significant does not mean that it is important.
+In addition to statistical significance, it is also important to consider practical significance.
+
 ### Practical Significance {#sec-practicalSignificance}
+
+*Practical significance* deals with how important the effect/association/difference is.
+It is based on the magnitude of the effect, called the *effect size*.
+Effect size can be quantified in various ways including:
+
+- Cohen's $d$
+- Standardized regression coefficient ($\beta$)
+- Correlation coefficient ($r$)
+- Coefficient of determination ($R^2$)
+- Eta squared ($\eta^2$)
+- Partial eta squared ($\eta_p^2$)
+
+| Effect Size Index                             | Small        | Medium       | Large        |
+|-----------------------------------------------|--------------|--------------|--------------|
+| Cohen's d                                     | $\ge |.20|$ | $\ge |.50|$ | $\ge |.80|$ |
+| Standardized regression coefficient ($\beta$) | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
+| Correlation coefficient ($r$)                 | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
+| Coefficient of determination ($R^2$)          | $\ge .01$   | $\ge .06$   | $\ge .14$   |
+| Eta squared ($\eta^2$)                        | $\ge .01$   | $\ge .06$   | $\ge .14$   |
+| Partial eta squared ($\eta_p^2$)              | $\ge .01$   | $\ge .06$   | $\ge .14$   |
+
+: Effect Size Thresholds for Small, Medium, and Large Effect Sizes. {#tbl-effectSize Thresholds}
 

### 2024-06-16 basic-statistics.qmd
20240616 - table formatting

@@
 | Effect Size Index                             | Small       | Medium      | Large       |
-|-----------------------------------------------|--------------|--------------|--------------|
+|:----------------------------------------------|:------------|:------------|:------------|
 | Cohen's d                                     | $\ge |.20|$ | $\ge |.50|$ | $\ge |.80|$ |

### 2024-06-16 basic-statistics.qmd
20240616 - clarification

@@
 - Cohen's $d$
-- Standardized regression coefficient ($\beta$)
+- Standardized regression coefficient (beta; $\beta$)
 - Correlation coefficient ($r$)

### 2024-06-17 basic-statistics.qmd
20240616 - clarifications

@@
 In science, statistical significance is evaluated with the *p*-value.
+The *p*-value does not represent the probability that you observed the result by chance.
 The *p*-value represents a conditional probability—it examines the probability of one event given another event.
-In particular, the *p*-value evaluates the likelihood that you would detect a result as extreme as (or more extreme than) the one observed (in terms of the magnitude of the difference or of the association) given that the null hypothesis ($H_0$) is true.
+In particular, the *p*-value evaluates the likelihood that you would detect a result as at least as extreme as the one observed (in terms of the magnitude of the difference or of the association) given that the null hypothesis ($H_0$) is true.
 
-This can be expressed as:
+This can be expressed in conditional probability notation, $Pr(A | B)$, which is the probability (likelihood) of event A occurring given that event B occurred (or given condition B).
 
-Left-tailed directional test (i.e., Quarterbacks have <u>shorter</u> careers than Running Backs; or number of carries is <u>negatively</u> associated with injury likelihood):
+Below is conditional probability notation for a left-tailed directional test (i.e., Quarterbacks have <u>shorter</u> careers than Running Backs; or number of carries is <u>negatively</u> associated with injury likelihood):
+
 $$
@@
 
-Right-tailed directional test (i.e., Quarterbacks have <u>longer</u> careers than Running Backs; or number of carries is <u>positively</u> associated with injury likelihood):
+where $T$ is the test statistic of interest (e.g., $t$-, $r-$, $F$-, or $B$-coefficient, depending on the test) and $t$ is the observed test statistic.
+
+Below is conditional probability notation for a right-tailed directional test (i.e., Quarterbacks have <u>longer</u> careers than Running Backs; or number of carries is <u>positively</u> associated with injury likelihood):
 
@@
 
-Two-tailed non-directional test (i.e., Quarterbacks <u>differ</u> in the length of their careers compared to Running Backs; or number of carries is <u>associated</u> with injury likelihood):
+Below is conditional probability notation for a two-tailed non-directional test (i.e., Quarterbacks <u>differ</u> in the length of their careers compared to Running Backs; or number of carries is <u>associated</u> with injury likelihood):
+
+$$
+p\text{-value} = 2 \times \text{min}(Pr(T \le t | H_0), Pr(T \ge t | H_0))
+$$
+
+where `min(a, b)` is the smaller number of `a` and `b`.
+
+If the distribution of the test statistic is symmetric around zero, the *p*-value for the two-tailed non-directional test simplifies to:
 
@@
 $$
-
-where $T$ is the test statistic of interest (e.g., $t$-, $r-$, $F$-, or $B$-coefficient, depending on the test), $t$ is the observed test statistic, and $Pr$ is the probability of the event.
 
@@
 In science, we typically consider an effect to be *statistically significant* if the *p*-value is less than .05.
-That is, there is a small chance (5%; or 1 in 20 times) that we would observe an effect as extreme as (or more extreme than) the effect observed, if the null hypothesis were true.
+That is, there is a small chance (5%; or 1 in 20 times) that we would observe an effect at least as extreme as the effect observed, if the null hypothesis were true.
 Statistical significance involves the *consistency* of an effect/association/difference; it suggests that the association/difference is reliably non-zero.

### 2024-06-17 basic-statistics.qmd
20240616 - number and cross-reference equations

@@
 
-Below is conditional probability notation for a left-tailed directional test (i.e., Quarterbacks have <u>shorter</u> careers than Running Backs; or number of carries is <u>negatively</u> associated with injury likelihood):
+The conditional probability notation for a left-tailed directional test (i.e., Quarterbacks have <u>shorter</u> careers than Running Backs; or number of carries is <u>negatively</u> associated with injury likelihood) is in @eq-pvalueLeftTailed.
 
@@
 p\text{-value} = Pr(T \le t | H_0)
-$$
+$$ {#eq-pvalueLeftTailed}
 
@@
 
-Below is conditional probability notation for a right-tailed directional test (i.e., Quarterbacks have <u>longer</u> careers than Running Backs; or number of carries is <u>positively</u> associated with injury likelihood):
+The conditional probability notation for a right-tailed directional test (i.e., Quarterbacks have <u>longer</u> careers than Running Backs; or number of carries is <u>positively</u> associated with injury likelihood) is in @eq-pvalueRightTailed.
 
@@
 p\text{-value} = Pr(T \ge t | H_0)
-$$
+$$ {#eq-pvalueRightTailed}
 
-Below is conditional probability notation for a two-tailed non-directional test (i.e., Quarterbacks <u>differ</u> in the length of their careers compared to Running Backs; or number of carries is <u>associated</u> with injury likelihood):
+The conditional probability notation for a two-tailed non-directional test (i.e., Quarterbacks <u>differ</u> in the length of their careers compared to Running Backs; or number of carries is <u>associated</u> with injury likelihood) is in @eq-pvalueTwoTailed.
 
@@
 p\text{-value} = 2 \times \text{min}(Pr(T \le t | H_0), Pr(T \ge t | H_0))
-$$
+$$ {#eq-pvalueTwoTailed}
 
@@
 
-If the distribution of the test statistic is symmetric around zero, the *p*-value for the two-tailed non-directional test simplifies to:
+If the distribution of the test statistic is symmetric around zero, the *p*-value for the two-tailed non-directional test simplifies to @eq-pvalueTwoTailedSimple.
 
@@
 p\text{-value} = 2 \times Pr(T \ge |t| | H_0)
-$$
+$$ {#eq-pvalueTwoTailedSimple}
 

### 2024-06-17 player-evaluation.qmd
20240613 - fix typo
20240616 - number and cross-reference equations

@@
 So, we can account for the player's mean fantasy points per game by dividing their game-to-game standard deviation of fantasy points ($\sigma$) by their mean fantasy points across games ($\mu$).
-This is known as the coefficient of variation (CV):
+This is known as the coefficient of variation (CV), which is provided in @eq-coefficientOfVariation.
 
@@
 CV = \frac{\sigma}{\mu}
-$$
+$$ {#eq-coefficientOfVariation}
 
@@
 playerA_cv <- playerA_mean / playerA_sd
-playerB_cv <- playerB_mean / playerb_sd
+playerB_cv <- playerB_mean / playerB_sd
 ```

### 2024-06-17 basic-statistics.qmd
20240616 - typo
20240616 - grammar
20240616 - emphasis

@@
 There are various indices of each.
-Indices to describe the center of a variable's data include:
+Indices to describe the *center* of a variable's data include:
 
@@
 
-Indices to describe the center of a variable's data include:
+Indices to describe the *variability* of a variable's data include:
 
@@
 
-Indices to describe the shape of a variable's data include:
+Indices to describe the *shape* of a variable's data include:
 
@@
 The alternative hypothesis ($H_1$) is the researcher's hypothesis that they want to evaluate.
-An alternative hypothesis ($H_1$) might directional or non-directional.
+An alternative hypothesis ($H_1$) might be directional (i.e., one-sided) or non-directional (i.e., two-sided).
 
@@
 Adapted from: <https://commons.wikimedia.org/wiki/File:InferentialStatisticalDecisionMakingTrees.pdf>.
-The original source is: Corston, R., & Colman, A.
+The original source is: Corston, R.
+& Colman, A.
 M.

### 2024-06-17 basic-statistics.qmd
20240616 - consistency
20240616 - clarification

@@
 
-Indices to describe the *variability* of a variable's data include:
+Indices to describe the *spread* (variability) of a variable's data include:
 
@@
 
-where $T$ is the test statistic of interest (e.g., $t$-, $r-$, $F$-, or $B$-coefficient, depending on the test) and $t$ is the observed test statistic.
+where $T$ is the test statistic of interest (e.g., the distribution of $t$-, $r-$, or $F$ values, depending on the test) and $t$ is the observed test statistic (e.g., $t$-, $r-$, or $F$-coefficient, depending on the test).
 

### 2024-06-17 basic-statistics.qmd
20240616 - range and min/max
20240617 - cite
20240617 - italics

@@
 - variance
+- range
+- minimum and maximum
 
@@
 
+Effect size thresholds for small, medium, and large effect sizes are in @tbl-effectSizeThresholds [@McGrath2006].
+
 | Effect Size Index                             | Small       | Medium      | Large       |
 |:----------------------------------------------|:------------|:------------|:------------|
-| Cohen's d                                     | $\ge |.20|$ | $\ge |.50|$ | $\ge |.80|$ |
+| Cohen's $d$                                   | $\ge |.20|$ | $\ge |.50|$ | $\ge |.80|$ |
 | Standardized regression coefficient ($\beta$) | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
@@
 
-: Effect Size Thresholds for Small, Medium, and Large Effect Sizes. {#tbl-effectSize Thresholds}
+: Effect Size Thresholds for Small, Medium, and Large Effect Sizes. {#tbl-effectSizeThresholds}
 

### 2024-06-17 basic-statistics.qmd
20240617 - table

@@
 | Effect Size Index                                   | Small       | Medium      | Large       |
-|:----------------------------------------------|:------------|:------------|:------------|
+|:----------------------------------------------------|:------------|:------------|:------------|
 | Cohen's $d$                                         | $\ge |.20|$ | $\ge |.50|$ | $\ge |.80|$ |
-| Standardized regression coefficient ($\beta$) | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
+| Standardized regression coefficient (beta; $\beta$) | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
 | Correlation coefficient ($r$)                       | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |

### 2024-06-18 basic-statistics.qmd
20240617 - placeholders

@@
 ```{r}
-
+library("petersenlab")
+library("tidyverse")
 ```
@@
 For instance, they may be used to describe the center, spread, or shape of the data.
+There are various indices of each.
 
-There are various indices of each.
-Indices to describe the *center* of a variable's data include:
+### Center {#sec-descriptiveStatisticsCenter}
+
+Indices to describe the *center* (central tendency) of a variable's data include:
 
@@
 - mode
+
+Below is R code to estimate each:
+
+```{r}
+
+```
+
+### Spread {#sec-descriptiveStatisticsSpread}
 
@@
 
+Below is R code to estimate each:
+
+```{r}
+
+```
+
+### Shape {#sec-descriptiveStatisticsShape}
+
 Indices to describe the *shape* of a variable's data include:
@@
 - kurtosis
+
+Below is R code to estimate each:
+
+```{r}
+
+```
+
+### Combination {#sec-descriptiveStatisticsCombination}
+
+To estimate a combination of indices of center, spread, and shape of the data, you can use the following code:
+
+```{r}
+#psych::describe(mydata)
+
+#mydata %>% 
+#  summarise(across(
+#      everything(),
+#      .fns = list(
+#        n = ~ length(na.omit(.)),
+#        missingness = ~ mean(is.na(.)) * 100,
+#        M = ~ mean(., na.rm = TRUE),
+#        SD = ~ sd(., na.rm = TRUE),
+#        min = ~ min(., na.rm = TRUE),
+#        max = ~ max(., na.rm = TRUE),
+#        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
+#        median = ~ median(., na.rm = TRUE),
+#        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
+#        skewness = ~ psych::skew(., na.rm = TRUE),
+#        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
+#      .names = "{.col}.{.fn}")) %>%
+#    pivot_longer(
+#      cols = everything(),
+#      names_to = c("variable","index"),
+#      names_sep = "\\.") %>% 
+#    pivot_wider(
+#      names_from = index,
+#      values_from = value)
+```
 

### 2024-06-18 correlation.qmd
20240605 - getting started/load packages
20240608 - getting started
20240617 - correlation coefficient

--- 2024-06-05 base-rates.qmd
@@
-# Base Rates {#sec-baseRates}
+# Correlation Analysis {#sec-correlation}
 
-## Getting Started
+## Getting Started {#sec-correlationGettingStarted}
 
-### Load Packages
+### Load Packages {#sec-correlationLoadPackages}
 
 ```{r}
+library("petersenlab")
+```
 
+## The Correlation Coefficient ($r$)
+
+The correlation coefficient ranges from −1.0 to +1.0.\index{correlation}
+The correlation coefficient ($r$) tells you two things: (1) the direction of the association (positive or negative) and (2) the magnitude of the association.\index{correlation}
+If the correlation coefficient is positive, the association is positive.\index{correlation}
+If the correlation coefficient is negative, the association is negative.\index{correlation}
+If the association is positive, as `X` increases, `Y` increases (or conversely, as `X` decreases, `Y` decreases).\index{correlation}
+If the association is negative, as `X` increases, `Y` decreases (or conversely, as `X` decreases, `Y` increases).\index{correlation}
+The smaller the absolute value of the correlation coefficient (i.e., the closer the $r$ value is to zero), the weaker the association and the flatter the slope of the best-fit line in a scatterplot.\index{correlation}
+The larger the absolute value of the correlation coefficient (i.e., the closer the absolute value of the $r$ value is to one), the stronger the association and the steeper the slope of the best-fit line in a scatterplot.\index{correlation}
+See @fig-rangeOfCorrelations for a range of different correlation coefficients and what some example data may look like for each direction and strength of association.\index{correlation}
+
+```{r}
+#| label: fig-rangeOfCorrelations
+#| fig-cap: "Correlation Coefficients."
+#| code-fold: true
+
+set.seed(52242)
+correlations <- data.frame(criterion = rnorm(1000))
+
+correlations$v1 <- complement(correlations$criterion, -1)
+correlations$v2 <- complement(correlations$criterion, -.9)
+correlations$v3 <- complement(correlations$criterion, -.8)
+correlations$v4 <- complement(correlations$criterion, -.7)
+correlations$v5 <- complement(correlations$criterion, -.6)
+correlations$v6 <- complement(correlations$criterion, -.5)
+correlations$v7 <- complement(correlations$criterion, -.4)
+correlations$v8 <- complement(correlations$criterion, -.3)
+correlations$v9 <- complement(correlations$criterion, -.2)
+correlations$v10 <-complement(correlations$criterion, -.1)
+correlations$v11 <-complement(correlations$criterion, 0)
+correlations$v12 <-complement(correlations$criterion, .1)
+correlations$v13 <-complement(correlations$criterion, .2)
+correlations$v14 <-complement(correlations$criterion, .3)
+correlations$v15 <-complement(correlations$criterion, .4)
+correlations$v16 <-complement(correlations$criterion, .5)
+correlations$v17 <-complement(correlations$criterion, .6)
+correlations$v18 <-complement(correlations$criterion, .7)
+correlations$v19 <-complement(correlations$criterion, .8)
+correlations$v20 <-complement(correlations$criterion, .9)
+correlations$v21 <-complement(correlations$criterion, 1)
+
+par(mfrow = c(7,3), mar = c(1, 0, 1, 0))
+
+# -1.0
+plot(correlations$criterion, correlations$v1, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v1)$estimate, 2))))
+abline(lm(v1 ~ criterion, data = correlations), col = "black")
+
+# -.9
+plot(correlations$criterion, correlations$v2, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v2)$estimate, 2))))
+abline(lm(v2 ~ criterion, data = correlations), col = "black")
+
+# -.8
+plot(correlations$criterion, correlations$v3, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v3)$estimate, 2))))
+abline(lm(v3 ~ criterion, data = correlations), col = "black")
+
+# -.7
+plot(correlations$criterion, correlations$v4, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v4)$estimate, 2))))
+abline(lm(v4 ~ criterion, data = correlations), col = "black")
+
+# -.6
+plot(correlations$criterion, correlations$v5, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v5)$estimate, 2))))
+abline(lm(v5 ~ criterion, data = correlations), col = "black")
+
+# -.5
+plot(correlations$criterion, correlations$v6, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v6)$estimate, 2))))
+abline(lm(v6 ~ criterion, data = correlations), col = "black")
+
+# -.4
+plot(correlations$criterion, correlations$v7, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v7)$estimate, 2))))
+abline(lm(v7 ~ criterion, data = correlations), col = "black")
+
+# -.3
+plot(correlations$criterion, correlations$v8, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v8)$estimate, 2))))
+abline(lm(v8 ~ criterion, data = correlations), col = "black")
+
+# -.2
+plot(correlations$criterion, correlations$v9, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v9)$estimate, 2))))
+abline(lm(v9 ~ criterion, data = correlations), col = "black")
+
+# -.1
+plot(correlations$criterion, correlations$v10, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v10)$estimate, 2))))
+abline(lm(v10 ~ criterion, data = correlations), col = "black")
+
+# 0.0
+plot(correlations$criterion, correlations$v11, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v11)$estimate, 2))))
+abline(lm(v11 ~ criterion, data = correlations), col = "black")
+
+# 0.1
+plot(correlations$criterion, correlations$v12, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v12)$estimate, 2))))
+abline(lm(v12 ~ criterion, data = correlations), col = "black")
+
+# 0.2
+plot(correlations$criterion, correlations$v13, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v13)$estimate, 2))))
+abline(lm(v13 ~ criterion, data = correlations), col = "black")
+
+# 0.3
+plot(correlations$criterion, correlations$v14, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v14)$estimate, 2))))
+abline(lm(v14 ~ criterion, data = correlations), col = "black")
+
+# 0.4
+plot(correlations$criterion, correlations$v15, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v15)$estimate, 2))))
+abline(lm(v15 ~ criterion, data = correlations), col = "black")
+
+# 0.5
+plot(correlations$criterion, correlations$v16, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v16)$estimate, 2))))
+abline(lm(v16 ~ criterion, data = correlations), col = "black")
+
+# 0.6
+plot(correlations$criterion, correlations$v17, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v17)$estimate, 2))))
+abline(lm(v17 ~ criterion, data = correlations), col = "black")
+
+# 0.7
+plot(correlations$criterion, correlations$v18, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v18)$estimate, 2))))
+abline(lm(v18 ~ criterion, data = correlations), col = "black")
+
+# 0.8
+plot(correlations$criterion, correlations$v19, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v19)$estimate, 2))))
+abline(lm(v19 ~ criterion, data = correlations), col = "black")
+
+# 0.9
+plot(correlations$criterion, correlations$v20, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v20)$estimate, 2))))
+abline(lm(v20 ~ criterion, data = correlations), col = "black")
+
+# 1.0
+plot(correlations$criterion, correlations$v21, xaxt = "n", yaxt = "n", xlab = "" , ylab = "",
+     main = substitute(paste(italic(r), " = ", x, sep = ""), list(x = round(cor.test(x = correlations$criterion, y = correlations$v21)$estimate, 2))))
+abline(lm(v21 ~ criterion, data = correlations), col = "black")
+
+dev.off() #par(mfrow = c(1,1))
 ```

### 2024-06-18 basic-statistics.qmd
20240617 - missing cite
20240617 - formulas

--- 2024-06-17 basic-statistics.qmd
@@
 - mode
+
+The mean of $X$ ($\bar{X}$) is calculated as:
+
+$$
+\bar{X} = \frac{\sum X_i}{n} = \frac{x_1 + x_2 + ... + x_n}{n}
+$$ {#eq-IQR}
+
+That is, to compute the mean, sum all of the values and divide by the number of values ($n$).
+
+The median is determined as the value at the 50th percentile (i.e., the value that is higher than 50% of the values and is lower than the other 50% of values).
+
+The mode is the most common value.
+
+Below is R code to estimate each:
+
+```{r}
+
+```
+
+### Spread {#sec-descriptiveStatisticsSpread}
 
@@
 - minimum and maximum
+- interquartile range (IQR)
+
+The (sample) variance of $X$ ($s^2$) is calculated as:
+
+$$
+s^2 = \frac{\sum (X_i - \bar{X})^2}{n-1}
+$$ {#eq-variance}
+
+The (sample) standard deviation of $X$ ($s$) is calculated as:
+
+$$
+s = \sqrt{\frac{\sum (X_i - \bar{X})^2}{n-1}}
+$$ {#eq-sd}
+
+The range is calculated of $X$ is calculated as:
+
+$$
+\text{range} = \text{maximum} - \text{minimum}
+$$ {#eq-range}
+
+The interquartile range (IQR) is calculated as:
+
+$$
+\text{IQR} = Q_3 - Q_1
+$$ {#eq-IQR}
+
+where $Q_3$ is the score at the third quartile (i.e., 75th percentile), and $Q_1$ is the score at the first quartile (i.e., 25th percentile).
+
+Below is R code to estimate each:
+
+```{r}
+
+```
+
+### Shape {#sec-descriptiveStatisticsShape}
 
@@
 - kurtosis
+
+Below is R code to estimate each:
+
+```{r}
+
+```
+
+### Combination {#sec-descriptiveStatisticsCombination}
+
+To estimate a combination of indices of center, spread, and shape of the data, you can use the following code:
+
+```{r}
+#psych::describe(mydata)
+
+#mydata %>% 
+#  summarise(across(
+#      everything(),
+#      .fns = list(
+#        n = ~ length(na.omit(.)),
+#        missingness = ~ mean(is.na(.)) * 100,
+#        M = ~ mean(., na.rm = TRUE),
+#        SD = ~ sd(., na.rm = TRUE),
+#        min = ~ min(., na.rm = TRUE),
+#        max = ~ max(., na.rm = TRUE),
+#        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
+#        IQR = ~ IQR(., na.rm = TRUE),
+#        median = ~ median(., na.rm = TRUE),
+#        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
+#        skewness = ~ psych::skew(., na.rm = TRUE),
+#        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
+#      .names = "{.col}.{.fn}")) %>%
+#    pivot_longer(
+#      cols = everything(),
+#      names_to = c("variable","index"),
+#      names_sep = "\\.") %>% 
+#    pivot_wider(
+#      names_from = index,
+#      values_from = value)
+```
 
@@
 
-Effect size thresholds for small, medium, and large effect sizes are in @tbl-effectSizeThresholds [@McGrath2006].
+Effect size thresholds [@McGrath2006] for small, medium, and large effect sizes are in @tbl-effectSizeThresholds [@McGrath2006].
 

### 2024-06-18 basic-statistics.qmd
20240617 - remove extra cite that isn't showing
20240617 - statistical significance

@@
 That is, there is a small chance (5%; or 1 in 20 times) that we would observe an effect at least as extreme as the effect observed, if the null hypothesis were true.
+If the *p*-value is less than .05, we reject the null hypothesis ($H_0$) that there was no difference or association.
+Thus, we conclude that there was a significant (non-zero) difference or association.
+If the *p*-value is greater than .05, we fail to reject the null hypothesis; the difference/association was not significant.
+Thus, we do not have confidence that there was a difference or association.
+However, we do not accept the null hypothesis; it could be there we did not observe an effect because we did not have adequate power to detect the effect—e.g., if the effect size was small, the data were noisy, and the sample size was small and/or unrepresentative.
+
 Statistical significance involves the *consistency* of an effect/association/difference; it suggests that the association/difference is reliably non-zero.
@@
 
-Effect size thresholds [@McGrath2006] for small, medium, and large effect sizes are in @tbl-effectSizeThresholds [@McGrath2006].
+Effect size thresholds [@McGrath2006] for small, medium, and large effect sizes are in @tbl-effectSizeThresholds.
 

### 2024-06-18 basic-statistics.qmd
20240618 - misconceptions about p-values

@@
 $$
-\bar{X} = \frac{\sum X_i}{n} = \frac{x_1 + x_2 + ... + x_n}{n}
+\bar{X} = \frac{\sum X_i}{n} = \frac{X_1 + X_2 + ... + X_n}{n}
 $$ {#eq-IQR}
@@
 
-To estimate a combination of indices of center, spread, and shape of the data, you can use the following code:
+To estimate multiple indices of center, spread, and shape of the data, you can use the following code:
 
@@
 
+Nevertheless, to be conservative (i.e., to avoid false positive/Type I errors), many researchers use two-tailed *p*-values regardless whether their hypothesis is one- or two-tailed.
+
 That is, for a test of group differences, the *p*-value evaluates the likelihood that you would observe a difference as large or larger than the one you observed between the groups if there were no systematic difference between the groups.
@@
 
-In science, we typically consider an effect to be *statistically significant* if the *p*-value is less than .05.
+In science, using what is called null-hypothesis significance testing (NHST), we typically consider an effect to be *statistically significant* if the *p*-value is less than .05.
 That is, there is a small chance (5%; or 1 in 20 times) that we would observe an effect at least as extreme as the effect observed, if the null hypothesis were true.
+So, you might expect around 5% of tests where the null hypothesis is true to be significant just by chance.
+
 If the *p*-value is less than .05, we reject the null hypothesis ($H_0$) that there was no difference or association.
@@
 However, we do not accept the null hypothesis; it could be there we did not observe an effect because we did not have adequate power to detect the effect—e.g., if the effect size was small, the data were noisy, and the sample size was small and/or unrepresentative.
+
+Interactive visualizations by Kristoffer Magnusson on *p*-values and null-hypothesis significance testing are below:
+
+- <https://rpsychologist.com/pvalue/> (archived at <https://perma.cc/JP9F-9ZVY>)
+- <https://rpsychologist.com/d3/pdist/> (archived at <https://perma.cc/BE96-8LSJ>)
+- <https://rpsychologist.com/d3/nhst/> (archived at <https://perma.cc/ZU9A-37F3>)
+
+Twelve misconceptions about *p*-values [@Goodman2008] are in @tbl-pValueMisconceptions.
+
+| Number                                   | Misconception
+|:----------------------------------------------------|:------------|
+| 1 If $p = .05$, the null hypothesis has only a 5% chance of being true.
+| 2 A nonsignificant difference (eg, $p > .05$) means there is no difference between groups.
+| 3 A statistically significant finding is clinically important.
+| 4 Studies with $p$-values on opposite sides of .05 are conflicting.
+| 5 Studies with the same $p$-value provide the same evidence against the null hypothesis.
+| 6 $p = .05$ means that we have observed data that would occur only 5% of the time under the null hypothesis.
+| 7 $p = .05$ and $p < .05$ mean the same thing.
+| 8 $p$-values are properly written as inequalities (e.g., "$p \le .05$" when $p = .015$)
+| 9 $p = .05$ means that if you reject the null hypothesis, the probability of a Type I error is only 5%.
+| 10 With a $p = .05$ threshold for significance, the chance of a Type I error will be 5%.
+| 11 You should use a one-sided $p$-value when you don't care about a result in one direction, or a difference in that direction is impossible.
+| 12 A scientific conclusion or treatment policy should be based on whether or not the $p$-value is significant.
+
+Twelve Misconceptions About *p*-Values from @Goodman2008. Goodman also provides a discussion about why each is false. {#tbl-pValueMisconceptions}
 

### 2024-06-18 basic-statistics.qmd
20240618 - NHST confusion matrix
20240618 - copyedit

@@
 
-That is, for a test of group differences, the *p*-value evaluates the likelihood that you would observe a difference as large or larger than the one you observed between the groups if there were no systematic difference between the groups.
+For a test of group differences, the *p*-value evaluates the likelihood that you would observe a difference as large or larger than the one you observed between the groups if there were no systematic difference between the groups.
 For instance, when evaluating whether Quarterbacks have <u>longer</u> careers than Running Backs, and you observed a mean difference of 0.03 years, the *p*-value evaluates the likelihood that you would observe a difference as larger or larger than 0.03 years between the groups if Quarterbacks do not differ from Running Backs in terms of the length of their career.
 
-That is, for a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed between the groups if there were no association between the variables.
+For a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed between the groups if there were no association between the variables.
 For instance, when evaluating whether number of carries is <u>positively</u> associated with injury likelihood, and you observed a correlation coefficient of $r = .25$ between number of carries and injury likelihood, the *p*-value evaluates the likelihood that you would observe a correlation as strong or stronger than $r = .25$ between the variables if number of carries is not associated with injury likelihood.
@@
 However, we do not accept the null hypothesis; it could be there we did not observe an effect because we did not have adequate power to detect the effect—e.g., if the effect size was small, the data were noisy, and the sample size was small and/or unrepresentative.
+
+A two-by-two confusion matrix for null-hypothesis significance testing is in @fig-nhstConfusionMatrix.
+
+::: {#fig-nhstConfusionMatrix}
+![](images/nhstConfusionMatrix.png)
+
+A Two-by-Two Confusion Matrix for Null-Hypothesis Significance Testing.
+:::
 

### 2024-06-18 basic-statistics.qmd
20240618 - fix table

@@
 
-| Number                                   | Misconception
-|:----------------------------------------------------|:------------|
-| 1 If $p = .05$, the null hypothesis has only a 5% chance of being true.
-| 2 A nonsignificant difference (eg, $p > .05$) means there is no difference between groups.
-| 3 A statistically significant finding is clinically important.
-| 4 Studies with $p$-values on opposite sides of .05 are conflicting.
-| 5 Studies with the same $p$-value provide the same evidence against the null hypothesis.
-| 6 $p = .05$ means that we have observed data that would occur only 5% of the time under the null hypothesis.
-| 7 $p = .05$ and $p < .05$ mean the same thing.
-| 8 $p$-values are properly written as inequalities (e.g., "$p \le .05$" when $p = .015$)
-| 9 $p = .05$ means that if you reject the null hypothesis, the probability of a Type I error is only 5%.
-| 10 With a $p = .05$ threshold for significance, the chance of a Type I error will be 5%.
-| 11 You should use a one-sided $p$-value when you don't care about a result in one direction, or a difference in that direction is impossible.
-| 12 A scientific conclusion or treatment policy should be based on whether or not the $p$-value is significant.
+| Number | Misconception                                                                                                                              |
+|:-------|:-------------------------------------------------------------------------------------------------------------------------------------------|
+| 1      | If $p = .05$, the null hypothesis has only a 5% chance of being true.                                                                      |
+| 2      | A nonsignificant difference (eg, $p > .05$) means there is no difference between groups.                                                   |
+| 3      | A statistically significant finding is clinically important.                                                                               |
+| 4      | Studies with $p$-values on opposite sides of .05 are conflicting.                                                                          |
+| 5      | Studies with the same $p$-value provide the same evidence against the null hypothesis.                                                     |
+| 6      | $p = .05$ means that we have observed data that would occur only 5% of the time under the null hypothesis.                                 |
+| 7      | $p = .05$ and $p < .05$ mean the same thing.                                                                                               |
+| 8      | $p$-values are properly written as inequalities (e.g., "$p \le .05$" when $p = .015$).                                                     |
+| 9      | $p = .05$ means that if you reject the null hypothesis, the probability of a Type I error is only 5%.                                      |
+| 10     | With a $p = .05$ threshold for significance, the chance of a Type I error will be 5%.                                                      |
+| 11     | You should use a one-sided $p$-value when you don't care about a result in one direction, or a difference in that direction is impossible. |
+| 12     | A scientific conclusion or treatment policy should be based on whether or not the $p$-value is significant.                                |
 

### 2024-06-18 basic-statistics.qmd
20240618 - edits

@@
 That is, there is a small chance (5%; or 1 in 20 times) that we would observe an effect at least as extreme as the effect observed, if the null hypothesis were true.
-So, you might expect around 5% of tests where the null hypothesis is true to be significant just by chance.
+So, you might expect around 5% of tests where the null hypothesis is true to be statistically significant just by chance.
 
 If the *p*-value is less than .05, we reject the null hypothesis ($H_0$) that there was no difference or association.
-Thus, we conclude that there was a significant (non-zero) difference or association.
-If the *p*-value is greater than .05, we fail to reject the null hypothesis; the difference/association was not significant.
+Thus, we conclude that there was a statistically significant (non-zero) difference or association.
+If the *p*-value is greater than .05, we fail to reject the null hypothesis; the difference/association was not statistically significant.
 Thus, we do not have confidence that there was a difference or association.
@@
 
-Twelve Misconceptions About *p*-Values from @Goodman2008. Goodman also provides a discussion about why each is false. {#tbl-pValueMisconceptions}
+: Twelve Misconceptions About *p*-Values from @Goodman2008. Goodman also provides a discussion about why each is false. {#tbl-pValueMisconceptions}
 

### 2024-06-18 basic-statistics.qmd
20240618 - tweaks

@@
 
-The mean of $X$ ($\bar{X}$) is calculated as:
+The mean of $X$ (written as: $\bar{X}$) is calculated as:
 
@@
 
-The mode is the most common value.
+The mode is the most common/frequent value.
 
@@
 
-The (sample) variance of $X$ ($s^2$) is calculated as:
+The (sample) variance of $X$ (written as: $s^2$) is calculated as:
 
@@
 
-The (sample) standard deviation of $X$ ($s$) is calculated as:
+The (sample) standard deviation of $X$ (written as: $s$) is calculated as:
 
@@
 
-: Twelve Misconceptions About *p*-Values from @Goodman2008. Goodman also provides a discussion about why each is false. {#tbl-pValueMisconceptions}
+: Twelve Misconceptions About *p*-Values from @Goodman2008. Goodman also provides a discussion about why each statement is false. {#tbl-pValueMisconceptions}
 

### 2024-06-18 index.qmd
20240618 - buy coffee

@@
 ```{=html}
+<a href="https://www.buymeacoffee.com/isaactpetersen" target="_blank"><img src="https://cdn.buymeacoffee.com/buttons/v2/default-yellow.png" alt="Buy Me A Coffee" style="height: 60px !important;width: 217px !important;" ></a>
+
+<script type="text/javascript" src="https://cdnjs.buymeacoffee.com/1.0.0/button.prod.min.js" data-name="bmc-button" data-slug="isaactpetersen" data-color="#FFDD00" data-emoji=""  data-font="Cookie" data-text="Buy me a coffee" data-outline-color="#000000" data-font-color="#000000" data-coffee-color="#ffffff" ></script>
+
 <form action="https://www.paypal.com/donate" method="post" target="_top">

### 2024-06-18 index.qmd
20240618 - donate links

--- 2024-06-16 index.qmd
@@
 ```{=html}
-<form action="https://www.paypal.com/donate" method="post" target="_top">
-<input type="hidden" name="business" value="Q4FMMAMFQYTX8" />
-<input type="hidden" name="no_recurring" value="0" />
-<input type="hidden" name="item_name" value="Support me in developing this resource for fantasy football analytics... Even a cup of coffee ($4.99) helps me stay awake!" />
-<input type="hidden" name="currency_code" value="USD" />
-<input type="image" src="https://www.paypalobjects.com/en_US/i/btn/btn_donate_LG.gif" border="0" name="submit" title="PayPal - The safer, easier way to pay online!" alt="Donate with PayPal button" />
-<img alt="" border="0" src="https://www.paypal.com/en_US/i/scr/pixel.gif" width="1" height="1" />
-</form>
+<a href="https://www.buymeacoffee.com/isaactpetersen" target="_blank"><img src="https://cdn.buymeacoffee.com/buttons/v2/default-yellow.png" alt="Buy Me A Coffee" style="height: 60px !important;width: 217px !important;" ></a>
+
+<br>
+
+<a href="https://www.paypal.com/donate/?business=Q4FMMAMFQYTX8&no_recurring=0&item_name=Support+me+in+developing+this+resource+for+fantasy+football+analytics...+Even+a+cup+of+coffee+%28%244.99%29+helps+me+stay+awake%21&currency_code=USD">(or use PayPal)</a>
 ```

### 2024-06-18 basic-statistics.qmd
202406187 - statistical power plot

@@
 library("petersenlab")
+library("grid")
 library("tidyverse")
@@
 
+```{r}
+#| label: fig-nhst
+#| fig-cap: "Statistical Power (Adapted from Kristoffer Magnusson): <https://rpsychologist.com/creating-a-typical-textbook-illustration-of-statistical-power-using-either-ggplot-or-base-graphics> (archived at <https://perma.cc/FG3J-85L6>)."
+#| code-fold: true
+
+m1 <- 0  # mu H0
+sd1 <- 1.5 # sigma H0
+m2 <- 3.5 # mu HA
+sd2 <- 1.5 # sigma HA
+
+z_crit <- qnorm(1-(0.05/2), m1, sd1)
+
+# set length of tails
+min1 <- m1-sd1*4
+max1 <- m1+sd1*4
+min2 <- m2-sd2*4
+max2 <- m2+sd2*4          
+# create x sequence
+x <- seq(min(min1,min2), max(max1, max2), .01)
+# generate normal dist #1
+y1 <- dnorm(x, m1, sd1)
+# put in data frame
+df1 <- data.frame("x" = x, "y" = y1)
+# generate normal dist #2
+y2 <- dnorm(x, m2, sd2)
+# put in data frame
+df2 <- data.frame("x" = x, "y" = y2)
+
+# Alpha polygon
+y.poly <- pmin(y1,y2)
+poly1 <- data.frame(x=x, y=y.poly)
+poly1 <- poly1[poly1$x >= z_crit, ] 
+poly1<-rbind(poly1, c(z_crit, 0))  # add lower-left corner
+
+# Beta polygon
+poly2 <- df2
+poly2 <- poly2[poly2$x <= z_crit,] 
+poly2<-rbind(poly2, c(z_crit, 0))  # add lower-left corner
+
+# power polygon; 1-beta
+poly3 <- df2
+poly3 <- poly3[poly3$x >= z_crit,] 
+poly3 <-rbind(poly3, c(z_crit, 0))  # add lower-left corner
+
+# combine polygons. 
+poly1$id <- 3 # alpha, give it the highest number to make it the top layer
+poly2$id <- 2 # beta
+poly3$id <- 1 # power; 1 - beta
+poly <- rbind(poly1, poly2, poly3)
+poly$id <- factor(poly$id,  labels=c("power","beta","alpha"))
+
+# plot with ggplot2
+ggplot(poly, aes(x,y, fill=id, group=id)) +
+  geom_polygon(show.legend=F, alpha=I(8/10)) +
+  # add line for treatment group
+  geom_line(data=df1, aes(x,y, color="H0", group=NULL, fill=NULL), linewidth=1.5, show_guide=F) + 
+  # add line for treatment group. These lines could be combined into one dataframe.
+  geom_line(data=df2, aes(color="HA", group=NULL, fill=NULL),linewidth=1.5, show_guide=F) +
+  # add vlines for z_crit
+  geom_vline(xintercept = z_crit, linewidth=1, linetype="dashed") +
+  # change colors 
+  scale_color_manual("Group", 
+                     values= c("HA" = "#981e0b","H0" = "black")) +
+  scale_fill_manual("test", values= c("alpha" = "#0d6374","beta" = "#be805e","power"="#7cecee")) +
+  # beta arrow
+  annotate("segment", x=0.1, y=0.045, xend=1.3, yend=0.01, arrow = arrow(length = unit(0.3, "cm")), linewidth=1) +
+  annotate("text", label="beta", x=0, y=0.05, parse=T, size=8) +
+  # alpha arrow
+  annotate("segment", x=4, y=0.043, xend=3.4, yend=0.01, arrow = arrow(length = unit(0.3, "cm")), linewidth=1) +
+  annotate("text", label="frac(alpha,2)", x=4.2, y=0.05, parse=T, size=8) +
+  # power arrow
+  annotate("segment", x=6, y=0.2, xend=4.5, yend=0.15, arrow = arrow(length = unit(0.3, "cm")), linewidth=1) +
+  annotate("text", label=expression(paste(1-beta, "  (\"power\")")), x=6.1, y=0.21, parse=T, size=8) +
+  # H_0 title
+  annotate("text", label="H[0]", x=m1, y=0.28, parse=T, size=8) +
+  # H_a title
+  annotate("text", label="H[1]", x=m2, y=0.28, parse=T, size=8) +
+  ggtitle("Statistical Power") +
+  # remove some elements
+  theme(
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.background = element_blank(),
+    plot.background = element_rect(fill="white"),
+    panel.border = element_blank(),
+    axis.line = element_blank(),
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    plot.title = element_text(size=22))
+```
+
 Interactive visualizations by Kristoffer Magnusson on *p*-values and null-hypothesis significance testing are below:

### 2024-06-18 basic-statistics.qmd
20240618 - clarifications

@@
 #| label: fig-nhst
-#| fig-cap: "Statistical Power (Adapted from Kristoffer Magnusson): <https://rpsychologist.com/creating-a-typical-textbook-illustration-of-statistical-power-using-either-ggplot-or-base-graphics> (archived at <https://perma.cc/FG3J-85L6>)."
+#| fig-cap: "Statistical Power (Adapted from Kristoffer Magnusson: <https://rpsychologist.com/creating-a-typical-textbook-illustration-of-statistical-power-using-either-ggplot-or-base-graphics>; archived at <https://perma.cc/FG3J-85L6>).
+The dashed line represents the critical value or threshold."
 #| code-fold: true

### 2024-06-18 index.qmd
20240618 - clarifications

@@
 <a href="https://www.paypal.com/donate/?business=Q4FMMAMFQYTX8&no_recurring=0&item_name=Support+me+in+developing+this+resource+for+fantasy+football+analytics...+Even+a+cup+of+coffee+%28%244.99%29+helps+me+stay+awake%21&currency_code=USD">(or use PayPal)</a>
+
+<br>
 ```

### 2024-06-18 index.qmd
20240618 - coffee

@@
   - [open an issue](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues) or [create a pull request](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/pulls) on the [book's GitHub repository](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook).\index{GitHub}
-  - [buy me a coffee](https://www.paypal.com/donate/?business=Q4FMMAMFQYTX8&no_recurring=0&item_name=Support+me+in+developing+this+resource+for+fantasy+football+analytics...+Even+a+cup+of+coffee+%28%244.99%29+helps+me+stay+awake%21&currency_code=USD)—Support me in developing this resource for fantasy football analytics...
-Even a cup of coffee ($4.99) helps me stay awake!
+  - [buy me a coffee](https://www.buymeacoffee.com/isaactpetersen)—Support me in developing this resource for fantasy football analytics... Even a cup of coffee ($4.99) helps me stay awake!
 
 ```{=html}
-<a href="https://www.buymeacoffee.com/isaactpetersen" target="_blank"><img src="https://cdn.buymeacoffee.com/buttons/v2/default-yellow.png" alt="Buy Me A Coffee" style="height: 60px !important;width: 217px !important;" ></a>
+<script type="text/javascript" src="https://cdnjs.buymeacoffee.com/1.0.0/button.prod.min.js" data-name="bmc-button" data-slug="isaactpetersen" data-color="#FFDD00" data-emoji=""  data-font="Lato" data-text="Buy me a coffee" data-outline-color="#000000" data-font-color="#000000" data-coffee-color="#ffffff" ></script>
 
@@
 
+<br>
 <br>

### 2024-06-19 basic-statistics.qmd
20240618 - NHST

@@
 
-### Null Hypothesis ($H_0$)
+### Null Hypothesis Significance Testing {#sec-nhst}
+
+#### Null Hypothesis ($H_0$) {#sec-nullHypothesis}
 
@@
 
-### Alternative Hypothesis ($H_1$)
+#### Alternative Hypothesis ($H_1$) {#sec-alternativeHypothesis}
 
@@
 
-### Statistical Significance {#sec-statisticalSignificance}
+#### Statistical Significance {#sec-statisticalSignificance}
 
@@
 A two-by-two confusion matrix for null-hypothesis significance testing is in @fig-nhstConfusionMatrix.
+There are four general possibilities:
+
+1. We can (correctly) reject the null hypothesis when it is in fact false ($1 - \beta$).
+This is a true positive.
+For instance, we may correctly determine that Quarterbacks have longer careers than Running Backs.
+1. We can (correctly) fail to reject the null hypothesis when it is in fact true ($1 - \alpha$).
+This is a true negative.
+For instance, we may correctly determine that Quarterbacks do not have longer careers than Running Backs.
+1. We can (incorrectly) reject the null hypothesis when it is in fact true ($\alpha$).
+This is a false positive.
+When performing a null hypothesis testing, a false positive is known as a Type I error.
+For instance, we may incorrectly determine that Quarterbacks have longer careers than Running Backs when, in fact, Quarterbacks and Running Backs do not differ in their career length.
+1. We can (incorrectly) fail to reject the null hypothesis when it is in fact false ($\beta$).
+This is a false positive.
+When performing a null hypothesis testing, a false negative is known as a Type II error.
+For instance, we may incorrectly determine that Quarterbacks and Running Backs do not differ in their career length when, in fact, Quarterbacks have longer careers than Running Backs.
 
@@
 :::
+
+A plot of statistical power is in @fig-nhst.
+In statistics, *power* is the probability of detecting an effect, if, in fact, the effect exists.
+Otherwise said, power is the probability of rejecting the null hypothesis, if, in fact, the null hypothesis is false.
+Power is influenced by several variables:
+
+- the sample size (*N*): the larger the *N*, the greater the power
+- the [effect size](sec-practicalSignificance): the larger the effect, the greater the power
+    - for group comparisons:
+        - larger between-group variance, and
+        - smaller within-group variance (i.e., strong measurement precision)
+- the alpha level (typically .05): the higher the alpha level, the greater the power; however, the higher we set the alpha level, the higher the likelihood of Type I errors (false positives)
+- one- versus two-tailed tests: one-tailed tests have higher power than two-tailed tests
+- within-subject versus between-subject comparisons: within-subject designs tend to have greater power than between-subject designs
 

### 2024-06-19 research-methods.qmd
20240616 - placeholders for code
20240616 - Kelvin temp
20240618 - NHST

@@
 That is, we could evaluate, when a given player has a sports drink (or more sports drinks), do they perform better than <u>when the same individual</u> has only water (or fewer sports drinks)?
+
+Within-subject designs tend to have greater statistical power than between-subject designs.
+However, within-subject designs often have *carryover effects*.
+For instance, consider the study in which we assign players to drink only water in the first and third quarters and to drink sports drink in the second and fourth quarters (an A-B-A-B design).
+Drinking sports drink in the second quarter could increase how much hydration a player has throughout the rest of the game, which could lead to altered performance in the third and fourth quarters that is not due to what they drink in third and fourth quarters.
 
@@
 
+- temperature on the Kelvin scale
 - height
@@
 
+```{r}
+
+```
+
 #### Inter-Rater Reliability {#sec-interraterReliability}
@@
 The consistency of scores across raters is commonly evaluated using the intraclass correlation coefficient (for continuous variables) and Cohen's kappa ($\kappa$; for categorical variables).
+
+```{r}
+
+```
 
@@
 
+```{r}
+
+```
+
 #### Internal Consistency {#sec-internalConsistency}
@@
 
+```{r}
+
+```
+
 #### Parallel-Forms Reliability {#sec-parallelFormsReliability}
@@
 Parallel-forms reliability is is commonly evaluated using the coefficient of equivalence (i.e., the Pearson correlation coefficent).
+
+```{r}
+
+```
 
@@
 
+```{r}
+
+```
+
 ##### Predictive (Criterion-Related) Validity {#sec-predictiveValidity}
@@
 For example, predictive validity could evaluate whether scores on the measure we developed to assess a player's aptitude to succeed in the NFL predicts later performance in the NFL.
+
+```{r}
+
+```
 
@@
 
+```{r}
+
+```
+
 #### Discriminant Validity {#sec-discriminantValidity}
@@
 For instance, if we develop a new measure to assess intelligence, its scores should be less strongly associated with measures of other constructs (e.g., measures of happiness).
+
+```{r}
+
+```
 
@@
 For instance, if we develop a strong measure of intelligence, it should result in increased predictive accuracy (for success in the NFL) compared to when just relying on the Wonderlic Contemporary Cognitive Ability Test.
+
+```{r}
+
+```
 

### 2024-06-19 basic-statistics.qmd
20240618 - tweaks

@@
 
-A two-by-two confusion matrix for null-hypothesis significance testing is in @fig-nhstConfusionMatrix.
-There are four general possibilities:
+There are four general possibilities of decision making outcomes when performing null-hypothesis significance testing:
 
-1. We can (correctly) reject the null hypothesis when it is in fact false ($1 - \beta$).
+1. We (correctly) reject the null hypothesis when it is in fact false ($1 - \beta$).
 This is a true positive.
 For instance, we may correctly determine that Quarterbacks have longer careers than Running Backs.
-1. We can (correctly) fail to reject the null hypothesis when it is in fact true ($1 - \alpha$).
+1. We (correctly) fail to reject the null hypothesis when it is in fact true ($1 - \alpha$).
 This is a true negative.
 For instance, we may correctly determine that Quarterbacks do not have longer careers than Running Backs.
-1. We can (incorrectly) reject the null hypothesis when it is in fact true ($\alpha$).
+1. We (incorrectly) reject the null hypothesis when it is in fact true ($\alpha$).
 This is a false positive.
-When performing a null hypothesis testing, a false positive is known as a Type I error.
+When performing null hypothesis testing, a false positive is known as a Type I error.
 For instance, we may incorrectly determine that Quarterbacks have longer careers than Running Backs when, in fact, Quarterbacks and Running Backs do not differ in their career length.
-1. We can (incorrectly) fail to reject the null hypothesis when it is in fact false ($\beta$).
+1. We (incorrectly) fail to reject the null hypothesis when it is in fact false ($\beta$).
 This is a false positive.
-When performing a null hypothesis testing, a false negative is known as a Type II error.
+When performing null hypothesis testing, a false negative is known as a Type II error.
 For instance, we may incorrectly determine that Quarterbacks and Running Backs do not differ in their career length when, in fact, Quarterbacks have longer careers than Running Backs.
+
+A two-by-two confusion matrix for null-hypothesis significance testing is in @fig-nhstConfusionMatrix.
 
@@
 
-A plot of statistical power is in @fig-nhst.
 In statistics, *power* is the probability of detecting an effect, if, in fact, the effect exists.
@@
 - the sample size (*N*): the larger the *N*, the greater the power
-- the [effect size](sec-practicalSignificance): the larger the effect, the greater the power
+- the [effect size](#sec-practicalSignificance): the larger the effect, the greater the power
     - for group comparisons:
@@
 - within-subject versus between-subject comparisons: within-subject designs tend to have greater power than between-subject designs
+
+A plot of statistical power is in @fig-nhst.
 
@@
 
-*Practical significance* deals with how important the effect/association/difference is.
+*Practical significance* deals with how large or important the effect/association/difference is.
 It is based on the magnitude of the effect, called the *effect size*.

### 2024-06-19 basic-statistics.qmd
20240619 - additions

@@
 - the sample size (*N*): the larger the *N*, the greater the power
+    - for group comparisons, the power depends on the sample size of each group
 - the [effect size](#sec-practicalSignificance): the larger the effect, the greater the power
-    - for group comparisons:
+    - for group comparisons, larger effect sizes reflect:
         - larger between-group variance, and
-        - smaller within-group variance (i.e., strong measurement precision)
-- the alpha level (typically .05): the higher the alpha level, the greater the power; however, the higher we set the alpha level, the higher the likelihood of Type I errors (false positives)
+        - smaller within-group variance (i.e., strong measurement precision, i.e., [reliability](#sec-reliability))
+- the alpha level: the researcher specifies the alpha level (though it is typically set at .05); the higher the alpha level, the greater the power; however, the higher we set the alpha level, the higher the likelihood of Type I errors (false positives)
 - one- versus two-tailed tests: one-tailed tests have higher power than two-tailed tests
-- within-subject versus between-subject comparisons: within-subject designs tend to have greater power than between-subject designs
+- [within-subject](#sec-withinSubject) versus [between-subject](#sec-betweenSubject) comparisons: [within-subject designs](#sec-withinSubject) tend to have greater power than [between-subject designs](#sec-betweenSubject)
 

### 2024-06-19 index.qmd
20240618 - make smaller
20240619 - tweaks
20240619 - tweak

@@
   - [open an issue](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues) or [create a pull request](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/pulls) on the [book's GitHub repository](https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook).\index{GitHub}
-  - [buy me a coffee](https://www.buymeacoffee.com/isaactpetersen)—Support me in developing this resource for fantasy football analytics... Even a cup of coffee ($4.99) helps me stay awake!
+  - [buy me a coffee](https://www.buymeacoffee.com/isaactpetersen)—Support me in developing this (free!) resource for fantasy football analytics... Even a cup of coffee helps me stay awake!
 
@@
 <script type="text/javascript" src="https://cdnjs.buymeacoffee.com/1.0.0/button.prod.min.js" data-name="bmc-button" data-slug="isaactpetersen" data-color="#FFDD00" data-emoji=""  data-font="Lato" data-text="Buy me a coffee" data-outline-color="#000000" data-font-color="#000000" data-coffee-color="#ffffff" ></script>
-
-<br>
 
@@
 
-![Creative Commons License](images/CC-BY.png){width=50%}
+![Creative Commons License](images/CC-BY.png){width=25% fig-align="left"}
 

### 2024-06-19 basic-statistics.qmd
20240619 - typo
20240619 - regression framework

@@
 1. We (incorrectly) fail to reject the null hypothesis when it is in fact false ($\beta$).
-This is a false positive.
+This is a false negative.
 When performing null hypothesis testing, a false negative is known as a Type II error.
@@
 
+However, many statistical tests can be re-formulated in a regression framework, as in @fig-statisticalDecisionTreeRegression.
+
+::: {#fig-statisticalDecisionTreeRegression}
+![](images/statisticalDecisionTreeRegression.png)
+
+A Statistical Decision Tree For Choosing an Appropriate Statistical Procedure, Re-Formulated in a Regression Framework.
+Adapted from: <https://commons.wikimedia.org/wiki/File:InferentialStatisticalDecisionMakingTrees.pdf>.
+The original source is: Corston, R.
+& Colman, A.
+M.
+(2000).
+*A crash course in SPSS for Windows*.
+Wiley-Blackwell.
+Changes were made to the original, including re-formulating the tests in a regression framework.
+:::
+
 For an online, interactive statistical decision tree to help you decide which statistical analysis to use, see here: <https://www.statsflowchart.co.uk>

### 2024-06-24 basic-statistics.qmd
20240623 - statistical tests

@@
 For an online, interactive statistical decision tree to help you decide which statistical analysis to use, see here: <https://www.statsflowchart.co.uk>
+
+## Statistical Tests {#sec-statisticalTests}
+
+### *t*-Test {#sec-tTest}
+
+There are several *t*-tests:
+
+- one-sample *t*-test
+- two-samples *t*-test
+    - independent samples *t*-test
+    - paired samples *t*-test
+
+A one-sample *t*-test is used to evaluate whether a sample mean differs systematically from a particular value.
+For instance, let's say you want to test out a new draft strategy.
+You could participate in a mock draft and draft players using the new strategy.
+Then, you could use a one-sample *t*-test to evaluate whether your new draft strategy yields players with more projected points than the average of players' projected points for other teams.
+
+Two-samples *t*-tests are used to test for differences between scores of two groups.
+If the two groups are independent, the independent samples *t*-test is used.
+If the two groups involve paired samples, the paired samples *t*-test is used.
+For instance, you could use an independent-samples *t*-test if you want to examine whether Quarterbacks tend to have have longer careers than Running Backs.
+By contrast, you could use a paired samples *t*-test if you want to examine whether Quarterbacks tend to score more points in the second year of their contract compared to their rookie year, because the same subjects were assessed twice (i.e., a [within-subject designs](#sec-withinSubject)).
+
+### Analysis of Variance {#sec-anova}
+
+Analysis of variance (ANOVA) allows examining whether groups differ systematically as a function of one or more factors.
+There are multiple variants of ANOVA:
+
+- one-way ANOVA
+- factorial ANOVA
+- repeated measures ANOVA (RM-ANOVA)
+- multivariate ANOVA (MANOVA)
+
+Like two-samples *t*-tests, ANOVA allows examining whether groups differ as a function of an [independent variable](#sec-experiment).
+However, unlike a *t*-test, ANOVA allows examining multiple multiple [independent variables](#sec-experiment) and more than two groups.
+
+A one-way ANOVA examines whether two or more groups differ as a function of an [independent variable](#sec-experiment).
+For instance, you could use a one-way ANOVA to evaluate if you want to evaluate whether multiple positions differ in their length of career.
+Factorial ANOVA examines whether two or more groups differ as a function of multiple [independent variables](#sec-experiment).
+For instance, you could use factorial ANOVA to evaluate whether one's length of career depends on one's position and weight.
+Repeated measures ANOVA examines whether scores differ across repeated measures (e.g., across time) for the same participants.
+For instance, you could use repeated-measures ANOVA to evaluate whether rookies score more points as the season progresses.
+Multivariate ANOVA examines whether multiple [dependent variables](#sec-experiment) differ as a function of one or more factor(s).
+For instance, you could use MANOVA to evaluate whether one's contract length and pay differ as a function of one's position.
+
+### Correlation {#sec-correlation}
+
+Correlation examines the association between a [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variable.
+
+### (Multiple) Regression {#sec-regression}
+
+Regression, like correlation, examines the association between a [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variable.
+However, unlike correlation, regression allows multiple [predictor variables](#sec-correlationalStudy).
+
+Regression with one predictor takes the following form:
+
+$$
+y = B_0 + B_1x_1 + \epsilon
+$$ {#eq-regression}
+
+where $y$ is the [outcome variable](#sec-correlationalStudy), $B_0$ is the intercept, $B_1$ is the slope, $x_1$ is the [predictor variable](#sec-correlationalStudy), and $\epsilon$ is the error term.
+
+Regression with multiple predictors takes the following form:
+
+$$
+y = B_0 + B_1x_1 + B_2x_2 + ... + B_px_p + \epsilon
+$$ {#eq-multipleRegression}
+
+where $p$ is the number of [predictor variables](#sec-correlationalStudy).
+
+### Formulating Statistical Tests in Terms of Partitioned Variance {#sec-statisticalTestsPartitionedVariance}
+
+Many statistical tests can be formulated in terms of partitioned variance.
+
+For instance, the *t* statistic from the independent-samples *t*-test and the *F* statistic from ANOVA can be thought of as the ratio of between-group variance to within-group variance:
+
+$$
+t \text{ or } F = \frac{\text{between-group variance}}{\text{within-group variance}}
+$$ {#eq-tTestPartitionedVariance}
+
+The correlation coefficient can be thought of as the ratio of shared variance (i.e., covariance) to total variance:
+
+$$
+r = \frac{\text{shared variance}}{\text{total variance}}
+$$ {#eq-correlationPartitionedVariance}

### 2024-06-24 basic-statistics.qmd
20240623 - tweaks

@@
 For instance, you could use an independent-samples *t*-test if you want to examine whether Quarterbacks tend to have have longer careers than Running Backs.
-By contrast, you could use a paired samples *t*-test if you want to examine whether Quarterbacks tend to score more points in the second year of their contract compared to their rookie year, because the same subjects were assessed twice (i.e., a [within-subject designs](#sec-withinSubject)).
+By contrast, you could use a paired samples *t*-test if you want to examine whether Quarterbacks tend to score more points in the second year of their contract compared to their rookie year, because the same subjects were assessed twice (i.e., a [within-subject design](#sec-withinSubject)).
 
@@
 
-Regression with multiple predictors takes the following form:
+Regression with multiple predictors—i.e., multiple regression—takes the following form:
 

### 2024-06-24 basic-statistics.qmd
20240623 - coefficient of determination

@@
 
-Regression with one predictor takes the following form:
+Regression with one predictor takes the form of @eq-regression:
 
@@
 
-Regression with multiple predictors—i.e., multiple regression—takes the following form:
+Regression with multiple predictors—i.e., multiple regression—takes the form of @eq-multipleRegression:
 
@@
 
-For instance, the *t* statistic from the independent-samples *t*-test and the *F* statistic from ANOVA can be thought of as the ratio of between-group variance to within-group variance:
+For instance, the *t* statistic from the independent-samples *t*-test and the *F* statistic from ANOVA can be thought of as the ratio of between-group variance to within-group variance, as in @eq-tTestPartitionedVariance:
 
@@
 
-The correlation coefficient can be thought of as the ratio of shared variance (i.e., covariance) to total variance:
+The correlation coefficient can be thought of as the ratio of shared variance (i.e., covariance) to total variance, as in @eq-correlationPartitionedVariance:
 
@@
 $$ {#eq-correlationPartitionedVariance}
+
+The coefficient of determination ($R^2$) is the proportion of variance in the [outcome variable](#sec-correlationalStudy) that is explained by the [predictor variables](#sec-correlationalStudy).
+The coefficient of determination can be expressed as the ratio of variance explained in the [outcome variable](#sec-correlationalStudy) to the total variance in the [outcome variable](#sec-correlationalStudy), as in @eq-coefficientOfDetermination:
+
+$$
+R^2 = \frac{\text{variance explained in the outcome variable}}{\text{total variance in the outcome variable}}
+$$ {#eq-coefficientOfDetermination}

### 2024-06-24 basic-statistics.qmd
20240623 - regression best-fit line

@@
 
+```{r}
+#| include: false
+#| eval: false
+
+set.seed(52242)
+regression <- data.frame(outcome = rnorm(40, mean = 5, sd = 2))
+
+regression$predictor <- complement(regression$outcome, .5)
+regression$predictor <- regression$predictor + abs(min(regression$predictor))
+
+lm(
+  outcome ~ predictor,
+  data = regression)
+
+ggplot2::ggplot(
+  data = regression,
+  ggplot2::aes(
+    x = predictor,
+    y = outcome,
+  )
+) +
+  ggplot2::geom_point() +
+  ggplot2::geom_smooth(
+    method = "lm",
+    linewidth = 2,
+    se = FALSE,
+    fullrange = TRUE) +
+  ggplot2::scale_x_continuous(
+    lim = c(0,8),
+    breaks = seq(from = 0, to = 8, by = 2),
+    expand = c(0,0)
+  ) +
+  ggplot2::scale_y_continuous(
+    lim = c(0,8),
+    breaks = seq(from = 0, to = 8, by = 2),
+    expand = c(0,0)
+  ) +
+  ggplot2::labs(
+    x = "Predictor Variable",
+    y = "Outcome Variable",
+    title = "Regression Best-Fit Line"
+  ) +
+  ggplot2::theme_classic(
+    base_size = 16) +
+  ggplot2::theme(legend.title = element_blank())
+
+ggsave("./images/regression.pdf", width = 6, height = 6)
+```
+
+A regression line is depicted in @fig-regression.
+
+::: {#fig-regression}
+![](images/regression.png)
+
+A Regression Best-Fit Line.
+:::
+
 Regression with multiple predictors—i.e., multiple regression—takes the form of @eq-multipleRegression:

### 2024-06-24 basic-statistics.qmd
20240624 - beta

@@
 $$
-y = B_0 + B_1x_1 + \epsilon
+y = \beta_0 + \beta_1x_1 + \epsilon
 $$ {#eq-regression}
 
-where $y$ is the [outcome variable](#sec-correlationalStudy), $B_0$ is the intercept, $B_1$ is the slope, $x_1$ is the [predictor variable](#sec-correlationalStudy), and $\epsilon$ is the error term.
+where $y$ is the [outcome variable](#sec-correlationalStudy), $\beta_0$ is the intercept, $\beta_1$ is the slope, $x_1$ is the [predictor variable](#sec-correlationalStudy), and $\epsilon$ is the error term.
 
@@
 $$
-y = B_0 + B_1x_1 + B_2x_2 + ... + B_px_p + \epsilon
+y = \beta_0 + \beta_1x_1 + \beta_2x_2 + ... + \beta_px_p + \epsilon
 $$ {#eq-multipleRegression}

### 2024-06-24 basic-statistics.qmd
20240624 - critical value

@@
 
-In science, using what is called null-hypothesis significance testing (NHST), we typically consider an effect to be *statistically significant* if the *p*-value is less than .05.
+Using what is called null-hypothesis significance testing (NHST), we consider an effect to be *statistically significant* if the *p*-value is less than some threshold, called the *alpha level*.
+In science, we typically want to be conservative because a false positive (i.e., Type I error) is considered more problematic than a false negative (i.e., Type II error).
+That is, we would rather say an effect does not exist when it really does than to say an effect does exist when it really does not.
+Thus, we typically set the alpha level to a low value, commonly .05.
+Then, we would consider an effect to be *statistically significant* if the *p*-value is less than .05.
 That is, there is a small chance (5%; or 1 in 20 times) that we would observe an effect at least as extreme as the effect observed, if the null hypothesis were true.
 So, you might expect around 5% of tests where the null hypothesis is true to be statistically significant just by chance.
+We could lower the rate of Type II (i.e., false negative) errors—i.e., we could detect more effects—if we set the alpha level to a higher value (e.g., .10); however, raising the alpha level would raise the possibility of Type I (false positive) errors.
 
@@
 $$ {#eq-coefficientOfDetermination}
+
+### Critical Value {#sec-statisticalTestsCriticalValue}
+
+The critical value is the test value for a given test, above which the effect is considered to be [statistically significant](#sec-statisticalSignificance).
+The critical value for [statistical significance](#sec-statisticalSignificance) for each test can be determined based on the degrees of freedom and alpha level.
+
+```{r}
+alpha <- .05
+N <- 200
+nGroup1 <- 150
+nGroup2 <- 150
+numGroups <- 4
+numLevelsFactorA <- 3
+numLevelsFactorB <- 4
+numPredictors <- 5
+```
+
+#### One-Sample *t*-Test {#sec-statisticalTestsCriticalValueOneSampleTtest}
+
+```{r}
+df_oneSampleTtest <- N - 1
+```
+
+One-tailed test:
+
+```{r}
+qt(1 - alpha, df_oneSampleTtest)
+```
+
+Two-tailed test:
+
+```{r}
+qt(1 - alpha/2, df_oneSampleTtest)
+```
+
+#### Independent-Samples *t*-Test {#sec-statisticalTestsCriticalValueIndependentSamplesTtest}
+
+```{r}
+df_independentSamplesTtest <- nGroup1 + nGroup2 - 2
+```
+
+One-tailed test:
+
+```{r}
+qt(1 - alpha, df_independentSamplesTtest)
+```
+
+Two-tailed test:
+
+```{r}
+qt(1 - alpha/2, df_independentSamplesTtest)
+```
+
+#### Paired-Samples *t*-Test {#sec-statisticalTestsCriticalValuePairedSamplesTtest}
+
+```{r}
+df_pairedSamplesTtest <- N - 1
+```
+
+One-tailed test:
+
+```{r}
+qt(1 - alpha, df_pairedSamplesTtest)
+```
+
+Two-tailed test:
+
+```{r}
+qt(1 - alpha/2, df_pairedSamplesTtest)
+```
+
+#### One-Way ANOVA {#sec-statisticalTestsCriticalValueOneWayANOVA}
+
+```{r}
+df_betweenOneWayANOVA <- numGroups - 1
+df_withinOneWayANOVA <- N - numGroups
+```
+
+One-tailed test:
+
+```{r}
+qf(1 - alpha, df_betweenOneWayANOVA, df_withinOneWayANOVA)
+```
+
+Two-tailed test:
+
+```{r}
+qf(1 - alpha/2, df_betweenOneWayANOVA, df_withinOneWayANOVA)
+```
+
+#### Factorial ANOVA {#sec-statisticalTestsCriticalValueFactorialANOVA}
+
+```{r}
+df_factorA <- numLevelsFactorA - 1
+df_factorB <- numLevelsFactorB - 1
+df_interaction <- df_factorA * df_factorB
+df_error <- numLevelsFactorA * numLevelsFactorB * (N - 1)
+```
+
+Factor A (one-tailed test):
+
+```{r}
+qf(1 - alpha, df_factorA, df_error)
+```
+
+Factor B (one-tailed test):
+
+```{r}
+qf(1 - alpha, df_factorB, df_error)
+```
+
+Interaction (one-tailed test):
+
+```{r}
+qf(1 - alpha, df_interaction, df_error)
+```
+
+Factor A (two-tailed test):
+
+```{r}
+qf(1 - alpha/2, df_factorA, df_error)
+```
+
+Factor B (two-tailed test):
+
+```{r}
+qf(1 - alpha/2, df_factorB, df_error)
+```
+
+Interaction (two-tailed test):
+
+```{r}
+qf(1 - alpha/2, df_interaction, df_error)
+```
+
+#### Repeated Measures ANOVA {#sec-statisticalTestsCriticalValueRMANOVA}
+
+```{r}
+df1_RMANOVA <- numGroups - 1
+df2_RMANOVA <- (numGroups - 1) * (N - 1)
+```
+
+One-tailed test:
+
+```{r}
+qf(1 - alpha, df1_RMANOVA, df2_RMANOVA)
+```
+
+Two-tailed test:
+
+```{r}
+qf(1 - alpha/2, df1_RMANOVA, df2_RMANOVA)
+```
+
+#### Correlation {#sec-statisticalTestsCriticalValueCorrelation}
+
+```{r}
+df_correlation <- N - 2
+```
+
+One-tailed test:
+
+```{r}
+qt(1 - alpha, df_correlation)
+```
+
+Two-tailed test:
+
+```{r}
+qt(1 - alpha/2, df_correlation)
+```
+
+#### Multiple Regression {#sec-statisticalTestsCriticalValueMultipleRegression}
+
+```{r}
+df1_regression <- numPredictors
+df2_regression <- N - numPredictors - 1
+```
+
+One-tailed test:
+
+```{r}
+qf(1 - alpha, df1_regression, df2_regression)
+```
+
+Two-tailed test:
+
+```{r}
+qf(1 - alpha/2, df1_regression, df2_regression)
+```

### 2024-06-24 basic-statistics.qmd
20240624 - df

@@
 The critical value for [statistical significance](#sec-statisticalSignificance) for each test can be determined based on the degrees of freedom and alpha level.
+The degrees of freedom (*df*) refer to the number of values in the calculation of a test statistic that are free to vary.
 
@@
 numLevelsFactorB <- 4
+numTimepoints <- 4
 numPredictors <- 5
@@
 #### One-Sample *t*-Test {#sec-statisticalTestsCriticalValueOneSampleTtest}
+
+For a one-sample *t*-test, the degrees of freedom is in @eq-dfOneSampleTtest:
+
+$$
+df = N - 1
+$$ {#eq-dfOneSampleTtest}
+
+where $N$ is sample size.
 
@@
 
+For an independent-samples *t*-test, the degrees of freedom is in @eq-dfIndependentSamplesTtest:
+
+$$
+df = n_1 + n_2 - 2
+$$ {#eq-dfIndependentSamplesTtest}
+
+where $n_1$ is the sample size of group 1 and $n_2$ is the sample size of group 2.
+
 ```{r}
@@
 #### Paired-Samples *t*-Test {#sec-statisticalTestsCriticalValuePairedSamplesTtest}
+
+For a paired-samples *t*-test, the degrees of freedom is in @eq-dfPairedSamplesTtest:
+
+$$
+df = N - 1
+$$ {#eq-dfPairedSamplesTtest}
+
+where $N$ is sample size (i.e., the number of paired observations).
 
@@
 
+For a one-way ANOVA, the degrees of freedom is in @eq-dfOneWayANOVA:
+
+$$
+df_\text{between} = g - 1
+df_\text{within} = N - g
+$$ {#eq-dfOneWayANOVA}
+
+where $N$ is sample size and $g$ is the number of groups.
+
 ```{r}
@@
 #### Factorial ANOVA {#sec-statisticalTestsCriticalValueFactorialANOVA}
+
+For a factorial two-way ANOVA, the degrees of freedom is in @eq-dfFactorialANOVA:
+
+$$
+df_\text{Factor A} = a - 1
+df_\text{Factor B} = b - 1
+df_\text{Interaction} = (a - 1)(b - 1)
+df_\text{error} = a * b * (N - 1)
+$$ {#eq-dfFactorialANOVA}
+
+where $N$ is sample size, $a$ is the number of levels for factor A, and $b$ is the number of levels for factor B.
 
@@
 
+For a repeated measures ANOVA, the degrees of freedom is in @eq-dfRepeatedMeasuresANOVA:
+
+$$
+df_1 = T - 1
+df_2 = (T - 1)(N - 1)
+$$ {#eq-dfRepeatedMeasuresANOVA}
+
+where $N$ is sample size and $T$ is the number of timepoints.
+
 ```{r}
-df1_RMANOVA <- numGroups - 1
-df2_RMANOVA <- (numGroups - 1) * (N - 1)
+df1_RMANOVA <- numTimepoints - 1
+df2_RMANOVA <- (numTimepoints - 1) * (N - 1)
 ```
@@
 #### Correlation {#sec-statisticalTestsCriticalValueCorrelation}
+
+For a correlation, the degrees of freedom is in @eq-dfCorrelation:
+
+$$
+df = N - 2
+$$ {#eq-dfCorrelation}
+
+where $N$ is sample size.
 
@@
 
+For multiple regression, the degrees of freedom is in @eq-dfMultipleRegression:
+
+$$
+df_1 = p
+df_2 = N - p - 1
+$$ {#eq-dfMultipleRegression}
+
+where $N$ is sample size and $p$ is the number of predictors.
+
 ```{r}

### 2024-06-24 basic-statistics.qmd
20240624 - align formulas

@@
 $$
-df_\text{between} = g - 1
-df_\text{within} = N - g
+\begin{aligned}
+  df_\text{between} &= g - 1 \\
+  df_\text{within} &= N - g
+\end{aligned}
 $$ {#eq-dfOneWayANOVA}
@@
 $$
-df_\text{Factor A} = a - 1
-df_\text{Factor B} = b - 1
-df_\text{Interaction} = (a - 1)(b - 1)
-df_\text{error} = a * b * (N - 1)
+\begin{aligned}
+  df_\text{Factor A} &= a - 1 \\
+  df_\text{Factor B} &= b - 1 \\
+  df_\text{Interaction} &= (a - 1)(b - 1) \\
+  df_\text{error} &= ab(N - 1)
+\end{aligned}
 $$ {#eq-dfFactorialANOVA}
@@
 $$
-df_1 = T - 1
-df_2 = (T - 1)(N - 1)
+\begin{aligned}
+  df_1 &= T - 1 \\
+  df_2 &= (T - 1)(N - 1)
+\end{aligned}
 $$ {#eq-dfRepeatedMeasuresANOVA}
 
-where $N$ is sample size and $T$ is the number of timepoints.
+where $N$ is sample size and $T$ is the number of timepoints (or conditions).
 
@@
 $$
-df_1 = p
-df_2 = N - p - 1
+\begin{aligned}
+  df_1 &= p \\
+  df_2 &= N - p - 1
+\end{aligned}
 $$ {#eq-dfMultipleRegression}

### 2024-06-24 basic-statistics.qmd
20240624 - chi-square test

@@
 A one-sample *t*-test is used to evaluate whether a sample mean differs systematically from a particular value.
+The null hypothesis is that the sample mean does not differ systematically from the pre-specified value.
+The alternative hypothesis is that the sample mean differs systematically from the pre-specified value.
 For instance, let's say you want to test out a new draft strategy.
@@
 If the two groups involve paired samples, the paired samples *t*-test is used.
+The null hypothesis is that the mean of group 1 does not differ systematically from the mean of group 2.
+The alternative hypothesis is that the mean of group 1 differs systematically from the mean of group 2.
 For instance, you could use an independent-samples *t*-test if you want to examine whether Quarterbacks tend to have have longer careers than Running Backs.
@@
 However, unlike a *t*-test, ANOVA allows examining multiple multiple [independent variables](#sec-experiment) and more than two groups.
+The null hypothesis is that the the groups' mean value does not differ systematically.
+The alternative hypothesis is that the groups' mean value differs systematically.
 
@@
 Correlation examines the association between a [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variable.
+The null hypothesis is that the the two variables are not associated.
+The alternative hypothesis is that the two variables are associated.
 
@@
 
+The null hypothesis is that the the [predictor variable(s)](#sec-correlationalStudy) are not associated with the [outcome variable](#sec-correlationalStudy).
+The alternative hypothesis is that the [predictor variable(s)](#sec-correlationalStudy) are associated with the [outcome variable](#sec-correlationalStudy).
+
+### Chi-Square Tests {#sec-chisquare}
+
+There are two primary types of chi-square tests:
+
+- chi-square goodness-of-fit test
+- chi-square test for association (aka test of independence)
+
+The chi-square goodness-of-fit test evaluates whether a set of categorical data came from a specified distribution.
+The null hypothesis is that the data came from the specified distribution.
+The alternative hypothesis is that the data did not come from the specified distribution.
+
+The chi-square test for association evaluates whether two categorical variables are associated.
+The null hypothesis is that the two variables are not associated.
+The alternative hypothesis is that the two variables are associated.
+
 ### Formulating Statistical Tests in Terms of Partitioned Variance {#sec-statisticalTestsPartitionedVariance}
@@
 numPredictors <- 5
+numCategories <- 6
+numRows <- 5
+numColumns <- 2
 ```
@@
 ```
+
+#### Chi-Square Goodness-of-Fit Test {#sec-statisticalTestsCriticalValueChiSquareGOF}
+
+For the chi-square goodness-of-fit test, the degrees of freedom is in @eq-dfChiSquareGOF:
+
+$$
+df = c - 1
+$$ {#eq-dfChiSquareGOF}
+
+where $c$ is the number of categories.
+
+```{r}
+df_chisquareGOF <- numCategories - 1
+```
+
+One-tailed test:
+
+```{r}
+qchisq(1 - alpha, df_chisquareGOF)
+```
+
+Two-tailed test:
+
+```{r}
+qchisq(1 - alpha/2, df_chisquareGOF)
+```
+
+#### Chi-Square Test for Association {#sec-statisticalTestsCriticalValueChiSquareAssociation}
+
+For the chi-square test for association, the degrees of freedom is in @eq-dfChiSquareAssociation:
+
+$$
+df = (r - 1) \times (c - 1)
+$$ {#eq-dfChiSquareAssociation}
+
+where $r$ is the number of rows in the contingency table and $c$ is the number of columns in the contingency table.
+
+```{r}
+df_chisquareAssociation <- (numRows - 1) * (numColumns - 1)
+```
+
+One-tailed test:
+
+```{r}
+qchisq(1 - alpha, df_chisquareAssociation)
+```
+
+Two-tailed test:
+
+```{r}
+qchisq(1 - alpha/2, df_chisquareAssociation)
+```

### 2024-06-24 basic-statistics.qmd
20240624 - statistical power

@@
 library("petersenlab")
+library("pwr")
 library("grid")
@@
 
+Cohen's $d$ is calculated as in @eq-cohensD:
+
+$$
+\begin{aligned}
+  d &= \frac{\text{mean difference}}{\text{pooled standard deviation}} \\
+   &= \frac{M_2 - M_1}{\text{pooled standard deviation}} \\
+\end{aligned}
+$$ {#eq-cohensD}
+
 Effect size thresholds [@McGrath2006] for small, medium, and large effect sizes are in @tbl-effectSizeThresholds.
@@
 
-### Chi-Square Tests {#sec-chisquare}
+### Chi-Square Test {#sec-chisquare}
 
@@
 ```
+
+### Statistical Power {#sec-statisticalPower}
+
+As described above, *statistical power* is the probability of detecting an effect, if, in fact, the effect exists.
+Statistical power for a given test can be calculated based on three factors:
+
+- [effect size](#sec-practicalSignificance)
+- sample size
+- [alpha level](#sec-statisticalSignificance)
+
+Knowing any three of the following, you can calculate the fourth: statistical power, effect size, sample size, and alpha level.
+
+```{r}
+power <- .8
+effectSize_d <- .5
+effectSize_r <- .24
+```
+
+#### One-Sample *t*-Test {#sec-statisticalPowerOneSampleTtest}
+
+Solving for statistical power achieved (given effect size, sample size, and alpha):
+
+```{r}
+pwr::pwr.t.test(
+  n = N,
+  d = effectSize_d,
+  sig.level = alpha,
+  type = "one.sample",
+  alternative = "two.sided")
+```
+
+Solving for sample size needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.t.test(
+  d = effectSize_d,
+  power = power,
+  sig.level = alpha,
+  type = "one.sample",
+  alternative = "two.sided")
+```
+
+Solving for minimum effect size needed to detect (given sample size, power, and alpha):
+
+```{r}
+pwr::pwr.t.test(
+  n = N,
+  power = power,
+  sig.level = alpha,
+  type = "one.sample",
+  alternative = "two.sided")
+```
+
+#### Independent-Samples *t*-Test {#sec-statisticalPowerIndependentSamplesTtest}
+
+Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+
+```{r}
+pwr::pwr.t.test(
+  n = N,
+  d = effectSize_d,
+  sig.level = alpha,
+  type = "two.sample",
+  alternative = "two.sided")
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.t.test(
+  d = effectSize_d,
+  power = power,
+  sig.level = alpha,
+  type = "two.sample",
+  alternative = "two.sided")
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+pwr::pwr.t.test(
+  n = N,
+  power = power,
+  sig.level = alpha,
+  type = "two.sample",
+  alternative = "two.sided")
+```
+
+#### Paired-Samples *t*-Test {#sec-statisticalPowerPairedSamplesTtest}
+
+Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+
+```{r}
+pwr::pwr.t.test(
+  n = N,
+  d = effectSize_d,
+  sig.level = alpha,
+  type = "paired",
+  alternative = "two.sided")
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.t.test(
+  d = effectSize_d,
+  power = power,
+  sig.level = alpha,
+  type = "paired",
+  alternative = "two.sided")
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+pwr::pwr.t.test(
+  n = N,
+  power = power,
+  sig.level = alpha,
+  type = "paired",
+  alternative = "two.sided")
+```
+
+#### One-Way ANOVA {#sec-statisticalPowerOneWayANOVA}
+
+#### Factorial ANOVA {#sec-statisticalPowerFactorialANOVA}
+
+#### Repeated Measures ANOVA {#sec-statisticalPowerRepeatedMeasuresANOVA}
+
+#### Correlation {#sec-statisticalPowerCorrelation}
+
+#### Multiple Regression {#sec-statisticalPowerMultipleRegression}
+
+#### Chi-Square Goodness-of-Fit Test {#sec-statisticalPowerChiSquareGOF}
+
+#### Chi-Square Test for Association {#sec-statisticalPowerChiSquareAssociation}

### 2024-06-24 basic-statistics.qmd
20240624 - power

@@
 ```{r}
-pwr::pwr.t.test(
+pwr::pwr.t.test( # balanced group sizes
   n = N,
@@
   alternative = "two.sided")
+
+pwr::pwr.t2n.test( # unbalanced group sizes
+  n1 = nGroup1,
+  n2 = nGroup2,
+  d = effectSize_d,
+  sig.level = alpha,
+  alternative = "two.sided")
 ```
@@
 ```{r}
-pwr::pwr.t.test(
+pwr::pwr.t.test( # balanced group sizes
   d = effectSize_d,
@@
   alternative = "two.sided")
+
+pwr::pwr.t2n.test( # unbalanced group sizes
+  n1 = nGroup1,
+  d = effectSize_d,
+  power = power,
+  sig.level = alpha,
+  alternative = "two.sided")
 ```
@@
   type = "two.sample",
+  alternative = "two.sided")
+
+pwr::pwr.t2n.test( # unbalanced group sizes
+  n1 = nGroup1,
+  n2 = nGroup2,
+  power = power,
+  sig.level = alpha,
   alternative = "two.sided")

### 2024-06-24 basic-statistics.qmd
20240624 - balanced vs unbalanced group sizes

--- 2024-06-24 basic-statistics.qmd
@@
 
+##### Balanced Group Sizes {#sec-statisticalPowerIndependentSamplesTtestBalanced}
+
 Solving for statistical power achieved (given effect size, sample size per group, and alpha):
@@
 
+##### Unbalanced Group Sizes {#sec-statisticalPowerIndependentSamplesTtestUnbalanced}
+
+Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+
+```{r}
+pwr::pwr.t2n.test(
+  n1 = nGroup1,
+  n2 = nGroup2,
+  d = effectSize_d,
+  sig.level = alpha,
+  alternative = "two.sided")
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.t2n.test(
+  n1 = nGroup1,
+  d = effectSize_d,
+  power = power,
+  sig.level = alpha,
+  alternative = "two.sided")
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+pwr::pwr.t2n.test(
+  n1 = nGroup1,
+  n2 = nGroup2,
+  power = power,
+  sig.level = alpha,
+  alternative = "two.sided")
+```
+
 #### Paired-Samples *t*-Test {#sec-statisticalPowerPairedSamplesTtest}

### 2024-06-24 basic-statistics.qmd
20240624 - coefficient formulas

@@
   d &= \frac{\text{mean difference}}{\text{pooled standard deviation}} \\
-   &= \frac{M_2 - M_1}{\text{pooled standard deviation}} \\
+   &= \frac{X_1 - X_2}{s} \\
 \end{aligned}
 $$ {#eq-cohensD}
+
+where:
+
+$$
+s = \sqrt{\frac{(n_1 - 1)s_1^2 + (n_2 - 1)s_2^2}{n_1 + n_2 - 2}}
+$$ {#eq-pooledStandardDeviation}
+
+where $n_1$ and $n_2$ is the sample size of group 1 and group 2, respectively, and $s_1$ and $s_2$ is the standard deviation of group 1 and group 2, respectively.
+
+Cohen's $f$ is commonly used in ANOVA, and is calculated as in @eq-cohensF:
+
+$$
+\begin{aligned}
+  f &= \sqrt{\frac{R^2}{1 - R^2}} \\
+    &= \sqrt{\frac{\eta^2}{1 - \eta^2}} \\
+\end{aligned}
+$$ {#eq-cohensF}
+
+Cohen's $f^2$ is commonly used in regression, and is calculated as in @eq-cohensFsquared:
+
+$$
+\begin{aligned}
+  f^2 &= \frac{R^2}{1 - R^2} \\
+    &= \frac{\eta^2}{1 - \eta^2} \\
+\end{aligned}
+$$ {#eq-cohensFsquared}
+
+$R^2$ is commonly used in regression, and is calculated as in @eq-rSquared:
+
+$$
+\begin{aligned}
+  R^2 &= 1 - \frac{\sum (Y_i - \hat{Y}_i)^2}{\sum (Y_i - \bar{Y})^2} \\
+      &= 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}} \\
+      &= 1 - \frac{\text{sum of squared residuals}}{\text{total sum of squares}}
+
+\end{aligned}
+$$ {#eq-rSquared}
+
+where $Y_i$ is the observed value of the [outcome variable](#sec-correlationalStudy) for the $i$th observation, $\hat{Y}_i$ is the model predicted value for the $i$th observation, $\bar{Y}$ is the mean of the observed values of the [outcome variable](#sec-correlationalStudy).
+The total sum of squares is an index of the total variation in the [outcome variable](#sec-correlationalStudy).
 
@@
 | Correlation coefficient ($r$)                       | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
+| Cohen's $f$                                         | $\ge .10$   | $\ge .25$   | $\ge .40$   |
+| Cohen's $f^2$                                       | $\ge .01$   | $\ge .06$   | $\ge .16$   |
 | Coefficient of determination ($R^2$)                | $\ge .01$   | $\ge .06$   | $\ge .14$   |
@@
 The alternative hypothesis is that the two variables are associated.
+
+The Pearson correlation coefficient ($r$) is calculated as in @eq-pearsonCorrelation:
+
+$$
+r = \frac{\sum (X_i - \bar{X})(Y_i - \bar{Y})}{\sqrt{\sum (X_i - \bar{X})^2 \sum (Y_i - \bar{Y})^2}}
+$$ {#eq-pearsonCorrelation}
+
+where $X$ is the [predictor variable](#sec-correlationalStudy) and $Y$ is the [outcome variable](#sec-correlationalStudy).
 

### 2024-06-24 basic-statistics.qmd
20240624 - attempt to fix error
20240624 - more power tests

@@
   f &= \sqrt{\frac{R^2}{1 - R^2}} \\
-    &= \sqrt{\frac{\eta^2}{1 - \eta^2}} \\
+    &= \sqrt{\frac{\eta^2}{1 - \eta^2}}
 \end{aligned}
@@
   f^2 &= \frac{R^2}{1 - R^2} \\
-    &= \frac{\eta^2}{1 - \eta^2} \\
+      &= \frac{\eta^2}{1 - \eta^2}
 \end{aligned}
@@
       &= 1 - \frac{\text{sum of squared residuals}}{\text{total sum of squares}}
-
 \end{aligned}
@@
 effectSize_r <- .24
+effectSize_f <- .25
+effectSize_fSquared <- .06
 ```
@@
 
+Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+
+```{r}
+pwr::pwr.anova.test(
+  n = N,
+  f = effectSize_f,
+  sig.level = alpha,
+  k = numGroups)
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.anova.test(
+  f = effectSize_f,
+  power = power,
+  sig.level = alpha,
+  k = numGroups)
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+pwr::pwr.anova.test(
+  n = N,
+  power = power,
+  sig.level = alpha,
+  k = numGroups)
+```
+
 #### Factorial ANOVA {#sec-statisticalPowerFactorialANOVA}
 
+Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+
+```{r}
+
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha):
+
+```{r}
+
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+
+```
+
 #### Repeated Measures ANOVA {#sec-statisticalPowerRepeatedMeasuresANOVA}
 
+Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+
+```{r}
+
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha):
+
+```{r}
+
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+
+```
+
 #### Correlation {#sec-statisticalPowerCorrelation}
 
+Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+
+```{r}
+pwr::pwr.r.test(
+  n = N,
+  r = effectSize_r,
+  sig.level = alpha,
+  alternative = "two.sided")
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.r.test(
+  r = effectSize_r,
+  power = power,
+  sig.level = alpha,
+  alternative = "two.sided")
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+pwr::pwr.r.test(
+  n = N,
+  power = power,
+  sig.level = alpha,
+  alternative = "two.sided")
+```
+
 #### Multiple Regression {#sec-statisticalPowerMultipleRegression}
 
+Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+
+```{r}
+pwr::pwr.f2.test(
+  f2 = effectSize_fSquared,
+  sig.level = alpha,
+  u = numPredictors,
+  v = N - numPredictors - 1)
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha)—$v = N - \text{numberOfPredictors} - 1$; thus, $N = v + \text{numberOfPredictors} + 1$:
+
+```{r}
+multipleRegressionSampleSizeModel <- pwr::pwr.f2.test(
+  f2 = effectSize_fSquared,
+  power = power,
+  sig.level = alpha,
+  u = numPredictors)
+
+multipleRegressionSampleSizeModel
+
+vNeeded <- multipleRegressionSampleSizeModel$v
+sampleSizeNeeded <- vNeeded + numPredictors + 1
+sampleSizeNeeded
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+pwr::pwr.f2.test(
+  power = power,
+  sig.level = alpha,
+  u = numPredictors,
+  v = N - numPredictors - 1)
+```
+
 #### Chi-Square Goodness-of-Fit Test {#sec-statisticalPowerChiSquareGOF}
 
+```{r}
+
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha):
+
+```{r}
+
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+
+```
+
 #### Chi-Square Test for Association {#sec-statisticalPowerChiSquareAssociation}
+
+```{r}
+
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha):
+
+```{r}
+
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+
+```

### 2024-06-24 basic-statistics.qmd
20240624 - free power tools

@@
 Knowing any three of the following, you can calculate the fourth: statistical power, effect size, sample size, and alpha level.
+Below is `R` code for calculating power for each of various statistical tests (i.e., a *power analysis*).
+For free point-and-click software for calculating statistical power, see G\*Power: <https://www.psychologie.hhu.de/arbeitsgruppen/allgemeine-psychologie-und-arbeitspsychologie/gpower.html>
 
@@
 
+Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+
 ```{r}
@@
 ```
+
+#### Multilevel Modeling {#sec-statisticalPowerMultilevelModeling}
+
+Power analysis for multilevel modeling approaches is more complicated than it is for other statistical analyses, such as correlation, multiple regression, *t*-tests, ANOVA, etc.
+
+There are free web applications for calculating power in multilevel modeling:
+
+- <https://aguinis.shinyapps.io/ml_power/>
+- <https://koumurayama.shinyapps.io/tmethod_mlm/>
+- <https://webpower.psychstat.org/wiki/models/index>
+
+#### Path Analysis, Factor Analysis, and Structural Equation Modeling {#sec-statisticalPowerSEM}
+
+Power analysis for latent variable modeling approaches like SEM is more complicated than it is for other statistical analyses, such as correlation, multiple regression, *t*-tests, ANOVA, etc.
+
+I provide an example of power analysis in SEM using Monte Carlo simulation in R here: <https://isaactpetersen.github.io/Principles-Psychological-Assessment/sem.html#monteCarloPowerAnalysis> [@PetersenPrinciplesPsychAssessment].
+
+There are also free web applications for calculating power in SEM:
+
+- <https://sjak.shinyapps.io/power4SEM/>
+- <https://sempower.shinyapps.io/sempower/>
+- <https://yilinandrewang.shinyapps.io/pwrSEM/>
+- <https://webpower.psychstat.org/wiki/models/index>
+
+#### Mediation and Moderation
+
+There are free tools for calculating power for tests of mediation and moderation:
+
+- <https://www.causalevaluation.org/power-analysis.html> (web application: <https://powerupr.shinyapps.io/index/>)
+- <https://webpower.psychstat.org/wiki/models/index>

### 2024-06-24 basic-statistics.qmd
20240624 - repeated measures ANOVA

@@
 library("pwr")
+library("WebPower")
 library("grid")
@@
       &= 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}} \\
-      &= 1 - \frac{\text{sum of squared residuals}}{\text{total sum of squares}}
+      &= 1 - \frac{\text{sum of squared residuals}}{\text{total sum of squares}} \\
+      &= \frac{f^2}{1 + f^2}
 \end{aligned}
@@
 numLevelsFactorB <- 4
-numTimepoints <- 4
+numMeasurements <- 4
 numPredictors <- 5
@@
 
-where $N$ is sample size and $T$ is the number of timepoints (or conditions).
-
-```{r}
-df1_RMANOVA <- numTimepoints - 1
-df2_RMANOVA <- (numTimepoints - 1) * (N - 1)
+where $N$ is sample size and $T$ is the number of measurements (i.e., the number of levels of the within-person factor: e.g., timepoints or conditions).
+
+```{r}
+df1_RMANOVA <- numMeasurements - 1
+df2_RMANOVA <- (numMeasurements - 1) * (N - 1)
 ```
@@
 ```{r}
-
+wp.rmanova(
+  n = N,
+  ng = numGroups,
+  nm = numMeasurements,
+  f = effectSize_f,
+  alpha = alpha,
+  type = 0)
+
+wp.rmanova(
+  n = N,
+  ng = numGroups,
+  nm = numMeasurements,
+  f = effectSize_f,
+  alpha = alpha,
+  type = 1)
+
+wp.rmanova(
+  n = N,
+  ng = numGroups,
+  nm = numMeasurements,
+  f = effectSize_f,
+  alpha = alpha,
+  type = 2)
 ```
@@
 ```{r}
-
+wp.rmanova(
+  power = power,
+  ng = numGroups,
+  nm = numMeasurements,
+  f = effectSize_f,
+  alpha = alpha,
+  type = 0)
+
+wp.rmanova(
+  power = power,
+  ng = numGroups,
+  nm = numMeasurements,
+  f = effectSize_f,
+  alpha = alpha,
+  type = 1)
+
+wp.rmanova(
+  power = power,
+  ng = numGroups,
+  nm = numMeasurements,
+  f = effectSize_f,
+  alpha = alpha,
+  type = 2)
 ```
@@
 ```{r}
-
+wp.rmanova(
+  n = N,
+  ng = numGroups,
+  nm = numMeasurements,
+  power = power,
+  alpha = alpha,
+  type = 0)
+
+wp.rmanova(
+  n = N,
+  ng = numGroups,
+  nm = numMeasurements,
+  power = power,
+  alpha = alpha,
+  type = 1)
+
+wp.rmanova(
+  n = N,
+  ng = numGroups,
+  nm = numMeasurements,
+  power = power,
+  alpha = alpha,
+  type = 2)
 ```

### 2024-06-25 basic-statistics.qmd
20240624 - more formulas

@@
 - Correlation coefficient ($r$)
+- Cohen's $\omega$ (omega)
+- Cohen's $f$
+- Cohen's $f^2$
 - Coefficient of determination ($R^2$)
@@
 - Partial eta squared ($\eta_p^2$)
+
+#### Cohen's $d$ {#sec-cohensD}
 
@@
 
+#### Standardized Regression Coefficient (Beta; $\beta$) {#sec-beta}
+
+The standardized regression coefficient (beta; $\beta$) is used in multiple regression, and is calculated as in @eq-beta:
+
+$$
+\beta_x = B_x \times \frac{s_x}{s_y}
+$$ {#eq-beta}
+
+where $B_x$ is the unstandardized regression coefficient of the [predictor variable](#sec-correlationalStudy) $x$ in predicting the [outcome variable](#sec-correlationalStudy) $y$, $s_x$ is the standard deviation of $x$, and $s_y$ is the standard deviation of $y$.
+
+#### Correlation Coefficient ($r$)
+
+The formula for the correlation coefficient is in @sec-correlation.
+
+#### Cohen's $\omega$ {#sec-cohensOmega}
+
+Cohen's $\omega$ is used in chi-square tests, and is calculated as in @eq-cohensOmega:
+
+$$
+\omega = \sqrt{\frac{\chi^2}{N} - \frac{df}{N}}
+$$ {#eq-cohensOmega}
+
+where $\chi^2$ is the chi-square statistic from the test, $N$ is the sample size, and $df$ is the degrees of freedom.
+
+#### Cohen's $f$ {#sec-cohensF}
+
 Cohen's $f$ is commonly used in ANOVA, and is calculated as in @eq-cohensF:
@@
 
+#### Cohen's $f^2$ {#sec-cohensFsquared}
+
 Cohen's $f^2$ is commonly used in regression, and is calculated as in @eq-cohensFsquared:
@@
 
+#### Coefficient of Determination ($R^2$) {#sec-rSquared}
+
+The coefficient of determination ($R^2$) reflects the proportion of variance in the [outcome variable](#sec-correlationalStudy) that is explained by the [predictor variable(s)](#sec-correlationalStudy).
 $R^2$ is commonly used in regression, and is calculated as in @eq-rSquared:
@@
       &= 1 - \frac{\text{sum of squared residuals}}{\text{total sum of squares}} \\
-      &= \frac{f^2}{1 + f^2}
+      &= \frac{f^2}{1 + f^2} \\
+      &= \eta^2
 \end{aligned}
@@
 The total sum of squares is an index of the total variation in the [outcome variable](#sec-correlationalStudy).
+
+#### Eta Squared ($\eta^2$) and Partial Eta Squared ($\eta_p^2$) {#sec-etaSquared}
+
+Like $R^2$, eta squared ($\eta^2$) reflects the proportion of variance in the [dependent variable](#sec-experiment) that is explained by the [independent variable(s)](#sec-experiment).
+$\eta^2$ is commonly used in ANOVA, and is calculated as in @eq-etaSquared:
+
+$$
+\begin{aligned}
+  \eta^2 &= \frac{SS_{\text{effect}}}{SS_{\text{total}}} \\
+      &= 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}} \\
+      &= 1 - \frac{\text{sum of squared residuals}}{\text{total sum of squares}} \\
+      &= \frac{f^2}{1 + f^2} \\
+      &= R^2
+\end{aligned}
+$$ {#eq-etaSquared}
+
+where $SS_{\text{effect}}$ is the sum of squares for the effect of interest and SS_{\text{total}} is the total sum of squares.
+
+Partial eta squared ($\eta_p^2$) reflects the proportion of variance in the [dependent variable](#sec-experiment) that is explained by the [independent variable](#sec-experiment) while controlling for the other [independent variables](#sec-experiment).
+$\eta_p^2$ is commonly used in ANOVA, and is calculated as in @eq-partialEtaSquared:
+
+$$
+\eta_p^2 = = \frac{SS_{\text{effect}}}{SS_{\text{effect}} + SS_{\text{error}}}
+$$ {#eq-partialEtaSquared}
+
+where $SS_{\text{effect}}$ is the sum of squares for the effect of interest and SS_{\text{error}} is the sum of squares for the residual error term.
+
+#### Effect Size Thresholds {#sec-effectSizeThresholds}
 
@@
 | Correlation coefficient ($r$)                       | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
+| Cohen's $\omega$                                    | $\ge .10$   | $\ge .30$   | $\ge .50$   |
 | Cohen's $f$                                         | $\ge .10$   | $\ge .25$   | $\ge .40$   |
@@
 The coefficient of determination ($R^2$) is the proportion of variance in the [outcome variable](#sec-correlationalStudy) that is explained by the [predictor variables](#sec-correlationalStudy).
-The coefficient of determination can be expressed as the ratio of variance explained in the [outcome variable](#sec-correlationalStudy) to the total variance in the [outcome variable](#sec-correlationalStudy), as in @eq-coefficientOfDetermination:
+\eta^2 is the proportion of variance in the [dependent variable](#sec-experiment) that is explained by the [independent variables](#sec-experiment).
+The coefficient of determination and \eta^2 can be expressed as the ratio of variance explained in the [outcome](#sec-correlationalStudy) or [dependent](#sec-experiment) variable to the total variance in the [outcome](#sec-correlationalStudy) or [dependent](#sec-experiment) variable, as in @eq-coefficientOfDeterminationPartitionedVariance and @eq-etaSquaredPartitionedVariance:
 
@@
 R^2 = \frac{\text{variance explained in the outcome variable}}{\text{total variance in the outcome variable}}
-$$ {#eq-coefficientOfDetermination}
+$$ {#eq-coefficientOfDeterminationPartitionedVariance}
+
+$$
+eta^2 = \frac{\text{variance explained in the dependent variable}}{\text{total variance in the dependent variable}}
+$$ {#eq-etaSquaredPartitionedVariance}
 

### 2024-06-25 basic-statistics.qmd
20240624 - chi-square goodness-of-fit test

@@
 effectSize_fSquared <- .06
+effectSize_omega <- .3
 ```
@@
 
-Solving for statistical power achieved (given effect size, sample size per group, and alpha):
-
-```{r}
-
-```
-
-Solving for sample size per group needed (given effect size, power, and alpha):
-
-```{r}
-
-```
-
-Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
-
-```{r}
-
+Solving for statistical power achieved (given effect size, sample size, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  N = N,
+  w = effectSize_omega,
+  df = numCategories - 1,
+  sig.level = alpha)
+```
+
+Solving for sample size needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  w = effectSize_omega,
+  power = power,
+  df = numCategories - 1,
+  sig.level = alpha)
+```
+
+Solving for minimum effect size needed to detect (given sample size, power, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  N = N,
+  power = power,
+  df = numCategories - 1,
+  sig.level = alpha)
 ```
@@
 
-Solving for statistical power achieved (given effect size, sample size per group, and alpha):
-
-```{r}
-
-```
-
-Solving for sample size per group needed (given effect size, power, and alpha):
-
-```{r}
-
-```
-
-Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
-
-```{r}
-
+Solving for statistical power achieved (given effect size, sample size, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  N = N,
+  w = effectSize_omega,
+  df = (numRows - 1)*(numColumns - 1),
+  sig.level = alpha)
+```
+
+Solving for sample size needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  w = effectSize_omega,
+  power = power,
+  df = (numRows - 1)*(numColumns - 1),
+  sig.level = alpha)
+```
+
+Solving for minimum effect size needed to detect (given sample size, power, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  N = N,
+  power = power,
+  df = (numRows - 1)*(numColumns - 1),
+  sig.level = alpha)
 ```

### 2024-06-25 basic-statistics.qmd
20240624 - rearrange for consistency

--- 2024-06-25 basic-statistics.qmd
@@
 library("pwr")
+library("pwrss")
 library("WebPower")
@@
 
+To calculate the effect size of a particular predictor, you can calculate $\Delta f^2$ as in @eq-deltaCohensFsquared:
+
+$$
+\begin{aligned}
+  \Delta f^2 &= \frac{R^2_{\text{model}} - R^2_{\text{reduced}}}{1 - R^2_{\text{model}}} \\
+             &= \frac{\eta^2_{\text{model}} - \eta^2_{\text{reduced}}}{1 - \eta^2_{\text{model}}}
+\end{aligned}
+$$ {#eq-deltaCohensFsquared}
+
+where $R^2_{\text{model}$ is the $R^2$ of the model with the [predictor variable](#sec-correlationalStudy) of interest and R^2_{\text{reduced}} is the $R^2$ of the model without the [predictor variable](#sec-correlationalStudy) of interest.
+
 #### Coefficient of Determination ($R^2$) {#sec-rSquared}
@@
 $$
-\eta_p^2 = = \frac{SS_{\text{effect}}}{SS_{\text{effect}} + SS_{\text{error}}}
+\eta_p^2 = \frac{SS_{\text{effect}}}{SS_{\text{effect}} + SS_{\text{error}}}
 $$ {#eq-partialEtaSquared}
@@
 effectSize_r <- .24
+effectSize_beta <- .24
 effectSize_f <- .25
 effectSize_fSquared <- .06
+effectSize_omega <- .3
 ```
@@
 pwr::pwr.t.test(
+  power = power,
   d = effectSize_d,
-  power = power,
   sig.level = alpha,
@@
 pwr::pwr.t.test(
-  n = N,
-  power = power,
+  power = power,
+  n = N,
   sig.level = alpha,
@@
 pwr::pwr.t2n.test(
+  power = power,
   n1 = nGroup1,
   d = effectSize_d,
-  power = power,
   sig.level = alpha,
@@
 pwr::pwr.t2n.test(
+  power = power,
   n1 = nGroup1,
   n2 = nGroup2,
-  power = power,
   sig.level = alpha,
@@
 pwr::pwr.anova.test(
+  power = power,
   f = effectSize_f,
-  power = power,
   sig.level = alpha,
@@
 pwr::pwr.anova.test(
-  n = N,
-  power = power,
+  power = power,
+  n = N,
   sig.level = alpha,
@@
 ```{r}
-wp.rmanova(
+WebPower::wp.rmanova(
   n = N,
@@
 
-wp.rmanova(
+WebPower::wp.rmanova(
   n = N,
@@
 ```{r}
-wp.rmanova(
+WebPower::wp.rmanova(
   power = power,
@@
 
-wp.rmanova(
+WebPower::wp.rmanova(
   power = power,
@@
 ```{r}
-wp.rmanova(
+WebPower::wp.rmanova(
+  power = power,
   n = N,
@@
   nm = numMeasurements,
-  power = power,
   alpha = alpha,
@@
 
-wp.rmanova(
+WebPower::wp.rmanova(
+  power = power,
   n = N,
@@
 pwr::pwr.r.test(
+  power = power,
   r = effectSize_r,
-  power = power,
   sig.level = alpha,
@@
 pwr::pwr.r.test(
-  n = N,
-  power = power,
+  power = power,
+  n = N,
   sig.level = alpha,
@@
 
-Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+Solving for statistical power achieved (given effect size, sample size, and alpha):
 
@@
   v = N - numPredictors - 1)
-```
-
-Solving for sample size per group needed (given effect size, power, and alpha)—$v = N - \text{numberOfPredictors} - 1$; thus, $N = v + \text{numberOfPredictors} + 1$:
+
+pwrss::pwrss.t.reg(
+  n = N,
+  beta1 = effectSize_beta,
+  k = numPredictors,
+  alpha = alpha,
+  alternative = "not equal")
+```
+
+Solving for sample size needed (given effect size, power, and alpha)—$v = N - \text{numberOfPredictors} - 1$; thus, $N = v + \text{numberOfPredictors} + 1$:
 
@@
 multipleRegressionSampleSizeModel <- pwr::pwr.f2.test(
+  power = power,
   f2 = effectSize_fSquared,
-  power = power,
   sig.level = alpha,
@@
 sampleSizeNeeded
-```
-
-Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+pwrss::pwrss.t.reg(
+  power = power,
+  beta1 = effectSize_beta,
+  k = numPredictors,
+  alpha = alpha,
+  alternative = "not equal")
+```
+
+Solving for minimum effect size needed to detect (given sample size, power, and alpha):
 
@@
 
-Solving for statistical power achieved (given effect size, sample size per group, and alpha):
-
-```{r}
-
-```
-
-Solving for sample size per group needed (given effect size, power, and alpha):
-
-```{r}
-
-```
-
-Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
-
-```{r}
-
+Solving for statistical power achieved (given effect size, sample size, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  N = N,
+  w = effectSize_omega,
+  df = numCategories - 1,
+  sig.level = alpha)
+```
+
+Solving for sample size needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  power = power,
+  w = effectSize_omega,
+  df = numCategories - 1,
+  sig.level = alpha)
+```
+
+Solving for minimum effect size needed to detect (given sample size, power, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  power = power,
+  N = N,
+  df = numCategories - 1,
+  sig.level = alpha)
 ```
@@
 
-Solving for statistical power achieved (given effect size, sample size per group, and alpha):
-
-```{r}
-
-```
-
-Solving for sample size per group needed (given effect size, power, and alpha):
-
-```{r}
-
-```
-
-Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
-
-```{r}
-
+Solving for statistical power achieved (given effect size, sample size, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  N = N,
+  w = effectSize_omega,
+  df = (numRows - 1)*(numColumns - 1),
+  sig.level = alpha)
+```
+
+Solving for sample size needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  power = power,
+  w = effectSize_omega,
+  df = (numRows - 1)*(numColumns - 1),
+  sig.level = alpha)
+```
+
+Solving for minimum effect size needed to detect (given sample size, power, and alpha):
+
+```{r}
+pwr::pwr.chisq.test(
+  power = power,
+  N = N,
+  df = (numRows - 1)*(numColumns - 1),
+  sig.level = alpha)
 ```

### 2024-06-25 basic-statistics.qmd
20240624 - tweaks

@@
 \eta^2 is the proportion of variance in the [dependent variable](#sec-experiment) that is explained by the [independent variables](#sec-experiment).
-The coefficient of determination and \eta^2 can be expressed as the ratio of variance explained in the [outcome](#sec-correlationalStudy) or [dependent](#sec-experiment) variable to the total variance in the [outcome](#sec-correlationalStudy) or [dependent](#sec-experiment) variable, as in @eq-coefficientOfDeterminationPartitionedVariance and @eq-etaSquaredPartitionedVariance:
-
-$$
-R^2 = \frac{\text{variance explained in the outcome variable}}{\text{total variance in the outcome variable}}
+The coefficient of determination and \eta^2 can be expressed as the ratio of variance explained in the [outcome](#sec-correlationalStudy) or [dependent](#sec-experiment) variable to the total variance in the [outcome](#sec-correlationalStudy) or [dependent](#sec-experiment) variable, as in @eq-coefficientOfDeterminationPartitionedVariance:
+
+$$
+R^2 \text{ or } \eta^2 = \frac{\text{variance explained in the outcome variable}}{\text{total variance in the outcome variable}}
 $$ {#eq-coefficientOfDeterminationPartitionedVariance}
-
-$$
-eta^2 = \frac{\text{variance explained in the dependent variable}}{\text{total variance in the dependent variable}}
-$$ {#eq-etaSquaredPartitionedVariance}
 
@@
 
-#### One-Way ANOVA {#sec-statisticalPowerOneWayANOVA}
+#### ANOVA {#sec-statisticalPowerANOVA}
 
@@
   k = numGroups)
-```
-
-#### Factorial ANOVA {#sec-statisticalPowerFactorialANOVA}
-
-Solving for statistical power achieved (given effect size, sample size per group, and alpha):
-
-```{r}
-
-```
-
-Solving for sample size per group needed (given effect size, power, and alpha):
-
-```{r}
-
-```
-
-Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
-
-```{r}
-
 ```

### 2024-06-25 basic-statistics.qmd
20240624 - fixes

@@
 
-where $R^2_{\text{model}$ is the $R^2$ of the model with the [predictor variable](#sec-correlationalStudy) of interest and R^2_{\text{reduced}} is the $R^2$ of the model without the [predictor variable](#sec-correlationalStudy) of interest.
+where $R^2_{\text{model}}$ is the $R^2$ of the model with the [predictor variable](#sec-correlationalStudy) of interest and $R^2_{\text{reduced}}$ is the $R^2$ of the model without the [predictor variable](#sec-correlationalStudy) of interest.
 
@@
 
-#### ANOVA {#sec-statisticalPowerANOVA}
+#### One-Way ANOVA {#sec-statisticalPowerOneWayANOVA}
 
@@
   k = numGroups)
+```
+
+#### Factorial ANOVA {#sec-statisticalPowerFactorialANOVA}
+
+Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+
+```{r}
+pwr::pwr.anova.test(
+  n = N,
+  f = effectSize_f,
+  sig.level = alpha,
+  k = numLevelsFactorA)
+
+pwr::pwr.anova.test(
+  n = N,
+  f = effectSize_f,
+  sig.level = alpha,
+  k = numLevelsFactorB)
+
+pwr::pwr.anova.test(
+  n = N,
+  f = effectSize_f,
+  sig.level = alpha,
+  k = numLevelsFactorA + numLevelsFactorB)
+```
+
+Solving for sample size per group needed (given effect size, power, and alpha):
+
+```{r}
+pwr::pwr.anova.test(
+  power = power,
+  f = effectSize_f,
+  sig.level = alpha,
+  k = numLevelsFactorA)
+
+pwr::pwr.anova.test(
+  power = power,
+  f = effectSize_f,
+  sig.level = alpha,
+  k = numLevelsFactorB)
+
+pwr::pwr.anova.test(
+  power = power,
+  f = effectSize_f,
+  sig.level = alpha,
+  k = numLevelsFactorA + numLevelsFactorB)
+```
+
+Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+
+```{r}
+pwr::pwr.anova.test(
+  power = power,
+  n = N,
+  sig.level = alpha,
+  k = numLevelsFactorA)
+
+pwr::pwr.anova.test(
+  power = power,
+  n = N,
+  sig.level = alpha,
+  k = numLevelsFactorB)
+
+pwr::pwr.anova.test(
+  power = power,
+  n = N,
+  sig.level = alpha,
+  k = numLevelsFactorA + numLevelsFactorB)
 ```

### 2024-06-25 basic-statistics.qmd
20240624 - power analysis simulation

@@
 
+The power analysis code above assumes the groups are of equal size (i.e., a balanced design).
+If the design is unbalanced (i.e., there are different numbers of participants in each group), it may be necessary to conduct a power analysis via a simulation.
+Below is an example of evaluating the statistical power for detecting an effect unbalanced designs via simulation:
+
+```{r}
+nSim <- 1000 # number of simulations
+
+# Function to generate data and perform ANOVA
+simulate_anova <- function(nGroup1, nGroup2, f, alpha) {
+  # Means for each group
+  mean1 <- 0
+  mean2 <- f * sqrt((nGroup1 + nGroup2) / 2)
+
+  # Generate data
+  group1 <- rnorm(nGroup1, mean = mean1, sd = 1)
+  group2 <- rnorm(nGroup2, mean = mean2, sd = 1)
+
+  # Combine data
+  data <- data.frame(
+    value = c(group1, group2),
+    group = factor(rep(c("Group1", "Group2"), c(nGroup1, nGroup2)))
+  )
+
+  # Perform ANOVA
+  aov_result <- aov(value ~ group, data = data)
+  p_value <- summary(aov_result)[[1]][["Pr(>F)"]][1]
+
+  # Check if p-value is less than alpha
+  return(p_value < alpha)
+}
+
+# Run simulations
+set.seed(52242) # for reproducibility
+powerSimulationOneWayAnova <- replicate(
+  nSim,
+  simulate_anova(
+    nGroup1 = 10,
+    nGroup2 = 25,
+    f = effectSize_f,
+    alpha = alpha))
+
+# Estimate power
+mean(powerSimulationOneWayAnova)
+```
+
 #### Factorial ANOVA {#sec-statisticalPowerFactorialANOVA}
+
+The power analysis code below assumes the groups are of equal size (i.e., a balanced design).
+If the design is unbalanced (i.e., there are different numbers of participants in each group), it may be necessary to conduct a power analysis via a simulation.
+See @sec-statisticalPowerOneWayANOVA for an example power analysis simulation for one-way ANOVA.
 

### 2024-06-25 basic-statistics.qmd
20240624 - tweaks

@@
   d &= \frac{\text{mean difference}}{\text{pooled standard deviation}} \\
-   &= \frac{X_1 - X_2}{s} \\
+   &= \frac{\bar{X_1} - \bar{X_2}}{s} \\
 \end{aligned}
@@
 The coefficient of determination ($R^2$) is the proportion of variance in the [outcome variable](#sec-correlationalStudy) that is explained by the [predictor variables](#sec-correlationalStudy).
-\eta^2 is the proportion of variance in the [dependent variable](#sec-experiment) that is explained by the [independent variables](#sec-experiment).
-The coefficient of determination and \eta^2 can be expressed as the ratio of variance explained in the [outcome](#sec-correlationalStudy) or [dependent](#sec-experiment) variable to the total variance in the [outcome](#sec-correlationalStudy) or [dependent](#sec-experiment) variable, as in @eq-coefficientOfDeterminationPartitionedVariance:
+$\eta^2$ is the proportion of variance in the [dependent variable](#sec-experiment) that is explained by the [independent variables](#sec-experiment).
+The coefficient of determination and $\eta^2$ can be expressed as the ratio of variance explained in the [outcome](#sec-correlationalStudy) or [dependent](#sec-experiment) variable to the total variance in the [outcome](#sec-correlationalStudy) or [dependent](#sec-experiment) variable, as in @eq-coefficientOfDeterminationPartitionedVariance:
 

### 2024-06-25 basic-statistics.qmd
20240624 - figures

@@
 
-For a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed between the groups if there were no association between the variables.
+For a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed between the groups if there were no association between the variables, as depicted in @fig-pValues.
 For instance, when evaluating whether number of carries is <u>positively</u> associated with injury likelihood, and you observed a correlation coefficient of $r = .25$ between number of carries and injury likelihood, the *p*-value evaluates the likelihood that you would observe a correlation as strong or stronger than $r = .25$ between the variables if number of carries is not associated with injury likelihood.
+
+```{r}
+#| layout-ncol: 2
+#| fig-cap: "*p*-Values"
+#| fig-subcap: 
+#|   - "What is the probability my data would look like this..."
+#|   - "...if in the population the association was really this?"
+#| label: fig-pValues
+#| code-fold: true
+
+set.seed(52242)
+correlations <- data.frame(criterion = rnorm(100))
+correlations$sample <- complement(correlations$criterion, .95)
+correlations$population <- complement(correlations$criterion, 0)
+
+ggplot2::ggplot(
+  data = correlations,
+  mapping = aes(
+    x = sample,
+    y = criterion
+  )
+) +
+  geom_point() +
+  ggplot2::labs(
+    x = "Predictor Variable",
+    y = "Outcome Variable",
+    title = "What is the probability my data would look like this..."
+  ) +
+  ggplot2::theme_classic(
+    base_size = 16) +
+  ggplot2::theme(
+    legend.title = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+ggplot2::ggplot(
+  data = correlations,
+  mapping = aes(
+    x = population,
+    y = criterion
+  )
+) +
+  geom_point() +
+  ggplot2::labs(
+    x = "Predictor Variable",
+    y = "Outcome Variable",
+    title = "...if in the population, the association was really this:"
+  ) +
+  ggplot2::theme_classic(
+    base_size = 16) +
+  ggplot2::theme(
+    legend.title = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+```
 

### 2024-06-25 multiple-regression.qmd
20240605 - getting started/load packages
20240608 - getting started
20240624 - figures

--- 2024-06-05 base-rates.qmd
@@
-# Base Rates {#sec-baseRates}
+# Multiple Regression {#sec-multipleRegression}
 
-## Getting Started
+## Getting Started {#sec-multipleRegressionGettingStarted}
 
-### Load Packages
+### Load Packages {#sec-multipleRegressionLoadPackages}
 
@@
 ```
+
+## Overview of Multiple Regression {#sec-multipleRegressionOverview}
+
+Multiple regression examines the association between multiple predictor variables and one outcome variable.
+It allows obtaining a more accurate estimate of the unique contribution of a given predictor, by controlling for other variables ([covariates](#sec-covariates)).
+
+## Components {#sec-multipleRegressionComponents}
+
+- $B$ = unstandardized coefficient: direction and magnitude of the estimate (original scale)
+- $\beta$ = standardized coefficient: direction and magnitude of the estimate (standard deviation scale)
+- $SE$ = standard error: uncertainty of unstandardized estimate
+
+## Coefficient of Determination ($R^2$) {#sec-multipleRegressionRSquared}
+
+::: {#fig-regression}
+![](images/multipleRegressionRSquared.png)
+
+Conceptual Depiction of Proportion of Variance Explained ($R^2$) in an Outcome Variable ($Y$) by Multiple Predictors ($X1$ and $X2$) in Multiple Regression.
+The size of each circle represents the variable's variance.
+The proportion of variance in $Y$ that is explained by the predictors is depicted by the areas in orange.
+The dark orange space ($G$) is where multiple predictors explain overlapping variance in the outcome.
+Overlapping variance that is explained in the outcome ($G$) will not be recovered in the regression coefficients when both predictors are included in the regression model.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
+:::
+
+## Covariates {#sec-covariates}
+
+Covariates are variables that you include in the statistical model to try to control for them so you can better isolate the unique contribution of the predictor variable(s) in relation to the outcome variable.
+Use of covariates examines the association between the predictor variable and the outcome variable when holding people's level constant on the covariates.
+Inclusion of confounds as covariates allows potentially gaining a more accurate estimate of the causal effect of the predictor variable on the outcome variable.
+Ideally, you want to include any and all confounds as covariates.
+As described in @sec-correlationCausation, confounds are third variables that influence both the predictor variable and the outcome variable and explain their association.
+Covariates are potentially (but not necessarily) confounds.
+For instance, you might include the player's age as a covariate in a model that examines whether a player's 40-yard dash time at the NFL Combine predicts their fantasy points in their rookie year, but it may not be a confound.
+
+## Multicollinearity {#sec-multipleRegressionMulticollinearity}
+
+::: {#fig-regression}
+![](images/multipleRegressionMulticollinearity.png)
+
+Conceptual Depiction of Multicollinearity in Multiple Regression. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
+:::

### 2024-06-25 research-methods.qmd
20240624 - figures

@@
 In this case, intelligence or financial resources may be a confound that influences both sports drink consumption and player performance, but sports drink consumptions—though correlated with player performance—does not influence player performance.
+
+For another example, consider that ice cream sales are associated with shark attacks.
+It is unlikely that more people eating ice creams leads to shark attacks.
+There is a likely a third variable—heat waves—that is a confound because it influences both ice cream sales and shark attacks and explains their association.
 

### 2024-06-25 basic-statistics.qmd
20240624 - p-value misconceptions
20240624 - tweak figure

--- 2024-06-25 basic-statistics.qmd
@@
 
-For a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed between the groups if there were no association between the variables.
+For a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed between the groups if there were no association between the variables, as depicted in @fig-pValues.
 For instance, when evaluating whether number of carries is <u>positively</u> associated with injury likelihood, and you observed a correlation coefficient of $r = .25$ between number of carries and injury likelihood, the *p*-value evaluates the likelihood that you would observe a correlation as strong or stronger than $r = .25$ between the variables if number of carries is not associated with injury likelihood.
+
+```{r}
+#| layout-ncol: 2
+#| fig-cap: "*p*-Values"
+#| fig-subcap: 
+#|   - "What is the probability my data would look like this..."
+#|   - "...if in the population the association was really this?"
+#| label: fig-pValues
+#| code-fold: true
+
+set.seed(52242)
+correlations <- data.frame(criterion = rnorm(2000))
+correlations$sample <- NA
+correlations$sample[1:100] <- complement(correlations$criterion[1:100], .9)
+correlations$population <- complement(correlations$criterion, 0)
+
+ggplot2::ggplot(
+  data = correlations,
+  mapping = aes(
+    x = sample,
+    y = criterion
+  )
+) +
+  geom_point() +
+  scale_x_continuous(
+    limits = c(-3.5,3)
+  ) +
+  ggplot2::labs(
+    x = "Predictor Variable",
+    y = "Outcome Variable",
+    title = "What is the probability my data would look like this..."
+  ) +
+  ggplot2::theme_classic(
+    base_size = 16) +
+  ggplot2::theme(
+    legend.title = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+ggplot2::ggplot(
+  data = correlations,
+  mapping = aes(
+    x = population,
+    y = criterion
+  )
+) +
+  geom_point() +
+  scale_x_continuous(
+    limits = c(-2.5,2.5)
+  ) +
+  ggplot2::labs(
+    x = "Predictor Variable",
+    y = "Outcome Variable",
+    title = "...if in the population, the association was really this:"
+  ) +
+  ggplot2::theme_classic(
+    base_size = 16) +
+  ggplot2::theme(
+    legend.title = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+```
 
@@
 
+That is, the *p*-value is <u>not</u>:
+
+- the probability that the effect was due to chance
+- the probability that the null hypothesis is true
+- the size of the effect
+- the importance of the effect
+- whether the effect is true, real, or causal
+
 Statistical significance involves the *consistency* of an effect/association/difference; it suggests that the association/difference is reliably non-zero.

### 2024-06-25 correlation.qmd
20240618 - NHST
20240625 - correlation

@@
 
+## Overview of Correlation {#sec-correlationOverview}
+
+Correlation is a metric of the association between variables.
+Covariance is the association between variables and in an unstandardized metric that differs for variables with different scales.
+By contrast, correlation is in a standarized metric that does not differ for variables with different scales.
+When examining the association between variables that are [interval](#sec-interval) or [ratio](#sec-ratio) levels of measurement, Pearson correlation is used.
+When examining the association between variables that are [ordinal](#sec-ordinal) in level of measurement, Spearman correlation is used.
+
 ## The Correlation Coefficient ($r$)
+
+The formula for the correlation coefficient\index{correlation} is in @eq-pearsonCorrelation.
 
@@
 ```
+
+Interactive visualizations by Kristoffer Magnusson on *p*-values and null-hypothesis significance testing are below:
+
+- <https://rpsychologist.com/correlation/> (archived at <https://perma.cc/G8YR-VCM4>)
+
+## Examples {#sec-correlationExamples}
+
+### Covariance {#sec-correlationExamplesCovariance}
+
+### Pearson Correlation {#sec-correlationExamplesPearson}
+
+### Spearman Correlation {#sec-correlationExamplesSpearman}
+
+### Nonlinear Correlation {#sec-correlationExamplesNonlinear}
+
+### Correlation Matrix {#sec-correlationExamplesMatrix}
+
+### Correlogram {#sec-correlationExamplesCorrelogram}

### 2024-06-25 correlation.qmd
20240625 - correlations

@@
 library("petersenlab")
+library("XICOR")
 ```
@@
 When examining the association between variables that are [ordinal](#sec-ordinal) in level of measurement, Spearman correlation is used.
+Pearson correlation is an index of the *linear* association between variables.
+If a nonlinear association is present, other indices like xi [$\xi$; @Chatterjee2021] are better suited to detecting the association.
 
@@
 #| fig-cap: "Correlation Coefficients."
+#| fig-height: 12
 #| code-fold: true
+#| message: false
 

### 2024-06-25 correlation.qmd
20240625 - suppress message
20240625 - distance correlation

@@
 Pearson correlation is an index of the *linear* association between variables.
-If a nonlinear association is present, other indices like xi [$\xi$; @Chatterjee2021] are better suited to detecting the association.
+If a nonlinear association is present, other indices like xi [$\xi$; @Chatterjee2021] and distance correlation coefficients are better suited to detect the association.
 
@@
 
-dev.off() #par(mfrow = c(1,1))
+invisible(dev.off()) #par(mfrow = c(1,1))
 ```

### 2024-06-25 basic-statistics.qmd
20240625 - power of mediation

@@
 
+- <https://schoemanna.shinyapps.io/mc_power_med/>
 - <https://www.causalevaluation.org/power-analysis.html> (web application: <https://powerupr.shinyapps.io/index/>)

### 2024-06-26 basic-statistics.qmd
20240626 - scores, scales, and regression

@@
 #      values_from = value)
+```
+
+## Scores and Scales {#sec-scoresAndScales}
+
+There are many different types of scores and scales.
+This book focuses on [raw scores](#sec-rawScores) and [*z*-scores](#sec-zScores).
+For information on other scores and scales, including percentile ranks, *T*-scores, standard scores, scaled scores, and stanine scores, see here: <https://isaactpetersen.github.io/Principles-Psychological-Assessment/scoresScales.html#scoreTransformation> [@PetersenPrinciplesPsychAssessment].
+
+### Raw Scores {#sec-rawScores}
+
+*Raw scores* are the original data on the original metric.
+Thus, raw scores are considered *unstandardized*.
+For example, raw scores that represent the players' age may range from 20 to 40.
+Raw scores depend on the construct and unit; thus raw scores may not be comparable across variables.
+
+### *z* Scores {#sec-zScores}
+
+*z* scores have a mean of zero and a standard deviation of one.
+*z* scores are frequently used to render scores across variables more comparable.
+Thus, *z* scores are considered a form of a *standardized* score.
+
+*z* scores are calculated using @eq-zScore:
+
+$$
+z = \frac{X - \bar{X}}{\sigma}
+$$ {#eq-zScore}
+
+where $X$ is the observed score, $\bar{X}$ is the mean observed score, and $\sigma$ is the standard deviation of the observed scores.
+
+You can easily convert a variable to a *z* score using the `scale()` function:
+
+```{r}
+#| eval: false
+
+scale(variable)
+```
+
+With a standard normal curve, 68% of scores fall within one standard deviation of the mean.
+95% of scores fall within two standard deviations of the mean.
+99.7% of scores fall within three standard deviations of the mean.
+
+The area under a normal curve within one standard deviation of the mean is calculated below using the `pnorm()` function, which calculates the cumulative density function for a normal curve.
+
+```{r}
+stdDeviations <- 1
+
+pnorm(stdDeviations) - pnorm(stdDeviations * -1)
+```
+
+The area under a normal curve within one standard deviation of the mean is depicted in @fig-zScoreDensity1SD.
+
+```{r}
+#| fig.cap: "Density of Standard Normal Distribution. The blue region represents the area within one standard deviation of the mean."
+#| fig.scap: "Density of Standard Normal Distribution: One Standard Deviation of the Mean."
+#| label: fig-zScoreDensity1SD
+#| code-fold: true
+
+x <- seq(-4, 4, length = 200)
+y <- dnorm(x, mean = 0, sd = 1)
+plot(x, y, type = "l",
+     xlab = "z Score",
+     ylab = "Normal Density")
+
+x <- seq(stdDeviations * -1, stdDeviations, length = 100)
+y <- dnorm(x, mean = 0, sd = 1)
+polygon(c(stdDeviations * -1, x, stdDeviations),
+        c(0, y, 0),
+        col = "blue")
+```
+
+The area under a normal curve within two standard deviations of the mean is calculated below:
+
+```{r}
+stdDeviations <- 2
+
+pnorm(stdDeviations) - pnorm(stdDeviations * -1)
+```
+
+The area under a normal curve within two standard deviations of the mean is depicted in @fig-zScoreDensity2SD.
+
+```{r}
+#| fig.cap: "Density of Standard Normal Distribution. The blue region represents the area within two standard deviations of the mean."
+#| fig.scap: "Density of Standard Normal Distribution: Two Standard Deviations of the Mean."
+#| label: fig-zScoreDensity2SD
+#| code-fold: true
+
+x <- seq(-4, 4, length = 200)
+y <- dnorm(x, mean = 0, sd = 1)
+plot(x, y, type = "l",
+     xlab = "z Score",
+     ylab = "Normal Density")
+
+x <- seq(stdDeviations * -1, stdDeviations, length = 100)
+y <- dnorm(x, mean = 0, sd = 1)
+polygon(c(stdDeviations * -1, x, stdDeviations),
+        c(0, y, 0),
+        col = "blue")
+```
+
+The area under a normal curve within three standard deviations of the mean is calculated below:
+
+```{r}
+stdDeviations <- 3
+
+pnorm(stdDeviations) - pnorm(stdDeviations * -1)
+```
+
+The area under a normal curve within three standard deviations of the mean is depicted in @fig-zScoreDensity3SD.
+
+```{r}
+#| fig.cap: "Density of Standard Normal Distribution. The blue region represents the area within three standard deviations of the mean."
+#| fig.scap: "Density of Standard Normal Distribution: Three Standard Deviations of the Mean."
+#| label: fig-zScoreDensity3SD
+#| code-fold: true
+
+x <- seq(-4, 4, length = 200)
+y <- dnorm(x, mean = 0, sd = 1)
+plot(x, y, type = "l",
+     xlab = "z Score",
+     ylab = "Normal Density")
+
+x <- seq(stdDeviations * -1, stdDeviations, length = 100)
+y <- dnorm(x, mean = 0, sd = 1)
+polygon(c(stdDeviations * -1, x, stdDeviations),
+        c(0, y, 0),
+        col = "blue")
+```
+
+If you want to determine the *z* score associated with a particular percentile in a normal distribution, you can use the `qnorm()` function.
+For instance, the *z* score associated with the 37th percentile is:
+
+```{r}
+qnorm(.37)
 ```
@@
 
-Regression with one predictor takes the form of @eq-regression:
-
-$$
-y = \beta_0 + \beta_1x_1 + \epsilon
-$$ {#eq-regression}
-
-where $y$ is the [outcome variable](#sec-correlationalStudy), $\beta_0$ is the intercept, $\beta_1$ is the slope, $x_1$ is the [predictor variable](#sec-correlationalStudy), and $\epsilon$ is the error term.
-
-```{r}
-#| include: false
-#| eval: false
-
-set.seed(52242)
-regression <- data.frame(outcome = rnorm(40, mean = 5, sd = 2))
-
-regression$predictor <- complement(regression$outcome, .5)
-regression$predictor <- regression$predictor + abs(min(regression$predictor))
-
-lm(
-  outcome ~ predictor,
-  data = regression)
-
-ggplot2::ggplot(
-  data = regression,
-  ggplot2::aes(
-    x = predictor,
-    y = outcome,
-  )
-) +
-  ggplot2::geom_point() +
-  ggplot2::geom_smooth(
-    method = "lm",
-    linewidth = 2,
-    se = FALSE,
-    fullrange = TRUE) +
-  ggplot2::scale_x_continuous(
-    lim = c(0,8),
-    breaks = seq(from = 0, to = 8, by = 2),
-    expand = c(0,0)
-  ) +
-  ggplot2::scale_y_continuous(
-    lim = c(0,8),
-    breaks = seq(from = 0, to = 8, by = 2),
-    expand = c(0,0)
-  ) +
-  ggplot2::labs(
-    x = "Predictor Variable",
-    y = "Outcome Variable",
-    title = "Regression Best-Fit Line"
-  ) +
-  ggplot2::theme_classic(
-    base_size = 16) +
-  ggplot2::theme(legend.title = element_blank())
-
-ggsave("./images/regression.pdf", width = 6, height = 6)
-```
-
+Regression with a single predictor takes the form in @eq-regression.
 A regression line is depicted in @fig-regression.
-
-::: {#fig-regression}
-![](images/regression.png)
-
-A Regression Best-Fit Line.
-:::
-
-Regression with multiple predictors—i.e., multiple regression—takes the form of @eq-multipleRegression:
-
-$$
-y = \beta_0 + \beta_1x_1 + \beta_2x_2 + ... + \beta_px_p + \epsilon
-$$ {#eq-multipleRegression}
-
-where $p$ is the number of [predictor variables](#sec-correlationalStudy).
+Multiple regression (i.e., regression with multiple predictors) takes the form in @eq-multipleRegression.
 

### 2024-06-26 multiple-regression.qmd
20240625 - multiple regression overview
20240626 - scores, scales, and regression

@@
 
+Regression with one predictor takes the form of @eq-regression:
+
+$$
+y = \beta_0 + \beta_1x_1 + \epsilon
+$$ {#eq-regression}
+
+where $y$ is the [outcome variable](#sec-correlationalStudy), $\beta_0$ is the intercept, $\beta_1$ is the slope, $x_1$ is the [predictor variable](#sec-correlationalStudy), and $\epsilon$ is the error term.
+
+A regression line is depicted in @fig-regression.
+
+```{r}
+#| include: false
+#| eval: false
+
+set.seed(52242)
+regression <- data.frame(outcome = rnorm(40, mean = 5, sd = 2))
+
+regression$predictor <- complement(regression$outcome, .5)
+regression$predictor <- regression$predictor + abs(min(regression$predictor))
+
+lm(
+  outcome ~ predictor,
+  data = regression)
+
+ggplot2::ggplot(
+  data = regression,
+  ggplot2::aes(
+    x = predictor,
+    y = outcome,
+  )
+) +
+  ggplot2::geom_point() +
+  ggplot2::geom_smooth(
+    method = "lm",
+    linewidth = 2,
+    se = FALSE,
+    fullrange = TRUE) +
+  ggplot2::scale_x_continuous(
+    lim = c(0,8),
+    breaks = seq(from = 0, to = 8, by = 2),
+    expand = c(0,0)
+  ) +
+  ggplot2::scale_y_continuous(
+    lim = c(0,8),
+    breaks = seq(from = 0, to = 8, by = 2),
+    expand = c(0,0)
+  ) +
+  ggplot2::labs(
+    x = "Predictor Variable",
+    y = "Outcome Variable",
+    title = "Regression Best-Fit Line"
+  ) +
+  ggplot2::theme_classic(
+    base_size = 16) +
+  ggplot2::theme(legend.title = element_blank())
+
+ggsave("./images/regression.pdf", width = 6, height = 6)
+```
+
+::: {#fig-regression}
+![](images/regression.png)
+
+A Regression Best-Fit Line.
+:::
+
+Regression with multiple predictors—i.e., multiple regression—takes the form of @eq-multipleRegression:
+
+$$
+y = \beta_0 + \beta_1x_1 + \beta_2x_2 + ... + \beta_px_p + \epsilon
+$$ {#eq-multipleRegression}
+
+where $p$ is the number of [predictor variables](#sec-correlationalStudy).
+
 ## Components {#sec-multipleRegressionComponents}
@@
 - $B$ = unstandardized coefficient: direction and magnitude of the estimate (original scale)
-- $\beta$ = standardized coefficient: direction and magnitude of the estimate (standard deviation scale)
+- $\beta$ (beta) = standardized coefficient: direction and magnitude of the estimate (standard deviation scale)
 - $SE$ = standard error: uncertainty of unstandardized estimate
+
+The unstandardized regression coefficient ($B$) is interpreted such that, for every unit change in the [predictor variable](#sec-correlationalStudy), there is a __ unit change in the outcome variable.
+For instance, when examining the association between age and fantasy points, if the unstandardized regression coefficient is 2.3, players score on average 2.3 more points for each additional year of age.
+(In reality, we might expect a nonlinear, inverted-U-shaped association between age and fantasy points such that players tend to reach their peak in the middle of their careers.)
+Unstandardized regression coefficients are tied to the metric of the raw data.
+Thus, a large unstandardized regression coefficient for two variables may mean completely different things.
+Holding the strength of the association constant, you tend to see larger unstandardized regression coefficients for variables with smaller units and smaller unstandardized regression coefficients for variables with larger units.
+
+Standardized regression coefficients can be obtained by standardizing the variables to [*z*-scores](#sec-zScores) so they all have a mean of zero and standard deviation of one.
+The standardized regression coefficient ($\beta$) is interpreted such that, for every standard deviation change in the [predictor variable](#sec-correlationalStudy), there is a __ standard deviation change in the outcome variable.
+For instance, when examining the association between age and fantasy points, if the standardized regression coefficient is 0.1, players score on average 0.1 standard deviation more points for each additional standard deviation of their year of age.
+Standardized regression coefficients—though not the case in all instances—tend to fall between [−1, 1].
+Thus, standardized regression coefficients tend to be more comparable across variables and models compared to unstandardized regression coefficients.
+In this way, standardized regression coefficients provide a meaningful index of [effect size](#sec-practicalSignificance).
 

### 2024-06-27 basic-statistics.qmd
20240607 - interpretation of p-values when examining group differences

@@
 
-For a test of group differences, the *p*-value evaluates the likelihood that you would observe a difference as large or larger than the one you observed between the groups if there were no systematic difference between the groups.
+For a test of group differences, the *p*-value evaluates the likelihood that you would observe a difference as large or larger than the one you observed between the groups if there were no systematic difference between the groups, as depicted in @fig-pValuesDifference.
 For instance, when evaluating whether Quarterbacks have <u>longer</u> careers than Running Backs, and you observed a mean difference of 0.03 years, the *p*-value evaluates the likelihood that you would observe a difference as larger or larger than 0.03 years between the groups if Quarterbacks do not differ from Running Backs in terms of the length of their career.
 
-For a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed between the groups if there were no association between the variables, as depicted in @fig-pValues.
-For instance, when evaluating whether number of carries is <u>positively</u> associated with injury likelihood, and you observed a correlation coefficient of $r = .25$ between number of carries and injury likelihood, the *p*-value evaluates the likelihood that you would observe a correlation as strong or stronger than $r = .25$ between the variables if number of carries is not associated with injury likelihood.
-
 ```{r}
 #| layout-ncol: 2
-#| fig-cap: "*p*-Values"
+#| fig-cap: "Interpretation of *p*-Values When Examining The Differences Between Groups."
 #| fig-subcap: 
 #|   - "What is the probability my data would look like this..."
-#|   - "...if in the population the association was really this?"
-#| label: fig-pValues
+#|   - "...if in the population, the groups were really this?"
+#| label: fig-pValuesDifference
+#| code-fold: true
+
+set.seed(52242)
+
+nObserved <- 1000
+nPopulation <- 1000000
+
+observedGroups <- data.frame(
+  score = c(rnorm(nObserved, mean = 47, sd = 3), rnorm(nObserved, mean = 52, sd = 3)),
+  group = as.factor(c(rep("Group 1", nObserved), rep("Group 2", nObserved)))
+)
+
+populationGroups <- data.frame(
+  score = c(rnorm(nPopulation, mean = 50, sd = 3.03), rnorm(nPopulation, mean = 50, sd = 3)),
+  group = as.factor(c(rep("Group 1", nPopulation), rep("Group 2", nPopulation)))
+)
+
+ggplot2::ggplot(
+  data = observedGroups,
+  mapping = aes(
+    x = score,
+    fill = group,
+    color = group
+  )
+) +
+  geom_density(alpha = 0.5) +
+  scale_color_manual(values = c("red", "blue")) +
+  scale_fill_manual(values = c("red","blue")) +
+  geom_vline(xintercept = mean(observedGroups$score[which(observedGroups$group == "Group 1")])) +
+  geom_vline(xintercept = mean(observedGroups$score[which(observedGroups$group == "Group 2")])) +
+  ggplot2::labs(
+    x = "Score",
+    y = "Frequency",
+    title = "What is the probability my data would look like this..."
+  ) +
+  ggplot2::theme_classic(
+    base_size = 16) +
+  ggplot2::theme(
+    legend.title = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+ggplot2::ggplot(
+  data = populationGroups,
+  mapping = aes(
+    x = score,
+    fill = group,
+    color = group
+  )
+) +
+  geom_density(alpha = 0.5) +
+  scale_color_manual(values = c("red", "blue")) +
+  scale_fill_manual(values = c("red","blue")) +
+  geom_vline(xintercept = mean(populationGroups$score[which(populationGroups$group == "Group 1")])) +
+  geom_vline(xintercept = mean(populationGroups$score[which(populationGroups$group == "Group 2")])) +
+  ggplot2::labs(
+    x = "Score",
+    y = "Frequency",
+    title = "...if in the population, the groups were really this:"
+  ) +
+  ggplot2::theme_classic(
+    base_size = 16) +
+  ggplot2::theme(
+    legend.title = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+```
+
+For a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed between the groups if there were no association between the variables, as depicted in @fig-pValuesAssociation.
+For instance, when evaluating whether number of carries is <u>positively</u> associated with injury likelihood, and you observed a correlation coefficient of $r = .25$ between number of carries and injury likelihood, the *p*-value evaluates the likelihood that you would observe a correlation as strong or stronger than $r = .25$ between the variables if number of carries is not associated with injury likelihood.
+
+```{r}
+#| layout-ncol: 2
+#| fig-cap: "Interpretation of *p*-Values When Examining The Association Between Variables."
+#| fig-subcap: 
+#|   - "What is the probability my data would look like this..."
+#|   - "...if in the population, the association was really this?"
+#| label: fig-pValuesAssociation
 #| code-fold: true

### 2024-06-27 basic-statistics.qmd
20240627 - duplicated section identifier
20240627 - pvalue interpretation

@@
 #| layout-ncol: 2
-#| fig-cap: "Interpretation of *p*-Values When Examining The Differences Between Groups."
+#| fig-cap: "Interpretation of *p*-Values When Examining The Differences Between Groups. The vertical black lines reflect the group means."
 #| fig-subcap: 
@@
 set.seed(52242)
+
+observedCorrelation <- 0.9
+
 correlations <- data.frame(criterion = rnorm(2000))
 correlations$sample <- NA
-correlations$sample[1:100] <- complement(correlations$criterion[1:100], .9)
+correlations$sample[1:100] <- complement(correlations$criterion[1:100], observedCorrelation)
 correlations$population <- complement(correlations$criterion, 0)
@@
   geom_point() +
+  geom_smooth(method = "lm") +
   scale_x_continuous(
@@
   ) +
+  annotate(
+    x = 0,
+    y = 4,
+    label = paste("italic(r) != ", 0, sep = ""),
+    parse = TRUE,
+    geom = "text",
+    size = 7) + 
   ggplot2::labs(
@@
   geom_point() +
+  geom_smooth(
+    method = "lm",
+    se = FALSE) +
   scale_x_continuous(
@@
   ) +
+  annotate(
+    x = 0,
+    y = 4,
+    label = paste("italic(r) == '", "0.00", "'", sep = ""),
+    parse = TRUE,
+    geom = "text",
+    size = 7) + 
   ggplot2::labs(
@@
 
-### Correlation {#sec-correlation}
+### Correlation {#sec-correlationBasic}
 

### 2024-06-27 basic-statistics.qmd
20240627 - legend location

@@
     axis.text.y = element_blank(),
-    axis.ticks.y = element_blank())
+    axis.ticks.y = element_blank(),
+    #plot.title.position = "plot"
+    legend.position = "inside",
+    legend.margin = margin(0, 0, 0, 0),
+    legend.justification.top = "left",
+    legend.justification.left = "top",
+    legend.justification.bottom = "right",
+    legend.justification.inside = c(1, 1),
+    legend.location = "plot")
 
@@
     axis.text.y = element_blank(),
-    axis.ticks.y = element_blank())
+    axis.ticks.y = element_blank(),
+    #plot.title.position = "plot",
+    legend.position = "inside",
+    legend.margin = margin(0, 0, 0, 0),
+    legend.justification.top = "left",
+    legend.justification.left = "top",
+    legend.justification.bottom = "right",
+    legend.justification.inside = c(1, 1),
+    legend.location = "plot")
 ```

### 2024-06-27 bugs.qmd
20240627 - sessionInfo() for reproducibility

@@
 
-- `players` database has a season column but it is only populated for some players, and it is only populated with 2020
-- `player_stats` database of weekly stats does not have a `game_id` column
+- `players` database has a season column but it is only populated for some players, and it is only populated with 2020 (<https://github.com/nflverse/nflreadr/issues/238>)
+- `player_stats` database of weekly stats does not have a `game_id` column (<https://github.com/nflverse/nflreadr/issues/238>)
+
+<https://github.com/nflverse/nflreadr/issues/237>

### 2024-06-27 getting-started.qmd
20240627 - sessionInfo() for reproducibility

@@
 ```
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-gettingStartedSessionInfo}
+
+At the end of each chapter in which `R` code is used, I provide the session information, which describes the system and operating system the code was run on and the versions of each package.
+That way, if you get different results from me, you can see which differ, to help with reproducibility.
+If you run the (all of) the exact same code as is provided in the text, in the exact same order, with the exact same setup (platform, operating system, package versions, etc.), you should get the exact same answer as is in the text.
+That is the idea of reproducibility—getting the exact same result with the exact same inputs.
+Reproducibility is crucial for studies to achieve greater confidence in their findings and to ensure better replicability of findings across studies.
+
+```{r}
+sessionInfo
+```
+
+:::

### 2024-06-27 basic-statistics.qmd
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240627 - cross-references

@@
 
-Knowing any three of the following, you can calculate the fourth: statistical power, effect size, sample size, and alpha level.
+Knowing any three of the following, you can calculate the fourth: statistical power, [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance).
 Below is `R` code for calculating power for each of various statistical tests (i.e., a *power analysis*).
@@
 
-Solving for statistical power achieved (given effect size, sample size, and alpha):
+Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance)):
 
@@
 
-Solving for sample size needed (given effect size, power, and alpha):
+Solving for sample size needed (given [effect size](#sec-practicalSignificance), power, and [alpha level](#sec-statisticalSignificance)):
 
@@
 
-Solving for minimum effect size needed to detect (given sample size, power, and alpha):
+Solving for the minimum detectable [effect size](#sec-practicalSignificance) (given sample size, power, and [alpha level](#sec-statisticalSignificance)):
 
@@
 
-Solving for statistical power achieved (given effect size, sample size per group, and alpha):
+Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):
 
@@
 
-Solving for sample size per group needed (given effect size, power, and alpha):
+Solving for sample size per group needed (given [effect size](#sec-practicalSignificance), power, and [alpha level](#sec-statisticalSignificance)):
 
@@
 
-Solving for minimum effect size needed to detect (given sample size per group, power, and alpha):
+Solving for the minimum detectable [effect size](#sec-practicalSignificance) (given sample size per group, power, and [alpha level](#sec-statisticalSignificance)):
 
@@
 
-Solving for sample size needed (given effect size, power, and alpha)—$v = N - \text{numberOfPredictors} - 1$; thus, $N = v + \text{numberOfPredictors} + 1$:
+Solving for sample size needed (given [effect size](#sec-practicalSignificance), power, and [alpha level](#sec-statisticalSignificance))—$v = N - \text{numberOfPredictors} - 1$; thus, $N = v + \text{numberOfPredictors} + 1$:
 
@@
 
-Power analysis for multilevel modeling approaches is more complicated than it is for other statistical analyses, such as correlation, multiple regression, *t*-tests, ANOVA, etc.
+Power analysis for multilevel modeling approaches is more complicated than it is for other statistical analyses, such as [correlation](#sec-correlation), [multiple regression](#multipleRegression), [*t*-tests](#sec-tTest), [ANOVA](@sec-anova), etc.
 
@@
 
-Power analysis for latent variable modeling approaches like SEM is more complicated than it is for other statistical analyses, such as correlation, multiple regression, *t*-tests, ANOVA, etc.
+Power analysis for latent variable modeling approaches like structural equation modeling (SEM) is more complicated than it is for other statistical analyses, such as [correlation](#sec-correlation), [multiple regression](#multipleRegression), [*t*-tests](#sec-tTest), [ANOVA](@sec-anova), etc.
 
@@
 
-There are free tools for calculating power for tests of mediation and moderation:
+There are free tools for calculating power for tests of [mediation](#mediation) and [moderation](#moderation):
 
@@
 - <https://webpower.psychstat.org/wiki/models/index>
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-basicStatsSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-06-29 correlation.qmd
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240629 - interpretation of correlation coefficients

@@
 
+See @fig-interepretingCorrelationCoefficients for the interpretation of the magnitude and direction (sign) of various correlation coefficients.
+
+```{r}
+#| label: fig-interepretingCorrelationCoefficients
+#| fig-cap: "Interpretation of the Magnitude and Direction (Sign) of Correlation Coefficients."
+#| code-fold: true
+#| message: false
+
+library("patchwork")
+
+set.seed(52242)
+correlations2 <- data.frame(criterion = rnorm(15))
+
+correlations2$v1 <- complement(correlations2$criterion, -1)
+correlations2$v2 <- complement(correlations2$criterion, -.9)
+correlations2$v3 <- complement(correlations2$criterion, -.8)
+correlations2$v4 <- complement(correlations2$criterion, -.7)
+correlations2$v5 <- complement(correlations2$criterion, -.6)
+correlations2$v6 <- complement(correlations2$criterion, -.5)
+correlations2$v7 <- complement(correlations2$criterion, -.4)
+correlations2$v8 <- complement(correlations2$criterion, -.3)
+correlations2$v9 <- complement(correlations2$criterion, -.2)
+correlations2$v10 <-complement(correlations2$criterion, -.1)
+correlations2$v11 <-complement(correlations2$criterion, 0)
+correlations2$v12 <-complement(correlations2$criterion, .1)
+correlations2$v13 <-complement(correlations2$criterion, .2)
+correlations2$v14 <-complement(correlations2$criterion, .3)
+correlations2$v15 <-complement(correlations2$criterion, .4)
+correlations2$v16 <-complement(correlations2$criterion, .5)
+correlations2$v17 <-complement(correlations2$criterion, .6)
+correlations2$v18 <-complement(correlations2$criterion, .7)
+correlations2$v19 <-complement(correlations2$criterion, .8)
+correlations2$v20 <-complement(correlations2$criterion, .9)
+correlations2$v21 <-complement(correlations2$criterion, 1)
+
+# -1.0
+p1 <- ggplot(
+  data = correlations2,
+  mapping = aes(
+    x = criterion,
+    y = v1
+  )
+) + 
+  geom_point() +
+  geom_smooth(
+    method = "lm",
+    se = FALSE) +
+  labs(
+    title = "Perfect Negative Association",
+    subtitle = expression(paste(italic("r"), " = ", "−1.0"))
+  ) +
+  theme_classic(
+    base_size = 16) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+# -0.9
+p2 <- ggplot(
+  data = correlations2,
+  mapping = aes(
+    x = criterion,
+    y = v2
+  )
+) + 
+  geom_point() +
+  geom_smooth(
+    method = "lm",
+    se = FALSE) +
+  labs(
+    title = "Strong Negative Association",
+    subtitle = expression(paste(italic("r"), " = ", "−0.9"))
+  ) +
+  theme_classic(
+    base_size = 16) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+# -0.5
+p3 <- ggplot(
+  data = correlations2,
+  mapping = aes(
+    x = criterion,
+    y = v6
+  )
+) + 
+  geom_point() +
+  geom_smooth(
+    method = "lm",
+    se = FALSE) +
+  labs(
+    title = "Moderate Negative Association",
+    subtitle = expression(paste(italic("r"), " = ", "−0.5"))
+  ) +
+  theme_classic(
+    base_size = 16) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+# -0.2
+p4 <- ggplot(
+  data = correlations2,
+  mapping = aes(
+    x = criterion,
+    y = v9
+  )
+) + 
+  geom_point() +
+  geom_smooth(
+    method = "lm",
+    se = FALSE) +
+  labs(
+    title = "Weak Negative Association",
+    subtitle = expression(paste(italic("r"), " = ", "−0.2"))
+  ) +
+  theme_classic(
+    base_size = 16) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+# 0.0
+p5 <- ggplot(
+  data = correlations2,
+  mapping = aes(
+    x = criterion,
+    y = v11
+  )
+) + 
+  geom_point() +
+  geom_smooth(
+    method = "lm",
+    se = FALSE) +
+  labs(
+    title = "No Association",
+    subtitle = expression(paste(italic("r"), " = ", "0.0"))
+  ) +
+  theme_classic(
+    base_size = 16) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+# 0.2
+p6 <- ggplot(
+  data = correlations2,
+  mapping = aes(
+    x = criterion,
+    y = v13
+  )
+) + 
+  geom_point() +
+  geom_smooth(
+    method = "lm",
+    se = FALSE) +
+  labs(
+    title = "Weak Positive Association",
+    subtitle = expression(paste(italic("r"), " = ", "0.2"))
+  ) +
+  theme_classic(
+    base_size = 16) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+# 0.5
+p7 <- ggplot(
+  data = correlations2,
+  mapping = aes(
+    x = criterion,
+    y = v16
+  )
+) + 
+  geom_point() +
+  geom_smooth(
+    method = "lm",
+    se = FALSE) +
+  labs(
+    title = "Moderate Positive Association",
+    subtitle = expression(paste(italic("r"), " = ", "0.5"))
+  ) +
+  theme_classic(
+    base_size = 16) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+# 0.9
+p8 <- ggplot(
+  data = correlations2,
+  mapping = aes(
+    x = criterion,
+    y = v20
+  )
+) + 
+  geom_point() +
+  geom_smooth(
+    method = "lm",
+    se = FALSE) +
+  labs(
+    title = "Strong Positive Association",
+    subtitle = expression(paste(italic("r"), " = ", "0.9"))
+  ) +
+  theme_classic(
+    base_size = 16) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+# 1.0
+p9 <- ggplot(
+  data = correlations2,
+  mapping = aes(
+    x = criterion,
+    y = v21
+  )
+) + 
+  geom_point() +
+  geom_smooth(
+    method = "lm",
+    se = FALSE) +
+  labs(
+    title = "Perfect Positive Association",
+    subtitle = expression(paste(italic("r"), " = ", "1.0"))
+  ) +
+  theme_classic(
+    base_size = 16) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+
+p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9
+```
+
 Interactive visualizations by Kristoffer Magnusson on *p*-values and null-hypothesis significance testing are below:
@@
 ### Correlogram {#sec-correlationExamplesCorrelogram}
+
+## Session Info {#sec-correlationSessionInfo}
+
+::: {.content-visible when-format="html"}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-06-29 correlation.qmd
20240629 - updates

@@
 library("XICOR")
+library("tidyverse")
 ```
@@
 The correlation coefficient ranges from −1.0 to +1.0.\index{correlation}
-The correlation coefficient ($r$) tells you two things: (1) the direction of the association (positive or negative) and (2) the magnitude of the association.\index{correlation}
+The correlation coefficient ($r$) tells you two things: (1) the direction (sign) of the association (positive or negative) and (2) the magnitude of the association.\index{correlation}
 If the correlation coefficient is positive, the association is positive.\index{correlation}

### 2024-06-29 correlation.qmd
20240629 - remove leading zeros

@@
     title = "Strong Negative Association",
-    subtitle = expression(paste(italic("r"), " = ", "−0.9"))
+    subtitle = expression(paste(italic("r"), " = ", "−.9"))
   ) +
@@
     title = "Moderate Negative Association",
-    subtitle = expression(paste(italic("r"), " = ", "−0.5"))
+    subtitle = expression(paste(italic("r"), " = ", "−.5"))
   ) +
@@
     title = "Weak Negative Association",
-    subtitle = expression(paste(italic("r"), " = ", "−0.2"))
+    subtitle = expression(paste(italic("r"), " = ", "−.2"))
   ) +
@@
     title = "No Association",
-    subtitle = expression(paste(italic("r"), " = ", "0.0"))
+    subtitle = expression(paste(italic("r"), " = ", ".0"))
   ) +
@@
     title = "Weak Positive Association",
-    subtitle = expression(paste(italic("r"), " = ", "0.2"))
+    subtitle = expression(paste(italic("r"), " = ", ".2"))
   ) +
@@
     title = "Moderate Positive Association",
-    subtitle = expression(paste(italic("r"), " = ", "0.5"))
+    subtitle = expression(paste(italic("r"), " = ", ".5"))
   ) +
@@
     title = "Strong Positive Association",
-    subtitle = expression(paste(italic("r"), " = ", "0.9"))
+    subtitle = expression(paste(italic("r"), " = ", ".9"))
   ) +

### 2024-06-29 correlation.qmd
20240629 - tweaks

@@
 #| fig-cap: "Interpretation of the Magnitude and Direction (Sign) of Correlation Coefficients."
+#| fig-height: 6
 #| code-fold: true
@@
   theme_classic(
-    base_size = 16) +
+    base_size = 12) +
   theme(

### 2024-06-29 correlation.qmd
20240629 - figure dimensions

@@
 #| fig-cap: "Interpretation of the Magnitude and Direction (Sign) of Correlation Coefficients."
-#| fig-height: 6
+#| fig-height: 9
+#| fig-width: 9
 #| code-fold: true
@@
 
-p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9
+p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9 +
+  plot_layout(
+    ncol = 3,
+    heights = 1,
+    widths = 1)
 ```

### 2024-06-30 basic-statistics.qmd
20240629 - coefficient of determination

@@
 However, just because something is statistically significant does not mean that it is important.
-In addition to statistical significance, it is also important to consider practical significance.
+For instance, consider that we discover that players who consume sports drink before a game tend to perform better than players who do not ($p < .05$).
+However, what if consumption of sports drinks is associated with an average improvement of 0.002 points per game.
+A small effect such as this might be detectable with a large sample size.
+This effect would be considered to be reliable/consistent because it is statistically significant.
+However, such an effect would not be practically important because it has an effect size so small that it results in differences that are not practically important.
+Thus, in addition to statistical significance, it is also important to consider practical significance.
 
@@
       &= \frac{f^2}{1 + f^2} \\
-      &= \eta^2
+      &= \eta^2 \\
+      &= \frac{\text{variance explained in }Y}{\text{total variance in }Y}
 \end{aligned}

### 2024-06-30 multiple-regression.qmd
20240626 - figure sizing
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240629 - coefficient of determination

@@
 
-Multiple regression examines the association between multiple predictor variables and one outcome variable.
+Multiple regression examines the association between multiple [predictor variables](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy).
 It allows obtaining a more accurate estimate of the unique contribution of a given predictor, by controlling for other variables ([covariates](#sec-covariates)).
@@
 
-The unstandardized regression coefficient ($B$) is interpreted such that, for every unit change in the [predictor variable](#sec-correlationalStudy), there is a __ unit change in the outcome variable.
+The unstandardized regression coefficient ($B$) is interpreted such that, for every unit change in the [predictor variable](#sec-correlationalStudy), there is a __ unit change in the [outcome variable](#sec-correlationalStudy).
 For instance, when examining the association between age and fantasy points, if the unstandardized regression coefficient is 2.3, players score on average 2.3 more points for each additional year of age.
@@
 Standardized regression coefficients can be obtained by standardizing the variables to [*z*-scores](#sec-zScores) so they all have a mean of zero and standard deviation of one.
-The standardized regression coefficient ($\beta$) is interpreted such that, for every standard deviation change in the [predictor variable](#sec-correlationalStudy), there is a __ standard deviation change in the outcome variable.
+The standardized regression coefficient ($\beta$) is interpreted such that, for every standard deviation change in the [predictor variable](#sec-correlationalStudy), there is a __ standard deviation change in the [outcome variable](#sec-correlationalStudy).
 For instance, when examining the association between age and fantasy points, if the standardized regression coefficient is 0.1, players score on average 0.1 standard deviation more points for each additional standard deviation of their year of age.
@@
 
+The coefficient of determination ($R^2$) reflects the proportion of variance in the [outcome (dependent) variable](#sec-correlationalStudy) that is explained by the model predictions: $R^2 = \frac{\text{variance explained in }Y}{\text{total variance in }Y}$.
+Various formulas for $R^2$ are in @eq-rSquared.
+Larger $R^2$ values indicate greater accuracy.
+Multiple regression can be conceptualized with overlapping circles (similar to a venn diagram), where the non-overlapping portions of the circles reflect nonshared variance and the overlapping portions of the circles reflect shared variance, as in @fig-regression.
+
 ::: {#fig-regression}
-![](images/multipleRegressionRSquared.png)
+![](images/multipleRegressionRSquared.png){width=50%}
 
@@
 
+One issue with $R^2$ is that it increases as the number of predictors increases, which can lead to [overfitting](#sec-overfitting) if using $R^2$ as an index to compare models for purposes of selecting the "best-fitting" model.
+Consider the following example (adapted from @PetersenPrinciplesPsychAssessment) in which you have one [predictor variable](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy), as shown in @tbl-regression1.
+
+```{r}
+#| echo: false
+library("apa")
+
+regression1 <- data.frame(
+  "y" = c(7, 13, 29, 10),
+  "x1" = c(1, 2, 7, 2))
+
+regression2 <- data.frame(
+  "y" = c(7, 13, 29, 10),
+  "x1" = c(1, 2, 7, 2),
+  "x2" = c(3, 5, 1, 2))
+
+regression1_model <- lm(
+  y ~ x1,
+  data = regression1)
+
+regression1_intercept <- regression1_model$coefficients[[1]]
+regression1_slope <- regression1_model$coefficients[[2]]
+regression1_rsquare <- summary(regression1_model)$r.squared
+
+regression2_model <- lm(y ~ x1 + x2, data = regression2)
+regression2_intercept <- regression2_model$coefficients[[1]]
+regression2_slope1 <- regression2_model$coefficients[[2]]
+regression2_slope2 <- regression2_model$coefficients[[3]]
+regression2_rsquare <- summary(regression2_model)$r.squared
+```
+
+```{r}
+#| label: tbl-regression1
+#| tbl-cap: "Example Data of Predictor (x1) and Outcome (y) Used for Regression Model."
+#| echo: false
+
+kable(regression1,
+      col.names = c("y","x1"),
+      booktabs = TRUE)
+```
+
+Using the data, the best fitting regression model is: $y = `r apa(regression1_intercept, decimals = 2)` + `r apa(regression1_slope, decimals = 2)` \cdot x_1$.
+In this example, the $R^2$ is $`r apa(regression1_rsquare, decimals = 2)`$.
+The equation is not a perfect prediction, but with a single [predictor variable](#sec-correlationalStudy), it captures the majority of the variance in the outcome.
+
+Now consider the following example where you add a second [predictor variable](#sec-correlationalStudy) to the data above, as shown in @tbl-regression2.
+
+```{r}
+#| label: tbl-regression2
+#| tbl-cap: "Example Data of Predictors (x1 and x2) and Outcome (y) Used for Regression Model."
+#| echo: false
+
+kable(regression2,
+      col.names = c("y","x1","x2"),
+      booktabs = TRUE)
+```
+
+With the second [predictor variable](#sec-correlationalStudy), the best fitting regression model is: $y = `r apa(regression2_intercept, decimals = 2)` + `r apa(regression2_slope1, decimals = 2)` \cdot x_1 + `r apa(regression2_slope2, decimals = 2)` \cdot x_2$.
+In this example, the $R^2$ is $`r apa(regression2_rsquare, decimals = 2)`$.
+The equation with the second [predictor variable](#sec-correlationalStudy) provides a perfect prediction of the outcome.
+
+Providing perfect prediction with the right set of [predictor variable](#sec-correlationalStudy)s is the dream of multiple regression.
+So, using multiple regression, we often add [predictor variables](#sec-correlationalStudy) to incrementally improve prediction.
+Knowing how much variance would be accounted for by random chance follows @eq-predictionByChance:
+
+$$
+E(R^2) = \frac{K}{n-1}
+$$ {#eq-predictionByChance}
+
+where $E(R^2)$ is the expected value of $R^2$ (the proportion of variance explained), $K$ is the number of [predictor variables](#sec-correlationalStudy), and $n$ is the sample size.
+The formula demonstrates that the more [predictor variables](#sec-correlationalStudy) in the regression model, the more variance will be accounted for by chance.
+With many [predictor variables](#sec-correlationalStudy) and a small sample, you can account for a large share of the variance merely by chance.
+
+As an example, consider that we have 13 [predictor variables](#sec-correlationalStudy) to predict fantasy performance for 43 players.
+Assume that, with 13 [predictor variables](#sec-correlationalStudy), we explain 38% of the variance ($R^2 = .38; r = .62$).
+We explained a lot of the variance in the outcome, but it is important to consider how much variance could have been explained by random chance: $E(R^2) = \frac{K}{n-1} = \frac{13}{43 - 1} = .31$.\
+We expect to explain 31% of the variance, by chance, in the outcome.
+So, 82% of the variance explained was likely spurious (i.e., $\frac{.31}{.38} = .82$).
+As the sample size increases, the spuriousness decreases.
+
+To account for the number of [predictor variables](#sec-correlationalStudy) in the model, we can use a modified version of $R^2$ called adjusted $R^2$ ($R^2_{adj}$).
+Adjusted $R^2$ ($R^2_{adj}$) accounts for the number of [predictor variables](#sec-correlationalStudy) in the model, based on how much would be expected to be accounted for by chance to penalize [overfitting](#sec-overfitting).
+Adjusted $R^2$ ($R^2_{adj}$) reflects the proportion of variance in the [outcome (dependent) variable](#sec-correlationalStudy) that is explained by the model predictions over and above what would be expected to be accounted for by chance, given the number of [predictor variables](#sec-correlationalStudy) in the model.
+The formula for adjusted $R^2$ ($R^2_{adj}$) is in @eq-adjustedRSquared:
+
+$$
+R^2_{adj} = 1 - (1 - R^2) \frac{n - 1}{n - p - 1}
+$$ {#eq-adjustedRSquared}
+
+where $p$ is the number of [predictor variables](#sec-correlationalStudy) in the model, and $n$ is the sample size.
+
+## Overfitting {#sec-overfitting}
+
 ## Covariates {#sec-covariates}
 
-Covariates are variables that you include in the statistical model to try to control for them so you can better isolate the unique contribution of the predictor variable(s) in relation to the outcome variable.
-Use of covariates examines the association between the predictor variable and the outcome variable when holding people's level constant on the covariates.
-Inclusion of confounds as covariates allows potentially gaining a more accurate estimate of the causal effect of the predictor variable on the outcome variable.
+Covariates are variables that you include in the statistical model to try to control for them so you can better isolate the unique contribution of the [predictor variable](#sec-correlationalStudy)(s) in relation to the [outcome variable](#sec-correlationalStudy).
+Use of covariates examines the association between the predictor variable and the [outcome variable](#sec-correlationalStudy) when holding people's level constant on the covariates.
+Inclusion of confounds as covariates allows potentially gaining a more accurate estimate of the causal effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy).
 Ideally, you want to include any and all confounds as covariates.
-As described in @sec-correlationCausation, confounds are third variables that influence both the predictor variable and the outcome variable and explain their association.
+As described in @sec-correlationCausation, confounds are third variables that influence both the [predictor variable](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy) and explain their association.
 Covariates are potentially (but not necessarily) confounds.
@@
 ::: {#fig-regression}
-![](images/multipleRegressionMulticollinearity.png)
+![](images/multipleRegressionMulticollinearity.png){width=50%}
 
@@
 :::
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-multipleRegressionSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-06-30 actuarial.qmd
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240629 - best actuarial approaches to prediction

@@
 # Human Judgment Versus Actuarial Approaches to Prediction {#sec-judgmentVsActuarial}
+
+## Getting Started {#sec-judgmentVsActuarialGettingStarted}
+
+### Load Packages {#sec-judgmentVsActuarialLoadPackages}
+
+```{r}
+
+```
+
+::: {.content-visible when-format="html"}
+
+## Best Actuarial Approaches to Prediction
+
+The best actuarial models tend relatively simple (parsimonious), that can account for one or several of the most important predictors and their optimal weightings, and that account for the base rate of the phenomenon.
+Even unit-weighted formulas (formulas whose [predictor variables](#sec-correlationalStudy) are equally weighted with a weight of one) can sometimes generalize better to other samples than complex weightings [@Garb2019].
+Differential weightings sometimes capture random variance and [over-fit](#sec-overfitting) the model, thus leading to predictive accuracy shrinkage in cross-validation samples [@Garb2019], as described below.
+The choice of [predictor variables](#sec-correlationalStudy) often matters more than their weighting.
+
+In general, there is often shrinkage of estimates from training data set to a test data set.
+*Shrinkage* is when variables with stronger predictive power in the original data set tend to show somewhat smaller predictive power (smaller regression coefficients) when applied to new groups.
+Shrinkage reflects a model [over-fitting](#sec-overfitting) (i.e., fitting to error by capitalizing on chance).
+Shrinkage is especially likely when the original sample is small and/or unrepresentative and the number of variables considered for inclusion is large.
+Cross-validation with large, representative samples can help evaluate the amount of shrinkage of estimates, particularly for more complex models such as machine learning models [@Ursenbach2019].
+Ideally, cross-validation would be conducted with a separate sample (external cross-validation) to see the generalizability of estimates.
+However, you can also do internal cross-validation.
+For example, you can perform *k*-fold cross-validation, where you:
+
+- split the data set into *k* groups
+- for each unique group:
+  - take the group as a hold-out data set (also called a test data set)
+  - take the remaining groups as a training data set
+  - fit a model on the training data set and evaluate it on the test data set
+  - after all *k*-folds have been used as the test data set, and all models have been fit, you average the estimates across the models, which presumably yields more robust, generalizable estimates
+
+An emerging technique that holds promise for increasing predictive accuracy of actuarial methods is machine learning [@Garb2019].
+However, one challenge of some machine learning techniques is that they are like a "black box" and are not transparent, which raises ethical concerns [@Garb2019].
+Machine learning may be most valuable when the data available are complex and there are many [predictor variables](#sec-correlationalStudy) [@Garb2019].
+
+## Session Info {#sec-judgmentVsActuarialSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-06-30 multiple-regression.qmd
20240629 - fix error

@@
 
-regression1_intercept <- regression1_model$coefficients[[1]]
-regression1_slope <- regression1_model$coefficients[[2]]
-regression1_rsquare <- summary(regression1_model)$r.squared
+regression1_intercept <- apa(regression1_model$coefficients[[1]], decimals = 2)
+regression1_slope <- apa(regression1_model$coefficients[[2]], decimals = 2)
+regression1_rsquare <- apa(summary(regression1_model)$r.squared, decimals = 2)
 
 regression2_model <- lm(y ~ x1 + x2, data = regression2)
-regression2_intercept <- regression2_model$coefficients[[1]]
-regression2_slope1 <- regression2_model$coefficients[[2]]
-regression2_slope2 <- regression2_model$coefficients[[3]]
-regression2_rsquare <- summary(regression2_model)$r.squared
+regression2_intercept <- apa(regression2_model$coefficients[[1]], decimals = 2)
+regression2_slope1 <- apa(regression2_model$coefficients[[2]], decimals = 2)
+regression2_slope2 <- apa(regression2_model$coefficients[[3]], decimals = 2)
+regression2_rsquare <- apa(summary(regression2_model)$r.squared, decimals = 2)
 ```
@@
 
-Using the data, the best fitting regression model is: $y = `r apa(regression1_intercept, decimals = 2)` + `r apa(regression1_slope, decimals = 2)` \cdot x_1$.
-In this example, the $R^2$ is $`r apa(regression1_rsquare, decimals = 2)`$.
+Using the data, the best fitting regression model is: $y = `{r} regression1_intercept` + `{r} regression1_slope` \cdot x_1$.
+In this example, the $R^2$ is $`{r} regression1_rsquare`$.
 The equation is not a perfect prediction, but with a single [predictor variable](#sec-correlationalStudy), it captures the majority of the variance in the outcome.
@@
 
-With the second [predictor variable](#sec-correlationalStudy), the best fitting regression model is: $y = `r apa(regression2_intercept, decimals = 2)` + `r apa(regression2_slope1, decimals = 2)` \cdot x_1 + `r apa(regression2_slope2, decimals = 2)` \cdot x_2$.
-In this example, the $R^2$ is $`r apa(regression2_rsquare, decimals = 2)`$.
+With the second [predictor variable](#sec-correlationalStudy), the best fitting regression model is: $y = `{r} regression2_intercept` + `{r} regression2_slope1` \cdot x_1 + `{r} regression2_slope2` \cdot x_2$.
+In this example, the $R^2$ is $`{r} regression2_rsquare`$.
 The equation with the second [predictor variable](#sec-correlationalStudy) provides a perfect prediction of the outcome.
@@
 Assume that, with 13 [predictor variables](#sec-correlationalStudy), we explain 38% of the variance ($R^2 = .38; r = .62$).
-We explained a lot of the variance in the outcome, but it is important to consider how much variance could have been explained by random chance: $E(R^2) = \frac{K}{n-1} = \frac{13}{43 - 1} = .31$.\
+We explained a lot of the variance in the outcome, but it is important to consider how much variance could have been explained by random chance: $E(R^2) = \frac{K}{n-1} = \frac{13}{43 - 1} = .31$.
 We expect to explain 31% of the variance, by chance, in the outcome.
@@
 
+Statistical models applied to big data (e.g., data with many [predictor variables](#sec-correlationalStudy)) can *over-fit* the data, which means that the statistical model accounts for error variance, which will not generalize to future samples.
+So, even though an over-fitting statistical model appears to be accurate because it is accounting for more variance, it is not actually that accurate—it will predict new data less accurately than how accurately it accounts for the data with which the model was built.
+In the case of fantasy football analytics, this is especially relevant because there are hundreds if not thousands of variables we could consider for inclusion and many, many players when considering historical data.
+
+Consider an example where you develop an algorithm to predict players' fantasy performance based on 2023 data using hundreds of [predictor variables](#sec-correlationalStudy).
+To some extent, these [predictor variables](#sec-correlationalStudy) will likely account for true variance (i.e., signal) and error variance (i.e., noise).
+If we were to apply the same algorithm based on the 2023 prediction model to 2024 data, the prediction model would likely predict less accurately than with 2023 data.
+The regression coefficients in the
+
+In @fig-overfittingModel, the blue line represents the true distribution of the data, and the red line is an over-fitting model:
+
+```{r overfittingModel, echo = FALSE, out.width = "100%", fig.align = "center", fig.height = 8, fig.width = 8, fig.cap = "Over-fitting Model in Red Relative to the True Distribution of the Data in Blue."}
+#| label: fig-overfittingModel
+#| fig-cap: "Over-fitting Model in Red Relative to the True Distribution of the Data in Blue. From @PetersenPrinciplesPsychAssessment."
+#| echo: false
+#| fig.height: 8
+#| fig.width: 8
+
+sampleSize <- 200
+quadraticX <- rnorm(sampleSize)
+quadraticY <- quadraticX ^ 2 + rnorm(sampleSize)
+quadraticData <- cbind(quadraticX, quadraticY) %>%
+  data.frame %>%
+  arrange(quadraticX)
+
+quadraticModel <- lm(quadraticY ~ quadraticX + I(quadraticX ^ 2), data = quadraticData)
+quadraticNewData <- data.frame(quadraticX = seq(from = min(quadraticData$quadraticX), to = max(quadraticData$quadraticY), length.out = sampleSize))
+quadraticNewData$quadraticY <- predict(quadraticModel, newdata = quadraticNewData)
+
+loessFit <- loess(quadraticY ~ quadraticX, data = quadraticData, span = 0.01, degree = 1)
+loessNewData <- data.frame(quadraticX = seq(from = min(quadraticData$quadraticX), to = max(quadraticData$quadraticY), length.out = sampleSize))
+quadraticNewData$loessY <- predict(loessFit, newdata = quadraticNewData)
+
+plot(x = quadraticData$quadraticX, y = quadraticData$quadraticY, xlab = "", ylab = "")
+lines(quadraticNewData$quadraticY ~ quadraticNewData$quadraticX, lwd = 2, col = "blue")
+lines(quadraticNewData$loessY ~ quadraticNewData$quadraticX, lwd = 2, col = "red")
+```
+
 ## Covariates {#sec-covariates}
@@
 
+*Multicollinearity* occurs when two or more [predictor variables](#sec-correlationalStudy) in a regression model are highly correlated.
+The problem of having multiple [predictor variables](#sec-correlationalStudy) that are highly correlated is that it makes it challenging to estimate the regression coefficients accurately.
+
+Multicollinearity in multiple regression is depicted conceptually in @fig-regression.
+
 ::: {#fig-regression}
@@
 :::
+
+Consider the following example adapted from @PetersenPrinciplesPsychAssessment where you have two [predictor variables](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy), as shown in @tbl-regression3.
+
+```{r}
+#| echo: false
+
+regression3 <- data.frame(
+  "y" = c(9, 11, 17, 3, 21, 13),
+  "x1" = c(2, 3, 4, 1, 5, 3.5),
+  "x2" = c(4, 6, 8, 2, 10, 7))
+```
+
+```{r}
+#| label: tbl-regression3
+#| tbl-cap: "Example Data of Predictors (x1 and x2) and Outcome (y) Used for Regression Model."
+#| echo: false
+
+kable(regression3,
+      col.names = c("y","x1","x2"),
+      booktabs = TRUE)
+```
+
+The second [predictor variable](#sec-correlationalStudy) is not very good—it is exactly twice the value of the first [predictor variable](#sec-correlationalStudy); thus, the two [predictor variables](#sec-correlationalStudy) are perfectly correlated (i.e., $r = 1.0$).
+This means that there are different prediction equation possibilities that are equally good—see Equations in @eq-multicollinearity:
+
+$$
+\begin{aligned}
+  2x_2 &= y \\
+  0x_1 + 2x_2 &= y \\
+  4x_1 &= y \\
+  4x_1 + 0x_2 &= y \\
+  2x_1 + 1x_2 &= y \\
+  5x_1 - 0.5x_2 &= y \\
+  ...
+&= y
+\end{aligned}
+$$ {#eq-multicollinearity}
+
+Then, what are the regression coefficients?
+We do not know what are the correct regression coefficients because each of the possibilities fits the data equally well.
+Thus, when estimating the regression model, we could obtain arbitrary estimates of the regression coefficients with an enormous standard error around each estimate.
+In general, multicollinearity increases the uncertainty (i.e., standard errors and confidence intervals) around the parameter estimates.
+Any [predictor variables](#sec-correlationalStudy) that have a correlation above ~ $r = .30$ with each other could have an impact on the confidence interval of the regression coefficient.
+As the correlations among the [predictor variables](#sec-correlationalStudy) increase, the chance of getting an arbitrary answer increases, sometimes called "bouncing betas."
+So, it is important to examine a correlation matrix of the [predictor variables](#sec-correlationalStudy) before putting them in the same regression model.
+You can also examine indices such as variance inflation factor (VIF).
+
+To address multicollinearity, you can drop a redundant predictor or you can also use principal component analysis or factor analysis of the predictors to reduce the predictors down to a smaller number of meaningful predictors.
+For a meaningful answer in a regression framework that is precise and confident, you need a low level of intercorrelation among predictors, unless you have a very large sample size.
 

### 2024-06-30 multiple-regression.qmd
20240629 - fix error

@@
 #| echo: false
-library("apa")
+
+library("MOTE")
 
@@
 
-```{r overfittingModel, echo = FALSE, out.width = "100%", fig.align = "center", fig.height = 8, fig.width = 8, fig.cap = "Over-fitting Model in Red Relative to the True Distribution of the Data in Blue."}
+```{r}
 #| label: fig-overfittingModel

### 2024-06-30 multiple-regression.qmd
20240630 - missing packages
20240630 - attempt to fix warning

@@
 ```{r}
-
+library("petersenlab")
+library("tidyverse")
+library("knitr")
 ```
@@
 
-library("MOTE")
-
 regression1 <- data.frame(
@@
 
-Using the data, the best fitting regression model is: $y = `{r} regression1_intercept` + `{r} regression1_slope` \cdot x_1$.
-In this example, the $R^2$ is $`{r} regression1_rsquare`$.
+Using the data, the best fitting regression model is: $y = \text{`{r} regression1_intercept`} + \text{`{r} regression1_slope`} \cdot x_1$.
+In this example, the $R^2$ is $\text{`{r} regression1_rsquare`}$.
 The equation is not a perfect prediction, but with a single [predictor variable](#sec-correlationalStudy), it captures the majority of the variance in the outcome.
@@
 
-With the second [predictor variable](#sec-correlationalStudy), the best fitting regression model is: $y = `{r} regression2_intercept` + `{r} regression2_slope1` \cdot x_1 + `{r} regression2_slope2` \cdot x_2$.
-In this example, the $R^2$ is $`{r} regression2_rsquare`$.
+With the second [predictor variable](#sec-correlationalStudy), the best fitting regression model is: $y = \text{`{r} regression2_intercept`} + \text{`{r} regression2_slope1`} \cdot x_1 + \text{`{r} regression2_slope2`} \cdot x_2$.
+In this example, the $R^2$ is $\text{`{r} regression2_rsquare`}$.
 The equation with the second [predictor variable](#sec-correlationalStudy) provides a perfect prediction of the outcome.

### 2024-06-30 multiple-regression.qmd
20240630 - formatting

@@
 
-kable(regression1,
+kable(
+  regression1,
   col.names = c("y","x1"),
@@
 
-Using the data, the best fitting regression model is: $y = \text{`{r} regression1_intercept`} + \text{`{r} regression1_slope`} \cdot x_1$.
-In this example, the $R^2$ is $\text{`{r} regression1_rsquare`}$.
+Using the data, the best fitting regression model is: $y =$ `{r} regression1_intercept` $+$ `{r} regression1_slope` $\cdot x_1$.
+In this example, the $R^2$ is `{r} regression1_rsquare`.
 The equation is not a perfect prediction, but with a single [predictor variable](#sec-correlationalStudy), it captures the majority of the variance in the outcome.
@@
 
-kable(regression2,
+kable(
+  regression2,
   col.names = c("y","x1","x2"),
@@
 
-With the second [predictor variable](#sec-correlationalStudy), the best fitting regression model is: $y = \text{`{r} regression2_intercept`} + \text{`{r} regression2_slope1`} \cdot x_1 + \text{`{r} regression2_slope2`} \cdot x_2$.
-In this example, the $R^2$ is $\text{`{r} regression2_rsquare`}$.
+With the second [predictor variable](#sec-correlationalStudy), the best fitting regression model is: $y =$ `{r} regression2_intercept` + `{r} regression2_slope1` $\cdot x_1 +$ `{r} regression2_slope2` $\cdot x_2$.
+In this example, the $R^2$ is `{r} regression2_rsquare`.
 The equation with the second [predictor variable](#sec-correlationalStudy) provides a perfect prediction of the outcome.
@@
 
-kable(regression3,
+kable(
+  regression3,
   col.names = c("y","x1","x2"),

### 2024-06-30 causal-inference.qmd
Built site for gh-pages
20240630 - causal inference

--- 2024-06-05 base-rates.qmd
@@
-# Base Rates {#sec-baseRates}
+# Causal Inference {#sec-causalInference}
 
-## Getting Started
+## Getting Started {#sec-causalInferenceGettingStarted}
 
-### Load Packages
+### Load Packages {#sec-causalInferenceLoadPackages}
 
@@
 ```
+
+## Correlation Does Not Imply Causality {#sec-correlationCausality}
+
+As described in @sec-correlationCausation, there are several reasons why two variables, $X$ and $Y$, might be correlated:
+
+- $X$ causes $Y$
+- $Y$ causes $X$
+- a third variable (i.e., confound), $Z$, influences both $X$ and $Y$
+- the association between $X$ and $Y$ is spurious
+
+## Criteria for Causality {#sec-conditionsForCausality}
+
+How do we know whether two processes are causally related?
+There are three criteria for establishing causality [@Shadish2002]:
+
+1. The cause (e.g., the independent or predictor variable) temporally precedes the effect (i.e., the dependent or outcome variable).
+1. The cause is related to (i.e., associated with) the effect.
+1. There are no other alternative explanations for the effect apart from the cause.
+
+The first criterion for establishing causality involves temporal precedence.
+In order for a cause to influence an effect, the cause must occur before the effect.
+For instance, if sports drink consumption influences player performance, the sports drink consumption (that is presumed to influence performance) must occur prior to the performance improvement.
+Establishing the first criterion eliminates the possibility that the association between the purported cause and effect reflects reverse causation.
+Reverse causation occurs when the purported effect is actually the cause of the purported cause, rather than the other way around.
+For instance, if sports drink consumption occurs only once, and it occurs only before and not after performance, then we have ruled out the possibility of reverse causation (i.e., that better performance causes players to consume sports drink).
+
+The second criterion involves association.
+The purported cause must be associated with the purported effect.
+Nevertheless, as the maxim goes, "correlation does not imply causality."
+Just because two variables are correlated does not necessarily mean that they are causally related.
+However, correlation is useful because causality requires that the two processes be correlated.
+That is, correlation is a necessary but insufficient condition for causality.
+For instance, if sports drink consumption influences player performance, sports drink consumption must be associated with performance improvement.
+
+The third criterion involves ruling out alternative reasons why the purported cause and effect may be related.
+As noted in @sec-correlationCausality, there are four reasons why $X$ may be correlated with $Y$.
+If we meet the first criterion of causality, we have removed the possibility that $Y$ causes $X$ (i.e., reverse causality).
+To meet the third criterion of causality, we need to remove the possibility that the association reflects a third variable (confound) that influences both the cause and effect, and we need to remove the possibility that the association is spurious—the possibility that the association between the purported cause and effect is due to random chance.
+There are multiple approaches to meeting the third criterion of causality, such as by use of [experiments](#sec-causalInferenceExperiment), [longitudinal designs](#sec-causalInferenceLongitudinal), [control variables](#sec-causalInferenceControlVariables), [within-subject designs](#sec-causalInferenceWithinSubject), and [genetically informed designs](#sec-causalInferenceGeneticallyInformed), as described in @sec-approachesCausalInference.
+
+## Causal Diagrams {#sec-causalDiagrams}
+
+## Approaches for Causal Inference {#sec-approachesCausalInference}
+
+### Experimental Designs {#sec-causalInferenceExperiment}
+
+As described in @sec-experiment, [Experimental designs](#sec-experiment) are designs in which participants are randomly assigned to one or more levels of the [independent variable](#sec-experiment) to observe its effects on the [dependent variable](#sec-experiment).
+[Experimental designs](#sec-experiment) provide the strongest tests of causality because they can rule out reverse causation and third variables.
+For instance, by manipulating sports drink consumption before the player performs, they can eliminate the possibility that reverse causation explains the effect of the [independent variable](#sec-experiment) on the [dependent variable](#sec-experiment).
+Second, through randomly assigning players to consume or not consume sports drink, this holds everything else constant (so long as the groups are evenly distributed according to other factors, such as their age, weight, etc.) and thus removes the possibility that third variable confounds explain the effect of the [independent variable](#sec-experiment) on the [dependent variable](#sec-experiment).
+
+### Quasi-Experimental Designs {#sec-causalInferenceQuasiExperiment}
+
+Although [experimental designs](#sec-causalInferenceExperiment) provide the strongest tests of causality, manytimes they are impossible, unethical, or impractical to do.
+For instance, it would not be practical to randomly assign National Football League (NFL) players to either consume or not consume sports drink before their games.
+Players have their pregame rituals and routines and many would likely not agree to participate in such a study.
+Thus, we often rely on quasi-experimental designs such as natural experiments and [observational/correlational designs](#sec-correlationalStudy).
+
+#### Longitudinal Designs {#sec-causalInferenceLongitudinal}
+
+#### Control Variables {#sec-causalInferenceControlVariables}
+
+#### Within-Subject Analyses {#sec-causalInferenceWithinSubject}
+
+#### Genetically Informed Designs {#sec-causalInferenceGeneticallyInformed}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-causalInferenceSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-06-30 causal-inference.qmd
20240630 - causal diagrams

@@
 ```{r}
-
+library("dagitty")
 ```
@@
 
+A key tool when describing a research question or hypothesis is to create a conceptual depiction of the hypothesized causal processes.
+A causal diagram depicts the hypothesized causal processes that link two or more variables.
+A common form of causal diagrams is the directed acyclic graph (DAG).
+DAGs provide a helpful tool to communicate about causal questions and help identify how to avoid bias (i.e., over-estimation) in associations between variables due to confounding (i.e., common causes) [@Digitale2022].
+You can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: https://dagitty.net (archived at https://perma.cc/U9BY-VZE2) to create DAGs.
+
+```{r}
+XCausesY <- dagitty::dagitty("dag{
+  X -> Y
+}")
+
+plot(dagitty::graphLayout(XCausesY))
+
+YCausesX <- dagitty::dagitty("dag{
+  Y -> X
+}")
+
+plot(dagitty::graphLayout(YCausesX))
+
+ZCausesXandY <- dagitty::dagitty("dag{
+  Z -> Y
+  Z -> X
+  X <-> Y
+}")
+
+plot(dagitty::graphLayout(ZCausesXandY))
+```
+
 ## Approaches for Causal Inference {#sec-approachesCausalInference}

### 2024-06-30 causal-inference.qmd
20240630 - longitudinal designs, within-subject designs, control variables

@@
 You can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: https://dagitty.net (archived at https://perma.cc/U9BY-VZE2) to create DAGs.
+Examples of various causal diagrams that could explain why *X* is associated with *Y* are in @fig-XCausesY, @fig-YCausesX, and @fig-YCausesX.
 
 ```{r}
+#| label: fig-XCausesY
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting *X* Causing *Y*."
+
 XCausesY <- dagitty::dagitty("dag{
@@
 plot(dagitty::graphLayout(XCausesY))
+```
+
+```{r}
+#| label: fig-YCausesX
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting Reverse Causation: *Y* Causing *X*."
 
@@
 
+```{r}
+#| label: fig-YCausesX
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, *Z*, Causing *X* and *Y*, Thus Explaining Why *X* and *Y* are associated."
+
+ZCausesXandY <- dagitty::dagitty("dag{
+  Z -> Y
+  Z -> X
+  X <-> Y
+}")
+
+plot(dagitty::graphLayout(ZCausesXandY))
+```
+
 ## Approaches for Causal Inference {#sec-approachesCausalInference}
@@
 
-As described in @sec-experiment, [Experimental designs](#sec-experiment) are designs in which participants are randomly assigned to one or more levels of the [independent variable](#sec-experiment) to observe its effects on the [dependent variable](#sec-experiment).
+As described in @sec-experiment, [experimental designs](#sec-experiment) are designs in which participants are randomly assigned to one or more levels of the [independent variable](#sec-experiment) to observe its effects on the [dependent variable](#sec-experiment).
 [Experimental designs](#sec-experiment) provide the strongest tests of causality because they can rule out reverse causation and third variables.
@@
 
-Although [experimental designs](#sec-causalInferenceExperiment) provide the strongest tests of causality, manytimes they are impossible, unethical, or impractical to do.
-For instance, it would not be practical to randomly assign National Football League (NFL) players to either consume or not consume sports drink before their games.
+Although [experimental designs](#sec-causalInferenceExperiment) provide the strongest tests of causality, manytimes they are impossible, unethical, or impractical to conduct.
+For instance, it would likely not be practical to randomly assign National Football League (NFL) players to either consume or not consume sports drink before their games.
 Players have their pregame rituals and routines and many would likely not agree to participate in such a study.
@@
 
+Research designs can be compared in terms of their [internal validity](#sec-internalValidity)—the extent to which we can be confident about causal inferences.
+A cross-sectional association is depicted in @fig-crossSectional:
+
+::: {#fig-crossSectional}
+![](images/Longitudinal-01.png)
+
+Cross-Sectional Association. T1 = Timepoint 1. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
+:::
+
+For instance, we might observe that sports drink consumptions is concurrently associated with better player performance.
+Among [observational/correlational research designs](#sec-correlationalStudy), [cross-sectional designs](#sec-crossSectional) tend to have the weakest [internal validity](#sec-internalValidity).
+For the reasons described in @sec-correlationCausality, if we observe a cross-sectional association between $X$ (e.g., sports drink consumption) and $Y$ (e.g., player performance), we have little confidence that $X$ causes $Y$.
+As a result, [longitudinal designs](#sec-longitudinal) can be valuable for more closely approximating causality if an [experimental designs](#sec-causalInferenceExperiment) is not possible.
+Consider a lagged association that might be observed in a [longitudinal design](#sec-longitudinal), as in @fig-laggedAssociation, which is a slightly better approach than relying on cross-sectional associations:
+
+::: {#fig-laggedAssociation}
+![](images/Longitudinal-02.png)
+
+Lagged Association. T1 = Timepoint 1. T2 = Timepoint 2. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
+:::
+
+For instance, we might observe that sports drink performance *before* the game is associated with better player performance *during* the game.
+A lagged association has somewhat better [internal validity](#sec-internalValidity) than a cross-sectional association because we have greater evidence of temporal precedence—that the influence of the predictor *precedes* the outcome because the predictor was assessed before the outcome and it shows a predictive association.
+However, part of the association between the predictor with later levels of the outcome could be due to prior levels of the outcome that are stable across time.
+That is, it could be that better player performance leads players to consume more sports drink and that player performance is relatively stable across time.
+In such a case, it may be observed that sports drink consumption predicts later player performance even though player performance influences sports drink consumption, rather than the other way around
+Thus, consider an even stronger alternative—a lagged association that controls for prior levels of the outcome, as in @fig-laggedAssociationControlPriorLevels:
+
+::: {#fig-laggedAssociationControlPriorLevels}
+![](images/Longitudinal-03.png)
+
+Lagged Association, Controlling for Prior Levels of the Outcome. T1 = Timepoint 1. T2 = Timepoint 2. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
+:::
+
+For instance, we might observe that sports drink performance *before* the game is associated with better player performance *during* the game, while controlling for prior player performance.
+A lagged association controlling for prior levels of the outcome has better [internal validity](#sec-internalValidity) than a lagged association that does not control for prior levels of the outcome.
+A lagged association that controls for prior levels further reduces the likelihood that the association owes to the reverse direction of effect, because earlier levels of the outcome are controlled.
+However, consider an even stronger alternative—lagged associations that control for prior levels of the outcome and that simultaneously test each direction of effect, as depicted in @fig-crossLaggedPanelModel:
+
+::: {#fig-crossLaggedPanelModel}
+![](images/Longitudinal-04.png)
+
+Lagged Association, Controlling for Prior Levels of the Outcome, Simultaneously Testing Both Directions Of Effect.
+T1 = Timepoint 1.
+T2 = Timepoint 2.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
+:::
+
+Lagged associations that control for prior levels of the outcome and that simultaneously test each direction of effect provide the strongest [internal validity](#sec-internalValidity) among [observational/correlational designs](#sec-correlationalStudy).
+Such a design can help better clarify which among the variables is the chicken and the egg—which variable is more likely to be the cause and which is more likely to be the effect.
+If there are bidirectional effects, such a design can also help clarify the magnitude of each direction of effect.
+For instance, we can simultaneously evaluate the extent to which sports drink predicts later player performance (while controlling for prior performance) and the reverse—player performance predicting later sports drink consumption (while contorlling for prior sports drink consumption).
+
+#### Within-Subject Analyses {#sec-causalInferenceWithinSubject}
+
+Another design feature of [longitudinal designs](#sec-causalInferenceLongitudinal) that can lead to greater [internal validity](#sec-internalValidity) is the use of within-subject analyses.
+Between-subject analyses, might examine, for instanc, whether players who consume more sports drink perform better on average compared to players who consume less sports drink.
+However, there are other between-person differences that could explain any observed between-subject associations between sports drink consumption and players performance.
+Another approach could be to apply within-subject analyses.
+For instance, you could examining whether, within the same individual, if a player consumes a sports drink, do they perform better compared to games in which they did not consume a sports drink.
+When we control for prior levels of the outcome in the prediction, we are evaluating whether the predictor is associated with *change* in the outcome.
+Predicting change provides stronger evidence consistent with causality because it uses the individual as their own control and controls for many time-invariant confounds (i.e., confounds that do not change across time).
+
 #### Control Variables {#sec-causalInferenceControlVariables}
 
-#### Within-Subject Analyses {#sec-causalInferenceWithinSubject}
+One of the plausible alternatives to the inference that $X$ causes $Y$ is that there are third variable confounds that influence both $X$ and $Y$, thus explaining why $X$ and $Y$ are associated, as depicted in @fig-correlationAndCausation3 and @fig-YCausesX.
+Thus, another approach that can help increase [internal validity](#sec-internalValidity) is to include plausible confounds as control variables.
+For instance, if a third variable such as education level might be a confound that influences both sports drink consumption and player performance, you could include education level as a [covariate](#sec-covariates) in the model.
+Inclusion of a [covariate](#sec-covariates) attempts to control for the variable by examining the association between the [predictor variable](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy) while holding the [covariate](#sec-covariates) variables constant.
+For instance, such a model would examine whether, when accounting for education level, there is an association between sports drink consumption and player performance.
+
+Failure to control for important third variables can lead to erroneous conclusions, as evidenced by the association depicted in @fig-simpsonParadox.
+In the example, if we did not control for gender, we would infer that there is a positive association between dosage and recovery probability.
+However, when we examine each men and women separately, we learn that the association between dosage and recovery probability is actually negative within each gender group.
+Thus, in this case, failure to control for gender would lead to false inferences about the association between dosage and recovery probability.
+
+::: {#fig-simpsonParadox}
+![](images/simpsonParadox.png)
+
+Example Where Failing to Control for a Variable (In This Case, Gender) Would Lead to False Inferences.
+In this example, the association between dosage and recovery probability is positive at the population level, but the association is negative among men and women separately.
+(Figure reprinted from @Kievit2013, Figure 1, p.
+2.
+Kievit, R., Frankenhuis, W., Waldorp, L., & Borsboom, D.
+(2013).
+Simpson's paradox in psychological science: a practical guide.
+*Frontiers in Psychology*, *4*(513).
+[https://doi.org/10.3389/fpsyg.2013.00513](https://doi.org/10.3389/fpsyg.2013.00513))
+:::
 

### 2024-06-30 multiple-regression.qmd
20240630 - causal inference
20240630 - longitudinal designs, within-subject designs, control variables

@@
 #| label: fig-overfittingModel
-#| fig-cap: "Over-fitting Model in Red Relative to the True Distribution of the Data in Blue. From @PetersenPrinciplesPsychAssessment."
+#| fig-cap: "Over-fitting Model in Red Relative to the True Distribution of the Data in Blue. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
 #| echo: false
@@
 Covariates are variables that you include in the statistical model to try to control for them so you can better isolate the unique contribution of the [predictor variable](#sec-correlationalStudy)(s) in relation to the [outcome variable](#sec-correlationalStudy).
-Use of covariates examines the association between the predictor variable and the [outcome variable](#sec-correlationalStudy) when holding people's level constant on the covariates.
+Use of covariates examines the association between the [predictor variable](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy) when holding people's level constant on the covariates.
 Inclusion of confounds as covariates allows potentially gaining a more accurate estimate of the causal effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy).

### 2024-06-30 causal-inference.qmd
20240630 - duplicate figure label

@@
 You can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: https://dagitty.net (archived at https://perma.cc/U9BY-VZE2) to create DAGs.
-Examples of various causal diagrams that could explain why *X* is associated with *Y* are in @fig-XCausesY, @fig-YCausesX, and @fig-YCausesX.
+Examples of various causal diagrams that could explain why *X* is associated with *Y* are in @fig-XCausesY, @fig-YCausesX, and @fig-ZCausesXandY.
 
@@
 ```{r}
-#| label: fig-YCausesX
+#| label: fig-ZCausesXandY
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, *Z*, Causing *X* and *Y*, Thus Explaining Why *X* and *Y* are associated."
@@
 
-One of the plausible alternatives to the inference that $X$ causes $Y$ is that there are third variable confounds that influence both $X$ and $Y$, thus explaining why $X$ and $Y$ are associated, as depicted in @fig-correlationAndCausation3 and @fig-YCausesX.
+One of the plausible alternatives to the inference that $X$ causes $Y$ is that there are third variable confounds that influence both $X$ and $Y$, thus explaining why $X$ and $Y$ are associated, as depicted in @fig-correlationAndCausation3 and @fig-ZCausesXandY.
 Thus, another approach that can help increase [internal validity](#sec-internalValidity) is to include plausible confounds as control variables.

### 2024-06-30 causal-inference.qmd
20240630 - figure sizing

@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting *X* Causing *Y*."
+#| out-width: 20%
 
@@
 plot(dagitty::graphLayout(YCausesX))
-
-ZCausesXandY <- dagitty::dagitty("dag{
-  Z -> Y
-  Z -> X
-  X <-> Y
-}")
-
-plot(dagitty::graphLayout(ZCausesXandY))
 ```
@@
 ::: {#fig-simpsonParadox}
-![](images/simpsonParadox.png)
+![](images/simpsonParadox.png){width=200%}
 

### 2024-06-30 causal-inference.qmd
20240630 - figure sizing
20240630 - transition

--- 2024-06-30 causal-inference.qmd
@@
 ::: {#fig-crossSectional}
-![](images/Longitudinal-01.png)
+![](images/Longitudinal-01.png){width=20%}
 
@@
 For instance, you could examining whether, within the same individual, if a player consumes a sports drink, do they perform better compared to games in which they did not consume a sports drink.
-When we control for prior levels of the outcome in the prediction, we are evaluating whether the predictor is associated with *change* in the outcome.
-Predicting change provides stronger evidence consistent with causality because it uses the individual as their own control and controls for many time-invariant confounds (i.e., confounds that do not change across time).
+When we control for prior levels of the outcome in the prediction, we are evaluating whether the predictor is associated with witin-person *change* in the outcome.
+Predicting within-person change provides stronger evidence consistent with causality because it uses the individual as their own control and controls for many time-invariant confounds (i.e., confounds that do not change across time).
+However, predicting within-person change does not, by itself, control for time-varying confounds.
+So, it can also be useful to control for time-varying confounds, such as by use of [control variables](#sec-causalInferenceControlVariables).
 

### 2024-07-01 causal-inference.qmd
20240630 - genetically informed designs

@@
 
+Another approach to control for variables is to use genetically informed designs.
+Genetically informed designs allow controlling for potential genetic effects in order to more closely approximate the contributions of various environmental effects.
+Genetically informed designs exploit differing degrees of genetic relatedness among participants to capture the extent to which genetic factors may contribute to an outcome.
+The average percent of DNA shared between people of varying relationships is provided in @tbl-geneticRelatednessRelativePairs (<https://isogg.org/wiki/Autosomal_DNA_statistics>; archived at <https://perma.cc/MK3D-DST8>):
+
+| Relationship                       | Average Percent of Autosomal DNA Shared by Pairs of Relatives |
+|:-----------------------------------|:--------------------------------------------------------------|
+| Monozygotic ("identical") twins    | 100%                                                          |
+| Dizygotic ("fraternal") twins      | 50%                                                           |
+| Parent/child                       | 50%                                                           |
+| Full siblings                      | 50%                                                           |
+| Grandparent/grandchild             | 25%                                                           |
+| Aunt-or-uncle/niece-or-nephew      | 25%                                                           |
+| Half-siblings                      | 25%                                                           |
+| First cousin                       | 12.5%                                                         |
+| Great-grandparent/great-grandchild | 12.5%                                                         |
+
+: Average Percent of Autosomal DNA Shared by Pairs of Relatives by Relationship Type. {#tbl-geneticRelatednessRelativePairs}
+
+For instance, researchers may compare monozygotic twins versus dizygotic twins in some outcome—a so-called "twin study".
+It is assumed that the trait/outcome is attributable to genetic factors to the extent that the monozygotic twins (who share 100% of their DNA) are more similar in the trait or outcome compared to the dizygotic twins (who share on average 50% of their DNA).
+Alternatively, researchers could compare full siblings versus half-siblings, or they could compare full siblings versus first cousins.
+
+Genetically informed designs are not as relevant for fantasy football analytics, but they are useful to present as one of various design features that researchers can draw upon to strengthen their ability to make causal inferences.
+
 ::: {.content-visible when-format="html"}

### 2024-07-01 causal-inference.qmd
20240701 - ggdag

@@
 library("dagitty")
+library("ggdag")
 ```
@@
 plot(dagitty::graphLayout(XCausesY))
+
+dagitty::impliedConditionalIndependencies(XCausesY)
+```
+
+```{r}
+#| eval: false
+
+XCausesY_alt <- ggdag::dagify(
+  Y ~ X
+)
+
+plot(XCausesY_alt)
+ggdag(XCausesY_alt)
 ```
@@
 plot(dagitty::graphLayout(YCausesX))
+
+dagitty::impliedConditionalIndependencies(YCausesX)
+```
+
+```{r}
+#| eval: false
+
+YCausesX_alt <- ggdag::dagify(
+  X ~ Y
+)
+
+plot(YCausesX_alt)
+ggdag(YCausesX_alt)
 ```
@@
 plot(dagitty::graphLayout(ZCausesXandY))
+```
+
+```{r}
+#| eval: false
+
+ZCausesXandY_alt <- ggdag::dagify(
+  X ~ Z,
+  Y ~ Z,
+  X ~~ Y
+)
+
+plot(ZCausesXandY_alt)
+ggdag(ZCausesXandY_alt)
+```
+
+Consider another example in @fig-dag:
+
+```{r}
+#| label: fig-dag
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph)."
+
+mediationDag <- dagitty::dagitty("dag{
+  X -> M
+  X -> A
+  M -> Y
+  Y -> Z
+  A <-> M
+}")
+
+plot(dagitty::graphLayout(mediationDag))
+
+dagitty::impliedConditionalIndependencies(mediationDag)
+
+dagitty::adjustmentSets(
+  mediationDag, 
+  exposure = "X",
+  outcome = "Z",
+  effect = "total")
+```
+
+```{r}
+#| eval: false
+
+mediationDag_alt <- ggdag::dagify(
+  M ~ X,
+  Y ~ M,
+  Z ~ Y,
+  A ~ X,
+  A ~~ M
+)
+
+plot(mediationDag_alt)
+ggdag(mediationDag_alt)
 ```

### 2024-07-01 causal-inference.qmd
20240701 - mediation DAG
20240701 - dag modifications
20240701 - causal diagrams and formatting

@@
 
-As described in @sec-correlationCausation, there are several reasons why two variables, $X$ and $Y$, might be correlated:
+As described in @sec-correlationCausation, there are several reasons why two variables, `X` and `Y`, might be correlated:
 
-- $X$ causes $Y$
-- $Y$ causes $X$
-- a third variable (i.e., confound), $Z$, influences both $X$ and $Y$
-- the association between $X$ and $Y$ is spurious
+- `X` causes `Y`
+- `Y` causes `X`
+- a third variable (i.e., confound), `Z`, influences both `X` and `Y`
+- the association between `X` and `Y` is spurious
 
@@
 The third criterion involves ruling out alternative reasons why the purported cause and effect may be related.
-As noted in @sec-correlationCausality, there are four reasons why $X$ may be correlated with $Y$.
-If we meet the first criterion of causality, we have removed the possibility that $Y$ causes $X$ (i.e., reverse causality).
+As noted in @sec-correlationCausality, there are four reasons why `X` may be correlated with `Y`.
+If we meet the first criterion of causality, we have removed the possibility that `Y` causes `X` (i.e., reverse causality).
 To meet the third criterion of causality, we need to remove the possibility that the association reflects a third variable (confound) that influences both the cause and effect, and we need to remove the possibility that the association is spurious—the possibility that the association between the purported cause and effect is due to random chance.
@@
 You can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: https://dagitty.net (archived at https://perma.cc/U9BY-VZE2) to create DAGs.
-Examples of various causal diagrams that could explain why *X* is associated with *Y* are in @fig-XCausesY, @fig-YCausesX, and @fig-ZCausesXandY.
+Examples of various causal diagrams that could explain why `X` is associated with `Y` are in @fig-XCausesY, @fig-YCausesX, and @fig-ZCausesXandY.
 
@@
 #| label: fig-XCausesY
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting *X* Causing *Y*."
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting `X` Causing `Y`."
 
@@
 
+Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
+
 ```{r}
-#| eval: false
+#| label: fig-XCausesY2
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting `X` Causing `Y`."
 
@@
 
-plot(XCausesY_alt)
-ggdag(XCausesY_alt)
+#plot(XCausesY_alt) # this creates the same plot as above
+ggdag::ggdag(XCausesY_alt) + theme_dag_blank()
 ```
@@
 #| label: fig-YCausesX
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting Reverse Causation: *Y* Causing *X*."
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting Reverse Causation: `Y` Causing `X`."
 
@@
 
+Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
+
 ```{r}
-#| eval: false
+#| label: fig-YCausesX2
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting Reverse Causation: `Y` Causing `X`."
 
@@
 
-plot(YCausesX_alt)
-ggdag(YCausesX_alt)
+#plot(YCausesX_alt) # this creates the same plot as above
+ggdag::ggdag(YCausesX_alt)
 ```
@@
 #| label: fig-ZCausesXandY
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, *Z*, Causing *X* and *Y*, Thus Explaining Why *X* and *Y* are associated."
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, `Z`, Causing `X` and `Y`, Thus Explaining Why `X` and `Y` are associated."
 
@@
 
+Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
+
 ```{r}
-#| eval: false
+#| label: fig-ZCausesXandY2
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, `Z`, Causing `X` and `Y`, Thus Explaining Why `X` and `Y` are associated."
 
@@
 
-plot(ZCausesXandY_alt)
-ggdag(ZCausesXandY_alt)
+#plot(ZCausesXandY_alt) # this creates the same plot as above
+ggdag::ggdag(ZCausesXandY_alt)
 ```
@@
 mediationDag <- dagitty::dagitty("dag{
-  X -> M
-  X -> A
-  M -> Y
-  Y -> Z
-  A <-> M
+  X -> M1
+  X -> M2
+  M1 -> Y
+  M2 -> Y
+  M1 <-> M2
 }")
@@
   mediationDag, 
-  exposure = "X",
-  outcome = "Z",
+  exposure = "M1",
+  outcome = "Y",
   effect = "total")
@@
 
+In this example, `X` influences `Y` via `M1` and `M2` (i.e., multiple mediatirs), and `M1` is also associated with `M2`.
+The `dagitty::impliedConditionalIndependencies()` function identifies variables in the causal diagram that are conditionally independent (i.e., uncorrelated) after controlling for other variables in the model.
+For this causal diagram, `X` is conditionally independent with `Y` because `X` is independent with `Y` when controlling for `M1` and `M2`.
+
+The `dagitty::adjustmentSets()` function identifies variables that would be necessary to control for (i.e., to include as covariates) in order to identify an unbiased estimate of the association (whether the total effect, i.e., `effect = "total"`; or the direct effect, i.e., `effect = "direct"`) between two variables (`exposure` and `outcome`).
+In this case, to identify the unbiased association between `M1` and `Y`, it is important to control for `M2.`
+
+Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
+
 ```{r}
-#| eval: false
+#| label: fig-dag2
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph)."
 
 mediationDag_alt <- ggdag::dagify(
-  M ~ X,
-  Y ~ M,
-  Z ~ Y,
-  A ~ X,
-  A ~~ M
+  M1 ~ X,
+  M2 ~ X,
+  Y ~ M1,
+  Y ~ M2,
+  M1 ~~ M2
 )
 
-plot(mediationDag_alt)
-ggdag(mediationDag_alt)
+#plot(mediationDag_alt) # this creates the same plot as above
+ggdag::ggdag(mediationDag_alt)
 ```
@@
 Among [observational/correlational research designs](#sec-correlationalStudy), [cross-sectional designs](#sec-crossSectional) tend to have the weakest [internal validity](#sec-internalValidity).
-For the reasons described in @sec-correlationCausality, if we observe a cross-sectional association between $X$ (e.g., sports drink consumption) and $Y$ (e.g., player performance), we have little confidence that $X$ causes $Y$.
+For the reasons described in @sec-correlationCausality, if we observe a cross-sectional association between `X` (e.g., sports drink consumption) and `Y` (e.g., player performance), we have little confidence that `X` causes `Y`.
 As a result, [longitudinal designs](#sec-longitudinal) can be valuable for more closely approximating causality if an [experimental designs](#sec-causalInferenceExperiment) is not possible.
@@
 
-One of the plausible alternatives to the inference that $X$ causes $Y$ is that there are third variable confounds that influence both $X$ and $Y$, thus explaining why $X$ and $Y$ are associated, as depicted in @fig-correlationAndCausation3 and @fig-ZCausesXandY.
+One of the plausible alternatives to the inference that `X` causes `Y` is that there are third variable confounds that influence both `X` and `Y`, thus explaining why `X` and `Y` are associated, as depicted in @fig-correlationAndCausation3 and @fig-ZCausesXandY.
 Thus, another approach that can help increase [internal validity](#sec-internalValidity) is to include plausible confounds as control variables.

### 2024-07-01 getting-started.qmd
20240627 - sessionInfo()
20240629 - fix error
20240701 - causal diagrams and formatting

@@
 
-A full outer join includes all rows in $x$ **or** $y$.
-It returns columns from $x$ and $y$.
+A full outer join includes all rows in `x` **or** `y`.
+It returns columns from `x` and `y`.
 Here is how to merge two data frames using a full outer join (i.e., "full join"):
@@
 
-A left outer join includes all rows in $x$.
-It returns columns from $x$ and $y$.
+A left outer join includes all rows in `x`.
+It returns columns from `x` and `y`.
 Here is how to merge two data frames using a left outer join ("left join"):
@@
 
-A right outer join includes all rows in $y$.
-It returns columns from $x$ and $y$.
+A right outer join includes all rows in `y`.
+It returns columns from `x` and `y`.
 Here is how to merge two data frames using a right outer join ("right join"):
@@
 
-An inner join includes all rows that are in **both** $x$ **and** $y$.
-An inner join will return one row of $x$ for each matching row of $y$, and can duplicate values of records on either side (left or right) if $x$ and $y$ have more than one matching record.
-It returns columns from $x$ and $y$.
+An inner join includes all rows that are in **both** `x` **and** `y`.
+An inner join will return one row of `x` for each matching row of `y`, and can duplicate values of records on either side (left or right) if `x` and `y` have more than one matching record.
+It returns columns from `x` and `y`.
 Here is how to merge two data frames using an inner join:
@@
 A semi join is a filter.
-A left semi join returns all rows from $x$ **with** a match in $y$.
-That is, it filters out records from $x$ that are not in $y$.
-Unlike an [inner join](#innerJoin), a left semi join will never duplicate rows of $x$, and it includes columns from only $x$ (not from $y$). 
+A left semi join returns all rows from `x` **with** a match in `y`.
+That is, it filters out records from `x` that are not in `y`.
+Unlike an [inner join](#innerJoin), a left semi join will never duplicate rows of `x`, and it includes columns from only `x` (not from `y`). 
 Here is how to merge two data frames using a left semi join:
@@
 An anti join is a filter.
-A left anti join returns all rows from $x$ **without** a match in $y$.
-That is, it filters out records from $x$ that are in $y$.
-It returns columns from only $x$ (not from $y$).
+A left anti join returns all rows from `x` **without** a match in `y`.
+That is, it filters out records from `x` that are in `y`.
+It returns columns from only `x` (not from `y`).
 Here is how to merge two data frames using a left anti join:
@@
 
-A cross join combines each row in $x$ with each row in $y$.
+A cross join combines each row in `x` with each row in `y`.
 
@@
 ```{r}
-sessionInfo
+sessionInfo()
 ```

### 2024-07-01 research-methods.qmd
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240701 - causal diagrams and formatting

@@
 
-Just because $X$ is associated with $Y$ does not mean that $X$ causes $Y$.
-Consider that you find an association between variables $X$ and $Y$.
+Just because `X` is associated with `Y` does not mean that `X` causes `Y`.
+Consider that you find an association between variables `X` and `Y`.
 
-There are several reasons why you might observe an association between $X$ and $Y$:
+There are several reasons why you might observe an association between `X` and `Y`:
 
-- $X$ causes $Y$
-- $Y$ causes $X$
-- a third variable (i.e., confound), $Z$, influences both $X$ and $Y$
-- the association between $X$ and $Y$ is spurious
+- `X` causes `Y`
+- `Y` causes `X`
+- a third variable (i.e., confound), `Z`, influences both `X` and `Y`
+- the association between `X` and `Y` is spurious
 
-For instance, one possibility is that the association we observed reflects our hypothesis that $X$ causes $Y$, as depicted in @fig-correlationAndCausation1.
+For instance, one possibility is that the association we observed reflects our hypothesis that `X` causes `Y`, as depicted in @fig-correlationAndCausation1.
 That is, consumption of more sports drink may improve players' performance.
@@
 
-Hypothesized Causal Effect Based on an Observed Association Between $X$ and $Y$, Such That $X$ Causes $Y$.
+Hypothesized Causal Effect Based on an Observed Association Between `X` and `Y`, Such That `X` Causes `Y`.
 :::
 
-However, a second possibility is that the association reflects the opposite direction of effect, where $Y$ actually causes $X$, as depicted in @fig-correlationAndCausation2.
+However, a second possibility is that the association reflects the opposite direction of effect, where `Y` actually causes `X`, as depicted in @fig-correlationAndCausation2.
 For instance, greater performance may lead players to drink more sports drink (rather than the reverse).
@@
 
-Reverse (Opposite) Direction of Effect From the Hypothesized Effect, Where $Y$ Causes $X$.
+Reverse (Opposite) Direction of Effect From the Hypothesized Effect, Where `Y` Causes `X`.
 :::
@@
 If a third variable is a common cause of each and accounts for their association, it is a *confound*.
-An observed association between $X$ and $Y$ could reflect a confound—i.e., a cause ($Z$) that influences both $X$ and $Y$, which explains why $X$ and $Y$ are correlated even though they are not causally related.
-A third variable confound that is a common cause of both $X$ and $Y$ is depicted in @fig-correlationAndCausation3.
+An observed association between `X` and `Y` could reflect a confound—i.e., a cause (`Z`) that influences both `X` and `Y`, which explains why `X` and `Y` are correlated even though they are not causally related.
+A third variable confound that is a common cause of both `X` and `Y` is depicted in @fig-correlationAndCausation3.
 For instance, it may not be that sport drink consumption per se influences player performance; rather, it may be that players who are more intelligent or have more financial resources tend to drink more sports drinks and also tend to perform better.
@@
 
-Confounded Association Between $X$ and $Y$ due to a Common Cause, $Z$.
+Confounded Association Between `X` and `Y` due to a Common Cause, `Z`.
 :::
@@
 :::
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-researchMethodsSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-01 causal-inference.qmd
20240701 - loops and reorganization

@@
 There are multiple approaches to meeting the third criterion of causality, such as by use of [experiments](#sec-causalInferenceExperiment), [longitudinal designs](#sec-causalInferenceLongitudinal), [control variables](#sec-causalInferenceControlVariables), [within-subject designs](#sec-causalInferenceWithinSubject), and [genetically informed designs](#sec-causalInferenceGeneticallyInformed), as described in @sec-approachesCausalInference.
-
-## Causal Diagrams {#sec-causalDiagrams}
-
-A key tool when describing a research question or hypothesis is to create a conceptual depiction of the hypothesized causal processes.
-A causal diagram depicts the hypothesized causal processes that link two or more variables.
-A common form of causal diagrams is the directed acyclic graph (DAG).
-DAGs provide a helpful tool to communicate about causal questions and help identify how to avoid bias (i.e., over-estimation) in associations between variables due to confounding (i.e., common causes) [@Digitale2022].
-You can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: https://dagitty.net (archived at https://perma.cc/U9BY-VZE2) to create DAGs.
-Examples of various causal diagrams that could explain why `X` is associated with `Y` are in @fig-XCausesY, @fig-YCausesX, and @fig-ZCausesXandY.
-
-```{r}
-#| label: fig-XCausesY
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting `X` Causing `Y`."
-
-XCausesY <- dagitty::dagitty("dag{
-  X -> Y
-}")
-
-plot(dagitty::graphLayout(XCausesY))
-
-dagitty::impliedConditionalIndependencies(XCausesY)
-```
-
-Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
-
-```{r}
-#| label: fig-XCausesY2
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting `X` Causing `Y`."
-
-XCausesY_alt <- ggdag::dagify(
-  Y ~ X
-)
-
-#plot(XCausesY_alt) # this creates the same plot as above
-ggdag::ggdag(XCausesY_alt) + theme_dag_blank()
-```
-
-```{r}
-#| label: fig-YCausesX
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting Reverse Causation: `Y` Causing `X`."
-
-YCausesX <- dagitty::dagitty("dag{
-  Y -> X
-}")
-
-plot(dagitty::graphLayout(YCausesX))
-
-dagitty::impliedConditionalIndependencies(YCausesX)
-```
-
-Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
-
-```{r}
-#| label: fig-YCausesX2
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting Reverse Causation: `Y` Causing `X`."
-
-YCausesX_alt <- ggdag::dagify(
-  X ~ Y
-)
-
-#plot(YCausesX_alt) # this creates the same plot as above
-ggdag::ggdag(YCausesX_alt)
-```
-
-```{r}
-#| label: fig-ZCausesXandY
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, `Z`, Causing `X` and `Y`, Thus Explaining Why `X` and `Y` are associated."
-
-ZCausesXandY <- dagitty::dagitty("dag{
-  Z -> Y
-  Z -> X
-  X <-> Y
-}")
-
-plot(dagitty::graphLayout(ZCausesXandY))
-```
-
-Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
-
-```{r}
-#| label: fig-ZCausesXandY2
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, `Z`, Causing `X` and `Y`, Thus Explaining Why `X` and `Y` are associated."
-
-ZCausesXandY_alt <- ggdag::dagify(
-  X ~ Z,
-  Y ~ Z,
-  X ~~ Y
-)
-
-#plot(ZCausesXandY_alt) # this creates the same plot as above
-ggdag::ggdag(ZCausesXandY_alt)
-```
-
-Consider another example in @fig-dag:
-
-```{r}
-#| label: fig-dag
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph)."
-
-mediationDag <- dagitty::dagitty("dag{
-  X -> M1
-  X -> M2
-  M1 -> Y
-  M2 -> Y
-  M1 <-> M2
-}")
-
-plot(dagitty::graphLayout(mediationDag))
-
-dagitty::impliedConditionalIndependencies(mediationDag)
-
-dagitty::adjustmentSets(
-  mediationDag, 
-  exposure = "M1",
-  outcome = "Y",
-  effect = "total")
-```
-
-In this example, `X` influences `Y` via `M1` and `M2` (i.e., multiple mediatirs), and `M1` is also associated with `M2`.
-The `dagitty::impliedConditionalIndependencies()` function identifies variables in the causal diagram that are conditionally independent (i.e., uncorrelated) after controlling for other variables in the model.
-For this causal diagram, `X` is conditionally independent with `Y` because `X` is independent with `Y` when controlling for `M1` and `M2`.
-
-The `dagitty::adjustmentSets()` function identifies variables that would be necessary to control for (i.e., to include as covariates) in order to identify an unbiased estimate of the association (whether the total effect, i.e., `effect = "total"`; or the direct effect, i.e., `effect = "direct"`) between two variables (`exposure` and `outcome`).
-In this case, to identify the unbiased association between `M1` and `Y`, it is important to control for `M2.`
-
-Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
-
-```{r}
-#| label: fig-dag2
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph)."
-
-mediationDag_alt <- ggdag::dagify(
-  M1 ~ X,
-  M2 ~ X,
-  Y ~ M1,
-  Y ~ M2,
-  M1 ~~ M2
-)
-
-#plot(mediationDag_alt) # this creates the same plot as above
-ggdag::ggdag(mediationDag_alt)
-```
 
@@
 
+## Causal Diagrams {#sec-causalDiagrams}
+
+A key tool when describing a research question or hypothesis is to create a conceptual depiction of the hypothesized causal processes.
+A causal diagram depicts the hypothesized causal processes that link two or more variables.
+A common form of causal diagrams is the directed acyclic graph (DAG).
+DAGs provide a helpful tool to communicate about causal questions and help identify how to avoid bias (i.e., over-estimation) in associations between variables due to confounding (i.e., common causes) [@Digitale2022].
+For instance, from a DAG, it is possible to determine what variables it is important to control for in order to get unbiased estimates of the association between two variables of interest.
+To create DAGs, you can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: https://dagitty.net (archived at https://perma.cc/U9BY-VZE2).
+Examples of various causal diagrams that could explain why `X` is associated with `Y` are in @fig-XCausesY, @fig-YCausesX, and @fig-ZCausesXandY.
+
+```{r}
+#| label: fig-XCausesY
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting `X` Causing `Y`."
+
+XCausesY <- dagitty::dagitty("dag{
+  X -> Y
+}")
+
+plot(dagitty::graphLayout(XCausesY))
+
+dagitty::impliedConditionalIndependencies(XCausesY)
+```
+
+Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
+
+```{r}
+#| label: fig-XCausesY2
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting `X` Causing `Y`."
+
+XCausesY_alt <- ggdag::dagify(
+  Y ~ X
+)
+
+#plot(XCausesY_alt) # this creates the same plot as above
+ggdag::ggdag(XCausesY_alt) + theme_dag_blank()
+```
+
+```{r}
+#| label: fig-YCausesX
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting Reverse Causation: `Y` Causing `X`."
+
+YCausesX <- dagitty::dagitty("dag{
+  Y -> X
+}")
+
+plot(dagitty::graphLayout(YCausesX))
+
+dagitty::impliedConditionalIndependencies(YCausesX)
+```
+
+Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
+
+```{r}
+#| label: fig-YCausesX2
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting Reverse Causation: `Y` Causing `X`."
+
+YCausesX_alt <- ggdag::dagify(
+  X ~ Y
+)
+
+#plot(YCausesX_alt) # this creates the same plot as above
+ggdag::ggdag(YCausesX_alt)
+```
+
+```{r}
+#| label: fig-ZCausesXandY
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, `Z`, Causing `X` and `Y`, Thus Explaining Why `X` and `Y` are associated."
+
+ZCausesXandY <- dagitty::dagitty("dag{
+  Z -> Y
+  Z -> X
+  X <-> Y
+}")
+
+plot(dagitty::graphLayout(ZCausesXandY))
+```
+
+Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
+
+```{r}
+#| label: fig-ZCausesXandY2
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, `Z`, Causing `X` and `Y`, Thus Explaining Why `X` and `Y` are associated."
+
+ZCausesXandY_alt <- ggdag::dagify(
+  X ~ Z,
+  Y ~ Z,
+  X ~~ Y
+)
+
+#plot(ZCausesXandY_alt) # this creates the same plot as above
+ggdag::ggdag(ZCausesXandY_alt)
+```
+
+Consider another example in @fig-dag:
+
+```{r}
+#| label: fig-dag
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph)."
+
+mediationDag <- dagitty::dagitty("dag{
+  X -> M1
+  X -> M2
+  M1 -> Y
+  M2 -> Y
+  M1 <-> M2
+}")
+
+plot(dagitty::graphLayout(mediationDag))
+
+dagitty::impliedConditionalIndependencies(mediationDag)
+
+dagitty::adjustmentSets(
+  mediationDag, 
+  exposure = "M1",
+  outcome = "Y",
+  effect = "total")
+```
+
+In this example, `X` influences `Y` via `M1` and `M2` (i.e., multiple mediatirs), and `M1` is also associated with `M2`.
+The `dagitty::impliedConditionalIndependencies()` function identifies variables in the causal diagram that are conditionally independent (i.e., uncorrelated) after controlling for other variables in the model.
+For this causal diagram, `X` is conditionally independent with `Y` because `X` is independent with `Y` when controlling for `M1` and `M2`.
+
+The `dagitty::adjustmentSets()` function identifies variables that would be necessary to control for (i.e., to include as covariates) in order to identify an unbiased estimate of the association (whether the total effect, i.e., `effect = "total"`; or the direct effect, i.e., `effect = "direct"`) between two variables (`exposure` and `outcome`).
+In this case, to identify the unbiased association between `M1` and `Y`, it is important to control for `M2.`
+
+Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
+
+```{r}
+#| label: fig-dag2
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph)."
+
+mediationDag_alt <- ggdag::dagify(
+  M1 ~ X,
+  M2 ~ X,
+  Y ~ M1,
+  Y ~ M2,
+  M1 ~~ M2
+)
+
+#plot(mediationDag_alt) # this creates the same plot as above
+ggdag::ggdag(mediationDag_alt)
+```
+
 ::: {.content-visible when-format="html"}

### 2024-07-01 getting-started.qmd
20240701 - loops and reorganization

@@
 
+## Loops {#sec-loops}
+
+If you want to perform the same computation multiple times, it can be faster to do it in a loop compared to writing out the same computation many times.
+For instance, here is a loop that prints each element of a vector and the loop index (`i`) that indicates where the loop is in terms of its iterations:
+
+```{r}
+players
+
+for(i in 1:length(players$ID)){
+  print(paste("The loop is at index:", i, sep = " "))
+  print(paste("My favorite player is:", players$name[i], sep = " "))
+}
+```
+
 ## Calculations {#sec-calculations}

### 2024-07-01 causal-inference.qmd
20240701 - clean up

@@
 #plot(YCausesX_alt) # this creates the same plot as above
-ggdag::ggdag(YCausesX_alt)
+ggdag::ggdag(YCausesX_alt) + theme_dag_blank()
 ```
@@
 #plot(ZCausesXandY_alt) # this creates the same plot as above
-ggdag::ggdag(ZCausesXandY_alt)
+ggdag::ggdag(ZCausesXandY_alt) + theme_dag_blank()
 ```
@@
 plot(dagitty::graphLayout(mediationDag))
+```
 
+```{r}
 dagitty::impliedConditionalIndependencies(mediationDag)
@@
 
-In this example, `X` influences `Y` via `M1` and `M2` (i.e., multiple mediatirs), and `M1` is also associated with `M2`.
+In this example, `X` influences `Y` via `M1` and `M2` (i.e., multiple mediators), and `M1` is also associated with `M2`.
 The `dagitty::impliedConditionalIndependencies()` function identifies variables in the causal diagram that are conditionally independent (i.e., uncorrelated) after controlling for other variables in the model.
@@
 #plot(mediationDag_alt) # this creates the same plot as above
-ggdag::ggdag(mediationDag_alt)
+ggdag::ggdag(mediationDag_alt) + theme_dag_blank()
 ```

### 2024-07-01 causal-inference.qmd
20240701 - clarifications

@@
 
+Which variables are important to be included as control variables can be informed by the use of [causal diagrams](#sec-causalDiagrams), as described in @sec-causalDiagrams.
+
 #### Genetically Informed Designs {#sec-causalInferenceGeneticallyInformed}
@@
 For instance, from a DAG, it is possible to determine what variables it is important to control for in order to get unbiased estimates of the association between two variables of interest.
-To create DAGs, you can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: https://dagitty.net (archived at https://perma.cc/U9BY-VZE2).
+To create DAGs, you can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: <https://dagitty.net> (archived at <https://perma.cc/U9BY-VZE2>).
 Examples of various causal diagrams that could explain why `X` is associated with `Y` are in @fig-XCausesY, @fig-YCausesX, and @fig-ZCausesXandY.

### 2024-07-01 getting-started.qmd
20240701 - clarifications

--- 2024-07-01 getting-started.qmd
@@
 
+## Loops {#sec-loops}
+
+If you want to perform the same computation multiple times, it can be faster to do it in a loop compared to writing out the same computation many times.
+For instance, here is a loop that runs from 1 to `{r} length(players$ID)` (the number of players in the `players` object), incrementing by 1 after each iteration.
+The loop prints each element of a vector (i.e., the player's name) and the loop index (`i`) that indicates where the loop is in terms of its iterations:
+
+```{r}
+for(i in 1:length(players$ID)){
+  print(paste("The loop is at index:", i, sep = " "))
+  print(paste("My favorite player is:", players$name[i], sep = " "))
+}
+```
+
 ## Calculations {#sec-calculations}

### 2024-07-01 causal-inference.qmd
20240701 - bias

@@
 
+### Confounding {#sec-causalDiagramConfounding}
+
+Confounding occurs when two variables have a spurious or noncausal association because both variables are caused by another variable(s) [@DOnofrio2020].
+That is, two variables share a common cause, and the common cause leads the variables to be associated even though they are not causally related.
+The common cause—i.e., the variable that influences the two variables—is known as a confound (or confounder).
+
+### Mediation {#sec-causalDiagramMediation}
+
+### Collider Bias {#sec-colliderBias}
+
+Collider bias occurs when two variables influence a third variable, the collider [@DOnofrio2020].
+That is, a collider is a variable that is caused by two other variables.
+
+#### M-Bias {#sec-causalDiagramMBias}
+
+Collider bias may also occur when neither variable of interest is a direct cause of the collider [@Lederer2019].
+M-bias is a form of collider bias that occurs when two variables that are not causally related, `A` and `B`, both influence a collider, `C`, and each (`A` and `B`) also influences a separate variable—e.g., `A` influences `D` and `B` influences `E`.
+M-bias is so-named from the M-shape of the DAG.
+
+### Butterfly Bias {#sec-colliderBias}
+
+Butterfly bias occurs when both confounding and M-bias are present.
+Butterfly bias is so-named from the butterfly shape of the DAG.
+
+### Selection Bias {#sec-causalDiagramSelectionBias}
+
+Selection bias occurs when the selection of participants or their inclusion in analyses depends on the variables being studied.
+
 ::: {.content-visible when-format="html"}

### 2024-07-02 causal-inference.qmd
20240701 - types of bias

@@
 
+### Overview {#sec-causalDiagramsOverview}
+
 A key tool when describing a research question or hypothesis is to create a conceptual depiction of the hypothesized causal processes.
@@
 The common cause—i.e., the variable that influences the two variables—is known as a confound (or confounder).
+An example of confounding is depicted in @fig-counfounding:
+
+```{r}
+#| label: fig-counfounding
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Confounding."
+
+confounding <- ggdag::confounder_triangle(
+  x = "Training Intensity",
+  y = "Fantasy Points",
+  z = "Age")
+
+confounding %>%
+  ggdag(
+    text = FALSE,
+    use_labels = "label") + 
+  theme_dag_blank()
+```
+
+```{r}
+dagitty::impliedConditionalIndependencies(confounding)
+
+dagitty::adjustmentSets(
+  confounding, 
+  exposure = "x",
+  outcome = "y",
+  effect = "total")
+```
 
 ### Mediation {#sec-causalDiagramMediation}
+
+An example of full mediation is depicted in @fig-fullMediation:
+
+```{r}
+#| label: fig-fullMediation
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Full Mediation."
+
+full_mediation <- ggdag::mediation_triangle(
+  x = "Coaching Quality",
+  y = "Fantasy Points",
+  m = "Player Preparation")
+
+full_mediation %>%
+  ggdag(
+    text = FALSE,
+    use_labels = "label") + 
+  theme_dag_blank()
+```
+
+```{r}
+dagitty::impliedConditionalIndependencies(full_mediation)
+
+dagitty::adjustmentSets(
+  full_mediation, 
+  exposure = "x",
+  outcome = "y",
+  effect = "total")
+
+dagitty::adjustmentSets(
+  full_mediation, 
+  exposure = "x",
+  outcome = "y",
+  effect = "direct")
+```
+
+An example of partial mediation is depicted in @fig-partialMediation:
+
+```{r}
+#| label: fig-partialMediation
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Partial Mediation."
+
+partial_mediation <- ggdag::mediation_triangle(
+  x = "Coaching Quality",
+  y = "Fantasy Points",
+  m = "Player Preparation",
+  x_y_associated = TRUE)
+
+partial_mediation %>%
+  ggdag(
+    text = FALSE,
+    use_labels = "label") + 
+  theme_dag_blank()
+```
+
+```{r}
+dagitty::impliedConditionalIndependencies(partial_mediation)
+
+dagitty::adjustmentSets(
+  partial_mediation, 
+  exposure = "x",
+  outcome = "y",
+  effect = "total")
+
+dagitty::adjustmentSets(
+  partial_mediation, 
+  exposure = "x",
+  outcome = "y",
+  effect = "direct")
+```
 
@@
 That is, a collider is a variable that is caused by two other variables.
+An example of collider bias is depicted in @fig-colliderBias:
+
+```{r}
+#| label: fig-colliderBias
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Collider Bias."
+
+colliderBias <- ggdag::collider_triangle(
+  x = "Diet",
+  y = "Coaching Strategy",
+  m = "Injury Status")
+
+colliderBias %>% 
+  ggdag(
+    text = FALSE,
+    use_labels = "label") + 
+  theme_dag_blank()
+```
+
+```{r}
+dagitty::impliedConditionalIndependencies(colliderBias)
+
+dagitty::adjustmentSets(
+  colliderBias, 
+  exposure = "x",
+  outcome = "y",
+  effect = "total")
+```
 
@@
 M-bias is so-named from the M-shape of the DAG.
+An example of M-bias is depicted in @fig-mBias:
+
+```{r}
+#| label: fig-mBias
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of M-Bias."
+
+mBias <- ggdag::m_bias(
+  x = "Number of Media Articles About the Team",
+  y = "Fantasy Points",
+  a = "Team Record",
+  b = "Coaching Quality",
+  m = "Fan Attendance at Game")
+
+mBias %>% 
+  ggdag(
+    text = FALSE,
+    use_labels = "label") + 
+  theme_dag_blank()
+```
+
+```{r}
+dagitty::impliedConditionalIndependencies(mBias)
+
+dagitty::adjustmentSets(
+  mBias, 
+  exposure = "x",
+  outcome = "y",
+  effect = "total")
+```
 
@@
 Butterfly bias occurs when both confounding and M-bias are present.
-Butterfly bias is so-named from the butterfly shape of the DAG.
+Butterfly bias (aka bow-tie bias) is so-named from the butterfly shape of the DAG.
+An example of butterfly bias is depicted in @fig-butterflyBias:
+
+```{r}
+#| label: fig-butterflyBias
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Butterfly Bias."
+
+butterflyBias <- ggdag::butterfly_bias(
+  x = "Off-field Behavior",
+  y = "Fantasy Points",
+  a = "Diet",
+  b = "Coaching Quality",
+  m = "Mental Health")
+
+butterflyBias %>% 
+  ggdag(
+    text = FALSE,
+    use_labels = "label") + 
+  theme_dag_blank()
+```
+
+```{r}
+dagitty::impliedConditionalIndependencies(butterflyBias)
+
+dagitty::adjustmentSets(
+  butterflyBias, 
+  exposure = "x",
+  outcome = "y",
+  effect = "total")
+```
 

### 2024-07-02 causal-inference.qmd
20240701 - bias

@@
 
-Confounding occurs when two variables have a spurious or noncausal association because both variables are caused by another variable(s) [@DOnofrio2020].
+Confounding occurs when two variables—that are both caused by another variable(s)—have a spurious or noncausal association [@DOnofrio2020].
 That is, two variables share a common cause, and the common cause leads the variables to be associated even though they are not causally related.
@@
 confounding <- ggdag::confounder_triangle(
-  x = "Training Intensity",
-  y = "Fantasy Points",
-  z = "Age")
+  x = "Player Endurance",
+  y = "Field Goals Made",
+  z = "Stadium Altitude")
 
@@
 Collider bias may also occur when neither variable of interest is a direct cause of the collider [@Lederer2019].
-M-bias is a form of collider bias that occurs when two variables that are not causally related, `A` and `B`, both influence a collider, `C`, and each (`A` and `B`) also influences a separate variable—e.g., `A` influences `D` and `B` influences `E`.
+M-bias is a form of collider bias that occurs when two variables that are not causally related, `A` and `B`, both influence a collider, `M`, and each (`A` and `B`) also influences a separate variable—e.g., `A` influences `X` and `B` influences `Y`.
 M-bias is so-named from the M-shape of the DAG.
@@
 
-### Butterfly Bias {#sec-colliderBias}
+#### Butterfly Bias {#sec-colliderBias}
 
-Butterfly bias occurs when both confounding and M-bias are present.
+Butterfly bias occurs when both confounding and [M-bias](#sec-causalDiagramMBias) are present.
 Butterfly bias (aka bow-tie bias) is so-named from the butterfly shape of the DAG.
+In butterfly bias, the following criteria are met:
+
+- Two variables (`A` and `B`) influence a collider (`M`).
+- The collider influences two variables, `X` and `Y`.
+- `A` also influences `X`.
+- `B` also influences `Y`.
+- `A` and `B` are not causally related
+- `X` and `Y` are not causally related
+
+Or, more succinctly:
+
+- `A` influences `M` and `X`.
+- `B` influences `M` and `Y`.
+- `M` influences `X` and `Y`.
+
+In butterfly bias, the collider (`M`) is also a confound.
+That is, a variable is both influenced by two variables and influences two variables.
 An example of butterfly bias is depicted in @fig-butterflyBias:

### 2024-07-02 causal-inference.qmd
20240702 - confounding and mediation

@@
 dagitty::impliedConditionalIndependencies(confounding)
+```
 
+The output indicates that player endurance (`X`) and field goals made (`Y`) are conditionally independent when accounting for stadium altitude (`Z`).
+
+```{r}
 dagitty::adjustmentSets(
@@
 ```
+
+The output indicates that, to obtain an unbiased estimate of the causal association between two variables, it is necessary to control for any confounding [@Lederer2019].
+That is, to obtain an unbiased estimate of the causal association between player endurance (`X`) and field goals made (`Y`), it is necessary to control for stadium altitude (`Z`).
 
@@
 dagitty::impliedConditionalIndependencies(full_mediation)
+```
 
-dagitty::adjustmentSets(
-  full_mediation, 
-  exposure = "x",
-  outcome = "y",
-  effect = "total")
+In full mediation, `X` and `Y` are conditionally independent when accounting for the mediator (`Z`).
+In this case, coaching quality (`X`) and fantasy points (`Y`) are conditionally independent when accounting for player preparation (`M`).
+In other words, in this example, player preparation is the mechanism that fully (i.e., 100%) accounts for the effect of coaching quality on players' fantasy points.
 
+```{r}
 dagitty::adjustmentSets(
@@
 ```
+
+The output indicates that, to obtain an unbiased estimate of the *direct* causal association between coaching quality (`X`) and fantasy points (`Y`) (i.e., the effect that is *not* mediated through intermediate processes), it is necessary to control for player preparation (`M`).
+
+```{r}
+dagitty::adjustmentSets(
+  full_mediation, 
+  exposure = "x",
+  outcome = "y",
+  effect = "total")
+```
+
+However, to obtain an unbiased estimate of the *total* causal association between coaching quality (`X`) and fantasy points (`Y`) (i.e., including the portion of the effect that is mediated through intermediate processes), it is important *not* to control for player preparation (`M`).
+When the goal is to understand the (total) causal effect of coaching quality (`X`) and fantasy points (`Y`), controlling for the mediator (player preparation; `M`) would be inappropriate because doing so would remove the causal effect, thus artificially reducing the estimate of the association between coaching quality (`X`) and fantasy points (`Y`) [@Lederer2019].
 

### 2024-07-02 causal-inference.qmd
20240702 - additions

@@
 
-Which variables are important to be included as control variables can be informed by the use of [causal diagrams](#sec-causalDiagrams), as described in @sec-causalDiagrams.
+However, it can be problematic to control for variables indiscriminantly.
+The use of [causal diagrams](#sec-causalDiagrams) can inform which variables are important to be included as control variables, and—just as important—which variables not to include as control variables, as described in @sec-causalDiagrams.
 
@@
 
+Mediation can be divided into two types: [full](#sec-causalDiagramMediationFull) and [partial](#sec-causalDiagramMediationPartial).
+In [full mediation](#sec-causalDiagramMediationFull), the mediator(s) fully account for the effect of the predictor variable on the outcome variable.
+In [partial mediation](#sec-causalDiagramMediationPartial), the mediator(s) partially but do not fully account for the the effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy).
+
+#### Full Mediation {#sec-causalDiagramMediationFull}
+
 An example of full mediation is depicted in @fig-fullMediation:
@@
 
+#### Partial Mediation {#sec-causalDiagramMediationPartial}
+
 An example of partial mediation is depicted in @fig-partialMediation:
@@
 dagitty::impliedConditionalIndependencies(partial_mediation)
+```
 
+```{r}
 dagitty::adjustmentSets(
@@
   effect = "total")
+```
 
+```{r}
 dagitty::adjustmentSets(

### 2024-07-02 multiple-regression.qmd
20240702 - additions

@@
 Multiple regression examines the association between multiple [predictor variables](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy).
-It allows obtaining a more accurate estimate of the unique contribution of a given predictor, by controlling for other variables ([covariates](#sec-covariates)).
+It allows obtaining a more accurate estimate of the unique contribution of a given [predictor variable](#sec-correlationalStudy), by controlling for other variables ([covariates](#sec-covariates)).
 
-Regression with one predictor takes the form of @eq-regression:
+Regression with one [predictor variable](#sec-correlationalStudy) takes the form of @eq-regression:
 

### 2024-07-02 causal-inference.qmd
20240702 - mediation causal diagrams

@@
 In [full mediation](#sec-causalDiagramMediationFull), the mediator(s) fully account for the effect of the predictor variable on the outcome variable.
-In [partial mediation](#sec-causalDiagramMediationPartial), the mediator(s) partially but do not fully account for the the effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy).
+In [partial mediation](#sec-causalDiagramMediationPartial), the mediator(s) partially but do not fully account for the effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy).
 
@@
 
+In partial mediation, `X` and `Y` are *not* conditionally independent when accounting for the mediator (`Z`).
+In this case, coaching quality (`X`) and fantasy points (`Y`) are still associated when accounting for player preparation (`M`).
+In other words, in this example, player preparation is a mechanism that partially but does not fully account for the effect of coaching quality on players' fantasy points.
+That is, there are likely other mechanisms, in addition to player preparation, that collectively account for the effect of coaching quality on fantasy points.
+For instance, coaching quality could also influence player fantasy points through better play-calling.
+
+```{r}
+dagitty::adjustmentSets(
+  partial_mediation, 
+  exposure = "x",
+  outcome = "y",
+  effect = "direct")
+```
+
+As with [full mediation](#sec-causalDiagramMediationPartial), the output indicates that, to obtain an unbiased estimate of the *direct* causal association between coaching quality (`X`) and fantasy points (`Y`) (i.e., the effect that is *not* mediated through intermediate processes), it is necessary to control for player preparation (`M`).
+
 ```{r}
@@
 
-```{r}
-dagitty::adjustmentSets(
-  partial_mediation, 
-  exposure = "x",
-  outcome = "y",
-  effect = "direct")
-```
+However, as with [full mediation](#sec-causalDiagramMediationPartial), to obtain an unbiased estimate of the *total* causal association between coaching quality (`X`) and fantasy points (`Y`) (i.e., including the portion of the effect that is mediated through intermediate processes), it is important *not* to control for player preparation (`M`).
+When the goal is to understand the (total) causal effect of coaching quality (`X`) and fantasy points (`Y`), controlling for a mediator (player preparation; `M`) would be inappropriate because doing so would remove the causal effect, thus artificially reducing the estimate of the association between coaching quality (`X`) and fantasy points (`Y`) [@Lederer2019].
 

### 2024-07-03 causal-inference.qmd
20240702 - collider bias

@@
 The output indicates that player endurance (`X`) and field goals made (`Y`) are conditionally independent when accounting for stadium altitude (`Z`).
+*Conditional independence* refers to two variables being unassociated when controlling for other variables.
 
@@
 
-Collider bias occurs when two variables influence a third variable, the collider [@DOnofrio2020].
-That is, a collider is a variable that is caused by two other variables.
-An example of collider bias is depicted in @fig-colliderBias:
+*Collision* occurs when two variables influence a third variable, the collider [@DOnofrio2020].
+That is, a collider is a variable that is caused by two other variables (i.e., a common effect).
+An example collision is depicted in @fig-colliderBias1 and @fig-colliderBias2:
 
 ```{r}
-#| label: fig-colliderBias
-#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Collider Bias."
+#| label: fig-colliderBias1
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of a Collision with a Collider (Injury Status)."
 
-colliderBias <- ggdag::collider_triangle(
+colliderBias1 <- ggdag::collider_triangle(
   x = "Diet",
@@
 
-colliderBias %>% 
+colliderBias1 %>% 
   ggdag(
@@
 ```{r}
-dagitty::impliedConditionalIndependencies(colliderBias)
+dagitty::impliedConditionalIndependencies(colliderBias1)
+```
 
+In this example collision, diet (`X`) and coaching strategy (`Y`) are independent.
+
+```{r}
 dagitty::adjustmentSets(
-  colliderBias, 
+  colliderBias1, 
   exposure = "x",
@@
 ```
+
+As the output indicates, we should not control for the collider when examining the association between the two causes of the collider.
+That is, we should not control for injury status (`M`) when examining the association between diet (`X`) and coaching strategy.
+Controlling for the collider leads to confounding and can artificially induce an association between the two causes of the collider despite no causal association between them [@Lederer2019].
+Obtaining a distorted or artificial association between two variables due to inappropriately controlling for a collider is known as *collider bias*.
+
+Consider another example:
+
+```{r}
+#| label: fig-colliderBias2
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Collider Bias."
+
+colliderBias2 <- ggdag::collider_triangle(
+  x = "Coaching Quality",
+  y = "Player Preparation",
+  m = "Fantasy Points",
+  x_y_associated = TRUE)
+
+colliderBias2 %>% 
+  ggdag(
+    text = FALSE,
+    use_labels = "label") + 
+  theme_dag_blank()
+```
+
+```{r}
+dagitty::impliedConditionalIndependencies(colliderBias2)
+```
+
+In this example of collider bias, there are no conditional independencies.
+
+```{r}
+dagitty::adjustmentSets(
+  colliderBias2, 
+  exposure = "x",
+  outcome = "y",
+  effect = "total")
+```
+
+Again, it would be important not to control for the collider, fantasy points (`M`), when examining the association between coaching quality (`X`) and player preparation (`Y`).
+In this case, controlling for the collider would remove some of the causal effect of coaching quality on player preparation and could lead to an artificially smaller estimate of the causal effect between coaching quality and player preparation.
 

### 2024-07-03 causal-inference.qmd
20240702 - M-bias and butterfly bias

@@
 
+In this example, fan attendance is the collider that is influenced separately by the team record and the coaching quality.
+This is a fictitious example for purposes of demonstration; in reality, coaching quality influences the team's record.
+
 ```{r}
 dagitty::impliedConditionalIndependencies(mBias)
+```
 
+As the output indicates, there are several conditional independencies.
+
+```{r}
 dagitty::adjustmentSets(
@@
 ```
+
+It is important not to control for the collider (fan attendance).
+If you control for the collider, you can induce an artificial association between team record and coaching quality.
+Moreover, because doing so induces an artificial association between team record and coaching quality, it can also induce an artifial association between the effects of team record and coaching quality: number of media articles about the team and fantasy points, respectively.
+That is, controlling for the collider can lead to an artificial association between `X` and `Y` that does not reflect a causal process.
 
@@
 
+In this case, players' mental health is a collider of their diet and the quality of the coaching they receive.
+In addition, players' mental health is a confound of their off-field behavior and fantasy points.
+
 ```{r}
 dagitty::impliedConditionalIndependencies(butterflyBias)
+```
 
+As the output indicates, there are several conditional independencies.
+
+```{r}
 dagitty::adjustmentSets(
@@
 ```
+
+When dealing with a collider that is also a confound, controlling for either set, `B` and `M` or `A` and `M`, will provide an unbiased estimate of the association between `X` and `Y`.
+In this case, controlling for either a) coaching quality and mental health or b) diet and mental health—but not both sets—will yield an unbiased estimate of the association between off-field behavior and fantasy points.
 

### 2024-07-03 causal-inference.qmd
20240702 - spelling

@@
 If you control for the collider, you can induce an artificial association between team record and coaching quality.
-Moreover, because doing so induces an artificial association between team record and coaching quality, it can also induce an artifial association between the effects of team record and coaching quality: number of media articles about the team and fantasy points, respectively.
+Moreover, because doing so induces an artificial association between team record and coaching quality, it can also induce an artificial association between the effects of team record and coaching quality: number of media articles about the team and fantasy points, respectively.
 That is, controlling for the collider can lead to an artificial association between `X` and `Y` that does not reflect a causal process.

### 2024-07-03 causal-inference.qmd
20240702 - cross-references

@@
 - `Y` causes `X`
-- a third variable (i.e., confound), `Z`, influences both `X` and `Y`
+- a third variable (i.e., [confound](#sec-causalDiagramConfounding)), `Z`, influences both `X` and `Y`
 - the association between `X` and `Y` is spurious
@@
 If we meet the first criterion of causality, we have removed the possibility that `Y` causes `X` (i.e., reverse causality).
-To meet the third criterion of causality, we need to remove the possibility that the association reflects a third variable (confound) that influences both the cause and effect, and we need to remove the possibility that the association is spurious—the possibility that the association between the purported cause and effect is due to random chance.
+To meet the third criterion of causality, we need to remove the possibility that the association reflects a third variable ([confound](#sec-causalDiagramConfounding)) that influences both the cause and effect, and we need to remove the possibility that the association is spurious—the possibility that the association between the purported cause and effect is due to random chance.
 There are multiple approaches to meeting the third criterion of causality, such as by use of [experiments](#sec-causalInferenceExperiment), [longitudinal designs](#sec-causalInferenceLongitudinal), [control variables](#sec-causalInferenceControlVariables), [within-subject designs](#sec-causalInferenceWithinSubject), and [genetically informed designs](#sec-causalInferenceGeneticallyInformed), as described in @sec-approachesCausalInference.
@@
 For instance, by manipulating sports drink consumption before the player performs, they can eliminate the possibility that reverse causation explains the effect of the [independent variable](#sec-experiment) on the [dependent variable](#sec-experiment).
-Second, through randomly assigning players to consume or not consume sports drink, this holds everything else constant (so long as the groups are evenly distributed according to other factors, such as their age, weight, etc.) and thus removes the possibility that third variable confounds explain the effect of the [independent variable](#sec-experiment) on the [dependent variable](#sec-experiment).
+Second, through randomly assigning players to consume or not consume sports drink, this holds everything else constant (so long as the groups are evenly distributed according to other factors, such as their age, weight, etc.) and thus removes the possibility that third variable [confounds](#sec-causalDiagramConfounding) explain the effect of the [independent variable](#sec-experiment) on the [dependent variable](#sec-experiment).
 
@@
 When we control for prior levels of the outcome in the prediction, we are evaluating whether the predictor is associated with witin-person *change* in the outcome.
-Predicting within-person change provides stronger evidence consistent with causality because it uses the individual as their own control and controls for many time-invariant confounds (i.e., confounds that do not change across time).
-However, predicting within-person change does not, by itself, control for time-varying confounds.
-So, it can also be useful to control for time-varying confounds, such as by use of [control variables](#sec-causalInferenceControlVariables).
+Predicting within-person change provides stronger evidence consistent with causality because it uses the individual as their own control and controls for many time-invariant [confounds](#sec-causalDiagramConfounding) (i.e., [confounds](#sec-causalDiagramConfounding) that do not change across time).
+However, predicting within-person change does not, by itself, control for time-varying [confounds](#sec-causalDiagramConfounding).
+So, it can also be useful to control for time-varying [confounds](#sec-causalDiagramConfounding), such as by use of [control variables](#sec-causalInferenceControlVariables).
 
@@
 
-One of the plausible alternatives to the inference that `X` causes `Y` is that there are third variable confounds that influence both `X` and `Y`, thus explaining why `X` and `Y` are associated, as depicted in @fig-correlationAndCausation3 and @fig-ZCausesXandY.
-Thus, another approach that can help increase [internal validity](#sec-internalValidity) is to include plausible confounds as control variables.
-For instance, if a third variable such as education level might be a confound that influences both sports drink consumption and player performance, you could include education level as a [covariate](#sec-covariates) in the model.
+One of the plausible alternatives to the inference that `X` causes `Y` is that there are third variable [confounds](#sec-causalDiagramConfounding) that influence both `X` and `Y`, thus explaining why `X` and `Y` are associated, as depicted in @fig-correlationAndCausation3 and @fig-ZCausesXandY.
+Thus, another approach that can help increase [internal validity](#sec-internalValidity) is to include plausible [confounds](#sec-causalDiagramConfounding) as control variables.
+For instance, if a third variable such as education level might be a [confound](#sec-causalDiagramConfounding) that influences both sports drink consumption and player performance, you could include education level as a [covariate](#sec-covariates) in the model.
 Inclusion of a [covariate](#sec-covariates) attempts to control for the variable by examining the association between the [predictor variable](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy) while holding the [covariate](#sec-covariates) variables constant.
@@
 A common form of causal diagrams is the directed acyclic graph (DAG).
-DAGs provide a helpful tool to communicate about causal questions and help identify how to avoid bias (i.e., over-estimation) in associations between variables due to confounding (i.e., common causes) [@Digitale2022].
+DAGs provide a helpful tool to communicate about causal questions and help identify how to avoid bias (i.e., over-estimation) in associations between variables due to [confounding](#sec-causalDiagramConfounding) (i.e., common causes) [@Digitale2022].
 For instance, from a DAG, it is possible to determine what variables it is important to control for in order to get unbiased estimates of the association between two variables of interest.
@@
 
-Collider bias may also occur when neither variable of interest is a direct cause of the collider [@Lederer2019].
-M-bias is a form of collider bias that occurs when two variables that are not causally related, `A` and `B`, both influence a collider, `M`, and each (`A` and `B`) also influences a separate variable—e.g., `A` influences `X` and `B` influences `Y`.
+[Collider bias](#sec-colliderBias) may also occur when neither variable of interest is a direct cause of the [collider](#sec-colliderBias) [@Lederer2019].
+M-bias is a form of [collider bias](#sec-colliderBias) that occurs when two variables that are not causally related, `A` and `B`, both influence a [collider](#sec-colliderBias), `M`, and each (`A` and `B`) also influences a separate variable—e.g., `A` influences `X` and `B` influences `Y`.
 M-bias is so-named from the M-shape of the DAG.
@@
 
-In this example, fan attendance is the collider that is influenced separately by the team record and the coaching quality.
+In this example, fan attendance is the [collider](#sec-colliderBias) that is influenced separately by the team record and the coaching quality.
 This is a fictitious example for purposes of demonstration; in reality, coaching quality influences the team's record.
@@
 
-It is important not to control for the collider (fan attendance).
-If you control for the collider, you can induce an artificial association between team record and coaching quality.
+It is important not to control for the [collider](#sec-colliderBias) (fan attendance).
+If you control for the [collider](#sec-colliderBias), you can induce an artificial association between team record and coaching quality.
 Moreover, because doing so induces an artificial association between team record and coaching quality, it can also induce an artificial association between the effects of team record and coaching quality: number of media articles about the team and fantasy points, respectively.
-That is, controlling for the collider can lead to an artificial association between `X` and `Y` that does not reflect a causal process.
+That is, controlling for the [collider](#sec-colliderBias) can lead to an artificial association between `X` and `Y` that does not reflect a causal process.
 
-#### Butterfly Bias {#sec-colliderBias}
+#### Butterfly Bias {#sec-butterflyBias}
 
-Butterfly bias occurs when both confounding and [M-bias](#sec-causalDiagramMBias) are present.
+Butterfly bias occurs when both [confounding](#sec-causalDiagramConfounding) and [M-bias](#sec-causalDiagramMBias) are present.
 Butterfly bias (aka bow-tie bias) is so-named from the butterfly shape of the DAG.
@@
 
-- Two variables (`A` and `B`) influence a collider (`M`).
-- The collider influences two variables, `X` and `Y`.
+- Two variables (`A` and `B`) influence a [collider](#sec-colliderBias) (`M`).
+- The [collider](#sec-colliderBias) influences two variables, `X` and `Y`.
 - `A` also influences `X`.
@@
 
-In butterfly bias, the collider (`M`) is also a confound.
+In butterfly bias, the [collider](#sec-colliderBias) (`M`) is also a [confound](#sec-causalDiagramConfounding).
 That is, a variable is both influenced by two variables and influences two variables.
@@
 
-In this case, players' mental health is a collider of their diet and the quality of the coaching they receive.
-In addition, players' mental health is a confound of their off-field behavior and fantasy points.
+In this case, players' mental health is a [collider](#sec-colliderBias) of their diet and the quality of the coaching they receive.
+In addition, players' mental health is a [confound](#sec-causalDiagramConfounding) of their off-field behavior and fantasy points.
 
@@
 
-When dealing with a collider that is also a confound, controlling for either set, `B` and `M` or `A` and `M`, will provide an unbiased estimate of the association between `X` and `Y`.
+When dealing with a [collider](#sec-colliderBias) that is also a [confound](#sec-causalDiagramConfounding), controlling for either set, `B` and `M` or `A` and `M`, will provide an unbiased estimate of the association between `X` and `Y`.
 In this case, controlling for either a) coaching quality and mental health or b) diet and mental health—but not both sets—will yield an unbiased estimate of the association between off-field behavior and fantasy points.

### 2024-07-03 causal-inference.qmd
20240703 - selection bias

@@
 Selection bias occurs when the selection of participants or their inclusion in analyses depends on the variables being studied.
+For instance, if you are conducting a study on the extent to which sports drink consumption influences fantasy points, there would be selection bias if players are less likely to participate in the study if they score fewer fantasy points.
+
+Now, consider a study in which you conduct a randomized controlled trial (RCT; i.e., an [experiment](#sec-experiment)) to evaluate the effect of a new medication on player performance.
+You randomly assign some players to take the medication and other players to take a placebo.
+Assume the new medication has side effects and leads many of the players who take it to drop out of the study.
+This is an example of attrition bias (i.e., systematic attrition).
+If you were to exclude these individuals from your analysis, it may make it appear that the medication led to better performance, because the players who experienced the side effect (and worse performance) dropped out of the study.
+Hence, conducting an analysis that excludes these players from the analysis would involve selection bias.
 

### 2024-07-03 basic-statistics.qmd
20240703 - conclusion section

@@
 Thus, we do not have confidence that there was a difference or association.
-However, we do not accept the null hypothesis; it could be there we did not observe an effect because we did not have adequate power to detect the effect—e.g., if the effect size was small, the data were noisy, and the sample size was small and/or unrepresentative.
+However, we do not accept the null hypothesis; it could be there we did not observe an effect because we did not have adequate power to detect the effect—e.g., if the [effect size](#sec-practicalSignificance) was small, the data were noisy, and the [sample size](#sec-sampleVsPopulation) was small and/or unrepresentative.
 
@@
 
-- the sample size (*N*): the larger the *N*, the greater the power
-    - for group comparisons, the power depends on the sample size of each group
+- the [sample size](#sec-sampleVsPopulation) (*N*): the larger the *N*, the greater the power
+    - for group comparisons, the power depends on the [sample size](#sec-sampleVsPopulation) of each group
 - the [effect size](#sec-practicalSignificance): the larger the effect, the greater the power
@@
 However, what if consumption of sports drinks is associated with an average improvement of 0.002 points per game.
-A small effect such as this might be detectable with a large sample size.
+A small effect such as this might be detectable with a large [sample size](#sec-sampleVsPopulation).
 This effect would be considered to be reliable/consistent because it is statistically significant.
-However, such an effect would not be practically important because it has an effect size so small that it results in differences that are not practically important.
-Thus, in addition to statistical significance, it is also important to consider practical significance.
+However, such an effect is so small that it results in differences that are not [practically important](#sec-practicalSignificance).
+Thus, in addition to statistical significance, it is also important to consider [practical significance](#sec-practicalSignificance).
 
@@
 - [effect size](#sec-practicalSignificance)
-- sample size
+- [sample size](#sec-sampleVsPopulation)
 - [alpha level](#sec-statisticalSignificance)
 
-Knowing any three of the following, you can calculate the fourth: statistical power, [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance).
+Knowing any three of the following, you can calculate the fourth: statistical power, [effect size](#sec-practicalSignificance), [sample size](#sec-sampleVsPopulation), and [alpha level](#sec-statisticalSignificance).
 Below is `R` code for calculating power for each of various statistical tests (i.e., a *power analysis*).
@@
 ```
+
+When designing a study, it is important to consider power and the [sample size](#sec-sampleVsPopulation) needed to detect the hypothesized effect size.
+If your [sample size](#sec-sampleVsPopulation) is too small and you do not detect an effect (i.e., $p > .05$), you do not know whether your failure to detect the effect was because a) the effect does not exist, or b) the effect exists but you did not have enough power to detect it.
 
@@
 
-There are free tools for calculating power for tests of [mediation](#mediation) and [moderation](#moderation):
+There are free tools for calculating power for tests of [mediation](#sec-mediation) and [moderation](#sec-moderation):
 
@@
 
+## Conclusion {#sec-basicStatsConclusion}
+
+[Descriptive statistics](#sec-descriptiveStatistics) are used to describe the data, including the center, spread, or shape of data.
+[Inferential statistics](#sec-inferentialStatistics) are used to draw inferences regarding differences between groups or associations between variables.
+Null hypothesis signficance testing is a framework for [inferential statistics](#sec-inferentialStatistics), in which there is a [null hypothesis](#sec-nullHypothesis) and [alternative hypothesis](##sec-alternativeHypothesis).
+The null hypothesis is that there is no difference between groups or that there is no association between variables.
+[Statistical significance](#sec-statisticalSignificance) is evaluated with a $p$-value, which represents the probability of obtaining a result at least as extreme as the result observed if the null hypothesis is true.
+Effects with *p*-values less than .05 are considered statistically significant.
+However, it is also important to consider [practical significance](#sec-practicalSignificance) and effect sizes.
+When designing a study, it is important to consider [statistical power](#sec-statisticalPower) and the [sample size](#sec-sampleVsPopulation) needed to detect the hypothesized effect size.
+You can determine a study's [power](#sec-statisticalPower) based on the [effect size](#sec-practicalSignificance), [sample size](#sec-sampleVsPopulation), and [alpha level](#sec-statisticalSignificance).
+
 ## Session Info {#sec-basicStatsSessionInfo}

### 2024-07-03 causal-inference.qmd
20240703 - conclusion section

@@
 
+## Conclusion {#sec-causalInferenceConclusion}
+
+There are three criteria for establishing causality: 1) the cause precedes the effect.
+2) The cause is related to the effect.
+3) There are no other alternative explanations for the effect apart from the cause.
+Various experimental and quasi-experimental designs and approaches can be leveraged to more closely approximate causal inferences.
+Longitudinal designs, within-subject analyses, inclusion of control variables, and genetically informed designs are all quasi-experimental designs that afford the researcher greater control over some possible third variable [confounds](#sec-causalDiagramConfounding).
+Causal diagrams can be a useful tool for identifying the proper variables to control for (and those not to control for).
+When [confounding](#sec-causalDiagramConfounding) exists, it is important to control for the [confound(s)](#sec-causalDiagramConfounding).
+It is important not to control for mediators when interested in the total effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy).
+When there is a [collision](#sec-colliderBias), it is important not to control for the [collider](#sec-colliderBias) unless the [collider](#sec-colliderBias) is also a [confound](#sec-causalDiagramConfounding).
+
 ::: {.content-visible when-format="html"}

### 2024-07-03 correlation.qmd
20240703 - conclusion section

@@
 
-Correlation is a metric of the association between variables.
+Correlation is an index of the association between variables.
 Covariance is the association between variables and in an unstandardized metric that differs for variables with different scales.
@@
 
+## Correlation Does Not Imply Causation
+
+As described in @sec-correlationCausation, correlation does not imply causation.
+There are several reasons (described in @sec-correlationCausation) that, just because `X` is correlated with `Y` does not necessarily mean that `X` causes `Y`.
+However, correlation can still be useful.
+In order for two processes to be causally related, they must be associated.
+That is, association is necessary but insufficient for causality.
+
+## Conclusion {#sec-correlationConclusion}
+
+Correlation is an index of the association between variables.
+The correlation coefficient ($r$) ranges from −1 to +1, and indicates the sign and magnitude of the association.
+Although correlation does not imply causality, identifying associations between variables can still be useful because association is a necessary (but insufficient) condition for causality.
+
 ## Session Info {#sec-correlationSessionInfo}

### 2024-07-03 multiple-regression.qmd
20240703 - conclusion section

@@
 
+## Conclusion {#sec-multipleRegressionConclusion}
+
+Multiple regression allows examining the association between multiple [predictor variables](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy).
+Inclusion of multiple predictors in the model allows for potentially greater predictive accuracy and identification of the extent to which each variable uniquely contributes to the outcome variable.
+As with [correlation](#sec-correlation), an association does not imply causation.
+However, identifiying associations is important because associations are a necessary (but insufficient) condition for causality.
+When developing a multiple regression model, it is important to pay attention for potential [multicollinearity](#sec-multipleRegressionMulticollinearity)—it may become difficult to detect a given [predictor variable](#sec-correlationalStudy) as [statistically significant](#sec-statisticalSignificance) due to the greater uncertainty around the parameter estimates.
+
 ## Session Info {#sec-multipleRegressionSessionInfo}

### 2024-07-03 research-methods.qmd
20240703 - conclusion section

@@
 The participants (i.e., people or players) who we successfully recruit to be in our study represent our *sample*.
+The number of participants in the study is our *sample size*.
 
@@
 
-A mnemonic to help remember the difference between [mediation](#mediation) and [moderation](#moderation) is in @fig-mediationModerationMnemonic.
+A mnemonic to help remember the difference between [mediation](#sec-mediation) and [moderation](#sec-moderation) is in @fig-mediationModerationMnemonic.
 
@@
 
-### Mediation {#mediation}
+### Mediation {#sec-mediation}
 
-#### Overview {#mediationOverview}
+#### Overview {#sec-mediationOverview}
 
@@
 
-#### Types of Mediation {#mediationTypes}
+#### Types of Mediation {#sec-mediationTypes}
 
-##### Full Mediation {#fullMediation}
+##### Full Mediation {#sec-fullMediation}
 
@@
 
-##### Partial Mediation {#partialMediation}
+##### Partial Mediation {#sec-partialMediation}
 
@@
 
-##### Multiple Mediators {#multipleMediators}
+##### Multiple Mediators {#sec-multipleMediators}
 
@@
 
-### Moderation (i.e., Interaction) {#moderation}
+### Moderation (i.e., Interaction) {#sec-moderation}
 
-#### Overview {#moderationOverview}
+#### Overview {#sec-moderationOverview}
 
@@
 
+## Conclusion {#sec-researchMethodsConclusion}
+
+There are various types of research designs.
+Each type of research design differs in the extent to which it supports the ability to draw causal inferences ([internal validity](#sec-internalValidity)) versus the extent to which it supports the ability to identify processes that generalize to the real-world ([external validity](#sec-externalValidity)).
+In addition, it is important to understand the distinction between [sample](#sec-sampleVsPopulation) and [population](#sec-sampleVsPopulation), and the distinction between [mediation](#sec-mediation) and [moderation](#sec-moderation).
+It is also important to consider the [levels of measurement](#sec-levelsOfMeasurement) used because they constrain the types of analyses that may be performed.
+In addition, it is important to consider the [psychometrics](#sec-psychometrics) of measurements, including multiple aspects of [reliability](#sec-reliability) (consistency) and [validity](#sec-validity) (accuracy).
+
 ::: {.content-visible when-format="html"}

### 2024-07-06 cognitive-bias.qmd
20240605 - getting started/load packages
Built site for gh-pages
20240627 - sessionInfo()
20240706 - cognitive biases overview

@@
 # Heuristics and Cognitive Biases in Prediction {#sec-cognitiveBias}
+
+## Getting Started {#sec-cognitiveBiasGettingStarted}
+
+### Load Packages {#sec-cognitiveBiasLoadPackages}
+
+```{r}
+
+```
+
+## Overview {#sec-cognitiveBiasOverview}
+
+When considering judgment and prediction, it is important to consider psychological concepts, including heuristics and cognitive biases.
+In the modern world of big data, research and society need people who know how to make sense of the information around us.
+Given humans' cognitive biases, it is valuable to leverage more objective approaches than relying on our "gut" and intutition.
+Statistical approaches can be a more objective way to identify systematic patterns.
+
+Statistical analysis—and science more generally—is a process to the pursuit of knowledge.
+Science is perhaps the best tool that society has to approximate the truth.
+It relies on empirical evidence and does not give undue weight to anecdotal evidence, intuition, tradition, or authority.
+Nevertheless, statistical analysis is not purely objective and is not a panacea.
+Science is a human enterprise—it is performed by humans each of whom has their own biases.
+For instance, cognitive biases can influence how people interpret statistics.
+As a result, the findings from any given study may be incorrect.
+Thus, it would be imprudent to make decisions based solely on the results of one study.
+That is why we wait for findings to be independently replicated by different groups of researchers using different methods.
+
+If a research team publishes flashy new and exciting findings, other researchers have an incentive to disprove the prior findings.
+Thus, we have more confidence if findings stand up to scrutiny from independent groups of researchers.
+We also draw upon meta-analyses—studies of many studies, to summarize the results of many studies and not just the findings from any single study that may not replicate.
+In this way, we can identify which findings are robust and most likely true versus the findings that fail to replicate.
+Thus, despite its flaws like any other human enterprise, science is a self-correcting process in which the long arc bends toward truth.
+
+In our everyday lives, humans are presented with overwhelming amounts of information.
+Because human minds cannot parse every piece of information equally, we tend to take mental shortcuts, called *heuristics*.
+These mental shortcuts can be helpful.
+They can help us make quick judgments to stay alive or to make complex decisions in the face of uncertainty.
+However, these mental shortcuts can also lead us astray and make systematic errors in our judgments and predictions.
+*Cognitive biases* are systematic errors in thinking.
+
+## Examples of Heuristics {#sec-heuristics}
+
+## Examples of Cognitive Biases {#sec-cognitiveBiases}
+
+## Conclusion {#sec-cognitiveBiasConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-cognitiveBiasSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-06 cognitive-bias.qmd
20240706 - epistemology

@@
 Statistical analysis—and science more generally—is a process to the pursuit of knowledge.
-Science is perhaps the best tool that society has to approximate the truth.
-It relies on empirical evidence and does not give undue weight to anecdotal evidence, intuition, tradition, or authority.
+An *epistemology* is an approach to knowledge.
+Science is perhaps the best approach (epistemology) that society has to approximate truth.
+Unlike other approaches to knowledge, science relies on empirical evidence and does not give undue weight to anecdotal evidence, intuition, tradition, or authority.
 Nevertheless, statistical analysis is not purely objective and is not a panacea.

### 2024-07-06 getting-started.qmd
20240706 - using functions and arguments

@@
 library("tidyverse")
+```
+
+## Using Functions and Arguments {#sec-functionsArguments}
+
+You can learn about a particular function and its arguments by entering a question mark before the name of the function:
+
+```{r}
+#| eval: false
+
+?NAME_OF_FUNCTION()
+```
+
+For instance, to learn about the `seq()` function, which creates a sequence of numbers, you can execute the following command:
+
+```{r}
+#| eval: false
+
+?seq()
+```
+
+```{r}
+#| eval: false
+
+seq(
+  from = 1,
+  to = 1,
+  by = ((to - from)/(length.out - 1)),
+  length.out = NULL,
+  along.with = NULL,
+  ...)
+```
+
+The `seq()` function takes the following arguments:
+
+- `from`
+- `to`
+- `by`
+- `length.out`
+- `along.with`
+- `...`
+
+The arguments have default values that are used if the user does not specify values for the arguments.
+The default values are provided in the `Usage` section and are in @tbl-seqFunction:
+
+| Argument     | Default Value for Argument       |
+|:-------------|:---------------------------------|
+| `from`       | `1`                              |
+| `to`         | `1`                              |
+| `by`         | `((to - from)/(length.out - 1))` |
+| `length.out` | `NULL`                           |
+| `along.with` | `NULL`                           |
+
+: Arguments and defaults for the `seq()` function. Arguments with a default of `NULL` are not used unless a value is provided by the user. {#tbl-seqFunction}
+
+You can specify a function and its arguments either by providing values for each argument in the order indicated by the function, or by naming its arguments.
+
+Here is an example of providing values to the arguments in the order indicated by the function:
+
+```{r}
+seq(1, 9)
+```
+
+Here is an example of providing values to the arguments by naming its arguments:
+
+```{r}
+seq(
+  from = 1,
+  to = 9,
+  by = 1)
+```
+
+If you provide values to arguments by naming the arguments, you can reorder the arguments and get the same answer:
+
+```{r}
+seq(
+  by = 1,
+  to = 9,
+  from = 1)
+```
+
+There are various combinations of arguments that one could use to obtain the same result.
+For instance, here is code to generate a sequence from 1 to 9 by 2:
+
+```{r}
+seq(
+  from = 1,
+  to = 9,
+  by = 2)
+```
+
+Or, alternatively, you could specify the length of the desired sequence (5 values):
+
+```{r}
+seq(
+  from = 1,
+  to = 9,
+  length.out = 5)
+```
+
+If you want to generate a series with gradually incrementing decimal places, you could specify a long desired sequence of 81 values:
+
+```{r}
+seq(
+  from = 1,
+  to = 9,
+  length.out = 81)
+```
+
+This is equivalent to specifying a sequence from 1 to 9 by 0.1:
+
+```{r}
+seq(
+  from = 1,
+  to = 9,
+  by = 0.1)
 ```

### 2024-07-07 causal-inference.qmd
20240706 - wording tweaks

@@
 
-## Correlation Does Not Imply Causality {#sec-correlationCausality}
+## Correlation Does Not Imply Causation {#sec-correlationCausality}
 
@@
 The purported cause must be associated with the purported effect.
-Nevertheless, as the maxim goes, "correlation does not imply causality."
+Nevertheless, as the maxim goes, "correlation does not imply causation."
 Just because two variables are correlated does not necessarily mean that they are causally related.

### 2024-07-07 correlation.qmd
20240706 - wording tweaks

@@
 The correlation coefficient ($r$) ranges from −1 to +1, and indicates the sign and magnitude of the association.
-Although correlation does not imply causality, identifying associations between variables can still be useful because association is a necessary (but insufficient) condition for causality.
+Although correlation does not imply causation, identifying associations between variables can still be useful because association is a necessary (but insufficient) condition for causality.
 

### 2024-07-07 getting-started.qmd
20240706 - clarifications

@@
 1. After installing `RStudio`, open `RStudio` and run the following code in the console to install several key `R` packages:
-
 ```{r}
@@
 install.packages(
-  c("petersenlab","ffanalytics","nflreadr","nflfastR","nfl4th","nflplotR",
+  c("petersenlab","remotes","nflreadr","nflfastR","nfl4th","nflplotR",
   "gsisdecoder","progressr","lubridate","tidyverse","psych"))
+```
+1. Some necessary packages, including the `ffanalytics` package, are hosted in GitHub and need to be installed using the following code (after installing the `remotes` package above):
+```{r}
+#| eval: false
+
+remotes::install_github("FantasyFootballAnalytics/ffanalytics")
 ```
@@
 
-For instance, to learn about the `seq()` function, which creates a sequence of numbers, you can execute the following command:
+Below, we provide examples for how to learn about and use functions and arguments, by using the `seq()` function as an example.
+The `seq()` function creates a sequence of numbers.
+To learn about the `seq()` function, which creates a sequence of numbers, you can execute the following command:
 
@@
 ```
+
+This is what the documentation shows for the `seq()` function in the `Usage` section:
 
@@
 
-The `seq()` function takes the following arguments:
+Based on this information, we know that the `seq()` function takes the following arguments:
 
@@
 
+What each argument represents (i.e., the meaning of `from`, `to`, `by`, etc.) is provided in the `Arguments` section of the documentation.
 You can specify a function and its arguments either by providing values for each argument in the order indicated by the function, or by naming its arguments.
 
-Here is an example of providing values to the arguments in the order indicated by the function:
+Here is an example of providing values to the arguments in the order indicated by the function, to create a sequence of numbers from 1 to 9:
 
@@
 
-If you want to generate a series with gradually incrementing decimal places, you could specify a long desired sequence of 81 values:
+If you want to generate a series with decimal values, you could specify a long desired sequence of 81 values:
 
@@
 
+Hopefully, that provides an example for how to learn about a particular function, its arguments, and how to use them.
+
 ## Download Football Data {#sec-downloadFootballData}
+
+Below, we provide examples for how to download various types of football data.
 

### 2024-07-07 causal-inference.qmd
20240706 - clarifications
20240706 - bidirectional effects

@@
 - `Y` causes `X`
+- `X` and `Y` are bidirectional: `X` causes `Y` and `Y` causes `X`
 - a third variable (i.e., [confound](#sec-causalDiagramConfounding)), `Z`, influences both `X` and `Y`
@@
 - `B` also influences `Y`.
-- `A` and `B` are not causally related
-- `X` and `Y` are not causally related
+- `A` and `B` are not causally related.
+- `X` and `Y` are not causally related.
 

### 2024-07-07 research-methods.qmd
20240706 - bidirectional effects

@@
 - `Y` causes `X`
+- `X` and `Y` are bidirectional: `X` causes `Y` and `Y` causes `X`
 - a third variable (i.e., confound), `Z`, influences both `X` and `Y`
@@
 
-A third possibility is that the association could reflect the influence of a third variable.
+A third possibility is that the association reflects a bidirectional effect, where `X` causes `Y` and `Y` causes `X`, as depicted in INSERT.
+For instance, consumption of more sports drink may improve players' performance, and greater performance in turn may lead players to drink more sports drink.
+
+A fourth possibility is that the association could reflect the influence of a third variable.
 If a third variable is a common cause of each and accounts for their association, it is a *confound*.

### 2024-07-07 cognitive-bias.qmd
20240706 - characteristics of science vs pseudoscience

@@
 Unlike other approaches to knowledge, science relies on empirical evidence and does not give undue weight to anecdotal evidence, intuition, tradition, or authority.
+
+Per @PetersenPrinciplesPsychAssessment, here are the characteristics of science that distinguish it from pseudoscience:
+
+> 1. Risky hypotheses are posed that are falsifiable.
+> The hypotheses can be shown to be wrong.
+> 1. Findings can be replicated independently by different research groups and different methods.
+> Evidence converges across studies and methods.
+> 1. Potential alternative explanations for findings are specified and examined empirically (with data).
+> 1. Steps are taken to guard against the undue influence of personal beliefs and biases.
+> 1. The strength of claims reflects the strength of evidence.
+> Findings and the ability to make judgments or predictions are not overstated.
+> For instance, it is important to present the degree of uncertainty from assessments with error bars or confidence intervals.
+> 1. Scientifically supported measurement strategies are used based on their psychometrics, including [reliability](#sec-reliability) and [validity](#sec-validity).
+
 Nevertheless, statistical analysis is not purely objective and is not a panacea.

### 2024-07-07 research-methods.qmd
20240707 - bidirectional effect

@@
 
-A third possibility is that the association reflects a bidirectional effect, where `X` causes `Y` and `Y` causes `X`, as depicted in INSERT.
+A third possibility is that the association reflects a bidirectional effect, where `X` causes `Y` and `Y` causes `X`, as depicted in @fig-correlationAndCausation3.
 For instance, consumption of more sports drink may improve players' performance, and greater performance in turn may lead players to drink more sports drink.
+
+::: {#fig-correlationAndCausation3}
+![](images/correlationAndCausation_3.png)
+
+Bidirectional Effect Between `X` and `Y`, such that `X` causes `Y` and `Y` causes `X`.
+:::
 
@@
 An observed association between `X` and `Y` could reflect a confound—i.e., a cause (`Z`) that influences both `X` and `Y`, which explains why `X` and `Y` are correlated even though they are not causally related.
-A third variable confound that is a common cause of both `X` and `Y` is depicted in @fig-correlationAndCausation3.
+A third variable confound that is a common cause of both `X` and `Y` is depicted in @fig-correlationAndCausation4.
 For instance, it may not be that sport drink consumption per se influences player performance; rather, it may be that players who are more intelligent or have more financial resources tend to drink more sports drinks and also tend to perform better.
@@
 
-::: {#fig-correlationAndCausation3}
-![](images/correlationAndCausation_3.png)
+::: {#fig-correlationAndCausation4}
+![](images/correlationAndCausation_4.png)
 

### 2024-07-07 causal-inference.qmd
20240707 - quasi-experimental designs

@@
 Thus, we often rely on quasi-experimental designs such as natural experiments and [observational/correlational designs](#sec-correlationalStudy).
+
+We cannot directly test or establish causality from a non-experimental research design.
+Nevertheless, we can leverage various design features that, in combination with other studies using different research methods, collectively strengthen our ability to make causal inferences.
+For instance, there are are no experiments in humans showing that smoking causes cancer—randomly assigning people to smoke or not smoke would not be ethical.
+The causal inference that smoking causes cancer was derived from a combination of experimental studies in rodents and observational studies in humans.
 

### 2024-07-08 cognitive-bias.qmd
20240707 - heuristics and cognitive biases

@@
 These mental shortcuts can be helpful.
-They can help us make quick judgments to stay alive or to make complex decisions in the face of uncertainty.
+They reduce our mental load and can help us make quick judgments to stay alive or to make complex decisions in the face of uncertainty.
 However, these mental shortcuts can also lead us astray and make systematic errors in our judgments and predictions.
@@
 
+Three important heuristics used in judgment and prediction in the face of uncertainty include [@Tversky1974]:
+
+- availability
+- representativeness
+- anchoring and adjustment
+
+The *availability heuristic* refers to the tendency for a person's judgments or predictions about the frequency or probability of something to be made based on how readily instances can be brought to mind.
+For instance, when making fantasy predictions about a player, more recent big performance games may more easily come to mind compared to lower-scoring games and games that occurred longer ago.
+Thus, a manager may be more inclined to pick players to start who had more recent, stronger performances rather than players who have higher long-term averages.
+
+The *representativess heuristic* refers to the tendency for a person's judgments or predictions about individuals to be made based on how similar the individual is to the person's existing mental prototypes.
+For instance, when coming out of college, Tight End Kyle Pitts [drew comparisons](https://247sports.com/article/kyle-pitts-lebron-james-2021-nfl-draft-florida-gators-football-163882176) to the "LeBron James" of Tight Ends (archived at <https://perma.cc/JQB5-XPVL>).
+The idea that his athletic profile leads him to be similar to the prototype of LeBron James may have led him to be too highly drafted by fantasy managers in his first seasons.
+
+::: {.content-visible when-format="html:js"}
+
+Here is a video of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends:
+
+![Video of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends.
+From: <https://x.com/GetUpESPN/status/1380165126108672001> (archived at <https://perma.cc/JW8E-KV2C>)](images/kylePitts.mp4){width=100%}
+
+:::
+
+The representativeness heuristic has been observed in gambling markets for predicting team wins in the National Football League (NFL) [@Woodland2015] and in decision making in fantasy soccer [@Kotrba2020].
+
+The *anchoring and adjustment heuristic* refers to the tendency for a person's judgments or predictions to be made with a reference point—an anchor—as a starting point from which they adjust their estimates upward or downward.
+The anchor is often inaccurate and given too much weight in the person's calculation, and too little adjustment is made to the anchor.
+For instance, a manager is trying to predict how many fantasy points a top Running Back may score.
+The player scored 300 fantasy points last season, but the team added a stronger backup Running Back and changed the Offensive Coordinator to be a more pass-heavy offense.
+The manager may use 300 fantasy points as an anchor (based on the player's performance last season), and may adjust downward 15 points to account for the offseason changes.
+However, it is possible that this downward adjustment is insufficient to account not only for the offseasons changes but also for potential regression effects.
+Regression effects are discussed further below.
+
 ## Examples of Cognitive Biases {#sec-cognitiveBiases}
+
+Examples of cognitive biases that result from one or more of these heuristics include:
+
+- base rate neglect
+- forgetting about regression to the mean
+- overconfidence
+- confirmation bias
+- hindsight bias
+- loss aversion bias
+- endowment bias
+- Dunning–Kruger effect
+
+## Examples of Fallacies {#sec-fallacies}
+
+Fallacies are mistaken beliefs that are often due to heuristics and cognitive biases.
+Examples of fallacies include:
+
+- sunk cost fallacy
+- hot hand fallacy
+- gambler's fallacy
 

### 2024-07-08 draft.qmd
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240707 - heuristics and cognitive biases

@@
 There is no one "right" draft strategy.
+As noted by @Lee2022 in their analysis of fantasy drafts, the effectiveness of any draft strategy depends on the strategies of the other managers in the league.
 Sometimes it works best to "zig" when everyone else is "zagging".
@@
 In contrast to your starting lineup, you may look to draft players on your bench who have greater [uncertainty](#sec-fantasyValueUncertainty) for their high reward potential in a low-risk selection given the lower price.
+
+An alternative to the "stars and scrubs" approach is to wait to draft more "high-value" players after other managers have over-paid for players.
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-draftSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-08 research-methods.qmd
20240707 - interval/ratio

@@
 - income
+- stock price
 - years of formal education

### 2024-07-08 basic-statistics.qmd
20240707 - interval/ratio
20240707 - Cohen cite

@@
 
-Effect size thresholds [@McGrath2006] for small, medium, and large effect sizes are in @tbl-effectSizeThresholds.
+Effect size thresholds [@Cohen1988; @McGrath2006] for small, medium, and large effect sizes are in @tbl-effectSizeThresholds.
 
@@
 Changes were made to the original, including the addition of several statistical tests.
+*Note*: "Interval" as a level of measurement includes data with an "[interval](#sec-interval)" or higher level of measurement; thus, it also includes data with a "[ratio](#sec-ratio)" level of measurement.
 :::

### 2024-07-08 basic-statistics.qmd
20240708 - robust indices of central tendency and variability

@@
 library("petersenlab")
+library("DescTools")
 library("pwr")
@@
 - median
+- Hodges-Lehmann statistic (aka pseudomedian)
 - mode
 
-The mean of $X$ (written as: $\bar{X}$) is calculated as:
+The mean of $X$ (written as: $\bar{X}$) is calculated as in @eq-IQR:
 
@@
 That is, to compute the mean, sum all of the values and divide by the number of values ($n$).
+One issue with the mean is that it is sensitive to extreme (outlying) values.
+For instance, the mean of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_mean`.
+
+```{r}
+#| code-fold: true
+
+exampleValues <- c(0, 0, 10, 15, 20, 30, 1000)
+exampleValues_mean <- mean(exampleValues)
+```
 
 The median is determined as the value at the 50th percentile (i.e., the value that is higher than 50% of the values and is lower than the other 50% of values).
+Compared to the mean, the median is less influenced by outliers.
+The median of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_median`.
+
+```{r}
+#| code-fold: true
+
+exampleValues_median <- median(exampleValues)
+```
+
+The Hodges-Lehmann statistic (aka pseudomedian) is computed as the median of all pairwise means, and it is also robust to outliers.
+The pseudomedian of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_pseudomedian`.
+
+```{r}
+#| code-fold: true
+
+exampleValues_pseudomedian <- DescTools::HodgesLehmann(exampleValues)
+```
 
 The mode is the most common/frequent value.
+The mode of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_mode`.
+
+```{r}
+#| code-fold: true
+
+exampleValues_mode <- petersenlab::Mode(exampleValues)
+```
 
@@
 - interquartile range (IQR)
-
-The (sample) variance of $X$ (written as: $s^2$) is calculated as:
+- median absolute deviation
+
+The (sample) variance of $X$ (written as: $s^2$) is calculated as in @eq-variance:
 
@@
 
-The (sample) standard deviation of $X$ (written as: $s$) is calculated as:
+where $X_i$ is each data point, $\bar{X}$ is the mean of $X$, and $n$ is the number of data points.
+
+The (sample) standard deviation of $X$ (written as: $s$) is calculated as in @eq-sd:
 
@@
 
-The range is calculated of $X$ is calculated as:
+The range is calculated of $X$ is calculated as in @eq-range:
 
@@
 
-The interquartile range (IQR) is calculated as:
+The interquartile range (IQR) is calculated as in @eq-IQR:
 
@@
 where $Q_3$ is the score at the third quartile (i.e., 75th percentile), and $Q_1$ is the score at the first quartile (i.e., 25th percentile).
+
+The median absolute deviation (MAD) is the median of all deviations from the median, and is calculated as in @eq-medianAbsoluteDeviation:
+
+$$
+\text{MAD} = \text{median}(|X_i - \tilde{X}|)
+$$ {#eq-medianAbsoluteDeviation}
+
+where $\tilde{X}$ is the median of `X`.
+Compared to the standard deviation, the median absolute deviation is more robust to outliers.
 
@@
 #        IQR = ~ IQR(., na.rm = TRUE),
+#        mad = ~ mad(., na.rm = TRUE),
 #        median = ~ median(., na.rm = TRUE),
+#        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
 #        mode = ~ petersenlab::Mode(., multipleModes = "mean"),

### 2024-07-08 basic-statistics.qmd
20240708 - reorder computations

@@
 
+```{r}
+#| code-fold: true
+
+exampleValues <- c(0, 0, 10, 15, 20, 30, 1000)
+exampleValues_mean <- apa(mean(exampleValues), 2)
+```
+
 That is, to compute the mean, sum all of the values and divide by the number of values ($n$).
@@
 
-exampleValues <- c(0, 0, 10, 15, 20, 30, 1000)
-exampleValues_mean <- mean(exampleValues)
+exampleValues_median <- median(exampleValues)
 ```
@@
 
-exampleValues_median <- median(exampleValues)
+exampleValues_pseudomedian <- DescTools::HodgesLehmann(exampleValues)
 ```
@@
 
-exampleValues_pseudomedian <- DescTools::HodgesLehmann(exampleValues)
+exampleValues_mode <- petersenlab::Mode(exampleValues)
 ```
@@
 The mode of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_mode`.
-
-```{r}
-#| code-fold: true
-
-exampleValues_mode <- petersenlab::Mode(exampleValues)
-```
 
@@
 #        IQR = ~ IQR(., na.rm = TRUE),
-#        mad = ~ mad(., na.rm = TRUE),
+#        MAD = ~ mad(., na.rm = TRUE),
 #        median = ~ median(., na.rm = TRUE),

### 2024-07-08 basic-statistics.qmd
20240708 - weighted mean, weighted median (or other quantile)

@@
 - mode
+- weighted mean
+- weighted median
 
@@
 
+If you want to give some values more weight to others, you can calculate a weighted mean and a weighted median (or other quantile), while assigning a weight to each value.
+
 Below is R code to estimate each:
@@
 ```{r}
-
+#mean(data, na.rm = TRUE)
+#median(data, na.rm = TRUE)
+#DescTools::HodgesLehmann(exampleValues, na.rm = TRUE)
+#petersenlab::Mode(exampleValues)
+#weighted.mean(data, weights, na.rm = TRUE)
+#DescTools::Quantile(data, weights, na.rm = TRUE)
 ```

### 2024-07-09 base-rates.qmd
20240608 - getting started
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240703 - conclusion section
20240708 - base rate fallacy

@@
 
-## Getting Started
+## Getting Started {#sec-baseRatesGettingStarted}
 
-### Load Packages
+### Load Packages {#sec-baseRatesLoadPackages}
 
@@
 ```
+
+## Overview {#sec-baseRatesConclusion}
+
+The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
+For instance, the base rate of HIV is the probability of developing HIV.
+In the U.S., [the prevalence rate of HIV is ~0.4% of the adult population](https://map.aidsvu.org/profiles/nation/usa/overview) (archived at <https://perma.cc/8GE6-GAPC>).
+
+## Conclusion {#sec-baseRatesConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-baseRatesSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-09 cognitive-bias.qmd
20240708 - base rate fallacy

@@
 They reduce our mental load and can help us make quick judgments to stay alive or to make complex decisions in the face of uncertainty.
-However, these mental shortcuts can also lead us astray and make systematic errors in our judgments and predictions.
+However, these mental shortcuts can also lead us astray and to make systematic errors in our judgments and predictions.
 *Cognitive biases* are systematic errors in thinking.
+*Fallacies* are forms of flawed reasoning.
 
@@
 
+### Availability Heuristic {#sec-heuristicsAvailability}
+
 The *availability heuristic* refers to the tendency for a person's judgments or predictions about the frequency or probability of something to be made based on how readily instances can be brought to mind.
@@
 Thus, a manager may be more inclined to pick players to start who had more recent, stronger performances rather than players who have higher long-term averages.
+
+### Representativeness Heuristic {#sec-heuristicsRepresentativeness}
 
@@
 
+### Anchoring and Adjustment Heuristic {#sec-heuristicsAnchoringAdjustment}
+
 The *anchoring and adjustment heuristic* refers to the tendency for a person's judgments or predictions to be made with a reference point—an anchor—as a starting point from which they adjust their estimates upward or downward.
@@
 
-- base rate neglect
-- forgetting about regression to the mean
-- overconfidence
+- overconfidence bias
 - confirmation bias
+- recency bias
 - hindsight bias
@@
 - endowment bias
-- Dunning–Kruger effect
+- bandwagon effect bias
+- Dunning–Kruger effect bias
+
+### Overconfidence Bias {#sec-cognitiveBiasesOverconfidence}
+
+### Confirmation Bias {#sec-cognitiveBiasesConfirmation}
+
+### Recency Bias {#sec-cognitiveBiasesRecency}
+
+### Hindsight Bias {#sec-cognitiveBiasesHindsight}
+
+### Loss Aversion Bias {#sec-cognitiveBiasesLossAversion}
+
+### Endowment Bias {#sec-cognitiveBiasesEndowment}
+
+### Bandwagon Effect Bias {#sec-cognitiveBiasesBandwagon}
+
+### Dunning–Kruger Effect Bias {#sec-cognitiveBiasesDunningKruger}
 
@@
 
+- base rate fallacy (aka base rate neglect)
+- regression fallacy
 - sunk cost fallacy
@@
 - gambler's fallacy
+
+### Base Rate Fallacy {#sec-fallaciesBaseRate}
+
+The *base rate fallacy* (aka base rate neglect) is the tendency to ignore information about the general probability of an event in favor of specific information about the event.
+The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
+For example, the base rate of work-related injury is the general probability of experiencing a work-related injury, irrespective of other factors (e.g., the type of job, the person's age, the person's sex).
+Among the working population in the U.S., the lifetime prevalence of work-related injuries (i.e., the percent of people who will experience a work-related injury at some point in their lives), is ~35% (<https://www.cdc.gov/mmwr/volumes/69/wr/mm6913a1.htm>; archived at <https://perma.cc/A2L6-WPEH>).
+Thus, the base rate of work-related injuries in the U.S. is ~35%.
+The probability of work-related injuries is higher for some occupations (e.g., construction) and for some groups (e.g., men, 55–64-year-olds, Black or Multiracial, who are self-employed and have less than high school education) than others.
+Nevertheless, is we ignore all of the interacting factors, the general probability of work-related injuries is 35%.
+If we made a prediction that someone would be highly likely (> 90%) to experience a work-related injury because they are male and self-employed, this would be ignoring the relatively lower base rate of work-related injury.
+Indeed, even men (36.7%) and self-employed individuals (41.2%) have less than a 50% chance of experiencing a work-related injury.
+
+As applied to fantasy football, consider that you read about a potential sleeper Wide Receiver who had a stellar performance in a preseason game.
+If you select this player early on in the draft based on this information, this would be ignoring the general probability that most players who have a strong performance in the preseason do not perform as well in the regular season.
+Performance in the preseason is not strongly predictive of performance in the regular season (<https://fivethirtyeight.com/features/the-nfl-preseason-is-not-predictive-but-it-can-often-seem-that-way/>; archived at <https://perma.cc/FSG2-6AXE>).
+
+### Regression Fallacy {#sec-fallaciesRegression}
+
+### Sunk Cost Fallacy {#sec-fallaciesSunkCost}
+
+### Hot Hand Fallacy {#sec-fallaciesHotHand}
+
+### Gambler's Fallacy {#sec-fallaciesGambler}
 

### 2024-07-09 cognitive-bias.qmd
20240708 - regression fallacy

@@
 
+The regression fallacy is the failure to account for the fact that things tend to naturally fluctuate around their mean and that, after an extreme fluctuation, subsequent scores tend to regress (or reverse) to the mean.
+An example of the regression fallacy is the so-called Sports Illustrated or Madden cover jinx curse.
+The Sports Illustrated or Madden cover jinx curse is the urban legend that players who appear on the cover of Sports Illustrated or Madden will perform poorly.
+But, such a phenomenon can be more simply explained by regression to the mean (<https://www.psychologytoday.com/us/blog/what-the-luck/201610/the-sports-illustrated-cover-jinx>; archived at <https://perma.cc/CZM9-TVFN>).
+When a player has a superb season, they likely benefited from some degree to good luck, and it is unlikely that they will repeat such a stellar season the following year.
+Instead, they are likely—at least based on random fluctuation—to regress to their long-term mean.
+
+Applied to fantasy football, consider that a Quarterback had a 5-touchdown game in Week 1.
+You are in need of a strong Quarterback, so you drop a solid player to pick him up.
+However, it is possible that the Quarterback benefited from playing against a week defense in the first game of the season.
+Future matchups may prove more difficult, and the player is unlikely to sustain such a solid performance consistently throughout the season.
+
 ### Sunk Cost Fallacy {#sec-fallaciesSunkCost}

### 2024-07-10 cognitive-bias.qmd
20240709 - specify

@@
 An example of the regression fallacy is the so-called Sports Illustrated or Madden cover jinx curse.
-The Sports Illustrated or Madden cover jinx curse is the urban legend that players who appear on the cover of Sports Illustrated or Madden will perform poorly.
+The Sports Illustrated or Madden cover jinx curse is the urban legend that players who appear on the cover of Sports Illustrated (the magazine) or Madden (the video game) will perform poorly.
 But, such a phenomenon can be more simply explained by regression to the mean (<https://www.psychologytoday.com/us/blog/what-the-luck/201610/the-sports-illustrated-cover-jinx>; archived at <https://perma.cc/CZM9-TVFN>).

### 2024-07-10 cognitive-bias.qmd
20240709 - fallacies

@@
 As applied to fantasy football, consider that you read about a potential sleeper Wide Receiver who had a stellar performance in a preseason game.
-If you select this player early on in the draft based on this information, this would be ignoring the general probability that most players who have a strong performance in the preseason do not perform as well in the regular season.
+If you select this player early on in the draft based on this information, this would be ignoring the general probability that most players who have a strong performance in the preseason do not perform as well in the regular season (i.e., base rate neglect).
 Performance in the preseason is not strongly predictive of performance in the regular season (<https://fivethirtyeight.com/features/the-nfl-preseason-is-not-predictive-but-it-can-often-seem-that-way/>; archived at <https://perma.cc/FSG2-6AXE>).
@@
 However, it is possible that the Quarterback benefited from playing against a week defense in the first game of the season.
-Future matchups may prove more difficult, and the player is unlikely to sustain such a solid performance consistently throughout the season.
+Future matchups may prove more difficult, and the player is unlikely to sustain such a solid performance consistently throughout the season (i.e., they are likely to regress toward their mean).
 
@@
 
+A *sunk cost* is a cost (e.g., in money, time, or effort) that has already been incurred and cannot be recovered.
+For instance, if a person orders an expensive meal at a restaurant, the order is a sunk cost.
+The *sunk cost fallacy* is the tendency to continue an endeavor when there is a sunk cost.
+For instance, when ordering the expensive meal at the restaurant, a person may over-eat so that they feel that they eat their money's worth of food.
+
+Applied to fantasy football, consider a situation in which you invest a lot of salary cap or a high draft pick to draft a promising player, but they repeatedly underperform.
+If you continue to start the player to justify your large investment, instead of benching him in favor of a higher-performing player, you are committing the sunk cost fallacy.
+
 ### Hot Hand Fallacy {#sec-fallaciesHotHand}
 
+The "hot hand" is the idea that a player who experiences a successful outcome will have greater chance of success in subsequent attempts.
+For instance, in basketball, it is widely claimed by coaches, players, and commentators that players who have the hot hand (i.e., who are "on fire") are more likely to make shots because they made previous shots.
+Evidence on the hot hand is mixed.
+Considerable evidence historically has suggested that there is no such thing as a "hot hand" [@Gilovich1985; @BarEli2006; @Avugos2013].
+Some recent research, however, has suggested that there may be a small hot hand effect in some contexts [@Bocskocsky2014; @Miller2014a].
+However, if any such effect exists, the hot hand may be limited to a small subset of players and the [effect size](#sec-practicalSignificance) of any hot hand effect appears to be small [@Pelechrinis2022].
+
+In football, when trying how to distribute the ball among multiple Running Backs, it is not uncommon to hear that a coach wants to give the ball to the Running Back with the "hot hand."
+In fantasy football, consider that a player just had a multiple touchdown game.
+Due to the hot hand fallacy, a manager might continue to start the player because they believe the player is "on fire" and is likely to continue to score at an unsustainable rate.
+
 ### Gambler's Fallacy {#sec-fallaciesGambler}
+
+The gambler's fallacy occurs due to an erroneous belief in the law of small numbers.
+The law of large numbers is a mathematical theorem that the average of a sufficiently large number of independent observations converges to the true value.
+For instance, if you flip a fair coin 1 million times, it is likely to land heads-up ~50% of the time.
+The law of *small* numbers (aka hasty generalization), by contrast, is an erroneous belief that small samples are representative of the populations from which they were drawn.
+For instance, if you flip a coin 10 times, belief in the law of small numbers would lead one to believe that the coin will flip heads-up exactly 5 times out of 10.
+However, in reality, the chance is less than 1 in 4 (24.6%) that exactly 5 of 10 coin flips turn up heads:
+
+```{r}
+dbinom(
+  x = 5,     # number of coins that flip heads-up
+  size = 10, # how many times you flip a coin
+  prob = 0.5 # probability of a coin flipping heads-up (i.e., fair coin = 50%)
+)
+```
+
+The *gambler's fallacy* is the erroneous belief that future probabilities are influenced by past events, even when the events are independent.
+For example, a gambler may pay close attention to a particular slot machine.
+If the slot machine has not paid out in a while, the gambler may believe that the slot machine is about to pay out soon, and may start putting coins in the slots.
+
+Applied to fantasy football, consider that a Quarterback has had several lousy games in a row.
+The gambler's fallacy might lead a manager to start the player under the belief that the player "is due" for a big game, expecting a strong performance from the player merely because they player has not had a good game in a while.
 

### 2024-07-10 cognitive-bias.qmd
20240709 - quote
20240709 - cross-references

@@
 The manager may use 300 fantasy points as an anchor (based on the player's performance last season), and may adjust downward 15 points to account for the offseason changes.
-However, it is possible that this downward adjustment is insufficient to account not only for the offseasons changes but also for potential regression effects.
-Regression effects are discussed further below.
+However, it is possible that this downward adjustment is insufficient to account not only for the offseasons changes but also for potential [regression effects](#sec-fallaciesRegression).
+[Regression effects](#sec-fallaciesRegression) are discussed further in @sec-fallaciesRegression.
 
@@
 ### Dunning–Kruger Effect Bias {#sec-cognitiveBiasesDunningKruger}
+
+> "The more you know, the more you know you don't know." – Anonymous
 

### 2024-07-10 getting-started.qmd
20240706 - try indentation
20240706 - fix indentation
20240707 - Cite Intro to NFL Analytics
20240710 - blog instructions

@@
 1. Download GitHub Desktop: <https://desktop.github.com>
+1. Make sure you are logged into your GitHub account on [GitHub.com](https://github.com).
+1. Go to the following GitHub repository: <https://github.com/isaactpetersen/QuartoBlogFantasyFootball> and complete the following steps:
+  1. Click "Use this Template" (in the top right of the screen) > "Create a new repository"
+  1. Make sure the checkbox is selected for the following option: "Include all branches"
+  1. Make sure your Owner account is selected
+  1. Specify the repository name to whatever you want, such as `FantasyFootballBlog`
+  1. Type a brief description, such as `Files for my fantasy football blog`
+  1. Keep the repository public (this is necessary for generating your blog)
+  1. Select "Create repository"
+1. After creating the new repository, make sure you are on the page of of your new repository and complete the following steps:
+  1. Click "Settings" (in the topof the screen)
+  1. Click "Actions" (in the left sidebar) > "General"
+  1. Make sure the following are selected:
+    - "Read and write permissions" (under "Workflow permissions")
+    - "Allow GitHub Actions to create and approve pull requests"
+    - then click "Save"
+  1. Click "Pages" (in the left sidebar)
+  1. Make sure the following are selected:
+    - "Deploy from a branch" (under "Source")
+    - "gh-pages/(root)" (under "Branch")
+    - then click "Save"
+1. Clone the repository to your local computer by clicking "Code" > "Open with GitHub Desktop", select the folder where you want the repository to be saved on your local computer, and click "Clone"
 :::
@@
 
-Below, we provide examples for how to download various types of football data.
+Below, we provide examples for how to download various types of National Football League (NFL) data.
+For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
 

### 2024-07-10 causal-inference.qmd
20240710 - counterfactuals

@@
 To meet the third criterion of causality, we need to remove the possibility that the association reflects a third variable ([confound](#sec-causalDiagramConfounding)) that influences both the cause and effect, and we need to remove the possibility that the association is spurious—the possibility that the association between the purported cause and effect is due to random chance.
+
 There are multiple approaches to meeting the third criterion of causality, such as by use of [experiments](#sec-causalInferenceExperiment), [longitudinal designs](#sec-causalInferenceLongitudinal), [control variables](#sec-causalInferenceControlVariables), [within-subject designs](#sec-causalInferenceWithinSubject), and [genetically informed designs](#sec-causalInferenceGeneticallyInformed), as described in @sec-approachesCausalInference.
+
+In general, to meet the third criterion of causality, one must consider the counterfactual.
+A *counterfactual* is a hypothetical scenario—it did not occur.
+However, when engaging in causal inference, it is important to consider what would have happened had the hypothetical scenario occurred.
+For instance, consider that we conduct an experiment to randomly assign some players to consume a sports drink before a game and other players to drink only water.
+In this case, our treatment/intervention group is the group of players that consumed a sports drink.
+The control group is the group players that drank only water.
+Now, consider that the players in the treatment group outperform the players in the the control group in their football game.
+In such a study, we observe what *did happen* when players received a treatment.
+The counterfactual is knowledge of what *would have happened* to those same players if they simultaneously had not received treatment [@Shadish2002].
+The true causal effect, then, is the difference between what did happen and what would have happened.
+However, we cannot observe a counterfactual.
+That is, we do not know for sure what would have happened to the players who received treatment if those same players had actually not received treatment.
+We have a control group, but the control group does not have the same players as the intervention group, and it is impossible for a person to simultaneously receive and not receive treatment.
+
+So, our goal in working toward causal inference as scientists is to create reasonable approximations to this impossible counterfactual [@Shadish2002].
+For instance, if using a [between-subject design](#sec-withinSubject), we want the two groups to be equivalent in every possible way except whether or not they receive the treatment, so we might stratify each group to be equivalent in terms of age, weight, position, experience, skill, etc.
+Or, we might test the same people using an A-B-A-B [within-subject design](#sec-withinSubject).
+In an A-B-A-B [within-subject design](#sec-withinSubject), players receive no treatment at baseline (timepoint 1: game 1), receive the treatment at timepoint 2 (game 2), receive no treatment at timepoint 3 (game 3), and receive the treatment at timepoint 4 (game 4).
+Neither of these approximations is a true counterfactual.
+In the [between-subject design](#sec-withinSubject), the players differ between the two groups, so we cannot know how the individuals who received the treatment would have performed if they had actually not received the treatment.
+In the A-B-A-B [within-subject design](#sec-withinSubject), the players are the same, but they timepoints that they receive or do not receive the treatment differ, and there can be [carryover effects](#sec-withinSubject) from one condition to the next.
+For instance, consuming sports drinks before game 2 might also help them be better hydrated in general, including, for subsequent games.
+Thus, we cannot know how a player would have performed in game 1 with treatment or in game 2 without treatment, etc.
+Nevertheless, it is important to be aware of the counterfactual and to engage in counterfactual reasoning to consider what would have happened if various hypothetical scenarios had occurred.
+Considering the counterfactual is important for designing closer approximations to the counterfactual in studies for stronger research designs and stronger causal inference.
 

### 2024-07-10 causal-inference.qmd
20240710 - tweaks

@@
 In general, to meet the third criterion of causality, one must consider the counterfactual.
-A *counterfactual* is a hypothetical scenario—it did not occur.
-However, when engaging in causal inference, it is important to consider what would have happened had the hypothetical scenario occurred.
+A *counterfactual* is what would have happened in the hypothetical scenario that the cause did not occur [i.e., what would have happened in the absence of the cause; @Shadish2002].
+When engaging in causal inference, it is important to consider what would have happened if the hypothetical cause had actually not occurred.
 For instance, consider that we conduct an experiment to randomly assign some players to consume a sports drink before a game and other players to drink only water.
@@
 Thus, we cannot know how a player would have performed in game 1 with treatment or in game 2 without treatment, etc.
-Nevertheless, it is important to be aware of the counterfactual and to engage in counterfactual reasoning to consider what would have happened if various hypothetical scenarios had occurred.
+Nevertheless, it is important to be aware of the counterfactual and to engage in counterfactual reasoning to consider what would have happened if the supposed cause had occurred.
 Considering the counterfactual is important for designing closer approximations to the counterfactual in studies for stronger research designs and stronger causal inference.

### 2024-07-10 causal-inference.qmd
20240710 - cross-reference

@@
 So, our goal in working toward causal inference as scientists is to create reasonable approximations to this impossible counterfactual [@Shadish2002].
-For instance, if using a [between-subject design](#sec-withinSubject), we want the two groups to be equivalent in every possible way except whether or not they receive the treatment, so we might stratify each group to be equivalent in terms of age, weight, position, experience, skill, etc.
+For instance, if using a [between-subject design](#sec-betweenSubject), we want the two groups to be equivalent in every possible way except whether or not they receive the treatment, so we might stratify each group to be equivalent in terms of age, weight, position, experience, skill, etc.
 Or, we might test the same people using an A-B-A-B [within-subject design](#sec-withinSubject).
@@
 Neither of these approximations is a true counterfactual.
-In the [between-subject design](#sec-withinSubject), the players differ between the two groups, so we cannot know how the individuals who received the treatment would have performed if they had actually not received the treatment.
+In the [between-subject design](#sec-betweenSubject), the players differ between the two groups, so we cannot know how the individuals who received the treatment would have performed if they had actually not received the treatment.
 In the A-B-A-B [within-subject design](#sec-withinSubject), the players are the same, but they timepoints that they receive or do not receive the treatment differ, and there can be [carryover effects](#sec-withinSubject) from one condition to the next.

### 2024-07-10 causal-inference.qmd
20240710 - counterfactual

@@
 Thus, we cannot know how a player would have performed in game 1 with treatment or in game 2 without treatment, etc.
-Nevertheless, it is important to be aware of the counterfactual and to engage in counterfactual reasoning to consider what would have happened if the supposed cause had occurred.
+Nevertheless, it is important to be aware of the counterfactual and to engage in counterfactual reasoning to consider what would have happened if the supposed cause had not occurred.
 Considering the counterfactual is important for designing closer approximations to the counterfactual in studies for stronger research designs and stronger causal inference.
@@
 3) There are no other alternative explanations for the effect apart from the cause.
+In general, it is important to be aware of the counterfactual and to consider what would have happened if the supposed cause had not occurred.
 Various experimental and quasi-experimental designs and approaches can be leveraged to more closely approximate causal inferences.

### 2024-07-10 cognitive-bias.qmd
20240710 - overconfidence

@@
 ### Overconfidence Bias {#sec-cognitiveBiasesOverconfidence}
+
+In general, people tend to be overconfident in their judgments and predictions.
+*Overconfidence bias* is the tendency for a person to have greater confidence in their abilities (including judgments and predictions) than is objectively warranted.
+There are three general ways that overconfidence has been identified [@Moore2008]:
+
+1. *overestimation* of one's actual performance
+1. *overplacement* of one's performance relative to others
+1. *overprecision* in one's beliefs/jugments/predictions
+
+*Overestimation* involves believing that one will perform better than one actually performs.
+Overestimation can be identified with a [calibration plot](#sec-calibration) of the predicted performance versus actual performance, where the person's predicted performance is systematically higher (in at least some cases) than their actual performance.
+Overestimation corresponds to the "[overprediction](#sec-calibration)" form of [miscalibration](#sec-calibration).
+
+*Overplacement* involves believing that one is better than others or will perform better than others, even when they do not.
+For instance, it is a common finding that more than half of people believe they are "above average" (i.e., above the median), even though that is statistically impossible.
+This calls to mind the fictitious Lake Wobegon in the radio show *A Praririe Home Companion*, "where all the women are strong, all the men are good-looking, and all the children are above average."
+
+*Overprecision* involves expressing excessive certainty regarding the accuracy of one's beliefs/judgments/predictions.
+For instance, if when a given weather forecaster says it will rain 80% of the time, it actually rains 30% of the time, the weather forecaster's predictions are overprecise.
+Likewise, if the weather forecast says it will rain 10% of the time and it actually rains 30% of the time, the predictions are also overprecise because the forecaster is expressing stronger confidence than is warranted that it will not rain.
+Overprecision can be identified with a [calibration plot](#sec-calibration) of the predicted probabilities versus the actual probabilities.
+Overprecision corresponds to the "[overextremity](#sec-calibration)" form of [miscalibration](#sec-calibration).
+
+A fantasy manager may be even more likely to exhibit overconfidence if they previously performed well or won their league, for which luck and random chance plays an important role.
+Indeed, it is estimated that nearly half (~45%) of the variability in fantasy football performance is estimated to be luck [and around 55% due to skill; @Getty2018].
+A manager who won their league in the prior season may believe they will perform better than they actually will (overestimation), will perform better than average (overplacement), and may hold excessive confidence regarding the accuracy of their predictions about which players will perform well or poorly (overprecision).
+These various types of overconfidence may lead them to draft high-risk players based on gut feeling, neglecting statistical analysis and expert consensus.
 

### 2024-07-10 evaluating-prediction-accuracy.qmd
20240605 - getting started/load packages
20240608 - getting started
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240703 - conclusion section
20240710 - overconfidence

@@
 # Evaluation of Prediction/Forecasting Accuracy {#sec-predictionAccuracy}
+
+## Getting Started {#sec-predictionAccuracyGettingStarted}
+
+### Load Packages {#sec-predictionAccuracyLoadPackages}
+
+```{r}
+
+```
+
+# Types of Accuracy {#sec-accuracyTypes}
+
+## Discrimination {#sec-discrimination}
+
+[Discuss options for discrete versus continuous data]
+
+## Calibration [#sec-calibration]
+
+[Discuss options for discrete versus continuous data]
+
+The various types of miscalibration are depicted in @fig-miscalibration:
+
+```{r}
+#| label: fig-miscalibration
+#| fig-cap: "Types of Miscalibration. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
+#| code-fold: true
+
+# Specify data
+examplePredictions <- seq(from = 0, to = 1, by = .1)
+exampleOutcomes <- c(0, .15, .3, .4, .45, .5, .55, .6, .7, .85, 1)
+
+overPrediction <- c(0, .02, .05, .1, .15, .2, .3, .4, .5, .7, 1)
+underPrediction <- c(0, .3, .5, .6, .7, .8, .85, .9, .95, .98, 1)
+overExtremity <- c(0, .3, .38, .42, .47, .5, .53, .58, .62, .7, 1)
+underExtremity <- c(0, .05, .08, .11, .2, .5, .8, .89, .92, .95, 1)
+
+# Plot
+par(
+  mfrow = c(2,2),
+  mar = c(5,4,1,1) + 0.1) #margins: bottom, left, top, right
+
+plot(
+  examplePredictions,
+  overExtremity,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Overextremity",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+
+plot(
+  examplePredictions,
+  underExtremity,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Underextremity",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+
+plot(
+  examplePredictions,
+  overPrediction,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Overprediction",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+
+plot(
+  examplePredictions,
+  underPrediction,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Underprediction",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+```
+
+## General Accuracy {#sec-generalAccuracy}
+
+[Discuss options for discrete versus continuous data]
+
+## Conclusion {#sec-predictionAccuracyConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-predictionAccuracySessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-10 cognitive-bias.qmd
20240710 - confirmation bias

@@
 
+*Confirmation bias* is the tendency for people to search for, interpret, and remember information that confirms one's beliefs, as opposed to information that might disconfirm one's beliefs.
+The result of confirmation bias is that people are unlikely to change their minds about something that they have a pre-existing belief about, because they tend to look only for information that supports their pre-existing beliefs.
+For instance, if you believe that a particular player is a strong breakout candidate to be a sleeper, you may be more likely to pay attention to evidence that supports that the player will breakout and may be less likely to pay attention to evidence that indicates the player may struggle.
+
 ### Recency Bias {#sec-cognitiveBiasesRecency}

### 2024-07-10 cognitive-bias.qmd
20240710 - recency bias

@@
 
+*Recency bias* is the tendency to weigh recent events more than earlier ones.
+For instance, a manager might observe that a Running Back on the waiver wire performed well in the last two games.
+*Recency bias* may lead the manager to pick up the player, overvaluing their recent performance.
+For instance, the manager may not have adequately weighed the player's overall season performance and the fact that the starting Running Back is returning to the lineup from injury, and that is why the player received more carries in the past two games (i.e., in place of the injured starter).
+
 ### Hindsight Bias {#sec-cognitiveBiasesHindsight}

### 2024-07-10 cognitive-bias.qmd
20240710 - hindsight bias

@@
 
+> "Hindsight is 20/20." – Idiom
+
+*Hindsight bias* is the tendency to perceive that past events were more predictable than they were.
+People tend to remember the succeess of their predictions and forget the failures of their predictions.
+For instance, if a third-string Quarterback has a breakout game, a fantasy manager may claim that they "knew it all along" that the player was going to breakout, despite not having picked up the player.
+That same manager may forget the many other predictions they had that did not come true.
+
 ### Loss Aversion Bias {#sec-cognitiveBiasesLossAversion}

### 2024-07-10 cognitive-bias.qmd
20240710 - Dunning-Krueger effect

@@
 
+Players' performance in fantasy football, and human behavior more generally, is complex and multiply determined.
+No one can predict accurately how all players will perform.
+Remain humble in your predictions; do not be more confident than you should be.
+
 ### Confirmation Bias {#sec-cognitiveBiasesConfirmation}
@@
 For instance, if you believe that a particular player is a strong breakout candidate to be a sleeper, you may be more likely to pay attention to evidence that supports that the player will breakout and may be less likely to pay attention to evidence that indicates the player may struggle.
+
+As a budding empiricist, you should actively seek out information that challenges or disconfirms your beliefs and work to incorporate it into your beliefs.
+Do your best to go into observation, data analysis, and data interpretation with an open mind.
 
@@
 
+*Loss aversion bias* is the tendency to avoid losses rather than acquiring equivalent gains.
+Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
+In fantasy football, loss aversion may lead managers to start or hold onto underperforming high drafts for too long instead of a starting a more promising player out of fear of losing potential value from their initial investment.
+
 ### Endowment Bias {#sec-cognitiveBiasesEndowment}
 
+*Endowment bias* is the tendency to overvalue merely because one owns it.
+For instance, a manager might overvalue a player they drafted in the first round, refusing to trade them even if they could get a better-performing player in return.
+
 ### Bandwagon Effect Bias {#sec-cognitiveBiasesBandwagon}
+
+The *bandwagon effect bias* is the tendency to do or believe things because other people are.
+It involves social conformity.
+For instance, consider if a rookie Wide Receiver has a breakout game and he is picked up in many fantasy leagues.
+A given manager might pick up the player because the player is frequently being picked up in many fantasy leagues, without evaluating whether the player's success is sustainable.
 
@@
 > "The more you know, the more you know you don't know." – Anonymous
+
+The *Dunning–Kruger effect bias* is the tendency for people with low ability/competency in a task to overestimate their ability.
+The Dunning–Kruger effect is depicted in @fig-dunningKrueger1 and @fig-dunningKrueger2.
+For instance, consider a new fantasy manager who experiences some initial wins (often called "beginner's luck").
+They may attribute their successes to their skill rather than to luck.
+Their [overconfidence](#sec-cognitiveBiasesOverconfidence) may lead them to believe they can win the league without much preparation.
+
+::: {#fig-dunningKrueger1}
+![](images/dunningKrueger_1.png)
+
+Dunning–Krueger Effect. Adapted from <https://commons.wikimedia.org/wiki/File:Effet_Dunning-Kruger.svg>.
+:::
+
+::: {#fig-dunningKrueger2}
+![](images/dunningKrueger_2.png)
+
+Dunning–Krueger Effect. Adapted from <https://commons.wikimedia.org/wiki/File:Dunning-kruger_effect_-_percentile.svg>.
+:::
 

### 2024-07-11 cognitive-bias.qmd
29249710 - figure labels
20240711 - conclusion

@@
 
+As described [above](#sec-cognitiveBiasOverview), heuristics are mental shortcuts that people use to handle the overwhelming amount of information to process.
 Three important heuristics used in judgment and prediction in the face of uncertainty include [@Tversky1974]:
@@
 
+As described [above](#sec-cognitiveBiasOverview), cognitive biases are systematic errors in thinking.
+Cognitive biases are often due to the use of [heuristics](#sec-heuristics).
 Examples of cognitive biases that result from one or more of these heuristics include:
@@
 No one can predict accurately how all players will perform.
-Remain humble in your predictions; do not be more confident than you should be.
+Remain humble in your predictions; do not be more confident than is warranted.
+If you approach the task of prediction with humility, you may be more able to be flexible and more willing to consider other players who you can draft for good value.
 
@@
 
-Dunning–Krueger Effect. Adapted from <https://commons.wikimedia.org/wiki/File:Effet_Dunning-Kruger.svg>.
+Dunning–Krueger Effect: Confidence as a Function of Competence. Adapted from <https://commons.wikimedia.org/wiki/File:Effet_Dunning-Kruger.svg>.
 :::
@@
 
-Dunning–Krueger Effect. Adapted from <https://commons.wikimedia.org/wiki/File:Dunning-kruger_effect_-_percentile.svg>.
+Dunning–Krueger Effect: Perceived Performance as a Function of Actual Performance. Adapted from <https://commons.wikimedia.org/wiki/File:Dunning-kruger_effect_-_percentile.svg>.
 :::
@@
 
-Fallacies are mistaken beliefs that are often due to heuristics and cognitive biases.
+As described [above](#sec-cognitiveBiasOverview), fallacies are mistaken beliefs and flawed reasoning.
+Fallacies are often due to the use of [heuristics](#sec-heuristics) and to [cognitive biases](#sec-cognitiveBiases).
 Examples of fallacies include:
@@
 
+We discuss ways to counteract the base rate fallacy in @sec-baseRates.
+
 ### Regression Fallacy {#sec-fallaciesRegression}
@@
 
-### Sunk Cost Fallacy {#sec-fallaciesSunkCost}
-
-A *sunk cost* is a cost (e.g., in money, time, or effort) that has already been incurred and cannot be recovered.
-For instance, if a person orders an expensive meal at a restaurant, the order is a sunk cost.
-The *sunk cost fallacy* is the tendency to continue an endeavor when there is a sunk cost.
-For instance, when ordering the expensive meal at the restaurant, a person may over-eat so that they feel that they eat their money's worth of food.
-
-Applied to fantasy football, consider a situation in which you invest a lot of salary cap or a high draft pick to draft a promising player, but they repeatedly underperform.
-If you continue to start the player to justify your large investment, instead of benching him in favor of a higher-performing player, you are committing the sunk cost fallacy.
-
 ### Hot Hand Fallacy {#sec-fallaciesHotHand}
@@
 Due to the hot hand fallacy, a manager might continue to start the player because they believe the player is "on fire" and is likely to continue to score at an unsustainable rate.
+
+It is important consider whether such a string of strong performances are outliers and if the player may, in future games, [regress to the mean](#sec-fallaciesRegression).
+When considering whether strong performances are outliers and may [regress to the mean](#sec-fallaciesRegression), it is valuable to consider whether the player's health, skill, or situation has appreciably changed (compared to the player's earlier, weaker performances).
+Is the player finally fully healthy?
+Has the player appreciably improved in some skill that will benefit them in future games?
+Has the player's long-term situation improved, such as moving up the depth chart, or receiving more carries/targets that is not tied to a specific opponent or game script?
+Or, alternatively, do the improvements appear to be driven by transient, game-specific factors, such as a the health of a teammate, the opponents they played, or the game script that ensued?
+If long-term outlook of the player has appreciably changed due to changes in the [fundamentals of a player's value](#sec-playerEvaluation), such as their [health](#sec-evalHealth), [skill](#sec-skill), or [situation](#sec-evalSituation), it is less likely that such performance improvements will [regress](#sec-fallaciesRegression) over the long run.
+
+### Sunk Cost Fallacy {#sec-fallaciesSunkCost}
+
+A *sunk cost* is a cost (e.g., in money, time, or effort) that has already been incurred and cannot be recovered.
+For instance, if a person orders an expensive meal at a restaurant, the order is a sunk cost.
+The *sunk cost fallacy* is the tendency to continue an endeavor when there is a sunk cost.
+For instance, when ordering the expensive meal at the restaurant, a person may over-eat so that they feel that they eat their money's worth of food.
+
+Applied to fantasy football, consider a situation in which you invest a lot of salary cap or a high draft pick to draft a promising player, but they repeatedly underperform.
+If you continue to start the player to justify your large investment, instead of benching him in favor of a higher-performing player, you are committing the sunk cost fallacy.
 
@@
 
+In conclusion, there are many [heuristics](#sec-heuristics), [cognitive bias](#sec-cognitiveBiases), and [fallacies](#sec-fallacies) that people engage in when making judgments and predictions.
+It is prudent to be aware of these common biases and to work to counteract them.
+For instance, look for information that challenges or disconfirms your beliefs, and work to incorporate this information into your beliefs.
+Do your best to pursue observation, data analysis, and data interpretation with an open mind.
+You never know what important information you might discover if you go in with an open mind.
+Pay attention to [fundamentals of a player's value](#sec-playerEvaluation), such as their [health](#sec-evalHealth), [skill](#sec-skill), or [situation](#sec-evalSituation) when considering whether a player's performance may [regress to the mean](#sec-fallaciesRegression).
+If the strong performances appear to be driven by transient, game-specific factors, such as a the health of a teammate, the opponents they played, or the game script that ensued, future performances may be more likely to [regress to the mean](#sec-fallaciesRegression).
+In general, people tend to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their predictions.
+Approach the task of prediction with humility; no one is consistently able to accurately predict how well players will perform.
+
 ::: {.content-visible when-format="html"}

### 2024-07-11 player-evaluation.qmd
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240711 - conclusion

@@
 There are also calculators that integrate information about body shape and information from the NFL Combine to determine a player's relative athletic score (RAS) for their position: <https://ras.football/ras-calculator/>
+
+## Skill {#sec-skill}
+
+When scouting players, scouts consider not only the player's [athletic profile](#sec-evalAthleticProfile), but also their position-relevant skill.
+For instance, how good are they are reading the defense, passing the ball, running routes, catching balls, making defenders miss tackles, taking care of the ball, consistency, etc.
+Scouting and evaluating skill is a complicated endeavor, and even the professional scouts frequently make mistakes in their evaluations and predictions.
+You can certainly read skill evaluations about various players; however, unlike metrics of athletic profile, we do not have direct access to the player's underlying skill.
+Some may say, "You know it when you see it."
+But, this is not particularly useful when trying to identify players who are undervalued or overvalued—because the skill evaluations are likely already "baked into" a player's projections.
+Because we do not have direct access to a player's skill, we tend to rely on indirect metrics of their ability, such as [historical performance](#sec-evalHistoricalPerformance).
 
@@
 By contrast, if you are predicted to lose to a team by a good margin, it may make sense to start the players with the highest upside.
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-playerEvaluationSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-11 causal-inference.qmd
20240711 - multiple figures

@@
 
-One of the plausible alternatives to the inference that `X` causes `Y` is that there are third variable [confounds](#sec-causalDiagramConfounding) that influence both `X` and `Y`, thus explaining why `X` and `Y` are associated, as depicted in @fig-correlationAndCausation3 and @fig-ZCausesXandY.
+One of the plausible alternatives to the inference that `X` causes `Y` is that there are third variable [confounds](#sec-causalDiagramConfounding) that influence both `X` and `Y`, thus explaining why `X` and `Y` are associated, as depicted in Figures [-@fig-correlationAndCausation3] and [-@fig-ZCausesXandY].
 Thus, another approach that can help increase [internal validity](#sec-internalValidity) is to include plausible [confounds](#sec-causalDiagramConfounding) as control variables.
@@
 To create DAGs, you can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: <https://dagitty.net> (archived at <https://perma.cc/U9BY-VZE2>).
-Examples of various causal diagrams that could explain why `X` is associated with `Y` are in @fig-XCausesY, @fig-YCausesX, and @fig-ZCausesXandY.
+Examples of various causal diagrams that could explain why `X` is associated with `Y` are in Figures [-@fig-XCausesY], [-@fig-YCausesX] and [-@fig-ZCausesXandY].
 
@@
 That is, a collider is a variable that is caused by two other variables (i.e., a common effect).
-An example collision is depicted in @fig-colliderBias1 and @fig-colliderBias2:
+An example collision is depicted in Figures [-@fig-colliderBias1] and [-@fig-colliderBias2]:
 

### 2024-07-11 research-methods.qmd
20240711 - multiple figures

@@
 
-For example, if the effect of consuming sports drinks on player performance differs for Quarterbacks and Wide Receivers, the interaction could be depicted in @fig-moderationExamplePathDiagram and @fig-moderationExampleInteractionGraph:
+For example, if the effect of consuming sports drinks on player performance differs for Quarterbacks and Wide Receivers, the interaction could be depicted in Figures [-@fig-moderationExamplePathDiagram] and [-@fig-moderationExampleInteractionGraph]:
 

### 2024-07-11 cognitive-bias.qmd
20240711 - multiple figures
20240711 - luck

@@
 
-Players' performance in fantasy football, and human behavior more generally, is complex and multiply determined.
-No one can predict accurately how all players will perform.
+Players' performance in fantasy football, and human behavior more generally, is complex and multiply determined (i.e., is influenced by many factors).
+Despite the bluster of so-called experts who pretend to know more than they can know, no one can consistently and accurately predict how all players will perform.
 Remain humble in your predictions; do not be more confident than is warranted.
@@
 The *Dunning–Kruger effect bias* is the tendency for people with low ability/competency in a task to overestimate their ability.
-The Dunning–Kruger effect is depicted in @fig-dunningKrueger1 and @fig-dunningKrueger2.
+The Dunning–Kruger effect is depicted in Figures [-@fig-dunningKrueger1] and [-@fig-dunningKrueger2].
 For instance, consider a new fantasy manager who experiences some initial wins (often called "beginner's luck").
@@
 In general, people tend to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their predictions.
+There is a lot of luck in fantasy football.
 Approach the task of prediction with humility; no one is consistently able to accurately predict how well players will perform.

### 2024-07-12 actuarial.qmd
20240711 - actuarial

@@
 
-::: {.content-visible when-format="html"}
+## Approaches to Prediction {#sec-approachesToPrediction}
 
-## Best Actuarial Approaches to Prediction
+There are two primary approaches to prediction: human judgment and the actuarial (i.e., statistical) method.
 
-The best actuarial models tend relatively simple (parsimonious), that can account for one or several of the most important predictors and their optimal weightings, and that account for the base rate of the phenomenon.
+### Human Judgment {#sec-humanJudgment}
+
+Using the judgment method of prediction, all gathered information is collected and formulated into a prediction in the person's mind.
+The person selects, measures, and combines information and produces projections solely according to their experience and judgment.
+
+### Actuarial/Statistical Method {#sec-actuarialPrediction}
+
+In the actuarial or statistical method of prediction, information is gathered and combined systematically in an evidence-based statistical prediction formula.
+The method is based on equations and data, so both are needed.
+
+An example of a statistical method of prediction is the Violence Risk Appraisal Guide [@Rice2013].
+The Violence Risk Appraisal Guide is used in an attempt to predict violence and is used for parole decisions.
+For instance, the equation might be something like @eq-actuarial:
+
+$$
+\text{Violence risk} = \beta \cdot \text{conduct disorder} + \beta \cdot \text{substance use} + \beta \cdot \text{suspended from school} + \beta \cdot \text{childhood aggression} + ...
+$$ {#eq-actuarial}
+
+Then, based on their score and the established cutoffs, a person is given a "low risk", "medium risk", or "high risk" designation.
+
+An actuarial formula for projecting a Running Backs rushing yards might be something like @eq-actuarialFantasy:
+
+$$
+\text{rushing yards} = \beta \cdot \text{rushing yards last season} + \beta \cdot \text{age} + \beta \cdot \text{injury history} + \beta \cdot \text{strength of offensive line} + ...
+$$ {#eq-actuarialFantasy}
+
+### Combining Human Judgment and Statistical Algorithms {#sec-combiningHumanJudgmentActuarialPrediction}
+
+There are numerous ways in which humans and statistical algorithms could be involved.
+On one extreme, humans make all judgments.
+On the other extreme, although humans may be involved in data collection, a statistical formula makes all decisions based on the input data, consistent with an actuarial approach.
+However, the human judgment and actuarial approaches can be combined in a hybrid way [@Dana2006].
+For example, to save time and money, a clinical psychologist might use an actuarial approach in all cases, but might only use a judgment approach when the actuarial approach gives a "positive" test.
+Or, the clinical psychologist might use both human judgment and an actuarial approach independently to see whether they agree.
+That is, the clinician may make a prediction based on their judgment and might also generate a prediction from an actuarial approach.
+
+The challenge is what to do when the human and the algorithm disagree.
+Hypothetically, humans reviewing and adjusting the results from the statistical algorithm could lead to more accurate prediction.
+However, human input also could lead to the possibility or exacerbation of biased predictions.
+In general, with very few exceptions, actuarial approaches are as accurate or more accurate than "expert" judgment [@Dawes1989; @Aegisdottir2006; @Baird2000; @DenHartigh2018; @Grove1996; @Grove2000].
+Moreover, the superiority of actuarial approaches to human judgment tends to hold even when the expert is given more information than the actuarial approach [@Dawes1989].
+Allowing experts to override actuarial predictions consistently leads to lower predictive accuracy [@Garb2019].
+
+There is sometimes a misconception that formulas cannot account for qualitative information.
+However, that is not true.
+Qualitative information can be scored or coded to be quantified so that it can be included in statistical formulas.
+That said, the quality of predictions rests on the quality and relevance of the assessment information for the particular prediction decision.
+If the assessment data are lousy, it is unlikely that a statistical algorithm (or a human for that matter) will make an accurate prediction:
+"Garbage in, garbage out".
+A statistical formula cannot rescue inaccurate assessment data.
+
+## Errors in Human Judgment {#sec-errorsInHumanJudgment}
+
+Human judgment is naturally subject to errors.
+Below, I describe a few errors to which human judgment seems particularly prone.
+
+When operating freely, clinicians tend to over-estimate exceptions to the established rules (i.e., the broken leg syndrome).
+@Meehl1957 acknowledged that there may be some situations where it is glaringly obvious that the statistical formula would be incorrect because it fails to account for an important factor.
+He called these special cases "broken leg" cases, in which the human should deviate from the formula (i.e., broken leg countervailing).
+The example goes like this:
+
+> "If a sociologist were predicting whether Professor X would go to the movies on a certain night, he might have an equation involving age, academic specialty, and introversion score.
+> The equation might yield a probability of .90 that Professor X goes to the movie tonight.
+> But if the family doctor announced that Professor X had just broken his leg, no sensible sociologist would stick with the equation.
+> Why didn't the factor of 'broken leg' appear in the formula?
+> Because broken legs are very rare, and in the sociologist's entire sample of 500 criterion cases plus 250 cross-validating cases, he did not come upon a single instance of it.
+> He uses the broken leg datum confidently, because 'broken leg' is a subclass of a larger class we may crudely denote as 'relatively immobilizing illness or injury,' and movie-attending is a subclass of a larger class of 'actions requiring moderate mobility.' "
+> [@Meehl1957, p. 269–270]
+
+However, people too often think that cases where they disagree with the statistical algorithm are broken leg cases.
+People too often think their case is an exception to the rule.
+As a result, they too often change the result of the statistical algorithm and are more likely to be wrong than right in doing so.
+Because actuarial methods are based on actual population levels (i.e., [base rates](#sec-baseRates)), unique exceptions are not over-estimated.
+
+Actuarial predictions are perfectly [reliable](#sec-reliability)—they will always return the same conclusion given an identical set of data.
+The human judge is likely to both disagree with others and with themselves given the same set of symptoms.
+
+The decision by an expert (all by all humans) is likely to be influenced by past experiences.
+Actuarial methods are based on objective algorithms, and past personal experience and personal biases do not factor into any decisions.
+Humans give weight to less relevant information, and often give too much weight to singular variables.
+Actuarial formulas do a better job of focusing on relevant variables.
+Computers are good at factoring in [base rates](#sec-baseRates).
+Humans ignore [base rates](#sec-baseRates) ([base rate neglect](#sec-fallaciesBaseRate)).
+
+Computers are better at accurately weighing predictors and calculating unbiased risk estimates.
+In an actuarial formula, the relevant predictors are weighted according to their predictive power.
+
+Humans are typically given no feedback on their judgments.
+To improve accuracy of judgments, it is important for feedback to be clear, consistent, and timely.
+
+## Humans Versus Computers {#sec-humansVsComputers}
+
+### Advantages of Computers {#sec-advantagesOfComputers}
+
+Here are some advantages of computers over humans:
+
+- Computers can process lots of information simultaneously.
+So can humans.
+But computers can to an even greater degree.
+- Computers are faster at making calculations.
+- Computations by computers are error-free (as long as the computations are programmed correctly).
+- Computers' judgments will not be biased by fatigue or emotional responses.
+- Computers' judgments will tend not to be biased in the way that humans' [cognitive biases](#sec-cognitiveBias) are.
+Computers are less likely to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their judgments.
+- Computers can more accurately weight the set of predictors based on large data sets.
+Humans tend to give too much weight to singular predictors.
+
+### Advantages of Humans {#sec-advantagesOfHumans}
+
+Computers are bad at some things too.
+Here are some advantages of humans over computers (as of now):
+
+- Humans can be better at identifying patterns in data (but also can mistakenly identify patterns where there are none—i.e., illusory correlation).
+- Humans can be flexible and take a different approach if a given approach is not working.
+- Humans are better at tasks requiring creativity and imagination, such as developing theories that explain phenomena.
+- Humans have the ability to reason, which is especially important when dealing with complex, abstract, or open-ended problems, or problems that have not been faced before (or for which we have insufficient data).
+- Humans are better able to learn.
+- Humans are better at holistic, gestalt processing, including facial and linguistic processing.
+
+There *may* be situations in which a human judgment would do better than an actuarial judgment.
+One situation where human judgment would be important is when no actuarial method exists for the judgment or prediction.
+For instance, when no actuarial method exists for the diagnosis or disorder (e.g., suicide), it is up to the clinician.
+However, we could collect data on the outcomes or on clinicians' judgments to develop an actuarial method that will be more reliable than the clinicians' judgments.
+That is, an actuarial method developed based on clinicians' judgments will be more accurate than clinicians' judgments.
+In other words, we do not necessarily need outcome data to develop an actuarial method.
+We could use the client's data as predictors of the clinicians' judgments to develop a structured approach to prediction that weighs factors similarly to clinicians, but with more [reliable](#sec-reliability) predictions.
+
+Another situation in which human judgment could outperform a statistical algorithm is in true "broken leg" cases, e.g., important and rare events (edge cases) that are not yet accounted for by the algorithm.
+
+Another situation in which human judgment could be preferable is if advanced, complex theories exist.
+Computers have a difficult time adhering to complex theories, so clinicians may be better suited.
+However, we do not have any of these complex theories in psychology that are accurate.
+We would need strong theory informed by data regarding causal influences, and [accurate](#validity) measures to assess them.
+However, no theories in psychology are that good.
+Nevertheless, predictive accuracy can be improved when considering theory [@Silver2012; @Garb2019].
+
+If the prediction requires complex configural relations that a computer will have a difficult time replicating, a clinician's judgment may be preferred.
+Although the likelihood that a person can accurately work through these complex relations is theoretically possible, it is highly unlikely.
+Holistic pattern recognition (such as language and faces) tends to be better by humans than computers.
+But computers are getting better with holistic pattern recognition through machine learning.
+
+In sum, the human seeks to integrate information to make a decision, but is biased.
+
+### Comparison of Evidence {#sec-evidenceOnHumanJudgmentVsActuarialPrediction}
+
+Hundreds of studies have examined clinical versus actuarial prediction methods across many disciplines.
+Findings consistently show that actuarial methods are as [accurate](#sec-validity) or more [accurate](#sec-validity) than human judgment/prediction methods.
+"There is no controversy in social science that shows such a large body of qualitatively diverse studies coming out so uniformly...as this one" [@Meehl1986, pp. 373–374].
+
+Actuarial methods are particularly valuable for criterion-referenced assessment tasks, in which the aim is to predict specific events or outcomes [@Garb2019].
+For instance, actuarial methods have shown promise in predicting violence, criminal recidivism, psychosis onset, course of mental disorders, treatment selection, treatment failure, suicide attempts, and suicide [@Garb2019].
+
+Moreover, actuarial methods are explicit; they can be transparent and lead to informed scientific criticism to improve them.
+By contrast, human judgment methods are not typically transparent; human judgment relies on mental processes that are often difficult to specify.
+
+## Why Judgment is More Widely Used Than Statistical Formulas {#sec-whyHumanJudgmentIsMoreWidelyUsed}
+
+Despite actuarial methods being generally more accurate than human judgment, judgment is much more widely used by clinicians.
+There are several reasons why actuarial methods have not caught on; one reason is professional traditions.
+Experts in any field do not like to think that a computer could outperform them.
+Some practitioners argue that judgment/prediction is an "art form" and that using a statistical formula is treating people like a number.
+However, using an approach (i.e., human judgment) that systematically leads to less [accurate](#sec-validity) decisions and predictions is an ethical problem.
+
+Some clinicians do not think that group averages (e.g., in terms of which treatment is most effective) apply to an individual client.
+This invokes the distinction between nomothetic (group-level) inferences and idiographic (individual-level) inferences.
+However, the scientific evidence and probability theory strongly indicate that it is better to generalize from group-level evidence than throwing out all the evidence and taking the approach of "anything goes."
+Clinicians frequently believe the broken leg fallacy, i.e., thinking that your client is an exception to the algorithmic prediction.
+In most cases, deviating from the statistical formula will result in less [accurate](#sec-validity) predictions.
+People tend to overestimate the probability of low [base rate](#sec-baseRates) conditions and events.
+
+Another reason why actuarial methods have not caught on is the belief that receiving a treatment is the only thing that matters.
+But it is an empirical question which treatment is most effective for whom.
+What if we could do better?
+For example, we could potentially use a formula to identify the most effective treatment for a client.
+Some treatments are no better than placebo; other treatments are actually harmful [@Lilienfeld2007; @Williams2021].
+
+Another reason why judgment methods are more widely used than actuarial methods is that so-called "experts" (and people in general) show [overconfidence](#sec-cognitiveBiasesOverconfidence) in their predictions—clinicians, experts, and humans in general think they are more accurate than they actually are.
+We see this when examining their calibration; their predictions tend to be miscalibrated.
+For example, things they report with 80% confidence occur less than 80% of the time, an example of [overprecision](#sec-cognitiveBiasesOverconfidence) in their predictions.
+Humans will sometimes be correct by chance, and they tend to mis-attribute that to their skill; humans tend to remember the successes and forget the failures.
+
+Another argument against using actuarial methods is that "no methods exist".
+In some cases, that is true—actuarial methods do not yet exist for some prediction problems.
+However, one can always create an algorithm of the experts' judgments, even if one does not have access to the outcome information.
+A model of clinicians' responses tends to be more [accurate](#sec-validity) than clinicians' judgments themselves because the model gives the same outcome with the same input data—i.e., it is perfectly [reliable](#sec-reliability).
+
+Another argument from some clinicians is that, "My job is to understand, not to predict".
+But what kind of understanding does not involve predictions?
+Accurate predictions help in understanding.
+Knowing how people would perform in different conditions is the same thing as good understanding.
+
+## Best Actuarial Approaches to Prediction {#sec-bestActuarialApproaches}
+
+The best actuarial models tend to be relatively simple (parsimonious), that can account for one or several of the most important predictors and their optimal weightings, and that account for the base rate of the phenomenon.
 Even unit-weighted formulas (formulas whose [predictor variables](#sec-correlationalStudy) are equally weighted with a weight of one) can sometimes generalize better to other samples than complex weightings [@Garb2019].
@@
 
+## Conclusion {#sec-conclusion-actuarial}
+
+In general, it is better to develop and use structured, actuarial approaches than informal approaches that rely on human judgment or judgment by "so-called" experts.
+Actuarial approaches to prediction tend to be as accurate or more accurate than expert judgment.
+Nevertheless, in many domains, human judgment tends to be much more widely used than actuarial approaches.
+
+::: {.content-visible when-format="html"}
+
 ## Session Info {#sec-judgmentVsActuarialSessionInfo}

### 2024-07-12 player-evaluation.qmd
20240711 - actuarial

@@
 
-Many aspects of a player's athletic profile are available from the National Football League (NFL) Combine, which is especially relevant for evaluating rookies.
+Many aspects of a player's athletic profile, including tests of speed (40-yard dash), strength (bench press), agility (2-yard shuttle run; three cone drill), and jumping ability (vertical jump; broad jump) are available from the National Football League (NFL) Combine, which is especially relevant for evaluating rookies.
 We demonstrate how to import data from the NFL Combine in @sec-combine.
@@
 
-It is important to consider a player's past performance.
+Compared to tests of speed, power, and agility at the NFL Combine, collegiate performance is a stronger predictor of performance in the NFL [@Lyons2011].
+That is, previous sports performance is the best predictor of future performance [for a review, see @DenHartigh2018].
+Thus, it is important to consider a player's past performance.
 However, the extent to which historical performance may predict future performance may depend on many factors such as (a) the similarity of the prior situation to the current situation, (b) how long ago the prior situation was, and (c) the extent to which the player (or situation) has changed in the interim.
@@
 An indicator that intends to be prognostic of future performance, and that accounts for past performance, is a player's draft position—that is, how early (or late) was a player selected in the NFL Draft.
-The earlier a player was selected in the NFL Draft, the greater likelihood that the player will perform well.
+The earlier a player was selected in the NFL Draft, the greater likelihood that the player will perform well; however, this is somewhat countered by the fact that the teams with the highest draft picks tend to be the worst based on the prior season's record.
 

### 2024-07-12 research-methods.qmd
20240711 - actuarial

@@
 - [incremental validity](#sec-incrementalValidity)
+- [ecological validity](#sec-ecologicalValidity)
 
@@
 
-Face validity evaluates the extent to which a measure "looks like" (on its face) it assesses the construct of interest.
+*Face validity* evaluates the extent to which a measure "looks like" (on its face) it assesses the construct of interest.
 For instance, if a measure is developed to assess aptitude of Wide Receivers for the position, it would be considered to have face validity if everyday (lay) people believe that it assesses aptitude for being a successful Wide Receiver.
@@
 
-Content validity evaluates the extent to which the measure assesses the full breadth of the content, as determined by context experts.
+*Content validity* evaluates the extent to which the measure assesses the full breadth of the content, as determined by context experts.
 For the measure to have content validity, it should not have gaps (missing content facets) or intrusions (facets of other constructs).
@@
 
-Criterion-related validity evaluates the extent to which the measure's scores are related to meaningful variables of interest.
+*Criterion-related validity* evaluates the extent to which the measure's scores are related to meaningful variables of interest.
 Criterion-related validity is commonly evaluated using a Pearson correlation or some form of regression.
@@
 
-Concurrent criterion-related validity evaluates the extent to which the measure's scores are related to meaningful variables of interest assessed at the same point in time.
+*Concurrent criterion-related validity* (aka concurrent validity) evaluates the extent to which the measure's scores are related to meaningful variables of interest assessed at the same point in time.
 That is, concurrent validity could evaluate whether current player statistics (e.g., passing yards) are associated with their fantasy points.
@@
 
-Predictive criterion-related validity evaluates the extent to which the measure's scores are related to meaningful variables of interest that are assessed at a later point in time.
+*Predictive criterion-related validity* (aks predictive validity) evaluates the extent to which the measure's scores are related to meaningful variables of interest that are assessed at a later point in time.
 For example, predictive validity could evaluate whether scores on the measure we developed to assess a player's aptitude to succeed in the NFL predicts later performance in the NFL.
@@
 
-Construct validity evaluates the extent to which the measure's scores accurately assess the construct of interest.
+*Construct validity* evaluates the extent to which the measure's scores accurately assess the construct of interest.
 If we develop a measure with intent to assess aptitude for being a successful Running Back, and it appears to more accurately assess aptitude for being a successful Wide Receiver, then our measure has poor construct validity for assessing aptitude for being a successful Running Back.
@@
 
-Convergent validity evaluates the extent to which the measure's scores are related to other measures of the same construct.
+*Convergent validity* evaluates the extent to which the measure's scores are related to other measures of the same construct.
 For instance, if we develop a new measure to assess intelligence, its scores should be related to scores from other measures designed to assess intelligence (e.g., Wonderlic Contemporary Cognitive Ability Test).
@@
 
-Discriminant validity evaluates the extent to which the measure's scores are unrelated to measures of the different constructs.
+*Discriminant validity* evaluates the extent to which the measure's scores are unrelated to measures of the different constructs.
 For instance, if we develop a new measure to assess intelligence, its scores should be less strongly associated with measures of other constructs (e.g., measures of happiness).
@@
 
-Incremental validity evalutes the extent to which the measure's scores provide an increase in predictive accuracy compared to other information that is easily and cheaply available.
+*Incremental validity* evalutes the extent to which the measure's scores provide an increase in predictive accuracy compared to other information that is easily and cheaply available.
 That is, in order to be useful, a strong measure should tell us something that we did not already know.
@@
 ```
+
+#### Ecological Validity {#sec-ecologicalValidity}
+
+*Ecological validity* evaluates the extent to which the measures' scores are indicative of the behavior of a person in the natural environment.
+For instance, measures of a players' speed during a game has higher ecological validity (and is more predictive of their performance) than their speed during the NFL Combine [@Lyons2011].
+For instance, compared to tests of speed, power, and agility at the NFL Combine, collegiate performance is a stronger predictor of performance in the NFL [@Lyons2011].
+That is, previous sports performance is the best predictor of future performance [for a review, see @DenHartigh2018].
 

### 2024-07-12 machine-learning.qmd
Built site for gh-pages
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240703 - conclusion section

--- 2024-06-05 base-rates.qmd
@@
-# Base Rates {#sec-baseRates}
+# Machine Learning {#sec-machineLearning}
 
-## Getting Started
+## Getting Started {#sec-machineLearningGettingStarted}
 
-### Load Packages
+### Load Packages {#sec-machineLearningLoadPackages}
 
@@
 ```
+
+## Fitting the Traditional Regression Models {#sec-fittingModels-regression}
+
+### Regression with One Predictor {#sec-regressionOnePredictor}
+
+### Regression with Multiple Predictors {#sec-regressionMultiplePredictors}
+
+## Fitting the Machine Learning Models {#sec-fittingModels-machineLearning}
+
+### Least Absolute Shrinkage and Selection Option (LASSO) {#sec-lasso}
+
+### Ridge Regression {#sec-ridgeRegression}
+
+### Elastic Net {#sec-elasticNet}
+
+### Random Forest Machine Learning {#sec-randomForest}
+
+### *k*-Fold Cross-Validation {#sec-kfoldCV}
+
+### Leave-One-Out (LOO) Cross-Validation {#sec-looCV}
+
+## Conclusion {#sec-machineLearningConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-machineLearningSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-12 actuarial.qmd
20240712 - actuarial

@@
-# Human Judgment Versus Actuarial Approaches to Prediction {#sec-judgmentVsActuarial}
+# Judgment Versus Actuarial Approaches to Prediction {#sec-judgmentVsActuarial}
 
@@
 The person selects, measures, and combines information and produces projections solely according to their experience and judgment.
+For instance, a proclaimed "fantasy expert" might use their experience, expertise, and judgment to make a prediction about how each player will perform by using whatever information and data they deem to be important, aggregating all of this information in their mind to make the prediction for each player.
 
@@
 $$
+\scriptsize
 \text{Violence risk} = \beta \cdot \text{conduct disorder} + \beta \cdot \text{substance use} + \beta \cdot \text{suspended from school} + \beta \cdot \text{childhood aggression} + ...
@@
 
-An actuarial formula for projecting a Running Backs rushing yards might be something like @eq-actuarialFantasy:
+An actuarial formula for projecting a Running Back's rushing yards might be something like @eq-actuarialFantasy:
 
 $$
+\scriptsize
 \text{rushing yards} = \beta \cdot \text{rushing yards last season} + \beta \cdot \text{age} + \beta \cdot \text{injury history} + \beta \cdot \text{strength of offensive line} + ...
 $$ {#eq-actuarialFantasy}
+
+The beta weights in the actuarial model reflect the relative weight to assign each predictor.
+For instance, in predicting rushing yards, a player's historical performance is likely the strongest predictor, whereas injury history might be a relatively weaker predictor.
+Thus, we might give historical performance a beta of 3 and injury history a beta of 1 to give a player's historical performance three times more weight than the player's injury history in predicting their rushing yards.
+For generating the actuarial model, you could obtain the beta weights for each predictor from [multiple regression](#sec-multipleRegression), from [machine learning](#sec-machineLearning), or from prior research on the relative importance of each predictor.
 
@@
 However, human input also could lead to the possibility or exacerbation of biased predictions.
-In general, with very few exceptions, actuarial approaches are as accurate or more accurate than "expert" judgment [@Dawes1989; @Aegisdottir2006; @Baird2000; @DenHartigh2018; @Grove1996; @Grove2000].
+In general, with very few exceptions, actuarial approaches are as accurate or more accurate than "expert" judgment [@Dawes1989; @Aegisdottir2006; @Baird2000; @Grove1996; @Grove2000].
+This is also likely true with respect to predicting player performance in sports [@DenHartigh2018].
 Moreover, the superiority of actuarial approaches to human judgment tends to hold even when the expert is given more information than the actuarial approach [@Dawes1989].
@@
 Qualitative information can be scored or coded to be quantified so that it can be included in statistical formulas.
+For instance, if an expert scout is able to meaningfully assess a player's [cognitive and motivational factors](#sec-evalCogMotivational) (i.e., the "[X factor](#sec-evalCogMotivational)" or "[intangibles](#sec-evalCogMotivational)"), the scout can score this across multiple players and include these data in the actuarial prediction formula.
+For instance, the scout could use a rating scale (e.g., 1 = "poor"; 2 = "fair"; 3 = "good"; 4 = "very good"; 5 = "excellent") to code (i.e., translate) their qualitative judgment into a quantifiable rating that can be integrated with other information in the actuarial formula.
 That said, the quality of predictions rests on the quality and relevance of the assessment information for the particular prediction decision.

### 2024-07-12 actuarial.qmd
20240712 - gambler's fallacy

@@
 \scriptsize
-\text{Violence risk} = \beta \cdot \text{conduct disorder} + \beta \cdot \text{substance use} + \beta \cdot \text{suspended from school} + \beta \cdot \text{childhood aggression} + ...
+\text{violence risk} = \beta \cdot \text{conduct disorder} + \beta \cdot \text{substance use} + \beta \cdot \text{suspended from school} + \beta \cdot \text{childhood aggression} + ...
 $$ {#eq-actuarial}

### 2024-07-12 cognitive-bias.qmd
20240711 - copyediting
20240712 - gambler's fallacy

@@
 ```{r}
-
+library("tidyverse")
 ```
@@
 For instance, if you flip a coin 10 times, belief in the law of small numbers would lead one to believe that the coin will flip heads-up exactly 5 times out of 10.
-However, in reality, the chance is less than 1 in 4 (24.6%) that exactly 5 of 10 coin flips turn up heads:
+However, in reality, the chance is less than 1 in 4 (24.6%) that exactly 5 of 10 coin flips turn up heads, as calculated below and as depicted in @fig-simulationOf10CoinFlips:
 
@@
 )
+```
+
+The `dbinom()` function in R provides the density of a binomial distribution.
+A binomial distribution is the probability of a particular number of successes (e.g., coins flipping heads-up) given a certain number of independent trials.
+
+```{r}
+#| fig-cap: "Histogram of number of coins that flip heads-up in a simulation of 10 coin flips (with 100,000 replications)."
+#| label: fig-simulationOf10CoinFlips
+#| code-fold: true
+
+set.seed(52242)
+
+numHeads <- rbinom(
+  n = 100000,
+  size = 10,
+  prob = .5
+)
+
+simulationOfFlipping10Coins <- data.frame(
+  numHeads = numHeads
+)
+
+simulationOfFlipping10Coins <- simulationOfFlipping10Coins %>% 
+  mutate(
+    highlight = ifelse(numHeads == 5, "yes", "no")
+  )
+
+ggplot(
+  data = simulationOfFlipping10Coins,
+  mapping = aes(
+    x = numHeads,
+    fill = highlight)
+) +
+  geom_histogram(
+    color = "#000000",
+    bins = 11
+  ) +
+  scale_x_continuous(
+    breaks = 0:10
+  ) +
+  scale_fill_manual(
+    values = c(
+      "yes" = "tomato",
+      "no" = "gray")
+  ) +
+  theme_classic() +
+  theme(legend.position = "none")
+```
+
+Although 5 is the modal count of coins that flip heads-up out of 10 flips (i.e., it was more common than any other number), it is less common than the aggregate probability of flipping any number of heads besides 5.
+The probability of getting any other number of coin flips turning up heads (other than 5) is:
+
+```{r}
+dbinom(
+  x = c(0:4, 6:10),
+  size = 10,
+  prob = 0.5
+) %>% sum()
 ```
@@
 In general, people tend to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their predictions.
-There is a lot of luck in fantasy football.
+There is considerable luck in fantasy football.
 Approach the task of prediction with humility; no one is consistently able to accurately predict how well players will perform.

### 2024-07-12 cognitive-bias.qmd
20240712 - plot title and axis titles

@@
 ```{r}
-#| fig-cap: "Histogram of number of coins that flip heads-up in a simulation of 10 coin flips (with 100,000 replications)."
+#| fig-cap: "Histogram of Number of Coins that Flip Heads-Up in a Simulation of 10 Coin Flips (with 100,000 Replications)."
 #| label: fig-simulationOf10CoinFlips
@@
 
-ggplot(
+ggplot2::ggplot(
   data = simulationOfFlipping10Coins,
@@
       "no" = "gray")
+  ) +
+  labs(
+    x = "Number of Coins Flipped Heads (out of 10 Coin Flips)",
+    y = "Frequency",
+    title = "Histogram of Number of Coins that Flip Heads-Up\nin a Simulation of 10 Coin Flips\n(with 100,000 Replications)."
   ) +

### 2024-07-13 base-rates.qmd
20240712 - base rates

@@
 
-## Overview {#sec-baseRatesConclusion}
+## Overview {#sec-baseRatesOverview}
+
+Player performance is a complex prediction task.
+Performance is probabilistically influenced by many processes, including processes internal to the player in addition to external processes.
+Moreover, people's performance occurs in the context of a dynamic system with nonlinear, probabilistic, and cascading influences that change across time.
+The ever-changing system makes behavior challenging to predict.
+And, similar to chaos theory, one small change in the system can lead to large differences later on.
+
+Let's consider a prediction example, assuming the following probabilities:
+
+- The probability of contracting HIV is .3%
+- The probability of a positive test for HIV is 1%
+- The probability of a positive test if you have HIV is 95%
+
+What is the probability of HIV if you have a positive test?
+
+As we will see, the probability is: $\frac{95\% \times .3\%}{1\%} = 28.5\%$.
+So based on the above probabilities, if you have a positive test, the probability that you have HIV is 28.5%.
+Most people tend to vastly overestimate the likelihood that the person has HIV in this example.
+Why?
+Because they do not pay enough attention to the base rate (in this example, the base rate of HIV is .3%).
+
+## Issues Around Probability {#sec-probability}
+
+### Types of Probabilities {#sec-probabilityTypes}
+
+It is important to distinguish between different types of probabilities: marginal probabilities, joint probabilities, and conditional probabilities.
+
+#### Base Rate (Marginal Probability) {#sec-baseRate}
 
@@
 In the U.S., [the prevalence rate of HIV is ~0.4% of the adult population](https://map.aidsvu.org/profiles/nation/usa/overview) (archived at <https://perma.cc/8GE6-GAPC>).
+
+For instance, we can consider the following marginal probabilities:
+
+$P(C_i)$ is the probability (i.e., base rate) of a classification, $C$, independent of other things.
+A base rate is often used as the "*prior probability*" in a Bayesian model.
+In our example above, $P(C_i)$ is the base rate (i.e., prevalence) of HIV in the population: $P(\text{HIV}) = .3\%$.
+$P(R_i)$ is the probability (base rate) of a response, $R$, independent of other things.
+In the example above, $P(R_i)$ is the base rate of a positive test for HIV: $P(\text{positive test}) = 1\%$.
+The base rate of a positive test is known as the *positivity rate* or *selection ratio*.
+
+#### Joint Probability {#sec-jointProbability}
+
+A *joint probability* is the probability of two (or more) events occurring simultaneously.
+For instance, the probability of events $A$ and $B$ both occurring together is $P(A, B)$.
+A joint probability can be calculated using the [marginal probability](#sec-baseRate) of each event, as in @eq-jointProbability:
+
+$$
+P(A, B) = P(A) \cdot P(B)
+$$ {#eq-jointProbability}
+
+Conversely (and rearranging the terms for the calculation of [conditional probability](#sec-conditionalProbability)), a [joint probability](#sec-jointProbability) can also be calculated using the [conditional probability](#sec-conditionalProbability) and [marginal probability](#sec-baseRate), as in @jointProbability2:
+
+$$
+P(A, B) = P(A | B) \cdot P(B)
+$$ {#eq-jointProbability2}
+
+#### Conditional Probability {#sec-conditionalProbability}
+
+A *conditional probability* is the probability of one event occurring given the occurrence of another event.
+Conditional probabilities are written as: $P(A | B)$.
+This is read as the probability that event $A$ occurs given that event $B$ occurred.
+For instance, we can consider the following conditional probabilities:
+
+$P(C | R)$ is the probability of a classification, $C$, given a response, $R$.
+In other words, $P(C | R)$ is the probability of having HIV given a positive test: $P(\text{HIV} | \text{positive test})$.
+$P(R | C)$ is the probability of a response, $R$, given a classification, $C$.
+In the example above, $P(R | C)$ is the probability of having a positive test given that a person has HIV: $P(\text{positive test} | \text{HIV}) = 95\%$.
+
+A conditional probability can be calculated using the [joint probability](#sec-jointProbability) and [marginal probability](#sec-baseRate) (base rate), as in @eq-conditionalProbability:
+
+$$
+P(A, B) = P(A | B) \cdot P(B)
+$$ {#eq-conditionalProbability}
+
+### Confusion of the Inverse {#sec-inverseFallacy}
+
+A [conditional probability](#sec-conditionalProbability) is not the same thing as its reverse (or inverse) [conditional probability](#sec-conditionalProbability).
+Unless the [base rate](#sec-baseRate) of the two events ($C$ and $R$) are the same, $P(C | R) \neq P(R | C)$.
+However, people frequently make the mistake of thinking that two inverse [conditional probabilities](#sec-conditionalProbability) are the same.
+This mistake is known as the "confusion of the inverse", or the "inverse fallacy", or the "conditional probability fallacy".
+The confusion of inverse probabilities is the logical error of representative thinking that leads people to assume that the probability of $C$ given $R$ is the same as the probability of $R$ given C, even though this is not true.
+As a few examples to demonstrate the logical fallacy, if 93% of breast cancers occur in high-risk women, this does not mean that 93% of high-risk women will eventually get breast cancer.
+As another example, if 77% of car accidents take place within 15 miles of a driver's home, this does not mean that you will get in an accident 77% of times you drive within 15 miles of your home.
+
+Which car is the most frequently stolen?
+It is often the Honda Accord or Honda Civic—probably because they are among the most popular/commonly available cars.
+The probability that the car is a Honda Accord given that a car was stolen ($p(\text{Honda Accord } | \text{ Stolen})$) is what the media reports and what the police care about.
+However, that is not what buyers and car insurance companies should care about.
+Instead, they care about the probability that the car will be stolen given that it is a Honda Accord ($p(\text{Stolen } | \text{ Honda Accord})$).
+
+Applied to fantasy football, the probability that a given player will be injured given that he is a Running Back ($p(\text{Injured } | \text{ RB})$) is not the same as the probability that a given player is a Running Back given that he is injured ($p(\text{RB } | \text{ Injured})$).
+
+### Bayes' Theorem {#sec-bayesTheorem}
+
+An alternative way of calculating a [conditional probability](#sec-conditionalProbability) is using the inverse [conditional probability](#sec-conditionalProbability) (instead of the [joint probability](#sec-jointProbability)).
+This is known as Bayes' theorem.
+Bayes' theorem can help us calculate a [conditional probability](#sec-conditionalProbability) of some classification, $C$, given some response, $R$, if we know the inverse [conditional probability](#sec-conditionalProbability) and the [base rate](#sec-baseRate) (marginal probability) of each.
+Bayes' theorem is in @eq-bayes1:
+
+$$
+\begin{aligned}
+  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)}
+\end{aligned}
+$$ {#eq-bayes1}
+
+Or, equivalently (rearranging the terms):
+
+$$
+\begin{aligned}
+  \frac{P(C | R)}{P(R | C)} = \frac{P(C_i)}{P(R_i)}
+\end{aligned}
+$$ {#eq-bayes2}
+
+Or, equivalently (rearranging the terms):
+
+$$
+\begin{aligned}
+  \frac{P(C | R)}{P(C_i)} = \frac{P(R | C)}{P(R_i)}
+\end{aligned}
+$$ {#eq-bayes3}
+
+More generally, Bayes' theorem has been described as:
+
+$$
+\begin{aligned}
+  P(H | E) &= \frac{P(E | H) \cdot P(H)}{P(E)} \\
+  \text{posterior probability} &= \frac{\text{likelihood} \times \text{prior probability}}{\text{model evidence}}
+\end{aligned}
+$$ {#eq-bayes6}
+
+where $H$ is the hypothesis, and $E$ is the evidence—the new information that was not used in computing the prior probability.
+
+In Bayesian terms, the *posterior probability* is the conditional probability of one event occurring given another event—it is the updated probability after the evidence is considered.
+In this case, the posterior probability is the probability of the classification occurring ($C$) given the response ($R$).
+The *likelihood* is the inverse conditional probability—the probability of the response ($R$) occurring given the classification ($C$).
+The *prior probability* is the marginal probability of the event (i.e., the classification) occurring, before we take into account any new information.
+The *model evidence* is the marginal probability of the other event occurring—i.e., the marginal probability of seeing the evidence.
+
+Bayes' theorem provides the foundation for a paradigm of statistics called Bayesian statistics, which (unlike frequentist statistics) does not use *p*-values.
+
+In the HIV example above, we can calculate the [conditional probability](#sec-conditionalProbability) of HIV given a positive test using three terms: the [conditional probability](#sec-conditionalProbability) of a positive test given HIV (i.e., the sensitivity of the test), the [base rate](#sec-baseRate) of HIV, and the [base rate](#sec-baseRate) of a positive test for HIV.
+The [conditional probability](#sec-conditionalProbability) of HIV given a positive test is in @eq-hivExample1:
+
+$$
+\begin{aligned}
+  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)} \\
+  P(\text{HIV} | \text{positive test}) &= \frac{P(\text{positive test} | \text{HIV}) \cdot P(\text{HIV})}{P(\text{positive test})} \\
+  &= \frac{\text{sensitivity of test} \times \text{base rate of HIV}}{\text{base rate of positive test}} \\
+  &= \frac{95\% \times .3\%}{1\%} = \frac{.95 \times .003}{.01}\\
+  &= 28.5\%
+\end{aligned}
+$$ {#eq-hivExample1}
+
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.
+
+```{r}
+petersenlab::pAgivenB(pBgivenA = .95, pA = .003, pB = .01)
+```
+
+Thus, assuming the probabilities in the example above, the [conditional probability](#conditionalProbability) of having HIV if a person has a positive test is 28.5%.
+Given a positive test, chances are higher than not that the person does not have HIV.
+
+Now let's see what happens if the person tests positive a second time.
+We would revise our "[prior probability](#sec-baseRate)" for HIV from the general prevalence in the population (0.3%) to be the "posterior probability" of HIV given a first positive test (28.5%).
+This is known as *Bayesian updating*.
+We would also update the "evidence" to be the [marginal probability](#sec-baseRate) of getting a second positive test.
+
+If we do not know a [marginal probability](#sec-baseRate) (i.e., base rate) of an event (e.g., getting a second positive test), we can calculate a [marginal probability](#sec-baseRate) with the *law of total probability* using [conditional probabilities](#sec-conditionalProbability) and the [marginal probability](#sec-baseRate) of another event (e.g., having HIV).
+According to the law of total probability, the probability of getting a positive test is the probability that a person with HIV gets a positive test (i.e., sensitivity) times the base rate of HIV plus the probability that a person without HIV gets a positive test (i.e., false positive rate) times the [base rate](#sec-baseRate) of not having HIV, as in @eq-lawOfTotalProbability:
+
+$$
+\begin{aligned}
+ P(\text{not } C_i) &= 1 - P(C_i) \\
+  P(R_i) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i) \\
+  1\% &= 95\% \times .3\% + P(R | \text{not } C) \times 99.7\% \\
+\end{aligned}
+$$ {#eq-lawOfTotalProbability}
+
+In this case, we know the [marginal probability](#sec-baseRate) ($P(R_i)$), and we can use that to solve for the unknown [conditional probability](#sec-conditionalProbability) that reflects the [false positive rate](#sec-falsePositiveRate) ($P(R | \text{not } C)$), as in @eq-conditionalProbabilityRevised:
+
+$$
+\scriptsize
+\begin{aligned}
+  P(R_i) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i) && \\
+  P(R_i) - [P(R | \text{not } C) \cdot P(\text{not } C_i)] &= P(R | C) \cdot P(C_i) && \text{Move } P(R | \text{not } C) \text{ to the left side} \\
+  - [P(R | \text{not } C) \cdot P(\text{not } C_i)] &= P(R | C) \cdot P(C_i) - P(R_i) && \text{Move } P(R_i) \text{ to the right side} \\
+  P(R | \text{not } C) \cdot P(\text{not } C_i) &= P(R_i) - [P(R | C) \cdot P(C_i)] && \text{Multiply by } -1 \\
+  P(R | \text{not } C) &= \frac{P(R_i) - [P(R | C) \cdot P(C_i)]}{P(\text{not } C_i)} && \text{Divide by } P(R | \text{not } C) \\
+  &= \frac{1\% - [95\% \times .3\%]}{99.7\%} = \frac{.01 - [.95 \times .003]}{.997}\\
+  &= .7171515\% \\
+\end{aligned}
+$$ {#eq-conditionalProbabilityRevised}
+
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pBgivenNotA()` function that estimates the probability of one event, $B$, given that another event, $A$, did not occur.
+
+```{r}
+petersenlab::pBgivenNotA(pBgivenA = .95, pA = .003, pB = .01)
+```
+
+With this [conditional probability](#sec-conditionalProbability) ($P(R | \text{not } C)$), the updated [marginal probability](#sec-baseRate) of having HIV ($P(C_i)$), and the updated marginal probability of not having HIV ($P(\text{not } C_i)$), we can now calculate an updated estimate of the [marginal probability](#sec-baseRate) of getting a second positive test.
+The probability of getting a second positive test is the probability that a person with HIV gets a second positive test (i.e., sensitivity) times the updated probability of HIV plus the probability that a person without HIV gets a second positive test (i.e., false positive rate) times the updated probability of not having HIV, as in @eq-baseRateUpdated:
+
+$$
+\begin{aligned}
+  P(R_{i}) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i) \\
+  &= 95\% \times 28.5\% + .7171515\% \times 71.5\% = .95 \times .285 + .007171515 \times .715 \\
+  &= 27.58776\%
+\end{aligned}
+$$ {#eq-baseRateUpdated}
+
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pB()` function that estimates the marginal probability of one event, $B$.
+
+```{r}
+petersenlab::pB(pBgivenA = .95, pA = .285, pBgivenNotA = .007171515)
+```
+
+We then substitute the updated [marginal probability](#sec-baseRate) of HIV ($P(C_i)$) and the updated [marginal probability](#sec-baseRate) of getting a second positive test ($P(R_i)$) into Bayes' theorem to get the probability that the person has HIV if they have a second positive test (assuming the errors of each test are independent, i.e., uncorrelated), as in @eq-baseRateUpdated2:
+
+$$
+\begin{aligned}
+  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)} \\
+  P(\text{HIV} | \text{a second positive test}) &= \frac{P(\text{a second positive test} | \text{HIV}) \cdot P(\text{HIV})}{P(\text{a second positive test})} \\
+  &= \frac{\text{sensitivity of test} \times \text{updated base rate of HIV}}{\text{updated base rate of positive test}} \\
+  &= \frac{95\% \times 28.5\%}{27.58776\%} \\
+  &= 98.14\%
+\end{aligned}
+$$ {#eq-baseRateUpdated2}
+
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.
+
+```{r}
+pAgivenB(pBgivenA = .95, pA = .285, pB = .2758776)
+```
+
+Thus, a second positive test greatly increases the posterior probability that the person has HIV from 28.5% to over 98%.
+
+As seen in the rearranged formula in @eq-bayes2, the ratio of the [conditional probabilities](#sec-conditionalProbability) is equal to the ratio of the [base rates](#sec-baseRate).
+Thus, it is important to consider [base rates](#sec-baseRate).
+People have a strong tendency to ignore (or give insufficient weight to) [base rates](#sec-baseRate) when making predictions.
+The failure to consider the [base rate](#sec-baseRate) when making predictions when given specific information about a case is known as the [base-rate fallacy](#sec-fallaciesBaseRate) or as [base rate neglect](#sec-fallaciesBaseRate).
+For example, people tend to say that the probability of a rare event is more likely than it actually is given specific information.
+
+As seen in the rearranged formula in @eq-bayes3, the inverse [conditional probabilities](#sec-conditionalProbability) ($P(C | R)$ and $P(R | C)$) are not equal unless the [base rates](#sec-baseRate) of $C$ and $R$ are the same.
+If the [base rates](#sec-baseRate) are not equal, we are making at least some prediction errors.
+If $P(C_i) > P(R_i)$, our predictions must include some false negatives.
+If $P(R_i) > P(C_i)$, our predictions must include some false positives.
+
+In sum, the [marginal probability](#sec-baseRate), including the [prior probability](#sec-baseRate) or [base rate](#sec-baseRate), should be weighed heavily in predictions unless there are sufficient data to indicate otherwise, i.e., to update the posterior probability based on new evidence.
+Bayes' theorem provides a powerful tool to anchor predictions to the [base rate](#sec-baseRate) unless sufficient evidence changes the posterior probability (by updating the evidence and [prior probability](#sec-baseRate)).
+
+## Base Rate of Rookie Performance {#sec-baseRateRookiePerformance}
+
+### Quarterbacks {#sec-baseRateRookiePerformanceQBs}
+
+### Running Backs {#sec-baseRateRookiePerformanceRBs}
+
+## How to Account for Base Rates {#sec-accountForBaseRates}
 

### 2024-07-13 cognitive-bias.qmd
20240712 - base rates

@@
 - gambler's fallacy
+- conditional probability fallacy
 
@@
 The probability of work-related injuries is higher for some occupations (e.g., construction) and for some groups (e.g., men, 55–64-year-olds, Black or Multiracial, who are self-employed and have less than high school education) than others.
-Nevertheless, is we ignore all of the interacting factors, the general probability of work-related injuries is 35%.
+Nevertheless, if we ignore all of the interacting factors, the general probability of work-related injuries is 35%.
 If we made a prediction that someone would be highly likely (> 90%) to experience a work-related injury because they are male and self-employed, this would be ignoring the relatively lower base rate of work-related injury.
@@
 
-We discuss ways to counteract the base rate fallacy in @sec-baseRates.
+More information on base rates and how to counteract the base rate fallacy in described in @sec-baseRates.
 
@@
 
+### Conditional Probability Fallacy {#sec-fallaciesConditionalProbability}
+
+We describe the [conditional probability fallacy](#sec-inverseFallacy) in @sec-inverseFallacy after introducing [conditional probability](#sec-conditionalProbability) in @sec-conditionalProbability.
+
 ## Conclusion {#sec-cognitiveBiasConclusion}

### 2024-07-13 evaluating-prediction-accuracy.qmd
20240712 - base rates

@@
 
-# Types of Accuracy {#sec-accuracyTypes}
+## Overview
 
-## Discrimination {#sec-discrimination}
+Predictions can come in different types.
+Some predictions involve categorical data, whereas other predictions involve continuous data.
+When dealing with categorical data, we can evaluate predictions using a 2x2 table known as a confusion matrix (see INSERT), or with logistic regression models.
+When dealing with continuous data, we can evaluate predictions using multiple regression or similar variants such as structural equation modeling and mixed models.
+
+## Types of Accuracy {#sec-accuracyTypes}
+
+### Discrimination {#sec-discrimination}
 
@@
 
-## Calibration [#sec-calibration]
+### Calibration [#sec-calibration]
 
@@
 
-## General Accuracy {#sec-generalAccuracy}
+### General Accuracy {#sec-generalAccuracy}
 

### 2024-07-13 base-rates.qmd
20240712 - clean up

@@
 ```{r}
-
+library("petersenlab")
 ```
@@
 ```{r}
-petersenlab::pAgivenB(pBgivenA = .95, pA = .003, pB = .01)
+petersenlab::pAgivenB(
+  pBgivenA = .95,
+  pA = .003,
+  pB = .01)
 ```
@@
 ```{r}
-petersenlab::pBgivenNotA(pBgivenA = .95, pA = .003, pB = .01)
+petersenlab::pBgivenNotA(
+  pBgivenA = .95,
+  pA = .003,
+  pB = .01)
 ```
@@
 ```{r}
-pAgivenB(pBgivenA = .95, pA = .285, pB = .2758776)
+petersenlab::pAgivenB(
+  pBgivenA = .95,
+  pA = .285,
+  pB = .2758776)
 ```
@@
 People have a strong tendency to ignore (or give insufficient weight to) [base rates](#sec-baseRate) when making predictions.
-The failure to consider the [base rate](#sec-baseRate) when making predictions when given specific information about a case is known as the [base-rate fallacy](#sec-fallaciesBaseRate) or as [base rate neglect](#sec-fallaciesBaseRate).
+The failure to consider the [base rate](#sec-baseRate) when making predictions when given specific information about a case is known as the [base rate fallacy](#sec-fallaciesBaseRate) or as [base rate neglect](#sec-fallaciesBaseRate).
 For example, people tend to say that the probability of a rare event is more likely than it actually is given specific information.

### 2024-07-13 base-rates.qmd
20240713 - posttest odds, posttest probability, probability nomogram

@@
 
-Player performance is a complex prediction task.
+Predicting player performance is a complex prediction task.
 Performance is probabilistically influenced by many processes, including processes internal to the player in addition to external processes.
@@
 And, similar to chaos theory, one small change in the system can lead to large differences later on.
+Moreover, there are important factors to keep in mind when making predictions.
 
@@
 
-In this case, we know the [marginal probability](#sec-baseRate) ($P(R_i)$), and we can use that to solve for the unknown [conditional probability](#sec-conditionalProbability) that reflects the [false positive rate](#sec-falsePositiveRate) ($P(R | \text{not } C)$), as in @eq-conditionalProbabilityRevised:
+In this case, we know the [marginal probability](#sec-baseRate) ($P(R_i)$), and we can use that to solve for the unknown [conditional probability](#sec-conditionalProbability) that reflects the false positive rate ($P(R | \text{not } C)$), as in @eq-conditionalProbabilityRevised:
 
@@
 
+There are various ways to account for [base rates](#sec-baseRate).
+One approach is to use [actuarial formulas](#sec-actuarialPrediction) (rather than human judgment) to make the predictions.
+[Actuarial formulas](#sec-actuarialPrediction) based on [multiple regression](#sec-multipleRegression) or [machine learning](#sec-machineLearning) can account for the [base rate](#sec-baseRate) of the event.
+
+Another approach is to leverage Bayes' theorem, using Bayesian updating.
+
+### Bayesian Updating {#bayesianUpdating}
+
+Bayesian updating involves comparing the relative probability of two outcomes, $P(C | R)$ versus $P(\text{not } C | R)$.
+If we want to compare the relative probability of two outcomes, we can use the odds form of Bayes' theorem, as in @eq-bayes5:
+$$
+\begin{aligned}
+  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)} \\
+  P(\text{not } C | R) &= \frac{P(R | \text{not } C) \cdot P(\text{not } C_i)}{P(R_i)} \\
+  \frac{P(C | R)}{P(\text{not } C | R)} &= \frac{\frac{P(R | C) \cdot P(C_i)}{P(R_i)}}{\frac{P(R | \text{not } C) \cdot P(\text{not } C_i)}{P(R_i)}} \\
+  &= \frac{P(R | C) \cdot P(C_i)}{P(R | \text{not } C) \cdot P(\text{not } C_i)} \\
+  &= \frac{P(C_i)}{P(\text{not } C_i)} \times \frac{P(R | C)}{P(R | \text{not } C)} \\
+  \text{posterior odds} &= \text{prior odds} \times \text{likelihood ratio}
+\end{aligned}
+$$ {#eq-bayes5}
+
+As presented in @eq-bayes5, the posttest (or posterior) odds are equal to the pretest odds multiplied by the likelihood ratio.
+Bayes' theorem is discussed in @sec-bayesTheorem.
+Using this formula, and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
+The calculation of posttest probability is described in INSERT.
+A *probability nomogram* is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the pretest (or prior) probability and likelihood ratio, as depicted in @fig-probabilityNomogram.
+To use a probability nomogram, connect the dots from the starting probability (left line) with the likelihood ratio (middle line) to see the updated probability.
+The updated (posttest) probability is where the connecting line crosses the third, right line.
+
+::: {#fig-probabilityNomogram}
+![](images/probabilityNomogram.png)
+
+Probability Nomogram.
+(Figure retrieved from [https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png](https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png)).
+:::
+
+For instance, if the starting probability is .5% and the likelihood ratio is 10 (e.g., sensitivity = .90, specificity = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$), the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the pretest probability and the likelihood ratio, or given the pretest probability and the sensitivity and specificity of the test.
+
+```{r}
+petersenlab::posttestProbability(
+  pretestProb = .005,
+  likelihoodRatio = 10)
+
+petersenlab::posttestProbability(
+  pretestProb = .005,
+  SN = .90,
+  SP = .91)
+
+petersenlab::posttestProbability(
+  TP = 450,
+  TN = 90545,
+  FP = 8955,
+  FN = 50)
+```
+
+::: {#fig-probabilityNomogramLine}
+![](images/probabilityNomogram.png)
+
+Probability Nomogram Example.
+(Figure adapted from [https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png](https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png).
+Also provided in: @Petersen2024a and @PetersenPrinciplesPsychAssessment.)
+:::
+
+A probability nomogram calculator can be found at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at https://perma.cc/X8TF-7YBX).
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot, as adapted from https://github.com/achekroud/nomogrammer:
+
+```{r}
+petersenlab::nomogrammer(
+  pretestProb = .005,
+  SN = 0.90,
+  SP = 0.91)
+
+petersenlab::nomogrammer(
+  pretestProb = .005,
+  PLR = 10,
+  NLR = 0.11)
+
+petersenlab::nomogrammer(
+  TP = 450,
+  TN = 90545,
+  FP = 8955,
+  FN = 50)
+```
+
 ## Conclusion {#sec-baseRatesConclusion}

### 2024-07-13 base-rates.qmd
20240713 - updates

@@
 
-Another approach is to leverage Bayes' theorem, using Bayesian updating.
+Another approach is to leverage Bayes' theorem, using Bayesian updating and the probability nomogram.
 
@@
 ::: {#fig-probabilityNomogram}
-![](images/probabilityNomogram.png)
+![](images/probabilityNomogram.png){width=50%}
 
@@
 
-For instance, if the starting probability is .5% and the likelihood ratio is 10 (e.g., sensitivity = .90, specificity = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$), the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.
+For instance, if the starting probability is 0.5% and the likelihood ratio is 10 (e.g., sensitivity = .90, specificity = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$) from a positive test, the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.
 The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the pretest probability and the likelihood ratio, or given the pretest probability and the sensitivity and specificity of the test.
@@
 
+If the starting probability is 0.5% and the likelihood ratio is 0.11 from a negative test, the updated probability is essentially nearly indistinguishable from zero (0.05%).
+```{r}
+petersenlab::posttestProbability(
+  pretestProb = .005,
+  likelihoodRatio = 0.11)
+```
+
 ::: {#fig-probabilityNomogramLine}
-![](images/probabilityNomogram.png)
+![](images/probabilityNomogramLine.png){width=50%}
 
@@
 
-A probability nomogram calculator can be found at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at https://perma.cc/X8TF-7YBX).
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot, as adapted from https://github.com/achekroud/nomogrammer:
+A probability nomogram calculator can be found at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot, as adapted from Adam Chekroud (<https://github.com/achekroud/nomogrammer>):
 
@@
 
+The blue line indicates the [posterior probability](#posttestProbability) of the condition given a positive test.
+The pink line indicates the [posterior probability](#posttestProbability) of the condition given a negative test.
+
 ## Conclusion {#sec-baseRatesConclusion}

### 2024-07-13 base-rates.qmd
20240713 - clarifications

@@
 
-### Bayesian Updating {#bayesianUpdating}
+### Bayesian Updating {#sec-bayesianUpdating}
 
@@
 
-As presented in @eq-bayes5, the posttest (or posterior) odds are equal to the pretest odds multiplied by the likelihood ratio.
-Bayes' theorem is discussed in @sec-bayesTheorem.
-Using this formula, and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
+As presented in @eq-bayes5, the posttest (or posterior) odds are equal to the pretest odds multiplied by the [likelihood ratio](#sec-diagnosticLikelihoodRatio).
+Below, we describe the [likelihood ratio](#sec-diagnosticLikelihoodRatio).
+
+#### Diagnostic Likelihood Ratio {#sec-diagnosticLikelihoodRatio}
+
+A likelihood ratio is the ratio of two probabilities.
+It can be used to compare the likelihood of two possibilities.
+The diagnostic likelihood ratio is an index of the predictive validity of an instrument: it is the ratio of the probability that a test result is correct to the probability that the test result is incorrect.
+The diagnostic likelihood ratio is also called the risk ratio.
+There are two types of diagnostic likelihood ratios: the [positive likelihood ratio](#sec-positiveLikelihoodRatio) and the [negative likelihood ratio](#sec-negativeLikelihoodRatio).
+
+##### Positive Likelihood Ratio (LR+) {#sec-positiveLikelihoodRatio}
+
+The positive likelihood ratio (LR+) compares the true positive rate to the false positive rate.
+Positive likelihood ratio values range from 1 to infinity.
+Higher values reflect greater accuracy, because it indicates the degree to which a true positive is more likely than a false positive.
+The formula for calculating the positive likelihood ratio is in @eq-positiveLikelihoodRatio.
+
+$$
+\begin{aligned}
+  \text{positive likelihood ratio (LR+)} &= \frac{\text{TPR}}{\text{FPR}} \\
+  &= \frac{P(R|C)}{P(R|\text{not } C)} \\
+  &= \frac{P(R|C)}{1 - P(\text{not } R|\text{not } C)} \\
+  &= \frac{\text{sensitivity}}{1 - \text{specificity}}
+\end{aligned}
+$$ {#eq-positiveLikelihoodRatio}
+
+##### Negative Likelihood Ratio (LR−) {#sec-negativeLikelihoodRatio}
+
+The negative likelihood ratio (LR−) compares the false negative rate to the true negative rate.
+Negative likelihood ratio values range from 0 to 1.
+Smaller values reflect greater accuracy, because it indicates that a false negative is less likely than a true negative.
+The formula for calculating the negative likelihood ratio is in @negativeLikelihoodRatio.
+
+$$
+\begin{aligned}
+  \text{negative likelihood ratio } (\text{LR}-) &= \frac{\text{FNR}}{\text{TNR}} \\
+  &= \frac{P(\text{not } R|C)}{P(\text{not } R|\text{not } C)} \\
+  &= \frac{1 - P(R|C)}{P(\text{not } R|\text{not } C)} \\
+  &= \frac{1 - \text{sensitivity}}{\text{specificity}}
+\end{aligned}
+$$ {#eq-positiveLikelihoodRatio}
+
+#### Probability Nomogram {#sec-probabilityNomogram}
+
+Using [Bayes's theorem](#sec-bayesTheorem) (described in @sec-bayesTheorem), solving for posttest odds (based on pretest odds and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), as in @eq-bayes5), and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
 The calculation of posttest probability is described in INSERT.
-A *probability nomogram* is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the pretest (or prior) probability and likelihood ratio, as depicted in @fig-probabilityNomogram.
-To use a probability nomogram, connect the dots from the starting probability (left line) with the likelihood ratio (middle line) to see the updated probability.
+A *probability nomogram* is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the [pretest (or prior) probability](#sec-baseRate) and [likelihood ratio](#sec-diagnosticLikelihoodRatio), as depicted in @fig-probabilityNomogram.
+To use a probability nomogram, connect the dots from the starting probability (left line) with the [likelihood ratio](#sec-diagnosticLikelihoodRatio) (middle line) to see the updated probability.
 The updated (posttest) probability is where the connecting line crosses the third, right line.
@@
 
-For instance, if the starting probability is 0.5% and the likelihood ratio is 10 (e.g., sensitivity = .90, specificity = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$) from a positive test, the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the pretest probability and the likelihood ratio, or given the pretest probability and the sensitivity and specificity of the test.
+For instance, if the starting probability is 0.5% and the [likelihood ratio](#sec-diagnosticLikelihoodRatio) is 10 (e.g., sensitivity = .90, specificity = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$) from a positive test (i.e., [positive likelihood ratio](#sec-positiveLikelihoodRatio)), the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the sensitivity (SN) and specificity (SP) of the test.
 
@@
   SP = .91)
+```
 
+The function can also estimate the posttest probability of an event given the number of true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN):
+
+```{r}
 petersenlab::posttestProbability(
@@
 
-If the starting probability is 0.5% and the likelihood ratio is 0.11 from a negative test, the updated probability is essentially nearly indistinguishable from zero (0.05%).
+We discuss true positives (TP), true negatives (TN), false positives (FP), false negatives (FN), sensitivity (SN), and specificity (SP) in INSERT.
+
+If the starting probability is 0.5% and the [likelihood ratio](#sec-diagnosticLikelihoodRatio) is 0.11 from a negative test (i.e., [negative likelihood ratio](#sec-negativeLikelihoodRatio)), the updated probability is nearly indistinguishable from zero (0.05%).
+
 ```{r}
@@
 A probability nomogram calculator can be found at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot, as adapted from Adam Chekroud (<https://github.com/achekroud/nomogrammer>):
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the sensitivity (SN) and specificity (SP) of the test, as adapted from Adam Chekroud (<https://github.com/achekroud/nomogrammer>):
 
@@
   NLR = 0.11)
+```
 
+The blue line indicates the [posterior probability](#posttestProbability) of the condition given a positive test.
+The pink line indicates the [posterior probability](#posttestProbability) of the condition given a negative test.
+
+The function can also create a nomogram plot using the true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN):
+
+```{r}
 petersenlab::nomogrammer(
@@
 ```
-
-The blue line indicates the [posterior probability](#posttestProbability) of the condition given a positive test.
-The pink line indicates the [posterior probability](#posttestProbability) of the condition given a negative test.
 

### 2024-07-13 base-rates.qmd
20240713 - more context

@@
 
-There are various ways to account for [base rates](#sec-baseRate).
-One approach is to use [actuarial formulas](#sec-actuarialPrediction) (rather than human judgment) to make the predictions.
+There are various ways to account for [base rates](#sec-baseRate), including the use of [actuarial formulas](#sec-accountForBaseRatesActuarial) and the use of [Bayesian updating](#sec-bayesianUpdating).
+
+### Actuarial Formula {#sec-accountForBaseRatesActuarial}
+
+One approach to account for [base rates](#sec-baseRate) is to use [actuarial formulas](#sec-actuarialPrediction) (rather than [human judgment](#sec-humanJudgment)) to make the predictions.
 [Actuarial formulas](#sec-actuarialPrediction) based on [multiple regression](#sec-multipleRegression) or [machine learning](#sec-machineLearning) can account for the [base rate](#sec-baseRate) of the event.
-
-Another approach is to leverage Bayes' theorem, using Bayesian updating and the probability nomogram.
 
@@
 
-Bayesian updating involves comparing the relative probability of two outcomes, $P(C | R)$ versus $P(\text{not } C | R)$.
+Another approach to account for [base rates](#sec-baseRate) is to leverage Bayes' theorem, using Bayesian updating and the [probability nomogram](#sec-probabilityNomogram).
+Bayesian updating is a form of [anchoring and adjustment](#sec-heuristicsAnchoringAdjustment); however, unlike the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment), it is a systematic approach to [anchoring and adjustment](#sec-heuristicsAnchoringAdjustment) that anchors one's predictions to the base rate, and then adjusts according to new information.
+That is, we start with a [pretest probability](#sec-baseRate) (i.e., [base rate](#sec-baseRate)) and update our predictions based on the extent of new information (i.e., the [likelihood ratio](#sec-diagnosticLikelihoodRatio)).
+
+To perform Bayesian updating involves comparing the relative probability of two outcomes, $P(C | R)$ versus $P(\text{not } C | R)$.
 If we want to compare the relative probability of two outcomes, we can use the odds form of Bayes' theorem, as in @eq-bayes5:

### 2024-07-13 basic-statistics.qmd
20240913 - additions and clean up

@@
 ### Null Hypothesis Significance Testing {#sec-nhst}
+
+To draw statistical inferences, the frequentist statistics paradigm leverages null hypothesis significance testing.
+Frequentist statistics is the most widely used statistical paradigm.
+However, frequentist statistics is not the only statistical paradigm.
+Other statistical paradigms exist, including [Bayesian statistics](#sec-bayesTheorem), which is based on [Bayes' theorem](#sec-bayesTheorem).
+This chapter focuses on the frequentist approach to hypothesis testing, known as null hypothesis significance testing.
+We discuss Bayesian statistics in @sec-baseRates.
 

### 2024-07-14 evaluating-prediction-accuracy.qmd
20240713 - evaluating prediction accuracy

@@
 ```{r}
-
+library("petersenlab")
+library("tidyverse")
+library("pROC")
+library("magrittr")
+library("viridis")
 ```
 
-## Overview
+## Overview {#sec-predictionAccuracyOverview}
 
@@
 Some predictions involve categorical data, whereas other predictions involve continuous data.
-When dealing with categorical data, we can evaluate predictions using a 2x2 table known as a confusion matrix (see INSERT), or with logistic regression models.
-When dealing with continuous data, we can evaluate predictions using multiple regression or similar variants such as structural equation modeling and mixed models.
+When dealing with a dichotomous (nominal data that are binary) predictor and outcome variable, we can evaluate predictions using a 2x2 table known as a confusion matrix (see INSERT), or with logistic regression models.
+When dealing with a continuous outcome variable (e.g., ordinal, interval, or ratio data), we can evaluate predictions using multiple regression or similar variants such as structural equation modeling and mixed models.
+
+## Prediction of Categorical Outcomes {#sec-predictionCategorical}
+
+### Decision Outcomes {#sec-decisionOutcomes}
+
+To consider how we can evaluate the accuracy of predictions, consider an example adapted from @Meehl1955.
+The military conducts a test of its prospective members to screen out applicants who would likely fail basic training.
+To evaluate the accuracy of our predictions using the test, we can examine a confusion matrix.
+A confusion matrix is a matrix that presents the predicted outcome on one dimension and the actual outcome (truth) on the other dimension.
+If the predictions and outcomes are dichotomous, the confusion matrix is a 2x2 matrix with two rows and two columns that represent four possible predicted-actual combinations (decision outcomes): true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN).
+
+When discussing the four decision outcomes, "true" means an accurate judgment, whereas "false" means an inaccurate judgment.
+"Positive" means that the judgment was that the person has the characteristic of interest, whereas "negative" means that the judgment was that the person does not have the characteristic of interest.
+A *true positive* is a correct judgment (or prediction) where the judgment was that the person has (or will have) the characteristic of interest, and, in truth, they actually have (or will have) the characteristic.
+A *true negative* is a correct judgment (or prediction) where the judgment was that the person does not have (or will not have) the characteristic of interest, and, in truth, they actually do not have (or will not have) the characteristic.
+A *false positive* is an incorrect judgment (or prediction) where the judgment was that the person has (or will have) the characteristic of interest, and, in truth, they actually do not have (or will not have) the characteristic.
+A *false negative* is an incorrect judgment (or prediction) where the judgment was that the person does not have (or will not have) the characteristic of interest, and, in truth, they actually do have (or will have) the characteristic.
+
+An example of a confusion matrix is in INSERT.
+
+With the information in the confusion matrix, we can calculate the marginal sums and the proportion of people in each cell (in parentheses), as depicted in INSERT.
+
+That is, we can sum across the rows and columns to identify how many people actually showed poor adjustment ($n = 100$) versus good adjustment ($n = 1,900$), and how many people were selected to reject ($n = 508$) versus retain ($n = 1,492$).
+If we sum the column of predicted marginal sums ($508 + 1,492$) or the row of actual marginal sums ($100 + 1,900$), we get the total number of people ($N = 2,000$).
+
+Based on the marginal sums, we can compute the [marginal probabilities](#sec-baseRate), as depicted in INSERT.
+
+The [marginal probability](#sec-baseRate) of the person having the characteristic of interest (i.e., showing poor adjustment) is called the [*base rate*](#baseRate) (BR).
+That is, the [base rate](#baseRate) is the proportion of people who have the characteristic.
+It is calculated by dividing the number of people with poor adjustment ($n = 100$) by the total number of people ($N = 2,000$): $BR = \frac{FN + TP}{N}$.
+Here, the [base rate](#baseRate) reflects the prevalence of poor adjustment.
+In this case, the [base rate](#baseRate) is .05, so there is a 5% chance that an applicant will be poorly adjusted.
+The [marginal probability](#baseRate) of good adjustment is equal to 1 minus the [base rate](#baseRate) of poor adjustment.
+
+The [marginal probability](#baseRate) of predicting that a person has the characteristic (i.e., rejecting a person) is called the *selection ratio* (SR).
+The selection ratio is the proportion of people who will be selected (in this case, rejected rather than retained); i.e., the proportion of people who are identified as having the characteristic.
+The selection ratio is calculated by dividing the number of people selected to reject ($n = 508$) by the total number of people ($N = 2,000$): $SR = \frac{TP + FP}{N}$.
+In this case, the selection ratio is .25, so 25% of people are rejected.
+The [marginal probability](#baseRate) of not selecting someone to reject (i.e., the [marginal probability](#baseRate) of retaining) is equal to 1 minus the selection ratio.
+
+The selection ratio might be something that the test dictates according to its cutoff score.
+Or, the selection ratio might be imposed by external factors that place limits on how many people you can assign a positive test value.
+For instance, when deciding whether to treat a client, the selection ratio may depend on how many therapists are available and how many cases can be treated.
+
+### Percent Accuracy {#sec-percentAccuracyOverview}
+
+Based on the confusion matrix, we can calculate the prediction accuracy based on the percent accuracy of the predictions.
+The percent accuracy is the number of correct predictions divided by the total number of predictions, and multiplied by 100.
+In the context of a confusion matrix, this is calculated as: $100\% \times \frac{\text{TP} + \text{TN}}{N}$.
+In this case, our percent accuracy was 78%—that is, 78% of our predictions were accurate, and 22% of our predictions were inaccurate.
+
+### Percent Accuracy by Chance {#sec-accuracyByChance}
+
+78% sounds pretty accurate.
+And it is much higher than 50%, so we are doing a pretty good job, right?
+Well, it is important to compare our accuracy to what accuracy we would expect to get by chance alone, if predictions were made by a random process rather than using a test's scores.
+Our selection ratio was 25.4%.
+How accurate would we be if we randomly selected 25.4% of people to reject?
+To determine what accuracy we could get by chance alone given the selection ratio and the base rate, we can calculate the chance probability of true positives and the chance probability of true negatives.
+The probability of a given cell in the confusion matrix is a [joint probability](#sec-jointProbability)—the probability of two events occurring simultaneously.
+To calculate a [joint probability](#sec-jointProbability), we multiply the probability of each event.
+
+So, to get the chance expectancies of true positives, we would multiply the respective [marginal probabilities](#baseRate), as in Equation @eq-truePositivesByChanceExample:
+
+$$
+\begin{aligned}
+  P(TP) &= P(\text{Poor adjustment}) \times P(\text{Reject})\\
+   &= BR \times SR \\
+  &= .05 \times .254 \\
+  &= .0127
+\end{aligned}
+$$ {#eq-truePositivesByChanceExample}
+
+To get the chance expectancies of true negatives, we would multiply the respective [marginal probabilities](#sec-baseRate), as in @trueNegativesByChanceExample:
+
+$$
+\begin{aligned}
+  P(TN) &= P(\text{Good adjustment}) \times P(\text{Retain})\\
+   &= (1 - BR) \times (1 - SR) \\
+  &= .95 \times .746 \\
+  &= .7087
+\end{aligned}
+$$ {#eq-trueNegativesByChanceExample}
+
+To get the percent accuracy by chance, we sum the chance expectancies for the correct predictions (TP and TN): $.0127 + .7087 = .7214$.
+Thus, the percent accuracy you can get by chance alone is 72%.
+This is because most of our predictions are to retain people, and the [base rate](#sec-baseRate) of poor adjustment is quite low (.05).
+Our measure with 78% accuracy provides only a 6% increment in correct predictions.
+Thus, you cannot judge how good your judgment or prediction is until you know how you would do by random chance.
+
+The chance expectancies for each cell of the confusion matrix are in INSERT
+
+### Predicting from the Base Rate {#sec-predictingFromBaseRate}
+
+Now, let us consider how well you would do if you were to predict from the [base rate](#sec-baseRate).
+Predicting from the [base rate](#sec-baseRate) is also called "betting from the [base rate](#sec-baseRate)", and it involves setting the selection ratio by taking advantage of the [base rate](#sec-baseRate) so that you go with the most likely outcome in every prediction.
+Because the [base rate](#sec-baseRate) is quite low (.05), we could predict from the [base rate](#sec-baseRate) by selecting no one to reject (i.e., setting the selection ratio at zero).
+Our percent accuracy by chance if we predict from the [base rate](#sec-baseRate) would be calculated by multiplying the [marginal probabilities](#sec-baseRate), as we did above, but with a new selection ratio, as in @eq-predictingFromBaseRateExample:
+
+$$
+\begin{aligned}
+  P(TP) &= P(\text{Poor adjustment}) \times P(\text{Reject})\\
+   &= BR \times SR \\
+  &= .05 \times 0 \\
+  &= 0 \\ \\
+  P(TN) &= P(\text{Good adjustment}) \times P(\text{Retain})\\
+   &= (1 - BR) \times (1 - SR) \\
+  &= .95 \times 1 \\
+  &= .95
+\end{aligned}
+$$ {#eq-predictingFromBaseRateExample}
+
+We sum the chance expectancies for the correct predictions (TP and TN): $0 + .95 = .95$.
+Thus, our percent accuracy by predicting from the [base rate](#sec-baseRate) is 95%.
+This is damning to our measure because it is a much higher accuracy than the accuracy of our measure.
+That is, we can be much more accurate than our measure simply by predicting from the [base rate](#sec-baseRate) and selecting no one to reject.
+
+Going with the most likely outcome in every prediction (predicting from the [base rate](#sec-baseRate)) can be highly accurate (in terms of percent accuracy) as noted by @Meehl1955, especially when the [base rate](#sec-baseRate) is very low or very high.
+This should serve as an important reminder that we need to compare the accuracy of our measures to the accuracy by (1) random chance and (2) predicting from the [base rate](#sec-baseRate).
+There are several important implications of the impact of [base rates](#sec-baseRate) on prediction accuracy.
+One implication is that using the same test in different settings with different [base rates](#sec-baseRate) will markedly change the accuracy of the test.
+Oftentimes, using a test will actually *decrease* the predictive accuracy when the [base rate](#sec-baseRate) deviates greatly from .50.
+But percent accuracy is not everything.
+Percent accuracy treats different kinds of errors as if they are equally important.
+However, the value we place on different kinds of errors may be different, as described next.
+
+### Different Kinds of Errors Have Different Costs {#sec-differentErrorsDifferentCosts}
+
+Some errors have a high cost, and some errors have a low cost.
+Among the four decision outcomes, there are two types of errors: false positives and false negatives.
+The extent to which false positives and false negatives are costly depends on the prediction problem.
+So, even though you can often be most accurate by going with the [base rate](#sec-baseRate), it may be advantageous to use a screening instrument despite lower overall accuracy because of the huge difference in costs of false positives versus false negatives in some cases.
+
+Consider the example of a screening instrument for HIV.
+False positives would be cases where we said that someone is at high risk of HIV when they are not, whereas false negatives are cases where we said that someone is not at high risk when they actually are.
+The costs of false positives include a shortage of blood, some follow-up testing, and potentially some anxiety, but that is about it.
+The costs of false negatives may be people getting HIV.
+In this case, the costs of false negatives greatly outweigh the costs of false positives, so we use a screening instrument to try to identify the cases at high risk for HIV because of the important consequences of failing to do so, even though using the screening instrument will lower our overall accuracy level.
+
+Another example is when the Central Intelligence Agency (CIA) used a screen for protective typists during wartime to try to detect spies.
+False positives would be cases where the CIA believes that a person is a spy when they are not, and the CIA does not hire them.
+False negatives would be cases where the CIA believes that a person is not a spy when they actually are, and the CIA hires them.
+In this case, a false positive would be fine, but a false negative would be really bad.
+
+How you weigh the costs of different errors depends considerably on the domain and context.
+Possible costs of false positives to society include: unnecessary and costly treatment with side effects and sending an innocent person to jail (despite our presumption of innocence in the United States criminal justice system that a person is innocent until proven guilty).
+Possible costs of false negatives to society include: setting a guilty person free, failing to detect a bomb or tumor, and preventing someone from getting treatment who needs it.
+
+The differential costs of different errors also depend on how much flexibility you have in the selection ratio in being able to set a stringent versus loose selection ratio.
+Consider if there is a high cost of getting rid of people during the selection process.
+For example, if you must hire 100 people and only 100 people apply for the position, you cannot lose people, so you need to hire even high-risk people.
+However, if you do not need to hire many people, then you can hire more conservatively.
+
+Any time the selection ratio differs from the [base rate](#sec-baseRate), you will make errors.
+For example, if you reject 25% of applicants, and the [base rate](#sec-baseRate) of poor adjustment is 5%, then you are making errors of over-rejecting (false positives).
+By contrast, if you reject 1% of applicants and the [base rate](#sec-baseRate) of poor adjustment is 5%, then you are making errors of under-rejecting or over-accepting (false negatives).
+
+A low [base rate](#sec-baseRate) makes it harder to make predictions, and tends to lead to less accurate predictions.
+For instance, it is very challenging to predict low [base rate](#sec-baseRate) behaviors, including suicide [@Kessler2020].
+
+### Sensitivity, Specificity, PPV, and NPV {#sec-sensitivitySpecificityPPVnpv}
+
+As described earlier, percent accuracy is not the only important aspect of accuracy.
+Percent accuracy can be misleading because it is highly influenced by [base rates](#sec-baseRate).
+You can have a high percent accuracy by predicting from the base rate and saying that no one has the condition (if the [base rate](#sec-baseRate) is low) or that everyone has the condition (if the [base rate](#sec-baseRate) is high).
+Thus, it is also important to consider other aspects of accuracy, including sensitivity (SN), specificity (SP), positive predictive value (PPV), and negative predictive value (NPV).
+We want our predictions to be sensitive to be able to detect the characteristic but also to be specific so that we classify only people actually with the characteristic as having the characteristic.
+
+Let us return to the confusion matrix in INSERT.
+If we know the frequency of each of the four predicted-actual combinations of the confusion matrix (TP, TN, FP, FN), we can calculate sensitivity, specificity, PPV, and NPV.
+
+Sensitivity is the proportion of those with the characteristic ($\text{TP} + \text{FN}$) that we identified with our measure ($\text{TP}$): $\frac{\text{TP}}{\text{TP} + \text{FN}} = \frac{86}{86 + 14} = .86$.
+Specificity is the proportion of those who do not have the characteristic ($\text{TN} + \text{FP}$) that we correctly classify as not having the characteristic ($\text{TN}$): $\frac{\text{TN}}{\text{TN} + \text{FP}} = \frac{1,478}{1,478 + 422} = .78$.
+PPV is the proportion of those who we classify as having the characteristic ($\text{TP} + \text{FP}$) who actually have the characteristic ($\text{TP}$): $\frac{\text{TP}}{\text{TP} + \text{FP}} = \frac{86}{86 + 422} = .17$
+NPV is the proportion of those we classify as not having the characteristic ($\text{TN} + \text{FN}$) who actually do not have the characteristic ($\text{TN}$): $\frac{\text{TN}}{\text{TN} + \text{FN}} = \frac{1,478}{1,478 + 14} = .99$.
+
+Sensitivity, specificity, PPV, and NPV are proportions, and their values therefore range from 0 to 1, where higher values reflect greater accuracy.
+With sensitivity, specificity, PPV, and NPV, we have a good snapshot of how accurate the measure is at a given cutoff.
+In our case, our measure is good at finding whom to reject (high sensitivity), but it is rejecting too many people who do not need to be rejected (lower PPV due to many FPs).
+Most people whom we classify as having the characteristic do not actually have the characteristic.
+However, the fact that we are over-rejecting could be okay depending on our goals, for instance, if we do not care about over-dropping (i.e., the PPV being low).
+
+#### Some Accuracy Estimates Depend on the Cutoff {#sec-accuracyCutoff}
+
+Sensitivity, specificity, PPV, and NPV differ based on the cutoff (i.e., threshold) for classification.
+Consider the following example.
+Aliens visit Earth, and they develop a test to determine whether a berry is edible or inedible.
+
+```{r}
+#| code-fold: true
+
+sampleSize <- 1000
+
+edibleScores <- rnorm(sampleSize, 50, 15)
+inedibleScores <- rnorm(sampleSize, 100, 15)
+
+edibleData <- data.frame(
+  score = c(
+    edibleScores,
+    inedibleScores),
+  type = c(
+    rep("edible", sampleSize),
+    rep("inedible", sampleSize)))
+
+cutoff <- 75
+
+hist_edible <- density(
+  edibleScores,
+  from = 0,
+  to = 150) %$% # from magrittr package
+  data.frame(
+    x = x,
+    y = y) %>%
+  mutate(area = x >= cutoff)
+
+hist_edible$type[hist_edible$area == TRUE] <- "edible_FP"
+hist_edible$type[hist_edible$area == FALSE] <- "edible_TN"
+
+hist_inedible <- density(
+  inedibleScores,
+  from = 0,
+  to = 150) %$% # from magrittr package
+  data.frame(
+    x = x,
+    y = y) %>%
+  mutate(area = x < cutoff)
+
+hist_inedible$type[hist_inedible$area == TRUE] <- "inedible_FN"
+hist_inedible$type[hist_inedible$area == FALSE] <- "inedible_TP"
+
+density_data <- bind_rows(
+  hist_edible,
+  hist_inedible)
+
+density_data$type <- factor(
+  density_data$type,
+  levels = c(
+    "edible_TN",
+    "inedible_TP",
+    "edible_FP",
+    "inedible_FN"))
+```
+
+@fig-classificationDistributions depicts the distributions of scores by berry type.
+Note how there are clearly two distinct distributions.
+However, the distributions overlap to some degree.
+Thus, any cutoff will have at least some inaccurate classifications.
+The extent of overlap of the distributions reflects the amount of measurement error of the measure with respect to the characteristic of interest.
+
+```{r}
+#| fig-cap: "Distribution of Test Scores by Berry Type."
+#| label: fig-classificationDistributions
+#| code-fold: true
+
+#No Cutoff
+ggplot(
+  data = edibleData,
+  aes(
+    x = score,
+    ymin = 0,
+    fill = type)) +
+  geom_density(alpha = .5) +
+  scale_fill_manual(
+    name = "Berry Type",
+    values = c(
+      viridis::viridis(2)[1],
+      viridis::viridis(2)[2])) +
+  scale_y_continuous(name = "Frequency") +
+  theme_bw() +
+  theme(
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+```
+
+@fig-classificationStandardCutoff depicts the distributions of scores by berry type with a cutoff.
+The red line indicates the cutoff—the level above which berries are classified by the test as inedible.
+There are errors on each side of the cutoff.
+Below the cutoff, there are some false negatives (blue): inedible berries that are inaccurately classified as edible.
+Above the cutoff, there are some false positives (green): edible berries that are inaccurately classified as inedible.
+Costs of false negatives could include sickness or death from eating the inedible berries.
+Costs of false positives could include taking longer to find food, finding insufficient food, and starvation.
+
+```{r}
+#| fig-cap: "Classifications Based on a Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
+#| label: fig-classificationStandardCutoff
+#| code-fold: true
+
+#Standard Cutoff
+ggplot(
+  data = density_data,
+  aes(
+    x = x,
+    ymin = 0,
+    ymax = y,
+    fill = type)) +
+  geom_ribbon(alpha = 1) +
+  scale_fill_manual(
+    name = "Berry Type",
+    values = c(
+      viridis::viridis(4)[4],
+      viridis::viridis(4)[1],
+      viridis::viridis(4)[3],
+      viridis::viridis(4)[2]),
+    breaks = c("edible_TN","inedible_TP","edible_FP","inedible_FN"),
+    labels = c("Edible: TN","Inedible: TP","Edible: FP","Inedible: FN")) +
+  geom_line(aes(y = y)) +
+  geom_vline(
+    xintercept = cutoff,
+    color = "red",
+    linewidth = 2) +
+  scale_x_continuous(name = "score") +
+  scale_y_continuous(name = "Frequency") +
+  theme_bw() +
+  theme(
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+```
+
+Based on our assessment goals, we might use a different selection ratio by changing the cutoff.
+@fig-classificationRaiseCutoff depicts the distributions of scores by berry type when we raise the cutoff.
+There are now more false negatives (blue) and fewer false positives (green).
+If we raise the cutoff (to be more conservative), the number of false negatives increases and the number of false positives decreases.
+Consequently, as the cutoff increases, sensitivity and NPV decrease (because we have more false negatives), whereas specificity and PPV increase (because we have fewer false positives).
+A higher cutoff could be optimal if the costs of false positives are considered greater than the costs of false negatives.
+For instance, if the aliens cannot risk eating the inedible berries because the berries are fatal, and there are sufficient edible berries that can be found to feed the alien colony.
+
+```{r}
+#| fig-cap: "Classifications Based on Raising the Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
+#| label: fig-classificationRaiseCutoff
+#| code-fold: true
+
+#Raise the cutoff
+cutoff <- 85
+
+hist_edible <- density(
+  edibleScores,
+  from = 0,
+  to = 150) %$% # from magrittr package
+  data.frame(
+    x = x,
+    y = y) %>%
+  mutate(area = x >= cutoff)
+
+hist_edible$type[hist_edible$area == TRUE] <- "edible_FP"
+hist_edible$type[hist_edible$area == FALSE] <- "edible_TN"
+
+hist_inedible <- density(
+  inedibleScores,
+  from = 0,
+  to = 150) %$% # from magrittr package
+  data.frame(
+    x = x,
+    y = y) %>%
+  mutate(area = x < cutoff)
+
+hist_inedible$type[hist_inedible$area == TRUE] <- "inedible_FN"
+hist_inedible$type[hist_inedible$area == FALSE] <- "inedible_TP"
+
+density_data <- bind_rows(
+  hist_edible,
+  hist_inedible)
+
+density_data$type <- factor(
+  density_data$type,
+  levels = c(
+    "edible_TN",
+    "inedible_TP",
+    "edible_FP",
+    "inedible_FN"))
+
+ggplot(
+  data = density_data,
+  aes(
+    x = x,
+    ymin = 0,
+    ymax = y,
+    fill = type)) +
+  geom_ribbon(alpha = 1) +
+  scale_fill_manual(
+    name = "Berry Type",
+    values = c(
+      viridis::viridis(4)[4],
+      viridis::viridis(4)[1],
+      viridis::viridis(4)[3],
+      viridis::viridis(4)[2]),
+    breaks = c("edible_TN","inedible_TP","edible_FP","inedible_FN"),
+    labels = c("Edible: TN","Inedible: TP","Edible: FP","Inedible: FN")) +
+  geom_line(aes(y = y)) +
+  geom_vline(
+    xintercept = cutoff,
+    color = "red",
+    linewidth = 2) +
+  scale_x_continuous(name = "score") +
+  scale_y_continuous(name = "Frequency") +
+  theme_bw() +
+  theme(
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+```
+
+@fig-classificationLowerCutoff depicts the distributions of scores by berry type when we lower the cutoff.
+There are now fewer false negatives (blue) and more false positives (green).
+If we lower the cutoff (to be more liberal), the number of false negatives decreases and the number of false positives increases.
+Consequently, as the cutoff decreases, sensitivity and NPV increase (because we have fewer false negatives), whereas specificity and PPV decrease (because we have more false positives).
+A lower cutoff could be optimal if the costs of false negatives are considered greater than the costs of false positives.
+For instance, if the aliens cannot risk missing edible berries because they are in short supply relative to the size of the alien colony, and eating the inedible berries would, at worst, lead to minor, temporary discomfort.
+
+```{r}
+#| fig-cap: "Classifications Based on Lowering the Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
+#| label: fig-classificationLowerCutoff
+#| code-fold: true
+
+#Lower the cutoff
+cutoff <- 65
+
+hist_edible <- density(
+  edibleScores,
+  from = 0,
+  to = 150) %$% # from magrittr package
+  data.frame(
+    x = x,
+    y = y) %>%
+  mutate(area = x >= cutoff)
+
+hist_edible$type[hist_edible$area == TRUE] <- "edible_FP"
+hist_edible$type[hist_edible$area == FALSE] <- "edible_TN"
+
+hist_inedible <- density(
+  inedibleScores,
+  from = 0,
+  to = 150) %$% # from magrittr package
+  data.frame(
+    x = x,
+    y = y) %>%
+  mutate(area = x < cutoff)
+
+hist_inedible$type[hist_inedible$area == TRUE] <- "inedible_FN"
+hist_inedible$type[hist_inedible$area == FALSE] <- "inedible_TP"
+
+density_data <- bind_rows(
+  hist_edible,
+  hist_inedible)
+
+density_data$type <- factor(
+  density_data$type,
+  levels = c(
+    "edible_TN",
+    "inedible_TP",
+    "edible_FP",
+    "inedible_FN"))
+
+ggplot(
+  data = density_data,
+  aes(
+    x = x,
+    ymin = 0,
+    ymax = y,
+    fill = type)) +
+  geom_ribbon(alpha = 1) +
+  scale_fill_manual(
+    name = "Berry Type",
+    values = c(
+      viridis::viridis(4)[4],
+      viridis::viridis(4)[1],
+      viridis::viridis(4)[3],
+      viridis::viridis(4)[2]),
+    breaks = c("edible_TN","inedible_TP","edible_FP","inedible_FN"),
+    labels = c("Edible: TN","Inedible: TP","Edible: FP","Inedible: FN")) +
+  geom_line(aes(y = y)) +
+  geom_vline(
+    xintercept = cutoff,
+    color = "red",
+    linewidth = 2) +
+  scale_x_continuous(name = "score") +
+  scale_y_continuous(name = "Frequency") +
+  theme_bw() +
+  theme(
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank())
+```
+
+In sum, sensitivity and specificity differ based on the cutoff for classification.
+if we raise the cutoff, sensitivity and PPV increase (due to fewer false positives), whereas sensitivity and NPV decrease (due to more false negatives).
+If we lower the cutoff, sensitivity and NPV increase (due to fewer false negatives), whereas specificity and PPV decrease (due to more false positives).
+Thus, the optimal cutoff depends on how costly each type of error is: false negatives and false positives.
+If false negatives are more costly than false positives, we would set a low cutoff.
+If false positives are more costly than false negatives, we would set a high cutoff.
+
+### Signal Detection Theory {#sec-sdt}
+
+Signal detection theory (SDT) is a probability-based theory for the detection of a given stimulus (signal) from a stimulus set that includes non-target stimuli (noise).
+SDT arose through the development of radar (**RA**dio **D**etection **A**nd **R**anging) and sonar (**SO**und **N**avigation **A**nd **R**anging) in World War II based on research on sensory-perception research.
+The military wanted to determine which objects on radar/sonar were enemy aircraft/submarines, and which were noise (e.g., different object in the environment or even just the weather itself).
+SDT allowed determining how many errors operators made (how accurate they were) and decomposing errors into different kinds of errors.
+SDT distinguishes between sensitivity and bias.
+In SDT, *sensitivity* (or discriminability) is how well an assessment distinguishes between a target stimulus and non-target stimuli (i.e., how well the assessment detects the target stimulus amid non-target stimuli).
+*Bias* is the extent to which the probability of a selection decision from the assessment is higher or lower than the true rate of the target stimulus.
+
+Some radar/sonar operators were not as sensitive to the differences between signal and noise, due to factors such as age, ability to distinguish gradations of a signal, etc.
+People who showed low sensitivity (i.e., who were not as successful at distinguishing between signal and noise) were screened out because the military perceived sensitivity as a skill that was not easily taught.
+By contrast, other operators could distinguish signal from noise, but their threshold was too low or high—they could take in information, but their decisions tended to be wrong due to systematic bias or poor calibration.
+That is, they systematically over-rejected or under-rejected stimuli.
+Over-rejecting leads to many false negatives (i.e., saying that a stimulus is safe when it is not).
+Under-rejecting leads to many false positives (i.e., saying that a stimulus is harmful when it is not).
+A person who showed good sensitivity but systematic bias was considered more teach-able than a person who showed low sensitivity.
+Thus, radar and sonar operators were selected based on their sensitivity to distinguish signal from noise, and then were trained to improve the calibration so they reduce their systematic bias and do not systematically over- or under-reject.
+
+Although SDT was originally developed for use in World War II, it now plays an important role in many areas of science and medicine.
+A medical application of SDT is tumor detection in radiology.
+SDT also plays an important role in psychology, especially cognitive psychology.
+For instance, research on social perception of sexual interest has shown that men tend to show lack of sensitivity to differences in women's affect—i.e., they have relative difficulties discriminating between friendliness and sexual interest [@Farris2008].
+Men also tend to show systematic bias (poor calibration) such that they tend to over-estimate women's sexual interest in them—i.e., men tend to have too low of a threshold for determining that a women is showing sexual interest in them [@Farris2006].
+
+SDT metrics of sensitivity include $d'$ ("$d$-prime"), $A$ (or $A'$), and the area under the receiver operating characteristic (ROC) curve.
+SDT metrics of bias include $\beta$ (beta), $c$, and $b$.
+
+#### Receiver Operating Characteristic (ROC) Curve {#sec-roc}
+
+The x-axis of the ROC curve is the false alarm rate or false positive rate ($1 -$ specificity).
+The y-axis is the hit rate or true positive rate (sensitivity).
+We can trace the ROC curve as the combination between sensitivity and specificity at every possible cutoff.
+At a cutoff of zero (top right of ROC curve), we calculate sensitivity (1.0) and specificity (0) and plot it.
+At a cutoff of zero, the assessment tells us to make an action for every stimulus (i.e., it is the most liberal).
+We then gradually increase the cutoff, and plot sensitivity and specificity at each cutoff.
+As the cutoff increases, sensitivity decreases and specificity increases.
+We end at the highest possible cutoff, where the sensitivity is 0 and the specificity is 1.0 (i.e., we never make an action; i.e., it is the most conservative).
+Each point on the ROC curve corresponds to a pair of hit and false alarm rates (sensitivity and specificity) resulting from a specific cutoff value.
+Then, we can draw lines or a curve to connect the points.
+
+INSERT depicts an empirical ROC plot where lines are drawn to connect the hit and false alarm rates.
+
+INSERT depicts an ROC curve where a smoothed and fitted curve is drawn to connect the hit and false alarm rates.
+
+##### Area Under the ROC Curve {#sec-auc}
+
+ROC methods can be used to compare and compute the discriminative power of measurement devices free from the influence of selection ratios, base rates, and costs and benefits.
+An ROC analysis yields a quantitative index of how well an index predicts a signal of interest or can discriminate between different signals.
+ROC analysis can help tell us how often our assessment would be correct.
+If we randomly pick two observations, and we were right once and wrong once, we were 50% accurate.
+But this would be a useless measure because it reflects chance responding.
+
+The geometrical area under the ROC curve reflects the discriminative accuracy of the measure.
+The index is called the **a**rea **u**nder the **c**urve (AUC) of an ROC curve.
+AUC quantifies the discriminative power of an assessment.
+AUC is the probability that a randomly selected target and a randomly selected non-target is ranked correctly by the assessment method.
+AUC values range from 0.0 to 1.0, where chance accuracy is 0.5 as indicated by diagonal line in the ROC curve.
+That is, a measure can be useful to the extent that its ROC curve is above the diagonal line (i.e., its discriminative accuracy is above chance).
+@fig-aucRange depicts ROC curves with a range of AUC values.
+
+```{r}
+#| fig-cap: "Receiver Operating Characteristic (ROC) Curves for Various Levels of Area Under The ROC Curve (AUC) for Various Measures."
+#| label: fig-aucRange
+#| code-fold: true
+
+set.seed(52242)
+
+auc60 <- petersenlab::simulateAUC(.60, 50000)
+auc70 <- petersenlab::simulateAUC(.70, 50000)
+auc80 <- petersenlab::simulateAUC(.80, 50000)
+auc90 <- petersenlab::simulateAUC(.90, 50000)
+auc95 <- petersenlab::simulateAUC(.95, 50000)
+auc99 <- petersenlab::simulateAUC(.99, 50000)
+
+plot(
+  pROC::roc(
+    y ~ x,
+    auc60,
+    smooth = TRUE),
+  legacy.axes = TRUE,
+  print.auc = TRUE,
+  print.auc.x = .52,
+  print.auc.y = .61,
+  print.auc.pattern = "%.2f")
+
+plot(
+  pROC::roc(
+    y ~ x,
+    auc70,
+    smooth = TRUE),
+  legacy.axes = TRUE,
+  print.auc = TRUE,
+  print.auc.x = .6,
+  print.auc.y = .67,
+  print.auc.pattern = "%.2f",
+  add = TRUE)
+
+plot(
+  pROC::roc(
+    y ~ x,
+    auc80,
+    smooth = TRUE),
+  legacy.axes = TRUE,
+  print.auc = TRUE,
+  print.auc.x = .695,
+  print.auc.y = .735,
+  print.auc.pattern = "%.2f",
+  add = TRUE)
+
+plot(
+  pROC::roc(
+    y ~ x,
+    auc90,
+    smooth = TRUE),
+  legacy.axes = TRUE,
+  print.auc = TRUE,
+  print.auc.x = .805,
+  print.auc.y = .815,
+  print.auc.pattern = "%.2f",
+  add = TRUE)
+
+plot(
+  pROC::roc(
+    y ~ x,
+    auc95,
+    smooth = TRUE),
+  legacy.axes = TRUE,
+  print.auc = TRUE,
+  print.auc.x = .875,
+  print.auc.y = .865,
+  print.auc.pattern = "%.2f",
+  add = TRUE)
+
+plot(
+  pROC::roc(
+    y ~ x,
+    auc99,
+    smooth = TRUE),
+  legacy.axes = TRUE,
+  print.auc = TRUE,
+  print.auc.x = .94,
+  print.auc.y = .94,
+  print.auc.pattern = "%.2f",
+  add = TRUE)
+```
+
+As an example, given an AUC of .75, this says that the overall score of an individual who has the characteristic in question will be higher 75% of the time than the overall score of an individual who does not have the characteristic.
+In lay terms, AUC provides the probability that we will classify correctly based on our instrument if we were to randomly pick one good and one bad outcome.
+AUC is a stronger index of accuracy than percent accuracy, because you can have high percent accuracy just by going with the base rate.
+AUC tells us how much better than chance a measure is at discriminating outcomes.
+AUC is useful as a measure of general discriminative accuracy, and it tells us how accurate a measure is at all possible cutoffs.
+Knowing the accuracy of a measure at all possible cutoffs can be helpful for selecting the optimal cutoff, given the goals of the assessment.
+In reality, however, we may not be interested in all cutoffs because not all errors are equal in their costs.
+
+If we lower the base rate, we would need a larger sample to get enough people to classify into each group.
+SDT/ROC methods are traditionally about dichotomous decisions (yes/no), not graded judgments.
+SDT/ROC methods can get messy with ordinal data that are more graded because you would have an AUC curve for each ordinal grouping.
+
+## Prediction of Continuous Outcomes {#sec-predictionCategorical}
 
@@
 
-### Calibration [#sec-calibration]
+### Calibration {#sec-calibration}
 

### 2024-07-14 evaluating-prediction-accuracy.qmd
20240713 - edits

@@
 
-So, to get the chance expectancies of true positives, we would multiply the respective [marginal probabilities](#baseRate), as in Equation @eq-truePositivesByChanceExample:
+So, to get the chance expectancies of true positives, we would multiply the respective [marginal probabilities](#baseRate), as in @eq-truePositivesByChanceExample:
 
@@
 
-To get the chance expectancies of true negatives, we would multiply the respective [marginal probabilities](#sec-baseRate), as in @trueNegativesByChanceExample:
+To get the chance expectancies of true negatives, we would multiply the respective [marginal probabilities](#sec-baseRate), as in @eq-trueNegativesByChanceExample:
 
@@
 
+@fig-classificationDistributions depicts the distributions of scores by berry type.
+Note how there are clearly two distinct distributions.
+However, the distributions overlap to some degree.
+Thus, any cutoff will have at least some inaccurate classifications.
+The extent of overlap of the distributions reflects the amount of measurement error of the measure with respect to the characteristic of interest.
+
 ```{r}
+#| fig-cap: "Distribution of Test Scores by Berry Type."
+#| label: fig-classificationDistributions
 #| code-fold: true
 
+#No Cutoff
 sampleSize <- 1000
@@
     "inedible_FN"))
-```
 
-@fig-classificationDistributions depicts the distributions of scores by berry type.
-Note how there are clearly two distinct distributions.
-However, the distributions overlap to some degree.
-Thus, any cutoff will have at least some inaccurate classifications.
-The extent of overlap of the distributions reflects the amount of measurement error of the measure with respect to the characteristic of interest.
-
-```{r}
-#| fig-cap: "Distribution of Test Scores by Berry Type."
-#| label: fig-classificationDistributions
-#| code-fold: true
-
-#No Cutoff
 ggplot(
@@
 #| label: fig-aucRange
+#| fig-width: 8
+#| fig-height: 8
 #| code-fold: true
@@
 #| fig-cap: "Types of Miscalibration. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
+#| fig-width: 8
+#| fig-height: 8
 #| code-fold: true

### 2024-07-14 base-rates.qmd
20240913 - additions and clean up
20240713 - calibration and base rates

@@
 ```{r}
-petersenlab::pB(pBgivenA = .95, pA = .285, pBgivenNotA = .007171515)
+petersenlab::pB(
+  pBgivenA = .95,
+  pA = .285,
+  pBgivenNotA = .007171515)
 ```
@@
 
-Using [Bayes's theorem](#sec-bayesTheorem) (described in @sec-bayesTheorem), solving for posttest odds (based on pretest odds and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), as in @eq-bayes5), and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
+Using [Bayes' theorem](#sec-bayesTheorem) (described in @sec-bayesTheorem), solving for posttest odds (based on pretest odds and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), as in @eq-bayes5), and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
 The calculation of posttest probability is described in INSERT.
@@
 
+People tend to ignore base rates ([base rate fallacy](#sec-fallaciesBaseRate)) when making predictions.
+They also tend to confuse inverse conditional probabilities ([conditional probability fallacy](#sec-inverseFallacy)).
+[Bayes' theorem](#sec-bayesTheorem) provides a way to convert from one conditional probability to its inverse conditional probability using the [base rate](#sec-baseRate) of each event.
+There are various ways to account for [base rates](#sec-baseRate) for more accurate predictions, including through the use of [actuarial formulas](#sec-accountForBaseRatesActuarial) and [Bayesian updating](#sec-bayesianUpdating).
+[Bayesian updating](#sec-bayesianUpdating) uses [Bayes' theorem](#sec-bayesTheorem) to calculate a posttest probability from a [pretest probability](#sec-baseRate) and a test result ([likelihood ratio](#sec-diagnosticLikelihoodRatio)).
+The [probability nomogram](#sec-probabilityNomogram) is a visual approach to [Bayesian updating](#sec-bayesianUpdating).
+
 ::: {.content-visible when-format="html"}

### 2024-07-14 cognitive-bias.qmd
20240913 - additions and clean up
20240713 - calibration and base rates

@@
 
-- availability
-- representativeness
-- anchoring and adjustment
+- availability heuristic
+- representativeness heuristic
+- anchoring and adjustment heuristic
 
@@
 *Overprecision* involves expressing excessive certainty regarding the accuracy of one's beliefs/judgments/predictions.
-For instance, if when a given weather forecaster says it will rain 80% of the time, it actually rains 30% of the time, the weather forecaster's predictions are overprecise.
+For instance, if when a given meteorologist says it will rain 80% of the time, it actually rains 30% of the time, the meteorologist's predictions are overprecise.
 Likewise, if the weather forecast says it will rain 10% of the time and it actually rains 30% of the time, the predictions are also overprecise because the forecaster is expressing stronger confidence than is warranted that it will not rain.

### 2024-07-14 evaluating-prediction-accuracy.qmd
20240713 - calibration and base rates

@@
   from = 0,
-  to = 150) %$% # from magrittr package
+  to = 150) %$% # exposition pipe magrittr::`%$%`
   data.frame(
@@
 
-[Discuss options for discrete versus continuous data]
+When dealing with a categorical outcome, calibration is the degree to which a probabilistic estimate of an event reflects the true underlying probability of the event.
+When dealing with a continuous outcome, calibration is the degree to which the predicted values are close in value to the outcome values.
+The importance of examining calibration (in addition to discrimination) is described by @Lindhiem2020.
+
+Calibration is relevant to all kinds of predictions, including weather forecasts.
+For instance, on the days that the meteorologist says there is a 60% chance of rain, it should rain about 60% of the time.
+Calibration is also important for fantasy football predictions.
+When projections state that a a group of players is each expected to score 200 points, their projections would be miscalibrated if those players scored only 150 points on average.
+
+Calibration can be examined in several ways, including Brier Scores, the Hosler-Lemeshow test, Spiegelhalter's $z$, and the mean difference between predicted and observed values at different binned thresholds as depicted graphically with a [calibration plot](#sec-calibrationPlot).
+
+#### Calibration Plot {#sec-calibrationPlot}
+
+Calibration plots can be helpful for identifying miscalibration.
+A calibration plot depicts the predicted probability of an event on the x-axis, and the actual (observed) probability of the event on the y-axis.
+The predictions are binned into a certain number of groups (commonly 10).
+The diagonal line reflects predictions that are perfectly calibrated.
+To the extent that predictions deviate from the diagonal line, the predictions are miscalibrated.
+
+There are four general patterns of miscalibration: overextremity, underextremity, overprediction, and underprediction (see @fig-miscalibration).
+*Overextremity* exists when the predicted probabilites are too close to the extremes (zero or one).
+*Underextremity* exists when the predicted probabilities are too far away from the extremes.
+*Overprediction* exists when the predicted probabilities are consistently greater than the observed probabilities.
+*Underprediction* exists when the predicted probabilities are consistently less than the observed probabilities.
+For a more thorough description of these types of miscalibration, see @Lindhiem2020.\index{calibration}
 

### 2024-07-14 base-rates.qmd
20240714 - cross-ref

@@
 
-Conversely (and rearranging the terms for the calculation of [conditional probability](#sec-conditionalProbability)), a [joint probability](#sec-jointProbability) can also be calculated using the [conditional probability](#sec-conditionalProbability) and [marginal probability](#sec-baseRate), as in @jointProbability2:
+Conversely (and rearranging the terms for the calculation of [conditional probability](#sec-conditionalProbability)), a [joint probability](#sec-jointProbability) can also be calculated using the [conditional probability](#sec-conditionalProbability) and [marginal probability](#sec-baseRate), as in @eq-jointProbability2:
 
@@
 Smaller values reflect greater accuracy, because it indicates that a false negative is less likely than a true negative.
-The formula for calculating the negative likelihood ratio is in @negativeLikelihoodRatio.
+The formula for calculating the negative likelihood ratio is in @eq-negativeLikelihoodRatio.
 
@@
 \end{aligned}
-$$ {#eq-positiveLikelihoodRatio}
+$$ {#eq-negativeLikelihoodRatio}
 

### 2024-07-14 index.qmd
20240706 - job title
20240713 - evaluating prediction accuracy
20240714 - subtitle and versioning

@@
 
-Petersen, I. T. (2024). *Fantasy football analytics*. Version 0.0.1. University of Iowa Libraries. <https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook>. [INSERT DOI LINK]
+Petersen, I.
+T.
+(2024).
+*Fantasy football analytics: Statistics, prediction, and empiricism Using R*.
+Version 0.0.1.
+University of Iowa Libraries.
+<https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook>.
+[INSERT DOI LINK]
 
@@
 @book{petersenFantasyFootballAnalytics,
-  title = {Fantasy football analytics},
+  title = {Fantasy football analytics: Statistics, prediction, and empiricism Using R},
   author = {Petersen, Isaac T.},
@@
 
-I am an Assistant Professor in the Department of Psychological and Brain Sciences at the University of Iowa.
+I am an Associate Professor in the Department of Psychological and Brain Sciences at the University of Iowa.
 I am a licensed psychologist with expertise in child clinical psychology.
@@
 In terms of my `R` and statistics background, I have published many [peer-reviewed publications](https://developmental-psychopathology.lab.uiowa.edu/publications) that employ advanced statistical methods, have published a [book on psychological assessment](https://www.routledge.com/9781032413068) [@Petersen2024a; @PetersenPrinciplesPsychAssessment] that includes applied examples in `R`, and have published the [`petersenlab` `R` package](https://cran.r-project.org/web/packages/petersenlab/index.html) [@R-petersenlab] on the Comprehensive R Archive Network (CRAN).
+Several sections in this book come from @PetersenPrinciplesPsychAssessment.
 I am also a co-author of the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) that provides free utilities for downloading fantasy football projections and additional fantasy-relevant data, and for calculating projected points given your league settings.

### 2024-07-14 evaluating-prediction-accuracy.qmd
20240713 - typo
20240714 - categorical

@@
 Some predictions involve categorical data, whereas other predictions involve continuous data.
-When dealing with a dichotomous (nominal data that are binary) predictor and outcome variable, we can evaluate predictions using a 2x2 table known as a confusion matrix (see INSERT), or with logistic regression models.
-When dealing with a continuous outcome variable (e.g., ordinal, interval, or ratio data), we can evaluate predictions using multiple regression or similar variants such as structural equation modeling and mixed models.
+When dealing with a dichotomous ([nominal data](#sec-nominal) that are binary) [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variable (or continuous data that have been dichotomized using a cutoff), we can evaluate predictions using a 2x2 table known as a confusion matrix (see INSERT), or with logistic regression models.
+When dealing with a continuous [outcome variable](#sec-correlationalStudy) (e.g., [ordinal](#sec-ordinal), [interval](#sec-interval), or [ratio](#sec-ratio) data), we can evaluate predictions using [multiple regression](#sec-multipleRegression) or similar variants such as structural equation modeling and mixed models.
+
+In fantasy football, we most commonly predict continuous [outcome variables](#sec-correlationalStudy) (e.g., fantasy points, rushing yards).
+Nevertheless, it is also important to understand principles in the prediction of categorical [outcomes variables](#sec-correlationalStudy).
 
@@
 For instance, it is very challenging to predict low [base rate](#sec-baseRate) behaviors, including suicide [@Kessler2020].
+For this reason, it is likely much more challenging to predict touchdowns—which happen relatively less often—than it is to predict passing/rushing/receiving yards—which are more frequent and continuously distributed.
+
+[EVALUATE EMPIRICALLY]
 
@@
 In sum, sensitivity and specificity differ based on the cutoff for classification.
-if we raise the cutoff, sensitivity and PPV increase (due to fewer false positives), whereas sensitivity and NPV decrease (due to more false negatives).
+If we raise the cutoff, sensitivity and PPV increase (due to fewer false positives), whereas sensitivity and NPV decrease (due to more false negatives).
 If we lower the cutoff, sensitivity and NPV increase (due to fewer false negatives), whereas specificity and PPV decrease (due to more false positives).
@@
 A medical application of SDT is tumor detection in radiology.
-SDT also plays an important role in psychology, especially cognitive psychology.
-For instance, research on social perception of sexual interest has shown that men tend to show lack of sensitivity to differences in women's affect—i.e., they have relative difficulties discriminating between friendliness and sexual interest [@Farris2008].
-Men also tend to show systematic bias (poor calibration) such that they tend to over-estimate women's sexual interest in them—i.e., men tend to have too low of a threshold for determining that a women is showing sexual interest in them [@Farris2006].
+Another application of SDT in society is using x-ray to detect bombs or other weapons.
+An example of applying SDT to fantasy football could be in the prediction (and evaluation) of whether or not a player scores a touchdown in a game.
 
@@
 
-## Prediction of Continuous Outcomes {#sec-predictionCategorical}
+## Prediction of Continuous Outcomes {#sec-predictionContinuous}
 

### 2024-07-15 base-rates.qmd
20240714 - accuracy indices

@@
 
-The positive likelihood ratio (LR+) compares the true positive rate to the false positive rate.
-Positive likelihood ratio values range from 1 to infinity.
-Higher values reflect greater accuracy, because it indicates the degree to which a true positive is more likely than a false positive.
+The positive likelihood ratio (LR+) compares the [true positive rate](#sec-sensitivity) to the [false positive rate](#sec-falsePositiveRate).
+Positive likelihood ratio values range from 1 to infinity.\index{positive likelihood ratio}
+Higher values reflect greater accuracy, because it indicates the degree to which a [true positive](#sec-truePositive) is more likely than a [false positive](#sec-falsePositive).
 The formula for calculating the positive likelihood ratio is in @eq-positiveLikelihoodRatio.
@@
 
-The negative likelihood ratio (LR−) compares the false negative rate to the true negative rate.
+The negative likelihood ratio (LR−) compares the [false negative rate](#sec-falseNegativeRate) to the [true negative rate](#sec-specificity).
 Negative likelihood ratio values range from 0 to 1.
-Smaller values reflect greater accuracy, because it indicates that a false negative is less likely than a true negative.
+Smaller values reflect greater accuracy, because it indicates that a [false negative](#sec-falseNegative) is less likely than a [true negative](#sec-trueNegative).
 The formula for calculating the negative likelihood ratio is in @eq-negativeLikelihoodRatio.

### 2024-07-15 evaluating-prediction-accuracy.qmd
20240714 - accuracy indices

@@
 
+### Accuracy Indices {#sec-accuracyIndicesCategorical}
+
+There are various accuracy indices we can use to evaluate the accuracy of predictions for categorical outcome variables.
+We have already described several accuracy indices, including percent accuracy, sensitivity, specificity, positive predictive value, negative predictive value, and area under the ROC curve.
+We describe these and other indices in greater detail below.
+
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `accuracyAtCutoff()` that computes many accuracy indices for the prediction of categorical outcome variables.
+
+```{r}
+#petersenlab::accuracyAtCutoff()
+```
+
+There are also test calculators available online:
+
+- http://araw.mede.uic.edu/cgi-bin/testcalc.pl
+- https://dlrs.shinyapps.io/shinyDLRs/
+
+### Confusion Matrix aka 2x2 Accuracy Table aka Cross-Tabulation aka Contingency Table {#sec-confusionMatrix}
+
+A confusion matrix (aka 2x2 accuracy table, cross-tabulation table, or contigency table) is a matrix for categorical data that presents the predicted outcome on one dimension and the actual outcome (truth) on the other dimension.
+If the predictions and outcomes are dichotomous, the confusion matrix is a 2x2 matrix with two rows and two columns that represent four possible predicted-actual combinations ([decision outcomes](#sec-decisionOutcomes)).
+In such a case, the confusion matrix provides a tabular count of each type of accurate cases ([true positives](#sec-truePositive) and [true negatives](#sec-trueNegative)) versus the number of each type of error ([false positives](#sec-falsePositive) and [false negatives](#sec-falseNegative)), as shown in INSERT.
+An example of a confusion matrix is in INSERT.
+
+#### Number {#sec-confusionMatrix-number}
+
+```{r}
+#table(mydata$diagnosisFactor, mydata$disorderFactor)
+```
+
+#### Number with margins added {#sec-confusionMatrix-numberMargins}
+
+```{r}
+#addmargins(table(mydata$diagnosisFactor, mydata$disorderFactor))
+```
+
+#### Proportions {#sec-confusionMatrix-numberMarginsProportions}
+
+```{r}
+#prop.table(table(mydata$diagnosisFactor, mydata$disorderFactor))
+```
+
+#### Proportions with margins added {#sec-confusionMatrix-numberMarginsProportionsMargins}
+
+```{r}
+#addmargins(prop.table(table(mydata$diagnosisFactor, mydata$disorderFactor)))
+```
+
+### True Positives (TP) {#sec-truePositive}
+
+True positives (TPs) are instances in which a positive classification (e.g., a disorder present) is correct—that is, the test says that a classification is present, and the classification is present.
+True positives are also called valid positives (VPs) or hits.
+Higher values reflect greater accuracy.
+The formula for true positives is in @eq-truePositive:
+
+$$
+\begin{aligned}
+  \text{TP} &= \text{BR} \times \text{SR} \times N
+\end{aligned}
+$$ {#eq-truePositive}
+
+### True Negatives (TN) {#sec-trueNegative}
+
+True negatives (TNs) are instances in which a negative classification (e.g., absence of a disorder) is correct—that is, the test says that a classification is not present, and the classification is actually not present.
+True negatives are also called valid negatives (VNs) or correct rejections.
+Higher values reflect greater accuracy.
+The formula for true negatives is in @eq-trueNegative:
+
+$$
+\begin{aligned}
+  \text{TN} &= (1 - \text{BR}) \times (1 - \text{SR}) \times N
+\end{aligned}
+$$ {#eq-trueNegative}
+
+### False Positives (FP) {#sec-falsePositive}
+
+False positives (FPs) are instances in which a positive classification (e.g., a disorder present) is incorrect—that is, the test says that a classification is present, and the classification is not present.
+False positives are also called false alarms (FAs).
+Lower values reflect greater accuracy.
+The formula for false positives is in Equation @eq-falsePositive:
+
+$$
+\begin{aligned}
+  \text{FP} &= (1 - \text{BR}) \times \text{SR} \times N
+\end{aligned}
+$$ {#eq-falsePositive}
+
+### False Negatives (FN) {#sec-falseNegative}
+
+False negatives (FNs) are instances in which a negative classification (e.g., absence of a disorder) is incorrect—that is, the test says that a classification is not present, and the classification is present.
+False negatives are also called misses.
+Lower values reflect greater accuracy.
+The formula for false negatives is in @eq-falseNegative:
+
+$$
+\begin{aligned}
+  \text{FN} &= \text{BR} \times (1 - \text{SR}) \times N
+\end{aligned}
+(\#eq:falseNegative)
+$$ {#eq-falseNegative}
+
+### Selection Ratio (SR) {#sec-selectionRatio}
+
+The selection ratio (SR) is the marginal probability of selection, independent of other things: $P(R_i)$.
+In clinical psychology, the selection ratio is the proportion of people who test positive for the disorder, as in @eq-selectionRatio:
+
+$$
+\begin{aligned}
+  \text{SR} &= P(R_i) \\
+  &= \frac{\text{TP} + \text{FP}}{N}
+\end{aligned}
+(\#eq:selectionRatio)
+$$ {#eq-selectionRatio}
+
+### Base Rate (BR) {#sec-pretestProbability}
+
+The [base rate](#sec-baseRate) (BR) of a classification is its [marginal probability](#sec-baseRate), independent of other things: $P(C_i)$.
+In clinical psychology, the base rate of a disorder is its prevalence in the population, as in @eq-baseRate.
+Without additional information, the [base rate](#sec-baseRate) is used as the initial *pretest probability*.
+
+$$
+\begin{aligned}
+  \text{BR} &= P(C_i) \\
+  &= \frac{\text{TP} + \text{FN}}{N}
+\end{aligned}
+$$ {#eq-baseRate}
+
+### Pretest Odds {#sec-pretestOdds}
+
+The pretest odds of a classification can be estimated using the pretest probability (i.e., [base rate](#sec-baseRate)).
+To convert a probability to odds, divide the probability by one minus that probability, as in @eq-pretestOdds.
+
+$$
+\begin{aligned}
+  \text{pretest odds} &= \frac{\text{pretest probability}}{1 - \text{pretest probability}} \\
+\end{aligned}
+(\#eq:pretestOdds)
+$$ {#eq-pretestOdds}
+
+### Percent Accuracy {#sec-percentAccuracy}
+
+Percent Accuracy is also called overall accuracy.
+Higher values reflect greater accuracy.
+The formula for percent accuracy is in @eq-percentAccuracy.
+Percent accuracy has several problems.
+First, it treats all errors ([FP](#sec-falsePositive) and [FN](#sec-falseNegative)) as equally important.
+However, in practice, it is rarely the case that [false positives](#sec-falsePositive) and [false negatives](#sec-falseNegative) are equally important.
+Second, percent accuracy can be misleading because it is highly influenced by [base rates](#sec-baseRate).
+You can have a high percent accuracy by predicting from the [base rate](#sec-baseRate) and saying that no one has the characteristic (if the [base rate](#sec-baseRate) is low) or that everyone has the characteristic (if the [base rate](#sec-baseRate) is high).
+Thus, it is also important to consider other aspects of accuracy.
+
+$$
+\text{Percent Accuracy} = 100\% \times \frac{\text{TP} + \text{TN}}{N}
+$$ {#eq-percentAccuracy}
+
+### Percent Accuracy by Chance {#sec-percentAccuracyByChance}
+
+The formula for calculating percent accuracy by chance is in @eq-percentAccuracyByChance.
+
+$$
+\begin{aligned}
+  \text{Percent Accuracy by Chance} &= 100\% \times [P(\text{TP}) + P(\text{TN})] \\
+  &= 100\% \times \{(\text{BR} \times {\text{SR}}) + [(1 - \text{BR}) \times (1 - \text{SR})]\}
+\end{aligned}
+$$ {#eq-percentAccuracyByChance}
+
+### Percent Accuracy Predicting from the Base Rate {#sec-percentAccuracyPredictingFromBaseRate}
+
+[*Predicting from the base rate*](#sec-predictingFromBaseRate) is going with the most likely outcome in every prediction.
+If the [base rate](#sec-baseRate) is less than .50, it would involve predicting that the condition is absent for every case.
+If the [base rate](#sec-baseRate) is .50 or above, it would involve predicting that the condition is present for every case.
+[Predicting from the base rate](#sec-predictingFromBaseRate) is a special case of [percent accuracy by chance](#sec-percentAccuracyByChance) when the [selection ratio](#sec-selectionRatio) is set to either one (if the [base rate](#sec-baseRate) $\geq$ .5) or zero (if the [base rate](#sec-baseRate) < .5).
+
+### Relative Improvement Over Chance (RIOC) {#sec-relativeImprovementOverChance}
+
+Relative improvement over chance (RIOC) is a prediction's improvement over chance as a proportion of the maximum possible improvement over chance, as described by @Farrington1989.
+Higher values reflect greater accuracy.
+The formula for calculating RIOC is in @eq-relativeImprovementOverChance.
+
+$$
+\begin{aligned}
+  \text{relative improvement over chance (RIOC)} &= \frac{\text{total correct} - \text{chance correct}}{\text{maximum correct} - \text{chance correct}} \\
+\end{aligned}
+$$ {#eq-relativeImprovementOverChance}
+
+### Relative Improvement Over Predicting from the Base Rate {#sec-relativeImprovementOverPredictingFromBaseRate}
+
+Relative improvement over [predicting from the base rate](#sec-predictingFromBaseRate) is a prediction's improvement over [predicting from the base rate](#sec-predictingFromBaseRate) as a proportion of the maximum possible improvement over [predicting from the base rate](#sec-predictingFromBaseRate).
+Higher values reflect greater accuracy.
+The formula for calculating relative improvement over predicting from the base rate is in @eq-relativeImprovementOverPredictingFromBaseRate.
+
+$$
+\scriptsize
+\begin{aligned}
+  \text{relative improvement over predicting from base rate} &= \frac{\text{total correct} - \text{correct by predicting from base rate}}{\text{maximum correct} - \text{correct by predicting from base rate}} \\
+\end{aligned}
+$$ {#eq-relativeImprovementOverPredictingFromBaseRate}
+
+### Sensitivity (SN) {#sec-sensitivity}
+
+Sensitivity (SN) is also called true positive rate (TPR), hit rate (HR), or recall.
+Sensitivity is the [conditional probability](#sec-conditionalProbability) of a positive test given that the person has the condition: $P(R|C)$.
+Higher values reflect greater accuracy.
+The formula for calculating sensitivity is in @eq-sensitivity.
+As described in Section @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), sensitivity decreases.
+As the cutoff decreases, sensitivity increases.
+
+$$
+\begin{aligned}
+  \text{sensitivity (SN)} &= P(R|C) \\
+  &= \frac{\text{TP}}{\text{TP} + \text{FN}} = \frac{\text{TP}}{N \times \text{BR}} = 1 - \text{FNR}
+\end{aligned}
+$$ {#eq-sensitivity}
+
+### Specificity (SP) {#sec-specificity}
+
+Specificity (SP) is also called true negative rate (TNR) or selectivity.
+Specificity is the [conditional probability](#sec-conditionalProbability) of a negative test given that the person does not have the condition: $P(\text{not } R|\text{not } C)$.
+Higher values reflect greater accuracy.
+The formula for calculating specificity is in @eq-specificity.
+As described in Section @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), specificity increases.
+As the cutoff decreases, specificity decreases.
+
+$$
+\begin{aligned}
+  \text{specificity (SP)} &= P(\text{not } R|\text{not } C) \\
+  &= \frac{\text{TN}}{\text{TN} + \text{FP}} = \frac{\text{TN}}{N (1 - \text{BR})} = 1 - \text{FPR}
+\end{aligned}
+$$ {#eq-specificity}
+
+### False Negative Rate (FNR) {#sec-falseNegativeRate}
+
+The false negative rate (FNR) is also called the miss rate.
+The false negative rate is the [conditional probability](#sec-conditionalProbability) of a negative test given that the person has the condition: $P(\text{not } R|C)$.
+Lower values reflect greater accuracy.
+The formula for calculating false negative rate is in @eq-falseNegativeRate.
+
+$$
+\begin{aligned}
+  \text{false negative rate (FNR)} &= P(\text{not } R|C) \\
+  &= \frac{\text{FN}}{\text{FN} + \text{TP}} = \frac{\text{FN}}{N \times \text{BR}} = 1 - \text{TPR}
+\end{aligned}
+$$ {#eq-falseNegativeRate}
+
+### False Positive Rate (FPR) {#sec-falsePositiveRate}
+
+The false positive rate (FPR) is also called the false alarm rate (FAR) or fall-out.
+The false positive rate is the [conditional probability](#sec-conditionalProbability) of a positive test given that the person does not have the condition: $P(R|\text{not } C)$.
+Lower values reflect greater accuracy.
+The formula for calculating false positive rate is in @eq-falsePositiveRate:
+
+$$
+\begin{aligned}
+  \text{false positive rate (FPR)} &= P(R|\text{not } C) \\
+  &= \frac{\text{FP}}{\text{FP} + \text{TN}} = \frac{\text{FP}}{N (1 - \text{BR})} = 1 - \text{TNR}
+\end{aligned}
+(\#eq:falsePositiveRate)
+$$ {#eq-falsePositiveRate}
+
+### Positive Predictive Value (PPV) {#sec-ppv}
+
+The positive predictive value (PPV) is also called the positive predictive power (PPP) or precision.
+Many people confuse [sensitivity](#sec-sensitivity) ($P(R|C)$) with its inverse [conditional probability](#sec-conditionalProbability), PPV ($P(C|R)$).
+PPV is the [conditional probability](#sec-conditionalProbability) of having the condition given a positive test: $P(C|R)$.
+Higher values reflect greater accuracy.
+The formula for calculating positive predictive value is in @eq-positivePredictiveValue.
+
+PPV can be low even when [sensitivity](#sec-sensitivity) is high because it depends not only on [sensitivity](#sec-sensitivity), but also on [specificity](#sec-specificity) and the [base rate](#sec-baseRate).
+Because PPV depends on the [base rate](#sec-baseRate), PPV is not an intrinsic property of a measure.
+The same measure will have a different PPV in different contexts with different [base rates](#sec-baseRate) [@Treat2023].
+As described in Section @sec-accuracyCutoff, as the [base rate](#sec-baseRate) increases, PPV increases.
+As the [base rate](#sec-baseRate) decreases, PPV decreases.
+PPV also differs as a function of the cutoff.
+As described in Section @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), PPV increases.
+As the cutoff decreases (becomes more liberal), PPV decreases.
+
+$$
+\small
+\begin{aligned}
+  \text{positive predictive value (PPV)} &= P(C|R) \\
+  &= \frac{\text{TP}}{\text{TP} + \text{FP}} = \frac{\text{TP}}{N \times \text{SR}}\\
+  &= \frac{\text{sensitivity} \times {\text{BR}}}{\text{sensitivity} \times {\text{BR}} + [(1 - \text{specificity}) \times (1 - \text{BR})]}
+\end{aligned}
+$$ {#eq-positivePredictiveValue}
+
+### Negative Predictive Value (NPV) {#sec-npv}
+
+The negative predictive value (NPV) is also called the negative predictive power (NPP).
+Many people confuse [specificity](#sec-specificity) ($P(\text{not } R|\text{not } C)$) with its inverse [conditional probability](#sec-conditionalProbability), NPV ($P(\text{not } C| \text{not } R)$).
+NPV is the [conditional probability](#sec-conditionalProbability) of not having the condition given a negative test: $P(\text{not } C| \text{not } R)$.
+Higher values reflect greater accuracy.
+The formula for calculating negative predictive value is in @eq-negativePredictiveValue.
+
+NPV can be low even when [specificity](#sec-specificity) is high because it depends not only on [specificity](#sec-specificity), but also on [sensitivity](#sec-sensitivity) and the [base rate](#sec-baseRate).
+Because NPV depends on the [base rate](#sec-baseRate), NPV is not an intrinsic property of a measure.
+The same measure will have a different NPV in different contexts with different [base rates](#sec-baseRate) [@Treat2023].
+As described in Section @sec-accuracyCutoff, as the [base rate](#sec-baseRate) increases, NPV decreases.
+As the [base rate](#sec-baseRate) decreases, NPV increases.
+NPV also differs as a function of the cutoff.
+As described in Section @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), NPV decreases.
+As the cutoff decreases (becomes more liberal), NPV decreases.
+
+$$
+\small
+\begin{aligned}
+  \text{negative predictive value (NPV)} &= P(\text{not } C|\text{not } R) \\
+  &= \frac{\text{TN}}{\text{TN} + \text{FN}} = \frac{\text{TN}}{N(\text{1 - SR})}\\
+  &= \frac{\text{specificity} \times (1-{\text{BR}})}{\text{specificity} \times (1-{\text{BR}}) + [(1 - \text{sensitivity}) \times \text{BR})]}
+\end{aligned}
+$$ {#eq-negativePredictiveValue}
+
+### False Discovery Rate (FDR) {#sec-falseDiscoveryRate}
+
+Many people confuse the false positive rate ($P(R|\text{not } C)$) with its inverse [conditional probability](#sec-conditionalProbability), the false discovery rate ($P(\text{not } C| R)$).
+The false discovery rate (FDR) is the [conditional probability](#sec-conditionalProbability) of not having the condition given a positive test: $P(\text{not } C| R)$.
+Lower values reflect greater accuracy.
+The formula for calculating false discovery rate is in @eq-falseDiscoveryRate.
+
+$$
+\begin{aligned}
+  \text{false discovery rate (FDR)} &= P(\text{not } C|R) \\
+  &= \frac{\text{FP}}{\text{FP} + \text{TP}} = 1 - \text{PPV}
+\end{aligned}
+$$ {#eq-falseDiscoveryRate}
+
+### False Omission Rate (FOR) {#sec-falseOmissionRate}
+
+Many people confuse the false negative rate ($P(\text{not } R|C)$) with its inverse [conditional probability](#sec-conditionalProbability), the false omission rate ($P(C|\text{not } R)$).
+The false omission rate (FOR) is the conditional probability of having the condition given a negative test: $P(C|\text{not } R)$.
+Lower values reflect greater accuracy.
+The formula for calculating false omission rate is in @sec-falseOmissionRate.
+
+$$
+\begin{aligned}
+  \text{false omission rate (FOR)} &= P(C|\text{not } R) \\
+  &= \frac{\text{FN}}{\text{FN} + \text{TN}} = 1 - \text{NPV}
+\end{aligned}
+$$ {#sec-falseOmissionRate}
+
+### Youden's J Statistic {#sec-youdenJ-example}
+
+Youden's J statistic is also called Youden's Index or informedness.
+Youden's J statistic is the sum of [sensitivity](#sec-sensitivity) and [specificity](#sec-specificity) (and subtracting one).
+Higher values reflect greater accuracy.
+The formula for calculating Youden's J statistic is in @eq-youdenIndex.
+
+$$
+\begin{aligned}
+  \text{Youden's J statistic} &= \text{sensitivity} + \text{specificity} - 1
+\end{aligned}
+$$ {#eq-youdenIndex}
+
+### Balanced Accuracy {#sec-balancedAccuracy}
+
+Balanced accuracy is the average of [sensitivity](#sec-sensitivity) and [specificity](#sec-specificity).
+Higher values reflect greater accuracy.
+The formula for calculating balanced accuracy is in @balancedAccuracy.
+
+$$
+\begin{aligned}
+  \text{balanced accuracy} &= \frac{\text{sensitivity} + \text{specificity}}{2}
+\end{aligned}
+$$ {#eq-balancedAccuracy}
+
+### F-Score {#sec-fScore}
+
+The F-score combines [precision](#sec-ppv) ([positive predictive value](#sec-ppv)) and [recall](#sec-sensitivity) ([sensitivity](#sec-sensitivity)), where $\beta$ indicates how many times more important [sensitivity](#sec-sensitivity) is than the [positive predictive value](#sec-ppv).
+If [sensitivity](#sec-sensitivity) and the [positive predictive value](#sec-ppv) are equally important, $\beta = 1$, and the F-score is called the $F_1$ score.
+Higher values reflect greater accuracy.
+The formula for calculating the F-score is in @eq-FScore.
+
+$$
+\begin{aligned}
+  F_\beta &= (1 + \beta^2) \cdot \frac{\text{positive predictive value} \cdot \text{sensitivity}}{(\beta^2 \cdot \text{positive predictive value}) + \text{sensitivity}} \\
+  &= \frac{(1 + \beta^2) \cdot \text{TP}}{(1 + \beta^2) \cdot \text{TP} + \beta^2 \cdot \text{FN} + \text{FP}}
+\end{aligned}
+$$ {#eq-FScore}
+
+The formula for calculating the $F_1$ score is in @eq-F1Score.
+
+$$
+\begin{aligned}
+  F_1 &= \frac{2 \cdot \text{positive predictive value} \cdot \text{sensitivity}}{(\text{positive predictive value}) + \text{sensitivity}} \\
+  &= \frac{2 \cdot \text{TP}}{2 \cdot \text{TP} + \text{FN} + \text{FP}}
+\end{aligned}
+$$ {#eq-F1Score}
+
+### Matthews Correlation Coefficient (MCC) {#sec-matthewsCorrelationCoefficient}
+
+The Matthews correlation coefficient (MCC) is also called the phi coefficient.
+It is a correlation coefficient between predicted and observed values from a binary classification.
+Higher values reflect greater accuracy.
+The formula for calculating the MCC is in @eq-matthewsCorrelationCoefficient.
+
+$$
+\begin{aligned}
+  \text{MCC} &= \frac{\text{TP} \times \text{TN} - \text{FP} \times \text{FN}}{\sqrt{(\text{TP} + \text{FP})(\text{TP} + \text{FN})(\text{TN} + \text{FP})(\text{TN} + \text{FN})}}
+\end{aligned}
+$$ {#eq-matthewsCorrelationCoefficient}
+
+### Diagnostic Odds Ratio {#sec-diagnosticOddsRatio}
+
+The diagnostic odds ratio is the odds of a positive test among people with the condition relative to the odds of a positive test among people without the condition.
+Higher values reflect greater accuracy.
+The formula for calculating the diagnostic odds ratio is in @eq-diagnosticOddsRatio.
+If the predictor is bad, the diagnostic odds ratio could be less than one, and values can go up from there.
+If the diagnostic odds ratio is greater than 2, we take the odds ratio seriously because we are twice as likely to predict accurately than inaccurately.
+However, the diagnostic odds ratio ignores/hides [base rates](#sec-baseRate).
+When interpreting the diagnostic odds ratio, it is important to keep in mind the [practical significance](#sec-practicalSignificance), because otherwise it is not very meaningful.
+Consider a risk factor that has a diagnostic odds ratio of 3 for tuberculosis, i.e., it puts you at 3 times as likely to develop tuberculosis.
+The prevalence of tuberculosis is relatively low.
+Assuming the prevalence of tuberculosis is less than 1/10th of 1%, your risk of developing tuberculosis is still very low even if the risk factor (with a diagnostic odds ratio of 3) is present.
+
+$$
+\begin{aligned}
+  \text{diagnostic odds ratio} &= \frac{\text{TP} \times \text{TN}}{\text{FP} \times \text{FN}} \\
+  &= \frac{\text{sensitivity} \times \text{specificity}}{(1 - \text{sensitivity}) \times (1 - \text{specificity})} \\
+  &= \frac{\text{PPV} \times \text{NPV}}{(1 - \text{PPV}) \times (1 - \text{NPV})} \\
+  &= \frac{\text{LR+}}{\text{LR}-}
+\end{aligned}
+$$ {#eq-diagnosticOddsRatio}
+
+### Diagnostic Likelihood Ratio {#sec-diagnosticLikelihoodRatio2}
+
+The diagnostic likelihood ratio is described in @sec-diagnosticLikelihoodRatio.
+There are two types of diagnostic likelihood ratios: the [positive likelihood ratio](#sec-positiveLikelihoodRatio) and the [negative likelihood ratio](#sec-negativeLikelihoodRatio).
+
+#### Positive Likelihood Ratio (LR+) {#sec-positiveLikelihoodRatio2}
+
+The positive likelihood ratio (LR+) is described in @sec-positiveLikelihoodRatio.
+The formula for calculating the positive likelihood ratio is in @eq-positiveLikelihoodRatio.
+
+#### Negative Likelihood Ratio (LR−) {#sec-negativeLikelihoodRatio2}
+
+The negative likelihood ratio (LR−) is described in @sec-negativeLikelihoodRatio.
+The formula for calculating the negative likelihood ratio is in @eq-positiveLikelihoodRatio.
+
+### Posttest Odds {#sec-posttestOdds}
+
+As presented in @eq-bayes5, the posttest (or posterior) odds are equal to the [pretest odds](#sec-pretestOdds) multiplied by the [likelihood ratio](#sec-diagnosticLikelihoodRatio).
+The posttest odds and [posttest probability](#sec-posttestProbability) can be useful to calculate when the [pretest probability](#sec-baseRate) is different from the [pretest probability](#sec-baseRate) (or prevalence) of the classification.
+For instance, you might use a different [pretest probability](#sec-baseRate) if a test result is already known and you want to know the updated [posttest probability](#sec-posttestProbability) after conducting a second test.
+The formula for calculating posttest odds is in @eq-posttestOdds.
+
+$$
+\begin{aligned}
+  \text{posttest odds} &= \text{pretest odds} \times \text{likelihood ratio} \\
+\end{aligned}
+$$ {#eq-posttestOdds}
+
+For calculating the posttest odds of a [true positive](#sec-truePositive) compared to a [false positive](#sec-falsePositive), we use the [positive likelihood ratio](#sec-positiveLikelihoodRatio) below.
+We would use the [negative likelihood ratio](#sec-negativeLikelihoodRatio) if we wanted to calculate the posttest odds of a [false negative](#sec-falseNegative) compared to a [true negative](#sec-trueNegative).
+
+### Posttest Probability {#sec-posttestProbability}
+
+The posttest probability is the probability of having the disorder given a test result.
+When the [base rate](#sec-baseRate) is used as the [pretest probability](#sec-baseRate), the posttest probability given a positive test is equal to [positive predictive value](#sec-ppv).
+To convert odds to a probability, divide the odds by one plus the odds, as is in @eq-posttestProbability.
+
+$$
+\begin{aligned}
+  \text{posttest probability} &= \frac{\text{posttest odds}}{1 + \text{posttest odds}}
+\end{aligned}
+$$ {#eq-posttestProbability}
+
 ## Prediction of Continuous Outcomes {#sec-predictionContinuous}

### 2024-07-15 evaluating-prediction-accuracy.qmd
20240714 - edits

@@
 
-- http://araw.mede.uic.edu/cgi-bin/testcalc.pl
-- https://dlrs.shinyapps.io/shinyDLRs/
-
-### Confusion Matrix aka 2x2 Accuracy Table aka Cross-Tabulation aka Contingency Table {#sec-confusionMatrix}
+- <http://araw.mede.uic.edu/cgi-bin/testcalc.pl>
+- <https://dlrs.shinyapps.io/shinyDLRs>
+
+#### Confusion Matrix aka 2x2 Accuracy Table aka Cross-Tabulation aka Contingency Table {#sec-confusionMatrix}
 
@@
 
-#### Number {#sec-confusionMatrix-number}
+##### Number {#sec-confusionMatrix-number}
 
@@
 
-#### Number with margins added {#sec-confusionMatrix-numberMargins}
+##### Number with margins added {#sec-confusionMatrix-numberMargins}
 
@@
 
-#### Proportions {#sec-confusionMatrix-numberMarginsProportions}
+##### Proportions {#sec-confusionMatrix-numberMarginsProportions}
 
@@
 
-#### Proportions with margins added {#sec-confusionMatrix-numberMarginsProportionsMargins}
+##### Proportions with margins added {#sec-confusionMatrix-numberMarginsProportionsMargins}
 
@@
 
-### True Positives (TP) {#sec-truePositive}
+#### True Positives (TP) {#sec-truePositive}
 
@@
 
-### True Negatives (TN) {#sec-trueNegative}
+#### True Negatives (TN) {#sec-trueNegative}
 
@@
 
-### False Positives (FP) {#sec-falsePositive}
+#### False Positives (FP) {#sec-falsePositive}
 
@@
 
-### False Negatives (FN) {#sec-falseNegative}
+#### False Negatives (FN) {#sec-falseNegative}
 
@@
 \end{aligned}
-(\#eq:falseNegative)
 $$ {#eq-falseNegative}
 
-### Selection Ratio (SR) {#sec-selectionRatio}
+#### Selection Ratio (SR) {#sec-selectionRatio}
 
@@
 \end{aligned}
-(\#eq:selectionRatio)
 $$ {#eq-selectionRatio}
 
-### Base Rate (BR) {#sec-pretestProbability}
+#### Base Rate (BR) {#sec-pretestProbability}
 
@@
 
-### Pretest Odds {#sec-pretestOdds}
+#### Pretest Odds {#sec-pretestOdds}
 
@@
 \end{aligned}
-(\#eq:pretestOdds)
 $$ {#eq-pretestOdds}
 
-### Percent Accuracy {#sec-percentAccuracy}
+#### Percent Accuracy {#sec-percentAccuracy}
 
@@
 
-### Percent Accuracy by Chance {#sec-percentAccuracyByChance}
+#### Percent Accuracy by Chance {#sec-percentAccuracyByChance}
 
@@
 
-### Percent Accuracy Predicting from the Base Rate {#sec-percentAccuracyPredictingFromBaseRate}
+#### Percent Accuracy Predicting from the Base Rate {#sec-percentAccuracyPredictingFromBaseRate}
 
@@
 
-### Relative Improvement Over Chance (RIOC) {#sec-relativeImprovementOverChance}
+#### Relative Improvement Over Chance (RIOC) {#sec-relativeImprovementOverChance}
 
@@
 
-### Relative Improvement Over Predicting from the Base Rate {#sec-relativeImprovementOverPredictingFromBaseRate}
+#### Relative Improvement Over Predicting from the Base Rate {#sec-relativeImprovementOverPredictingFromBaseRate}
 
@@
 
-### Sensitivity (SN) {#sec-sensitivity}
+#### Sensitivity (SN) {#sec-sensitivity}
 
@@
 
-### Specificity (SP) {#sec-specificity}
+#### Specificity (SP) {#sec-specificity}
 
@@
 
-### False Negative Rate (FNR) {#sec-falseNegativeRate}
+#### False Negative Rate (FNR) {#sec-falseNegativeRate}
 
@@
 
-### False Positive Rate (FPR) {#sec-falsePositiveRate}
+#### False Positive Rate (FPR) {#sec-falsePositiveRate}
 
@@
 \end{aligned}
-(\#eq:falsePositiveRate)
 $$ {#eq-falsePositiveRate}
 
-### Positive Predictive Value (PPV) {#sec-ppv}
+#### Positive Predictive Value (PPV) {#sec-ppv}
 
@@
 
-### Negative Predictive Value (NPV) {#sec-npv}
+#### Negative Predictive Value (NPV) {#sec-npv}
 
@@
 
-### False Discovery Rate (FDR) {#sec-falseDiscoveryRate}
+#### False Discovery Rate (FDR) {#sec-falseDiscoveryRate}
 
@@
 
-### False Omission Rate (FOR) {#sec-falseOmissionRate}
+#### False Omission Rate (FOR) {#sec-falseOmissionRate}
 
@@
 
-### Youden's J Statistic {#sec-youdenJ-example}
+#### Youden's J Statistic {#sec-youdenJ-example}
 
@@
 
-### Balanced Accuracy {#sec-balancedAccuracy}
+#### Balanced Accuracy {#sec-balancedAccuracy}
 
@@
 Higher values reflect greater accuracy.
-The formula for calculating balanced accuracy is in @balancedAccuracy.
+The formula for calculating balanced accuracy is in @eq-balancedAccuracy.
 
@@
 
-### F-Score {#sec-fScore}
+#### F-Score {#sec-fScore}
 
@@
 
-### Matthews Correlation Coefficient (MCC) {#sec-matthewsCorrelationCoefficient}
+#### Matthews Correlation Coefficient (MCC) {#sec-matthewsCorrelationCoefficient}
 
@@
 
-### Diagnostic Odds Ratio {#sec-diagnosticOddsRatio}
+#### Diagnostic Odds Ratio {#sec-diagnosticOddsRatio}
 
@@
 
-### Diagnostic Likelihood Ratio {#sec-diagnosticLikelihoodRatio2}
+#### Diagnostic Likelihood Ratio {#sec-diagnosticLikelihoodRatio2}
 
@@
 
-#### Positive Likelihood Ratio (LR+) {#sec-positiveLikelihoodRatio2}
+##### Positive Likelihood Ratio (LR+) {#sec-positiveLikelihoodRatio2}
 
@@
 
-#### Negative Likelihood Ratio (LR−) {#sec-negativeLikelihoodRatio2}
+##### Negative Likelihood Ratio (LR−) {#sec-negativeLikelihoodRatio2}
 
@@
 
-### Posttest Odds {#sec-posttestOdds}
+#### Posttest Odds {#sec-posttestOdds}
 
@@
 
-### Posttest Probability {#sec-posttestProbability}
+#### Posttest Probability {#sec-posttestProbability}
 

### 2024-07-15 evaluating-prediction-accuracy.qmd
20240715 - ways to improve predictions

@@
 
+## Ways to Improve Prediction Accuracy {#waysToImprovePredictionAccuracy}
+
+On the whole, experts' predictions are inaccurate.
+Experts' predictions from many different domains tend to be inaccurate, including political scientists [@Tetlock2017], physicians [@Koehler2002], clinical psychologists [@Oskamp1965], stock market traders and corporate financial officers [@Skala2008], seismologists' predictions of earthquakes [@Hough2016], economists' predictions about the economy [@Makridakis2009], lawyers [@Koehler2002], and business managers [@Russo1992].
+The most common pattern of experts' predictions is that they show overextremity, that is, their predictions have probability judgments that tend to be too extreme, as described in Section @sec-calibration.
+Overextremity of experts' predictions reflects the [overprecision](#sec-cognitiveBiasesOverconfidence) type of [overconfidence bias](#sec-cognitiveBiasesOverconfidence).
+The degree of confidence of a person's predictions is often not a good indicator of the accuracy of their predictions [and confidence and prediction accuracy are sometimes inversely associated; @Silver2012].
+[Heuristics](#sec-heuristics) such as the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment), cognitive biases such as [confirmation bias](#sec-cognitiveBiasesConfirmation) [@Hoch1985; @Koriat1980], fallacies such as the [base rate fallacy](#sec-fallaciesBaseRate) [@Eddy1982; @Koehler2002] could contribute to overconfidence of predictions.
+[Poorly calibrated](#sec-calibration) predictions are especially likely when the [base rate](#sec-baseRate) is very low (e.g., suicide) or when the [base rate](#sec-baseRate) is very high [@Koehler2002].
+
+Nevertheless, there are some domains that have shown greater predictive accuracy, from which we may learn what practices may lead to greater accuracy.
+For instance, experts have shown stronger predictive accuracy in weather forecasting [@Murphy1984], horse race betting [@Johnson2001], and playing the card game of bridge [@Keren1987], but see @Koehler2002 for exceptions.
+
+Here are some potential ways to improve the accuracy (and honesty) of predictions and judgments:
+
+- Provide appropriate [anchoring of your predictions to the base rate](#sec-accountForBaseRates) of the phenomenon you are predicting.
+To the extent that the [base rate](#sec-baseRate) of the event you are predicting is low, more extreme evidence should be necessary to consistently and accurately predict that the event will occur.
+Applying [actuarial formulas](#sec-actuarialPrediction) and [Bayes' theorem](#sec-bayesTheorem) can help you appropriately weigh the [base rate](#sec-baseRate) and evidence.
+- Include multiple predictors, ideally from different measures and measurement methods.
+Include the predictors with the strongest validity based on theory of the causal process and based on [criterion-related validity](#sec-criterionValidity).
+- When possible, aggregate multiple perspectives of predictions, especially predictions made independently (from different people/methods/etc.).
+The "wisdom of the crowd" is often more accurate than individuals' predictions, including predictions by so-called "experts" [@Silver2012].
+- A goal of prediction is to capture as much signal as possible and as little noise (error) as possible [@Silver2012].
+Parsimony (i.e., not having too many predictors) can help reduce the amount of error variance captured by the prediction model.
+However, to accurately model complex systems like human behavior, complex models may be necessary.
+However, strong theory of the causal processes and dynamics may be necessary to develop accurate complex models.
+- Although incorporating theory can be helpful, provide more weight to empiricism than to theory, until our theories and measures are stronger.
+Ideally, we would use theory to design a model that mirrors the causal system, with accurate measures of each process in the system, so we could make accurate predictions.
+However, as described in Section \@ref(theoryEmpiricism), our psychological theories of the causal processes that influence behavior are not yet very strong.
+Until we have stronger theories that specify the causal process for a given outcome, and until we have accurate measures of those causal processes, [actuarial approaches](#sec-actuarialPrediction) are likely to be most accurate, as discussed in @sec-judgmentVsActuarial.
+At the same time, keep in mind that measures involving human behavior, and their resulting data, are often noisy.
+As a result, theoretically (conceptually) informed empirical approaches may lead to more accuracy than empiricism alone.
+- Use an empirically validated and cross-validated [statistical algorithm](#sec-actuarialPrediction) to combine information from the predictors in a formalized way.
+Give each predictor appropriate weight in the statistical algorithm, according to its strength of association with the outcome.
+Use measures with strong [reliability](#sec-reliability) and [validity](#sec-validity) for assessing these processes to be used in the algorithm.
+Cross-validation will help reduce the likelihood that your model is fitting to noise and will maximize the likelihood that the model predicts accurately when applied to new data (i.e., the model's predictions accurately generalize), as described in @sec-bestActuarialApproaches.
+- When presenting your predictions, acknowledge what you do not know.
+- Express your predictions in terms of probabilistic estimates and present the uncertainty in your predictions with confidence intervals [even though bolder, more extreme predictions tend to receive stronger television ratings; @Silver2012].
+- Qualify your predictions by identifying and noting counter-examples that would not be well fit by your prediction model, such as extreme cases, edge cases, and "broken leg" [@Meehl1957] cases.
+- Provide clear, consistent, and timely feedback on the outcomes of the predictions to the people making the predictions [@Bolger2004].
+- Be self-critical about your predictions.
+Update your judgments based on their accuracy, rather than trying to confirm your beliefs [@Atanasov2020].
+- In addition to considering the accuracy of the prediction, consider the quality of the prediction *process*, especially when random chance is involved to a degree (such as in poker) [@Silver2012].
+- Work to identify and mitigate potential blindspots; be aware of cognitive biases and fallacies, such as [confirmation bias](#sec-cognitiveBiasesConfirmation) and the [base rate fallacy](#sec-fallaciesBaseRate).
+
 ## Conclusion {#sec-predictionAccuracyConclusion}

### 2024-07-15 evaluating-prediction-accuracy.qmd
20240715 - edits

@@
 ```{r}
-#table(mydata$diagnosisFactor, mydata$disorderFactor)
+#table(mydata$diagnosisFactor, mydata$diseaseFactor)
 ```
@@
 ```{r}
-#addmargins(table(mydata$diagnosisFactor, mydata$disorderFactor))
+#addmargins(table(mydata$diagnosisFactor, mydata$diseaseFactor))
 ```
@@
 ```{r}
-#prop.table(table(mydata$diagnosisFactor, mydata$disorderFactor))
+#prop.table(table(mydata$diagnosisFactor, mydata$diseaseFactor))
 ```
@@
 ```{r}
-#addmargins(prop.table(table(mydata$diagnosisFactor, mydata$disorderFactor)))
+#addmargins(prop.table(table(mydata$diagnosisFactor, mydata$diseaseFactor)))
 ```
@@
 
-True positives (TPs) are instances in which a positive classification (e.g., a disorder present) is correct—that is, the test says that a classification is present, and the classification is present.
+True positives (TPs) are instances in which a positive classification (e.g., stating that a disease is present for a person) is correct—that is, the test says that a classification is present, and the classification is present.
 True positives are also called valid positives (VPs) or hits.
@@
 
-True negatives (TNs) are instances in which a negative classification (e.g., absence of a disorder) is correct—that is, the test says that a classification is not present, and the classification is actually not present.
+True negatives (TNs) are instances in which a negative classification (e.g., stating that a disease is absent for a person) is correct—that is, the test says that a classification is not present, and the classification is actually not present.
 True negatives are also called valid negatives (VNs) or correct rejections.
@@
 
-False positives (FPs) are instances in which a positive classification (e.g., a disorder present) is incorrect—that is, the test says that a classification is present, and the classification is not present.
+False positives (FPs) are instances in which a positive classification (e.g., stating that a disease is present for a person) is incorrect—that is, the test says that a classification is present, and the classification is not present.
 False positives are also called false alarms (FAs).
@@
 
-False negatives (FNs) are instances in which a negative classification (e.g., absence of a disorder) is incorrect—that is, the test says that a classification is not present, and the classification is present.
+False negatives (FNs) are instances in which a negative classification (e.g., stating that a disease is absent for a person) is incorrect—that is, the test says that a classification is not present, and the classification is present.
 False negatives are also called misses.
@@
 The selection ratio (SR) is the marginal probability of selection, independent of other things: $P(R_i)$.
-In clinical psychology, the selection ratio is the proportion of people who test positive for the disorder, as in @eq-selectionRatio:
+It is not an aspect of accuracy, per se.
+In medicine, the selection ratio is the proportion of people who test positive for the disease.
+In fantasy football, the selection ratio is the proportion of players who you predict will show a given outcome.
+For example, if you are trying to predict the players who will score a touchdown in a game, the selection ratio is the proportion of players who you predict will score a touchdown.
+The formula for calculating the selection ratio is in @eq-selectionRatio.
 
@@
 The [base rate](#sec-baseRate) (BR) of a classification is its [marginal probability](#sec-baseRate), independent of other things: $P(C_i)$.
-In clinical psychology, the base rate of a disorder is its prevalence in the population, as in @eq-baseRate.
+In medicine, the base rate of a disease is its prevalence in the population, as in @eq-baseRate.
 Without additional information, the [base rate](#sec-baseRate) is used as the initial *pretest probability*.
+In fantasy football, the [base rate](#sec-baseRate) is the proportion of players who actually show the particular outcome.
+For example, if you are trying to predict the players who will score a touchdown in a game, the [base rate](#sec-baseRate) is the proportion of players who actually score a touchdown in the game.
+The formula for calculating the selection ratio is in @eq-baseRate.
 
@@
 
-The posttest probability is the probability of having the disorder given a test result.
+The posttest probability is the probability of having the characteristic given a test result.
 When the [base rate](#sec-baseRate) is used as the [pretest probability](#sec-baseRate), the posttest probability given a positive test is equal to [positive predictive value](#sec-ppv).

### 2024-07-15 actuarial.qmd
20240715 - theory and empiricism

@@
 Human judgment is naturally subject to errors.
+Common [heuristics](#sec-heuristics), [cognitive biases](#sec-cognitiveBiases), and [fallacies](sec-fallacies) are described in @sec-cognitiveBias.
 Below, I describe a few errors to which human judgment seems particularly prone.
 
-When operating freely, clinicians tend to over-estimate exceptions to the established rules (i.e., the broken leg syndrome).
+When operating freely, clinicians and medical experts (and humans more generally) tend to overestimate exceptions to the established rules (i.e., the broken leg syndrome).
 @Meehl1957 acknowledged that there may be some situations where it is glaringly obvious that the statistical formula would be incorrect because it fails to account for an important factor.
@@
 As a result, they too often change the result of the statistical algorithm and are more likely to be wrong than right in doing so.
-Because actuarial methods are based on actual population levels (i.e., [base rates](#sec-baseRates)), unique exceptions are not over-estimated.
+Because actuarial methods are based on actual population levels (i.e., [base rates](#sec-baseRates)), unique exceptions are not overestimated.
 
@@
 *Shrinkage* is when variables with stronger predictive power in the original data set tend to show somewhat smaller predictive power (smaller regression coefficients) when applied to new groups.
-Shrinkage reflects a model [over-fitting](#sec-overfitting) (i.e., fitting to error by capitalizing on chance).
+Shrinkage reflects a model [overfitting](#sec-overfitting) (i.e., fitting to error by capitalizing on chance).
 Shrinkage is especially likely when the original sample is small and/or unrepresentative and the number of variables considered for inclusion is large.
-Cross-validation with large, representative samples can help evaluate the amount of shrinkage of estimates, particularly for more complex models such as machine learning models [@Ursenbach2019].
+Cross-validation with large, representative samples can help evaluate the amount of shrinkage of estimates, particularly for more complex models such as [machine learning](#sec-machineLearning) models [@Ursenbach2019].
 Ideally, cross-validation would be conducted with a separate sample (external cross-validation) to see the generalizability of estimates.
@@
 
-An emerging technique that holds promise for increasing predictive accuracy of actuarial methods is machine learning [@Garb2019].
-However, one challenge of some machine learning techniques is that they are like a "black box" and are not transparent, which raises ethical concerns [@Garb2019].
-Machine learning may be most valuable when the data available are complex and there are many [predictor variables](#sec-correlationalStudy) [@Garb2019].
+An emerging technique that holds promise for increasing predictive accuracy of actuarial methods is [machine learning](#sec-machineLearning) [@Garb2019].
+However, one challenge of some [machine learning](#sec-machineLearning) techniques is that they are like a "black box" and are not transparent, which raises ethical concerns [@Garb2019].
+[machine learning](#sec-machineLearning) may be most valuable when the data available are complex and there are many [predictor variables](#sec-correlationalStudy) [@Garb2019].
 

### 2024-07-15 evaluating-prediction-accuracy.qmd
20240715 - theory and empiricism

@@
 
-## Ways to Improve Prediction Accuracy {#waysToImprovePredictionAccuracy}
+### Theory Versus Empiricism {#sec-theoryVsEmpiricism}
+
+One question that inevitably arises when making predictions is the extent to which one should leverage theory versus empiricism.
+Theory involves conceptual claims of understanding how the causal system works (i.e., what influences what).
+For example, use of theory in prediction might involve specification of the causal system that influences player performance, measurement of those factors, and the integration of that information to make a prediction.
+Empiricism involves "letting the data speak for themselves" and is an atheoretical approach.
+For example, empiricism might involve examining how thousands of variables are associated with the criterion of interest (e.g., fantasy points) and developing the best-fitting model based on those thousands of predictor variables.
+
+Although the atheoretical approach can perform reasonably well, it can be improved by making better use of theory.
+An empirical result (e.g., a correlation) might not necessarily have a lot of meaning associated with it.
+As the maxim goes, [correlation does not imply causation](#sec-correlationCausality).
+Moreover, empiricism can lead to [overfitting](#sec-overfitting).
+So, empiricism is often not enough.
+
+As @Silver2012 notes, "The numbers have no way of speaking for themselves.
+We speak for them.
+We imbue them with meaning." (p. 9).
+If we *understand* the variables in the system and how they influence each other, we can predict things more accurately than predicting for the sake of predicting.
+For instance, [we have made great strides in the last decades when it comes to more accurate weather forecasts](https://www.npr.org/sections/money/2023/07/11/1186458991/should-we-invest-more-in-weather-forecasting-it-may-save-your-life) (archived at https://perma.cc/PF8P-BT3D), including extreme weather events like hurricanes.
+These great strides have more to do with a better causal understanding of the weather system and the ability to conduct simulations of the atmosphere than merely because of big data [@Silver2012].
+By contrast, other events are still incredibly difficult to predict, including earthquakes, in large part because we do not have a strong understanding of the system (and because we do not have ways of precisely measuring those causes because they occur at a depth below which we are realistically able to drill) [@Silver2012].
+
+At the same time, in the social and behavioral sciences, our theories of the causal processes that influence outcomes are not yet very strong.
+Indeed, I have misgivings calling them theories because they do not meet the traditional scientific standard for a theory.
+A scientific theory is an explanation of the natural world that is testable and falsifiable, and that has withstood rigorous scientific testing and scrutiny.
+In psychology (and other areas of social and behavioral sciences), our "theories" are more like conceptual frameworks.
+And these conceptual frameworks are often vague, do not make specific predictions of effects *and* noneffects, and do not hold up consistently when rigorously tested.
+As described by @Meehl1978:
+
+> I consider it unnecessary to persuade you that most so-called "theories" in the soft areas of psychology (clinical, counseling, social, personality, community, and school psychology) are scientifically unimpressive and technologically worthless ...
+Perhaps the easiest way to convince yourself is by scanning the literature of soft psychology over the last 30 years and noticing what happens to theories.
+> Most of them suffer the fate that General MacArthur ascribed to old generals—They never die, they just slowly fade away.
+> In the developed sciences, theories tend either to become widely accepted and built into the larger edifice of well-tested human knowledge or else they suffer destruction in the face of recalcitrant facts and are abandoned, perhaps regretfully as a "nice try."
+> But in fields like personology and social psychology, this seems not to happen.
+> There is a period of enthusiasm about a new theory, a period of attempted application to several fact domains, a period of disillusionment as the negative data come in, a growing bafflement about inconsistent and unreplicable empirical results, multiple resort to ad hoc excuses, and then finally people just sort of lose interest in the thing and pursue other endeavors.
+(pp.
+806–807).
+
+Even if we had strong theoretical understanding of the causal system that influences behavior, we would likely still have difficulty making accurate predictions because the field has largely relied on relatively crude instruments.
+According to one philosophical perspective known as LaPlace's demon, if we were able to know the exact conditions of everything in the universe, we would be able to know how the conditions would be in the future.
+This is an example of scientific determinism, where if you know the initial conditions, you also know the future.
+Other perspectives, such as quantum mechanics and chaos theory, would say that, even if we knew the initial conditions with 100% certainty, there would still be uncertainty in our understanding of the future.
+But assume, for a moment, that LaPlace's demon is true.
+A challenge in the social and behavioral sciences is that we have a relatively poor understanding of the initial conditions of the universe.
+Thus, our predictions would necessarily be probabilistic, similar to weather forecasts.
+Despite having a strong understanding of how weather systems behave, we have imperfect understanding of the initial conditions (e.g., the position and movement of all molecules) [@Silver2012].
+
+Theories tend to make grand conceptual claims that one observed variable influences another observed variable through a complex chain of intervening processes that are unobservable.
+Empiricism provides rich lower-level information, but lacks the broader picture.
+So, it seems, that we need both theory *and* empiricism.
+Theory and empiricism can—and should—inform each other.
+
+## Ways to Improve Prediction Accuracy {#sec-waysToImprovePredictionAccuracy}
 
@@
 The degree of confidence of a person's predictions is often not a good indicator of the accuracy of their predictions [and confidence and prediction accuracy are sometimes inversely associated; @Silver2012].
-[Heuristics](#sec-heuristics) such as the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment), cognitive biases such as [confirmation bias](#sec-cognitiveBiasesConfirmation) [@Hoch1985; @Koriat1980], fallacies such as the [base rate fallacy](#sec-fallaciesBaseRate) [@Eddy1982; @Koehler2002] could contribute to overconfidence of predictions.
+[Heuristics](#sec-heuristics) such as the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment), [cognitive biases](#sec-cognitiveBiases) such as [confirmation bias](#sec-cognitiveBiasesConfirmation) [@Hoch1985; @Koriat1980], [fallacies](#sec-fallacies) such as the [base rate fallacy](#sec-fallaciesBaseRate) [@Eddy1982; @Koehler2002] could contribute to overconfidence of predictions.
 [Poorly calibrated](#sec-calibration) predictions are especially likely when the [base rate](#sec-baseRate) is very low (e.g., suicide) or when the [base rate](#sec-baseRate) is very high [@Koehler2002].
@@
 Ideally, we would use theory to design a model that mirrors the causal system, with accurate measures of each process in the system, so we could make accurate predictions.
-However, as described in Section \@ref(theoryEmpiricism), our psychological theories of the causal processes that influence behavior are not yet very strong.
+However, as described in @sec-theoryVsEmpiricism, our psychological theories of the causal processes that influence behavior are not yet very strong.
 Until we have stronger theories that specify the causal process for a given outcome, and until we have accurate measures of those causal processes, [actuarial approaches](#sec-actuarialPrediction) are likely to be most accurate, as discussed in @sec-judgmentVsActuarial.
@@
 - In addition to considering the accuracy of the prediction, consider the quality of the prediction *process*, especially when random chance is involved to a degree (such as in poker) [@Silver2012].
-- Work to identify and mitigate potential blindspots; be aware of cognitive biases and fallacies, such as [confirmation bias](#sec-cognitiveBiasesConfirmation) and the [base rate fallacy](#sec-fallaciesBaseRate).
+- Work to identify and mitigate potential blindspots; be aware of [cognitive biases](#sec-cognitiveBiases) and [fallacies](#sec-fallacies), such as [confirmation bias](#sec-cognitiveBiasesConfirmation) and the [base rate fallacy](#sec-fallaciesBaseRate).
 

### 2024-07-15 multiple-regression.qmd
20240715 - theory and empiricism

@@
 
-Statistical models applied to big data (e.g., data with many [predictor variables](#sec-correlationalStudy)) can *over-fit* the data, which means that the statistical model accounts for error variance, which will not generalize to future samples.
-So, even though an over-fitting statistical model appears to be accurate because it is accounting for more variance, it is not actually that accurate—it will predict new data less accurately than how accurately it accounts for the data with which the model was built.
+Statistical models applied to big data (e.g., data with many [predictor variables](#sec-correlationalStudy)) can *overfit* the data, which means that the statistical model accounts for error variance, which will not generalize to future samples.
+So, even though an overfitting statistical model appears to be accurate because it is accounting for more variance, it is not actually that accurate—it will predict new data less accurately than how accurately it accounts for the data with which the model was built.
 In the case of fantasy football analytics, this is especially relevant because there are hundreds if not thousands of variables we could consider for inclusion and many, many players when considering historical data.
@@
 
-In @fig-overfittingModel, the blue line represents the true distribution of the data, and the red line is an over-fitting model:
+In @fig-overfittingModel, the blue line represents the true distribution of the data, and the red line is an overfitting model:
 
@@
 #| fig-cap: "Over-fitting Model in Red Relative to the True Distribution of the Data in Blue. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
-#| echo: false
-#| fig.height: 8
-#| fig.width: 8
+#| code-fold: true
+
+set.seed(52242)
 
@@
 
-quadraticModel <- lm(quadraticY ~ quadraticX + I(quadraticX ^ 2), data = quadraticData)
-quadraticNewData <- data.frame(quadraticX = seq(from = min(quadraticData$quadraticX), to = max(quadraticData$quadraticY), length.out = sampleSize))
-quadraticNewData$quadraticY <- predict(quadraticModel, newdata = quadraticNewData)
+quadraticModel <- lm(
+  quadraticY ~ quadraticX + I(quadraticX ^ 2),
+  data = quadraticData)
 
-loessFit <- loess(quadraticY ~ quadraticX, data = quadraticData, span = 0.01, degree = 1)
-loessNewData <- data.frame(quadraticX = seq(from = min(quadraticData$quadraticX), to = max(quadraticData$quadraticY), length.out = sampleSize))
-quadraticNewData$loessY <- predict(loessFit, newdata = quadraticNewData)
+quadraticNewData <- data.frame(
+  quadraticX = seq(
+    from = min(quadraticData$quadraticX),
+    to = max(quadraticData$quadraticY),
+    length.out = sampleSize))
 
-plot(x = quadraticData$quadraticX, y = quadraticData$quadraticY, xlab = "", ylab = "")
-lines(quadraticNewData$quadraticY ~ quadraticNewData$quadraticX, lwd = 2, col = "blue")
-lines(quadraticNewData$loessY ~ quadraticNewData$quadraticX, lwd = 2, col = "red")
+quadraticNewData$quadraticY <- predict(
+  quadraticModel,
+  newdata = quadraticNewData)
+
+loessFit <- loess(
+  quadraticY ~ quadraticX,
+  data = quadraticData,
+  span = 0.01,
+  degree = 1)
+
+loessNewData <- data.frame(
+  quadraticX = seq(
+    from = min(quadraticData$quadraticX),
+    to = max(quadraticData$quadraticY),
+    length.out = sampleSize))
+
+quadraticNewData$loessY <- predict(
+  loessFit,
+  newdata = quadraticNewData)
+
+plot(
+  x = quadraticData$quadraticX,
+  y = quadraticData$quadraticY,
+  xlab = "",
+  ylab = "")
+
+lines(
+  quadraticNewData$quadraticY ~ quadraticNewData$quadraticX,
+  lwd = 2,
+  col = "blue")
+
+lines(
+  quadraticNewData$loessY ~ quadraticNewData$quadraticX,
+  lwd = 2,
+  col = "red")
 ```

### 2024-07-16 evaluating-prediction-accuracy.qmd
20240715 - accuracy indices
20240715 - level of hierarchy
20240716 - edits

@@
 The selection ratio (SR) is the marginal probability of selection, independent of other things: $P(R_i)$.
-It is not an aspect of accuracy, per se.
+It is not an index of accuracy, per se.
 In medicine, the selection ratio is the proportion of people who test positive for the disease.
@@
 The [base rate](#sec-baseRate) (BR) of a classification is its [marginal probability](#sec-baseRate), independent of other things: $P(C_i)$.
+It is not an index of accuracy, per se.
 In medicine, the base rate of a disease is its prevalence in the population, as in @eq-baseRate.
@@
 
-### Theory Versus Empiricism {#sec-theoryVsEmpiricism}
+## Theory Versus Empiricism {#sec-theoryVsEmpiricism}
 
@@
 If we *understand* the variables in the system and how they influence each other, we can predict things more accurately than predicting for the sake of predicting.
-For instance, [we have made great strides in the last decades when it comes to more accurate weather forecasts](https://www.npr.org/sections/money/2023/07/11/1186458991/should-we-invest-more-in-weather-forecasting-it-may-save-your-life) (archived at https://perma.cc/PF8P-BT3D), including extreme weather events like hurricanes.
+For instance, [we have made great strides in the last decades when it comes to more accurate weather forecasts](https://www.npr.org/sections/money/2023/07/11/1186458991/should-we-invest-more-in-weather-forecasting-it-may-save-your-life) (archived at <https://perma.cc/PF8P-BT3D>), including extreme weather events like hurricanes.
 These great strides have more to do with a better causal understanding of the weather system and the ability to conduct simulations of the atmosphere than merely because of big data [@Silver2012].
@@
 Update your judgments based on their accuracy, rather than trying to confirm your beliefs [@Atanasov2020].
-- In addition to considering the accuracy of the prediction, consider the quality of the prediction *process*, especially when random chance is involved to a degree (such as in poker) [@Silver2012].
+- In addition to considering the accuracy of the prediction, consider the quality of the prediction *process*, especially when random chance is involved to a degree, such as in poker and fantasy football [@Silver2012].
 - Work to identify and mitigate potential blindspots; be aware of [cognitive biases](#sec-cognitiveBiases) and [fallacies](#sec-fallacies), such as [confirmation bias](#sec-cognitiveBiasesConfirmation) and the [base rate fallacy](#sec-fallaciesBaseRate).

### 2024-07-17 basic-statistics.qmd
20240714 - cite package
20240716 - conditional probability notation

@@
 The mode of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_mode`.
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `Mode()` function for computing the mode of a set of data.
 
@@
 
-This can be expressed in conditional probability notation, $Pr(A | B)$, which is the probability (likelihood) of event A occurring given that event B occurred (or given condition B).
+This can be expressed in conditional probability notation, $P(A | B)$, which is the probability (likelihood) of event A occurring given that event B occurred (or given condition B).
 
@@
 $$
-p\text{-value} = Pr(T \le t | H_0)
+p\text{-value} = P(T \le t | H_0)
 $$ {#eq-pvalueLeftTailed}
@@
 $$
-p\text{-value} = Pr(T \ge t | H_0)
+p\text{-value} = P(T \ge t | H_0)
 $$ {#eq-pvalueRightTailed}
@@
 $$
-p\text{-value} = 2 \times \text{min}(Pr(T \le t | H_0), Pr(T \ge t | H_0))
+p\text{-value} = 2 \times \text{min}(P(T \le t | H_0), P(T \ge t | H_0))
 $$ {#eq-pvalueTwoTailed}
@@
 $$
-p\text{-value} = 2 \times Pr(T \ge |t| | H_0)
+p\text{-value} = 2 \times P(T \ge |t| | H_0)
 $$ {#eq-pvalueTwoTailedSimple}

### 2024-07-17 evaluating-prediction-accuracy.qmd
20240716 - reorder
20240717 - add more info

@@
 
+## Types of Accuracy {#sec-accuracyTypes}
+
+There are two primary dimensions of accuracy: (1) [discrimination](#sec-discrimination) and (2) [calibration](#sec-calibration).
+Some general indexes of accuracy combine discrimination and calibration, as described in INSERT.
+
+### Discrimination {#sec-discrimination}
+
+When dealing with a categorical outcome, discrimination is the ability to separate events from non-events.
+When dealing with a continuous outcome, discrimination is the strength of the association between the predictor and the outcome.
+Aspects of discrimination at a particular cutoff (e.g., sensitivity, specificity, area under the ROC curve) are described in INSERT.
+
+### Calibration {#sec-calibration}
+
+When dealing with a categorical outcome, calibration is the degree to which a probabilistic estimate of an event reflects the true underlying probability of the event.
+When dealing with a continuous outcome, calibration is the degree to which the predicted values are close in value to the outcome values.
+The importance of examining calibration (in addition to discrimination) is described by @Lindhiem2020.
+
+Calibration is relevant to all kinds of predictions, including weather forecasts.
+For instance, on the days that the meteorologist says there is a 60% chance of rain, it should rain about 60% of the time.
+Calibration is also important for fantasy football predictions.
+When projections state that a a group of players is each expected to score 200 points, their projections would be miscalibrated if those players scored only 150 points on average.
+
+Calibration can be examined in several ways, including Brier Scores, the Hosler-Lemeshow test, Spiegelhalter's $z$, and the mean difference between predicted and observed values at different binned thresholds as depicted graphically with a [calibration plot](#sec-calibrationPlot).
+
+#### Calibration Plot {#sec-calibrationPlot}
+
+Calibration plots can be helpful for identifying miscalibration.
+A calibration plot depicts the predicted probability of an event on the x-axis, and the actual (observed) probability of the event on the y-axis.
+The predictions are binned into a certain number of groups (commonly 10).
+The diagonal line reflects predictions that are perfectly calibrated.
+To the extent that predictions deviate from the diagonal line, the predictions are miscalibrated.
+
+There are four general patterns of miscalibration: overextremity, underextremity, overprediction, and underprediction (see @fig-miscalibration).
+*Overextremity* exists when the predicted probabilites are too close to the extremes (zero or one).
+*Underextremity* exists when the predicted probabilities are too far away from the extremes.
+*Overprediction* exists when the predicted probabilities are consistently greater than the observed probabilities.
+*Underprediction* exists when the predicted probabilities are consistently less than the observed probabilities.
+For a more thorough description of these types of miscalibration, see @Lindhiem2020.
+
+The various types of miscalibration are depicted in @fig-miscalibration:
+
+```{r}
+#| label: fig-miscalibration
+#| fig-cap: "Types of Miscalibration. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
+#| fig-width: 8
+#| fig-height: 8
+#| code-fold: true
+
+# Specify data
+examplePredictions <- seq(from = 0, to = 1, by = .1)
+exampleOutcomes <- c(0, .15, .3, .4, .45, .5, .55, .6, .7, .85, 1)
+
+overPrediction <- c(0, .02, .05, .1, .15, .2, .3, .4, .5, .7, 1)
+underPrediction <- c(0, .3, .5, .6, .7, .8, .85, .9, .95, .98, 1)
+overExtremity <- c(0, .3, .38, .42, .47, .5, .53, .58, .62, .7, 1)
+underExtremity <- c(0, .05, .08, .11, .2, .5, .8, .89, .92, .95, 1)
+
+# Plot
+par(
+  mfrow = c(2,2),
+  mar = c(5,4,1,1) + 0.1) #margins: bottom, left, top, right
+
+plot(
+  examplePredictions,
+  overExtremity,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Overextremity",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+
+plot(
+  examplePredictions,
+  underExtremity,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Underextremity",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+
+plot(
+  examplePredictions,
+  overPrediction,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Overprediction",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+
+plot(
+  examplePredictions,
+  underPrediction,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Underprediction",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+```
+
+### General Accuracy {#sec-generalAccuracy}
+
+[Discuss options for discrete versus continuous data]
+
 ## Prediction of Categorical Outcomes {#sec-predictionCategorical}
@@
 
-To consider how we can evaluate the accuracy of predictions, consider an example adapted from @Meehl1955.
+To consider how we can evaluate the accuracy of predictions for a categorical outcome, consider an example adapted from @Meehl1955.
 The military conducts a test of its prospective members to screen out applicants who would likely fail basic training.
@@
 Specificity is the proportion of those who do not have the characteristic ($\text{TN} + \text{FP}$) that we correctly classify as not having the characteristic ($\text{TN}$): $\frac{\text{TN}}{\text{TN} + \text{FP}} = \frac{1,478}{1,478 + 422} = .78$.
-PPV is the proportion of those who we classify as having the characteristic ($\text{TP} + \text{FP}$) who actually have the characteristic ($\text{TP}$): $\frac{\text{TP}}{\text{TP} + \text{FP}} = \frac{86}{86 + 422} = .17$
+PPV is the proportion of those who we classify as having the characteristic ($\text{TP} + \text{FP}$) who actually have the characteristic ($\text{TP}$): $\frac{\text{TP}}{\text{TP} + \text{FP}} = \frac{86}{86 + 422} = .17$.
 NPV is the proportion of those we classify as not having the characteristic ($\text{TN} + \text{FN}$) who actually do not have the characteristic ($\text{TN}$): $\frac{\text{TN}}{\text{TN} + \text{FN}} = \frac{1,478}{1,478 + 14} = .99$.
@@
 
-## Types of Accuracy {#sec-accuracyTypes}
-
-### Discrimination {#sec-discrimination}
-
-[Discuss options for discrete versus continuous data]
-
-### Calibration {#sec-calibration}
-
-When dealing with a categorical outcome, calibration is the degree to which a probabilistic estimate of an event reflects the true underlying probability of the event.
-When dealing with a continuous outcome, calibration is the degree to which the predicted values are close in value to the outcome values.
-The importance of examining calibration (in addition to discrimination) is described by @Lindhiem2020.
-
-Calibration is relevant to all kinds of predictions, including weather forecasts.
-For instance, on the days that the meteorologist says there is a 60% chance of rain, it should rain about 60% of the time.
-Calibration is also important for fantasy football predictions.
-When projections state that a a group of players is each expected to score 200 points, their projections would be miscalibrated if those players scored only 150 points on average.
-
-Calibration can be examined in several ways, including Brier Scores, the Hosler-Lemeshow test, Spiegelhalter's $z$, and the mean difference between predicted and observed values at different binned thresholds as depicted graphically with a [calibration plot](#sec-calibrationPlot).
-
-#### Calibration Plot {#sec-calibrationPlot}
-
-Calibration plots can be helpful for identifying miscalibration.
-A calibration plot depicts the predicted probability of an event on the x-axis, and the actual (observed) probability of the event on the y-axis.
-The predictions are binned into a certain number of groups (commonly 10).
-The diagonal line reflects predictions that are perfectly calibrated.
-To the extent that predictions deviate from the diagonal line, the predictions are miscalibrated.
-
-There are four general patterns of miscalibration: overextremity, underextremity, overprediction, and underprediction (see @fig-miscalibration).
-*Overextremity* exists when the predicted probabilites are too close to the extremes (zero or one).
-*Underextremity* exists when the predicted probabilities are too far away from the extremes.
-*Overprediction* exists when the predicted probabilities are consistently greater than the observed probabilities.
-*Underprediction* exists when the predicted probabilities are consistently less than the observed probabilities.
-For a more thorough description of these types of miscalibration, see @Lindhiem2020.\index{calibration}
-
-The various types of miscalibration are depicted in @fig-miscalibration:
-
-```{r}
-#| label: fig-miscalibration
-#| fig-cap: "Types of Miscalibration. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
-#| fig-width: 8
-#| fig-height: 8
-#| code-fold: true
-
-# Specify data
-examplePredictions <- seq(from = 0, to = 1, by = .1)
-exampleOutcomes <- c(0, .15, .3, .4, .45, .5, .55, .6, .7, .85, 1)
-
-overPrediction <- c(0, .02, .05, .1, .15, .2, .3, .4, .5, .7, 1)
-underPrediction <- c(0, .3, .5, .6, .7, .8, .85, .9, .95, .98, 1)
-overExtremity <- c(0, .3, .38, .42, .47, .5, .53, .58, .62, .7, 1)
-underExtremity <- c(0, .05, .08, .11, .2, .5, .8, .89, .92, .95, 1)
-
-# Plot
-par(
-  mfrow = c(2,2),
-  mar = c(5,4,1,1) + 0.1) #margins: bottom, left, top, right
-
-plot(
-  examplePredictions,
-  overExtremity,
-  xlim = c(0,1),
-  ylim = c(0,1),
-  main = "Overextremity",
-  xlab = "Predicted Probability",
-  ylab = "Observed Proportion",
-  bty = "l",
-  cex = 1.5,
-  col = "#e41a1c",
-  type = "o")
-
-lines(
-  c(0,1),
-  c(0,1),
-  lwd = 2,
-  col = "#377eb8")
-
-plot(
-  examplePredictions,
-  underExtremity,
-  xlim = c(0,1),
-  ylim = c(0,1),
-  main = "Underextremity",
-  xlab = "Predicted Probability",
-  ylab = "Observed Proportion",
-  bty = "l",
-  cex = 1.5,
-  col = "#e41a1c",
-  type = "o")
-
-lines(
-  c(0,1),
-  c(0,1),
-  lwd = 2,
-  col = "#377eb8")
-
-plot(
-  examplePredictions,
-  overPrediction,
-  xlim = c(0,1),
-  ylim = c(0,1),
-  main = "Overprediction",
-  xlab = "Predicted Probability",
-  ylab = "Observed Proportion",
-  bty = "l",
-  cex = 1.5,
-  col = "#e41a1c",
-  type = "o")
-
-lines(
-  c(0,1),
-  c(0,1),
-  lwd = 2,
-  col = "#377eb8")
-
-plot(
-  examplePredictions,
-  underPrediction,
-  xlim = c(0,1),
-  ylim = c(0,1),
-  main = "Underprediction",
-  xlab = "Predicted Probability",
-  ylab = "Observed Proportion",
-  bty = "l",
-  cex = 1.5,
-  col = "#e41a1c",
-  type = "o")
-
-lines(
-  c(0,1),
-  c(0,1),
-  lwd = 2,
-  col = "#377eb8")
-```
-
-### General Accuracy {#sec-generalAccuracy}
-
-[Discuss options for discrete versus continuous data]
-
 ## Theory Versus Empiricism {#sec-theoryVsEmpiricism}

### 2024-07-18 evaluating-prediction-accuracy.qmd
20240717 - threshold-independent accuracy indices

@@
 There are two primary dimensions of accuracy: (1) [discrimination](#sec-discrimination) and (2) [calibration](#sec-calibration).
-Some general indexes of accuracy combine discrimination and calibration, as described in INSERT.
+Some general indexes of accuracy combine discrimination and calibration, as described in @sec-generalAccurac.
+In addition, accuracy indices can be [threshold-dependent](#sec-thresholdDependentAccuracy) or [-independent](#sec-thresholdIndependentAccuracy) and can be scale-dependent or -independent.
+[Threshold-dependent accuracy indices](#sec-thresholdDependentAccuracy) differ based on the cutoff (i.e., threshold), whereas [threshold-independent accuracy indices](#sec-thresholdIndependentAccuracy) do not.
+Thus, raising or lowering the cutoff will change [threshold-dependent](#sec-thresholdDependentAccuracy) accuracy indices.
+Scale-dependent accuracy indices depend on the metric/scale of the data, whereas scale-independent accuracy indices do not.
+Thus, scale-dependent accuracy indices cannot be directly compared when using measures of differing scales, whereas scale-independent accuracy indices can be compared across data of differing scales.
 
@@
 When projections state that a a group of players is each expected to score 200 points, their projections would be miscalibrated if those players scored only 150 points on average.
-
-Calibration can be examined in several ways, including Brier Scores, the Hosler-Lemeshow test, Spiegelhalter's $z$, and the mean difference between predicted and observed values at different binned thresholds as depicted graphically with a [calibration plot](#sec-calibrationPlot).
-
-#### Calibration Plot {#sec-calibrationPlot}
-
-Calibration plots can be helpful for identifying miscalibration.
-A calibration plot depicts the predicted probability of an event on the x-axis, and the actual (observed) probability of the event on the y-axis.
-The predictions are binned into a certain number of groups (commonly 10).
-The diagonal line reflects predictions that are perfectly calibrated.
-To the extent that predictions deviate from the diagonal line, the predictions are miscalibrated.
 
@@
 
-The various types of miscalibration are depicted in @fig-miscalibration:
-
-```{r}
-#| label: fig-miscalibration
-#| fig-cap: "Types of Miscalibration. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
-#| fig-width: 8
-#| fig-height: 8
-#| code-fold: true
-
-# Specify data
-examplePredictions <- seq(from = 0, to = 1, by = .1)
-exampleOutcomes <- c(0, .15, .3, .4, .45, .5, .55, .6, .7, .85, 1)
-
-overPrediction <- c(0, .02, .05, .1, .15, .2, .3, .4, .5, .7, 1)
-underPrediction <- c(0, .3, .5, .6, .7, .8, .85, .9, .95, .98, 1)
-overExtremity <- c(0, .3, .38, .42, .47, .5, .53, .58, .62, .7, 1)
-underExtremity <- c(0, .05, .08, .11, .2, .5, .8, .89, .92, .95, 1)
-
-# Plot
-par(
-  mfrow = c(2,2),
-  mar = c(5,4,1,1) + 0.1) #margins: bottom, left, top, right
-
-plot(
-  examplePredictions,
-  overExtremity,
-  xlim = c(0,1),
-  ylim = c(0,1),
-  main = "Overextremity",
-  xlab = "Predicted Probability",
-  ylab = "Observed Proportion",
-  bty = "l",
-  cex = 1.5,
-  col = "#e41a1c",
-  type = "o")
-
-lines(
-  c(0,1),
-  c(0,1),
-  lwd = 2,
-  col = "#377eb8")
-
-plot(
-  examplePredictions,
-  underExtremity,
-  xlim = c(0,1),
-  ylim = c(0,1),
-  main = "Underextremity",
-  xlab = "Predicted Probability",
-  ylab = "Observed Proportion",
-  bty = "l",
-  cex = 1.5,
-  col = "#e41a1c",
-  type = "o")
-
-lines(
-  c(0,1),
-  c(0,1),
-  lwd = 2,
-  col = "#377eb8")
-
-plot(
-  examplePredictions,
-  overPrediction,
-  xlim = c(0,1),
-  ylim = c(0,1),
-  main = "Overprediction",
-  xlab = "Predicted Probability",
-  ylab = "Observed Proportion",
-  bty = "l",
-  cex = 1.5,
-  col = "#e41a1c",
-  type = "o")
-
-lines(
-  c(0,1),
-  c(0,1),
-  lwd = 2,
-  col = "#377eb8")
-
-plot(
-  examplePredictions,
-  underPrediction,
-  xlim = c(0,1),
-  ylim = c(0,1),
-  main = "Underprediction",
-  xlab = "Predicted Probability",
-  ylab = "Observed Proportion",
-  bty = "l",
-  cex = 1.5,
-  col = "#e41a1c",
-  type = "o")
-
-lines(
-  c(0,1),
-  c(0,1),
-  lwd = 2,
-  col = "#377eb8")
-```
+Indices for evaluating calibration are described in @sec-calibrationIndices.
 
@@
 
-[Discuss options for discrete versus continuous data]
+General accuracy indices combine estimates of [discrimination](#sec-discrimination) and [calibration](#sec-calibration).
 
 ## Prediction of Categorical Outcomes {#sec-predictionCategorical}
+
+To evaluate the accuracy of our predictions for categorical outcome variables (e.g., binary, dichotomous, or [nominal](#sec-nominal) data), we can use either [threshold-dependent](#sec-thresholdDependentAccuracy) or [threshold-independent](#sec-thresholdIndependentAccuracy) accuracy indices.
+
+## Prediction of Continuous Outcomes {#sec-predictionContinuous}
+
+To evaluate the accuracy of our predictions for continuous outcome variables (e.g., [ordinal](#sec-ordinal), [interval](#sec-interval), or [ratio](#sec-ratio) data), the outcome variable does not have cutoffs, so we would use [threshold-independent accuracy indices](#sec-thresholdIndependentAccuracy).
+
+## Threshold-Dependent Accuracy Indices {#sec-thresholdDependentAccuracy}
 
@@
 
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `accuracyAtCutoff()` that computes many accuracy indices for the prediction of categorical outcome variables.
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `accuracyAtCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables.
 
@@
 #petersenlab::accuracyAtCutoff()
+```
+
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `accuracyAtEachCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables at each possible cutoff.
+
+```{r}
+#petersenlab::accuracyAtEachCutoff()
 ```
@@
 
-## Prediction of Continuous Outcomes {#sec-predictionContinuous}
+## Threshold-Independent Accuracy Indices {#sec-thresholdIndependentAccuracy}
+
+This section describes threshold-independent indexes of accuracy.
+That is, each index of accuracy described in this section provides a single numerical index of accuracy that aggregates the accuracy across all possible cutoffs.
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `accuracyOverall()` function that computes many threshold-independent accuracy indices.
+
+### General Prediction Accuracy {#sec-generalPredictionAccuracy}
+
+There are many metrics of general prediction accuracy.
+When thinking about which metric(s) may be best for a given problem, it is important to consider the purpose of the assessment.
+The estimates of general prediction accuracy are separated below into [scale-dependent](#sec-scaleDependentAccuracy) and [scale-independent](#sec-scaleIndependentAccuracy) accuracy estimates.
+
+#### Scale-Dependent Accuracy Estimates {#sec-scaleDependentAccuracy}
+
+The estimates of prediction accuracy described in this section are scale-dependent.
+These accuracy estimates depend on the unit of measurement and therefore cannot be compared across measures with different scales or across data sets.
+
+##### Mean Error {#sec-meanError}
+
+Here, "error" ($e$) is the difference between the predicted and observed value for a given individual ($i$).
+Mean error (ME; also known as bias) is the mean difference between the predicted and observed values across individuals ($i$), that is, the mean of the errors across individuals ($e_i$).
+Values closer to zero reflect greater accuracy.
+If mean error is above zero, it indicates that predicted values are, on average, greater than observed values (i.e., overestimating errors).
+If mean error is below zero, it indicates that predicted values are, on average, less than observed values (i.e., underestimating errors).
+If both over-estimating and under-estimating errors are present, however, they can cancel each other out.
+As a result, even with a mean error of zero, there can still be considerable error present.
+Thus, although mean error can be helpful for examining whether predictions systematically under- or over-estimate the actual scores, other forms of accuracy are necessary to examine the *extent* of error.
+The formula for mean error is in @eq-meanError:
+
+##### Mean Absolute Error (MAE) {#sec-meanAbsoluteError}
+
+Mean absolute error (MAE) is the mean of the absolute value of differences between the predicted and observed values across individuals, that is, the mean of the absolute value of errors.
+Smaller MAE values (closer to zero) reflect greater accuracy.
+MAE is preferred over [root mean squared error](#sec-rootMeanSquaredError) (RMSE) when you want to give equal weight to all errors and when the outliers have considerable impact.
+The formula for MAE is in @eq-meanAbsoluteError:
+
+$$
+\begin{aligned}
+  \text{mean absolute error (MAE)} &= \frac{\sum\limits_{i = 1}^n|\text{predicted}_i - \text{observed}_i|}{n} \\
+  &= \text{mean}(|e_i|)
+\end{aligned}
+$$ {#eq-meanAbsoluteError}
+
+##### Mean Squared Error (MSE) {#sec-meanSquaredError}
+
+Mean squared error (MSE) is the mean of the square of the differences between the predicted and observed values across individuals, that is, the mean of the squared value of errors.
+Smaller MSE values (closer to zero) reflect greater accuracy.
+MSE penalizes larger errors more heavily than smaller errors (unlike [MAE](#sec-meanAbsoluteError)).
+However, MSE is sensitive to outliers and can be impacted if the errors are skewed.
+The formula for MSE is in @eq-meanSquaredError:
+
+$$
+\begin{aligned}
+  \text{mean squared error (MSE)} &= \frac{\sum\limits_{i = 1}^n(\text{predicted}_i - \text{observed}_i)^2}{n} \\
+  &= \text{mean}(e_i^2)
+\end{aligned}
+$$ {#eq-meanSquaredError}
+
+##### Root Mean Squared Error (RMSE) {#sec-rootMeanSquaredError}
+
+Root mean squared error (RMSE) is the square root of the mean of the square of the differences between the predicted and observed values across individuals, that is, the root mean squared value of errors.
+Smaller RMSE values (closer to zero) reflect greater accuracy.
+RMSE penalizes larger errors more heavily than smaller errors (unlike [MAE](#sec-meanAbsoluteError)).
+However, RMSE is sensitive to outliers and can be impacted if the errors are skewed.
+The formula for RMSE is in @eq-rootMeanSquaredError:
+
+$$
+\begin{aligned}
+  \text{root mean squared error (RMSE)} &= \sqrt{\frac{\sum\limits_{i = 1}^n(\text{predicted}_i - \text{observed}_i)^2}{n}} \\
+  &= \sqrt{\text{mean}(e_i^2)}
+\end{aligned}
+$$ {#eq-rootMeanSquaredError}
+
+#### Scale-Independent Accuracy Estimates {#sec-scaleIndependentAccuracy}
+
+The estimates of prediction accuracy described in this section are intended to be scale-*independent* (unit-free) so the accuracy estimates can be compared across measures with different scales or across data sets [@Hyndman2018].
+
+##### Mean Percentage Error (MPE) {#sec-meanPercentageError}
+
+Mean percentage error (MPE) values closer to zero reflect greater accuracy.
+The formula for percentage error is in @eq-percentageError:
+
+$$
+\begin{aligned}
+  \text{percentage error }(p_i) = \frac{100\% \times (\text{observed}_i - \text{predicted}_i)}{\text{observed}_i}
+\end{aligned}
+$$ {#eq-percentageError}
+
+We then take the mean of the percentage errors to get MPE.
+The formula for MPE is in @eq-meanPercentageError:
+
+$$
+\begin{aligned}
+  \text{mean percentage error (MPE)} &= \frac{100\%}{n} \sum\limits_{i = 1}^n \frac{\text{observed}_i - \text{predicted}_i}{\text{observed}_i} \\
+  &= \text{mean(percentage error)} \\
+  &= \text{mean}(p_i)
+\end{aligned}
+$$ {#eq-meanPercentageError}
+
+Note: MPE is undefined when one or more of the observed values equals zero, due to division by zero.
+The `accuracyOverall()` function of the [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+
+##### Mean Absolute Percentage Error (MAPE) {#sec-meanAbsolutePercentageError}
+
+Smaller mean absolute percentage error (MAPE) values (closer to zero) reflect greater accuracy.
+The formula for MAPE is in @eq-meanAbsolutePercentageError.
+MAPE is asymmetric because it overweights underestimates and underweights overestimates.
+MAPE can be preferable to [symmetric mean absolute percentage error](#sec-symmetricMeanAbsolutePercentageError) (sMAPE) if there are no observed values of zero and if you want to emphasize the importance of underestimates (relative to overestimates).
+
+$$
+\begin{aligned}
+  \text{mean absolute percentage error (MAPE)} &= \frac{100\%}{n} \sum\limits_{i = 1}^n \Bigg|\frac{\text{observed}_i - \text{predicted}_i}{\text{observed}_i}\Bigg| \\
+  &= \text{mean(|percentage error|)} \\
+  &= \text{mean}(|p_i|)
+\end{aligned}
+$$ {#eq-meanAbsolutePercentageError}
+
+Note: MAPE is undefined when one or more of the observed values equals zero, due to division by zero.
+The `accuracyOverall()` function of the [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+
+##### Symmetric Mean Absolute Percentage Error (sMAPE) {#sec-symmetricMeanAbsolutePercentageError}
+
+Unlike MAPE, symmetric mean absolute percentage error (sMAPE) is symmetric because it equally weights underestimates and overestimates.
+Smaller sMAPE values (closer to zero) reflect greater accuracy.
+The formula for sMAPE is in @eq-symmetricMeanAbsolutePercentageError:
+
+$$
+\small
+\begin{aligned}
+  \text{symmetric mean absolute percentage error (sMAPE)} = \frac{100\%}{n} \sum\limits_{i = 1}^n \frac{|\text{predicted}_i - \text{observed}_i|}{|\text{predicted}_i| + |\text{observed}_i|}
+\end{aligned}
+$$ {#eq-symmetricMeanAbsolutePercentageError}
+
+Note: sMAPE is undefined when one or more of the individuals has a prediction–observed combination such that the sum of the absolute value of the predicted value and the absolute value of the observed value equals zero ($|\text{predicted}_i| + |\text{observed}_i|$), due to division by zero.
+The `accuracyOverall()` function of the [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+
+##### Mean Absolute Scaled Error (MASE) {#sec-meanAbsoluteScaledError}
+
+Mean absolute scaled error (MASE) is described by [@Hyndman2018].
+Values closer to zero reflect greater accuracy.
+
+The adapted formula for MASE with non-time series data is described [here (https://stats.stackexchange.com/a/108963/20338)](https://stats.stackexchange.com/a/108963/20338) (archived at https://perma.cc/G469-8NAJ).
+Scaled errors are calculated using @eq-scaledError:
+
+$$
+\begin{aligned}
+  \text{scaled error}(q_i) &= \frac{\text{observed}_i - \text{predicted}_i}{\text{scaling factor}} \\
+  &= \frac{\text{observed}_i - \text{predicted}_i}{\frac{1}{n} \sum\limits_{i = 1}^n |\text{observed}_i - \overline{\text{observed}}|}
+\end{aligned}
+$$ {#eq-scaledError}
+
+Then, we calculate the mean of the absolute value of the scaled errors to get MASE, as in @eq-meanAbsoluteScaledError:
+
+$$
+\begin{aligned}
+  \text{mean absolute scaled error (MASE)} &= \frac{1}{n} \sum\limits_{i = 1}^n |q_i| \\
+  &= \text{mean(|scaled error|)} \\
+  &= \text{mean}(|q_i|)
+\end{aligned}
+$$ {#eq-meanAbsoluteScaledError}
+
+Note: MASE is undefined when the scaling factor is zero, due to division by zero.
+With non-time series data, the scaling factor is the average of the absolute value of individuals' observed scores minus the average observed score ($\frac{1}{n} \sum\limits_{i = 1}^n |\text{observed}_i - \overline{\text{observed}}|$).
+
+##### Root Mean Squared Log Error (RMSLE) {#sec-rootMeanSquaredLogError}
+
+The squared log of the accuracy ratio is described by @Tofallis2015.
+The accuracy ratio is in @eq-accuracyRatio1:
+
+$$
+\begin{aligned}
+  \text{accuracy ratio} &= \frac{\text{predicted}_i}{\text{observed}_i}
+\end{aligned}
+$$ {#eq-accuracyRatio1}
+
+However, the accuracy ratio is undefined with observed or predicted values of zero, so it is common to modify it by adding 1 to the predictor and denominator, as in @eq-accuracyRatio2:
+
+$$
+\begin{aligned}
+  \text{accuracy ratio} &= \frac{\text{predicted}_i + 1}{\text{observed}_i + 1}
+\end{aligned}
+$$ {#eq-accuracyRatio2}
+
+Squaring the log values keeps the values positive, such that smaller values (values closer to zero) reflect greater accuracy.
+Then we take the mean of the squared log values, which keeps the values positive, and calculate the square root of the mean squared log values to put them back on the (pre-squared) log metric.
+This is known as the root mean squared log error (RMSLE).
+Division inside the log is equal to subtraction outside the log.
+So, the formula can be reformulated with the subtraction of two logs, as in @eq-rootMeanSquaredLogError:
+
+$$
+\begin{aligned}
+  \text{root mean squared log error (RMSLE)} &= \sqrt{\sum\limits_{i = 1}^n log\bigg(\frac{\text{predicted}_i + 1}{\text{observed}_i + 1}\bigg)^2} \\
+  &= \sqrt{\text{mean}\Bigg[log\bigg(\frac{\text{predicted}_i + 1}{\text{observed}_i + 1}\bigg)^2\Bigg]} \\
+  &= \sqrt{\text{mean}\big[log(\text{accuracy ratio})^2\big]} = \sqrt{\text{mean}\Big\{\big[log(\text{predicted}_i + 1) - log(\text{actual}_i + 1)\big]^2\Big\}}
+\end{aligned}
+$$ {#eq-rootMeanSquaredLogError}
+
+RMSLE can be preferable when the scores have a wide range of values and are skewed.
+RMSLE can help to reduce the impact of outliers.
+RMSLE gives more weight to smaller errors in the prediction of small observed values, while also penalizing larger errors in the prediction of larger observed values.
+It overweights underestimates and underweights overestimates.
+
+There are other variations of prediction accuracy metrics that use the log of the accuracy ratio.
+One variation makes it similar to median symmetric percentage error [@Morley2018].
+
+Note: Root mean squared log error is undefined when one or more predicted values or actual values equals −1.
+When predicted or actual values are -1, this leads to $log(0)$, which is undefined.
+The `accuracyOverall()` function of the [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+
+##### Coefficient of Determination ($R^2$) {#sec-coefficientOfDetermination}
+
+The [coefficient of determination](#sec-multipleRegressionRSquared) ($R^2$) reflects the proportion of variance in the outcome (dependent) variable that is explained by the model predictions: $R^2 = \frac{\text{variance explained in }Y}{\text{total variance in }Y}$.
+Larger values indicate greater accuracy.
+
+$R^2$ is commonly estimated in [multiple regression](#sec-multipleRegression), in which multiple predictors are allowed to predict one outcome.
+
+###### Adjusted $R^2$ ($R^2_{adj}$) {#sec-adjustedRsquared}
+
+Adjusted $R^2$ is similar to the [coefficient of determination](#sec-coefficientOfDetermination), but it accounts for the number of predictors included in the regression model to penalize [overfitting](#sec-overfitting).
+Adjusted $R^2$ reflects the proportion of variance in the outcome (dependent) variable that is explained by the model predictions over and above what would be expected to be accounted for by chance, given the number of predictors in the model.
+Larger values indicate greater accuracy.
+The formula for adjusted $R^2$ is in @eq-adjustedRSquared.
+Adjusted $R^2$ is described further in @sec-multipleRegressionRSquared.
+
+###### Predictive $R^2$ {#sec-predictiveRsquared}
+
+Predictive $R^2$ is described here: https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/ (archived at https://perma.cc/BK8J-HFUK).
+Predictive $R^2$ penalizes [overfitting](#sec-overfitting), unlike traditional $R^2$.
+Larger values indicate greater accuracy.
+
+### Discrimination {#sec-discriminationIndices}
+
+When dealing with a categorical outcome, discrimination is the ability to separate events from non-events.
+When dealing with a continuous outcome, discrimination is the strength of the association between the predictor and the outcome.
+[Threshold-dependent](#sec-thresholdDependentAccuracy) aspects of [discrimination](#sec-discrimination) at a particular cutoff (e.g., [sensitivity](#sec-sensitivity), [specificity](#sec-specificity)) are described in @sec-thresholdDependentAccuracy.
+
+#### Area under the ROC curve (AUC) {#sec-aucROC}
+
+The [area under the ROC curve (AUC)](#sec-auc) is a general index of discrimination accuracy for a categorical outcome.
+It is also called the concordance ($c$) statistic.
+Larger values reflect greater [discrimination accuracy](#sec-discrimination).
+[AUC](#auc) was estimated using the `pROC` package [@R-pROC].
+
+#### Effect Size ($\beta$) of Regression {#sec-standardizedRegressionCoefficient}
+
+The effect size of a predictor, i.e., the standardized regression coefficient is called a beta ($\beta$) coefficient, is a general index of [discrimination accuracy](#sec-discrimination) for a continuous outcome.
+Larger values reflect greater accuracy.
+We can obtain standardized regression coefficients by standardizing the predictors and outcome using the `scale()` function in R.
+
+### Calibration {#sec-calibrationIndices}
+
+When dealing with a categorical outcome, calibration is the degree to which a probabilistic estimate of an event reflects the true underlying probability of the event.
+When dealing with a continuous outcome, calibration is the degree to which the predicted values are close in value to the outcome values.
+The importance of examining calibration (in addition to [discrimination](#sec-discrimination)) is described by @Lindhiem2020.
+Calibration can be examined in several ways, including [Spiegelhalter's $z$](#sec-spiegelhalterZ) (see @sec-spiegelhalterZ), and the [mean difference between predicted and observed values](#sec-miscalibration) at different binned thresholds as depicted graphically with a [calibration plot](#sec-calibrationPlot) (see @fig-miscalibration).
+
+#### Calibration Plot {#sec-calibrationPlot}
+
+Calibration plots can be helpful for identifying miscalibration.
+A calibration plot depicts the predicted probability of an event on the x-axis, and the actual (observed) probability of the event on the y-axis.
+The predictions are binned into a certain number of groups (commonly 10).
+The diagonal line reflects predictions that are perfectly calibrated.
+To the extent that predictions deviate from the diagonal line, the predictions are miscalibrated.
+
+The various types of miscalibration are depicted in @fig-miscalibration:
+
+```{r}
+#| label: fig-miscalibration
+#| fig-cap: "Types of Miscalibration. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
+#| fig-width: 8
+#| fig-height: 8
+#| code-fold: true
+
+# Specify data
+examplePredictions <- seq(from = 0, to = 1, by = .1)
+exampleOutcomes <- c(0, .15, .3, .4, .45, .5, .55, .6, .7, .85, 1)
+
+overPrediction <- c(0, .02, .05, .1, .15, .2, .3, .4, .5, .7, 1)
+underPrediction <- c(0, .3, .5, .6, .7, .8, .85, .9, .95, .98, 1)
+overExtremity <- c(0, .3, .38, .42, .47, .5, .53, .58, .62, .7, 1)
+underExtremity <- c(0, .05, .08, .11, .2, .5, .8, .89, .92, .95, 1)
+
+# Plot
+par(
+  mfrow = c(2,2),
+  mar = c(5,4,1,1) + 0.1) #margins: bottom, left, top, right
+
+plot(
+  examplePredictions,
+  overExtremity,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Overextremity",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+
+plot(
+  examplePredictions,
+  underExtremity,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Underextremity",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+
+plot(
+  examplePredictions,
+  overPrediction,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Overprediction",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+
+plot(
+  examplePredictions,
+  underPrediction,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  main = "Underprediction",
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  cex = 1.5,
+  col = "#e41a1c",
+  type = "o")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+```
+
+#### Spiegelhalter's z {#sec-spiegelhalterZ}
+
+Spiegelhalter's $z$ was calculated using the `rms` package [@R-rms].
+Smaller $z$ values (and larger associated $p$-values) reflect greater [calibration accuracy](#calibration).
+A statistically significant Spiegelhalter's $z$ (*p* < .05) indicates a significant degree of miscalibration.
+
+#### Calibration for predicting a continuous outcome {#sec-calibrationContinuous}
+
+When predicting a continuous outcome, [calibration](#sec-calibration) of the predicted values in relation to the outcome values can be examined in multiple ways including:
+
+- in a [calibration plot](#sec-calibrationPlot), the extent to which the intercept is near zero and the slope is near one
+- in a [calibration plot](#sec-calibrationPlot), the extent to which the 95% confidence interval of the observed value, across all values of the predicted values, includes the diagonal reference line with an intercept of zero and a slope of one
+- [mean error](#sec-meanError)
+- [mean absolute error](#sec-meanAbsoluteError)
+- [mean squared error](#sec-meanSquaredError)
+- [root mean squared error](#sec-rootMeanSquaredError)
+
+With a plot of the predictions on the x-axis, and the outcomes on the y-axis (i.e., a [calibration plot](#sec-calibrationPlot)), [calibration](#sec-calibration) can be examined graphically as the extent to which the best-fit regression line has an intercept (alpha) close to zero and a slope (beta) close to one [@Steyerberg2014; @Stevens2020].
+The intercept is also called "calibration-in-the-large", whereas "calibration-in-the-small" refers to the extent to which the predicted values match the observed values at a specific predicted value (e.g., when the weather forecaster says that there is a 10% chance of rain, does it actually rain 10% of the time?).
+For predictions to be well [calibrated](#sec-calibration), the intercept should be close to zero and the slope should be close to one.
+If the slope is close to one but the intercept is not close to zero (or the intercept is close to zero but the slope is not close to one), the predictions would not be considered well [calibrated](#sec-calibration).
+The 95% confidence interval of the observed value, across all values of the predicted values, should include the diagonal reference line whose intercept is zero and whose slope is one.
+
+For instance, based on the intercept and slope of the [calibration plot](#sec-calibrationPlot) in Figure INSERT, the predictions are not well calibrated, despite having a slope near one, because the 95% confidence interval of the intercept does not include zero.
+The best-fit line is the yellow line.
+The intercept from the best-fit line is positive, as shown in the regression equation.
+This is a case of underprediction, where the predicted values are consistently less than the observed values.
+The confidence interval of the observed value (i.e., the purple band) is the interval within which we have 95% confidence that the true observed value would lie for a given predicted value, based on the model
+The 95% prediction interval of the observed value (i.e., the dashed red lines) is the interval within which we would expect that 95% of future observations would lie for a given predicted value.
+The black diagonal line indicates the reference line with an intercept of zero and a slope of one.
+The predictions would be significantly [miscalibrated](#sec-calibration) at a given level of the predicted values if the 95% confidence interval of the observed value does not include the reference line at that level of the predicted value.
+In this case, the 95% confidence interval of the observed value does not include the reference line (i.e., the actual observed value) at lower levels of the predicted values, so the predictions are [miscalibrated](#sec-calibration) lower levels of the predicted values.
+
+Gold-standard recommendations include examining the predicted values in relation to the observed values using locally estimated scatterplot smoothing (LOESS) [@Austin2014], such as in Figure INSERT.
+We can examine whether the LOESS-based 95% confidence interval of the observed value at every level of the predicted values includes the diagonal reference line (i.e., the actual observed value).
+In this case, the 95% confidence interval of the observed value does not include the reference line at lower levels of the predicted values, so the predictions are [miscalibrated](#sec-calibration) at lower levels of the predicted values.
 
@@
 Theory and empiricism can—and should—inform each other.
+
+## Test Bias {#sec-testBias}
+
+Test bias refers to systematic error (in measurement, prediction, etc.) as a function of group membership that leads the same score to have different meaning for different groups.
+For instance, if the Wonderlic Contemporary Cognitive Ability Test is a strong predictor of performance for Quarterbacks but not for Running Backs, the test is biased.
+Test bias, including how to identify and address it, is described in [Petersen (2024)](https://isaactpetersen.github.io/Principles-Psychological-Assessment/bias.html).
 
@@
 - Work to identify and mitigate potential blindspots; be aware of [cognitive biases](#sec-cognitiveBiases) and [fallacies](#sec-fallacies), such as [confirmation bias](#sec-cognitiveBiasesConfirmation) and the [base rate fallacy](#sec-fallaciesBaseRate).
+- Evaluate for the possibility of [test bias](#sec-testBias).
+Correct for any [test bias](#sec-testBias).
 

### 2024-07-18 evaluating-prediction-accuracy.qmd
20240717 - package bibliography

@@
 There are two primary dimensions of accuracy: (1) [discrimination](#sec-discrimination) and (2) [calibration](#sec-calibration).
-Some general indexes of accuracy combine discrimination and calibration, as described in @sec-generalAccurac.
+Some general indexes of accuracy combine discrimination and calibration, as described in @sec-generalAccuracy.
 In addition, accuracy indices can be [threshold-dependent](#sec-thresholdDependentAccuracy) or [-independent](#sec-thresholdIndependentAccuracy) and can be scale-dependent or -independent.
@@
 
+$$
+\begin{aligned}
+  \text{mean error} &= \frac{\sum\limits_{i = 1}^n(\text{predicted}_i - \text{observed}_i)}{n} \\
+  &= \text{mean}(e_i)
+\end{aligned}
+$$ {#eq-meanError}
+
 ##### Mean Absolute Error (MAE) {#sec-meanAbsoluteError}
@@
 For instance, if the Wonderlic Contemporary Cognitive Ability Test is a strong predictor of performance for Quarterbacks but not for Running Backs, the test is biased.
-Test bias, including how to identify and address it, is described in [Petersen (2024)](https://isaactpetersen.github.io/Principles-Psychological-Assessment/bias.html).
+Test bias, including how to identify and address it, is described in [@PetersenPrinciplesPsychAssessment](https://isaactpetersen.github.io/Principles-Psychological-Assessment/bias.html).
 

### 2024-07-18 index.qmd
20240717 - package bibliography

@@
 # Preface {#sec-preface .unnumbered .unlisted}
+
+```{r}
+#| include: false
+
+# Automatically create a bib database for R packages
+packagesForBibliography <- c(
+  "petersenlab", "ffanalytics", "pROC", "rms"
+)
+
+knitr::write_bib(
+  c(.packages(), packagesForBibliography),
+  "packages_new.bib")
+```
 
@@
 Several sections in this book come from @PetersenPrinciplesPsychAssessment.
-I am also a co-author of the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) that provides free utilities for downloading fantasy football projections and additional fantasy-relevant data, and for calculating projected points given your league settings.
+I am also a co-author of the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] that provides free utilities for downloading fantasy football projections and additional fantasy-relevant data, and for calculating projected points given your league settings.
 

### 2024-07-18 evaluating-prediction-accuracy.qmd
20240718 - update Hyndman book edition

@@
 
-The estimates of prediction accuracy described in this section are intended to be scale-*independent* (unit-free) so the accuracy estimates can be compared across measures with different scales or across data sets [@Hyndman2018].
+The estimates of prediction accuracy described in this section are intended to be scale-*independent* (unit-free) so the accuracy estimates can be compared across measures with different scales or across data sets [@Hyndman2021].
 
@@
 
-Mean absolute scaled error (MASE) is described by [@Hyndman2018].
+Mean absolute scaled error (MASE) is described by [@Hyndman2021].
 Values closer to zero reflect greater accuracy.

### 2024-07-18 evaluating-prediction-accuracy.qmd
20240718 - integrating accuracy indices

@@
 There are two primary dimensions of accuracy: (1) [discrimination](#sec-discrimination) and (2) [calibration](#sec-calibration).
+[Discrimination](#sec-discrimination) and [calibration](#sec-calibration) are distinct forms of accuracy.
+Just because predictions are high in one form of accuracy does not mean that they will be high in the other form of accuracy.
+As described by @Lindhiem2020, predictions can follow any of the following configurations (and anywhere in between):
+
+- high [discrimination](#sec-discrimination), high [calibration](#sec-calibration)
+- high [discrimination](#sec-discrimination), low [calibration](#sec-calibration)
+- low [discrimination](#sec-discrimination), high [calibration](#sec-calibration)
+- low [discrimination](#sec-discrimination), low [calibration](#sec-calibration)
+
 Some general indexes of accuracy combine discrimination and calibration, as described in @sec-generalAccuracy.
+
 In addition, accuracy indices can be [threshold-dependent](#sec-thresholdDependentAccuracy) or [-independent](#sec-thresholdIndependentAccuracy) and can be scale-dependent or -independent.
@@
 In this case, the 95% confidence interval of the observed value does not include the reference line at lower levels of the predicted values, so the predictions are [miscalibrated](#sec-calibration) at lower levels of the predicted values.
+
+## Integrating the Accuracy Indices {#sec-integratingAccuracy}
+
+After computing the accuracy indices of [discrimination](#sec-discrimination) and (2) [calibration](#sec-calibration), it is then the task to integrate the indices to determine (a) which are the most accurate predictions for the given goals, and (b) whether additional improvements and refinements to the predictions need to be made.
+Each of the accuracy indices is computed differently and thus reward (and penalize) predictive (in)accuracy differently.
+Sometimes, the the accuracy indices will paint a consistent picture regarding which predictions are the most accurate.
+Other times, the accuracy indices may disagree about which predictions are most accurate.
+
+In fantasy football, when evaluating the accuracy of seasonal projections, we care most about accurately distinguishing between higher levels of points (e.g., 200 vs 150) as opposed to lower levels of points (e.g., 0 vs 10).
+Thus, it can be helpful to punish larger errors more heavily than smaller errors, as [RMSE](#sec-rootMeanSquaredError) (unlike [MAE](#sec-meanAbsoluteError)).
+
+Thus, we would emphasize the following metrics:
+
+- [discrimination](#sec-discrimination):
+  - [adjusted $R^2$](#sec-adjustedRsquared)
+- [calibration](#sec-calibration):
+  - [calibration plot](#sec-calibrationPlot)
+- [general accuracy](#sec-calibration):
+  - [RMSE](#sec-rootMeanSquaredError)
+
+If you focus on only one accuracy index, [RMSE](#sec-rootMeanSquaredError) would be a good choice.
+However, I would also examine a [calibration plot](#sec-calibrationPlot) to evaluate whether predictions are poorly calibrated at higher levels of points.
+I would also examine [ME](#sec-meanError)—not to compare the accuracy of various predictions per se—but to determine whether predictions are systematically under- or overestimating actual points.
+If so, predictions may be able to be refined by adding or subtracting a constant to the predictions; however, this could worsen other accuracy indices, so it is important to conduct an interative process of modifying then evaluating, then further modifiying and evaluating, etc.
+
+If we are making predictions about a categorical variable, we would emphasize the following metrics:
+
+- [discrimination](#sec-discrimination):
+  - area under the receiver operating curve
+  - and, secondarily—depending on the particulary cutoff and the relative costs of false positives versus false negatives:
+    - sensitivity
+    - specificicity
+    - positive predictive value
+    - negative predictive value
+- [calibration](#sec-calibration):
+  - [calibration plot](#sec-calibrationPlot)
 

### 2024-07-18 evaluating-prediction-accuracy.qmd
20240718 - cross-refs

@@
 That is, a measure can be useful to the extent that its ROC curve is above the diagonal line (i.e., its discriminative accuracy is above chance).
+
+AUC is a [threshold-independent accuracy index](#sec-thresholdIndependentAccuracy) that applies across all possible cutoff values.
+
 @fig-aucRange depicts ROC curves with a range of AUC values.
@@
 
-Spiegelhalter's $z$ was calculated using the `rms` package [@R-rms].
-Smaller $z$ values (and larger associated $p$-values) reflect greater [calibration accuracy](#calibration).
-A statistically significant Spiegelhalter's $z$ (*p* < .05) indicates a significant degree of miscalibration.
+Spiegelhalter's *z* was calculated using the `rms` package [@R-rms].
+Smaller *z* values (and larger associated *p*-values) reflect greater [calibration accuracy](#calibration).
+A statistically significant Spiegelhalter's *z* (*p* < .05) indicates a significant degree of miscalibration.
 
@@
 If so, predictions may be able to be refined by adding or subtracting a constant to the predictions; however, this could worsen other accuracy indices, so it is important to conduct an interative process of modifying then evaluating, then further modifiying and evaluating, etc.
+It may also be valuable to evaluate the accuracy of various subsets of the predictions.
+For instance, you might examine the predictive accuracy of players whose project points are greater than 100, to evaluate the accuracy of predictions specifically to distinguish between players at higher levels of points.
 
@@
 - [discrimination](#sec-discrimination):
-  - area under the receiver operating curve
+  - [area under the receiver operating curve](#sec-auc)
   - and, secondarily—depending on the particulary cutoff and the relative costs of false positives versus false negatives:
-    - sensitivity
-    - specificicity
-    - positive predictive value
-    - negative predictive value
+    - [sensitivity](#sec-sensitivity)
+    - [specificicity](#sec-specificity)
+    - [positive predictive value](#sec-ppv)
+    - [negative predictive value](#sec-npv)
 - [calibration](#sec-calibration):
   - [calibration plot](#sec-calibrationPlot)
+  - [Spiegelhalter's *z*](#sec-spiegelhalterZ)
 

### 2024-07-18 evaluating-prediction-accuracy.qmd
20240718 - miscalibration

@@
 
+#### {#sec-miscalibration}
+
+The mean difference between predicted values versus observed values at a given cutoff is an index of [miscalibration](#sec-calibration) of predictions at that cutoff.
+It is called "calibration-in-the-small" (as opposed to calibration-in-the-large, which spans all cutoffs).
+Values closer to zero reflect greater accuracy.
+Values above zero indicate that the predicted values are, on average, greater than the observed values.
+Values below zero indicate that the observed values are, on average, greater than the predicted values.
+
 ## Threshold-Independent Accuracy Indices {#sec-thresholdIndependentAccuracy}
@@
 I would also examine [ME](#sec-meanError)—not to compare the accuracy of various predictions per se—but to determine whether predictions are systematically under- or overestimating actual points.
-If so, predictions may be able to be refined by adding or subtracting a constant to the predictions; however, this could worsen other accuracy indices, so it is important to conduct an interative process of modifying then evaluating, then further modifiying and evaluating, etc.
+If so, predictions may be able to be refined by adding or subtracting a constant to the predictions; however, this could worsen other accuracy indices, so it is important to conduct an iterative process of modifying then evaluating, then further modifiying and evaluating, etc.
 It may also be valuable to evaluate the accuracy of various subsets of the predictions.
@@
   - [Spiegelhalter's *z*](#sec-spiegelhalterZ)
+  - [Mean difference between observed and predicted values](#sec-miscalibration)
 

### 2024-07-18 cognitive-bias.qmd
20240718 - miscalibration

@@
 *Overestimation* involves believing that one will perform better than one actually performs.
-Overestimation can be identified with a [calibration plot](#sec-calibration) of the predicted performance versus actual performance, where the person's predicted performance is systematically higher (in at least some cases) than their actual performance.
+Overestimation can be identified with a [calibration plot](#sec-calibrationPlot) of the predicted performance versus actual performance, where the person's predicted performance is systematically higher (in at least some cases) than their actual performance.
 Overestimation corresponds to the "[overprediction](#sec-calibration)" form of [miscalibration](#sec-calibration).
@@
 Likewise, if the weather forecast says it will rain 10% of the time and it actually rains 30% of the time, the predictions are also overprecise because the forecaster is expressing stronger confidence than is warranted that it will not rain.
-Overprecision can be identified with a [calibration plot](#sec-calibration) of the predicted probabilities versus the actual probabilities.
+Overprecision can be identified with a [calibration plot](#sec-calibrationPlot) of the predicted probabilities versus the actual probabilities.
 Overprecision corresponds to the "[overextremity](#sec-calibration)" form of [miscalibration](#sec-calibration).
+
+Overestimation and overprecision are studied in various ways.
+Typically, people are asked about (a) whether an event will occur or (b) the likelihood that the event will occur, across many events.
+For the former (approach "a"), people may be asked to make a dichotomous judgment or prediction, by responding to the question: e.g., "Will it rain tomorrow? [YES/NO]".
+They will then rate their confidence (as a percentage) in their answer (0–100%).
+They would make each of these two ratings for each of many events.
+Then, we can evaluate, for a given respondent, the degree to which the probabilistic estimate of an event reflects the true underlying probability of the event.
+For instance, for a given respondent (and for respondents in general), for the events when the respondent says they are 80% confident an event (e.g., rain) will occur, does the event actually occur around 80% of the time?
+For the latter (approach "b"), people may indicate the likelihood that the event will occur, by responding to the question: "How likely is it that it will rain tomorrow? (0–100%)".
+Then, we can evaluate, for instance, for the events when the respondent says that an event (e.g., rain) is 80% likely to occur, does the event actually occur around 80% of the time?
 

### 2024-07-18 evaluating-prediction-accuracy.qmd
20240718 - miscalibration

@@
 Calibration is also important for fantasy football predictions.
-When projections state that a a group of players is each expected to score 200 points, their projections would be miscalibrated if those players scored only 150 points on average.
+When projections state that a group of players is each expected to score 200 points, their projections would be miscalibrated if those players scored only 150 points on average.
 
@@
 
+Well-calibrated predictions are depicted in @fig-calibrationPlot:
+
+```{r}
+#| label: fig-calibrationPlot
+#| fig-cap: "Predictions that are Well-Calibrated. That is, the predicted values are close to the observed values."
+#| fig-width: 8
+#| fig-height: 8
+#| code-fold: true
+
+# Specify data
+examplePredictionsWellCalibrated <- seq(from = 0, to = 1, by = .1)
+exampleOutcomesWellCalibrated <- seq(from = 0, to = 1, by = .1)
+
+# Plot
+plot(
+  examplePredictionsWellCalibrated,
+  exampleOutcomesWellCalibrated,
+  xlim = c(0,1),
+  ylim = c(0,1),
+  xlab = "Predicted Probability",
+  ylab = "Observed Proportion",
+  bty = "l",
+  type = "n")
+
+lines(
+  c(0,1),
+  c(0,1),
+  lwd = 2,
+  col = "#377eb8")
+
+points(
+  examplePredictionsWellCalibrated,
+  exampleOutcomesWellCalibrated,
+  cex = 1.5,
+  col = "#e41a1c",
+  lwd = 2,
+  type = "p")
+```
+
 The various types of miscalibration are depicted in @fig-miscalibration:
@@
 
-#### Spiegelhalter's z {#sec-spiegelhalterZ}
+#### Spiegelhalter's *z* {#sec-spiegelhalterZ}
 

### 2024-07-18 evaluating-prediction-accuracy.qmd
20240718 - more info on miscalibration

@@
 
-The various types of miscalibration are depicted in @fig-miscalibration:
+The various types of general miscalibration are depicted in @fig-miscalibration:
 
@@
 ```
+
+However, predictions could also be miscalibrated in more specific ways.
+For instance, predictions could be well-calibrated at all predicted probabilities except for a given predicted probability (e.g., 20%).
+Or, the predictions could be miscalibrated but not systematically over- or underpredicted.
+Thus, it is important to evaluate a calibration plot to evaluate the extent to which the predictions are miscalibrated and the pattern of that miscalibration.
 

### 2024-07-18 evaluating-prediction-accuracy.qmd
20240718 - conclusion

@@
 
+When the [base rate](#baseRate) of a behavior is very low or very high, you can be highly accurate in predicting the behavior by [predicting from the base rate](#sec-predictingFromBaseRate).
+Thus, you cannot judge how accurate your prediction is until you know how accurate your predictions would be by [random chance](#sec-accuracyByChance).
+Moreover, maximizing [percent accuracy](#sec-percentAccuracy) may not be the ultimate goal because [different errors have different costs](#sec-differentErrorsDifferentCosts).
+Though there are many indices of accuracy, there are two general types of accuracy: [discrimination](#sec-discrimination) and [calibration](#sec-calibration).
+[Discrimination](#sec-discrimination) accuracy is frequently evaluated with the [area under the receiver operating characteristic curve](#sec-auc), or with [sensitivity](#sec-sensitivity) and [specificity](#sec-specificity), or with [standardized regression coefficients](#sec-standardizedRegressionCoefficient) or the [coefficient of determination](#sec-multipleRegressionRSquared).
+[Calibration](#sec-calibration) accuracy is frequently evaluated graphically and with various indices.
+[Sensitivity](#sec-sensitivity) and [specificity](#sec-specificity) [depend on the cutoff](#sec-accuracyCutoff).
+It is important to evaluate both [discrimination](#sec-discrimination) and [calibration](#sec-calibration) when evaluating prediction accuracy.
+
 ::: {.content-visible when-format="html"}

### 2024-07-19 base-rates.qmd
20240718 - test saving and loading file across chapters

@@
 
+```{r}
+#| include: false
+
+# Test saving file to the repo for use in other chapters
+test <- data.frame(
+  v1 = "This is a Test"
+)
+
+write.csv(
+  x = test,
+  file = "./data/test.txt",
+  row.names = FALSE
+)
+```
+
 ## Conclusion {#sec-baseRatesConclusion}
 
+Fantasy performance—and behavior more generally—is challenging to predict.
+People commonly demonstrate [biases](#sec-cognitiveBiases) and [fallacies](#sec-fallacies) when making predictions.
 People tend to ignore base rates ([base rate fallacy](#sec-fallaciesBaseRate)) when making predictions.

### 2024-07-19 evaluating-prediction-accuracy.qmd
20240718 - header
20240718 - test saving and loading file across chapters

@@
 Nevertheless, it is also important to understand principles in the prediction of categorical [outcomes variables](#sec-correlationalStudy).
+
+In any domain, it is important to evaluate the accuracy of predictions, so we can know how (in)accurate we are, and we can strive to continually improve our predictions.
+Fantasy performance—and human behavior more general—is incredibly challenging to predict.
+In fantasy football, there is considerable luck/chance/randomness.
+There are relatively few (i.e. 17) games, and there is a sizeable injury risk for each player in a given game.
+These and other factors combine to render fantasy football predictions not highly accurate.
+But, first, let's learn about the various ways we can evaluate the accuracy of predictions.
 
@@
 
-#### {#sec-miscalibration}
+#### Mean Difference Between Predicted and Observed Values {#sec-miscalibration}
 
@@
 
+```{r}
+#| include: false
+
+# Test reading file from the repo that was saved from another chapter
+test <- read.csv(
+  file = "./data/test.txt"
+)
+```
+
+```{r}
+#| echo: false
+
+print(test)
+```
+
 ::: {.content-visible when-format="html"}

### 2024-07-19 getting-started.qmd
20240718 - save big data files for use in other chapters
20240718 - load 4th down data
20240718 - callout note for 4th down data b/c it takes a long time to run

@@
 ```{r}
-#| eval: false
 #| include: false
@@
 
+load(file = here("data", "nfl_pbp.RData"))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
@@
 
-```{r}
-#| eval: false
-
+::: {#nte-download4thDownData .callout-note title="Downloading 4th down data"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
 nfl_4thdown <- nfl4th::load_4th_pbp(seasons = 2014:2023)
@@
 )
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = here("data", "nfl_4thdown.RData"))
 ```
@@
 )
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = here("data", "nfl_participation.RData"))
 ```

### 2024-07-19 mythbusters.qmd
20240605 - getting started/load packages
20240608 - getting started
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240703 - conclusion section
20240718 - skeleton for mythbusters
20240719 - fix errors and cleanup

--- 2024-05-18 mythbusters.qmd
@@
 # Mythbusters: Putting Fantasy Football Beliefs/Anecdotes to the Test {#sec-mythbusters}
+
+## Getting Started {#sec-mythbustersGettingStarted}
+
+### Load Packages {#sec-mythbustersLoadPackages}
+
+```{r}
+library("here")
+library("tidyverse")
+```
+
+### Load Data {#sec-mythbustersLoadData}
+
+```{r}
+load(file = here("data", "nfl_playerContracts.RData"))
+load(file = here("data", "nfl_pbp.RData"))
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
+```
+
+## Do Players Score More Fantasy Points in their Contract Year? {#sec-contractYear}
+
+```{r}
+# Subset to remove players without a year signed
+nfl_playerContracts_subset <- nfl_playerContracts %>% 
+  dplyr::filter(!is.na(year_signed) & year_signed != 0)
+
+# Determine the contract year for a given contract
+nfl_playerContracts_subset$contractYear <- nfl_playerContracts_subset$year_signed + nfl_playerContracts_subset$years - 1
+
+# Arrange contracts by player and year_signed
+nfl_playerContracts_subset <- nfl_playerContracts_subset %>%
+  dplyr::group_by(player, position) %>% 
+  dplyr::arrange(player, position, -year_signed) %>% 
+  dplyr::ungroup()
+
+# Determine if the player played in the original contract year
+nfl_playerContracts_subset <- nfl_playerContracts_subset %>%
+  dplyr::group_by(player, position) %>%
+  dplyr::mutate(
+    next_contract_start = lag(year_signed)) %>%
+  dplyr::ungroup() %>%
+  dplyr::mutate(
+    played_in_contract_year = ifelse(
+      is.na(next_contract_start) | contractYear < next_contract_start,
+      TRUE,
+      FALSE))
+
+# Check individual players
+nfl_playerContracts_subset %>% 
+  dplyr::filter(player == "Aaron Rodgers") %>% 
+  dplyr::select(player:years, contractYear, next_contract_start, played_in_contract_year)
+
+nfl_playerContracts_subset %>% 
+  dplyr::filter(player %in% c("Jared Allen", "Aaron Rodgers")) %>% 
+  dplyr::select(player:years, contractYear, next_contract_start, played_in_contract_year)
+
+# Merge with seasonal fantasy points data
+```
+
+## Conclusion {#sec-mythbustersConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-mythbustersSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-19 evaluating-prediction-accuracy.qmd
20240719 - clean up
20240719 - links and formatting

@@
 $$
+\scriptsize
 \begin{aligned}
@@
 
-Predictive $R^2$ is described here: https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/ (archived at https://perma.cc/BK8J-HFUK).
+Predictive $R^2$ is described here: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>).
 Predictive $R^2$ penalizes [overfitting](#sec-overfitting), unlike traditional $R^2$.
@@
 Larger values reflect greater accuracy.
-We can obtain standardized regression coefficients by standardizing the predictors and outcome using the `scale()` function in R.
+We can obtain standardized regression coefficients by standardizing the predictors and outcome using the `scale()` function in `R`.
 
@@
 
-```{r}
-#| include: false
-
-# Test reading file from the repo that was saved from another chapter
-test <- read.csv(
-  file = "./data/test.txt"
-)
-```
-
-```{r}
-#| echo: false
-
-print(test)
-```
-
 ::: {.content-visible when-format="html"}

### 2024-07-19 getting-started.qmd
20240719 - fix errors and cleanup
20240719 - clean up
20240719 - remove `here` package dependency

--- 2024-07-10 getting-started.qmd
@@
 
-library("here")
-
 # Downloaded Data
-load(file = here("data", "nfl_players.RData"))
-load(file = here("data", "nfl_teams.RData"))
-load(file = here("data", "nfl_rosters.RData"))
-load(file = here("data", "nfl_rosters_weekly.RData"))
-load(file = here("data", "nfl_schedules.RData"))
-load(file = here("data", "nfl_combine.RData"))
-load(file = here("data", "nfl_draftPicks.RData"))
-load(file = here("data", "nfl_depthCharts.RData"))
+load(file = "./data/nfl_players.RData")
+load(file = "./data/nfl_teams.RData")
+load(file = "./data/nfl_rosters.RData")
+load(file = "./data/nfl_rosters_weekly.RData")
+load(file = "./data/nfl_schedules.RData")
+load(file = "./data/nfl_combine.RData")
+load(file = "./data/nfl_draftPicks.RData")
+load(file = "./data/nfl_depthCharts.RData")
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
-load(file = here("data", "nfl_actualStats_weekly.RData"))
-load(file = here("data", "nfl_injuries.RData"))
-load(file = here("data", "nfl_snapCounts.RData"))
-load(file = here("data", "nfl_espnQBR.RData"))
-load(file = here("data", "nfl_nextGenStats_weekly.RData"))
-load(file = here("data", "nfl_advancedStatsPFR.RData"))
-load(file = here("data", "nfl_playerContracts.RData"))
-load(file = here("data", "nfl_ftnCharting.RData"))
-load(file = here("data", "nfl_playerIDs.RData"))
-load(file = here("data", "nfl_rankings.RData"))
-load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))
-load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
+load(file = "./data/nfl_actualStats_weekly.RData")
+load(file = "./data/nfl_injuries.RData")
+load(file = "./data/nfl_snapCounts.RData")
+load(file = "./data/nfl_espnQBR.RData")
+load(file = "./data/nfl_nextGenStats_weekly.RData")
+load(file = "./data/nfl_advancedStatsPFR.RData")
+load(file = "./data/nfl_playerContracts.RData")
+load(file = "./data/nfl_ftnCharting.RData")
+load(file = "./data/nfl_playerIDs.RData")
+load(file = "./data/nfl_rankings.RData")
+load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
+load(file = "./data/nfl_expectedFantasyPoints_offense.RData")
 
 # Calculated Data
-load(file = here("data", "nfl_actualStats_career.RData"))
-load(file = here("data", "nfl_actualStats_seasonal.RData"))
+load(file = "./data/nfl_actualStats_career.RData")
+load(file = "./data/nfl_actualStats_seasonal.RData")
 ```
@@
   nfl_players,
-  file = here("data", "nfl_players.RData")
-)
-```
-
-```{r}
-#| include: false
-
-library("here")
-
-load(file = here("data", "nfl_players.RData"))
+  file = "./data/nfl_players.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_players.RData")
 ```
@@
   nfl_teams,
-  file = here("data", "nfl_teams.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_teams.RData"))
+  file = "./data/nfl_teams.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_teams.RData")
 ```
@@
   nfl_rosters,
-  file = here("data", "nfl_rosters.RData")
+  file = "./data/nfl_rosters.RData"
 )
@@
   nfl_rosters_weekly,
-  file = here("data", "nfl_rosters_weekly.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_rosters.RData"))
-load(file = here("data", "nfl_rosters_weekly.RData"))
+  file = "./data/nfl_rosters_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_rosters.RData")
+load(file = "./data/nfl_rosters_weekly.RData")
 ```
@@
   nfl_schedules,
-  file = here("data", "nfl_schedules.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_schedules.RData"))
+  file = "./data/nfl_schedules.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_schedules.RData")
 ```
@@
   nfl_combine,
-  file = here("data", "nfl_combine.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_combine.RData"))
+  file = "./data/nfl_combine.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_combine.RData")
 ```
@@
   nfl_draftPicks,
-  file = here("data", "nfl_draftPicks.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_draftPicks.RData"))
+  file = "./data/nfl_draftPicks.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_draftPicks.RData")
 ```
@@
   nfl_depthCharts,
-  file = here("data", "nfl_depthCharts.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_depthCharts.RData"))
+  file = "./data/nfl_depthCharts.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_depthCharts.RData")
 ```
@@
   nfl_pbp,
-  file = here("data", "nfl_pbp.RData")
-)
+  file = "./data/nfl_pbp.RData"
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_pbp.RData")
 ```
@@
 
-```{r}
-#| eval: false
+::: {#nte-download4thDownData .callout-note title="Downloading 4th down data"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| warning: false
+#| message: false
 
@@
   nfl_4thdown,
-  file = here("data", "nfl_4thdown.RData")
-)
+  file = "./data/nfl_4thdown.RData"
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_4thdown.RData")
 ```
@@
   nfl_participation,
-  file = here("data", "nfl_participation.RData")
-)
+  file = "./data/nfl_participation.RData"
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_participation.RData")
 ```
@@
   nfl_actualStats_offense_weekly, nfl_actualStats_defense_weekly, nfl_actualStats_kicking_weekly,
-  file = here("data", "nfl_actualStats_weekly.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_actualStats_weekly.RData"))
+  file = "./data/nfl_actualStats_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_weekly.RData")
 ```
@@
   nfl_injuries,
-  file = here("data", "nfl_injuries.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_injuries.RData"))
+  file = "./data/nfl_injuries.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_injuries.RData")
 ```
@@
   nfl_snapCounts,
-  file = here("data", "nfl_snapCounts.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_snapCounts.RData"))
+  file = "./data/nfl_snapCounts.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_snapCounts.RData")
 ```
@@
   nfl_espnQBR,
-  file = here("data", "nfl_espnQBR.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_espnQBR.RData"))
+  file = "./data/nfl_espnQBR.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_espnQBR.RData")
 ```
@@
   nfl_nextGenStats_weekly,
-  file = here("data", "nfl_nextGenStats_weekly.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_nextGenStats_weekly.RData"))
+  file = "./data/nfl_nextGenStats_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_nextGenStats_weekly.RData")
 ```
@@
   nfl_advancedStatsPFR,
-  file = here("data", "nfl_advancedStatsPFR.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_advancedStatsPFR.RData"))
+  file = "./data/nfl_advancedStatsPFR.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_advancedStatsPFR.RData")
 ```
@@
   nfl_playerContracts,
-  file = here("data", "nfl_playerContracts.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_playerContracts.RData"))
+  file = "./data/nfl_playerContracts.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_playerContracts.RData")
 ```
@@
   nfl_ftnCharting,
-  file = here("data", "nfl_ftnCharting.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_ftnCharting.RData"))
+  file = "./data/nfl_ftnCharting.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_ftnCharting.RData")
 ```
@@
   nfl_playerIDs,
-  file = here("data", "nfl_playerIDs.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_playerIDs.RData"))
+  file = "./data/nfl_playerIDs.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_playerIDs.RData")
 ```
@@
   nfl_rankings,
-  file = here("data", "nfl_rankings.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_rankings.RData"))
+  file = "./data/nfl_rankings.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_rankings.RData")
 ```
@@
   nfl_expectedFantasyPoints_weekly,
-  file = here("data", "nfl_expectedFantasyPoints_weekly.RData")
+  file = "./data/nfl_expectedFantasyPoints_weekly.RData"
 )
@@
   nfl_expectedFantasyPoints_offense,
-  file = here("data", "nfl_expectedFantasyPoints_offense.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))
-load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
+  file = "./data/nfl_expectedFantasyPoints_offense.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
+load(file = "./data/nfl_expectedFantasyPoints_offense.RData")
 ```
@@
 
+Number of rows:
+
+```{r}
+nrow(nfl_players)
+```
+
+Number of columns:
+
+```{r}
+ncol(nfl_players)
+```
+
 ### Number of Elements {#sec-dataCharacteristicsLength}
@@
   nfl_actualStats_offense_career, nfl_actualStats_defense_career, nfl_actualStats_kicking_career,
-  file = here("data", "nfl_actualStats_career.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_actualStats_career.RData"))
+  file = "./data/nfl_actualStats_career.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_career.RData")
 ```
@@
   nfl_actualStats_offense_seasonal, nfl_actualStats_defense_seasonal, nfl_actualStats_kicking_seasonal,
-  file = here("data", "nfl_actualStats_seasonal.RData")
-)
-```
-
-```{r}
-#| include: false
-
-load(file = here("data", "nfl_actualStats_seasonal.RData"))
+  file = "./data/nfl_actualStats_seasonal.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_seasonal.RData")
 ```

### 2024-07-19 sports-cognitive-psychology.qmd
20240605 - getting started/load packages
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240703 - conclusion section
20240719 - remove `here` package dependency

@@
 
-## Load Packages {#sec-loadPackages}
+## Getting Started {#sec-sportsCognitivePsychologyGettingStarted}
+
+### Load Packages {#sec-sportsCognitivePsychologyLoadPackages}
 
@@
 
-## Download Football Data {#sec-downloadFootballData}
+### Download Football Data {#sec-sportsCognitivePsychologyDownloadFootballData}
 
@@
 ```
+
+## Conclusion {#sec-sportsCognitivePsychologyConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-sportsCognitivePsychologySessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-19 basic-statistics.qmd
20240719 - summary of statistical decision trees

@@
 
+A statistical decision tree is a flowchart or decision tree that depicts which statistical test to use given the purpose of analysis, the type of data, etc.
 An example statistical decision tree is depicted in @fig-statisticalDecisionTree.
@@
 
+This statistical decision tree can be generally summarized such that associations are examined with the correlation/regression family, and differences are examined with the *t*-test/ANOVA family, as depicted in @fig-statisticalDecisionTreeSummary.
+
+::: {#fig-statisticalDecisionTreeSummary}
+![](images/statisticalDecisionTreeSummary.png)
+
+Summary of A Statistical Decision Tree For Choosing an Appropriate Statistical Procedure.
+:::
+
 However, many statistical tests can be re-formulated in a regression framework, as in @fig-statisticalDecisionTreeRegression.
@@
 :::
+
+Both associations and differences can be examined with the regression family, which greatly simplifies our summary of the statistical decision tree, as depicted in @fig-statisticalDecisionTreeSummaryRegression.
+
+::: {#fig-statisticalDecisionTreeSummaryRegression}
+![](images/statisticalDecisionTreeSummaryRegression.png)
+
+Summary of A Statistical Decision Tree For Choosing an Appropriate Statistical Procedure.
+:::
+
+Thus, in most cases, the regression framework can be used to examine most questions regarding associations between variables or differences between groups.
 

### 2024-07-20 mythbusters.qmd
20240719 - remove `here` package dependency
20240719 - contract year

@@
 ```{r}
-library("here")
+library("petersenlab")
+library("lme4")
+library("lmerTest")
 library("tidyverse")
@@
 ```{r}
-load(file = here("data", "nfl_playerContracts.RData"))
-load(file = here("data", "nfl_pbp.RData"))
+load(file = "./data/nfl_playerContracts.RData")
+load(file = "./data/nfl_actualStats_weekly.RData")
+load(file = "./data/nfl_actualStats_seasonal.RData")
+load(file = "./data/nfl_espnQBR.RData")
 ```
 
-```{r}
-#| eval: false
-#| include: false
-
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
-```
-
-## Do Players Score More Fantasy Points in their Contract Year? {#sec-contractYear}
+## Do Players Perform Better in their Contract Year? {#sec-contractYear}
 
 ```{r}
+#| code-fold: true
+
 # Subset to remove players without a year signed
@@
 # Check individual players
-nfl_playerContracts_subset %>% 
-  dplyr::filter(player == "Aaron Rodgers") %>% 
-  dplyr::select(player:years, contractYear, next_contract_start, played_in_contract_year)
+#nfl_playerContracts_subset %>% 
+#  dplyr::filter(player == "Aaron Rodgers") %>% 
+#  dplyr::select(player:years, contractYear, next_contract_start, played_in_contract_year)
+#
+#nfl_playerContracts_subset %>% 
+#  dplyr::filter(player %in% c("Jared Allen", "Aaron Rodgers")) %>% 
+#  dplyr::select(player:years, contractYear, next_contract_start, played_in_contract_year)
 
-nfl_playerContracts_subset %>% 
-  dplyr::filter(player %in% c("Jared Allen", "Aaron Rodgers")) %>% 
-  dplyr::select(player:years, contractYear, next_contract_start, played_in_contract_year)
+# Subset data
+nfl_playerContractYears <- nfl_playerContracts_subset %>% 
+  dplyr::filter(played_in_contract_year == TRUE) %>% 
+  dplyr::filter(position %in% c("QB","RB","WR","TE")) %>% 
+  dplyr::select(player, position, team, contractYear) %>% 
+  dplyr::mutate(playerMerge = toupper(str_replace_all(player, "[[:punct:]]", ""))) %>% 
+  dplyr::rename(season = contractYear) %>% 
+  dplyr::mutate(contractYear = 1)
+
+# Merge with weekly and seasonal stats data
+nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly %>% 
+  dplyr::mutate(playerMerge = toupper(str_replace_all(player_display_name, "[[:punct:]]", "")))
+#nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal %>% 
+#  mutate(playerMerge = toupper(str_replace_all(player_display_name, "[[:punct:]]", "")))
+
+nfl_actualStatsContracts_offense_weekly <- dplyr::full_join(
+  nfl_actualStats_offense_weekly,
+  nfl_playerContractYears,
+  by = c("playerMerge", "position_group" = "position", "season")
+) %>% 
+  dplyr::filter(position_group %in% c("QB","RB","WR","TE"))
+
+#nfl_actualStatsContracts_offense_seasonal <- full_join(
+#  nfl_actualStats_offense_seasonal,
+#  nfl_playerContractYears,
+#  by = c("playerMerge", "position_group" = "position", "season")
+#) %>% 
+#  filter(position_group %in% c("QB","RB","WR","TE"))
+
+nfl_actualStatsContracts_offense_weekly$contractYear[which(is.na(nfl_actualStatsContracts_offense_weekly$contractYear))] <- 0
+#nfl_actualStatsContracts_offense_seasonal$contractYear[which(is.na(nfl_actualStatsContracts_offense_seasonal$contractYear))] <- 0
+
+nfl_actualStatsContracts_offense_weekly <- nfl_actualStatsContracts_offense_weekly %>% 
+  dplyr::arrange(playerMerge, season, season_type, week)
+
+#nfl_actualStatsContracts_offense_seasonal <- nfl_actualStatsContracts_offense_seasonal %>% 
+#  arrange(playerMerge, season)
+
+nfl_actualStatsContractsSubset_offense_weekly <- nfl_actualStatsContracts_offense_weekly %>% 
+  dplyr::filter(season_type == "REG")
+
+# Calculate Per Game Totals
+nfl_actualStatsContracts_seasonal <- nfl_actualStatsContractsSubset_offense_weekly %>% 
+  dplyr::group_by(playerMerge, season) %>% 
+  dplyr::summarise(
+    position_group = petersenlab::Mode(position_group),
+    rushing_yards = sum(rushing_yards, na.rm = TRUE), # season total
+    carries = sum(carries, na.rm = TRUE), # season total
+    rushing_epa = mean(rushing_epa, na.rm = TRUE),
+    receiving_yards = mean(receiving_yards, na.rm = TRUE),
+    receiving_epa = mean(receiving_epa, na.rm = TRUE),
+    contractYear = mean(contractYear, na.rm = TRUE),
+    games = n(),
+    .groups = "drop_last"
+  ) %>% 
+  dplyr::mutate(ypc = rushing_yards / carries)
 
 # Merge with seasonal fantasy points data
+```
+
+### QB
+
+First, we prepare the data by merging and performing additional processing:
+
+```{r}
+#| code-fold: true
+
+# Merge with QBR data
+nfl_espnQBR$playerMerge <- paste(nfl_espnQBR$name_first, nfl_espnQBR$name_last, sep = " ") %>% 
+  stringr::str_replace_all(., "[[:punct:]]", "") %>% 
+  toupper()
+
+nfl_contractYearQBR_weekly <- nfl_playerContractYears %>% 
+  dplyr::filter(position == "QB") %>% 
+  dplyr::full_join(
+    .,
+    nfl_espnQBR,
+    by = c("playerMerge","team","season")
+  )
+
+nfl_contractYearQBR_weekly$contractYear[which(is.na(nfl_contractYearQBR_weekly$contractYear))] <- 0
+
+nfl_contractYearQBR_weekly <- nfl_contractYearQBR_weekly %>% 
+  dplyr::arrange(playerMerge, season, season_type, game_week)
+
+nfl_contractYearQBRsubset_weekly <- nfl_contractYearQBR_weekly %>% 
+  dplyr::filter(
+    season_type == "Regular",
+    game_week != "Season Total") %>% 
+  dplyr::mutate(game_week = as.numeric(game_week)) %>% 
+  dplyr::arrange(playerMerge, season, season_type, game_week)
+
+#hist(nfl_contractYearQBRsubset_weekly$qb_plays) # players have at least 20 dropbacks per game
+
+# Calculate Per Game Totals
+nfl_contractYearQBR_seasonal <- nfl_contractYearQBRsubset_weekly %>% 
+  dplyr::group_by(playerMerge, season) %>% 
+  dplyr::summarise(
+    qbr = mean(qbr_total, na.rm = TRUE),
+    pts_added = mean(pts_added, na.rm = TRUE),
+    epa_pass = mean(pass, na.rm = TRUE),
+    qb_plays = sum(qb_plays, na.rm = TRUE), # season total
+    contractYear = mean(contractYear, na.rm = TRUE),
+    games = n(),
+    .groups = "drop_last"
+  )
+
+nfl_contractYearQBRsubset_seasonal <- nfl_contractYearQBR_seasonal %>% 
+  dplyr::filter(games >= 5) # keep only player-season combinations in which QBs played at least 5 games
+```
+
+Then, we analyze the data.
+Below is a mixed model that examines whether a player has a higher QBR per game when they are in a contract year compared to when they are not in a contract year.
+
+```{r}
+mixedModel_qbr <- lmerTest::lmer(
+  qbr ~ contractYear + (1 | playerMerge),
+  data = nfl_contractYearQBR_seasonal
+)
+
+summary(mixedModel_qbr)
+```
+
+```{r}
+mixedModel_ptsAdded <- lmerTest::lmer(
+  pts_added ~ contractYear + (1 | playerMerge),
+  data = nfl_contractYearQBR_seasonal
+)
+
+summary(mixedModel_ptsAdded)
+```
+
+```{r}
+mixedModel_epaPass <- lmerTest::lmer(
+  epa_pass ~ contractYear + (1 | playerMerge),
+  data = nfl_contractYearQBR_seasonal
+)
+
+summary(mixedModel_epaPass)
+```
+
+```{r}
+# Placeholder for model predicting fantasy points
+```
+
+### RB
+
+```{r}
+#| code-fold: true
+
+nfl_actualStatsContractsRB_seasonal <- nfl_actualStatsContracts_seasonal %>% 
+  dplyr::filter(
+    position_group == "RB",
+    games >= 5)
+```
+
+```{r}
+mixedModel_ypc <- lmerTest::lmer(
+  ypc ~ contractYear + (1 | playerMerge),
+  data = nfl_actualStatsContractsRB_seasonal
+)
+
+summary(mixedModel_ypc)
+```
+
+```{r}
+mixedModel_epaRush <- lmerTest::lmer(
+  rushing_epa ~ contractYear + (1 | playerMerge),
+  data = nfl_actualStatsContractsRB_seasonal
+)
+
+summary(mixedModel_epaRush)
+```
+
+```{r}
+# Placeholder for model predicting fantasy points
+```
+
+### WR/TE
+
+```{r}
+#| code-fold: true
+
+nfl_actualStatsContractsWRTE_seasonal <- nfl_actualStatsContracts_seasonal %>% 
+  dplyr::filter(
+    position_group %in% c("WR","TE"),
+    games >= 5)
+```
+
+```{r}
+mixedModel_receivingYards <- lmerTest::lmer(
+  receiving_yards ~ contractYear + (1 | playerMerge),
+  data = nfl_actualStatsContractsWRTE_seasonal
+)
+
+summary(mixedModel_receivingYards)
+```
+
+```{r}
+mixedModel_epaReceiving <- lmerTest::lmer(
+  receiving_epa ~ contractYear + (1 | playerMerge),
+  data = nfl_actualStatsContractsWRTE_seasonal
+)
+
+summary(mixedModel_epaReceiving)
+```
+
+```{r}
+# Placeholder for model predicting fantasy points
+```
+
+### QB/RB/WR/TE
+
+```{r}
+# Placeholder for model predicting fantasy points
+# Include player position as a covariate
 ```

### 2024-07-20 mythbusters.qmd
20240720 - contract year

@@
 ## Do Players Perform Better in their Contract Year? {#sec-contractYear}
+
+Considerable speculation exists regarding whether players perform better in their last year of their contract (i.e., their "contract year").
+Fantasy football talking heads and commentators frequently discuss the benefit of selecting players who are in their contract year, because it supposedly means that player has more motivation to perform well so they get a new contract and get paid more.
+To our knowledge, no peer-reviewed studies have examined this question for football players.
+One study found that National Basketball Association (NBA) players improved in field goal percentage, points, and player efficiency rating (but not other statistics: rebounds, assists, steals, or blocks) from their pre-contract year to their contract year, and that Major League Baseball (MLB) players improved in runs batted in (RBIs; but not other statistics: batting average, slugging percentage, on base percentage, home runs, fielding percentage) from their pre-contract year to their contract year [White2014a].
+Other casual analyses have been examined contract-year performance of National Football League (NFL) players, including articles in [2012](https://www.4for4.com/2012/preseason/2012-contract-year-players-and-myth-increased-production) (archived [here](https://perma.cc/CT3F-QN5E)) and [2022](https://www.4for4.com/2022/preseason/do-players-perform-better-fantasy-football-contract-year) (archived [here](https://perma.cc/F4F5-7RQZ)).
+
+Let's examine the question empirically.
+In order to do that, we have to make some assumptions/constraints.
+In this example, we will make the following constraints:
+
+- We will examine seasons since 2011, beginning when most data for player contracts are available.
+- For maximum [statistical power](#sec-statisticalPower) to detect an effect if a contract year effect exists, we will examine all seasons for a player (since 2011), not just their contract year and their pre-contract year.
+- To ensure a more fair, apples-to-apples comparison of the games in which players played, we will examine *per-game* performance (except for yards per carry, which is based on $\frac{\text{rushing yards}}{\text{carries}}$ from the entire season).
+- We will examine regular season games only (no postseason).
+- To ensure we do not make generalization about a player's performance in a season from a small sample, the player has to play at least 5 games in a given season for that player–season combination to be included in analysis.
 
@@
 
+#table(nfl_playerContracts$year_signed) # most contract data is available beginning in 2011
+
 # Calculate Per Game Totals
@@
 nfl_contractYearQBRsubset_seasonal <- nfl_contractYearQBR_seasonal %>% 
-  dplyr::filter(games >= 5) # keep only player-season combinations in which QBs played at least 5 games
+  dplyr::filter(
+    games >= 5, # keep only player-season combinations in which QBs played at least 5 games
+    season >= 2011) # keep only seasons since 2011 (when most contract data are available)
 ```
@@
     position_group == "RB",
-    games >= 5)
+    games >= 5, # keep only player-season combinations in which QBs played at least 5 games
+    season >= 2011) # keep only seasons since 2011 (when most contract data are available)
 ```
@@
     position_group %in% c("WR","TE"),
-    games >= 5)
+    games >= 5, # keep only player-season combinations in which QBs played at least 5 games
+    season >= 2011) # keep only seasons since 2011 (when most contract data are available)
 ```

### 2024-07-20 mythbusters.qmd
20240720 - effect size

@@
 library("lmerTest")
+library("MuMIn")
 library("tidyverse")
@@
 summary(mixedModel_qbr)
+MuMIn::r.squaredGLMM(mixedModel_qbr)
 ```
@@
 summary(mixedModel_ptsAdded)
+MuMIn::r.squaredGLMM(mixedModel_ptsAdded)
 ```
@@
 summary(mixedModel_epaPass)
+MuMIn::r.squaredGLMM(mixedModel_epaPass)
 ```
@@
 summary(mixedModel_ypc)
+MuMIn::r.squaredGLMM(mixedModel_ypc)
 ```
@@
 summary(mixedModel_epaRush)
+MuMIn::r.squaredGLMM(mixedModel_epaRush)
 ```
@@
 summary(mixedModel_receivingYards)
+MuMIn::r.squaredGLMM(mixedModel_receivingYards)
 ```
@@
 summary(mixedModel_epaReceiving)
+MuMIn::r.squaredGLMM(mixedModel_epaReceiving)
 ```

### 2024-07-20 mythbusters.qmd
20240720 - effect sizes

--- 2024-07-20 mythbusters.qmd
@@
 library("lmerTest")
+library("MuMIn")
+library("emmeans")
 library("tidyverse")
+```
+
+```{r}
+#| include: false
+
+emm_options(lmerTest.limit = 100000)
+emm_options(pbkrtest.limit = 100000)
 ```
@@
 
+#nfl_actualStatsContracts_offense_weekly$contractYear <- factor(
+#  nfl_actualStatsContracts_offense_weekly$contractYear,
+#  levels = c(0, 1),
+#  labels = c("no", "yes"))
+
+#nfl_actualStatsContracts_offense_seasonal$contractYear <- factor(
+#  nfl_actualStatsContracts_offense_seasonal$contractYear,
+#  levels = c(0, 1),
+#  labels = c("no", "yes"))
+
 nfl_actualStatsContracts_offense_weekly <- nfl_actualStatsContracts_offense_weekly %>% 
@@
   ) %>% 
-  dplyr::mutate(ypc = rushing_yards / carries)
+  dplyr::mutate(
+    ypc = rushing_yards / carries,
+    contractYear = factor(
+      contractYear,
+      levels = c(0, 1),
+      labels = c("no", "yes")
+    ))
 
@@
 nfl_contractYearQBR_weekly$contractYear[which(is.na(nfl_contractYearQBR_weekly$contractYear))] <- 0
+#nfl_contractYearQBR_weekly$contractYear <- factor(
+#  nfl_contractYearQBR_weekly$contractYear,
+#  levels = c(0, 1),
+#  labels = c("no", "yes"))
 
@@
     .groups = "drop_last"
-  )
+  ) %>% 
+  dplyr::mutate(
+    contractYear = factor(
+      contractYear,
+      levels = c(0, 1),
+      labels = c("no", "yes")
+    ))
 
@@
 summary(mixedModel_qbr)
+MuMIn::r.squaredGLMM(mixedModel_qbr)
+emmeans::emmeans(mixedModel_qbr, "contractYear")
 ```
@@
 summary(mixedModel_ptsAdded)
+MuMIn::r.squaredGLMM(mixedModel_ptsAdded)
+emmeans::emmeans(mixedModel_ptsAdded, "contractYear")
 ```
@@
 summary(mixedModel_epaPass)
+MuMIn::r.squaredGLMM(mixedModel_epaPass)
+emmeans::emmeans(mixedModel_epaPass, "contractYear")
 ```
@@
 summary(mixedModel_ypc)
+MuMIn::r.squaredGLMM(mixedModel_ypc)
+emmeans::emmeans(mixedModel_ypc, "contractYear")
 ```
@@
 summary(mixedModel_epaRush)
+MuMIn::r.squaredGLMM(mixedModel_epaRush)
+emmeans::emmeans(mixedModel_epaRush, "contractYear")
 ```
@@
 summary(mixedModel_receivingYards)
+MuMIn::r.squaredGLMM(mixedModel_receivingYards)
+emmeans::emmeans(mixedModel_receivingYards, "contractYear")
 ```
@@
 summary(mixedModel_epaReceiving)
+MuMIn::r.squaredGLMM(mixedModel_epaReceiving)
+emmeans::emmeans(mixedModel_epaReceiving, "contractYear")
 ```

### 2024-07-20 mythbusters.qmd
20240720 - additional caveats

@@
 
-- We will examine seasons since 2011, beginning when most data for player contracts are available.
+- We will determine a player's contract year programmatically based on the year the contract was signed.
+For instance, if a player signed a 3-year contract in 2015, their contract would expire in 2018, and thus their contract year would be 2017.
+Note: this is a coarse way of determining a player's contract year because it could depend on when during the year the player's contract is signed.
+If we were submitting this analysis as a paper to a scientific journal, it would be important to verify each player's contract year.
+- We will examine performance in all seasons since 2011, beginning when most data for player contracts are available.
 - For maximum [statistical power](#sec-statisticalPower) to detect an effect if a contract year effect exists, we will examine all seasons for a player (since 2011), not just their contract year and their pre-contract year.

### 2024-07-20 multiple-regression.qmd
20240720 - assumptions

@@
 In this way, standardized regression coefficients provide a meaningful index of [effect size](#sec-practicalSignificance).
+
+## Assumptions of Multiple Regression {#sec-assumptionsRegression}
+
+Linear regression models make the following assumptions:
+
+- there is a linear association between the predictor variables and the outcome variable
+- there is homoscedasticity of the residuals; the residuals do not differ as a function of the predictor variables or as a function of the outcome variable
+- the residuals are independent; they are uncorrelated with each other
+- the residuals are normally distributed
 
@@
 
+## Impact of Oultiers {#sec-multipleRegressionOutliers}
+
+[As with correlation](#sec-correlationOutliers), multiple regression can be strongly impacted by outliers.
+
 ## Conclusion {#sec-multipleRegressionConclusion}

### 2024-07-20 mythbusters.qmd
20240720 - assumptions

@@
 - To ensure we do not make generalization about a player's performance in a season from a small sample, the player has to play at least 5 games in a given season for that player–season combination to be included in analysis.
+
+For analysis, the same player contributes multiple observations of performance (i.e., multiple seasons) due to the longitudinal nature of the data.
+Inclusion of multiple data points from the same player would violate the [assumption of multiple regression](#sec-assumptionsRegression) that all observations are independent.
+Thus, we use mixed-effects models that allow nonindependent observations.
+In our mixed-effects models, we include a random intercept for each player, to allow our model to account for players' differing level of performance.
 
@@
 
-### QB
+### QB {#sec-contractYearQB}
 
@@
 
-### RB
+### RB {#sec-contractYearRB}
 
@@
 
-### WR/TE
+### WR/TE {#sec-contractYearWRTE}
 
@@
 
-### QB/RB/WR/TE
+### QB/RB/WR/TE {#sec-contractYearQBRBWRTE}
 

### 2024-07-20 mythbusters.qmd
20240720 - player_id

@@
 nfl_actualStatsContracts_seasonal <- nfl_actualStatsContractsSubset_offense_weekly %>% 
-  dplyr::group_by(playerMerge, season) %>% 
+  dplyr::group_by(player_id, season) %>% 
   dplyr::summarise(
+    player_display_name = petersenlab::Mode(player_display_name),
     position_group = petersenlab::Mode(position_group),
@@
   dplyr::mutate(
+    player_id = as.factor(player_id),
     ypc = rushing_yards / carries,
@@
 
+nfl_contractYearQBR_seasonal <- nfl_contractYearQBR_seasonal %>% 
+  group_by(playerMerge) %>%
+  mutate(player_id = as.factor(as.character(cur_group_id())))
+
 nfl_contractYearQBRsubset_seasonal <- nfl_contractYearQBR_seasonal %>% 
@@
 mixedModel_qbr <- lmerTest::lmer(
-  qbr ~ contractYear + (1 | playerMerge),
+  qbr ~ contractYear + (1 | player_id),
   data = nfl_contractYearQBR_seasonal
@@
 mixedModel_ptsAdded <- lmerTest::lmer(
-  pts_added ~ contractYear + (1 | playerMerge),
+  pts_added ~ contractYear + (1 | player_id),
   data = nfl_contractYearQBR_seasonal
@@
 mixedModel_epaPass <- lmerTest::lmer(
-  epa_pass ~ contractYear + (1 | playerMerge),
+  epa_pass ~ contractYear + (1 | player_id),
   data = nfl_contractYearQBR_seasonal
@@
 mixedModel_ypc <- lmerTest::lmer(
-  ypc ~ contractYear + (1 | playerMerge),
+  ypc ~ contractYear + (1 | player_id),
   data = nfl_actualStatsContractsRB_seasonal
@@
 mixedModel_epaRush <- lmerTest::lmer(
-  rushing_epa ~ contractYear + (1 | playerMerge),
+  rushing_epa ~ contractYear + (1 | player_id),
   data = nfl_actualStatsContractsRB_seasonal
@@
 mixedModel_receivingYards <- lmerTest::lmer(
-  receiving_yards ~ contractYear + (1 | playerMerge),
+  receiving_yards ~ contractYear + (1 | player_id),
   data = nfl_actualStatsContractsWRTE_seasonal
@@
 mixedModel_epaReceiving <- lmerTest::lmer(
-  receiving_epa ~ contractYear + (1 | playerMerge),
+  receiving_epa ~ contractYear + (1 | player_id),
   data = nfl_actualStatsContractsWRTE_seasonal

### 2024-07-21 getting-started.qmd
20240719 - remove output
20240721 - save player age and experience

@@
 ```{r}
-#| warning: false
-#| message: false
+#| output: false
 
@@
 
-### Player Age {#sec-calculatePlayerAge}
+### Player Age and Experience {#sec-calculatePlayerAge}
+
+#### Weekly {#sec-calculatePlayerAgeWeekly}
 
@@
 
+# Calculate ageCentered and ageCenteredQuadratic
+player_game_birthdate_gameday_offense$ageCentered20 <- player_game_birthdate_gameday_offense$age - 20
+player_game_birthdate_gameday_offense$ageCentered20Quadratic <- player_game_birthdate_gameday_offense$ageCentered20 ^ 2
+
+player_game_birthdate_gameday_defense$ageCentered20 <- player_game_birthdate_gameday_defense$age - 20
+player_game_birthdate_gameday_defense$ageCentered20Quadratic <- player_game_birthdate_gameday_defense$ageCentered20 ^ 2
+
+player_game_birthdate_gameday_kicking$ageCentered20 <- player_game_birthdate_gameday_kicking$age - 20
+player_game_birthdate_gameday_kicking$ageCentered20Quadratic <- player_game_birthdate_gameday_kicking$ageCentered20 ^ 2
+
 # Merge with player info
@@
 # Merge with player weekly stats
-player_age_stats_offense <- left_join(
+player_stats_weekly_offense <- full_join(
   player_age_offense %>% select(-position, -position_group),
   nfl_actualStats_game_offense_weekly,
-  by = c(c("season","week","player_id","game_id")))
-
-player_age_stats_defense <- left_join(
+  by = c("season","week","player_id","game_id"))
+
+player_stats_weekly_defense <- full_join(
   player_age_defense %>% select(-position, -position_group),
   nfl_actualStats_game_defense_weekly,
-  by = c(c("season","week","player_id","game_id")))
-
-player_age_stats_kicking <- left_join(
+  by = c("season","week","player_id","game_id"))
+
+player_stats_weekly_kicking <- full_join(
   player_age_kicking %>% select(-position, -position_group),
   nfl_actualStats_game_kicking_weekly,
-  by = c(c("season","week","player_id","game_id")))
-
-player_age_stats_offense$years_of_experience <- as.integer(player_age_stats_offense$years_of_experience)
-player_age_stats_defense$years_of_experience <- as.integer(player_age_stats_defense$years_of_experience)
-player_age_stats_kicking$years_of_experience <- as.integer(player_age_stats_kicking$years_of_experience)
-
+  by = c("season","week","player_id","game_id"))
+
+player_stats_weekly_offense$total_years_of_experience <- as.integer(player_stats_weekly_offense$years_of_experience)
+player_stats_weekly_defense$total_years_of_experience <- as.integer(player_stats_weekly_defense$years_of_experience)
+player_stats_weekly_kicking$total_years_of_experience <- as.integer(player_stats_weekly_kicking$years_of_experience)
+
+player_stats_weekly_offense$years_of_experience <- NULL
+player_stats_weekly_defense$years_of_experience <- NULL
+player_stats_weekly_kicking$years_of_experience <- NULL
+
+distinct_seasons_offense <- player_stats_weekly_offense %>%
+  select(player_id, season) %>%
+  distinct() %>% 
+  left_join(
+    nfl_players[,c("gsis_id","years_of_experience")],
+    by = c("player_id" = "gsis_id")
+  ) %>% 
+  mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
+  select(-years_of_experience)
+
+distinct_seasons_defense <- player_stats_weekly_defense %>%
+  select(player_id, season) %>%
+  distinct() %>% 
+  left_join(
+    nfl_players[,c("gsis_id","years_of_experience")],
+    by = c("player_id" = "gsis_id")
+  ) %>% 
+  mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
+  select(-years_of_experience)
+
+distinct_seasons_kicking <- player_stats_weekly_kicking %>%
+  select(player_id, season) %>%
+  distinct() %>% 
+  left_join(
+    nfl_players[,c("gsis_id","years_of_experience")],
+    by = c("player_id" = "gsis_id")
+  ) %>% 
+  mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
+  select(-years_of_experience)
+
+years_of_experience_offense <- distinct_seasons_offense %>% 
+  arrange(player_id, -season) %>% 
+  group_by(player_id) %>%
+  mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
+  ungroup()
+
+years_of_experience_defense <- distinct_seasons_defense %>% 
+  arrange(player_id, -season) %>% 
+  group_by(player_id) %>%
+  mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
+  ungroup()
+
+years_of_experience_kicking <- distinct_seasons_kicking %>% 
+  arrange(player_id, -season) %>% 
+  group_by(player_id) %>%
+  mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
+  ungroup()
+
+years_of_experience_offense$years_of_experience[which(years_of_experience_offense$years_of_experience < 0)] <- 0
+years_of_experience_defense$years_of_experience[which(years_of_experience_defense$years_of_experience < 0)] <- 0
+years_of_experience_kicking$years_of_experience[which(years_of_experience_kicking$years_of_experience < 0)] <- 0
+
+player_stats_weekly_offense <- player_stats_weekly_offense %>% 
+  left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+
+player_stats_weekly_defense <- player_stats_weekly_defense %>% 
+  left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+
+player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
+  left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+
+# Save data
+save(
+  player_stats_weekly_offense, player_stats_weekly_defense, player_stats_weekly_kicking,
+  file = "./data/player_stats_weekly.RData"
+)
+```
+
+#### Seasonal {#sec-calculatePlayerAgeSeasonal}
+
+```{r}
 # Merge player info with seasonal stats
-player_seasonal_offense <- left_join(
+player_stats_seasonal_offense <- full_join(
   nfl_actualStats_offense_seasonal,
@@
 
-player_seasonal_defense <- left_join(
+player_stats_seasonal_defense <- full_join(
   nfl_actualStats_defense_seasonal,
@@
 
-player_seasonal_kicking <- left_join(
+player_stats_seasonal_kicking <- full_join(
   nfl_actualStats_kicking_seasonal,
@@
 
-player_seasonal_offense <- player_seasonal_offense %>% 
+player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
   left_join(
@@
 
-player_seasonal_defense <- player_seasonal_defense %>% 
+player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
   left_join(
@@
 
-player_seasonal_kicking <- player_seasonal_kicking %>% 
+player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
   left_join(
@@
 
-player_seasonal_offense$age <- interval(
-  start = player_seasonal_offense$birth_date,
-  end = player_seasonal_offense$startdate
+player_stats_seasonal_offense$age <- interval(
+  start = player_stats_seasonal_offense$birth_date,
+  end = player_stats_seasonal_offense$startdate
 ) %>% 
@@
 
-player_seasonal_defense$age <- interval(
-  start = player_seasonal_defense$birth_date,
-  end = player_seasonal_defense$startdate
+player_stats_seasonal_defense$age <- interval(
+  start = player_stats_seasonal_defense$birth_date,
+  end = player_stats_seasonal_defense$startdate
 ) %>% 
@@
 
-player_seasonal_kicking$age <- interval(
-  start = player_seasonal_kicking$birth_date,
-  end = player_seasonal_kicking$startdate
+player_stats_seasonal_kicking$age <- interval(
+  start = player_stats_seasonal_kicking$birth_date,
+  end = player_stats_seasonal_kicking$startdate
 ) %>% 
   time_length(unit = "years")
+
+# Calculate ageCentered and ageCenteredQuadratic
+player_stats_seasonal_offense$ageCentered20 <- player_stats_seasonal_offense$age - 20
+player_stats_seasonal_offense$ageCentered20Quadratic <- player_stats_seasonal_offense$ageCentered20 ^ 2
+
+player_stats_seasonal_defense$ageCentered20 <- player_stats_seasonal_defense$age - 20
+player_stats_seasonal_defense$ageCentered20Quadratic <- player_stats_seasonal_defense$ageCentered20 ^ 2
+
+player_stats_seasonal_kicking$ageCentered20 <- player_stats_seasonal_kicking$age - 20
+player_stats_seasonal_kicking$ageCentered20Quadratic <- player_stats_seasonal_kicking$ageCentered20 ^ 2
+
+# Years of experience
+player_stats_seasonal_offense$years_of_experience <- NULL
+player_stats_seasonal_defense$years_of_experience <- NULL
+player_stats_seasonal_kicking$years_of_experience <- NULL
+
+player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+  left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+
+player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
+  left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+
+player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
+  left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+
+# Save data
+save(
+  player_stats_seasonal_offense, player_stats_seasonal_defense, player_stats_seasonal_kicking,
+  file = "./data/player_stats_seasonal.RData"
+)
 ```

### 2024-07-21 mythbusters.qmd
20240721 - account for player age and experience

@@
 load(file = "./data/nfl_playerContracts.RData")
-load(file = "./data/nfl_actualStats_weekly.RData")
-load(file = "./data/nfl_actualStats_seasonal.RData")
+load(file = "./data/player_stats_weekly.RData") #load(file = "./data/nfl_actualStats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData") #load(file = "./data/nfl_actualStats_seasonal.RData")
 load(file = "./data/nfl_espnQBR.RData")
 ```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+```
+
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
 
@@
 In our mixed-effects models, we include a random intercept for each player, to allow our model to account for players' differing level of performance.
+We examine two mixed-effects models for each outcome variable.
+One that accounts for the effects of age and experience, and one that does not.
+The model that accounts for the effects of age and experience includes random intercepts and random linear slopes, along with a fixed quadratic effect of age to allow for curvilinear effects and a fixed effect of experience.
 
@@
 # Merge with weekly and seasonal stats data
-nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly %>% 
+player_stats_weekly_offense <- player_stats_weekly_offense %>% 
   dplyr::mutate(playerMerge = toupper(str_replace_all(player_display_name, "[[:punct:]]", "")))
@@
 
-nfl_actualStatsContracts_offense_weekly <- dplyr::full_join(
-  nfl_actualStats_offense_weekly,
+player_statsContracts_offense_weekly <- dplyr::full_join(
+  player_stats_weekly_offense,
   nfl_playerContractYears,
@@
 
-#nfl_actualStatsContracts_offense_seasonal <- full_join(
-#  nfl_actualStats_offense_seasonal,
+#player_statsContracts_offense_seasonal <- full_join(
+#  player_stats_seasonal_offense,
 #  nfl_playerContractYears,
@@
 
-nfl_actualStatsContracts_offense_weekly$contractYear[which(is.na(nfl_actualStatsContracts_offense_weekly$contractYear))] <- 0
-#nfl_actualStatsContracts_offense_seasonal$contractYear[which(is.na(nfl_actualStatsContracts_offense_seasonal$contractYear))] <- 0
+player_statsContracts_offense_weekly$contractYear[which(is.na(player_statsContracts_offense_weekly$contractYear))] <- 0
+#player_statsContracts_offense_seasonal$contractYear[which(is.na(player_statsContracts_offense_seasonal$contractYear))] <- 0
 
-#nfl_actualStatsContracts_offense_weekly$contractYear <- factor(
-#  nfl_actualStatsContracts_offense_weekly$contractYear,
+#player_statsContracts_offense_weekly$contractYear <- factor(
+#  player_statsContracts_offense_weekly$contractYear,
 #  levels = c(0, 1),
@@
 
-#nfl_actualStatsContracts_offense_seasonal$contractYear <- factor(
-#  nfl_actualStatsContracts_offense_seasonal$contractYear,
+#player_statsContracts_offense_seasonal$contractYear <- factor(
+#  player_statsContracts_offense_seasonal$contractYear,
 #  levels = c(0, 1),
@@
 
-nfl_actualStatsContracts_offense_weekly <- nfl_actualStatsContracts_offense_weekly %>% 
+player_statsContracts_offense_weekly <- player_statsContracts_offense_weekly %>% 
   dplyr::arrange(playerMerge, season, season_type, week)
 
-#nfl_actualStatsContracts_offense_seasonal <- nfl_actualStatsContracts_offense_seasonal %>% 
+#player_statsContracts_offense_seasonal <- player_statsContracts_offense_seasonal %>% 
 #  arrange(playerMerge, season)
 
-nfl_actualStatsContractsSubset_offense_weekly <- nfl_actualStatsContracts_offense_weekly %>% 
+player_statsContractsSubset_offense_weekly <- player_statsContracts_offense_weekly %>% 
   dplyr::filter(season_type == "REG")
@@
 # Calculate Per Game Totals
-nfl_actualStatsContracts_seasonal <- nfl_actualStatsContractsSubset_offense_weekly %>% 
+player_statsContracts_seasonal <- player_statsContractsSubset_offense_weekly %>% 
   dplyr::group_by(player_id, season) %>% 
@@
     position_group = petersenlab::Mode(position_group),
+    age = min(age, na.rm = TRUE),
+    years_of_experience = min(years_of_experience, na.rm = TRUE),
     rushing_yards = sum(rushing_yards, na.rm = TRUE), # season total
@@
     ))
+
+player_statsContracts_seasonal$age[which(is.infinite((player_statsContracts_seasonal$age)))] <- NA
+player_statsContracts_seasonal$years_of_experience[which(is.infinite((player_statsContracts_seasonal$years_of_experience)))] <- NA
+
+player_statsContracts_seasonal$ageCentered20 <- player_statsContracts_seasonal$age - 20
+player_statsContracts_seasonal$ageCentered20Quadratic <- player_statsContracts_seasonal$ageCentered20 ^ 2
 
@@
 
+# Merge with age and experience
+nfl_contractYearQBRsubset_weekly <- player_statsContractsSubset_offense_weekly %>% 
+  dplyr::filter(position == "QB") %>% 
+  dplyr::select(playerMerge, season, week, age, years_of_experience) %>% 
+  full_join(
+    nfl_contractYearQBRsubset_weekly,
+    by = c("playerMerge","season", c("week" = "game_week"))
+  ) %>% 
+  arrange(player_id, season, week)
+
 #hist(nfl_contractYearQBRsubset_weekly$qb_plays) # players have at least 20 dropbacks per game
@@
   dplyr::summarise(
+    age = min(age, na.rm = TRUE),
+    years_of_experience = min(years_of_experience, na.rm = TRUE),
     qbr = mean(qbr_total, na.rm = TRUE),
@@
     ))
+
+nfl_contractYearQBR_seasonal$age[which(is.infinite((nfl_contractYearQBR_seasonal$age)))] <- NA
+nfl_contractYearQBR_seasonal$years_of_experience[which(is.infinite((nfl_contractYearQBR_seasonal$years_of_experience)))] <- NA
+
+nfl_contractYearQBR_seasonal$ageCentered20 <- nfl_contractYearQBR_seasonal$age - 20
+nfl_contractYearQBR_seasonal$ageCentered20Quadratic <- nfl_contractYearQBR_seasonal$ageCentered20 ^ 2
 
@@
   qbr ~ contractYear + (1 | player_id),
-  data = nfl_contractYearQBR_seasonal
+  data = nfl_contractYearQBR_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
 )
@@
 emmeans::emmeans(mixedModel_qbr, "contractYear")
+
+mixedModelAge_qbr <- lmerTest::lmer(
+  qbr ~ contractYear + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
+  data = nfl_contractYearQBR_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_qbr)
+MuMIn::r.squaredGLMM(mixedModelAge_qbr)
+emmeans::emmeans(mixedModelAge_qbr, "contractYear")
 ```
@@
   pts_added ~ contractYear + (1 | player_id),
-  data = nfl_contractYearQBR_seasonal
+  data = nfl_contractYearQBR_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
 )
@@
 emmeans::emmeans(mixedModel_ptsAdded, "contractYear")
+
+mixedModelAge_ptsAdded <- lmerTest::lmer(
+  pts_added ~ contractYear + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
+  data = nfl_contractYearQBR_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_ptsAdded)
+MuMIn::r.squaredGLMM(mixedModelAge_ptsAdded)
+emmeans::emmeans(mixedModelAge_ptsAdded, "contractYear")
 ```
@@
   epa_pass ~ contractYear + (1 | player_id),
-  data = nfl_contractYearQBR_seasonal
+  data = nfl_contractYearQBR_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
 )
@@
 emmeans::emmeans(mixedModel_epaPass, "contractYear")
+
+mixedModelAge_epaPass <- lmerTest::lmer(
+  epa_pass ~ contractYear + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 | player_id), # removed random slopes to address convergence issue
+  data = nfl_contractYearQBR_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_epaPass)
+MuMIn::r.squaredGLMM(mixedModelAge_epaPass)
+emmeans::emmeans(mixedModelAge_epaPass, "contractYear")
 ```
@@
 
-nfl_actualStatsContractsRB_seasonal <- nfl_actualStatsContracts_seasonal %>% 
+player_statsContractsRB_seasonal <- player_statsContracts_seasonal %>% 
   dplyr::filter(
@@
   ypc ~ contractYear + (1 | player_id),
-  data = nfl_actualStatsContractsRB_seasonal
+  data = player_statsContractsRB_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
 )
@@
 emmeans::emmeans(mixedModel_ypc, "contractYear")
+
+mixedModelAge_ypc <- lmerTest::lmer(
+  ypc ~ contractYear + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
+  data = player_statsContractsRB_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_ypc)
+MuMIn::r.squaredGLMM(mixedModelAge_ypc)
+emmeans::emmeans(mixedModelAge_ypc, "contractYear")
 ```
@@
   rushing_epa ~ contractYear + (1 | player_id),
-  data = nfl_actualStatsContractsRB_seasonal
+  data = player_statsContractsRB_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
 )
@@
 emmeans::emmeans(mixedModel_epaRush, "contractYear")
+
+mixedModelAge_epaRush <- lmerTest::lmer(
+  rushing_epa ~ contractYear + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
+  data = player_statsContractsRB_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_epaRush)
+MuMIn::r.squaredGLMM(mixedModelAge_epaRush)
+emmeans::emmeans(mixedModelAge_epaRush, "contractYear")
 ```
@@
 
-nfl_actualStatsContractsWRTE_seasonal <- nfl_actualStatsContracts_seasonal %>% 
+player_statsContractsWRTE_seasonal <- player_statsContracts_seasonal %>% 
   dplyr::filter(
@@
   receiving_yards ~ contractYear + (1 | player_id),
-  data = nfl_actualStatsContractsWRTE_seasonal
+  data = player_statsContractsWRTE_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
 )
@@
 emmeans::emmeans(mixedModel_receivingYards, "contractYear")
+
+mixedModelAge_receivingYards <- lmerTest::lmer(
+  receiving_yards ~ contractYear + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
+  data = player_statsContractsWRTE_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_receivingYards)
+MuMIn::r.squaredGLMM(mixedModelAge_receivingYards)
+emmeans::emmeans(mixedModelAge_receivingYards, "contractYear")
 ```
@@
   receiving_epa ~ contractYear + (1 | player_id),
-  data = nfl_actualStatsContractsWRTE_seasonal
+  data = player_statsContractsWRTE_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
 )
@@
 emmeans::emmeans(mixedModel_epaReceiving, "contractYear")
+
+mixedModelAge_epaReceiving <- lmerTest::lmer(
+  receiving_epa ~ contractYear + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
+  data = player_statsContractsWRTE_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_epaReceiving)
+MuMIn::r.squaredGLMM(mixedModelAge_epaReceiving)
+emmeans::emmeans(mixedModelAge_epaReceiving, "contractYear")
 ```

### 2024-07-21 mixed-models.qmd
20240721 - mixed models

# Mixed Models {#sec-mixedModels}

## Getting Started {#sec-mixedModelsGettingStarted}

### Load Packages {#sec-mixedModelsLoadPackages}

```{r}

```

## Overview of Mixed Models {#sec-mixedModelsOverview}

## Conclusion {#sec-mixedModelsConclusion}

::: {.content-visible when-format="html"}

## Session Info {#sec-mixedModelsSessionInfo}

```{r}
sessionInfo()
```

:::

### 2024-07-21 mythbusters.qmd
20240721 - model specification

@@
 In our mixed-effects models, we include a random intercept for each player, to allow our model to account for players' differing level of performance.
-We examine two mixed-effects models for each outcome variable.
-One that accounts for the effects of age and experience, and one that does not.
-The model that accounts for the effects of age and experience includes random intercepts and random linear slopes, along with a fixed quadratic effect of age to allow for curvilinear effects and a fixed effect of experience.
+We examine two mixed-effects models for each outcome variable: one model that accounts for the effects of age and experience, and one model that does not.
+
+The model that does not account for the effects of age and experience includes:
+
+a) random intercepts to allow the model to estimate a different starting point for each player
+a) a fixed effect for whether the player is in a contract year
+
+The model that accounts for the effects of age and experience includes:
+
+a) random intercepts to allow the model to estimate a different starting point for each player
+a) random linear slopes (i.e., random effect of linear age) to allow the model to estimate a different form of change for each player
+a) a fixed quadratic effect of age to allow for curvilinear effects
+a) a fixed effect of experience
+a) a fixed effect for whether the player is in a contract year
 

### 2024-07-21 getting-started.qmd
20240721 - cleanup

--- 2024-07-19 getting-started.qmd
@@
   nfl_pbp_seasonalList[[i]] <- nfl_pbp %>% 
-    filter(season == seasons[i])
+    dplyr::filter(season == seasons[i])
 
@@
 nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonalList %>% 
-  bind_rows()
+  dplyr::bind_rows()
 nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonalList %>% 
-  bind_rows()
+  dplyr::bind_rows()
 nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonalList %>% 
-  bind_rows()
+  dplyr::bind_rows()
 ```
@@
 nfl_actualStats_offense_weekly_long <- nfl_actualStats_offense_weekly %>% 
-  pivot_longer(
+  tidyr::pivot_longer(
     cols = c(recent_team, opponent_team),
@@
 # Perform separate inner join operations for the home_team and away_team
-nfl_actualStats_offense_weekly_home <- inner_join(
+nfl_actualStats_offense_weekly_home <- dplyr::inner_join(
   nfl_actualStats_offense_weekly_long,
@@
 
-nfl_actualStats_offense_weekly_away <- inner_join(
+nfl_actualStats_offense_weekly_away <- dplyr::inner_join(
   nfl_actualStats_offense_weekly_long,
@@
 
-nfl_actualStats_defense_weekly_home <- inner_join(
+nfl_actualStats_defense_weekly_home <- dplyr::inner_join(
   nfl_actualStats_defense_weekly,
@@
 
-nfl_actualStats_defense_weekly_away <- inner_join(
+nfl_actualStats_defense_weekly_away <- dplyr::inner_join(
   nfl_actualStats_defense_weekly,
@@
 
-nfl_actualStats_kicking_weekly_home <- inner_join(
+nfl_actualStats_kicking_weekly_home <- dplyr::inner_join(
   nfl_actualStats_kicking_weekly,
@@
 
-nfl_actualStats_kicking_weekly_away <- inner_join(
+nfl_actualStats_kicking_weekly_away <- dplyr::inner_join(
   nfl_actualStats_kicking_weekly,
@@
 # Combine the results of the join operations
-nfl_actualStats_offense_weekly_schedules_long <- bind_rows(
+nfl_actualStats_offense_weekly_schedules_long <- dplyr::bind_rows(
   nfl_actualStats_offense_weekly_home,
@@
 
-nfl_actualStats_defense_weekly_schedules_long <- bind_rows(
+nfl_actualStats_defense_weekly_schedules_long <- dplyr::bind_rows(
   nfl_actualStats_defense_weekly_home,
@@
 
-nfl_actualStats_kicking_weekly_schedules_long <- bind_rows(
+nfl_actualStats_kicking_weekly_schedules_long <- dplyr::bind_rows(
   nfl_actualStats_kicking_weekly_home,
@@
 player_game_gameday_offense <- nfl_actualStats_offense_weekly_schedules_long %>%
-  distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
-  pivot_wider(
+  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
+  tidyr::pivot_wider(
     names_from = home_away,
@@
 player_game_gameday_defense <- nfl_actualStats_defense_weekly_schedules_long %>%
-  distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
-  pivot_wider(
+  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
+  tidyr::pivot_wider(
     names_from = home_away,
@@
 player_game_gameday_kicking <- nfl_actualStats_kicking_weekly_schedules_long %>%
-  distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
-  pivot_wider(
+  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
+  tidyr::pivot_wider(
     names_from = home_away,
@@
 # Merge player birthdate and the game date
-player_game_birthdate_gameday_offense <- left_join(
+player_game_birthdate_gameday_offense <- dplyr::left_join(
   player_game_gameday_offense,
@@
 
-player_game_birthdate_gameday_defense <- left_join(
+player_game_birthdate_gameday_defense <- dplyr::left_join(
   player_game_gameday_defense,
@@
 
-player_game_birthdate_gameday_kicking <- left_join(
+player_game_birthdate_gameday_kicking <- dplyr::left_join(
   player_game_gameday_kicking,
@@
 
-player_game_birthdate_gameday_offense$birth_date <- ymd(player_game_birthdate_gameday_offense$birth_date)
-player_game_birthdate_gameday_offense$gameday <- ymd(player_game_birthdate_gameday_offense$gameday)
-
-player_game_birthdate_gameday_defense$birth_date <- ymd(player_game_birthdate_gameday_defense$birth_date)
-player_game_birthdate_gameday_defense$gameday <- ymd(player_game_birthdate_gameday_defense$gameday)
-
-player_game_birthdate_gameday_kicking$birth_date <- ymd(player_game_birthdate_gameday_kicking$birth_date)
-player_game_birthdate_gameday_kicking$gameday <- ymd(player_game_birthdate_gameday_kicking$gameday)
+player_game_birthdate_gameday_offense$birth_date <- lubridate::ymd(player_game_birthdate_gameday_offense$birth_date)
+player_game_birthdate_gameday_offense$gameday <- lubridate::ymd(player_game_birthdate_gameday_offense$gameday)
+
+player_game_birthdate_gameday_defense$birth_date <- lubridate::ymd(player_game_birthdate_gameday_defense$birth_date)
+player_game_birthdate_gameday_defense$gameday <- lubridate::ymd(player_game_birthdate_gameday_defense$gameday)
+
+player_game_birthdate_gameday_kicking$birth_date <- lubridate::ymd(player_game_birthdate_gameday_kicking$birth_date)
+player_game_birthdate_gameday_kicking$gameday <- lubridate::ymd(player_game_birthdate_gameday_kicking$gameday)
 
 # Calculate player's age for a given week as the difference between their birthdate and the game date
-player_game_birthdate_gameday_offense$age <- interval(
+player_game_birthdate_gameday_offense$age <- lubridate::interval(
   start = player_game_birthdate_gameday_offense$birth_date,
@@
 ) %>% 
-  time_length(unit = "years")
-
-player_game_birthdate_gameday_defense$age <- interval(
+  lubridate::time_length(unit = "years")
+
+player_game_birthdate_gameday_defense$age <- lubridate::interval(
   start = player_game_birthdate_gameday_defense$birth_date,
@@
 ) %>% 
-  time_length(unit = "years")
-
-player_game_birthdate_gameday_kicking$age <- interval(
+  lubridate::time_length(unit = "years")
+
+player_game_birthdate_gameday_kicking$age <- lubridate::interval(
   start = player_game_birthdate_gameday_kicking$birth_date,
@@
 ) %>% 
-  time_length(unit = "years")
+  lubridate::time_length(unit = "years")
+
+# Calculate ageCentered and ageCenteredQuadratic
+player_game_birthdate_gameday_offense$ageCentered20 <- player_game_birthdate_gameday_offense$age - 20
+player_game_birthdate_gameday_offense$ageCentered20Quadratic <- player_game_birthdate_gameday_offense$ageCentered20 ^ 2
+
+player_game_birthdate_gameday_defense$ageCentered20 <- player_game_birthdate_gameday_defense$age - 20
+player_game_birthdate_gameday_defense$ageCentered20Quadratic <- player_game_birthdate_gameday_defense$ageCentered20 ^ 2
+
+player_game_birthdate_gameday_kicking$ageCentered20 <- player_game_birthdate_gameday_kicking$age - 20
+player_game_birthdate_gameday_kicking$ageCentered20Quadratic <- player_game_birthdate_gameday_kicking$ageCentered20 ^ 2
 
 # Merge with player info
-player_age_offense <-  left_join(
+player_age_offense <- dplyr::left_join(
   player_game_birthdate_gameday_offense,
@@
 
-player_age_defense <-  left_join(
+player_age_defense <- dplyr::left_join(
   player_game_birthdate_gameday_defense,
@@
 
-player_age_kicking <-  left_join(
+player_age_kicking <- dplyr::left_join(
   player_game_birthdate_gameday_kicking,
@@
 nfl_actualStats_game_offense_weekly <- nfl_actualStats_offense_weekly %>% 
-  left_join(
+  dplyr::left_join(
     player_age_offense[,c("season","week","player_id","game_id")],
@@
 nfl_actualStats_game_defense_weekly <- nfl_actualStats_defense_weekly %>% 
-  left_join(
+  dplyr::left_join(
     player_age_offense[,c("season","week","player_id","game_id")],
@@
 nfl_actualStats_game_kicking_weekly <- nfl_actualStats_kicking_weekly %>% 
-  left_join(
+  dplyr::left_join(
     player_age_offense[,c("season","week","player_id","game_id")],
@@
 # Merge with player weekly stats
-player_age_stats_offense <- left_join(
+player_stats_weekly_offense <- dplyr::full_join(
   player_age_offense %>% select(-position, -position_group),
   nfl_actualStats_game_offense_weekly,
-  by = c(c("season","week","player_id","game_id")))
-
-player_age_stats_defense <- left_join(
+  by = c("season","week","player_id","game_id"))
+
+player_stats_weekly_defense <- dplyr::full_join(
   player_age_defense %>% select(-position, -position_group),
   nfl_actualStats_game_defense_weekly,
-  by = c(c("season","week","player_id","game_id")))
-
-player_age_stats_kicking <- left_join(
+  by = c("season","week","player_id","game_id"))
+
+player_stats_weekly_kicking <- dplyr::full_join(
   player_age_kicking %>% select(-position, -position_group),
   nfl_actualStats_game_kicking_weekly,
-  by = c(c("season","week","player_id","game_id")))
-
-player_age_stats_offense$years_of_experience <- as.integer(player_age_stats_offense$years_of_experience)
-player_age_stats_defense$years_of_experience <- as.integer(player_age_stats_defense$years_of_experience)
-player_age_stats_kicking$years_of_experience <- as.integer(player_age_stats_kicking$years_of_experience)
-
+  by = c("season","week","player_id","game_id"))
+
+player_stats_weekly_offense$total_years_of_experience <- as.integer(player_stats_weekly_offense$years_of_experience)
+player_stats_weekly_defense$total_years_of_experience <- as.integer(player_stats_weekly_defense$years_of_experience)
+player_stats_weekly_kicking$total_years_of_experience <- as.integer(player_stats_weekly_kicking$years_of_experience)
+
+player_stats_weekly_offense$years_of_experience <- NULL
+player_stats_weekly_defense$years_of_experience <- NULL
+player_stats_weekly_kicking$years_of_experience <- NULL
+
+distinct_seasons_offense <- player_stats_weekly_offense %>%
+  dplyr::select(player_id, season) %>%
+  dplyr::distinct() %>% 
+  dplyr::left_join(
+    nfl_players[,c("gsis_id","years_of_experience")],
+    by = c("player_id" = "gsis_id")
+  ) %>% 
+  dplyr::mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
+  dplyr::select(-years_of_experience)
+
+distinct_seasons_defense <- player_stats_weekly_defense %>%
+  dplyr::select(player_id, season) %>%
+  dplyr::distinct() %>% 
+  dplyr::left_join(
+    nfl_players[,c("gsis_id","years_of_experience")],
+    by = c("player_id" = "gsis_id")
+  ) %>% 
+  dplyr::mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
+  dplyr::select(-years_of_experience)
+
+distinct_seasons_kicking <- player_stats_weekly_kicking %>%
+  dplyr::select(player_id, season) %>%
+  dplyr::distinct() %>% 
+  dplyr::left_join(
+    nfl_players[,c("gsis_id","years_of_experience")],
+    by = c("player_id" = "gsis_id")
+  ) %>% 
+  dplyr::mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
+  dplyr::select(-years_of_experience)
+
+years_of_experience_offense <- distinct_seasons_offense %>% 
+  dplyr::arrange(player_id, -season) %>% 
+  dplyr::group_by(player_id) %>%
+  dplyr::mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
+  dplyr::ungroup()
+
+years_of_experience_defense <- distinct_seasons_defense %>% 
+  dplyr::arrange(player_id, -season) %>% 
+  dplyr::group_by(player_id) %>%
+  dplyr::mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
+  dplyr::ungroup()
+
+years_of_experience_kicking <- distinct_seasons_kicking %>% 
+  dplyr::arrange(player_id, -season) %>% 
+  dplyr::group_by(player_id) %>%
+  dplyr::mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
+  dplyr::ungroup()
+
+years_of_experience_offense$years_of_experience[which(years_of_experience_offense$years_of_experience < 0)] <- 0
+years_of_experience_defense$years_of_experience[which(years_of_experience_defense$years_of_experience < 0)] <- 0
+years_of_experience_kicking$years_of_experience[which(years_of_experience_kicking$years_of_experience < 0)] <- 0
+
+player_stats_weekly_offense <- player_stats_weekly_offense %>% 
+  dplyr::left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+
+player_stats_weekly_defense <- player_stats_weekly_defense %>% 
+  dplyr::left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+
+player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
+  dplyr::left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+```
+
+```{r}
+# Save data
+save(
+  player_stats_weekly_offense, player_stats_weekly_defense, player_stats_weekly_kicking,
+  file = "./data/player_stats_weekly.RData"
+)
+```
+
+#### Seasonal {#sec-calculatePlayerAgeSeasonal}
+
+```{r}
 # Merge player info with seasonal stats
-player_seasonal_offense <- left_join(
+player_stats_seasonal_offense <- dplyr::full_join(
   nfl_actualStats_offense_seasonal,
@@
 
-player_seasonal_defense <- left_join(
+player_stats_seasonal_defense <- dplyr::full_join(
   nfl_actualStats_defense_seasonal,
@@
 
-player_seasonal_kicking <- left_join(
+player_stats_seasonal_kicking <- dplyr::full_join(
   nfl_actualStats_kicking_seasonal,
@@
 season_startdate <- nfl_schedules %>% 
-  group_by(season) %>% 
-  summarise(startdate = min(gameday, na.rm = TRUE))
-
-player_seasonal_offense <- player_seasonal_offense %>% 
-  left_join(
+  dplyr::group_by(season) %>% 
+  dplyr::summarise(startdate = min(gameday, na.rm = TRUE))
+
+player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+  dplyr::left_join(
     season_startdate,
@@
 
-player_seasonal_defense <- player_seasonal_defense %>% 
-  left_join(
+player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
+  dplyr::left_join(
     season_startdate,
@@
 
-player_seasonal_kicking <- player_seasonal_kicking %>% 
-  left_join(
+player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
+  dplyr::left_join(
     season_startdate,
@@
 
-player_seasonal_offense$age <- interval(
-  start = player_seasonal_offense$birth_date,
-  end = player_seasonal_offense$startdate
+player_stats_seasonal_offense$age <- lubridate::interval(
+  start = player_stats_seasonal_offense$birth_date,
+  end = player_stats_seasonal_offense$startdate
 ) %>% 
-  time_length(unit = "years")
-
-player_seasonal_defense$age <- interval(
-  start = player_seasonal_defense$birth_date,
-  end = player_seasonal_defense$startdate
+  lubridate::time_length(unit = "years")
+
+player_stats_seasonal_defense$age <- lubridate::interval(
+  start = player_stats_seasonal_defense$birth_date,
+  end = player_stats_seasonal_defense$startdate
 ) %>% 
-  time_length(unit = "years")
-
-player_seasonal_kicking$age <- interval(
-  start = player_seasonal_kicking$birth_date,
-  end = player_seasonal_kicking$startdate
+  lubridate::time_length(unit = "years")
+
+player_stats_seasonal_kicking$age <- lubridate::interval(
+  start = player_stats_seasonal_kicking$birth_date,
+  end = player_stats_seasonal_kicking$startdate
 ) %>% 
-  time_length(unit = "years")
+  lubridate::time_length(unit = "years")
+
+# Calculate ageCentered and ageCenteredQuadratic
+player_stats_seasonal_offense$ageCentered20 <- player_stats_seasonal_offense$age - 20
+player_stats_seasonal_offense$ageCentered20Quadratic <- player_stats_seasonal_offense$ageCentered20 ^ 2
+
+player_stats_seasonal_defense$ageCentered20 <- player_stats_seasonal_defense$age - 20
+player_stats_seasonal_defense$ageCentered20Quadratic <- player_stats_seasonal_defense$ageCentered20 ^ 2
+
+player_stats_seasonal_kicking$ageCentered20 <- player_stats_seasonal_kicking$age - 20
+player_stats_seasonal_kicking$ageCentered20Quadratic <- player_stats_seasonal_kicking$ageCentered20 ^ 2
+
+# Years of experience
+player_stats_seasonal_offense$years_of_experience <- NULL
+player_stats_seasonal_defense$years_of_experience <- NULL
+player_stats_seasonal_kicking$years_of_experience <- NULL
+
+player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+  dplyr::left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+
+player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
+  dplyr::left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+
+player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
+  dplyr::left_join(
+    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    by = c("player_id","season")
+  )
+```
+
+```{r}
+# Save data
+save(
+  player_stats_seasonal_offense, player_stats_seasonal_defense, player_stats_seasonal_kicking,
+  file = "./data/player_stats_seasonal.RData"
+)
 ```
@@
 # Subset Data
-rb_seasonal <- player_seasonal_offense %>% 
+rb_seasonal <- player_stats_seasonal_offense %>% 
   filter(position_group == "RB")

### 2024-07-21 mythbusters.qmd
20240721 - cleanup

@@
 load(file = "./data/nfl_playerContracts.RData")
-load(file = "./data/player_stats_weekly.RData") #load(file = "./data/nfl_actualStats_weekly.RData")
-load(file = "./data/player_stats_seasonal.RData") #load(file = "./data/nfl_actualStats_seasonal.RData")
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
 load(file = "./data/nfl_espnQBR.RData")

### 2024-07-21 data-visualization.qmd
Built site for gh-pages
20240627 - sessionInfo() for reproducibility
20240703 - conclusion section
20240721 - data visualization

--- 2024-06-05 base-rates.qmd
@@
-# Base Rates {#sec-baseRates}
+# Data Visualization {#sec-dataVisualization}
 
-## Getting Started
+## Getting Started {#sec-dataVisualizationGettingStarted}
 
-### Load Packages
+### Load Packages {#sec-dataVisualizationLoadPackages}
 
 ```{r}
+library("tidyverse")
+library("nflplotR")
+```
 
+### Load Data {#sec-dataVisualizationLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
 ```
+
+```{r}
+load(file = "./data/nfl_pbp.RData")
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+```
+
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
+
+## Scatterplot {#sec-scatterplot}
+
+First, we prepare the data:
+
+```{r}
+# Subset Data
+rb_seasonal <- player_stats_seasonal_offense %>% 
+  filter(position_group == "RB")
+```
+
+Second, we create the base plot using the `ggplot()` function of the `ggplot2` package.
+We specify the data object and the variables in the data object that are associated with the x- and y-axes:
+
+```{r}
+ggplot2::ggplot(
+  data = rb_seasonal, # specify data object
+  aes(
+    x = age, # specify variable on x-axis
+    y = rushing_yards)) # specify variable on y-axis
+```
+
+Third, we create a scatterplot using the `geom_point()` function from the `ggplot2` package:
+
+```{r}
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point() # add points for scatterplot
+```
+
+Fourth, we add a linear best-fit line using the `geom_smooth()`:
+
+```{r}
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point() +
+  geom_smooth(method = "lm") # add linear best-fit line
+```
+
+We could also estimate a quadratic polynomial best-fit line:
+
+```{r}
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point() +
+  geom_smooth(
+    method = "lm",
+    formula = y ~ poly(x, 2)) # add quadratic best-fit line
+```
+
+Or, we could estimate a smooth best-fit (LOESS) line to allow for any form of nonlinearity:
+
+```{r}
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point() +
+  geom_smooth(method = "loess") # add smooth best-fit (LOESS) line
+```
+
+By default, the best-fit line is based on a generalized additive model, which allows for nonlinearity:
+
+```{r}
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point() +
+  geom_smooth() # add GAM best-fit line; same as specifying method = "gam"
+```
+
+Then, we can change the axes:
+
+```{r}
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point() +
+  geom_smooth() +
+  scale_x_continuous(
+    expand = c(0,0), # set origin of x-axis to 0
+    lim = c(20,40), # set limits of x-axis
+    breaks = seq(from = 20, to = 40, by = 5) # specify x-axis labels
+  ) +
+  scale_y_continuous(
+    expand = c(0,0), # set origin of y-axis to 0
+    lim = c(0,NA), # set limits of y-axis
+    breaks = seq(from = 0, to = 2500, by = 250) # specify y-axis labels
+  )
+```
+
+Then, we can add plot labels:
+
+```{r}
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point() +
+  geom_smooth() +
+  scale_x_continuous(
+    expand = c(0,0),
+    lim = c(20,40),
+    breaks = seq(from = 20, to = 40, by = 5)
+  ) +
+  scale_y_continuous(
+    expand = c(0,0),
+    lim = c(0,NA),
+    breaks = seq(from = 0, to = 2500, by = 250)
+  ) +
+  labs( # add plot labels
+    x = "Rushing Back Age (years)",
+    y = "Rushing Yards per season",
+    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
+    subtitle = "(minimum 50 rushing attempts)"
+  )
+```
+
+Then, we can use a theme such as the classic theme (`theme_classic()`) to make it more visually presentable:
+
+```{r}
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point() +
+  geom_smooth() +
+  scale_x_continuous(
+    expand = c(0,0),
+    lim = c(20,40),
+    breaks = seq(from = 20, to = 40, by = 5)
+  ) +
+  scale_y_continuous(
+    expand = c(0,0),
+    lim = c(0,NA),
+    breaks = seq(from = 0, to = 2500, by = 250)
+  ) +
+  labs(
+    x = "Rushing Back Age (years)",
+    y = "Rushing Yards per season",
+    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
+    subtitle = "(minimum 50 rushing attempts)"
+  ) +
+  theme_classic() # use the classic theme
+```
+
+Or, we could use a different theme, such as `theme_dark()`.
+For a list of themes available in `ggplot2`, see here: <https://ggplot2-book.org/themes#sec-themes>.
+
+```{r}
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point() +
+  geom_smooth() +
+  scale_x_continuous(
+    expand = c(0,0),
+    lim = c(20,40),
+    breaks = seq(from = 20, to = 40, by = 5)
+  ) +
+  scale_y_continuous(
+    expand = c(0,0),
+    lim = c(0,NA),
+    breaks = seq(from = 0, to = 2500, by = 250)
+  ) +
+  labs(
+    x = "Rushing Back Age (years)",
+    y = "Rushing Yards per season",
+    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
+    subtitle = "(minimum 50 rushing attempts)"
+  ) +
+  theme_dark() # use the dark theme
+```
+
+## Examples {#sec-scatterplot}
+
+### Players {#sec-scatterplotPlayers}
+
+#### Running Back Performance By Player Age {#sec-plotRBperformanceByAge}
+
+```{r}
+# Prepare Data
+rushing_attempts <- nfl_pbp %>% 
+  dplyr::filter(season_type == "REG") %>% 
+  dplyr::filter(
+    rush == 1,
+    rush_attempt == 1,
+    qb_scramble == 0,
+    qb_dropback == 0,
+    !is.na(rushing_yards))
+
+rb_yardsPerCarry <- rushing_attempts %>% 
+  dplyr::group_by(rusher_id, season) %>% 
+  dplyr::summarise(
+    ypc = mean(rushing_yards, na.rm = TRUE),
+    rush_attempts = n(),
+    .groups = "drop") %>% 
+  dplyr::ungroup() %>% 
+  dplyr::left_join(
+    player_stats_seasonal_offense,
+    by = c("rusher_id" = "player_id", "season")
+  ) %>% 
+  dplyr::filter(
+    position_group == "RB",
+    rush_attempts >= 50)
+```
+
+##### Rushing Yards Per Carry {#sec-plotYPCbyAge}
+
+```{r}
+#| label: fig-rushYPCperSeasonByAge
+#| fig-cap: "NFL Rushing Yards Per Carry per Season by Player Age"
+
+ggplot2::ggplot(
+  data = rb_yardsPerCarry,
+  aes(
+    x = age,
+    y = ypc)) +
+  geom_point() +
+  geom_smooth() +
+  labs(
+    x = "Rushing Back Age (years)",
+    y = "Rushing Yards per Carry/season",
+    title = "NFL Rushing Yards Per Carry per Season by Player Age",
+    subtitle = "(minimum 50 rushing attempts)"
+  ) +
+  theme_classic()
+```
+
+##### Rushing EPA Per Season {#sec-plotRushEPAbyAge}
+
+```{r}
+#| label: fig-rushEPAperSeasonByAge
+#| fig-cap: "NFL Rushing Expected Points Added (EPA) per Season by Player Age"
+
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_epa)) +
+  geom_point() +
+  geom_smooth() +
+  labs(
+    x = "Rushing Back Age (years)",
+    y = "Rushing EPA/season",
+    title = "NFL Rushing Expected Points Added (EPA) per Season by Player Age"
+  ) +
+  theme_classic()
+```
+
+### Teams {#sec-scatterplotTeams}
+
+#### Defensive and Offensive EPA per Play {#sec-plotDefOffEPA}
+
+Expected points added (EPA) per play by the team with possession.
+
+```{r}
+#| label: fig-plotDefOffEPA
+#| fig-cap: "2023 NFL Offensive and Defensive EPA per Play"
+
+pbp_regularSeason <- nfl_pbp %>% 
+  dplyr::filter(
+    season == 2023,
+    season_type == "REG") %>%
+  dplyr::filter(!is.na(posteam) & (rush == 1 | pass == 1))
+
+epa_offense <- pbp_regularSeason %>%
+  dplyr::group_by(team = posteam) %>%
+  dplyr::summarise(off_epa = mean(epa, na.rm = TRUE))
+
+epa_defense <- pbp_regularSeason %>%
+  dplyr::group_by(team = defteam) %>%
+  dplyr::summarise(def_epa = mean(epa, na.rm = TRUE))
+
+epa_combined <- epa_offense %>%
+  dplyr::inner_join(
+    epa_defense,
+    by = "team")
+
+ggplot2::ggplot(
+  data = epa_combined,
+  aes(
+    x = off_epa,
+    y = def_epa)) +
+  nflplotR::geom_mean_lines(
+    aes(
+      x0 = off_epa ,
+      y0 = def_epa)) +
+  nflplotR::geom_nfl_logos(
+    aes(
+      team_abbr = team),
+      width = 0.065,
+      alpha = 0.7) +
+  labs(
+    x = "Offense EPA/play",
+    y = "Defense EPA/play",
+    title = "2023 NFL Offensive and Defensive EPA per Play"
+  ) +
+  theme_classic() +
+  scale_y_reverse()
+```
+
+## Conclusion {#sec-dataVisualizationConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-dataVisualizationSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-21 getting-started.qmd
20240721 - data visualization

@@
 For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
+We save each data file after downloading it, so we can use the data in subsequent chapters.
 
@@
 #| eval: false
-#| include: false
 
@@
 ```{r}
-#| include: false
-
 save(
@@
 
-## Plotting {#sec-plotting}
-
-### Rushing Yards per Carry By Player Age {#sec-plotRushEPAbyAge}
-
-```{r}
-#| label: fig-rushYPCperSeasonByAge
-#| fig-cap: "2023 NFL Rushing Yards Per Carry per Season by Player Age"
-
-# Prepare Data
-rushing_attempts <- nfl_pbp %>% 
-  dplyr::filter(
-    season_type == "REG") %>% 
-    filter(
-      rush == 1,
-      rush_attempt == 1,
-      qb_scramble == 0,
-      qb_dropback == 0,
-      !is.na(rushing_yards))
-
-rb_yardsPerCarry <- rushing_attempts %>% 
-  group_by(rusher_id, season) %>% 
-  summarise(
-    ypc = mean(rushing_yards, na.rm = TRUE),
-    rush_attempts = n(),
-    .groups = "drop") %>% 
-  ungroup() %>% 
-  left_join(
-    nfl_players %>% select(-season),
-    by = c("rusher_id" = "gsis_id")
-  ) %>% 
-  filter(
-    position_group == "RB",
-    rush_attempts >= 50) %>% 
-  left_join(
-    season_startdate,
-    by = "season"
-  )
-
-rb_yardsPerCarry$age <- interval(
-  start = rb_yardsPerCarry$birth_date,
-  end = rb_yardsPerCarry$startdate
-) %>% 
-  time_length(unit = "years")
-
-# Create Plot
-ggplot2::ggplot(
-  data = rb_yardsPerCarry,
-  ggplot2::aes(
-    x = age,
-    y = ypc)) +
-  ggplot2::geom_point() +
-  ggplot2::geom_smooth() +
-  ggplot2::labs(
-    x = "Rushing Back Age (years)",
-    y = "Rushing Yards per Carry/season",
-    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
-    subtitle = "(minimum 50 rushing attempts)"
-  ) +
-  ggplot2::theme_classic()
-```
-
-```{r}
-#| label: fig-rushEPAperSeasonByAge
-#| fig-cap: "2023 NFL Rushing Expected Points Added (EPA) per Season by Player Age"
-
-# Subset Data
-rb_seasonal <- player_stats_seasonal_offense %>% 
-  filter(position_group == "RB")
-
-# Create Plot
-ggplot2::ggplot(
-  data = rb_seasonal,
-  ggplot2::aes(
-    x = age,
-    y = rushing_epa)) +
-  ggplot2::geom_point() +
-  ggplot2::geom_smooth() +
-  ggplot2::labs(
-    x = "Rushing Back Age (years)",
-    y = "Rushing EPA/season",
-    title = "2023 NFL Rushing Expected Points Added (EPA) per Season by Player Age"
-  ) +
-  ggplot2::theme_classic()
-```
-
-### Defensive and Offensive EPA per Play {#sec-plotDefOffEPA}
-
-Expected points added (EPA) per play by the team with possession.
-
-```{r}
-#| label: fig-plotDefOffEPA
-#| fig-cap: "2023 NFL Offensive and Defensive EPA per Play"
-
-pbp_regularSeason <- nfl_pbp %>% 
-  dplyr::filter(
-    season == 2023,
-    season_type == "REG") %>%
-  dplyr::filter(!is.na(posteam) & (rush == 1 | pass == 1))
-
-epa_offense <- pbp_regularSeason %>%
-  dplyr::group_by(team = posteam) %>%
-  dplyr::summarise(off_epa = mean(epa, na.rm = TRUE))
-
-epa_defense <- pbp_regularSeason %>%
-  dplyr::group_by(team = defteam) %>%
-  dplyr::summarise(def_epa = mean(epa, na.rm = TRUE))
-
-epa_combined <- epa_offense %>%
-  dplyr::inner_join(epa_defense, by = "team")
-
-ggplot2::ggplot(
-  data = epa_combined,
-  ggplot2::aes(
-    x = off_epa,
-    y = def_epa)) +
-  nflplotR::geom_mean_lines(
-    ggplot2::aes(
-      x0 = off_epa ,
-      y0 = def_epa)) +
-  nflplotR::geom_nfl_logos(
-    ggplot2::aes(
-      team_abbr = team),
-      width = 0.065,
-      alpha = 0.7) +
-  ggplot2::labs(
-    x = "Offense EPA/play",
-    y = "Defense EPA/play",
-    title = "2023 NFL Offensive and Defensive EPA per Play"
-  ) +
-  ggplot2::theme_classic() +
-  ggplot2::scale_y_reverse()
-```
-
 ::: {.content-visible when-format="html"}

### 2024-07-21 multiple-regression.qmd
20240721 - fix typo
20240721 - additions
20240721 - clean up

@@
   data = regression,
-  ggplot2::aes(
+  aes(
     x = predictor,
@@
 ) +
-  ggplot2::geom_point() +
-  ggplot2::geom_smooth(
+  geom_point() +
+  geom_smooth(
     method = "lm",
@@
     fullrange = TRUE) +
-  ggplot2::scale_x_continuous(
+  scale_x_continuous(
     lim = c(0,8),
@@
   ) +
-  ggplot2::scale_y_continuous(
+  scale_y_continuous(
     lim = c(0,8),
@@
   ) +
-  ggplot2::labs(
+  labs(
     x = "Predictor Variable",
@@
   ) +
-  ggplot2::theme_classic(
+  theme_classic(
     base_size = 16) +
-  ggplot2::theme(legend.title = element_blank())
+  theme(legend.title = element_blank())
 
@@
 
-::: {.content-visible when-format="html"}
-
 ## Impact of Oultiers {#sec-multipleRegressionOutliers}
@@
 [As with correlation](#sec-correlationOutliers), multiple regression can be strongly impacted by outliers.
+
+## Bayesian Multiple Regression {#sec-multipleRegressionBayesian}
 
@@
 
+::: {.content-visible when-format="html"}
+
 ## Session Info {#sec-multipleRegressionSessionInfo}

### 2024-07-21 mythbusters.qmd
20240721 - data visualization
20240721 - clean up

@@
 
+### Specify Package Options {#sec-mythbustersPackageOptions}
 ```{r}
-#| include: false
-
 emm_options(lmerTest.limit = 100000)
@@
 ```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_playerContracts.RData")
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+load(file = "./data/nfl_espnQBR.RData")
+```
+
+```{r}
 load(file = "./data/nfl_playerContracts.RData")
@@
 load(file = "./data/nfl_espnQBR.RData")
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
 ```

### 2024-07-21 data-visualization.qmd
20240721 - figure labels

@@
 
-Second, we create the base plot using the `ggplot()` function of the `ggplot2` package.
+Second, we create the base plot using the `ggplot()` function of the `ggplot2` package, as in @fig-scatterplot1.
 We specify the data object and the variables in the data object that are associated with the x- and y-axes:
@@
 ```{r}
+#| label: fig-scatterplot1
+#| fig-cap: "Base Plot."
+
 ggplot2::ggplot(
@@
 
-Third, we create a scatterplot using the `geom_point()` function from the `ggplot2` package:
+Third, we create a scatterplot using the `geom_point()` function from the `ggplot2` package, as in @fig-scatterplot2:
 
 ```{r}
+#| label: fig-scatterplot2
+#| fig-cap: "Scatterplot."
+
 ggplot2::ggplot(
@@
 
-Fourth, we add a linear best-fit line using the `geom_smooth()`:
+Fourth, we add a linear best-fit line using the `geom_smooth()`, as in @fig-scatterplot3:
 
 ```{r}
+#| label: fig-scatterplot3
+#| fig-cap: "Scatterplot with Linear Best-Fit Line."
+
 ggplot2::ggplot(
@@
 
-We could also estimate a quadratic polynomial best-fit line:
+We could also estimate a quadratic polynomial best-fit line, as in @fig-scatterplot4:
 
 ```{r}
+#| label: fig-scatterplot4
+#| fig-cap: "Scatterplot with Quadratic Best-Fit Line."
+
 ggplot2::ggplot(
@@
 
-Or, we could estimate a smooth best-fit (LOESS) line to allow for any form of nonlinearity:
+Or, we could estimate a smooth best-fit line using locally estimated scatterplot smoothing (LOESS) to allow for any form of nonlinearity, as in @fig-scatterplot5:
 
 ```{r}
+#| label: fig-scatterplot5
+#| fig-cap: "Scatterplot with Best-Fit Line Using Locally Estimated Scatterplot Smoothing (LOESS)."
+
 ggplot2::ggplot(
@@
 
-By default, the best-fit line is based on a generalized additive model, which allows for nonlinearity:
+By default, the best-fit line is based on a generalized additive model, which allows for nonlinearity, as in @fig-scatterplot6:
 
 ```{r}
+#| label: fig-scatterplot6
+#| fig-cap: "Scatterplot with Best-Fit Line from Generalized Additive Model."
+
 ggplot2::ggplot(
@@
 
-Then, we can change the axes:
+Then, we can change the axes, as in @fig-scatterplot7:
 
 ```{r}
+#| label: fig-scatterplot7
+#| fig-cap: "Scatterplot with Modified Axes."
+
 ggplot2::ggplot(
@@
 
-Then, we can add plot labels:
+Then, we can add plot labels, as in @fig-scatterplot8:
 
 ```{r}
+#| label: fig-scatterplot8
+#| fig-cap: "Scatterplot with Plot Labels."
+
 ggplot2::ggplot(
@@
 
-Then, we can use a theme such as the classic theme (`theme_classic()`) to make it more visually presentable:
+Then, we can use a theme such as the classic theme (`theme_classic()`) to make it more visually presentable, as in @fig-scatterplot9:
 
 ```{r}
+#| label: fig-scatterplot9
+#| fig-cap: "Scatterplot with Classic Theme."
+
 ggplot2::ggplot(
@@
 
-Or, we could use a different theme, such as `theme_dark()`.
+Or, we could use a different theme, such as the dark theme (`theme_dark()`) in @fig-scatterplot10.
 For a list of themes available in `ggplot2`, see here: <https://ggplot2-book.org/themes#sec-themes>.
@@
 ```{r}
+#| label: fig-scatterplot10
+#| fig-cap: "Scatterplot with Dark Theme."
+
 ggplot2::ggplot(

### 2024-07-21 data-visualization.qmd
20240721 - interactive plotly plot

@@
 library("nflplotR")
+library("plotly")
 ```
@@
 
+```{r}
+#| label: fig-scatterplot11
+#| fig-cap: "Interactive Scatterplot Using Plotly."
+
+plot_ypcByPlayerAge <- ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point(aes(label = player_display_name)) + # add player name for mouse over tooltip
+  geom_smooth() +
+  scale_x_continuous(
+    expand = c(0,0),
+    lim = c(20,40),
+    breaks = seq(from = 20, to = 40, by = 5)
+  ) +
+  scale_y_continuous(
+    expand = c(0,0),
+    lim = c(0,NA),
+    breaks = seq(from = 0, to = 2500, by = 250)
+  ) +
+  labs(
+    x = "Running Back Age (years)",
+    y = "Rushing Yards Per Season",
+    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
+    subtitle = "(minimum 50 rushing attempts)"
+  ) +
+  theme_classic()
+
+ggplotly(plot_ypcByPlayerAge)
+```
+
 ## Examples {#sec-scatterplot}

### 2024-07-21 data-visualization.qmd
20240721 - update interactive plot labels

--- 2024-07-21 data-visualization.qmd
@@
 
+```{r}
+#| label: fig-scatterplot11
+#| fig-cap: "Interactive Scatterplot Using Plotly."
+
+plot_ypcByPlayerAge <- ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point(aes(
+    text = player_display_name, # add player name for mouse over tooltip
+    label = season)) + # add season for mouse over tooltip
+  geom_smooth() +
+  scale_x_continuous(
+    expand = c(0,0),
+    lim = c(20,40),
+    breaks = seq(from = 20, to = 40, by = 5)
+  ) +
+  scale_y_continuous(
+    expand = c(0,0),
+    lim = c(0,NA),
+    breaks = seq(from = 0, to = 2500, by = 250)
+  ) +
+  labs(
+    x = "Running Back Age (years)",
+    y = "Rushing Yards Per Season",
+    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
+    subtitle = "(minimum 50 rushing attempts)"
+  ) +
+  theme_classic()
+
+ggplotly(plot_ypcByPlayerAge)
+```
+
 ## Examples {#sec-scatterplot}

### 2024-07-21 data-visualization.qmd
20240721 - conditional content
20240721 - attempt to fix error

--- 2024-07-21 data-visualization.qmd
@@
 
+::: {.content-visible when-format="html:js"}
+
+After creating our plot, we can make the plot interactive using the `ggplotly()` function from the plotly package, as in @fig-scatterplot11.
+
+```{r}
+#| label: fig-scatterplot11
+#| fig-cap: "Interactive Scatterplot Using Plotly."
+
+plot_ypcByPlayerAge <- ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season)) + # add season for mouse over tooltip
+  geom_smooth() +
+  scale_x_continuous(
+    expand = c(0,0),
+    lim = c(20,40),
+    breaks = seq(from = 20, to = 40, by = 5)
+  ) +
+  scale_y_continuous(
+    expand = c(0,0),
+    lim = c(0,NA),
+    breaks = seq(from = 0, to = 2500, by = 250)
+  ) +
+  labs(
+    x = "Running Back Age (years)",
+    y = "Rushing Yards Per Season",
+    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
+    subtitle = "(minimum 50 rushing attempts)"
+  ) +
+  theme_classic()
+
+ggplotly(plot_ypcByPlayerAge)
+```
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
+After creating our plot, we can make the plot interactive using the `ggplotly()` function from the plotly package.
+
+```{r}
+#| eval: false
+
+plot_ypcByPlayerAge <- ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season)) + # add season for mouse over tooltip
+  geom_smooth() +
+  scale_x_continuous(
+    expand = c(0,0),
+    lim = c(20,40),
+    breaks = seq(from = 20, to = 40, by = 5)
+  ) +
+  scale_y_continuous(
+    expand = c(0,0),
+    lim = c(0,NA),
+    breaks = seq(from = 0, to = 2500, by = 250)
+  ) +
+  labs(
+    x = "Running Back Age (years)",
+    y = "Rushing Yards Per Season",
+    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
+    subtitle = "(minimum 50 rushing attempts)"
+  ) +
+  theme_classic()
+
+ggplotly(plot_ypcByPlayerAge)
+```
+
+:::
+
 ## Examples {#sec-scatterplot}
@@
     y = ypc)) +
-  geom_point() +
+  geom_point(
+    aes(
+      text = player_display_name,
+      label = season)) +
   geom_smooth() +
@@
     y = rushing_epa)) +
-  geom_point() +
+  geom_point(
+    aes(
+      text = player_display_name,
+      label = season)) +
   geom_smooth() +

### 2024-07-21 data-visualization.qmd
20240721 - interactive plots

@@
   labs( # add plot labels
-    x = "Rushing Back Age (years)",
-    y = "Rushing Yards per season",
-    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
-    subtitle = "(minimum 50 rushing attempts)"
+    x = "Running Back's Age (years)",
+    y = "Rushing Yards Per Season",
+    title = "NFL Rushing Yards (Season) by Player Age",
+    subtitle = "(Among Running Backs)"
   )
@@
   labs(
-    x = "Rushing Back Age (years)",
-    y = "Rushing Yards per season",
-    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
-    subtitle = "(minimum 50 rushing attempts)"
+    x = "Running Back's Age (years)",
+    y = "Rushing Yards Per Season",
+    title = "NFL Rushing Yards (Season) by Player Age",
+    subtitle = "(Among Running Backs)"
   ) +
@@
   labs(
-    x = "Running Back Age (years)",
+    x = "Running Back's Age (years)",
     y = "Rushing Yards Per Season",
-    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
-    subtitle = "(minimum 50 rushing attempts)"
+    title = "NFL Rushing Yards (Season) by Player Age",
+    subtitle = "(Among Running Backs)"
   ) +
@@
 
+::: {.content-visible when-format="html:js"}
+
 ```{r}
 #| label: fig-rushYPCperSeasonByAge
-#| fig-cap: "NFL Rushing Yards Per Carry per Season by Player Age"
+#| fig-cap: "NFL Rushing Yards Per Carry Per Season by Player Age"
+
+plot_ypcByPlayerAge2 <- ggplot2::ggplot(
+  data = rb_yardsPerCarry,
+  aes(
+    x = age,
+    y = ypc)) +
+  geom_point(
+    aes(
+      text = player_display_name,
+      label = season)) +
+  geom_smooth() +
+  labs(
+    x = "Running Back's Age (years)",
+    y = "Rushing Yards Per Carry (Season)",
+    title = "NFL Rushing Yards Per Carry (Season) by Player Age",
+    subtitle = "(minimum 50 rushing attempts)"
+  ) +
+  theme_classic()
+
+ggplotly(plot_ypcByPlayerAge2)
+```
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
+```{r}
+#| label: fig-rushYPCperSeasonByAge
+#| fig-cap: "NFL Rushing Yards Per Carry Per Season by Player Age"
 
@@
   labs(
-    x = "Rushing Back Age (years)",
-    y = "Rushing Yards per Carry/season",
-    title = "NFL Rushing Yards Per Carry per Season by Player Age",
+    x = "Running Back's Age (years)",
+    y = "Rushing Yards Per Carry (Season)",
+    title = "NFL Rushing Yards Per Carry (Season) by Player Age",
     subtitle = "(minimum 50 rushing attempts)"
@@
 
+:::
+
 ##### Rushing EPA Per Season {#sec-plotRushEPAbyAge}
+
+::: {.content-visible when-format="html:js"}
+
+```{r}
+#| label: fig-rushEPAperSeasonByAge
+#| fig-cap: "NFL Rushing Expected Points Added (EPA) per Season by Player Age"
+
+plot_rushEPAbyPlayerAge <- ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_epa)) +
+  geom_point(
+    aes(
+      text = player_display_name,
+      label = season)) +
+  geom_smooth() +
+  labs(
+    x = "Running Back's Age (years)",
+    y = "Rushing EPA (Season)",
+    title = "NFL Rushing Expected Points Added (Season) by Player Age"
+  ) +
+  theme_classic()
+
+ggplotly(plot_rushEPAbyPlayerAge)
+```
+
+:::
+
+::: {.content-visible when-format="pdf"}
 
@@
   labs(
-    x = "Rushing Back Age (years)",
-    y = "Rushing EPA/season",
-    title = "NFL Rushing Expected Points Added (EPA) per Season by Player Age"
+    x = "Running Back's Age (years)",
+    y = "Rushing EPA (Season)",
+    title = "NFL Rushing Expected Points Added (Season) by Player Age"
   ) +
@@
 ```
+
+:::
 
@@
 ```{r}
-#| label: fig-plotDefOffEPA
-#| fig-cap: "2023 NFL Offensive and Defensive EPA per Play"
-
 pbp_regularSeason <- nfl_pbp %>% 
@@
     by = "team")
+```
+
+```{r}
+#| label: fig-plotDefOffEPA
+#| fig-cap: "2023 NFL Offensive and Defensive EPA per Play"
 

### 2024-07-21 data-visualization.qmd
20240721 - fix error

@@
 
+Rushing yards per carry over the course of the season is depicted as a function of the Running Back's age in @fig-rushYPCperSeasonByAge.
+
 ```{r}
 #| label: fig-rushYPCperSeasonByAge
-#| fig-cap: "NFL Rushing Yards Per Carry Per Season by Player Age"
+#| fig-cap: "Rushing Yards Per Carry (Season) by Player Age."
 
@@
 
+Rushing yards per carry over the course of the season is depicted as a function of the Running Back's age in @fig-rushYPCperSeasonByAgePDF.
+
 ```{r}
-#| label: fig-rushYPCperSeasonByAge
-#| fig-cap: "NFL Rushing Yards Per Carry Per Season by Player Age"
+#| label: fig-rushYPCperSeasonByAgePDF
+#| fig-cap: "NFL Rushing Yards Per Carry Per Season by Player Age."
 
@@
 
+Rushing expected points added (EPA) over the course of the season is depicted as a function of the Running Back's age in @fig-rushEPAperSeasonByAge.
+
 ```{r}
 #| label: fig-rushEPAperSeasonByAge
-#| fig-cap: "NFL Rushing Expected Points Added (EPA) per Season by Player Age"
+#| fig-cap: "Rushing Expected Points Added (Season) by Player Age."
 
@@
 
+Rushing expected points added (EPA) over the course of the season is depicted as a function of the Running Back's age in @fig-rushEPAperSeasonByAgePDF.
+
 ```{r}
-#| label: fig-rushEPAperSeasonByAge
-#| fig-cap: "NFL Rushing Expected Points Added (EPA) per Season by Player Age"
+#| label: fig-rushEPAperSeasonByAgePDF
+#| fig-cap: "NFL Rushing Expected Points Added (EPA) per Season by Player Age."
 
@@
 
+Defensive EPA per play during the 2023 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA.
+
 ```{r}
 #| label: fig-plotDefOffEPA
-#| fig-cap: "2023 NFL Offensive and Defensive EPA per Play"
+#| fig-cap: "2023 NFL Offensive and Defensive EPA Per Play."
 

### 2024-07-22 data-visualization.qmd
20240721 - additional figures

@@
 We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
+
+## Overview {#sec-dataVisualizationOverview}
+
+The R Graph Gallery provides examples of various types of plots: <https://r-graph-gallery.com>.
+
+## Univariate Distribution {#sec-univariateDistribution}
+
+### Histogram {#sec-histogram}
+
+A histogram of fantasy points is depicted in @fig-histogram.
+
+```{r}
+#| label: fig-histogram
+#| fig-cap: "Histogram of Fantasy Points."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasy_points)
+) +
+  geom_histogram(
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  theme_classic()
+```
+
+### Density Plot {#sec-densityPlot}
+
+A histogram of fantasy points is depicted in @fig-density.
+
+```{r}
+#| label: fig-density
+#| fig-cap: "Density Plot of Fantasy Points."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasy_points,
+    fill = position_group)
+) +
+  geom_density(alpha = 0.7) + # add transparency
+  theme_classic()
+```
+
+### Histogram with Overlaid Density Plot {#sec-histogramDensity}
+
+A histogram of fantasy points with an overlaid density plot is depicted in @fig-histogramDensity.
+
+```{r}
+#| label: fig-histogramDensity
+#| fig-cap: "Histogram with Overlaid Density Plot."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasy_points)
+) +
+  geom_histogram(
+    aes(y = after_stat(density)),
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  geom_density(
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  theme_classic()
+```
+
+### Box-and-Whisker Plot {#sec-boxAndWhiskerPlot}
+
+A box-and-whisker plot of fantasy points is depicted in @fig-boxAndWhisker.
+
+```{r}
+#| label: fig-boxAndWhisker
+#| fig-cap: "Box-and-Whisker Plot."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = position_group,
+    y = fantasy_points,
+    fill = position_group)
+) +
+  geom_boxplot() +
+  theme_classic() + 
+  theme(legend.position = "none")
+```
+
+### Violin Plot {#sec-violinPlot}
+
+A violin plot of fantasy points is depicted in @fig-boxAndWhisker.
+
+```{r}
+#| label: fig-violin
+#| fig-cap: "Violin Plot."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = position_group,
+    y = fantasy_points,
+    fill = position_group)
+) +
+  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
+  theme_classic() + 
+  theme(legend.position = "none")
+```
 
@@
 
+## Line Chart {#sec-lineChart}
+
+## Bar Plot {#sec-barPlot}
+
+A bar plot of fantasy points by position is depicted in @fig-barPlot.
+
+```{r}
+#| label: fig-barPlot
+#| fig-cap: "Bar Plot."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = position_group,
+    y = fantasy_points,
+    fill = position_group
+    )
+) +
+  geom_bar(
+    stat = "identity",
+    position = position_dodge(width = 0)) +
+  theme_classic()
+```
+
+To display bars corresponding to a 95% confidence interval, we first compute summary statistics:
+
+```{r}
+player_stats_seasonal_offense_summary <- player_stats_seasonal_offense %>%
+    filter(position_group %in% c("QB","RB","WR","TE")) %>% 
+  group_by(position_group) %>%
+  summarise( 
+    n = sum(!is.na(fantasy_points)),
+    mean = mean(fantasy_points, na.rm = TRUE),
+    sd = sd(fantasy_points, na.rm = TRUE)
+  ) %>%
+  mutate(se = sd/sqrt(n)) %>%
+  mutate(
+    ci_lower = mean - qt(0.975, df = n - 1) * se,
+    ci_upper = mean + qt(0.975, df = n - 1) * se
+  )
+```
+
+The summary statistics are in @tbl-summaryStats.
+
+```{r}
+#| label: tbl-summaryStats
+#| tbl-cap: "Table of Summary Statistics."
+
+player_stats_seasonal_offense_summary
+```
+
+Based on the summary statistics in @tbl-summaryStats, we create a bar plot with bars representing the 95% confidence interval in @fig-barPlotConfidenceInterval.
+
+```{r}
+#| label: fig-barPlotConfidenceInterval
+#| fig-cap: "Bar Plot with Bars Representing the 95% Confidence Interval."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense_summary %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = position_group,
+    y = mean,
+    fill = position_group
+    )
+) +
+  geom_bar(
+    stat = "identity",
+    position = position_dodge(width = 0)) +
+  geom_errorbar(
+    aes(ymin = ci_lower, ymax = ci_upper),
+    width = 0.2,
+    position = position_dodge(width = 0)
+  ) +
+  theme_classic()
+```
+
 ## Examples {#sec-scatterplot}

### 2024-07-22 data-visualization.qmd
20240721 - line chart

@@
 
+A bar plot of Tom Brady's fantasy points by season is depicted in @fig-lineChart.
+
+```{r}
+#| label: fig-lineChart
+#| fig-cap: "Line Chart."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense %>%
+    filter(player_display_name == "Tom Brady"),
+  mapping = aes(
+    x = season,
+    y = fantasy_points,
+    fill = position_group
+    )
+) +
+  geom_line(
+    linewidth = 1.5,
+    color = "blue"
+  ) +
+  theme_classic()
+```
+
 ## Bar Plot {#sec-barPlot}

### 2024-07-22 data-visualization.qmd
20240721 - fix bar plot

@@
 
-A bar plot of fantasy points by position is depicted in @fig-barPlot.
-
-```{r}
-#| label: fig-barPlot
-#| fig-cap: "Bar Plot."
-
-ggplot2::ggplot(
-  data = player_stats_seasonal_offense %>%
-    filter(position_group %in% c("QB","RB","WR","TE")),
-  mapping = aes(
-    x = position_group,
-    y = fantasy_points,
-    fill = position_group
-    )
-) +
-  geom_bar(
-    stat = "identity",
-    position = position_dodge(width = 0)) +
-  theme_classic()
-```
-
-To display bars corresponding to a 95% confidence interval, we first compute summary statistics:
+To create a bar plot, we first compute summary statistics:
 
@@
 
+A bar plot of fantasy points by position is depicted in @fig-barPlot.
+
+```{r}
+#| label: fig-barPlot
+#| fig-cap: "Bar Plot."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense_summary,
+  mapping = aes(
+    x = position_group,
+    y = mean,
+    fill = position_group,
+    color = position_group
+    )
+) +
+  geom_bar(
+    stat = "identity") +
+  theme_classic()
+```
+
 Based on the summary statistics in @tbl-summaryStats, we create a bar plot with bars representing the 95% confidence interval in @fig-barPlotConfidenceInterval.
@@
     y = mean,
-    fill = position_group
+    fill = position_group,
+    color = position_group
     )
@@
   geom_bar(
-    stat = "identity",
-    position = position_dodge(width = 0)) +
+    stat = "identity") +
   geom_errorbar(
-    aes(ymin = ci_lower, ymax = ci_upper),
+    aes(
+      ymin = ci_lower,
+      ymax = ci_upper),
     width = 0.2,
-    position = position_dodge(width = 0)
+    color = "black"
   ) +

### 2024-07-22 causal-inference.qmd
20240721 - ancestors and descendants

@@
 A common form of causal diagrams is the directed acyclic graph (DAG).
-DAGs provide a helpful tool to communicate about causal questions and help identify how to avoid bias (i.e., over-estimation) in associations between variables due to [confounding](#sec-causalDiagramConfounding) (i.e., common causes) [@Digitale2022].
+DAGs provide a helpful tool to communicate about causal questions and help identify how to avoid bias (i.e., overestimation) in associations between variables due to [confounding](#sec-causalDiagramConfounding) (i.e., common causes) [@Digitale2022].
 For instance, from a DAG, it is possible to determine what variables it is important to control for in order to get unbiased estimates of the association between two variables of interest.
@@
 
+### Ancestors and Descendants {#sec-ancestorsDescendants}
+
+In a causal model, an *ancestor* is a variable that influences another variable, either directly or indirectly via another variable [@Rohrer2018].
+A *descendant* is a variable that is influenced by another variable [@Rohrer2018].
+In general, one should not control for descendants of the outcome variable, because doing so could eliminate the apparent effect of a legitimate cause on the outcome variable [@Digitale2022].
+Moreover, as described above, when trying to understand the total causal effect of a predictor variable on an outcome variable, one should not control for descendants of the predictor variable that are also antecedents of the outcome variable (i.e., mediators of the effect of the predictor variable on the outcome variable) [@Digitale2022].
+
+Consider the example in @fig-dagDescendant:
+
+```{r}
+#| label: fig-dagDescendant
+#| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Ancestor (Mediation) and Descendant of Y."
+
+descendentDag <- dagitty::dagitty("dag{
+  X -> M
+  M -> Y
+  X -> Y
+  Y -> Z
+}")
+
+#plot(dagitty::graphLayout(descendentDag))
+ggdag::ggdag(descendentDag) + theme_dag_blank()
+```
+
+```{r}
+dagitty::impliedConditionalIndependencies(descendentDag)
+```
+
+In this example, `X` and `M` are conditionally independent with `Z` when accounting for the mediator (`Y`).
+
+```{r}
+dagitty::adjustmentSets(
+  descendentDag, 
+  exposure = "X",
+  outcome = "Y",
+  effect = "direct")
+
+dagitty::adjustmentSets(
+  descendentDag, 
+  exposure = "X",
+  outcome = "Y",
+  effect = "total")
+```
+
+As indicated above, one should not control for the descendant of the outcome variable; thus, one should not control for `Z` when examining the association between `X` or `M` and `Y`.
+
 ### Collider Bias {#sec-colliderBias}
@@
 Various experimental and quasi-experimental designs and approaches can be leveraged to more closely approximate causal inferences.
-Longitudinal designs, within-subject analyses, inclusion of control variables, and genetically informed designs are all quasi-experimental designs that afford the researcher greater control over some possible third variable [confounds](#sec-causalDiagramConfounding).
+[Longitudinal designs](#sec-causalInferenceLongitudinal), [within-subject analyses](#sec-causalInferenceWithinSubject), [inclusion of control variables](#sec-causalInferenceControlVariables), and [genetically informed designs](#sec-causalInferenceGeneticallyInformed) are all quasi-experimental designs that afford the researcher greater control over some possible third variable [confounds](#sec-causalDiagramConfounding).
 Causal diagrams can be a useful tool for identifying the proper variables to control for (and those not to control for).
@@
 It is important not to control for mediators when interested in the total effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy).
-When there is a [collision](#sec-colliderBias), it is important not to control for the [collider](#sec-colliderBias) unless the [collider](#sec-colliderBias) is also a [confound](#sec-causalDiagramConfounding).
+In addition, it is important not to control for [descendants](#sec-ancestorsDescendants) of the [outcome variable](#sec-correlationalStudy).
+When there is a [collision](#sec-colliderBias), it is important not to control for the [collider](#sec-colliderBias) (unless the [collider](#sec-colliderBias) is also a [confound](#sec-causalDiagramConfounding)).
 

### 2024-07-22 data-visualization.qmd
20240722 - add rug plot

@@
 
-### Histogram with Overlaid Density Plot {#sec-histogramDensity}
+### Histogram with Overlaid Density and Rug Plot {#sec-histogramDensity}
 
@@
   ) +
+  geom_rug() +
   theme_classic()

### 2024-07-22 mythbusters.qmd
20240722 - fix citation

@@
 To our knowledge, no peer-reviewed studies have examined this question for football players.
-One study found that National Basketball Association (NBA) players improved in field goal percentage, points, and player efficiency rating (but not other statistics: rebounds, assists, steals, or blocks) from their pre-contract year to their contract year, and that Major League Baseball (MLB) players improved in runs batted in (RBIs; but not other statistics: batting average, slugging percentage, on base percentage, home runs, fielding percentage) from their pre-contract year to their contract year [White2014a].
+One study found that National Basketball Association (NBA) players improved in field goal percentage, points, and player efficiency rating (but not other statistics: rebounds, assists, steals, or blocks) from their pre-contract year to their contract year, and that Major League Baseball (MLB) players improved in runs batted in (RBIs; but not other statistics: batting average, slugging percentage, on base percentage, home runs, fielding percentage) from their pre-contract year to their contract year [@White2014a].
 Other casual analyses have been examined contract-year performance of National Football League (NFL) players, including articles in [2012](https://www.4for4.com/2012/preseason/2012-contract-year-players-and-myth-increased-production) (archived [here](https://perma.cc/CT3F-QN5E)) and [2022](https://www.4for4.com/2022/preseason/do-players-perform-better-fantasy-football-contract-year) (archived [here](https://perma.cc/F4F5-7RQZ)).

### 2024-07-23 data-visualization.qmd
20240722 - confidence interval

@@
 
-A histogram of fantasy points with an overlaid density plot is depicted in @fig-histogramDensity.
+A histogram of fantasy points with an overlaid density and rug plot is depicted in @fig-histogramDensity.
 
@@
 #| label: fig-histogramDensity
-#| fig-cap: "Histogram with Overlaid Density Plot."
+#| fig-cap: "Histogram with Overlaid Density and Rug Plot."
 
@@
 ```{r}
+confidenceLevel <- .95 # for 95% confidence interval
+
 player_stats_seasonal_offense_summary <- player_stats_seasonal_offense %>%
@@
   mutate(
-    ci_lower = mean - qt(0.975, df = n - 1) * se,
-    ci_upper = mean + qt(0.975, df = n - 1) * se
+    ci_lower = mean - qt(p = 1 - (1 - confidenceLevel) / 2, df = n - 1) * se,
+    ci_upper = mean + qt(p = 1 - (1 - confidenceLevel) / 2, df = n - 1) * se
   )

### 2024-07-23 modern-portfolio-theory.qmd
Built site for gh-pages
20240608 - getting started
20240627 - sessionInfo()

@@
 # Modern Portfolio Theory {#sec-modernPortfolioTheory}
+
+## Getting Started {#sec-modernPortfolioTheoryGettingStarted}
+
+### Load Packages {#sec-modernPortfolioTheoryLoadPackages}
+
+```{r}
+library("quantmod")
+library("fPortfolio")
+library("tidyverse")
+```
+
+## Overview {#sec-modernPortfolioTheoryOverview}
+
+<https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league/> (archived at <https://perma.cc/NE35-G6LR>)
+
+## Download Historical Stock Prices {#sec-downloadStockPrices}
+
+```{r}
+symbols <- c(
+  "AAPL",  # Apple
+  "MSFT",  # Microsoft
+  "GOOGL", # Google
+  "AMZN",  # Amazon
+  "META",  # Meta/Facebook
+  "V",     # Visa
+  "DIS",   # Disney
+  "NKE",   # Nike
+  "TSLA")  # Tesla
+
+quantmod::getSymbols(symbols)
+```
+
+## Calculate Stock Returns {#sec-calculateStockReturns}
+
+```{r}
+prices <- do.call(
+  merge,
+  lapply(
+    symbols,
+    function(sym) quantmod::Cl(get(sym))))
+
+returns <- na.omit(
+  TTR::ROC(
+    prices,
+    type = "discrete"))
+
+returns_ts <- timeSeries::as.timeSeries(returns)
+```
+
+## Create Portfolio {#sec-createPortfolio}
+
+```{r}
+portfolioSpec <- fPortfolio::portfolioSpec()
+
+fPortfolio::setNFrontierPoints(portfolioSpec) <- 1000
+```
+
+## Determine the Efficient Frontier {#sec-efficientFrontier}
+
+```{r}
+efficientFrontier <- portfolioFrontier(
+  returns_ts,
+  spec = portfolioSpec)
+
+efficientFrontier
+```
+
+```{r}
+# Extract the coordinates of individual assets
+asset_means <- colMeans(returns)
+asset_sd <- apply(returns, 2, sd)
+
+# Add some padding to plot limits (so ticker symbols don't get cut off)
+xlim <- range(asset_sd) * c(0.9, 1.1)
+ylim <- range(asset_means) * c(0.9, 1.1)
+
+xlim[1] <- 0
+ylim[1] <- 0
+
+# Set scientific notation penalty
+options(scipen = 999)
+
+plot(
+  efficientFrontier,
+  which = c(
+    1,  # efficient frontier
+    3,  # tangency portfolio
+    4), # risk/return of individual assets
+  control = list(
+    xlim = xlim,
+    ylim = ylim
+  ))
+
+# Add text labels for individual assets
+points(
+  asset_sd,
+  asset_means,
+  col = "red",
+  pch = 19)
+
+text(
+  asset_sd,
+  asset_means,
+  labels = symbols,
+  pos = 4,
+  cex = 0.8,
+  col = "black")
+```
+
+## Identify the Optimal Weights {#sec-portfolioOptimalWeights}
+
+### Tangency Portfolio {#sec-tangencyPortfolio}
+
+The tangency portfolio is the portfolio with the highest Sharpe ratio (i.e., the highest ratio of return to risk).
+In other words, it is the portfolio with the greatest risk-adjusted returns.
+
+```{r}
+# Find the tangency portfolio (portfolio with the highest Sharpe ratio)
+tangencyPortfolio <- fPortfolio::tangencyPortfolio(
+  data = returns_ts,
+  spec = portfolioSpec)
+
+# Extract optimal weights
+tangencyPortfolio_optimalWeights <- fPortfolio::getWeights(tangencyPortfolio)
+tangencyPortfolio_optimalWeights
+
+# Output the results
+summary(tangencyPortfolio)
+```
+
+### Portfolio with Max Return at a Given Risk-Level {#sec-tangencyPortfolio}
+
+```{r}
+#| eval: false
+
+# Define target risk levels
+targetRisks <- seq(0, 0.3, by = 0.01)
+
+# Initialize storage for optimal portfolios
+optimalPortfolios <- list()
+optimalWeights_list <- list()
+
+# Find optimal weightings for each target risk level
+for (risk in targetRisks) {
+  # Create a portfolio optimization specification with the target risk
+  portfolioSpec <- fPortfolio::portfolioSpec()
+  fPortfolio::setTargetRisk(portfolioSpec) <- risk
+
+  # Solve for the maximum return at this target risk
+  optimal_portfolio <- fPortfolio::maxreturnPortfolio(
+    returns_ts,
+    spec = portfolioSpec)
+
+  # Store the optimal portfolio
+  optimalPortfolios[[as.character(risk)]] <- optimal_portfolio
+
+  # Store the optimal portfolio weights with risk level
+  optimal_weights <- fPortfolio::getWeights(optimal_portfolio)
+  optimalWeights_list[[as.character(risk)]] <- c(RiskLevel = risk, optimal_weights)
+}
+
+optimalWeightsByRisk <- dplyr::bind_rows(optimalWeights_list)
+optimalWeightsByRisk
+```
+
+## Conclusion {#sec-modernPortfolioTheoryConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-modernPortfolioTheorySessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-23 cognitive-bias.qmd
20240723 - draft and in-group bias

@@
 - confirmation bias
+- in-group bias
 - recency bias
@@
 Do your best to go into observation, data analysis, and data interpretation with an open mind.
+
+### In-Group Bias {#sec-cognitiveBiasesInGroup}
+
+*In-group bias* is the tendency for people to prefer others who are members of their same (in-)group.
+For instance, a manager whose favorite team is the Dallas Cowboys might try to draft mostly Cowboys players, regardless of their historical performance.
 

### 2024-07-23 draft.qmd
20240723 - draft and in-group bias

@@
 
+One factor that is *not* included above is whether a player is on your favorite team.
+Managers commonly like to draft players on their favorite teams (e.g., Cowboys, Eagles).
+That is fine—fantasy football is a game.
+Do what is fun for you.
+However, if your goal is to select the best players, leave your allegiances at the door.
+Selecting players based on their playing for your favorite team—rather than based on performance—is a form of [cognitive bias](#sec-cognitiveBias).
+
 ### Snake Draft {#sec-draftStrategySnake}
@@
 An alternative to the "stars and scrubs" approach is to wait to draft more "high-value" players after other managers have over-paid for players.
+In any case, having some small amount of cap left over toward the end of the draft can help you draft good value players to fill out your bench spots for cheap (e.g., $2).
 

### 2024-07-23 modern-portfolio-theory.qmd
20240723 - diversification

@@
 ## Overview {#sec-modernPortfolioTheoryOverview}
+
+- diversify; do not put all your eggs in one basket
+
+In terms of diversification, it can be helpful to diversify in multiple ways.
+First, it can be helpful not to rely on just one or two "stud" players.
+If they are on bye or have a down week, your team is more likely to suffer.
+Also, there are risks in picking multiple offensive players from the same team.
+For instance, if you have the Quarterback and Wide Receiver from the same team, and the team has a poor offensive outing, that will have a greater impact.
+Having multiple players on a juggernaut offense can be a boon, but it can be challenging to predict which offense will lead the league.
 

### 2024-07-24 modern-portfolio-theory.qmd
Built site for gh-pages

@@
 
+Selecting players for your fantasy team is like picking stocks.
+Your goal is to pick assets (i.e., players/stocks) that will perform best and that others undervalue.
+But what is the best way to do that?
+You could do lots of research to pick players/stocks with strong fundamentals that you think will do particularly well next year.
+By picking these players/stocks, you are predicting that they will *outperform* their expectations.
+However, all of your information is likely already reflected in the current valuation of the player/stock, so your prediction is basically a gamble.
+This is evidenced by the fact that people do not reliably beat the crowd/market.
+
+Even so-called experts do not beat the market reliably.
+There is little consistency in the performance of mutual fund managers over time.
+The following charts are from Leonard Mlodinow's book, "The Drunkard's Walk: How Randomness Rules Our Lives".
+The chart on the left depicts the performance of the top mutual funds from 1991 to 1995.
+The chart on the right depicts the performance of the same funds in the same order over the subsequent 5 years (1996 to 2000):
+
+The best funds from 1991–1995 weren not necessarily the best funds from 1996–2000.
+This suggests that mutual fund managers differ in great part because of luck or chance rather than reliable skill.
+That is likely why [a cat beat professional investors in a stock market challenge](http://www.npr.org/blogs/money/2013/01/14/169326326/housecat-beats-investors-in-stock-market-challenge) (archived at <https://perma.cc/R3XU-K6J8>).
+Although our sample size is much smaller with fantasy football projections, there also appears to be little consistency in fantasy football sites' rank in accuracy over time [INSERT], suggesting that the projection sources are not reliably better than each other (or the crowd) over time.
+
+The market reflects all of the knowledge of the crowd.
+One common misconception is that if you go with the market, you will receive "average" returns (by "average", I mean that you will be in the 50th percentile among investors).
+This is not true—it has been shown that most mutual funds (about 80%) underperform the average returns of the stock market.
+So, by going with the market average, you will likely perform better than the "average" fund/investor.
+Consistent with this, crowd-averaged fantasy football projections tend to be more accurate than any individual's projection: INSERT
+This evidence is consistent with the notion of the wisdom of the crowd.
+
 - diversify; do not put all your eggs in one basket
 
+Another important lesson from investing is diversification.
+If you have too much money in one asset and that asset tanks, you will lose your money.
+In other words, you do not want to put all of your eggs in one basket.
+By owning different asset classes (e.g., domestic and international stocks and bonds), you can limit your downside risk without sacrificing much in terms of expected return.
+
+This lesson can also apply to fantasy football.
 In terms of diversification, it can be helpful to diversify in multiple ways.
@@
 Also, there are risks in picking multiple offensive players from the same team.
+If you draft your starting Quarterback and Wide Receiver from the same team (e.g., the Cowboys), you are exposing your fantasy team to considerable risk.
 For instance, if you have the Quarterback and Wide Receiver from the same team, and the team has a poor offensive outing, that will have a greater impact.
+You can limit your downside risk by diversifying—drafting players from different teams.
+That way if the Cowboys' offense does poorly in a given week, your fantasy team will not be as affected.
 Having multiple players on a juggernaut offense can be a boon, but it can be challenging to predict which offense will lead the league.
+
+However, sometimes having two players on the same team might be beneficial because some positions may be uncorrelated or even negatively correlated, which can also reduce risk.
+For instance, the performance of the Tight End and Running Back on the same team tends to be slightly negatively correlated, so it might not be a bad idea to start the Tight End and Running Back from the same team.
+For a correlation matrix of all positions on the team, see INSERT
 

### 2024-07-24 modern-portfolio-theory.qmd
Built site for gh-pages

@@
 ## Overview {#sec-modernPortfolioTheoryOverview}
+
+## Fantasy Football is Like Stock Picking
 
@@
 
+Okay, fantasy football might be similar to stock picking, so what?
+You are most likely to pick the best players if you go with the wisdom of the crowd (e.g., average projections) and diversify.
+Most projections are public information, so you might wonder whether using crowd projections gains you anything because everybody else has access to public information.
+However, this is also the case with stocks, and people still consistently perform best over time when they go with the market.
+
 <https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league/> (archived at <https://perma.cc/NE35-G6LR>)
+
+## The Efficient Frontier
+
+The ultimate goal is to draft players for your starting lineup that provide the most projected points and the smallest downside risk.
+This is similar to the notion in investing of the efficient frontier, where your goal is to pick funds that have the greatest expected returns for the least risk (where risk is the variability in returns over time).
 

### 2024-07-24 modern-portfolio-theory.qmd
Built site for gh-pages
20240724 - modern portfolio theory

--- 2024-07-24 modern-portfolio-theory.qmd
@@
 
+## Fantasy Football is Like Stock Picking {#sec-likeStockPicking}
+
 Selecting players for your fantasy team is like picking stocks.
-Your goal is to pick assets (i.e., players/stocks) that will perform best and that others undervalue.
+In both fantasy football and the stock market, your goal is to pick assets (i.e., players/stocks) that will perform best and that others undervalue.
 But what is the best way to do that?
+Below, we discuss approaches to picking players/stocks.
+
+### The Wisdom of the Crowd (or Market) {#sec-wisdomOfTheCrowd}
+
+In picking players, there are various approaches one could take.
 You could do lots of research to pick players/stocks with strong fundamentals that you think will do particularly well next year.
@@
 
-The best funds from 1991–1995 weren not necessarily the best funds from 1996–2000.
+The best funds from 1991–1995 were not necessarily the best funds from 1996–2000.
 This suggests that mutual fund managers differ in great part because of luck or chance rather than reliable skill.
@@
 This evidence is consistent with the notion of the wisdom of the crowd.
+Moreover, even if the stock market is relatively efficient in terms of valuing stocks based on the available information, your fantasy football league is likely not.
+Thus, it may be effective to use crowd-based projections to identify players who are undervalued by your league.
 
-- diversify; do not put all your eggs in one basket
+### Diversification {#sec-diversification}
 
-Another important lesson from investing is diversification.
+Modern portfolio theory is a framework for determining the optimal composition of an investment portfolio to maximize expected returns for a given level of risk.
+Here, *risk* refers to the [variability](#sec-descriptiveStatisticsSpread) (e.g., standard deviation or variance) of returns across time.
+Given two portfolios with the same expected returns over time, people will prefer the "safer" portfolio—that is, the portfolio with less [variability](#sec-descriptiveStatisticsSpread)/volatility across time.
+One of the powerful notions of modern portfolio theory is that, through diversification, one can achieve lower risk with the same expected returns.
+In investing, *diversification* involves owning multiple asset classes (e.g., domestic and international stocks and bonds), with the goal of having asset classes that are either uncorrelated or negatively correlated.
+That is, owning different types of assets is safer than owning only one type.
 If you have too much money in one asset and that asset tanks, you will lose your money.
 In other words, you do not want to put all of your eggs in one basket.
-By owning different asset classes (e.g., domestic and international stocks and bonds), you can limit your downside risk without sacrificing much in terms of expected return.
+By owning different asset classes, you can limit your downside risk without sacrificing much in terms of expected return.
 
 This lesson can also apply to fantasy football.
+When assembling a team, you are essentially putting together a portfolio of assets (i.e., team of players).
+As with stocks, each player has an expected return (i.e., projection) and a degree of risk.
+In fantasy football, a player's risk might be quantified in terms of the [variability](#sec-descriptiveStatisticsSpread) of projected scores for a player across projection sources (e.g., Projection Source A, Source B, Souce C, etc.), or as historical game-to-game [variability](#sec-descriptiveStatisticsSpread).
+All things equal, we want to minimize our risk for a given level of expected returns.
+That way, we have the best chance of winning any given week.
+For the same level of expected returns, higher risk teams might have a few amazing games, but their teams might fall flat in other weeks.
+That is, for a given (high) rate of return, [you are best off in the long run (i.e., over the course of a season) with a lower risk team](https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league/) compared to a higher risk team (archived at <https://perma.cc/NE35-G6LR>).
+
 In terms of diversification, it can be helpful to diversify in multiple ways.
@@
 For instance, the performance of the Tight End and Running Back on the same team tends to be slightly negatively correlated, so it might not be a bad idea to start the Tight End and Running Back from the same team.
-For a correlation matrix of all positions on the team, see INSERT
+For a correlation matrix of all positions on the team, see: <https://assets-global.website-files.com/5f1af76ed86d6771ad48324b/607a4434a565aa7763bd1312_AndyAsh-Sharpstack-RPpaper.pdf> (archived at <https://perma.cc/JQ6G-KSRT>).
 
-<https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league/> (archived at <https://perma.cc/NE35-G6LR>)
+Another important idea from modern portfolio theory is that, if you want to achieve higher returns, you may be able to by accepting additional—and the right combination of—risk.
+That is, higher returns generally requires taking on additional risk—at least as long as we stay along the [efficient frontier](#sec-efficientFrontier), described next.
 
-## Download Historical Stock Prices {#sec-downloadStockPrices}
+## The Efficient Frontier of a Stock Portfolio {#sec-efficientFrontier}
+
+The ultimate goal in fantasy football is to draft players for your starting lineup that provide the most projected points and the smallest downside risk.
+That is, your goal is to achieve the optimal portfolio at a given level of risk, depending on how much risk you are willing to tolerate.
+One of the key tools in modern portfolio theory for identifying the optimal portfolio (for a given risk level) is the efficient frontier.
+The efficient frontier is a visual depiction of the maximum expected returns for a given level of risk (where risk is the [variability](#sec-descriptiveStatisticsSpread) in returns over time).
+The efficient frontier is helpful for identifying the optimal portfolio—the optimal combination and weighting of assets—for a given risk level.
+
+In the example below, we use historical returns (since YEAR) as the expected future returns.
+However, using historical returns as the expected future returns is risky because, as described in the common disclaimer, "Past performance does not guarantee future results."
+If you select a relatively short period of historical returns, you may be selecting a period when the stock performed particularly well.
+When evaluating historical returns it is preferable to evaluate long time horizons and to evaluate how the stock performed during period of both boom (i.e., "bull markets") and bust (i.e., "bear markets", such as in a recession).
+
+### Download Historical Stock Prices {#sec-downloadStockPrices}
 
@@
 
-## Calculate Stock Returns {#sec-calculateStockReturns}
+### Calculate Stock Returns {#sec-calculateStockReturns}
 
@@
 
-## Create Portfolio {#sec-createPortfolio}
+### Create Portfolio {#sec-createPortfolio}
 
@@
 
-## Determine the Efficient Frontier {#sec-efficientFrontier}
+### Determine the Efficient Frontier {#sec-efficientFrontier}
 
@@
 
-## Identify the Optimal Weights {#sec-portfolioOptimalWeights}
+### Identify the Optimal Weights {#sec-portfolioOptimalWeights}
 
-### Tangency Portfolio {#sec-tangencyPortfolio}
+#### Tangency Portfolio {#sec-tangencyPortfolio}
 
@@
 
-### Portfolio with Max Return at a Given Risk-Level {#sec-tangencyPortfolio}
+#### Portfolio with Max Return at a Given Risk Level {#sec-tangencyPortfolio}
 
@@
 
+## The Efficient Frontier of a Fantasy Team {#sec-efficientFrontierFantasy}
+
+<https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league/> (archived at <https://perma.cc/JQ6G-KSRT>)
+
 ## Conclusion {#sec-modernPortfolioTheoryConclusion}
+
+In summary, fantasy football is similar to stock picking.
+You are most likely to pick the best players if you go with the wisdom of the crowd (e.g., average projections across projection sources) and diversify.
+Most projections are public information, so you might wonder whether using crowd projections gains you anything because everybody else has access to public information.
+However, this is also the case with stocks, and people still consistently perform best over time when they go with the market.
 

### 2024-07-24 modern-portfolio-theory.qmd
20240724 - add modern portfolio theory

@@
 ```{r}
+#| label: fig-efficientFrontier
+#| fig-cap: "Efficient Frontier for a Stock Portfolio."
+
 # Extract the coordinates of individual assets

### 2024-07-25 modern-portfolio-theory.qmd
20240724 - duplicate section labels
20240724 - add package to function call

@@
 
-### Determine the Efficient Frontier {#sec-efficientFrontier}
+### Determine the Efficient Frontier {#sec-determineEfficientFrontier}
 
 ```{r}
-efficientFrontier <- portfolioFrontier(
+efficientFrontier <- fPortfolio::portfolioFrontier(
   returns_ts,
@@
 
-#### Portfolio with Max Return at a Given Risk Level {#sec-tangencyPortfolio}
+#### Portfolio with Max Return at a Given Risk Level {#sec-maxReturnPortfolio}
 

### 2024-07-25 cluster-analysis.qmd
20240605 - getting started/load packages
20240608 - getting started
Built site for gh-pages
20240703 - conclusion section
20240725 - cluster analysis

--- 2024-06-05 base-rates.qmd
@@
-# Base Rates {#sec-baseRates}
+# Cluster Analysis {#sec-clusterAnalysis}
 
-## Getting Started
+## Getting Started {#sec-clusterAnalysisGettingStarted}
 
-### Load Packages
+### Load Packages {#sec-clusterAnalysisLoadPackages}
+
+```{r}
+library("mclust")
+library("tidyverse")
+```
+
+### Load Data {#sec-clusterAnalysisLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+```
+
+### Overview {#sec-clusterAnalysisOverview}
 
@@
 ```
+
+### Example {#sec-clusterAnalysisExample}
+
+#### Prepare Data {#sec-clusterAnalysisExamplePrepareData}
+
+```{r}
+recentSeason <- max(player_stats_seasonal_offense$season, na.rm = TRUE)
+recentSeason
+
+player_stats_seasonal_offense_recent <- player_stats_seasonal_offense %>% 
+  filter(season == recentSeason)
+
+player_stats_seasonal_offense_recentQB <- player_stats_seasonal_offense_recent %>% 
+  filter(position_group == "QB")
+
+player_stats_seasonal_offense_recentRB <- player_stats_seasonal_offense_recent %>% 
+  filter(position_group == "RB")
+
+player_stats_seasonal_offense_recentWR <- player_stats_seasonal_offense_recent %>% 
+  filter(position_group == "WR")
+
+player_stats_seasonal_offense_recentTE <- player_stats_seasonal_offense_recent %>% 
+  filter(position_group == "TE")
+```
+
+#### Identify the Optimal Number of Tiers by Position {#sec-clusterAnalysisExampleNumTiers}
+
+##### Quarterbacks {#sec-clusterAnalysisExampleNumTiersQBs}
+
+```{r}
+tiersQB_bic <- mclustBIC(
+  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  G = 1:9
+)
+
+tiersQB_bic
+summary(tiersQB_bic)
+plot(tiersQB_bic)
+
+tiersQB_icl <- mclustICL(
+  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  G = 1:9
+)
+
+tiersQB_icl
+summary(tiersQB_icl)
+plot(tiersQB_icl)
+
+tiersQB_boostrap <- mclustBootstrapLRT(
+  player_stats_seasonal_offense_recentQB$fantasy_points,
+  modelName = "V") # variable/unequal variance (for univariate data)
+
+numTiersQB <- as.numeric(summary(tiersQB_boostrap)[,"Length"][1]) # or could specify the number of teams manually
+
+tiersQB_boostrap
+plot(
+  tiersQB_boostrap,
+  G = numTiersQB - 1)
+```
+
+##### Running Backs {#sec-clusterAnalysisExampleNumTiersRBs}
+
+```{r}
+tiersRB_bic <- mclustBIC(
+  data = player_stats_seasonal_offense_recentRB$fantasy_points,
+  G = 1:9
+)
+
+tiersRB_bic
+summary(tiersRB_bic)
+plot(tiersRB_bic)
+
+tiersRB_icl <- mclustICL(
+  data = player_stats_seasonal_offense_recentRB$fantasy_points,
+  G = 1:9
+)
+
+tiersRB_icl
+summary(tiersRB_icl)
+plot(tiersRB_icl)
+
+tiersRB_boostrap <- mclustBootstrapLRT(
+  player_stats_seasonal_offense_recentRB$fantasy_points,
+  modelName = "V") # variable/unequal variance (for univariate data)
+
+numTiersRB <- as.numeric(summary(tiersRB_boostrap)[,"Length"][1]) # or could specify the number of teams manually
+
+tiersRB_boostrap
+plot(
+  tiersRB_boostrap,
+  G = numTiersRB - 1)
+```
+
+##### Wide Receivers {#sec-clusterAnalysisExampleNumTiersWRs}
+
+```{r}
+tiersWR_bic <- mclustBIC(
+  data = player_stats_seasonal_offense_recentWR$fantasy_points,
+  G = 1:9
+)
+
+tiersWR_bic
+summary(tiersWR_bic)
+plot(tiersWR_bic)
+
+tiersWR_icl <- mclustICL(
+  data = player_stats_seasonal_offense_recentWR$fantasy_points,
+  G = 1:9
+)
+
+tiersWR_icl
+summary(tiersWR_icl)
+plot(tiersWR_icl)
+
+tiersWR_boostrap <- mclustBootstrapLRT(
+  player_stats_seasonal_offense_recentWR$fantasy_points,
+  modelName = "V") # variable/unequal variance (for univariate data)
+
+numTiersWR <- as.numeric(summary(tiersWR_boostrap)[,"Length"][1]) # or could specify the number of teams manually
+
+tiersWR_boostrap
+plot(
+  tiersWR_boostrap,
+  G = numTiersWR - 1)
+```
+
+##### Tight Ends {#sec-clusterAnalysisExampleNumTiersTEs}
+
+```{r}
+tiersTE_bic <- mclustBIC(
+  data = player_stats_seasonal_offense_recentTE$fantasy_points,
+  G = 1:9
+)
+
+tiersTE_bic
+summary(tiersTE_bic)
+plot(tiersTE_bic)
+
+tiersTE_icl <- mclustICL(
+  data = player_stats_seasonal_offense_recentTE$fantasy_points,
+  G = 1:9
+)
+
+tiersTE_icl
+summary(tiersTE_icl)
+plot(tiersTE_icl)
+
+tiersTE_boostrap <- mclustBootstrapLRT(
+  player_stats_seasonal_offense_recentTE$fantasy_points,
+  modelName = "V") # variable/unequal variance (for univariate data)
+
+numTiersTE <- as.numeric(summary(tiersTE_boostrap)[,"Length"][1]) # or could specify the number of teams manually
+
+tiersTE_boostrap
+plot(
+  tiersTE_boostrap,
+  G = numTiersTE - 1)
+```
+
+#### Fit the Cluster Model to the Optimal Number of Tiers {#sec-clusterAnalysisExampleNumTiers}
+
+##### Quarterbacks {#sec-clusterAnalysisExampleNumTiersQBs}
+
+In our data, all of the following models are equivalent—i.e., they result in the same unequal variance model with a 4-cluster solution—but they arrive there in different ways.
+
+```{r}
+#| eval: false
+
+Mclust(
+  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  G = numTiersQB,
+)
+
+Mclust(
+  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  G = 4,
+)
+
+Mclust(
+  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+)
+
+Mclust(
+  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  x = tiersQB_bic
+)
+```
+
+Let's fit one of these:
+
+```{r}
+clusterModelQBs <- Mclust(
+  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  G = numTiersQB,
+)
+```
+
+Here are the number of players that are in each of the four clusters (i.e., tiers):
+
+```{r}
+table(clusterModelQBs$classification)
+```
+
+We can merge the player's classification into the dataset and plot each player's classification:
+
+```{r}
+#| label: fig-qbTiers
+#| fig-cap: "Quarterback Fantasy Points by Tier."
+
+player_stats_seasonal_offense_recentQB$tier <- clusterModelQBs$classification
+
+player_stats_seasonal_offense_recentQB$position_rank <- rank(
+  player_stats_seasonal_offense_recentQB$fantasy_points * -1,
+  na.last = "keep",
+  ties.method = "min")
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense_recentQB,
+  mapping = aes(
+    x = fantasy_points,
+    y = position_rank,
+    color = as.factor(tier)
+  )) +
+  geom_label(
+    aes(
+      x = fantasy_points,
+      label = round(fantasy_points, 0),
+      fill = as.factor(tier)
+      ),
+    color = "white",
+    size = 5,
+    #show.legend = FALSE
+    ) +
+  geom_label(
+    aes(
+      x = fantasy_points,
+      label = player_display_name,
+      fill = as.factor(tier)),
+    color = "white",
+    hjust = 1.1,
+    size = 4,
+    show.legend = FALSE
+    ) +
+  scale_y_continuous(trans = "reverse") +
+  coord_cartesian(clip = "off") +
+  theme_classic() +
+  labs(
+    x = "Projected Points",
+    y = "Rank",
+    title = "Quarterback Fantasy Points by Tier",
+    fill = "Tier")
+```
+
+##### Running Backs {#sec-clusterAnalysisExampleNumTiersRBs}
+
+##### Wide Receivers {#sec-clusterAnalysisExampleNumTiersWRs}
+
+##### Tight Ends {#sec-clusterAnalysisExampleNumTiersTEs}
+
+## Conclusion {#sec-clusterAnalysisConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-clusterAnalysisSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-07-26 modern-portfolio-theory.qmd
20240725 - cluster analysis
20240725 - updates

@@
 
-Modern portfolio theory is a framework for determining the optimal composition of an investment portfolio to maximize expected returns for a given level of risk.
+Modern portfolio theory (mean-variance theory) is a framework for determining the optimal composition of an investment portfolio to maximize expected returns for a given level of risk.
 Here, *risk* refers to the [variability](#sec-descriptiveStatisticsSpread) (e.g., standard deviation or variance) of returns across time.
@@
 Another important idea from modern portfolio theory is that, if you want to achieve higher returns, you may be able to by accepting additional—and the right combination of—risk.
-That is, higher returns generally requires taking on additional risk—at least as long as we stay along the [efficient frontier](#sec-efficientFrontier), described next.
+In general, risk is positively correlated with return.
+That is, receiving higher returns generally requires taking on additional risk—at least as long as we stay along the [efficient frontier](#sec-efficientFrontier), described next.
 
@@
 The efficient frontier is helpful for identifying the optimal portfolio—the optimal combination and weighting of assets—for a given risk level.
+Anything below the efficient frontier is considered inefficient (i.e., lower-than-maximum returns for a given level of risk).
 
@@
 
+### Based on Variability Across Projection Sources {#sec-efficientFrontierFantasyProjectionsUncertainty}
+
+### Based on Historical Game-to-Game Variability {#sec-efficientFrontierFantasyProjectionsUncertainty}
+
 <https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league/> (archived at <https://perma.cc/JQ6G-KSRT>)

### 2024-07-26 sports-cognitive-psychology.qmd
20240725 - updates

@@
 ```
+
+## Overview {#sec-sportsCognitivePsychologyOverview}
+
+<https://stacker.com/sports/15-ways-analytics-has-changed-sports> (archived at <https://perma.cc/PQ5R-TWFA>)
+
+- Bill James, "Moneyball", Billy Beane, Theo Epstein, and role of Sabermetrics in helping MLB teams with limited budgets (e.g., Oakland Athletics) compete with larger budgets: using statistics to identify player value more accurately, especially for identifying undervalued players
+- development of advanced metrics (e.g., value over replacement player, wins above replacement, fieling independent pitching)
+- focus on on-base percentage and slugging percentage (and on-base plus slugging percentage) instead of batting average; defensive shifting; fewer attempts to steal bases; fewer bunts; batters taking more pitches; more frequent pitching changes (for pitcher–batter matchups, e.g., righty–lefty); focus on velocity and spin rate
+- other sports, including basketball and football are making more use of analytics
+- basketball: types of shots (focusing on three pointers and layups); star players resting more games
+- football: going for it on fourth down; greater emphasis on the passing game; drafting Running Backs later in the draft (and, overall, valuing Running Backs less); trading down in the draft for more low picks than fewer high picks [because top picks are frequently overvalued; @Massey2013]
 

### 2024-07-26 fantasy-football.qmd
20240608 - cleanup
20240726 - additional context

@@
 
-An example formation is depicted in Figure @fig-footballFormation.
+An example formation is depicted in @fig-footballFormation.
 
@@
 Behind each end zone is a field goal post.
-A diagram of a football field is depicted in Figure @fig-footballField.
+A diagram of a football field is depicted in @fig-footballField.
 
@@
 Managers make decisions such as selecting which players to draft, selecting which players to play (i.e., "start") on a weekly basis, identifying players to pick up from the remaining pool of available players (i.e., waiver wire), and making trades with other teams.
+
+There are variety of types of fantasy football leagues.
+In standard re-draft leagues, managers re-draft players each season.
+In keeper leagues, managers are allowed to keep one or more players from one season to the next, possibly for some cost (e.g., toward a keeper cap, loss of future draft pick).
+In dynasty leagues, managers act like a general manager and keep most of their roster from year to year.
+They may involve player contracts, salary caps, and free agent drafts.
+In best ball leagues, the manager's best possible lineup (in terms of the highest-scoring players for the necessary roster positions) are automatically selected for that week's lineup.
+
 Fantasy football relies heavily on prediction—trying to predict which players will perform best and selecting them accordingly.
@@
 Managers receive points for only those players who are on their starting lineup (not players on their bench).
-A manager's goal is to outscore their opponent each week to win matches and ultimately claim victory in the league.
+In a standard league, a manager's goal is to outscore their opponent each week to win matches and ultimately claim victory in the league.
+At the end of the regular season, many leagues have a playoffs to determine the league champion.
+In total points leagues, the league champion is determined by how many points they score throughout the entire season, rather than based on weekly matchups and playoffs.
+
 Scoring settings can differ from league to league.
-
 Below are common scoring settings for fantasy leagues.

### 2024-07-26 modern-portfolio-theory.qmd
20240726 - additional context

@@
 There is little consistency in the performance of mutual fund managers over time.
-The following charts are from Leonard Mlodinow's book, "The Drunkard's Walk: How Randomness Rules Our Lives".
-The chart on the left depicts the performance of the top mutual funds from 1991 to 1995.
-The chart on the right depicts the performance of the same funds in the same order over the subsequent 5 years (1996 to 2000):
-
-The best funds from 1991–1995 were not necessarily the best funds from 1996–2000.
+In the book, "The Drunkard's Walk: How Randomness Rules Our Lives", @Mlodinow2008 reported essentially no correlation between performance of the top mutual funds in a five-year period with their performance over the subsequent five years.
+That is, the best funds in a one period were not necessarily the best funds in another period.
 This suggests that mutual fund managers differ in great part because of luck or chance rather than reliable skill.
+In any given year, some mutual funds will do better than other mutual funds.
+But this overperformance in a given year likely reflects more randomness than skill.
 That is likely why [a cat beat professional investors in a stock market challenge](http://www.npr.org/blogs/money/2013/01/14/169326326/housecat-beats-investors-in-stock-market-challenge) (archived at <https://perma.cc/R3XU-K6J8>).
@@
 This evidence is consistent with the notion of the wisdom of the crowd.
-Moreover, even if the stock market is relatively efficient in terms of valuing stocks based on the available information, your fantasy football league is likely not.
+Moreover, even if the stock market is relatively accurate ("efficient") in terms of valuing stocks based on all (publicly) available information (i.e., the efficient market hypothesis), your fantasy football league is likely not.
 Thus, it may be effective to use crowd-based projections to identify players who are undervalued by your league.
@@
 In fantasy football, a player's risk might be quantified in terms of the [variability](#sec-descriptiveStatisticsSpread) of projected scores for a player across projection sources (e.g., Projection Source A, Source B, Souce C, etc.), or as historical game-to-game [variability](#sec-descriptiveStatisticsSpread).
+Variability of projected scores for a player across projection sources could reflect the *uncertainty* of projections for a player.
+Variability of historical (actual) fantasy points across games could reflect many factors, including risks due to injuries, situational changes (e.g., being traded to a new team or changes in team composition such as due to the acquisition of new players on the team), game scripts, and the tendency for the player to be "boom-or-bust" (e.g., if they are highly dependent on scoring touchdowns or long receptions for fantasy points).
 All things equal, we want to minimize our risk for a given level of expected returns.
@@
 
-The ultimate goal in fantasy football is to draft players for your starting lineup that provide the most projected points and the smallest downside risk.
+The ultimate goal in fantasy football is to draft players for your starting lineup that provide the most projected points (i.e., the highest returns) and the smallest downside risk.
 That is, your goal is to achieve the optimal portfolio at a given level of risk, depending on how much risk you are willing to tolerate.
@@
 
+In fantasy football, the efficient frontier can be helpful for identifying the optimal players to draft for a given risk level (and potentially within the salary cap).
+It can also be helpful for identifying potential trades.
+In this way, modern portfolio theory and the efficient frontier can be helpful for arbitrage—buying and selling the same asset (in this case, player) to take advantage of different prices for the same asset.
+
 ### Based on Variability Across Projection Sources {#sec-efficientFrontierFantasyProjectionsUncertainty}

### 2024-07-27 index.qmd
20240727 - mixed models

@@
 
-```
+```{bibtex}
 @book{petersenFantasyFootballAnalytics,
-  title = {Fantasy football analytics: Statistics, prediction, and empiricism Using R},
+  title = {Fantasy football analytics: Statistics, prediction, and empiricism using {R}},
   author = {Petersen, Isaac T.},

### 2024-07-27 mixed-models.qmd
20240721 - additions
20240727 - mixed models

@@
 ```{r}
-
+library("petersenlab")
+library("lme4")
+library("lmerTest")
+library("MuMIn")
+library("emmeans")
+library("sjstats")
+library("mgcv")
+library("tidyverse")
 ```
 
+### Specify Package Options {#sec-mixedModelsPackageOptions}
+
+```{r}
+emm_options(lmerTest.limit = 100000)
+emm_options(pbkrtest.limit = 100000)
+```
+
+### Load Data {#sec-mixedModelsLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+```
+
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
+
 ## Overview of Mixed Models {#sec-mixedModelsOverview}
+
+## Fantasy Points Per Season by Position, Age, and Experience {#sec-fantasyPointsByAgeExperience}
+
+```{r}
+player_stats_seasonal_offense_subset <- player_stats_seasonal_offense %>% 
+  filter(position_group %in% c("QB","RB","WR","TE"))
+
+player_stats_seasonal_offense_subset$position[which(player_stats_seasonal_offense_subset$position == "HB")] <- "RB"
+
+player_stats_seasonal_kicking_subset <- player_stats_seasonal_kicking %>% 
+  filter(position == "K")
+
+player_stats_seasonal_offense_subset <- bind_rows(
+  player_stats_seasonal_offense_subset,
+  player_stats_seasonal_kicking_subset
+)
+
+player_stats_seasonal_offense_subset$player_idFactor <- factor(player_stats_seasonal_offense_subset$player_id)
+```
+
+### Models {#sec-fantasyPointsByAgeExperienceModels}
+
+```{r}
+pointsPerSeason_nullModel <- lmerTest::lmer(
+  fantasy_points ~ 1 + (1 | player_id),
+  data = player_stats_seasonal_offense_subset,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_nullModel)
+MuMIn::r.squaredGLMM(pointsPerSeason_nullModel)
+performance::icc(pointsPerSeason_nullModel)
+
+pointsPerSeason_position <- lmerTest::lmer(
+  fantasy_points ~ as.factor(position) + (1 | player_id),
+  data = player_stats_seasonal_offense_subset,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_position)
+MuMIn::r.squaredGLMM(pointsPerSeason_position)
+emmeans::emmeans(pointsPerSeason_position, "position")
+performance::icc(pointsPerSeason_position)
+
+pointsPerSeason_positionAgeFixedLinearSlopes <- lmerTest::lmer(
+  fantasy_points ~ as.factor(position) + ageCentered20 + (1 | player_id),
+  data = player_stats_seasonal_offense_subset,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_positionAgeFixedLinearSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
+emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "position")
+emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "ageCentered20")
+performance::icc(pointsPerSeason_positionAgeFixedLinearSlopes)
+
+pointsPerSeason_positionAgeRandomLinearSlopes <- lmerTest::lmer(
+  fantasy_points ~ as.factor(position) + ageCentered20 + (1 + ageCentered20 | player_id),
+  data = player_stats_seasonal_offense_subset,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_positionAgeRandomLinearSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "position")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "ageCentered20")
+performance::icc(pointsPerSeason_positionAgeRandomLinearSlopes)
+
+pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes <- lmerTest::lmer(
+  fantasy_points ~ as.factor(position) + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 | player_id),
+  data = player_stats_seasonal_offense_subset,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "position")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
+performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+
+pointsPerSeason_positionAgeRandomLinearRandomQuadraticSlopes <- lmerTest::lmer(
+  fantasy_points ~ as.factor(position) + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 + ageCentered20Quadratic | player_id),
+  data = player_stats_seasonal_offense_subset,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction <- lmerTest::lmer(
+  fantasy_points ~ as.factor(position) + ageCentered20 + ageCentered20Quadratic + as.factor(position):ageCentered20 + as.factor(position):ageCentered20Quadratic + (1 + ageCentered20 | player_id),
+  data = player_stats_seasonal_offense_subset,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "position")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "ageCentered20")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
+performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+
+pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
+  fantasy_points ~ as.factor(position) + ageCentered20 + ageCentered20Quadratic + as.factor(position):ageCentered20 + as.factor(position):ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
+  data = player_stats_seasonal_offense_subset,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "position")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20Quadratic")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "years_of_experience")
+performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+
+pointsPerSeason_gam <- gam(
+  fantasy_points ~ as.factor(position) + ageCentered20 + as.factor(position):ageCentered20 + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
+  data = player_stats_seasonal_offense_subset
+)
+
+summary(pointsPerSeason_gam)
+```
+
+Compare models:
+
+```{r}
+anova(
+  pointsPerSeason_nullModel,
+  pointsPerSeason_position,
+)
+
+anova(
+  pointsPerSeason_positionAgeFixedLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
+)
+```
+
+### Plots of Model-Implied Fantasy Points by Position, Age, and Experience {#sec-fantasyPointsByAgeExperiencePlots}
+
+```{r}
+pointsPerSeason_positionAge_newData <- expand.grid(
+  position = factor(c("FB","QB","RB","TE","WR")), #,"K"
+  age = seq(from = 20, to = 40, length.out = 10000)
+)
+
+pointsPerSeason_positionAge_newData$ageCentered20 <- pointsPerSeason_positionAge_newData$age - 20
+pointsPerSeason_positionAge_newData$ageCentered20Quadratic <- pointsPerSeason_positionAge_newData$ageCentered20 ^ 2
+pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$ageCentered20)
+
+pointsPerSeason_positionAge_newData$fantasyPoints <- predict(
+  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
+  newdata = pointsPerSeason_positionAge_newData,
+  re.form = NA
+)
+
+pointsPerSeason_positionAge_newData$fantasyPoints_gam <- predict(
+  object = pointsPerSeason_gam,
+  newdata = pointsPerSeason_positionAge_newData,
+  newdata.guaranteed = TRUE,
+  exclude = "s(player_idFactor,ageCentered20)"
+)
+```
+
+#### Model-Implied Fantasy Points by Position and Age {#sec-fantasyPointsByPositionAgePlot}
+
+```{r}
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints,
+    color = position
+  )
+) + 
+  geom_smooth() +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position"
+  ) +
+  theme_classic()
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_gam,
+    color = position
+  )
+) + 
+  geom_smooth() +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position"
+  ) +
+  theme_classic()
+```
+
+#### Model-Implied Fantasy Points by Experience {#sec-fantasyPointsByExperiencePlot}
 

### 2024-07-27 basic-statistics.qmd
20240721 - typo
20240727 - cleanup

@@
     size = 7) + 
-  ggplot2::labs(
+  labs(
     x = "Predictor Variable",
@@
   ) +
-  ggplot2::theme_classic(
+  theme_classic(
     base_size = 16) +
-  ggplot2::theme(
+  theme(
     legend.title = element_blank(),
@@
 
-::: {.content-visible when-format="html"}
-
 ## Conclusion {#sec-basicStatsConclusion}
@@
 
+::: {.content-visible when-format="html"}
+
 ## Session Info {#sec-basicStatsSessionInfo}

### 2024-07-27 mixed-models.qmd
20240727 - mixed models

--- 2024-07-21 mixed-models.qmd
@@
 ```{r}
-
+library("petersenlab")
+library("lme4")
+library("lmerTest")
+library("MuMIn")
+library("emmeans")
+library("sjstats")
+library("mgcv")
+library("AICcmodavg")
+library("parallel")
+library("tidyverse")
 ```
 
+### Specify Package Options {#sec-mixedModelsPackageOptions}
+
+```{r}
+emm_options(lmerTest.limit = 100000)
+emm_options(pbkrtest.limit = 100000)
+```
+
+### Load Data {#sec-mixedModelsLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_depthCharts.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/nfl_depthCharts.RData")
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+```
+
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
+
 ## Overview of Mixed Models {#sec-mixedModelsOverview}
+
+## Fantasy Points Per Season by Position, Age, and Experience {#sec-fantasyPointsByAgeExperience}
+
+```{r}
+player_stats_seasonal_offense_subset <- player_stats_seasonal_offense %>% 
+  filter(position_group %in% c("QB","RB","WR","TE"))
+
+player_stats_seasonal_offense_subset$position[which(player_stats_seasonal_offense_subset$position == "HB")] <- "RB"
+
+player_stats_seasonal_kicking_subset <- player_stats_seasonal_kicking %>% 
+  filter(position == "K")
+
+player_stats_seasonal_offense_subset <- bind_rows(
+  player_stats_seasonal_offense_subset,
+  player_stats_seasonal_kicking_subset
+)
+
+player_stats_seasonal_offense_subset$player_idFactor <- factor(player_stats_seasonal_offense_subset$player_id)
+player_stats_seasonal_offense_subset$positionFactor <- factor(player_stats_seasonal_offense_subset$position)
+
+#for(i in unique(nfl_depthCharts$season)){
+#  print(i)
+#  nfl_depthCharts %>%
+#    filter(season == i) %>% 
+#    select(week) %>% 
+#    arrange(week) %>% 
+#    unique() %>% 
+#    pull() %>% 
+#    print()
+#}
+
+seasons17week <- 2001:2020
+seasons18week <- 2021:max(nfl_depthCharts$season, na.rm = TRUE)
+
+endOfSeasonDepthCharts <- nfl_depthCharts %>% 
+  filter((season %in% seasons17week & week == 18) | (season %in% seasons18week & week == 19)) # get end-of-season depth charts
+
+qb1s <- endOfSeasonDepthCharts %>% 
+  filter(position == "QB", depth_team == 1)
+
+fb1s <- endOfSeasonDepthCharts %>% 
+  filter(position == "FB", depth_team == 1)
+
+k1s <- endOfSeasonDepthCharts %>% 
+  filter(position == "K", depth_team == 1)
+
+rb1s <- endOfSeasonDepthCharts %>% 
+  filter(position == "RB", depth_team == 1)
+
+wr1s <- endOfSeasonDepthCharts %>% 
+  filter(position == "WR", depth_team == 1)
+
+te1s <- endOfSeasonDepthCharts %>% 
+  filter(position == "TE", depth_team == 1)
+
+player_stats_seasonal_offense_subsetDepth <- player_stats_seasonal_offense_subset %>% 
+  filter(player_id %in% c(
+    qb1s$gsis_id,
+    fb1s$gsis_id,
+    k1s$gsis_id,
+    rb1s$gsis_id,
+    wr1s$gsis_id,
+    te1s$gsis_id
+    ))
+```
+
+### Plots of Raw Data {#sec-fantasyPointsByAgeExperienceRawData}
+
+Plots of players' raw fantasy points data by age are in Figures [-@fig-rawFantasyPointsByAgeQB]–[-@fig-rawFantasyPointsByAgeTE].
+
+```{r}
+#| label: fig-rawFantasyPointsByAgeQB
+#| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Quarterbacks."
+
+ggplot(
+  data = player_stats_seasonal_offense_subset %>% filter(position == "QB"),
+  mapping = aes(
+    x = age,
+    y = fantasy_points,
+    group = player_id)) +
+  geom_line()  +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Quarterbacks"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-rawFantasyPointsByAgeFB
+#| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Fullbacks."
+
+ggplot(
+  data = player_stats_seasonal_offense_subset %>% filter(position == "FB"),
+  mapping = aes(
+    x = age,
+    y = fantasy_points,
+    group = player_id)) +
+  geom_line()  +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Fullbacks"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-rawFantasyPointsByAgeRB
+#| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Running Backs."
+
+ggplot(
+  data = player_stats_seasonal_offense_subset %>% filter(position == "RB"),
+  mapping = aes(
+    x = age,
+    y = fantasy_points,
+    group = player_id)) +
+  geom_line()  +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Running Backs"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-rawFantasyPointsByAgeWR
+#| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Wide Receivers."
+
+ggplot(
+  data = player_stats_seasonal_offense_subset %>% filter(position == "WR"),
+  mapping = aes(
+    x = age,
+    y = fantasy_points,
+    group = player_id)) +
+  geom_line()  +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Wide Receivers"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-rawFantasyPointsByAgeTE
+#| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Tight Ends."
+
+ggplot(
+  data = player_stats_seasonal_offense_subset %>% filter(position == "TE"),
+  mapping = aes(
+    x = age,
+    y = fantasy_points,
+    group = player_id)) +
+  geom_line()  +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Tight Ends"
+  ) +
+  theme_classic()
+```
+
+### Models {#sec-fantasyPointsByAgeExperienceModels}
+
+```{r}
+pointsPerSeason_nullModel <- lmerTest::lmer(
+  fantasy_points ~ 1 + (1 | player_idFactor),
+  data = player_stats_seasonal_offense_subset,
+  REML = FALSE,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_nullModel)
+MuMIn::r.squaredGLMM(pointsPerSeason_nullModel)
+performance::icc(pointsPerSeason_nullModel)
+AIC(pointsPerSeason_nullModel)
+
+pointsPerSeason_position <- lmerTest::lmer(
+  fantasy_points ~ positionFactor + (1 | player_idFactor),
+  data = player_stats_seasonal_offense_subset,
+  REML = FALSE,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_position)
+MuMIn::r.squaredGLMM(pointsPerSeason_position)
+emmeans::emmeans(pointsPerSeason_position, "positionFactor")
+performance::icc(pointsPerSeason_position)
+AIC(pointsPerSeason_position)
+
+pointsPerSeason_positionAgeFixedLinearSlopes <- lmerTest::lmer(
+  fantasy_points ~ positionFactor + ageCentered20 + (1 | player_idFactor),
+  data = player_stats_seasonal_offense_subset,
+  REML = FALSE,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_positionAgeFixedLinearSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
+emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "positionFactor")
+emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "ageCentered20")
+performance::icc(pointsPerSeason_positionAgeFixedLinearSlopes)
+AIC(pointsPerSeason_positionAgeFixedLinearSlopes)
+
+pointsPerSeason_positionAgeRandomLinearSlopes <- lmerTest::lmer(
+  fantasy_points ~ positionFactor + ageCentered20 + (1 + ageCentered20 | player_idFactor),
+  data = player_stats_seasonal_offense_subset,
+  REML = FALSE,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_positionAgeRandomLinearSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "positionFactor")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "ageCentered20")
+performance::icc(pointsPerSeason_positionAgeRandomLinearSlopes)
+AIC(pointsPerSeason_positionAgeRandomLinearSlopes)
+
+pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes <- lmerTest::lmer(
+  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
+  data = player_stats_seasonal_offense_subset,
+  REML = FALSE,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "positionFactor")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
+performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+
+pointsPerSeason_positionAgeRandomLinearRandomQuadraticSlopes <- lmerTest::lmer(
+  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 + ageCentered20Quadratic | player_idFactor),
+  data = player_stats_seasonal_offense_subset,
+  REML = FALSE,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction <- lmerTest::lmer(
+  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
+  data = player_stats_seasonal_offense_subset,
+  REML = FALSE,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "positionFactor")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "ageCentered20")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "ageCentered20Quadratic")
+performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+
+pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
+  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
+  data = player_stats_seasonal_offense_subset,
+  REML = FALSE,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "positionFactor")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20Quadratic")
+emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "years_of_experience")
+performance::icc(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+```
+
+Compare models:
+
+```{r}
+anova(
+  pointsPerSeason_nullModel,
+  pointsPerSeason_position
+)
+
+anova(
+  pointsPerSeason_positionAgeFixedLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
+)
+
+mixedModels <- list(
+  "null" = pointsPerSeason_nullModel,
+  "position" = pointsPerSeason_position,
+  "fixedLinear" = pointsPerSeason_positionAgeFixedLinearSlopes,
+  "randomLinear" = pointsPerSeason_positionAgeRandomLinearSlopes,
+  "randomLinearFixedQuadratic" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  "randomLinearFixedQuadraticInteraction" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
+)
+
+aictab(mixedModels)
+```
+
+Generalized Additive Model:
+
+```{r}
+num_cores <- detectCores()
+
+pointsPerSeason_gam <- bam( # using bam() instead of gam() for faster estimation due to large size of data
+  fantasy_points ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
+  data = player_stats_seasonal_offense_subset,
+  nthreads = num_cores
+)
+
+pointsPerSeason_gamSummary <- summary(pointsPerSeason_gam)
+
+pointsPerSeason_gamSummary
+pointsPerSeason_gamSummary$r.sq
+MuMIn::r.squaredGLMM(pointsPerSeason_gam)
+AIC(pointsPerSeason_gam)
+```
+
+Players at the top of the end-of-season depth chart:
+
+```{r}
+pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
+  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
+  data = player_stats_seasonal_offense_subset,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+MuMIn::r.squaredGLMM(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "positionFactor")
+emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20")
+emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "ageCentered20Quadratic")
+emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "years_of_experience")
+performance::icc(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+
+pointsPerSeasonDepth_gam <- bam( # using bam() instead of gam() for faster estimation due to large size of data
+  fantasy_points ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
+  data = player_stats_seasonal_offense_subsetDepth,
+  nthreads = num_cores
+)
+
+pointsPerSeasonDepth_gamSummary <- summary(pointsPerSeasonDepth_gam)
+
+pointsPerSeasonDepth_gamSummary
+pointsPerSeasonDepth_gamSummary$r.sq
+MuMIn::r.squaredGLMM(pointsPerSeasonDepth_gam)
+AIC(pointsPerSeasonDepth_gam)
+```
+
+### Plots of Model-Implied Fantasy Points by Position and Age {#sec-fantasyPointsByAgePlots}
+
+```{r}
+# Create newdata object
+pointsPerSeason_positionAge_newData <- expand.grid(
+  positionFactor = factor(c("FB","QB","RB","TE","WR")), #,"K"
+  age = seq(from = 20, to = 40, length.out = 10000)
+)
+
+pointsPerSeason_positionAge_newData$ageCentered20 <- pointsPerSeason_positionAge_newData$age - 20
+pointsPerSeason_positionAge_newData$ageCentered20Quadratic <- pointsPerSeason_positionAge_newData$ageCentered20 ^ 2
+pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$ageCentered20)
+
+# From Quadratic Model: All Players
+pointsPerSeason_positionAge_newData$fantasyPoints_quadratic <- predict(
+  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
+  newdata = pointsPerSeason_positionAge_newData,
+  re.form = NA
+)
+
+# From Quadratic Model: Players at Top of End-of-Season Depth Chart
+pointsPerSeason_positionAge_newData$fantasyPoints_depthQuadratic <- predict(
+  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
+  newdata = pointsPerSeason_positionAge_newData,
+  re.form = NA
+)
+
+# From GAM Model: All Players
+pointsPerSeason_positionAge_newData$fantasyPoints_gam <- predict(
+  object = pointsPerSeason_gam,
+  newdata = pointsPerSeason_positionAge_newData,
+  newdata.guaranteed = TRUE,
+  exclude = "s(player_idFactor,ageCentered20)"
+)
+
+# From GAM Model: Players at Top of End-of-Season Depth Chart
+pointsPerSeason_positionAge_newData$fantasyPoints_depthGAM <- predict(
+  object = pointsPerSeasonDepth_gam,
+  newdata = pointsPerSeason_positionAge_newData,
+  newdata.guaranteed = TRUE,
+  exclude = "s(player_idFactor,ageCentered20)"
+)
+```
+
+Plots of model-implied fantasy points by position and age are in Figures [-@fig-fantasyPointsByAgeQuadratic]–[-@fig-fantasyPointsByAgeGAM].
+
+```{r}
+#| label: fig-fantasyPointsByAgeQuadratic
+#| fig-cap: "Plot of Model-Implied Quadratic Trajectories of Fantasy Points by Age."
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_quadratic,
+    color = positionFactor
+  )
+) + 
+  geom_smooth() +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Quadratic Model with All Players"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-fantasyPointsByAgeQuadraticDepth
+#| fig-cap: "Plot of Model-Implied Quadratic Trajectories of Fantasy Points by Age For Players Who Were Once at the Top of the End-of-Season Depth Chart."
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_depthQuadratic,
+    color = positionFactor
+  )
+) + 
+  geom_smooth() +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Quadratic Model with Players Who Were Once at Top of End-of-Season Depth Chart"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-fantasyPointsByAgeGAM
+#| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age from a Generalized Additive Model."
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_gam,
+    color = positionFactor
+  )
+) + 
+  geom_smooth() +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Generalized Additive Model with All Players"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-fantasyPointsByAgeGAM
+#| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, For Players Who Were Once at the Top of the End-of-Season Depth Chart."
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_depthGAM,
+    color = positionFactor
+  )
+) + 
+  geom_smooth() +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Generalized Additive Model with Players Who Were Once at Top of End-of-Season Depth Chart"
+  ) +
+  theme_classic()
+```
+
+### Plots of Individuals' Model-Implied Fantasy Points by Age {#sec-fantasyPointsIndividualByAgePlots}
+
+```{r}
+player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subset %>%
+  filter(!is.na(player_idFactor), !is.na(fantasy_points), !is.na(positionFactor), !is.na(ageCentered20), !is.na(ageCentered20Quadratic), !is.na(years_of_experience))
+
+player_stats_seasonal_offense_subsetCC$fantasyPoints_quadratic <- predict(
+  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
+  newdata = player_stats_seasonal_offense_subsetCC
+)
+
+player_stats_seasonal_offense_subsetCC$fantasyPoints_gam <- predict(
+  object = pointsPerSeason_gam,
+  newdata = player_stats_seasonal_offense_subsetCC
+)
+```
+
+Plots of players' model-implied fantasy points by age are in Figures [-@fig-individualFantasyPointsByAgeQB]–[-@fig-individualFantasyPointsByAgeTE].
+
+```{r}
+#| label: fig-individualFantasyPointsByAgeQB
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Quarterbacks. Overlaid with the Model-Implied Trajectory for Quarterbacks."
+
+ggplot(
+  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "QB"),
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_gam,
+    group = player_id)) +
+  geom_smooth(
+    se = FALSE,
+    linewidth = 0.5,
+    color = "black") +
+  geom_smooth(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_gam
+    ),
+    data = pointsPerSeason_positionAge_newData %>% filter(positionFactor == "QB"),
+    inherit.aes = FALSE,
+    se = TRUE,
+    linewidth = 2
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Quarterbacks"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-individualFantasyPointsByAgeFB
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Fullbacks. Overlaid with the Model-Implied Trajectory for Fullbacks."
+
+ggplot(
+  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "FB"),
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_gam,
+    group = player_id)) +
+  geom_smooth(
+    se = FALSE,
+    linewidth = 0.5,
+    color = "black") +
+  geom_smooth(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_gam
+    ),
+    data = pointsPerSeason_positionAge_newData %>% filter(positionFactor == "FB"),
+    inherit.aes = FALSE,
+    se = TRUE,
+    linewidth = 2
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Fullbacks"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-individualFantasyPointsByAgeRB
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Running Backs. Overlaid with the Model-Implied Trajectory for Running Backs."
+
+ggplot(
+  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "RB"),
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_gam,
+    group = player_id)) +
+  geom_smooth(
+    se = FALSE,
+    linewidth = 0.5,
+    color = "black") +
+  geom_smooth(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_gam
+    ),
+    data = pointsPerSeason_positionAge_newData %>% filter(positionFactor == "RB"),
+    inherit.aes = FALSE,
+    se = TRUE,
+    linewidth = 2
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Running Backs"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-individualFantasyPointsByAgeWR
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Wide Receivers. Overlaid with the Model-Implied Trajectory for Wide Receivers."
+
+ggplot(
+  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "WR"),
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_gam,
+    group = player_id)) +
+  geom_smooth(
+    se = FALSE,
+    linewidth = 0.5,
+    color = "black") +
+  geom_smooth(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_gam
+    ),
+    data = pointsPerSeason_positionAge_newData %>% filter(positionFactor == "WR"),
+    inherit.aes = FALSE,
+    se = TRUE,
+    linewidth = 2
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Wide Receivers"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-individualFantasyPointsByAgeTE
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Tight Ends. Overlaid with the Model-Implied Trajectory for Wide Tight Ends."
+
+ggplot(
+  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "TE"),
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_gam,
+    group = player_id)) +
+  geom_smooth(
+    se = FALSE,
+    linewidth = 0.5,
+    color = "black") +
+  geom_smooth(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_gam
+    ),
+    data = pointsPerSeason_positionAge_newData %>% filter(positionFactor == "TE"),
+    inherit.aes = FALSE,
+    se = TRUE,
+    linewidth = 2
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Tight Ends"
+  ) +
+  theme_classic()
+```
 

### 2024-07-27 mixed-models.qmd
20240727 - duplicate label

@@
 
-Plots of model-implied fantasy points by position and age are in Figures [-@fig-fantasyPointsByAgeQuadratic]–[-@fig-fantasyPointsByAgeGAM].
+Plots of model-implied fantasy points by position and age are in Figures [-@fig-fantasyPointsByAgeQuadratic]–[-@fig-fantasyPointsByAgeGAMdepth].
 
@@
 ```{r}
-#| label: fig-fantasyPointsByAgeGAM
+#| label: fig-fantasyPointsByAgeGAMdepth
 #| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, For Players Who Were Once at the Top of the End-of-Season Depth Chart."

### 2024-07-27 mixed-models.qmd
20240727 - fixes

@@
 num_cores <- detectCores()
+num_cores
 
@@
 performance::icc(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+AIC(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
 

### 2024-07-27 cluster-analysis.qmd
20240727 - cluster analysis

@@
 library("tidyverse")
+library("plotly")
 ```
@@
 
-### Example {#sec-clusterAnalysisExample}
+### Example: Tiers of Prior Season Fantasy Points {#sec-clusterAnalysisExample}
 
@@
 
-tiersRB_boostrap <- mclustBootstrapLRT(
-  player_stats_seasonal_offense_recentRB$fantasy_points,
-  modelName = "V") # variable/unequal variance (for univariate data)
+numTiersRB <- 3
 
-numTiersRB <- as.numeric(summary(tiersRB_boostrap)[,"Length"][1]) # or could specify the number of teams manually
-
-tiersRB_boostrap
-plot(
-  tiersRB_boostrap,
-  G = numTiersRB - 1)
+# The following code errors out:
+#tiersRB_boostrap <- mclustBootstrapLRT(
+#  player_stats_seasonal_offense_recentRB$fantasy_points,
+#  modelName = "V") # variable/unequal variance (for univariate data)
+#
+#numTiersRB <- as.numeric(summary(tiersRB_boostrap)[,"Length"][1]) # or could specify the number of teams manually
+#
+#tiersRB_boostrap
+#plot(
+#  tiersRB_boostrap,
+#  G = numTiersRB - 1)
 ```
@@
 
-We can merge the player's classification into the dataset and plot each player's classification:
+##### Running Backs {#sec-clusterAnalysisExampleNumTiersRBs}
+
+```{r}
+clusterModelRBs <- Mclust(
+  data = player_stats_seasonal_offense_recentRB$fantasy_points,
+  G = numTiersRB,
+)
+```
+
+Here are the number of players that are in each of the four clusters (i.e., tiers):
+
+```{r}
+table(clusterModelRBs$classification)
+```
+
+##### Wide Receivers {#sec-clusterAnalysisExampleNumTiersWRs}
+
+```{r}
+clusterModelWRs <- Mclust(
+  data = player_stats_seasonal_offense_recentWR$fantasy_points,
+  G = numTiersWR,
+)
+```
+
+Here are the number of players that are in each of the four clusters (i.e., tiers):
+
+```{r}
+table(clusterModelWRs$classification)
+```
+
+##### Tight Ends {#sec-clusterAnalysisExampleNumTiersTEs}
+
+```{r}
+clusterModelTEs <- Mclust(
+  data = player_stats_seasonal_offense_recentTE$fantasy_points,
+  G = numTiersTE,
+)
+```
+
+Here are the number of players that are in each of the four clusters (i.e., tiers):
+
+```{r}
+table(clusterModelTEs$classification)
+```
+
+#### Plot the Tiers {#sec-clusterAnalysisExamplePlotTiers}
+
+We can merge the player's classification into the dataset and plot each player's classification.
+
+##### Quarterbacks {#sec-clusterAnalysisExamplePlotTiersQB}
 
@@
 
+player_stats_seasonal_offense_recentQB <- player_stats_seasonal_offense_recentQB %>%
+  mutate(
+    tier = factor(case_match( # recode so 1 is highest tier
+      tier,
+      1 ~ 4,
+      2 ~ 3,
+      3 ~ 2,
+      4 ~ 1)))
+
 player_stats_seasonal_offense_recentQB$position_rank <- rank(
@@
 
-ggplot2::ggplot(
+plot_qbTiers <- ggplot2::ggplot(
   data = player_stats_seasonal_offense_recentQB,
@@
     y = position_rank,
-    color = as.factor(tier)
+    color = tier
   )) +
-  geom_label(
+  geom_point(
     aes(
-      x = fantasy_points,
-      label = round(fantasy_points, 0),
-      fill = as.factor(tier)
-      ),
-    color = "white",
-    size = 5,
-    #show.legend = FALSE
-    ) +
-  geom_label(
-    aes(
-      x = fantasy_points,
-      label = player_display_name,
-      fill = as.factor(tier)),
-    color = "white",
-    hjust = 1.1,
-    size = 4,
-    show.legend = FALSE
-    ) +
+      text = player_display_name # add player name for mouse over tooltip
+  )) +
   scale_y_continuous(trans = "reverse") +
   coord_cartesian(clip = "off") +
-  theme_classic() +
   labs(
     x = "Projected Points",
-    y = "Rank",
+    y = "Position Rank",
     title = "Quarterback Fantasy Points by Tier",
-    fill = "Tier")
+    color = "Tier") +
+  theme_classic() +
+  theme(legend.position = "top")
+
+ggplotly(plot_qbTiers)
 ```
 
-##### Running Backs {#sec-clusterAnalysisExampleNumTiersRBs}
+##### Running Backs {#sec-clusterAnalysisExamplePlotTiersRBs}
 
-##### Wide Receivers {#sec-clusterAnalysisExampleNumTiersWRs}
+```{r}
+#| label: fig-rbTiers
+#| fig-cap: "Running Back Fantasy Points by Tier."
 
-##### Tight Ends {#sec-clusterAnalysisExampleNumTiersTEs}
+player_stats_seasonal_offense_recentRB$tier <- clusterModelRBs$classification
+
+player_stats_seasonal_offense_recentRB <- player_stats_seasonal_offense_recentRB %>%
+  mutate(
+    tier = factor(case_match( # recode so 1 is highest tier
+      tier,
+      1 ~ 3,
+      2 ~ 2,
+      3 ~ 1)))
+
+player_stats_seasonal_offense_recentRB$position_rank <- rank(
+  player_stats_seasonal_offense_recentRB$fantasy_points * -1,
+  na.last = "keep",
+  ties.method = "min")
+
+plot_rbTiers <- ggplot2::ggplot(
+  data = player_stats_seasonal_offense_recentRB,
+  mapping = aes(
+    x = fantasy_points,
+    y = position_rank,
+    color = tier
+  )) +
+  geom_point(
+    aes(
+      text = player_display_name # add player name for mouse over tooltip
+  )) +
+  scale_y_continuous(trans = "reverse") +
+  coord_cartesian(clip = "off") +
+  labs(
+    x = "Projected Points",
+    y = "Position Rank",
+    title = "Running Back Fantasy Points by Tier",
+    color = "Tier") +
+  theme_classic() +
+  theme(legend.position = "top")
+
+ggplotly(plot_rbTiers)
+```
+
+##### Wide Receivers {#sec-clusterAnalysisExamplePlotTiersWRs}
+
+```{r}
+#| label: fig-wrTiers
+#| fig-cap: "Quarterback Fantasy Points by Tier."
+
+player_stats_seasonal_offense_recentWR$tier <- clusterModelWRs$classification
+
+player_stats_seasonal_offense_recentWR <- player_stats_seasonal_offense_recentWR %>%
+  mutate(
+    tier = factor(case_match( # recode so 1 is highest tier
+      tier,
+      1 ~ 5,
+      2 ~ 4,
+      3 ~ 3,
+      4 ~ 2,
+      5 ~ 1)))
+
+player_stats_seasonal_offense_recentWR$position_rank <- rank(
+  player_stats_seasonal_offense_recentWR$fantasy_points * -1,
+  na.last = "keep",
+  ties.method = "min")
+
+plot_wrTiers <- ggplot2::ggplot(
+  data = player_stats_seasonal_offense_recentWR,
+  mapping = aes(
+    x = fantasy_points,
+    y = position_rank,
+    color = tier
+  )) +
+  geom_point(
+    aes(
+      text = player_display_name # add player name for mouse over tooltip
+  )) +
+  scale_y_continuous(trans = "reverse") +
+  coord_cartesian(clip = "off") +
+  labs(
+    x = "Projected Points",
+    y = "Position Rank",
+    title = "Wide Receiver Fantasy Points by Tier",
+    color = "Tier") +
+  theme_classic() +
+  theme(legend.position = "top")
+
+ggplotly(plot_wrTiers)
+```
+
+##### Tight Ends {#sec-clusterAnalysisExamplePlotTiersTEs}
+
+```{r}
+#| label: fig-teTiers
+#| fig-cap: "Tight End Fantasy Points by Tier."
+
+player_stats_seasonal_offense_recentTE$tier <- clusterModelTEs$classification
+
+player_stats_seasonal_offense_recentTE <- player_stats_seasonal_offense_recentTE %>%
+  mutate(
+    tier = factor(case_match( # recode so 1 is highest tier
+      tier,
+      1 ~ 4,
+      2 ~ 3,
+      3 ~ 2,
+      4 ~ 1)))
+
+player_stats_seasonal_offense_recentTE$position_rank <- rank(
+  player_stats_seasonal_offense_recentTE$fantasy_points * -1,
+  na.last = "keep",
+  ties.method = "min")
+
+plot_teTiers <- ggplot2::ggplot(
+  data = player_stats_seasonal_offense_recentTE,
+  mapping = aes(
+    x = fantasy_points,
+    y = position_rank,
+    color = tier
+  )) +
+  geom_point(
+    aes(
+      text = player_display_name # add player name for mouse over tooltip
+  )) +
+  scale_y_continuous(trans = "reverse") +
+  coord_cartesian(clip = "off") +
+  labs(
+    x = "Projected Points",
+    y = "Position Rank",
+    title = "Tight End Fantasy Points by Tier",
+    color = "Tier") +
+  theme_classic() +
+  theme(legend.position = "top")
+
+ggplotly(plot_teTiers)
+```
 

### 2024-07-27 mixed-models.qmd
20240727 - interactive plots

@@
 library("parallel")
+library("plotly")
+library("viridis")
 library("tidyverse")
@@
 
-ggplot(
-  data = player_stats_seasonal_offense_subset %>% filter(position == "QB"),
+plot_rawFantasyPointsByAgeQB <- ggplot(
+  data = player_stats_seasonal_offense_subset %>% 
+    filter(position == "QB") %>% 
+    mutate(
+      age = round(age, 2),
+      fantasy_points = round(fantasy_points, 2)
+    ),
   mapping = aes(
@@
     y = fantasy_points,
-    group = player_id)) +
-  geom_line()  +
+    color = player_id)) +
+  geom_line(
+    aes(
+      text = player_display_name # add player name for mouse over tooltip
+  )) +
+  scale_color_viridis(discrete = TRUE) +
   labs(
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age",
-    subtitle = "Quarterbacks"
+    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
   ) +
-  theme_classic()
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggplotly(
+  plot_rawFantasyPointsByAgeQB,
+  tooltip = c("age","fantasy_points","text"))
 ```
@@
 
-ggplot(
-  data = player_stats_seasonal_offense_subset %>% filter(position == "FB"),
+plot_rawFantasyPointsByAgeFB <- ggplot(
+  data = player_stats_seasonal_offense_subset %>% 
+    filter(position == "FB") %>% 
+    mutate(
+      age = round(age, 2),
+      fantasy_points = round(fantasy_points, 2)
+    ),
   mapping = aes(
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age",
-    subtitle = "Fullbacks"
+    title = "Fantasy Points (Season) by Player Age: Fullbacks"
   ) +
-  theme_classic()
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggplotly(
+  plot_rawFantasyPointsByAgeFB,
+  tooltip = c("age","fantasy_points","text"))
 ```
@@
 
-ggplot(
-  data = player_stats_seasonal_offense_subset %>% filter(position == "RB"),
+plot_rawFantasyPointsByAgeRB <- ggplot(
+  data = player_stats_seasonal_offense_subset %>% 
+    filter(position == "RB") %>% 
+    mutate(
+      age = round(age, 2),
+      fantasy_points = round(fantasy_points, 2)
+    ),
   mapping = aes(
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age",
-    subtitle = "Running Backs"
+    title = "Fantasy Points (Season) by Player Age: Running Backs"
   ) +
-  theme_classic()
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggplotly(
+  plot_rawFantasyPointsByAgeRB,
+  tooltip = c("age","fantasy_points","text"))
 ```
@@
 
-ggplot(
-  data = player_stats_seasonal_offense_subset %>% filter(position == "WR"),
+plot_rawFantasyPointsByAgeWR <- ggplot(
+  data = player_stats_seasonal_offense_subset %>% 
+    filter(position == "WR") %>% 
+    mutate(
+      age = round(age, 2),
+      fantasy_points = round(fantasy_points, 2)
+    ),
   mapping = aes(
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age",
-    subtitle = "Wide Receivers"
+    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
   ) +
-  theme_classic()
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggplotly(
+  plot_rawFantasyPointsByAgeWR,
+  tooltip = c("age","fantasy_points","text"))
 ```
@@
 
-ggplot(
-  data = player_stats_seasonal_offense_subset %>% filter(position == "TE"),
+plot_rawFantasyPointsByAgeTE <- ggplot(
+  data = player_stats_seasonal_offense_subset %>% 
+    filter(position == "TE") %>% 
+    mutate(
+      age = round(age, 2),
+      fantasy_points = round(fantasy_points, 2)
+    ),
   mapping = aes(
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age",
-    subtitle = "Tight Ends"
+    title = "Fantasy Points (Season) by Player Age: Tight Ends"
   ) +
-  theme_classic()
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggplotly(
+  plot_rawFantasyPointsByAgeTE,
+  tooltip = c("age","fantasy_points","text"))
 ```
@@
 
-ggplot(
+plot_individualFantasyPointsByAgeQB <- ggplot(
   data = player_stats_seasonal_offense_subsetCC %>% filter(position == "QB"),
@@
   geom_smooth(
+    aes(
+      text = player_display_name # add player name for mouse over tooltip
+      ),
     se = FALSE,
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age",
-    subtitle = "Quarterbacks"
+    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
   ) +
   theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsByAgeQB,
+  tooltip = c("age","fantasyPoints_gam","text"))
 ```
@@
 
-ggplot(
+plot_individualFantasyPointsByAgeFB <- ggplot(
   data = player_stats_seasonal_offense_subsetCC %>% filter(position == "FB"),
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age",
-    subtitle = "Fullbacks"
+    title = "Fantasy Points (Season) by Player Age: Fullbacks"
   ) +
   theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsByAgeFB,
+  tooltip = c("age","fantasyPoints_gam","text"))
 ```
@@
 
-ggplot(
+plot_individualFantasyPointsByAgeRB <- ggplot(
   data = player_stats_seasonal_offense_subsetCC %>% filter(position == "RB"),
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age",
-    subtitle = "Running Backs"
+    title = "Fantasy Points (Season) by Player Age: Running Backs"
   ) +
   theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsByAgeRB,
+  tooltip = c("age","fantasyPoints_gam","text"))
 ```
@@
 
-ggplot(
+plot_individualFantasyPointsByAgeWR <- ggplot(
   data = player_stats_seasonal_offense_subsetCC %>% filter(position == "WR"),
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age",
-    subtitle = "Wide Receivers"
+    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
   ) +
   theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsByAgeWR,
+  tooltip = c("age","fantasyPoints_gam","text"))
 ```
@@
 
-ggplot(
+plot_individualFantasyPointsByAgeTE <- ggplot(
   data = player_stats_seasonal_offense_subsetCC %>% filter(position == "TE"),
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age",
-    subtitle = "Tight Ends"
+    title = "Fantasy Points (Season) by Player Age: Tight Ends"
   ) +
   theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsByAgeTE,
+  tooltip = c("age","fantasyPoints_gam","text"))
 ```

### 2024-07-27 mixed-models.qmd
20240727 - peak ages and age of zero expected fantasy points

@@
 
-Plots of players' raw fantasy points data by age are in Figures [-@fig-rawFantasyPointsByAgeQB]–[-@fig-rawFantasyPointsByAgeTE].
+#### Quarterbacks {#sec-fantasyPointsByAgeExperienceRawDataQB}
+
+A plot of Quarterbacks' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeQB.
 
@@
 
+#### Fullbacks {#sec-fantasyPointsByAgeExperienceRawDataFB}
+
+A plot of Fullbacks' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeFB.
+
 ```{r}
@@
 ```
+
+#### Running Backs {#sec-fantasyPointsByAgeExperienceRawDataRB}
+
+A plot of Running Backs' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeRB.
 
@@
 
+#### Wide Receivers {#sec-fantasyPointsByAgeExperienceRawDataWR}
+
+A plot of Wide Receivers' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeWR.
+
 ```{r}
@@
 ```
+
+#### Tight Ends {#sec-fantasyPointsByAgeExperienceRawDataTE}
+
+A plot of Tight Ends' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeTE.
 
@@
 pointsPerSeason_positionAge_newData$ageCentered20Quadratic <- pointsPerSeason_positionAge_newData$ageCentered20 ^ 2
-pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$ageCentered20)
+pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$age - 22) # assuming that most players start at age 22 (i.e., rookie year) and thus have 1 year of experience at age 23
+pointsPerSeason_positionAge_newData$years_of_experience[which(pointsPerSeason_positionAge_newData$years_of_experience < 0)] <- 0
 
@@
 
+#### Quadratic Model {#sec-fantasyPointsByAgePlotsQuadratic}
+
 ```{r}
@@
 ```
+
+#### Quadratic Model: Top of Depth Chart {#sec-fantasyPointsByAgePlotsQuadraticDepth}
 
@@
 
+#### Generalized Additive Model {#sec-fantasyPointsByAgePlotsGAM}
+
 ```{r}
@@
 ```
+
+#### Generalized Additive Model: Top of Depth Chart {#sec-fantasyPointsByAgePlotsGAMdepth}
 
@@
 )
+
+zeroAge <- pointsPerSeason_positionAge_newData %>% 
+  group_by(positionFactor) %>% 
+  filter(fantasyPoints_gam < 0) %>% 
+  slice(which.min(age))
+
+peakAge <- pointsPerSeason_positionAge_newData %>% 
+  group_by(positionFactor) %>% 
+  slice(which.max(fantasyPoints_gam))
+
+peakAge2 <- pointsPerSeason_positionAge_newData %>% 
+  filter(age > 22) %>% 
+  group_by(positionFactor) %>% 
+  slice(which.max(fantasyPoints_gam))
+
+qbPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "QB")], 0)
+fbPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "FB")], 0)
+rbPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "RB")], 0)
+wrPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "WR")], 0)
+wrPeakAge2 <- round(peakAge2$age[which(peakAge$positionFactor == "WR")], 0)
+tePeakAge <- round(peakAge$age[which(peakAge$positionFactor == "TE")], 0)
+
+qbZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "QB")], 0)
+fbZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "FB")], 0)
+rbZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "RB")], 0)
+wrZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "WR")], 0)
+teZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "TE")], 0)
 ```
 
-Plots of players' model-implied fantasy points by age are in Figures [-@fig-individualFantasyPointsByAgeQB]–[-@fig-individualFantasyPointsByAgeTE].
+#### Quarterbacks {#sec-fantasyPointsIndividualByAgePlotsQB}
+
+A plot of Quarterbacks' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeQB.
+The model-implied peak of Quarterbacks' fantasy points is at age `{r} qbPeakAge`.
+The model-predicted value of zero fantasy points for Quarterbacks is at `{r} qbZeroAge`.
 
@@
 
+#### Fullbacks {#sec-fantasyPointsIndividualByAgePlotsFB}
+
+A plot of Fullbacks' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeFB.
+The model-implied peak of Fullbacks' fantasy points is at age `{r} fbPeakAge`.
+The model-predicted value of zero fantasy points for Quarterbacks is at `{r} fbZeroAge`.
+
 ```{r}
@@
 ```
+
+#### Running Backs {#sec-fantasyPointsIndividualByAgePlotsRB}
+
+A plot of Running Backs' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeRB.
+The model-implied peak of Running Backs' fantasy points is at age `{r} rbPeakAge`.
+The model-predicted value of zero fantasy points for Quarterbacks is at `{r} rbZeroAge`.
 
@@
 
+#### Wide Receivers {#sec-fantasyPointsIndividualByAgePlotsWR}
+
+A plot of Wide Receivers' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeWR.
+The model-implied peaks of Running Backs' fantasy points are at ages `{r} wrPeakAge` and `{r} wrPeakAge2`.
+The model-predicted value of zero fantasy points for Quarterbacks is at `{r} wrZeroAge`.
+
 ```{r}
@@
 ```
+
+#### Tight Ends {#sec-fantasyPointsIndividualByAgePlotsTE}
+
+A plot of Tight Ends' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeTE.
+The model-implied peak of Running Backs' fantasy points is at age `{r} tePeakAge`.
+The model-predicted value of zero fantasy points for Quarterbacks is at `{r} teZeroAge`.
 

### 2024-07-27 mixed-models.qmd
20240727 - organize headings

@@
 
+#### Null Model {#sec-fantasyPointsByAgeExperienceModelsNull}
+
 ```{r}
@@
 AIC(pointsPerSeason_nullModel)
+```
 
+#### Model with Position as Fixed-Effect Predictor {#sec-fantasyPointsByAgeExperienceModelsPosition}
+
+```{r}
 pointsPerSeason_position <- lmerTest::lmer(
@@
 AIC(pointsPerSeason_position)
+```
 
+#### Identify the Best-Fitting Functional Form of Age {#sec-fantasyPointsByAgeExperienceModelsAge}
+
+##### Linear Models {#sec-fantasyPointsByAgeExperienceModelsAgeLinear}
+
+```{r}
 pointsPerSeason_positionAgeFixedLinearSlopes <- lmerTest::lmer(
@@
 AIC(pointsPerSeason_positionAgeRandomLinearSlopes)
+```
 
+##### Quadratic Models {#sec-fantasyPointsByAgeExperienceModelsAgeQuadratic}
+
+```{r}
 pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes <- lmerTest::lmer(
@@
 
-Compare models:
+##### Compare Models {#sec-fantasyPointsByAgeExperienceModelsAgeCompare}
 
@@
 
-Generalized Additive Model:
+##### Generalized Additive Model {#sec-fantasyPointsByAgeExperienceModelsAgeGAM}
 
@@
 
-Players at the top of the end-of-season depth chart:
+##### Players Who Were (at Least Once) at the Top of the End-of-Season Depth Chart {#sec-fantasyPointsByAgeExperienceModelsAgeDepth}
 

### 2024-07-28 cognitive-bias.qmd
20240727 - ecological fallacy

@@
 
+### Ecological Fallacy {#sec-fallaciesEcological}
+
+The ecological fallacy is the error of drawing inferences about an individual from group-level data.
+For instance, a manager is interested in how age is associated with fantasy performance.
+They examine the correlation between age and fantasy points and find that age is positively correlated with fantasy points among Quarterbacks.
+The ecological fallacy would then take this finding to infer that an individual Quarterback is likely to increase in their performance with age.
+We describe the ecological fallacy more in @sec-ecologicalFallacy.
+
 ## Conclusion {#sec-cognitiveBiasConclusion}

### 2024-07-28 mixed-models.qmd
20240727 - ecological fallacy

@@
 
+We will discuss a modeling framework that goes by many terms, including mixed models, mixed-effects models, multilevel models, hierarchical linear models.
+
+### Ecological Fallacy {#sec-ecologicalFallacy}
+
+### Simpson's Paradox {#sec-simpsonsParadox}
+
 ## Fantasy Points Per Season by Position, Age, and Experience {#sec-fantasyPointsByAgeExperience}
@@
 
-### Plots of Raw Data {#sec-fantasyPointsByAgeExperienceRawData}
+### Scatterplots of Fantasy Points by Age and Position {#sec-fantasyPointsByAgeScatterplot}
+
+[Scatterplots](#sec-scatterplot) are a helpful tool for quickly examining the association between two variables.
+However, [scatterplots](#sec-scatterplot)—as well as correlation and multiple regression—can hide meaningful associations that differ across units of analysis.
+
+#### Quarterbacks {#sec-fantasyPointsByAgeScatterplotQB}
+
+A [scatterplot](#sec-scatterplot) of Quarterbacks' fantasy points by age is in @fig-fantasyPointsByAgeScatterplotQB.
+
+```{r}
+#| label: fig-fantasyPointsByAgeScatterplotQB
+#| fig-cap: "Scatterplot of Fantasy Points by Age for Quarterbacks."
+
+plot_scatterplotFantasyPointsByAgeQB <- ggplot(
+  data = player_stats_seasonal_offense_subset %>% 
+    filter(position == "QB") %>% 
+    mutate(
+      age = round(age, 2),
+      fantasy_points = round(fantasy_points, 2)
+    ),
+  mapping = aes(
+    x = age,
+    y = fantasy_points,
+    color = player_id)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+  )) +
+  geom_smooth(
+    mapping = aes(
+    x = age,
+    y = fantasy_points),
+    inherit.aes = FALSE
+  ) +
+  scale_color_viridis(discrete = TRUE) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
+  ) +
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggplotly(
+  plot_scatterplotFantasyPointsByAgeQB,
+  tooltip = c("age","fantasy_points","text","label"))
+```
+
+Based on the [scatterplot](#sec-scatterplot) (and the bivariate association below), Quarterbacks' fantasy points appear to increase with age.
+
+```{r}
+cor.test(
+  formula = ~ age + fantasy_points,
+  data = player_stats_seasonal_offense_subset %>% filter(position == "QB")
+)
+```
+
+#### Fullbacks {#sec-fantasyPointsByAgeScatterplotFB}
+
+A [scatterplot](#sec-scatterplot) of Fullbacks' fantasy points by age is in @fig-fantasyPointsByAgeScatterplotFB.
+
+```{r}
+#| label: fig-fantasyPointsByAgeScatterplotFB
+#| fig-cap: "Scatterplot of Fantasy Points by Age for Fullbacks."
+
+plot_scatterplotFantasyPointsByAgeFB <- ggplot(
+  data = player_stats_seasonal_offense_subset %>% 
+    filter(position == "FB") %>% 
+    mutate(
+      age = round(age, 2),
+      fantasy_points = round(fantasy_points, 2)
+    ),
+  mapping = aes(
+    x = age,
+    y = fantasy_points,
+    color = player_id)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+  )) +
+  geom_smooth(
+    mapping = aes(
+    x = age,
+    y = fantasy_points),
+    inherit.aes = FALSE
+  ) +
+  scale_color_viridis(discrete = TRUE) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age: Fullbacks"
+  ) +
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggplotly(
+  plot_scatterplotFantasyPointsByAgeFB,
+  tooltip = c("age","fantasy_points","text","label"))
+```
+
+Based on the [scatterplot](#sec-scatterplot), Fullbacks' fantasy points appear to be relatively stable across ages.
+
+#### Running Backs {#sec-fantasyPointsByAgeScatterplotRB}
+
+A [scatterplot](#sec-scatterplot) of Running Backs' fantasy points by age is in @fig-fantasyPointsByAgeScatterplotRB.
+
+```{r}
+#| label: fig-fantasyPointsByAgeScatterplotRB
+#| fig-cap: "Scatterplot of Fantasy Points by Age for Running Backs."
+
+plot_scatterplotFantasyPointsByAgeRB <- ggplot(
+  data = player_stats_seasonal_offense_subset %>% 
+    filter(position == "RB") %>% 
+    mutate(
+      age = round(age, 2),
+      fantasy_points = round(fantasy_points, 2)
+    ),
+  mapping = aes(
+    x = age,
+    y = fantasy_points,
+    color = player_id)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+  )) +
+  geom_smooth(
+    mapping = aes(
+    x = age,
+    y = fantasy_points),
+    inherit.aes = FALSE
+  ) +
+  scale_color_viridis(discrete = TRUE) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age: Running Backs"
+  ) +
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggplotly(
+  plot_scatterplotFantasyPointsByAgeRB,
+  tooltip = c("age","fantasy_points","text","label"))
+```
+
+Based on the [scatterplot](#sec-scatterplot), Running Backs' fantasy points appear to be relatively stable across ages.
+
+#### Wide Receivers {#sec-fantasyPointsByAgeScatterplotWR}
+
+A [scatterplot](#sec-scatterplot) of Wide Receivers' fantasy points by age is in @fig-fantasyPointsByAgeScatterplotWR.
+
+```{r}
+#| label: fig-fantasyPointsByAgeScatterplotWR
+#| fig-cap: "Scatterplot of Fantasy Points by Age for Wide Receivers."
+
+plot_scatterplotFantasyPointsByAgeWR <- ggplot(
+  data = player_stats_seasonal_offense_subset %>% 
+    filter(position == "WR") %>% 
+    mutate(
+      age = round(age, 2),
+      fantasy_points = round(fantasy_points, 2)
+    ),
+  mapping = aes(
+    x = age,
+    y = fantasy_points,
+    color = player_id)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+  )) +
+  geom_smooth(
+    mapping = aes(
+    x = age,
+    y = fantasy_points),
+    inherit.aes = FALSE
+  ) +
+  scale_color_viridis(discrete = TRUE) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
+  ) +
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggplotly(
+  plot_scatterplotFantasyPointsByAgeWR,
+  tooltip = c("age","fantasy_points","text","label"))
+```
+
+Based on the [scatterplot](#sec-scatterplot), Wide Receivers' fantasy points appear to be relatively stable across ages.
+
+#### Tight Ends {#sec-fantasyPointsByAgeScatterplotTE}
+
+A [scatterplot](#sec-scatterplot) of Tight Ends' fantasy points by age is in @fig-fantasyPointsByAgeScatterplotTE.
+
+```{r}
+#| label: fig-fantasyPointsByAgeScatterplotTE
+#| fig-cap: "Scatterplot of Fantasy Points by Age for Tight Ends."
+
+plot_scatterplotFantasyPointsByAgeTE <- ggplot(
+  data = player_stats_seasonal_offense_subset %>% 
+    filter(position == "TE") %>% 
+    mutate(
+      age = round(age, 2),
+      fantasy_points = round(fantasy_points, 2)
+    ),
+  mapping = aes(
+    x = age,
+    y = fantasy_points,
+    color = player_id)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+  )) +
+  geom_smooth(
+    mapping = aes(
+    x = age,
+    y = fantasy_points),
+    inherit.aes = FALSE
+  ) +
+  scale_color_viridis(discrete = TRUE) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age: Tight Ends"
+  ) +
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggplotly(
+  plot_scatterplotFantasyPointsByAgeTE,
+  tooltip = c("age","fantasy_points","text","label"))
+```
+
+Based on the [scatterplot](#sec-scatterplot) (and the bivariate association below), Tight Ends' fantasy points appear to increase with age.
+
+```{r}
+cor.test(
+  formula = ~ age + fantasy_points,
+  data = player_stats_seasonal_offense_subset %>% filter(position == "TE")
+)
+```
+
+### Plots of Raw Trajectories of Fantasy Points By Age and Player {#sec-fantasyPointsByAgeExperienceRawData}
+
+[Scatterplots](#sec-scatterplot) can be helpful for quickly visualizing the association between two variables.
+However, as mentioned earlier, [scatterplots](#sec-scatterplot) can hide the association between variables at different units of analysis.
+For instance, consider that we are trying to predict how a player will perform based on their age.
+We are interested not only in what the association is between age and fantasy points between players (i.e., a between-person association).
+We are also interested in what the association is between age and fantasy points *within* a given player (and within each player; i.e., a within-individual association).
+Arguably, the within-individual association between age and fantasy points is more relevant to the prediction of performance than the association between age and fantasy points between players.
+Assuming that the between-player association between age and fantasy points is the same as the within-player association when it is not is an example of the [ecological fallacy](#sec-fallaciesEcological).
+
+Below, we depict players' raw trajectories of fantasy points as a function of age.
+These are known as spaghetti plots.
+By examining the trajectory for each player, we can get a better understanding of hor performance changes (within an individual) as a function of age.
 
@@
 
-### Models {#sec-fantasyPointsByAgeExperienceModels}
+### Mixed Models {#sec-fantasyPointsByAgeExperienceModels}
+
+By accounting for which player each observation comes from using mixed models, we can examine the association between age and fantasy points in a more meaningful way, without violating the [assumption in multiple regression](#sec-assumptionsRegression) that the observations are independent (i.e., that the residuals are uncorrelated).
 
@@
 
+Mixed models allow accounting for multiple levels or units of analysis and to include both fixed and random effects.
+Inclusion of random effects allows effects of a predictor to vary as a function of individuals in the grouping level.
+This allows for more accurately predicting phenomena.
+We applied mixed models with random intercepts and random slopes to allow our model estimates to account for the fact that different players have different starting points (intercepts) and different changes over time (slopes) in fantasy points.
+A quadratic, inverted-U-shaped form as a function of age fit better than a linear form as a function of age in predicting players' fantasy points.
+A generalized additive model that allows further nonlinearity fit even better than the quadratic model.
+
+Based on the bivariate scatterplots between age and fantasy points, we might conclude that players tend to stay stable or even increase in fantasy points with age.
+However, we would be wrong.
+When we account for the longitudinal data (i.e., multiple observations over time for the same player) using mixed models, we observe that fantasy points tend to decrease with age, with the timing and rate of decline differing for each position.
+In other words, the association between age and fantasy points differs at the person level versus the group level.
+This is an example of Simpson's paradox.
+The discrepancy between the positive or null association between age and fantasy points at the group level, and the negative association at the person level may be due, in part, to the selective attrition of players with age.
+The highest performing players will tend to play the longest, whereas the poorest performing players will retire or get dropped from the team at the youngest ages.
+Thus, the selective attrition may make it seem that there is no association between age and performance (or even a positive one!), when in fact, players' performance tends to decrease with age after age 26 or so (with the timing differing from position to position).
+
+In sum, mixed models are valuable for examining associations between variables when there are multiple levels (i.e., clustering or nesting).
+It is important not to confuse the association at one level (e.g., group level) with the association at another level (e.g., person level), which is an example of the [ecological fallacy](#sec-fallaciesEcological).
+
 ::: {.content-visible when-format="html"}

### 2024-07-28 mixed-models.qmd
20240727 - fix labels

@@
 The model-implied peak of Fullbacks' fantasy points is at age `{r} fbPeakAge`.
-The model-predicted value of zero fantasy points for Quarterbacks is at `{r} fbZeroAge`.
+The model-predicted value of zero fantasy points for Fullbacks is at `{r} fbZeroAge`.
 
@@
 The model-implied peak of Running Backs' fantasy points is at age `{r} rbPeakAge`.
-The model-predicted value of zero fantasy points for Quarterbacks is at `{r} rbZeroAge`.
+The model-predicted value of zero fantasy points for Running Backs is at `{r} rbZeroAge`.
 
@@
 A plot of Wide Receivers' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeWR.
-The model-implied peaks of Running Backs' fantasy points are at ages `{r} wrPeakAge` and `{r} wrPeakAge2`.
-The model-predicted value of zero fantasy points for Quarterbacks is at `{r} wrZeroAge`.
+The model-implied peaks of Wide Receivers' fantasy points are at ages `{r} wrPeakAge` and `{r} wrPeakAge2`.
+The model-predicted value of zero fantasy points for Wide Receivers is at `{r} wrZeroAge`.
 
@@
 A plot of Tight Ends' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeTE.
-The model-implied peak of Running Backs' fantasy points is at age `{r} tePeakAge`.
-The model-predicted value of zero fantasy points for Quarterbacks is at `{r} teZeroAge`.
+The model-implied peak of Tight Ends' fantasy points is at age `{r} tePeakAge`.
+The model-predicted value of zero fantasy points for Tight Ends is at `{r} teZeroAge`.
 

### 2024-07-28 mixed-models.qmd
20240727 - fold code

@@
 player_stats_seasonal_offense_subset$positionFactor <- factor(player_stats_seasonal_offense_subset$position)
-
-#for(i in unique(nfl_depthCharts$season)){
-#  print(i)
-#  nfl_depthCharts %>%
-#    filter(season == i) %>% 
-#    select(week) %>% 
-#    arrange(week) %>% 
-#    unique() %>% 
-#    pull() %>% 
-#    print()
-#}
-
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+for(i in unique(nfl_depthCharts$season)){
+  print(i)
+  nfl_depthCharts %>%
+    filter(season == i) %>% 
+    select(week) %>% 
+    arrange(week) %>% 
+    unique() %>% 
+    pull() %>% 
+    print()
+}
+```
+
+```{r}
 seasons17week <- 2001:2020
@@
 #| fig-cap: "Scatterplot of Fantasy Points by Age for Quarterbacks."
+#| code-fold: true
 
@@
 #| fig-cap: "Scatterplot of Fantasy Points by Age for Fullbacks."
+#| code-fold: true
 
@@
 #| fig-cap: "Scatterplot of Fantasy Points by Age for Running Backs."
+#| code-fold: true
 
@@
 #| fig-cap: "Scatterplot of Fantasy Points by Age for Wide Receivers."
+#| code-fold: true
 
@@
 #| fig-cap: "Scatterplot of Fantasy Points by Age for Tight Ends."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Quarterbacks."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Fullbacks."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Running Backs."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Wide Receivers."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Tight Ends."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Model-Implied Quadratic Trajectories of Fantasy Points by Age."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Model-Implied Quadratic Trajectories of Fantasy Points by Age For Players Who Were Once at the Top of the End-of-Season Depth Chart."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age from a Generalized Additive Model."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, For Players Who Were Once at the Top of the End-of-Season Depth Chart."
+#| code-fold: true
 
@@
 )
+```
+
+```{r}
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Quarterbacks. Overlaid with the Model-Implied Trajectory for Quarterbacks."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Fullbacks. Overlaid with the Model-Implied Trajectory for Fullbacks."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Running Backs. Overlaid with the Model-Implied Trajectory for Running Backs."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Wide Receivers. Overlaid with the Model-Implied Trajectory for Wide Receivers."
+#| code-fold: true
 
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Tight Ends. Overlaid with the Model-Implied Trajectory for Wide Tight Ends."
+#| code-fold: true
 

### 2024-07-28 mixed-models.qmd
20240727 - selective attrition

@@
 This is an example of Simpson's paradox.
+
 The discrepancy between the positive or null association between age and fantasy points at the group level, and the negative association at the person level may be due, in part, to the selective attrition of players with age.
 The highest performing players will tend to play the longest, whereas the poorest performing players will retire or get dropped from the team at the youngest ages.
-Thus, the selective attrition may make it seem that there is no association between age and performance (or even a positive one!), when in fact, players' performance tends to decrease with age after age 26 or so (with the timing differing from position to position).
+Thus, the selective attrition may make it seem that there is no association between age and performance (or even a positive one!), when in fact, players' performance tends to decrease with age after age 26 or so (with the timing differing from position to position), until the player eventually retires or is dropped from the team.
+Selective attrition is common in longitudinal studies and intervention studies.
+For instance, attrition may be more likely for individuals from lower socioeconomic status backgrounds because they may face more challenges in continuing in longitudinal studies such as fewer financial resources, greater life stressors, etc.
+In addition, people who experience side effects or lack of improvement may be more likely to drop out of intervention studies.
+Examining only those who completed treatment (an example of [selection bias](#sec-causalDiagramSelectionBias)) would make the intervention look more effective than it actually was because the people who stay in the study are those who experience the greatest improvement.
+Thus, it is important to use approaches such as mixed models or other approaches that account for the multiple observations from the same person, that use all available information, and that do not exclude people who do not complete all portions of the study.
 

### 2024-07-28 mixed-models.qmd
20240727 - include season in plots

@@
     aes(
-      text = player_display_name # add player name for mouse over tooltip
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
   )) +
@@
   plot_rawFantasyPointsByAgeQB,
-  tooltip = c("age","fantasy_points","text"))
+  tooltip = c("age","fantasy_points","text","label"))
 ```
@@
   plot_rawFantasyPointsByAgeFB,
-  tooltip = c("age","fantasy_points","text"))
+  tooltip = c("age","fantasy_points","text","label"))
 ```
@@
   plot_rawFantasyPointsByAgeRB,
-  tooltip = c("age","fantasy_points","text"))
+  tooltip = c("age","fantasy_points","text","label"))
 ```
@@
   plot_rawFantasyPointsByAgeWR,
-  tooltip = c("age","fantasy_points","text"))
+  tooltip = c("age","fantasy_points","text","label"))
 ```
@@
   plot_rawFantasyPointsByAgeTE,
-  tooltip = c("age","fantasy_points","text"))
+  tooltip = c("age","fantasy_points","text","label"))
 ```
@@
     aes(
-      text = player_display_name # add player name for mouse over tooltip
+      x = age,
+      y = fantasyPoints_gam,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
     ),
@@
   ) +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_gam,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
   labs(
@@
   plot_individualFantasyPointsByAgeQB,
-  tooltip = c("age","fantasyPoints_gam","text"))
+  tooltip = c("age","fantasyPoints_gam","text","label")
+)
 ```
@@
   plot_individualFantasyPointsByAgeFB,
-  tooltip = c("age","fantasyPoints_gam","text"))
+  tooltip = c("age","fantasyPoints_gam","text","label")
+)
 ```
@@
   plot_individualFantasyPointsByAgeRB,
-  tooltip = c("age","fantasyPoints_gam","text"))
+  tooltip = c("age","fantasyPoints_gam","text","label")
+)
 ```
@@
   plot_individualFantasyPointsByAgeWR,
-  tooltip = c("age","fantasyPoints_gam","text"))
+  tooltip = c("age","fantasyPoints_gam","text","label")
+)
 ```
@@
   plot_individualFantasyPointsByAgeTE,
-  tooltip = c("age","fantasyPoints_gam","text"))
+  tooltip = c("age","fantasyPoints_gam","text","label")
+)
 ```

### 2024-07-28 mixed-models.qmd
20240728 - intervention studies

@@
 #| label: fig-individualFantasyPointsByAgeQB
-#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Quarterbacks. Overlaid with the Model-Implied Trajectory for Quarterbacks."
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Quarterbacks.
+Overlaid with the Model-Implied Trajectory for Quarterbacks in Blue."
 #| code-fold: true
@@
 #| label: fig-individualFantasyPointsByAgeFB
-#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Fullbacks. Overlaid with the Model-Implied Trajectory for Fullbacks."
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Fullbacks. Overlaid with the Model-Implied Trajectory for Fullbacks in Blue."
 #| code-fold: true
@@
 #| label: fig-individualFantasyPointsByAgeRB
-#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Running Backs. Overlaid with the Model-Implied Trajectory for Running Backs."
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Running Backs.
+Overlaid with the Model-Implied Trajectory for Running Backs in Blue."
 #| code-fold: true
@@
 #| label: fig-individualFantasyPointsByAgeWR
-#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Wide Receivers. Overlaid with the Model-Implied Trajectory for Wide Receivers."
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Wide Receivers.
+Overlaid with the Model-Implied Trajectory for Wide Receivers in Blue."
 #| code-fold: true
@@
 #| label: fig-individualFantasyPointsByAgeTE
-#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Tight Ends. Overlaid with the Model-Implied Trajectory for Wide Tight Ends."
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Tight Ends.
+Overlaid with the Model-Implied Trajectory for Wide Tight Ends in Blue."
 #| code-fold: true
@@
 
-## Conclusion {#sec-mixedModelsConclusion}
-
-Mixed models allow accounting for multiple levels or units of analysis and to include both fixed and random effects.
-Inclusion of random effects allows effects of a predictor to vary as a function of individuals in the grouping level.
-This allows for more accurately predicting phenomena.
+### Summary of Findings {#sec-fantasyPointsByAgeExperienceSummary}
+
 We applied mixed models with random intercepts and random slopes to allow our model estimates to account for the fact that different players have different starting points (intercepts) and different changes over time (slopes) in fantasy points.
 A quadratic, inverted-U-shaped form as a function of age fit better than a linear form as a function of age in predicting players' fantasy points.
-A generalized additive model that allows further nonlinearity fit even better than the quadratic model.
+A generalized additive model that allowed further nonlinearity fit even better than the quadratic model.
 
 Based on the bivariate scatterplots between age and fantasy points, we might conclude that players tend to stay stable or even increase in fantasy points with age.
-However, we would be wrong.
+However, this conclusion would be wrong.
 When we account for the longitudinal data (i.e., multiple observations over time for the same player) using mixed models, we observe that fantasy points tend to decrease with age, with the timing and rate of decline differing for each position.
 In other words, the association between age and fantasy points differs at the person level versus the group level.
-This is an example of Simpson's paradox.
-
-The discrepancy between the positive or null association between age and fantasy points at the group level, and the negative association at the person level may be due, in part, to the selective attrition of players with age.
-The highest performing players will tend to play the longest, whereas the poorest performing players will retire or get dropped from the team at the youngest ages.
-Thus, the selective attrition may make it seem that there is no association between age and performance (or even a positive one!), when in fact, players' performance tends to decrease with age after age 26 or so (with the timing differing from position to position), until the player eventually retires or is dropped from the team.
-Selective attrition is common in longitudinal studies and intervention studies.
+This is an example of [Simpson's paradox](#sec-simpsonsParadox).
+
+The discrepancy between the positive or null association between age and players' fantasy points at the group level, and the negative association at the person level may be due, in part, to the selective attrition of players with age.
+The players who play the longest will tend to be the highest performing players, whereas the poorest performing players will retire or get dropped from the team at younger ages.
+Thus, the selective attrition of weaker players may make it seem that there is no association between age and performance (or even a positive one!), when in fact, players' performance tends to decrease with age after age 26 or so (with the timing differing from position to position), until the player eventually retires or is dropped from the team.
+Selective attrition is common in [longitudinal studies](#sec-longitudinal) (such as this one) and [intervention studies](#sec-intervention).
 For instance, attrition may be more likely for individuals from lower socioeconomic status backgrounds because they may face more challenges in continuing in longitudinal studies such as fewer financial resources, greater life stressors, etc.
-In addition, people who experience side effects or lack of improvement may be more likely to drop out of intervention studies.
+In addition, people who experience side effects or lack of improvement may be more likely to drop out of [intervention studies](#sec-intervention).
 Examining only those who completed treatment (an example of [selection bias](#sec-causalDiagramSelectionBias)) would make the intervention look more effective than it actually was because the people who stay in the study are those who experience the greatest improvement.
@@
 
-In sum, mixed models are valuable for examining associations between variables when there are multiple levels (i.e., clustering or nesting).
+## Conclusion {#sec-mixedModelsConclusion}
+
+Mixed models allow accounting for multiple levels or units of analysis and to include both fixed and random effects.
+Inclusion of random effects allows the association between the predictor variables (the intercept and age) and the outcome variable (fantasy points) to differ for each individual in the grouping level (in this case, each player).
+This allows for more accurately predicting phenomena.
+Based on the bivariate scatterplots between age and fantasy points, we might conclude that players tend to stay stable or even increase in fantasy points with age.
+However, this conclusion would be wrong.
+When we account for the longitudinal data using mixed models, we observe that players' fantasy points tend to decrease with age, with the timing and rate of decline differing for each position.
+In other words, the association between age and fantasy points differs at the person level versus the group level, which is an example of [Simpson's paradox](#sec-simpsonsParadox).
+In sum, mixed models are valuable for examining associations between variables when there are multiple levels of data (i.e., multiple observations within the same unit, known as clustering or nesting).
 It is important not to confuse the association at one level (e.g., group level) with the association at another level (e.g., person level), which is an example of the [ecological fallacy](#sec-fallaciesEcological).

### 2024-07-28 research-methods.qmd
20240728 - intervention studies

@@
 The variable that we manipulate is the *independent variable*.
-By contast, the *dependent variable* is the variable that we evaluate to determine whether it was influenced by the manipulation (i.e., by the independent variable).Besides the independent and dependent variables, the researcher attempts to hold everything else constant through processes including standardization and random assignment.
+By contrast, the *dependent variable* is the variable that we evaluate to determine whether it was influenced by the manipulation (i.e., by the independent variable).Besides the independent and dependent variables, the researcher attempts to hold everything else constant through processes including standardization and random assignment.
 *Standardization* involves using the same procedures to assess each participant, so that scores can be fairly compared across participants (and groups).
 Random assignment involves randomly assigning participants to conditions of the independent variable, so the people in each condition are comparable and do not differ systematically.
+
+#### Intervention Study {#sec-intervention}
+
+An intervention study is a study that involves some modification (e.g., a treatment) with the intent to improve people's standing on the dependent variable (e.g., depression).
+Some intervention studies have a control group, whereas intervention studies do not.
+Inclusion of a control group is valuable; without a control group, you do not know whether any apparent gains in the treatment condition were due to the treatment per se versus just the mere passage of time, regression effects, or other things that were going on in the participants' lives.
+An intervention that includes random assignment (e.g., to the intervention or control group) is an experiment.
+A randomized controlled trial (RCT) is an example of an experiment because it is an intervention with random assignment.
 

### 2024-07-28 mixed-models.qmd
20240728 - compare model fit

@@
 library("AICcmodavg")
+library("bbmle")
 library("parallel")
@@
 
-#### Null Model {#sec-fantasyPointsByAgeExperienceModelsNull}
-
-```{r}
-pointsPerSeason_nullModel <- lmerTest::lmer(
+#### Linear Regression Model {#sec-fantasyPointsByAgeExperienceModelsRegression}
+
+```{r}
+pointsPerSeason_regressionModel <- lm(
+  fantasy_points ~ 1,
+  data = player_stats_seasonal_offense_subset,
+  na.action = "na.exclude"
+)
+
+summary(pointsPerSeason_regressionModel)
+summary(pointsPerSeason_regressionModel)$r.squared
+AIC(pointsPerSeason_regressionModel)
+MuMIn::AICc(pointsPerSeason_regressionModel)
+```
+
+#### Random Intercepts Model {#sec-fantasyPointsByAgeExperienceModelsIntercepts}
+
+```{r}
+pointsPerSeason_randomIntercepts <- lmerTest::lmer(
   fantasy_points ~ 1 + (1 | player_idFactor),
@@
 
-summary(pointsPerSeason_nullModel)
-MuMIn::r.squaredGLMM(pointsPerSeason_nullModel)
-performance::icc(pointsPerSeason_nullModel)
-AIC(pointsPerSeason_nullModel)
-```
-
-#### Model with Position as Fixed-Effect Predictor {#sec-fantasyPointsByAgeExperienceModelsPosition}
+summary(pointsPerSeason_randomIntercepts)
+MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
+performance::icc(pointsPerSeason_randomIntercepts)
+AIC(pointsPerSeason_randomIntercepts)
+AICc(pointsPerSeason_randomIntercepts)
+```
+
+#### Random Intercepts Model with Position as Fixed-Effect Predictor {#sec-fantasyPointsByAgeExperienceModelsPosition}
 
@@
 
+Groups of models:
+
+```{r}
+lmVsMixedModel <- list(
+  "linearRegression" = pointsPerSeason_regressionModel,
+  "randomIntercepts" = pointsPerSeason_randomIntercepts
+)
+
+lmAndmixedModels <- list(
+  "linearRegression" = pointsPerSeason_regressionModel,
+  "randomIntercepts" = pointsPerSeason_randomIntercepts,
+  "position" = pointsPerSeason_position,
+  "fixedLinear" = pointsPerSeason_positionAgeFixedLinearSlopes,
+  "randomLinear" = pointsPerSeason_positionAgeRandomLinearSlopes,
+  "randomLinearFixedQuadratic" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  "randomLinearFixedQuadraticInteraction" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
+)
+
+mixedModels <- list(
+  "randomIntercepts" = pointsPerSeason_randomIntercepts,
+  "position" = pointsPerSeason_position,
+  "fixedLinear" = pointsPerSeason_positionAgeFixedLinearSlopes,
+  "randomLinear" = pointsPerSeason_positionAgeRandomLinearSlopes,
+  "randomLinearFixedQuadratic" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  "randomLinearFixedQuadraticInteraction" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
+)
+
+mixedModels1 <- list(
+  "randomIntercepts" = pointsPerSeason_randomIntercepts,
+  "position" = pointsPerSeason_position
+)
+
+mixedModels2 <- list(
+  "fixedLinear" = pointsPerSeason_positionAgeFixedLinearSlopes,
+  "randomLinear" = pointsPerSeason_positionAgeRandomLinearSlopes,
+  "randomLinearFixedQuadratic" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  "randomLinearFixedQuadraticInteraction" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
+)
+```
+
+###### Likelihood Ratio Test {#sec-fantasyPointsByAgeExperienceModelsAgeCompareLRT}
+
 ```{r}
 anova(
-  pointsPerSeason_nullModel,
+  pointsPerSeason_randomIntercepts,
   pointsPerSeason_position
@@
 )
-
-mixedModels <- list(
-  "null" = pointsPerSeason_nullModel,
-  "position" = pointsPerSeason_position,
-  "fixedLinear" = pointsPerSeason_positionAgeFixedLinearSlopes,
-  "randomLinear" = pointsPerSeason_positionAgeRandomLinearSlopes,
-  "randomLinearFixedQuadratic" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
-  "randomLinearFixedQuadraticInteraction" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
-)
-
-aictab(mixedModels)
+```
+
+###### Akaike Information Criterion (AIC) {#sec-fantasyPointsByAgeExperienceModelsAgeCompareAIC}
+
+```{r}
+AIC(
+  pointsPerSeason_regressionModel,
+  pointsPerSeason_randomIntercepts,
+  pointsPerSeason_positionAgeFixedLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
+  )
+
+#AICcmodavg::aictab(lmAndmixedModels) # throws error (can't mix lm with lmer)
+bbmle::AICtab(lmAndmixedModels)
+
+AICcmodavg::aictab(mixedModels)
+bbmle::AICtab(mixedModels)
+```
+
+###### Corrected Akaike Information Criterion (AICc) {#sec-fantasyPointsByAgeExperienceModelsAgeCompareAICc}
+
+```{r}
+#AICcmodavg::aictab(lmVsMixedModel) # throws error (can't mix lm with lmer)
+bbmle::AICctab(lmVsMixedModel)
+
+AICcmodavg::aictab(mixedModels) # throws error (can't mix lm with lmer)
+#bbmle::AICctab(mixedModels) # throws error (different numbers of observations)
+
+AICcmodavg::aictab(mixedModels1)
+bbmle::AICctab(mixedModels1)
+
+AICcmodavg::aictab(mixedModels2)
+bbmle::AICctab(mixedModels2)
+```
+
+###### Bayesian Information Criterion (BIC) {#sec-fantasyPointsByAgeExperienceModelsAgeCompareBIC}
+
+```{r}
+BIC(
+  pointsPerSeason_regressionModel,
+  pointsPerSeason_randomIntercepts,
+  pointsPerSeason_positionAgeFixedLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
+  )
+```
+
+###### $R^2$ {#sec-fantasyPointsByAgeExperienceModelsAgeCompareRsquared}
+
+```{r}
+summary(pointsPerSeason_regressionModel)$r.squared
+MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+```
+
+###### Deviance {#sec-fantasyPointsByAgeExperienceModelsAgeCompareDeviance}
+
+```{r}
+deviance(pointsPerSeason_regressionModel)
+deviance(pointsPerSeason_randomIntercepts)
+deviance(pointsPerSeason_positionAgeFixedLinearSlopes)
+deviance(pointsPerSeason_positionAgeRandomLinearSlopes)
+deviance(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+deviance(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
 ```
@@
 AIC(pointsPerSeason_gam)
+```
+
+###### Compare Models
+
+```{r}
+linearMixedModelsVsGAM <- list(
+  "fixedLinear" = pointsPerSeason_positionAgeFixedLinearSlopes,
+  "randomLinear" = pointsPerSeason_positionAgeRandomLinearSlopes,
+  "randomLinearFixedQuadratic" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  "randomLinearFixedQuadraticInteraction" = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction,
+  "gam" = pointsPerSeason_gam
+)
+
+AIC(
+  pointsPerSeason_positionAgeFixedLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction,
+  pointsPerSeason_gam
+)
+
+BIC(
+  pointsPerSeason_positionAgeFixedLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction,
+  pointsPerSeason_gam
+)
+
+#AICcmodavg::aictab(linearMixedModelsVsGAM) # throws error (can't mix bam with lmer)
+bbmle::AICctab(linearMixedModelsVsGAM)
+
+summary(pointsPerSeason_regressionModel)$r.squared
+MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+MuMIn::r.squaredGLMM(pointsPerSeason_gam)
+
+deviance(pointsPerSeason_regressionModel)
+deviance(pointsPerSeason_randomIntercepts)
+deviance(pointsPerSeason_positionAgeFixedLinearSlopes)
+deviance(pointsPerSeason_positionAgeRandomLinearSlopes)
+deviance(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+deviance(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+deviance(pointsPerSeason_gam)
+
+logLik(pointsPerSeason_regressionModel)
+logLik(pointsPerSeason_randomIntercepts)
+logLik(pointsPerSeason_positionAgeFixedLinearSlopes)
+logLik(pointsPerSeason_positionAgeRandomLinearSlopes)
+logLik(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+logLik(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+logLik(pointsPerSeason_gam)
 ```

### 2024-07-28 mixed-models.qmd
20240728 - compare models

@@
 
-lmAndmixedModels <- list(
+lmAndMixedModels <- list(
   "linearRegression" = pointsPerSeason_regressionModel,
@@
 
-#AICcmodavg::aictab(lmAndmixedModels) # throws error (can't mix lm with lmer)
-bbmle::AICtab(lmAndmixedModels)
-
-AICcmodavg::aictab(mixedModels)
-bbmle::AICtab(mixedModels)
+bbmle::AICtab(lmAndMixedModels)
 ```
@@
   )
+
+#AICcmodavg::bictab(lmAndMixedModels) # throws error (can't mix lm with lmer)
+bbmle::BICtab(lmVsMixedModel)
+
+AICcmodavg::bictab(mixedModels) # throws error (can't mix lm with lmer)
+#bbmle::AICctab(mixedModels) # throws error (different numbers of observations)
+
+AICcmodavg::bictab(mixedModels1)
+bbmle::BICtab(mixedModels1)
+
+AICcmodavg::bictab(mixedModels2)
+bbmle::BICtab(mixedModels2)
 ```
@@
 
-###### Compare Models
+###### Compare Models {#sec-fantasyPointsByAgeExperienceModelsAgeGAMcompare}
 
@@
 
+#AICcmodavg::aictab(linearMixedModelsVsGAM) # throws error (can't mix bam with lmer)
+bbmle::AICctab(linearMixedModelsVsGAM)
+
 BIC(
@@
 
-#AICcmodavg::aictab(linearMixedModelsVsGAM) # throws error (can't mix bam with lmer)
-bbmle::AICctab(linearMixedModelsVsGAM)
+#AICcmodavg::bictab(linearMixedModelsVsGAM) # throws error (can't mix bam with lmer)
+bbmle::BICtab(linearMixedModelsVsGAM)
 

### 2024-07-28 mixed-models.qmd
20240728 - compare fit

@@
 
-Groups of models:
+After fitting several models, we now must compare their fit to determine which model fits "best" while also considering parsimony.
+Parsimonious models are more likely to be true and more likely to generalize to other samples, because more complex models are more likely to overfit the data.
+Thus, more complex models will almost always fit better than simpler models.
+Thus, we are not just interested in whether a more complex model fits better than the simpler model; we also care about whether the more complex model fits *significantly* better than the simpler model given its additional complexity.
+For evaluating and comparing models, we examine the [likelihood ratio test](#sec-fantasyPointsByAgeExperienceModelsAgeCompareLRT), the [Akaike Information Criterion](#sec-fantasyPointsByAgeExperienceModelsAgeCompareAIC) (AIC), the [corrected AIC](#sec-fantasyPointsByAgeExperienceModelsAgeCompareAICc) (AICc), the [Bayesian Information Criterion](#sec-fantasyPointsByAgeExperienceModelsAgeCompareBIC) (BIC), [$R^2$](#sec-fantasyPointsByAgeExperienceModelsAgeCompareRsquared), and [deviance](#sec-fantasyPointsByAgeExperienceModelsAgeCompareDeviance).
+
+The BIC penalizes model complexity more than the AIC does.
+The BIC is preferable when there is a "true" model, and one intends to identify the true model.
+The AIC is preferable when we are concerned more about predictive accuracy and when overfitting is less of a concern.
+Because we are more concerned about predictive accuracy and we do not believe one of these models is the "true" model per se of age-related changes in fantasy performance, we will give more weight to AIC than BIC.
+
+Below, we specify various groups of models for the model fit comparisons:
 

### 2024-07-28 mixed-models.qmd
20240728 - fit indices

@@
 
-### Mixed Models {#sec-fantasyPointsByAgeExperienceModels}
-
-By accounting for which player each observation comes from using mixed models, we can examine the association between age and fantasy points in a more meaningful way, without violating the [assumption in multiple regression](#sec-assumptionsRegression) that the observations are independent (i.e., that the residuals are uncorrelated).
-
-#### Linear Regression Model {#sec-fantasyPointsByAgeExperienceModelsRegression}
-
-```{r}
-pointsPerSeason_regressionModel <- lm(
+### Linear Regression Models {#sec-fantasyPointsByAgeExperienceModelsRegression}
+
+#### Null Model {#sec-fantasyPointsByAgeExperienceModelsRegressionNull}
+
+```{r}
+pointsPerSeason_nullModel <- lm(
   fantasy_points ~ 1,
@@
 
-summary(pointsPerSeason_regressionModel)
-summary(pointsPerSeason_regressionModel)$r.squared
-AIC(pointsPerSeason_regressionModel)
-MuMIn::AICc(pointsPerSeason_regressionModel)
-```
+summary(pointsPerSeason_nullModel)
+summary(pointsPerSeason_nullModel)$r.squared
+AIC(pointsPerSeason_nullModel)
+MuMIn::AICc(pointsPerSeason_nullModel)
+```
+
+#### Linear Model {#sec-fantasyPointsByAgeExperienceModelsRegressionLinear}
+
+```{r}
+pointsPerSeason_linearRegression <- lm(
+  fantasy_points ~ positionFactor + ageCentered20 + positionFactor:ageCentered20,
+  data = player_stats_seasonal_offense_subset,
+  na.action = "na.exclude"
+)
+
+summary(pointsPerSeason_linearRegression)
+summary(pointsPerSeason_linearRegression)$r.squared
+AIC(pointsPerSeason_linearRegression)
+MuMIn::AICc(pointsPerSeason_linearRegression)
+```
+
+#### Quadratic Model {#sec-fantasyPointsByAgeExperienceModelsRegressionLinear}
+
+```{r}
+pointsPerSeason_quadraticRegression <- lm(
+  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic,
+  data = player_stats_seasonal_offense_subset,
+  na.action = "na.exclude"
+)
+
+summary(pointsPerSeason_quadraticRegression)
+summary(pointsPerSeason_quadraticRegression)$r.squared
+AIC(pointsPerSeason_quadraticRegression)
+MuMIn::AICc(pointsPerSeason_quadraticRegression)
+```
+
+#### Compare Models {#sec-fantasyPointsByAgeExperienceModelsRegressionCompare}
+
+```{r}
+anova(
+  pointsPerSeason_linearRegression,
+  pointsPerSeason_quadraticRegression
+)
+
+AIC(
+  pointsPerSeason_nullModel,
+  pointsPerSeason_linearRegression,
+  pointsPerSeason_quadraticRegression
+  )
+
+lmModels <- list(
+  "nullModel" = pointsPerSeason_nullModel,
+  "linearRegression" = pointsPerSeason_linearRegression,
+  "quadraticRegression" = pointsPerSeason_quadraticRegression
+)
+
+bbmle::AICtab(lmModels)
+
+summary(pointsPerSeason_nullModel)$r.squared
+summary(pointsPerSeason_linearRegression)$r.squared
+summary(pointsPerSeason_quadraticRegression)$r.squared
+
+deviance(pointsPerSeason_nullModel)
+deviance(pointsPerSeason_linearRegression)
+deviance(pointsPerSeason_quadraticRegression)
+
+logLik(pointsPerSeason_nullModel)
+logLik(pointsPerSeason_linearRegression)
+logLik(pointsPerSeason_quadraticRegression)
+```
+
+### Mixed Models {#sec-fantasyPointsByAgeExperienceModels}
+
+By accounting for which player each observation comes from using mixed models, we can examine the association between age and fantasy points in a more meaningful way, without violating the [assumption in multiple regression](#sec-assumptionsRegression) that the observations are independent (i.e., that the residuals are uncorrelated).
 
@@
 Thus, we are not just interested in whether a more complex model fits better than the simpler model; we also care about whether the more complex model fits *significantly* better than the simpler model given its additional complexity.
-For evaluating and comparing models, we examine the [likelihood ratio test](#sec-fantasyPointsByAgeExperienceModelsAgeCompareLRT), the [Akaike Information Criterion](#sec-fantasyPointsByAgeExperienceModelsAgeCompareAIC) (AIC), the [corrected AIC](#sec-fantasyPointsByAgeExperienceModelsAgeCompareAICc) (AICc), the [Bayesian Information Criterion](#sec-fantasyPointsByAgeExperienceModelsAgeCompareBIC) (BIC), [$R^2$](#sec-fantasyPointsByAgeExperienceModelsAgeCompareRsquared), and [deviance](#sec-fantasyPointsByAgeExperienceModelsAgeCompareDeviance).
+For evaluating and comparing models, we examine the [likelihood ratio test](#sec-fantasyPointsByAgeExperienceModelsAgeCompareLRT), the [Akaike Information Criterion](#sec-fantasyPointsByAgeExperienceModelsAgeCompareAIC) (AIC), the [corrected AIC](#sec-fantasyPointsByAgeExperienceModelsAgeCompareAICc) (AICc), the [Bayesian Information Criterion](#sec-fantasyPointsByAgeExperienceModelsAgeCompareBIC) (BIC), [$R^2$](#sec-fantasyPointsByAgeExperienceModelsAgeCompareRsquared), [deviance](#sec-fantasyPointsByAgeExperienceModelsAgeCompareDeviance), and [log likelihood](#sec-fantasyPointsByAgeExperienceModelsAgeCompareLogLik).
 
@@
 lmVsMixedModel <- list(
-  "linearRegression" = pointsPerSeason_regressionModel,
+  "nullModel" = pointsPerSeason_nullModel,
   "randomIntercepts" = pointsPerSeason_randomIntercepts
@@
 lmAndMixedModels <- list(
-  "linearRegression" = pointsPerSeason_regressionModel,
+  "nullModel" = pointsPerSeason_nullModel,
+  "linearRegression" = pointsPerSeason_linearRegression,
+  "quadraticRegression" = pointsPerSeason_quadraticRegression,
   "randomIntercepts" = pointsPerSeason_randomIntercepts,
@@
 AIC(
-  pointsPerSeason_regressionModel,
+  pointsPerSeason_nullModel,
+  pointsPerSeason_linearRegression,
+  pointsPerSeason_quadraticRegression,
   pointsPerSeason_randomIntercepts,
@@
 BIC(
-  pointsPerSeason_regressionModel,
+  pointsPerSeason_nullModel,
+  pointsPerSeason_linearRegression,
+  pointsPerSeason_quadraticRegression,
   pointsPerSeason_randomIntercepts,
@@
 
-AICcmodavg::bictab(mixedModels) # throws error (can't mix lm with lmer)
+AICcmodavg::bictab(mixedModels)
 #bbmle::AICctab(mixedModels) # throws error (different numbers of observations)
@@
 ```{r}
-summary(pointsPerSeason_regressionModel)$r.squared
+summary(pointsPerSeason_nullModel)$r.squared
+summary(pointsPerSeason_linearRegression)$r.squared
+summary(pointsPerSeason_quadraticRegression)$r.squared
 MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
@@
 ```{r}
-deviance(pointsPerSeason_regressionModel)
+deviance(pointsPerSeason_nullModel)
+deviance(pointsPerSeason_linearRegression)
+deviance(pointsPerSeason_quadraticRegression)
 deviance(pointsPerSeason_randomIntercepts)
@@
 deviance(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+```
+
+###### Log Likelihood {#sec-fantasyPointsByAgeExperienceModelsAgeCompareLogLik}
+
+```{r}
+logLik(pointsPerSeason_nullModel)
+logLik(pointsPerSeason_linearRegression)
+logLik(pointsPerSeason_quadraticRegression)
+logLik(pointsPerSeason_randomIntercepts)
+logLik(pointsPerSeason_positionAgeFixedLinearSlopes)
+logLik(pointsPerSeason_positionAgeRandomLinearSlopes)
+logLik(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+logLik(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
 ```
@@
 
-summary(pointsPerSeason_regressionModel)$r.squared
+summary(pointsPerSeason_nullModel)$r.squared
+summary(pointsPerSeason_linearRegression)$r.squared
+summary(pointsPerSeason_quadraticRegression)$r.squared
 MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
@@
 
-deviance(pointsPerSeason_regressionModel)
+deviance(pointsPerSeason_nullModel)
+deviance(pointsPerSeason_linearRegression)
+deviance(pointsPerSeason_quadraticRegression)
 deviance(pointsPerSeason_randomIntercepts)
@@
 
-logLik(pointsPerSeason_regressionModel)
+logLik(pointsPerSeason_nullModel)
+logLik(pointsPerSeason_linearRegression)
+logLik(pointsPerSeason_quadraticRegression)
 logLik(pointsPerSeason_randomIntercepts)

### 2024-07-28 mixed-models.qmd
20240728 - plots

@@
 
+Create a `newdata` object for generating the plots of model-implied fantasy points by age and position:
+
+```{r}
+pointsPerSeason_positionAge_newData <- expand.grid(
+  positionFactor = factor(c("FB","QB","RB","TE","WR")), #,"K"
+  age = seq(from = 20, to = 40, length.out = 10000)
+)
+
+pointsPerSeason_positionAge_newData$ageCentered20 <- pointsPerSeason_positionAge_newData$age - 20
+pointsPerSeason_positionAge_newData$ageCentered20Quadratic <- pointsPerSeason_positionAge_newData$ageCentered20 ^ 2
+pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$age - 22) # assuming that most players start at age 22 (i.e., rookie year) and thus have 1 year of experience at age 23
+pointsPerSeason_positionAge_newData$years_of_experience[which(pointsPerSeason_positionAge_newData$years_of_experience < 0)] <- 0
+```
+
+Create an object with complete cases for generating the plots of individuals' -implied fantasy points by age and position:
+
+```{r}
+player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subset %>%
+  filter(!is.na(player_idFactor), !is.na(fantasy_points), !is.na(positionFactor), !is.na(ageCentered20), !is.na(ageCentered20Quadratic), !is.na(years_of_experience))
+```
+
 ### Scatterplots of Fantasy Points by Age and Position {#sec-fantasyPointsByAgeScatterplot}
@@
 
+```{r}
+#| label: fig-fantasyPointsNull
+#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Null Model."
+#| code-fold: true
+
+pointsPerSeason_positionAge_newData$fantasyPoints_nullModel <- predict(
+  object = pointsPerSeason_nullModel,
+  newdata = pointsPerSeason_positionAge_newData
+)
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_nullModel
+  )
+) + 
+  geom_line(linewidth = 2) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Null Model"
+  ) +
+  theme_classic()
+```
+
 #### Linear Model {#sec-fantasyPointsByAgeExperienceModelsRegressionLinear}
@@
 
+```{r}
+#| label: fig-fantasyPointsLinearRegression
+#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Linear Regression Model."
+#| code-fold: true
+
+pointsPerSeason_positionAge_newData$fantasyPoints_linearRegression <- predict(
+  object = pointsPerSeason_linearRegression,
+  newdata = pointsPerSeason_positionAge_newData
+)
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_linearRegression,
+    color = positionFactor
+  )
+) + 
+  geom_line(linewidth = 2) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Linear Regression Model"
+  ) +
+  theme_classic()
+```
+
 #### Quadratic Model {#sec-fantasyPointsByAgeExperienceModelsRegressionLinear}
@@
 MuMIn::AICc(pointsPerSeason_quadraticRegression)
+```
+
+```{r}
+#| label: fig-fantasyPointsQuadraticRegression
+#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Quadratic Regression Model."
+#| code-fold: true
+
+pointsPerSeason_positionAge_newData$fantasyPoints_quadraticRegression <- predict(
+  object = pointsPerSeason_quadraticRegression,
+  newdata = pointsPerSeason_positionAge_newData
+)
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_quadraticRegression,
+    color = positionFactor
+  )
+) + 
+  geom_line(linewidth = 2) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Quadratic Regression Model"
+  ) +
+  theme_classic()
 ```
@@
 
+```{r}
+#| label: fig-fantasyPointsLinearRegression
+#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Random Intercepts Mixed Model."
+#| code-fold: true
+
+pointsPerSeason_positionAge_newData$fantasyPoints_randomIntercepts <- predict(
+  object = pointsPerSeason_randomIntercepts,
+  newdata = pointsPerSeason_positionAge_newData,
+  re.form = NA
+)
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_randomIntercepts
+  )
+) + 
+  geom_line(linewidth = 2) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "Random Intercepts Model"
+  ) +
+  theme_classic()
+```
+
 #### Random Intercepts Model with Position as Fixed-Effect Predictor {#sec-fantasyPointsByAgeExperienceModelsPosition}
@@
 AIC(pointsPerSeason_position)
+```
+
+```{r}
+#| label: fig-fantasyPointsPosition
+#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Random Intercepts Mixed Model With Position as a Fixed-Effect Predictor."
+#| code-fold: true
+
+pointsPerSeason_positionAge_newData$fantasyPoints_position <- predict(
+  object = pointsPerSeason_position,
+  newdata = pointsPerSeason_positionAge_newData,
+  re.form = NA
+)
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_position,
+    color = positionFactor
+  )
+) + 
+  geom_line(linewidth = 2) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Random Intercepts Model With Position as Fixed-Effect Predictor"
+  ) +
+  theme_classic()
 ```
@@
 AIC(pointsPerSeason_positionAgeFixedLinearSlopes)
-
+```
+
+```{r}
+#| label: fig-fantasyPointsFixedLinearSlopes
+#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts and Fixed Linear Slopes."
+#| code-fold: true
+
+pointsPerSeason_positionAge_newData$fantasyPoints_fixedLinearSlopes <- predict(
+  object = pointsPerSeason_positionAgeFixedLinearSlopes,
+  newdata = pointsPerSeason_positionAge_newData,
+  re.form = NA
+)
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_fixedLinearSlopes,
+    color = positionFactor
+  )
+) + 
+  geom_line(linewidth = 2) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Mixed Model with Random Intercepts and Fixed Linear Slopes"
+  ) +
+  theme_classic()
+```
+
+```{r}
 pointsPerSeason_positionAgeRandomLinearSlopes <- lmerTest::lmer(
@@
 AIC(pointsPerSeason_positionAgeRandomLinearSlopes)
+```
+
+```{r}
+#| label: fig-fantasyPointsRandomLinearSlopes
+#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts and Random Linear Slopes."
+#| code-fold: true
+
+pointsPerSeason_positionAge_newData$fantasyPoints_randomLinearSlopes <- predict(
+  object = pointsPerSeason_positionAgeRandomLinearSlopes,
+  newdata = pointsPerSeason_positionAge_newData,
+  re.form = NA
+)
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_randomLinearSlopes,
+    color = positionFactor
+  )
+) + 
+  geom_line(linewidth = 2) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Mixed Model with Random Intercepts and Random Linear Slopes"
+  ) +
+  theme_classic()
 ```
@@
 AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
-
+```
+
+```{r}
+#| label: fig-fantasyPointsRandomLinearFixedQuadracticSlopes
+#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes."
+#| code-fold: true
+
+pointsPerSeason_positionAge_newData$fantasyPoints_randomLinearFixedQuadraticSlopes <- predict(
+  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  newdata = pointsPerSeason_positionAge_newData,
+  re.form = NA
+)
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_randomLinearFixedQuadraticSlopes,
+    color = positionFactor
+  )
+) + 
+  geom_line(linewidth = 2) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Mixed Model with Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes"
+  ) +
+  theme_classic()
+```
+
+```{r}
 pointsPerSeason_positionAgeRandomLinearRandomQuadraticSlopes <- lmerTest::lmer(
@@
 )
-
+```
+
+```{r}
 pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction <- lmerTest::lmer(
@@
 AIC(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
-
+```
+
+```{r}
+#| label: fig-fantasyPointsRandomLinearFixedQuadraticSlopesInteraction
+#| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position."
+#| code-fold: true
+
+pointsPerSeason_positionAge_newData$fantasyPoints_randomLinearFixedQuadraticSlopesInteraction <- predict(
+  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction,
+  newdata = pointsPerSeason_positionAge_newData,
+  re.form = NA
+)
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_randomLinearFixedQuadraticSlopesInteraction,
+    color = positionFactor
+  )
+) + 
+  geom_line(linewidth = 2) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Mixed Model with Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position"
+  ) +
+  theme_classic()
+```
+
+```{r}
 pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
@@
 ```{r}
-# Create newdata object
-pointsPerSeason_positionAge_newData <- expand.grid(
-  positionFactor = factor(c("FB","QB","RB","TE","WR")), #,"K"
-  age = seq(from = 20, to = 40, length.out = 10000)
-)
-
-pointsPerSeason_positionAge_newData$ageCentered20 <- pointsPerSeason_positionAge_newData$age - 20
-pointsPerSeason_positionAge_newData$ageCentered20Quadratic <- pointsPerSeason_positionAge_newData$ageCentered20 ^ 2
-pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$age - 22) # assuming that most players start at age 22 (i.e., rookie year) and thus have 1 year of experience at age 23
-pointsPerSeason_positionAge_newData$years_of_experience[which(pointsPerSeason_positionAge_newData$years_of_experience < 0)] <- 0
-
 # From Quadratic Model: All Players
@@
 ```{r}
-player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subset %>%
-  filter(!is.na(player_idFactor), !is.na(fantasy_points), !is.na(positionFactor), !is.na(ageCentered20), !is.na(ageCentered20Quadratic), !is.na(years_of_experience))
-
 player_stats_seasonal_offense_subsetCC$fantasyPoints_quadratic <- predict(

### 2024-07-28 mixed-models.qmd
20240728 - fix duplicate chunk label

@@
 ```{r}
-#| label: fig-fantasyPointsLinearRegression
+#| label: fig-fantasyPointsRandomIntercepts
 #| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Random Intercepts Mixed Model."

### 2024-07-29 mixed-models.qmd
20240728 - individuals' model-implied trajectories

@@
 ```{r}
-library("petersenlab")
 library("lme4")
@@
 player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subset %>%
-  filter(!is.na(player_idFactor), !is.na(fantasy_points), !is.na(positionFactor), !is.na(ageCentered20), !is.na(ageCentered20Quadratic), !is.na(years_of_experience))
+  filter(
+    !is.na(player_idFactor),
+    !is.na(fantasy_points),
+    !is.na(positionFactor),
+    !is.na(ageCentered20),
+    !is.na(ageCentered20Quadratic),
+    !is.na(years_of_experience))
 ```
@@
 
+A plot of the model-implied trajectories of fantasy points by age from the null model is in @fig-fantasyPointsNull.
+
 ```{r}
@@
 
+A plot of the model-implied trajectories of fantasy points by age from the linear regression model is in @fig-fantasyPointsLinearRegression.
+
 ```{r}
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Linear Regression Model"
+    subtitle = "Linear Regression Model",
+    color = "Position"
   ) +
@@
 
+A plot of the model-implied trajectories of fantasy points by age from the regression model with a quadratic term for age is in @fig-fantasyPointsQuadraticRegression.
+
 ```{r}
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Quadratic Regression Model"
+    subtitle = "Quadratic Regression Model",
+    color = "Position"
   ) +
@@
 
+A plot of the model-implied trajectories of fantasy points by age from the mixed model with random intercepts is in @fig-fantasyPointsRandomIntercepts.
+
 ```{r}
@@
   theme_classic()
+```
+
+A plot of individuals' model-implied trajectories of fantasy points by age from the mixed model with random intercepts is in @fig-individualFantasyPointsRandomIntercepts.
+
+```{r}
+#| label: fig-individualFantasyPointsRandomIntercepts
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Mixed Model with Random Intercepts. Overlaid with the Model-Implied Trajectory."
+#| code-fold: true
+
+player_stats_seasonal_offense_subsetCC$fantasyPoints_randomIntercepts <- predict(
+  object = pointsPerSeason_randomIntercepts,
+  newdata = player_stats_seasonal_offense_subsetCC
+)
+
+plot_individualFantasyPointsRandomIntercepts <- ggplot(
+  data = player_stats_seasonal_offense_subsetCC,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_randomIntercepts,
+    #color = positionFactor,
+    group = player_id)) +
+  geom_smooth(
+    aes(
+      x = age,
+      y = fantasyPoints_randomIntercepts,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    se = FALSE,
+    linewidth = 0.5,
+    color = "black") +
+  geom_line(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_randomIntercepts
+    ),
+    data = pointsPerSeason_positionAge_newData,
+    inherit.aes = FALSE,
+    se = TRUE,
+    color = "#3366FF",
+    linewidth = 2
+  ) +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_randomIntercepts,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position: Random Intercepts Model",
+    #color = "Position"
+  ) +
+  theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsRandomIntercepts,
+  tooltip = c("age","fantasyPoints_randomIntercepts","text","label")
+)
 ```
@@
 
+A plot of the model-implied trajectories of fantasy points by age and position from the mixed model with random intercepts and a fixed effect of position is in @fig-fantasyPointsPosition.
+
 ```{r}
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Random Intercepts Model With Position as Fixed-Effect Predictor"
+    subtitle = "Random Intercepts Model With Position as Fixed-Effect Predictor",
+    color = "Position"
   ) +
@@
 
+A plot of individuals' model-implied trajectories of fantasy points by age and position from the mixed model with random intercepts and a fixed effect of position is in @fig-individualFantasyPointsPosition.
+
+```{r}
+#| label: fig-individualFantasyPointsPosition
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Mixed Model With Random Intercepts and a Fixed-Effect of Position.
+Overlaid with the Model-Implied Trajectory by Position."
+#| code-fold: true
+
+player_stats_seasonal_offense_subsetCC$fantasyPoints_position <- predict(
+  object = pointsPerSeason_position,
+  newdata = player_stats_seasonal_offense_subsetCC
+)
+
+plot_individualFantasyPointsPosition <- ggplot(
+  data = player_stats_seasonal_offense_subsetCC,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_position,
+    color = positionFactor,
+    group = player_id)) +
+  geom_smooth(
+    aes(
+      x = age,
+      y = fantasyPoints_position,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    se = FALSE,
+    linewidth = 0.5) +
+  geom_line(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_position,
+      color = positionFactor
+    ),
+    data = pointsPerSeason_positionAge_newData,
+    inherit.aes = FALSE,
+    se = TRUE,
+    linewidth = 2
+  ) +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_position,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position: Random Intercepts Model With Position As Predictor",
+    color = "Position"
+  ) +
+  theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsRandomIntercepts,
+  tooltip = c("age","fantasyPoints_position","text","label")
+)
+```
+
 #### Identify the Best-Fitting Functional Form of Age {#sec-fantasyPointsByAgeExperienceModelsAge}
@@
 ##### Linear Models {#sec-fantasyPointsByAgeExperienceModelsAgeLinear}
+
+###### Random Intercepts, Fixed Linear Slopes {#sec-fantasyPointsByAgeExperienceModelsAgeLinearFixed}
 
@@
 
+A plot of the model-implied trajectories of fantasy points by age and position from the mixed model with random intercepts and fixed linear slopes is in @fig-fantasyPointsFixedLinearSlopes.
+
 ```{r}
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Mixed Model with Random Intercepts and Fixed Linear Slopes"
+    subtitle = "Mixed Model with Random Intercepts and Fixed Linear Slopes",
+    color = "Position"
   ) +
@@
 ```
+
+A plot of individuals model-implied trajectories of fantasy points by age and position from the mixed model with random intercepts and fixed linear slopes is in @fig-individualFantasyPointsFixedLinearSlopes.
+
+```{r}
+#| label: fig-individualFantasyPointsFixedLinearSlopes
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts and Fixed Slopes.
+Overlaid with the Model-Implied Trajectory by Position."
+#| code-fold: true
+
+player_stats_seasonal_offense_subsetCC$fantasyPoints_fixedLinearSlopes <- predict(
+  object = pointsPerSeason_positionAgeFixedLinearSlopes,
+  newdata = player_stats_seasonal_offense_subsetCC
+)
+
+plot_individualFantasyPointsFixedLinearSlopes <- ggplot(
+  data = player_stats_seasonal_offense_subsetCC,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_fixedLinearSlopes,
+    color = positionFactor,
+    group = player_id)) +
+  geom_smooth(
+    aes(
+      x = age,
+      y = fantasyPoints_fixedLinearSlopes,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    se = FALSE,
+    linewidth = 0.5) +
+  geom_line(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_fixedLinearSlopes,
+      color = positionFactor
+    ),
+    data = pointsPerSeason_positionAge_newData,
+    inherit.aes = FALSE,
+    se = TRUE,
+    linewidth = 2
+  ) +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_fixedLinearSlopes,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position: Model With Random Intercepts and Fixed Slopes",
+    color = "Position"
+  ) +
+  theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsFixedLinearSlopes,
+  tooltip = c("age","fantasyPoints_fixedLinearSlopes","text","label")
+)
+```
+
+###### Random Intercepts, Random Linear Slopes {#sec-fantasyPointsByAgeExperienceModelsAgeLinearRandom}
 
@@
 
+A plot of the model-implied trajectories of fantasy points by age and position from the mixed model with random intercepts and random linear slopes is in @fig-fantasyPointsRandomLinearSlopes.
+
 ```{r}
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Mixed Model with Random Intercepts and Random Linear Slopes"
+    subtitle = "Mixed Model with Random Intercepts and Random Linear Slopes",
+    color = "Position"
   ) +
@@
 
+A plot of individuals' model-implied trajectories of fantasy points by age and position from the mixed model with random intercepts and random linear slopes is in @fig-fantasyPointsRandomLinearSlopes.
+
+```{r}
+#| label: fig-individualFantasyPointsRandomLinearSlopes
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts and Random Linear Slopes.
+Overlaid with the Model-Implied Trajectory by Position."
+#| code-fold: true
+
+player_stats_seasonal_offense_subsetCC$fantasyPoints_randomLinearSlopes <- predict(
+  object = pointsPerSeason_positionAgeRandomLinearSlopes,
+  newdata = player_stats_seasonal_offense_subsetCC
+)
+
+plot_individualFantasyPointsRandomLinearSlopes <- ggplot(
+  data = player_stats_seasonal_offense_subsetCC,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_randomLinearSlopes,
+    color = positionFactor,
+    group = player_id)) +
+  geom_smooth(
+    aes(
+      x = age,
+      y = fantasyPoints_randomLinearSlopes,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    se = FALSE,
+    linewidth = 0.5) +
+  geom_line(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_randomLinearSlopes,
+      color = positionFactor
+    ),
+    data = pointsPerSeason_positionAge_newData,
+    inherit.aes = FALSE,
+    se = TRUE,
+    linewidth = 2
+  ) +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_randomLinearSlopes,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position: Model With Random Intercepts and Random Linear Slopes",
+    color = "Position"
+  ) +
+  theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsRandomLinearSlopes,
+  tooltip = c("age","fantasyPoints_randomLinearSlopes","text","label")
+)
+```
+
 ##### Quadratic Models {#sec-fantasyPointsByAgeExperienceModelsAgeQuadratic}
+
+###### Random Intercepts, Random Linear Slopes, Fixed Quadratic Slopes {#sec-fantasyPointsByAgeExperienceModelsAgeQuadraticFixed}
 
@@
 
+A plot of the model-implied trajectories of fantasy points by age and position from the mixed model with random intercepts, random linear slopes, and fixed quadratic slopes is in @fig-fantasyPointsRandomLinearFixedQuadracticSlopes.
+
 ```{r}
@@
 ```
+
+A plot of individuals' model-implied trajectories of fantasy points by age and position from the mixed model with random intercepts, random linear slopes, and fixed quadratic slopes is in @fig-individualFantasyPointsRandomLinearFixedQuadracticSlopes.
+
+```{r}
+#| label: fig-individualFantasyPointsRandomLinearFixedQuadracticSlopes
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes.
+Overlaid with the Model-Implied Trajectory by Position."
+#| code-fold: true
+
+player_stats_seasonal_offense_subsetCC$fantasyPoints_randomLinearFixedQuadraticSlopes <- predict(
+  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  newdata = player_stats_seasonal_offense_subsetCC
+)
+
+plot_individualFantasyPointsRandomLinearFixedQuadracticSlopes <- ggplot(
+  data = player_stats_seasonal_offense_subsetCC,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_randomLinearFixedQuadraticSlopes,
+    color = positionFactor,
+    group = player_id)) +
+  geom_smooth(
+    aes(
+      x = age,
+      y = fantasyPoints_randomLinearFixedQuadraticSlopes,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    se = FALSE,
+    linewidth = 0.5) +
+  geom_line(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_randomLinearFixedQuadraticSlopes,
+      color = positionFactor
+    ),
+    data = pointsPerSeason_positionAge_newData,
+    inherit.aes = FALSE,
+    se = TRUE,
+    linewidth = 2
+  ) +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_randomLinearFixedQuadraticSlopes,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position: Model With Random Intercepts and Random Linear Slopes",
+    color = "Position"
+  ) +
+  theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsRandomLinearFixedQuadracticSlopes,
+  tooltip = c("age","fantasyPoints_randomLinearFixedQuadraticSlopes","text","label")
+)
+```
+
+###### Random Intercepts, Random Linear Slopes, Random Quadratic Slopes {#sec-fantasyPointsByAgeExperienceModelsAgeQuadraticRandom}
 
@@
 ```
+
+###### Random Intercepts, Random Linear Slopes, Fixed Quadratic Slopes in Interaction With Position {#sec-fantasyPointsByAgeExperienceModelsAgeQuadraticFixedInteraction}
 
@@
 
+A plot of the model-implied trajectories of fantasy points by age and position from the mixed model with random intercepts, random linear slopes, and fixed quadratic slopes in interaction with position is in @fig-fantasyPointsRandomLinearFixedQuadraticSlopesInteraction.
+
 ```{r}
@@
 ```
+
+A plot of individuals' model-implied trajectories of fantasy points by age and position from the mixed model with random intercepts, random linear slopes, and fixed quadratic slopes in interaction with position is in @fig-individualFantasyPointsRandomLinearFixedQuadraticSlopesInteraction.
+
+```{r}
+#| label: fig-individualFantasyPointsRandomLinearFixedQuadraticSlopesInteraction
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position.
+Overlaid with the Model-Implied Trajectory by Position."
+#| code-fold: true
+
+player_stats_seasonal_offense_subsetCC$fantasyPoints_randomLinearFixedQuadraticSlopesInteraction <- predict(
+  object = pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction,
+  newdata = player_stats_seasonal_offense_subsetCC
+)
+
+plot_individualFantasyPointsRandomLinearFixedQuadraticSlopesInteraction <- ggplot(
+  data = player_stats_seasonal_offense_subsetCC,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_randomLinearFixedQuadraticSlopesInteraction,
+    color = positionFactor,
+    group = player_id)) +
+  geom_smooth(
+    aes(
+      x = age,
+      y = fantasyPoints_randomLinearFixedQuadraticSlopesInteraction,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    se = FALSE,
+    linewidth = 0.5) +
+  geom_line(
+    mapping = aes(
+      x = age,
+      y = fantasyPoints_randomLinearFixedQuadraticSlopesInteraction,
+      color = positionFactor
+    ),
+    data = pointsPerSeason_positionAge_newData,
+    inherit.aes = FALSE,
+    se = TRUE,
+    linewidth = 2
+  ) +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_randomLinearFixedQuadraticSlopesInteraction,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position:\nModel With Random Intercepts, Random Linear Slopes, and\nFixed Quadratic Slopes in Interaction With Position",
+    color = "Position"
+  ) +
+  theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsRandomLinearFixedQuadraticSlopesInteraction,
+  tooltip = c("age","fantasyPoints_randomLinearFixedQuadraticSlopesInteraction","text","label")
+)
+```
+
+###### Adding Fixed-Effect Predictor of Experience {#sec-fantasyPointsByAgeExperienceModelsAgeQuadraticFixedInteractionExperience}
 

### 2024-07-29 data-visualization.qmd
20240724 - duplicate section labels
20240728 - data viz edits

@@
 
+As a tutorial, we walk through some of the (many) modifications that can be made to create an advanced, customized plot in `ggplot2`.
+
 First, we prepare the data:
@@
 
-Second, we create the base plot using the `ggplot()` function of the `ggplot2` package, as in @fig-scatterplot1.
+### Base Layer {#sec-scatterplotBaseLayer}
+
+Second, we create the base layer of the plot using the `ggplot()` function of the `ggplot2` package, as in @fig-scatterplot1.
 We specify the data object and the variables in the data object that are associated with the x- and y-axes:
@@
 ```
+
+### Add Points {#sec-scatterplotAddPoints}
 
@@
 ```
+
+### Best-Fit Line {#sec-scatterplotBestFitLine}
 
@@
 
+### Modify Axes {#sec-scatterplotAxes}
+
 Then, we can change the axes, as in @fig-scatterplot7:
@@
 ```
+
+### Plot Labels {#sec-scatterplotPlotLabels}
 
@@
     x = "Running Back's Age (years)",
-    y = "Rushing Yards Per Season",
+    y = "Rushing Yards (Season)",
     title = "NFL Rushing Yards (Season) by Player Age",
@@
 ```
+
+### Theme {#sec-scatterplotTheme}
 
@@
 
+### Interactive {#sec-scatterplotInteractive}
+
 ::: {.content-visible when-format="html:js"}
 
-After creating our plot, we can make the plot interactive using the `ggplotly()` function from the plotly package, as in @fig-scatterplot11.
+After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package, as in @fig-scatterplot11.
 
@@
 
-After creating our plot, we can make the plot interactive using the `ggplotly()` function from the plotly package.
+After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package.
 
@@
 
-## Examples {#sec-scatterplot}
+## Examples {#sec-scatterplotExamples}
 
@@
 #| label: fig-rushEPAperSeasonByAgePDF
-#| fig-cap: "NFL Rushing Expected Points Added (EPA) per Season by Player Age."
+#| fig-cap: "NFL Rushing Expected Points Added (EPA) During a Season by Player Age."
 

### 2024-07-29 mixed-models.qmd
20240728 - fixes

@@
 
-#### Quadratic Model {#sec-fantasyPointsByAgeExperienceModelsRegressionLinear}
+#### Quadratic Model {#sec-fantasyPointsByAgeExperienceModelsRegressionQuadratic}
 
@@
 plot_individualFantasyPointsRandomIntercepts <- ggplot(
-  data = player_stats_seasonal_offense_subsetCC,
+  data = player_stats_seasonal_offense_subsetCC %>% 
+    mutate(
+      age = round(age, 2),
+      fantasyPoints_randomIntercepts = round(fantasyPoints_randomIntercepts, 2)
+    ),
   mapping = aes(
@@
     y = fantasyPoints_randomIntercepts,
-    #color = positionFactor,
     group = player_id)) +
-  geom_smooth(
+  geom_line(
     aes(
@@
     ),
-    se = FALSE,
     linewidth = 0.5,
@@
     ),
-    data = pointsPerSeason_positionAge_newData,
+    data = pointsPerSeason_positionAge_newData %>% 
+      mutate(
+        age = round(age, 2),
+        fantasy_points = round(fantasyPoints_randomIntercepts, 2)
+        ),
     inherit.aes = FALSE,
@@
     linewidth = 2
-  ) +
-  geom_point(
-    aes(
-      x = age,
-      y = fantasyPoints_randomIntercepts,
-      text = player_display_name, # add player name for mouse over tooltip
-      label = season # add season for mouse over tooltip
-    ),
-    size = 1,
-    color = "transparent" # make points invisible but keep tooltips
   ) +
@@
 plot_individualFantasyPointsPosition <- ggplot(
-  data = player_stats_seasonal_offense_subsetCC,
+  data = player_stats_seasonal_offense_subsetCC %>% 
+    mutate(
+      age = round(age, 2),
+      fantasyPoints_position = round(fantasyPoints_position, 2)
+    ),
   mapping = aes(
@@
     group = player_id)) +
-  geom_smooth(
+  geom_line(
     aes(
@@
     ),
-    se = FALSE,
     linewidth = 0.5) +
@@
     ),
-    data = pointsPerSeason_positionAge_newData,
+    data = pointsPerSeason_positionAge_newData %>% 
+      mutate(
+        age = round(age, 2),
+        fantasyPoints_position = round(fantasyPoints_position, 2)
+        ),
     inherit.aes = FALSE,
@@
   ) +
-  geom_point(
-    aes(
-      x = age,
-      y = fantasyPoints_position,
-      text = player_display_name, # add player name for mouse over tooltip
-      label = season # add season for mouse over tooltip
-    ),
-    size = 1,
-    color = "transparent" # make points invisible but keep tooltips
-  ) +
   labs(
@@
 ggplotly(
-  plot_individualFantasyPointsRandomIntercepts,
+  plot_individualFantasyPointsPosition,
   tooltip = c("age","fantasyPoints_position","text","label")
@@
 plot_individualFantasyPointsFixedLinearSlopes <- ggplot(
-  data = player_stats_seasonal_offense_subsetCC,
+  data = player_stats_seasonal_offense_subsetCC %>% 
+    mutate(
+      age = round(age, 2),
+      fantasyPoints_fixedLinearSlopes = round(fantasyPoints_fixedLinearSlopes, 2)
+    ),
   mapping = aes(
@@
     linewidth = 2
-  ) +
-  geom_point(
-    aes(
-      x = age,
-      y = fantasyPoints_fixedLinearSlopes,
-      text = player_display_name, # add player name for mouse over tooltip
-      label = season # add season for mouse over tooltip
-    ),
-    size = 1,
-    color = "transparent" # make points invisible but keep tooltips
   ) +
@@
 plot_individualFantasyPointsRandomLinearSlopes <- ggplot(
-  data = player_stats_seasonal_offense_subsetCC,
+  data = player_stats_seasonal_offense_subsetCC %>% 
+    mutate(
+      age = round(age, 2),
+      fantasyPoints_randomLinearSlopes = round(fantasyPoints_randomLinearSlopes, 2)
+    ),
   mapping = aes(
@@
     ),
-    data = pointsPerSeason_positionAge_newData,
+    data = pointsPerSeason_positionAge_newData %>% 
+      mutate(
+        age = round(age, 2),
+        fantasyPoints_randomLinearSlopes = round(fantasyPoints_randomLinearSlopes, 2)
+        ),
     inherit.aes = FALSE,
@@
     linewidth = 2
-  ) +
-  geom_point(
-    aes(
-      x = age,
-      y = fantasyPoints_randomLinearSlopes,
-      text = player_display_name, # add player name for mouse over tooltip
-      label = season # add season for mouse over tooltip
-    ),
-    size = 1,
-    color = "transparent" # make points invisible but keep tooltips
   ) +
@@
 plot_individualFantasyPointsRandomLinearFixedQuadracticSlopes <- ggplot(
-  data = player_stats_seasonal_offense_subsetCC,
+  data = player_stats_seasonal_offense_subsetCC %>% 
+    mutate(
+      age = round(age, 2),
+      fantasyPoints_randomLinearFixedQuadraticSlopes = round(fantasyPoints_randomLinearFixedQuadraticSlopes, 2)
+    ),
   mapping = aes(
@@
     ),
-    data = pointsPerSeason_positionAge_newData,
+    data = pointsPerSeason_positionAge_newData %>% 
+      mutate(
+        age = round(age, 2),
+        fantasyPoints_randomLinearFixedQuadraticSlopes = round(fantasyPoints_randomLinearFixedQuadraticSlopes, 2)
+        ),
     inherit.aes = FALSE,
@@
     linewidth = 2
-  ) +
-  geom_point(
-    aes(
-      x = age,
-      y = fantasyPoints_randomLinearFixedQuadraticSlopes,
-      text = player_display_name, # add player name for mouse over tooltip
-      label = season # add season for mouse over tooltip
-    ),
-    size = 1,
-    color = "transparent" # make points invisible but keep tooltips
   ) +
@@
 plot_individualFantasyPointsRandomLinearFixedQuadraticSlopesInteraction <- ggplot(
-  data = player_stats_seasonal_offense_subsetCC,
+  data = player_stats_seasonal_offense_subsetCC %>% 
+    mutate(
+      age = round(age, 2),
+      fantasyPoints_randomLinearFixedQuadraticSlopesInteraction = round(fantasyPoints_randomLinearFixedQuadraticSlopesInteraction, 2)
+    ),
   mapping = aes(
@@
     ),
-    data = pointsPerSeason_positionAge_newData,
+    data = pointsPerSeason_positionAge_newData %>% 
+      mutate(
+        age = round(age, 2),
+        fantasyPoints_randomLinearFixedQuadraticSlopesInteraction = round(fantasyPoints_randomLinearFixedQuadraticSlopesInteraction, 2)
+        ),
     inherit.aes = FALSE,
@@
     linewidth = 2
-  ) +
-  geom_point(
-    aes(
-      x = age,
-      y = fantasyPoints_randomLinearFixedQuadraticSlopesInteraction,
-      text = player_display_name, # add player name for mouse over tooltip
-      label = season # add season for mouse over tooltip
-    ),
-    size = 1,
-    color = "transparent" # make points invisible but keep tooltips
   ) +

### 2024-07-29 mixed-models.qmd
20240729 - separate plot title across multiple lines

@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age and Position: Random Intercepts Model With Position As Predictor",
+    title = "Fantasy Points (Season) by Player Age and Position:\nRandom Intercepts Model With Position As Predictor",
     color = "Position"
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age and Position: Model With Random Intercepts and Fixed Slopes",
+    title = "Fantasy Points (Season) by Player Age and Position:\nModel With Random Intercepts and Fixed Slopes",
     color = "Position"
@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age and Position: Model With Random Intercepts and Random Linear Slopes",
+    title = "Fantasy Points (Season) by Player Age and Position:\nModel With Random Intercepts and Random Linear Slopes",
     color = "Position"

### 2024-07-29 multiple-regression.qmd
20240721 - placeholders
20240729 - moderation

@@
 
+## Moderated Multiple Regression {#sec-moderatedMultipleRegression}
+
+When examining moderation in multiple regression, several steps are important:
+
+- When computing the interaction term, first mean-center the predictor variables.
+Calculate the interaction term as the multiplication of the mean-centered predictor variables.
+Mean-centering the predictor variables when computing the interaction term is important for addressing issues regarding [multicollinearity](#sec-multipleRegressionMulticollinearity) [@Iacobucci2016].
+- When including an interaction term in the model, make sure also to include the main effects.
+
+## Mediation {#sec-multipleRegressionMediation}
+
 ## Bayesian Multiple Regression {#sec-multipleRegressionBayesian}
@@
 As with [correlation](#sec-correlation), an association does not imply causation.
-However, identifiying associations is important because associations are a necessary (but insufficient) condition for causality.
+However, identifying associations is important because associations are a necessary (but insufficient) condition for causality.
 When developing a multiple regression model, it is important to pay attention for potential [multicollinearity](#sec-multipleRegressionMulticollinearity)—it may become difficult to detect a given [predictor variable](#sec-correlationalStudy) as [statistically significant](#sec-statisticalSignificance) due to the greater uncertainty around the parameter estimates.

### 2024-07-29 mixed-models.qmd
20240729 - overview of mixed models

@@
 We will discuss a modeling framework that goes by many terms, including mixed models, mixed-effects models, multilevel models, hierarchical linear models.
+They are sometimes called multilevel models and hiearchical linear models, whose name emphasizes the hierarchical structure of the data because the data are nonindependent.
+When observations (i.e., data points) are collected from multiple lower-level units (e.g., people) in an upper-level unit (e.g., married couple, family, classroom, school, neighborhood, team), the data from the lower-level units are considered "nested" within the upper-level unit.
+The data from the lower-level unit are likely to be correlated, to some degree, because they come from the same upper-level unit.
+For example, multiple players may come from the same team, and the players' performance on that team is likely interrelated because they share common experiences and influence one another.
+Thus, data from multiple players on a given team are considered nested within that team.
+Longitudinal data can also be considered nested data, in which time points are nested within the person (i.e., the same player provides an observation across multiple time points).
+As we will discuss, it is important to account for levels of nesting when the observations are nonindependent.
+
+These models are also sometimes called mixed models or mixed-effects models because the models can include a mix of fixed and random effects.
+Fixed effects are effects that are constant across individuals (i.e., upper-level units).
+Random effects are effects that vary across individuals (i.e., upper-level units).
+For instance, consider a longitudinal study of fantasy performance as a function of age.
+If we have longitudinal data for multiple players, the time points are nested within players.
+Examining the association between age as a fixed effect in relation to fantasy performace would examine the assocication between a player's age and their fantasy performance while holding the association between age and fantasy performance constant across all players.
+That is, it would assume that all players show the same trajectory such as increase for 4 years then decrease.
+Examining the association between age as a random effect in relation to fantasy performace would examine the assocication between a player's age and their fantasy performance while allowing the association between age and fantasy performance to vary across players.
+That is, it would allow the possibility that some players improve with age, whereas other players decline in performance with age.
+
+When including random effects of a variable (e.g., age) in a mixed model, it is also important to include fixed effects of that variable in the model.
+This is because random effects have a mean of zero.
+Fixed effects allow the mean to differ from zero.
+Thus, inclusion of random effects without the corresponding fixed effect can lead to bias in estimation of the association between the predictor variables and the outcome variable.
 

### 2024-07-29 cluster-analysis.qmd
20240728 - fixes
20240729 - overview of cluster analysis

@@
 
-```{r}
+Whereas [factor analysis](#sec-factorAnalysis) evaluates how *variables* do or do not hang together—in terms of their associations and non-associations, cluster analysis evaluates how *people* are or or not similar—in terms of their scores on one or more variables.
+The goal of cluster analysis is to identify distinguishable subgroups of people.
+The people within a subgroup are expected to be more similar to each other than they are to people in other subgroups.
+For instance, we might expect that there are distinguishable subtypes of Wide Receivers: possession, deep threats, and slot-type Wide Receivers.
+Possession Wide Receivers tend to be taller and heavier, with good hands who catch the ball at a high rate.
+Deep threat Wide Receivers tend to be fast.
+Slot-type Wide Receivers tend to be small, quick, and agile.
+In order to identify these clusters of Wide Receivers, we might conduct a cluster analysis with variables relating to the players' height, weight, percent of (catchable) targets caught, air yards received, and various metrics from the National Football League (NFL) Combine, including their times in the 40-yard dash, 20-yard shuttle run, and three cone drill.
 
-```
+There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, etc.
+In this chapter, we focus on examples using model-based clustering with the `mclust` package, which uses Gaussian finite mixture modeling.
 
@@
 ```{r}
-recentSeason <- max(player_stats_seasonal_offense$season, na.rm = TRUE)
+recentSeason <- max(player_stats_seasonal_offense$season, na.rm = TRUE) # also works: nflreadr::most_recent_season()
 recentSeason
@@
 
-#### Fit the Cluster Model to the Optimal Number of Tiers {#sec-clusterAnalysisExampleNumTiers}
+#### Fit the Cluster Model to the Optimal Number of Tiers {#sec-clusterAnalysisExampleModel}
 
-##### Quarterbacks {#sec-clusterAnalysisExampleNumTiersQBs}
+##### Quarterbacks {#sec-clusterAnalysisExampleModelQBs}
 
@@
 
-##### Running Backs {#sec-clusterAnalysisExampleNumTiersRBs}
+##### Running Backs {#sec-clusterAnalysisExampleModelRBs}
 
@@
 
-##### Wide Receivers {#sec-clusterAnalysisExampleNumTiersWRs}
+##### Wide Receivers {#sec-clusterAnalysisExampleModelWRs}
 
@@
 
-##### Tight Ends {#sec-clusterAnalysisExampleNumTiersTEs}
+##### Tight Ends {#sec-clusterAnalysisExampleModelTEs}
 

### 2024-07-29 player-evaluation.qmd
20240729 - overview of cluster analysis

@@
 
-Many aspects of a player's athletic profile, including tests of speed (40-yard dash), strength (bench press), agility (2-yard shuttle run; three cone drill), and jumping ability (vertical jump; broad jump) are available from the National Football League (NFL) Combine, which is especially relevant for evaluating rookies.
+Many aspects of a player's athletic profile, including tests of speed (40-yard dash), strength (bench press), agility (20-yard shuttle run; three cone drill), and jumping ability (vertical jump; broad jump) are available from the National Football League (NFL) Combine, which is especially relevant for evaluating rookies.
 We demonstrate how to import data from the NFL Combine in @sec-combine.

### 2024-07-29 mixed-models.qmd
20240729 - fix labels

@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age and Position:\nModel With Random Intercepts and Random Linear Slopes",
+    title = "Fantasy Points (Season) by Player Age and Position:\nModel With Random Intercepts, Random Linear Slopes, and\nFixed Quadratic Slopes",
     color = "Position"
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Mixed Model with Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position"
+    subtitle = "Mixed Model with Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes\nin Interaction With Position"
   ) +
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Quadratic Model with All Players"
+    subtitle = "Quadratic Model with All Players",
+    color = "Position"
   ) +
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Quadratic Model with Players Who Were Once at Top of End-of-Season Depth Chart"
+    subtitle = "Quadratic Model with Players Who Were Once at Top of End-of-Season Depth Chart",
+    color = "Position"
   ) +
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Generalized Additive Model with All Players"
+    subtitle = "Generalized Additive Model with All Players",
+    color = "Position"
   ) +
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Generalized Additive Model with Players Who Were Once at Top of End-of-Season Depth Chart"
+    subtitle = "Generalized Additive Model with Players Who Were Once at Top of End-of-Season Depth Chart",
+    color = "Position"
   ) +

### 2024-07-30 getting-started.qmd
20240722 - fix nested list
20240730 - separate weekly and seasonal PFR; compute more player ages from PFR

@@
 load(file = "./data/nfl_nextGenStats_weekly.RData")
-load(file = "./data/nfl_advancedStatsPFR.RData")
+load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
+load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
 load(file = "./data/nfl_playerContracts.RData")
@@
 load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
-load(file = "./data/nfl_expectedFantasyPoints_offense.RData")
+load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
 
@@
 
-### Advanced Stats from PFR {#sec-advancedStatsPFR}
-
-A Data Dictionary for PFR passing data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html>
+### Advanced Stats from Pro Football Reference {#sec-advancedStatsPFR}
+
+A Data Dictionary for Pro Football Reference passing data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html>
 
@@
 
-nfl_advancedStatsPFR <- bind_rows(
+nfl_advancedStatsPFR_seasonal <- bind_rows(
   nfl_advancedStatsPFR_pass_seasonal,
+  nfl_advancedStatsPFR_rush_seasonal,
+  nfl_advancedStatsPFR_rec_seasonal,
+  nfl_advancedStatsPFR_def_seasonal
+)
+
+nfl_advancedStatsPFR_weekly <- bind_rows(
   nfl_advancedStatsPFR_pass_weekly,
-  nfl_advancedStatsPFR_rush_seasonal,
   nfl_advancedStatsPFR_rush_weekly,
-  nfl_advancedStatsPFR_rec_seasonal,
   nfl_advancedStatsPFR_rec_weekly,
-  nfl_advancedStatsPFR_def_seasonal,
-  nfl_advancedStatsPFR_def_weekly,
-)
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_advancedStatsPFR,
-  file = "./data/nfl_advancedStatsPFR.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_advancedStatsPFR.RData")
+  nfl_advancedStatsPFR_def_weekly
+)
+```
+
+Merge with `gsis_id` for merging with other datasets:
+
+```{r}
+#| eval: false
+
+nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly %>% 
+  rename(pfr_id = pfr_player_id)
+
+nfl_advancedStatsPFR_seasonal <- left_join(
+  nfl_advancedStatsPFR_seasonal,
+  nfl_playerIDs %>% 
+    filter(!is.na(pfr_id)) %>% 
+    filter(gsis_id != "00-0039137") %>% # drop DL Byron Young, keep OLB Byron Young
+    select(pfr_id, gsis_id) %>% 
+    unique(),
+  by = "pfr_id"
+)
+
+nfl_advancedStatsPFR_weekly <- left_join(
+  nfl_advancedStatsPFR_weekly,
+  nfl_playerIDs %>% 
+    filter(!is.na(pfr_id)) %>% 
+    filter(gsis_id != "00-0039137") %>% # drop DL Byron Young, keep OLB Byron Young
+    select(pfr_id, gsis_id) %>% 
+    unique(),
+  by = "pfr_id"
+)
+
+nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0035665" & nfl_advancedStatsPFR_seasonal$pos %in% c("LB","LILB","RILB"))] <- NA # drop LB David Young, keep DB David Young
+#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0035665" & nfl_advancedStatsPFR_weekly$team %in% c("TEN","MIA"))] <- NA # drop LB David Young, keep DB David Young
+
+nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0035292" & nfl_advancedStatsPFR_seasonal$pos %in% c("LB","LILB","RILB"))] <- NA # drop LB David Young, keep DB David Young
+nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0035292" & nfl_advancedStatsPFR_weekly$team %in% c("TEN","MIA"))] <- NA # drop LB David Young, keep DB David Young
+
+nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0033894" & nfl_advancedStatsPFR_seasonal$pos == "DB")] <- NA # drop S Marcus Williams, keep DB David Young
+#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0033894" & nfl_advancedStatsPFR_weekly$pos == "DB")] <- NA # drop S Marcus Williams
+
+nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0038407" & nfl_advancedStatsPFR_seasonal$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones
+#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0038407" & nfl_advancedStatsPFR_weekly$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones
+
+nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0037106" & nfl_advancedStatsPFR_seasonal$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones
+#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0037106" & nfl_advancedStatsPFR_weekly$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones
+
+nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0038549" & nfl_advancedStatsPFR_seasonal$pos == "WR")] <- NA # drop WR DJ TUrner, keep CB DJ Turner
+#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0038549" & nfl_advancedStatsPFR_weekly$pos == "WR")] <- NA # drop WR DJ TUrner, keep CB DJ Turner
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_advancedStatsPFR_seasonal,
+  file = "./data/nfl_advancedStatsPFR_seasonal.RData"
+)
+
+save(
+  nfl_advancedStatsPFR_weekly,
+  file = "./data/nfl_advancedStatsPFR_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
+load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
 ```
@@
 
-nfl_expectedFantasyPoints_offense <- bind_rows(
+nfl_expectedFantasyPoints_pbp <- bind_rows(
   nfl_expectedFantasyPoints_pass,
@@
 save(
-  nfl_expectedFantasyPoints_offense,
-  file = "./data/nfl_expectedFantasyPoints_offense.RData"
+  nfl_expectedFantasyPoints_pbp,
+  file = "./data/nfl_expectedFantasyPoints_pbp.RData"
 )
@@
 load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
-load(file = "./data/nfl_expectedFantasyPoints_offense.RData")
+load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
 ```
@@
 
+# Merge with Pro Football Reference Data on Player Age by Season
+player_game_birthdate_gameday_offense <- player_game_birthdate_gameday_offense %>% 
+  dplyr::left_join(
+    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
+    by = c("player_id" = "gsis_id", "season")
+  )
+
+player_game_birthdate_gameday_defense <- player_game_birthdate_gameday_defense %>% 
+  dplyr::left_join(
+    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
+    by = c("player_id" = "gsis_id", "season")
+  )
+
+player_game_birthdate_gameday_kicking <- player_game_birthdate_gameday_kicking %>% 
+  dplyr::left_join(
+    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
+    by = c("player_id" = "gsis_id", "season")
+  )
+
+# Set age as first non-missing value from calculation above or from PFR
+player_game_birthdate_gameday_offense <- player_game_birthdate_gameday_offense %>% 
+  mutate(age = coalesce(age.x, age.y)) %>% 
+  select(-age.x, -age.y)
+
+player_game_birthdate_gameday_defense <- player_game_birthdate_gameday_defense %>% 
+  mutate(age = coalesce(age.x, age.y)) %>% 
+  select(-age.x, -age.y)
+
+player_game_birthdate_gameday_kicking <- player_game_birthdate_gameday_kicking %>% 
+  mutate(age = coalesce(age.x, age.y)) %>% 
+  select(-age.x, -age.y)
+
 # Calculate ageCentered and ageCenteredQuadratic
@@
 
+# Merge with Pro Football Reference Data on Player Age by Season
+player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+  dplyr::left_join(
+    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
+    by = c("player_id" = "gsis_id", "season")
+  )
+
+player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
+  dplyr::left_join(
+    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
+    by = c("player_id" = "gsis_id", "season")
+  )
+
+player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
+  dplyr::left_join(
+    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
+    by = c("player_id" = "gsis_id", "season")
+  )
+
+# Set age as first non-missing value from calculation above or from PFR
+player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+  mutate(age = coalesce(age.x, age.y)) %>% 
+  select(-age.x, -age.y)
+
+player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
+  mutate(age = coalesce(age.x, age.y)) %>% 
+  select(-age.x, -age.y)
+
+player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
+  mutate(age = coalesce(age.x, age.y)) %>% 
+  select(-age.x, -age.y)
+
 # Calculate ageCentered and ageCenteredQuadratic

### 2024-07-30 getting-started.qmd
20240730 - clarification comment

@@
 
+# Remove distinct players who were given the same `pfr_id` (to allow merging)
 nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0035665" & nfl_advancedStatsPFR_seasonal$pos %in% c("LB","LILB","RILB"))] <- NA # drop LB David Young, keep DB David Young

### 2024-07-30 machine-learning.qmd
20240729 - machine learning methods for longitudinal data
Built site for gh-pages

@@
 ```{r}
+library("LongituRF")
+```
 
+### Load Data {#sec-machineLearningLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+# Downloaded Data
+load(file = "./data/nfl_players.RData")
+load(file = "./data/nfl_teams.RData")
+load(file = "./data/nfl_rosters.RData")
+load(file = "./data/nfl_rosters_weekly.RData")
+load(file = "./data/nfl_schedules.RData")
+load(file = "./data/nfl_combine.RData")
+load(file = "./data/nfl_draftPicks.RData")
+load(file = "./data/nfl_depthCharts.RData")
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
+load(file = "./data/nfl_actualStats_weekly.RData")
+load(file = "./data/nfl_injuries.RData")
+load(file = "./data/nfl_snapCounts.RData")
+load(file = "./data/nfl_espnQBR.RData")
+load(file = "./data/nfl_nextGenStats_weekly.RData")
+load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
+load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
+load(file = "./data/nfl_playerContracts.RData")
+load(file = "./data/nfl_ftnCharting.RData")
+load(file = "./data/nfl_playerIDs.RData")
+load(file = "./data/nfl_rankings.RData")
+load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
+load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
+
+# Calculated Data
+load(file = "./data/nfl_actualStats_career.RData")
+load(file = "./data/nfl_actualStats_seasonal.RData")
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+```
+
+```{r}
+# Downloaded Data
+load(file = "./data/nfl_players.RData")
+load(file = "./data/nfl_teams.RData")
+load(file = "./data/nfl_rosters.RData")
+load(file = "./data/nfl_rosters_weekly.RData")
+load(file = "./data/nfl_schedules.RData")
+load(file = "./data/nfl_combine.RData")
+load(file = "./data/nfl_draftPicks.RData")
+load(file = "./data/nfl_depthCharts.RData")
+load(file = "./data/nfl_pbp.RData")
+load(file = "./data/nfl_4thdown.RData")
+load(file = "./data/nfl_participation.RData")
+load(file = "./data/nfl_actualStats_weekly.RData")
+load(file = "./data/nfl_injuries.RData")
+load(file = "./data/nfl_snapCounts.RData")
+load(file = "./data/nfl_espnQBR.RData")
+load(file = "./data/nfl_nextGenStats_weekly.RData")
+load(file = "./data/nfl_advancedStatsPFR.RData")
+load(file = "./data/nfl_playerContracts.RData")
+load(file = "./data/nfl_ftnCharting.RData")
+load(file = "./data/nfl_playerIDs.RData")
+load(file = "./data/nfl_rankings.RData")
+load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
+load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
+
+# Calculated Data
+load(file = "./data/nfl_actualStats_career.RData")
+load(file = "./data/nfl_actualStats_seasonal.RData")
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+```
+
+## Overview of Machine Learning {#sec-machineLearningOverview}
+
+Machine learning can be useful for leveraging big data and lots of predictor variable to develop predictive models with greater accuracy.
+However, many machine learning techniques are black boxes—it is often unclear
+Machine learning tends to be a data-driven, atheoretical technique.
+This can result in overfitting.
+Thus, when estimating machine learning models, it is common to keep a hold-out sample for use in cross-validation to evaluate the extent of shrinkage of model coefficients.
+Shrinkage occurs when predictor variables explain some random error variance in the original model.
+When the model is applied to an independent sample, the predictive model will likely not perform quite as well, and the regressions coefficients will tend to get smaller (i.e., shrink).
+
+## Data Processing {#sec-machineLearningDataProcessing}
+
+```{r}
+#| eval: false
+#| include: false
+
+varNames <- c(
+  names(nfl_players),
+  names(nfl_teams),
+  names(nfl_rosters),
+  names(nfl_rosters_weekly),
+  names(nfl_schedules),
+  names(nfl_combine),
+  names(nfl_draftPicks),
+  names(nfl_depthCharts),
+  names(nfl_pbp),
+  names(nfl_4thdown),
+  names(nfl_participation),
+  names(nfl_actualStats_offense_weekly),
+  names(nfl_injuries),
+  names(nfl_snapCounts),
+  names(nfl_espnQBR),
+  names(nfl_nextGenStats_weekly),
+  names(nfl_advancedStatsPFR_seasonal),
+  names(nfl_advancedStatsPFR_weekly),
+  names(nfl_playerContracts),
+  names(nfl_ftnCharting),
+  names(nfl_playerIDs),
+  names(nfl_rankings),
+  names(nfl_expectedFantasyPoints_weekly),
+  names(nfl_expectedFantasyPoints_pbp)
+)
+
+varNames <- unique(varNames)
+
+write.csv(
+  varNames,
+  file = "./data/varNames.csv",
+  row.names = FALSE
+)
+
+nfl_players$gsis_id
+nfl_rosters$gsis_id
+nfl_rosters_weekly$gsis_id
+nfl_draftPicks$gsis_id
+nfl_depthCharts$gsis_id
+nfl_advancedStatsPFR_seasonal$gsis_id
+
+nfl_actualStats_offense_weekly$player_id
+nfl_expectedFantasyPoints_weekly$player_id
+
+nfl_rankings$id
+
+nfl_combine$pfr_id
+nfl_advancedStatsPFR$pfr_id
+nfl_advancedStatsPFR$pfr_player_id
+
+nfl_playerIDs$gsis_id
+nfl_playerIDs$pfr_id
+```
+
+### Merge Data {#sec-machineLearningMergeData}
+
+```{r}
+# Use common name for the same (gsis_id) ID variable
+nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly %>% 
+  rename(gsis_id = player_id)
+
+nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal %>% 
+  rename(gsis_id = player_id)
+
+player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+  rename(gsis_id = player_id)
+
+player_stats_weekly_offense <- player_stats_weekly_offense %>% 
+  rename(gsis_id = player_id)
+
+nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
+  rename(gsis_id = player_id)
+
+player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+  rename(gsis_id = player_id)
+
+# Create lists of objects to merge, depending on data structure: id; or id-season; or id-season-week
+#-todo: remove redundant variables
+idListToMerge <- list(
+  nfl_players %>% filter(!is.na(gsis_id)) %>% select(-season),
+  nfl_draftPicks %>% filter(!is.na(gsis_id)) %>% select(-season, -age)
+)
+
+idSeasonListToMerge <- list(
+  nfl_rosters %>% filter(!is.na(gsis_id), !is.na(season)) %>% select(-week),
+  player_stats_seasonal_offense %>% filter(!is.na(gsis_id), !is.na(season)), #nfl_actualStats_offense_seasonal %>% filter(!is.na(gsis_id), !is.na(season)),
+  nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season))
+)
+
+idSeasonWeekToMerge <- list(
+  nfl_rosters_weekly %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week)),
+  nfl_depthCharts %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week)),
+  #nfl_actualStats_offense_weekly,
+  nfl_expectedFantasyPoints_weekly %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week))
+  #nfl_advancedStatsPFR_weekly,
+)
+
+# Merge data
+idMerged <- idListToMerge %>% 
+  reduce(full_join, by = c("gsis_id"))
+
+idSeasonMerged <- idSeasonListToMerge %>% 
+  reduce(full_join, by = c("gsis_id","season"))
+
+idSeasonWeekMerged <- idSeasonWeekToMerge %>% 
+  reduce(full_join, by = c("gsis_id","season","week"))
+
+seasonalData <- full_join(
+  idMerged,
+  idSeasonMerged,
+  by = "gsis_id"
+)
+
+seasonalAndWeeklyData <- full_join( # need to troubleshoot error
+  seasonalData,
+  idSeasonWeekMerged,
+  by = c("gsis_id","season")
+)
 ```
@@
 
+#### Cross-Sectional Data {#sec-randomForestCrossSectionalData}
+
+#### Longitudinal Data {#sec-randomForestLongitudinalData}
+
+[@Hu2023]
+
 ### *k*-Fold Cross-Validation {#sec-kfoldCV}

### 2024-07-30 mixed-models.qmd
20240730 - fix model fit comparisons

@@
 
+MuMIn::AICc(
+  pointsPerSeason_nullModel,
+  pointsPerSeason_linearRegression,
+  pointsPerSeason_quadraticRegression
+)
+
 summary(pointsPerSeason_nullModel)$r.squared
@@
 bbmle::AICctab(mixedModels2)
+
+MuMIn::AICc(
+  pointsPerSeason_nullModel,
+  pointsPerSeason_linearRegression,
+  pointsPerSeason_quadraticRegression,
+  pointsPerSeason_randomIntercepts,
+  pointsPerSeason_positionAgeFixedLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction
+)
 ```
@@
 AIC(
+  #pointsPerSeason_nullModel,
+  #pointsPerSeason_linearRegression,
+  #pointsPerSeason_quadraticRegression,
+  #pointsPerSeason_randomIntercepts,
   pointsPerSeason_positionAgeFixedLinearSlopes,
@@
 #AICcmodavg::aictab(linearMixedModelsVsGAM) # throws error (can't mix bam with lmer)
-bbmle::AICctab(linearMixedModelsVsGAM)
-
-BIC(
+#bbmle::AICctab(linearMixedModelsVsGAM) # different numbers of observations
+
+MuMIn::AICc(
+  pointsPerSeason_nullModel,
+  pointsPerSeason_linearRegression,
+  pointsPerSeason_quadraticRegression,
+  pointsPerSeason_randomIntercepts,
   pointsPerSeason_positionAgeFixedLinearSlopes,
@@
 
+BIC(
+  #pointsPerSeason_nullModel,
+  #pointsPerSeason_linearRegression,
+  #pointsPerSeason_quadraticRegression,
+  #pointsPerSeason_randomIntercepts,
+  pointsPerSeason_positionAgeFixedLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes,
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction,
+  pointsPerSeason_gam
+)
+
 #AICcmodavg::bictab(linearMixedModelsVsGAM) # throws error (can't mix bam with lmer)
-bbmle::BICtab(linearMixedModelsVsGAM)
+#bbmle::AICctab(linearMixedModelsVsGAM) # different numbers of observations
 

### 2024-07-30 getting-started.qmd
20240730 - identify duplicates

@@
 
+# Prepare data for merging
 nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly %>% 
@@
 
+# Identify duplicates
+nfl_advancedStatsPFR_seasonal %>% 
+  select(pfr_id, season, age) %>% 
+  na.omit() %>% 
+  unique() %>% 
+  group_by(pfr_id, season) %>% 
+  filter(n() > 1) %>% 
+  arrange(pfr_id, season)
+
+# Merge seasonal data with the player IDs
 nfl_advancedStatsPFR_seasonal <- left_join(
@@
 
+# Merge weekly data with the player IDs
 nfl_advancedStatsPFR_weekly <- left_join(

### 2024-07-30 machine-learning.qmd
20240730 - merging, data prep, imputation

@@
 library("LongituRF")
+library("missForest")
+library("powerjoin")
+library("tidyverse")
 ```
@@
 ```{r}
-# Use common name for the same (gsis_id) ID variable
+# Prepare data for merging
+## Use common name for the same (gsis_id) ID variable
 nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly %>% 
@@
 
+## Ensure variables with the same name have the same type
+nfl_players <- nfl_players %>% 
+  mutate(
+    birth_date = as.Date(birth_date),
+    jersey_number = as.character(jersey_number),
+    gsis_it_id = as.character(gsis_it_id),
+    years_of_experience = as.integer(years_of_experience))
+
 player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
-  rename(gsis_id = player_id)
+  mutate(
+    birth_date = as.Date(birth_date),
+    jersey_number = as.character(jersey_number),
+    gsis_it_id = as.character(gsis_it_id))
+
+nfl_rosters <- nfl_rosters %>% 
+  mutate(
+    draft_number = as.integer(draft_number))
+
+nfl_rosters_weekly <- nfl_rosters %>% 
+  mutate(
+    draft_number = as.integer(draft_number))
+
+nfl_depthCharts <- nfl_depthCharts %>% 
+  mutate(
+    season = as.integer(season))
+
+nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
+  mutate(
+    receptions = as.integer(receptions))
+
+## Rename variables
+#-todo: rename variables in expected fantasy points so they don't get coalesced with actual points
+nfl_draftPicks <- nfl_draftPicks %>%
+  rename(
+    games_career = games,
+    pass_completions_career = pass_completions,
+    pass_attempts_career = pass_attempts,
+    pass_yards_career = pass_yards,
+    pass_tds_career = pass_tds,
+    pass_ints_career = pass_ints,
+    rush_atts_career = rush_atts,
+    rush_yards_career = rush_yards,
+    rush_tds_career = rush_tds,
+    receptions_career = receptions,
+    rec_yards_career = rec_yards,
+    rec_tds_career = rec_tds,
+    def_solo_tackles_career = def_solo_tackles,
+    def_ints_career = def_ints,
+    def_sacks_career = def_sacks
+  )
 
@@
 
-idSeasonWeekToMerge <- list(
+idSeasonWeekListToMerge <- list(
   nfl_rosters_weekly %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week)),
@@
 idMerged <- idListToMerge %>% 
-  reduce(full_join, by = c("gsis_id"))
+  reduce(
+    powerjoin::power_full_join,
+    by = c("gsis_id"),
+    conflict = coalesce_xy)
 
 idSeasonMerged <- idSeasonListToMerge %>% 
-  reduce(full_join, by = c("gsis_id","season"))
+  reduce(
+    powerjoin::power_full_join,
+    by = c("gsis_id","season"),
+    conflict = coalesce_xy)
 
-idSeasonWeekMerged <- idSeasonWeekToMerge %>% 
-  reduce(full_join, by = c("gsis_id","season","week"))
+idSeasonWeekMerged <- idSeasonWeekListToMerge %>% 
+  reduce(
+    powerjoin::power_full_join,
+    by = c("gsis_id","season","week"),
+    conflict = coalesce_xy)
 
-seasonalData <- full_join(
+seasonalData <- powerjoin::power_full_join(
+  idSeasonMerged,
   idMerged,
-  idSeasonMerged,
-  by = "gsis_id"
+  by = "gsis_id",
+  conflict = coalesce_xy
 )
 
-seasonalAndWeeklyData <- full_join( # need to troubleshoot error
+seasonalAndWeeklyData <- powerjoin::power_full_join(
+  idSeasonWeekMerged,
   seasonalData,
-  idSeasonWeekMerged,
-  by = c("gsis_id","season")
+  by = c("gsis_id","season"),
+  conflict = coalesce_xy
 )
+```
+
+### Subset to Predictor Variables and Outcome Variable {#sec-subsetToPredictorsAndOutcome}
+
+```{r}
+seasonalData %>% select_if(~class(.) == "Date")
+seasonalData %>% select_if(is.character)
+seasonalData %>% select_if(is.factor)
+seasonalData %>% select_if(is.logical)
+
+seasonalData_subset <- seasonalData %>% 
+  dplyr::select(-c(
+    birth_date, loaded, full_name, player_name, player_display_name, display_name, suffix, headshot_url, player, pos,
+    gsis_id, espn_id, sportradar_id, yahoo_id, rotowire_id, pff_id, fantasy_data_id, sleeper_id, pfr_id,
+    pfr_player_id, cfb_player_id, pfr_player_name, esb_id, gsis_it_id, smart_id,
+    college, college_name, team_abbr, current_team_id, college_conference, draft_club, status_description_abbr,
+    status_short_description, short_name, headshot, uniform_number, jersey_number, first_name, last_name,
+    football_name, team
+    ))
+```
+
+### Convert Character Predictors to Factors {#sec-convertCharacterToFactor}
+
+```{r}
+seasonalData_subset <- seasonalData_subset %>% mutate(across(where(is.character), as.factor))
+seasonalData_subset <- seasonalData_subset %>% mutate(across(where(is.logical), as.factor))
+
+seasonalData_subset %>% select_if(~class(.) == "Date") %>% str()
+seasonalData_subset %>% select_if(is.character) %>% str()
+seasonalData_subset %>% select_if(is.factor) %>% str()
+seasonalData_subset %>% select_if(is.logical) %>% str()
+```
+
+### Impute the Missing Data {#sec-missingDataImputation}
+
+```{r}
+seasonalData_subset_imp <- missForest::missForest(seasonalData_subset %>% data.frame)
 ```
@@
 
+```{r}
+smerf <- LongituRF::MERF(
+  X = seasonalData[,c("ageCentered20")] %>% as.matrix(),
+  Y = seasonalData$fantasy_points,
+  Z = seasonalData[,c("ageCentered20")] %>% as.matrix(),
+  id = seasonalData$gsis_id,
+  time = seasonalData$ageCentered20,
+  mtry = 2,
+  ntree = 500,
+  sto = "BM")
+
+smerf$forest # is the fitted random forest (obtained at the last iteration)
+smerf$random_effects # are the predicted random effects for each individual
+smerf$omega # are the predicted stochastic processes
+plot(smerf$Vraisemblance) # evolution of the log-likelihood
+smerf$OOB # OOB error at each iteration
+```
+
 ### *k*-Fold Cross-Validation {#sec-kfoldCV}

### 2024-07-30 machine-learning.qmd
20240730 - parallelization

@@
 library("LongituRF")
+library("doParallel")
 library("missForest")
@@
 ```{r}
-seasonalData_subset_imp <- missForest::missForest(seasonalData_subset %>% data.frame)
+num_cores <- detectCores()
+num_cores
+
+registerDoParallel(cores = num_cores)
+
+seasonalData_subset_imp <- missForest::missForest(
+  seasonalData_subset %>% data.frame,
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "variables")
 ```

### 2024-07-31 getting-started.qmd
20240731 - duplicate instances

@@
 
+The `nfl_players` object is in `id` form.
+That is, each row should be uniquely identified by `gsis_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_players <- nfl_players %>% 
+  select(gsis_id, everything()) %>% 
+  arrange(display_name)
+```
+
+Let's check for duplicate `id` instances:
+
+```{r}
+nfl_players %>% 
+  group_by(gsis_id) %>% 
+  filter(n() > 1)
+```
+
+Let's do some data cleanup:
+
+```{r}
+nfl_players[nfl_players == ""] <- NA
+
+nfl_players <- nfl_players %>% 
+  filter(!is.na(gsis_id))
+```
+
 ```{r}
@@
 
+The `nfl_teams` object is in `id` form.
+That is, each row should be uniquely identified by `team_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_teams <- nfl_teams %>% 
+  select(team_id, everything())
+```
+
+Let's check for duplicate `id` instances:
+
+```{r}
+nfl_teams %>% 
+  group_by(team_id) %>% 
+  filter(n() > 1)
+```
+
 ```{r}
@@
 
+The `nfl_rosters` object is in `id`-`season`-`team` form.
+That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `team`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_rosters <- nfl_rosters %>% 
+  select(gsis_id, season, team, week, everything()) %>% 
+  arrange(full_name, gsis_id, season, team, week)
+```
+
+Let's check for duplicate `id` instances:
+
+```{r}
+nfl_rosters %>% 
+  group_by(gsis_id, season, team) %>% 
+  filter(n() > 1)
+```
+
+Let's do some data cleanup:
+
+```{r}
+# Drop players with missing values for gsis_id
+nfl_rosters <- nfl_rosters %>% 
+  filter(!is.na(gsis_id))
+
+# Fill in missing values for a player in their duplicate instances, and then keep only the first of the duplicate instances
+nfl_rosters <- nfl_rosters %>% 
+  group_by(gsis_id, season, team) %>% 
+  fill(names(.), .direction = "downup") %>% 
+  slice_head(n = 1) %>% 
+  ungroup()
+```
+
+Let's check again for duplicate `id` instances:
+
+```{r}
+nfl_rosters %>% 
+  group_by(gsis_id, season, team) %>% 
+  filter(n() > 1)
+```
+
+The `nfl_rosters_weekly` object is in `id`-`season`-`week` form.
+That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `week`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_rosters_weekly <- nfl_rosters_weekly %>% 
+  select(gsis_id, season, week, everything()) %>% 
+  arrange(full_name, gsis_id, season, week)
+```
+
+Let's check for duplicate `id` instances:
+
+```{r}
+nfl_rosters_weekly %>% 
+  group_by(gsis_id, season, week) %>% 
+  filter(n() > 1)
+```
+
+Let's do some data cleanup:
+
+```{r}
+# Drop players with missing values for gsis_id
+nfl_rosters_weekly <- nfl_rosters_weekly %>% 
+  filter(!is.na(gsis_id))
+
+# Fill in missing values for a player in their duplicate instances, and then keep only the first of the duplicate instances
+nfl_rosters_weekly <- nfl_rosters_weekly %>% 
+  group_by(gsis_id, season, week) %>% 
+  fill(names(.), .direction = "downup") %>% 
+  slice_head(n = 1) %>% 
+  ungroup()
+```
+
+Let's check again for duplicate `id` instances:
+
+```{r}
+nfl_rosters_weekly %>% 
+  group_by(gsis_id, season, week) %>% 
+  filter(n() > 1)
+```
+
 ```{r}
@@
 
+The `nfl_schedules` object is in `id` form.
+That is, each row should be uniquely identified by `game_id`.
+
+Let's check for duplicate `id` instances:
+
+```{r}
+nfl_schedules %>% 
+  group_by(game_id) %>% 
+  filter(n() > 1)
+```
+
 ```{r}
@@
 
+The `nfl_combine` object is in `id` form.
+That is, each row should be uniquely identified by their `id`.
+However, there is no `gsis_id` variable to merge it easily with other datasets.
+Some of the players have other `id` variables, including `pfr_id` and `cfb_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_combine <- nfl_combine %>% 
+  select(pfr_id, cfb_id, everything()) %>% 
+  arrange(season, player_name)
+```
+
+Let's check for duplicate `id` instances:
+
+```{r}
+nfl_combine %>% 
+  group_by(pfr_id) %>% 
+  filter(n() > 1, !is.na(pfr_id)) %>% 
+  arrange(pfr_id)
+
+nfl_combine %>% 
+  group_by(cfb_id) %>% 
+  filter(n() > 1, !is.na(cfb_id)) %>% 
+  arrange(cfb_id)
+```
+
+However, these apparent duplicates appear to be different players:
+
+```{r}
+nfl_combine %>% 
+  group_by(season, pfr_id, pos) %>% 
+  filter(n() > 1, !is.na(pfr_id)) %>% 
+  arrange(pfr_id)
+
+nfl_combine %>% 
+  group_by(season, cfb_id, pos) %>% 
+  filter(n() > 1, !is.na(cfb_id)) %>% 
+  arrange(cfb_id)
+```
+
 ```{r}
@@
 
+The `nfl_draftPicks` object is in `id` form.
+That is, each row should be uniquely identified by `gsis_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_draftPicks <- nfl_draftPicks %>% 
+  select(gsis_id, everything()) %>% 
+  arrange(pfr_player_name)
+```
+
+Let's check for duplicate `id` instances:
+
+```{r}
+nfl_draftPicks %>% 
+  group_by(gsis_id) %>% 
+  filter(n() > 1)
+```
+
+Let's do some data cleanup:
+
+```{r}
+nfl_draftPicks[nfl_draftPicks == ""] <- NA
+
+nfl_draftPicks <- nfl_draftPicks %>% 
+  filter(!is.na(gsis_id))
+```
+
+Let's check again for duplicate `id` instances:
+
+```{r}
+nfl_draftPicks %>% 
+  group_by(gsis_id) %>% 
+  filter(n() > 1)
+```
+
 ```{r}
@@
   nflreadr::load_depth_charts(seasons = TRUE))
+```
+
+The `nfl_depthCharts` object is in `id`-`season`-`week`-`position` form.
+That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, `week`, and `depth_position`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_depthCharts <- nfl_depthCharts %>% 
+  select(gsis_id, season, week, depth_position, everything()) %>% 
+  arrange(full_name, gsis_id, season, week, depth_position)
+```
+
+Let's check for duplicate `id` instances:
+
+```{r}
+nfl_depthCharts %>% 
+  group_by(gsis_id, season, week, depth_position) %>% 
+  filter(n() > 1)
+```
+
+Let's do some data cleanup:
+
+```{r}
+# Drop players with missing values for gsis_id
+nfl_depthCharts <- nfl_depthCharts %>% 
+  filter(!is.na(gsis_id))
+
+# Fill in missing values for a player in their duplicate instances, and then keep only the first of the duplicate instances
+nfl_depthCharts <- nfl_depthCharts %>% 
+  group_by(gsis_id, season, week, depth_position) %>% 
+  fill(names(.), .direction = "downup") %>% 
+  slice_head(n = 1) %>% 
+  ungroup()
+```
+
+Let's check again for duplicate `id` instances:
+
+```{r}
+nfl_depthCharts %>% 
+  group_by(gsis_id, season, week, depth_position) %>% 
+  filter(n() > 1)
 ```
@@
 nfl_pbp <- nflreadr::load_pbp(seasons = TRUE)
+```
+
+The `nfl_pbp` object is in `game_id`-`drive`-`play_id` form.
+That is, each row should be uniquely identified by the combination of `game_id`, `drive`, `play_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_pbp <- nfl_pbp %>% 
+  select(game_id, drive, play_id, everything()) %>% 
+  arrange(game_id, drive, play_id)
+```
+
+Let's check for duplicate `id` instances:
+
+```{r}
+nfl_pbp %>% 
+  group_by(game_id, drive, play_id) %>% 
+  filter(n() > 1)
 ```
@@
   )
+
+# Arrange data
+player_stats_weekly_offense <- player_stats_weekly_offense %>% 
+  arrange(player_display_name, season, week)
+
+player_stats_weekly_defense <- player_stats_weekly_defense %>% 
+  arrange(player_display_name, season, week)
+
+player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
+  arrange(player_display_name, season, week)
 ```
@@
   )
+
+# Arrange data
+player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+  relocate(season, .after = player_id) %>% 
+  arrange(player_display_name, season)
+
+player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
+  relocate(season, .after = player_id) %>% 
+  arrange(player_display_name, season)
+
+player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
+  relocate(season, .after = player_id) %>% 
+  arrange(player_display_name, season)
 ```

### 2024-07-31 machine-learning.qmd
20240731 - duplicate instances

@@
 library("powerjoin")
+library("mice")
+library("miceadds")
 library("tidyverse")
@@
 
-### Merge Data {#sec-machineLearningMergeData}
+### Prepare Data for Merging {#sec-machineLearningPrepareDataForMerging}
 
@@
 # Prepare data for merging
+#-todo: calculate years_of_experience
 ## Use common name for the same (gsis_id) ID variable
@@
   )
+```
 
+### Merge Data {#sec-machineLearningMergeData}
+
+```{r}
 # Create lists of objects to merge, depending on data structure: id; or id-season; or id-season-week
@@
 idListToMerge <- list(
-  nfl_players %>% filter(!is.na(gsis_id)) %>% select(-season),
+  nfl_players %>% filter(!is.na(gsis_id)) %>% select(-season, -years_of_experience),
   nfl_draftPicks %>% filter(!is.na(gsis_id)) %>% select(-season, -age)
@@
 )
+
+# check on duplicate id-by-season data (nfl_rosters, nfl_advancedStatsPFR_seasonal)
+seasonalData %>% 
+  select(gsis_id, season)
+
+seasonalData %>% 
+  select(gsis_id, season) %>% 
+  unique()
 ```
@@
     birth_date, loaded, full_name, player_name, player_display_name, display_name, suffix, headshot_url, player, pos,
-    gsis_id, espn_id, sportradar_id, yahoo_id, rotowire_id, pff_id, fantasy_data_id, sleeper_id, pfr_id,
+    espn_id, sportradar_id, yahoo_id, rotowire_id, pff_id, fantasy_data_id, sleeper_id, pfr_id,
     pfr_player_id, cfb_player_id, pfr_player_name, esb_id, gsis_it_id, smart_id,
@@
 
+(CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION)
+
+#### `missForest`
+
 ```{r}
@@
 seasonalData_subset_imp <- missForest::missForest(
-  seasonalData_subset %>% data.frame,
+  seasonalData_subset %>% select(ageCentered20, ageCentered20Quadratic, fantasy_points) %>% data.frame,
   verbose = TRUE,
   mtry = 100,
-  parallelize = "variables")
+  parallelize = "forests")
+```
+
+#### `MICE`
+
+```{r}
+# Specify data to impute
+dataToImpute <- seasonalData_subset %>% 
+  select(gsis_id, season, ageCentered20, ageCentered20Quadratic, fantasy_points, years_of_experience) %>% 
+  filter(!is.na(gsis_id), !is.na(season)) %>% 
+  mutate(ID = as.integer(as.factor(as.character(gsis_id)))) %>% 
+  arrange(gsis_id, season)
+
+# Specify variables to impute
+varsToImpute <- c("fantasy_points")
+Y <- varsToImpute
+
+# Specify predictors of missing data
+clusterVariable <- "ID"
+fixedEffectPredictors <- c("ageCentered20Quadratic")
+randomEffectPrectors <- c("ageCentered20")
+
+# Specify number of imputations
+numImputations <- 1
+
+# Specify the imputation method
+meth <- make.method(dataToImpute)
+meth[1:length(meth)] <- ""
+meth[Y] <- "2l.pmm" # specify the imputation method here; this can differ by outcome variable
+
+# Specify the predictor matrix
+
+pred <- make.predictorMatrix(dataToImpute)
+pred[1:nrow(pred), 1:ncol(pred)] <- 0
+pred[Y, clusterVariable] <- (-2) # cluster variable
+pred[Y, randomEffectPrectors] <- 2 # random effect predictor
+#pred[Y, fixedEffectPredictors] <- 1 # fixed effect predictor
+pred[Y, Y] <- 1 # fixed effect predictor of each predictor in predicting the other predictors
+
+diag(pred) <- 0
+pred
+
+# Imputation
+mi_mice <- mice(
+  as.data.frame(dataToImpute),
+  method = meth,
+  predictorMatrix = pred,
+  m = numImputations,
+  maxit = 100, # generally use 100 maximum iterations; this example uses 5 for speed
+  seed = 52242)
+
+mi_mice$loggedEvents
+
+seasonalData_subset_imp <- complete(
+  mi_mice,
+  action = "stacked",
+  include = FALSE)
 ```

### 2024-08-01 getting-started.qmd
20240731 - add raw data files

@@
 
-nfl_players <- progressr::with_progress(
+nfl_players_raw <- progressr::with_progress(
   nflreadr::load_players())
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_players_raw,
+  file = "./data/nfl_players_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_players_raw.RData")
 ```
@@
 ```{r}
-nfl_players <- nfl_players %>% 
+nfl_players <- nfl_players_raw %>% 
   select(gsis_id, everything()) %>% 
@@
 ```{r}
+# Convert missing values to NA
 nfl_players[nfl_players == ""] <- NA
 
+# Drop players with missing values for gsis_id
 nfl_players <- nfl_players %>% 
@@
 
-nfl_teams <- progressr::with_progress(
+nfl_teams_raw <- progressr::with_progress(
   nflreadr::load_teams(current = TRUE))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_teams_raw,
+  file = "./data/nfl_teams_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_teams_raw.RData")
 ```
@@
 ```{r}
-nfl_teams <- nfl_teams %>% 
+nfl_teams <- nfl_teams_raw %>% 
   select(team_id, everything())
@@
 
-nfl_rosters <- progressr::with_progress(
+nfl_rosters_raw <- progressr::with_progress(
   nflreadr::load_rosters(seasons = TRUE))
 
-nfl_rosters_weekly <- progressr::with_progress(
+nfl_rosters_weekly_raw <- progressr::with_progress(
   nflreadr::load_rosters_weekly(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_rosters_raw,
+  file = "./data/nfl_rosters_raw.RData"
+)
+
+save(
+  nfl_rosters_weekly_raw,
+  file = "./data/nfl_rosters_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_rosters_raw.RData")
+load(file = "./data/nfl_rosters_weekly_raw.RData")
 ```
@@
 ```{r}
-nfl_rosters <- nfl_rosters %>% 
+nfl_rosters <- nfl_rosters_raw %>% 
   select(gsis_id, season, team, week, everything()) %>% 
@@
 ```{r}
-nfl_rosters_weekly <- nfl_rosters_weekly %>% 
+nfl_rosters_weekly <- nfl_rosters_weekly_raw %>% 
   select(gsis_id, season, week, everything()) %>% 
@@
 
-nfl_schedules <- progressr::with_progress(
+nfl_schedules_raw <- progressr::with_progress(
   nflreadr::load_schedules(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_schedules_raw,
+  file = "./data/nfl_schedules_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_schedules_raw.RData")
 ```
@@
 That is, each row should be uniquely identified by `game_id`.
+
+```{r}
+nfl_schedules <- nfl_schedules_raw
+```
 
@@
 
-nfl_combine <- progressr::with_progress(
+nfl_combine_raw <- progressr::with_progress(
   nflreadr::load_combine(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_combine_raw,
+  file = "./data/nfl_combine_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_combine_raw.RData")
 ```
@@
 ```{r}
-nfl_combine <- nfl_combine %>% 
+nfl_combine <- nfl_combine_raw %>% 
   select(pfr_id, cfb_id, everything()) %>% 
@@
 
-nfl_draftPicks <- progressr::with_progress(
+nfl_draftPicks_raw <- progressr::with_progress(
   nflreadr::load_draft_picks(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_draftPicks_raw,
+  file = "./data/nfl_draftPicks_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_draftPicks_raw.RData")
 ```
@@
 ```{r}
-nfl_draftPicks <- nfl_draftPicks %>% 
+nfl_draftPicks <- nfl_draftPicks_raw %>% 
   select(gsis_id, everything()) %>% 
@@
 ```{r}
+# Convert missing values to NA
 nfl_draftPicks[nfl_draftPicks == ""] <- NA
 
+# Drop players with missing values for gsis_id
 nfl_draftPicks <- nfl_draftPicks %>% 
@@
 
-nfl_depthCharts <- progressr::with_progress(
+nfl_depthCharts_raw <- progressr::with_progress(
   nflreadr::load_depth_charts(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_depthCharts_raw,
+  file = "./data/nfl_depthCharts_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_depthCharts_raw.RData")
 ```
@@
 ```{r}
-nfl_depthCharts <- nfl_depthCharts %>% 
+nfl_depthCharts <- nfl_depthCharts_raw %>% 
   select(gsis_id, season, week, depth_position, everything()) %>% 
@@
 
-nfl_pbp <- progressr::with_progress(
+nfl_pbp_raw <- progressr::with_progress(
   nflreadr::load_pbp(seasons = TRUE))
@@
 
-nfl_pbp <- nflreadr::load_pbp(seasons = TRUE)
+nfl_pbp_raw <- nflreadr::load_pbp(seasons = TRUE)
+```
+
+```{r}
+save(
+  nfl_pbp_raw,
+  file = "./data/nfl_pbp_raw.RData"
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_pbp_raw.RData")
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp_raw.RData", fsep = ""))
 ```
@@
 ```{r}
-nfl_pbp <- nfl_pbp %>% 
+nfl_pbp <- nfl_pbp_raw %>% 
   select(game_id, drive, play_id, everything()) %>% 

### 2024-08-01 getting-started.qmd
20240801 - add raw data files

@@
 We save each data file after downloading it, so we can use the data in subsequent chapters.
-
-```{r}
-#| eval: false
-#| include: false
-
-# Downloaded Data
+Guidance for how to merge the various data files is provided at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md>
+
+```{r}
+#| eval: false
+#| include: false
+
+# Downloaded Data - Raw
+load(file = "./data/nfl_players_raw.RData")
+load(file = "./data/nfl_teams_raw.RData")
+load(file = "./data/nfl_rosters_raw.RData")
+load(file = "./data/nfl_rosters_weekly_raw.RData")
+load(file = "./data/nfl_schedules_raw.RData")
+load(file = "./data/nfl_combine_raw.RData")
+load(file = "./data/nfl_draftPicks_raw.RData")
+load(file = "./data/nfl_depthCharts_raw.RData")
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp_raw.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown_raw.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation_raw.RData", fsep = ""))
+load(file = "./data/nfl_actualStats_weekly_raw.RData")
+load(file = "./data/nfl_injuries_raw.RData")
+load(file = "./data/nfl_snapCounts_raw.RData")
+load(file = "./data/nfl_espnQBR_seasonal_raw.RData")
+load(file = "./data/nfl_espnQBR_weekly_raw.RData")
+load(file = "./data/nfl_nextGenStats_weekly_raw.RData")
+load(file = "./data/nfl_advancedStatsPFR_seasonal_raw.RData")
+load(file = "./data/nfl_advancedStatsPFR_weekly_raw.RData")
+load(file = "./data/nfl_playerContracts_raw.RData")
+load(file = "./data/nfl_ftnCharting_raw.RData")
+load(file = "./data/nfl_playerIDs_raw.RData")
+load(file = "./data/nfl_rankings_draft_raw.RData")
+load(file = "./data/nfl_rankings_weekly_raw.RData")
+load(file = "./data/nfl_expectedFantasyPoints_weekly_raw.RData")
+load(file = "./data/nfl_expectedFantasyPoints_pbp_raw.RData")
+
+# Calculated Data - Raw
+load(file = "./data/nfl_actualStats_career_raw.RData")
+load(file = "./data/nfl_actualStats_seasonal_raw.RData")
+
+# Downloaded Data - Processed
 load(file = "./data/nfl_players.RData")
@@
 load(file = "./data/nfl_snapCounts.RData")
-load(file = "./data/nfl_espnQBR.RData")
+load(file = "./data/nfl_espnQBR_seasonal.RData")
+load(file = "./data/nfl_espnQBR_weekly.RData")
 load(file = "./data/nfl_nextGenStats_weekly.RData")
@@
 load(file = "./data/nfl_playerIDs.RData")
-load(file = "./data/nfl_rankings.RData")
+load(file = "./data/nfl_rankings_draft.RData")
+load(file = "./data/nfl_rankings_weekly.RData")
 load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
@@
 
-# Calculated Data
+# Calculated Data - Processed
 load(file = "./data/nfl_actualStats_career.RData")
@@
 
-The `nfl_players` object is in `id` form.
+The `nfl_players` object is in `player` form.
 That is, each row should be uniquely identified by `gsis_id`.
@@
 
-Let's check for duplicate `id` instances:
+Let's check for duplicate `player` instances:
 
@@
 
-The `nfl_teams` object is in `id` form.
+The `nfl_teams` object is in `team` form.
 That is, each row should be uniquely identified by `team_id`.
@@
 
-Let's check for duplicate `id` instances:
+Let's check for duplicate `team` instances:
 
@@
 
-The `nfl_rosters` object is in `id`-`season`-`team` form.
+The `nfl_rosters` object is in `player`-`season`-`team` form.
 That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `team`.
@@
 
-Let's check for duplicate `id` instances:
+Let's check for duplicate `player`-`season`-`team` instances:
 
@@
 
-Let's check again for duplicate `id` instances:
+Let's check again for duplicate `player`-`season`-`team` instances:
 
@@
 
-The `nfl_rosters_weekly` object is in `id`-`season`-`week` form.
+The `nfl_rosters_weekly` object is in `player`-`season`-`week` form.
 That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `week`.
@@
 
-Let's check for duplicate `id` instances:
+Let's check for duplicate `player`-`season`-`week` instances:
 
@@
 
-Let's check again for duplicate `id` instances:
+Let's check again for duplicate `player`-`season`-`week` instances:
 
@@
 
-The `nfl_schedules` object is in `id` form.
+The `nfl_schedules` object is in `game` form.
 That is, each row should be uniquely identified by `game_id`.
@@
 
-Let's check for duplicate `id` instances:
+Let's check for duplicate `game` instances:
 
@@
 
-The `nfl_combine` object is in `id` form.
-That is, each row should be uniquely identified by their `id`.
+The `nfl_combine` object is in `player` form.
+That is, each row should be uniquely identified by the player's `id`.
 However, there is no `gsis_id` variable to merge it easily with other datasets.
@@
 
-Let's check for duplicate `id` instances:
+Let's check for duplicate `pfr_id` and `cfb_id` instances:
 
@@
 
-The `nfl_draftPicks` object is in `id` form.
+The `nfl_draftPicks` object is in `player` form.
 That is, each row should be uniquely identified by `gsis_id`.
@@
 
-Let's check again for duplicate `id` instances:
+Let's check again for duplicate `player` instances:
 
@@
 
-Let's check for duplicate `id` instances:
+Let's check for duplicate `id`-`season`-`week`-`position` instances:
 
@@
 
-Let's check again for duplicate `id` instances:
+Let's check again for duplicate `id`-`season`-`week`-`position` instances:
 
@@
 
-The `nfl_pbp` object is in `game_id`-`drive`-`play_id` form.
+The `nfl_pbp` object is in `game`-`drive`-`play` form.
 That is, each row should be uniquely identified by the combination of `game_id`, `drive`, `play_id`.
@@
 
-Let's check for duplicate `id` instances:
+Let's check for duplicate `game`-`drive`-`play` instances:
 
@@
 
-nfl_4thdown <- nfl4th::load_4th_pbp(seasons = 2014:2023)
+nfl_4thdown_raw <- nfl4th::load_4th_pbp(seasons = 2014:nflreadr::most_recent_season())
+```
+
+```{r}
+save(
+  nfl_4thdown_raw,
+  file = "./data/nfl_4thdown_raw.RData"
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_4thdown_raw.RData")
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown_raw.RData", fsep = ""))
+```
+
+The `nfl_4thdown` object is in `game`-`drive`-`play` form.
+That is, each row should be uniquely identified by the combination of `game_id`, `drive`, `play_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_4thdown <- nfl_4thdown_raw %>% 
+  select(game_id, drive, play_id, everything()) %>% 
+  arrange(game_id, drive, play_id)
+```
+
+Let's check for duplicate `game`-`drive`-`play` instances:
+
+```{r}
+nfl_4thdown %>% 
+  group_by(game_id, drive, play_id) %>% 
+  filter(n() > 1)
 ```
@@
 ```{r}
-nfl_participation <- progressr::with_progress(
+nfl_participation_raw <- progressr::with_progress(
   nflreadr::load_participation(
@@
 save(
+  nfl_participation_raw,
+  file = "./data/nfl_participation_raw.RData"
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_participation_raw.RData")
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation_raw.RData", fsep = ""))
+```
+
+The `nfl_participation` object is in `game`-`drive`-`play` form.
+That is, each row should be uniquely identified by the combination of `nflverse_game_id`, `drive`, `play_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_participation <- nfl_participation_raw %>% 
+  select(nflverse_game_id, drive, play_id, everything()) %>% 
+  arrange(nflverse_game_id, drive, play_id)
+```
+
+Let's check for duplicate `game`-`drive`-`play` instances:
+
+```{r}
+nfl_participation %>% 
+  group_by(nflverse_game_id, drive, play_id) %>% 
+  filter(n() > 1)
+```
+
+```{r}
+save(
   nfl_participation,
@@
 
-nfl_actualStats_offense_weekly <- progressr::with_progress(
+nfl_actualStats_offense_weekly_raw <- progressr::with_progress(
   nflreadr::load_player_stats(
@@
 
-nfl_actualStats_defense_weekly <- progressr::with_progress(
+nfl_actualStats_defense_weekly_raw <- progressr::with_progress(
   nflreadr::load_player_stats(
@@
 
-nfl_actualStats_kicking_weekly <- progressr::with_progress(
+nfl_actualStats_kicking_weekly_raw <- progressr::with_progress(
   nflreadr::load_player_stats(
@@
 save(
+  nfl_actualStats_offense_weekly_raw, nfl_actualStats_defense_weekly_raw, nfl_actualStats_kicking_weekly_raw,
+  file = "./data/nfl_actualStats_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_weekly_raw.RData")
+```
+
+The `nfl_actualStats_weekly` objects are in `player`-`season`-`week` form.
+That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly_raw %>% 
+  select(player_id, season, week, everything()) %>% 
+  arrange(player_display_name, player_id, season, week)
+
+nfl_actualStats_defense_weekly <- nfl_actualStats_defense_weekly_raw %>% 
+  select(player_id, season, week, everything()) %>% 
+  arrange(player_display_name, player_id, season, week)
+
+nfl_actualStats_kicking_weekly <- nfl_actualStats_kicking_weekly_raw %>% 
+  select(player_id, season, week, everything()) %>% 
+  arrange(player_display_name, player_id, season, week)
+```
+
+Let's check for duplicate `player`-`season`-`week` instances:
+
+```{r}
+nfl_actualStats_offense_weekly %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1)
+
+nfl_actualStats_defense_weekly %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1)
+
+nfl_actualStats_kicking_weekly %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1)
+```
+
+```{r}
+#| eval: false
+
+save(
   nfl_actualStats_offense_weekly, nfl_actualStats_defense_weekly, nfl_actualStats_kicking_weekly,
@@
 
-nfl_injuries <- progressr::with_progress(
+nfl_injuries_raw <- progressr::with_progress(
   nflreadr::load_injuries(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_injuries_raw,
+  file = "./data/nfl_injuries_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_injuries_raw.RData")
+```
+
+The `nfl_injuries` object is in `player`-`season`-`week` form.
+That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `week`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_injuries <- nfl_injuries_raw %>% 
+  select(gsis_id, season, week, everything()) %>% 
+  arrange(full_name, gsis_id, season, week)
+```
+
+Let's check for duplicate `player`-`season`-`week` instances:
+
+```{r}
+nfl_injuries %>% 
+  group_by(gsis_id, season, week) %>% 
+  filter(n() > 1)
 ```
@@
 
-nfl_snapCounts <- progressr::with_progress(
+nfl_snapCounts_raw <- progressr::with_progress(
   nflreadr::load_snap_counts(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_snapCounts_raw,
+  file = "./data/nfl_snapCounts_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_snapCounts_raw.RData")
+```
+
+The `nfl_snapCounts` object is in `game`-`player` form.
+That is, each row should be uniquely identified by the combination of `game_id` and `pfr_player_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_snapCounts <- nfl_snapCounts_raw %>% 
+  select(game_id, pfr_player_id, everything()) %>% 
+  arrange(game_id, pfr_player_id)
+```
+
+Let's check for duplicate `game` instances:
+
+```{r}
+nfl_snapCounts %>% 
+  group_by(game_id, pfr_player_id) %>% 
+  filter(n() > 1)
 ```
@@
 
-nfl_espnQBR_seasonal <- progressr::with_progress(
+nfl_espnQBR_seasonal_raw <- progressr::with_progress(
   nflreadr::load_espn_qbr(
@@
 
-nfl_espnQBR_weekly <- progressr::with_progress(
+nfl_espnQBR_weekly_raw <- progressr::with_progress(
   nflreadr::load_espn_qbr(
     seasons = TRUE,
-    summary_type = c("weekly")))
-
-nfl_espnQBR_weekly$game_week <- as.character(nfl_espnQBR_weekly$game_week)
-
-nfl_espnQBR <- bind_rows(
+    summary_type = c("week")))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_espnQBR_seasonal_raw,
+  file = "./data/nfl_espnQBR_seasonal_raw.RData"
+)
+
+save(
+  nfl_espnQBR_weekly_raw,
+  file = "./data/nfl_espnQBR_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_espnQBR_seasonal_raw.RData")
+load(file = "./data/nfl_espnQBR_weekly_raw.RData")
+```
+
+The `nfl_espnQBR_seasonal` object is in `player`-`season`-`season type` form, where `season type` refers to regular season versus postseason.
+That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `season_type`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_espnQBR_seasonal <- nfl_espnQBR_seasonal_raw %>% 
+  select(player_id, season, season_type, everything()) %>% 
+  arrange(name_display, player_id, season, season_type)
+```
+
+Let's check for duplicate `player`-`season`-`team` instances:
+
+```{r}
+nfl_espnQBR_seasonal %>% 
+  group_by(player_id, season, season_type) %>% 
+  filter(n() > 1)
+```
+
+The `nfl_espnQBR_weekly` object is in both `game`-`player` form and `player`-`season`-`season type`-`week` form, where `season type` refers to regular season versus postseason.
+That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `week` or by the combination of `player_id`, `season`, `season_type`, and `week_num`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_espnQBR_weekly <- nfl_espnQBR_weekly_raw %>% 
+  select(player_id, season, season_type, week_num, everything()) %>% 
+  arrange(name_display, player_id, season, season_type, week_num)
+```
+
+Let's check for duplicate `game`-`player` or `player`-`season`-`season type`-`week` instances:
+
+```{r}
+nfl_espnQBR_weekly %>% 
+  arrange(game_id, player_id) %>% 
+  group_by(game_id, player_id) %>% 
+  filter(n() > 1)
+
+nfl_espnQBR_weekly %>% 
+  arrange(player_id, season, season_type, week_num) %>% 
+  group_by(player_id, season, season_type, week_num) %>% 
+  filter(n() > 1)
+```
+
+```{r}
+#| eval: false
+
+save(
   nfl_espnQBR_seasonal,
-  nfl_espnQBR_weekly
-)
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_espnQBR,
-  file = "./data/nfl_espnQBR.RData"
+  file = "./data/nfl_espnQBR_seasonal.RData"
+)
+
+save(
+  nfl_espnQBR_weekly,
+  file = "./data/nfl_espnQBR_weekly.RData"
 )
@@
 
-nfl_nextGenStats_pass_weekly <- progressr::with_progress(
+nfl_nextGenStats_pass_weekly_raw <- progressr::with_progress(
   nflreadr::load_nextgen_stats(
@@
 
-nfl_nextGenStats_rush_weekly <- progressr::with_progress(
+nfl_nextGenStats_rush_weekly_raw <- progressr::with_progress(
   nflreadr::load_nextgen_stats(
@@
 
-nfl_nextGenStats_rec_weekly <- progressr::with_progress(
+nfl_nextGenStats_rec_weekly_raw <- progressr::with_progress(
   nflreadr::load_nextgen_stats(
@@
 
-nfl_nextGenStats_weekly <- bind_rows(
-  nfl_nextGenStats_pass_weekly,
-  nfl_nextGenStats_rush_weekly,
-  nfl_nextGenStats_rec_weekly
-)
+nfl_nextGenStats_weekly_raw <- bind_rows(
+  nfl_nextGenStats_pass_weekly_raw,
+  nfl_nextGenStats_rush_weekly_raw,
+  nfl_nextGenStats_rec_weekly_raw
+)
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_nextGenStats_weekly_raw,
+  file = "./data/nfl_nextGenStats_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_nextGenStats_weekly_raw.RData")
+```
+
+The `nfl_nextGenStats_weekly` object is in `player`-`season`-`season type`-`week` form, where `season type` refers to regular season versus postseason.
+That is, each row should be uniquely identified by the combination of `player_gsis_id`, `season`, `season_type`, and `week`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_nextGenStats_weekly <- nfl_nextGenStats_weekly_raw %>% 
+  select(player_gsis_id, season, season_type, week, everything()) %>% 
+  arrange(player_display_name, player_gsis_id, season, season_type)
+```
+
+Let's check for duplicate `player`-`season`-`season type`-`week` instances:
+
+```{r}
+nfl_nextGenStats_weekly %>% 
+  group_by(player_gsis_id, season, season_type, week) %>% 
+  filter(n() > 1)
 ```
@@
 
-nfl_advancedStatsPFR_pass_seasonal <- progressr::with_progress(
+nfl_advancedStatsPFR_pass_seasonal_raw <- progressr::with_progress(
   nflreadr::load_pfr_advstats(
@@
 
-nfl_advancedStatsPFR_pass_weekly <- progressr::with_progress(
+nfl_advancedStatsPFR_pass_weekly_raw <- progressr::with_progress(
   nflreadr::load_pfr_advstats(
@@
 
-nfl_advancedStatsPFR_rush_seasonal <- progressr::with_progress(
+nfl_advancedStatsPFR_rush_seasonal_raw <- progressr::with_progress(
   nflreadr::load_pfr_advstats(
@@
 
-nfl_advancedStatsPFR_rush_weekly <- progressr::with_progress(
+nfl_advancedStatsPFR_rush_weekly_raw <- progressr::with_progress(
   nflreadr::load_pfr_advstats(
@@
 
-nfl_advancedStatsPFR_rec_seasonal <- progressr::with_progress(
+nfl_advancedStatsPFR_rec_seasonal_raw <- progressr::with_progress(
   nflreadr::load_pfr_advstats(
@@
 
-nfl_advancedStatsPFR_rec_weekly <- progressr::with_progress(
+nfl_advancedStatsPFR_rec_weekly_raw <- progressr::with_progress(
   nflreadr::load_pfr_advstats(
@@
 
-nfl_advancedStatsPFR_def_seasonal <- progressr::with_progress(
+nfl_advancedStatsPFR_def_seasonal_raw <- progressr::with_progress(
   nflreadr::load_pfr_advstats(
@@
 
-nfl_advancedStatsPFR_def_weekly <- progressr::with_progress(
+nfl_advancedStatsPFR_def_weekly_raw <- progressr::with_progress(
   nflreadr::load_pfr_advstats(
@@
     summary_level = c("week")))
-
-nfl_advancedStatsPFR_seasonal <- bind_rows(
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_advancedStatsPFR_pass_seasonal_raw,
+  nfl_advancedStatsPFR_rush_seasonal_raw,
+  nfl_advancedStatsPFR_rec_seasonal_raw,
+  nfl_advancedStatsPFR_def_seasonal_raw,
+  file = "./data/nfl_advancedStatsPFR_seasonal_raw.RData"
+)
+
+save(
+  nfl_advancedStatsPFR_pass_weekly_raw,
+  nfl_advancedStatsPFR_rush_weekly_raw,
+  nfl_advancedStatsPFR_rec_weekly_raw,
+  nfl_advancedStatsPFR_def_weekly_raw,
+  file = "./data/nfl_advancedStatsPFR_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_advancedStatsPFR_seasonal_raw.RData")
+load(file = "./data/nfl_advancedStatsPFR_weekly_raw.RData")
+```
+
+```{r}
+nfl_advancedStatsPFR_pass_seasonal <- nfl_advancedStatsPFR_pass_seasonal_raw %>% 
+  mutate(
+    nameMerge = toupper(str_remove_all(player, "[[:punct:]]"))
+  ) %>% 
+  rename(
+    player_name.pass = player
+  )
+
+nfl_advancedStatsPFR_rush_seasonal <- nfl_advancedStatsPFR_rush_seasonal_raw %>% 
+  mutate(
+    nameMerge = toupper(str_remove_all(player, "[[:punct:]]"))
+  ) %>% 
+  rename(
+    player_name.rush = player,
+    age.rush = age,
+    team = tm
+  )
+
+nfl_advancedStatsPFR_rec_seasonal <- nfl_advancedStatsPFR_rec_seasonal_raw %>% 
+  mutate(
+    nameMerge = toupper(str_remove_all(player, "[[:punct:]]"))
+  ) %>% 
+  rename(
+    player_name.rec = player,
+    age.rec = age,
+    team = tm
+  )
+
+nfl_advancedStatsPFR_def_seasonal <- nfl_advancedStatsPFR_def_seasonal_raw %>% 
+  mutate(
+    nameMerge = toupper(str_remove_all(player, "[[:punct:]]"))
+  ) %>% 
+  rename(
+    player_name.def = player,
+    age.def = age,
+    team = tm
+  )
+
+nfl_advancedStatsPFR_pass_weekly <- nfl_advancedStatsPFR_pass_weekly_raw %>% 
+  mutate(
+    nameMerge = toupper(str_remove_all(pfr_player_name, "[[:punct:]]"))
+  ) %>% 
+  rename(
+    player_name.pass = pfr_player_name,
+    season.pass = season,
+    week.pass = week,
+    team.pass = team
+  )
+
+nfl_advancedStatsPFR_rush_weekly <- nfl_advancedStatsPFR_rush_weekly_raw %>% 
+  mutate(
+    nameMerge = toupper(str_remove_all(pfr_player_name, "[[:punct:]]"))
+  ) %>% 
+  rename(
+    player_name.rush = pfr_player_name,
+    season.rush = season,
+    week.rush = week,
+    team.rush = team
+  )
+
+nfl_advancedStatsPFR_rec_weekly <- nfl_advancedStatsPFR_rec_weekly_raw %>% 
+  mutate(
+    nameMerge = toupper(str_remove_all(pfr_player_name, "[[:punct:]]"))
+  ) %>% 
+  rename(
+    player_name.rec = pfr_player_name,
+    season.rec = season,
+    week.rec = week,
+    team.rec = team
+  )
+
+nfl_advancedStatsPFR_def_weekly <- nfl_advancedStatsPFR_def_weekly_raw %>% 
+  mutate(
+    nameMerge = toupper(str_remove_all(pfr_player_name, "[[:punct:]]"))
+  ) %>% 
+  rename(
+    player_name.def = pfr_player_name,
+    season.def = season,
+    week.def = week,
+    team.def = team
+  )
+
+nfl_advancedStatsPFR_seasonal_list <- list(
   nfl_advancedStatsPFR_pass_seasonal,
@@
   nfl_advancedStatsPFR_rec_seasonal,
-  nfl_advancedStatsPFR_def_seasonal
-)
-
-nfl_advancedStatsPFR_weekly <- bind_rows(
+  nfl_advancedStatsPFR_def_seasonal)
+
+nfl_advancedStatsPFR_weekly_list <- list(
   nfl_advancedStatsPFR_pass_weekly,
@@
   nfl_advancedStatsPFR_rec_weekly,
-  nfl_advancedStatsPFR_def_weekly
-)
+  nfl_advancedStatsPFR_def_weekly)
+
+nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal_list %>% 
+  purrr::reduce(
+    full_join,
+    by = c("pfr_id","nameMerge","season","team"))
+
+nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly_list %>% 
+  purrr::reduce(
+    full_join,
+    by = c("game_id","pfr_player_id")) #nameMerge
+
+#nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly_list %>% 
+#  purrr::reduce(
+#    full_join,
+#    by = c("pfr_player_id","nameMerge","season","week"))
+
+nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal %>% 
+  mutate(
+    player_name = coalesce(
+      player_name.pass,
+      player_name.rush,
+      player_name.rec,
+      player_name.def
+    ),
+    age = coalesce(
+      #age.pass,
+      age.rush,
+      age.rec,
+      age.def
+    )
+  ) %>% 
+  select(-c(starts_with("player_name.")))
+
+nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly %>% 
+  mutate(
+    player_name = coalesce(
+      player_name.pass,
+      player_name.rush,
+      player_name.rec,
+      player_name.def
+    ),
+    ,
+    season = coalesce(
+      season.pass,
+      season.rush,
+      season.rec,
+      season.def
+    ),
+    week = coalesce(
+      week.pass,
+      week.rush,
+      week.rec,
+      week.def
+    ),
+    team = coalesce(
+      team.pass,
+      team.rush,
+      team.rec,
+      team.def
+    )
+  ) %>% 
+  select(-c(starts_with("player_name.")))
+
+# Todo:
+#-separate variables from each pass/rush/rec/def object
+```
+
+The `nfl_advancedStatsPFR_seasonal` object is in `player`-`season`-`team` form.
+That is, each row should be uniquely identified by the combination of `pfr_id`, `season`, and `team`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal %>% 
+  select(pfr_id, season, team, player_name, everything()) %>% 
+  arrange(player_name, pfr_id, season, team)
+```
+
+Let's check for duplicate `player`-`season`-`team` instances:
+
+```{r}
+nfl_advancedStatsPFR_seasonal %>% 
+  group_by(pfr_id, season, team) %>% 
+  filter(n() > 1)
+```
+
+The `nfl_advancedStatsPFR_weekly` object is in both `game`-`player` form and `player`-`season`-`week` form.
+That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `week` or by the combination of `player_id`, `season`, `season_type`, and `week_num`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly %>% 
+  select(pfr_player_id, season, week, everything()) %>% 
+  arrange(player_name, pfr_player_id, season, week)
+```
+
+Let's check for duplicate `game`-`player` or `player`-`season`-`week` instances:
+
+```{r}
+nfl_advancedStatsPFR_weekly %>% 
+  arrange(game_id, pfr_player_id) %>% 
+  group_by(game_id, pfr_player_id) %>% 
+  filter(n() > 1)
+
+nfl_advancedStatsPFR_weekly %>% 
+  arrange(pfr_player_id, season, week) %>% 
+  group_by(pfr_player_id, season, week) %>% 
+  filter(n() > 1)
 ```
@@
 ```{r}
-#| eval: false
-
 # Prepare data for merging
@@
 
+Let's check again for duplicate `game`-`player` or `player`-`season`-`week` instances:
+
+```{r}
+nfl_advancedStatsPFR_seasonal %>% 
+  group_by(pfr_id, season, team) %>% 
+  filter(n() > 1)
+
+nfl_advancedStatsPFR_weekly %>% 
+  arrange(game_id, pfr_id) %>% 
+  group_by(game_id, pfr_id) %>% 
+  filter(n() > 1)
+
+nfl_advancedStatsPFR_weekly %>% 
+  arrange(pfr_id, season, week) %>% 
+  group_by(pfr_id, season, week) %>% 
+  filter(n() > 1)
+```
+
 ```{r}
@@
 
-nfl_playerContracts <- progressr::with_progress(
+nfl_playerContracts_raw <- progressr::with_progress(
   nflreadr::load_contracts())
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_playerContracts_raw,
+  file = "./data/nfl_playerContracts_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_playerContracts_raw.RData")
+```
+
+The `nfl_playerContracts` object is in `player`-`year`-`team`-`value` form.
+That is, each row should be uniquely identified by the combination of `otc_id`, `year_signed`, `team`, and `value`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_playerContracts <- nfl_playerContracts_raw %>% 
+  select(otc_id, year_signed, team, everything()) %>% 
+  arrange(player, otc_id, year_signed, team)
+```
+
+Let's check for duplicate `player`-`year`-`team`-`value` instances:
+
+```{r}
+nfl_playerContracts %>% 
+  group_by(otc_id, year_signed, team, value) %>% 
+  filter(n() > 1)
 ```
@@
 
-nfl_ftnCharting <- progressr::with_progress(
+nfl_ftnCharting_raw <- progressr::with_progress(
   nflreadr::load_ftn_charting(seasons = TRUE))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_ftnCharting_raw,
+  file = "./data/nfl_ftnCharting_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_ftnCharting_raw.RData")
+```
+
+The `nfl_ftnCharting` object is in `game`-`play` form.
+That is, each row should be uniquely identified by the combination of `nflverse_game_id` and `play_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_ftnCharting <- nfl_ftnCharting_raw %>% 
+  select(nflverse_game_id, nflverse_play_id, everything()) %>% 
+  arrange(nflverse_game_id, nflverse_play_id)
+```
+
+Let's check for duplicate `game`-`drive`-`play` instances:
+
+```{r}
+nfl_ftnCharting %>% 
+  group_by(nflverse_game_id, nflverse_play_id) %>% 
+  filter(n() > 1)
 ```
@@
 
-nfl_playerIDs <- progressr::with_progress(
+nfl_playerIDs_raw <- progressr::with_progress(
   nflreadr::load_ff_playerids())
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_playerIDs_raw,
+  file = "./data/nfl_playerIDs_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_playerIDs_raw.RData")
+```
+
+The `nfl_playerIDs` object is in `player` form.
+That is, each row should be uniquely identified by `mfl_id`.
+
+```{r}
+nfl_playerIDs <- nfl_playerIDs_raw %>% 
+  arrange(name, mfl_id)
+```
+
+Let's check for duplicate `player` instances:
+
+```{r}
+nfl_playerIDs %>% 
+  group_by(mfl_id) %>% 
+  filter(n() > 1)
+
+nfl_playerIDs %>% 
+  filter(!is.na(gsis_id)) %>% 
+  group_by(gsis_id) %>% 
+  filter(n() > 1)
 ```
@@
 
-#nfl_rankings <- progressr::with_progress( # currently throws error
+#nfl_rankings_raw <- progressr::with_progress( # currently throws error
 #  nflreadr::load_ff_rankings(type = "all"))
 
-nfl_rankings_draft <- progressr::with_progress(
+nfl_rankings_draft_raw <- progressr::with_progress(
   nflreadr::load_ff_rankings(type = "draft"))
 
-nfl_rankings_weekly <- progressr::with_progress(
+nfl_rankings_weekly_raw <- progressr::with_progress(
   nflreadr::load_ff_rankings(type = "week"))
-
-nfl_rankings <- bind_rows(
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_rankings_draft_raw,
+  file = "./data/nfl_rankings_draft_raw.RData"
+)
+
+save(
+  nfl_rankings_weekly_raw,
+  file = "./data/nfl_rankings_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_rankings_draft_raw.RData")
+load(file = "./data/nfl_rankings_weekly_raw.RData")
+```
+
+The `nfl_rankings_draft` object is in `player`-`page_type` form.
+That is, each row should be uniquely identified by the player's `id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_rankings_draft <- nfl_rankings_draft_raw %>% 
+  select(id, page_type, player, pos, team, everything()) %>% 
+  arrange(player, id, pos, page_type)
+```
+
+Let's check for duplicate `player`-`page_type` instances:
+
+```{r}
+nfl_rankings_draft %>% 
+  group_by(id, page_type) %>% 
+  filter(n() > 1)
+```
+
+The `nfl_rankings_weekly` object is in `player`-`page` form.
+That is, each row should be uniquely identified by `fantasypros_id` and `page`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_rankings_weekly <- nfl_rankings_weekly_raw %>% 
+  select(fantasypros_id, page, player_name, pos, team, everything()) %>% 
+  arrange(player_name, fantasypros_id, page, pos)
+```
+
+Let's check for duplicate `player`-`page` instances:
+
+```{r}
+nfl_rankings_weekly %>% 
+  group_by(fantasypros_id, page) %>% 
+  filter(n() > 1)
+```
+
+```{r}
+#| eval: false
+
+save(
   nfl_rankings_draft,
-  nfl_rankings_weekly
-)
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_rankings,
-  file = "./data/nfl_rankings.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_rankings.RData")
+  file = "./data/nfl_rankings_draft.RData"
+)
+
+save(
+  nfl_rankings_weekly,
+  file = "./data/nfl_rankings_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_rankings_draft.RData")
+load(file = "./data/nfl_rankings_weekly.RData")
 ```
@@
 
-nfl_expectedFantasyPoints_weekly <- progressr::with_progress(
+nfl_expectedFantasyPoints_weekly_raw <- progressr::with_progress(
   nflreadr::load_ff_opportunity(
@@
 
-nfl_expectedFantasyPoints_pass <- progressr::with_progress(
+nfl_expectedFantasyPoints_pass_raw <- progressr::with_progress(
   nflreadr::load_ff_opportunity(
@@
 
-nfl_expectedFantasyPoints_rush <- progressr::with_progress(
+nfl_expectedFantasyPoints_rush_raw <- progressr::with_progress(
   nflreadr::load_ff_opportunity(
@@
   ))
-
-nfl_expectedFantasyPoints_weekly$season <- as.integer(nfl_expectedFantasyPoints_weekly$season)
-
-nfl_expectedFantasyPoints_pbp <- bind_rows(
-  nfl_expectedFantasyPoints_pass,
-  nfl_expectedFantasyPoints_rush
-)
-```
-
-```{r}
-#| eval: false
-
+```
+
+```{r}
+nfl_expectedFantasyPoints_weekly_raw <- nflreadr::load_ff_opportunity(
+  seasons = TRUE,
+  stat_type = "weekly",
+  model_version = "latest"
+  )
+
+nfl_expectedFantasyPoints_pass_raw <- nflreadr::load_ff_opportunity(
+  seasons = TRUE,
+  stat_type = "pbp_pass",
+  model_version = "latest"
+  )
+
+nfl_expectedFantasyPoints_rush_raw <- nflreadr::load_ff_opportunity(
+  seasons = TRUE,
+  stat_type = "pbp_rush",
+  model_version = "latest"
+  )
+```
+
+```{r}
+save(
+  nfl_expectedFantasyPoints_weekly_raw,
+  file = "./data/nfl_expectedFantasyPoints_weekly_raw.RData"
+)
+
+save(
+  nfl_expectedFantasyPoints_pass_raw,
+  nfl_expectedFantasyPoints_rush_raw,
+  file = "./data/nfl_expectedFantasyPoints_pbp_raw.RData"
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_expectedFantasyPoints_weekly_raw.RData")
+load(file = "./data/nfl_expectedFantasyPoints_pbp_raw.RData")
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly_raw.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp_raw.RData", fsep = ""))
+```
+
+```{r}
+nfl_expectedFantasyPoints_pbp_list <- list(
+  nfl_expectedFantasyPoints_pass_raw,
+  nfl_expectedFantasyPoints_rush_raw)
+
+nfl_expectedFantasyPoints_pbp <- full_join(
+  nfl_expectedFantasyPoints_pass_raw,
+  nfl_expectedFantasyPoints_rush_raw,
+  by = c("game_id","fixed_drive","play_id"),
+  suffix = c(".pass", ".rush")
+)
+
+# Todo:
+#-separate variables from each pass/rush object
+```
+
+The `nfl_expectedFantasyPoints_weekly` object is in `game`-`player` form and in `season`-`week`-`player` form.
+That is, each row should be uniquely identified by the combination of `game_id` and `player_id`.
+Each row should also be uniquely identified by the combination of `season`, `week`, and player_id.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly_raw %>% 
+  select(season, week, game_id, player_id, full_name, posteam, position, everything()) %>% 
+  arrange(full_name, player_id, season, week)
+```
+
+Let's check for duplicate `game`-`player` instances and `season`-`week`-`player` instances:
+
+```{r}
+nfl_expectedFantasyPoints_weekly %>% 
+  group_by(game_id, player_id) %>% 
+  filter(n() > 1)
+
+nfl_expectedFantasyPoints_weekly %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1)
+```
+
+Let's do some data cleanup:
+
+```{r}
+# Drop players with missing values for player_id
+nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
+  filter(!is.na(player_id))
+```
+
+Let's check again for duplicate `game`-`player` instances and `season`-`week`-`player` instances:
+
+```{r}
+nfl_expectedFantasyPoints_weekly %>% 
+  group_by(game_id, player_id) %>% 
+  filter(n() > 1)
+
+nfl_expectedFantasyPoints_weekly %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1)
+```
+
+The `nfl_expectedFantasyPoints_pbp` object is in `game`-`drive`-`play` form.
+That is, each row should be uniquely identified by the combination of `game_id`, `fixed_drive`, and `play_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_expectedFantasyPoints_pbp <- nfl_expectedFantasyPoints_pbp %>% 
+  select(game_id, fixed_drive, play_id, everything()) %>% 
+  arrange(game_id, fixed_drive, play_id)
+```
+
+Let's check for duplicate `game`-`player` instances and `season`-`week`-`player` instances:
+
+```{r}
+nfl_expectedFantasyPoints_pbp %>% 
+  group_by(game_id, fixed_drive, play_id) %>% 
+  filter(n() > 1)
+```
+
+```{r}
 save(
@@
 ```{r}
+#| eval: false
 #| include: false
@@
 load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
 ```
@@
 
-nfl_actualStats_offense_career <- nflfastR::calculate_player_stats(
+nfl_actualStats_offense_career_raw <- nflfastR::calculate_player_stats(
   nfl_pbp,
@@
 
-nfl_actualStats_defense_career <- nflfastR::calculate_player_stats_def(
+nfl_actualStats_defense_career_raw <- nflfastR::calculate_player_stats_def(
   nfl_pbp,
@@
 
-nfl_actualStats_kicking_career <- nflfastR::calculate_player_stats_kicking(
+nfl_actualStats_kicking_career_raw <- nflfastR::calculate_player_stats_kicking(
   nfl_pbp,
   weekly = FALSE)
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_offense_career_raw, nfl_actualStats_defense_career_raw, nfl_actualStats_kicking_career_raw,
+  file = "./data/nfl_actualStats_career_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load("./data/nfl_actualStats_career_raw.RData")
+```
+
+The `nfl_actualStats_offense_career`, `nfl_actualStats_defense_career`, and `nfl_actualStats_kicking_career` objects are in `player` form.
+That is, each row should be uniquely identified by the combination of `player_id`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_offense_career <- nfl_actualStats_offense_career_raw %>% 
+  arrange(player_display_name, player_id)
+
+nfl_actualStats_defense_career <- nfl_actualStats_defense_career_raw %>% 
+  arrange(player_display_name, player_id)
+
+nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career_raw %>% 
+  arrange(player_display_name, player_id)
+```
+
+Let's check for duplicate `player` instances:
+
+```{r}
+nfl_actualStats_offense_career %>% 
+  group_by(player_id) %>% 
+  filter(n() > 1)
+
+nfl_actualStats_defense_career %>% 
+  group_by(player_id) %>% 
+  filter(n() > 1)
+
+nfl_actualStats_kicking_career %>% 
+  group_by(player_id) %>% 
+  filter(n() > 1)
+```
+
+Let's do some data cleanup:
+
+```{r}
+# Sum statistics across a player's years with multiple teams
+nfl_actualStats_defense_career <- nfl_actualStats_defense_career %>% 
+  group_by(player_id) %>% 
+  summarise(
+    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
+    .groups = "drop"
+  )
+
+nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career %>% 
+  group_by(player_id) %>% 
+  summarise(
+    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
+    .groups = "drop"
+  )
+
+# Re-calculate percentage stats
+nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career %>% 
+  mutate(
+    fg_pct = fg_made / fg_att,
+    pat_pct = pat_made / pat_att
+  )
+
+# Merge data back with player info
+nfl_actualStats_defense_career_raw_playerInfo <- nfl_actualStats_defense_career_raw %>% 
+  select(player_id, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
+  unique() %>% 
+  group_by(player_id) %>% 
+  slice_head(n = 1) %>% 
+  ungroup()
+
+nfl_actualStats_kicking_career_raw_playerInfo <- nfl_actualStats_kicking_career_raw %>% 
+  select(player_id, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
+  unique() %>% 
+  group_by(player_id) %>% 
+  slice_head(n = 1) %>% 
+  ungroup()
+
+nfl_actualStats_defense_career <- nfl_actualStats_defense_career_raw_playerInfo %>% 
+  right_join(
+    nfl_actualStats_defense_career,
+    by = "player_id"
+  )
+
+nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career_raw_playerInfo %>% 
+  right_join(
+    nfl_actualStats_kicking_career,
+    by = "player_id"
+  )
+
+# Rearrange data
+nfl_actualStats_defense_career <- nfl_actualStats_defense_career %>% 
+  arrange(player_display_name, player_id)
+
+nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career %>% 
+  arrange(player_display_name, player_id)
+```
+
+Let's check again for duplicate `player` instances:
+
+```{r}
+nfl_actualStats_offense_career %>% 
+  group_by(player_id) %>% 
+  filter(n() > 1)
+
+nfl_actualStats_defense_career %>% 
+  group_by(player_id) %>% 
+  filter(n() > 1)
+
+nfl_actualStats_kicking_career %>% 
+  group_by(player_id) %>% 
+  filter(n() > 1)
 ```

### 2024-08-02 getting-started.qmd
20240801 - fix error
20240801 - print less to screen to reduce resource usage

@@
 load(file = "./data/nfl_rankings_weekly_raw.RData")
-load(file = "./data/nfl_expectedFantasyPoints_weekly_raw.RData")
-load(file = "./data/nfl_expectedFantasyPoints_pbp_raw.RData")
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly_raw.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp_raw.RData", fsep = ""))
 
@@
 load(file = "./data/nfl_rankings_weekly.RData")
-load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
-load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
 
@@
   group_by(gsis_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(team_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(gsis_id, season, team) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(gsis_id, season, week) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(game_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   filter(n() > 1, !is.na(pfr_id)) %>% 
-  arrange(pfr_id)
+  arrange(pfr_id) %>% 
+  head()
 
@@
   filter(n() > 1, !is.na(cfb_id)) %>% 
-  arrange(cfb_id)
+  arrange(cfb_id) %>% 
+  head()
 ```
@@
   group_by(gsis_id, season, week, depth_position) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(game_id, drive, play_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(nflverse_game_id, drive, play_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(player_id, season, week) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 
@@
   group_by(player_id, season, week) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(game_id, pfr_player_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(player_id, season, season_type) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(game_id, player_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 
@@
   group_by(player_id, season, season_type, week_num) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
 
-load(file = "./data/nfl_espnQBR.RData")
+load(file = "./data/nfl_espnQBR_seasonal.RData")
+load(file = "./data/nfl_espnQBR_weekly.RData")
 ```
@@
   group_by(player_gsis_id, season, season_type, week) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(pfr_id, season, team) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(game_id, pfr_player_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 
@@
   group_by(pfr_player_id, season, week) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   filter(n() > 1) %>% 
-  arrange(pfr_id, season)
+  arrange(pfr_id, season) %>% 
+  head()
 
@@
   group_by(pfr_id, season, team) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 
@@
   group_by(game_id, pfr_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 
@@
   group_by(pfr_id, season, week) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(otc_id, year_signed, team, value) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(nflverse_game_id, nflverse_play_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(mfl_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 
@@
   group_by(id, page_type) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(fantasypros_id, page) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(game_id, fixed_drive, play_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   group_by(player_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 
@@
   group_by(player_id) %>% 
-  filter(n() > 1)
+  filter(n() > 1) %>% 
+  head()
 ```

### 2024-08-02 getting-started.qmd
20240801 - garbage collection

@@
 ```{r}
+#| include: false
+
+# garbage collection
+rm(nfl_pbp_raw)
+gc()
+```
+
+```{r}
 #| eval: false
@@
 )
+```
+
+```{r}
+#| include: false
+
+# garbage collection
+rm(nfl_4thdown_raw)
+rm(nfl_4thdown)
+gc()
 ```
@@
 )
+```
+
+```{r}
+#| include: false
+
+# garbage collection
+rm(nfl_participation_raw)
+rm(nfl_participation)
+gc()
 ```
@@
 ```{r}
+#| include: false
+
+# garbage collection
+rm(nfl_expectedFantasyPoints_weekly_raw)
+rm(nfl_expectedFantasyPoints_pbp_raw)
+rm(nfl_expectedFantasyPoints_weekly)
+rm(nfl_expectedFantasyPoints_pbp)
+gc()
+```
+
+```{r}
 #| eval: false
@@
 
+```{r}
+#| include: false
+
+# garbage collection
+rm(nfl_pbp)
+gc()
+```
+
 ### Historical Actual Fantasy Points {#sec-calculateActualPoints}

### 2024-08-02 getting-started.qmd
20240802 - move player IDs earlier so they can be used for later code
20240802 - actual stats

@@
 
+### Fantasy Player IDs {#sec-fantasyPlayerIDs}
+
+A Data Dictionary for fantasy player ID data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_playerids.html>
+
+```{r}
+#| eval: false
+
+nfl_playerIDs_raw <- progressr::with_progress(
+  nflreadr::load_ff_playerids())
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_playerIDs_raw,
+  file = "./data/nfl_playerIDs_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_playerIDs_raw.RData")
+```
+
+The `nfl_playerIDs` object is in `player` form.
+That is, each row should be uniquely identified by `mfl_id`.
+
+```{r}
+nfl_playerIDs <- nfl_playerIDs_raw %>% 
+  arrange(name, mfl_id)
+```
+
+Let's check for duplicate `player` instances:
+
+```{r}
+nfl_playerIDs %>% 
+  group_by(mfl_id) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_playerIDs %>% 
+  filter(!is.na(gsis_id)) %>% 
+  group_by(gsis_id) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_playerIDs,
+  file = "./data/nfl_playerIDs.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_playerIDs.RData")
+```
+
 ### Player Info {#sec-downloadPlayerInfo}
@@
 
-### Fantasy Player IDs {#sec-fantasyPlayerIDs}
-
-A Data Dictionary for fantasy player ID data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_playerids.html>
-
-```{r}
-#| eval: false
-
-nfl_playerIDs_raw <- progressr::with_progress(
-  nflreadr::load_ff_playerids())
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_playerIDs_raw,
-  file = "./data/nfl_playerIDs_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_playerIDs_raw.RData")
-```
-
-The `nfl_playerIDs` object is in `player` form.
-That is, each row should be uniquely identified by `mfl_id`.
-
-```{r}
-nfl_playerIDs <- nfl_playerIDs_raw %>% 
-  arrange(name, mfl_id)
-```
-
-Let's check for duplicate `player` instances:
-
-```{r}
-nfl_playerIDs %>% 
-  group_by(mfl_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_playerIDs %>% 
-  filter(!is.na(gsis_id)) %>% 
-  group_by(gsis_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_playerIDs,
-  file = "./data/nfl_playerIDs.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_playerIDs.RData")
-```
-
 ### FantasyPros Rankings {#sec-fantasyProsRankings}
@@
 
-The `nfl_actualStats_offense_career`, `nfl_actualStats_defense_career`, and `nfl_actualStats_kicking_career` objects are in `player` form.
+The `nfl_actualStats_career` objects are in `player` form.
 That is, each row should be uniquely identified by the combination of `player_id`.
@@
 
-nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonalList %>% 
+nfl_actualStats_offense_seasonal_raw <- nfl_actualStats_offense_seasonalList %>% 
   dplyr::bind_rows()
-nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonalList %>% 
+nfl_actualStats_defense_seasonal_raw <- nfl_actualStats_defense_seasonalList %>% 
   dplyr::bind_rows()
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonalList %>% 
+nfl_actualStats_kicking_seasonal_raw <- nfl_actualStats_kicking_seasonalList %>% 
   dplyr::bind_rows()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_offense_seasonal_raw, nfl_actualStats_defense_seasonal_raw, nfl_actualStats_kicking_seasonal_raw,
+  file = "./data/nfl_actualStats_seasonal_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load("./data/nfl_actualStats_seasonal_raw.RData")
+```
+
+The `nfl_actualStats_seasonal` objects are in `player`-`season` form.
+That is, each row should be uniquely identified by the combination of `player_id` and `season`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal_raw %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal_raw %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal_raw %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+```
+
+Let's check for duplicate `player` instances:
+
+```{r}
+nfl_actualStats_offense_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_defense_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_kicking_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+Let's do some data cleanup:
+
+```{r}
+# Sum statistics across a player's years with multiple teams
+nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal %>% 
+  group_by(player_id, season) %>% 
+  summarise(
+    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
+    .groups = "drop"
+  )
+
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
+  group_by(player_id, season) %>% 
+  summarise(
+    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
+    .groups = "drop"
+  )
+
+# Re-calculate percentage stats
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
+  mutate(
+    fg_pct = fg_made / fg_att,
+    pat_pct = pat_made / pat_att
+  )
+
+# Merge data back with player info
+nfl_actualStats_defense_seasonal_raw_playerInfo <- nfl_actualStats_defense_seasonal_raw %>% 
+  select(player_id, season, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
+  unique() %>% 
+  group_by(player_id) %>% 
+  slice_head(n = 1) %>% 
+  ungroup()
+
+nfl_actualStats_kicking_seasonal_raw_playerInfo <- nfl_actualStats_kicking_seasonal_raw %>% 
+  select(player_id, season, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
+  unique() %>% 
+  group_by(player_id) %>% 
+  slice_head(n = 1) %>% 
+  ungroup()
+
+nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal_raw_playerInfo %>% 
+  right_join(
+    nfl_actualStats_defense_seasonal,
+    by = c("player_id","season")
+  )
+
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal_raw_playerInfo %>% 
+  right_join(
+    nfl_actualStats_kicking_seasonal,
+    by = c("player_id","season")
+  )
+
+# Rearrange data
+nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+```
+
+Let's check again for duplicate `player` instances:
+
+```{r}
+nfl_actualStats_offense_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_defense_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_kicking_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
 ```
@@
   )
-
-# Arrange data
+```
+
+The `player_stats_weekly` objects are in `player`-`season`-`week` form.
+That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
+Let's rearrange the data accordingly:
+
+```{r}
 player_stats_weekly_offense <- player_stats_weekly_offense %>% 
-  arrange(player_display_name, season, week)
+  arrange(player_display_name, player_id, season, week)
 
 player_stats_weekly_defense <- player_stats_weekly_defense %>% 
-  arrange(player_display_name, season, week)
+  arrange(player_display_name, player_id, season, week)
 
 player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
-  arrange(player_display_name, season, week)
+  arrange(player_display_name, player_id, season, week)
+```
+
+Let's check for duplicate `player`-`season`-`week` instances:
+
+```{r}
+player_stats_weekly_offense %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+player_stats_weekly_defense %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1) %>% 
+  head() #todo: fix duplicate instances
+
+player_stats_weekly_kicking %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1) %>% 
+  head() #todo: fix duplicate instances
 ```
@@
   )
-
-# Arrange data
+```
+
+The `player_stats_seasonal` objects are in `player`-`season` form.
+That is, each row should be uniquely identified by the combination of `player_id` and `season`.
+Let's rearrange the data accordingly:
+
+```{r}
 player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
-  relocate(season, .after = player_id) %>% 
-  arrange(player_display_name, season)
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
 
 player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  relocate(season, .after = player_id) %>% 
-  arrange(player_display_name, season)
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
 
 player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  relocate(season, .after = player_id) %>% 
-  arrange(player_display_name, season)
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+```
+
+Let's check for duplicate `player`-`season` instances:
+
+```{r}
+player_stats_seasonal_offense %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+player_stats_seasonal_defense %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+player_stats_seasonal_kicking %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
 ```

### 2024-08-02 download-football-data.qmd
20240802 - separate chapter for downloading football data

--- 2024-08-02 getting-started.qmd
@@
-# Getting Started with `R` for Data Analysis {#sec-gettingStarted}
-
-The book uses `R` for statistical analyses (<http://www.r-project.org>).
-R is a free software environment; you can download it at no charge here: <https://cran.r-project.org>.
-
-## Initial Setup {#sec-initialSetup}
-
-To get started, follow the following steps:
-
-1. Install `R`: <https://cran.r-project.org>
-1. Install `RStudio Desktop`: <https://posit.co/download/rstudio-desktop>
-1. After installing `RStudio`, open `RStudio` and run the following code in the console to install several key `R` packages:
-   ```{r}
-   #| eval: false
-
-   install.packages(
-     c("petersenlab","remotes","nflreadr","nflfastR","nfl4th","nflplotR",
-     "gsisdecoder","progressr","lubridate","tidyverse","psych"))
-   ```
-1. Some necessary packages, including the `ffanalytics` package, are hosted in GitHub and need to be installed using the following code (after installing the `remotes` package above):
-   ```{r}
-   #| eval: false
-
-   remotes::install_github("FantasyFootballAnalytics/ffanalytics")
-   ```
-
-::: {#nte-gettingStarted .callout-note title="If you are in Dr. Petersen's class"}
-If you are in Dr. Petersen's class, also perform the following steps:
-
-1. Set up a free account on [GitHub.com](https://github.com).
-1. Download GitHub Desktop: <https://desktop.github.com>
-1. Make sure you are logged into your GitHub account on [GitHub.com](https://github.com).
-1. Go to the following GitHub repository: <https://github.com/isaactpetersen/QuartoBlogFantasyFootball> and complete the following steps:
-    1. Click "Use this Template" (in the top right of the screen) > "Create a new repository"
-    1. Make sure the checkbox is selected for the following option: "Include all branches"
-    1. Make sure your Owner account is selected
-    1. Specify the repository name to whatever you want, such as `FantasyFootballBlog`
-    1. Type a brief description, such as `Files for my fantasy football blog`
-    1. Keep the repository public (this is necessary for generating your blog)
-    1. Select "Create repository"
-1. After creating the new repository, make sure you are on the page of of your new repository and complete the following steps:
-    1. Click "Settings" (in the topof the screen)
-    1. Click "Actions" (in the left sidebar) > "General"
-    1. Make sure the following are selected:
-        - "Read and write permissions" (under "Workflow permissions")
-        - "Allow GitHub Actions to create and approve pull requests"
-        - then click "Save"
-    1. Click "Pages" (in the left sidebar)
-    1. Make sure the following are selected:
-        - "Deploy from a branch" (under "Source")
-        - "gh-pages/(root)" (under "Branch")
-        - then click "Save"
-1. Clone the repository to your local computer by clicking "Code" > "Open with GitHub Desktop", select the folder where you want the repository to be saved on your local computer, and click "Clone"
-:::
-
-## Installing Packages {#sec-installingPackages}
-
-You can install `R` packages using the following syntax:
-
-```{r}
-#| eval: false
-
-install.packages("INSERT_PACKAGE_NAME_HERE")
-```
-
-For instance, you can use the following code to install the `nflreadr` package:
-
-```{r}
-#| eval: false
-
-install.packages("nflreadr")
-```
+# Download and Process NFL Football Data {#sec-downloadFootballData}
 
@@
 
-## Using Functions and Arguments {#sec-functionsArguments}
-
-You can learn about a particular function and its arguments by entering a question mark before the name of the function:
-
-```{r}
-#| eval: false
-
-?NAME_OF_FUNCTION()
-```
-
-Below, we provide examples for how to learn about and use functions and arguments, by using the `seq()` function as an example.
-The `seq()` function creates a sequence of numbers.
-To learn about the `seq()` function, which creates a sequence of numbers, you can execute the following command:
-
-```{r}
-#| eval: false
-
-?seq()
-```
-
-This is what the documentation shows for the `seq()` function in the `Usage` section:
-
-```{r}
-#| eval: false
-
-seq(
-  from = 1,
-  to = 1,
-  by = ((to - from)/(length.out - 1)),
-  length.out = NULL,
-  along.with = NULL,
-  ...)
-```
-
-Based on this information, we know that the `seq()` function takes the following arguments:
-
-- `from`
-- `to`
-- `by`
-- `length.out`
-- `along.with`
-- `...`
-
-The arguments have default values that are used if the user does not specify values for the arguments.
-The default values are provided in the `Usage` section and are in @tbl-seqFunction:
-
-| Argument     | Default Value for Argument       |
-|:-------------|:---------------------------------|
-| `from`       | `1`                              |
-| `to`         | `1`                              |
-| `by`         | `((to - from)/(length.out - 1))` |
-| `length.out` | `NULL`                           |
-| `along.with` | `NULL`                           |
-
-: Arguments and defaults for the `seq()` function. Arguments with a default of `NULL` are not used unless a value is provided by the user. {#tbl-seqFunction}
-
-What each argument represents (i.e., the meaning of `from`, `to`, `by`, etc.) is provided in the `Arguments` section of the documentation.
-You can specify a function and its arguments either by providing values for each argument in the order indicated by the function, or by naming its arguments.
-
-Here is an example of providing values to the arguments in the order indicated by the function, to create a sequence of numbers from 1 to 9:
-
-```{r}
-seq(1, 9)
-```
-
-Here is an example of providing values to the arguments by naming its arguments:
-
-```{r}
-seq(
-  from = 1,
-  to = 9,
-  by = 1)
-```
-
-If you provide values to arguments by naming the arguments, you can reorder the arguments and get the same answer:
-
-```{r}
-seq(
-  by = 1,
-  to = 9,
-  from = 1)
-```
-
-There are various combinations of arguments that one could use to obtain the same result.
-For instance, here is code to generate a sequence from 1 to 9 by 2:
-
-```{r}
-seq(
-  from = 1,
-  to = 9,
-  by = 2)
-```
-
-Or, alternatively, you could specify the length of the desired sequence (5 values):
-
-```{r}
-seq(
-  from = 1,
-  to = 9,
-  length.out = 5)
-```
-
-If you want to generate a series with decimal values, you could specify a long desired sequence of 81 values:
-
-```{r}
-seq(
-  from = 1,
-  to = 9,
-  length.out = 81)
-```
-
-This is equivalent to specifying a sequence from 1 to 9 by 0.1:
-
-```{r}
-seq(
-  from = 1,
-  to = 9,
-  by = 0.1)
-```
-
-Hopefully, that provides an example for how to learn about a particular function, its arguments, and how to use them.
-
-## Download Football Data {#sec-downloadFootballData}
-
-Below, we provide examples for how to download various types of National Football League (NFL) data.
+## Data Dictionaries of NFL Data {#sec-dataDictionary}
+
+Data Dictionaries are metadata that describe the meaning of the variables in a dataset.
+You can find Data Dictionaries for the various National Football League (NFL) datasets at the following link: <https://nflreadr.nflverse.com/articles/index.html>.
+
+## Types of NFL Data {#sec-downloadFootballDataOverview}
+
+Below, we provide examples for how to download various types of NFL data.
 For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
@@
 
-## Data Dictionary {#sec-dataDictionary}
-
-Data Dictionaries are metadata that describe the meaning of the variables in a datset.
-You can find Data Dictionaries for the various NFL datasets at the following link: <https://nflreadr.nflverse.com/articles/index.html>.
-
-## Create a Data Frame {#sec-createDF}
-
-Here is an example of creating a data frame:
-
-```{r}
-players <- data.frame(
-  ID = 1:12,
-  name = c(
-    "Ken Cussion",
-    "Ben Sacked",
-    "Chuck Downfield",
-    "Ron Ingback",
-    "Rhonda Ball",
-    "Hugo Long",
-    "Lionel Scrimmage",
-    "Drew Blood",
-    "Chase Emdown",
-    "Justin Time",
-    "Spike D'Ball",
-    "Isac Ulooz"),
-  position = c("QB","QB","QB","RB","RB","WR","WR","WR","WR","TE","TE","LB"),
-  age = c(40, 30, 24, 20, 18, 23, 27, 32, 26, 23, NA, 37)
-  )
-
-fantasyPoints <- data.frame(
-  ID = c(2, 7, 13, 14),
-  fantasyPoints = c(250, 170, 65, 15)
-)
-```
-
-## Variable Names {#sec-variableNames}
-
-To see the names of variables in a data frame, use the following syntax:
-
-```{r}
-names(nfl_players)
-
-names(players)
-
-names(fantasyPoints)
-```
-
-## Logical Operators {#sec-logicalOperators}
-
-### Is Equal To: `==` {#sec-logicalOperatorEqual}
-
-```{r}
-players$position == "RB"
-```
-
-### Is Not Equal To: `!=` {#sec-logicalOperatorNotEqual}
-
-```{r}
-players$position != "RB"
-```
-
-### Is Greater Than: `>` {#sec-logicalOperatorGreaterThan}
-
-```{r}
-players$age > 30
-```
-
-### Is Less Than: `<` {#sec-logicalOperatorLessThan}
-
-```{r}
-players$age < 30
-```
-
-### Is Greater Than or Equal To: `>=` {#sec-logicalOperatorGreaterThanEqual}
-
-```{r}
-players$age >= 30
-```
-
-### Is Less Than or Equal To: `<=` {#sec-logicalOperatorLessThanEqual}
-
-```{r}
-players$age <= 30
-```
-
-### Is In a Value of Another Vector: `%in%` {#sec-logicalOperatorInVector}
-
-```{r}
-players$position %in% c("RB","WR")
-```
-
-### Is Not In a Value of Another Vector: `!(%in%)` {#sec-logicalOperatorNotInVector}
-
-```{r}
-!(players$position %in% c("RB","WR"))
-```
-
-### Is Missing: `is.na()` {#sec-logicalOperatorMissing}
-
-```{r}
-is.na(players$age)
-```
-
-### Is Not Missing: `!is.na()` {#sec-logicalOperatorNotMissing}
-
-```{r}
-!is.na(players$age)
-```
-
-### And: `&`{#sec-logicalOperatorAnd}
-
-```{r}
-players$position == "WR" & players$age > 26
-```
-
-### Or: `|` {#sec-logicalOperatorOr}
-
-```{r}
-players$position == "WR" | players$age > 23
-```
-
-## Subset {#sec-subset}
-
-To subset a data frame, use brackets to specify the subset of rows and columns to keep, where the value/vector before the comma specifies the rows to keep, and the value/vector after the comma specifies the columns to keep:
-
-```{r}
-#| eval: false
-
-dataframe[rowsToKeep, columnsToKeep]
-```
-
-You can subset by using any of the following:
-
-- numeric indices of the rows/columns to keep (or drop)
-- names of the rows/columns to keep (or drop)
-- values of `TRUE` and `FALSE` corresponding to which rows/columns to keep
-
-### One Variable {#sec-subsetOneVariable}
-
-To subset one variable, use the following syntax:
-
-```{r}
-players$name
-```
-
-or:
-
-```{r}
-players[,"name"]
-```
-
-### Particular Rows of One Variable {#sec-subsetParticularRowsOneVariable}
-
-To subset one variable, use the following syntax:
-
-```{r}
-players$name[which(players$position == "RB")]
-```
-
-or:
-
-```{r}
-players[which(players$position == "RB"), "name"]
-```
-
-### Particular Columns (Variables) {#sec-subsetParticularColumns}
-
-To subset particular columns/variables, use the following syntax:
-
-#### Base `R` {#sec-subsetParticularColumnsBaseR}
-
-```{r}
-subsetVars <- c("name","age")
-
-players[,c(2,4)]
-players[,c("name","age")]
-players[,subsetVars]
-```
-
-Or, to drop columns:
-
-```{r}
-dropVars <- c("name","age")
-
-players[,-c(2,4)]
-players[,!(names(players) %in% c("name","age"))]
-players[,!(names(players) %in% dropVars)]
-```
-
-#### Tidyverse {#sec-subsetParticularColumnsTidyverse}
-
-```{r}
-players %>%
-  select(name, age)
-
-players %>%
-  select(name:age)
-
-players %>%
-  select(all_of(subsetVars))
-```
-
-Or, to drop columns:
-
-```{r}
-players %>%
-  select(-name, -age)
-
-players %>%
-  select(-c(name:age))
-
-players %>%
-  select(-all_of(dropVars))
-```
-
-### Particular Rows {#sec-subsetParticularRows}
-
-To subset particular rows, use the following syntax:
-
-#### Base `R` {#sec-subsetParticularRowsBaseR}
-
-```{r}
-subsetRows <- c(4,5)
-
-players[c(4,5),]
-players[subsetRows,]
-players[which(players$position == "RB"),]
-```
-
-#### Tidyverse {#sec-subsetParticularRowsTidyverse}
-
-```{r}
-players %>%
-  filter(position == "WR")
-
-players %>%
-  filter(position == "WR", age <= 26)
-
-players %>%
-  filter(position == "WR" | age >= 26)
-```
-
-### Particular Rows and Columns {#sec-subsetParticularRowsAndColumns}
-
-To subset particular rows and columns, use the following syntax:
-
-#### Base `R` {#sec-subsetParticularRowsAndColumnsBaseR}
-
-```{r}
-players[c(4,5), c(2,4)]
-players[subsetRows, subsetVars]
-players[which(players$position == "RB"), subsetVars]
-```
-
-#### Tidyverse {#sec-subsetParticularRowsAndColumnsTidyverse}
-
-```{r}
-players %>%
-  filter(position == "RB") %>%
-  select(all_of(subsetVars))
-```
-
-## View Data {#sec-viewData}
-
-### All Data {#sec-viewDataAll}
-
-To view data, use the following syntax:
-
-```{r}
-#| eval: false
-
-View(players)
-```
-
-### First 6 Rows/Elements {#sec-viewDataHead}
-
-To view only the first six rows (if a data frame) or elements (if a vector), use the following syntax:
-
-```{r}
-head(nfl_players)
-head(nfl_players$display_name)
-```
-
-## Data Characteristics {#sec-dataCharacteristics}
-
-### Data Structure {#sec-dataCharacteristicsStructure}
-
-```{r}
-str(nfl_players)
-```
-
-### Data Dimensions {#sec-dataCharacteristicsDimensions}
-
-Number of rows and columns:
-
-```{r}
-dim(nfl_players)
-```
-
-Number of rows:
-
-```{r}
-nrow(nfl_players)
-```
-
-Number of columns:
-
-```{r}
-ncol(nfl_players)
-```
-
-### Number of Elements {#sec-dataCharacteristicsLength}
-
-```{r}
-length(nfl_players$display_name)
-```
-
-### Number of Missing Elements {#sec-dataCharacteristicsMissingElements}
-
-```{r}
-length(nfl_players$college_name[which(is.na(nfl_players$college_name))])
-```
-
-### Number of Non-Missing Elements {#sec-dataCharacteristicsNonMissingElements}
-
-```{r}
-length(nfl_players$college_name[which(!is.na(nfl_players$college_name))])
-length(na.omit(nfl_players$college_name))
-```
-
-## Create New Variables {#sec-createNewVars}
-
-To create a new variable, use the following syntax:
-
-```{r}
-players$newVar <- NA
-```
-
-Here is an example of creating a new variable:
-
-```{r}
-players$newVar <- 1:nrow(players)
-```
-
-```{r}
-#| include: false
-
-players$newVar <- NULL
-```
-
-## Recode Variables {#sec-recodeVars}
-
-Here is an example of recoding a variable:
-
-```{r}
-players$oldVar1 <- NA
-players$oldVar1[which(players$position == "QB")] <- "quarterback"
-players$oldVar1[which(players$position == "RB")] <- "running back"
-players$oldVar1[which(players$position == "WR")] <- "wide receiver"
-players$oldVar1[which(players$position == "TE")] <- "tight end"
-
-players$oldVar2 <- NA
-players$oldVar2[which(players$age < 30)] <- "young"
-players$oldVar2[which(players$age >= 30)] <- "old"
-```
-
-Recode multiple variables:
-
-```{r}
-players %>%
-  mutate(across(c(
-    oldVar1:oldVar2),
-    ~ case_match(
-      .,
-      c("quarterback","old","running back") ~ 0,
-      c("wide receiver","tight end","young") ~ 1)))
-```
-
-## Rename Variables {#renameVars}
-
-```{r}
-players <- players %>% 
-  rename(
-    newVar1 = oldVar1,
-    newVar2 = oldVar2)
-```
-
-Using a vector of variable names:
-
-```{r}
-#| eval: false
-
-varNamesFrom <- c("oldVar1","oldVar2")
-varNamesTo <- c("newVar1","newVar2")
-
-players <- players %>% 
-  rename_with(~ varNamesTo, all_of(varNamesFrom))
-```
-
-## Convert the Types of Variables {#convertVarTypes}
-
-One variable:
-
-```{r}
-players$factorVar <- factor(players$ID)
-players$numericVar <- as.numeric(players$age)
-players$integerVar <- as.integer(players$newVar1)
-players$characterVar <- as.character(players$newVar2)
-```
-
-Multiple variables:
-
-```{r}
-players %>%
-  mutate(across(c(
-    ID,
-    age),
-    as.numeric))
-
-players %>%
-  mutate(across(
-    age:newVar1,
-    as.character))
-
-players %>%
-  mutate(across(where(is.factor), as.character))
-```
-
-## Merging/Joins {#sec-merging}
-
-### Overview {#sec-mergingOverview}
-
-Merging (also called joining) merges two data objects using a shared set of variables called "keys."
-The keys are the variable(s) that uniquely identify each row (i.e., they account for the levels of nesting).
-In some data objects, the key might be the player's identification number (e.g., `player_id`).
-However, some data objects have multiple keys.
-For instance, in long form data objects, each participant may have multiple rows corresponding to multiple seasons.
-In this case, the keys may be `player_id` and `season`.
-If a participant has multiple rows corresponding to seasons and games/weeks, the keys are `player_id`, `season`, and `week`.
-In general, each row should have a value on each of the keys; there should be no missingness in the keys.
-
-To merge two objects, the key(s) that will be used to match the records must be present in both objects.
-The keys are used to merge the variables in object 1 (`x`) with the variables in object 2 (`y`).
-Different merge types select different rows to merge.
-
-Note: if the two objects include variables with the same name (apart from the keys), `R` will not know how you want each to appear in the merged object.
-So, it will add a suffix (e.g., `.x`, `.y`) to each common variable to indicate which object (i.e., object `x` or object `y`) the variable came from, where object `x` is the first object—i.e., the object to which object `y` (the second object) is merged.
-In general, apart from the keys, you should not include variables with the same name in two objects to be merged.
-To prevent this, either remove or rename the shared variable in one of the objects, or include the shared variable as a key.
-However, as described above, you should include it as a key ***only*** if it uniquely identifies each row in terms of levels of nesting.
-
-### Data Before Merging {#sec-mergingData}
-
-Here are the data in the `players` object:
-
-```{r}
-players
-
-dim(players)
-```
-
-The data are structured in ID form.
-That is, every row in the dataset is uniquely identified by the variable, `ID`.
-
-Here are the data in the `fantasyPoints` object:
-
-```{r}
-fantasyPoints
-
-dim(fantasyPoints)
-```
-
-### Types of Joins {#sec-mergeTypes}
-
-#### Visual Overview of Join Types {#sec-mergeTypesVisual}
-
-Below is a visual that depicts various types of merges/joins.
-Object `x` is the circle labeled as `x`.
-Object `y` is the circle labeled as `y`.
-The area of overlap in the Venn diagram indicates the rows on the keys that are shared between the two objects (e.g., the same `player_id`, `season`, and `week`).
-The non-overlapping area indicates the rows on the keys that are unique to each object.
-The shaded blue area indicates which rows (on the keys) are kept in the merged object from each of the two objects, when using each of the merge types.
-For instance, a left outer join keeps the shared rows and the rows that are unique to object `x`, but it drops the rows that are unique to object `y`.
-
-![Types of merges/joins](images/joinTypes.png)
-
-#### Full Outer Join {#sec-fullJoin}
-
-A full outer join includes all rows in `x` **or** `y`.
-It returns columns from `x` and `y`.
-Here is how to merge two data frames using a full outer join (i.e., "full join"):
-
-```{r}
-fullJoinData <- full_join(
-  players,
-  fantasyPoints,
-  by = "ID")
-
-fullJoinData
-
-dim(fullJoinData)
-```
-
-#### Left Outer Join {#sec-leftJoin}
-
-A left outer join includes all rows in `x`.
-It returns columns from `x` and `y`.
-Here is how to merge two data frames using a left outer join ("left join"):
-
-```{r}
-leftJoinData <- left_join(
-  players,
-  fantasyPoints,
-  by = "ID")
-
-leftJoinData
-
-dim(leftJoinData)
-```
-
-#### Right Outer Join {#sec-rightJoin}
-
-A right outer join includes all rows in `y`.
-It returns columns from `x` and `y`.
-Here is how to merge two data frames using a right outer join ("right join"):
-
-```{r}
-rightJoinData <- right_join(
-  players,
-  fantasyPoints,
-  by = "ID")
-
-rightJoinData
-
-dim(rightJoinData)
-```
-
-#### Inner Join {#sec-innerJoin}
-
-An inner join includes all rows that are in **both** `x` **and** `y`.
-An inner join will return one row of `x` for each matching row of `y`, and can duplicate values of records on either side (left or right) if `x` and `y` have more than one matching record.
-It returns columns from `x` and `y`.
-Here is how to merge two data frames using an inner join:
-
-```{r}
-innerJoinData <- inner_join(
-  players,
-  fantasyPoints,
-  by = "ID")
-
-innerJoinData
-
-dim(innerJoinData)
-```
-
-#### Semi Join {#sec-semiJoin}
-
-A semi join is a filter.
-A left semi join returns all rows from `x` **with** a match in `y`.
-That is, it filters out records from `x` that are not in `y`.
-Unlike an [inner join](#innerJoin), a left semi join will never duplicate rows of `x`, and it includes columns from only `x` (not from `y`). 
-Here is how to merge two data frames using a left semi join:
-
-```{r}
-semiJoinData <- semi_join(
-  players,
-  fantasyPoints,
-  by = "ID")
-
-semiJoinData
-
-dim(semiJoinData)
-```
-
-#### Anti Join {#sec-antiJoin}
-
-An anti join is a filter.
-A left anti join returns all rows from `x` **without** a match in `y`.
-That is, it filters out records from `x` that are in `y`.
-It returns columns from only `x` (not from `y`).
-Here is how to merge two data frames using a left anti join:
-
-```{r}
-antiJoinData <- anti_join(
-  players,
-  fantasyPoints,
-  by = "ID")
-
-antiJoinData
-
-dim(antiJoinData)
-```
-
-#### Cross Join {#sec-crossJoin}
-
-A cross join combines each row in `x` with each row in `y`.
-
-```{r}
-crossJoinData <- cross_join(
-  players,
-  fantasyPoints)
-
-crossJoinData
-
-dim(crossJoinData)
-```
-
-## Transform Data from Long to Wide {#sec-longToWide}
-
-Depending on the analysis, it may be important to restructure the data to be in long or wide form.
-When the data are in wide form, each player has only one row.
-When the data are in long form, each player has multiple rows—e.g., a row for each game.
-The data structure is called wide or long form because a dataset in wide form has more columns and fewer rows (i.e., it appears wider and shorter), whereas a dataset in long form has more rows and fewer columns (i.e., it appears narrower and taller).
-
-Here are the data in the `nfl_actualStats_offense_weekly` object.
-The data are structured in "player-season-week form".
-That is, every row in the dataset is uniquely identified by the variables, `player_id`, `season`, and `week`.
-This is an example of long form, because each player has multiple rows.
-
-Original data:
-
-```{r}
-dataLong <- nfl_actualStats_offense_weekly %>% 
-  select(player_id, player_display_name, season, week, fantasy_points)
-
-dim(dataLong)
-
-names(dataLong)
-```
-
-Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "player form" (where each row is uniquely identified by the `player_id` variable):
-
-```{r}
-dataWide <- dataLong %>% 
-  pivot_wider(
-    names_from = c(season, week),
-    names_glue = "{.value}_{season}_week{week}",
-    values_from = fantasy_points)
-
-dim(dataWide)
-
-names(dataWide)
-```
-
-## Transform Data from Wide to Long {#sec-wideToLong}
-
-Conversely, we can also restructure data from wide to long.
-
-Original data:
-
-```{r}
-dataWide <- nfl_actualStats_offense_weekly %>% 
-  select(player_id, player_display_name, season, week, recent_team, opponent_team)
-
-dim(dataWide)
-
-names(dataWide)
-```
-
-Data in long form, transformed from wide form using `tidyverse`:
-
-```{r}
-dataLong <- dataWide %>% 
-  pivot_longer(
-    cols = c(recent_team, opponent_team),
-    names_to = "role",
-    values_to = "team")
-
-dim(dataLong)
-
-names(dataLong)
-```
-
-## Loops {#sec-loops}
-
-If you want to perform the same computation multiple times, it can be faster to do it in a loop compared to writing out the same computation many times.
-For instance, here is a loop that runs from 1 to `{r} length(players$ID)` (the number of players in the `players` object), incrementing by 1 after each iteration.
-The loop prints each element of a vector (i.e., the player's name) and the loop index (`i`) that indicates where the loop is in terms of its iterations:
-
-```{r}
-for(i in 1:length(players$ID)){
-  print(paste("The loop is at index:", i, sep = " "))
-  print(paste("My favorite player is:", players$name[i], sep = " "))
-}
-```
-
 ## Calculations {#sec-calculations}
@@
 
-## Session Info {#sec-gettingStartedSessionInfo}
-
-At the end of each chapter in which `R` code is used, I provide the session information, which describes the system and operating system the code was run on and the versions of each package.
-That way, if you get different results from me, you can see which differ, to help with reproducibility.
-If you run the (all of) the exact same code as is provided in the text, in the exact same order, with the exact same setup (platform, operating system, package versions, etc.), you should get the exact same answer as is in the text.
-That is the idea of reproducibility—getting the exact same result with the exact same inputs.
-Reproducibility is crucial for studies to achieve greater confidence in their findings and to ensure better replicability of findings across studies.
+## Session Info {#sec-downloadFootballDataSessionInfo}
 

### 2024-08-02 getting-started.qmd
20240802 - separate chapter for downloading football data

--- 2024-08-02 getting-started.qmd
@@
 
-## Download Football Data {#sec-downloadFootballData}
-
-Below, we provide examples for how to download various types of National Football League (NFL) data.
-For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
-We save each data file after downloading it, so we can use the data in subsequent chapters.
-Guidance for how to merge the various data files is provided at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md>
-
-```{r}
-#| eval: false
-#| include: false
-
-# Downloaded Data - Raw
-load(file = "./data/nfl_players_raw.RData")
-load(file = "./data/nfl_teams_raw.RData")
-load(file = "./data/nfl_rosters_raw.RData")
-load(file = "./data/nfl_rosters_weekly_raw.RData")
-load(file = "./data/nfl_schedules_raw.RData")
-load(file = "./data/nfl_combine_raw.RData")
-load(file = "./data/nfl_draftPicks_raw.RData")
-load(file = "./data/nfl_depthCharts_raw.RData")
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp_raw.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown_raw.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation_raw.RData", fsep = ""))
-load(file = "./data/nfl_actualStats_weekly_raw.RData")
-load(file = "./data/nfl_injuries_raw.RData")
-load(file = "./data/nfl_snapCounts_raw.RData")
-load(file = "./data/nfl_espnQBR_seasonal_raw.RData")
-load(file = "./data/nfl_espnQBR_weekly_raw.RData")
-load(file = "./data/nfl_nextGenStats_weekly_raw.RData")
-load(file = "./data/nfl_advancedStatsPFR_seasonal_raw.RData")
-load(file = "./data/nfl_advancedStatsPFR_weekly_raw.RData")
-load(file = "./data/nfl_playerContracts_raw.RData")
-load(file = "./data/nfl_ftnCharting_raw.RData")
-load(file = "./data/nfl_playerIDs_raw.RData")
-load(file = "./data/nfl_rankings_draft_raw.RData")
-load(file = "./data/nfl_rankings_weekly_raw.RData")
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly_raw.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp_raw.RData", fsep = ""))
-
-# Calculated Data - Raw
-load(file = "./data/nfl_actualStats_career_raw.RData")
-load(file = "./data/nfl_actualStats_seasonal_raw.RData")
-
-# Downloaded Data - Processed
-load(file = "./data/nfl_players.RData")
-load(file = "./data/nfl_teams.RData")
-load(file = "./data/nfl_rosters.RData")
-load(file = "./data/nfl_rosters_weekly.RData")
-load(file = "./data/nfl_schedules.RData")
-load(file = "./data/nfl_combine.RData")
-load(file = "./data/nfl_draftPicks.RData")
-load(file = "./data/nfl_depthCharts.RData")
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
-load(file = "./data/nfl_actualStats_weekly.RData")
-load(file = "./data/nfl_injuries.RData")
-load(file = "./data/nfl_snapCounts.RData")
-load(file = "./data/nfl_espnQBR_seasonal.RData")
-load(file = "./data/nfl_espnQBR_weekly.RData")
-load(file = "./data/nfl_nextGenStats_weekly.RData")
-load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
-load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
-load(file = "./data/nfl_playerContracts.RData")
-load(file = "./data/nfl_ftnCharting.RData")
-load(file = "./data/nfl_playerIDs.RData")
-load(file = "./data/nfl_rankings_draft.RData")
-load(file = "./data/nfl_rankings_weekly.RData")
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
-
-# Calculated Data - Processed
-load(file = "./data/nfl_actualStats_career.RData")
-load(file = "./data/nfl_actualStats_seasonal.RData")
-```
-
-### Players {#sec-downloadPlayers}
-
-```{r}
-#| eval: false
-
-nfl_players_raw <- progressr::with_progress(
-  nflreadr::load_players())
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_players_raw,
-  file = "./data/nfl_players_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_players_raw.RData")
-```
-
-The `nfl_players` object is in `player` form.
-That is, each row should be uniquely identified by `gsis_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_players <- nfl_players_raw %>% 
-  select(gsis_id, everything()) %>% 
-  arrange(display_name)
-```
-
-Let's check for duplicate `player` instances:
-
-```{r}
-nfl_players %>% 
-  group_by(gsis_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's do some data cleanup:
-
-```{r}
-# Convert missing values to NA
-nfl_players[nfl_players == ""] <- NA
-
-# Drop players with missing values for gsis_id
-nfl_players <- nfl_players %>% 
-  filter(!is.na(gsis_id))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_players,
-  file = "./data/nfl_players.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_players.RData")
-```
-
-### Teams {#sec-downloadTeams}
-
-```{r}
-#| eval: false
-
-nfl_teams_raw <- progressr::with_progress(
-  nflreadr::load_teams(current = TRUE))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_teams_raw,
-  file = "./data/nfl_teams_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_teams_raw.RData")
-```
-
-The `nfl_teams` object is in `team` form.
-That is, each row should be uniquely identified by `team_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_teams <- nfl_teams_raw %>% 
-  select(team_id, everything())
-```
-
-Let's check for duplicate `team` instances:
-
-```{r}
-nfl_teams %>% 
-  group_by(team_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_teams,
-  file = "./data/nfl_teams.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_teams.RData")
-```
-
-### Player Info {#sec-downloadPlayerInfo}
-
-### Rosters {#sec-downloadRosters}
-
-A Data Dictionary for rosters is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_rosters.html>
-
-```{r}
-#| eval: false
-
-nfl_rosters_raw <- progressr::with_progress(
-  nflreadr::load_rosters(seasons = TRUE))
-
-nfl_rosters_weekly_raw <- progressr::with_progress(
-  nflreadr::load_rosters_weekly(seasons = TRUE))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_rosters_raw,
-  file = "./data/nfl_rosters_raw.RData"
-)
-
-save(
-  nfl_rosters_weekly_raw,
-  file = "./data/nfl_rosters_weekly_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_rosters_raw.RData")
-load(file = "./data/nfl_rosters_weekly_raw.RData")
-```
-
-The `nfl_rosters` object is in `player`-`season`-`team` form.
-That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `team`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_rosters <- nfl_rosters_raw %>% 
-  select(gsis_id, season, team, week, everything()) %>% 
-  arrange(full_name, gsis_id, season, team, week)
-```
-
-Let's check for duplicate `player`-`season`-`team` instances:
-
-```{r}
-nfl_rosters %>% 
-  group_by(gsis_id, season, team) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's do some data cleanup:
-
-```{r}
-# Drop players with missing values for gsis_id
-nfl_rosters <- nfl_rosters %>% 
-  filter(!is.na(gsis_id))
-
-# Fill in missing values for a player in their duplicate instances, and then keep only the first of the duplicate instances
-nfl_rosters <- nfl_rosters %>% 
-  group_by(gsis_id, season, team) %>% 
-  fill(names(.), .direction = "downup") %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-```
-
-Let's check again for duplicate `player`-`season`-`team` instances:
-
-```{r}
-nfl_rosters %>% 
-  group_by(gsis_id, season, team) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-The `nfl_rosters_weekly` object is in `player`-`season`-`week` form.
-That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `week`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_rosters_weekly <- nfl_rosters_weekly_raw %>% 
-  select(gsis_id, season, week, everything()) %>% 
-  arrange(full_name, gsis_id, season, week)
-```
-
-Let's check for duplicate `player`-`season`-`week` instances:
-
-```{r}
-nfl_rosters_weekly %>% 
-  group_by(gsis_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's do some data cleanup:
-
-```{r}
-# Drop players with missing values for gsis_id
-nfl_rosters_weekly <- nfl_rosters_weekly %>% 
-  filter(!is.na(gsis_id))
-
-# Fill in missing values for a player in their duplicate instances, and then keep only the first of the duplicate instances
-nfl_rosters_weekly <- nfl_rosters_weekly %>% 
-  group_by(gsis_id, season, week) %>% 
-  fill(names(.), .direction = "downup") %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-```
-
-Let's check again for duplicate `player`-`season`-`week` instances:
-
-```{r}
-nfl_rosters_weekly %>% 
-  group_by(gsis_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_rosters,
-  file = "./data/nfl_rosters.RData"
-)
-
-save(
-  nfl_rosters_weekly,
-  file = "./data/nfl_rosters_weekly.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_rosters.RData")
-load(file = "./data/nfl_rosters_weekly.RData")
-```
-
-### Game Schedules {#sec-downloadSchedules}
-
-A Data Dictionary for game schedules data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_schedules.html>
-
-```{r}
-#| eval: false
-
-nfl_schedules_raw <- progressr::with_progress(
-  nflreadr::load_schedules(seasons = TRUE))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_schedules_raw,
-  file = "./data/nfl_schedules_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_schedules_raw.RData")
-```
-
-The `nfl_schedules` object is in `game` form.
-That is, each row should be uniquely identified by `game_id`.
-
-```{r}
-nfl_schedules <- nfl_schedules_raw
-```
-
-Let's check for duplicate `game` instances:
-
-```{r}
-nfl_schedules %>% 
-  group_by(game_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_schedules,
-  file = "./data/nfl_schedules.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_schedules.RData")
-```
-
-### The Combine {#sec-combine}
-
-A Data Dictionary for data from the combine is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_combine.html>
-
-```{r}
-#| eval: false
-
-nfl_combine_raw <- progressr::with_progress(
-  nflreadr::load_combine(seasons = TRUE))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_combine_raw,
-  file = "./data/nfl_combine_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_combine_raw.RData")
-```
-
-The `nfl_combine` object is in `player` form.
-That is, each row should be uniquely identified by the player's `id`.
-However, there is no `gsis_id` variable to merge it easily with other datasets.
-Some of the players have other `id` variables, including `pfr_id` and `cfb_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_combine <- nfl_combine_raw %>% 
-  select(pfr_id, cfb_id, everything()) %>% 
-  arrange(season, player_name)
-```
-
-Let's check for duplicate `pfr_id` and `cfb_id` instances:
-
-```{r}
-nfl_combine %>% 
-  group_by(pfr_id) %>% 
-  filter(n() > 1, !is.na(pfr_id)) %>% 
-  arrange(pfr_id) %>% 
-  head()
-
-nfl_combine %>% 
-  group_by(cfb_id) %>% 
-  filter(n() > 1, !is.na(cfb_id)) %>% 
-  arrange(cfb_id) %>% 
-  head()
-```
-
-However, these apparent duplicates appear to be different players:
-
-```{r}
-nfl_combine %>% 
-  group_by(season, pfr_id, pos) %>% 
-  filter(n() > 1, !is.na(pfr_id)) %>% 
-  arrange(pfr_id) %>% 
-  head()
-
-nfl_combine %>% 
-  group_by(season, cfb_id, pos) %>% 
-  filter(n() > 1, !is.na(cfb_id)) %>% 
-  arrange(cfb_id) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_combine,
-  file = "./data/nfl_combine.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_combine.RData")
-```
-
-### Draft Picks {#sec-draftPicks}
-
-A Data Dictionary for draft picks data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_draft_picks.html>
-
-```{r}
-#| eval: false
-
-nfl_draftPicks_raw <- progressr::with_progress(
-  nflreadr::load_draft_picks(seasons = TRUE))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_draftPicks_raw,
-  file = "./data/nfl_draftPicks_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_draftPicks_raw.RData")
-```
-
-The `nfl_draftPicks` object is in `player` form.
-That is, each row should be uniquely identified by `gsis_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_draftPicks <- nfl_draftPicks_raw %>% 
-  select(gsis_id, everything()) %>% 
-  arrange(pfr_player_name)
-```
-
-Let's check for duplicate `player` instances:
-
-```{r}
-nfl_draftPicks %>% 
-  group_by(gsis_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's do some data cleanup:
-
-```{r}
-# Convert missing values to NA
-nfl_draftPicks[nfl_draftPicks == ""] <- NA
-
-# Drop players with missing values for gsis_id
-nfl_draftPicks <- nfl_draftPicks %>% 
-  filter(!is.na(gsis_id))
-```
-
-Let's check again for duplicate `player` instances:
-
-```{r}
-nfl_draftPicks %>% 
-  group_by(gsis_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_draftPicks,
-  file = "./data/nfl_draftPicks.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_draftPicks.RData")
-```
-
-### Depth Charts {#sec-depthCharts}
-
-A Data Dictionary for data from weekly depth charts is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_depth_charts.html>
-
-```{r}
-#| eval: false
-
-nfl_depthCharts_raw <- progressr::with_progress(
-  nflreadr::load_depth_charts(seasons = TRUE))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_depthCharts_raw,
-  file = "./data/nfl_depthCharts_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_depthCharts_raw.RData")
-```
-
-The `nfl_depthCharts` object is in `id`-`season`-`week`-`position` form.
-That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, `week`, and `depth_position`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_depthCharts <- nfl_depthCharts_raw %>% 
-  select(gsis_id, season, week, depth_position, everything()) %>% 
-  arrange(full_name, gsis_id, season, week, depth_position)
-```
-
-Let's check for duplicate `id`-`season`-`week`-`position` instances:
-
-```{r}
-nfl_depthCharts %>% 
-  group_by(gsis_id, season, week, depth_position) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's do some data cleanup:
-
-```{r}
-# Drop players with missing values for gsis_id
-nfl_depthCharts <- nfl_depthCharts %>% 
-  filter(!is.na(gsis_id))
-
-# Fill in missing values for a player in their duplicate instances, and then keep only the first of the duplicate instances
-nfl_depthCharts <- nfl_depthCharts %>% 
-  group_by(gsis_id, season, week, depth_position) %>% 
-  fill(names(.), .direction = "downup") %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-```
-
-Let's check again for duplicate `id`-`season`-`week`-`position` instances:
-
-```{r}
-nfl_depthCharts %>% 
-  group_by(gsis_id, season, week, depth_position) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_depthCharts,
-  file = "./data/nfl_depthCharts.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_depthCharts.RData")
-```
-
-### Play-By-Play Data {#sec-downloadPlayByPlay}
-
-To download play-by-play data from prior weeks and seasons, we can use the `load_pbp()` function of the `nflreadr` package.
-We add a progress bar using the `with_progress()` function from the `progressr` package because it takes a while to run.
-A Data Dictionary for the play-by-play data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>
-
-::: {#nte-downloadPlayByPlay .callout-note title="Downloading play-by-play data"}
-Note: the following code takes a while to run.
-:::
-
-```{r}
-#| eval: false
-
-nfl_pbp_raw <- progressr::with_progress(
-  nflreadr::load_pbp(seasons = TRUE))
-```
-
-```{r}
-#| include: false
-
-nfl_pbp_raw <- nflreadr::load_pbp(seasons = TRUE)
-```
-
-```{r}
-save(
-  nfl_pbp_raw,
-  file = "./data/nfl_pbp_raw.RData"
-)
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = "./data/nfl_pbp_raw.RData")
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp_raw.RData", fsep = ""))
-```
-
-The `nfl_pbp` object is in `game`-`drive`-`play` form.
-That is, each row should be uniquely identified by the combination of `game_id`, `drive`, `play_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_pbp <- nfl_pbp_raw %>% 
-  select(game_id, drive, play_id, everything()) %>% 
-  arrange(game_id, drive, play_id)
-```
-
-Let's check for duplicate `game`-`drive`-`play` instances:
-
-```{r}
-nfl_pbp %>% 
-  group_by(game_id, drive, play_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-save(
-  nfl_pbp,
-  file = "./data/nfl_pbp.RData"
-)
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = "./data/nfl_pbp.RData")
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
-```
-
-### 4th Down Data {#sec-download4thDown}
-
-::: {#nte-download4thDownData .callout-note title="Downloading 4th down data"}
-Note: the following code takes a while to run.
-:::
-
-```{r}
-#| output: false
-
-nfl_4thdown_raw <- nfl4th::load_4th_pbp(seasons = 2014:nflreadr::most_recent_season())
-```
-
-```{r}
-save(
-  nfl_4thdown_raw,
-  file = "./data/nfl_4thdown_raw.RData"
-)
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = "./data/nfl_4thdown_raw.RData")
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown_raw.RData", fsep = ""))
-```
-
-The `nfl_4thdown` object is in `game`-`drive`-`play` form.
-That is, each row should be uniquely identified by the combination of `game_id`, `drive`, `play_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_4thdown <- nfl_4thdown_raw %>% 
-  select(game_id, drive, play_id, everything()) %>% 
-  arrange(game_id, drive, play_id)
-```
-
-Let's check for duplicate `game`-`drive`-`play` instances:
-
-```{r}
-nfl_4thdown %>% 
-  group_by(game_id, drive, play_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-save(
-  nfl_4thdown,
-  file = "./data/nfl_4thdown.RData"
-)
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = "./data/nfl_4thdown.RData")
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
-```
-
-### Participation {#sec-downloadParticipation}
-
-A Data Dictionary for the participation data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_participation.html>
-
-```{r}
-nfl_participation_raw <- progressr::with_progress(
-  nflreadr::load_participation(
-    seasons = TRUE,
-    include_pbp = TRUE))
-```
-
-```{r}
-save(
-  nfl_participation_raw,
-  file = "./data/nfl_participation_raw.RData"
-)
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = "./data/nfl_participation_raw.RData")
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation_raw.RData", fsep = ""))
-```
-
-The `nfl_participation` object is in `game`-`drive`-`play` form.
-That is, each row should be uniquely identified by the combination of `nflverse_game_id`, `drive`, `play_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_participation <- nfl_participation_raw %>% 
-  select(nflverse_game_id, drive, play_id, everything()) %>% 
-  arrange(nflverse_game_id, drive, play_id)
-```
-
-Let's check for duplicate `game`-`drive`-`play` instances:
-
-```{r}
-nfl_participation %>% 
-  group_by(nflverse_game_id, drive, play_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-save(
-  nfl_participation,
-  file = "./data/nfl_participation.RData"
-)
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = "./data/nfl_participation.RData")
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
-```
-
-### Historical Weekly Actual Player Statistics {#sec-downloadActualStats}
-
-We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package.
-A Data Dictionary for statistics for offensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
-A Data Dictionary for statistics for defensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.
-
-```{r}
-#| eval: false
-
-nfl_actualStats_offense_weekly_raw <- progressr::with_progress(
-  nflreadr::load_player_stats(
-    seasons = TRUE,
-    stat_type = "offense"))
-
-nfl_actualStats_defense_weekly_raw <- progressr::with_progress(
-  nflreadr::load_player_stats(
-    seasons = TRUE,
-    stat_type = "defense"))
-
-nfl_actualStats_kicking_weekly_raw <- progressr::with_progress(
-  nflreadr::load_player_stats(
-    seasons = TRUE,
-    stat_type = "kicking"))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_offense_weekly_raw, nfl_actualStats_defense_weekly_raw, nfl_actualStats_kicking_weekly_raw,
-  file = "./data/nfl_actualStats_weekly_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_weekly_raw.RData")
-```
-
-The `nfl_actualStats_weekly` objects are in `player`-`season`-`week` form.
-That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly_raw %>% 
-  select(player_id, season, week, everything()) %>% 
-  arrange(player_display_name, player_id, season, week)
-
-nfl_actualStats_defense_weekly <- nfl_actualStats_defense_weekly_raw %>% 
-  select(player_id, season, week, everything()) %>% 
-  arrange(player_display_name, player_id, season, week)
-
-nfl_actualStats_kicking_weekly <- nfl_actualStats_kicking_weekly_raw %>% 
-  select(player_id, season, week, everything()) %>% 
-  arrange(player_display_name, player_id, season, week)
-```
-
-Let's check for duplicate `player`-`season`-`week` instances:
-
-```{r}
-nfl_actualStats_offense_weekly %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_defense_weekly %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_kicking_weekly %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_offense_weekly, nfl_actualStats_defense_weekly, nfl_actualStats_kicking_weekly,
-  file = "./data/nfl_actualStats_weekly.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_weekly.RData")
-```
-
-### Injuries {#sec-injuries}
-
-A Data Dictionary for injury data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_injuries.html>
-
-```{r}
-#| eval: false
-
-nfl_injuries_raw <- progressr::with_progress(
-  nflreadr::load_injuries(seasons = TRUE))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_injuries_raw,
-  file = "./data/nfl_injuries_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_injuries_raw.RData")
-```
-
-The `nfl_injuries` object is in `player`-`season`-`week` form.
-That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `week`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_injuries <- nfl_injuries_raw %>% 
-  select(gsis_id, season, week, everything()) %>% 
-  arrange(full_name, gsis_id, season, week)
-```
-
-Let's check for duplicate `player`-`season`-`week` instances:
-
-```{r}
-nfl_injuries %>% 
-  group_by(gsis_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_injuries,
-  file = "./data/nfl_injuries.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_injuries.RData")
-```
-
-### Snap Counts {#sec-snapCounts}
-
-A Data Dictionary for snap counts data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_snap_counts.html>
-
-```{r}
-#| eval: false
-
-nfl_snapCounts_raw <- progressr::with_progress(
-  nflreadr::load_snap_counts(seasons = TRUE))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_snapCounts_raw,
-  file = "./data/nfl_snapCounts_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_snapCounts_raw.RData")
-```
-
-The `nfl_snapCounts` object is in `game`-`player` form.
-That is, each row should be uniquely identified by the combination of `game_id` and `pfr_player_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_snapCounts <- nfl_snapCounts_raw %>% 
-  select(game_id, pfr_player_id, everything()) %>% 
-  arrange(game_id, pfr_player_id)
-```
-
-Let's check for duplicate `game` instances:
-
-```{r}
-nfl_snapCounts %>% 
-  group_by(game_id, pfr_player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_snapCounts,
-  file = "./data/nfl_snapCounts.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_snapCounts.RData")
-```
-
-### ESPN QBR {#sec-espnQBR}
-
-A Data Dictionary for ESPN QBR data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_espn_qbr.html>
-
-```{r}
-#| eval: false
-
-nfl_espnQBR_seasonal_raw <- progressr::with_progress(
-  nflreadr::load_espn_qbr(
-    seasons = TRUE,
-    summary_type = c("season")))
-
-nfl_espnQBR_weekly_raw <- progressr::with_progress(
-  nflreadr::load_espn_qbr(
-    seasons = TRUE,
-    summary_type = c("week")))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_espnQBR_seasonal_raw,
-  file = "./data/nfl_espnQBR_seasonal_raw.RData"
-)
-
-save(
-  nfl_espnQBR_weekly_raw,
-  file = "./data/nfl_espnQBR_weekly_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_espnQBR_seasonal_raw.RData")
-load(file = "./data/nfl_espnQBR_weekly_raw.RData")
-```
-
-The `nfl_espnQBR_seasonal` object is in `player`-`season`-`season type` form, where `season type` refers to regular season versus postseason.
-That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `season_type`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_espnQBR_seasonal <- nfl_espnQBR_seasonal_raw %>% 
-  select(player_id, season, season_type, everything()) %>% 
-  arrange(name_display, player_id, season, season_type)
-```
-
-Let's check for duplicate `player`-`season`-`team` instances:
-
-```{r}
-nfl_espnQBR_seasonal %>% 
-  group_by(player_id, season, season_type) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-The `nfl_espnQBR_weekly` object is in both `game`-`player` form and `player`-`season`-`season type`-`week` form, where `season type` refers to regular season versus postseason.
-That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `week` or by the combination of `player_id`, `season`, `season_type`, and `week_num`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_espnQBR_weekly <- nfl_espnQBR_weekly_raw %>% 
-  select(player_id, season, season_type, week_num, everything()) %>% 
-  arrange(name_display, player_id, season, season_type, week_num)
-```
-
-Let's check for duplicate `game`-`player` or `player`-`season`-`season type`-`week` instances:
-
-```{r}
-nfl_espnQBR_weekly %>% 
-  arrange(game_id, player_id) %>% 
-  group_by(game_id, player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_espnQBR_weekly %>% 
-  arrange(player_id, season, season_type, week_num) %>% 
-  group_by(player_id, season, season_type, week_num) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_espnQBR_seasonal,
-  file = "./data/nfl_espnQBR_seasonal.RData"
-)
-
-save(
-  nfl_espnQBR_weekly,
-  file = "./data/nfl_espnQBR_weekly.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_espnQBR_seasonal.RData")
-load(file = "./data/nfl_espnQBR_weekly.RData")
-```
-
-### NFL Next Gen Stats {#sec-nextGenStats}
-
-A Data Dictionary for NFL Next Gen Stats data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_nextgen_stats.html>
-
-```{r}
-#| eval: false
-
-nfl_nextGenStats_pass_weekly_raw <- progressr::with_progress(
-  nflreadr::load_nextgen_stats(
-    seasons = TRUE,
-    stat_type = c("passing")))
-
-nfl_nextGenStats_rush_weekly_raw <- progressr::with_progress(
-  nflreadr::load_nextgen_stats(
-    seasons = TRUE,
-    stat_type = c("rushing")))
-
-nfl_nextGenStats_rec_weekly_raw <- progressr::with_progress(
-  nflreadr::load_nextgen_stats(
-    seasons = TRUE,
-    stat_type = c("receiving")))
-
-nfl_nextGenStats_weekly_raw <- bind_rows(
-  nfl_nextGenStats_pass_weekly_raw,
-  nfl_nextGenStats_rush_weekly_raw,
-  nfl_nextGenStats_rec_weekly_raw
-)
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_nextGenStats_weekly_raw,
-  file = "./data/nfl_nextGenStats_weekly_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_nextGenStats_weekly_raw.RData")
-```
-
-The `nfl_nextGenStats_weekly` object is in `player`-`season`-`season type`-`week` form, where `season type` refers to regular season versus postseason.
-That is, each row should be uniquely identified by the combination of `player_gsis_id`, `season`, `season_type`, and `week`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_nextGenStats_weekly <- nfl_nextGenStats_weekly_raw %>% 
-  select(player_gsis_id, season, season_type, week, everything()) %>% 
-  arrange(player_display_name, player_gsis_id, season, season_type)
-```
-
-Let's check for duplicate `player`-`season`-`season type`-`week` instances:
-
-```{r}
-nfl_nextGenStats_weekly %>% 
-  group_by(player_gsis_id, season, season_type, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_nextGenStats_weekly,
-  file = "./data/nfl_nextGenStats_weekly.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_nextGenStats_weekly.RData")
-```
-
-### Advanced Stats from Pro Football Reference {#sec-advancedStatsPFR}
-
-A Data Dictionary for Pro Football Reference passing data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html>
-
-```{r}
-#| eval: false
-
-nfl_advancedStatsPFR_pass_seasonal_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("pass"),
-    summary_level = c("season")))
-
-nfl_advancedStatsPFR_pass_weekly_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("pass"),
-    summary_level = c("week")))
-
-nfl_advancedStatsPFR_rush_seasonal_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("rush"),
-    summary_level = c("season")))
-
-nfl_advancedStatsPFR_rush_weekly_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("rush"),
-    summary_level = c("week")))
-
-nfl_advancedStatsPFR_rec_seasonal_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("rec"),
-    summary_level = c("season")))
-
-nfl_advancedStatsPFR_rec_weekly_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("rec"),
-    summary_level = c("week")))
-
-nfl_advancedStatsPFR_def_seasonal_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("def"),
-    summary_level = c("season")))
-
-nfl_advancedStatsPFR_def_weekly_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("def"),
-    summary_level = c("week")))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_advancedStatsPFR_pass_seasonal_raw,
-  nfl_advancedStatsPFR_rush_seasonal_raw,
-  nfl_advancedStatsPFR_rec_seasonal_raw,
-  nfl_advancedStatsPFR_def_seasonal_raw,
-  file = "./data/nfl_advancedStatsPFR_seasonal_raw.RData"
-)
-
-save(
-  nfl_advancedStatsPFR_pass_weekly_raw,
-  nfl_advancedStatsPFR_rush_weekly_raw,
-  nfl_advancedStatsPFR_rec_weekly_raw,
-  nfl_advancedStatsPFR_def_weekly_raw,
-  file = "./data/nfl_advancedStatsPFR_weekly_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_advancedStatsPFR_seasonal_raw.RData")
-load(file = "./data/nfl_advancedStatsPFR_weekly_raw.RData")
-```
-
-```{r}
-nfl_advancedStatsPFR_pass_seasonal <- nfl_advancedStatsPFR_pass_seasonal_raw %>% 
-  mutate(
-    nameMerge = toupper(str_remove_all(player, "[[:punct:]]"))
-  ) %>% 
-  rename(
-    player_name.pass = player
-  )
-
-nfl_advancedStatsPFR_rush_seasonal <- nfl_advancedStatsPFR_rush_seasonal_raw %>% 
-  mutate(
-    nameMerge = toupper(str_remove_all(player, "[[:punct:]]"))
-  ) %>% 
-  rename(
-    player_name.rush = player,
-    age.rush = age,
-    team = tm
-  )
-
-nfl_advancedStatsPFR_rec_seasonal <- nfl_advancedStatsPFR_rec_seasonal_raw %>% 
-  mutate(
-    nameMerge = toupper(str_remove_all(player, "[[:punct:]]"))
-  ) %>% 
-  rename(
-    player_name.rec = player,
-    age.rec = age,
-    team = tm
-  )
-
-nfl_advancedStatsPFR_def_seasonal <- nfl_advancedStatsPFR_def_seasonal_raw %>% 
-  mutate(
-    nameMerge = toupper(str_remove_all(player, "[[:punct:]]"))
-  ) %>% 
-  rename(
-    player_name.def = player,
-    age.def = age,
-    team = tm
-  )
-
-nfl_advancedStatsPFR_pass_weekly <- nfl_advancedStatsPFR_pass_weekly_raw %>% 
-  mutate(
-    nameMerge = toupper(str_remove_all(pfr_player_name, "[[:punct:]]"))
-  ) %>% 
-  rename(
-    player_name.pass = pfr_player_name,
-    season.pass = season,
-    week.pass = week,
-    team.pass = team
-  )
-
-nfl_advancedStatsPFR_rush_weekly <- nfl_advancedStatsPFR_rush_weekly_raw %>% 
-  mutate(
-    nameMerge = toupper(str_remove_all(pfr_player_name, "[[:punct:]]"))
-  ) %>% 
-  rename(
-    player_name.rush = pfr_player_name,
-    season.rush = season,
-    week.rush = week,
-    team.rush = team
-  )
-
-nfl_advancedStatsPFR_rec_weekly <- nfl_advancedStatsPFR_rec_weekly_raw %>% 
-  mutate(
-    nameMerge = toupper(str_remove_all(pfr_player_name, "[[:punct:]]"))
-  ) %>% 
-  rename(
-    player_name.rec = pfr_player_name,
-    season.rec = season,
-    week.rec = week,
-    team.rec = team
-  )
-
-nfl_advancedStatsPFR_def_weekly <- nfl_advancedStatsPFR_def_weekly_raw %>% 
-  mutate(
-    nameMerge = toupper(str_remove_all(pfr_player_name, "[[:punct:]]"))
-  ) %>% 
-  rename(
-    player_name.def = pfr_player_name,
-    season.def = season,
-    week.def = week,
-    team.def = team
-  )
-
-nfl_advancedStatsPFR_seasonal_list <- list(
-  nfl_advancedStatsPFR_pass_seasonal,
-  nfl_advancedStatsPFR_rush_seasonal,
-  nfl_advancedStatsPFR_rec_seasonal,
-  nfl_advancedStatsPFR_def_seasonal)
-
-nfl_advancedStatsPFR_weekly_list <- list(
-  nfl_advancedStatsPFR_pass_weekly,
-  nfl_advancedStatsPFR_rush_weekly,
-  nfl_advancedStatsPFR_rec_weekly,
-  nfl_advancedStatsPFR_def_weekly)
-
-nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal_list %>% 
-  purrr::reduce(
-    full_join,
-    by = c("pfr_id","nameMerge","season","team"))
-
-nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly_list %>% 
-  purrr::reduce(
-    full_join,
-    by = c("game_id","pfr_player_id")) #nameMerge
-
-#nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly_list %>% 
-#  purrr::reduce(
-#    full_join,
-#    by = c("pfr_player_id","nameMerge","season","week"))
-
-nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal %>% 
-  mutate(
-    player_name = coalesce(
-      player_name.pass,
-      player_name.rush,
-      player_name.rec,
-      player_name.def
-    ),
-    age = coalesce(
-      #age.pass,
-      age.rush,
-      age.rec,
-      age.def
-    )
-  ) %>% 
-  select(-c(starts_with("player_name.")))
-
-nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly %>% 
-  mutate(
-    player_name = coalesce(
-      player_name.pass,
-      player_name.rush,
-      player_name.rec,
-      player_name.def
-    ),
-    ,
-    season = coalesce(
-      season.pass,
-      season.rush,
-      season.rec,
-      season.def
-    ),
-    week = coalesce(
-      week.pass,
-      week.rush,
-      week.rec,
-      week.def
-    ),
-    team = coalesce(
-      team.pass,
-      team.rush,
-      team.rec,
-      team.def
-    )
-  ) %>% 
-  select(-c(starts_with("player_name.")))
-
-# Todo:
-#-separate variables from each pass/rush/rec/def object
-```
-
-The `nfl_advancedStatsPFR_seasonal` object is in `player`-`season`-`team` form.
-That is, each row should be uniquely identified by the combination of `pfr_id`, `season`, and `team`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal %>% 
-  select(pfr_id, season, team, player_name, everything()) %>% 
-  arrange(player_name, pfr_id, season, team)
-```
-
-Let's check for duplicate `player`-`season`-`team` instances:
-
-```{r}
-nfl_advancedStatsPFR_seasonal %>% 
-  group_by(pfr_id, season, team) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-The `nfl_advancedStatsPFR_weekly` object is in both `game`-`player` form and `player`-`season`-`week` form.
-That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `week` or by the combination of `player_id`, `season`, `season_type`, and `week_num`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly %>% 
-  select(pfr_player_id, season, week, everything()) %>% 
-  arrange(player_name, pfr_player_id, season, week)
-```
-
-Let's check for duplicate `game`-`player` or `player`-`season`-`week` instances:
-
-```{r}
-nfl_advancedStatsPFR_weekly %>% 
-  arrange(game_id, pfr_player_id) %>% 
-  group_by(game_id, pfr_player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_advancedStatsPFR_weekly %>% 
-  arrange(pfr_player_id, season, week) %>% 
-  group_by(pfr_player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Merge with `gsis_id` for merging with other datasets:
-
-```{r}
-# Prepare data for merging
-nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly %>% 
-  rename(pfr_id = pfr_player_id)
-
-# Identify duplicates
-nfl_advancedStatsPFR_seasonal %>% 
-  select(pfr_id, season, age) %>% 
-  na.omit() %>% 
-  unique() %>% 
-  group_by(pfr_id, season) %>% 
-  filter(n() > 1) %>% 
-  arrange(pfr_id, season) %>% 
-  head()
-
-# Merge seasonal data with the player IDs
-nfl_advancedStatsPFR_seasonal <- left_join(
-  nfl_advancedStatsPFR_seasonal,
-  nfl_playerIDs %>% 
-    filter(!is.na(pfr_id)) %>% 
-    filter(gsis_id != "00-0039137") %>% # drop DL Byron Young, keep OLB Byron Young
-    select(pfr_id, gsis_id) %>% 
-    unique(),
-  by = "pfr_id"
-)
-
-# Merge weekly data with the player IDs
-nfl_advancedStatsPFR_weekly <- left_join(
-  nfl_advancedStatsPFR_weekly,
-  nfl_playerIDs %>% 
-    filter(!is.na(pfr_id)) %>% 
-    filter(gsis_id != "00-0039137") %>% # drop DL Byron Young, keep OLB Byron Young
-    select(pfr_id, gsis_id) %>% 
-    unique(),
-  by = "pfr_id"
-)
-
-# Remove distinct players who were given the same `pfr_id` (to allow merging)
-nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0035665" & nfl_advancedStatsPFR_seasonal$pos %in% c("LB","LILB","RILB"))] <- NA # drop LB David Young, keep DB David Young
-#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0035665" & nfl_advancedStatsPFR_weekly$team %in% c("TEN","MIA"))] <- NA # drop LB David Young, keep DB David Young
-
-nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0035292" & nfl_advancedStatsPFR_seasonal$pos %in% c("LB","LILB","RILB"))] <- NA # drop LB David Young, keep DB David Young
-nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0035292" & nfl_advancedStatsPFR_weekly$team %in% c("TEN","MIA"))] <- NA # drop LB David Young, keep DB David Young
-
-nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0033894" & nfl_advancedStatsPFR_seasonal$pos == "DB")] <- NA # drop S Marcus Williams, keep DB David Young
-#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0033894" & nfl_advancedStatsPFR_weekly$pos == "DB")] <- NA # drop S Marcus Williams
-
-nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0038407" & nfl_advancedStatsPFR_seasonal$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones
-#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0038407" & nfl_advancedStatsPFR_weekly$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones
-
-nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0037106" & nfl_advancedStatsPFR_seasonal$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones
-#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0037106" & nfl_advancedStatsPFR_weekly$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones
-
-nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0038549" & nfl_advancedStatsPFR_seasonal$pos == "WR")] <- NA # drop WR DJ TUrner, keep CB DJ Turner
-#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0038549" & nfl_advancedStatsPFR_weekly$pos == "WR")] <- NA # drop WR DJ TUrner, keep CB DJ Turner
-```
-
-Let's check again for duplicate `game`-`player` or `player`-`season`-`week` instances:
-
-```{r}
-nfl_advancedStatsPFR_seasonal %>% 
-  group_by(pfr_id, season, team) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_advancedStatsPFR_weekly %>% 
-  arrange(game_id, pfr_id) %>% 
-  group_by(game_id, pfr_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_advancedStatsPFR_weekly %>% 
-  arrange(pfr_id, season, week) %>% 
-  group_by(pfr_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_advancedStatsPFR_seasonal,
-  file = "./data/nfl_advancedStatsPFR_seasonal.RData"
-)
-
-save(
-  nfl_advancedStatsPFR_weekly,
-  file = "./data/nfl_advancedStatsPFR_weekly.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
-load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
-```
-
-### Player Contracts {#sec-playerContracts}
-
-A Data Dictionary for player contracts data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_contracts.html>
-
-```{r}
-#| eval: false
-
-nfl_playerContracts_raw <- progressr::with_progress(
-  nflreadr::load_contracts())
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_playerContracts_raw,
-  file = "./data/nfl_playerContracts_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_playerContracts_raw.RData")
-```
-
-The `nfl_playerContracts` object is in `player`-`year`-`team`-`value` form.
-That is, each row should be uniquely identified by the combination of `otc_id`, `year_signed`, `team`, and `value`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_playerContracts <- nfl_playerContracts_raw %>% 
-  select(otc_id, year_signed, team, everything()) %>% 
-  arrange(player, otc_id, year_signed, team)
-```
-
-Let's check for duplicate `player`-`year`-`team`-`value` instances:
-
-```{r}
-nfl_playerContracts %>% 
-  group_by(otc_id, year_signed, team, value) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_playerContracts,
-  file = "./data/nfl_playerContracts.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_playerContracts.RData")
-```
-
-### FTN Charting Data {#sec-ftnCharting}
-
-A Data Dictionary for FTN Charting data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ftn_charting.html>
-
-```{r}
-#| eval: false
-
-nfl_ftnCharting_raw <- progressr::with_progress(
-  nflreadr::load_ftn_charting(seasons = TRUE))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_ftnCharting_raw,
-  file = "./data/nfl_ftnCharting_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_ftnCharting_raw.RData")
-```
-
-The `nfl_ftnCharting` object is in `game`-`play` form.
-That is, each row should be uniquely identified by the combination of `nflverse_game_id` and `play_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_ftnCharting <- nfl_ftnCharting_raw %>% 
-  select(nflverse_game_id, nflverse_play_id, everything()) %>% 
-  arrange(nflverse_game_id, nflverse_play_id)
-```
-
-Let's check for duplicate `game`-`drive`-`play` instances:
-
-```{r}
-nfl_ftnCharting %>% 
-  group_by(nflverse_game_id, nflverse_play_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_ftnCharting,
-  file = "./data/nfl_ftnCharting.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_ftnCharting.RData")
-```
-
-### Fantasy Player IDs {#sec-fantasyPlayerIDs}
-
-A Data Dictionary for fantasy player ID data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_playerids.html>
-
-```{r}
-#| eval: false
-
-nfl_playerIDs_raw <- progressr::with_progress(
-  nflreadr::load_ff_playerids())
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_playerIDs_raw,
-  file = "./data/nfl_playerIDs_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_playerIDs_raw.RData")
-```
-
-The `nfl_playerIDs` object is in `player` form.
-That is, each row should be uniquely identified by `mfl_id`.
-
-```{r}
-nfl_playerIDs <- nfl_playerIDs_raw %>% 
-  arrange(name, mfl_id)
-```
-
-Let's check for duplicate `player` instances:
-
-```{r}
-nfl_playerIDs %>% 
-  group_by(mfl_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_playerIDs %>% 
-  filter(!is.na(gsis_id)) %>% 
-  group_by(gsis_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_playerIDs,
-  file = "./data/nfl_playerIDs.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_playerIDs.RData")
-```
-
-### FantasyPros Rankings {#sec-fantasyProsRankings}
-
-A Data Dictionary for FantasyPros ranking data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_rankings.html>
-
-```{r}
-#| eval: false
-
-#nfl_rankings_raw <- progressr::with_progress( # currently throws error
-#  nflreadr::load_ff_rankings(type = "all"))
-
-nfl_rankings_draft_raw <- progressr::with_progress(
-  nflreadr::load_ff_rankings(type = "draft"))
-
-nfl_rankings_weekly_raw <- progressr::with_progress(
-  nflreadr::load_ff_rankings(type = "week"))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_rankings_draft_raw,
-  file = "./data/nfl_rankings_draft_raw.RData"
-)
-
-save(
-  nfl_rankings_weekly_raw,
-  file = "./data/nfl_rankings_weekly_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_rankings_draft_raw.RData")
-load(file = "./data/nfl_rankings_weekly_raw.RData")
-```
-
-The `nfl_rankings_draft` object is in `player`-`page_type` form.
-That is, each row should be uniquely identified by the player's `id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_rankings_draft <- nfl_rankings_draft_raw %>% 
-  select(id, page_type, player, pos, team, everything()) %>% 
-  arrange(player, id, pos, page_type)
-```
-
-Let's check for duplicate `player`-`page_type` instances:
-
-```{r}
-nfl_rankings_draft %>% 
-  group_by(id, page_type) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-The `nfl_rankings_weekly` object is in `player`-`page` form.
-That is, each row should be uniquely identified by `fantasypros_id` and `page`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_rankings_weekly <- nfl_rankings_weekly_raw %>% 
-  select(fantasypros_id, page, player_name, pos, team, everything()) %>% 
-  arrange(player_name, fantasypros_id, page, pos)
-```
-
-Let's check for duplicate `player`-`page` instances:
-
-```{r}
-nfl_rankings_weekly %>% 
-  group_by(fantasypros_id, page) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_rankings_draft,
-  file = "./data/nfl_rankings_draft.RData"
-)
-
-save(
-  nfl_rankings_weekly,
-  file = "./data/nfl_rankings_weekly.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_rankings_draft.RData")
-load(file = "./data/nfl_rankings_weekly.RData")
-```
-
-### Expected Fantasy Points {#sec-expectedFantasyPoints}
-
-A Data Dictionary for expected fantasy points data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_opportunity.html>
-
-```{r}
-#| eval: false
-
-nfl_expectedFantasyPoints_weekly_raw <- progressr::with_progress(
-  nflreadr::load_ff_opportunity(
-    seasons = TRUE,
-    stat_type = "weekly",
-    model_version = "latest"
-  ))
-
-nfl_expectedFantasyPoints_pass_raw <- progressr::with_progress(
-  nflreadr::load_ff_opportunity(
-    seasons = TRUE,
-    stat_type = "pbp_pass",
-    model_version = "latest"
-  ))
-
-nfl_expectedFantasyPoints_rush_raw <- progressr::with_progress(
-  nflreadr::load_ff_opportunity(
-    seasons = TRUE,
-    stat_type = "pbp_rush",
-    model_version = "latest"
-  ))
-```
-
-```{r}
-nfl_expectedFantasyPoints_weekly_raw <- nflreadr::load_ff_opportunity(
-  seasons = TRUE,
-  stat_type = "weekly",
-  model_version = "latest"
-  )
-
-nfl_expectedFantasyPoints_pass_raw <- nflreadr::load_ff_opportunity(
-  seasons = TRUE,
-  stat_type = "pbp_pass",
-  model_version = "latest"
-  )
-
-nfl_expectedFantasyPoints_rush_raw <- nflreadr::load_ff_opportunity(
-  seasons = TRUE,
-  stat_type = "pbp_rush",
-  model_version = "latest"
-  )
-```
-
-```{r}
-save(
-  nfl_expectedFantasyPoints_weekly_raw,
-  file = "./data/nfl_expectedFantasyPoints_weekly_raw.RData"
-)
-
-save(
-  nfl_expectedFantasyPoints_pass_raw,
-  nfl_expectedFantasyPoints_rush_raw,
-  file = "./data/nfl_expectedFantasyPoints_pbp_raw.RData"
-)
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = "./data/nfl_expectedFantasyPoints_weekly_raw.RData")
-load(file = "./data/nfl_expectedFantasyPoints_pbp_raw.RData")
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly_raw.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp_raw.RData", fsep = ""))
-```
-
-```{r}
-nfl_expectedFantasyPoints_pbp_list <- list(
-  nfl_expectedFantasyPoints_pass_raw,
-  nfl_expectedFantasyPoints_rush_raw)
-
-nfl_expectedFantasyPoints_pbp <- full_join(
-  nfl_expectedFantasyPoints_pass_raw,
-  nfl_expectedFantasyPoints_rush_raw,
-  by = c("game_id","fixed_drive","play_id"),
-  suffix = c(".pass", ".rush")
-)
-
-# Todo:
-#-separate variables from each pass/rush object
-```
-
-The `nfl_expectedFantasyPoints_weekly` object is in `game`-`player` form and in `season`-`week`-`player` form.
-That is, each row should be uniquely identified by the combination of `game_id` and `player_id`.
-Each row should also be uniquely identified by the combination of `season`, `week`, and player_id.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly_raw %>% 
-  select(season, week, game_id, player_id, full_name, posteam, position, everything()) %>% 
-  arrange(full_name, player_id, season, week)
-```
-
-Let's check for duplicate `game`-`player` instances and `season`-`week`-`player` instances:
-
-```{r}
-nfl_expectedFantasyPoints_weekly %>% 
-  group_by(game_id, player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_expectedFantasyPoints_weekly %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's do some data cleanup:
-
-```{r}
-# Drop players with missing values for player_id
-nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
-  filter(!is.na(player_id))
-```
-
-Let's check again for duplicate `game`-`player` instances and `season`-`week`-`player` instances:
-
-```{r}
-nfl_expectedFantasyPoints_weekly %>% 
-  group_by(game_id, player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_expectedFantasyPoints_weekly %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-The `nfl_expectedFantasyPoints_pbp` object is in `game`-`drive`-`play` form.
-That is, each row should be uniquely identified by the combination of `game_id`, `fixed_drive`, and `play_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_expectedFantasyPoints_pbp <- nfl_expectedFantasyPoints_pbp %>% 
-  select(game_id, fixed_drive, play_id, everything()) %>% 
-  arrange(game_id, fixed_drive, play_id)
-```
-
-Let's check for duplicate `game`-`player` instances and `season`-`week`-`player` instances:
-
-```{r}
-nfl_expectedFantasyPoints_pbp %>% 
-  group_by(game_id, fixed_drive, play_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-save(
-  nfl_expectedFantasyPoints_weekly,
-  file = "./data/nfl_expectedFantasyPoints_weekly.RData"
-)
-
-save(
-  nfl_expectedFantasyPoints_pbp,
-  file = "./data/nfl_expectedFantasyPoints_pbp.RData"
-)
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
-load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
-```
-
-## Data Dictionary {#sec-dataDictionary}
-
-Data Dictionaries are metadata that describe the meaning of the variables in a datset.
-You can find Data Dictionaries for the various NFL datasets at the following link: <https://nflreadr.nflverse.com/articles/index.html>.
-
 ## Create a Data Frame {#sec-createDF}
@@
 
-Here are the data in the `nfl_actualStats_offense_weekly` object.
+Here are the data in the `nfl_actualStats_offense_weekly` [UPDATE] object.
 The data are structured in "player-season-week form".
@@
 ```{r}
-dataLong <- nfl_actualStats_offense_weekly %>% 
-  select(player_id, player_display_name, season, week, fantasy_points)
-
-dim(dataLong)
-
-names(dataLong)
+# UPDATE
+
+#dataLong <- nfl_actualStats_offense_weekly %>% 
+#  select(player_id, player_display_name, season, week, fantasy_points)
+#
+#dim(dataLong)
+#
+#names(dataLong)
 ```
@@
 ```{r}
-dataWide <- dataLong %>% 
-  pivot_wider(
-    names_from = c(season, week),
-    names_glue = "{.value}_{season}_week{week}",
-    values_from = fantasy_points)
-
-dim(dataWide)
-
-names(dataWide)
+# UPDATE
+
+#dataWide <- dataLong %>% 
+#  pivot_wider(
+#    names_from = c(season, week),
+#    names_glue = "{.value}_{season}_week{week}",
+#    values_from = fantasy_points)
+#
+#dim(dataWide)
+#
+#names(dataWide)
 ```
@@
 ```{r}
-dataWide <- nfl_actualStats_offense_weekly %>% 
-  select(player_id, player_display_name, season, week, recent_team, opponent_team)
-
-dim(dataWide)
-
-names(dataWide)
+# UPDATE
+
+#dataWide <- nfl_actualStats_offense_weekly %>% 
+#  select(player_id, player_display_name, season, week, recent_team, opponent_team)
+#
+#dim(dataWide)
+#
+#names(dataWide)
 ```
@@
 ```{r}
-dataLong <- dataWide %>% 
-  pivot_longer(
-    cols = c(recent_team, opponent_team),
-    names_to = "role",
-    values_to = "team")
-
-dim(dataLong)
-
-names(dataLong)
+# UPDATE
+
+#dataLong <- dataWide %>% 
+#  pivot_longer(
+#    cols = c(recent_team, opponent_team),
+#    names_to = "role",
+#    values_to = "team")
+#
+#dim(dataLong)
+#
+#names(dataLong)
 ```
@@
 }
-```
-
-## Calculations {#sec-calculations}
-
-### Historical Actual Player Statistics {#sec-calculateActualStats}
-
-In addition to week-by-week actual player statistics, we can also compute historical actual player statistics as a function of different timeframes, including season-by-season and career statistics.
-
-#### Career Statistics {#sec-calculateActualCareerStats}
-
-First, we can compute the players' career statistics using the `calculate_player_stats()`, `calculate_player_stats_def()`, and `calculate_player_stats_kicking()` functions from the `nflfastR` package for offensive players, defensive players, and kickers, respectively.
-
-::: {#nte-calculateActualCareerStats .callout-note title="Calculating players' career statistics"}
-Note: the following code takes a while to run.
-:::
-
-```{r}
-#| eval: false
-
-nfl_actualStats_offense_career_raw <- nflfastR::calculate_player_stats(
-  nfl_pbp,
-  weekly = FALSE)
-
-nfl_actualStats_defense_career_raw <- nflfastR::calculate_player_stats_def(
-  nfl_pbp,
-  weekly = FALSE)
-
-nfl_actualStats_kicking_career_raw <- nflfastR::calculate_player_stats_kicking(
-  nfl_pbp,
-  weekly = FALSE)
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_offense_career_raw, nfl_actualStats_defense_career_raw, nfl_actualStats_kicking_career_raw,
-  file = "./data/nfl_actualStats_career_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load("./data/nfl_actualStats_career_raw.RData")
-```
-
-The `nfl_actualStats_offense_career`, `nfl_actualStats_defense_career`, and `nfl_actualStats_kicking_career` objects are in `player` form.
-That is, each row should be uniquely identified by the combination of `player_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_actualStats_offense_career <- nfl_actualStats_offense_career_raw %>% 
-  arrange(player_display_name, player_id)
-
-nfl_actualStats_defense_career <- nfl_actualStats_defense_career_raw %>% 
-  arrange(player_display_name, player_id)
-
-nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career_raw %>% 
-  arrange(player_display_name, player_id)
-```
-
-Let's check for duplicate `player` instances:
-
-```{r}
-nfl_actualStats_offense_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_defense_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_kicking_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's do some data cleanup:
-
-```{r}
-# Sum statistics across a player's years with multiple teams
-nfl_actualStats_defense_career <- nfl_actualStats_defense_career %>% 
-  group_by(player_id) %>% 
-  summarise(
-    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
-    .groups = "drop"
-  )
-
-nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career %>% 
-  group_by(player_id) %>% 
-  summarise(
-    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
-    .groups = "drop"
-  )
-
-# Re-calculate percentage stats
-nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career %>% 
-  mutate(
-    fg_pct = fg_made / fg_att,
-    pat_pct = pat_made / pat_att
-  )
-
-# Merge data back with player info
-nfl_actualStats_defense_career_raw_playerInfo <- nfl_actualStats_defense_career_raw %>% 
-  select(player_id, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
-  unique() %>% 
-  group_by(player_id) %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-
-nfl_actualStats_kicking_career_raw_playerInfo <- nfl_actualStats_kicking_career_raw %>% 
-  select(player_id, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
-  unique() %>% 
-  group_by(player_id) %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-
-nfl_actualStats_defense_career <- nfl_actualStats_defense_career_raw_playerInfo %>% 
-  right_join(
-    nfl_actualStats_defense_career,
-    by = "player_id"
-  )
-
-nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career_raw_playerInfo %>% 
-  right_join(
-    nfl_actualStats_kicking_career,
-    by = "player_id"
-  )
-
-# Rearrange data
-nfl_actualStats_defense_career <- nfl_actualStats_defense_career %>% 
-  arrange(player_display_name, player_id)
-
-nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career %>% 
-  arrange(player_display_name, player_id)
-```
-
-Let's check again for duplicate `player` instances:
-
-```{r}
-nfl_actualStats_offense_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_defense_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_kicking_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-save(
-  nfl_actualStats_offense_career, nfl_actualStats_defense_career, nfl_actualStats_kicking_career,
-  file = "./data/nfl_actualStats_career.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_career.RData")
-```
-
-#### Season-by-Season Statistics {#sec-calculateActualSeasonalStats}
-
-Second, we can compute the players' season-by-season statistics.
-
-```{r}
-seasons <- unique(nfl_pbp$season)
-
-nfl_pbp_seasonalList <- list()
-nfl_actualStats_offense_seasonalList <- list()
-nfl_actualStats_defense_seasonalList <- list()
-nfl_actualStats_kicking_seasonalList <- list()
-```
-
-::: {#nte-calculateActualSeasonalStats .callout-note title="Calculating players' season-by-season statistics"}
-Note: the following code takes a while to run.
-:::
-
-```{r}
-#| eval: false
-
-pb <- txtProgressBar(
-  min = 0,
-  max = length(seasons),
-  style = 3)
-
-for(i in 1:length(seasons)){
-  # Subset play-by-play data by season
-  nfl_pbp_seasonalList[[i]] <- nfl_pbp %>% 
-    dplyr::filter(season == seasons[i])
-
-  # Compute actual statistics by season
-  nfl_actualStats_offense_seasonalList[[i]] <- 
-    nflfastR::calculate_player_stats(
-      nfl_pbp_seasonalList[[i]],
-      weekly = FALSE)
-
-  nfl_actualStats_defense_seasonalList[[i]] <- 
-    nflfastR::calculate_player_stats_def(
-      nfl_pbp_seasonalList[[i]],
-      weekly = FALSE)
-
-  nfl_actualStats_kicking_seasonalList[[i]] <- 
-    nflfastR::calculate_player_stats_kicking(
-      nfl_pbp_seasonalList[[i]],
-      weekly = FALSE)
-
-  nfl_actualStats_offense_seasonalList[[i]]$season <- seasons[i]
-  nfl_actualStats_defense_seasonalList[[i]]$season <- seasons[i]
-  nfl_actualStats_kicking_seasonalList[[i]]$season <- seasons[i]
-
-  print(
-    paste("Completed computing projections for season: ", seasons[i], sep = ""))
-
-  # Update the progress bar
-  setTxtProgressBar(pb, i)
-}
-
-# Close the progress bar
-close(pb)
-
-nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonalList %>% 
-  dplyr::bind_rows()
-nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonalList %>% 
-  dplyr::bind_rows()
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonalList %>% 
-  dplyr::bind_rows()
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-save(
-  nfl_actualStats_offense_seasonal, nfl_actualStats_defense_seasonal, nfl_actualStats_kicking_seasonal,
-  file = "./data/nfl_actualStats_seasonal.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_seasonal.RData")
-```
-
-#### Week-by-Week Statistics {#sec-calculateActualWeeklyStats}
-
-We already load players' week-by-week statistics [above](#sec-downloadActualStats).
-Nevertheless, we could compute players' weekly statistics from the play-by-play data using the following syntax:
-
-```{r}
-#| eval: false
-
-nfl_actualStats_offense_weekly <- nflfastR::calculate_player_stats(
-  nfl_pbp,
-  weekly = TRUE)
-
-nfl_actualStats_defense_weekly <- nflfastR::calculate_player_stats_def(
-  nfl_pbp,
-  weekly = TRUE)
-
-nfl_actualStats_kicking_weekly <- nflfastR::calculate_player_stats_kicking(
-  nfl_pbp,
-  weekly = TRUE)
-```
-
-### Historical Actual Fantasy Points {#sec-calculateActualPoints}
-
-Specify scoring settings:
-
-#### Weekly {#sec-calculateActualPointsWeekly}
-
-#### Seasonal {#sec-calculateActualPointsSeasonal}
-
-#### Career {#sec-calculateActualPointsCareer}
-
-### Player Age and Experience {#sec-calculatePlayerAge}
-
-#### Weekly {#sec-calculatePlayerAgeWeekly}
-
-```{r}
-# Reshape from wide to long format
-nfl_actualStats_offense_weekly_long <- nfl_actualStats_offense_weekly %>% 
-  tidyr::pivot_longer(
-    cols = c(recent_team, opponent_team),
-    names_to = "role",
-    values_to = "team")
-
-# Perform separate inner join operations for the home_team and away_team
-nfl_actualStats_offense_weekly_home <- dplyr::inner_join(
-  nfl_actualStats_offense_weekly_long,
-  nfl_schedules,
-  by = c("season","week","team" = "home_team")) %>% 
-  mutate(home_away = "home_team")
-
-nfl_actualStats_offense_weekly_away <- dplyr::inner_join(
-  nfl_actualStats_offense_weekly_long,
-  nfl_schedules,
-  by = c("season","week","team" = "away_team")) %>% 
-  mutate(home_away = "away_team")
-
-nfl_actualStats_defense_weekly_home <- dplyr::inner_join(
-  nfl_actualStats_defense_weekly,
-  nfl_schedules,
-  by = c("season","week","team" = "home_team")) %>% 
-  mutate(home_away = "home_team")
-
-nfl_actualStats_defense_weekly_away <- dplyr::inner_join(
-  nfl_actualStats_defense_weekly,
-  nfl_schedules,
-  by = c("season","week","team" = "away_team")) %>% 
-  mutate(home_away = "away_team")
-
-nfl_actualStats_kicking_weekly_home <- dplyr::inner_join(
-  nfl_actualStats_kicking_weekly,
-  nfl_schedules,
-  by = c("season","week","team" = "home_team")) %>% 
-  mutate(home_away = "home_team")
-
-nfl_actualStats_kicking_weekly_away <- dplyr::inner_join(
-  nfl_actualStats_kicking_weekly,
-  nfl_schedules,
-  by = c("season","week","team" = "away_team")) %>% 
-  mutate(home_away = "away_team")
-
-# Combine the results of the join operations
-nfl_actualStats_offense_weekly_schedules_long <- dplyr::bind_rows(
-  nfl_actualStats_offense_weekly_home,
-  nfl_actualStats_offense_weekly_away)
-
-nfl_actualStats_defense_weekly_schedules_long <- dplyr::bind_rows(
-  nfl_actualStats_defense_weekly_home,
-  nfl_actualStats_defense_weekly_away)
-
-nfl_actualStats_kicking_weekly_schedules_long <- dplyr::bind_rows(
-  nfl_actualStats_kicking_weekly_home,
-  nfl_actualStats_kicking_weekly_away)
-
-# Reshape from long to wide
-player_game_gameday_offense <- nfl_actualStats_offense_weekly_schedules_long %>%
-  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
-  tidyr::pivot_wider(
-    names_from = home_away,
-    values_from = team)
-
-player_game_gameday_defense <- nfl_actualStats_defense_weekly_schedules_long %>%
-  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
-  tidyr::pivot_wider(
-    names_from = home_away,
-    values_from = team)
-
-player_game_gameday_kicking <- nfl_actualStats_kicking_weekly_schedules_long %>%
-  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
-  tidyr::pivot_wider(
-    names_from = home_away,
-    values_from = team)
-
-# Merge player birthdate and the game date
-player_game_birthdate_gameday_offense <- dplyr::left_join(
-  player_game_gameday_offense,
-  unique(nfl_players[,c("gsis_id","birth_date")]),
-  by = c("player_id" = "gsis_id")
-)
-
-player_game_birthdate_gameday_defense <- dplyr::left_join(
-  player_game_gameday_defense,
-  unique(nfl_players[,c("gsis_id","birth_date")]),
-  by = c("player_id" = "gsis_id")
-)
-
-player_game_birthdate_gameday_kicking <- dplyr::left_join(
-  player_game_gameday_kicking,
-  unique(nfl_players[,c("gsis_id","birth_date")]),
-  by = c("player_id" = "gsis_id")
-)
-
-player_game_birthdate_gameday_offense$birth_date <- lubridate::ymd(player_game_birthdate_gameday_offense$birth_date)
-player_game_birthdate_gameday_offense$gameday <- lubridate::ymd(player_game_birthdate_gameday_offense$gameday)
-
-player_game_birthdate_gameday_defense$birth_date <- lubridate::ymd(player_game_birthdate_gameday_defense$birth_date)
-player_game_birthdate_gameday_defense$gameday <- lubridate::ymd(player_game_birthdate_gameday_defense$gameday)
-
-player_game_birthdate_gameday_kicking$birth_date <- lubridate::ymd(player_game_birthdate_gameday_kicking$birth_date)
-player_game_birthdate_gameday_kicking$gameday <- lubridate::ymd(player_game_birthdate_gameday_kicking$gameday)
-
-# Calculate player's age for a given week as the difference between their birthdate and the game date
-player_game_birthdate_gameday_offense$age <- lubridate::interval(
-  start = player_game_birthdate_gameday_offense$birth_date,
-  end = player_game_birthdate_gameday_offense$gameday
-) %>% 
-  lubridate::time_length(unit = "years")
-
-player_game_birthdate_gameday_defense$age <- lubridate::interval(
-  start = player_game_birthdate_gameday_defense$birth_date,
-  end = player_game_birthdate_gameday_defense$gameday
-) %>% 
-  lubridate::time_length(unit = "years")
-
-player_game_birthdate_gameday_kicking$age <- lubridate::interval(
-  start = player_game_birthdate_gameday_kicking$birth_date,
-  end = player_game_birthdate_gameday_kicking$gameday
-) %>% 
-  lubridate::time_length(unit = "years")
-
-# Merge with Pro Football Reference Data on Player Age by Season
-player_game_birthdate_gameday_offense <- player_game_birthdate_gameday_offense %>% 
-  dplyr::left_join(
-    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
-    by = c("player_id" = "gsis_id", "season")
-  )
-
-player_game_birthdate_gameday_defense <- player_game_birthdate_gameday_defense %>% 
-  dplyr::left_join(
-    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
-    by = c("player_id" = "gsis_id", "season")
-  )
-
-player_game_birthdate_gameday_kicking <- player_game_birthdate_gameday_kicking %>% 
-  dplyr::left_join(
-    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
-    by = c("player_id" = "gsis_id", "season")
-  )
-
-# Set age as first non-missing value from calculation above or from PFR
-player_game_birthdate_gameday_offense <- player_game_birthdate_gameday_offense %>% 
-  mutate(age = coalesce(age.x, age.y)) %>% 
-  select(-age.x, -age.y)
-
-player_game_birthdate_gameday_defense <- player_game_birthdate_gameday_defense %>% 
-  mutate(age = coalesce(age.x, age.y)) %>% 
-  select(-age.x, -age.y)
-
-player_game_birthdate_gameday_kicking <- player_game_birthdate_gameday_kicking %>% 
-  mutate(age = coalesce(age.x, age.y)) %>% 
-  select(-age.x, -age.y)
-
-# Calculate ageCentered and ageCenteredQuadratic
-player_game_birthdate_gameday_offense$ageCentered20 <- player_game_birthdate_gameday_offense$age - 20
-player_game_birthdate_gameday_offense$ageCentered20Quadratic <- player_game_birthdate_gameday_offense$ageCentered20 ^ 2
-
-player_game_birthdate_gameday_defense$ageCentered20 <- player_game_birthdate_gameday_defense$age - 20
-player_game_birthdate_gameday_defense$ageCentered20Quadratic <- player_game_birthdate_gameday_defense$ageCentered20 ^ 2
-
-player_game_birthdate_gameday_kicking$ageCentered20 <- player_game_birthdate_gameday_kicking$age - 20
-player_game_birthdate_gameday_kicking$ageCentered20Quadratic <- player_game_birthdate_gameday_kicking$ageCentered20 ^ 2
-
-# Merge with player info
-player_age_offense <- dplyr::left_join(
-  player_game_birthdate_gameday_offense,
-  nfl_players %>% select(-birth_date, -season),
-  by = c("player_id" = "gsis_id"))
-
-player_age_defense <- dplyr::left_join(
-  player_game_birthdate_gameday_defense,
-  nfl_players %>% select(-birth_date, -season),
-  by = c("player_id" = "gsis_id"))
-
-player_age_kicking <- dplyr::left_join(
-  player_game_birthdate_gameday_kicking,
-  nfl_players %>% select(-birth_date, -season),
-  by = c("player_id" = "gsis_id"))
-
-# Add game_id to weekly stats to facilitate merging
-nfl_actualStats_game_offense_weekly <- nfl_actualStats_offense_weekly %>% 
-  dplyr::left_join(
-    player_age_offense[,c("season","week","player_id","game_id")],
-    by = c("season","week","player_id"))
-
-nfl_actualStats_game_defense_weekly <- nfl_actualStats_defense_weekly %>% 
-  dplyr::left_join(
-    player_age_offense[,c("season","week","player_id","game_id")],
-    by = c("season","week","player_id"))
-
-nfl_actualStats_game_kicking_weekly <- nfl_actualStats_kicking_weekly %>% 
-  dplyr::left_join(
-    player_age_offense[,c("season","week","player_id","game_id")],
-    by = c("season","week","player_id"))
-
-# Merge with player weekly stats
-player_stats_weekly_offense <- dplyr::full_join(
-  player_age_offense %>% select(-position, -position_group),
-  nfl_actualStats_game_offense_weekly,
-  by = c("season","week","player_id","game_id"))
-
-player_stats_weekly_defense <- dplyr::full_join(
-  player_age_defense %>% select(-position, -position_group),
-  nfl_actualStats_game_defense_weekly,
-  by = c("season","week","player_id","game_id"))
-
-player_stats_weekly_kicking <- dplyr::full_join(
-  player_age_kicking %>% select(-position, -position_group),
-  nfl_actualStats_game_kicking_weekly,
-  by = c("season","week","player_id","game_id"))
-
-player_stats_weekly_offense$total_years_of_experience <- as.integer(player_stats_weekly_offense$years_of_experience)
-player_stats_weekly_defense$total_years_of_experience <- as.integer(player_stats_weekly_defense$years_of_experience)
-player_stats_weekly_kicking$total_years_of_experience <- as.integer(player_stats_weekly_kicking$years_of_experience)
-
-player_stats_weekly_offense$years_of_experience <- NULL
-player_stats_weekly_defense$years_of_experience <- NULL
-player_stats_weekly_kicking$years_of_experience <- NULL
-
-distinct_seasons_offense <- player_stats_weekly_offense %>%
-  dplyr::select(player_id, season) %>%
-  dplyr::distinct() %>% 
-  dplyr::left_join(
-    nfl_players[,c("gsis_id","years_of_experience")],
-    by = c("player_id" = "gsis_id")
-  ) %>% 
-  dplyr::mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
-  dplyr::select(-years_of_experience)
-
-distinct_seasons_defense <- player_stats_weekly_defense %>%
-  dplyr::select(player_id, season) %>%
-  dplyr::distinct() %>% 
-  dplyr::left_join(
-    nfl_players[,c("gsis_id","years_of_experience")],
-    by = c("player_id" = "gsis_id")
-  ) %>% 
-  dplyr::mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
-  dplyr::select(-years_of_experience)
-
-distinct_seasons_kicking <- player_stats_weekly_kicking %>%
-  dplyr::select(player_id, season) %>%
-  dplyr::distinct() %>% 
-  dplyr::left_join(
-    nfl_players[,c("gsis_id","years_of_experience")],
-    by = c("player_id" = "gsis_id")
-  ) %>% 
-  dplyr::mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
-  dplyr::select(-years_of_experience)
-
-years_of_experience_offense <- distinct_seasons_offense %>% 
-  dplyr::arrange(player_id, -season) %>% 
-  dplyr::group_by(player_id) %>%
-  dplyr::mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
-  dplyr::ungroup()
-
-years_of_experience_defense <- distinct_seasons_defense %>% 
-  dplyr::arrange(player_id, -season) %>% 
-  dplyr::group_by(player_id) %>%
-  dplyr::mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
-  dplyr::ungroup()
-
-years_of_experience_kicking <- distinct_seasons_kicking %>% 
-  dplyr::arrange(player_id, -season) %>% 
-  dplyr::group_by(player_id) %>%
-  dplyr::mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
-  dplyr::ungroup()
-
-years_of_experience_offense$years_of_experience[which(years_of_experience_offense$years_of_experience < 0)] <- 0
-years_of_experience_defense$years_of_experience[which(years_of_experience_defense$years_of_experience < 0)] <- 0
-years_of_experience_kicking$years_of_experience[which(years_of_experience_kicking$years_of_experience < 0)] <- 0
-
-player_stats_weekly_offense <- player_stats_weekly_offense %>% 
-  dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
-    by = c("player_id","season")
-  )
-
-player_stats_weekly_defense <- player_stats_weekly_defense %>% 
-  dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
-    by = c("player_id","season")
-  )
-
-player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
-  dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
-    by = c("player_id","season")
-  )
-
-# Arrange data
-player_stats_weekly_offense <- player_stats_weekly_offense %>% 
-  arrange(player_display_name, season, week)
-
-player_stats_weekly_defense <- player_stats_weekly_defense %>% 
-  arrange(player_display_name, season, week)
-
-player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
-  arrange(player_display_name, season, week)
-```
-
-```{r}
-# Save data
-save(
-  player_stats_weekly_offense, player_stats_weekly_defense, player_stats_weekly_kicking,
-  file = "./data/player_stats_weekly.RData"
-)
-```
-
-#### Seasonal {#sec-calculatePlayerAgeSeasonal}
-
-```{r}
-# Merge player info with seasonal stats
-player_stats_seasonal_offense <- dplyr::full_join(
-  nfl_actualStats_offense_seasonal,
-  nfl_players %>% select(-position, -position_group, -season),
-  by = c("player_id" = "gsis_id")
-)
-
-player_stats_seasonal_defense <- dplyr::full_join(
-  nfl_actualStats_defense_seasonal,
-  nfl_players %>% select(-position, -position_group, -season),
-  by = c("player_id" = "gsis_id")
-)
-
-player_stats_seasonal_kicking <- dplyr::full_join(
-  nfl_actualStats_kicking_seasonal,
-  nfl_players %>% select(-position, -position_group, -season),
-  by = c("player_id" = "gsis_id")
-)
-
-# Calculate age
-season_startdate <- nfl_schedules %>% 
-  dplyr::group_by(season) %>% 
-  dplyr::summarise(startdate = min(gameday, na.rm = TRUE))
-
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
-  dplyr::left_join(
-    season_startdate,
-    by = "season"
-  )
-
-player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  dplyr::left_join(
-    season_startdate,
-    by = "season"
-  )
-
-player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  dplyr::left_join(
-    season_startdate,
-    by = "season"
-  )
-
-player_stats_seasonal_offense$age <- lubridate::interval(
-  start = player_stats_seasonal_offense$birth_date,
-  end = player_stats_seasonal_offense$startdate
-) %>% 
-  lubridate::time_length(unit = "years")
-
-player_stats_seasonal_defense$age <- lubridate::interval(
-  start = player_stats_seasonal_defense$birth_date,
-  end = player_stats_seasonal_defense$startdate
-) %>% 
-  lubridate::time_length(unit = "years")
-
-player_stats_seasonal_kicking$age <- lubridate::interval(
-  start = player_stats_seasonal_kicking$birth_date,
-  end = player_stats_seasonal_kicking$startdate
-) %>% 
-  lubridate::time_length(unit = "years")
-
-# Merge with Pro Football Reference Data on Player Age by Season
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
-  dplyr::left_join(
-    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
-    by = c("player_id" = "gsis_id", "season")
-  )
-
-player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  dplyr::left_join(
-    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
-    by = c("player_id" = "gsis_id", "season")
-  )
-
-player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  dplyr::left_join(
-    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
-    by = c("player_id" = "gsis_id", "season")
-  )
-
-# Set age as first non-missing value from calculation above or from PFR
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
-  mutate(age = coalesce(age.x, age.y)) %>% 
-  select(-age.x, -age.y)
-
-player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  mutate(age = coalesce(age.x, age.y)) %>% 
-  select(-age.x, -age.y)
-
-player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  mutate(age = coalesce(age.x, age.y)) %>% 
-  select(-age.x, -age.y)
-
-# Calculate ageCentered and ageCenteredQuadratic
-player_stats_seasonal_offense$ageCentered20 <- player_stats_seasonal_offense$age - 20
-player_stats_seasonal_offense$ageCentered20Quadratic <- player_stats_seasonal_offense$ageCentered20 ^ 2
-
-player_stats_seasonal_defense$ageCentered20 <- player_stats_seasonal_defense$age - 20
-player_stats_seasonal_defense$ageCentered20Quadratic <- player_stats_seasonal_defense$ageCentered20 ^ 2
-
-player_stats_seasonal_kicking$ageCentered20 <- player_stats_seasonal_kicking$age - 20
-player_stats_seasonal_kicking$ageCentered20Quadratic <- player_stats_seasonal_kicking$ageCentered20 ^ 2
-
-# Years of experience
-player_stats_seasonal_offense$years_of_experience <- NULL
-player_stats_seasonal_defense$years_of_experience <- NULL
-player_stats_seasonal_kicking$years_of_experience <- NULL
-
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
-  dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
-    by = c("player_id","season")
-  )
-
-player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
-    by = c("player_id","season")
-  )
-
-player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
-    by = c("player_id","season")
-  )
-
-# Arrange data
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
-  relocate(season, .after = player_id) %>% 
-  arrange(player_display_name, season)
-
-player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  relocate(season, .after = player_id) %>% 
-  arrange(player_display_name, season)
-
-player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  relocate(season, .after = player_id) %>% 
-  arrange(player_display_name, season)
-```
-
-```{r}
-# Save data
-save(
-  player_stats_seasonal_offense, player_stats_seasonal_defense, player_stats_seasonal_kicking,
-  file = "./data/player_stats_seasonal.RData"
-)
 ```

### 2024-08-02 download-football-data.qmd
20240802 - fix indentation
20240802 - indentation and cleanup

@@
 
-The `nfl_depthCharts` object is in `id`-`season`-`week`-`position` form.
+The `nfl_depthCharts` object is in `player`-`season`-`week`-`position` form.
 That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, `week`, and `depth_position`.
@@
 
-Let's check for duplicate `id`-`season`-`week`-`position` instances:
+Let's check for duplicate `player`-`season`-`week`-`position` instances:
 
@@
 
-Let's check again for duplicate `id`-`season`-`week`-`position` instances:
+Let's check again for duplicate `player`-`season`-`week`-`position` instances:
 
@@
 
-The `nfl_expectedFantasyPoints_weekly` object is in `game`-`player` form and in `season`-`week`-`player` form.
+The `nfl_expectedFantasyPoints_weekly` object is in `game`-`player` form and in `player`-`season`-`week` form.
 That is, each row should be uniquely identified by the combination of `game_id` and `player_id`.
-Each row should also be uniquely identified by the combination of `season`, `week`, and player_id.
+Each row should also be uniquely identified by the combination of `player_id`, `season`, and `week`.
 Let's rearrange the data accordingly:
@@
 nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly_raw %>% 
-  select(season, week, game_id, player_id, full_name, posteam, position, everything()) %>% 
+  select(player_id, season, week, game_id, full_name, posteam, position, everything()) %>% 
   arrange(full_name, player_id, season, week)
@@
 
-Let's check for duplicate `game`-`player` instances and `season`-`week`-`player` instances:
+Let's check for duplicate `game`-`player` instances and `player`-`season`-`week` instances:
 

### 2024-08-02 machine-learning.qmd
20240802 - indentation and cleanup

@@
 
-# Downloaded Data
+# Downloaded Data - Processed
 load(file = "./data/nfl_players.RData")
@@
 load(file = "./data/nfl_playerIDs.RData")
-load(file = "./data/nfl_rankings.RData")
-load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
-load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
+load(file = "./data/nfl_rankings_draft.RData")
+load(file = "./data/nfl_rankings_weekly.RData")
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
 
-# Calculated Data
+# Calculated Data - Processed
 load(file = "./data/nfl_actualStats_career.RData")
@@
 ```{r}
-# Downloaded Data
+# Downloaded Data - Processed
 load(file = "./data/nfl_players.RData")
@@
 load(file = "./data/nfl_snapCounts.RData")
-load(file = "./data/nfl_espnQBR.RData")
+load(file = "./data/nfl_espnQBR_seasonal.RData")
+load(file = "./data/nfl_espnQBR_weekly.RData")
 load(file = "./data/nfl_nextGenStats_weekly.RData")
-load(file = "./data/nfl_advancedStatsPFR.RData")
+load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
+load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
 load(file = "./data/nfl_playerContracts.RData")
@@
 
+## Rename other variables to ensure common names
+player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+  rename(team = team_abbr)
+
 ## Ensure variables with the same name have the same type
@@
   )
+
+# Check duplicate ids
+player_stats_seasonal_offense %>% 
+  group_by(gsis_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_advancedStatsPFR_seasonal %>% 
+  group_by(gsis_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
 ```
@@
 #-todo: remove redundant variables
-idListToMerge <- list(
+playerListToMerge <- list(
   nfl_players %>% filter(!is.na(gsis_id)) %>% select(-season, -years_of_experience),
@@
 
-idSeasonListToMerge <- list(
-  nfl_rosters %>% filter(!is.na(gsis_id), !is.na(season)) %>% select(-week),
-  player_stats_seasonal_offense %>% filter(!is.na(gsis_id), !is.na(season)), #nfl_actualStats_offense_seasonal %>% filter(!is.na(gsis_id), !is.na(season)),
-  nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season))
+playerSeasonListToMerge <- list(
+  player_stats_seasonal_offense %>% filter(!is.na(gsis_id), !is.na(season), !is.na(team))
 )
 
-idSeasonWeekListToMerge <- list(
+playerSeasonTeamListToMerge <- list(
+  nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(team))
+)
+
+playerSeasonWeekListToMerge <- list(
   nfl_rosters_weekly %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week)),
-  nfl_depthCharts %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week)),
   #nfl_actualStats_offense_weekly,
@@
 
+playerSeasonWeekPositionListToMerge <- list(
+  nfl_depthCharts %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week)),
+)
+
 # Merge data
-idMerged <- idListToMerge %>% 
+playerMerged <- playerListToMerge %>% 
   reduce(
@@
 
-idSeasonMerged <- idSeasonListToMerge %>% 
+playerSeasonMerged <- playerSeasonListToMerge %>% 
   reduce(
@@
 
-idSeasonWeekMerged <- idSeasonWeekListToMerge %>% 
+playerSeasonTeamMerged <- playerSeasonTeamListToMerge %>% 
+  reduce(
+    powerjoin::power_full_join,
+    by = c("gsis_id","season"),
+    conflict = coalesce_xy)
+
+playerSeasonWeekMerged <- playerSeasonWeekListToMerge %>% 
   reduce(

### 2024-08-02 download-football-data.qmd
20240802 - fix duplicate identifier
20240802 - update indentation

@@
 
-## Load Packages {#sec-loadPackages}
+## Load Packages {#sec-downloadFootballDataLoadPackages}
 
@@
 
-nfl_4thdown_raw <- nfl4th::load_4th_pbp(seasons = 2014:nflreadr::most_recent_season())
+nfl_4thdown_raw <- nfl4th::load_4th_pbp(
+  seasons = 2014:nflreadr::most_recent_season())
 ```
@@
     ),
-    ,
     season = coalesce(

### 2024-08-03 getting-started.qmd
20240802 - remove packages that are no longer needed for this chapter
20240802 - transform between long and wide form

@@
 
-For instance, you can use the following code to install the `nflreadr` package:
+For instance, you can use the following code to install the `tidyverse` package:
 
@@
 
-install.packages("nflreadr")
+install.packages("tidyverse")
 ```
@@
 ```{r}
-library("ffanalytics")
-library("nflreadr")
-library("nflfastR")
-library("nfl4th")
-library("nflplotR")
-library("progressr")
-library("lubridate")
 library("tidyverse")
@@
   fantasyPoints = c(250, 170, 65, 15)
+)
+
+fantasyPoints_weekly = expand.grid(
+  ID = 1:12,
+  season = c(2022, 2023),
+  week = 1:17
+)
+
+set.seed(52242)
+fantasyPoints_weekly$fantasyPoints <- sample(
+  0:35,
+  size = nrow(fantasyPoints_weekly),
+  replace = TRUE
 )
@@
 
-Here are the data in the `nfl_actualStats_offense_weekly` [UPDATE] object.
+Here are the original data in long form.
 The data are structured in "player-season-week form".
-That is, every row in the dataset is uniquely identified by the variables, `player_id`, `season`, and `week`.
+That is, every row in the dataset is uniquely identified by the combination of variables, `ID`, `season`, and `week`.
 This is an example of long form, because each player has multiple rows.
 
-Original data:
+```{r}
+dataLong <- full_join(
+  players %>% select(-age),
+  fantasyPoints_weekly,
+  by = c("ID")
+)
+
+dataLong
+
+dim(dataLong)
+
+names(dataLong)
+```
+
+Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "player form" (where each row is uniquely identified by the `ID` variable):
 
 ```{r}
-# UPDATE
+dataWide <- dataLong %>% 
+  pivot_wider(
+    names_from = c(season, week),
+    names_glue = "{.value}_{season}_week{week}",
+    values_from = fantasyPoints)
 
-#dataLong <- nfl_actualStats_offense_weekly %>% 
-#  select(player_id, player_display_name, season, week, fantasy_points)
-#
-#dim(dataLong)
-#
-#names(dataLong)
-```
+dataWide
 
-Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "player form" (where each row is uniquely identified by the `player_id` variable):
+dim(dataWide)
 
-```{r}
-# UPDATE
-
-#dataWide <- dataLong %>% 
-#  pivot_wider(
-#    names_from = c(season, week),
-#    names_glue = "{.value}_{season}_week{week}",
-#    values_from = fantasy_points)
-#
-#dim(dataWide)
-#
-#names(dataWide)
+names(dataWide)
 ```
@@
 Conversely, we can also restructure data from wide to long.
-
-Original data:
+Here are the data in long form, after they have been transformed from wide form using `tidyverse`:
 
 ```{r}
-# UPDATE
+dataLong <- dataWide %>% 
+  pivot_longer(
+    cols = fantasyPoints_2022_week1:fantasyPoints_2023_week17,
+    names_to = c("season", "week"),
+    names_pattern = "fantasyPoints_(.*)_week(.*)",
+    values_to = "fantasyPoints")
 
-#dataWide <- nfl_actualStats_offense_weekly %>% 
-#  select(player_id, player_display_name, season, week, recent_team, opponent_team)
-#
-#dim(dataWide)
-#
-#names(dataWide)
-```
+dataLong
 
-Data in long form, transformed from wide form using `tidyverse`:
+dim(dataLong)
 
-```{r}
-# UPDATE
-
-#dataLong <- dataWide %>% 
-#  pivot_longer(
-#    cols = c(recent_team, opponent_team),
-#    names_to = "role",
-#    values_to = "team")
-#
-#dim(dataLong)
-#
-#names(dataLong)
+names(dataLong)
 ```

### 2024-08-03 download-football-data.qmd
20240802 - fix indentation
20240802 - clean up player names for merging

@@
 library("ffanalytics")
+library("petersenlab")
 library("nflreadr")
@@
   mutate(
-    nameMerge = toupper(str_remove_all(player, "[[:punct:]]"))
+    nameMerge = petersenlab::cleanUpNames(player)
   ) %>% 
@@
   mutate(
-    nameMerge = toupper(str_remove_all(pfr_player_name, "[[:punct:]]"))
+    nameMerge = petersenlab::cleanUpNames(pfr_player_name)
   ) %>% 

### 2024-08-04 basic-statistics.qmd
20240803 - fixes

@@
 
-where $SS_{\text{effect}}$ is the sum of squares for the effect of interest and SS_{\text{total}} is the total sum of squares.
+where $SS_{\text{effect}}$ is the sum of squares for the effect of interest and $SS_{\text{total}}$ is the total sum of squares.
 
@@
 
-where $SS_{\text{effect}}$ is the sum of squares for the effect of interest and SS_{\text{error}} is the sum of squares for the residual error term.
+where $SS_{\text{effect}}$ is the sum of squares for the effect of interest and $SS_{\text{error}}$ is the sum of squares for the residual error term.
 

### 2024-08-05 mixed-models.qmd
20240731 - placeholder
20240805 - bayesian mixed model

@@
 library("bbmle")
+library("rstan")
+library("brms")
+library("cmdstanr") # todo: install.packages("cmdstanr", repos = c('https://stan-dev.r-universe.dev', getOption("repos"))); check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
 library("parallel")
+library("parallely")
 library("plotly")
@@
 
+### Bayesian Mixed Models {#sec-fantasyPointsByAgeExperienceModelsBayesian}
+
+#### Determine Response Distribution {#sec-fantasyPointsByAgeExperienceModelsBayesianDistribution}
+
+```{r}
+#| label: fig-histogramFantasyPoints
+#| fig-cap: "Histogram of Fantasy Points with Overlaid Density Plot and Rug Plot."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense_subset,
+  mapping = aes(
+    x = fantasy_points)
+) +
+  geom_histogram(
+    aes(y = after_stat(density)),
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  geom_density(
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  geom_rug() +
+  theme_classic()
+```
+
+Because the response distribution (i.e., fantasy points) is positively skewed, we will use a gamma response distribution.
+
+#### Prepare Data {#sec-fantasyPointsByAgeExperienceModelsBayesianPrepareData}
+
+```{r}
+player_stats_seasonal_offense_subset$fantasy_points_posOnly <- player_stats_seasonal_offense_subset$fantasy_points
+player_stats_seasonal_offense_subset$fantasy_points_posOnly[which(player_stats_seasonal_offense_subset$fantasy_points_posOnly < 0)] <- 0
+```
+
+#### Specify Model Formula {#sec-fantasyPointsByAgeExperienceModelsBayesianFormula}
+
+Specify model formula:
+
+```{r}
+#| eval: false
+
+#mgcv smooth terms: https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/smooth.terms.html
+
+#bayesianMixedModelFormula <- brms::bf(
+#  fantasy_points ~ positionFactor + s(ageCentered20, by = #positionFactor) + years_of_experience + s(ageCentered20, #player_idFactor, bs = "fs") + (1 | player_idFactor)
+#)
+
+#bayesianMixedModelFormula <- brms::bf(
+#  fantasy_points ~ positionFactor + s(ageCentered20, by = #positionFactor) + years_of_experience + (1 | player_idFactor)
+#)
+
+bayesianMixedModelFormula <- brms::bf(
+  fantasy_points_posOnly ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(ageCentered20, player_idFactor, bs = "re") + (1 | player_idFactor)
+)
+```
+
+#### Run Model {#sec-fantasyPointsByAgeExperienceModelsBayesianRunModel}
+
+Now, we can run the model.
+
+::: {#nte-bayesianMixedModel .callout-note title="Bayesian mixed model"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+
+bayesianMixedModelFit <- brms::brm(
+  formula = bayesianMixedModelFormula,
+  data = player_stats_seasonal_offense_subset,
+  family = hurdle_gamma(),
+  cores = 4,
+  save_pars = save_pars(latent = FALSE, all = FALSE),
+  threads = threading(parallelly::availableCores()),
+  backend = "cmdstanr",
+  seed = 52242,
+  silent = 0
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  bayesianMixedModelFit,
+  file = "./data/bayesianMixedModelFit.RData")
+```
+
+```{r}
+#| include: false
+
+load("./data/bayesianMixedModelFit.RData")
+```
+
+#### Model Summary {#sec-fantasyPointsByAgeExperienceModelsBayesianModelSummary}
+
+```{r}
+summary(bayesianMixedModelFit)
+brms::prior_summary(bayesianMixedModelFit)
+```
+
+#### Trace Plots {#sec-fantasyPointsByAgeExperienceModelsBayesianTrace}
+
+```{r}
+#| label: fig-bayesianTracePlots
+#| fig-cap: "Trace Plots from Bayesian Mixed Model."
+
+plot(bayesianMixedModelFit, ask = FALSE)
+```
+
+#### Posterior Predictive Check {#sec-fantasyPointsByAgeExperienceModelsBayesianPPcheck}
+
+```{r}
+#| label: fig-bayesianPosteriorPredictiveCheck
+#| fig-cap: "Posterior Predictive Check from Bayesian Mixed Model."
+
+pp_check(bayesianMixedModelFit) + 
+  ggplot2::xlim(0, 600)
+```
+
 ### Plots of Model-Implied Fantasy Points by Position and Age {#sec-fantasyPointsByAgePlots}

### 2024-08-05 mixed-models.qmd
20240805 - load data object from OSF
20240805 - load bayesian mixed model

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/bayesianMixedModelFit.RData", fsep = ""))
 ```
@@
 ```{r}
+#| eval: false
 #| include: false
 
-load("./data/bayesianMixedModelFit.RData")
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/bayesianMixedModelFit.RData", fsep = ""))
+```
+
+```{r}
+#| include: false
+
+load(url("https://osf.io/download/q6rjf/"))
 ```

### 2024-08-05 mixed-models.qmd
20240805 - misspelled package name

@@
 library("parallel")
-library("parallely")
+library("parallelly")
 library("plotly")

### 2024-08-06 cluster-analysis.qmd
20240730 - REF
20240806 - overview of cluster analysis approaches

@@
 
-There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, etc.
-In this chapter, we focus on examples using model-based clustering with the `mclust` package, which uses Gaussian finite mixture modeling.
+There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, hierarchical clustering (aka connectivity-based clustering), etc.
+An overview of approaches to cluster analysis in `R` is provided by @Kassambara2017a.
+In this chapter, we focus on examples using model-based clustering with the `mclust` package [@Scrucca2023], which uses Gaussian finite mixture modeling.
 

### 2024-08-07 cognitive-bias.qmd
20240803 - Miller (2013)
20240807 - primacy effect bias; role of cognitive biases

@@
 As described [above](#sec-cognitiveBiasOverview), cognitive biases are systematic errors in thinking.
+They lead us to be less accurate in our judgments and predictions.
+Inaccuracy may not always be bad; several cognitive biases may serve to help people feel better about themselves, including [confirmation bias](#sec-cognitiveBiasesConfirmation), [overconfidence bias](#sec-cognitiveBiasesOverconfidence), optimism bias, self-serving bias, and actor-observer bias.
+In fantasy football, however, cognitive biases and inaccurate judgments can lead to worse performance.
 Cognitive biases are often due to the use of [heuristics](#sec-heuristics).
-Examples of cognitive biases that result from one or more of these heuristics include:
+
+@Miller2013c describes cognitive biases in fantasy sports.
+Examples of cognitive biases relevant to fantasy football that result from one or more heuristics include:
 
@@
 - in-group bias
-- recency bias
+- primacy effect bias
+- recency effect bias
 - hindsight bias
@@
 For instance, if you believe that a particular player is a strong breakout candidate to be a sleeper, you may be more likely to pay attention to evidence that supports that the player will breakout and may be less likely to pay attention to evidence that indicates the player may struggle.
+That is, people tend to look for confirmation that the players they want to draft (or start) are preferred by the experts [@Miller2013c].
+Confirmation bias may lead to "going with your gut" (i.e., making a decision based on your intuition).
 
@@
 
-### Recency Bias {#sec-cognitiveBiasesRecency}
+### Primacy Effect Bias {#sec-cognitiveBiasesPrimacy}
 
-*Recency bias* is the tendency to weigh recent events more than earlier ones.
+*Primacy effect bias* is the tendency to recall the first information that one encounters better than information presented later on.
+For instance, primacy effect bias might lead a manager to draft a player based on the first information they encountered about the player.
+
+### Recency Effect Bias {#sec-cognitiveBiasesRecency}
+
+*Recency effect bias* is the tendency to weigh recent events more than earlier ones.
 For instance, a manager might observe that a Running Back on the waiver wire performed well in the last two games.
-*Recency bias* may lead the manager to pick up the player, overvaluing their recent performance.
+Recency effect bias may lead the manager to pick up the player, overvaluing their recent performance.
 For instance, the manager may not have adequately weighed the player's overall season performance and the fact that the starting Running Back is returning to the lineup from injury, and that is why the player received more carries in the past two games (i.e., in place of the injured starter).
+
+Recency effect bias seems a bit at odds with primacy effect bias; however, they can both coexist.
+People tend to remember the earliest information they encounter in addition to the most recent information, and they tend to forget the information encountered in between.
 

### 2024-08-07 cluster-analysis.qmd
20240807 - error code

@@
 numTiersRB <- 3
+```
 
-# The following code errors out:
-#tiersRB_boostrap <- mclustBootstrapLRT(
-#  player_stats_seasonal_offense_recentRB$fantasy_points,
-#  modelName = "V") # variable/unequal variance (for univariate data)
-#
-#numTiersRB <- as.numeric(summary(tiersRB_boostrap)[,"Length"][1]) # or could specify the number of teams manually
-#
-#tiersRB_boostrap
-#plot(
-#  tiersRB_boostrap,
-#  G = numTiersRB - 1)
+The model-based bootstrap clustering of Running Backs' fantasy points is unable to run due to an error:
+
+```{r}
+tiersRB_boostrap <- mclustBootstrapLRT(
+  player_stats_seasonal_offense_recentRB$fantasy_points,
+  modelName = "V") # variable/unequal variance (for univariate data)
+```
+
+Thus, we cannot use the following code, which would otherwise summarize the model results, specify the number of tiers, and plot model comparisons:
+
+```{r}
+#| eval: false
+
+numTiersRB <- as.numeric(summary(tiersRB_boostrap)[,"Length"][1]) # or could specify the number of teams manually
+
+tiersRB_boostrap
+plot(
+  tiersRB_boostrap,
+  G = numTiersRB - 1)
 ```

### 2024-08-07 cognitive-bias.qmd
20240807 - additional cognitive biases

@@
 They lead us to be less accurate in our judgments and predictions.
-Inaccuracy may not always be bad; several cognitive biases may serve to help people feel better about themselves, including [confirmation bias](#sec-cognitiveBiasesConfirmation), [overconfidence bias](#sec-cognitiveBiasesOverconfidence), optimism bias, self-serving bias, and actor-observer bias.
+Inaccuracy may not always be bad; several cognitive biases may serve to help people feel better about themselves [@Miller2013c], including [confirmation bias](#sec-cognitiveBiasesConfirmation), [overconfidence bias](#sec-cognitiveBiasesOverconfidence), [optimism bias](#sec-cognitiveBiasesOptimism), and [self-serving bias](#sec-cognitiveBiasesSelfServing).
 In fantasy football, however, cognitive biases and inaccurate judgments can lead to worse performance.
@@
 - overconfidence bias
+- optimism bias
 - confirmation bias
 - in-group bias
+- hindsight bias
+- outcome bias
+- omission bias
+- loss aversion bias
 - primacy effect bias
 - recency effect bias
-- hindsight bias
-- loss aversion bias
-- endowment bias
+- framing effect bias
+- endowment effect bias
 - bandwagon effect bias
@@
 
+### Optimism Bias {#sec-cognitiveBiasesOptimism}
+
+*Optimism bias* is the tendency for people to overestimate one's likelihood of experiencing positive events and underestimate one's likelihood of experiencing negative events.
+For instance a manager might overestimate the likelihood that their team will win the championship and may underestimate the likelihood that they may lose a player to injury.
+
 ### Confirmation Bias {#sec-cognitiveBiasesConfirmation}
@@
 
-As a budding empiricist, you should actively seek out information that challenges or disconfirms your beliefs and work to incorporate it into your beliefs.
+As a budding empiricist, you should actively seek out information that challenges or disconfirms your beliefs—particularly from trustworthy, reputable sources—and work to incorporate the information into your beliefs.
 Do your best to go into observation, data analysis, and data interpretation with an open mind.
@@
 For instance, a manager whose favorite team is the Dallas Cowboys might try to draft mostly Cowboys players, regardless of their historical performance.
+
+### Hindsight Bias {#sec-cognitiveBiasesHindsight}
+
+> "Hindsight is 20/20." – Idiom
+
+*Hindsight bias* is the tendency to perceive that past events were more predictable than they were.
+People tend to remember the succeess of their predictions and forget the failures of their predictions.
+For instance, if a third-string Quarterback has a breakout game, a fantasy manager may claim that they "knew it all along" that the player was going to breakout, despite not having picked up the player.
+That same manager may forget the many other predictions they had that did not come true.
+
+### Outcome Bias {#sec-cognitiveBiasesOutcome}
+
+*Outcome bias* is the tendency to evaluate the quality of a decision based on the eventual outcome of that decision, rather than based on the quality of the decision based on the information that was known at the time of the decision.
+For instance, consider that a manager starts a lower-ranked player over a higher-ranked player because the lower-ranked player has a more favorable matchup.
+Now, consider that the lower-ranked player scores fewer points than the higher-ranked player because the lower-ranked player got injured.
+If the manager concludes that their decision-making process for who to start was flawed and thus not pay attention to matchups in subsequent decisions, this would reflect outcome bias.
+Likewise, if the manager makes a poor-quality decision but gets lucky, and they attribute their success to a sound decision-making process, this would also reflect outcome bias.
+
+Focus on the decision-making process when evaluating the quality of decisions.
+Yes, also evaluate the outcomes of decisions, but do not give yourself too much credit for lucky decisions or too much blame for unlucky decisions (like injuries).
+That said, one thing you can count on is that players get injured.
+So, it is to your benefit to draft and compose your team to have some bench depth so you can handle injuries when they inevitably occur.
+
+### Self-Serving Bias {#sec-cognitiveBiasesSelfServing}
+
+*Self-serving bias* is the tendency to perceive oneself in an overly positive manner.
+Doing so may function to bolster people's self-esteem.
+One way that self-serving bias can commonly manifest is the tendency people to attribute successes to internal factors—such as one's character, ability, effort, or actions—but to attribute failures to external factors beyond one's control—such as others, unfairness, bad luck, etc.
+That is, self-serving bias involves claiming more reponsibility for one's success than for one's failure [@Miller2013c].
+For instance, a manager who wins the league may attribute their success to their fantasy football skill and smart decisions (rather than luck).
+However, if they get last place, they may attribute their failure not to themselves but instead to factors outside of their control, such as bad luck, player injuries, and poor performance by players due to circumstances out of their control (e.g., bad weather conditions, coaching decisions).
+
+Although the self-serving bias may help us feel better about ourselves, it can lead us to [overestimate our abilities](#sec-cognitiveBiasesOverconfidence) when things go well and to fail to learn from mistakes when things go poorly.
+Remember that nearly half of the variability in fantasy football performance is estimated to be luck, and the other half likely due to skill [@Getty2018].
+Thus, successes in fantasy football likely involve a good amount of luck; likewise, failures in fantasy football can also reflect mistakes and poor decision making that arise from lower skill.
+
+### Omission Bias {#sec-cognitiveBiasesOmission}
+
+*Omission bias* is the tendency to prefer inaction over action.
+For instance, consider a manager who is on the fence to drop an underperforming player and to pick up someone else to start.
+If the manager does not drop the player and the player scores few points, the manager is likely to be less critical of their decision than if they had picked up another player who then scored the same (low) number of points.
+
+Consider another example adapted from @Miller2013c: two players scored the same number of points in a given week.
+A Wide Receiver scored 1.6 points from 16 receiving yards on 1 target.
+A Running Back scored 1.6 points from 56 receiving yards and two dropped fumbles.
+The manager may harbor more blame for the player who performed a harmful action (losing two fumbles) than the player who received only one target (harmful inaction).
+
+Pay attention to players' usage and opportunities, not just how many points a player scored and whether they had more fumbles or interceptions than other players.
+The more times a player touches the ball, the more points they are likely to score.
+
+### Loss Aversion Bias {#sec-cognitiveBiasesLossAversion}
+
+*Loss aversion bias* is the tendency to avoid losses rather than acquiring equivalent gains.
+That is, loss aversion (similar to punishment sensitivity, as opposed to reward sensitivity) is the tendency to weigh losses (or punishment) more heavily than gains (or rewards).
+Loss aversion is different from risk aversion.
+Risk aversion is the tendency to prefer outcomes with low uncertainty (compared to outcomes with greater uncertainty).
+Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
+In fantasy football, loss aversion may lead managers to start or hold onto underperforming high drafts for too long instead of a starting a more promising player out of fear of losing potential value from their initial investment.
 
@@
 
-### Hindsight Bias {#sec-cognitiveBiasesHindsight}
+### Framing Effect Bias {#sec-cognitiveBiasesFraming}
 
-> "Hindsight is 20/20." – Idiom
+*Framing effect bias* is the tendency to make decisions based on how information is framed (e.g., based on positive or negative connotations), rather than just the information itself.
+For instance, consider a weekly column written by two different fantasy football expert commentators.
+Expert A may write that a Running Back has scored at least 15 fantasy points in 70% of games this season (i.e., positive framing).
+Expert B may write that the same Running Back has scored fewer than 15 fantasy points in 30% of games this season (i.e., negative framing).
+Both statements convey the same statistical information; however, a manager might be more inclined to start the player when reading the statement from Expert A than from Expert B because of how the information was framed (i.e., presented).
+People prefer to make decisions to avoid loss, consistent with [loss aversion bias](#sec-cognitiveBiasesLossAversion), so the negative framing might lead a person to be less likely to start the player.
 
-*Hindsight bias* is the tendency to perceive that past events were more predictable than they were.
-People tend to remember the succeess of their predictions and forget the failures of their predictions.
-For instance, if a third-string Quarterback has a breakout game, a fantasy manager may claim that they "knew it all along" that the player was going to breakout, despite not having picked up the player.
-That same manager may forget the many other predictions they had that did not come true.
+Note how information is framed, knowing that the framing itself could lead you astray.
+Thus, try to focus on the information itself rather than the framing when making decisions.
 
-### Loss Aversion Bias {#sec-cognitiveBiasesLossAversion}
+### Endowment Effect Bias {#sec-cognitiveBiasesEndowment}
 
-*Loss aversion bias* is the tendency to avoid losses rather than acquiring equivalent gains.
-Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
-In fantasy football, loss aversion may lead managers to start or hold onto underperforming high drafts for too long instead of a starting a more promising player out of fear of losing potential value from their initial investment.
+*Endowment effect bias* is the tendency to overvalue merely because one owns it.
+That is, a manager tends to value the same player more when they play for their team rather than an opponent's team.
+When trading, the endowment effect bias might lead a manager to demand more to trade away a player than to acquire the same player [@@Miller2013c].
+For instance, a manager might overvalue a player they drafted in the first round, refusing to trade them even if they could get a better-performing player in return.
 
-### Endowment Bias {#sec-cognitiveBiasesEndowment}
-
-*Endowment bias* is the tendency to overvalue merely because one owns it.
-For instance, a manager might overvalue a player they drafted in the first round, refusing to trade them even if they could get a better-performing player in return.
+A player's trade value is not fixed.
+It makes sense to adjust a player's trade value based on many factors such as your team's needs, the needs of your trade partner's team, the player's remaining strength of schedule, etc.
+However, you want to avoid adjusting a player's value based merely on the fact that he is on your team.
+Evaluate a variety of trade analyzers to more objectively evaluate players' worth.
 
@@
 If you continue to start the player to justify your large investment, instead of benching him in favor of a higher-performing player, you are committing the sunk cost fallacy.
+It is important to own up to our mistakes; when you make a mistake, the sooner you realize it, own it—for instance, by either benching or dropping the weak-performing player—and move on, the sooner you will be able to replace him with a better-performing player.
+At the same time, you probably do not want to give too much weight to a player's performance in any given week, due to random chance that may lead a player to greatly under- or overperform; and whose future peformance is likely to [regress (whether positively or negatively) to their mean](#sec-fallaciesRegression).
 

### 2024-08-07 cognitive-bias.qmd
20240807 - cross-references

@@
 
-- availability heuristic
-- representativeness heuristic
-- anchoring and adjustment heuristic
+- [availability heuristic](#sec-heuristicsAvailability)
+- [representativeness heuristic](#sec-heuristicsRepresentativeness)
+- [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment)
 
@@
 
-- overconfidence bias
-- optimism bias
-- confirmation bias
-- in-group bias
-- hindsight bias
-- outcome bias
-- omission bias
-- loss aversion bias
-- primacy effect bias
-- recency effect bias
-- framing effect bias
-- endowment effect bias
-- bandwagon effect bias
-- Dunning–Kruger effect bias
+- [overconfidence bias](#sec-cognitiveBiasesOverconfidence)
+- [optimism bias](#sec-cognitiveBiasesOptimism)
+- [confirmation bias](#sec-cognitiveBiasesConfirmation)
+- [in-group bias](#sec-cognitiveBiasesInGroup)
+- [hindsight bias](#sec-cognitiveBiasesHindsight)
+- [outcome bias](#sec-cognitiveBiasesOutcome)
+- [self-serving bias](#sec-cognitiveBiasesSelfServing)
+- [omission bias](#sec-cognitiveBiasesOmission)
+- [loss aversion bias](#sec-cognitiveBiasesLossAversion)
+- [primacy effect bias](#sec-cognitiveBiasesPrimacy)
+- [recency effect bias](#sec-cognitiveBiasesRecency)
+- [framing effect bias](#sec-cognitiveBiasesFraming)
+- [endowment effect bias](#sec-cognitiveBiasesEndowment)
+- [bandwagon effect bias](#sec-cognitiveBiasesBandwagon)
+- [Dunning–Kruger effect bias](#sec-cognitiveBiasesDunningKruger)
 
@@
 
-- base rate fallacy (aka base rate neglect)
-- regression fallacy
-- sunk cost fallacy
-- hot hand fallacy
-- gambler's fallacy
-- conditional probability fallacy
+- [base rate fallacy](#sec-fallaciesBaseRate) (aka base rate neglect)
+- [regression fallacy](#sec-fallaciesRegression)
+- [hot hand fallacy](#sec-fallaciesHotHand)
+- [sunk cost fallacy](#sec-fallaciesSunkCost)
+- [gambler's fallacy](#sec-fallaciesGambler)
+- [conditional probability fallacy](#sec-fallaciesConditionalProbability)
+- [ecological fallacy](#sec-fallaciesEcological)
 

### 2024-08-08 basic-statistics.qmd
20240807 - weighted quantile/median

@@
 If you want to give some values more weight to others, you can calculate a weighted mean and a weighted median (or other quantile), while assigning a weight to each value.
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains various functions for computing the weighted median (i.e., a weighted quantile at the 0.5 quantile, which is equivalent to the 50th percentile) based on @Akinshin2023.
+Because some projections are outliers, we use a trimmed version of the weighted Harrell-Davis quantile estimator for greater robustness.
 
@@
 #weighted.mean(data, weights, na.rm = TRUE)
-#DescTools::Quantile(data, weights, na.rm = TRUE)
+#petersenlab::wthdquantile(data, weights, probs = 0.5)
 ```

### 2024-08-08 getting-started.qmd
20240808 - why R, learning R, getting help with R

@@
 R is a free software environment; you can download it at no charge here: <https://cran.r-project.org>.
+
+## Learning `R` {#sec-learningR}
+
+Here are a various resources for learning `R`:
+
+- Intro to `R`: <https://www.statmethods.net>
+- Video training courses in `R` skills: <https://www.pluralsight.com/search?q=R>
+- Browse the `Cookbook for R` to find solutions to common tasks and problems: <http://www.cookbook-r.com>
+- Browse the `R Graph Gallery` to find examples of various graphs: <https://r-graph-gallery.com>
+- Free `Codeacademy` course on `R`: <https://www.codecademy.com/learn/learn-r>
+- Free `Coursera` courses on `R`: <https://www.coursera.org/search?query=R>
+- Watch these videos from `Coursera`: <https://blog.revolutionanalytics.com/2012/12/coursera-videos.html>
+- `Posit`/`Rstudio` Webinars: <https://posit.co/resources/videos/>
+- UCLA Stats Website: <https://stats.idre.ucla.edu/r/>
+- Introduction to `R` course on `Datacamp`: <https://www.datacamp.com/courses/free-introduction-to-r>
+- Teaching `R` in a Kinder, Gentler, More Effective Manner: <https://github.com/matloff/TidyverseSkeptic>
+- Learn `R` interactively with `swirl`: <https://swirlstats.com>
+- Use the `learnr` package: <https://rstudio.github.io/learnr/>
+- Resources for learning `tidyverse`, which is a collection of `R` packages for data management: <https://www.tidyverse.org/learn/>
+- You will sometimes find relevant articles on `R-bloggers`: <https://www.r-bloggers.com>
+
+## Getting Help with `R` {#sec-gettingHelpR}
+
+If you have `R` questions, you can ask them in a number of places:
+
+- Forums:
+    - `Posit`: <https://forum.posit.co>
+    - `StackOverflow`: <https://stackoverflow.com/questions/tagged/r>
+    - `Reddit`: <https://www.reddit.com/r/rstats/>
+- The `R` mailing list: <https://stat.ethz.ch/mailman/listinfo/r-help>
+
+The following article provides additional resources and good guidance: <https://www.r-bloggers.com/where-to-get-help-with-your-r-question/>.
+
+When posting a question on forums or mailing lists, keep a few things in mind:
+
+- Read the posting guidelines before posting!
+- Be respectful of other people and their time.
+`R` is free software.
+People are offering their free time to help.
+They are under no obligation to help you.
+If you are disrespectful or act like they owe you anything, you will rub people the wrong way and will be less likely to get help.
+- Provide a minimal, reproducible example.
+Providing a minimal, reproducible example can be crucial for getting a helpful response.
+By going to the trouble of creating a minimal reproducible example and identifying the minimum conditions necessary to reproduce the issue, you will often figure out how to resolve it.
+Here are guidelines on providing a minimal reproducible example: <https://stackoverflow.com/help/minimal-reproducible-example> (archived at <https://perma.cc/6NUB-UTYF>).
+My strong recommendation is to provide a `reprex` whenever possible: <https://reprex.tidyverse.org>.
 

### 2024-08-08 intro.qmd
20240527 - statistical programming
20240527 - formatting
20240808 - why R, learning R, getting help with R

@@
 The book provides an introduction to modern analytical techniques used to make informed predictions, test theories, and draw conclusions from a given dataset.
+The book leverages the software `R` for providing applied data analysis examples.
 
@@
 
+## Why R? {#sec-whyR}
+
+The book provides data analysis examples using the statistical analysis software, `R`.
+Why `R`?
+
+- `R` is free!
+Anyone can use it.
+- `R` is open source—it is not a black box.
+You can see what is going on "under the hood" and can examine the code for any function or computation you perform.
+You can even modify and improve these functions by changing the code, and you can create your own functions.
+- `R` is open platform—you can use it on multiple platforms, including Windows, MacOS, and Linux.
+- `R` has advanced statistics capabilities.
+It was designed for statistical analysis and has strong capabalities for data wrangling.
+- `R` has capabilities for state-of-the-art graphics.
+It has advanced capabilities for creating statistical graphics.
+- `R` is widely used—there is a large community of people who use `R` for data analysis that you can draw upon for help from others.
+- `R` analyses are based on code (rather than a graphical user interface), which allows reproducibility—with the same data, code, and setup (platform, `R` version, package versions, etc.), you should get the same answer every time.
+There are strong resources available for ensuring your analyses in `R` are reproducible by others [@Gandrud2020].
+- Anyone (including you) can contribute `R` packages to the community to improve its functionality.
+Statistical experts from all over the world have contributed open source packages to `R` for specialized tasks.
+In the chance there is not an `R` package that does what you need to do, you can write a function to perform the task and can contribute it as a package to the community for others to use and improve.
+The number of `R` packages contributed to the community is growing at a rapid rate.
+As of this writing, over 20,000 packages have been contributed to the [Comprehensive R Archive Network](https://cran.r-project.org/web/packages/) (CRAN).
+And many more are stored on publicly available version control repositories like GitHub and GitLab.
+Chances are, if there is an analysis you need to do, an `R` package exists to do it.
+
 ## Educational Value {#sec-educationalValue}
 
-Skills in data analysis are highly valuable.
+Skills in statistics, statistical programming, and data analysis are highly valuable.
 This book includes practical and conceptual tools that build a foundation for critical thinking.
@@
 - Describe the strengths and weaknesses of humans versus computers in prediction scenarios.
-- Manipulate and summarize datasets.
+- Apply basic skills in statistical programming using `R` to manipulate and summarize datasets and to conduct data analysis.
 - Critically evaluate the strengths and limitations of different statistical models and methodologies used in predicting uncertain events, enhancing their understanding of statistical inference and model selection.

### 2024-08-08 getting-started.qmd
20240808 - minimal reproducible example

@@
 Providing a minimal, reproducible example can be crucial for getting a helpful response.
-By going to the trouble of creating a minimal reproducible example and identifying the minimum conditions necessary to reproduce the issue, you will often figure out how to resolve it.
-Here are guidelines on providing a minimal reproducible example: <https://stackoverflow.com/help/minimal-reproducible-example> (archived at <https://perma.cc/6NUB-UTYF>).
+By going to the trouble of creating a minimal, reproducible example and identifying the minimum conditions necessary to reproduce the issue, you will often figure out how to resolve it.
+Here are guidelines on providing a minimal, reproducible example: <https://stackoverflow.com/help/minimal-reproducible-example> (archived at <https://perma.cc/6NUB-UTYF>).
+Here are a good example and guidelines for providing a minimal, reproducible example in `R`: <https://stackoverflow.com/a/5963610> (archived at <https://perma.cc/PC9L-DQZG>).
 My strong recommendation is to provide a `reprex` whenever possible: <https://reprex.tidyverse.org>.

### 2024-08-08 download-football-data.qmd
20240806 - game schedules
20240808 - separate variables from different position objects for Advanced PFR stats

@@
 
-The `nfl_schedules` object is in `game` form.
+The `nfl_schedules` object is in `game` form and in `season`-`week` (and -`game type`) form.
 That is, each row should be uniquely identified by `game_id`.
+Each row should also be uniquely identified by the combination of `season` and `week` (and `game type`).
 
@@
 ```{r}
+# Clean up player name for merging; name variables based on which data object they're from
+
+## Seasonal Data
 nfl_advancedStatsPFR_pass_seasonal <- nfl_advancedStatsPFR_pass_seasonal_raw %>% 
@@
   ) %>% 
-  rename(
-    player_name.pass = player
-  )
+  rename_with(
+    ~ paste0(., ".pass"),
+    -c(pfr_id, nameMerge, season, team))
 
@@
   rename(
-    player_name.rush = player,
-    age.rush = age,
     team = tm
-  )
+  ) %>% 
+  rename_with(
+    ~ paste0(., ".rush"),
+    -c(pfr_id, nameMerge, season, team))
 
@@
   rename(
-    player_name.rec = player,
-    age.rec = age,
     team = tm
-  )
+  ) %>% 
+  rename_with(
+    ~ paste0(., ".rec"),
+    -c(pfr_id, nameMerge, season, team))
 
@@
   rename(
-    player_name.def = player,
-    age.def = age,
     team = tm
-  )
-
+  ) %>% 
+  rename_with(
+    ~ paste0(., ".def"),
+    -c(pfr_id, nameMerge, season, team))
+
+## Weekly Data
 nfl_advancedStatsPFR_pass_weekly <- nfl_advancedStatsPFR_pass_weekly_raw %>% 
@@
   ) %>% 
-  rename(
-    player_name.pass = pfr_player_name,
-    season.pass = season,
-    week.pass = week,
-    team.pass = team
-  )
+  rename_with(
+    ~ paste0(., ".pass"),
+    -c(game_id, pfr_player_id))
 
@@
   ) %>% 
-  rename(
-    player_name.rush = pfr_player_name,
-    season.rush = season,
-    week.rush = week,
-    team.rush = team
-  )
+  rename_with(
+    ~ paste0(., ".rush"),
+    -c(game_id, pfr_player_id))
 
@@
   ) %>% 
-  rename(
-    player_name.rec = pfr_player_name,
-    season.rec = season,
-    week.rec = week,
-    team.rec = team
-  )
+  rename_with(
+    ~ paste0(., ".rec"),
+    -c(game_id, pfr_player_id))
 
@@
   ) %>% 
-  rename(
-    player_name.def = pfr_player_name,
-    season.def = season,
-    week.def = week,
-    team.def = team
-  )
-
+  rename_with(
+    ~ paste0(., ".def"),
+    -c(game_id, pfr_player_id))
+
+## Merge across positions
 nfl_advancedStatsPFR_seasonal_list <- list(
@@
   mutate(
-    player_name = coalesce(
-      player_name.pass,
-      player_name.rush,
-      player_name.rec,
-      player_name.def
+    pfr_player_name = coalesce(
+      player.pass,
+      player.rush,
+      player.rec,
+      player.def
     ),
@@
       age.def
+    ),
+    pos = coalesce(
+      #pos.pass,
+      pos.rush,
+      pos.rec,
+      pos.def
     )
   ) %>% 
-  select(-c(starts_with("player_name.")))
+  select(-c(
+    starts_with("player."),
+    starts_with("age."),
+    starts_with("pos.")))
 
@@
   mutate(
-    player_name = coalesce(
-      player_name.pass,
-      player_name.rush,
-      player_name.rec,
-      player_name.def
+    pfr_player_name = coalesce(
+      pfr_player_name.pass,
+      pfr_player_name.rush,
+      pfr_player_name.rec,
+      pfr_player_name.def
     ),
@@
       team.def
+    ),
+    nameMerge = coalesce(
+      nameMerge.pass,
+      nameMerge.rush,
+      nameMerge.rec,
+      nameMerge.def
+    ),
+    pfr_game_id = coalesce(
+      pfr_game_id.pass,
+      pfr_game_id.rush,
+      pfr_game_id.rec,
+      pfr_game_id.def
+    ),
+    game_type = coalesce(
+      game_type.pass,
+      game_type.rush,
+      game_type.rec,
+      game_type.def
+    ),
+    opponent = coalesce(
+      opponent.pass,
+      opponent.rush,
+      opponent.rec,
+      opponent.def
     )
   ) %>% 
-  select(-c(starts_with("player_name.")))
-
-# Todo:
-#-separate variables from each pass/rush/rec/def object
+  select(-c(
+    starts_with("pfr_player_name."),
+    starts_with("season."),
+    starts_with("week."),
+    starts_with("team."),
+    starts_with("nameMerge."),
+    starts_with("pfr_game_id."),
+    starts_with("game_type."),
+    starts_with("opponent."),
+    ))
 ```
@@
 nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal %>% 
-  select(pfr_id, season, team, player_name, everything()) %>% 
-  arrange(player_name, pfr_id, season, team)
+  select(pfr_id, season, team, pfr_player_name, everything()) %>% 
+  arrange(pfr_player_name, pfr_id, season, team)
 ```
@@
 
+```{r}
+# Todo:
+#-sum/avg variables within each pass/rush/rec/def object by team for seasonal data (so seasonal data are in player-season form, not player-season-team form)
+```
+
 The `nfl_advancedStatsPFR_weekly` object is in both `game`-`player` form and `player`-`season`-`week` form.
-That is, each row should be uniquely identified by the combination of `gsis_id`, `season`, and `week` or by the combination of `player_id`, `season`, `season_type`, and `week_num`.
+That is, each row should be uniquely identified by the combination of `pfr_player_id`, `season`, and `week` or by the combination of `pfr_player_id`, `season`, `game_type`, and `week`.
 Let's rearrange the data accordingly:
@@
 nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly %>% 
-  select(pfr_player_id, season, week, everything()) %>% 
-  arrange(player_name, pfr_player_id, season, week)
+  select(pfr_player_id, season, week, game_type, game_id, pfr_player_name, everything()) %>% 
+  arrange(pfr_player_name, pfr_player_id, season, week)
 ```
@@
 
+Now, each row of the `nfl_advancedStatsPFR_seasonal` object should be uniquely identified by the combination of `gsis_id` (or `pfr_id`), `season`, and `team`.
+Each row of the `nfl_advancedStatsPFR_weekly` object should be uniquely identified by the combination of `gsis_id` (or `pfr_id`), `season`, and `week` or by the combination of `gsis_id` (or `pfr_id`), `season`, `game_type`, and `week`.
+
 Let's check again for duplicate `game`-`player` or `player`-`season`-`week` instances:
@@
 ```{r}
+# Based on gsis_id
+nfl_advancedStatsPFR_seasonal %>% 
+  select(gsis_id, everything()) %>% 
+  filter(!is.na(gsis_id)) %>% 
+  group_by(gsis_id, season, team) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_advancedStatsPFR_weekly %>% 
+  select(gsis_id, everything()) %>% 
+  filter(!is.na(gsis_id)) %>% 
+  arrange(game_id, gsis_id) %>% 
+  group_by(game_id, gsis_id) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_advancedStatsPFR_weekly %>% 
+  select(gsis_id, everything()) %>% 
+  filter(!is.na(gsis_id)) %>% 
+  arrange(gsis_id, season, week) %>% 
+  group_by(gsis_id, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+# Based on pfr_id
 nfl_advancedStatsPFR_seasonal %>% 

### 2024-08-08 download-football-data.qmd
20240808 - convert PFR advanced stats to player-season form by aggregation of data across teams for a given player-season combination

@@
 
-A Data Dictionary for Pro Football Reference passing data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html>
+A Data Dictionary for Pro Football Reference passing data is located at the following links:
+
+- <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html>
+- <https://www.pro-football-reference.com/about/advanced_stats.htm>
+
+Advanced stats from the 2023 season are available at the following link: <https://www.pro-football-reference.com/years/2023/advanced.htm>
 
@@
 
-nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal_list %>% 
+nfl_advancedStatsPFR_seasonalByTeam <- nfl_advancedStatsPFR_seasonal_list %>% 
   purrr::reduce(
@@
 
-nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal %>% 
+nfl_advancedStatsPFR_seasonalByTeam <- nfl_advancedStatsPFR_seasonalByTeam %>% 
   mutate(
@@
       pos.def
+    ),
+    g = coalesce(
+      #g.pass,
+      g.rush,
+      g.rec,
+      g.def
+    ),
+    gs = coalesce(
+      #gs.pass,
+      gs.rush,
+      gs.rec,
+      gs.def
     )
@@
     starts_with("age."),
-    starts_with("pos.")))
+    starts_with("pos."),
+    starts_with("g."),
+    starts_with("gs.")))
 
@@
 
-The `nfl_advancedStatsPFR_seasonal` object is in `player`-`season`-`team` form.
+The `nfl_advancedStatsPFR_seasonalByTeam` object is in `player`-`season`-`team` form.
 That is, each row should be uniquely identified by the combination of `pfr_id`, `season`, and `team`.
@@
 ```{r}
-nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal %>% 
+nfl_advancedStatsPFR_seasonalByTeam <- nfl_advancedStatsPFR_seasonalByTeam %>% 
   select(pfr_id, season, team, pfr_player_name, everything()) %>% 
@@
 ```{r}
+nfl_advancedStatsPFR_seasonalByTeam %>% 
+  group_by(pfr_id, season, team) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+Aggregate variables  within each pass/rush/rec/def object by team for seasonal data (so seasonal data are in player-season form, not player-season-team form).
+Depending on the variable, aggregation was performed using a sum, weighted mean (weighted by the number of games played for each team), or a recomputed percentage.
+
+```{r}
+pfrVars <- nfl_advancedStatsPFR_seasonalByTeam %>% 
+  select(pass_attempts.pass:m_tkl_percent.def, g, gs) %>% 
+  names()
+
+weightedAverageVars <- c(
+  "pocket_time.pass",
+  "ybc_att.rush","yac_att.rush",
+  "ybc_r.rec","yac_r.rec","adot.rec","rat.rec",
+  "yds_cmp.def","yds_tgt.def","dadot.def","m_tkl_percent.def","rat.def"
+)
+
+recomputeVars <- c(
+  "drop_pct.pass", # drops.pass / pass_attempts.pass
+  "bad_throw_pct.pass", # bad_throws.pass / pass_attempts.pass
+  "on_tgt_pct.pass", # on_tgt_throws.pass / pass_attempts.pass
+  "pressure_pct.pass", # times_pressured.pass / pass_attempts.pass
+  "drop_percent.rec", # drop.rec / tgt.rec
+  "rec_br.rec", # rec.rec / brk_tkl.rec
+  "cmp_percent.def" # cmp.def / tgt.def
+)
+
+sumVars <- pfrVars[pfrVars %ni% c(
+  weightedAverageVars, recomputeVars,
+  "nameMerge", "loaded.pass", "loaded.rush", "loaded.rec", "loaded.def")]
+
+nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonalByTeam %>% 
+  group_by(pfr_id, nameMerge, season) %>% 
+  summarise(
+    across(all_of(weightedAverageVars), ~ weighted.mean(.x, w = g, na.rm = TRUE)),
+    across(all_of(sumVars), ~ sum(.x, na.rm = TRUE)),
+    .groups = "drop") %>% 
+  mutate(
+    drop_pct.pass = drops.pass / pass_attempts.pass,
+    bad_throw_pct.pass = bad_throws.pass / pass_attempts.pass,
+    on_tgt_pct.pass = on_tgt_throws.pass / pass_attempts.pass,
+    pressure_pct.pass = times_pressured.pass / pass_attempts.pass,
+    drop_percent.rec = drop.rec / tgt.rec,
+    rec_br.rec = drop.rec / tgt.rec,
+    cmp_percent.def = cmp.def / tgt.def
+  )
+
+# Merge with other player info
+nfl_advancedStatsPFR_seasonalByTeam_1stTeam <- nfl_advancedStatsPFR_seasonalByTeam %>% 
+  group_by(pfr_id, season) %>% 
+  slice(1)
+
+nfl_advancedStatsPFR_seasonalByTeam_1stTeam_mergeVars <- nfl_advancedStatsPFR_seasonalByTeam_1stTeam %>% 
+  select(pfr_id, season, team, pfr_player_name, age, pos) #, g, gs
+
+nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal %>% 
+  left_join(
+    nfl_advancedStatsPFR_seasonalByTeam_1stTeam_mergeVars,
+    by = c("pfr_id","season")
+  ) %>% 
+  select(
+    pfr_id, season, pfr_player_name, pos, age, team, g, gs, 
+    contains(".pass"), contains(".rush"), contains(".rec"), contains(".def"), 
+    everything())
+```
+
+Let's check for duplicate `player`-`season` instances:
+
+```{r}
 nfl_advancedStatsPFR_seasonal %>% 
-  group_by(pfr_id, season, team) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-# Todo:
-#-sum/avg variables within each pass/rush/rec/def object by team for seasonal data (so seasonal data are in player-season form, not player-season-team form)
+  group_by(pfr_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
 ```
@@
 
-Now, each row of the `nfl_advancedStatsPFR_seasonal` object should be uniquely identified by the combination of `gsis_id` (or `pfr_id`), `season`, and `team`.
+Now, each row of the `nfl_advancedStatsPFR_seasonal` object should be uniquely identified by the combination of `gsis_id` (or `pfr_id`), and `season`.
 Each row of the `nfl_advancedStatsPFR_weekly` object should be uniquely identified by the combination of `gsis_id` (or `pfr_id`), `season`, and `week` or by the combination of `gsis_id` (or `pfr_id`), `season`, `game_type`, and `week`.
@@
   filter(!is.na(gsis_id)) %>% 
-  group_by(gsis_id, season, team) %>% 
+  group_by(gsis_id, season) %>% 
   filter(n() > 1) %>% 
@@
 nfl_advancedStatsPFR_seasonal %>% 
-  group_by(pfr_id, season, team) %>% 
+  group_by(pfr_id, season) %>% 
   filter(n() > 1) %>% 

### 2024-08-08 machine-learning.qmd
20240802 - check for duplicates
20240808 - notes

@@
 nfl_advancedStatsPFR_seasonal %>% 
-  group_by(gsis_id, season) %>% 
-  filter(n() > 1) %>% 
+  group_by(gsis_id, season, team) %>% 
+  filter(n() > 1, !is.na(gsis_id)) %>% 
+  select(gsis_id, pfr_id, season, team, everything()) %>% 
   head()
@@
 
-(CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION)
+NOTES:
+
+- CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
+- IMPUTE THE TRAINING AND TEST DATA SEPARATELY
 

### 2024-08-09 mythbusters.qmd
20240727 - mixed models
20240808 - fixes

@@
 ### Specify Package Options {#sec-mythbustersPackageOptions}
+
 ```{r}
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
-load(file = "./data/nfl_espnQBR.RData")
+load(file = "./data/nfl_espnQBR_seasonal.RData")
 ```
@@
 load(file = "./data/player_stats_seasonal.RData")
-load(file = "./data/nfl_espnQBR.RData")
+load(file = "./data/nfl_espnQBR_seasonal.RData")
 ```
@@
   dplyr::select(player, position, team, contractYear) %>% 
-  dplyr::mutate(playerMerge = toupper(str_replace_all(player, "[[:punct:]]", ""))) %>% 
+  dplyr::mutate(nameMerge = petersenlab::cleanUpNames(player)) %>% 
   dplyr::rename(season = contractYear) %>% 
@@
 player_stats_weekly_offense <- player_stats_weekly_offense %>% 
-  dplyr::mutate(playerMerge = toupper(str_replace_all(player_display_name, "[[:punct:]]", "")))
+  dplyr::mutate(nameMerge = petersenlab::cleanUpNames(player_display_name, ))
 #nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal %>% 
-#  mutate(playerMerge = toupper(str_replace_all(player_display_name, "[[:punct:]]", "")))
+#  mutate(nameMerge = petersenlab::cleanUpNames(player_display_name, ))
 
@@
   nfl_playerContractYears,
-  by = c("playerMerge", "position_group" = "position", "season")
+  by = c("nameMerge", "position_group" = "position", "season")
 ) %>% 
@@
 #  nfl_playerContractYears,
-#  by = c("playerMerge", "position_group" = "position", "season")
+#  by = c("nameMerge", "position_group" = "position", "season")
 #) %>% 
@@
 player_statsContracts_offense_weekly <- player_statsContracts_offense_weekly %>% 
-  dplyr::arrange(playerMerge, season, season_type, week)
+  dplyr::arrange(nameMerge, season, season_type, week)
 
 #player_statsContracts_offense_seasonal <- player_statsContracts_offense_seasonal %>% 
-#  arrange(playerMerge, season)
+#  arrange(nameMerge, season)
 
@@
 # Merge with QBR data
-nfl_espnQBR$playerMerge <- paste(nfl_espnQBR$name_first, nfl_espnQBR$name_last, sep = " ") %>% 
-  stringr::str_replace_all(., "[[:punct:]]", "") %>% 
-  toupper()
+nfl_espnQBR_seasonal$nameMerge <- paste(nfl_espnQBR_seasonal$name_first, nfl_espnQBR_seasonal$name_last, sep = " ") %>% 
+  petersenlab::cleanUpNames(.)
 
@@
     .,
-    nfl_espnQBR,
-    by = c("playerMerge","team","season")
+    nfl_espnQBR_seasonal,
+    by = c("nameMerge","team","season")
   )
@@
 nfl_contractYearQBR_weekly <- nfl_contractYearQBR_weekly %>% 
-  dplyr::arrange(playerMerge, season, season_type, game_week)
+  dplyr::arrange(nameMerge, season, season_type, game_week)
 
@@
   dplyr::mutate(game_week = as.numeric(game_week)) %>% 
-  dplyr::arrange(playerMerge, season, season_type, game_week)
+  dplyr::arrange(nameMerge, season, season_type, game_week)
 
@@
   dplyr::filter(position == "QB") %>% 
-  dplyr::select(playerMerge, season, week, age, years_of_experience) %>% 
+  dplyr::select(nameMerge, season, week, age, years_of_experience) %>% 
   full_join(
     nfl_contractYearQBRsubset_weekly,
-    by = c("playerMerge","season", c("week" = "game_week"))
+    by = c("nameMerge","season", c("week" = "game_week"))
   ) %>% 
@@
 nfl_contractYearQBR_seasonal <- nfl_contractYearQBRsubset_weekly %>% 
-  dplyr::group_by(playerMerge, season) %>% 
+  dplyr::group_by(nameMerge, season) %>% 
   dplyr::summarise(
@@
 nfl_contractYearQBR_seasonal <- nfl_contractYearQBR_seasonal %>% 
-  group_by(playerMerge) %>%
+  group_by(nameMerge) %>%
   mutate(player_id = as.factor(as.character(cur_group_id())))

### 2024-08-09 sports-cognitive-psychology.qmd
20240808 - fixes

--- 2024-07-26 sports-cognitive-psychology.qmd
@@
 library("tidyverse")
+library("ggtext")
 ```
@@
 
-# Downloaded Data
-load(file = "./data/nfl_players.RData")
-load(file = "./data/nfl_teams.RData")
-load(file = "./data/nfl_rosters.RData")
-load(file = "./data/nfl_rosters_weekly.RData")
-load(file = "./data/nfl_schedules.RData")
-load(file = "./data/nfl_combine.RData")
-load(file = "./data/nfl_draftPicks.RData")
-load(file = "./data/nfl_depthCharts.RData")
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
-load(file = "./data/nfl_actualStats_weekly.RData")
-load(file = "./data/nfl_injuries.RData")
-load(file = "./data/nfl_snapCounts.RData")
-load(file = "./data/nfl_espnQBR.RData")
-load(file = "./data/nfl_nextGenStats_weekly.RData")
-load(file = "./data/nfl_advancedStatsPFR.RData")
-load(file = "./data/nfl_playerContracts.RData")
-load(file = "./data/nfl_ftnCharting.RData")
-load(file = "./data/nfl_playerIDs.RData")
-load(file = "./data/nfl_rankings.RData")
-load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
-load(file = "./data/nfl_expectedFantasyPoints_offense.RData")
+```
 
-# Calculated Data
-load(file = "./data/nfl_actualStats_career.RData")
-load(file = "./data/nfl_actualStats_seasonal.RData")
+```{r}
+load(file = "./data/nfl_pbp.RData")
+load(file = "./data/nfl_4thdown.RData")
 ```
@@
   ggplot(aes(go_boost, go, color = as.factor(season))) + 
-  geom_richtext(data = text_df,   
+  ggtext::geom_richtext(
+    data = text_df,   
     aes(x, y, label = label, angle = angle), 
-                color = "black", fill = NA, label.color = NA, size = 5) + 
+    color = "black",
+    fill = NA,
+    label.color = NA,
+    size = 5) + 
   geom_vline(xintercept = 0) +
@@
     plot.subtitle = element_markdown(size = 14, hjust = 0.5),
-    axis.title.x = element_text(size = 14, face="bold"),
-    axis.title.y = element_text(size = 14, face="bold")
+    axis.title.x = element_text(size = 14, face = "bold"),
+    axis.title.y = element_text(size = 14, face = "bold")
   ) +
@@
     arrow = arrow(length = unit(0.05, "npc")),
-    color = "black", size = 2
+    color = "black", linewidth = 2
     ) +
@@
     arrow = arrow(length = unit(0.05, "npc")),
-    color = "black", size = 2
+    color = "black", linewidth = 2
   )

### 2024-08-09 mythbusters.qmd
20240809 - updates based on updated data object

@@
 load(file = "./data/nfl_espnQBR_seasonal.RData")
+load(file = "./data/nfl_espnQBR_weekly.RData")
 ```
@@
 player_stats_weekly_offense <- player_stats_weekly_offense %>% 
-  dplyr::mutate(nameMerge = petersenlab::cleanUpNames(player_display_name, ))
+  dplyr::mutate(nameMerge = petersenlab::cleanUpNames(player_display_name))
 #nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal %>% 
@@
 
-player_statsContracts_seasonal$age[which(is.infinite((player_statsContracts_seasonal$age)))] <- NA
-player_statsContracts_seasonal$years_of_experience[which(is.infinite((player_statsContracts_seasonal$years_of_experience)))] <- NA
+player_statsContracts_seasonal[sapply(player_statsContracts_seasonal, is.infinite)] <- NA
 
@@
 # Merge with QBR data
-nfl_espnQBR_seasonal$nameMerge <- paste(nfl_espnQBR_seasonal$name_first, nfl_espnQBR_seasonal$name_last, sep = " ") %>% 
+nfl_espnQBR_weekly$nameMerge <- paste(nfl_espnQBR_weekly$name_first, nfl_espnQBR_weekly$name_last, sep = " ") %>% 
   petersenlab::cleanUpNames(.)
@@
     .,
-    nfl_espnQBR_seasonal,
+    nfl_espnQBR_weekly,
     by = c("nameMerge","team","season")
@@
 nfl_contractYearQBRsubset_weekly <- nfl_contractYearQBR_weekly %>% 
-  dplyr::filter(
-    season_type == "Regular",
-    game_week != "Season Total") %>% 
-  dplyr::mutate(game_week = as.numeric(game_week)) %>% 
-  dplyr::arrange(nameMerge, season, season_type, game_week)
+  dplyr::filter(season_type == "Regular") %>% 
+  dplyr::arrange(nameMerge, season, season_type, game_week) %>% 
+  mutate(
+    player = coalesce(player, name_display),
+    position = "QB") %>% 
+  group_by(nameMerge, player_id) %>% 
+  fill(player, .direction = "downup")
 
@@
     by = c("nameMerge","season", c("week" = "game_week"))
-  ) %>% 
+  ) %>% select(player_id, season, week, player, everything()) %>% 
   arrange(player_id, season, week)
@@
 
-nfl_contractYearQBR_seasonal$age[which(is.infinite((nfl_contractYearQBR_seasonal$age)))] <- NA
-nfl_contractYearQBR_seasonal$years_of_experience[which(is.infinite((nfl_contractYearQBR_seasonal$years_of_experience)))] <- NA
+nfl_contractYearQBR_seasonal[sapply(nfl_contractYearQBR_seasonal, is.infinite)] <- NA
 

### 2024-08-10 machine-learning.qmd
20240809 - updates

@@
   mutate(
+    season = as.integer(season),
     receptions = as.integer(receptions))
@@
 nfl_advancedStatsPFR_seasonal %>% 
-  group_by(gsis_id, season, team) %>% 
+  group_by(gsis_id, season) %>% 
   filter(n() > 1, !is.na(gsis_id)) %>% 
@@
 playerSeasonListToMerge <- list(
-  player_stats_seasonal_offense %>% filter(!is.na(gsis_id), !is.na(season), !is.na(team))
-)
-
-playerSeasonTeamListToMerge <- list(
-  nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(team))
+  player_stats_seasonal_offense %>% filter(!is.na(gsis_id), !is.na(season)),
+  nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season))
 )
@@
 playerSeasonWeekPositionListToMerge <- list(
-  nfl_depthCharts %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week)),
+  nfl_depthCharts %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week))
 )
@@
 
-playerSeasonTeamMerged <- playerSeasonTeamListToMerge %>% 
-  reduce(
-    powerjoin::power_full_join,
-    by = c("gsis_id","season"),
-    conflict = coalesce_xy)
-
 playerSeasonWeekMerged <- playerSeasonWeekListToMerge %>% 
@@
 seasonalData <- powerjoin::power_full_join(
-  idSeasonMerged,
-  idMerged,
+  playerSeasonMerged,
+  playerMerged,
   by = "gsis_id",
@@
 seasonalAndWeeklyData <- powerjoin::power_full_join(
-  idSeasonWeekMerged,
+  playerSeasonWeekMerged,
   seasonalData,
@@
 # check on duplicate id-by-season data (nfl_rosters, nfl_advancedStatsPFR_seasonal)
-seasonalData %>% 
-  select(gsis_id, season)
 
 seasonalData %>% 
-  select(gsis_id, season) %>% 
-  unique()
+  group_by(gsis_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
 ```
@@
 seasonalData_subset <- seasonalData %>% 
-  dplyr::select(-c(
-    birth_date, loaded, full_name, player_name, player_display_name, display_name, suffix, headshot_url, player, pos,
-    espn_id, sportradar_id, yahoo_id, rotowire_id, pff_id, fantasy_data_id, sleeper_id, pfr_id,
-    pfr_player_id, cfb_player_id, pfr_player_name, esb_id, gsis_it_id, smart_id,
-    college, college_name, team_abbr, current_team_id, college_conference, draft_club, status_description_abbr,
-    status_short_description, short_name, headshot, uniform_number, jersey_number, first_name, last_name,
-    football_name, team
-    ))
+  dplyr::select(-any_of(c(
+    "birth_date", "loaded", "full_name", "player_name", "player_display_name", "display_name", "suffix", "headshot_url", "player", "pos",
+    "espn_id", "sportradar_id", "yahoo_id", "rotowire_id", "pff_id", "fantasy_data_id", "sleeper_id", "pfr_id",
+    "pfr_player_id", "cfb_player_id", "pfr_player_name", "esb_id", "gsis_it_id", "smart_id",
+    "college", "college_name", "team_abbr", "current_team_id", "college_conference", "draft_club", "status_description_abbr",
+    "status_short_description", "short_name", "headshot", "uniform_number", "jersey_number", "first_name", "last_name",
+    "football_name", "team"
+  ))) %>% 
+  filter(!is.na(season))
 ```
@@
 seasonalData_subset %>% select_if(is.logical) %>% str()
+```
+
+### Lag Fantasy Points {#sec-lagFantasyPoints}
+
+```{r}
+seasonalData_subset <- seasonalData_subset %>% 
+  arrange(gsis_id, season) %>% 
+  group_by(gsis_id) %>% 
+  mutate(
+    fantasyPoints_lag = lead(fantasy_points)
+  ) %>% 
+  ungroup()
 ```
@@
 seasonalData_subset_imp <- missForest::missForest(
-  seasonalData_subset %>% select(ageCentered20, ageCentered20Quadratic, fantasy_points) %>% data.frame,
+  seasonalData_subset %>% select(ageCentered20, ageCentered20Quadratic, fantasyPoints_lag) %>% data.frame,
   verbose = TRUE,
@@
 smerf <- LongituRF::MERF(
-  X = seasonalData[,c("ageCentered20")] %>% as.matrix(),
-  Y = seasonalData$fantasy_points,
-  Z = seasonalData[,c("ageCentered20")] %>% as.matrix(),
-  id = seasonalData$gsis_id,
-  time = seasonalData$ageCentered20,
+  X = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
+  Y = seasonalData_subset_imp$ximp[,c("fantasyPoints_lag")],
+  Z = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
+  id = seasonalData_subset$gsis_id,
+  time = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
   mtry = 2,

### 2024-08-10 machine-learning.qmd
20240809 - updates

@@
 
+### Split into Test and Training Data {#sec-splitTestTraining}
+
 ### Impute the Missing Data {#sec-missingDataImputation}
@@
 seasonalData_subset_imp <- missForest::missForest(
-  seasonalData_subset %>% select(ageCentered20, ageCentered20Quadratic, fantasyPoints_lag) %>% data.frame,
+  seasonalData_subset %>% select(ageCentered20, ageCentered20Quadratic, years_of_experience, position_group, fantasyPoints_lag) %>% data.frame,
   verbose = TRUE,
@@
 smerf <- LongituRF::MERF(
-  X = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
+  X = seasonalData_subset_imp$ximp[,c("position_group")] %>% as.matrix(),
   Y = seasonalData_subset_imp$ximp[,c("fantasyPoints_lag")],
-  Z = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
+  Z = seasonalData_subset_imp$ximp[,c("years_of_experience")] %>% as.matrix(),
   id = seasonalData_subset$gsis_id,

### 2024-08-10 machine-learning.qmd
Built site for gh-pages

@@
   conflict = coalesce_xy
-)
+) %>% 
+  filter(!is.na(season))
 
@@
   conflict = coalesce_xy
-)
+) %>% 
+  filter(!is.na(week))
 
@@
   head()
+
+seasonalAndWeeklyData %>% 
+  group_by(gsis_id, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+### Lag Fantasy Points {#sec-lagFantasyPoints}
+
+```{r}
+seasonalData <- seasonalData %>% 
+  arrange(gsis_id, season) %>% 
+  group_by(gsis_id) %>% 
+  mutate(
+    fantasyPoints_lag = lead(fantasy_points)
+  ) %>% 
+  ungroup()
+
+seasonalData$fantasyPoints_lag[is.na(seasonalData$fantasyPoints_lag)] <- 0
+```
+
+### Fill in Missing Data {#sec-fillMissingData}
+
+```{r}
+seasonalData <- seasonalData %>% 
+  arrange(gsis_id, season) %>% 
+  group_by(gsis_id) %>% 
+  fill(
+    player_name, player_display_name, pos, position, position_group,
+    .direction = "downup") %>% 
+  ungroup()
 ```
@@
     "football_name", "team"
-  ))) %>% 
-  filter(!is.na(season))
+  )))
 ```
@@
 
-### Lag Fantasy Points {#sec-lagFantasyPoints}
+### Separate by Position {#sec-separateByPosition}
 
 ```{r}
-seasonalData_subset <- seasonalData_subset %>% 
-  arrange(gsis_id, season) %>% 
-  group_by(gsis_id) %>% 
-  mutate(
-    fantasyPoints_lag = lead(fantasy_points)
-  ) %>% 
-  ungroup()
+seasonalData_subsetQB <- seasonalData_subset %>% 
+  filter(position == "QB") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    completions:rushing_2pt_conversions, contains(".pass"))
+
+seasonalData_subsetRB <- seasonalData_subset %>% 
+  filter(position == "RB") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    carries:special_teams_tds, contains(".rush"), contains(".rec"))
+
+seasonalData_subsetWR <- seasonalData_subset %>% 
+  filter(position == "WR") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    carries:special_teams_tds, contains(".rush"), contains(".rec"))
+
+seasonalData_subsetTE <- seasonalData_subset %>% 
+  filter(position == "TE") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    carries:special_teams_tds, contains(".rush"), contains(".rec"))
 ```
@@
 
+- SEPARATE BY POSITION
 - CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
@@
 
-seasonalData_subset_imp <- missForest::missForest(
-  seasonalData_subset %>% select(ageCentered20, ageCentered20Quadratic, years_of_experience, position_group, fantasyPoints_lag) %>% data.frame,
+seasonalData_subsetQB_imp <- missForest::missForest(
+  seasonalData_subsetQB %>% 
+    select(ageCentered20, ageCentered20Quadratic, years_of_experience, passing_epa, pacr, dakota, rushing_epa) %>% 
+    data.frame(),
   verbose = TRUE,
@@
 dataToImpute <- seasonalData_subset %>% 
-  select(gsis_id, season, ageCentered20, ageCentered20Quadratic, fantasy_points, years_of_experience) %>% 
+  select(gsis_id, season, ageCentered20, ageCentered20Quadratic, fantasyPoints_lag, years_of_experience, passing_epa) %>% 
   filter(!is.na(gsis_id), !is.na(season)) %>% 
@@
 smerf <- LongituRF::MERF(
-  X = seasonalData_subset_imp$ximp[,c("position_group")] %>% as.matrix(),
-  Y = seasonalData_subset_imp$ximp[,c("fantasyPoints_lag")],
-  Z = seasonalData_subset_imp$ximp[,c("years_of_experience")] %>% as.matrix(),
-  id = seasonalData_subset$gsis_id,
-  time = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
+  X = seasonalData_subsetQB_imp$ximp[,c("passing_epa")] %>% as.matrix(),
+  Y = seasonalData_subsetQB$fantasyPoints_lag,
+  Z = seasonalData_subsetQB_imp$ximp[,c("years_of_experience")] %>% as.matrix(),
+  id = seasonalData_subsetQB$gsis_id,
+  time = seasonalData_subsetQB_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
   mtry = 2,

### 2024-08-10 mixed-models.qmd
20240810 - clean up

@@
 
+Information about smooth terms in the `mgcv` package is provided at the following link: <https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/smooth.terms.html>.
+
 Specify model formula:
@@
 #| eval: false
-
-#mgcv smooth terms: https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/smooth.terms.html
-
-#bayesianMixedModelFormula <- brms::bf(
-#  fantasy_points ~ positionFactor + s(ageCentered20, by = #positionFactor) + years_of_experience + s(ageCentered20, #player_idFactor, bs = "fs") + (1 | player_idFactor)
-#)
-
-#bayesianMixedModelFormula <- brms::bf(
-#  fantasy_points ~ positionFactor + s(ageCentered20, by = #positionFactor) + years_of_experience + (1 | player_idFactor)
-#)
 

### 2024-08-10 machine-learning.qmd
20240810 - random forests

--- 2024-08-10 machine-learning.qmd
@@
 library("miceadds")
+library("caret")
 library("tidyverse")
@@
   conflict = coalesce_xy
-)
+) %>% 
+  filter(!is.na(season)) %>% 
+  select(gsis_id, season, player_display_name, position, recent_team, games, everything()) #-todo: deduplicate team from prior merged objects
 
@@
   head()
+
+seasonalAndWeeklyData %>% 
+  group_by(gsis_id, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+### Fill in Missing Data {#sec-fillMissingData}
+
+```{r}
+seasonalData <- seasonalData %>% 
+  arrange(gsis_id, season) %>% 
+  group_by(gsis_id) %>% 
+  fill(
+    player_name, player_display_name, pos, position, position_group,
+    .direction = "downup") %>% 
+  ungroup()
+```
+
+### Lag Fantasy Points {#sec-lagFantasyPoints}
+
+```{r}
+seasonalData_lag <- seasonalData %>% 
+  arrange(gsis_id, season) %>% 
+  group_by(gsis_id) %>% 
+  mutate(
+    fantasyPoints_lag = lead(fantasy_points)
+  ) %>% 
+  ungroup()
+
+seasonalData_lag %>% 
+  select(gsis_id, player_display_name, season, fantasy_points, fantasyPoints_lag)
+```
+
+### Create New Data Object {#sec-newDataObject}
+
+```{r}
+newData_seasonal <- seasonalData_lag %>% 
+  filter(season == max(season, na.rm = TRUE))
+
+# Drop rows with missing fantasyPoints_lag (i.e., the player's last season), so we don't impute criterion (fantasy points)
+seasonalData_lag <- seasonalData_lag %>% 
+  filter(!is.na(fantasyPoints_lag))
 ```
@@
 ```{r}
-seasonalData %>% select_if(~class(.) == "Date")
-seasonalData %>% select_if(is.character)
-seasonalData %>% select_if(is.factor)
-seasonalData %>% select_if(is.logical)
+seasonalData_lag %>% select_if(~class(.) == "Date")
+seasonalData_lag %>% select_if(is.character)
+seasonalData_lag %>% select_if(is.factor)
+seasonalData_lag %>% select_if(is.logical)
 
-seasonalData_subset <- seasonalData %>% 
+seasonalData_lag_subset <- seasonalData_lag %>% 
   dplyr::select(-any_of(c(
@@
 ```{r}
-seasonalData_subset <- seasonalData_subset %>% mutate(across(where(is.character), as.factor))
-seasonalData_subset <- seasonalData_subset %>% mutate(across(where(is.logical), as.factor))
+seasonalData_lag_subset <- seasonalData_lag_subset %>% mutate(across(where(is.character), as.factor))
+seasonalData_lag_subset <- seasonalData_lag_subset %>% mutate(across(where(is.logical), as.factor))
 
-seasonalData_subset %>% select_if(~class(.) == "Date") %>% str()
-seasonalData_subset %>% select_if(is.character) %>% str()
-seasonalData_subset %>% select_if(is.factor) %>% str()
-seasonalData_subset %>% select_if(is.logical) %>% str()
+seasonalData_lag_subset %>% select_if(~class(.) == "Date") %>% str()
+seasonalData_lag_subset %>% select_if(is.character) %>% str()
+seasonalData_lag_subset %>% select_if(is.factor) %>% str()
+seasonalData_lag_subset %>% select_if(is.logical) %>% str()
 ```
 
-### Lag Fantasy Points {#sec-lagFantasyPoints}
+### Separate by Position {#sec-separateByPosition}
 
 ```{r}
-seasonalData_subset <- seasonalData_subset %>% 
-  arrange(gsis_id, season) %>% 
-  group_by(gsis_id) %>% 
-  mutate(
-    fantasyPoints_lag = lead(fantasy_points)
-  ) %>% 
-  ungroup()
+# Training Data
+seasonalData_lag_subsetQB <- seasonalData_lag_subset %>% 
+  filter(position == "QB") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    completions:rushing_2pt_conversions, contains(".pass"))
+
+seasonalData_lag_subsetRB <- seasonalData_lag_subset %>% 
+  filter(position == "RB") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    carries:special_teams_tds, contains(".rush"), contains(".rec"))
+
+seasonalData_lag_subsetWR <- seasonalData_lag_subset %>% 
+  filter(position == "WR") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    carries:special_teams_tds, contains(".rush"), contains(".rec"))
+
+seasonalData_lag_subsetTE <- seasonalData_lag_subset %>% 
+  filter(position == "TE") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    carries:special_teams_tds, contains(".rush"), contains(".rec"))
+
+# Out-of-Sample Prediction
+newData_seasonalQB <- newData_seasonal %>% 
+  filter(position == "QB") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    completions:rushing_2pt_conversions, contains(".pass"))
+
+newData_seasonalRB <- newData_seasonal %>% 
+  filter(position == "RB") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    carries:special_teams_tds, contains(".rush"), contains(".rec"))
+
+newData_seasonalWR <- newData_seasonal %>% 
+  filter(position == "WR") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    carries:special_teams_tds, contains(".rush"), contains(".rec"))
+
+newData_seasonalTE <- newData_seasonal %>% 
+  filter(position == "TE") %>% 
+  select(
+    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
+    height, weight, rookie_year, draft_number,
+    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    carries:special_teams_tds, contains(".rush"), contains(".rec"))
 ```
@@
 
-#### `missForest`
+Here is a vignette demonstrating how to impute missing data using `missForest()`: <https://rpubs.com/lmorgan95/MissForest> (archived at: <https://perma.cc/6GB4-2E22>).
 
 ```{r}
+#| eval: false
+
+num_cores <- detectCores() - 1
+```
+
+```{r}
+#| include: false
+
 num_cores <- detectCores()
+```
+
+```{r}
 num_cores
 
-registerDoParallel(cores = num_cores)
+cl <- makeCluster(num_cores)
+registerDoParallel(cl)
 
-seasonalData_subset_imp <- missForest::missForest(
-  seasonalData_subset %>% select(ageCentered20, ageCentered20Quadratic, years_of_experience, position_group, fantasyPoints_lag) %>% data.frame,
+set.seed(52242)
+
+# QBs
+seasonalData_lag_subsetQB_imp <- missForest::missForest(
+  seasonalData_lag_subsetQB %>% 
+    select(-gsis_id) %>%
+    data.frame(),
   verbose = TRUE,
@@
   parallelize = "forests")
-```
 
-#### `MICE`
+seasonalData_lag_subsetQB_imp$OOBerror
 
-```{r}
-# Specify data to impute
-dataToImpute <- seasonalData_subset %>% 
-  select(gsis_id, season, ageCentered20, ageCentered20Quadratic, fantasy_points, years_of_experience) %>% 
-  filter(!is.na(gsis_id), !is.na(season)) %>% 
-  mutate(ID = as.integer(as.factor(as.character(gsis_id)))) %>% 
-  arrange(gsis_id, season)
+newData_seasonalQB_imp <- missForest::missForest(
+  newData_seasonalQB %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
 
-# Specify variables to impute
-varsToImpute <- c("fantasy_points")
-Y <- varsToImpute
+newData_seasonalQB_imp$OOBerror
 
-# Specify predictors of missing data
-clusterVariable <- "ID"
-fixedEffectPredictors <- c("ageCentered20Quadratic")
-randomEffectPrectors <- c("ageCentered20")
+# RBs
+seasonalData_lag_subsetRB_imp <- missForest::missForest(
+  seasonalData_lag_subsetRB %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
 
-# Specify number of imputations
-numImputations <- 1
+seasonalData_lag_subsetRB_imp$OOBerror
 
-# Specify the imputation method
-meth <- make.method(dataToImpute)
-meth[1:length(meth)] <- ""
-meth[Y] <- "2l.pmm" # specify the imputation method here; this can differ by outcome variable
+newData_seasonalRB_imp <- missForest::missForest(
+  newData_seasonalRB %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
 
-# Specify the predictor matrix
+newData_seasonalRB_imp$OOBerror
 
-pred <- make.predictorMatrix(dataToImpute)
-pred[1:nrow(pred), 1:ncol(pred)] <- 0
-pred[Y, clusterVariable] <- (-2) # cluster variable
-pred[Y, randomEffectPrectors] <- 2 # random effect predictor
-#pred[Y, fixedEffectPredictors] <- 1 # fixed effect predictor
-pred[Y, Y] <- 1 # fixed effect predictor of each predictor in predicting the other predictors
+# WRs
+seasonalData_lag_subsetWR_imp <- missForest::missForest(
+  seasonalData_lag_subsetWR %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
 
-diag(pred) <- 0
-pred
+seasonalData_lag_subsetWR_imp$OOBerror
 
-# Imputation
-mi_mice <- mice(
-  as.data.frame(dataToImpute),
-  method = meth,
-  predictorMatrix = pred,
-  m = numImputations,
-  maxit = 100, # generally use 100 maximum iterations; this example uses 5 for speed
-  seed = 52242)
+newData_seasonalWR_imp <- missForest::missForest(
+  newData_seasonalWR %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
 
-mi_mice$loggedEvents
+newData_seasonalWR_imp$OOBerror
 
-seasonalData_subset_imp <- complete(
-  mi_mice,
-  action = "stacked",
-  include = FALSE)
+# TEs
+seasonalData_lag_subsetTE_imp <- missForest::missForest(
+  seasonalData_lag_subsetTE %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
+
+seasonalData_lag_subsetTE_imp$OOBerror
+
+newData_seasonalTE_imp <- missForest::missForest(
+  newData_seasonalTE %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
+
+newData_seasonalTE_imp$OOBerror
+
+stopCluster(cl)
 ```
@@
 
+```{r}
+cl <- makeCluster(num_cores)
+registerDoParallel(cl)
+
+set.seed(52242)
+
+randomForest_qb <- caret::train(
+  fantasyPoints_lag ~ ., # use all predictors
+  data = seasonalData_lag_subsetQB_imp$ximp,
+  method = "rf",
+  trControl = trainControl(
+    method = "cv",
+    number = 10)) # 10-fold cross-validation
+
+randomForest_rb <- caret::train(
+  fantasyPoints_lag ~ ., # use all predictors
+  data = seasonalData_lag_subsetRB_imp$ximp,
+  method = "rf",
+  trControl = trainControl(
+    method = "cv",
+    number = 10)) # 10-fold cross-validation
+
+randomForest_wr <- caret::train(
+  fantasyPoints_lag ~ ., # use all predictors
+  data = seasonalData_lag_subsetWR_imp$ximp,
+  method = "rf",
+  trControl = trainControl(
+    method = "cv",
+    number = 10)) # 10-fold cross-validation
+
+randomForest_te <- caret::train(
+  fantasyPoints_lag ~ ., # use all predictors
+  data = seasonalData_lag_subsetTE_imp$ximp,
+  method = "rf",
+  trControl = trainControl(
+    method = "cv",
+    number = 10)) # 10-fold cross-validation
+
+stopCluster(cl)
+
+print(randomForest_qb)
+print(randomForest_rb)
+print(randomForest_wr)
+print(randomForest_te)
+
+newData_seasonalQB_imp$ximp$fantasyPoints_lag <- predict(
+  randomForest_qb,
+  newdata = newData_seasonalQB_imp$ximp
+)
+
+newData_seasonalRB_imp$ximp$fantasyPoints_lag <- predict(
+  randomForest_rb,
+  newdata = newData_seasonalRB_imp$ximp
+)
+
+newData_seasonalWR_imp$ximp$fantasyPoints_lag <- predict(
+  randomForest_wr,
+  newdata = newData_seasonalWR_imp$ximp
+)
+
+newData_seasonalTE_imp$ximp$fantasyPoints_lag <- predict(
+  randomForest_te,
+  newdata = newData_seasonalTE_imp$ximp
+)
+
+newData_seasonalQB$fantasyPoints_lag <- newData_seasonalQB_imp$ximp$fantasyPoints_lag
+newData_seasonalRB$fantasyPoints_lag <- newData_seasonalRB_imp$ximp$fantasyPoints_lag
+newData_seasonalWR$fantasyPoints_lag <- newData_seasonalWR_imp$ximp$fantasyPoints_lag
+newData_seasonalTE$fantasyPoints_lag <- newData_seasonalTE_imp$ximp$fantasyPoints_lag
+
+newData_seasonalQB <- left_join(
+  newData_seasonalQB,
+  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
+  by = "gsis_id"
+)
+
+newData_seasonalRB <- left_join(
+  newData_seasonalRB,
+  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
+  by = "gsis_id"
+)
+
+newData_seasonalWR <- left_join(
+  newData_seasonalWR,
+  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
+  by = "gsis_id"
+)
+
+newData_seasonalTE <- left_join(
+  newData_seasonalTE,
+  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
+  by = "gsis_id"
+)
+
+newData_seasonalQB %>%
+  arrange(-fantasyPoints_lag) %>% 
+  select(gsis_id, player_display_name, fantasyPoints_lag)
+
+newData_seasonalRB %>%
+  arrange(-fantasyPoints_lag) %>% 
+  select(gsis_id, player_display_name, fantasyPoints_lag)
+
+newData_seasonalWR %>%
+  arrange(-fantasyPoints_lag) %>% 
+  select(gsis_id, player_display_name, fantasyPoints_lag)
+
+newData_seasonalTE %>%
+  arrange(-fantasyPoints_lag) %>% 
+  select(gsis_id, player_display_name, fantasyPoints_lag)
+```
+
 #### Longitudinal Data {#sec-randomForestLongitudinalData}
@@
 ```{r}
+#| eval: false
+
 smerf <- LongituRF::MERF(
-  X = seasonalData_subset_imp$ximp[,c("position_group")] %>% as.matrix(),
-  Y = seasonalData_subset_imp$ximp[,c("fantasyPoints_lag")],
-  Z = seasonalData_subset_imp$ximp[,c("years_of_experience")] %>% as.matrix(),
-  id = seasonalData_subset$gsis_id,
-  time = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
-  mtry = 2,
+  X = seasonalData_subsetQB_imp$ximp[,c("passing_epa")] %>% as.matrix(),
+  Y = seasonalData_subsetQB$fantasyPoints_lag,
+  Z = seasonalData_subsetQB_imp$ximp[,c("pacr")] %>% as.matrix(),
+  id = seasonalData_subsetQB$gsis_id,
+  time = seasonalData_subsetQB_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
   ntree = 500,
@@
 
-smerf$forest # is the fitted random forest (obtained at the last iteration)
-smerf$random_effects # are the predicted random effects for each individual
-smerf$omega # are the predicted stochastic processes
+smerf$forest # the fitted random forest (obtained at the last iteration)
+smerf$random_effects # the predicted random effects for each player
+smerf$omega # the predicted stochastic processes
 plot(smerf$Vraisemblance) # evolution of the log-likelihood

### 2024-08-10 mixed-models.qmd
20240810 - select best-fitting response distribution

@@
 library("cmdstanr") # todo: install.packages("cmdstanr", repos = c('https://stan-dev.r-universe.dev', getOption("repos"))); check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
+library("fitdistrplus")
 library("parallel")
@@
 
-Create an object with complete cases for generating the plots of individuals' -implied fantasy points by age and position:
+Create an object with complete cases for generating the plots of individuals' model-implied fantasy points by age and position:
 
@@
 ```{r}
+fantasyPointsVector <- player_stats_seasonal_offense_subset$fantasy_points %>% 
+    na.omit() %>% 
+    as.vector()
+
+player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros <- player_stats_seasonal_offense_subset$fantasyPointsPos <- player_stats_seasonal_offense_subset$fantasy_points
+player_stats_seasonal_offense_subset$fantasyPointsPos[player_stats_seasonal_offense_subset$fantasy_points < 0] <- 0
+player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros[player_stats_seasonal_offense_subset$fantasy_points <= 0] <- 0.01
+
+fantasyPointsVectorPosNoZeros <- fantasyPointsVectorPos <- fantasyPointsVector
+fantasyPointsVectorPos[fantasyPointsVector < 0] <- 0
+fantasyPointsVectorPosNoZeros[fantasyPointsVector <= 0] <- 0.01
+```
+
+```{r}
 #| label: fig-histogramFantasyPoints
@@
   theme_classic()
+```
+
+```{r}
+#| label: fig-cullenFreyGraphy
+#| fig-cap: "Cullen and Frey Graph."
+
+fitdistrplus::descdist(fantasyPointsVector)
+```
+
+```{r}
+# all values
+fit.norm <- fitdistrplus::fitdist(fantasyPointsVector, "norm")
+
+# positive-only
+fit.exp <- fitdist(fantasyPointsVectorPos, "exp")
+
+# positive and no zeros
+fit.gamma <- fitdistrplus::fitdist(fantasyPointsVectorPosNoZeros, "gamma")
+fit.lognormal <- fitdistrplus::fitdist(fantasyPointsVectorPosNoZeros, "lnorm")
+fit.weibull <- fitdistrplus::fitdist(fantasyPointsVectorPosNoZeros, "weibull")
+
+# Model fit
+AIC(fit.norm)
+AIC(fit.exp)
+AIC(fit.gamma) # fits best
+AIC(fit.lognormal)
+AIC(fit.weibull)
+```
+
+```{r}
+plot(fit.norm)
+```
+
+```{r}
+plot(fit.lognormal)
+```
+
+```{r}
+plot(fit.gamma)
+```
+
+```{r}
+plot(fit.exp)
+```
+
+```{r}
+plot(fit.weibull)
+```
+
+```{r}
+model_normal <- glm(
+  formula = fantasy_points ~ 1,
+  family = gaussian(),
+  data = player_stats_seasonal_offense_subset,
+  maxit = 100000)
+
+model_gamma <- glm(
+  formula = fantasyPointsPosNoZeros ~ 1,
+  family = Gamma(),
+  data = player_stats_seasonal_offense_subset,
+  maxit = 100000)
+
+sum(resid(model_normal)^2)
+sum(resid(model_gamma)^2)
+
+AIC(model_normal)
+AIC(model_gamma)
 ```

### 2024-08-11 download-football-data.qmd
20240810 - updates

@@
 nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly_raw %>% 
+  rename(team = recent_team) %>% 
   select(player_id, season, week, everything()) %>% 
@@
 nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal_raw %>% 
+  rename(team = recent_team) %>% 
   select(player_id, season, everything()) %>% 
@@
   tidyr::pivot_longer(
-    cols = c(recent_team, opponent_team),
+    cols = c(team, opponent_team),
     names_to = "role",
@@
   player_game_birthdate_gameday_offense,
-  nfl_players %>% select(-birth_date, -season),
+  nfl_players %>% select(-birth_date, -season, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id"))
@@
   player_game_birthdate_gameday_defense,
-  nfl_players %>% select(-birth_date, -season),
+  nfl_players %>% select(-birth_date, -season, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id"))
@@
   player_game_birthdate_gameday_kicking,
-  nfl_players %>% select(-birth_date, -season),
+  nfl_players %>% select(-birth_date, -season, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id"))
@@
   nfl_actualStats_offense_seasonal,
-  nfl_players %>% select(-position, -position_group, -season),
+  nfl_players %>% select(-position, -position_group, -season, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id")
@@
   nfl_actualStats_defense_seasonal,
-  nfl_players %>% select(-position, -position_group, -season),
+  nfl_players %>% select(-position, -position_group, -season, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id")
@@
   nfl_actualStats_kicking_seasonal,
-  nfl_players %>% select(-position, -position_group, -season),
+  nfl_players %>% select(-position, -position_group, -season, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id")

### 2024-08-11 machine-learning.qmd
20240810 - updates

@@
 ```{r}
-library("LongituRF")
 library("doParallel")
@@
 library("powerjoin")
-library("mice")
-library("miceadds")
 library("caret")
@@
 ## Rename other variables to ensure common names
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
-  rename(team = team_abbr)
 
@@
   playerSeasonMerged,
-  playerMerged,
+  playerMerged %>% select(-team, -team_abbr, -team_seq, -current_team_id),
   by = "gsis_id",
@@
   filter(!is.na(season)) %>% 
-  select(gsis_id, season, player_display_name, position, recent_team, games, everything()) #-todo: deduplicate team from prior merged objects
+  select(gsis_id, season, player_display_name, position, team, games, everything())
 
@@
 
-# check on duplicate id-by-season data (nfl_rosters, nfl_advancedStatsPFR_seasonal)
-
+# Duplicate cases
 seasonalData %>% 
@@
   newData_seasonalQB,
-  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
+  newData_seasonal %>% select(gsis_id, player_display_name, team),
   by = "gsis_id"
@@
   newData_seasonalRB,
-  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
+  newData_seasonal %>% select(gsis_id, player_display_name, team),
   by = "gsis_id"
@@
   newData_seasonalWR,
-  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
+  newData_seasonal %>% select(gsis_id, player_display_name, team),
   by = "gsis_id"
@@
   newData_seasonalTE,
-  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
+  newData_seasonal %>% select(gsis_id, player_display_name, team),
   by = "gsis_id"
@@
 #| eval: false
+
+library("LongituRF")
 

### 2024-08-11 mixed-models.qmd
20240810 - updates

@@
 ```{r}
+player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros <- player_stats_seasonal_offense_subset$fantasyPointsPos <- player_stats_seasonal_offense_subset$fantasy_points
+player_stats_seasonal_offense_subset$fantasyPointsPos[player_stats_seasonal_offense_subset$fantasy_points < 0] <- 0
+player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros[player_stats_seasonal_offense_subset$fantasy_points <= 0] <- 0.01
+
 fantasyPointsVector <- player_stats_seasonal_offense_subset$fantasy_points %>% 
@@
     as.vector()
-
-player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros <- player_stats_seasonal_offense_subset$fantasyPointsPos <- player_stats_seasonal_offense_subset$fantasy_points
-player_stats_seasonal_offense_subset$fantasyPointsPos[player_stats_seasonal_offense_subset$fantasy_points < 0] <- 0
-player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros[player_stats_seasonal_offense_subset$fantasy_points <= 0] <- 0.01
 
@@
 ```{r}
+#| label: fig-normalDistribution
+#| fig-cap: "Fit of normal distribution to fantasy points."
+
 plot(fit.norm)
@@
 ```{r}
+#| label: fig-logNormalDistribution
+#| fig-cap: "Fit of log normal distribution to fantasy points."
+
 plot(fit.lognormal)
@@
 ```{r}
+#| label: fig-gammaDistribution
+#| fig-cap: "Fit of gamma distribution to fantasy points."
+
 plot(fit.gamma)
@@
 ```{r}
+#| label: fig-exponentialDistribution
+#| fig-cap: "Fit of exponential distribution to fantasy points."
+
 plot(fit.exp)
@@
 ```{r}
+#| label: fig-weibullDistribution
+#| fig-cap: "Fit of Weibull distribution to fantasy points."
+
 plot(fit.weibull)

### 2024-08-11 machine-learning.qmd
20240811 - drop variables from second step

@@
 playerListToMerge <- list(
-  nfl_players %>% filter(!is.na(gsis_id)) %>% select(-season, -years_of_experience),
-  nfl_draftPicks %>% filter(!is.na(gsis_id)) %>% select(-season, -age)
+  nfl_players %>% filter(!is.na(gsis_id)) %>% select(-season),
+  nfl_draftPicks %>% filter(!is.na(gsis_id)) %>% select(-season)
 )
@@
   playerSeasonMerged,
-  playerMerged %>% select(-team, -team_abbr, -team_seq, -current_team_id),
+  playerMerged %>% select(-age, -years_of_experience, -team, -team_abbr, -team_seq, -current_team_id),
   by = "gsis_id",

### 2024-08-11 time-series-analysis.qmd
20240605 - getting started/load packages
20240608 - getting started
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240703 - conclusion section
20240811 - time series analysis

--- 2024-06-05 base-rates.qmd
@@
-# Base Rates {#sec-baseRates}
+# Time Series Analysis {#sec-timeSeries}
 
-## Getting Started
+## Getting Started {#sec-timeSeriesGettingStarted}
 
-### Load Packages
+### Load Packages {#sec-timeSeriesLoadPackages}
 
 ```{r}
+library("xts")
+library("zoo")
+library("forecast")
+library("tidyverse")
+```
 
+### Load Data {#sec-timeSeriesLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
 ```
+
+```{r}
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+```
+
+## Overview of Time Series Analysis {#sec-timeSeriesOverview}
+
+Time series analysis is useful when trying to generate forecasts from longitudinal data.
+There many different types of time series analyses.
+For simplicity, in this chapter, we use autoregressive integrated moving average (ARIMA) models to demonstrate one approach to time series analysis.
+ARIMA models perform best when there is a clear pattern where later values are influenced by earlier values (i.e., autoregressive effects and moving average effects).
+This may not strongly be the case in fantasy football, so ARIMA models may not be particularly useful in forecasting fantasy football performance.
+Nevertheless, ARIMA models are widely used in forecasting financial markets and economic indicators.
+Thus, it is a useful technique to learn.
+
+Adapted from: <https://rc2e.com/timeseriesanalysis> (archived at <https://perma.cc/U5P6-2VWC>).
+
+## Create the Time Series Objects {#sec-createTimeSeriesObjects}
+
+```{r}
+weeklyFantasyPoints_tomBrady <- player_stats_weekly_offense %>% 
+  filter(
+    player_id == "00-0019596" | player_display_name == "Tom Brady")
+
+weeklyFantasyPoints_peytonManning <- player_stats_weekly_offense %>% 
+  filter(
+    player_id == "00-0010346" | player_display_name == "Peyton Manning")
+
+ts_tomBrady <- xts::xts(
+  x = weeklyFantasyPoints_tomBrady["fantasy_points"],
+  order.by = weeklyFantasyPoints_tomBrady$gameday)
+
+ts_peytonManning <- xts::xts(
+  x = weeklyFantasyPoints_peytonManning["fantasy_points"],
+  order.by = weeklyFantasyPoints_peytonManning$gameday)
+
+ts_tomBrady
+ts_peytonManning
+
+ts_combined <- merge(
+  ts_tomBrady,
+  ts_peytonManning
+)
+
+names(ts_combined) <- c("Tom Brady","Peyton Manning")
+```
+
+## Plot the Time Series {#sec-plotTimeSeries}
+
+```{r}
+plot(
+  ts_tomBrady,
+  main = "Tom Brady's Fantasy Points by Game")
+```
+
+```{r}
+plot(
+  ts_combined,
+  legend,
+  legend.loc = "topright",
+  main = "Fantasy Points by Game")
+```
+
+## Rolling Mean/Median {#sec-rollingMeanMedian}
+
+```{r}
+zoo::rollmean(
+  x = ts_tomBrady,
+  k = 5)
+
+zoo::rollmedian(
+  x = ts_tomBrady,
+  k = 5)
+```
+
+## Autocorrelation {#sec-autoCorrelation}
+
+```{r}
+acf(ts_tomBrady)
+```
+
+```{r}
+Box.test(ts_tomBrady)
+```
+
+## Fit an Autoregressive Integrated Moving Average Model {#sec-arimaModel}
+
+```{r}
+forecast::auto.arima(ts_tomBrady)
+forecast::auto.arima(ts_peytonManning)
+```
+
+```{r}
+arima_tomBrady <- arima(
+  ts_tomBrady,
+  order = c(5, 1, 4))
+
+summary(arima_tomBrady)
+confint(arima_tomBrady)
+forecast::checkresiduals(arima_tomBrady)
+
+arima_tomBrady_removeNonSigTerms <- arima(
+  ts_tomBrady,
+  order = c(5, 1, 4),
+  fixed = c(NA, NA, 0, NA, NA, NA, NA, NA, NA))
+
+summary(arima_tomBrady_removeNonSigTerms)
+confint(arima_tomBrady_removeNonSigTerms)
+forecast::checkresiduals(arima_tomBrady_removeNonSigTerms)
+
+arima_peytonManning <- arima(
+  ts_peytonManning,
+  order = c(1, 0, 1))
+
+summary(arima_peytonManning)
+confint(arima_peytonManning)
+forecast::checkresiduals(arima_peytonManning)
+```
+
+## Generate the Model Forecasts {#sec-arimaForecasts}
+
+```{r}
+forecast_tomBrady <- forecast::forecast(
+  arima_tomBrady,
+  level = c(80, 95)) # 80% and 95% confidence intervals
+
+forecast_peytonManning <- forecast::forecast(
+  arima_peytonManning,
+  level = c(80, 95)) # 80% and 95% confidence intervals
+
+forecast_tomBrady
+forecast_peytonManning
+```
+
+## Plot the Model Forecasts {#sec-plotARIMAforecast}
+
+```{r}
+forecast::autoplot(forecast_tomBrady) + 
+  labs(
+    x = "Game Number",
+    y = "Fantasy Points",
+    title = "Tom Brady's Historical and Projected Fantasy Points by Game"
+  ) +
+  theme_classic()
+```
+
+## Conclusion {#sec-timeSeriesConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-timeSeriesSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-08-11 download-football-data.qmd
20240811 - fantasy football projections

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp_raw.RData", fsep = ""))
+load(file = "./data/players_projections_seasonal_raw.RData")
 
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
+load(file = "./data/players_projections_seasonal_average.RData")
 
@@
 
+#### Processing {#sec-processingPlayers}
+
 Let's do some data cleanup:
@@
 ```
+
+#### Processing {#sec-processingRosters}
 
@@
 
+#### Processing {#sec-processingDraftPicks}
+
 Let's do some data cleanup:
@@
 ```
+
+#### Processing {#sec-processingDepthCharts}
 
@@
 ```
+
+#### Processing {#sec-processingAdvancedStatsPFR}
 
@@
 )
-
-# Todo:
-#-separate variables from each pass/rush object
 ```
@@
 ```
+
+#### Processing {#sec-processingExpectedFantasyPoints}
 
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
+```
+
+### Fantasy Football Projections {#sec-fantasyFootballProjections}
+
+#### Download Players' Projections {#sec-scrapeProjections}
+
+::: {#nte-downloadProjections .callout-note title="Downloading players' seasonal projections"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+
+players_projections_seasonal_raw <- ffanalytics::scrape_data(
+  season = NULL, # NULL grabs the current season
+  week = 0) # NULL grabs seasonal projections; NULL grabs the current week
+```
+
+```{r}
+#| eval: false
+
+save(
+  players_projections_seasonal_raw,
+  file = "./data/players_projections_seasonal_raw.RData"
+)
+```
+
+The data file is saved in the project repository and can be loaded using the following command:
+
+```{r}
+#| eval: false
+
+load(file = "./data/players_projections_seasonal_raw.RData")
+```
+
+The `players_projections_seasonal_raw` object is in `player`-`position`-`projection source` form.
+That is, each row should be uniquely identified by the combination of `id`, `pos`  and `data_src`.
+Each row should also be uniquely identified by the combination of `player`, `pos`, and `data_src`.
+
+Let's check for duplicate `player`-`position`-`projection source` instances:
+
+```{r}
+players_projections_seasonal_raw %>% 
+  bind_rows() %>% 
+  select(id, pos, data_src, everything()) %>%
+  arrange(player, id, pos, data_src) %>% 
+  group_by(id, pos, data_src) %>% 
+  filter(n() > 1, !is.na(id)) %>% 
+  head()
+
+players_projections_seasonal_raw %>% 
+  bind_rows() %>% 
+  select(player, pos, data_src, everything()) %>%
+  arrange(player, id, pos, data_src) %>% 
+  group_by(id, pos, data_src) %>% 
+  filter(n() > 1, !is.na(id)) %>% 
+  head()
+```
+
+#### Specify League Scoring Settings {#sec-leagueScoringSettings}
+
+First, create a scoring object using the default scoring object:
+
+```{r}
+scoring_obj_default <- ffanalytics::scoring
+```
+
+View the default scoring settings:
+
+```{r}
+scoring_obj_default
+```
+
+Now, modify the scoring settings to match our league settings:
+
+```{r}
+scoring_obj <- scoring_obj_default
+
+scoring_obj$pass$pass_int <- -2
+scoring_obj$misc$fumbles_lost <- -2
+scoring_obj$idp$idp_fum_force <- -2
+```
+
+View our scoring settings:
+
+```{r}
+scoring_obj
+```
+
+#### Calculate Projected Points {#sec-calculateProjectedPoints}
+
+```{r}
+players_projectedStats_seasonal <- ffanalytics::projections_table(
+  players_projections_seasonal_raw,
+  scoring_rules = scoring_obj,
+  return_raw_stats = TRUE)
+
+players_projectedPoints_seasonal <- ffanalytics::projections_table(
+  players_projections_seasonal_raw,
+  scoring_rules = scoring_obj,
+  return_raw_stats = FALSE)
+```
+
+The `players_projectedStats_seasonal` and `players_projectedPoints_seasonal` objects are in `player`-`average type`-`position` form.
+That is, each row should be uniquely identified by the combination of `id`, `avg_type`  and `pos` (or `position`).
+
+Let's check for duplicate `player`-`position`-`projection source` instances:
+
+```{r}
+players_projectedPoints_seasonal %>% 
+  group_by(id, avg_type, pos) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+players_projectedStats_seasonal %>% 
+  group_by(id, avg_type, position) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+Let's merge the two objects.
+
+```{r}
+players_projections_seasonal_average <- full_join(
+  players_projectedPoints_seasonal,
+  players_projectedStats_seasonal,
+  by = c("id","avg_type","pos" = "position")
+)
+```
+
+Let's again check for duplicate `player`-`position`-`projection source` instances:
+
+```{r}
+players_projections_seasonal_average %>% 
+  group_by(id, avg_type, pos) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+#### Add Additional Player Information {#sec-addAdditionalInfo}
+
+```{r}
+#| eval: false
+
+players_projections_seasonal_average <- players_projections_seasonal_average %>% 
+  add_ecr() %>% 
+  add_adp() %>% 
+  add_aav() %>%
+  add_uncertainty() %>% 
+  add_player_info()
+```
+
+```{r}
+#| eval: false
+
+save(
+  players_projections_seasonal_average,
+  file = "./data/players_projections_seasonal_average.RData"
+)
+```
+
+The data file is saved in the project repository and can be loaded using the following command:
+
+```{r}
+load(file = "./data/players_projections_seasonal_average.RData")
 ```

### 2024-08-11 fantasy-football.qmd
20240811 - fantasy football projections

@@
 |-------------------------------|--------|
-| Defensive or special teams TD | 3      |
+| Defensive or special teams TD | 6      |
 | Interception                  | 2      |

### 2024-08-12 download-football-data.qmd
20240811 - fantasy football projections

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
+load(file = "./data/players_projectedPoints_seasonal.RData")
 load(file = "./data/players_projections_seasonal_average.RData")
@@
 ```{r}
-#| eval: false
-
 load(file = "./data/players_projections_seasonal_raw.RData")
@@
 
-```{r}
-players_projectedStats_seasonal <- ffanalytics::projections_table(
+Calculate projected points by source:
+
+```{r}
+players_projectedPoints_seasonal <- ffanalytics:::impute_and_score_sources(
+  data_result = players_projections_seasonal_raw,
+  scoring_rules = scoring_obj)
+```
+
+Calculate projected statistics and points, averaged across sources:
+
+```{r}
+players_projectedStatsAverage_seasonal <- ffanalytics::projections_table(
   players_projections_seasonal_raw,
@@
 
-players_projectedPoints_seasonal <- ffanalytics::projections_table(
+players_projectedPointsAverage_seasonal <- ffanalytics::projections_table(
   players_projections_seasonal_raw,
@@
 
-The `players_projectedStats_seasonal` and `players_projectedPoints_seasonal` objects are in `player`-`average type`-`position` form.
+The `players_projectedPoints_seasonal`, `players_projectedStatsAverage_seasonal`, and `players_projectedPointsAverage_seasonal` objects are in `player`-`average type`-`position` form.
 That is, each row should be uniquely identified by the combination of `id`, `avg_type`  and `pos` (or `position`).
@@
 players_projectedPoints_seasonal %>% 
+  bind_rows() %>% 
+  select(id, pos, data_src, everything()) %>%
+  arrange(player, id, pos, data_src) %>% 
+  group_by(id, pos, data_src) %>% 
+  filter(n() > 1, !is.na(id)) %>% 
+  head()
+
+players_projectedStatsAverage_seasonal %>% 
+  group_by(id, avg_type, position) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+players_projectedPointsAverage_seasonal %>% 
   group_by(id, avg_type, pos) %>% 
@@
   head()
-
-players_projectedStats_seasonal %>% 
-  group_by(id, avg_type, position) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's merge the two objects.
+```
+
+Let's merge the two averaged projected statistics and points objects:
 
@@
 players_projections_seasonal_average <- full_join(
-  players_projectedPoints_seasonal,
-  players_projectedStats_seasonal,
+  players_projectedPointsAverage_seasonal,
+  players_projectedStatsAverage_seasonal,
   by = c("id","avg_type","pos" = "position")
@@
 save(
+  players_projectedPoints_seasonal,
+  file = "./data/players_projectedPoints_seasonal.RData"
+)
+
+save(
   players_projections_seasonal_average,
@@
 ```{r}
+load(file = "./data/players_projectedPoints_seasonal.RData")
 load(file = "./data/players_projections_seasonal_average.RData")

### 2024-08-12 time-series-analysis.qmd
20240811 - fantasy football projections

@@
     y = "Fantasy Points",
-    title = "Tom Brady's Historical and Projected Fantasy Points by Game"
+    title = "Tom Brady's Historical and Projected Fantasy Points by Game",
+    subtitle = "(if he were to have continued playing additional seasons)"
   ) +

### 2024-08-12 simulation.qmd
20240605 - getting started/load packages
20240608 - getting started
Built site for gh-pages
20240703 - conclusion section
20240811 - fantasy football projections

@@
 # Simulation: the Monte Carlo Method {#sec-simulation}
+
+## Getting Started {#sec-simulationGettingStarted}
+
+### Load Packages {#sec-simulationLoadPackages}
+
+```{r}
+library("tidyverse")
+```
+
+### Load Data {#sec-simulationLoadData}
+
+```{r}
+load(file = "./data/players_projectedPoints_seasonal.RData")
+```
+
+## Overview {#sec-simulationOverview}
+
+## Simulation of Projected Statistics and Points {#sec-simulationProjectedPoints}
+
+```{r}
+calculate_fantasyPoints <- function(df, scoring_rules){
+  #fantasyPoints <-
+
+  #return(fantasyPoints)
+}
+
+monte_carlo_simulation <- function(position_df, num_simulations = 10000){
+  player_ids <- unique(position_df$id)
+
+  results <- lapply(player_ids, function(player_id){
+    player_data <- position_df %>% filter(id == player_id)
+
+    # Run the simulation for this player
+    simulations <- replicate(num_simulations, {
+      sampled_data <- player_data %>%
+        summarise(across(
+          c(pass_yds, pass_tds, rush_yds, rush_tds, rec_yds, rec_tds), 
+          ~ if (all(is.na(.))) NA else sample(.[!is.na(.)], 1, replace = TRUE)))
+
+      sampled_data
+    }, simplify = FALSE)
+
+    simulation_df <- bind_rows(simulations)
+    #simulation_df$fantasyPoints <- calculate_fantasyPoints(simulation_df)
+
+    ci <- simulation_df %>%
+      summarise(across(where(is.numeric), list(
+        lower = \(x) quantile(x, 0.025, na.rm = TRUE),
+        upper = \(x) quantile(x, 0.975, na.rm = TRUE))))
+
+    ci <- ci %>% mutate(id = player_id)
+
+    return(ci)
+    })
+
+    # Combine results for all players
+    final_ci <- bind_rows(results)
+
+    return(final_ci)
+}
+
+monte_carlo_simulation(players_projectedPoints_seasonal$QB, num_simulations = 10)
+
+QB_ci <- monte_carlo_simulation(players_projectedPoints_seasonal$QB)
+```
+
+## Conclusion {#sec-simulationConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-simulationSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-08-12 simulation.qmd
Built site for gh-pages

@@
 
-  results <- lapply(player_ids, function(player_id){
+  results <- lapply(player_ids, function(player_id){ # use curly brackets when there are multiple expressions
     player_data <- position_df %>% filter(id == player_id)
@@
     # Run the simulation for this player
-    simulations <- replicate(num_simulations, {
+    simulations <- replicate(
+      num_simulations, {
         sampled_data <- player_data %>%
           summarise(across(
-          c(pass_yds, pass_tds, rush_yds, rush_tds, rec_yds, rec_tds), 
+            c(pass_yds, pass_tds, rush_yds, rush_tds, rec_yds, rec_tds), # add more statistical categories
             ~ if (all(is.na(.))) NA else sample(.[!is.na(.)], 1, replace = TRUE)))
 
-      sampled_data
-    }, simplify = FALSE)
+        sampled_data # return last expression
+        }, 
+      simplify = FALSE)
 
-    simulation_df <- bind_rows(simulations)
-    #simulation_df$fantasyPoints <- calculate_fantasyPoints(simulation_df)
+    simulation_df <- dplyr::bind_rows(simulations)
+    #simulation_df$fantasyPoints <- calculate_fantasyPoints(simulation_df) # write function to calculate fantasy points
 
@@
 
-    ci <- ci %>% mutate(id = player_id)
+    ci <- ci %>% 
+      mutate(id = player_id)
 

### 2024-08-13 time-series-analysis.qmd
20240812 - ARIMA additions

--- 2024-08-11 time-series-analysis.qmd
@@
 Time series analysis is useful when trying to generate forecasts from longitudinal data.
+That is, time series analysis seeks to evaluate change over time to predict future values.
+
 There many different types of time series analyses.
 For simplicity, in this chapter, we use autoregressive integrated moving average (ARIMA) models to demonstrate one approach to time series analysis.
-ARIMA models perform best when there is a clear pattern where later values are influenced by earlier values (i.e., autoregressive effects and moving average effects).
+
+## Autoregressive Integrated Moving Average (ARIMA) Models {#sec-timeSeriesARIMA}
+
+@Hyndman2021 provide a nice overview of ARIMA models.
+As noted by @Hyndman2021, ARIMA models aim to describe how a variable is correlated with itself over time (autocorrelation)—i.e., how earlier levels of a variable are correlated with later levels of the same variable.
+ARIMA models perform best when there is a clear pattern where later values are influenced by earlier values.
+ARIMA models incorporate autoregression effects, moving average effects, and differencing.
+
+ARIMA models can have various numbers of terms and model complexity.
+They are specified in the following form: $\text{ARIMA}(p,d,q)$, where:
+
+- $p =$ the number of autoregressive terms
+- $d =$ the number of differences between consecutive scores (to make the time series stationary by reducing trends and seasonality)
+- $q =$ the number of moving average terms
+
+ARIMA models assume that the data are stationary (i.e., there are no long-term trends), are non-seasonal (i.e., there is no consistency of the timing of the peaks or troughs in the line), and that earlier values influence later values.
 This may not strongly be the case in fantasy football, so ARIMA models may not be particularly useful in forecasting fantasy football performance.
+Other approaches, such as exponential smoothing, may be useful for data that show longer-term trends and seasonality [@Hyndman2021].
 Nevertheless, ARIMA models are widely used in forecasting financial markets and economic indicators.
@@
 ```{r}
+#| label: fig-timeSeriesHistoricalTomBrady
+#| fig-cap: "Tom Brady's Historical Fantasy Points by Game."
+
 plot(
@@
 ```{r}
+#| label: fig-timeSeriesHistoricalBradyManning
+#| fig-cap: "Historical Fantasy Points by Game for Tom Brady and Peyone Manning."
+
 plot(
@@
 
+The autocorrelation function (ACF) plot depicts the autocorrelation of scores as a function of the length of the lag.
+Significant autocorrelation is detected when the autocorrelation exceeds the dashed blue lines, as is depicted in @fig-timeSeriesACFtomBrady.
+
 ```{r}
+#| label: fig-timeSeriesACFtomBrady
+#| fig-cap: "Autocorrelation Function (ACF) Plot of Tom Brady's Historical Fantasy Points by Game."
+
 acf(ts_tomBrady)
@@
 ```{r}
+#| label: fig-arimaTomBrady
+#| fig-cap: "Model Summary of Autoregressive Integrated Moving Average Model fit to Tom Brady's Historical Performance by Game."
+
 arima_tomBrady <- arima(
@@
 forecast::checkresiduals(arima_tomBrady)
+```
+
+```{r}
+#| label: fig-arimaModifiedTomBrady
+#| fig-cap: "Model Summary of modified Autoregressive Integrated Moving Average Model fit to Tom Brady's Historical Performance by Game."
 
@@
 forecast::checkresiduals(arima_tomBrady_removeNonSigTerms)
+```
+
+```{r}
+#| label: fig-arimaPeytonManning
+#| fig-cap: "Model Summary of Autoregressive Integrated Moving Average Model fit to Peyton Manning's Historical Performance by Game."
 
@@
 ```{r}
+#| label: fig-timeSeriesProjectionTomBrady
+#| fig-cap: "Tom Brady's Historical and Projected Fantasy Points by Game."
+
 forecast::autoplot(forecast_tomBrady) + 
@@
     y = "Fantasy Points",
-    title = "Tom Brady's Historical and Projected Fantasy Points by Game"
+    title = "Tom Brady's Historical and Projected Fantasy Points by Game",
+    subtitle = "(if he were to have continued playing additional seasons)"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-timeSeriesProjectionPeytonManning
+#| fig-cap: "Peyton Manning's Historical and Projected Fantasy Points by Game."
+
+forecast::autoplot(forecast_peytonManning) + 
+  labs(
+    x = "Game Number",
+    y = "Fantasy Points",
+    title = "Peyton Manning's Historical and Projected Fantasy Points by Game",
+    subtitle = "(if he were to have continued playing additional seasons)"
   ) +

### 2024-08-13 getting-started.qmd
20240812 - load data to fix error
20240812 - create a vector and list

@@
 
+## Create a Vector {#sec-createVector}
+
+A vector is a series of elements that can be numeric or character.
+It has one dimension (length).
+To create a vector, use the `c()` to combine elements into a vector:
+
+```{r}
+exampleVector <- c(40, 30, 24, 20, 18, 23, 27, 32, 26, 23, NA, 37)
+```
+
 ## Create a Data Frame {#sec-createDF}
 
+A data frame has two dimensions: rows and columns.
 Here is an example of creating a data frame:
@@
 )
+```
+
+## Create a List {#sec-createList}
+
+A list can store multiple data frames in one object:
+
+```{r}
+exampleList <- list(players, fantasyPoints, fantasyPoints_weekly)
+```
+
+## Load a Data Frame {#sec-gettingStartedLoadData}
+
+```{r}
+load(file = "./data/nfl_players.RData")
 ```

### 2024-08-13 cluster-analysis.qmd
20240813 - WR types

@@
 ```{r}
+library("petersenlab")
 library("mclust")
@@
 
+load(file = "./data/nfl_players.RData")
+load(file = "./data/nfl_combine.RData")
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
+load(file = "./data/nfl_actualStats_career.RData")
 ```
@@
 ```{r}
+load(file = "./data/nfl_players.RData")
+load(file = "./data/nfl_combine.RData")
 load(file = "./data/player_stats_weekly.RData")
 load(file = "./data/player_stats_seasonal.RData")
+load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
+load(file = "./data/nfl_actualStats_career.RData")
 ```
@@
 
-### Example: Tiers of Prior Season Fantasy Points {#sec-clusterAnalysisExample}
+### Tiers of Prior Season Fantasy Points {#sec-clusterAnalysisExample}
 
@@
 ```{r}
-tiersQB_bic <- mclustBIC(
+tiersQB_bic <- mclust::mclustBIC(
   data = player_stats_seasonal_offense_recentQB$fantasy_points,
@@
 
-tiersQB_icl <- mclustICL(
+tiersQB_icl <- mclust::mclustICL(
   data = player_stats_seasonal_offense_recentQB$fantasy_points,
@@
 
-tiersQB_boostrap <- mclustBootstrapLRT(
-  player_stats_seasonal_offense_recentQB$fantasy_points,
+tiersQB_boostrap <- mclust::mclustBootstrapLRT(
+  data = player_stats_seasonal_offense_recentQB$fantasy_points,
   modelName = "V") # variable/unequal variance (for univariate data)
@@
 ```{r}
-tiersRB_bic <- mclustBIC(
+tiersRB_bic <- mclust::mclustBIC(
   data = player_stats_seasonal_offense_recentRB$fantasy_points,
@@
 
-tiersRB_icl <- mclustICL(
+tiersRB_icl <- mclust::mclustICL(
   data = player_stats_seasonal_offense_recentRB$fantasy_points,
@@
 ```{r}
-tiersRB_boostrap <- mclustBootstrapLRT(
-  player_stats_seasonal_offense_recentRB$fantasy_points,
+tiersRB_boostrap <- mclust::mclustBootstrapLRT(
+  data = player_stats_seasonal_offense_recentRB$fantasy_points,
   modelName = "V") # variable/unequal variance (for univariate data)
@@
 ```{r}
-tiersWR_bic <- mclustBIC(
+tiersWR_bic <- mclust::mclustBIC(
   data = player_stats_seasonal_offense_recentWR$fantasy_points,
@@
 
-tiersWR_icl <- mclustICL(
+tiersWR_icl <- mclust::mclustICL(
   data = player_stats_seasonal_offense_recentWR$fantasy_points,
@@
 
-tiersWR_boostrap <- mclustBootstrapLRT(
-  player_stats_seasonal_offense_recentWR$fantasy_points,
+tiersWR_boostrap <- mclust::mclustBootstrapLRT(
+  data = player_stats_seasonal_offense_recentWR$fantasy_points,
   modelName = "V") # variable/unequal variance (for univariate data)
@@
 ```{r}
-tiersTE_bic <- mclustBIC(
+tiersTE_bic <- mclust::mclustBIC(
   data = player_stats_seasonal_offense_recentTE$fantasy_points,
@@
 
-tiersTE_icl <- mclustICL(
+tiersTE_icl <- mclust::mclustICL(
   data = player_stats_seasonal_offense_recentTE$fantasy_points,
@@
 
-tiersTE_boostrap <- mclustBootstrapLRT(
-  player_stats_seasonal_offense_recentTE$fantasy_points,
+tiersTE_boostrap <- mclust::mclustBootstrapLRT(
+  data = player_stats_seasonal_offense_recentTE$fantasy_points,
   modelName = "V") # variable/unequal variance (for univariate data)
@@
 
-Mclust(
+mclust::Mclust(
   data = player_stats_seasonal_offense_recentQB$fantasy_points,
@@
 ```{r}
-clusterModelQBs <- Mclust(
+clusterModelQBs <- mclust::Mclust(
   data = player_stats_seasonal_offense_recentQB$fantasy_points,
@@
 ```{r}
-clusterModelRBs <- Mclust(
+clusterModelRBs <- mclust::Mclust(
   data = player_stats_seasonal_offense_recentRB$fantasy_points,
@@
 ```{r}
-clusterModelWRs <- Mclust(
+clusterModelWRs <- mclust::Mclust(
   data = player_stats_seasonal_offense_recentWR$fantasy_points,
@@
 ```{r}
-clusterModelTEs <- Mclust(
+clusterModelTEs <- mclust::Mclust(
   data = player_stats_seasonal_offense_recentTE$fantasy_points,
@@
 
+### Types of Wide Receivers {#sec-clusterAnalysisWRtypes}
+
+```{r}
+#| eval: false
+#| include: false
+
+#names(nfl_players) #gsis_id: height, weight
+#names(nfl_combine) #gsis_id: vertical, forty, ht, wt
+#names(player_stats_seasonal_offense) #player_id, season: receptions, receiving_air_yards, air_yards_share, target_share
+#names(nfl_advancedStatsPFR_seasonal) #gsis_id, season: adot.rec, rec.rec, brk_tkl.rec, drop.rec, drop_percent.rec
+#names(nfl_actualStats_offense_career) #player_id: receptions, targets, receiving_air_yards, air_yards_share, target_share
+```
+
+```{r}
+# Compute Advanced PFR Stats by Career
+pfrVars <- nfl_advancedStatsPFR_seasonal %>% 
+  select(pocket_time.pass:cmp_percent.def, g, gs) %>% 
+  names()
+
+weightedAverageVars <- c(
+  "pocket_time.pass",
+  "ybc_att.rush","yac_att.rush",
+  "ybc_r.rec","yac_r.rec","adot.rec","rat.rec",
+  "yds_cmp.def","yds_tgt.def","dadot.def","m_tkl_percent.def","rat.def"
+)
+
+recomputeVars <- c(
+  "drop_pct.pass", # drops.pass / pass_attempts.pass
+  "bad_throw_pct.pass", # bad_throws.pass / pass_attempts.pass
+  "on_tgt_pct.pass", # on_tgt_throws.pass / pass_attempts.pass
+  "pressure_pct.pass", # times_pressured.pass / pass_attempts.pass
+  "drop_percent.rec", # drop.rec / tgt.rec
+  "rec_br.rec", # rec.rec / brk_tkl.rec
+  "cmp_percent.def" # cmp.def / tgt.def
+)
+
+sumVars <- pfrVars[pfrVars %ni% c(
+  weightedAverageVars, recomputeVars,
+  "nameMerge", "loaded.pass", "loaded.rush", "loaded.rec", "loaded.def")]
+
+nfl_advancedStatsPFR_career <- nfl_advancedStatsPFR_seasonal %>% 
+  group_by(pfr_id, nameMerge) %>% 
+  summarise(
+    across(all_of(weightedAverageVars), ~ weighted.mean(.x, w = g, na.rm = TRUE)),
+    across(all_of(sumVars), ~ sum(.x, na.rm = TRUE)),
+    .groups = "drop") %>% 
+  mutate(
+    drop_pct.pass = drops.pass / pass_attempts.pass,
+    bad_throw_pct.pass = bad_throws.pass / pass_attempts.pass,
+    on_tgt_pct.pass = on_tgt_throws.pass / pass_attempts.pass,
+    pressure_pct.pass = times_pressured.pass / pass_attempts.pass,
+    drop_percent.rec = drop.rec / tgt.rec,
+    rec_br.rec = drop.rec / tgt.rec,
+    cmp_percent.def = cmp.def / tgt.def
+  )
+
+uniqueCases <- nfl_advancedStatsPFR_seasonal %>% select(pfr_id, nameMerge, gsis_id) %>% unique()
+
+uniqueCases %>%
+  group_by(pfr_id) %>% 
+  filter(n() > 1)
+
+nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal %>% 
+  filter(pfr_id != "WillMa06" | nameMerge != "MARCUSWILLIAMS" | !is.na(gsis_id))
+
+nfl_advancedStatsPFR_career <- left_join(
+  nfl_advancedStatsPFR_career,
+  nfl_advancedStatsPFR_seasonal %>% select(pfr_id, nameMerge, gsis_id) %>% unique(),
+  by = c("pfr_id", "nameMerge")
+)
+
+# Compute Player Stats Per Season
+player_stats_seasonal_careerWRs <- player_stats_seasonal_offense %>% 
+  filter(position == "WR") %>% 
+  group_by(player_id) %>% 
+  summarise(
+    across(all_of(c("targets", "receptions", "receiving_air_yards")), ~ weighted.mean(.x, w = games, na.rm = TRUE)),
+    .groups = "drop")
+
+# Drop players with no receiving air yards
+player_stats_seasonal_careerWRs <- player_stats_seasonal_careerWRs %>% 
+  filter(receiving_air_yards != 0) %>% 
+  rename(
+    targets_per_season = targets,
+    receptions_per_season = receptions,
+    receiving_air_yards_per_season = receiving_air_yards
+  )
+
+# Merge
+playerListToMerge <- list(
+  nfl_players %>% select(gsis_id, display_name, position, height, weight),
+  nfl_combine %>% select(gsis_id, vertical, forty, ht, wt),
+  player_stats_seasonal_careerWRs %>% select(player_id, targets_per_season, receptions_per_season, receiving_air_yards_per_season) %>% 
+    rename(gsis_id = player_id),
+  nfl_actualStats_offense_career %>% select(player_id, receptions, targets, receiving_air_yards, air_yards_share, target_share) %>% 
+    rename(gsis_id = player_id),
+  nfl_advancedStatsPFR_career %>% select(gsis_id, adot.rec, rec.rec, brk_tkl.rec, drop.rec, drop_percent.rec)
+)
+
+merged_data <- playerListToMerge %>% 
+  reduce(
+    full_join,
+    by = c("gsis_id"),
+    na_matches = "never")
+```
+
+Additional processing:
+
+```{r}
+merged_data <- merged_data %>% 
+  separate(ht, into = c("feet", "inches"), sep = "-") %>%
+  mutate(
+    feet = as.numeric(feet),
+    inches = as.numeric(inches),
+    ht_numeric = feet * 12 + inches
+  ) %>%
+  select(-feet, -inches)
+
+merged_data <- merged_data %>% 
+  mutate(
+    height_coalesced = coalesce(height, ht_numeric),
+    weight_coalesced = coalesce(weight, wt),
+    receptions_coalesced = pmax(receptions, rec.rec, na.rm = TRUE),
+    receiving_air_yards_per_rec = receiving_air_yards / receptions
+  )
+
+merged_data$receiving_air_yards_per_rec[which(merged_data$receptions == 0)] <- 0
+
+merged_dataWRs <- merged_data %>% 
+  filter(position == "WR")
+
+merged_dataWRs_cluster <- merged_dataWRs %>% 
+  filter(receptions_coalesced >= 100) %>% # keep WRs with at least 100 receptions
+  select(gsis_id, display_name, vertical, forty, height_coalesced, weight_coalesced, targets_per_season, receptions_per_season, receiving_air_yards_per_season, air_yards_share, target_share, adot.rec, brk_tkl.rec, drop_percent.rec, receiving_air_yards_per_rec) %>%
+  na.omit()
+```
+
+#### Identify the Number of WR Types {#sec-clusterAnalysisNumWRtypes}
+
+```{r}
+wrTypes_bic <- mclust::mclustBIC(
+  data = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
+  G = 1:9
+)
+
+wrTypes_bic
+summary(wrTypes_bic)
+plot(wrTypes_bic)
+
+wrTypes_icl <- mclust::mclustICL(
+  data = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
+  G = 1:9
+)
+
+wrTypes_icl
+summary(wrTypes_icl)
+plot(wrTypes_icl)
+```
+
+```{r}
+#| eval: false
+
+wrTypes_boostrap <- mclust::mclustBootstrapLRT(
+  data = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
+  modelName = "VVV") # ellipsoidal (variable volume, variable shape), variable orientation (for multivariate data)
+
+wrTypes_boostrap
+plot(
+  wrTypes_boostrap,
+  G = numTypesWR - 1)
+```
+
+#### Fit the Cluster Model to the Optimal Number of WR Types {#sec-clusterAnalysisModelWRtypes}
+
+```{r}
+numTypesWR <- 3
+
+clusterModelWRtypes <- mclust::Mclust(
+  data = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
+  G = numTypesWR,
+)
+
+summary(clusterModelWRtypes)
+```
+
+```{r}
+plot(
+  clusterModelWRtypes,
+  what = "BIC")
+```
+
+```{r}
+plot(
+  clusterModelWRtypes,
+  what = "classification")
+```
+
+```{r}
+plot(
+  clusterModelWRtypes,
+  what = "uncertainty")
+```
+
+```{r}
+plot(
+  clusterModelWRtypes,
+  what = "density")
+```
+
+```{r}
+table(clusterModelWRtypes$classification)
+
+merged_dataWRs_cluster$type <- clusterModelWRtypes$classification
+
+merged_dataWRs_cluster %>% 
+  group_by(type) %>% 
+  summarise(across(
+    where(is.numeric),
+    ~ mean(., na.rm = TRUE)
+    )) %>% 
+  t() %>% 
+  round(., 2)
+```
+
 ## Conclusion {#sec-clusterAnalysisConclusion}

### 2024-08-13 download-football-data.qmd
20240813 - WR types

@@
 
+Let's do some data processing to help with merging the dataset with other datasets:
+
+```{r}
+nfl_playerIDs <- nfl_playerIDs %>% 
+  mutate(
+    nameMerge = petersenlab::cleanUpNames(name)
+  )
+```
+
 ```{r}
@@
 
-Let's check for duplicate `pfr_id` and `cfb_id` instances:
-
-```{r}
+Let's do some data processing to help with merging the dataset with other datasets:
+
+```{r}
+nfl_combine <- nfl_combine %>% 
+  mutate(
+    nameMerge = petersenlab::cleanUpNames(player_name)
+  )
+
+# First, merge on both pfr_id and cfb_id
+merged_data1 <- left_join(
+  nfl_combine,
+  nfl_playerIDs %>% select(pfr_id, cfbref_id, gsis_id),
+  by = c("pfr_id", "cfb_id" = "cfbref_id"),
+  na_matches = "never"
+)
+
+# Second, merge on pfr_id
+merged_data2 <- left_join(
+  nfl_combine,
+  nfl_playerIDs %>% select(pfr_id, nameMerge, position, gsis_id),
+  by = c("pfr_id","nameMerge","pos" = "position"),
+  na_matches = "never"
+)
+
+# Third, merge on cfb_id
+merged_data3 <- left_join(
+  nfl_combine,
+  nfl_playerIDs %>% select(cfbref_id, nameMerge, position, gsis_id),
+  by = c("cfb_id" = "cfbref_id","nameMerge","pos" = "position"),
+  na_matches = "never"
+)
+
+# Combine gsis_id across merges
+nfl_combine$gsis_id <- coalesce(
+  merged_data1$gsis_id,
+  merged_data2$gsis_id,
+  merged_data3$gsis_id
+  )
+
+# Rearrange the data
+nfl_combine <- nfl_combine %>% 
+  select(gsis_id, pfr_id, cfb_id, player_name, everything()) %>% 
+  arrange(season, player_name)
+```
+
+Let's check for duplicate `gsis_id`, `pfr_id`, and `cfb_id` instances:
+
+```{r}
+nfl_combine %>% 
+  group_by(gsis_id) %>% 
+  filter(n() > 1, !is.na(gsis_id)) %>% 
+  arrange(gsis_id) %>% 
+  head()
+
 nfl_combine %>% 
@@
 
-However, these apparent duplicates appear to be different players:
-
-```{r}
+Let's do some additional data processing:
+
+```{r}
+# Drop Stanford Samuels Jr.
+nfl_combine$gsis_id[which(nfl_combine$cfb_id == "stanford-samuels-1")] <- NA
+```
+
+However, these apparent duplicates appear to be different players at different positions:
+
+```{r}
+nfl_combine %>% 
+  group_by(season, gsis_id, pos) %>% 
+  filter(n() > 1, !is.na(gsis_id)) %>% 
+  arrange(gsis_id) %>% 
+  head()
+
 nfl_combine %>% 

### 2024-08-13 modern-portfolio-theory.qmd
20240727 - fixes
20240813 - WR types

@@
 In this way, modern portfolio theory and the efficient frontier can be helpful for arbitrage—buying and selling the same asset (in this case, player) to take advantage of different prices for the same asset.
+That is, you could buy low and, for players who outperform expectations, sell high—in the form of a trade.
 
@@
 
-### Based on Historical Game-to-Game Variability {#sec-efficientFrontierFantasyProjectionsUncertainty}
+### Based on Historical Game-to-Game Variability {#sec-efficientFrontierFantasyProjectionsGameToGameVariability}
 

### 2024-08-14 cluster-analysis.qmd
20240813 - move processing earlier

@@
 merged_data <- merged_data %>% 
-  separate(ht, into = c("feet", "inches"), sep = "-") %>%
-  mutate(
-    feet = as.numeric(feet),
-    inches = as.numeric(inches),
-    ht_numeric = feet * 12 + inches
-  ) %>%
-  select(-feet, -inches)
-
-merged_data <- merged_data %>% 
   mutate(

### 2024-08-14 download-football-data.qmd
20240813 - move processing earlier

@@
 nfl_combine$gsis_id[which(nfl_combine$cfb_id == "stanford-samuels-1")] <- NA
+
+nfl_combine <- nfl_combine %>% 
+  separate_wider_delim(
+    ht,
+    names = c("feet", "inches"),
+    delim = "-") %>%
+  mutate(
+    feet = as.numeric(feet),
+    inches = as.numeric(inches),
+    ht = feet * 12 + inches
+  ) %>%
+  select(-feet, -inches)
 ```

### 2024-08-14 data-visualization.qmd
20240813 - plot titles and highlighting

@@
 ```{r}
-library("tidyverse")
 library("nflplotR")
 library("plotly")
+library("gghighlight")
+library("tidyverse")
 ```
@@
 
+A key principle of graphic design and data visualization is the importance of contrast.
+Each visual component (e.g., line) that is important to see should be easy to distinguish.
+For instance, you can highlight lines or points of interest to draw people's attention to the target of interest.
+For examples of highlighting in figures, see Figures [-@fig-lineChartHighlighting] (@sec-lineChartHighlighting) and [-@fig-simulationOf10CoinFlips].
+
 ## Univariate Distribution {#sec-univariateDistribution}
@@
   ) +
+  labs(
+    x = "Fantasy Points",
+    title = "Histogram of Fantasy Points"
+  ) +
   theme_classic()
@@
 #| label: fig-density
-#| fig-cap: "Density Plot of Fantasy Points."
+#| fig-cap: "Density Plot of Fantasy Points by Position."
 
@@
   geom_density(alpha = 0.7) + # add transparency
+  labs(
+    x = "Fantasy Points",
+    fill = "Position",
+    title = "Density Plot of Fantasy Points by Position"
+  ) +
   theme_classic()
@@
   geom_rug() +
+  labs(
+    x = "Fantasy Points",
+    title = "Histogram of Fantasy Points with Overlaid Density and Rug Plot"
+  ) +
   theme_classic()
@@
   geom_boxplot() +
+  labs(
+    x = "Position",
+    y = "Fantasy Points",
+    title = "Box-and-Whisker Plot of Fantasy Points by Position"
+  ) +
   theme_classic() + 
@@
 #| label: fig-violin
-#| fig-cap: "Violin Plot."
+#| fig-cap: "Violin Plot. Lines represent the 25th, 50th, and 75th quantiles."
 
@@
   geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
+  labs(
+    x = "Position",
+    y = "Fantasy Points",
+    title = "Violin Plot by Position",
+    subtitle = "Lines represent the 25th, 50th, and 75th quantiles"
+  ) +
   theme_classic() + 
@@
     x = season,
-    y = fantasy_points,
-    fill = position_group
+    y = fantasy_points
     )
@@
     color = "blue"
+  ) +
+  labs(
+    x = "Season",
+    y = "Fantasy Points",
+    title = "Bar Plot of Tom Brady's Fantasy Points by Season"
+  ) +
+  theme_classic()
+```
+
+### With Highlighting {#sec-lineChartHighlighting}
+
+```{r}
+#| label: fig-lineChartHighlighting
+#| fig-cap: "Line Chart with Highlighting."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense,
+  mapping = aes(
+    x = season,
+    y = fantasy_points,
+    group = player_id,
+    color = player_display_name)
+) +
+  geom_line() +
+  gghighlight::gghighlight(
+    player_display_name == "Tom Brady",
+    label_key = player_display_name) +
+  labs(
+    x = "Season",
+    y = "Fantasy Points",
+    title = "Fantasy Points by Season and Player",
+    subtitle = "(Tom Brady in Red)"
   ) +
@@
     stat = "identity") +
-  theme_classic()
+  labs(
+    x = "Position",
+    y = "Fantasy Points",
+    title = "Bar Plot of Fantasy Points by Position"
+  ) +
+  theme_classic() + 
+  theme(legend.position = "none")
 ```
@@
   ) +
-  theme_classic()
+  labs(
+    x = "Position",
+    y = "Fantasy Points",
+    title = "Bar Plot of Fantasy Points by Position"
+  ) +
+  theme_classic() + 
+  theme(legend.position = "none")
 ```

### 2024-08-14 cluster-analysis.qmd
20240814 - WR types

@@
 library("mclust")
+library("plotly")
 library("tidyverse")
-library("plotly")
 ```
@@
   mutate(
-    height_coalesced = coalesce(height, ht_numeric),
+    height_coalesced = coalesce(height, ht),
     weight_coalesced = coalesce(weight, wt),
@@
   data = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
-  modelName = "VVV") # ellipsoidal (variable volume, variable shape), variable orientation (for multivariate data)
+  modelName = "VVE") # ellipsoidal (variable volume, variable shape), equal orientation (for multivariate data)
 
@@
 
+Type 1 WRs:
+
+```{r}
+merged_dataWRs_cluster %>% 
+  filter(type == 1) %>% 
+  select(display_name)
+```
+
+Type 2 WRs:
+
+```{r}
+merged_dataWRs_cluster %>% 
+  filter(type == 2) %>% 
+  select(display_name)
+```
+
+Type 3 WRs:
+
+```{r}
+merged_dataWRs_cluster %>% 
+  filter(type == 3) %>% 
+  select(display_name)
+```
+
 ## Conclusion {#sec-clusterAnalysisConclusion}

### 2024-08-14 data-visualization.qmd
20240814 - color scheme

@@
 The R Graph Gallery provides examples of various types of plots: <https://r-graph-gallery.com>.
+In this chapter, we will examine how to create statistical graphics to visualize data.
+We will create the plots using the `ggplot2` package.
+When creating plots in `ggplot2` with multiple points or lines (e.g., multiple players or levels of a predictor variable), it is easiest to do so with the data in [long form](#sec-wideToLong) (as opposed to [wide form](#sec-longToWide)).
 
@@
 For examples of highlighting in figures, see Figures [-@fig-lineChartHighlighting] (@sec-lineChartHighlighting) and [-@fig-simulationOf10CoinFlips].
+It is also important to use color schemes with distinguishable colors.
+Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package, as demonstrated in @fig-barPlotConfidenceInterval (@sec-barPlotModifiedColorScheme).
 
@@
 
+### With Error Bars {#sec-barPlotErrorBars}
+
 Based on the summary statistics in @tbl-summaryStats, we create a bar plot with bars representing the 95% confidence interval in @fig-barPlotConfidenceInterval.
@@
     stat = "identity") +
+  geom_errorbar(
+    aes(
+      ymin = ci_lower,
+      ymax = ci_upper),
+    width = 0.2,
+    color = "black"
+  ) +
+  labs(
+    x = "Position",
+    y = "Fantasy Points",
+    title = "Bar Plot of Fantasy Points by Position"
+  ) +
+  theme_classic() + 
+  theme(legend.position = "none")
+```
+
+### Modified Color Scheme {#sec-barPlotModifiedColorScheme}
+
+We can also modify the color scheme, as in @fig-barPlotConfidenceInterval.
+
+```{r}
+#| label: fig-barPlotConfidenceInterval
+#| fig-cap: "Bar Plot with Bars Representing the 95% Confidence Interval."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense_summary %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = position_group,
+    y = mean,
+    fill = position_group,
+    color = position_group
+    )
+) +
+  geom_bar(
+    stat = "identity") +
+  scale_fill_brewer(palette = "Set1") +
   geom_errorbar(

### 2024-08-14 data-visualization.qmd
20240804 - fix error due to duplicate identifier

@@
 It is also important to use color schemes with distinguishable colors.
-Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package, as demonstrated in @fig-barPlotConfidenceInterval (@sec-barPlotModifiedColorScheme).
+Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package, as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
 
@@
 
-We can also modify the color scheme, as in @fig-barPlotConfidenceInterval.
+We can also modify the color scheme, as in @fig-barPlotColorScheme
 
 ```{r}
-#| label: fig-barPlotConfidenceInterval
+#| label: fig-barPlotColorScheme
 #| fig-cap: "Bar Plot with Bars Representing the 95% Confidence Interval."

### 2024-08-14 cluster-analysis.qmd
20240814 - interpreting the clusters

@@
 In this chapter, we focus on examples using model-based clustering with the `mclust` package [@Scrucca2023], which uses Gaussian finite mixture modeling.
+The various types of `mclust` models are provided here:
+<https://mclust-org.github.io/mclust/reference/mclustModelNames.html>.
 
@@
   filter(receptions_coalesced >= 100) %>% # keep WRs with at least 100 receptions
-  select(gsis_id, display_name, vertical, forty, height_coalesced, weight_coalesced, targets_per_season, receptions_per_season, receiving_air_yards_per_season, air_yards_share, target_share, adot.rec, brk_tkl.rec, drop_percent.rec, receiving_air_yards_per_rec) %>%
+  select(gsis_id, display_name, vertical, forty, height_coalesced, weight_coalesced, adot.rec, drop_percent.rec, receiving_air_yards_per_rec, brk_tkl.rec, receptions_per_season) %>% #targets_per_season, receiving_air_yards_per_season, air_yards_share, target_share
   na.omit()
@@
 
+Based on the cluster analyses, it appears that three clusters are the best fit to the data.
+
+```{r}
+numTypesWR <- 3
+```
+
 ```{r}
@@
   data = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
-  modelName = "VVE") # ellipsoidal (variable volume, variable shape), equal orientation (for multivariate data)
+  modelName = "EVE") # ellipsoidal with equal volume, variable shape, and equal orientation (for multivariate data)
 
@@
 ```{r}
-numTypesWR <- 3
-
 clusterModelWRtypes <- mclust::Mclust(
@@
 ```
+
+#### Plots of the Cluster Model {#sec-clusterAnalysisPlotsWRtypes}
 
@@
 
+#### Interpreting the Clusters {#sec-clusterAnalysisInterpretationWRtypes}
+
 ```{r}
@@
 
-Type 1 WRs:
+Based on this analysis (and the variables included), there appear to be three types of Wide Receivers.
+Type 1 Wide Receivers includes the Elite WR1s who are strong possession receivers (note: not all players in a given cluster map on perfectly to the typology—i.e., not all Type 1 Wide Receivers are elite WR1s).
+They tend to have the lowest drop percentage, the shortest average depth of target, and the fewest receiving air yards per reception.
+They tend to have the most receptions per season and break the most tackles.
+
+Type 2 Wide Receivers includes the consistent contributor, WR2 types.
+They had fewer receptions and fewer broken tackles than Type 1 Wide Receivers.
+Their average depth of target was longer than Type 1, and they had more receiving air yards per reception than Type 1.
+
+Type 3 Wide Receivers includes the deep threats.
+They have the greatest average depth of target and the most receiving yards per reception.
+However, they also have the fewest receptions, the highest drop percentage, and the fewest broken tackles.
+Thus, they may be considered the boom-or-bust Wide Receivers.
+
+The tiers were not particularly distinguishable based on their height, weight, vertical jump, or forty-yard dash time.
+
+Type 1 ("Elite/WR1") WRs:
 
@@
 
-Type 2 WRs:
+Type 2 ("Consistent Contributor/WR2") WRs:
 
@@
 
-Type 3 WRs:
+Type 3 ("Deep Threat/Boom-or-Bust") WRs:
 

### 2024-08-14 data-visualization.qmd
20240814 - principles of design and data visualization

@@
 ## Overview {#sec-dataVisualizationOverview}
+
+### Principles of Graphic Design {#sec-dataVisualizationOverviewGraphicDesign}
+
+When designing graphics, it is important to understand general principles of graphic design.
+An article that describes important principles about graphic design is at the following link: <https://www.adobe.com/express/learn/blog/8-basic-design-principles-to-help-you-create-better-graphics> (archived at <https://perma.cc/29P9-NNSK>).
+The important principles include:
+
+> 1. Focus on alignment.
+> 2. Use hierarchy to help focus your design.
+> 3. Leverage contrast to accentuate important design elements.
+> 4. Use repetition to your advantage.
+> 5. Consider proximity when organizing your graphic elements.
+> 6. Make sure your designs have balance.
+> 7. Optimize color to support your design.
+> 8. Leave negative space.
+
+### Principles of Data Visualization {#sec-dataVisualizationOverviewPrinciples}
+
+Data visualization involves graphic design in a particular domain—the visualization of data (numeric-derived information).
+@Schwabish2021 describes five principles in data visualization:
+
+> 1. Show the data.
+> 2. Reduce the clutter.
+> 3. Integrate the graphics and text.
+> 4. Avoid the spaghetti chart.
+> 5. Start with gray.
+
+In his classic book, @Tufte2001 states that effective data visualizations should follow principles of graphical excellence and integrity.
+He notes that "Graphical excellence is that which gives to the viewer the greatest number of ideas in the shortest time with the least ink in the smallest space." (p. 51).
+That is, data visualizations should seek to maximize the data-to-ink ratio (within reason), and should spend less space on "fluff" (i.e., non-data things that can be erased without losing meaning, such as grid lines, redundancies, etc.).
+@Tufte2001 describes six principles of graphical integrity:
+
+> 1. The representation of numbers, as physically measured on the surface of the graphic itself, should be directly proportional to the numerical quantities represented.
+> 2. Clear, detailed, and thorough labeling should be used to defeat graphic distortion and ambiguity. Write out explanations of the data on the graphic itself. Label important events in the data.
+> 3. Show data variation, not design variation.
+> 4. In time-series displays of money, deflated and standardized units of monetary measurement are nearly always better than nominal units.
+> 5. The number of information-carrying (variable) dimensions depicted should not exceed the number of dimensions in the data.
+> 6. Graphics must not quote data out of context.
+> – Tufte (2001, p. 77)
+
+### Creating Data Visualizations in `R` {#sec-dataVisualizationOverviewInR}
 

### 2024-08-14 data-visualization.qmd
20240814 - friendly, accessible graphics

@@
 > – Tufte (2001, p. 77)
+
+@Tufte2001 also provides recommendations for friendly, accessible graphics, including:
+
+- spell words out (rather than using abbreviations)
+- have words run from left to write (including the y-axis title)
+- include little messages to help explain the data
+- place labels on the graphic so no legend is needed
+- avoid elaborately encoded shadings, cross-hatching, and colors
+- avoid "chartjunk"—i.e., unnecessary or distracting elements (e.g., excessive decoration, overly complex graphics, graphical effects, and irrelevant information such as moiré vibration, heavy grids, and self-promoting graphs) that do not improve viewers' understanding of the data
+- if colors are used, use colors that are distinguishable by color-deficient and color-blind viewers (red–green is a common form of color-blindness)
+- use type (i.e., of the text) that is clear, precise, and modest
+- use text that is upper-and-lower case, not all capitals
 

### 2024-08-14 getting-started.qmd
20240812 - typo
20240814 - placeholder for piping
20240814 - piping

@@
 1. After creating the new repository, make sure you are on the page of of your new repository and complete the following steps:
-    1. Click "Settings" (in the topof the screen)
+    1. Click "Settings" (in the top of the screen)
     1. Click "Actions" (in the left sidebar) > "General"
@@
 players$position == "WR" | players$age > 23
+```
+
+## Piping {#sec-piping}
+
+In base `R`, if you want to perform multiple operations, it is common to either a) nest the operations, or b) save the object at each step.
+
+Below is an example of nested operations:
+
+```{r}
+length(names(nfl_players))
+```
+
+Below is an example of saving the intermediate object at each step:
+
+```{r}
+variableNames <- names(nfl_players)
+variableNames
+
+lengthOfVariableNames <- length(variableNames)
+lengthOfVariableNames
+```
+
+Code for performing nested operations can be challenging to read.
+Saving the intermediate object can be a waste of time to do if you are not interested in the intermediate object, and can take up unnecessary memory and computational resources.
+An alternative approach is to use piping.
+Piping allows taking the result from one computation and sending it to the next computation, thus allowing a chain of computations without saving the intermediate object at each step.
+
+In base `R`, you can perform piping with the `|>` expression.
+In `tidyverse` you can perform piping with the `%>%` expression.
+
+#### Base `R` {#sec-pipingBaseR}
+
+```{r}
+nfl_players |>
+  names() |>
+  length()
+```
+
+#### Tidyverse {#sec-pipingTidyverse}
+
+```{r}
+nfl_players %>%
+  names() %>%
+  length()
 ```

### 2024-08-15 machine-learning.qmd
Built site for gh-pages

@@
 ```{r}
+library("petersenlab")
+library("parallel")
 library("doParallel")
@@
 library("caret")
+library("gpboost")
 library("tidyverse")
@@
 Thus, when estimating machine learning models, it is common to keep a hold-out sample for use in cross-validation to evaluate the extent of shrinkage of model coefficients.
+The data that the model is trained on is known as the "training data".
+The data that the model was not trained on but is then is independently tested on—i.e., the hold-out sample—is the "test data".
 Shrinkage occurs when predictor variables explain some random error variance in the original model.
-When the model is applied to an independent sample, the predictive model will likely not perform quite as well, and the regressions coefficients will tend to get smaller (i.e., shrink).
+When the model is applied to an independent sample (i.e., the test data), the predictive model will likely not perform quite as well, and the regressions coefficients will tend to get smaller (i.e., shrink).
+
+If the test data were collected as part of the same processes as the original data and were merely held out for purposes of analysis, this is called internal cross-validation.
+If the test data were collected separately from the original data used to train the model, this is called external cross-validation.
+
+Most machine learning methods were developed with cross-sectional data in mind.
+That is, they assume that each person has only one observation on the outcome variable.
+However, with longitudinal data, each person has multiple observations on the outcome variable.
+
+When performing machine learning, there may be multiple approaches you can take to address this:
+
+- transforming data from [long to wide](#sec-longToWide) form, so that each person has only one row
+- when designing the training and test sets, keep all measurements from the same person in the same data object (either the training or test set); do not have some measurements from a given person in the training set and other measurements from the same person in the test set
+- use a machine learning approach that accounts for the clustered/nested nature of the data
 
@@
 
+### Additional Processing {#sec-mlAdditionalProcessing}
+
+```{r}
+# Convert character and logical variables to factors
+seasonalData <- seasonalData %>% 
+  mutate(
+    across(
+      where(is.character),
+      as.factor
+    ),
+    across(
+      where(is.logical),
+      as.factor
+    )
+  )
+```
+
 ### Fill in Missing Data {#sec-fillMissingData}
@@
 
-### Create New Data Object {#sec-newDataObject}
-
-```{r}
-newData_seasonal <- seasonalData_lag %>% 
-  filter(season == max(season, na.rm = TRUE))
-
-# Drop rows with missing fantasyPoints_lag (i.e., the player's last season), so we don't impute criterion (fantasy points)
-seasonalData_lag <- seasonalData_lag %>% 
-  filter(!is.na(fantasyPoints_lag))
-```
-
 ### Subset to Predictor Variables and Outcome Variable {#sec-subsetToPredictorsAndOutcome}
@@
 
-seasonalData_lag_subset <- seasonalData_lag %>% 
-  dplyr::select(-any_of(c(
+dropVars <- c(
   "birth_date", "loaded", "full_name", "player_name", "player_display_name", "display_name", "suffix", "headshot_url", "player", "pos",
@@
   "status_short_description", "short_name", "headshot", "uniform_number", "jersey_number", "first_name", "last_name",
-    "football_name", "team"
-  )))
-```
+  "football_name", "team")
 
-### Convert Character Predictors to Factors {#sec-convertCharacterToFactor}
-
-```{r}
-seasonalData_lag_subset <- seasonalData_lag_subset %>% mutate(across(where(is.character), as.factor))
-seasonalData_lag_subset <- seasonalData_lag_subset %>% mutate(across(where(is.logical), as.factor))
-
-seasonalData_lag_subset %>% select_if(~class(.) == "Date") %>% str()
-seasonalData_lag_subset %>% select_if(is.character) %>% str()
-seasonalData_lag_subset %>% select_if(is.factor) %>% str()
-seasonalData_lag_subset %>% select_if(is.logical) %>% str()
+seasonalData_lag_subset <- seasonalData_lag %>% 
+  dplyr::select(-any_of(dropVars))
 ```
@@
 ```{r}
-# Training Data
 seasonalData_lag_subsetQB <- seasonalData_lag_subset %>% 
@@
     carries:special_teams_tds, contains(".rush"), contains(".rec"))
-
-# Out-of-Sample Prediction
-newData_seasonalQB <- newData_seasonal %>% 
-  filter(position == "QB") %>% 
-  select(
-    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
-    height, weight, rookie_year, draft_number,
-    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
-    completions:rushing_2pt_conversions, contains(".pass"))
-
-newData_seasonalRB <- newData_seasonal %>% 
-  filter(position == "RB") %>% 
-  select(
-    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
-    height, weight, rookie_year, draft_number,
-    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
-    carries:special_teams_tds, contains(".rush"), contains(".rec"))
-
-newData_seasonalWR <- newData_seasonal %>% 
-  filter(position == "WR") %>% 
-  select(
-    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
-    height, weight, rookie_year, draft_number,
-    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
-    carries:special_teams_tds, contains(".rush"), contains(".rec"))
-
-newData_seasonalTE <- newData_seasonal %>% 
-  filter(position == "TE") %>% 
-  select(
-    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
-    height, weight, rookie_year, draft_number,
-    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
-    carries:special_teams_tds, contains(".rush"), contains(".rec"))
 ```
@@
 ### Split into Test and Training Data {#sec-splitTestTraining}
+
+CURRENTLY (WILL CHANGE):
+
+- seasonalData: 1999-2023
+- seasonalData_lag: 1999-2022 (predicting fantasy points in 2023)
+- newData_seasonal: 2023 (to be used for predicting fantasy points in 2024)
+
+to create:
+
+- seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
+- seasonalData_lag_train: 1999-2022 (predicting fantasy points in 2023), most players
+- seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), some retired players
+- (eventually, newData_seasonal, which is derived from the imputed version of seasonalData_lag_all, and then keeps only data from 2023, thus predicting fantasy points in 2024, and removing the fantasy_points_lag variable)
+
+To impute:
+
+- seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
+- seasonalData_lag_train: 1999-2022 (predicting fantasy points in 2023), most players
+- seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), some retired players
+
+IMPUTATION AND MODEL PROCESS:
+
+IMPUTATION:
+- training data
+- test data
+- all data (training and test data), used for generating next year predictions
+
+MODEL:
+- training data (imputed version of seasonalData_lag_all): all seasons (except 2023 predicting 2024), most players
+
+EVALUATE MODEL PREDICTIONS:
+- test data (imputed version of seasonalData_lag_test): all seasons (except 2023 predicting 2024), some retired players
+
+GENERATE MODEL PREDICTIONS
+- next year data (newData_seasonal): 2023 (predicting 2024)
+
+```{r}
+seasonalData_lag_qb_all <- seasonalData_lag_subsetQB
+seasonalData_lag_rb_all <- seasonalData_lag_subsetRB
+seasonalData_lag_wr_all <- seasonalData_lag_subsetWR
+seasonalData_lag_te_all <- seasonalData_lag_subsetTE
+
+set.seed(52242)
+
+activeQBs <- unique(seasonalData_lag_qb_all$gsis_id[which(seasonalData_lag_qb_all$season == max(seasonalData_lag_qb_all$season, na.rm = TRUE))])
+retiredQBs <- unique(seasonalData_lag_qb_all$gsis_id[which(seasonalData_lag_qb_all$gsis_id %ni% activeQBs)])
+numQBs <- length(unique(seasonalData_lag_qb_all$gsis_id))
+qbHoldoutIDs <- sample(retiredQBs, size = ceiling(.2 * numQBs)) # holdout 20% of players
+
+activeRBs <- unique(seasonalData_lag_rb_all$gsis_id[which(seasonalData_lag_rb_all$season == max(seasonalData_lag_rb_all$season, na.rm = TRUE))])
+retiredRBs <- unique(seasonalData_lag_rb_all$gsis_id[which(seasonalData_lag_rb_all$gsis_id %ni% activeRBs)])
+numRBs <- length(unique(seasonalData_lag_rb_all$gsis_id))
+rbHoldoutIDs <- sample(retiredRBs, size = ceiling(.2 * numRBs)) # holdout 20% of players
+
+activeWRs <- unique(seasonalData_lag_wr_all$gsis_id[which(seasonalData_lag_wr_all$season == max(seasonalData_lag_wr_all$season, na.rm = TRUE))])
+retiredWRs <- unique(seasonalData_lag_wr_all$gsis_id[which(seasonalData_lag_wr_all$gsis_id %ni% activeWRs)])
+numWRs <- length(unique(seasonalData_lag_wr_all$gsis_id))
+wrHoldoutIDs <- sample(retiredWRs, size = ceiling(.2 * numWRs)) # holdout 20% of players
+
+activeTEs <- unique(seasonalData_lag_te_all$gsis_id[which(seasonalData_lag_te_all$season == max(seasonalData_lag_te_all$season, na.rm = TRUE))])
+retiredTEs <- unique(seasonalData_lag_te_all$gsis_id[which(seasonalData_lag_te_all$gsis_id %ni% activeTEs)])
+numTEs <- length(unique(seasonalData_lag_te_all$gsis_id))
+teHoldoutIDs <- sample(retiredTEs, size = ceiling(.2 * numTEs)) # holdout 20% of players
+
+seasonalData_lag_qb_train <- seasonalData_lag_qb_all %>% 
+  filter(gsis_id %ni% qbHoldoutIDs)
+seasonalData_lag_qb_test <- seasonalData_lag_qb_all %>% 
+  filter(gsis_id %in% qbHoldoutIDs)
+
+seasonalData_lag_rb_train <- seasonalData_lag_rb_all %>% 
+  filter(gsis_id %ni% rbHoldoutIDs)
+seasonalData_lag_rb_test <- seasonalData_lag_rb_all %>% 
+  filter(gsis_id %in% rbHoldoutIDs)
+
+seasonalData_lag_wr_train <- seasonalData_lag_wr_all %>% 
+  filter(gsis_id %ni% wrHoldoutIDs)
+seasonalData_lag_wr_test <- seasonalData_lag_wr_all %>% 
+  filter(gsis_id %in% wrHoldoutIDs)
+
+seasonalData_lag_te_train <- seasonalData_lag_te_all %>% 
+  filter(gsis_id %ni% teHoldoutIDs)
+seasonalData_lag_te_test <- seasonalData_lag_te_all %>% 
+  filter(gsis_id %in% teHoldoutIDs)
+```
 
@@
 
-- SEPARATE BY POSITION
+- CONSIDER USING RETIRED PLAYERS AS THE HOLDOUT SAMPLE
 - CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
-- IMPUTE THE TRAINING AND TEST DATA SEPARATELY
+- CONSIDER IMPUTING THE TRAINING AND TEST DATA SEPARATELY BY POSITION
 
 Here is a vignette demonstrating how to impute missing data using `missForest()`: <https://rpubs.com/lmorgan95/MissForest> (archived at: <https://perma.cc/6GB4-2E22>).
+Below, we impute the training data and new data separately by position.
+
+```{r}
+#| include: false
+
+num_cores <- parallel::detectCores()
+num_true_cores <- parallel::detectCores(logical = FALSE)
+```
 
@@
 num_cores <- detectCores() - 1
-```
-
-```{r}
-#| include: false
-
-num_cores <- detectCores()
+num_true_cores <- parallel::detectCores(logical = FALSE) - 1
 ```
@@
 
-cl <- makeCluster(num_cores)
-registerDoParallel(cl)
+cl <- parallel::makeCluster(num_cores)
+doParallel::registerDoParallel(cl)
 
@@
 seasonalData_lag_subsetQB_imp$OOBerror
+data_train_qb <- cbind(seasonalData_lag_subsetQB[,"gsis_id"], seasonalData_lag_subsetQB_imp$ximp)
+data_train_qb_matrix <- data_train_qb %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
 
@@
 newData_seasonalQB_imp$OOBerror
+newData_qb <- cbind(newData_seasonalQB[,"gsis_id"], newData_seasonalQB_imp$ximp)
+newData_qb_matrix <- newData_qb %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
 
@@
 ```{r}
-cl <- makeCluster(num_cores)
-registerDoParallel(cl)
+cl <- parallel::makeCluster(num_cores)
+doParallel::registerDoParallel(cl)
 
@@
 
+### Combining Tree-Boosting with Mixed Models {#sec-treeBoosting}
+
+Adapted from here:
+<https://towardsdatascience.com/mixed-effects-machine-learning-for-longitudinal-panel-data-with-gpboost-part-iii-523bb38effc>
+
+```{r}
+# Gamma distribution requires positive-only values
+data_train_qb_matrix[,"fantasyPoints_lag"][data_train_qb_matrix[,"fantasyPoints_lag"] <= 0] <- 0.01
+
+# Predictor variables
+pred_vars_qb <- data_train_qb_matrix %>% 
+  as_tibble() %>% 
+  select(-gsis_id, -fantasyPoints_lag, -ageCentered20, ageCentered20Quadratic) %>% 
+  names()
+
+#pred_vars_qb_categorical <- ""
+
+gp_model_qb <- GPModel(
+  group_data = data_train_qb_matrix[,"gsis_id"],
+  likelihood = "gamma", # "gaussian"
+  group_rand_coef_data = cbind(
+    data_train_qb_matrix[,"ageCentered20"],
+    data_train_qb_matrix[,"ageCentered20Quadratic"]),
+  ind_effect_group_rand_coef = c(1,1))
+
+gp_data_qb <- gpb.Dataset(
+  data = data_train_qb_matrix[,pred_vars_qb],
+  label = data_train_qb_matrix[,"fantasyPoints_lag"])
+
+#gpb.Dataset.set.categorical(gp_data_qb, pred_vars_qb_categorical)
+
+params_qb <- list(
+  learning_rate = 0.01,
+  max_depth = 2,
+  num_leaves = 2^10,
+  min_data_in_leaf = 10,
+  lambda_l2 = 10,
+  num_threads = num_cores)
+  #optimizer_cov = "nelder_mead")
+
+nrounds_qb <- 379
+
+gp_model_qb$set_optim_params(params = list(optimizer_cov = "nelder_mead"))
+summary(gp_model_qb) # Estimated random effects model
+
+gp_model_fit_qb <- gpb.train(
+  data = gp_data_qb,
+  gp_model = gp_model_qb,
+  nrounds = nrounds_qb,
+  params = params_qb,
+  verbose = 0)
+
+# Model predictions
+pred_qb <- predict(
+  gp_model_fit_qb,
+  data = newData_qb_matrix[,pred_vars_qb],
+  group_data_pred = newData_qb_matrix[,"gsis_id"],
+  group_rand_coef_data_pred = cbind(
+    newData_qb_matrix[,"ageCentered20"],
+    newData_qb_matrix[,"ageCentered20Quadratic"]),
+  predict_var = FALSE,
+  pred_latent = FALSE)
+
+y_pred_qb <- pred_qb[["response_mean"]]
+mean((newData_qb[,"fantasyPoints_lag"] - y_pred_qb)^2)
+```
+
 ## Conclusion {#sec-machineLearningConclusion}

### 2024-08-15 cluster-analysis.qmd
20240814 - merge_name

@@
 library("petersenlab")
+library("nflreadr")
 library("mclust")
@@
   weightedAverageVars, recomputeVars,
-  "nameMerge", "loaded.pass", "loaded.rush", "loaded.rec", "loaded.def")]
+  "merge_name", "loaded.pass", "loaded.rush", "loaded.rec", "loaded.def")]
 
 nfl_advancedStatsPFR_career <- nfl_advancedStatsPFR_seasonal %>% 
-  group_by(pfr_id, nameMerge) %>% 
+  group_by(pfr_id, merge_name) %>% 
   summarise(
@@
 
-uniqueCases <- nfl_advancedStatsPFR_seasonal %>% select(pfr_id, nameMerge, gsis_id) %>% unique()
+uniqueCases <- nfl_advancedStatsPFR_seasonal %>% select(pfr_id, merge_name, gsis_id) %>% unique()
 
@@
 nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonal %>% 
-  filter(pfr_id != "WillMa06" | nameMerge != "MARCUSWILLIAMS" | !is.na(gsis_id))
+  filter(pfr_id != "WillMa06" | merge_name != "MARCUSWILLIAMS" | !is.na(gsis_id))
 
@@
   nfl_advancedStatsPFR_career,
-  nfl_advancedStatsPFR_seasonal %>% select(pfr_id, nameMerge, gsis_id) %>% unique(),
-  by = c("pfr_id", "nameMerge")
+  nfl_advancedStatsPFR_seasonal %>% select(pfr_id, merge_name, gsis_id) %>% unique(),
+  by = c("pfr_id", "merge_name")
 )

### 2024-08-15 download-football-data.qmd
20240814 - merge_name

@@
 ```{r}
+nfl_playerIDs$name[which(nfl_playerIDs$name == "Bennett,Michael")] <- "Michael Bennett"
+
 nfl_playerIDs <- nfl_playerIDs %>% 
   mutate(
-    nameMerge = petersenlab::cleanUpNames(name)
+    merge_name = nflreadr::clean_player_names(name, lowercase = TRUE)
   )
@@
   mutate(
-    nameMerge = petersenlab::cleanUpNames(player_name)
+    merge_name = nflreadr::clean_player_names(player_name, lowercase = TRUE)
   )
@@
   nfl_combine,
-  nfl_playerIDs %>% select(pfr_id, nameMerge, position, gsis_id),
-  by = c("pfr_id","nameMerge","pos" = "position"),
+  nfl_playerIDs %>% select(pfr_id, merge_name, position, gsis_id),
+  by = c("pfr_id","merge_name","pos" = "position"),
   na_matches = "never"
@@
   nfl_combine,
-  nfl_playerIDs %>% select(cfbref_id, nameMerge, position, gsis_id),
-  by = c("cfb_id" = "cfbref_id","nameMerge","pos" = "position"),
+  nfl_playerIDs %>% select(cfbref_id, merge_name, position, gsis_id),
+  by = c("cfb_id" = "cfbref_id","merge_name","pos" = "position"),
   na_matches = "never"
@@
   mutate(
-    nameMerge = petersenlab::cleanUpNames(player)
+    merge_name = nflreadr::clean_player_names(player, lowercase = TRUE)
   ) %>% 
@@
     ~ paste0(., ".pass"),
-    -c(pfr_id, nameMerge, season, team))
+    -c(pfr_id, merge_name, season, team))
 
@@
     ~ paste0(., ".rush"),
-    -c(pfr_id, nameMerge, season, team))
+    -c(pfr_id, merge_name, season, team))
 
@@
     ~ paste0(., ".rec"),
-    -c(pfr_id, nameMerge, season, team))
+    -c(pfr_id, merge_name, season, team))
 
@@
   mutate(
-    nameMerge = petersenlab::cleanUpNames(player)
+    merge_name = nflreadr::clean_player_names(player, lowercase = TRUE)
   )  %>% 
@@
     ~ paste0(., ".def"),
-    -c(pfr_id, nameMerge, season, team))
+    -c(pfr_id, merge_name, season, team))
 
@@
   mutate(
-    nameMerge = petersenlab::cleanUpNames(pfr_player_name)
+    merge_name = nflreadr::clean_player_names(pfr_player_name, lowercase = TRUE)
   ) %>% 
@@
     full_join,
-    by = c("pfr_id","nameMerge","season","team"))
+    by = c("pfr_id","merge_name","season","team"))
 
@@
     full_join,
-    by = c("game_id","pfr_player_id")) #nameMerge
+    by = c("game_id","pfr_player_id")) #merge_name
 
@@
 #    full_join,
-#    by = c("pfr_player_id","nameMerge","season","week"))
+#    by = c("pfr_player_id","merge_name","season","week"))
 
@@
     ),
-    nameMerge = coalesce(
-      nameMerge.pass,
-      nameMerge.rush,
-      nameMerge.rec,
-      nameMerge.def
+    merge_name = coalesce(
+      merge_name.pass,
+      merge_name.rush,
+      merge_name.rec,
+      merge_name.def
     ),
@@
     starts_with("team."),
-    starts_with("nameMerge."),
+    starts_with("merge_name."),
     starts_with("pfr_game_id."),
@@
   weightedAverageVars, recomputeVars,
-  "nameMerge", "loaded.pass", "loaded.rush", "loaded.rec", "loaded.def")]
+  "merge_name", "loaded.pass", "loaded.rush", "loaded.rec", "loaded.def")]
 
 nfl_advancedStatsPFR_seasonal <- nfl_advancedStatsPFR_seasonalByTeam %>% 
-  group_by(pfr_id, nameMerge, season) %>% 
+  group_by(pfr_id, merge_name, season) %>% 
   summarise(

### 2024-08-15 mythbusters.qmd
20240814 - merge_name

@@
 library("petersenlab")
+library("nflreadr")
 library("lme4")
@@
   dplyr::select(player, position, team, contractYear) %>% 
-  dplyr::mutate(nameMerge = petersenlab::cleanUpNames(player)) %>% 
+  dplyr::mutate(merge_name = nflreadr::clean_player_names(player, lowercase = TRUE)) %>% 
   dplyr::rename(season = contractYear) %>% 
@@
 player_stats_weekly_offense <- player_stats_weekly_offense %>% 
-  dplyr::mutate(nameMerge = petersenlab::cleanUpNames(player_display_name))
+  dplyr::mutate(merge_name = nflreadr::clean_player_names(player_display_name, lowercase = TRUE))
 #nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal %>% 
-#  mutate(nameMerge = petersenlab::cleanUpNames(player_display_name, ))
+#  mutate(merge_name = nflreadr::clean_player_names(player_display_name, lowercase = TRUE))
 
@@
   nfl_playerContractYears,
-  by = c("nameMerge", "position_group" = "position", "season")
+  by = c("merge_name", "position_group" = "position", "season")
 ) %>% 
@@
 #  nfl_playerContractYears,
-#  by = c("nameMerge", "position_group" = "position", "season")
+#  by = c("merge_name", "position_group" = "position", "season")
 #) %>% 
@@
 player_statsContracts_offense_weekly <- player_statsContracts_offense_weekly %>% 
-  dplyr::arrange(nameMerge, season, season_type, week)
+  dplyr::arrange(merge_name, season, season_type, week)
 
 #player_statsContracts_offense_seasonal <- player_statsContracts_offense_seasonal %>% 
-#  arrange(nameMerge, season)
+#  arrange(merge_name, season)
 
@@
 # Merge with QBR data
-nfl_espnQBR_weekly$nameMerge <- paste(nfl_espnQBR_weekly$name_first, nfl_espnQBR_weekly$name_last, sep = " ") %>% 
-  petersenlab::cleanUpNames(.)
+nfl_espnQBR_weekly$merge_name <- paste(nfl_espnQBR_weekly$name_first, nfl_espnQBR_weekly$name_last, sep = " ") %>% 
+  nflreadr::clean_player_names(., lowercase = TRUE)
 
@@
     nfl_espnQBR_weekly,
-    by = c("nameMerge","team","season")
+    by = c("merge_name","team","season")
   )
@@
 nfl_contractYearQBR_weekly <- nfl_contractYearQBR_weekly %>% 
-  dplyr::arrange(nameMerge, season, season_type, game_week)
+  dplyr::arrange(merge_name, season, season_type, game_week)
 
@@
   dplyr::filter(season_type == "Regular") %>% 
-  dplyr::arrange(nameMerge, season, season_type, game_week) %>% 
+  dplyr::arrange(merge_name, season, season_type, game_week) %>% 
   mutate(
@@
     position = "QB") %>% 
-  group_by(nameMerge, player_id) %>% 
+  group_by(merge_name, player_id) %>% 
   fill(player, .direction = "downup")
@@
   dplyr::filter(position == "QB") %>% 
-  dplyr::select(nameMerge, season, week, age, years_of_experience) %>% 
+  dplyr::select(merge_name, season, week, age, years_of_experience) %>% 
   full_join(
     nfl_contractYearQBRsubset_weekly,
-    by = c("nameMerge","season", c("week" = "game_week"))
+    by = c("merge_name","season", c("week" = "game_week"))
   ) %>% select(player_id, season, week, player, everything()) %>% 
@@
 nfl_contractYearQBR_seasonal <- nfl_contractYearQBRsubset_weekly %>% 
-  dplyr::group_by(nameMerge, season) %>% 
+  dplyr::group_by(merge_name, season) %>% 
   dplyr::summarise(
@@
 nfl_contractYearQBR_seasonal <- nfl_contractYearQBR_seasonal %>% 
-  group_by(nameMerge) %>%
+  group_by(merge_name) %>%
   mutate(player_id = as.factor(as.character(cur_group_id())))

### 2024-08-15 actuarial.qmd
20240814 - shrinkage and cross-validation

@@
 
-## Best Actuarial Approaches to Prediction {#sec-bestActuarialApproaches}
+## Challenges of Data-Driven Approaches and How to Address {#sec-challengesDataDrivenApproaches}
 
-The best actuarial models tend to be relatively simple (parsimonious), that can account for one or several of the most important predictors and their optimal weightings, and that account for the base rate of the phenomenon.
-Even unit-weighted formulas (formulas whose [predictor variables](#sec-correlationalStudy) are equally weighted with a weight of one) can sometimes generalize better to other samples than complex weightings [@Garb2019].
-Differential weightings sometimes capture random variance and [over-fit](#sec-overfitting) the model, thus leading to predictive accuracy shrinkage in cross-validation samples [@Garb2019], as described below.
-The choice of [predictor variables](#sec-correlationalStudy) often matters more than their weighting.
+There are various challenges of data-driven approaches.
+First, they are sometimes not interpretable or consistent with theory.
+Second, they tend to [overfit](#sec-overfitting) the data.
+[Overfitting](#sec-overfitting) is described in @sec-overfitting.
+Third, as a result of [overfitting](#sec-overfitting) the data, they tend to show [shrinkage](#sec-shrinkage).
+
+### Shrinkage {#sec-shrinkage}
 
@@
 *Shrinkage* is when variables with stronger predictive power in the original data set tend to show somewhat smaller predictive power (smaller regression coefficients) when applied to new groups.
-Shrinkage reflects a model [overfitting](#sec-overfitting) (i.e., fitting to error by capitalizing on chance).
+Shrinkage reflects a model [overfitting](#sec-overfitting)—i.e., when the model explains error variance by capitalizing on chance.
 Shrinkage is especially likely when the original sample is small and/or unrepresentative and the number of variables considered for inclusion is large.
+To help minimize the extent of shrinkage, it is recommended to apply [cross-validation](#sec-crossValidation).
+
+### Cross-Validation {#sec-crossValidation}
+
 Cross-validation with large, representative samples can help evaluate the amount of shrinkage of estimates, particularly for more complex models such as [machine learning](#sec-machineLearning) models [@Ursenbach2019].
@@
 
+## Best Actuarial Approaches to Prediction {#sec-bestActuarialApproaches}
+
+The best actuarial models tend to be relatively simple (parsimonious), that can account for one or several of the most important predictors and their optimal weightings, and that account for the base rate of the phenomenon.
+Even unit-weighted formulas (formulas whose [predictor variables](#sec-correlationalStudy) are equally weighted with a weight of one) can sometimes generalize better to other samples than complex weightings [@Garb2019].
+Differential weightings sometimes capture random variance and [over-fit](#sec-overfitting) the model, thus leading to predictive accuracy shrinkage in cross-validation samples [@Garb2019], as described below.
+The choice of [predictor variables](#sec-correlationalStudy) often matters more than their weighting.
+
 An emerging technique that holds promise for increasing predictive accuracy of actuarial methods is [machine learning](#sec-machineLearning) [@Garb2019].
 However, one challenge of some [machine learning](#sec-machineLearning) techniques is that they are like a "black box" and are not transparent, which raises ethical concerns [@Garb2019].
-[machine learning](#sec-machineLearning) may be most valuable when the data available are complex and there are many [predictor variables](#sec-correlationalStudy) [@Garb2019].
+Moreover, [machine learning](#sec-machineLearning) also tends to lead to [overfitting](#sec-overfitting) and [shrinkage](#sec-shrinkage).
+[machine learning](#sec-machineLearning) may be most valuable when the data available are complex and there are many [predictor variables](#sec-correlationalStudy) [@Garb2019], and when the model is validated with [cross-validation](#sec-crossValidation).
 

### 2024-08-15 machine-learning.qmd
Built site for gh-pages

@@
 
+Machine learning takes us away from focusing on [causal inference](#sec-causalInference).
+Machine learning does not care about which processes are causal—i.e., which processes influence the outcome.
+Instead, machine learning cares about prediction—it cares about a predictor variable to the extent that it increases predictive accuracy regardless of whether it is causally related to the outcome.
+
 Machine learning can be useful for leveraging big data and lots of predictor variable to develop predictive models with greater accuracy.
-However, many machine learning techniques are black boxes—it is often unclear
+However, many machine learning techniques are black boxes—it is often unclear how or why certain predictions are made, which can make it difficult to interpret the model's decisions and understand the underlying relationships between variables.
 Machine learning tends to be a data-driven, atheoretical technique.
@@
 
-### Fill in Missing Data {#sec-fillMissingData}
+### Fill in Missing Data for Static Variables {#sec-fillMissingData}
 
@@
     fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
-    completions:rushing_2pt_conversions, contains(".pass"))
+    completions:rushing_2pt_conversions, special_teams_tds, contains(".pass"), contains(".rush"))
 
@@
 
+::: {#nte-machineLearningImputeMissingData .callout-note title="Impute missing data for machine learning"}
+Note: the following code takes a while to run.
+:::
+
 ```{r}
@@
 # QBs
-seasonalData_lag_subsetQB_imp <- missForest::missForest(
-  seasonalData_lag_subsetQB %>% 
+seasonalData_lag_qb_all_imp <- missForest::missForest(
+  seasonalData_lag_qb_all %>% 
     select(-gsis_id) %>%
@@
 
-seasonalData_lag_subsetQB_imp$OOBerror
-data_train_qb <- cbind(seasonalData_lag_subsetQB[,"gsis_id"], seasonalData_lag_subsetQB_imp$ximp)
-data_train_qb_matrix <- data_train_qb %>%
+seasonalData_lag_qb_all_imp$OOBerror
+data_all_qb <- cbind(seasonalData_lag_qb_all[,"gsis_id"], seasonalData_lag_qb_all_imp$ximp)
+data_all_qb_matrix <- data_all_qb %>%
   mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
   as.matrix()
+newData_qb <- data_all_qb %>% 
+  filter(season == max(season, na.rm = TRUE)) %>% 
+  select(-fantasyPoints_lag)
+newData_qb_matrix <- data_all_qb_matrix[
+  data_all_qb_matrix[, "season"] == max(data_all_qb_matrix[, "season"], na.rm = TRUE), # keep only rows with the most recent season
+  , # all columns
+  drop = FALSE]
 
-newData_seasonalQB_imp <- missForest::missForest(
-  newData_seasonalQB %>% 
+dropCol_qb <- which(colnames(newData_qb_matrix) == "fantasyPoints_lag")
+newData_qb_matrix <- newData_qb_matrix[, -dropCol_qb, drop = FALSE]
+
+seasonalData_lag_qb_train_imp <- missForest::missForest(
+  seasonalData_lag_qb_train %>% 
     select(-gsis_id) %>%
@@
 
-newData_seasonalQB_imp$OOBerror
-newData_qb <- cbind(newData_seasonalQB[,"gsis_id"], newData_seasonalQB_imp$ximp)
-newData_qb_matrix <- newData_qb %>%
+seasonalData_lag_qb_train_imp$OOBerror
+data_train_qb <- cbind(seasonalData_lag_qb_train[,"gsis_id"], seasonalData_lag_qb_train_imp$ximp)
+data_train_qb_matrix <- data_train_qb %>%
   mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
@@
 
-# RBs
-seasonalData_lag_subsetRB_imp <- missForest::missForest(
-  seasonalData_lag_subsetRB %>% 
+seasonalData_lag_qb_test_imp <- missForest::missForest(
+  seasonalData_lag_qb_test %>% 
     select(-gsis_id) %>%
@@
 
-seasonalData_lag_subsetRB_imp$OOBerror
+seasonalData_lag_qb_test_imp$OOBerror
+data_test_qb <- cbind(seasonalData_lag_qb_test[,"gsis_id"], seasonalData_lag_qb_test_imp$ximp)
+data_test_qb_matrix <- data_test_qb %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
 
-newData_seasonalRB_imp <- missForest::missForest(
-  newData_seasonalRB %>% 
+# RBs
+seasonalData_lag_rb_all_imp <- missForest::missForest(
+  seasonalData_lag_rb_all %>% 
     select(-gsis_id) %>%
@@
 
-newData_seasonalRB_imp$OOBerror
+seasonalData_lag_rb_all_imp$OOBerror
+data_all_rb <- cbind(seasonalData_lag_rb_all[,"gsis_id"], seasonalData_lag_rb_all_imp$ximp)
+data_all_rb_matrix <- data_all_rb %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
+newData_rb <- data_all_rb %>% 
+  filter(season == max(season, na.rm = TRUE)) %>% 
+  select(-fantasyPoints_lag)
+newData_rb_matrix <- data_all_rb_matrix[
+  data_all_rb_matrix[, "season"] == max(data_all_rb_matrix[, "season"], na.rm = TRUE), # keep only rows with the most recent season
+  , # all columns
+  drop = FALSE]
 
-# WRs
-seasonalData_lag_subsetWR_imp <- missForest::missForest(
-  seasonalData_lag_subsetWR %>% 
+dropCol_rb <- which(colnames(newData_rb_matrix) == "fantasyPoints_lag")
+newData_rb_matrix <- newData_rb_matrix[, -dropCol_rb, drop = FALSE]
+
+seasonalData_lag_rb_train_imp <- missForest::missForest(
+  seasonalData_lag_rb_train %>% 
     select(-gsis_id) %>%
@@
 
-seasonalData_lag_subsetWR_imp$OOBerror
+seasonalData_lag_rb_train_imp$OOBerror
+data_train_rb <- cbind(seasonalData_lag_rb_train[,"gsis_id"], seasonalData_lag_rb_train_imp$ximp)
+data_train_rb_matrix <- data_train_rb %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
 
-newData_seasonalWR_imp <- missForest::missForest(
-  newData_seasonalWR %>% 
+seasonalData_lag_rb_test_imp <- missForest::missForest(
+  seasonalData_lag_rb_test %>% 
     select(-gsis_id) %>%
@@
 
-newData_seasonalWR_imp$OOBerror
+seasonalData_lag_rb_test_imp$OOBerror
+data_test_rb <- cbind(seasonalData_lag_rb_test[,"gsis_id"], seasonalData_lag_rb_test_imp$ximp)
+data_test_rb_matrix <- data_test_rb %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
 
-# TEs
-seasonalData_lag_subsetTE_imp <- missForest::missForest(
-  seasonalData_lag_subsetTE %>% 
+# WRs
+seasonalData_lag_wr_all_imp <- missForest::missForest(
+  seasonalData_lag_wr_all %>% 
     select(-gsis_id) %>%
@@
 
-seasonalData_lag_subsetTE_imp$OOBerror
+seasonalData_lag_wr_all_imp$OOBerror
+data_all_wr <- cbind(seasonalData_lag_wr_all[,"gsis_id"], seasonalData_lag_wr_all_imp$ximp)
+data_all_wr_matrix <- data_all_wr %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
+newData_wr <- data_all_wr %>% 
+  filter(season == max(season, na.rm = TRUE)) %>% 
+  select(-fantasyPoints_lag)
+newData_wr_matrix <- data_all_wr_matrix[
+  data_all_wr_matrix[, "season"] == max(data_all_wr_matrix[, "season"], na.rm = TRUE), # keep only rows with the most recent season
+  , # all columns
+  drop = FALSE]
 
-newData_seasonalTE_imp <- missForest::missForest(
-  newData_seasonalTE %>% 
+dropCol_wr <- which(colnames(newData_wr_matrix) == "fantasyPoints_lag")
+newData_wr_matrix <- newData_wr_matrix[, -dropCol_wr, drop = FALSE]
+
+seasonalData_lag_wr_train_imp <- missForest::missForest(
+  seasonalData_lag_wr_train %>% 
     select(-gsis_id) %>%
@@
 
-newData_seasonalTE_imp$OOBerror
+seasonalData_lag_wr_train_imp$OOBerror
+data_train_wr <- cbind(seasonalData_lag_wr_train[,"gsis_id"], seasonalData_lag_wr_train_imp$ximp)
+data_train_wr_matrix <- data_train_wr %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
+
+seasonalData_lag_wr_test_imp <- missForest::missForest(
+  seasonalData_lag_wr_test %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
+
+seasonalData_lag_wr_test_imp$OOBerror
+data_test_wr <- cbind(seasonalData_lag_wr_test[,"gsis_id"], seasonalData_lag_wr_test_imp$ximp)
+data_test_wr_matrix <- data_test_wr %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
+
+# TEs
+seasonalData_lag_te_all_imp <- missForest::missForest(
+  seasonalData_lag_te_all %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
+
+seasonalData_lag_te_all_imp$OOBerror
+data_all_te <- cbind(seasonalData_lag_te_all[,"gsis_id"], seasonalData_lag_te_all_imp$ximp)
+data_all_te_matrix <- data_all_te %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
+newData_te <- data_all_te %>% 
+  filter(season == max(season, na.rm = TRUE)) %>% 
+  select(-fantasyPoints_lag)
+newData_te_matrix <- data_all_te_matrix[
+  data_all_te_matrix[, "season"] == max(data_all_te_matrix[, "season"], na.rm = TRUE), # keep only rows with the most recent season
+  , # all columns
+  drop = FALSE]
+
+dropCol_te <- which(colnames(newData_te_matrix) == "fantasyPoints_lag")
+newData_te_matrix <- newData_te_matrix[, -dropCol_te, drop = FALSE]
+
+seasonalData_lag_te_train_imp <- missForest::missForest(
+  seasonalData_lag_te_train %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
+
+seasonalData_lag_te_train_imp$OOBerror
+data_train_te <- cbind(seasonalData_lag_te_train[,"gsis_id"], seasonalData_lag_te_train_imp$ximp)
+data_train_te_matrix <- data_train_te %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
+
+seasonalData_lag_te_test_imp <- missForest::missForest(
+  seasonalData_lag_te_test %>% 
+    select(-gsis_id) %>%
+    data.frame(),
+  verbose = TRUE,
+  mtry = 100,
+  parallelize = "forests")
+
+seasonalData_lag_te_test_imp$OOBerror
+data_test_te <- cbind(seasonalData_lag_te_test[,"gsis_id"], seasonalData_lag_te_test_imp$ximp)
+data_test_te_matrix <- data_test_te %>%
+  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
+  as.matrix()
 
@@
 
-#pred_vars_qb_categorical <- ""
+#pred_vars_qb_categorical <- "" # to specify categorical predictors
 
@@
 
-#gpb.Dataset.set.categorical(gp_data_qb, pred_vars_qb_categorical)
+#gpb.Dataset.set.categorical(gp_data_qb, pred_vars_qb_categorical) # to specify categorical predictors
 
@@
   num_threads = num_cores)
-  #optimizer_cov = "nelder_mead")
 
@@
 gp_model_qb$set_optim_params(params = list(optimizer_cov = "nelder_mead"))
-summary(gp_model_qb) # Estimated random effects model
 
@@
   nrounds = nrounds_qb,
-  params = params_qb,
-  verbose = 0)
+  params = params_qb) # verbose = 0
 
-# Model predictions
-pred_qb <- predict(
+summary(gp_model_qb) # Estimated random effects model
+
+# Test Model on Test Data
+pred_test_qb <- predict(
+  gp_model_fit_qb,
+  data = data_test_qb_matrix[,pred_vars_qb],
+  group_data_pred = data_test_qb_matrix[,"gsis_id"],
+  group_rand_coef_data_pred = cbind(
+    data_test_qb_matrix[,"ageCentered20"],
+    data_test_qb_matrix[,"ageCentered20Quadratic"]),
+  predict_var = FALSE,
+  pred_latent = FALSE)
+
+y_pred_test_qb <- pred_test_qb[["response_mean"]]
+mean((data_test_qb_matrix[,"fantasyPoints_lag"] - y_pred_test_qb)^2)
+
+# Generate model predictions for next season
+pred_nextYear_qb <- predict(
   gp_model_fit_qb,
@@
 
-y_pred_qb <- pred_qb[["response_mean"]]
-mean((newData_qb[,"fantasyPoints_lag"] - y_pred_qb)^2)
+newData_qb$fantasyPoints_lag <- pred_nextYear_qb$response_mean
+
+# Merge with player names
+newData_qb <- left_join(
+  newData_qb,
+  nfl_playerIDs %>% select(gsis_id, name),
+  by = "gsis_id"
+)
+
+newData_qb %>% 
+  arrange(-fantasyPoints_lag) %>% 
+  select(name, fantasyPoints_lag)
 ```

### 2024-08-15 machine-learning.qmd
20240815 - tree-boosting QBs

@@
 
+### Specify Options {#machineLearningSpecifyOptions}
+
+```{r}
+options(scipen = 999) # prevent scientific notation
+```
+
 ## Overview of Machine Learning {#sec-machineLearningOverview}
@@
 
-```{r}
-# Gamma distribution requires positive-only values
-data_train_qb_matrix[,"fantasyPoints_lag"][data_train_qb_matrix[,"fantasyPoints_lag"] <= 0] <- 0.01
-
-# Predictor variables
+#### Process Data {#sec-treeBoostingProcessData}
+
+If using a gamma distribution, it requires positive-only values:
+
+```{r}
+#data_train_qb_matrix[,"fantasyPoints_lag"][data_train_qb_matrix[,"fantasyPoints_lag"] <= 0] <- 0.01
+```
+
+#### Specify Predictor Variables {#sec-treeBoostingSpecifyPredictors}
+
+```{r}
 pred_vars_qb <- data_train_qb_matrix %>% 
   as_tibble() %>% 
-  select(-gsis_id, -fantasyPoints_lag, -ageCentered20, ageCentered20Quadratic) %>% 
+  select(-fantasyPoints_lag, -ageCentered20, ageCentered20Quadratic) %>% # -gsis_id
   names()
 
-#pred_vars_qb_categorical <- "" # to specify categorical predictors
+pred_vars_qb_categorical <- "gsis_id" # to specify categorical predictors
+```
+
+#### Specify General Model Options {#sec-treeBoostingSpecifyGeneralModelOptions}
+
+```{r}
+model_likelihood <- "gaussian" # gamma
+nrounds <- 1000
+```
+
+#### Identify Optimal Tuning Parameters {#sec-treeBoostingOptimalTuningParameters}
+
+```{r}
+# Partition training data into inner training data and validation data
+ntrain_qb <- dim(data_train_qb_matrix)[1]
+
+set.seed(52242)
+valid_tune_idx_qb <- sample.int(ntrain_qb, as.integer(0.2*ntrain_qb))
+
+folds_qb <- list(valid_tune_idx_qb)
+
+# Specify parameter grid, gp_model, and gpb.Dataset
+param_grid_qb <- list(
+  "learning_rate" = c(1,0.1,0.01),
+  "max_depth" = c(1,2,3,5,10),
+  "min_data_in_leaf" = c(10,100,1000),
+  "lambda_l2" = c(0,1,10))
+
+other_params_qb <- list(num_leaves = 2^10)
 
@@
   group_data = data_train_qb_matrix[,"gsis_id"],
-  likelihood = "gamma", # "gaussian"
+  likelihood = model_likelihood,
   group_rand_coef_data = cbind(
@@
   data = data_train_qb_matrix[,pred_vars_qb],
+  categorical_feature = pred_vars_qb_categorical,
   label = data_train_qb_matrix[,"fantasyPoints_lag"])
 
-#gpb.Dataset.set.categorical(gp_data_qb, pred_vars_qb_categorical) # to specify categorical predictors
+# Find optimal tuning parameters
+opt_params_qb <- gpb.grid.search.tune.parameters(
+  param_grid = param_grid_qb,
+  params = other_params_qb,
+  num_try_random = NULL,
+  folds = folds_qb,
+  data = gp_data_qb,
+  gp_model = gp_model_qb,
+  nrounds = nrounds,
+  early_stopping_rounds = 10,
+  verbose_eval = 1,
+  metric = "mse")
+
+opt_params_qb
+```
+
+#### Specify Model and Tuning Parameters {#sec-treeBoostingSpecifyModel}
+
+```{r}
+gp_model_qb <- GPModel(
+  group_data = data_train_qb_matrix[,"gsis_id"],
+  likelihood = model_likelihood,
+  group_rand_coef_data = cbind(
+    data_train_qb_matrix[,"ageCentered20"],
+    data_train_qb_matrix[,"ageCentered20Quadratic"]),
+  ind_effect_group_rand_coef = c(1,1))
+
+gp_data_qb <- gpb.Dataset(
+  data = data_train_qb_matrix[,pred_vars_qb],
+  categorical_feature = pred_vars_qb_categorical,
+  label = data_train_qb_matrix[,"fantasyPoints_lag"])
 
 params_qb <- list(
-  learning_rate = 0.01,
-  max_depth = 2,
+  learning_rate = 1,
+  max_depth = 10,
   num_leaves = 2^10,
   min_data_in_leaf = 10,
-  lambda_l2 = 10,
+  lambda_l2 = 0,
   num_threads = num_cores)
 
-nrounds_qb <- 379
-
-gp_model_qb$set_optim_params(params = list(optimizer_cov = "nelder_mead"))
-
+#gp_model_qb$set_optim_params(params = list(optimizer_cov = "nelder_mead")) # to speed up model estimation
+```
+
+#### Fit Model {#sec-treeBoostingFitModel}
+
+```{r}
 gp_model_fit_qb <- gpb.train(
@@
   gp_model = gp_model_qb,
-  nrounds = nrounds_qb,
+  nrounds = nrounds,
   params = params_qb) # verbose = 0
@@
 summary(gp_model_qb) # Estimated random effects model
-
+```
+
+#### Evaluate Accuracy of Model on Test Data {#sec-treeBoostingModelAccuracy}
+
+```{r}
 # Test Model on Test Data
@@
 y_pred_test_qb <- pred_test_qb[["response_mean"]]
+cbind(y_pred_test_qb, data_test_qb_matrix[,"fantasyPoints_lag"])
 mean((data_test_qb_matrix[,"fantasyPoints_lag"] - y_pred_test_qb)^2)
-
+```
+
+#### Generate Predictions for Next Season {#sec-treeBoostingModelPredictions}
+
+```{r}
 # Generate model predictions for next season
@@
   arrange(-fantasyPoints_lag) %>% 
-  select(name, fantasyPoints_lag)
+  select(name, fantasyPoints_lag, fantasy_points)
 ```

### 2024-08-15 machine-learning.qmd
20240815 - evaluate accuracy

@@
 
+nrounds_qb <- 7000 # identify optimal number of trees through iteration and cross-validation
+
 #gp_model_qb$set_optim_params(params = list(optimizer_cov = "nelder_mead")) # to speed up model estimation
@@
   gp_model = gp_model_qb,
-  nrounds = nrounds,
+  nrounds = nrounds_qb,
   params = params_qb) # verbose = 0
@@
 cbind(y_pred_test_qb, data_test_qb_matrix[,"fantasyPoints_lag"])
-mean((data_test_qb_matrix[,"fantasyPoints_lag"] - y_pred_test_qb)^2)
+
+petersenlab::accuracyOverall(
+  predicted = y_pred_test_qb,
+  actual = data_test_qb_matrix[,"fantasyPoints_lag"],
+  dropUndefined = TRUE
+)
 ```

### 2024-08-15 cognitive-bias.qmd
20240815 - clean up citations

@@
 
-@Miller2013c describes cognitive biases in fantasy sports.
+Miller [-@Miller2013c] describes cognitive biases in fantasy sports.
 Examples of cognitive biases relevant to fantasy football that result from one or more heuristics include:
@@
 
-Consider another example adapted from @Miller2013c: two players scored the same number of points in a given week.
+Consider another example adapted from Miller [-@Miller2013c]: two players scored the same number of points in a given week.
 A Wide Receiver scored 1.6 points from 16 receiving yards on 1 target.
@@
 That is, a manager tends to value the same player more when they play for their team rather than an opponent's team.
-When trading, the endowment effect bias might lead a manager to demand more to trade away a player than to acquire the same player [@@Miller2013c].
+When trading, the endowment effect bias might lead a manager to demand more to trade away a player than to acquire the same player [@Miller2013c].
 For instance, a manager might overvalue a player they drafted in the first round, refusing to trade them even if they could get a better-performing player in return.

### 2024-08-15 mixed-models.qmd
20240815 - Simpson's paradox

@@
 We will discuss a modeling framework that goes by many terms, including mixed models, mixed-effects models, multilevel models, hierarchical linear models.
-They are sometimes called multilevel models and hiearchical linear models, whose name emphasizes the hierarchical structure of the data because the data are nonindependent.
+They are sometimes called multilevel models and hierarchical linear models, whose name emphasizes the hierarchical structure of the data because the data are nonindependent.
 When observations (i.e., data points) are collected from multiple lower-level units (e.g., people) in an upper-level unit (e.g., married couple, family, classroom, school, neighborhood, team), the data from the lower-level units are considered "nested" within the upper-level unit.
-The data from the lower-level unit are likely to be correlated, to some degree, because they come from the same upper-level unit.
+In this context, *nested data* refers to multiple observations from the same upper-level unit.
+For instance, longitudinal data are nested within the same participant.
+Students are nested within classrooms, which are nested within schools.
+Players are nested within teams.
+
+When data are nested, the data from the lower-level unit are likely to be correlated, to some degree, because they come from the same upper-level unit.
 For example, multiple players may come from the same team, and the players' performance on that team is likely interrelated because they share common experiences and influence one another.
@@
 
+As described in @sec-fallaciesEcological, the ecological fallacy is the error of drawing inferences about an individual from group-level data.
+An type of ecological fallacy is [Simpson's paradox](#sec-simpsonsParadox).
+
 ### Simpson's Paradox {#sec-simpsonsParadox}
+
+Examples of Simpson's Paradox are depicted in Figures [-@fig-simpsonParadox] and [-@fig-simpsonsParadox].
+In the example below, there is a positive association between the predictor variable and outcome variable for each group.
+However, when the groups are combined, there is a *negative* association between predictor variable and outcome variable.
+That is, the sign of an association can differ at different levels of analysis (e.g., group level versus person level).
+
+::: {#fig-simpsonsParadox}
+![](images/simpsonsParadox.png)
+
+An Example of Simpson's Paradox.
+In this example, there is a positive association between `x` and `y` within each group (i.e., red or blue), but there is a negative association between `x` and `y` when the groups are combined.
+(Figure retrieved from <https://en.wikipedia.org/wiki/File:Simpson%27s_paradox_continuous.svg>)
+:::
+
+Consider that we observe a between-person association between a predictor, for example, how much sports drink a player drinks and their performance, such that players who drink more sports drink before games tend to perform better during games.
+However, based on this association, if we draw the inference that sports drink consumption leads to better performance, this could be a faulty inference.
+It is possible that, at the *within*-person level, there is no association or even a negative association between sports drink consumption and performance.
+For helping to approximate causality, a much stronger test than relying on the between-person association would be to examine the association within the individual.
+That is, examining the association within the individual would examine: when a player consumes more sports drink, whether they perform better than when the same player consumes less sports drink.
+We describe within-person analyses to approximate causal inferences in @sec-causalInferenceWithinSubject.
+
+In short, we often need to account for the groups or "levels" within which people are nested.
+When multiple observations are nested within the same upper-level unit, they are often correlated.
+That is, some variance is attributable to the lower-level unit (e.g., the player) and some variance is attributable to the upper-level unit (e.g., the team).
+It is important to evaluate how much variance is attributable to the upper-level unit.
+If substantial variance is attributable to the upper-level unit, it is important to account for the nested structure of the data.
+Mixed models are a useful approach to account for data with a nested structure so you can avoid committing the ecological fallacy.
 

### 2024-08-16 cognitive-bias.qmd
20240816 - performance heuristic; anecdotal fallacy

@@
 As described [above](#sec-cognitiveBiasOverview), heuristics are mental shortcuts that people use to handle the overwhelming amount of information to process.
-Three important heuristics used in judgment and prediction in the face of uncertainty include [@Tversky1974]:
+Important heuristics used in judgment and prediction in the face of uncertainty include [@Tversky1974; @Critcher2014]:
 
@@
 - [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment)
+- [performance heuristic](#sec-heuristicsPerformance)
 
@@
 [Regression effects](#sec-fallaciesRegression) are discussed further in @sec-fallaciesRegression.
+
+### Performance Heuristic {#sec-heuristicsPerformance}
+
+The *performance heuristic* refers to the tendency for people to predict their improvement in a task based on their previous performance [@Critcher2014].
+Although past performance tends to be correlated with future performance, people tend to show regression to the mean, as described in @sec-fallaciesRegression.
+Thus, people who show strong performance in their first few games may in many cases show worse performance in subsequent games.
+Moreover, initial performance does not necessarily indicate the extent of one's future *improvement*.
+For instance, if a manager has a successful first season, the performance heuristic may lead them to overestimate the extent to which they will improve in their skill or performance in the future.
 
@@
 - [gambler's fallacy](#sec-fallaciesGambler)
+- [anecdotal fallacy](#sec-fallaciesAnecdotal)
 - [conditional probability fallacy](#sec-fallaciesConditionalProbability)
@@
 
+### Anecdotal Fallacy {#sec-fallaciesAnecdotal}
+
+The *anecdotal fallacy* is the error of drawing inferences based on personal experience or singular examples rather than based on stronger forms of evidence.
+Although people tend to find stories and anecdotes compelling, they are among the weakest forms of evidence, as described in @#sec-internalExternalValidityTradeoffs.
+Consider that a fantasy football commentator writes an article about how they won their league by drafting a Quarterback and Wide Receiver by drafting with their first two picks, respectively.
+The anecdotal fallacy might lead a manager who reads the article to believe that they are more likely to win their league if they follow a similar strategy, despite more compelling evidence otherwise.
+
 ### Conditional Probability Fallacy {#sec-fallaciesConditionalProbability}
@@
 
-The ecological fallacy is the error of drawing inferences about an individual from group-level data.
+The *ecological fallacy* is the error of drawing inferences about an individual from group-level data.
 For instance, a manager is interested in how age is associated with fantasy performance.

### 2024-08-16 decision-making.qmd
20240605 - getting started/load packages
20240608 - getting started
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240703 - conclusion section
20240816 - wisdom of the crowd

@@
 # Decision Making in the Context of Uncertainty {#sec-decisionMaking}
+
+## Getting Started {#sec-decisionMakingGettingStarted}
+
+### Load Packages {#sec-decisionMakingLoadPackages}
+
+```{r}
+
+```
+
+## Wisdom of the Crowd {#sec-wisdomOfCrowd}
+
+In many domains, the average of forecasters' predictions is more accurate than the accuracy of the constituent individuals.
+In some domains, the average of non-expert forecasts is more accurate than the forecasts by individual experts.
+This phenomenon is called "collective intelligence", the "wisdom of the crowd", or the "wisdom of crowds" [@Rader2017; @Simoiu2019; @Surowiecki2005; @Wagner2010].
+
+Aggregation of predictions from multiple people leverages several important features, including cognitive diversity, error cancellation, INSERT.
+
+Cognitive diversity refers to the representation of individuals with different perspectives because of their "differences in knowledge, training, experience, or thinking styles" [@Rader2017, p. 8].
+Error cancellation deals with the idea that some individuals' predictions will fall above the truth and some individuals' predictions will fall below the truth—i.e., individual judgments will "bracket" the truth [@Mannes2014].
+So, the average of the predictions will "cancel out" some of the errors so that the predictions average out to more closely approximate the truth.
+
+Averaging projections from individuals tends to yield predictions that are more accurate than the accuracy of most forecasters in the group [@Mannes2014].
+Moreover, averaged projections tends to be more accurate than consensus-based judgments from groups of people that interact and discuss, due to cognitive biases associated with the social interaction among groups, such as herding in which people align their behavior with others [@Mannes2014; @Simoiu2019].
+
+Crowd-averaged projections tend to be most accurate when:
+
+- the crowd consists of individuals who hold expertise in the domain such that they will make predictions that fall close to the truth
+- there is relatively low variability in the expertise of the individual forecasters in terms of their ability to make accurate forecasts
+- there is cognitive diversity among the forecasters
+- the projections are made independently
+- the bracketing rate (i.e., the frequency with which any two forecasters' predictions fall on opposite sides of the truth) is high
+
+However, the crowd is not accurate in all situations or domains.
+For instance, the crowd tends to be less accurate than the (prospectively identified) best forecaster when there is great variability in forecasters' expertise (in terms of the forecasters' ability to forecast accurately) and when the bracketing rate is low  [@Mannes2014].
+Some forecasters may provide terrible projections; thus, including them in an average may make the average projections substantially less accurate.
+Thus, it may be necessary to examine the average of a "select crowd", by aggregating the projections of the most consistently accurate forecasters [@Mannes2014].
+
+## Accuracy of Fantasy Football Crowd Projections {#sec-accuracyOfCrowd}
+
+Even though the crowd tends to be more accurate than individual forecasters, crowd-averaged projections (at least among experts) are not necessarily highly accurate.
+In fantasy football, [DESCRIBE THEIR ACCURACY].
+The relatively modest accuracy of the projections by so-called fantasy "experts'" and of their average of their projections could occur for a number of reasons.
+One possibility is that the level of expertise of the "expert" forecasters in terms of being able to provide accurate forecasts is not strong.
+That is, because football performance and injuries are so challenging to predict, individual forecasters' projections may not be particularly close to the truth.
+
+A second possibility is that the bracketing rate of the predictions is not particularly high [@Mannes2014].
+Even if the individual forecasters' projections are not close to the truth, if ~50% of them overestimate the truth and the other 50% of the underestimate the truth, the average will more closely approximate the truth.
+However, if most forecasters overestimate the truth, averaging the projections will not necessarily lead to more accurate projections.
+
+A third possibility is that the forecasts of the different experts are not independent.
+
+Each of these possibilities is likely true to some degree.
+First, individuals' predictions are unlikely to be highly accurate consistently.
+Second, there are many players who are systematically *over*predicted (e.g., due to their injury) or *under*predicted (e.g., due to their becoming the starter after a teammate becomes injured, is traded, etc.)—an example of [overextremity miscalibration](#sec-calibration).
+In general, it is likely for players who are projected to score more points to be overpredicted and for players who are projected to score fewer points to be underpredicted [PRESENT CALIBRATION STATS].
+Third, the experts may interact and discuss with one another.
+In any case, they are able to see each other's projections and make change their projections, accordingly.
+
+## Conclusion {#sec-decisionMakingConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-decisionMakingSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2024-08-16 modern-portfolio-theory.qmd
20240816 - wisdom of the crowd

@@
 Consistent with this, crowd-averaged fantasy football projections tend to be more accurate than any individual's projection: INSERT
-This evidence is consistent with the notion of the wisdom of the crowd.
+This evidence is consistent with the notion of the [wisdom of the crowd](#sec-wisdomOfCrowd), described in @sec-wisdomOfCrowd.
 Moreover, even if the stock market is relatively accurate ("efficient") in terms of valuing stocks based on all (publicly) available information (i.e., the efficient market hypothesis), your fantasy football league is likely not.
@@
 In summary, fantasy football is similar to stock picking.
-You are most likely to pick the best players if you go with the wisdom of the crowd (e.g., average projections across projection sources) and diversify.
+You are most likely to pick the best players if you go with the [wisdom of the crowd](#sec-wisdomOfTheCrowd) (e.g., average projections across projection sources) and diversify.
 Most projections are public information, so you might wonder whether using crowd projections gains you anything because everybody else has access to public information.
 However, this is also the case with stocks, and people still consistently perform best over time when they go with the market.
+Nevertheless, crowd projections are not highly accurate.
+And fantasy football is a game, so feel free to have fun and deviate from the crowd!
+However, you may be just as (if not more) likely to be wrong by deviating from the crowd.
 

### 2024-08-16 machine-learning.qmd
Built site for gh-pages

@@
 
-When performing machine learning, there may be multiple approaches you can take to address this:
-
-- transforming data from [long to wide](#sec-longToWide) form, so that each person has only one row
+When performing machine learning, various approaches may help address this:
+
+- transform data from [long to wide](#sec-longToWide) form, so that each person has only one row
 - when designing the training and test sets, keep all measurements from the same person in the same data object (either the training or test set); do not have some measurements from a given person in the training set and other measurements from the same person in the test set

### 2024-08-16 cognitive-bias.qmd
20240816 - copyediting

@@
 The *anecdotal fallacy* is the error of drawing inferences based on personal experience or singular examples rather than based on stronger forms of evidence.
-Although people tend to find stories and anecdotes compelling, they are among the weakest forms of evidence, as described in @#sec-internalExternalValidityTradeoffs.
+Although people tend to find stories and anecdotes compelling, they are among the weakest forms of evidence, as described in @sec-internalExternalValidityTradeoffs.
 Consider that a fantasy football commentator writes an article about how they won their league by drafting a Quarterback and Wide Receiver by drafting with their first two picks, respectively.
-The anecdotal fallacy might lead a manager who reads the article to believe that they are more likely to win their league if they follow a similar strategy, despite more compelling evidence otherwise.
+The anecdotal fallacy might lead a manager who reads the article to believe that they are more likely to win their league if they follow a similar strategy, despite stronger evidence otherwise.
 

### 2024-08-16 decision-making.qmd
20240816 - updates

@@
 Cognitive diversity refers to the representation of individuals with different perspectives because of their "differences in knowledge, training, experience, or thinking styles" [@Rader2017, p. 8].
-Error cancellation deals with the idea that some individuals' predictions will fall above the truth and some individuals' predictions will fall below the truth—i.e., individual judgments will "bracket" the truth [@Mannes2014].
-So, the average of the predictions will "cancel out" some of the errors so that the predictions average out to more closely approximate the truth.
+Cognitive diversity is important because judgments from a cognitively homogeneous group will tend to err *systematically*.
+That is, they tend to err in the same direction—either consistently above or below the truth.
+By contrast, a cognitively diverse group will not tend to err systematically.
+For people of a cognitively diverse group, their judgments will tend to err *randomly*—where some people's predictions fall above the truth and some people's predictions fall below the truth—i.e., individual judgments "bracket" the truth [@Mannes2014].
+Error cancellation deals with the idea that, when individuals' judgments bracket the truth and show random rather than systematic error, the average of the predictions will "cancel out" some of the errors so that the predictions average out to more closely approximate the truth.
 
@@
 - there is cognitive diversity among the forecasters
-- the projections are made independently
-- the bracketing rate (i.e., the frequency with which any two forecasters' predictions fall on opposite sides of the truth) is high
+- the projections are made independently—i.e., the forecasters are not aware of others' forecasts and do not discuss or interact with the other forecasters
+- the bracketing rate—i.e., the frequency with which any two forecasters' predictions fall on opposite sides of the truth—is high
 
-However, the crowd is not accurate in all situations or domains.
+However, the crowd is not more accurate than the expert or best forecaster in all situations or domains.
 For instance, the crowd tends to be less accurate than the (prospectively identified) best forecaster when there is great variability in forecasters' expertise (in terms of the forecasters' ability to forecast accurately) and when the bracketing rate is low  [@Mannes2014].

### 2024-08-17 decision-making.qmd
20240817 - updates

@@
 In some domains, the average of non-expert forecasts is more accurate than the forecasts by individual experts.
-This phenomenon is called "collective intelligence", the "wisdom of the crowd", or the "wisdom of crowds" [@Rader2017; @Simoiu2019; @Surowiecki2005; @Wagner2010].
+This phenomenon is called "collective intelligence", the "wisdom of the crowd", or the "wisdom of crowds" [@Larrick2024; @Rader2017; @Simoiu2019; @Surowiecki2005; @Wagner2010].
 
-Aggregation of predictions from multiple people leverages several important features, including cognitive diversity, error cancellation, INSERT.
+Aggregation of predictions from multiple people leverages several important features, including cognitive diversity and error cancellation.
 
@@
 Error cancellation deals with the idea that, when individuals' judgments bracket the truth and show random rather than systematic error, the average of the predictions will "cancel out" some of the errors so that the predictions average out to more closely approximate the truth.
+However, when individuals' judgments do not bracket the truth, the average of the predictions will not cancel out the errors.
 
 Averaging projections from individuals tends to yield predictions that are more accurate than the accuracy of most forecasters in the group [@Mannes2014].
+Indeed, averaged predictions must be as (if not more) accurate than the average forecaster (i.e., the forecaster at the 50th percentile in terms of accuracy), and averaged predictions are often much more accurate [@Larrick2024].
 Moreover, averaged projections tends to be more accurate than consensus-based judgments from groups of people that interact and discuss, due to cognitive biases associated with the social interaction among groups, such as herding in which people align their behavior with others [@Mannes2014; @Simoiu2019].
@@
 
+When forecasters' predictions perfect bracket the truth (i.e., 50% of predictions are above the truth and 50% of predictive are below the truth), the average predictions must be more accurate than the accuracy of the average forecaster [@Larrick2024].
+
 However, the crowd is not more accurate than the expert or best forecaster in all situations or domains.
@@
 Even though the crowd tends to be more accurate than individual forecasters, crowd-averaged projections (at least among experts) are not necessarily highly accurate.
-In fantasy football, [DESCRIBE THEIR ACCURACY].
+In fantasy football, [DESCRIBE THEIR ACCURACY--BOTH FOR ALL PLAYERS AND WHEN SUBSETTING TO THE TOP X PLAYERS].
+Nevertheless, individual sources tend to be even less accurate. [DESCRIBE THEIR ACCURACY].
+
+Even though some sources are more accurate than the average in a given year, they are not *consistently* more accurate than the average.
+Prediction involves a combination of luck and skill.
+In some years, a prediction will invariably do better than others, in party, based on luck.
+However, luck is unlikely to continue systematically into future years, so a source that got lucky in a given year is unlikely to continue to be lucky in subsequent years.
+That is, determinining the most accurate source in a given year, after the fact, is not necessarily the same as identifying the most skilled forecaster.
+It requires a large sample of predictions to determine whether a given forecaster is reliably (i.e., consistently) more accurate than other forecasters and to identify the most accurate forecaster [@Larrick2024].
+Thus, it can be challenging to know, in advance, who the most accurate forecasters will be.
+Because average projections are as or more accurate than the average forecaster's prediction, averaging projections across all forecasters is superior to choosing individual forecasters when the forecasters are roughly similar in forecasting ability or when it is hard to distinguish their ability in advance [@Larrick2024].
+
 The relatively modest accuracy of the projections by so-called fantasy "experts'" and of their average of their projections could occur for a number of reasons.
@@
 Third, the experts may interact and discuss with one another.
+Interaction and discussion among experts may lead them to follow the herd and conform their projections to what each other predict.
 In any case, they are able to see each other's projections and make change their projections, accordingly.

### 2024-08-18 decision-making.qmd
20240817 - updates

@@
 By contrast, a cognitively diverse group will not tend to err systematically.
-For people of a cognitively diverse group, their judgments will tend to err *randomly*—where some people's predictions fall above the truth and some people's predictions fall below the truth—i.e., individual judgments "bracket" the truth [@Mannes2014].
+Individuals of a cognitively diverse group will bring different areas of expertise (e.g., determination of player skill, player opportunity, matchup strength, etc.) to bear in making their predictions and will thus make different mistakes [@Larrick2024].
+Consequently, for the people comprising a cognitively diverse group, their judgments will tend to err *randomly*—where some people's predictions fall above the truth and some people's predictions fall below the truth—i.e., individual judgments "bracket" the truth [@Mannes2014].
+
 Error cancellation deals with the idea that, when individuals' judgments bracket the truth and show random rather than systematic error, the average of the predictions will "cancel out" some of the errors so that the predictions average out to more closely approximate the truth.
@@
 Averaging projections from individuals tends to yield predictions that are more accurate than the accuracy of most forecasters in the group [@Mannes2014].
-Indeed, averaged predictions must be as (if not more) accurate than the average forecaster (i.e., the forecaster at the 50th percentile in terms of accuracy), and averaged predictions are often much more accurate [@Larrick2024].
-Moreover, averaged projections tends to be more accurate than consensus-based judgments from groups of people that interact and discuss, due to cognitive biases associated with the social interaction among groups, such as herding in which people align their behavior with others [@Mannes2014; @Simoiu2019].
+Indeed, when at least some of of the projections bracket the truth, averaged predictions must be more accurate than the average individual forecaster—in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE) (not in terms of the forecaster at the 50th percentile)—and averaged predictions are often much more accurate [@Larrick2024].
+If none of the projections bracket the truth—e.g., all projections overestimate the truth—averaged predictions will be as acurate than the average individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError) [@Larrick2024].
+In sum, "Averaging the answers of a crowd, therefore, ensures a level of accuracy no worse than the average member of the crowd and, in some cases, a level better than nearly all members" [@Larrick2024, p.
+126].
+Moreover, averaged projections tend to be more accurate than consensus-based judgments from groups of people that interact and discuss, due to cognitive biases associated with the social interaction among groups, such as herding in which people align their behavior with others [@Mannes2014; @Simoiu2019].
+
+There are well-known prediction markets that allow you to see the crowd-average prediction for various events:
+
+- <https://www.predictit.org>
+- <https://polymarket.com>
 
@@
 - the bracketing rate—i.e., the frequency with which any two forecasters' predictions fall on opposite sides of the truth—is high
-
-When forecasters' predictions perfect bracket the truth (i.e., 50% of predictions are above the truth and 50% of predictive are below the truth), the average predictions must be more accurate than the accuracy of the average forecaster [@Larrick2024].
 
@@
 
+The [`petersenlab` package](https://cran.r-project.org/web/packages/petersenlab/index.html) [@R-petersenlab] has the `wisdomOfCrowd()` function that computes the overall accuracy of the crowd-averaged projections, including the bracketing rate of the individual projections.
+
+```{r}
+# insert example using: wisdomOfCrowd()
+```
+
 Even though some sources are more accurate than the average in a given year, they are not *consistently* more accurate than the average.
@@
 That is, determinining the most accurate source in a given year, after the fact, is not necessarily the same as identifying the most skilled forecaster.
+It is easy to identify the most accurate source after the fact, but it is challenging to predict, in advance, who the best forecaster will be [@@Larrick2024].
 It requires a large sample of predictions to determine whether a given forecaster is reliably (i.e., consistently) more accurate than other forecasters and to identify the most accurate forecaster [@Larrick2024].
@@
 Even if the individual forecasters' projections are not close to the truth, if ~50% of them overestimate the truth and the other 50% of the underestimate the truth, the average will more closely approximate the truth.
-However, if most forecasters overestimate the truth, averaging the projections will not necessarily lead to more accurate projections.
+However, if all forecasters overestimate the truth for a given player, averaging the projections will not necessarily lead to more accurate projections.
 
@@
 Interaction and discussion among experts may lead them to follow the herd and conform their projections to what each other predict.
+This has been terms informational influence and may reflect the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment) [@Larrick2024].
 In any case, they are able to see each other's projections and make change their projections, accordingly.

### 2024-08-18 decision-making.qmd
20240817 - decision making in the context of uncertainty

@@
 ```
+
+## Overview of Decision Making in Uncertainty {#sec-decisionMakingOverview}
 
@@
 Averaging projections from individuals tends to yield predictions that are more accurate than the accuracy of most forecasters in the group [@Mannes2014].
-Indeed, when at least some of of the projections bracket the truth, averaged predictions must be more accurate than the average individual forecaster—in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE) (not in terms of the forecaster at the 50th percentile)—and averaged predictions are often much more accurate [@Larrick2024].
-If none of the projections bracket the truth—e.g., all projections overestimate the truth—averaged predictions will be as acurate than the average individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError) [@Larrick2024].
+Indeed, when at least some of of the projections bracket the truth, averaged predictions must be more accurate than the average individual forecaster—in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE)—and averaged predictions are often much more accurate [@Larrick2024].
+When referring to the accuracy of the "average individual forecaster", we are referring to accuracy in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE)—not to the accuracy of the forecaster at the 50th percentile.
+If none of the projections bracket the truth—e.g., all projections overestimate the truth—averaged predictions will be as accurate than the average individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError) [@Larrick2024].
 In sum, "Averaging the answers of a crowd, therefore, ensures a level of accuracy no worse than the average member of the crowd and, in some cases, a level better than nearly all members" [@Larrick2024, p.
 126].
-Moreover, averaged projections tend to be more accurate than consensus-based judgments from groups of people that interact and discuss, due to cognitive biases associated with the social interaction among groups, such as herding in which people align their behavior with others [@Mannes2014; @Simoiu2019].
+Moreover, averaged projections tend to be more accurate than consensus-based judgments from groups of people that interact and discuss, due to cognitive biases associated with the social interaction among groups, such as herding in which people align their behavior with others [@Mannes2014; @Simoiu2019], though discussion can be helpful in some contexts [@Larrick2024].
 
-There are well-known prediction markets that allow you to see the crowd-average prediction for various events:
+There are well-known prediction markets, in which people bet money to make predictions for various events, which allows determining the crowd-averaged prediction for events:
 
@@
 - the bracketing rate—i.e., the frequency with which any two forecasters' predictions fall on opposite sides of the truth—is high
+- there are at least 5–10 sources of projections
 
@@
 Thus, it may be necessary to examine the average of a "select crowd", by aggregating the projections of the most consistently accurate forecasters [@Mannes2014].
+Incorporating at least 5–10 forecasters leverages most of the benefits of the crowd; adding additional forecasters tends to result in diminishing returns [@Larrick2024].
+However, to the extent that those who are most accurate in a given period reflects luck, you are better off averaging the predictions of all forecasters than selecting the forecasters who were most accurate in the most recent period [@Larrick2024].
 
@@
 Prediction involves a combination of luck and skill.
-In some years, a prediction will invariably do better than others, in party, based on luck.
-However, luck is unlikely to continue systematically into future years, so a source that got lucky in a given year is unlikely to continue to be lucky in subsequent years.
+In some years, a prediction will invariably do better than others, in part, based on luck.
+However, luck is unlikely to continue systematically into future years, so a source that got lucky in a given year is likely to [regress to the mean](#sec-fallaciesRegression) in subsequent years.
 That is, determinining the most accurate source in a given year, after the fact, is not necessarily the same as identifying the most skilled forecaster.
-It is easy to identify the most accurate source after the fact, but it is challenging to predict, in advance, who the best forecaster will be [@@Larrick2024].
+It is easy to identify the most accurate source after the fact, but it is challenging to predict, in advance, who the best forecaster will be [@Larrick2024].
 It requires a large sample of predictions to determine whether a given forecaster is reliably (i.e., consistently) more accurate than other forecasters and to identify the most accurate forecaster [@Larrick2024].
@@
 
+## How Well Do People Incorporate Advice From Others? {#sec-incoporateAdvice}
+
+An important question is how well people incorporate advice from others.
+In the context of fantasy football, advice might be sources of projections.
+In general, evidence from social psychology suggests that people tend to underweight how much weight they give to advice from others relative to their own opinions, a phenomeon called egocentric discounting [@Larrick2024; @Rader2017].
+People tend to weight others' advice around 30% in terms of the proportion of a shift a person makes toward another person's perspective, though this depends on the perceived accuracy of the advisor.
+Moreover, people frequently ignores others' advice entirely.
+In general, people make use of crowds too little—they put too much weight on their own prediction and not enough weight on others' predictions whose diversity can be leveraged for error cancellation [@Larrick2024].
+
+## Suggestions {#sec-decisionMakingSuggestions}
+
+Based on the above dicussion, here are some suggestions for decision making in the context of uncertainty:
+
+- seek advice from diverse perspectives and incorporate it into your decision making
+- try to get opinions from others *before* you state your perspective and before the various sources of advice discuss with each other, to ensure independence of advice
+- if some sources of advice (or projections) are clearly more skilled and accurate than others, you can average this "select" crowd of projections or give them greater weight in a weighted average
+- if it is unclear whether or which sources are reliably more accurate than others, using a simple average across all sources (i.e., crowd-averaged projections) can be a useful approach that is as accurate as—if not more accurate than—the average individual forecaster
+- incorporate at least 5–10 sources of projections
+
 ## Conclusion {#sec-decisionMakingConclusion}

### 2024-08-18 time-series-analysis.qmd
20240818 - Bayesian model forecasts

@@
 ```{r}
+library("petersenlab")
 library("xts")
@@
 library("forecast")
+library("brms")
+library("plotly")
 library("tidyverse")
@@
 load(file = "./data/player_stats_seasonal.RData")
+load(url("https://osf.io/download/q6rjf/"))
 ```
@@
 There many different types of time series analyses.
-For simplicity, in this chapter, we use autoregressive integrated moving average (ARIMA) models to demonstrate one approach to time series analysis.
+For simplicity, in this chapter, we use [autoregressive integrated moving average](#sec-timeSeriesARIMA) (ARIMA) models to demonstrate one approach to time series analysis.
+We also leverage [Bayesian mixed models](#sec-fantasyPointsByAgeExperienceModelsBayesian) to generate forecasts of future performance and plots of individuals model-implied performance by age and position.
 
@@
 
-## Create the Time Series Objects {#sec-createTimeSeriesObjects}
+### Create the Time Series Objects {#sec-createTimeSeriesObjects}
 
@@
 
-## Plot the Time Series {#sec-plotTimeSeries}
+### Plot the Time Series {#sec-plotTimeSeries}
 
@@
 
-## Rolling Mean/Median {#sec-rollingMeanMedian}
+### Rolling Mean/Median {#sec-rollingMeanMedian}
 
@@
 
-## Autocorrelation {#sec-autoCorrelation}
+### Autocorrelation {#sec-autoCorrelation}
 
@@
 
-## Fit an Autoregressive Integrated Moving Average Model {#sec-arimaModel}
+### Fit an Autoregressive Integrated Moving Average Model {#sec-arimaModel}
 
@@
 
-## Generate the Model Forecasts {#sec-arimaForecasts}
+### Generate the Model Forecasts {#sec-arimaForecasts}
 
@@
 
-## Plot the Model Forecasts {#sec-plotARIMAforecast}
+### Plot the Model Forecasts {#sec-plotARIMAforecast}
 
@@
 
+## Bayesian Mixed Models {#sec-timeSeriesBayesian}
+
+The Bayesian longitudinal mixed models were estimated in @sec-fantasyPointsByAgeExperienceModelsBayesian.
+
+### Prepare New Data Object {#sec-timeSeriesBayesianNewData}
+
+```{r}
+player_stats_seasonal_offense_subset <- player_stats_seasonal_offense %>% 
+  filter(position_group %in% c("QB","RB","WR","TE"))
+
+player_stats_seasonal_offense_subset$position[which(player_stats_seasonal_offense_subset$position == "HB")] <- "RB"
+
+player_stats_seasonal_kicking_subset <- player_stats_seasonal_kicking %>% 
+  filter(position == "K")
+
+player_stats_seasonal_offense_subset <- bind_rows(
+  player_stats_seasonal_offense_subset,
+  player_stats_seasonal_kicking_subset
+)
+
+player_stats_seasonal_offense_subset$player_idFactor <- factor(player_stats_seasonal_offense_subset$player_id)
+player_stats_seasonal_offense_subset$positionFactor <- factor(player_stats_seasonal_offense_subset$position)
+```
+
+```{r}
+player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subset %>%
+  filter(
+    !is.na(player_idFactor),
+    !is.na(fantasy_points),
+    !is.na(positionFactor),
+    !is.na(ageCentered20),
+    !is.na(ageCentered20Quadratic),
+    !is.na(years_of_experience))
+
+player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subsetCC %>% 
+  filter(player_id %in% bayesianMixedModelFit$data$player_idFactor) %>% 
+  mutate(positionFactor = droplevels(positionFactor))
+
+player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subsetCC %>%
+  group_by(player_id) %>% 
+  group_modify(~ add_row(.x, season = max(player_stats_seasonal_offense_subsetCC$season) + 1)) %>% 
+  fill(player_display_name, player_idFactor, position, position_group, positionFactor, team, .direction = "downup") %>% 
+  ungroup
+
+player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subsetCC %>% 
+  left_join(
+    player_stats_seasonal_offense_subsetCC %>% 
+      filter(season == max(player_stats_seasonal_offense_subsetCC$season) - 1) %>% 
+      select(player_id, age_lastYear = age, years_of_experience_lastYear = years_of_experience), 
+    by = "player_id") %>%
+  mutate(
+    age = if_else(season == max(player_stats_seasonal_offense_subsetCC$season), age_lastYear + 1, age), # increment age by 1
+    ageCentered20 = age - 20,
+    years_of_experience = if_else(season == max(player_stats_seasonal_offense_subsetCC$season), years_of_experience_lastYear + 1, years_of_experience)) # increment experience by 1
+
+activePlayers <- unique(player_stats_seasonal_offense_subsetCC[c("player_id","season")]) %>% 
+  filter(season == max(player_stats_seasonal_offense_subsetCC$season) - 1) %>% 
+  select(player_id) %>% 
+  pull()
+
+inactivePlayers <- player_stats_seasonal_offense_subsetCC$player_id[which(player_stats_seasonal_offense_subsetCC$player_id %ni% activePlayers)]
+
+player_stats_seasonal_offense_subsetCC <- player_stats_seasonal_offense_subsetCC %>% 
+  filter(player_id %in% activePlayers | (player_id %in% inactivePlayers & season < max(player_stats_seasonal_offense_subsetCC$season) - 1)) %>% 
+  mutate(
+    player_idFactor = droplevels(player_idFactor) 
+  )
+```
+
+### Generate Predictions {#sec-timeSeriesBayesianGeneratePredictions}
+
+```{r}
+player_stats_seasonal_offense_subsetCC$fantasyPoints_bayesian <- predict(
+  bayesianMixedModelFit,
+  newdata = player_stats_seasonal_offense_subsetCC
+)[,"Estimate"]
+```
+
+### Table of Next Season Predictions {#sec-timeSeriesBayesianGeneratePredictions}
+
+```{r}
+player_stats_seasonal_offense_subsetCC %>% 
+  filter(season == max(player_stats_seasonal_offense_subsetCC$season), position == "QB") %>%
+  arrange(-fantasyPoints_bayesian) %>% 
+  select(player_display_name, fantasyPoints_bayesian)
+
+player_stats_seasonal_offense_subsetCC %>% 
+  filter(season == max(player_stats_seasonal_offense_subsetCC$season), position == "RB") %>%
+  arrange(-fantasyPoints_bayesian) %>% 
+  select(player_display_name, fantasyPoints_bayesian)
+
+player_stats_seasonal_offense_subsetCC %>% 
+  filter(season == max(player_stats_seasonal_offense_subsetCC$season), position == "WR") %>%
+  arrange(-fantasyPoints_bayesian) %>% 
+  select(player_display_name, fantasyPoints_bayesian)
+
+player_stats_seasonal_offense_subsetCC %>% 
+  filter(season == max(player_stats_seasonal_offense_subsetCC$season), position == "TE") %>%
+  arrange(-fantasyPoints_bayesian) %>% 
+  select(player_display_name, fantasyPoints_bayesian)
+```
+
+### Plot of Individuals' Model-Implied Predictions {#sec-timeSeriesBayesianGeneratePredictions}
+
+#### Quarterbacks {#sec-fantasyPointsIndividualByAgePlotsQB}
+
+```{r}
+#| label: fig-individualFantasyPointsBayesianByAgeQB
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Quarterbacks."
+#| code-fold: true
+
+plot_individualFantasyPointsByAgeQB <- ggplot(
+  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "QB"),
+  mapping = aes(
+    x = round(age, 2),
+    y = round(fantasyPoints_bayesian, 2),
+    group = player_id)) +
+  geom_smooth(
+    aes(
+      x = age,
+      y = fantasyPoints_bayesian,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    se = FALSE,
+    linewidth = 0.5,
+    color = "black") +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_bayesian,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age: Quarterbacks"
+  ) +
+  theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsByAgeQB,
+  tooltip = c("age","fantasyPoints_bayesian","text","label")
+)
+```
+
+#### Running Backs {#sec-fantasyPointsIndividualByAgePlotsRB}
+
+```{r}
+#| label: fig-individualFantasyPointsBayesianByAgeRB
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Running Backs."
+#| code-fold: true
+
+plot_individualFantasyPointsByAgeRB <- ggplot(
+  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "RB"),
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_bayesian,
+    group = player_id)) +
+  geom_smooth(
+    aes(
+      x = age,
+      y = fantasyPoints_bayesian,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    se = FALSE,
+    linewidth = 0.5,
+    color = "black") +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_bayesian,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age: Running Backs"
+  ) +
+  theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsByAgeRB,
+  tooltip = c("age","fantasyPoints_bayesian","text","label")
+)
+```
+
+#### Wide Receivers {#sec-fantasyPointsIndividualByAgePlotsWR}
+
+```{r}
+#| label: fig-individualFantasyPointsBayesianByAgeWR
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Wide Receivers."
+#| code-fold: true
+
+plot_individualFantasyPointsByAgeWR <- ggplot(
+  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "WR"),
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_bayesian,
+    group = player_id)) +
+  geom_smooth(
+    aes(
+      x = age,
+      y = fantasyPoints_bayesian,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    se = FALSE,
+    linewidth = 0.5,
+    color = "black") +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_bayesian,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
+  ) +
+  theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsByAgeWR,
+  tooltip = c("age","fantasyPoints_bayesian","text","label")
+)
+```
+
+#### Tight Ends {#sec-fantasyPointsIndividualByAgePlotsTE}
+
+```{r}
+#| label: fig-individualFantasyPointsBayesianByAgeTE
+#| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Tight Ends."
+#| code-fold: true
+
+plot_individualFantasyPointsByAgeTE <- ggplot(
+  data = player_stats_seasonal_offense_subsetCC %>% filter(position == "TE"),
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_bayesian,
+    group = player_id)) +
+  geom_smooth(
+    aes(
+      x = age,
+      y = fantasyPoints_bayesian,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    se = FALSE,
+    linewidth = 0.5,
+    color = "black") +
+  geom_point(
+    aes(
+      x = age,
+      y = fantasyPoints_bayesian,
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season # add season for mouse over tooltip
+    ),
+    size = 1,
+    color = "transparent" # make points invisible but keep tooltips
+  ) +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
+  ) +
+  theme_classic()
+
+ggplotly(
+  plot_individualFantasyPointsByAgeTE,
+  tooltip = c("age","fantasyPoints_bayesian","text","label")
+)
+```
+
 ## Conclusion {#sec-timeSeriesConclusion}

### 2024-08-23 cognitive-bias.qmd
20240822 - affect heuristic

@@
 As described [above](#sec-cognitiveBiasOverview), heuristics are mental shortcuts that people use to handle the overwhelming amount of information to process.
-Important heuristics used in judgment and prediction in the face of uncertainty include [@Tversky1974; @Critcher2014]:
+Important heuristics used in judgment and prediction in the face of uncertainty include [@Kahneman2011; @Tversky1974; @Critcher2014]:
 
@@
 - [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment)
+- [affect heuristic](#sec-heuristicsAffect)
 - [performance heuristic](#sec-heuristicsPerformance)
@@
 
-The *representativess heuristic* refers to the tendency for a person's judgments or predictions about individuals to be made based on how similar the individual is to the person's existing mental prototypes.
+The *representativess heuristic* refers to the tendency for a person's judgments or predictions about individuals to be made based on how similar the individual is to (i.e., how closely the individual resembles) the person's existing mental prototypes.
 For instance, when coming out of college, Tight End Kyle Pitts [drew comparisons](https://247sports.com/article/kyle-pitts-lebron-james-2021-nfl-draft-florida-gators-football-163882176) to the "LeBron James" of Tight Ends (archived at <https://perma.cc/JQB5-XPVL>).
@@
 [Regression effects](#sec-fallaciesRegression) are discussed further in @sec-fallaciesRegression.
+
+### Affect Heuristic {#sec-heuristicsAffect}
+
+The *affect heuristic* refers to the tendency for a person to make judgments or predictions based on their emotions and feelings—such as whether they like or dislike something—with little deliberation or reasoning [@Kahneman2011].
+For instance, if a manager's favorite team is the Dallas Cowboys, they may hold a distaste for players on teams that are rivals of the Cowboys, such as the Philadelphia Eagles.
+The affect heuristic may lead them to avoid drafting Eagles players even though some Eagles players may perform well.
 

### 2024-08-23 machine-learning.qmd
20240823 - change imputation procedure from missForest to missRanger for speed and out-of-sample prediction

@@
 library("doParallel")
-library("missForest")
+library("missRanger")
 library("powerjoin")
@@
 Here is a vignette demonstrating how to impute missing data using `missForest()`: <https://rpubs.com/lmorgan95/MissForest> (archived at: <https://perma.cc/6GB4-2E22>).
-Below, we impute the training data and new data separately by position.
-
-```{r}
-#| include: false
-
-num_cores <- parallel::detectCores()
-num_true_cores <- parallel::detectCores(logical = FALSE)
-```
-
-```{r}
-#| eval: false
-
-num_cores <- detectCores() - 1
-num_true_cores <- parallel::detectCores(logical = FALSE) - 1
-```
+Below, we impute the training data (and all data) separately by position.
+We then use the imputed training data to make out-of-sample predictions to fill in the missing data for the testing data.
+We do not want to impute the training and testing data together so that we can keep them separate for the purposes of cross-validation.
+However, we impute all data (training and test data together) for purposes of making out-of-sample predictions from the machine learning models to predict players' performance next season (when actuals are not yet available for evaluating their accuracy).
 
@@
 ```{r}
-num_cores
-
-cl <- parallel::makeCluster(num_cores)
-doParallel::registerDoParallel(cl)
-
-set.seed(52242)
-
 # QBs
-seasonalData_lag_qb_all_imp <- missForest::missForest(
-  seasonalData_lag_qb_all %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_qb_all_imp$OOBerror
-data_all_qb <- cbind(seasonalData_lag_qb_all[,"gsis_id"], seasonalData_lag_qb_all_imp$ximp)
+seasonalData_lag_qb_all_imp <- missRanger::missRanger(
+  seasonalData_lag_qb_all,
+  pmm.k = 5,
+  verbose = 2,
+  seed = 52242,
+  keep_forests = TRUE)
+
+seasonalData_lag_qb_all_imp
+
+data_all_qb <- seasonalData_lag_qb_all_imp$data
 data_all_qb_matrix <- data_all_qb %>%
@@
 
-seasonalData_lag_qb_train_imp <- missForest::missForest(
-  seasonalData_lag_qb_train %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_qb_train_imp$OOBerror
-data_train_qb <- cbind(seasonalData_lag_qb_train[,"gsis_id"], seasonalData_lag_qb_train_imp$ximp)
+seasonalData_lag_qb_train_imp <- missRanger::missRanger(
+  seasonalData_lag_qb_train,
+  pmm.k = 5,
+  verbose = 2,
+  seed = 52242,
+  keep_forests = TRUE)
+
+seasonalData_lag_qb_train_imp
+
+data_train_qb <- seasonalData_lag_qb_train_imp$data
 data_train_qb_matrix <- data_train_qb %>%
@@
 
-seasonalData_lag_qb_test_imp <- missForest::missForest(
-  seasonalData_lag_qb_test %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_qb_test_imp$OOBerror
-data_test_qb <- cbind(seasonalData_lag_qb_test[,"gsis_id"], seasonalData_lag_qb_test_imp$ximp)
+seasonalData_lag_qb_test_imp <- predict(
+  object = seasonalData_lag_qb_train_imp,
+  newdata = seasonalData_lag_qb_test,
+  seed = 52242)
+
+data_test_qb <- seasonalData_lag_qb_test_imp
 data_test_qb_matrix <- data_test_qb %>%
@@
 # RBs
-seasonalData_lag_rb_all_imp <- missForest::missForest(
-  seasonalData_lag_rb_all %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_rb_all_imp$OOBerror
-data_all_rb <- cbind(seasonalData_lag_rb_all[,"gsis_id"], seasonalData_lag_rb_all_imp$ximp)
+seasonalData_lag_rb_all_imp <- missRanger::missRanger(
+  seasonalData_lag_rb_all,
+  pmm.k = 5,
+  verbose = 2,
+  seed = 52242,
+  keep_forests = TRUE)
+
+seasonalData_lag_rb_all_imp
+
+data_all_rb <- seasonalData_lag_rb_all_imp$data
 data_all_rb_matrix <- data_all_rb %>%
@@
 
-seasonalData_lag_rb_train_imp <- missForest::missForest(
-  seasonalData_lag_rb_train %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_rb_train_imp$OOBerror
-data_train_rb <- cbind(seasonalData_lag_rb_train[,"gsis_id"], seasonalData_lag_rb_train_imp$ximp)
+seasonalData_lag_rb_train_imp <- missRanger::missRanger(
+  seasonalData_lag_rb_train,
+  pmm.k = 5,
+  verbose = 2,
+  seed = 52242,
+  keep_forests = TRUE)
+
+seasonalData_lag_rb_train_imp
+
+data_train_rb <- seasonalData_lag_rb_train_imp$data
 data_train_rb_matrix <- data_train_rb %>%
@@
 
-seasonalData_lag_rb_test_imp <- missForest::missForest(
-  seasonalData_lag_rb_test %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_rb_test_imp$OOBerror
-data_test_rb <- cbind(seasonalData_lag_rb_test[,"gsis_id"], seasonalData_lag_rb_test_imp$ximp)
+seasonalData_lag_rb_test_imp <- predict(
+  object = seasonalData_lag_rb_train_imp,
+  newdata = seasonalData_lag_rb_test,
+  seed = 52242)
+
+data_test_rb <- seasonalData_lag_rb_test_imp
 data_test_rb_matrix <- data_test_rb %>%
@@
 # WRs
-seasonalData_lag_wr_all_imp <- missForest::missForest(
-  seasonalData_lag_wr_all %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_wr_all_imp$OOBerror
-data_all_wr <- cbind(seasonalData_lag_wr_all[,"gsis_id"], seasonalData_lag_wr_all_imp$ximp)
+seasonalData_lag_wr_all_imp <- missRanger::missRanger(
+  seasonalData_lag_wr_all,
+  pmm.k = 5,
+  verbose = 2,
+  seed = 52242,
+  keep_forests = TRUE)
+
+seasonalData_lag_wr_all_imp
+
+data_all_wr <- seasonalData_lag_wr_all_imp$data
 data_all_wr_matrix <- data_all_wr %>%
@@
 
-seasonalData_lag_wr_train_imp <- missForest::missForest(
-  seasonalData_lag_wr_train %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_wr_train_imp$OOBerror
-data_train_wr <- cbind(seasonalData_lag_wr_train[,"gsis_id"], seasonalData_lag_wr_train_imp$ximp)
+seasonalData_lag_wr_train_imp <- missRanger::missRanger(
+  seasonalData_lag_wr_train,
+  pmm.k = 5,
+  verbose = 2,
+  seed = 52242,
+  keep_forests = TRUE)
+
+seasonalData_lag_wr_train_imp
+
+data_train_wr <- seasonalData_lag_wr_train_imp$data
 data_train_wr_matrix <- data_train_wr %>%
@@
 
-seasonalData_lag_wr_test_imp <- missForest::missForest(
-  seasonalData_lag_wr_test %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_wr_test_imp$OOBerror
-data_test_wr <- cbind(seasonalData_lag_wr_test[,"gsis_id"], seasonalData_lag_wr_test_imp$ximp)
+seasonalData_lag_wr_test_imp <- predict(
+  object = seasonalData_lag_wr_train_imp,
+  newdata = seasonalData_lag_wr_test,
+  seed = 52242)
+
+data_test_wr <- seasonalData_lag_wr_test_imp
 data_test_wr_matrix <- data_test_wr %>%
@@
 # TEs
-seasonalData_lag_te_all_imp <- missForest::missForest(
-  seasonalData_lag_te_all %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_te_all_imp$OOBerror
-data_all_te <- cbind(seasonalData_lag_te_all[,"gsis_id"], seasonalData_lag_te_all_imp$ximp)
+seasonalData_lag_te_all_imp <- missRanger::missRanger(
+  seasonalData_lag_te_all,
+  pmm.k = 5,
+  verbose = 2,
+  seed = 52242,
+  keep_forests = TRUE)
+
+seasonalData_lag_te_all_imp
+
+data_all_te <- seasonalData_lag_te_all_imp$data
 data_all_te_matrix <- data_all_te %>%
@@
 
-seasonalData_lag_te_train_imp <- missForest::missForest(
-  seasonalData_lag_te_train %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_te_train_imp$OOBerror
-data_train_te <- cbind(seasonalData_lag_te_train[,"gsis_id"], seasonalData_lag_te_train_imp$ximp)
+seasonalData_lag_te_train_imp <- missRanger::missRanger(
+  seasonalData_lag_te_train,
+  pmm.k = 5,
+  verbose = 2,
+  seed = 52242,
+  keep_forests = TRUE)
+
+seasonalData_lag_te_train_imp
+
+data_train_te <- seasonalData_lag_te_train_imp$data
 data_train_te_matrix <- data_train_te %>%
@@
 
-seasonalData_lag_te_test_imp <- missForest::missForest(
-  seasonalData_lag_te_test %>% 
-    select(-gsis_id) %>%
-    data.frame(),
-  verbose = TRUE,
-  mtry = 100,
-  parallelize = "forests")
-
-seasonalData_lag_te_test_imp$OOBerror
-data_test_te <- cbind(seasonalData_lag_te_test[,"gsis_id"], seasonalData_lag_te_test_imp$ximp)
+seasonalData_lag_te_test_imp <- predict(
+  object = seasonalData_lag_te_train_imp,
+  newdata = seasonalData_lag_te_test,
+  seed = 52242)
+
+data_test_te <- seasonalData_lag_te_test_imp
 data_test_te_matrix <- data_test_te %>%
@@
   as.matrix()
-
-stopCluster(cl)
+```
+
+## Identify Cores for Parallel Processing {#sec-coresParallel}
+
+```{r}
+#| include: false
+
+num_cores <- parallel::detectCores()
+num_true_cores <- parallel::detectCores(logical = FALSE)
+```
+
+```{r}
+#| eval: false
+
+num_cores <- detectCores() - 1
+num_true_cores <- parallel::detectCores(logical = FALSE) - 1
+```
+
+```{r}
+num_cores
 ```

### 2024-08-29 sports-cognitive-psychology.qmd
20240829 - copyediting

@@
 - Bill James, "Moneyball", Billy Beane, Theo Epstein, and role of Sabermetrics in helping MLB teams with limited budgets (e.g., Oakland Athletics) compete with larger budgets: using statistics to identify player value more accurately, especially for identifying undervalued players
-- development of advanced metrics (e.g., value over replacement player, wins above replacement, fieling independent pitching)
+- development of advanced metrics (e.g., value over replacement player, wins above replacement, fielding independent pitching)
 - focus on on-base percentage and slugging percentage (and on-base plus slugging percentage) instead of batting average; defensive shifting; fewer attempts to steal bases; fewer bunts; batters taking more pitches; more frequent pitching changes (for pitcher–batter matchups, e.g., righty–lefty); focus on velocity and spin rate

### 2024-08-29 actuarial.qmd
20240829 - media examples

@@
 For instance, a proclaimed "fantasy expert" might use their experience, expertise, and judgment to make a prediction about how each player will perform by using whatever information and data they deem to be important, aggregating all of this information in their mind to make the prediction for each player.
+Professional scouts and coaches use judgment when making predictions or selecting players based on their impressions of the players [@DenHartigh2018].
+As an example in popular media, in the movie, "Trouble with the Curve", a professional scout makes a judgment about how well a baseball hitter will do in the major leagues from his impressions of the hitter's ability based on the sound of the ball off the player's bat.
 
@@
 For generating the actuarial model, you could obtain the beta weights for each predictor from [multiple regression](#sec-multipleRegression), from [machine learning](#sec-machineLearning), or from prior research on the relative importance of each predictor.
+
+As an example of using the actuarial approach, Billy Beane, who was the general manager of the Oakland Athletics at the time, wanted to find ways for his team—which had less finanical resources than its competitors—to compete with teams that hard more money to sign players.
+Because the team did not have the resources to sign the best players, they had to find to find other ways to find the optimal players that they could afford.
+So, he used statistical formulas that weight variables, such as on-base percentage and slugging percentage, according to their value for winning games, for the use of selecting players.
+His approach became well-known based on the Michael Lewis book, "Moneyball: The Art of Winning an Unfair Game", and the eventual movie, "Moneyball".
 

### 2024-08-29 player-evaluation.qmd
20240828 - additions
20240829 - signs and samples

@@
 
-Compared to tests of speed, power, and agility at the NFL Combine, collegiate performance is a stronger predictor of performance in the NFL [@Lyons2011].
-That is, previous sports performance is the best predictor of future performance [for a review, see @DenHartigh2018].
-Thus, it is important to consider a player's past performance.
+It is important to consider a player's past performance.
 However, the extent to which historical performance may predict future performance may depend on many factors such as (a) the similarity of the prior situation to the current situation, (b) how long ago the prior situation was, and (c) the extent to which the player (or situation) has changed in the interim.
@@
 How efficient were they given the number of opportunities they had?
+This is important to consider because different players have different opportunities.
+For example, some Running Backs had more opportunity than others (e.g., more carries), so comparing players merely on rushing yards would be misleading.
 If they were relatively more efficient, they will likely score more points than many of their peers when given more opportunities.
@@
 
+## Signs Versus Samples {#sec-signsSamples}
+
+When thinking about the most important domains to assess for evaluating players and predicting their performance, it is important to distinguish between signs and samples [@DenHartigh2018].
+*Signs* are indicators of underlying states.
+The signs approach to assessment involves assessing processes that may predict performance, where the emphasis is on what the sign indicates about the underlying attribute, rather than on the specific behavior itself.
+In football, the signs approach might involve measuring skills using separate tests where players' skills are tested in isolation.
+For instance, the NFL Combine takes a signs approach to assessment, in which players have a separate test to assess for speed (e.g., 40-yard dash) and jumping ability (e.g., vertical jump).
+
+*Samples* reflect behaviors that are close to the behavior of interest.
+The samples approach to asesssment tries to assess the person's performance in a behavior, situation, and context that are as reflective as possible of the kinds of behaviors, situations, and contexts that the person would have to perform in for the position or occupation.
+In the samples approach, the behavior itself is the main emphasis, rather than an underlying attribute.
+In football, the samples approach might involve measuring the player's performance during games or game-like situations.
+For instance, to assess the skills of a Wide Receiver, you might observe—during games or game-like situations—how well they are able to catch passes when closely guarded or double-teamed, or how well they are able to catch poorly thrown passes.
+
+In general, samples tend to be stronger than signs for predicting player performance in sports [@DenHartigh2018].
+For instance, compared to tests of speed, power, and agility at the NFL Combine, collegiate performance is a stronger predictor of performance in the NFL [@Lyons2011].
+That is, previous sports performance is the best predictor of future performance [for a review, see @DenHartigh2018].
+
+According to the theoretical perspective to sports performance known as the ecological dynamics approach, successful performance in sports involves the coordination of multiple, intertwined skills that are contextually embedded [@DenHartigh2018].
+For instance, a Wide Receiver needs to coordinate skills in speed, route running, jumping, and good hands to create space from defenders and to catch poorly thrown passes in a game.
+The ecological dynamics approach considers the important interaction of the player, task, and environment.
+Thus, to assess players in a way that is likely to be most predictive of their future performance, it is important for the assessment to retain the player–task–environment interaction [@DenHartigh2018].
+For example, it would be valuable to use assessments from games or game-like situations in which players must perform tasks that leverage multiple, intertwined skills and that are similar to the tasks you want to predict.
+
 ## Putting it Altogether {#sec-evalIntegration}
 
-After performing an evaluation of the relevant domain(s) for a given player, then one must integrate the evaluation information across domains to make a judgment about a player's overall value.
+After performing an evaluation of the relevant domain(s) for a given player, one must integrate the evaluation information across domains to make a judgment about a player's overall value.
+When considering how much weight to give to each of various factors, it is important to evaluate how much [predictive validity](#sec-predictiveValidity) each factor has for predicting a player's successful performance in the NFL.
+Then, one can weight each variable according to its [predictive validity](#sec-predictiveValidity) using an [actuarial approach](#sec-actuarialPrediction).
+[Actuarial approaches](#sec-actuarialPrediction) are described in a later chapter, in @sec-actuarialPrediction.
+For now, suffice it to say that [actuarial approaches](#sec-actuarialPrediction) leverage statistical formulas, as opposed to using [judgment](#sec-humanJudgment) alone.
+As described in @sec-cognitiveBias, people's judgment—including judgments by experts (e.g., professional scouts and coaches)—tends to be riddled with biases [@DenHartigh2018].
+Moreover, experts tend to disagree in terms of their judgments/predictions about players [@DenHartigh2018].
+In addition, there are benefits to leveraging multiple perpectives, consistent with the "[wisdom of the crowd](#sec-wisdomOfCrowd)", as described in @sec-wisdomOfCrowd.
+In general, as noted in @sec-signsSamples, give more weight to [samples](#sec-signsSamples) of relevant behavior—such as past performance in games—than to [signs](#sec-signsSamples) such as NFL Combine metrics (e.g., bench press, 40-yard dash, etc.).
+Where you can, use assessments from games or game-like situations in which players must perform tasks that leverage multiple, intertwined skills and that are similar to the tasks you want to predict.
+
 When thinking about a player's value, it can be worth thinking of a player's upside and a player's downside.
-Player that are more consistent may show higher downside but a lower upside.
+Players that are more consistent may show higher downside but a lower upside.
 Younger, less experienced players may show a higher upside but a lower downside.

### 2024-08-29 actuarial.qmd
20240829 - steps to conduct actuarial approaches

@@
 
+## Steps to Conduct Actuarial Approaches {#sec-stepsActuarial}
+
+Here are several steps to conduct actuarial approaches [@DenHartigh2018; @Kahneman2011]:
+
+1. Determine a set of relevant variables to measure
+1. Determine how you will combine the variables
+    - Do some variables have more weight than other variables?
+1. Determine how the variables will be scored
+    - e.g., a 7-point likert scale
+1. Combine the scores based on the pre-specified formula
+1. Use the final score to make your prediction (for selecting players)
+
 ## Challenges of Data-Driven Approaches and How to Address {#sec-challengesDataDrivenApproaches}

### 2024-08-30 cognitive-bias.qmd
20240823 - use of heuristics
20240829 - Kahneman WYSIATI; independent judgments; overconfidence

@@
 - [performance heuristic](#sec-heuristicsPerformance)
+- ["What You See Is All There Is" (WYSIATI) heuristic](#sec-heuristicsWYSIATI)
+
+Use of heuristics has been observed among fantasy sports managers [@Smith2006a].
 
@@
 For instance, if a manager has a successful first season, the performance heuristic may lead them to overestimate the extent to which they will improve in their skill or performance in the future.
+
+### What You See Is All There Is (WYSIATI) Heuristic {#sec-heuristicsWYSIATI}
+
+The "What You See Is All There Is" (WYSIATI) heuristic refers to the tendency for a person to make a decision or judgment based solely on the information that is immediately available to them (i.e., "right in front of them"), without considering the possibility that there may be additional important information that they do not have [@Enke2020].
+The WYSIATI heuristic involves jumping to conclusions on the basis of limited information [@Kahneman2011].
+For instance, a manager might start a player based on the fact that the player scored lots of points in their most recent game.
+However, the manager may not consider other important factors that they did not know about, such as the fact that the player filled in for an injured player, and will get fewer touches in the upcoming game now that the original starter is back to full health.
 
@@
 
+The confidence that people tend to have in their predictions depends less and the quantity and quality of the evidence for their predictions and more on the quality of the narrative they can tell about the (little) information they have immediate access to [@Kahneman2011].
+For instance, people tend not to consider other relevant, important factors that they do not have immediate access to, consistent with the [WYSIATI heuristic](#sec-heuristicsWYSIATI).
+
 A fantasy manager may be even more likely to exhibit overconfidence if they previously performed well or won their league, for which luck and random chance plays an important role.
@@
 *Confirmation bias* is the tendency for people to search for, interpret, and remember information that confirms one's beliefs, as opposed to information that might disconfirm one's beliefs.
+The search for confirming evidence is called "positive test strategy"; it runs counter to advice by philosophers of science who advise testing hypotheses by trying to refute them [i.e., by searching for *disconfirming* evidence; @Kahneman2011].
 The result of confirmation bias is that people are unlikely to change their minds about something that they have a pre-existing belief about, because they tend to look only for information that supports their pre-existing beliefs.

### 2024-08-30 decision-making.qmd
20240829 - wisdom of the crowd
20240829 - Kahneman WYSIATI; independent judgments; overconfidence

@@
 Cognitive diversity is important because judgments from a cognitively homogeneous group will tend to err *systematically*.
-That is, they tend to err in the same direction—either consistently above or below the truth.
+That is, they tend to err in the same direction—either consistently above or below the truth; thus, their errors are correlated.
 By contrast, a cognitively diverse group will not tend to err systematically.
@@
 Consequently, for the people comprising a cognitively diverse group, their judgments will tend to err *randomly*—where some people's predictions fall above the truth and some people's predictions fall below the truth—i.e., individual judgments "bracket" the truth [@Mannes2014].
+Thus, judgments from a cognitively diverse group tend to have uncorrelated errors.
 
@@
 - <https://polymarket.com>
+- <https://www.rotowire.com/betting/nfl/player-futures.php>
 
@@
 In fantasy football, [DESCRIBE THEIR ACCURACY--BOTH FOR ALL PLAYERS AND WHEN SUBSETTING TO THE TOP X PLAYERS].
-Nevertheless, individual sources tend to be even less accurate. [DESCRIBE THEIR ACCURACY].
+Nevertheless, individual sources tend to be even less accurate.
+[DESCRIBE THEIR ACCURACY].
 
@@
 ```{r}
-# insert example using: wisdomOfCrowd()
+# insert example using historical projections: wisdomOfCrowd()
+```
+
+```{r}
+predictedValues <- c(10,20,30,40,60,70,80,90)
+actualValue <- 50
+errors <- predictedValues - actualValue
+
+errors
+median(errors)
+mean(predictedValues)
+
+petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+```
+
+```{r}
+predictedValues <- c(10,29,29,40,60,70,80,90)
+actualValue <- 71
+errors <- predictedValues - actualValue
+
+errors
+median(errors)
+mean(predictedValues)
+
+petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+```
+
+```{r}
+predictedValues <- c(10,20,30,40,60,70,80,90)
+actualValue <- 100
+errors <- predictedValues - actualValue
+
+errors
+median(errors)
+mean(predictedValues)
+
+petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
 ```
@@
 - try to get opinions from others *before* you state your perspective and before the various sources of advice discuss with each other, to ensure independence of advice
+  - As noted by @Kahneman2011, "before an issue is discussed, all members of the committee should be asked to write a very brief summary of their position...
+The standard practice of open discussion gives too much weight to the opinions of those who speak early and assertively, causing others to line up behind them." (p.
+85).
 - if some sources of advice (or projections) are clearly more skilled and accurate than others, you can average this "select" crowd of projections or give them greater weight in a weighted average

### 2024-08-30 draft.qmd
20240830 - draft and player eval

@@
 Sometimes it works best to "zig" when everyone else is "zagging".
+For instance, @Lee2022 found that the most successful draft strategies in terms of roster composition (i.e., the number of players for each position) were not the most common draft strategies.
 For instance, if you notice that everyone else is drafting Wide Receivers, this may mean that other managers are over-valuing Wide Receivers, and this could be a nice opportunity to draft a Running Back for good value.
@@
 
+As described in @sec-wisdomOfCrowd, there can be benefits of leveraging the [wisdom of the crowd](#sec-wisdomOfCrowd) by using rankings or projections that are averaged across many people and perspectives.
 @sec-fantasyValueSources describes where to obtain [aggregated rankings](#sec-fantasyValueRankings), [aggregated projections](#sec-fantasyValueProjections), [ADP](#sec-fantasyValueADPAAV), and [AAV](#sec-fantasyValueADPAAV) data.
+As described in @sec-judgmentVsActuarial, there are also benefits to using the [actuarial approach](#sec-judgmentVsActuarial) to prediction rather than (merely) using judgment.
+
+It is not sufficient to compare players in terms of projected points because some positions have more depth than other positions.
+Some positions show *positional scarcity*—that is, a limited number of high performing players.
 
@@
 In this example, there is a much greater [dropoff](#sec-fantasyValueDropoff) for Tight Ends than there is for Quarterbacks.
-Thus, even though "Quarterback A" is projected to score more points than "Tight End A", "Tight End A" may be more valuable because there is still a good Quarterback available if someone else drafts "Quarterback A".
+Thus, even though "Quarterback A" is projected to score more points than "Tight End A", "Tight End A" may be more valuable, relatively, because there is still a good Quarterback available if someone else drafts "Quarterback A".
+In general, Kickers and Defenses tend to have the lowest dropoff (i.e., the lowest expected drop in fantasy points) by positional rank, so it makes sense to draft Kickers and Defenses late in the draft [@Lee2022].
+Defenses, in particular, appear to be among the least predictable of the positions [@Lee2022].
 
 Another important concept is a player's [value over a typical replacement player](#sec-fantasyValueVORP) at that position (shortened to "value over replacement player"; VORP), which is described in @sec-fantasyValueVORP.
+A player's [value over a typical replacement player](#sec-fantasyValueVORP) provides a way to more fairly compare (and thus rank) players across different positions.
 
@@
 
+In general, the most consequential decisions tend to be those made early in the draft regarding the top projected players [@Lee2022].
+That is, the earlier selections tend to have the greatest impact on the success of one's fantasy season.
+So, make sure to spend time in your draft preparation to identify the players you want to select early in the draft.
+Moreover, some teams like to hedge their bet on their top Running Backs in the case that the player were to get injured, by drafting the player's backup, a strategy known as *handcuffing*.
+However, there is not strong evidence that handcuffing leads to better outcomes [@Lee2022].
+
+In general, @Lee2022 found that teams that drafted more Running Backs and Wide Receivers tended to outperform other teams.
+
 ### Snake Draft {#sec-draftStrategySnake}
@@
 For instance, if a particular Quarterback has a slightly higher [VORP](#sec-fantasyValueVORP) than a particular Running Back, but the Quarterback is likely to be available at the manager's next pick but the Running Back is likely to be unavailable at their next pick, it might make more sense to draft the Running Back.
+
+Herding behavior is common in snake drafts.
+As described in @sec-wisdomOfCrowd, herding occurs when people align their behavior with others.
+For instance, @Lee2022 found evidence that when Quarterbacks, Kickers, and Defenses were drafted by one team, subsequent teams were more likely to draft a player of that position.
+The same was not the case for Running Backs, Wide Receivers, and Tight Ends.
 

### 2024-08-30 player-evaluation.qmd
20240829 - Vegas lines
20240830 - draft and player eval

@@
 
+## Availability {#sec-evalAvailability}
+
+In addition to [injuries](#sec-evalHealth), other factors can affect a player's availability, including league suspensions, for instance, due to violating the league's conduct policy.
+
 ## Age and Career Stage {#sec-evalAgeCareerStage}
@@
 Projections that are aggregated across sources are provided by [Fantasy Football Analytics](https://fantasyfootballanalytics.net) (see the webapp here: <https://apps.fantasyfootballanalytics.net>) and by [FantasyPros](https://www.fantasypros.com/nfl/auction-values/calculator.php).
+Moreover, sites also provide season-long player futures for the over/under betting lines, which are based on Vegas lines and how people bet: <https://www.rotowire.com/betting/nfl/player-futures.php>.
 

### 2024-08-30 download-football-data.qmd
20240830 - merging

@@
 We save each data file after downloading it, so we can use the data in subsequent chapters.
-Guidance for how to merge the various data files is provided at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md>
+This chapter uses extensive use of merging to process the data for later use.
+See @sec-merging for a reminder of how to perform merging, the types of merges, and what you can expect when you merge data objects with different formats.
+Guidance for how to merge the various NFL-related data files is provided at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md>.
 

### 2024-08-30 getting-started.qmd
20240823 - git
20240830 - merging

@@
 
+1. Download and install `git`: <https://git-scm.com/downloads>
 1. Set up a free account on [GitHub.com](https://github.com).
-1. Download GitHub Desktop: <https://desktop.github.com>
+1. Download and install `GitHub Desktop`: <https://desktop.github.com>
 1. Make sure you are logged into your GitHub account on [GitHub.com](https://github.com).
@@
 For instance, in long form data objects, each participant may have multiple rows corresponding to multiple seasons.
-In this case, the keys may be `player_id` and `season`.
-If a participant has multiple rows corresponding to seasons and games/weeks, the keys are `player_id`, `season`, and `week`.
+In this case, the keys may be `player_id` and `season`—that is, the data are in `player`-`season` form.
+If a participant has multiple rows corresponding to seasons and games/weeks, the keys are `player_id`, `season`, and `week`—in this case, the data are in `player`-`season`-`week` form.
 In general, each row should have a value on each of the keys; there should be no missingness in the keys.
@@
 However, as described above, you should include it as a key ***only*** if it uniquely identifies each row in terms of levels of nesting.
+
+When two objects are merged that have different formats, the resulting data object inherits the format of the data object that has more levels of nesting.
+For instance, consider that you want to merge two objects, object A and object B.
+Object A is in `player` form and object B is in `player`-`season`-`week` form.
+When you merge them, the resulting data object will be in `player`-`season`-`week` form.
 
@@
 Here are the original data in long form.
-The data are structured in "player-season-week form".
-That is, every row in the dataset is uniquely identified by the combination of variables, `ID`, `season`, and `week`.
+The data are structured in "`player`-`season`-`week` form".
+That is, every row in the dataset is uniquely identified by the combination of variables, `ID`, `season`, and `week`—these are the keys.
 This is an example of long form, because each player has multiple rows.
@@
 
-Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "player form" (where each row is uniquely identified by the `ID` variable):
+Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable):
 

### 2024-08-30 download-football-data.qmd
20240830 - additional NFL data

@@
 
-A Data Dictionary for rosters is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_rosters.html>
+A Data Dictionary for rosters is located at the following links:
+
+- <https://nflreadr.nflverse.com/articles/dictionary_rosters.html>
+- <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#rosters>
 
@@
   nflreadr::load_rosters_weekly(seasons = TRUE))
+
+rosters <- read_csv("https://raw.githubusercontent.com/leesharpe/nfldata/master/data/rosters.csv")
 ```
@@
   nfl_rosters_raw,
+  rosters,
   file = "./data/nfl_rosters_raw.RData"
@@
 nfl_rosters <- nfl_rosters_raw %>% 
+  left_join(
+    rosters %>% select(playerid, season, team, side, category, games, starts, years, av),
+    by = c("season","team","pfr_id" = "playerid"),
+    na_matches = "never"
+  ) %>% 
   select(gsis_id, season, team, week, everything()) %>% 
@@
 load(file = "./data/nfl_schedules.RData")
+```
+
+### Standings {#sec-standings}
+
+A Data Dictionary for game schedules data is located at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#standings>
+
+```{r}
+#| eval: false
+
+nfl_standings_raw <- read_csv("http://www.habitatring.com/standings.csv")
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_standings_raw,
+  file = "./data/nfl_standings_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_standings_raw.RData")
+```
+
+The `nfl_standings` object is in `season`-`team` form.
+That is, each row should be uniquely identified by the combination of `season` and `team`.
+
+```{r}
+nfl_standings <- nfl_standings_raw
+```
+
+Let's check for duplicate `season`-`team` instances:
+
+```{r}
+nfl_standings %>% 
+  group_by(season, team) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_standings,
+  file = "./data/nfl_standings.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_standings.RData")
 ```
@@
 
+```{r}
+#| eval: false
+
+save(
+  nfl_draftPicks,
+  file = "./data/nfl_draftPicks.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_draftPicks.RData")
+```
+
+### Draft Values {#sec-draftValues}
+
+A Data Dictionary for draft values data is located at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#draft_values>
+
+```{r}
+#| eval: false
+
+nfl_draftValues_raw <- read_csv("https://raw.githubusercontent.com/leesharpe/nfldata/master/data/draft_values.csv")
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_draftValues_raw,
+  file = "./data/nfl_draftValues_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_draftValues_raw.RData")
+```
+
+The `nfl_draftValues` object is in `pick` form.
+That is, each row should be uniquely identified by `pick`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_draftValues <- nfl_draftValues_raw %>% 
+  select(pick, everything()) %>% 
+  arrange(pick)
+```
+
+Let's check for duplicate `pick` instances:
+
+```{r}
+nfl_draftValues %>% 
+  group_by(pick) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_draftValues,
+  file = "./data/nfl_draftValues.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_draftValues.RData")
+```
+
 #### Processing {#sec-processingDraftPicks}
@@
 load(file = "./data/nfl_depthCharts.RData")
+```
+
+### Trades {#sec-trades}
+
+A Data Dictionary for game schedules data is located at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#trades>
+
+```{r}
+#| eval: false
+
+nfl_trades_raw <- read_csv("https://raw.githubusercontent.com/leesharpe/nfldata/master/data/trades.csv")
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_trades_raw,
+  file = "./data/nfl_trades_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_trades_raw.RData")
+```
+
+The `nfl_trades` object is in `trade`-`player` form.
+That is, each row should be uniquely identified by the combination of `trade_id` and `pfr_id`.
+
+```{r}
+nfl_trades <- nfl_trades_raw
+```
+
+Let's check for duplicate `trade`-`player` instances:
+
+```{r}
+nfl_trades %>%
+  filter(!is.na(pfr_id)) %>% 
+  group_by(trade_id, pfr_id) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_trades,
+  file = "./data/nfl_trades.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_trades.RData")
 ```

### 2024-08-30 download-football-data.qmd
20240831 - OSF data

@@
 
-Below, we provide examples for how to download various types of NFL data.
+Below, we provide examples for how to download various types of NFL data using the `nflreadr` package.
 For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
 We save each data file after downloading it, so we can use the data in subsequent chapters.
-This chapter uses extensive use of merging to process the data for later use.
+If you have difficulty downloading the data files using the `nflreadr` package, we also saved the data files so they are publicly available on the Open Science Framework: <https://osf.io/z6pg4>.
+
+This chapter extensively uses merging to process the data for later use.
 See @sec-merging for a reminder of how to perform merging, the types of merges, and what you can expect when you merge data objects with different formats.

### 2024-08-31 base-rates.qmd
20240719 - clean up
20240830 - base rates and heuristics

--- 2024-07-15 base-rates.qmd
@@
 Because they do not pay enough attention to the base rate (in this example, the base rate of HIV is .3%).
+
+In general, people tend to overestimate the likelihood of low base-rate events.
+That is, if the base rate of an event or condition—such as schizophrenia—is low (e.g., ~0.5%), people overestimate the likelihood that a person has schizophrenia when given specific information about the person such as their symptoms and history.
 
@@
 
+Fantasy performance—and behavior more generally—is challenging to predict.
+People commonly demonstrate [biases](#sec-cognitiveBiases) and [fallacies](#sec-fallacies) when making predictions.
 People tend to ignore base rates ([base rate fallacy](#sec-fallaciesBaseRate)) when making predictions.

### 2024-08-31 cognitive-bias.qmd
20240830 - base rates and heuristics

@@
 They reduce our mental load and can help us make quick judgments to stay alive or to make complex decisions in the face of uncertainty.
+For instance, snap judgments can help us rapidly differentiate between a friend and a foe, for survival.
 However, these mental shortcuts can also lead us astray and to make systematic errors in our judgments and predictions.
@@
 
-The *affect heuristic* refers to the tendency for a person to make judgments or predictions based on their emotions and feelings—such as whether they like or dislike something—with little deliberation or reasoning [@Kahneman2011].
+The *affect heuristic* refers to the tendency for a person to make judgments or predictions based on their emotions and feelings—such as whether they like or dislike something—with little deliberation or reasoning.
+That is, the affect heuristic involves letting one's likes and dislikes determine one's beliefs about the world [@Kahneman2011].
 For instance, if a manager's favorite team is the Dallas Cowboys, they may hold a distaste for players on teams that are rivals of the Cowboys, such as the Philadelphia Eagles.
-The affect heuristic may lead them to avoid drafting Eagles players even though some Eagles players may perform well.
+The affect heuristic may lead them to believe that Eagles players are overrated and to avoid drafting Eagles players even though some Eagles players may perform well.
 
@@
 The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
+In general, people tend to overestimate the likelihood of low base-rate events.
+
 For example, the base rate of work-related injury is the general probability of experiencing a work-related injury, irrespective of other factors (e.g., the type of job, the person's age, the person's sex).

### 2024-09-01 player-evaluation.qmd
20240901 - contract hold-outs

@@
 
-In addition to [injuries](#sec-evalHealth), other factors can affect a player's availability, including league suspensions, for instance, due to violating the league's conduct policy.
+In addition to [injuries](#sec-evalHealth), other factors can affect a player's availability, including contract hold-outs or league suspensions, for instance, due to violating the league's conduct policy.
+Some players may be "holding out", which means that they may refuse to play until they sign a more desirable contract with the team (e.g., for more money or for a longer-term contract).
 

### 2024-09-01 cognitive-bias.qmd
20240901 - law of small numbers

@@
 The law of *small* numbers (aka hasty generalization), by contrast, is an erroneous belief that small samples are representative of the populations from which they were drawn.
+In general, large samples yield more precise results than small samples, whereas extreme results (in either direction—e.g., much higher or much lower than 50% of coins that flip heads-up) are more likely to be obtained from small samples [@Kahneman2011].
 For instance, if you flip a coin 10 times, belief in the law of small numbers would lead one to believe that the coin will flip heads-up exactly 5 times out of 10.
@@
 
-numHeads <- rbinom(
-  n = 100000,
-  size = 10,
+smallNumFlips <- 10
+largeNumFlips <- 1000000
+numSimulations <- 100000
+
+numHeadsSmallNumFlips <- rbinom(
+  n = numSimulations,
+  size = smallNumFlips,
   prob = .5
@@
 
-simulationOfFlipping10Coins <- data.frame(
-  numHeads = numHeads
+numHeadsLargeNumFlips <- rbinom(
+  n = numSimulations,
+  size = largeNumFlips,
+  prob = .5
 )
 
-simulationOfFlipping10Coins <- simulationOfFlipping10Coins %>% 
+simulationOfFlippingCoins <- data.frame(
+  numHeadsSmallNumFlips = numHeadsSmallNumFlips,
+  numHeadsLargeNumFlips = numHeadsLargeNumFlips
+)
+
+simulationOfFlippingCoins <- simulationOfFlippingCoins %>% 
   mutate(
-    highlight = ifelse(numHeads == 5, "yes", "no")
+    proportionHeadsSmallNumFlips = numHeadsSmallNumFlips / smallNumFlips,
+    proportionHeadsLargeNumFlips = numHeadsLargeNumFlips / largeNumFlips,
+    highlight = ifelse(numHeads == 5, "yes", "no"),
+    extremeSmallNumFlips = ifelse(proportionHeadsSmallNumFlips <= 0.2 | proportionHeadsSmallNumFlips >= 0.8, TRUE, FALSE),
+    extremeLargeNumFlips = ifelse(proportionHeadsLargeNumFlips <= 0.2 | proportionHeadsLargeNumFlips >= 0.8, TRUE, FALSE),
   )
@@
 ggplot2::ggplot(
-  data = simulationOfFlipping10Coins,
+  data = simulationOfFlippingCoins,
   mapping = aes(
-    x = numHeads,
+    x = numHeadsSmallNumFlips,
     fill = highlight)
@@
 ) %>% sum()
+```
+
+Moreover, among the simulations with only 10 coin flips, nearly 11% of sets had extreme results (≤ 20% heads or ≥ 80% heads):
+
+```{r}
+prop.table(table(simulationOfFlippingCoins$extremeSmallNumFlips))
+```
+
+By contrast, among the simulations with 1 million coin flips, none of the sets had extreme results:
+
+```{r}
+prop.table(table(simulationOfFlippingCoins$extremeLargeNumFlips))
 ```

### 2024-09-02 cognitive-bias.qmd
20240902 - fix error

@@
     proportionHeadsLargeNumFlips = numHeadsLargeNumFlips / largeNumFlips,
-    highlight = ifelse(numHeads == 5, "yes", "no"),
+    highlight = ifelse(numHeadsSmallNumFlips == 5, "yes", "no"),
     extremeSmallNumFlips = ifelse(proportionHeadsSmallNumFlips <= 0.2 | proportionHeadsSmallNumFlips >= 0.8, TRUE, FALSE),

### 2024-09-02 cognitive-bias.qmd
20240902 - anchoring and adjustment

@@
 The *anchoring and adjustment heuristic* refers to the tendency for a person's judgments or predictions to be made with a reference point—an anchor—as a starting point from which they adjust their estimates upward or downward.
-The anchor is often inaccurate and given too much weight in the person's calculation, and too little adjustment is made to the anchor.
-For instance, a manager is trying to predict how many fantasy points a top Running Back may score.
+The anchor provides a suggestion to the person—even without their paying attention, leading to a priming effect which evokes information that is compatible with the suggestion (i.e., a suggestive influence).
+We are more suggestible than we would like to think we are.
+The anchor is often inaccurate and given too much weight in the person's calculation, and too little adjustment is made away from the anchor.
+As an example, insufficient adjustment from an anchor may lead someone to drive too fast after exiting a highway [@Kahneman2011].
+The anchoring effect occurs even when the anchor (i.e., suggestion) is a random number and is completely uninformative of the correct answer [@Kahneman2011]!
+
+Applied to fantasy football, a manager is trying to predict how many fantasy points a top Running Back may score.
 The player scored 300 fantasy points last season, but the team added a stronger backup Running Back and changed the Offensive Coordinator to be a more pass-heavy offense.
@@
 [Regression effects](#sec-fallaciesRegression) are discussed further in @sec-fallaciesRegression.
+
+To counter the effect of the anchoring heuristic, @Kahneman2011 suggests deliberately thinking the opposite of the anchor.
+And, if a person you are negotiating with makes an outrageous initial offer, instead of making an equally outrageous counter-offer, @Kahneman2011 suggests storming out or threatening to leave, and indicating that you are unwilling to negotiate with that number on the table.
+This technique could be helpful during trade negotiations for players in fantasy football.
+Other techniques include, instead of focusing on the anchor, focusing on the minimal offer that the opponent would accept, or on the costs to the opponent of failing to reach an agreement [@Kahneman2011].
 

### 2024-09-04 download-football-data.qmd
20240904 - data update before start of 2024 season

@@
 ```{r}
-library("ffanalytics")
-library("petersenlab")
+library("ffanalytics") # to install: install.packages("remotes"); remotes::install_github("FantasyFootballAnalytics/ffanalytics")
+library("petersenlab") # to install: install.packages("remotes"); remotes::install_github("DevPsyLab/petersenlab")
 library("nflreadr")
@@
 
+#### Processing {#sec-processingDraftPicks}
+
+Let's do some data cleanup:
+
+```{r}
+# Convert missing values to NA
+nfl_draftPicks[nfl_draftPicks == ""] <- NA
+
+# Drop players with missing values for gsis_id
+nfl_draftPicks <- nfl_draftPicks %>% 
+  filter(!is.na(gsis_id))
+```
+
+Let's check again for duplicate `player` instances:
+
+```{r}
+nfl_draftPicks %>% 
+  group_by(gsis_id) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
 ```{r}
@@
 load(file = "./data/nfl_draftValues.RData")
-```
-
-#### Processing {#sec-processingDraftPicks}
-
-Let's do some data cleanup:
-
-```{r}
-# Convert missing values to NA
-nfl_draftPicks[nfl_draftPicks == ""] <- NA
-
-# Drop players with missing values for gsis_id
-nfl_draftPicks <- nfl_draftPicks %>% 
-  filter(!is.na(gsis_id))
-```
-
-Let's check again for duplicate `player` instances:
-
-```{r}
-nfl_draftPicks %>% 
-  group_by(gsis_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_draftPicks,
-  file = "./data/nfl_draftPicks.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_draftPicks.RData")
 ```
@@
 
+# Seasonal Data
 nfl_advancedStatsPFR_pass_seasonal_raw <- progressr::with_progress(
@@
 
+nfl_advancedStatsPFR_rush_seasonal_raw <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("rush"),
+    summary_level = c("season")))
+
+nfl_advancedStatsPFR_rec_seasonal_raw <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("rec"),
+    summary_level = c("season")))
+
+nfl_advancedStatsPFR_def_seasonal_raw <- progressr::with_progress(
+  nflreadr::load_pfr_advstats(
+    seasons = TRUE,
+    stat_type = c("def"),
+    summary_level = c("season")))
+
+# Weekly Data
 nfl_advancedStatsPFR_pass_weekly_raw <- progressr::with_progress(
@@
 
-nfl_advancedStatsPFR_rush_seasonal_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("rush"),
-    summary_level = c("season")))
-
 nfl_advancedStatsPFR_rush_weekly_raw <- progressr::with_progress(
@@
 
-nfl_advancedStatsPFR_rec_seasonal_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("rec"),
-    summary_level = c("season")))
-
 nfl_advancedStatsPFR_rec_weekly_raw <- progressr::with_progress(
@@
     summary_level = c("week")))
-
-nfl_advancedStatsPFR_def_seasonal_raw <- progressr::with_progress(
-  nflreadr::load_pfr_advstats(
-    seasons = TRUE,
-    stat_type = c("def"),
-    summary_level = c("season")))
 
@@
 
+# Seasonal Projections
 players_projections_seasonal_raw <- ffanalytics::scrape_data(
@@
   week = 0) # NULL grabs seasonal projections; NULL grabs the current week
+
+# Weekly Projections
+players_projections_weekly_raw <- ffanalytics::scrape_data(
+  season = NULL, # NULL grabs the current season
+  week = NULL) # NULL grabs the current week
 ```
@@
 )
+
+save(
+  players_projections_weekly_raw,
+  file = "./data/players_projections_weekly_raw.RData"
+)
 ```
@@
 load(file = "./data/players_projections_seasonal_raw.RData")
-```
-
-The `players_projections_seasonal_raw` object is in `player`-`position`-`projection source` form.
+load(file = "./data/players_projections_weekly_raw.RData")
+```
+
+The `players_projections_seasonal_raw` and `players_projections_weekly_raw` object is in `player`-`position`-`projection source` form.
 That is, each row should be uniquely identified by the combination of `id`, `pos`  and `data_src`.
@@
   head()
+
+players_projections_weekly_raw %>% 
+  bind_rows() %>% 
+  select(id, pos, data_src, everything()) %>%
+  arrange(player, id, pos, data_src) %>% 
+  group_by(id, pos, data_src) %>% 
+  filter(n() > 1, !is.na(id)) %>% 
+  head()
+
+players_projections_weekly_raw %>% 
+  bind_rows() %>% 
+  select(player, pos, data_src, everything()) %>%
+  arrange(player, id, pos, data_src) %>% 
+  group_by(id, pos, data_src) %>% 
+  filter(n() > 1, !is.na(id)) %>% 
+  head()
 ```
@@
 
-Now, modify the scoring settings to match our league settings:
+Now, modify the scoring settings to match our league settings (based on [scoring settings](https://support.nfl.com/hc/en-us/articles/4989179237404-Scoring) for fantasy leagues on NFL.com):
 
@@
 
+# Offense
 scoring_obj$pass$pass_int <- -2
+scoring_obj$rec$rec <- 1
 scoring_obj$misc$fumbles_lost <- -2
+
+# Kickers
+scoring_obj$kick$fg_4049 <- 3
+
+# Defense/Special Teams
+scoring_obj$pts_bracket[[3]]$threshold <- 13
+scoring_obj$pts_bracket[[3]]$points <- 4
+
+scoring_obj$pts_bracket[[4]]$threshold <- 20
+scoring_obj$pts_bracket[[4]]$points <- 1
+
+scoring_obj$pts_bracket[[5]]$threshold <- 27
+scoring_obj$pts_bracket[[5]]$points <- 0
+
+scoring_obj$pts_bracket[[6]] <- list(
+  threshold = 34,
+  points = -1
+)
+
+scoring_obj$pts_bracket[[7]] <- list(
+  threshold = 99,
+  points = -4
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+# My Yahoo League
+scoring_obj <- scoring_obj_default
+
+scoring_obj$pass$pass_int <- -3
+scoring_obj$misc$fumbles_lost <- -3
 scoring_obj$idp$idp_fum_force <- -2
@@
   scoring_rules = scoring_obj)
+
+players_projectedPoints_weekly <- ffanalytics:::impute_and_score_sources(
+  data_result = players_projections_weekly_raw,
+  scoring_rules = scoring_obj)
 ```
@@
 ```{r}
+# Seasonal Projections
 players_projectedStatsAverage_seasonal <- ffanalytics::projections_table(
@@
   return_raw_stats = FALSE)
-```
-
-The `players_projectedPoints_seasonal`, `players_projectedStatsAverage_seasonal`, and `players_projectedPointsAverage_seasonal` objects are in `player`-`average type`-`position` form.
+
+# Weekly Projections
+players_projectedStatsAverage_weekly <- ffanalytics::projections_table(
+  players_projections_weekly_raw,
+  scoring_rules = scoring_obj,
+  return_raw_stats = TRUE)
+
+players_projectedPointsAverage_weekly <- ffanalytics::projections_table(
+  players_projections_weekly_raw,
+  scoring_rules = scoring_obj,
+  return_raw_stats = FALSE)
+```
+
+The `players_projectedPoints_seasonal`, `players_projectedStatsAverage_seasonal`, `players_projectedPointsAverage_seasonal`, `players_projectedPoints_weekly`, `players_projectedStatsAverage_weekly`, and `players_projectedPointsAverage_weekly` objects are in `player`-`average type`-`position` form.
 That is, each row should be uniquely identified by the combination of `id`, `avg_type`  and `pos` (or `position`).
@@
   head()
+
+players_projectedPoints_weekly %>% 
+  bind_rows() %>% 
+  select(id, pos, data_src, everything()) %>%
+  arrange(player, id, pos, data_src) %>% 
+  group_by(id, pos, data_src) %>% 
+  filter(n() > 1, !is.na(id)) %>% 
+  head()
+
+players_projectedStatsAverage_weekly %>% 
+  group_by(id, avg_type, position) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+players_projectedPointsAverage_weekly %>% 
+  group_by(id, avg_type, pos) %>% 
+  filter(n() > 1) %>% 
+  head()
 ```
@@
 )
+
+players_projections_weekly_average <- full_join(
+  players_projectedPointsAverage_weekly,
+  players_projectedStatsAverage_weekly,
+  by = c("id","avg_type","pos" = "position")
+)
 ```
@@
 players_projections_seasonal_average %>% 
+  group_by(id, avg_type, pos) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+players_projections_weekly_average %>% 
   group_by(id, avg_type, pos) %>% 
@@
   add_player_info()
-```
-
-```{r}
-#| eval: false
-
-save(
-  players_projectedPoints_seasonal,
+
+players_projections_weekly_average <- players_projections_weekly_average %>% 
+  add_ecr() %>% 
+  #add_uncertainty() %>% # currently throws an error
+  add_player_info()
+```
+
+```{r}
+#| eval: false
+
+save(
+  players_projectedPoints_seasonal, players_projections_seasonal_average,
   file = "./data/players_projectedPoints_seasonal.RData"
@@
 save(
-  players_projections_seasonal_average,
-  file = "./data/players_projections_seasonal_average.RData"
+  players_projectedPoints_weekly, players_projections_weekly_average,
+  file = "./data/players_projectedPoints_weekly.RData"
 )
@@
 load(file = "./data/players_projectedPoints_seasonal.RData")
-load(file = "./data/players_projections_seasonal_average.RData")
+load(file = "./data/players_projectedPoints_weekly.RData")
 ```
@@
   player_game_birthdate_gameday_offense,
-  nfl_players %>% select(-birth_date, -season, -team_abbr, - team_seq),
+  nfl_players %>% select(-birth_date, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id"))
@@
   player_game_birthdate_gameday_defense,
-  nfl_players %>% select(-birth_date, -season, -team_abbr, - team_seq),
+  nfl_players %>% select(-birth_date, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id"))
@@
   player_game_birthdate_gameday_kicking,
-  nfl_players %>% select(-birth_date, -season, -team_abbr, - team_seq),
+  nfl_players %>% select(-birth_date, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id"))
@@
   nfl_actualStats_offense_seasonal,
-  nfl_players %>% select(-position, -position_group, -season, -team_abbr, - team_seq),
+  nfl_players %>% select(-position, -position_group, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id")
@@
   nfl_actualStats_defense_seasonal,
-  nfl_players %>% select(-position, -position_group, -season, -team_abbr, - team_seq),
+  nfl_players %>% select(-position, -position_group, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id")
@@
   nfl_actualStats_kicking_seasonal,
-  nfl_players %>% select(-position, -position_group, -season, -team_abbr, - team_seq),
+  nfl_players %>% select(-position, -position_group, -team_abbr, - team_seq),
   by = c("player_id" = "gsis_id")

### 2024-09-05 download-football-data.qmd
20240904 - fix comment

@@
   season = NULL, # NULL grabs the current season
-  week = 0) # NULL grabs seasonal projections; NULL grabs the current week
+  week = 0) # 0 grabs seasonal projections
 

### 2024-09-05 download-football-data.qmd
20240905 - D/ST points bracket

@@
 # Defense/Special Teams
-scoring_obj$pts_bracket[[3]]$threshold <- 13
-scoring_obj$pts_bracket[[3]]$points <- 4
-
-scoring_obj$pts_bracket[[4]]$threshold <- 20
-scoring_obj$pts_bracket[[4]]$points <- 1
-
-scoring_obj$pts_bracket[[5]]$threshold <- 27
-scoring_obj$pts_bracket[[5]]$points <- 0
-
-scoring_obj$pts_bracket[[6]] <- list(
-  threshold = 34,
-  points = -1
-)
-
-scoring_obj$pts_bracket[[7]] <- list(
-  threshold = 99,
-  points = -4
+scoring_obj$pts_bracket <- list(
+  list(threshold = 0, points = 10),
+  list(threshold = 6, points = 7),
+  list(threshold = 13, points = 4),
+  list(threshold = 20, points = 1),
+  list(threshold = 27, points = 0),
+  list(threshold = 34, points = -1),
+  list(threshold = 99, points = -4)
 )

### 2024-09-06 base-rates.qmd
20240906 - base rates

@@
 In sum, the [marginal probability](#sec-baseRate), including the [prior probability](#sec-baseRate) or [base rate](#sec-baseRate), should be weighed heavily in predictions unless there are sufficient data to indicate otherwise, i.e., to update the posterior probability based on new evidence.
-Bayes' theorem provides a powerful tool to anchor predictions to the [base rate](#sec-baseRate) unless sufficient evidence changes the posterior probability (by updating the evidence and [prior probability](#sec-baseRate)).
+Bayes' theorem specifies how prior beliefs (i.e., [base rate](#sec-baseRate) informations) should be integrated with the [predictive accuracy](#sec-predictiveValidity) of the evidence to make predictions.
+It thus provides a powerful tool to [anchor](#sec-heuristicsAnchoringAdjustment) predictions to the [base rate](#sec-baseRate) unless sufficient evidence changes the posterior probability (by updating the evidence and [prior probability](#sec-baseRate)).
+In general, you should [anchor](#sec-heuristicsAnchoringAdjustment) your predictions to the [base rate](#sec-baseRate) and [adjust](#sec-heuristicsAnchoringAdjustment) from there.
+As noted by @Kahneman2011, if you have doubts about the quality of the evidence for a particular prediction question, keep your predictions close to the [base rate](#sec-baseRate), and modify them only modifying them based on the new information.
 

### 2024-09-06 cognitive-bias.qmd
20240906 - base rates

@@
 The *representativess heuristic* refers to the tendency for a person's judgments or predictions about individuals to be made based on how similar the individual is to (i.e., how closely the individual resembles) the person's existing mental prototypes.
-For instance, when coming out of college, Tight End Kyle Pitts [drew comparisons](https://247sports.com/article/kyle-pitts-lebron-james-2021-nfl-draft-florida-gators-football-163882176) to the "LeBron James" of Tight Ends (archived at <https://perma.cc/JQB5-XPVL>).
+In this way, the representativeness heuristic involves stereotyping, which is not a statistically optimal approach to predictions [@Kahneman2011].
+Moreover, the representativness heuristic does not adequately consider the quality of the evidence [@Kahneman2011].
+It is thus important to question the [predictive accuracy](#sec-predictiveValidity) of the evidence for the particular prediction question.
+
+Consider the following example of the representativeness heuristic.
+When coming out of college, Tight End Kyle Pitts [drew comparisons](https://247sports.com/article/kyle-pitts-lebron-james-2021-nfl-draft-florida-gators-football-163882176) to the "LeBron James" of Tight Ends (archived at <https://perma.cc/JQB5-XPVL>).
 The idea that his athletic profile leads him to be similar to the prototype of LeBron James may have led him to be too highly drafted by fantasy managers in his first seasons.
@@
 The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
-In general, people tend to overestimate the likelihood of low base-rate events.
+Focusing on the specific information about the event (and thus underweighting the base rate) involves the [representativeness heuristic](#sec-heuristicsRepresentativeness) [@Kahneman2011]—making judgments based on how similar the event is to a mental prototype regardless of how frequent the event the event actually is.
+In general, people tend to overestimate the likelihood of low base-rate events [@Kahneman2011].
 

### 2024-09-06 cognitive-bias.qmd
20240906 - expand example

@@
 
-The *availability heuristic* refers to the tendency for a person's judgments or predictions about the frequency or probability of something to be made based on how readily instances can be brought to mind.
-For instance, when making fantasy predictions about a player, more recent big performance games may more easily come to mind compared to lower-scoring games and games that occurred longer ago.
+The *availability heuristic* refers to the tendency for a person to make judgments or predictions about the frequency or probability of something based on how readily instances can be brought to mind.
+For instance, when making fantasy predictions about a player, more recent big performance games (or games that receive more news coverage) may more easily come to mind compared to lower-scoring games and games that occurred longer ago.
 Thus, a manager may be more inclined to pick players to start who had more recent, stronger performances rather than players who have higher long-term averages.

### 2024-09-07 base-rates.qmd
20240907 - base rate example

@@
 
+We can then estimate the marginal probability of the event, substititing in $P(R | \text{not } C)$, using the law of total probability.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pA()` function that estimates the marginal probability of one event, $A$.
+
+```{r}
+petersenlab::pA(
+  pAgivenB = .95,
+  pB = .003,
+  pAgivenNotB = .007171515)
+```
+
 The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pBgivenNotA()` function that estimates the probability of one event, $B$, given that another event, $A$, did not occur.
@@
 As noted by @Kahneman2011, if you have doubts about the quality of the evidence for a particular prediction question, keep your predictions close to the [base rate](#sec-baseRate), and modify them only modifying them based on the new information.
+
+## Cab Example {#sec-cabExample}
+
+Below is an example:
+
+> A cab was involved in a hit-and-run accident at night.
+> Two cab companies, the Green and the Blue, operate in the city.
+> You are given the following data:
+> 
+> - 85% of the cabs in the city are Green and 15% are Blue.
+> - A witness identified the cab as Blue.
+> The court tested the reliability of the witness under the circumstances that existed on the night of the accident and concluded that the witness correctly identified each one of the two colors 80% of the time and failed 20% of the time.
+>
+> What is the probability that the cab involved in the accident was Blue rather than Green?
+> [@Kahneman2011, p. 166]
+
+Thus, we know the following:
+
+$$
+\begin{aligned}
+  P(\text{Blue}) &= .15 && \text{prior probability of a Blue cab}\\
+  P(\text{Green}) &= .85 && \text{prior probability of a Green cab}\\
+  P(\text{Correct}|\text{Blue}) &= .80 && \text{probability the witness correctly identifies a Blue cab}\\
+  P(\text{Correct}|\text{Green}) &= .80 && \text{probability the witness correctly identifies a Green cab}\\
+  P(\text{Incorrect}|\text{Blue}) &= .20 && \text{probability the witness incorrectly identifies a Blue cab}\\
+  P(\text{Incorrect}|\text{Green}) &= .20 && \text{probability the witness incorrectly identifies a Green cab}\\
+\end{aligned}
+$$
+
+We want to know the probability that the cab involved in the accident was Blue, given that the witness identified it as Blue ($P(\text{Blue}|\text{Identified as Blue})$).
+To estimate this probability, we can apply [Bayes' theorem](#sec-bayesTheorem) to estimate the posterior probability:
+
+$$
+\begin{aligned}
+  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)}\\
+  P(\text{Blue}|\text{Identified as Blue}) &= \frac{P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue})}{P(\text{Identified as Blue})}
+\end{aligned}
+$$
+
+We can compute the term in the denominator ($P(\text{Identified as Blue})$) using the law of total probability (described in @sec-bayesTheorem).
+$$
+\begin{aligned}
+  P(R_i) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i) \\
+  P(R_i) &= P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue}) + P(\text{Identified as Blue}|\text{Green}) \cdot P(\text{Green})
+  0.29 &= (.80 \times .15) + (.20 \times .85) \\
+\end{aligned}
+$$
+
+```{r}
+petersenlab::pA(
+  pAgivenB = .80,
+  pB = .15,
+  pAgivenNotB = .20)
+```
+
+We can now substitute this value into the denominator of Bayes' theorem to estimate the posterior probability:
+
+$$
+\begin{aligned}
+  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)}\\
+  P(\text{Blue}|\text{Identified as Blue}) &= \frac{P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue})}{P(\text{Identified as Blue})}\\
+  0.414 &= \frac{0.80 \times 0.15}{0.29}
+\end{aligned}
+$$
+
+```{r}
+petersenlab::pAgivenB(
+  pBgivenA = .80,
+  pA = .15,
+  pB = .29)
+```
+
+Thus, there was a 41.4% probability that the car involved in the accident was Blue rather than Green.
+However, when faced with this problem, people tend to ignore the base rate and go with the witness [@Kahneman2011].
+According to @Kahneman2011, the most frequent response to this question regarding is that there is an 80% that the car was Blue.
 

### 2024-09-07 cognitive-bias.qmd
20240907 - base rate example
20240907 - updates

@@
 In this way, the representativeness heuristic involves stereotyping, which is not a statistically optimal approach to predictions [@Kahneman2011].
-Moreover, the representativness heuristic does not adequately consider the quality of the evidence [@Kahneman2011].
+Stereotypes are statements about groups or categories that are at least tentatively accepted as facts about every member of that group or category [@Kahneman2011].
+That is, a (perhaps common) characteristic of a group is attributed to each member of that group.
+In addition, the representativness heuristic does not adequately consider the quality of the evidence [@Kahneman2011].
 It is thus important to question the [predictive accuracy](#sec-predictiveValidity) of the evidence for the particular prediction question.
@@
 For instance, people tend not to consider other relevant, important factors that they do not have immediate access to, consistent with the [WYSIATI heuristic](#sec-heuristicsWYSIATI).
+Instead, people tend to focus on the cohesiveness and plausibility of the story, which differs from its probability [@Kahneman2011].
 

### 2024-09-07 base-rates.qmd
20240907 - updates
20240907 - copyedit

@@
 \end{aligned}
-$$
+$$ {#eq-cabExampleGivenInfo}
 
@@
 \end{aligned}
-$$
+$$ {#eq-cabExampleBayesTheorem}
 
@@
 \begin{aligned}
-  P(R_i) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i) \\
-  P(R_i) &= P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue}) + P(\text{Identified as Blue}|\text{Green}) \cdot P(\text{Green})
+  P(R_i) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i)\\
+  P(R_i) &= P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue}) + P(\text{Identified as Blue}|\text{Green}) \cdot P(\text{Green})\\
   0.29 &= (.80 \times .15) + (.20 \times .85) \\
 \end{aligned}
-$$
+$$ {#eq-cabExampleLawOfTotalProbability}
 
@@
 \end{aligned}
-$$
+$$ {#eq-cabExamplePosterior}
 
@@
 Thus, there was a 41.4% probability that the car involved in the accident was Blue rather than Green.
-However, when faced with this problem, people tend to ignore the base rate and go with the witness [@Kahneman2011].
+However, when faced with this problem, people tend to [ignore the base rate](#sec-fallaciesBaseRate) and go with the witness [@Kahneman2011].
 According to @Kahneman2011, the most frequent response to this question regarding is that there is an 80% that the car was Blue.

### 2024-09-08 actuarial.qmd
20270907 - citations in block quotes

@@
 
-> "If a sociologist were predicting whether Professor X would go to the movies on a certain night, he might have an equation involving age, academic specialty, and introversion score.
+> If a sociologist were predicting whether Professor X would go to the movies on a certain night, he might have an equation involving age, academic specialty, and introversion score.
 > The equation might yield a probability of .90 that Professor X goes to the movie tonight.
@@
 > Because broken legs are very rare, and in the sociologist's entire sample of 500 criterion cases plus 250 cross-validating cases, he did not come upon a single instance of it.
-> He uses the broken leg datum confidently, because 'broken leg' is a subclass of a larger class we may crudely denote as 'relatively immobilizing illness or injury,' and movie-attending is a subclass of a larger class of 'actions requiring moderate mobility.' "
-> [@Meehl1957, p. 269–270]
+> He uses the broken leg datum confidently, because 'broken leg' is a subclass of a larger class we may crudely denote as 'relatively immobilizing illness or injury,' and movie-attending is a subclass of a larger class of 'actions requiring moderate mobility.'
+> 
+> --- Meehl [-@Meehl1957, p. 269–270]
 

### 2024-09-08 data-visualization.qmd
20240814 - formatting
20240824 - data viz books
20270907 - citations in block quotes
20240907 - ridgeline plots

@@
 library("gghighlight")
+library("ggridges")
 library("tidyverse")
@@
 > 6. Graphics must not quote data out of context.
-> – Tufte (2001, p. 77)
+> 
+> --- Tufte [-@Tufte2001, p. 77]
 
@@
 The R Graph Gallery provides examples of various types of plots: <https://r-graph-gallery.com>.
+Books on data visualization in `R` include [*`ggplot2`: Elegant Graphics for Data Analysis*](https://ggplot2-book.org) [@ggplot3rdEd] and [*`R` Graphics Cookbook: Practical Recipes for Visualizing Data*](https://r-graphics.org) [@Chang2018].
 In this chapter, we will examine how to create statistical graphics to visualize data.
@@
 
-A violin plot of fantasy points is depicted in @fig-boxAndWhisker.
+A violin plot of fantasy points is depicted in @fig-violin.
 
@@
     y = "Fantasy Points",
-    title = "Violin Plot by Position",
+    title = "Violin Plot of Fantasy Points by Position",
     subtitle = "Lines represent the 25th, 50th, and 75th quantiles"
@@
   theme(legend.position = "none")
+```
+
+### Ridgeline Plot {#sec-ridgelinePlot}
+
+A ridgeline plot of fantasy points is depicted in @fig-ridgeline1.
+
+```{r}
+#| label: fig-ridgeline1
+#| fig-cap: "Ridgeline Plot."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasy_points,
+    y = position_group,
+    group = position_group,
+    fill = position_group)
+) +
+  geom_density_ridges(
+    rel_min_height = 0.0085, # remove trailing tails
+  ) +
+  labs(
+    x = "Fantasy Points",
+    y = "Position",
+    title = "Ridgeline Plot of Fantasy Points by Position"
+  ) +
+  theme_classic() + 
+  theme(legend.position = "none")
+```
+
+We can add lines at the quartiles, as depicted in @fig-ridgeline2.
+
+```{r}
+#| label: fig-ridgeline2
+#| fig-cap: "Ridgeline Plot. Lines represent the 25th, 50th, and 75th quantiles."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasy_points,
+    y = position_group,
+    group = position_group,
+    fill = factor(stat(quantile)))
+) +
+  stat_density_ridges(
+    geom = "density_ridges_gradient",
+    calc_ecdf = TRUE,
+    quantiles = 4,
+    quantile_lines = TRUE
+  ) +
+  # scale_fill_viridis_d() # if want to use viridis color scheme
+  labs(
+    x = "Fantasy Points",
+    y = "Position",
+    title = "Ridgeline Plot by Position",
+    subtitle = "Lines represent the 25th, 50th, and 75th quantiles",
+    fill = "Quartiles"
+  ) +
+  theme_classic()
 ```

### 2024-09-08 data-visualization.qmd
20240907 - specify package

@@
 ) +
-  geom_density_ridges(
+  ggridges::geom_density_ridges(
     rel_min_height = 0.0085, # remove trailing tails
@@
 ) +
-  stat_density_ridges(
+  ggridges::stat_density_ridges(
     geom = "density_ridges_gradient",

### 2024-09-08 base-rates.qmd
20270907 - citations in block quotes
20240907 - base rate in fantasy football

@@
 > What is the probability that the cab involved in the accident was Blue rather than Green?
-> [@Kahneman2011, p. 166]
+> 
+> --- Kahneman [-@Kahneman2011, p. 166]
 
@@
 According to @Kahneman2011, the most frequent response to this question regarding is that there is an 80% that the car was Blue.
+
+## Base Rates Applied to Fantasy Football {#sec-baseRateFantasyFootball}
+
+Base rates are also relevant to fantasy football.
+Unlike yardage (e.g., passing yards, rushing yards, receiving yards), touchdowns occur *relatively* less frequently.
+Whereas a solid Wide Receiver may log 100+ receptions and 1,200+ yards in a season, they may have "only" 8–14 receiving touchdowns in a given season.
+As noted in @sec-differentErrorsDifferentCosts, lower base-rate events—including touchdowns—are harder to predict accurately.
+As noted by [Harris (2012)](https://www.espn.com/fantasy/football/ffl/story?page=nfldk2k12_vbdwork): "NFL statistical projections are basically impossible to get right.
+(Take it from someone who helps create them for a living.) Yes, we can do a passable job with yardage totals for players who don't suffer unexpected injuries or depth-chart pratfalls.
+But so much of fantasy football hinges on touchdowns, and touchdowns are impossibly difficult to predict from season to season (let alone week to week)." (archived at <https://perma.cc/4QNH-J2LD>).
+Thus, it is important not to lend too much credence to predictions of touchdowns.
+Focus on other things that may be more predictable (and that may be indirectly prognostic of touchdowns) such as yards, carries/targets, receptions, depth of targets, red zone carries/targets, short distance carries/targets, etc.
 

### 2024-09-08 data-visualization.qmd
20240908 - updates

@@
     group = position_group,
-    fill = factor(stat(quantile)))
+    fill = factor(after_stat(quantile)))
 ) +
   ggridges::stat_density_ridges(
+    rel_min_height = 0.0085, # remove trailing tails
     geom = "density_ridges_gradient",
@@
   ) +
-  # scale_fill_viridis_d() # if want to use viridis color scheme
+   scale_fill_viridis_d() + # use viridis color scheme
   labs(
@@
     y = "Position",
-    title = "Ridgeline Plot by Position",
-    subtitle = "Lines represent the 25th, 50th, and 75th quantiles",
-    fill = "Quartiles"
+    title = "Ridgeline Plot of Fantasy Points by Position",
+    subtitle = "Vertical lines represent the 25th, 50th, and 75th quantiles",
+    fill = "Quartile"
   ) +

### 2024-09-09 cognitive-bias.qmd
20240908 - regression to the mean

@@
 The regression fallacy is the failure to account for the fact that things tend to naturally fluctuate around their mean and that, after an extreme fluctuation, subsequent scores tend to regress (or reverse) to the mean.
+Regression to the mean occurs because performance—both success and failure—is influenced by talent and by luck [@Kahneman2011].
+And, unlike talent (or one's level one the construct of interest), luck is fickle.
+Regression to the mean can explain why those with depression tend to improve over time without doing anything, and can explain why highly intelligent women tend to marry men who are less intelligent than they are [@Kahneman2011].
+Despite peoples' tendency to ascribe such changes to causal phenomena, regression-related changes do not need a causal explanation—they are a mathematically inevitable consequence that arises from the fact that an outcome is influenced, to some degree, both by one's level on the construct and by error [@Kahneman2011].
+
 An example of the regression fallacy is the so-called Sports Illustrated or Madden cover jinx curse.
 The Sports Illustrated or Madden cover jinx curse is the urban legend that players who appear on the cover of Sports Illustrated (the magazine) or Madden (the video game) will perform poorly.
-But, such a phenomenon can be more simply explained by regression to the mean (<https://www.psychologytoday.com/us/blog/what-the-luck/201610/the-sports-illustrated-cover-jinx>; archived at <https://perma.cc/CZM9-TVFN>).
+But, such a phenomenon can be more simply explained by regression to the mean (<https://www.psychologytoday.com/us/blog/what-the-luck/201610/the-sports-illustrated-cover-jinx>; archived at <https://perma.cc/CZM9-TVFN>) [@Kahneman2011].
 When a player has a superb season, they likely benefited from some degree to good luck, and it is unlikely that they will repeat such a stellar season the following year.

### 2024-09-09 getting-started.qmd
20240909 - assignment operator

@@
 It has one dimension (length).
-To create a vector, use the `c()` to combine elements into a vector:
+To create a vector, use the `c()` to combine elements into a vector.
+And, we use the assignment operator (`<-`) to assign the vector to an object named `exampleVector`, so we can access it later.
 
@@
 
+We can then access the contents of the object by calling its name:
+
+```{r}
+exampleVector
+```
+
 ## Create a Data Frame {#sec-createDF}
@@
 A data frame has two dimensions: rows and columns.
-Here is an example of creating a data frame:
+Here is an example of creating a data frame, while using the assignment operator (`<-`) to assign the data frame to an object so we can access it later:
 
@@
 
-fantasyPoints_weekly = expand.grid(
+fantasyPoints_weekly <- expand.grid(
   ID = 1:12,

### 2024-09-09 getting-started.qmd
20240909 - more info on merging/joins

@@
 Merging (also called joining) merges two data objects using a shared set of variables called "keys."
-The keys are the variable(s) that uniquely identify each row (i.e., they account for the levels of nesting).
-In some data objects, the key might be the player's identification number (e.g., `player_id`).
-However, some data objects have multiple keys.
-For instance, in long form data objects, each participant may have multiple rows corresponding to multiple seasons.
-In this case, the keys may be `player_id` and `season`—that is, the data are in `player`-`season` form.
-If a participant has multiple rows corresponding to seasons and games/weeks, the keys are `player_id`, `season`, and `week`—in this case, the data are in `player`-`season`-`week` form.
+The keys are the variable(s) that are used to align the rows from the two objects.
+The data for the given key(s) in the first object get paired with (i.e., get placed in the same row as) the data for that same key in the second object.
 In general, each row should have a value on each of the keys; there should be no missingness in the keys.
-
 To merge two objects, the key(s) that will be used to match the records must be present in both objects.
@@
 Different merge types select different rows to merge.
+
+For some data objects, you might want to combine information for the same player from multiple data objects.
+If each data object is in `player` form (i.e., `player_id` uniquely identifies each row), you might merge by the player's identification number (e.g., `player_id`).
+In this case, the key uniquely identifies each row.
+
+However, some data objects have multiple keys.
+For instance, in long form data objects, each player may have multiple rows corresponding to multiple seasons.
+In this case, the keys may be `player_id` and `season`—that is, the data are in `player`-`season` form.
+If object 1 and object 2 are both in `player`-`season` form, we would use `player_id` and `season` as the keys to merge the two objects.
+In this case, the keys uniquely identify each row; that is, they account for the levels of nesting.
+
+However, if the data objects are of different form, we would select the keys as the variable(s) that represent the lowest common denominator of variables used to join the data objects that are present in both objects.
+For instance, assume that object 1 is in `player`-`season` form.
+For object 2, each player has multiple rows corresponding to seasons and games/weeks—in this case, object 2 is in `player`-`season`-`week` form.
+Object 1 does not have the `week` variable, so it cannot be used to join the objects.
+Thus, we would use `player_id` and `season` as the keys to merge the two objects, because both variables are present in both objects.
+
+It is important not to have rows with duplicate values on the keys.
+For instance, if there is more than one row with the same `player_id` in each object (or multiple rows in object 2 with the same combination of `player_id`, `season`, and `week`), then each row with that `player_id` in object 1 gets paired with each row with that `player_id` in object 2.
+The many possible combinations can lead to the resulting object greatly expanding in terms of the number of rows.
+Thus, you want the keys to uniquely identify each row.
+In the example below, `player` is present in each object, so we can merge by `player`; however, each object has multiple rows with the same player.
+For example, `mergeExample1A` has three rows for `player` A; `mergeExample1B` has two rows for `player` A.
+Thus, when we merge them, the resulting object has many more rows than each respective object (even though neither object has players that the other object does not).
+
+```{r}
+mergeExample1A <- data.frame(
+  player = c("A","A","A","B","B"),
+  age = c(20,22,24,26,28)
+)
+
+mergeExample1B <- data.frame(
+  player = c("A","A","B","B"),
+  points = c(10,15,20,25)
+)
+
+mergeExample1 <- full_join(
+  mergeExample1A,
+  mergeExample1B,
+  by = "player")
+
+mergeExample1
+
+dim(mergeExample1)
+```
 
@@
 To prevent this, either remove or rename the shared variable in one of the objects, or include the shared variable as a key.
-However, as described above, you should include it as a key ***only*** if it uniquely identifies each row in terms of levels of nesting.
+However, as described above, you should include it as a key ***only*** if you want to use its values to align the rows from each object.
+Below is an example of merging two objects with the same variable name (i.e., points) that is not used as a key.
+
+```{r}
+mergeExample2A <- data.frame(
+  player = c("A","B","C","D","E"),
+  points = c(20,22,24,26,28)
+)
+
+mergeExample2B <- data.frame(
+  player = c("A","B","C","F"),
+  points = c(10,15,20,25)
+)
+
+mergeExample2 <- full_join(
+  mergeExample2A,
+  mergeExample2B,
+  by = "player")
+
+mergeExample2
+```
 
@@
 When you merge them, the resulting data object will be in `player`-`season`-`week` form.
+
+```{r}
+mergeExample3A <- data.frame(
+  player = c("A","B","C","D","E"),
+  weight = c(225,250,275,300,325)
+)
+
+mergeExample3B <- data.frame(
+  player = c("A","A","B","B"),
+  season = c(2023,2023,2024,2024,2024,2024),
+  week = c(1,2,1,2,3,4)
+  points = c(10,15,20,25,30,35)
+)
+
+mergeExample3 <- full_join(
+  mergeExample3A,
+  mergeExample3B,
+  by = "player")
+
+mergeExample3
+```
 

### 2024-09-09 data-visualization.qmd
20240909 - horizontal y-axis title

--- 2024-09-08 data-visualization.qmd
@@
   ) +
-  theme_classic()
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
 ```
@@
   theme_classic() + 
-  theme(legend.position = "none")
+  theme(
+    legend.position = "none",
+    axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
 ```
@@
     group = position_group,
-    fill = factor(stat(quantile)))
-) +
-  stat_density_ridges(
+    fill = factor(after_stat(quantile)))
+) +
+  ggridges::stat_density_ridges(
+    rel_min_height = 0.0085, # remove trailing tails
     geom = "density_ridges_gradient",
@@
     y = "Position",
-    title = "Ridgeline Plot by Position",
-    subtitle = "Lines represent the 25th, 50th, and 75th quantiles",
-    fill = "Quartiles"
-  ) +
-  theme_classic()
+    title = "Ridgeline Plot of Fantasy Points by Position",
+    subtitle = "Vertical lines represent the 25th, 50th, and 75th quantiles",
+    fill = "Quartile"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
 ```
@@
   theme_classic() + 
-  theme(legend.position = "none")
+  theme(legend.position = "none") +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
 ```
@@
     stat = "identity") +
-  scale_fill_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Dark2") +
   geom_errorbar(
@@
   theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) + # horizontal y-axis title
   scale_y_reverse()

### 2024-09-09 data-visualization.qmd
20240909 - make border color the same as fill color

@@
     y = mean,
-    fill = position_group,
-    color = position_group
+    fill = position_group
     )

### 2024-09-10 base-rates.qmd
20240908 - updates
20240910 - nomogram plots
20240910 - probability nomogram

@@
 
+[PROVIDE ACCURACY OF PROJECTIONS OF TOUCHDOWNS VS YARDAGE]
+
 ## Base Rate of Rookie Performance {#sec-baseRateRookiePerformance}
@@
 The calculation of posttest probability is described in INSERT.
-A *probability nomogram* is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the [pretest (or prior) probability](#sec-baseRate) and [likelihood ratio](#sec-diagnosticLikelihoodRatio), as depicted in @fig-probabilityNomogram.
+A *probability nomogram* (aka Fagan nomogram) is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the [pretest (or prior) probability](#sec-baseRate) and [likelihood ratio](#sec-diagnosticLikelihoodRatio), as depicted in @fig-probabilityNomogram.
 To use a probability nomogram, connect the dots from the starting probability (left line) with the [likelihood ratio](#sec-diagnosticLikelihoodRatio) (middle line) to see the updated probability.
@@
 For instance, if the starting probability is 0.5% and the [likelihood ratio](#sec-diagnosticLikelihoodRatio) is 10 (e.g., sensitivity = .90, specificity = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$) from a positive test (i.e., [positive likelihood ratio](#sec-positiveLikelihoodRatio)), the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.
+
+An interactive probability nomogram is available at the following link: <https://jamaevidence.mhmedical.com/data/calculators/LR_nomogram.html> (archived at <https://perma.cc/Z3SW-QMJ3>).
+
 The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the sensitivity (SN) and specificity (SP) of the test.
@@
   SP = 0.91)
+```
 
+The blue line indicates the [posterior probability](#posttestProbability) of the condition given a positive test.
+The pink line indicates the [posterior probability](#posttestProbability) of the condition given a negative test.
+
+```{r}
 petersenlab::nomogrammer(
@@
   FN = 50)
+```
+
+The function can also create a nomogram plot using the sensitivity (SN) and selection rate of the test.
+Here is a nomogram plot from the [HIV example](#sec-baseRatesOverview):
+
+```{r}
+petersenlab::nomogrammer(
+  pretestProb = .003,
+  SN = .95,
+  selectionRate = .01
+)
+```
+
+Here is a nomogram plot from the [cab example](#sec-cabExample):
+
+```{r}
+petersenlab::nomogrammer(
+  pretestProb = .15,
+  SN = .80,
+  SP = .80
+)
 ```

### 2024-09-11 getting-started.qmd
20240911 - fix error

@@
   season = c(2023,2023,2024,2024,2024,2024),
-  week = c(1,2,1,2,3,4)
+  week = c(1,2,1,2,3,4),
   points = c(10,15,20,25,30,35)

### 2024-09-11 getting-started.qmd
20240911 - loading and saving data

@@
 
+Here is how you load a `.RData` file using a *relative path* (i.e., a path relative to the working directory, where the working directory is represented by a period):
+
 ```{r}
 load(file = "./data/nfl_players.RData")
+```
+
+The following code loads a file from an *absolute path*:
+
+```{r}
+#| eval: false
+
+nfl_players <- read.csv("C:/Users/myusername/nfl_players.RData")
+```
+
+Here is how you load a `.csv` file:
+
+```{r}
+#| eval: false
+
+nfl_players <- read.csv("./data/nfl_players.csv") # relative path
+nfl_players <- read.csv("C:/Users/myusername/nfl_players.csv") # absolute path
+```
+
+## Save a Data Frame {#sec-gettingStartedSaveData}
+
+Here is how you save a `.RData` file using a *relative path*:
+
+```{r}
+#| eval: false
+
+save(
+  nfl_players,
+  file = "./data/nfl_players.RData")
+```
+
+The following code saves a file to an *absolute path*:
+
+```{r}
+#| eval: false
+
+save(
+  nfl_players,
+  file = "C:/Users/myusername/nfl_players.RData")
+```
+
+Here is how you save a `.csv` file:
+
+```{r}
+#| eval: false
+
+write.csv(
+  nfl_players,
+  file = "./data/nfl_players.csv") # relative path
+
+write.csv(
+  nfl_players,
+  file = "C:/Users/myusername/nfl_players.csv") # absolute path
 ```

### 2024-09-11 basic-statistics.qmd
20240911 - skewness and kurtosis

@@
 
-- mean
+- mean (aka "average")
 - median
@@
 - kurtosis
+
+Positive skewness (right-skewed) reflects a longer or heavier right-tailed distribution, whereas negative skewness (left-skewed) reflects a longer or heavier left-tailed distribution.
+Fantasy points tend to be positively skewed.
+
+The kurtosis reflects the extent of extreme (outlying) values in a distribution relative to a normal distribution (or bell curve).
+A mesokurtic distribution (with a kurtosis value near zero) reflects a normal amount of tailedness.
+Positive kurtosis values reflect a leptokurtic distribution, where there are lighter tails and a sharper peak than a normal distribution.
+Negative kurtosis values reflect a platykurtic distribution, where there are heavier tails and a flatter peak than a normal distribution.
+Fantasy points tend to have a leptokurtic distribution.
 

### 2024-09-11 base-rates.qmd
20240911 - informal updating

@@
 Focus on other things that may be more predictable (and that may be indirectly prognostic of touchdowns) such as yards, carries/targets, receptions, depth of targets, red zone carries/targets, short distance carries/targets, etc.
+When dealing with numeric predictions (rather than categorical outcomes), the equivalent of the base rate is the average value.
+For instance, the "base rate" of fantasy points for a given position is the average number of fantasy points for that position.
 
@@
 If we want to compare the relative probability of two outcomes, we can use the odds form of Bayes' theorem, as in @eq-bayes5:
+
 $$
@@
 
+#### Informal Updating {#sec-informalUpdating}
+
+@Kahneman2011 provides the following guidance for an informal approach to updating that anchors predictions to the base rate:
+
+1. Start with the "baseline prediction" (i.e., base rate or average outcome).
+1. Generate or identify your "intuitive prediction"—the number that matches your impression of the evidence.
+1. Your posterior prediction should fall somewhere between the baseline prediction and the intuitive prediction.
+"In the default case of no useful evidence, you stay with the baseline [prediction].
+At the other extreme, you also stay with your initial [i.e., intuitive] prediction.
+This will happen, of course, only if you remain completely confident in your initial prediction after a critical review of the evidence that supports it.
+In most cases you will find some reasons to doubt that the correlation between your intuitive judgment and the truth is perfect, and you will end up somewhere between the two poles." (pp. 191–192).
+Base the extent of adjustment (from the baseline prediction) on the magnitude of the correlation between your prediction/evidence and the truth, which acts similar to the [likelihood ratio](#sec-diagnosticLikelihoodRatio).
+For instance, if the correlation between your prediction/evidence and the truth is .5, move 50% of the difference from the baseline prediction to the intuitive prediction.
+
 ## Conclusion {#sec-baseRatesConclusion}

### 2024-09-11 getting-started.qmd
20240911 - fix error

@@
 mergeExample3B <- data.frame(
-  player = c("A","A","B","B"),
+  player = c("A","A","A","A","B","B"),
   season = c(2023,2023,2024,2024,2024,2024),

### 2024-09-11 basic-statistics.qmd
20240911 - clarifications

@@
 
-For a test of group differences, the *p*-value evaluates the likelihood that you would observe a difference as large or larger than the one you observed between the groups if there were no systematic difference between the groups, as depicted in @fig-pValuesDifference.
-For instance, when evaluating whether Quarterbacks have <u>longer</u> careers than Running Backs, and you observed a mean difference of 0.03 years, the *p*-value evaluates the likelihood that you would observe a difference as larger or larger than 0.03 years between the groups if Quarterbacks do not differ from Running Backs in terms of the length of their career.
+For a test of group differences, the *p*-value evaluates the likelihood that you would observe a difference as large or larger than the one you observed between the groups if there were no systematic difference between the groups in the population, as depicted in @fig-pValuesDifference.
+For instance, when evaluating whether Quarterbacks have <u>longer</u> careers than Running Backs, and you observed a mean difference of 0.03 years, the *p*-value evaluates the likelihood that you would observe a difference as large or larger than 0.03 years between the groups if, in truth among all Quarterbacks and Running Backs in the NFL, Quarterbacks do not differ from Running Backs in terms of the length of their career.
 
@@
 
-For a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed between the groups if there were no association between the variables, as depicted in @fig-pValuesAssociation.
-For instance, when evaluating whether number of carries is <u>positively</u> associated with injury likelihood, and you observed a correlation coefficient of $r = .25$ between number of carries and injury likelihood, the *p*-value evaluates the likelihood that you would observe a correlation as strong or stronger than $r = .25$ between the variables if number of carries is not associated with injury likelihood.
+For a test of whether two variables are associated, the *p*-value evaluates the likelihood that you would observe an association as strong or stronger than the one you observed if there were no actual association between the variables in the population, as depicted in @fig-pValuesAssociation.
+For instance, when evaluating whether number of carries is <u>positively</u> associated with injury likelihood, and you observed a correlation coefficient of $r = .25$ between number of carries and injury likelihood, the *p*-value evaluates the likelihood that you would observe a correlation as strong or stronger than $r = .25$ between the variables if, in truth among all NFL Running Backs, number of carries is not associated with injury likelihood.
 

### 2024-09-11 cognitive-bias.qmd
20240911 - narrative fallacy

@@
 However, these mental shortcuts can also lead us astray and to make systematic errors in our judgments and predictions.
+For instance, heuristics can lead us to identify patterns out of randomness.
 *Cognitive biases* are systematic errors in thinking.
@@
 
-*Hindsight bias* is the tendency to perceive that past events were more predictable than they were.
+*Hindsight bias* is the tendency to perceive that past events were more predictable than they were (i.e., the "I-knew-it-all-along" effect).
 People tend to remember the succeess of their predictions and forget the failures of their predictions.
@@
 Likewise, if the manager makes a poor-quality decision but gets lucky, and they attribute their success to a sound decision-making process, this would also reflect outcome bias.
+In general, people tend to blame decision makers for good decisions that turned out badly, and to give them too little credit for good decisions that turned out well [because the decisions appear obvious but only after the fact; @Kahneman2011].
+[Hindsight bias](#sec-cognitiveBiasesHindsight) and outcome bias tend to lead to risk aversion [@Kahneman2011].
 
@@
 
+### Narrative Fallacy {#sec-fallaciesNarrative}
+
+The *narrative fallacy* is the error of creating a story with cause-and-effect explanations from random events.
+According to @Kahneman2011, people find stories more compelling if they a) are simple (as opposed to complex), b) are concrete (as opposed to abstract), c) provide a greater attribution to talent, stupidity, and intentions (compared to luck), and d) focus on a few salient events that occured (rather than the many events that failed to occur).
+For instance, consider that a player undergoes a challenging start to the season; however, they have higher-performing games in the middle of the season.
+A fantasy football manager may construct a "causal narrative" that the player overcame obstacles and is now trending toward strong performance for the rest of the season, even though the stronger performance could reflect some natural and random fluctuation.
+
 ### Conditional Probability Fallacy {#sec-fallaciesConditionalProbability}

### 2024-09-11 cognitive-bias.qmd
20240911 - cross-references

@@
 In the modern world of big data, research and society need people who know how to make sense of the information around us.
-Given humans' cognitive biases, it is valuable to leverage more objective approaches than relying on our "gut" and intutition.
+Given humans' [cognitive biases](#sec-cognitiveBiases), it is valuable to leverage more objective approaches than relying on our "gut" and intutition.
 Statistical approaches can be a more objective way to identify systematic patterns.
@@
 Science is a human enterprise—it is performed by humans each of whom has their own biases.
-For instance, cognitive biases can influence how people interpret statistics.
+For instance, [cognitive biases](#sec-cognitiveBiases) can influence how people interpret statistics.
 As a result, the findings from any given study may be incorrect.
@@
 In our everyday lives, humans are presented with overwhelming amounts of information.
-Because human minds cannot parse every piece of information equally, we tend to take mental shortcuts, called *heuristics*.
+Because human minds cannot parse every piece of information equally, we tend to take mental shortcuts, called [*heuristics*](#sec-heuristics).
 These mental shortcuts can be helpful.
@@
 However, these mental shortcuts can also lead us astray and to make systematic errors in our judgments and predictions.
-For instance, heuristics can lead us to identify patterns out of randomness.
-*Cognitive biases* are systematic errors in thinking.
-*Fallacies* are forms of flawed reasoning.
+For instance, [heuristics](#sec-heuristics) can lead us to identify patterns out of randomness.
+[*Cognitive biases*](#sec-cognitiveBiases) are systematic errors in thinking.
+[*Fallacies*](#sec-fallacies) are forms of flawed reasoning.
+
+Below, we provide examples of [heuristics](#sec-heuristics), [*cognitive biases*](#sec-cognitiveBiases), and [*fallacies*](#sec-fallacies).
 

### 2024-09-12 cognitive-bias.qmd
20240911 - additions

@@
 Below, we provide examples of [heuristics](#sec-heuristics), [*cognitive biases*](#sec-cognitiveBiases), and [*fallacies*](#sec-fallacies).
+These are not exhaustive lists of all the [heuristics](#sec-heuristics), [*cognitive biases*](#sec-cognitiveBiases), and [*fallacies*](#sec-fallacies).
+They are some examples of key ones that are relevant fantasy football.
 
@@
 For instance, people tend not to consider other relevant, important factors that they do not have immediate access to, consistent with the [WYSIATI heuristic](#sec-heuristicsWYSIATI).
-Instead, people tend to focus on the cohesiveness and plausibility of the story, which differs from its probability [@Kahneman2011].
+Instead, people tend to focus on the coherence and plausibility of the story, which differs from its probability [@Kahneman2011].
+That is, one's confidence is not necessarily related to the likelihood that the judgment is correct—it often reflects the "illusion of validity" [@Kahneman2011].
+As one gains more knowledge in a domain, the person also tends to become unrealistically overconfident in their ability to make accurate judgments and predictions [@Kahneman2011].
 
@@
 *Hindsight bias* is the tendency to perceive that past events were more predictable than they were (i.e., the "I-knew-it-all-along" effect).
+Everything seems to make sense in hindsight [@Kahneman2011].
 People tend to remember the succeess of their predictions and forget the failures of their predictions.

### 2024-09-12 cognitive-bias.qmd
20240912 - add to bulleted list

@@
 - [anecdotal fallacy](#sec-fallaciesAnecdotal)
+- [narrative fallacy](#sec-fallaciesNarrative)
 - [conditional probability fallacy](#sec-fallaciesConditionalProbability)

### 2024-09-13 actuarial.qmd
20240912 - actuarial approches

@@
 Moreover, the superiority of actuarial approaches to human judgment tends to hold even when the expert is given more information than the actuarial approach [@Dawes1989].
+In addition, actuarial predictions outperform human judgment even when the human is given the result of the actuarial prediction [@Kahneman2011].
 Allowing experts to override actuarial predictions consistently leads to lower predictive accuracy [@Garb2019].
@@
 
-Here are some advantages of computers over humans:
+Here are some advantages of computers over humans, including "experts":
 
@@
 - Computers are faster at making calculations.
+- Given the same input, a formula will give the exact same result everytime.
+Humans' judgment tends to be inconsistent both across raters and within rater across time, when trying to make judgments or predictions from complex information [@Kahneman2011].
+As noted in @sec-reliabilityVsValidity, [reliability](#sec-reliability) sets the upper bound for [validity](#sec-validity), so unreliable judgments cannot be accurate (i.e., valid).
 - Computations by computers are error-free (as long as the computations are programmed correctly).
@@
 Humans tend to give too much weight to singular predictors.
+Experts may attempt to be clever and to consider complex combinations of predictors, but doing so often reduces validity [@Kahneman2011].
+Simple combinations of predictions often outperform more complex combinations [@Kahneman2011].
 
@@
 For instance, actuarial methods have shown promise in predicting violence, criminal recidivism, psychosis onset, course of mental disorders, treatment selection, treatment failure, suicide attempts, and suicide [@Garb2019].
+Actuarial methods are especially important to use in low-validity environments (like fantasy football) in which there is considerable uncertainty and unpredictability [@Kahneman2011].
 
@@
 
-The best actuarial models tend to be relatively simple (parsimonious), that can account for one or several of the most important predictors and their optimal weightings, and that account for the base rate of the phenomenon.
-Even unit-weighted formulas (formulas whose [predictor variables](#sec-correlationalStudy) are equally weighted with a weight of one) can sometimes generalize better to other samples than complex weightings [@Garb2019].
-Differential weightings sometimes capture random variance and [over-fit](#sec-overfitting) the model, thus leading to predictive accuracy shrinkage in cross-validation samples [@Garb2019], as described below.
+The best actuarial models tend to be relatively simple (parsimonious), that can account for one or several of the most important predictors and their optimal weightings, and that account for the [base rate](#sec-baseRate) of the phenomenon.
+[Multiple regression](#sec-multipleRegression) and/or prior literature can be used to identify the weights of various predictors.
+Even unit-weighted formulas (formulas whose [predictor variables](#sec-correlationalStudy) are equally weighted with a weight of one) can sometimes generalize better to other samples than complex weightings [@Garb2019; @Kahneman2011].
+Differential weightings sometimes capture random variance and [over-fit](#sec-overfitting) the model, thus leading to predictive accuracy [shrinkage](#sec-shrinkage) in [cross-validation](#sec-crossValidation) samples [@Garb2019], as described below.
 The choice of [predictor variables](#sec-correlationalStudy) often matters more than their weighting.

### 2024-09-13 evaluating-prediction-accuracy.qmd
20240803 - Miller (2013)
20270907 - citations in block quotes
20240911 - additions
20240912 - actuarial approches

@@
 # Evaluation of Prediction/Forecasting Accuracy {#sec-predictionAccuracy}
+
+> "Nothing ruins fantasy like reality." – Renee Miller
 
@@
 Fantasy performance—and human behavior more general—is incredibly challenging to predict.
+Indeed, many things in the world, in particular long-term trends, are unpredictable [@Kahneman2011].
 In fantasy football, there is considerable luck/chance/randomness.
@@
 These and other factors combine to render fantasy football predictions not highly accurate.
+Domains with high uncertainty and unpredictability are considered "low-validity environments" [@Kahneman2011, p. 223].
 But, first, let's learn about the various ways we can evaluate the accuracy of predictions.
@@
 > There is a period of enthusiasm about a new theory, a period of attempted application to several fact domains, a period of disillusionment as the negative data come in, a growing bafflement about inconsistent and unreplicable empirical results, multiple resort to ad hoc excuses, and then finally people just sort of lose interest in the thing and pursue other endeavors.
-(pp.
-806–807).
+> 
+> Meehl [-@Meehl1978, pp. 806–807]
 
@@
 Experts' predictions from many different domains tend to be inaccurate, including political scientists [@Tetlock2017], physicians [@Koehler2002], clinical psychologists [@Oskamp1965], stock market traders and corporate financial officers [@Skala2008], seismologists' predictions of earthquakes [@Hough2016], economists' predictions about the economy [@Makridakis2009], lawyers [@Koehler2002], and business managers [@Russo1992].
+Thus, I would not put much confidence in the predictions by fantasy football fundits.
 The most common pattern of experts' predictions is that they show overextremity, that is, their predictions have probability judgments that tend to be too extreme, as described in Section @sec-calibration.

### 2024-09-14 actuarial.qmd
20240913 - experts

@@
 To improve accuracy of judgments, it is important for feedback to be clear, consistent, and timely.
+Intuition is a form of recognition-based judgment (i.e., recognizing cues that provide access to information in memory).
+Development of strong intuition depends on the quality and speed of feedback, in addition to having adequate opportunities to practice [i.e., sufficient opportunities to learn the cues; @Kahneman2011].
+The quality and speed of the feedback tend to benefit anesthesiologists who often quickly learn the results of their actions.
+By contrast, radiologists tend not to receive quality feedback about the accuracy of their diagnoses, including their [false-positive](#sec-falsePositive) and [false-negative](#sec-falseNegative) decisions [@Kahneman2011].
+
+In general, many so-called experts are "pseudo-experts" who do not know the boundaries of their competence—that is, they do not know what they do not know; they have the illusion of validity of their predictions and are [overconfident](#sec-cognitiveBiasesOverconfidence) about their predictions [@Kahneman2011].
+Yet, many people arrogantly proclaim to have predictive powers, including in low-validity environments such as fantasy football.
+Indeed, people are more likely to watch a show if the pundits are opinionated, clear, and (overly) confident, because they are more entertaining [even though they less to be less accurate than individuals who thinking is more complex and less decisive; @Kahneman2011].
+True experts know their limits in terms of knowledge and ability to predict.
+
+Intuitions tend to be skilled when a) the environment is regular and predictable, and b) there is opportunity to learn the regularities, cues, and contingencies through extensive practice [@Kahneman2011].
+Example domains that meet these conditions supporting intuition include activities such as chess, bridge, and poker, and occupations such as medical providers, athletes, and firefighters.
+By contrast, fantasy football and other domains such as stock-picking, clinical psychology, and other long-terms forecasts are low-validity environments that are irregular and unpredictable.
+In environments that do not have stable regularities, intuition cannot be trusted [@Kahneman2011].
 

### 2024-09-14 actuarial.qmd
20240913 - tv pundits

@@
 Yet, many people arrogantly proclaim to have predictive powers, including in low-validity environments such as fantasy football.
-Indeed, people are more likely to watch a show if the pundits are opinionated, clear, and (overly) confident, because they are more entertaining [even though they less to be less accurate than individuals who thinking is more complex and less decisive; @Kahneman2011].
+Indeed, pundits are more likely to be television guests if they are opinionated, clear, and (overly) confident and make big, bold predictions, because they are more entertaining [even though they tend to be less accurate than individuals whose thinking is more complex and less decisive; @Kahneman2011; @Silver2012].
 True experts know their limits in terms of knowledge and ability to predict.

### 2024-09-15 actuarial.qmd
20240915 - optimism and overconfidence

@@
 Indeed, pundits are more likely to be television guests if they are opinionated, clear, and (overly) confident and make big, bold predictions, because they are more entertaining [even though they tend to be less accurate than individuals whose thinking is more complex and less decisive; @Kahneman2011; @Silver2012].
-True experts know their limits in terms of knowledge and ability to predict.
+[Optimism](#sec-cognitiveBiasesOptimism) and [(over)confidence](#sec-cognitiveBiasesOverconfidence) are valued by society [@Kahneman2011].
+Nevertheless, true experts know their limits in terms of knowledge and ability to predict.
 

### 2024-09-15 cognitive-bias.qmd
20240914 - base rates
20240915 - optimism and overconfidence

@@
 This calls to mind the fictitious Lake Wobegon in the radio show *A Praririe Home Companion*, "where all the women are strong, all the men are good-looking, and all the children are above average."
+As an example, 90% of drivers believe they are better than average [@Kahneman2011].
 
@@
 As one gains more knowledge in a domain, the person also tends to become unrealistically overconfident in their ability to make accurate judgments and predictions [@Kahneman2011].
+In general, people tend to focus on what they know and to neglect what they do not know leading to overconfidence in one's beliefs [@Kahneman2011].
 
@@
 
+People tend to focus on the role of skill and to neglect the role of luck when explaining the past and predicting the future, giving people an illusion of control [@Kahneman2011]. 
 Players' performance in fantasy football, and human behavior more generally, is complex and multiply determined (i.e., is influenced by many factors).
@@
 For instance a manager might overestimate the likelihood that their team will win the championship and may underestimate the likelihood that they may lose a player to injury.
+In terms of its key consequences for decisions, optimism bias may be considered the most important cognitive bias [@Kahneman2011].
+In general, people show an optimism bias.
+However, some people ("optimists") are even more optimistic than others [@Kahneman2011].
+The optimism bias—in particular a delusional optimism—can lead to [overconfidence](#sec-cognitiveBiasesOverconfidence) [@Kahneman2011].
+Although optimism is associated with many advantages in terms of well-being, it can also lead to excessive risk-taking and to costly persistence [@Kahneman2011].
 
@@
 Focusing on the specific information about the event (and thus underweighting the base rate) involves the [representativeness heuristic](#sec-heuristicsRepresentativeness) [@Kahneman2011]—making judgments based on how similar the event is to a mental prototype regardless of how frequent the event the event actually is.
+As noted by @Kahneman, "people who have information about an individual case rarely feel the need to know the statistics of the class which the case belongs." (p. 249).
 In general, people tend to overestimate the likelihood of low base-rate events [@Kahneman2011].

### 2024-09-15 actuarial.qmd
20240915 - pundits
20240915 - videos

@@
 Indeed, pundits are more likely to be television guests if they are opinionated, clear, and (overly) confident and make big, bold predictions, because they are more entertaining [even though they tend to be less accurate than individuals whose thinking is more complex and less decisive; @Kahneman2011; @Silver2012].
+Consider sports pundits like Stephen A. Smith and Skip Bayless who make bold predictions with uber confidence.
 [Optimism](#sec-cognitiveBiasesOptimism) and [(over)confidence](#sec-cognitiveBiasesOverconfidence) are valued by society [@Kahneman2011].
 Nevertheless, true experts know their limits in terms of knowledge and ability to predict.
+
+::: {.content-visible when-format="html:js"}
+
+Here is a video of sports pundits, Stephen A. Smith and Skip Bayless, making bold statements and incorrect predictions:
+
+![Video of Sports Pundits, Stephen A.
+Smith and Skip Bayless, Making Bold Statements and Incorrect Predictions.
+From: <https://www.youtube.com/watch?v=lTjBuEPcLlc>.](images/sportsPundits.mp4){width=100%}
+
+:::
 

### 2024-09-15 fantasy-football.qmd
20240915 - videos

@@
 
-![Video that Provides An Introduction to American Football.](images/football.mp4){width=100%}
+![Video that Provides An Introduction to American Football. From: <https://www.youtube.com/watch?v=Ddwp1HyEFRE>.](images/football.mp4){width=100%}
 
@@
 
-![Video that Provides an Introduction to Fantasy Football.](images/fantasyFootball.mp4){width=100%}
+![Video that Provides an Introduction to Fantasy Football. From: <https://www.youtube.com/watch?v=XhrBapdhLEc>.](images/fantasyFootball.mp4){width=100%}
 

### 2024-09-17 base-rates.qmd
20240914 - base rates
20240916 - Nate Silver examples

@@
 ### Bayes' Theorem {#sec-bayesTheorem}
+
+#### Standard Formulation {#sec-bayesTheoremStandard}
 
@@
 
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.
 
@@
 If $P(R_i) > P(C_i)$, our predictions must include some false positives.
+
+#### Alternative Formulation {#sec-bayesTheoremAlternative}
+
+Using the law of total probability, we can substitute the calculation of the [marginal probability](#sec-baseRate) ($P(R_i)$) into Bayes' theorem to get an alternative formulation of Bayes' theorem, as in @eq-baseRateUpdated3:
+
+$$
+\begin{aligned}
+  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)} \\
+  &= \frac{P(R | C) \cdot P(C_i)}{P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i)} \\
+  &= \frac{P(R | C) \cdot P(C_i)}{P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot [1 - P(C_i)]}
+\end{aligned}
+$$ {#eq-baseRateUpdated3}
+
+Instead of using [marginal probability](#sec-baseRate) ([base rate](#sec-baseRate)) of $R$, as in the original formulation of Bayes' theorem, it uses the [conditional probability](#sec-conditionalProbability), $P(R|\text{not } C)$.
+Thus, it uses three terms: two [conditional probabilities](#sec-conditionalProbability)—$P(R|C)$ and $P(R|\text{not } C)$—and one [marginal probability](#sec-baseRate), $P(C_i)$.
+
+Let us see how the alternative formulation of Bayes' theorem applies to the HIV example above.
+We can calculate the probability of HIV given a positive test using three terms: the [conditional probability](#sec-conditionalProbability) that a person with HIV gets a positive test (i.e., [sensitivity](#sec-sensitivity)), the [conditional probability](#sec-conditionalProbability) that a person without HIV gets a positive test (i.e., [false positive rate](#sec-falsePositiveRate)), and the [base rate](#sec-baseRate) of HIV.
+Using the $P(R|\text{not } C)$ calculated in @eq-conditionalProbabilityRevised, the [conditional probability](#sec-conditionalProbability) of HIV given a single positive test is in @eq-bayes4:
+
+$$
+\small
+\begin{aligned}
+  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot [1 - P(C_i)]} \\
+  &= \frac{\text{sensitivity of test} \times \text{base rate of HIV}}{\text{sensitivity of test} \times \text{base rate of HIV} + \text{false positive rate of test} \times (1 - \text{base rate of HIV})} \\
+  &= \frac{95\% \times .3\%}{95\% \times .3\% + .7171515\% \times (1 - .3\%)} = \frac{.95 \times .003}{.95 \times .003 + .007171515 \times (1 - .003)}\\
+  &= 28.5\%
+\end{aligned}
+$$ {#eq-bayes4}
+
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.
+
+```{r}
+pAgivenB(
+  pBgivenA = .95,
+  pA = .003,
+  pBgivenNotA = .007171515)
+
+pAgivenB(
+  pBgivenA = .95,
+  pA = .003,
+  pBgivenNotA = pBgivenNotA(
+    pBgivenA = .95,
+    pA = .003,
+    pB = .01))
+```
+
+To calculate the [conditional probability](#sec-conditionalProbability) of HIV given a second positive test, we update our priors because the person has now tested positive for HIV.
+We update the [prior probability](#sec-baseRate) of HIV ($P(C_i)$) based on the posterior probability of HIV after a positive test ($P(C | R)$) that we calculated above.
+We can calculate the [conditional probability](#sec-conditionalProbability) of HIV given a second positive test using three terms: the [conditional probability](#sec-conditionalProbability) that a person with HIV gets a positive test (i.e., [sensitivity](#sec-sensitivity); which stays the same), the [conditional probability](#sec-conditionalProbability) that a person without HIV gets a positive test (i.e., [false positive rate](#sec-falsePositiveRate); which stays the same), and the updated [marginal probability](#sec-baseRate) of HIV.
+The [conditional probability](#sec-conditionalProbability) of HIV given a second positive test is in @eq-baseRateUpdated4:
+
+$$
+\scriptsize
+\begin{aligned}
+  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot [1 - P(C_i)]} \\
+  &= \frac{\text{sensitivity of test} \times \text{updated base rate of HIV}}{\text{sensitivity of test} \times \text{updated base rate of HIV} + \text{false positive rate of test} \times (1 - \text{updated base rate of HIV})} \\
+  &= \frac{95\% \times 28.5\%}{95\% \times 28.5\% + .7171515\% \times (1 - 28.5\%)} = \frac{.95 \times .285}{.95 \times .285 + .007171515 \times (1 - .285)}\\
+  &= 98.14\%
+\end{aligned}
+$$ {#eq-baseRateUpdated4}
+
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.
+
+```{r}
+pAgivenB(
+  pBgivenA = .95,
+  pA = .285,
+  pBgivenNotA = .007171515)
+
+pAgivenB(
+  pBgivenA = .95,
+  pA = .285,
+  pBgivenNotA = pBgivenNotA(
+    pBgivenA = .95,
+    pA = .003,
+    pB = .01))
+```
+
+#### Interim Summary
 
@@
 We can compute the term in the denominator ($P(\text{Identified as Blue})$) using the law of total probability (described in @sec-bayesTheorem).
+
 $$
@@
 
+## Nate Silver Examples {#sec-nateSilverExamples}
+
+@Silver2012 provides several examples that leverage the [alternative formulation of Bayes' theorem](#sec-bayesTheoremAlternative) provided in @eq-baseRateUpdated3 and summarized below:
+
+$$
+\begin{aligned}
+  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot [1 - P(C_i)]}
+\end{aligned}
+$$ {#eq-baseRateUpdated3}
+
+In each case, the formula uses three elements to calculate the probability that the hypothesis is true:
+
+1. the [conditional probability](#sec-conditionalProbability) the likelihood of observing the evidence, $R$, given that the hypothesis, $C$, is true (i.e., $P(R|C)$; [sensitivity](#sec-sensitivity))
+1.
+the [conditional probability](#sec-conditionalProbability) the likelihood of observing the evidence, $R$, given that the hypothesis, $C$, is false (i.e., $P(R | \text{not } C)$; [false positive rate](#sec-falsePositiveRate))
+1. the [marginal probability](#sec-baseRate) (base rate) of the event occurring (i.e., the prior probability of the hypothesis, $C$ being true; $P(C_i)$)
+
+Example 1: You just came home and a strange pair of underwear in your drawer.
+What is the probability that your partner is cheating on you?
+
+- the [prior probability](#sec-baseRate) that your spouse is cheating on you: 4%
+- the [conditional probability](#sec-conditionalProbability) of underwear appearing given that your spouse is cheating on you: 50%
+- the [conditional probability](#sec-conditionalProbability) of underwear appearing given that your spouse is *not* cheating on you: 5%
+
+```{r}
+pAgivenB(
+  pBgivenA = .50,
+  pA = .04,
+  pBgivenNotA = .05)
+```
+
+Example 2: What is the probability that a woman in her 40s has breast cancer if she tested positive on a mammogram?
+
+- the [prior probability](#sec-baseRate) that she has cancer: 1.4%
+- the [conditional probability](#sec-conditionalProbability) that she has a positive test given that she has cancer: 75%
+- the [conditional probability](#sec-conditionalProbability) that she has a positive test given that she does *not* have cancer: 10%
+
+```{r}
+pAgivenB(
+  pBgivenA = .75,
+  pA = .014,
+  pBgivenNotA = .10)
+```
+
+Example 3A: Consider the information we had on 9/11 when the first plane hit the World Trade Center.
+What is the probability that a terror attack occurred given that the first plane hit the World Trade Center?
+
+- the [prior probability](#sec-baseRate) that terrorists crash a plane into a Manhattan skyscraper: 0.005%
+- the [conditional probability](#sec-conditionalProbability) that a plane crashes into a Manhattan skyscraper if terrorists are attacking Manhattan skyscrapers: 100%
+- the [conditional probability](#sec-conditionalProbability) that a plane crashes into a Manhattan skyscraper if terrorists are *not* attacking Manhattan skyscrapers (i.e., it is an accident): 0.008%
+
+```{r}
+pAgivenB(
+  pBgivenA = 1,
+  pA = .00005,
+  pBgivenNotA = .00008)
+```
+
+Example 3A: Now, consider that a second plan just hit the World Trade Center.
+What is the probability that a terror attack occurred given that a second plane hit the World Trade Center?
+
+- the revised [prior probability](#sec-baseRate) that terrorists crash a plane into a Manhattan skyscraper (from Example 3A): 38.46272%
+- the [conditional probability](#sec-conditionalProbability) that a plane crashes into a Manhattan skyscraper if terrorists are attacking Manhattan skyscrapers: 100%
+- the [conditional probability](#sec-conditionalProbability) that a plane crashes into a Manhattan skyscraper if terrorists are *not* attacking Manhattan skyscrapers (i.e., it is an accident): 0.008%
+
+```{r}
+pAgivenB(
+  pBgivenA = 1,
+  pA = .3846272,
+  pBgivenNotA = .00008)
+```
+
 ## Base Rates Applied to Fantasy Football {#sec-baseRateFantasyFootball}
@@
 Focus on other things that may be more predictable (and that may be indirectly prognostic of touchdowns) such as yards, carries/targets, receptions, depth of targets, red zone carries/targets, short distance carries/targets, etc.
+
+[PROVIDE ACCURACY OF PROJECTIONS OF TOUCHDOWNS VS YARDAGE]
+
 When dealing with numeric predictions (rather than categorical outcomes), the equivalent of the base rate is the average value.
 For instance, the "base rate" of fantasy points for a given position is the average number of fantasy points for that position.
-
-[PROVIDE ACCURACY OF PROJECTIONS OF TOUCHDOWNS VS YARDAGE]
+We could subdivide even further to identify, for instance, the "base rate" of fantasy points for the Wide Receiver at the top of the depth chart on a team.
 

### 2024-09-17 actuarial.qmd
20240916 - compelling

@@
 Yet, many people arrogantly proclaim to have predictive powers, including in low-validity environments such as fantasy football.
-Indeed, pundits are more likely to be television guests if they are opinionated, clear, and (overly) confident and make big, bold predictions, because they are more entertaining [even though they tend to be less accurate than individuals whose thinking is more complex and less decisive; @Kahneman2011; @Silver2012].
+Indeed, pundits are more likely to be television guests if they are opinionated, clear, and (overly) confident and make big, bold predictions, because they are more entertaining and their predictions seem more compelling [even though they tend to be less accurate than individuals whose thinking is more complex and less decisive; @Kahneman2011; @Silver2012].
 Consider sports pundits like Stephen A. Smith and Skip Bayless who make bold predictions with uber confidence.

### 2024-09-17 base-rates.qmd
20240916 - copyediting

@@
 
-Example 1: You just came home and a strange pair of underwear in your drawer.
+Example 1: You came home and found a strange pair of underwear in your underwear drawer.
 What is the probability that your partner is cheating on you?
 
-- the [prior probability](#sec-baseRate) that your spouse is cheating on you: 4%
-- the [conditional probability](#sec-conditionalProbability) of underwear appearing given that your spouse is cheating on you: 50%
-- the [conditional probability](#sec-conditionalProbability) of underwear appearing given that your spouse is *not* cheating on you: 5%
+- the [prior probability](#sec-baseRate) that your partner is cheating on you: 4%
+- the [conditional probability](#sec-conditionalProbability) of underwear appearing given that your partner is cheating on you: 50%
+- the [conditional probability](#sec-conditionalProbability) of underwear appearing given that your partner is *not* cheating on you: 5%
 

### 2024-09-17 cognitive-bias.qmd
20240916 - risk aversion bias

@@
 - [loss aversion bias](#sec-cognitiveBiasesLossAversion)
+- [risk aversion bias](#sec-cognitiveBiasesRiskAversion)
 - [primacy effect bias](#sec-cognitiveBiasesPrimacy)
@@
 That is, loss aversion (similar to punishment sensitivity, as opposed to reward sensitivity) is the tendency to weigh losses (or punishment) more heavily than gains (or rewards).
-Loss aversion is different from risk aversion.
-Risk aversion is the tendency to prefer outcomes with low uncertainty (compared to outcomes with greater uncertainty).
+Loss aversion is different from [risk aversion](#sec-cognitiveBiasesRiskAversion).
 Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
 In fantasy football, loss aversion may lead managers to start or hold onto underperforming high drafts for too long instead of a starting a more promising player out of fear of losing potential value from their initial investment.
+
+### Risk Aversion Bias {#sec-cognitiveBiasesRiskAversion}
+
+Risk aversion is the tendency to prefer outcomes with low uncertainty (compared to outcomes with greater uncertainty).
+For instance, a fantasy manager may prefer to draft players who have demonstrated performance compared to players right out of college with no proven history in the NFL.
+
+Risk seeking, by contrast, has been observed in entrepreneurs and in generals when all of their options are poor [@Kahneman2011].
+In fantasy football, risk seeking may be more likely when a manager has a team full of underperforming players.
 
@@
 
-Note how information is framed, knowing that the framing itself could lead you astray.
+Pay attention to how information is framed, knowing that the framing itself could lead you astray.
 Thus, try to focus on the information itself rather than the framing when making decisions.

### 2024-09-17 base-rates.qmd
20240916 - copyediting

@@
 \end{aligned}
-$$ {#eq-baseRateUpdated3}
+$$ {#eq-baseRateAlternative}
 
-In each case, the formula uses three elements to calculate the probability that the hypothesis is true:
+In each example, the formula uses three elements to calculate the probability that the hypothesis is true:
 
@@
 the [conditional probability](#sec-conditionalProbability) the likelihood of observing the evidence, $R$, given that the hypothesis, $C$, is false (i.e., $P(R | \text{not } C)$; [false positive rate](#sec-falsePositiveRate))
-1. the [marginal probability](#sec-baseRate) (base rate) of the event occurring (i.e., the prior probability of the hypothesis, $C$ being true; $P(C_i)$)
+1. the [marginal probability](#sec-baseRate) (base rate) of the event occurring (i.e., the prior probability of the hypothesis, $C$, being true; $P(C_i)$)
 
@@
 
-Example 3A: Now, consider that a second plan just hit the World Trade Center.
+Example 3B: Now, consider that a second plan just hit the World Trade Center.
 What is the probability that a terror attack occurred given that a second plane hit the World Trade Center?

### 2024-09-17 base-rates.qmd
20240917 - nomogram for Nate Silver examples

@@
 
-1. the [conditional probability](#sec-conditionalProbability) the likelihood of observing the evidence, $R$, given that the hypothesis, $C$, is true (i.e., $P(R|C)$; [sensitivity](#sec-sensitivity))
+1.
+the [conditional probability](#sec-conditionalProbability) the likelihood of observing the evidence, $R$, given that the hypothesis, $C$, is true (i.e., $P(R|C)$; [true positive rate](#sec-sensitivity))
 1.
@@
 1. the [marginal probability](#sec-baseRate) (base rate) of the event occurring (i.e., the prior probability of the hypothesis, $C$, being true; $P(C_i)$)
+
+Thus, the formula uses the [base rate](#sec-baseRate), the [true positive rate](#sec-sensitivity) (sensitivity), and the [false positive rate](#sec-falsePositiveRate).
+The ratio of the [true positive rate](#sec-sensitivity) to the [false positive rate](#sec-falsePositiveRate) is called the [positive likelihood ratio](#sec-positiveLikelihoodRatio), and is used in [Bayesian updating](#sec-bayesianUpdating).
+
+### Example 1: Is Your Partner Cheating on You? {#sec-nateSilverExample1}
 
@@
 
+### Example 2: Does a Woman Have Cancer? {#sec-nateSilverExample2}
+
 Example 2: What is the probability that a woman in her 40s has breast cancer if she tested positive on a mammogram?
@@
 ```
+
+### Example 3: Was it a Terrorist Attack? {#sec-nateSilverExample3}
+
+#### Example 3A {#sec-nateSilverExample3A}
 
@@
 ```
+
+#### Example 3B {#sec-nateSilverExample3B}
 
@@
 
-The function can also estimate the posttest probability of an event given the number of true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN):
+The function can also estimate the posttest probability of an event given the number of [true positives](#sec-truePositive) (TP), [true negatives](#sec-trueNegative) (TN), [false positives](#sec-falsePositive) (FP), and [false negatives](#sec-falseNegative) (FN):
 
@@
 
-We discuss true positives (TP), true negatives (TN), false positives (FP), false negatives (FN), sensitivity (SN), and specificity (SP) in INSERT.
+We discuss [true positives](#sec-truePositive) (TP), [true negatives](#sec-trueNegative) (TN), [false positives](#sec-falsePositive) (FP), [false negatives](#sec-falseNegative) (FN), [sensitivity](#sec-sensitivity) (SN), and [specificity](#sec-specificity) (SP) in @sec-thresholdDependentAccuracy (@sec-decisionOutcomes and @sec-sensitivitySpecificityPPVnpv).
 
@@
 A probability nomogram calculator can be found at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the sensitivity (SN) and specificity (SP) of the test, as adapted from Adam Chekroud (<https://github.com/achekroud/nomogrammer>):
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from Adam Chekroud (<https://github.com/achekroud/nomogrammer>):
 
@@
 
-The function can also create a nomogram plot using the true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN):
+The function can also create a nomogram plot using the [true positives](#sec-truePositive) (TP), [true negatives](#sec-trueNegative) (TN), [false positives](#sec-falsePositive) (FP), and [false negatives](#sec-falseNegative) (FN):
 
@@
 
-The function can also create a nomogram plot using the sensitivity (SN) and selection rate of the test.
+The function can also create a nomogram plot using the [sensitivity](#sec-sensitivity) (SN) and [selection rate](#sec-selectionRatio) of the test.
 Here is a nomogram plot from the [HIV example](#sec-baseRatesOverview):
@@
 
-Here is a nomogram plot from the [cab example](#sec-cabExample):
+Here is a nomogram plot from the [cab example](#sec-cabExample) [@Kahneman2011]:
 
@@
   SP = .80
+)
+```
+
+Here is a nomogram plot from [Example 1](#sec-nateSilverExample1) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):
+
+```{r}
+petersenlab::nomogrammer(
+  pretestProb = .04,
+  SN = .50,
+  SP = 1 - .05
+)
+```
+
+Here is a nomogram plot from [Example 2](#sec-nateSilverExample2) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):
+
+```{r}
+petersenlab::nomogrammer(
+  pretestProb = .014,
+  SN = .75,
+  SP = 1 - .10
+)
+```
+
+Here is a nomogram plot from [Example 3A](#sec-nateSilverExample3A) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):
+
+```{r}
+petersenlab::nomogrammer(
+  pretestProb = .00005,
+  SN = 1,
+  SP = 1 - .00008
+)
+```
+
+Here is a nomogram plot from [Example 3B](#sec-nateSilverExample3B) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):
+
+```{r}
+petersenlab::nomogrammer(
+  pretestProb = .3846272,
+  SN = 1,
+  SP = 1 - .00008
 )

### 2024-09-17 base-rates.qmd
20240917 - update nomogram plots based on FPR

@@
 
-For instance, if the starting probability is 0.5% and the [likelihood ratio](#sec-diagnosticLikelihoodRatio) is 10 (e.g., sensitivity = .90, specificity = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$) from a positive test (i.e., [positive likelihood ratio](#sec-positiveLikelihoodRatio)), the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.
+For instance, if the starting probability is 0.5% and the [likelihood ratio](#sec-diagnosticLikelihoodRatio) is 10 (e.g., [sensitivity](#sec-sensitivity) = .90, [specificity](#sec-specificity) = .91: $\text{likelihood ratio} = \frac{\text{sensitivity}}{1 - \text{specificity}} = \frac{.9}{1-.91} = 10$) from a positive test (i.e., [positive likelihood ratio](#sec-positiveLikelihoodRatio)), the updated probability is less than 5%, as depicted in @fig-probabilityNomogramLine.
 
@@
 
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the sensitivity (SN) and specificity (SP) of the test.
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test.
 
@@
 
+The function can also create a nomogram plot using the [sensitivity](#sec-sensitivity) (SN) and [false positive rate](#sec-falsePositiveRate) of the test.
 Here is a nomogram plot from [Example 1](#sec-nateSilverExample1) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):
@@
   SN = .50,
-  SP = 1 - .05
+  FPR = .05
 )
@@
   SN = .75,
-  SP = 1 - .10
+  FPR = .10
 )
@@
   SN = 1,
-  SP = 1 - .00008
+  FPR = .00008
 )

### 2024-09-17 cognitive-bias.qmd
20240917 - risk aversion

@@
 
-Risk aversion is the tendency to prefer outcomes with low uncertainty (compared to outcomes with greater uncertainty).
-For instance, a fantasy manager may prefer to draft players who have demonstrated performance compared to players right out of college with no proven history in the NFL.
+*Risk aversion bias* is the tendency to prefer outcomes with low uncertainty (i.e., risk), even if they offer lower potential rewards compared to outcomes with greater uncertainty but potentially greater rewards.
+Risk aversion leads people to select safer options but may lead them to miss out on higher-gain opportunities.
+For instance, risk aversion may lead a fantasy manager to start players who are more steady (i.e., show greater game-to-game [consistency](#sec-evalHistoricalConsistency)) over players who are more volatile (i.e., show greater game-to-game variability) but have higher upside potential.
 

### 2024-09-17 base-rates.qmd
20240917 - updates

@@
 The function can also create a nomogram plot using the [sensitivity](#sec-sensitivity) (SN) and [false positive rate](#sec-falsePositiveRate) of the test.
-Here is a nomogram plot from [Example 1](#sec-nateSilverExample1) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):
+Here is a nomogram plot from [Example 1](#sec-nateSilverExample1) from @Silver2012:
 
@@
 
-Here is a nomogram plot from [Example 2](#sec-nateSilverExample2) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):
+Here is a nomogram plot from [Example 2](#sec-nateSilverExample2) from @Silver2012:
 
@@
 
-Here is a nomogram plot from [Example 3A](#sec-nateSilverExample3A) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):
+Here is a nomogram plot from [Example 3A](#sec-nateSilverExample3A) from @Silver2012:
 
@@
 
-Here is a nomogram plot from [Example 3B](#sec-nateSilverExample3B) from @Silver2012 (where [specificity](#sec-specificity) equals 1 minus the [false positive rate](#sec-falsePositiveRate)):
+Here is a nomogram plot from [Example 3B](#sec-nateSilverExample3B) from @Silver2012:
 

### 2024-09-17 base-rates.qmd
20240917 - clarifications

@@
 
-### Example 2: Does a Woman Have Cancer? {#sec-nateSilverExample2}
+### Example 2: Does a Person Have Breast Cancer? {#sec-nateSilverExample2}
 
@@
 
-- the [prior probability](#sec-baseRate) that she has cancer: 1.4%
-- the [conditional probability](#sec-conditionalProbability) that she has a positive test given that she has cancer: 75%
-- the [conditional probability](#sec-conditionalProbability) that she has a positive test given that she does *not* have cancer: 10%
+- the [prior probability](#sec-baseRate) that she has breast cancer: 1.4%
+- the [conditional probability](#sec-conditionalProbability) that she has a positive test given that she has breast cancer: 75%
+- the [conditional probability](#sec-conditionalProbability) that she has a positive test given that she does *not* have breast cancer: 10%
 
@@
 [Bayes' theorem](#sec-bayesTheorem) provides a way to convert from one conditional probability to its inverse conditional probability using the [base rate](#sec-baseRate) of each event.
-There are various ways to account for [base rates](#sec-baseRate) for more accurate predictions, including through the use of [actuarial formulas](#sec-accountForBaseRatesActuarial) and [Bayesian updating](#sec-bayesianUpdating).
+There are various ways to account for [base rates](#sec-baseRate) for more accurate predictions, including through the use of [actuarial formulas](#sec-accountForBaseRatesActuarial), [Bayesian updating](#sec-bayesianUpdating), and more [informal approaches](#sec-informalUpdating).
 [Bayesian updating](#sec-bayesianUpdating) uses [Bayes' theorem](#sec-bayesTheorem) to calculate a posttest probability from a [pretest probability](#sec-baseRate) and a test result ([likelihood ratio](#sec-diagnosticLikelihoodRatio)).

### 2024-09-17 base-rates.qmd
20240917 - copyediting

@@
 
-Example 3B: Now, consider that a second plan just hit the World Trade Center.
+Example 3B: Now, consider that a second plane just hit the World Trade Center.
 What is the probability that a terror attack occurred given that a second plane hit the World Trade Center?
 
-- the revised [prior probability](#sec-baseRate) that terrorists crash a plane into a Manhattan skyscraper (from Example 3A): 38.46272%
+- the revised [prior probability](#sec-baseRate) that terrorists crash a plane into a Manhattan skyscraper (from [Example 3A](#sec-nateSilverExample3A)): 38.46272%
 - the [conditional probability](#sec-conditionalProbability) that a plane crashes into a Manhattan skyscraper if terrorists are attacking Manhattan skyscrapers: 100%

### 2024-09-17 cognitive-bias.qmd
20240917 - missing cite

@@
 Focusing on the specific information about the event (and thus underweighting the base rate) involves the [representativeness heuristic](#sec-heuristicsRepresentativeness) [@Kahneman2011]—making judgments based on how similar the event is to a mental prototype regardless of how frequent the event the event actually is.
-As noted by @Kahneman, "people who have information about an individual case rarely feel the need to know the statistics of the class which the case belongs." (p. 249).
+As noted by @Kahneman2011, "people who have information about an individual case rarely feel the need to know the statistics of the class which the case belongs." (p. 249).
 In general, people tend to overestimate the likelihood of low base-rate events [@Kahneman2011].

### 2024-09-17 evaluating-prediction-accuracy.qmd
20240917 - updates

@@
 - [general accuracy](#sec-calibration):
+  - [MAE](#sec-meanAbsoluteError)
   - [RMSE](#sec-rootMeanSquaredError)
 
-If you focus on only one accuracy index, [RMSE](#sec-rootMeanSquaredError) would be a good choice.
+If you focus on only one accuracy index, [MAE](#sec-meanAbsoluteError) or [RMSE](#sec-rootMeanSquaredError) would be a good choice.
 However, I would also examine a [calibration plot](#sec-calibrationPlot) to evaluate whether predictions are poorly calibrated at higher levels of points.
 I would also examine [ME](#sec-meanError)—not to compare the accuracy of various predictions per se—but to determine whether predictions are systematically under- or overestimating actual points.
-If so, predictions may be able to be refined by adding or subtracting a constant to the predictions; however, this could worsen other accuracy indices, so it is important to conduct an iterative process of modifying then evaluating, then further modifiying and evaluating, etc.
+If so, predictions may be able to be refined by adding or subtracting a constant to the predictions (or to a subset of the predictions); however, this could worsen other accuracy indices, so it is important to conduct an iterative process of modifying then evaluating, then further modifying and evaluating, etc.
 It may also be valuable to evaluate the accuracy of various subsets of the predictions.
-For instance, you might examine the predictive accuracy of players whose project points are greater than 100, to evaluate the accuracy of predictions specifically to distinguish between players at higher levels of points.
+For instance, you might examine the predictive accuracy of players whose projected points are greater than 100, to evaluate the accuracy of predictions specifically to distinguish between players at higher levels of points, which is one of the key goals when selecting which players to draft.
 
@@
   - [area under the receiver operating curve](#sec-auc)
-  - and, secondarily—depending on the particulary cutoff and the relative costs of false positives versus false negatives:
+  - and, secondarily—depending on the particulary cutoff and the relative costs of [false positives](#sec-falsePositive) versus [false negatives](#sec-falseNegative):
     - [sensitivity](#sec-sensitivity)

### 2024-09-18 cognitive-bias.qmd
20240917 - loss aversion & prospect theory

@@
 That is, loss aversion (similar to punishment sensitivity, as opposed to reward sensitivity) is the tendency to weigh losses (or punishment) more heavily than gains (or rewards).
+Loss aversion is a key aspect of prospect theory [@Kahneman2011].
+The loss aversion ratio indicates the extent to which a person weighs losses greater than gains, and for many people is between 1.5 to 2.5 [@Kahneman2011].
 Loss aversion is different from [risk aversion](#sec-cognitiveBiasesRiskAversion).
 Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
-In fantasy football, loss aversion may lead managers to start or hold onto underperforming high drafts for too long instead of a starting a more promising player out of fear of losing potential value from their initial investment.
+In fantasy football, loss aversion may lead managers to start or hold onto underperforming high drafted players for too long instead of starting a more promising player out of fear of losing potential value from their initial investment.
 
@@
 
-Risk seeking, by contrast, has been observed in entrepreneurs and in generals when all of their options are poor [@Kahneman2011].
-In fantasy football, risk seeking may be more likely when a manager has a team full of underperforming players.
+In mixed gambles, in which it is possible for a person to experience either a gain or a loss, [loss aversion](#sec-cognitiveBiasesLossAversion) tends to lead to risk-averse choices [@Kahneman2011]. 
+By contrast, when all of a person's options are poor, people tend to engage in risk seeking, as has been observed in entrepreneurs and in generals [@Kahneman2011].
+In fantasy football, risk seeking may be more likely when a manager has a team full of underperforming players and a weak record.
 

### 2024-09-18 cognitive-bias.qmd
20240918 - copyediting

@@
 Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
-In fantasy football, loss aversion may lead managers to start or hold onto underperforming high drafted players for too long instead of starting a more promising player out of fear of losing potential value from their initial investment.
+In fantasy football, loss aversion may lead managers to start or hold onto players for too long who were highly drafted yet are underperforming instead of starting a more promising player out of fear of losing potential value from their initial investment.
 

### 2024-09-18 download_weekly_projections.R
20240918 - separate script to download weekly projections

@@ -1,0 +1,77 @@
+# Download Weekly Projections
+
+# Libraries ----
+library("ffanalytics")
+library("tidyverse")
+
+# Download Weekly Projections ----
+players_projections_weekly_raw <- ffanalytics::scrape_data(
+  season = NULL, # NULL grabs the current season
+  week = NULL) # NULL grabs the current week
+
+# Save Raw Projections ----
+save(
+  players_projections_weekly_raw,
+  file = "./data/players_projections_weekly_raw.RData"
+)
+
+# Scoring Settings ----
+scoring_obj_default <- ffanalytics::scoring
+
+scoring_obj <- scoring_obj_default
+
+## Offense ----
+scoring_obj$pass$pass_int <- -2
+scoring_obj$rec$rec <- 1
+scoring_obj$misc$fumbles_lost <- -2
+
+## Kickers ----
+scoring_obj$kick$fg_4049 <- 3
+
+## Defense/Special Teams ----
+scoring_obj$pts_bracket <- list(
+  list(threshold = 0, points = 10),
+  list(threshold = 6, points = 7),
+  list(threshold = 13, points = 4),
+  list(threshold = 20, points = 1),
+  list(threshold = 27, points = 0),
+  list(threshold = 34, points = -1),
+  list(threshold = 99, points = -4)
+)
+
+# Calculate Projected Points ----
+
+## By Source ----
+players_projectedPoints_weekly <- ffanalytics:::impute_and_score_sources(
+  data_result = players_projections_weekly_raw,
+  scoring_rules = scoring_obj)
+
+## Averaged Across Sources ----
+players_projectedStatsAverage_weekly <- ffanalytics::projections_table(
+  players_projections_weekly_raw,
+  scoring_rules = scoring_obj,
+  return_raw_stats = TRUE)
+
+players_projectedPointsAverage_weekly <- ffanalytics::projections_table(
+  players_projections_weekly_raw,
+  scoring_rules = scoring_obj,
+  return_raw_stats = FALSE)
+
+# Merge ----
+players_projections_weekly_average <- full_join(
+  players_projectedPointsAverage_weekly,
+  players_projectedStatsAverage_weekly,
+  by = c("id","avg_type","pos" = "position")
+)
+
+# Add Additional Player Information ----
+players_projections_weekly_average <- players_projections_weekly_average %>% 
+  add_ecr() %>% 
+  #add_uncertainty() %>% # currently throws an error
+  add_player_info()
+
[…]

### 2024-09-18 download_weekly_projections.R
20240918 - script to download weekly projections

@@ -1,5 +1,14 @@
 # Download Weekly Projections
 
+# Specify Season and Week Number ----
+
+season <- 2024
+weekNumber <- 3
+
+# Processing ----
+season <- as.character(season)
+weekNumber <- as.character(weekNumber)
+
 # Libraries ----
 library("ffanalytics")
 library("tidyverse")
@@ -71,7 +80,13 @@ players_projections_weekly_average <- players_projections_weekly_average %>%
   add_player_info()
 
 # Save Projected Points ----
+
 save(
   players_projectedPoints_weekly, players_projections_weekly_average,
   file = "./data/players_projectedPoints_weekly.RData"
+)
+
+save(
+  players_projectedPoints_weekly, players_projections_weekly_average,
+  file = paste("./data/players_projectedPoints_weekly_", season, "_", "week", weekNumber, ".RData", sep = "")
 )

### 2024-09-18 download_weekly_projections.R
20240918 - save weekly projections

@@ -88,5 +88,5 @@ save(
 
 save(
   players_projectedPoints_weekly, players_projections_weekly_average,
-  file = paste("./data/players_projectedPoints_weekly_", season, "_", "week", weekNumber, ".RData", sep = "")
+  file = paste("./data/by_year/players_projectedPoints_weekly_", season, "_", "week", weekNumber, ".RData", sep = "")
 )

### 2024-09-18 decision-making.qmd
20240918 - betting markets for football

@@
 - <https://polymarket.com>
+
+There are also betting markets for football:
+
 - <https://www.rotowire.com/betting/nfl/player-futures.php>
+- <https://vegasprojections.com>
+- <https://www.actionnetwork.com/nfl/props>
 

### 2024-09-19 cognitive-bias.qmd
20240918 - loss aversion and endowment effect

@@
 That is, loss aversion (similar to punishment sensitivity, as opposed to reward sensitivity) is the tendency to weigh losses (or punishment) more heavily than gains (or rewards).
-Loss aversion is a key aspect of prospect theory [@Kahneman2011].
+Thus, the target state matters less than the *change* (i.e., gains and/or losses) from one's reference point [e.g., their current roster; @Kahneman2011].
+Loss aversion is a key aspect of prospect theory and suggests that individuals tend to be biased in favor of their reference situation and toward making smaller rather than larger changes [@Kahneman2011].
+Loss aversion explains why people tend to favor the favoring the status quo; the negative consequences of a change often loom larger than the positive consequences of a change [@Kahneman2011].
+
 The loss aversion ratio indicates the extent to which a person weighs losses greater than gains, and for many people is between 1.5 to 2.5 [@Kahneman2011].
@@
 
-*Endowment effect bias* is the tendency to overvalue merely because one owns it.
+*Endowment effect bias* is the tendency to overvalue merely because one owns it—owning it increases its value to the owner [especially for goods that are not regularly traded; @Kahneman2011].
 That is, a manager tends to value the same player more when they play for their team rather than an opponent's team.
+It is related to [loss aversion bias](#sec-cognitiveBiasesLossAversion) and prospect theory; when owning a good (i.e., a person's reference point), they focus on the *loss* of the good if they were they to sell or trade it away rather than on the gains from the sale or trade.
 When trading, the endowment effect bias might lead a manager to demand more to trade away a player than to acquire the same player [@Miller2013c].

### 2024-09-19 data-visualization.qmd
20240919 - notes from Schwabish (2021)

@@
 
+"Showing the data" involves showing the data that matters the most.
+"Reducing the clutter" involves removing non-data things that obscure the data—for example, extraneous gridlines, tick marks, data markers (e.g., symbols to distinguish between series), and complex shadings (e.g., textured or filled gradients).
+"Integrating the graphics and text" involves using headline titles, clear and useful labels (instead of legends), and helpful annotations.
+Headline or newspaper-like titles are titles that are succinct with active phasing and that indicate the take-away message (e.g., "Quarterbacks Threw Fewer Touchdowns in 2024 than in Previous Years").
+In terms of labels, @Schwabish2021 advocates to label the data directly instead of using a legend.
+In terms of helpful annotations, you can provide additional text that helps explain the data (e.g., peaks or valleys, outliers, or other variations that deserve explanation), including how to interpret the chart.
+"Avoiding the spaghetti chart" means avoiding packed charts with too much information that makes them difficult to interpret.
+Spaghetti charts are lines with many lines that, make the plot look like a bunch of spaghetti.
+However, @Schwabish2021 also advocates against using charts of other types that are complicated and difficult to interpret due to too much information, such as complicated maps or bar plots with too many colors, icons, or bars.
+If there are too many lines or series, @Schwabish2021 advocates breaking it up into multiple charts (i.e., facets, trellis charts, or small multiples).
+"Starting with gray" refers to the idea of using gray as the default color for most lines/points/bars, so that you can use a color to highlight the lines/points/bars of interest.
+In addition, as noted by @Schwabish2021, it is important to treat data as objectively as possible and not to present figures in a biased way as to mislead.
+
 In his classic book, @Tufte2001 states that effective data visualizations should follow principles of graphical excellence and integrity.
@@
 That is, data visualizations should seek to maximize the data-to-ink ratio (within reason), and should spend less space on "fluff" (i.e., non-data things that can be erased without losing meaning, such as grid lines, redundancies, etc.).
+This is consistent with Schwabish's -@Schwabish2021 principles of showing the data and reducing the clutter.
 @Tufte2001 describes six principles of graphical integrity:
@@
 Each visual component (e.g., line) that is important to see should be easy to distinguish.
-For instance, you can highlight lines or points of interest to draw people's attention to the target of interest.
+For instance, you can highlight lines or points of interest to draw people's attention to the target of interest [@Schwabish2021].
 For examples of highlighting in figures, see Figures [-@fig-lineChartHighlighting] (@sec-lineChartHighlighting) and [-@fig-simulationOf10CoinFlips].

### 2024-09-19 data-visualization.qmd
20240919 - box-and-whisker plot

@@
 
+In a box-and-whisker plot, the box is created using the 1st and 3rd quartiles (i.e., the 25th and 75th percentiles, respectively).
+The length of the box is equal to the interquartile range, which is calculated as: $\text{IQR} = Q_3 - Q_1$, where $Q_3$ and $Q_1$ are the third and first quartiles, respectively.
+The line in the middle of the box is located at the median (i.e., the 2nd quartile or 50th percentile).
+The whiskers commonly extend $1.5 \times \text{IQR}$ units from the box.
+That is, the upper whisker is commonly located at $1.5 \times \text{IQR}$ units above the third quartile.
+The lower whisker is commonly located at $1.5 \times \text{IQR}$ units below the first quartile.
+The points represent extreme values (i.e., outliers) that are outside the whiskers.
+
 A box-and-whisker plot of fantasy points is depicted in @fig-boxAndWhisker.
@@
 ) +
-  geom_boxplot() +
+  geom_boxplot(staplewidth = 0.25) +
   labs(

### 2024-09-20 cognitive-bias.qmd
20240920 - loss aversion

@@
 Loss aversion explains why people tend to favor the favoring the status quo; the negative consequences of a change often loom larger than the positive consequences of a change [@Kahneman2011].
+Aversion to losses and potential threats may be related to humans' survival instinct [@Kahneman2011].
+With par as a reference point, this may explain why professional golfers tend to putt more accurately for par than for birdie, even for putts of equivalent distance [@Kahneman2011].
 
-The loss aversion ratio indicates the extent to which a person weighs losses greater than gains, and for many people is between 1.5 to 2.5 [@Kahneman2011].
+The loss aversion ratio indicates the extent to which a person weighs losses greater than gains, and for many people is around 2 (typically between 1.5 to 2.5; [@Kahneman2011].
+That is, people tend to give as much twice the weight to losses as to gains.
 Loss aversion is different from [risk aversion](#sec-cognitiveBiasesRiskAversion).
@@
 In fantasy football, loss aversion may lead managers to start or hold onto players for too long who were highly drafted yet are underperforming instead of starting a more promising player out of fear of losing potential value from their initial investment.
+Loss aversion can also influence trade negotiations.
 

### 2024-09-20 cognitive-bias.qmd
20240920 - prospect theory

@@
 
+Principles of prospect theory and loss aversion bias are depicted in @fig-prospectTheory.
+
+::: {#fig-prospectTheory}
+![](images/prospectTheory.png)
+
+Prospect Theory and Loss Aversion: People Tend to Weigh Losses More Heavily Than Gains. Adapted from <https://commons.wikimedia.org/wiki/File:Graphique_th%C3%A9orie_des_perspectives.svg>.
+:::
+
 The loss aversion ratio indicates the extent to which a person weighs losses greater than gains, and for many people is around 2 (typically between 1.5 to 2.5; [@Kahneman2011].
@@
 
-Dunning–Krueger Effect: Confidence as a Function of Competence. Adapted from <https://commons.wikimedia.org/wiki/File:Effet_Dunning-Kruger.svg>.
+Dunning–Krueger Effect: Confidence as a Function of Competence.
+People with low competency in a task tend to overestimate their ability.
+Adapted from <https://commons.wikimedia.org/wiki/File:Effet_Dunning-Kruger.svg>.
 :::
@@
 
-Dunning–Krueger Effect: Perceived Performance as a Function of Actual Performance. Adapted from <https://commons.wikimedia.org/wiki/File:Dunning-kruger_effect_-_percentile.svg>.
+Dunning–Krueger Effect: Perceived Performance as a Function of Actual Performance.
+People who perform poorly in a task tend to overestimate their performance.
+Adapted from <https://commons.wikimedia.org/wiki/File:Dunning-kruger_effect_-_percentile.svg>.
 :::

### 2024-09-20 data-visualization.qmd
20240920 - placeholder for example
20240920 - example data visualization

@@
 - use text that is upper-and-lower case, not all capitals
+
+An example figure that applies these principles of data visualization is in @fig-exampleDataVisualization.
+
+```{r}
+#| label: fig-exampleDataVisualization
+#| fig-cap: "Example Figure that Applies the Principles of Data Visualization."
+#| code-fold: true
+
+confidenceLevel <- .95 # for 95% confidence interval
+
+player_stats_seasonal_offense_summary <- player_stats_seasonal_offense %>%
+    filter(position_group %in% c("QB","RB","WR","TE")) %>% 
+  group_by(position_group) %>%
+  summarise( 
+    n = sum(!is.na(fantasy_points)),
+    mean = mean(fantasy_points, na.rm = TRUE),
+    sd = sd(fantasy_points, na.rm = TRUE)
+  ) %>%
+  mutate(se = sd/sqrt(n)) %>%
+  mutate(
+    ci_lower = mean - qt(p = 1 - (1 - confidenceLevel) / 2, df = n - 1) * se,
+    ci_upper = mean + qt(p = 1 - (1 - confidenceLevel) / 2, df = n - 1) * se,
+    positionLabel = case_match(
+      position_group,
+      "QB" ~ "Quarterback",
+      "RB" ~ "Running Back",
+      "WR" ~ "Wide Receiver",
+      "TE" ~ "Tight End"
+      )
+  )
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense_summary %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = positionLabel,
+    y = mean,
+    fill = positionLabel
+    )
+) +
+  geom_bar(
+    stat = "identity") +
+  geom_errorbar(
+    aes(
+      ymin = ci_lower,
+      ymax = ci_upper),
+    width = 0.2,
+    color = "black"
+  ) +
+  gghighlight::gghighlight(
+    positionLabel == "Quarterback",
+    label_key = positionLabel) +
+  labs(
+    x = "Position",
+    y = "Fantasy Points",
+    title = "Quarterbacks Score More Fantasy Points than Other Positions"
+  ) +
+  annotate(
+    "segment",
+    x = 3.5,
+    xend = 3.2,
+    y = 70,
+    yend = 35,
+    color = "blue",
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow()) +
+  annotate(
+    "text",
+    x = 2.75,
+    y = 75,
+    label = "Tight Ends score fewer points\nthan other positions",
+    hjust = 0) + # left-justify
+  theme_classic() + 
+  theme(legend.position = "none") +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
 

### 2024-09-20 data-visualization.qmd
20240920 - minor edit

@@
     y = 75,
-    label = "Tight Ends score fewer points\nthan other positions",
+    label = "Tight Ends score fewer fantasy\npoints than other positions",
     hjust = 0) + # left-justify

### 2024-09-21 cognitive-bias.qmd
20240921 - base rates

@@
 
-Prospect Theory and Loss Aversion: People Tend to Weigh Losses More Heavily Than Gains. Adapted from <https://commons.wikimedia.org/wiki/File:Graphique_th%C3%A9orie_des_perspectives.svg>.
+Prospect Theory and Loss Aversion: People Tend to Weigh Losses More Heavily Than Gains.
+That is, the utility function is steeper in the loss (red) region than in the gain (blue) region.
+Adapted from <https://commons.wikimedia.org/wiki/File:Graphique_th%C3%A9orie_des_perspectives.svg>.
 :::
@@
 The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
+The, a "low base-rate" event is an event with a low probability of occurring (i.e., an unlikely event).
 Focusing on the specific information about the event (and thus underweighting the base rate) involves the [representativeness heuristic](#sec-heuristicsRepresentativeness) [@Kahneman2011]—making judgments based on how similar the event is to a mental prototype regardless of how frequent the event the event actually is.
 As noted by @Kahneman2011, "people who have information about an individual case rarely feel the need to know the statistics of the class which the case belongs." (p. 249).
-In general, people tend to overestimate the likelihood of low base-rate events [@Kahneman2011].
+In general, people tend to a) overestimate the likelihood of low base-rate events and b) overweight low base-rate events in their decisions [@Kahneman2011].
+This is especially true if the unlikely event attracts attention, because of the [confirmatory bias](#sec-cognitiveBiasesConfirmation) of memory [@Kahneman2011].
 
@@
 Indeed, even men (36.7%) and self-employed individuals (41.2%) have less than a 50% chance of experiencing a work-related injury.
+
+The [framing](#sec-cognitiveBiasesFraming) of an event can also alter its weighting.
+For instance, concrete representations tend to be weighted more heavily than abstract representations [@Kahneman2011].
+Abstract representations are framed in terms of "how likely" the event will occur in in terms of chance, risk, probability, or likelihood.
+Concrete representations are framed in terms of  "how many" people will be affected by the event.
+An example abstract representation is that there is a 0.001% chance of dying from this particular disease.
+When framed in this way, the risk appears small.
+Now, consider the same information framed in terms of a concrete representation: 1 in 100,000 people will die from this disease.
+Due to a phenomenon called *denominator neglect*, low base-rate events are more heavily weighted when they are framed using a concrete representation than an abstract representation [@Kahneman2011].
+
+There are low base-rate events that tend to be underweighted when making decisions, such as whether to make anticipatory preparations for a future earthquake, housing bubble, or financial crisis [@Kahneman2011].
+One major cause of underweighting is if a person has never experienced the rare event (i.e., the earthquake or finanical crisis) [@Kahneman2011].
+However, for low base-rate events that people have experienced, the [availability](#sec-heuristicsAvailability) and [representativeness](#sec-heuristicsRepresentativeness) heuristics tend to lead people to overestimate the likelihood that the unlikely event occurs.
+In general, many factors—including obsessive concerns about the event, vivid images of the event, concrete representations of the consequences of the event (e.g., "a disease kills 1 person out of every 1,000"), and explicit reminders of the event—contribute to the overweighting of unlikely events [@Kahneman2011].
 

### 2024-09-22 cognitive-bias.qmd
20240922 - omission bias and regret

@@
 
-*Omission bias* is the tendency to prefer inaction over action.
+*Omission bias* is the tendency to prefer inaction over action, or to prefer the default option over changes from the default option [@Kahneman2011].
 For instance, consider a manager who is on the fence to drop an underperforming player and to pick up someone else to start.
-If the manager does not drop the player and the player scores few points, the manager is likely to be less critical of their decision than if they had picked up another player who then scored the same (low) number of points.
+If the manager does not drop the player and the player scores few points, the manager is likely to be less critical and regretful of their decision than if they had picked up another player who then scored the same (low) number of points.
+Thus, a person's decisions may be influenced by the anticipation of regret so as to keep the status quo, avoid losses, and minimize the expectation of regret—such as choosing the same order each time at a restaurant to avoid regret [@Kahneman2011].
+However, people may anticipate more regret than they actually experience, because people tend to underestimate the protective impact of the psychological defenses they will deploy against regret [@Kahneman2011].
+In general, @Kahneman2011 recommends not to put too much weight on potential future regret when making decisions: "even if you have some [regret], it will hurt less than you now think." (p. 352).
 
@@
 For instance, when ordering the expensive meal at the restaurant, a person may over-eat so that they feel that they eat their money's worth of food.
+The sunk cost fallacy is an example of the commitment bias, an escalation of commitment by which a person continues to commit additional time, effort, or resources to a failing project.
+It is important to be willing to identify and terminate failing projects to cut one's losses.
 

### 2024-09-22 cognitive-bias.qmd
20240922 - additional info

@@
 They lead us to be less accurate in our judgments and predictions.
-Inaccuracy may not always be bad; several cognitive biases may serve to help people feel better about themselves [@Miller2013c], including [confirmation bias](#sec-cognitiveBiasesConfirmation), [overconfidence bias](#sec-cognitiveBiasesOverconfidence), [optimism bias](#sec-cognitiveBiasesOptimism), and [self-serving bias](#sec-cognitiveBiasesSelfServing).
+Inaccuracy may not always be bad; several cognitive biases may serve to help people feel better about themselves [@Miller2013c], including [confirmation bias](#sec-cognitiveBiasesConfirmation), [overconfidence bias](#sec-cognitiveBiasesOverconfidence), [optimism bias](#sec-cognitiveBiasesOptimism), and [self-serving bias](#sec-cognitiveBiasesSelfServing); or to prevent feelings of regret (e.g., [omission bias](#sec-cognitiveBiasesOmission))—or to prevent having to admit failure (e.g., [commitment bias](#sec-fallaciesSunkCost)).
 In fantasy football, however, cognitive biases and inaccurate judgments can lead to worse performance.
@@
 If the manager does not drop the player and the player scores few points, the manager is likely to be less critical and regretful of their decision than if they had picked up another player who then scored the same (low) number of points.
-Thus, a person's decisions may be influenced by the anticipation of regret so as to keep the status quo, avoid losses, and minimize the expectation of regret—such as choosing the same order each time at a restaurant to avoid regret [@Kahneman2011].
+Thus, a person's decisions may be influenced by the anticipation or fear of regret so as to keep the status quo, avoid losses, and minimize the expectation of regret—such as choosing the same order each time at a restaurant to avoid regret [@Kahneman2011].
 However, people may anticipate more regret than they actually experience, because people tend to underestimate the protective impact of the psychological defenses they will deploy against regret [@Kahneman2011].
@@
 The more times a player touches the ball, the more points they are likely to score.
+And do not let the anticipation of regret prevent you from making important decisions.
 
@@
 For instance, when ordering the expensive meal at the restaurant, a person may over-eat so that they feel that they eat their money's worth of food.
-The sunk cost fallacy is an example of the commitment bias, an escalation of commitment by which a person continues to commit additional time, effort, or resources to a failing project.
+The sunk cost fallacy is an example of the *commitment bias*, an *escalation of commitment* by which a person continues to commit additional time, effort, or resources to a failing project.
+People tend to refuse to cut their losses when doing so would admit failure [@Kahneman2011].
 It is important to be willing to identify and terminate failing projects to cut one's losses.

### 2024-09-23 data-visualization.qmd
20240920 - remove extra line
20240923 - faceting

@@
 However, @Schwabish2021 also advocates against using charts of other types that are complicated and difficult to interpret due to too much information, such as complicated maps or bar plots with too many colors, icons, or bars.
-If there are too many lines or series, @Schwabish2021 advocates breaking it up into multiple charts (i.e., facets, trellis charts, or small multiples).
+If there are too many lines or series, @Schwabish2021 advocates [breaking it up into multiple charts](#sec-faceting) (i.e., facets, trellis charts, or small multiples).
+An example of [faceted charts](#sec-faceting) is depicted in @fig-facetedScatterplot.
 "Starting with gray" refers to the idea of using gray as the default color for most lines/points/bars, so that you can use a color to highlight the lines/points/bars of interest.
@@
 
+## Faceting {#sec-faceting}
+
+```{r}
+#| label: fig-facetedScatterplot
+#| fig-cap: "Faceted Scatterplot."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal_offense %>% 
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  aes(
+    x = age,
+    y = fantasy_points)) +
+  geom_point() +
+  geom_smooth() +
+  scale_x_continuous(
+    expand = c(0,0)
+  ) +
+  scale_y_continuous(
+    expand = c(0,0),
+    lim = c(0,NA)
+  ) +
+  labs(
+    x = "Player's Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age"
+  ) +
+  theme_bw() +
+  facet_wrap(vars(position_group)) # facet by position_group
+```
+
 ## Examples {#sec-scatterplotExamples}

### 2024-09-25 download_weekly_projections.R
20240925 - download week 4 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2024
-weekNumber <- 3
+weekNumber <- 4
 
 # Processing ----
 season <- as.character(season)

### 2024-09-29 cognitive-bias.qmd
20240929 - addressing cognitive biases

@@
 - [Dunning–Kruger effect bias](#sec-cognitiveBiasesDunningKruger)
+
+In general, when faced with a situation in which bias is likely, @Kahneman2011 advises slowing down, being thoughtful and deliberate, questioning your intuitions, and imposing orderly procedures, such as checklists, reference-class forecasting, and a "premortem".
+A premortem is when you consider, in advance of the decision, what likely went wrong if the decision were to work out terribly.
+Reference-class forecasting involves examining a reference class of similar projects to identify the most likely range of outcomes for the project.
 

### 2024-10-08 data-visualization.qmd
20241008 - data visualization catalogue

@@
 
-The R Graph Gallery provides examples of various types of plots: <https://r-graph-gallery.com>.
+The Data Visualization Catalogue provides examples of various types of plots depending on one's goal: <https://datavizcatalogue.com/search.html>.
+The R Graph Gallery provides examples of various types of plots and how to create them in `R`: <https://r-graph-gallery.com>.
 Books on data visualization in `R` include [*`ggplot2`: Elegant Graphics for Data Analysis*](https://ggplot2-book.org) [@ggplot3rdEd] and [*`R` Graphics Cookbook: Practical Recipes for Visualizing Data*](https://r-graphics.org) [@Chang2018].

### 2024-10-09 download_weekly_projections.R
20241009 - week 6

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2024
-weekNumber <- 5
+weekNumber <- 6
 
 # Processing ----
 season <- as.character(season)

### 2024-10-10 data-visualization.qmd
20241010 - update season from 2023 to 2024

@@
   dplyr::filter(
-    season == 2023,
+    season == 2024,
     season_type == "REG") %>%
@@
 
-Defensive EPA per play during the 2023 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA.
+Defensive EPA per play during the 2024 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA.
 
@@
 #| label: fig-plotDefOffEPA
-#| fig-cap: "2023 NFL Offensive and Defensive EPA Per Play."
+#| fig-cap: "2024 NFL Offensive and Defensive EPA Per Play."
 
@@
     y = "Defense EPA/play",
-    title = "2023 NFL Offensive and Defensive EPA per Play"
+    title = "2024 NFL Offensive and Defensive EPA per Play"
   ) +

### 2024-10-10 sports-cognitive-psychology.qmd
20241010 - update season from 2023 to 2024

@@
 text_df <- tibble(
-  label = c("NFL coaches<br>in <span style='color:#00BFC4'>**2023**</span>", "NFL coaches<br>in <span style='color:#F8766D'>**2014**</span>"),
+  label = c("NFL coaches<br>in <span style='color:#00BFC4'>**2024**</span>", "NFL coaches<br>in <span style='color:#F8766D'>**2014**</span>"),
   x = c(6, 8.2),
@@
 nfl_4thdown %>%
-  filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2014, 2023)) %>%
+  filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2014, 2024)) %>%
   ggplot(aes(go_boost, go, color = as.factor(season))) + 
@@
        y = "Go-for-it percentage",
-       subtitle = "4th down decisions in 2023 versus 2014, win prob. > 20%",
+       subtitle = "4th down decisions in 2024 versus 2014, win prob. > 20%",
        title = glue::glue("How <span style='color:red'>math</span> is changing football")) +

### 2024-10-16 download_weekly_projections.R
20241016 - week 7 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2024
-weekNumber <- 6
+weekNumber <- 7
 
 # Processing ----
 season <- as.character(season)

### 2024-10-16 download-football-data.qmd
20241010 - update season from 2023 to 2024
20241016 - delete obsolete file reference; use fixed_drive instead of drive

@@
 load(file = "./data/players_projectedPoints_seasonal.RData")
-load(file = "./data/players_projections_seasonal_average.RData")
 
@@
 The `nfl_pbp` object is in `game`-`drive`-`play` form.
-That is, each row should be uniquely identified by the combination of `game_id`, `drive`, `play_id`.
+That is, each row should be uniquely identified by the combination of `game_id`, `fixed_drive`, `play_id`.
 Let's rearrange the data accordingly:
@@
 nfl_pbp %>% 
-  group_by(game_id, drive, play_id) %>% 
+  group_by(game_id, fixed_drive, play_id) %>% 
   filter(n() > 1) %>% 
@@
 
-Advanced stats from the 2023 season are available at the following link: <https://www.pro-football-reference.com/years/2023/advanced.htm>
+Advanced stats from the 2024 season are available at the following link: <https://www.pro-football-reference.com/years/2024/advanced.htm>
 

### 2024-10-16 base-rates.qmd
20241013 - specify heading
20241014 - reduce use of abbreviations
20241016 - render the figure caption as alt text

@@
 
-#### Example 3A {#sec-nateSilverExample3A}
+#### Example 3A: The First Plane Hit the World Trade Center {#sec-nateSilverExample3A}
 
@@
 
-#### Example 3B {#sec-nateSilverExample3B}
+#### Example 3B: The Second Plane Hit the World Trade Center {#sec-nateSilverExample3B}
 
@@
 ::: {#fig-probabilityNomogram}
-![](images/probabilityNomogram.png){width=50%}
+![](images/probabilityNomogram.png){width=50%}\
 
@@
 ::: {#fig-probabilityNomogramLine}
-![](images/probabilityNomogramLine.png){width=50%}
+![](images/probabilityNomogramLine.png){width=50%}\
 

### 2024-10-16 causal-inference.qmd
20240815 - Simpson's paradox
20240816 - clean up references
20241016 - render the figure caption as alt text

@@
 ::: {#fig-crossSectional}
-![](images/Longitudinal-01.png){width=20%}
+![](images/Longitudinal-01.png){width=20%}\
 
@@
 ::: {#fig-laggedAssociation}
-![](images/Longitudinal-02.png)
+![](images/Longitudinal-02.png)\
 
@@
 ::: {#fig-laggedAssociationControlPriorLevels}
-![](images/Longitudinal-03.png)
+![](images/Longitudinal-03.png)\
 
@@
 ::: {#fig-crossLaggedPanelModel}
-![](images/Longitudinal-04.png)
+![](images/Longitudinal-04.png)\
 
@@
 ::: {#fig-simpsonParadox}
-![](images/simpsonParadox.png){width=200%}
+![](images/simpsonParadox.png){width=200%}\
 
@@
 (2013).
-Simpson's paradox in psychological science: a practical guide.
+Simpson's paradox in psychological science: A practical guide.
 *Frontiers in Psychology*, *4*(513).
@@
 
-However, it can be problematic to control for variables indiscriminantly.
+However, it can be problematic to control for variables indiscriminantly [@Spector2010; @Wysocki2022].
 The use of [causal diagrams](#sec-causalDiagrams) can inform which variables are important to be included as control variables, and—just as important—which variables not to include as control variables, as described in @sec-causalDiagrams.

### 2024-10-16 fantasy-football.qmd
20241016 - render the figure caption as alt text

@@
 
-![Video that Provides An Introduction to American Football. From: <https://www.youtube.com/watch?v=Ddwp1HyEFRE>.](images/football.mp4){width=100%}
+![Video that Provides An Introduction to American Football. From: <https://www.youtube.com/watch?v=Ddwp1HyEFRE>.](images/football.mp4){width=100%}\
 
@@
 ::: {#fig-footballFormation}
-![](images/footballFormation.png)
+![](images/footballFormation.png)\
 
@@
 ::: {#fig-footballField}
-![](images/footballField.png){width=50%}
+![](images/footballField.png){width=50%}\
 
@@
 
-![Video that Provides an Introduction to Fantasy Football. From: <https://www.youtube.com/watch?v=XhrBapdhLEc>.](images/fantasyFootball.mp4){width=100%}
+![Video that Provides an Introduction to Fantasy Football. From: <https://www.youtube.com/watch?v=XhrBapdhLEc>.](images/fantasyFootball.mp4){width=100%}\
 

### 2024-10-16 index.qmd
20241016 - render the figure caption as alt text

@@
 
-![Creative Commons License](images/CC-BY.png){width=25% fig-align="left"}
+![Creative Commons License](images/CC-BY.png){width=25% fig-align="left"}\
 
@@
 <!--
-[![DOI](images/zenodo_doiBadge.png)](https://doi.org/10.5281/zenodo.6466589)
+[![DOI](images/zenodo_doiBadge.png)](https://doi.org/10.5281/zenodo.6466589)\
 -->

### 2024-10-16 multiple-regression.qmd
20241016 - render the figure caption as alt text

@@
 ::: {#fig-regression}
-![](images/regression.png)
+![](images/regression.png)\
 
@@
 ::: {#fig-regression}
-![](images/multipleRegressionRSquared.png){width=50%}
+![](images/multipleRegressionRSquared.png){width=50%}\
 
@@
 ::: {#fig-regression}
-![](images/multipleRegressionMulticollinearity.png){width=50%}
+![](images/multipleRegressionMulticollinearity.png){width=50%}\
 

### 2024-10-16 basic-statistics.qmd
20241016 - render the figure caption as alt text
20241016 - add alt-text to figures

@@
 #| fig-cap: "Interpretation of *p*-Values When Examining The Differences Between Groups. The vertical black lines reflect the group means."
+#| fig-alt: "Interpretation of *p*-Values When Examining The Differences Between Groups. The vertical black lines reflect the group means."
 #| fig-subcap: 
@@
 #| fig-cap: "Interpretation of *p*-Values When Examining The Association Between Variables."
+#| fig-alt: "Interpretation of *p*-Values When Examining The Association Between Variables."
 #| fig-subcap: 
@@
 ::: {#fig-nhstConfusionMatrix}
-![](images/nhstConfusionMatrix.png)
+![](images/nhstConfusionMatrix.png)\
 
@@
 #| fig-cap: "Statistical Power (Adapted from Kristoffer Magnusson: <https://rpsychologist.com/creating-a-typical-textbook-illustration-of-statistical-power-using-either-ggplot-or-base-graphics>; archived at <https://perma.cc/FG3J-85L6>).
+The dashed line represents the critical value or threshold."
+#| fig-alt: "Statistical Power (Adapted from Kristoffer Magnusson: <https://rpsychologist.com/creating-a-typical-textbook-illustration-of-statistical-power-using-either-ggplot-or-base-graphics>; archived at <https://perma.cc/FG3J-85L6>).
 The dashed line represents the critical value or threshold."
@@
 ::: {#fig-statisticalDecisionTree}
-![](images/statisticalDecisionTree.png)
+![](images/statisticalDecisionTree.png)\
 
@@
 ::: {#fig-statisticalDecisionTreeSummary}
-![](images/statisticalDecisionTreeSummary.png)
+![](images/statisticalDecisionTreeSummary.png)\
 
@@
 ::: {#fig-statisticalDecisionTreeRegression}
-![](images/statisticalDecisionTreeRegression.png)
+![](images/statisticalDecisionTreeRegression.png)\
 
@@
 ::: {#fig-statisticalDecisionTreeSummaryRegression}
-![](images/statisticalDecisionTreeSummaryRegression.png)
+![](images/statisticalDecisionTreeSummaryRegression.png)\
 

### 2024-10-16 causal-inference.qmd
20241016 - add alt-text to figures

@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting `X` Causing `Y`."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Depicting `X` Causing `Y`."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting Reverse Causation: `Y` Causing `X`."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Depicting Reverse Causation: `Y` Causing `X`."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, `Z`, Causing `X` and `Y`, Thus Explaining Why `X` and `Y` are associated."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Depicting a Third Variable Confound, `Z`, Causing `X` and `Y`, Thus Explaining Why `X` and `Y` are associated."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph)."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph)."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Confounding."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Example of Confounding."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Full Mediation."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Example of Full Mediation."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Partial Mediation."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Example of Partial Mediation."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Ancestor (Mediation) and Descendant of Y."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Example of Ancestor (Mediation) and Descendant of Y."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of a Collision with a Collider (Injury Status)."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Example of a Collision with a Collider (Injury Status)."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Collider Bias."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Example of Collider Bias."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of M-Bias."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Example of M-Bias."
 
@@
 #| fig-cap: "Causal Diagram (Directed Acyclic Graph) Example of Butterfly Bias."
+#| fig-alt: "Causal Diagram (Directed Acyclic Graph) Example of Butterfly Bias."
 

### 2024-10-16 cluster-analysis.qmd
20241016 - add alt-text to figures

@@
 #| fig-cap: "Quarterback Fantasy Points by Tier."
+#| fig-alt: "Quarterback Fantasy Points by Tier."
 
@@
 #| fig-cap: "Running Back Fantasy Points by Tier."
+#| fig-alt: "Running Back Fantasy Points by Tier."
 
@@
 #| fig-cap: "Tight End Fantasy Points by Tier."
+#| fig-alt: "Tight End Fantasy Points by Tier."
 

### 2024-10-16 cognitive-bias.qmd
20241016 - render the figure caption as alt text
20241016 - add alt-text to figures

@@
 ![Video of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends.
-From: <https://x.com/GetUpESPN/status/1380165126108672001> (archived at <https://perma.cc/JW8E-KV2C>)](images/kylePitts.mp4){width=100%}
+From: <https://x.com/GetUpESPN/status/1380165126108672001> (archived at <https://perma.cc/JW8E-KV2C>)](images/kylePitts.mp4){width=100%}\
 
@@
 ::: {#fig-prospectTheory}
-![](images/prospectTheory.png)
+![](images/prospectTheory.png)\
 
@@
 ::: {#fig-dunningKrueger1}
-![](images/dunningKrueger_1.png)
+![](images/dunningKrueger_1.png)\
 
@@
 ::: {#fig-dunningKrueger2}
-![](images/dunningKrueger_2.png)
+![](images/dunningKrueger_2.png)\
 
@@
 #| fig-cap: "Histogram of Number of Coins that Flip Heads-Up in a Simulation of 10 Coin Flips (with 100,000 Replications)."
+#| fig-alt: "Histogram of Number of Coins that Flip Heads-Up in a Simulation of 10 Coin Flips (with 100,000 Replications)."
 #| label: fig-simulationOf10CoinFlips

### 2024-10-16 correlation.qmd
20240720 - assumptions
20241016 - add alt-text to figures

@@
 #| fig-cap: "Correlation Coefficients."
+#| fig-alt: "Correlation Coefficients."
 #| fig-height: 12
@@
 #| fig-cap: "Interpretation of the Magnitude and Direction (Sign) of Correlation Coefficients."
+#| fig-alt: "Interpretation of the Magnitude and Direction (Sign) of Correlation Coefficients."
 #| fig-height: 9
@@
 
-## Correlation Does Not Imply Causation
+## Impact of Outliers {#sec-correlationOutliers}
+
+## Correlation Does Not Imply Causation {#sec-correlation-correlationAndCausation}
 

### 2024-10-16 data-visualization.qmd
20241016 - add alt-text to figures

@@
 #| fig-cap: "Example Figure that Applies the Principles of Data Visualization."
+#| fig-alt: "Example Figure that Applies the Principles of Data Visualization."
 #| code-fold: true
@@
 #| fig-cap: "Histogram of Fantasy Points."
+#| fig-alt: "Histogram of Fantasy Points."
 
@@
 #| fig-cap: "Density Plot of Fantasy Points by Position."
+#| fig-alt: "Density Plot of Fantasy Points by Position."
 
@@
 #| fig-cap: "Histogram with Overlaid Density and Rug Plot."
+#| fig-alt: "Histogram with Overlaid Density and Rug Plot."
 
@@
 #| fig-cap: "Box-and-Whisker Plot."
+#| fig-alt: "Box-and-Whisker Plot."
 
@@
 #| fig-cap: "Violin Plot. Lines represent the 25th, 50th, and 75th quantiles."
+#| fig-alt: "Violin Plot. Lines represent the 25th, 50th, and 75th quantiles."
 
@@
 #| fig-cap: "Ridgeline Plot."
+#| fig-alt: "Ridgeline Plot."
 
@@
 #| fig-cap: "Ridgeline Plot. Lines represent the 25th, 50th, and 75th quantiles."
+#| fig-alt: "Ridgeline Plot. Lines represent the 25th, 50th, and 75th quantiles."
 
@@
 #| fig-cap: "Base Plot."
+#| fig-alt: "Base Plot."
 
@@
 #| fig-cap: "Scatterplot."
+#| fig-alt: "Scatterplot."
 
@@
 #| fig-cap: "Scatterplot with Linear Best-Fit Line."
+#| fig-alt: "Scatterplot with Linear Best-Fit Line."
 
@@
 #| fig-cap: "Scatterplot with Quadratic Best-Fit Line."
+#| fig-alt: "Scatterplot with Quadratic Best-Fit Line."
 
@@
 #| fig-cap: "Scatterplot with Best-Fit Line Using Locally Estimated Scatterplot Smoothing (LOESS)."
+#| fig-alt: "Scatterplot with Best-Fit Line Using Locally Estimated Scatterplot Smoothing (LOESS)."
 
@@
 #| fig-cap: "Scatterplot with Best-Fit Line from Generalized Additive Model."
+#| fig-alt: "Scatterplot with Best-Fit Line from Generalized Additive Model."
 
@@
 #| fig-cap: "Scatterplot with Modified Axes."
+#| fig-alt: "Scatterplot with Modified Axes."
 
@@
 #| fig-cap: "Scatterplot with Plot Labels."
+#| fig-alt: "Scatterplot with Plot Labels."
 
@@
 #| fig-cap: "Scatterplot with Classic Theme."
+#| fig-alt: "Scatterplot with Classic Theme."
 
@@
 #| fig-cap: "Scatterplot with Dark Theme."
+#| fig-alt: "Scatterplot with Dark Theme."
 
@@
 #| fig-cap: "Interactive Scatterplot Using Plotly."
+#| fig-alt: "Interactive Scatterplot Using Plotly."
 
@@
 #| fig-cap: "Line Chart."
+#| fig-alt: "Line Chart."
 
@@
 #| fig-cap: "Line Chart with Highlighting."
+#| fig-alt: "Line Chart with Highlighting."
 
@@
 #| fig-cap: "Bar Plot."
+#| fig-alt: "Bar Plot."
 
@@
 #| fig-cap: "Bar Plot with Bars Representing the 95% Confidence Interval."
+#| fig-alt: "Bar Plot with Bars Representing the 95% Confidence Interval."
 
@@
 #| fig-cap: "Faceted Scatterplot."
+#| fig-alt: "Faceted Scatterplot."
 
@@
 #| fig-cap: "Rushing Yards Per Carry (Season) by Player Age."
+#| fig-alt: "Rushing Yards Per Carry (Season) by Player Age."
 
@@
 #| fig-cap: "NFL Rushing Yards Per Carry Per Season by Player Age."
+#| fig-alt: "NFL Rushing Yards Per Carry Per Season by Player Age."
 
@@
 #| fig-cap: "Rushing Expected Points Added (Season) by Player Age."
+#| fig-alt: "Rushing Expected Points Added (Season) by Player Age."
 
@@
 #| fig-cap: "NFL Rushing Expected Points Added (EPA) During a Season by Player Age."
+#| fig-alt: "NFL Rushing Expected Points Added (EPA) During a Season by Player Age."
 
@@
 #| fig-cap: "2024 NFL Offensive and Defensive EPA Per Play."
+#| fig-alt: "2024 NFL Offensive and Defensive EPA Per Play."
 

### 2024-10-16 evaluating-prediction-accuracy.qmd
20241016 - add alt-text to figures

@@
 #| fig-cap: "Distribution of Test Scores by Berry Type."
+#| fig-alt: "Distribution of Test Scores by Berry Type."
 #| label: fig-classificationDistributions
@@
 #| fig-cap: "Classifications Based on a Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
+#| fig-alt: "Classifications Based on a Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
 #| label: fig-classificationStandardCutoff
@@
 #| fig-cap: "Classifications Based on Raising the Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
+#| fig-alt: "Classifications Based on Raising the Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
 #| label: fig-classificationRaiseCutoff
@@
 #| fig-cap: "Classifications Based on Lowering the Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
+#| fig-alt: "Classifications Based on Lowering the Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
 #| label: fig-classificationLowerCutoff
@@
 #| fig-cap: "Receiver Operating Characteristic (ROC) Curves for Various Levels of Area Under The ROC Curve (AUC) for Various Measures."
+#| fig-alt: "Receiver Operating Characteristic (ROC) Curves for Various Levels of Area Under The ROC Curve (AUC) for Various Measures."
 #| label: fig-aucRange
@@
 #| fig-cap: "Predictions that are Well-Calibrated. That is, the predicted values are close to the observed values."
+#| fig-alt: "Predictions that are Well-Calibrated. That is, the predicted values are close to the observed values."
 #| fig-width: 8
@@
 #| fig-cap: "Types of Miscalibration. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
+#| fig-alt: "Types of Miscalibration. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
 #| fig-width: 8

### 2024-10-16 mixed-models.qmd
20241016 - render the figure caption as alt text
20241016 - add alt-text to figures

@@
 ::: {#fig-simpsonsParadox}
-![](images/simpsonsParadox.png)
+![](images/simpsonsParadox.png)\
 
@@
 #| fig-cap: "Scatterplot of Fantasy Points by Age for Quarterbacks."
+#| fig-alt: "Scatterplot of Fantasy Points by Age for Quarterbacks."
 #| code-fold: true
@@
 #| fig-cap: "Scatterplot of Fantasy Points by Age for Fullbacks."
+#| fig-alt: "Scatterplot of Fantasy Points by Age for Fullbacks."
 #| code-fold: true
@@
 #| fig-cap: "Scatterplot of Fantasy Points by Age for Running Backs."
+#| fig-alt: "Scatterplot of Fantasy Points by Age for Running Backs."
 #| code-fold: true
@@
 #| fig-cap: "Scatterplot of Fantasy Points by Age for Wide Receivers."
+#| fig-alt: "Scatterplot of Fantasy Points by Age for Wide Receivers."
 #| code-fold: true
@@
 #| fig-cap: "Scatterplot of Fantasy Points by Age for Tight Ends."
+#| fig-alt: "Scatterplot of Fantasy Points by Age for Tight Ends."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Quarterbacks."
+#| fig-alt: "Plot of Raw Trajectories of Fantasy Points by Age for Quarterbacks."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Fullbacks."
+#| fig-alt: "Plot of Raw Trajectories of Fantasy Points by Age for Fullbacks."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Running Backs."
+#| fig-alt: "Plot of Raw Trajectories of Fantasy Points by Age for Running Backs."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Wide Receivers."
+#| fig-alt: "Plot of Raw Trajectories of Fantasy Points by Age for Wide Receivers."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Raw Trajectories of Fantasy Points by Age for Tight Ends."
+#| fig-alt: "Plot of Raw Trajectories of Fantasy Points by Age for Tight Ends."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Null Model."
+#| fig-alt: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Null Model."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Linear Regression Model."
+#| fig-alt: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Linear Regression Model."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Quadratic Regression Model."
+#| fig-alt: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Quadratic Regression Model."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Random Intercepts Mixed Model."
+#| fig-alt: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Random Intercepts Mixed Model."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Mixed Model with Random Intercepts. Overlaid with the Model-Implied Trajectory."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Mixed Model with Random Intercepts. Overlaid with the Model-Implied Trajectory."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Random Intercepts Mixed Model With Position as a Fixed-Effect Predictor."
+#| fig-alt: "Plot of Model-Implied Trajectories of Fantasy Points by Age in Random Intercepts Mixed Model With Position as a Fixed-Effect Predictor."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Mixed Model With Random Intercepts and a Fixed-Effect of Position.
+Overlaid with the Model-Implied Trajectory by Position."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Mixed Model With Random Intercepts and a Fixed-Effect of Position.
 Overlaid with the Model-Implied Trajectory by Position."
@@
 #| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts and Fixed Linear Slopes."
+#| fig-alt: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts and Fixed Linear Slopes."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts and Fixed Slopes.
+Overlaid with the Model-Implied Trajectory by Position."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts and Fixed Slopes.
 Overlaid with the Model-Implied Trajectory by Position."
@@
 #| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts and Random Linear Slopes."
+#| fig-alt: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts and Random Linear Slopes."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts and Random Linear Slopes.
+Overlaid with the Model-Implied Trajectory by Position."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts and Random Linear Slopes.
 Overlaid with the Model-Implied Trajectory by Position."
@@
 #| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes."
+#| fig-alt: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes.
+Overlaid with the Model-Implied Trajectory by Position."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes.
 Overlaid with the Model-Implied Trajectory by Position."
@@
 #| fig-cap: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position."
+#| fig-alt: "Plot of Model-Implied Trajectories of Fantasy Points by Age and Position in Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position.
+Overlaid with the Model-Implied Trajectory by Position."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age and Position, from a Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position.
 Overlaid with the Model-Implied Trajectory by Position."
@@
 #| fig-cap: "Histogram of Fantasy Points with Overlaid Density Plot and Rug Plot."
+#| fig-alt: "Histogram of Fantasy Points with Overlaid Density Plot and Rug Plot."
 
@@
 #| fig-cap: "Cullen and Frey Graph."
+#| fig-alt: "Cullen and Frey Graph."
 
@@
 #| fig-cap: "Fit of normal distribution to fantasy points."
+#| fig-alt: "Fit of normal distribution to fantasy points."
 
@@
 #| fig-cap: "Fit of log normal distribution to fantasy points."
+#| fig-alt: "Fit of log normal distribution to fantasy points."
 
@@
 #| fig-cap: "Fit of gamma distribution to fantasy points."
+#| fig-alt: "Fit of gamma distribution to fantasy points."
 
@@
 #| fig-cap: "Fit of exponential distribution to fantasy points."
+#| fig-alt: "Fit of exponential distribution to fantasy points."
 
@@
 #| fig-cap: "Fit of Weibull distribution to fantasy points."
+#| fig-alt: "Fit of Weibull distribution to fantasy points."
 
@@
 #| fig-cap: "Trace Plots from Bayesian Mixed Model."
+#| fig-alt: "Trace Plots from Bayesian Mixed Model."
 
@@
 #| fig-cap: "Posterior Predictive Check from Bayesian Mixed Model."
+#| fig-alt: "Posterior Predictive Check from Bayesian Mixed Model."
 
@@
 #| fig-cap: "Plot of Model-Implied Quadratic Trajectories of Fantasy Points by Age."
+#| fig-alt: "Plot of Model-Implied Quadratic Trajectories of Fantasy Points by Age."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Model-Implied Quadratic Trajectories of Fantasy Points by Age For Players Who Were Once at the Top of the End-of-Season Depth Chart."
+#| fig-alt: "Plot of Model-Implied Quadratic Trajectories of Fantasy Points by Age For Players Who Were Once at the Top of the End-of-Season Depth Chart."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age from a Generalized Additive Model."
+#| fig-alt: "Plot of Implied Trajectories of Fantasy Points by Age from a Generalized Additive Model."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, For Players Who Were Once at the Top of the End-of-Season Depth Chart."
+#| fig-alt: "Plot of Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, For Players Who Were Once at the Top of the End-of-Season Depth Chart."
 #| code-fold: true
@@
 Overlaid with the Model-Implied Trajectory for Quarterbacks in Blue."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Quarterbacks.
+Overlaid with the Model-Implied Trajectory for Quarterbacks in Blue."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Fullbacks. Overlaid with the Model-Implied Trajectory for Fullbacks in Blue."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Fullbacks. Overlaid with the Model-Implied Trajectory for Fullbacks in Blue."
 #| code-fold: true
@@
 Overlaid with the Model-Implied Trajectory for Running Backs in Blue."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Running Backs.
+Overlaid with the Model-Implied Trajectory for Running Backs in Blue."
 #| code-fold: true
@@
 Overlaid with the Model-Implied Trajectory for Wide Receivers in Blue."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Wide Receivers.
+Overlaid with the Model-Implied Trajectory for Wide Receivers in Blue."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Tight Ends.
+Overlaid with the Model-Implied Trajectory for Wide Tight Ends in Blue."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Generalized Additive Model, for Tight Ends.
 Overlaid with the Model-Implied Trajectory for Wide Tight Ends in Blue."

### 2024-10-16 modern-portfolio-theory.qmd
20240911 - additions
20241016 - add alt-text to figures

@@
 That is likely why [a cat beat professional investors in a stock market challenge](http://www.npr.org/blogs/money/2013/01/14/169326326/housecat-beats-investors-in-stock-market-challenge) (archived at <https://perma.cc/R3XU-K6J8>).
+That is, "few stock pickers, if any, have the skill needed to beat the market consistently, year after year" [@Kahneman2011, p. 214].
 Although our sample size is much smaller with fantasy football projections, there also appears to be little consistency in fantasy football sites' rank in accuracy over time [INSERT], suggesting that the projection sources are not reliably better than each other (or the crowd) over time.
@@
 #| fig-cap: "Efficient Frontier for a Stock Portfolio."
+#| fig-alt: "Efficient Frontier for a Stock Portfolio."
 

### 2024-10-16 player-evaluation.qmd
20240914 - Vegas lines
20240917 - updates
20240917 - loss aversion & prospect theory
20241016 - add alt-text to figures

@@
 Moreover, sites also provide season-long player futures for the over/under betting lines, which are based on Vegas lines and how people bet: <https://www.rotowire.com/betting/nfl/player-futures.php>.
+In addition, sites also provide weekly projections based on Vegas lines: <https://vegasprojections.com>; <https://www.actionnetwork.com/nfl/props>.
 
@@
 #| fig-cap: "Density Plot of Projected Points for Three Players"
+#| fig-alt: "Density Plot of Projected Points for Three Players"
 #| code-fold: true

### 2024-10-16 time-series-analysis.qmd
20241016 - add alt-text to figures

@@
 #| fig-cap: "Tom Brady's Historical Fantasy Points by Game."
+#| fig-alt: "Tom Brady's Historical Fantasy Points by Game."
 
@@
 #| fig-cap: "Historical Fantasy Points by Game for Tom Brady and Peyone Manning."
+#| fig-alt: "Historical Fantasy Points by Game for Tom Brady and Peyone Manning."
 
@@
 #| fig-cap: "Autocorrelation Function (ACF) Plot of Tom Brady's Historical Fantasy Points by Game."
+#| fig-alt: "Autocorrelation Function (ACF) Plot of Tom Brady's Historical Fantasy Points by Game."
 
@@
 #| fig-cap: "Model Summary of Autoregressive Integrated Moving Average Model fit to Tom Brady's Historical Performance by Game."
+#| fig-alt: "Model Summary of Autoregressive Integrated Moving Average Model fit to Tom Brady's Historical Performance by Game."
 
@@
 #| fig-cap: "Model Summary of modified Autoregressive Integrated Moving Average Model fit to Tom Brady's Historical Performance by Game."
+#| fig-alt: "Model Summary of modified Autoregressive Integrated Moving Average Model fit to Tom Brady's Historical Performance by Game."
 
@@
 #| fig-cap: "Model Summary of Autoregressive Integrated Moving Average Model fit to Peyton Manning's Historical Performance by Game."
+#| fig-alt: "Model Summary of Autoregressive Integrated Moving Average Model fit to Peyton Manning's Historical Performance by Game."
 
@@
 #| fig-cap: "Tom Brady's Historical and Projected Fantasy Points by Game."
+#| fig-alt: "Tom Brady's Historical and Projected Fantasy Points by Game."
 
@@
 #| fig-cap: "Peyton Manning's Historical and Projected Fantasy Points by Game."
+#| fig-alt: "Peyton Manning's Historical and Projected Fantasy Points by Game."
 
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Quarterbacks."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Quarterbacks."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Running Backs."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Running Backs."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Wide Receivers."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Wide Receivers."
 #| code-fold: true
@@
 #| fig-cap: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Tight Ends."
+#| fig-alt: "Plot of Individuals' Implied Trajectories of Fantasy Points by Age, from a Bayesian Generalized Additive Model, for Tight Ends."
 #| code-fold: true

### 2024-10-16 basic-statistics.qmd
20241016 - reorder figure labels

@@
 ```{r}
+#| label: fig-pValuesDifference
 #| layout-ncol: 2
@@
 #|   - "...if in the population, the groups were really this?"
-#| label: fig-pValuesDifference
 #| code-fold: true
@@
 ```{r}
+#| label: fig-pValuesAssociation
 #| layout-ncol: 2
@@
 #|   - "...if in the population, the association was really this?"
-#| label: fig-pValuesAssociation
 #| code-fold: true

### 2024-10-16 cognitive-bias.qmd
20241016 - reorder figure labels

@@
 ```{r}
+#| label: fig-simulationOf10CoinFlips
 #| fig-cap: "Histogram of Number of Coins that Flip Heads-Up in a Simulation of 10 Coin Flips (with 100,000 Replications)."
 #| fig-alt: "Histogram of Number of Coins that Flip Heads-Up in a Simulation of 10 Coin Flips (with 100,000 Replications)."
-#| label: fig-simulationOf10CoinFlips
 #| code-fold: true

### 2024-10-16 evaluating-prediction-accuracy.qmd
20241016 - reorder figure labels

@@
 ```{r}
+#| label: fig-classificationDistributions
 #| fig-cap: "Distribution of Test Scores by Berry Type."
 #| fig-alt: "Distribution of Test Scores by Berry Type."
-#| label: fig-classificationDistributions
 #| code-fold: true
@@
 ```{r}
+#| label: fig-classificationStandardCutoff
 #| fig-cap: "Classifications Based on a Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
 #| fig-alt: "Classifications Based on a Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
-#| label: fig-classificationStandardCutoff
 #| code-fold: true
@@
 ```{r}
+#| label: fig-classificationRaiseCutoff
 #| fig-cap: "Classifications Based on Raising the Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
 #| fig-alt: "Classifications Based on Raising the Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
-#| label: fig-classificationRaiseCutoff
 #| code-fold: true
@@
 ```{r}
+#| label: fig-classificationLowerCutoff
 #| fig-cap: "Classifications Based on Lowering the Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
 #| fig-alt: "Classifications Based on Lowering the Cutoff. Note that some true negatives and true positives are hidden behind the false positives and false negatives."
-#| label: fig-classificationLowerCutoff
 #| code-fold: true
@@
 ```{r}
+#| label: fig-aucRange
 #| fig-cap: "Receiver Operating Characteristic (ROC) Curves for Various Levels of Area Under The ROC Curve (AUC) for Various Measures."
 #| fig-alt: "Receiver Operating Characteristic (ROC) Curves for Various Levels of Area Under The ROC Curve (AUC) for Various Measures."
-#| label: fig-aucRange
 #| fig-width: 8

### 2024-10-17 time-series-analysis.qmd
20241016 - duplicate identifiers

@@
 
-### Table of Next Season Predictions {#sec-timeSeriesBayesianGeneratePredictions}
+### Table of Next Season Predictions {#sec-timeSeriesBayesianNextSeasonPredictions}
 
@@
 
-### Plot of Individuals' Model-Implied Predictions {#sec-timeSeriesBayesianGeneratePredictions}
+### Plot of Individuals' Model-Implied Predictions {#sec-timeSeriesBayesianModelImpliedPredictions}
 
-#### Quarterbacks {#sec-fantasyPointsIndividualByAgePlotsQB}
+#### Quarterbacks {#sec-fantasyPointsIndividualByAgeTimeSeriesPlotsQB}
 
@@
 
-#### Running Backs {#sec-fantasyPointsIndividualByAgePlotsRB}
+#### Running Backs {#sec-fantasyPointsIndividualByAgeTimeSeriesPlotsRB}
 
@@
 
-#### Wide Receivers {#sec-fantasyPointsIndividualByAgePlotsWR}
+#### Wide Receivers {#sec-fantasyPointsIndividualByAgeTimeSeriesPlotsWR}
 
@@
 
-#### Tight Ends {#sec-fantasyPointsIndividualByAgePlotsTE}
+#### Tight Ends {#sec-fantasyPointsIndividualByAgeTimeSeriesPlotsTE}
 

### 2024-10-17 base-rates.qmd
20241016 - figure alt text

@@
 ::: {#fig-probabilityNomogram}
-![](images/probabilityNomogram.png){width=50%}\
+![](images/probabilityNomogram.png){width=50% fig-alt="Probability Nomogram.
+(Figure retrieved from https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png)"}
 
@@
 ::: {#fig-probabilityNomogramLine}
-![](images/probabilityNomogramLine.png){width=50%}\
+![](images/probabilityNomogramLine.png){width=50% fig-alt="Probability Nomogram.
+(Figure retrieved from https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png).
+Also provided in: @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 

### 2024-10-17 basic-statistics.qmd
20241016 - figure alt text

@@
 ::: {#fig-nhstConfusionMatrix}
-![](images/nhstConfusionMatrix.png)\
+![](images/nhstConfusionMatrix.png){fig-alt="A Two-by-Two Confusion Matrix for Null-Hypothesis Significance Testing."}
 
@@
 ::: {#fig-statisticalDecisionTree}
-![](images/statisticalDecisionTree.png)\
+![](images/statisticalDecisionTree.png){fig-alt="A Statistical Decision Tree For Choosing an Appropriate Statistical Procedure.
+Adapted from: <https://commons.wikimedia.org/wiki/File:InferentialStatisticalDecisionMakingTrees.pdf>.
+The original source is: Corston, R.
+& Colman, A.
+M.
+(2000).
+*A crash course in SPSS for Windows*.
+Wiley-Blackwell.
+Changes were made to the original, including the addition of several statistical tests."}
 
@@
 ::: {#fig-statisticalDecisionTreeSummary}
-![](images/statisticalDecisionTreeSummary.png)\
+![](images/statisticalDecisionTreeSummary.png){fig-alt="Summary of A Statistical Decision Tree For Choosing an Appropriate Statistical Procedure."}
 
@@
 ::: {#fig-statisticalDecisionTreeRegression}
-![](images/statisticalDecisionTreeRegression.png)\
+![](images/statisticalDecisionTreeRegression.png){fig-alt="A Statistical Decision Tree For Choosing an Appropriate Statistical Procedure, Re-Formulated in a Regression Framework.
+Adapted from: <https://commons.wikimedia.org/wiki/File:InferentialStatisticalDecisionMakingTrees.pdf>.
+The original source is: Corston, R.
+& Colman, A.
+M.
+(2000).
+*A crash course in SPSS for Windows*.
+Wiley-Blackwell.
+Changes were made to the original, including re-formulating the tests in a regression framework."}
 
@@
 ::: {#fig-statisticalDecisionTreeSummaryRegression}
-![](images/statisticalDecisionTreeSummaryRegression.png)\
+![](images/statisticalDecisionTreeSummaryRegression.png){fig-alt="Summary of A Statistical Decision Tree For Choosing an Appropriate Statistical Procedure."}
 

### 2024-10-17 causal-inference.qmd
20241016 - figure alt text

@@
 ::: {#fig-crossSectional}
-![](images/Longitudinal-01.png){width=20%}\
+![](images/Longitudinal-01.png){width=20% fig-alt="Cross-Sectional Association. T1 = Timepoint 1. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
@@
 ::: {#fig-laggedAssociation}
-![](images/Longitudinal-02.png)\
+![](images/Longitudinal-02.png){fig-alt="Lagged Association. T1 = Timepoint 1. T2 = Timepoint 2. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
@@
 ::: {#fig-laggedAssociationControlPriorLevels}
-![](images/Longitudinal-03.png)\
+![](images/Longitudinal-03.png){fig-alt="Lagged Association, Controlling for Prior Levels of the Outcome.
+T1 = Timepoint 1.
+T2 = Timepoint 2.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
@@
 ::: {#fig-crossLaggedPanelModel}
-![](images/Longitudinal-04.png)\
+![](images/Longitudinal-04.png){fig-alt="Lagged Association, Controlling for Prior Levels of the Outcome, Simultaneously Testing Both Directions Of Effect.
+T1 = Timepoint 1.
+T2 = Timepoint 2.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
@@
 ::: {#fig-simpsonParadox}
-![](images/simpsonParadox.png){width=200%}\
+![](images/simpsonParadox.png){width=200% fig-alt="Example Where Failing to Control for a Variable (In This Case, Gender) Would Lead to False Inferences.
+In this example, the association between dosage and recovery probability is positive at the population level, but the association is negative among men and women separately.
+(Figure reprinted from @Kievit2013, Figure 1, p.
+2.
+Kievit, R., Frankenhuis, W., Waldorp, L., & Borsboom, D.
+(2013).
+Simpson's paradox in psychological science: A practical guide.
+*Frontiers in Psychology*, *4*(513).
+https://doi.org/10.3389/fpsyg.2013.00513)"}
 

### 2024-10-17 cognitive-bias.qmd
20241016 - figure alt text

@@
 ::: {#fig-prospectTheory}
-![](images/prospectTheory.png)\
+![](images/prospectTheory.png){fig-alt="Prospect Theory and Loss Aversion: People Tend to Weigh Losses More Heavily Than Gains.
+That is, the utility function is steeper in the loss (red) region than in the gain (blue) region.
+Adapted from <https://commons.wikimedia.org/wiki/File:Graphique_th%C3%A9orie_des_perspectives.svg>."}
 
@@
 ::: {#fig-dunningKrueger1}
-![](images/dunningKrueger_1.png)\
+![](images/dunningKrueger_1.png){fig-alt="Dunning–Krueger Effect: Confidence as a Function of Competence.
+People with low competency in a task tend to overestimate their ability.
+Adapted from <https://commons.wikimedia.org/wiki/File:Effet_Dunning-Kruger.svg>."}
 
@@
 ::: {#fig-dunningKrueger2}
-![](images/dunningKrueger_2.png)\
+![](images/dunningKrueger_2.png){fig-alt="Dunning–Krueger Effect: Perceived Performance as a Function of Actual Performance.
+People who perform poorly in a task tend to overestimate their performance.
+Adapted from <https://commons.wikimedia.org/wiki/File:Dunning-kruger_effect_-_percentile.svg>."}
 

### 2024-10-17 fantasy-football.qmd
20241016 - figure alt text

@@
 ::: {#fig-footballFormation}
-![](images/footballFormation.png)\
+![](images/footballFormation.png){fig-alt="An Example Football Formation for the Offense and Defense.
+The solid line indicates the line of scrimmage.
+The arrow indicates the direction the offense tries to advance the ball."}
 
@@
 ::: {#fig-footballField}
-![](images/footballField.png){width=50%}\
+![](images/footballField.png){width=50% fig-alt="A Diagram of a Football Field.
+The yard markers depict the distance from the nearest end zone.
+The orange shaded area is called the "red zone", where chances of scoring points are highest.
+The original figure was modified to depict field goal posts.
+(Figure retrieved from <https://commons.wikimedia.org/wiki/File:American_football_field.svg>"}
 

### 2024-10-17 multiple-regression.qmd
20241016 - add alt-text to figures
20241016 - figure alt text

--- 2024-07-29 multiple-regression.qmd
@@
 ::: {#fig-regression}
-![](images/regression.png)
+![](images/regression.png){fig-alt="A Regression Best-Fit Line."}
 
@@
 ::: {#fig-regression}
-![](images/multipleRegressionRSquared.png){width=50%}
+![](images/multipleRegressionRSquared.png){width=50% fig-alt="Conceptual Depiction of Proportion of Variance Explained ($R^2$) in an Outcome Variable ($Y$) by Multiple Predictors ($X1$ and $X2$) in Multiple Regression.
+The size of each circle represents the variable's variance.
+The proportion of variance in $Y$ that is explained by the predictors is depicted by the areas in orange.
+The dark orange space ($G$) is where multiple predictors explain overlapping variance in the outcome.
+Overlapping variance that is explained in the outcome ($G$) will not be recovered in the regression coefficients when both predictors are included in the regression model.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
@@
 #| fig-cap: "Over-fitting Model in Red Relative to the True Distribution of the Data in Blue. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
+#| fig-alt: "Over-fitting Model in Red Relative to the True Distribution of the Data in Blue. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."
 #| code-fold: true
@@
 ::: {#fig-regression}
-![](images/multipleRegressionMulticollinearity.png){width=50%}
+![](images/multipleRegressionMulticollinearity.png){width=50% fig-alt="Conceptual Depiction of Multicollinearity in Multiple Regression. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 

### 2024-10-17 research-methods.qmd
20241016 - render the figure caption as alt text
20241016 - figure alt text

@@
 ::: {#fig-correlationAndCausation1}
-![](images/correlationAndCausation_1.png)
+![](images/correlationAndCausation_1.png){fig-alt="Hypothesized Causal Effect Based on an Observed Association Between `X` and `Y`, Such That `X` Causes `Y`."}
 
@@
 ::: {#fig-correlationAndCausation2}
-![](images/correlationAndCausation_2.png)
+![](images/correlationAndCausation_2.png){fig-alt="Reverse (Opposite) Direction of Effect From the Hypothesized Effect, Where `Y` Causes `X`."}
 
@@
 ::: {#fig-correlationAndCausation3}
-![](images/correlationAndCausation_3.png)
+![](images/correlationAndCausation_3.png){fig-alt="Bidirectional Effect Between `X` and `Y`, such that `X` causes `Y` and `Y` causes `X`."}
 
@@
 ::: {#fig-correlationAndCausation4}
-![](images/correlationAndCausation_4.png)
+![](images/correlationAndCausation_4.png){fig-alt="Confounded Association Between `X` and `Y` due to a Common Cause, `Z`."}
 
@@
 ::: {#fig-mediationModerationMnemonic}
-![](images/mediationModerationMnemonic.png)
+![](images/mediationModerationMnemonic.png){fig-alt="Mediation Versus Moderation Mnemonic."}
 
@@
 ::: {#fig-mediation}
-![](images/mediation.png)
+![](images/mediation.png){fig-alt="Mediation."}
 
@@
 ::: {#fig-mediationExample}
-![](images/mediationExample.png)
+![](images/mediationExample.png){fig-alt="Mediation Example."}
 
@@
 ::: {#fig-fullMediation}
-![](images/mediation.png)
+![](images/mediation.png){fig-alt="Full Mediation."}
 
@@
 ::: {#fig-partialMediation}
-![](images/mediationPartial.png)
+![](images/mediationPartial.png){fig-alt="Partial Mediation."}
 
@@
 ::: {#fig-multipleMediators}
-![](images/mediationMultiple.png)
+![](images/mediationMultiple.png){fig-alt="Multiple Mediators."}
 
@@
 ::: {#fig-moderation}
-![](images/moderation.png)
+![](images/moderation.png){fig-alt="Moderation."}
 
@@
 ::: {#fig-moderationExamplePathDiagram}
-![](images/moderationExample1.png)
+![](images/moderationExample1.png){fig-alt="Moderation Example: Path Diagram."}
 
@@
 ::: {#fig-moderationExampleInteractionGraph}
-![](images/moderationExample2.png)
+![](images/moderationExample2.png){fig-alt="Moderation Example: Interaction Graph."}
 
@@
 ::: {#fig-reliabilityVsValidity}
-![](images/reliabilityVsValidity.png)
+![](images/reliabilityVsValidity.png){fig-alt="Reliability Versus Validity."}
 

### 2024-10-19 research-methods.qmd
20241019 - hypothesis and prediction

@@
 
+## Research Questions, Hypotheses, and Predictions {#sec-hypothesisVsPrediction}
+
+A research question is a question that the investigator (you!) want to know the answer to.
+For example, a research question might be: "Does consumption of sports drink improve player performance?"
+A hypothesis is a proposed explanation.
+A prediction is "the expected result of a test that is derived, by deduction, from a hypothesis or theory" [@Eastwell2014, p. 17; archived at <https://perma.cc/8EX4-8JYN>].
+Here is an example of a hypothesis and the resulting prediction:
+
+> The present study evaluates whether consumption of sports drink improves player performance.
+> I <u>*hypothesize*</u> that consumption of sports drink leads football players to perform better in games because of greater endurance owing to restoration of electrolytes.
+> If the hypothesis is true, I <u>*predict*</u> that players who consume sports drink during a game will score more fantasy points than players who do not consume sports drink during the game.
+
 ## Research Designs {#sec-researchDesigns}
@@
 For instance, we may be interested to evaluate whether players perform better (e.g., run faster) if they drink a sports drink compared when they drink only water.
-Our hypothesis might be that players will be expected to perform better when they drink a sports drink (compared to when they drink only water).
+Our hypothesis might be that players will be expected to perform better when they drink a sports drink (compared to when they drink only water), for the reasons specified in @sec-hypothesisVsPrediction.
 To this this research question and hypothesis, we might conduct an experiment by randomly assigning some players during practice to receive a sports drink and some players to receive only water.

### 2024-10-23 download_weekly_projections.R
20241023 - week 8 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2024
-weekNumber <- 7
+weekNumber <- 8
 
 # Processing ----
 season <- as.character(season)

### 2024-10-25 machine-learning.qmd
Built site for gh-pages

@@
 
+## Types of Machine Learning {#sec-machineLearningTypes}
+
+There are many approaches to machine learning.
+This chapter discusses several key ones:
+
+- supervised learning
+    - continuous outcome (i.e., regression)
+      - linear regression
+      - lasso regression
+      - ridge regression
+      - elastic net regression
+    - categorical outcome (i.e., classification)
+      - logistic regression
+      - support vector machine
+      - random forest
+      - extreme gradient boosting
+- unsupervised learning
+    - clustering
+    - principal component analysis
+- semi-supervised learning
+- reinforcement learning
+    - deep learning
+
+### Supervised Learning {#sec-machineLearningTypesSupervised}
+
+[DEFINE SUPERVISED LEARNING]
+
+Unlike linear and logistic regression, various machine learning techniques can handle multicollinearity, including LASSO regression, ridge regression, and elastic net regression.
+Least absolute shrinkage and selection operator (LASSO) regression helps perform selection of which predictor variables to keep in the model by shrinking some coefficients to zero.
+Ridge regression shrinks the coefficients of predictor variables toward zero, but not to zero, so it does not perform selection of which predictor variables to retain; this allows it to allow nonzero coefficients for multiple correlated predictor variables in the context of multicollinearity.
+Elastic net involves a combination of LASSO and ridge regression; it performs selection of which predictor variables to keep by shrinking the coefficients of some predictor variables to zero, and it shrinks the coefficients of some predictor variables toward zero, to address multicollinearity.
+
+Unless interactions or nonlinear terms are specified, linear, logistic, LASSO, ridge, and elastic net regresstion do not account for interactions among the predictor variables or for nonlinear associations between the predictor variables and the outcome variable.
+By contrast, random forests and extreme gradient boosting do account for interactions among the predictor variables and for nonlinear associations between the predictor variables and the outcome variable.
+
+### Unsupervised Learning {#sec-machineLearningTypesUnsupervised}
+
+[DEFINE UNSUPERVISED LEARNING]
+
+We describe [cluster analysis](#sec-clusterAnalysis) in @sec-clusterAnalysis.
+We describe [principal component analysis](#sec-pca) in @sec-pca.
+
+### Semi-supervised Learning {#sec-machineLearningTypesSemisupervised}
+
+[DEFINE SEMI-SUPERVISED LEARNING]
+
+### Reinforcement Learning {#sec-machineLearningTypesReinforcement}
+
+[DEFINE REINFORCEMENT LEARNING]
+
 ## Data Processing {#sec-machineLearningDataProcessing}

### 2024-11-06 download_weekly_projections.R
20241106 - week 10 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2024
-weekNumber <- 9
+weekNumber <- 10
 
 # Processing ----
 season <- as.character(season)

### 2024-11-27 bugs.qmd
20241126 - calculate_stats() function from nflfastR, TODOs, and bugs

--- 2024-05-25 bugs.qmd
@@
 
-Bugs to report (`nflreadr`):
+Bugs to report (`nflreadr` or `nflfastR`):
 
-- `players` database has a season column but it is only populated for some players, and it is only populated with 2020
-- `player_stats` database of weekly stats does not have a `game_id` column
+- `players` database has a season column but it is only populated for some players, and it is only populated with 2020 (<https://github.com/nflverse/nflreadr/issues/238>)
+- `player_stats` database of weekly stats does not have a `game_id` column (<https://github.com/nflverse/nflreadr/issues/238>)
+- common naming convention for key variables, for purposes of merging (<https://github.com/nflverse/nflreadr/issues/237>)
+- use of deprecated `summarise()` for returning more than 1 row per group; use `reframe()` instead (<https://github.com/nflverse/nflfastR/issues/480>)
+- calculate career stats using `calculate_stats()` function

### 2024-11-27 download-football-data.qmd
20241126 - calculate_stats() function from nflfastR, TODOs, and bugs

@@
 
+First, we can compute the players' career statistics using the `calculate_stats()` function from the `nflfastR` package.
+
+TODO: Insert code example.
+
+##### OLD (DELETE CODE BELOW)
+
 First, we can compute the players' career statistics using the `calculate_player_stats()`, `calculate_player_stats_def()`, and `calculate_player_stats_kicking()` functions from the `nflfastR` package for offensive players, defensive players, and kickers, respectively.
@@
 
+TODO: Save/update data file in repo with the data generated from this code, and follow the data file through the rest of the book, updating code as necessary.
+
+```{r}
+#| eval: false
+nfl_actualStats_seasonal_player_raw <- calculate_stats(
+  seasons = TRUE,
+  summary_level = "season",
+  season_type = "REG")
+
+nfl_actualStats_seasonal_team_raw <- calculate_stats(
+  seasons = TRUE,
+  summary_level = "season",
+  stat_type = "team",
+  season_type = "REG")
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_seasonal_player_raw, nfl_actualStats_seasonal_team_raw,
+  file = "./data/nfl_actualStats_seasonal_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load("./data/nfl_actualStats_seasonal_raw.RData")
+```
+
+The `nfl_actualStats_seasonal_player_raw` object is in `player`-`season` form.
+That is, each row should be uniquely identified by the combination of `player_id` and `season`.
+The `nfl_actualStats_seasonal_team_raw` object is in `team`-`season` form.
+That is, each row should be uniquely identified by the combination of `team` and `season`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_seasonal_player <- nfl_actualStats_seasonal_player_raw %>% 
+  rename(team = recent_team) %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_seasonal_team <- nfl_actualStats_seasonal_team_raw %>% 
+  select(season, team, everything()) %>% 
+  arrange(season, team)
+load(file = "./data/nfl_actualStats_seasonal.RData")
+```
+
+##### OLD (DELETE CODE BELOW)
+
 ```{r}
@@
 Nevertheless, we could compute players' weekly statistics from the play-by-play data using the following syntax:
+
+TODO:
+
+- Decide whether to use `calculate_stats()` from `nflfastR` or `load_player_stats()` from `nflreadr`.
+- Save/update data file in repo with the data generated from the relevant code, and follow the data file through the rest of the book, updating code as necessary.
+
+```{r}
+#| eval: false
+nfl_actualStats_weekly_raw <- calculate_stats(
+  seasons = TRUE,
+  summary_level = "week")
+
+nfl_actualStats_weekly_defense_raw <- calculate_stats(
+  seasons = TRUE,
+  summary_level = "week",
+  stat_type = "team")
+```
+
+```{r}
+#| include: false
+
+# garbage collection
+rm(nfl_pbp)
+gc()
+```
+
+##### OLD (DELETE CODE BELOW)
 

### 2024-11-27 download-football-data.qmd
20241126 - calculate_stats() function from nflfastR

@@
   arrange(season, team)
+```
+
+Let's check for duplicate `player` instances:
+
+```{r}
+nfl_actualStats_seasonal_player %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_seasonal_team %>% 
+  group_by(team, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_actualStats_seasonal_player, nfl_actualStats_seasonal_team,
+  file = "./data/nfl_actualStats_seasonal.RData"
+)
+```
+
+```{r}
+#| include: false
+
 load(file = "./data/nfl_actualStats_seasonal.RData")

### 2024-11-27 bugs.qmd
20241126 - bugs

--- 2024-05-25 bugs.qmd
@@
 
-Bugs to report (`nflreadr`):
+Bugs to report (`nflreadr` or `nflfastR`):
 
-- `players` database has a season column but it is only populated for some players, and it is only populated with 2020
-- `player_stats` database of weekly stats does not have a `game_id` column
+- `players` database has a season column but it is only populated for some players, and it is only populated with 2020 (<https://github.com/nflverse/nflreadr/issues/238>)
+- `player_stats` database of weekly stats does not have a `game_id` column (<https://github.com/nflverse/nflreadr/issues/238>)
+- common naming convention for key variables, for purposes of merging (<https://github.com/nflverse/nflreadr/issues/237>)
+- use of deprecated `summarise()` for returning more than 1 row per group; use `reframe()` instead (<https://github.com/nflverse/nflfastR/issues/480>)
+- calculate career stats using `calculate_stats()` function (as was formerly able to do with the `calculate_player_stats()` function)
+- will `nflreadr` switch to using the objects from the `calculate_stats()` function in `nflfastR`?

### 2024-11-27 download-football-data.qmd
20241126 - Data Dictionary
20241127 - include postseason in download

--- 2024-11-27 download-football-data.qmd
@@
   season_type = "REG")
+
+nfl_actualStats_seasonal_player_inclPost_raw <- calculate_stats(
+  seasons = TRUE,
+  summary_level = "season",
+  season_type = "REG+POST")
+
+nfl_actualStats_seasonal_team_inclPost_raw <- calculate_stats(
+  seasons = TRUE,
+  summary_level = "season",
+  stat_type = "team",
+  season_type = "REG+POST")
 ```
@@
   nfl_actualStats_seasonal_player_raw, nfl_actualStats_seasonal_team_raw,
+  nfl_actualStats_seasonal_player_inclPost_raw, nfl_actualStats_seasonal_team_inclPost_raw,
   file = "./data/nfl_actualStats_seasonal_raw.RData"
@@
 load("./data/nfl_actualStats_seasonal_raw.RData")
+```
+
+A Data Dictionary for the variables is available in the `nfl_stats_variables` object that is returned when running the `calculate_stats()` function:
+
+```{r}
+nfl_stats_variables
 ```
@@
   arrange(season, team)
+
+nfl_actualStats_seasonal_player_inclPost <- nfl_actualStats_seasonal_player_inclPost_raw %>% 
+  rename(team = recent_team) %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_seasonal_team_inclPost <- nfl_actualStats_seasonal_team_inclPost_raw %>% 
+  select(season, team, everything()) %>% 
+  arrange(season, team)
+```
+
+Let's check for duplicate `player` instances:
+
+```{r}
+nfl_actualStats_seasonal_player %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_seasonal_team %>% 
+  group_by(team, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_seasonal_player_inclPost %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_seasonal_team_inclPost %>% 
+  group_by(team, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_actualStats_seasonal_player, nfl_actualStats_seasonal_team,
+  nfl_actualStats_seasonal_player_inclPost, nfl_actualStats_seasonal_team_inclPost,
+  file = "./data/nfl_actualStats_seasonal.RData"
+)
+```
+
+```{r}
+#| include: false
+
 load(file = "./data/nfl_actualStats_seasonal.RData")

### 2024-12-18 download_weekly_projections.R
20241218 - week 16 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2024
-weekNumber <- 15
+weekNumber <- 16
 
 # Processing ----
 season <- as.character(season)

### 2024-12-19 data-visualization.qmd
20241218 - visualizing univariate distributions
20241219 - color schemes for color-blind viewers

@@
 For examples of highlighting in figures, see Figures [-@fig-lineChartHighlighting] (@sec-lineChartHighlighting) and [-@fig-simulationOf10CoinFlips].
+
 It is also important to use color schemes with distinguishable colors.
 Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package, as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
+There are a variety of resources for color schemes that are accessible to color-blind viewers:
+
+- <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> (archived at <https://perma.cc/7VTA-Y8YS>)
+- <https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html> (archived at <https://perma.cc/NK9K-HL7L>)
+- @Nunez2018
 
 ## Univariate Distribution {#sec-univariateDistribution}
+
+For ways of visualizing univariate distributions, see [here](https://z3tt.github.io/beyond-bar-and-box-plots) (archived at <https://perma.cc/EEJ8-LND2>).
 

### 2024-12-25 basic-statistics.qmd
20241225 - convert to calculate_stats() from nflfastR v5.0

@@
 
+### Load Data {#sec-basicStatsLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/player_stats_seasonal.RData")
+```
+
+We created the `player_stats_seasonal.RData` object in @sec-calculatePlayerAge.
+
 ## Descriptive Statistics {#sec-descriptiveStatistics}
@@
 ```{r}
-#mean(data, na.rm = TRUE)
-#median(data, na.rm = TRUE)
-#DescTools::HodgesLehmann(exampleValues, na.rm = TRUE)
-#petersenlab::Mode(exampleValues)
-#weighted.mean(data, weights, na.rm = TRUE)
-#petersenlab::wthdquantile(data, weights, probs = 0.5)
+mean(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+median(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+DescTools::HodgesLehmann(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+petersenlab::Mode(player_stats_seasonal$fantasyPoints)
+
+weighted.mean(
+  player_stats_seasonal$fantasyPoints,
+  weights = sample( # randomly generate weights (could specify them manually)
+    x = 1:3,
+    size = length(player_stats_seasonal$fantasyPoints),
+    replace = TRUE),
+  na.rm = TRUE)
+
+petersenlab::wthdquantile(
+  player_stats_seasonal$fantasyPoints,
+  w = sample( # randomly generate weights (could specify them manually)
+    x = 1:3,
+    size = length(player_stats_seasonal$fantasyPoints),
+    replace = TRUE),
+  probs = 0.5)
 ```
@@
 ```{r}
-#psych::describe(mydata)
-
-#mydata %>% 
-#  summarise(across(
-#      everything(),
-#      .fns = list(
-#        n = ~ length(na.omit(.)),
-#        missingness = ~ mean(is.na(.)) * 100,
-#        M = ~ mean(., na.rm = TRUE),
-#        SD = ~ sd(., na.rm = TRUE),
-#        min = ~ min(., na.rm = TRUE),
-#        max = ~ max(., na.rm = TRUE),
-#        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
-#        IQR = ~ IQR(., na.rm = TRUE),
-#        MAD = ~ mad(., na.rm = TRUE),
-#        median = ~ median(., na.rm = TRUE),
-#        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
-#        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
-#        skewness = ~ psych::skew(., na.rm = TRUE),
-#        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
-#      .names = "{.col}.{.fn}")) %>%
-#    pivot_longer(
-#      cols = everything(),
-#      names_to = c("variable","index"),
-#      names_sep = "\\.") %>% 
-#    pivot_wider(
-#      names_from = index,
-#      values_from = value)
+psych::describe(player_stats_seasonal["fantasyPoints"])
+
+player_stats_seasonal %>% 
+  select(age, years_of_experience, fantasyPoints) %>% 
+  summarise(across(
+      everything(),
+      .fns = list(
+        n = ~ length(na.omit(.)),
+        missingness = ~ mean(is.na(.)) * 100,
+        M = ~ mean(., na.rm = TRUE),
+        SD = ~ sd(., na.rm = TRUE),
+        min = ~ min(., na.rm = TRUE),
+        max = ~ max(., na.rm = TRUE),
+        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
+        IQR = ~ IQR(., na.rm = TRUE),
+        MAD = ~ mad(., na.rm = TRUE),
+        median = ~ median(., na.rm = TRUE),
+        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
+        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
+        skewness = ~ psych::skew(., na.rm = TRUE),
+        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
+      .names = "{.col}.{.fn}")) %>%
+    pivot_longer(
+      cols = everything(),
+      names_to = c("variable","index"),
+      names_sep = "\\.") %>% 
+    pivot_wider(
+      names_from = index,
+      values_from = value)
 ```

### 2024-12-25 bugs.qmd
20241216 - link to github issue for bug report
20241225 - convert to calculate_stats() from nflfastR v5.0

--- 2024-06-27 bugs.qmd
@@
 
-Bugs to report (`nflreadr`):
+Bugs to report (`nflreadr` or `nflfastR`):
 
@@
 - `player_stats` database of weekly stats does not have a `game_id` column (<https://github.com/nflverse/nflreadr/issues/238>)
-
-<https://github.com/nflverse/nflreadr/issues/237>
+- common naming convention for key variables, for purposes of merging (<https://github.com/nflverse/nflreadr/issues/237>)
+- use of deprecated `summarise()` for returning more than 1 row per group; use `reframe()` instead (<https://github.com/nflverse/nflfastR/issues/480>)
+- calculate career stats using `calculate_stats()` function (as was formerly able to do with the `calculate_player_stats()` function) (<https://github.com/nflverse/nflfastR/issues/501>)
+- will `nflreadr` switch to using the objects from the `calculate_stats()` function in `nflfastR`? (<https://github.com/nflverse/nflreadr/issues/261>)

### 2024-12-25 cluster-analysis.qmd
20241225 - convert to calculate_stats() from nflfastR v5.0

@@
 ```{r}
-recentSeason <- max(player_stats_seasonal_offense$season, na.rm = TRUE) # also works: nflreadr::most_recent_season()
+recentSeason <- max(player_stats_seasonal$season, na.rm = TRUE) # also works: nflreadr::most_recent_season()
 recentSeason
 
-player_stats_seasonal_offense_recent <- player_stats_seasonal_offense %>% 
-  filter(season == recentSeason)
+player_stats_seasonal_offense_recent <- player_stats_seasonal %>% 
+  filter(season == recentSeason) %>% 
+  filter(position_group %in% c("QB","RB","WR","TE"))
 
@@
 tiersQB_bic <- mclust::mclustBIC(
-  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  data = player_stats_seasonal_offense_recentQB$fantasyPoints,
   G = 1:9
@@
 tiersQB_icl <- mclust::mclustICL(
-  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  data = player_stats_seasonal_offense_recentQB$fantasyPoints,
   G = 1:9
@@
 tiersQB_boostrap <- mclust::mclustBootstrapLRT(
-  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  data = player_stats_seasonal_offense_recentQB$fantasyPoints,
   modelName = "V") # variable/unequal variance (for univariate data)
@@
 tiersRB_bic <- mclust::mclustBIC(
-  data = player_stats_seasonal_offense_recentRB$fantasy_points,
+  data = player_stats_seasonal_offense_recentRB$fantasyPoints,
   G = 1:9
@@
 tiersRB_icl <- mclust::mclustICL(
-  data = player_stats_seasonal_offense_recentRB$fantasy_points,
+  data = player_stats_seasonal_offense_recentRB$fantasyPoints,
   G = 1:9
@@
 tiersRB_boostrap <- mclust::mclustBootstrapLRT(
-  data = player_stats_seasonal_offense_recentRB$fantasy_points,
+  data = player_stats_seasonal_offense_recentRB$fantasyPoints,
   modelName = "V") # variable/unequal variance (for univariate data)
@@
 tiersWR_bic <- mclust::mclustBIC(
-  data = player_stats_seasonal_offense_recentWR$fantasy_points,
+  data = player_stats_seasonal_offense_recentWR$fantasyPoints,
   G = 1:9
@@
 tiersWR_icl <- mclust::mclustICL(
-  data = player_stats_seasonal_offense_recentWR$fantasy_points,
+  data = player_stats_seasonal_offense_recentWR$fantasyPoints,
   G = 1:9
@@
 tiersWR_boostrap <- mclust::mclustBootstrapLRT(
-  data = player_stats_seasonal_offense_recentWR$fantasy_points,
+  data = player_stats_seasonal_offense_recentWR$fantasyPoints,
   modelName = "V") # variable/unequal variance (for univariate data)
@@
 tiersTE_bic <- mclust::mclustBIC(
-  data = player_stats_seasonal_offense_recentTE$fantasy_points,
+  data = player_stats_seasonal_offense_recentTE$fantasyPoints,
   G = 1:9
@@
 tiersTE_icl <- mclust::mclustICL(
-  data = player_stats_seasonal_offense_recentTE$fantasy_points,
+  data = player_stats_seasonal_offense_recentTE$fantasyPoints,
   G = 1:9
@@
 tiersTE_boostrap <- mclust::mclustBootstrapLRT(
-  data = player_stats_seasonal_offense_recentTE$fantasy_points,
+  data = player_stats_seasonal_offense_recentTE$fantasyPoints,
   modelName = "V") # variable/unequal variance (for univariate data)
@@
 mclust::Mclust(
-  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  data = player_stats_seasonal_offense_recentQB$fantasyPoints,
   G = numTiersQB,
@@
 mclust::Mclust(
-  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  data = player_stats_seasonal_offense_recentQB$fantasyPoints,
   G = 4,
@@
 mclust::Mclust(
-  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  data = player_stats_seasonal_offense_recentQB$fantasyPoints,
 )
@@
 mclust::Mclust(
-  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  data = player_stats_seasonal_offense_recentQB$fantasyPoints,
   x = tiersQB_bic
@@
 clusterModelQBs <- mclust::Mclust(
-  data = player_stats_seasonal_offense_recentQB$fantasy_points,
+  data = player_stats_seasonal_offense_recentQB$fantasyPoints,
   G = numTiersQB,
@@
 clusterModelRBs <- mclust::Mclust(
-  data = player_stats_seasonal_offense_recentRB$fantasy_points,
+  data = player_stats_seasonal_offense_recentRB$fantasyPoints,
   G = numTiersRB,
@@
 clusterModelWRs <- mclust::Mclust(
-  data = player_stats_seasonal_offense_recentWR$fantasy_points,
+  data = player_stats_seasonal_offense_recentWR$fantasyPoints,
   G = numTiersWR,
@@
 clusterModelTEs <- mclust::Mclust(
-  data = player_stats_seasonal_offense_recentTE$fantasy_points,
+  data = player_stats_seasonal_offense_recentTE$fantasyPoints,
   G = numTiersTE,
@@
   mutate(
-    tier = factor(case_match( # recode so 1 is highest tier
-      tier,
-      1 ~ 4,
-      2 ~ 3,
-      3 ~ 2,
-      4 ~ 1)))
+    tier = factor(max(tier, na.rm = TRUE) + 1 - tier)
+  )
 
 player_stats_seasonal_offense_recentQB$position_rank <- rank(
-  player_stats_seasonal_offense_recentQB$fantasy_points * -1,
+  player_stats_seasonal_offense_recentQB$fantasyPoints * -1,
   na.last = "keep",
@@
   mapping = aes(
-    x = fantasy_points,
+    x = fantasyPoints,
     y = position_rank,
@@
   mutate(
-    tier = factor(case_match( # recode so 1 is highest tier
-      tier,
-      1 ~ 3,
-      2 ~ 2,
-      3 ~ 1)))
+    tier = factor(max(tier, na.rm = TRUE) + 1 - tier)
+  )
 
 player_stats_seasonal_offense_recentRB$position_rank <- rank(
-  player_stats_seasonal_offense_recentRB$fantasy_points * -1,
+  player_stats_seasonal_offense_recentRB$fantasyPoints * -1,
   na.last = "keep",
@@
   mutate(
-    tier = factor(case_match( # recode so 1 is highest tier
-      tier,
-      1 ~ 5,
-      2 ~ 4,
-      3 ~ 3,
-      4 ~ 2,
-      5 ~ 1)))
+    tier = factor(max(tier, na.rm = TRUE) + 1 - tier)
+  )
 
 player_stats_seasonal_offense_recentWR$position_rank <- rank(
-  player_stats_seasonal_offense_recentWR$fantasy_points * -1,
+  player_stats_seasonal_offense_recentWR$fantasyPoints * -1,
   na.last = "keep",
@@
   mutate(
-    tier = factor(case_match( # recode so 1 is highest tier
-      tier,
-      1 ~ 4,
-      2 ~ 3,
-      3 ~ 2,
-      4 ~ 1)))
+    tier = factor(max(tier, na.rm = TRUE) + 1 - tier)
+  )
 
 player_stats_seasonal_offense_recentTE$position_rank <- rank(
-  player_stats_seasonal_offense_recentTE$fantasy_points * -1,
+  player_stats_seasonal_offense_recentTE$fantasyPoints * -1,
   na.last = "keep",
@@
 # Compute Player Stats Per Season
-player_stats_seasonal_careerWRs <- player_stats_seasonal_offense %>% 
+player_stats_seasonal_careerWRs <- player_stats_seasonal %>% 
   filter(position == "WR") %>% 

### 2024-12-25 data-visualization.qmd
20241225 - convert to calculate_stats() from nflfastR v5.0

@@
 
-player_stats_seasonal_offense_summary <- player_stats_seasonal_offense %>%
+player_stats_seasonal_offense_summary <- player_stats_seasonal %>%
   filter(position_group %in% c("QB","RB","WR","TE")) %>% 
@@
   summarise( 
-    n = sum(!is.na(fantasy_points)),
-    mean = mean(fantasy_points, na.rm = TRUE),
-    sd = sd(fantasy_points, na.rm = TRUE)
+    n = sum(!is.na(fantasyPoints)),
+    mean = mean(fantasyPoints, na.rm = TRUE),
+    sd = sd(fantasyPoints, na.rm = TRUE)
   ) %>%
@@
 ggplot2::ggplot(
-  data = player_stats_seasonal_offense %>%
+  data = player_stats_seasonal %>%
     filter(position_group %in% c("QB","RB","WR","TE")),
   mapping = aes(
-    x = fantasy_points)
+    x = fantasyPoints)
 ) +
@@
 ggplot2::ggplot(
-  data = player_stats_seasonal_offense %>%
+  data = player_stats_seasonal %>%
     filter(position_group %in% c("QB","RB","WR","TE")),
   mapping = aes(
-    x = fantasy_points,
+    x = fantasyPoints,
     fill = position_group)
@@
 ggplot2::ggplot(
-  data = player_stats_seasonal_offense %>%
+  data = player_stats_seasonal %>%
     filter(position_group %in% c("QB","RB","WR","TE")),
@@
     x = position_group,
-    y = fantasy_points,
+    y = fantasyPoints,
     fill = position_group)
@@
 ggplot2::ggplot(
-  data = player_stats_seasonal_offense %>%
+  data = player_stats_seasonal %>%
     filter(position_group %in% c("QB","RB","WR","TE")),
   mapping = aes(
-    x = fantasy_points,
+    x = fantasyPoints,
     y = position_group,
@@
 ggplot2::ggplot(
-  data = player_stats_seasonal_offense %>%
+  data = player_stats_seasonal %>%
     filter(player_display_name == "Tom Brady"),
@@
     x = season,
-    y = fantasy_points
+    y = fantasyPoints
     )
@@
 ggplot2::ggplot(
-  data = player_stats_seasonal_offense,
+  data = player_stats_seasonal %>% 
+    filter(position_group %in% c("QB")),
   mapping = aes(
     x = season,
-    y = fantasy_points,
+    y = fantasyPoints,
     group = player_id,
@@
 ) +
-  geom_line() +
+  geom_line(linewidth = 2) +
   gghighlight::gghighlight(
     player_display_name == "Tom Brady",
-    label_key = player_display_name) +
+    label_key = player_display_name,
+    unhighlighted_params = list(linewidth = 0.5)) +
   labs(
@@
 
-player_stats_seasonal_offense_summary <- player_stats_seasonal_offense %>%
+player_stats_seasonal_offense_summary <- player_stats_seasonal %>%
     filter(position_group %in% c("QB","RB","WR","TE")) %>% 
@@
 ggplot2::ggplot(
-  data = player_stats_seasonal_offense %>% 
+  data = player_stats_seasonal %>% 
     filter(position_group %in% c("QB","RB","WR","TE")),
@@
     x = age,
-    y = fantasy_points)) +
+    y = fantasyPoints)) +
   geom_point() +

### 2024-12-25 download-football-data.qmd
20241225 - convert to calculate_stats() from nflfastR v5.0

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation_raw.RData", fsep = ""))
+load(file = "./data/nfl_actualStats_position_weekly_raw.RData")
 load(file = "./data/nfl_actualStats_weekly_raw.RData")
@@
 load(file = "./data/nfl_actualStats_seasonal_raw.RData")
+load(file = "./data/nfl_actualFantasyPoints_career_raw.RData")
+load(file = "./data/nfl_actualFantasyPoints_seasonal_raw.RData")
+load(file = "./data/nfl_actualFantasyPoints_weekly_raw.RData")
 
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
+load(file = "./data/nfl_actualStats_position_weekly.RData")
 load(file = "./data/nfl_actualStats_weekly.RData")
@@
 load(file = "./data/nfl_actualStats_seasonal.RData")
+load(file = "./data/nfl_actualFantasyPoints_career.RData")
+load(file = "./data/nfl_actualFantasyPoints_seasonal.RData")
+load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
 ```
@@
 
-### Historical Weekly Actual Player Statistics {#sec-downloadActualStats}
+### Historical Actual Player Statistics {#sec-downloadActualStats}
+
+#### Career Statistics {#sec-downloadActualCareerStats}
+
+#### Season-by-Season Statistics {#sec-downloadActualSeasonalStats}
+
+#### Week-by-Week Statistics {#sec-downloadActualWeeklyStats}
 
@@
   nfl_actualStats_offense_weekly_raw, nfl_actualStats_defense_weekly_raw, nfl_actualStats_kicking_weekly_raw,
-  file = "./data/nfl_actualStats_weekly_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_weekly_raw.RData")
+  file = "./data/nfl_actualStats_position_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_position_weekly_raw.RData")
 ```
@@
   nfl_actualStats_offense_weekly, nfl_actualStats_defense_weekly, nfl_actualStats_kicking_weekly,
-  file = "./data/nfl_actualStats_weekly.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_weekly.RData")
+  file = "./data/nfl_actualStats_position_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_position_weekly.RData")
 ```
@@
 
-Now, modify the scoring settings to match our league settings (based on [scoring settings](https://support.nfl.com/hc/en-us/articles/4989179237404-Scoring) for fantasy leagues on NFL.com):
+Now, modify the scoring settings to match your league settings.
+Below, we use the [scoring settings](https://support.nfl.com/hc/en-us/articles/4989179237404-Scoring) for fantasy leagues on NFL.com, which happen to be point-per-reception leagues (i.e., PPR leagues):
 
@@
 
-load("./data/nfl_actualStats_career_raw.RData")
+load(file = "./data/nfl_actualStats_career_raw.RData")
 ```
@@
 
+::: {#nte-downloadSeasonalActualStats .callout-note title="Downloading season-by-season statistics"}
+Note: the following code takes a while to run.
+:::
+
 ```{r}
@@
 
-load("./data/nfl_actualStats_seasonal_raw.RData")
+load(file = "./data/nfl_actualStats_seasonal_raw.RData")
 ```
@@
 
-```{r}
-#| eval: false
-nfl_actualStats_weekly_raw <- calculate_stats(
+::: {#nte-downloadWeeklyActualStats .callout-note title="Downloading week-by-week statistics"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+nfl_actualStats_weekly_player_raw <- calculate_stats(
   seasons = TRUE,
@@
 
-nfl_actualStats_weekly_defense_raw <- calculate_stats(
+nfl_actualStats_weekly_team_raw <- calculate_stats(
   seasons = TRUE,
@@
 
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_weekly_player_raw, nfl_actualStats_weekly_team_raw,
+  file = "./data/nfl_actualStats_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_weekly_raw.RData")
+```
+
+The `nfl_actualStats_weekly` objects are in `player`-`season`-`week` form.
+That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_weekly_player <- nfl_actualStats_weekly_player_raw %>% 
+  select(player_id, season, week, everything()) %>% 
+  arrange(player_display_name, player_id, season, week)
+
+nfl_actualStats_weekly_team <- nfl_actualStats_weekly_team_raw %>% 
+  select(season, week, team, everything()) %>% 
+  arrange(season, week, team)
+```
+
+Let's check for duplicate `player`-`season`-`week` instances:
+
+```{r}
+nfl_actualStats_weekly_player %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_weekly_team %>% 
+  group_by(team, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_weekly_player, nfl_actualStats_weekly_team,
+  file = "./data/nfl_actualStats_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_weekly.RData")
+```
+
 ##### OLD (DELETE CODE BELOW)
@@
 
-#### Weekly {#sec-calculateActualPointsWeekly}
-
-#### Seasonal {#sec-calculateActualPointsSeasonal}
+#### Week-by-Week {#sec-calculateActualPointsWeekly}
+
+```{r}
+nfl_actualFantasyPoints_player_weekly_seasonalList <- list()
+nfl_actualFantasyPoints_dst_weekly_seasonalList <- list()
+```
+
+::: {#nte-calculateActualFantasyPointsWeekly .callout-note title="Calculating players' week-by-week fantasy points"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+
+#nfl_actualFantasyPoints_weekly_raw <- ffanalytics:::actual_points_scoring(
+#  season = 2023,
+#  summary_level = c("week"),
+#  stat_type = c("player", "dst", "team"), 
+#  season_type = c("REG", "POST", "REG+POST"),
+#  scoring_rules = scoring_obj,
+#  vor_baseline = NULL,
+#  rename_colums = FALSE
+#)
+
+pb <- txtProgressBar(
+  min = 0,
+  max = length(seasons),
+  style = 3)
+
+for(i in 1:length(seasons)){
+  # Compute actual statistics by season
+  nfl_actualFantasyPoints_player_weekly_seasonalList[[i]] <- 
+    ffanalytics:::actual_points_scoring(
+      season = seasons[i],
+      summary_level = c("week"),
+      stat_type = c("player"), 
+      #season_type = c("REG"),
+      scoring_rules = scoring_obj,
+      vor_baseline = NULL,
+      rename_colums = FALSE)
+
+  nfl_actualFantasyPoints_dst_weekly_seasonalList[[i]] <- 
+    ffanalytics:::actual_points_scoring(
+      season = seasons[i],
+      summary_level = c("week"),
+      stat_type = c("dst"), 
+      #season_type = c("REG"),
+      scoring_rules = scoring_obj,
+      vor_baseline = NULL,
+      rename_colums = FALSE)
+
+  nfl_actualFantasyPoints_player_weekly_seasonalList[[i]]$season <- seasons[i]
+  nfl_actualFantasyPoints_dst_weekly_seasonalList[[i]]$season <- seasons[i]
+
+  print(
+    paste("Completed computing actual fantasy points for season: ", seasons[i], sep = ""))
+
+  # Update the progress bar
+  setTxtProgressBar(pb, i)
+}
+
+# Close the progress bar
+close(pb)
+
+nfl_actualFantasyPoints_player_weekly_raw <- nfl_actualFantasyPoints_player_weekly_seasonalList %>% 
+  dplyr::bind_rows()
+
+nfl_actualFantasyPoints_dst_weekly_raw <- nfl_actualFantasyPoints_dst_weekly_seasonalList %>% 
+  dplyr::bind_rows()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualFantasyPoints_player_weekly_raw, nfl_actualFantasyPoints_dst_weekly_raw,
+  file = "./data/nfl_actualFantasyPoints_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualFantasyPoints_weekly_raw.RData")
+```
+
+The `nfl_actualFantasyPoints_weekly` objects are in `player`-`season`-`week` (or `team`-`season`-`week`) form.
+That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week` (or `team`-`season`-`week`).
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualFantasyPoints_player_weekly <- nfl_actualFantasyPoints_player_weekly_raw %>% 
+  rename(
+    fantasyPoints = raw_points,
+    team = recent_team) %>% 
+  select(player_id, season, week, everything()) %>% 
+  arrange(player_display_name, player_id, season, week)
+
+nfl_actualFantasyPoints_dst_weekly <- nfl_actualFantasyPoints_dst_weekly_raw %>% 
+  rename(
+    fantasyPoints = raw_points,
+    team = recent_team) %>% 
+  select(season, week, team, everything()) %>% 
+  arrange(season, week, team)
+```
+
+Let's check for duplicate `player`-`season`-`week` instances:
+
+```{r}
+nfl_actualFantasyPoints_player_weekly %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualFantasyPoints_dst_weekly %>% 
+  group_by(team, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualFantasyPoints_player_weekly, nfl_actualFantasyPoints_dst_weekly,
+  file = "./data/nfl_actualFantasyPoints_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
+```
+
+##### OLD (DELETE CODE BELOW)
+
+```{r}
+#| eval: false
+
+nfl_actualStats_offense_weekly
+nfl_actualStats_defense_weekly
+nfl_actualStats_kicking_weekly
+
+test_weekly_offense <- ffanalytics:::actual_points_scoring(
+  nflr_player_stats_df = nfl_actualStats_offense_weekly %>% filter(season == 2023),
+  nflr_pbp_data = nfl_pbp %>% filter(season == 2023),
+  scoring_rules = scoring_obj,
+  vor_baseline = NULL,
+  rename_colums = TRUE)
+
+test_weekly_defense <- ffanalytics:::actual_points_scoring(
+  nflr_player_stats_df = nfl_actualStats_defense_weekly %>% filter(season == 2023),
+  nflr_pbp_data = nfl_pbp %>% filter(season == 2023),
+  scoring_rules = scoring_obj,
+  vor_baseline = NULL,
+  rename_colums = TRUE)
+
+test_weekly_kicking <- ffanalytics:::actual_points_scoring(
+  nflr_player_stats_df = nfl_actualStats_kicking_weekly %>% filter(season == 2023),
+  nflr_pbp_data = nfl_pbp %>% filter(season == 2023),
+  scoring_rules = scoring_obj,
+  vor_baseline = NULL,
+  rename_colums = TRUE)
+```
+
+#### Season-by-Season {#sec-calculateActualPointsSeasonal}
+
+```{r}
+nfl_actualFantasyPoints_player_seasonal_seasonalList <- list()
+nfl_actualFantasyPoints_dst_seasonal_seasonalList <- list()
+```
+
+::: {#nte-calculateActualFantasyPointsSeasonal .callout-note title="Calculating players' season-by-season fantasy points"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+
+#nfl_actualFantasyPoints_seasonal_raw <- ffanalytics:::actual_points_scoring(
+#  season = 2023,
+#  summary_level = c("season"),
+#  stat_type = c("player", "dst", "team"), 
+#  season_type = c("REG"),
+#  scoring_rules = scoring_obj,
+#  vor_baseline = NULL,
+#  rename_colums = TRUE
+#)
+
+pb <- txtProgressBar(
+  min = 0,
+  max = length(seasons),
+  style = 3)
+
+for(i in 1:length(seasons)){
+  # Compute actual statistics by season
+  nfl_actualFantasyPoints_player_seasonal_seasonalList[[i]] <- 
+    ffanalytics:::actual_points_scoring(
+      season = seasons[i],
+      summary_level = c("season"),
+      stat_type = c("player"), 
+      season_type = c("REG"),
+      scoring_rules = scoring_obj,
+      vor_baseline = NULL,
+      rename_colums = FALSE)
+
+  nfl_actualFantasyPoints_dst_seasonal_seasonalList[[i]] <- 
+    ffanalytics:::actual_points_scoring(
+      season = seasons[i],
+      summary_level = c("season"),
+      stat_type = c("dst"), 
+      season_type = c("REG"),
+      scoring_rules = scoring_obj,
+      vor_baseline = NULL,
+      rename_colums = FALSE)
+
+  nfl_actualFantasyPoints_player_seasonal_seasonalList[[i]]$season <- seasons[i]
+  nfl_actualFantasyPoints_dst_seasonal_seasonalList[[i]]$season <- seasons[i]
+
+  print(
+    paste("Completed computing actual fantasy points for season: ", seasons[i], sep = ""))
+
+  # Update the progress bar
+  setTxtProgressBar(pb, i)
+}
+
+# Close the progress bar
+close(pb)
+
+nfl_actualFantasyPoints_player_seasonal_raw <- nfl_actualFantasyPoints_player_seasonal_seasonalList %>% 
+  dplyr::bind_rows()
+
+nfl_actualFantasyPoints_dst_seasonal_raw <- nfl_actualFantasyPoints_dst_seasonal_seasonalList %>% 
+  dplyr::bind_rows()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualFantasyPoints_player_seasonal_raw, nfl_actualFantasyPoints_dst_seasonal_raw,
+  file = "./data/nfl_actualFantasyPoints_seasonal_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualFantasyPoints_seasonal_raw.RData")
+```
+
+The `nfl_actualFantasyPoints_seasonal` objects are in `player`-`season` (or `team`-`season`) form.
+That is, each row should be uniquely identified by the combination of `player_id` and `season` (or `team`-`season`).
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualFantasyPoints_player_seasonal <- nfl_actualFantasyPoints_player_seasonal_raw %>% 
+  rename(fantasyPoints = raw_points) %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualFantasyPoints_dst_seasonal <- nfl_actualFantasyPoints_dst_seasonal_raw %>% 
+  rename(
+    fantasyPoints = raw_points,
+    team = recent_team) %>% 
+  select(season, team, everything()) %>% 
+  arrange(season, team)
+```
+
+Let's check for duplicate `player`-`season`-`week` instances:
+
+```{r}
+nfl_actualFantasyPoints_player_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualFantasyPoints_dst_seasonal %>% 
+  group_by(team, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualFantasyPoints_player_seasonal, nfl_actualFantasyPoints_dst_seasonal,
+  file = "./data/nfl_actualFantasyPoints_seasonal.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualFantasyPoints_seasonal.RData")
+```
 
@@
 
+```{r}
+nfl_actualFantasyPoints_player_career_raw <- nfl_actualFantasyPoints_player_seasonal %>% 
+  group_by(player_id) %>% 
+  summarise(fantasyPoints = sum(fantasyPoints, na.rm = TRUE))
+
+nfl_actualFantasyPoints_player_career_raw <- full_join(
+  nfl_actualFantasyPoints_player_career_raw,
+  nfl_players %>% select(gsis_id, display_name, first_name, last_name, position_group, position),
+  by = c("player_id" = "gsis_id")
+)
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualFantasyPoints_player_career,
+  file = "./data/nfl_actualFantasyPoints_career_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualFantasyPoints_career_raw.RData")
+```
+
+```{r}
+nfl_actualFantasyPoints_player_career <- nfl_actualFantasyPoints_player_career_raw %>% 
+  select(player_id, display_name, first_name, last_name, position_group, position, fantasyPoints)
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualFantasyPoints_player_career,
+  file = "./data/nfl_actualFantasyPoints_career.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualFantasyPoints_career.RData")
+```
+
 ### Player Age and Experience {#sec-calculatePlayerAge}
@@
 # Reshape from wide to long format
-nfl_actualStats_offense_weekly_long <- nfl_actualStats_offense_weekly %>% 
+nfl_actualFantasyPoints_player_weekly_long <- nfl_actualFantasyPoints_player_weekly %>% 
   tidyr::pivot_longer(
@@
 # Perform separate inner join operations for the home_team and away_team
-nfl_actualStats_offense_weekly_home <- dplyr::inner_join(
-  nfl_actualStats_offense_weekly_long,
+nfl_actualFantasyPoints_player_weekly_home <- dplyr::inner_join(
+  nfl_actualFantasyPoints_player_weekly_long,
   nfl_schedules,
@@
 
-nfl_actualStats_offense_weekly_away <- dplyr::inner_join(
-  nfl_actualStats_offense_weekly_long,
+nfl_actualFantasyPoints_player_weekly_away <- dplyr::inner_join(
+  nfl_actualFantasyPoints_player_weekly_long,
   nfl_schedules,
@@
 
-nfl_actualStats_defense_weekly_home <- dplyr::inner_join(
-  nfl_actualStats_defense_weekly,
-  nfl_schedules,
-  by = c("season","week","team" = "home_team")) %>% 
-  mutate(home_away = "home_team")
-
-nfl_actualStats_defense_weekly_away <- dplyr::inner_join(
-  nfl_actualStats_defense_weekly,
-  nfl_schedules,
-  by = c("season","week","team" = "away_team")) %>% 
-  mutate(home_away = "away_team")
-
-nfl_actualStats_kicking_weekly_home <- dplyr::inner_join(
-  nfl_actualStats_kicking_weekly,
-  nfl_schedules,
-  by = c("season","week","team" = "home_team")) %>% 
-  mutate(home_away = "home_team")
-
-nfl_actualStats_kicking_weekly_away <- dplyr::inner_join(
-  nfl_actualStats_kicking_weekly,
-  nfl_schedules,
-  by = c("season","week","team" = "away_team")) %>% 
-  mutate(home_away = "away_team")
-
 # Combine the results of the join operations
-nfl_actualStats_offense_weekly_schedules_long <- dplyr::bind_rows(
-  nfl_actualStats_offense_weekly_home,
-  nfl_actualStats_offense_weekly_away)
-
-nfl_actualStats_defense_weekly_schedules_long <- dplyr::bind_rows(
-  nfl_actualStats_defense_weekly_home,
-  nfl_actualStats_defense_weekly_away)
-
-nfl_actualStats_kicking_weekly_schedules_long <- dplyr::bind_rows(
-  nfl_actualStats_kicking_weekly_home,
-  nfl_actualStats_kicking_weekly_away)
+nfl_actualFantasyPoints_player_weekly_schedules_long <- dplyr::bind_rows(
+  nfl_actualFantasyPoints_player_weekly_home,
+  nfl_actualFantasyPoints_player_weekly_away)
 
 # Reshape from long to wide
-player_game_gameday_offense <- nfl_actualStats_offense_weekly_schedules_long %>%
+player_game_gameday <- nfl_actualFantasyPoints_player_weekly_schedules_long %>%
   dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
@@
 
-player_game_gameday_defense <- nfl_actualStats_defense_weekly_schedules_long %>%
-  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
-  tidyr::pivot_wider(
-    names_from = home_away,
-    values_from = team)
-
-player_game_gameday_kicking <- nfl_actualStats_kicking_weekly_schedules_long %>%
-  dplyr::distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
-  tidyr::pivot_wider(
-    names_from = home_away,
-    values_from = team)
-
 # Merge player birthdate and the game date
-player_game_birthdate_gameday_offense <- dplyr::left_join(
-  player_game_gameday_offense,
+player_game_birthdate_gameday <- dplyr::left_join(
+  player_game_gameday,
   unique(nfl_players[,c("gsis_id","birth_date")]),
@@
 
-player_game_birthdate_gameday_defense <- dplyr::left_join(
-  player_game_gameday_defense,
-  unique(nfl_players[,c("gsis_id","birth_date")]),
-  by = c("player_id" = "gsis_id")
-)
-
-player_game_birthdate_gameday_kicking <- dplyr::left_join(
-  player_game_gameday_kicking,
-  unique(nfl_players[,c("gsis_id","birth_date")]),
-  by = c("player_id" = "gsis_id")
-)
-
-player_game_birthdate_gameday_offense$birth_date <- lubridate::ymd(player_game_birthdate_gameday_offense$birth_date)
-player_game_birthdate_gameday_offense$gameday <- lubridate::ymd(player_game_birthdate_gameday_offense$gameday)
-
-player_game_birthdate_gameday_defense$birth_date <- lubridate::ymd(player_game_birthdate_gameday_defense$birth_date)
-player_game_birthdate_gameday_defense$gameday <- lubridate::ymd(player_game_birthdate_gameday_defense$gameday)
-
-player_game_birthdate_gameday_kicking$birth_date <- lubridate::ymd(player_game_birthdate_gameday_kicking$birth_date)
-player_game_birthdate_gameday_kicking$gameday <- lubridate::ymd(player_game_birthdate_gameday_kicking$gameday)
+player_game_birthdate_gameday$birth_date <- lubridate::ymd(player_game_birthdate_gameday$birth_date)
+player_game_birthdate_gameday$gameday <- lubridate::ymd(player_game_birthdate_gameday$gameday)
 
 # Calculate player's age for a given week as the difference between their birthdate and the game date
-player_game_birthdate_gameday_offense$age <- lubridate::interval(
-  start = player_game_birthdate_gameday_offense$birth_date,
-  end = player_game_birthdate_gameday_offense$gameday
+player_game_birthdate_gameday$age <- lubridate::interval(
+  start = player_game_birthdate_gameday$birth_date,
+  end = player_game_birthdate_gameday$gameday
 ) %>% 
@@
 
-player_game_birthdate_gameday_defense$age <- lubridate::interval(
-  start = player_game_birthdate_gameday_defense$birth_date,
-  end = player_game_birthdate_gameday_defense$gameday
-) %>% 
-  lubridate::time_length(unit = "years")
-
-player_game_birthdate_gameday_kicking$age <- lubridate::interval(
-  start = player_game_birthdate_gameday_kicking$birth_date,
-  end = player_game_birthdate_gameday_kicking$gameday
-) %>% 
-  lubridate::time_length(unit = "years")
-
 # Merge with Pro Football Reference Data on Player Age by Season
-player_game_birthdate_gameday_offense <- player_game_birthdate_gameday_offense %>% 
+player_game_birthdate_gameday <- player_game_birthdate_gameday %>% 
   dplyr::left_join(
@@
 
-player_game_birthdate_gameday_defense <- player_game_birthdate_gameday_defense %>% 
-  dplyr::left_join(
-    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
-    by = c("player_id" = "gsis_id", "season")
-  )
-
-player_game_birthdate_gameday_kicking <- player_game_birthdate_gameday_kicking %>% 
-  dplyr::left_join(
-    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
-    by = c("player_id" = "gsis_id", "season")
-  )
-
 # Set age as first non-missing value from calculation above or from PFR
-player_game_birthdate_gameday_offense <- player_game_birthdate_gameday_offense %>% 
+player_game_birthdate_gameday <- player_game_birthdate_gameday %>% 
   mutate(age = coalesce(age.x, age.y)) %>% 
@@
 
-player_game_birthdate_gameday_defense <- player_game_birthdate_gameday_defense %>% 
-  mutate(age = coalesce(age.x, age.y)) %>% 
-  select(-age.x, -age.y)
-
-player_game_birthdate_gameday_kicking <- player_game_birthdate_gameday_kicking %>% 
-  mutate(age = coalesce(age.x, age.y)) %>% 
-  select(-age.x, -age.y)
-
 # Calculate ageCentered and ageCenteredQuadratic
-player_game_birthdate_gameday_offense$ageCentered20 <- player_game_birthdate_gameday_offense$age - 20
-player_game_birthdate_gameday_offense$ageCentered20Quadratic <- player_game_birthdate_gameday_offense$ageCentered20 ^ 2
-
-player_game_birthdate_gameday_defense$ageCentered20 <- player_game_birthdate_gameday_defense$age - 20
-player_game_birthdate_gameday_defense$ageCentered20Quadratic <- player_game_birthdate_gameday_defense$ageCentered20 ^ 2
-
-player_game_birthdate_gameday_kicking$ageCentered20 <- player_game_birthdate_gameday_kicking$age - 20
-player_game_birthdate_gameday_kicking$ageCentered20Quadratic <- player_game_birthdate_gameday_kicking$ageCentered20 ^ 2
+player_game_birthdate_gameday$ageCentered20 <- player_game_birthdate_gameday$age - 20
+player_game_birthdate_gameday$ageCentered20Quadratic <- player_game_birthdate_gameday$ageCentered20 ^ 2
 
 # Merge with player info
-player_age_offense <- dplyr::left_join(
-  player_game_birthdate_gameday_offense,
+player_age <- dplyr::left_join(
+  player_game_birthdate_gameday,
   nfl_players %>% select(-birth_date, -team_abbr, - team_seq),
@@
 
-player_age_defense <- dplyr::left_join(
-  player_game_birthdate_gameday_defense,
-  nfl_players %>% select(-birth_date, -team_abbr, - team_seq),
-  by = c("player_id" = "gsis_id"))
-
-player_age_kicking <- dplyr::left_join(
-  player_game_birthdate_gameday_kicking,
-  nfl_players %>% select(-birth_date, -team_abbr, - team_seq),
-  by = c("player_id" = "gsis_id"))
-
 # Add game_id to weekly stats to facilitate merging
-nfl_actualStats_game_offense_weekly <- nfl_actualStats_offense_weekly %>% 
+nfl_actualFantasyPoints_player_weekly <- nfl_actualFantasyPoints_player_weekly %>% 
   dplyr::left_join(
-    player_age_offense[,c("season","week","player_id","game_id")],
+    player_age[,c("season","week","player_id","game_id")],
     by = c("season","week","player_id"))
 
-nfl_actualStats_game_defense_weekly <- nfl_actualStats_defense_weekly %>% 
-  dplyr::left_join(
-    player_age_offense[,c("season","week","player_id","game_id")],
-    by = c("season","week","player_id"))
-
-nfl_actualStats_game_kicking_weekly <- nfl_actualStats_kicking_weekly %>% 
-  dplyr::left_join(
-    player_age_offense[,c("season","week","player_id","game_id")],
-    by = c("season","week","player_id"))
-
 # Merge with player weekly stats
-player_stats_weekly_offense <- dplyr::full_join(
-  player_age_offense %>% select(-position, -position_group),
-  nfl_actualStats_game_offense_weekly,
+player_stats_weekly <- dplyr::full_join(
+  player_age %>% select(-position, -position_group),
+  nfl_actualFantasyPoints_player_weekly,
   by = c("season","week","player_id","game_id"))
 
-player_stats_weekly_defense <- dplyr::full_join(
-  player_age_defense %>% select(-position, -position_group),
-  nfl_actualStats_game_defense_weekly,
-  by = c("season","week","player_id","game_id"))
-
-player_stats_weekly_kicking <- dplyr::full_join(
-  player_age_kicking %>% select(-position, -position_group),
-  nfl_actualStats_game_kicking_weekly,
-  by = c("season","week","player_id","game_id"))
-
-player_stats_weekly_offense$total_years_of_experience <- as.integer(player_stats_weekly_offense$years_of_experience)
-player_stats_weekly_defense$total_years_of_experience <- as.integer(player_stats_weekly_defense$years_of_experience)
-player_stats_weekly_kicking$total_years_of_experience <- as.integer(player_stats_weekly_kicking$years_of_experience)
-
-player_stats_weekly_offense$years_of_experience <- NULL
-player_stats_weekly_defense$years_of_experience <- NULL
-player_stats_weekly_kicking$years_of_experience <- NULL
-
-distinct_seasons_offense <- player_stats_weekly_offense %>%
+player_stats_weekly$total_years_of_experience <- as.integer(player_stats_weekly$years_of_experience)
+
+player_stats_weekly$years_of_experience <- NULL
+
+distinct_seasons <- player_stats_weekly %>%
   dplyr::select(player_id, season) %>%
@@
 
-distinct_seasons_defense <- player_stats_weekly_defense %>%
-  dplyr::select(player_id, season) %>%
-  dplyr::distinct() %>% 
-  dplyr::left_join(
-    nfl_players[,c("gsis_id","years_of_experience")],
-    by = c("player_id" = "gsis_id")
-  ) %>% 
-  dplyr::mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
-  dplyr::select(-years_of_experience)
-
-distinct_seasons_kicking <- player_stats_weekly_kicking %>%
-  dplyr::select(player_id, season) %>%
-  dplyr::distinct() %>% 
-  dplyr::left_join(
-    nfl_players[,c("gsis_id","years_of_experience")],
-    by = c("player_id" = "gsis_id")
-  ) %>% 
-  dplyr::mutate(total_years_of_experience = as.integer(years_of_experience)) %>% 
-  dplyr::select(-years_of_experience)
-
-years_of_experience_offense <- distinct_seasons_offense %>% 
+years_of_experience <- distinct_seasons %>% 
   dplyr::arrange(player_id, -season) %>% 
@@
 
-years_of_experience_defense <- distinct_seasons_defense %>% 
-  dplyr::arrange(player_id, -season) %>% 
-  dplyr::group_by(player_id) %>%
-  dplyr::mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
-  dplyr::ungroup()
-
-years_of_experience_kicking <- distinct_seasons_kicking %>% 
-  dplyr::arrange(player_id, -season) %>% 
-  dplyr::group_by(player_id) %>%
-  dplyr::mutate(years_of_experience = first(total_years_of_experience) - (row_number() - 1)) %>%
-  dplyr::ungroup()
-
-years_of_experience_offense$years_of_experience[which(years_of_experience_offense$years_of_experience < 0)] <- 0
-years_of_experience_defense$years_of_experience[which(years_of_experience_defense$years_of_experience < 0)] <- 0
-years_of_experience_kicking$years_of_experience[which(years_of_experience_kicking$years_of_experience < 0)] <- 0
-
-player_stats_weekly_offense <- player_stats_weekly_offense %>% 
+years_of_experience$years_of_experience[which(years_of_experience$years_of_experience < 0)] <- 0
+
+player_stats_weekly <- player_stats_weekly %>% 
   dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
-    by = c("player_id","season")
-  )
-
-player_stats_weekly_defense <- player_stats_weekly_defense %>% 
-  dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
-    by = c("player_id","season")
-  )
-
-player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
-  dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    years_of_experience[,c("player_id","season","years_of_experience")],
     by = c("player_id","season")
@@
 ```{r}
-player_stats_weekly_offense <- player_stats_weekly_offense %>% 
+player_stats_weekly <- player_stats_weekly %>% 
   arrange(player_display_name, player_id, season, week)
-
-player_stats_weekly_defense <- player_stats_weekly_defense %>% 
-  arrange(player_display_name, player_id, season, week)
-
-player_stats_weekly_kicking <- player_stats_weekly_kicking %>% 
-  arrange(player_display_name, player_id, season, week)
 ```
@@
 ```{r}
-player_stats_weekly_offense %>% 
+player_stats_weekly %>% 
   group_by(player_id, season, week) %>% 
@@
   head()
-
-player_stats_weekly_defense %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head() #todo: fix duplicate instances
-
-player_stats_weekly_kicking %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head() #todo: fix duplicate instances
 ```
@@
 save(
-  player_stats_weekly_offense, player_stats_weekly_defense, player_stats_weekly_kicking,
+  player_stats_weekly,
   file = "./data/player_stats_weekly.RData"
@@
 # Merge player info with seasonal stats
-player_stats_seasonal_offense <- dplyr::full_join(
-  nfl_actualStats_offense_seasonal,
-  nfl_players %>% select(-position, -position_group, -team_abbr, - team_seq),
-  by = c("player_id" = "gsis_id")
-)
-
-player_stats_seasonal_defense <- dplyr::full_join(
-  nfl_actualStats_defense_seasonal,
-  nfl_players %>% select(-position, -position_group, -team_abbr, - team_seq),
-  by = c("player_id" = "gsis_id")
-)
-
-player_stats_seasonal_kicking <- dplyr::full_join(
-  nfl_actualStats_kicking_seasonal,
+player_stats_seasonal <- dplyr::full_join(
+  nfl_actualFantasyPoints_player_seasonal,
   nfl_players %>% select(-position, -position_group, -team_abbr, - team_seq),
@@
 
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+player_stats_seasonal <- player_stats_seasonal %>% 
   dplyr::left_join(
@@
 
-player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  dplyr::left_join(
-    season_startdate,
-    by = "season"
-  )
-
-player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  dplyr::left_join(
-    season_startdate,
-    by = "season"
-  )
-
-player_stats_seasonal_offense$age <- lubridate::interval(
-  start = player_stats_seasonal_offense$birth_date,
-  end = player_stats_seasonal_offense$startdate
+player_stats_seasonal$age <- lubridate::interval(
+  start = player_stats_seasonal$birth_date,
+  end = player_stats_seasonal$startdate
 ) %>% 
@@
 
-player_stats_seasonal_defense$age <- lubridate::interval(
-  start = player_stats_seasonal_defense$birth_date,
-  end = player_stats_seasonal_defense$startdate
-) %>% 
-  lubridate::time_length(unit = "years")
-
-player_stats_seasonal_kicking$age <- lubridate::interval(
-  start = player_stats_seasonal_kicking$birth_date,
-  end = player_stats_seasonal_kicking$startdate
-) %>% 
-  lubridate::time_length(unit = "years")
-
 # Merge with Pro Football Reference Data on Player Age by Season
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+player_stats_seasonal <- player_stats_seasonal %>% 
   dplyr::left_join(
@@
 
-player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  dplyr::left_join(
-    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
-    by = c("player_id" = "gsis_id", "season")
-  )
-
-player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  dplyr::left_join(
-    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
-    by = c("player_id" = "gsis_id", "season")
-  )
-
 # Set age as first non-missing value from calculation above or from PFR
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+player_stats_seasonal <- player_stats_seasonal %>% 
   mutate(age = coalesce(age.x, age.y)) %>% 
@@
 
-player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  mutate(age = coalesce(age.x, age.y)) %>% 
-  select(-age.x, -age.y)
-
-player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  mutate(age = coalesce(age.x, age.y)) %>% 
-  select(-age.x, -age.y)
-
 # Calculate ageCentered and ageCenteredQuadratic
-player_stats_seasonal_offense$ageCentered20 <- player_stats_seasonal_offense$age - 20
-player_stats_seasonal_offense$ageCentered20Quadratic <- player_stats_seasonal_offense$ageCentered20 ^ 2
-
-player_stats_seasonal_defense$ageCentered20 <- player_stats_seasonal_defense$age - 20
-player_stats_seasonal_defense$ageCentered20Quadratic <- player_stats_seasonal_defense$ageCentered20 ^ 2
-
-player_stats_seasonal_kicking$ageCentered20 <- player_stats_seasonal_kicking$age - 20
-player_stats_seasonal_kicking$ageCentered20Quadratic <- player_stats_seasonal_kicking$ageCentered20 ^ 2
+player_stats_seasonal$ageCentered20 <- player_stats_seasonal$age - 20
+player_stats_seasonal$ageCentered20Quadratic <- player_stats_seasonal$ageCentered20 ^ 2
 
 # Years of experience
-player_stats_seasonal_offense$years_of_experience <- NULL
-player_stats_seasonal_defense$years_of_experience <- NULL
-player_stats_seasonal_kicking$years_of_experience <- NULL
-
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+player_stats_seasonal$years_of_experience <- NULL
+
+player_stats_seasonal <- player_stats_seasonal %>% 
   dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
-    by = c("player_id","season")
-  )
-
-player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
-    by = c("player_id","season")
-  )
-
-player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  dplyr::left_join(
-    years_of_experience_offense[,c("player_id","season","years_of_experience")],
+    years_of_experience[,c("player_id","season","years_of_experience")],
     by = c("player_id","season")
@@
 ```{r}
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+player_stats_seasonal <- player_stats_seasonal %>% 
   select(player_id, season, everything()) %>% 
   arrange(player_display_name, player_id, season)
-
-player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-
-player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
 ```
@@
 ```{r}
-player_stats_seasonal_offense %>% 
+player_stats_seasonal %>% 
   group_by(player_id, season) %>% 
@@
   head()
-
-player_stats_seasonal_defense %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-player_stats_seasonal_kicking %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
 ```
@@
 save(
-  player_stats_seasonal_offense, player_stats_seasonal_defense, player_stats_seasonal_kicking,
+  player_stats_seasonal,
   file = "./data/player_stats_seasonal.RData"

### 2024-12-25 machine-learning.qmd
20241025 - ensemble methods
20241225 - convert to calculate_stats() from nflfastR v5.0

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
-load(file = "./data/nfl_actualStats_weekly.RData")
+load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
 load(file = "./data/nfl_injuries.RData")
@@
 load(file = "./data/nfl_participation.RData")
-load(file = "./data/nfl_actualStats_weekly.RData")
+load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
 load(file = "./data/nfl_injuries.RData")
@@
 
+*Ensemble* machine learning methods combine multiple machine learning approaches with the goal that combining multiple approaches might lead to more accurate predictions that any one method might be able to achieve on its own.
+
 ### Supervised Learning {#sec-machineLearningTypesSupervised}
@@
   names(nfl_participation),
-  names(nfl_actualStats_offense_weekly),
+  names(nfl_actualFantasyPoints_player_weekly),
   names(nfl_injuries),
   names(nfl_snapCounts),
-  names(nfl_espnQBR),
+  names(nfl_espnQBR_seasonal),
+  names(nfl_espnQBR_weekly),
   names(nfl_nextGenStats_weekly),
@@
   names(nfl_playerIDs),
-  names(nfl_rankings),
+  names(nfl_rankings_draft),
+  names(nfl_rankings_weekly),
   names(nfl_expectedFantasyPoints_weekly),
@@
 nfl_combine$pfr_id
-nfl_advancedStatsPFR$pfr_id
-nfl_advancedStatsPFR$pfr_player_id
+nfl_advancedStatsPFR_seasonal$pfr_id
+#nfl_advancedStatsPFR_seasonal$pfr_player_id
 
@@
 ## Use common name for the same (gsis_id) ID variable
-nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly %>% 
+nfl_actualFantasyPoints_player_weekly <- nfl_actualFantasyPoints_player_weekly %>% 
   rename(gsis_id = player_id)
 
-nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal %>% 
+nfl_actualFantasyPoints_player_seasonal <- nfl_actualFantasyPoints_player_seasonal %>% 
   rename(gsis_id = player_id)
 
-player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
+player_stats_seasonal_offense <- player_stats_seasonal %>% 
+  filter(position_group %in% c("QB","RB","WR","TE")) %>% 
   rename(gsis_id = player_id)
 
-player_stats_weekly_offense <- player_stats_weekly_offense %>% 
+player_stats_weekly_offense <- player_stats_weekly %>% 
+  filter(position_group %in% c("QB","RB","WR","TE")) %>% 
   rename(gsis_id = player_id)
@@
 playerListToMerge <- list(
-  nfl_players %>% filter(!is.na(gsis_id)) %>% select(-season),
+  nfl_players %>% filter(!is.na(gsis_id)),
   nfl_draftPicks %>% filter(!is.na(gsis_id)) %>% select(-season)
@@
   mutate(
-    fantasyPoints_lag = lead(fantasy_points)
+    fantasyPoints_lag = lead(fantasyPoints)
   ) %>% 
@@
 seasonalData_lag %>% 
-  select(gsis_id, player_display_name, season, fantasy_points, fantasyPoints_lag)
+  select(gsis_id, player_display_name, season, fantasyPoints, fantasyPoints_lag)
 ```
@@
     height, weight, rookie_year, draft_number,
-    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
     completions:rushing_2pt_conversions, special_teams_tds, contains(".pass"), contains(".rush"))
@@
     height, weight, rookie_year, draft_number,
-    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
+    fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
     carries:special_teams_tds, contains(".rush"), contains(".rec"))
@@
   arrange(-fantasyPoints_lag) %>% 
-  select(name, fantasyPoints_lag, fantasy_points)
+  select(name, fantasyPoints_lag, fantasyPoints)
 ```

### 2024-12-25 mixed-models.qmd
20241016 - figure alt text
20241225 - convert to calculate_stats() from nflfastR v5.0

@@
 ::: {#fig-simpsonsParadox}
-![](images/simpsonsParadox.png)\
+![](images/simpsonsParadox.png){fig-alt="An Example of Simpson's Paradox.
+In this example, there is a positive association between `x` and `y` within each group (i.e., red or blue), but there is a negative association between `x` and `y` when the groups are combined.
+(Figure retrieved from <https://en.wikipedia.org/wiki/File:Simpson%27s_paradox_continuous.svg>)"}
 
@@
 ```{r}
-player_stats_seasonal_offense_subset <- player_stats_seasonal_offense %>% 
-  filter(position_group %in% c("QB","RB","WR","TE"))
+player_stats_seasonal_offense_subset <- player_stats_seasonal %>% 
+  dplyr::filter(position_group %in% c("QB","RB","WR","TE") | position %in% c("K"))
 
 player_stats_seasonal_offense_subset$position[which(player_stats_seasonal_offense_subset$position == "HB")] <- "RB"
-
-player_stats_seasonal_kicking_subset <- player_stats_seasonal_kicking %>% 
-  filter(position == "K")
-
-player_stats_seasonal_offense_subset <- bind_rows(
-  player_stats_seasonal_offense_subset,
-  player_stats_seasonal_kicking_subset
-)
 
@@
     !is.na(player_idFactor),
-    !is.na(fantasy_points),
+    !is.na(fantasyPoints),
     !is.na(positionFactor),
@@
       age = round(age, 2),
-      fantasy_points = round(fantasy_points, 2)
-    ),
-  mapping = aes(
-    x = age,
-    y = fantasy_points,
+      fantasyPoints = round(fantasyPoints, 2)
+    ),
+  mapping = aes(
+    x = age,
+    y = fantasyPoints,
     color = player_id)) +
@@
     x = age,
-    y = fantasy_points),
+    y = fantasyPoints),
     inherit.aes = FALSE
@@
   plot_scatterplotFantasyPointsByAgeQB,
-  tooltip = c("age","fantasy_points","text","label"))
+  tooltip = c("age","fantasyPoints","text","label"))
 ```
@@
 cor.test(
-  formula = ~ age + fantasy_points,
+  formula = ~ age + fantasyPoints,
   data = player_stats_seasonal_offense_subset %>% filter(position == "QB")
@@
   plot_scatterplotFantasyPointsByAgeFB,
-  tooltip = c("age","fantasy_points","text","label"))
+  tooltip = c("age","fantasyPoints","text","label"))
 ```
@@
   plot_scatterplotFantasyPointsByAgeRB,
-  tooltip = c("age","fantasy_points","text","label"))
+  tooltip = c("age","fantasyPoints","text","label"))
 ```
@@
   plot_scatterplotFantasyPointsByAgeWR,
-  tooltip = c("age","fantasy_points","text","label"))
+  tooltip = c("age","fantasyPoints","text","label"))
 ```
@@
   plot_scatterplotFantasyPointsByAgeTE,
-  tooltip = c("age","fantasy_points","text","label"))
+  tooltip = c("age","fantasyPoints","text","label"))
 ```
@@
 cor.test(
-  formula = ~ age + fantasy_points,
+  formula = ~ age + fantasyPoints,
   data = player_stats_seasonal_offense_subset %>% filter(position == "TE")
@@
   plot_rawFantasyPointsByAgeQB,
-  tooltip = c("age","fantasy_points","text","label"))
+  tooltip = c("age","fantasyPoints","text","label"))
 ```
@@
   plot_rawFantasyPointsByAgeFB,
-  tooltip = c("age","fantasy_points","text","label"))
+  tooltip = c("age","fantasyPoints","text","label"))
 ```
@@
   plot_rawFantasyPointsByAgeRB,
-  tooltip = c("age","fantasy_points","text","label"))
+  tooltip = c("age","fantasyPoints","text","label"))
 ```
@@
   plot_rawFantasyPointsByAgeWR,
-  tooltip = c("age","fantasy_points","text","label"))
+  tooltip = c("age","fantasyPoints","text","label"))
 ```
@@
   plot_rawFantasyPointsByAgeTE,
-  tooltip = c("age","fantasy_points","text","label"))
+  tooltip = c("age","fantasyPoints","text","label"))
 ```
@@
 pointsPerSeason_nullModel <- lm(
-  fantasy_points ~ 1,
+  fantasyPoints ~ 1,
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeason_linearRegression <- lm(
-  fantasy_points ~ positionFactor + ageCentered20 + positionFactor:ageCentered20,
+  fantasyPoints ~ positionFactor + ageCentered20 + positionFactor:ageCentered20,
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeason_quadraticRegression <- lm(
-  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic,
+  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic,
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeason_randomIntercepts <- lmerTest::lmer(
-  fantasy_points ~ 1 + (1 | player_idFactor),
+  fantasyPoints ~ 1 + (1 | player_idFactor),
   data = player_stats_seasonal_offense_subset,
@@
         age = round(age, 2),
-        fantasy_points = round(fantasyPoints_randomIntercepts, 2)
+        fantasyPoints = round(fantasyPoints_randomIntercepts, 2)
         ),
@@
 pointsPerSeason_position <- lmerTest::lmer(
-  fantasy_points ~ positionFactor + (1 | player_idFactor),
+  fantasyPoints ~ positionFactor + (1 | player_idFactor),
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeason_positionAgeFixedLinearSlopes <- lmerTest::lmer(
-  fantasy_points ~ positionFactor + ageCentered20 + (1 | player_idFactor),
+  fantasyPoints ~ positionFactor + ageCentered20 + (1 | player_idFactor),
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeason_positionAgeRandomLinearSlopes <- lmerTest::lmer(
-  fantasy_points ~ positionFactor + ageCentered20 + (1 + ageCentered20 | player_idFactor),
+  fantasyPoints ~ positionFactor + ageCentered20 + (1 + ageCentered20 | player_idFactor),
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes <- lmerTest::lmer(
-  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
+  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeason_positionAgeRandomLinearRandomQuadraticSlopes <- lmerTest::lmer(
-  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 + ageCentered20Quadratic | player_idFactor),
+  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + (1 + ageCentered20 + ageCentered20Quadratic | player_idFactor),
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction <- lmerTest::lmer(
-  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
+  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + (1 + ageCentered20 | player_idFactor),
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
-  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
+  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeason_gam <- bam( # using bam() instead of gam() for faster estimation due to large size of data
-  fantasy_points ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
+  fantasyPoints ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
-  fantasy_points ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
+  fantasyPoints ~ positionFactor + ageCentered20 + ageCentered20Quadratic + positionFactor:ageCentered20 + positionFactor:ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_idFactor),
   data = player_stats_seasonal_offense_subset,
@@
 pointsPerSeasonDepth_gam <- bam( # using bam() instead of gam() for faster estimation due to large size of data
-  fantasy_points ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
+  fantasyPoints ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(player_idFactor, ageCentered20, bs = "re"),
   data = player_stats_seasonal_offense_subsetDepth,
@@
 ```{r}
-player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros <- player_stats_seasonal_offense_subset$fantasyPointsPos <- player_stats_seasonal_offense_subset$fantasy_points
-player_stats_seasonal_offense_subset$fantasyPointsPos[player_stats_seasonal_offense_subset$fantasy_points < 0] <- 0
-player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros[player_stats_seasonal_offense_subset$fantasy_points <= 0] <- 0.01
-
-fantasyPointsVector <- player_stats_seasonal_offense_subset$fantasy_points %>% 
+player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros <- player_stats_seasonal_offense_subset$fantasyPointsPos <- player_stats_seasonal_offense_subset$fantasyPoints
+player_stats_seasonal_offense_subset$fantasyPointsPos[player_stats_seasonal_offense_subset$fantasyPoints < 0] <- 0
+player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros[player_stats_seasonal_offense_subset$fantasyPoints <= 0] <- 0.01
+
+fantasyPointsVector <- player_stats_seasonal_offense_subset$fantasyPoints %>% 
     na.omit() %>% 
@@
   mapping = aes(
-    x = fantasy_points)
+    x = fantasyPoints)
 ) +
@@
 model_normal <- glm(
-  formula = fantasy_points ~ 1,
+  formula = fantasyPoints ~ 1,
   family = gaussian(),
@@
 ```{r}
-player_stats_seasonal_offense_subset$fantasy_points_posOnly <- player_stats_seasonal_offense_subset$fantasy_points
-player_stats_seasonal_offense_subset$fantasy_points_posOnly[which(player_stats_seasonal_offense_subset$fantasy_points_posOnly < 0)] <- 0
+player_stats_seasonal_offense_subset$fantasyPoints_posOnly <- player_stats_seasonal_offense_subset$fantasyPoints
+player_stats_seasonal_offense_subset$fantasyPoints_posOnly[which(player_stats_seasonal_offense_subset$fantasyPoints_posOnly < 0)] <- 0
 ```
@@
 bayesianMixedModelFormula <- brms::bf(
-  fantasy_points_posOnly ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(ageCentered20, player_idFactor, bs = "re") + (1 | player_idFactor)
+  fantasyPoints_posOnly ~ positionFactor + s(ageCentered20, by = positionFactor) + years_of_experience + s(ageCentered20, player_idFactor, bs = "re") + (1 | player_idFactor)
 )

### 2024-12-25 mythbusters.qmd
20241225 - convert to calculate_stats() from nflfastR v5.0

@@
 # Merge with weekly and seasonal stats data
-player_stats_weekly_offense <- player_stats_weekly_offense %>% 
+player_stats_weekly_offense <- player_stats_weekly %>% 
+  dplyr::filter(position_group %in% c("QB","RB","WR","TE")) %>% 
   dplyr::mutate(merge_name = nflreadr::clean_player_names(player_display_name, lowercase = TRUE))
@@
     receiving_epa = mean(receiving_epa, na.rm = TRUE),
+    fantasyPoints = sum(fantasyPoints, na.rm = TRUE), # season total
     contractYear = mean(contractYear, na.rm = TRUE),
@@
   dplyr::filter(position == "QB") %>% 
-  dplyr::select(merge_name, season, week, age, years_of_experience) %>% 
+  dplyr::select(merge_name, season, week, age, years_of_experience, fantasyPoints) %>% 
   full_join(
@@
     qb_plays = sum(qb_plays, na.rm = TRUE), # season total
+    fantasyPoints = sum(fantasyPoints, na.rm = TRUE), # season total
     contractYear = mean(contractYear, na.rm = TRUE),
@@
 ```{r}
-# Placeholder for model predicting fantasy points
+mixedModel_fantasyPtsPass <- lmerTest::lmer(
+  fantasyPoints ~ contractYear + (1 | player_id),
+  data = nfl_contractYearQBR_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModel_fantasyPtsPass)
+MuMIn::r.squaredGLMM(mixedModel_fantasyPtsPass)
+emmeans::emmeans(mixedModel_fantasyPtsPass, "contractYear")
+
+mixedModelAge_fantasyPtsPass <- lmerTest::lmer(
+  fantasyPoints ~ contractYear + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 | player_id), # removed random slopes to address convergence issue
+  data = nfl_contractYearQBR_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_fantasyPtsPass)
+MuMIn::r.squaredGLMM(mixedModelAge_fantasyPtsPass)
+emmeans::emmeans(mixedModelAge_fantasyPtsPass, "contractYear")
 ```
@@
 ```{r}
-# Placeholder for model predicting fantasy points
+mixedModel_fantasyPtsRush <- lmerTest::lmer(
+  fantasyPoints ~ contractYear + (1 | player_id),
+  data = player_statsContractsRB_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModel_fantasyPtsRush)
+MuMIn::r.squaredGLMM(mixedModel_fantasyPtsRush)
+emmeans::emmeans(mixedModel_fantasyPtsRush, "contractYear")
+
+mixedModelAge_fantasyPtsRush <- lmerTest::lmer(
+  fantasyPoints ~ contractYear + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
+  data = player_statsContractsRB_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_fantasyPtsRush)
+MuMIn::r.squaredGLMM(mixedModelAge_fantasyPtsRush)
+emmeans::emmeans(mixedModelAge_fantasyPtsRush, "contractYear")
 ```
@@
 ```{r}
-# Placeholder for model predicting fantasy points
+mixedModel_fantasyPtsReceiving <- lmerTest::lmer(
+  fantasyPoints ~ contractYear + (1 | player_id),
+  data = player_statsContractsWRTE_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModel_fantasyPtsReceiving)
+MuMIn::r.squaredGLMM(mixedModel_fantasyPtsReceiving)
+emmeans::emmeans(mixedModel_fantasyPtsReceiving, "contractYear")
+
+mixedModelAge_fantasyPtsReceiving <- lmerTest::lmer(
+  fantasyPoints ~ contractYear + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
+  data = player_statsContractsWRTE_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_fantasyPtsReceiving)
+MuMIn::r.squaredGLMM(mixedModelAge_fantasyPtsReceiving)
+emmeans::emmeans(mixedModelAge_fantasyPtsReceiving, "contractYear")
 ```

### 2024-12-25 research-methods.qmd
20241019 - copyediting
20241225 - convert to calculate_stats() from nflfastR v5.0

@@
 
-A research question is a question that the investigator (you!) want to know the answer to.
+A research question is a question that the investigator (you!) wants to know the answer to.
 For example, a research question might be: "Does consumption of sports drink improve player performance?"
@@
 
-*Incremental validity* evalutes the extent to which the measure's scores provide an increase in predictive accuracy compared to other information that is easily and cheaply available.
+*Incremental validity* evaluates the extent to which the measure's scores provide an increase in predictive accuracy compared to other information that is easily and cheaply available.
 That is, in order to be useful, a strong measure should tell us something that we did not already know.

### 2024-12-25 time-series-analysis.qmd
20241225 - convert to calculate_stats() from nflfastR v5.0

@@
 library("brms")
+library("rstan")
 library("plotly")
@@
 ```{r}
-weeklyFantasyPoints_tomBrady <- player_stats_weekly_offense %>% 
+weeklyFantasyPoints_tomBrady <- player_stats_weekly %>% 
   filter(
@@
 
-weeklyFantasyPoints_peytonManning <- player_stats_weekly_offense %>% 
+weeklyFantasyPoints_peytonManning <- player_stats_weekly %>% 
   filter(
@@
 ts_tomBrady <- xts::xts(
-  x = weeklyFantasyPoints_tomBrady["fantasy_points"],
+  x = weeklyFantasyPoints_tomBrady["fantasyPoints"],
   order.by = weeklyFantasyPoints_tomBrady$gameday)
@@
 ts_peytonManning <- xts::xts(
-  x = weeklyFantasyPoints_peytonManning["fantasy_points"],
+  x = weeklyFantasyPoints_peytonManning["fantasyPoints"],
   order.by = weeklyFantasyPoints_peytonManning$gameday)

### 2024-12-25 download-football-data.qmd
20241225 - fix error

@@
 nfl_actualFantasyPoints_player_seasonal <- nfl_actualFantasyPoints_player_seasonal_raw %>% 
-  rename(fantasyPoints = raw_points) %>% 
+  rename(
+    fantasyPoints = raw_points,
+    team = recent_team) %>% 
   select(player_id, season, everything()) %>% 

### 2024-12-25 mixed-models.qmd
20241225 - plot tweaks

@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Mixed Model with Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes"
+    subtitle = "Mixed Model with Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes",
+    color = "Position"
   ) +
@@
     title = "Fantasy Points (Season) by Player Age and Position",
-    subtitle = "Mixed Model with Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes\nin Interaction With Position"
+    subtitle = "Mixed Model with Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes\nin Interaction With Position",
+    color = "Position"
   ) +
@@
   ) +
-  theme_classic()
+  theme_classic() +
+  guides(color = guide_legend(override.aes = list(fill = NA))) # transparent legend background
 ```

### 2024-12-28 data-visualization.qmd
20241228 - fix error

@@
 # Subset Data
-rb_seasonal <- player_stats_seasonal_offense %>% 
+rb_seasonal <- player_stats_seasonal %>% 
   filter(position_group == "RB")
@@
   dplyr::left_join(
-    player_stats_seasonal_offense,
+    player_stats_seasonal,
     by = c("rusher_id" = "player_id", "season")

### 2024-12-29 mixed-models.qmd
20241229 - clean up code

@@
 ```{r}
-player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros <- player_stats_seasonal_offense_subset$fantasyPointsPos <- player_stats_seasonal_offense_subset$fantasyPoints
-player_stats_seasonal_offense_subset$fantasyPointsPos[player_stats_seasonal_offense_subset$fantasyPoints < 0] <- 0
-player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros[player_stats_seasonal_offense_subset$fantasyPoints <= 0] <- 0.01
+player_stats_seasonal_offense_subset$fantasyPoints_posOnlyNoZeros <- player_stats_seasonal_offense_subset$fantasyPoints_posOnly <- player_stats_seasonal_offense_subset$fantasyPoints
+player_stats_seasonal_offense_subset$fantasyPoints_posOnly[player_stats_seasonal_offense_subset$fantasyPoints < 0] <- 0
+player_stats_seasonal_offense_subset$fantasyPoints_posOnlyNoZeros[player_stats_seasonal_offense_subset$fantasyPoints <= 0] <- 0.01
 
@@
 
-fantasyPointsVectorPosNoZeros <- fantasyPointsVectorPos <- fantasyPointsVector
-fantasyPointsVectorPos[fantasyPointsVector < 0] <- 0
-fantasyPointsVectorPosNoZeros[fantasyPointsVector <= 0] <- 0.01
+fantasyPointsVector_posOnlyNoZeros <- fantasyPointsVector_posOnly <- fantasyPointsVector
+fantasyPointsVector_posOnly[fantasyPointsVector < 0] <- 0
+fantasyPointsVector_posOnlyNoZeros[fantasyPointsVector <= 0] <- 0.01
 ```
@@
 # positive-only
-fit.exp <- fitdist(fantasyPointsVectorPos, "exp")
+fit.exp <- fitdist(fantasyPointsVector_posOnly, "exp")
 
 # positive and no zeros
-fit.gamma <- fitdistrplus::fitdist(fantasyPointsVectorPosNoZeros, "gamma")
-fit.lognormal <- fitdistrplus::fitdist(fantasyPointsVectorPosNoZeros, "lnorm")
-fit.weibull <- fitdistrplus::fitdist(fantasyPointsVectorPosNoZeros, "weibull")
+fit.gamma <- fitdistrplus::fitdist(fantasyPointsVector_posOnlyNoZeros, "gamma")
+fit.lognormal <- fitdistrplus::fitdist(fantasyPointsVector_posOnlyNoZeros, "lnorm")
+fit.weibull <- fitdistrplus::fitdist(fantasyPointsVector_posOnlyNoZeros, "weibull")
 
@@
 model_gamma <- glm(
-  formula = fantasyPointsPosNoZeros ~ 1,
+  formula = fantasyPoints_posOnlyNoZeros ~ 1,
   family = Gamma(),
@@
 Because the response distribution (i.e., fantasy points) is positively skewed, we will use a gamma response distribution.
-
-#### Prepare Data {#sec-fantasyPointsByAgeExperienceModelsBayesianPrepareData}
-
-```{r}
-player_stats_seasonal_offense_subset$fantasyPoints_posOnly <- player_stats_seasonal_offense_subset$fantasyPoints
-player_stats_seasonal_offense_subset$fantasyPoints_posOnly[which(player_stats_seasonal_offense_subset$fantasyPoints_posOnly < 0)] <- 0
-```
 

### 2025-01-01 download_weekly_projections.R
20250101 - week 18 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2024
-weekNumber <- 16
+weekNumber <- 18
 
 # Processing ----
 season <- as.character(season)

### 2025-01-08 download_weekly_projections.R
20250108 - minor edits

@@ -2,8 +2,8 @@
 
 # Specify Season and Week Number ----
 
-season <- 2024
-weekNumber <- 18
+season <- 2025
+weekNumber <- 1
 
 # Processing ----
 season <- as.character(season)

### 2025-01-12 getting-started.qmd
20241016 - render the figure caption as alt text
20250112 - archived link

@@
 
-The following article provides additional resources and good guidance: <https://www.r-bloggers.com/where-to-get-help-with-your-r-question/>.
+The following article provides additional resources and good guidance: <https://www.r-bloggers.com/where-to-get-help-with-your-r-question/> (archived at <https://perma.cc/4RRE-KL33>).
 
@@
 
-![Types of merges/joins](images/joinTypes.png)
+![Types of merges/joins](images/joinTypes.png)\
 

### 2025-01-17 index.qmd
20250108 - minor edits
20250117 - 2025 citation

@@
 T.
-(2024).
-*Fantasy football analytics: Statistics, prediction, and empiricism Using R*.
+(2025).
+*Fantasy football analytics: Statistics, prediction, and empiricism using R*.
 Version 0.0.1.
@@
   author = {Petersen, Isaac T.},
-  year = {2024},
+  year = {2025},
   publisher = {{University of Iowa Libraries}},

### 2025-02-07 mixed-models.qmd
20241229 - bootstrap Cullen and Frey graph
20250207 - challenges of trajectories

@@
 
+In this chapter, we use mixed models to demonstrate how to model trajectories (growth curves) of players' fantasy points as a function of age.
+One of the challenges when modeling players' fantasy points as a function of age is that fantasy points are a function of both ability and opportunity.
+With age, the player's ability and opportunity may both decline, but it is difficult to disentangle the extent to which a player's decline is related to declines in ability versus opportunity.
+As players get older, they may be supplanted by younger, more talented players.
+With age, despite still having latent (unobserved) ability, players will get fewer and, eventually, no opportunity.
+And we do not know the counterfactual—how many fantasy points they would have scored had they been given the full opportunity each season.
+Thus, players may go from scoring 100+ points in a season to all of a sudden scoring way fewer points, with little in the way of intermediate steps.
+Thus, this should inform how you interpret the resulting trajectories.
+For instance, the steep declines can make it look like the model-implied fantasy points go well below zero at later ages, even though it is obviously not likely for players to obtain such a magnitude of negative fantasy points.
+These negative model-implied values of fantasy points at later ages is likely an artifact of the decreasing opportunity that players tend to get as they age.
+
 ```{r}
@@
 
-fitdistrplus::descdist(fantasyPointsVector)
+fitdistrplus::descdist(
+  fantasyPointsVector,
+  boot = 1000)
 ```

### 2025-02-07 mixed-models.qmd
20250207 - copyedit

@@
 For instance, the steep declines can make it look like the model-implied fantasy points go well below zero at later ages, even though it is obviously not likely for players to obtain such a magnitude of negative fantasy points.
-These negative model-implied values of fantasy points at later ages is likely an artifact of the decreasing opportunity that players tend to get as they age.
+The negative model-implied values of fantasy points at later ages are likely an artifact of the decreasing opportunity that players tend to get as they age.
 

### 2025-02-07 mixed-models.qmd
20250207 - clarifications

@@
 The negative model-implied values of fantasy points at later ages are likely an artifact of the decreasing opportunity that players tend to get as they age.
+The crossover point, where the positive values becomes negative might indicate, for instance, that players at that age tend to have retired and thus have no opportunity.
+Regardless, the form/shape/steepness of the decline is still potentially meaningful even if the precise negative number at a given age is not.
 
@@
 The model-implied peak of Quarterbacks' fantasy points is at age `{r} qbPeakAge`.
-The model-predicted value of zero fantasy points for Quarterbacks is at `{r} qbZeroAge`.
+The model-predicted value of zero fantasy points for Quarterbacks is at age `{r} qbZeroAge`.
 
@@
 The model-implied peak of Fullbacks' fantasy points is at age `{r} fbPeakAge`.
-The model-predicted value of zero fantasy points for Fullbacks is at `{r} fbZeroAge`.
+The model-predicted value of zero fantasy points for Fullbacks is at age `{r} fbZeroAge`.
 
@@
 The model-implied peak of Running Backs' fantasy points is at age `{r} rbPeakAge`.
-The model-predicted value of zero fantasy points for Running Backs is at `{r} rbZeroAge`.
+The model-predicted value of zero fantasy points for Running Backs is at age `{r} rbZeroAge`.
 
@@
 The model-implied peaks of Wide Receivers' fantasy points are at ages `{r} wrPeakAge` and `{r} wrPeakAge2`.
-The model-predicted value of zero fantasy points for Wide Receivers is at `{r} wrZeroAge`.
+The model-predicted value of zero fantasy points for Wide Receivers is at age `{r} wrZeroAge`.
 
@@
 The model-implied peak of Tight Ends' fantasy points is at age `{r} tePeakAge`.
-The model-predicted value of zero fantasy points for Tight Ends is at `{r} teZeroAge`.
+The model-predicted value of zero fantasy points for Tight Ends is at age `{r} teZeroAge`.
 

### 2025-02-08 mixed-models.qmd
20250208 - clean up

@@
 library("brms")
-library("cmdstanr") # todo: install.packages("cmdstanr", repos = c('https://stan-dev.r-universe.dev', getOption("repos"))); check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
+library("cmdstanr") # todo: install.packages("cmdstanr", repos = c("https://stan-dev.r-universe.dev", getOption("repos"))); check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
 library("fitdistrplus")

### 2025-02-09 data-visualization.qmd
20250208 - zoom in; don't remove data from plot

@@
   geom_smooth() +
+  coord_cartesian(
+    xlim = c(20,40), # set limits of x-axis
+    ylim = c(0,NA), # set limits of y-axis
+    expand = FALSE) + # don't add space between axes and data
   scale_x_continuous(
-    expand = c(0,0), # set origin of x-axis to 0
-    lim = c(20,40), # set limits of x-axis
     breaks = seq(from = 20, to = 40, by = 5) # specify x-axis labels
@@
   scale_y_continuous(
-    expand = c(0,0), # set origin of y-axis to 0
-    lim = c(0,NA), # set limits of y-axis
     breaks = seq(from = 0, to = 2500, by = 250) # specify y-axis labels
@@
   geom_smooth() +
+  coord_cartesian(
+    xlim = c(20,40),
+    ylim = c(0,NA),
+    expand = FALSE) +
   scale_x_continuous(
-    expand = c(0,0),
-    lim = c(20,40),
     breaks = seq(from = 20, to = 40, by = 5) 
@@
   scale_y_continuous(
-    expand = c(0,0),
-    lim = c(0,NA),
     breaks = seq(from = 0, to = 2500, by = 250)

### 2025-02-09 data-visualization.qmd
20250208 - marginal density/histogram
20250208 - specify package

@@
 library("ggridges")
+library("ggExtra")
 library("tidyverse")
@@
 
+### Marginal Density/Histogram {#sec-scatterplotMarginalDensigram}
+
+```{r}
+ggExtra::ggMarginal(
+  plot_ypcByPlayerAge,
+  type = "densigram")
+```
+
 ## Line Chart {#sec-lineChart}

### 2025-02-09 data-visualization.qmd
20250208 - figure caption

@@
 
-```{r}
+We can also add a marginal density/histogram, as in @fig-scatterplot12:
+
+```{r}
+#| label: fig-scatterplot12
+#| fig-cap: "Scatterplot With Marginal Density/Histogram."
+#| fig-alt: "Scatterplot With Marginal Density/Histogram."
+
 ggExtra::ggMarginal(

### 2025-02-09 data-visualization.qmd
20250209 - histogram with overlaid density plot

@@
     x = "Fantasy Points",
+    y = "Count",
     title = "Histogram of Fantasy Points"
@@
     x = "Fantasy Points",
+    y = "Density",
     fill = "Position",
@@
 
-### Histogram with Overlaid Density and Rug Plot {#sec-histogramDensity}
-
-A histogram of fantasy points with an overlaid density and rug plot is depicted in @fig-histogramDensity.
+### Density Plot with Histogram and Rug Plot {#sec-densityHistogram}
+
+A density plot of fantasy points with a histogram and rug plot is depicted in @fig-histogramDensity.
 
@@
     x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Fantasy Points with Histogram and Rug Plot"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+### Histogram with Overlaid Density and Rug Plot {#sec-histogramDensity}
+
+A histogram of fantasy points with an overlaid density and rug plot is depicted in @fig-histogramDensity.
+
+```{r}
+#| label: fig-histogramDensity
+#| fig-cap: "Histogram with Overlaid Density and Rug Plot."
+#| fig-alt: "Histogram with Overlaid Density and Rug Plot."
+
+binWidth <- 15
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasyPoints)
+) +
+  geom_histogram(
+    color = "#000000",
+    fill = "#0099F8",
+    binwidth = binWidth
+  ) +
+  geom_density(
+    aes(y = after_stat(count) * binWidth),
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  geom_rug() +
+  labs(
+    x = "Fantasy Points",
+    y = "Count",
     title = "Histogram of Fantasy Points with Overlaid Density and Rug Plot"

### 2025-02-09 data-visualization.qmd
20250209 - density plot with histogram
20250209 - use coord_cartesian to zoom in plot w/o removing data

@@
 #| label: fig-histogramDensity
-#| fig-cap: "Histogram with Overlaid Density and Rug Plot."
-#| fig-alt: "Histogram with Overlaid Density and Rug Plot."
+#| fig-cap: "Density Plot with Histogram and Rug Plot."
+#| fig-alt: "Density Plot with Histogram and Rug Plot."
 
@@
   geom_smooth() +
-  scale_x_continuous(
-    expand = c(0,0)
-  ) +
-  scale_y_continuous(
-    expand = c(0,0),
-    lim = c(0,NA)
-  ) +
+  coord_cartesian(
+    ylim = c(0,NA), # set limits of y-axis
+    expand = FALSE) + # don't add space between axes and data
   labs(

### 2025-02-09 data-visualization.qmd
20250209 - fix figure labels

@@
 
-A density plot of fantasy points with a histogram and rug plot is depicted in @fig-histogramDensity.
-
-```{r}
-#| label: fig-histogramDensity
+A density plot of fantasy points with a histogram and rug plot is depicted in @fig-densityHistogram.
+
+```{r}
+#| label: fig-densityHistogram
 #| fig-cap: "Density Plot with Histogram and Rug Plot."

### 2025-02-09 decision-making.qmd
20250209 - sports betting

@@
 However, luck is unlikely to continue systematically into future years, so a source that got lucky in a given year is likely to [regress to the mean](#sec-fallaciesRegression) in subsequent years.
-That is, determinining the most accurate source in a given year, after the fact, is not necessarily the same as identifying the most skilled forecaster.
+That is, determining the most accurate source in a given year, after the fact, is not necessarily the same as identifying the most skilled forecaster.
 It is easy to identify the most accurate source after the fact, but it is challenging to predict, in advance, who the best forecaster will be [@Larrick2024].
@@
 In the context of fantasy football, advice might be sources of projections.
-In general, evidence from social psychology suggests that people tend to underweight how much weight they give to advice from others relative to their own opinions, a phenomeon called egocentric discounting [@Larrick2024; @Rader2017].
+In general, evidence from social psychology suggests that people tend to underweight how much weight they give to advice from others relative to their own opinions, a phenomenon called egocentric discounting [@Larrick2024; @Rader2017].
 People tend to weight others' advice around 30% in terms of the proportion of a shift a person makes toward another person's perspective, though this depends on the perceived accuracy of the advisor.
@@
 
+## Strategies for Managing Risk and Uncertainty {#sec-managinRiskUncertainty}
+
+## Risk Management Principles from Cognitive Psychology {#sec-riskManagementCognitivePsych}
+
+## Sports Betting/Gambling {#sec-sportsBetting}
+
+[overconfidence; confirmation bias; variable ratio reinforcement; addiction/delay discounting; wisdom of the crowd, and the challenge of beating "the market"]
+
+There is substantial [overconfidence](#sec-cognitiveBiasesOverconfidence) (in particular overestimation) in sports betting/gambling.
+[A study](https://mattbrownecon.github.io/assets/papers/jmp/sportsbetting.pdf) (archived at <https://perma.cc/X2AW-SUBZ>) of frequent sports bettors found that they tended to predict that they would gain 0.3 cents for every dollar wagered, but in fact lost 7.5 cents for every dollar wagered [@Brown2025].
+[Overconfidence](#sec-cognitiveBiasesOverconfidence) was greatest among those who frequently wagered multi-leg bets (parlays), losing ~25 cents for every dollar wagered [@Brown2025].
+
 ## Suggestions {#sec-decisionMakingSuggestions}
 
-Based on the above dicussion, here are some suggestions for decision making in the context of uncertainty:
+Based on the above discussion, here are some suggestions for decision making in the context of uncertainty:
 

### 2025-02-11 player-evaluation.qmd
20250210 - embrace the chaos

@@
 ```
+
+However, another perspective is to embrace the chaos and uncertainty that is part of fantasy football; one is better positioned to win a given week if you have one or a few players that "go off"—i.e., that score lots of points.
+Thus, starting boom-or-bust players can make it more likely that you win in the "boom weeks"; however, this may also mean having some "bust weeks".
 

### 2025-02-15 causal-inference.qmd
20250214 - control variables

@@
 Thus, another approach that can help increase [internal validity](#sec-internalValidity) is to include plausible [confounds](#sec-causalDiagramConfounding) as control variables.
-For instance, if a third variable such as education level might be a [confound](#sec-causalDiagramConfounding) that influences both sports drink consumption and player performance, you could include education level as a [covariate](#sec-covariates) in the model.
+For instance, if a third variable such as education level might be a [confound](#sec-causalDiagramConfounding) that influences both sports drink consumption and player performance, you could include control for a player's education level.
+
+There are several ways to control for a variable:
+
+1. randomization (when possible)
+1. restriction
+1. matching
+1. covariate
+1. stratification
+
+One way to control for a variable—and what is typically considered the most rigorous control—when possible, is to randomize.
+For example, you could randomly assign players to drink sports drink during a game versus water.
+By randomly assigning players to each group, it is expected that, if the groups are large enough, the differences between the groups on all variables besides the [independent](#sec-experiment) and [dependent](#sec-experiment) variables will be approximately equal across both group; thus, any difference in the [dependent variable](#sec-experiment) is thought to be attributable to the manipulation of the [independent variable](#sec-experiment).
+However, sometimes a variable is not able to be manipulated for practical or ethical reasons.
+For instance, if you wanted to control for education level when examining the association between sports drink consumption and performance, you cannot randomly assign players to have lower educational levels (though you could possibly provide an intervention to increase the education levels of a subset of players).
+
+A second way to control for a variable is restriction—e.g., you can select your sample so that it only includes portions/subgroups of a variable.
+For example, you could recruit—as participants—only those players who are college graduates.
+
+Matching means that you match people based on the control variable.
+If you some people consume sports drink whereas others do not, matching would mean finding—for every player included in the study who consumes sports drink—an equivalent player in terms of education who does not consume sports drink.
+
+Another way to control for a variable is to include the variable as a [covariate](#sec-covariates) in the model.
 Inclusion of a [covariate](#sec-covariates) attempts to control for the variable by examining the association between the [predictor variable](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy) while holding the [covariate](#sec-covariates) variables constant.
-For instance, such a model would examine whether, when accounting for education level, there is an association between sports drink consumption and player performance.
+For example, if you want to control for a player's education level in examining the association between sports drink consumption and performance, you could statistically control for player education by including the player's education level as a [covariate](#sec-covariates) in predicting performance.
+Such a model would examine whether, when accounting for education level, there is an association between sports drink consumption and player performance.
 
+Another way to control for a variable is to stratify based on the variable—e.g., you can examine the association between the variables of interest within subgroups of the control variable.
+For instance, taking the example of controlling for a player's education in examining the association between sports drink consumption and performance, you could examine the association between sports drink consumption and performance separately for various education levels: a) high school or lower, b) some college, c) college degree, d) graduate degree.
+
+In each of these examples, you are controlling for player education for purposes of examining the effect of sports drink consumption on performance.
 Failure to control for important third variables can lead to erroneous conclusions, as evidenced by the association depicted in @fig-simpsonParadox.
@@
 DAGs provide a helpful tool to communicate about causal questions and help identify how to avoid bias (i.e., overestimation) in associations between variables due to [confounding](#sec-causalDiagramConfounding) (i.e., common causes) [@Digitale2022].
-For instance, from a DAG, it is possible to determine what variables it is important to control for in order to get unbiased estimates of the association between two variables of interest.
+For instance, from a DAG, it is possible to determine what variables it is important to control for (and how)—and just as importantly, what not to control for—in order to get unbiased estimates of the effect of one variable on another variable or of the association between two variables of interest.
 To create DAGs, you can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: <https://dagitty.net> (archived at <https://perma.cc/U9BY-VZE2>).
@@
 
-The `dagitty::adjustmentSets()` function identifies variables that would be necessary to control for (i.e., to include as covariates) in order to identify an unbiased estimate of the association (whether the total effect, i.e., `effect = "total"`; or the direct effect, i.e., `effect = "direct"`) between two variables (`exposure` and `outcome`).
+The `dagitty::adjustmentSets()` function identifies variables that would be necessary to control for in order to identify an unbiased estimate of the association (whether the total effect, i.e., `effect = "total"`; or the direct effect, i.e., `effect = "direct"`) between two variables (`exposure` and `outcome`).
 In this case, to identify the unbiased association between `M1` and `Y`, it is important to control for `M2.`

### 2025-02-15 research-methods.qmd
20250214 - control variables

@@
 The variable that we manipulate is the *independent variable*.
-By contrast, the *dependent variable* is the variable that we evaluate to determine whether it was influenced by the manipulation (i.e., by the independent variable).Besides the independent and dependent variables, the researcher attempts to hold everything else constant through processes including standardization and random assignment.
+By contrast, the *dependent variable* is the variable that we evaluate to determine whether it was influenced by the manipulation (i.e., by the independent variable).
+Besides the independent and dependent variables, the researcher attempts to hold everything else constant through processes including standardization and random assignment.
 *Standardization* involves using the same procedures to assess each participant, so that scores can be fairly compared across participants (and groups).

### 2025-02-15 causal-inference.qmd
20250214 - copyediting

@@
 
-One way to control for a variable—and what is typically considered the most rigorous control—when possible, is to randomize.
-For example, you could randomly assign players to drink sports drink during a game versus water.
-By randomly assigning players to each group, it is expected that, if the groups are large enough, the differences between the groups on all variables besides the [independent](#sec-experiment) and [dependent](#sec-experiment) variables will be approximately equal across both group; thus, any difference in the [dependent variable](#sec-experiment) is thought to be attributable to the manipulation of the [independent variable](#sec-experiment).
+One way to control for a variable—and what is typically considered the most rigorous control—when possible, is to randomize (which would make it an [experiment](#sec-causalInferenceExperiment)).
+For example, you could randomly assign players to consume sports drink during a game versus water.
+By randomly assigning players to each group, it is expected that, if the groups are large enough, the differences between the groups on all variables besides the [independent](#sec-experiment) and [dependent](#sec-experiment) variables will be approximately equal across both groups; thus, any difference in the [dependent variable](#sec-experiment) is thought to be attributable to the manipulation of the [independent variable](#sec-experiment).
 However, sometimes a variable is not able to be manipulated for practical or ethical reasons.
@@
 Matching means that you match people based on the control variable.
-If you some people consume sports drink whereas others do not, matching would mean finding—for every player included in the study who consumes sports drink—an equivalent player in terms of education who does not consume sports drink.
+If some people consume sports drink whereas others do not, matching would mean finding—for every player included in the study who consumes sports drink—an equivalent player in terms of education who does not consume sports drink.
 

### 2025-02-20 causal-inference.qmd
20250215 - counterfactual
20250219 - causal inference

@@
 ```
+
+### Causation
+
+A causal effect is an effect in which the level of one input variable results in different levels on the outcome variable.
+For instance, for a given player, the extent to which consumption of sports drink influences their performance reflects the difference between how well the player would perform if they consume sports drink compared to how well they would perform if they do not consume sports drink.
 
@@
 
-In general, to meet the third criterion of causality, one must consider the counterfactual.
+In general, to meet the third criterion of causality, one must consider the counterfactual—what would have happened "counter to the fact".
 A *counterfactual* is what would have happened in the hypothetical scenario that the cause did not occur [i.e., what would have happened in the absence of the cause; @Shadish2002].
 When engaging in causal inference, it is important to consider what would have happened if the hypothetical cause had actually not occurred.
+It is like an alternate universe where the cause did not happen.
 For instance, consider that we conduct an experiment to randomly assign some players to consume a sports drink before a game and other players to drink only water.
@@
 However, we cannot observe a counterfactual.
+This is the fundamental problem of causal inference—we only observe one potential world/outcome/universe.
 That is, we do not know for sure what would have happened to the players who received treatment if those same players had actually not received treatment.
@@
 Thus, another approach that can help increase [internal validity](#sec-internalValidity) is to include plausible [confounds](#sec-causalDiagramConfounding) as control variables.
-For instance, if a third variable such as education level might be a [confound](#sec-causalDiagramConfounding) that influences both sports drink consumption and player performance, you could include control for a player's education level.
+For instance, if a third variable such as education level might be a [confound](#sec-causalDiagramConfounding) that influences both sports drink consumption and player performance, you could control for a player's education level.
 
@@
 For example, you could randomly assign players to consume sports drink during a game versus water.
-By randomly assigning players to each group, it is expected that, if the groups are large enough, the differences between the groups on all variables besides the [independent](#sec-experiment) and [dependent](#sec-experiment) variables will be approximately equal across both groups; thus, any difference in the [dependent variable](#sec-experiment) is thought to be attributable to the manipulation of the [independent variable](#sec-experiment).
+By randomly assigning players to each group, it is expected that, if the groups are large enough, the differences between the groups on all variables besides the [independent](#sec-experiment) and [dependent](#sec-experiment) variables will be approximately equal across both groups; thus, such an approach controls for both observed and unobserved [confounds](#sec-causalDiagramConfounding), and any difference in the [dependent variable](#sec-experiment) is thought to be attributable to the manipulation of the [independent variable](#sec-experiment).
 However, sometimes a variable is not able to be manipulated for practical or ethical reasons.
 For instance, if you wanted to control for education level when examining the association between sports drink consumption and performance, you cannot randomly assign players to have lower educational levels (though you could possibly provide an intervention to increase the education levels of a subset of players).
+Thus, observational studies are also necessary.
+
+However, in observational studies, the groups may differ on the [confounding factors](#sec-causalDiagramConfounding); people may choose the level of the they receive.
+Thus, we need to attempt to control for the [confounds](#sec-causalDiagramConfounding); however, in an observational study, we can only potentially adjust for the observed [confounds](#sec-causalDiagramConfounding), unlike in an [experiment](#sec-causalInferenceExperiment) in which we can control for both observed and unobserved [confounds](#sec-causalDiagramConfounding).
 
@@
 The use of [causal diagrams](#sec-causalDiagrams) can inform which variables are important to be included as control variables, and—just as important—which variables not to include as control variables, as described in @sec-causalDiagrams.
+Another approach that is sometimes used for matching, as a sample weight, or with stratification is the use of propensity scores.
 
@@
 
-Selection bias occurs when the selection of participants or their inclusion in analyses depends on the variables being studied.
+Selection bias occurs when the selection of participants or their inclusion in analyses depends on the variables being studied (or on their [colliders](#sec-colliderBias)).
 For instance, if you are conducting a study on the extent to which sports drink consumption influences fantasy points, there would be selection bias if players are less likely to participate in the study if they score fewer fantasy points.

### 2025-02-26 sports-cognitive-psychology.qmd
20250226 - clean up code

@@
 
-Adapted from Ben Baldwin: https://www.nfl4th.com/articles/4th-down-research.html
+Adapted from Ben Baldwin: https://www.nfl4th.com/articles/4th-down-research.html (archived at <https://perma.cc/S5D8-3NCU>)
 
@@
   filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2014, 2024)) %>%
-  ggplot(aes(go_boost, go, color = as.factor(season))) + 
+  ggplot(
+    aes(
+      x = go_boost,
+      y = go,
+      color = as.factor(season))) + 
   ggtext::geom_richtext(
     data = text_df,   
-    aes(x, y, label = label, angle = angle), 
+    aes(
+      x,
+      y,
+      label = label,
+      angle = angle), 
     color = "black",
@@
   geom_vline(xintercept = 0) +
-  stat_smooth(method = "gam", method.args = list(gamma = 1), formula = y ~ s(x, bs = "cs", k = 10), show.legend = F, se = F, linewidth = 4) +
+  stat_smooth(
+    method = "gam",
+    method.args = list(gamma = 1),
+    formula = y ~ s(x, bs = "cs", k = 10),
+    show.legend = FALSE,
+    se = FALSE,
+    linewidth = 4) +
   # this is just to get the plot to draw the full 0 to 100 range
-  geom_hline(yintercept = 100, alpha = 0) +
-  geom_hline(yintercept = 0, alpha = 0) +
+  geom_hline(
+    yintercept = 100,
+    alpha = 0) +
+  geom_hline(
+    yintercept = 0,
+    alpha = 0) +
   ggplot2::theme_classic() +
-  labs(x = "Gain in win probability by going for it",
+  labs(
+    x = "Gain in win probability by going for it",
     y = "Go-for-it percentage",
@@
   ) +
-  scale_y_continuous(breaks = scales::pretty_breaks(n = 4), limits = c(0, 100), expand = c(0,0)) +
-  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(-10, 10), expand = c(0,0)) +
-  annotate("text", x= -1.2, y= 70, label = "Should\nkick", color="black", size = 5) +
-  annotate("text", x= 1.2, y= 70, label = "Should\ngo for it", color="black", size = 5) +
+  scale_y_continuous(
+    breaks = scales::pretty_breaks(n = 4),
+    limits = c(0, 100),
+    expand = c(0,0)) +
+  scale_x_continuous(
+    breaks = scales::pretty_breaks(n = 10),
+    limits = c(-10, 10),
+    expand = c(0,0)) +
+  annotate(
+    "text",
+    x = -1.2,
+    y = 70,
+    label = "Should\nkick",
+    color = "black",
+    size = 5) +
+  annotate(
+    "text",
+    x = 1.2,
+    y = 70,
+    label = "Should\ngo for it",
+    color = "black",
+    size = 5) +
   geom_segment(
-    aes(x = -.1, y = 80, xend = -2, yend = 80),
+    aes(
+      x = -.1,
+      y = 80,
+      xend = -2,
+      yend = 80),
     arrow = arrow(length = unit(0.05, "npc")),
-    color = "black", linewidth = 2
-    ) +
+    color = "black",
+    linewidth = 2) +
   geom_segment(
-    aes(x = .1, y = 80, xend = 2, yend = 80),
+    aes(
+      x = .1,
+      y = 80,
+      xend = 2,
+      yend = 80),
     arrow = arrow(length = unit(0.05, "npc")),
-    color = "black", linewidth = 2
-  )
+    color = "black",
+    linewidth = 2)
 ```

### 2025-02-26 pca.qmd
Built site for gh-pages
20250226 - PCA

@@
 # Data Reduction: Principal Component Analysis {#sec-pca}
+
+## Getting Started {#sec-pcaGettingStarted}
+
+### Load Packages {#sec-pcaLoadPackages}
+
+```{r}
+
+```
+
+## Overview of Principal Component Analysis {#sec-pcaOverview}
+
+Principal component analysis (PCA) is used if you want to reduce your data matrix.\index{principal component analysis}
+PCA composites represent the variances of an observed measure in as economical a fashion as possible, with no latent underlying variables.\index{principal component analysis}
+The goal of PCA is to identify a smaller number of components that explain as much variance in a set of variables as possible.\index{principal component analysis}\index{data!reduction}
+It is an atheoretical way to decompose a matrix.\index{principal component analysis}
+PCA involves decomposition of a data matrix into a set of eigenvectors, which are transformations of the old variables.\index{principal component analysis}\index{eigenvector}
+
+The eigenvectors attempt to simplify the data in the matrix.\index{parsimony}\index{principal component analysis}\index{eigenvector}
+PCA takes the data matrix and identifies the weighted sum of all variables that does the best job at explaining variance: these are the principal components, also called eigenvectors.\index{principal component analysis}\index{eigenvector}
+Principal components reflect optimally weighted sums.\index{principal component analysis}
+
+PCA decomposes the data matrix into any number of components—as many as the number of variables, which will always account for all variance.\index{principal component analysis}
+After the PCA is performed, you can look at the results and discard the components which likely reflect error variance.\index{principal component analysis}
+Judgments about which components to retain are based on empirical criteria in conjunction with theory to select a parsimonious number of components that account for the majority of variance.\index{principal component analysis}
+
+The eigenvalue reflects the amount of variance explained by the component (eigenvector).\index{principal component analysis}\index{eigenvalue}
+When using a varimax (orthogonal) rotation, an eigenvalue for a component is calculated as the sum of squared standardized component loadings on that component.\index{principal component analysis}\index{orthogonal rotation}\index{eigenvalue}
+When using oblique rotation, however, the items explain more variance than is attributable to their factor loadings because the factors are correlated.\index{principal component analysis}\index{oblique rotation}
+
+PCA pulls the first principal component out (i.e., the eigenvector that explains the most variance) and makes a new data matrix: i.e., new correlation matrix.\index{principal component analysis}\index{eigenvector}
+Then the PCA pulls out the component that explains the next most variance—i.e., the eigenvector with the next largest eigenvalue, and it does this for all components, equal to the same number of variables.\index{principal component analysis}\index{eigenvector}\index{eigenvalue}
+For instance, if there are six variables, it will iteratively extract an additional component up to six components.\index{principal component analysis}
+You can extract as many eigenvectors as there are variables.\index{principal component analysis}\index{eigenvector}
+If you extract all six components, the data matrix left over will be the same as the correlation matrix in Figure \@ref(fig:correlationMatrix2).\index{principal component analysis}
+That is, the remaining variables (as part of the leftover data matrix) will be entirely uncorrelated with each other, because six components explain 100% of the variance from six variables.\index{principal component analysis}
+In other words, you can explain (6) variables with (6) new things!\index{principal component analysis}
+
+However, it does no good if you have to use all (6) components because there is no data reduction from the original number of variables.\index{principal component analysis}\index{data!reduction}
+When the goal is data reducation (as in PCA), the hope is that the first few components will explain most of the variance, so we can explain the variability in the data with fewer components than there are variables.\index{principal component analysis}\index{data!reduction}
+
+The sum of all eigenvalues is equal to the number of variables in the analysis.\index{principal component analysis}\index{eigenvalue}
+PCA does not have the same assumptions as [factor analysis](#sec-factorAnalysis), which assumes that measures are partly from common variance and error.\index{principal component analysis}\index{factor analysis}\index{measurement error}
+But if you estimate (6) eigenvectors and only keep (2), the model is a two-component model and whatever left becomes error.\index{principal component analysis}\index{measurement error}\index{eigenvector}
+Therefore, PCA does not have the same assumptions as [factor analysis](#sec-factorAnalysis), but it often ends up in the same place.\index{principal component analysis}\index{factor analysis}
+
+## Component Rotation {#sec-pcaRotation}
+
+Similar considerations as in [factor analysis](#sec-factorAnalysis) can be used to determine whether and how to rotate components in PCA.
+The considerations for determining whether and how to rotate factors in [factor analysis](#sec-factorAnalysis) are described in @sec-factorAnalysisRotation.
+
+## Determining the Number of Components to Retain {#sec-pcaNumComponents}
+
+Similar criteria as in [factor analysis](#sec-factorAnalysis) can be used to determine the number of components to retain in PCA.
+The criteria for determining the number of factors to retain in [factor analysis](#sec-factorAnalysis) are described in @sec-factorAnalysisNumFactors.
+
+## Interpreting and Using PCA Components {#sec-interpretingAndUsingComponentsPCA}
+
+The next step is interpreting the PCA components.\index{principal component analysis}
+Use theory to interpret and label the components.\index{principal component analysis}
+
+## PCA Versus Factor Analysis {#sec-pcaVsFactorAnalysis}
+
+The key distinction between [factor analysis](#sec-factorAnalysis) and PCA is depicted in @fig-pcaVsFactorAnalysis.\index{factor analysis}\index{principal component analysis}
+
+::: {#fig-pcaVsFactorAnalysis}
+![](images/pcaVsFactorAnalysis.png){fig-alt="Distinction Between Factor Analysis and Principal Component Analysis."}
+
+Distinction Between Factor Analysis and Principal Component Analysis.
+:::
+
+There are several differences between [factor analysis](#sec-factorAnalysis) and PCA.\index{factor analysis}\index{principal component analysis}
+[Factor analysis](#sec-factorAnalysis) has greater sophistication than PCA, but greater sophistication often results in greater assumptions.\index{factor analysis}\index{principal component analysis}
+[Factor analysis](#sec-factorAnalysis) does not always work; the data may not always fit to a [factor analysis](#sec-factorAnalysis) model.\index{factor analysis}\index{principal component analysis}
+However, PCA can decompose any data matrix; it always works.\index{principal component analysis}
+PCA is okay if you are not interested in the factor structure.\index{principal component analysis}
+PCA uses all variance of variables and assumes variables have no error, so it does not account for measurement error.\index{principal component analysis}\index{measurement error}
+PCA is good if you just want to form a linear composite and perform data reduction.\index{principal component analysis}\index{linear composite}\index{construct!formative}\index{construct!reflective}
+However, if you are interested in the factor structure, use [factor analysis](#sec-factorAnalysis), which estimates a latent variable that accounts for the common variance and discards error variance.\index{factor analysis}
+[Factor analysis](#sec-factorAnalysis) better handles error than PCA—[factor analysis](#sec-factorAnalysis) assumes that what is in the variable is the combination of common construct variance and error.\index{principal component analysis}\index{factor analysis}
+By contrast, PCA assumes that the measures have no measurement error.\index{principal component analysis}\index{measurement error}
+[Factor analysis](#sec-factorAnalysis) is useful for the identification of latent constructs—i.e., underlying dimensions or factors that explain (cause) observed scores.\index{factor analysis}
+
+## Conclusion {#sec-pcaConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-pcaSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2025-02-26 factor-analysis.qmd
20240605 - getting started/load packages
Built site for gh-pages
20240703 - conclusion section
20250226 - factor analysis

--- 2024-06-05 base-rates.qmd
@@
-# Base Rates {#sec-baseRates}
+# Factor Analysis {#sec-factorAnalysis}
 
-## Getting Started
+## Getting Started {#sec-factorAnalysisGettingStarted}
 
-### Load Packages
+### Load Packages {#sec-factorAnalysisLoadPackages}
 
@@
 ```
+
+## Overview of Factor Analysis {#sec-factorAnalysisOverview}
+
+Factor analysis involves the estimation of latent variables.
+Latent variables are ways of studying and operationalizing theoretical constructs that cannot be directly observed or quantified.
+Factor analysis is a class of latent variable models that is designated to identify the structure of a measure or set of measures, and ideally, a construct or set of constructs.\index{factor analysis}
+It aims to identify the optimal latent structure for a group of variables.\index{factor analysis}
+The goal of factor analysis is to identify simple, parsimonious factors that underlie the "junk" (i.e., scores filled with measurement error) that we observe.\index{parsimony}\index{factor analysis!confirmatory}
+
+Factor analysis encompasses two general types: [confirmatory factor analysis](#sec-cfa) and [exploratory factor analysis](#sec-efa).\index{factor analysis}\index{factor analysis!confirmatory}\index{factor analysis!exploratory}
+[*Exploratory factor analysis*](#sec-efa) (EFA) is a latent variable modeling approach that is used when the researcher has no a priori hypotheses about how a set of variables is structured.\index{factor analysis!exploratory}
+[EFA](#sec-efa) seeks to identify the empirically optimal-fitting model in ways that balance accuracy (i.e., variance accounted for) and parsimony (i.e., simplicity).\index{parsimony}\index{factor analysis!exploratory}\index{latent variable}
+[*Confirmatory factor analysis*](#sec-cfa) (CFA) is a latent variable modeling approach that is used when a researcher wants to evaluate how well a hypothesized model fits, and the model can be examined in comparison to alternative models.\index{factor analysis!confirmatory}\index{latent variable}
+Using a [CFA](#sec-cfa) approach, the researcher can pit models representing two theoretical frameworks against each other to see which better accounts for the observed data.\index{factor analysis!confirmatory}
+
+Factor analysis involves observed (manifest) variables and unobserved (latent) factors.\index{factor analysis}\index{latent variable}
+Factor analysis assumes that the latent factor influences the manifest variables, and the latent factor therefore reflects the common variance among the variables.\index{factor analysis}\index{latent variable}\index{latent variable}
+A factor model potentially includes factor loadings, residuals (errors or disturbances), intercepts/means, covariances, and regression paths.\index{factor analysis}
+A regression path indicates a hypothesis that one variable (or factor) influences another.\index{factor analysis}
+The standardized regression coefficient represents the strength of association between the variables or factors.\index{factor analysis}\index{standardized regression coefficient}
+A factor loading is a regression path from a latent factor to an observed (manifest) variable.\index{factor analysis}\index{factor analysis!factor loading}\index{latent variable}
+The standardized factor loading represents the strength of association between the variable and the latent factor.\index{factor analysis}\index{factor analysis!factor loading}\index{latent variable}\index{standardized regression coefficient}
+A residual is variance in a variable (or factor) that is unexplained by other variables or factors.\index{factor analysis}\index{factor analysis!residual}
+An indicator's intercept is the expected value of the variable when the factor(s) (onto which it loads) is equal to zero.\index{factor analysis}\index{factor analysis!intercept}
+Covariances are the associations between variables (or factors).\index{factor analysis}\index{factor analysis!covariance}
+A covariance path between two variables represents omitted shared cause(s) of the variables.\index{factor analysis}\index{factor analysis!covariance}
+For instance, if you depict a covariance path between two variables, it means that there is a shared cause of the two variables that is omitted from the model (for instance, if the common cause is not known or was not assessed).\index{factor analysis}\index{factor analysis!covariance}
+
+In factor analysis, the relation between an indicator ($\text{X}$) and its underlying latent factor(s) ($\text{F}$) can be represented with a regression formula as in @eq-indicatorLatentAssociation:\index{factor analysis}\index{factor analysis!factor loading}\index{factor analysis!intercept}
+
+$$
+\bar{X} = \lambda \cdot \text{F} + \text{Item Intercept} + \text{Error Term}
+$$ {#eq-indicatorLatentAssociation}
+
+where:
+
+- $\text{X}$ is the observed value of the indicator
+- $\lambda$ is the factor loading, indicating the strength of the association between the indicator and the latent factor(s)
+- $\text{F}$ is the person's value on the latent factor(s)
+- $\text{Item Intercept}$ represents the constant term that accounts for the expected value of the indicator when the latent factor(s) are zero
+- $\text{Error Term}$ is the residual, indicating the extent of variance in the indicator that is not explained by the latent factor(s)
+
+When the latent factors are uncorrelated, the (standardized) error term for an indicator is calculated as 1 minus the sum of squared standardized factor loadings for a given item (including cross-loadings).\index{factor analysis}\index{factor analysis!residual}\index{factor analysis!factor loading}\index{cross-loading}
+A cross-loadings is when a variable loads onto more than one latent factor.\index{factor analysis}\index{cross-loading}
+
+Factor analysis is a powerful technique to help identify the factor structure that underlies a measure or construct.\index{factor analysis}
+However, given the extensive method variance that influences scores on measure, factor analysis (and [principal component analysis](#sec-pca)) tends to extract method factors.\index{factor analysis}\index{principal component analysis}\index{method bias}\index{factor analysis!method factor}
+Method factors are factors that are related to the methods being assessed rather than the construct of interest.\index{factor analysis}\index{method bias}\index{factor analysis!method factor}
+To better estimate construct factors, it is sometimes necessary to estimate both construct and method factors.\index{factor analysis}\index{factor analysis!method factor}
+
+## Exploratory Factor Analysis {#sec-cfa}
+
+Exploratory factor analysis (EFA) is used if you have no a priori hypotheses about the factor structure of the model, but you would like to understand the latent variables represented by your items.\index{factor analysis!exploratory}
+
+EFA is partly induced from the data.\index{factor analysis!exploratory}
+You feed in the data and let the program build the factor model.\index{factor analysis!exploratory}
+You can set some parameters going in, including how to extract or rotate the factors.\index{factor analysis!exploratory}
+The factors are extracted from the data without specifying the number and pattern of loadings between the items and the latent factors [@Bollen2002].\index{factor analysis!exploratory}
+All cross-loadings are freely estimated.\index{cross-loading}\index{factor analysis!exploratory}
+
+### Factor Rotation {#sec-factorAnalysisRotation}
+
+When using EFA or [PCA](#sec-pca), an important step is, possibly, to rotate the factors to make them more interpretable and simple, which is the whole goal.\index{factor analysis}\index{principal component analysis}\index{rotation}
+To interpret the results of a factor analysis, we examine the factor matrix.\index{factor analysis}\index{rotation}
+The columns refer to the different factors; the rows refer to the different observed variables.\index{factor analysis}\index{rotation}
+The cells in the table are the factor loadings—they are basically the correlation between the variable and the factor.\index{factor analysis}\index{rotation}\index{factor analysis!factor loading}
+Our goal is to achieve a model with simple structure because it is easily interpretable.\index{factor analysis}\index{simple structure}
+*Simple structure* means that every variable loads perfectly on one and only one factor, as operationalized by a matrix of factor loadings with values of one and zero and nothing else.\index{factor analysis}\index{simple structure}
+An example of a factor matrix that follows simple structure is depicted in @fig-simpleStructure.\index{factor analysis}\index{simple structure}
+
+::: {#fig-simpleStructure}
+![](images/simpleStructure.png){fig-alt="Example of a Factor Matrix That Follows Simple Structure."}
+
+Example of a Factor Matrix That Follows Simple Structure.
+:::
+
+An example of a measurement model that follows simple structure is depicted in @fig-factorSolutionSimpleStructure.\index{factor analysis}\index{simple structure}
+Each variable loads onto one and only one factor, which makes it easy to interpret the meaning of each factor, because a given factor represents the common variance among the items that load onto it.\index{factor analysis}
+
+::: {#fig-factorSolutionSimpleStructure}
+![](images/factorSolutionSimpleStructure.png){fig-alt="Example of a Measurement Model That Follows Simple Structure.
+'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems."}
+
+Example of a Measurement Model That Follows Simple Structure. 'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
+:::
+
+However, pure simple structure only occurs in simulations, not in real-life data.\index{factor analysis}\index{simple structure}
+In reality, our [measurement model](#measurementModel-sem) in an unrotated factor analysis model might look like the model in @fig-factorSolutionUnrotatedExample.\index{factor analysis}
+In this example, the [measurement model](#measurementModel-sem) does not show simple structure because the items have cross-loadings—that is, the items load onto more than one factor.\index{factor analysis}\index{cross-loading}\index{simple structure}\index{structural equation modeling!measurement model}
+The cross-loadings make it difficult to interpret the factors, because all of the items load onto all of the factors, so the factors are not very distinct from each other, which makes it difficult to interpret what the factors mean.\index{cross-loading}\index{factor analysis}\index{cross-loading}\index{simple structure}\index{structural equation modeling!measurement model}
+
+::: {#fig-factorSolutionUnrotatedExample}
+![](images/factorSolutionUnrotatedExample.png){fig-alt="Example of a Measurement Model That Does Not Follow Simple Structure.
+'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems."}
+
+Example of a Measurement Model That Does Not Follow Simple Structure. 'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
+:::
+
+As a result of the challenges of interpretability caused by cross-loadings, factor rotations are often performed.\index{cross-loading}\index{factor analysis}\index{rotation}
+An example of an unrotated factor matrix is in @fig-factorMatrix.\index{factor analysis}\index{rotation}
+
+::: {#fig-factorMatrix}
+![](images/factorMatrix.png){fig-alt="Example of a Factor Matrix."}
+
+Example of a Factor Matrix.
+:::
+
+In the example factor matrix in @fig-factorSolutionUnrotated, the factor analysis is not very helpful—it tells us very little because it did not distinguish between the two factors.\index{factor analysis}\index{rotation}
+The variables have similar loadings on Factor 1 and Factor 2.\index{factor analysis}\index{rotation}
+An example of a unrotated factor solution is in Figure \@ref(fig:factorSolutionUnrotated).\index{factor analysis}\index{rotation}
+In the figure, all of the variables are in the midst of the quadrants—they are not on the factors' axes.\index{factor analysis}\index{rotation}
+Thus, the factors are not very informative.\index{factor analysis}\index{rotation}
+
+::: {#fig-factorSolutionUnrotated}
+![](images/factorSolutionUnrotated.png){fig-alt="Example of an Unrotated Factor Solution."}
+
+Example of an Unrotated Factor Solution.
+:::
+
+As a result, to improve the interpretability of the factor analysis, we can do what is called rotation.\index{factor analysis}\index{rotation}
+Rotation leverages the idea that there are infinite solutions to the factor analysis model that fit equally well.\index{factor analysis}\index{rotation}
+*Rotation* involves changing the orientation of the factors by changing the axes so that variables end up with very high (close to one or negative one) or very low (close to zero) loadings, so that it is clear which factors include which variables.\index{factor analysis}\index{rotation}
+That is, rotation rescales the factors and tries to identify the ideal solution (factor) for each variable.\index{factor analysis}\index{rotation}
+It searches for simple structure and keeps searching until it finds a minimum.\index{factor analysis}\index{rotation}\index{simple structure}
+After rotation, if the rotation was successful for imposing simple structure, each factor will have loadings close to one (or negative one) for some variables and close to zero for other variables.\index{factor analysis}\index{rotation}\index{simple structure}
+The goal of factor rotation is to achieve simple structure, to help make it easier to interpret the meaning of the factors.\index{factor analysis}\index{rotation}\index{simple structure}
+
+To perform factor rotation, orthogonal rotations are often used.\index{factor analysis}\index{rotation}\index{simple structure}\index{orthogonal rotation}
+Orthogonal rotations make the rotated factors uncorrelated.\index{factor analysis}\index{rotation}\index{simple structure}\index{orthogonal rotation}
+An example of a commonly used orthogonal rotation is varimax rotation.\index{factor analysis}\index{rotation}\index{simple structure}\index{orthogonal rotation}
+Varimax rotation maximizes the sum of the variance of the squared loadings (i.e., so that items have either a very high or very low loading on a factor) and yields axes with a 90-degree angle.\index{factor analysis}\index{rotation}\index{simple structure}\index{orthogonal rotation}
+
+An example of a factor matrix following an orthogonal rotation is depicted in @fig-factorMatrixRotated.\index{factor analysis}\index{rotation}\index{simple structure}\index{orthogonal rotation}
+An example of a factor solution following an orthogonal rotation is depicted in @fig-factorSolutionRotated.\index{factor analysis}\index{rotation}\index{simple structure}\index{orthogonal rotation}
+
+::: {#fig-factorMatrixRotated}
+![](images/factorMatrixRotated.png){fig-alt="Example of a Rotated Factor Matrix."}
+
+Example of a Rotated Factor Matrix.
+:::
+
+::: {#fig-factorSolutionRotated}
+![](images/factorSolutionRotated.png){fig-alt="Example of a Rotated Factor Solution."}
+
+Example of a Rotated Factor Solution.
+:::
+
+An example of a factor matrix from SPSS following an orthogonal rotation is depicted in @fig-rotatedFactorMatrix.\index{factor analysis}\index{rotation}\index{simple structure}\index{orthogonal rotation}
+
+::: {#fig-rotatedFactorMatrix}
+![](images/rotatedFactorMatrix.png){fig-alt="Example of a Rotated Factor Matrix From SPSS."}
+
+Example of a Rotated Factor Matrix From SPSS.
+:::
+
+An example of a factor structure from an orthogonal rotation is in @fig-orthogonalRotation.\index{factor analysis}\index{rotation}\index{simple structure}\index{orthogonal rotation}
+
+::: {#fig-orthogonalRotation}
+![](images/FactorAnalysis-08.png){fig-alt="Example of a Factor Structure From an Orthogonal Rotation."}
+
+Example of a Factor Structure From an Orthogonal Rotation.
+:::
+
+Sometimes, however, the two factors and their constituent variables may be correlated.\index{factor analysis}\index{rotation}\index{orthogonal rotation}\index{oblique rotation}
+Examples of two correlated factors may be depression and anxiety.\index{factor analysis}\index{rotation}\index{orthogonal rotation}\index{oblique rotation}
+When the two factors are correlated in reality, if we make them uncorrelated, this would result in an inaccurate model.\index{factor analysis}\index{rotation}\index{orthogonal rotation}\index{oblique rotation}
+Oblique rotation allows for factors to be correlated and yields axes with less an angle of less than 90 degrees.\index{factor analysis}\index{rotation}\index{oblique rotation}
+However, if the factors have low correlation (e.g., .2 or less), you can likely continue with orthogonal rotation.\index{factor analysis}\index{rotation}\index{oblique rotation}
+Nevertheless, just because an oblique rotation allows for correlated factors does not mean that the factors will be correlated, so oblique rotation provides greater flexibility than orthogonal rotation.\index{factor analysis}\index{rotation}\index{orthogonal rotation}\index{oblique rotation}
+An example of a factor structure from an oblique rotation is in @fig-obliqueRotation.\index{factor analysis}\index{rotation}\index{oblique rotation}
+Results from an oblique rotation are more complicated than orthogonal rotation—they provide lots of output and are more complicated to interpret.\index{factor analysis}\index{rotation}\index{oblique rotation}
+In addition, oblique rotation might not yield a smooth answer if you have a relatively small sample size.\index{factor analysis}\index{rotation}\index{oblique rotation}
+
+::: {#fig-obliqueRotation}
+![](images/FactorAnalysis-09.png){fig-alt="Example of a Factor Structure From an Oblique Rotation."}
+
+Example of a Factor Structure From an Oblique Rotation.
+:::
+
+As an example of rotation based on interpretability, consider the Five-Factor Model of Personality (the Big Five), which goes by the acronym, OCEAN: **O**penness, **C**onscientiousness, **E**xtraversion, **A**greeableness, and **N**euroticism.\index{factor analysis}\index{rotation}\index{oblique rotation}
+Although the five factors of personality are somewhat correlated, we can use rotation to ensure they are maximally independent.\index{factor analysis}\index{rotation}
+Upon rotation, extraversion and neuroticism are essentially uncorrelated, as depicted in @fig-factorRotation.\index{factor analysis}\index{rotation}
+The other pole of extraversion is intraversion and the other pole of neuroticism might be emotional stability or calmness.
+
+::: {#fig-factorRotation}
+![](images/factorRotation.png){fig-alt="Example of a Factor Rotation of Neuroticism and Extraversion."}
+
+Example of a Factor Rotation of Neuroticism and Extraversion.
+:::
+
+Simple structure is achieved when each variable loads highly onto as few factors as possible (i.e., each item has only one significant or primary loading).\index{simple structure}
+Oftentimes this is not the case, so we choose our rotation method in order to decide if the factors can be correlated (an oblique rotation) or if the factors will be uncorrelated (an orthogonal rotation).\index{factor analysis}\index{rotation}\index{orthogonal rotation}\index{oblique rotation}
+If the factors are not correlated with each other, use an orthogonal rotation.\index{factor analysis}\index{rotation}\index{orthogonal rotation}
+The correlation between an item and a factor is a factor loading, which is simply a way to ask how much a variable is correlated with the underlying factor.\index{factor analysis}\index{structural equation modeling!factor loading}
+However, its interpretation is more complicated if there are correlated factors!\index{factor analysis}\index{rotation}\index{oblique rotation}
+
+An orthogonal rotation (e.g., varimax) can help with simplicity of interpretation because it seeks to yield simple structure without cross-loadings.\index{simple structure}\index{cross-loading}\index{factor analysis}\index{rotation}\index{orthogonal rotation}
+Cross-loadings are instances where a variable loads onto multiple factors.\index{cross-loading}
+My recommendation would always be to use an orthogonal rotation if you have reason to believe that finding simple structure in your data is possible; otherwise, the factors are extremely difficult to interpret—what exactly does a cross-loading even mean?\index{cross-loading}\index{factor analysis}\index{rotation}\index{orthogonal rotation}
+However, you should always try an oblique rotation, too, to see how strongly the factors are correlated.\index{factor analysis}\index{rotation}\index{oblique rotation}
+Examples of oblique rotations include oblimin and promax.\index{factor analysis}\index{rotation}\index{oblique rotation}
+
+## Confirmatory Factor Analysis {#sec-cfa}
+
+Confirmatory factor analysis (CFA) is used to (dis)confirm a priori hypotheses about the factor structure of the model.\index{factor analysis!confirmatory}
+CFA is a test of the hypothesis.\index{factor analysis!confirmatory}
+In CFA, you specify the model and ask how well this model represents the data.\index{factor analysis!confirmatory}
+The researcher specifies the number, meaning, associations, and pattern of free parameters in the factor loading matrix [@Bollen2002].\index{factor analysis!confirmatory}
+A key advantage of CFA is the ability to directly compare alternative models (i.e., factor structures), which is valuable for theory testing [@Strauss2009].\index{factor analysis!confirmatory}
+For instance, you could use [CFA](#cfa) to test whether the variance in several measures' scores is best explained with one factor or two factors.\index{factor analysis!confirmatory}
+In CFA, cross-loadings are not estimated unless the researcher specifies them.\index{cross-loading}\index{factor analysis!confirmatory}
+
+## Determining the Number of Factors to Retain {#sec-factorAnalysisNumFactors}
+
+A goal of factor analysis and [PCA](#pca) is simplification or parsimony, while still explaining as much variance as possible.\index{parsimony}\index{factor analysis}\index{principal component analysis}
+The hope is that you can have fewer factors that explain the associations between the variables than the number of observed variables.\index{factor analysis}
+But how do you decide on the number of factors?\index{factor analysis}
+
+There are a number of criteria that one can use to help determine how many factors/components to keep:\index{factor analysis}
+
+- Kaiser-Guttman criterion: in [PCA](#pca), components with eigenvalues greater than 1\index{principal component analysis}\index{eigenvalue}
+    - or, for factor analysis, factors with eigenvalues greater than zero\index{factor analysis}\index{eigenvalue}
+- Cattell's scree test: the "elbow" in a scree plot minus one; sometimes operationalized with optimal coordinates (OC) or the acceleration factor (AF)\index{factor analysis}\index{scree plot}
+- Parallel analysis: factors that explain more variance than randomly simulated data\index{factor analysis}\index{parallel analysis}
+- [Very simple structure (VSS)](#vssPlot) criterion: larger is better\index{factor analysis}
+- Velicer's minimum average partial (MAP) test: smaller is better\index{factor analysis}
+- Akaike information criterion (AIC): smaller is better\index{factor analysis}
+- Bayesian information criterion (BIC): smaller is better\index{factor analysis}
+- Sample size-adjusted BIC (SABIC): smaller is better\index{factor analysis}
+- Root mean square error of approximation (RMSEA): smaller is better\index{factor analysis}
+- Chi-square difference test: smaller is better; a significant test indicates that the more complex model is significantly better fitting than the less complex model\index{factor analysis}
+- Standardized root mean square residual (SRMR): smaller is better\index{factor analysis}
+- Comparative Fit Index (CFI): larger is better\index{factor analysis}
+- Tucker Lewis Index (TLI): larger is better\index{factor analysis}
+
+There is not necessarily a "correct" criterion to use in determining how many factors to keep, so it is generally recommended that researchers use multiple criteria in combination with theory and interpretability.\index{factor analysis}
+
+A scree plot provides lots of information.\index{factor analysis}\index{scree plot}
+A scree plot has the factor number on the x-axis and the eigenvalue on the y-axis.\index{factor analysis}\index{scree plot}\index{eigenvalue}
+The eigenvalue is the variance accounted for by a factor; when using a varimax (orthogonal) rotation, an eigenvalue (or factor variance) is calculated as the sum of squared standardized factor (or component) loadings on that factor.\index{factor analysis}\index{scree plot}\index{eigenvalue}
+An example of a scree plot is in @fig-screePlot.\index{factor analysis}\index{scree plot}
+
+::: {#fig-screePlot}
+![](images/screePlot.png){fig-alt="Example of a Scree Plot."}
+
+Example of a Scree Plot.
+:::
+
+The total variance is equal to the number of variables you have, so one eigenvalue is approximately one variable's worth of variance.\index{eigenvalue}\index{factor analysis}
+The first factor  accounts for the most variance, the second factor accounts for the second-most variance, and so on.\index{factor analysis}
+The more factors you add, the less variance is explained by the additional factor.\index{factor analysis}
+
+One criterion for how many factors to keep is the Kaiser-Guttman criterion.
+According to the Kaiser-Guttman criterion, you should keep any factors whose eigenvalue is greater than 1.\index{factor analysis}
+That is, for the sake of simplicity, parsimony, and data reduction, you should take any factors that explain more than a single variable would explain.\index{parsimony}\index{data!reduction}\index{factor analysis}
+According to the Kaiser-Guttman criterion, we would keep three factors from Figure \@ref(fig:screePlot) that have eigenvalues greater than 1.\index{factor analysis}\index{eigenvalue}
+The default in SPSS is to retain factors with eigenvalues greater than 1.\index{factor analysis}\index{eigenvalue}
+However, keeping factors whose eigenvalue is greater than 1 is not the most correct rule.\index{factor analysis}\index{eigenvalue}
+If you let SPSS do this, you may get many factors with eigenvalues around 1 (e.g., factors with an eigenvalue ~ 1.0001) that are not adding so much that it is worth the added complexity.\index{factor analysis}\index{eigenvalue}
+The Kaiser-Guttman criterion usually results in keeping too many factors.\index{factor analysis}
+Factors with small eigenvalues around 1 could reflect error shared across variables.\index{factor analysis}\index{eigenvalue}
+For instance, factors with small eigenvalues could reflect method variance (i.e., method factor), such as a self-report factor that turns up as a factor in factor analysis, but that may be useless to you as a conceptual factor of a construct of interest.\index{factor analysis!method factor}\index{factor analysis}\index{eigenvalue}
+
+Another criterion is Cattell's scree test, which involves selecting the number of factors from looking at the scree plot.\index{factor analysis}\index{scree plot}
+"Scree" refers to the rubble of stones at the bottom of a mountain.\index{factor analysis}\index{scree plot}
+According to Cattell's scree test, you should keep the factors before the last steep drop in eigenvalues—i.e., the factors before the rubble, where the slope approaches zero.\index{factor analysis}\index{eigenvalue}\index{scree plot}
+The beginning of the scree (or rubble), where the slope approaches zero, is called the "elbow" of a scree plot.\index{factor analysis}\index{scree plot}
+Using Cattell's scree test, you retain the number of factors that explain the most variance prior to the explained variance drop-off, because, ultimately, you want to include only as many factors in which you gain substantially more by the inclusion of these factors.\index{factor analysis}\index{scree plot}
+That is, you would keep the number of factors at the elbow of the scree plot minus one.\index{factor analysis}\index{scree plot}
+If the last steep drop occurs from Factor 4 to Factor 5 and the elbow is at Factor 5, we would keep four factors.\index{factor analysis}\index{scree plot}
+In Figure \@ref(fig:screePlot), the last steep drop in eigenvalues occurs from Factor 3 to Factor 4; the elbow of the scree plot occurs at Factor 4.\index{factor analysis}\index{scree plot}
+We would keep the number of factors at the elbow minus one.\index{factor analysis}\index{scree plot}
+Thus, using Cattell's scree test, we would keep three factors based on Figure \@ref(fig:screePlot).\index{factor analysis}\index{scree plot}
+
+There are more sophisticated ways of using a scree plot, but they usually end up at a similar decision.\index{factor analysis}\index{scree plot}
+Examples of more sophisticated tests include parallel analysis and [very simple structure (VSS) plots](#vssPlot).\index{factor analysis}\index{very simple structure plot}\index{parallel analysis}
+In a parallel analysis, you examine where the eigenvalues from observed data and random data converge, so you do not retain a factor that explains less variance than would be expected by random chance.\index{factor analysis}\index{parallel analysis}\index{eigenvalue}
+
+In general, my recommendation is to use Cattell's scree test, and then test the factor solutions with plus or minus one factor.\index{factor analysis}\index{scree plot}
+You should never accept [PCA](#pca) components with eigenvalues less than one (or factors with eigenvalues less than zero), because they are likely to be largely composed of error.\index{factor analysis}\index{eigenvalue}
+If you are using maximum likelihood factor analysis, you can compare the fit of various models with model fit criteria to see which model fits best for its parsimony.\index{parsimony}\index{factor analysis}
+A model will always fit better when you add additional parameters or factors, so you examine if there is *significant* improvement in model fit when adding the additional factor—that is, we keep adding complexity until additional complexity does not buy us much.\index{parsimony}\index{factor analysis}
+Always try a factor solution that is one less and one more than suggested by Cattell's scree test to buffer your final solution because the purpose of factor analysis is to explain things and to have interpretability.\index{factor analysis}\index{scree plot}
+Even if all rules or indicators suggest to keep X number of factors, maybe $\pm$ one factor helps clarify things.\index{factor analysis}
+Even though factor analysis is empirical, theory and interpretatability should also inform decisions.\index{factor analysis}
+
+## Interpreting and Using Latent Factors {#sec-interpretingAndUsingLatentFactors}
+
+The next step is interpreting the model and latent factors.\index{factor analysis}
+One data matrix can lead to many different (correct) models—you must choose one based on the factor structure and theory.\index{factor analysis}
+Use theory to interpret the model and label the factors.\index{factor analysis}
+
+## Conclusion {#sec-factorAnalysisConclusion}
+
+::: {.content-visible when-format="html"}
+
+## Session Info {#sec-factorAnalysisSessionInfo}
+
+```{r}
+sessionInfo()
+```
+
+:::

### 2025-02-26 factor-analysis.qmd
20250226 - edits

@@
 A factor model potentially includes factor loadings, residuals (errors or disturbances), intercepts/means, covariances, and regression paths.\index{factor analysis}
-A regression path indicates a hypothesis that one variable (or factor) influences another.\index{factor analysis}
-The standardized regression coefficient represents the strength of association between the variables or factors.\index{factor analysis}\index{standardized regression coefficient}
+When depicting a factor analysis model, rectangles represent variables we observe (i.e., manifest variables), and circles represent latent (i.e., unobserved) variables.\index{factor analysis}\index{latent variable}
+A regression path indicates a hypothesis that one variable (or factor) influences another, and it is depicted using a single-headed arrow.\index{factor analysis}
+The standardized regression coefficient (i.e., beta or $\beta$) represents the strength of association between the variables or factors.\index{factor analysis}\index{standardized regression coefficient}
 A factor loading is a regression path from a latent factor to an observed (manifest) variable.\index{factor analysis}\index{factor analysis!factor loading}\index{latent variable}
@@
 A residual is variance in a variable (or factor) that is unexplained by other variables or factors.\index{factor analysis}\index{factor analysis!residual}
-An indicator's intercept is the expected value of the variable when the factor(s) (onto which it loads) is equal to zero.\index{factor analysis}\index{factor analysis!intercept}
-Covariances are the associations between variables (or factors).\index{factor analysis}\index{factor analysis!covariance}
+A variable's intercept is the expected value of the variable when the factor(s) (onto which it loads) is equal to zero.\index{factor analysis}\index{factor analysis!intercept}
+A covariance is the unstandardized index of the strength of association between between variables (or factors), and it is depicted with a double-headed arrow.\index{factor analysis}\index{factor analysis!covariance}
+Because a covariance is unstandardized, its scale depends on the scale of the variables.\index{factor analysis!covariance}
 A covariance path between two variables represents omitted shared cause(s) of the variables.\index{factor analysis}\index{factor analysis!covariance}
@@
 
-An example of a measurement model that follows simple structure is depicted in @fig-factorSolutionSimpleStructure.\index{factor analysis}\index{simple structure}
+An example of a factor analysis model that follows simple structure is depicted in @fig-factorSolutionSimpleStructure.\index{factor analysis}\index{simple structure}
 Each variable loads onto one and only one factor, which makes it easy to interpret the meaning of each factor, because a given factor represents the common variance among the items that load onto it.\index{factor analysis}
@@
 ::: {#fig-factorSolutionSimpleStructure}
-![](images/factorSolutionSimpleStructure.png){fig-alt="Example of a Measurement Model That Follows Simple Structure.
+![](images/factorSolutionSimpleStructure.png){fig-alt="Example of a Factor Analysis Model That Follows Simple Structure.
 'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems."}
 
-Example of a Measurement Model That Follows Simple Structure. 'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
+Example of a Factor Analysis Model That Follows Simple Structure. 'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
 :::
@@
 However, pure simple structure only occurs in simulations, not in real-life data.\index{factor analysis}\index{simple structure}
-In reality, our [measurement model](#measurementModel-sem) in an unrotated factor analysis model might look like the model in @fig-factorSolutionUnrotatedExample.\index{factor analysis}
-In this example, the [measurement model](#measurementModel-sem) does not show simple structure because the items have cross-loadings—that is, the items load onto more than one factor.\index{factor analysis}\index{cross-loading}\index{simple structure}\index{structural equation modeling!measurement model}
-The cross-loadings make it difficult to interpret the factors, because all of the items load onto all of the factors, so the factors are not very distinct from each other, which makes it difficult to interpret what the factors mean.\index{cross-loading}\index{factor analysis}\index{cross-loading}\index{simple structure}\index{structural equation modeling!measurement model}
+In reality, our unrotated factor analysis model might look like the model in @fig-factorSolutionUnrotatedExample.\index{factor analysis}
+In this example, the factor analysis model does not show simple structure because the items have cross-loadings—that is, the items load onto more than one factor.\index{factor analysis}\index{cross-loading}\index{simple structure}
+The cross-loadings make it difficult to interpret the factors, because all of the items load onto all of the factors, so the factors are not very distinct from each other, which makes it difficult to interpret what the factors mean.\index{cross-loading}\index{factor analysis}\index{cross-loading}\index{simple structure}
 
 ::: {#fig-factorSolutionUnrotatedExample}
-![](images/factorSolutionUnrotatedExample.png){fig-alt="Example of a Measurement Model That Does Not Follow Simple Structure.
+![](images/factorSolutionUnrotatedExample.png){fig-alt="Example of a Factor Analysis Model That Does Not Follow Simple Structure.
 'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems."}
 
-Example of a Measurement Model That Does Not Follow Simple Structure. 'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
+Example of a Factor Analysis Model That Does Not Follow Simple Structure. 'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
 :::

### 2025-02-26 pca.qmd
20250226 - edits
20250226 - ensuring the variables are on the same scale

@@
 
+## Ensuring the Variables are on the Same Scale {#sec-pcaStandardize}
+
+Before performing a PCA, it is important to ensure that the variables included in the PCA are on the same scale.
+PCA seeks to identify components that explain variance in the data, so if the variables are not on the same scale, some variables may contribute considerably more variance than others.
+A common way of ensuring that variables are on the same scale is to standardize them using, for example,  [*z*-scores](#sec-zScores).
+
 ## Component Rotation {#sec-pcaRotation}
@@
 
+Both [factor analysis](#sec-factorAnalysis) and PCA can be used for data reducation.
 The key distinction between [factor analysis](#sec-factorAnalysis) and PCA is depicted in @fig-pcaVsFactorAnalysis.\index{factor analysis}\index{principal component analysis}

### 2025-02-26 factor-analysis.qmd
20250226 - edits

@@
 
-When using EFA or [PCA](#sec-pca), an important step is, possibly, to rotate the factors to make them more interpretable and simple, which is the whole goal.\index{factor analysis}\index{principal component analysis}\index{rotation}
+When using EFA or [principal component analysis](#sec-pca), an important step is, possibly, to rotate the factors to make them more interpretable and simple, which is the whole goal.\index{factor analysis}\index{principal component analysis}\index{rotation}
 To interpret the results of a factor analysis, we examine the factor matrix.\index{factor analysis}\index{rotation}
@@
 The variables have similar loadings on Factor 1 and Factor 2.\index{factor analysis}\index{rotation}
-An example of a unrotated factor solution is in Figure \@ref(fig:factorSolutionUnrotated).\index{factor analysis}\index{rotation}
+An example of a unrotated factor solution is in @fig-factorSolutionUnrotated.\index{factor analysis}\index{rotation}
 In the figure, all of the variables are in the midst of the quadrants—they are not on the factors' axes.\index{factor analysis}\index{rotation}
@@
 
-A goal of factor analysis and [PCA](#pca) is simplification or parsimony, while still explaining as much variance as possible.\index{parsimony}\index{factor analysis}\index{principal component analysis}
+A goal of factor analysis and [principal component analysis](#sec-pca) is simplification or parsimony, while still explaining as much variance as possible.\index{parsimony}\index{factor analysis}\index{principal component analysis}
 The hope is that you can have fewer factors that explain the associations between the variables than the number of observed variables.\index{factor analysis}
@@
 
-- Kaiser-Guttman criterion: in [PCA](#pca), components with eigenvalues greater than 1\index{principal component analysis}\index{eigenvalue}
-    - or, for factor analysis, factors with eigenvalues greater than zero\index{factor analysis}\index{eigenvalue}
+- Kaiser-Guttman criterion: factors with eigenvalues greater than zero\index{factor analysis}\index{eigenvalue}
+    - or, for [principal component analysis](#sec-pca), components with eigenvalues greater than 1\index{principal component analysis}\index{eigenvalue}
 - Cattell's scree test: the "elbow" in a scree plot minus one; sometimes operationalized with optimal coordinates (OC) or the acceleration factor (AF)\index{factor analysis}\index{scree plot}
 - Parallel analysis: factors that explain more variance than randomly simulated data\index{factor analysis}\index{parallel analysis}
-- [Very simple structure (VSS)](#vssPlot) criterion: larger is better\index{factor analysis}
+- Very simple structure (VSS) criterion: larger is better\index{factor analysis}
 - Velicer's minimum average partial (MAP) test: smaller is better\index{factor analysis}
@@
 That is, for the sake of simplicity, parsimony, and data reduction, you should take any factors that explain more than a single variable would explain.\index{parsimony}\index{data!reduction}\index{factor analysis}
-According to the Kaiser-Guttman criterion, we would keep three factors from Figure \@ref(fig:screePlot) that have eigenvalues greater than 1.\index{factor analysis}\index{eigenvalue}
+According to the Kaiser-Guttman criterion, we would keep three factors from @fig-screePlot that have eigenvalues greater than 1.\index{factor analysis}\index{eigenvalue}
 The default in SPSS is to retain factors with eigenvalues greater than 1.\index{factor analysis}\index{eigenvalue}
@@
 If the last steep drop occurs from Factor 4 to Factor 5 and the elbow is at Factor 5, we would keep four factors.\index{factor analysis}\index{scree plot}
-In Figure \@ref(fig:screePlot), the last steep drop in eigenvalues occurs from Factor 3 to Factor 4; the elbow of the scree plot occurs at Factor 4.\index{factor analysis}\index{scree plot}
+In @fig-screePlot, the last steep drop in eigenvalues occurs from Factor 3 to Factor 4; the elbow of the scree plot occurs at Factor 4.\index{factor analysis}\index{scree plot}
 We would keep the number of factors at the elbow minus one.\index{factor analysis}\index{scree plot}
-Thus, using Cattell's scree test, we would keep three factors based on Figure \@ref(fig:screePlot).\index{factor analysis}\index{scree plot}
+Thus, using Cattell's scree test, we would keep three factors based on @fig-screePlot.\index{factor analysis}\index{scree plot}
 
 There are more sophisticated ways of using a scree plot, but they usually end up at a similar decision.\index{factor analysis}\index{scree plot}
-Examples of more sophisticated tests include parallel analysis and [very simple structure (VSS) plots](#vssPlot).\index{factor analysis}\index{very simple structure plot}\index{parallel analysis}
+Examples of more sophisticated tests include parallel analysis and very simple structure (VSS) plots.\index{factor analysis}\index{very simple structure plot}\index{parallel analysis}
 In a parallel analysis, you examine where the eigenvalues from observed data and random data converge, so you do not retain a factor that explains less variance than would be expected by random chance.\index{factor analysis}\index{parallel analysis}\index{eigenvalue}
@@
 In general, my recommendation is to use Cattell's scree test, and then test the factor solutions with plus or minus one factor.\index{factor analysis}\index{scree plot}
-You should never accept [PCA](#pca) components with eigenvalues less than one (or factors with eigenvalues less than zero), because they are likely to be largely composed of error.\index{factor analysis}\index{eigenvalue}
+You should never accept factors with eigenvalues less than zero (or components from [principal component analysis](#sec-pca) with eigenvalues less than one), because they are likely to be largely composed of error.\index{factor analysis}\index{eigenvalue}
 If you are using maximum likelihood factor analysis, you can compare the fit of various models with model fit criteria to see which model fits best for its parsimony.\index{parsimony}\index{factor analysis}

### 2025-02-26 pca.qmd
20250226 - edits

@@
 You can extract as many eigenvectors as there are variables.\index{principal component analysis}\index{eigenvector}
-If you extract all six components, the data matrix left over will be the same as the correlation matrix in Figure \@ref(fig:correlationMatrix2).\index{principal component analysis}
+If you extract all six components, the data matrix left over will be the same as the correlation matrix in @fig-correlationMatrix2.\index{principal component analysis}
 That is, the remaining variables (as part of the leftover data matrix) will be entirely uncorrelated with each other, because six components explain 100% of the variance from six variables.\index{principal component analysis}
 In other words, you can explain (6) variables with (6) new things!\index{principal component analysis}
+
+::: {#fig-correlationMatrix2}
+![](images/correlationMatrix2.png){fig-alt="Example Correlation Matrix 2."}
+
+Example Correlation Matrix 2.
+:::
 

### 2025-02-27 pca.qmd
20250226 - typos

@@
 However, it does no good if you have to use all (6) components because there is no data reduction from the original number of variables.\index{principal component analysis}\index{data!reduction}
-When the goal is data reducation (as in PCA), the hope is that the first few components will explain most of the variance, so we can explain the variability in the data with fewer components than there are variables.\index{principal component analysis}\index{data!reduction}
+When the goal is data reduction (as in PCA), the hope is that the first few components will explain most of the variance, so we can explain the variability in the data with fewer components than there are variables.\index{principal component analysis}\index{data!reduction}
 
@@
 
-Both [factor analysis](#sec-factorAnalysis) and PCA can be used for data reducation.
+Both [factor analysis](#sec-factorAnalysis) and PCA can be used for data reduction.
 The key distinction between [factor analysis](#sec-factorAnalysis) and PCA is depicted in @fig-pcaVsFactorAnalysis.\index{factor analysis}\index{principal component analysis}

### 2025-02-27 sports-cognitive-psychology.qmd
20250227 - sports and cognitive psychology

@@
 
-<https://stacker.com/sports/15-ways-analytics-has-changed-sports> (archived at <https://perma.cc/PQ5R-TWFA>)
+There are a number of [ways in which analytics have changed sports](https://stacker.com/sports/15-ways-analytics-has-changed-sports) (archived at <https://perma.cc/PQ5R-TWFA>).
+Here is an summary of some of the key ways that analytics has led to changes.
 
-- Bill James, "Moneyball", Billy Beane, Theo Epstein, and role of Sabermetrics in helping MLB teams with limited budgets (e.g., Oakland Athletics) compete with larger budgets: using statistics to identify player value more accurately, especially for identifying undervalued players
-- development of advanced metrics (e.g., value over replacement player, wins above replacement, fielding independent pitching)
-- focus on on-base percentage and slugging percentage (and on-base plus slugging percentage) instead of batting average; defensive shifting; fewer attempts to steal bases; fewer bunts; batters taking more pitches; more frequent pitching changes (for pitcher–batter matchups, e.g., righty–lefty); focus on velocity and spin rate
-- other sports, including basketball and football are making more use of analytics
-- basketball: types of shots (focusing on three pointers and layups); star players resting more games
-- football: going for it on fourth down; greater emphasis on the passing game; drafting Running Backs later in the draft (and, overall, valuing Running Backs less); trading down in the draft for more low picks than fewer high picks [because top picks are frequently overvalued; @Massey2013]
+Much of the history of analytics in sports traces back to Bill James.
+James was a member of the Society for American Baseball Research (SABR).
+He published a series of Baseball Abstracts, which included players' statistics.
+James published the first Baseball Abstract in 1977 (<https://www.pbs.org/thinktank/transcript1197.html>; archived at <https://www.pbs.org/thinktank/transcript1197.html>).
+He coined the term Sabermetrics (which was originally named SABRmetrics, based on the name of the society) to refer to advanced metrics and statistical/empirical analysis of baseball.
+James believed that traditional statistics like batting average, runs batted in (RBIs), and pitchers' wins did not reflect a player's true value—i.e., their contribution to the team's success.
+Instead, he advocated for the use of more advanced metrics that could provide a deeper understanding of a player's performance and value such as on-base percentage (OBP), slugging percentage (SLG), wins above replacement (WAR), and fielding independent pitching (FIP).
+He also developed advanced metrics such as runs created (RC) and defense efficiency rating (DER).
+
+James' ideas were slow to catch on among those in baseball.
+Neverthless, some people eventually caught on to his ideas—and to good success.
+Billy Beane, a general manager, used Sabermetrics to help the Oakland Athletics, a small market team with a limited budget, better compete with teams with larger budgets.
+He used statistics such as on-base percentage to identify player value more accurately, especially for identifying undervalued players.
+The story was described in Michael Lewis' book, Moneyball, which was turned into a movie.
+Following publication of Moneyball, Theo Epstein, who was president of the Boston Red Sox and then of the Chicago Cubs, used sabermetrics to help each win the World Series.
+
+In addition to teams using Sabermetrics to evaluate player talent, teams also began to frequently use statistical analysis to inform decision making during games, which led to key changes in the style of play.
+For instace, defensive shifts—where defensive players moved to locations on the field where particular hitters were most likely to hit the ball—became more common, attempts to steal bases became less common, there were fewer bunts, batters took more pitches (i.e., watched more pitches without swinging), there were more frequent pitching changes (for particular pitcher–batter matchups; such as to have a right-handed pitcher face a left-handed batter or vice versa), and a greater focus on velocity and spin rate among pitchers.
+Some of these analytics-drive changes in play style eventually led Major League Baseball (MLB) to make rule changes to make the game more exciting to watch, including outlawing defensive shifts, reducing the number of pitching changes allowed, and making the bases larger and easier to steal.
+
+Although baseball was one of the first major sports to embrace analytics, other sports have been transformed by analytics, as well.
+For instance, in basketball, there has been a greater focus on three points and buckets close to the rim (e.g., dunks and layups), with way fewer midrange shots.
+Moreover, star players rest more games.
+
+Football has also seen greater use of analytics, although its uptake has been somewhat slower than in many other sports.
+When asked about how he makes the decision about whether to go for two after a touchdown, Steelers Head Coach Mike Tomlin stated:
+
+> We work a menu of plays in that area over the course of the week.
+> We rank them at the latter part of the week, and then we get into the stadium and we play it by ear.
+> A lot of it has to do wiht the feel or the flow of the game.
+> Maybe what personnel group we think they're going to match our personnel group with.
+> As we start to play and work the ball down the field on the drives that produce the touchdowns before the point after, we have a little inclination of what their personality might be at least in terms of matching our personnel.
+> All of those things weigh into the decision.
+> It legitimately is a feel thing.
+> I think that's why you play the game.
+> You can take analytics to baseball and things like that but football is always going to be football.
+> I got a lot of respect for analytics and numbers, but I'm not going to make judgements based on those numbers.
+> The game is the game.
+> It's an emotional one played by emotional and driven men.
+> That's an element of the game you can't measure.
+> Often times decisions such as that weigh heavily into the equation.
+> --- [Steelers Head Coach Mike Tomlin, 2015](https://steelersdepot.com/2015/09/tomlin-prefers-feel-over-analytics/) (archived at <https://perma.cc/7CHJ-BTWX>)
+
+That is, Mike Tomlin, suggested that he does not take into account analytics when making decisions.
+Perhaps that is why the Pittsburgh Steelers were voted [one of the least analytically advanced teams](https://www.espn.com/nfl/story/_/id/29939438/2020-nfl-analytics-survey-which-teams-most-least-analytically-inclined) in the NFL in 2020 (archived at <https://perma.cc/R7FA-HGGB>).
+
+A player talent evaluator for the NFL noted that [many people around the league believe that the use of analytics is better suited for baseball than football](https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html) because baseball involves more games, players, and one-on-one matchups (archived at <https://web.archive.org/web/20200803205803/https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html>).
+By contrast, player performance in football may be more dependent on teammates and playcalls.
+The [player talent evaluator noted](https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html), "In football they don't know how to use all the numbers yet...They have the data, they have all these different stats but the people I have talked to aren't completely sure how to use it." (archived at <https://web.archive.org/web/20200803205803/https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html>).
+
+Nevertheless, more and more people around the league are using analytics; for examples of articles, see [here](https://www.theringer.com/2018/12/19/nfl/nfl-analytics-revolution) (archived at <https://perma.cc/7BKP-A4GJ>)  and [here](https://www.samford.edu/sports-analytics/fans/2020/The-Future-of-NFL-Data-Analytics) (archived at <https://perma.cc/WXE3-53E6>).
+Head Coach Doug Peterson heavily relied on analytics with the Philadelphia Eagles to help them win the 2017 Super Bowl; for more information, see [here](https://www.forbes.com/sites/liamfox/2021/08/12/how-the-nfl-uses-analytics-according-to-the-lead-analyst-of-a-super-bowl-champion/) (archived at <https://perma.cc/49KQ-R785>) and [here](https://www.nfl.com/news/super-bowl-lii-how-the-2017-philadelphia-eagles-were-built-0ap3000000912753) (archived at <https://perma.cc/2GRF-8Z5K>).
+
+Examples of changes in football due to the use of analytics include more often going for it on fourth down, a greater emphasis on the passing game, drafting Running Backs later in the draft (and, more generally, valuing Running Backs less), and trading down in the draft to obtain more low picks (rather than having fewer high picks) because top picks are frequently overvalued [@Massey2013].
 
 ## Coaching and Risk Aversion {#sec-coachingRiskAversion}
+
+It had been known for a long time that going for it on fourth down would frequently increase a team's chances of winning.
+Despite that, historically, teams rarely went for it on fourth down and elected to punt or kick a field goal instead.
+It is curious that there was such a discrepancy between.
+One potential explanation for the discrepancy was because of coaches' risk aversion.
+According to this idea, in the case of failing to successfully convert on fourth down, coaches do not want to have to defend their decision to go for it to the media or the owner or general manager.
+That is, they may often play not to lose, rather than to win, in order to keep their job.
 

### 2025-02-27 sports-cognitive-psychology.qmd
20250227 - typos

@@
 There are a number of [ways in which analytics have changed sports](https://stacker.com/sports/15-ways-analytics-has-changed-sports) (archived at <https://perma.cc/PQ5R-TWFA>).
-Here is an summary of some of the key ways that analytics has led to changes.
+Here is a summary of some of the key ways that analytics has led to changes.
 
@@
 For instace, defensive shifts—where defensive players moved to locations on the field where particular hitters were most likely to hit the ball—became more common, attempts to steal bases became less common, there were fewer bunts, batters took more pitches (i.e., watched more pitches without swinging), there were more frequent pitching changes (for particular pitcher–batter matchups; such as to have a right-handed pitcher face a left-handed batter or vice versa), and a greater focus on velocity and spin rate among pitchers.
-Some of these analytics-drive changes in play style eventually led Major League Baseball (MLB) to make rule changes to make the game more exciting to watch, including outlawing defensive shifts, reducing the number of pitching changes allowed, and making the bases larger and easier to steal.
+Some of these analytics-driven changes in play style eventually led Major League Baseball (MLB) to make rule changes in an attempt to make the game more exciting to watch, including banning defensive shifts, reducing the number of pitching changes allowed, and making the bases larger and easier to steal.
 

### 2025-02-27 sports-cognitive-psychology.qmd
20250227 - edits

@@
 > Often times decisions such as that weigh heavily into the equation.
+> 
 > --- [Steelers Head Coach Mike Tomlin, 2015](https://steelersdepot.com/2015/09/tomlin-prefers-feel-over-analytics/) (archived at <https://perma.cc/7CHJ-BTWX>)
@@
 Despite that, historically, teams rarely went for it on fourth down and elected to punt or kick a field goal instead.
-It is curious that there was such a discrepancy between.
-One potential explanation for the discrepancy was because of coaches' risk aversion.
+It is curious that there was such a discrepancy between the decisions that would maximize teams' winning percentage and the decisions coaches actually made.
+One potential explanation for the discrepancy is because of coaches' risk aversion.
+As noted in @sec-cognitiveBiasesRiskAversion, when it is possible to experience either a gain or a loss from a decision, [loss aversion bias](#sec-cognitiveBiasesLossAversion) tends to lead people to make risk-averse decisions [@Kahneman2011].
 According to this idea, in the case of failing to successfully convert on fourth down, coaches do not want to have to defend their decision to go for it to the media or the owner or general manager.
-That is, they may often play not to lose, rather than to win, in order to keep their job.
 

### 2025-02-27 basic-statistics.qmd
20250227 - clean up

@@
 
-# combine polygons. 
+# combine polygons
 poly1$id <- 3 # alpha, give it the highest number to make it the top layer
@@
 poly <- rbind(poly1, poly2, poly3)
-poly$id <- factor(poly$id,  labels=c("power","beta","alpha"))
+poly$id <- factor(poly$id,  labels = c("power", "beta", "alpha"))
 
@@
   # add line for treatment group
-  geom_line(data=df1, aes(x,y, color="H0", group=NULL, fill=NULL), linewidth=1.5, show_guide=F) + 
+  geom_line(data=df1, aes(x, y, color = "H0", group = NULL, fill = NULL), linewidth = 1.5, show_guide = FALSE) + 
   # add line for treatment group. These lines could be combined into one dataframe.
-  geom_line(data=df2, aes(color="HA", group=NULL, fill=NULL),linewidth=1.5, show_guide=F) +
+  geom_line(data=df2, aes(color = "HA", group = NULL, fill = NULL), linewidth = 1.5, show_guide = FALSE) +
   # add vlines for z_crit
@@
   # change colors
-  scale_color_manual("Group", 
-                     values= c("HA" = "#981e0b","H0" = "black")) +
-  scale_fill_manual("test", values= c("alpha" = "#0d6374","beta" = "#be805e","power"="#7cecee")) +
+  scale_color_manual(
+    "Group",
+    values = c("HA" = "#981e0b", "H0" = "black")) +
+  scale_fill_manual("test", values= c("alpha" = "#0d6374", "beta" = "#be805e", "power"="#7cecee")) +
   # beta arrow
-  annotate("segment", x=0.1, y=0.045, xend=1.3, yend=0.01, arrow = arrow(length = unit(0.3, "cm")), linewidth=1) +
+  annotate("segment", x = 0.1, y = 0.045, xend = 1.3, yend = 0.01, arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
   annotate("text", label="beta", x=0, y=0.05, parse=T, size=8) +
   # alpha arrow
-  annotate("segment", x=4, y=0.043, xend=3.4, yend=0.01, arrow = arrow(length = unit(0.3, "cm")), linewidth=1) +
+  annotate("segment", x = 4, y = 0.043, xend = 3.4, yend = 0.01, arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
   annotate("text", label="frac(alpha,2)", x=4.2, y=0.05, parse=T, size=8) +
   # power arrow
-  annotate("segment", x=6, y=0.2, xend=4.5, yend=0.15, arrow = arrow(length = unit(0.3, "cm")), linewidth=1) +
-  annotate("text", label=expression(paste(1-beta, "  (\"power\")")), x=6.1, y=0.21, parse=T, size=8) +
+  annotate("segment", x = 6, y = 0.2, xend = 4.5, yend = 0.15, arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
+  annotate("text", label = expression(paste(1-beta, "  (\"power\")")), x = 6.1, y = 0.21, parse = TRUE, size = 8) +
   # H_0 title
-  annotate("text", label="H[0]", x=m1, y=0.28, parse=T, size=8) +
+  annotate("text", label = "H[0]", x = m1, y = 0.28, parse = TRUE, size = 8) +
   # H_a title
-  annotate("text", label="H[1]", x=m2, y=0.28, parse=T, size=8) +
+  annotate("text", label = "H[1]", x = m2, y = 0.28, parse = TRUE, size = 8) +
   ggtitle("Statistical Power") +

### 2025-02-27 data-visualization.qmd
20250227 - clean up

@@
     aes(
-      x0 = off_epa ,
+      x0 = off_epa,
       y0 = def_epa)) +

### 2025-02-28 sports-cognitive-psychology.qmd
20250227 - edits
20250227 - clean up
20250227 - update links

@@
 He published a series of Baseball Abstracts, which included players' statistics.
-James published the first Baseball Abstract in 1977 (<https://www.pbs.org/thinktank/transcript1197.html>; archived at <https://www.pbs.org/thinktank/transcript1197.html>).
+James published the first Baseball Abstract in 1977 (<https://www.pbs.org/thinktank/transcript1197.html>; archived at <https://perma.cc/Y7J7-GB9V>).
 He coined the term Sabermetrics (which was originally named SABRmetrics, based on the name of the society) to refer to advanced metrics and statistical/empirical analysis of baseball.
@@
 According to this idea, in the case of failing to successfully convert on fourth down, coaches do not want to have to defend their decision to go for it to the media or the owner or general manager.
+That is, they may often play *not to lose*, rather than to win, in order to keep their job.
 
@@
 
-Adapted from Ben Baldwin: https://www.nfl4th.com/articles/4th-down-research.html (archived at <https://perma.cc/S5D8-3NCU>)
+Adapted from Ben Baldwin: <https://www.nfl4th.com/articles/4th-down-research.html> (archived at <https://perma.cc/S5D8-3NCU>)
 

### 2025-02-28 sports-cognitive-psychology.qmd
20250227 - video and examples

@@
 Moreover, star players rest more games.
+Here is an article on how Shane Battier, who was not the quickest, fastest, or most athletic player, used analytics successfully to guard players who were quicker or more athletic than him: <https://www.nytimes.com/2009/02/15/magazine/15Battier-t.html> (archived at <https://web.archive.org/web/20250219225627/https://www.nytimes.com/2009/02/15/magazine/15Battier-t.html>).
 
@@
 > We rank them at the latter part of the week, and then we get into the stadium and we play it by ear.
-> A lot of it has to do wiht the feel or the flow of the game.
+> A lot of it has to do with the feel or the flow of the game.
 > Maybe what personnel group we think they're going to match our personnel group with.
@@
 > You can take analytics to baseball and things like that but football is always going to be football.
-> I got a lot of respect for analytics and numbers, but I'm not going to make judgements based on those numbers.
+> I got a lot of respect for analytics and numbers, but I'm not going to make judgments based on those numbers.
 > The game is the game.
@@
 A player talent evaluator for the NFL noted that [many people around the league believe that the use of analytics is better suited for baseball than football](https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html) because baseball involves more games, players, and one-on-one matchups (archived at <https://web.archive.org/web/20200803205803/https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html>).
-By contrast, player performance in football may be more dependent on teammates and playcalls.
+By contrast, player performance in football may be more dependent on teammates and play calls.
 The [player talent evaluator noted](https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html), "In football they don't know how to use all the numbers yet...They have the data, they have all these different stats but the people I have talked to aren't completely sure how to use it." (archived at <https://web.archive.org/web/20200803205803/https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html>).
+Some dismiss analytics entirely.
 
-Nevertheless, more and more people around the league are using analytics; for examples of articles, see [here](https://www.theringer.com/2018/12/19/nfl/nfl-analytics-revolution) (archived at <https://perma.cc/7BKP-A4GJ>)  and [here](https://www.samford.edu/sports-analytics/fans/2020/The-Future-of-NFL-Data-Analytics) (archived at <https://perma.cc/WXE3-53E6>).
+::: {.content-visible when-format="html:js"}
+
+For instance, here is a video clip of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops."
+
+![Football Commentator Making Fun of Analytics in Football. From: <https://streamable.com/1hw8>.](images/commentatorOnAnalytics.mp4){width=100%}\
+
+:::
+
+Nevertheless, more and more people around the league are using analytics—likely because it works; for examples of articles, see [here](https://www.theringer.com/2018/12/19/nfl/nfl-analytics-revolution) (archived at <https://perma.cc/7BKP-A4GJ>) and [here](https://www.samford.edu/sports-analytics/fans/2020/The-Future-of-NFL-Data-Analytics) (archived at <https://perma.cc/WXE3-53E6>).
 Head Coach Doug Peterson heavily relied on analytics with the Philadelphia Eagles to help them win the 2017 Super Bowl; for more information, see [here](https://www.forbes.com/sites/liamfox/2021/08/12/how-the-nfl-uses-analytics-according-to-the-lead-analyst-of-a-super-bowl-champion/) (archived at <https://perma.cc/49KQ-R785>) and [here](https://www.nfl.com/news/super-bowl-lii-how-the-2017-philadelphia-eagles-were-built-0ap3000000912753) (archived at <https://perma.cc/2GRF-8Z5K>).

### 2025-02-28 sports-cognitive-psychology.qmd
20250227 - updates

@@
 James' ideas were slow to catch on among those in baseball.
-Neverthless, some people eventually caught on to his ideas—and to good success.
+Nevertheless, some people eventually caught on to his ideas—and to good success.
 Billy Beane, a general manager, used Sabermetrics to help the Oakland Athletics, a small market team with a limited budget, better compete with teams with larger budgets.
@@
 In addition to teams using Sabermetrics to evaluate player talent, teams also began to frequently use statistical analysis to inform decision making during games, which led to key changes in the style of play.
-For instace, defensive shifts—where defensive players moved to locations on the field where particular hitters were most likely to hit the ball—became more common, attempts to steal bases became less common, there were fewer bunts, batters took more pitches (i.e., watched more pitches without swinging), there were more frequent pitching changes (for particular pitcher–batter matchups; such as to have a right-handed pitcher face a left-handed batter or vice versa), and a greater focus on velocity and spin rate among pitchers.
+For instance, defensive shifts—where defensive players moved to locations on the field where particular hitters were most likely to hit the ball—became more common, attempts to steal bases became less common, there were fewer bunts, batters took more pitches (i.e., watched more pitches without swinging), there were more frequent pitching changes (for particular pitcher–batter matchups; such as to have a right-handed pitcher face a left-handed batter or vice versa), and a greater focus on velocity and spin rate among pitchers.
 Some of these analytics-driven changes in play style eventually led Major League Baseball (MLB) to make rule changes in an attempt to make the game more exciting to watch, including banning defensive shifts, reducing the number of pitching changes allowed, and making the bases larger and easier to steal.
@@
 Perhaps that is why the Pittsburgh Steelers were voted [one of the least analytically advanced teams](https://www.espn.com/nfl/story/_/id/29939438/2020-nfl-analytics-survey-which-teams-most-least-analytically-inclined) in the NFL in 2020 (archived at <https://perma.cc/R7FA-HGGB>).
+Nevertheless, the [Steelers did eventually hire a person focused on analytics](https://www.espn.com/blog/pittsburgh-steelers/post/_/id/14521/why-the-steelers-hired-a-carnegie-mellon-professor-for-advanced-analytics) (archived at <https://perma.cc/W7VE-JZQX>).
 

### 2025-02-28 sports-cognitive-psychology.qmd
20250228 - updates

@@
 Football has also seen greater use of analytics, although its uptake has been somewhat slower than in many other sports.
-When asked about how he makes the decision about whether to go for two after a touchdown, Steelers Head Coach Mike Tomlin stated:
+When asked about how he makes the decision about whether to go for a two-point conversion after a touchdown, Steelers Head Coach Mike Tomlin stated:
 
@@
 > We rank them at the latter part of the week, and then we get into the stadium and we play it by ear.
-> A lot of it has to do with the feel or the flow of the game.
-> Maybe what personnel group we think they're going to match our personnel group with.
+> A lot of it has to do with the feel or the flow of the game—maybe what personnel group we think they're going to match our personnel group with.
 > As we start to play and work the ball down the field on the drives that produce the touchdowns before the point after, we have a little inclination of what their personality might be at least in terms of matching our personnel.
@@
 > I think that's why you play the game.
-> You can take analytics to baseball and things like that but football is always going to be football.
+> You can take analytics to baseball and things like that, but football is always going to be football.
 > I got a lot of respect for analytics and numbers, but I'm not going to make judgments based on those numbers.
@@
 > That's an element of the game you can't measure.
-> Often times decisions such as that weigh heavily into the equation.
+> Oftentimes, decisions such as that weigh heavily into the equation.
 > 
@@
 
-For instance, here is a video clip of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops."
+For instance, here is a video clip of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops", suggesting that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
 
@@
 
+Rates of going for it on fourth down were low until 2017.
+After the 2017 season, rates of going for it on fourth down increased dramatically, as depicted in @fig-4thDownAttemptsByYear.
+
 ```{r}
+#| label: fig-4thDownAttemptsByYear
+#| fig-cap: "The Proportion of Fourth Downs That are Attempts to go for it (Rather Than Punts or Field Goals)."
+#| fig-alt: "The Proportion of Fourth Downs That are Attempts to go for it (Rather Than Punts or Field Goals)."
+
 nfl_pbp4thDown <- nfl_pbp %>% 
@@
   ) +
+  scale_x_continuous(
+    minor_breaks = seq(1995, 3000, 1),
+    breaks = seq(1995, 3000, 5)
+  ) +
   ggplot2::labs(
@@
   ) +
-  ggplot2::theme_classic()
+  ggplot2::theme_classic() +
+  theme()
 ```
 
-Adapted from Ben Baldwin: <https://www.nfl4th.com/articles/4th-down-research.html> (archived at <https://perma.cc/S5D8-3NCU>)
+As depicted in the @fig-goForItPercentageByWinProbChange (which was adapted from a [figure by Ben Baldwin](https://www.nfl4th.com/articles/4th-down-research.html); archived at <https://perma.cc/S5D8-3NCU>), teams went for it on fourth down about twice as often in 2021 compared to 2017, when teams would increase their win probably by going for it.
 
 ```{r}
+#| label: fig-goForItPercentageByWinProbChange
+#| fig-cap: "Go-For-It Percentage on Fourth Downs in 2021 Versus 2017 as a Function of the Change in Win Probability for Going for it.
+To remove desperation fourth down attempts, the figure focuses on just plays in which the estimated win probability for the offensive team (given the current situation at the start of the given play and incorporating the pre-game Vegas line) was greater than 20%."
+#| fig-alt: "Go-For-It Percentage on Fourth Downs in 2021 Versus 2017 as a Function of the Change in Win Probability for Going for it.
+To remove desperation fourth down attempts, the figure focuses on just plays in which the estimated win probability for the offensive team (given the current situation at the start of the given play and incorporating the pre-game Vegas line) was greater than 20%."
+
 # labels on the plot
 text_df <- tibble(
-  label = c("NFL coaches<br>in <span style='color:#00BFC4'>**2024**</span>", "NFL coaches<br>in <span style='color:#F8766D'>**2014**</span>"),
+  label = c("NFL coaches<br>in <span style='color:#00BFC4'>**2021**</span>", "NFL coaches<br>in <span style='color:#F8766D'>**2017**</span>"),
   x = c(6, 8.2),
@@
 nfl_4thdown %>%
-  filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2014, 2024)) %>%
+  filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2017, 2021)) %>%
   ggplot(
@@
       color = as.factor(season))) + 
-  ggtext::geom_richtext(
-    data = text_df,
-    aes(
-      x,
-      y,
-      label = label,
-      angle = angle),
-    color = "black",
-    fill = NA,
-    label.color = NA,
-    size = 5) + 
   geom_vline(xintercept = 0) +
@@
     alpha = 0) +
+  ggtext::geom_richtext(
+    data = text_df,
+    aes(
+      x,
+      y,
+      label = label,
+      angle = angle),
+    color = "black",
+    fill = NA,
+    label.color = NA,
+    size = 5) + 
   ggplot2::theme_classic() +
@@
     y = "Go-for-it percentage",
-    subtitle = "4th down decisions in 2024 versus 2014, win prob. > 20%",
+    subtitle = "4th down decisions in 2021 versus 2017, win prob. > 20%",
     title = glue::glue("How <span style='color:red'>math</span> is changing football")) +

### 2025-02-28 time-series-analysis.qmd
20250227 - clean up
20250228 - updates

@@
 #| label: fig-timeSeriesHistoricalBradyManning
-#| fig-cap: "Historical Fantasy Points by Game for Tom Brady and Peyone Manning."
-#| fig-alt: "Historical Fantasy Points by Game for Tom Brady and Peyone Manning."
+#| fig-cap: "Historical Fantasy Points by Game for Tom Brady and Peyton Manning."
+#| fig-alt: "Historical Fantasy Points by Game for Tom Brady and Peyton Manning."
 

### 2025-02-28 sports-cognitive-psychology.qmd
20250228 - updates

@@
 
-As depicted in the @fig-goForItPercentageByWinProbChange (which was adapted from a [figure by Ben Baldwin](https://www.nfl4th.com/articles/4th-down-research.html); archived at <https://perma.cc/S5D8-3NCU>), teams went for it on fourth down about twice as often in 2021 compared to 2017, when teams would increase their win probably by going for it.
+As depicted in the @fig-goForItPercentageByWinProbChange (which was adapted from a [figure by Ben Baldwin](https://www.nfl4th.com/articles/4th-down-research.html); archived at <https://perma.cc/S5D8-3NCU>), teams went for it on fourth down about twice as often in 2021 compared to 2017, when when their win probability was at least 20% (given the current situation at the start of the given play) and they would increase their win probably by going for it.
 

### 2025-02-28 sports-cognitive-psychology.qmd
20250228 - update plots

@@
     title = "4th Down Attempts (Proportion) by Season",
-  ) +
-  ggplot2::theme_classic() +
-  theme()
+  )
 ```
@@
   x = c(6, 8.2),
-  y = c(80, 37),
+  y = c(90, 37),
   angle = c(10, 10),
@@
     size = 5) + 
-  ggplot2::theme_classic() +
+  theme_classic() +
   labs(

### 2025-02-28 sports-cognitive-psychology.qmd
20250228 - figure updates

@@
   ) +
-  ggplot2::labs(
+  labs(
     x = "Season",
-    y = "Proportion of 4th Down Plays that are Attempts (to Get the First Down)",
+    y = "Proportion of 4th Down Plays that\nare Attempts (to Get the First Down)",
     title = "4th Down Attempts (Proportion) by Season",
-  )
+  ) +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
 ```
 
-As depicted in the @fig-goForItPercentageByWinProbChange (which was adapted from a [figure by Ben Baldwin](https://www.nfl4th.com/articles/4th-down-research.html); archived at <https://perma.cc/S5D8-3NCU>), teams went for it on fourth down about twice as often in 2021 compared to 2017, when when their win probability was at least 20% (given the current situation at the start of the given play) and they would increase their win probably by going for it.
+As depicted in the @fig-goForItPercentageByWinProbChange (which was adapted from a [figure by Ben Baldwin](https://www.nfl4th.com/articles/4th-down-research.html); archived at <https://perma.cc/S5D8-3NCU>), teams went for it on fourth down nearly twice as often in 2021 compared to 2017, when when their win probability was at least 20% (given the current situation at the start of the given play) and they would increase their win probably by going for it.
 
@@
     axis.title.x = element_text(size = 14, face = "bold"),
-    axis.title.y = element_text(size = 14, face = "bold")
+    axis.title.y = element_text(size = 14, face = "bold", angle = 0, vjust = 0.5) # horizontal y-axis title
   ) +

### 2025-02-28 sports-cognitive-psychology.qmd
20250228 - clean up code and reduce warnings

@@
   scale_x_continuous(
-    minor_breaks = seq(1995, 3000, 1),
-    breaks = seq(1995, 3000, 5)
+    minor_breaks = seq(0, 3000, 1),
+    breaks = seq(0, 3000, 5)
   ) +
@@
 text_df <- tibble(
-  label = c("NFL coaches<br>in <span style='color:#00BFC4'>**2021**</span>", "NFL coaches<br>in <span style='color:#F8766D'>**2017**</span>"),
+  label = c(
+    "NFL coaches<br>in <span style='color:#00BFC4'>**2021**</span>",
+    "NFL coaches<br>in <span style='color:#F8766D'>**2017**</span>"),
   x = c(6, 8.2),
@@
 nfl_4thdown %>%
-  filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2017, 2021)) %>%
+  filter(
+    vegas_wp > .2,
+    between(go_boost, -10, 10),
+    season %in% c(2017, 2021)) %>%
   ggplot(
@@
     legend.position = "none",
-    plot.title = element_markdown(size = 22, hjust = 0.5),
-    plot.subtitle = element_markdown(size = 14, hjust = 0.5),
-    axis.title.x = element_text(size = 14, face = "bold"),
-    axis.title.y = element_text(size = 14, face = "bold", angle = 0, vjust = 0.5) # horizontal y-axis title
+    plot.title = element_markdown(
+      size = 22,
+      hjust = 0.5),
+    plot.subtitle = element_markdown(
+      size = 14,
+      hjust = 0.5),
+    axis.title.x = element_text(
+      size = 14,
+      face = "bold"),
+    axis.title.y = element_text(
+      size = 14,
+      face = "bold",
+      angle = 0, # horizontal y-axis title
+      vjust = 0.5)
   ) +
@@
     limits = c(0, 100),
-    expand = c(0,0)) +
+    expand = c(0, 0)) +
   scale_x_continuous(
@@
     limits = c(-10, 10),
-    expand = c(0,0)) +
+    expand = c(0, 0)) +
   annotate(
@@
     size = 5) +
-  geom_segment(
-    aes(
+  annotate(
+    "segment",
     x = -.1,
@@
     xend = -2,
-      yend = 80),
+    yend = 80,
     arrow = arrow(length = unit(0.05, "npc")),
@@
     linewidth = 2) +
-  geom_segment(
-    aes(
+  annotate(
+    "segment",
     x = .1,
@@
     xend = 2,
-      yend = 80),
+    yend = 80,
     arrow = arrow(length = unit(0.05, "npc")),

### 2025-03-01 sports-cognitive-psychology.qmd
20250228 - plot tweaks

@@
     "NFL coaches<br>in <span style='color:#F8766D'>**2017**</span>"),
-  x = c(6, 8.2),
-  y = c(90, 37),
+  x = c(5, 7),
+  y = c(90, 20),
   angle = c(10, 10),
@@
     plot.subtitle = element_markdown(
-      size = 14,
+      size = 13.5,
       hjust = 0.5),
@@
     "text",
-    x = -1.2,
+    x = -1.5,
     y = 70,
@@
     "text",
-    x = 1.2,
+    x = 1.75,
     y = 70,

### 2025-03-01 sports-cognitive-psychology.qmd
20250228 - plot annotation

@@
   ) +
-  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) + # horizontal y-axis title
+  annotate(
+    "segment",
+    x = 2015,
+    xend = 2017,
+    y = 0.08,
+    yend = 0.123,
+    color = "blue",
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow()) +
+  annotate(
+    "text",
+    x = 2015,
+    y = 0.07,
+    label = "Rapid increases in 4th down\nattempts after the 2017 season",
+    hjust = 0.5) # center
 ```
@@
   x = c(5, 7),
-  y = c(90, 20),
+  y = c(87.5, 22.5),
   angle = c(10, 10),

### 2025-03-01 cognitive-bias.qmd
20250227 - clean up
20250301 - punting on 4th down

@@
 That is, people tend to give as much twice the weight to losses as to gains.
+The late baseball manager Sparky Anderson once said, "Losing hurts twice as bad as winning feels good."
 Loss aversion is different from [risk aversion](#sec-cognitiveBiasesRiskAversion).
@@
 Loss aversion can also influence trade negotiations.
+
+An example of loss aversion bias in coaching basketball is removing a star player from a game so you can keep the player for the end of the game [@Moskowitz2011].
+As described in @sec-coachingGoingForIt, an example of loss aversion bias in football is punting on fourth down when, in many cases, it would be advantageous to go for it [@Moskowitz2011].
 

### 2025-03-01 sports-cognitive-psychology.qmd
20250301 - punting on 4th down

@@
 
+### Going for It on Fourth Down {#sec-coachingGoingForIt}
+
 It had been known for a long time that going for it on fourth down would frequently increase a team's chances of winning.
@@
 It is curious that there was such a discrepancy between the decisions that would maximize teams' winning percentage and the decisions coaches actually made.
+In many fourth down situations, by punting the ball, coaches actively and systematically made decisions that reduced their team's chances of winning.
 One potential explanation for the discrepancy is because of coaches' risk aversion.
@@
 That is, they may often play *not to lose*, rather than to win, in order to keep their job.
+However, in many cases, this is the wrong choice [@Moskowitz2011].
 
-Rates of going for it on fourth down were low until 2017.
+Rates of going for it on fourth down in the NFL were low until 2017.
 After the 2017 season, rates of going for it on fourth down increased dramatically, as depicted in @fig-4thDownAttemptsByYear.
@@
 #| label: fig-4thDownAttemptsByYear
-#| fig-cap: "The Proportion of Fourth Downs That are Attempts to go for it (Rather Than Punts or Field Goals)."
-#| fig-alt: "The Proportion of Fourth Downs That are Attempts to go for it (Rather Than Punts or Field Goals)."
+#| fig-cap: "The Proportion of Fourth Downs That are Attempts to Go for it (Rather Than Punts or Field Goals)."
+#| fig-alt: "The Proportion of Fourth Downs That are Attempts to Go for it (Rather Than Punts or Field Goals)."
 
@@
 
+NFL teams were later to implement analytics into their decision making than even teams at lower levels of the sport.
+Some high school teams, such as the Pulaski Academy Bruins (Little Rock, AR), coached by Kevin Kelley, consistently went for it on fourth down well before 2017 [@Moskowitz2011].
+
+> [Bruins Head Football Coach Kevin] Kelley believes that the "quant jocks" don't go far enough to validate the no-punting worldview and, more generally, the virtues of risk-taking.
+> "The math guys, the astrophysicist guys, they just do the raw numbers and they don't figure emotion into it—and that's the biggest thing of all," he says.
+> "The built-in emotion involved in football is unbelievable, and that's where the benefits really pay off."
+> What he means is this: A defense that stops an opponent on third down is usually ecstatic.
+> They've done their job.
+> The punting unit comes on, and the offense takes over.
+> When that defense instead gives up a fourth-down conversion, it has a hugely deflating effect.
+> At Pulaski's games, you can see the shoulders of the opposing defensive players slump and their eyes look down when they fail to stop the Bruins on fourth down.
+> 
+> [@Moskowitz2011, p. 37]
+
+Based on Romer's [-@Romer2006] analysis of third down plays (because few teams went for it on fourth down), focusing on plays in the first quarter (to remove desperation plays), he identified several general conclusions about fourth-down situations:
+
+1. Inside the opponent's 45-yard line, a team is better off going for it than punting with 6 (or less) yards to go (for a first down).
+1. Inside the opponent's 33-yard line, a team is better off going for it with 10 (or less) yards to go (unless little time remains and a field goal would decide the game).
+1. Once reaching the opponent's 5-yard line, a team is better off going for it.
+1. Regardless of field position, a team is always better off going for it with 3 (or less) yards to go.
+
+Nevertheless, out of the fourth down plays @Romer2006 identified between 1998–2000 in which it would have been advantageous to go for it, the team made the suboptimal decision (punting or kicking) 90% of the time.
+"Inasmuch as an academic paper can become a cult hit, Romer's made the rounds in NFL executive offices, but most NFL coaches seemed to dismiss his findings as the handiwork of an egghead, polluting art with science." [@Moskowitz2011, p.
+39].
+
+Here is an analysis by the New York Times on when to go for it on fourth down: <https://www.nytimes.com/2014/09/05/upshot/4th-down-when-to-go-for-it-and-why.html> (archived at <https://perma.cc/KA9Y-BRUD>).
+Here is a fourth down calculator: <https://rbsdm.com/stats/fourth_calculator/>.
+
+::: {#fig-4thDownAnalysis}
+![](images/4thDownAnalysis.png){fig-alt="Analysis of When to Go for it on Fourth Down Versus What Coaches Actually Do on 4th Down (From 2014)."}
+
+Analysis of When to Go for it on Fourth Down Versus What Coaches Actually Do on 4th Down (From 2014).
+Retrieved from <https://www.nytimes.com/2014/09/05/upshot/4th-down-when-to-go-for-it-and-why.html> (archived at <https://perma.cc/KA9Y-BRUD>).
+:::
+
+## Other Coaching Absurdities {#sec-coachingAbsurdities}
+
+### Icing the Kicker {#sec-icingTheKicker}
+
 ## Conclusion {#sec-sportsCognitivePsychologyConclusion}

### 2025-03-01 sports-cognitive-psychology.qmd
20250301 - icing the kicker

@@
 
-## Coaching and Risk Aversion {#sec-coachingRiskAversion}
+## Coaching and Loss/Risk Aversion {#sec-coachingRiskAversion}
 
@@
 
+In a 2008 game between the Jets and Raiders, Jets Kicker Jay Feely missed a game-tying field goal to send the game to overtime.
+However, moments before the snap, the Raiders called a timeout in an attempt to "ice the kicker."
+The idea behind icing the kicker is making the kicker think about the pressure of upcoming kick so he gets "cold feet" (metaphorically speaking) and misses the kick.
+After the brief timeout, Feely got a second chance and made the field goal to send the game into overtime.
+
+After the game, [Feely noted](https://www.nfl.com/news/janikowski-gives-raiders-win-over-jets-in-overtime-09000d5d80bc3910), "I heard the whistle before I started, which is an advantage to the kicker...If you're going to do that, do that before he kicks.
+I can kick it down the middle, see what the wind does and adjust.
+It helps the kicker tremendously." (archived at <https://perma.cc/U7SV-UE2E>).
+
+However, does icing the kicker work?
+There is mixed evidence on icing the kicker.
+Some studies find some evidence for lower percentage of iced kicks made compared to non-iced kicks [@Goldschmied2024], whereas other studies do not [e.g., @GonzalezSanchez2024; @Moskowitz2011].
+This suggests that, if there is such an effect, the [effect size](#sec-practicalSignificance) is likely very small, and it may better suit the team to use your timeouts in other situations.
+
 ## Conclusion {#sec-sportsCognitivePsychologyConclusion}

### 2025-03-01 cognitive-bias.qmd
20250301 - hot hand fallacy

@@
 The "hot hand" is the idea that a player who experiences a successful outcome will have greater chance of success in subsequent attempts.
-For instance, in basketball, it is widely claimed by coaches, players, and commentators that players who have the hot hand (i.e., who are "on fire") are more likely to make shots because they made previous shots.
+For instance, in basketball, it is widely claimed by coaches, players, and commentators that players who have the hot hand (i.e., who are "on fire") are more likely to make shots because they made previous shots and are gaining confidence.
+The hot hand is a supposed example of momentum.
+However, there is not strong evidence of momentum in sports either at the player or team level; rather, there may be evidence of reversals, the opposite of momentum [@Moskowitz2011].
+Streaks certainly exist, but momentum suggests that if a person performs well (or poorly) in a given instance, their performance will influence (and thus predict) their future performance.
+In general, streaks appear to be largely random variation (i.e., luck/chance) around a player's mean performance ability.
+
 Evidence on the hot hand is mixed.
-Considerable evidence historically has suggested that there is no such thing as a "hot hand" [@Gilovich1985; @BarEli2006; @Avugos2013].
+Considerable evidence historically has suggested that there is no such thing as a "hot hand" even though many players and fans believe in it [@Gilovich1985; @BarEli2006; @Moskowitz2011; @Avugos2013; @Wetzels2016].
+Even though streaks may occur, making a shot may not increase the likelihood of making the next shot; that is, the likelihood of making a field goal seems to be independent of past success (or failure) on recent attempts.
+Defenses may adjust to guard a "hot" player more closely.
+However, the findings suggesting no hot hand tend to hold even whether there is no defense [@Gilovich1985].
 Some recent research, however, has suggested that there may be a small hot hand effect in some contexts [@Bocskocsky2014; @Miller2014a].
@@
 The gambler's fallacy occurs due to an erroneous belief in the law of small numbers.
-The law of large numbers is a mathematical theorem that the average of a sufficiently large number of independent observations converges to the true value.
+The law of *large* numbers is a mathematical theorem that the average of a sufficiently large number of independent observations converges to the true value.
 For instance, if you flip a fair coin 1 million times, it is likely to land heads-up ~50% of the time.
@@
 If the slot machine has not paid out in a while, the gambler may believe that the slot machine is about to pay out soon, and may start putting coins in the slots.
+Or, a gambler on a losing streak may (wrongly) believe that they "are due" and keeps gambling as a result.
 

### 2025-03-01 time-series-analysis.qmd
20250301 - fix typo

@@
     y = "Fantasy Points (Season)",
-    title = "Fantasy Points (Season) by Player Age: Wide Receivers"
+    title = "Fantasy Points (Season) by Player Age: Tight Ends"
   ) +

### 2025-03-02 mixed-models.qmd
20250302 - Bayesian model object

@@
 ::: {#nte-bayesianMixedModel .callout-note title="Bayesian mixed model"}
-Note: the following code takes a while to run.
+Note: the following code that runs the model takes a while.
+If you just want to save time and load the model object instead of running the model, you can load the model object (which has already been fit) using this code:
+
+```{r}
+load(url("https://osf.io/download/q6rjf/"))
+```
 :::

### 2025-03-02 time-series-analysis.qmd
20250302 - Bayesian model object

@@
 load(file = "./data/player_stats_seasonal.RData")
-load(url("https://osf.io/download/q6rjf/"))
+```
+
+The following code loads the Bayesian model object that was fit in @sec-fantasyPointsByAgeExperienceModelsBayesian.
+
+```{r}
+load(url("https://osf.io/download/q6rjf/")) # Bayesian model object
 ```

### 2025-03-04 sports-cognitive-psychology.qmd
20250301 - hot hand fallacy
20250304 - cites

@@
 
-## Other Coaching Absurdities {#sec-coachingAbsurdities}
+## Other Beliefs in Sports {#sec-beliefsInSports}
 
@@
 There is mixed evidence on icing the kicker.
-Some studies find some evidence for lower percentage of iced kicks made compared to non-iced kicks [@Goldschmied2024], whereas other studies do not [e.g., @GonzalezSanchez2024; @Moskowitz2011].
+Some studies find some evidence for lower percentage of iced kicks made compared to non-iced kicks [@Goldschmied2024], whereas other studies do not [e.g., @Berry2004; @GonzalezSanchez2024; @Moskowitz2011].
 This suggests that, if there is such an effect, the [effect size](#sec-practicalSignificance) is likely very small, and it may better suit the team to use your timeouts in other situations.
+
+### Hot Hand {#sec-hotHand}
+
+The belief in the "hot hand" is described in @sec-fallaciesHotHand.
 

### 2025-03-08 causal-inference.qmd
20250220 - copyediting
20250227 - clean up
20250308 - R packages

@@
 The use of [causal diagrams](#sec-causalDiagrams) can inform which variables are important to be included as control variables, and—just as important—which variables not to include as control variables, as described in @sec-causalDiagrams.
-Another approach that is sometimes used for matching, as a sample weight, or with stratification is the use of propensity scores.
+Another approach that is sometimes used to control for variables—either by using matching, or as a sample weight, or with stratification—is the use of propensity scores.
 
@@
 For instance, from a DAG, it is possible to determine what variables it is important to control for (and how)—and just as importantly, what not to control for—in order to get unbiased estimates of the effect of one variable on another variable or of the association between two variables of interest.
-To create DAGs, you can use the `R` package `dagitty` [@Textor2017] or the associated browser-based extension, DAGitty: <https://dagitty.net> (archived at <https://perma.cc/U9BY-VZE2>).
+To create DAGs, you can use the `R` package `dagitty` [@R-daggity; @Textor2016_packages] or the associated browser-based extension, DAGitty: <https://dagitty.net> (archived at <https://perma.cc/U9BY-VZE2>).
 Examples of various causal diagrams that could explain why `X` is associated with `Y` are in Figures [-@fig-XCausesY], [-@fig-YCausesX] and [-@fig-ZCausesXandY].
@@
 
-Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
+Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax) using the `ggdag` package [@R-ggdag]:
 

### 2025-03-08 cluster-analysis.qmd
20250308 - R packages

@@
 An overview of approaches to cluster analysis in `R` is provided by @Kassambara2017a.
-In this chapter, we focus on examples using model-based clustering with the `mclust` package [@Scrucca2023], which uses Gaussian finite mixture modeling.
+In this chapter, we focus on examples using model-based clustering with the `R` package `mclust` [@R-mclust; @Scrucca2023_packages], which uses Gaussian finite mixture modeling.
 The various types of `mclust` models are provided here:

### 2025-03-08 data-visualization.qmd
20250308 - R packages

@@
 In this chapter, we will examine how to create statistical graphics to visualize data.
-We will create the plots using the `ggplot2` package.
+We will create the plots using the `ggplot2` package [@R-ggplot2; @Wickham2016_packages].
 When creating plots in `ggplot2` with multiple points or lines (e.g., multiple players or levels of a predictor variable), it is easiest to do so with the data in [long form](#sec-wideToLong) (as opposed to [wide form](#sec-longToWide)).
@@
 It is also important to use color schemes with distinguishable colors.
-Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package, as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
+Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
 There are a variety of resources for color schemes that are accessible to color-blind viewers:
@@
 - <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> (archived at <https://perma.cc/7VTA-Y8YS>)
+- the `viridis` package [@R-viridis; @Garnier2024a_packages]
   - <https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html> (archived at <https://perma.cc/NK9K-HL7L>)
@@
 
-A ridgeline plot of fantasy points is depicted in @fig-ridgeline1.
+A ridgeline plot of fantasy points is depicted in @fig-ridgeline1 using the `ggridges` package [@R-ggridges].
 
@@
 
-Second, we create the base layer of the plot using the `ggplot()` function of the `ggplot2` package, as in @fig-scatterplot1.
+Second, we create the base layer of the plot using the `ggplot()` function of the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot1.
 We specify the data object and the variables in the data object that are associated with the x- and y-axes:
@@
 
-Third, we create a scatterplot using the `geom_point()` function from the `ggplot2` package, as in @fig-scatterplot2:
+Third, we create a scatterplot using the `geom_point()` function from the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot2:
 
@@
 
-After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package, as in @fig-scatterplot11.
+After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages], as in @fig-scatterplot11.
 
@@
 
-After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package.
+After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages].
 
@@
 
-We can also add a marginal density/histogram, as in @fig-scatterplot12:
+We can also add a marginal density/histogram, as in @fig-scatterplot12 using the `ggExtra` package [@R-ggExtra]:
 
@@
 ### With Highlighting {#sec-lineChartHighlighting}
+
+We use the `gghighlight` package [@R-gghighlight] to highlight particular elements:
 
@@
 
-Defensive EPA per play during the 2024 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA.
+Defensive EPA per play during the 2024 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA using the `nflplotR` package [@R-nflplotR].
 

### 2025-03-08 download-football-data.qmd
20250227 - clean up
20250308 - R packages

@@
 
-Below, we provide examples for how to download various types of NFL data using the `nflreadr` package.
+Below, we provide examples for how to download various types of NFL data using the `nflreadr` package [@R-nflreadr].
 For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
 We save each data file after downloading it, so we can use the data in subsequent chapters.
-If you have difficulty downloading the data files using the `nflreadr` package, we also saved the data files so they are publicly available on the Open Science Framework: <https://osf.io/z6pg4>.
+If you have difficulty downloading the data files using the `nflreadr` package [@R-nflreadr], we also saved the data files so they are publicly available on the Open Science Framework: <https://osf.io/z6pg4>.
 
@@
 
-To download play-by-play data from prior weeks and seasons, we can use the `load_pbp()` function of the `nflreadr` package.
-We add a progress bar using the `with_progress()` function from the `progressr` package because it takes a while to run.
+To download play-by-play data from prior weeks and seasons, we can use the `load_pbp()` function of the `nflreadr` package [@R-nflreadr].
+We add a progress bar using the `with_progress()` function from the `progressr` package [@R-progressr] because it takes a while to run.
 A Data Dictionary for the play-by-play data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>
@@
 
-We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package.
+We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
 A Data Dictionary for statistics for offensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
@@
 
-First, we can compute the players' career statistics using the `calculate_stats()` function from the `nflfastR` package.
+First, we can compute the players' career statistics using the `calculate_stats()` function from the `nflfastR` package [@R-nflfastR].
 
@@
 
-First, we can compute the players' career statistics using the `calculate_player_stats()`, `calculate_player_stats_def()`, and `calculate_player_stats_kicking()` functions from the `nflfastR` package for offensive players, defensive players, and kickers, respectively.
+First, we can compute the players' career statistics using the `calculate_player_stats()`, `calculate_player_stats_def()`, and `calculate_player_stats_kicking()` functions from the `nflfastR` package [@R-nflfastR] for offensive players, defensive players, and kickers, respectively.
 
@@
 
-- Decide whether to use `calculate_stats()` from `nflfastR` or `load_player_stats()` from `nflreadr`.
+- Decide whether to use `calculate_stats()` from `nflfastR` [@R-nflfastR] or `load_player_stats()` from `nflreadr` [@R-nflreadr].
 - Save/update data file in repo with the data generated from the relevant code, and follow the data file through the rest of the book, updating code as necessary.
@@
 
+We calculate the player's age based on the difference between dates using the `lubridate` package [@R-lubridate]:
+
 ```{r}

### 2025-03-08 getting-started.qmd
20250226 - reprex do's and don'ts
20250227 - clean up
20250308 - R packages

@@
 
-The book uses `R` for statistical analyses (<http://www.r-project.org>).
+The book uses the software `R` [@R-base] for statistical analyses (<http://www.r-project.org>).
 R is a free software environment; you can download it at no charge here: <https://cran.r-project.org>.
@@
 - Learn `R` interactively with `swirl`: <https://swirlstats.com>
-- Use the `learnr` package: <https://rstudio.github.io/learnr/>
-- Resources for learning `tidyverse`, which is a collection of `R` packages for data management: <https://www.tidyverse.org/learn/>
+- Use the `learnr` package [@R-learnr]: <https://rstudio.github.io/learnr/>
+- Resources for learning `tidyverse` [@R-tidyverse; @Wickham2019_packages], which is a collection of `R` packages for data management: <https://www.tidyverse.org/learn/>
 - You will sometimes find relevant articles on `R-bloggers`: <https://www.r-bloggers.com>
@@
 My strong recommendation is to provide a `reprex` whenever possible: <https://reprex.tidyverse.org>.
+Here are Reprex do's and don'ts: <https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html>.
 
@@
    ```
-1. Some necessary packages, including the `ffanalytics` package, are hosted in GitHub and need to be installed using the following code (after installing the `remotes` package above):
+1.
+Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub and need to be installed using the following code (after installing the `remotes` package [@R-remotes] above):
    ```{r}
@@
 
-For instance, you can use the following code to install the `tidyverse` package:
+For instance, you can use the following code to install the `tidyverse` package [@R-tidyverse]:
 

### 2025-03-08 index.qmd
20250308 - R packages

@@
 packagesForBibliography <- c(
-  "petersenlab", "ffanalytics", "pROC", "rms"
+  "petersenlab", "ffanalytics",
+  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
+  "nflreadr", "nflplotR", "nflfastR", "gsisdecoder", "progressr", "DescTools",
+  "pwr", "pwrss", "WebPower", "XICOR", "dagitty", "ggdag", "lubridate",
+  "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", "pROC", "lme4",
+  "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", "AICcmodavg", "rstan",
+  "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", "viridis", "plotly",
+  "quantmod", "fPortfolio", "xts", "zoo", "forecast", "stringi", #"webshot2", 
+  "parallel", "parallelly", "ggridges", "cmdstanr",
+  "learnr", "remotes" #"mgcv",
+
 )

### 2025-03-08 intro.qmd
20240915 - videos
20241016 - render the figure caption as alt text
20250308 - R packages

@@
 
-The book provides data analysis examples using the statistical analysis software, `R`.
+The book provides data analysis examples using the statistical analysis software, `R` [@R-base].
 Why `R`?
@@
 - `R` has advanced statistics capabilities.
-It was designed for statistical analysis and has strong capabalities for data wrangling.
+It was designed for statistical analysis and has strong capabilities for data wrangling.
 - `R` has capabilities for state-of-the-art graphics.
@@
 
-![Video of Prof. Benjamin Motz Discussing Statistics and His Fantasy Football Course.](images/motzVideo.mp4){width=100%}
+![Video of Prof. Benjamin Motz Discussing Statistics and His Fantasy Football Course. From: <https://www.facebook.com/watch/?v=10155572257183615>.](images/motzVideo.mp4){width=100%}\
 

### 2025-03-08 mixed-models.qmd
20250308 - R packages

@@
 
+We can estimate model fit using the `MuMIn` package [@R-MuMIn]:
+
 ```{r}
@@
 
+We compare models using the `bbmle` [@R-bbmle] and `MuMIn` [@R-MuMIn] packages.
+
 ```{r}
@@
 #### Random Intercepts Model {#sec-fantasyPointsByAgeExperienceModelsIntercepts}
+
+We estimate the multilevel models using the `lmerTest` package [@R-lmerTest], which is an extension of the `lme4` package [@R-lme4; @Bates2015_packages]:
 
@@
 #### Random Intercepts Model with Position as Fixed-Effect Predictor {#sec-fantasyPointsByAgeExperienceModelsPosition}
+
+We use the `emmeans` package [@R-emmeans] to determine the model-implied number of points per season by position.
 
@@
 
+We compute model fit using the using the `AICcmodavg` [@R-AICcmodavg], `bbmle` [@R-bbmle], and `MuMIn` [@R-MuMIn] packages
+
 ```{r}
@@
 
+We can use the `fitdistrplus` package [@R-fitdistrplus; @DelignetteMuller2015_packages]:
+
 ```{r}
@@
 
-Information about smooth terms in the `mgcv` package is provided at the following link: <https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/smooth.terms.html>.
+Information about smooth terms in the `mgcv` package [@R-mgcv; @Wood2017] is provided at the following link: <https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/smooth.terms.html>.
+We estimate the model using the `brms` package [@R-brms; @Buerkner2017_packages; @Buerkner2018_packages], which provides an interface to the `rstan` package [@R-rstan].
 
@@
 :::
+
+We use the `cmdstanr` backend [@R-cmdstanr] and threading using the `parellely` package [@R-parallelly] for parallel (faster) processing:
 

### 2025-03-08 download-football-data.qmd
20250308 - update citations

@@
 Data Dictionaries are metadata that describe the meaning of the variables in a dataset.
-You can find Data Dictionaries for the various National Football League (NFL) datasets at the following link: <https://nflreadr.nflverse.com/articles/index.html>.
+@Ho2025 provide Data Dictionaries for the various National Football League (NFL) datasets at the following link: <https://nflreadr.nflverse.com/articles/index.html>.
 
@@
 See @sec-merging for a reminder of how to perform merging, the types of merges, and what you can expect when you merge data objects with different formats.
-Guidance for how to merge the various NFL-related data files is provided at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md>.
+Guidance for how to merge the various NFL-related data files is provided by @Sharpe2020: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md>.
 

### 2025-03-08 fantasy-football.qmd
20241017 - fix width of image of field?
20241018 - figure sizing
20250308 - update citations

@@
 
-Football is the most widely watched sport in the United States.^[<https://news.gallup.com/poll/610046/football-retains-dominant-position-favorite-sport.aspx> (archived at <https://perma.cc/X2UG-RAAK>); <https://www.statista.com/statistics/1430289/most-watched-sports-leagues-usa/> (archived at <https://perma.cc/JNU6-S96A>)]
+Football is the most widely watched sport in the United States [archived at <https://perma.cc/X2UG-RAAK>; @Jones2024; archived at <https://perma.cc/JNU6-S96A>; @statistica2023].
 
@@
 
-Here is a video that provides a brief introduction to American football:
+Here is a video that provides a brief introduction to American football [@Ly2015]:
 
@@
 The yard markers depict the distance from the nearest end zone.
-The orange shaded area is called the "red zone", where chances of scoring points are highest.
+The orange shaded area is called the 'red zone', where chances of scoring points are highest.
 The original figure was modified to depict field goal posts.
-(Figure retrieved from <https://commons.wikimedia.org/wiki/File:American_football_field.svg>"}
+(Figure retrieved from <https://commons.wikimedia.org/wiki/File:American_football_field.svg>)"}
 
@@
 Fantasy football is one of the most widely played games in the history of games.
-It is estimated that around 62 million people play fantasy sports^[<https://thefsga.org/industry-demographics/> (archived at <https://perma.cc/9PB8-ZDJJ>)], of whom around 29 million play fantasy football.^[<https://www.statista.com/topics/10895/fantasy-sports-in-the-us/> (archived at <https://perma.cc/8YSN-UUNT>)]
+It is estimated that around 62 million people play fantasy sports [archived at <https://perma.cc/9PB8-ZDJJ>; @FSGA2023], of whom around 29 million play fantasy football [archived at <https://perma.cc/8YSN-UUNT>; @statistica2023a].
 As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available for [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]

### 2025-03-08 getting-started.qmd
20250308 - update citations

@@
 
-The following article provides additional resources and good guidance: <https://www.r-bloggers.com/where-to-get-help-with-your-r-question/> (archived at <https://perma.cc/4RRE-KL33>).
+@Salmon2018 provides additional resources and good guidance for getting help with `R`: <https://masalmon.eu/2018/07/22/wheretogethelp/> (archived at <https://perma.cc/4RRE-KL33>).
 
@@
 By going to the trouble of creating a minimal, reproducible example and identifying the minimum conditions necessary to reproduce the issue, you will often figure out how to resolve it.
-Here are guidelines on providing a minimal, reproducible example: <https://stackoverflow.com/help/minimal-reproducible-example> (archived at <https://perma.cc/6NUB-UTYF>).
-Here are a good example and guidelines for providing a minimal, reproducible example in `R`: <https://stackoverflow.com/a/5963610> (archived at <https://perma.cc/PC9L-DQZG>).
+@StackOverflow2025 offers guidelines on providing a minimal, reproducible example: <https://stackoverflow.com/help/minimal-reproducible-example> (archived at <https://perma.cc/6NUB-UTYF>).
+@StackOverflow2018 offers a good example and guidelines for providing a minimal, reproducible example in `R`: <https://stackoverflow.com/a/5963610> (archived at <https://perma.cc/PC9L-DQZG>).
 My strong recommendation is to provide a `reprex` whenever possible: <https://reprex.tidyverse.org>.
-Here are Reprex do's and don'ts: <https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html>.
+@Bryan2025 provide Reprex do's and don'ts: <https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html>.
 

### 2025-03-08 index.qmd
20250308 - update citations

@@
 
-I strive to follow [principles of accessibility](https://bookdown.org/yihui/rmarkdown-cookbook/html-accessibility.html) (archived at <https://perma.cc/8XJ9-Q6QJ>) to make the book content accessible to people with visual impairments and physical disabilities.
+I strive to follow [principles of accessibility](https://bookdown.org/yihui/rmarkdown-cookbook/html-accessibility.html) [archived at <https://perma.cc/8XJ9-Q6QJ>; @Xie2024; @Xie2020_packages] to make the book content accessible to people with visual impairments and physical disabilities.
 If there are additional ways I can make the content more accessible, please let me know.

### 2025-03-08 download-football-data.qmd
20250308 - data dictionaries

@@
 Data Dictionaries are metadata that describe the meaning of the variables in a dataset.
-@Ho2025 provide Data Dictionaries for the various National Football League (NFL) datasets at the following link: <https://nflreadr.nflverse.com/articles/index.html>.
+@HoDataDictionaries provide Data Dictionaries for the various National Football League (NFL) datasets at the following link: <https://nflreadr.nflverse.com/articles/index.html>.
 
@@
 See @sec-merging for a reminder of how to perform merging, the types of merges, and what you can expect when you merge data objects with different formats.
-Guidance for how to merge the various NFL-related data files is provided by @Sharpe2020: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md>.
+Guidance for how to merge the various NFL-related data files is provided by @SharpeNFLDataSets: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md>.
 
@@
 
-A Data Dictionary for fantasy player ID data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_playerids.html>
+@HoFFPlayerIDs provide a Data Dictionary for fantasy player ID data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_playerids.html>
 
@@
 
-- <https://nflreadr.nflverse.com/articles/dictionary_rosters.html>
-- <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#rosters>
+- <https://nflreadr.nflverse.com/articles/dictionary_rosters.html> [@HoRosters]
+- <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#rosters> [@SharpeRosters]
 
@@
 
-A Data Dictionary for game schedules data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_schedules.html>
+@HoSchedules provide a Data Dictionary for game schedules data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_schedules.html>
 
@@
 
-A Data Dictionary for game schedules data is located at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#standings>
+@SharpeStandings provides a Data Dictionary for standings data at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#standings>
 
@@
 
-A Data Dictionary for data from the combine is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_combine.html>
+@HoCombine provide a Data Dictionary for data from the combine at the following link: <https://nflreadr.nflverse.com/articles/dictionary_combine.html>
 
@@
 
-A Data Dictionary for draft picks data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_draft_picks.html>
+@HoDraftPicks provide a Data Dictionary for draft picks data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_draft_picks.html>
 
@@
 
-A Data Dictionary for draft values data is located at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#draft_values>
+@SharpeDraftValues provides a Data Dictionary for draft values data at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#draft_values>
 
@@
 
-A Data Dictionary for data from weekly depth charts is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_depth_charts.html>
+@HoDepthCharts provide a Data Dictionary for data from weekly depth charts at the following link: <https://nflreadr.nflverse.com/articles/dictionary_depth_charts.html>
 
@@
 
-A Data Dictionary for game schedules data is located at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#trades>
+@SharpeTrades provides a Data Dictionary for NFL trade data at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#trades>
 
@@
 We add a progress bar using the `with_progress()` function from the `progressr` package [@R-progressr] because it takes a while to run.
-A Data Dictionary for the play-by-play data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>
+@HoPBP provide a Data Dictionary for the play-by-play data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>
 
@@
 
-A Data Dictionary for the participation data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_participation.html>
+@HoParticipation provide a Data Dictionary for the participation data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_participation.html>
 
@@
 We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
-A Data Dictionary for statistics for offensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
-A Data Dictionary for statistics for defensive players is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.
+@HoPlayerStats provide a Data Dictionary for statistics for offensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
+@HoPlayerStatsDefense provide a Data Dictionary for statistics for defensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.
 
@@
 
-A Data Dictionary for injury data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_injuries.html>
+@HoInjuries provide a Data Dictionary for injury data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_injuries.html>
 
@@
 
-A Data Dictionary for snap counts data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_snap_counts.html>
+@HoSnapCounts provide a Data Dictionary for snap counts data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_snap_counts.html>
 
@@
 
-A Data Dictionary for ESPN QBR data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_espn_qbr.html>
+@HoQBR provide a Data Dictionary for ESPN QBR data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_espn_qbr.html>
 
@@
 
-A Data Dictionary for NFL Next Gen Stats data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_nextgen_stats.html>
+@HoNextGenStats provide a Data Dictionary for NFL Next Gen Stats data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_nextgen_stats.html>
 
@@
 
-A Data Dictionary for Pro Football Reference passing data is located at the following links:
-
-- <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html>
-- <https://www.pro-football-reference.com/about/advanced_stats.htm>
-
-Advanced stats from the 2024 season are available at the following link: <https://www.pro-football-reference.com/years/2024/advanced.htm>
+There is a Data Dictionary for Pro Football Reference passing data at the following links:
+
+- <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html> [@HoPFRPassing]
+- <https://www.pro-football-reference.com/about/advanced_stats.htm> [archived at <https://perma.cc/LKH5-92PQ>; @PFR2025]
+
+@PFR2024 provides advanced stats from the 2024 season at the following link: <https://www.pro-football-reference.com/years/2024/advanced.htm>
 
@@
 
-A Data Dictionary for player contracts data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_contracts.html>
+@HoContracts provide a Data Dictionary for player contracts data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_contracts.html>
 
@@
 
-A Data Dictionary for FTN Charting data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ftn_charting.html>
+@HoFTNCharting provide a Data Dictionary for FTN Charting data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ftn_charting.html>
 
@@
 
-A Data Dictionary for FantasyPros ranking data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_rankings.html>
+@HoFantasyProsRankings provide a Data Dictionary for FantasyPros ranking data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_rankings.html>
 
@@
 
-A Data Dictionary for expected fantasy points data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_opportunity.html>
+@HoOpportunity provide a Data Dictionary for expected fantasy points data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_opportunity.html>
 
@@
 Now, modify the scoring settings to match your league settings.
-Below, we use the [scoring settings](https://support.nfl.com/hc/en-us/articles/4989179237404-Scoring) for fantasy leagues on NFL.com, which happen to be point-per-reception leagues (i.e., PPR leagues):
+Below, we use the [scoring settings](https://support.nfl.com/hc/en-us/articles/4989179237404-Scoring) for fantasy leagues on [NFL.com](https://www.nfl.com), which happen to be point-per-reception leagues (i.e., PPR leagues):
 

### 2025-03-09 base-rates.qmd
20250308 - citations

@@
 For instance, the base rate of HIV is the probability of developing HIV.
-In the U.S., [the prevalence rate of HIV is ~0.4% of the adult population](https://map.aidsvu.org/profiles/nation/usa/overview) (archived at <https://perma.cc/8GE6-GAPC>).
+In the U.S., the prevalence rate of HIV is ~0.4% of the adult population [@AIDSVu2022; archived at <https://perma.cc/8GE6-GAPC>].
 
@@
 As noted in @sec-differentErrorsDifferentCosts, lower base-rate events—including touchdowns—are harder to predict accurately.
-As noted by [Harris (2012)](https://www.espn.com/fantasy/football/ffl/story?page=nfldk2k12_vbdwork): "NFL statistical projections are basically impossible to get right.
+As noted by @Harris2012: "NFL statistical projections are basically impossible to get right.
 (Take it from someone who helps create them for a living.) Yes, we can do a passable job with yardage totals for players who don't suffer unexpected injuries or depth-chart pratfalls.
@@
 
-An interactive probability nomogram is available at the following link: <https://jamaevidence.mhmedical.com/data/calculators/LR_nomogram.html> (archived at <https://perma.cc/Z3SW-QMJ3>).
+An interactive probability nomogram is available from @Altarejos2025 at the following link: <https://jamaevidence.mhmedical.com/data/calculators/LR_nomogram.html> (archived at <https://perma.cc/Z3SW-QMJ3>).
 
@@
 
-A probability nomogram calculator can be found at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from Adam Chekroud (<https://github.com/achekroud/nomogrammer>):
+A probability nomogram calculator is available from @Schwartz2006 at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from @Chekroud2017:
 

### 2025-03-09 basic-statistics.qmd
20250308 - R packages
20250308 - citations

@@
 The Hodges-Lehmann statistic (aka pseudomedian) is computed as the median of all pairwise means, and it is also robust to outliers.
+The pseudomedian can be computed using the `DescTools` package [@R-DescTools].
 The pseudomedian of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_pseudomedian`.
@@
 
-A plot of statistical power is in @fig-nhst.
+A plot of statistical power is in @fig-nhst, as adapted from @Magnusson2013 (archived at <https://perma.cc/FG3J-85L6>).
 
@@
 
-- <https://rpsychologist.com/pvalue/> (archived at <https://perma.cc/JP9F-9ZVY>)
-- <https://rpsychologist.com/d3/pdist/> (archived at <https://perma.cc/BE96-8LSJ>)
-- <https://rpsychologist.com/d3/nhst/> (archived at <https://perma.cc/ZU9A-37F3>)
+- <https://rpsychologist.com/pvalue> [@Magnusson2021; archived at <https://perma.cc/JP9F-9ZVY>]
+- <https://rpsychologist.com/d3/pdist> [@Magnusson2015; archived at <https://perma.cc/BE96-8LSJ>]
+- <https://rpsychologist.com/d3/nhst> [@Magnusson2014; archived at <https://perma.cc/ZU9A-37F3>]
 
@@
 
-For an online, interactive statistical decision tree to help you decide which statistical analysis to use, see here: <https://www.statsflowchart.co.uk>
+@JacksonWood2017 provides an online, interactive statistical decision tree to help you decide which statistical analysis to use: <https://www.statsflowchart.co.uk>
 
@@
 Below is `R` code for calculating power for each of various statistical tests (i.e., a *power analysis*).
+We use the `pwr` [@R-pwr] and `pwrss` [@R-pwrss] packages to estimate statistical power.
 For free point-and-click software for calculating statistical power, see G\*Power: <https://www.psychologie.hhu.de/arbeitsgruppen/allgemeine-psychologie-und-arbeitspsychologie/gpower.html>
@@
 
-- <https://aguinis.shinyapps.io/ml_power/>
-- <https://koumurayama.shinyapps.io/tmethod_mlm/>
-- <https://webpower.psychstat.org/wiki/models/index>
+- <https://aguinis.shinyapps.io/ml_power> [@Mathieu2012]
+- <https://koumurayama.shinyapps.io/summary_statistics_based_power> [@Murayama2022]
+- <https://webpower.psychstat.org/wiki/models/index> [@Zhang2018]
 
@@
 
-- <https://sjak.shinyapps.io/power4SEM/>
-- <https://sempower.shinyapps.io/sempower/>
-- <https://yilinandrewang.shinyapps.io/pwrSEM/>
-- <https://webpower.psychstat.org/wiki/models/index>
+- <https://sjak.shinyapps.io/power4SEM> [@Jak2020]
+- <https://sempower.shinyapps.io/sempower> [@Moshagen2024]
+- <https://yilinandrewang.shinyapps.io/pwrSEM> [@Wang2021]
+- <https://webpower.psychstat.org/wiki/models/index> [@Zhang2018]
 
@@
 
-- <https://schoemanna.shinyapps.io/mc_power_med/>
-- <https://www.causalevaluation.org/power-analysis.html> (web application: <https://powerupr.shinyapps.io/index/>)
-- <https://webpower.psychstat.org/wiki/models/index>
+- <https://schoemanna.shinyapps.io/mc_power_med> [@Schoemann2017]
+- <https://www.causalevaluation.org/power-analysis.html> (web application: <https://powerupr.shinyapps.io/index/>) [@Ataneka2023], based on the `PowerUpR` package [@Bulus2021]
+- <https://webpower.psychstat.org/wiki/models/index> [@Zhang2018]
 

### 2025-03-09 causal-inference.qmd
20250308 - citations

@@
 Genetically informed designs exploit differing degrees of genetic relatedness among participants to capture the extent to which genetic factors may contribute to an outcome.
-The average percent of DNA shared between people of varying relationships is provided in @tbl-geneticRelatednessRelativePairs (<https://isogg.org/wiki/Autosomal_DNA_statistics>; archived at <https://perma.cc/MK3D-DST8>):
+The average percent of DNA shared between people of varying relationships is provided by @ISOGG2022 in @tbl-geneticRelatednessRelativePairs (archived at <https://perma.cc/MK3D-DST8>):
 

### 2025-03-09 cognitive-bias.qmd
20250305 - hot hand citations
20250308 - citations

@@
 Consider the following example of the representativeness heuristic.
-When coming out of college, Tight End Kyle Pitts [drew comparisons](https://247sports.com/article/kyle-pitts-lebron-james-2021-nfl-draft-florida-gators-football-163882176) to the "LeBron James" of Tight Ends (archived at <https://perma.cc/JQB5-XPVL>).
+When coming out of college, Tight End Kyle Pitts drew comparisons to the "LeBron James" of Tight Ends [@Nivison2021; archived at <https://perma.cc/JQB5-XPVL>].
 The idea that his athletic profile leads him to be similar to the prototype of LeBron James may have led him to be too highly drafted by fantasy managers in his first seasons.
@@
 
-Here is a video of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends:
+Here is a video from @GetUpESPN2021 of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends:
 
@@
 For example, the base rate of work-related injury is the general probability of experiencing a work-related injury, irrespective of other factors (e.g., the type of job, the person's age, the person's sex).
-Among the working population in the U.S., the lifetime prevalence of work-related injuries (i.e., the percent of people who will experience a work-related injury at some point in their lives), is ~35% (<https://www.cdc.gov/mmwr/volumes/69/wr/mm6913a1.htm>; archived at <https://perma.cc/A2L6-WPEH>).
+Among the working population in the U.S., the lifetime prevalence of work-related injuries (i.e., the percent of people who will experience a work-related injury at some point in their lives), is ~35% [@Free2020; archived at <https://perma.cc/A2L6-WPEH>].
 Thus, the base rate of work-related injuries in the U.S. is ~35%.
@@
 If you select this player early on in the draft based on this information, this would be ignoring the general probability that most players who have a strong performance in the preseason do not perform as well in the regular season (i.e., base rate neglect).
-Performance in the preseason is not strongly predictive of performance in the regular season (<https://fivethirtyeight.com/features/the-nfl-preseason-is-not-predictive-but-it-can-often-seem-that-way/>; archived at <https://perma.cc/FSG2-6AXE>).
+Performance in the preseason is not strongly predictive of performance in the regular season [@Schalter2022; archived at <https://perma.cc/FSG2-6AXE>].
 
@@
 The Sports Illustrated or Madden cover jinx curse is the urban legend that players who appear on the cover of Sports Illustrated (the magazine) or Madden (the video game) will perform poorly.
-But, such a phenomenon can be more simply explained by regression to the mean (<https://www.psychologytoday.com/us/blog/what-the-luck/201610/the-sports-illustrated-cover-jinx>; archived at <https://perma.cc/CZM9-TVFN>) [@Kahneman2011].
+But, such a phenomenon can be more simply explained by regression to the mean [@Kahneman2011; @Smith2016a; archived at <https://perma.cc/CZM9-TVFN>].
 When a player has a superb season, they likely benefited from some degree to good luck, and it is unlikely that they will repeat such a stellar season the following year.
@@
 However, the findings suggesting no hot hand tend to hold even whether there is no defense [@Gilovich1985].
-Some recent research, however, has suggested that there may be a small hot hand effect in some contexts [@Bocskocsky2014; @Miller2014a].
+Some recent research, however, has suggested that there may be a small hot hand effect in some contexts [@Bocskocsky2014; @Miller2014a; @Miller2024].
 However, if any such effect exists, the hot hand may be limited to a small subset of players and the [effect size](#sec-practicalSignificance) of any hot hand effect appears to be small [@Pelechrinis2022].

### 2025-03-09 correlation.qmd
20250308 - R packages
20250308 - citations

@@
 
-Interactive visualizations by Kristoffer Magnusson on *p*-values and null-hypothesis significance testing are below:
-
-- <https://rpsychologist.com/correlation/> (archived at <https://perma.cc/G8YR-VCM4>)
+An interactive visualization by @Magnusson2020 on interpreting correlations is at the following link: <https://rpsychologist.com/correlation/> (archived at <https://perma.cc/G8YR-VCM4>)
 
@@
 ### Nonlinear Correlation {#sec-correlationExamplesNonlinear}
+
+A nonlinear correlation can be estimated using the `XICOR` package [@R-XICOR; @Chatterjee2023_packages]:
 

### 2025-03-09 data-visualization.qmd
20250308 - citations

@@
 When designing graphics, it is important to understand general principles of graphic design.
-An article that describes important principles about graphic design is at the following link: <https://www.adobe.com/express/learn/blog/8-basic-design-principles-to-help-you-create-better-graphics> (archived at <https://perma.cc/29P9-NNSK>).
+@AdobeExpress2020 describes important principles about graphic design at the following link: <https://www.adobe.com/express/learn/blog/8-basic-design-principles-to-help-you-create-better-graphics> (archived at <https://perma.cc/29P9-NNSK>).
 The important principles include:
@@
 In this chapter, we will examine how to create statistical graphics to visualize data.
-We will create the plots using the `ggplot2` package [@R-ggplot2; @Wickham2016_packages].
+We will create the plots using the `ggplot2` package [@R-ggplot2; @ggplot3rdEd].
 When creating plots in `ggplot2` with multiple points or lines (e.g., multiple players or levels of a predictor variable), it is easiest to do so with the data in [long form](#sec-wideToLong) (as opposed to [wide form](#sec-longToWide)).
@@
 
-- <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> (archived at <https://perma.cc/7VTA-Y8YS>)
+- <https://www.datylon.com/blog/data-visualization-for-colorblind-readers> [@Kilin2022; archived at <https://perma.cc/7VTA-Y8YS>]
 - the `viridis` package [@R-viridis; @Garnier2024a_packages]
@@
 
-For ways of visualizing univariate distributions, see [here](https://z3tt.github.io/beyond-bar-and-box-plots) (archived at <https://perma.cc/EEJ8-LND2>).
+@Scherer2021 describes various ways of visualizing univariate distributions; see [here](https://z3tt.github.io/beyond-bar-and-box-plots) (archived at <https://perma.cc/EEJ8-LND2>).
 
@@
 Or, we could use a different theme, such as the dark theme (`theme_dark()`) in @fig-scatterplot10.
-For a list of themes available in `ggplot2`, see here: <https://ggplot2-book.org/themes#sec-themes>.
+For a list of themes available in `ggplot2`, see here: <https://ggplot2-book.org/themes#sec-themes> [@ggplot3rdEd].
 

### 2025-03-09 draft.qmd
20240830 - depth
20250227 - clean up
20250308 - citations

@@
 
-The Spurs in the National Basketball Association (NBA) were well-reputed for [excelling in this draft strategy](https://harvardsportsanalysis.org/2013/11/beating-the-nba-draft-does-any-team-outperform-expectations/) (archived at <https://perma.cc/X7NW-WZC6>).
+The Spurs in the National Basketball Association (NBA) were well-reputed for [excelling in this draft strategy](https://harvardsportsanalysis.org/2013/11/beating-the-nba-draft-does-any-team-outperform-expectations/) [@Ryan2013a; archived at <https://perma.cc/X7NW-WZC6>].
 They frequently used their second-round picks to draft high-risk, high-reward players.
@@
 
-According to [an analysis](https://harvardsportsanalysis.wordpress.com/wp-content/uploads/2012/04/fantasyfootballdraftanalysis1.pdf) by the Harvard Sports Analysis Collective (archived at <https://perma.cc/P7RX-92UU>), the majority of the manager's salary cap should be spent on the starting lineup, and you should spend less on bench players.
+According to [an analysis](https://harvardsportsanalysis.wordpress.com/wp-content/uploads/2012/04/fantasyfootballdraftanalysis1.pdf) by the Harvard Sports Analysis Collective, the majority of the manager's salary cap should be spent on the starting lineup, and you should spend less on bench players [@Chakravarthy2012; archived at <https://perma.cc/P7RX-92UU>].
 This is known as the "stars and scrubs" draft strategy.
@@
 In any case, having some small amount of cap left over toward the end of the draft can help you draft good value players to fill out your bench spots for cheap (e.g., $2).
+Having some depth can help you offset the risk of injuries and the possibility that some of your starters may underperform their expectation, which is quite likely given how [challenging it is to predict fantasy performance](#sec-accuracyOfCrowd).
 

### 2025-03-09 evaluating-prediction-accuracy.qmd
20241105 - equation number
20250308 - R packages
20250308 - citations

@@
 
-- <http://araw.mede.uic.edu/cgi-bin/testcalc.pl>
-- <https://dlrs.shinyapps.io/shinyDLRs>
+- <http://araw.mede.uic.edu/cgi-bin/testcalc.pl> [@Schwartz2006; archived at <https://perma.cc/X8TF-7YBX>]
+- <https://dlrs.shinyapps.io/shinyDLRs> [@Goodman2022]
 
@@
 \end{aligned}
-$$ {#sec-falseOmissionRate}
+$$ {#eq-falseOmissionRate}
 
@@
 
-The adapted formula for MASE with non-time series data is described [here (https://stats.stackexchange.com/a/108963/20338)](https://stats.stackexchange.com/a/108963/20338) (archived at https://perma.cc/G469-8NAJ).
+The adapted formula for MASE with non-time series data is described by @Hyndman2014 [here (https://stats.stackexchange.com/a/108963/20338)](https://stats.stackexchange.com/a/108963/20338) (archived at https://perma.cc/G469-8NAJ).
 Scaled errors are calculated using @eq-scaledError:
@@
 
-Predictive $R^2$ is described here: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>).
+Predictive $R^2$ is described by @Hopper2014 here: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>).
 Predictive $R^2$ penalizes [overfitting](#sec-overfitting), unlike traditional $R^2$.
@@
 Larger values reflect greater [discrimination accuracy](#sec-discrimination).
-[AUC](#auc) was estimated using the `pROC` package [@R-pROC].
+[AUC](#auc) was estimated using the `pROC` package [@R-pROC; @Robin2011_packages].
 
@@
 If we *understand* the variables in the system and how they influence each other, we can predict things more accurately than predicting for the sake of predicting.
-For instance, [we have made great strides in the last decades when it comes to more accurate weather forecasts](https://www.npr.org/sections/money/2023/07/11/1186458991/should-we-invest-more-in-weather-forecasting-it-may-save-your-life) (archived at <https://perma.cc/PF8P-BT3D>), including extreme weather events like hurricanes.
+For instance, [we have made great strides in the last decades when it comes to more accurate weather forecasts](https://www.npr.org/sections/money/2023/07/11/1186458991/should-we-invest-more-in-weather-forecasting-it-may-save-your-life) [@Rosalsky2023; archived at <https://perma.cc/PF8P-BT3D>], including extreme weather events like hurricanes.
 These great strides have more to do with a better causal understanding of the weather system and the ability to conduct simulations of the atmosphere than merely because of big data [@Silver2012].

### 2025-03-09 fantasy-football.qmd
20250308 - citations

@@
 
-Football is the most widely watched sport in the United States [archived at <https://perma.cc/X2UG-RAAK>; @Jones2024; archived at <https://perma.cc/JNU6-S96A>; @statistica2023].
+Football is the most widely watched sport in the United States [@Jones2024; archived at <https://perma.cc/X2UG-RAAK>; @statistica2023; archived at <https://perma.cc/JNU6-S96A>].
 
@@
 
-Here is a video that provides a brief introduction to fantasy football:
+Here is a video that provides a brief introduction to fantasy football [@NFLFantasy2020]:
 

### 2025-03-09 index.qmd
20250308 - citations

--- 2025-01-17 index.qmd
@@
 packagesForBibliography <- c(
-  "petersenlab", "ffanalytics", "pROC", "rms"
+  "petersenlab", "ffanalytics",
+  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
+  "nflreadr", "nflplotR", "nflfastR", "nfl4th", "gsisdecoder", "progressr",
+  "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", "ggdag",
+  "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", "pROC",
+  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", "AICcmodavg",
+  "rstan", "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", "viridis",
+  "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", "stringi", #"webshot2", 
+  "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin",
+  "caret", "gpboost", "cmdstanr", "learnr", "remotes" #"mgcv",
 )

### 2025-03-09 machine-learning.qmd
20250308 - citations

@@
 
+To merge data, we use the `powerjoin` package [@R-powerjoin]:
+
 ```{r}
@@
 However, we impute all data (training and test data together) for purposes of making out-of-sample predictions from the machine learning models to predict players' performance next season (when actuals are not yet available for evaluating their accuracy).
+To impute data, we use the `missRanger` package [@R-missRanger].
 
@@
 
+We use the `caret` package [@R-caret].
+We use the `parallel` [@R-parallel] and `doParallel` [@R-doParallel] packages for parallel (faster) processing.
+
 ```{r}
@@
 
+To combine tree-boosting with mixed models, we use the `gpboost` package [@gpboost].
+
 Adapted from here:
@@
 
-gp_model_qb <- GPModel(
+gp_model_qb <- gpboost::GPModel(
   group_data = data_train_qb_matrix[,"gsis_id"],
@@
 
-gp_data_qb <- gpb.Dataset(
+gp_data_qb <- gpboost::gpb.Dataset(
   data = data_train_qb_matrix[,pred_vars_qb],
@@
 # Find optimal tuning parameters
-opt_params_qb <- gpb.grid.search.tune.parameters(
+opt_params_qb <- gpboost::gpb.grid.search.tune.parameters(
   param_grid = param_grid_qb,
@@
 ```{r}
-gp_model_qb <- GPModel(
+gp_model_qb <- gpboost::GPModel(
   group_data = data_train_qb_matrix[,"gsis_id"],
@@
 ```{r}
-gp_model_fit_qb <- gpb.train(
+gp_model_fit_qb <- gpboost::gpb.train(
   data = gp_data_qb,

### 2025-03-09 modern-portfolio-theory.qmd
20250227 - clean up
20250308 - R packages
20250308 - citations

@@
 But this overperformance in a given year likely reflects more randomness than skill.
-That is likely why [a cat beat professional investors in a stock market challenge](http://www.npr.org/blogs/money/2013/01/14/169326326/housecat-beats-investors-in-stock-market-challenge) (archived at <https://perma.cc/R3XU-K6J8>).
+That is likely why [a cat beat professional investors in a stock market challenge](http://www.npr.org/blogs/money/2013/01/14/169326326/housecat-beats-investors-in-stock-market-challenge) [@Goldstein2013; archived at <https://perma.cc/R3XU-K6J8>].
 That is, "few stock pickers, if any, have the skill needed to beat the market consistently, year after year" [@Kahneman2011, p. 214].
@@
 For the same level of expected returns, higher risk teams might have a few amazing games, but their teams might fall flat in other weeks.
-That is, for a given (high) rate of return, [you are best off in the long run (i.e., over the course of a season) with a lower risk team](https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league/) compared to a higher risk team (archived at <https://perma.cc/NE35-G6LR>).
+That is, for a given (high) rate of return, [you are best off in the long run (i.e., over the course of a season) with a lower risk team](https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league) compared to a higher risk team [@Hitchings2012; archived at <https://perma.cc/NE35-G6LR>].
 
@@
 For instance, the performance of the Tight End and Running Back on the same team tends to be slightly negatively correlated, so it might not be a bad idea to start the Tight End and Running Back from the same team.
-For a correlation matrix of all positions on the team, see: <https://assets-global.website-files.com/5f1af76ed86d6771ad48324b/607a4434a565aa7763bd1312_AndyAsh-Sharpstack-RPpaper.pdf> (archived at <https://perma.cc/JQ6G-KSRT>).
+For a correlation matrix of all positions on the team, see: <https://assets-global.website-files.com/5f1af76ed86d6771ad48324b/607a4434a565aa7763bd1312_AndyAsh-Sharpstack-RPpaper.pdf> [@Sherman2021; archived at <https://perma.cc/JQ6G-KSRT>].
 
@@
 ### Download Historical Stock Prices {#sec-downloadStockPrices}
+
+We download historical stock prices using the `quantmod` package [@R-quantmod]:
 
@@
 ### Create Portfolio {#sec-createPortfolio}
+
+We use the `fPortfolio` package [@R-fPortfolio] to determine the optimal portfolio.
 
@@
 
-<https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league/> (archived at <https://perma.cc/JQ6G-KSRT>)
+<https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league> [@Hitchings2012; archived at <https://perma.cc/JQ6G-KSRT>]
 

### 2025-03-09 mythbusters.qmd
20250308 - citations

@@
 One study found that National Basketball Association (NBA) players improved in field goal percentage, points, and player efficiency rating (but not other statistics: rebounds, assists, steals, or blocks) from their pre-contract year to their contract year, and that Major League Baseball (MLB) players improved in runs batted in (RBIs; but not other statistics: batting average, slugging percentage, on base percentage, home runs, fielding percentage) from their pre-contract year to their contract year [@White2014a].
-Other casual analyses have been examined contract-year performance of National Football League (NFL) players, including articles in [2012](https://www.4for4.com/2012/preseason/2012-contract-year-players-and-myth-increased-production) (archived [here](https://perma.cc/CT3F-QN5E)) and [2022](https://www.4for4.com/2022/preseason/do-players-perform-better-fantasy-football-contract-year) (archived [here](https://perma.cc/F4F5-7RQZ)).
+Other casual analyses have been examined contract-year performance of National Football League (NFL) players, including articles in [2012](https://www.4for4.com/2012/preseason/2012-contract-year-players-and-myth-increased-production) [@Bales2012; archived [here](https://perma.cc/CT3F-QN5E)] and [2022](https://www.4for4.com/2022/preseason/do-players-perform-better-fantasy-football-contract-year) [@Niles2022; archived [here](https://perma.cc/F4F5-7RQZ)].
 

### 2025-03-09 sports-cognitive-psychology.qmd
20250308 - R packages
20250308 - citations

@@
 
-There are a number of [ways in which analytics have changed sports](https://stacker.com/sports/15-ways-analytics-has-changed-sports) (archived at <https://perma.cc/PQ5R-TWFA>).
+There are a number of [ways in which analytics have changed sports](https://stacker.com/sports/15-ways-analytics-has-changed-sports) [@Underwood2019; archived at <https://perma.cc/PQ5R-TWFA>].
 Here is a summary of some of the key ways that analytics has led to changes.
@@
 Moreover, star players rest more games.
-Here is an article on how Shane Battier, who was not the quickest, fastest, or most athletic player, used analytics successfully to guard players who were quicker or more athletic than him: <https://www.nytimes.com/2009/02/15/magazine/15Battier-t.html> (archived at <https://web.archive.org/web/20250219225627/https://www.nytimes.com/2009/02/15/magazine/15Battier-t.html>).
+Here is an article on how Shane Battier, who was not the quickest, fastest, or most athletic player, used analytics successfully to guard players who were quicker or more athletic than him: <https://www.nytimes.com/2009/02/15/magazine/15Battier-t.html> [@Lewis2009b; archived at <https://web.archive.org/web/20250219225627/https://www.nytimes.com/2009/02/15/magazine/15Battier-t.html>].
 
@@
 That is, Mike Tomlin, suggested that he does not take into account analytics when making decisions.
-Perhaps that is why the Pittsburgh Steelers were voted [one of the least analytically advanced teams](https://www.espn.com/nfl/story/_/id/29939438/2020-nfl-analytics-survey-which-teams-most-least-analytically-inclined) in the NFL in 2020 (archived at <https://perma.cc/R7FA-HGGB>).
-Nevertheless, the [Steelers did eventually hire a person focused on analytics](https://www.espn.com/blog/pittsburgh-steelers/post/_/id/14521/why-the-steelers-hired-a-carnegie-mellon-professor-for-advanced-analytics) (archived at <https://perma.cc/W7VE-JZQX>).
+Perhaps that is why the Pittsburgh Steelers were voted [one of the least analytically advanced teams](https://www.espn.com/nfl/story/_/id/29939438/2020-nfl-analytics-survey-which-teams-most-least-analytically-inclined) in the NFL in 2020 [@Walder2020; archived at <https://perma.cc/R7FA-HGGB>].
+Nevertheless, the [Steelers did eventually hire a person focused on analytics](https://www.espn.com/blog/pittsburgh-steelers/post/_/id/14521/why-the-steelers-hired-a-carnegie-mellon-professor-for-advanced-analytics) [@Fowler2015; archived at <https://perma.cc/W7VE-JZQX>].
 
-A player talent evaluator for the NFL noted that [many people around the league believe that the use of analytics is better suited for baseball than football](https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html) because baseball involves more games, players, and one-on-one matchups (archived at <https://web.archive.org/web/20200803205803/https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html>).
+A player talent evaluator for the NFL noted that [many people around the league believe that the use of analytics is better suited for baseball than football](https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html) because baseball involves more games, players, and one-on-one matchups [@Reed2016; archived at <https://web.archive.org/web/20200803205803/https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html>].
 By contrast, player performance in football may be more dependent on teammates and play calls.
-The [player talent evaluator noted](https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html), "In football they don't know how to use all the numbers yet...They have the data, they have all these different stats but the people I have talked to aren't completely sure how to use it." (archived at <https://web.archive.org/web/20200803205803/https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html>).
+The [player talent evaluator noted](https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html), "In football they don't know how to use all the numbers yet...They have the data, they have all these different stats but the people I have talked to aren't completely sure how to use it." [@Reed2016; archived at <https://web.archive.org/web/20200803205803/https://www.cleveland.com/browns/2016/01/in_an_nfl_divided_over_analyti.html>].
 Some dismiss analytics entirely.
@@
 
-For instance, here is a video clip of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops", suggesting that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
+For instance, here is a 2016 video clip of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops", suggesting that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
 
-![Football Commentator Making Fun of Analytics in Football. From: <https://streamable.com/1hw8>.](images/commentatorOnAnalytics.mp4){width=100%}\
+![Football Commentator Making Fun of Analytics in Football. From: <https://streamable.com/1hw8> (2016).](images/commentatorOnAnalytics.mp4){width=100%}\
 
@@
 
-Nevertheless, more and more people around the league are using analytics—likely because it works; for examples of articles, see [here](https://www.theringer.com/2018/12/19/nfl/nfl-analytics-revolution) (archived at <https://perma.cc/7BKP-A4GJ>) and [here](https://www.samford.edu/sports-analytics/fans/2020/The-Future-of-NFL-Data-Analytics) (archived at <https://perma.cc/WXE3-53E6>).
-Head Coach Doug Peterson heavily relied on analytics with the Philadelphia Eagles to help them win the 2017 Super Bowl; for more information, see [here](https://www.forbes.com/sites/liamfox/2021/08/12/how-the-nfl-uses-analytics-according-to-the-lead-analyst-of-a-super-bowl-champion/) (archived at <https://perma.cc/49KQ-R785>) and [here](https://www.nfl.com/news/super-bowl-lii-how-the-2017-philadelphia-eagles-were-built-0ap3000000912753) (archived at <https://perma.cc/2GRF-8Z5K>).
+Nevertheless, more and more people around the league are using analytics—likely because it works; for examples of articles, see [here](https://www.theringer.com/2018/12/19/nfl/nfl-analytics-revolution) [@Clark2018; archived at <https://perma.cc/7BKP-A4GJ>] and [here](https://www.samford.edu/sports-analytics/fans/2020/The-Future-of-NFL-Data-Analytics) [@Awbrey2020; archived at <https://perma.cc/WXE3-53E6>].
+Head Coach Doug Peterson heavily relied on analytics with the Philadelphia Eagles to help them win the 2017 Super Bowl; for more information, see [here](https://www.forbes.com/sites/liamfox/2021/08/12/how-the-nfl-uses-analytics-according-to-the-lead-analyst-of-a-super-bowl-champion/) [@Fox2021; archived at <https://perma.cc/49KQ-R785>] and [here](https://www.nfl.com/news/super-bowl-lii-how-the-2017-philadelphia-eagles-were-built-0ap3000000912753) [@Rosenthal2018; archived at <https://perma.cc/2GRF-8Z5K>].
 
@@
 
-As depicted in the @fig-goForItPercentageByWinProbChange (which was adapted from a [figure by Ben Baldwin](https://www.nfl4th.com/articles/4th-down-research.html); archived at <https://perma.cc/S5D8-3NCU>), teams went for it on fourth down nearly twice as often in 2021 compared to 2017, when when their win probability was at least 20% (given the current situation at the start of the given play) and they would increase their win probably by going for it.
+As depicted in the @fig-goForItPercentageByWinProbChange [which was adapted from a [figure by Ben Baldwin](https://www.nfl4th.com/articles/4th-down-research.html); @R-nfl4th; archived at <https://perma.cc/S5D8-3NCU>], teams went for it on fourth down nearly twice as often in 2021 compared to 2017, when when their win probability was at least 20% (given the current situation at the start of the given play) and they would increase their win probably by going for it.
+We use the `ggtext` package [@R-ggtext] to annotate text:
 
@@
 
-Here is an analysis by the New York Times on when to go for it on fourth down: <https://www.nytimes.com/2014/09/05/upshot/4th-down-when-to-go-for-it-and-why.html> (archived at <https://perma.cc/KA9Y-BRUD>).
+Here is an analysis by the New York Times on when to go for it on fourth down: <https://www.nytimes.com/2014/09/05/upshot/4th-down-when-to-go-for-it-and-why.html> [@NYT2014; archived at <https://perma.cc/KA9Y-BRUD>].
 Here is a fourth down calculator: <https://rbsdm.com/stats/fourth_calculator/>.
@@
 I can kick it down the middle, see what the wind does and adjust.
-It helps the kicker tremendously." (archived at <https://perma.cc/U7SV-UE2E>).
+It helps the kicker tremendously." [@AP2008; archived at <https://perma.cc/U7SV-UE2E>].
 

### 2025-03-09 download-football-data.qmd
20250308 - citations
20250308 - reorder archive link

@@
 
+We use the `nfl4th` package [@R-nfl4th] to download fourth down data.
+
 ::: {#nte-download4thDownData .callout-note title="Downloading 4th down data"}
@@
 - <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html> [@HoPFRPassing]
-- <https://www.pro-football-reference.com/about/advanced_stats.htm> [archived at <https://perma.cc/LKH5-92PQ>; @PFR2025]
+- <https://www.pro-football-reference.com/about/advanced_stats.htm> [@PFR2025; archived at <https://perma.cc/LKH5-92PQ>]
 

### 2025-03-09 fantasy-football.qmd
20250308 - reorder archive link

@@
 Fantasy football is one of the most widely played games in the history of games.
-It is estimated that around 62 million people play fantasy sports [archived at <https://perma.cc/9PB8-ZDJJ>; @FSGA2023], of whom around 29 million play fantasy football [archived at <https://perma.cc/8YSN-UUNT>; @statistica2023a].
+It is estimated that around 62 million people play fantasy sports [@FSGA2023; archived at <https://perma.cc/9PB8-ZDJJ>], of whom around 29 million play fantasy football [@statistica2023a; archived at <https://perma.cc/8YSN-UUNT>].
 As noted in the [Introduction](intro.qmd), fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life National Football League (NFL) players.^[Fantasy leagues are also available for [baseball](https://en.wikipedia.org/wiki/Fantasy_baseball), [basketball](https://en.wikipedia.org/wiki/Fantasy_basketball), and many other sports.]

### 2025-03-09 index.qmd
20250308 - reorder archive link

@@
 
-I strive to follow [principles of accessibility](https://bookdown.org/yihui/rmarkdown-cookbook/html-accessibility.html) [archived at <https://perma.cc/8XJ9-Q6QJ>; @Xie2024; @Xie2020_packages] to make the book content accessible to people with visual impairments and physical disabilities.
+I strive to follow [principles of accessibility](https://bookdown.org/yihui/rmarkdown-cookbook/html-accessibility.html) [@Xie2024; @Xie2020_packages; archived at <https://perma.cc/8XJ9-Q6QJ>] to make the book content accessible to people with visual impairments and physical disabilities.
 If there are additional ways I can make the content more accessible, please let me know.

### 2025-03-11 evaluating-prediction-accuracy.qmd
202503011 - clean up link

@@
 
-The adapted formula for MASE with non-time series data is described by @Hyndman2014 [here (https://stats.stackexchange.com/a/108963/20338)](https://stats.stackexchange.com/a/108963/20338) (archived at https://perma.cc/G469-8NAJ).
+The adapted formula for MASE with non-time series data is described by @Hyndman2014 at the following link: <https://stats.stackexchange.com/a/108963/20338> (archived at <https://perma.cc/G469-8NAJ>).
 Scaled errors are calculated using @eq-scaledError:

### 2025-03-18 sports-cognitive-psychology.qmd
20250318 - midrange shots in basketball

@@
 Although baseball was one of the first major sports to embrace analytics, other sports have been transformed by analytics, as well.
-For instance, in basketball, there has been a greater focus on three points and buckets close to the rim (e.g., dunks and layups), with way fewer midrange shots.
+For instance, in basketball, there has been a greater focus on three points and buckets close to the rim (e.g., dunks and layups), with way fewer midrange shots [@Partnow2021].
 Moreover, star players rest more games.

### 2025-03-19 player-evaluation.qmd
20250319 - changes in player situation

@@
 For players who have played in the NFL, past performance becomes more relevant because, presumably, the prior situation is more similar (than was their situation in college) to their current situation.
-Nevertheless, lots of things change from game to game and season to season: injuries, coaches, coaching strategies, teammates, etc.
-So just because a player performed well or poorly in a given game or season does not necessarily mean that they will perform similarly in subsequent games/seasons.
+Nevertheless, as described below, lots of things change from game to game and season to season, and such changes are important to monitor because they can render prior situations less relevant to the player's current situation.
+Thus, it is important not to rely just on a player's historical performance from last season.
 Nevertheless, historical performance is one of the best indicators we have.
@@
 We demonstrate how to calculate historical fantasy points in @sec-calculateActualPoints.
+
+#### Changes in Situational Factors {#sec-evalChanges}
+
+Lots of things change from game to game and season to season: injuries, coaches, coaching strategies, teammates, etc.
+Just because a player performed well or poorly in a given game or season does not necessarily mean that they will perform similarly in subsequent games/seasons.
+Thus, it is crucial to consider the player's current situation—and what has changed since the prior game or season.
+Consider what has changed in terms of [situational factors](#sec-evalSituation):
+
+- Is the player on a new team?
+- Do they have a new head coach or coordinator?
+- How will the team's offensive scheme change?
+- Does the team have a better offensive line?
+- Does the team have better receiving targets?
+- Will the player's position change on the depth chart?
+- Will there be more competition for targets?
+
+There is often greater uncertainty in selecting a player who has changed teams or whose situation has greatly changed, but such risk can be highly rewarded when the player's new team or situation is better suited to the player.
+If a player underachieved on a given team, they may perform better for another team.
 

### 2025-03-23 simulation.qmd
20250227 - clean up
20250323 - simulation

@@
 
+A simulation is an "imitative representation" of a phenomenon that could exist the real world.
+In statistics, simulations are computer-driven investigations to better understand a phenomenon by studying its behavior under different conditions.
+For instance, we might want to determine the likely range of outcomes for a player, in terms of the range of fantasy points that a player might score over the course of a season.
+Simulations can be conducted in various ways.
+Two common ways of conducting simulations are via [bootstrapping](#sec-simulationOverviewBootstrapping) and via [Monte Carlo simulation](#sec-simulationOverviewMonteCarlo).
+
+### Bootstrapping {#sec-simulationOverviewBootstrapping}
+
+Bootstrapping involves repeated resampling (with replacement) from observed data.
+For instance, if we have 100 sources provide projections for a player, we could estimate the most likely range of fantasy points for the player by repeatedly sampling from the 100 projections.
+
+### Monte Carlo Simulation {#sec-simulationOverviewMonteCarlo}
+
+Monte Carlo simulation involves repeated random sampling from a known distribution [@Sigal2016].
+For instance, if we know the population distribution for the likely outcomes for a player—e.g., a normal distribution with a known mean and standard deviation—we can repeatedly sample randomly from this distribution.
+
 ## Simulation of Projected Statistics and Points {#sec-simulationProjectedPoints}
+
+### Bootstrapping {#sec-simulationProjectedPointsBootstrapping}
+
+### Monte Carlo Simulation {#sec-simulationProjectedPointsMonteCarlo}
 

### 2025-03-25 sports-cognitive-psychology.qmd
20250324 - conservative play-calling
20250325 - expert performance and automaticity

@@
 
+In addition to punting on fourth down, [loss aversion bias](#sec-cognitiveBiasesLossAversion) may lead coaches to make more conservative play-calling, in general, such as running the ball when passing would tend to be more advantageous.
+
 ## Other Beliefs in Sports {#sec-beliefsInSports}
@@
 
+## Psychological Factors in Player Performance {#sec-psychologicalFactorsPerformance}
+
+## Expert Performance and Automaticity {#sec-expertPerformanceAutomaticity}
+
+This section was adapted from course materials from @Motz2018.
+
+Automaticity involves fast and effortless processing that requires little or no focused attention.
+The more automatic a process is, the less likely a person is to be consciously aware of it.
+For instance, breathing is a fairly automatic process—people can breathe even when they are not concentrating on breathing.
+Automaticity helps free up resources for more demanding tasks.
+Football players and others often face situations requiring multitasking—i.e., performing two (or more) tasks simultaneously.
+For instance, a Quarterback must look for open receivers, monitor approaching defenders, move to open space when necessary, and make an accurate, well-timed throw.
+
+Multitasking is common in everyday life, but some tasks may suffer more from divided attention than others.
+Driving, for example, is a complex task that may be impacted by cognitive load from secondary tasks like phone conversations.
+It is an important question of how well people are able to multitask.
+For instance, we could ask whether talking on a cell phone and driving simultaneously drains cognitive resources.
+If so, does talking on a phone while driving impair driving performance or conversation skills?
+Both tasks (conversing or driving) are highly practiced, so we might expect there to be little cost to dividing attention across these two tasks.
+This provides a nice opportunity to examine automaticity.
+
+Many countries have banned hand-held cell phone use while driving.
+Some countries have banned all use of cell phones, including hands-free, while driving.
+However, states in the United States differ in their cell phone laws.
+
+Almost all cell phone bans restrict the use of handheld cell phones while driving but do not restrict the use of hands-free devices.
+This is implicitly based on the assumption that any impairment to driving while talking on a cell phone is due to the use of a handheld device.
+However, an alternative possibility is that driving impairment with cell phone use is due to dividing one's attention.
+If this is the case, it would impact driving with both handheld and hands-free devices.
+
+A study by @Redelmeier1997 found that 25% of a sample of 699 accidents involved drives that had spoken on a cell phone within 10 minutes prior to the accident.
+The use of handheld and hands-free phone were equally associated with accidents suggesting divided attention is the problem, not the use of a peripheral device.
+However, this association was observed from a [correlational design](#sec-correlationalStudy); thus, we cannot conclude that cell phone use directly caused the accidents.
+Thus, we need studies to examine the effects of multitasking using [experimental designs](#sec-experiment) to establish causality.
+
+@Strayer2001 conducted an experimental study on multitasking.
+For the first task, the tracking task, participants tracked the position of a moving target on a computer screen.
+Every 10–20 seconds, the moving target flashed either red or green.
+If the target turned red (such as one might see when a traffic light turns red), the participant was supposed to push a "brake button", which might simulate the act of needing to brake while driving.
+If the target turned green (such as one might see when a traffic light turns green), the participant was not supposed to do anything.
+During a dual-task phase, the participant was supposed to perform a second task while completing the tracking task.
+For the second task, participants were randomly assigned to one of four conditions: 1) conversing with an experimenter using a hand-held mobile device, 2) conversing with an experimenter using a hands-free device, 3) listening to a radio program of their choice, or 4) listening to a book on tape.
+The researchers found that the probability of missing a color change more than doubled for the cell phone group when performing both the tracking and conversation tasks simultaneously (relative to performing only the tracking task).
+Reaction time also increased for the cell phone group when performing the dual tasks compared to when performing the single tracking task.
+However, there was no difference—in terms of the probability of missing a color change or reaction time—between the dual tasks and single task for the radio or book-on-tape control groups, suggesting that listening, by itself, does not interfere with the tracking task—rather, it appears that actively having a conversation (that includes talking and listening) interferes with tracking changes.
+Moreover, there was no difference between handheld or hands-free phones in terms of the probability of missing a color change or reaction time, suggesting that it was not the use of a hands-occupying device that caused their weaker performance in the tracking task—it was the act of dividing one's attention across tasks (i.e., conversation task and tracking task) that led to weaker performance in tracking changes.
+
+Many people talk on the phone while driving.
+What does it mean that people are (somehow) able to perform these two tasks simultaneously?
+It suggests that those who use the cell phone while driving are treating one of the tasks—driving—as an automatic process and not giving it focused attention.
+Evidence suggests that talking on the phone while driving increases the risk of accident to levels comparable to driving drunk [i.e., a block alcohol level above the legal limit; @Redelmeier1997].
+The cognitive task of talking on the phone uses attentional resources that would otherwise be used for driving a car.
+
+@Strayer2004a also replicated their findings using a driving simulator instead of a tracking task.
+The authors found that use of a cell phone led to 18% slower brake onset time, 12% greater following distance, 17% longer half-recovery time (the time to recover 50% of the speed that was lost during braking), and twice as many rear-end collisions—all despite no general differences in speed.
+Although older participants were poorer overall in driving performance, the effects of cell phone use did not differ for younger (~20 years of age) and older (~70 years of age) adults, suggesting that the driving performance for both younger and older adults is impaired by cell phone use.
+
+The findings described above might suggest that more attention leads to better performance.
+However, this is not always the case.
+@Beilock2001 found that once a task becomes automatic, focusing too much on it can actually hinder performance.
+When a task is highly practiced or automatic, excessive attention can hurt performance.
+That is, more attention can lead to worse performance.
+This is called "choking under pressure" and is based on explicit monitoring theory.
+Highly practiced tasks might include tasks like putting in golf, pitching a baseball, throwing/kicking a football, etc.
+
+By contrast, for tasks that require conscious control, decision-making, and real-time adjustments, more focused attention tends to lead to better performance.
+Tasks that require real-time adjustment include tasks such as avoiding unexpected hazards while driving or trying to avoid rushers while operating as a Quarterback.
+In sum, devoting explicit attention to a task appears to improve performance when the task requires deliberate cognitive processing but can hurt performance when the task is highly automatic.
+
 ## Conclusion {#sec-sportsCognitivePsychologyConclusion}

### 2025-03-26 sports-cognitive-psychology.qmd
20250326 - expert performance

@@
 For instance, breathing is a fairly automatic process—people can breathe even when they are not concentrating on breathing.
+Chewing food can also be automatic.
 Automaticity helps free up resources for more demanding tasks.
@@
 Almost all cell phone bans restrict the use of handheld cell phones while driving but do not restrict the use of hands-free devices.
-This is implicitly based on the assumption that any impairment to driving while talking on a cell phone is due to the use of a handheld device.
-However, an alternative possibility is that driving impairment with cell phone use is due to dividing one's attention.
+This is implicitly based on the assumption, from the *peripheral-interference hypothesis*, that any impairment to driving while talking on a cell phone is due to the use of a handheld device.
+However, an alternative possibility, the *attentional hypothesis*, posits that that driving impairment with cell phone use is due to dividing one's attention.
 If this is the case, it would impact driving with both handheld and hands-free devices.
@@
 However, this association was observed from a [correlational design](#sec-correlationalStudy); thus, we cannot conclude that cell phone use directly caused the accidents.
+It may be, for instance, that risk-taking individuals or individuals in a strongly emotional state may be more likely to both a) have car accidents and b) to use cell phones while driving, which might account for such an association.
 Thus, we need studies to examine the effects of multitasking using [experimental designs](#sec-experiment) to establish causality.
@@
 For the first task, the tracking task, participants tracked the position of a moving target on a computer screen.
-Every 10–20 seconds, the moving target flashed either red or green.
+Every 10–20 seconds, the moving target flashed either red or green to simulate a traffic signal.
 If the target turned red (such as one might see when a traffic light turns red), the participant was supposed to push a "brake button", which might simulate the act of needing to brake while driving.
@@
 For the second task, participants were randomly assigned to one of four conditions: 1) conversing with an experimenter using a hand-held mobile device, 2) conversing with an experimenter using a hands-free device, 3) listening to a radio program of their choice, or 4) listening to a book on tape.
-The researchers found that the probability of missing a color change more than doubled for the cell phone group when performing both the tracking and conversation tasks simultaneously (relative to performing only the tracking task).
+The researchers found that the probability of missing a color change (i.e., a simulated traffic signal) more than doubled for the cell phone group when performing both the tracking and conversation tasks simultaneously (relative to performing only the tracking task).
 Reaction time also increased for the cell phone group when performing the dual tasks compared to when performing the single tracking task.
@@
 Moreover, there was no difference between handheld or hands-free phones in terms of the probability of missing a color change or reaction time, suggesting that it was not the use of a hands-occupying device that caused their weaker performance in the tracking task—it was the act of dividing one's attention across tasks (i.e., conversation task and tracking task) that led to weaker performance in tracking changes.
+Thus, the findings were consistent with the attentional hypothesis, and suggest that divided attention can interfere with performance—in addition to any deficits incurred due to peripheral factors such as handling the phone while dialing.
 
@@
 It suggests that those who use the cell phone while driving are treating one of the tasks—driving—as an automatic process and not giving it focused attention.
-Evidence suggests that talking on the phone while driving increases the risk of accident to levels comparable to driving drunk [i.e., a block alcohol level above the legal limit; @Redelmeier1997].
+Evidence suggests that talking on the phone while driving increases the risk of accident to levels comparable to driving drunk [i.e., a blood alcohol level above the legal limit; @Redelmeier1997].
 The cognitive task of talking on the phone uses attentional resources that would otherwise be used for driving a car.
 
-@Strayer2004a also replicated their findings using a driving simulator instead of a tracking task.
-The authors found that use of a cell phone led to 18% slower brake onset time, 12% greater following distance, 17% longer half-recovery time (the time to recover 50% of the speed that was lost during braking), and twice as many rear-end collisions—all despite no general differences in speed.
-Although older participants were poorer overall in driving performance, the effects of cell phone use did not differ for younger (~20 years of age) and older (~70 years of age) adults, suggesting that the driving performance for both younger and older adults is impaired by cell phone use.
+@Strayer2004a also replicated their findings using a driving simulator, which is a more [ecologically valid](#sec-ecologicalValidity) assessment of driving performance than the tracking task.
+In the study, using the driving simulator, participants drove on a highway in both single-task (i.e., driving only) and dual-task (i.e., driving while talking on a cell phone) conditions.
+Participants drove while following a car that braked at various times.
+The authors found that use of a cell phone led drivers to show a 18% slower brake onset time, 12% greater following distance, 17% longer half-recovery time (the time to recover 50% of the speed that was lost during braking), and twice as many rear-end collisions—all despite no general differences in speed.
+Although older drivers were poorer overall in driving performance, the effects of cell phone use did not differ for younger (~20 years of age) and older (~70 years of age) drivers, suggesting that the driving performance for both younger and older adults is impaired by cell phone use.
+Indeed, the reaction time of younger drivers who were talking on a cell phone was equivalent to the reaction time of older drivers who were not using a cell phone.
 
 The findings described above might suggest that more attention leads to better performance.
-However, this is not always the case.
-@Beilock2001 found that once a task becomes automatic, focusing too much on it can actually hinder performance.
+Based on that assumption, we should tell a Quarterback who has thrown multiple interceptions during a game to pay more attention to his throwing mechanics when throwing the ball, right?
+However, it is not always the case that devoting more attention to a task leads to better performance.
+There are two primary competing theories for "choking" in pressure situations in sports.
+"Choking" refers to underperforming relative to one's ability [@Beilock2001].
+Distraction theories posit that the pressure situation creates a distracting environment that shifts one's attention to cues that are irrelevant to the task, such as one's worries about the situation and its consequences [@Beilock2001].
+An alternative possibility, based on explicit monitoring theory, is that pressure situations raise one's self-awareness and anxiety about performing well, which increases one's attention to the step-by-step procedures involved in skill processes, which may disrupt well-learned or proceduralized performance [@Beilock2001].
+@Beilock2001 found that, among automatic tasks (e.g.
+putting in golf among highly trained individuals), focusing too much on the task can actually hinder performance, consistent with explicit monitoring theory.
 When a task is highly practiced or automatic, excessive attention can hurt performance.
@@
 This is called "choking under pressure" and is based on explicit monitoring theory.
-Highly practiced tasks might include tasks like putting in golf, pitching a baseball, throwing/kicking a football, etc.
+
+Explicit monitoring theories posit that pressure situations raises one's self-awareness and anxiety about performing well, which increases one's attention to the stop-by-step procedures involved in skill processes [@Beilock2001].
+For highly practiced experts, automatic skill tasks might include tasks like hitting/putting a golf ball, pitching a baseball, throwing/kicking a football, shooting free throws in basketball, shooting arrows in archery, etc.
+Sports players often refer to highly practiced tasks as involving "muscle memory."
 
@@
 Tasks that require real-time adjustment include tasks such as avoiding unexpected hazards while driving or trying to avoid rushers while operating as a Quarterback.
+Unpredictable events occur when driving and when operating as a Quarterback; when unpredictable things happen and one's attentions is diverted, one's performance is likely to be impaired [@Strayer2001].
 In sum, devoting explicit attention to a task appears to improve performance when the task requires deliberate cognitive processing but can hurt performance when the task is highly automatic.

### 2025-03-27 sports-cognitive-psychology.qmd
20250327 - Yerkes-Dodson law

@@
 
+The Yerkes–Dodson law describes the association between arousal and performance, as depicted in @fig-yerkesDodsonLaw.
+For simple tasks, performance tends to increase with arousal.
+For difficult tasks, there is an inverted-U-shaped association between arousal and performance, as depicted in @fig-hebbianYerkesDodsonLaw.
+For difficult tasks, too little arousal is bad because it leads to insufficient attention on the task at hand; however, too much arousal is also bad because anxiety can impair performance through its negative effects on attention (e.g., "tunnel vision"), memory, decision making, problem-solving, and multitasking.
+That is, there is an optimal range of arousal for challenging tasks, such as those in sports.
+
+::: {#fig-yerkesDodsonLaw}
+![](images/YerkesDodson.png){width=50% fig-alt="Yerkes–Dodson law. (Figure retrieved from https://en.wikipedia.org/wiki/File:OriginalYerkesDodson.svg)"}
+
+Yerkes–Dodson law.
+For simple tasks, performance increases with arousal.
+For difficult tasks, there is an inverted-U-shaped association between arousal and performance such that too little and too much arousal impairs performance.
+(Figure from [https://en.wikipedia.org/wiki/File:OriginalYerkesDodson.svg](https://en.wikipedia.org/wiki/File:OriginalYerkesDodson.svg).)
+:::
+
+::: {#fig-hebbianYerkesDodsonLaw}
+![](images/HebbianYerkesDodson.png){width=50% fig-alt="Yerkes–Dodson law. (Figure retrieved from https://en.wikipedia.org/wiki/File:HebbianYerkesDodson.svg)"}
+
+Yerkes–Dodson law for difficult tasks.
+For difficult tasks, there is an inverted-U-shaped association between arousal and performance.
+Too little arousal is bad because it leads to insufficient attention on the task at hand; by contrast, too much arousal is also bad because anxiety can impair performance through its negative effects on attention (e.g., "tunnel vision"), memory, decision making, problem-solving, and multitasking.
+(Figure from [https://en.wikipedia.org/wiki/File:HebbianYerkesDodson.svg](https://en.wikipedia.org/wiki/File:HebbianYerkesDodson.svg).)
+:::
+
 ## Expert Performance and Automaticity {#sec-expertPerformanceAutomaticity}

### 2025-03-28 sports-cognitive-psychology.qmd
20250328 - copyediting and figure width

@@
 For difficult tasks, too little arousal is bad because it leads to insufficient attention on the task at hand; however, too much arousal is also bad because anxiety can impair performance through its negative effects on attention (e.g., "tunnel vision"), memory, decision making, problem-solving, and multitasking.
-That is, there is an optimal range of arousal for challenging tasks, such as those in sports.
+That is, there is an optimal range of arousal for complex and challenging tasks, such as those in sports, which often require divided attention and multitasking.
 
 ::: {#fig-yerkesDodsonLaw}
-![](images/YerkesDodson.png){width=50% fig-alt="Yerkes–Dodson law. (Figure retrieved from https://en.wikipedia.org/wiki/File:OriginalYerkesDodson.svg)"}
+![](images/YerkesDodson.png){width=100% fig-alt="Yerkes–Dodson law. (Figure retrieved from https://en.wikipedia.org/wiki/File:OriginalYerkesDodson.svg)"}
 
@@
 ::: {#fig-hebbianYerkesDodsonLaw}
-![](images/HebbianYerkesDodson.png){width=50% fig-alt="Yerkes–Dodson law. (Figure retrieved from https://en.wikipedia.org/wiki/File:HebbianYerkesDodson.svg)"}
+![](images/HebbianYerkesDodson.png){width=100% fig-alt="Yerkes–Dodson law. (Figure retrieved from https://en.wikipedia.org/wiki/File:HebbianYerkesDodson.svg)"}
 

### 2025-03-30 base-rates.qmd
20250330 - update links

@@
 ![](images/probabilityNomogramLine.png){width=50% fig-alt="Probability Nomogram.
-(Figure retrieved from https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png).
+(Figure retrieved from https://commons.wikimedia.org/wiki/File:Fagan_nomogram.svg).
 Also provided in: @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
@@
 Probability Nomogram Example.
-(Figure adapted from [https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png](https://upload.wikimedia.org/wikipedia/commons/thumb/6/66/Fagan_nomogram.svg/945px-Fagan_nomogram.svg.png).
+(Figure adapted from [https://commons.wikimedia.org/wiki/File:Fagan_nomogram.svg](https://commons.wikimedia.org/wiki/File:Fagan_nomogram.svg).
 Also provided in: @Petersen2024a and @PetersenPrinciplesPsychAssessment.)

### 2025-03-31 decision-making.qmd
20250330 - reinforcement schedules

@@
 
+### Reinforcement Schedules {#sec-reinforcementSchedules}
+
+Reinforcement schedules are depicted in @fig-reinforcementSchedules, as adapted from @Spielman2020.
+
+::: {#fig-reinforcementSchedules}
+![](images/reinforcementSchedules.png){width=100% fig-alt="Reinforcement Schedules.
+The four reinforcement schedules yield different response patterns.
+The variable ratio schedule is unpredictable and yields high and steady response rates, with little if any pause after reinforcement (e.g., gambler who plays the slot machines).
+A fixed ratio schedule is predictable and produces a high response rate, with a short pause after reinforcement (e.g., salesperson paid on commission).
+The variable interval schedule is unpredictable and produces a moderate, steady response rate (e.g., restaurant manager who is paid a bonus if the food inspector, who comes at unpredictable times, gives the restaurant a good rating).
+The fixed interval schedule yields a scallop-shaped response pattern, reflecting a significant pause after reinforcement (e.g., surgery patient who can receive a painkiller after pressing a button, but only up to once per hour).
+(Figure and caption adapted from Spielman, Lovett, & Jenkins (2020); https://openstax.org/books/psychology-2e/pages/6-3-operant-conditioning; Spielman, R.
+M., Jenkins, W.
+J., & Lovett, M.
+D.
+(2000).
+*Psychology*.
+OpenStax.
+https://openstax.org/details/books/psychology-2e; archived at https://perma.cc/9DZ8-34X2)"}
+
+Reinforcement Schedules.
+The four reinforcement schedules yield different response patterns.
+The variable ratio schedule is unpredictable and yields high and steady response rates, with little if any pause after reinforcement (e.g., gambler who plays the slot machines).
+A fixed ratio schedule is predictable and produces a high response rate, with a short pause after reinforcement (e.g., salesperson paid on commission).
+The variable interval schedule is unpredictable and produces a moderate, steady response rate (e.g., restaurant manager who is paid a bonus if the food inspector, who comes at unpredictable times, gives the restaurant a good rating).
+The fixed interval schedule yields a scallop-shaped response pattern, reflecting a significant pause after reinforcement (e.g., surgery patient who can receive a painkiller after pressing a button, but only up to once per hour).
+(Figure and caption adapted from @Spielman2020; [https://openstax.org/books/psychology-2e/pages/6-3-operant-conditioning](https://openstax.org/books/psychology-2e/pages/6-3-operant-conditioning); Spielman, R.
+M., Jenkins, W.
+J., & Lovett, M.
+D.
+(2020).
+*Psychology*.
+OpenStax.
+[https://openstax.org/details/books/psychology-2e](https://openstax.org/details/books/psychology-2e); archived at [https://perma.cc/9DZ8-34X2](https://perma.cc/9DZ8-34X2)).
+:::
+
 ## Suggestions {#sec-decisionMakingSuggestions}

### 2025-03-31 decision-making.qmd
20250331 - reinforcement

@@
 
+The National Football League (NFL) used to be adamantly against sports gambling.
+They wanted to protect the integrity of the game.
+They felt that sports gambling might lead to match fixing, point shaving, and other forms of corruption that would undermine the legitimacy of competition in the game.
+Match fixing occurs when players, referees, coaches, or others conspire to manipulate the outcome of the game.
+Point shaving involves manipulating the score—commonly by a player intentionally underperforming (even if it does not lead to "throwing the game" and causing the team to lose).
+Match fixing and point shaving can occur for a variety of reasons, but a common reason is to ensure that particular gamblers win their bet.
+Another reason the NFL was against sports betting was that, prior to the U.S.
+Supreme Court overturning the Professional and Amateur Sports Protection Act (PASPA) in 2018, there were legal restrictions on gambling that prevented sports betting in many parts of the country.
+
+However, after the legalization of sports betting following the overturning of PASPA in 2018, the NFL has changed their tune and are now strongly in favor of sports betting.
+The rise of legal sports betting has created massive revenue streams for the NFL through partnerships, sponsorships, and advertising deals with sportsbooks.
+
 There is substantial [overconfidence](#sec-cognitiveBiasesOverconfidence) (in particular overestimation) in sports betting/gambling.
@@
 
-### Reinforcement Schedules {#sec-reinforcementSchedules}
+### Reinforcement {#sec-reinforcement}
 
-Reinforcement schedules are depicted in @fig-reinforcementSchedules, as adapted from @Spielman2020.
+Reinforcement involves a (typically appetitive) stimulus that increases the frequency of a behavior.
+For instance, reinforcers could include things like food, praise, attention, money, etc.
+Thus, sports gambling can provide reinforcement through the receipt of money.
+
+#### Continuous Reinforcement {#sec-reinforcementContinuous}
+
+In general, when you want to train a new behavior, continuous reinforcement is the fastest way to do so.
+Continuous reinforcement means rewarding the behavior each time that it occurs.
+For instance, a parent might use continuous reinforcement to train their child to put their toys away—the parent may praise them or give them physical affection each time they put their toys away.
+However, continuous reinforcement is susceptible to extinction.
+Extinction means cessation of the target behavior (putting the toys away, in this example).
+That is, for a previously continusouly reinforced behavior, if the behavior stops being rewarded, the person is less likely to continue the behavior.
+Thus, to make a behavior less susceptible to extinction, [intermittent reinforcement](#sec-reinforcementIntermittent) may be used.
+
+#### Intermittent Reinforcement {#sec-reinforcementIntermittent}
+
+Intermittent reinforcement (also called partial reinforcement) means rewarding the behavior sometimes but not everytime.
+There are various approaches to intermittent reinforcement, called schedules of reinforcement.
+The four primary reinforcement schedules are: fixed interval, fixed ratio, variable interval, adn variable ratio.
+The reinforcement schedules are depicted in @fig-reinforcementSchedules, as adapted from @Spielman2020.
 
@@
 
+A *fixed interval* reinforcement schedule occurs when the person is rewarded after a set amount of time.
+For instance, if the reward becomes available every 30 minutes, the person receives the reward the first time they engage in the behavior after the reward becomes available.
+As an example, consider a surgery patient who can receive a painkiller after pressing a button, but only up to once per hour.
+
+A *variable interval* reinforcement schedule occurs when the person is rewarded after a varying (and unpredictable) amount of time.
+As an example, consider a restaurant manager who is paid a bonus if the food inspector, who comes at unpredictable times, gives the restaurant a good rating.
+
+A *fixed ratio* reinforcement schedule occurs when the person is rewarded after a set number of responses.
+As an example, consider a coffee shop that offers a free coffee after every 10 purchases that a person makes.
+
+A *variable ratio* reinforcement schedule occurs when the person is rewarded after a varying (and unpredictable) number of responses.
+As an example, consider a slot machine that provides a large monetary reward (paired with the machine lighting up and bells going off) after a varying number of attempts.
+Many forms of gambling involve variable ratio reinforcement.
+
+The variable ratio reinforcement schedule is the least susceptible to extinction.
+People can go without reinforcement many times and may still continue to gamble out of the hope for a large reward on the next attempt.
+The low susceptibility of variable ratio reinforcement to extinction is, in part, why gambling (which often involves variable ratio reinforcement) can be highly addictive despite people's tendency to lose money gambling.
+
+### Does Fantasy Football Luck or Skill? {#sec-luckOrSkill}
+
+Fantasy football (and sports betting, more generally) is not 100% luck.
+Skill can be involved in identifying undervalued assets—whether in terms of stocks in the stock market or professional football players, where inside information can be especially valuable (and thus illegal to profit from).
+To evaluate the percentage of variability in fantasy football that is attributable to luck versus skill, @Getty2018 evaluated the extent to which one's performance in the first half of the season was correlated with one's performance in the second half of the season, under the assumption that underlying skill would lead to persistence of performance across time.
+The authors examined FanDuel daily fantasy sports (DFS) leagues, which allow the user to set an entirely new lineup each week from all players, based on salary constraints ( so a user cannot just select the best player at every position).
+The authors estimated that ~55% of the variability in fantasy football performance across time was due to skill and 45% was due to luck [@Getty2018].
+Thus, performance in fantasy football is around half and half luck versus skill.
+
+Because fantasy football is not 100% luck, it is not the same as a slot machine.
+However, there are still unpredictable elements (e.g., injuries), and there is considerable luck involved.
+As a result, sports betting and gambling on fantasy football still involve components of [variable ratio reinforcement](#sec-reinforcementIntermittent) that lend them to being potentially highly addictive.
+
 ## Suggestions {#sec-decisionMakingSuggestions}

### 2025-03-31 decision-making.qmd
20250331 - bracketing

@@
 
+Below are examples where the predictions a) are centered on the truth, b) bracket the truth, and c) do not bracket the truth.
+They are illustrated in @fig-bracketing.
+
+::: {#fig-bracketing}
+![](images/bracketing.png){width=100% fig-alt="Three distributions of individual judgments in which judgments bracket and are centered on the truth (top left panel), bracket but are not centered on the truth (top right panel), or do not bracket the truth (i.e., are biased below the truth; bottom panel).
+The distributions depict the predictions.
+The dashed lines represent the true value.
+(Figure and caption adapted from Larrick, R.
+P., Mannes, A.
+E., & Soll, J.
+B.
+(2024).
+The social psychology of the wisdom of crowds (with a new section on recent advances).
+In F.
+M.
+Federspiel, G.
+Montibeller, & M.
+Seifert (Eds.), *Behavioral decision analysis* (pp.
+121–143).
+Springer.
+https://doi.org/10.1007/978-3-031-44424-1_7)"}
+
+Three distributions of individual judgments in which judgments bracket and are centered on the truth (top left panel), bracket but are not centered on the truth (top right panel), or do not bracket the truth (i.e., are biased below the truth; bottom panel).
+The distributions depict the predictions.
+The dashed lines represent the true value.
+(Figure and caption adapted from @Larrick2024; Larrick, R.
+P., Mannes, A.
+E., & Soll, J.
+B.
+(2024).
+The social psychology of the wisdom of crowds (with a new section on recent advances).
+In F.
+M.
+Federspiel, G.
+Montibeller, & M.
+Seifert (Eds.), *Behavioral decision analysis* (pp.
+121–143).
+Springer.
+[https://doi.org/10.1007/978-3-031-44424-1_7](https://doi.org/10.1007/978-3-031-44424-1_7)).
+:::
+
+Here is a code example of when the crowd-averaged predictions are centered on the truth for a player:
+
 ```{r}
@@
 ```
+
+Here is a code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player:
 
@@
 
+Here is a code example of when the crowd-averaged predictions do not bracket the truth for a player:
+
 ```{r}
@@
 ```
+
+As demonstrated from these examples, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+When the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+When the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
 

### 2025-03-31 decision-making.qmd
20250331 - bracketing

@@
 
+As demonstrated above, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+
 Here is a code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player:
@@
 ```
+
+As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+
+Here is another code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player:
+
+```{r}
+predictedValues <- c(10,29,29,40,60,70,80,90)
+actualValue <- 15
+errors <- predictedValues - actualValue
+
+errors
+median(errors)
+mean(predictedValues)
+
+petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+```
+
+As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
 
@@
 
-As demonstrated from these examples, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
-When the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
-When the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+As demonstrated above, when the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+
+Here is another code example of when the crowd-averaged predictions do not bracket the truth for a player:
+
+```{r}
+predictedValues <- c(0,65,70,75,80,85,90,95)
+actualValue <- 100
+errors <- predictedValues - actualValue
+
+errors
+median(errors)
+mean(predictedValues)
+
+petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+```
+
+As demonstrated above, when the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError), but they can be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
+
+In conclusion, as demonstrated from these examples, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+When the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
+When the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError), but they can be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
+
+In sum, going with crowd-averaged projections will always yield predictions that are as or more accurate (in terms of [mean absolute error](#sec-meanAbsoluteError)) than the (collective) accuracy of the individuals' projections.
+However, the crowd-averaged projections may be less accurate than any individual projection, including the projection of the forecaster who is at the 50th percentile in accuracy.
+That is, depending on the bracketing rate, the crowd-averaged projections can still be less accurate than most individual projections.
 
@@
 
-### Does Fantasy Football Luck or Skill? {#sec-luckOrSkill}
+### Is Fantasy Football a Game of Luck or Skill? {#sec-luckOrSkill}
 

### 2025-04-01 decision-making.qmd
20250331 - bracketing

@@
 library("petersenlab")
+library("tidyverse")
 ```
@@
 
-Here is a code example of when the crowd-averaged predictions are centered on the truth for a player:
+Here is a code example of when the crowd-averaged predictions are centered on the truth for a player.
+The distribution of projected points versus truth is depicted in @fig-bracketingCentered.
 
@@
 
-As demonstrated above, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+```{r}
+#| label: fig-bracketingCentered
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Brackets and is Centered on the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Brackets and is Centered on the Truth."
+#| code-fold: true
 
-Here is a code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player:
+bracketingCentered <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+maeBracketingCentered_individual <- bracketingCentered[which(row.names(bracketingCentered) == "individual"), "MAE"]
+maeBracketingCentered_crowd <- bracketingCentered[which(row.names(bracketingCentered) == "crowdAveraged"), "MAE"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5) +
+  #annotate(
+  #  "segment",
+  #  x = mean(predictedValues),
+  #  xend = mean(predictedValues) + maeBracketingCentered_crowd,
+  #  y = 0,
+  #  yend = 0,
+  #  linewidth = 1.5,
+  #  alpha = 0.6,
+  #  arrow = arrow(ends = "both")) +
+  annotate(
+    "text",
+    x = mean(predictedValues) + (maeBracketingCentered_crowd / 2) + 1,
+    y = 0.0003,
+    label = paste("MAE: ", maeBracketingCentered_crowd, sep = ""),
+    hjust = 0) + # left-justify
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs. Truth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate (MAE = `{r} maeBracketingCentered_crowd`) than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingCentered_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError).
+
+Here is a code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player.
+The distribution of projected points versus truth is depicted in @fig-bracketingNotCentered1.
 
@@
 
-As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+```{r}
+#| label: fig-bracketingNotCentered1
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
+#| code-fold: true
 
-Here is another code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player:
+bracketingNotCentered1 <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+maeBracketingNotCentered1_individual <- bracketingNotCentered1[which(row.names(bracketingNotCentered1) == "individual"), "MAE"]
+maeBracketingNotCentered1_crowd <- bracketingNotCentered1[which(row.names(bracketingNotCentered1) == "crowdAveraged"), "MAE"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5) +
+  annotate(
+    "segment",
+    x = mean(predictedValues),
+    xend = mean(predictedValues) + maeBracketingNotCentered1_crowd,
+    y = 0,
+    yend = 0,
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow(ends = "both")) +
+  annotate(
+    "text",
+    x = mean(predictedValues) + (maeBracketingNotCentered1_crowd / 2),
+    y = 0.0003,
+    label = paste("MAE: ", maeBracketingNotCentered1_crowd, sep = ""),
+    hjust = 0.5) + # center-justify
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs. Truth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions (MAE = `{r} maeBracketingNotCentered1_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError).
+
+Here is another code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player.
+The distribution of projected points versus truth is depicted in @fig-bracketingNotCentered2.
 
@@
 
-As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
+```{r}
+#| label: fig-bracketingNotCentered2
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
+#| code-fold: true
 
-Here is a code example of when the crowd-averaged predictions do not bracket the truth for a player:
+bracketingNotCentered2 <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+maeBracketingNotCentered2_individual <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "individual"), "MAE"]
+maeBracketingNotCentered2_crowd <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "crowdAveraged"), "MAE"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5) +
+  annotate(
+    "segment",
+    x = mean(predictedValues),
+    xend = mean(predictedValues) - maeBracketingNotCentered2_crowd,
+    y = 0,
+    yend = 0,
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow(ends = "both")) +
+  annotate(
+    "text",
+    x = mean(predictedValues) - (maeBracketingNotCentered2_crowd / 2),
+    y = 0.0003,
+    label = paste("MAE: ", maeBracketingNotCentered2_crowd, sep = ""),
+    hjust = 0.5) + # center-justify
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs. Truth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions (MAE = `{r} maeBracketingNotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered2_crowd`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
+
+Here is a code example of when the crowd-averaged predictions do not bracket the truth for a player.
+The distribution of projected points versus truth is depicted in @fig-noBracketing1.
 
@@
 
-As demonstrated above, when the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+```{r}
+#| label: fig-noBracketing1
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Does Not Bracket the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Does Not Bracket the Truth."
+#| code-fold: true
 
-Here is another code example of when the crowd-averaged predictions do not bracket the truth for a player:
+noBracketing1 <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+maeNoBracketing1_individual <- noBracketing1[which(row.names(noBracketing1) == "individual"), "MAE"]
+maeNoBracketing1_crowd <- noBracketing1[which(row.names(noBracketing1) == "crowdAveraged"), "MAE"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5) +
+  annotate(
+    "segment",
+    x = mean(predictedValues),
+    xend = mean(predictedValues) + maeNoBracketing1_crowd,
+    y = 0,
+    yend = 0,
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow(ends = "both")) +
+  annotate(
+    "text",
+    x = mean(predictedValues) + (maeNoBracketing1_crowd / 2),
+    y = 0.0003,
+    label = paste("MAE: ", maeNoBracketing1_crowd, sep = ""),
+    hjust = 0.5) + # center-justify
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs. Truth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions (MAE = `{r} maeNoBracketing1_crowd`) are as accurate as the accuracy of the typical individual forecaster (MAE = `{r} maeNoBracketing1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError).
+
+Here is another code example of when the crowd-averaged predictions do not bracket the truth for a player.
+The distribution of projected points versus truth is depicted in @fig-noBracketing2.
 
@@
 
-As demonstrated above, when the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError), but they can be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
+```{r}
+#| label: fig-noBracketing2
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Does Not Bracket the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Does Not Bracket the Truth."
+#| code-fold: true
+
+noBracketing2 <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+maeNoBracketing2_individual <- noBracketing2[which(row.names(noBracketing2) == "individual"), "MAE"]
+maeNoBracketing2_crowd <- noBracketing2[which(row.names(noBracketing2) == "crowdAveraged"), "MAE"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5) +
+  annotate(
+    "segment",
+    x = mean(predictedValues),
+    xend = mean(predictedValues) + maeNoBracketing2_crowd,
+    y = 0,
+    yend = 0,
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow(ends = "both")) +
+  annotate(
+    "text",
+    x = mean(predictedValues) + (maeNoBracketing2_crowd / 2),
+    y = 0.0006,
+    label = paste("MAE: ", maeNoBracketing2_crowd, sep = ""),
+    hjust = 0.5) + # center-justify
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs. Truth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions (MAE = `{r} maeNoBracketing2_crowd`) are as accurate as the accuracy of the typical individual forecaster (MAE = `{r} maeNoBracketing2_crowd`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
 
@@
 That is, depending on the bracketing rate, the crowd-averaged projections can still be less accurate than most individual projections.
+In general, the higher the bracketing rate and the more closely that the projections are centered on the truth, the more accurate the crowd-averaged projections will be.
 

### 2025-04-01 decision-making.qmd
20250401 - median (absolute) error

@@
     linetype = "dashed",
-    linewidth = 1.5) +
+    linewidth = 1.5
+  ) +
   #annotate(
@@
     label = paste("MAE: ", maeBracketingCentered_crowd, sep = ""),
-    hjust = 0) + # left-justify
+    hjust = 0 # left-justify
+  ) +
   labs(
@@
     title = "Density Plot of Predicted Fantasy Points for a Player",
-    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs. Truth, i.e., Actual Fantasy Points (Dashed Line)"
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs.\nTruth, i.e., Actual Fantasy Points (Dashed Line)"
   ) +
@@
 errors
-median(errors)
+median(abs(errors))
 mean(predictedValues)
@@
     linetype = "dashed",
-    linewidth = 1.5) +
+    linewidth = 1.5
+  ) +
   annotate(
@@
     alpha = 0.6,
-    arrow = arrow(ends = "both")) +
+    arrow = arrow(ends = "both")
+  ) +
   annotate(
@@
     label = paste("MAE: ", maeBracketingNotCentered1_crowd, sep = ""),
-    hjust = 0.5) + # center-justify
+    hjust = 0.5 # center-justify
+  ) +
   labs(
@@
     label = paste("MAE: ", maeBracketingNotCentered2_crowd, sep = ""),
-    hjust = 0.5) + # center-justify
+    hjust = 0.5 # center-justify
+  ) +
   labs(
@@
     label = paste("MAE: ", maeNoBracketing1_crowd, sep = ""),
-    hjust = 0.5) + # center-justify
+    hjust = 0.5 # center-justify
+  ) +
   labs(
@@
     label = paste("MAE: ", maeNoBracketing2_crowd, sep = ""),
-    hjust = 0.5) + # center-justify
+    hjust = 0.5 # center-justify
+  ) +
   labs(

### 2025-04-01 decision-making.qmd
20250401 - median absolute error

--- 2025-03-31 decision-making.qmd
@@
 errors
-median(errors)
 mean(predictedValues)
@@
 
-As demonstrated above, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+```{r}
+#| label: fig-bracketingCentered
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Brackets and is Centered on the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Brackets and is Centered on the Truth."
+#| code-fold: true
 
-Here is a code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player:
+bracketingCentered <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+maeBracketingCentered_individual <- bracketingCentered[which(row.names(bracketingCentered) == "individual"), "MAE"]
+maeBracketingCentered_crowd <- bracketingCentered[which(row.names(bracketingCentered) == "crowdAveraged"), "MAE"]
+
+mdaeBracketingCentered_individual <- bracketingCentered[which(row.names(bracketingCentered) == "individual"), "MdAE"]
+meBracketingCentered_crowd <- bracketingCentered[which(row.names(bracketingCentered) == "crowdAveraged"), "ME"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5
+  ) +
+  #annotate(
+  #  "segment",
+  #  x = mean(predictedValues),
+  #  xend = mean(predictedValues) + maeBracketingCentered_crowd,
+  #  y = 0,
+  #  yend = 0,
+  #  linewidth = 1.5,
+  #  alpha = 0.6,
+  #  arrow = arrow(ends = "both")) +
+  annotate(
+    "text",
+    x = mean(predictedValues) + (maeBracketingCentered_crowd / 2) + 1,
+    y = 0.0003,
+    label = paste("Mean Error: ", meBracketingCentered_crowd, sep = ""),
+    hjust = 0 # left-justify
+  ) +
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs.\nTruth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate (MAE/MdAE = `{r} maeBracketingCentered_crowd`) than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingCentered_individual`; MdAE = `{r} mdaeBracketingCentered_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
+
+Here is a code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player, with a high bracketing rate.
+The distribution of projected points versus truth is depicted in @fig-bracketingNotCentered1.
 
@@
 
-As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+```{r}
+#| label: fig-bracketingNotCentered1
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
+#| code-fold: true
 
-Here is another code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player:
+bracketingNotCentered1 <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+maeBracketingNotCentered1_individual <- bracketingNotCentered1[which(row.names(bracketingNotCentered1) == "individual"), "MAE"]
+maeBracketingNotCentered1_crowd <- bracketingNotCentered1[which(row.names(bracketingNotCentered1) == "crowdAveraged"), "MAE"]
+
+mdaeBracketingNotCentered1_individual <- bracketingNotCentered1[which(row.names(bracketingNotCentered1) == "crowdAveraged"), "MdAE"]
+meBracketingNotCentered1_crowd <- bracketingNotCentered1[which(row.names(bracketingNotCentered1) == "crowdAveraged"), "ME"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5
+  ) +
+  annotate(
+    "segment",
+    x = mean(predictedValues),
+    xend = actualValue,
+    y = 0,
+    yend = 0,
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow(ends = "both")
+  ) +
+  annotate(
+    "text",
+    x = actualValue + (meBracketingNotCentered1_crowd / 2),
+    y = 0.0007,
+    label = paste("Mean Error: ", meBracketingNotCentered1_crowd, sep = ""),
+    hjust = 0.5 # center-justify
+  ) +
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs.\nTruth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth, and the bracketing rate is high, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingNotCentered1_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered1_individual`; MdAE = `{r} mdaeBracketingNotCentered1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
+
+Here is another code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player, with a low bracketing rate.
+The distribution of projected points versus truth is depicted in @fig-bracketingNotCentered2.
 
@@
 
-As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
+```{r}
+#| label: fig-bracketingNotCentered2
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
+#| code-fold: true
 
-Here is a code example of when the crowd-averaged predictions do not bracket the truth for a player:
+bracketingNotCentered2 <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+maeBracketingNotCentered2_individual <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "individual"), "MAE"]
+maeBracketingNotCentered2_crowd <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "crowdAveraged"), "MAE"]
+
+mdaeBracketingNotCentered2_individual <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "crowdAveraged"), "MdAE"]
+meBracketingNotCentered2_crowd <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "crowdAveraged"), "ME"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5
+  ) +
+  annotate(
+    "segment",
+    x = mean(predictedValues),
+    xend = actualValue,
+    y = 0,
+    yend = 0,
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow(ends = "both")
+  ) +
+  annotate(
+    "text",
+    x = actualValue + (meBracketingNotCentered2_crowd / 2),
+    y = 0.0003,
+    label = paste("Mean Error: ", meBracketingNotCentered2_crowd, sep = ""),
+    hjust = 0.5 # center-justify
+  ) +
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs.\nTruth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth and the bracketing rate is low, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingNotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy (MdAE = `{r} mdaeBracketingNotCentered2_individual`).
+
+Here is a code example of when the crowd-averaged predictions do not bracket the truth for a player, but the distribution is still close to truth.
+The distribution of projected points versus truth is depicted in @fig-noBracketing1.
 
@@
 
-As demonstrated above, when the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+```{r}
+#| label: fig-noBracketing1
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Does Not Bracket the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Does Not Bracket the Truth."
+#| code-fold: true
 
-Here is another code example of when the crowd-averaged predictions do not bracket the truth for a player:
+noBracketing1 <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+maeNoBracketing1_individual <- noBracketing1[which(row.names(noBracketing1) == "individual"), "MAE"]
+maeNoBracketing1_crowd <- noBracketing1[which(row.names(noBracketing1) == "crowdAveraged"), "MAE"]
+
+mdaeNoBracketing1_individual <- noBracketing1[which(row.names(noBracketing1) == "crowdAveraged"), "MdAE"]
+meNoBracketing1_crowd <- noBracketing1[which(row.names(noBracketing1) == "crowdAveraged"), "ME"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5
+  ) +
+  annotate(
+    "segment",
+    x = mean(predictedValues),
+    xend = actualValue,
+    y = 0,
+    yend = 0,
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow(ends = "both")
+  ) +
+  annotate(
+    "text",
+    x = actualValue + (meNoBracketing1_crowd / 2),
+    y = 0.0003,
+    label = paste("Mean Error: ", meNoBracketing1_crowd, sep = ""),
+    hjust = 0.5 # center-justify
+  ) +
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs.\nTruth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the crowd-averaged predictions do not bracket the truth but are still close to the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeNoBracketing1_crowd`) are as accurate as the accuracy of the typical individual forecaster (MAE = `{r} maeNoBracketing1_individual`; MdAE = `{r} mdaeNoBracketing1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
+
+Here is another code example of when the crowd-averaged predictions do not bracket the truth for a player and there is an outlier prediction that is far away from the truth.
+The distribution of projected points versus truth is depicted in @fig-noBracketing2.
 
@@
 
-As demonstrated above, when the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError), but they can be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
+```{r}
+#| label: fig-noBracketing2
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Does Not Bracket the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Does Not Bracket the Truth."
+#| code-fold: true
+
+noBracketing2 <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+maeNoBracketing2_individual <- noBracketing2[which(row.names(noBracketing2) == "individual"), "MAE"]
+maeNoBracketing2_crowd <- noBracketing2[which(row.names(noBracketing2) == "crowdAveraged"), "MAE"]
+
+mdaeNoBracketing2_individual <- noBracketing2[which(row.names(noBracketing2) == "crowdAveraged"), "MdAE"]
+meNoBracketing2_crowd <- noBracketing2[which(row.names(noBracketing2) == "crowdAveraged"), "ME"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5
+  ) +
+  annotate(
+    "segment",
+    x = mean(predictedValues),
+    xend = actualValue,
+    y = 0,
+    yend = 0,
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow(ends = "both")
+  ) +
+  annotate(
+    "text",
+    x = actualValue + (meNoBracketing2_crowd / 2),
+    y = 0.0015,
+    label = paste("Mean Error: ", meNoBracketing2_crowd, sep = ""),
+    hjust = 0.5 # center-justify
+  ) +
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs.\nTruth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the crowd-averaged predictions do not bracket the truth and there are one or more predictions that are far away from the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeNoBracketing2_crowd`) are as accurate as the accuracy of the typical individual forecaster (MAE = `{r} maeNoBracketing2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy (MdAE = `{r} mdaeNoBracketing2_individual`).
 

### 2025-04-01 evaluating-prediction-accuracy.qmd
20250311 - quote attribution
20250311 - add citation
20250401 - median absolute error

@@
 
-> "Nothing ruins fantasy like reality." – Renee Miller
+> "Nothing ruins fantasy more than reality." – [Renee Miller](https://www.youtube.com/watch?v=gmpLFWs5ae0&t=1430s) [@Yahoo2024]
 
@@
 $$ {#eq-meanAbsoluteError}
+
+##### Median Absolute Error (MdAE) {#sec-medianAbsoluteError}
+
+Median absolute error (MdAE) is the median of the absolute value of differences between the predicted and observed values across individuals, that is, the median of the absolute value of errors.
+Smaller MdAE values (closer to zero) reflect greater accuracy.
+MdAE indicates the accuracy of the "typical" prediction (i.e., the prediction that is at the 50th percentile in terms of accuracy).
+The formula for MdAE is in @eq-medianAbsoluteError:
+
+$$
+\begin{aligned}
+  \text{median absolute error (MdAE)} &= \frac{\text{median}(|\text{predicted}_i - \text{observed}_i|)}{n} \\
+  &= \text{median}(|e_i|)
+\end{aligned}
+$$ {#eq-medianAbsoluteError}
 

### 2025-04-01 decision-making.qmd
20250401 - correct values

@@
 
-mdaeBracketingNotCentered1_individual <- bracketingNotCentered1[which(row.names(bracketingNotCentered1) == "crowdAveraged"), "MdAE"]
+mdaeBracketingNotCentered1_individual <- bracketingNotCentered1[which(row.names(bracketingNotCentered1) == "individual"), "MdAE"]
 meBracketingNotCentered1_crowd <- bracketingNotCentered1[which(row.names(bracketingNotCentered1) == "crowdAveraged"), "ME"]
@@
 
-mdaeBracketingNotCentered2_individual <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "crowdAveraged"), "MdAE"]
+mdaeBracketingNotCentered2_individual <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "individual"), "MdAE"]
 meBracketingNotCentered2_crowd <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "crowdAveraged"), "ME"]
@@
 
-mdaeNoBracketing1_individual <- noBracketing1[which(row.names(noBracketing1) == "crowdAveraged"), "MdAE"]
+mdaeNoBracketing1_individual <- noBracketing1[which(row.names(noBracketing1) == "individual"), "MdAE"]
 meNoBracketing1_crowd <- noBracketing1[which(row.names(noBracketing1) == "crowdAveraged"), "ME"]
@@
 
-mdaeNoBracketing2_individual <- noBracketing2[which(row.names(noBracketing2) == "crowdAveraged"), "MdAE"]
+mdaeNoBracketing2_individual <- noBracketing2[which(row.names(noBracketing2) == "individual"), "MdAE"]
 meNoBracketing2_crowd <- noBracketing2[which(row.names(noBracketing2) == "crowdAveraged"), "ME"]
@@
 
-In conclusion, as demonstrated from these examples, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
-When the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
-When the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError), but they can be less accurate than the individual forecaster who is at the 50th percentile in accuracy.
+In conclusion, as demonstrated from these examples, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError) and of most individual forecasters in terms of [median absolute error](#sec-medianAbsoluteError).
+When the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+If the bracketing rate is high, the crowd-averaged predictions will be more accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than most individual forecasters; however, if the bracketing rate is low, the crowd-averaged predictions can be less accurate than than most individual forecasters.
+When the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
+Depending on how far some of the individual projections are from the truth, the crowd-averaged projections can be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than most individual forecasters.
 
 In sum, going with crowd-averaged projections will always yield predictions that are as or more accurate (in terms of [mean absolute error](#sec-meanAbsoluteError)) than the (collective) accuracy of the individuals' projections.
-However, the crowd-averaged projections may be less accurate than any individual projection, including the projection of the forecaster who is at the 50th percentile in accuracy.
+However, the crowd-averaged projections may be less accurate than any *individual* projection, including the projection of the forecaster who is at the 50th percentile in accuracy.
 That is, depending on the bracketing rate, the crowd-averaged projections can still be less accurate than most individual projections.

### 2025-04-01 decision-making.qmd
20250401 - closed arrowheads

@@
   #  x = mean(predictedValues),
-  #  xend = mean(predictedValues) + maeBracketingCentered_crowd,
+  #  xend = actualValue,
   #  y = 0,
@@
   #  linewidth = 1.5,
-  #  alpha = 0.6,
-  #  arrow = arrow(ends = "both")) +
+  #  arrow = arrow(
+  #    ends = "both",
+  #    type = "closed")
+  #)
   annotate(
@@
     linewidth = 1.5,
-    alpha = 0.6,
-    arrow = arrow(ends = "both")
+    arrow = arrow(
+      ends = "both",
+      type = "closed")
   ) +

### 2025-04-01 decision-making.qmd
20250401 - smaller arrowheads

@@
   #  arrow = arrow(
+  #    angle = 20,
   #    ends = "both",
@@
     arrow = arrow(
+      angle = 20,
       ends = "both",
@@
     linewidth = 1.5,
-    alpha = 0.6,
-    arrow = arrow(ends = "both")
+    arrow = arrow(
+      angle = 20,
+      ends = "both",
+      type = "closed")
   ) +

### 2025-04-02 R-scripts/process_projections.R
20250402 - process projections

@@ -1,0 +1,191 @@
+# Process Projections
+
+library("stringr")
+library("tidyverse")
+library("purrr")
+
+# Weekly Projections ----
+
+filenames <- list.files(
+  path = "./data/by_year",
+  pattern = "players_projectedPoints_weekly_.*\\.RData",
+  recursive = TRUE,
+  full.names = TRUE)
+
+filenames_season <- str_extract(filenames, "\\d{4}")
+filenames_week <- str_extract(filenames, "week\\d+") %>% 
+  str_remove("week") %>% 
+  as.integer()
+
+objectNames_projectedPoints <- paste("players_projectedPoints_weekly_", filenames_season, "_", "week", filenames_week, sep = "")
+objectNames_projectedPoints_combined <- paste("players_projectedPoints_weekly_combined_", filenames_season, "_", "week", filenames_week, sep = "")
+objectNames_projections <- paste("players_projections_weekly_average_", filenames_season, "_", "week", filenames_week, sep = "")
+
+## Loop by Season ----
+for(i in unique(filenames_season)){
+  filenames_subset <- filenames[which(filenames_season %in% i)]
+  filenames_week_subset <- filenames_week[which(filenames_season %in% i)]
+  
+  objectNames_projectedPoints_subset <- objectNames_projectedPoints[which(filenames_season %in% i)]
+  objectNames_projectedPoints_combined_subset <- objectNames_projectedPoints_combined[which(filenames_season %in% i)]
+  objectNames_projections_subset <- objectNames_projections[which(filenames_season %in% i)]
+  
+  ### Loop by Week ----
+  for(j in 1:length(filenames_week_subset)){
+    rm(players_projectedPoints_weekly)
+    rm(players_projectedPoints_weekly_combined)
+    rm(players_projections_weekly_average)
+    
+    objectName_projectedPoints <- objectNames_projectedPoints_subset[j] #paste("players_projectedPoints_weekly_", i, "_", "week", j, sep = "")
+    objectName_projectedPoints_combined <- objectNames_projectedPoints_combined_subset[j]
+    objectName_projections <- objectNames_projections_subset[j] #paste("players_projections_weekly_average_", i, "_", "week", j, sep = "")
+    fullFilepath <- filenames_subset[j] #paste("./data/by_year/", objectName_projectedPoints, ".RData", sep = "")
+    
+    load(file = fullFilepath)
+    
+    # Projected Points
[…]

### 2025-04-05 decision-making.qmd
20250405 - wisdom of the crowd

@@
 
-Here is a code example of when the crowd-averaged predictions are centered on the truth for a player.
+Here is a code example of when the individual predictions (from which the crowd-averaged predictions are derived) are centered on the truth for a player.
 The distribution of projected points versus truth is depicted in @fig-bracketingCentered.
@@
 )
+
+bracketingCentered_bracketingRate <- bracketingCentered[which(row.names(bracketingCentered) == "crowdAveraged"), "bracketingRate"] * 100
 
@@
 
-As demonstrated above, when the crowd-averaged predictions are centered on the truth, the crowd-averaged predictions are perfectly accurate and are more accurate (MAE/MdAE = `{r} maeBracketingCentered_crowd`) than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingCentered_individual`; MdAE = `{r} mdaeBracketingCentered_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
+As demonstrated above, when the individual predictions are centered on the truth, the bracketing rate is very high (in this case `{r} petersenlab::apa(bracketingCentered_bracketingRate, decimals = 2)`%), and the crowd-averaged predictions are perfectly accurate and are more accurate (MAE/MdAE = `{r} maeBracketingCentered_crowd`) than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingCentered_individual`; MdAE = `{r} mdaeBracketingCentered_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
 
-Here is a code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player, with a high bracketing rate.
+Here is a code example of when the individual predictions bracket but are not centered on the truth for a player, with a high bracketing rate and no strong outliers.
 The distribution of projected points versus truth is depicted in @fig-bracketingNotCentered1.
@@
 )
+
+bracketingNotCentered1_bracketingRate <- bracketingNotCentered1[which(row.names(bracketingNotCentered1) == "crowdAveraged"), "bracketingRate"] * 100
 
@@
 
-As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth, and the bracketing rate is high, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingNotCentered1_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered1_individual`; MdAE = `{r} mdaeBracketingNotCentered1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
+As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case `{r} petersenlab::apa(bracketingNotCentered1_bracketingRate, decimals = 2)`%) with no strong outliers, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingNotCentered1_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered1_individual`; MdAE = `{r} mdaeBracketingNotCentered1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
 
-Here is another code example of when the crowd-averaged predictions bracket but are not centered on the truth for a player, with a low bracketing rate.
-The distribution of projected points versus truth is depicted in @fig-bracketingNotCentered2.
+Here is a code example of when the individual predictions bracket but are not centered on the truth for a player, with a high bracketing rate but with an outlier projection that is far away from the truth.
+The distribution of projected points versus truth is depicted in @fig-bracketingnotCentered2.
 
 ```{r}
-predictedValues <- c(10,29,29,40,60,70,80,90)
-actualValue <- 15
+predictedValues <- c(40,60,70,80,90,100,110,250)
+actualValue <- 74
 errors <- predictedValues - actualValue
@@
 ```{r}
-#| label: fig-bracketingNotCentered2
+#| label: fig-bracketingnotCentered2
 #| fig-cap: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
@@
 
-bracketingNotCentered2 <- petersenlab::wisdomOfCrowd(
+bracketingnotCentered2 <- petersenlab::wisdomOfCrowd(
   predicted = predictedValues,
@@
 
-maeBracketingNotCentered2_individual <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "individual"), "MAE"]
-maeBracketingNotCentered2_crowd <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "crowdAveraged"), "MAE"]
+bracketingnotCentered2_bracketingRate <- bracketingnotCentered2[which(row.names(bracketingnotCentered2) == "crowdAveraged"), "bracketingRate"] * 100
 
-mdaeBracketingNotCentered2_individual <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "individual"), "MdAE"]
-meBracketingNotCentered2_crowd <- bracketingNotCentered2[which(row.names(bracketingNotCentered2) == "crowdAveraged"), "ME"]
+maeBracketingnotCentered2_individual <- bracketingnotCentered2[which(row.names(bracketingnotCentered2) == "individual"), "MAE"]
+maeBracketingnotCentered2_crowd <- bracketingnotCentered2[which(row.names(bracketingnotCentered2) == "crowdAveraged"), "MAE"]
+
+mdaeBracketingnotCentered2_individual <- bracketingnotCentered2[which(row.names(bracketingnotCentered2) == "individual"), "MdAE"]
+meBracketingnotCentered2_crowd <- bracketingnotCentered2[which(row.names(bracketingnotCentered2) == "crowdAveraged"), "ME"]
 
@@
     "text",
-    x = actualValue + (meBracketingNotCentered2_crowd / 2),
-    y = 0.0003,
-    label = paste("Mean Error: ", meBracketingNotCentered2_crowd, sep = ""),
+    x = actualValue + (meBracketingnotCentered2_crowd / 2),
+    y = 0.0007,
+    label = paste("Mean Error: ", meBracketingnotCentered2_crowd, sep = ""),
     hjust = 0.5 # center-justify
@@
 
-As demonstrated above, when the crowd-averaged predictions bracket but are not centered on the truth and the bracketing rate is low, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingNotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy (MdAE = `{r} mdaeBracketingNotCentered2_individual`).
+As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case `{r} petersenlab::apa(bracketingnotCentered2_bracketingRate, decimals = 2)`%) but with an outlier projection that is far away from the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy ( MdAE = `{r} mdaeBracketingnotCentered2_individual`).
 
-Here is a code example of when the crowd-averaged predictions do not bracket the truth for a player, but the distribution is still close to truth.
+Here is another code example of when the individual predictions bracket but are not centered on the truth for a player, with a low bracketing rate.
+The distribution of projected points versus truth is depicted in @fig-bracketingnotCentered3.
+
+```{r}
+predictedValues <- c(10,29,29,40,60,70,80,90)
+actualValue <- 14
+errors <- predictedValues - actualValue
+
+errors
+mean(predictedValues)
+
+petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+```
+
+```{r}
+#| label: fig-bracketingnotCentered3
+#| fig-cap: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
+#| fig-alt: "Example Distribution of Projected Fantasy Points that Brackets But is Not Centered on the Truth."
+#| code-fold: true
+
+bracketingnotCentered3 <- petersenlab::wisdomOfCrowd(
+  predicted = predictedValues,
+  actual = actualValue
+)
+
+bracketingnotCentered3_bracketingRate <- bracketingnotCentered3[which(row.names(bracketingnotCentered3) == "crowdAveraged"), "bracketingRate"] * 100
+
+maeBracketingnotCentered3_individual <- bracketingnotCentered3[which(row.names(bracketingnotCentered3) == "individual"), "MAE"]
+maeBracketingnotCentered3_crowd <- bracketingnotCentered3[which(row.names(bracketingnotCentered3) == "crowdAveraged"), "MAE"]
+
+mdaeBracketingnotCentered3_individual <- bracketingnotCentered3[which(row.names(bracketingnotCentered3) == "individual"), "MdAE"]
+meBracketingnotCentered3_crowd <- bracketingnotCentered3[which(row.names(bracketingnotCentered3) == "crowdAveraged"), "ME"]
+
+ggplot2::ggplot(,
+  mapping = aes(
+    x = predictedValues)
+) +
+  geom_density(
+    fill = "lightgray"
+  ) +
+  geom_vline(
+    xintercept = mean(predictedValues),
+    color = "darkgray",
+    linewidth = 1.5) +
+    geom_vline(
+    xintercept = actualValue,
+    linetype = "dashed",
+    linewidth = 1.5
+  ) +
+  annotate(
+    "segment",
+    x = mean(predictedValues),
+    xend = actualValue,
+    y = 0,
+    yend = 0,
+    linewidth = 1.5,
+    arrow = arrow(
+      angle = 20,
+      ends = "both",
+      type = "closed")
+  ) +
+  annotate(
+    "text",
+    x = actualValue + (meBracketingnotCentered3_crowd / 2),
+    y = 0.0003,
+    label = paste("Mean Error: ", meBracketingnotCentered3_crowd, sep = ""),
+    hjust = 0.5 # center-justify
+  ) +
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Predicted Fantasy Points for a Player",
+    subtitle = "Crowd-Averaged Projection (Vertical Gray Line) Vs.\nTruth, i.e., Actual Fantasy Points (Dashed Line)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+As demonstrated above, when the individual predictions bracket but are not centered on the truth and the bracketing rate is low (in this case `{r} petersenlab::apa(bracketingnotCentered3_bracketingRate, decimals = 2)`%), the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered3_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered3_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy (MdAE = `{r} mdaeBracketingnotCentered3_individual`).
+
+Here is a code example of when the individual predictions do not bracket the truth for a player, and there are no strong outliers.
 The distribution of projected points versus truth is depicted in @fig-noBracketing1.
@@
 
-As demonstrated above, when the crowd-averaged predictions do not bracket the truth but are still close to the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeNoBracketing1_crowd`) are as accurate as the accuracy of the typical individual forecaster (MAE = `{r} maeNoBracketing1_individual`; MdAE = `{r} mdaeNoBracketing1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
+As demonstrated above, when the individual predictions do not bracket the truth (i.e., the bracketing rate is 0%) and there are no strong outliers, the crowd-averaged predictions (MAE/MdAE = `{r} maeNoBracketing1_crowd`) are as accurate as the accuracy of the typical individual forecaster (MAE = `{r} maeNoBracketing1_individual`; MdAE = `{r} mdaeNoBracketing1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
 
@@
 ```{r}
-predictedValues <- c(0,65,70,75,80,85,90,95)
+predictedValues <- c(0,65,67,70,80,85,90,95)
 actualValue <- 100
@@
 When the crowd-averaged predictions bracket but are not centered on the truth, the crowd-averaged predictions are more accurate than the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
-If the bracketing rate is high, the crowd-averaged predictions will be more accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than most individual forecasters; however, if the bracketing rate is low, the crowd-averaged predictions can be less accurate than than most individual forecasters.
+If the bracketing rate is high and there are no strong outliers, the crowd-averaged predictions will also be more accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than most individual forecasters; however, if the bracketing rate is low or there are strong outliers, the crowd-averaged predictions can be less accurate than than most individual forecasters.
+Outliers could make the crowd-averaged prediction more or less accurate, depending on whether the outlying prediction is closer to or farther away from the truth, compared to the other predictions.
 When the crowd-averaged predictions do not bracket the truth, the crowd-averaged predictions are as accurate as the accuracy of the typical individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError).
-Depending on how far some of the individual projections are from the truth, the crowd-averaged projections can be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than most individual forecasters.
+When the crowd-averaged predictions do not bracket the truth and there are outliers, the crowd-averaged projections can be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than most individual forecasters (again, depending on whether the outlying prediction is closer to or farther away from the truth, compared to the other predictions).
 
@@
 However, the crowd-averaged projections may be less accurate than any *individual* projection, including the projection of the forecaster who is at the 50th percentile in accuracy.
-That is, depending on the bracketing rate, the crowd-averaged projections can still be less accurate than most individual projections.
+That is, if there is a low bracketing rate or inaccurate outliers, the crowd-averaged projections can be less accurate than most individual projections.
 In general, the higher the bracketing rate and the more closely that the projections are centered on the truth, the more accurate the crowd-averaged projections will be.
+
+There may be multiple ways of handling outlier predictions.
+For instance, instead of computing a simple average of the predictions, one could calculate a [weighted average](#sec-descriptiveStatisticsCenter) that weights each prediction according to the historical accuracy of the prediction source.
+Alternatively, one could calculate the crowd-averaged prediction using an index of the center of a distribution that is less sensitive to outliers, such as a [median](#sec-descriptiveStatisticsCenter) or robust average such as the [Hodges-Lehmann statistic](#sec-descriptiveStatisticsCenter) (aka pseudomedian).
 

### 2025-04-05 basic-statistics.qmd
20250405 - descriptive stats

@@
 exampleValues <- c(0, 0, 10, 15, 20, 30, 1000)
-exampleValues_mean <- apa(mean(exampleValues), 2)
+exampleValues_mean <- mean(exampleValues)
 ```
@@
 One issue with the mean is that it is sensitive to extreme (outlying) values.
-For instance, the mean of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_mean`.
+For instance, the mean of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} petersenlab::apa(exampleValues_mean, decimals = 2)`.
 
@@
 The pseudomedian of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_pseudomedian`.
+The pseudomedian is an example of a robust average—an index of the center of a distribution that is less sensitive to outliers.
 
@@
 
+- mean:
+
 ```{r}
 mean(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- median:
+
+```{r}
 median(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- Hodges-Lehmann statistic (aka pseudomedian):
+
+```{r}
 DescTools::HodgesLehmann(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- mode:
+
+```{r}
 petersenlab::Mode(player_stats_seasonal$fantasyPoints)
-
+```
+
+- weighted mean:
+
+```{r}
 weighted.mean(
@@
   na.rm = TRUE)
-
+```
+
+- weighted median:
+
+```{r}
 petersenlab::wthdquantile(
@@
 
-```{r}
-
+- standard deviation:
+
+```{r}
+sd(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- variance:
+
+```{r}
+var(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- range:
+
+```{r}
+range(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+
+max(player_stats_seasonal$fantasyPoints, na.rm = TRUE) - min(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- minimum and maximum:
+
+```{r}
+min(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+max(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- interquartile range (IQR):
+
+```{r}
+IQR(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- median absolute deviation:
+
+```{r}
+mad(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
 ```
@@
 
-```{r}
-
+- skewness:
+
+```{r}
+psych::skew(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- kurtosis:
+
+```{r}
+psych::kurtosi(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
 ```
@@
 player_stats_seasonal %>% 
-  select(age, years_of_experience, fantasyPoints) %>% 
-  summarise(across(
+  dplyr::select(age, years_of_experience, fantasyPoints) %>% 
+  dplyr::summarise(across(
       everything(),
@@
       .names = "{.col}.{.fn}")) %>%
-    pivot_longer(
+    tidyr::pivot_longer(
       cols = everything(),
@@
       names_sep = "\\.") %>% 
-    pivot_wider(
+    tidyr::pivot_wider(
       names_from = index,

### 2025-04-05 getting-started.qmd
20250405 - specify package for function

@@
 players %>%
-  select(name, age)
-
-players %>%
-  select(name:age)
-
-players %>%
-  select(all_of(subsetVars))
+  dplyr::select(name, age)
+
+players %>%
+  dplyr::select(name:age)
+
+players %>%
+  dplyr::select(all_of(subsetVars))
 ```
@@
 players %>%
-  select(-name, -age)
-
-players %>%
-  select(-c(name:age))
-
-players %>%
-  select(-all_of(dropVars))
+  dplyr::select(-name, -age)
+
+players %>%
+  dplyr::select(-c(name:age))
+
+players %>%
+  dplyr::select(-all_of(dropVars))
 ```
@@
 players %>%
-  filter(position == "RB") %>%
-  select(all_of(subsetVars))
+  dplyr::filter(position == "RB") %>%
+  dplyr::select(all_of(subsetVars))
 ```
@@
 players %>%
-  mutate(across(c(
+  dplyr::mutate(across(c(
     oldVar1:oldVar2),
@@
 players <- players %>% 
-  rename(
+  dplyr::rename(
     newVar1 = oldVar1,
@@
 players <- players %>% 
-  rename_with(~ varNamesTo, all_of(varNamesFrom))
+  dplyr::rename_with(~ varNamesTo, all_of(varNamesFrom))
 ```
@@
 players %>%
-  mutate(across(c(
+  dplyr::mutate(across(c(
     ID,
@@
 players %>%
-  mutate(across(
+  dplyr::mutate(across(
     age:newVar1,
@@
 players %>%
-  mutate(across(where(is.factor), as.character))
+  dplyr::mutate(across(
+    where(is.factor),
+    as.character))
 ```
@@
 
-mergeExample1 <- full_join(
+mergeExample1 <- dplyr::full_join(
   mergeExample1A,
@@
 
-mergeExample2 <- full_join(
+mergeExample2 <- dplyr::full_join(
   mergeExample2A,
@@
 
-mergeExample3 <- full_join(
+mergeExample3 <- dplyr::full_join(
   mergeExample3A,
@@
 ```{r}
-fullJoinData <- full_join(
+fullJoinData <- dplyr::full_join(
   players,
@@
 ```{r}
-leftJoinData <- left_join(
+leftJoinData <- dplyr::left_join(
   players,
@@
 ```{r}
-rightJoinData <- right_join(
+rightJoinData <- dplyr::right_join(
   players,
@@
 ```{r}
-innerJoinData <- inner_join(
+innerJoinData <- dplyr::inner_join(
   players,
@@
 ```{r}
-semiJoinData <- semi_join(
+semiJoinData <- dplyr::semi_join(
   players,
@@
 ```{r}
-antiJoinData <- anti_join(
+antiJoinData <- dplyr::anti_join(
   players,
@@
 ```{r}
-crossJoinData <- cross_join(
+crossJoinData <- dplyr::cross_join(
   players,
@@
 ```{r}
-dataLong <- full_join(
-  players %>% select(-age),
+dataLong <- dplyr::full_join(
+  players %>% dplyr::select(-age),
   fantasyPoints_weekly,
@@
 dataWide <- dataLong %>% 
-  pivot_wider(
+  tidyr::pivot_wider(
     names_from = c(season, week),
@@
 dataLong <- dataWide %>% 
-  pivot_longer(
+  tidyr::pivot_longer(
     cols = fantasyPoints_2022_week1:fantasyPoints_2023_week17,

### 2025-04-05 decision-making.qmd
20250405 - clarifications

@@
 
-As demonstrated above, when the individual predictions are centered on the truth, the bracketing rate is very high (in this case `{r} petersenlab::apa(bracketingCentered_bracketingRate, decimals = 2)`%), and the crowd-averaged predictions are perfectly accurate and are more accurate (MAE/MdAE = `{r} maeBracketingCentered_crowd`) than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingCentered_individual`; MdAE = `{r} mdaeBracketingCentered_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
+As demonstrated above, when the individual predictions are centered on the truth, the bracketing rate is very high (in this case, `{r} petersenlab::apa(bracketingCentered_bracketingRate, decimals = 2)`%), and the crowd-averaged predictions are perfectly accurate and are more accurate (MAE/MdAE = `{r} maeBracketingCentered_crowd`) than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingCentered_individual`; MdAE = `{r} mdaeBracketingCentered_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
 
@@
 
-As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case `{r} petersenlab::apa(bracketingNotCentered1_bracketingRate, decimals = 2)`%) with no strong outliers, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingNotCentered1_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered1_individual`; MdAE = `{r} mdaeBracketingNotCentered1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
+As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case, `{r} petersenlab::apa(bracketingNotCentered1_bracketingRate, decimals = 2)`%) with no strong outliers, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingNotCentered1_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingNotCentered1_individual`; MdAE = `{r} mdaeBracketingNotCentered1_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError) and [median absolute error](#sec-medianAbsoluteError).
 
@@
 
-As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case `{r} petersenlab::apa(bracketingnotCentered2_bracketingRate, decimals = 2)`%) but with an outlier projection that is far away from the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy ( MdAE = `{r} mdaeBracketingnotCentered2_individual`).
+As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case, `{r} petersenlab::apa(bracketingnotCentered2_bracketingRate, decimals = 2)`%) but with an outlier projection that is far away from the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy ( MdAE = `{r} mdaeBracketingnotCentered2_individual`).
 
@@
 
-As demonstrated above, when the individual predictions bracket but are not centered on the truth and the bracketing rate is low (in this case `{r} petersenlab::apa(bracketingnotCentered3_bracketingRate, decimals = 2)`%), the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered3_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered3_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy (MdAE = `{r} mdaeBracketingnotCentered3_individual`).
+As demonstrated above, when the individual predictions bracket but are not centered on the truth and the bracketing rate is low (in this case, `{r} petersenlab::apa(bracketingnotCentered3_bracketingRate, decimals = 2)`%), the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered3_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered3_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy (MdAE = `{r} mdaeBracketingnotCentered3_individual`).
 

### 2025-04-05 evaluating-prediction-accuracy.qmd
20250405 - clarifications

@@
 
-Mean absolute error (MAE) is the mean of the absolute value of differences between the predicted and observed values across individuals, that is, the mean of the absolute value of errors.
+Mean absolute error (MAE) is the mean of the absolute value of differences between the predicted and observed values across individuals.
+That is, MAE is the mean of the absolute value of errors (i.e., the mean distance between predicted and observed values).
 Smaller MAE values (closer to zero) reflect greater accuracy.
@@
 
-Median absolute error (MdAE) is the median of the absolute value of differences between the predicted and observed values across individuals, that is, the median of the absolute value of errors.
+Median absolute error (MdAE) is the median of the absolute value of differences between the predicted and observed values across individuals.
+That is, MdAE is the median of the absolute value of errors (i.e., the median distance between predicted and observed values).
 Smaller MdAE values (closer to zero) reflect greater accuracy.

### 2025-04-05 machine-learning.qmd
20250405 - clarifications

@@
     - deep learning
+- ensemble
 

### 2025-04-05 sports-cognitive-psychology.qmd
20250405 - video citation

@@
 
-For instance, here is a 2016 video clip of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops", suggesting that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
+For instance, here is a [2015 video clip](https://streamable.com/1hw8) of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops" [@ABC2015], suggesting that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
+The full game is [here](https://www.youtube.com/watch?v=1L9D2rp0mk8) [@ABC2015a].
 
-![Football Commentator Making Fun of Analytics in Football. From: <https://streamable.com/1hw8> (2016).](images/commentatorOnAnalytics.mp4){width=100%}\
+![Football Commentator Making Fun of Analytics in Football. From: <https://streamable.com/1hw8> (2015).](images/commentatorOnAnalytics.mp4){width=100%}\
 
@@
 
-Nevertheless, more and more people around the league are using analytics—likely because it works; for examples of articles, see [here](https://www.theringer.com/2018/12/19/nfl/nfl-analytics-revolution) [@Clark2018; archived at <https://perma.cc/7BKP-A4GJ>] and [here](https://www.samford.edu/sports-analytics/fans/2020/The-Future-of-NFL-Data-Analytics) [@Awbrey2020; archived at <https://perma.cc/WXE3-53E6>].
+Nevertheless, more and more people around the league use analytics—likely because it works; for examples of articles, see [here](https://www.theringer.com/2018/12/19/nfl/nfl-analytics-revolution) [@Clark2018; archived at <https://perma.cc/7BKP-A4GJ>] and [here](https://www.samford.edu/sports-analytics/fans/2020/The-Future-of-NFL-Data-Analytics) [@Awbrey2020; archived at <https://perma.cc/WXE3-53E6>].
 Head Coach Doug Peterson heavily relied on analytics with the Philadelphia Eagles to help them win the 2017 Super Bowl; for more information, see [here](https://www.forbes.com/sites/liamfox/2021/08/12/how-the-nfl-uses-analytics-according-to-the-lead-analyst-of-a-super-bowl-champion/) [@Fox2021; archived at <https://perma.cc/49KQ-R785>] and [here](https://www.nfl.com/news/super-bowl-lii-how-the-2017-philadelphia-eagles-were-built-0ap3000000912753) [@Rosenthal2018; archived at <https://perma.cc/2GRF-8Z5K>].

### 2025-04-05 decision-making.qmd
20250405 - additions

@@
 
-As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case, `{r} petersenlab::apa(bracketingnotCentered2_bracketingRate, decimals = 2)`%) but with an outlier projection that is far away from the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy ( MdAE = `{r} mdaeBracketingnotCentered2_individual`).
+As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case, `{r} petersenlab::apa(bracketingnotCentered2_bracketingRate, decimals = 2)`%) but with an outlier projection that is far away from the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy (MdAE = `{r} mdaeBracketingnotCentered2_individual`).
 
@@
 
-## Strategies for Managing Risk and Uncertainty {#sec-managinRiskUncertainty}
+## Strategies for Managing Risk and Uncertainty {#sec-managingRiskUncertainty}
 
@@
 
+The question of whether fantasy football is a game of chance or skill is important because such considerations help determine whether betting on one's performance is considered gambling, which might be illegal or regulated in many jurisdictions.
 Fantasy football (and sports betting, more generally) is not 100% luck.
@@
 To evaluate the percentage of variability in fantasy football that is attributable to luck versus skill, @Getty2018 evaluated the extent to which one's performance in the first half of the season was correlated with one's performance in the second half of the season, under the assumption that underlying skill would lead to persistence of performance across time.
-The authors examined FanDuel daily fantasy sports (DFS) leagues, which allow the user to set an entirely new lineup each week from all players, based on salary constraints ( so a user cannot just select the best player at every position).
+The authors examined FanDuel daily fantasy sports (DFS) leagues, which allow the user to set an entirely new lineup each week from all players, based on salary constraints (so a user cannot just select the best player at every position).
 The authors estimated that ~55% of the variability in fantasy football performance across time was due to skill and 45% was due to luck [@Getty2018].
@@
 
-- seek advice from diverse perspectives and incorporate it into your decision making
-- try to get opinions from others *before* you state your perspective and before the various sources of advice discuss with each other, to ensure independence of advice
+- Seek advice from diverse perspectives and incorporate it into your decision making.
+- Try to get opinions from others *before* you state your perspective and before the various sources of advice discuss with each other, to ensure independence of advice.
   - As noted by @Kahneman2011, "before an issue is discussed, all members of the committee should be asked to write a very brief summary of their position...
@@
 85).
-- if some sources of advice (or projections) are clearly more skilled and accurate than others, you can average this "select" crowd of projections or give them greater weight in a weighted average
-- if it is unclear whether or which sources are reliably more accurate than others, using a simple average across all sources (i.e., crowd-averaged projections) can be a useful approach that is as accurate as—if not more accurate than—the average individual forecaster
-- incorporate at least 5–10 sources of projections
+- If some sources of advice (or projections) are clearly more skilled and accurate than others, you can average this "select" crowd of projections or give them greater weight in a [weighted average]().
+- If it is unclear whether or which sources are reliably more accurate than others, using a simple average across all sources (i.e., crowd-averaged projections) can be a useful approach that is as accurate as—if not more accurate than—the average individual forecaster.
+- Incorporate at least 5–10 sources of projections.
+Use a [weighted](#sec-descriptiveStatisticsCenter) or [robust](#sec-descriptiveStatisticsCenter) average to account for outlier projections.
+- Do not bet on fantasy football (or anything else for that matter) unless you are willing to lose the money.
+Sports bettors tend to be [overconfident](#sec-cognitiveBiasesOverconfidence); on average, they lose 7.5 cents for every dollar wagered [@Brown2025].
+If you are going to bet, only bet a small portion of your money and never more than you are willing to lose.
+And do not make multi-leg bets (parlays); people lose on average 25 cents for every dollar wagered on parlays [@Brown2025].
 

### 2025-04-05 causal-inference.qmd
20250405 - edits

@@
 In each of these examples, you are controlling for player education for purposes of examining the effect of sports drink consumption on performance.
-Failure to control for important third variables can lead to erroneous conclusions, as evidenced by the association depicted in @fig-simpsonParadox.
+Failure to control for important third variables can lead to erroneous conclusions, as evidenced by the association depicted in @fig-simpsonParadox, which is an example of [Simpson's paradox](#sec-simpsonsParadox) described in @sec-simpsonsParadox.
 In the example, if we did not control for gender, we would infer that there is a positive association between dosage and recovery probability.
-However, when we examine each men and women separately, we learn that the association between dosage and recovery probability is actually negative within each gender group.
+However, when we examine each men and women separately, we learn that the association between dosage and recovery probability is actually *negative* within each gender group.
 Thus, in this case, failure to control for gender would lead to false inferences about the association between dosage and recovery probability.

### 2025-04-05 cognitive-bias.qmd
20250318 - increasing shot difficulty
20250405 - edits

@@
 > 1. Scientifically supported measurement strategies are used based on their psychometrics, including [reliability](#sec-reliability) and [validity](#sec-validity).
+
+By contrast, according to @Lilienfeld2015a, some of the frequent features of pseudoscience include:
+
+> 1. An overuse of ad hoc hypotheses designed to immunize claims from falsification
+> 1. Absence of self-correction
+> 1. Evasion of peer review
+> 1. Emphasis on confirmation rather than refutation
+> 1. Reversed burden of proof
+> 1. Absence of connectivity
+> 1. Overreliance on testimonial and anecdotal evidence
+> 1. Use of obscurantist language
+> 1. Absence of boundary conditions
+> 1. The mantra of holism
 
@@
 However, if any such effect exists, the hot hand may be limited to a small subset of players and the [effect size](#sec-practicalSignificance) of any hot hand effect appears to be small [@Pelechrinis2022].
+Indeed, the slight potential increase in shooting ability tends to be offset by the tendency that "hot" players have to increase their shot difficulty after making a previous shot, so the net hot hand effect appears to be "vanishingly small" [@Partnow2021].
 

### 2025-04-05 correlation.qmd
20250405 - edits

@@
 The formula for the correlation coefficient\index{correlation} is in @eq-pearsonCorrelation.
+As noted in @sec-statisticalTestsPartitionedVariance, the correlation coefficient can be thought of as the ratio of shared variance (i.e., covariance) to total variance, as in @eq-correlationPartitionedVariance.
 
@@
 
-A nonlinear correlation can be estimated using the `XICOR` package [@R-XICOR; @Chatterjee2023_packages]:
+A nonlinear correlation can be estimated using the $\xi$ coefficient from the `XICOR` package [@R-XICOR; @Chatterjee2023_packages]:
 

### 2025-04-05 multiple-regression.qmd
20250227 - clean up
20250405 - edits

@@
 
+### Evaluating the Assumptions of Multiple Regression {#sec-assumptionsRegressionEvaluating}
+
 ## Coefficient of Determination ($R^2$) {#sec-multipleRegressionRSquared}
 
-The coefficient of determination ($R^2$) reflects the proportion of variance in the [outcome (dependent) variable](#sec-correlationalStudy) that is explained by the model predictions: $R^2 = \frac{\text{variance explained in }Y}{\text{total variance in }Y}$.
+As noted in @sec-statisticalTestsPartitionedVariance, 
+The coefficient of determination ($R^2$) reflects the proportion of variance in the [outcome (dependent) variable](#sec-correlationalStudy) that is explained by the model predictions, as in @eq-coefficientOfDeterminationPartitionedVariance: $R^2 = \frac{\text{variance explained in }Y}{\text{total variance in }Y}$.
 Various formulas for $R^2$ are in @eq-rSquared.
@@
 If we were to apply the same algorithm based on the 2023 prediction model to 2024 data, the prediction model would likely predict less accurately than with 2023 data.
-The regression coefficients in the
+The regression coefficients in the [FILL IN]
 

### 2025-04-05 correlation.qmd
20250405 - updates

@@
 
-An interactive visualization by @Magnusson2020 on interpreting correlations is at the following link: <https://rpsychologist.com/correlation/> (archived at <https://perma.cc/G8YR-VCM4>)
+An interactive visualization by @Magnusson2023 on interpreting correlations is at the following link: <https://rpsychologist.com/correlation/> (archived at <https://perma.cc/G8YR-VCM4>)
 
@@
 
-A nonlinear correlation can be estimated using the $\xi$ coefficient from the `XICOR` package [@R-XICOR; @Chatterjee2023_packages]:
+A nonlinear correlation can be estimated using the xi ($\xi$) coefficient from the `XICOR` package [@R-XICOR; @Chatterjee2023_packages]:
 

### 2025-04-05 getting-started.qmd
20250405 - edits
20250405 - updates

@@
 
-Below is a visual that depicts various types of merges/joins.
+@fig-joinTypes depicts various types of merges/joins.
 Object `x` is the circle labeled as `x`.
@@
 
-![Types of merges/joins](images/joinTypes.png)\
+::: {#fig-joinTypes}
+![](images/joinTypes.png){width=100% fig-alt="Types of Merges/Joins."}
+
+Types of Merges/Joins.
+:::
 
@@
 At the end of each chapter in which `R` code is used, I provide the session information, which describes the system and operating system the code was run on and the versions of each package.
-That way, if you get different results from me, you can see which differ, to help with reproducibility.
-If you run the (all of) the exact same code as is provided in the text, in the exact same order, with the exact same setup (platform, operating system, package versions, etc.), you should get the exact same answer as is in the text.
+That way, if you get different results from me, you can see which session settings differ, to help with reproducibility.
+If you run the (all of) the exact same code as is provided in the text, in the exact same order, with the exact same setup (platform, operating system, package versions, etc.) and the exact same data, you should get the exact same answer as is in the text.
 That is the idea of reproducibility—getting the exact same result with the exact same inputs.

### 2025-04-06 simulation.qmd
Built site for gh-pages

@@
 Monte Carlo simulation involves repeated random sampling from a known distribution [@Sigal2016].
-For instance, if we know the population distribution for the likely outcomes for a player—e.g., a normal distribution with a known mean and standard deviation—we can repeatedly sample randomly from this distribution.
+For instance, if we know the population distribution for the likely outcomes for a player—e.g., a normal distribution with a known mean (e.g., 150 points) and standard deviation (e.g., 20 points)—we can repeatedly sample randomly from this distribution.
 

### 2025-04-06 correlation.qmd
20250405 - edits

@@
 Covariance is the association between variables and in an unstandardized metric that differs for variables with different scales.
-By contrast, correlation is in a standarized metric that does not differ for variables with different scales.
+By contrast, correlation is in a standardized metric that does not differ for variables with different scales.
 When examining the association between variables that are [interval](#sec-interval) or [ratio](#sec-ratio) levels of measurement, Pearson correlation is used.
@@
 
-A nonlinear correlation can be estimated using the xi ($\xi$) coefficient from the `XICOR` package [@R-XICOR; @Chatterjee2023_packages]:
+A nonlinear correlation can be estimated using the xi ($\xi$) coefficient from the `XICOR` package [@R-XICOR; @Chatterjee2021]:
 
@@
 ## Impact of Outliers {#sec-correlationOutliers}
+
+## Impact of Restricted Range {#sec-correlationRestrictedRange}
 

### 2025-04-06 download-football-data.qmd
20250406 - career stats

@@
 
+#### Season-by-Season Statistics {#sec-calculateActualSeasonalStats}
+
+First, we can compute the players' season-by-season statistics.
+
+TODO: Save/update data file in repo with the data generated from this code, and follow the data file through the rest of the book, updating code as necessary.
+
+::: {#nte-downloadSeasonalActualStats .callout-note title="Downloading season-by-season statistics"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+nfl_actualStats_seasonal_player_raw <- nflfastR::calculate_stats(
+  seasons = TRUE,
+  summary_level = "season",
+  season_type = "REG")
+
+nfl_actualStats_seasonal_team_raw <- nflfastR::calculate_stats(
+  seasons = TRUE,
+  summary_level = "season",
+  stat_type = "team",
+  season_type = "REG")
+
+nfl_actualStats_seasonal_player_inclPost_raw <- nflfastR::calculate_stats(
+  seasons = TRUE,
+  summary_level = "season",
+  season_type = "REG+POST")
+
+nfl_actualStats_seasonal_team_inclPost_raw <- nflfastR::calculate_stats(
+  seasons = TRUE,
+  summary_level = "season",
+  stat_type = "team",
+  season_type = "REG+POST")
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_seasonal_player_raw, nfl_actualStats_seasonal_team_raw,
+  nfl_actualStats_seasonal_player_inclPost_raw, nfl_actualStats_seasonal_team_inclPost_raw,
+  file = "./data/nfl_actualStats_seasonal_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_seasonal_raw.RData")
+```
+
+A Data Dictionary for the variables is available in the `nfl_stats_variables` object that is returned when running the `calculate_stats()` function:
+
+```{r}
+nfl_stats_variables
+```
+
+The `nfl_actualStats_seasonal_player_raw` object is in `player`-`season` form.
+That is, each row should be uniquely identified by the combination of `player_id` and `season`.
+The `nfl_actualStats_seasonal_team_raw` object is in `team`-`season` form.
+That is, each row should be uniquely identified by the combination of `team` and `season`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_seasonal_player <- nfl_actualStats_seasonal_player_raw %>% 
+  rename(team = recent_team) %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_seasonal_team <- nfl_actualStats_seasonal_team_raw %>% 
+  select(season, team, everything()) %>% 
+  arrange(season, team)
+
+nfl_actualStats_seasonal_player_inclPost <- nfl_actualStats_seasonal_player_inclPost_raw %>% 
+  rename(team = recent_team) %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_seasonal_team_inclPost <- nfl_actualStats_seasonal_team_inclPost_raw %>% 
+  select(season, team, everything()) %>% 
+  arrange(season, team)
+```
+
+Let's check for duplicate `player` instances:
+
+```{r}
+nfl_actualStats_seasonal_player %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_seasonal_team %>% 
+  group_by(team, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_seasonal_player_inclPost %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_seasonal_team_inclPost %>% 
+  group_by(team, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_actualStats_seasonal_player, nfl_actualStats_seasonal_team,
+  nfl_actualStats_seasonal_player_inclPost, nfl_actualStats_seasonal_team_inclPost,
+  file = "./data/nfl_actualStats_seasonal.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_seasonal.RData")
+```
+
+##### OLD (DELETE CODE BELOW)
+
+```{r}
+seasons <- unique(nfl_pbp$season)
+
+nfl_pbp_seasonalList <- list()
+nfl_actualStats_offense_seasonalList <- list()
+nfl_actualStats_defense_seasonalList <- list()
+nfl_actualStats_kicking_seasonalList <- list()
+```
+
+::: {#nte-calculateActualSeasonalStats .callout-note title="Calculating players' season-by-season statistics"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+
+pb <- txtProgressBar(
+  min = 0,
+  max = length(seasons),
+  style = 3)
+
+for(i in 1:length(seasons)){
+  # Subset play-by-play data by season
+  nfl_pbp_seasonalList[[i]] <- nfl_pbp %>% 
+    dplyr::filter(season == seasons[i])
+
+  # Compute actual statistics by season
+  nfl_actualStats_offense_seasonalList[[i]] <- 
+    nflfastR::calculate_player_stats(
+      nfl_pbp_seasonalList[[i]],
+      weekly = FALSE)
+
+  nfl_actualStats_defense_seasonalList[[i]] <- 
+    nflfastR::calculate_player_stats_def(
+      nfl_pbp_seasonalList[[i]],
+      weekly = FALSE)
+
+  nfl_actualStats_kicking_seasonalList[[i]] <- 
+    nflfastR::calculate_player_stats_kicking(
+      nfl_pbp_seasonalList[[i]],
+      weekly = FALSE)
+
+  nfl_actualStats_offense_seasonalList[[i]]$season <- seasons[i]
+  nfl_actualStats_defense_seasonalList[[i]]$season <- seasons[i]
+  nfl_actualStats_kicking_seasonalList[[i]]$season <- seasons[i]
+
+  print(
+    paste("Completed computing projections for season: ", seasons[i], sep = ""))
+
+  # Update the progress bar
+  setTxtProgressBar(pb, i)
+}
+
+# Close the progress bar
+close(pb)
+
+nfl_actualStats_offense_seasonal_raw <- nfl_actualStats_offense_seasonalList %>% 
+  dplyr::bind_rows()
+nfl_actualStats_defense_seasonal_raw <- nfl_actualStats_defense_seasonalList %>% 
+  dplyr::bind_rows()
+nfl_actualStats_kicking_seasonal_raw <- nfl_actualStats_kicking_seasonalList %>% 
+  dplyr::bind_rows()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_offense_seasonal_raw, nfl_actualStats_defense_seasonal_raw, nfl_actualStats_kicking_seasonal_raw,
+  file = "./data/nfl_actualStats_seasonal_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_seasonal_raw.RData")
+```
+
+The `nfl_actualStats_seasonal` objects are in `player`-`season` form.
+That is, each row should be uniquely identified by the combination of `player_id` and `season`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal_raw %>% 
+  rename(team = recent_team) %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal_raw %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal_raw %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+```
+
+Let's check for duplicate `player` instances:
+
+```{r}
+nfl_actualStats_offense_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_defense_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_kicking_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+Let's do some data cleanup:
+
+```{r}
+# Sum statistics across a player's years with multiple teams
+nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal %>% 
+  group_by(player_id, season) %>% 
+  summarise(
+    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
+    .groups = "drop"
+  )
+
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
+  group_by(player_id, season) %>% 
+  summarise(
+    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
+    .groups = "drop"
+  )
+
+# Re-calculate percentage stats
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
+  mutate(
+    fg_pct = fg_made / fg_att,
+    pat_pct = pat_made / pat_att
+  )
+
+# Merge data back with player info
+nfl_actualStats_defense_seasonal_raw_playerInfo <- nfl_actualStats_defense_seasonal_raw %>% 
+  select(player_id, season, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
+  unique() %>% 
+  group_by(player_id) %>% 
+  slice_head(n = 1) %>% 
+  ungroup()
+
+nfl_actualStats_kicking_seasonal_raw_playerInfo <- nfl_actualStats_kicking_seasonal_raw %>% 
+  select(player_id, season, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
+  unique() %>% 
+  group_by(player_id) %>% 
+  slice_head(n = 1) %>% 
+  ungroup()
+
+nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal_raw_playerInfo %>% 
+  right_join(
+    nfl_actualStats_defense_seasonal,
+    by = c("player_id","season")
+  )
+
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal_raw_playerInfo %>% 
+  right_join(
+    nfl_actualStats_kicking_seasonal,
+    by = c("player_id","season")
+  )
+
+# Rearrange data
+nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+```
+
+Let's check again for duplicate `player` instances:
+
+```{r}
+nfl_actualStats_offense_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_defense_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_kicking_seasonal %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  nfl_actualStats_offense_seasonal, nfl_actualStats_defense_seasonal, nfl_actualStats_kicking_seasonal,
+  file = "./data/nfl_actualStats_seasonal.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_seasonal.RData")
+```
+
 #### Career Statistics {#sec-calculateActualCareerStats}
 
-First, we can compute the players' career statistics using the `calculate_stats()` function from the `nflfastR` package [@R-nflfastR].
-
-TODO: Insert code example.
+Second, we can compute the players' career statistics using the `calculate_stats()` function from the `nflfastR` package [@R-nflfastR].
+
+```{r}
+# Compute Player Stats by Career
+playerVars <- nfl_actualStats_seasonal_player %>% 
+  select(games:fantasy_points_ppr) %>% 
+  names()
+
+weightedAverageVars_byGames <- c(
+  "passing_epa","rushing_epa","receiving_epa",
+  "target_share","air_yards_share"
+)
+
+weightedAverageVars_byAttempts <- c(
+  "passing_cpoe"
+)
+
+weightedAverageVars_byTargets <- c(
+  "wopr"
+)
+
+concatenateVars <- c(
+  "fg_made_list","fg_missed_list","fg_blocked_list","gwfg_distance_list"
+)
+
+recomputeVars <- c(
+  "pacr", # passing_yards / passing_air_yards
+  "racr", # receiving_yards / receiving_air_yards
+  "fg_pct", # fg_made / fg_att
+  "pat_pct" # pat_att / pat_made
+)
+
+sumVars <- playerVars[playerVars %ni% c(
+  weightedAverageVars_byGames, weightedAverageVars_byAttempts, weightedAverageVars_byTargets, concatenateVars, recomputeVars)]
+
+nfl_actualStats_career_player <- nfl_actualStats_seasonal_player %>% 
+  group_by(player_id) %>% 
+  summarise(
+    across(all_of(weightedAverageVars_byGames), ~ weighted.mean(.x, w = games, na.rm = TRUE)),
+    across(all_of(weightedAverageVars_byAttempts), ~ weighted.mean(.x, w = attempts, na.rm = TRUE)),
+    across(all_of(weightedAverageVars_byTargets), ~ weighted.mean(.x, w = targets, na.rm = TRUE)),
+    across(all_of(concatenateVars), ~ paste(.x[!is.na(.x)], collapse = ";")),
+    across(all_of(sumVars), ~ sum(.x, na.rm = TRUE)),
+    .groups = "drop") %>% 
+  mutate(
+    pacr = passing_yards / passing_air_yards,
+    racr = receiving_yards / receiving_air_yards,
+    fg_pct = fg_made / fg_att,
+    pat_pct = pat_att / pat_made
+  )
+
+nfl_actualStats_career_player$fg_made_list[which(nfl_actualStats_career_player$fg_made_list == "")] <- NA
+nfl_actualStats_career_player$fg_missed_list[which(nfl_actualStats_career_player$fg_missed_list == "")] <- NA
+nfl_actualStats_career_player$fg_blocked_list[which(nfl_actualStats_career_player$fg_blocked_list == "")] <- NA
+nfl_actualStats_career_player$gwfg_distance_list[which(nfl_actualStats_career_player$gwfg_distance_list == "")] <- NA
+
+nfl_actualStats_career_player_inclPost <- nfl_actualStats_seasonal_player_inclPost %>% 
+  group_by(player_id) %>% 
+  summarise(
+    across(all_of(weightedAverageVars_byGames), ~ weighted.mean(.x, w = games, na.rm = TRUE)),
+    across(all_of(weightedAverageVars_byAttempts), ~ weighted.mean(.x, w = attempts, na.rm = TRUE)),
+    across(all_of(weightedAverageVars_byTargets), ~ weighted.mean(.x, w = targets, na.rm = TRUE)),
+    across(all_of(concatenateVars), ~ paste(.x[!is.na(.x)], collapse = ";")),
+    across(all_of(sumVars), ~ sum(.x, na.rm = TRUE)),
+    .groups = "drop") %>% 
+  mutate(
+    pacr = passing_yards / passing_air_yards,
+    racr = receiving_yards / receiving_air_yards,
+    fg_pct = fg_made / fg_att,
+    pat_pct = pat_att / pat_made
+  )
+
+nfl_actualStats_career_player_inclPost$fg_made_list[which(nfl_actualStats_career_player_inclPost$fg_made_list == "")] <- NA
+nfl_actualStats_career_player_inclPost$fg_missed_list[which(nfl_actualStats_career_player_inclPost$fg_missed_list == "")] <- NA
+nfl_actualStats_career_player_inclPost$fg_blocked_list[which(nfl_actualStats_career_player_inclPost$fg_blocked_list == "")] <- NA
+nfl_actualStats_career_player_inclPost$gwfg_distance_list[which(nfl_actualStats_career_player_inclPost$gwfg_distance_list == "")] <- NA
+```
+
+Let's check for duplicate `player` instances:
+
+```{r}
+nfl_actualStats_career_player %>% 
+  group_by(player_id) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_career_player_inclPost %>% 
+  group_by(player_id) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_career_player, nfl_actualStats_career_player_inclPost,
+  file = "./data/nfl_actualStats_career.RData"
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_actualStats_career.RData")
+```
 
@@
 
-#### Season-by-Season Statistics {#sec-calculateActualSeasonalStats}
-
-Second, we can compute the players' season-by-season statistics.
-
-TODO: Save/update data file in repo with the data generated from this code, and follow the data file through the rest of the book, updating code as necessary.
-
-::: {#nte-downloadSeasonalActualStats .callout-note title="Downloading season-by-season statistics"}
-Note: the following code takes a while to run.
-:::
-
-```{r}
-#| eval: false
-nfl_actualStats_seasonal_player_raw <- calculate_stats(
-  seasons = TRUE,
-  summary_level = "season",
-  season_type = "REG")
-
-nfl_actualStats_seasonal_team_raw <- calculate_stats(
-  seasons = TRUE,
-  summary_level = "season",
-  stat_type = "team",
-  season_type = "REG")
-
-nfl_actualStats_seasonal_player_inclPost_raw <- calculate_stats(
-  seasons = TRUE,
-  summary_level = "season",
-  season_type = "REG+POST")
-
-nfl_actualStats_seasonal_team_inclPost_raw <- calculate_stats(
-  seasons = TRUE,
-  summary_level = "season",
-  stat_type = "team",
-  season_type = "REG+POST")
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_seasonal_player_raw, nfl_actualStats_seasonal_team_raw,
-  nfl_actualStats_seasonal_player_inclPost_raw, nfl_actualStats_seasonal_team_inclPost_raw,
-  file = "./data/nfl_actualStats_seasonal_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_seasonal_raw.RData")
-```
-
-A Data Dictionary for the variables is available in the `nfl_stats_variables` object that is returned when running the `calculate_stats()` function:
-
-```{r}
-nfl_stats_variables
-```
-
-The `nfl_actualStats_seasonal_player_raw` object is in `player`-`season` form.
-That is, each row should be uniquely identified by the combination of `player_id` and `season`.
-The `nfl_actualStats_seasonal_team_raw` object is in `team`-`season` form.
-That is, each row should be uniquely identified by the combination of `team` and `season`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_actualStats_seasonal_player <- nfl_actualStats_seasonal_player_raw %>% 
-  rename(team = recent_team) %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-
-nfl_actualStats_seasonal_team <- nfl_actualStats_seasonal_team_raw %>% 
-  select(season, team, everything()) %>% 
-  arrange(season, team)
-
-nfl_actualStats_seasonal_player_inclPost <- nfl_actualStats_seasonal_player_inclPost_raw %>% 
-  rename(team = recent_team) %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-
-nfl_actualStats_seasonal_team_inclPost <- nfl_actualStats_seasonal_team_inclPost_raw %>% 
-  select(season, team, everything()) %>% 
-  arrange(season, team)
-```
-
-Let's check for duplicate `player` instances:
-
-```{r}
-nfl_actualStats_seasonal_player %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_seasonal_team %>% 
-  group_by(team, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_seasonal_player_inclPost %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_seasonal_team_inclPost %>% 
-  group_by(team, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-save(
-  nfl_actualStats_seasonal_player, nfl_actualStats_seasonal_team,
-  nfl_actualStats_seasonal_player_inclPost, nfl_actualStats_seasonal_team_inclPost,
-  file = "./data/nfl_actualStats_seasonal.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_seasonal.RData")
-```
-
-##### OLD (DELETE CODE BELOW)
-
-```{r}
-seasons <- unique(nfl_pbp$season)
-
-nfl_pbp_seasonalList <- list()
-nfl_actualStats_offense_seasonalList <- list()
-nfl_actualStats_defense_seasonalList <- list()
-nfl_actualStats_kicking_seasonalList <- list()
-```
-
-::: {#nte-calculateActualSeasonalStats .callout-note title="Calculating players' season-by-season statistics"}
-Note: the following code takes a while to run.
-:::
-
-```{r}
-#| eval: false
-
-pb <- txtProgressBar(
-  min = 0,
-  max = length(seasons),
-  style = 3)
-
-for(i in 1:length(seasons)){
-  # Subset play-by-play data by season
-  nfl_pbp_seasonalList[[i]] <- nfl_pbp %>% 
-    dplyr::filter(season == seasons[i])
-
-  # Compute actual statistics by season
-  nfl_actualStats_offense_seasonalList[[i]] <- 
-    nflfastR::calculate_player_stats(
-      nfl_pbp_seasonalList[[i]],
-      weekly = FALSE)
-
-  nfl_actualStats_defense_seasonalList[[i]] <- 
-    nflfastR::calculate_player_stats_def(
-      nfl_pbp_seasonalList[[i]],
-      weekly = FALSE)
-
-  nfl_actualStats_kicking_seasonalList[[i]] <- 
-    nflfastR::calculate_player_stats_kicking(
-      nfl_pbp_seasonalList[[i]],
-      weekly = FALSE)
-
-  nfl_actualStats_offense_seasonalList[[i]]$season <- seasons[i]
-  nfl_actualStats_defense_seasonalList[[i]]$season <- seasons[i]
-  nfl_actualStats_kicking_seasonalList[[i]]$season <- seasons[i]
-
-  print(
-    paste("Completed computing projections for season: ", seasons[i], sep = ""))
-
-  # Update the progress bar
-  setTxtProgressBar(pb, i)
-}
-
-# Close the progress bar
-close(pb)
-
-nfl_actualStats_offense_seasonal_raw <- nfl_actualStats_offense_seasonalList %>% 
-  dplyr::bind_rows()
-nfl_actualStats_defense_seasonal_raw <- nfl_actualStats_defense_seasonalList %>% 
-  dplyr::bind_rows()
-nfl_actualStats_kicking_seasonal_raw <- nfl_actualStats_kicking_seasonalList %>% 
-  dplyr::bind_rows()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_offense_seasonal_raw, nfl_actualStats_defense_seasonal_raw, nfl_actualStats_kicking_seasonal_raw,
-  file = "./data/nfl_actualStats_seasonal_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_seasonal_raw.RData")
-```
-
-The `nfl_actualStats_seasonal` objects are in `player`-`season` form.
-That is, each row should be uniquely identified by the combination of `player_id` and `season`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal_raw %>% 
-  rename(team = recent_team) %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-
-nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal_raw %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal_raw %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-```
-
-Let's check for duplicate `player` instances:
-
-```{r}
-nfl_actualStats_offense_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_defense_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_kicking_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's do some data cleanup:
-
-```{r}
-# Sum statistics across a player's years with multiple teams
-nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal %>% 
-  group_by(player_id, season) %>% 
-  summarise(
-    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
-    .groups = "drop"
-  )
-
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
-  group_by(player_id, season) %>% 
-  summarise(
-    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
-    .groups = "drop"
-  )
-
-# Re-calculate percentage stats
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
-  mutate(
-    fg_pct = fg_made / fg_att,
-    pat_pct = pat_made / pat_att
-  )
-
-# Merge data back with player info
-nfl_actualStats_defense_seasonal_raw_playerInfo <- nfl_actualStats_defense_seasonal_raw %>% 
-  select(player_id, season, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
-  unique() %>% 
-  group_by(player_id) %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-
-nfl_actualStats_kicking_seasonal_raw_playerInfo <- nfl_actualStats_kicking_seasonal_raw %>% 
-  select(player_id, season, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
-  unique() %>% 
-  group_by(player_id) %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-
-nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal_raw_playerInfo %>% 
-  right_join(
-    nfl_actualStats_defense_seasonal,
-    by = c("player_id","season")
-  )
-
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal_raw_playerInfo %>% 
-  right_join(
-    nfl_actualStats_kicking_seasonal,
-    by = c("player_id","season")
-  )
-
-# Rearrange data
-nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-```
-
-Let's check again for duplicate `player` instances:
-
-```{r}
-nfl_actualStats_offense_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_defense_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_kicking_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-save(
-  nfl_actualStats_offense_seasonal, nfl_actualStats_defense_seasonal, nfl_actualStats_kicking_seasonal,
-  file = "./data/nfl_actualStats_seasonal.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_seasonal.RData")
-```
-
 #### Week-by-Week Statistics {#sec-calculateActualWeeklyStats}
@@
 #| eval: false
-nfl_actualStats_weekly_player_raw <- calculate_stats(
+nfl_actualStats_weekly_player_raw <- nflfastR::calculate_stats(
   seasons = TRUE,
@@
 
-nfl_actualStats_weekly_team_raw <- calculate_stats(
+nfl_actualStats_weekly_team_raw <- nflfastR::calculate_stats(
   seasons = TRUE,

### 2025-04-07 download-football-data.qmd
20250406 - fix calculation

@@
   "fg_pct", # fg_made / fg_att
-  "pat_pct" # pat_att / pat_made
+  "pat_pct" # pat_made / pat_att
 )
@@
     fg_pct = fg_made / fg_att,
-    pat_pct = pat_att / pat_made
+    pat_pct = pat_made / pat_att
   )

### 2025-04-07 sports-cognitive-psychology.qmd
20250407 - Lochbaum meta-analysis

@@
 
+Meta-analyses have demonstrated that there are a number of psychological factors that influence player performance in sports.
+For instance, a meta analysis by @Lochbaum2022 found that psychological processes such as confidence and mindfulness were associated with better performance in sports.
+
+One process that may have a nonlinear association with performance is arousal or anxiety.
 The Yerkes–Dodson law describes the association between arousal and performance, as depicted in @fig-yerkesDodsonLaw.

### 2025-04-07 download-football-data.qmd
20250407 - updates

@@
 
+Now, let's identify the unique seasons of play-by-play data for use in later calculations:
+
+```{r}
+seasons <- unique(nfl_pbp$season)
+```
+
 ### 4th Down Data {#sec-download4thDown}
@@
   nflreadr::load_participation(
-    seasons = TRUE,
+    seasons = 2016:2023, # participation data are no longer available after 2023
     include_pbp = TRUE))
@@
 
+We set the `seasons` variable in @sec-downloadPlayByPlay.
+
 ```{r}

### 2025-04-07 correlation.qmd
20250407 - correlation

@@
 library("XICOR")
+library("psych")
+library("DescTools")
 library("tidyverse")
 ```
+
+### Load Data {#sec-correlationLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+```
+
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
 
@@
 
+Is there an association between a player's height and the number of fantasy points they score?
+We can examine this possibility separately by position.
+
+First, let's examine descriptive statistics (height is measured in inches):
+
+```{r}
+player_stats_seasonal %>% 
+  dplyr::select(height, fantasyPoints) %>% 
+  dplyr::summarise(across(
+      everything(),
+      .fns = list(
+        n = ~ length(na.omit(.)),
+        missingness = ~ mean(is.na(.)) * 100,
+        M = ~ mean(., na.rm = TRUE),
+        SD = ~ sd(., na.rm = TRUE),
+        min = ~ min(., na.rm = TRUE),
+        max = ~ max(., na.rm = TRUE),
+        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
+        IQR = ~ IQR(., na.rm = TRUE),
+        MAD = ~ mad(., na.rm = TRUE),
+        median = ~ median(., na.rm = TRUE),
+        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
+        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
+        skewness = ~ psych::skew(., na.rm = TRUE),
+        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
+      .names = "{.col}.{.fn}")) %>%
+    tidyr::pivot_longer(
+      cols = everything(),
+      names_to = c("variable","index"),
+      names_sep = "\\.") %>% 
+    tidyr::pivot_wider(
+      names_from = index,
+      values_from = value)
+```
+
 ### Covariance {#sec-correlationExamplesCovariance}
 
+The following syntax shows the variance-covariance of a set of variables, where the covariance of a variable with itself is the variable's variance.
+
+```{r}
+cov(
+  player_stats_seasonal[,c("height","weight","fantasyPoints")],
+  use = "pairwise.complete.obs")
+```
+
+If you just want the covariance between two variables, you can use the following syntax.
+
+```{r}
+cov(
+  player_stats_seasonal$height,
+  player_stats_seasonal$fantasyPoints,
+  use = "pairwise.complete.obs")
+```
+
+Thus, it appears there is a negative covariance between height and fantasy points in the whole population of National Football League (NFL) players.
+
+As shown below, the negative covariance between height and fantasy points holds when examining just Quarterbacks, Running Backs, Wide Receivers, and Tight Ends.
+
+```{r}
+cov(
+  player_stats_seasonal %>% filter(position %in% c("QB","RB","WR","TE")) %>% select(height,fantasyPoints),
+  use = "pairwise.complete.obs")
+```
+
+We can also examine the association separately by position.
+Here is the association for Quarterbacks:
+
+```{r}
+cov(
+  player_stats_seasonal %>% filter(position == "QB") %>% select(height,fantasyPoints),
+  use = "pairwise.complete.obs")
+```
+
+Here is the association for Running Backs:
+
+```{r}
+cov(
+  player_stats_seasonal %>% filter(position == "RB") %>% select(height,fantasyPoints),
+  use = "pairwise.complete.obs")
+```
+
+Here is the association for Wide Receivers:
+
+```{r}
+cov(
+  player_stats_seasonal %>% filter(position == "WR") %>% select(height,fantasyPoints),
+  use = "pairwise.complete.obs")
+```
+
+Here is the association for Tight Ends:
+
+```{r}
+cov(
+  player_stats_seasonal %>% filter(position == "TE") %>% select(height,fantasyPoints),
+  use = "pairwise.complete.obs")
+```
+
+Interestingly, there is a *positive* covariance between height and fantasy points when examining each position separately.
+This is an example of [Simpson's paradox](#sec-simpsonsParadox), where the sign of an association differs at different levels of analysis, as described in @sec-simpsonsParadox.
+There is a *negative* covariance between height and fantasy points when examining Quarterbacks, Running Backs, Wide Receivers, and Tight Ends altogether, but there is a *positive* covariance between height and fantasy points when examining the association between each position separately.
+That is, if you were to examine all positions together, you might assume that being taller might be disadvantageous to scoring fantasy fantasy points; however, once we examine the association within a given position, it emerges that height appears to be somewhat advantageous, if anything, for scoring fantasy points.
+
+The covariance is an unstandardized index of the association between two variables.
+However, it can be helpful to use a standardized index to examine the [effect size](#sec-practicalSignificance)—i.e., the strength of the association.
+The [Pearson correlation coefficient](Pearson Correlation), $r$, is an example of a standardized index of the covariance between two variables.
+
 ### Pearson Correlation {#sec-correlationExamplesPearson}
+
+Pearson correlation assumes that the variables are [interval](#sec-interval) or [ratio](#sec-interval) level of measurement.
+If the data are [ordinal](#sec-ordinal), [Spearman correlation](#sec-correlationExamplesSpearman) is more appropriate.
+
+The following syntax shows the Pearson correlation matrix of a set of variables.
+
+```{r}
+cor(
+  player_stats_seasonal[,c("height","weight","fantasyPoints")],
+  use = "pairwise.complete.obs")
+```
+
+If you just want the Pearson correlation between two variables, you can use the following syntax.
+
+```{r}
+cor.test(
+  ~ height + fantasyPoints,
+  data = player_stats_seasonal
+)
+```
+
+```{r}
+#| include: false
+
+pearson_test <- cor.test(
+  ~ height + fantasyPoints,
+  data = player_stats_seasonal
+)
+
+pearson_r <- pearson_test$estimate
+pearson_df <- pearson_test$parameter
+pearson_p <- pearson_test$p.value
+```
+
+The $r$ value (`{r} pearson_r`) is slightly negative.
+The [effect size](#sec-practicalSignificance) of the association is small—for what is considered a small, medium, and large effect size for $r$ values, see @sec-effectSizeThresholds.
+The $p$-value is less than .05, so it is a statistically significant correlation.
+We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[$`{r} pearson_df`$] =$ `{r} pearson_r`, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.
+
+Now let's examine the association separately by position.
+Here is the association for Quarterbacks:
+
+```{r}
+cor.test(
+  ~ height + fantasyPoints,
+  data = player_stats_seasonal %>% 
+    filter(position == "QB")
+)
+```
+
+Here is the association for Running Backs:
+
+```{r}
+cor.test(
+  ~ height + fantasyPoints,
+  data = player_stats_seasonal %>% 
+    filter(position == "RB")
+)
+```
+
+Here is the association for Wide Receivers:
+
+```{r}
+cor.test(
+  ~ height + fantasyPoints,
+  data = player_stats_seasonal %>% 
+    filter(position == "WR")
+)
+```
+
+Here is the association for Tight Ends:
+
+```{r}
+cor.test(
+  ~ height + fantasyPoints,
+  data = player_stats_seasonal %>% 
+    filter(position == "TE")
+)
+```
+
+The sign of the association was positive for each position, suggesting that, for a given position (among Quarterbacks, Running Backs, Wide Receivers, and Tight Ends), taller players tend to score more fantasy points.
+However, the [effect size](#sec-practicalSignificance) is small.
+Moreover, as described in @sec-correlation-correlationAndCausation, just because there is an association between variables does not mean that the association reflects a causal effect.
+
+Please note that the $p$-value for a correlation assumes that the observations are independent—in particular, that the residuals are not correlated.
+However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
+This non-independence violates the traditional assumptions of the significance test of a correlation.
+We could address this assumption by analyzing only one season from each player.
+However, for simplicity, we present results from the whole dataframe.
+In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression).
 

### 2025-04-07 multiple-regression.qmd
20250407 - correlation

@@
 
-Multiple regression examines the association between multiple [predictor variables](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy).
+Multiple regression is an extension of [correlation](#sec-correlation).
+[Correlation](#sec-correlation) examines the association between one [predictor variables](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy).
+Multiple regression examines the association between *multiple* [predictor variables](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy).
 It allows obtaining a more accurate estimate of the unique contribution of a given [predictor variable](#sec-correlationalStudy), by controlling for other variables ([covariates](#sec-covariates)).

### 2025-04-08 correlation.qmd
20250407 - correlation

@@
 
+Notice the diagonal values are all 1.0.
+That is because a variable is perfectly correlated with itself.
+Also notice that the values above the diagonal are a mirror image of (i.e., they are the same as) the respective values below the diagonal.
+The association between two variables is the same regardless of the order (i.e., which is the [predictor variable](#sec-correlationalStudy) and which is the [outcome variable](#sec-correlationalStudy)); that is, the association between variables A and B is the same as the association between variables B and A.
+
 If you just want the Pearson correlation between two variables, you can use the following syntax.
@@
 
-The $r$ value (`{r} pearson_r`) is slightly negative.
+The $r$ value (`{r} petersenlab::apa(pearson_r, digits = 2)`) is slightly negative.
 The [effect size](#sec-practicalSignificance) of the association is small—for what is considered a small, medium, and large effect size for $r$ values, see @sec-effectSizeThresholds.
 The $p$-value is less than .05, so it is a statistically significant correlation.
-We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[$`{r} pearson_df`$] =$ `{r} pearson_r`, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.
+We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[$ `{r} pearson_df` $]=$ `{r} petersenlab::apa(pearson_r, digits = 2)`, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.
 
@@
 However, for simplicity, we present results from the whole dataframe.
-In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression).
+In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @#sec-assumptionsRegression.
 

### 2025-04-08 download-football-data.qmd
20250407 - clarification of key variable for sorting vs merging

@@
 That is, each row should be uniquely identified by `gsis_id`.
-Let's rearrange the data accordingly:
+Let's rearrange the data accordingly by player.
 
@@
 ```
+
+In this case, I arrange the data by player using `display_name`.
+You could arrange by `gsis_id` (or any other variable, for that matter), but `gsis_id` is fairly unintelligible by itself.
+For instance, here are the first 6 players by ID according to `gsis_id`:
+
+```{r}
+head(nfl_players_raw$gsis_id)
+```
+
+Note how it is unclear who these players are until you combine this column with other, more relevant information.
+Thus, I prefer to sort by a variable that is more interpretable, such as a player's name.
+Here are the first six players by name according to `display_name`:
+
+```{r}
+head(nfl_players_raw$display_name)
+```
+
+Note how it is more clear, at a glance and without additional, who these players are.
+
+In data analysis, the variable(s) that are used to sort the dataframe are primarily chosen for aesthetic or usability purposes.
+Many data analysis approaches do not depend on the order of the rows in the data.
+However, the sorting of the dataframe may influence data wrangling operations.
+Thus, when performing data wrangling operations it may be prudent to make sure to re-sort the data to the desired format before performing such operations.
+
+Regardless of which variable(s) are used to sort the dataframe, it is important to know which variable(s) uniquely identify the rows because that will influence approaches to [merge](#sec-merging) the data with other dataframes.
+For instance, it would be preferable to merge two data objects based on the player ID (`gsis_id`) rather than the player name (`display_name`), because multiple players share the same name, which can lead the [merge](#sec-merging) operation not to know which player from the first dataframe goes with which player from the second dataframe when both players have the same name (despite different having different player IDs).
 

### 2025-04-08 download-football-data.qmd
20250407 - delete old code

@@
 
-##### OLD (DELETE CODE BELOW)
-
-```{r}
-seasons <- unique(nfl_pbp$season)
-
-nfl_pbp_seasonalList <- list()
-nfl_actualStats_offense_seasonalList <- list()
-nfl_actualStats_defense_seasonalList <- list()
-nfl_actualStats_kicking_seasonalList <- list()
-```
-
-::: {#nte-calculateActualSeasonalStats .callout-note title="Calculating players' season-by-season statistics"}
-Note: the following code takes a while to run.
-:::
-
-```{r}
-#| eval: false
-
-pb <- txtProgressBar(
-  min = 0,
-  max = length(seasons),
-  style = 3)
-
-for(i in 1:length(seasons)){
-  # Subset play-by-play data by season
-  nfl_pbp_seasonalList[[i]] <- nfl_pbp %>% 
-    dplyr::filter(season == seasons[i])
-
-  # Compute actual statistics by season
-  nfl_actualStats_offense_seasonalList[[i]] <- 
-    nflfastR::calculate_player_stats(
-      nfl_pbp_seasonalList[[i]],
-      weekly = FALSE)
-
-  nfl_actualStats_defense_seasonalList[[i]] <- 
-    nflfastR::calculate_player_stats_def(
-      nfl_pbp_seasonalList[[i]],
-      weekly = FALSE)
-
-  nfl_actualStats_kicking_seasonalList[[i]] <- 
-    nflfastR::calculate_player_stats_kicking(
-      nfl_pbp_seasonalList[[i]],
-      weekly = FALSE)
-
-  nfl_actualStats_offense_seasonalList[[i]]$season <- seasons[i]
-  nfl_actualStats_defense_seasonalList[[i]]$season <- seasons[i]
-  nfl_actualStats_kicking_seasonalList[[i]]$season <- seasons[i]
-
-  print(
-    paste("Completed computing projections for season: ", seasons[i], sep = ""))
-
-  # Update the progress bar
-  setTxtProgressBar(pb, i)
-}
-
-# Close the progress bar
-close(pb)
-
-nfl_actualStats_offense_seasonal_raw <- nfl_actualStats_offense_seasonalList %>% 
-  dplyr::bind_rows()
-nfl_actualStats_defense_seasonal_raw <- nfl_actualStats_defense_seasonalList %>% 
-  dplyr::bind_rows()
-nfl_actualStats_kicking_seasonal_raw <- nfl_actualStats_kicking_seasonalList %>% 
-  dplyr::bind_rows()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_offense_seasonal_raw, nfl_actualStats_defense_seasonal_raw, nfl_actualStats_kicking_seasonal_raw,
-  file = "./data/nfl_actualStats_seasonal_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_seasonal_raw.RData")
-```
-
-The `nfl_actualStats_seasonal` objects are in `player`-`season` form.
-That is, each row should be uniquely identified by the combination of `player_id` and `season`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal_raw %>% 
-  rename(team = recent_team) %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-
-nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal_raw %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal_raw %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-```
-
-Let's check for duplicate `player` instances:
-
-```{r}
-nfl_actualStats_offense_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_defense_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_kicking_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's do some data cleanup:
-
-```{r}
-# Sum statistics across a player's years with multiple teams
-nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal %>% 
-  group_by(player_id, season) %>% 
-  summarise(
-    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
-    .groups = "drop"
-  )
-
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
-  group_by(player_id, season) %>% 
-  summarise(
-    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
-    .groups = "drop"
-  )
-
-# Re-calculate percentage stats
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
-  mutate(
-    fg_pct = fg_made / fg_att,
-    pat_pct = pat_made / pat_att
-  )
-
-# Merge data back with player info
-nfl_actualStats_defense_seasonal_raw_playerInfo <- nfl_actualStats_defense_seasonal_raw %>% 
-  select(player_id, season, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
-  unique() %>% 
-  group_by(player_id) %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-
-nfl_actualStats_kicking_seasonal_raw_playerInfo <- nfl_actualStats_kicking_seasonal_raw %>% 
-  select(player_id, season, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
-  unique() %>% 
-  group_by(player_id) %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-
-nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal_raw_playerInfo %>% 
-  right_join(
-    nfl_actualStats_defense_seasonal,
-    by = c("player_id","season")
-  )
-
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal_raw_playerInfo %>% 
-  right_join(
-    nfl_actualStats_kicking_seasonal,
-    by = c("player_id","season")
-  )
-
-# Rearrange data
-nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonal %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-
-nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonal %>% 
-  select(player_id, season, everything()) %>% 
-  arrange(player_display_name, player_id, season)
-```
-
-Let's check again for duplicate `player` instances:
-
-```{r}
-nfl_actualStats_offense_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_defense_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_kicking_seasonal %>% 
-  group_by(player_id, season) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-save(
-  nfl_actualStats_offense_seasonal, nfl_actualStats_defense_seasonal, nfl_actualStats_kicking_seasonal,
-  file = "./data/nfl_actualStats_seasonal.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_seasonal.RData")
-```
-
 #### Career Statistics {#sec-calculateActualCareerStats}
@@
 
-##### OLD (DELETE CODE BELOW)
-
-First, we can compute the players' career statistics using the `calculate_player_stats()`, `calculate_player_stats_def()`, and `calculate_player_stats_kicking()` functions from the `nflfastR` package [@R-nflfastR] for offensive players, defensive players, and kickers, respectively.
-
-::: {#nte-calculateActualCareerStats .callout-note title="Calculating players' career statistics"}
-Note: the following code takes a while to run.
-:::
-
-```{r}
-#| eval: false
-
-nfl_actualStats_offense_career_raw <- nflfastR::calculate_player_stats(
-  nfl_pbp,
-  weekly = FALSE)
-
-nfl_actualStats_defense_career_raw <- nflfastR::calculate_player_stats_def(
-  nfl_pbp,
-  weekly = FALSE)
-
-nfl_actualStats_kicking_career_raw <- nflfastR::calculate_player_stats_kicking(
-  nfl_pbp,
-  weekly = FALSE)
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_offense_career_raw, nfl_actualStats_defense_career_raw, nfl_actualStats_kicking_career_raw,
-  file = "./data/nfl_actualStats_career_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_career_raw.RData")
-```
-
-The `nfl_actualStats_career` objects are in `player` form.
-That is, each row should be uniquely identified by the combination of `player_id`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_actualStats_offense_career <- nfl_actualStats_offense_career_raw %>% 
-  arrange(player_display_name, player_id)
-
-nfl_actualStats_defense_career <- nfl_actualStats_defense_career_raw %>% 
-  arrange(player_display_name, player_id)
-
-nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career_raw %>% 
-  arrange(player_display_name, player_id)
-```
-
-Let's check for duplicate `player` instances:
-
-```{r}
-nfl_actualStats_offense_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_defense_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_kicking_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Let's do some data cleanup:
-
-```{r}
-# Sum statistics across a player's years with multiple teams
-nfl_actualStats_defense_career <- nfl_actualStats_defense_career %>% 
-  group_by(player_id) %>% 
-  summarise(
-    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
-    .groups = "drop"
-  )
-
-nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career %>% 
-  group_by(player_id) %>% 
-  summarise(
-    across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
-    .groups = "drop"
-  )
-
-# Re-calculate percentage stats
-nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career %>% 
-  mutate(
-    fg_pct = fg_made / fg_att,
-    pat_pct = pat_made / pat_att
-  )
-
-# Merge data back with player info
-nfl_actualStats_defense_career_raw_playerInfo <- nfl_actualStats_defense_career_raw %>% 
-  select(player_id, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
-  unique() %>% 
-  group_by(player_id) %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-
-nfl_actualStats_kicking_career_raw_playerInfo <- nfl_actualStats_kicking_career_raw %>% 
-  select(player_id, player_name, player_display_name, position, position_group, team, headshot_url) %>% 
-  unique() %>% 
-  group_by(player_id) %>% 
-  slice_head(n = 1) %>% 
-  ungroup()
-
-nfl_actualStats_defense_career <- nfl_actualStats_defense_career_raw_playerInfo %>% 
-  right_join(
-    nfl_actualStats_defense_career,
-    by = "player_id"
-  )
-
-nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career_raw_playerInfo %>% 
-  right_join(
-    nfl_actualStats_kicking_career,
-    by = "player_id"
-  )
-
-# Rearrange data
-nfl_actualStats_defense_career <- nfl_actualStats_defense_career %>% 
-  arrange(player_display_name, player_id)
-
-nfl_actualStats_kicking_career <- nfl_actualStats_kicking_career %>% 
-  arrange(player_display_name, player_id)
-```
-
-Let's check again for duplicate `player` instances:
-
-```{r}
-nfl_actualStats_offense_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_defense_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_kicking_career %>% 
-  group_by(player_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-save(
-  nfl_actualStats_offense_career, nfl_actualStats_defense_career, nfl_actualStats_kicking_career,
-  file = "./data/nfl_actualStats_career.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_career.RData")
-```
-
 #### Week-by-Week Statistics {#sec-calculateActualWeeklyStats}
@@
 load(file = "./data/nfl_actualStats_weekly.RData")
-```
-
-##### OLD (DELETE CODE BELOW)
-
-```{r}
-#| eval: false
-
-nfl_actualStats_offense_weekly <- nflfastR::calculate_player_stats(
-  nfl_pbp,
-  weekly = TRUE)
-
-nfl_actualStats_defense_weekly <- nflfastR::calculate_player_stats_def(
-  nfl_pbp,
-  weekly = TRUE)
-
-nfl_actualStats_kicking_weekly <- nflfastR::calculate_player_stats_kicking(
-  nfl_pbp,
-  weekly = TRUE)
-```
-
-```{r}
-#| include: false
-
-# garbage collection
-rm(nfl_pbp)
-gc()
 ```
@@
 
-##### OLD (DELETE CODE BELOW)
-
-```{r}
-#| eval: false
-
-nfl_actualStats_offense_weekly
-nfl_actualStats_defense_weekly
-nfl_actualStats_kicking_weekly
-
-test_weekly_offense <- ffanalytics:::actual_points_scoring(
-  nflr_player_stats_df = nfl_actualStats_offense_weekly %>% filter(season == 2023),
-  nflr_pbp_data = nfl_pbp %>% filter(season == 2023),
-  scoring_rules = scoring_obj,
-  vor_baseline = NULL,
-  rename_colums = TRUE)
-
-test_weekly_defense <- ffanalytics:::actual_points_scoring(
-  nflr_player_stats_df = nfl_actualStats_defense_weekly %>% filter(season == 2023),
-  nflr_pbp_data = nfl_pbp %>% filter(season == 2023),
-  scoring_rules = scoring_obj,
-  vor_baseline = NULL,
-  rename_colums = TRUE)
-
-test_weekly_kicking <- ffanalytics:::actual_points_scoring(
-  nflr_player_stats_df = nfl_actualStats_kicking_weekly %>% filter(season == 2023),
-  nflr_pbp_data = nfl_pbp %>% filter(season == 2023),
-  scoring_rules = scoring_obj,
-  vor_baseline = NULL,
-  rename_colums = TRUE)
-```
-
 #### Season-by-Season {#sec-calculateActualPointsSeasonal}

### 2025-04-08 correlation.qmd
20250407 - fix error

@@
 
-The $r$ value (`{r} petersenlab::apa(pearson_r, digits = 2)`) is slightly negative.
+The $r$ value (`{r} petersenlab::apa(pearson_r, decimals = 2)`) is slightly negative.
 The [effect size](#sec-practicalSignificance) of the association is small—for what is considered a small, medium, and large effect size for $r$ values, see @sec-effectSizeThresholds.
 The $p$-value is less than .05, so it is a statistically significant correlation.
-We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[$ `{r} pearson_df` $]=$ `{r} petersenlab::apa(pearson_r, digits = 2)`, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.
+We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[$ `{r} pearson_df` $]=$ `{r} petersenlab::apa(pearson_r, decimals = 2)`, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.
 

### 2025-04-08 download-football-data.qmd
20250407 - edits

@@
 You could arrange by `gsis_id` (or any other variable, for that matter), but `gsis_id` is fairly unintelligible by itself.
-For instance, here are the first 6 players by ID according to `gsis_id`:
+For instance, here are the first six players by ID according to `gsis_id`:
 
@@
 
-Note how it is more clear, at a glance and without additional, who these players are.
+Note how it is more clear, at a glance and without additional information, who these players are.
 

### 2025-04-08 correlation.qmd
20250407 - fix formula

@@
 The $p$-value is less than .05, so it is a statistically significant correlation.
-We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[$ `{r} pearson_df` $]=$ `{r} petersenlab::apa(pearson_r, decimals = 2)`, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.
+We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[`{r} pearson_df`] = `{r} petersenlab::apa(pearson_r, decimals = 2)`$, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.
 
@@
 However, for simplicity, we present results from the whole dataframe.
-In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @#sec-assumptionsRegression.
+In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @sec-assumptionsRegression.
 

### 2025-04-08 correlation.qmd
20250408 - noquote

@@
 pearson_p <- pearson_test$p.value
+
+#pValue <- function(value, digits = 3){
+#    if(value < .001){
+#      newValue <- " < .001"
+#    } else {
+#      newValue <- paste(
+#        " = ",
+#        petersenlab::suppressLeadingZero(specify_decimal(value, digits)),
+#        sep = "")
+#    }
+#  
+#    return(newValue)
+#  }
+#
+#textCorrelation <- function(x, y, decimals = 2){
+#  substitute(
+#  paste(
+#    italic(r), "(", valueDF, ")",  " = ", valueR,
+#    ", ", italic(p), valueP,
+#    sep = ""),
+#  list(
+#    valueR = noquote(sub("0.", ".", sprintf("%.2f", round(cor.test(x, y)$estimate[[1]], decimals)))),
+#    valueDF = cor.test(x, y)$parameter[[1]],
+#    valueP = pValue(cor.test(x, y)$p.value[[1]])
+#    )
+#  )
+#}
+#
+#textCorrelation <- function(x, y, decimals = 2){
+#  substitute(
+#  paste(
+#    r, "(", valueDF, ")",  " = ", valueR,
+#    ", ", p, valueP,
+#    sep = ""),
+#  list(
+#    valueR = noquote(sub("0.", ".", sprintf("%.2f", round(cor.test(x, y)$estimate[[1]], decimals)))),
+#    valueDF = cor.test(x, y)$parameter[[1]],
+#    valueP = pValue(cor.test(x, y)$p.value[[1]])
+#    )
+#  )
+#}
+#
+#textCorrelation(player_stats_seasonal$height, player_stats_seasonal$fantasyPoints)
 ```
 
-The $r$ value (`{r} petersenlab::apa(pearson_r, decimals = 2)`) is slightly negative.
+The $r$ value (`{r} petersenlab::apa(pearson_r, decimals = 2, leading = FALSE)`) is slightly negative.
 The [effect size](#sec-practicalSignificance) of the association is small—for what is considered a small, medium, and large effect size for $r$ values, see @sec-effectSizeThresholds.
 The $p$-value is less than .05, so it is a statistically significant correlation.
-We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[`{r} pearson_df`] = `{r} petersenlab::apa(pearson_r, decimals = 2)`$, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.
+We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r(`{r} pearson_df`) = `{r} noquote(petersenlab::apa(pearson_r, decimals = 2, leading = FALSE))`$, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.
 

### 2025-04-08 correlation.qmd
20250408 - additions

--- 2025-04-08 correlation.qmd
@@
 
-The $r$ value (`{r} petersenlab::apa(pearson_r, decimals = 2)`) is slightly negative.
+The $r$ value (`{r} petersenlab::apa(pearson_r, decimals = 2, leading = FALSE)`) is slightly negative.
 The [effect size](#sec-practicalSignificance) of the association is small—for what is considered a small, medium, and large effect size for $r$ values, see @sec-effectSizeThresholds.
 The $p$-value is less than .05, so it is a statistically significant correlation.
-We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[`{r} pearson_df`] = `{r} petersenlab::apa(pearson_r, decimals = 2)`$, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.
+We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r(`r pearson_df`) = `r petersenlab::apa(pearson_r, decimals = 2, leading = FALSE)`$, $p `r petersenlab::pValue(pearson_p)`$); the taller a player was, the fewer fantasy points they tended to score.
 
@@
 
-Please note that the $p$-value for a correlation assumes that the observations are independent—in particular, that the residuals are not correlated.
-However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
-This non-independence violates the traditional assumptions of the significance test of a correlation.
-We could address this assumption by analyzing only one season from each player.
-However, for simplicity, we present results from the whole dataframe.
-In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @sec-assumptionsRegression.
+### Rank Correlation: Spearman's Rho ($\rho$) {#sec-correlationExamplesSpearman}
 
-### Spearman Correlation {#sec-correlationExamplesSpearman}
+Spearman's rho ($\rho$) is a rank correlation.
+Spearman's rho does not assume that the data are [interval](#sec-interval) or [ratio](#sec-ratio) level of measurement.
+Unlike [Pearson correlation](sec-correlationExamplesPearson), Spearman's rho allows estimating associations among variables that are [ordinal](#sec-ordinal) level of measurement.
+
+For instance, using Spearman's rho, we could examine the association between a player's draft round and their fantasy points, because draft round is an [ordinal](#sec-ordinal) variable:
+
+```{r}
+cor(
+  player_stats_seasonal[,c("draftround","fantasyPoints")],
+  use = "pairwise.complete.obs",
+  method = "spearman")
+
+cor.test(
+  ~ draftround + fantasyPoints,
+  data = player_stats_seasonal,
+  method = "spearman"
+)
+```
+
+```{r}
+#| include: false
+
+spearman_test <- cor.test(
+  ~ draftround + fantasyPoints,
+  data = player_stats_seasonal,
+  method = "spearman"
+)
+
+spearman_r <- spearman_test$estimate
+spearman_df <- cor.test(
+  ~ draftround + fantasyPoints,
+  data = player_stats_seasonal)$parameter
+spearman_p <- spearman_test$p.value
+```
+
+Using Spearman's rho, there is a negative association between draft round and fantasy points ($r(`r spearman_df`) = `r petersenlab::apa(spearman_r, decimals = 2, leading = FALSE)`$, $p `r petersenlab::pValue(spearman_p)`$); unsurprisingly, the earlier the round a player is drafted, the more fantasy points they tend to score.
+
+### Partial Correlation {#sec-correlationExamplesNonlinear}
 
@@
 
+```{r}
+set.seed(52242) # for reproducibility
+
+XICOR::xicor(
+  player_stats_seasonal$height,
+  player_stats_seasonal$fantasyPoints,
+  pvalue = TRUE
+)
+
+set.seed(52242) # for reproducibility
+
+XICOR::calculateXI(
+  player_stats_seasonal$height,
+  player_stats_seasonal$fantasyPoints)
+```
+
 ### Correlation Matrix {#sec-correlationExamplesMatrix}
+
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `cor.table()` function that generates a correlation matrix of variables, including the $r$ value, number of observations ($n$), and $p$-value for each association.
+Here is a Pearson correlation matrix:
+
+```{r}
+petersenlab::cor.table(
+  player_stats_seasonal[,c("height","weight","fantasyPoints")]
+)
+```
+
+Here is a Spearman correlation matrix:
+
+```{r}
+petersenlab::cor.table(
+  player_stats_seasonal[,c("height","weight","fantasyPoints")],
+  correlation = "spearman"
+)
+```
+
+Here is a correlation matrix of variables that includes just the $r$ value and asterisks denoting the statistical significance of the association, for greater concision: $^\dagger p < .1$; $^* p < .05$; $^{**} p < .01$; $^{***} p < .001$.
+
+```{r}
+petersenlab::cor.table(
+  player_stats_seasonal[,c("height","weight","fantasyPoints")],
+  type = "manuscript" # include r values and asterisks
+)
+```
+
+Here is a correlation matrix of variables that includes just the $r$ value for even greater concision.
+
+```{r}
+petersenlab::cor.table(
+  player_stats_seasonal[,c("height","weight","fantasyPoints")],
+  type = "manuscriptBig" # include just r values
+)
+```
+
+Here is a partial correlation matrix, controlling for the player's age:
+
+```{r}
+petersenlab::partialcor.table(
+  player_stats_seasonal[,c("height","weight","fantasyPoints")],
+  z = player_stats_seasonal$age # control for the player's age
+)
+```
 
@@
 
+## Addressing Non-Independence of Observations {#sec-correlationNonIndependence}
+
+Please note that the $p$-value for a correlation assumes that the observations are independent—in particular, that the residuals are not correlated.
+However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
+This non-independence violates the traditional assumptions of the significance test of a correlation.
+We could address this assumption by analyzing only one season from each player or by estimating the significance of the correlation coefficient using cluster-robust standard errors.
+For simplicity, we present results above from the whole dataframe.
+In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @sec-assumptionsRegression.
+Below, we demonstrate how to account for non-independence of observations using cluster-robust standard errors.
+
+```{r}
+
+```
+
 ## Impact of Outliers {#sec-correlationOutliers}
+
+### Examining Robust Correlation {#sec-correlationRobust}
 

### 2025-04-08 correlation.qmd
20250408 - correlogram

@@
 library("DescTools")
+library("corrplot")
 library("tidyverse")
@@
 
+We can depict a correlogram using the `corrplot()` function of the `corrplot` package [@R-corrplot], as shown in @fig-correlogram.
+
+```{r}
+#| label: fig-correlogram
+#| fig-cap: "Correlogram."
+#| fig-alt: "Correlogram."
+
+corrplot(cor(
+  player_stats_seasonal[,c("height","weight","fantasyPoints")],
+  use = "pairwise.complete.obs"))
+```
+
 ## Addressing Non-Independence of Observations {#sec-correlationNonIndependence}

### 2025-04-08 index.qmd
20250316 - package update
20250405 - updates
20250408 - correlogram

@@
   "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin",
-  "caret", "gpboost", "cmdstanr", "learnr", "remotes" #"mgcv",
+  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot" #"mgcv",
 )
 
-knitr::write_bib(
+xfun::pkg_bib(
   c(.packages(), packagesForBibliography),
@@
 <a href="https://www.paypal.com/donate/?business=Q4FMMAMFQYTX8&no_recurring=0&item_name=Support+me+in+developing+this+resource+for+fantasy+football+analytics...+Even+a+cup+of+coffee+%28%244.99%29+helps+me+stay+awake%21&currency_code=USD">(or use PayPal)</a>
+
+<p>
+
+You can also sponsor my open source work using <a href="https://github.com/sponsors/isaactpetersen">GitHub Sponsors</a>.
 

### 2025-04-09 multiple-regression.qmd
20250407 - correlation
20250408 - addressing nonindependence of observations

@@
 library("petersenlab")
+library("rms")
 library("tidyverse")
@@
 ```
+
+### Load Data {#sec-correlationLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+```
+
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
 
@@
 
-## Impact of Oultiers {#sec-multipleRegressionOutliers}
+## Addressing Non-Independence of Observations {#sec-multipleRegressionNonIndependence}
+
+```{r}
+player_stats_seasonal_subset <- player_stats_seasonal %>% 
+  filter(!is.na(player_id))
+
+rms::robcov(rms::ols(
+  fantasyPoints ~ height,
+  data = player_stats_seasonal_subset,
+  x = TRUE,
+  y = TRUE),
+  cluster = player_stats_seasonal_subset$player_id) #account for nested data within player
+```
+
+## Impact of Outliers {#sec-multipleRegressionOutliers}
 

### 2025-04-09 correlation.qmd
20250408 - interactive scatterplot with best-fit line

@@
 
+### Scatterplot with Best-Fit Line
+
+A scatterplot with best-fit line
+
+```{r}
+#| label: fig-scatterplotWithBestFitLine
+#| fig-cap: "Scatterplot of Fantasy Points (Season) by Player Height With Best-Fit Line."
+#| fig-alt: "Scatterplot of Fantasy Points (Season) by Player Height With Best-Fit Line."
+
+plot_scatterplot <- ggplot2::ggplot(
+  data = player_stats_seasonal %>% filter(position %in% c("QB","RB","WR","TE")),
+  aes(
+    x = height,
+    y = fantasyPoints)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season), # add season for mouse over tooltip
+    alpha = 0.05) +
+  geom_smooth(
+    method = "lm",
+    color = "black") +
+  geom_smooth() +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Player Height (Inches)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Height",
+    subtitle = "(Among QBs, RBs, WRs, and TEs)"
+  ) +
+  theme_classic()
+
+plotly::ggplotly(plot_scatterplot)
+```
+
+```{r}
+#| label: fig-scatterplotByPositionWithBestFitLine
+#| fig-cap: "Scatterplot of Fantasy Points (Season) by Player Height and Position With Best-Fit Line."
+#| fig-alt: "Scatterplot of Fantasy Points (Season) by Player Height and Position With Best-Fit Line."
+
+plot_scatterplotByPosition <- ggplot2::ggplot(
+  data = player_stats_seasonal %>% filter(position %in% c("QB","RB","WR","TE")),
+  aes(
+    x = height,
+    y = fantasyPoints,
+    color = position,
+    fill = position)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season), # add season for mouse over tooltip
+    alpha = 0.7) +
+  geom_smooth(
+    method = "lm",
+    color = "black") +
+  geom_smooth(
+    method = "loess",
+    span = 0.5) +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Player Height (Inches)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Height and Position"
+  ) +
+  theme_classic()
+
+plotly::ggplotly(plot_scatterplotByPosition)
+```
+
 ### Rank Correlation: Spearman's Rho ($\rho$) {#sec-correlationExamplesSpearman}
@@
 In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @sec-assumptionsRegression.
-Below, we demonstrate how to account for non-independence of observations using cluster-robust standard errors.
-
-```{r}
-
-```
+In section @sec-multipleRegressionNonIndependence, we demonstrate how to account for non-independence of observations using cluster-robust standard errors.
 

### 2025-04-09 mixed-models.qmd
20250408 - plotly

@@
 
-ggplotly(
+plotly::ggplotly(
   plot_scatterplotFantasyPointsByAgeQB,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_scatterplotFantasyPointsByAgeFB,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_scatterplotFantasyPointsByAgeRB,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_scatterplotFantasyPointsByAgeWR,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_scatterplotFantasyPointsByAgeTE,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_rawFantasyPointsByAgeQB,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_rawFantasyPointsByAgeFB,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_rawFantasyPointsByAgeRB,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_rawFantasyPointsByAgeWR,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_rawFantasyPointsByAgeTE,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsRandomIntercepts,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsPosition,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsFixedLinearSlopes,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsRandomLinearSlopes,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsRandomLinearFixedQuadracticSlopes,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsRandomLinearFixedQuadraticSlopesInteraction,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsByAgeQB,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsByAgeFB,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsByAgeRB,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsByAgeWR,
@@
 
-ggplotly(
+plotly::ggplotly(
   plot_individualFantasyPointsByAgeTE,

### 2025-04-09 time-series-analysis.qmd
20250308 - R packages
20250308 - citations
20250408 - plotly

@@
 
-Adapted from: <https://rc2e.com/timeseriesanalysis> (archived at <https://perma.cc/U5P6-2VWC>).
+Adapted from: <https://rc2e.com/timeseriesanalysis> [@Long2019; archived at <https://perma.cc/U5P6-2VWC>].
 
@@
 ### Fit an Autoregressive Integrated Moving Average Model {#sec-arimaModel}
+
+Using the `forecast` package [@R-forecast; @Hyndman2008_packages]:
 

### 2025-04-09 multiple-regression.qmd
20250408 - non-independence

@@
 
+Please note that the $p$-value for regression coefficients assumes that the observations are independent—in particular, that the residuals are not correlated.
+However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
+This non-independence violates the traditional assumptions of the significance of regression coefficients.
+We could address this assumption by analyzing only one season from each player or by estimating the significance of the regression coefficients using cluster-robust standard errors.
+For simplicity in the models above, we present results above from the whole dataframe.
+In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence.
+Below, we demonstrate how to account for non-independence of observations using cluster-robust standard errors.
+
 ```{r}

### 2025-04-09 correlation.qmd
20250408 - visualization and partial correlation

--- 2025-04-08 correlation.qmd
@@
 library("DescTools")
+library("psych")
 library("corrplot")
@@
 
+### Scatterplot with Best-Fit Line
+
+A scatterplot with best-fit line is in @fig-scatterplotWithBestFitLine.
+
+```{r}
+#| label: fig-scatterplotWithBestFitLine
+#| fig-cap: "Scatterplot of Fantasy Points (Season) by Player Height With Best-Fit Line. The linear best-fit line is in black. The nonlinear best-fit line is in blue."
+#| fig-alt: "Scatterplot of Fantasy Points (Season) by Player Height With Best-Fit Line. The linear best-fit line is in black. The nonlinear best-fit line is in blue."
+
+plot_scatterplot <- ggplot2::ggplot(
+  data = player_stats_seasonal %>% filter(position %in% c("QB","RB","WR","TE")),
+  aes(
+    x = height,
+    y = fantasyPoints)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season), # add season for mouse over tooltip
+    alpha = 0.05) +
+  geom_smooth(
+    method = "lm",
+    color = "black") +
+  geom_smooth() +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Player Height (Inches)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Height",
+    subtitle = "(Among QBs, RBs, WRs, and TEs)"
+  ) +
+  theme_classic()
+
+plotly::ggplotly(plot_scatterplot)
+```
+
+Examining the linear line of best fit, we would see that there is a slight negative slope, consistent with a weak negative association.
+However, the nonlinear best-fit line suggests that there may be an inverted-U-shaped association, which might suggest that there may be an "optimal range of height," where being too tall or too short may be a disadvantage.
+We now examine the association between height and fantasy points separately by position in @fig-scatterplotByPositionWithBestFitLine.
+
+```{r}
+#| label: fig-scatterplotByPositionWithBestFitLine
+#| fig-cap: "Scatterplot of Fantasy Points (Season) by Player Height and Position With Best-Fit Line.
+The linear best-fit line for each position is in black.
+The nonlinear best-fit line for each group is represented by the color line."
+#| fig-alt: "Scatterplot of Fantasy Points (Season) by Player Height and Position With Best-Fit Line.
+The linear best-fit line for each position is in black.
+The nonlinear best-fit line for each group is represented by the color line."
+
+plot_scatterplotByPosition <- ggplot2::ggplot(
+  data = player_stats_seasonal %>% filter(position %in% c("QB","RB","WR","TE")),
+  aes(
+    x = height,
+    y = fantasyPoints,
+    color = position,
+    fill = position)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season), # add season for mouse over tooltip
+    alpha = 0.7) +
+  geom_smooth(
+    method = "lm",
+    color = "black") +
+  geom_smooth(
+    method = "loess",
+    span = 0.5) +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Player Height (Inches)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Height and Position"
+  ) +
+  guides(fill = "none") +
+  theme_classic()
+
+plotly::ggplotly(plot_scatterplotByPosition)
+```
+
+Examining the association between height and fantasy points separately by position, it appears that there is an inverted-U-shaped associations for Wide Receivers, a relatively flat association for Running Backs and Tight Ends, and a nonlinear association among Quarterbacks.
+Interestingly, the best-fit lines suggest that the different positions have different height ranges, which is confirmed in @fig-ridgelinePlotPlayerHeightByPosition.
+
+```{r}
+#| label: fig-ridgelinePlotPlayerHeightByPosition
+#| fig-cap: "Ridgeline Plot of Player Height by Position."
+#| fig-alt: "Ridgeline Plot of Player Height by Position."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = height,
+    y = position,
+    group = position,
+    fill = position)
+) +
+  ggridges::geom_density_ridges(
+    bandwidth = 0.6
+  ) +
+  labs(
+    x = "Height (inches)",
+    y = "Position",
+    title = "Ridgeline Plot of Player Height by Position"
+  ) +
+  theme_classic() + 
+  theme(
+    legend.position = "none",
+    axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+According to the distributions of player height by position (among Quarterbacks, Running Backs, Wide Receivers, and Tight Ends), Running Backs tend to be the shortest, followed by Wide Receivers; Tight Ends tend to be the tallest, followed by Quarterbacks.
+Wide Receivers showed the greatest variability; some were quite short, whereas others were quite tall.
+
 ### Rank Correlation: Spearman's Rho ($\rho$) {#sec-correlationExamplesSpearman}
@@
 
-### Partial Correlation {#sec-correlationExamplesNonlinear}
+### Partial Correlation {#sec-correlationExamplesPartial}
+
+A partial correlation examines the association between variables after controlling for one or more variables.
+Below, we examine a partial [Pearson correlation](#sec-correlationExamplesPearson) between height and fantasy points after controlling for age.
+
+```{r}
+psych::partial.r(
+  data = player_stats_seasonal,
+  x = c("height","fantasyPoints"),
+  y = "age"
+)
+```
+
+Below, we examine a partial [Spearman's rho rank correlation](#sec-correlationExamplesSpearman) between height and fantasy points after controlling for age.
+
+```{r}
+psych::partial.r(
+  data = player_stats_seasonal,
+  x = c("draftround","fantasyPoints"),
+  y = "age",
+  method = "spearman"
+)
+```
 

### 2025-04-09 correlation.qmd
20250409 - psych package

@@
 
-First, let's examine descriptive statistics (height is measured in inches):
+First, let's examine [descriptive statistics](#sec-descriptiveStatistics) of height and fantasy points (height is measured in inches):
 
@@
 A partial correlation examines the association between variables after controlling for one or more variables.
-Below, we examine a partial [Pearson correlation](#sec-correlationExamplesPearson) between height and fantasy points after controlling for age.
+Below, we examine a partial [Pearson correlation](#sec-correlationExamplesPearson) between height and fantasy points after controlling for age using the the `partial.r()` function from the `psych` package [@R-psych].
 
@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `cor.table()` function that generates a correlation matrix of variables, including the $r$ value, number of observations ($n$), and $p$-value for each association.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `cor.table()` function that generates a correlation matrix of variables, including the $r$ value, number of observations ($n$), asterisks denoting statistical significance, and $p$-value for each association.
+The asterisks follow the following traditional convention for statistical significance: $^\dagger p < .1$; $^* p < .05$; $^{**} p < .01$; $^{***} p < .001$.
 Here is a Pearson correlation matrix:
@@
 
-Here is a correlation matrix of variables that includes just the $r$ value and asterisks denoting the statistical significance of the association, for greater concision: $^\dagger p < .1$; $^* p < .05$; $^{**} p < .01$; $^{***} p < .001$.
+Here is a correlation matrix of variables that includes just the $r$ value and asterisks denoting the statistical significance of the association, for greater concision.
 
@@
 
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `partialcor.table()` function that generates a partial correlation matrix.
 Here is a partial correlation matrix, controlling for the player's age:

### 2025-04-09 cluster-analysis.qmd
20250408 - plotly
20250409 - fix error

@@
 
-ggplotly(plot_qbTiers)
+plotly::ggplotly(plot_qbTiers)
 ```
@@
 
-ggplotly(plot_rbTiers)
+plotly::ggplotly(plot_rbTiers)
 ```
@@
 
-ggplotly(plot_wrTiers)
+plotly::ggplotly(plot_wrTiers)
 ```
@@
 
-ggplotly(plot_teTiers)
+plotly::ggplotly(plot_teTiers)
 ```
@@
 
-#names(nfl_players) #gsis_id: height, weight
-#names(nfl_combine) #gsis_id: vertical, forty, ht, wt
-#names(player_stats_seasonal_offense) #player_id, season: receptions, receiving_air_yards, air_yards_share, target_share
-#names(nfl_advancedStatsPFR_seasonal) #gsis_id, season: adot.rec, rec.rec, brk_tkl.rec, drop.rec, drop_percent.rec
-#names(nfl_actualStats_offense_career) #player_id: receptions, targets, receiving_air_yards, air_yards_share, target_share
+# names(nfl_players) #gsis_id: height, weight
+# names(nfl_combine) #gsis_id: vertical, forty, ht, wt
+# names(player_stats_seasonal_offense) #player_id, season: receptions, receiving_air_yards, air_yards_share, target_share
+# names(nfl_advancedStatsPFR_seasonal) #gsis_id, season: adot.rec, rec.rec, brk_tkl.rec, drop.rec, drop_percent.rec
+# names(nfl_actualStats_career_player_inclPost) #player_id: receptions, targets, receiving_air_yards, air_yards_share, target_share
 ```
@@
     rename(gsis_id = player_id),
-  nfl_actualStats_offense_career %>% select(player_id, receptions, targets, receiving_air_yards, air_yards_share, target_share) %>% 
+  nfl_actualStats_career_player_inclPost %>% select(player_id, receptions, targets, receiving_air_yards, air_yards_share, target_share) %>% 
     rename(gsis_id = player_id),

### 2025-04-09 mixed-models.qmd
20250409 - minor tweaks

@@
 
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_depthCharts.RData", fsep = ""))
+load(file = "./data/nfl_depthCharts.RData")
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))

### 2025-04-09 correlation.qmd
20250409 - Anscombe's quartet

@@
 
-The formula for the correlation coefficient\index{correlation} is in @eq-pearsonCorrelation.
+The formula for the ([Pearson](#sec-correlationExamplesPearson)) correlation coefficient\index{correlation} is in @eq-pearsonCorrelation.
 As noted in @sec-statisticalTestsPartitionedVariance, the correlation coefficient can be thought of as the ratio of shared variance (i.e., covariance) to total variance, as in @eq-correlationPartitionedVariance.
@@
 An interactive visualization by @Magnusson2023 on interpreting correlations is at the following link: <https://rpsychologist.com/correlation/> (archived at <https://perma.cc/G8YR-VCM4>)
+
+Keep in mind that the ([Pearson](#sec-correlationExamplesPearson)) correlation examines the strength of the *linear* association between two variables.\index{correlation}
+If the association between two variables is nonlinear, the ([Pearson](#sec-correlationExamplesPearson)) correlation provides the strength of the linear trend and may not provide a meaningful index of the strength of the association between the variables.\index{correlation}
+For instance, Anscombe's quartet includes four sets of data that have nearly identical basic descriptive statistics (see Tables [-@tbl-anscombe] and [-@tbl-anscombeStats]), including the same bivariate correlation, yet have very different distributions and whose association takes very different forms (see Figure \@ref(fig:anscombeQuartet)).\index{correlation}
+
+| x1 | y1    | x2 | y2   | x3 | y3    | x4 | y4   |
+|----|-------|----|------|----|-------|----|------|
+| 10 | 8.04  | 10 | 9.14 | 10 | 7.46  | 8  | 6.58 |
+| 8  | 6.95  | 8  | 8.14 | 8  | 6.77  | 8  | 5.76 |
+| 13 | 7.58  | 13 | 8.74 | 13 | 12.74 | 8  | 7.71 |
+| 9  | 8.81  | 9  | 8.77 | 9  | 7.11  | 8  | 8.84 |
+| 11 | 8.33  | 11 | 9.26 | 11 | 7.81  | 8  | 8.47 |
+| 14 | 9.96  | 14 | 8.1  | 14 | 8.84  | 8  | 7.04 |
+| 6  | 7.24  | 6  | 6.13 | 6  | 6.08  | 8  | 5.25 |
+| 4  | 4.26  | 4  | 3.1  | 4  | 5.39  | 19 | 12.5 |
+| 12 | 10.84 | 12 | 9.13 | 12 | 8.15  | 8  | 5.56 |
+| 7  | 4.82  | 7  | 7.26 | 7  | 6.42  | 8  | 7.91 |
+| 5  | 5.68  | 5  | 4.74 | 5  | 5.73  | 8  | 6.89 |
+
+: Anscombe's Quartet {#tbl-anscombe}
+
+| Property                     | Value        |
+|------------------------------|--------------|
+| Sample size                  | 11           |
+| Mean of X                    | 9.0          |
+| Mean of Y                    | ~7.5         |
+| Variance of X                | 11.0         |
+| Variance of Y                | ~4.1         |
+| Equation of regression line  | Y = 3 + 0.5X |
+| Standard error of slope      | 0.118        |
+| One-sample t-statistic       | 4.24         |
+| Sum of squares of X          | 110.0        |
+| Regression sum of squares    | 27.50        |
+| Residual sum of squares of Y | 13.75        |
+| Correlation coefficient      | .816         |
+| Coefficient of determination | .67          |
+
+: Descriptive Statistics of Anscombe's Quartet {#tbl-anscombeStats}
+
+```{r}
+#| label: fig-anscombeQuartet
+#| fig-cap: "Anscombe's Quartet."
+#| fig-alt: "Anscombe's Quartet."
+#| code-fold: true
+
+anscombe
+
+par(mfrow = c(2,2))
+
+plot(
+  anscombe$x1,
+  anscombe$y1,
+  xlab = "x",
+  ylab = "y",
+  xlim = c(0,20),
+  ylim = c(0,15),
+  main = 
+    substitute(
+      paste(italic(r), " = ", x, sep = ""),
+      list(x = round(cor.test(x = anscombe$x1, y = anscombe$y1)$estimate, 2))))
+
+abline(lm(
+  y1 ~ x1,
+  data = anscombe),
+  col = "black",
+  lty = 2)
+
+plot(
+  anscombe$x2,
+  anscombe$y2,
+  xlab = "x",
+  ylab = "y",
+  xlim = c(0,20),
+  ylim = c(0,15),
+  main = substitute(
+    paste(italic(r), " = ", x, sep = ""),
+    list(x = round(cor.test(x = anscombe$x2, y = anscombe$y2)$estimate, 2))))
+
+abline(lm(
+  y2 ~ x2,
+  data = anscombe),
+  col = "black",
+  lty = 2)
+
+plot(
+  anscombe$x3,
+  anscombe$y3,
+  xlab = "x",
+  ylab = "y",
+  xlim = c(0,20),
+  ylim = c(0,15),
+  main = substitute(
+    paste(italic(r), " = ", x, sep = ""),
+    list(x = round(cor.test(x = anscombe$x3, y = anscombe$y3)$estimate, 2))))
+
+abline(lm(
+  y3 ~ x3,
+  data = anscombe),
+  col = "black",
+  lty = 2)
+
+plot(
+  anscombe$x4,
+  anscombe$y4,
+  xlab = "x",
+  ylab = "y",
+  xlim = c(0,20),
+  ylim = c(0,15),
+  main = substitute(
+    paste(italic(r), " = ", x, sep = ""),
+    list(x = round(cor.test(x = anscombe$x4, y = anscombe$y4)$estimate, 2))))
+
+abline(lm(
+  y4 ~ x4,
+  data = anscombe),
+  col = "black",
+  lty = 2)
+```
 

### 2025-04-09 download-football-data.qmd
20250409 - remove implausible values

@@
 
+Let's remove infinite (implausible) values:
+
+```{r}
+nfl_actualStats_seasonal_player[sapply(nfl_actualStats_seasonal_player, is.infinite)] <- NA
+nfl_actualStats_seasonal_team[sapply(nfl_actualStats_seasonal_team, is.infinite)] <- NA
+nfl_actualStats_seasonal_player_inclPost[sapply(nfl_actualStats_seasonal_player_inclPost, is.infinite)] <- NA
+nfl_actualStats_seasonal_team_inclPost[sapply(nfl_actualStats_seasonal_team_inclPost, is.infinite)] <- NA
+```
+
 ```{r}

### 2025-04-09 correlation.qmd
20250409 - nonlinear correlation, robust correlation, outliers, restricted range

@@
 library("corrplot")
+library("correlation")
 library("tidyverse")
@@
 
+An example of a nonlinear correlation is a distance correlation and the xi ($\xi$) coefficient.
+These estimates of nonlinear correlation only range from 0–1 (where 0 represents no association and 1 represents a strong association), so they are interpreted differently than a traditional correlation coefficient (which can be negative).
+We can estimate a distance correlation using the `correlation()` function of the correlation package [@R-correlation; @Makowski2020].
+
+```{r}
+correlation(
+  data = player_stats_seasonal %>% select(draftround, fantasyPoints),
+  method = "distance"
+)
+```
+
 A nonlinear correlation can be estimated using the xi ($\xi$) coefficient from the `XICOR` package [@R-XICOR; @Chatterjee2021]:
@@
 
+The correlation coefficient is strongly impacted by outliers—i.e., extreme (i.e., very large or very small) values relative to the other observations.
+Consider the following example:
+
+```{r}
+set.seed(52242)
+
+v1 <- rnorm(10)
+v2 <- rnorm(10)
+
+cor.test(
+  v1,
+  v2
+)
+```
+
+The associated scatterplot and best-fit line is in @fig-plotWithoutOutliers.
+
+```{r}
+#| label: fig-plotWithoutOutliers
+#| fig-cap: "Scatterplot Without Strong Outliers."
+#| fig-alt: "Scatterplot Without Strong Outliers."
+#| code-fold: true
+
+plot(
+  v1,
+  v2
+)
+
+petersenlab::addText(
+  v1,
+  v2,
+  ycoord = 0.7
+)
+
+abline(lm(
+  v2 ~ v1),
+  col = "black")
+```
+
+Now, let's add one outlier.
+
+```{r}
+v1[length(v1) + 1] <- 4
+v2[length(v2) + 1] <- 4
+
+cor.test(
+  v1,
+  v2
+)
+```
+
+The associated scatterplot and best-fit line for the updated data are in @fig-plotWithOutlier.
+
+```{r}
+#| label: fig-plotWithOutlier
+#| fig-cap: "Scatterplot With Outlier."
+#| fig-alt: "Scatterplot With Outlier."
+#| code-fold: true
+
+plot(
+  v1,
+  v2
+)
+
+petersenlab::addText(
+  v1,
+  v2
+)
+
+points(
+  v1[length(v1)],
+  v2[length(v2)],
+  pch = 19,
+  col = "red"
+)
+
+text(
+  x = 3.8,
+  y = 3.8,
+  labels = "outlier")
+
+abline(lm(
+  v2 ~ v1),
+  col = "black")
+```
+
+Note how the association was not close to being statistically significant with the outlier, but with the outlier added, the association is statistically significant.
+One way to combat this is to use methods for estimating the association between variables that are robust to (i.e., less impacted by) outliers.
+In the next section, we describe robust correlation methods.
+
 ### Examining Robust Correlation {#sec-correlationRobust}
 
+There are various approaches to estimating correlation in the presence of outliers—so-called *robust correlation* methods.
+
+One approach is the biweight midcorrelation, which is based on the median rather than the mean, and is thus less sensitive to outliers.
+Another approach is the percentage bend correlation, which gives less weight to observations that are farther away from the median.
+We can estimate the each using the `correlation()` function of the correlation package [@R-correlation; @Makowski2020].
+
+```{r}
+correlation(
+  data = data.frame(v1, v2),
+  method = "biweight"
+)
+
+correlation(
+  data = data.frame(v1, v2),
+  method = "percentage"
+)
+```
+
 ## Impact of Restricted Range {#sec-correlationRestrictedRange}
+
+In addition to being [impacted by outliers](#sec-correlationOutliers), correlations can also be greatly impacted by restricted variability or restricted range [@Petersen2024a; @PetersenPrinciplesPsychAssessment].
+Correlation depends on variability; if there is no or limited variability, it can be difficult to detect an association with another variable.
+Thus, if a variable has restricted range—such as owing to a floor effect or ceiling effect—that tends to artificially weaken associations.
+For instance, a floor effect is one in which many of the scores are the lowest possible score.
+By contrast, a ceiling effect is one in which many of the scores are the highest possible score.
+
+Consider the following association between air yards share and receiving yards:
+
+```{r}
+cor.test(
+  player_stats_seasonal$air_yards_share,
+  player_stats_seasonal$receiving_yards
+)
+```
+
+There is a statistically significant, moderate positive association between air yards share and receiving yards, as depicted in @fig-scatterplotWithoutRangeRestriction.
+
+```{r}
+#| label: fig-scatterplotWithoutRangeRestriction
+#| fig-cap: "Scatterplot of Player Receiving Yards (Season) by Air Yard Share. The best-fit line is in black."
+#| fig-alt: "Scatterplot of Player Receiving Yards (Season) by Air Yard Share. The best-fit line is in black."
+#| code-fold: true
+
+plot_scatterplotWithoutRangeRestriction <- ggplot2::ggplot(
+  data = player_stats_seasonal %>% filter(air_yards_share > 0),
+  aes(
+    x = air_yards_share,
+    y = receiving_yards)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season), # add season for mouse over tooltip
+    alpha = 0.10) +
+  geom_smooth(
+    method = "lm",
+    color = "black") +
+  #geom_smooth() +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Air Yards Share",
+    y = "Receiving Yards",
+    title = "Receiving Yards (Season) by Air Yards Share",
+    subtitle = "(Among Players with Air Yards Share > 0)"
+  ) +
+  theme_classic()
+
+plotly::ggplotly(plot_scatterplotWithoutRangeRestriction)
+```
+
+Now, consider the same association when we restrict the range to examine only those players whose air yards share was more than 25 (thus resulting in less variability):
+
+```{r}
+cor.test(
+  player_stats_seasonal %>% 
+    filter(air_yards_share > 25) %>% 
+    select(air_yards_share) %>% 
+    pull(),
+  player_stats_seasonal %>% 
+    filter(air_yards_share > 25) %>% 
+    select(receiving_yards) %>% 
+    pull()
+)
+```
+
+The association is no longer statistically significant, and the association is weak in terms of [effect size](#sec-effectSizeThresholds), as depicted in @fig-scatterplotWithRangeRestriction.
+
+```{r}
+#| label: fig-scatterplotWithRangeRestriction
+#| fig-cap: "Scatterplot of Player Receiving Yards (Season) by Target Share. The linear best-fit line is in black. The nonlinear best-fit line is in blue."
+#| fig-alt: "Scatterplot of Player Receiving Yards (Season) by Target Share. The linear best-fit line is in black. The nonlinear best-fit line is in blue."
+#| code-fold: true
+
+plot_scatterplotWithRangeRestriction <- ggplot2::ggplot(
+  data = player_stats_seasonal %>% filter(air_yards_share > 25),
+  aes(
+    x = air_yards_share,
+    y = receiving_yards)) +
+  geom_point(
+    aes(
+      text = player_display_name, # add player name for mouse over tooltip
+      label = season)) + # add season for mouse over tooltip)
+  geom_smooth(
+    method = "lm",
+    color = "black") +
+  #geom_smooth() +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Air Yards Share",
+    y = "Receiving Yards",
+    title = "Receiving Yards (Season) by Air Yards Share",
+    subtitle = "(Among Players with Air Yards Share > 25)"
+  ) +
+  theme_classic()
+
+plotly::ggplotly(plot_scatterplotWithRangeRestriction)
+```
 

### 2025-04-09 index.qmd
20250409 - nonlinear correlation, robust correlation, outliers, restricted range

@@
   "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin",
-  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot" #"mgcv",
+  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation" #"mgcv",
 )

### 2025-04-09 correlation.qmd
20250409 - clean up and cross references

@@
 By contrast, correlation is in a standardized metric that does not differ for variables with different scales.
-When examining the association between variables that are [interval](#sec-interval) or [ratio](#sec-ratio) levels of measurement, Pearson correlation is used.
-When examining the association between variables that are [ordinal](#sec-ordinal) in level of measurement, Spearman correlation is used.
-Pearson correlation is an index of the *linear* association between variables.
+When examining the association between variables that are [interval](#sec-interval) or [ratio](#sec-ratio) levels of measurement, [Pearson correlation](#sec-correlationExamplesPearson) is used.
+When examining the association between variables that are [ordinal](#sec-ordinal) in level of measurement, [Spearman correlation](#sec-correlationExamplesSpearman) is used.
+[Pearson correlation](#sec-correlationExamplesPearson) is an index of the *linear* association between variables.
 If a nonlinear association is present, other indices like xi [$\xi$; @Chatterjee2021] and distance correlation coefficients are better suited to detect the association.
@@
 If the association between two variables is nonlinear, the ([Pearson](#sec-correlationExamplesPearson)) correlation provides the strength of the linear trend and may not provide a meaningful index of the strength of the association between the variables.\index{correlation}
-For instance, Anscombe's quartet includes four sets of data that have nearly identical basic descriptive statistics (see Tables [-@tbl-anscombe] and [-@tbl-anscombeStats]), including the same bivariate correlation, yet have very different distributions and whose association takes very different forms (see Figure \@ref(fig:anscombeQuartet)).\index{correlation}
+For instance, Anscombe's quartet includes four sets of data that have nearly identical basic descriptive statistics (see Tables [-@tbl-anscombe] and [-@tbl-anscombeStats]), including the same bivariate correlation, yet have very different distributions and whose association takes very different forms (see @fig-anscombeQuartet).\index{correlation}
 
 | x1 | y1    | x2 | y2   | x3 | y3    | x4 | y4    |
-|----|-------|----|------|----|-------|----|------|
+|---:|------:|---:|-----:|---:|------:|---:|------:|
 | 10 | 8.04  | 10 | 9.14 | 10 | 7.46  | 8  | 6.58  |
@@
 | 11 | 8.33  | 11 | 9.26 | 11 | 7.81  | 8  | 8.47  |
-| 14 | 9.96  | 14 | 8.1  | 14 | 8.84  | 8  | 7.04 |
+| 14 | 9.96  | 14 | 8.10 | 14 | 8.84  | 8  | 7.04  |
 | 6  | 7.24  | 6  | 6.13 | 6  | 6.08  | 8  | 5.25  |
-| 4  | 4.26  | 4  | 3.1  | 4  | 5.39  | 19 | 12.5 |
+| 4  | 4.26  | 4  | 3.10 | 4  | 5.39  | 19 | 12.50 |
 | 12 | 10.84 | 12 | 9.13 | 12 | 8.15  | 8  | 5.56  |
@@
 | Property                     | Value        |
-|------------------------------|--------------|
+|:-----------------------------|:-------------|
 | Sample size                  | 11           |
@@
 ```{r}
+#| output: false
+
+anscombe
+```
+
+```{r}
 #| label: fig-anscombeQuartet
@@
 #| code-fold: true
-
-anscombe
 
@@
 However, it can be helpful to use a standardized index to examine the [effect size](#sec-practicalSignificance)—i.e., the strength of the association.
-The [Pearson correlation coefficient](Pearson Correlation), $r$, is an example of a standardized index of the covariance between two variables.
+The [Pearson correlation coefficient](#sec-correlationExamplesPearson), $r$, is an example of a standardized index of the covariance between two variables.
 
@@
 The asterisks follow the following traditional convention for statistical significance: $^\dagger p < .1$; $^* p < .05$; $^{**} p < .01$; $^{***} p < .001$.
-Here is a Pearson correlation matrix:
+Here is a [Pearson correlation](#sec-correlationExamplesPearson) matrix:
 
@@
 
-Here is a Spearman correlation matrix:
+Here is a [Spearman correlation](#sec-correlationExamplesSpearman) matrix:
 
@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `partialcor.table()` function that generates a partial correlation matrix.
-Here is a partial correlation matrix, controlling for the player's age:
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `partialcor.table()` function that generates a [partial correlation](#sec-correlationExamplesPartial) matrix.
+Here is a [partial correlation](#sec-correlationExamplesPartial) matrix, controlling for the player's age:
 
@@
 #| label: fig-plotWithOutlier
-#| fig-cap: "Scatterplot With Outlier."
-#| fig-alt: "Scatterplot With Outlier."
+#| fig-cap: "Scatterplot With Outlier (in Red)."
+#| fig-alt: "Scatterplot With Outlier (in Red)."
 #| code-fold: true
@@
 
-Note how the association was not close to being statistically significant with the outlier, but with the outlier added, the association is statistically significant.
+Note how the association was not close to being statistically significant without the outlier, but with the outlier added, the association is statistically significant.
 One way to combat this is to use methods for estimating the association between variables that are robust to (i.e., less impacted by) outliers.

### 2025-04-10 simulation.qmd
Built site for gh-pages

@@
 ### Bootstrapping {#sec-simulationProjectedPointsBootstrapping}
+
+```{r}
+#| eval: false
+
+# https://github.com/FantasyFootballAnalytics/ffanalytics/blob/08cab3cbe6eb645b853bb814e671e186b78a0837/R/calc_projections.R#L218-L253
+
+# This gives an example for the data format accepted by the source_points() function
+l = ffanalytics:::scrape_espn(season = 2024, week = 0)
+
+# The two important things for the "data_result" object going into the source_points()
+# function are the format of the list (with positions as names) and the attributes:
+# season and week
+
+# Before sending the data_result to the source_points() function make sure it has these attributes set 
+# (e.g., this is for 2024 week 0)
+
+attr(data_result, "season") <- 2024
+attr(data_result, "week") <- 0
+
+df = ffanalytics:::source_points(l, scoring_rules = ffanalytics::scoring)
+```
 

### 2025-04-10 multiple-regression.qmd
20250410 - placeholder

@@
 library("rms")
+library("lavaan")
+library("mice")
+library("brms")
 library("tidyverse")
@@
 
-Consider an example where you develop an algorithm to predict players' fantasy performance based on 2023 data using hundreds of [predictor variables](#sec-correlationalStudy).
+Consider an example where you develop an algorithm to predict players' fantasy performance based on 2024 data using hundreds of [predictor variables](#sec-correlationalStudy).
 To some extent, these [predictor variables](#sec-correlationalStudy) will likely account for true variance (i.e., signal) and error variance (i.e., noise).
-If we were to apply the same algorithm based on the 2023 prediction model to 2024 data, the prediction model would likely predict less accurately than with 2023 data.
+If we were to apply the same algorithm based on the 2024 prediction model to 2025 data, the prediction model would likely predict less accurately than with 2024 data.
+The regression coefficients tend to become weaker when applied to new data, a phenomenon called [shrinkage](#sec-shrinkage), which is described in @sec-shrinkage.
 The regression coefficients in the [FILL IN]
@@
 Use of covariates examines the association between the [predictor variable](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy) when holding people's level constant on the covariates.
-Inclusion of confounds as covariates allows potentially gaining a more accurate estimate of the causal effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy).
-Ideally, you want to include any and all confounds as covariates.
-As described in @sec-correlationCausation, confounds are third variables that influence both the [predictor variable](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy) and explain their association.
-Covariates are potentially (but not necessarily) confounds.
-For instance, you might include the player's age as a covariate in a model that examines whether a player's 40-yard dash time at the NFL Combine predicts their fantasy points in their rookie year, but it may not be a confound.
+Inclusion of [confounds](#sec-causalDiagramConfounding) as covariates allows potentially gaining a more accurate estimate of the causal effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy).
+Ideally, you want to include any and all [confounds](#sec-causalDiagramConfounding) as covariates.
+As described in @sec-correlationCausation, [confounds](#sec-causalDiagramConfounding) are third variables that influence both the [predictor variable](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy) and explain their association.
+Covariates are potentially (but not necessarily) [confounds](#sec-causalDiagramConfounding).
+For instance, you might include the player's age as a covariate in a model that examines whether a player's 40-yard dash time at the NFL Combine predicts their fantasy points in their rookie year, but it may not be a [confound](#sec-causalDiagramConfounding).
+
+## Examples {#sec-multipleRegressionExamples}
 
@@
 
+## Handling of Missing Data {#sec-multipleRegressionMissingness}
+
+An important consideration in multiple regression is how missing data are handled.
+Multiple regression in `R` using the `lm()` function applies listwise deletion.
+*Listwise deletion* removes any row (in the data file) from analysis that has a missing value on the [outcome variable](#sec-correlationalStudy) or any of the [predictor variables](#sec-correlationalStudy).
+Removing all rows from analysis that have any missingness in the model variables can be a problem because missingness is often not completely at random—missingness often occurs systematically (i.e., for a reason).
+For instance, participants may be less likely to have data for all variables if they are from a lower socioeconomic status background and do not have the time to participate in all study procedures.
+Thus, applying listwise deletion, we might systematically exclude participants from lower socioeconomic status backgrounds (or other groups), which could lead to less generalizable inferences.
+
+It is thus important to consider approaches to handle missingness.
+Various approaches to handle missingness include pairwise deletion (aka available-case analysis), multiple imputation, and full information maximum likelihood (FIML).
+
+### Pairwise Deletion {#sec-multipleRegressionPairwiseDeletion}
+
+### Multiple Imputation {#sec-multipleRegressionMultipleImputation}
+
+### Full Information Maximum Likelihood {#sec-multipleRegressionFIML}
+
 ## Addressing Non-Independence of Observations {#sec-multipleRegressionNonIndependence}

### 2025-04-10 correlation.qmd
20250410 - placeholder
20250410 - update figures

@@
 #| fig-alt: "Anscombe's Quartet."
+#| fig-height: 6
+#| fig-width: 6
 #| code-fold: true
@@
 
-Consider the following association between air yards share and receiving yards:
+Consider the following association between passing attempts and expected points added (EPA) via passing.
+Expected points added from a given play is calculated as the difference between a team's expected points before the play and the team's expected points after the play.
+This can be summed across all passing plays to determine a player's EPA from passing during a given season.
 
@@
 cor.test(
-  player_stats_seasonal$air_yards_share,
-  player_stats_seasonal$receiving_yards
-)
-```
-
-There is a statistically significant, moderate positive association between air yards share and receiving yards, as depicted in @fig-scatterplotWithoutRangeRestriction.
+  player_stats_seasonal$attempts,
+  player_stats_seasonal$passing_epa
+)
+```
+
+There is a statistically significant, moderate positive association between passing attempts and expected points added (EPA) via passing, as depicted in @fig-scatterplotWithoutRangeRestriction.
 
@@
 #| label: fig-scatterplotWithoutRangeRestriction
-#| fig-cap: "Scatterplot of Player Receiving Yards (Season) by Air Yard Share. The best-fit line is in black."
-#| fig-alt: "Scatterplot of Player Receiving Yards (Season) by Air Yard Share. The best-fit line is in black."
+#| fig-cap: "Scatterplot of Player Passing Expected Points Added (Season) by Passing Attempts, Without Range Restriction. The best-fit line is in blue."
+#| fig-alt: "Scatterplot of Player Passing Expected Points Added (Season) by Passing Attempts, Without Range Restriction. The best-fit line is in blue."
 #| code-fold: true
@@
 plot_scatterplotWithoutRangeRestriction <- ggplot2::ggplot(
-  data = player_stats_seasonal %>% filter(air_yards_share > 0),
+  data = player_stats_seasonal,
   aes(
-    x = air_yards_share,
-    y = receiving_yards)) +
+    x = attempts,
+    y = passing_epa)) +
   geom_point(
@@
       label = season), # add season for mouse over tooltip
-    alpha = 0.10) +
+    alpha = 0.3) +
   geom_smooth(
-    method = "lm",
-    color = "black") +
+    method = "lm") +
   #geom_smooth() +
   coord_cartesian(
-    ylim = c(0,NA),
     expand = FALSE) +
   labs(
-    x = "Air Yards Share",
-    y = "Receiving Yards",
-    title = "Receiving Yards (Season) by Air Yards Share",
-    subtitle = "(Among Players with Air Yards Share > 0)"
+    x = "Player's Passing Attempts",
+    y = "Player's Expected Points Added (EPA) via Passing",
+    title = "Player's Passing Expected Points Added (EPA)\nby Passing Attempts (Season)",
+    #subtitle = ""
   ) +
@@
 
-Now, consider the same association when we restrict the range to examine only those players whose air yards share was more than 25 (thus resulting in less variability):
+Now, consider the same association when we restrict the range to examine only those players who had fewer than 450 pass attempts in a season (thus resulting in less variability):
 
@@
   player_stats_seasonal %>% 
-    filter(air_yards_share > 25) %>% 
-    select(air_yards_share) %>% 
+    filter(attempts < 450) %>% 
+    select(attempts) %>% 
     pull(),
   player_stats_seasonal %>% 
-    filter(air_yards_share > 25) %>% 
-    select(receiving_yards) %>% 
+    filter(attempts < 450) %>% 
+    select(passing_epa) %>% 
     pull()
@@
 
-The association is no longer statistically significant, and the association is weak in terms of [effect size](#sec-effectSizeThresholds), as depicted in @fig-scatterplotWithRangeRestriction.
+The association is no longer positive and is weak in terms of [effect size](#sec-effectSizeThresholds), as depicted in @fig-scatterplotWithRangeRestriction.
 
@@
 #| label: fig-scatterplotWithRangeRestriction
-#| fig-cap: "Scatterplot of Player Receiving Yards (Season) by Target Share. The linear best-fit line is in black. The nonlinear best-fit line is in blue."
-#| fig-alt: "Scatterplot of Player Receiving Yards (Season) by Target Share. The linear best-fit line is in black. The nonlinear best-fit line is in blue."
+#| fig-cap: "Scatterplot of Player Passing Expected Points Added (Season) by Passing Attempts, With Range Restriction. The best-fit line is in blue."
+#| fig-alt: "Scatterplot of Player Passing Expected Points Added (Season) by Passing Attempts, With Range Restriction. The best-fit line is in blue."
 #| code-fold: true
@@
 plot_scatterplotWithRangeRestriction <- ggplot2::ggplot(
-  data = player_stats_seasonal %>% filter(air_yards_share > 25),
+  data = player_stats_seasonal %>% filter(attempts < 450),
   aes(
-    x = air_yards_share,
-    y = receiving_yards)) +
+    x = attempts,
+    y = passing_epa)) +
   geom_point(
@@
       text = player_display_name, # add player name for mouse over tooltip
-      label = season)) + # add season for mouse over tooltip)
+      label = season), # add season for mouse over tooltip
+    alpha = 0.3) +
   geom_smooth(
-    method = "lm",
-    color = "black") +
+    method = "lm") +
   #geom_smooth() +
   coord_cartesian(
-    ylim = c(0,NA),
     expand = FALSE) +
   labs(
-    x = "Air Yards Share",
-    y = "Receiving Yards",
-    title = "Receiving Yards (Season) by Air Yards Share",
-    subtitle = "(Among Players with Air Yards Share > 25)"
+    x = "Player's Passing Attempts",
+    y = "Player's Expected Points Added (EPA) via Passing",
+    title = "Player's Passing Expected Points Added (EPA)\nby Passing Attempts (Season)",
+    subtitle = "(Among Players with < 450 Passing Attempts)"
   ) +

### 2025-04-10 multiple-regression.qmd
20250410 - multiple regression

@@
 library("rms")
+library("car")
 library("lavaan")
@@
 
+The *standard error* of a regression coefficient represents the uncertainty of the parameter.
+If we have less uncertainty (i.e., more confidence) about the parameter, the standard error will be small.
+If we have more uncertainty (i.e., less confidence) about the parameter, the standard error will be large.
+If we used the same sampling procedure repeatedly and calculated the regression coefficient each time, the true parameter in the population would fall 68% of the time within the interval of:  $[\text{model parameter estimate for the regression coefficient} \pm 1 \text{ standard error}]$ .
+
+A *confidence interval* represents a range of plausible values such that, with repeated sampling, the true value falls within a given interval with some confidence.
+Our parameter estimate for the regression coefficient, plus or minus 1 standard error, reflects the 68% confidence interval for the coefficient.
+The 95% confidence interval is computed as the parameter estimated plus or minus 1.96 standard errors.
+For instance, if the parameter estimate for the regression coefficient is 0.50, and the standard error is 0.10, the 95% confidence interval is [0.30, 0.70]: $0.5 - (1.96 \times 0.10) = 0.3$; $0.5 + (1.96 \times 0.10) = 0.7$.
+That is, if we used the same sample procedure repeatedly, the true value of the regression coefficient would be expected to be 95% of the time somewhere between 0.30 to 0.70.
+
 ## Assumptions of Multiple Regression {#sec-assumptionsRegression}
@@
 
-### Evaluating the Assumptions of Multiple Regression {#sec-assumptionsRegressionEvaluating}
+Homoscedasticity of the residuals means that the variance of the residuals does not differ as a function of the outcome variable or as a function of the predictor variables (i.e., the residuals show constant variance as a function of outcome/predictors).
+
+### Evaluating Assumptions of Multiple Regression {#sec-assumptionsRegressionEvaluating}
 
@@
 To some extent, these [predictor variables](#sec-correlationalStudy) will likely account for true variance (i.e., signal) and error variance (i.e., noise).
-If we were to apply the same algorithm based on the 2024 prediction model to 2025 data, the prediction model would likely predict less accurately than with 2024 data.
+If we were to apply the same algorithm based on the `r as.integer(format(Sys.Date(), "%Y")) - 1` prediction model to `r as.integer(format(Sys.Date(), "%Y"))` data, the prediction model would likely predict less accurately than with `r as.integer(format(Sys.Date(), "%Y")) - 1` data.
 The regression coefficients tend to become weaker when applied to new data, a phenomenon called [shrinkage](#sec-shrinkage), which is described in @sec-shrinkage.
@@
 ## Examples {#sec-multipleRegressionExamples}
+
+Let's say we want to use a number of variables to predict a wide receiver's fantasy performance.
+We want to consider several predictors, including the player's age, height, weight, and target share.
+We have only a few predictors and our sample size is large enough such that [overfitting](#sec-overfitting) is not likely a concern.
+
+Let's first examine the bivariate association of each using a scatterplot to evaluate for any potential nonlinearity, as depicted in @fig-scatterplots.
+
+```{r}
+#| layout-ncol: 2
+#| label: fig-scatterplots
+#| fig-cap: "Scatterplots With Fantasy Points (Season) Among Wide Receivers. The linear best-fit line is in black. The nonlinear best-fit line is in blue."
+#| fig-alt: "Scatterplots With Fantasy Points (Season) Among Wide Receivers. The linear best-fit line is in black. The nonlinear best-fit line is in blue."
+#| fig-subcap:
+#|   - "Age"
+#|   - "Height"
+#|   - "Weight"
+#|   - "Target Share"
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  aes(
+    x = age,
+    y = fantasyPoints)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth(
+    method = "lm",
+    color = "black") +
+  geom_smooth() +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Player Age (Years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age",
+    subtitle = "(Among Wide Receivers)"
+  ) +
+  theme_classic()
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  aes(
+    x = height,
+    y = fantasyPoints)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth(
+    method = "lm",
+    color = "black") +
+  geom_smooth() +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Player Height (Inches)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Height",
+    subtitle = "(Among Wide Receivers)"
+  ) +
+  theme_classic()
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  aes(
+    x = weight,
+    y = fantasyPoints)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth(
+    method = "lm",
+    color = "black") +
+  geom_smooth() +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Player Weight (Pounds)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Weight",
+    subtitle = "(Among Wide Receivers)"
+  ) +
+  theme_classic()
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  aes(
+    x = target_share,
+    y = fantasyPoints)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth(
+    method = "lm",
+    color = "black") +
+  geom_smooth() +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Target Share",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Target Share",
+    subtitle = "(Among Wide Receivers)"
+  ) +
+  theme_classic()
+```
+
+There are some suggestions of potential nonlinearity, such as an inverted-U-shaped association between height and fantasy points, suggesting that there may an optimal range for height among Wide Receivers—being too short or too tall could be a disadvantage.
+In addition, target share shows a weakening association as target share increases.
+Thus, after evaluating the linear association between the predictors and outcome, we will also examine the possibility for curvilinear associations.
+
+First, let's estimate a multiple regression model with only linear terms:
+
+```{r}
+linearRegressionModel <- lm(
+  fantasyPoints ~ age + height + weight + target_share,
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  na.action = "na.exclude"
+)
+```
+
+Here are the model results:
+
+```{r}
+summary(linearRegressionModel)
+```
+
+The only terms that are significantly associated with fantasy performance among Wide Receivers are weight and target share, both of which showed a positive association with fantasy points.
+If we want to obtain standardized regression coefficients, we can standardize the outcome variable and each predictor variable using the `scale()` function:
+
+```{r}
+linearRegressionModelStandardized <- lm(
+  scale(fantasyPoints) ~ scale(age) + scale(height) + scale(weight) + scale(target_share),
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  na.action = "na.exclude"
+)
+
+summary(linearRegressionModelStandardized)
+```
+
+Target share has a large [effect size](#sec-practicalSignificance).
+All of the other predictors have small effect sizes.
+
+Now let's consider whether any of the terms show curvilinear associations with fantasy points by adding quadratic terms:
+
+```{r}
+quadraticTermsRegressionModel <- lm(
+  fantasyPoints ~ age + I(age^2) + height + I(height^2) + weight + I(weight^2) + target_share + I(target_share^2),
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  na.action = "na.exclude"
+)
+
+summary(quadraticTermsRegressionModel)
+```
+
+Only the quadratic term for target share shows a significant association.
+If we wanted to visualize the shape of the model-implied association between target share and fantasy points, we could generate the model-implied predictions using the data range that we want to visualize.
+
+```{r}
+newdata <- data.frame(
+  target_share = seq(
+    from = min(player_stats_seasonal$target_share[which(player_stats_seasonal$position == "WR")], na.rm = TRUE),
+    to = max(player_stats_seasonal$target_share[which(player_stats_seasonal$position == "WR")], na.rm = TRUE),
+    length.out = 1000
+  )
+)
+
+newdata$age <- mean(player_stats_seasonal$age[which(player_stats_seasonal$position == "WR")], na.rm = TRUE)
+newdata$height <- mean(player_stats_seasonal$height[which(player_stats_seasonal$position == "WR")], na.rm = TRUE)
+newdata$weight <- mean(player_stats_seasonal$weight[which(player_stats_seasonal$position == "WR")], na.rm = TRUE)
+```
+
+```{r}
+newdata$fantasyPoints <- predict(
+  quadraticTermsRegressionModel,
+  newdata = newdata
+)
+```
+
+We can depict the model-implied predictions of fantasy points as a function of target share, as shown in @fig-modelImpliedPredictionsQuadraticRegression.
+
+```{r}
+#| label: fig-modelImpliedPredictionsQuadraticRegression
+#| fig-cap: "Model-Implied Predictions of A Wide Receiver's Fantasy Points as a Function of Target Share. The model-implied predictions were estimated based on a multiple regression model."
+#| fig-alt: "Model-Implied Predictions of A Wide Receiver's Fantasy Points as a Function of Target Share. The model-implied predictions were estimated based on a multiple regression model."
+
+ggplot2::ggplot(
+  data = newdata,
+  aes(
+    x = target_share,
+    y = fantasyPoints)) +
+  geom_smooth() +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Target Share",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Target Share",
+    subtitle = "(Among Wide Receivers)"
+  ) +
+  theme_classic()
+```
+
+### Evaluating and Addressing Assumptions {#sec-multipleRegressionExamplesAddressingAssumptions}
+
+The assumptions for multiple regression are described in @sec-assumptionsRegression.
+I describe ways to evaluate assumptions in @sec-assumptionsRegressionEvaluating.
+
+As a reminder, here are four assumptions:
+
+- there is a linear association between the predictor variables and the outcome variable
+- there is homoscedasticity of the residuals; the residuals do not differ as a function of the predictor variables or as a function of the outcome variable
+- the residuals are independent; they are uncorrelated with each other
+- the residuals are normally distributed
+
+#### Linear Association {#sec-multipleRegressionExamplesAddressingAssumptionsLinearAssociation}
+
+We evaluated the shape of the association between the predictor variables and the outcome variables using scatterplots.
+We accounted for potential curvilinearity with a quadratic term.
+Other ways to account for nonlinearity, in addition to polynomials (e.g., quadratic, cubic, quartic, or higher-degree terms), include transforming predictors (e.g., log, square root, inverse, exponential transform), splines/piecewise regression, and generalized additive models.
+
+We can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), and added-variable plots (@fig-addedVariablePlots), to evaluate for potential nonlinearity.
+For evaluating linearity, we would expect minimal bend/curvature in the lines.
+
+```{r}
+#| label: fig-marginalModelPlots
+#| fig-cap: "Marginal Model Plots."
+#| fig-alt: "Marginal Model Plots."
+
+car::marginalModelPlots(
+  quadraticTermsRegressionModel,
+  sd = TRUE,
+  id = TRUE)
+```
+
+```{r}
+#| label: fig-addedVariablePlots
+#| fig-cap: "Added-Variable Plots."
+#| fig-alt: "Added-Variable Plots."
+
+car::avPlots(
+  quadraticTermsRegressionModel,
+  id = TRUE)
+```
+
+The marginal model plots (@fig-marginalModelPlots) and residual plots (@fig-residualPlots) suggest that the nonlinearity of the association between target share and fantasy points may not be fully captured by the quadratic term.
+Thus, we may need to apply a different approach to handling the nonlinear association between target share and fantasy points.
+
+One approach we can take is to transform the `target_shares` variable to be more normally distributed.
+
+The histogram for the raw `target_shares` variable is in @fig-histogramTargetShare.
+
+```{r}
+#| label: fig-histogramTargetShare
+#| fig-cap: "Histogram of Target Share."
+#| fig-alt: "Histogram of Target Share."
+
+hist(player_stats_seasonal$target_share[which(player_stats_seasonal$position == "WR")])
+```
+
+The variable shows a strong positive skew.
+To address a strong positive skew, we can use a log transformation.
+The histogram of the log-transformed variable is in @fig-histogramTargetSharesLog.
+
+```{r}
+#| label: fig-histogramTargetSharesTransformed
+#| fig-cap: "Histogram of Target Share, Transformed."
+#| fig-alt: "Histogram of Target Share, Transformed."
+
+hist(log(player_stats_seasonal$target_share[which(player_stats_seasonal$position == "WR")] + 1))
+```
+
+Now we can re-fit the model with the log-transformed variable.
+
+```{r}
+linearRegressionModel_logTargetShare <- lm(
+  fantasyPoints ~ age + I(age^2) + height + I(height^2) + weight + I(weight^2) + I(log(target_share + 1)),
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  na.action = "na.exclude"
+)
+
+summary(linearRegressionModel_logTargetShare)
+```
+
+Target share shows a more linear association with fantasy points after log-transforming it (albeit still not perfect), as depicted in Figures [-@fig-marginalModelPlotsLogTransformed], [-@fig-addedVariablePlotsLogTransformed], and [-@fig-residualPlotsLogTransformed].
+
+```{r}
+#| label: fig-marginalModelPlotsLogTransformed
+#| fig-cap: "Marginal Model Plots After Log Transformation of Target Share."
+#| fig-alt: "Marginal Model Plots After Log Transformation of Target Share."
+
+car::marginalModelPlots(
+  linearRegressionModel_logTargetShare,
+  sd = TRUE,
+  id = TRUE)
+```
+
+```{r}
+#| label: fig-addedVariablePlotsLogTransformed
+#| fig-cap: "Added-Variable Plots After Log Transformation of Target Share."
+#| fig-alt: "Added-Variable Plots After Log Transformation of Target Share."
+
+car::avPlots(
+  linearRegressionModel_logTargetShare,
+  id = TRUE)
+```
+
+```{r}
+#| label: fig-residualPlotsLogTransformed
+#| fig-cap: "Residual Plots After Log Transformation of Target Share."
+#| fig-alt: "Residual Plots After Log Transformation of Target Share."
+
+car::residualPlots(
+  linearRegressionModel_logTargetShare,
+  id = TRUE)
+```
+
+#### Homoscedasticity {#sec-multipleRegressionExamplesAddressingAssumptionsHomoscedasticity}
+
+To evaluate homoscedasticity, we can evaluate residual plots (@fig-residualPlots).
+In residual plots, you want a constant spread of residuals versus fitted values—you do not want the residuals to show a fan or cone shape.
+
+```{r}
+#| label: fig-residualPlots
+#| fig-cap: "Residual Plots."
+#| fig-alt: "Residual Plots."
+
+car::residualPlots(
+  quadraticTermsRegressionModel,
+  id = TRUE)
+```
+
+In this example, the residuals appear to increase as a function of the fitted values.
+To handle this, we may need to transform the outcome variable to be more normally distributed.
+
+The histogram for target share is in @fig-histogramFantasyPoints
+
+```{r}
+#| label: fig-histogramFantasyPoints
+#| fig-cap: "Histogram of Fantasy Points (Among Wide Receivers)."
+#| fig-alt: "Histogram of Fantasy Points (Among Wide Receivers)."
+
+hist(player_stats_seasonal$fantasyPoints[which(player_stats_seasonal$position == "WR")])
+```
+
+The histogram of the transformed variable is in @fig-histogramFantasyPointsTransformed.
+
+```{r}
+#| label: fig-histogramFantasyPointsTransformed
+#| fig-cap: "Histogram of Fantasy Points (Among Wide Receivers), Transformed."
+#| fig-alt: "Histogram of Fantasy Points (Among Wide Receivers), Transformed."
+
+hist(player_stats_seasonal$fantasyPoints[which(player_stats_seasonal$position == "WR")])
+```
+
+Now we can refit the model.
+
+```{r}
+linearRegressionModel_outcomeTransformed <- lm(
+  fantasyPoints ~ age + height + weight + I(log(target_share + 1)),
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  na.action = "na.exclude"
+)
+
+summary(linearRegressionModel_outcomeTransformed)
+```
+
+The residual plots are in @fig-residualPlotsOutcomeTransformed.
+
+```{r}
+#| label: fig-residualPlotsOutcomeTransformed
+#| fig-cap: "Residual Plots After Transformation of Fantasy Points."
+#| fig-alt: "Residual Plots After Transformation of Fantasy Points."
+
+car::residualPlots(
+  linearRegressionModel_outcomeTransformed,
+  id = TRUE)
+```
+
+#### Uncorrelated Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsUncorrelatedResiduals}
+
+#### Normally Distributed Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsNormallyDistributedResiduals}
 

### 2025-04-11 multiple-regression.qmd
20250410 - multiple regression

@@
 library("car")
+library("caret")
+library("lme4")
+library("performance")
 library("lavaan")
 library("mice")
+library("miceadds")
 library("brms")
@@
 ### Evaluating Assumptions of Multiple Regression {#sec-assumptionsRegressionEvaluating}
+
+To evaluate the shape of the association between the predictor variables and the outcome variable, we can examine scatterplots, residual plots, marginal model plots, and added-variable plots.
+
+To evaluate homoscedasticity, we can evaluate residual plots (@fig-residualPlots).
+In residual plots, homoscedasticity is identified as a constant spread of residuals versus fitted values—the residuals do not show a fan or cone shape.
+
+To determine if residuals are correlated by a grouping level, we can examine the proportion of variance that is attributable to the grouping level using the intraclass correlation coefficient (ICC) from a [mixed model](#sec-mixedModels).
+The greater the ICC value, the more variance is accounted for by the grouping level, and the more the residuals are intercorrelated.
+
+We can examine whether residuals are normally distributed using quantile–quantile (QQ) plots and probability–probability (PP) plots.
+If the residuals are normally distributed, they should stay close to the diagonal reference line.
 
@@
 
-The histogram for target share is in @fig-histogramFantasyPoints
+The histogram for raw fantasy points is in @fig-histogramFantasyPoints.
 
@@
 
+Let's use a Yeo-Johnson transformation of fantasy points:
+
+```{r}
+yjTransformed <- caret::preProcess(
+  player_stats_seasonal["fantasyPoints"],
+  method = c("YeoJohnson"))
+
+yjTransformed
+
+player_stats_seasonal$fantasyPoints_transformed <- predict(
+  yjTransformed,
+  newdata = player_stats_seasonal["fantasyPoints"])$fantasyPoints
+```
+
 The histogram of the transformed variable is in @fig-histogramFantasyPointsTransformed.
@@
 
-hist(player_stats_seasonal$fantasyPoints[which(player_stats_seasonal$position == "WR")])
+hist(player_stats_seasonal$fantasyPoints_transformed[which(player_stats_seasonal$position == "WR")])
 ```
@@
 linearRegressionModel_outcomeTransformed <- lm(
-  fantasyPoints ~ age + height + weight + I(log(target_share + 1)),
+  fantasyPoints_transformed ~ age + height + weight + I(log(target_share + 1)),
   data = player_stats_seasonal %>% filter(position %in% c("WR")),
@@
 
+The residuals show more constant variance after transforming the outcome variable.
+
 #### Uncorrelated Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsUncorrelatedResiduals}
 
+To determine if residuals are correlated given the nested structure of the data, we can examine the proportion of variance that is attributable to the particular player.
+To do this, we can estimate the intraclass correlation coefficient (ICC) from a [mixed model](#sec-mixedModels).
+
+```{r}
+mixedModel <- lmer(
+  fantasyPoints_transformed ~ 1 + (1 | player_id),
+  data = player_stats_seasonal)
+
+performance::icc(mixedModel)
+```
+
+The ICC indicates that over half of the variance is attribute to between-player variance, so it would be important to account for the player-specific variance using a [mixed model](#sec-mixedModels).
+For simplicity, we focus on multiple regression models in this chapter; mixed models are described in @sec-mixedModels.
+
 #### Normally Distributed Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsNormallyDistributedResiduals}
+
+We can examine whether residuals are normally distributed using quantile–quantile (QQ) plots and probability–probability (PP) plots, as in Figures [-@fig-qqPlot] and [-@fig-ppPlot].
+If the residuals are normally distributed, they should stay close to the diagonal reference line.
+
+```{r}
+#| label: fig-qqPlot
+#| fig-cap: "Residual Plots After Transformation of Fantasy Points."
+#| fig-alt: "Residual Plots After Transformation of Fantasy Points."
+
+car::qqPlot(
+  linearRegressionModel_outcomeTransformed,
+  main = "QQ Plot",
+  id = TRUE)
+```
+
+```{r}
+#| label: fig-ppPlot
+#| fig-cap: "Residual Plots After Transformation of Fantasy Points."
+#| fig-alt: "Residual Plots After Transformation of Fantasy Points."
+
+petersenlab::ppPlot(
+  linearRegressionModel_outcomeTransformed)
+```
 
@@
 So, it is important to examine a correlation matrix of the [predictor variables](#sec-correlationalStudy) before putting them in the same regression model.
-You can also examine indices such as variance inflation factor (VIF).
+You can also examine indices such as variance inflation factor (VIF), where a value greater than 5 or 10 indicates multicollinearity.
+
+Here is the VIF from our earlier model:
+
+```{r}
+car::vif(linearRegressionModel_outcomeTransformed)
+```
 
@@
 
+```{r}
+player_stats_seasonal$target_share_log <- log(player_stats_seasonal$target_share + 1)
+
+modelData <- player_stats_seasonal %>% 
+  filter(position %in% c("WR")) %>% 
+  select(fantasyPoints_transformed, age, height, weight, target_share_log)
+
+numObs <- sum(complete.cases(modelData))
+varMeans <- colMeans(modelData, na.rm = TRUE)
+varCovariances <- cov(modelData, use = "pairwise.complete.obs")
+
+pairwiseRegression_syntax <- '
+  fantasyPoints_transformed ~ age + height + weight + target_share_log
+  fantasyPoints_transformed ~~ fantasyPoints_transformed
+  fantasyPoints_transformed ~ 1
+'
+
+pairwiseRegression_fit <- lavaan(
+  pairwiseRegression_syntax,
+  sample.mean = varMeans,
+  sample.cov = varCovariances,
+  sample.nobs = numObs
+)
+
+summary(
+  pairwiseRegression_fit,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
 ### Multiple Imputation {#sec-multipleRegressionMultipleImputation}
 
+```{r}
+numImputations <- 5
+
+dataToImpute <- player_stats_seasonal %>% 
+  filter(position %in% c("WR")) %>% 
+  select(player_id, where(is.numeric)) %>% 
+  select(
+    player_id:games, carries:wopr, fantasy_points, fantasy_points_ppr,
+    rush_40_yds, rec_40_yds, fumbles, two_pts, return_yds,
+    rush_100_yds:draftround, height:target_share_log) %>% 
+  select(-c(fantasy_points_ppr, ageCentered20, target_share_log)) # drop collinear variables
+
+predictors <- c("targets","receiving_yards","receiving_air_yards","receiving_yards_after_catch","receiving_first_downs","racr")
+
+dataToImpute$player_id_integer <- as.integer(as.factor(dataToImpute$player_id))
+
+varsToImpute <- c("age","height","weight","target_share")
+Y <- varsToImpute
+```
+
+Now, let's specify the imputation method:
+
+```{r}
+meth <- make.method(dataToImpute)
+meth[1:length(meth)] <- ""
+meth[Y] <- "2l.pmm" # specify the imputation method here; this can differ by outcome variable
+```
+
+Now, let's specify the prediction matrix.
+A predictor matrix is a matrix of values, where:
+
+- columns with non-zero values are predictors of the variable specified in the given row
+- the diagonal of the predictor matrix should be zero because a variable cannot predict itself
+
+The values are:
+
+- NOT a predictor of the outcome: `0`
+- cluster variable: `-2`
+- fixed effect of predictor: `1`
+- fixed effect and random effect of predictor: `2`
+- include cluster mean of predictor in addition to fixed effect of predictor: `3`
+- include cluster mean of predictor in addition to fixed effect and random effect of predictor: `4`
+
+```{r}
+pred <- make.predictorMatrix(dataToImpute)
+pred[1:nrow(pred), 1:ncol(pred)] <- 0
+pred[Y, "player_id_integer"] <- (-2) # cluster variable
+pred[Y, predictors] <- 1 # fixed effect predictors
+pred[Y, "age"] <- 2 # random effect predictor
+pred[Y, Y] <- 1 # fixed effect predictor
+
+diag(pred) <- 0
+pred
+```
+
+Now, let's run the imputation:
+
+```{r}
+imp <- mice(
+  as.data.frame(dataToImpute),
+  method = meth,
+  predictorMatrix = pred,
+  m = numImputations,
+  maxit = 5, # generally use 100 maximum iterations; this example uses 5 for speed
+  seed = 52242)
+```
+
+Below are some imputation diagnostics.
+Trace plots are in @fig-miTracePlots.
+
+```{r}
+#| label: fig-miTracePlots
+#| fig-cap: "Trace plots from multiple imputation."
+#| fig-alt: "Trace plots from multiple imputation."
+
+plot(imp, c("target_share"))
+```
+
+A density plot is in @fig-miDensityPlot.
+
+```{r}
+#| label: fig-miDensityPlot
+#| fig-cap: "Density plot from multiple imputation."
+#| fig-alt: "Density plot from multiple imputation."
+
+densityplot(imp, ~ target_share)
+```
+
+The imputated data does not match well the distribution of the observed data.
+Thus, it may be necessary to select a different imputation method for more accurate imputation.
+
+Now, let's do some post-processing:
+
+```{r}
+imp_long <- complete(
+  imp,
+  action = "long",
+  include = TRUE)
+
+imp_long$target_share_log <- log(imp_long$target_share + 1)
+
+imp_long$fantasyPoints_transformed <- predict(
+  yjTransformed,
+  newdata = imp_long["fantasyPoints"])$fantasyPoints
+
+imp_mids <- as.mids(imp_long)
+```
+
+Now let's estimate multiple regression with the multiply imputed data:
+
+```{r}
+imp_regression <- with(
+  imp_mids,
+  lm(
+    fantasyPoints_transformed ~ age + height + weight + target_share_log)
+  )
+
+summary(pool(imp_regression))
+```
+
 ### Full Information Maximum Likelihood {#sec-multipleRegressionFIML}
+
+```{r}
+fimlRegression_syntax <- '
+  fantasyPoints_transformed ~ age + height + weight + target_share_log
+  fantasyPoints_transformed ~~ fantasyPoints_transformed
+  fantasyPoints_transformed ~ 1
+'
+
+fimlRegression_fit <- sem(
+  fimlRegression_syntax,
+  data = player_stats_seasonal,
+  missing = "ML",
+  fixed.x = FALSE
+)
+
+summary(
+  fimlRegression_fit,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
 
@@
 rms::robcov(rms::ols(
-  fantasyPoints ~ height,
+  fantasyPoints_transformed ~ age + height + weight + target_share_log,
   data = player_stats_seasonal_subset,
@@
 
+```{r}
+bayesianMultipleRegressionModel <- brm(
+  formula = fantasyPoints_transformed ~ age + height + weight + I(log(target_share + 1)),
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  family = gaussian()
+)
+
+summary(bayesianMultipleRegressionModel)
+```
+
 ## Conclusion {#sec-multipleRegressionConclusion}

### 2025-04-11 index.qmd
20250411 - add packages

--- 2025-04-08 index.qmd
@@
   "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin",
-  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot" #"mgcv",
+  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation", 
+  "mgcv", "rms", "car", "performance", "lavaan", "mice", "miceadds"
 )

### 2025-04-11 multiple-regression.qmd
20250411 - add packages

@@
 
-We can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), and added-variable plots (@fig-addedVariablePlots), to evaluate for potential nonlinearity.
+We can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), and added-variable plots (@fig-addedVariablePlots) from the `car` package [@Fox2019; @R-car], to evaluate for potential nonlinearity.
 For evaluating linearity, we would expect minimal bend/curvature in the lines.
@@
 To determine if residuals are correlated given the nested structure of the data, we can examine the proportion of variance that is attributable to the particular player.
-To do this, we can estimate the intraclass correlation coefficient (ICC) from a [mixed model](#sec-mixedModels).
+To do this, we can estimate the intraclass correlation coefficient (ICC) from a [mixed model](#sec-mixedModels) using the `performance` package [@Ludecke2021; @R-performance].
 
@@
 
+We can estimate a regression model that uses pairwise deletion using the `lavaan` package [@Rosseel2012; @R-lavaan].
+
 ```{r}
@@
 
-pairwiseRegression_fit <- lavaan(
+pairwiseRegression_fit <- lavaan::lavaan(
   pairwiseRegression_syntax,
@@
 ### Multiple Imputation {#sec-multipleRegressionMultipleImputation}
+
+We can multiply impute data using the `mice` package [@vanBuuren2011; @R-mice].
 
@@
 
-Now, let's specify the imputation method:
+Now, let's specify the imputation method—we use the two-level predictive mean matching (`2l.pmm`) method from the `miceadds` package [@R-miceadds] to account for the nonindependent data (owing to multiple seasons per player):
 
@@
 ```{r}
-imp <- mice(
+imp <- mice::mice(
   as.data.frame(dataToImpute),
@@
 
+We can estimate a regression model that uses full information maximum likelihood using the `lavaan` package [@Rosseel2012; @R-lavaan].
+
 ```{r}
@@
 In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence.
-Below, we demonstrate how to account for non-independence of observations using cluster-robust standard errors.
+Below, we demonstrate how to account for non-independence of observations using cluster-robust standard errors with the `rms` package [@R-rms].
 

### 2025-04-11 download-football-data.qmd
20250411 - add packages
20250411 - update data to remove infinite (implausible) values

@@
 
+#### Season-by-Season Statistics {#sec-downloadActualSeasonalStats}
+
+We calculate historical actual season-by-season statistics in @sec-calculateActualSeasonalStats.
+
 #### Career Statistics {#sec-downloadActualCareerStats}
 
-#### Season-by-Season Statistics {#sec-downloadActualSeasonalStats}
+We calculate historical actual career statistics in @sec-calculateActualCareerStats.
 
@@
 nfl_actualStats_career_player_inclPost$gwfg_distance_list[which(nfl_actualStats_career_player_inclPost$gwfg_distance_list == "")] <- NA
+
+nfl_actualStats_career_player[sapply(nfl_actualStats_career_player, is.infinite)] <- NA
+nfl_actualStats_career_player_inclPost[sapply(nfl_actualStats_career_player_inclPost, is.infinite)] <- NA
 ```
@@
 
+Let's remove infinite (implausible) values:
+
+```{r}
+nfl_actualFantasyPoints_player_weekly_raw[sapply(nfl_actualFantasyPoints_player_weekly_raw, is.infinite)] <- NA
+nfl_actualFantasyPoints_dst_weekly_raw[sapply(nfl_actualFantasyPoints_dst_weekly_raw, is.infinite)] <- NA
+```
+
 ```{r}
@@
 
+Let's remove infinite (implausible) values:
+
+```{r}
+nfl_actualFantasyPoints_player_seasonal[sapply(nfl_actualFantasyPoints_player_seasonal, is.infinite)] <- NA
+nfl_actualFantasyPoints_dst_seasonal[sapply(nfl_actualFantasyPoints_dst_seasonal, is.infinite)] <- NA
+```
+
 ```{r}
@@
 ```
+
+Let's rearrange the data:
 

### 2025-04-11 index.qmd
20250411 - add packages

--- 2025-04-08 index.qmd
@@
   "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin",
-  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot" #"mgcv",
+  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation", 
+  "mgcv", "rms", "car", "performance", "lavaan", "mice", "miceadds","interactions"
 )

### 2025-04-11 multiple-regression.qmd
20250411 - moderated multiple regression

@@
 library("miceadds")
+library("interactions")
 library("brms")
@@
 diag(pred) <- 0
-pred
 ```
@@
 
+First, we mean-center the predictors.
+In this case, we center the predictors around the mean of height and weight for Wide Receivers:
+
+```{r}
+player_stats_seasonal$height_centered <- player_stats_seasonal$height - mean(player_stats_seasonal$height[which(player_stats_seasonal$position == "WR")], na.rm = TRUE)
+player_stats_seasonal$weight_centered <- player_stats_seasonal$weight - mean(player_stats_seasonal$weight[which(player_stats_seasonal$position == "WR")], na.rm = TRUE)
+```
+
+Then, we compute the interaction term as the multiplication of the two centered predictors:
+
+```{r}
+player_stats_seasonal$heightXweight <- player_stats_seasonal$height_centered * player_stats_seasonal$weight_centered
+```
+
+Then, we fit the moderated multiple regression model:
+
+```{r}
+moderationModel <- lm(
+  fantasyPoints_transformed ~ height_centered + weight_centered + height_centered:weight_centered,
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  na.action = "na.exclude"
+)
+
+summary(moderationModel)
+```
+
+This model is equivalent to the model that includes the interaction term explicitly:
+
+```{r}
+#| eval: false
+
+moderationModel <- lm(
+  fantasyPoints_transformed ~ height_centered + weight_centered + heightXweight,
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  na.action = "na.exclude"
+)
+
+summary(moderationModel)
+```
+
+Now, we can visualize the interaction to understand it.
+We create an interaction plot (@fig-interactionPlot) and Johnson-Neyman plot (@fig-johnsonNeymanPlot) using the `interactions` package [@R-interactions].
+
+```{r}
+#| label: fig-interactionPlot
+#| fig-cap: "Interaction Plot from Moderated Multiple Regression."
+#| fig-alt: "Interaction Plot from Moderated Multiple Regression."
+
+interactions::interact_plot(
+  moderationModel,
+  pred = height_centered,
+  modx = weight_centered)
+```
+
+```{r}
+#| label: fig-johnsonNeymanPlot
+#| fig-cap: "Johnson-Neyman Plot from Moderated Multiple Regression."
+#| fig-alt: "Johnson-Neyman Plot from Moderated Multiple Regression."
+
+interactions::johnson_neyman(
+  moderationModel,
+  pred = height_centered,
+  modx = weight_centered,
+  alpha = .05)
+```
+
+Here is a simple slopes analysis:
+
+```{r}
+interactions::sim_slopes(
+  moderationModel,
+  pred = height_centered,
+  modx = weight_centered,
+  johnson_neyman = FALSE)
+```
+
 ## Mediation {#sec-multipleRegressionMediation}

### 2025-04-11 correlation.qmd
20250411 - clean up

@@
 #| code-fold: true
-#| message: false
 

### 2025-04-11 multiple-regression.qmd
20250411 - clean up

@@
 ```{r}
+#| output: false
+
 imp <- mice::mice(

### 2025-04-11 multiple-regression.qmd
20250411 - clean up

@@
 ```{r}
+#| output: false
+
 bayesianMultipleRegressionModel <- brm(

### 2025-04-12 multiple-regression.qmd
20250411 - clean up

@@
 
-Here is the VIF from our earlier model:
+Here are the VIFs from our earlier model:
 
@@
 )
-
+```
+
+```{r}
 summary(bayesianMultipleRegressionModel)

### 2025-04-13 basic-statistics.qmd
20250413 - distribution visualizer

@@
 ```
+
+::: {.content-visible when-format="html"}
+
+You can use the following web application to visualize various distributions according to a specified [center](#sec-descriptiveStatisticsCenter) (mean), [spread](#sec-descriptiveStatisticsSpread) (standard deviation), and [shape](#sec-descriptiveStatisticsShape) (skewness and kurtosis):
+
+```{=html}
+<iframe id="distributionVisualizer" src="https://ffashiny.shinyapps.io/distributionvisualizer/" style="border: none; width: 100%; height: 100%" frameborder="0"></iframe>
+```
+
+:::
 

### 2025-04-13 basic-statistics.qmd
20250413 - distribution visualizer

--- 2025-04-05 basic-statistics.qmd
@@
 ```
+
+::: {.content-visible when-format="html"}
+
+You can use the following web application to visualize various distributions according to a specified [center](#sec-descriptiveStatisticsCenter) (mean), [spread](#sec-descriptiveStatisticsSpread) (standard deviation), and [shape](#sec-descriptiveStatisticsShape) (skewness and kurtosis).
+Try varying the standard deviation, skewness, and kurtosis to see the impact of each.
+
+```{=html}
+<iframe id="distributionVisualizer" src="https://ffashiny.shinyapps.io/distributionvisualizer/" style="border: none; width: 100%; height: 720px" frameborder="0"></iframe>
+```
+
+:::
 

### 2025-04-13 basic-statistics.qmd
20250413 - update frame size

@@
 ```{=html}
-<iframe id="distributionVisualizer" src="https://ffashiny.shinyapps.io/distributionvisualizer/" style="border: none; width: 100%; height: 720px" frameborder="0"></iframe>
+<iframe id="distributionVisualizer" src="https://ffashiny.shinyapps.io/distributionvisualizer/" style="border: none; width: 100%; height: 1000px" frameborder="0"></iframe>
 ```

### 2025-04-13 basic-statistics.qmd
20250413 - add header

@@
 
-You can use the following web application to visualize various distributions according to a specified [center](#sec-descriptiveStatisticsCenter) (mean), [spread](#sec-descriptiveStatisticsSpread) (standard deviation), and [shape](#sec-descriptiveStatisticsShape) (skewness and kurtosis).
+### Interactive Visualization {#sec-descriptiveStatisticsVisualization}
+
+You can use the following interactive graph to visualize various distributions according to a specified [center](#sec-descriptiveStatisticsCenter) (mean), [spread](#sec-descriptiveStatisticsSpread) (standard deviation), and [shape](#sec-descriptiveStatisticsShape) (skewness and kurtosis).
 Try varying the standard deviation, skewness, and kurtosis to see the impact of each.

### 2025-04-13 basic-statistics.qmd
20250413 - fixes

@@
 A mesokurtic distribution (with a kurtosis value near zero) reflects a normal amount of tailedness.
-Positive kurtosis values reflect a leptokurtic distribution, where there are lighter tails and a sharper peak than a normal distribution.
-Negative kurtosis values reflect a platykurtic distribution, where there are heavier tails and a flatter peak than a normal distribution.
+Positive kurtosis values reflect a leptokurtic distribution, where there are heavier tails and a sharper peak than a normal distribution, indicating more frequent extreme values.
+Negative kurtosis values reflect a platykurtic distribution, where there are lighter tails and a flatter peak than a normal distribution, indicating fewer extreme values.
 Fantasy points tend to have a leptokurtic distribution.

### 2025-04-14 index.qmd
20250414 - add packages and DOIs

--- 2025-04-09 index.qmd
@@
   "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin",
-  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation" #"mgcv",
+  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation", 
+  "mgcv", "rms", "car", "performance", "lavaan", "mice", "miceadds","interactions",
+  "shiny","PearsonDS"
 )

### 2025-04-15 evaluating-prediction-accuracy.qmd
20250415 - R-squared and cross-refs

@@
 
-##### Mean Error {#sec-meanError}
+##### Mean Error (ME) {#sec-meanError}
 
@@
 
-The [coefficient of determination](#sec-multipleRegressionRSquared) ($R^2$) reflects the proportion of variance in the outcome (dependent) variable that is explained by the model predictions: $R^2 = \frac{\text{variance explained in }Y}{\text{total variance in }Y}$.
+The [coefficient of determination](#sec-multipleRegressionRSquared) ($R^2$) is described in @sec-multipleRegressionRSquared.
+The [coefficient of determination](#sec-multipleRegressionRSquared) reflects the proportion of variance in the outcome (dependent) variable that is explained by the model predictions: $R^2 = \frac{\text{variance explained in }Y}{\text{total variance in }Y}$.
 Larger values indicate greater accuracy.
@@
 
+Adjusted $R^2$ ($R^2_{adj}$) is described in @sec-multipleRegressionAdjustedRsquared.
 Adjusted $R^2$ is similar to the [coefficient of determination](#sec-coefficientOfDetermination), but it accounts for the number of predictors included in the regression model to penalize [overfitting](#sec-overfitting).
@@
 The formula for adjusted $R^2$ is in @eq-adjustedRSquared.
-Adjusted $R^2$ is described further in @sec-multipleRegressionRSquared.
 

### 2025-04-15 multiple-regression.qmd
20250415 - R-squared and cross-refs

@@
 
-Regression with multiple predictors—i.e., multiple regression—takes the form of @eq-multipleRegression:
+Regression with multiple [predictors](#sec-correlationalStudy)—i.e., multiple regression—takes the form of @eq-multipleRegression:
 
@@
 
-- there is a linear association between the predictor variables and the outcome variable
-- there is homoscedasticity of the residuals; the residuals do not differ as a function of the predictor variables or as a function of the outcome variable
+- there is a linear association between the [predictor variable](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy)
+- there is homoscedasticity of the residuals; the residuals do not differ as a function of the [predictor variable](#sec-correlationalStudy) or as a function of the [outcome variable](#sec-correlationalStudy)
 - the residuals are independent; they are uncorrelated with each other
@@
 
-Homoscedasticity of the residuals means that the variance of the residuals does not differ as a function of the outcome variable or as a function of the predictor variables (i.e., the residuals show constant variance as a function of outcome/predictors).
+Homoscedasticity of the residuals means that the variance of the residuals does not differ as a function of the [outcome variable](#sec-correlationalStudy) or as a function of the [predictor variable](#sec-correlationalStudy) (i.e., the residuals show constant variance as a function of [outcome](#sec-correlationalStudy)/[predictors](#sec-correlationalStudy)).
+
+Those are some of the key assumptions of multiple regression.
+However, there are additional assumptions of multiple regression, including ones discussed in the chapter on [Causal Inference](#sec-causalInference).
+For instance, the variables included should reflect causal process such that the [predictor variables](#sec-correlationalStudy) influence the [outcome variable](#sec-correlationalStudy), and not the other way around.
+That is, the [outcome variable](#sec-correlationalStudy) should not influence the [predictor variables](#sec-correlationalStudy) (i.e., there should be no reverse causation).
+In addition, it is important control for any [confound(s)](#sec-causalDiagramConfounding).
+If a [confound](#sec-causalDiagramConfounding) is not controlled for, this is called omitted-variable bias, and it leads the researcher to incorrectly attribute the effects of the omitted variable to the included variables.
 
@@
 
-To evaluate the shape of the association between the predictor variables and the outcome variable, we can examine scatterplots, residual plots, marginal model plots, and added-variable plots.
+To evaluate the shape of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy), we can examine scatterplots, residual plots, marginal model plots, and added-variable plots.
 
@@
 
-## Coefficient of Determination ($R^2$) {#sec-multipleRegressionRSquared}
+## How Much Variance the Model Explains {#sec-multipleRegressionExplainedVariance}
+
+When estimating a multiple regression model, it can be useful to evaluate how much variance in the [outcome variable](#sec-correlationalStudy) that the [predictor variable(s)](#sec-correlationalStudy) explain (i.e., account for).
+If the variables collectively explain only a small amount of variance, it suggest that the [predictor variables](#sec-correlationalStudy) have a small [effect size](#sec-practicalSignificance) and that other [predictor variables](#sec-correlationalStudy) will be necessary to account for the majority of variability in the [outcome variable](#sec-correlationalStudy).
+There are two primary indices of how much how much variance in the [outcome variable](#sec-correlationalStudy) is explained by the [predictor variable(s)](#sec-correlationalStudy): the [coefficient of determination](#sec-multipleRegressionRSquared) ($R^2$) and [adjusted $R^2$](#sec-multipleRegressionAdjustedRsquared) ($R^2_{adj}$), described below.
+
+### Coefficient of Determination ($R^2$) {#sec-multipleRegressionRSquared}
 
@@
 
-One issue with $R^2$ is that it increases as the number of predictors increases, which can lead to [overfitting](#sec-overfitting) if using $R^2$ as an index to compare models for purposes of selecting the "best-fitting" model.
+One issue with $R^2$ is that it increases as the number of [predictors](#sec-correlationalStudy) increases, which can lead to [overfitting](#sec-overfitting) if using $R^2$ as an index to compare models for purposes of selecting the "best-fitting" model.
 Consider the following example (adapted from @PetersenPrinciplesPsychAssessment) in which you have one [predictor variable](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy), as shown in @tbl-regression1.
@@
 In this example, the $R^2$ is `{r} regression1_rsquare`.
-The equation is not a perfect prediction, but with a single [predictor variable](#sec-correlationalStudy), it captures the majority of the variance in the outcome.
+The equation is not a perfect prediction, but with a single [predictor variable](#sec-correlationalStudy), it captures the majority of the variance in the [outcome](#sec-correlationalStudy).
 
@@
 In this example, the $R^2$ is `{r} regression2_rsquare`.
-The equation with the second [predictor variable](#sec-correlationalStudy) provides a perfect prediction of the outcome.
+The equation with the second [predictor variable](#sec-correlationalStudy) provides a perfect prediction of the [outcome](#sec-correlationalStudy).
 
@@
 Assume that, with 13 [predictor variables](#sec-correlationalStudy), we explain 38% of the variance ($R^2 = .38; r = .62$).
-We explained a lot of the variance in the outcome, but it is important to consider how much variance could have been explained by random chance: $E(R^2) = \frac{K}{n-1} = \frac{13}{43 - 1} = .31$.
-We expect to explain 31% of the variance, by chance, in the outcome.
+We explained a lot of the variance in the [outcome](#sec-correlationalStudy), but it is important to consider how much variance could have been explained by random chance: $E(R^2) = \frac{K}{n-1} = \frac{13}{43 - 1} = .31$.
+We expect to explain 31% of the variance, by chance, in the [outcome](#sec-correlationalStudy).
 So, 82% of the variance explained was likely spurious (i.e., $\frac{.31}{.38} = .82$).
 As the sample size increases, the spuriousness decreases.
-
-To account for the number of [predictor variables](#sec-correlationalStudy) in the model, we can use a modified version of $R^2$ called adjusted $R^2$ ($R^2_{adj}$).
+To account for the number of [predictor variables](#sec-correlationalStudy) in the model, we can use a modified version of $R^2$ called adjusted $R^2$ ($R^2_{adj}$), described next.
+
+### Adjusted $R^2$ ($R^2_{adj}$) {#sec-multipleRegressionAdjustedRsquared}
+
 Adjusted $R^2$ ($R^2_{adj}$) accounts for the number of [predictor variables](#sec-correlationalStudy) in the model, based on how much would be expected to be accounted for by chance to penalize [overfitting](#sec-overfitting).
@@
 
+### Predicting Wide Receivers' Fantasy Points {#sec-multipleRegressionFantasyPointsWRs}
+
 Let's say we want to use a number of variables to predict a wide receiver's fantasy performance.
-We want to consider several predictors, including the player's age, height, weight, and target share.
-We have only a few predictors and our sample size is large enough such that [overfitting](#sec-overfitting) is not likely a concern.
+We want to consider several [predictors](#sec-correlationalStudy), including the player's age, height, weight, and target share.
+We have only a few [predictors](#sec-correlationalStudy) and our sample size is large enough such that [overfitting](#sec-overfitting) is not likely a concern.
+
+#### Examine Bivariate Associations {#sec-multipleRegressionFantasyPointsWRsBivariate}
 
@@
 In addition, target share shows a weakening association as target share increases.
-Thus, after evaluating the linear association between the predictors and outcome, we will also examine the possibility for curvilinear associations.
+Thus, after evaluating the linear association between the [predictors](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy), we will also examine the possibility for curvilinear associations.
+
+#### Estimate Multiple Regression Model {#sec-multipleRegressionFantasyPointsWRsModel}
 
@@
 The only terms that are significantly associated with fantasy performance among Wide Receivers are weight and target share, both of which showed a positive association with fantasy points.
-If we want to obtain standardized regression coefficients, we can standardize the outcome variable and each predictor variable using the `scale()` function:
+
+We can obtain the [coefficient of determination](#sec-multipleRegressionRSquared) ($R^2$) and [adjusted $R^2$](#sec-multipleRegressionAdjustedRsquared) ($R^2_{adj}$) using the following code:
+
+```{r}
+summary(linearRegressionModel)$r.squared
+summary(linearRegressionModel)$adj.r.squared
+```
+
+```{r}
+#| include: false
+
+linearRegressionModel_rsquared <- summary(linearRegressionModel)$r.squared
+```
+
+The model explained `r petersenlab::apa(linearRegressionModel_rsquared * 100, decimals = 0)`% of the variability in fantasy points (i.e., $R^2 = `r petersenlab::apa(linearRegressionModel_rsquared, decimals = 2, leading = FALSE)`$).
+
+If we want to obtain standardized regression coefficients, we can standardize the [outcome variable](#sec-correlationalStudy) and each [predictor variable](#sec-correlationalStudy) using the `scale()` function:
 
@@
 Target share has a large [effect size](#sec-practicalSignificance).
-All of the other predictors have small effect sizes.
+All of the other [predictors](#sec-correlationalStudy) have small [effect size](#sec-practicalSignificance).
 
@@
 
-### Evaluating and Addressing Assumptions {#sec-multipleRegressionExamplesAddressingAssumptions}
+#### Evaluating and Addressing Assumptions {#sec-multipleRegressionExamplesAddressingAssumptions}
 
@@
 
-- there is a linear association between the predictor variables and the outcome variable
-- there is homoscedasticity of the residuals; the residuals do not differ as a function of the predictor variables or as a function of the outcome variable
+- there is a linear association between the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy)
+- there is homoscedasticity of the residuals; the residuals do not differ as a function of the [predictor variables](#sec-correlationalStudy) or as a function of the [outcome variable](#sec-correlationalStudy)
 - the residuals are independent; they are uncorrelated with each other
@@
 
-#### Linear Association {#sec-multipleRegressionExamplesAddressingAssumptionsLinearAssociation}
-
-We evaluated the shape of the association between the predictor variables and the outcome variables using scatterplots.
+##### Linear Association {#sec-multipleRegressionExamplesAddressingAssumptionsLinearAssociation}
+
+We evaluated the shape of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variables](#sec-correlationalStudy) using scatterplots.
 We accounted for potential curvilinearity with a quadratic term.
-Other ways to account for nonlinearity, in addition to polynomials (e.g., quadratic, cubic, quartic, or higher-degree terms), include transforming predictors (e.g., log, square root, inverse, exponential transform), splines/piecewise regression, and generalized additive models.
+Other ways to account for nonlinearity, in addition to polynomials (e.g., quadratic, cubic, quartic, or higher-degree terms), include transforming [predictors](#sec-correlationalStudy) (e.g., log, square root, inverse, exponential transform), splines/piecewise regression, and generalized additive models.
 
@@
 
-#### Homoscedasticity {#sec-multipleRegressionExamplesAddressingAssumptionsHomoscedasticity}
+##### Homoscedasticity {#sec-multipleRegressionExamplesAddressingAssumptionsHomoscedasticity}
 
@@
 In this example, the residuals appear to increase as a function of the fitted values.
-To handle this, we may need to transform the outcome variable to be more normally distributed.
+To handle this, we may need to transform the [outcome variable](#sec-correlationalStudy) to be more normally distributed.
 
@@
 
-The residuals show more constant variance after transforming the outcome variable.
-
-#### Uncorrelated Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsUncorrelatedResiduals}
+The residuals show more constant variance after transforming the [outcome variable](#sec-correlationalStudy).
+
+##### Uncorrelated Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsUncorrelatedResiduals}
 
@@
 
-#### Normally Distributed Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsNormallyDistributedResiduals}
+##### Normally Distributed Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsNormallyDistributedResiduals}
 
@@
 
-To address multicollinearity, you can drop a redundant predictor or you can also use principal component analysis or factor analysis of the predictors to reduce the predictors down to a smaller number of meaningful predictors.
-For a meaningful answer in a regression framework that is precise and confident, you need a low level of intercorrelation among predictors, unless you have a very large sample size.
+To address multicollinearity, you can drop a redundant predictor or you can also use [principal component analysis](#sec-pca) or [factor analysis](#sec-factorAnalysis) of the [predictors](#sec-correlationalStudy) to reduce the [predictors](#sec-correlationalStudy) down to a smaller number of meaningful [predictors](#sec-correlationalStudy).
+For a meaningful answer in a regression framework that is precise and confident, you need a low level of intercorrelation among [predictors](#sec-correlationalStudy), unless you have a very large sample size.
 
@@
 
-- columns with non-zero values are predictors of the variable specified in the given row
+- columns with non-zero values are [predictors](#sec-correlationalStudy) of the variable specified in the given row
 - the diagonal of the predictor matrix should be zero because a variable cannot predict itself
@@
 
-- When computing the interaction term, first mean-center the predictor variables.
-Calculate the interaction term as the multiplication of the mean-centered predictor variables.
-Mean-centering the predictor variables when computing the interaction term is important for addressing issues regarding [multicollinearity](#sec-multipleRegressionMulticollinearity) [@Iacobucci2016].
+- When computing the interaction term, first mean-center the [predictor variables](#sec-correlationalStudy).
+Calculate the interaction term as the multiplication of the mean-centered [predictor variables](#sec-correlationalStudy).
+Mean-centering the [predictor variables](#sec-correlationalStudy) when computing the interaction term is important for addressing issues regarding [multicollinearity](#sec-multipleRegressionMulticollinearity) [@Iacobucci2016].
 - When including an interaction term in the model, make sure also to include the main effects.
 
-First, we mean-center the predictors.
-In this case, we center the predictors around the mean of height and weight for Wide Receivers:
+First, we mean-center the [predictors](#sec-correlationalStudy).
+In this case, we center the [predictors](#sec-correlationalStudy) around the mean of height and weight for Wide Receivers:
 
@@
 
-Then, we compute the interaction term as the multiplication of the two centered predictors:
+Then, we compute the interaction term as the multiplication of the two centered [predictors](#sec-correlationalStudy):
 
@@
 Multiple regression allows examining the association between multiple [predictor variables](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy).
-Inclusion of multiple predictors in the model allows for potentially greater predictive accuracy and identification of the extent to which each variable uniquely contributes to the outcome variable.
+Inclusion of multiple [predictors](#sec-correlationalStudy) in the model allows for potentially greater predictive accuracy and identification of the extent to which each variable uniquely contributes to the [outcome variable](#sec-correlationalStudy).
 As with [correlation](#sec-correlation), an association does not imply causation.

### 2025-04-15 causal-inference.qmd
20250409 - minor tweaks
20250415 - reorganization

@@
 
-### Causation
+## Causation {#sec-causation}
 
@@
 Establishing the first criterion eliminates the possibility that the association between the purported cause and effect reflects reverse causation.
-Reverse causation occurs when the purported effect is actually the cause of the purported cause, rather than the other way around.
+*Reverse causation* occurs when the purported effect is actually the cause of the purported cause, rather than the other way around.
 For instance, if sports drink consumption occurs only once, and it occurs only before and not after performance, then we have ruled out the possibility of reverse causation (i.e., that better performance causes players to consume sports drink).
@@
 For instance, from a DAG, it is possible to determine what variables it is important to control for (and how)—and just as importantly, what not to control for—in order to get unbiased estimates of the effect of one variable on another variable or of the association between two variables of interest.
-To create DAGs, you can use the `R` package `dagitty` [@R-daggity; @Textor2016_packages] or the associated browser-based extension, DAGitty: <https://dagitty.net> (archived at <https://perma.cc/U9BY-VZE2>).
+To create DAGs, you can use the `R` package `dagitty` [@R-dagitty; @Textor2016_packages] or the associated browser-based extension, DAGitty: <https://dagitty.net> (archived at <https://perma.cc/U9BY-VZE2>).
 Examples of various causal diagrams that could explain why `X` is associated with `Y` are in Figures [-@fig-XCausesY], [-@fig-YCausesX] and [-@fig-ZCausesXandY].

### 2025-04-15 correlation.qmd
20250415 - reorganization

@@
 library("correlation")
+library("ggridges")
 library("tidyverse")
@@
 ## Examples {#sec-correlationExamples}
+
+### Player Height and Fantasy Points {#sec-correlationHeightFantasyPoints}
 
@@
 
-### Covariance {#sec-correlationExamplesCovariance}
+#### Covariance {#sec-correlationExamplesCovariance}
 
@@
 
-### Pearson Correlation {#sec-correlationExamplesPearson}
+#### Pearson Correlation {#sec-correlationExamplesPearson}
 
@@
 
-### Scatterplot with Best-Fit Line
+#### Scatterplot with Best-Fit Line
 
@@
 
-### Rank Correlation: Spearman's Rho ($\rho$) {#sec-correlationExamplesSpearman}
+#### Rank Correlation: Spearman's Rho ($\rho$) {#sec-correlationExamplesSpearman}
 
@@
 
-### Partial Correlation {#sec-correlationExamplesPartial}
+#### Partial Correlation {#sec-correlationExamplesPartial}
 
@@
 
-### Nonlinear Correlation {#sec-correlationExamplesNonlinear}
+#### Nonlinear Correlation {#sec-correlationExamplesNonlinear}
 
@@
 
-### Correlation Matrix {#sec-correlationExamplesMatrix}
+#### Correlation Matrix {#sec-correlationExamplesMatrix}
 
@@
 
-### Correlogram {#sec-correlationExamplesCorrelogram}
+#### Correlogram {#sec-correlationExamplesCorrelogram}
 

### 2025-04-15 multiple-regression.qmd
20250415 - reorganization

@@
 In addition, it is important control for any [confound(s)](#sec-causalDiagramConfounding).
-If a [confound](#sec-causalDiagramConfounding) is not controlled for, this is called omitted-variable bias, and it leads the researcher to incorrectly attribute the effects of the omitted variable to the included variables.
+If a [confound](#sec-causalDiagramConfounding) is not controlled for, this is called *omitted-variable bias*, and it leads the researcher to incorrectly attribute the effects of the omitted variable to the included variables.
 
@@
 
+#### Examine Descriptive Statistics {#sec-multipleRegressionFantasyPointsWRsDescriptive}
+
+Let's first examine [descriptive statistics](#sec-descriptiveStatistics) of the predictor and outcome variables.
+
+```{r}
+player_stats_seasonal %>% 
+  dplyr::select(fantasyPoints, age, height, weight, target_share) %>% 
+  dplyr::summarise(across(
+      everything(),
+      .fns = list(
+        n = ~ length(na.omit(.)),
+        missingness = ~ mean(is.na(.)) * 100,
+        M = ~ mean(., na.rm = TRUE),
+        SD = ~ sd(., na.rm = TRUE),
+        min = ~ min(., na.rm = TRUE),
+        max = ~ max(., na.rm = TRUE),
+        range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
+        IQR = ~ IQR(., na.rm = TRUE),
+        MAD = ~ mad(., na.rm = TRUE),
+        median = ~ median(., na.rm = TRUE),
+        pseudomedian = ~ DescTools::HodgesLehmann(., na.rm = TRUE),
+        mode = ~ petersenlab::Mode(., multipleModes = "mean"),
+        skewness = ~ psych::skew(., na.rm = TRUE),
+        kurtosis = ~ psych::kurtosi(., na.rm = TRUE)),
+      .names = "{.col}.{.fn}")) %>%
+    tidyr::pivot_longer(
+      cols = everything(),
+      names_to = c("variable","index"),
+      names_sep = "\\.") %>% 
+    tidyr::pivot_wider(
+      names_from = index,
+      values_from = value)
+```
+
+Let's also examine the distributions of the variables using a [density plot](#sec-ridgelinePlot), as depicted in @fig-densityPlotMultipleRegressionVars.
+
+```{r}
+#| layout: [[100], [50, 50], [50, 50]]
+#| label: fig-densityPlotMultipleRegressionVars
+#| fig-cap: "Density Plot of Model Variables (i.e., Predictor and Outcome Variables)."
+#| fig-alt: "Density Plot of Model Variables (i.e., Predictor and Outcome Variables)."
+#| fig-subcap:
+#|   - "Fantasy Points"
+#|   - "Age"
+#|   - "Height"
+#|   - "Weight"
+#|   - "Target Share"
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("WR")),
+  mapping = aes(
+    x = fantasyPoints)
+) +
+  geom_histogram(
+    aes(y = after_stat(density)),
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  geom_density(
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  geom_rug() +
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Fantasy Points with Histogram and Rug Plot"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("WR")),
+  mapping = aes(
+    x = age)
+) +
+  geom_histogram(
+    aes(y = after_stat(density)),
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  geom_density(
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  geom_rug() +
+  labs(
+    x = "Age (years)",
+    y = "Density",
+    title = "Density Plot of Player Age with Histogram and Rug Plot"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("WR")),
+  mapping = aes(
+    x = height)
+) +
+  geom_histogram(
+    aes(y = after_stat(density)),
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  geom_density(
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  geom_rug() +
+  labs(
+    x = "Fantasy Points",
+    y = "Height (inches)",
+    title = "Density Plot of Player Height with Histogram and Rug Plot"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("WR")),
+  mapping = aes(
+    x = weight)
+) +
+  geom_histogram(
+    aes(y = after_stat(density)),
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  geom_density(
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  geom_rug() +
+  labs(
+    x = "Weight (pounds)",
+    y = "Density",
+    title = "Density Plot of Player Weight with Histogram and Rug Plot"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("WR")),
+  mapping = aes(
+    x = target_share)
+) +
+  geom_histogram(
+    aes(y = after_stat(density)),
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  geom_density(
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  geom_rug() +
+  labs(
+    x = "Target Share",
+    y = "Density",
+    title = "Density Plot of Target Share with Histogram and Rug Plot"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
 #### Examine Bivariate Associations {#sec-multipleRegressionFantasyPointsWRsBivariate}
 
-Let's first examine the bivariate association of each using a scatterplot to evaluate for any potential nonlinearity, as depicted in @fig-scatterplots.
+Then, let's examine the bivariate association of each using a scatterplot to evaluate for any potential nonlinearity, as depicted in @fig-scatterplots.
 
@@
 
-First, let's estimate a multiple regression model with only linear terms:
+Now that we have examined [descriptive statistics](#sec-descriptiveStatistics) and bivariate associations, let's first estimate a multiple regression model with only linear terms:
 
@@
 
+We could also generate the model-implied prediction of fantasy points for any value of the [predictor variables](#sec-correlationalStudy).
+For instance, here are the number of fantasy points expected for a Wide Receiver who is 23 years old, who is 6'2" tall (72 inches), and weighs
+
 #### Evaluating and Addressing Assumptions {#sec-multipleRegressionExamplesAddressingAssumptions}

### 2025-04-15 multiple-regression.qmd
20250415 - mediation

@@
 
+A mediation model takes the following general form in the `lavaan` package [@Rosseel2012; @R-lavaan].
+
+```{r}
+#| eval: false
+
+mediationModel <- '
+  # direct effect (cPrime)
+  Y ~ direct*X
+
+  # mediator
+  M ~ a*X
+  Y ~ b*M
+
+  # indirect effect = a*b
+  indirect := a*b
+
+  # total effect (c)
+  total := direct + indirect
+'
+```
+
+Let's substitute in our [predictor](#sec-correlationalStudy), [outcome](#sec-correlationalStudy), and hypothesized [mediator](#sec-mediation).
+In this case, we predict that receiving touchdowns partially accounts for the association between Wide Receiver's target share and their fantasy points.
+This is a silly example because fantasy points are derived, in part, from touchdowns, so of course touchdowns will partially account for almost any effect on Wide Receivers' fantasy points.
+This example is merely for demonstrating the process of developing and examining a mediation model.
+
+```{r}
+mediationModel <- '
+  # direct effect (cPrime)
+  fantasyPoints_transformed ~ direct*target_share
+
+  # mediator
+  receiving_tds ~ a*target_share
+  fantasyPoints_transformed ~ b*receiving_tds
+
+  # indirect effect = a*b
+  indirect := a*b
+
+  # total effect (c)
+  total := direct + indirect
+'
+```
+
+To get a robust estimate of the indirect effect, we obtain bootstrapped estimates from 1,000 bootstrap draws.
+Typically, we would obtain bootstrapped estimates from 10,000 bootstrap draws, but this example uses only 1,000 bootstrap draws for a shorter runtime.
+
+```{r}
+mediationFit <- sem(
+  mediationModel,
+  data = player_stats_seasonal,
+  se = "bootstrap",
+  bootstrap = 1000, # generally use 10,000 bootstrap draws; this example uses 1,000 for speed
+  parallel = "multicore", # parallelization for speed: use "multicore" for Mac/Linux; "snow" for PC
+  iseed = 52242, # for reproducibility
+  missing = "ML",
+  estimator = "ML",
+  std.lv = TRUE,
+  fixed.x = FALSE)
+```
+
+Here are the model results:
+
+```{r}
+summary(
+  mediationFit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+We can also estimate a model with multiple hypothesized [mediator](#sec-mediations):
+
+```{r}
+multipleMediatorModel <- '
+  # direct effect (cPrime)
+  fantasyPoints_transformed ~ direct*target_share
+
+  # mediator
+  receiving_tds ~ a1*target_share
+  receiving_yards ~ a2*target_share
+
+  fantasyPoints_transformed ~ b1*receiving_tds + b1*receiving_yards
+
+  # indirect effect = a*b
+  indirect1 := a1*b1
+  indirect2 := a2*b2
+  indirectTotal := indirect1 + indirect2
+
+  # total effect (c)
+  total := direct + indirectTotal
+'
+```
+
+```{r}
+multipleMediatorFit <- sem(
+  multipleMediatorModel,
+  data = player_stats_seasonal,
+  se = "bootstrap",
+  bootstrap = 1000, # generally use 10,000 bootstrap draws; this example uses 1,000 for speed
+  parallel = "multicore", # parallelization for speed: use "multicore" for Mac/Linux; "snow" for PC
+  iseed = 52242, # for reproducibility
+  missing = "ML",
+  estimator = "ML",
+  std.lv = TRUE,
+  fixed.x = FALSE)
+```
+
+Here are the model results:
+
+```{r}
+summary(
+  multipleMediatorFit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
 ## Bayesian Multiple Regression {#sec-multipleRegressionBayesian}

### 2025-04-16 correlation.qmd
20250415 - update packages

@@
 These estimates of nonlinear correlation only range from 0–1 (where 0 represents no association and 1 represents a strong association), so they are interpreted differently than a traditional correlation coefficient (which can be negative).
-We can estimate a distance correlation using the `correlation()` function of the correlation package [@R-correlation; @Makowski2020].
+We can estimate a distance correlation using the `correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].
 
@@
 Another approach is the percentage bend correlation, which gives less weight to observations that are farther away from the median.
-We can estimate the each using the `correlation()` function of the correlation package [@R-correlation; @Makowski2020].
+We can estimate the each using the `correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].
 

### 2025-04-16 multiple-regression.qmd
20250415 - fix
20250415 - update packages

@@
 
-We can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), and added-variable plots (@fig-addedVariablePlots) from the `car` package [@Fox2019; @R-car], to evaluate for potential nonlinearity.
+We can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), and added-variable plots (@fig-addedVariablePlots) from the `car` package [@Fox2019_packages; @R-car], to evaluate for potential nonlinearity.
 For evaluating linearity, we would expect minimal bend/curvature in the lines.
@@
 To determine if residuals are correlated given the nested structure of the data, we can examine the proportion of variance that is attributable to the particular player.
-To do this, we can estimate the intraclass correlation coefficient (ICC) from a [mixed model](#sec-mixedModels) using the `performance` package [@Ludecke2021; @R-performance].
+To do this, we can estimate the intraclass correlation coefficient (ICC) from a [mixed model](#sec-mixedModels) using the `performance` package [@Ludecke2021_packages; @R-performance].
 
@@
 
-We can estimate a regression model that uses pairwise deletion using the `lavaan` package [@Rosseel2012; @R-lavaan].
+We can estimate a regression model that uses pairwise deletion using the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
 
@@
 
-We can multiply impute data using the `mice` package [@vanBuuren2011; @R-mice].
+We can multiply impute data using the `mice` package [@vanBuuren2011_packages; @R-mice].
 
@@
 
-We can estimate a regression model that uses full information maximum likelihood using the `lavaan` package [@Rosseel2012; @R-lavaan].
+We can estimate a regression model that uses full information maximum likelihood using the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
 
@@
 
-A mediation model takes the following general form in the `lavaan` package [@Rosseel2012; @R-lavaan].
+A mediation model takes the following general form in the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
 
@@
 
-  fantasyPoints_transformed ~ b1*receiving_tds + b1*receiving_yards
+  fantasyPoints_transformed ~ b1*receiving_tds + b2*receiving_yards
 

### 2025-04-16 multiple-regression.qmd
20250415 - fixes

@@
 To address a strong positive skew, we can use a log transformation.
-The histogram of the log-transformed variable is in @fig-histogramTargetSharesLog.
+The histogram of the log-transformed variable is in @fig-histogramTargetSharesTransformed.
 
@@
   mediationModel,
-  data = player_stats_seasonal,
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
   se = "bootstrap",
@@
   estimator = "ML",
-  std.lv = TRUE,
   fixed.x = FALSE)
@@
   multipleMediatorModel,
-  data = player_stats_seasonal,
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
   se = "bootstrap",

### 2025-04-16 multiple-regression.qmd
20250416 - fix and newdata prediction

@@
   labs(
-    x = "Fantasy Points",
-    y = "Height (inches)",
+    x = "Height (inches)",
+    y = "Density",
     title = "Density Plot of Player Height with Histogram and Rug Plot"
@@
 We could also generate the model-implied prediction of fantasy points for any value of the [predictor variables](#sec-correlationalStudy).
-For instance, here are the number of fantasy points expected for a Wide Receiver who is 23 years old, who is 6'2" tall (72 inches), and weighs
+For instance, here are the number of fantasy points expected for a Wide Receiver who is 23 years old, is 6'2" tall (72 inches), weighs 200 pounds, and has a 3% target share:
+
+```{r}
+predict(
+  quadraticTermsRegressionModel,
+  newdata = data.frame(
+    age = 23,
+    height = 72,
+    weight = 200,
+    target_share = 3
+  )
+)
+```
+
+```{r}
+#| include: false
+
+predictedValue <- predict(
+  quadraticTermsRegressionModel,
+  newdata = data.frame(
+    age = 23,
+    height = 72,
+    weight = 200,
+    target_share = 3
+  )
+)
+```
+
+The player would be expected to score `r petersenlab::apa(predictedValue, decimals = 0)` fantasy points.
 

### 2025-04-16 multiple-regression.qmd
20250416 - model formulas

@@
 
+The model formula is in @eq-linearRegressionModel:
+
+$$
+\text{fantasy points} = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{height} + \beta_3 \cdot \text{weight} + \beta_4 \cdot \text{target share} + \epsilon
+$$ {#eq-linearRegressionModel}
+
 Here are the model results:
@@
 )
-
+```
+
+The model formula is in @eq-quadraticTermsRegressionModel:
+
+$$
+\begin{aligned}
+  \text{fantasy points} &= \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{age}^2 + \beta_3 \cdot \text{height} + \beta_4 \cdot \text{height}^2 + \\
+  &\phantom{=} \beta_5 \cdot \text{weight} + \beta_6 \cdot \text{weight}^2 + \beta_7 \cdot \text{target share} + \beta_8 \cdot \text{target share}^2 + \epsilon
+\end{aligned}
+$$ {#eq-quadraticTermsRegressionModel}
+
+Here are the model results:
+
+```{r}
 summary(quadraticTermsRegressionModel)
@@
 
+We can calculate this by substituting in the regression coefficients and predictor values:
+
+Here is the computation:
+
+```{r}
+age <- 23
+height <- 72
+weight <- 200
+target_share <- 3
+
+beta0 <- coef(quadraticTermsRegressionModel)["(Intercept)"]
+beta1 <- coef(quadraticTermsRegressionModel)["age"]
+beta2 <- coef(quadraticTermsRegressionModel)["I(age^2)"]
+beta3 <- coef(quadraticTermsRegressionModel)["height"]
+beta4 <- coef(quadraticTermsRegressionModel)["I(height^2)"]
+beta5 <- coef(quadraticTermsRegressionModel)["weight"]
+beta6 <- coef(quadraticTermsRegressionModel)["I(weight^2)"]
+beta7 <- coef(quadraticTermsRegressionModel)["target_share"]
+beta8 <- coef(quadraticTermsRegressionModel)["I(target_share^2)"]
+
+predictedFantasyPoints <- beta0 + beta1*age + beta2*(age^2) + beta3*height + beta4*(height^2) + beta5*weight + beta6*(weight^2) + beta7*target_share + beta8*(target_share^2)
+predictedFantasyPoints
+```
+
+The model formula with substituted values is in @eq-quadraticTermsRegressionModelSubstitute:
+
+$$
+\begin{aligned}
+  \text{fantasy points} &= \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{age}^2 + \beta_3 \cdot \text{height} + \beta_4 \cdot \text{height}^2 + \\
+  &\phantom{=} \beta_5 \cdot \text{weight} + \beta_6 \cdot \text{weight}^2 + \beta_7 \cdot \text{target share} + \beta_8 \cdot \text{target share}^2 + \epsilon \\
+  &= `r petersenlab::apa(beta0, decimals = 2)` + `r petersenlab::apa(beta1, decimals = 2)` \cdot \text{age} + `r petersenlab::apa(beta2, decimals = 2)` \cdot \text{age}^2 + `r petersenlab::apa(beta3, decimals = 2)` \cdot \text{height} + `r petersenlab::apa(beta4, decimals = 2)` \cdot \text{height}^2 + \\
+  &\phantom{=} `r petersenlab::apa(beta5, decimals = 2)` \cdot \text{weight} + `r petersenlab::apa(beta6, decimals = 2)` \cdot \text{weight}^2 + `r petersenlab::apa(beta7, decimals = 2)` \cdot \text{target share} + `r petersenlab::apa(beta8, decimals = 2)` \cdot \text{target share}^2 + \epsilon \\
+  &= `r petersenlab::apa(predictedFantasyPoints, decimals = 2)`
+\end{aligned}
+$$ {#eq-quadraticTermsRegressionModelSubstitute}
+
 #### Evaluating and Addressing Assumptions {#sec-multipleRegressionExamplesAddressingAssumptions}

### 2025-04-16 multiple-regression.qmd
20250416 - update formulas

--- 2025-04-16 multiple-regression.qmd
@@
 where $p$ is the number of [predictor variables](#sec-correlationalStudy).
+Multiple regression is basically a weighted sum of the [predictor variables](#sec-correlationalStudy), and adding an intercept.
+Under the hood, multiple regression seeks to identify the best weight for each [predictor](#sec-correlationalStudy).
 
@@
 The model explained `r petersenlab::apa(linearRegressionModel_rsquared * 100, decimals = 0)`% of the variability in fantasy points (i.e., $R^2 = `r petersenlab::apa(linearRegressionModel_rsquared, decimals = 2, leading = FALSE)`$).
+
+```{r}
+#| include: false
+
+linearRegressionModel_beta0 <- coef(quadraticTermsRegressionModel)[["(Intercept)"]]
+linearRegressionModel_beta1 <- coef(quadraticTermsRegressionModel)[["age"]]
+linearRegressionModel_beta2 <- coef(quadraticTermsRegressionModel)[["height"]]
+linearRegressionModel_beta3 <- coef(quadraticTermsRegressionModel)[["weight"]]
+linearRegressionModel_beta4 <- coef(quadraticTermsRegressionModel)[["target_share"]]
+```
+
+The model formula with substituted values is in @eq-linearRegressionModelSubstitute:
+
+$$
+\begin{aligned}
+  \text{fantasy points} = &\beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{height} + \beta_3 \cdot \text{weight} + \beta_4 \cdot \text{target share} + \epsilon \\
+  = &`r petersenlab::apa(linearRegressionModel_beta0, decimals = 2)` + `r petersenlab::apa(linearRegressionModel_beta1, decimals = 2)` \cdot \text{age} + `r petersenlab::apa(linearRegressionModel_beta2, decimals = 2)` \cdot \text{height} + `r petersenlab::apa(linearRegressionModel_beta3, decimals = 2)` \cdot \text{weight} + `r petersenlab::apa(linearRegressionModel_beta4, decimals = 2)` \cdot \text{target share} + \epsilon
+\end{aligned}
+$$ {#eq-linearRegressionModelSubstitute}
 
@@
 )
-
+```
+
+The model formula is in @eq-quadraticTermsRegressionModel:
+
+$$
+\begin{aligned}
+  \text{fantasy points} \; = \; &\beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{age}^2 + \beta_3 \cdot \text{height} + \beta_4 \cdot \text{height}^2 + \\
+  &\beta_5 \cdot \text{weight} + \beta_6 \cdot \text{weight}^2 + \beta_7 \cdot \text{target share} + \beta_8 \cdot \text{target share}^2 + \epsilon
+\end{aligned}
+$$ {#eq-quadraticTermsRegressionModel}
+
+Here are the model results:
+
+```{r}
 summary(quadraticTermsRegressionModel)
@@
 
+We can calculate this by substituting in the regression coefficients and predictor values:
+
+Here is the computation:
+
+```{r}
+age <- 23
+height <- 72
+weight <- 200
+target_share <- 3
+
+beta0 <- coef(quadraticTermsRegressionModel)[["(Intercept)"]]
+beta1 <- coef(quadraticTermsRegressionModel)[["age"]]
+beta2 <- coef(quadraticTermsRegressionModel)[["I(age^2)"]]
+beta3 <- coef(quadraticTermsRegressionModel)[["height"]]
+beta4 <- coef(quadraticTermsRegressionModel)[["I(height^2)"]]
+beta5 <- coef(quadraticTermsRegressionModel)[["weight"]]
+beta6 <- coef(quadraticTermsRegressionModel)[["I(weight^2)"]]
+beta7 <- coef(quadraticTermsRegressionModel)[["target_share"]]
+beta8 <- coef(quadraticTermsRegressionModel)[["I(target_share^2)"]]
+
+predictedFantasyPoints <- beta0 + beta1*age + beta2*(age^2) + beta3*height + beta4*(height^2) + beta5*weight + beta6*(weight^2) + beta7*target_share + beta8*(target_share^2)
+predictedFantasyPoints
+```
+
+The model formula with substituted values is in @eq-quadraticTermsRegressionModelSubstitute:
+
+$$
+\begin{aligned}
+  \text{fantasy points} \; = \; &\beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{age}^2 + \beta_3 \cdot \text{height} + \beta_4 \cdot \text{height}^2 + \\
+  &\beta_5 \cdot \text{weight} + \beta_6 \cdot \text{weight}^2 + \beta_7 \cdot \text{target share} + \beta_8 \cdot \text{target share}^2 + \epsilon \\
+  = &`r petersenlab::apa(beta0, decimals = 2)` + `r petersenlab::apa(beta1, decimals = 2)` \cdot \text{age} + `r petersenlab::apa(beta2, decimals = 2)` \cdot \text{age}^2 + `r petersenlab::apa(beta3, decimals = 2)` \cdot \text{height} + `r petersenlab::apa(beta4, decimals = 2)` \cdot \text{height}^2 + \\
+  &`r petersenlab::apa(beta5, decimals = 2)` \cdot \text{weight} + `r petersenlab::apa(beta6, decimals = 2)` \cdot \text{weight}^2 + `r petersenlab::apa(beta7, decimals = 2)` \cdot \text{target share} + `r petersenlab::apa(beta8, decimals = 2)` \cdot \text{target share}^2 + \epsilon \\
+  = &`r petersenlab::apa(predictedFantasyPoints, decimals = 2)`
+\end{aligned}
+$$ {#eq-quadraticTermsRegressionModelSubstitute}
+
 #### Evaluating and Addressing Assumptions {#sec-multipleRegressionExamplesAddressingAssumptions}

### 2025-04-16 multiple-regression.qmd
20250416 - fix

@@
 
-linearRegressionModel_beta0 <- coef(quadraticTermsRegressionModel)[["(Intercept)"]]
-linearRegressionModel_beta1 <- coef(quadraticTermsRegressionModel)[["age"]]
-linearRegressionModel_beta2 <- coef(quadraticTermsRegressionModel)[["height"]]
-linearRegressionModel_beta3 <- coef(quadraticTermsRegressionModel)[["weight"]]
-linearRegressionModel_beta4 <- coef(quadraticTermsRegressionModel)[["target_share"]]
+linearRegressionModel_beta0 <- coef(linearRegressionModel)[["(Intercept)"]]
+linearRegressionModel_beta1 <- coef(linearRegressionModel)[["age"]]
+linearRegressionModel_beta2 <- coef(linearRegressionModel)[["height"]]
+linearRegressionModel_beta3 <- coef(linearRegressionModel)[["weight"]]
+linearRegressionModel_beta4 <- coef(linearRegressionModel)[["target_share"]]
 ```

### 2025-04-16 multiple-regression.qmd
20250416 - substitute values

@@
   &\beta_5 \cdot \text{weight} + \beta_6 \cdot \text{weight}^2 + \beta_7 \cdot \text{target share} + \beta_8 \cdot \text{target share}^2 + \epsilon \\
-  = &`r petersenlab::apa(beta0, decimals = 2)` + `r petersenlab::apa(beta1, decimals = 2)` \cdot \text{age} + `r petersenlab::apa(beta2, decimals = 2)` \cdot \text{age}^2 + `r petersenlab::apa(beta3, decimals = 2)` \cdot \text{height} + `r petersenlab::apa(beta4, decimals = 2)` \cdot \text{height}^2 + \\
-  &`r petersenlab::apa(beta5, decimals = 2)` \cdot \text{weight} + `r petersenlab::apa(beta6, decimals = 2)` \cdot \text{weight}^2 + `r petersenlab::apa(beta7, decimals = 2)` \cdot \text{target share} + `r petersenlab::apa(beta8, decimals = 2)` \cdot \text{target share}^2 + \epsilon \\
+  = &`r petersenlab::apa(beta0, decimals = 2)` + `r petersenlab::apa(beta1, decimals = 2)` \cdot `r petersenlab::apa(age, decimals = 0)` + `r petersenlab::apa(beta2, decimals = 2)` \cdot `r petersenlab::apa(age, decimals = 0)`^2 + `r petersenlab::apa(beta3, decimals = 2)` \cdot `r petersenlab::apa(height, decimals = 0)` + `r petersenlab::apa(beta4, decimals = 2)` \cdot `r petersenlab::apa(height, decimals = 0)`^2 + \\
+  &`r petersenlab::apa(beta5, decimals = 2)` \cdot `r petersenlab::apa(weight, decimals = 0)` + `r petersenlab::apa(beta6, decimals = 2)` \cdot `r petersenlab::apa(weight, decimals = 0)`^2 + `r petersenlab::apa(beta7, decimals = 2)` \cdot `r petersenlab::apa(target_share, decimals = 0)` + `r petersenlab::apa(beta8, decimals = 2)` \cdot `r petersenlab::apa(target_share, decimals = 0)`^2 + \epsilon \\
   = &`r petersenlab::apa(predictedFantasyPoints, decimals = 2)`

### 2025-04-16 multiple-regression.qmd
20250416 - absolute value of direct and indirect effects

@@
 
-We can calculate this by substituting in the regression coefficients and predictor values:
-
+We can calculate this by substituting in the regression coefficients and predictor values.
 Here is the computation:
@@
   # total effect (c)
-  total := direct + indirect
+  total := abs(direct) + abs(indirect)
 '
@@
   # total effect (c)
-  total := direct + indirectTotal
+  total := abs(direct) + abs(indirectTotal)
 '

### 2025-04-17 multiple-regression.qmd
20250417 - diagnostic plots

@@
 For instance, if the parameter estimate for the regression coefficient is 0.50, and the standard error is 0.10, the 95% confidence interval is [0.30, 0.70]: $0.5 - (1.96 \times 0.10) = 0.3$; $0.5 + (1.96 \times 0.10) = 0.7$.
-That is, if we used the same sample procedure repeatedly, the true value of the regression coefficient would be expected to be 95% of the time somewhere between 0.30 to 0.70.
+That is, if we used the same sampling procedure repeatedly, the true value of the regression coefficient would be expected to be 95% of the time somewhere between 0.30 to 0.70.
 
@@
 Homoscedasticity of the residuals means that the variance of the residuals does not differ as a function of the [outcome variable](#sec-correlationalStudy) or as a function of the [predictor variable](#sec-correlationalStudy) (i.e., the residuals show constant variance as a function of [outcome](#sec-correlationalStudy)/[predictors](#sec-correlationalStudy)).
+If the residuals differ as a function of the [outcome](#sec-correlationalStudy) or [predictor](#sec-correlationalStudy) variable, this is called heterotoscedasticity.
 
@@
 However, there are additional assumptions of multiple regression, including ones discussed in the chapter on [Causal Inference](#sec-causalInference).
-For instance, the variables included should reflect causal process such that the [predictor variables](#sec-correlationalStudy) influence the [outcome variable](#sec-correlationalStudy), and not the other way around.
+For instance, the variables included should reflect a causal process such that the [predictor variables](#sec-correlationalStudy) influence the [outcome variable](#sec-correlationalStudy), and not the other way around.
 That is, the [outcome variable](#sec-correlationalStudy) should not influence the [predictor variables](#sec-correlationalStudy) (i.e., there should be no reverse causation).
@@
 
-### Evaluating Assumptions of Multiple Regression {#sec-assumptionsRegressionEvaluating}
-
-To evaluate the shape of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy), we can examine scatterplots, residual plots, marginal model plots, and added-variable plots.
+### Evaluating and Addressing Assumptions of Multiple Regression {#sec-assumptionsRegressionEvaluating}
+
+#### Linear Association {#sec-assumptionsRegressionEvaluatingLinearAssociation}
+
+To evaluate the shape of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy), we can examine [scatterplots](#sec-scatterplot), residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), added-variable plots (@fig-addedVariablePlots), and component-plus-residual plots (@fig-componentPlusResidualPlots).
+Residual plots depict the residuals (errors) on the y-axis as a function of the fitted values or a specific [predictor](#sec-correlationalStudy) on the x-axis.
+Marginal model plots are basically glorified [scatterplots](#sec-scatterplot) that depict the [outcome variable](#sec-correlationalStudy) (both in terms of observed values and model-fitted values) on the y-axis and the [predictor variables](#sec-correlationalStudy) on the x-axis.
+Added-variable plots depict the *unique* association of each [predictor variables](#sec-correlationalStudy) with the [outcome variable](#sec-correlationalStudy) when controlling for all the other [predictor variables](#sec-correlationalStudy) in the model.
+Component-plus-residual plots depict partial residuals on the y-axis as a function of each [predictor variable](#sec-correlationalStudy) on the x-axis, where a partial residual for a given [predictor](#sec-correlationalStudy) is the effect of a given predictor (thus controlling for all the other [predictor variables](#sec-correlationalStudy) in the model) plus the residual from the full model.
+
+If the shape of the association is nonlinear (as indicated by any of these plots), various approaches may be necessary such as including nonlinear terms (e.g., polynomial terms such as quadratic, cubic, quartic, or higher-degree terms), transforming the [predictors](#sec-correlationalStudy) (e.g., log, square root, inverse, exponential, Box-Cox, Yeo-Johnson transform), use of splines/piecewise regression, and generalized additive models.
+
+#### Homoscedasticity {#sec-assumptionsRegressionEvaluatingHomoscedasticity}
 
@@
 In residual plots, homoscedasticity is identified as a constant spread of residuals versus fitted values—the residuals do not show a fan or cone shape.
+If there is heteroscedasticity, it may be necessary to transform the [outcome variable](#sec-correlationalStudy) to be more normally distributed.
+
+#### Uncorrelated Residuals {#sec-assumptionsRegressionEvaluatingUncorrelatedResiduals}
 
@@
 The greater the ICC value, the more variance is accounted for by the grouping level, and the more the residuals are intercorrelated.
-
-We can examine whether residuals are normally distributed using quantile–quantile (QQ) plots and probability–probability (PP) plots.
-If the residuals are normally distributed, they should stay close to the diagonal reference line.
+If the residuals are intercorrelated, it may be necessary to account for the grouping structure of the data using a [mixed model](#sec-mixedModels).
+
+#### Normally Distributed Residuals {#sec-assumptionsRegressionEvaluatingNormallyDistributedResiduals}
+
+To examine whether residuals are normally distributed, we can examine quantile–quantile (QQ) plots and probability–probability (PP) plots.
+QQ plots are particularly useful for identifying deviations from normality in the tails of the distribution; PP plots are particularly useful for identifying deviations from normality in the center of the distribution.
+Researchers tend to be more concerned about the tails of the distribution, because extreme values tend to have a greater impact on inferences.
+Various examples of QQ plots and deviations from normality are depicted in @fig-distributionQQplots.
+If the residuals are normally distributed, they will stay close to the diagonal reference line of the QQ and PP plots.
+
+```{r}
+#| label: fig-distributionQQplots
+#| fig-cap: "Quantile–Quantile (QQ) Plots of Various Distributions."
+#| fig-alt: "Quantile–Quantile (QQ) Plots of Various Distributions."
+#| fig-subcap:
+#|   - "Normal Distribution"
+#|   - "Normal Distribution"
+#|   - "Bimodal Distribution"
+#|   - "Bimodal Distribution"
+#|   - "Negatively Skewed Distribution"
+#|   - "Negatively Skewed Distribution"
+#|   - "Positively Skewed Distribution"
+#|   - "Positively Skewed Distribution"
+#|   - "Platykurtic (Light-Tailed) Distribution"
+#|   - "Platykurtic (Light-Tailed) Distribution"
+#|   - "Lepokurtic (Heavy-Tailed) Distribution"
+#|   - "Lepokurtic (Heavy-Tailed) Distribution"
+#| code-fold: true
+
+set.seed(52242)
+
+sampleSize <- 10000
+
+distribution_normal <- rnorm(
+  sampleSize,
+  mean = 0,
+  sd = 1)
+
+distribution_bimodal <- c(
+  rnorm(
+    sampleSize/2,
+    mean = -2,
+    sd = 1),
+  rnorm(
+    sampleSize/2,
+    mean = 2,
+    sd = 1))
+
+distribution_negativeSkew <- -rlnorm(
+  sampleSize,
+  meanlog = 0,
+  sdlog = 1)
+
+distribution_positiveSkew <- rlnorm(
+  sampleSize,
+  meanlog = 0,
+  sdlog = 1)
+
+distribution_lightTailed <- runif(
+  sampleSize,
+  min = -2,
+  max = 2)
+
+distribution_heavyTailed <- rt(
+  sampleSize,
+  df = 2)
+
+hist(
+  distribution_normal,
+  main = "Normal Distribution",
+  col = "skyblue")
+
+car::qqPlot(
+  distribution_normal,
+  main = "QQ Plot",
+  id = TRUE)
+
+hist(
+  distribution_bimodal,
+  main = "Bimodal",
+  col = "skyblue")
+
+car::qqPlot(
+  distribution_bimodal,
+  main = "QQ Plot",
+  id = TRUE)
+
+hist(
+  distribution_negativeSkew,
+  main = "Negatively Skewed",
+  col = "skyblue")
+
+car::qqPlot(
+  distribution_negativeSkew,
+  main = "QQ Plot",
+  id = TRUE)
+
+hist(
+  distribution_positiveSkew,
+  main = "Positively Skewed",
+  col = "skyblue")
+
+car::qqPlot(
+  distribution_positiveSkew,
+  main = "QQ Plot",
+  id = TRUE)
+
+hist(
+  distribution_lightTailed,
+  main = "Platykurtic (Light Tailed)",
+  col = "skyblue")
+
+car::qqPlot(
+  distribution_lightTailed,
+  main = "QQ Plot",
+  id = TRUE)
+
+hist(
+  distribution_heavyTailed,
+  main = "Leptokurtic (Heavy Tailed)",
+  col = "skyblue")
+
+car::qqPlot(
+  distribution_heavyTailed,
+  main = "QQ Plot",
+  id = TRUE)
+```
+
+If the residuals are not normally distributed (i.e., they do not stay close to the diagonal reference line of the QQ and PP plots), it may be necessary to transform the [outcome variable](#sec-correlationalStudy) to be more normally distributed or to use a generalized linear model (GLM) that more closely matches the distribution of the [outcome variable](#sec-correlationalStudy) (e.g., Poisson, binomial, gamma).
 
@@
 
-Then, let's examine the bivariate association of each using a scatterplot to evaluate for any potential nonlinearity, as depicted in @fig-scatterplots.
+Then, let's examine the bivariate association of each using a [scatterplot](#sec-scatterplot) to evaluate for any potential nonlinearity, as depicted in @fig-scatterplots.
 
@@
 
-We evaluated the shape of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variables](#sec-correlationalStudy) using scatterplots.
-We accounted for potential curvilinearity with a quadratic term.
-Other ways to account for nonlinearity, in addition to polynomials (e.g., quadratic, cubic, quartic, or higher-degree terms), include transforming [predictors](#sec-correlationalStudy) (e.g., log, square root, inverse, exponential transform), splines/piecewise regression, and generalized additive models.
-
-We can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), and added-variable plots (@fig-addedVariablePlots) from the `car` package [@Fox2019_packages; @R-car], to evaluate for potential nonlinearity.
+We evaluated the shape of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variables](#sec-correlationalStudy) using [scatterplots](#sec-scatterplot).
+We accounted for potential curvilinearity in the associations with a quadratic term.
+Other ways to account for nonlinearity, in addition to polynomials, include transforming [predictors](#sec-correlationalStudy), use of splines/piecewise regression, and generalized additive models.
+
+To evaluate for potential nonlinearity in the associations, we can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), and added-variable plots (@fig-addedVariablePlots), and component-plus-residual plots (@fig-componentPlusResidualPlots) from the `car` package [@Fox2019_packages; @R-car].
 For evaluating linearity, we would expect minimal bend/curvature in the lines.
@@
 
-The marginal model plots (@fig-marginalModelPlots) and residual plots (@fig-residualPlots) suggest that the nonlinearity of the association between target share and fantasy points may not be fully captured by the quadratic term.
+```{r}
+#| label: fig-componentPlusResidualPlots
+#| fig-cap: "Component-Plus-Residual Plots."
+#| fig-alt: "Component-Plus-Residual Plots."
+
+car::crPlots(
+  quadraticTermsRegressionModel,
+  sd = TRUE,
+  id = TRUE)
+```
+
+The marginal model plots (@fig-marginalModelPlots), residual plots (@fig-residualPlots), and  component-plus-residual plots (@fig-componentPlusResidualPlots) suggest that the nonlinearity of the association between target share and fantasy points may not be fully captured by the quadratic term.
 Thus, we may need to apply a different approach to handling the nonlinear association between target share and fantasy points.
@@
 
-hist(player_stats_seasonal$target_share[which(player_stats_seasonal$position == "WR")])
+hist(
+  player_stats_seasonal$target_share[which(player_stats_seasonal$position == "WR")],
+  main = "Histogram of Target Share")
 ```
@@
 
-hist(log(player_stats_seasonal$target_share[which(player_stats_seasonal$position == "WR")] + 1))
+hist(
+  log(player_stats_seasonal$target_share[which(player_stats_seasonal$position == "WR")] + 1),
+  "Histogram of Target Share (Log Transformed)")
 ```
@@
 
-Target share shows a more linear association with fantasy points after log-transforming it (albeit still not perfect), as depicted in Figures [-@fig-marginalModelPlotsLogTransformed], [-@fig-addedVariablePlotsLogTransformed], and [-@fig-residualPlotsLogTransformed].
+Target share shows a more linear association with fantasy points after log-transforming it (albeit still not perfect), as depicted in Figures [-@fig-marginalModelPlotsLogTransformed], [-@fig-addedVariablePlotsLogTransformed], [-@fig-componentPlusResidualPlotsLogTransformed], and [-@fig-residualPlotsLogTransformed].
 
@@
 ```{r}
+#| label: fig-componentPlusResidualPlotsLogTransformed
+#| fig-cap: "Component-Plus-Residual Plots After Log Transformation of Target Share."
+#| fig-alt: "Component-Plus-Residual Plots After Log Transformation of Target Share."
+
+car::crPlots(
+  linearRegressionModel_logTargetShare,
+  id = TRUE)
+```
+
+```{r}
 #| label: fig-residualPlotsLogTransformed
@@
 
-hist(player_stats_seasonal$fantasyPoints[which(player_stats_seasonal$position == "WR")])
-```
-
+hist(
+  player_stats_seasonal$fantasyPoints[which(player_stats_seasonal$position == "WR")],
+  main = "Histogram of Fantasy Points")
+```
+
+We can apply a Yeo-Johnson transformation to the [outcome variable](#sec-correlationalStudy) to generate a more normally distributed variable.
+A Yeo-Johnson transformation estimates the optimal transformation to make the variable more normally distributed.
 Let's use a Yeo-Johnson transformation of fantasy points:
@@
 
-hist(player_stats_seasonal$fantasyPoints_transformed[which(player_stats_seasonal$position == "WR")])
+hist(
+  player_stats_seasonal$fantasyPoints_transformed[which(player_stats_seasonal$position == "WR")],
+  main = "Histogram of Fantasy Points (Transformed)")
 ```
@@
 
-fimlRegression_fit <- sem(
+fimlRegression_fit <- lavaan::sem(
   fimlRegression_syntax,
@@
 ```{r}
-mediationFit <- sem(
+mediationFit <- lavaan::sem(
   mediationModel,
@@
 ```{r}
-multipleMediatorFit <- sem(
+multipleMediatorFit <- lavaan::sem(
   multipleMediatorModel,

### 2025-04-17 multiple-regression.qmd
20250417 - plots

@@
 ```{r}
+#| layout-ncol: 2
 #| label: fig-distributionQQplots
@@
 #|   - "Platykurtic (Light-Tailed) Distribution"
-#|   - "Lepokurtic (Heavy-Tailed) Distribution"
-#|   - "Lepokurtic (Heavy-Tailed) Distribution"
+#|   - "Leptokurtic (Heavy-Tailed) Distribution"
+#|   - "Leptokurtic (Heavy-Tailed) Distribution"
 #| code-fold: true
@@
   main = "Normal Distribution",
-  col = "skyblue")
+  col = "#0099F8")
 
@@
   distribution_bimodal,
-  main = "Bimodal",
-  col = "skyblue")
+  main = "Bimodal Distribution",
+  col = "#0099F8")
 
@@
   distribution_negativeSkew,
-  main = "Negatively Skewed",
-  col = "skyblue")
+  main = "Negatively Skewed Distribution",
+  col = "#0099F8")
 
@@
   distribution_positiveSkew,
-  main = "Positively Skewed",
-  col = "skyblue")
+  main = "Positively Skewed Distribution",
+  col = "#0099F8")
 
@@
   distribution_lightTailed,
-  main = "Platykurtic (Light Tailed)",
-  col = "skyblue")
+  main = "Platykurtic (Light Tailed) Distribution",
+  col = "#0099F8")
 
@@
   distribution_heavyTailed,
-  main = "Leptokurtic (Heavy Tailed)",
-  col = "skyblue")
+  main = "Leptokurtic (Heavy Tailed) Distribution",
+  col = "#0099F8")
 
@@
 To address multicollinearity, you can drop a redundant predictor or you can also use [principal component analysis](#sec-pca) or [factor analysis](#sec-factorAnalysis) of the [predictors](#sec-correlationalStudy) to reduce the [predictors](#sec-correlationalStudy) down to a smaller number of meaningful [predictors](#sec-correlationalStudy).
-For a meaningful answer in a regression framework that is precise and confident, you need a low level of intercorrelation among [predictors](#sec-correlationalStudy), unless you have a very large sample size.
+For a meaningful answer regarding [predictors](#sec-correlationalStudy) in a regression framework that is precise and confident, you need a low level of intercorrelation among [predictors](#sec-correlationalStudy), unless you have a very large sample size.
+However, if you are merely interested in prediction—and are not interested in interpreting the regression coefficients of individual predictors—multicollinearity poses less of a problem.
+For instance, [machine learning](#sec-machineLearning) cares more about achieving the greatest predictive accuracy possible and cares less about explaining which [predictors](#sec-correlationalStudy) are causally related to the [outcome](#sec-correlationalStudy).
+So, multicollinearity is less of a concern for [machine learning](#sec-machineLearning) approaches.
 
@@
 However, identifying associations is important because associations are a necessary (but insufficient) condition for causality.
-When developing a multiple regression model, it is important to pay attention for potential [multicollinearity](#sec-multipleRegressionMulticollinearity)—it may become difficult to detect a given [predictor variable](#sec-correlationalStudy) as [statistically significant](#sec-statisticalSignificance) due to the greater uncertainty around the parameter estimates.
+When developing a multiple regression model, there are various [assumptions](#sec-assumptionsRegression) that are important to evaluate.
+In addition, it is important to pay attention for potential [multicollinearity](#sec-multipleRegressionMulticollinearity)—it may become difficult to detect a given [predictor variable](#sec-correlationalStudy) as [statistically significant](#sec-statisticalSignificance) due to the greater uncertainty around the parameter estimates.
 

### 2025-04-17 mixed-models.qmd
20250417 - cross-references

@@
 [Scatterplots](#sec-scatterplot) are a helpful tool for quickly examining the association between two variables.
-However, [scatterplots](#sec-scatterplot)—as well as correlation and multiple regression—can hide meaningful associations that differ across units of analysis.
+However, [scatterplots](#sec-scatterplot)—as well as [correlation](#sec-correlation) and [multiple regression](#sec-multipleRegression)—can hide meaningful associations that differ across units of analysis.
 

### 2025-04-17 sports-cognitive-psychology.qmd
20250408 - placeholder
20250417 - flow

@@
 
+Athletes sometimes describe periods of optimal performance where they are "in the zone."
+These periods of mental state are called "flow", and are characterized by the following subjective characteristics [@Nakamura2021]:
+
+- intense and focused concentration on the present moment
+- merging of action and awareness
+- loss of reflective self-consciousness (i.e., loss of awareness of oneself as a social actor)
+- a sense that one can control one's actions and deal with the situation because one knows how to respond to whatever happens next
+- distorted sense of time (often, a sense that time has passed faster than usual)
+- experiencing the activity as intrinsically rewarding
+
+Because these periods often coincide with strong performance, athletes commonly strive to find ways to achieve flow.
+
 One process that may have a nonlinear association with performance is arousal or anxiety.

### 2025-04-17 multiple-regression.qmd
20250417 - spread-level plot

@@
 
-To evaluate homoscedasticity, we can evaluate residual plots (@fig-residualPlots).
-In residual plots, homoscedasticity is identified as a constant spread of residuals versus fitted values—the residuals do not show a fan or cone shape.
+To evaluate homoscedasticity, we can evaluate a residual plot (@fig-residualPlots) and spread-level plot (@fig-spreadLevelPlot).
+A residual plot depicts the residuals on the y-axis as a function of the model's fitted values on the x-axis.
+Homoscedasticity in a residual plot is identified as a constant spread of residuals versus fitted values—the residuals do not show a fan or cone shape; a fan or cone shape indicates heteroscedasticity.
+A spread-level plot depicts the log of the absolute value of studentized residuals on the y-axis as a function of the log of the model's fitted values on the x-axis.
+Homoscedasticity in a spread-level plot is identified as a flat slope; a slope that differs from zero indicates heteroscedasticity.
+
 If there is heteroscedasticity, it may be necessary to transform the [outcome variable](#sec-correlationalStudy) to be more normally distributed.
+The spread-level plot provides a suggested power transformation to transform the [outcome variable](#sec-correlationalStudy) so that the spread of residuals becomes more uniform across the fitted values.
 
@@
 
-To evaluate homoscedasticity, we can evaluate residual plots (@fig-residualPlots).
-In residual plots, you want a constant spread of residuals versus fitted values—you do not want the residuals to show a fan or cone shape.
+To evaluate homoscedasticity, we can evaluate a residual plot (@fig-residualPlots) and spread-level plot (@fig-spreadLevelPlots).
+In a residual plot, you want a constant spread of residuals versus fitted values—you do not want the residuals to show a fan or cone shape.
 
@@
 car::residualPlots(
+  quadraticTermsRegressionModel,
+  id = TRUE)
+```
+
+In a spread-level plot, you want a flat (zero) slope—you do not want a positive or negative slope.
+
+```{r}
+#| label: fig-spreadLevelPlot
+#| fig-cap: "Spread-Level Plot."
+#| fig-alt: "Spread-Level Plot."
+
+car::spreadLevelPlot(
   quadraticTermsRegressionModel,
@@
 
-The residual plots are in @fig-residualPlotsOutcomeTransformed.
+The residual plot is in @fig-residualPlotsOutcomeTransformed.
 
@@
 car::residualPlots(
+  linearRegressionModel_outcomeTransformed,
+  id = TRUE)
+```
+
+The spread-level plot is in @fig-spreadLevelPlotOutcomeTransformed.
+
+```{r}
+#| label: fig-spreadLevelPlotOutcomeTransformed
+#| fig-cap: "Spread-Level Plot After Transformation of Fantasy Points."
+#| fig-alt: "Spread-Level Plot After Transformation of Fantasy Points."
+
+car::spreadLevelPlot(
   linearRegressionModel_outcomeTransformed,

### 2025-04-17 multiple-regression.qmd
20250417 - fixes and references

@@
 
-To evaluate for potential nonlinearity in the associations, we can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), and added-variable plots (@fig-addedVariablePlots), and component-plus-residual plots (@fig-componentPlusResidualPlots) from the `car` package [@Fox2019_packages; @R-car].
+To evaluate for potential nonlinearity in the associations, we can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), added-variable plots (@fig-addedVariablePlots), and component-plus-residual plots (@fig-componentPlusResidualPlots) from the `car` package [@Fox2019_packages; @R-car].
 For evaluating linearity, we would expect minimal bend/curvature in the lines.
@@
   log(player_stats_seasonal$target_share[which(player_stats_seasonal$position == "WR")] + 1),
-  "Histogram of Target Share (Log Transformed)")
+  main = "Histogram of Target Share (Log Transformed)")
 ```
@@
 
+When creating a residual plot, the `car` package [@Fox2019_packages; @R-car] also provides a test of the nonlinearity of each [predictor](#sec-correlationalStudy).
+
 ```{r}
@@
 
-To evaluate homoscedasticity, we can evaluate a residual plot (@fig-residualPlots) and spread-level plot (@fig-spreadLevelPlots).
+To evaluate homoscedasticity, we can evaluate a residual plot (@fig-residualPlots) and spread-level plot (@fig-spreadLevelPlots) from the `car` package [@Fox2019_packages; @R-car].
 In a residual plot, you want a constant spread of residuals versus fitted values—you do not want the residuals to show a fan or cone shape.

### 2025-04-18 index.qmd
20250417 - robust regression

--- 2025-04-11 index.qmd
@@
   "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation", 
-  "mgcv", "rms", "car", "performance", "lavaan", "mice", "miceadds"
+  "mgcv", "rms", "car", "performance", "lavaan", "mice", "miceadds","interactions",
+  "shiny","PearsonDS","robustbase"
 )

### 2025-04-18 multiple-regression.qmd
20250417 - robust regression

@@
 library("brms")
+library("robustbase")
 library("tidyverse")
@@
   filter(position %in% c("WR")) %>% 
-  select(player_id, where(is.numeric)) %>% 
+  select(player_id, position, where(is.numeric)) %>% 
   select(
@@
   fimlRegression_syntax,
-  data = player_stats_seasonal,
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
   missing = "ML",
@@
 player_stats_seasonal_subset <- player_stats_seasonal %>% 
-  filter(!is.na(player_id))
+  filter(!is.na(player_id)) %>% 
+  filter(position %in% c("WR"))
 
 rms::robcov(rms::ols(
-  fantasyPoints_transformed ~ age + height + weight + target_share_log,
+  fantasyPoints_transformed ~ age + height + weight + I(log(target_share + 1)),
   data = player_stats_seasonal_subset,
@@
 [As with correlation](#sec-correlationOutliers), multiple regression can be strongly impacted by outliers.
+
+### Robust Regression {#sec-multipleRegressionRobustRegression}
+
+To address outliers, there are various approaches to robust regression.
+One approach is to use an MM-type estimator, such as is used in the `lmrob()` and `glmrob()` functions of the `robustbase` package [@R-robustbase; @Todorov2009_packages].
+
+```{r}
+robustRegression <- lmrob(
+  fantasyPoints_transformed ~ age + height + weight + I(log(target_share + 1)),
+  data = player_stats_seasonal %>% filter(position %in% c("WR"))
+)
+
+summary(robustRegression)
+```
 

### 2025-04-18 multiple-regression.qmd
20250417 - figures

@@
 #| label: fig-distributionQQplots
-#| fig-cap: "Quantile–Quantile (QQ) Plots of Various Distributions."
-#| fig-alt: "Quantile–Quantile (QQ) Plots of Various Distributions."
+#| fig-cap: "Quantile–Quantile (QQ) Plots of Various Distributions (Right Side) with the Histogram of the Associated Distribution (Left Side)."
+#| fig-alt: "Quantile–Quantile (QQ) Plots of Various Distributions (Right Side) with the Histogram of the Associated Distribution (Left Side)."
 #| fig-subcap:
@@
   distribution_normal,
-  main = "QQ Plot",
-  id = TRUE)
+  main = "QQ Plot")
 
@@
   distribution_bimodal,
-  main = "QQ Plot",
-  id = TRUE)
+  main = "QQ Plot")
 
@@
   distribution_negativeSkew,
-  main = "QQ Plot",
-  id = TRUE)
+  main = "QQ Plot")
 
@@
   distribution_positiveSkew,
-  main = "QQ Plot",
-  id = TRUE)
+  main = "QQ Plot")
 
@@
   distribution_lightTailed,
-  main = "QQ Plot",
-  id = TRUE)
+  main = "QQ Plot")
 
@@
   distribution_heavyTailed,
-  main = "QQ Plot",
-  id = TRUE)
+  main = "QQ Plot")
 ```

### 2025-04-18 research-methods.qmd
20250219 - causal inference
20250417 - attenuation of associations due to unreliability

@@
 Random assignment involves randomly assigning participants to conditions of the independent variable, so the people in each condition are comparable and do not differ systematically.
+However, not all things are feasible or ethical to manipulate.
+For instance, it would not be ethical to randomly assign some players to receive a head trauma.
+In addition, it would not be feasible to manipulate the weather of a game or the performance of players.
 
@@
 For more information on these and other aspects of reliability, see <https://isaactpetersen.github.io/Principles-Psychological-Assessment/reliability.html> [@Petersen2024a; @PetersenPrinciplesPsychAssessment].
+Reliability is important to consider because random measurement error (i.e., unreliability) weakens the associations between variables [@Goodwin2006; @Schmidt1996].
 

### 2025-04-18 multiple-regression.qmd
20250417 - nonlinearity and heteroscedasticity

@@
 
-To evaluate the shape of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy), we can examine [scatterplots](#sec-scatterplot), residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), added-variable plots (@fig-addedVariablePlots), and component-plus-residual plots (@fig-componentPlusResidualPlots).
+To evaluate the shape of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy), we can examine [scatterplots](#sec-scatterplot) (@fig-linearNonlinearAssociations), residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), added-variable plots (@fig-addedVariablePlots), and component-plus-residual plots (@fig-componentPlusResidualPlots).
 Residual plots depict the residuals (errors) on the y-axis as a function of the fitted values or a specific [predictor](#sec-correlationalStudy) on the x-axis.
@@
 
+Examples of linear and nonlinear associations are depicted with [scatterplots](#sec-scatterplot) in @fig-linearNonlinearAssociations.
+
+```{r}
+#| layout-ncol: 2
+#| label: fig-linearNonlinearAssociations
+#| fig-cap: "Example Associations Depicted With Scatterplots."
+#| fig-alt: "Example Associations Depicted With Scatterplots."
+#| fig-subcap:
+#|   - "Example of a Linear Association"
+#|   - "Example of a Nonlinear Association"
+#| code-fold: true
+
+set.seed(52242)
+
+sampleSize <- 1000
+quadraticX <- runif(
+  sampleSize,
+  min = -4,
+  max = 4)
+linearY <- quadraticX + rnorm(sampleSize, mean = 0, sd = 0.5)
+quadraticY <- quadraticX ^ 2 + rnorm(sampleSize)
+quadraticData <- cbind(quadraticX, quadraticY) %>%
+  data.frame %>%
+  arrange(quadraticX)
+
+quadraticModel <- lm(
+  quadraticY ~ quadraticX + I(quadraticX ^ 2),
+  data = quadraticData)
+
+quadraticNewData <- data.frame(
+  quadraticX = seq(
+    from = min(quadraticData$quadraticX),
+    to = max(quadraticData$quadraticY),
+    length.out = sampleSize))
+
+quadraticNewData$quadraticY <- predict(
+  quadraticModel,
+  newdata = quadraticNewData)
+
+plot(
+  x = quadraticData$quadraticX,
+  y = quadraticData$quadraticY,
+  xlab = "",
+  ylab = "",
+  main = "Nonlinear Association")
+
+lines(
+  quadraticNewData$quadraticY ~ quadraticNewData$quadraticX,
+  lwd = 2,
+  col = "blue")
+
+plot(
+  x = quadraticX,
+  y = linearY,
+  xlab = "",
+  ylab = "",
+  main = "Linear Association")
+
+abline(
+  lm(linearY ~ quadraticX),
+  lwd = 2,
+  col = "blue")
+```
+
 If the shape of the association is nonlinear (as indicated by any of these plots), various approaches may be necessary such as including nonlinear terms (e.g., polynomial terms such as quadratic, cubic, quartic, or higher-degree terms), transforming the [predictors](#sec-correlationalStudy) (e.g., log, square root, inverse, exponential, Box-Cox, Yeo-Johnson transform), use of splines/piecewise regression, and generalized additive models.
@@
 A residual plot depicts the residuals on the y-axis as a function of the model's fitted values on the x-axis.
-Homoscedasticity in a residual plot is identified as a constant spread of residuals versus fitted values—the residuals do not show a fan or cone shape; a fan or cone shape indicates heteroscedasticity.
+Homoscedasticity in a residual plot is identified as a constant spread of residuals versus fitted values—the residuals do not show a fan, cone, or bow-tie shape; a fan, cone, or bow-tie shape indicates heteroscedasticity.
+In a residual plot, a fan or cone shape indicates increasing or decreasing variance in the residuals as a function of the fitted values; a bow-tie shape indicates that the residuals are smallest in the middle of the fitted values and greatest on the extremes of the fitted values.
 A spread-level plot depicts the log of the absolute value of studentized residuals on the y-axis as a function of the log of the model's fitted values on the x-axis.
 Homoscedasticity in a spread-level plot is identified as a flat slope; a slope that differs from zero indicates heteroscedasticity.
+
+```{r}
+#| layout: [[100], [50, 50]]
+#| label: fig-homscedasticityHeteroscedasticity
+#| fig-cap: "Example of Homoscedasticity and Heteroscedasticity in Residual Plots."
+#| fig-alt: "Example of Homoscedasticity and Heteroscedasticity in Residual Plots."
+#| fig-subcap:
+#|   - "Homoscedasticity"
+#|   - "Fan-Shaped Heteroscedasticity"
+#|   - "Bow-Tie-Shaped Heteroscedasticity"
+#| code-fold: true
+
+set.seed(52242)
+
+# 1. Homoscedasticity
+sampleSize <- 1000
+x1 <- runif(sampleSize, 0, 10)
+y1 <- 3 + 2 * x1 + rnorm(sampleSize, mean = 0, sd = 2)
+fit1 <- lm(y1 ~ x1)
+res1 <- resid(fit1)
+fitted1 <- fitted(fit1)
+
+# 2. Fan-shaped heteroscedasticity
+x2 <- runif(sampleSize, 0, 10)
+y2 <- 3 + 2 * x2 + rnorm(sampleSize, mean = 0, sd = 0.5 * x2) # increasing variance
+fit2 <- lm(y2 ~ x2)
+res2 <- resid(fit2)
+fitted2 <- fitted(fit2)
+
+# 3. Bow-tie-shaped heteroscedasticity
+x3 <- runif(sampleSize, 0, 10)
+sd3 <- abs(x3 - 5) + 0.5  # variance smallest in middle, higher on edges
+y3 <- 3 + 2 * x3 + rnorm(sampleSize, mean = 0, sd = sd3)
+fit3 <- lm(y3 ~ x3)
+res3 <- resid(fit3)
+fitted3 <- fitted(fit3)
+
+plot(
+  fitted1,
+  res1,
+  main = "Homoscedasticity",
+  xlab = "Fitted value",
+  ylab = "Residual")
+
+abline(
+  h = 0,
+  lwd = 2,
+  col = "blue")
+
+plot(
+  fitted2,
+  res2,
+  main = "Fan-Shaped Heteroscedasticity",
+  xlab = "Fitted value",
+  ylab = "Residual")
+
+abline(
+  h = 0,
+  lwd = 2,
+  col = "blue")
+
+plot(
+  fitted3,
+  res3,
+  main = "Bow-Tie-Shaped Heteroscedasticity",
+  xlab = "Fitted value",
+  ylab = "Residual")
+
+abline(
+  h = 0,
+  lwd = 2,
+  col = "blue")
+```
 

### 2025-04-18 multiple-regression.qmd
20250417 - clean up figures

@@
   distribution_normal,
-  main = "QQ Plot")
+  main = "QQ Plot",
+  id = FALSE)
 
@@
   distribution_bimodal,
-  main = "QQ Plot")
+  main = "QQ Plot",
+  id = FALSE)
 
@@
   distribution_negativeSkew,
-  main = "QQ Plot")
+  main = "QQ Plot",
+  id = FALSE)
 
@@
   distribution_positiveSkew,
-  main = "QQ Plot")
+  main = "QQ Plot",
+  id = FALSE)
 
@@
   distribution_lightTailed,
-  main = "QQ Plot")
+  main = "QQ Plot",
+  id = FALSE)
 
@@
   distribution_heavyTailed,
-  main = "QQ Plot")
+  main = "QQ Plot",
+  id = FALSE)
 ```

### 2025-04-18 mixed-models.qmd
20250418 - diagnostic plots

@@
 library("fitdistrplus")
+library("performance")
 library("parallel")
@@
 
+#### Diagnostic Plots {#sec-fantasyPointsByAgeExperienceModelsDiagnosticPlots}
+
+##### Plot of Residuals {#sec-fantasyPointsByAgeExperienceModelsResidualPlot}
+
+```{r}
+plot(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+
+plot(
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
+  type = c("p","smooth"),
+  col.line = 1)
+```
+
+##### Scale-Location Plot {#sec-fantasyPointsByAgeExperienceModelsScaleLocationPlot}
+
+```{r}
+plot(
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
+  sqrt(abs(resid(.))) ~ fitted(.),
+  type = c("p","smooth"),
+  col.line = 1)
+```
+
+##### Quantile–Quantile (QQ) Plot {#sec-fantasyPointsByAgeExperienceModelsQQPlot}
+
+```{r}
+qqnorm(resid(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience))
+qqline(resid(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience))
+```
+
+##### Probability–Probability (PP) Plots {#sec-fantasyPointsByAgeExperienceModelsDiagnosticPPPlots}
+
+```{r}
+petersenlab::ppPlot(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+```
+
+##### Residuals Versus Leverage Plot {#sec-fantasyPointsByAgeExperienceModelsResidualsVsLeveragePlot}
+
+```{r}
+plot(
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
+  rstudent(.) ~ hatvalues(.))
+```
+
+##### Other Diagnostic Plots {#sec-fantasyPointsByAgeExperienceModelsOtherDiagnosticPlots}
+
+```{r}
+performance::check_model(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+```
+
 ##### Generalized Additive Model {#sec-fantasyPointsByAgeExperienceModelsAgeGAM}

### 2025-04-18 mixed-models.qmd
20250418 - fix plots
20250418 - figure labels

@@
 ```{r}
-plot(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+#| label: fig-residualPlotMixedModel
+#| fig-cap: "Plot of Residuals."
+#| fig-alt: "Plot of Residuals."
 
@@
 ```{r}
+#| label: fig-scaleLocationPlotMixedModel
+#| fig-cap: "Scale-Location Plot."
+#| fig-alt: "Scale-Location Plot."
+
 plot(
@@
 ```{r}
+#| label: fig-qqPlotMixedModel
+#| fig-cap: "Quantile–Quantile (QQ) Plot."
+#| fig-alt: "Quantile–Quantile (QQ) Plot."
+
 qqnorm(resid(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience))
@@
 
-##### Probability–Probability (PP) Plots {#sec-fantasyPointsByAgeExperienceModelsDiagnosticPPPlots}
-
-```{r}
+##### Probability–Probability (PP) Plot {#sec-fantasyPointsByAgeExperienceModelsDiagnosticPPPlots}
+
+```{r}
+#| label: fig-qqPlotMixedModel
+#| fig-cap: "Probability–Probability (PP) Plot."
+#| fig-alt: "Probability–Probability (PP) Plot."
+
 petersenlab::ppPlot(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
@@
 ```{r}
+#| label: fig-residualsVsLeveragePlotMixedModel
+#| fig-cap: "Residuals Versus Leverage Plot."
+#| fig-alt: "Residuals Versus Leverage Plot."
+
 plot(
@@
 ```{r}
+#| label: fig-otherDiagnosticPlotsMixedModel
+#| fig-cap: "Other Diagnostic Plots."
+#| fig-alt: "Other Diagnostic Plots."
+#| fig-height: 10
+
 performance::check_model(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)

### 2025-04-18 base-rates.qmd
20250418 - rookie performance

@@
 library("petersenlab")
+library("psych")
+library("tidyverse")
 ```
+
+### Load Data {#sec-baseRatesLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_depthCharts.RData")
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+```
+
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
 
@@
 
+We examine the base rates of rookie performance compared to performance of non-rookies.
+Rookies who play tend to be better players who were drafted in early rounds of the National Football League (NFL) draft.
+Thus, to more fairly compare rookies and non-rookies, we examine only players who were drafted in the first two rounds of the NFL Draft.
+
+```{r}
+rookiesDraftedEarly <- player_stats_seasonal %>% 
+  filter(years_of_experience == 0) %>% 
+  filter(draftround %in% c(1,2)) %>% 
+  arrange(player_id, season) %>% 
+  group_by(player_id) %>% 
+  slice_head(n = 1) %>% 
+  ungroup() %>% 
+  arrange(player_display_name)
+
+nonRookiesDraftedEarly <- player_stats_seasonal %>% 
+  filter(years_of_experience > 0) %>% 
+  filter(draftround %in% c(1,2)) %>% 
+  arrange(player_display_name)
+```
+
+[EXAMINE FOR ROOKIES WHO WERE DRAFTED IN FIRST TWO ROUNDS OF NFL DRAFT]
+
+[ALSO, FOR A SECONDARY ANALYSIS, SUBSET TO PLAYERS WHO PLAYED A FULL SEASON]
+
 ### Quarterbacks {#sec-baseRateRookiePerformanceQBs}
 
+```{r}
+nonRookiesDraftedEarly %>% 
+  filter(position == "QB") %>%
+  select(fantasyPoints) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "QB") %>%
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+rookiesDraftedEarly %>% 
+  filter(position == "QB") %>%
+  select(fantasyPoints) %>% 
+  hist()
+```
+
+Among Quarterbacks who played a full season:
+
+```{r}
+nonRookiesDraftedEarly %>% 
+  filter(position == "QB") %>%
+  filter(games >= 16) %>% 
+  select(fantasyPoints) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "QB") %>%
+  filter(games >= 16) %>% 
+  select(fantasyPoints) %>% 
+  psych::describe()
+```
+
 ### Running Backs {#sec-baseRateRookiePerformanceRBs}
+
+```{r}
+nonRookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+nonRookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
+  select(rushing_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+rookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
+  select(fantasyPoints) %>% 
+  hist()
+
+rookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
+  select(rushing_tds) %>% 
+  psych::describe()
+```
+
+Among Running Backs who played a full season:
+
+```{r}
+nonRookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
+  filter(games >= 16) %>% 
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+nonRookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
+  filter(games >= 16) %>% 
+  select(rushing_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
+  filter(games >= 16) %>% 
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+rookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
+  filter(games >= 16) %>% 
+  select(rushing_tds) %>% 
+  psych::describe()
+```
+
+### Wide Receivers {#sec-baseRateRookiePerformanceWRs}
+
+```{r}
+nonRookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+nonRookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+rookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
+  select(fantasyPoints) %>% 
+  hist()
+
+rookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+Among Wide Receivers who played a full season:
+
+```{r}
+nonRookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
+  filter(games >= 16) %>% 
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+nonRookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
+  filter(games >= 16) %>% 
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
+  filter(games >= 16) %>% 
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+rookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
+  filter(games >= 16) %>% 
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+### Tight Ends {#sec-baseRateRookiePerformanceTEs}
+
+```{r}
+nonRookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+nonRookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+rookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
+  select(fantasyPoints) %>% 
+  hist()
+
+rookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+Among Tight Ends who played a full season:
+
+```{r}
+nonRookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
+  filter(games >= 16) %>% 
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+nonRookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
+  filter(games >= 16) %>% 
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
+  filter(games >= 16) %>% 
+  select(fantasyPoints) %>% 
+  psych::describe()
+
+rookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
+  filter(games >= 16) %>% 
+  select(receiving_tds) %>% 
+  psych::describe()
+```
 

### 2025-04-18 base-rates.qmd
20250418 - fixes

@@
   psych::describe()
-
+```
+
+```{r}
 rookiesDraftedEarly %>% 
@@
   filter(position == "RB") %>%
+  select(rushing_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
   select(fantasyPoints) %>% 
   hist()
-
-rookiesDraftedEarly %>% 
-  filter(position == "RB") %>%
-  select(rushing_tds) %>% 
-  psych::describe()
 ```
@@
   filter(position == "WR") %>%
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
   select(fantasyPoints) %>% 
   hist()
-
-rookiesDraftedEarly %>% 
-  filter(position == "WR") %>%
-  select(receiving_tds) %>% 
-  psych::describe()
 ```
@@
   filter(position == "TE") %>%
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+rookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
   select(fantasyPoints) %>% 
   hist()
-
-rookiesDraftedEarly %>% 
-  filter(position == "TE") %>%
-  select(receiving_tds) %>% 
-  psych::describe()
 ```

### 2025-04-18 base-rates.qmd
20250418 - figure captions

--- 2025-04-18 base-rates.qmd
@@
   psych::describe()
+```
+
+```{r}
+#| label: fig-histogramRookieQBs
+#| fig-cap: "Histogram of Fantasy Points Among Rookie Quarterbacks Who Were Drafted in the First Two Rounds of the National Football League Draft."
+#| fig-alt: "Histogram of Fantasy Points Among Rookie Quarterbacks Who Were Drafted in the First Two Rounds of the National Football League Draft."
 
@@
   filter(position == "RB") %>%
+  select(rushing_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+#| label: fig-histogramRookieRBs
+#| fig-cap: "Histogram of Fantasy Points Among Rookie Running Backs Who Were Drafted in the First Two Rounds of the National Football League Draft."
+#| fig-alt: "Histogram of Fantasy Points Among Rookie Running Backs Who Were Drafted in the First Two Rounds of the National Football League Draft."
+
+rookiesDraftedEarly %>% 
+  filter(position == "RB") %>%
   select(fantasyPoints) %>% 
   hist()
-
-rookiesDraftedEarly %>% 
-  filter(position == "RB") %>%
-  select(rushing_tds) %>% 
-  psych::describe()
 ```
@@
   filter(position == "WR") %>%
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+#| label: fig-histogramRookieWRs
+#| fig-cap: "Histogram of Fantasy Points Among Rookie Wide Receivers Who Were Drafted in the First Two Rounds of the National Football League Draft."
+#| fig-alt: "Histogram of Fantasy Points Among Rookie Wide Receivers Who Were Drafted in the First Two Rounds of the National Football League Draft."
+
+rookiesDraftedEarly %>% 
+  filter(position == "WR") %>%
   select(fantasyPoints) %>% 
   hist()
-
-rookiesDraftedEarly %>% 
-  filter(position == "WR") %>%
-  select(receiving_tds) %>% 
-  psych::describe()
 ```
@@
   filter(position == "TE") %>%
+  select(receiving_tds) %>% 
+  psych::describe()
+```
+
+```{r}
+#| label: fig-histogramRookieTEs
+#| fig-cap: "Histogram of Fantasy Points Among Rookie Tight Ends Who Were Drafted in the First Two Rounds of the National Football League Draft."
+#| fig-alt: "Histogram of Fantasy Points Among Rookie Tight Ends Who Were Drafted in the First Two Rounds of the National Football League Draft."
+
+rookiesDraftedEarly %>% 
+  filter(position == "TE") %>%
   select(fantasyPoints) %>% 
   hist()
-
-rookiesDraftedEarly %>% 
-  filter(position == "TE") %>%
-  select(receiving_tds) %>% 
-  psych::describe()
 ```

### 2025-04-19 base-rates.qmd
20250419 - density plot of rookies vs non-rookies by position

@@
 
-load(file = "./data/nfl_depthCharts.RData")
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
@@
   ungroup() %>% 
+  mutate(
+    rookie = 0,
+    rookieLabel = "Rookie"
+  ) %>% 
   arrange(player_display_name)
@@
   filter(draftround %in% c(1,2)) %>% 
+  mutate(
+    rookie = 1,
+    rookieLabel = "Non-rookie"
+  ) %>% 
   arrange(player_display_name)
-```
-
-[EXAMINE FOR ROOKIES WHO WERE DRAFTED IN FIRST TWO ROUNDS OF NFL DRAFT]
-
-[ALSO, FOR A SECONDARY ANALYSIS, SUBSET TO PLAYERS WHO PLAYED A FULL SEASON]
+
+playersDraftedEarly <- bind_rows(
+  rookiesDraftedEarly,
+  nonRookiesDraftedEarly
+) %>% 
+  mutate(rookieLabel = factor(rookieLabel, levels = c("Rookie","Non-rookie")))
+```
+
+```{r}
+#| label: fig-densityPlotRookieVsNonRookieByPosition
+#| fig-cap: "Density Plot of Fantasy Points for Rookies Versus Non-Rookies by Position Among Players Who Were Drafted in the First Two Rounds of the National Football League Draft."
+#| fig-alt: "Density Plot of Fantasy Points for Rookies Versus Non-Rookies by Position Among Players Who Were Drafted in the First Two Rounds of the National Football League Draft."
+
+ggplot2::ggplot(
+  data = playersDraftedEarly %>% filter(position %in% c("QB","RB","WR","TE")),
+  aes(
+    x = fantasyPoints)) + 
+  geom_density(
+    aes(
+      fill = rookieLabel),
+    alpha = 0.5) +
+  facet_wrap( ~ position) +
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Density Plot of Fantasy Points for Rookies Versus Non-Rookies by Position"
+  ) +
+  theme_bw() + 
+  theme(
+    axis.title.y = element_text(angle = 0, vjust = 0.5), # horizontal y-axis title
+    legend.title = element_blank()) # remove legend title
+```
 

### 2025-04-19 evaluating-prediction-accuracy.qmd
20250419 - calibration plots

@@
 library("viridis")
+library("viridisLite")
+library("msir")
+```
+
+### Load Data {#sec-predictionAccuracyLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projections_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projections_seasonal.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+load(file = "./data/players_projections_weekly.RData")
+load(file = "./data/players_projections_seasonal.RData")
 ```
@@
 Test bias, including how to identify and address it, is described in [@PetersenPrinciplesPsychAssessment](https://isaactpetersen.github.io/Principles-Psychological-Assessment/bias.html).
+
+## Accuracy of Fantasy Projections {#sec-accuracyProjections}
+
+### Seasonal Projections {#sec-accuracyProjectionsSeasonal}
+
+To evaluate the accuracy of projections, we must first merge projections with actual performance.
+
+```{r}
+player_stats_seasonal_subset <- player_stats_seasonal %>% 
+  filter(!is.na(player_id))
+
+nfl_playerIDs_subset <- nfl_playerIDs %>% 
+  filter(!is.na(gsis_id)) %>% 
+  distinct(gsis_id, .keep_all = TRUE) %>% # keep only rows that do not have duplicate values of gsis_id
+  select(-all_of(c("team", "position", "height", "weight", "age")))
+
+players_projectedPoints_seasonal_combined$season <- as.integer(players_projectedPoints_seasonal_combined$season)
+
+player_stats_seasonal_subset_IDs <- left_join(
+  player_stats_seasonal_subset,
+  nfl_playerIDs_subset,
+  by = c("player_id" = "gsis_id")
+) %>% 
+  filter(!is.na(mfl_id))
+
+projectionsWithActuals_seasonal <- full_join(
+  player_stats_seasonal_subset_IDs,
+  players_projectedPoints_seasonal_combined,
+  by = c("mfl_id" = "id", "season"),
+  suffix = c("", "_proj"),
+)
+
+projectionsWithActuals_seasonal <- projectionsWithActuals_seasonal %>% 
+  unite(
+    "player_id_season",
+    player_id,
+    season,
+    remove = FALSE
+  )
+
+playersWithHighProjectedOrActualPoints <- projectionsWithActuals_seasonal %>% 
+  filter(raw_points > 100 | fantasyPoints > 100) %>% 
+  select(player_id_season) %>% 
+  pull()
+
+projectionsWithActuals_seasonal_qb <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "QB")
+
+projectionsWithActuals_seasonal_rb <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "RB")
+
+projectionsWithActuals_seasonal_wr <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "WR")
+
+projectionsWithActuals_seasonal_te <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "TE")
+
+projectionsWithActuals_seasonal_k <- projectionsWithActuals_seasonal %>% 
+  filter(position == "K")
+
+projectionsWithActuals_seasonal_dl <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "DL")
+
+projectionsWithActuals_seasonal_lb <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "LB")
+
+projectionsWithActuals_seasonal_db <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "DB")
+```
+
+```{r}
+player_stats_seasonal_subset_IDs %>% 
+  filter(!is.na(gsis_id)) %>% 
+  group_by(gsis_id) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+#### Overall {#sec-accuracyProjectionsSeasonalOverall}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal$raw_points,
+  actual = projectionsWithActuals_seasonal$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotSeasonalOverall
+#| fig-cap: "Calibration Plot for Seasonal Projections.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Seasonal Projections.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 8
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_seasonal$raw_points,
+  y = projectionsWithActuals_seasonal$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMax <- max(c(
+  projectionsWithActuals_seasonal$raw_points,
+  projectionsWithActuals_seasonal$fantasyPoints), na.rm = TRUE) + 50
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_seasonal,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(0, plotRangeMax),
+    expand = FALSE) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "loess",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of LOESS model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Seasonal Projections"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+#### Quarterbacks {#sec-accuracyProjectionsSeasonalQB}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_qb$raw_points,
+  actual = projectionsWithActuals_seasonal_qb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotSeasonalQBs
+#| fig-cap: "Calibration Plot for Seasonal Projections of Quarterbacks.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Seasonal Projections of Quarterbacks.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 8
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_seasonal_qb$raw_points,
+  y = projectionsWithActuals_seasonal_qb$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMax <- max(c(
+  projectionsWithActuals_seasonal_qb$raw_points,
+  projectionsWithActuals_seasonal_qb$fantasyPoints), na.rm = TRUE) + 50
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_seasonal_qb,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(0, plotRangeMax),
+    expand = FALSE) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "loess",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of LOESS model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Seasonal Projections of Quarterbacks"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_qb$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_qb$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+#### Running Backs {#sec-accuracyProjectionsSeasonalRB}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_rb$raw_points,
+  actual = projectionsWithActuals_seasonal_rb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotSeasonalRBs
+#| fig-cap: "Calibration Plot for Seasonal Projections of Running Backs.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Seasonal Projections of Running Backs.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 8
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_seasonal_rb$raw_points,
+  y = projectionsWithActuals_seasonal_rb$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMax <- max(c(
+  projectionsWithActuals_seasonal_rb$raw_points,
+  projectionsWithActuals_seasonal_rb$fantasyPoints), na.rm = TRUE) + 0
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_seasonal_rb,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(0, plotRangeMax),
+    expand = FALSE) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "loess",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of LOESS model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Seasonal Projections of Running Backs"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_rb$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_rb$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+#### Wide Receivers {#sec-accuracyProjectionsSeasonalWR}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_wr$raw_points,
+  actual = projectionsWithActuals_seasonal_wr$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotSeasonalWRs
+#| fig-cap: "Calibration Plot for Seasonal Projections of Wide Receivers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Seasonal Projections of Wide Receivers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 8
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_seasonal_wr$raw_points,
+  y = projectionsWithActuals_seasonal_wr$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMax <- max(c(
+  projectionsWithActuals_seasonal_wr$raw_points,
+  projectionsWithActuals_seasonal_wr$fantasyPoints), na.rm = TRUE) + 40
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_seasonal_wr,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(0, plotRangeMax),
+    expand = FALSE) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "loess",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of LOESS model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Seasonal Projections of Wide Receivers"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_wr$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_wr$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+#### Tight Ends {#sec-accuracyProjectionsSeasonalTE}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_te$raw_points,
+  actual = projectionsWithActuals_seasonal_te$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotSeasonalTEs
+#| fig-cap: "Calibration Plot for Seasonal Projections of Tight Ends.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Seasonal Projections of Tight Ends.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 8
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_seasonal_te$raw_points,
+  y = projectionsWithActuals_seasonal_te$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMax <- max(c(
+  projectionsWithActuals_seasonal_te$raw_points,
+  projectionsWithActuals_seasonal_te$fantasyPoints), na.rm = TRUE) + 0
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_seasonal_te,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(0, plotRangeMax),
+    expand = FALSE) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "loess",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of LOESS model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Seasonal Projections of Tight Ends"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_te$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_te$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+#### Kickers {#sec-accuracyProjectionsSeasonalK}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_k$raw_points,
+  actual = projectionsWithActuals_seasonal_k$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotSeasonalKs
+#| fig-cap: "Calibration Plot for Seasonal Projections of Kickers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Seasonal Projections of Kickers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 8
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_seasonal_k$raw_points,
+  y = projectionsWithActuals_seasonal_k$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMax <- max(c(
+  projectionsWithActuals_seasonal_k$raw_points,
+  projectionsWithActuals_seasonal_k$fantasyPoints), na.rm = TRUE) + 100
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_seasonal_k,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(0, plotRangeMax),
+    expand = FALSE) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "loess",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of LOESS model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Seasonal Projections of Kickers"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_k$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_k$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+#### Defensive Linemen {#sec-accuracyProjectionsSeasonalDL}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_dl$raw_points,
+  actual = projectionsWithActuals_seasonal_dl$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotSeasonalDLs
+#| fig-cap: "Calibration Plot for Seasonal Projections of Defensive Linemen.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Seasonal Projections of Defensive Linemen.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 8
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_seasonal_dl$raw_points,
+  y = projectionsWithActuals_seasonal_dl$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMax <- max(c(
+  projectionsWithActuals_seasonal_dl$raw_points,
+  projectionsWithActuals_seasonal_dl$fantasyPoints), na.rm = TRUE) + 0
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_seasonal_dl,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(0, plotRangeMax),
+    expand = FALSE) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "loess",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of LOESS model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Seasonal Projections of Defensive Linemen"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_dl$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_dl$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+#### Linebackers {#sec-accuracyProjectionsSeasonalDL}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_lb$raw_points,
+  actual = projectionsWithActuals_seasonal_lb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotSeasonalLBs
+#| fig-cap: "Calibration Plot for Seasonal Projections of Linebackers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Seasonal Projections of Linebackers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 8
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_seasonal_lb$raw_points,
+  y = projectionsWithActuals_seasonal_lb$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMax <- max(c(
+  projectionsWithActuals_seasonal_lb$raw_points,
+  projectionsWithActuals_seasonal_lb$fantasyPoints), na.rm = TRUE) + 0
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_seasonal_lb,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(0, plotRangeMax),
+    expand = FALSE) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "loess",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of LOESS model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Seasonal Projections of Linebackers"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_lb$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_lb$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+#### Defensive Backs {#sec-accuracyProjectionsSeasonalDB}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_db$raw_points,
+  actual = projectionsWithActuals_seasonal_db$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotSeasonalDBs
+#| fig-cap: "Calibration Plot for Seasonal Projections of Defensive Backs.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Seasonal Projections of Defensive Backs.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 8
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_seasonal_db$raw_points,
+  y = projectionsWithActuals_seasonal_db$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMax <- max(c(
+  projectionsWithActuals_seasonal_db$raw_points,
+  projectionsWithActuals_seasonal_db$fantasyPoints), na.rm = TRUE) + 0
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_seasonal_db,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(0, plotRangeMax),
+    expand = FALSE) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "loess",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of LOESS model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Seasonal Projections of Defensive Backs"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_db$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_db$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+### Weekly Projections {#sec-accuracyProjectionsWeekly}
+
+To evaluate the accuracy of projections, we must first merge projections with actual performance.
+
+```{r}
+
+```
 

### 2025-04-19 evaluating-prediction-accuracy.qmd
20250419 - calibration plot tweaks

@@
 
-plotRangeMax <- max(c(
-  projectionsWithActuals_seasonal$raw_points,
-  projectionsWithActuals_seasonal$fantasyPoints), na.rm = TRUE) + 50
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
 
@@
     xlim = c(0, plotRangeMax),
-    ylim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
     expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 100)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 100)) +
   geom_line(
@@
 
-plotRangeMax <- max(c(
-  projectionsWithActuals_seasonal_qb$raw_points,
-  projectionsWithActuals_seasonal_qb$fantasyPoints), na.rm = TRUE) + 50
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
 
@@
 
-plotRangeMax <- max(c(
-  projectionsWithActuals_seasonal_rb$raw_points,
-  projectionsWithActuals_seasonal_rb$fantasyPoints), na.rm = TRUE) + 0
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
 
@@
 
-plotRangeMax <- max(c(
-  projectionsWithActuals_seasonal_wr$raw_points,
-  projectionsWithActuals_seasonal_wr$fantasyPoints), na.rm = TRUE) + 40
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
 
@@
 
-plotRangeMax <- max(c(
-  projectionsWithActuals_seasonal_te$raw_points,
-  projectionsWithActuals_seasonal_te$fantasyPoints), na.rm = TRUE) + 0
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
 
@@
     xlim = c(0, plotRangeMax),
-    ylim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
     expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 50)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 50)) +
   geom_line(
@@
 
-plotRangeMax <- max(c(
-  projectionsWithActuals_seasonal_k$raw_points,
-  projectionsWithActuals_seasonal_k$fantasyPoints), na.rm = TRUE) + 100
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
 
@@
 
-plotRangeMax <- max(c(
-  projectionsWithActuals_seasonal_dl$raw_points,
-  projectionsWithActuals_seasonal_dl$fantasyPoints), na.rm = TRUE) + 0
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
 
@@
     xlim = c(0, plotRangeMax),
-    ylim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
     expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 20)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 20)) +
   geom_line(
@@
 
-plotRangeMax <- max(c(
-  projectionsWithActuals_seasonal_lb$raw_points,
-  projectionsWithActuals_seasonal_lb$fantasyPoints), na.rm = TRUE) + 0
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
 
@@
 
-plotRangeMax <- max(c(
-  projectionsWithActuals_seasonal_db$raw_points,
-  projectionsWithActuals_seasonal_db$fantasyPoints), na.rm = TRUE) + 0
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
 

### 2025-04-19 evaluating-prediction-accuracy.qmd
20250419 - fix

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projections_seasonal.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_playerIDs.RData", fsep = ""))
 ```
@@
 load(file = "./data/players_projections_seasonal.RData")
+load(file = "./data/nfl_playerIDs.RData")
 ```

### 2025-04-20 evaluating-prediction-accuracy.qmd
20250419 - fixes

@@
 
-```{r}
-player_stats_seasonal_subset_IDs %>% 
-  filter(!is.na(gsis_id)) %>% 
-  group_by(gsis_id) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
 #### Overall {#sec-accuracyProjectionsSeasonalOverall}
@@
 The dashed red lines are the 95% prediction interval of actual fantasy points."
-#| fig-width: 8
-#| fig-height: 8
+#| fig-width: 6
+#| fig-height: 6
 
@@
 
-#### Linebackers {#sec-accuracyProjectionsSeasonalDL}
+#### Linebackers {#sec-accuracyProjectionsSeasonalLB}
 

### 2025-04-20 multiple-regression.qmd
20250418 - fix plot order
20250419 - fixes

@@
 plot(
+  x = quadraticX,
+  y = linearY,
+  xlab = "",
+  ylab = "",
+  main = "Linear Association")
+
+abline(
+  lm(linearY ~ quadraticX),
+  lwd = 2,
+  col = "blue")
+
+plot(
   x = quadraticData$quadraticX,
@@
   quadraticNewData$quadraticY ~ quadraticNewData$quadraticX,
-  lwd = 2,
-  col = "blue")
-
-plot(
-  x = quadraticX,
-  y = linearY,
-  xlab = "",
-  ylab = "",
-  main = "Linear Association")
-
-abline(
-  lm(linearY ~ quadraticX),
   lwd = 2,
@@
 
-To evaluate homoscedasticity, we can evaluate a residual plot (@fig-residualPlots) and spread-level plot (@fig-spreadLevelPlots) from the `car` package [@Fox2019_packages; @R-car].
+To evaluate homoscedasticity, we can evaluate a residual plot (@fig-residualPlots) and spread-level plot (@fig-spreadLevelPlot) from the `car` package [@Fox2019_packages; @R-car].
 In a residual plot, you want a constant spread of residuals versus fitted values—you do not want the residuals to show a fan or cone shape.

### 2025-04-21 evaluating-prediction-accuracy.qmd
20250419 - figure sizing
20250420 - cleanup

@@
 library("petersenlab")
-library("tidyverse")
 library("pROC")
@@
 library("msir")
+library("tidyverse")
 ```
@@
 The dashed red lines are the 95% prediction interval of actual fantasy points."
-#| fig-width: 6
+#| fig-width: 8
 #| fig-height: 6

### 2025-04-21 evaluating-prediction-accuracy.qmd
20250421 - accuracy of weekly projections

@@
 ```{r}
-
+player_stats_weekly_subset <- player_stats_weekly %>% 
+  filter(!is.na(player_id))
+
+nfl_playerIDs_subset <- nfl_playerIDs %>% 
+  filter(!is.na(gsis_id)) %>% 
+  distinct(gsis_id, .keep_all = TRUE) %>% # keep only rows that do not have duplicate values of gsis_id
+  select(-all_of(c("team", "position", "height", "weight", "age")))
+
+players_projectedPoints_weekly_combined$season <- as.integer(players_projectedPoints_weekly_combined$season)
+
+player_stats_weekly_subset_IDs <- left_join(
+  player_stats_weekly_subset,
+  nfl_playerIDs_subset,
+  by = c("player_id" = "gsis_id")
+) %>% 
+  filter(!is.na(mfl_id))
+
+projectionsWithActuals_weekly <- full_join(
+  player_stats_weekly_subset_IDs,
+  players_projectedPoints_weekly_combined,
+  by = c("mfl_id" = "id", "season", "week"),
+  suffix = c("", "_proj"),
+)
+
+projectionsWithActuals_weekly <- projectionsWithActuals_weekly %>% 
+  unite(
+    "player_id_season_week",
+    player_id,
+    season,
+    week,
+    remove = FALSE
+  )
+
+playersWithHighProjectedOrActualPoints_weekly <- projectionsWithActuals_weekly %>% 
+  filter(raw_points > 6 | fantasyPoints > 6) %>% 
+  select(player_id_season_week) %>% 
+  pull()
+
+projectionsWithActuals_weekly_qb <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "QB")
+
+projectionsWithActuals_weekly_rb <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "RB")
+
+projectionsWithActuals_weekly_wr <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "WR")
+
+projectionsWithActuals_weekly_te <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "TE")
+
+projectionsWithActuals_weekly_k <- projectionsWithActuals_weekly %>% 
+  filter(position == "K")
+
+projectionsWithActuals_weekly_dl <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "DL")
+
+projectionsWithActuals_weekly_lb <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "LB")
+
+projectionsWithActuals_weekly_db <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "DB")
+```
+
+#### Overall {#sec-accuracyProjectionsWeeklyOverall}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly$raw_points,
+  actual = projectionsWithActuals_weekly$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotWeeklyOverall
+#| fig-cap: "Calibration Plot for Weekly Projections.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Weekly Projections.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 6
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_weekly$raw_points,
+  y = projectionsWithActuals_weekly$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_weekly,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
+    expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "gam",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of GAM model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Weekly Projections"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+#### Quarterbacks {#sec-accuracyProjectionsWeeklyQB}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_qb$raw_points,
+  actual = projectionsWithActuals_weekly_qb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotWeeklyQBs
+#| fig-cap: "Calibration Plot for Weekly Projections of Quarterbacks.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Weekly Projections of Quarterbacks.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 6
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_weekly_qb$raw_points,
+  y = projectionsWithActuals_weekly_qb$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_weekly_qb,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
+    expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "gam",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of GAM model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Weekly Projections of Quarterbacks"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_qb$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_qb$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+#### Running Backs {#sec-accuracyProjectionsWeeklyRB}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_rb$raw_points,
+  actual = projectionsWithActuals_weekly_rb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotWeeklyRBs
+#| fig-cap: "Calibration Plot for Weekly Projections of Running Backs.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Weekly Projections of Running Backs.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 6
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_weekly_rb$raw_points,
+  y = projectionsWithActuals_weekly_rb$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_weekly_rb,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
+    expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "gam",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of GAM model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Weekly Projections of Running Backs"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_rb$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_rb$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+#### Wide Receivers {#sec-accuracyProjectionsWeeklyWR}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_wr$raw_points,
+  actual = projectionsWithActuals_weekly_wr$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotWeeklyWRs
+#| fig-cap: "Calibration Plot for Weekly Projections of Wide Receivers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Weekly Projections of Wide Receivers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 6
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_weekly_wr$raw_points,
+  y = projectionsWithActuals_weekly_wr$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_weekly_wr,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
+    expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "gam",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of GAM model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Weekly Projections of Wide Receivers"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_wr$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_wr$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+#### Tight Ends {#sec-accuracyProjectionsWeeklyTE}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_te$raw_points,
+  actual = projectionsWithActuals_weekly_te$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotWeeklyTEs
+#| fig-cap: "Calibration Plot for Weekly Projections of Tight Ends.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Weekly Projections of Tight Ends.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 6
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_weekly_te$raw_points,
+  y = projectionsWithActuals_weekly_te$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_weekly_te,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
+    expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "gam",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of GAM model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Weekly Projections of Tight Ends"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_te$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_te$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+#### Kickers {#sec-accuracyProjectionsWeeklyK}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_k$raw_points,
+  actual = projectionsWithActuals_weekly_k$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotWeeklyKs
+#| fig-cap: "Calibration Plot for Weekly Projections of Kickers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Weekly Projections of Kickers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 6
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_weekly_k$raw_points,
+  y = projectionsWithActuals_weekly_k$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_weekly_k,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
+    expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "gam",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of GAM model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Weekly Projections of Kickers"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_k$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_k$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+#### Defensive Linemen {#sec-accuracyProjectionsWeeklyDL}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_dl$raw_points,
+  actual = projectionsWithActuals_weekly_dl$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotWeeklyDLs
+#| fig-cap: "Calibration Plot for Weekly Projections of Defensive Linemen.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Weekly Projections of Defensive Linemen.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 6
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_weekly_dl$raw_points,
+  y = projectionsWithActuals_weekly_dl$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_weekly_dl,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
+    expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "gam",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of GAM model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Weekly Projections of Defensive Linemen"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_dl$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_dl$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+#### Linebackers {#sec-accuracyProjectionsWeeklyLB}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_lb$raw_points,
+  actual = projectionsWithActuals_weekly_lb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotWeeklyLBs
+#| fig-cap: "Calibration Plot for Weekly Projections of Linebackers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Weekly Projections of Linebackers.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 6
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_weekly_lb$raw_points,
+  y = projectionsWithActuals_weekly_lb$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_weekly_lb,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
+    expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "gam",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of GAM model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Weekly Projections of Linebackers"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_lb$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_lb$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+#### Defensive Backs {#sec-accuracyProjectionsWeeklyDB}
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_db$raw_points,
+  actual = projectionsWithActuals_weekly_db$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-calibrationPlotWeeklyDBs
+#| fig-cap: "Calibration Plot for Weekly Projections of Defensive Backs.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-alt: "Calibration Plot for Weekly Projections of Defensive Backs.
+The yellow line is the best-fit line based on LOESS.
+The purple band is the 95% confidence interval of actual fantasy points.
+The dashed red lines are the 95% prediction interval of actual fantasy points."
+#| fig-width: 8
+#| fig-height: 6
+
+#95 prediction interval based on LOESS model
+calibrationLoessModel <- msir::loess.sd(
+  x = projectionsWithActuals_weekly_db$raw_points,
+  y = projectionsWithActuals_weekly_db$fantasyPoints,
+  nsigma = qnorm(.975),
+  na.action = "na.exclude")
+
+calibrationLoessPredictionInterval <- data.frame(
+  x = calibrationLoessModel$x,
+  y = calibrationLoessModel$y,
+  lower = calibrationLoessModel$lower,
+  upper = calibrationLoessModel$upper)
+
+plotRangeMin <- min(calibrationLoessModel$lower, na.rm = TRUE)
+
+plotRangeMax <- max(calibrationLoessModel$upper, na.rm = TRUE)
+
+ggplot2::ggplot(
+  data = projectionsWithActuals_weekly_db,
+  aes(
+    x = raw_points,
+    y = fantasyPoints)) + 
+  coord_cartesian(
+    xlim = c(0, plotRangeMax),
+    ylim = c(plotRangeMin, plotRangeMax),
+    expand = FALSE) +
+  scale_x_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  scale_y_continuous(breaks = seq(0, plotRangeMax, by = 5)) +
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = lower),
+    color = "red",
+    linetype = "dashed") + #Lower estimate of 95% prediction interval of linear model
+  geom_line(
+    data = calibrationLoessPredictionInterval,
+    aes(
+      x = x,
+      y = upper),
+    color = "red",
+    linetype = "dashed") + #Upper estimate of 95% prediction interval of linear model
+  geom_smooth(
+    method = "gam",
+    color = viridisLite::viridis(3)[3],
+    fill = viridisLite::viridis(3)[1],
+    alpha = 0.7) + #95% confidence interval of GAM model
+  geom_abline(
+    slope = 1,
+    intercept = 0) +
+  labs(
+    x = "Projected Fantasy Points",
+    y = "Actual Fantasy Points",
+    title = "Calibration Plot for Weekly Projections of Defensive Backs"
+  ) +
+  theme_classic() + 
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_db$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_db$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
 ```

### 2025-04-21 download-football-data.qmd
20250421 - nflfastR package and target share

@@
 
-Below, we provide examples for how to download various types of NFL data using the `nflreadr` package [@R-nflreadr].
+Below, we provide examples for how to download various types of NFL data using the `nflreadr` [@R-nflreadr] and `nflfastR` [@R-nflfastR] packages.
 For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
 We save each data file after downloading it, so we can use the data in subsequent chapters.
-If you have difficulty downloading the data files using the `nflreadr` package [@R-nflreadr], we also saved the data files so they are publicly available on the Open Science Framework: <https://osf.io/z6pg4>.
+If you have difficulty downloading the data files using the `nflreadr` [@R-nflreadr] or `nflfastR` [@R-nflfastR] packages, we also saved the data files so they are publicly available on the Open Science Framework: <https://osf.io/z6pg4>.
 
@@
 
-First, we can compute the players' season-by-season statistics.
+First, we can compute the players' season-by-season statistics using the `nflfastR` [@R-nflfastR] package.
 

### 2025-04-21 multiple-regression.qmd
20250421 - nflfastR package and target share

@@
 We have only a few [predictors](#sec-correlationalStudy) and our sample size is large enough such that [overfitting](#sec-overfitting) is not likely a concern.
+
+On a weekly basis, target share is computed as the number of targets a player receives divided by the team's total number of targets.
+It would be nice to compute it this way for seasonal data, as well; however, for calculating seasonal target share, the `nflfastR` [@R-nflfastR] package currently sums up the player's weekly target share.
+Thus, seasonal target share (unlike weekly target share)—as it is currently calculated—does not range from 0–1.
+For example, if a player has a weekly target share of 15%, that would be a seasonal (cross-week-summed) target share of 2.55 ($0.15 \times 17 \; \text{games} = 2.55$.
 
@@
 We could also generate the model-implied prediction of fantasy points for any value of the [predictor variables](#sec-correlationalStudy).
-For instance, here are the number of fantasy points expected for a Wide Receiver who is 23 years old, is 6'2" tall (72 inches), weighs 200 pounds, and has a 3% target share:
+For instance, here are the number of fantasy points expected for a Wide Receiver who is 23 years old, is 6'2" tall (72 inches), weighs 200 pounds, and has a (cross-week-summed) target share of 3 (which reflects an approximate target share of 17.6% percent per game; i.e., $0.176 = \frac{3}{17 \; \text{games}}$):
 

### 2025-04-21 evaluating-prediction-accuracy.qmd
20250421 - example of prediction of categorical outcome

@@
 Test bias, including how to identify and address it, is described in [@PetersenPrinciplesPsychAssessment](https://isaactpetersen.github.io/Principles-Psychological-Assessment/bias.html).
+
+## Example of Predicting a Categorical outcome {#sec-predictionCategoricalExample}
+
+Let's consider an example of predicting a categorical outcome.
+For this example, let's predict whether or not a Running Back rushes for 200 yards in a game.
+To predict that, we will use the number of rushing yards the player had in the previous game.
+
+```{r}
+player_stats_weekly_example <- player_stats_weekly %>% 
+  filter(position == "RB") %>% 
+  arrange(player_id, season, week) %>% 
+  group_by(player_id, season) %>% 
+  mutate(rushing_yards_priorGame = lag(rushing_yards)) %>% 
+  ungroup()
+```
+
+```{r}
+player_stats_weekly_example %>% 
+  select(player_id, season, week, rushing_yards, rushing_yards_priorGame) %>% 
+  head()
+```
+
+Here are our accuracy indices at a cutoff of 100.
+That is, we predict that each player who rushed more than 100 yards in a given game will rush for 200 yards in their next game.
+
+```{r}
+petersenlab::accuracyAtCutoff(
+  predicted = player_stats_weekly_example$rushing_yards_priorGame,
+  actual = player_stats_weekly_example$rush_200_yds,
+  cutoff = 100
+)
+```
+
+Now, let's lower the cutoff to 50 and see how that impacts our predictive accuracy.
+
+```{r}
+petersenlab::accuracyAtCutoff(
+  predicted = player_stats_weekly_example$rushing_yards_priorGame,
+  actual = player_stats_weekly_example$rush_200_yds,
+  cutoff = 50
+)
+```
+
+Now, let's raise the cutoff to 150 and see how that impacts our predictive accuracy.
+
+```{r}
+petersenlab::accuracyAtCutoff(
+  predicted = player_stats_weekly_example$rushing_yards_priorGame,
+  actual = player_stats_weekly_example$rush_200_yds,
+  cutoff = 150
+)
+```
+
+We can also examine the accuracy at each possible cutoff.
+
+```{r}
+petersenlab::accuracyAtEachCutoff(
+  predicted = player_stats_weekly_example$rushing_yards_priorGame,
+  actual = player_stats_weekly_example$rush_200_yds
+)
+```
+
+We can also create a receiver operating characteristic (ROC) curve, as in @fig-rocCurveExample.
+
+```{r}
+rocCurve <- pROC::roc(
+  data = player_stats_weekly_example,
+  response = rush_200_yds,
+  predictor = rushing_yards_priorGame,
+  smooth = FALSE)
+```
+
+Here is the area under the ROC curve (AUC):
+
+```{r}
+rocCurve$auc
+```
+
+```{r}
+#| label: fig-rocCurveExample
+#| fig-cap: "Example Receiver Operating Characteristic (ROC) Curve."
+#| fig-alt: "Example Receiver Operating Characteristic (ROC) Curve."
+#| fig-width: 8
+#| fig-height: 6
+
+plot(
+  rocCurve,
+  legacy.axes = TRUE,
+  print.auc = TRUE)
+```
 

### 2025-04-23 evaluating-prediction-accuracy.qmd
20250421 - update figure caption for calibration plots
20250423 - reorganize
20250423 - prediction of touchdowns vs yards

@@
 load(file = "./data/nfl_playerIDs.RData")
+```
+
+### Prepare Data {#sec-predictionAccuracyPrepareData}
+
+#### Seasonal Projections {#sec-predictionAccuracyPrepareDataSeasonal}
+
+To evaluate the accuracy of projections, we must first merge projections with actual performance.
+Below, we merge seasonal projections with actual performance.
+
+```{r}
+player_stats_seasonal_subset <- player_stats_seasonal %>% 
+  filter(!is.na(player_id))
+
+nfl_playerIDs_subset <- nfl_playerIDs %>% 
+  filter(!is.na(gsis_id)) %>% 
+  distinct(gsis_id, .keep_all = TRUE) %>% # keep only rows that do not have duplicate values of gsis_id
+  select(-all_of(c("team", "position", "height", "weight", "age")))
+
+players_projectedPoints_seasonal_combined$season <- as.integer(players_projectedPoints_seasonal_combined$season)
+
+player_stats_seasonal_subset_IDs <- left_join(
+  player_stats_seasonal_subset,
+  nfl_playerIDs_subset,
+  by = c("player_id" = "gsis_id")
+) %>% 
+  filter(!is.na(mfl_id))
+
+projectionsWithActuals_seasonal <- full_join(
+  player_stats_seasonal_subset_IDs,
+  players_projectedPoints_seasonal_combined,
+  by = c("mfl_id" = "id", "season"),
+  suffix = c("", "_proj"),
+)
+
+projectionsWithActuals_seasonal <- projectionsWithActuals_seasonal %>% 
+  unite(
+    "player_id_season",
+    player_id,
+    season,
+    remove = FALSE
+  )
+
+playersWithHighProjectedOrActualPoints <- projectionsWithActuals_seasonal %>% 
+  filter(raw_points > 100 | fantasyPoints > 100) %>% 
+  select(player_id_season) %>% 
+  pull()
+
+projectionsWithActuals_seasonal_qb <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "QB")
+
+projectionsWithActuals_seasonal_rb <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "RB")
+
+projectionsWithActuals_seasonal_wr <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "WR")
+
+projectionsWithActuals_seasonal_te <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "TE")
+
+projectionsWithActuals_seasonal_k <- projectionsWithActuals_seasonal %>% 
+  filter(position == "K")
+
+projectionsWithActuals_seasonal_dl <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "DL")
+
+projectionsWithActuals_seasonal_lb <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "LB")
+
+projectionsWithActuals_seasonal_db <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "DB")
+```
+
+### Weekly Projections {#sec-predictionAccuracyPrepareDataWeekly}
+
+Below, we merge weekly projections with actual performance.
+
+```{r}
+player_stats_weekly_subset <- player_stats_weekly %>% 
+  filter(!is.na(player_id))
+
+nfl_playerIDs_subset <- nfl_playerIDs %>% 
+  filter(!is.na(gsis_id)) %>% 
+  distinct(gsis_id, .keep_all = TRUE) %>% # keep only rows that do not have duplicate values of gsis_id
+  select(-all_of(c("team", "position", "height", "weight", "age")))
+
+players_projectedPoints_weekly_combined$season <- as.integer(players_projectedPoints_weekly_combined$season)
+
+player_stats_weekly_subset_IDs <- left_join(
+  player_stats_weekly_subset,
+  nfl_playerIDs_subset,
+  by = c("player_id" = "gsis_id")
+) %>% 
+  filter(!is.na(mfl_id))
+
+projectionsWithActuals_weekly <- full_join(
+  player_stats_weekly_subset_IDs,
+  players_projectedPoints_weekly_combined,
+  by = c("mfl_id" = "id", "season", "week"),
+  suffix = c("", "_proj"),
+)
+
+projectionsWithActuals_weekly <- projectionsWithActuals_weekly %>% 
+  unite(
+    "player_id_season_week",
+    player_id,
+    season,
+    week,
+    remove = FALSE
+  )
+
+playersWithHighProjectedOrActualPoints_weekly <- projectionsWithActuals_weekly %>% 
+  filter(raw_points > 6 | fantasyPoints > 6) %>% 
+  select(player_id_season_week) %>% 
+  pull()
+
+projectionsWithActuals_weekly_qb <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "QB")
+
+projectionsWithActuals_weekly_rb <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "RB")
+
+projectionsWithActuals_weekly_wr <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "WR")
+
+projectionsWithActuals_weekly_te <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "TE")
+
+projectionsWithActuals_weekly_k <- projectionsWithActuals_weekly %>% 
+  filter(position == "K")
+
+projectionsWithActuals_weekly_dl <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "DL")
+
+projectionsWithActuals_weekly_lb <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "LB")
+
+projectionsWithActuals_weekly_db <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "DB")
 ```
@@
 
-[EVALUATE EMPIRICALLY]
+Here is the accuracy of the prediction of passing touchdowns versus passing yards among Quarterbacks:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_qb$pass_tds,
+  actual = projectionsWithActuals_weekly_qb$passing_tds,
+  dropUndefined = TRUE
+)
+
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_qb$pass_yds,
+  actual = projectionsWithActuals_weekly_qb$passing_yards,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+rSquared_passTds <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_qb$pass_tds,
+  actual = projectionsWithActuals_weekly_qb$passing_tds,
+  dropUndefined = TRUE
+)$rsquared
+
+rSquared_passYds <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_qb$pass_yds,
+  actual = projectionsWithActuals_weekly_qb$passing_yards,
+  dropUndefined = TRUE
+)$rsquared
+```
+
+The accuracy for predicting passing yards ($R^2 = `r petersenlab::apa(rSquared_passYds, decimals = 2, leading = FALSE)`$) is higher than the accuracy for predicting passing touchdowns ($R^2 = `r petersenlab::apa(rSquared_passTds, decimals = 2, leading = FALSE)`$).
+
+Here is the accuracy of the prediction of rushing touchdowns versus rushing yards among Running Backs:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_rb$rush_tds,
+  actual = projectionsWithActuals_weekly_rb$rushing_tds,
+  dropUndefined = TRUE
+)
+
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_rb$rush_yds,
+  actual = projectionsWithActuals_weekly_rb$rushing_yards,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+rSquared_rushTds <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_rb$rush_tds,
+  actual = projectionsWithActuals_weekly_rb$rushing_tds,
+  dropUndefined = TRUE
+)$rsquared
+
+rSquared_rushYds <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_rb$rush_yds,
+  actual = projectionsWithActuals_weekly_rb$rushing_yards,
+  dropUndefined = TRUE
+)$rsquared
+```
+
+The accuracy for predicting rushing yards ($R^2 = `r petersenlab::apa(rSquared_rushYds, decimals = 2, leading = FALSE)`$) is higher than the accuracy for predicting rushing touchdowns ($R^2 = `r petersenlab::apa(rSquared_rushTds, decimals = 2, leading = FALSE)`$).
+
+Here is the accuracy of the prediction of receiving touchdowns versus receiving yards among Wide Receivers:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_wr$rec_tds,
+  actual = projectionsWithActuals_weekly_wr$receiving_tds,
+  dropUndefined = TRUE
+)
+
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_wr$rec_yds,
+  actual = projectionsWithActuals_weekly_wr$receiving_yards,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+rSquared_recTds <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_wr$rec_tds,
+  actual = projectionsWithActuals_weekly_wr$receiving_tds,
+  dropUndefined = TRUE
+)$rsquared
+
+rSquared_recYds <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_wr$rec_yds,
+  actual = projectionsWithActuals_weekly_wr$receiving_yards,
+  dropUndefined = TRUE
+)$rsquared
+```
+
+The accuracy for predicting receiving yards ($R^2 = `r petersenlab::apa(rSquared_recYds, decimals = 2, leading = FALSE)`$) is higher than the accuracy for predicting rushing touchdowns ($R^2 = `r petersenlab::apa(rSquared_recTds, decimals = 2, leading = FALSE)`$) among Wide Receivers.
+
+Here is the accuracy of the prediction of receiving touchdowns versus receiving yards among Tight Ends:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_te$rec_tds,
+  actual = projectionsWithActuals_weekly_te$receiving_tds,
+  dropUndefined = TRUE
+)
+
+petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_te$rec_yds,
+  actual = projectionsWithActuals_weekly_te$receiving_yards,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+rSquared_recTds_te <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_te$rec_tds,
+  actual = projectionsWithActuals_weekly_te$receiving_tds,
+  dropUndefined = TRUE
+)$rsquared
+
+rSquared_recYds_te <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_te$rec_yds,
+  actual = projectionsWithActuals_weekly_te$receiving_yards,
+  dropUndefined = TRUE
+)$rsquared
+```
+
+The accuracy for predicting receiving yards ($R^2 = `r petersenlab::apa(rSquared_recYds_te, decimals = 2, leading = FALSE)`$) is higher than the accuracy for predicting rushing touchdowns ($R^2 = `r petersenlab::apa(rSquared_recTds_te, decimals = 2, leading = FALSE)`$) among Tight Ends.
+
+In sum, the proportion of variance explained by the prediction (i.e., $R^2$) is much higher for passing/rushing/receiving yards than it is for touchdowns.
+This is consistent with the notion that prediction accuracy tends to be lower for lower [base rate](#sec-baseRate) events (like touchdowns) compared to higher [base rate](#sec-baseRate) events.
 
@@
 
-To evaluate the accuracy of projections, we must first merge projections with actual performance.
-
-```{r}
-player_stats_seasonal_subset <- player_stats_seasonal %>% 
-  filter(!is.na(player_id))
-
-nfl_playerIDs_subset <- nfl_playerIDs %>% 
-  filter(!is.na(gsis_id)) %>% 
-  distinct(gsis_id, .keep_all = TRUE) %>% # keep only rows that do not have duplicate values of gsis_id
-  select(-all_of(c("team", "position", "height", "weight", "age")))
-
-players_projectedPoints_seasonal_combined$season <- as.integer(players_projectedPoints_seasonal_combined$season)
-
-player_stats_seasonal_subset_IDs <- left_join(
-  player_stats_seasonal_subset,
-  nfl_playerIDs_subset,
-  by = c("player_id" = "gsis_id")
-) %>% 
-  filter(!is.na(mfl_id))
-
-projectionsWithActuals_seasonal <- full_join(
-  player_stats_seasonal_subset_IDs,
-  players_projectedPoints_seasonal_combined,
-  by = c("mfl_id" = "id", "season"),
-  suffix = c("", "_proj"),
-)
-
-projectionsWithActuals_seasonal <- projectionsWithActuals_seasonal %>% 
-  unite(
-    "player_id_season",
-    player_id,
-    season,
-    remove = FALSE
-  )
-
-playersWithHighProjectedOrActualPoints <- projectionsWithActuals_seasonal %>% 
-  filter(raw_points > 100 | fantasyPoints > 100) %>% 
-  select(player_id_season) %>% 
-  pull()
-
-projectionsWithActuals_seasonal_qb <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "QB")
-
-projectionsWithActuals_seasonal_rb <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "RB")
-
-projectionsWithActuals_seasonal_wr <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "WR")
-
-projectionsWithActuals_seasonal_te <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "TE")
-
-projectionsWithActuals_seasonal_k <- projectionsWithActuals_seasonal %>% 
-  filter(position == "K")
-
-projectionsWithActuals_seasonal_dl <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "DL")
-
-projectionsWithActuals_seasonal_lb <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "LB")
-
-projectionsWithActuals_seasonal_db <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "DB")
-```
-
 #### Overall {#sec-accuracyProjectionsSeasonalOverall}
@@
 #| fig-cap: "Calibration Plot for Seasonal Projections.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-alt: "Calibration Plot for Seasonal Projections.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-cap: "Calibration Plot for Seasonal Projections of Quarterbacks.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-alt: "Calibration Plot for Seasonal Projections of Quarterbacks.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-cap: "Calibration Plot for Seasonal Projections of Running Backs.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-alt: "Calibration Plot for Seasonal Projections of Running Backs.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-cap: "Calibration Plot for Seasonal Projections of Wide Receivers.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-alt: "Calibration Plot for Seasonal Projections of Wide Receivers.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-cap: "Calibration Plot for Seasonal Projections of Tight Ends.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-alt: "Calibration Plot for Seasonal Projections of Tight Ends.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-cap: "Calibration Plot for Seasonal Projections of Kickers.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-alt: "Calibration Plot for Seasonal Projections of Kickers.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-cap: "Calibration Plot for Seasonal Projections of Defensive Linemen.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-alt: "Calibration Plot for Seasonal Projections of Defensive Linemen.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-cap: "Calibration Plot for Seasonal Projections of Linebackers.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-alt: "Calibration Plot for Seasonal Projections of Linebackers.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-cap: "Calibration Plot for Seasonal Projections of Defensive Backs.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 #| fig-alt: "Calibration Plot for Seasonal Projections of Defensive Backs.
+The diagonal black line is the reference line representing perfect calibration.
 The yellow line is the best-fit line based on LOESS.
@@
 
-To evaluate the accuracy of projections, we must first merge projections with actual performance.
-
-```{r}
-player_stats_weekly_subset <- player_stats_weekly %>% 
-  filter(!is.na(player_id))
-
-nfl_playerIDs_subset <- nfl_playerIDs %>% 
-  filter(!is.na(gsis_id)) %>% 
-  distinct(gsis_id, .keep_all = TRUE) %>% # keep only rows that do not have duplicate values of gsis_id
-  select(-all_of(c("team", "position", "height", "weight", "age")))
-
-players_projectedPoints_weekly_combined$season <- as.integer(players_projectedPoints_weekly_combined$season)
-
-player_stats_weekly_subset_IDs <- left_join(
-  player_stats_weekly_subset,
-  nfl_playerIDs_subset,
-  by = c("player_id" = "gsis_id")
-) %>% 
-  filter(!is.na(mfl_id))
-
-projectionsWithActuals_weekly <- full_join(
-  player_stats_weekly_subset_IDs,
-  players_projectedPoints_weekly_combined,
-  by = c("mfl_id" = "id", "season", "week"),
-  suffix = c("", "_proj"),
-)
-
-projectionsWithActuals_weekly <- projectionsWithActuals_weekly %>% 
-  unite(
-    "player_id_season_week",
-    player_id,
-    season,
-    week,
-    remove = FALSE
-  )
-
-playersWithHighProjectedOrActualPoints_weekly <- projectionsWithActuals_weekly %>% 
-  filter(raw_points > 6 | fantasyPoints > 6) %>% 
-  select(player_id_season_week) %>% 
-  pull()
-
-projectionsWithActuals_weekly_qb <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "QB")
-
-projectionsWithActuals_weekly_rb <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "RB")
-
-projectionsWithActuals_weekly_wr <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "WR")
-
-projectionsWithActuals_weekly_te <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "TE")
-
-projectionsWithActuals_weekly_k <- projectionsWithActuals_weekly %>% 
-  filter(position == "K")
-
-projectionsWithActuals_weekly_dl <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "DL")
-
-projectionsWithActuals_weekly_lb <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "LB")
-
-projectionsWithActuals_weekly_db <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "DB")
-```
-
 #### Overall {#sec-accuracyProjectionsWeeklyOverall}
@@
 #| fig-cap: "Calibration Plot for Weekly Projections.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-alt: "Calibration Plot for Weekly Projections.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-cap: "Calibration Plot for Weekly Projections of Quarterbacks.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-alt: "Calibration Plot for Weekly Projections of Quarterbacks.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-cap: "Calibration Plot for Weekly Projections of Running Backs.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-alt: "Calibration Plot for Weekly Projections of Running Backs.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-cap: "Calibration Plot for Weekly Projections of Wide Receivers.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-alt: "Calibration Plot for Weekly Projections of Wide Receivers.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-cap: "Calibration Plot for Weekly Projections of Tight Ends.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-alt: "Calibration Plot for Weekly Projections of Tight Ends.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-cap: "Calibration Plot for Weekly Projections of Kickers.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-alt: "Calibration Plot for Weekly Projections of Kickers.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-cap: "Calibration Plot for Weekly Projections of Defensive Linemen.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-alt: "Calibration Plot for Weekly Projections of Defensive Linemen.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-cap: "Calibration Plot for Weekly Projections of Linebackers.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-alt: "Calibration Plot for Weekly Projections of Linebackers.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-cap: "Calibration Plot for Weekly Projections of Defensive Backs.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 #| fig-alt: "Calibration Plot for Weekly Projections of Defensive Backs.
-The yellow line is the best-fit line based on LOESS.
+The diagonal black line is the reference line representing perfect calibration.
+The yellow line is the best-fit line based on a generalized additive model.
 The purple band is the 95% confidence interval of actual fantasy points.

### 2025-04-25 correlation.qmd
20250417 - diagnostic plots
20250425 - xi correlation coefficient

@@
 
-A scatterplot with best-fit line is in @fig-scatterplotWithBestFitLine.
+A [scatterplot](#sec-scatterplot) with best-fit line is in @fig-scatterplotWithBestFitLine.
 
@@
   player_stats_seasonal$fantasyPoints,
+  method = "permutation",
   pvalue = TRUE
 )
-
-set.seed(52242) # for reproducibility
-
-XICOR::calculateXI(
-  player_stats_seasonal$height,
-  player_stats_seasonal$fantasyPoints)
 ```

### 2025-04-25 correlation.qmd
20250425 - organization

@@
 
-## Examples {#sec-correlationExamples}
-
-### Player Height and Fantasy Points {#sec-correlationHeightFantasyPoints}
+## Example: Player Height and Fantasy Points {#sec-correlationHeightFantasyPoints}
 
@@
 
-#### Covariance {#sec-correlationExamplesCovariance}
+### Covariance {#sec-correlationExamplesCovariance}
 
@@
 
-#### Pearson Correlation {#sec-correlationExamplesPearson}
+### Pearson Correlation {#sec-correlationExamplesPearson}
 
@@
 
-#### Scatterplot with Best-Fit Line
+### Scatterplot with Best-Fit Line
 
@@
 
-#### Rank Correlation: Spearman's Rho ($\rho$) {#sec-correlationExamplesSpearman}
+### Rank Correlation: Spearman's Rho ($\rho$) {#sec-correlationExamplesSpearman}
 
@@
 
-#### Partial Correlation {#sec-correlationExamplesPartial}
+### Partial Correlation {#sec-correlationExamplesPartial}
 
@@
 
-#### Nonlinear Correlation {#sec-correlationExamplesNonlinear}
+### Nonlinear Correlation {#sec-correlationExamplesNonlinear}
 
@@
 
-#### Correlation Matrix {#sec-correlationExamplesMatrix}
+### Correlation Matrix {#sec-correlationExamplesMatrix}
 
@@
 
-#### Correlogram {#sec-correlationExamplesCorrelogram}
+### Correlogram {#sec-correlationExamplesCorrelogram}
 

### 2025-04-25 data-visualization.qmd
20250408 - plotly
20250425 - scatterplot with rug plot/marginal densigram

@@
 library("ggExtra")
+library("patchwork")
 library("tidyverse")
@@
 
-ggplotly(plot_ypcByPlayerAge)
+plotly::ggplotly(plot_ypcByPlayerAge)
 ```
@@
 
+### Scatterplot with Rug Plot {#sec-scatterplotRugPlot}
+
+We can also add a rug plot, as in @fig-scatterplot12.
+
+```{r}
+#| label: fig-scatterplot12
+#| fig-cap: "Scatterplot With Rug Plot."
+#| fig-alt: "Scatterplot With Rug Plot."
+
+ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age,
+    y = rushing_yards)) +
+  geom_point() +
+  geom_smooth() +
+  geom_rug(
+    col = "#800000",
+    alpha = 0.2) +
+  coord_cartesian(
+    xlim = c(20,40),
+    ylim = c(0,NA)) +
+  scale_x_continuous(
+    breaks = seq(from = 20, to = 40, by = 5)
+  ) +
+  scale_y_continuous(
+    breaks = seq(from = 0, to = 2500, by = 250)
+  ) +
+  labs(
+    x = "Running Back's Age (years)",
+    y = "Rushing Yards (Season)",
+    title = "NFL Rushing Yards (Season) by Player Age",
+    subtitle = "(Among Running Backs)"
+  ) +
+  theme_classic()
+```
+
 ### Marginal Density/Histogram {#sec-scatterplotMarginalDensigram}
 
-We can also add a marginal density/histogram, as in @fig-scatterplot12 using the `ggExtra` package [@R-ggExtra]:
-
-```{r}
-#| label: fig-scatterplot12
+We can also add a marginal density/histogram, as in @fig-scatterplot13 using the `ggExtra` package [@R-ggExtra]:
+
+```{r}
+#| label: fig-scatterplot13
 #| fig-cap: "Scatterplot With Marginal Density/Histogram."
@@
   type = "densigram")
+```
+
+Or we can create the plots manually, as in @fig-scatterplot14 using the `patchwork` package [@R-patchwork]:
+
+```{r}
+#| label: fig-scatterplot14
+#| fig-cap: "Scatterplot With Marginal Density/Histogram."
+#| fig-alt: "Scatterplot With Marginal Density/Histogram."
+
+dens1 <- ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = age)) +
+  coord_cartesian(
+    xlim = c(20,40),
+    expand = FALSE) +
+  geom_histogram(
+    aes(y = after_stat(density)),
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  geom_density(
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  theme_void()
+
+dens2 <- ggplot2::ggplot(
+  data = rb_seasonal,
+  aes(
+    x = rushing_yards)) +
+  geom_histogram(
+    aes(y = after_stat(density)),
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  geom_density(
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  theme_void() +
+  coord_flip(
+    ylim = c(0, NA),
+    expand = FALSE)
+
+dens1 + patchwork::plot_spacer() + plot_ypcByPlayerAge + dens2 + 
+  patchwork::plot_layout(
+    ncol = 2, 
+    nrow = 2, 
+    widths = c(4, 1),
+    heights = c(1, 4)
+  )
 ```

### 2025-04-25 multiple-regression.qmd
20250425 - reorganize

@@
 
-## Examples {#sec-multipleRegressionExamples}
-
-### Predicting Wide Receivers' Fantasy Points {#sec-multipleRegressionFantasyPointsWRs}
+## Example: Predicting Wide Receivers' Fantasy Points {#sec-multipleRegressionFantasyPointsWRs}
 
@@
 
-#### Examine Descriptive Statistics {#sec-multipleRegressionFantasyPointsWRsDescriptive}
+### Examine Descriptive Statistics {#sec-multipleRegressionFantasyPointsWRsDescriptive}
 
@@
 
-#### Examine Bivariate Associations {#sec-multipleRegressionFantasyPointsWRsBivariate}
+### Examine Bivariate Associations {#sec-multipleRegressionFantasyPointsWRsBivariate}
 
@@
 
-#### Estimate Multiple Regression Model {#sec-multipleRegressionFantasyPointsWRsModel}
+### Estimate Multiple Regression Model {#sec-multipleRegressionFantasyPointsWRsModel}
 
@@
 
-#### Evaluating and Addressing Assumptions {#sec-multipleRegressionExamplesAddressingAssumptions}
+### Evaluating and Addressing Assumptions {#sec-multipleRegressionExamplesAddressingAssumptions}
 
@@
 
-##### Linear Association {#sec-multipleRegressionExamplesAddressingAssumptionsLinearAssociation}
+#### Linear Association {#sec-multipleRegressionExamplesAddressingAssumptionsLinearAssociation}
 
@@
 
-##### Homoscedasticity {#sec-multipleRegressionExamplesAddressingAssumptionsHomoscedasticity}
+#### Homoscedasticity {#sec-multipleRegressionExamplesAddressingAssumptionsHomoscedasticity}
 
@@
 
-##### Uncorrelated Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsUncorrelatedResiduals}
+#### Uncorrelated Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsUncorrelatedResiduals}
 
@@
 
-##### Normally Distributed Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsNormallyDistributedResiduals}
+#### Normally Distributed Residuals {#sec-multipleRegressionExamplesAddressingAssumptionsNormallyDistributedResiduals}
 

### 2025-04-25 base-rates.qmd
20250425 - structure

@@
 Whereas a solid Wide Receiver may log 100+ receptions and 1,200+ yards in a season, they may have "only" 8–14 receiving touchdowns in a given season.
-As noted in @sec-differentErrorsDifferentCosts, lower base-rate events—including touchdowns—are harder to predict accurately.
+As noted in @sec-difficultyPredictingLowBRevents, lower base-rate events—including touchdowns—are harder to predict accurately.
 As noted by @Harris2012: "NFL statistical projections are basically impossible to get right.
@@
 
-[PROVIDE ACCURACY OF PROJECTIONS OF TOUCHDOWNS VS YARDAGE]
+In @sec-difficultyPredictingLowBRevents, we evaluate the accuracy of predicting touchdowns versus yardage.
 

### 2025-05-05 evaluating-prediction-accuracy.qmd
20250425 - structure
20250505 - predictive R-squared

@@
 library("msir")
+library("caret")
 library("tidyverse")
@@
 By contrast, if you reject 1% of applicants and the [base rate](#sec-baseRate) of poor adjustment is 5%, then you are making errors of under-rejecting or over-accepting (false negatives).
+
+### Difficulty Predicting Low Base Rate Events {#sec-difficultyPredictingLowBRevents}
 
@@
 
+An issue with $R^2$ is that it can be artificially inflated by adding more predictors to the model, even if those predictors do not improve the model fit.
+$R^2$ never decreases when adding predictors to a model, even if the predictors do not provide any predictive power.
+Thus, to account for the number of predictors in the model, we can use [adjusted $R^2$](#sec-adjustedRsquared).
+
 ###### Adjusted $R^2$ ($R^2_{adj}$) {#sec-adjustedRsquared}
@@
 Adjusted $R^2$ is similar to the [coefficient of determination](#sec-coefficientOfDetermination), but it accounts for the number of predictors included in the regression model to penalize [overfitting](#sec-overfitting).
+Specifically, adjusted $R^2$ penalizes the number of predictors in the model.
 Adjusted $R^2$ reflects the proportion of variance in the outcome (dependent) variable that is explained by the model predictions over and above what would be expected to be accounted for by chance, given the number of predictors in the model.
+Thus, adjusted $R^2$ will increase only if the new predictor improves the model (i.e., explains additional variance in the outcome variable) more than would be expected by chance.
 Larger values indicate greater accuracy.
@@
 
-Predictive $R^2$ is described by @Hopper2014 here: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>).
+Predictive $R^2$ is described by @Hopper2014 and @Tuzen2025: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>) and <https://mfatihtuzen.netlify.app/posts/2025-04-30_rsquared/> (archived at <https://perma.cc/EPP8-VG23>).
 Predictive $R^2$ penalizes [overfitting](#sec-overfitting), unlike traditional $R^2$.
 Larger values indicate greater accuracy.
+The formula for predictive $R^2$ is in in @eq-predictiveRSquared:
+
+$$
+\begin{aligned}
+  R^2_{pred} &= 1 - \frac{\text{PRESS}}{\text{SS}_\text{total}}
+\end{aligned}
+$$ {#eq-predictiveRSquared}
+
+where PRESS is the predictive residual sum of squares, and $\text{SS}_\text{total}$ is the total sum of squares.
+
+It is preferable to examine predictive $R^2$ using new data that the model was not trained on—i.e., cross-validation, either in an independent sample (i.e., external cross-validation) or in a hold-out sample (i.e., internal cross-validation on test data).\index{predictive $R^2$}\index{cross-validation}
+
+For example, here is an example of estimating predictive $R^2$ 10-fold cross-validation using the `caret` package [@R-caret].\index{predictive $R^2$}\index{cross-validation}
+
+```{r}
+#| eval: false
+
+set.seed(52242)
+
+# Set up 10-fold cross-validation
+train_control <- caret::trainControl(
+  method = "cv",
+  number = 10)
+
+# Fit the model with cross-validation
+model_cv <- caret::train(
+  outcomeVariable ~ predictorVariable1 + predictorVariable2,
+  data = mydata,
+  method = "lm",
+  na.action = na.exclude,
+  trControl = train_control,
+  metric = "Rsquared")
+
+# View R-squared across the folds
+model_cv$results$Rsquared
+```
 

### 2025-05-05 evaluating-prediction-accuracy.qmd
20250505 - remove subject indexing for now

@@
 
-It is preferable to examine predictive $R^2$ using new data that the model was not trained on—i.e., cross-validation, either in an independent sample (i.e., external cross-validation) or in a hold-out sample (i.e., internal cross-validation on test data).\index{predictive $R^2$}\index{cross-validation}
-
-For example, here is an example of estimating predictive $R^2$ 10-fold cross-validation using the `caret` package [@R-caret].\index{predictive $R^2$}\index{cross-validation}
+It is preferable to examine predictive $R^2$ using new data that the model was not trained on—i.e., cross-validation, either in an independent sample (i.e., external cross-validation) or in a hold-out sample (i.e., internal cross-validation on test data).
+
+For example, here is an example of estimating predictive $R^2$ 10-fold cross-validation using the `caret` package [@R-caret].
 

### 2025-05-05 research-methods.qmd
20250505 - addition and typos

@@
 The consistency of scores across time can be examined in terms of relative or absolute test–retest reliability.
-Relative test–retest reliability—i.e., the consistency of individual differences across time—is commonly evaluated using the coefficient of stability (i.e., the Pearson correlation coefficent).
-Absolute test–retest reliability—i.e., the absolute consistency of people's scores across time—is commonly evaluated using the coefficient of repeatability.
+Relative test–retest reliability—i.e., the consistency of individual differences across time—is commonly evaluated using the coefficient of stability (i.e., the Pearson correlation coefficient).
+Absolute test–retest reliability—i.e., the absolute consistency of people's scores across time—is commonly evaluated using the coefficient of repeatability and the coefficient of agreement (i.e., intraclass correlation coefficient).
 
@@
 If we develop two equivalent versions of the Wonderlic Contemporary Cognitive Ability Test (Form A and Form B) so that players sitting next to each other do not receive the same items, we would expect a player's score on Form A would be similar to their score on Form B.
-Parallel-forms reliability is is commonly evaluated using the coefficient of equivalence (i.e., the Pearson correlation coefficent).
+Parallel-forms reliability is is commonly evaluated using the coefficient of equivalence (i.e., the Pearson correlation coefficient).
 

### 2025-05-06 evaluating-prediction-accuracy.qmd
20250505 - predictive R-squared

@@
 model_cv$results$Rsquared
+```
+
+Here is an example of estimating predictive $R^2$ with a training and testing data set using the `caret` package [@R-caret].
+
+```{r}
+#| eval: false
+
+set.seed(52242)
+
+inTrain <- caret::createDataPartition(
+  y = as.factor(mydata$outcomeVariable),
+  p = .75, # the proportion of data for the training data set
+  list = FALSE
+)
+
+training <- mydata[ inTrain,]
+testing  <- mydata[-inTrain,]
+
+training_model <- lm(
+  outcomeVariable ~ predictorVariable1 + predictorVariable2,
+  data = training)
+
+testing_predictions <- predict(
+  training_model,
+  newdata = testing)
+
+petersenlab::accuracyOverall(
+  predicted = testing_predictions,
+  actual = testing$outcomeVariable)$rsquared
 ```

### 2025-05-06 evaluating-prediction-accuracy.qmd
20250505 - predictive R-squared

@@
 
-###### Predictive $R^2$ {#sec-predictiveRsquared}
-
-Predictive $R^2$ is described by @Hopper2014 and @Tuzen2025: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>) and <https://mfatihtuzen.netlify.app/posts/2025-04-30_rsquared/> (archived at <https://perma.cc/EPP8-VG23>).
+###### Predictive $R^2$ ($R^2_{\text{pred}}$) {#sec-predictiveRsquared}
+
+Predictive $R^2$ ($R^2_{\text{pred}}$) is described by @Hopper2014 and @Tuzen2025: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>) and <https://mfatihtuzen.netlify.app/posts/2025-04-30_rsquared/> (archived at <https://perma.cc/EPP8-VG23>).
 Predictive $R^2$ penalizes [overfitting](#sec-overfitting), unlike traditional $R^2$.
@@
 \begin{aligned}
-  R^2_{pred} &= 1 - \frac{\text{PRESS}}{\text{SS}_\text{total}}
+  R^2_{\text{pred}} &= 1 - \frac{\text{PRESS}}{\text{SS}_\text{total}}
 \end{aligned}
@@
 
-For example, here is an example of estimating predictive $R^2$ 10-fold cross-validation using the `caret` package [@R-caret].
+For example, here is an example of estimating predictive $R^2$ using 10-fold cross-validation using the `caret` package [@R-caret].
 

### 2025-05-15 download-football-data.qmd
20250514 - NFL standings

@@
 library("nflfastR")
+library("nflseedR")
 library("nfl4th")
@@
 
-```{r}
-#| eval: false
-
-nfl_standings_raw <- read_csv("http://www.habitatring.com/standings.csv")
+We can determine NFL standings based on the `nfl_schedules` object using the `nflseedR` [@R-nflseedR] package.
+
+```{r}
+#| eval: false
+
+nfl_standings_raw <- nflseedR::nfl_standings(nfl_schedules) #read_csv("http://www.habitatring.com/standings.csv"); archived at: https://perma.cc/5QLN-VK7V
 ```
@@
 ```{r}
-nfl_standings <- nfl_standings_raw
+nfl_standings <- data.frame(nfl_standings_raw)
 ```

### 2025-05-15 index.qmd
20250512 - OpenHawks grant
20250514 - NFL standings

--- 2025-04-11 index.qmd
@@
   "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
-  "nflreadr", "nflplotR", "nflfastR", "nfl4th", "gsisdecoder", "progressr",
-  "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", "ggdag",
-  "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", "pROC",
-  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", "AICcmodavg",
-  "rstan", "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", "viridis",
-  "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", "stringi", #"webshot2", 
-  "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin",
-  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation", 
-  "mgcv", "rms", "car", "performance", "lavaan", "mice", "miceadds","interactions"
+  "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", "gsisdecoder", 
+  "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", 
+  "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
+  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
+  "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", 
+  "viridis", "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", #"webshot2", 
+  "stringi", "parallel", "parallelly", "doParallel", "missRanger", "ggridges", 
+  "powerjoin", "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", 
+  "correlation", "mgcv", "rms", "car", "performance", "lavaan", "mice", 
+  "miceadds","interactions", "shiny","PearsonDS","robustbase"
 )
@@
 
+This book was supported by an OpenHawks Open Educational Resources grants program from the University of Iowa Libraries.
+
 I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.

### 2025-05-21 basic-statistics.qmd
20250414 - add packages and DOIs
20250520 - update package list

@@
 library("grid")
+library("psych")
 library("tidyverse")
@@
 Try varying the standard deviation, skewness, and kurtosis to see the impact of each.
+The interactive graph was created using the `shiny` [@R-shiny] and `PearsonDS` [@R-PearsonDS] packages.
+The source code for interactive graph is here: <https://github.com/isaactpetersen/DistributionVisualizer>.
 

### 2025-05-21 getting-started.qmd
20250409 - working directory
20250520 - update package list

@@
 1. Install `RStudio Desktop`: <https://posit.co/download/rstudio-desktop>
-1. After installing `RStudio`, open `RStudio` and run the following code in the console to install several key `R` packages:
+1. After installing `RStudio`, open `RStudio` and run the following code in the console to install the `R` packages used in this book (note: this will take a while):
    ```{r}
@@
    install.packages(
-     c("petersenlab","remotes","nflreadr","nflfastR","nfl4th","nflplotR",
-     "gsisdecoder","progressr","lubridate","tidyverse","psych"))
+     c("petersenlab","remotes","knitr","rmarkdown","dplyr",
+       "ggplot2","tidyverse","psych","xfun","downlit","xml2",
+       "nflreadr","nflplotR","nflfastR","nflseedR","nfl4th",
+       "gsisdecoder","progressr","DescTools","pwr","pwrss",
+       "WebPower","XICOR","dagitty","ggdag","lubridate",
+       "ggtext","gghighlight","ggExtra","grid","patchwork",
+       "pROC","lme4","lmerTest","MuMIn","emmeans","pbkrtest",
+       "sjstats","AICcmodavg","rstan","brms","bbmle",
+       "fitdistrplus","mclust","magrittr","viridis","viridisLite",
+       "msir","plotly","webshot2","quantmod","fPortfolio",
+       "xts","zoo","forecast","stringi","parallel","parallelly",
+       "doParallel","missRanger","powerjoin","caret","gpboost",
+       "ggridges","corrplot","rms","car","caret","performance",
+       "correlation","interactions","lavaan","mice","miceadds",
+       "robustbase","see"))
    ```
 1.
-Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub and need to be installed using the following code (after installing the `remotes` package [@R-remotes] above):
-   ```{r}
-   #| eval: false
-
+Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub (and are not hosted on the Comprehensive R Archive Network [CRAN]) and thus need to be installed using the following code (after installing the `remotes` package [@R-remotes] above)[^petersenlabPackageGitHub]:
+   ```{r}
+   #| eval: false
+
+   remotes::install_github("DevPsyLab/petersenlab")
    remotes::install_github("FantasyFootballAnalytics/ffanalytics")
-   ```
+   remotes::install_github("stan-dev/cmdstanr")
+   ```
+
+[^petersenlabPackageGitHub]: Although the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] is hosted on CRAN, installing from GitHub will ensure you have the latest version.
 
@@
 
+To determine where you working directory is, you can type:
+
+```{r}
+getwd()
+```
+
+To change your working directory, you can type:
+
+```{r}
+#| eval: false
+
+setwd("C:/Users/myusername/")
+```
+
 The following code loads a file from an *absolute path*:
@@
 
-Here is how you save a `.RData` file using a *relative path*:
+Here is how you save a `.RData` file using a *relative path* (i.e., relative to your working directory):
 

### 2025-05-21 basic-statistics.qmd
20250520 - sports analytics

@@
 
-#### Mediation and Moderation
+#### Mediation and Moderation {#sec-statisticalPowerMediationModeration}
 
@@
 
+## Ethics of Reporting Statistics {#sec-ethicsReportingStatistics}
+
+It is important to report statistics ethically and responsibly.
+People can use statistics to lie [@Huff2023].
+It is important to present statistics fairly and accurately even if they are not convenient to your preferred narrative.
+In addition, it is important to consider [practical significance](#sec-practicalSignificance) in addition to [statistical significance](#sec-statisticalSignificance).
+
 ## Conclusion {#sec-basicStatsConclusion}
@@
 [Inferential statistics](#sec-inferentialStatistics) are used to draw inferences regarding differences between groups or associations between variables.
-Null hypothesis signficance testing is a framework for [inferential statistics](#sec-inferentialStatistics), in which there is a [null hypothesis](#sec-nullHypothesis) and [alternative hypothesis](##sec-alternativeHypothesis).
+Null hypothesis signficance testing is a framework for [inferential statistics](#sec-inferentialStatistics), in which there is a [null hypothesis](#sec-nullHypothesis) and [alternative hypothesis](#sec-alternativeHypothesis).
 The null hypothesis is that there is no difference between groups or that there is no association between variables.

### 2025-05-21 cognitive-bias.qmd
20250520 - sports analytics

@@
 
-Focus on the decision-making process when evaluating the quality of decisions.
+Focus on the decision-making process when evaluating the quality of decisions, rather than the outcome [@Tanney2021].
 Yes, also evaluate the outcomes of decisions, but do not give yourself too much credit for lucky decisions or too much blame for unlucky decisions (like injuries).
+Luck is going to happen.
+If you are making optimal decisions that maximize the expected value of returns (e.g., projected fantasy points), you are following a strong process.
 That said, one thing you can count on is that players get injured.

### 2025-05-21 data-visualization.qmd
20250520 - sports analytics

@@
 ## Overview {#sec-dataVisualizationOverview}
+
+Graphics serve an important way of efficiently communicating findings to diverse audiences.
+Effective graphics can help people with less technical knowledge better understand your findings.
+As noted by @Tanney2021, it is important to communicate your work clearly to stakeholders (e.g., coaches, companies, or prospective clients); if you cannot explain your work clearly, you cannot expect a stakeholder to "buy in" to what you are selling.
 

### 2025-05-21 decision-making.qmd
20250520 - sports analytics

@@
 
+Sports analyst @Tanney2021 encourages people to consider framing your decisions as bets, to use a probabilistic mindset (rather than a determininistic mindset).
+Even if something is perceived as likely to occur with 100% certainty, it is not guaranteed that such an event will occur.
+Other things can and do happen, and [overconfidence](#sec-cognitiveBiasesOverconfidence) can get in the way of effective decision-making.
+
+A key goal is to identify decisions that provide a net advantage in expected value.
+Although there is considerable variability and the outcome is not guaranteed, small advantages can yield considerable gains in the long run @Tanney2021; however, one must be able to survive the costs of losses to obtain enough long-term trials.
+So, do not bet more than you are willing to lose.
+
 ## Risk Management Principles from Cognitive Psychology {#sec-riskManagementCognitivePsych}

### 2025-05-21 sports-cognitive-psychology.qmd
20250520 - citation
20250520 - sports analytics

@@
 For instance, in basketball, there has been a greater focus on three points and buckets close to the rim (e.g., dunks and layups), with way fewer midrange shots [@Partnow2021].
-Moreover, star players rest more games.
+Moreover, star players rest more games ("load management").
 Here is an article on how Shane Battier, who was not the quickest, fastest, or most athletic player, used analytics successfully to guard players who were quicker or more athletic than him: <https://www.nytimes.com/2009/02/15/magazine/15Battier-t.html> [@Lewis2009b; archived at <https://web.archive.org/web/20250219225627/https://www.nytimes.com/2009/02/15/magazine/15Battier-t.html>].
 
-Football has also seen greater use of analytics, although its uptake has been somewhat slower than in many other sports.
+Football has also seen greater use of analytics, although its uptake has been somewhat slower than in many other sports [@Carroll2023].
 When asked about how he makes the decision about whether to go for a two-point conversion after a touchdown, Steelers Head Coach Mike Tomlin stated:
@@
 
+## Challenging a Call {#challengingCall}
+
+A coach can challenge a call (and have it reviewed) by throwing a red flag onto the field.
+If, after the review, the play is reversed, the team does not lose a timeout; however, if the play stands as originally called, the team loses a timeout.
+One task is for coaches to determine when it is optimal to challenge a call.
+NFL.com provides an opportunity for you to determine whether you should challenge a call: <https://operations.nfl.com/learn-the-game/making-the-call/you-make-the-call/>.
+
+@Tanney2021 provides a framework for determining whether a coach should challenge a call, based on the expected value of the change in winning probability of a reversal versus a non-reversal:
+
+$$
+\begin{aligned}
+  \text{expected value} &= \text{Sum}(\text{Probabilities} \times \text{Payouts}) \\
+  \text{expected value} &= (\text{Reversal \%})(\text{Positive change in Winning Probability of a Reversal}) + (1 - \text{Reversal \%})(\text{Negative change in Winning Probability of Losing a Timeout})
+\end{aligned}
+$$ {#eq-challengingCall}
+
 ## Other Beliefs in Sports {#sec-beliefsInSports}

### 2025-05-25 sports-cognitive-psychology.qmd
20250524 - clean up

@@
   \text{expected value} &= \text{Sum}(\text{Probabilities} \times \text{Payouts}) \\
-  \text{expected value} &= (\text{Reversal \%})(\text{Positive change in Winning Probability of a Reversal}) + (1 - \text{Reversal \%})(\text{Negative change in Winning Probability of Losing a Timeout})
+  &= (\text{Reversal \%})(\text{Positive change in Winning Probability of a Reversal}) \\
+  &\quad + (1 - \text{Reversal \%})(\text{Negative change in Winning Probability of Losing a Timeout})
 \end{aligned}

### 2025-05-26 sports-cognitive-psychology.qmd
20250525 - tweak1

@@
   \text{expected value} &= \text{Sum}(\text{Probabilities} \times \text{Payouts}) \\
-  &= (\text{Reversal \%})(\text{Positive change in Winning Probability of a Reversal}) \\
-  &\quad + (1 - \text{Reversal \%})(\text{Negative change in Winning Probability of Losing a Timeout})
+  &= (\text{Reversal \%})(\text{Positive change in Winning Probability of a Reversal}) + \\
+  &\phantom{{}={}} (1 - \text{Reversal \%})(\text{Negative change in Winning Probability of Losing a Timeout})
 \end{aligned}

### 2025-05-26 sports-cognitive-psychology.qmd
20250525 - tweak2

--- 2025-05-25 sports-cognitive-psychology.qmd
@@
 \begin{aligned}
-  \text{expected value} &= \text{Sum}(\text{Probabilities} \times \text{Payouts}) \\
-  &= (\text{Reversal \%})(\text{Positive change in Winning Probability of a Reversal}) \\
-  &\quad + (1 - \text{Reversal \%})(\text{Negative change in Winning Probability of Losing a Timeout})
+  \text{expected value} = \; &\text{Sum}(\text{Probabilities} \times \text{Payouts}) \\
+  = \; &(\text{Reversal \%})(\text{Positive change in Winning Probability of a Reversal}) + \\
+  &(1 - \text{Reversal \%})(\text{Negative change in Winning Probability of Losing a Timeout})
 \end{aligned}

### 2025-06-05 correlation.qmd
20250605 - correlation examples

@@
 The smaller the absolute value of the correlation coefficient (i.e., the closer the $r$ value is to zero), the weaker the association and the flatter the slope of the best-fit line in a scatterplot.\index{correlation}
-The larger the absolute value of the correlation coefficient (i.e., the closer the absolute value of the $r$ value is to one), the stronger the association and the steeper the slope of the best-fit line in a scatterplot.\index{correlation}
+The larger the absolute value of the correlation coefficient (i.e., the closer the absolute value of the $r$ value is to one), the stronger the association.\index{correlation}
 See @fig-rangeOfCorrelations for a range of different correlation coefficients and what some example data may look like for each direction and strength of association.\index{correlation}
@@
 ```
+
+Also note that, although the the ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength of the (linear) association between two variables, it does not reflect the slope of that association, as depicted in @fig-correlationExamples.
+
+::: {#fig-correlationExamples}
+![](images/correlationExamples.png){fig-alt="Several sets of ($x$, $y$) points, with the the ([Pearson](#sec-correlationExamplesPearson)) correlation coefficient of $x$ and $y$ for each set.
+The the ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength and direction of a linear association (top row), but not the slope of that association (middle), nor many aspects of nonlinear associations (bottom).
+N.B.: the figure in the center has a slope of 0 but in that case the correlation coefficient is undefined because the variance of $y$ is zero.
+From <https://en.wikipedia.org/wiki/File:Correlation_examples2.svg>."}
+
+Several sets of ($x$, $y$) points, with the the ([Pearson](#sec-correlationExamplesPearson)) correlation coefficient of $x$ and $y$ for each set.
+The the ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength and direction of a linear association (top row), but not the slope of that association (middle), nor many aspects of nonlinear associations (bottom).
+N.B.: the figure in the center has a slope of 0 but in that case the correlation coefficient is undefined because the variance of $y$ is zero.
+From <https://en.wikipedia.org/wiki/File:Correlation_examples2.svg>.
+:::
 

### 2025-06-05 correlation.qmd
20250605 - typos

@@
 
-Also note that, although the the ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength of the (linear) association between two variables, it does not reflect the slope of that association, as depicted in @fig-correlationExamples.
+Also note that, although the ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength of the (linear) association between two variables, it does not reflect the slope of that association, as depicted in @fig-correlationExamples.
 
 ::: {#fig-correlationExamples}
-![](images/correlationExamples.png){fig-alt="Several sets of ($x$, $y$) points, with the the ([Pearson](#sec-correlationExamplesPearson)) correlation coefficient of $x$ and $y$ for each set.
-The the ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength and direction of a linear association (top row), but not the slope of that association (middle), nor many aspects of nonlinear associations (bottom).
+![](images/correlationExamples.png){fig-alt="Several sets of ($x$, $y$) points, with the ([Pearson](#sec-correlationExamplesPearson)) correlation coefficient of $x$ and $y$ for each set.
+The ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength and direction of a linear association (top row), but not the slope of that association (middle), nor many aspects of nonlinear associations (bottom).
 N.B.: the figure in the center has a slope of 0 but in that case the correlation coefficient is undefined because the variance of $y$ is zero.
@@
 
-Several sets of ($x$, $y$) points, with the the ([Pearson](#sec-correlationExamplesPearson)) correlation coefficient of $x$ and $y$ for each set.
-The the ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength and direction of a linear association (top row), but not the slope of that association (middle), nor many aspects of nonlinear associations (bottom).
+Several sets of ($x$, $y$) points, with the ([Pearson](#sec-correlationExamplesPearson)) correlation coefficient of $x$ and $y$ for each set.
+The ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength and direction of a linear association (top row), but not the slope of that association (middle), nor many aspects of nonlinear associations (bottom).
 N.B.: the figure in the center has a slope of 0 but in that case the correlation coefficient is undefined because the variance of $y$ is zero.

### 2025-06-05 correlation.qmd
20250605 - converting a covariance to a correlation

@@
 The [Pearson correlation coefficient](#sec-correlationExamplesPearson), $r$, is an example of a standardized index of the covariance between two variables.
+We can convert a covariance to a [correlation](#sec-correlationExamplesPearson) by standardizing it; i.e., by dividing by the multiplication of the standard deviation of $x$ and the standard deviation of $y$, as in @eq-covToCor:
+
+$$
+r_{xy} = \frac{\text{Cov}(x, y)}{\sigma_x \cdot \sigma_y}
+$$ {#eq-covToCor}
+
+For instance, for the covariance between height and fantasy points, we can convert it to a [correlation](#sec-correlationExamplesPearson) (we subset to just those rows that have data on both $x$ and $y$ to ensure we compute the standard deviations using the same values that are included in the estimate of the covariance):
+
+```{r}
+complete_cases <- player_stats_seasonal[complete.cases(player_stats_seasonal$height, player_stats_seasonal$fantasyPoints),]
+
+cov_value <- cov(
+  complete_cases$height,
+  complete_cases$fantasyPoints,
+  use = "pairwise.complete.obs")
+
+cor_value <- cov_value / (sd(complete_cases$height, na.rm = TRUE) * sd(complete_cases$fantasyPoints, na.rm = TRUE))
+
+cov_value
+
+cor_value
+
+cor(
+  complete_cases$height,
+  complete_cases$fantasyPoints,
+  use = "pairwise.complete.obs")
+```
 

### 2025-06-06 basic-statistics.qmd
20250606 - update formulas

@@
 
-The mean of $X$ (written as: $\bar{X}$) is calculated as in @eq-IQR:
-
-$$
-\bar{X} = \frac{\sum X_i}{n} = \frac{X_1 + X_2 + ... + X_n}{n}
+The mean of $x$ (written as: $\bar{x}$) is calculated as in @eq-IQR:
+
+$$
+\bar{x} = \frac{\sum x_i}{n} = \frac{x_1 + x_2 + ... + x_n}{n}
 $$ {#eq-IQR}
@@
 
-The (sample) variance of $X$ (written as: $s^2$) is calculated as in @eq-variance:
-
-$$
-s^2 = \frac{\sum (X_i - \bar{X})^2}{n-1}
+The (sample) variance of $x$ (written as: $s^2$) is calculated as in @eq-variance:
+
+$$
+s^2 = \frac{\sum (x_i - \bar{x})^2}{n-1}
 $$ {#eq-variance}
 
-where $X_i$ is each data point, $\bar{X}$ is the mean of $X$, and $n$ is the number of data points.
-
-The (sample) standard deviation of $X$ (written as: $s$) is calculated as in @eq-sd:
-
-$$
-s = \sqrt{\frac{\sum (X_i - \bar{X})^2}{n-1}}
+where $x_i$ is each data point, $\bar{x}$ is the mean of $x$, and $n$ is the number of data points.
+
+The (sample) standard deviation of $x$ (written as: $s$) is calculated as in @eq-sd:
+
+$$
+s = \sqrt{\frac{\sum (x_i - \bar{x})^2}{n-1}}
 $$ {#eq-sd}
 
-The range is calculated of $X$ is calculated as in @eq-range:
+The range is calculated of $x$ is calculated as in @eq-range:
 
@@
 $$
-\text{MAD} = \text{median}(|X_i - \tilde{X}|)
+\text{MAD} = \text{median}(|x_i - \tilde{x}|)
 $$ {#eq-medianAbsoluteDeviation}
 
-where $\tilde{X}$ is the median of `X`.
+where $\tilde{x}$ is the median of $x$.
 Compared to the standard deviation, the median absolute deviation is more robust to outliers.
@@
 $$
-z = \frac{X - \bar{X}}{\sigma}
+z = \frac{x - \bar{x}}{\sigma}
 $$ {#eq-zScore}
 
-where $X$ is the observed score, $\bar{X}$ is the mean observed score, and $\sigma$ is the standard deviation of the observed scores.
+where $x$ is the observed score, $\bar{x}$ is the mean observed score, and $\sigma$ is the standard deviation of the observed scores.
 
@@
 \begin{aligned}
-  R^2 &= 1 - \frac{\sum (Y_i - \hat{Y}_i)^2}{\sum (Y_i - \bar{Y})^2} \\
+  R^2 &= 1 - \frac{\sum (y_i - \hat{y}_i)^2}{\sum (y_i - \bar{y})^2} \\
       &= 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}} \\
@@
       &= \eta^2 \\
-      &= \frac{\text{variance explained in }Y}{\text{total variance in }Y}
+      &= \frac{\text{variance explained in }y}{\text{total variance in }y}
 \end{aligned}
@@
 
-where $Y_i$ is the observed value of the [outcome variable](#sec-correlationalStudy) for the $i$th observation, $\hat{Y}_i$ is the model predicted value for the $i$th observation, $\bar{Y}$ is the mean of the observed values of the [outcome variable](#sec-correlationalStudy).
+where $y$ is the [outcome variable](#sec-correlationalStudy), $y_i$ is the observed value of the [outcome variable](#sec-correlationalStudy) for the $i$th observation, $\hat{y}_i$ is the model predicted value for the $i$th observation, $\bar{y}$ is the mean of the observed values of the [outcome variable](#sec-correlationalStudy).
 The total sum of squares is an index of the total variation in the [outcome variable](#sec-correlationalStudy).
@@
 
-where $X$ is the [predictor variable](#sec-correlationalStudy) and $Y$ is the [outcome variable](#sec-correlationalStudy).
+where $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), and $\text{Cov}(x, y)$ is the covariance between $x$ and $y$.
 

### 2025-06-06 basic-statistics.qmd
20250606 - formulas for correlation and covariance

@@
 $$
-r = \frac{\sum (X_i - \bar{X})(Y_i - \bar{Y})}{\sqrt{\sum (X_i - \bar{X})^2 \sum (Y_i - \bar{Y})^2}}
+\begin{aligned}
+  r &= \frac{\frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})}
+{\sqrt{\frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})^2} \sqrt{\frac{1}{n - 1} \sum_{i=1}^{n} (y_i - \bar{y})^2}} 
+  && \quad \text{the numerator is the covariance; the denominator is the standard deviations} \\
+    &= \frac{\sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})}
+{\sqrt{\sum_{i=1}^{n} (x_i - \bar{x})^2} \sqrt{\sum_{i=1}^{n} (y_i - \bar{y})^2}} 
+  && \quad \text{$n-1$ cancels out in numerator and denominator (typical formula for Pearson correlation)} \\
+    &= \frac{\frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})}{\sigma_x \cdot \sigma_y} 
+  && \quad \text{adding back $n-1$, replacing denominator from line 1 with standard deviation symbols} \\
+    &= \frac{\text{Cov}(x, y)}{\sigma_x \cdot \sigma_y} 
+  && \quad \text{replacing numerator with covariance notation}
+\end{aligned}
 $$ {#eq-pearsonCorrelation}

### 2025-06-06 correlation.qmd
20250606 - formulas for correlation and covariance

--- 2025-06-05 correlation.qmd
@@
 
-The following syntax shows the variance-covariance of a set of variables, where the covariance of a variable with itself is the variable's variance.
+The covariance is an unstandardized index of the association between two variables.
+The formula for (sample) covariance is in @eq-covariance and is the numerator to the formula for [correlation](#sec-correlationExamplesPearson).
+
+$$
+\text{Cov}(x, y) = \frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})
+$$ {#eq-covariance}
+
+For instance, we can calculate the covariance between two variables manually (we subset to just those rows that have data on both $x$ and $y$ to ensure we compute the means using the same values that are included in the estimate of the covariance):
+
+```{r}
+complete_cases <- player_stats_seasonal[complete.cases(player_stats_seasonal$height, player_stats_seasonal$fantasyPoints),]
+
+n <- nrow(complete_cases)
+x <- complete_cases$height
+y <- complete_cases$fantasyPoints
+x_mean <- mean(complete_cases$height)
+y_mean <- mean(complete_cases$fantasyPoints)
+
+cov_manual <- sum((x - x_mean) * (y - y_mean)) / (n - 1)
+cov_manual
+```
+
+The following syntax shows the variance-covariance matrix of a set of variables, where the covariance of a variable with itself is the variable's variance.
+The variances are the values on the diagonal; the covariances are the values on the off-diagonal.
 
@@
 
-The covariance is an unstandardized index of the association between two variables.
-However, it can be helpful to use a standardized index to examine the [effect size](#sec-practicalSignificance)—i.e., the strength of the association.
+The covariance is unstandardized—its metric depends on the metrics of the two variables examined in the association.
+However, it can be helpful to use a standardized index of the association between the variables.
+A standardized index means that the index is on a common metric and does not depend on the metrics of the two variables.
+This is valuable because it allows fairly comparing the strength of associations between sets of variables with different metrics, which is meaningful for estimating the [effect size](#sec-practicalSignificance)—i.e., the strength of the association.
 The [Pearson correlation coefficient](#sec-correlationExamplesPearson), $r$, is an example of a standardized index of the covariance between two variables.
+We can convert a covariance to a [correlation](#sec-correlationExamplesPearson) by standardizing it; i.e., by dividing by the multiplication of the standard deviation of $x$ and the standard deviation of $y$, as in @eq-pearsonCorrelation.
+
+For instance, for the covariance between height and fantasy points, we can convert it to a [correlation](#sec-correlationExamplesPearson):
+
+```{r}
+cov_value <- cov(
+  complete_cases$height,
+  complete_cases$fantasyPoints,
+  use = "pairwise.complete.obs")
+
+cor_value <- cov_value / (sd(complete_cases$height, na.rm = TRUE) * sd(complete_cases$fantasyPoints, na.rm = TRUE))
+
+cov_value
+
+cor_value
+
+cor(
+  complete_cases$height,
+  complete_cases$fantasyPoints,
+  use = "pairwise.complete.obs")
+```
 
@@
 
+Correlation is a standardized index of the association between two variables.
+Pearson correlation is the most common type of corrrelation.
+The formula for Pearson correlation is in @eq-pearsonCorrelation.
 Pearson correlation assumes that the variables are [interval](#sec-interval) or [ratio](#sec-interval) level of measurement.

### 2025-06-06 correlation.qmd
20250606 - tweaks

@@
 This is valuable because it allows fairly comparing the strength of associations between sets of variables with different metrics, which is meaningful for estimating the [effect size](#sec-practicalSignificance)—i.e., the strength of the association.
-The [Pearson correlation coefficient](#sec-correlationExamplesPearson), $r$, is an example of a standardized index of the covariance between two variables.
+The [Pearson correlation coefficient](#sec-correlationExamplesPearson), $r$, is an example of a standardized index of the covariance between two variables because it has a common metric (whose possible values range from −1 to +1, regardless of the variables examined in the association).
 We can convert a covariance to a [correlation](#sec-correlationExamplesPearson) by standardizing it; i.e., by dividing by the multiplication of the standard deviation of $x$ and the standard deviation of $y$, as in @eq-pearsonCorrelation.
@@
 Correlation is a standardized index of the association between two variables.
-Pearson correlation is the most common type of corrrelation.
+Pearson correlation is the most common type of correlation.
 The formula for Pearson correlation is in @eq-pearsonCorrelation.

### 2025-06-07 download-football-data.qmd
20250607 - fixes

@@
 
-nfl_standings_raw <- nflseedR::nfl_standings(nfl_schedules) #read_csv("http://www.habitatring.com/standings.csv"); archived at: https://perma.cc/5QLN-VK7V
+nfl_standings_raw <- nflseedR::nfl_standings(
+  nfl_schedules%>% filter(!is.na(result))) #read_csv("http://www.habitatring.com/standings.csv"); archived at: https://perma.cc/5QLN-VK7V
 ```
@@
 save(
-  nfl_actualFantasyPoints_player_career,
+  nfl_actualFantasyPoints_player_career_raw,
   file = "./data/nfl_actualFantasyPoints_career_raw.RData"

### 2025-06-08 mixed-models.qmd
20250418 - fixes
20250608 - typos

@@
 If we have longitudinal data for multiple players, the time points are nested within players.
-Examining the association between age as a fixed effect in relation to fantasy performace would examine the assocication between a player's age and their fantasy performance while holding the association between age and fantasy performance constant across all players.
+Examining the association between age as a fixed effect in relation to fantasy performance would examine the association between a player's age and their fantasy performance while holding the association between age and fantasy performance constant across all players.
 That is, it would assume that all players show the same trajectory such as increase for 4 years then decrease.
-Examining the association between age as a random effect in relation to fantasy performace would examine the assocication between a player's age and their fantasy performance while allowing the association between age and fantasy performance to vary across players.
+Examining the association between age as a random effect in relation to fantasy performance would examine the association between a player's age and their fantasy performance while allowing the association between age and fantasy performance to vary across players.
 That is, it would allow the possibility that some players improve with age, whereas other players decline in performance with age.
@@
 ```{r}
-#| label: fig-qqPlotMixedModel
+#| label: fig-ppPlotMixedModel
 #| fig-cap: "Probability–Probability (PP) Plot."

### 2025-06-09 mixed-models.qmd
20250609 - tweak

@@
 library("brms")
-library("cmdstanr") # todo: install.packages("cmdstanr", repos = c("https://stan-dev.r-universe.dev", getOption("repos"))); check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
+library("cmdstanr") # todo: install.packages("cmdstanr", repos = c("https://stan-dev.r-universe.dev", getOption("repos"))); cmdstanr::check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
 library("fitdistrplus")

### 2025-06-09 research-methods.qmd
20250609 - spurious correlations

@@
 
-Lastly, the association might be spurious.
+Lastly, the association might be spurious (pure coincidence).
 It might just reflect random variation (i.e., chance), and that when tested on an independent sample, what appeared as an association in the original dataset may not hold when testing the association in a new dataset.
+@Vigen2024 provides a collection of spurious correlations: <https://www.tylervigen.com/spurious-correlations> (archived at <https://perma.cc/L443-JPRN>).
+As @Vigen2024 notes, spurious correlations may be particularly common in the context of outliers (extreme values), a small sample size, and when performing data dredging (or *p*-hacking)—i.e., examining the associations among many, many variables.
 

### 2025-06-09 basic-statistics.qmd
20250609 - coefficient of variation

@@
 - median absolute deviation
+- coefficient of variation
 
@@
 
+The coefficient of variation (CV) accounts for the fact that variables with larger means also tend to have larger standard deviations; it is calculated as the standard deviation divided by the mean, as in @eq-coefficientOfVariation:
+
+$$
+\text{CV} = \frac{SD(x)}{\bar{x}}
+$$ {#eq-coefficientOfVariation}
+
+where $SD(x)$ is the standard deviation of $x$.
+
 Below is R code to estimate each:
@@
 mad(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- coefficient of variation:
+
+```{r}
+sd(player_stats_seasonal$fantasyPoints, na.rm = TRUE) / mean(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
 ```
@@
         MAD = ~ mad(., na.rm = TRUE),
+        CV = ~ sd(., na.rm = TRUE) / mean(., na.rm = TRUE),
         median = ~ median(., na.rm = TRUE),

### 2025-06-09 player-evaluation.qmd
20250609 - coefficient of variation

@@
 
-However, the players differ considerably in their uncertainty (i.e., the source-to-source variability in their projections), as operationalized with the standard deviation and coefficient variation of projected points across sources for a given player.
+However, the players differ considerably in their uncertainty (i.e., the source-to-source variability in their projections), as operationalized with the [standard deviation](#sec-descriptiveStatisticsSpread) and [coefficient of variation](#sec-descriptiveStatisticsSpread) of projected points across sources for a given player.
 

### 2025-06-09 simulation.qmd
20250609 - bootstrapping

@@
-# Simulation: the Monte Carlo Method {#sec-simulation}
+# Simulation: Bootstrapping and the Monte Carlo Method {#sec-simulation}
 
@@
 ```{r}
+library("ffanalytics")
+library("data.table")
+library("future")
+library("future.apply")
 library("tidyverse")
@@
 ```{r}
-load(file = "./data/players_projectedPoints_seasonal.RData")
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projectedPoints_seasonal.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_playerIDs.RData", fsep = ""))
 ```
@@
 ```{r}
+all_proj <- bind_rows(players_projectedPoints_seasonal)
+```
+
+```{r}
+vars_by_pos <- list(
+  QB = c(
+    "games",
+    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
+    "rush_att", "rush_yds", "rush_tds",
+    "fumbles_lost"#,
+    #"sacks",
+    #"pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",  
+    #"pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_40_yds", "pass_400_yds",
+    #"rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds"
+    ),
+  RB = c(
+    "games",
+    "rush_att", "rush_yds", "rush_tds",
+    "rec_tgt", "rec", "rec_yds", "rec_tds",
+    "fumbles_lost",
+    "return_yds", "return_tds"#,
+    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
+    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
+    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
+  ),
+  WR = c(
+    "games",
+    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
+    "rush_att", "rush_yds", "rush_tds",
+    "rec_tgt", "rec", "rec_yds", "rec_tds",
+    "fumbles_lost",
+    "return_yds", "return_tds"#,
+    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
+    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
+    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
+  ),
+  TE = c(
+    "games",
+    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
+    "rush_att", "rush_yds", "rush_tds",
+    "rec_tgt", "rec", "rec_yds", "rec_tds",
+    "fumbles_lost",
+    "return_yds", "return_tds"#,
+    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
+    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
+    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
+  )
+)
+```
+
+```{r}
+bootstrapSimulation <- function(projectedStats, vars_by_pos, n_iter = 10000, seed = NULL) {
+  dt <- data.table::as.data.table(all_proj)
+  all_ids <- unique(dt$id)
+
+  if (!is.null(seed)) set.seed(seed)
+
+  future::plan(multisession)
+
+  results <- future.apply::future_lapply(
+    all_ids,
+    function(player_id) {
+      player_data <- dt[id == player_id]
+      player_pos <- unique(player_data$pos)
+
+      if (length(player_pos) != 1 || !player_pos %in% names(vars_by_pos)) return(NULL)
+
+      stat_vars <- vars_by_pos[[player_pos]]
+      out <- data.table::data.table(iteration = 1:n_iter, id = player_id, pos = player_pos)
+
+      for (var in stat_vars) {
+        if (var %in% names(player_data)) {
+          non_na_values <- player_data[[var]][!is.na(player_data[[var]])]
+
+          if (length(non_na_values) > 0) {
+            out[[var]] <- sample(non_na_values, n_iter, replace = TRUE)
+          } else {
+            out[[var]] <- NA_real_
+          }
+        } else {
+          out[[var]] <- NA_real_
+        }
+      }
+
+      return(out)
+    },
+    future.seed = TRUE
+  )
+
+  data.table::rbindlist(results, use.names = TRUE, fill = TRUE)
+}
+
+bootstappedStats <- bootstrapSimulation(
+  projectedStats = all_proj,
+  vars_by_pos = vars_by_pos,
+  n_iter = 10000,
+  seed = 52242)
+
+bootstappedStats$data_src <- bootstappedStats$iteration
+
+bootstappedStatsByPosition <- split(
+  bootstappedStats,
+  by = "pos",
+  keep.by = TRUE)
+
+bootstappedStatsByPosition <- lapply(
+  bootstappedStatsByPosition,
+  as_tibble)
+
+attr(bootstappedStatsByPosition, "season") <- 2024
+attr(bootstappedStatsByPosition, "week") <- 0
+
+bootstrappedFantasyPoints <- ffanalytics:::source_points(
+  data_result = bootstappedStatsByPosition,
+  scoring_rules = ffanalytics::scoring)
+
+bootstrappedFantasyPoints$iteration <- bootstrappedFantasyPoints$data_src
+bootstrappedFantasyPoints$data_src <- NULL
+
+bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
+  left_join(
+    nfl_playerIDs[,c("mfl_id","name","merge_name","team")],
+    by = c("id" = "mfl_id")
+  )
+```
+
+```{r}
 #| eval: false
+#| include: false
 
@@
 
-df = ffanalytics:::source_points(l, scoring_rules = ffanalytics::scoring)
+df <- ffanalytics:::source_points(data_result = l, scoring_rules = ffanalytics::scoring)
+```
+
+```{r}
+bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints %>% 
+  group_by(id) %>% 
+  summarise(
+    mean = mean(raw_points, na.rm = TRUE),
+    SD = sd(raw_points, na.rm = TRUE),
+    min = min(raw_points, na.rm = TRUE),
+    max = max(raw_points, na.rm = TRUE),
+    range = max(raw_points, na.rm = TRUE) - min(raw_points, na.rm = TRUE),
+    IQR = IQR(raw_points, na.rm = TRUE),
+    MAD = mad(raw_points, na.rm = TRUE),
+    CV = SD/mean,
+    median = median(raw_points, na.rm = TRUE),
+    pseudomedian = DescTools::HodgesLehmann(raw_points, na.rm = TRUE),
+    mode = petersenlab::Mode(raw_points, multipleModes = "mean"),
+    skewness = psych::skew(raw_points, na.rm = TRUE),
+    kurtosis = psych::kurtosi(raw_points, na.rm = TRUE)
+  )
+
+bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints_summary %>% 
+  left_join(
+    nfl_playerIDs[,c("mfl_id","name","merge_name","team","position")],
+    by = c("id" = "mfl_id")
+  ) %>% 
+  select(name, team, position, mean:kurtosis, everything())
+
+bootstrappedFantasyPoints_summary
+```
+
+```{r}
+#| eval: false
+
+bootstrappedFantasyPoints_summary %>% filter(mean > 100 & position == "QB") %>% arrange(CV) %>% head
+bootstrappedFantasyPoints_summary %>% filter(mean > 100 & position == "QB") %>% arrange(-CV) %>% head
+```
+
+```{r}
+#| label: fig-bootstrapping1
+#| fig-cap: "Distribution of Projected Fantasy Points for Patrick Mahomes from Bootstrapping."
+#| fig-alt: "Distribution of Projected Fantasy Points for Patrick Mahomes from Bootstrapping."
+
+ggplot2::ggplot(
+  data = bootstrappedFantasyPoints %>%
+    filter(pos == "QB" & name == "Patrick Mahomes"),
+  mapping = aes(
+    x = raw_points)
+) +
+  geom_histogram(
+    aes(y = after_stat(density)),
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  geom_density(
+    color = "#000000",
+    fill = "#F85700",
+    alpha = 0.6 # add transparency
+  ) +
+  geom_rug() +
+  #coord_cartesian(
+  #  xlim = c(0,400)) +
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    title = "Distribution of Projected Fantasy Points for Patrick Mahomes"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+#| label: fig-bootstrapping1
+#| fig-cap: "Distribution of Projected Fantasy Points for Two Players from Bootstrapping.
+There is relatively narrow uncertainty around projected fantasy points for Dak Prescott, whereas there is relatively wide uncertainty around the projected fantasy points for Drake Maye."
+#| fig-alt: "Distribution of Projected Fantasy Points for Two Players from Bootstrapping.
+There is relatively narrow uncertainty around projected fantasy points for Dak Prescott, whereas there is relatively wide uncertainty around the projected fantasy points for Drake Maye."
+
+ggplot2::ggplot(
+  data = bootstrappedFantasyPoints %>%
+    filter(pos == "QB" & (name %in% c("Dak Prescott", "Drake Maye"))),
+  mapping = aes(
+    x = raw_points,
+    group = name,
+    #color = name,
+    fill = name)
+) +
+  geom_histogram(
+    aes(y = after_stat(density))
+  ) +
+  geom_density(
+    alpha = 0.6 # add transparency
+  ) +
+  coord_cartesian(
+    xlim = c(0,NA),
+    expand = FALSE) +
+  #geom_rug() +
+  labs(
+    x = "Fantasy Points",
+    y = "Density",
+    fill = "",
+    color = "",
+    title = "Distribution of Projected Fantasy Points"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
 ```
@@
 ```{r}
+#| eval: false
+#| include: false
+
 calculate_fantasyPoints <- function(df, scoring_rules){

### 2025-06-10 simulation.qmd
20250609 - fix

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_playerIDs.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/players_projectedPoints_seasonal.RData")
+load(file = "./data/nfl_playerIDs.RData")
 ```
@@
 
+An example distribution of projected fantasy points is in @fig-bootstrapping1.
+
 ```{r}
@@
 
+Projections of two players—one with relatively narrow uncertainty and one with relatively wide uncertainty—are depicted in @fig-bootstrapping2.
+
 ```{r}
-#| label: fig-bootstrapping1
+#| label: fig-bootstrapping2
 #| fig-cap: "Distribution of Projected Fantasy Points for Two Players from Bootstrapping.

### 2025-06-10 evaluating-prediction-accuracy.qmd
20250609 - tweaks

@@
 - [discrimination](#sec-discrimination):
-  - [area under the receiver operating curve](#sec-auc)
+  - [area under the receiver operating characteristic curve](#sec-auc) (AUC)
   - and, secondarily—depending on the particulary cutoff and the relative costs of [false positives](#sec-falsePositive) versus [false negatives](#sec-falseNegative):
     - [sensitivity](#sec-sensitivity)
-    - [specificicity](#sec-specificity)
-    - [positive predictive value](#sec-ppv)
-    - [negative predictive value](#sec-npv)
+    - [specificity](#sec-specificity)
+    - [positive predictive value](#sec-ppv) (PPV)
+    - [negative predictive value](#sec-npv) (NPV)
 - [calibration](#sec-calibration):

### 2025-06-10 index.qmd
20250609 - update package list

@@
   "correlation", "mgcv", "rms", "car", "performance", "lavaan", "mice", 
-  "miceadds","interactions", "shiny","PearsonDS","robustbase"
+  "miceadds", "interactions", "shiny", "PearsonDS", "robustbase", "data.table", 
+  "future", "future.apply"
 )

### 2025-06-10 index.qmd
20250609 - copyediting

@@
 The GitHub repository for the book is located here:\index{GitHub} <https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook>.
-If you have data or analysis examples that are you willing to share and include in the book, feel free to contact me.
+If you have data or analysis examples that are you are willing to share and include in the book, feel free to contact me.
 
@@
 Through my training, I have learned the value of statistics for answering important questions that I find interesting.
-In graduate training, I came to the realization that statistics are relevant not only for psychology and science, but also for domains that I enjoy as hobbies, including sports and fantasy sports.
+In graduate training, I came to the realization that statistics is relevant not only for psychology and science, but also for domains that I enjoy as hobbies, including sports and fantasy sports.
 I have played in a longstanding fantasy football league for over 20 years (since my junior year of high school) with old friends from high school.
@@
 So I began blogging online about the value of applying statistics to improve decision making in fantasy football.
-Apparently, many people were interested in learning statistics when they could apply them to a domain that they find interesting like fantasy football.
+Apparently, many people were interested in learning statistics when they could apply them to a domain that they find interesting, like fantasy football.
 My blog eventually became [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), a website that uses advanced statistics to help people win their fantasy football leagues.
@@
 
-This book was supported by an OpenHawks Open Educational Resources grants program from the University of Iowa Libraries.
+This book was supported by an OpenHawks Open Educational Resources grant program from the University of Iowa Libraries.
 

### 2025-06-10 simulation.qmd
20250609 - fix

@@
 ```{r}
-all_proj <- bind_rows(players_projectedPoints_seasonal)
+all_proj <- dplyr::bind_rows(players_projectedPoints_seasonal)
 ```
@@
 bootstrapSimulation <- function(projectedStats, vars_by_pos, n_iter = 10000, seed = NULL) {
-  dt <- data.table::as.data.table(all_proj)
+  dt <- data.table::as.data.table(projectedStats)
   all_ids <- unique(dt$id)
@@
 
-  future::plan(multisession)
+  future::plan(future::multisession)
 
@@
       stat_vars <- vars_by_pos[[player_pos]]
-      out <- data.table::data.table(iteration = 1:n_iter, id = player_id, pos = player_pos)
+      out <- data.table(iteration = seq_len(n_iter), id = player_id, pos = player_pos)
 

### 2025-06-10 intro.qmd
20250308 - update citations
20250609 - copyediting

@@
 
-This book was originally written for a undergraduate-level course entitled, "Fantasy Football: Predictive Analytics and Empiricism".
-The chapters provide an overview of topics that each could have its own class and textbook, such as [causal inference](causal-inference.qmd), [factor analysis](factor-analysis.qmd), [cluster analysis](cluster-analysis.qmd), [principal component analysis](pca.qmd), [machine learning](machine-learning.qmd), [cognitive biases](cognitive-bias.qmd), [modern portfolio theory](modern-portfolio-theory.qmd), [data visualization](data-visualization.qmd), [simulation](simulation.qmd), etc.
+This book was originally written for an undergraduate-level course entitled, "Fantasy Football: Predictive Analytics and Empiricism".
+The chapters provide an overview of topics—each of which could have its own class and textbook, such as [causal inference](causal-inference.qmd), [factor analysis](factor-analysis.qmd), [cluster analysis](cluster-analysis.qmd), [principal component analysis](pca.qmd), [machine learning](machine-learning.qmd), [cognitive biases](cognitive-bias.qmd), [modern portfolio theory](modern-portfolio-theory.qmd), [data visualization](data-visualization.qmd), [simulation](simulation.qmd), etc.
 The book gives readers an overview of the breadth of the approaches to prediction and empiricism.
@@
 The book aims to help readers evaluate theory in the light of evidence (and vice versa) and to refine decision making in the context of uncertainty.
-Readers will learn about the ways that psychological science (and related disciplines) poses questions, formulates hypotheses, designs studies to test those questions, and interprets the findings, collectively with the aim to answer questions, improve decision making, and solve problems.
+Readers will learn about the ways that psychological science (and related disciplines) poses questions, formulates hypotheses, designs studies to test those questions, and interprets the findings, collectively with the aim of answering questions, improving decision making, and solving problems.
 
@@
 
-Here is a video of a Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football:
+Here is a video of Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football [@NFL2014]:
 

### 2025-06-10 intro.qmd
20250609 - learning goals and outcomes

@@
 
-## Learning Objectives {#sec-objectives}
+## Goals {#learningGoals}
 
-This book aims to help readers accomplish the following learning objectives:
+The book has three overarching goals; to help students:
 
-- Apply empirical inference and appreciate the value it provides over speculative supposition.
-- Ask educated questions when confronted with decisions in the face of uncertainty.
-- Understand human decision making, including common heuristics and cognitive biases and how to mitigate them analytically.
-- Engage in critical thinking about causality, including devising plausible alternative explanations for observed effects.
-- Understand causal inference including confounding, causal pathways, and counterfactuals.
-- Think empirically about human behavior and performance.
-- Describe the strengths and weaknesses of humans versus computers in prediction scenarios.
-- Apply basic skills in statistical programming using `R` to manipulate and summarize datasets and to conduct data analysis.
-- Critically evaluate the strengths and limitations of different statistical models and methodologies used in predicting uncertain events, enhancing their understanding of statistical inference and model selection.
-- Use various analytical techniques for predicting the outcome of uncertain events, and for uncovering latent causes of patterns in observed data.
-- Interpret findings from various statistical approaches and evaluate the accuracy of predictions.
-- Engage in iterative problem-solving processes, refining analytical approaches based on feedback and outcomes, and adapting strategies accordingly.
-- Communicate statistical findings and analyses in both written and oral formats, demonstrating proficiency in presenting complex information to diverse audiences.
-- Make sense of big data.
+1. Appreciate the value of statistics for answering questions you find interesting
+1. Think critically and be a critical consumer of information in society
+1. Get excited about learning stats!
+
+## Learning Outcomes {#learningOutcomes}
+
+Through reading the book, readers will be able to:
+
+1. Think critically and empirically about human behavior and performance
+1. Make sense of big data
+1. Use, evaluate, and interpret statistical analyses
+1. Communicate data-driven insights
+
+### How Readers Will Achieve the Learning Outcomes {#learningOutcomesHowAchieve}
+
+Readers will achieve the learning outcomes through the following specific outcomes:
+
+#### 1. Think critically and empirically about human behavior and performance {#learningOutcomesThinkCritically}
+
+- Apply empirical inference and articulate its advantages over speculative supposition.
+- Formulate research questions, hypotheses, and predictions.
+- Provide well-reasoned guidance for decision-making under conditions of uncertainty.
+- Describe common flaws in human judgment and decision making, including heuristics and cognitive biases, and explain how they can be mitigated analytically.
+- Critically evaluate causal claims by generating plausible alternative explanations.
+- Apply core concepts in causal inference, including confounding, causal pathways, and counterfactuals.
+- Compare the strengths and limitations of human and machine approaches to prediction.
+
+#### 2. Make sense of big data {#learningOutcomesBigData}
+
+- Apply foundational skills in statistical programming using `R` to combine, manipulate, clean, and summarize large datasets.
+
+#### 3. Use, evaluate, and interpret statistical analyses {#learningOutcomesAnalyses}
+
+- Conduct data analyses using `R`, including appropriate application of statistical models.
+- Critically evaluate the strengths and limitations of statistical models and methods used for predicting uncertain events.
+- Apply analytical techniques to predict outcomes and uncover latent causes in observed data.
+- Interpret results from statistical analyses and evaluate the accuracy of predictions.
+- Engage in iterative problem-solving, refining analytical strategies based on results and feedback.
 - Use practical analytical skills that can be applied in future research and job settings.
+
+#### 4. Communicate data-driven insights {#learningOutcomesCommunication}
+
+- Communicate statistical findings in writing.
+- Identify characteristics of effective data visualizations.
+- Create data visualizations.
 

### 2025-06-10 simulation.qmd
20250610 - updates

@@
     "rush_att", "rush_yds", "rush_tds",
-    "fumbles_lost"#,
-    #"sacks",
-    #"pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",  
-    #"pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_40_yds", "pass_400_yds",
-    #"rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds"
+    "fumbles_lost",
+    "sacks",
+    "pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",  
+    "pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_40_yds", "pass_400_yds",
+    "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds"
     ),
@@
     "fumbles_lost",
-    "return_yds", "return_tds"#,
-    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
-    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
-    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
+    "return_yds", "return_tds",
+    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
+    "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
+    "rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds"
   ),
@@
     "fumbles_lost",
-    "return_yds", "return_tds"#,
-    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
-    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
-    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
-  )
+    "return_yds", "return_tds",
+    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
+    "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
+    "rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds"
+  ),
+  K = c(
+    "fg_50", "fg_50_att", "fg_49", "fg_49_att", "fg_39", "fg_att_39", "fg", "fg_att", "xp", "xp_att"
+  ),
+  DST = c(
+    "dst_fum_recvr", "dst_sacks", "dst_fumbles", "dst_tackles", "dst_td", "dst_yds_against", "dst_pts_against"
+  ) ### ADD IDP VARS
 )
@@
 bootstrapSimulation <- function(projectedStats, vars_by_pos, n_iter = 10000, seed = NULL) {
-  dt <- data.table::as.data.table(projectedStats)
+  dt <- data.table::as.data.table(projectedStats) # using data.table for speed
   all_ids <- unique(dt$id)
@@
 
-  future::plan(future::multisession)
+  future::plan(future::multisession) # parallelize tasks across multiple background R sessions using multiple cores to speed up simulation
 
   results <- future.apply::future_lapply(
-    all_ids,
+    all_ids, # apply the function below to each player using a parallelized loop
     function(player_id) {
@@
 
-      stat_vars <- vars_by_pos[[player_pos]]
+      stat_vars <- vars_by_pos[[player_pos]] # pulls the relevant stat variables to simulate for this player's position
       out <- data.table(iteration = seq_len(n_iter), id = player_id, pos = player_pos)
 
-      for (var in stat_vars) {
+      for (var in stat_vars) { # Loop over each stat variable that should be simulated for the player's position
         if (var %in% names(player_data)) {
-          non_na_values <- player_data[[var]][!is.na(player_data[[var]])]
+          non_na_values <- player_data[[var]][!is.na(player_data[[var]])] # pulls non-missing values of the stat for the player (from all projection sources)
 
           if (length(non_na_values) > 0) {
-            out[[var]] <- sample(non_na_values, n_iter, replace = TRUE)
+            out[[var]] <- sample(non_na_values, n_iter, replace = TRUE) # If there are valid values, sample with replacement to simulate n_iter values
           } else {
-            out[[var]] <- NA_real_
+            out[[var]] <- NA_real_ # specifies a numeric missing value (if all values were missing)
           }
         } else {
-          out[[var]] <- NA_real_
+          out[[var]] <- NA_real_ # specifies a numeric missing value (if the stat variable doesn't exist)
         }
@@
     },
-    future.seed = TRUE
+    future.seed = TRUE # ensures that each parallel process gets a reproducible random seed
   )
 
-  data.table::rbindlist(results, use.names = TRUE, fill = TRUE)
+  data.table::rbindlist(results, use.names = TRUE, fill = TRUE) # combines all the individual player results into one large data table, aligning columns by name; fill = TRUE ensures that missing columns are filled with NA where necessary
 }
+```
 
+```{r}
 bootstappedStats <- bootstrapSimulation(
@@
   seed = 52242)
+```
 
+```{r}
 bootstappedStats$data_src <- bootstappedStats$iteration
@@
   )
+
+bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
+  rename(projectedPoints = raw_points) %>% 
+  arrange(-projectedPoints)
 ```
@@
   summarise(
-    mean = mean(raw_points, na.rm = TRUE),
-    SD = sd(raw_points, na.rm = TRUE),
-    min = min(raw_points, na.rm = TRUE),
-    max = max(raw_points, na.rm = TRUE),
-    range = max(raw_points, na.rm = TRUE) - min(raw_points, na.rm = TRUE),
-    IQR = IQR(raw_points, na.rm = TRUE),
-    MAD = mad(raw_points, na.rm = TRUE),
+    mean = mean(projectedPoints, na.rm = TRUE),
+    SD = sd(projectedPoints, na.rm = TRUE),
+    min = min(projectedPoints, na.rm = TRUE),
+    max = max(projectedPoints, na.rm = TRUE),
+    range = max(projectedPoints, na.rm = TRUE) - min(projectedPoints, na.rm = TRUE),
+    IQR = IQR(projectedPoints, na.rm = TRUE),
+    MAD = mad(projectedPoints, na.rm = TRUE),
     CV = SD/mean,
-    median = median(raw_points, na.rm = TRUE),
-    pseudomedian = DescTools::HodgesLehmann(raw_points, na.rm = TRUE),
-    mode = petersenlab::Mode(raw_points, multipleModes = "mean"),
-    skewness = psych::skew(raw_points, na.rm = TRUE),
-    kurtosis = psych::kurtosi(raw_points, na.rm = TRUE)
+    median = median(projectedPoints, na.rm = TRUE),
+    pseudomedian = DescTools::HodgesLehmann(projectedPoints, na.rm = TRUE),
+    mode = petersenlab::Mode(projectedPoints, multipleModes = "mean"),
+    skewness = psych::skew(projectedPoints, na.rm = TRUE),
+    kurtosis = psych::kurtosi(projectedPoints, na.rm = TRUE)
   )
@@
   ) %>% 
-  select(name, team, position, mean:kurtosis, everything())
+  select(name, team, position, mean:kurtosis, everything()) %>% 
+  arrange(-mean)
+```
 
-bootstrappedFantasyPoints_summary
+```{r}
+bootstrappedFantasyPoints_summary %>% 
+  filter(position == "QB") %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
+
+bootstrappedFantasyPoints_summary %>% 
+  filter(position == "RB") %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
+
+bootstrappedFantasyPoints_summary %>% 
+  filter(position == "WR") %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
+
+bootstrappedFantasyPoints_summary %>% 
+  filter(position == "TE") %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
+
+bootstrappedFantasyPoints_summary %>% 
+  filter(position == "PK") %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
 ```
@@
 #| eval: false
+#| include: false
 

### 2025-06-10 index.qmd
20250609 - typo
20250610 - acknowledgment

@@
 The GitHub repository for the book is located here:\index{GitHub} <https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook>.
-If you have data or analysis examples that are you are willing to share and include in the book, feel free to contact me.
+If you have data or analysis examples that you are willing to share and include in the book, feel free to contact me.
 
@@
 I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, and Adam Peterson, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
+In particular, Andrew Tungate made important additions to the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] that were necessary for conducting a [bootstrap simulation](#sec-simulationProjectedPointsBootstrapping) of players' projected fantasy points.
 I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.

### 2025-06-10 simulation.qmd
20250610 - fix

@@
   mapping = aes(
-    x = raw_points)
+    x = projectedPoints)
 ) +
@@
   mapping = aes(
-    x = raw_points,
+    x = projectedPoints,
     group = name,

### 2025-06-10 correlation.qmd
20250610 - cor to cov

@@
 $$
-\text{Cov}(x, y) = \frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})
+\begin{aligned}
+  \text{Cov}(x, y) &= \frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})\\
+  &= r_{xy} \cdot \sigma_x \cdot \sigma_y
+\end{aligned}
 $$ {#eq-covariance}
+
+where $n$ is the sample size, $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), $x_i$ is the ith observation on the [predictor variable](#sec-correlationalStudy), $y_i$ is the ith observation of the [outcome variable](#sec-correlationalStudy), $\bar{x}$ is the mean of the [predictor variable](#sec-correlationalStudy), $\bar{y}$ is the mean of the [outcome variable](#sec-correlationalStudy), $r_{xy}$ is the correlation between the [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variables, $\sigma_x$ is the standard deviation of the [predictor variable](#sec-correlationalStudy), and $\sigma_y$ is the standard deviation of the [outcome variable](#sec-correlationalStudy).
 
@@
   use = "pairwise.complete.obs")
+```
+
+We can also convert a correlation back to covariance using the formula in @eq-covariance.
+
+```{r}
+cor_value * sd(complete_cases$height, na.rm = TRUE) * sd(complete_cases$fantasyPoints, na.rm = TRUE)
 ```

### 2025-06-12 basic-statistics.qmd
20250612 - coefficient of determination

@@
 
-where $y$ is the [outcome variable](#sec-correlationalStudy), $y_i$ is the observed value of the [outcome variable](#sec-correlationalStudy) for the $i$th observation, $\hat{y}_i$ is the model predicted value for the $i$th observation, $\bar{y}$ is the mean of the observed values of the [outcome variable](#sec-correlationalStudy).
+where $y$ is the [outcome variable](#sec-correlationalStudy), $y_i$ is the observed value of the [outcome variable](#sec-correlationalStudy) for the $i$th observation, $\hat{y}_i$ is the model predicted value for the $i$th observation, and $\bar{y}$ is the mean of the observed values of the [outcome variable](#sec-correlationalStudy).
 The total sum of squares is an index of the total variation in the [outcome variable](#sec-correlationalStudy).
+
+The coefficient of determination typically ranges from 0 to 1, but it can be negative.
+A negative coefficient of determination indicates that the predicted values are less accurate than using the mean of the actual values as the predicted values (i.e., predicting the mean of the observed values for every case).
 

### 2025-06-12 decision-making.qmd
20250612 - tweaks

@@
 - Seek advice from diverse perspectives and incorporate it into your decision making.
-- Try to get opinions from others *before* you state your perspective and before the various sources of advice discuss with each other, to ensure independence of advice.
+- Get opinions from others *before* you state your perspective and before the various sources of advice discuss with each other, to ensure independence of advice.
   - As noted by @Kahneman2011, "before an issue is discussed, all members of the committee should be asked to write a very brief summary of their position...
@@
 85).
-- If some sources of advice (or projections) are clearly more skilled and accurate than others, you can average this "select" crowd of projections or give them greater weight in a [weighted average]().
+- If some sources of advice (or projections) are clearly more skilled and accurate than others, you can average this "select" crowd of projections or give them greater weight in a [weighted average](#sec-descriptiveStatisticsCenter).
 - If it is unclear whether or which sources are reliably more accurate than others, using a simple average across all sources (i.e., crowd-averaged projections) can be a useful approach that is as accurate as—if not more accurate than—the average individual forecaster.

### 2025-06-15 simulation.qmd
20250615 - progress bar

@@
 library("future.apply")
+library("progressr")
 library("tidyverse")
@@
 ```{r}
-bootstrapSimulation <- function(projectedStats, vars_by_pos, n_iter = 10000, seed = NULL) {
-  dt <- data.table::as.data.table(projectedStats) # using data.table for speed
+bootstrapSimulation <- function(
+    projectedStats,
+    vars_by_pos,
+    n_iter = 10000,
+    seed = NULL,
+    progress = TRUE) {
+
+  dt <- data.table::as.data.table(projectedStats) # use data.table for speed
   all_ids <- unique(dt$id)
@@
 
-  results <- future.apply::future_lapply(
+  if (progress) progressr::handlers("txtprogressbar") # specify progress-bar style
+
+  results <- progressr::with_progress({ # wrap in with_progress for progress bar
+    p <- if (progress) progressr::progressor(along = all_ids) else NULL # create progressor for progress bar
+
+    future.apply::future_lapply(
       all_ids, # apply the function below to each player using a parallelized loop
       function(player_id) {
+        if (!is.null(p)) p() # advance progress bar
+
         player_data <- dt[id == player_id]
@@
 
-      stat_vars <- vars_by_pos[[player_pos]] # pulls the relevant stat variables to simulate for this player's position
+        stat_vars <- vars_by_pos[[player_pos]] # pull the relevant stat variables to simulate for this player's position
         out <- data.table(iteration = seq_len(n_iter), id = player_id, pos = player_pos)
 
-      for (var in stat_vars) { # Loop over each stat variable that should be simulated for the player's position
+        for (var in stat_vars) { # loop over each stat variable that should be simulated for the player's position
           if (var %in% names(player_data)) {
-          non_na_values <- player_data[[var]][!is.na(player_data[[var]])] # pulls non-missing values of the stat for the player (from all projection sources)
+            non_na_values <- player_data[[var]][!is.na(player_data[[var]])] # pull non-missing values of the stat for the player (from all projection sources)
 
             if (length(non_na_values) > 0) {
-            out[[var]] <- sample(non_na_values, n_iter, replace = TRUE) # If there are valid values, sample with replacement to simulate n_iter values
+              out[[var]] <- sample(non_na_values, n_iter, replace = TRUE) # if there are valid values, sample with replacement to simulate n_iter values
             } else {
-            out[[var]] <- NA_real_ # specifies a numeric missing value (if all values were missing)
+              out[[var]] <- NA_real_ # specify a numeric missing value (if all values were missing)
             }
           } else {
-          out[[var]] <- NA_real_ # specifies a numeric missing value (if the stat variable doesn't exist)
+            out[[var]] <- NA_real_ # specify a numeric missing value (if the stat variable doesn't exist)
           }
@@
     )
+  })
 
@@
 ```{r}
+#| eval: false
+
 bootstappedStats <- bootstrapSimulation(
@@
   seed = 52242)
+```
+
+```{r}
+#| include: false
+
+bootstappedStats <- bootstrapSimulation(
+  projectedStats = all_proj,
+  vars_by_pos = vars_by_pos,
+  n_iter = 10000,
+  seed = 52242,
+  progress = FALSE)
 ```
@@
     SD = sd(projectedPoints, na.rm = TRUE),
+    floor = quantile(projectedPoints, .10), # 10th quantile
+    ceiling = quantile(projectedPoints, .90), # 90th quantile
     min = min(projectedPoints, na.rm = TRUE),

### 2025-06-17 simulation.qmd
20250616 - add vars

@@
     "games",
-    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
+    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
     "rush_att", "rush_yds", "rush_tds",
-    "fumbles_lost",
+    "fumbles_lost", "fumbles_total", "two_pts",
     "sacks",
     "pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",
-    "pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_40_yds", "pass_400_yds",
-    "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds"
+    "pass_40_yds", "pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_400_yds",
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
     ),
@@
     "rush_att", "rush_yds", "rush_tds",
-    "rec_tgt", "rec", "rec_yds", "rec_tds",
-    "fumbles_lost",
+    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
+    "fumbles_lost", "fumbles_total", "two_pts",
     "return_yds", "return_tds",
     "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
-    "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
-    "rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds"
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
+    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
   ),
@@
     "games",
-    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
+    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
     "rush_att", "rush_yds", "rush_tds",
-    "rec_tgt", "rec", "rec_yds", "rec_tds",
-    "fumbles_lost",
+    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
+    "fumbles_lost", "fumbles_total", "two_pts",
     "return_yds", "return_tds",
     "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
-    "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
-    "rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds"
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
+    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
   ),
@@
     "games",
-    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
+    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
     "rush_att", "rush_yds", "rush_tds",
-    "rec_tgt", "rec", "rec_yds", "rec_tds",
-    "fumbles_lost",
+    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
+    "fumbles_lost", "fumbles_total", "two_pts",
     "return_yds", "return_tds",
     "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
-    "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
-    "rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds"
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
+    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
   ),
   K = c(
-    "fg_50", "fg_50_att", "fg_49", "fg_49_att", "fg_39", "fg_att_39", "fg", "fg_att", "xp", "xp_att"
+    "fg_0019", "fg_2029", "fg_3039", "fg_4049", "fg_50", "fg_50_att",
+    "fg_39", "fg_att_39", "fg_49", "fg_49_att",
+    "fg", "fg_att", "fg_miss", "xp", "xp_att"
+  ),
+  IDP = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  DL = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  LB = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  DB = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
   ),
   DST = c(
-    "dst_fum_recvr", "dst_sacks", "dst_fumbles", "dst_tackles", "dst_td", "dst_yds_against", "dst_pts_against"
-  ) ### ADD IDP VARS
+    "dst_fum_recvr", "dst_fum_rec", "dst_int", "dst_safety", "dst_sacks", "dst_td", "dst_blk",
+    "dst_fumbles", "dst_tackles", "dst_yds_against", "dst_pts_against", "dst_pts_allowed", "dst_ret_yds"
+  )
 )

### 2025-06-17 simulation.qmd
20250616 - add IDP

--- 2025-06-15 simulation.qmd
@@
     "games",
-    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
+    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
     "rush_att", "rush_yds", "rush_tds",
-    "fumbles_lost",
+    "fumbles_lost", "fumbles_total", "two_pts",
     "sacks",
     "pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",
-    "pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_40_yds", "pass_400_yds",
-    "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds"
+    "pass_40_yds", "pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_400_yds",
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds"
     ),
@@
     "games",
-    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
+    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
     "rush_att", "rush_yds", "rush_tds",
-    "rec_tgt", "rec", "rec_yds", "rec_tds",
-    "fumbles_lost",
+    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
+    "fumbles_lost", "fumbles_total", "two_pts",
     "return_yds", "return_tds",
     "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
-    "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
-    "rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds"
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
+    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
   ),
   K = c(
-    "fg_50", "fg_50_att", "fg_49", "fg_49_att", "fg_39", "fg_att_39", "fg", "fg_att", "xp", "xp_att"
+    "fg_0019", "fg_2029", "fg_3039", "fg_4049", "fg_50", "fg_50_att",
+    "fg_39", "fg_att_39", "fg_49", "fg_49_att",
+    "fg", "fg_att", "fg_miss", "xp", "xp_att"
+  ),
+  D = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  DL = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  LB = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  DB = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
   ),
   DST = c(
-    "dst_fum_recvr", "dst_sacks", "dst_fumbles", "dst_tackles", "dst_td", "dst_yds_against", "dst_pts_against"
-  ) ### ADD IDP VARS
+    "dst_fum_recvr", "dst_fum_rec", "dst_int", "dst_safety", "dst_sacks", "dst_td", "dst_blk",
+    "dst_fumbles", "dst_tackles", "dst_yds_against", "dst_pts_against", "dst_pts_allowed", "dst_ret_yds"
+  )
 )
@@
       \(x) round(x, digits = 2)))
+
+bootstrappedFantasyPoints_summary %>% 
+  filter(position %in% c("DL","DT","DE")) %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
+
+bootstrappedFantasyPoints_summary %>% 
+  filter(position %in% c("LB","MLB","OLB")) %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
+
+bootstrappedFantasyPoints_summary %>% 
+  filter(position %in% c("DB","S","CB")) %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
 ```

### 2025-06-17 basic-statistics.qmd
20250616 - update formulas

@@
 
-- mean (aka "average")
+- mean (aka "average") (i.e., $M$ or $\mu$)
 - median
@@
 
-The mean of $x$ (written as: $\bar{x}$) is calculated as in @eq-IQR:
+The mean of $x$ (written as: $\bar{x}$ or $\mu_x$) is calculated as in @eq-mean:
 
@@
 \bar{x} = \frac{\sum x_i}{n} = \frac{x_1 + x_2 + ... + x_n}{n}
-$$ {#eq-IQR}
+$$ {#eq-mean}
 
@@
 
-- standard deviation
-- variance
+- standard deviation ($s$ or $\sigma$)
+- variance ($s^2$ or $\sigma^2$)
 - range
@@
 where $x_i$ is each data point, $\bar{x}$ is the mean of $x$, and $n$ is the number of data points.
+The population variance divides by $n$ rather than $n-1$ and is written as $\sigma^2$.
 
@@
 
+The population standard deviation divides by $n$ rather than $n-1$ and is written as $\sigma$.
+
 The range is calculated of $x$ is calculated as in @eq-range:
@@
 
-The coefficient of variation (CV) accounts for the fact that variables with larger means also tend to have larger standard deviations; it is calculated as the standard deviation divided by the mean, as in @eq-coefficientOfVariation:
-
-$$
-\text{CV} = \frac{SD(x)}{\bar{x}}
-$$ {#eq-coefficientOfVariation}
-
-where $SD(x)$ is the standard deviation of $x$.
+The coefficient of variation (CV) accounts for the fact that variables with larger means also tend to have larger standard deviations; it is calculated as the standard deviation divided by the mean, as in @eq-coefficientOfVariation.
 
@@
 $$
-z = \frac{x - \bar{x}}{\sigma}
+z = \frac{x - \bar{x}}{s_x}
 $$ {#eq-zScore}
 
-where $x$ is the observed score, $\bar{x}$ is the mean observed score, and $\sigma$ is the standard deviation of the observed scores.
+where $x$ is the observed score, $\bar{x}$ is the mean observed score, and $s_x$ is the standard deviation of the observed scores.
 
@@
   && \quad \text{$n-1$ cancels out in numerator and denominator (typical formula for Pearson correlation)} \\
-    &= \frac{\frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})}{\sigma_x \cdot \sigma_y} 
+    &= \frac{\frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})}{s_x \cdot s_y} 
   && \quad \text{adding back $n-1$, replacing denominator from line 1 with standard deviation symbols} \\
-    &= \frac{\text{Cov}(x, y)}{\sigma_x \cdot \sigma_y} 
+    &= \frac{\text{Cov}(x, y)}{s_x \cdot s_y} 
   && \quad \text{replacing numerator with covariance notation}
@@
 
-where $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), and $\text{Cov}(x, y)$ is the covariance between $x$ and $y$.
+where $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), $\text{Cov}(x, y)$ is the covariance between $x$ and $y$, $s_x$ is the standard deviation of $x$, and $s_y$ is the standard deviation of $y$.
 

### 2025-06-17 correlation.qmd
20250616 - update formulas

@@
   \text{Cov}(x, y) &= \frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})\\
-  &= r_{xy} \cdot \sigma_x \cdot \sigma_y
+  &= r_{xy} \cdot s_x \cdot s_y
 \end{aligned}
@@
 
-where $n$ is the sample size, $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), $x_i$ is the ith observation on the [predictor variable](#sec-correlationalStudy), $y_i$ is the ith observation of the [outcome variable](#sec-correlationalStudy), $\bar{x}$ is the mean of the [predictor variable](#sec-correlationalStudy), $\bar{y}$ is the mean of the [outcome variable](#sec-correlationalStudy), $r_{xy}$ is the correlation between the [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variables, $\sigma_x$ is the standard deviation of the [predictor variable](#sec-correlationalStudy), and $\sigma_y$ is the standard deviation of the [outcome variable](#sec-correlationalStudy).
+where $n$ is the sample size, $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), $x_i$ is the ith observation on the [predictor variable](#sec-correlationalStudy), $y_i$ is the ith observation of the [outcome variable](#sec-correlationalStudy), $\bar{x}$ is the mean of the [predictor variable](#sec-correlationalStudy), $\bar{y}$ is the mean of the [outcome variable](#sec-correlationalStudy), $r_{xy}$ is the correlation between the [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variables, $s_x$ is the standard deviation of the [predictor variable](#sec-correlationalStudy), and $s_y$ is the standard deviation of the [outcome variable](#sec-correlationalStudy).
 

### 2025-06-17 player-evaluation.qmd
20250616 - update formulas

@@
 However, the standard deviation tends to be upwardly biased as the mean increases.
-So, we can account for the player's mean fantasy points per game by dividing their game-to-game standard deviation of fantasy points ($\sigma$) by their mean fantasy points across games ($\mu$).
+So, we can account for the player's mean fantasy points per game by dividing their game-to-game standard deviation of fantasy points ($s_x$ or $\sigma_x$) by their mean fantasy points across games ($\bar{x}$ or $\mu_x$).
 This is known as the coefficient of variation (CV), which is provided in @eq-coefficientOfVariation.
@@
 $$
-CV = \frac{\sigma}{\mu}
+\begin{aligned}
+  \text{CV} &= \frac{s_x}{\bar{x}}\\
+  &= \frac{\sigma_x}{\mu_x}
+\end{aligned}
 $$ {#eq-coefficientOfVariation}

### 2025-06-17 simulation.qmd
20250617 - temporarily drop IDP

--- 2025-06-15 simulation.qmd
@@
     "games",
-    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
+    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
     "rush_att", "rush_yds", "rush_tds",
-    "rec_tgt", "rec", "rec_yds", "rec_tds",
-    "fumbles_lost",
+    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
+    "fumbles_lost", "fumbles_total", "two_pts",
     "return_yds", "return_tds",
     "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
-    "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
-    "rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds"
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
+    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
   ),
   K = c(
-    "fg_50", "fg_50_att", "fg_49", "fg_49_att", "fg_39", "fg_att_39", "fg", "fg_att", "xp", "xp_att"
-  ),
-  DST = c(
-    "dst_fum_recvr", "dst_sacks", "dst_fumbles", "dst_tackles", "dst_td", "dst_yds_against", "dst_pts_against"
-  ) ### ADD IDP VARS
+    "fg_0019", "fg_2029", "fg_3039", "fg_4049", "fg_50", "fg_50_att",
+    "fg_39", "fg_att_39", "fg_49", "fg_49_att",
+    "fg", "fg_att", "fg_miss", "xp", "xp_att"
+  )#,
+  #D = c(
+  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  #),
+  #DL = c(
+  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  #),
+  #LB = c(
+  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  #),
+  #DB = c(
+  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  #),
+  #DST = c(
+  #  "dst_fum_recvr", "dst_fum_rec", "dst_int", "dst_safety", "dst_sacks", "dst_td", "dst_blk",
+  #  "dst_fumbles", "dst_tackles", "dst_yds_against", "dst_pts_against", "dst_pts_allowed", "dst_ret_yds"
+  #)
 )

### 2025-06-17 simulation.qmd
20250617 - chunk labels for troubleshooting

--- 2025-06-17 simulation.qmd
@@
 ```{r}
+#| label: vars-to-bootstrap-by-position
+
 vars_by_pos <- list(
@@
     "pass_40_yds", "pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_400_yds",
-    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds"
     ),
@@
     "fg", "fg_att", "fg_miss", "xp", "xp_att"
-  ),
-  IDP = c(
-    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  ),
-  DL = c(
-    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  ),
-  LB = c(
-    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  ),
-  DB = c(
-    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  ),
-  DST = c(
-    "dst_fum_recvr", "dst_fum_rec", "dst_int", "dst_safety", "dst_sacks", "dst_td", "dst_blk",
-    "dst_fumbles", "dst_tackles", "dst_yds_against", "dst_pts_against", "dst_pts_allowed", "dst_ret_yds"
-  )
+  )#,
+  #D = c(
+  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  #),
+  #DL = c(
+  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  #),
+  #LB = c(
+  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  #),
+  #DB = c(
+  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  #),
+  #DST = c(
+  #  "dst_fum_recvr", "dst_fum_rec", "dst_int", "dst_safety", "dst_sacks", "dst_td", "dst_blk",
+  #  "dst_fumbles", "dst_tackles", "dst_yds_against", "dst_pts_against", "dst_pts_allowed", "dst_ret_yds"
+  #)
 )
@@
 ```{r}
+#| label: function-for-bootstrap-simulation
+
 bootstrapSimulation <- function(
@@
 ```{r}
+#| label: code-to-run-bootstrap-simulation
 #| eval: false
@@
 ```{r}
+#| label: run-bootstrap-simulation
 #| include: false
@@
 ```{r}
+#| label: score-bootstrap-simulation
+
 bootstappedStats$data_src <- bootstappedStats$iteration
@@
 ```{r}
+#| label: compute-fantasy-points-example
 #| eval: false
@@
 ```{r}
+#| label: summarize-bootstrap-simulation-scores
+
 bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints %>% 
@@
 ```{r}
+#| label: display-bootstrap-simulation-summaries
+
 bootstrappedFantasyPoints_summary %>% 
@@
       \(x) round(x, digits = 2)))
+
+#bootstrappedFantasyPoints_summary %>% 
+#  filter(position %in% c("DL","DT","DE")) %>% 
+#  mutate(
+#    across(
+#      where(is.numeric),
+#      \(x) round(x, digits = 2)))
+#
+#bootstrappedFantasyPoints_summary %>% 
+#  filter(position %in% c("LB","MLB","OLB")) %>% 
+#  mutate(
+#    across(
+#      where(is.numeric),
+#      \(x) round(x, digits = 2)))
+#
+#bootstrappedFantasyPoints_summary %>% 
+#  filter(position %in% c("DB","S","CB")) %>% 
+#  mutate(
+#    across(
+#      where(is.numeric),
+#      \(x) round(x, digits = 2)))
 ```
@@
 ```{r}
+#| label: identify-players-with-lo-or-hi-cv
 #| eval: false

### 2025-06-17 simulation.qmd
20250617 - better memory handling

@@
 ```{r}
+#| label: code-to-specify-vars-to-bootstrap-by-position
+#| eval: false
+
+vars_by_pos <- list(
+  QB = c(
+    "games",
+    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
+    "rush_att", "rush_yds", "rush_tds",
+    "fumbles_lost", "fumbles_total", "two_pts",
+    "sacks",
+    "pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",
+    "pass_40_yds", "pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_400_yds",
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds"
+    ),
+  RB = c(
+    "games",
+    "rush_att", "rush_yds", "rush_tds",
+    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
+    "fumbles_lost", "fumbles_total", "two_pts",
+    "return_yds", "return_tds",
+    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
+    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
+  ),
+  WR = c(
+    "games",
+    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
+    "rush_att", "rush_yds", "rush_tds",
+    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
+    "fumbles_lost", "fumbles_total", "two_pts",
+    "return_yds", "return_tds",
+    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
+    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
+  ),
+  TE = c(
+    "games",
+    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
+    "rush_att", "rush_yds", "rush_tds",
+    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
+    "fumbles_lost", "fumbles_total", "two_pts",
+    "return_yds", "return_tds",
+    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
+    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
+    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
+  ),
+  K = c(
+    "fg_0019", "fg_2029", "fg_3039", "fg_4049", "fg_50", "fg_50_att",
+    "fg_39", "fg_att_39", "fg_49", "fg_49_att",
+    "fg", "fg_att", "fg_miss", "xp", "xp_att"
+  ),
+  D = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  DL = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  LB = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  DB = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  DST = c(
+    "dst_fum_recvr", "dst_fum_rec", "dst_int", "dst_safety", "dst_sacks", "dst_td", "dst_blk",
+    "dst_fumbles", "dst_tackles", "dst_yds_against", "dst_pts_against", "dst_pts_allowed", "dst_ret_yds"
+  )
+)
+```
+
+```{r}
 #| label: vars-to-bootstrap-by-position
@@
     "fg", "fg_att", "fg_miss", "xp", "xp_att"
-  )#,
-  #D = c(
-  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  #),
-  #DL = c(
-  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  #),
-  #LB = c(
-  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  #),
-  #DB = c(
-  #  "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  #),
-  #DST = c(
-  #  "dst_fum_recvr", "dst_fum_rec", "dst_int", "dst_safety", "dst_sacks", "dst_td", "dst_blk",
-  #  "dst_fumbles", "dst_tackles", "dst_yds_against", "dst_pts_against", "dst_pts_allowed", "dst_ret_yds"
-  #)
+  ),
+  D = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  DL = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  LB = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  DB = c(
+    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
+  ),
+  DST = c(
+    "dst_fum_recvr", "dst_fum_rec", "dst_int", "dst_safety", "dst_sacks", "dst_td", "dst_blk",
+    "dst_fumbles", "dst_tackles", "dst_yds_against", "dst_pts_against", "dst_pts_allowed", "dst_ret_yds"
+  )
 )
@@
 ```{r}
-#| label: score-bootstrap-simulation
+#| label: free-up-memory1
+#| include: false
+
+rm(all_proj)
+gc()
+```
+
+```{r}
+#| label: put-simulated-data-in-format-for-scoring
 
@@
 attr(bootstappedStatsByPosition, "week") <- 0
+```
+
+```{r}
+#| label: free-up-memory2
+#| include: false
+
+rm(bootstappedStats)
+gc()
+```
+
+```{r}
+#| label: score-bootstrap-simulation
 
@@
   scoring_rules = ffanalytics::scoring)
+```
+
+```{r}
+#| label: free-up-memory3
+#| include: false
+
+rm(bootstappedStatsByPosition)
+gc()
+```
+
+```{r}
+#| label: postprocessing-bootstrap-simulation
 
@@
 bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
-  left_join(
-    nfl_playerIDs[,c("mfl_id","name","merge_name","team")],
-    by = c("id" = "mfl_id")
-  )
+  rename(projectedPoints = raw_points)
+```
 
-bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
-  rename(projectedPoints = raw_points) %>% 
-  arrange(-projectedPoints)
+```{r}
+#| label: free-up-memory4
+#| include: false
+
+gc()
 ```
@@
   )
+```
+
+```{r}
+#| label: free-up-memory5
+#| include: false
+
+rm(bootstrappedFantasyPoints)
+gc()
+```
+
+```{r}
+#| label: merge-player-info
 
@@
 ```{r}
-#| label: display-bootstrap-simulation-summaries
+#| label: free-up-memory6
+#| include: false
+
+gc()
+```
+
+```{r}
+#| label: display-bootstrap-simulation-summaries-qb
 
@@
       \(x) round(x, digits = 2)))
+```
+
+```{r}
+#| label: display-bootstrap-simulation-summaries-rb
 
@@
       \(x) round(x, digits = 2)))
+```
+
+```{r}
+#| label: display-bootstrap-simulation-summaries-wr
 
@@
       \(x) round(x, digits = 2)))
+```
+
+```{r}
+#| label: display-bootstrap-simulation-summaries-te
 
@@
       \(x) round(x, digits = 2)))
+```
+
+```{r}
+#| label: display-bootstrap-simulation-summaries-k
 
@@
       \(x) round(x, digits = 2)))
+```
 
-#bootstrappedFantasyPoints_summary %>% 
-#  filter(position %in% c("DL","DT","DE")) %>% 
-#  mutate(
-#    across(
-#      where(is.numeric),
-#      \(x) round(x, digits = 2)))
-#
-#bootstrappedFantasyPoints_summary %>% 
-#  filter(position %in% c("LB","MLB","OLB")) %>% 
-#  mutate(
-#    across(
-#      where(is.numeric),
-#      \(x) round(x, digits = 2)))
-#
-#bootstrappedFantasyPoints_summary %>% 
-#  filter(position %in% c("DB","S","CB")) %>% 
-#  mutate(
-#    across(
-#      where(is.numeric),
-#      \(x) round(x, digits = 2)))
+```{r}
+#| label: display-bootstrap-simulation-summaries-dl
+
+bootstrappedFantasyPoints_summary %>% 
+  filter(position %in% c("DL","DT","DE")) %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
+```
+
+```{r}
+#| label: display-bootstrap-simulation-summaries-lb
+
+bootstrappedFantasyPoints_summary %>% 
+  filter(position %in% c("LB","MLB","OLB")) %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
+```
+
+```{r}
+#| label: display-bootstrap-simulation-summaries-db
+
+bootstrappedFantasyPoints_summary %>% 
+  filter(position %in% c("DB","S","CB")) %>% 
+  mutate(
+    across(
+      where(is.numeric),
+      \(x) round(x, digits = 2)))
 ```

### 2025-06-17 basic-statistics.qmd
20250617 - additional documentation

@@
 
-The (sample) variance of $x$ (written as: $s^2$) is calculated as in @eq-variance:
+The ([sample](#sec-sampleVsPopulation)) variance of $x$ (written as: $s^2$) is calculated as in @eq-variance:
 
@@
 where $x_i$ is each data point, $\bar{x}$ is the mean of $x$, and $n$ is the number of data points.
-The population variance divides by $n$ rather than $n-1$ and is written as $\sigma^2$.
-
-The (sample) standard deviation of $x$ (written as: $s$) is calculated as in @eq-sd:
+The [population](#sec-sampleVsPopulation) variance divides by $n$ rather than $n-1$ and is written as $\sigma^2$.
+
+The ([sample](#sec-sampleVsPopulation)) standard deviation of $x$ (written as: $s$) is calculated as in @eq-sd:
 
@@
 
-The population standard deviation divides by $n$ rather than $n-1$ and is written as $\sigma$.
+The [population](#sec-sampleVsPopulation) standard deviation divides by $n$ rather than $n-1$ and is written as $\sigma$.
 

### 2025-06-17 getting-started.qmd
20250617 - additional documentation

@@
 
+You can load a given library using the following syntax:
+
+```{r}
+#| eval: false
+
+library("INSERT_PACKAGE_NAME_HERE")
+```
+
+For instance, you can use the following code to load the `tidyverse` package [@R-tidyverse]:
+
 ```{r}
@@
 
+A function often takes particular input(s) and produces some form of output.
+The name of the function is followed by parentheses; the inputs go in between the parentheses.
 You can learn about a particular function and its arguments by entering a question mark before the name of the function:
@@
 It has one dimension (length).
-To create a vector, use the `c()` to combine elements into a vector.
+To create a vector, use the `c()` function to combine elements into a vector.
 And, we use the assignment operator (`<-`) to assign the vector to an object named `exampleVector`, so we can access it later.
+Anything on the right side of an assignment operator gets assigned to the object name on the left side of the assignment operator.
 
@@
 
+Loops are fine for basic computations, but other approaches (such as the `apply()` family of functions) can be even faster than loops.
+
 ## Session Info {#sec-gettingStartedSessionInfo}

### 2025-06-17 simulation.qmd
20250617 - additional documentation

@@
 For instance, if we know the population distribution for the likely outcomes for a player—e.g., a normal distribution with a known mean (e.g., 150 points) and standard deviation (e.g., 20 points)—we can repeatedly sample randomly from this distribution.
+The distribution could be, as examples, a normal distribution, a log-normal distribution, a binomial distribution, a chi-square distribution, etc. [@Sigal2016].
+The distribution provides a probability density function, which indicates the probability that any particular value would be observed if the data arose from that distribution.
 

### 2025-06-17 index.qmd
20250617 - SimDesign

@@
   "miceadds", "interactions", "shiny", "PearsonDS", "robustbase", "data.table", 
-  "future", "future.apply"
+  "future", "future.apply", "SimDesign"
 )

### 2025-06-17 simulation.qmd
20250617 - SimDesign

@@
 library("progressr")
+library("SimDesign")
 library("tidyverse")
@@
 
+You can generate a template for Monte Carlo simulations in the `SimDesign` package using the following code:
+
+```{r}
+SimDesign::SimFunctions()
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+#-------------------------------------------------------------------
+
+Design <- SimDesign::createDesign(
+  factor1 = NA,
+  factor2 = NA)
+
+#-------------------------------------------------------------------
+
+Generate <- function(condition, fixed_objects) {
+  dat <- data.frame()
+  dat
+}
+
+Analyse <- function(condition, dat, fixed_objects) {
+  ret <- nc(stat1 = NaN, stat2 = NaN)
+  ret
+}
+
+Summarise <- function(condition, results, fixed_objects) {
+  ret <- c(bias = NaN, RMSE = NaN)
+  ret
+}
+
+#-------------------------------------------------------------------
+
+res <- SimDesign::runSimulation(
+  design = Design,
+  replications = 2,
+  generate = Generate,
+  analyse = Analyse,
+  summarise = Summarise)
+
+res
+```
+
 ```{r}

### 2025-06-17 simulation.qmd
20250617 - memory handling

@@
 ```{r}
-#| label: put-simulated-data-in-format-for-scoring
+#| label: put-simulated-data-in-format-for-scoring1
 
-bootstappedStats$data_src <- bootstappedStats$iteration
+data.table::setnames(bootstappedStats, "iteration", "data_src") # data.table equivalent to: bootstappedStats$data_src <- bootstappedStats$iteration
 
@@
   keep.by = TRUE)
-
-bootstappedStatsByPosition <- lapply(
-  bootstappedStatsByPosition,
-  as_tibble)
-
-attr(bootstappedStatsByPosition, "season") <- 2024
-attr(bootstappedStatsByPosition, "week") <- 0
 ```
@@
 rm(bootstappedStats)
+gc()
+```
+
+```{r}
+#| label: put-simulated-data-in-format-for-scoring1
+
+bootstappedStatsByPosition <- lapply(
+  bootstappedStatsByPosition,
+  setDF)
+
+attr(bootstappedStatsByPosition, "season") <- 2024
+attr(bootstappedStatsByPosition, "week") <- 0
+```
+
+```{r}
+#| label: free-up-memory3
+#| include: false
+
 gc()
@@
 ```{r}
-#| label: free-up-memory3
+#| label: free-up-memory4
 #| include: false
@@
 ```{r}
-#| label: free-up-memory4
+#| label: free-up-memory5
 #| include: false
@@
 ```{r}
-#| label: free-up-memory5
+#| label: free-up-memory6
 #| include: false
@@
 ```{r}
-#| label: free-up-memory6
+#| label: free-up-memory7
 #| include: false

### 2025-06-18 simulation.qmd
20250617 - fix duplicate chunk label
20250617 - reduce iterations to fix memory issue

@@
   vars_by_pos = vars_by_pos,
-  n_iter = 10000,
+  n_iter = 5000,
   seed = 52242)
@@
   vars_by_pos = vars_by_pos,
-  n_iter = 10000,
+  n_iter = 5000,
   seed = 52242,
@@
 ```{r}
-#| label: put-simulated-data-in-format-for-scoring1
+#| label: put-simulated-data-in-format-for-scoring2
 

### 2025-06-18 simulation.qmd
20250618 - fix
20250618 - citations

@@
 all_proj <- dplyr::bind_rows(players_projectedPoints_seasonal)
-```
-
-```{r}
-#| label: code-to-specify-vars-to-bootstrap-by-position
-#| eval: false
-
-vars_by_pos <- list(
-  QB = c(
-    "games",
-    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
-    "rush_att", "rush_yds", "rush_tds",
-    "fumbles_lost", "fumbles_total", "two_pts",
-    "sacks",
-    "pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",
-    "pass_40_yds", "pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_400_yds",
-    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds"
-    ),
-  RB = c(
-    "games",
-    "rush_att", "rush_yds", "rush_tds",
-    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
-    "fumbles_lost", "fumbles_total", "two_pts",
-    "return_yds", "return_tds",
-    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
-    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
-    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
-  ),
-  WR = c(
-    "games",
-    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
-    "rush_att", "rush_yds", "rush_tds",
-    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
-    "fumbles_lost", "fumbles_total", "two_pts",
-    "return_yds", "return_tds",
-    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
-    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
-    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
-  ),
-  TE = c(
-    "games",
-    "pass_att", "pass_comp", "pass_inc", "pass_yds", "pass_tds", "pass_int",
-    "rush_att", "rush_yds", "rush_tds",
-    "rec_tgt", "rec", "rec_yds", "rec_tds", "rec_rz_tgt",
-    "fumbles_lost", "fumbles_total", "two_pts",
-    "return_yds", "return_tds",
-    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
-    "rush_40_yds", "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds",
-    "rec_40_yds", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds"
-  ),
-  K = c(
-    "fg_0019", "fg_2029", "fg_3039", "fg_4049", "fg_50", "fg_50_att",
-    "fg_39", "fg_att_39", "fg_49", "fg_49_att",
-    "fg", "fg_att", "fg_miss", "xp", "xp_att"
-  ),
-  D = c(
-    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  ),
-  DL = c(
-    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  ),
-  LB = c(
-    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  ),
-  DB = c(
-    "idp_solo", "idp_asst", "idp_sack", "idp_int", "idp_fum_force", "idp_fum_rec", "idp_pd", "idp_td", "idp_safety"
-  ),
-  DST = c(
-    "dst_fum_recvr", "dst_fum_rec", "dst_int", "dst_safety", "dst_sacks", "dst_td", "dst_blk",
-    "dst_fumbles", "dst_tackles", "dst_yds_against", "dst_pts_against", "dst_pts_allowed", "dst_ret_yds"
-  )
-)
 ```
@@
 ```
+
+For performing the bootstrapping, we leverage the `data.table` [@R-data.table], `future` [@R-future], and `future.apply` [@R-future.apply] packages for speed (by using parallel processing) and memory efficiency.
+We use the `progressr` [@R-progressr] package to create a progress bar.
 
@@
 
-rm(bootstrappedFantasyPoints)
 gc()
@@
 
-You can generate a template for Monte Carlo simulations in the `SimDesign` package using the following code:
+You can generate a template for Monte Carlo simulations in the `SimDesign` [@R-SimDesign; @Chalmers2020] package using the following code:
 

### 2025-06-18 simulation.qmd
20250618 - free up memory for Monte Carlo simulation
20250618 - fix

@@
 bootstrappedFantasyPoints_summary %>% 
-  filter(position == "QB") %>% 
+  dplyr::filter(position == "QB") %>% 
   mutate(
@@
 bootstrappedFantasyPoints_summary %>% 
-  filter(position == "RB") %>% 
+  dplyr::filter(position == "RB") %>% 
   mutate(
@@
 bootstrappedFantasyPoints_summary %>% 
-  filter(position == "WR") %>% 
+  dplyr::filter(position == "WR") %>% 
   mutate(
@@
 bootstrappedFantasyPoints_summary %>% 
-  filter(position == "TE") %>% 
+  dplyr::filter(position == "TE") %>% 
   mutate(
@@
 bootstrappedFantasyPoints_summary %>% 
-  filter(position == "PK") %>% 
+  dplyr::filter(position == "PK") %>% 
   mutate(
@@
 bootstrappedFantasyPoints_summary %>% 
-  filter(position %in% c("DL","DT","DE")) %>% 
+  dplyr::filter(position %in% c("DL","DT","DE")) %>% 
   mutate(
@@
 bootstrappedFantasyPoints_summary %>% 
-  filter(position %in% c("LB","MLB","OLB")) %>% 
+  dplyr::filter(position %in% c("LB","MLB","OLB")) %>% 
   mutate(
@@
 bootstrappedFantasyPoints_summary %>% 
-  filter(position %in% c("DB","S","CB")) %>% 
+  dplyr::filter(position %in% c("DB","S","CB")) %>% 
   mutate(
@@
 
-bootstrappedFantasyPoints_summary %>% filter(mean > 100 & position == "QB") %>% arrange(CV) %>% head
-bootstrappedFantasyPoints_summary %>% filter(mean > 100 & position == "QB") %>% arrange(-CV) %>% head
+bootstrappedFantasyPoints_summary %>% dplyr::filter(mean > 100 & position == "QB") %>% arrange(CV) %>% head
+bootstrappedFantasyPoints_summary %>% dplyr::filter(mean > 100 & position == "QB") %>% arrange(-CV) %>% head
 ```
@@
   data = bootstrappedFantasyPoints %>%
-    filter(pos == "QB" & name == "Patrick Mahomes"),
+    dplyr::filter(pos == "QB" & name == "Patrick Mahomes"),
   mapping = aes(
@@
   data = bootstrappedFantasyPoints %>%
-    filter(pos == "QB" & (name %in% c("Dak Prescott", "Drake Maye"))),
+    dplyr::filter(pos == "QB" & (name %in% c("Dak Prescott", "Drake Maye"))),
   mapping = aes(
@@
   theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
+```
+
+```{r}
+#| label: free-up-memory8
+#| include: false
+
+rm(bootstrappedFantasyPoints)
+gc()
 ```

### 2025-06-18 basic-statistics.qmd
20250618 - quantiles

@@
 - range
-- minimum and maximum
+- minimum and maximum (or some other quantiles)
 - interquartile range (IQR)
@@
 max(player_stats_seasonal$fantasyPoints, na.rm = TRUE)
+```
+
+- quantiles (in this case, the 10th and 90th quantiles):
+
+```{r}
+quantile(player_stats_seasonal$fantasyPoints, 0.10, na.rm = TRUE)
+quantile(player_stats_seasonal$fantasyPoints, 0.90, na.rm = TRUE)
 ```

### 2025-06-18 simulation.qmd
20250618 - fix
20250618 - quantiles

--- 2025-06-18 simulation.qmd
@@
 bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
+  left_join(
+    nfl_playerIDs[,c("mfl_id","name","merge_name","team")],
+    by = c("id" = "mfl_id")
+  )
+
+bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
   rename(projectedPoints = raw_points)
@@
     SD = sd(projectedPoints, na.rm = TRUE),
-    floor = quantile(projectedPoints, .10), # 10th quantile
     ceiling = quantile(projectedPoints, .90), # 90th quantile
@@
 
+```{r}
+#| label: free-up-memory8
+#| include: false
+
+rm(bootstrappedFantasyPoints)
+gc()
+```
+
 ### Monte Carlo Simulation {#sec-simulationProjectedPointsMonteCarlo}

### 2025-06-18 simulation.qmd
20250618 - headings for organization

@@
 ### Bootstrapping {#sec-simulationProjectedPointsBootstrapping}
+
+#### Prepare Data {#sec-simulationProjectedPointsBootstrappingPrepareData}
 
@@
 
+#### Bootstrapping Function {#sec-simulationProjectedPointsBootstrappingFunction}
+
 For performing the bootstrapping, we leverage the `data.table` [@R-data.table], `future` [@R-future], and `future.apply` [@R-future.apply] packages for speed (by using parallel processing) and memory efficiency.
@@
 
+#### Run the Bootsrapping Simulation {#sec-simulationProjectedPointsBootstrappingRun}
+
 ```{r}
@@
 ```
+
+#### Score Fantasy Points from the Simulation {#sec-simulationProjectedPointsBootstrappingScore}
 
@@
 
+#### Summarize Players' Distribution of Projected Fantasy Points {#sec-simulationProjectedPointsBootstrappingSummarize}
+
 ```{r}
@@
     SD = sd(projectedPoints, na.rm = TRUE),
-    ceiling = quantile(projectedPoints, .90), # 90th quantile
+    q10 = quantile(projectedPoints, .10), # 10th quantile
+    q90 = quantile(projectedPoints, .90), # 90th quantile
     min = min(projectedPoints, na.rm = TRUE),
@@
 ```
+
+#### View Players' Distribution of Projected Fantasy Points {#sec-simulationProjectedPointsBootstrappingView}
 
@@
 
+#### `SimDesign` Package {#sec-simulationProjectedPointsMonteCarloSimDesign}
+
 You can generate a template for Monte Carlo simulations in the `SimDesign` [@R-SimDesign; @Chalmers2020] package using the following code:

### 2025-06-18 index.qmd
20250618 - add package for fitting skew-normal distribution

--- 2025-06-10 index.qmd
@@
   "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
-  "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", 
+  "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "sn", "mclust", "magrittr", 
   "viridis", "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", #"webshot2", 

### 2025-06-18 simulation.qmd
20250618 - add package for fitting skew-normal distribution

@@
 library("SimDesign")
+library("fitdistrplus")
+library("sn")
 library("tidyverse")

### 2025-06-19 basic-statistics.qmd
20250618 - fix formulas

@@
   d &= \frac{\text{mean difference}}{\text{pooled standard deviation}} \\
-   &= \frac{\bar{X_1} - \bar{X_2}}{s} \\
+   &= \frac{\bar{x_1} - \bar{x_2}}{s} \\
 \end{aligned}
@@
 
-where:
+where $\bar{x_1}$ and $\bar{x_2}$ is the mean for group 1 and group 2, respectively, and:
 

### 2025-06-19 basic-statistics.qmd
20250618 - include quantiles

@@
         max = ~ max(., na.rm = TRUE),
+        q10 = quantile(., .10, na.rm = TRUE), # 10th quantile
+        q90 = quantile(., .90, na.rm = TRUE), # 90th quantile
         range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),
@@
 
-The formula for the correlation coefficient is in @sec-correlation.
+The formula for the correlation coefficient is in @eq-pearsonCorrelation.
 

### 2025-06-19 multiple-regression.qmd
20250617 - fix duplicate identifier
20250618 - include quantiles

@@
 
-### Load Data {#sec-correlationLoadData}
+### Load Data {#sec-multipleRegressionLoadData}
 

### 2025-06-19 simulation.qmd
20250618 - typo
20250618 - include quantiles

@@
 
-#### Run the Bootsrapping Simulation {#sec-simulationProjectedPointsBootstrappingRun}
+#### Run the Bootstrapping Simulation {#sec-simulationProjectedPointsBootstrappingRun}
 
@@
     SD = sd(projectedPoints, na.rm = TRUE),
-    q10 = quantile(projectedPoints, .10), # 10th quantile
-    q90 = quantile(projectedPoints, .90), # 90th quantile
     min = min(projectedPoints, na.rm = TRUE),
     max = max(projectedPoints, na.rm = TRUE),
+    q10 = quantile(projectedPoints, .10, na.rm = TRUE), # 10th quantile
+    q90 = quantile(projectedPoints, .90, na.rm = TRUE), # 90th quantile
     range = max(projectedPoints, na.rm = TRUE) - min(projectedPoints, na.rm = TRUE),

### 2025-06-19 basic-statistics.qmd
20250619 - quantiles

@@
 - minimum and maximum (or some other quantiles)
+- weighted quantiles
 - interquartile range (IQR)
@@
 
+If you want to give some values more weight to others, you can calculate weighted quantiles, while assigning a weight to each value.
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains various functions for computing weighted quantiles based on @Akinshin2023.
+Because some projections are outliers, we use a trimmed version of the weighted Harrell-Davis quantile estimator for greater robustness.
+
 Below is R code to estimate each:
@@
 ```{r}
-quantile(player_stats_seasonal$fantasyPoints, 0.10, na.rm = TRUE)
-quantile(player_stats_seasonal$fantasyPoints, 0.90, na.rm = TRUE)
+quantile(player_stats_seasonal$fantasyPoints, c(0.10, 0.90), na.rm = TRUE)
+```
+
+- weighted quantiles (in this case, the 10th and 90th quantiles):
+
+```{r}
+petersenlab::wthdquantile(
+  player_stats_seasonal$fantasyPoints,
+  w = sample( # randomly generate weights (could specify them manually)
+    x = 1:3,
+    size = length(player_stats_seasonal$fantasyPoints),
+    replace = TRUE),
+  probs = c(0.10, 0.90))
 ```
@@
         max = ~ max(., na.rm = TRUE),
-        q10 = quantile(., .10, na.rm = TRUE), # 10th quantile
-        q90 = quantile(., .90, na.rm = TRUE), # 90th quantile
+        q10 = ~ quantile(., .10, na.rm = TRUE), # 10th quantile
+        q90 = ~ quantile(., .90, na.rm = TRUE), # 90th quantile
         range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),

### 2025-06-19 correlation.qmd
20250619 - quantiles

--- 2025-06-17 correlation.qmd
@@
         max = ~ max(., na.rm = TRUE),
+        q10 = ~ quantile(., .10, na.rm = TRUE), # 10th quantile
+        q90 = ~ quantile(., .90, na.rm = TRUE), # 90th quantile
         range = ~ max(., na.rm = TRUE) - min(., na.rm = TRUE),

### 2025-06-19 simulation.qmd
20250619 - Monte Carlo simulation

@@
 bootstrappedFantasyPoints_summary %>% 
-  filter(position == "PK") %>% 
+  filter(position == c("K","PK")) %>% 
   mutate(
@@
 
+#### Prepare Data {#sec-simulationProjectedPointsMonteCarloPrepareData}
+
+```{r}
+#| include: false
+
+all_proj <- dplyr::bind_rows(players_projectedPoints_seasonal)
+```
+
+```{r}
+all_proj <- all_proj %>% 
+  rename(projectedPoints = raw_points)
+```
+
+#### Optimal Distribution for Each Player {#sec-simulationProjectedPointsMonteCarloOptimalDistribution}
+
+```{r}
+# Function to identify the "best" distribution (Normal vs Skew‑Normal) for every player (tries both families and picks by AIC; uses empirical distribution if fewer than 2 unique scores)
+fit_best <- function(x) {
+  # Basic checks
+  if (length(unique(x)) < 2 || all(is.na(x))) { # Use empirical distribution if there are fewer than 2 unique scores
+    return(list(type = "empirical", empirical = x))
+  }
+
+  # Try Normal Distribution
+  fit_norm <- tryCatch(
+    fitdistrplus::fitdist(x, distr = "norm"),
+    error = function(e) NULL
+  )
+
+  # Try Skew-Normal Distribution
+  fit_skew <- tryCatch(
+    sn::selm(x ~ 1),
+    error = function(e) NULL
+  )
+
+  # Handle bad fits: sd = NA, etc.
+  if (!is.null(fit_norm) && any(is.na(fit_norm$estimate))) {
+    fit_norm <- NULL
+  }
+
+  if (!is.null(fit_skew)) {
+    pars <- tryCatch(sn::coef(fit_skew, param.type = "dp"), error = function(e) NULL)
+    if (is.null(pars) || any(is.na(pars))) {
+      fit_skew <- NULL
+    }
+  }
+
+  # Choose best available
+  if (!is.null(fit_norm) && !is.null(fit_skew)) {
+    aic_norm <- AIC(fit_norm)
+    aic_skew <- AIC(fit_skew)
+    if (aic_skew + 2 < aic_norm) { # skew-normal is more complex (has more parameters) than normal distribution, so only select a skew-normal distribution if it fits substantially better than a normal distribution
+      pars <- sn::coef(fit_skew, param.type = "dp")
+      return(list(
+        type  = "skewnorm",
+        xi    = pars["dp.location"],
+        omega = pars["dp.scale"],
+        alpha = pars["dp.shape"]))
+    } else {
+      return(list(
+        type = "norm",
+        mean = fit_norm$estimate["mean"],
+        sd   = fit_norm$estimate["sd"]))
+    }
+  } else if (!is.null(fit_norm)) {
+    return(list(
+      type = "norm",
+      mean = fit_norm$estimate["mean"],
+      sd   = fit_norm$estimate["sd"]))
+  } else {
+    return(list(type = "empirical", empirical = x))
+  }
+}
+```
+
+```{r}
+proj_dists_tbl <- all_proj |>
+  filter(!is.na(id) & id != "") |>
+  group_by(id) |>
+  summarise(
+    dist_info = list(fit_best(projectedPoints)),
+    n_proj = n(), # record of how many sources they have
+    .groups = "drop"
+  )
+
+proj_dists <- proj_dists_tbl |>
+  filter(!is.na(id) & id != "") |>
+  distinct(id, .keep_all = TRUE) |>
+  (\(x) setNames(x$dist_info, x$id))()
+```
+
+```{r}
+proj_dists_tbl |>
+  dplyr::mutate(
+    dist_type = purrr::map_chr(dist_info, ~ .x$type)
+  ) |>
+  dplyr::count(dist_type)
+```
+
+#### `SimDesign` Step 1: Design Grid {#sec-simulationProjectedPointsMonteCarloDesign}
+
+Now we build the `SimDesign` design grid based on the number of projections that each player had.
+
+```{r}
+Design <- proj_dists_tbl |>
+  transmute(
+    id,
+    n_sources = n_proj # how many projections the player had
+  )
+```
+
+```{r}
+#| eval: false
+
+missing_ids <- setdiff(Design$id, names(proj_dists))
+length(missing_ids) # should be 0
+
+any(is.na(proj_dists_tbl$id)) # should be FALSE
+any(is.na(Design$id))         # should be FALSE
+any(is.na(names(proj_dists))) # should be FALSE
+```
+
+#### `SimDesign` Step 2: Generate {#sec-simulationProjectedPointsMonteCarloGenerate}
+
+```{r}
+Generate <- function(condition, fixed_objects = NULL) {
+
+  dist_info <- fixed_objects$proj_dists[[as.character(condition$id)]]
+  n_sources <- condition$n_sources
+
+  sim_points <- switch(
+    dist_info$type,
+    empirical = sample(
+      dist_info$empirical,
+      n_sources,
+      replace = TRUE),
+
+    norm = rnorm(
+      n_sources,
+      mean = dist_info$mean,
+      sd = dist_info$sd),
+
+    skewnorm = sn::rsn(
+      n_sources,
+      xi = dist_info$xi,
+      omega = dist_info$omega,
+      alpha = dist_info$alpha),
+
+    stop("Unknown distribution type: ", dist_info$type)
+  )
+
+  data.frame(
+    id = condition$id,
+    sim_points = sim_points)
+}
+```
+
+#### `SimDesign` Step 3: Analyze {#sec-simulationProjectedPointsMonteCarloAnalyze}
+
+```{r}
+Analyse <- function(condition, dat, fixed_objects = NULL) {
+  tibble::tibble(
+    id        = condition$id,
+    mean_pts  = mean(dat$sim_points, na.rm = TRUE),
+    sd_pts    = sd(dat$sim_points, na.rm = TRUE),
+    q10       = quantile(dat$sim_points, 0.10, na.rm = TRUE),
+    q90       = quantile(dat$sim_points, 0.90, na.rm = TRUE),
+    p100      = mean(dat$sim_points >= 100, na.rm = TRUE),
+    p150      = mean(dat$sim_points >= 150, na.rm = TRUE),
+    p200      = mean(dat$sim_points >= 200, na.rm = TRUE),
+    p250      = mean(dat$sim_points >= 250, na.rm = TRUE),
+    p300      = mean(dat$sim_points >= 300, na.rm = TRUE),
+    p350      = mean(dat$sim_points >= 300, na.rm = TRUE)
+  )
+}
+```
+
+#### `SimDesign` Step 4: Summarize {#sec-simulationProjectedPointsMonteCarloSummarize}
+
+```{r}
+Summarise <- function(condition, results, fixed_objects = NULL) {
+  dplyr::summarise(
+    results,
+    across(
+      where(is.numeric),
+      list(
+        mean = ~mean(.x, na.rm = TRUE),
+        sd   = ~sd(.x,  na.rm = TRUE)),
+      .names = "{.col}_{.fn}"
+    ),
+    .groups = "drop"
+  )
+}
+```
+
+#### `SimDesign` Step 5: Run the Simulation {#sec-simulationProjectedPointsMonteCarloRun}
+
+Now, we can run the model.
+
+::: {#nte-monteCarloSimulation .callout-note title="Monte Carlo Simulation"}
+Note: the following code that runs the simulation takes a while.
+If you just want to save time and load the results object instead of running the simulation, you can load the results object of the simulation (which has already been run) using this code:
+
+```{r}
+load(url("https://osf.io/download/ues7n/"))
+```
+:::
+
+```{r}
+#| eval: false
+
+monteCarloSim_results <- SimDesign::runSimulation(
+  design = Design,
+  replications = 1000,
+  generate = Generate,
+  analyse = Analyse,
+  summarise = Summarise,
+  fixed_objects = list(proj_dists = proj_dists),
+  seed = genSeeds(Design, iseed = 52242), # for reproducibility
+  parallel = TRUE # for faster (parallel) processing
+)
+```
+
+#### Simulation Results {#sec-simulationProjectedPointsMonteCarloResults}
+
+```{r}
+monteCarloSim_results <- monteCarloSim_results %>% 
+  left_join(
+    nfl_playerIDs[,c("mfl_id","name","merge_name","position","team")],
+    by = c("id" = "mfl_id")
+  ) %>% 
+  select(name, team, position, everything()) %>% 
+  arrange(-mean_pts_mean)
+```
+
 ```{r}
@@
 
-calculate_fantasyPoints <- function(df, scoring_rules){
-  #fantasyPoints <-
+save(
+  monteCarloSim_results,
+  file = "./data/monteCarloSim_results.RData")
+```
 
-  #return(fantasyPoints)
-}
+```{r}
+#| include: false
 
-monte_carlo_simulation <- function(position_df, num_simulations = 10000){
-  player_ids <- unique(position_df$id)
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/monteCarloSim_results.RData", fsep = ""))
+```
 
-  results <- lapply(player_ids, function(player_id){ # use curly brackets when there are multiple expressions
-    player_data <- position_df %>% filter(id == player_id)
+```{r}
+monteCarloSim_results
+```
 
-    # Run the simulation for this player
-    simulations <- replicate(
-      num_simulations, {
-        sampled_data <- player_data %>%
-          summarise(across(
-            c(pass_yds, pass_tds, rush_yds, rush_tds, rec_yds, rec_tds), # add more statistical categories
-            ~ if (all(is.na(.))) NA else sample(.[!is.na(.)], 1, replace = TRUE)))
+```{r}
+monteCarloSim_results %>% 
+  filter(position == "QB")
+```
 
-        sampled_data # return last expression
-        },
-      simplify = FALSE)
+```{r}
+monteCarloSim_results %>% 
+  filter(position == "RB")
+```
 
-    simulation_df <- dplyr::bind_rows(simulations)
-    #simulation_df$fantasyPoints <- calculate_fantasyPoints(simulation_df) # write function to calculate fantasy points
+```{r}
+monteCarloSim_results %>% 
+  filter(position == "WR")
+```
 
-    ci <- simulation_df %>%
-      summarise(across(where(is.numeric), list(
-        lower = \(x) quantile(x, 0.025, na.rm = TRUE),
-        upper = \(x) quantile(x, 0.975, na.rm = TRUE))))
+```{r}
+monteCarloSim_results %>% 
+  filter(position == "TE")
+```
 
-    ci <- ci %>% 
-      mutate(id = player_id)
+```{r}
+monteCarloSim_results %>% 
+  filter(position %in% c("K","PK"))
+```
 
-    return(ci)
-    })
+```{r}
+monteCarloSim_results %>% 
+  filter(position %in% c("DL","DT","DE"))
+```
 
-    # Combine results for all players
-    final_ci <- bind_rows(results)
+```{r}
+monteCarloSim_results %>% 
+  filter(position %in% c("LB","MLB","OLB"))
+```
 
-    return(final_ci)
-}
-
-monte_carlo_simulation(players_projectedPoints_seasonal$QB, num_simulations = 10)
-
-QB_ci <- monte_carlo_simulation(players_projectedPoints_seasonal$QB)
+```{r}
+monteCarloSim_results %>% 
+  filter(position %in% c("DB","S","CB"))
 ```

### 2025-06-19 simulation.qmd
20250619 - fix

@@
     p300      = mean(dat$sim_points >= 300, na.rm = TRUE),
-    p350      = mean(dat$sim_points >= 300, na.rm = TRUE)
+    p350      = mean(dat$sim_points >= 350, na.rm = TRUE)
   )
@@
 ```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/monteCarloSim_results.RData")
+```
+
+```{r}
+#| eval: false
 #| include: false
@@
 ```
+
+The `pX` variable represent the probability that a player scoring more than X number of points.
+For example, the `p300` variable represents the probability that each player scores more than 300 points.
+However, it is important to note that this is based on the distribution of *projected* points.
 

### 2025-06-19 actuarial.qmd
20241016 - render the figure caption as alt text
20250308 - citations
20250619 - calibration accuracy of projections

@@
 
-Here is a video of sports pundits, Stephen A. Smith and Skip Bayless, making bold statements and incorrect predictions:
+@TPS2017 provides a video of sports pundits, Stephen A. Smith and Skip Bayless, making bold statements and incorrect predictions:
 
@@
 Smith and Skip Bayless, Making Bold Statements and Incorrect Predictions.
-From: <https://www.youtube.com/watch?v=lTjBuEPcLlc>.](images/sportsPundits.mp4){width=100%}
+From: <https://www.youtube.com/watch?v=lTjBuEPcLlc>.](images/sportsPundits.mp4){width=100%}\
 
@@
 Another reason why judgment methods are more widely used than actuarial methods is that so-called "experts" (and people in general) show [overconfidence](#sec-cognitiveBiasesOverconfidence) in their predictions—clinicians, experts, and humans in general think they are more accurate than they actually are.
-We see this when examining their calibration; their predictions tend to be miscalibrated.
+We see this when examining their calibration; their predictions tend to be miscalibrated (this is also the case with fantasy football projections, as described in Section @sec-accuracyProjections).
 For example, things they report with 80% confidence occur less than 80% of the time, an example of [overprecision](#sec-cognitiveBiasesOverconfidence) in their predictions.

### 2025-06-19 decision-making.qmd
20250619 - calibration accuracy of projections

@@
 Second, there are many players who are systematically *over*predicted (e.g., due to their injury) or *under*predicted (e.g., due to their becoming the starter after a teammate becomes injured, is traded, etc.)—an example of [overextremity miscalibration](#sec-calibration).
-In general, it is likely for players who are projected to score more points to be overpredicted and for players who are projected to score fewer points to be underpredicted [PRESENT CALIBRATION STATS].
+In general, it is likely for players who are projected to score more points to be overpredicted and for players who are projected to score fewer points to be underpredicted, as described in Section @sec-accuracyProjections.
 Third, the experts may interact and discuss with one another.

### 2025-06-19 evaluating-prediction-accuracy.qmd
20250619 - ROC curve
20250619 - calibration accuracy of projections

@@
 
+### Specify Options {#predictionAccuracySpecifyOptions}
+
+```{r}
+options(scipen = 999) # prevent scientific notation
+```
+
 ### Prepare Data {#sec-predictionAccuracyPrepareData}
@@
 There are four general patterns of miscalibration: overextremity, underextremity, overprediction, and underprediction (see @fig-miscalibration).
-*Overextremity* exists when the predicted probabilites are too close to the extremes (zero or one).
+*Overextremity* exists when the predicted probabilities are too close to the extremes (zero or one).
 *Underextremity* exists when the predicted probabilities are too far away from the extremes.
@@
 
+par(pty = "s")
+
 plot(
@@
 ## Accuracy of Fantasy Projections {#sec-accuracyProjections}
+
+Below are estimates of projection accuracy and [calibration plots](#sec-calibrationPlot) for [seasonal](#sec-accuracyProjectionsSeasonal) and [weekly](#sec-accuracyProjectionsWeekly) projections.
+Projections were poorly [calibrated](#sec-calibrationIndices) with respect to actual fantasy points.
+In general, projected showed a modest tendency to underpredict performance at lower levels of projected points and a moderate-to-strong tendency overpredict performance at higher levels of projected points, an example of the overextremity form of [miscalibration](#sec-calibration).
+As a reminder, the overextremity form of [miscalibration](#sec-calibration) is an example of the overprecision form of [overconfidence](#sec-cognitiveBiasesOverconfidence).
+That is, projection sources appear to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their projections of player performance.
 

### 2025-06-20 evaluating-prediction-accuracy.qmd
20250619 - unamalgating multi-item scales
20250619 - ROC curve

@@
 #| fig-width: 8
-#| fig-height: 6
+#| fig-height: 8
 
@@
 The "wisdom of the crowd" is often more accurate than individuals' predictions, including predictions by so-called "experts" [@Silver2012].
+- Instead of aggregating scores from multiple variables, another option is to "unamalgamate" the predictors, that is, to use each variable as a separate predictor, to maximize predictive power [@TrafimowInPress].
 - A goal of prediction is to capture as much signal as possible and as little noise (error) as possible [@Silver2012].

### 2025-06-20 simulation.qmd
20250619 - update monte carlo simulation

@@
 ```{r}
+#| eval: false
+
 monteCarloSim_results <- monteCarloSim_results %>% 

### 2025-06-20 index.qmd
20250619 - efficient frontier

--- 2025-06-10 index.qmd
@@
   "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
-  "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", 
-  "viridis", "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", #"webshot2", 
-  "stringi", "parallel", "parallelly", "doParallel", "missRanger", "ggridges", 
-  "powerjoin", "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", 
-  "correlation", "mgcv", "rms", "car", "performance", "lavaan", "mice", 
-  "miceadds", "interactions", "shiny", "PearsonDS", "robustbase", "data.table", 
-  "future", "future.apply"
+  "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "sn", "mclust", 
+  "magrittr", "viridis", "plotly", "quantmod", "fPortfolio", "NMOF", "xts", #"webshot2",
+  "zoo", "forecast", "stringi", "parallel", "parallelly", "doParallel", 
+  "missRanger", "ggridges", "powerjoin", "caret", "gpboost", "cmdstanr", 
+  "learnr", "remotes", "corrplot", "correlation", "mgcv", "rms", "car", 
+  "performance", "lavaan", "mice", "miceadds", "interactions", "shiny", 
+  "PearsonDS", "robustbase", "data.table", "future", "future.apply", "SimDesign"
 )

### 2025-06-20 modern-portfolio-theory.qmd
20250619 - efficient frontier

@@
 library("fPortfolio")
+library("NMOF")
 library("tidyverse")
+```
+
+### Load Data {#sec-modernPortfolioTheoryLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projectedPoints_seasonal.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_playerIDs.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/players_projectedPoints_seasonal.RData")
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/nfl_playerIDs.RData")
 ```
@@
 
-In the example below, we use historical returns (since YEAR) as the expected future returns.
+In the example below, we use historical returns (since 2012) as the expected future returns.
 However, using historical returns as the expected future returns is risky because, as described in the common disclaimer, "Past performance does not guarantee future results."
@@
 returns_ts <- timeSeries::as.timeSeries(returns)
+```
+
+```{r}
+#| eval: false
+
+minYear <- format(as.Date(row.names(returns_ts)), "%Y") %>% 
+  na.omit %>% 
+  as.integer %>% 
+  min()
 ```
@@
 
+```{r}
+all_proj <- dplyr::bind_rows(players_projectedPoints_seasonal)
+
+all_proj <- all_proj %>% 
+  rename(projectedPoints = raw_points)
+```
+
+```{r}
+all_proj_summary <- all_proj %>% 
+  group_by(id) %>% 
+  summarise(
+    mean = mean(projectedPoints, na.rm = TRUE),
+    sd = sd(projectedPoints, na.rm = TRUE),
+    var = var(projectedPoints, na.rm = TRUE)
+  )
+
+all_proj_summary <- all_proj_summary %>% 
+  left_join(
+    nfl_playerIDs[,c("mfl_id","name","merge_name","position","team")],
+    by = c("id" = "mfl_id")
+  ) %>% 
+  select(name, team, position, everything()) %>% 
+  arrange(-mean)
+```
+
+::: {.content-visible when-format="html:js"}
+
+```{r}
+#| label: fig-projectedPointsMeanVsSD
+#| fig-cap: "Mean of Projected Fantasy Points For Each Player as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."
+#| fig-alt: "Mean of Projected Fantasy Points For Each Player as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."
+
+plot_projectedPointsMeanVsSD <- ggplot2::ggplot(
+  data = all_proj_summary,
+  aes(
+    x = sd,
+    y = mean)) +
+  geom_point(
+    aes(
+      text = name, # add player name for mouse over tooltip
+      label = position)) + # add season for mouse over tooltip
+  #geom_smooth() +
+  coord_cartesian(
+    xlim = c(0,NA),
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Standard Deviation of Player's Projected Fantasy Points (Season) Across Sources",
+    y = "Mean of Player's Projected Fantasy Points (Season) Across Sources",
+    title = "Mean vs Standard Deviation of Players' Projected Fantasy Points"
+  ) +
+  theme_classic()
+
+plotly::ggplotly(plot_projectedPointsMeanVsSD)
+```
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
+```{r}
+#| label: fig-projectedPointsMeanVsSD
+#| fig-cap: "Mean of Projected Fantasy Points For Each Player as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."
+#| fig-alt: "Mean of Projected Fantasy Points For Each Player as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."
+
+ggplot2::ggplot(
+  data = all_proj_summary,
+  aes(
+    x = sd,
+    y = mean)) +
+  geom_point(
+    aes(
+      text = name, # add player name for mouse over tooltip
+      label = position)) + # add season for mouse over tooltip
+  #geom_smooth() +
+  coord_cartesian(
+    xlim = c(0,NA),
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Standard Deviation of Player's Projected Fantasy Points (Season) Across Sources",
+    y = "Mean of Player's Projected Fantasy Points (Season) Across Sources",
+    title = "Mean vs Standard Deviation of Players' Projected Fantasy Points"
+  ) +
+  theme_classic()
+```
+
+:::
+
+```{r}
+all_proj_summary_noNA <- all_proj_summary %>% 
+  filter(!is.na(id) & !is.na(mean) & !is.na(var))
+
+all_proj_summary_noNA_removeZeroVar <- all_proj_summary_noNA %>% 
+  filter(var > 0)
+```
+
+```{r}
+const_cor <- function(rho, numAssets) {
+  C <- array(rho, dim = c(numAssets, numAssets))
+  diag(C) <- 1
+  C
+}
+
+varCovMatrix <- diag(all_proj_summary_noNA_removeZeroVar$sd) %*% const_cor(0, nrow(all_proj_summary_noNA_removeZeroVar)) %*% diag(all_proj_summary_noNA_removeZeroVar$sd) # create a variance-covariance matrix assuming assets/players are uncorrelated
+
+efficientFrontierData <- NMOF::mvFrontier(
+  all_proj_summary_noNA_removeZeroVar$mean,
+  varCovMatrix,
+  wmin = 0,
+  wmax = 1,
+  n = 50)
+```
+
+```{r}
+#| label: fig-efficientFrontierAcrossProjections
+#| fig-cap: "Efficient Frontier of Projected Fantasy Points Across Sources."
+#| fig-alt: "Efficient Frontier of Projected Fantasy Points Across Sources."
+
+plot(
+  efficientFrontierData$volatility,
+  efficientFrontierData$return,
+  pch = 19,
+  cex = 0.5,
+  type = "o",
+  xlab = "Variance in Projected Fantasy Points Across Sources",
+  ylab = "Projected Fantasy Points")
+```
+
 ### Based on Historical Game-to-Game Variability {#sec-efficientFrontierFantasyProjectionsGameToGameVariability}
@@
 <https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league> [@Hitchings2012; archived at <https://perma.cc/JQ6G-KSRT>]
+
+```{r}
+player_stats_weekly_recent <- player_stats_weekly %>% 
+  filter(season == max(season))
+
+player_stats_weekly_recent_summary <- player_stats_weekly_recent %>% 
+  group_by(player_id) %>% 
+  summarise(
+    mean = mean(fantasyPoints, na.rm = TRUE),
+    sd = sd(fantasyPoints, na.rm = TRUE),
+    var = var(fantasyPoints, na.rm = TRUE)
+  )
+
+player_stats_weekly_recent_summary <- player_stats_weekly_recent_summary %>% 
+  left_join(
+    nfl_playerIDs[,c("gsis_id","name","merge_name","position","team")],
+    by = c("player_id" = "gsis_id")
+  ) %>% 
+  select(name, team, position, everything()) %>% 
+  arrange(-mean)
+```
+
+::: {.content-visible when-format="html:js"}
+
+```{r}
+#| label: fig-fantasyPointsMeanVsSD
+#| fig-cap: "Mean of Weekly Fantasy Points For Each Player as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."
+#| fig-alt: "Mean of Weekly Fantasy Points For Each Player as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."
+
+plot_fantasyPointsMeanVsSD <- ggplot2::ggplot(
+  data = player_stats_weekly_recent_summary,
+  aes(
+    x = sd,
+    y = mean)) +
+  geom_point(
+    aes(
+      text = name, # add player name for mouse over tooltip
+      label = position)) + # add season for mouse over tooltip
+  #geom_smooth() +
+  coord_cartesian(
+    xlim = c(0,NA),
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Standard Deviation of Player's Fantasy Points From Week-to-Week",
+    y = "Mean of Player's Fantasy Points Across Weeks",
+    title = "Mean vs Standard Deviation of Players' Fantasy Points"
+  ) +
+  theme_classic()
+
+plotly::ggplotly(plot_fantasyPointsMeanVsSD)
+```
+
+:::
+
+::: {.content-visible when-format="pdf"}
+
+```{r}
+#| label: fig-fantasyPointsMeanVsSD
+#| fig-cap: "Mean of Weekly Fantasy Points For Each Player as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."
+#| fig-alt: "Mean of Weekly Fantasy Points For Each Player as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."
+
+ggplot2::ggplot(
+  data = player_stats_weekly_recent_summary,
+  aes(
+    x = sd,
+    y = mean)) +
+  geom_point(
+    aes(
+      text = name, # add player name for mouse over tooltip
+      label = position)) + # add season for mouse over tooltip
+  #geom_smooth() +
+  coord_cartesian(
+    xlim = c(0,NA),
+    ylim = c(0,NA),
+    expand = FALSE) +
+  labs(
+    x = "Standard Deviation of Player's Fantasy Points From Week-to-Week",
+    y = "Mean of Player's Fantasy Points Across Weeks",
+    title = "Mean vs Standard Deviation of Players' Fantasy Points"
+  ) +
+  theme_classic()
+```
+
+:::
+
+```{r}
+player_stats_weekly_recent_summary_noNA <- player_stats_weekly_recent_summary %>% 
+  filter(!is.na(player_id) & !is.na(mean) & !is.na(var))
+
+player_stats_weekly_recent_summary_noNA_removeZeroVar <- player_stats_weekly_recent_summary_noNA %>% 
+  filter(var > 0)
+```
+
+```{r}
+varCovMatrix2 <- diag(player_stats_weekly_recent_summary_noNA_removeZeroVar$sd) %*% const_cor(0, nrow(player_stats_weekly_recent_summary_noNA_removeZeroVar)) %*% diag(player_stats_weekly_recent_summary_noNA_removeZeroVar$sd) # create a variance-covariance matrix assuming assets/players are uncorrelated
+
+efficientFrontierData2 <- NMOF::mvFrontier(
+  player_stats_weekly_recent_summary_noNA_removeZeroVar$mean,
+  varCovMatrix2,
+  wmin = 0,
+  wmax = 1,
+  n = 50)
+```
+
+```{r}
+#| label: fig-efficientFrontierAcrossWeeks
+#| fig-cap: "Efficient Frontier of Fantasy Points Across Weeks."
+#| fig-alt: "Efficient Frontier of Fantasy Points Across Weeks."
+
+plot(
+  efficientFrontierData2$volatility,
+  efficientFrontierData2$return,
+  pch = 19,
+  cex = 0.5,
+  type = "o",
+  xlab = "Variance in Fantasy Points Across Weeks",
+  ylab = "Weekly Fantasy Points")
+```
 

### 2025-06-20 multiple-regression.qmd
20250619 - efficient frontier

--- 2025-04-25 multiple-regression.qmd
@@
 QQ plots are particularly useful for identifying deviations from normality in the tails of the distribution; PP plots are particularly useful for identifying deviations from normality in the center of the distribution.
-Researchers tend to be more concerned about the tails of the distribution, because extreme values tend to have a greater impact on inferences.
+Researchers tend to be more concerned about the tails of the distribution, because extreme values tend to have a greater impact on inferences, so researchers tend to use QQ plots more often than PP plots.
 Various examples of QQ plots and deviations from normality are depicted in @fig-distributionQQplots.

### 2025-06-20 modern-portfolio-theory.qmd
20250620 - fix duplicate labels

@@
 ```{r}
-#| label: fig-projectedPointsMeanVsSD
+#| label: fig-projectedPointsMeanVsSDPDF
 #| fig-cap: "Mean of Projected Fantasy Points For Each Player as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."
@@
 ```{r}
-#| label: fig-fantasyPointsMeanVsSD
+#| label: fig-fantasyPointsMeanVsSDPDF
 #| fig-cap: "Mean of Weekly Fantasy Points For Each Player as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."

### 2025-06-20 modern-portfolio-theory.qmd
20250620 - sum of players' variances

@@
   type = "o",
-  xlab = "Variance in Projected Fantasy Points Across Sources",
+  xlab = "Standard Deviation of Projected Fantasy Points Across Sources",
   ylab = "Projected Fantasy Points")
 ```
+
+The formula for the variance (and standard deviation) of the sum of two variables is in @eq-varianceSumTwoVars:
+
+$$
+\begin{aligned}
+  \text{Var}(x + y) &= \text{Var}(x) + \text{Var}(y) + 2\text{Cov}(x, y)\\
+  s^2_{(x + y)} &= s^2_x + s^2_y +  2\text{Cov}(x, y)\\
+  \text{Std Dev}(x + y) &= \sqrt{\text{Var}(x) + \text{Var}(y) + 2\text{Cov}(x, y)}\\
+  s_{(x + y)} &= \sqrt{s^2_x + s^2_y + 2\text{Cov}(x, y)}\\
+\end{aligned}
+$$ {#eq-varianceSumTwoVars}
+
+If we assume that players' performance is uncorrelated with one another, the covariance between the players is zero, so the formula simplifies to @eq-varianceSumTwoVarsNoCov:
+
+$$
+\begin{aligned}
+  \text{Var}(x + y) &= \text{Var}(x) + \text{Var}(y)\\
+  s^2_{(x + y)} &= s^2_x + s^2_y\\
+  \text{Std Dev}(x + y) &= \sqrt{\text{Var}(x) + \text{Var}(y)}\\
+  s_{(x + y)} &= \sqrt{s^2_x + s^2_y}\\
+\end{aligned}
+$$ {#eq-varianceSumTwoVarsNoCov}
+
+That is, if players' performance is independent of each other, the variance of two players' points is merely the sum of their variances; their standard deviation is then the square root of that.
+In reality, players' performance is not truly independent—particularly for players on the same team or, for a given game, for players on opposing teams.
+However, for players not playing in the same game, it is a reasonable assumption to make and it greatly simplifies the math.
+If you wanted to, you could account for players' covariance by using @eq-varianceSumTwoVars.
+An example variance-covariance matrix of players' performance is in @sec-efficientFrontierFantasyProjectionsGameToGameVariability.
 
@@
   arrange(-mean)
+```
+
+Here is the variance-covariance matrix of players' performance (it is very large because there are many players):
+
+```{r}
+dataForVarCovMatrix <- player_stats_weekly_recent %>% 
+  select(player_id, week, fantasyPoints) %>% 
+  pivot_wider(
+    names_from = player_id,
+    values_from = fantasyPoints
+  )
+
+cov(
+  dataForVarCovMatrix,
+  use = "pairwise.complete.obs")
 ```
@@
   type = "o",
-  xlab = "Variance in Fantasy Points Across Weeks",
+  xlab = "Standard Deviation of Fantasy Points Across Weeks",
   ylab = "Weekly Fantasy Points")

### 2025-06-20 modern-portfolio-theory.qmd
20250620 - remove covariance matrix b/c too big

@@
 
-Here is the variance-covariance matrix of players' performance (it is very large because there are many players):
+Here is the code to generate the variance-covariance matrix of players' performance (it is very large because there are many players):
 
@@
   )
+```
+
+```{r}
+#| eval: false
 

### 2025-06-21 modern-portfolio-theory.qmd
20250621 - efficient frontier

@@
 
+As shown below, the correlation between mean and standard deviation of projected points is positive.
+That is, greater expected returns (i.e., mean of projected fantasy points) are associated with greater risk or volatility (i.e., standard deviation of projected fantasy points).
+Thus, to obtain a lineup that scores greater fantasy points, it may be necessary to take on greater risk.
+Moreover, you would not want take take on more risk for the same number of expected fantasy points; in addition, you would not want to score fewer fantasy points for the same amount of risk [@Hitchings2012].
+
+```{r}
+cor.test(
+  ~ mean + sd,
+  data = all_proj_summary
+)
+```
+
 ::: {.content-visible when-format="html:js"}
@@
 ```{r}
+#| eval: false
+#| include: false
+
+all_proj_summary %>% 
+  filter(position == "QB" & mean > 300) %>% 
+  arrange(-sd)
+
+all_proj_summary %>% 
+  filter(position == "QB" & mean > 150) %>% 
+  arrange(sd)
+
+all_proj_summary %>% 
+  filter(position == "RB" & mean > 200) %>% 
+  arrange(-sd)
+
+all_proj_summary %>% 
+  filter(position == "RB" & mean > 100) %>% 
+  arrange(sd)
+
+all_proj_summary %>% 
+  filter(position == "WR" & mean > 200) %>% 
+  arrange(-sd)
+
+all_proj_summary %>% 
+  filter(position == "WR" & mean > 100) %>% 
+  arrange(sd)
+```
+
+```{r}
+hiLo_permutations <- expand.grid(
+  rep(list(c("Hi", "Lo")), 3),
+  stringsAsFactors = FALSE
+)
+
+qbRbWr_projections <- data.frame(
+  condition = c("Hi", "Lo"),
+  qb_name = c("Josh Allen", "Matthew Stafford"),
+  rb_name = c("Saquon Barkley", "Joe Mixon"),
+  wr_name = c("Ja'Marr Chase", "Courtland Sutton")
+)
+
+qbRbWr_projections$qb_mean <- c(
+  all_proj_summary$mean[which(all_proj_summary$position == "QB" & all_proj_summary$name == qbRbWr_projections$qb_name[1])],
+  all_proj_summary$mean[which(all_proj_summary$position == "QB" & all_proj_summary$name == qbRbWr_projections$qb_name[2])])
+
+qbRbWr_projections$qb_sd <- c(
+  all_proj_summary$sd[which(all_proj_summary$position == "QB" & all_proj_summary$name == qbRbWr_projections$qb_name[1])],
+  all_proj_summary$sd[which(all_proj_summary$position == "QB" & all_proj_summary$name == qbRbWr_projections$qb_name[2])])
+
+qbRbWr_projections$rb_mean <- c(
+  all_proj_summary$mean[which(all_proj_summary$position == "RB" & all_proj_summary$name == qbRbWr_projections$rb_name[1])],
+  all_proj_summary$mean[which(all_proj_summary$position == "RB" & all_proj_summary$name == qbRbWr_projections$rb_name[2])])
+
+qbRbWr_projections$rb_sd <- c(
+  all_proj_summary$sd[which(all_proj_summary$position == "RB" & all_proj_summary$name == qbRbWr_projections$rb_name[1])],
+  all_proj_summary$sd[which(all_proj_summary$position == "RB" & all_proj_summary$name == qbRbWr_projections$rb_name[2])])
+
+qbRbWr_projections$wr_mean <- c(
+  all_proj_summary$mean[which(all_proj_summary$position == "WR" & all_proj_summary$name == qbRbWr_projections$wr_name[1])],
+  all_proj_summary$mean[which(all_proj_summary$position == "WR" & all_proj_summary$name == qbRbWr_projections$wr_name[2])])
+
+qbRbWr_projections$wr_sd <- c(
+  all_proj_summary$sd[which(all_proj_summary$position == "WR" & all_proj_summary$name == qbRbWr_projections$wr_name[1])],
+  all_proj_summary$sd[which(all_proj_summary$position == "WR" & all_proj_summary$name == qbRbWr_projections$wr_name[2])])
+
+qbRbWr_projections$qb_lastName <- sapply(strsplit(qbRbWr_projections$qb_name, " "), function(x) tail(x, 1))
+qbRbWr_projections$rb_lastName <- sapply(strsplit(qbRbWr_projections$rb_name, " "), function(x) tail(x, 1))
+qbRbWr_projections$wr_lastName <- sapply(strsplit(qbRbWr_projections$wr_name, " "), function(x) tail(x, 1))
+
+team_projections <- data.frame(
+  team = apply(hiLo_permutations, 1, paste0, collapse = ""),
+  team_name = NA,
+  team_mean = NA,
+  team_sd = NA,
+  qb = hiLo_permutations$Var1,
+  rb = hiLo_permutations$Var2,
+  wr = hiLo_permutations$Var3,
+  qb_name = NA,
+  qb_lastName = NA,
+  qb_mean = NA,
+  qb_sd = NA,
+  rb_name = NA,
+  rb_lastName = NA,
+  rb_mean = NA,
+  rb_sd = NA,
+  wr_name = NA,
+  wr_lastName = NA,
+  wr_mean = NA,
+  wr_sd = NA
+)
+
+team_projections$qb_name[which(team_projections$qb == "Hi")] <- qbRbWr_projections$qb_name[which(qbRbWr_projections$condition == "Hi")]
+team_projections$qb_name[which(team_projections$qb == "Lo")] <- qbRbWr_projections$qb_name[which(qbRbWr_projections$condition == "Lo")]
+team_projections$qb_lastName[which(team_projections$qb == "Hi")] <- qbRbWr_projections$qb_lastName[which(qbRbWr_projections$condition == "Hi")]
+team_projections$qb_lastName[which(team_projections$qb == "Lo")] <- qbRbWr_projections$qb_lastName[which(qbRbWr_projections$condition == "Lo")]
+team_projections$qb_mean[which(team_projections$qb == "Hi")] <- qbRbWr_projections$qb_mean[which(qbRbWr_projections$condition == "Hi")]
+team_projections$qb_mean[which(team_projections$qb == "Lo")] <- qbRbWr_projections$qb_mean[which(qbRbWr_projections$condition == "Lo")]
+team_projections$qb_sd[which(team_projections$qb == "Hi")] <- qbRbWr_projections$qb_sd[which(qbRbWr_projections$condition == "Hi")]
+team_projections$qb_sd[which(team_projections$qb == "Lo")] <- qbRbWr_projections$qb_sd[which(qbRbWr_projections$condition == "Lo")]
+
+team_projections$rb_name[which(team_projections$rb == "Hi")] <- qbRbWr_projections$rb_name[which(qbRbWr_projections$condition == "Hi")]
+team_projections$rb_name[which(team_projections$rb == "Lo")] <- qbRbWr_projections$rb_name[which(qbRbWr_projections$condition == "Lo")]
+team_projections$rb_lastName[which(team_projections$rb == "Hi")] <- qbRbWr_projections$rb_lastName[which(qbRbWr_projections$condition == "Hi")]
+team_projections$rb_lastName[which(team_projections$rb == "Lo")] <- qbRbWr_projections$rb_lastName[which(qbRbWr_projections$condition == "Lo")]
+team_projections$rb_mean[which(team_projections$rb == "Hi")] <- qbRbWr_projections$rb_mean[which(qbRbWr_projections$condition == "Hi")]
+team_projections$rb_mean[which(team_projections$rb == "Lo")] <- qbRbWr_projections$rb_mean[which(qbRbWr_projections$condition == "Lo")]
+team_projections$rb_sd[which(team_projections$rb == "Hi")] <- qbRbWr_projections$rb_sd[which(qbRbWr_projections$condition == "Hi")]
+team_projections$rb_sd[which(team_projections$rb == "Lo")] <- qbRbWr_projections$rb_sd[which(qbRbWr_projections$condition == "Lo")]
+
+team_projections$wr_name[which(team_projections$wr == "Hi")] <- qbRbWr_projections$wr_name[which(qbRbWr_projections$condition == "Hi")]
+team_projections$wr_name[which(team_projections$wr == "Lo")] <- qbRbWr_projections$wr_name[which(qbRbWr_projections$condition == "Lo")]
+team_projections$wr_lastName[which(team_projections$wr == "Hi")] <- qbRbWr_projections$wr_lastName[which(qbRbWr_projections$condition == "Hi")]
+team_projections$wr_lastName[which(team_projections$wr == "Lo")] <- qbRbWr_projections$wr_lastName[which(qbRbWr_projections$condition == "Lo")]
+team_projections$wr_mean[which(team_projections$wr == "Hi")] <- qbRbWr_projections$wr_mean[which(qbRbWr_projections$condition == "Hi")]
+team_projections$wr_mean[which(team_projections$wr == "Lo")] <- qbRbWr_projections$wr_mean[which(qbRbWr_projections$condition == "Lo")]
+team_projections$wr_sd[which(team_projections$wr == "Hi")] <- qbRbWr_projections$wr_sd[which(qbRbWr_projections$condition == "Hi")]
+team_projections$wr_sd[which(team_projections$wr == "Lo")] <- qbRbWr_projections$wr_sd[which(qbRbWr_projections$condition == "Lo")]
+
+team_projections$team_name <- apply(
+  team_projections[, c("qb_lastName", "rb_lastName", "wr_lastName")],
+  1,
+  paste,
+  collapse = "/"
+)
+
+team_projections$team_mean <- rowSums(team_projections[,c("qb_mean","rb_mean","wr_mean")])
+team_projections$team_sd <- rowSums(team_projections[,c("qb_sd","rb_sd","wr_sd")])
+```
+
+In @fig-teamPortfolio1, we depict the combined mean of projected fantasy points for a Quarterback/Running Back/Wide Receiver combination as a function of the standard deviation of their projected fantasy points across sources.
+
+```{r}
+#| label: fig-teamPortfolio1
+#| fig-cap: "Mean of Projected Fantasy Points For a Quarterback/Running Back/Wide Receiver Combination as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."
+#| fig-alt: "Mean of Projected Fantasy Points For a Quarterback/Running Back/Wide Receiver Combination as a Function of the Standard Deviation of the Projected Fantasy Points Across Sources."
+
+ggplot2::ggplot(
+  data = team_projections,
+  aes(
+    x = team_sd,
+    y = team_mean,
+    label = team_name)) +
+  geom_point() +
+  ggrepel::geom_text_repel(aes(label = team_name)) +
+  #geom_text(
+  #  hjust = 0,
+  #  vjust = 0) +
+  #geom_smooth() +
+  coord_cartesian(
+    xlim = c(NA,NA),
+    ylim = c(NA,NA),
+    expand = TRUE) +
+  labs(
+    x = "Standard Deviation of Players' Projected Fantasy Points (Season) Across Sources",
+    y = "Mean of Players' Projected Fantasy Points (Season) Across Sources",
+    title = "Mean vs Standard Deviation of Players' Projected Fantasy Points"
+  ) +
+  theme_classic()
+```
+
+```{r}
 all_proj_summary_noNA <- all_proj_summary %>% 
@@
 
+When you evaluate all possible combinations of players, you can obtain the expected returns and standard deviations for each player combination.
+Based on that, you can determine the maximum number of fantasy points (i.e., the best possible team) at each level of risk.
+The best expected returns at each level of risk is known as the efficient frontier.
+The efficient frontier of projected fantasy points across sources is in @fig-efficientFrontierAcrossProjections.
+
 ```{r}
@@
   xlab = "Standard Deviation of Projected Fantasy Points Across Sources",
-  ylab = "Projected Fantasy Points")
+  ylab = "Projected Fantasy Points",
+  main = "Efficient Frontier of Projected Fantasy Points Across Sources")
 ```
@@
   arrange(-mean)
+```
+
+```{r}
+cor.test(
+  ~ mean + sd,
+  data = player_stats_weekly_recent_summary
+)
 ```
@@
 ```{r}
+#| eval: false
+#| include: false
+
+player_stats_weekly_recent_summary %>% 
+  filter(position == "QB" & mean > 20) %>% 
+  arrange(-sd)
+
+player_stats_weekly_recent_summary %>% 
+  filter(position == "QB" & mean > 10) %>% 
+  arrange(sd)
+
+player_stats_weekly_recent_summary %>% 
+  filter(position == "RB" & mean > 17) %>% 
+  arrange(-sd)
+
+player_stats_weekly_recent_summary %>% 
+  filter(position == "RB" & mean > 10) %>% 
+  arrange(sd)
+
+player_stats_weekly_recent_summary %>% 
+  filter(position == "WR" & mean > 17) %>% 
+  arrange(-sd)
+
+player_stats_weekly_recent_summary %>% 
+  filter(position == "WR" & mean > 10) %>% 
+  arrange(sd)
+```
+
+```{r}
+qbRbWr_projections2 <- data.frame(
+  condition = c("Hi", "Lo"),
+  qb_name = c("Josh Allen", "Patrick Mahomes"),
+  rb_name = c("Saquon Barkley", "David Montgomery"),
+  wr_name = c("Ja'Marr Chase", "Stefon Diggs")
+)
+
+qbRbWr_projections2$qb_mean <- c(
+  player_stats_weekly_recent_summary$mean[which(player_stats_weekly_recent_summary$position == "QB" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$qb_name[1])],
+  player_stats_weekly_recent_summary$mean[which(player_stats_weekly_recent_summary$position == "QB" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$qb_name[2])])
+
+qbRbWr_projections2$qb_sd <- c(
+  player_stats_weekly_recent_summary$sd[which(player_stats_weekly_recent_summary$position == "QB" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$qb_name[1])],
+  player_stats_weekly_recent_summary$sd[which(player_stats_weekly_recent_summary$position == "QB" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$qb_name[2])])
+
+qbRbWr_projections2$rb_mean <- c(
+  player_stats_weekly_recent_summary$mean[which(player_stats_weekly_recent_summary$position == "RB" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$rb_name[1])],
+  player_stats_weekly_recent_summary$mean[which(player_stats_weekly_recent_summary$position == "RB" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$rb_name[2])])
+
+qbRbWr_projections2$rb_sd <- c(
+  player_stats_weekly_recent_summary$sd[which(player_stats_weekly_recent_summary$position == "RB" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$rb_name[1])],
+  player_stats_weekly_recent_summary$sd[which(player_stats_weekly_recent_summary$position == "RB" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$rb_name[2])])
+
+qbRbWr_projections2$wr_mean <- c(
+  player_stats_weekly_recent_summary$mean[which(player_stats_weekly_recent_summary$position == "WR" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$wr_name[1])],
+  player_stats_weekly_recent_summary$mean[which(player_stats_weekly_recent_summary$position == "WR" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$wr_name[2])])
+
+qbRbWr_projections2$wr_sd <- c(
+  player_stats_weekly_recent_summary$sd[which(player_stats_weekly_recent_summary$position == "WR" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$wr_name[1])],
+  player_stats_weekly_recent_summary$sd[which(player_stats_weekly_recent_summary$position == "WR" & player_stats_weekly_recent_summary$name == qbRbWr_projections2$wr_name[2])])
+
+qbRbWr_projections2$qb_lastName <- sapply(strsplit(qbRbWr_projections2$qb_name, " "), function(x) tail(x, 1))
+qbRbWr_projections2$rb_lastName <- sapply(strsplit(qbRbWr_projections2$rb_name, " "), function(x) tail(x, 1))
+qbRbWr_projections2$wr_lastName <- sapply(strsplit(qbRbWr_projections2$wr_name, " "), function(x) tail(x, 1))
+
+team_projections2 <- data.frame(
+  team = apply(hiLo_permutations, 1, paste0, collapse = ""),
+  team_name = NA,
+  team_mean = NA,
+  team_sd = NA,
+  qb = hiLo_permutations$Var1,
+  rb = hiLo_permutations$Var2,
+  wr = hiLo_permutations$Var3,
+  qb_name = NA,
+  qb_lastName = NA,
+  qb_mean = NA,
+  qb_sd = NA,
+  rb_name = NA,
+  rb_lastName = NA,
+  rb_mean = NA,
+  rb_sd = NA,
+  wr_name = NA,
+  wr_lastName = NA,
+  wr_mean = NA,
+  wr_sd = NA
+)
+
+team_projections2$qb_name[which(team_projections2$qb == "Hi")] <- qbRbWr_projections2$qb_name[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$qb_name[which(team_projections2$qb == "Lo")] <- qbRbWr_projections2$qb_name[which(qbRbWr_projections2$condition == "Lo")]
+team_projections2$qb_lastName[which(team_projections2$qb == "Hi")] <- qbRbWr_projections2$qb_lastName[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$qb_lastName[which(team_projections2$qb == "Lo")] <- qbRbWr_projections2$qb_lastName[which(qbRbWr_projections2$condition == "Lo")]
+team_projections2$qb_mean[which(team_projections2$qb == "Hi")] <- qbRbWr_projections2$qb_mean[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$qb_mean[which(team_projections2$qb == "Lo")] <- qbRbWr_projections2$qb_mean[which(qbRbWr_projections2$condition == "Lo")]
+team_projections2$qb_sd[which(team_projections2$qb == "Hi")] <- qbRbWr_projections2$qb_sd[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$qb_sd[which(team_projections2$qb == "Lo")] <- qbRbWr_projections2$qb_sd[which(qbRbWr_projections2$condition == "Lo")]
+
+team_projections2$rb_name[which(team_projections2$rb == "Hi")] <- qbRbWr_projections2$rb_name[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$rb_name[which(team_projections2$rb == "Lo")] <- qbRbWr_projections2$rb_name[which(qbRbWr_projections2$condition == "Lo")]
+team_projections2$rb_lastName[which(team_projections2$rb == "Hi")] <- qbRbWr_projections2$rb_lastName[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$rb_lastName[which(team_projections2$rb == "Lo")] <- qbRbWr_projections2$rb_lastName[which(qbRbWr_projections2$condition == "Lo")]
+team_projections2$rb_mean[which(team_projections2$rb == "Hi")] <- qbRbWr_projections2$rb_mean[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$rb_mean[which(team_projections2$rb == "Lo")] <- qbRbWr_projections2$rb_mean[which(qbRbWr_projections2$condition == "Lo")]
+team_projections2$rb_sd[which(team_projections2$rb == "Hi")] <- qbRbWr_projections2$rb_sd[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$rb_sd[which(team_projections2$rb == "Lo")] <- qbRbWr_projections2$rb_sd[which(qbRbWr_projections2$condition == "Lo")]
+
+team_projections2$wr_name[which(team_projections2$wr == "Hi")] <- qbRbWr_projections2$wr_name[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$wr_name[which(team_projections2$wr == "Lo")] <- qbRbWr_projections2$wr_name[which(qbRbWr_projections2$condition == "Lo")]
+team_projections2$wr_lastName[which(team_projections2$wr == "Hi")] <- qbRbWr_projections2$wr_lastName[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$wr_lastName[which(team_projections2$wr == "Lo")] <- qbRbWr_projections2$wr_lastName[which(qbRbWr_projections2$condition == "Lo")]
+team_projections2$wr_mean[which(team_projections2$wr == "Hi")] <- qbRbWr_projections2$wr_mean[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$wr_mean[which(team_projections2$wr == "Lo")] <- qbRbWr_projections2$wr_mean[which(qbRbWr_projections2$condition == "Lo")]
+team_projections2$wr_sd[which(team_projections2$wr == "Hi")] <- qbRbWr_projections2$wr_sd[which(qbRbWr_projections2$condition == "Hi")]
+team_projections2$wr_sd[which(team_projections2$wr == "Lo")] <- qbRbWr_projections2$wr_sd[which(qbRbWr_projections2$condition == "Lo")]
+
+team_projections2$team_name <- apply(
+  team_projections2[, c("qb_lastName", "rb_lastName", "wr_lastName")],
+  1,
+  paste,
+  collapse = "/"
+)
+
+team_projections2$team_mean <- rowSums(team_projections2[,c("qb_mean","rb_mean","wr_mean")])
+team_projections2$team_sd <- rowSums(team_projections2[,c("qb_sd","rb_sd","wr_sd")])
+```
+
+In @fig-teamPortfolio2, we depict the combined mean of projected fantasy points for a Quarterback/Running Back/Wide Receiver combination as a function of the standard deviation of their projected fantasy points across sources.
+
+```{r}
+#| label: fig-teamPortfolio2
+#| fig-cap: "Mean of Weekly Fantasy Points For a Quarterback/Running Back/Wide Receiver Combination as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."
+#| fig-alt: "Mean of Weekly Fantasy Points For a Quarterback/Running Back/Wide Receiver Combination as a Function of the Week-to-Week Standard Deviation of their Fantasy Points."
+
+ggplot2::ggplot(
+  data = team_projections2,
+  aes(
+    x = team_sd,
+    y = team_mean,
+    label = team_name)) +
+  geom_point() +
+  ggrepel::geom_text_repel(aes(label = team_name)) +
+  #geom_text(
+  #  hjust = 0,
+  #  vjust = 0) +
+  #geom_smooth() +
+  coord_cartesian(
+    xlim = c(NA,NA),
+    ylim = c(NA,NA),
+    expand = TRUE) +
+  labs(
+    x = "Standard Deviation of Players' Projected Fantasy Points (Season) Across Sources",
+    y = "Mean of Players' Projected Fantasy Points (Season) Across Sources",
+    title = "Mean vs Standard Deviation of Players' Projected Fantasy Points"
+  ) +
+  theme_classic()
+```
+
+```{r}
 player_stats_weekly_recent_summary_noNA <- player_stats_weekly_recent_summary %>% 
@@
 
+The efficient frontier of fantasy points across weeks is in @fig-efficientFrontierAcrossWeeks.
+
 ```{r}
@@
   xlab = "Standard Deviation of Fantasy Points Across Weeks",
-  ylab = "Weekly Fantasy Points")
+  ylab = "Weekly Fantasy Points",
+  main = "Efficient Frontier of Fantasy Points Across Weeks")
 ```

### 2025-06-21 download-football-data.qmd
20250621 - variable names

@@
 load(file = "./data/nfl_schedules_raw.RData")
+load(file = "./data/nfl_standings_raw.RData")
 load(file = "./data/nfl_combine_raw.RData")
 load(file = "./data/nfl_draftPicks_raw.RData")
+load(file = "./data/nfl_draftValues_raw.RData")
 load(file = "./data/nfl_depthCharts_raw.RData")
+load(file = "./data/nfl_trades_raw.RData")
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp_raw.RData", fsep = ""))
@@
 load(file = "./data/players_projections_seasonal_raw.RData")
+load(file = "./data/players_projections_weekly_raw.RData")
 
@@
 load(file = "./data/nfl_schedules.RData")
+load(file = "./data/nfl_standings.RData")
 load(file = "./data/nfl_combine.RData")
 load(file = "./data/nfl_draftPicks.RData")
+load(file = "./data/nfl_draftValues.RData")
 load(file = "./data/nfl_depthCharts.RData")
+load(file = "./data/nfl_trades.RData")
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
@@
 load(file = "./data/players_projectedPoints_seasonal.RData")
+load(file = "./data/players_projectedPoints_weekly.RData")
 
@@
 load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_players) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_teams) %>% as.data.frame()
+```
+
 ```{r}
@@
   )
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_playerIDs) %>% as.data.frame()
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_rosters) %>% as.data.frame()
+
+names(nfl_rosters_weekly) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_schedules) %>% as.data.frame()
+```
+
 ```{r}
@@
   head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_standings) %>% as.data.frame()
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_combine) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_draftPicks) %>% as.data.frame()
+```
+
 ```{r}
@@
   head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_draftValues) %>% as.data.frame()
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_depthCharts) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_trades) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_pbp) %>% as.data.frame()
+```
+
 ```{r}
@@
   head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_4thdown) %>% as.data.frame()
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_participation) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_actualStats_offense_weekly) %>% as.data.frame()
+
+names(nfl_actualStats_defense_weekly) %>% as.data.frame()
+
+names(nfl_actualStats_kicking_weekly) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_injuries) %>% as.data.frame()
+```
+
 ```{r}
@@
   head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_snapCounts) %>% as.data.frame()
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_espnQBR_seasonal) %>% as.data.frame()
+
+names(nfl_espnQBR_weekly) %>% as.data.frame()
+```
+
 ```{r}
@@
   head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_nextGenStats_weekly) %>% as.data.frame()
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_advancedStatsPFR_seasonal) %>% as.data.frame()
+
+names(nfl_advancedStatsPFR_weekly) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_playerContracts) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_ftnCharting) %>% as.data.frame()
+```
+
 ```{r}
@@
   head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_rankings_weekly) %>% as.data.frame()
 ```
@@
   head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_expectedFantasyPoints_pbp) %>% as.data.frame()
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(dplyr::bind_rows(players_projectedPoints_seasonal)) %>% as.data.frame()
+
+names(players_projections_seasonal_average) %>% as.data.frame()
+
+names(dplyr::bind_rows(players_projectedPoints_weekly)) %>% as.data.frame()
+
+names(players_projections_weekly_average) %>% as.data.frame()
+```
+
 ```{r}
@@
 nfl_actualStats_seasonal_team_inclPost[sapply(nfl_actualStats_seasonal_team_inclPost, is.infinite)] <- NA
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_actualStats_seasonal_player) %>% as.data.frame()
+
+names(nfl_actualStats_seasonal_team) %>% as.data.frame()
+
+names(nfl_actualStats_seasonal_player_inclPost) %>% as.data.frame()
+
+names(nfl_actualStats_seasonal_team_inclPost) %>% as.data.frame()
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_actualStats_career_player) %>% as.data.frame()
+
+names(nfl_actualStats_career_player_inclPost) %>% as.data.frame()
+```
+
 ```{r}
@@
   head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_actualStats_weekly_player) %>% as.data.frame()
+
+names(nfl_actualStats_weekly_team) %>% as.data.frame()
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_actualFantasyPoints_player_weekly) %>% as.data.frame()
+
+names(nfl_actualFantasyPoints_dst_weekly) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(nfl_actualFantasyPoints_player_seasonal) %>% as.data.frame()
+
+names(nfl_actualFantasyPoints_dst_seasonal) %>% as.data.frame()
+```
+
 ```{r}
@@
   select(player_id, display_name, first_name, last_name, position_group, position, fantasyPoints)
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_actualFantasyPoints_player_career) %>% as.data.frame()
 ```
@@
 
+Here are the variable names:
+
+```{r}
+names(player_stats_weekly) %>% as.data.frame()
+```
+
 ```{r}
@@
 
+Here are the variable names:
+
+```{r}
+names(player_stats_seasonal) %>% as.data.frame()
+```
+
 ```{r}

### 2025-06-21 simulation.qmd
20250621 - clean up code

@@
 ```{r}
-proj_dists_tbl <- all_proj |>
-  filter(!is.na(id) & id != "") |>
-  group_by(id) |>
+proj_dists_tbl <- all_proj %>%
+  filter(!is.na(id) & id != "") %>%
+  group_by(id) %>%
   summarise(
@@
 
-proj_dists <- proj_dists_tbl |>
-  filter(!is.na(id) & id != "") |>
-  distinct(id, .keep_all = TRUE) |>
+proj_dists <- proj_dists_tbl %>%
+  filter(!is.na(id) & id != "") %>%
+  distinct(id, .keep_all = TRUE) %>%
   (\(x) setNames(x$dist_info, x$id))()
@@
 ```{r}
-proj_dists_tbl |>
+proj_dists_tbl %>%
   dplyr::mutate(
     dist_type = purrr::map_chr(dist_info, ~ .x$type)
-  ) |>
+  ) %>%
   dplyr::count(dist_type)
@@
 ```{r}
-Design <- proj_dists_tbl |>
-  transmute(
+Design <- proj_dists_tbl %>%
+  dplyr::mutate(
     id,
-    n_sources = n_proj # how many projections the player had
+    n_sources = n_proj,
+    .keep = "none"
   )

### 2025-06-21 sports-cognitive-psychology.qmd
20250621 - update REF

@@
 There is mixed evidence on icing the kicker.
-Some studies find some evidence for lower percentage of iced kicks made compared to non-iced kicks [@Goldschmied2024], whereas other studies do not [e.g., @Berry2004; @GonzalezSanchez2024; @Moskowitz2011].
+Some studies find some evidence for lower percentage of iced kicks made compared to non-iced kicks [@Goldschmied2025], whereas other studies do not [e.g., @Berry2004; @GonzalezSanchez2024; @Moskowitz2011].
 This suggests that, if there is such an effect, the [effect size](#sec-practicalSignificance) is likely very small, and it may better suit the team to use your timeouts in other situations.

### 2025-06-22 modern-portfolio-theory.qmd
20250621 - reorder
20250621 - fold code

@@
 
+The formula for the variance (and standard deviation) of the sum of two variables is in @eq-varianceSumTwoVars:
+
+$$
+\begin{aligned}
+  \text{Var}(x + y) &= \text{Var}(x) + \text{Var}(y) + 2\text{Cov}(x, y)\\
+  s^2_{(x + y)} &= s^2_x + s^2_y +  2\text{Cov}(x, y)\\
+  \text{Std Dev}(x + y) &= \sqrt{\text{Var}(x) + \text{Var}(y) + 2\text{Cov}(x, y)}\\
+  s_{(x + y)} &= \sqrt{s^2_x + s^2_y + 2\text{Cov}(x, y)}\\
+\end{aligned}
+$$ {#eq-varianceSumTwoVars}
+
+If we assume that players' performance is uncorrelated with one another, the covariance between the players is zero, so the formula simplifies to @eq-varianceSumTwoVarsNoCov:
+
+$$
+\begin{aligned}
+  \text{Var}(x + y) &= \text{Var}(x) + \text{Var}(y)\\
+  s^2_{(x + y)} &= s^2_x + s^2_y\\
+  \text{Std Dev}(x + y) &= \sqrt{\text{Var}(x) + \text{Var}(y)}\\
+  s_{(x + y)} &= \sqrt{s^2_x + s^2_y}\\
+\end{aligned}
+$$ {#eq-varianceSumTwoVarsNoCov}
+
+That is, if players' performance is independent of each other, the variance of two players' points is merely the sum of their variances; their standard deviation is then the square root of that.
+In reality, players' performance is not truly independent—particularly for players on the same team or, for a given game, for players on opposing teams.
+However, for players not playing in the same game, it is a reasonable assumption to make and it greatly simplifies the math.
+If you wanted to, you could account for players' covariance by using @eq-varianceSumTwoVars.
+An example variance-covariance matrix of players' performance is in @sec-efficientFrontierFantasyProjectionsGameToGameVariability.
+
 ```{r}
@@
 ```{r}
+#| code-fold: true
+
 hiLo_permutations <- expand.grid(
@@
 
-The formula for the variance (and standard deviation) of the sum of two variables is in @eq-varianceSumTwoVars:
-
-$$
-\begin{aligned}
-  \text{Var}(x + y) &= \text{Var}(x) + \text{Var}(y) + 2\text{Cov}(x, y)\\
-  s^2_{(x + y)} &= s^2_x + s^2_y +  2\text{Cov}(x, y)\\
-  \text{Std Dev}(x + y) &= \sqrt{\text{Var}(x) + \text{Var}(y) + 2\text{Cov}(x, y)}\\
-  s_{(x + y)} &= \sqrt{s^2_x + s^2_y + 2\text{Cov}(x, y)}\\
-\end{aligned}
-$$ {#eq-varianceSumTwoVars}
-
-If we assume that players' performance is uncorrelated with one another, the covariance between the players is zero, so the formula simplifies to @eq-varianceSumTwoVarsNoCov:
-
-$$
-\begin{aligned}
-  \text{Var}(x + y) &= \text{Var}(x) + \text{Var}(y)\\
-  s^2_{(x + y)} &= s^2_x + s^2_y\\
-  \text{Std Dev}(x + y) &= \sqrt{\text{Var}(x) + \text{Var}(y)}\\
-  s_{(x + y)} &= \sqrt{s^2_x + s^2_y}\\
-\end{aligned}
-$$ {#eq-varianceSumTwoVarsNoCov}
-
-That is, if players' performance is independent of each other, the variance of two players' points is merely the sum of their variances; their standard deviation is then the square root of that.
-In reality, players' performance is not truly independent—particularly for players on the same team or, for a given game, for players on opposing teams.
-However, for players not playing in the same game, it is a reasonable assumption to make and it greatly simplifies the math.
-If you wanted to, you could account for players' covariance by using @eq-varianceSumTwoVars.
-An example variance-covariance matrix of players' performance is in @sec-efficientFrontierFantasyProjectionsGameToGameVariability.
-
 ### Based on Historical Game-to-Game Variability {#sec-efficientFrontierFantasyProjectionsGameToGameVariability}
@@
 ```{r}
+#| code-fold: true
+
 qbRbWr_projections2 <- data.frame(

### 2025-06-22 modern-portfolio-theory.qmd
20250622 - clarifying text

@@
 
+A scatterplot of the expected returns versus risk, based on projected fantasy points, is in @fig-projectedPointsMeanVsSD.
+
 ```{r}
@@
 
+A scatterplot of the expected returns versus risk, based on projected fantasy points, is in @fig-projectedPointsMeanVsSDPDF.
+
 ```{r}
@@
 
+Now let's consider the expected returns versus risk for a combination of players.
 The formula for the variance (and standard deviation) of the sum of two variables is in @eq-varianceSumTwoVars:
@@
 ```
+
+Below is code for obtaining the expected returns and risk for various combinations of Quarterback, Running Back, and Wide Receiver.
 
@@
 
+Now let's generate the efficient frontier across all players.
+
 ```{r}
@@
 
+A scatterplot of the expected returns versus risk, based on weekly fantasy points, is in @fig-fantasyPointsMeanVsSD.
+
 ```{r}
@@
 
+A scatterplot of the expected returns versus risk, based on weekly fantasy points, is in @fig-fantasyPointsMeanVsSDPDF.
+
 ```{r}
@@
 :::
+
+Now let's consider the expected returns versus risk for a combination of players.
+Below is code for obtaining the expected returns and risk for various combinations of Quarterback, Running Back, and Wide Receiver.
 
@@
   labs(
-    x = "Standard Deviation of Players' Projected Fantasy Points (Season) Across Sources",
-    y = "Mean of Players' Projected Fantasy Points (Season) Across Sources",
+    x = "Standard Deviation of Players' Fantasy Points From Week-to-Week",
+    y = "Mean of Players' Fantasy Points Across Weeks",
     title = "Mean vs Standard Deviation of Players' Projected Fantasy Points"
@@
 ```
+
+Now let's generate the efficient frontier across all players.
 
@@
 
+In these examples, we have computed the efficient frontier across *all* players.
+However, it would be more relevant to establish the efficient frontier for the possible combinations of players (e.g., one Quarterback, two Running Backs, two Wide Receivers, one Tight End, one Kicker, etc.).
+Nevertheless, the above examples illustrate how one might do this; one could extend the examples to consider whole teams rather than just three players at a time.
+
 ## Conclusion {#sec-modernPortfolioTheoryConclusion}

### 2025-06-22 modern-portfolio-theory.qmd
20250622 - simplify code; add clarifying text

@@
 
+We can examine the efficient frontier of fantasy performance based on the mean and variability of players' projected fantasy points across projection sources.
+
 ```{r}
@@
 Thus, to obtain a lineup that scores greater fantasy points, it may be necessary to take on greater risk.
-Moreover, you would not want take take on more risk for the same number of expected fantasy points; in addition, you would not want to score fewer fantasy points for the same amount of risk [@Hitchings2012].
+Moreover, you would not want take take on more risk for the same number of expected fantasy points; in addition, you would not want to score fewer fantasy points for the same amount of risk [@Hitchings2012; archived at <https://perma.cc/NE35-G6LR>].
 
@@
   geom_point() +
-  ggrepel::geom_text_repel(aes(label = team_name)) +
-  #geom_text(
-  #  hjust = 0,
-  #  vjust = 0) +
-  #geom_smooth() +
-  coord_cartesian(
-    xlim = c(NA,NA),
-    ylim = c(NA,NA),
-    expand = TRUE) +
+  ggrepel::geom_text_repel() +
   labs(
@@
 
-<https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league> [@Hitchings2012; archived at <https://perma.cc/JQ6G-KSRT>]
+We can also examine the efficient frontier of fantasy performance based on the mean and variability of players' projected week-to-week fantasy points [@Hitchings2012; archived at <https://perma.cc/JQ6G-KSRT>].
 

### 2025-06-23 evaluating-prediction-accuracy.qmd
20250622 - context

@@
 Below are estimates of projection accuracy and [calibration plots](#sec-calibrationPlot) for [seasonal](#sec-accuracyProjectionsSeasonal) and [weekly](#sec-accuracyProjectionsWeekly) projections.
-Projections were poorly [calibrated](#sec-calibrationIndices) with respect to actual fantasy points.
+In general, projections showed moderate accuracy in predicting players' fantasy performance.
+Projections showed moderate accuracy in terms of [discrimination](#sec-discrimination), based on [$R^2$](#sec-coefficientOfDetermination).
+However, projections were not highly accurate.
+Projections showed some [miscalibration](#sec-calibrationIndices) with respect to actual fantasy points.
 In general, projected showed a modest tendency to underpredict performance at lower levels of projected points and a moderate-to-strong tendency overpredict performance at higher levels of projected points, an example of the overextremity form of [miscalibration](#sec-calibration).
@@
 That is, projection sources appear to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their projections of player performance.
+Relatedly, crowd-averaged projections tend to be more accurate than individual sources of projections [CITATION].
 

### 2025-06-23 simulation.qmd
20250622 - context

@@
 ## Simulation of Projected Statistics and Points {#sec-simulationProjectedPoints}
+
+Below, we perform [bootstrapping](#sec-simulationProjectedPointsBootstrapping) and [Monte Carlo](#sec-simulationProjectedPointsMonteCarlo) simulations of projected statistics and points.
+However, it is worth noting that—as for any simulation—the quality of the results depend on the quality of the inputs.
+In this case, the quality of the simulation depends on the quality of the projections.
+If the projections are no good, the simulation results will not be trustworthy.
+Garbage in, garbage out.
+As we evaluated in @sec-accuracyProjections, projections tend to show moderate accuracy for fantasy performance, but they are not highly accurate.
+Thus, we should treat simulation results arising from fantasy projections with a good dose of skepticism.
 

### 2025-06-24 machine-learning.qmd
20250624 - updates

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
-load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
+#load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
 load(file = "./data/nfl_injuries.RData")
@@
 load(file = "./data/nfl_participation.RData")
-load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
+#load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
 load(file = "./data/nfl_injuries.RData")
@@
   names(nfl_participation),
-  names(nfl_actualFantasyPoints_player_weekly),
+  #names(nfl_actualFantasyPoints_player_weekly),
   names(nfl_injuries),
@@
 ## Use common name for the same (gsis_id) ID variable
-nfl_actualFantasyPoints_player_weekly <- nfl_actualFantasyPoints_player_weekly %>% 
-  rename(gsis_id = player_id)
-
-nfl_actualFantasyPoints_player_seasonal <- nfl_actualFantasyPoints_player_seasonal %>% 
-  rename(gsis_id = player_id)
+
+#nfl_actualFantasyPoints_player_weekly <- nfl_actualFantasyPoints_player_weekly %>% 
+#  rename(gsis_id = player_id)
+#
+#nfl_actualFantasyPoints_player_seasonal <- nfl_actualFantasyPoints_player_seasonal %>% 
+#  rename(gsis_id = player_id)
 
@@
 
-nfl_rosters_weekly <- nfl_rosters %>% 
+nfl_rosters_weekly <- nfl_rosters_weekly %>% 
   mutate(
@@
     season = as.integer(season),
-    receptions = as.integer(receptions))
+    receptions = as.integer(receptions)) %>% 
+  distinct(gsis_id, season, week, .keep_all = TRUE) # drop duplicated rows
 
@@
 ) %>% 
-  filter(!is.na(week))
+  filter(!is.na(week)) %>% 
+  select(gsis_id, season, week, full_name, position, team, everything())
 
@@
 
-- seasonalData: 1999-2023
-- seasonalData_lag: 1999-2022 (predicting fantasy points in 2023)
-- newData_seasonal: 2023 (to be used for predicting fantasy points in 2024)
+- seasonalData: 1999-2024: i.e. all data currently available
+- seasonalData_lag: 1999-2023 (predicting fantasy points in 2024)
+- newData_seasonal: 2024 (to be used for predicting fantasy points in 2025)
 
@@
 - seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
-- seasonalData_lag_train: 1999-2022 (predicting fantasy points in 2023), most players
-- seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), some retired players
-- (eventually, newData_seasonal, which is derived from the imputed version of seasonalData_lag_all, and then keeps only data from 2023, thus predicting fantasy points in 2024, and removing the fantasy_points_lag variable)
+- seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most players
+- seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some of the) retired players; 2023 rather than 2024 because don't have fantasy points for retired players in most recent year
+- (eventually, newData_seasonal, which is derived from the imputed version of seasonalData_lag_all, and then keeps only data from 2024, thus predicting fantasy points in 2025, and removing the fantasy_points_lag variable)
 
@@
 - seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
-- seasonalData_lag_train: 1999-2022 (predicting fantasy points in 2023), most players
-- seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), some retired players
+- seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most players
+- seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some of the) retired players
 
@@
 MODEL:
-- training data (imputed version of seasonalData_lag_all): all seasons (except 2023 predicting 2024), most players
+- training data (imputed version of seasonalData_lag_all): all seasons (except 2024 predicting 2025), most players
 
 EVALUATE MODEL PREDICTIONS:
-- test data (imputed version of seasonalData_lag_test): all seasons (except 2023 predicting 2024), some retired players
+- test data (imputed version of seasonalData_lag_test): all seasons (except 2024 predicting 2025), some retired players
 
 GENERATE MODEL PREDICTIONS
-- next year data (newData_seasonal): 2023 (predicting 2024)
+- next year data (newData_seasonal): 2024 (predicting 2025)
 

### 2025-06-29 machine-learning.qmd
20250629 - notes to self

@@
 
-CURRENTLY (WILL CHANGE):
-
-- seasonalData: 1999-2024: i.e. all data currently available
-- seasonalData_lag: 1999-2023 (predicting fantasy points in 2024)
-- newData_seasonal: 2024 (to be used for predicting fantasy points in 2025)
-
-to create:
-
-- seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
-- seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most players
-- seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some of the) retired players; 2023 rather than 2024 because don't have fantasy points for retired players in most recent year
-- (eventually, newData_seasonal, which is derived from the imputed version of seasonalData_lag_all, and then keeps only data from 2024, thus predicting fantasy points in 2025, and removing the fantasy_points_lag variable)
-
-To impute:
-
-- seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
-- seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most players
-- seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some of the) retired players
-
-IMPUTATION AND MODEL PROCESS:
-
-IMPUTATION:
-- training data
-- test data
-- all data (training and test data), used for generating next year predictions
-
-MODEL:
-- training data (imputed version of seasonalData_lag_all): all seasons (except 2024 predicting 2025), most players
-
-EVALUATE MODEL PREDICTIONS:
-- test data (imputed version of seasonalData_lag_test): all seasons (except 2024 predicting 2025), some retired players
-
-GENERATE MODEL PREDICTIONS
-- next year data (newData_seasonal): 2024 (predicting 2025)
+```{r}
+#| eval: false
+#| include: false
+
+# Notes to self
+
+# CURRENTLY (WILL CHANGE):
+# 
+# - seasonalData: 1999-2024: i.e. all data currently available
+# - seasonalData_lag: 1999-2023 (predicting fantasy points in 2024)
+# - newData_seasonal: 2024 (to be used for predicting fantasy points in 2025)
+
+# to create:
+# 
+# - seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
+# - seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most #players
+# - seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some #of the) retired players; 2023 rather than 2024 because don't have fantasy #points for retired players in most recent year
+# - (eventually, newData_seasonal, which is derived from the imputed version of #seasonalData_lag_all, and then keeps only data from 2024, thus predicting #fantasy points in 2025, and removing the fantasy_points_lag variable)
+
+# To impute:
+# 
+# - seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
+# - seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most #players
+# - seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some #of the) retired players
+
+# IMPUTATION AND MODEL PROCESS:
+# 
+# IMPUTATION:
+# - training data
+# - test data
+# - all data (training and test data), used for generating next year #predictions
+# 
+# MODEL:
+# - training data (imputed version of seasonalData_lag_all): all seasons #(except 2024 predicting 2025), most players
+# 
+# EVALUATE MODEL PREDICTIONS:
+# - test data (imputed version of seasonalData_lag_test): all seasons (except #2024 predicting 2025), some retired players
+# 
+# GENERATE MODEL PREDICTIONS
+# - next year data (newData_seasonal): 2024 (predicting 2025)
+```
 
@@
 
-NOTES:
-
-- CONSIDER USING RETIRED PLAYERS AS THE HOLDOUT SAMPLE
-- CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
-- CONSIDER IMPUTING THE TRAINING AND TEST DATA SEPARATELY BY POSITION
+```{r}
+#| eval: false
+#| include: false
+
+# Notes to self
+
+# - CONSIDER USING RETIRED PLAYERS AS THE HOLDOUT SAMPLE
+# - CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
+# - CONSIDER IMPUTING THE TRAINING AND TEST DATA SEPARATELY BY POSITION
+```
 
@@
   as.matrix()
-
+```
+
+```{r}
 # RBs
@@
   as.matrix()
-
+```
+
+```{r}
 # WRs
@@
   as.matrix()
-
+```
+
+```{r}
 # TEs

### 2025-06-30 machine-learning.qmd
20250630 - early draft for machine learning

@@
 ## Rename variables
-#-todo: rename variables in expected fantasy points so they don't get coalesced with actual points
 nfl_draftPicks <- nfl_draftPicks %>%
@@
 
+## Subset variables
+nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
+  select(gsis_id:position, contains("_exp"), contains("_diff"), contains("_team")) #drop "raw stats" variables (e.g., rec_yards_gained) so they don't get coalesced with actual stats
+
 # Check duplicate ids
@@
 
+Identify objects with shared variable names:
+
+```{r}
+dplyr::intersect(
+  names(nfl_players),
+  names(nfl_draftPicks))
+
+length(na.omit(nfl_players$position)) # use by default (more cases)
+length(na.omit(nfl_draftPicks$position))
+
+dplyr::intersect(
+  names(player_stats_seasonal_offense),
+  names(nfl_advancedStatsPFR_seasonal))
+
+length(na.omit(player_stats_seasonal_offense$season)) # use by default (more cases)
+length(na.omit(nfl_advancedStatsPFR_seasonal$season))
+
+length(na.omit(player_stats_seasonal_offense$team)) # use by default (more cases)
+length(na.omit(nfl_advancedStatsPFR_seasonal$team))
+
+length(na.omit(player_stats_seasonal_offense$age)) # use by default (more cases)
+length(na.omit(nfl_advancedStatsPFR_seasonal$age))
+
+dplyr::intersect(
+  names(nfl_rosters_weekly),
+  names(nfl_expectedFantasyPoints_weekly))
+
+length(na.omit(nfl_rosters_weekly$season)) # use by default (more cases)
+length(na.omit(nfl_expectedFantasyPoints_weekly$season))
+
+length(na.omit(nfl_rosters_weekly$week)) # use by default (more cases)
+length(na.omit(nfl_expectedFantasyPoints_weekly$week))
+
+length(na.omit(nfl_rosters_weekly$position)) # use by default (more cases)
+length(na.omit(nfl_expectedFantasyPoints_weekly$position))
+
+length(na.omit(nfl_rosters_weekly$full_name)) # use by default (more cases)
+length(na.omit(nfl_expectedFantasyPoints_weekly$full_name))
+```
+
 ### Merge Data {#sec-machineLearningMergeData}
@@
     by = c("gsis_id"),
-    conflict = coalesce_xy)
+    conflict = coalesce_xy) # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
 
@@
     by = c("gsis_id","season"),
-    conflict = coalesce_xy)
+    conflict = coalesce_xy) # where the objects have the same variable name (e.g., team), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
 
@@
     by = c("gsis_id","season","week"),
-    conflict = coalesce_xy)
-
+    conflict = coalesce_xy) # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
+```
+
+Identify objects with shared variable names:
+
+```{r}
+dplyr::intersect(
+  names(playerSeasonMerged),
+  names(playerMerged))
+```
+
+```{r}
 seasonalData <- powerjoin::power_full_join(
   playerSeasonMerged,
-  playerMerged %>% select(-age, -years_of_experience, -team, -team_abbr, -team_seq, -current_team_id),
+  playerMerged %>% select(-age, -years_of_experience, -team, -team_abbr, -team_seq, -current_team_id), # drop variables from id objects that change from year to year (and thus are not necessarily accurate for a given season)
   by = "gsis_id",
-  conflict = coalesce_xy
+  conflict = coalesce_xy # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
 ) %>% 
@@
   select(gsis_id, season, player_display_name, position, team, games, everything())
-
+```
+
+```{r}
+dplyr::intersect(
+  names(playerSeasonWeekMerged),
+  names(seasonalData))
+```
+
+```{r}
 seasonalAndWeeklyData <- powerjoin::power_full_join(
@@
   by = c("gsis_id","season"),
-  conflict = coalesce_xy
+  conflict = coalesce_xy # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
 ) %>% 
@@
   select(gsis_id, season, week, full_name, position, team, everything())
-
+```
+
+```{r}
 # Duplicate cases
@@
 seasonalData_lag %>% 
-  select(gsis_id, player_display_name, season, fantasyPoints, fantasyPoints_lag)
+  select(gsis_id, player_display_name, season, fantasyPoints, fantasyPoints_lag) # verify that lagging worked as expected
 ```
@@
 
-# Notes to self
+# Notes to self:
+# - Using only retired players as the holdout sample, so that we can include all current/active players in the training model
 
@@
 # - seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
-# - seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most #players
-# - seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some #of the) retired players; 2023 rather than 2024 because don't have fantasy #points for retired players in most recent year
-# - (eventually, newData_seasonal, which is derived from the imputed version of #seasonalData_lag_all, and then keeps only data from 2024, thus predicting #fantasy points in 2025, and removing the fantasy_points_lag variable)
+# - seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most players
+# - seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some of the) retired players; 2023 rather than 2024 because don't have fantasy points for retired players in most recent year
+# - (eventually, newData_seasonal, which is derived from the imputed version of seasonalData_lag_all, and then keeps only data from 2024, thus predicting fantasy points in 2025, and removing the fantasy_points_lag variable)
 
@@
 # - seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
-# - seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most #players
-# - seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some #of the) retired players
+# - seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most players
+# - seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some of the) retired players
 
@@
 # - test data
-# - all data (training and test data), used for generating next year #predictions
+# - all data (training and test data), used for generating next year predictions
 # 
 # MODEL:
-# - training data (imputed version of seasonalData_lag_all): all seasons #(except 2024 predicting 2025), most players
+# - training data (imputed version of seasonalData_lag_all): all seasons (except 2024 predicting 2025), most players
 # 
 # EVALUATE MODEL PREDICTIONS:
-# - test data (imputed version of seasonalData_lag_test): all seasons (except #2024 predicting 2025), some retired players
+# - test data (imputed version of seasonalData_lag_test): all seasons (except 2024 predicting 2025), (some of the) retired players
 # 
@@
 
-set.seed(52242)
+set.seed(52242) # for reproducibility (to keep the same train/holdout players)
 
@@
 
+set.seed(52242) # for reproducibility (to keep the same train/holdout players); added here to prevent a downstream error with predict.missRanger() due to missingness; this suggests that an error can arise from including a player in the holdout sample who has missingness in particular variables; would be good to identify which player(s) in the holdout sample evoke that error to identify the kinds of missingness that yield the error
+
 activeWRs <- unique(seasonalData_lag_wr_all$gsis_id[which(seasonalData_lag_wr_all$season == max(seasonalData_lag_wr_all$season, na.rm = TRUE))])
@@
 
-# - CONSIDER USING RETIRED PLAYERS AS THE HOLDOUT SAMPLE
 # - CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
@@
 
-newData_seasonalQB <- left_join(
+newData_seasonalQB <- dplyr::left_join(
   newData_seasonalQB,
@@
 
-newData_seasonalRB <- left_join(
+newData_seasonalRB <- dplyr::left_join(
   newData_seasonalRB,
@@
 
-newData_seasonalWR <- left_join(
+newData_seasonalWR <- dplyr::left_join(
   newData_seasonalWR,
@@
 
-newData_seasonalTE <- left_join(
+newData_seasonalTE <- dplyr::left_join(
   newData_seasonalTE,
@@
 model_likelihood <- "gaussian" # gamma
-nrounds <- 1000
+nrounds <- 2000
 ```
@@
 #### Identify Optimal Tuning Parameters {#sec-treeBoostingOptimalTuningParameters}
+
+For identifying the optimal tuning parameters for boosting, we partition the training data into inner training data and validation data.
+We randomly split the training data into 80% inner training data and 20% held-out validation data.
+We then use the [mean absolute error](#sec-meanAbsoluteError) as our index of prediction accuracy on the held-out validation data.
 
@@
 set.seed(52242)
-valid_tune_idx_qb <- sample.int(ntrain_qb, as.integer(0.2*ntrain_qb))
+valid_tune_idx_qb <- sample.int(ntrain_qb, as.integer(0.2*ntrain_qb)) #
 
@@
 param_grid_qb <- list(
-  "learning_rate" = c(1,0.1,0.01),
-  "max_depth" = c(1,2,3,5,10),
-  "min_data_in_leaf" = c(10,100,1000),
-  "lambda_l2" = c(0,1,10))
-
-other_params_qb <- list(num_leaves = 2^10)
+  "learning_rate" = c(0.2, 0.1, 0.05, 0.01),
+  "max_depth" = c(3, 5, 7),
+  "min_data_in_leaf" = c(10, 50, 100),
+  "lambda_l2" = c(0, 1, 5))
+
+other_params_qb <- list(
+  num_leaves = 2^6) # 2^n, where n is smaller than the largest max_depth
 
@@
   nrounds = nrounds,
-  early_stopping_rounds = 10,
+  early_stopping_rounds = 50,
   verbose_eval = 1,
-  metric = "mse")
+  metric = "mae")
 
@@
 ```
+
+A learning rate of 1 is very high for boosting.
+Even if a learning rate of 1 did well in tuning, I use a lower learning rate (0.2) to avoid overfitting.
+I also added some light regularization (`lambda_l2`) for better generalization.
+I also set the maximum tree depth (`max_depth`) at 7 to capture up to 3-way interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^5 (32).
+I set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 50.
 
@@
 params_qb <- list(
-  learning_rate = 1,
-  max_depth = 10,
-  num_leaves = 2^10,
-  min_data_in_leaf = 10,
-  lambda_l2 = 0,
+  learning_rate = 0.2, # 0.1,
+  max_depth = 7, #3,
+  min_data_in_leaf = 10, #50
+  lambda_l2 = 0, # 1,
+  num_leaves = 2^6, #2^5,
   num_threads = num_cores)
 
-nrounds_qb <- 7000 # identify optimal number of trees through iteration and cross-validation
+nrounds_qb <- 2000 # identify optimal number of trees through iteration and cross-validation
 

### 2025-07-01 machine-learning.qmd
20250630 - free up memory

@@
 
+```{r}
+#| label: free-up-memory1-ml
+#| include: false
+
+gc()
+```
+
 ### Specify Options {#machineLearningSpecifyOptions}
@@
 ```{r}
+#| label: prepare-data-merging-machine-learning
+
 # Prepare data for merging
@@
 
+```{r}
+#| label: free-up-memory2-ml
+#| include: false
+
+gc()
+```
+
 Identify objects with shared variable names:
@@
 ```{r}
+#| label: merge-data1-machine-learning
+
 # Create lists of objects to merge, depending on data structure: id; or id-season; or id-season-week
@@
 
+```{r}
+#| label: free-up-memory3-ml
+#| include: false
+
+gc()
+```
+
 Identify objects with shared variable names:
@@
 ```{r}
+#| label: merge-data2-machine-learning
+
 seasonalData <- powerjoin::power_full_join(
@@
 ```{r}
+#| label: free-up-memory4-ml
+#| include: false
+
+gc()
+```
+
+```{r}
 dplyr::intersect(
@@
 ```{r}
+#| label: merge-data3-machine-learning
+
 seasonalAndWeeklyData <- powerjoin::power_full_join(
@@
 ```{r}
+#| label: free-up-memory5-ml
+#| include: false
+
+gc()
+```
+
+```{r}
 # Duplicate cases
@@
 ```{r}
+#| label: additional-processing-machine-learning
+
 # Convert character and logical variables to factors
@@
 
+```{r}
+#| label: free-up-memory6-ml
+#| include: false
+
+gc()
+```
+
 ### Fill in Missing Data for Static Variables {#sec-fillMissingData}
@@
 ```{r}
+#| label: fill-in-missing-data-machine-learning
+
 seasonalData <- seasonalData %>% 
@@
 
+```{r}
+#| label: free-up-memory7-ml
+#| include: false
+
+gc()
+```
+
 ### Lag Fantasy Points {#sec-lagFantasyPoints}
@@
 ```{r}
+#| label: lag-fantasy-points-machine-learning
+
 seasonalData_lag <- seasonalData %>% 
@@
 
+```{r}
+#| label: free-up-memory8-ml
+#| include: false
+
+gc()
+```
+
 ### Subset to Predictor Variables and Outcome Variable {#sec-subsetToPredictorsAndOutcome}
@@
 ```{r}
+#| label: subset-data-machine-learning
+
 seasonalData_lag %>% select_if(~class(.) == "Date")
@@
 
+```{r}
+#| label: free-up-memory9-ml
+#| include: false
+
+gc()
+```
+
 ### Separate by Position {#sec-separateByPosition}
@@
 ```{r}
+#| label: separate-by-position-machine-learning
+
 seasonalData_lag_subsetQB <- seasonalData_lag_subset %>% 
@@
     carries:special_teams_tds, contains(".rush"), contains(".rec"))
+```
+
+```{r}
+#| label: free-up-memory10-ml
+#| include: false
+
+gc()
 ```
@@
 ```{r}
+#| label: split-test-training-machine-learning
+
 seasonalData_lag_qb_all <- seasonalData_lag_subsetQB
@@
   filter(gsis_id %in% teHoldoutIDs)
+```
+
+```{r}
+#| label: free-up-memory11-ml
+#| include: false
+
+rm(nfl_players)
+rm(nfl_teams)
+rm(nfl_rosters)
+rm(nfl_rosters_weekly)
+rm(nfl_schedules)
+rm(nfl_combine)
+rm(nfl_draftPicks)
+rm(nfl_depthCharts)
+rm(nfl_pbp)
+rm(nfl_4thdown)
+rm(nfl_participation)
+rm(nfl_actualFantasyPoints_weekly)
+rm(nfl_injuries)
+rm(nfl_snapCounts)
+rm(nfl_espnQBR_seasonal)
+rm(nfl_espnQBR_weekly)
+rm(nfl_nextGenStats_weekly)
+rm(nfl_advancedStatsPFR_seasonal)
+rm(nfl_advancedStatsPFR_weekly)
+rm(nfl_playerContracts)
+rm(nfl_ftnCharting)
+rm(nfl_playerIDs)
+rm(nfl_rankings_draft)
+rm(nfl_rankings_weekly)
+rm(nfl_expectedFantasyPoints_weekly)
+rm(nfl_expectedFantasyPoints_pbp)
+rm(nfl_actualStats_career)
+rm(nfl_actualStats_career_player)
+rm(nfl_actualStats_career_player_inclPost)
+rm(nfl_actualStats_seasonal)
+rm(nfl_actualStats_seasonal_player)
+rm(nfl_actualStats_seasonal_player_inclPost)
+rm(nfl_actualStats_seasonal_team)
+rm(nfl_actualStats_seasonal_team_inclPost)
+rm(player_stats_weekly)
+rm(player_stats_weekly_offense)
+rm(player_stats_seasonal)
+rm(player_stats_seasonal_offense)
+
+rm(playerListToMerge)
+rm(playerMerged)
+rm(playerSeasonListToMerge)
+rm(playerSeasonMerged)
+rm(playerSeasonWeekListToMerge)
+rm(playerSeasonWeekMerged)
+rm(playerSeasonWeekPositionListToMerge)
+rm(seasonalAndWeeklyData)
+
+gc()
 ```
@@
 ```{r}
+#| label: impute-qb-machine-learning
+
 # QBs
@@
 ```{r}
+#| label: free-up-memory12-ml
+#| include: false
+
+gc()
+```
+
+```{r}
+#| label: impute-rb-machine-learning
+
 # RBs
@@
 ```{r}
+#| label: free-up-memory13-ml
+#| include: false
+
+gc()
+```
+
+```{r}
+#| label: impute-wr-machine-learning
+
 # WRs
@@
 ```{r}
+#| label: free-up-memory14-ml
+#| include: false
+
+gc()
+```
+
+```{r}
+#| label: impute-te-machine-learning
+
 # TEs
@@
 
+```{r}
+#| label: free-up-memory15-ml
+#| include: false
+
+gc()
+```
+
 ## Identify Cores for Parallel Processing {#sec-coresParallel}
@@
 ```{r}
+#| label: parallel-cores-machine-learning
 #| include: false
@@
 ```{r}
+#| label: cross-sectional-machine-learning
+
 cl <- parallel::makeCluster(num_cores)
@@
 ```{r}
+#| label: specify-predictors-machine-learning
+
 pred_vars_qb <- data_train_qb_matrix %>% 
@@
 ```{r}
+#| label: specify-general-model-options-machine-learning
+
 model_likelihood <- "gaussian" # gamma
@@
 ```{r}
+#| label: tuning-parameters-machine-learning
+
 # Partition training data into inner training data and validation data
@@
 
+```{r}
+#| label: free-up-memory16-ml
+#| include: false
+
+gc()
+```
+
 A learning rate of 1 is very high for boosting.
@@
 ```{r}
+#| label: model-parameters-machine-learning
+
 gp_model_qb <- gpboost::GPModel(
@@
 ```{r}
+#| label: fit-model-machine-learning
+
 gp_model_fit_qb <- gpboost::gpb.train(
@@
 
+```{r}
+#| label: free-up-memory17-ml
+#| include: false
+
+gc()
+```
+
 #### Evaluate Accuracy of Model on Test Data {#sec-treeBoostingModelAccuracy}
@@
 ```{r}
+#| label: evaluate-model-machine-learning
+
 # Test Model on Test Data
@@
 ```{r}
+#| label: next-season-predictions-machine-learning
+
 # Generate model predictions for next season

### 2025-07-01 machine-learning.qmd
20250630 - gamma distribution

@@
 rm(nfl_ftnCharting)
-rm(nfl_playerIDs)
+#rm(nfl_playerIDs)
 rm(nfl_rankings_draft)
@@
 data_all_qb <- seasonalData_lag_qb_all_imp$data
+data_all_qb$fantasyPointsMC_lag <- scale(data_all_qb$fantasyPoints_lag, scale = FALSE) # mean-centered
 data_all_qb_matrix <- data_all_qb %>%
@@
   filter(season == max(season, na.rm = TRUE)) %>% 
-  select(-fantasyPoints_lag)
+  select(-fantasyPoints_lag, -fantasyPointsMC_lag)
 newData_qb_matrix <- data_all_qb_matrix[
@@
 
-dropCol_qb <- which(colnames(newData_qb_matrix) == "fantasyPoints_lag")
+dropCol_qb <- which(colnames(newData_qb_matrix) %in% c("fantasyPoints_lag","fantasyPointsMC_lag"))
 newData_qb_matrix <- newData_qb_matrix[, -dropCol_qb, drop = FALSE]
@@
 data_train_qb <- seasonalData_lag_qb_train_imp$data
+data_train_qb$fantasyPointsMC_lag <- scale(data_train_qb$fantasyPoints_lag, scale = FALSE) # mean-centered
 data_train_qb_matrix <- data_train_qb %>%
@@
 data_all_rb <- seasonalData_lag_rb_all_imp$data
+data_all_rb$fantasyPointsMC_lag <- scale(data_all_rb$fantasyPoints_lag, scale = FALSE) # mean-centered
 data_all_rb_matrix <- data_all_rb %>%
@@
   filter(season == max(season, na.rm = TRUE)) %>% 
-  select(-fantasyPoints_lag)
+  select(-fantasyPoints_lag, -fantasyPointsMC_lag)
 newData_rb_matrix <- data_all_rb_matrix[
@@
 
-dropCol_rb <- which(colnames(newData_rb_matrix) == "fantasyPoints_lag")
+dropCol_rb <- which(colnames(newData_rb_matrix) %in% c("fantasyPoints_lag","fantasyPointsMC_lag"))
 newData_rb_matrix <- newData_rb_matrix[, -dropCol_rb, drop = FALSE]
@@
 data_train_rb <- seasonalData_lag_rb_train_imp$data
+data_train_rb$fantasyPointsMC_lag <- scale(data_train_rb$fantasyPoints_lag, scale = FALSE) # mean-centered
 data_train_rb_matrix <- data_train_rb %>%
@@
 data_all_wr <- seasonalData_lag_wr_all_imp$data
+data_all_wr$fantasyPointsMC_lag <- scale(data_all_wr$fantasyPoints_lag, scale = FALSE) # mean-centered
 data_all_wr_matrix <- data_all_wr %>%
@@
   filter(season == max(season, na.rm = TRUE)) %>% 
-  select(-fantasyPoints_lag)
+  select(-fantasyPoints_lag, -fantasyPointsMC_lag)
 newData_wr_matrix <- data_all_wr_matrix[
@@
 
-dropCol_wr <- which(colnames(newData_wr_matrix) == "fantasyPoints_lag")
+dropCol_wr <- which(colnames(newData_wr_matrix) %in% c("fantasyPoints_lag","fantasyPointsMC_lag"))
 newData_wr_matrix <- newData_wr_matrix[, -dropCol_wr, drop = FALSE]
@@
 data_train_wr <- seasonalData_lag_wr_train_imp$data
+data_train_wr$fantasyPointsMC_lag <- scale(data_train_wr$fantasyPoints_lag, scale = FALSE) # mean-centered
 data_train_wr_matrix <- data_train_wr %>%
@@
 data_all_te <- seasonalData_lag_te_all_imp$data
+data_all_te$fantasyPointsMC_lag <- scale(data_all_te$fantasyPoints_lag, scale = FALSE) # mean-centered
 data_all_te_matrix <- data_all_te %>%
@@
   filter(season == max(season, na.rm = TRUE)) %>% 
-  select(-fantasyPoints_lag)
+  select(-fantasyPoints_lag, -fantasyPointsMC_lag)
 newData_te_matrix <- data_all_te_matrix[
@@
 
-dropCol_te <- which(colnames(newData_te_matrix) == "fantasyPoints_lag")
+dropCol_te <- which(colnames(newData_te_matrix) %in% c("fantasyPoints_lag","fantasyPointsMC_lag"))
 newData_te_matrix <- newData_te_matrix[, -dropCol_te, drop = FALSE]
@@
 data_train_te <- seasonalData_lag_te_train_imp$data
+data_train_te$fantasyPointsMC_lag <- scale(data_train_te$fantasyPoints_lag, scale = FALSE) # mean-centered
 data_train_te_matrix <- data_train_te %>%
@@
 ```{r}
-#data_train_qb_matrix[,"fantasyPoints_lag"][data_train_qb_matrix[,"fantasyPoints_lag"] <= 0] <- 0.01
+data_train_qb_matrix[,"fantasyPoints_lag"][data_train_qb_matrix[,"fantasyPoints_lag"] <= 0] <- 0.01
 ```
@@
   as_tibble() %>% 
-  select(-fantasyPoints_lag, -ageCentered20, ageCentered20Quadratic) %>% # -gsis_id
+  select(-fantasyPoints_lag, -fantasyPointsMC_lag, -ageCentered20, ageCentered20Quadratic) %>% # -gsis_id
   names()
@@
 
-model_likelihood <- "gaussian" # gamma
-nrounds <- 2000
+model_likelihood <- "gamma" # gaussian
+nrounds <- 2000 # maximum number of boosting iterations (i.e., number of trees built sequentially); more rounds = potentially better learning, but also greater risk of overfitting
 ```
@@
 param_grid_qb <- list(
-  "learning_rate" = c(0.2, 0.1, 0.05, 0.01),
-  "max_depth" = c(3, 5, 7),
-  "min_data_in_leaf" = c(10, 50, 100),
-  "lambda_l2" = c(0, 1, 5))
+  "learning_rate" = c(0.2, 0.1, 0.05, 0.01), # the step size used when updating predictions after each boosting round (high values make big updates, which can speed up learning but risk overshooting; low values are usually more accurate but require more rounds)
+  "max_depth" = c(3, 5, 7), # maximum depth (levels) of each decision tree; deeper trees capture more complex patterns and interactions but risk overfitting; shallower trees tend to generalize better
+  "min_data_in_leaf" = c(10, 50, 100), # minimum number of training examples in a leaf node; higher values = more regularization (simpler trees)
+  "lambda_l2" = c(0, 1, 5)) # L2 regularization penalty for large weights in tree splits; adds a "cost" for complexity; helps prevent overfitting by shrinking the contribution of each tree
 
 other_params_qb <- list(
-  num_leaves = 2^6) # 2^n, where n is smaller than the largest max_depth
+  num_leaves = 2^6) # maximum number of leaves per tree; controls the maximum complexity of each tree (along with max_depth); more leaves = more expressive models, but can overfit if min_data_in_leaf is too small; num_leaves must be consistent with max_depth, because deeper trees naturally support more leaves; max is: 2^n, where n is the largest max_depth
 
@@
   categorical_feature = pred_vars_qb_categorical,
-  label = data_train_qb_matrix[,"fantasyPoints_lag"])
+  label = data_train_qb_matrix[,"fantasyPoints_lag"]) #use mean-centered variable and add mean back afterward
 
@@
   nrounds = nrounds,
-  early_stopping_rounds = 50,
+  early_stopping_rounds = 50, # stops training early if the model hasn’t improved on the validation set in 50 rounds; prevents overfitting and saves time
   verbose_eval = 1,
@@
 A learning rate of 1 is very high for boosting.
-Even if a learning rate of 1 did well in tuning, I use a lower learning rate (0.2) to avoid overfitting.
+Even if a learning rate of 1 did well in tuning, I use a lower learning rate (0.1) to avoid overfitting.
 I also added some light regularization (`lambda_l2`) for better generalization.
-I also set the maximum tree depth (`max_depth`) at 7 to capture up to 3-way interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^5 (32).
-I set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 50.
+I also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^5 (32).
+I set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 10.
 
@@
 params_qb <- list(
-  learning_rate = 0.2, # 0.1,
-  max_depth = 7, #3,
-  min_data_in_leaf = 10, #50
-  lambda_l2 = 0, # 1,
-  num_leaves = 2^6, #2^5,
+  learning_rate = 0.1,
+  max_depth = 5,
+  min_data_in_leaf = 10,
+  lambda_l2 = 1,
+  num_leaves = 2^5,
   num_threads = num_cores)
 
-nrounds_qb <- 2000 # identify optimal number of trees through iteration and cross-validation
+nrounds_qb <- 123 # identify optimal number of trees through iteration and cross-validation
 
@@
   params = params_qb) # verbose = 0
-
-summary(gp_model_qb) # Estimated random effects model
 ```
@@
 gc()
+```
+
+#### Model Results {#sec-treeBoostingModelResults}
+
+```{r}
+summary(gp_model_qb) # estimated random effects model
+
+gp_model_qb_importance <- gpboost::gpb.importance(gp_model_fit_qb)
+gp_model_qb_importance
+```
+
+```{r}
+#| label: fig-featureImportanceTreeBoosting
+#| fig-cap: "Importance of Features (Predictors) in Tree Boosting Machine Learning Model."
+#| fig-alt: "Importance of Features (Predictors) in Tree Boosting Machine Learning Model."
+
+gpboost::gpb.plot.importance(gp_model_qb_importance)
 ```
@@
 
-y_pred_test_qb <- pred_test_qb[["response_mean"]]
+y_pred_test_qb <- pred_test_qb[["response_mean"]] # if outcome is mean-centered, add mean(data_train_qb_matrix[,"fantasyPoints_lag"])
 cbind(y_pred_test_qb, data_test_qb_matrix[,"fantasyPoints_lag"])

### 2025-07-01 machine-learning.qmd
20250701 - citations

@@
 We use the `caret` package [@R-caret].
-We use the `parallel` [@R-parallel] and `doParallel` [@R-doParallel] packages for parallel (faster) processing.
+We use the `doParallel` [@R-doParallel] package for parallel (faster) processing.
 
@@
 
-To combine tree-boosting with mixed models, we use the `gpboost` package [@gpboost].
+To combine tree-boosting with mixed models, we use the `gpboost` package [@R-gpboost].
 

### 2025-07-01 machine-learning.qmd
20250701 - fix errors

@@
 gc()
+```
+
+```{r}
+#| label: create-new-data-machine-learning
+
+newData_seasonal <- seasonalData %>% 
+  filter(season == max(season, na.rm = TRUE))
 ```
@@
   fantasyPoints_lag ~ ., # use all predictors
-  data = seasonalData_lag_subsetQB_imp$ximp,
+  data = seasonalData_lag_qb_train_imp$data,
   method = "rf",
@@
   fantasyPoints_lag ~ ., # use all predictors
-  data = seasonalData_lag_subsetRB_imp$ximp,
+  data = seasonalData_lag_rb_train_imp$data,
   method = "rf",
@@
   fantasyPoints_lag ~ ., # use all predictors
-  data = seasonalData_lag_subsetWR_imp$ximp,
+  data = seasonalData_lag_wr_train_imp$data,
   method = "rf",
@@
   fantasyPoints_lag ~ ., # use all predictors
-  data = seasonalData_lag_subsetTE_imp$ximp,
+  data = seasonalData_lag_te_train_imp$data,
   method = "rf",
@@
 
-newData_seasonalQB_imp$ximp$fantasyPoints_lag <- predict(
+newData_qb$fantasyPoints_lag <- predict(
   randomForest_qb,
-  newdata = newData_seasonalQB_imp$ximp
-)
-
-newData_seasonalRB_imp$ximp$fantasyPoints_lag <- predict(
+  newdata = newData_qb
+)
+
+newData_rb$fantasyPoints_lag <- predict(
   randomForest_rb,
-  newdata = newData_seasonalRB_imp$ximp
-)
-
-newData_seasonalWR_imp$ximp$fantasyPoints_lag <- predict(
+  newdata = newData_rb
+)
+
+newData_wr$fantasyPoints_lag <- predict(
   randomForest_wr,
-  newdata = newData_seasonalWR_imp$ximp
-)
-
-newData_seasonalTE_imp$ximp$fantasyPoints_lag <- predict(
+  newdata = newData_wr
+)
+
+newData_te$fantasyPoints_lag <- predict(
   randomForest_te,
-  newdata = newData_seasonalTE_imp$ximp
-)
-
-newData_seasonalQB$fantasyPoints_lag <- newData_seasonalQB_imp$ximp$fantasyPoints_lag
-newData_seasonalRB$fantasyPoints_lag <- newData_seasonalRB_imp$ximp$fantasyPoints_lag
-newData_seasonalWR$fantasyPoints_lag <- newData_seasonalWR_imp$ximp$fantasyPoints_lag
-newData_seasonalTE$fantasyPoints_lag <- newData_seasonalTE_imp$ximp$fantasyPoints_lag
-
-newData_seasonalQB <- dplyr::left_join(
-  newData_seasonalQB,
+  newdata = newData_te
+)
+
+newData_qb$fantasyPoints_lag <- newData_qb$fantasyPoints_lag
+newData_rb$fantasyPoints_lag <- newData_rb$fantasyPoints_lag
+newData_wr$fantasyPoints_lag <- newData_wr$fantasyPoints_lag
+newData_te$fantasyPoints_lag <- newData_te$fantasyPoints_lag
+
+newData_qb <- dplyr::left_join(
+  newData_qb,
   newData_seasonal %>% select(gsis_id, player_display_name, team),
@@
 
-newData_seasonalRB <- dplyr::left_join(
-  newData_seasonalRB,
+newData_rb <- dplyr::left_join(
+  newData_rb,
   newData_seasonal %>% select(gsis_id, player_display_name, team),
@@
 
-newData_seasonalWR <- dplyr::left_join(
-  newData_seasonalWR,
+newData_wr <- dplyr::left_join(
+  newData_wr,
   newData_seasonal %>% select(gsis_id, player_display_name, team),
@@
 
-newData_seasonalTE <- dplyr::left_join(
-  newData_seasonalTE,
+newData_te <- dplyr::left_join(
+  newData_te,
   newData_seasonal %>% select(gsis_id, player_display_name, team),
@@
 
-newData_seasonalQB %>%
+newData_qb %>%
   arrange(-fantasyPoints_lag) %>% 
@@
 
-newData_seasonalRB %>%
+newData_rb %>%
   arrange(-fantasyPoints_lag) %>% 
@@
 
-newData_seasonalWR %>%
+newData_wr %>%
   arrange(-fantasyPoints_lag) %>% 
@@
 
-newData_seasonalTE %>%
+newData_te %>%
   arrange(-fantasyPoints_lag) %>% 
@@
 y_pred_test_qb <- pred_test_qb[["response_mean"]] # if outcome is mean-centered, add mean(data_train_qb_matrix[,"fantasyPoints_lag"])
-cbind(y_pred_test_qb, data_test_qb_matrix[,"fantasyPoints_lag"])
+
+predictedVsActual <- data.frame(
+  predictedPoints = y_pred_test_qb,
+  actualPoints = data_test_qb_matrix[,"fantasyPoints_lag"]
+)
+
+predictedVsActual
 
 petersenlab::accuracyOverall(
-  predicted = y_pred_test_qb,
-  actual = data_test_qb_matrix[,"fantasyPoints_lag"],
+  predicted = predictedVsActual$predictedPoints,
+  actual = predictedVsActual$actualPoints,
   dropUndefined = TRUE

### 2025-07-01 index.qmd
20250701 - visualizing regression results

--- 2025-06-17 index.qmd
@@
   "petersenlab", "ffanalytics",
-  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
-  "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", "gsisdecoder", 
-  "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", 
-  "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
-  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
-  "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", 
-  "viridis", "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", #"webshot2", 
-  "stringi", "parallel", "parallelly", "doParallel", "missRanger", "ggridges", 
-  "powerjoin", "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", 
-  "correlation", "mgcv", "rms", "car", "performance", "lavaan", "mice", 
-  "miceadds", "interactions", "shiny", "PearsonDS", "robustbase", "data.table", 
-  "future", "future.apply", "SimDesign"
+  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "broom", "downlit", 
+  "xml2", "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", 
+  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
+  "dagitty", "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", 
+  "patchwork", "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", 
+  "sjstats", "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "sn", 
+  "mclust", "magrittr", "viridis", "plotly", "quantmod", "fPortfolio", "NMOF", 
+  "xts", "zoo", "forecast", "stringi", "parallel", "parallelly", "doParallel", #"webshot2"
+  "missRanger", "ggridges", "powerjoin", "caret", "gpboost", "cmdstanr", "learnr", 
+  "remotes", "corrplot", "correlation", "mgcv", "rms", "car", "performance", 
+  "lavaan", "mice", "miceadds", "interactions", "shiny", "PearsonDS", 
+  "robustbase", "data.table", "future", "future.apply", "QuantPsyc", "SimDesign"
 )

### 2025-07-01 multiple-regression.qmd
20250701 - visualizing regression results

@@
 library("robustbase")
+library("QuantPsyc")
+library("broom")
 library("tidyverse")
@@
 
-If we want to obtain standardized regression coefficients, we can standardize the [outcome variable](#sec-correlationalStudy) and each [predictor variable](#sec-correlationalStudy) using the `scale()` function:
+If we want to obtain standardized regression coefficients, we can use the `lm.beta()` function of the `QuantPsyc` package.
+
+```{r}
+QuantPsyc::lm.beta(linearRegressionModel)
+```
+
+Or, we can standardize the [outcome variable](#sec-correlationalStudy) and each [predictor variable](#sec-correlationalStudy) using the `scale()` function:
 
@@
 
-Only the quadratic term for target share shows a significant association.
+Only the quadratic term for target share (not height or weight) shows a significant association.
+
+Here are the standardized coefficients:
+
+```{r}
+QuantPsyc::lm.beta(quadraticTermsRegressionModel)
+```
+
+```{r}
+quadraticTermsRegressionModelStandardized <- lm(
+  scale(fantasyPoints) ~ scale(age) + scale(I(age^2)) + scale(height) + scale(I(height^2)) + scale(weight) + scale(I(weight^2)) + scale(target_share) + scale(I(target_share^2)),
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  na.action = "na.exclude"
+)
+
+summary(quadraticTermsRegressionModelStandardized)
+```
+
+### Visualizing Regression Results {#sec-multipleRegressionVisualizingResults}
+
+#### Regression Coefficients {#sec-multipleRegressionVisualizingRegressionCoefficients}
+
+To visualize the regression coefficients, we can use the `tidy()` function of the `broom` package.
+
+```{r}
+quadraticTermsRegressionModel_tidy <- tidy(
+  quadraticTermsRegressionModel,
+  conf.int = TRUE)
+
+quadraticTermsRegressionModelStandardized_tidy <- tidy(
+  quadraticTermsRegressionModelStandardized,
+  conf.int = TRUE)
+```
+
+```{r}
+#| label: fig-regressionCoefficients
+#| fig-cap: "Regression Coefficients with 95% Confidence Interval."
+#| fig-alt: "Regression Coefficients with 95% Confidence Interval."
+
+ggplot(
+  quadraticTermsRegressionModel_tidy,
+  aes(
+    x = term,
+    y = estimate)) +
+  geom_point() +
+  geom_errorbar(
+    aes(
+      ymin = conf.low,
+      ymax = conf.high),
+    width = 0.2) +
+  coord_flip() + # flip axes for readability
+  labs(
+    title = "Regression Coefficients with 95% CI",
+    x = "Predictor",
+    y = "Coefficient Estimate") +
+  theme_minimal()
+```
+
+```{r}
+#| label: fig-regressionCoefficientsStandardized
+#| fig-cap: "Standardized Regression Coefficients with 95% Confidence Interval."
+#| fig-alt: "Standardized Regression Coefficients with 95% Confidence Interval."
+
+ggplot(
+  quadraticTermsRegressionModelStandardized_tidy,
+  aes(
+    x = term,
+    y = estimate)) +
+  geom_point() +
+  geom_errorbar(
+    aes(
+      ymin = conf.low,
+      ymax = conf.high),
+    width = 0.2) +
+  coord_flip() + # flip axes for readability
+  labs(
+    title = "Regression Coefficients with 95% CI",
+    x = "Predictor",
+    y = "Standardized Coefficient Estimate") +
+  theme_minimal()
+```
+
+#### Model-Implied Association {#sec-multipleRegressionVisualizingModelImpliedAssociation}
+
 If we wanted to visualize the shape of the model-implied association between target share and fantasy points, we could generate the model-implied predictions using the data range that we want to visualize.

### 2025-07-01 index.qmd
20250701 - visualize regression coefficients

--- 2025-06-18 index.qmd
@@
   "petersenlab", "ffanalytics",
-  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
-  "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", "gsisdecoder", 
-  "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", 
-  "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
-  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
-  "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "sn", "mclust", "magrittr", 
-  "viridis", "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", #"webshot2", 
-  "stringi", "parallel", "parallelly", "doParallel", "missRanger", "ggridges", 
-  "powerjoin", "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", 
-  "correlation", "mgcv", "rms", "car", "performance", "lavaan", "mice", 
-  "miceadds", "interactions", "shiny", "PearsonDS", "robustbase", "data.table", 
-  "future", "future.apply", "SimDesign"
+  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "broom", "broom.mixed", 
+  "downlit", "xml2", "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", 
+  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
+  "dagitty", "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", 
+  "patchwork", "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", 
+  "sjstats", "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", 
+  "fitdistrplus", "sn", "mclust", "magrittr", "viridis", "plotly", "quantmod", 
+  "fPortfolio", "NMOF", "xts", "zoo", "forecast", "stringi", "parallel", #"webshot2"
+  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", "caret", 
+  "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation", "mgcv", 
+  "rms", "car", "performance", "lavaan", "mice", "miceadds", "interactions", 
+  "shiny", "PearsonDS", "robustbase", "data.table", "future", "future.apply", 
+  "QuantPsyc", "SimDesign"
 )

### 2025-07-01 mixed-models.qmd
20250701 - visualize regression coefficients

--- 2025-06-08 mixed-models.qmd
@@
 library("parallelly")
+library("broom.mixed")
+library("tidybayes")
 library("plotly")
@@
 
+##### Regression Coefficients {#sec-fantasyPointsByAgeExperienceModelsRegressionCoefficients}
+
+To visualize the regression coefficients, we can use the `tidy()` function of the `broom.mixed` package, as in @fig-individualFantasyPointsRandomLinearFixedQuadraticSlopesInteractionCoefficients.
+
+```{r}
+#| label: fig-individualFantasyPointsRandomLinearFixedQuadraticSlopesInteractionCoefficients
+#| fig-cap: "Regression Coefficients with 95% Confidence Interval from a Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position.
+Overlaid with the Model-Implied Trajectory by Position."
+#| fig-alt: "Regression Coefficients with 95% Confidence Interval from a Mixed Model With Random Intercepts, Random Linear Slopes, and Fixed Quadratic Slopes in Interaction With Position.
+Overlaid with the Model-Implied Trajectory by Position."
+#| code-fold: true
+
+pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience_tidy <- broom.mixed::tidy(
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience,
+  effects = "fixed",
+  conf.int = TRUE)
+
+ggplot(
+  pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience_tidy,
+  aes(
+    x = term,
+    y = estimate)) +
+  geom_point() +
+  geom_errorbar(
+    aes(
+      ymin = conf.low,
+      ymax = conf.high),
+    width = 0.2) +
+  coord_flip() + # flip axes for readability
+  labs(
+    title = "Regression Coefficients with 95% CI",
+    y = "Coefficient Estimate",
+    x = "") +
+  theme_minimal()
+```
+
 ##### Plot of Residuals {#sec-fantasyPointsByAgeExperienceModelsResidualPlot}
@@
 
+#### Plot of Regression Coefficients {#sec-fantasyPointsByAgeExperienceModelsBayesianRegressionCoefficients}
+
+To visualize the regression coefficients, we can use the `tidy()` function of the `broom.mixed` package, as in @fig-bayesianRegressionCoefficients.
+
+```{r}
+#| label: fig-bayesianRegressionCoefficients
+#| fig-cap: "Regression Coefficients with 95% Credible Interval from Bayesian Mixed Model."
+#| fig-alt: "Regression Coefficients with 95% Credible Interval from Bayesian Mixed Model."
+
+bayesianMixedModelFit_tidy <- broom.mixed::tidy(
+  bayesianMixedModelFit,
+  effects = "fixed",
+  conf.int = TRUE)
+
+ggplot(
+  bayesianMixedModelFit_tidy,
+  aes(
+    x = term,
+    y = estimate)) +
+  geom_point() +
+  geom_errorbar(
+    aes(
+      ymin = conf.low,
+      ymax = conf.high),
+    width = 0.2) +
+  coord_flip() + # flip axes for readability
+  labs(
+    title = "Posterior Distributions and 95% Credible Intervals",
+    x = "Predictor",
+    y = "Posterior Estimate") +
+  theme_minimal()
+```
+
+To visualize the distribution of regression coefficients, we can use the `tidybayes` package, as in @fig-bayesianRegressionCoefficientsDistributions.
+
+```{r}
+#| label: fig-bayesianRegressionCoefficientsDistributions
+#| fig-cap: "Distributions of Regression Coefficients with 66% (Thin Line) and 95% (Thin Line) Credible Intervals from Bayesian Mixed Model."
+#| fig-alt: "Distributions of Regression Coefficients with 66% (Thin Line) and 95% (Thin Line) Credible Intervals from Bayesian Mixed Model."
+
+bayesianMixedModelFit %>%
+  tidybayes::tidy_draws() %>%
+  select(starts_with("b_")) %>%
+  pivot_longer(
+    everything(),
+    names_to = "term",
+    values_to = "estimate") %>%
+  mutate(term = sub("^b_", "", term)) %>%
+  ggplot(
+    aes(
+      x = estimate,
+      y = term)) +
+  stat_halfeye() + # by default, .width shows the 66% (thick line) and 95% (thin line) credible intervals
+  labs(
+    title = "Posterior Distributions of Regression Coefficients with 66% and 95% Credible Intervals",
+    x = "Predictor",
+    y = "Posterior Estimate") +
+  theme_minimal()
+```
+
 ### Plots of Model-Implied Fantasy Points by Position and Age {#sec-fantasyPointsByAgePlots}

### 2025-07-01 multiple-regression.qmd
20250701 - visualize regression coefficients

@@
 
-To visualize the regression coefficients, we can use the `tidy()` function of the `broom` package.
-
-```{r}
-quadraticTermsRegressionModel_tidy <- tidy(
+To visualize the regression coefficients, we can use the `tidy()` function of the `broom` package, as in Figures [-@fig-regressionCoefficients] and [-@fig-regressionCoefficientsStandardized].
+
+```{r}
+quadraticTermsRegressionModel_tidy <- broom::tidy(
   quadraticTermsRegressionModel,
@@
 
-quadraticTermsRegressionModelStandardized_tidy <- tidy(
+quadraticTermsRegressionModelStandardized_tidy <- broom::tidy(
   quadraticTermsRegressionModelStandardized,

### 2025-07-01 mixed-models.qmd
20250701 - citations

@@
 
-To visualize the regression coefficients, we can use the `tidy()` function of the `broom.mixed` package, as in @fig-individualFantasyPointsRandomLinearFixedQuadraticSlopesInteractionCoefficients.
+To visualize the regression coefficients, we can use the `tidy()` function of the `broom.mixed` package [@R-broom.mixed], as in @fig-individualFantasyPointsRandomLinearFixedQuadraticSlopesInteractionCoefficients.
 
@@
 
-To visualize the regression coefficients, we can use the `tidy()` function of the `broom.mixed` package, as in @fig-bayesianRegressionCoefficients.
+To visualize the regression coefficients, we can use the `tidy()` function of the `broom.mixed` package [@R-broom.mixed], as in @fig-bayesianRegressionCoefficients.
 
@@
 
-To visualize the distribution of regression coefficients, we can use the `tidybayes` package, as in @fig-bayesianRegressionCoefficientsDistributions.
+To visualize the distribution of regression coefficients, we can use the `tidybayes` package [@R-tidybayes], as in @fig-bayesianRegressionCoefficientsDistributions.
 

### 2025-07-01 multiple-regression.qmd
20250701 - citations

@@
 
-If we want to obtain standardized regression coefficients, we can use the `lm.beta()` function of the `QuantPsyc` package.
+If we want to obtain standardized regression coefficients, we can use the `lm.beta()` function of the `QuantPsyc` package [R-QuantPsyc].
 
@@
 
-To visualize the regression coefficients, we can use the `tidy()` function of the `broom` package, as in Figures [-@fig-regressionCoefficients] and [-@fig-regressionCoefficientsStandardized].
+To visualize the regression coefficients, we can use the `tidy()` function of the `broom` package [@R-broom], as in Figures [-@fig-regressionCoefficients] and [-@fig-regressionCoefficientsStandardized].
 

### 2025-07-01 simulation.qmd
20250701 - citations

@@
 
-You can generate a template for Monte Carlo simulations in the `SimDesign` [@R-SimDesign; @Chalmers2020] package using the following code:
+You can generate a template for Monte Carlo simulations in the `SimDesign` [@R-SimDesign; @Chalmers2020_packages] package using the following code:
 
@@
 #### Optimal Distribution for Each Player {#sec-simulationProjectedPointsMonteCarloOptimalDistribution}
+
+For each player, we identify the optimal distribution as either a normal distribution , or as a skew-normal distribution.
+The normal distribution was fit using the `fitdist()` function of the `fitdistrplus` package [@R-fitdistrplus; @DelignetteMuller2015_packages].
+The skew-normal distribution was fit using the `selm()` function of the `sn` package [@R-sn; @Azzalini2023_packages].
 

### 2025-07-01 mixed-models.qmd
20250701 - standardized regression coefficients

@@
 ```{r}
+library("petersenlab")
 library("lme4")
@@
 summary(pointsPerSeason_nullModel)
+
 summary(pointsPerSeason_nullModel)$r.squared
@@
 
+To obtain standardized regression coefficients, we use the `lm.beta.lmer()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab].
 We use the `emmeans` package [@R-emmeans] to determine the model-implied number of points per season by position.
@@
 summary(pointsPerSeason_position)
+petersenlab::lm.beta.lmer(pointsPerSeason_position)
+
 MuMIn::r.squaredGLMM(pointsPerSeason_position)
@@
 summary(pointsPerSeason_positionAgeFixedLinearSlopes)
+petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeFixedLinearSlopes)
+
 MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
@@
 summary(pointsPerSeason_positionAgeRandomLinearSlopes)
+petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeFixedLinearSlopes)
+
 MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
@@
 summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+
 MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
@@
 summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+
 MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
@@
 summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+
 MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
@@
 summary(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+petersenlab::lm.beta.lmer(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+
 MuMIn::r.squaredGLMM(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)

### 2025-07-01 index.qmd
20250701 - standardized regression coefficients

--- 2025-06-20 index.qmd
@@
   "petersenlab", "ffanalytics",
-  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
-  "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", "gsisdecoder", 
-  "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", 
-  "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
-  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
-  "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "sn", "mclust", 
-  "magrittr", "viridis", "plotly", "quantmod", "fPortfolio", "NMOF", "xts", #"webshot2",
-  "zoo", "forecast", "stringi", "parallel", "parallelly", "doParallel", 
-  "missRanger", "ggridges", "powerjoin", "caret", "gpboost", "cmdstanr", 
-  "learnr", "remotes", "corrplot", "correlation", "mgcv", "rms", "car", 
-  "performance", "lavaan", "mice", "miceadds", "interactions", "shiny", 
-  "PearsonDS", "robustbase", "data.table", "future", "future.apply", "SimDesign"
+  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "broom", "broom.mixed", 
+  "downlit", "xml2", "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", 
+  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
+  "dagitty", "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", 
+  "patchwork", "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", 
+  "sjstats", "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", 
+  "fitdistrplus", "sn", "mclust", "magrittr", "viridis", "plotly", "quantmod", 
+  "fPortfolio", "NMOF", "xts", "zoo", "forecast", "stringi", "parallel", #"webshot2"
+  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", "caret", 
+  "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation", "mgcv", 
+  "rms", "car", "performance", "lavaan", "mice", "miceadds", "interactions", 
+  "shiny", "PearsonDS", "robustbase", "data.table", "future", "future.apply", 
+  "QuantPsyc", "SimDesign", "effectsize"
 )

### 2025-07-01 mixed-models.qmd
20250701 - standardized regression coefficients

@@
 ```{r}
-library("petersenlab")
 library("lme4")
 library("lmerTest")
+library("effectsize")
 library("MuMIn")
@@
 summary(pointsPerSeason_nullModel)
+print(effectsize::standardize_parameters(pointsPerSeason_nullModel, method = "refit"), digits = 2)
 
@@
 summary(pointsPerSeason_linearRegression)
+print(effectsize::standardize_parameters(pointsPerSeason_linearRegression, method = "refit"), digits = 2)
+
 summary(pointsPerSeason_linearRegression)$r.squared
@@
 summary(pointsPerSeason_quadraticRegression)
+print(effectsize::standardize_parameters(pointsPerSeason_linearRegression, method = "refit"), digits = 2)
+
 summary(pointsPerSeason_quadraticRegression)$r.squared
@@
 summary(pointsPerSeason_randomIntercepts)
+print(effectsize::standardize_parameters(pointsPerSeason_randomIntercepts, method = "refit"), digits = 2)
+
 MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
@@
 
-To obtain standardized regression coefficients, we use the `lm.beta.lmer()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab].
+To obtain standardized regression coefficients, we use the `standardize_parameters()` function of the `effectsize` package [@R-effectsize; @BenShachar2020_packages].
 We use the `emmeans` package [@R-emmeans] to determine the model-implied number of points per season by position.
@@
 summary(pointsPerSeason_position)
-petersenlab::lm.beta.lmer(pointsPerSeason_position)
+print(effectsize::standardize_parameters(pointsPerSeason_position, method = "refit"), digits = 2)
 
@@
 summary(pointsPerSeason_positionAgeFixedLinearSlopes)
-petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeFixedLinearSlopes)
+print(effectsize::standardize_parameters(pointsPerSeason_positionAgeFixedLinearSlopes, method = "refit"), digits = 2)
 
@@
 summary(pointsPerSeason_positionAgeRandomLinearSlopes)
-petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeFixedLinearSlopes)
+print(effectsize::standardize_parameters(pointsPerSeason_positionAgeRandomLinearSlopes, method = "refit"), digits = 2)
 
@@
 summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
-petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+print(effectsize::standardize_parameters(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, method = "refit"), digits = 2)
 
@@
 summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
-petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+print(effectsize::standardize_parameters(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, method = "refit"), digits = 2)
 
@@
 summary(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
-petersenlab::lm.beta.lmer(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+print(effectsize::standardize_parameters(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, method = "refit"), digits = 2)
 

### 2025-07-01 multiple-regression.qmd
20250701 - standardized regression coefficients

@@
 library("robustbase")
-library("QuantPsyc")
 library("broom")
@@
 
-If we want to obtain standardized regression coefficients, we can use the `lm.beta()` function of the `QuantPsyc` package [R-QuantPsyc].
-
-```{r}
-QuantPsyc::lm.beta(linearRegressionModel)
+If we want to obtain standardized regression coefficients, we can use the `standardize_parameters()` function of the `effectsize` package [@R-effectsize; @BenShachar2020_packages].
+
+```{r}
+print(effectsize::standardize_parameters(linearRegressionModel, method = "basic"), digits = 2)
 ```
@@
 summary(linearRegressionModelStandardized)
+
+print(effectsize::standardize_parameters(linearRegressionModel, method = "refit"), digits = 2)
 ```
@@
 ```{r}
-QuantPsyc::lm.beta(quadraticTermsRegressionModel)
+print(effectsize::standardize_parameters(quadraticTermsRegressionModel, method = "basic"), digits = 2)
 ```
@@
 summary(quadraticTermsRegressionModelStandardized)
+
+print(effectsize::standardize_parameters(quadraticTermsRegressionModel, method = "refit"), digits = 2)
 ```

### 2025-07-01 machine-learning.qmd
20250701 - heading
20250701 - tweaks

@@
 ```
+
+### Create New Data Object for Merging with Later Predictions {#sec-newDataObject}
 
@@
   categorical_feature = pred_vars_qb_categorical,
-  label = data_train_qb_matrix[,"fantasyPoints_lag"]) #use mean-centered variable and add mean back afterward
+  label = data_train_qb_matrix[,"fantasyPoints_lag"]) # could use mean-centered variable and add mean back afterward
 

### 2025-07-01 multiple-regression.qmd
20250701 - tweaks

@@
   labs(
-    title = "Regression Coefficients with 95% CI",
+    title = "Standardized Regression Coefficients with 95% CI",
     x = "Predictor",

### 2025-07-01 machine-learning.qmd
20250701 - clarify

@@
   categorical_feature = pred_vars_qb_categorical,
-  label = data_train_qb_matrix[,"fantasyPoints_lag"]) # could use mean-centered variable and add mean back afterward
+  label = data_train_qb_matrix[,"fantasyPoints_lag"]) # could instead use mean-centered variable (fantasyPointsMC_lag) and add mean back afterward
 

### 2025-07-02 index.qmd
20250701 - citations & longitudinal random forest model

--- 2025-07-01 index.qmd
@@
   "petersenlab", "ffanalytics",
-  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "broom", "downlit", 
-  "xml2", "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", 
+  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "broom", "broom.mixed", 
+  "downlit", "xml2", "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", 
   "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
@@
   "patchwork", "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", 
-  "sjstats", "AICcmodavg", "rstan", "brms", "bbmle", "fitdistrplus", "sn", 
-  "mclust", "magrittr", "viridis", "plotly", "quantmod", "fPortfolio", "NMOF", 
-  "xts", "zoo", "forecast", "stringi", "parallel", "parallelly", "doParallel", #"webshot2"
-  "missRanger", "ggridges", "powerjoin", "caret", "gpboost", "cmdstanr", "learnr", 
-  "remotes", "corrplot", "correlation", "mgcv", "rms", "car", "performance", 
-  "lavaan", "mice", "miceadds", "interactions", "shiny", "PearsonDS", 
-  "robustbase", "data.table", "future", "future.apply", "QuantPsyc", "SimDesign"
+  "sjstats", "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", 
+  "fitdistrplus", "sn", "mclust", "magrittr", "viridis", "plotly", "quantmod", 
+  "fPortfolio", "NMOF", "xts", "zoo", "forecast", "stringi", "parallel", #"webshot2"
+  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", "caret", 
+  "tidymodels", "LongituRF", "gpboost", "cmdstanr", "learnr", "remotes", 
+  "corrplot", "correlation", "mgcv", "rms", "car", "performance", "lavaan", 
+  "mice", "miceadds", "interactions", "shiny", "PearsonDS", "robustbase", 
+  "data.table", "future", "future.apply", "QuantPsyc", "SimDesign", "effectsize"
 )

### 2025-07-02 machine-learning.qmd
20250701 - citations & longitudinal random forest model

@@
 library("powerjoin")
-library("caret")
+library("tidymodels")
+library("LongituRF")
+#library("caret")
 library("gpboost")
@@
 #| label: cross-sectional-machine-learning
+#| eval: false
 
@@
 
-[@Hu2023]
-
-```{r}
-#| eval: false
-
-library("LongituRF")
-
+Approaches to estimating random forest models with longitudinal data are described by @@Hu2023.
+Below, we fit longitudinal random forest models using the `MERF()` function of the `LongituRF` package [@R-LongituRF].
+
+```{r}
 smerf <- LongituRF::MERF(
-  X = seasonalData_subsetQB_imp$ximp[,c("passing_epa")] %>% as.matrix(),
-  Y = seasonalData_subsetQB$fantasyPoints_lag,
-  Z = seasonalData_subsetQB_imp$ximp[,c("pacr")] %>% as.matrix(),
-  id = seasonalData_subsetQB$gsis_id,
-  time = seasonalData_subsetQB_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
+  X = data_train_qb_matrix %>% as_tibble() %>% dplyr::select(season:att_br.rush) %>% as.matrix(),
+  Y = data_train_qb_matrix[,c("fantasyPoints_lag")] %>% as.matrix(),
+  Z = data_train_qb_matrix[,c("pacr")] %>% as.matrix(),
+  id = data_train_qb_matrix[,c("gsis_id")] %>% as.matrix(),
+  time = data_train_qb_matrix[,c("ageCentered20")] %>% as.matrix(),
   ntree = 500,
@@
 smerf$omega # the predicted stochastic processes
-plot(smerf$Vraisemblance) # evolution of the log-likelihood
 smerf$OOB # OOB error at each iteration
+```
+
+```{r}
+#| label: fig-randomForestLongitudinalLogLikelihood
+#| fig-cap: "Evolution of the Log-Likelihood."
+#| fig-alt: "Evolution of the Log-Likelihood."
+
+plot(smerf$Vraisemblance)
 ```

### 2025-07-02 machine-learning.qmd
20250702 - machine learning

@@
 library("parallel")
-library("doParallel")
+library("future")
 library("missRanger")
@@
 library("LongituRF")
-#library("caret")
 library("gpboost")
@@
 
-num_cores <- detectCores() - 1
+num_cores <- parallel::detectCores() - 1
 num_true_cores <- parallel::detectCores(logical = FALSE) - 1
@@
 
+We use the `future` [@R-future] package for parallel (faster) processing.
+
+```{r}
+future::plan(future::multisession, workers = num_cores)
+```
+
 ## Fitting the Traditional Regression Models {#sec-fittingModels-regression}
@@
 
+```{r}
+# Set seed for reproducibility
+set.seed(52242)
+
+# Set up cross-validation
+folds <- rsample::group_vfold_cv(
+  data_train_qb,
+  group = gsis_id,
+  v = 10) # 10-fold cross-validation
+
+# Define Recipe (Formula)
+rec <- recipes::recipe(
+  fantasyPoints_lag ~ ageCentered20,
+  data = data_train_qb)
+
+# Define Model
+lm_spec <- parsnip::linear_reg() %>%
+  parsnip::set_engine("lm") %>%
+  parsnip::set_mode("regression")
+
+# Workflow
+lm_wf <- workflows::workflow() %>%
+  workflows::add_recipe(rec) %>%
+  workflows::add_model(lm_spec)
+
+# Fit Model with Cross-Validation
+cv_results <- tune::fit_resamples(
+  lm_wf,
+  resamples = folds,
+  metrics = metric_set(rmse, mae, rsq),
+  control = control_resamples(save_pred = TRUE)
+)
+
+# View Cross-Validation metrics
+tune::collect_metrics(cv_results)
+
+# Fit Final Model on Training Data
+final_model <- workflows::fit(
+  lm_wf,
+  data = data_train_qb)
+
+# View Coefficients
+final_model %>% 
+  workflows::extract_fit_parsnip() %>% 
+  broom::tidy()
+
+# Predict on Test Data
+predict(final_model, data_test_qb)
+
+df <- data_test_qb %>%
+  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)
+
+# Evaluate Accuracy of Predictions
+petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-regressionOnePredictor
+#| fig-cap: "Predicted Versus Actual Fantasy Points for Regression Model with One Predictor (Player Age)."
+#| fig-alt: "Predicted Versus Actual Fantasy Points for Regression Model with One Predictor (Player Age)."
+
+# Calculate combined range for axes
+axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)
+
+ggplot(
+  df,
+  aes(
+    x = pred,
+    y = fantasyPoints_lag)) +
+  geom_point(
+    size = 2,
+    alpha = 0.6) +
+  geom_abline(
+    slope = 1,
+    intercept = 0,
+    color = "blue",
+    linetype = "dashed") +
+  coord_equal(
+    xlim = axis_limits,
+    ylim = axis_limits) +
+  labs(
+    title = "Predicted vs Actual Fantasy Points (Test Data)",
+    x = "Predicted Fantasy Points",
+    y = "Actual Fantasy Points"
+  ) +
+  theme_classic()
+```
+
+```{r}
+newData_qb %>%
+  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
+  left_join(
+    .,
+    nfl_playerIDs %>% select(gsis_id, name),
+    by = "gsis_id"
+  ) %>% 
+  select(name, fantasyPoints_lag) %>% 
+  arrange(-fantasyPoints_lag)
+```
+
 ### Regression with Multiple Predictors {#sec-regressionMultiplePredictors}
 
+```{r}
+# Set seed for reproducibility
+set.seed(52242)
+
+# Set up cross-validation
+folds <- rsample::group_vfold_cv(
+  data_train_qb,
+  group = gsis_id,
+  v = 10) # 10-fold cross-validation
+
+# Define Recipe (Formula)
+rec <- recipes::recipe(
+  fantasyPoints_lag ~ .,
+  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))
+
+# Define Model
+lm_spec <- parsnip::linear_reg() %>%
+  parsnip::set_engine("lm") %>%
+  parsnip::set_mode("regression")
+
+# Workflow
+lm_wf <- workflows::workflow() %>%
+  workflows::add_recipe(rec) %>%
+  workflows::add_model(lm_spec)
+
+# Fit Model with Cross-Validation
+cv_results <- tune::fit_resamples(
+  lm_wf,
+  resamples = folds,
+  metrics = metric_set(rmse, mae, rsq),
+  control = control_resamples(save_pred = TRUE)
+)
+
+# View Cross-Validation metrics
+tune::collect_metrics(cv_results)
+
+# Fit Final Model on Training Data
+final_model <- workflows::fit(
+  lm_wf,
+  data = data_train_qb)
+
+# View Coefficients
+final_model %>% 
+  workflows::extract_fit_parsnip() %>% 
+  broom::tidy()
+
+# Predict on Test Data
+predict(final_model, data_test_qb)
+
+df <- data_test_qb %>%
+  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)
+
+# Evaluate Accuracy of Predictions
+petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-regressionMultiplePredictors
+#| fig-cap: "Predicted Versus Actual Fantasy Points for Regression Model with Multiple Predictors."
+#| fig-alt: "Predicted Versus Actual Fantasy Points for Regression Model with Multiple Predictors."
+
+# Calculate combined range for axes
+axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)
+
+ggplot(
+  df,
+  aes(
+    x = pred,
+    y = fantasyPoints_lag)) +
+  geom_point(
+    size = 2,
+    alpha = 0.6) +
+  geom_abline(
+    slope = 1,
+    intercept = 0,
+    color = "blue",
+    linetype = "dashed") +
+  coord_equal(
+    xlim = axis_limits,
+    ylim = axis_limits) +
+  labs(
+    title = "Predicted vs Actual Fantasy Points (Test Data)",
+    x = "Predicted Fantasy Points",
+    y = "Actual Fantasy Points"
+  ) +
+  theme_classic()
+```
+
+```{r}
+newData_qb %>%
+  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
+  left_join(
+    .,
+    nfl_playerIDs %>% select(gsis_id, name),
+    by = "gsis_id"
+  ) %>% 
+  select(name, fantasyPoints_lag) %>% 
+  arrange(-fantasyPoints_lag)
+```
+
 ## Fitting the Machine Learning Models {#sec-fittingModels-machineLearning}
@@
 
+```{r}
+# Set seed for reproducibility
+set.seed(52242)
+
+# Set up cross-validation
+folds <- rsample::group_vfold_cv(
+  data_train_qb,
+  group = gsis_id,
+  v = 10) # 10-fold cross-validation
+
+# Define Recipe (Formula)
+rec <- recipes::recipe(
+  fantasyPoints_lag ~ .,
+  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))
+
+# Define Model
+lasso_spec <- 
+  parsnip::linear_reg(
+    penalty = tune(),
+    mixture = 1) %>%
+  set_engine("glmnet")
+
+# Workflow
+lasso_wf <- workflows::workflow() %>%
+  workflows::add_recipe(rec) %>%
+  workflows::add_model(lasso_spec)
+
+# Define grid of penalties to try (log scale is typical)
+penalty_grid <- dials::grid_regular(
+  dials::penalty(range = c(-4, -1)),
+  levels = 20)
+
+# Tune the Penalty Parameter
+cv_results <- tune::tune_grid(
+  lasso_wf,
+  resamples = folds,
+  grid = penalty_grid,
+  metrics = metric_set(rmse, mae, rsq),
+  control = control_grid(save_pred = TRUE)
+)
+
+# View Cross-Validation metrics
+tune::collect_metrics(cv_results)
+
+# Identify best penalty
+select_best(cv_results, metric = "rmse")
+select_best(cv_results, metric = "mae")
+select_best(cv_results, metric = "rsq")
+
+best_penalty <- select_best(cv_results, metric = "mae")
+
+# Finalize Workflow with Best Penalty
+final_wf <- finalize_workflow(
+  lasso_wf,
+  best_penalty)
+
+# Fit Final Model on Training Data
+final_model <- workflows::fit(
+  final_wf,
+  data = data_train_qb)
+
+# View Coefficients
+final_model %>% 
+  workflows::extract_fit_parsnip() %>% 
+  broom::tidy()
+
+# Predict on Test Data
+predict(final_model, data_test_qb)
+
+df <- data_test_qb %>%
+  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)
+
+# Evaluate Accuracy of Predictions
+petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-lassoModel
+#| fig-cap: "Predicted Versus Actual Fantasy Points for Least Absolute Shrinkage and Selection Option (LASSO) Model."
+#| fig-alt: "Predicted Versus Actual Fantasy Points for Least Absolute Shrinkage and Selection Option (LASSO) Model."
+
+# Calculate combined range for axes
+axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)
+
+ggplot(
+  df,
+  aes(
+    x = pred,
+    y = fantasyPoints_lag)) +
+  geom_point(
+    size = 2,
+    alpha = 0.6) +
+  geom_abline(
+    slope = 1,
+    intercept = 0,
+    color = "blue",
+    linetype = "dashed") +
+  coord_equal(
+    xlim = axis_limits,
+    ylim = axis_limits) +
+  labs(
+    title = "Predicted vs Actual Fantasy Points (Test Data)",
+    x = "Predicted Fantasy Points",
+    y = "Actual Fantasy Points"
+  ) +
+  theme_classic()
+```
+
+```{r}
+newData_qb %>%
+  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
+  left_join(
+    .,
+    nfl_playerIDs %>% select(gsis_id, name),
+    by = "gsis_id"
+  ) %>% 
+  select(name, fantasyPoints_lag) %>% 
+  arrange(-fantasyPoints_lag)
+```
+
 ### Ridge Regression {#sec-ridgeRegression}
 
+```{r}
+# Set seed for reproducibility
+set.seed(52242)
+
+# Set up cross-validation
+folds <- rsample::group_vfold_cv(
+  data_train_qb,
+  group = gsis_id,
+  v = 10) # 10-fold cross-validation
+
+# Define Recipe (Formula)
+rec <- recipes::recipe(
+  fantasyPoints_lag ~ .,
+  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))
+
+# Define Model
+ridge_spec <- 
+  linear_reg(
+    penalty = tune(),
+    mixture = 0) %>%
+  set_engine("glmnet")
+
+# Workflow
+ridge_wf <- workflows::workflow() %>%
+  workflows::add_recipe(rec) %>%
+  workflows::add_model(ridge_spec)
+
+# Define grid of penalties to try (log scale is typical)
+penalty_grid <- dials::grid_regular(
+  dials::penalty(range = c(-4, -1)),
+  levels = 20)
+
+# Tune the Penalty Parameter
+cv_results <- tune::tune_grid(
+  ridge_wf,
+  resamples = folds,
+  grid = penalty_grid,
+  metrics = metric_set(rmse, mae, rsq),
+  control = control_grid(save_pred = TRUE)
+)
+
+# View Cross-Validation metrics
+tune::collect_metrics(cv_results)
+
+# Identify best penalty
+select_best(cv_results, metric = "rmse")
+select_best(cv_results, metric = "mae")
+select_best(cv_results, metric = "rsq")
+
+best_penalty <- select_best(cv_results, metric = "mae")
+
+# Finalize Workflow with Best Penalty
+final_wf <- finalize_workflow(
+  ridge_wf,
+  best_penalty)
+
+# Fit Final Model on Training Data
+final_model <- workflows::fit(
+  final_wf,
+  data = data_train_qb)
+
+# View Coefficients
+final_model %>% 
+  workflows::extract_fit_parsnip() %>% 
+  broom::tidy()
+
+# Predict on Test Data
+predict(final_model, data_test_qb)
+
+df <- data_test_qb %>%
+  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)
+
+# Evaluate Accuracy of Predictions
+petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-ridgeRegressionModel
+#| fig-cap: "Predicted Versus Actual Fantasy Points for Ridge Regression Model."
+#| fig-alt: "Predicted Versus Actual Fantasy Points for Ridge Regression Model."
+
+# Calculate combined range for axes
+axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)
+
+ggplot(
+  df,
+  aes(
+    x = pred,
+    y = fantasyPoints_lag)) +
+  geom_point(
+    size = 2,
+    alpha = 0.6) +
+  geom_abline(
+    slope = 1,
+    intercept = 0,
+    color = "blue",
+    linetype = "dashed") +
+  coord_equal(
+    xlim = axis_limits,
+    ylim = axis_limits) +
+  labs(
+    title = "Predicted vs Actual Fantasy Points (Test Data)",
+    x = "Predicted Fantasy Points",
+    y = "Actual Fantasy Points"
+  ) +
+  theme_classic()
+```
+
+```{r}
+newData_qb %>%
+  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
+  left_join(
+    .,
+    nfl_playerIDs %>% select(gsis_id, name),
+    by = "gsis_id"
+  ) %>% 
+  select(name, fantasyPoints_lag) %>% 
+  arrange(-fantasyPoints_lag)
+```
+
 ### Elastic Net {#sec-elasticNet}
 
+```{r}
+# Set seed for reproducibility
+set.seed(52242)
+
+# Set up cross-validation
+folds <- rsample::group_vfold_cv(
+  data_train_qb,
+  group = gsis_id,
+  v = 10) # 10-fold cross-validation
+
+# Define Recipe (Formula)
+rec <- recipes::recipe(
+  fantasyPoints_lag ~ .,
+  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))
+
+# Define Model
+enet_spec <- 
+  linear_reg(
+    penalty = tune(),
+    mixture = tune()) %>%
+  set_engine("glmnet")
+
+# Workflow
+enet_wf <- workflows::workflow() %>%
+  workflows::add_recipe(rec) %>%
+  workflows::add_model(enet_spec)
+
+# Define a regular grid for both penalty and mixture
+grid_enet <- dials::grid_regular(
+  dials::penalty(range = c(-4, -1)),
+  dials::mixture(range = c(0, 1)),
+  levels = c(20, 5) # 20 penalty values × 5 mixture values
+)
+
+# Tune the Grid
+cv_results <- tune::tune_grid(
+  enet_wf,
+  resamples = folds,
+  grid = grid_enet,
+  metrics = metric_set(rmse, mae, rsq),
+  control = control_grid(save_pred = TRUE)
+)
+
+# View Cross-Validation metrics
+tune::collect_metrics(cv_results)
+
+# Identify best penalty
+select_best(cv_results, metric = "rmse")
+select_best(cv_results, metric = "mae")
+select_best(cv_results, metric = "rsq")
+
+best_penalty <- select_best(cv_results, metric = "mae")
+
+# Finalize Workflow with Best Penalty
+final_wf <- finalize_workflow(
+  enet_wf,
+  best_penalty)
+
+# Fit Final Model on Training Data
+final_model <- workflows::fit(
+  final_wf,
+  data = data_train_qb)
+
+# View Coefficients
+final_model %>% 
+  workflows::extract_fit_parsnip() %>% 
+  broom::tidy()
+
+# Predict on Test Data
+predict(final_model, data_test_qb)
+
+df <- data_test_qb %>%
+  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)
+
+# Evaluate Accuracy of Predictions
+petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-elasticNetModel
+#| fig-cap: "Predicted Versus Actual Fantasy Points for Elastic Net Model."
+#| fig-alt: "Predicted Versus Actual Fantasy Points for Elastic Net Model."
+
+# Calculate combined range for axes
+axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)
+
+ggplot(
+  df,
+  aes(
+    x = pred,
+    y = fantasyPoints_lag)) +
+  geom_point(
+    size = 2,
+    alpha = 0.6) +
+  geom_abline(
+    slope = 1,
+    intercept = 0,
+    color = "blue",
+    linetype = "dashed") +
+  coord_equal(
+    xlim = axis_limits,
+    ylim = axis_limits) +
+  labs(
+    title = "Predicted vs Actual Fantasy Points (Test Data)",
+    x = "Predicted Fantasy Points",
+    y = "Actual Fantasy Points"
+  ) +
+  theme_classic()
+```
+
+```{r}
+newData_qb %>%
+  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
+  left_join(
+    .,
+    nfl_playerIDs %>% select(gsis_id, name),
+    by = "gsis_id"
+  ) %>% 
+  select(name, fantasyPoints_lag) %>% 
+  arrange(-fantasyPoints_lag)
+```
+
 ### Random Forest Machine Learning {#sec-randomForest}
@@
 
-We use the `caret` package [@R-caret].
-We use the `doParallel` [@R-doParallel] package for parallel (faster) processing.
-
-```{r}
-#| label: cross-sectional-machine-learning
-#| eval: false
-
-cl <- parallel::makeCluster(num_cores)
-doParallel::registerDoParallel(cl)
-
+```{r}
+# Set seed for reproducibility
 set.seed(52242)
 
-randomForest_qb <- caret::train(
-  fantasyPoints_lag ~ ., # use all predictors
-  data = seasonalData_lag_qb_train_imp$data,
-  method = "rf",
-  trControl = trainControl(
-    method = "cv",
-    number = 10)) # 10-fold cross-validation
-
-randomForest_rb <- caret::train(
-  fantasyPoints_lag ~ ., # use all predictors
-  data = seasonalData_lag_rb_train_imp$data,
-  method = "rf",
-  trControl = trainControl(
-    method = "cv",
-    number = 10)) # 10-fold cross-validation
-
-randomForest_wr <- caret::train(
-  fantasyPoints_lag ~ ., # use all predictors
-  data = seasonalData_lag_wr_train_imp$data,
-  method = "rf",
-  trControl = trainControl(
-    method = "cv",
-    number = 10)) # 10-fold cross-validation
-
-randomForest_te <- caret::train(
-  fantasyPoints_lag ~ ., # use all predictors
-  data = seasonalData_lag_te_train_imp$data,
-  method = "rf",
-  trControl = trainControl(
-    method = "cv",
-    number = 10)) # 10-fold cross-validation
-
-stopCluster(cl)
-
-print(randomForest_qb)
-print(randomForest_rb)
-print(randomForest_wr)
-print(randomForest_te)
-
-newData_qb$fantasyPoints_lag <- predict(
-  randomForest_qb,
-  newdata = newData_qb
-)
-
-newData_rb$fantasyPoints_lag <- predict(
-  randomForest_rb,
-  newdata = newData_rb
-)
-
-newData_wr$fantasyPoints_lag <- predict(
-  randomForest_wr,
-  newdata = newData_wr
-)
-
-newData_te$fantasyPoints_lag <- predict(
-  randomForest_te,
-  newdata = newData_te
-)
-
-newData_qb$fantasyPoints_lag <- newData_qb$fantasyPoints_lag
-newData_rb$fantasyPoints_lag <- newData_rb$fantasyPoints_lag
-newData_wr$fantasyPoints_lag <- newData_wr$fantasyPoints_lag
-newData_te$fantasyPoints_lag <- newData_te$fantasyPoints_lag
-
-newData_qb <- dplyr::left_join(
-  newData_qb,
-  newData_seasonal %>% select(gsis_id, player_display_name, team),
+# Set up cross-validation
+folds <- rsample::group_vfold_cv(
+  data_train_qb,
+  group = gsis_id,
+  v = 10) # 10-fold cross-validation
+
+# Define Recipe (Formula)
+rec <- recipes::recipe(
+  fantasyPoints_lag ~ .,
+  data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))
+
+# Define Model
+rf_spec <- 
+  parsnip::rand_forest(
+    mtry = tune::tune(),
+    min_n = tune::tune(),
+    trees = 500) %>%
+  parsnip::set_mode("regression") %>%
+  parsnip::set_engine("ranger", importance = "impurity")
+
+# Workflow
+rf_wf <- workflows::workflow() %>%
+  workflows::add_recipe(rec) %>%
+  workflows::add_model(rf_spec)
+
+# Create Grid
+n_predictors <- recipes::prep(rec) %>%
+  recipes::juice() %>%
+  dplyr::select(-fantasyPoints_lag) %>%
+  ncol()
+
+# Dynamically define ranges based on data
+rf_params <- hardhat::extract_parameter_set_dials(rf_spec) %>%
+  dials:::update.parameters(
+    mtry = dials::mtry(range = c(1L, n_predictors)),
+    min_n = dials::min_n(range = c(2L, 10L))
+  )
+
+rf_grid <- dials::grid_random(rf_params, size = 15) #dials::grid_regular(rf_params, levels = 5)
+
+# Tune the Grid
+cv_results <- tune::tune_grid(
+  rf_wf,
+  resamples = folds,
+  grid = rf_grid,
+  metrics = metric_set(rmse, mae, rsq),
+  control = control_grid(save_pred = TRUE)
+)
+
+# View Cross-Validation metrics
+tune::collect_metrics(cv_results)
+
+# Identify best penalty
+select_best(cv_results, metric = "rmse")
+select_best(cv_results, metric = "mae")
+select_best(cv_results, metric = "rsq")
+
+best_penalty <- select_best(cv_results, metric = "mae")
+
+# Finalize Workflow with Best Penalty
+final_wf <- finalize_workflow(
+  rf_wf,
+  best_penalty)
+
+# Fit Final Model on Training Data
+final_model <- workflows::fit(
+  final_wf,
+  data = data_train_qb)
+
+# View Feature Importance
+rf_fit <- final_model %>% 
+  workflows::extract_fit_parsnip()
+
+rf_fit
+
+ranger_obj <- rf_fit$fit
+
+ranger_obj$variable.importance
+
+# Predict on Test Data
+predict(final_model, data_test_qb)
+
+df <- data_test_qb %>%
+  mutate(pred = predict(final_model, new_data = data_test_qb)$.pred)
+
+# Evaluate Accuracy of Predictions
+petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| label: fig-randomForestModel
+#| fig-cap: "Predicted Versus Actual Fantasy Points for Random Forest Model."
+#| fig-alt: "Predicted Versus Actual Fantasy Points for Random Forest Model."
+
+# Calculate combined range for axes
+axis_limits <- range(c(df$pred, df$fantasyPoints_lag), na.rm = TRUE)
+
+ggplot(
+  df,
+  aes(
+    x = pred,
+    y = fantasyPoints_lag)) +
+  geom_point(
+    size = 2,
+    alpha = 0.6) +
+  geom_abline(
+    slope = 1,
+    intercept = 0,
+    color = "blue",
+    linetype = "dashed") +
+  coord_equal(
+    xlim = axis_limits,
+    ylim = axis_limits) +
+  labs(
+    title = "Predicted vs Actual Fantasy Points (Test Data)",
+    x = "Predicted Fantasy Points",
+    y = "Actual Fantasy Points"
+  ) +
+  theme_classic()
+```
+
+```{r}
+newData_qb %>%
+  mutate(fantasyPoints_lag = predict(final_model, new_data = newData_qb)$.pred) %>% 
+  left_join(
+    .,
+    nfl_playerIDs %>% select(gsis_id, name),
     by = "gsis_id"
-)
-
-newData_rb <- dplyr::left_join(
-  newData_rb,
-  newData_seasonal %>% select(gsis_id, player_display_name, team),
-  by = "gsis_id"
-)
-
-newData_wr <- dplyr::left_join(
-  newData_wr,
-  newData_seasonal %>% select(gsis_id, player_display_name, team),
-  by = "gsis_id"
-)
-
-newData_te <- dplyr::left_join(
-  newData_te,
-  newData_seasonal %>% select(gsis_id, player_display_name, team),
-  by = "gsis_id"
-)
-
-newData_qb %>%
-  arrange(-fantasyPoints_lag) %>% 
-  select(gsis_id, player_display_name, fantasyPoints_lag)
-
-newData_rb %>%
-  arrange(-fantasyPoints_lag) %>% 
-  select(gsis_id, player_display_name, fantasyPoints_lag)
-
-newData_wr %>%
-  arrange(-fantasyPoints_lag) %>% 
-  select(gsis_id, player_display_name, fantasyPoints_lag)
-
-newData_te %>%
-  arrange(-fantasyPoints_lag) %>% 
-  select(gsis_id, player_display_name, fantasyPoints_lag)
+  ) %>% 
+  select(name, fantasyPoints_lag) %>% 
+  arrange(-fantasyPoints_lag)
+```
+
+Now we can stop the parallel backend:
+
+```{r}
+future::plan(future::sequential)
+```
+
+```{r}
+#| include: false
+
+gc()
 ```

### 2025-07-02 machine-learning.qmd
20250702 - clean up

@@
 
-# Set up cross-validation
+# Set up Cross-Validation
 folds <- rsample::group_vfold_cv(
   data_train_qb,
-  group = gsis_id,
+  group = gsis_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
   v = 10) # 10-fold cross-validation
@@
   resamples = folds,
-  metrics = metric_set(rmse, mae, rsq),
-  control = control_resamples(save_pred = TRUE)
+  metrics = yardstick::metric_set(rmse, mae, rsq),
+  control = tune::control_resamples(save_pred = TRUE)
 )
@@
 # Predict on Test Data
-predict(final_model, data_test_qb)
-
 df <- data_test_qb %>%
@@
 rec <- recipes::recipe(
-  fantasyPoints_lag ~ .,
+  fantasyPoints_lag ~ ., # use all predictors
   data = data_train_qb %>% select(-gsis_id, -fantasyPointsMC_lag))
@@
   parsnip::linear_reg(
-    penalty = tune(),
+    penalty = tune::tune(),
     mixture = 1) %>%
-  set_engine("glmnet")
+  parsnip::set_engine("glmnet")
 
@@
   grid = penalty_grid,
-  metrics = metric_set(rmse, mae, rsq),
-  control = control_grid(save_pred = TRUE)
+  metrics = yardstick::metric_set(rmse, mae, rsq),
+  control = tune::control_grid(save_pred = TRUE)
 )
@@
 # Identify best penalty
-select_best(cv_results, metric = "rmse")
-select_best(cv_results, metric = "mae")
-select_best(cv_results, metric = "rsq")
-
-best_penalty <- select_best(cv_results, metric = "mae")
+tune::select_best(cv_results, metric = "rmse")
+tune::select_best(cv_results, metric = "mae")
+tune::select_best(cv_results, metric = "rsq")
+
+best_penalty <- tune::select_best(cv_results, metric = "mae")
 
 # Finalize Workflow with Best Penalty
-final_wf <- finalize_workflow(
+final_wf <- tune::finalize_workflow(
   lasso_wf,
@@
 ridge_spec <- 
-  linear_reg(
-    penalty = tune(),
+  parnsnip::linear_reg(
+    penalty = tune::tune(),
     mixture = 0) %>%
-  set_engine("glmnet")
+  parsnip::set_engine("glmnet")
 
@@
 # Identify best penalty
-select_best(cv_results, metric = "rmse")
-select_best(cv_results, metric = "mae")
-select_best(cv_results, metric = "rsq")
-
-best_penalty <- select_best(cv_results, metric = "mae")
+tune::select_best(cv_results, metric = "rmse")
+tune::select_best(cv_results, metric = "mae")
+tune::select_best(cv_results, metric = "rsq")
+
+best_penalty <- tune::select_best(cv_results, metric = "mae")
 
 # Finalize Workflow with Best Penalty
-final_wf <- finalize_workflow(
+final_wf <- tune::finalize_workflow(
   ridge_wf,
@@
 enet_spec <- 
-  linear_reg(
-    penalty = tune(),
-    mixture = tune()) %>%
-  set_engine("glmnet")
+  parsnip::linear_reg(
+    penalty = tune::tune(),
+    mixture = tune::tune()) %>%
+  parsnip::set_engine("glmnet")
 
@@
   grid = grid_enet,
-  metrics = metric_set(rmse, mae, rsq),
-  control = control_grid(save_pred = TRUE)
+  metrics = yardstick::metric_set(rmse, mae, rsq),
+  control = tune::control_grid(save_pred = TRUE)
 )
@@
 # Identify best penalty
-select_best(cv_results, metric = "rmse")
-select_best(cv_results, metric = "mae")
-select_best(cv_results, metric = "rsq")
-
-best_penalty <- select_best(cv_results, metric = "mae")
+tune::select_best(cv_results, metric = "rmse")
+tune::select_best(cv_results, metric = "mae")
+tune::select_best(cv_results, metric = "rsq")
+
+best_penalty <- tune::select_best(cv_results, metric = "mae")
 
 # Finalize Workflow with Best Penalty
-final_wf <- finalize_workflow(
+final_wf <- tune::finalize_workflow(
   enet_wf,
@@
   parsnip::set_mode("regression") %>%
-  parsnip::set_engine("ranger", importance = "impurity")
+  parsnip::set_engine(
+    "ranger",
+    importance = "impurity")
 
@@
   grid = rf_grid,
-  metrics = metric_set(rmse, mae, rsq),
-  control = control_grid(save_pred = TRUE)
+  metrics = yardstick::metric_set(rmse, mae, rsq),
+  control = tune::control_grid(save_pred = TRUE)
 )
@@
 # Identify best penalty
-select_best(cv_results, metric = "rmse")
-select_best(cv_results, metric = "mae")
-select_best(cv_results, metric = "rsq")
-
-best_penalty <- select_best(cv_results, metric = "mae")
+tune::select_best(cv_results, metric = "rmse")
+tune::select_best(cv_results, metric = "mae")
+tune::select_best(cv_results, metric = "rsq")
+
+best_penalty <- tune::select_best(cv_results, metric = "mae")
 
 # Finalize Workflow with Best Penalty
-final_wf <- finalize_workflow(
+final_wf <- tune::finalize_workflow(
   rf_wf,
@@
 smerf$forest # the fitted random forest (obtained at the last iteration)
-smerf$random_effects # the predicted random effects for each player
-smerf$omega # the predicted stochastic processes
+smerf$random_effects %>% data.frame() # the predicted random effects for each player
+smerf$omega %>% data.frame() # the predicted stochastic processes
 smerf$OOB # OOB error at each iteration
+
+# Predict on Test Data
+predict(
+  smerf,
+  X = data_test_qb %>% as_tibble() %>% dplyr::select(season:att_br.rush) %>% as.matrix(),
+  Z = data_test_qb[,c("pacr")] %>% as.matrix(),
+  id = data_test_qb[,c("gsis_id")] %>% as.matrix(),
+  time = data_test_qb[,c("ageCentered20")] %>% as.matrix())
 ```

### 2025-07-02 machine-learning.qmd
20250702 - standardized regression coefficients

@@
 library("gpboost")
+library("effectsize")
 library("tidyverse")
@@
 
+final_model %>%
+  workflows::extract_fit_parsnip() %>% 
+  effectsize::standardize_parameters()
+
 # Predict on Test Data
@@
 ridge_spec <- 
-  parnsnip::linear_reg(
+  parsnip::linear_reg(
     penalty = tune::tune(),

### 2025-07-02 machine-learning.qmd
20250702 - longitudinal random forest

@@
 smerf <- LongituRF::MERF(
-  X = data_train_qb_matrix %>% as_tibble() %>% dplyr::select(season:att_br.rush) %>% as.matrix(),
-  Y = data_train_qb_matrix[,c("fantasyPoints_lag")] %>% as.matrix(),
-  Z = data_train_qb_matrix[,c("pacr")] %>% as.matrix(),
-  id = data_train_qb_matrix[,c("gsis_id")] %>% as.matrix(),
-  time = data_train_qb_matrix[,c("ageCentered20")] %>% as.matrix(),
+  X = data_train_qb %>% dplyr::select(season:att_br.rush) %>% as.matrix(), # predictors of the fixed effects
+  Y = data_train_qb[,c("fantasyPoints_lag")] %>% as.matrix(), # outcome variable
+  Z = data_train_qb %>% dplyr::mutate(constant = 1) %>% dplyr::select(constant, passing_yards, passing_tds, passing_interceptions, passing_epa, pacr) %>% as.matrix(), # predictors of the random effects
+  id = data_train_qb[,c("gsis_id")] %>% as.matrix(), # player ID (for nesting)
+  time = data_train_qb[,c("ageCentered20")] %>% as.matrix(), # time variable
   ntree = 500,
@@
   smerf,
-  X = data_test_qb %>% as_tibble() %>% dplyr::select(season:att_br.rush) %>% as.matrix(),
-  Z = data_test_qb[,c("pacr")] %>% as.matrix(),
+  X = data_test_qb %>% dplyr::select(season:att_br.rush) %>% as.matrix(),
+  Z = data_train_qb %>% dplyr::mutate(constant = 1) %>% dplyr::select(constant, passing_yards, passing_tds, passing_interceptions, passing_epa, pacr) %>% as.matrix(),
   id = data_test_qb[,c("gsis_id")] %>% as.matrix(),

### 2025-07-02 machine-learning.qmd
20250702 - cross-validation

@@
 
-We use the `future` [@R-future] package for parallel (faster) processing.
+We use the `future` package [@R-future] for parallel (faster) processing.
 
@@
 
-## Fitting the Traditional Regression Models {#sec-fittingModels-regression}
-
-### Regression with One Predictor {#sec-regressionOnePredictor}
-
-```{r}
-# Set seed for reproducibility
-set.seed(52242)
-
-# Set up Cross-Validation
-folds <- rsample::group_vfold_cv(
+## Set up the Cross-Validation Folds {#sec-folds}
+
+### *k*-Fold Cross-Validation {#sec-kFoldCrossValidation}
+
+*k*-fold cross-validation partitions the data into *k* folds (subsets).
+In each of the *k* iterations, the model is trained on $k - 1$ folds and is evaluated on the remaining fold.
+For example, in a 10-fold cross-validation (i.e., $k = 10$), as used below, the model is trained 10 times, each time leaving out a different 10% of the data for validation.
+*k*-fold cross-validation is widely used because it tends to yield stable estimates of model performance, by balancing bias and variance.
+It is also computationally efficient, requiring only *k* model fits to evaluate model performance.
+
+```{r}
+set.seed(52242) # for reproducibility
+
+folds_kFold <- rsample::group_vfold_cv(
   data_train_qb,
@@
   v = 10) # 10-fold cross-validation
+```
+
+### Leave-One-Out (LOO) Cross-Validation {#sec-looCrossValidation}
+
+Leave-one-out (LOO) cross-validation partitions the data into *n* folds, where *n* is the sample size.
+In each of the *n* iterations, the model is trained on $n - 1$ observations and is evaluated on the one left out.
+For example, in a LOO cross-validation with 100 players, the model is trained 100 times, each time leaving out a different player for validation.
+LOO cross-validation is especially useful when the dataset is small—too small to form reliable training sets in [*k*-fold cross-validation](#sec-kFoldCrossValidation) (e.g., with $k = 5$ or $k = 10$, which divide the sample into 5 or 10 folds, respectively).
+However, LOO tends to be less computationally efficient because it requires more model fits than [*k*-fold cross-validation](#sec-kFoldCrossValidation).
+LOO tends to have low bias, producing performance estimates closer to those obtained when fitting the model to the full dataset, because each model is trained on nearly all the data.
+However, LOO also tends to have high variance in its error estimates, because each validation fold contains only a single observation, making those estimates more sensitive to individual data points.
+
+```{r}
+set.seed(52242) # for reproducibility
+
+folds_loo <- rsample::loo_cv(data_train_qb)
+```
+
+## Fitting the Traditional Regression Models {#sec-fittingModels-regression}
+
+### Regression with One Predictor {#sec-regressionOnePredictor}
+
+```{r}
+# Set seed for reproducibility
+set.seed(52242)
+
+# Set up Cross-Validation
+folds <- folds_kFold
 
@@
 # Set up Cross-Validation
-folds <- rsample::group_vfold_cv(
-  data_train_qb,
-  group = gsis_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
-  v = 10) # 10-fold cross-validation
+folds <- folds_kFold
 
@@
 
-### *k*-Fold Cross-Validation {#sec-kfoldCV}
-
-### Leave-One-Out (LOO) Cross-Validation {#sec-looCV}
-
 ### Combining Tree-Boosting with Mixed Models {#sec-treeBoosting}

### 2025-07-02 machine-learning.qmd
20250702 - describe model types

@@
     - continuous outcome (i.e., regression)
-      - linear regression
-      - lasso regression
-      - ridge regression
-      - elastic net regression
+      - [linear regression](#sec-fittingModels-regression)
+      - [lasso regression](#sec-lasso)
+      - [ridge regression](#sec-ridgeRegression)
+      - [elastic net regression](#sec-elasticNet)
+      - [random forest](#sec-randomForest)
     - categorical outcome (i.e., classification)
-      - logistic regression
+      - [logistic regression](#sec-multipleRegressionTypes)
       - support vector machine
-      - random forest
-      - extreme gradient boosting
+      - [random forest](#sec-randomForest)
+      - boosting
+        - [tree boosting](#sec-treeBoosting)
 - unsupervised learning
-    - clustering
-    - principal component analysis
+    - [cluster analysis](#sec-clusterAnalysis)
+    - [principal component analysis](#sec-pca)
+    - [factor analysis](#sec-factorAnalysis)
 - semi-supervised learning
@@
 
-*Ensemble* machine learning methods combine multiple machine learning approaches with the goal that combining multiple approaches might lead to more accurate predictions that any one method might be able to achieve on its own.
+*Ensemble* machine learning methods combine multiple machine learning approaches with the goal that combining multiple approaches might lead to more accurate predictions than any one method might be able to achieve on its own.
 
@@
 
-[DEFINE SUPERVISED LEARNING]
-
-Unlike linear and logistic regression, various machine learning techniques can handle multicollinearity, including LASSO regression, ridge regression, and elastic net regression.
-Least absolute shrinkage and selection operator (LASSO) regression helps perform selection of which predictor variables to keep in the model by shrinking some coefficients to zero.
-Ridge regression shrinks the coefficients of predictor variables toward zero, but not to zero, so it does not perform selection of which predictor variables to retain; this allows it to allow nonzero coefficients for multiple correlated predictor variables in the context of multicollinearity.
-Elastic net involves a combination of LASSO and ridge regression; it performs selection of which predictor variables to keep by shrinking the coefficients of some predictor variables to zero, and it shrinks the coefficients of some predictor variables toward zero, to address multicollinearity.
-
-Unless interactions or nonlinear terms are specified, linear, logistic, LASSO, ridge, and elastic net regresstion do not account for interactions among the predictor variables or for nonlinear associations between the predictor variables and the outcome variable.
-By contrast, random forests and extreme gradient boosting do account for interactions among the predictor variables and for nonlinear associations between the predictor variables and the outcome variable.
+Supervised learning involves learning from data where the correct classification or outcome is known.
+For instance, predicting how many points a player will score is a supervised learning task, because there is a ground truth—the actual number of points scored—that can be used to train and evaluate the model.
+
+Unlike linear and logistic regression, various machine learning techniques can handle multicollinearity, including lasso regression, ridge regression, and elastic net regression.
+Least absolute shrinkage and selection operator (lasso) regression performs selection of which predictor variables to keep in the model by shrinking some coefficients to zero, effectively removing them from the model.
+Ridge regression shrinks the coefficients of predictor variables toward zero, but not to zero, so it does not perform selection of which predictor variables to retain; this allows it to yield stable estimates for multiple correlated predictor variables in the context of multicollinearity.
+Elastic net involves a combination of lasso and ridge regression; it performs selection of which predictor variables to keep by shrinking the coefficients of some predictor variables to zero (like lasso, for variable selection), and it shrinks the coefficients of some predictor variables toward zero (like ridge, for handling multicollinearity among correlated predictors).
+
+Unless interactions or nonlinear terms are specified, linear, logistic, lasso, ridge, and elastic net regression assume additive and linear associations between the predictors and outcome.
+That is, they do not automatically account for interactions among the predictor variables or for nonlinear associations between the predictor variables and the outcome variable (unless interaction terms or nonlinear transformations are explicitly included).
+By contrast, random forests and tree boosting methods automatically account for interactions and nonlinear associations between predictors and the outcome variable.
+These models recursively partition the data in ways that capture complex patterns without the need to manually specify interaction or polynomial terms.
 
@@
 
-[DEFINE UNSUPERVISED LEARNING]
+Unsupervised learning involves learning from data without known classifications.
+Unsupervised learning is used to discover hidden patterns, groupings, or structures in the data.
+For instance, if we want to identify different subtypes of Wide Receivers based on their playing style or performance metrics, or uncover underlying dimensions in a large dataset, we would use an unsupervised learning approach.
 
@@
 
-[DEFINE SEMI-SUPERVISED LEARNING]
+Semi-supervised learning combines supervised learning and unsupervised learning by training the model on some data for which the classification is known and some data for which the classification is not known.
 
@@
 
-[DEFINE REINFORCEMENT LEARNING]
+Reinforcement learning involves an agent learning to make decisions by interacting with the environment.
+Through trial and error, the agent receives feedback in the form of rewards or penalties and learns a strategy that maximizes the cumulative reward over time.
 
@@
 
+Many of the machine learning approaches described in this chapter require no missing observations in order for a case to be included in the analysis.
+In this section, we demonstrate one approach to imputing missing data.
 Here is a vignette demonstrating how to impute missing data using `missForest()`: <https://rpubs.com/lmorgan95/MissForest> (archived at: <https://perma.cc/6GB4-2E22>).
@@
 
-## Fitting the Traditional Regression Models {#sec-fittingModels-regression}
+## Fitting the Traditional Linear Regression Models {#sec-fittingModels-regression}
+
+We describe linear regression in @sec-multipleRegression.
 
@@
 
-### Least Absolute Shrinkage and Selection Option (LASSO) {#sec-lasso}
+### Least Absolute Shrinkage and Selection Option (Lasso) {#sec-lasso}
 
@@
 #| label: fig-lassoModel
-#| fig-cap: "Predicted Versus Actual Fantasy Points for Least Absolute Shrinkage and Selection Option (LASSO) Model."
-#| fig-alt: "Predicted Versus Actual Fantasy Points for Least Absolute Shrinkage and Selection Option (LASSO) Model."
+#| fig-cap: "Predicted Versus Actual Fantasy Points for Least Absolute Shrinkage and Selection Option (Lasso) Model."
+#| fig-alt: "Predicted Versus Actual Fantasy Points for Least Absolute Shrinkage and Selection Option (Lasso) Model."
 

### 2025-07-02 multiple-regression.qmd
20250702 - describe model types

@@
 
+## Different Types of Regression {#sec-multipleRegressionTypes}
+
+When the outcome variable is continuous, linear regression is common.
+However, there are other types of regression depending on type and distribution of the [outcome variable](#sec-correlationalStudy).
+For instance, if the [outcome variable](#sec-correlationalStudy) is binary, logistic regression would be used.
+If the [outcome variable](#sec-correlationalStudy) is an ordered categorical variable, ordinal regression would be used.
+If the [outcome variable](#sec-correlationalStudy) is a count, Poisson or negative binomial regression would be preferable.
+If the [outcome variable](#sec-correlationalStudy) is a proportion, beta regression would be used.
+
+[INSERT EXAMPLES]
+
 ## Conclusion {#sec-multipleRegressionConclusion}

### 2025-07-03 machine-learning.qmd
20250702 - tweak
20250703 - descriptions

@@
 
+Several data processing steps are necessary to get the data in the form necessary for machine learning.
+
 ```{r}
@@
 
+First, we apply several steps.
+We subset to the positions and variables of interest.
+We also rename columns and change variable types to make sure they match the column names and types across objects, which will be important later when we merge the data.
+
 ```{r}
@@
 
-Identify objects with shared variable names:
+Below, we identify shared variable names across objects to be merged to make sure we account for them in merging:
 
@@
 
-To merge data, we use the `powerjoin` package [@R-powerjoin]:
+To perform machine learning, we need all of the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy) in the same data file.
+Thus, we must merge data files.
+To merge data, we use the `powerjoin` package [@R-powerjoin], which allows coalescing variables with the same name from two different objects.
+We specify `coalesce_xy`, which means that—for variables that have the same name across both objects—it keeps the value from object 1 (if present); if not, it keeps the value from object 2.
+We first merge variables from objects that have the same structure—player data (i.e., `id` form), seasonal data (i.e., `id`-`season` form), or weekly data (i.e., `id`-`season`-`week` form).
 
@@
 
-Identify objects with shared variable names:
+To prepare for merging player data with seasonal data, we identify shared variable names across the objects:
 
@@
 ```
+
+Then we merge the player data with the seasonal data:
 
@@
 
+To prepare for merging player and seasonal data with weekly data, we identify shared variable names across the objects:
+
 ```{r}
@@
 ```
+
+Then we merge the player and seasonal data with the weekly data:
 
@@
 ### Additional Processing {#sec-mlAdditionalProcessing}
+
+For purposes of machine learning, we set all character and logical columns to factors.
 
@@
 
+For variables that are not expected to change, such as a player's name and position, we fill in missing values by using a player's value on those variables from other rows in the data.
+
 ```{r}
@@
 
+We create a new data object that contains the latest seasonal data, for merging with later predictions.
+
 ```{r}
@@
 ### Lag Fantasy Points {#sec-lagFantasyPoints}
+
+To develop a machine learning model that uses a player's performance metrics in a given season for predicting the player's fantasy points in the subsequent season, we need to include the player's fantasy points from the subsequent season in the same row as the previous season's performance metrics.
+Thus, we need to create a lagged variable for fantasy points.
+That way, 2024 fantasy points are in the same row as 2023 performance metrics, 2023 fantasy points are in the same row as 2023 performance metrics, and so on.
+We call this the lagged fantasy points variable (`fantasyPoints_lag`).
+We also retain the original same-year fantasy points variable (`fantasyPoints`) so it can be used as predictor of their subsequent-year fantasy points.
 
@@
 
+Then, we drop variables that we do not want to include in the model as our [predictor](#sec-correlationalStudy) or [outcome](#sec-correlationalStudy) variable.
+Thus, all of the variables in the object are our [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variables.
+
 ```{r}
@@
 
+Then, we separate the objects by position, so we can develop different machine learning models for each position.
+
 ```{r}
@@
 ### Split into Test and Training Data {#sec-splitTestTraining}
+
+Because machine learning can leverage many predictors, it is at high risk of [overfitting](#sec-overfitting)—explaining error variance that would not generalize to new data, such as data for new players or future seasons.
+Thus, it is important to develop and tune the machine learning model so as not to [overfit](sec-overfitting) the model.
+In machine learning, it is common to use [cross-validation](#sec-crossValidation) where we train the model on a subset of the observations, and we evaluate how well the model generalizes to unseen (e.g., "hold-out") observations.
+Then, we select the model parameters by how well the model generalizes to the hold-out data, so we are selecting  a model that maximizes accuracy and generalizability (i.e., parsimony).
+
+For internal [cross-validation](#sec-crossValidation), it is common to divide the data into three subsets:
+
+1. training data
+1. validation data
+1. test data
+
+The training set is used to fit the model.
+It is usually the largest portion of the data.
+We fit various models to the training set based on which parameters we want to evaluate (e.g., how many trees to use in a tree-boosting model).
+
+The models fit with the training set are then evaluated using the unseen observations in the validation set.
+The validation set is used to tune the model parameters and prevent [overfitting](#sec-overfitting).
+We select the model parameters that yield the greatest accuracy in the validation set.
+In [*k*-fold cross-validation](#sec-kFoldCrossValidation), the validation set rotates across folds, thus replacing the need for a separate validation set.
+
+The test set is used after model training and tuning to evaluate the model's generalizability to unseen data.
+
+Below, we split the data into test and training data.
+Our ultimate goal is to predict next year's fantasy points.
+However, to do that effectively, we must first develop a model for which we can evaluate its accuracy against historical fantasy points (because we do not yet know players will score in the future).
+We want to include all current/active players in our training data, so that our predictions of their future performance can be accounted for by including their prior data in the model.
+Thus, we use retired players as our hold-out (test) data.
+We split our data into 80% training data and 20% testing data.
+The 20% testing data thus includes all retired players, but not all retired players are in the testing data.
+
+Then, for the analysis, we can either a) use rotating folds (as the case for [*k*-fold](#sec-kFoldCrossValidation) and [leave-one-out [LOO]](#sec-looCrossValidation) cross-validation) for which a separate validation set (from the training set) is not needed, as we do in @sec-folds, or we can b) subdivide the training set into an inner training set and validation set, as we do in @sec-treeBoostingOptimalTuningParameters.
 
@@
 
+Below, we fit a linear regression model with one predictor and evaluate it with cross-validation.
+We also evaluate its accuracy on the hold-out (test) data.
+
 ```{r}
@@
 ```
+
+@fig-regressionOnePredictor depicts the predicted versus actual fantasy points for the model on the test data.
 
@@
 
+Below are the model predictions for next year's fantasy points:
+
 ```{r}
@@
 
+Below, we fit a linear regression model with multiple predictors and evaluate it with cross-validation.
+We also evaluate its accuracy on the hold-out (test) data.
+
 ```{r}
@@
 ```
+
+@fig-regressionMultiplePredictors depicts the predicted versus actual fantasy points for the model on the test data.
 
@@
 
+Below, we fit a lasso model.
+We evaluate it and tune its penalty parameter with cross-validation.
+We also evaluate its accuracy on the hold-out (test) data.
+
 ```{r}
@@
 ```
+
+@fig-lassoModel depicts the predicted versus actual fantasy points for the model on the test data.
 
@@
 
+Below, we fit a ridge regression model.
+We evaluate it and tune its penalty parameter with cross-validation.
+We also evaluate its accuracy on the hold-out (test) data.
+
 ```{r}
@@
 ```
+
+@fig-ridgeRegressionModel depicts the predicted versus actual fantasy points for the model on the test data.
 
@@
 
+Below, we fit an elastic net model.
+We evaluate it and tune its penalty and mixture parameters with cross-validation.
+We also evaluate its accuracy on the hold-out (test) data.
+
 ```{r}
@@
 ```
+
+@fig-elasticNetModel depicts the predicted versus actual fantasy points for the model on the test data.
 
@@
 ```
+
+@fig-randomForestModel depicts the predicted versus actual fantasy points for the model on the test data.
 
@@
 
-Approaches to estimating random forest models with longitudinal data are described by @@Hu2023.
+Approaches to estimating random forest models with longitudinal data are described by @Hu2023.
 Below, we fit longitudinal random forest models using the `MERF()` function of the `LongituRF` package [@R-LongituRF].
@@
 set.seed(52242)
-valid_tune_idx_qb <- sample.int(ntrain_qb, as.integer(0.2*ntrain_qb)) #
+valid_tune_idx_qb <- sample.int(ntrain_qb, as.integer(0.2*ntrain_qb)) # validation set
 

### 2025-07-03 machine-learning.qmd
20250703 - describe parameters

@@
       - [linear regression](#sec-fittingModels-regression)
-      - [lasso regression](#sec-lasso)
+      - [least absolute shrinkage and selection option (LASSO) regression](#sec-lasso)
       - [ridge regression](#sec-ridgeRegression)
@@
 
-Unlike linear and logistic regression, various machine learning techniques can handle multicollinearity, including lasso regression, ridge regression, and elastic net regression.
-Least absolute shrinkage and selection operator (lasso) regression performs selection of which predictor variables to keep in the model by shrinking some coefficients to zero, effectively removing them from the model.
-Ridge regression shrinks the coefficients of predictor variables toward zero, but not to zero, so it does not perform selection of which predictor variables to retain; this allows it to yield stable estimates for multiple correlated predictor variables in the context of multicollinearity.
-Elastic net involves a combination of lasso and ridge regression; it performs selection of which predictor variables to keep by shrinking the coefficients of some predictor variables to zero (like lasso, for variable selection), and it shrinks the coefficients of some predictor variables toward zero (like ridge, for handling multicollinearity among correlated predictors).
-
-Unless interactions or nonlinear terms are specified, linear, logistic, lasso, ridge, and elastic net regression assume additive and linear associations between the predictors and outcome.
+Unlike linear and logistic regression, various machine learning techniques can handle [multicollinearity](#sec-multipleRegressionMulticollinearity), including [LASSO regression](#sec-lasso), [ridge regression](#sec-ridgeRegression), and [elastic net regression](#sec-elasticNet).
+[Least absolute shrinkage and selection option (LASSO) regression](#sec-lasso) performs selection of which predictor variables to keep in the model by shrinking some coefficients to zero, effectively removing them from the model.
+[Ridge regression](#sec-ridgeRegression) shrinks the coefficients of predictor variables toward zero, but not to zero, so it does not perform selection of which predictor variables to retain; this allows it to yield stable estimates for multiple correlated predictor variables in the context of [multicollinearity](#sec-multipleRegressionMulticollinearity).
+[Elastic net](#sec-elasticNet) involves a combination of [LASSO](#sec-lasso) and [ridge](#sec-ridgeRegression) regression; it performs selection of which predictor variables to keep by shrinking the coefficients of some predictor variables to zero (like [LASSO](#sec-lasso), for variable selection), and it shrinks the coefficients of some predictor variables toward zero (like [ridge](#sec-ridgeRegression), for handling [multicollinearity](#sec-multipleRegressionMulticollinearity) among correlated predictors).
+
+Unless interactions or nonlinear terms are specified, linear, logistic, [LASSO](#sec-lasso), [ridge](#sec-ridgeRegression), and [elastic net](#sec-elasticNet) regression assume additive and linear associations between the predictors and outcome.
 That is, they do not automatically account for interactions among the predictor variables or for nonlinear associations between the predictor variables and the outcome variable (unless interaction terms or nonlinear transformations are explicitly included).
@@
 
-### Least Absolute Shrinkage and Selection Option (Lasso) {#sec-lasso}
-
-Below, we fit a lasso model.
+### Least Absolute Shrinkage and Selection Option (LASSO) {#sec-lasso}
+
+Below, we fit a LASSO model.
 We evaluate it and tune its penalty parameter with cross-validation.
-We also evaluate its accuracy on the hold-out (test) data.
+The penalty parameter in a LASSO model controls the strength of regularization applied to the model coefficients.
+Smaller penalty values result in less regularization, allowing the model to retain more predictors with larger (nonzero) coefficients.
+This typically reduces bias but increases variance of the model's predictions, as the model may [overfit](#sec-overfitting) to the training data by including irrelevant or weak predictors.
+Larger penalty values apply stronger regularization, shrinking more coefficients exactly to zero.
+This encourages a sparser model that may increase bias (by excluding useful predictors) but reduces variance and improves generalizability by simplifying the model and reducing [overfitting](#sec-overfitting).
+
+After tuning the model, we evaluate its accuracy on the hold-out (test) data.
 
@@
 We evaluate it and tune its penalty parameter with cross-validation.
-We also evaluate its accuracy on the hold-out (test) data.
+The penalty parameter in a ridge regression model controls the amount of regularization applied to the model's coefficients.
+Smaller penalty values allow coefficients to remain large, resulting in a model that closely fits the training data.
+This may reduce bias but increases the risk of [overfitting](#sec-overfitting), especially in the presence of [multicollinearity](#sec-multipleRegressionMulticollinearity) or many weak predictors.
+Larger penalty values shrink coefficients toward zero (though not exactly to zero), which reduces model complexity.
+This typically increases bias slightly but reduces variance of the model's predictions, making the model more stable and better suited for generalization to new data.
+
+After tuning the model, we also evaluate its accuracy on the hold-out (test) data.
 
@@
 We evaluate it and tune its penalty and mixture parameters with cross-validation.
-We also evaluate its accuracy on the hold-out (test) data.
+
+The penalty parameter controls the overall strength of regularization applied to the model’s coefficients.
+Smaller penalty values allow coefficients to remain large, which can reduce bias but increase variance of the model's predictions and can increase the risk of [overfitting](#sec-overfitting).
+Larger penalty values shrink coefficients more aggressively, leading to simpler models with potentially higher bias but lower variance of predictions.
+This regularization helps prevent [overfitting](#sec-overfitting), especially when the model includes many predictors or [multicollinearity](#sec-multipleRegressionMulticollinearity).
+
+The mixture parameter controls the balance between [LASSO](#sec-lasso) and [ridge](#sec-ridgeRegression) regularization.
+It ranges from 0 to 1:
+A value of 0 applies pure [ridge regression](#sec-ridgeRegression), which shrinks all coefficients but keeps them in the model.
+A value of 1 applies pure [LASSO](#sec-lasso), which can shrink some coefficients exactly to zero, effectively performing variable selection.
+Values between 0 and 1 apply a combination: [ridge](#sec-ridgeRegression)-like smoothing and [LASSO](#sec-lasso)-like sparsity.
+Smaller mixture values favor shrinkage without variable elimination, whereas larger values favor sparser solutions by excluding weak predictors.
+
+After tuning the model, we also evaluate its accuracy on the hold-out (test) data.
 
@@
 
+Below, we fit a random forest model.
+We evaluate it and tune two parameters with cross-validation.
+The first parameter is `mtry`, which controls the number of predictors randomly sampled at each split in a decision tree.
+Smaller `mtry` values increase tree diversity by forcing trees to consider different subsets of predictors.
+This typically reduces the variance of the overall model's predictions (because the final prediction is averaged over more diverse trees) but may increase bias if strong predictors are frequently excluded.
+Larger `mtry` allow more predictors to be considered at each split, making trees more similar to each other.
+This can reduce bias but often increases variance of the model's predictions, because the trees are more correlated and less effective at error cancellation when averaged.
+
+The second parameter is `min_n`, which controls the minimum number of observations that must be present in a node for a split to be attempted.
+Smaller `min_n` values allow trees to grow deeper and capture more fine-grained patterns in the training data.
+This typically reduces bias but increases variance of the overall model's predictions, because deeper trees are more likely to [overfit](#sec-overfitting) to noise in the training data.
+Larger `min_n` values restrict the depth of the trees by requiring more data to justify a split.
+This can reduce variance by producing simpler, more generalizable trees—but may increase bias if the trees are unable to capture important structure in the data.
+
+After tuning the model, we evaluate its accuracy on the hold-out (test) data.
+
 ```{r}
@@
 
+The above approaches to machine learning assume that the observations are independent across rows.
+However, in our case, this assumption does not hold because the data are longitudinal—each player has multiple seasons, and each row corresponds to a unique player–season combination.
+In the approaches below, using a random forest model (this section) and a [tree-boosting model](#sec-treeBoosting) (in @sec-treeBoosting), we address this by explicitly accounting for the nesting of longitudinal observations within players.
+
 Approaches to estimating random forest models with longitudinal data are described by @Hu2023.

### 2025-07-03 index.qmd
20250703 - types of regression

--- 2025-07-01 index.qmd
@@
   "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", "caret", 
-  "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation", "mgcv", 
-  "rms", "car", "performance", "lavaan", "mice", "miceadds", "interactions", 
-  "shiny", "PearsonDS", "robustbase", "data.table", "future", "future.apply", 
-  "QuantPsyc", "SimDesign"
+  "tidymodels", "LongituRF", "gpboost", "cmdstanr", "learnr", "remotes", 
+  "corrplot", "correlation", "mgcv", "rms", "car", "performance", "lavaan", 
+  "mice", "miceadds", "interactions", "shiny", "PearsonDS", "robustbase", 
+  "ordinal", "MASS", "data.table", "future", "future.apply", "QuantPsyc", 
+  "SimDesign", "effectsize"
 )

### 2025-07-03 multiple-regression.qmd
20250703 - types of regression

--- 2025-07-01 multiple-regression.qmd
@@
 library("robustbase")
+library("ordinal")
+library("MASS")
 library("broom")
@@
 That is, if we used the same sampling procedure repeatedly, the true value of the regression coefficient would be expected to be 95% of the time somewhere between 0.30 to 0.70.
+
+## Types of Regression {#sec-multipleRegressionTypes}
+
+When the outcome variable is continuous, [linear regression](#sec-multipleRegressionLinear) is common.
+However, there are other types of regression depending on type and distribution of the [outcome variable](#sec-correlationalStudy).
+For instance, if the [outcome variable](#sec-correlationalStudy) is binary, [logistic regression](#sec-multipleRegressionLogistic) would be used.
+If the [outcome variable](#sec-correlationalStudy) is an ordered categorical variable, [ordinal regression](#sec-multipleRegressionOrdinal) would be used.
+If the [outcome variable](#sec-correlationalStudy) is a count, [Poisson](#sec-multipleRegressionPoisson) or [negative binomial](#sec-multipleRegressionNegativeBinomial) regression would be preferable.
+If the [outcome variable](#sec-correlationalStudy) is a proportion, [beta regression](#sec-multipleRegressionBeta) would be used.
+
+Here are examples of each:
+
+### Linear Regression {#sec-multipleRegressionLinear}
+
+```{r}
+linearRegression <- lm(
+  fantasyPoints ~ age + height + weight + target_share,
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  na.action = "na.exclude"
+)
+
+summary(linearRegression)
+```
+
+### Logistic Regression {#sec-multipleRegressionLogistic}
+
+```{r}
+newdata <- player_stats_weekly %>%
+  mutate( # create a binary variable to be used as the outcome variable
+    receiving_td = case_when(
+      is.na(receiving_tds) ~ NA_real_, # keep NA
+      receiving_tds >= 1   ~ 1,
+      receiving_tds == 0   ~ 0
+    )
+  )
+
+logisticRegression <- glm(
+  receiving_td ~ age + height + weight + target_share,
+  data = newdata %>% filter(position %in% c("WR")),
+  family = binomial(),
+  na.action = "na.exclude"
+)
+
+summary(logisticRegression)
+```
+
+### Ordinal Regression {#sec-multipleRegressionOrdinal}
+
+```{r}
+newdata <- player_stats_weekly
+newdata$receiving_tdsFactor <- factor(newdata$receiving_tds, ordered = TRUE)
+table(newdata$receiving_tdsFactor)
+
+ordinalRegression <- ordinal::clm(
+  receiving_tdsFactor ~ age + height + target_share,
+  data = newdata %>% filter(position %in% c("WR")),
+  na.action = "na.exclude"
+)
+
+summary(ordinalRegression)
+```
+
+### Poisson Regression {#sec-multipleRegressionPoisson}
+
+```{r}
+poissonRegression <- glm(
+  receiving_tds ~ age + height + weight + target_share,
+  data = newdata %>% filter(position %in% c("WR")),
+  family = poisson(),
+  na.action = na.exclude
+)
+
+summary(poissonRegression)
+```
+
+### Negative Binomial Regression {#sec-multipleRegressionNegativeBinomial}
+
+```{r}
+negativeBinomialRegression <- MASS::glm.nb(
+  receiving_tds ~ age + height + weight + target_share,
+  data = newdata %>% filter(position %in% c("WR")),
+  na.action = na.exclude,
+  control = glm.control(maxit = 10000)
+)
+
+summary(negativeBinomialRegression)
+```
+
+### Beta Regression {#sec-multipleRegressionBeta}
+
+```{r}
+#| output: false
+
+betaRegression <- brm(
+  formula = target_share ~ age + height + weight,
+  data = newdata %>% filter(position %in% c("WR")),
+  family = zero_one_inflated_beta(),
+)
+```
+
+```{r}
+summary(betaRegression)
+```
 

### 2025-07-04 multiple-regression.qmd
20250704 - effect sizes

@@
 summary(linearRegression)
+print(effectsize::standardize_parameters(linearRegression, method = "refit"), digits = 2)
 ```
@@
 summary(logisticRegression)
+print(effectsize::standardize_parameters(logisticRegression, method = "refit"), digits = 2)
 ```
@@
 summary(ordinalRegression)
+print(effectsize::standardize_parameters(ordinalRegression, method = "refit"), digits = 2)
 ```
@@
 summary(poissonRegression)
+print(effectsize::standardize_parameters(poissonRegression, method = "refit"), digits = 2)
 ```
@@
 summary(negativeBinomialRegression)
+print(effectsize::standardize_parameters(negativeBinomialRegression, method = "refit"), digits = 2)
 ```

### 2025-07-04 index.qmd
20250704 - track chunk execution times

@@
+```{r setup}
+#| include: false
+
+# Enable automatic chunk timing logging to GitHub Actions
+knitr::opts_chunk$set(time = TRUE)  # Apply to ALL chunks globally
+
+# Null coalescing operator
+`%||%` <- function(a, b) if (!is.null(a)) a else b
+
+# Define timing hook
+knitr::knit_hooks$set(time = function(before, options) {
+  if (before) {
+    options$start_time <- Sys.time()
+  } else {
+    elapsed <- difftime(Sys.time(), options$start_time, units = "secs")
+    msg <- sprintf("::notice title=Chunk Time::Chunk `%s` ran in %.2f seconds.", 
+                   options$label %||% "<unnamed>", as.numeric(elapsed))
+    message(msg)  # Send to GitHub Actions log
+  }
+})
+```
+
 # Preface {#sec-preface .unnumbered .unlisted}

### 2025-07-04 multiple-regression.qmd
20250704 - code chunk labels

@@
 ```{r}
+#| label: linear-regression
+
 linearRegression <- lm(
@@
 ```{r}
+#| label: logistic-regression
+
 newdata <- player_stats_weekly %>%
@@
 ```{r}
+#| label: ordinal-regression
+
 newdata <- player_stats_weekly
@@
 ```{r}
+#| label: poisson-regression
+
 poissonRegression <- glm(
@@
 ```{r}
+#| label: negative-binomial-regression
+
 negativeBinomialRegression <- MASS::glm.nb(
@@
 ```{r}
+#| label: beta-regression
 #| output: false
@@
 ```{r}
+#| label: beta-regression-summmary
+
 summary(betaRegression)

### 2025-07-04 causal-inference.qmd
20250704 - edits for grammar and clarity

@@
 In the [between-subject design](#sec-betweenSubject), the players differ between the two groups, so we cannot know how the individuals who received the treatment would have performed if they had actually not received the treatment.
-In the A-B-A-B [within-subject design](#sec-withinSubject), the players are the same, but they timepoints that they receive or do not receive the treatment differ, and there can be [carryover effects](#sec-withinSubject) from one condition to the next.
-For instance, consuming sports drinks before game 2 might also help them be better hydrated in general, including, for subsequent games.
-Thus, we cannot know how a player would have performed in game 1 with treatment or in game 2 without treatment, etc.
+In the A-B-A-B [within-subject design](#sec-withinSubject), all players receive the same sequence of alternating conditions—no treatment, treatment, no treatment, treatment—across four timepoints.
+Although the A-B-A-B [within-subject design](#sec-withinSubject) keeps participants constant, the specific occasion of measurement may differ from participant to participant (thus introducing time or context of measurement as a potential [confound](#sec-causalDiagramConfounding)), and there can be [carryover effects](#sec-withinSubject) from one condition to the next.
+For instance, consuming sports drinks before game 2 might also help the player be better hydrated in general, including for subsequent games.
+Thus, we cannot know how a player would have performed in game 1 with treatment or in game 2 without treatment, and so on.
 Nevertheless, it is important to be aware of the counterfactual and to engage in counterfactual reasoning to consider what would have happened if the supposed cause had not occurred.

### 2025-07-04 causal-inference.qmd
20250704 - grammar edit

@@
 Another approach could be to apply within-subject analyses.
-For instance, you could examining whether, within the same individual, if a player consumes a sports drink, do they perform better compared to games in which they did not consume a sports drink.
+For instance, you could examine whether, within the same individual, if a player consumes a sports drink, do they perform better compared to games in which they did not consume a sports drink.
 When we control for prior levels of the outcome in the prediction, we are evaluating whether the predictor is associated with witin-person *change* in the outcome.

### 2025-07-04 index.qmd
20250704 - log chunk execution times

--- 2025-07-03 index.qmd
@@
+```{r setup}
+#| include: false
+
+# Enable automatic chunk timing logging to GitHub Actions
+knitr::opts_chunk$set(time = TRUE) # Apply to ALL chunks globally
+
+# Null coalescing operator
+`%||%` <- function(a, b) if (!is.null(a)) a else b
+
+# Setup output file
+log_file <- "chunk-times.txt"
+if (file.exists(log_file)) file.remove(log_file)
+
+# Define hook
+knitr::knit_hooks$set(time = function(before, options) {
+  if (before) {
+    options$start_time <- Sys.time()
+  } else {
+    elapsed <- difftime(Sys.time(), options$start_time, units = "secs")
+    label <- options$label %||% "<unnamed>"
+    msg <- sprintf("Chunk `%s` ran in %.2f seconds.", label, as.numeric(elapsed))
+
+    # Log to GitHub Actions via stdout (so it's not suppressed)
+    cat(sprintf("::notice title=Chunk Time::%s\n", msg))
+
+    # Also log to file
+    cat(msg, file = log_file, append = TRUE, sep = "\n")
+  }
+})
+```
+
 # Preface {#sec-preface .unnumbered .unlisted}

### 2025-07-05 machine-learning.qmd
20250705 - machine learning summary?

@@
 library("tidyverse")
+library("knitr")
 ```
@@
 rec <- recipes::recipe(
-  fantasyPoints_lag ~ ageCentered20,
+  fantasyPoints_lag ~ fantasyPoints,
   data = data_train_qb)
@@
 )
+```
+
+```{r}
+#| include: false
+
+rsquare_onePredictor <- petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)$rsquared %>% round(., digits = 2)
 ```
@@
 
+```{r}
+#| include: false
+
+rsquare_multiplePredictors <- petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)$rsquared %>% round(., digits = 2)
+```
+
 @fig-regressionMultiplePredictors depicts the predicted versus actual fantasy points for the model on the test data.
@@
 
+```{r}
+#| include: false
+
+rsquare_lasso <- petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)$rsquared %>% round(., digits = 2)
+```
+
 @fig-lassoModel depicts the predicted versus actual fantasy points for the model on the test data.
@@
 
+```{r}
+#| include: false
+
+rsquare_ridge <- petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)$rsquared %>% round(., digits = 2)
+```
+
 @fig-ridgeRegressionModel depicts the predicted versus actual fantasy points for the model on the test data.
@@
 
+```{r}
+#| include: false
+
+rsquare_elasticNet <- petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)$rsquared %>% round(., digits = 2)
+```
+
 @fig-elasticNetModel depicts the predicted versus actual fantasy points for the model on the test data.
@@
 
-#### Cross-Sectional Data {#sec-randomForestCrossSectionalData}
+#### Assuming Independent Observations {#sec-randomForestIndependent}
 
@@
 )
+```
+
+```{r}
+#| include: false
+
+rsquare_randomForest <- petersenlab::accuracyOverall(
+  predicted = df$pred,
+  actual = df$fantasyPoints_lag,
+  dropUndefined = TRUE
+)$rsquared %>% round(., digits = 2)
 ```
@@
 
-#### Longitudinal Data {#sec-randomForestLongitudinalData}
+#### Accounting for Longitudinal Data {#sec-randomForestLongitudinalData}
 
@@
 smerf$OOB # OOB error at each iteration
-
+```
+
+```{r}
+#| label: fig-randomForestLongitudinalLogLikelihood
+#| fig-cap: "Evolution of the Log-Likelihood."
+#| fig-alt: "Evolution of the Log-Likelihood."
+
+plot(smerf$Vraisemblance)
+```
+
+The following code generates an error, for some reason.
+This issue has been posted on the GitHub repository: <https://github.com/sistm/LongituRF/issues/5>.
+Hopefully the package maintainer will help address the issue.
+
+```{r}
 # Predict on Test Data
@@
   time = data_test_qb[,c("ageCentered20")] %>% as.matrix())
-```
-
-```{r}
-#| label: fig-randomForestLongitudinalLogLikelihood
-#| fig-cap: "Evolution of the Log-Likelihood."
-#| fig-alt: "Evolution of the Log-Likelihood."
-
-plot(smerf$Vraisemblance)
 ```
@@
 
+```{r}
+#| include: false
+
+rsquare_treeBoosting <- petersenlab::accuracyOverall(
+  predicted = predictedVsActual$predictedPoints,
+  actual = predictedVsActual$actualPoints,
+  dropUndefined = TRUE
+)$rsquared %>% round(., digits = 2)
+```
+
 #### Generate Predictions for Next Season {#sec-treeBoostingModelPredictions}
@@
 
+## Summarizing the Machine Learning Models {#sec-summarizingModels-machineLearning}
+
+```{r}
+#| include: false
+
+machineLearningTable <- data.frame(
+  model = c(
+    "Regression with One Predictor",
+    "Regressions with Multiple Predictors",
+    "LASSO",
+    "Ridge Regression",
+    "Elastic Net",
+    "Random Forest",
+    "Tree-Boosting"
+  ), 
+  rsquared = c(
+    rsquare_onePredictor,
+    rsquare_multiplePredictors,
+    rsquare_lasso,
+    rsquare_ridge,
+    rsquare_elasticNet,
+    rsquare_randomForest,
+    rsquare_treeBoosting
+  )
+)
+```
+
+```{r}
+#| label: tbl-machineLearningModelSummary
+#| tbl-cap: "Summary of Machine Learning Models."
+#| echo: false
+
+knitr::kable(
+  machineLearningTable,
+  col.names = c("Model","$R^2$"),
+  booktabs = TRUE)
+```
+
 ## Conclusion {#sec-machineLearningConclusion}

### 2025-07-05 multiple-regression.qmd
20250705 - machine learning summary?

@@
 
-kable(
+knitr::kable(
   regression1,
@@
 
-kable(
+knitr::kable(
   regression2,
@@
 
-kable(
+knitr::kable(
   regression3,

### 2025-07-06 index.qmd
20250705 - tweaks
20250705 - log chunk execution times

--- 2025-07-03 index.qmd
@@
+```{r}
+#| label: setup
+#| include: false
+#| message: !expr NA
+
+# Preface
+
+# Globally apply message = NA (https://github.com/quarto-dev/quarto-cli/discussions/7443)
+knitr::opts_chunk$set(
+  time = TRUE,
+  message = NA # show messages in console but not book
+)
+
+# Null coalescing operator
+`%||%` <- function(a, b) if (!is.null(a)) a else b
+
+# Prepare log file
+log_file <- "chunk-times.txt"
+if (file.exists(log_file)) file.remove(log_file)
+
+# Timing hook
+knitr::knit_hooks$set(time = function(before, options) {
+  if (before) {
+    options$start_time <- Sys.time()
+  } else {
+    elapsed <- difftime(Sys.time(), options$start_time, units = "secs")
+    label <- options$label %||% "<unnamed>"
+    msg <- sprintf("Chunk `%s` ran in %.2f seconds.", label, as.numeric(elapsed))
+
+    # Log to GitHub Actions console
+    message(sprintf("::notice title=Chunk Time::%s", msg))
+
+    # Also save to file
+    cat(msg, file = log_file, append = TRUE, sep = "\n")
+  }
+})
+```
+
 # Preface {#sec-preface .unnumbered .unlisted}

### 2025-07-07 index.qmd
20250705 - message = NA
20250706 - track chunk execution times attempt 1

--- 2025-07-04 index.qmd
@@
-```{r setup}
+```{r}
+#| label: setup
 #| include: false
 
-# Enable automatic chunk timing logging to GitHub Actions
-knitr::opts_chunk$set(time = TRUE)  # Apply to ALL chunks globally
+# Global knitr options
+knitr::opts_chunk$set(
+  time = TRUE,
+  message = NA # Show in console, not in rendered output
+)
 
@@
 
-# Define timing hook
+# Log file setup
+log_file <- "chunk-times.txt"
+if (file.exists(log_file)) file.remove(log_file)
+
+# Environment to track start times
+.chunk_times_env <- new.env()
+
+# Timing hook
 knitr::knit_hooks$set(time = function(before, options) {
+  label <- options$label %||% "<unnamed>"
+
   if (before) {
-    options$start_time <- Sys.time()
+    .chunk_times_env[[label]] <- Sys.time()
   } else {
-    elapsed <- difftime(Sys.time(), options$start_time, units = "secs")
-    msg <- sprintf("::notice title=Chunk Time::Chunk `%s` ran in %.2f seconds.", 
-                   options$label %||% "<unnamed>", as.numeric(elapsed))
-    message(msg)  # Send to GitHub Actions log
+    start_time <- .chunk_times_env[[label]]
+    if (!is.null(start_time)) {
+      elapsed <- difftime(Sys.time(), start_time, units = "secs")
+      msg <- sprintf("Chunk `%s` ran in %.2f seconds.", label, as.numeric(elapsed))
+
+      # Send message to GitHub Actions log
+      message(sprintf("::notice title=Chunk Time::%s", msg))
+
+      # Also save to file
+      cat(msg, file = log_file, append = TRUE, sep = "\n")
+    }
   }

### 2025-07-07 getting-started.qmd
20250707 - additional resources

@@
 
+## Learning Statistics {#sec-learningStats}
+
+Here are resources for learning statistics:
+
+- OpenIntro Statistics: <https://www.openintro.org/book/os/>
+- Introductory Statistics: <https://openstax.org/details/books/introductory-statistics-2e>
+
 ## Learning `R` {#sec-learningR}
@@
 - Video training courses in `R` skills: <https://www.pluralsight.com/search?q=R>
+- `R` for Data Science: <https://r4ds.hadley.nz>
 - Browse the `Cookbook for R` to find solutions to common tasks and problems: <http://www.cookbook-r.com>
@@
 - Introduction to `R` course on `Datacamp`: <https://www.datacamp.com/courses/free-introduction-to-r>
+- Learning Statistics with `R`: <https://learningstatisticswithr.com>
 - Teaching `R` in a Kinder, Gentler, More Effective Manner: <https://github.com/matloff/TidyverseSkeptic>
@@
 My strong recommendation is to provide a `reprex` whenever possible: <https://reprex.tidyverse.org>.
-@Bryan2025 provide Reprex do's and don'ts: <https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html>.
+@Bryan2025 provide `reprex` do's and don'ts: <https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html>.
 

### 2025-07-07 machine-learning.qmd
20250705 - tweaks
20250707 - citations

@@
 
+The LASSO models were fit using the `glmnet` package [@R-glmnet; @Friedman2010_packages; @Tay2023_packages].
+
 ```{r}
@@
 
+The ridge regression models were fit using the `glmnet` package [@R-glmnet; @Friedman2010_packages; @Tay2023_packages].
+
 ```{r}
@@
 
+The elastic net models were fit using the `glmnet` package [@R-glmnet; @Friedman2010_packages; @Tay2023_packages].
+
 ```{r}
@@
 
+The random forest models were fit using the `ranger` package [@R-ranger; @Wright2017_packages].
+
 ```{r}
@@
 
+A summary of the accuracy of the machine learning models is in @tbl-machineLearningModelSummary.
+
 ```{r}

### 2025-07-07 multiple-regression.qmd
20250707 - citations

@@
 
+We fit a linear regression model using the `lm()` function.
+
 ```{r}
@@
 ### Logistic Regression {#sec-multipleRegressionLogistic}
+
+We fit a logistic regression model using the `glm()` function and specifying `family = binomial()`.
 
@@
 
+We fit an ordinal regression model using the `clm()` function of the `ordinal` package [@R-ordinal].
+
 ```{r}
@@
 ### Poisson Regression {#sec-multipleRegressionPoisson}
+
+We fit a Poisson regression model using the `glm()` function and specifying `family = poisson()`.
 
@@
 
+We fit a negative binomial regression model using the `glm.nb()` function of the `MASS` package [@R-MASS].
+
 ```{r}
@@
 
+We fit a (Bayesian) zero-one-inflated beta regression model (to allow for zeros and ones) using the `brm()` function of the `brms` package [@R-brms] and specifying `family = zero_one_inflated_beta()`.
+
+::: {#nte-bayesianBetaRegression .callout-note title="Bayesian beta regression"}
+Note: the following code that runs the model takes a while.
+:::
+
 ```{r}
@@
 
-betaRegression <- brm(
+betaRegression <- brms::brm(
   formula = target_share ~ age + height + weight,

### 2025-07-07 player-evaluation.qmd
20250617 - clean up formulas
20250707 - tweaks

@@
 This is important to consider because different players have different opportunities.
-For example, some Running Backs had more opportunity than others (e.g., more carries), so comparing players merely on rushing yards would be misleading.
+For example, some Running Backs have more opportunity than others (e.g., more carries), so comparing players merely on rushing yards would be misleading.
 If they were relatively more efficient, they will likely score more points than many of their peers when given more opportunities.
-If they were relativelly inefficient, their capacity to score fantasy points may be more dependent on touches/opportunities.
-Efficiency might be operationalized by indicators such as yards per passing attempt, yards per rushing attempt, yards per target, yards per reception, etc.
+If they were relatively inefficient, their capacity to score fantasy points may be more dependent on touches/opportunities.
+Efficiency might be operationalized by indicators such as yards per passing attempt, yards per rushing attempt, yards per target, yards per reception, yards per route run, etc.
 
@@
 $$
-\begin{aligned}
-  \text{CV} &= \frac{s_x}{\bar{x}}\\
-  &= \frac{\sigma_x}{\mu_x}
-\end{aligned}
+\text{CV} = \frac{s_x}{\bar{x}}
 $$ {#eq-coefficientOfVariation}
@@
 In terms of the player's opportunity and usage, how many snaps do they get?
+How many routes do they run?
 How many touches and/or targets do they receive?
-Being on the field for more snaps and receiving more touches and/or targets means that the player has more opportunities to score fantasy points.
+Being on the field for more snaps, running more routes, and receiving more touches and/or targets means that the player has more opportunities to score fantasy points.
 Are they targeted in the red zone?
@@
 Another aspect to consider is how challenging their matchup(s) and strength of schedule is.
-For a Quarterback, it is valuable to consider how strong the oppenent's passing defense is.
+For a Quarterback, it is valuable to consider how strong the opponent's passing defense is.
 For a Running Back, how strong is the running defense?

### 2025-07-07 machine-learning.qmd
20250707 - accuracy indices and summary

@@
 This can result in overfitting.
-Thus, when estimating machine learning models, it is common to keep a hold-out sample for use in cross-validation to evaluate the extent of shrinkage of model coefficients.
+Thus, when estimating machine learning models, it is common to keep a hold-out sample for use in [cross-validation](#sec-crossValidation) to evaluate the extent of shrinkage of model coefficients.
 The data that the model is trained on is known as the "training data".
@@
 
-If the test data were collected as part of the same processes as the original data and were merely held out for purposes of analysis, this is called internal cross-validation.
-If the test data were collected separately from the original data used to train the model, this is called external cross-validation.
+If the test data were collected as part of the same processes as the original data and were merely held out for purposes of analysis, this is called internal [cross-validation](#sec-crossValidation).
+If the test data were collected separately from the original data used to train the model, this is called external [cross-validation](#sec-crossValidation).
 
@@
 We then use the imputed training data to make out-of-sample predictions to fill in the missing data for the testing data.
-We do not want to impute the training and testing data together so that we can keep them separate for the purposes of cross-validation.
+We do not want to impute the training and testing data together so that we can keep them separate for the purposes of [cross-validation](#sec-crossValidation).
 However, we impute all data (training and test data together) for purposes of making out-of-sample predictions from the machine learning models to predict players' performance next season (when actuals are not yet available for evaluating their accuracy).
@@
 
+In the examples below, we predict the future fantasy points of Quarterbacks.
+However, the examples could be applied to any of the positions.
+There are various approaches to [cross-validation](#sec-crossValidation).
+In the examples below, we use [*k*-fold cross-validation](#sec-kFoldCrossValidation).
+However, we also provide the code to apply [leave-one-out (LOO) cross-validation](#sec-looCrossValidation).
+
 ### *k*-Fold Cross-Validation {#sec-kFoldCrossValidation}
@@
 
-Below, we fit a linear regression model with one predictor and evaluate it with cross-validation.
+Below, we fit a linear regression model with one predictor and evaluate it with [cross-validation](#sec-crossValidation).
 We also evaluate its accuracy on the hold-out (test) data.
@@
 
-rsquare_onePredictor <- petersenlab::accuracyOverall(
+accuracyIndices_onePredictor <- petersenlab::accuracyOverall(
   predicted = df$pred,
@@
   dropUndefined = TRUE
-)$rsquared %>% round(., digits = 2)
+)
+
+rsquare_onePredictor <- round(accuracyIndices_onePredictor$rsquared, digits = 2)
+me_onePredictor <- round(accuracyIndices_onePredictor$ME, digits = 2)
+mae_onePredictor <- round(accuracyIndices_onePredictor$MAE, digits = 2)
 ```
@@
 
-Below, we fit a linear regression model with multiple predictors and evaluate it with cross-validation.
+Below, we fit a linear regression model with multiple predictors and evaluate it with [cross-validation](#sec-crossValidation).
 We also evaluate its accuracy on the hold-out (test) data.
@@
 
-rsquare_multiplePredictors <- petersenlab::accuracyOverall(
+accuracyIndices_multiplePredictors <- petersenlab::accuracyOverall(
   predicted = df$pred,
@@
   dropUndefined = TRUE
-)$rsquared %>% round(., digits = 2)
+)
+
+rsquare_multiplePredictors <- round(accuracyIndices_multiplePredictors$rsquared, digits = 2)
+me_multiplePredictors <- round(accuracyIndices_multiplePredictors$ME, digits = 2)
+mae_multiplePredictors <- round(accuracyIndices_multiplePredictors$MAE, digits = 2)
 ```
@@
 Below, we fit a LASSO model.
-We evaluate it and tune its penalty parameter with cross-validation.
+We evaluate it and tune its penalty parameter with [cross-validation](#sec-crossValidation).
 The penalty parameter in a LASSO model controls the strength of regularization applied to the model coefficients.
@@
 
-rsquare_lasso <- petersenlab::accuracyOverall(
+accuracyIndices_lasso <- petersenlab::accuracyOverall(
   predicted = df$pred,
@@
   dropUndefined = TRUE
-)$rsquared %>% round(., digits = 2)
+)
+
+rsquare_lasso <- round(accuracyIndices_lasso$rsquared, digits = 2)
+me_lasso <- round(accuracyIndices_lasso$ME, digits = 2)
+mae_lasso <- round(accuracyIndices_lasso$MAE, digits = 2)
 ```
@@
 Below, we fit a ridge regression model.
-We evaluate it and tune its penalty parameter with cross-validation.
+We evaluate it and tune its penalty parameter with [cross-validation](#sec-crossValidation).
 The penalty parameter in a ridge regression model controls the amount of regularization applied to the model's coefficients.
@@
 
-rsquare_ridge <- petersenlab::accuracyOverall(
+accuracyIndices_ridge <- petersenlab::accuracyOverall(
   predicted = df$pred,
@@
   dropUndefined = TRUE
-)$rsquared %>% round(., digits = 2)
+)
+
+rsquare_ridge <- round(accuracyIndices_ridge$rsquared, digits = 2)
+me_ridge <- round(accuracyIndices_ridge$ME, digits = 2)
+mae_ridge <- round(accuracyIndices_ridge$MAE, digits = 2)
 ```
@@
 Below, we fit an elastic net model.
-We evaluate it and tune its penalty and mixture parameters with cross-validation.
+We evaluate it and tune its penalty and mixture parameters with [cross-validation](#sec-crossValidation).
 
@@
 
-rsquare_elasticNet <- petersenlab::accuracyOverall(
+accuracyIndices_elasticNet <- petersenlab::accuracyOverall(
   predicted = df$pred,
@@
   dropUndefined = TRUE
-)$rsquared %>% round(., digits = 2)
+)
+
+rsquare_elasticNet <- round(accuracyIndices_elasticNet$rsquared, digits = 2)
+me_elasticNet <- round(accuracyIndices_elasticNet$ME, digits = 2)
+mae_elasticNet <- round(accuracyIndices_elasticNet$MAE, digits = 2)
 ```
@@
 Below, we fit a random forest model.
-We evaluate it and tune two parameters with cross-validation.
+We evaluate it and tune two parameters with [cross-validation](#sec-crossValidation).
 The first parameter is `mtry`, which controls the number of predictors randomly sampled at each split in a decision tree.
@@
 
-rsquare_randomForest <- petersenlab::accuracyOverall(
+accuracyIndices_randomForest <- petersenlab::accuracyOverall(
   predicted = df$pred,
@@
   dropUndefined = TRUE
-)$rsquared %>% round(., digits = 2)
+)
+
+rsquare_randomForest <- round(accuracyIndices_randomForest$rsquared, digits = 2)
+me_randomForest <- round(accuracyIndices_randomForest$ME, digits = 2)
+mae_randomForest <- round(accuracyIndices_randomForest$MAE, digits = 2)
 ```
@@
 
-rsquare_treeBoosting <- petersenlab::accuracyOverall(
+accuracyIndices_treeBoosting <- petersenlab::accuracyOverall(
   predicted = predictedVsActual$predictedPoints,
@@
   dropUndefined = TRUE
-)$rsquared %>% round(., digits = 2)
+)
+
+rsquare_treeBoosting <- round(accuracyIndices_treeBoosting$rsquared, digits = 2)
+me_treeBoosting <- round(accuracyIndices_treeBoosting$ME, digits = 2)
+mae_treeBoosting <- round(accuracyIndices_treeBoosting$MAE, digits = 2)
 ```
@@
     rsquare_treeBoosting
-  )
-)
-```
-
+  ),
+  me = c(
+    me_onePredictor,
+    me_multiplePredictors,
+    me_lasso,
+    me_ridge,
+    me_elasticNet,
+    me_randomForest,
+    me_treeBoosting
+  ),
+  mae = c(
+    mae_onePredictor,
+    mae_multiplePredictors,
+    mae_lasso,
+    mae_ridge,
+    mae_elasticNet,
+    mae_randomForest,
+    mae_treeBoosting
+  )
+)
+
+rsquare_maximum <- max(
+  c(rsquare_onePredictor, rsquare_multiplePredictors,
+    rsquare_lasso, rsquare_ridge, rsquare_elasticNet,
+    rsquare_randomForest, rsquare_treeBoosting),
+  na.rm = TRUE
+)
+```
+
+We examined multiple machine learning models to identify the model that generates the most accurate prediction of Quarterbacks' future fantasy points.
 A summary of the accuracy of the machine learning models is in @tbl-machineLearningModelSummary.
@@
   machineLearningTable,
-  col.names = c("Model","$R^2$"),
+  col.names = c("Model","$R^2$","Mean Error","Mean Absolute Error"),
   booktabs = TRUE)
@@
 
+The most accurate model had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using [regression with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
+The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
+This suggests that the other predictors examined did not have [incremental validity](#sec-incrementalValidity)—they did not show utility above and beyond a player's fantasy points in predicting their future fantasy points.
+If anything, considering the additional predictor variables showed *decremental* validity—the inclusion of the additional variables worsened prediction, possibly by weakening the predictive association of the previous season's fantasy points and misattributing its predictive effect to other, related predictors in the model.
+
+Among the machine learning models, the [random forest model](#sec-randomForest) was most accurate.
+However, it was still less accurate than a [simple regression model with one predictor](#sec-regressionOnePredictor).
+This is consistent with the common finding that simple models (and parsimony) are often preferable.
+As an example, @Youngstrom2018 found that simple models did just as well and in some cases better than [LASSO models](#sec-lasso) in classifying bipolar disorder.
+
+The most accurate of models all explained less than half of the variance in players' future fantasy points.
+Thus, there remains considerable variance to be explained.
+The considerable unexplained variance suggests that we are missing key variables that are important predictors of future fantasy performance.
+It is also consistent with the conclusion from many other domains in psychology that human behavior is challenging to predict [@PetersenPrinciplesPsychAssessment].
+
 ## Conclusion {#sec-machineLearningConclusion}

### 2025-07-07 machine-learning.qmd
20250707 - additional summary

@@
 A learning rate of 1 is very high for boosting.
-Even if a learning rate of 1 did well in tuning, I use a lower learning rate (0.1) to avoid overfitting.
-I also added some light regularization (`lambda_l2`) for better generalization.
-I also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^5 (32).
-I set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 10.
+Even if a learning rate of 1 did well in tuning, we use a lower learning rate (0.1) to avoid overfitting.
+We also added some light regularization (`lambda_l2`) for better generalization.
+We also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^5 (32).
+We set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 10.
 
@@
 )
+
+mae_minimum <- min(
+  c(mae_onePredictor, mae_multiplePredictors,
+    mae_lasso, mae_ridge, mae_elasticNet,
+    mae_randomForest, mae_treeBoosting),
+  na.rm = TRUE
+)
 ```
@@
 
-The most accurate model had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using [regression with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
-The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
+The most accurate model, in terms of $R^2$, had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using [regression with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
+The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
 This suggests that the other predictors examined did not have [incremental validity](#sec-incrementalValidity)—they did not show utility above and beyond a player's fantasy points in predicting their future fantasy points.
@@
 
-Among the machine learning models, the [random forest model](#sec-randomForest) was most accurate.
+Among the machine learning models, the [random forest model](#sec-randomForest) was most accurate according to $R^2$.
 However, it was still less accurate than a [simple regression model with one predictor](#sec-regressionOnePredictor).
@@
 As an example, @Youngstrom2018 found that simple models did just as well and in some cases better than [LASSO models](#sec-lasso) in classifying bipolar disorder.
+
+The most accurate model in terms of [mean absolute error](#sec-meanAbsoluteError) was the [tree-boosting model](#sec-treeBoosting), which had a [mean absolute error](#sec-meanAbsoluteError) of `{r} mae_minimum`.
+However, even in the most accurate model, the predicted values were over `{r} petersenlab::apa(mae_minimum, decimals = 0)` points away from the actual values, suggesting that the model-predicted values were not particularly accurate.
+In sum, the [regression model with one predictor](#sec-regressionOnePredictor) was the most accurate in terms of differentiating between players ($R^2$), whereas the [tree-boosting model](#sec-treeBoosting) was the most accurate in terms of how close the predicted values were to the actual values.
 

### 2025-07-08 pca.qmd
20250707 - PCA example

@@
 ```{r}
+library("psych")
+library("nFactors")
+library("tidyverse")
+```
 
+### Load Data {#sec-pcaLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_nextGenStats_weekly.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/nfl_nextGenStats_weekly.RData")
+```
+
+### Prepare Data {#sec-pcaPrepareData}
+
+#### Merge Data {#sec-pcaPrepareDataMerge}
+
+```{r}
+dataMerged <- full_join(
+  player_stats_weekly,
+  nfl_nextGenStats_weekly %>% select(-any_of(c("player_display_name","completions","attempts","receptions","targets"))),
+  by = c("player_id" = "player_gsis_id","season","season_type","week"),
+)
+```
+
+#### Specify Variables {#sec-pcaPrepareDataSpecifyVars}
+
+```{r}
+pcaVars <- c(
+  "completions","attempts","passing_yards","passing_tds","passing_interceptions",
+  "sacks_suffered","sack_yards_lost","sack_fumbles","sack_fumbles_lost",
+  "passing_air_yards","passing_yards_after_catch","passing_first_downs",
+  "passing_epa","passing_cpoe","passing_2pt_conversions","pacr","pass_40_yds",
+  "pass_inc","pass_comp_pct","fumbles","two_pts",
+  "avg_time_to_throw","avg_completed_air_yards","avg_intended_air_yards",
+  "avg_air_yards_differential","aggressiveness","max_completed_air_distance",
+  "avg_air_yards_to_sticks","passer_rating", #,"completion_percentage"
+  "expected_completion_percentage","completion_percentage_above_expectation",
+  "avg_air_distance","max_air_distance")
+```
+
+#### Standardize Variables {#sec-pcaPrepareDataStandardize}
+
+```{r}
+dataForPCA <- dataMerged
+dataForPCA[pcaVars] <- scale(dataForPCA[pcaVars])
 ```
@@
 
+## Example of Principal Component Analysis {#sec-pcaExample}
+
+```{r}
+#| label: fig-pcaScreePlot1
+#| fig-cap: "Scree Plot."
+#| fig-alt: "Scree Plot."
+
+psych::fa.parallel(
+  x = dataForPCA[pcaVars],
+  fa = "pc"
+)
+```
+
+```{r}
+#| label: fig-pcaScreePlot2
+#| fig-cap: "Scree Plot."
+#| fig-alt: "Scree Plot."
+
+screeDataPCA <- nFactors::nScree(
+  x = cor(
+    dataForPCA[pcaVars],
+    use = "pairwise.complete.obs"),
+  model = "components")
+
+nFactors::plotnScree(screeDataPCA)
+```
+
+```{r}
+psych::vss(
+  dataForPCA[pcaVars],
+  rotate = "oblimin",
+  fm = "pc")
+```
+
+```{r}
+pca1ComponentOblique <- psych::principal(
+  dataForPCA[pcaVars],
+  nfactors = 1,
+  rotate = "oblimin")
+
+pca2ComponentOblique <- psych::principal(
+  dataForPCA[pcaVars],
+  nfactors = 2,
+  rotate = "oblimin")
+
+pca3ComponentOblique <- psych::principal(
+  dataForPCA[pcaVars],
+  nfactors = 3,
+  rotate = "oblimin")
+
+pca4ComponentOblique <- psych::principal(
+  dataForPCA[pcaVars],
+  nfactors = 4,
+  rotate = "oblimin")
+
+pca5ComponentOblique <- psych::principal(
+  dataForPCA[pcaVars],
+  nfactors = 5,
+  rotate = "oblimin")
+
+pca6ComponentOblique <- psych::principal(
+  dataForPCA[pcaVars],
+  nfactors = 6,
+  rotate = "oblimin")
+
+pca7ComponentOblique <- psych::principal(
+  dataForPCA[pcaVars],
+  nfactors = 7,
+  rotate = "oblimin")
+
+pca8ComponentOblique <- psych::principal(
+  dataForPCA[pcaVars],
+  nfactors = 8,
+  rotate = "oblimin")
+
+pca9ComponentOblique <- psych::principal(
+  dataForPCA[pcaVars],
+  nfactors = 9,
+  rotate = "oblimin")
+```
+
+```{r}
+pca1ComponentOblique
+pca2ComponentOblique
+pca3ComponentOblique
+pca4ComponentOblique
+pca5ComponentOblique
+pca6ComponentOblique
+pca7ComponentOblique
+pca8ComponentOblique
+pca9ComponentOblique
+```
+
+Here are the variables that had a standardized component loading of 0.4 or greater on each component:
+
+```{r}
+component1vars <- c(
+  "completions","attempts","passing_yards","passing_tds","passing_interceptions",
+  "sacks_suffered","sack_yards_lost","sack_fumbles","sack_fumbles_lost",
+  "passing_air_yards","passing_yards_after_catch","passing_first_downs",
+  #"passing_epa","passing_cpoe","passing_2pt_conversions","pacr","pass_40_yds",
+  "pass_inc","fumbles")#,"two_pts","pass_comp_pct",
+  #"avg_time_to_throw","avg_completed_air_yards","avg_intended_air_yards",
+  #"avg_air_yards_differential","aggressiveness","max_completed_air_distance",
+  #"avg_air_yards_to_sticks","passer_rating", #,"completion_percentage"
+  #"expected_completion_percentage","completion_percentage_above_expectation",
+  #"avg_air_distance","max_air_distance")
+
+component2vars <- c(
+  #"completions","attempts","passing_yards","passing_tds","passing_interceptions",
+  #"sacks_suffered","sack_yards_lost","sack_fumbles","sack_fumbles_lost",
+  "passing_air_yards",#"passing_yards_after_catch","passing_first_downs",
+  "pacr",#"passing_epa","passing_cpoe","passing_2pt_conversions","pass_40_yds",
+  #"pass_inc","pass_comp_pct","fumbles","two_pts",
+  "avg_completed_air_yards","avg_intended_air_yards",#"avg_time_to_throw",
+  "aggressiveness","max_completed_air_distance",#"avg_air_yards_differential",
+  "avg_air_yards_to_sticks",#"passer_rating", #,"completion_percentage"
+  "expected_completion_percentage",#"completion_percentage_above_expectation",
+  "avg_air_distance","max_air_distance")
+
+component3vars <- c(
+  "passing_tds",#"completions","attempts","passing_yards","passing_interceptions",
+  #"sacks_suffered","sack_yards_lost","sack_fumbles","sack_fumbles_lost",
+  #"passing_air_yards","passing_yards_after_catch","passing_first_downs",
+  "passing_epa","passing_cpoe","pass_40_yds",#"passing_2pt_conversions","pacr",
+  "pass_comp_pct",#"fumbles","two_pts","pass_inc",
+  "avg_air_yards_differential","max_completed_air_distance",
+  "avg_time_to_throw","avg_completed_air_yards","avg_intended_air_yards",
+  #"aggressiveness",
+  "passer_rating",#"avg_air_yards_to_sticks", #,"completion_percentage"
+  "completion_percentage_above_expectation")#,"expected_completion_percentage",
+  #"avg_air_distance","max_air_distance")
+```
+
+```{r}
+pca3ComponentOblique$scores %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+pca3ComponentOblique$scores[which.max(pca3ComponentOblique$scores[,1]),]
+dataMerged[which.max(pca3ComponentOblique$scores[,1]), c("player_display_name", "season", "week", component1vars)]
+
+pca3ComponentOblique$scores[which.min(pca3ComponentOblique$scores[,1]),]
+dataMerged[which.min(pca3ComponentOblique$scores[,1]), c("player_display_name", "season", "week", component1vars)]
+```
+
+```{r}
+pca3ComponentOblique$scores[which.max(pca3ComponentOblique$scores[,2]),]
+dataMerged[which.max(pca3ComponentOblique$scores[,2]), c("player_display_name", "season", "week", component2vars)]
+
+pca3ComponentOblique$scores[which.min(pca3ComponentOblique$scores[,2]),]
+dataMerged[which.min(pca3ComponentOblique$scores[,2]), c("player_display_name", "season", "week", component2vars)]
+```
+
+```{r}
+pca3ComponentOblique$scores[which.max(pca3ComponentOblique$scores[,3]),]
+dataMerged[which.max(pca3ComponentOblique$scores[,3]), c("player_display_name", "season", "week", component3vars)]
+
+pca3ComponentOblique$scores[which.min(pca3ComponentOblique$scores[,3]),]
+dataMerged[which.min(pca3ComponentOblique$scores[,3]), c("player_display_name", "season", "week", component3vars)]
+```
+
 ## Conclusion {#sec-pcaConclusion}

### 2025-07-08 index.qmd
20250707 - additional resources
20250707 - PCA fixes and citations

@@
   "fitdistrplus", "sn", "mclust", "magrittr", "viridis", "plotly", "quantmod", 
-  "fPortfolio", "NMOF", "xts", "zoo", "forecast", "stringi", "parallel", #"webshot2"
-  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", "caret", 
-  "tidymodels", "LongituRF", "gpboost", "cmdstanr", "learnr", "remotes", 
-  "corrplot", "correlation", "mgcv", "rms", "car", "performance", "lavaan", 
-  "mice", "miceadds", "interactions", "shiny", "PearsonDS", "robustbase", 
-  "ordinal", "MASS", "data.table", "future", "future.apply", "QuantPsyc", 
-  "SimDesign", "effectsize"
+  "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast", "stringi", #"webshot2"
+  "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
+  "caret", "tidymodels", "LongituRF", "gpboost", "cmdstanr", "learnr", 
+  "remotes", "corrplot", "correlation", "mgcv", "rms", "car", "performance", 
+  "lavaan", "mice", "miceadds", "interactions", "shiny", "PearsonDS", 
+  "robustbase", "ordinal", "MASS", "data.table", "future", "future.apply", 
+  "QuantPsyc", "SimDesign", "effectsize"
 )
@@
 
+## Example Syllabus {#sec-exampleSyllabus}
+
+If you are interested in teaching a course using this textbook, here is an example syllabus based on my class: <https://isaactpetersen.github.io/psy-3170-2025-fall/syllabus.html>
+
 ## Acknowledgments {#sec-acknowledgments}

### 2025-07-08 pca.qmd
20250707 - PCA fixes and citations

@@
 
+We generated the scree plot in @fig-pcaScreePlot1 using the `fa.parallel()` function of the `psych` package [@R-psych].
+
 ```{r}
@@
 ```
+
+We generated the scree plot in @fig-pcaScreePlot2 using the `nScree()` and `plotnScree()` functions of the `nFactors` package [@R-nFactors].
 
@@
 
+We generated the very simple structure (VSS) plot in @fig-pcaVSSPlot using the `vss()` function of the `psych` package [@R-psych].
+
 ```{r}
+#| label: fig-pcaVSSPlot
+#| fig-cap: "Very Simple Structure Plot."
+#| fig-alt: "Very Simple Structure Plot."
+
 psych::vss(

### 2025-07-08 factor-analysis.qmd
20250228 - remove duplicate label
20250618 - fix formulas
20250707 - updates

@@
 ```{r}
+library("psych")
+library("nFactors")
+library("lavaan")
+library("lavaangui")
+library("tidyverse")
+```
 
+### Load Data {#sec-factorAnalysisLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_nextGenStats_weekly.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/nfl_nextGenStats_weekly.RData")
+```
+
+### Prepare Data {#sec-factorAnalysisPrepareData}
+
+#### Merge Data {#sec-factorAnalysisPrepareDataMerge}
+
+```{r}
+dataMerged <- full_join(
+  player_stats_weekly,
+  nfl_nextGenStats_weekly %>% select(-any_of(c("player_display_name","completions","attempts","receptions","targets"))),
+  by = c("player_id" = "player_gsis_id","season","season_type","week"),
+)
+```
+
+#### Specify Variables {#sec-pcaPrepareDataSpecifyVars}
+
+```{r}
+faVars <- c(
+  "completions","attempts","passing_yards","passing_tds","passing_interceptions",
+  "sacks_suffered","sack_yards_lost","sack_fumbles","sack_fumbles_lost",
+  "passing_air_yards","passing_yards_after_catch","passing_first_downs",
+  "passing_epa","passing_cpoe","passing_2pt_conversions","pacr","pass_40_yds",
+  "pass_inc","pass_comp_pct","fumbles","two_pts",
+  "avg_time_to_throw","avg_completed_air_yards","avg_intended_air_yards",
+  "avg_air_yards_differential","aggressiveness","max_completed_air_distance",
+  "avg_air_yards_to_sticks","passer_rating", #,"completion_percentage"
+  "expected_completion_percentage","completion_percentage_above_expectation",
+  "avg_air_distance","max_air_distance")
 ```
@@
 $$
-\bar{X} = \lambda \cdot \text{F} + \text{Item Intercept} + \text{Error Term}
+X = \lambda \cdot \text{F} + \text{Item Intercept} + \text{Error Term}
 $$ {#eq-indicatorLatentAssociation}
@@
 
-## Exploratory Factor Analysis {#sec-cfa}
+## Exploratory Factor Analysis {#sec-efa}
 
@@
 
+### Example {#sec-efaExample}
+
+We generated the scree plot in @fig-efaScreePlot1 using the `fa.parallel()` function of the `psych` package [@R-psych].
+
+```{r}
+#| label: fig-efaScreePlot1
+#| fig-cap: "Scree Plot."
+#| fig-alt: "Scree Plot."
+
+psych::fa.parallel(
+  x = dataMerged[faVars],
+  fm = "minres", # errors out when using "ml"
+  fa = "fa"
+)
+```
+
+We generated the scree plot in @fig-efaScreePlot2 using the `nScree()` and `plotnScree()` functions of the `nFactors` package [@R-nFactors].
+
+```{r}
+#| label: fig-efaScreePlot2
+#| fig-cap: "Scree Plot."
+#| fig-alt: "Scree Plot."
+
+screeDataEFA <- nFactors::nScree(
+  x = cor(
+    dataMerged[faVars],
+    use = "pairwise.complete.obs"),
+  model = "factors")
+
+nFactors::plotnScree(screeDataEFA)
+```
+
+We generated the very simple structure (VSS) plots in Figures [-@fig-efaVSSPlot1] and [-@fig-efaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
+
+```{r}
+#| label: fig-efaVSSPlot1
+#| fig-cap: "Very Simple Structure Plot."
+#| fig-alt: "Very Simple Structure Plot."
+
+psych::vss(
+  dataMerged[faVars],
+  rotate = "oblimin",
+  fm = "minres") # errors out when using "mle"
+```
+
+```{r}
+#| label: fig-efaVSSPlot2
+#| fig-cap: "Very Simple Structure Plot."
+#| fig-alt: "Very Simple Structure Plot."
+
+psych::nfactors(
+  dataMerged[faVars],
+  rotate = "oblimin",
+  fm = "minres") # errors out when using "ml"
+```
+
+We fit EFA models using the `efa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
+
+```{r}
+#| eval: false
+
+efaOblique_fit <- lavaan::efa(
+  data = dataMerged,
+  ov.names = faVars,
+  nfactors = 1:9,
+  rotation = "geomin",
+  missing = "ML",
+  estimator = "MLR",
+  meanstructure = TRUE
+)
+```
+
+A path diagram of the EFA model was created using the `plot_lavaan()` function of the `lavaangui` package [@R-lavaangui; @Karch2025_packages].
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(efaOblique_fit)
+```
+
 ## Confirmatory Factor Analysis {#sec-cfa}
@@
 In CFA, cross-loadings are not estimated unless the researcher specifies them.\index{cross-loading}\index{factor analysis!confirmatory}
+
+### Example {#sec-cfaExample}
+
+We fit CFA models using the `cfa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
+
+```{r}
+cfaModel1_syntax <- '
+ #Factor loadings
+ F1 =~ completions + attempts + passing_yards + passing_tds + passing_interceptions + sacks_suffered + sack_yards_lost + sack_fumbles + sack_fumbles_lost + passing_air_yards + passing_yards_after_catch + passing_first_downs + pass_inc + fumbles
+ F2 =~ passing_air_yards + pacr + avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + avg_air_yards_to_sticks + expected_completion_percentage + avg_air_distance + max_air_distance
+ F3 =~ passing_tds + passing_epa + passing_cpoe + pass_40_yds + pass_comp_pct + avg_air_yards_differential + max_completed_air_distance + avg_time_to_throw + avg_completed_air_yards + avg_intended_air_yards + passer_rating + completion_percentage_above_expectation
+'
+```
+
+```{r}
+#| eval: false
+
+cfaModel1_fit <- lavaan::cfa(
+  cfaModel1_syntax,
+  data = dataMerged,
+  missing = "ML",
+  estimator = "MLR",
+  std.lv = TRUE)
+```
+
+```{r}
+#| eval: false
+
+summary(
+  cfaModel1_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+```{r}
+#| eval: false
+
+lavaan::fitMeasures(
+  cfaModel1_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+#| eval: false
+
+lavaan::residuals(
+  cfaModel1_fit,
+  type = "cor")
+```
+
+```{r}
+#| eval: false
+
+lavaan::modificationindices(
+  cfaModel1_fit,
+  sort. = TRUE)
+```
+
+```{r}
+#| eval: false
+
+lavaan::lavPredict(cfaModel1_fit)
+```
+
+A path diagram of the CFA model was created using the `plot_lavaan()` function of the `lavaangui` package [@R-lavaangui; @Karch2025_packages].
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(cfaModel1_fit)
+```
 

### 2025-07-08 index.qmd
20250707 - updates

@@
   "remotes", "corrplot", "correlation", "mgcv", "rms", "car", "performance", 
-  "lavaan", "mice", "miceadds", "interactions", "shiny", "PearsonDS", 
-  "robustbase", "ordinal", "MASS", "data.table", "future", "future.apply", 
-  "QuantPsyc", "SimDesign", "effectsize"
+  "lavaan", "lavaangui", "mice", "miceadds", "interactions", "shiny", 
+  "PearsonDS", "robustbase", "ordinal", "MASS", "data.table", "future", 
+  "future.apply", "QuantPsyc", "SimDesign", "effectsize"
 )

### 2025-07-08 factor-analysis.qmd
20250708 - tweaks

@@
 #| label: fig-efaScreePlot1
-#| fig-cap: "Scree Plot."
-#| fig-alt: "Scree Plot."
+#| fig-cap: "Scree Plot: With Comparisons to Simulated and Resampled Data."
+#| fig-alt: "Scree Plot: With Comparisons to Simulated and Resampled Data."
 
@@
 #| label: fig-efaScreePlot2
-#| fig-cap: "Scree Plot."
-#| fig-alt: "Scree Plot."
+#| fig-cap: "Scree Plot with Parallel Analysis."
+#| fig-alt: "Scree Plot with Parallel Analysis."
 
@@
 #| label: fig-efaVSSPlot2
-#| fig-cap: "Very Simple Structure Plot."
-#| fig-alt: "Very Simple Structure Plot."
+#| fig-cap: "Model Indices by Number of Factors."
+#| fig-alt: "Model Indices by Number of Factors."
 

### 2025-07-08 pca.qmd
20250708 - tweaks

@@
 #| label: fig-pcaScreePlot1
-#| fig-cap: "Scree Plot."
-#| fig-alt: "Scree Plot."
+#| fig-cap: "Scree Plot: With Comparisons to Simulated and Resampled Data."
+#| fig-alt: "Scree Plot: With Comparisons to Simulated and Resampled Data."
 
@@
 #| label: fig-pcaScreePlot2
-#| fig-cap: "Scree Plot."
-#| fig-alt: "Scree Plot."
+#| fig-cap: "Scree Plot with Parallel Analysis."
+#| fig-alt: "Scree Plot with Parallel Analysis."
 
@@
 
-We generated the very simple structure (VSS) plot in @fig-pcaVSSPlot using the `vss()` function of the `psych` package [@R-psych].
+We generated the very simple structure (VSS) plots in Figures [-@fig-pcaVSSPlot1] and [-@fig-pcaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
 
 ```{r}
-#| label: fig-pcaVSSPlot
+#| label: fig-pcaVSSPlot1
 #| fig-cap: "Very Simple Structure Plot."
@@
 psych::vss(
+  dataForPCA[pcaVars],
+  rotate = "oblimin",
+  fm = "pc")
+```
+
+```{r}
+#| label: fig-pcaVSSPlot2
+#| fig-cap: "Model Indices by Number of Components."
+#| fig-alt: "Model Indices by Number of Components."
+
+psych::nfactors(
   dataForPCA[pcaVars],

### 2025-07-08 factor-analysis.qmd
20250708 - tweaks

@@
   "avg_air_distance","max_air_distance")
+```
+
+#### Standardize Variables {#sec-pcaPrepareDataStandardize}
+
+Although standardizing variables is not strictly necessary in factor analysis, it can be helpful to prevent some variables from having considerably larger variances than others.
+
+```{r}
+dataForFA <- dataMerged
+dataForFA[faVars] <- scale(dataForFA[faVars])
 ```
@@
 psych::fa.parallel(
-  x = dataMerged[faVars],
+  x = dataForFA[faVars],
   fm = "minres", # errors out when using "ml"
@@
   x = cor(
-    dataMerged[faVars],
+    dataForFA[faVars],
     use = "pairwise.complete.obs"),
@@
 psych::vss(
-  dataMerged[faVars],
+  dataForFA[faVars],
   rotate = "oblimin",
@@
 psych::nfactors(
-  dataMerged[faVars],
+  dataForFA[faVars],
   rotate = "oblimin",
@@
 efaOblique_fit <- lavaan::efa(
-  data = dataMerged,
+  data = dataForFA,
   ov.names = faVars,
@@
   estimator = "MLR",
+  cluster = "player_id",
   meanstructure = TRUE
@@
   cfaModel1_syntax,
-  data = dataMerged,
+  data = dataForFA,
   missing = "ML",
   estimator = "MLR",
-  std.lv = TRUE)
+  std.lv = TRUE,
+  cluster = "player_id",
+  em.iter.max = 2000000)
 ```

### 2025-07-08 machine-learning.qmd
20250708 - tweaks

@@
 It is also consistent with the conclusion from many other domains in psychology that human behavior is challenging to predict [@PetersenPrinciplesPsychAssessment].
+We could likely improve our predictive accuracy by including projections, but it is unclear what information projections are or are not based on.
+For instance, if projections already account for the prior season's statistics and fantasy points, then including both may not buy you much in terms of increased predictive accuracy.
 

### 2025-07-08 _chunk-timing.qmd
20250708 - track chunk execution times

```{r}
#| label: setup-timing
#| include: false

# Set knitr options globally for all chunks
knitr::opts_chunk$set(time_it = TRUE)

# Get chapter-specific log file name
current_file <- knitr::current_input()
log_file <- sprintf("chunk-times-%s.txt", tools::file_path_sans_ext(basename(current_file)))

# Clear previous log
if (file.exists(log_file)) file.remove(log_file)

# Create timing hook
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    label <- if (!is.null(options$label)) options$label else "<unnamed>"
    if (before) {
      now <<- Sys.time()
    } else {
      elapsed <- difftime(Sys.time(), now, units = "secs")
      msg <- sprintf("This chunk (`%s`) ran in %.2f seconds.", label, as.numeric(elapsed))

      # Print to GitHub Actions log
      message(msg)

      # Save to chapter-specific log file
      cat(msg, file = log_file, append = TRUE, sep = "\n")
    }
  }
}))
```

### 2025-07-08 index.qmd
20250708 - track chunk execution times

@@
-```{r}
-#| label: setup
-#| include: false
-
-# Global knitr options
-knitr::opts_chunk$set(
-  time = TRUE,
-  message = NA # Show in console, not in rendered output
-)
-
-# Null coalescing operator
-`%||%` <- function(a, b) if (!is.null(a)) a else b
-
-# Log file setup
-log_file <- "chunk-times.txt"
-if (file.exists(log_file)) file.remove(log_file)
-
-# Environment to track start times
-.chunk_times_env <- new.env()
-
-# Timing hook
-knitr::knit_hooks$set(time = function(before, options) {
-  label <- options$label %||% "<unnamed>"
-
-  if (before) {
-    .chunk_times_env[[label]] <- Sys.time()
-  } else {
-    start_time <- .chunk_times_env[[label]]
-    if (!is.null(start_time)) {
-      elapsed <- difftime(Sys.time(), start_time, units = "secs")
-      msg <- sprintf("Chunk `%s` ran in %.2f seconds.", label, as.numeric(elapsed))
-
-      # Send message to GitHub Actions log
-      message(sprintf("::notice title=Chunk Time::%s", msg))
-
-      # Also save to file
-      cat(msg, file = log_file, append = TRUE, sep = "\n")
-    }
-  }
-})
-```
+{{< include _chunk-timing.qmd >}}
 

### 2025-07-08 evaluating-prediction-accuracy.qmd
20250703 - fix heading hierarchy (https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues/7)
To address issue https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues/7
20250708 - track chunk execution times
20250708 - packages and citations

@@
+{{< include _chunk-timing.qmd >}}
+
 # Evaluation of Prediction/Forecasting Accuracy {#sec-predictionAccuracy}
@@
 
-### Weekly Projections {#sec-predictionAccuracyPrepareDataWeekly}
+#### Weekly Projections {#sec-predictionAccuracyPrepareDataWeekly}
 
@@
 
+We calculate the 95% prediction interval based on a LOESS model, which allows nonlinear associations, using the `loess.sd()` function of the `msir` package [@R-msir; @Scrucca2011_packages].
+
 ```{r}
@@
 The diagonal black line is the reference line representing perfect calibration.
-The yellow line is the best-fit line based on LOESS.
+The yellow line is the best-fit line based on a locally estimated scatterplot smoothing (LOESS) model.
 The purple band is the 95% confidence interval of actual fantasy points.
@@
 The purple band is the 95% confidence interval of actual fantasy points.
-The dashed red lines are the 95% prediction interval of actual fantasy points."
+The dashed red lines are the 95% prediction interval of actual fantasy points based on a locally estimated scatterplot smoothing (LOESS) model."
 #| fig-alt: "Calibration Plot for Weekly Projections.
@@
 The purple band is the 95% confidence interval of actual fantasy points.
-The dashed red lines are the 95% prediction interval of actual fantasy points."
+The dashed red lines are the 95% prediction interval of actual fantasy points based on a locally estimated scatterplot smoothing (LOESS) model."
 #| fig-width: 8
@@
 The purple band is the 95% confidence interval of actual fantasy points.
-The dashed red lines are the 95% prediction interval of actual fantasy points."
+The dashed red lines are the 95% prediction interval of actual fantasy points based on a locally estimated scatterplot smoothing (LOESS) model."
 #| fig-alt: "Calibration Plot for Weekly Projections of Quarterbacks.
@@
 The purple band is the 95% confidence interval of actual fantasy points.
-The dashed red lines are the 95% prediction interval of actual fantasy points."
+The dashed red lines are the 95% prediction interval of actual fantasy points based on a locally estimated scatterplot smoothing (LOESS) model."
 #| fig-alt: "Calibration Plot for Weekly Projections of Running Backs.
@@
 The purple band is the 95% confidence interval of actual fantasy points.
-The dashed red lines are the 95% prediction interval of actual fantasy points."
+The dashed red lines are the 95% prediction interval of actual fantasy points based on a locally estimated scatterplot smoothing (LOESS) model."
 #| fig-alt: "Calibration Plot for Weekly Projections of Wide Receivers.
@@
 The purple band is the 95% confidence interval of actual fantasy points.
-The dashed red lines are the 95% prediction interval of actual fantasy points."
+The dashed red lines are the 95% prediction interval of actual fantasy points based on a locally estimated scatterplot smoothing (LOESS) model."
 #| fig-alt: "Calibration Plot for Weekly Projections of Tight Ends.
@@
 The purple band is the 95% confidence interval of actual fantasy points.
-The dashed red lines are the 95% prediction interval of actual fantasy points."
+The dashed red lines are the 95% prediction interval of actual fantasy points based on a locally estimated scatterplot smoothing (LOESS) model."
 #| fig-alt: "Calibration Plot for Weekly Projections of Kickers.
@@
 The purple band is the 95% confidence interval of actual fantasy points.
-The dashed red lines are the 95% prediction interval of actual fantasy points."
+The dashed red lines are the 95% prediction interval of actual fantasy points based on a locally estimated scatterplot smoothing (LOESS) model."
 #| fig-alt: "Calibration Plot for Weekly Projections of Defensive Linemen.
@@
 The purple band is the 95% confidence interval of actual fantasy points.
-The dashed red lines are the 95% prediction interval of actual fantasy points."
+The dashed red lines are the 95% prediction interval of actual fantasy points based on a locally estimated scatterplot smoothing (LOESS) model."
 #| fig-alt: "Calibration Plot for Weekly Projections of Linebackers.
@@
 The purple band is the 95% confidence interval of actual fantasy points.
-The dashed red lines are the 95% prediction interval of actual fantasy points."
+The dashed red lines are the 95% prediction interval of actual fantasy points based on a locally estimated scatterplot smoothing (LOESS) model."
 #| fig-alt: "Calibration Plot for Weekly Projections of Defensive Backs.

### 2025-07-08 getting-started.qmd
20250708 - track chunk execution times
20250708 - packages and citations

@@
+{{< include _chunk-timing.qmd >}}
+
 # Getting Started with `R` for Data Analysis {#sec-gettingStarted}
@@
 
-   install.packages(
-     c("petersenlab","remotes","knitr","rmarkdown","dplyr",
-       "ggplot2","tidyverse","psych","xfun","downlit","xml2",
-       "nflreadr","nflplotR","nflfastR","nflseedR","nfl4th",
-       "gsisdecoder","progressr","DescTools","pwr","pwrss",
-       "WebPower","XICOR","dagitty","ggdag","lubridate",
-       "ggtext","gghighlight","ggExtra","grid","patchwork",
-       "pROC","lme4","lmerTest","MuMIn","emmeans","pbkrtest",
-       "sjstats","AICcmodavg","rstan","brms","bbmle",
-       "fitdistrplus","mclust","magrittr","viridis","viridisLite",
-       "msir","plotly","webshot2","quantmod","fPortfolio",
-       "xts","zoo","forecast","stringi","parallel","parallelly",
-       "doParallel","missRanger","powerjoin","caret","gpboost",
-       "ggridges","corrplot","rms","car","caret","performance",
-       "correlation","interactions","lavaan","mice","miceadds",
-       "robustbase","see"))
+   install.packages(c(
+     "petersenlab","remotes","knitr","rmarkdown","dplyr","ggplot2","tidyverse",
+     "broom","broom.mixed","psych","downlit","xml2","nflreadr","nflplotR",
+     "nflfastR","nflseedR","nfl4th","gsisdecoder","progressr","DescTools",
+     "pwr","pwrss","WebPower","XICOR","dagitty","ggdag","lubridate","ggtext",
+     "gghighlight","ggExtra","grid","patchwork","pROC", "lme4","lmerTest",
+     "MuMIn","emmeans","pbkrtest","sjstats","AICcmodavg","rstan","brms",
+     "tidybayes","bbmle","fitdistrplus","sn","mclust","magrittr","viridis",
+     "viridisLite","msir","plotly","webshot2","quantmod","fPortfolio","NMOF",
+     "nFactors","xts","zoo","forecast","stringi","parallelly","doParallel",
+     "missRanger","ggridges","powerjoin","caret","tidymodels","LongituRF",
+     "gpboost","corrplot","correlation","mgcv","rms","car","performance",
+     "lavaan","lavaangui","mice","miceadds","interactions","robustbase",
+     "ordinal","MASS","data.table","future","future.apply","see","SimDesign",
+     "effectsize"))
    ```

### 2025-07-08 index.qmd
20250708 - packages and citations

@@
   "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "broom", "broom.mixed", 
-  "downlit", "xml2", "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", 
-  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
-  "dagitty", "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", 
-  "patchwork", "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", 
-  "sjstats", "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", 
-  "fitdistrplus", "sn", "mclust", "magrittr", "viridis", "plotly", "quantmod", 
-  "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast", "stringi", #"webshot2"
-  "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
-  "caret", "tidymodels", "LongituRF", "gpboost", "cmdstanr", "learnr", 
-  "remotes", "corrplot", "correlation", "mgcv", "rms", "car", "performance", 
-  "lavaan", "lavaangui", "mice", "miceadds", "interactions", "shiny", 
-  "PearsonDS", "robustbase", "ordinal", "MASS", "data.table", "future", 
-  "future.apply", "QuantPsyc", "SimDesign", "effectsize"
+  "psych", "downlit", "xml2", "nflreadr", "nflplotR", "nflfastR", "nflseedR", 
+  "nfl4th", "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", 
+  "XICOR", "dagitty", "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", 
+  "grid", "patchwork", "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", 
+  "pbkrtest", "sjstats", "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", 
+  "fitdistrplus", "sn", "mclust", "magrittr", "viridis", "viridisLite", "msir", 
+  "plotly", "webshot2", "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", 
+  "zoo", "forecast", "stringi", "parallelly", "doParallel", "missRanger", 
+  "ggridges", "powerjoin", "caret", "tidymodels", "LongituRF", "gpboost", 
+  "cmdstanr", "corrplot", "correlation", "mgcv", "rms", "car", "performance", 
+  "lavaan", "lavaangui", "mice", "miceadds", "interactions", "robustbase", 
+  "ordinal", "MASS", "data.table", "future", "future.apply", "see", "SimDesign", 
+  "effectsize", 
+  "learnr", "remotes", "shiny", "PearsonDS" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
 )

### 2025-07-08 machine-learning.qmd
20250708 - track chunk execution times
20250708 - packages and citations

@@
+{{< include _chunk-timing.qmd >}}
+
 # Machine Learning {#sec-machineLearning}
@@
 library("petersenlab")
-library("parallel")
 library("future")
@@
 
-num_cores <- parallel::detectCores()
-num_true_cores <- parallel::detectCores(logical = FALSE)
+num_cores <- parallelly::availableCores()
+num_true_cores <- parallelly::availableCores(logical = FALSE)
 ```
@@
 
-num_cores <- parallel::detectCores() - 1
-num_true_cores <- parallel::detectCores(logical = FALSE) - 1
+num_cores <- parallelly::availableCores() - 1
+num_true_cores <- parallelly::availableCores(logical = FALSE) - 1
 ```

### 2025-07-08 mixed-models.qmd
20250701 - tweaks
20250708 - track chunk execution times
20250708 - packages and citations

@@
+{{< include _chunk-timing.qmd >}}
+
 # Mixed Models {#sec-mixedModels}
@@
 library("performance")
-library("parallel")
 library("parallelly")
@@
 
+#### Multicollinearity {#sec-mixedModelsMulticollinearity}
+
+As with [multiple regression](#sec-multipleRegression), [multicollinearity](#sec-multipleRegressionMulticollinearity) is also a potential concern for mixed models.
+Here are the variance inflation factors (VIFs) from an earlier model (VIF is described in @sec-multipleRegressionMulticollinearity):
+
+```{r}
+car::vif(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+```
+
 #### Diagnostic Plots {#sec-fantasyPointsByAgeExperienceModelsDiagnosticPlots}

### 2025-07-08 mixed-models.qmd
20250708 - fix and labels

@@
 ```{r}
-num_cores <- detectCores()
-num_cores
+#| label: parallel-cores-mixed-models
+#| include: false
+
+num_cores <- parallelly::availableCores()
+num_true_cores <- parallelly::availableCores(logical = FALSE)
+```
+
+```{r}
+#| eval: false
+
+num_cores <- parallelly::availableCores() - 1
+num_true_cores <- parallelly::availableCores(logical = FALSE) - 1
+```
+
+```{r}
+#| label: gam-mixed-models
 
@@
 ```{r}
+#| label: players-top-depth-chart-mixed-models
+
 pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience <- lmerTest::lmer(
@@
 AIC(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+```
+
+```{r}
+#| label: gam-players-top-depth-chart-mixed-models
 

### 2025-07-09 factor-analysis.qmd
20250708 - track chunk execution times
20250708 - tweaks

@@
+{{< include _chunk-timing.qmd >}}
+
 # Factor Analysis {#sec-factorAnalysis}
@@
 
+#### Subset Data {#sec-pcaPrepareDataSubset}
+
+```{r}
+dataMerged_subset <- dataMerged %>% 
+  filter(!is.na(player_id))
+```
+
 #### Standardize Variables {#sec-pcaPrepareDataStandardize}
@@
 ```{r}
-dataForFA <- dataMerged
+dataForFA <- dataMerged_subset
 dataForFA[faVars] <- scale(dataForFA[faVars])
@@
   cluster = "player_id",
-  em.iter.max = 2000000)
+  em.h1.iter.max = 2000000)
 ```

### 2025-07-09 intro.qmd
20250708 - track chunk execution times
20250708 - videos and quotes

@@
+{{< include _chunk-timing.qmd >}}
+
 # Introduction {#sec-intro}
@@
 
+::: {.content-visible when-format="html:js"}
+
+::: {#fig-video-motz}
+
+Here is a video of Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football [@NFL2014]:
+
+![Video of Prof. Benjamin Motz Discussing Statistics and His Fantasy Football Course. From: <https://www.facebook.com/watch/?v=10155572257183615>.](images/motzVideo.mp4){width=100%}\
+
+:::
+
+:::
+
 ## Goals {#learningGoals}
@@
 
-"*This material probably won't win you fantasy football championships.
-You could take what we learn and apply it to fantasy football and you might become 5 percent more likely to win.
-Or...
-Consider the broader relevance of this.
-You could learn data analysis and figure out ways to apply it to other systems.
-And you could be making a six-figure salary within the next five years.*"
-– Benjamin Motz, Ph.D.
-
-::: {.content-visible when-format="html:js"}
-
-Here is a video of Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football [@NFL2014]:
-
-![Video of Prof. Benjamin Motz Discussing Statistics and His Fantasy Football Course. From: <https://www.facebook.com/watch/?v=10155572257183615>.](images/motzVideo.mp4){width=100%}\
-
-:::
+> "*This material probably won't win you fantasy football championships.
+> You could take what we learn and apply it to fantasy football and you might become 5 percent more likely to win.
+> Or...
+> Consider the broader relevance of this.
+> You could learn data analysis and figure out ways to apply it to other systems.
+> And you could be making a six-figure salary within the next five years.*" – [Benjamin Motz, Ph.D.](#fig-video-motz) [@NFL2014]

### 2025-07-09 getting-started.qmd
20250709 - consolidate packages using package ecosystems

--- 2025-07-07 getting-started.qmd
@@
 
-   install.packages(
-     c("petersenlab","remotes","knitr","rmarkdown","dplyr",
-       "ggplot2","tidyverse","psych","xfun","downlit","xml2",
-       "nflreadr","nflplotR","nflfastR","nflseedR","nfl4th",
-       "gsisdecoder","progressr","DescTools","pwr","pwrss",
-       "WebPower","XICOR","dagitty","ggdag","lubridate",
-       "ggtext","gghighlight","ggExtra","grid","patchwork",
-       "pROC","lme4","lmerTest","MuMIn","emmeans","pbkrtest",
-       "sjstats","AICcmodavg","rstan","brms","bbmle",
-       "fitdistrplus","mclust","magrittr","viridis","viridisLite",
-       "msir","plotly","webshot2","quantmod","fPortfolio",
-       "xts","zoo","forecast","stringi","parallel","parallelly",
-       "doParallel","missRanger","powerjoin","caret","gpboost",
-       "ggridges","corrplot","rms","car","caret","performance",
-       "correlation","interactions","lavaan","mice","miceadds",
-       "robustbase","see"))
+   install.packages(c(
+     "petersenlab","remotes","knitr","rmarkdown","tidyverse","nflverse",
+     "tidymodels","easystats","broom","broom.mixed","psych","downlit","xml2",
+     "gsisdecoder","progressr","DescTools","pwr","pwrss","WebPower","XICOR",
+     "dagitty","ggdag","ggtext","gghighlight","ggExtra","grid","patchwork",
+     "pROC", "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats",
+     "AICcmodavg","rstan","brms","tidybayes","bbmle","fitdistrplus","sn",
+     "mclust","magrittr","viridis","viridisLite","msir","plotly","webshot2",
+     "quantmod","fPortfolio","NMOF","nFactors","xts","zoo","forecast","stringi",
+     "parallelly","doParallel","missRanger","ggridges","powerjoin","caret",
+     "LongituRF","gpboost","corrplot","mgcv","rms","car","lavaan","lavaangui",
+     "mice","miceadds","interactions","robustbase","ordinal","MASS",
+     "data.table","future","future.apply","SimDesign"))
    ```

### 2025-07-09 index.qmd
20250709 - consolidate packages using package ecosystems

--- 2025-07-08 index.qmd
@@
 
+# Packages to drop: stringi (replace with stringr from tidyverse); caret (replace with tidymodels)
+
 # Automatically create a bib database for R packages
@@
   "petersenlab", "ffanalytics",
-  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "broom", "broom.mixed", 
-  "downlit", "xml2", "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", 
+  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
   "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
-  "dagitty", "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", 
-  "patchwork", "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", 
-  "sjstats", "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", 
-  "fitdistrplus", "sn", "mclust", "magrittr", "viridis", "plotly", "quantmod", 
-  "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast", "stringi", #"webshot2"
-  "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
-  "caret", "tidymodels", "LongituRF", "gpboost", "cmdstanr", "learnr", 
-  "remotes", "corrplot", "correlation", "mgcv", "rms", "car", "performance", 
-  "lavaan", "lavaangui", "mice", "miceadds", "interactions", "shiny", 
-  "PearsonDS", "robustbase", "ordinal", "MASS", "data.table", "future", 
-  "future.apply", "QuantPsyc", "SimDesign", "effectsize"
+  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
+  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
+  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
+  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
+  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
+  "stringi", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
+  "caret", "LongituRF", "gpboost", "cmdstanr", "corrplot", "mgcv", "rms", 
+  "car", "lavaan", "lavaangui", "mice", "miceadds", "interactions", 
+  "robustbase", "ordinal", "MASS", "data.table", "future", "future.apply", 
+  "SimDesign",
+  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
+  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
+  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
+  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
+  "learnr", "remotes", "shiny", "PearsonDS" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
 )

### 2025-07-09 getting-started.qmd
20250709 - r-squared and dominance analysis

--- 2025-07-07 getting-started.qmd
@@
 
-   install.packages(
-     c("petersenlab","remotes","knitr","rmarkdown","dplyr",
-       "ggplot2","tidyverse","psych","xfun","downlit","xml2",
-       "nflreadr","nflplotR","nflfastR","nflseedR","nfl4th",
-       "gsisdecoder","progressr","DescTools","pwr","pwrss",
-       "WebPower","XICOR","dagitty","ggdag","lubridate",
-       "ggtext","gghighlight","ggExtra","grid","patchwork",
-       "pROC","lme4","lmerTest","MuMIn","emmeans","pbkrtest",
-       "sjstats","AICcmodavg","rstan","brms","bbmle",
-       "fitdistrplus","mclust","magrittr","viridis","viridisLite",
-       "msir","plotly","webshot2","quantmod","fPortfolio",
-       "xts","zoo","forecast","stringi","parallel","parallelly",
-       "doParallel","missRanger","powerjoin","caret","gpboost",
-       "ggridges","corrplot","rms","car","caret","performance",
-       "correlation","interactions","lavaan","mice","miceadds",
-       "robustbase","see"))
+   install.packages(c(
+     "petersenlab","remotes","knitr","rmarkdown","tidyverse","nflverse",
+     "tidymodels","easystats","broom","broom.mixed","psych","downlit","xml2",
+     "gsisdecoder","progressr","DescTools","pwr","pwrss","WebPower","XICOR",
+     "dagitty","ggdag","ggtext","gghighlight","ggExtra","grid","patchwork",
+     "pROC", "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats",
+     "AICcmodavg","rstan","brms","tidybayes","bbmle","fitdistrplus","sn",
+     "mclust","magrittr","viridis","viridisLite","msir","plotly","webshot2",
+     "quantmod","fPortfolio","NMOF","nFactors","xts","zoo","forecast","stringi",
+     "parallelly","doParallel","missRanger","ggridges","powerjoin","caret",
+     "LongituRF","gpboost","corrplot","mgcv","rms","car","lavaan","lavaangui",
+     "mice","miceadds","interactions","robustbase","ordinal","MASS",
+     "data.table","future","future.apply","SimDesign","domir"))
    ```

### 2025-07-09 index.qmd
20250709 - r-squared and dominance analysis

--- 2025-07-08 index.qmd
@@
   "petersenlab", "ffanalytics",
-  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "broom", "broom.mixed", 
-  "downlit", "xml2", "nflreadr", "nflplotR", "nflfastR", "nflseedR", "nfl4th", 
+  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
   "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
-  "dagitty", "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", 
-  "patchwork", "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", 
-  "sjstats", "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", 
-  "fitdistrplus", "sn", "mclust", "magrittr", "viridis", "plotly", "quantmod", 
-  "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast", "stringi", #"webshot2"
-  "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
-  "caret", "tidymodels", "LongituRF", "gpboost", "cmdstanr", "learnr", 
-  "remotes", "corrplot", "correlation", "mgcv", "rms", "car", "performance", 
-  "lavaan", "lavaangui", "mice", "miceadds", "interactions", "shiny", 
-  "PearsonDS", "robustbase", "ordinal", "MASS", "data.table", "future", 
-  "future.apply", "QuantPsyc", "SimDesign", "effectsize"
+  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
+  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
+  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
+  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
+  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
+  "stringi", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
+  "caret", "LongituRF", "gpboost", "cmdstanr", "corrplot", "mgcv", "rms", 
+  "car", "lavaan", "lavaangui", "mice", "miceadds", "interactions", 
+  "robustbase", "ordinal", "MASS", "data.table", "future", "future.apply", 
+  "SimDesign", "domir",
+  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
+  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
+  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
+  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
+  "learnr", "remotes", "shiny", "PearsonDS" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
 )

### 2025-07-09 mixed-models.qmd
20250709 - r-squared and dominance analysis

@@
 
-MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
+performance::r2(pointsPerSeason_randomIntercepts)
 performance::icc(pointsPerSeason_randomIntercepts)
@@
 
-MuMIn::r.squaredGLMM(pointsPerSeason_position)
+performance::r2(pointsPerSeason_position)
 emmeans::emmeans(pointsPerSeason_position, "positionFactor")
@@
 
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
+performance::r2(pointsPerSeason_positionAgeFixedLinearSlopes)
 emmeans::emmeans(pointsPerSeason_positionAgeFixedLinearSlopes, "positionFactor")
@@
 
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
+performance::r2(pointsPerSeason_positionAgeRandomLinearSlopes)
 emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearSlopes, "positionFactor")
@@
 
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+performance::r2(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
 emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes, "positionFactor")
@@
 
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+performance::r2(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
 emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction, "positionFactor")
@@
 
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+performance::r2(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
 emmeans::emmeans(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "positionFactor")
@@
 summary(pointsPerSeason_quadraticRegression)$r.squared
-MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+performance::r2(pointsPerSeason_randomIntercepts)
+performance::r2(pointsPerSeason_positionAgeFixedLinearSlopes)
+performance::r2(pointsPerSeason_positionAgeRandomLinearSlopes)
+performance::r2(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+performance::r2(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
 ```
@@
 )
+```
+
+```{r}
+#| label: gam-mixed-models-summary
 
@@
 pointsPerSeason_gamSummary$r.sq
-MuMIn::r.squaredGLMM(pointsPerSeason_gam)
+performance::r2(pointsPerSeason_gam)
 AIC(pointsPerSeason_gam)
@@
 summary(pointsPerSeason_quadraticRegression)$r.squared
-MuMIn::r.squaredGLMM(pointsPerSeason_randomIntercepts)
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeFixedLinearSlopes)
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearSlopes)
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
-MuMIn::r.squaredGLMM(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
-MuMIn::r.squaredGLMM(pointsPerSeason_gam)
+performance::r2(pointsPerSeason_randomIntercepts)
+performance::r2(pointsPerSeason_positionAgeFixedLinearSlopes)
+performance::r2(pointsPerSeason_positionAgeRandomLinearSlopes)
+performance::r2(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopes)
+performance::r2(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteraction)
+performance::r2(pointsPerSeason_gam)
 
@@
 
-MuMIn::r.squaredGLMM(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
+performance::r2(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
 emmeans::emmeans(pointsPerSeasonDepth_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience, "positionFactor")
@@
 )
+```
+
+```{r}
+#| label: gam-players-top-depth-chart-mixed-models-summary
 
@@
 pointsPerSeasonDepth_gamSummary$r.sq
-MuMIn::r.squaredGLMM(pointsPerSeasonDepth_gam)
+performance::r2(pointsPerSeasonDepth_gam)
 AIC(pointsPerSeasonDepth_gam)
@@
 
+::: {#nte-bayesianMixedModelR2 .callout-note title="Bayesian mixed model R-sqaured"}
+Note: the following code that estimates the $R^2$ for the Bayesian mixed model takes a while to run.
+:::
+
+```{r}
+#| eval: false
+
+performance::r2(bayesianMixedModelFit)
+```
+
 #### Trace Plots {#sec-fantasyPointsByAgeExperienceModelsBayesianTrace}

### 2025-07-09 multiple-regression.qmd
20250708 - track chunk execution times
20250709 - r-squared and dominance analysis

@@
+{{< include _chunk-timing.qmd >}}
+
 # Multiple Regression {#sec-multipleRegression}
@@
 library("broom")
+library("effectsize")
 library("tidyverse")
@@
 summary(linearRegression)
+
 print(effectsize::standardize_parameters(linearRegression, method = "refit"), digits = 2)
@@
 We fit a logistic regression model using the `glm()` function and specifying `family = binomial()`.
+To calculate the model $R^2$, we use the `r2()` function of the `performance` package [@R-performance; @Ludecke2021_packages].
 
@@
 summary(logisticRegression)
+performance::r2(logisticRegression)
 print(effectsize::standardize_parameters(logisticRegression, method = "refit"), digits = 2)
@@
 summary(ordinalRegression)
+performance::r2(ordinalRegression)
 print(effectsize::standardize_parameters(ordinalRegression, method = "refit"), digits = 2)
@@
 summary(poissonRegression)
+performance::r2(poissonRegression)
 print(effectsize::standardize_parameters(poissonRegression, method = "refit"), digits = 2)
@@
 summary(negativeBinomialRegression)
+performance::r2(negativeBinomialRegression)
 print(effectsize::standardize_parameters(negativeBinomialRegression, method = "refit"), digits = 2)
@@
 summary(betaRegression)
+performance::r2(betaRegression)
 ```
@@
 
+### Dominance Analysis {#sec-multipleRegressionDominanceAnalysis}
+
+We can perform a dominance analysis to evaluate the relative importance of the predictors in the regression model.
+To perform the dominance analysis, we use the `dominance_analysis()` function of the `parameters` package [@R-parameters; @Luedecke2020_packages], which leverages the `domir` package [@R-domir].
+
+```{r}
+parameters::dominance_analysis(linearRegressionModel)
+```
+
 ### Visualizing Regression Results {#sec-multipleRegressionVisualizingResults}
@@
 ```{r}
-meth <- make.method(dataToImpute)
+meth <- mice::make.method(dataToImpute)
 meth[1:length(meth)] <- ""
@@
 ```{r}
-pred <- make.predictorMatrix(dataToImpute)
+pred <- mice::make.predictorMatrix(dataToImpute)
 pred[1:nrow(pred), 1:ncol(pred)] <- 0
@@
 ```{r}
-imp_long <- complete(
+imp_long <- mice::complete(
   imp,
@@
 
-imp_mids <- as.mids(imp_long)
+imp_mids <- mice::as.mids(imp_long)
 ```
@@
 
-summary(pool(imp_regression))
+summary(mice::pool(imp_regression))
 ```
@@
 
-rms::robcov(rms::ols(
+regressionWithClusterVariable <- rms::robcov(rms::ols(
   fantasyPoints_transformed ~ age + height + weight + I(log(target_share + 1)),
@@
   cluster = player_stats_seasonal_subset$player_id) #account for nested data within player
+
+regressionWithClusterVariable
+performance::r2(regressionWithClusterVariable)
 ```
@@
 ```{r}
-robustRegression <- lmrob(
+robustRegression <- robustbase::lmrob(
   fantasyPoints_transformed ~ age + height + weight + I(log(target_share + 1)),
@@
 
-We can also estimate a model with multiple hypothesized [mediator](#sec-mediations):
+We can also estimate a model with multiple hypothesized [mediators](#sec-mediations):
 
@@
 summary(bayesianMultipleRegressionModel)
+performance::r2(bayesianMultipleRegressionModel)
 ```

### 2025-07-09 actuarial.qmd
20250702 - cross-validation
20250708 - track chunk execution times
20250709 - videos

@@
+{{< include _chunk-timing.qmd >}}
+
 # Judgment Versus Actuarial Approaches to Prediction {#sec-judgmentVsActuarial}
@@
 
-@TPS2017 provides a video of sports pundits, Stephen A. Smith and Skip Bayless, making bold statements and incorrect predictions:
+::: {#fig-video-sports-pundits}
 
-![Video of Sports Pundits, Stephen A.
-Smith and Skip Bayless, Making Bold Statements and Incorrect Predictions.
-From: <https://www.youtube.com/watch?v=lTjBuEPcLlc>.](images/sportsPundits.mp4){width=100%}\
+@TPS2017 provides a video of sports pundits, Stephen A. Smith and Skip Bayless, making bold statements and incorrect predictions, in @fig-video-sports-pundits:
+
+{{< video images/sportsPundits.mp4 >}}
+
+:::
 
@@
 However, you can also do internal cross-validation.
-For example, you can perform *k*-fold cross-validation, where you:
+For example, you can perform [*k*-fold cross-validation](#sec-kFoldCrossValidation), where you:
 

### 2025-07-09 cognitive-bias.qmd
20250619 - calibration accuracy of projections
20250708 - track chunk execution times
20250709 - videos

@@
+{{< include _chunk-timing.qmd >}}
+
 # Heuristics and Cognitive Biases in Prediction {#sec-cognitiveBias}
@@
 
-Here is a video from @GetUpESPN2021 of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends:
+::: {#fig-kyle-pitts}
 
-![Video of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends.
-From: <https://x.com/GetUpESPN/status/1380165126108672001> (archived at <https://perma.cc/JW8E-KV2C>)](images/kylePitts.mp4){width=100%}\
+A video from @GetUpESPN2021 (archived at <https://perma.cc/JW8E-KV2C>) of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends is in @fig-kyle-pitts:
+
+{{< video images/kylePitts.mp4 >}}
+
+:::
 
@@
 Overprecision corresponds to the "[overextremity](#sec-calibration)" form of [miscalibration](#sec-calibration).
+
+With respect to fantasy football, projections of players' fantasy points tend to be overprecise, as described in @sec-accuracyProjections.
+However, predictions about how well oneself will do in a fantasy football league may involve overestimation and overplacement.
 

### 2025-07-09 evaluating-prediction-accuracy.qmd
20250708 - videos and quotes
20250709 - videos

@@
 
-> "Nothing ruins fantasy more than reality." – [Renee Miller](https://www.youtube.com/watch?v=gmpLFWs5ae0&t=1430s) [@Yahoo2024]
+> "Nothing ruins fantasy more than reality." – [Renee Miller, Ph.D.](@fig-fantasy-vs-reality) [@Yahoo2024]
+
+::: {.content-visible when-format="html:js"}
+
+::: {#fig-fantasy-vs-reality}
+
+{{< video https://www.youtube.com/watch?v=gmpLFWs5ae0 start="1430" >}}
+
+:::
+
+:::
 

### 2025-07-09 fantasy-football.qmd
20250708 - track chunk execution times
20250709 - videos

@@
+{{< include _chunk-timing.qmd >}}
+
 # Intro to Football and Fantasy {#sec-introFantasy}
@@
 
-Here is a video that provides a brief introduction to American football [@Ly2015]:
+::: {#fig-football-intro}
 
-![Video that Provides An Introduction to American Football. From: <https://www.youtube.com/watch?v=Ddwp1HyEFRE>.](images/football.mp4){width=100%}\
+A video that provides a brief introduction to American football is in @fig-football-intro [@Ly2015]:
+
+{{< video images/football.mp4 >}}
+
+:::
 
@@
 
-Here is a video that provides a brief introduction to fantasy football [@NFLFantasy2020]:
+::: {#fig-fantasy-football-intro}
 
-![Video that Provides an Introduction to Fantasy Football. From: <https://www.youtube.com/watch?v=XhrBapdhLEc>.](images/fantasyFootball.mp4){width=100%}\
+A video that provides a brief introduction to fantasy football is in @fig-fantasy-football-intro [@NFLFantasy2020]:
+
+{{< video images/fantasyFootball.mp4 >}}
 

### 2025-07-09 getting-started.qmd
20250709 - videos

@@
 
-::: {#nte-gettingStarted .callout-note title="If you are in Dr. Petersen's class"}
-If you are in Dr. Petersen's class, also perform the following steps:
+::: {#nte-gettingStarted .callout-note title="If you are in Professor Petersen's class"}
+If you are in Professor Petersen's class, also perform the following steps:
 
@@
 
+Loops are fine for basic computations, but other approaches (such as the `apply()` family of functions) can be even faster than loops.
+
 ::: {.content-visible when-format="html"}
-
-Loops are fine for basic computations, but other approaches (such as the `apply()` family of functions) can be even faster than loops.
 

### 2025-07-09 intro.qmd
20250709 - videos

@@
 
-Here is a video of Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football [@NFL2014]:
+A video of Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football is in @fig-video-motz [@NFL2014]:
 
-![Video of Prof. Benjamin Motz Discussing Statistics and His Fantasy Football Course. From: <https://www.facebook.com/watch/?v=10155572257183615>.](images/motzVideo.mp4){width=100%}\
+{{< video images/motzVideo.mp4 >}}
 

### 2025-07-09 mixed-models.qmd
20250709 - videos

@@
 
-::: {#nte-bayesianMixedModelR2 .callout-note title="Bayesian mixed model R-sqaured"}
+::: {#nte-bayesianMixedModelR2 .callout-note title="Bayesian mixed model R-squared"}
 Note: the following code that estimates the $R^2$ for the Bayesian mixed model takes a while to run.

### 2025-07-09 sports-cognitive-psychology.qmd
20250708 - track chunk execution times
20250709 - videos

@@
+{{< include _chunk-timing.qmd >}}
+
 # Sports and Cognitive Psychology {#sec-sportsCognitivePsychology}
@@
 
-For instance, here is a [2015 video clip](https://streamable.com/1hw8) of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops" [@ABC2015], suggesting that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
+::: {#fig-football-commentator}
+
+For instance, there is a [2015 video clip](https://streamable.com/1hw8) of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops" in @fig-football-commentator [@ABC2015], suggesting that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
 The full game is [here](https://www.youtube.com/watch?v=1L9D2rp0mk8) [@ABC2015a].
 
-![Football Commentator Making Fun of Analytics in Football. From: <https://streamable.com/1hw8> (2015).](images/commentatorOnAnalytics.mp4){width=100%}\
+{{< video images/commentatorOnAnalytics.mp4 >}}
+
+:::
 

### 2025-07-10 actuarial.qmd
20250709 - videos
20250710 - video citations

@@
 
+@fig-video-sports-pundits is a video of sports pundits, Stephen A. Smith and Skip Bayless, making bold statements and incorrect predictions [@TPS2017]:
+
 ::: {#fig-video-sports-pundits}
-
-@TPS2017 provides a video of sports pundits, Stephen A. Smith and Skip Bayless, making bold statements and incorrect predictions, in @fig-video-sports-pundits:
 

### 2025-07-10 cognitive-bias.qmd
20250709 - videos
20250710 - video citations

@@
 
+@fig-kyle-pitts is a video from @GetUpESPN2021 (archived at <https://perma.cc/JW8E-KV2C>) of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends:
+
 ::: {#fig-kyle-pitts}
-
-A video from @GetUpESPN2021 (archived at <https://perma.cc/JW8E-KV2C>) of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends is in @fig-kyle-pitts:
 

### 2025-07-10 fantasy-football.qmd
20250709 - videos
20250710 - video citations

@@
 
+@fig-football-intro is a video that provides a brief introduction to American football [@Ly2015]:
+
 ::: {#fig-football-intro}
-
-A video that provides a brief introduction to American football is in @fig-football-intro [@Ly2015]:
 
@@
 
+@fig-fantasy-football-intro is a video that provides a brief introduction to fantasy football [@NFLFantasy2020]:
+
 ::: {#fig-fantasy-football-intro}
-
-A video that provides a brief introduction to fantasy football is in @fig-fantasy-football-intro [@NFLFantasy2020]:
 

### 2025-07-10 intro.qmd
20250709 - videos
20250710 - video citations

@@
 
+@fig-video-motz is a video of Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football [@NFL2014]:
+
 ::: {#fig-video-motz}
-
-A video of Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football is in @fig-video-motz [@NFL2014]:
 

### 2025-07-10 sports-cognitive-psychology.qmd
20250709 - videos
20250710 - video citations

@@
 
+As an example of people dismissing analytics, @fig-football-commentator is a [2015 video clip](https://streamable.com/1hw8) of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops" [@ABC2015].
+His characterization suggests he believes that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
+The full game is [here](https://www.youtube.com/watch?v=1L9D2rp0mk8) [@ABC2015a].
+
 ::: {#fig-football-commentator}
-
-For instance, there is a [2015 video clip](https://streamable.com/1hw8) of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops" in @fig-football-commentator [@ABC2015], suggesting that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
-The full game is [here](https://www.youtube.com/watch?v=1L9D2rp0mk8) [@ABC2015a].
 

### 2025-07-10 factor-analysis.qmd
20250710 - factor analysis

--- 2025-07-08 factor-analysis.qmd
@@
 
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
@@
 ```{r}
+load(file = "./data/player_stats_seasonal.RData")
 load(file = "./data/player_stats_weekly.RData")
@@
 
+#### Season-Averages {#sec-factorAnalysisPrepareDataSeasonAvg}
+
+```{r}
+player_stats_seasonal_avgPerGame <- player_stats_seasonal %>% 
+  mutate(
+    completionsPerGame = completions / games,
+    attemptsPerGame = attempts / games,
+    passing_yardsPerGame = passing_yards / games,
+    passing_tdsPerGame = passing_tds / games,
+    passing_interceptionsPerGame = passing_interceptions / games,
+    sacks_sufferedPerGame = sacks_suffered / games,
+    sack_yards_lostPerGame = sack_yards_lost / games,
+    sack_fumblesPerGame = sack_fumbles / games,
+    sack_fumbles_lostPerGame = sack_fumbles_lost / games,
+    passing_air_yardsPerGame = passing_air_yards / games,
+    passing_yards_after_catchPerGame = passing_yards_after_catch / games,
+    passing_first_downsPerGame = passing_first_downs / games,
+    #passing_epaPerGame = passing_epa / games,
+    #passing_cpoePerGame = passing_cpoe / games,
+    passing_2pt_conversionsPerGame = passing_2pt_conversions / games,
+    #pacrPerGame = pacr / games
+    pass_40_ydsPerGame = pass_40_yds / games,
+    pass_incPerGame = pass_inc / games,
+    #pass_comp_pctPerGame = pass_comp_pct / games,
+    fumblesPerGame = fumbles / games
+  )
+
+nfl_nextGenStats_seasonal <- nfl_nextGenStats_weekly %>% 
+  filter(season_type == "REG") %>% 
+  select(-any_of(c("week","season_type","player_display_name","player_position","team_abbr","player_first_name","player_last_name","player_jersey_number","player_short_name"))) %>% 
+  group_by(player_gsis_id, season) %>% 
+  summarise(
+    across(everything(),
+      ~ mean(.x, na.rm = TRUE)),
+    .groups = "drop")
+```
+
 #### Merge Data {#sec-factorAnalysisPrepareDataMerge}
@@
 dataMerged <- full_join(
-  player_stats_weekly,
-  nfl_nextGenStats_weekly %>% select(-any_of(c("player_display_name","completions","attempts","receptions","targets"))),
-  by = c("player_id" = "player_gsis_id","season","season_type","week"),
+  player_stats_seasonal_avgPerGame,
+  nfl_nextGenStats_seasonal %>% select(-any_of(c("player_display_name","completions","attempts","receptions","targets"))),
+  by = c("player_id" = "player_gsis_id","season"),
 )
@@
 ```{r}
-faVars <- c(
-  "completions","attempts","passing_yards","passing_tds","passing_interceptions",
-  "sacks_suffered","sack_yards_lost","sack_fumbles","sack_fumbles_lost",
-  "passing_air_yards","passing_yards_after_catch","passing_first_downs",
-  "passing_epa","passing_cpoe","passing_2pt_conversions","pacr","pass_40_yds",
-  "pass_inc","pass_comp_pct","fumbles","two_pts",
+#| eval: false
+
+c(
+  "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame","passing_interceptionsPerGame",
+  "sacks_sufferedPerGame","sack_yards_lostPerGame","sack_fumblesPerGame","sack_fumbles_lostPerGame",
+  "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame",
+  "passing_epa","passing_cpoe","passing_2pt_conversionsPerGame","pacr","pass_40_ydsPerGame",
+  "pass_incPerGame","pass_comp_pct","fumblesPerGame", #"two_pts",
   "avg_time_to_throw","avg_completed_air_yards","avg_intended_air_yards",
@@
 
+```{r}
+#efaVars <- c(
+#  "completionsPerGame","passing_yardsPerGame","passing_tdsPerGame",
+#  "avg_completed_air_yards","avg_intended_air_yards","avg_air_distance",
+#  "passing_cpoe","pass_comp_pct","passer_rating","completion_percentage_above_expectation")
+
+#efaVars <- c(
+# "passing_yardsPerGame","passing_tdsPerGame",
+# "avg_completed_air_yards","avg_air_distance",
+# "pass_comp_pct","passer_rating")
+#
+#cfaVars <- c(
+# "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
+# "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame",
+# "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
+# "avg_air_distance","max_air_distance","avg_air_yards_to_sticks","passing_cpoe","pass_comp_pct",
+# "passer_rating","completion_percentage_above_expectation"
+#)
+
+#faVars <- unique(c(cfaVars, efaVars))
+
+efaVars <- c(
+ "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
+ "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame",
+ "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
+ "avg_air_distance","max_air_distance","avg_air_yards_to_sticks","passing_cpoe","pass_comp_pct",
+ "passer_rating","completion_percentage_above_expectation"
+)
+```
+
+#### Subset Data {#sec-pcaPrepareDataSubset}
+
+```{r}
+dataMerged_subset <- dataMerged %>% 
+  filter(!is.na(player_id)) %>% 
+  filter(position == "QB")
+```
+
 #### Standardize Variables {#sec-pcaPrepareDataStandardize}
 
-Although standardizing variables is not strictly necessary in factor analysis, it can be helpful to prevent some variables from having considerably larger variances than others.
+Standardizing variables is not strictly necessary in factor analysis; however, standardization can be helpful to prevent some variables from having considerably larger variances than others.
 
 ```{r}
-dataForFA <- dataMerged
-dataForFA[faVars] <- scale(dataForFA[faVars])
+dataForFA <- dataMerged_subset
+dataForFA_standardized <- dataForFA
+
+dataForFA_standardized[faVars] <- scale(dataForFA_standardized[faVars])
+#dataForFA[faVars] <- scale(dataForFA[faVars])
 ```
@@
 
+```{r}
+#| eval: false
+#| include: false
+
+cor_mat <- cor(dataForFA[faVars], use = "pairwise.complete.obs")
+
+cor_tidy <- cor_mat %>%
+  as.data.frame() %>%
+  rownames_to_column("var1") %>%
+  pivot_longer(-var1, names_to = "var2", values_to = "correlation") %>%
+  filter(var1 != var2) %>% # remove self-correlations
+  mutate(pair = pmap_chr(list(var1, var2), ~ paste(sort(c(...)), collapse = "___"))) %>% 
+  distinct(pair, .keep_all = TRUE) # remove duplicate pairs (symmetry)
+
+high_corrs <- cor_tidy %>%
+  filter(abs(correlation) > 0.85) %>%
+  arrange(desc(abs(correlation)))
+
+high_corrs
+
+cov_mat <- cov(dataForFA[, faVars], use = "pairwise.complete.obs")
+eigen_cov_mat <- eigen(cov_mat)
+print(eigen_cov_mat$values)
+
+# Index of smallest (most problematic) eigenvalue
+idx <- which.min(eigen_cov_mat$values)
+
+# Extract the associated eigenvector
+problem_vector <- eigen_cov_mat$vectors[, idx]
+names(problem_vector) <- faVars
+
+# Sort by absolute magnitude to identify most influential variables
+sort(abs(problem_vector), decreasing = TRUE)
+```
+
 We generated the scree plot in @fig-efaScreePlot1 using the `fa.parallel()` function of the `psych` package [@R-psych].
@@
 
+#screeDataEFA <- nFactors::nScree(
+#  x = cor( # throws error with correlation matrix, so use covariance instead (below)
+#    dataForFA[faVars],
+#    use = "pairwise.complete.obs"),
+#  model = "factors")
+
 screeDataEFA <- nFactors::nScree(
-  x = cor(
+  x = cov(
     dataForFA[faVars],
     use = "pairwise.complete.obs"),
+  cor = FALSE,
   model = "factors")
@@
 ```{r}
-#| eval: false
+#| label: efa
 
-efaOblique_fit <- lavaan::efa(
+efa_fit <- lavaan::efa(
   data = dataForFA,
-  ov.names = faVars,
-  nfactors = 1:9,
+  ov.names = efaVars,
+  nfactors = 1:7,
   rotation = "geomin",
@@
   estimator = "MLR",
-  cluster = "player_id",
-  meanstructure = TRUE
-)
+  bounds = "standard",
+  meanstructure = TRUE,
+  em.h1.iter.max = 2000000)
 ```
 
-A path diagram of the EFA model was created using the `plot_lavaan()` function of the `lavaangui` package [@R-lavaangui; @Karch2025_packages].
+```{r}
+#| label: efa-summary
+
+summary(efa_fit)
+```
+
+A path diagram of the EFA model was created using the `lavaanPlot()` function of the `lavaanPlot` package [@R-lavaanPlot].
+
+```{r}
+lavaanPlot::lavaanPlot(
+  efa_fit$nf3,
+  coefs = TRUE,
+  stand = TRUE)
+```
+
+Here is the syntax for estimating a three-factor EFA using exploratory structural equation modeling (ESEM).
+Estimating the model in a ESEM framework allows us to make modifications to the model, such as adding correlated residuals, and adding predictors or outcomes of the latent factors.
+
+```{r}
+#| label: efa-3factor-syntax
+
+efa3factor_syntax <- '
+ # EFA Factor Loadings
+ efa("efa1")*f1 + 
+ efa("efa1")*f2 + 
+ efa("efa1")*f3 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
+ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
+ passer_rating + completion_percentage_above_expectation
+'
+```
+
+To fit the ESEM model, we use the `sem()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
+
+```{r}
+#| label: efa-3factor-fit
+
+efa3factor_fit <- sem(
+  efa3factor_syntax,
+  data = dataForFA,
+  information = "observed",
+  missing = "ML",
+  estimator = "MLR",
+  rotation = "geomin",
+  bounds = "standard",
+  meanstructure = TRUE,
+  em.h1.iter.max = 2000000)
+```
+
+```{r}
+summary(
+  efa3factor_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+```{r}
+lavaan::fitMeasures(
+  efa3factor_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+lavaan::residuals(
+  efa3factor_fit,
+  type = "cor")
+```
+
+```{r}
+lavaan::modificationindices(
+  efa3factor_fit,
+  sort. = TRUE)
+```
+
+```{r}
+lavaan::lavPredict(efa3factor_fit)
+```
+
+```{r}
+lavaanPlot::lavaanPlot(
+  efa3factor_fit,
+  coefs = TRUE,
+  stand = TRUE)
+```
+
+To make the plot interactive for editing, you can use the `plot_lavaan()` function of the `lavaangui` package [@R-lavaangui; @Karch2025_packages]:
 
@@
 
-lavaangui::plot_lavaan(efaOblique_fit)
+lavaangui::plot_lavaan(efa3factor_fit)
+```
+
+Here is a modification of the 3-factor model with correlated residuals:
+
+```{r}
+#| label: efa-3factorModified-syntax
+
+efa3factorModified_syntax <- '
+ # EFA Factor Loadings
+ efa("efa1")*f1 + 
+ efa("efa1")*f2 + 
+ efa("efa1")*f3 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
+ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
+ passer_rating + completion_percentage_above_expectation
+
+ # Correlated Residuals
+ completionsPerGame ~~ attemptsPerGame
+ passing_yardsPerGame ~~ passing_yards_after_catchPerGame
+ attemptsPerGame ~~ passing_yardsPerGame
+ attemptsPerGame ~~ passing_air_yardsPerGame
+ passing_air_yardsPerGame ~~ passing_yards_after_catchPerGame
+ passing_yards_after_catchPerGame ~~ avg_completed_air_yards
+ passing_tdsPerGame ~~ passer_rating
+ completionsPerGame ~~ passing_air_yardsPerGame
+ max_completed_air_distance ~~ max_air_distance
+ aggressiveness ~~ completion_percentage_above_expectation
+'
+```
+
+```{r}
+#| label: efa-3factorModified-fit
+
+efa3factorModified_fit <- sem(
+  efa3factorModified_syntax,
+  data = dataForFA,
+  information = "observed",
+  missing = "ML",
+  estimator = "MLR",
+  rotation = "geomin",
+  bounds = "standard",
+  meanstructure = TRUE,
+  em.h1.iter.max = 2000000)
+```
+
+```{r}
+summary(
+  efa3factorModified_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+```{r}
+lavaan::fitMeasures(
+  efa3factorModified_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+lavaan::residuals(
+  efa3factorModified_fit,
+  type = "cor")
+```
+
+```{r}
+lavaan::modificationindices(
+  efa3factorModified_fit,
+  sort. = TRUE)
+```
+
+```{r}
+lavaan::lavPredict(efa3factorModified_fit)
+```
+
+```{r}
+lavaanPlot::lavaanPlot(
+  efa3factorModified_fit,
+  coefs = TRUE,
+  stand = TRUE)
+```
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(efa3factorModified_fit)
+```
+
+```{r}
+anova(
+  efa3factorModified_fit,
+  efa3factor_fit
+)
 ```
@@
 
+```{r}
+#| eval: false
+#| include: false
+
+psych::principal(
+  dataForFA_standardized[faVars],
+  nfactors = 3,
+  rotate = "oblimin")
+
+'
+ #Factor loadings
+ F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + passing_interceptionsPerGame + passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
+ F2 =~ pacr + avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + avg_air_yards_to_sticks + expected_completion_percentage + avg_air_distance + max_air_distance
+ F3 =~ passing_epa + passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
+'
+
+factor1_syntax <- '
+ # Factor loadings
+ F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
+
+ # Correlated residuals
+ passing_air_yardsPerGame ~~ passing_yards_after_catchPerGame
+ completionsPerGame ~~ attemptsPerGame
+ attemptsPerGame ~~ passing_yards_after_catchPerGame
+ passing_yards_after_catchPerGame ~~ passing_first_downsPerGame
+ completionsPerGame ~~ passing_first_downsPerGame
+ attemptsPerGame ~~ passing_tdsPerGame
+ completionsPerGame ~~ passing_tdsPerGame
+ passing_tdsPerGame ~~ passing_air_yardsPerGame
+'
+
+factor1_fit <- lavaan::cfa(
+  factor1_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+
+summary(
+  factor1_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+
+lavaan::modificationindices(
+  factor1_fit,
+  sort. = TRUE)
+
+factor2_syntax <- '
+ # Factor loadings
+ F2 =~ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + avg_air_distance + max_air_distance + avg_air_yards_to_sticks #expected_completion_percentage + pacr
+
+ # Correlated residuals
+ max_completed_air_distance ~~ max_air_distance
+ avg_completed_air_yards ~~ max_completed_air_distance
+ avg_completed_air_yards ~~ avg_air_distance
+ avg_completed_air_yards ~~ max_air_distance
+ avg_intended_air_yards ~~ max_completed_air_distance
+'
+
+factor2_fit <- lavaan::cfa(
+  factor2_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+
+summary(
+  factor2_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+
+lavaan::modificationindices(
+  factor2_fit,
+  sort. = TRUE)
+
+factor3_syntax <- '
+ #Factor loadings
+ F3 =~ passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
+'
+
+factor3_fit <- lavaan::cfa(
+  factor3_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+
+summary(
+  factor3_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+
+lavaan::modificationindices(
+  factor3_fit,
+  sort. = TRUE)
+```
+
 We fit CFA models using the `cfa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
@@
 ```{r}
-cfaModel1_syntax <- '
+cfa1factor_syntax <- '
  #Factor loadings
- F1 =~ completions + attempts + passing_yards + passing_tds + passing_interceptions + sacks_suffered + sack_yards_lost + sack_fumbles + sack_fumbles_lost + passing_air_yards + passing_yards_after_catch + passing_first_downs + pass_inc + fumbles
- F2 =~ passing_air_yards + pacr + avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + avg_air_yards_to_sticks + expected_completion_percentage + avg_air_distance + max_air_distance
- F3 =~ passing_tds + passing_epa + passing_cpoe + pass_40_yds + pass_comp_pct + avg_air_yards_differential + max_completed_air_distance + avg_time_to_throw + avg_completed_air_yards + avg_intended_air_yards + passer_rating + completion_percentage_above_expectation
+ F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
+ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
+ passer_rating + completion_percentage_above_expectation
 '
@@
 ```{r}
-#| eval: false
+#| label: cfa-1factor-fit
 
-cfaModel1_fit <- lavaan::cfa(
-  cfaModel1_syntax,
+cfa1factor_fit <- lavaan::cfa(
+  cfa1factor_syntax,
   data = dataForFA,
@@
   estimator = "MLR",
+  bounds = "standard",
   std.lv = TRUE,
-  cluster = "player_id",
-  em.iter.max = 2000000)
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
 ```
@@
 ```{r}
-#| eval: false
-
 summary(
-  cfaModel1_fit,
+  cfa1factor_fit,
   fit.measures = TRUE,
@@
 ```{r}
-#| eval: false
-
 lavaan::fitMeasures(
-  cfaModel1_fit,
+  cfa1factor_fit,
   fit.measures = c(
@@
 ```{r}
-#| eval: false
+lavaan::residuals(
+  cfa1factor_fit,
+  type = "cor")
+```
 
-lavaan::residuals(
-  cfaModel1_fit,
-  type = "cor")
+```{r}
+lavaan::modificationindices(
+  cfa1factor_fit,
+  sort. = TRUE)
+```
+
+```{r}
+lavaan::lavPredict(cfa1factor_fit)
+```
+
+```{r}
+lavaanPlot::lavaanPlot(
+  cfa1factor_fit,
+  coefs = TRUE,
+  stand = TRUE)
 ```
@@
 
+lavaangui::plot_lavaan(cfa1factor_fit)
+```
+
+```{r}
+cfa2factor_syntax <- '
+ #Factor loadings
+ F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
+
+ F2 =~ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
+ passer_rating + completion_percentage_above_expectation
+'
+```
+
+```{r}
+#| label: cfa-2factor-fit
+
+cfa2factor_fit <- lavaan::cfa(
+  cfa2factor_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+```
+
+```{r}
+summary(
+  cfa2factor_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+```{r}
+lavaan::fitMeasures(
+  cfa2factor_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+lavaan::residuals(
+  cfa2factor_fit,
+  type = "cor")
+```
+
+```{r}
 lavaan::modificationindices(
-  cfaModel1_fit,
+  cfa2factor_fit,
   sort. = TRUE)
+```
+
+```{r}
+lavaan::lavPredict(cfa2factor_fit)
+```
+
+```{r}
+lavaanPlot::lavaanPlot(
+  cfa2factor_fit,
+  coefs = TRUE,
+  stand = TRUE)
 ```
@@
 
-lavaan::lavPredict(cfaModel1_fit)
+lavaangui::plot_lavaan(cfa2factor_fit)
 ```
 
-A path diagram of the CFA model was created using the `plot_lavaan()` function of the `lavaangui` package [@R-lavaangui; @Karch2025_packages].
+```{r}
+anova(
+  cfa2factor_fit,
+  cfa1factor_fit
+)
+```
+
+```{r}
+cfa3factor_syntax <- '
+ #Factor loadings
+ F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
+
+ F2 =~ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks
+
+ F3 =~ passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
+'
+```
+
+```{r}
+#| label: cfa-3factor-fit
+
+cfa3factor_fit <- lavaan::cfa(
+  cfa3factor_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+```
+
+```{r}
+summary(
+  cfa3factor_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+```{r}
+lavaan::fitMeasures(
+  cfa3factor_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+lavaan::residuals(
+  cfa3factor_fit,
+  type = "cor")
+```
+
+```{r}
+lavaan::modificationindices(
+  cfa3factor_fit,
+  sort. = TRUE)
+```
+
+```{r}
+lavaan::lavPredict(cfa3factor_fit)
+```
+
+```{r}
+lavaanPlot::lavaanPlot(
+  cfa3factor_fit,
+  coefs = TRUE,
+  stand = TRUE)
+```
 
@@
 
-lavaangui::plot_lavaan(cfaModel1_fit)
+lavaangui::plot_lavaan(cfa3factor_fit)
+```
+
+```{r}
+anova(
+  cfa3factor_fit,
+  cfa2factor_fit
+)
+```
+
+```{r}
+cfa3factorModified_syntax <- '
+ # Factor loadings
+ F1 =~ NA*completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
+
+ F2 =~ NA*avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks
+
+ F3 =~ NA*passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
+
+ # Cross loadings
+ #F3 =~ attemptsPerGame
+
+ # Correlated residuals
+ passing_air_yardsPerGame ~~ passing_yards_after_catchPerGame
+ completionsPerGame ~~ attemptsPerGame
+ attemptsPerGame ~~ passing_yards_after_catchPerGame
+ passing_yards_after_catchPerGame ~~ passing_first_downsPerGame
+ completionsPerGame ~~ passing_first_downsPerGame
+ attemptsPerGame ~~ passing_tdsPerGame
+ completionsPerGame ~~ passing_tdsPerGame
+ passing_tdsPerGame ~~ passing_air_yardsPerGame
+
+ max_completed_air_distance ~~ max_air_distance
+ avg_completed_air_yards ~~ max_completed_air_distance
+ avg_completed_air_yards ~~ avg_air_distance
+ avg_completed_air_yards ~~ max_air_distance
+ avg_intended_air_yards ~~ max_completed_air_distance
+
+ completionsPerGame ~~ pass_comp_pct
+ passing_yardsPerGame ~~ avg_completed_air_yards
+ passing_yardsPerGame ~~ max_completed_air_distance
+ passing_tdsPerGame ~~ passer_rating
+ aggressiveness ~~ completion_percentage_above_expectation
+
+ # Variances
+ F1 ~~ 1*F1
+ F2 ~~ 1*F2
+ #F3 ~~ 1*F3
+'
+```
+
+```{r}
+#| label: cfa-3factorModified-fit
+
+cfa3factorModified_fit <- lavaan::cfa(
+  cfa3factorModified_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  #std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+```
+
+```{r}
+summary(
+  cfa3factorModified_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+```{r}
+lavaan::fitMeasures(
+  cfa3factorModified_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+lavaan::residuals(
+  cfa3factorModified_fit,
+  type = "cor")
+```
+
+```{r}
+lavaan::modificationindices(
+  cfa3factorModified_fit,
+  sort. = TRUE)
+```
+
+```{r}
+lavaan::lavPredict(cfa3factorModified_fit)
+```
+
+```{r}
+lavaanPlot::lavaanPlot(
+  cfa3factorModified_fit,
+  coefs = TRUE,
+  stand = TRUE)
+```
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(cfa3factorModified_fit)
+```
+
+```{r}
+anova(
+  cfa3factorModified_fit,
+  cfa3factor_fit
+)
 ```

### 2025-07-10 getting-started.qmd
20250710 - factor analysis

@@
      "parallelly","doParallel","missRanger","ggridges","powerjoin","caret",
-     "LongituRF","gpboost","corrplot","mgcv","rms","car","lavaan","lavaangui",
-     "mice","miceadds","interactions","robustbase","ordinal","MASS",
+     "LongituRF","gpboost","corrplot","mgcv","rms","car","lavaan","lavaanPlot",
+     "lavaangui","mice","miceadds","interactions","robustbase","ordinal","MASS",
      "data.table","future","future.apply","SimDesign","domir"))

### 2025-07-10 index.qmd
20250710 - factor analysis

--- 2025-07-08 index.qmd
@@
   "petersenlab", "ffanalytics",
-  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "broom", "broom.mixed", 
-  "psych", "downlit", "xml2", "nflreadr", "nflplotR", "nflfastR", "nflseedR", 
-  "nfl4th", "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", 
-  "XICOR", "dagitty", "ggdag", "lubridate", "ggtext", "gghighlight", "ggExtra", 
-  "grid", "patchwork", "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", 
-  "pbkrtest", "sjstats", "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", 
-  "fitdistrplus", "sn", "mclust", "magrittr", "viridis", "viridisLite", "msir", 
-  "plotly", "webshot2", "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", 
-  "zoo", "forecast", "stringi", "parallelly", "doParallel", "missRanger", 
-  "ggridges", "powerjoin", "caret", "tidymodels", "LongituRF", "gpboost", 
-  "cmdstanr", "corrplot", "correlation", "mgcv", "rms", "car", "performance", 
-  "lavaan", "lavaangui", "mice", "miceadds", "interactions", "robustbase", 
-  "ordinal", "MASS", "data.table", "future", "future.apply", "see", "SimDesign", 
-  "effectsize", 
-  "learnr", "remotes", "shiny", "PearsonDS" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
+  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
+  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
+  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
+  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
+  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
+  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
+  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
+  "stringi", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
+  "caret", "LongituRF", "gpboost", "cmdstanr", "corrplot", "mgcv", "rms", 
+  "car", "lavaan", "lavaanPlot", "mice", "miceadds", "interactions", 
+  "robustbase", "ordinal", "MASS", "data.table", "future", "future.apply", 
+  "SimDesign", "domir",
+  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
+  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
+  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
+  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
+  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
 )

### 2025-07-10 factor-analysis.qmd
20250710 - clean up code

@@
 ```{r}
-#| eval: false
+#| include: false
 
@@
 ```{r}
-#efaVars <- c(
-#  "completionsPerGame","passing_yardsPerGame","passing_tdsPerGame",
-#  "avg_completed_air_yards","avg_intended_air_yards","avg_air_distance",
-#  "passing_cpoe","pass_comp_pct","passer_rating","completion_percentage_above_expectation")
-
-#efaVars <- c(
-# "passing_yardsPerGame","passing_tdsPerGame",
-# "avg_completed_air_yards","avg_air_distance",
-# "pass_comp_pct","passer_rating")
-#
-#cfaVars <- c(
-# "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
-# "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame",
-# "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
-# "avg_air_distance","max_air_distance","avg_air_yards_to_sticks","passing_cpoe","pass_comp_pct",
-# "passer_rating","completion_percentage_above_expectation"
-#)
-
-#faVars <- unique(c(cfaVars, efaVars))
-
-efaVars <- c(
+faVars <- c(
  "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
@@
 dataForFA_standardized[faVars] <- scale(dataForFA_standardized[faVars])
-#dataForFA[faVars] <- scale(dataForFA[faVars])
 ```
@@
   data = dataForFA,
-  ov.names = efaVars,
+  ov.names = faVars,
   nfactors = 1:7,

### 2025-07-10 factor-analysis.qmd
20250710 - fix library loading

@@
 ```{r}
+#| eval: false
+
 library("psych")
@@
 library("lavaan")
+library("lavaanPlot")
 library("lavaangui")
+library("tidyverse")
+```
+
+```{r}
+#| include: false
+
+# Don't load lavaangui (b/c not installed)
+library("psych")
+library("nFactors")
+library("lavaan")
+library("lavaanPlot")
 library("tidyverse")

### 2025-07-10 actuarial.qmd
20250710 - quotes

@@
 
-> If a sociologist were predicting whether Professor X would go to the movies on a certain night, he might have an equation involving age, academic specialty, and introversion score.
+> *If a sociologist were predicting whether Professor X would go to the movies on a certain night, he might have an equation involving age, academic specialty, and introversion score.
 > The equation might yield a probability of .90 that Professor X goes to the movie tonight.
@@
 > Because broken legs are very rare, and in the sociologist's entire sample of 500 criterion cases plus 250 cross-validating cases, he did not come upon a single instance of it.
-> He uses the broken leg datum confidently, because 'broken leg' is a subclass of a larger class we may crudely denote as 'relatively immobilizing illness or injury,' and movie-attending is a subclass of a larger class of 'actions requiring moderate mobility.'
+> He uses the broken leg datum confidently, because 'broken leg' is a subclass of a larger class we may crudely denote as 'relatively immobilizing illness or injury,' and movie-attending is a subclass of a larger class of 'actions requiring moderate mobility.'*
 > 

### 2025-07-10 base-rates.qmd
20250708 - track chunk execution times
20250710 - quotes

@@
+{{< include _chunk-timing.qmd >}}
+
 # Base Rates {#sec-baseRates}
@@
 
-> A cab was involved in a hit-and-run accident at night.
+> *A cab was involved in a hit-and-run accident at night.
 > Two cab companies, the Green and the Blue, operate in the city.
@@
 > 
-> What is the probability that the cab involved in the accident was Blue rather than Green?
+> What is the probability that the cab involved in the accident was Blue rather than Green?*
 > 

### 2025-07-10 evaluating-prediction-accuracy.qmd
20250709 - videos
20250710 - video citations
20250710 - quotes

@@
 
-> "Nothing ruins fantasy more than reality." – [Renee Miller, Ph.D.](@fig-fantasy-vs-reality) [@Yahoo2024]
+> "*Nothing ruins fantasy more than reality.*" – [Renee Miller, Ph.D.](#fig-fantasy-vs-reality) [@Yahoo2024]
 
@@
 
-> I consider it unnecessary to persuade you that most so-called "theories" in the soft areas of psychology (clinical, counseling, social, personality, community, and school psychology) are scientifically unimpressive and technologically worthless ...
+> *I consider it unnecessary to persuade you that most so-called "theories" in the soft areas of psychology (clinical, counseling, social, personality, community, and school psychology) are scientifically unimpressive and technologically worthless ...
 Perhaps the easiest way to convince yourself is by scanning the literature of soft psychology over the last 30 years and noticing what happens to theories.
@@
 > But in fields like personology and social psychology, this seems not to happen.
-> There is a period of enthusiasm about a new theory, a period of attempted application to several fact domains, a period of disillusionment as the negative data come in, a growing bafflement about inconsistent and unreplicable empirical results, multiple resort to ad hoc excuses, and then finally people just sort of lose interest in the thing and pursue other endeavors.
+> There is a period of enthusiasm about a new theory, a period of attempted application to several fact domains, a period of disillusionment as the negative data come in, a growing bafflement about inconsistent and unreplicable empirical results, multiple resort to ad hoc excuses, and then finally people just sort of lose interest in the thing and pursue other endeavors.*
 > 
-> Meehl [-@Meehl1978, pp. 806–807]
+> --- Meehl [-@Meehl1978, pp. 806–807]
 

### 2025-07-10 intro.qmd
20250710 - quotes

@@
 
-> "*This material probably won't win you fantasy football championships.
+> *This material probably won't win you fantasy football championships.
 > You could take what we learn and apply it to fantasy football and you might become 5 percent more likely to win.
@@
 > You could learn data analysis and figure out ways to apply it to other systems.
-> And you could be making a six-figure salary within the next five years.*" – [Benjamin Motz, Ph.D.](#fig-video-motz) [@NFL2014]
+> And you could be making a six-figure salary within the next five years.*
+> 
+> --- [Benjamin Motz, Ph.D.](#fig-video-motz) [@NFL2014]

### 2025-07-10 player-evaluation.qmd
20250708 - track chunk execution times
20250710 - quotes

@@
+{{< include _chunk-timing.qmd >}}
+
 # Player Evaluation {#sec-playerEvaluation}
@@
 
-> "The best predictor of future behavior is past behavior." – Unknown
+> "*The best predictor of future behavior is past behavior.*" – Unknown
 
-> "Past performance does not guarantee future results." – A common disclaimer about investments.
+> "*Past performance does not guarantee future results.*" – A common disclaimer about investments.
 

### 2025-07-10 sports-cognitive-psychology.qmd
20250710 - quotes

@@
 
-> We work a menu of plays in that area over the course of the week.
+> "*We work a menu of plays in that area over the course of the week.
 > We rank them at the latter part of the week, and then we get into the stadium and we play it by ear.
@@
 > That's an element of the game you can't measure.
-> Oftentimes, decisions such as that weigh heavily into the equation.
+> Oftentimes, decisions such as that weigh heavily into the equation.*"
 > 
@@
 
-> [Bruins Head Football Coach Kevin] Kelley believes that the "quant jocks" don't go far enough to validate the no-punting worldview and, more generally, the virtues of risk-taking.
-> "The math guys, the astrophysicist guys, they just do the raw numbers and they don't figure emotion into it—and that's the biggest thing of all," he says.
-> "The built-in emotion involved in football is unbelievable, and that's where the benefits really pay off."
+> [Bruins Head Football Coach Kevin] Kelley believes that the "*quant jocks*" don't go far enough to validate the no-punting worldview and, more generally, the virtues of risk-taking.
+> "*The math guys, the astrophysicist guys, they just do the raw numbers and they don't figure emotion into it—and that's the biggest thing of all,*" he says.
+> "*The built-in emotion involved in football is unbelievable, and that's where the benefits really pay off.*"
 > What he means is this: A defense that stops an opponent on third down is usually ecstatic.
@@
 > 
-> [@Moskowitz2011, p. 37]
+> --- [@Moskowitz2011, p. 37]
 

### 2025-07-10 factor-analysis.qmd
20250710 - players with highest levels on latent factors

@@
 dataMerged <- full_join(
-  player_stats_seasonal_avgPerGame,
+  player_stats_seasonal_avgPerGame %>% select(-any_of("week")),
   nfl_nextGenStats_seasonal %>% select(-any_of(c("player_display_name","completions","attempts","receptions","targets"))),
@@
 ```{r}
-lavaan::lavPredict(efa3factor_fit)
+efa3factor_factorScores <- lavaan::lavPredict(efa3factor_fit)
+
+head(efa3factor_factorScores)
 ```
@@
  # EFA Factor Loadings
- efa("efa1")*f1 + 
- efa("efa1")*f2 + 
- efa("efa1")*f3 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ efa("efa1")*F1 + 
+ efa("efa1")*F2 + 
+ efa("efa1")*F3 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
  passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
@@
 ```{r}
-lavaan::lavPredict(efa3factorModified_fit)
+efa3factorModified_factorScores <- lavaan::lavPredict(efa3factorModified_fit)
+
+head(efa3factorModified_factorScores)
 ```
@@
 )
+```
+
+Here are the variables that had a standardized loading greater than 0.3 on each of the factors:
+
+```{r}
+factor1vars <- c(
+  "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
+  "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame")
+
+factor2vars <- c(
+  "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
+  "avg_air_distance","max_air_distance","avg_air_yards_to_sticks")
+
+factor3vars <- c(
+  "passing_cpoe","pass_comp_pct","passer_rating","completion_percentage_above_expectation")
+```
+
+```{r}
+efa3factorModified_factorScores[which.max(efa3factorModified_factorScores[,"F1"]),]
+```
+
+```{r}
+dataForFA[which.max(efa3factorModified_factorScores[,"F1"]), c("player_display_name", "season", factor1vars)]
+```
+
+```{r}
+efa3factorModified_factorScores[which.max(efa3factorModified_factorScores[,"F2"]),]
+```
+
+```{r}
+dataForFA[which.max(efa3factorModified_factorScores[,"F2"]), c("player_display_name", "season", factor2vars)]
+```
+
+```{r}
+efa3factorModified_factorScores[which.max(efa3factorModified_factorScores[,"F3"]),]
+```
+
+```{r}
+dataForFA[which.max(cfa3factorModified_factorScores[,"F3"]), c("player_display_name", "season", factor3vars)]
 ```
@@
 ```{r}
-lavaan::lavPredict(cfa1factor_fit)
+cfa1factor_factorScores <- lavaan::lavPredict(cfa1factor_fit)
+
+head(cfa1factor_factorScores)
 ```
@@
 ```{r}
-lavaan::lavPredict(cfa2factor_fit)
+cfa2factor_factorScores <- lavaan::lavPredict(cfa2factor_fit)
+
+head(cfa2factor_factorScores)
 ```
@@
 ```{r}
-lavaan::lavPredict(cfa3factor_fit)
+cfa3factor_factorScores <- lavaan::lavPredict(cfa3factor_fit)
+
+head(cfa3factor_factorScores)
 ```
@@
  F2 ~~ 1*F2
- #F3 ~~ 1*F3
 '
@@
 ```{r}
-lavaan::lavPredict(cfa3factorModified_fit)
+cfa3factorModified_factorScores <- lavaan::lavPredict(cfa3factorModified_fit)
+
+head(cfa3factorModified_factorScores)
 ```
@@
 )
+```
+
+Here are the variables that loaded onto each of the factors:
+
+```{r}
+factor1vars <- c(
+  "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
+  "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame")
+
+factor2vars <- c(
+  "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
+  "avg_air_distance","max_air_distance","avg_air_yards_to_sticks")
+
+factor3vars <- c(
+  "passing_cpoe","pass_comp_pct","passer_rating","completion_percentage_above_expectation")
+```
+
+```{r}
+cfa3factorModified_factorScores[which.max(cfa3factorModified_factorScores[,"F1"]),]
+```
+
+```{r}
+dataForFA[which.max(cfa3factorModified_factorScores[,"F1"]), c("player_display_name", "season", factor1vars)]
+```
+
+```{r}
+cfa3factorModified_factorScores[which.max(cfa3factorModified_factorScores[,"F2"]),]
+```
+
+```{r}
+dataForFA[which.max(cfa3factorModified_factorScores[,"F2"]), c("player_display_name", "season", factor2vars)]
+```
+
+```{r}
+cfa3factorModified_factorScores[which.max(cfa3factorModified_factorScores[,"F3"]),]
+```
+
+```{r}
+dataForFA[which.max(cfa3factorModified_factorScores[,"F3"]), c("player_display_name", "season", factor3vars)]
 ```

### 2025-07-11 factor-analysis.qmd
20250710 - figure labels

@@
 
-A path diagram of the EFA model was created using the `lavaanPlot()` function of the `lavaanPlot` package [@R-lavaanPlot].
-
-```{r}
+A path diagram of the EFA model in @fig-efaPathDiagram-3factor was created using the `lavaanPlot()` function of the `lavaanPlot` package [@R-lavaanPlot].
+
+```{r}
+#| label: fig-efaPathDiagram-3factor
+#| fig-cap: "Path Diagram of the Three-Factor Exploratory Factor Analysis Model."
+#| fig-alt: "Path Diagram of the Three-Factor Exploratory Factor Analysis Model."
+
 lavaanPlot::lavaanPlot(
@@
 
-```{r}
+A path diagram of the three-factor ESEM model is in @fig-esemPathDiagram-3factor.
+
+```{r}
+#| label: fig-esemPathDiagram-3factor
+#| fig-cap: "Path Diagram of the Three-Factor Exploratory Structural Equation Model."
+#| fig-alt: "Path Diagram of the Three-Factor Exploratory Structural Equation Model."
+
 lavaanPlot::lavaanPlot(
@@
 
-```{r}
+The path diagram of the modified ESEM model with correlated residuals is in @fig-esemPathDiagram-3factorModified.
+
+```{r}
+#| label: fig-esemPathDiagram-3factorModified
+#| fig-cap: "Path Diagram of the Three-Factor Exploratory Structural Equation Model With Correlated Residuals."
+#| fig-alt: "Path Diagram of the Three-Factor Exploratory Structural Equation Model With Correlated Residuals."
+
 lavaanPlot::lavaanPlot(
@@
 
-```{r}
+The path diagram of the one-factor CFA model is in @fig-cfaPathDiagram-1factor.
+
+```{r}
+#| label: fig-cfaPathDiagram-1factor
+#| fig-cap: "Path Diagram of the One-Factor Confirmatory Factor Analysis Model."
+#| fig-alt: "Path Diagram of the One-Factor Confirmatory Factor Analysis Model."
+
 lavaanPlot::lavaanPlot(
@@
 
-```{r}
+The path diagram of the two-factor CFA model is in @fig-cfaPathDiagram-2factor.
+
+```{r}
+#| label: fig-cfaPathDiagram-2factor
+#| fig-cap: "Path Diagram of the Two-Factor Confirmatory Factor Analysis Model."
+#| fig-alt: "Path Diagram of the Two-Factor Confirmatory Factor Analysis Model."
+
 lavaanPlot::lavaanPlot(
@@
 
-```{r}
+The path diagram of the three-factor CFA model is in @fig-cfaPathDiagram-3factor.
+
+```{r}
+#| label: fig-cfaPathDiagram-3factor
+#| fig-cap: "Path Diagram of the Three-Factor Confirmatory Factor Analysis Model."
+#| fig-alt: "Path Diagram of the Three-Factor Confirmatory Factor Analysis Model."
+
 lavaanPlot::lavaanPlot(
@@
 
-```{r}
+The path diagram of the modified three-factor CFA model with correlated residuals is in @fig-cfaPathDiagram-3factorModified.
+
+```{r}
+#| label: fig-cfaPathDiagram-3factorModified
+#| fig-cap: "Path Diagram of the Modified Three-Factor Confirmatory Factor Analysis Model With Correlated Residuals."
+#| fig-alt: "Path Diagram of the Modified Three-Factor Confirmatory Factor Analysis Model With Correlated Residuals."
+
 lavaanPlot::lavaanPlot(

### 2025-07-11 factor-analysis.qmd
20250710 - lavaanPlot2()

@@
 
-A path diagram of the EFA model in @fig-efaPathDiagram-3factor was created using the `lavaanPlot()` function of the `lavaanPlot` package [@R-lavaanPlot].
+A path diagram of the EFA model in @fig-efaPathDiagram-3factor was created using the `lavaanPlot2()` function of the `lavaanPlot` package [@R-lavaanPlot].
 
@@
 
-lavaanPlot::lavaanPlot(
+lavaanPlot::lavaanPlot2(
   efa_fit$nf3,
-  coefs = TRUE,
+  include = "covs"
   stand = TRUE)
@@
 
-lavaanPlot::lavaanPlot(
+lavaanPlot::lavaanPlot2(
   efa3factor_fit,
-  coefs = TRUE,
+  include = "covs"
   stand = TRUE)
@@
 
-lavaanPlot::lavaanPlot(
+lavaanPlot::lavaanPlot2(
   efa3factorModified_fit,
-  coefs = TRUE,
+  include = "covs"
   stand = TRUE)
@@
 ```{r}
-dataForFA[which.max(cfa3factorModified_factorScores[,"F3"]), c("player_display_name", "season", factor3vars)]
+dataForFA[which.max(efa3factorModified_factorScores[,"F3"]), c("player_display_name", "season", factor3vars)]
 ```
@@
 
-lavaanPlot::lavaanPlot(
+lavaanPlot::lavaanPlot2(
   cfa1factor_fit,
-  coefs = TRUE,
+  include = "covs"
   stand = TRUE)
@@
 
-lavaanPlot::lavaanPlot(
+lavaanPlot::lavaanPlot2(
   cfa2factor_fit,
-  coefs = TRUE,
+  include = "covs"
   stand = TRUE)
@@
 
-lavaanPlot::lavaanPlot(
+lavaanPlot::lavaanPlot2(
   cfa3factor_fit,
-  coefs = TRUE,
+  include = "covs"
   stand = TRUE)
@@
 
-lavaanPlot::lavaanPlot(
+lavaanPlot::lavaanPlot2(
   cfa3factorModified_fit,
-  coefs = TRUE,
+  include = "covs"
   stand = TRUE)

### 2025-07-11 factor-analysis.qmd
20250710 - fix plots

--- 2025-07-11 factor-analysis.qmd
@@
   coefs = TRUE,
+  #covs = TRUE,
   stand = TRUE)
@@
   coefs = TRUE,
+  covs = TRUE,
   stand = TRUE)

### 2025-07-11 cognitive-bias.qmd
20250710 - quotes
20250711 - replacement of video content for PDF/ePub

@@
 :::
+
+:::
+
+::: {.content-visible unless-format="html:js"}
+
+@GetUpESPN2021 provides a video (archived at <https://perma.cc/JW8E-KV2C>) of Kyle Pitts drawing comparisons to the LeBron James of Tight Ends.
 
@@
 
-> "Hindsight is 20/20." – Idiom
+> "*Hindsight is 20/20.*" – Idiom
 
@@
 
-> "The more you know, the more you know you don't know." – Anonymous
+> "*The more you know, the more you know you don't know.*" – Anonymous
 

### 2025-07-11 sports-cognitive-psychology.qmd
20250711 - replacement of video content for PDF/ePub

@@
 His characterization suggests he believes that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
-The full game is [here](https://www.youtube.com/watch?v=1L9D2rp0mk8) [@ABC2015a].
+The full game is available to watch [here](https://www.youtube.com/watch?v=1L9D2rp0mk8) [@ABC2015a].
 
@@
 :::
+
+:::
+
+::: {.content-visible unless-format="html:js"}
+
+As an example of people dismissing analytics, there is a 2015 video clip of a football commentator, Mike Patrick, making fun of people who compile statistics for informing decision-making in football, calling the people who compile the numbers "guys who wear socks with flip-flops" [@ABC2015].
+His characterization suggests he believes that people who compile analytics to make recommendations in football are out of touch with the realities of football and do not have the same on-the-field understanding as the coaches and players.
+The full game is available to watch here: <https://www.youtube.com/watch?v=1L9D2rp0mk8> [@ABC2015a].
 

### 2025-07-11 pca.qmd
20250708 - track chunk execution times
20250711 - players highest and loading on each PCA component

@@
+{{< include _chunk-timing.qmd >}}
+
 # Data Reduction: Principal Component Analysis {#sec-pca}
@@
 
+```{r}
+dataMergedWithPCA <- cbind(dataMerged, pca3ComponentOblique$scores)
+```
+
 Here are the variables that had a standardized component loading of 0.4 or greater on each component:
@@
 ```{r}
-pca3ComponentOblique$scores[which.max(pca3ComponentOblique$scores[,1]),]
-dataMerged[which.max(pca3ComponentOblique$scores[,1]), c("player_display_name", "season", "week", component1vars)]
-
-pca3ComponentOblique$scores[which.min(pca3ComponentOblique$scores[,1]),]
-dataMerged[which.min(pca3ComponentOblique$scores[,1]), c("player_display_name", "season", "week", component1vars)]
+dataMergedWithPCA %>% 
+  arrange(-TC1) %>% 
+  head %>% 
+  select(player_display_name, season, week, all_of(component1vars))
 ```
@@
 ```{r}
-pca3ComponentOblique$scores[which.max(pca3ComponentOblique$scores[,2]),]
-dataMerged[which.max(pca3ComponentOblique$scores[,2]), c("player_display_name", "season", "week", component2vars)]
-
-pca3ComponentOblique$scores[which.min(pca3ComponentOblique$scores[,2]),]
-dataMerged[which.min(pca3ComponentOblique$scores[,2]), c("player_display_name", "season", "week", component2vars)]
+dataMergedWithPCA %>% 
+  arrange(TC1) %>% 
+  head %>% 
+  select(player_display_name, season, week, all_of(component1vars))
 ```
@@
 ```{r}
-pca3ComponentOblique$scores[which.max(pca3ComponentOblique$scores[,3]),]
-dataMerged[which.max(pca3ComponentOblique$scores[,3]), c("player_display_name", "season", "week", component3vars)]
+dataMergedWithPCA %>% 
+  arrange(-TC2) %>% 
+  head %>% 
+  select(player_display_name, season, week, all_of(component2vars))
+```
 
-pca3ComponentOblique$scores[which.min(pca3ComponentOblique$scores[,3]),]
-dataMerged[which.min(pca3ComponentOblique$scores[,3]), c("player_display_name", "season", "week", component3vars)]
+```{r}
+dataMergedWithPCA %>% 
+  arrange(TC2) %>% 
+  head %>% 
+  select(player_display_name, season, week, all_of(component2vars))
+```
+
+```{r}
+dataMergedWithPCA %>% 
+  arrange(-TC3) %>% 
+  head %>% 
+  select(player_display_name, season, week, all_of(component3vars))
+```
+
+```{r}
+dataMergedWithPCA %>% 
+  arrange(TC3) %>% 
+  head %>% 
+  select(player_display_name, season, week, all_of(component3vars))
 ```

### 2025-07-11 factor-analysis.qmd
20250711 - players with lowest/highest levels on components/factors

@@
 head(efa3factorModified_factorScores)
+
+dataForFA_efa <- cbind(dataForFA, efa3factorModified_factorScores)
 ```
@@
 ```{r}
-efa3factorModified_factorScores[which.max(efa3factorModified_factorScores[,"F1"]),]
-```
-
-```{r}
-dataForFA[which.max(efa3factorModified_factorScores[,"F1"]), c("player_display_name", "season", factor1vars)]
-```
-
-```{r}
-efa3factorModified_factorScores[which.max(efa3factorModified_factorScores[,"F2"]),]
-```
-
-```{r}
-dataForFA[which.max(efa3factorModified_factorScores[,"F2"]), c("player_display_name", "season", factor2vars)]
-```
-
-```{r}
-efa3factorModified_factorScores[which.max(efa3factorModified_factorScores[,"F3"]),]
-```
-
-```{r}
-dataForFA[which.max(efa3factorModified_factorScores[,"F3"]), c("player_display_name", "season", factor3vars)]
+dataForFA_efa %>% 
+  arrange(-F1) %>% 
+  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+dataForFA_efa %>% 
+  arrange(F1) %>% 
+  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+dataForFA_efa %>% 
+  arrange(-F2) %>% 
+  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+dataForFA_efa %>% 
+  arrange(F2) %>% 
+  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+dataForFA_efa %>% 
+  arrange(-F3) %>% 
+  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+dataForFA_efa %>% 
+  arrange(F3) %>% 
+  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
 ```
@@
 head(cfa3factorModified_factorScores)
+
+dataForFA_cfa <- cbind(dataForFA, cfa3factorModified_factorScores)
 ```
@@
 ```{r}
-cfa3factorModified_factorScores[which.max(cfa3factorModified_factorScores[,"F1"]),]
-```
-
-```{r}
-dataForFA[which.max(cfa3factorModified_factorScores[,"F1"]), c("player_display_name", "season", factor1vars)]
-```
-
-```{r}
-cfa3factorModified_factorScores[which.max(cfa3factorModified_factorScores[,"F2"]),]
-```
-
-```{r}
-dataForFA[which.max(cfa3factorModified_factorScores[,"F2"]), c("player_display_name", "season", factor2vars)]
-```
-
-```{r}
-cfa3factorModified_factorScores[which.max(cfa3factorModified_factorScores[,"F3"]),]
-```
-
-```{r}
-dataForFA[which.max(cfa3factorModified_factorScores[,"F3"]), c("player_display_name", "season", factor3vars)]
+dataForFA_cfa %>% 
+  arrange(-F1) %>% 
+  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(F1) %>% 
+  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(-F2) %>% 
+  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(F2) %>% 
+  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(-F3) %>% 
+  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(F3) %>% 
+  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
 ```

### 2025-07-11 pca.qmd
20250711 - players with lowest/highest levels on components/factors

@@
   arrange(-TC1) %>% 
-  head %>% 
-  select(player_display_name, season, week, all_of(component1vars))
+  select(player_display_name, season, week, TC1, all_of(component1vars)) %>% 
+  na.omit() %>% 
+  head()
 ```
@@
   arrange(TC1) %>% 
-  head %>% 
-  select(player_display_name, season, week, all_of(component1vars))
+  select(player_display_name, season, week, TC1, all_of(component1vars)) %>% 
+  na.omit() %>% 
+  head()
 ```
@@
   arrange(-TC2) %>% 
-  head %>% 
-  select(player_display_name, season, week, all_of(component2vars))
+  select(player_display_name, season, week, TC2, all_of(component2vars)) %>% 
+  na.omit() %>% 
+  head()
 ```
@@
   arrange(TC2) %>% 
-  head %>% 
-  select(player_display_name, season, week, all_of(component2vars))
+  select(player_display_name, season, week, TC2, all_of(component2vars)) %>% 
+  na.omit() %>% 
+  head()
 ```
@@
   arrange(-TC3) %>% 
-  head %>% 
-  select(player_display_name, season, week, all_of(component3vars))
+  select(player_display_name, season, week, TC3, all_of(component3vars)) %>% 
+  na.omit() %>% 
+  head()
 ```
@@
   head %>% 
-  select(player_display_name, season, week, all_of(component3vars))
+  select(player_display_name, season, week, TC3, all_of(component3vars)) %>% 
+  na.omit() %>% 
+  head()
 ```

### 2025-07-11 pca.qmd
20250711 - PCA interpretation/labelling the components

@@
 
+The variables in component 1 appear to reflect Quarterback usage: completions, incompletions, passing attempts, passing yards, passing touchdowns, interceptions thrown, fumbles, times sacked, sack yards lost ("reversed"—i.e., negatively associated with the component), sack fumbles, sack fumbles lost, passing air yards (total horizontal distance the ball travels on all pass attempts), passing yards after the catch, and first downs gained by passing.
+Quarterbacks who tend to throw more tend to have higher levels on those variables.
+Thus, we label component 1 as "Usage", which reflects total Quarterback involvement, regardless of efficiency or outcome.
+
+The variables in component 2 appear to reflect Quarterback aggressiveness: passing air yards, passing air conversion ratio (reversed; ratio of passing yards to passing air yards), average air yards on completed passes, average air yards on all attempted passes, aggressiveness (percentage of passing attempts thrown into tight windows, where there is a defender within one yard or less of the receiver at the time of the completion or incompletion), expected completion percentage (reversed; based on air distance, receiver separation, Quarterback/Wide Receiver movement, pass location, whether there was pressure on the Quarterback when throwing, the throw angle and trajectory, receiver and defender positioning at the catch point, and defensive coverage scheme), average amount of air yards ahead of or behind the first down marker on passing attempts, average air distance (the true three-dimensional distance the ball travels in the air), maximum air distance, and maximum air distance on completed passes.
+Quarterbacks who throw the ball farther and into tighter windows tend to have higher values on those variables.
+Thus, we label component 2 as "Aggressiveness", which reflects throwing longer, more difficult passes with a tight window.
+
+The variables in component 3 appear to reflect Quarterback performance: passing touchdowns, passing expected points added, passing completion percentage above expectation, passes completed of 40 yards or more, pass completion percentage, air yards differential (intended air yards $-$ completed air yards; attempting deeper passes than he on average completes); maximum completed air distance, average time to throw, average completed air yards, average intended air yards, and passer rating.
+Quarterbacks who perform better tend to have higher values on those variables.
+Thus, we label component 3 as "Performance".
+
 ```{r}
@@
 ```
+
+Here are the players and weeks that showed the highest levels of Quarterback "Usage":
 
@@
 
+Here are the players and weeks that showed the lowest levels of Quarterback "Usage":
+
 ```{r}
@@
 ```
+
+Here are the players and weeks that showed the highest levels of Quarterback "Aggressiveness":
 
@@
 
+Here are the players and weeks that showed the lowest levels of Quarterback "Aggressiveness":
+
 ```{r}
@@
 
+Here are the players and weeks that showed the highest levels of Quarterback "Performance":
+
 ```{r}
@@
 ```
+
+Here are the players and weeks that showed the lowest levels of Quarterback "Performance":
 

### 2025-07-12 factor-analysis.qmd
20250711 - interpretation of factors/components

@@
 
+The variables that loaded most strongly onto factor 1 appear to reflect Quarterback usage: completions per game, passing attempts per game, passing yards per game, passing touchdowns per game, passing air yards (total horizontal distance the ball travels on all pass attempts) per game, passing yards after the catch per game, and first downs gained per game by passing.
+Quarterbacks who tend to throw more tend to have higher levels on those variables.
+Thus, we label component 1 as "Usage", which reflects total Quarterback involvement, regardless of efficiency or outcome.
+
+The variables that loaded most strongly onto factor 2 appear to reflect Quarterback aggressiveness: average air yards on completed passes, average air yards on all attempted passes, aggressiveness (percentage of passing attempts thrown into tight windows, where there is a defender within one yard or less of the receiver at the time of the completion or incompletion), average amount of air yards ahead of or behind the first down marker on passing attempts, average air distance (the true three-dimensional distance the ball travels in the air), maximum air distance, and maximum air distance on completed passes.
+Quarterbacks who throw the ball farther and into tighter windows tend to have higher values on those variables.
+Thus, we label component 2 as "Aggressiveness", which reflects throwing longer, more difficult passes with a tight window.
+
+The variables that loaded most strongly onto factor 3 appear to reflect Quarterback performance: passing completion percentage above expectation, pass completion percentage, and passer rating.
+Quarterbacks who perform better tend to have higher values on those variables.
+Thus, we label component 3 as "Performance".
+
+Here are the players and seasons that showed the highest levels of Quarterback "Usage":
+
 ```{r}
@@
 
+Here are the players and seasons that showed the lowest levels of Quarterback "Usage":
+
 ```{r}
@@
 
+Here are the players and seasons that showed the highest levels of Quarterback "Aggressiveness":
+
 ```{r}
@@
 
+Here are the players and seasons that showed the lowest levels of Quarterback "Aggressiveness":
+
 ```{r}
@@
 
+Here are the players and seasons that showed the highest levels of Quarterback "Performance":
+
 ```{r}
@@
   arrange(-F3) %>% 
+  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the highest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_efa %>% 
+  arrange(-F3) %>% 
+  filter(games >= 10) %>% 
+  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the lowest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_efa %>% 
+  arrange(F3) %>% 
   select(player_display_name, season, F3, all_of(factor3vars)) %>% 
@@
 
+If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the lowest levels of Quarterback "Performance":
+
 ```{r}
@@
   arrange(F3) %>% 
-  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
+  filter(games >= 10) %>% 
+  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
   na.omit() %>% 
@@
 
+If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the highest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(-F3) %>% 
+  filter(games >= 10) %>% 
+  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the lowest levels of Quarterback "Performance":
+
 ```{r}
@@
   select(player_display_name, season, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the lowest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(F3) %>% 
+  filter(games >= 10) %>% 
+  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
   na.omit() %>% 

### 2025-07-12 pca.qmd
20250711 - interpretation of factors/components

--- 2025-07-11 pca.qmd
@@
 
+The variables that loaded most strongly onto component 1 appear to reflect Quarterback usage: completions, incompletions, passing attempts, passing yards, passing touchdowns, interceptions thrown, fumbles, times sacked, sack yards lost ("reversed"—i.e., negatively associated with the component), sack fumbles, sack fumbles lost, passing air yards (total horizontal distance the ball travels on all pass attempts), passing yards after the catch, and first downs gained by passing.
+Quarterbacks who tend to throw more tend to have higher levels on those variables.
+Thus, we label component 1 as "Usage", which reflects total Quarterback involvement, regardless of efficiency or outcome.
+
+The variables that loaded most strongly onto component 2 appear to reflect Quarterback aggressiveness: passing air yards, passing air conversion ratio (reversed; ratio of passing yards to passing air yards), average air yards on completed passes, average air yards on all attempted passes, aggressiveness (percentage of passing attempts thrown into tight windows, where there is a defender within one yard or less of the receiver at the time of the completion or incompletion), expected completion percentage (reversed; based on air distance, receiver separation, Quarterback/Wide Receiver movement, pass location, whether there was pressure on the Quarterback when throwing, the throw angle and trajectory, receiver and defender positioning at the catch point, and defensive coverage scheme), average amount of air yards ahead of or behind the first down marker on passing attempts, average air distance (the true three-dimensional distance the ball travels in the air), maximum air distance, and maximum air distance on completed passes.
+Quarterbacks who throw the ball farther and into tighter windows tend to have higher values on those variables.
+Thus, we label component 2 as "Aggressiveness", which reflects throwing longer, more difficult passes with a tight window.
+
+The variables that loaded most strongly onto component 3 appear to reflect Quarterback performance: passing touchdowns, passing expected points added, passing completion percentage above expectation, passes completed of 40 yards or more, pass completion percentage, air yards differential (intended air yards $-$ completed air yards; attempting deeper passes than he on average completes), maximum completed air distance, average time to throw, average completed air yards, average intended air yards, and passer rating.
+Quarterbacks who perform better tend to have higher values on those variables.
+Thus, we label component 3 as "Performance".
+
 ```{r}

### 2025-07-12 multiple-regression.qmd
20250711 - exponentiate regression coefficients

@@
 Under the hood, multiple regression seeks to identify the best weight for each [predictor](#sec-correlationalStudy).
+The intercept is the expected value of the outcome variable when all of the predictor variables have a value of zero.
 
@@
 print(effectsize::standardize_parameters(logisticRegression, method = "refit"), digits = 2)
+
+broom::tidy(
+  logisticRegression,
+  exponentiate = TRUE,
+  conf.int = TRUE)
 ```
@@
 print(effectsize::standardize_parameters(ordinalRegression, method = "refit"), digits = 2)
+
+broom::tidy(
+  ordinalRegression,
+  exponentiate = TRUE,
+  conf.int = TRUE)
 ```
@@
 print(effectsize::standardize_parameters(poissonRegression, method = "refit"), digits = 2)
+
+broom::tidy(
+  poissonRegression,
+  exponentiate = TRUE,
+  conf.int = TRUE)
 ```
@@
 print(effectsize::standardize_parameters(negativeBinomialRegression, method = "refit"), digits = 2)
+
+broom::tidy(
+  negativeBinomialRegression,
+  exponentiate = TRUE,
+  conf.int = TRUE)
 ```

### 2025-07-12 factor-analysis.qmd
20250712 - add interpretation

@@
 # Factor Analysis {#sec-factorAnalysis}
+
+> "All models are wrong, but some are useful."
+>
+> --- George Box [-@Box1979, p. 202]
 
@@
 A factor loading is a regression path from a latent factor to an observed (manifest) variable.\index{factor analysis}\index{factor analysis!factor loading}\index{latent variable}
-The standardized factor loading represents the strength of association between the variable and the latent factor.\index{factor analysis}\index{factor analysis!factor loading}\index{latent variable}\index{standardized regression coefficient}
+The standardized factor loading represents the strength of association between the variable and the latent factor, where conceptually, it is intended to reflect the magnitude that the latent factor influences the observed variable.\index{factor analysis}\index{factor analysis!factor loading}\index{latent variable}\index{standardized regression coefficient}
 A residual is variance in a variable (or factor) that is unexplained by other variables or factors.\index{factor analysis}\index{factor analysis!residual}
 A variable's intercept is the expected value of the variable when the factor(s) (onto which it loads) is equal to zero.\index{factor analysis}\index{factor analysis!intercept}
-A covariance is the unstandardized index of the strength of association between between variables (or factors), and it is depicted with a double-headed arrow.\index{factor analysis}\index{factor analysis!covariance}
+A [*covariance*](#sec-correlationExamplesCovariance) is the unstandardized index of the strength of association between between two variables (or factors), and it is depicted with a double-headed arrow.\index{factor analysis}\index{factor analysis!covariance}
+
+Because a covariance is unstandardized, its scale depends on the scale of the variables.
+The covariance between two variables is the average product of their deviations from their respective means, as in @eq-covariance.
+The covariance of a variable with itself is equivalent to its variance, as in @eq-covarianceToVariance.
+By contrast, a [*correlation*](#sec-correlation) is a standardized index of the strength of association between two variables.
+Because a correlation is standardized (fixed between [−1,1]), its scale does not depend on the scales of the variables.
 Because a covariance is unstandardized, its scale depends on the scale of the variables.\index{factor analysis!covariance}
@@
 When the latent factors are uncorrelated, the (standardized) error term for an indicator is calculated as 1 minus the sum of squared standardized factor loadings for a given item (including cross-loadings).\index{factor analysis}\index{factor analysis!residual}\index{factor analysis!factor loading}\index{cross-loading}
-A cross-loadings is when a variable loads onto more than one latent factor.\index{factor analysis}\index{cross-loading}
+A cross-loading is when a variable loads onto more than one latent factor.\index{factor analysis}\index{cross-loading}
 
@@
 
-## Exploratory Factor Analysis {#sec-efa}
+## Path Diagrams {#sec-pathDiagrams}
+
+A key tool when designing a structural equation model is a conceptual depiction of the hypothesized causal processes.
+A path diagram depicts the hypothesized causal processes that link two or more variables.
+@Karch2025_packages provides a tool to create `lavaan` `R` syntax from a path analytic diagram: <https://lavaangui.org>.
+
+In a path analysis diagram, rectangles represent variables we observe, and circles represent latent (i.e., unobserved) variables.
+Single-headed arrows indicate regression paths, where conceptually, one variable is thought to influence another variable.
+Double-headed arrows indicate covariance paths, where conceptually, two variables are associated for some unknown reason (i.e., an omitted shared cause).
+
+## Decisions in Factor Analysis {#sec-factorAnalysisDecisions}
+
+There are five primary decisions to make in factor analysis:
+
+1. what variables to include in the model and how to scale them
+1. method of factor extraction: whether to use exploratory or confirmatory factor analysis
+1. if using exploratory factor analysis, whether and how to rotate factors
+1. how many factors to retain (and what variables load onto which factors)
+1. how to interpret and use the factors
+
+The answer you get can differ highly depending on the decisions you make.
+Below, we provide guidance for each of these decisions.
+
+### 1. Variables to Include and their Scaling {#sec-variablesToInclude-factorAnalysis}
+
+The first decision when conducting a factor analysis is which variables to include and the scaling of those variables.
+What factors you extract can differ widely depending on what variables you include in the analysis.
+For example, if you include many variables from the same source (e.g., self-report), it is possible that you will extract a factor that represents the common variance among the variables from that source (i.e., the self-reported variables).
+This would be considered a method factor, which works against the goal of estimating latent factors that represent the constructs of interest (as opposed to the measurement methods used to estimate those constructs).
+
+An additional consideration is the scaling of the variables: whether to use raw variables, standardized variables, or dividing some variables by a constant to make the variables' variances more similar.
+Before performing a [principal component analysis](#sec-pca) (PCA), it is generally important to ensure that the variables included in the PCA are on the same scale.
+PCA seeks to identify components that explain variance in the data, so if the variables are not on the same scale, some variables may contribute considerably more variance than others.
+A common way of ensuring that variables are on the same scale is to standardize them using, for example,  [*z*-scores](#zScores) that have a mean of zero and standard deviation of one.
+By contrast, factor analysis can better accommodate variables that are on different scales.
+
+### 2. Method of Factor Extraction {#sec-methodOfFactorExtraction}
+
+#### Exploratory Factor Analysis {#sec-efa}
 
@@
 
-### Factor Rotation {#sec-factorAnalysisRotation}
+#### Confirmatory Factor Analysis {#sec-cfa}
+
+Confirmatory factor analysis (CFA) is used to (dis)confirm a priori hypotheses about the factor structure of the model.\index{factor analysis!confirmatory}
+CFA is a test of the hypothesis.\index{factor analysis!confirmatory}
+In CFA, you specify the model and ask how well this model represents the data.\index{factor analysis!confirmatory}
+The researcher specifies the number, meaning, associations, and pattern of free parameters in the factor loading matrix [@Bollen2002].\index{factor analysis!confirmatory}
+A key advantage of CFA is the ability to directly compare alternative models (i.e., factor structures), which is valuable for theory testing [@Strauss2009].\index{factor analysis!confirmatory}
+For instance, you could use [CFA](#cfa) to test whether the variance in several measures' scores is best explained with one factor or two factors.\index{factor analysis!confirmatory}
+In CFA, cross-loadings are not estimated unless the researcher specifies them.\index{cross-loading}\index{factor analysis!confirmatory}
+
+### 3. Factor Rotation {#sec-factorAnalysisRotation}
 
@@
 
-### Example {#sec-efaExample}
-
-```{r}
-#| eval: false
-#| include: false
-
-cor_mat <- cor(dataForFA[faVars], use = "pairwise.complete.obs")
-
-cor_tidy <- cor_mat %>%
-  as.data.frame() %>%
-  rownames_to_column("var1") %>%
-  pivot_longer(-var1, names_to = "var2", values_to = "correlation") %>%
-  filter(var1 != var2) %>% # remove self-correlations
-  mutate(pair = pmap_chr(list(var1, var2), ~ paste(sort(c(...)), collapse = "___"))) %>% 
-  distinct(pair, .keep_all = TRUE) # remove duplicate pairs (symmetry)
-
-high_corrs <- cor_tidy %>%
-  filter(abs(correlation) > 0.85) %>%
-  arrange(desc(abs(correlation)))
-
-high_corrs
-
-cov_mat <- cov(dataForFA[, faVars], use = "pairwise.complete.obs")
-eigen_cov_mat <- eigen(cov_mat)
-print(eigen_cov_mat$values)
-
-# Index of smallest (most problematic) eigenvalue
-idx <- which.min(eigen_cov_mat$values)
-
-# Extract the associated eigenvector
-problem_vector <- eigen_cov_mat$vectors[, idx]
-names(problem_vector) <- faVars
-
-# Sort by absolute magnitude to identify most influential variables
-sort(abs(problem_vector), decreasing = TRUE)
-```
-
-We generated the scree plot in @fig-efaScreePlot1 using the `fa.parallel()` function of the `psych` package [@R-psych].
-
-```{r}
-#| label: fig-efaScreePlot1
-#| fig-cap: "Scree Plot: With Comparisons to Simulated and Resampled Data."
-#| fig-alt: "Scree Plot: With Comparisons to Simulated and Resampled Data."
-
-psych::fa.parallel(
-  x = dataForFA[faVars],
-  fm = "minres", # errors out when using "ml"
-  fa = "fa"
-)
-```
-
-We generated the scree plot in @fig-efaScreePlot2 using the `nScree()` and `plotnScree()` functions of the `nFactors` package [@R-nFactors].
-
-```{r}
-#| label: fig-efaScreePlot2
-#| fig-cap: "Scree Plot with Parallel Analysis."
-#| fig-alt: "Scree Plot with Parallel Analysis."
-
-#screeDataEFA <- nFactors::nScree(
-#  x = cor( # throws error with correlation matrix, so use covariance instead (below)
-#    dataForFA[faVars],
-#    use = "pairwise.complete.obs"),
-#  model = "factors")
-
-screeDataEFA <- nFactors::nScree(
-  x = cov(
-    dataForFA[faVars],
-    use = "pairwise.complete.obs"),
-  cor = FALSE,
-  model = "factors")
-
-nFactors::plotnScree(screeDataEFA)
-```
-
-We generated the very simple structure (VSS) plots in Figures [-@fig-efaVSSPlot1] and [-@fig-efaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
-
-```{r}
-#| label: fig-efaVSSPlot1
-#| fig-cap: "Very Simple Structure Plot."
-#| fig-alt: "Very Simple Structure Plot."
-
-psych::vss(
-  dataForFA[faVars],
-  rotate = "oblimin",
-  fm = "minres") # errors out when using "mle"
-```
-
-```{r}
-#| label: fig-efaVSSPlot2
-#| fig-cap: "Model Indices by Number of Factors."
-#| fig-alt: "Model Indices by Number of Factors."
-
-psych::nfactors(
-  dataForFA[faVars],
-  rotate = "oblimin",
-  fm = "minres") # errors out when using "ml"
-```
-
-We fit EFA models using the `efa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
-
-```{r}
-#| label: efa
-
-efa_fit <- lavaan::efa(
-  data = dataForFA,
-  ov.names = faVars,
-  nfactors = 1:7,
-  rotation = "geomin",
-  missing = "ML",
-  estimator = "MLR",
-  bounds = "standard",
-  meanstructure = TRUE,
-  em.h1.iter.max = 2000000)
-```
-
-```{r}
-#| label: efa-summary
-
-summary(efa_fit)
-```
-
-A path diagram of the EFA model in @fig-efaPathDiagram-3factor was created using the `lavaanPlot()` function of the `lavaanPlot` package [@R-lavaanPlot].
-
-```{r}
-#| label: fig-efaPathDiagram-3factor
-#| fig-cap: "Path Diagram of the Three-Factor Exploratory Factor Analysis Model."
-#| fig-alt: "Path Diagram of the Three-Factor Exploratory Factor Analysis Model."
-
-lavaanPlot::lavaanPlot(
-  efa_fit$nf3,
-  coefs = TRUE,
-  #covs = TRUE,
-  stand = TRUE)
-```
-
-Here is the syntax for estimating a three-factor EFA using exploratory structural equation modeling (ESEM).
-Estimating the model in a ESEM framework allows us to make modifications to the model, such as adding correlated residuals, and adding predictors or outcomes of the latent factors.
-
-```{r}
-#| label: efa-3factor-syntax
-
-efa3factor_syntax <- '
- # EFA Factor Loadings
- efa("efa1")*f1 + 
- efa("efa1")*f2 + 
- efa("efa1")*f3 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
- passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
- avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
- avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
- passer_rating + completion_percentage_above_expectation
-'
-```
-
-To fit the ESEM model, we use the `sem()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
-
-```{r}
-#| label: efa-3factor-fit
-
-efa3factor_fit <- sem(
-  efa3factor_syntax,
-  data = dataForFA,
-  information = "observed",
-  missing = "ML",
-  estimator = "MLR",
-  rotation = "geomin",
-  bounds = "standard",
-  meanstructure = TRUE,
-  em.h1.iter.max = 2000000)
-```
-
-```{r}
-summary(
-  efa3factor_fit,
-  fit.measures = TRUE,
-  standardized = TRUE,
-  rsquare = TRUE)
-```
-
-```{r}
-lavaan::fitMeasures(
-  efa3factor_fit,
-  fit.measures = c(
-    "chisq", "df", "pvalue",
-    "chisq.scaled", "df.scaled", "pvalue.scaled",
-    "chisq.scaling.factor",
-    "baseline.chisq","baseline.df","baseline.pvalue",
-    "rmsea", "cfi", "tli", "srmr",
-    "rmsea.robust", "cfi.robust", "tli.robust"))
-```
-
-```{r}
-lavaan::residuals(
-  efa3factor_fit,
-  type = "cor")
-```
-
-```{r}
-lavaan::modificationindices(
-  efa3factor_fit,
-  sort. = TRUE)
-```
-
-```{r}
-efa3factor_factorScores <- lavaan::lavPredict(efa3factor_fit)
-
-head(efa3factor_factorScores)
-```
-
-A path diagram of the three-factor ESEM model is in @fig-esemPathDiagram-3factor.
-
-```{r}
-#| label: fig-esemPathDiagram-3factor
-#| fig-cap: "Path Diagram of the Three-Factor Exploratory Structural Equation Model."
-#| fig-alt: "Path Diagram of the Three-Factor Exploratory Structural Equation Model."
-
-lavaanPlot::lavaanPlot(
-  efa3factor_fit,
-  coefs = TRUE,
-  covs = TRUE,
-  stand = TRUE)
-```
-
-To make the plot interactive for editing, you can use the `plot_lavaan()` function of the `lavaangui` package [@R-lavaangui; @Karch2025_packages]:
-
-```{r}
-#| eval: false
-
-lavaangui::plot_lavaan(efa3factor_fit)
-```
-
-Here is a modification of the 3-factor model with correlated residuals:
-
-```{r}
-#| label: efa-3factorModified-syntax
-
-efa3factorModified_syntax <- '
- # EFA Factor Loadings
- efa("efa1")*F1 + 
- efa("efa1")*F2 + 
- efa("efa1")*F3 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
- passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
- avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
- avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
- passer_rating + completion_percentage_above_expectation
-
- # Correlated Residuals
- completionsPerGame ~~ attemptsPerGame
- passing_yardsPerGame ~~ passing_yards_after_catchPerGame
- attemptsPerGame ~~ passing_yardsPerGame
- attemptsPerGame ~~ passing_air_yardsPerGame
- passing_air_yardsPerGame ~~ passing_yards_after_catchPerGame
- passing_yards_after_catchPerGame ~~ avg_completed_air_yards
- passing_tdsPerGame ~~ passer_rating
- completionsPerGame ~~ passing_air_yardsPerGame
- max_completed_air_distance ~~ max_air_distance
- aggressiveness ~~ completion_percentage_above_expectation
-'
-```
-
-```{r}
-#| label: efa-3factorModified-fit
-
-efa3factorModified_fit <- sem(
-  efa3factorModified_syntax,
-  data = dataForFA,
-  information = "observed",
-  missing = "ML",
-  estimator = "MLR",
-  rotation = "geomin",
-  bounds = "standard",
-  meanstructure = TRUE,
-  em.h1.iter.max = 2000000)
-```
-
-```{r}
-summary(
-  efa3factorModified_fit,
-  fit.measures = TRUE,
-  standardized = TRUE,
-  rsquare = TRUE)
-```
-
-```{r}
-lavaan::fitMeasures(
-  efa3factorModified_fit,
-  fit.measures = c(
-    "chisq", "df", "pvalue",
-    "chisq.scaled", "df.scaled", "pvalue.scaled",
-    "chisq.scaling.factor",
-    "baseline.chisq","baseline.df","baseline.pvalue",
-    "rmsea", "cfi", "tli", "srmr",
-    "rmsea.robust", "cfi.robust", "tli.robust"))
-```
-
-```{r}
-lavaan::residuals(
-  efa3factorModified_fit,
-  type = "cor")
-```
-
-```{r}
-lavaan::modificationindices(
-  efa3factorModified_fit,
-  sort. = TRUE)
-```
-
-```{r}
-efa3factorModified_factorScores <- lavaan::lavPredict(efa3factorModified_fit)
-
-head(efa3factorModified_factorScores)
-
-dataForFA_efa <- cbind(dataForFA, efa3factorModified_factorScores)
-```
-
-The path diagram of the modified ESEM model with correlated residuals is in @fig-esemPathDiagram-3factorModified.
-
-```{r}
-#| label: fig-esemPathDiagram-3factorModified
-#| fig-cap: "Path Diagram of the Three-Factor Exploratory Structural Equation Model With Correlated Residuals."
-#| fig-alt: "Path Diagram of the Three-Factor Exploratory Structural Equation Model With Correlated Residuals."
-
-lavaanPlot::lavaanPlot(
-  efa3factorModified_fit,
-  coefs = TRUE,
-  #covs = TRUE,
-  stand = TRUE)
-```
-
-```{r}
-#| eval: false
-
-lavaangui::plot_lavaan(efa3factorModified_fit)
-```
-
-```{r}
-anova(
-  efa3factorModified_fit,
-  efa3factor_fit
-)
-```
-
-Here are the variables that had a standardized loading greater than 0.3 on each of the factors:
-
-```{r}
-factor1vars <- c(
-  "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
-  "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame")
-
-factor2vars <- c(
-  "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
-  "avg_air_distance","max_air_distance","avg_air_yards_to_sticks")
-
-factor3vars <- c(
-  "passing_cpoe","pass_comp_pct","passer_rating","completion_percentage_above_expectation")
-```
-
-The variables that loaded most strongly onto factor 1 appear to reflect Quarterback usage: completions per game, passing attempts per game, passing yards per game, passing touchdowns per game, passing air yards (total horizontal distance the ball travels on all pass attempts) per game, passing yards after the catch per game, and first downs gained per game by passing.
-Quarterbacks who tend to throw more tend to have higher levels on those variables.
-Thus, we label component 1 as "Usage", which reflects total Quarterback involvement, regardless of efficiency or outcome.
-
-The variables that loaded most strongly onto factor 2 appear to reflect Quarterback aggressiveness: average air yards on completed passes, average air yards on all attempted passes, aggressiveness (percentage of passing attempts thrown into tight windows, where there is a defender within one yard or less of the receiver at the time of the completion or incompletion), average amount of air yards ahead of or behind the first down marker on passing attempts, average air distance (the true three-dimensional distance the ball travels in the air), maximum air distance, and maximum air distance on completed passes.
-Quarterbacks who throw the ball farther and into tighter windows tend to have higher values on those variables.
-Thus, we label component 2 as "Aggressiveness", which reflects throwing longer, more difficult passes with a tight window.
-
-The variables that loaded most strongly onto factor 3 appear to reflect Quarterback performance: passing completion percentage above expectation, pass completion percentage, and passer rating.
-Quarterbacks who perform better tend to have higher values on those variables.
-Thus, we label component 3 as "Performance".
-
-Here are the players and seasons that showed the highest levels of Quarterback "Usage":
-
-```{r}
-dataForFA_efa %>% 
-  arrange(-F1) %>% 
-  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-Here are the players and seasons that showed the lowest levels of Quarterback "Usage":
-
-```{r}
-dataForFA_efa %>% 
-  arrange(F1) %>% 
-  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-Here are the players and seasons that showed the highest levels of Quarterback "Aggressiveness":
-
-```{r}
-dataForFA_efa %>% 
-  arrange(-F2) %>% 
-  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-Here are the players and seasons that showed the lowest levels of Quarterback "Aggressiveness":
-
-```{r}
-dataForFA_efa %>% 
-  arrange(F2) %>% 
-  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-Here are the players and seasons that showed the highest levels of Quarterback "Performance":
-
-```{r}
-dataForFA_efa %>% 
-  arrange(-F3) %>% 
-  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the highest levels of Quarterback "Performance":
-
-```{r}
-dataForFA_efa %>% 
-  arrange(-F3) %>% 
-  filter(games >= 10) %>% 
-  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-Here are the players and seasons that showed the lowest levels of Quarterback "Performance":
-
-```{r}
-dataForFA_efa %>% 
-  arrange(F3) %>% 
-  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the lowest levels of Quarterback "Performance":
-
-```{r}
-dataForFA_efa %>% 
-  arrange(F3) %>% 
-  filter(games >= 10) %>% 
-  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-## Confirmatory Factor Analysis {#sec-cfa}
-
-Confirmatory factor analysis (CFA) is used to (dis)confirm a priori hypotheses about the factor structure of the model.\index{factor analysis!confirmatory}
-CFA is a test of the hypothesis.\index{factor analysis!confirmatory}
-In CFA, you specify the model and ask how well this model represents the data.\index{factor analysis!confirmatory}
-The researcher specifies the number, meaning, associations, and pattern of free parameters in the factor loading matrix [@Bollen2002].\index{factor analysis!confirmatory}
-A key advantage of CFA is the ability to directly compare alternative models (i.e., factor structures), which is valuable for theory testing [@Strauss2009].\index{factor analysis!confirmatory}
-For instance, you could use [CFA](#cfa) to test whether the variance in several measures' scores is best explained with one factor or two factors.\index{factor analysis!confirmatory}
-In CFA, cross-loadings are not estimated unless the researcher specifies them.\index{cross-loading}\index{factor analysis!confirmatory}
-
-### Example {#sec-cfaExample}
-
-```{r}
-#| eval: false
-#| include: false
-
-psych::principal(
-  dataForFA_standardized[faVars],
-  nfactors = 3,
-  rotate = "oblimin")
-
-'
- #Factor loadings
- F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + passing_interceptionsPerGame + passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
- F2 =~ pacr + avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + avg_air_yards_to_sticks + expected_completion_percentage + avg_air_distance + max_air_distance
- F3 =~ passing_epa + passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
-'
-
-factor1_syntax <- '
- # Factor loadings
- F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
-
- # Correlated residuals
- passing_air_yardsPerGame ~~ passing_yards_after_catchPerGame
- completionsPerGame ~~ attemptsPerGame
- attemptsPerGame ~~ passing_yards_after_catchPerGame
- passing_yards_after_catchPerGame ~~ passing_first_downsPerGame
- completionsPerGame ~~ passing_first_downsPerGame
- attemptsPerGame ~~ passing_tdsPerGame
- completionsPerGame ~~ passing_tdsPerGame
- passing_tdsPerGame ~~ passing_air_yardsPerGame
-'
-
-factor1_fit <- lavaan::cfa(
-  factor1_syntax,
-  data = dataForFA,
-  missing = "ML",
-  estimator = "MLR",
-  bounds = "standard",
-  std.lv = TRUE,
-  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
-  em.h1.iter.max = 2000000)
-
-summary(
-  factor1_fit,
-  fit.measures = TRUE,
-  standardized = TRUE,
-  rsquare = TRUE)
-
-lavaan::modificationindices(
-  factor1_fit,
-  sort. = TRUE)
-
-factor2_syntax <- '
- # Factor loadings
- F2 =~ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + avg_air_distance + max_air_distance + avg_air_yards_to_sticks #expected_completion_percentage + pacr
-
- # Correlated residuals
- max_completed_air_distance ~~ max_air_distance
- avg_completed_air_yards ~~ max_completed_air_distance
- avg_completed_air_yards ~~ avg_air_distance
- avg_completed_air_yards ~~ max_air_distance
- avg_intended_air_yards ~~ max_completed_air_distance
-'
-
-factor2_fit <- lavaan::cfa(
-  factor2_syntax,
-  data = dataForFA,
-  missing = "ML",
-  estimator = "MLR",
-  bounds = "standard",
-  std.lv = TRUE,
-  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
-  em.h1.iter.max = 2000000)
-
-summary(
-  factor2_fit,
-  fit.measures = TRUE,
-  standardized = TRUE,
-  rsquare = TRUE)
-
-lavaan::modificationindices(
-  factor2_fit,
-  sort. = TRUE)
-
-factor3_syntax <- '
- #Factor loadings
- F3 =~ passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
-'
-
-factor3_fit <- lavaan::cfa(
-  factor3_syntax,
-  data = dataForFA,
-  missing = "ML",
-  estimator = "MLR",
-  bounds = "standard",
-  std.lv = TRUE,
-  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
-  em.h1.iter.max = 2000000)
-
-summary(
-  factor3_fit,
-  fit.measures = TRUE,
-  standardized = TRUE,
-  rsquare = TRUE)
-
-lavaan::modificationindices(
-  factor3_fit,
-  sort. = TRUE)
-```
-
-We fit CFA models using the `cfa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
-
-```{r}
-cfa1factor_syntax <- '
- #Factor loadings
- F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
- passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
- avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
- avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
- passer_rating + completion_percentage_above_expectation
-'
-```
-
-```{r}
-#| label: cfa-1factor-fit
-
-cfa1factor_fit <- lavaan::cfa(
-  cfa1factor_syntax,
-  data = dataForFA,
-  missing = "ML",
-  estimator = "MLR",
-  bounds = "standard",
-  std.lv = TRUE,
-  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
-  em.h1.iter.max = 2000000)
-```
-
-```{r}
-summary(
-  cfa1factor_fit,
-  fit.measures = TRUE,
-  standardized = TRUE,
-  rsquare = TRUE)
-```
-
-```{r}
-lavaan::fitMeasures(
-  cfa1factor_fit,
-  fit.measures = c(
-    "chisq", "df", "pvalue",
-    "chisq.scaled", "df.scaled", "pvalue.scaled",
-    "chisq.scaling.factor",
-    "baseline.chisq","baseline.df","baseline.pvalue",
-    "rmsea", "cfi", "tli", "srmr",
-    "rmsea.robust", "cfi.robust", "tli.robust"))
-```
-
-```{r}
-lavaan::residuals(
-  cfa1factor_fit,
-  type = "cor")
-```
-
-```{r}
-lavaan::modificationindices(
-  cfa1factor_fit,
-  sort. = TRUE)
-```
-
-```{r}
-cfa1factor_factorScores <- lavaan::lavPredict(cfa1factor_fit)
-
-head(cfa1factor_factorScores)
-```
-
-The path diagram of the one-factor CFA model is in @fig-cfaPathDiagram-1factor.
-
-```{r}
-#| label: fig-cfaPathDiagram-1factor
-#| fig-cap: "Path Diagram of the One-Factor Confirmatory Factor Analysis Model."
-#| fig-alt: "Path Diagram of the One-Factor Confirmatory Factor Analysis Model."
-
-lavaanPlot::lavaanPlot(
-  cfa1factor_fit,
-  coefs = TRUE,
-  #covs = TRUE,
-  stand = TRUE)
-```
-
-```{r}
-#| eval: false
-
-lavaangui::plot_lavaan(cfa1factor_fit)
-```
-
-```{r}
-cfa2factor_syntax <- '
- #Factor loadings
- F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
- passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
-
- F2 =~ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
- avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
- passer_rating + completion_percentage_above_expectation
-'
-```
-
-```{r}
-#| label: cfa-2factor-fit
-
-cfa2factor_fit <- lavaan::cfa(
-  cfa2factor_syntax,
-  data = dataForFA,
-  missing = "ML",
-  estimator = "MLR",
-  bounds = "standard",
-  std.lv = TRUE,
-  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
-  em.h1.iter.max = 2000000)
-```
-
-```{r}
-summary(
-  cfa2factor_fit,
-  fit.measures = TRUE,
-  standardized = TRUE,
-  rsquare = TRUE)
-```
-
-```{r}
-lavaan::fitMeasures(
-  cfa2factor_fit,
-  fit.measures = c(
-    "chisq", "df", "pvalue",
-    "chisq.scaled", "df.scaled", "pvalue.scaled",
-    "chisq.scaling.factor",
-    "baseline.chisq","baseline.df","baseline.pvalue",
-    "rmsea", "cfi", "tli", "srmr",
-    "rmsea.robust", "cfi.robust", "tli.robust"))
-```
-
-```{r}
-lavaan::residuals(
-  cfa2factor_fit,
-  type = "cor")
-```
-
-```{r}
-lavaan::modificationindices(
-  cfa2factor_fit,
-  sort. = TRUE)
-```
-
-```{r}
-cfa2factor_factorScores <- lavaan::lavPredict(cfa2factor_fit)
-
-head(cfa2factor_factorScores)
-```
-
-The path diagram of the two-factor CFA model is in @fig-cfaPathDiagram-2factor.
-
-```{r}
-#| label: fig-cfaPathDiagram-2factor
-#| fig-cap: "Path Diagram of the Two-Factor Confirmatory Factor Analysis Model."
-#| fig-alt: "Path Diagram of the Two-Factor Confirmatory Factor Analysis Model."
-
-lavaanPlot::lavaanPlot(
-  cfa2factor_fit,
-  coefs = TRUE,
-  covs = TRUE,
-  stand = TRUE)
-```
-
-```{r}
-#| eval: false
-
-lavaangui::plot_lavaan(cfa2factor_fit)
-```
-
-```{r}
-anova(
-  cfa2factor_fit,
-  cfa1factor_fit
-)
-```
-
-```{r}
-cfa3factor_syntax <- '
- #Factor loadings
- F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
- passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
-
- F2 =~ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
- avg_air_distance + max_air_distance + avg_air_yards_to_sticks
-
- F3 =~ passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
-'
-```
-
-```{r}
-#| label: cfa-3factor-fit
-
-cfa3factor_fit <- lavaan::cfa(
-  cfa3factor_syntax,
-  data = dataForFA,
-  missing = "ML",
-  estimator = "MLR",
-  bounds = "standard",
-  std.lv = TRUE,
-  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
-  em.h1.iter.max = 2000000)
-```
-
-```{r}
-summary(
-  cfa3factor_fit,
-  fit.measures = TRUE,
-  standardized = TRUE,
-  rsquare = TRUE)
-```
-
-```{r}
-lavaan::fitMeasures(
-  cfa3factor_fit,
-  fit.measures = c(
-    "chisq", "df", "pvalue",
-    "chisq.scaled", "df.scaled", "pvalue.scaled",
-    "chisq.scaling.factor",
-    "baseline.chisq","baseline.df","baseline.pvalue",
-    "rmsea", "cfi", "tli", "srmr",
-    "rmsea.robust", "cfi.robust", "tli.robust"))
-```
-
-```{r}
-lavaan::residuals(
-  cfa3factor_fit,
-  type = "cor")
-```
-
-```{r}
-lavaan::modificationindices(
-  cfa3factor_fit,
-  sort. = TRUE)
-```
-
-```{r}
-cfa3factor_factorScores <- lavaan::lavPredict(cfa3factor_fit)
-
-head(cfa3factor_factorScores)
-```
-
-The path diagram of the three-factor CFA model is in @fig-cfaPathDiagram-3factor.
-
-```{r}
-#| label: fig-cfaPathDiagram-3factor
-#| fig-cap: "Path Diagram of the Three-Factor Confirmatory Factor Analysis Model."
-#| fig-alt: "Path Diagram of the Three-Factor Confirmatory Factor Analysis Model."
-
-lavaanPlot::lavaanPlot(
-  cfa3factor_fit,
-  coefs = TRUE,
-  covs = TRUE,
-  stand = TRUE)
-```
-
-```{r}
-#| eval: false
-
-lavaangui::plot_lavaan(cfa3factor_fit)
-```
-
-```{r}
-anova(
-  cfa3factor_fit,
-  cfa2factor_fit
-)
-```
-
-```{r}
-cfa3factorModified_syntax <- '
- # Factor loadings
- F1 =~ NA*completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
- passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
-
- F2 =~ NA*avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
- avg_air_distance + max_air_distance + avg_air_yards_to_sticks
-
- F3 =~ NA*passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
-
- # Cross loadings
- #F3 =~ attemptsPerGame
-
- # Correlated residuals
- passing_air_yardsPerGame ~~ passing_yards_after_catchPerGame
- completionsPerGame ~~ attemptsPerGame
- attemptsPerGame ~~ passing_yards_after_catchPerGame
- passing_yards_after_catchPerGame ~~ passing_first_downsPerGame
- completionsPerGame ~~ passing_first_downsPerGame
- attemptsPerGame ~~ passing_tdsPerGame
- completionsPerGame ~~ passing_tdsPerGame
- passing_tdsPerGame ~~ passing_air_yardsPerGame
-
- max_completed_air_distance ~~ max_air_distance
- avg_completed_air_yards ~~ max_completed_air_distance
- avg_completed_air_yards ~~ avg_air_distance
- avg_completed_air_yards ~~ max_air_distance
- avg_intended_air_yards ~~ max_completed_air_distance
-
- completionsPerGame ~~ pass_comp_pct
- passing_yardsPerGame ~~ avg_completed_air_yards
- passing_yardsPerGame ~~ max_completed_air_distance
- passing_tdsPerGame ~~ passer_rating
- aggressiveness ~~ completion_percentage_above_expectation
-
- # Variances
- F1 ~~ 1*F1
- F2 ~~ 1*F2
-'
-```
-
-```{r}
-#| label: cfa-3factorModified-fit
-
-cfa3factorModified_fit <- lavaan::cfa(
-  cfa3factorModified_syntax,
-  data = dataForFA,
-  missing = "ML",
-  estimator = "MLR",
-  bounds = "standard",
-  #std.lv = TRUE,
-  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
-  em.h1.iter.max = 2000000)
-```
-
-```{r}
-summary(
-  cfa3factorModified_fit,
-  fit.measures = TRUE,
-  standardized = TRUE,
-  rsquare = TRUE)
-```
-
-```{r}
-lavaan::fitMeasures(
-  cfa3factorModified_fit,
-  fit.measures = c(
-    "chisq", "df", "pvalue",
-    "chisq.scaled", "df.scaled", "pvalue.scaled",
-    "chisq.scaling.factor",
-    "baseline.chisq","baseline.df","baseline.pvalue",
-    "rmsea", "cfi", "tli", "srmr",
-    "rmsea.robust", "cfi.robust", "tli.robust"))
-```
-
-```{r}
-lavaan::residuals(
-  cfa3factorModified_fit,
-  type = "cor")
-```
-
-```{r}
-lavaan::modificationindices(
-  cfa3factorModified_fit,
-  sort. = TRUE)
-```
-
-```{r}
-cfa3factorModified_factorScores <- lavaan::lavPredict(cfa3factorModified_fit)
-
-head(cfa3factorModified_factorScores)
-
-dataForFA_cfa <- cbind(dataForFA, cfa3factorModified_factorScores)
-```
-
-The path diagram of the modified three-factor CFA model with correlated residuals is in @fig-cfaPathDiagram-3factorModified.
-
-```{r}
-#| label: fig-cfaPathDiagram-3factorModified
-#| fig-cap: "Path Diagram of the Modified Three-Factor Confirmatory Factor Analysis Model With Correlated Residuals."
-#| fig-alt: "Path Diagram of the Modified Three-Factor Confirmatory Factor Analysis Model With Correlated Residuals."
-
-lavaanPlot::lavaanPlot(
-  cfa3factorModified_fit,
-  coefs = TRUE,
-  #covs = TRUE,
-  stand = TRUE)
-```
-
-```{r}
-#| eval: false
-
-lavaangui::plot_lavaan(cfa3factorModified_fit)
-```
-
-```{r}
-anova(
-  cfa3factorModified_fit,
-  cfa3factor_fit
-)
-```
-
-Here are the variables that loaded onto each of the factors:
-
-```{r}
-factor1vars <- c(
-  "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
-  "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame")
-
-factor2vars <- c(
-  "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
-  "avg_air_distance","max_air_distance","avg_air_yards_to_sticks")
-
-factor3vars <- c(
-  "passing_cpoe","pass_comp_pct","passer_rating","completion_percentage_above_expectation")
-```
-
-The variables that loaded most strongly onto factor 1 appear to reflect Quarterback usage: completions per game, passing attempts per game, passing yards per game, passing touchdowns per game, passing air yards (total horizontal distance the ball travels on all pass attempts) per game, passing yards after the catch per game, and first downs gained per game by passing.
-Quarterbacks who tend to throw more tend to have higher levels on those variables.
-Thus, we label component 1 as "Usage", which reflects total Quarterback involvement, regardless of efficiency or outcome.
-
-The variables that loaded most strongly onto factor 2 appear to reflect Quarterback aggressiveness: average air yards on completed passes, average air yards on all attempted passes, aggressiveness (percentage of passing attempts thrown into tight windows, where there is a defender within one yard or less of the receiver at the time of the completion or incompletion), average amount of air yards ahead of or behind the first down marker on passing attempts, average air distance (the true three-dimensional distance the ball travels in the air), maximum air distance, and maximum air distance on completed passes.
-Quarterbacks who throw the ball farther and into tighter windows tend to have higher values on those variables.
-Thus, we label component 2 as "Aggressiveness", which reflects throwing longer, more difficult passes with a tight window.
-
-The variables that loaded most strongly onto factor 3 appear to reflect Quarterback performance: passing completion percentage above expectation, pass completion percentage, and passer rating.
-Quarterbacks who perform better tend to have higher values on those variables.
-Thus, we label component 3 as "Performance".
-
-Here are the players and seasons that showed the highest levels of Quarterback "Usage":
-
-```{r}
-dataForFA_cfa %>% 
-  arrange(-F1) %>% 
-  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-Here are the players and seasons that showed the lowest levels of Quarterback "Usage":
-
-```{r}
-dataForFA_cfa %>% 
-  arrange(F1) %>% 
-  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-Here are the players and seasons that showed the highest levels of Quarterback "Aggressiveness":
-
-```{r}
-dataForFA_cfa %>% 
-  arrange(-F2) %>% 
-  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-Here are the players and seasons that showed the lowest levels of Quarterback "Aggressiveness":
-
-```{r}
-dataForFA_cfa %>% 
-  arrange(F2) %>% 
-  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-Here are the players and seasons that showed the highest levels of Quarterback "Performance":
-
-```{r}
-dataForFA_cfa %>% 
-  arrange(-F3) %>% 
-  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the highest levels of Quarterback "Performance":
-
-```{r}
-dataForFA_cfa %>% 
-  arrange(-F3) %>% 
-  filter(games >= 10) %>% 
-  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-Here are the players and seasons that showed the lowest levels of Quarterback "Performance":
-
-```{r}
-dataForFA_cfa %>% 
-  arrange(F3) %>% 
-  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the lowest levels of Quarterback "Performance":
-
-```{r}
-dataForFA_cfa %>% 
-  arrange(F3) %>% 
-  filter(games >= 10) %>% 
-  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
-  na.omit() %>% 
-  head()
-```
-
-## Determining the Number of Factors to Retain {#sec-factorAnalysisNumFactors}
+### 4. Determining the Number of Factors to Retain {#sec-factorAnalysisNumFactors}
 
@@
 The hope is that you can have fewer factors that explain the associations between the variables than the number of observed variables.\index{factor analysis}
+It does not make sense to replace 18 variables with 18 latent factors because that would not result in any simplification.
 But how do you decide on the number of factors?\index{factor analysis}
@@
 In a parallel analysis, you examine where the eigenvalues from observed data and random data converge, so you do not retain a factor that explains less variance than would be expected by random chance.\index{factor analysis}\index{parallel analysis}\index{eigenvalue}
-
-In general, my recommendation is to use Cattell's scree test, and then test the factor solutions with plus or minus one factor.\index{factor analysis}\index{scree plot}
-You should never accept factors with eigenvalues less than zero (or components from [principal component analysis](#sec-pca) with eigenvalues less than one), because they are likely to be largely composed of error.\index{factor analysis}\index{eigenvalue}
+Using the VSS criterion, the optimal number of factors to retain is the number of factors that maximizes the VSS criterion [@Revelle1979].
+The VSS criterion is evaluated with models in which factor loadings for a given item that are less than the maximum factor loading for that item are suppressed to zero, thus forcing simple structure (i.e., no cross-loadings).
+The goal is finding a factor structure with interpretability so that factors are clearly distinguishable.
+Thus, we want to identify the number of factors with the highest VSS criterion.
+
+In general, my recommendation is to use Cattell's scree test, and then test the factor solutions with plus or minus one factor, in addition to examining [model fit](#sec-modelFitIndices).\index{factor analysis}\index{scree plot}
+You should never accept factors with eigenvalues less than zero (or components from [PCA](#sec-pca) with eigenvalues less than one), because they are likely to be largely composed of error.\index{factor analysis}\index{eigenvalue}
 If you are using maximum likelihood factor analysis, you can compare the fit of various models with model fit criteria to see which model fits best for its parsimony.\index{parsimony}\index{factor analysis}
@@
 
-## Interpreting and Using Latent Factors {#sec-interpretingAndUsingLatentFactors}
+#### Model Fit Indices {#sec-modelFitIndices}
+
+In factor analysis, we fit a model to observed data, or to the variance-covariance matrix, and we evaluate the degree of model misfit.
+That is, fit indices evaluate how likely it is that a given causal model gave rise to the observed data.
+Various model fit indices can be used for evaluating how well a model fits the data and for comparing the fit of two competing models.
+Fit indices known as absolute fit indices compare whether the model fits better than the best-possible fitting model (i.e., a saturated model).
+Examples of absolute fit indices include the chi-square test, root mean square error of approximation (RMSEA), and the standardized root mean square residual (SRMR).
+
+The chi-square test evaluates whether the model has a significant degree of misfit relative to the best-possible fitting model (a saturated model that fits as many parameters as possible; i.e., as many parameters as there are degrees of freedom); the null hypothesis of a chi-square test is that there is no difference between the predicted data (i.e., the data that would be observed if the model were true) and the observed data.
+Thus, a non-significant chi-square test indicates good model fit.
+However, because the null hypothesis of the chi-square test is that the model-implied covariance matrix is exactly equal to the observed covariance matrix (i.e., a model of perfect fit), this may be an unrealistic comparison.
+Models are simplifications of reality, and our models are virtually never expected to be a perfect description of reality.
+Thus, we would say a model is "useful" and partially validated if "it helps us to understand the relation between variables and does a 'reasonable' job of matching the data...A perfect fit may be an inappropriate standard, and a high chi-square estimate may indicate what we already know—that the hypothesized model holds approximately, not perfectly." [@Bollen1989, p.
+268].
+The power of the chi-square test depends on sample size, and a large sample will likely detect small differences as significantly worse than the best-possible fitting model [@Bollen1989].
+
+RMSEA is an index of absolute fit.
+Lower values indicate better fit.
+
+SRMR is an index of absolute fit with no penalty for model complexity.
+Lower values indicate better fit.
+
+There are also various fit indices known as incremental, comparative, or relative fit indices that compare whether the model fits better than the worst-possible fitting model (i.e., a "baseline" or "null" model).
+Incremental fit indices include a chi-square difference test, the comparative fit index (CFI), and the Tucker-Lewis index (TLI).
+Unlike the chi-square test comparing the model to the best-possible fitting model, a significant chi-square test of the relative fit index indicates better fit—i.e., that the model fits better than the worst-possible fitting model.
+
+CFI is another relative fit index that compares the model to the worst-possible fitting model.
+Higher values indicate better fit.
+
+TLI is another relative fit index.
+Higher values indicate better fit.
+
+Parsimony fit include fit indices that use information criteria fit indices, including the Akaike Information Criterion (AIC) and the Bayesian Information Criterion (BIC).\index{parsimony}
+BIC penalizes model complexity more so than AIC.
+Lower AIC and BIC values indicate better fit.
+
+Chi-square difference tests and CFI can be used to compare two nested models.
+AIC and BIC can be used to compare two non-nested models.
+
+Criteria for acceptable fit and good fit of SEM models are in @semFitIndices.
+In addition, dynamic fit indexes have been proposed based on simulation to identify fit index cutoffs that are tailored to the characteristics of the specific model and data [@McNeish2023]; dynamic fit indexes are available via the `dynamic` package [@R-dynamic] or with a [webapp](https://dynamicfit.app).
+
+| SEM Fit Index | Acceptable Fit | Good Fit   |
+|---------------|----------------|------------|
+| RMSEA         | $\leq$ .08     | $\leq$ .05 |
+| CFI           | $\geq$ .90     | $\geq$ .95 |
+| TLI           | $\geq$ .90     | $\geq$ .95 |
+| SRMR          | $\leq$ .10     | $\leq$ .08 |
+
+: Criteria for Acceptable and Good Fit of Structural Equation Models Based on Fit Indices. {#tbl-semFitIndices}
+
+However, good model fit does not necessarily indicate a true model.
+
+In addition to global fit indices, it can also be helpful to examine evidence of local fit, such as the residual covariance matrix.
+The residual covariance matrix represents the difference between the observed covariance matrix and the model-implied covariance matrix (the observed covariance matrix minus the model-implied covariance matrix).
+These difference values are called *covariance residuals*.
+Standardizing the covariance matrix by converting each to a correlation matrix can be helpful for interpreting the magnitude of any local misfit.
+This is known as a residual correlation matrix, which is composed of *correlation residuals*.
+Correlation residuals greater than |.10| are possible evidence for poor local fit [@Kline2023].
+If a correlation residual is positive, it suggests that the model underpredicts the observed association between the two variables (i.e., the observed covariance is greater than the model-implied covariance).
+If a correlation residual is negative, it suggests that the model overpredicts their observed association between the two variables (i.e., the observed covariance is smaller than the model-implied covariance).
+If the two variables are connected by only indirect pathways, it may be helpful to respecify the model with direct pathways between the two variables, such as a direct effect (i.e., regression path) or a covariance path.
+For guidance on evaluating local fit, see @Kline2024.
+
+### 5. Interpreting and Using Latent Factors {#sec-interpretingAndUsingLatentFactors}
 
@@
 Use theory to interpret the model and label the factors.\index{factor analysis}
+In latent variable models, factors have meaning.
+You can use them as predictors, mediators, moderators, or outcomes.
+When possible, it is preferable to use the factors by examining their associations with other variables in the same model.
+You can extract factor scores, if necessary, for use in other analyses; however, it is preferable to examine the associations between factors and other variables in the same model if possible.
+And, using latent factors helps disattenuate associations for measurement error, to identify what the association is between variables when removing random measurement error.
+
+## Example of Exploratory Factor Analysis {#sec-efaExample}
+
+```{r}
+#| eval: false
+#| include: false
+
+cor_mat <- cor(dataForFA[faVars], use = "pairwise.complete.obs")
+
+cor_tidy <- cor_mat %>%
+  as.data.frame() %>%
+  rownames_to_column("var1") %>%
+  pivot_longer(-var1, names_to = "var2", values_to = "correlation") %>%
+  filter(var1 != var2) %>% # remove self-correlations
+  mutate(pair = pmap_chr(list(var1, var2), ~ paste(sort(c(...)), collapse = "___"))) %>% 
+  distinct(pair, .keep_all = TRUE) # remove duplicate pairs (symmetry)
+
+high_corrs <- cor_tidy %>%
+  filter(abs(correlation) > 0.85) %>%
+  arrange(desc(abs(correlation)))
+
+high_corrs
+
+cov_mat <- cov(dataForFA[, faVars], use = "pairwise.complete.obs")
+eigen_cov_mat <- eigen(cov_mat)
+print(eigen_cov_mat$values)
+
+# Index of smallest (most problematic) eigenvalue
+idx <- which.min(eigen_cov_mat$values)
+
+# Extract the associated eigenvector
+problem_vector <- eigen_cov_mat$vectors[, idx]
+names(problem_vector) <- faVars
+
+# Sort by absolute magnitude to identify most influential variables
+sort(abs(problem_vector), decreasing = TRUE)
+```
+
+We generated the scree plot in @fig-efaScreePlot1 using the `fa.parallel()` function of the `psych` package [@R-psych].
+The optimal coordinates and the acceleration factor attempt to operationalize the Cattell scree test: i.e., the "elbow" of the scree plot [@Ruscio2012].
+The optimal coordinators factor is quantified using a series of linear equations to determine whether observed eigenvalues exceed the predicted values.
+The acceleration factor is quantified using the acceleration of the curve, that is, the second derivative.
+The Kaiser-Guttman rule states to keep principal components whose eigenvalues are greater than 1.
+However, for [exploratory factor analysis](#efa) (as opposed to [PCA](#pca)), the criterion is to keep the factors whose eigenvalues are greater than zero (i.e., not the factors whose eigenvalues are greater than 1) [@Dinno2014].
+
+The number of factors to keep would depend on which criteria one uses.
+Based on the rule to keep factors whose eigenvalues are greater than zero and based on the parallel test, we would keep five factors.
+However, based on the Cattell scree test (the "elbow" of the screen plot minus one), we would keep three factors.
+If using the optimal coordinates, we would keep eight factors; if using the acceleration factor, we would keep one factor.
+Therefore, interpretability of the factors would be important for deciding how many factors to keep.
+
+```{r}
+#| label: fig-efaScreePlot1
+#| fig-cap: "Scree Plot: With Comparisons to Simulated and Resampled Data."
+#| fig-alt: "Scree Plot: With Comparisons to Simulated and Resampled Data."
+
+psych::fa.parallel(
+  x = dataForFA[faVars],
+  fm = "minres", # errors out when using "ml"
+  fa = "fa"
+)
+```
+
+We generated the scree plot in @fig-efaScreePlot2 using the `nScree()` and `plotnScree()` functions of the `nFactors` package [@R-nFactors].
+
+```{r}
+#| label: fig-efaScreePlot2
+#| fig-cap: "Scree Plot with Parallel Analysis."
+#| fig-alt: "Scree Plot with Parallel Analysis."
+
+#screeDataEFA <- nFactors::nScree(
+#  x = cor( # throws error with correlation matrix, so use covariance instead (below)
+#    dataForFA[faVars],
+#    use = "pairwise.complete.obs"),
+#  model = "factors")
+
+screeDataEFA <- nFactors::nScree(
+  x = cov(
+    dataForFA[faVars],
+    use = "pairwise.complete.obs"),
+  cor = FALSE,
+  model = "factors")
+
+nFactors::plotnScree(screeDataEFA)
+```
+
+We generated the very simple structure (VSS) plots in Figures [-@fig-efaVSSPlot1] and [-@fig-efaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
+In addition to VSS plots, the output also provides additional criteria by which to determine the optimal number of factors, each for which lower values are better, including the Velicer minimum average partial (MAP) test, the Bayesian information criterion (BIC), the sample size-adjusted BIC (SABIC), and the root mean square error of approximation (RMSEA).
+Depending on the criterion, the optimal number of factors extracted varies between 2 and 8 factors.
+
+```{r}
+#| label: fig-efaVSSPlot1
+#| fig-cap: "Very Simple Structure Plot."
+#| fig-alt: "Very Simple Structure Plot."
+
+psych::vss(
+  dataForFA[faVars],
+  rotate = "oblimin",
+  fm = "minres") # errors out when using "mle"
+```
+
+```{r}
+#| label: fig-efaVSSPlot2
+#| fig-cap: "Model Indices by Number of Factors."
+#| fig-alt: "Model Indices by Number of Factors."
+
+psych::nfactors(
+  dataForFA[faVars],
+  rotate = "oblimin",
+  fm = "minres") # errors out when using "ml"
+```
+
+We fit EFA models using the `efa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
+
+```{r}
+#| label: efa
+
+efa_fit <- lavaan::efa(
+  data = dataForFA,
+  ov.names = faVars,
+  nfactors = 1:7,
+  rotation = "geomin",
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  meanstructure = TRUE,
+  em.h1.iter.max = 2000000)
+```
+
+The model fits well according to CFI with 5 or more factors; the model fits well according to RMSEA with 6 or more factors.
+However, 5+ factors does not represent much of a simplification (relative to the 18 variables included).
+Moreover, in the model with four factors, only one variable had a significant loading on Factor 1.
+Thus, even with four factors, one of the factors does not seem to represent the aggregation of multiple variables.
+For these reasons, and because the "elbow test" for the scree plot suggested three factors, we decided to retain three factors and to see if we could achieve better fit by making additional model modifications (e.g., correlated residuals).
+Correlated residuals may be necessary, for example, when variables are correlated for reasons other than the latent factors.
+
+```{r}
+#| label: efa-summary
+
+summary(efa_fit)
+```
+
+A path diagram of the three-factor EFA model in @fig-efaPathDiagram-3factor was created using the `lavaanPlot()` function of the `lavaanPlot` package [@R-lavaanPlot].
+
+```{r}
+#| label: fig-efaPathDiagram-3factor
+#| fig-cap: "Path Diagram of the Three-Factor Exploratory Factor Analysis Model."
+#| fig-alt: "Path Diagram of the Three-Factor Exploratory Factor Analysis Model."
+
+lavaanPlot::lavaanPlot(
+  efa_fit$nf3,
+  coefs = TRUE,
+  #covs = TRUE,
+  stand = TRUE)
+```
+
+Here is the syntax for estimating a three-factor EFA using exploratory structural equation modeling (ESEM).
+Estimating the model in a ESEM framework allows us to make modifications to the model, such as adding correlated residuals, and adding predictors or outcomes of the latent factors. 
+The syntax below represents the same model (with the same fit indices) as the three-factor EFA model above.
+
+```{r}
+#| label: efa-3factor-syntax
+
+efa3factor_syntax <- '
+ # EFA Factor Loadings
+ efa("efa1")*f1 + 
+ efa("efa1")*f2 + 
+ efa("efa1")*f3 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
+ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
+ passer_rating + completion_percentage_above_expectation
+'
+```
+
+To fit the ESEM model, we use the `sem()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
+
+```{r}
+#| label: efa-3factor-fit
+
+efa3factor_fit <- sem(
+  efa3factor_syntax,
+  data = dataForFA,
+  information = "observed",
+  missing = "ML",
+  estimator = "MLR",
+  rotation = "geomin",
+  bounds = "standard",
+  meanstructure = TRUE,
+  em.h1.iter.max = 2000000)
+```
+
+The fit indices suggests that the model does not fit well to the data and that additional model modifications are necessary.
+The fit indices are below.
+
+```{r}
+summary(
+  efa3factor_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+```{r}
+lavaan::fitMeasures(
+  efa3factor_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+lavaan::residuals(
+  efa3factor_fit,
+  type = "cor")
+```
+
+We can examine the model modification indices to identify parameters that, if estimated, would substantially improve model fit.
+For instance, the modification indices below indicate additional correlated residuals that could substantially improve model fit.
+However, it is generally not recommended to blindly estimate additional parameters solely based on modification indices, which can lead to data dredging and [overfitting](#sec-overfitting).
+Rather, it is generally advised to consider modification indices in light of theory.
+Based on the modification indices, we will add several correlated residuals to the model, to help account for why variables are associated with each other for reasons other than their underlying latent factors.
+
+```{r}
+lavaan::modificationindices(
+  efa3factor_fit,
+  sort. = TRUE)
+```
+
+Below are factor scores from the model for the first six players:
+
+```{r}
+efa3factor_factorScores <- lavaan::lavPredict(efa3factor_fit)
+
+head(efa3factor_factorScores)
+```
+
+A path diagram of the three-factor ESEM model is in @fig-esemPathDiagram-3factor.
+
+```{r}
+#| label: fig-esemPathDiagram-3factor
+#| fig-cap: "Path Diagram of the Three-Factor Exploratory Structural Equation Model."
+#| fig-alt: "Path Diagram of the Three-Factor Exploratory Structural Equation Model."
+
+lavaanPlot::lavaanPlot(
+  efa3factor_fit,
+  coefs = TRUE,
+  covs = TRUE,
+  stand = TRUE)
+```
+
+To make the plot interactive for editing, you can use the `plot_lavaan()` function of the `lavaangui` package [@R-lavaangui; @Karch2025_packages]:
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(efa3factor_fit)
+```
+
+Below is a modification of the three-factor model with correlated residuals.
+For instance, it makes sense that passing completions and attempts are related to each other (even after accounting for their latent factor).
+
+```{r}
+#| label: efa-3factorModified-syntax
+
+efa3factorModified_syntax <- '
+ # EFA Factor Loadings
+ efa("efa1")*F1 + 
+ efa("efa1")*F2 + 
+ efa("efa1")*F3 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
+ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
+ passer_rating + completion_percentage_above_expectation
+
+ # Correlated Residuals
+ completionsPerGame ~~ attemptsPerGame
+ passing_yardsPerGame ~~ passing_yards_after_catchPerGame
+ attemptsPerGame ~~ passing_yardsPerGame
+ attemptsPerGame ~~ passing_air_yardsPerGame
+ passing_air_yardsPerGame ~~ passing_yards_after_catchPerGame
+ passing_yards_after_catchPerGame ~~ avg_completed_air_yards
+ passing_tdsPerGame ~~ passer_rating
+ completionsPerGame ~~ passing_air_yardsPerGame
+ max_completed_air_distance ~~ max_air_distance
+ aggressiveness ~~ completion_percentage_above_expectation
+'
+```
+
+```{r}
+#| label: efa-3factorModified-fit
+
+efa3factorModified_fit <- sem(
+  efa3factorModified_syntax,
+  data = dataForFA,
+  information = "observed",
+  missing = "ML",
+  estimator = "MLR",
+  rotation = "geomin",
+  bounds = "standard",
+  meanstructure = TRUE,
+  em.h1.iter.max = 2000000)
+```
+
+```{r}
+summary(
+  efa3factorModified_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+The model fits substantially better (though not perfectly) with the additional correlated residuals.
+Below are the model fit indices:
+
+```{r}
+lavaan::fitMeasures(
+  efa3factorModified_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+lavaan::residuals(
+  efa3factorModified_fit,
+  type = "cor")
+```
+
+```{r}
+lavaan::modificationindices(
+  efa3factorModified_fit,
+  sort. = TRUE)
+```
+
+Below are factor scores from the model for the first six players:
+
+```{r}
+efa3factorModified_factorScores <- lavaan::lavPredict(efa3factorModified_fit)
+
+head(efa3factorModified_factorScores)
+
+dataForFA_efa <- cbind(dataForFA, efa3factorModified_factorScores)
+```
+
+The path diagram of the modified ESEM model with correlated residuals is in @fig-esemPathDiagram-3factorModified.
+
+```{r}
+#| label: fig-esemPathDiagram-3factorModified
+#| fig-cap: "Path Diagram of the Three-Factor Exploratory Structural Equation Model With Correlated Residuals."
+#| fig-alt: "Path Diagram of the Three-Factor Exploratory Structural Equation Model With Correlated Residuals."
+
+lavaanPlot::lavaanPlot(
+  efa3factorModified_fit,
+  coefs = TRUE,
+  #covs = TRUE,
+  stand = TRUE)
+```
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(efa3factorModified_fit)
+```
+
+Model fit of nested models can be compared with a chi-square difference test.
+This allows us to evaluate whether the more complex model fits signficantly better than the less complex model.
+In this case, our more complex model is the three-factor model with correlated residuals; the less complex model is the three-factor model without correlated residuals.
+The modified model with the correlated residuals and the original model are considered "nested" models.
+The original model is nested within the modified model because the modified model includes all of the terms of the original model along with additional terms.
+
+```{r}
+anova(
+  efa3factorModified_fit,
+  efa3factor_fit
+)
+```
+
+In this case, the model with the correlated residuals fit significantly better (i.e., has a significantly smaller chi-square value) than the model without the correlated residuals.
+
+Here are the variables that had a standardized loading greater than 0.3 on each of the factors:
+
+```{r}
+factor1vars <- c(
+  "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
+  "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame")
+
+factor2vars <- c(
+  "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
+  "avg_air_distance","max_air_distance","avg_air_yards_to_sticks")
+
+factor3vars <- c(
+  "passing_cpoe","pass_comp_pct","passer_rating","completion_percentage_above_expectation")
+```
+
+The variables that loaded most strongly onto factor 1 appear to reflect Quarterback usage: completions per game, passing attempts per game, passing yards per game, passing touchdowns per game, passing air yards (total horizontal distance the ball travels on all pass attempts) per game, passing yards after the catch per game, and first downs gained per game by passing.
+Quarterbacks who tend to throw more tend to have higher levels on those variables.
+Thus, we label component 1 as "Usage", which reflects total Quarterback involvement, regardless of efficiency or outcome.
+
+The variables that loaded most strongly onto factor 2 appear to reflect Quarterback aggressiveness: average air yards on completed passes, average air yards on all attempted passes, aggressiveness (percentage of passing attempts thrown into tight windows, where there is a defender within one yard or less of the receiver at the time of the completion or incompletion), average amount of air yards ahead of or behind the first down marker on passing attempts, average air distance (the true three-dimensional distance the ball travels in the air), maximum air distance, and maximum air distance on completed passes.
+Quarterbacks who throw the ball farther and into tighter windows tend to have higher values on those variables.
+Thus, we label component 2 as "Aggressiveness", which reflects throwing longer, more difficult passes with a tight window.
+
+The variables that loaded most strongly onto factor 3 appear to reflect Quarterback performance: passing completion percentage above expectation, pass completion percentage, and passer rating.
+Quarterbacks who perform better tend to have higher values on those variables.
+Thus, we label component 3 as "Performance".
+
+Here are the players and seasons that showed the highest levels of Quarterback "Usage":
+
+```{r}
+dataForFA_efa %>% 
+  arrange(-F1) %>% 
+  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the lowest levels of Quarterback "Usage":
+
+```{r}
+dataForFA_efa %>% 
+  arrange(F1) %>% 
+  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the highest levels of Quarterback "Aggressiveness":
+
+```{r}
+dataForFA_efa %>% 
+  arrange(-F2) %>% 
+  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the lowest levels of Quarterback "Aggressiveness":
+
+```{r}
+dataForFA_efa %>% 
+  arrange(F2) %>% 
+  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the highest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_efa %>% 
+  arrange(-F3) %>% 
+  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the highest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_efa %>% 
+  arrange(-F3) %>% 
+  filter(games >= 10) %>% 
+  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the lowest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_efa %>% 
+  arrange(F3) %>% 
+  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the lowest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_efa %>% 
+  arrange(F3) %>% 
+  filter(games >= 10) %>% 
+  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+## Example of Confirmatory Factor Analysis {#sec-cfaExample}
+
+```{r}
+#| eval: false
+#| include: false
+
+psych::principal(
+  dataForFA_standardized[faVars],
+  nfactors = 3,
+  rotate = "oblimin")
+
+'
+ #Factor loadings
+ F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + passing_interceptionsPerGame + passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
+ F2 =~ pacr + avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + avg_air_yards_to_sticks + expected_completion_percentage + avg_air_distance + max_air_distance
+ F3 =~ passing_epa + passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
+'
+
+factor1_syntax <- '
+ # Factor loadings
+ F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
+
+ # Correlated residuals
+ passing_air_yardsPerGame ~~ passing_yards_after_catchPerGame
+ completionsPerGame ~~ attemptsPerGame
+ attemptsPerGame ~~ passing_yards_after_catchPerGame
+ passing_yards_after_catchPerGame ~~ passing_first_downsPerGame
+ completionsPerGame ~~ passing_first_downsPerGame
+ attemptsPerGame ~~ passing_tdsPerGame
+ completionsPerGame ~~ passing_tdsPerGame
+ passing_tdsPerGame ~~ passing_air_yardsPerGame
+'
+
+factor1_fit <- lavaan::cfa(
+  factor1_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+
+summary(
+  factor1_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+
+lavaan::modificationindices(
+  factor1_fit,
+  sort. = TRUE)
+
+factor2_syntax <- '
+ # Factor loadings
+ F2 =~ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + avg_air_distance + max_air_distance + avg_air_yards_to_sticks #expected_completion_percentage + pacr
+
+ # Correlated residuals
+ max_completed_air_distance ~~ max_air_distance
+ avg_completed_air_yards ~~ max_completed_air_distance
+ avg_completed_air_yards ~~ avg_air_distance
+ avg_completed_air_yards ~~ max_air_distance
+ avg_intended_air_yards ~~ max_completed_air_distance
+'
+
+factor2_fit <- lavaan::cfa(
+  factor2_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+
+summary(
+  factor2_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+
+lavaan::modificationindices(
+  factor2_fit,
+  sort. = TRUE)
+
+factor3_syntax <- '
+ #Factor loadings
+ F3 =~ passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
+'
+
+factor3_fit <- lavaan::cfa(
+  factor3_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+
+summary(
+  factor3_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+
+lavaan::modificationindices(
+  factor3_fit,
+  sort. = TRUE)
+```
+
+We fit CFA models using the `cfa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
+We compare one-, two-, and three-factor CFA models.
+
+Below is the syntax for the one-factor CFA model:
+
+```{r}
+cfa1factor_syntax <- '
+ #Factor loadings
+ F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
+ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
+ passer_rating + completion_percentage_above_expectation
+'
+```
+
+```{r}
+#| label: cfa-1factor-fit
+
+cfa1factor_fit <- lavaan::cfa(
+  cfa1factor_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+```
+
+```{r}
+summary(
+  cfa1factor_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+The one-factor model did not fit well according to the fit indices:
+
+```{r}
+lavaan::fitMeasures(
+  cfa1factor_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+lavaan::residuals(
+  cfa1factor_fit,
+  type = "cor")
+```
+
+```{r}
+lavaan::modificationindices(
+  cfa1factor_fit,
+  sort. = TRUE)
+```
+
+Below are factor scores from the model for the first six players:
+
+```{r}
+cfa1factor_factorScores <- lavaan::lavPredict(cfa1factor_fit)
+
+head(cfa1factor_factorScores)
+```
+
+The path diagram of the one-factor CFA model is in @fig-cfaPathDiagram-1factor.
+
+```{r}
+#| label: fig-cfaPathDiagram-1factor
+#| fig-cap: "Path Diagram of the One-Factor Confirmatory Factor Analysis Model."
+#| fig-alt: "Path Diagram of the One-Factor Confirmatory Factor Analysis Model."
+
+lavaanPlot::lavaanPlot(
+  cfa1factor_fit,
+  coefs = TRUE,
+  #covs = TRUE,
+  stand = TRUE)
+```
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(cfa1factor_fit)
+```
+
+Below is the syntax for the two-factor CFA model:
+
+```{r}
+cfa2factor_syntax <- '
+ #Factor loadings
+ F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
+
+ F2 =~ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
+ passer_rating + completion_percentage_above_expectation
+'
+```
+
+```{r}
+#| label: cfa-2factor-fit
+
+cfa2factor_fit <- lavaan::cfa(
+  cfa2factor_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+```
+
+```{r}
+summary(
+  cfa2factor_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+The two-factor model did not fit well according to the fit indices:
+
+```{r}
+lavaan::fitMeasures(
+  cfa2factor_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+lavaan::residuals(
+  cfa2factor_fit,
+  type = "cor")
+```
+
+```{r}
+lavaan::modificationindices(
+  cfa2factor_fit,
+  sort. = TRUE)
+```
+
+Below are factor scores from the model for the first six players:
+
+```{r}
+cfa2factor_factorScores <- lavaan::lavPredict(cfa2factor_fit)
+
+head(cfa2factor_factorScores)
+```
+
+The path diagram of the two-factor CFA model is in @fig-cfaPathDiagram-2factor.
+
+```{r}
+#| label: fig-cfaPathDiagram-2factor
+#| fig-cap: "Path Diagram of the Two-Factor Confirmatory Factor Analysis Model."
+#| fig-alt: "Path Diagram of the Two-Factor Confirmatory Factor Analysis Model."
+
+lavaanPlot::lavaanPlot(
+  cfa2factor_fit,
+  coefs = TRUE,
+  covs = TRUE,
+  stand = TRUE)
+```
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(cfa2factor_fit)
+```
+
+Because the one-factor model is nested within the two-factor model, we can compare them using a chi-square difference test.
+The two factor model fit considerably better than the one-factor model in terms of a lower chi-square value:
+
+```{r}
+anova(
+  cfa2factor_fit,
+  cfa1factor_fit
+)
+```
+
+Below is the syntax for the three-factor model:
+
+```{r}
+cfa3factor_syntax <- '
+ #Factor loadings
+ F1 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
+
+ F2 =~ avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks
+
+ F3 =~ passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
+'
+```
+
+```{r}
+#| label: cfa-3factor-fit
+
+cfa3factor_fit <- lavaan::cfa(
+  cfa3factor_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+```
+
+```{r}
+summary(
+  cfa3factor_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+The three-factor model did not fit well according to fit indices:
+```{r}
+lavaan::fitMeasures(
+  cfa3factor_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+However, we know that the three-factor model fit improved considerably when accounting for correlated residuals.
+So, we plan to examine modification indices to see if we can account for covariances between variables that were not explained by their underlying latent factors.
+
+```{r}
+lavaan::residuals(
+  cfa3factor_fit,
+  type = "cor")
+```
+
+Below are the modification indices, suggesting potential model modifications that would improve model fit such as correlated residuals and cross-loadings.
+
+```{r}
+lavaan::modificationindices(
+  cfa3factor_fit,
+  sort. = TRUE)
+```
+
+Below are factor scores from the model for the first six players:
+
+```{r}
+cfa3factor_factorScores <- lavaan::lavPredict(cfa3factor_fit)
+
+head(cfa3factor_factorScores)
+```
+
+The path diagram of the three-factor CFA model is in @fig-cfaPathDiagram-3factor.
+
+```{r}
+#| label: fig-cfaPathDiagram-3factor
+#| fig-cap: "Path Diagram of the Three-Factor Confirmatory Factor Analysis Model."
+#| fig-alt: "Path Diagram of the Three-Factor Confirmatory Factor Analysis Model."
+
+lavaanPlot::lavaanPlot(
+  cfa3factor_fit,
+  coefs = TRUE,
+  covs = TRUE,
+  stand = TRUE)
+```
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(cfa3factor_fit)
+```
+
+Because the two-factor model is nested within the three-factor model, we can compare them using a chi-square difference test.
+The three factor model fit considerably better than the two-factor model in terms of a lower chi-square value:
+
+```{r}
+anova(
+  cfa3factor_fit,
+  cfa2factor_fit
+)
+```
+
+Based on the model modifications suggested by the modification indices for the three-factor model, we modify the model to account for correlated residuals, using the syntax below:
+
+```{r}
+cfa3factorModified_syntax <- '
+ # Factor loadings
+ F1 =~ NA*completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
+ passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame
+
+ F2 =~ NA*avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
+ avg_air_distance + max_air_distance + avg_air_yards_to_sticks
+
+ F3 =~ NA*passing_cpoe + pass_comp_pct + passer_rating + completion_percentage_above_expectation
+
+ # Cross loadings
+ #F3 =~ attemptsPerGame
+
+ # Correlated residuals
+ passing_air_yardsPerGame ~~ passing_yards_after_catchPerGame
+ completionsPerGame ~~ attemptsPerGame
+ attemptsPerGame ~~ passing_yards_after_catchPerGame
+ passing_yards_after_catchPerGame ~~ passing_first_downsPerGame
+ completionsPerGame ~~ passing_first_downsPerGame
+ attemptsPerGame ~~ passing_tdsPerGame
+ completionsPerGame ~~ passing_tdsPerGame
+ passing_tdsPerGame ~~ passing_air_yardsPerGame
+
+ max_completed_air_distance ~~ max_air_distance
+ avg_completed_air_yards ~~ max_completed_air_distance
+ avg_completed_air_yards ~~ avg_air_distance
+ avg_completed_air_yards ~~ max_air_distance
+ avg_intended_air_yards ~~ max_completed_air_distance
+
+ completionsPerGame ~~ pass_comp_pct
+ passing_yardsPerGame ~~ avg_completed_air_yards
+ passing_yardsPerGame ~~ max_completed_air_distance
+ passing_tdsPerGame ~~ passer_rating
+ aggressiveness ~~ completion_percentage_above_expectation
+
+ # Variances
+ F1 ~~ 1*F1
+ F2 ~~ 1*F2
+'
+```
+
+```{r}
+#| label: cfa-3factorModified-fit
+
+cfa3factorModified_fit <- lavaan::cfa(
+  cfa3factorModified_syntax,
+  data = dataForFA,
+  missing = "ML",
+  estimator = "MLR",
+  bounds = "standard",
+  #std.lv = TRUE,
+  cluster = "player_id", # account for nested data within the same player (i.e., longitudinal data)
+  em.h1.iter.max = 2000000)
+```
+
+```{r}
+summary(
+  cfa3factorModified_fit,
+  fit.measures = TRUE,
+  standardized = TRUE,
+  rsquare = TRUE)
+```
+
+The three-factor model with correlated residuals fit was acceptable according to CFI and TLI.
+The model fit was subpar for RMSEA and SRMR.
+
+```{r}
+lavaan::fitMeasures(
+  cfa3factorModified_fit,
+  fit.measures = c(
+    "chisq", "df", "pvalue",
+    "chisq.scaled", "df.scaled", "pvalue.scaled",
+    "chisq.scaling.factor",
+    "baseline.chisq","baseline.df","baseline.pvalue",
+    "rmsea", "cfi", "tli", "srmr",
+    "rmsea.robust", "cfi.robust", "tli.robust"))
+```
+
+```{r}
+lavaan::residuals(
+  cfa3factorModified_fit,
+  type = "cor")
+```
+
+```{r}
+lavaan::modificationindices(
+  cfa3factorModified_fit,
+  sort. = TRUE)
+```
+
+Below are factor scores from the model for the first six players:
+
+```{r}
+cfa3factorModified_factorScores <- lavaan::lavPredict(cfa3factorModified_fit)
+
+head(cfa3factorModified_factorScores)
+
+dataForFA_cfa <- cbind(dataForFA, cfa3factorModified_factorScores)
+```
+
+The path diagram of the modified three-factor CFA model with correlated residuals is in @fig-cfaPathDiagram-3factorModified.
+
+```{r}
+#| label: fig-cfaPathDiagram-3factorModified
+#| fig-cap: "Path Diagram of the Modified Three-Factor Confirmatory Factor Analysis Model With Correlated Residuals."
+#| fig-alt: "Path Diagram of the Modified Three-Factor Confirmatory Factor Analysis Model With Correlated Residuals."
+
+lavaanPlot::lavaanPlot(
+  cfa3factorModified_fit,
+  coefs = TRUE,
+  #covs = TRUE,
+  stand = TRUE)
+```
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(cfa3factorModified_fit)
+```
+
+```{r}
+anova(
+  cfa3factorModified_fit,
+  cfa3factor_fit
+)
+```
+
+Here are the variables that loaded onto each of the factors:
+
+```{r}
+factor1vars <- c(
+  "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
+  "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame")
+
+factor2vars <- c(
+  "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
+  "avg_air_distance","max_air_distance","avg_air_yards_to_sticks")
+
+factor3vars <- c(
+  "passing_cpoe","pass_comp_pct","passer_rating","completion_percentage_above_expectation")
+```
+
+The variables that loaded most strongly onto factor 1 appear to reflect Quarterback usage: completions per game, passing attempts per game, passing yards per game, passing touchdowns per game, passing air yards (total horizontal distance the ball travels on all pass attempts) per game, passing yards after the catch per game, and first downs gained per game by passing.
+Quarterbacks who tend to throw more tend to have higher levels on those variables.
+Thus, we label component 1 as "Usage", which reflects total Quarterback involvement, regardless of efficiency or outcome.
+
+The variables that loaded most strongly onto factor 2 appear to reflect Quarterback aggressiveness: average air yards on completed passes, average air yards on all attempted passes, aggressiveness (percentage of passing attempts thrown into tight windows, where there is a defender within one yard or less of the receiver at the time of the completion or incompletion), average amount of air yards ahead of or behind the first down marker on passing attempts, average air distance (the true three-dimensional distance the ball travels in the air), maximum air distance, and maximum air distance on completed passes.
+Quarterbacks who throw the ball farther and into tighter windows tend to have higher values on those variables.
+Thus, we label component 2 as "Aggressiveness", which reflects throwing longer, more difficult passes with a tight window.
+
+The variables that loaded most strongly onto factor 3 appear to reflect Quarterback performance: passing completion percentage above expectation, pass completion percentage, and passer rating.
+Quarterbacks who perform better tend to have higher values on those variables.
+Thus, we label component 3 as "Performance".
+
+Here are the players and seasons that showed the highest levels of Quarterback "Usage":
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(-F1) %>% 
+  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the lowest levels of Quarterback "Usage":
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(F1) %>% 
+  select(player_display_name, season, F1, all_of(factor1vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the highest levels of Quarterback "Aggressiveness":
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(-F2) %>% 
+  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the lowest levels of Quarterback "Aggressiveness":
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(F2) %>% 
+  select(player_display_name, season, F2, all_of(factor2vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the highest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(-F3) %>% 
+  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the highest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(-F3) %>% 
+  filter(games >= 10) %>% 
+  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+Here are the players and seasons that showed the lowest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(F3) %>% 
+  select(player_display_name, season, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
+
+If we restrict it to Quarterbacks who played at least 10 games in the season, here are the players and seasons that showed the lowest levels of Quarterback "Performance":
+
+```{r}
+dataForFA_cfa %>% 
+  arrange(F3) %>% 
+  filter(games >= 10) %>% 
+  select(player_display_name, season, games, F3, all_of(factor3vars)) %>% 
+  na.omit() %>% 
+  head()
+```
 

### 2025-07-12 index.qmd
20250712 - add interpretation

--- 2025-07-09 index.qmd
@@
   "caret", "LongituRF", "gpboost", "cmdstanr", "corrplot", "mgcv", "rms", 
-  "car", "lavaan", "lavaangui", "mice", "miceadds", "interactions", 
+  "car", "lavaan", "lavaanPlot", "mice", "miceadds", "interactions", 
   "robustbase", "ordinal", "MASS", "data.table", "future", "future.apply", 
-  "SimDesign",
+  "SimDesign", "domir",
   "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
@@
   "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
-  "learnr", "remotes", "shiny", "PearsonDS" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
+  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
 )

### 2025-07-12 multiple-regression.qmd
20250712 - add interpretation

@@
 
-The *standard error* of a regression coefficient represents the uncertainty of the parameter.
-If we have less uncertainty (i.e., more confidence) about the parameter, the standard error will be small.
-If we have more uncertainty (i.e., less confidence) about the parameter, the standard error will be large.
-If we used the same sampling procedure repeatedly and calculated the regression coefficient each time, the true parameter in the population would fall 68% of the time within the interval of:  $[\text{model parameter estimate for the regression coefficient} \pm 1 \text{ standard error}]$ .
+The *standard error* of a regression coefficient represents the imprecision or uncertainty of the parameter.
+If we have less uncertainty (i.e., more confidence) about the parameter, the standard error will be small, reflecting greater precision of the regression coefficient.
+If we have more uncertainty (i.e., less confidence) about the parameter, the standard error will be large, reflecting less precision of the regression coefficient.
+If we used the same sampling procedure repeatedly and calculated the regression coefficient each time, the true parameter in the population would fall 68% of the time within the interval of:  $[\text{model parameter estimate for the regression coefficient} \pm 1 \text{ standard error}]$.
+The standard error is related to the sample size—the larger the sample size, the smaller the standard error (the greater the precision of our estimate of the regression coefficient).
+Otherwise said, having more data gives more precise estimates and thus increases [statistical power](#sec-statisticalPower).
 

### 2025-07-12 pca.qmd
20250712 - add interpretation

@@
 
-## Ensuring the Variables are on the Same Scale {#sec-pcaStandardize}
+## Decisions in Principal Component Analysis {#sec-pcaDecisions}
 
-Before performing a PCA, it is important to ensure that the variables included in the PCA are on the same scale.
+There are four primary decisions to make in PCA:
+
+1. what variables to include in the model and how to scale them
+1. whether and how to rotate components
+1. how many components to retain
+1. how to interpret and use the components
+
+As in [factor analysis](#sec-factorAnalysis), the answer you get can differ highly depending on the decisions you make.
+We provide guidance on each of these decisions below and in @sec-factorAnalysisDecisions.
+
+### 1. Variables to Include and their Scaling {#sec-variablesToInclude-pca}
+
+As in [factor analysis](#sec-factorAnalysis).
+The first decision when conducting a factor analysis is which variables to include and the scaling of those variables.
+Guidance on which variables to include is in @sec-variablesToInclude-factorAnalysis.
+
+In addition, before performing a PCA, it is important to ensure that the variables included in the PCA are on the same scale.
 PCA seeks to identify components that explain variance in the data, so if the variables are not on the same scale, some variables may contribute considerably more variance than others.
@@
 
-## Component Rotation {#sec-pcaRotation}
+### 2. Component Rotation {#sec-pcaRotation}
 
@@
 
-## Determining the Number of Components to Retain {#sec-pcaNumComponents}
+### 3. Determining the Number of Components to Retain {#sec-pcaNumComponents}
 
@@
 
-## Interpreting and Using PCA Components {#sec-interpretingAndUsingComponentsPCA}
+### 4. Interpreting and Using PCA Components {#sec-interpretingAndUsingComponentsPCA}
 
@@
 
+The number of components to keep would depend on which criteria one uses.
+Based on the rule to keep factors whose eigenvalues are greater than one and based on the parallel test, we would keep nine components.
+However, based on the Cattell scree test (the "elbow" of the screen plot minus one), we would keep three components.
+If using the optimal coordinates, we would keep four components; if using the acceleration factor, we would keep one component.
+Therefore, interpretability of the components would be important for deciding how many components to keep.
+
 ```{r}
@@
 We generated the very simple structure (VSS) plots in Figures [-@fig-pcaVSSPlot1] and [-@fig-pcaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
+The optimal number of components based on VSS criterion is three components.
 
@@
 
+Based on the component solutions, the three-component solution maps onto our three-factor solution from [factor analysis](#sec-factorAnalysis).
+The three-component solution explains more than half of the variance in the variables.
+The fourth component in a four-component solution is not particularly interpretable, explains relatively little variance, and seems to be related to sacks: sacks suffered, sack yards lost, sack fumbles, and sack fumbles lost.
+However, two of the sack-related variables (sacks suffered and sack yards lost) load more strongly onto the first component, suggesting that these sack-related variables are better captured by another component.
+Moreover, sacks might be considered a scoring methodological component rather than a particular concept of interest.
+For these resources we prefer the three-component solution and choose to retain three components.
+
 ```{r}
@@
 Thus, we label component 3 as "Performance".
+
+Below are component scores from the PCA for the first six players:
 

### 2025-07-12 machine-learning.qmd
20250712 - documentation

@@
 
+We set up the *k* folds using the `group_vfold_cv()` function of the `rsample` package [@R-sample].
+
 ```{r}
@@
 
+We set up the LOO folds using the `loo_cv()` function of the `rsample` package [@R-sample].
+
 ```{r}
@@
 We also evaluate its accuracy on the hold-out (test) data.
+For each of the models, we fit and evaluate the models using the `tidymodels` ecosystem of packages [@R-tidymodels; @Kuhn2020_packages].
+we specify our model formula using the `recipe()` function of the `recipes` package [@R-recipes].
+We define the model using the `linear_reg()`, `set_engine()`, and `set_mode()` functions of the `parsnip` package [@R-parsnip].
+We specify the workflow using the `workflow()`, `add_recipe()`, and `add_model()` functions of the `workflows` package [@R-workflows].
+We fit the cross-validation model using the `fit_resamples()` function of the `tune` package [@R-tune].
+We specify the accuracy metrics to evaluate using the `metric_set()` function of the `yardstick` package [@R-yardstick].
+We fit the final model using the `fit()` function of the `workflows` package [@R-workflows].
+We evaluate the accuracy of the model's predictions on the test data using the `accuracyOverall()` of the [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab].
 
@@
 
+For the machine learning models, we perform the parameter tuning using the `tune()`, `tune_grid()`, `select_best()`, and `finalize_workflow()` functions of the `tune` package [@R-tune].
+We specify the grid of possible parameter values using the `grid_regular()` function of the `dials` package [@R-dials].
+
 ```{r}
@@
 The random forest models were fit using the `ranger` package [@R-ranger; @Wright2017_packages].
+We specify the grid of possible parameter values using the `grid_random()`, `update.parameters()`, `mtry()`, and `min_n()` functions of the `dials` package [@R-dials] and the `extract_parameter_set_dials()` function of the `hardhat` package [@R-hardhat].
 

### 2025-07-12 multiple-regression.qmd
20250712 - heteroscedasticity types

@@
 library("effectsize")
+library("tidymodels")
 library("tidyverse")
@@
 ```{r}
-#| layout: [[100], [50, 50]]
+#| layout: [[100], [50, 50], [50, 50]]
 #| label: fig-homscedasticityHeteroscedasticity
@@
 #|   - "Bow-Tie-Shaped Heteroscedasticity"
+#|   - "Diamond-Shaped Heteroscedasticity"
+#|   - "Triangle-Shaped Heteroscedasticity"
 #| code-fold: true
@@
 set.seed(52242)
+sampleSize <- 1000
 
 # 1. Homoscedasticity
-sampleSize <- 1000
 x1 <- runif(sampleSize, 0, 10)
@@
 
+# 4. Diamond-shaped heteroscedasticity
+x4 <- runif(sampleSize, 0, 10)
+sd4 <- -abs(x4 - 5) + 5.5  # inverted bow-tie
+y4 <- 3 + 2 * x4 + rnorm(sampleSize, mean = 0, sd = sd4)
+fit4 <- lm(y4 ~ x4)
+res4 <- resid(fit4)
+fitted4 <- fitted(fit4)
+
+#5. Triangle-shaped heteroscedasticity
+height <- 10
+
+sample_triangle <- function(n, v1, v2, v3) {
+  u <- runif(n)
+  v <- runif(n)
+  is_flip <- u + v > 1
+  u[is_flip] <- 1 - u[is_flip]
+  v[is_flip] <- 1 - v[is_flip]
+
+  x <- (1 - u - v) * v1[1] + u * v2[1] + v * v3[1]
+  y <- (1 - u - v) * v1[2] + u * v2[2] + v * v3[2]
+
+  cbind(x, y)
+}
+
+# Triangle vertices
+v1 <- c(0, height)  # top left
+v2 <- c(10, height) # top right
+v3 <- c(5, 0)       # bottom center
+
+# Sample points inside triangle
+tri_points <- sample_triangle(
+  sampleSize,
+  v1,
+  v2,
+  v3)
+
+# Extract x and residuals
+x5 <- tri_points[, 1]
+residuals5 <- tri_points[, 2]
+
+# Generate outcome
+y5 <- 3 + 2 * x5 + residuals5
+
+# Fit model
+fit5 <- lm(y5 ~ x5)
+res5 <- resid(fit5)
+fitted5 <- fitted(fit5)
+
+# Plots
 plot(
@@
   main = "Bow-Tie-Shaped Heteroscedasticity",
+  xlab = "Fitted value",
+  ylab = "Residual")
+
+abline(
+  h = 0,
+  lwd = 2,
+  col = "blue")
+
+plot(
+  fitted4,
+  res4,
+  main = "Diamond-Shaped Heteroscedasticity",
+  xlab = "Fitted value",
+  ylab = "Residual")
+
+abline(
+  h = 0,
+  lwd = 2,
+  col = "blue")
+
+plot(
+  fitted5,
+  res5,
+  main = "Triangle-Shaped Heteroscedasticity",
   xlab = "Fitted value",

### 2025-07-12 multiple-regression.qmd
20250712 - bootstrap regression

@@
 
+Another approach to handling outliers is to use boostrapping, which involves fitting models to various bootstrap resamples of the data.
+Bootstrap samples are datasets generated by sampling repeatedly with replacement.
+
+We set up the bootstrap folds using the `bootstraps()` function of the `rsample` package [@R-sample].
+
+```{r}
+set.seed(52242)
+bootstrapSamples <- 2000
+
+# Create bootstrap resamples (with apparent = TRUE)
+boots <- rsample::bootstraps(
+  data = player_stats_seasonal %>% filter(position %in% c("WR")),
+  times = bootstrapSamples,
+  apparent = TRUE
+)
+
+# Define recipe
+rec <- recipes::recipe(
+  fantasyPoints_transformed ~ age + height + weight + target_share,
+  data = player_stats_seasonal %>% filter(position %in% c("WR"))
+) %>%
+  recipes::step_log(target_share, offset = 1) # replace I(log(target_share + 1))
+
+# Define model spec and workflow
+lm_spec <- parsnip::linear_reg() %>%
+  parsnip::set_engine("lm") %>%
+  parsnip::set_mode("regression")
+
+lm_workflow <- workflows::workflow() %>%
+  workflows::add_recipe(rec) %>%
+  workflows::add_model(lm_spec)
+
+# Function for fitting the models on the bootstrap samples
+fit_lm <- function(split) {
+  analysis_data <- rsample::analysis(split)
+  fit <- workflows::fit(lm_workflow, data = analysis_data)
+  broom::tidy(fit)
+}
+
+# Fit bootstrap models and save results
+boot_models <- boots %>%
+  mutate(coef_info = purrr::map(splits, fit_lm))
+
+# Extract bootstrapped coefficients
+boot_coefs <- 
+  boot_models %>% 
+  tidyr::unnest(coef_info)
+
+# Compute percentile intervals for the regression coefficients
+percentile_intervals <- rsample::int_pctl(boot_models, coef_info)
+#percentile_intervals <- rsample::int_pctl(boot_results, results)
+
+percentile_intervals
+```
+
+```{r}
+#| label: fig-miDensityPlot
+#| fig-cap: "Histogram of Parameter Estimates Across Bootstraps."
+#| fig-alt: "Histogram of Parameter Estimates Across Bootstraps."
+
+ggplot(
+  boot_coefs,
+  aes(estimate)) +
+  geom_histogram(
+    fill = "gray80",
+    color = "black") +
+  facet_wrap(
+    ~ term,
+    scales = "free") +
+  geom_vline(
+    aes(
+      xintercept = .lower),
+    data = percentile_intervals,
+    col = "blue",
+    linetype = "dashed") +
+  geom_vline(
+    aes(
+      xintercept = .upper),
+    data = percentile_intervals,
+    col = "blue",
+    linetype = "dashed") +
+  labs(
+    title = "Bootstrap Distributions of Coefficients",
+    x = "Coefficient Estimate",
+    y = "Count") +
+  theme_classic()
+```
+
 ## Moderated Multiple Regression {#sec-moderatedMultipleRegression}

### 2025-07-12 multiple-regression.qmd
20250712 - fix

@@
 percentile_intervals <- rsample::int_pctl(boot_models, coef_info)
-#percentile_intervals <- rsample::int_pctl(boot_results, results)
 
@@
 
-```{r}
-#| label: fig-miDensityPlot
+The distributions of the regression coefficients across boostraps are depicted in @fig-bootstrapRegressionCoefficients.
+
+```{r}
+#| label: fig-bootstrapRegressionCoefficients
 #| fig-cap: "Histogram of Parameter Estimates Across Bootstraps."

### 2025-07-12 multiple-regression.qmd
20250712 - BCa confidence interval

@@
 # Function for fitting the models on the bootstrap samples
-fit_lm <- function(split) {
+fit_lm <- function(split, ...) {
   analysis_data <- rsample::analysis(split)
@@
 
-# Extract bootstrapped coefficients
-boot_coefs <- 
-  boot_models %>% 
+# Extract bootstrapped coefficient estimates
+boot_coefs <- boot_models %>%
   tidyr::unnest(coef_info)
 
-# Compute percentile intervals for the regression coefficients
-percentile_intervals <- rsample::int_pctl(boot_models, coef_info)
-
+# Percentile confidence intervals
+percentile_intervals <- rsample::int_pctl(
+  .data = boot_models,
+  statistics = coef_info)
+
+# Bias-corrected and accelerated confidence intervals
+bca_intervals <- rsample::int_bca(
+  .data = boot_models,
+  statistics = coef_info,
+  .fn = fit_lm
+)
+
+# View confidence intervals
 percentile_intervals
+bca_intervals
 ```

### 2025-07-13 factor-analysis.qmd
20250712 - tweaks
20250713 - tweaks

@@
 
-> "All models are wrong, but some are useful."
+> "*All models are wrong, but some are useful.*"
 >
@@
 
-Criteria for acceptable fit and good fit of SEM models are in @semFitIndices.
+Criteria for acceptable fit and good fit of SEM models are in @tbl-semFitIndices.
 In addition, dynamic fit indexes have been proposed based on simulation to identify fit index cutoffs that are tailored to the characteristics of the specific model and data [@McNeish2023]; dynamic fit indexes are available via the `dynamic` package [@R-dynamic] or with a [webapp](https://dynamicfit.app).

### 2025-07-13 machine-learning.qmd
20250712 - tweaks
20250713 - tweaks

@@
 
-We set up the *k* folds using the `group_vfold_cv()` function of the `rsample` package [@R-sample].
+We set up the *k* folds using the `group_vfold_cv()` function of the `rsample` package [@R-rsample].
 
@@
 
-We set up the LOO folds using the `loo_cv()` function of the `rsample` package [@R-sample].
+We set up the LOO folds using the `loo_cv()` function of the `rsample` package [@R-rsample].
 

### 2025-07-13 mixed-models.qmd
20250713 - typo

@@
 These are known as spaghetti plots.
-By examining the trajectory for each player, we can get a better understanding of hor performance changes (within an individual) as a function of age.
+By examining the trajectory for each player, we can get a better understanding of how performance changes (within an individual) as a function of age.
 

### 2025-07-14 player-evaluation.qmd
20250713 - Bohr quote
20250714 - volume

@@
 # Player Evaluation {#sec-playerEvaluation}
+
+> "*It is very difficult to predict—especially the future.*"
+>
+> --- Neils Bohr
 
@@
 
-It is also important to consider a player's opportunity and usage, which are influenced by many factors, including the skill of the player, the skill of their teammates, the role of the player on the team, the coaching style, the strategy of the opposing team, game scripts, etc.
+It is also important to consider a player's opportunity, usage, and volume, which are influenced by many factors, including the skill of the player, the skill of their teammates, the role of the player on the team, the coaching style, the strategy of the opposing team, game scripts, etc.
 In terms of the player's opportunity and usage, how many snaps do they get?
@@
 Being on the field for more snaps, running more routes, and receiving more touches and/or targets means that the player has more opportunities to score fantasy points.
+What is the depth of their targets?
+Are they receiving short passes or long passes?
+Players who receive longer passes may have an opportunity for more fantasy points but also may be more volatile (i.e., less [consistent](#sec-evalHistoricalConsistency)).
 Are they targeted in the red zone?

### 2025-07-15 machine-learning.qmd
20250715 - don't load all data or run all code chunks for efficiency

@@
 load(file = "./data/nfl_depthCharts.RData")
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
+#load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
+#load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
+#load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
 #load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
-load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
+#load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
 
@@
 load(file = "./data/nfl_depthCharts.RData")
-load(file = "./data/nfl_pbp.RData")
-load(file = "./data/nfl_4thdown.RData")
-load(file = "./data/nfl_participation.RData")
+#load(file = "./data/nfl_pbp.RData")
+#load(file = "./data/nfl_4thdown.RData")
+#load(file = "./data/nfl_participation.RData")
 #load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
@@
 load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
-load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
+#load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
 
@@
   names(nfl_depthCharts),
-  names(nfl_pbp),
-  names(nfl_4thdown),
-  names(nfl_participation),
+  #names(nfl_pbp),
+  #names(nfl_4thdown),
+  #names(nfl_participation),
   #names(nfl_actualFantasyPoints_player_weekly),
@@
   names(nfl_rankings_weekly),
-  names(nfl_expectedFantasyPoints_weekly),
-  names(nfl_expectedFantasyPoints_pbp)
+  names(nfl_expectedFantasyPoints_weekly)#,
+  #names(nfl_expectedFantasyPoints_pbp)
 )
@@
 #| label: impute-rb-machine-learning
+#| eval: false
 
@@
 #| label: impute-wr-machine-learning
+#| eval: false
 
@@
 #| label: impute-te-machine-learning
+#| eval: false
 

### 2025-07-15 multiple-regression.qmd
20250712 - tweaks
20250715 - load beta regression model from OSF

@@
 Note: the following code that runs the model takes a while.
+If you just want to save time and load the model object instead of running the model, you can load the model object (which has already been fit) using this code:
+
+```{r}
+load(url("https://osf.io/download/fe37j/"))
+```
 :::
@@
 #| output: false
+#| eval: false
 
@@
 )
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+save(
+  betaRegression,
+  file = "./data/betaRegression.RData")
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/betaRegression.RData", fsep = ""))
+```
+
+```{r}
+#| include: false
+
+load(url("https://osf.io/download/fe37j/"))
 ```
@@
 
-We set up the bootstrap folds using the `bootstraps()` function of the `rsample` package [@R-sample].
+We set up the bootstrap folds using the `bootstraps()` function of the `rsample` package [@R-rsample].
 

### 2025-07-15 factor-analysis.qmd
20250715 - factor analysis and SEM

@@
 
+## Factor Analysis and Structural Equation Modeling
+
+Factor analysis forms the measurement model component of a structural equation model (SEM).
+The *measurement model* is what we settle on as the estimation of each construct before we add the structural component to estimate the relations among latent variables.
+Basically, in a structural equation model, we add the structural component onto the measurement model.
+For instance, our measurement model (i.e., based on factor analyis) might be to estimate, from a set of items, three latent factors: usage, aggressiveness, and performance.
+Our structural model then may examine what processes (e.g., sport drink consumption or sleep) influence these latent factors, how the latent factors influence each other, and what the latent factors influence (e.g., fantasy points).
+SEM is [confirmatory factor analysis](#sec-cfa) with regression paths that specify hypothesized causal relations between the latent variables (the structural component of the model).\index{latent variable}\index{structural equation modeling}\index{factor analysis!confirmatory}
+Exploratory structural equation modeling (ESEM) is a form of SEM that allows for a combination of [exploratory factor analysis](#sec-efa) and [confirmatory factor analysis](#sec-cfa) to estimate latent variables and the relations between them.\index{latent variable}\index{structural equation modeling}\index{factor analysis!exploratory}\index{factor analysis!confirmatory}\index{structural equation modeling!exploratory}
+
 ## Path Diagrams {#sec-pathDiagrams}
 
-A key tool when designing a structural equation model is a conceptual depiction of the hypothesized causal processes.
-A path diagram depicts the hypothesized causal processes that link two or more variables.
+A key tool when designing a factor analysis or structural equation model is a conceptual depiction of the hypothesized causal processes.
+A *path diagram* depicts the hypothesized causal processes that link two or more variables.
+Path diagrams are an example of a causal diagram and are similar to [directed acyclic graphs](#sec-causalDiagrams) discussed in @sec-causalDiagrams.
 @Karch2025_packages provides a tool to create `lavaan` `R` syntax from a path analytic diagram: <https://lavaangui.org>.
@@
 If the factors are not correlated with each other, use an orthogonal rotation.\index{factor analysis}\index{rotation}\index{orthogonal rotation}
-The correlation between an item and a factor is a factor loading, which is simply a way to ask how much a variable is correlated with the underlying factor.\index{factor analysis}\index{structural equation modeling!factor loading}
+The correlation between an item and a factor is a factor loading, which is simply a way to ask how much a variable is correlated with the underlying factor.\index{factor analysis!factor loading}
 However, its interpretation is more complicated if there are correlated factors!\index{factor analysis}\index{rotation}\index{oblique rotation}
@@
 
-: Criteria for Acceptable and Good Fit of Structural Equation Models Based on Fit Indices. {#tbl-semFitIndices}
+: Criteria for Acceptable and Good Fit of Factor Analysis and Structural Equation Models Based on Fit Indices. {#tbl-semFitIndices}
 

### 2025-07-15 download-football-data.qmd
20250708 - track chunk execution times
20250715 - add acknowledgments and tweaks

@@
+{{< include _chunk-timing.qmd >}}
+
 # Download and Process NFL Football Data {#sec-downloadFootballData}
@@
 nfl_standings_raw <- nflseedR::nfl_standings(
-  nfl_schedules%>% filter(!is.na(result))) #read_csv("http://www.habitatring.com/standings.csv"); archived at: https://perma.cc/5QLN-VK7V
+  nfl_schedules %>% filter(!is.na(result))) #read_csv("http://www.habitatring.com/standings.csv"); archived at: https://perma.cc/5QLN-VK7V
 ```
@@
 
-TODO: Save/update data file in repo with the data generated from this code, and follow the data file through the rest of the book, updating code as necessary.
-
 ::: {#nte-downloadSeasonalActualStats .callout-note title="Downloading season-by-season statistics"}
@@
 
-We already load players' week-by-week statistics [above](#sec-downloadActualStats).
-Nevertheless, we could compute players' weekly statistics from the play-by-play data using the following syntax:
-
-TODO:
-
-- Decide whether to use `calculate_stats()` from `nflfastR` [@R-nflfastR] or `load_player_stats()` from `nflreadr` [@R-nflreadr].
-- Save/update data file in repo with the data generated from the relevant code, and follow the data file through the rest of the book, updating code as necessary.
+We already load players' week-by-week statistics [above](#sec-downloadActualStats) using the `load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
+Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:
 

### 2025-07-15 index.qmd
20250715 - add acknowledgments and tweaks

@@
 
+This work would not be possible without the important contributions of many people.
+This book was made possible by the hard work of the package developers of the `nflverse` ecosystem of R packages, who made football data available, including Sebastian Carl, Ben Baldwin, Lee Sharpe, Tan Ho, and John Edwards.
 I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.

### 2025-07-15 bugs.qmd
20250715 - bug list

--- 2024-11-27 bugs.qmd
@@
 - common naming convention for key variables, for purposes of merging (<https://github.com/nflverse/nflreadr/issues/237>)
-- use of deprecated `summarise()` for returning more than 1 row per group; use `reframe()` instead (<https://github.com/nflverse/nflfastR/issues/480>)
-- calculate career stats using `calculate_stats()` function
+- use of deprecated `summarise()` for returning more than 1 row per group; use `reframe()` instead (issue 1: <https://github.com/nflverse/nflfastR/issues/480>; issue 2: <https://github.com/nflverse/nflfastR/issues/520>)
+- calculate career stats using `calculate_stats()` function (as was formerly able to do with the `calculate_player_stats()` function) (<https://github.com/nflverse/nflfastR/issues/501>)
+- will `nflreadr` switch to using the objects from the `calculate_stats()` function in `nflfastR`? (<https://github.com/nflverse/nflreadr/issues/261>)

### 2025-07-16 multiple-regression.qmd
20250716 - beta regression

@@
 library("brms")
+library("parallelly")
 library("robustbase")
@@
   family = zero_one_inflated_beta(),
+  cores = 4,
+  threads = threading(parallelly::availableCores()),
+  seed = 52242,
+  silent = 0
 )

### 2025-07-17 correlation.qmd
20250708 - track chunk execution times
20250712 - add interpretation
20250717 - Conclusion sections

@@
+{{< include _chunk-timing.qmd >}}
+
 # Correlation Analysis {#sec-correlation}
+
+This chapter provides an overview of correlation analysis.
 
@@
 The covariance is an unstandardized index of the association between two variables.
+It represents the average product of two variables' deviations from their respective means.
 The formula for (sample) covariance is in @eq-covariance and is the numerator to the formula for [correlation](#sec-correlationExamplesPearson).
@@
 ```
+
+The covariance of a variable with itself is the variable's variance:
+
+$$
+\begin{aligned}
+  \text{Cov}(x, x) &= \frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})(x_i - \bar{x})\\
+  &= \frac{1}{n - 1} \sum_{i=1}^{n} (x_i - \bar{x})^2 \\
+  &= \frac{\sum (x_i - \bar{x})^2}{n-1}\\
+  &= s^2
+\end{aligned}
+$$ {#eq-covarianceToVariance}
 
@@
 
-Correlation is an index of the association between variables.
+Correlation is a standardized index of the association between variables.
 The correlation coefficient ($r$) ranges from −1 to +1, and indicates the sign and magnitude of the association.
 Although correlation does not imply causation, identifying associations between variables can still be useful because association is a necessary (but insufficient) condition for causality.
+The correlation coefficient can be heavily impacted by [outliers](#sec-correlationOutliers) and [restricted range](#sec-correlationRestrictedRange).
 

### 2025-07-17 data-visualization.qmd
20250708 - track chunk execution times
20250717 - Conclusion sections

@@
+{{< include _chunk-timing.qmd >}}
+
 # Data Visualization {#sec-dataVisualization}
+
+This chapter provides an overview of principles of data visualization and how to implement them using `R`.
 
@@
 
+This chapter provided an overview of principles of data visualization and how to implement them using `R`.
+Key principles of data visualization include: ensuring adequate contrast, highlighting key elements of interest (and using gray for other elements), reducing clutter, and adding helpful titles and annotations.
+We provided examples of data visualizations, including [histograms](#sec-histogram), [density plots](#sec-densityPlot), [rug plots](#sec-densityHistogram), [box-and-whisker plots](#sec-boxAndWhiskerPlot), [violin plots](#sec-violinPlot), [ridgeline plots](#sec-ridgelinePlot), [scatterplots](#sec-scatterplot), [line charts](#sec-lineChart), and [bar plots](#sec-barPlot).
+
 ::: {.content-visible when-format="html"}

### 2025-07-17 download-football-data.qmd
20250717 - Conclusion sections

@@
 # Download and Process NFL Football Data {#sec-downloadFootballData}
+
+This chapter provides an overview of how to download and process a wide range of football and fantasy football-related data.
+We leverage the data in subsequent chapters for data analysis.
 
@@
 
+# Conclusion {#sec-downloadFootballDataConclusion}
+
+This chapter provided an overview of how to download and process a wide range of football and fantasy football-related data.
+Now that we have downloaded and processed a wide range of data, we can use the data in subsequent chapters for data analysis.
+
 ::: {.content-visible when-format="html"}

### 2025-07-17 draft.qmd
20250708 - track chunk execution times
20250717 - Conclusion sections

@@
+{{< include _chunk-timing.qmd >}}
+
 # The Fantasy Draft {#sec-draft}
+
+This chapter provides an overview of the major types of drafts in fantasy football and various draft strategies.
 
@@
 
+## Conclusion {#sec-draftConclusion}
+
+The two major draft types are [snake drafts](#sec-draftSnake) and [auction drafts](#sec-draftAuction).
+[Auction drafts](#sec-draftAuction) take longer than [snake drafts](#sec-draftSnake) but are more fair because every player is available to every manager, so long as the manager is able and willing to bid enough.
+There is no one right draft strategy.
+In general, it is helpful to draft players with the best value over other players at their position.
+
 ::: {.content-visible when-format="html"}

### 2025-07-17 fantasy-football.qmd
20250711 - replacement of video content for PDF/ePub
20250717 - Conclusion sections

@@
 :::
+
+:::
+
+::: {.content-visible unless-format="html:js"}
+
+@Ly2015 provides a video of a brief introduction to American football.
 
@@
 
+:::
+
+::: {.content-visible unless-format="html:js"}
+
+@NFLFantasy2020 provides a video of a brief introduction to fantasy football.
+
+:::
+
 ### The Fantasy League {#sec-league}
@@
 - "PAT" = point after touchdown (i.e., extra point/point after attempt)
+
+# Conclusion {#sec-fantasyFootballConclusion}
+
+This chapter provided a brief primer on (American) football and fantasy football.
+Football is the most widely watched sport in the United States.
+The goal in football is for a team to score more points than their opponent.
+In football, points can be scored by scoring a touchdown (6 points) followed by successfully kicking a point-after-touchdown (1 additional point) or successfully completing a two-point conversion (2 additional points), kicking a field goal (3 points), or getting a safety (2 points).
+
+Fantasy football is an online game where participants assemble (i.e., "draft") imaginary teams composed of real-life NFL players.
+The participants or "managers" make decisions such as selecting which players to draft, selecting which players start on a weekly basis, identifying players to pick up from the waiver wire, and making trades with other teams.
+Fantasy football relies heavily on prediction—trying to predict which players will perform best and selecting them accordingly.
+In the game of fantasy football, managers accumulate points on a weekly basis based on players' actual statistical performances in NFL games.
+A manager's goal is to outscore their opponent each week to win matches and ultimately claim victory in the league.

### 2025-07-17 getting-started.qmd
20250717 - Conclusion sections

@@
 R is a free software environment; you can download it at no charge here: <https://cran.r-project.org>.
+This chapter provides an overview of how to install and learn the software `R`, how to troubleshoot code, and how to perform various data management operations.
 
@@
 
+# Conclusion {#sec-gettingStartedConclusion}
+
+This chapter provided an overview of how to install and learn the software `R`, how to troubleshoot code, and how to perform various data management operations.
+Most of the code throughout the book that performs data management operations leverages the techniques implemented in this chapter.
+
 ::: {.content-visible when-format="html"}

### 2025-07-17 index.qmd
20250717 - Conclusion sections

@@
 I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, and Adam Peterson, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
-In particular, Andrew Tungate made important additions to the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] that were necessary for conducting a [bootstrap simulation](#sec-simulationProjectedPointsBootstrapping) of players' projected fantasy points.
+In particular, Andrew Tungate and Dennis Andersen helped develop the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], which is useful for scraping fantasy football projections and calculating projections averaged across sources.
+Andrew Tungate also made important additions to the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] to allow computing fantasy points given projected (or actual) player statistics, which was necessary for conducting a [bootstrap simulation](#sec-simulationProjectedPointsBootstrapping) of players' projected fantasy points.
 I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.

### 2025-07-17 intro.qmd
20250711 - replacement of video content for PDF/ePub
20250717 - Conclusion sections

@@
 
+::: {.content-visible unless-format="html:js"}
+
+@NFL2014 provides a video of Professor Benjamin Motz that describes the value of teaching statistics through the lens of fantasy football.
+
+:::
+
 ## Goals {#learningGoals}
@@
 > --- [Benjamin Motz, Ph.D.](#fig-video-motz) [@NFL2014]
+
+## Conclusion {#sec-introConclusion}
+
+In conclusion, fantasy football provides an exciting lens through which to learn statistics, and `R` is a useful (and free!) software for conducting statistical analysis and creating statistical graphics.

### 2025-07-17 multiple-regression.qmd
20250717 - Conclusion sections

@@
 # Multiple Regression {#sec-multipleRegression}
+
+This chapter provides an overview of multiple regression.
 
@@
 Thus, seasonal target share (unlike weekly target share)—as it is currently calculated—does not range from 0–1.
-For example, if a player has a weekly target share of 15%, that would be a seasonal (cross-week-summed) target share of 2.55 ($0.15 \times 17 \; \text{games} = 2.55$.
+For example, if a player has a weekly target share of 15%, that would be a seasonal (cross-week-summed) target share of 2.55 ($0.15 \times 17 \; \text{games} = 2.55$).
 
@@
 
-The only terms that are significantly associated with fantasy performance among Wide Receivers are weight and target share, both of which showed a positive association with fantasy points.
+The only terms that were significantly associated with fantasy performance among Wide Receivers are weight and target share, both of which showed a positive association with fantasy points.
 
@@
 Target share has a large [effect size](#sec-practicalSignificance).
-All of the other [predictors](#sec-correlationalStudy) have small [effect size](#sec-practicalSignificance).
+All of the other [predictors](#sec-correlationalStudy) have a small [effect size](#sec-practicalSignificance).
 

### 2025-07-17 player-evaluation.qmd
20250717 - Conclusion sections

@@
 > --- Neils Bohr
+
+This chapter provides an overview of some of the many factors in player evaluation.
+Ultimately, player evaluation involves prediction—predicting how well the player will do in future games.
 
@@
 For example, it would be valuable to use assessments from games or game-like situations in which players must perform tasks that leverage multiple, intertwined skills and that are similar to the tasks you want to predict.
+The extent to which the assessments are more reflective of games or game-like situations, we consider the assessment to have better [ecological validity](#sec-ecologicalValidity).
 
@@
 
+## Conclusion {#sec-playerEvaluationConclusion}
+
+In conclusion, there are many factors to consider when evaluating players.
+You will have less direct access to some factors (e.g., cognitive and motivational factors) than others (e.g., historical performance).
+When considering how much weight to give to each of various factors, it is important to evaluate how much [predictive validity](#sec-predictiveValidity) each factor has for predicting a player's successful performance, and to weight it accordingly.
+In general, samples tend to be stronger than signs for predicting player performance in sports.
+For instance, compared to tests of speed, power, and agility at the NFL Combine, collegiate performance is a stronger predictor of performance in the NFL.
+That is, previous sports performance is the best predictor of future performance.
+When thinking about a player's value, it can be worth thinking their potential upside and downside.
+Whether you value a player with higher upside or higher downside may depend on many factors such as, when drafting, whether you are seeking a starter or bench player, and, when in-season, whether you are projected to win by a large margin in your matchpup.
+
 ::: {.content-visible when-format="html"}

### 2025-07-18 player-evaluation.qmd
20250717 - sites with Vegas lines

@@
 Moreover, sites also provide season-long player futures for the over/under betting lines, which are based on Vegas lines and how people bet: <https://www.rotowire.com/betting/nfl/player-futures.php>.
-In addition, sites also provide weekly projections based on Vegas lines: <https://vegasprojections.com>; <https://www.actionnetwork.com/nfl/props>.
+In addition, sites also provide weekly projections based on Vegas lines: <https://vegasprojections.com>; <https://www.actionnetwork.com/nfl/props>; <https://tools.32beatwriters.com>; <https://the-odds-api.com>.
 

### 2025-07-18 actuarial.qmd
20250711 - replacement of video content for PDF/ePub
20250717 - chapter overviews

@@
 # Judgment Versus Actuarial Approaches to Prediction {#sec-judgmentVsActuarial}
+
+This chapter compares [judgment](#sec-humanJudgment) versus [actuarial approaches](#sec-actuarialPrediction) to prediction in fantasy football (and other domains).
 
@@
 :::
+
+:::
+
+::: {.content-visible unless-format="html:js"}
+
+@TPS2017 provides a video of sports pundits, Stephen A. Smith and Skip Bayless, making bold statements and incorrect predictions.
 

### 2025-07-18 cognitive-bias.qmd
20250717 - chapter overviews

@@
 # Heuristics and Cognitive Biases in Prediction {#sec-cognitiveBias}
+
+This chapter provides an overview of many [heuristics](#sec-heuristics), [cognitive biases](#sec-cognitiveBiases), and [fallacies](#sec-fallacies) that people engage in when making judgments and predictions in fantasy football.
 
@@
 
-In conclusion, there are many [heuristics](#sec-heuristics), [cognitive bias](#sec-cognitiveBiases), and [fallacies](#sec-fallacies) that people engage in when making judgments and predictions.
+In conclusion, there are many [heuristics](#sec-heuristics), [cognitive biases](#sec-cognitiveBiases), and [fallacies](#sec-fallacies) that people engage in when making judgments and predictions.
 It is prudent to be aware of these common biases and to work to counteract them.

### 2025-07-18 evaluating-prediction-accuracy.qmd
20250711 - replacement of video content for PDF/ePub
20250717 - chapter overviews

@@
 
+::: {.content-visible when-format="html:js"}
+
 > "*Nothing ruins fantasy more than reality.*" – [Renee Miller, Ph.D.](#fig-fantasy-vs-reality) [@Yahoo2024]
 
-::: {.content-visible when-format="html:js"}
-
 ::: {#fig-fantasy-vs-reality}
@@
 :::
+
+::: {.content-visible unless-format="html:js"}
+
+> "*Nothing ruins fantasy more than reality.*" – [Renee Miller, Ph.D.](https://www.youtube.com/watch?v=gmpLFWs5ae0&t=1430) [@Yahoo2024]
+
+:::
+
+This chapter provides an overview of ways to evaluate the accuracy of predictions.
+In addition, we evaluate the accuracy of fantasy football projections.
 
@@
 
+In terms of the accuracy of historical fantasy football projections, projections showed moderate accuracy in terms of [discrimination](#sec-discrimination) in predicting players' fantasy performance.
+However, projections were not highly accurate.
+Projections showed some [miscalibration](#sec-calibrationIndices) with respect to actual fantasy points.
+Projection sources appeared to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their projections of player performance.
+In general, projected showed a modest tendency to underpredict performance at lower levels of projected points and a moderate-to-strong tendency overpredict performance at higher levels of projected points, an example of the overextremity, which is an example of the overprecision form of [overconfidence](#sec-cognitiveBiasesOverconfidence).
+Crowd-averaged projections tend to be more accurate than individual sources of projections.
+
 ::: {.content-visible when-format="html"}

### 2025-07-18 mythbusters.qmd
20250708 - track chunk execution times
20250717 - chapter overviews

@@
+{{< include _chunk-timing.qmd >}}
+
 # Mythbusters: Putting Fantasy Football Beliefs/Anecdotes to the Test {#sec-mythbusters}
+
+In this chapter, we put a popular fantasy football belief to the test.
+We evaluate whether the widely held belief is true that players perform better during a contract year.
 
@@
 library("lmerTest")
-library("MuMIn")
+library("performance")
 library("emmeans")
@@
 summary(mixedModel_qbr)
-MuMIn::r.squaredGLMM(mixedModel_qbr)
+performance::r2(mixedModel_qbr)
 emmeans::emmeans(mixedModel_qbr, "contractYear")
@@
 summary(mixedModelAge_qbr)
-MuMIn::r.squaredGLMM(mixedModelAge_qbr)
+performance::r2(mixedModelAge_qbr)
 emmeans::emmeans(mixedModelAge_qbr, "contractYear")
@@
 summary(mixedModel_ptsAdded)
-MuMIn::r.squaredGLMM(mixedModel_ptsAdded)
+performance::r2(mixedModel_ptsAdded)
 emmeans::emmeans(mixedModel_ptsAdded, "contractYear")
@@
 summary(mixedModelAge_ptsAdded)
-MuMIn::r.squaredGLMM(mixedModelAge_ptsAdded)
+performance::r2(mixedModelAge_ptsAdded)
 emmeans::emmeans(mixedModelAge_ptsAdded, "contractYear")
@@
 summary(mixedModel_epaPass)
-MuMIn::r.squaredGLMM(mixedModel_epaPass)
+performance::r2(mixedModel_epaPass)
 emmeans::emmeans(mixedModel_epaPass, "contractYear")
@@
 summary(mixedModelAge_epaPass)
-MuMIn::r.squaredGLMM(mixedModelAge_epaPass)
+performance::r2(mixedModelAge_epaPass)
 emmeans::emmeans(mixedModelAge_epaPass, "contractYear")
@@
 summary(mixedModel_fantasyPtsPass)
-MuMIn::r.squaredGLMM(mixedModel_fantasyPtsPass)
+performance::r2(mixedModel_fantasyPtsPass)
 emmeans::emmeans(mixedModel_fantasyPtsPass, "contractYear")
@@
 summary(mixedModelAge_fantasyPtsPass)
-MuMIn::r.squaredGLMM(mixedModelAge_fantasyPtsPass)
+performance::r2(mixedModelAge_fantasyPtsPass)
 emmeans::emmeans(mixedModelAge_fantasyPtsPass, "contractYear")
@@
 summary(mixedModel_ypc)
-MuMIn::r.squaredGLMM(mixedModel_ypc)
+performance::r2(mixedModel_ypc)
 emmeans::emmeans(mixedModel_ypc, "contractYear")
@@
 summary(mixedModelAge_ypc)
-MuMIn::r.squaredGLMM(mixedModelAge_ypc)
+performance::r2(mixedModelAge_ypc)
 emmeans::emmeans(mixedModelAge_ypc, "contractYear")
@@
 summary(mixedModel_epaRush)
-MuMIn::r.squaredGLMM(mixedModel_epaRush)
+performance::r2(mixedModel_epaRush)
 emmeans::emmeans(mixedModel_epaRush, "contractYear")
@@
 summary(mixedModelAge_epaRush)
-MuMIn::r.squaredGLMM(mixedModelAge_epaRush)
+performance::r2(mixedModelAge_epaRush)
 emmeans::emmeans(mixedModelAge_epaRush, "contractYear")
@@
 summary(mixedModel_fantasyPtsRush)
-MuMIn::r.squaredGLMM(mixedModel_fantasyPtsRush)
+performance::r2(mixedModel_fantasyPtsRush)
 emmeans::emmeans(mixedModel_fantasyPtsRush, "contractYear")
@@
 summary(mixedModelAge_fantasyPtsRush)
-MuMIn::r.squaredGLMM(mixedModelAge_fantasyPtsRush)
+performance::r2(mixedModelAge_fantasyPtsRush)
 emmeans::emmeans(mixedModelAge_fantasyPtsRush, "contractYear")
@@
 summary(mixedModel_receivingYards)
-MuMIn::r.squaredGLMM(mixedModel_receivingYards)
+performance::r2(mixedModel_receivingYards)
 emmeans::emmeans(mixedModel_receivingYards, "contractYear")
@@
 summary(mixedModelAge_receivingYards)
-MuMIn::r.squaredGLMM(mixedModelAge_receivingYards)
+performance::r2(mixedModelAge_receivingYards)
 emmeans::emmeans(mixedModelAge_receivingYards, "contractYear")
@@
 summary(mixedModel_epaReceiving)
-MuMIn::r.squaredGLMM(mixedModel_epaReceiving)
+performance::r2(mixedModel_epaReceiving)
 emmeans::emmeans(mixedModel_epaReceiving, "contractYear")
@@
 summary(mixedModelAge_epaReceiving)
-MuMIn::r.squaredGLMM(mixedModelAge_epaReceiving)
+performance::r2(mixedModelAge_epaReceiving)
 emmeans::emmeans(mixedModelAge_epaReceiving, "contractYear")
@@
 summary(mixedModel_fantasyPtsReceiving)
-MuMIn::r.squaredGLMM(mixedModel_fantasyPtsReceiving)
+performance::r2(mixedModel_fantasyPtsReceiving)
 emmeans::emmeans(mixedModel_fantasyPtsReceiving, "contractYear")
@@
 summary(mixedModelAge_fantasyPtsReceiving)
-MuMIn::r.squaredGLMM(mixedModelAge_fantasyPtsReceiving)
+performance::r2(mixedModelAge_fantasyPtsReceiving)
 emmeans::emmeans(mixedModelAge_fantasyPtsReceiving, "contractYear")
@@
 ### QB/RB/WR/TE {#sec-contractYearQBRBWRTE}
+
+```{r}
+player_statsContractsQBRBWRTE_seasonal <- player_statsContracts_seasonal %>% 
+  dplyr::filter(
+    position_group %in% c("QB","RB","WR","TE"),
+    games >= 5, # keep only player-season combinations in which QBs played at least 5 games
+    season >= 2011) # keep only seasons since 2011 (when most contract data are available)
+```
+
+```{r}
+mixedModel_fantasyPts <- lmerTest::lmer(
+  fantasyPoints ~ contractYear + position_group + (1 | player_id),
+  data = player_statsContractsQBRBWRTE_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModel_fantasyPts)
+performance::r2(mixedModel_fantasyPts)
+emmeans::emmeans(mixedModel_fantasyPts, "contractYear")
+
+mixedModelAge_fantasyPts <- lmerTest::lmer(
+  fantasyPoints ~ contractYear + position_group + ageCentered20 + ageCentered20Quadratic + years_of_experience + (1 + ageCentered20 | player_id),
+  data = player_statsContractsQBRBWRTE_seasonal,
+  control = lmerControl(optimizer = "bobyqa")
+)
+
+summary(mixedModelAge_fantasyPts)
+performance::r2(mixedModelAge_fantasyPts)
+emmeans::emmeans(mixedModelAge_fantasyPts, "contractYear")
+```
 

### 2025-07-18 index.qmd
20250718 - acknowledgments

@@
 I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.
-I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, and Adam Peterson, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
-In particular, Andrew Tungate and Dennis Andersen helped develop the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], which is useful for scraping fantasy football projections and calculating projections averaged across sources.
+I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, Adam Peterson, Jesse Kartes, and Chris Russo, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
+In particular, Andrew Tungate and Dennis Andersen helped develop the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], which scrapes fantasy football projections from the web and calculates projections averaged across sources.
 Andrew Tungate also made important additions to the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] to allow computing fantasy points given projected (or actual) player statistics, which was necessary for conducting a [bootstrap simulation](#sec-simulationProjectedPointsBootstrapping) of players' projected fantasy points.
+I am fortunate to work with such a talented, passionate, and creative team.
 I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.

### 2025-07-18 mythbusters.qmd
20250718 - summary text

@@
 In this chapter, we put a popular fantasy football belief to the test.
-We evaluate whether the widely held belief is true that players perform better during a contract year.
+We evaluate the widely held belief that players perform better during a contract year.
 
@@
 Then, we analyze the data.
+
+#### Quarterback Rating {#sec-contractYearQB-QBR}
+
 Below is a mixed model that examines whether a player has a higher QBR per game when they are in a contract year compared to when they are not in a contract year.
+The first model includes just contract year as a predictor.
+The second model includes additional covariates, including player age and experience.
+In terms of Quarterback Rating (QBR), findings from the models indicate that Quarterbacks did not perform significantly better in their contract year.
 
@@
 
+#### Points Added {#sec-contractYearQB-PointsAdded}
+
+In terms of points added, Quarterbacks did not perform better in their contract year.
+
 ```{r}
@@
 
+#### Expected Points Added {#sec-contractYearQB-PointsAdded}
+
+In terms of expected points added (EPA) from passing plays, when not controlling for player age and experience, Quarterbacks performed better in their contract year.
+However, when controlling for player age and experience, Quarterbacks did not perform significantly better in their contract year.
+
 ```{r}
@@
 ```
+
+#### Fantasy Points {#sec-contractYearQB-FantasyPoints}
+
+In terms of fantasy points, Quarterbacks performed significantly worse in their contract year, even controlling for player age and experience.
 
@@
 
+#### Yards Per Carry {#sec-contractYearRB-YPC}
+
+In terms of yards per carry (YPC), Running Backs did not perform significantly better in their contract year.
+
 ```{r}
@@
 
+#### Expected Points Added {#sec-contractYearRB-EPA}
+
+In terms of expected points added (EPA) from rushing plays, Running Backs did not perform significantly better in their contract year.
+
 ```{r}
@@
 ```
+
+#### Fantasy Points {#sec-contractYearRB-FantasyPoints}
+
+In terms of fantasy points, Running Backs performed significantly worse in their contract year, even controlling for player age and experience.
 
@@
 
+#### Receiving Yards {#sec-contractYearWRTE-ReceivingYards}
+
+In terms of receiving yards, Wide Receivers/Tight Ends performed significantly worse in their contract year, even controlling for player age and experience.
+
 ```{r}
@@
 
+#### Expected Points Added {#sec-contractYearWRTE-EPA}
+
+In terms of expected points added (EPA) from receiving plays, Wide Receivers/Tight Ends performed significantly worse in their contract year, even controlling for player age and experience.
+
 ```{r}
@@
 ```
+
+#### Fantasy Points {#sec-contractYearWRTE-FantasyPoints}
+
+In terms of fantasy points, Wide Receivers/Tight Ends performed significantly worse in their contract year, even controlling for player age and experience.
 
@@
 
+#### Fantasy Points {#sec-contractYearQBRBWRTE-FantasyPoints}
+
+In terms of fantasy points, Quarterbacks/Running Backs/Wide Receivers/Tight Ends performed significantly worse in their contract year, even controlling for player age and experience.
+
 ```{r}
@@
 
-```{r}
-# Placeholder for model predicting fantasy points
-# Include player position as a covariate
-```
+## Conclusion {#sec-mythbustersConclusion}
 
-## Conclusion {#sec-mythbustersConclusion}
+There is a widely held belief that NFL players perform better in the last year of the contract because they are motivated to gain another contract.
+There is some evidence in the NBA and MLB that players tend to perform better in their contract year.
+We evaluated this possibility among NFL players who were Quarterbacks, Running Backs, Wide Receivers, or Tight Ends.
+We evaluated a wide range of performance indexes, including Quarterback Rating, yards per carry, points added, expected points added, receiving yards, and fantasy points.
+None of the positions showed significantly better performance in their contract year for any of the performance indexes.
+By contrast, if anything, players tended to perform more poorly during their contract year, as operationalized by fantasy points, receiving yards (WR/TE), and EPA from receiving plays (WR/TE), even when controlling for player and age experience.
+In sum, we did not find evidence in support of the contract year hypothesis and consider this myth debunked.
+However, we are open to this possibility being reexamined in new ways or with additional performance
 

### 2025-07-18 base-rates.qmd
20250717 - chapter overviews
20250718 - cross-references

@@
 # Base Rates {#sec-baseRates}
+
+This chapter provides an overview of base rates and the important roles they play in prediction.
 
@@
 Using [Bayes' theorem](#sec-bayesTheorem) (described in @sec-bayesTheorem), solving for posttest odds (based on pretest odds and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), as in @eq-bayes5), and converting odds to probabilities, we can use a Fagan probability nomogram to determine the posttest probability following a test result.
-The calculation of posttest probability is described in INSERT.
+The calculation of posttest (posterior) probability is described in @sec-posttestProbability.
 A *probability nomogram* (aka Fagan nomogram) is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the [pretest (or prior) probability](#sec-baseRate) and [likelihood ratio](#sec-diagnosticLikelihoodRatio), as depicted in @fig-probabilityNomogram.
@@
 
-The blue line indicates the [posterior probability](#posttestProbability) of the condition given a positive test.
-The pink line indicates the [posterior probability](#posttestProbability) of the condition given a negative test.
+The blue line indicates the [posterior probability](#sec-posttestProbability) of the condition given a positive test.
+The pink line indicates the [posterior probability](#sec-posttestProbability) of the condition given a negative test.
 

### 2025-07-18 getting-started.qmd
20250718 - create a function

@@
 
-## Installing Packages {#sec-installingPackages}
+## Install Packages {#sec-installingPackages}
 
@@
 install.packages("tidyverse")
+```
+
+You can also install multiple packages with one line of code (replacing each package with its respective name):
+
+```{r}
+#| eval: false
+
+install.packages(c("PACKAGE1","PACKAGE2","PACKAGE3"))
 ```
@@
 The name of the function is followed by parentheses; the inputs go in between the parentheses.
+The possible inputs that a function can accept are called "arguments".
 You can learn about a particular function and its arguments by entering a question mark before the name of the function:
@@
 Hopefully, that provides an example for how to learn about a particular function, its arguments, and how to use them.
+
+## Create a Function {#sec-createFunction}
+
+Functions are useful if you want to perform an operation multiple times.
+Use of a function can save you time without needed to retype out all of the code each time.
+For instance, let's say you want to convert temperature between Fahrenheit and Celsius, you could create a function to do that.
+
+```{r}
+convert_temperature <- function(temperature, unit = "F"){
+  if(unit == "F"){ # if the input temperature(s) in Fahrenheit
+    newtemp <- (temperature - 32) / (9/5)
+  } else if(unit == "C"){ # if the input temperature(s) in Celsius
+    newtemp <- (temperature * (9/5)) + 32
+  }
+
+  return(newtemp)
+}
+```
+
+Now we can use the function to convert temperatures between Fahrenheit and Celsius.
+A temperature of 32°F is equal to 0°C.
+A temperature of 0°C is equal to 89.6°F.
+
+```{r}
+convert_temperature(
+  temperature = 32,
+  unit = "F"
+)
+
+convert_temperature(
+  temperature = 32,
+  unit = "C"
+)
+```
+
+We can also convert the temperature for a vector of values at once:
+
+```{r}
+convert_temperature(
+  temperature = c(0, 10, 20, 30, 40, 50),
+  unit = "F"
+)
+
+convert_temperature(
+  temperature = c(0, 10, 20, 30, 40, 50),
+  unit = "C"
+)
+```
+
+Because the default unit is "F", we do not need to specify the unit if our input temperatures are in Fahrenheit:
+
+```{r}
+convert_temperature(
+  c(0, 10, 20, 30, 40, 50)
+)
+```
 

### 2025-07-19 evaluating-prediction-accuracy.qmd
20250718 - confusion matrix

@@
 Some predictions involve categorical data, whereas other predictions involve continuous data.
-When dealing with a dichotomous ([nominal data](#sec-nominal) that are binary) [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variable (or continuous data that have been dichotomized using a cutoff), we can evaluate predictions using a 2x2 table known as a confusion matrix (see INSERT), or with logistic regression models.
+When dealing with a dichotomous ([nominal data](#sec-nominal) that are binary) [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variable (or continuous data that have been dichotomized using a cutoff), we can evaluate predictions using a 2x2 table known as a [confusion matrix](#sec-confusionMatrix) (see @fig-confusionMatrix), or with logistic regression models.
 When dealing with a continuous [outcome variable](#sec-correlationalStudy) (e.g., [ordinal](#sec-ordinal), [interval](#sec-interval), or [ratio](#sec-ratio) data), we can evaluate predictions using [multiple regression](#sec-multipleRegression) or similar variants such as structural equation modeling and mixed models.
@@
 The military conducts a test of its prospective members to screen out applicants who would likely fail basic training.
-To evaluate the accuracy of our predictions using the test, we can examine a confusion matrix.
-A confusion matrix is a matrix that presents the predicted outcome on one dimension and the actual outcome (truth) on the other dimension.
-If the predictions and outcomes are dichotomous, the confusion matrix is a 2x2 matrix with two rows and two columns that represent four possible predicted-actual combinations (decision outcomes): true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN).
+To evaluate the accuracy of our predictions using the test, we can examine a [confusion matrix](#sec-confusionMatrix).
+A [confusion matrix](#sec-confusionMatrix) is a matrix that presents the predicted outcome on one dimension and the actual outcome (truth) on the other dimension.
+If the predictions and outcomes are dichotomous, the [confusion matrix](#sec-confusionMatrix) is a 2x2 matrix with two rows and two columns that represent four possible predicted-actual combinations (decision outcomes), as in @fig-confusionMatrix: true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN).
+
+::: {#fig-confusionMatrix}
+![](images/confusionMatrix.png){fig-alt="A Two-by-Two Confusion Matrix."}
+
+A Two-by-Two Confusion Matrix.
+:::
 
@@
 
-An example of a confusion matrix is in INSERT.
-
-With the information in the confusion matrix, we can calculate the marginal sums and the proportion of people in each cell (in parentheses), as depicted in INSERT.
+An example of a [confusion matrix](#sec-confusionMatrix) is in @fig-confusionMatrix1.
+
+::: {#fig-confusionMatrix1}
+![](images/confusionMatrix1.png){fig-alt="Example of a Two-by-Two Confusion Matrix. TP = true positives; TN = true negatives; FP = false positives; FN = false negatives."}
+
+Example of a Two-by-Two Confusion Matrix. TP = true positives; TN = true negatives; FP = false positives; FN = false negatives.
+:::
+
+With the information in the [confusion matrix](#sec-confusionMatrix), we can calculate the marginal sums and the proportion of people in each cell (in parentheses), as depicted in @fig-confusionMatrix2.
+
+::: {#fig-confusionMatrix2}
+![](images/confusionMatrix2.png){fig-alt="Example of a Two-by-Two Confusion Matrix With Marginal Sums. TP = true positives; TN = true negatives; FP = false positives; FN = false negatives."}
+
+Example of a Two-by-Two Confusion Matrix With Marginal Sums. TP = true positives; TN = true negatives; FP = false positives; FN = false negatives.
+:::
 
@@
 
-Based on the marginal sums, we can compute the [marginal probabilities](#sec-baseRate), as depicted in INSERT.
+Based on the marginal sums, we can compute the [marginal probabilities](#sec-baseRate), as depicted in @fig-confusionMatrix3.
+
+::: {#fig-confusionMatrix3}
+![](images/confusionMatrix3.png){fig-alt="Example of a Two-by-Two Confusion Matrix With Marginal Sums and Marginal Probabilities.
+TP = true positives; TN = true negatives; FP = false positives; FN = false negatives; BR = base rate; SR = selection ratio."}
+
+Example of a Two-by-Two Confusion Matrix With Marginal Sums and Marginal Probabilities.
+TP = true positives; TN = true negatives; FP = false positives; FN = false negatives; BR = base rate; SR = selection ratio.
+:::
 
@@
 
-Based on the confusion matrix, we can calculate the prediction accuracy based on the percent accuracy of the predictions.
+Based on the [confusion matrix](#sec-confusionMatrix), we can calculate the prediction accuracy based on the percent accuracy of the predictions.
 The percent accuracy is the number of correct predictions divided by the total number of predictions, and multiplied by 100.
-In the context of a confusion matrix, this is calculated as: $100\% \times \frac{\text{TP} + \text{TN}}{N}$.
+In the context of a [confusion matrix](#sec-confusionMatrix), this is calculated as: $100\% \times \frac{\text{TP} + \text{TN}}{N}$.
 In this case, our percent accuracy was 78%—that is, 78% of our predictions were accurate, and 22% of our predictions were inaccurate.
@@
 To determine what accuracy we could get by chance alone given the selection ratio and the base rate, we can calculate the chance probability of true positives and the chance probability of true negatives.
-The probability of a given cell in the confusion matrix is a [joint probability](#sec-jointProbability)—the probability of two events occurring simultaneously.
+The probability of a given cell in the [confusion matrix](#sec-confusionMatrix) is a [joint probability](#sec-jointProbability)—the probability of two events occurring simultaneously.
 To calculate a [joint probability](#sec-jointProbability), we multiply the probability of each event.
@@
 
-The chance expectancies for each cell of the confusion matrix are in INSERT
+The chance expectancies for each cell of the [confusion matrix](#sec-confusionMatrix) are in @fig-confusionMatrix4.
+
+::: {#fig-confusionMatrix4}
+![](images/confusionMatrix4.png){fig-alt="Chance Expectancies in Two-by-Two Confusion Matrix. BR = base rate; SR = selection ratio."}
+
+Chance Expectancies in Two-by-Two Confusion Matrix. BR = base rate; SR = selection ratio.
+:::
 
@@
 
-Let us return to the confusion matrix in INSERT.
-If we know the frequency of each of the four predicted-actual combinations of the confusion matrix (TP, TN, FP, FN), we can calculate sensitivity, specificity, PPV, and NPV.
+Let us return to the [confusion matrix](#sec-confusionMatrix) in @fig-confusionMatrix3.
+If we know the frequency of each of the four predicted-actual combinations of the [confusion matrix](#sec-confusionMatrix) (TP, TN, FP, FN), we can calculate sensitivity, specificity, PPV, and NPV.
 

### 2025-07-19 getting-started.qmd
20250718 - updates

@@
 
-## Create a Function {#sec-createFunction}
-
-Functions are useful if you want to perform an operation multiple times.
-Use of a function can save you time without needed to retype out all of the code each time.
-For instance, let's say you want to convert temperature between Fahrenheit and Celsius, you could create a function to do that.
-
-```{r}
-convert_temperature <- function(temperature, unit = "F"){
-  if(unit == "F"){ # if the input temperature(s) in Fahrenheit
-    newtemp <- (temperature - 32) / (9/5)
-  } else if(unit == "C"){ # if the input temperature(s) in Celsius
-    newtemp <- (temperature * (9/5)) + 32
-  }
-
-  return(newtemp)
-}
-```
-
-Now we can use the function to convert temperatures between Fahrenheit and Celsius.
-A temperature of 32°F is equal to 0°C.
-A temperature of 0°C is equal to 89.6°F.
-
-```{r}
-convert_temperature(
-  temperature = 32,
-  unit = "F"
-)
-
-convert_temperature(
-  temperature = 32,
-  unit = "C"
-)
-```
-
-We can also convert the temperature for a vector of values at once:
-
-```{r}
-convert_temperature(
-  temperature = c(0, 10, 20, 30, 40, 50),
-  unit = "F"
-)
-
-convert_temperature(
-  temperature = c(0, 10, 20, 30, 40, 50),
-  unit = "C"
-)
-```
-
-Because the default unit is "F", we do not need to specify the unit if our input temperatures are in Fahrenheit:
-
-```{r}
-convert_temperature(
-  c(0, 10, 20, 30, 40, 50)
-)
-```
-
 ## Create a Vector {#sec-createVector}
@@
 
+## If...Else Conditions {#sec-ifelse}
+
+We can use the construction, `if()...else if()...else()` if we want to perform conditional operations.
+The construction, `if()...else if()...else()` can only be used on one value at a time.
+
+```{r}
+player_rank <- 15
+
+if(player_rank <= 10){ # check this condition first
+  tier <- 1
+  print(tier)
+} else if(player_rank <= 20){ # if first condition was not met, check this condition next
+  tier <- 2
+  print(tier)
+} else if(player_rank <= 30){ # if first two conditions were not met, check this condition next
+  tier <- 3
+  print(tier)
+} else{ # if all other conditions were not met, then do this
+  print("Don't draft this player!")
+}
+```
+
+To apply conditional operations to a vector, we can use the `ifelse()` function.
+
+```{r}
+player_rank <- c(1, 10, 20, 40, 100)
+
+tier <- ifelse(
+  player_rank <= 10, # check this condition
+  1, # assign this value if true
+  2) # assign this value if false
+
+tier
+```
+
 ## Piping {#sec-piping}
@@
 
+## Create a Function {#sec-createFunction}
+
+Now, let's put together what we have learned to create a useful function.
+Functions are useful if you want to perform an operation multiple times.
+Any operation that you want to perform multiple times, you can create a function to accomplish.
+Use of a function can save you time without needed to retype out all of the code each time.
+For instance, let's say you want to convert temperature between Fahrenheit and Celsius, you could create a function to do that.
+In this case, our function has two arguments: `temperature` (in degrees) and `unit` of the original temperature (`F` for Fahrenheit or `C` for Celsius, where the default unit is Fahrenheit).
+
+```{r}
+convert_temperature <- function(temperature, unit = "F"){
+  if(unit == "F"){ # if the input temperature(s) in Fahrenheit
+    newtemp <- (temperature - 32) / (9/5)
+  } else if(unit == "C"){ # if the input temperature(s) in Celsius
+    newtemp <- (temperature * (9/5)) + 32
+  }
+
+  return(newtemp)
+}
+```
+
+Now we can use the function to convert temperatures between Fahrenheit and Celsius.
+A temperature of 32°F is equal to 0°C.
+A temperature of 0°C is equal to 89.6°F.
+
+```{r}
+convert_temperature(
+  temperature = 32,
+  unit = "F"
+)
+
+convert_temperature(
+  temperature = 32,
+  unit = "C"
+)
+```
+
+We can also convert the temperature for a vector of values at once:
+
+```{r}
+convert_temperature(
+  temperature = c(0, 10, 20, 30, 40, 50),
+  unit = "F"
+)
+
+convert_temperature(
+  temperature = c(0, 10, 20, 30, 40, 50),
+  unit = "C"
+)
+```
+
+Because the default unit is "F", we do not need to specify the unit if our input temperatures are in Fahrenheit:
+
+```{r}
+convert_temperature(
+  c(0, 10, 20, 30, 40, 50)
+)
+```
+
 # Conclusion {#sec-gettingStartedConclusion}

### 2025-07-19 modern-portfolio-theory.qmd
20250701 - citations
20250708 - track chunk execution times
20250717 - chapter overviews
20250718 - updates

@@
+{{< include _chunk-timing.qmd >}}
+
 # Modern Portfolio Theory {#sec-modernPortfolioTheory}
+
+This chapter provides an overview of modern portfolio theory.
 
@@
 #### Portfolio with Max Return at a Given Risk Level {#sec-maxReturnPortfolio}
+
+If the `maxreturnPortfolio()` function worked as expected, the following could should work for determining the portfolio with the maximum return at each of various risk levels.
+However, there is a known bug in the `fportfolio` package [@R-fPortfolio] that the optimal portfolio (based on maximum returns) does not change when changing the target risk level, suggesting that it is not taking into account the target risk level specified by the user (see here: <https://stackoverflow.com/q/78784306/2029527>).
 
@@
 
+Below, we use the `mvFrontier()` function from the `NMOF` package [@R-NMOF; @Gilli2019_packages; @Schumann20112024_packages] to compute the efficient frontier.
+
 ```{r}

### 2025-07-19 evaluating-prediction-accuracy.qmd
20250718 - confusion matrix

@@
 
+Players are uniquely identified by `player_id`-`season`-`data_src`:
+
+```{r}
+projectionsWithActuals_seasonal %>% 
+  filter(!is.na(player_id)) %>% 
+  group_by(player_id, season, data_src) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
 #### Weekly Projections {#sec-predictionAccuracyPrepareDataWeekly}
@@
   filter(position_group == "DB")
+```
+
+Players are uniquely identified by `player_id`-`season`-`week`-`data_src`:
+
+```{r}
+projectionsWithActuals_weekly %>% 
+  filter(!is.na(player_id)) %>% 
+  group_by(player_id, season, week, data_src) %>% 
+  filter(n() > 1) %>% 
+  head()
 ```
@@
 If the predictions and outcomes are dichotomous, the confusion matrix is a 2x2 matrix with two rows and two columns that represent four possible predicted-actual combinations ([decision outcomes](#sec-decisionOutcomes)).
-In such a case, the confusion matrix provides a tabular count of each type of accurate cases ([true positives](#sec-truePositive) and [true negatives](#sec-trueNegative)) versus the number of each type of error ([false positives](#sec-falsePositive) and [false negatives](#sec-falseNegative)), as shown in INSERT.
-An example of a confusion matrix is in INSERT.
+In such a case, the confusion matrix provides a tabular count of each type of accurate cases ([true positives](#sec-truePositive) and [true negatives](#sec-trueNegative)) versus the number of each type of error ([false positives](#sec-falsePositive) and [false negatives](#sec-falseNegative)), as shown in @fig-confusionMatrix.
+An example of a confusion matrix is in @fig-confusionMatrix1.
+
+To generate a 2x2 confusion matrix, we need to make a dichotomous decision/prediction.
+For this example, our goal is to predict which Running Backs or Wide Receivers will score 300 or more fantasy points.
+We will predict a player as scoring 300 or more fantasy points for a given projection source if that projection source projected the player to score 300 or more fantasy points.
+Thus, the same player has multiple rows in the data file—one for each projection source.
+
+```{r}
+projectionsWithActuals_seasonal_subset <- projectionsWithActuals_seasonal %>% 
+  filter(position_group %in% c("RB","WR")) %>% 
+  filter(season == 2024)
+
+projectionsWithActuals_seasonal_subset$prediction <- ifelse(projectionsWithActuals_seasonal_subset$raw_points >= 300, 1, 0)
+projectionsWithActuals_seasonal_subset$truth <- ifelse(projectionsWithActuals_seasonal_subset$fantasyPoints >= 300, 1, 0)
+
+projectionsWithActuals_seasonal_subset$predictionFactor <- factor(
+  projectionsWithActuals_seasonal_subset$prediction,
+  levels = c(1,0),
+  labels = c("Prediction: Will Score 300+ Points", "Prediction: Will Not Score 300+ Points"))
+
+projectionsWithActuals_seasonal_subset$truthFactor <- factor(
+  projectionsWithActuals_seasonal_subset$truth,
+  levels = c(1,0),
+  labels = c("Truth: Scored 300+ Points", "Truth: Did Not Score 300+ Points")
+)
+
+table(projectionsWithActuals_seasonal_subset$predictionFactor)
+
+table(projectionsWithActuals_seasonal_subset$truthFactor)
+```
 
@@
 ```{r}
-#table(mydata$diagnosisFactor, mydata$diseaseFactor)
-```
-
-##### Number with margins added {#sec-confusionMatrix-numberMargins}
-
-```{r}
-#addmargins(table(mydata$diagnosisFactor, mydata$diseaseFactor))
+table(
+  projectionsWithActuals_seasonal_subset$predictionFactor,
+  projectionsWithActuals_seasonal_subset$truthFactor)
+```
+
+##### Number With Margins Added {#sec-confusionMatrix-numberMargins}
+
+```{r}
+addmargins(table(
+  projectionsWithActuals_seasonal_subset$predictionFactor,
+  projectionsWithActuals_seasonal_subset$truthFactor))
 ```
@@
 ```{r}
-#prop.table(table(mydata$diagnosisFactor, mydata$diseaseFactor))
-```
-
-##### Proportions with margins added {#sec-confusionMatrix-numberMarginsProportionsMargins}
-
-```{r}
-#addmargins(prop.table(table(mydata$diagnosisFactor, mydata$diseaseFactor)))
+prop.table(table(
+  projectionsWithActuals_seasonal_subset$predictionFactor,
+  projectionsWithActuals_seasonal_subset$truthFactor))
+```
+
+##### Proportions With Margins Added {#sec-confusionMatrix-numberMarginsProportionsMargins}
+
+```{r}
+addmargins(prop.table(table(
+  projectionsWithActuals_seasonal_subset$predictionFactor,
+  projectionsWithActuals_seasonal_subset$truthFactor)))
 ```

### 2025-07-19 modern-portfolio-theory.qmd
20250718 - citations

@@
 That is, "few stock pickers, if any, have the skill needed to beat the market consistently, year after year" [@Kahneman2011, p. 214].
-Although our sample size is much smaller with fantasy football projections, there also appears to be little consistency in fantasy football sites' rank in accuracy over time [INSERT], suggesting that the projection sources are not reliably better than each other (or the crowd) over time.
+Although our sample size is much smaller with fantasy football projections, there also appears to be little consistency in fantasy football sites' rank in accuracy over time [@Kartes2024; archived at <https://perma.cc/69F7-LLTN>; @Petersen2017; archived at <https://perma.cc/BG2W-ANUF>], suggesting that the projection sources are not reliably better than each other (or the crowd) over time.
 
@@
 So, by going with the market average, you will likely perform better than the "average" fund/investor.
-Consistent with this, crowd-averaged fantasy football projections tend to be more accurate than any individual's projection: INSERT
+Consistent with this, crowd-averaged fantasy football projections tend to be more accurate than any individual's projection [@Kartes2024; archived at <https://perma.cc/69F7-LLTN>; @Petersen2017; archived at <https://perma.cc/BG2W-ANUF>].
 This evidence is consistent with the notion of the [wisdom of the crowd](#sec-wisdomOfCrowd), described in @sec-wisdomOfCrowd.

### 2025-07-19 evaluating-prediction-accuracy.qmd
20250719 - citations

@@
 That is, projection sources appear to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their projections of player performance.
-Relatedly, crowd-averaged projections tend to be more accurate than individual sources of projections [CITATION].
+Relatedly, crowd-averaged projections tend to be more accurate than individual sources of projections [@Kartes2024; archived at <https://perma.cc/69F7-LLTN>; @Petersen2017; archived at <https://perma.cc/BG2W-ANUF>].
 

### 2025-07-19 player-evaluation.qmd
20250719 - citations

@@
 Projected statistics (e.g., projected passing touchdowns) are agnostic to league settings and can thus be used to generate league-specific fantasy projections and rankings.
-Thus, projected statisitics may be more useful than rankings because they can be used to generate rankings for your particular league settings.
+Thus, projected statistics may be more useful than rankings because they can be used to generate rankings for your particular league settings.
 For instance, if you know how many touchdowns, yards, and interceptions a Quarterback is a projected to throw (in addition to any other relevant categories for the player, e.g., rushing yards and touchdowns), you can calculate how many fantasy points the Quarterback is expected to gain in your league (or in any league).
 Thus, you can calculate ranking from projections, but you cannot reverse engineer projections from rankings.
+Moreover, projections have been shown to be more accurate than rankings [@Petersen2016d; archived at <https://perma.cc/KEZ9-WRB6>].
 
@@
 Thus, players with greater uncertainty are risky and tend to have a higher upside (or ceiling) and a lower downside (or floor).
+For this reason, identifying players with a high uncertainty can be useful for identifying potential "sleepers"—i.e., late-round draft picks who outperform expectations [@Petersen2014; archived at <https://perma.cc/K6LX-85A4>].
 

### 2025-07-19 evaluating-prediction-accuracy.qmd
20250719 - positional bias

@@
 Relatedly, crowd-averaged projections tend to be more accurate than individual sources of projections [@Kartes2024; archived at <https://perma.cc/69F7-LLTN>; @Petersen2017; archived at <https://perma.cc/BG2W-ANUF>].
+
+The association between projected and actual fantasy points tended to flatten at the higher levels of projected points, for some positions, including Running Backs, Kickers, and defensive players.
+A consequence of this pattern, from the perspective of draft strategy, is that it would not make sense to draft players projected to score more than the flattening of the projected–actual curve.
+For instance, drafting a highly routed Wide Receiver may be safer than drafting a highly ranked Running Back, who may be more prone to injury.
+This finding was echoed by @Kartes2025 (archived at <https://perma.cc/R3WS-CAKD>) whose evaluation of the accuracy of projections by position found that:
+
+a.
+Quarterbacks with preseason ranks of 6–10 showed the greatest systematic underperformance relative to their projections, suggesting that you should either draft an elite Quarterback (ranked 1–5) or wait until later rounds to draft multiple lower-ranked Quarterbacks.
+a. Elite Running Backs (with preseason ranks of 1–5) showed strong underperformance (relative to projections), suggesting that you should draft Running Backs ranked 6 and below.
+a. Wide Receivers showed among the more consistent and predictable patterns in projections, suggesting that it may be safer to draft a top Wide Receiver than top-ranked players at other positions.
+a.
+Tight Ends showed relative inaccuracy of projections with limited upside concentrated among a few players, which means it is difficult, at the outset, to predict which Tight Ends will excel; based on this, @Kartes2025 suggested avoiding overpaying for mid-tier Tight Ends; to either draft a Tight End early or to wait until later rounds.
+
+This does *not* mean that you should intentionally draft a 6th ranked Running Back instead of the top-ranked Running Back.
+Instead, what it may mean is that, instead of drafting the top-ranked Running Back, you may be better off drafting the top-ranked Wide Receiver and to get a value player at Running Back in a subsequent pick.
 

### 2025-07-19 evaluating-prediction-accuracy.qmd
20250719 - descriptive text

@@
 
+Here are the seasonal accuracy indexes for players across all positions, but only among those who had high projected or actual points (i.e., the players that are most important to differentiate between):
 ```{r}
@@
 
+Here are the seasonal accuracy indexes for all Quarterbacks:
 ```{r}
@@
 
+A calibration plot of the seasonal projections for Quarterbacks is in @fig-calibrationPlotSeasonalQBs.
 ```{r}
@@
 
+Here are the seasonal accuracy indexes for only those Quarterbacks who had high projected or actual points (i.e., the players that are most important to differentiate between):
 ```{r}
@@
 
+Here are the seasonal accuracy indexes for all Running Backs:
 ```{r}
@@
 
+A calibration plot of the seasonal projections for Running Backs is in @fig-calibrationPlotSeasonalRBs.
 ```{r}
@@
 
+Here are the seasonal accuracy indexes for only those Running Backs who had high projected or actual points (i.e., the players that are most important to differentiate between):
 ```{r}
@@
 
+Here are the seasonal accuracy indexes for all Wide Receivers:
+
 ```{r}
@@
 ```
+
+A calibration plot of the seasonal projections for Wide Receivers is in @fig-calibrationPlotSeasonalWRs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Wide Receivers who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the seasonal accuracy indexes for all Tight Ends:
+
 ```{r}
@@
 ```
+
+A calibration plot of the seasonal projections for Tight Ends is in @fig-calibrationPlotSeasonalTEs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Tight Ends who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the seasonal accuracy indexes for all Kickers:
+
 ```{r}
@@
 ```
+
+A calibration plot of the seasonal projections for Kickers is in @fig-calibrationPlotSeasonalKs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Kickers who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the seasonal accuracy indexes for all Defensive Linemen:
+
 ```{r}
@@
 ```
+
+A calibration plot of the seasonal projections for Defensive Linemen is in @fig-calibrationPlotSeasonalDLs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Defensive Linemen who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the seasonal accuracy indexes for all Linebackers:
+
 ```{r}
@@
 ```
+
+A calibration plot of the seasonal projections for Linebackers is in @fig-calibrationPlotSeasonalLBs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Linebackers who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the seasonal accuracy indexes for all Defensive Backs:
+
 ```{r}
@@
 ```
+
+A calibration plot of the seasonal projections for Defensive backs is in @fig-calibrationPlotSeasonalDBs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Defensive Backs who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the weekly accuracy indexes for all players (across all positions):
+
 ```{r}
@@
 ```
+
+A calibration plot of the weekly projections for all players (across all positions) is in @fig-calibrationPlotWeeklyOverall.
 
@@
 
+Here are the weekly accuracy indexes for players across all positions, but only among those who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the weekly accuracy indexes for all Quarterbacks:
+
 ```{r}
@@
 ```
+
+A calibration plot of the weekly projections for Quarterbacks is in @fig-calibrationPlotWeeklyQBs.
 
@@
 
+Here are the weekly accuracy indexes for only those Quarterbacks who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the weekly accuracy indexes for all Running Backs:
+
 ```{r}
@@
 ```
+
+A calibration plot of the weekly projections for Running Backs is in @fig-calibrationPlotWeeklyRBs.
 
@@
 
+Here are the weekly accuracy indexes for only those Running Backs who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the weekly accuracy indexes for all Wide Receivers:
+
 ```{r}
@@
 ```
+
+A calibration plot of the weekly projections for Wide Receivers is in @fig-calibrationPlotWeeklyWRs.
 
@@
 
+Here are the weekly accuracy indexes for only those Wide Receivers who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the weekly accuracy indexes for all Tight Ends:
+
 ```{r}
@@
 ```
+
+A calibration plot of the weekly projections for Tight Ends is in @fig-calibrationPlotWeeklyTEs.
 
@@
 
+Here are the weekly accuracy indexes for only those Tight Ends who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the weekly accuracy indexes for all Kickers:
+
 ```{r}
@@
 ```
+
+A calibration plot of the weekly projections for Kickers is in @fig-calibrationPlotWeeklyKs.
 
@@
 
+Here are the weekly accuracy indexes for only those Kickers who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the weekly accuracy indexes for all Defensive Linemen:
+
 ```{r}
@@
 ```
+
+A calibration plot of the weekly projections for Defensive Linemen is in @fig-calibrationPlotWeeklyDLs.
 
@@
 
+Here are the weekly accuracy indexes for only those Defensive Linemen who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the weekly accuracy indexes for all Linebackers:
+
 ```{r}
@@
 ```
+
+A calibration plot of the weekly projections for Linebackers is in @fig-calibrationPlotWeeklyLBs.
 
@@
 
+Here are the weekly accuracy indexes for only those Linebackers who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 
+Here are the weekly accuracy indexes for all Defensive Backs:
+
 ```{r}
@@
 ```
+
+A calibration plot of the weekly projections for Linebackers is in @fig-calibrationPlotWeeklyDBs.
 
@@
 ```
+
+Here are the weekly accuracy indexes for only those Defensive Backs who had high projected or actual points (i.e., the players that are most important to differentiate between):
 

### 2025-07-19 evaluating-prediction-accuracy.qmd
20250719 - updats

@@
 
+Here are the seasonal accuracy indexes for all players (across all positions):
+
 ```{r}
@@
 
+A calibration plot of the seasonal projections for all players (across all positions) is in @fig-calibrationPlotSeasonalOverall.
 We calculate the 95% prediction interval based on a LOESS model, which allows nonlinear associations, using the `loess.sd()` function of the `msir` package [@R-msir; @Scrucca2011_packages].
@@
 Here are the seasonal accuracy indexes for players across all positions, but only among those who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 Here are the seasonal accuracy indexes for all Quarterbacks:
+
 ```{r}
@@
 A calibration plot of the seasonal projections for Quarterbacks is in @fig-calibrationPlotSeasonalQBs.
+
 ```{r}
@@
 Here are the seasonal accuracy indexes for only those Quarterbacks who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}
@@
 Here are the seasonal accuracy indexes for all Running Backs:
+
 ```{r}
@@
 A calibration plot of the seasonal projections for Running Backs is in @fig-calibrationPlotSeasonalRBs.
+
 ```{r}
@@
 Here are the seasonal accuracy indexes for only those Running Backs who had high projected or actual points (i.e., the players that are most important to differentiate between):
+
 ```{r}

### 2025-07-19 evaluating-prediction-accuracy.qmd
20250719 - summarizing accuracy by position

--- 2025-07-19 evaluating-prediction-accuracy.qmd
@@
 For instance, drafting a highly routed Wide Receiver may be safer than drafting a highly ranked Running Back, who may be more prone to injury.
-This finding was echoed by @Kartes2025 (archived at <https://perma.cc/R3WS-CAKD>) whose evaluation of the accuracy of projections by position found that:
+This finding is consistent with findings from @Petersen2017 [archived at <https://perma.cc/BG2W-ANUF>], who found that projections were more accurate for Quarterbacks and Wide Receivers than for Running Backs; projections were the least accurate for Kickers, Defensive Backs, and Defense/Special Teams.
+These finding were also extended by @Kartes2025 (archived at <https://perma.cc/R3WS-CAKD>) whose evaluation of the accuracy of projections by position found that:
 
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_overall <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal$raw_points,
+  actual = projectionsWithActuals_seasonal$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_overall <- round(accuracyIndices_seasonal_overall$rsquared, digits = 2)
+me_seasonal_overall <- round(accuracyIndices_seasonal_overall$ME, digits = 2)
+mae_seasonal_overall <- round(accuracyIndices_seasonal_overall$MAE, digits = 2)
+```
+
+A calibration plot of the seasonal projections for all players (across all positions) is in @fig-calibrationPlotSeasonalOverall.
 We calculate the 95% prediction interval based on a LOESS model, which allows nonlinear associations, using the `loess.sd()` function of the `msir` package [@R-msir; @Scrucca2011_packages].
@@
 
+Here are the seasonal accuracy indexes for players across all positions, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_overall_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_overall_subset <- round(accuracyIndices_seasonal_overall_subset$rsquared, digits = 2)
+me_seasonal_overall_subset <- round(accuracyIndices_seasonal_overall_subset$ME, digits = 2)
+mae_seasonal_overall_subset <- round(accuracyIndices_seasonal_overall_subset$MAE, digits = 2)
+```
+
 #### Quarterbacks {#sec-accuracyProjectionsSeasonalQB}
+
+Here are the seasonal accuracy indexes for all Quarterbacks:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_qb <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_qb$raw_points,
+  actual = projectionsWithActuals_seasonal_qb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_qb <- round(accuracyIndices_seasonal_qb$rsquared, digits = 2)
+me_seasonal_qb <- round(accuracyIndices_seasonal_qb$ME, digits = 2)
+mae_seasonal_qb <- round(accuracyIndices_seasonal_qb$MAE, digits = 2)
+```
+
+A calibration plot of the seasonal projections for Quarterbacks is in @fig-calibrationPlotSeasonalQBs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Quarterbacks who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_qb_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_qb$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_qb$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_qb_subset <- round(accuracyIndices_seasonal_qb_subset$rsquared, digits = 2)
+me_seasonal_qb_subset <- round(accuracyIndices_seasonal_qb_subset$ME, digits = 2)
+mae_seasonal_qb_subset <- round(accuracyIndices_seasonal_qb_subset$MAE, digits = 2)
+```
+
 #### Running Backs {#sec-accuracyProjectionsSeasonalRB}
+
+Here are the seasonal accuracy indexes for all Running Backs:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_rb <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_rb$raw_points,
+  actual = projectionsWithActuals_seasonal_rb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_rb <- round(accuracyIndices_seasonal_rb$rsquared, digits = 2)
+me_seasonal_rb <- round(accuracyIndices_seasonal_rb$ME, digits = 2)
+mae_seasonal_rb <- round(accuracyIndices_seasonal_rb$MAE, digits = 2)
+```
+
+A calibration plot of the seasonal projections for Running Backs is in @fig-calibrationPlotSeasonalRBs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Running Backs who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_rb_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_rb$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_rb$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_rb_subset <- round(accuracyIndices_seasonal_rb_subset$rsquared, digits = 2)
+me_seasonal_rb_subset <- round(accuracyIndices_seasonal_rb_subset$ME, digits = 2)
+mae_seasonal_rb_subset <- round(accuracyIndices_seasonal_rb_subset$MAE, digits = 2)
+```
+
 #### Wide Receivers {#sec-accuracyProjectionsSeasonalWR}
+
+Here are the seasonal accuracy indexes for all Wide Receivers:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_wr <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_wr$raw_points,
+  actual = projectionsWithActuals_seasonal_wr$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_wr <- round(accuracyIndices_seasonal_wr$rsquared, digits = 2)
+me_seasonal_wr <- round(accuracyIndices_seasonal_wr$ME, digits = 2)
+mae_seasonal_wr <- round(accuracyIndices_seasonal_wr$MAE, digits = 2)
+```
+
+A calibration plot of the seasonal projections for Wide Receivers is in @fig-calibrationPlotSeasonalWRs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Wide Receivers who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_wr_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_wr$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_wr$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_wr_subset <- round(accuracyIndices_seasonal_wr_subset$rsquared, digits = 2)
+me_seasonal_wr_subset <- round(accuracyIndices_seasonal_wr_subset$ME, digits = 2)
+mae_seasonal_wr_subset <- round(accuracyIndices_seasonal_wr_subset$MAE, digits = 2)
+```
+
 #### Tight Ends {#sec-accuracyProjectionsSeasonalTE}
+
+Here are the seasonal accuracy indexes for all Tight Ends:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_te <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_te$raw_points,
+  actual = projectionsWithActuals_seasonal_te$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_te <- round(accuracyIndices_seasonal_te$rsquared, digits = 2)
+me_seasonal_te <- round(accuracyIndices_seasonal_te$ME, digits = 2)
+mae_seasonal_te <- round(accuracyIndices_seasonal_te$MAE, digits = 2)
+```
+
+A calibration plot of the seasonal projections for Tight Ends is in @fig-calibrationPlotSeasonalTEs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Tight Ends who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_te_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_te$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_te$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_te_subset <- round(accuracyIndices_seasonal_te_subset$rsquared, digits = 2)
+me_seasonal_te_subset <- round(accuracyIndices_seasonal_te_subset$ME, digits = 2)
+mae_seasonal_te_subset <- round(accuracyIndices_seasonal_te_subset$MAE, digits = 2)
+```
+
 #### Kickers {#sec-accuracyProjectionsSeasonalK}
+
+Here are the seasonal accuracy indexes for all Kickers:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_k <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_k$raw_points,
+  actual = projectionsWithActuals_seasonal_k$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_k <- round(accuracyIndices_seasonal_k$rsquared, digits = 2)
+me_seasonal_k <- round(accuracyIndices_seasonal_k$ME, digits = 2)
+mae_seasonal_k <- round(accuracyIndices_seasonal_k$MAE, digits = 2)
+```
+
+A calibration plot of the seasonal projections for Kickers is in @fig-calibrationPlotSeasonalKs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Kickers who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_k_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_k$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_k$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_k_subset <- round(accuracyIndices_seasonal_k_subset$rsquared, digits = 2)
+me_seasonal_k_subset <- round(accuracyIndices_seasonal_k_subset$ME, digits = 2)
+mae_seasonal_k_subset <- round(accuracyIndices_seasonal_k_subset$MAE, digits = 2)
+```
+
 #### Defensive Linemen {#sec-accuracyProjectionsSeasonalDL}
+
+Here are the seasonal accuracy indexes for all Defensive Linemen:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_dl <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_dl$raw_points,
+  actual = projectionsWithActuals_seasonal_dl$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_dl <- round(accuracyIndices_seasonal_dl$rsquared, digits = 2)
+me_seasonal_dl <- round(accuracyIndices_seasonal_dl$ME, digits = 2)
+mae_seasonal_dl <- round(accuracyIndices_seasonal_dl$MAE, digits = 2)
+```
+
+A calibration plot of the seasonal projections for Defensive Linemen is in @fig-calibrationPlotSeasonalDLs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Defensive Linemen who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_dl_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_dl$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_dl$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_dl_subset <- round(accuracyIndices_seasonal_dl_subset$rsquared, digits = 2)
+me_seasonal_dl_subset <- round(accuracyIndices_seasonal_dl_subset$ME, digits = 2)
+mae_seasonal_dl_subset <- round(accuracyIndices_seasonal_dl_subset$MAE, digits = 2)
+```
+
 #### Linebackers {#sec-accuracyProjectionsSeasonalLB}
+
+Here are the seasonal accuracy indexes for all Linebackers:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_lb <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_lb$raw_points,
+  actual = projectionsWithActuals_seasonal_lb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_lb <- round(accuracyIndices_seasonal_lb$rsquared, digits = 2)
+me_seasonal_lb <- round(accuracyIndices_seasonal_lb$ME, digits = 2)
+mae_seasonal_lb <- round(accuracyIndices_seasonal_lb$MAE, digits = 2)
+```
+
+A calibration plot of the seasonal projections for Linebackers is in @fig-calibrationPlotSeasonalLBs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Linebackers who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_lb_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_lb$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_lb$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_lb_subset <- round(accuracyIndices_seasonal_lb_subset$rsquared, digits = 2)
+me_seasonal_lb_subset <- round(accuracyIndices_seasonal_lb_subset$ME, digits = 2)
+mae_seasonal_lb_subset <- round(accuracyIndices_seasonal_lb_subset$MAE, digits = 2)
+```
+
 #### Defensive Backs {#sec-accuracyProjectionsSeasonalDB}
+
+Here are the seasonal accuracy indexes for all Defensive Backs:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_db <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_db$raw_points,
+  actual = projectionsWithActuals_seasonal_db$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_db <- round(accuracyIndices_seasonal_db$rsquared, digits = 2)
+me_seasonal_db <- round(accuracyIndices_seasonal_db$ME, digits = 2)
+mae_seasonal_db <- round(accuracyIndices_seasonal_db$MAE, digits = 2)
+```
+
+A calibration plot of the seasonal projections for Defensive backs is in @fig-calibrationPlotSeasonalDBs.
 
@@
 
+Here are the seasonal accuracy indexes for only those Defensive Backs who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_db_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_seasonal_db$raw_points[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = projectionsWithActuals_seasonal_db$fantasyPoints[which(projectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_db_subset <- round(accuracyIndices_seasonal_db_subset$rsquared, digits = 2)
+me_seasonal_db_subset <- round(accuracyIndices_seasonal_db_subset$ME, digits = 2)
+mae_seasonal_db_subset <- round(accuracyIndices_seasonal_db_subset$MAE, digits = 2)
+```
+
 ### Weekly Projections {#sec-accuracyProjectionsWeekly}
@@
 #### Overall {#sec-accuracyProjectionsWeeklyOverall}
+
+Here are the weekly accuracy indexes for all players (across all positions):
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_overall <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly$raw_points,
+  actual = projectionsWithActuals_weekly$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_overall <- round(accuracyIndices_weekly_overall$rsquared, digits = 2)
+me_weekly_overall <- round(accuracyIndices_weekly_overall$ME, digits = 2)
+mae_weekly_overall <- round(accuracyIndices_weekly_overall$MAE, digits = 2)
+```
+
+A calibration plot of the weekly projections for all players (across all positions) is in @fig-calibrationPlotWeeklyOverall.
 
@@
 
+Here are the weekly accuracy indexes for players across all positions, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_weekly_overall_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_overall_subset <- round(accuracyIndices_weekly_overall_subset$rsquared, digits = 2)
+me_weekly_overall_subset <- round(accuracyIndices_weekly_overall_subset$ME, digits = 2)
+mae_weekly_overall_subset <- round(accuracyIndices_weekly_overall_subset$MAE, digits = 2)
+```
+
 #### Quarterbacks {#sec-accuracyProjectionsWeeklyQB}
+
+Here are the weekly accuracy indexes for all Quarterbacks:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_qb <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_qb$raw_points,
+  actual = projectionsWithActuals_weekly_qb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_qb <- round(accuracyIndices_weekly_qb$rsquared, digits = 2)
+me_weekly_qb <- round(accuracyIndices_weekly_qb$ME, digits = 2)
+mae_weekly_qb <- round(accuracyIndices_weekly_qb$MAE, digits = 2)
+```
+
+A calibration plot of the weekly projections for Quarterbacks is in @fig-calibrationPlotWeeklyQBs.
 
@@
 
+Here are the weekly accuracy indexes for only those Quarterbacks who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_weekly_qb_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_qb$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_qb$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_qb_subset <- round(accuracyIndices_weekly_qb_subset$rsquared, digits = 2)
+me_weekly_qb_subset <- round(accuracyIndices_weekly_qb_subset$ME, digits = 2)
+mae_weekly_qb_subset <- round(accuracyIndices_weekly_qb_subset$MAE, digits = 2)
+```
+
 #### Running Backs {#sec-accuracyProjectionsWeeklyRB}
+
+Here are the weekly accuracy indexes for all Running Backs:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_rb <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_rb$raw_points,
+  actual = projectionsWithActuals_weekly_rb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_rb <- round(accuracyIndices_weekly_rb$rsquared, digits = 2)
+me_weekly_rb <- round(accuracyIndices_weekly_rb$ME, digits = 2)
+mae_weekly_rb <- round(accuracyIndices_weekly_rb$MAE, digits = 2)
+```
+
+A calibration plot of the weekly projections for Running Backs is in @fig-calibrationPlotWeeklyRBs.
 
@@
 
+Here are the weekly accuracy indexes for only those Running Backs who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_weekly_rb_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_rb$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_rb$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_rb_subset <- round(accuracyIndices_weekly_rb_subset$rsquared, digits = 2)
+me_weekly_rb_subset <- round(accuracyIndices_weekly_rb_subset$ME, digits = 2)
+mae_weekly_rb_subset <- round(accuracyIndices_weekly_rb_subset$MAE, digits = 2)
+```
+
 #### Wide Receivers {#sec-accuracyProjectionsWeeklyWR}
+
+Here are the weekly accuracy indexes for all Wide Receivers:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_wr <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_wr$raw_points,
+  actual = projectionsWithActuals_weekly_wr$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_wr <- round(accuracyIndices_weekly_wr$rsquared, digits = 2)
+me_weekly_wr <- round(accuracyIndices_weekly_wr$ME, digits = 2)
+mae_weekly_wr <- round(accuracyIndices_weekly_wr$MAE, digits = 2)
+```
+
+A calibration plot of the weekly projections for Wide Receivers is in @fig-calibrationPlotWeeklyWRs.
 
@@
 
+Here are the weekly accuracy indexes for only those Wide Receivers who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_weekly_wr_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_wr$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_wr$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_wr_subset <- round(accuracyIndices_weekly_wr_subset$rsquared, digits = 2)
+me_weekly_wr_subset <- round(accuracyIndices_weekly_wr_subset$ME, digits = 2)
+mae_weekly_wr_subset <- round(accuracyIndices_weekly_wr_subset$MAE, digits = 2)
+```
+
 #### Tight Ends {#sec-accuracyProjectionsWeeklyTE}
+
+Here are the weekly accuracy indexes for all Tight Ends:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_te <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_te$raw_points,
+  actual = projectionsWithActuals_weekly_te$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_te <- round(accuracyIndices_weekly_te$rsquared, digits = 2)
+me_weekly_te <- round(accuracyIndices_weekly_te$ME, digits = 2)
+mae_weekly_te <- round(accuracyIndices_weekly_te$MAE, digits = 2)
+```
+
+A calibration plot of the weekly projections for Tight Ends is in @fig-calibrationPlotWeeklyTEs.
 
@@
 
+Here are the weekly accuracy indexes for only those Tight Ends who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_weekly_te_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_te$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_te$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_te_subset <- round(accuracyIndices_weekly_te_subset$rsquared, digits = 2)
+me_weekly_te_subset <- round(accuracyIndices_weekly_te_subset$ME, digits = 2)
+mae_weekly_te_subset <- round(accuracyIndices_weekly_te_subset$MAE, digits = 2)
+```
+
 #### Kickers {#sec-accuracyProjectionsWeeklyK}
+
+Here are the weekly accuracy indexes for all Kickers:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_k <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_k$raw_points,
+  actual = projectionsWithActuals_weekly_k$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_k <- round(accuracyIndices_weekly_k$rsquared, digits = 2)
+me_weekly_k <- round(accuracyIndices_weekly_k$ME, digits = 2)
+mae_weekly_k <- round(accuracyIndices_weekly_k$MAE, digits = 2)
+```
+
+A calibration plot of the weekly projections for Kickers is in @fig-calibrationPlotWeeklyKs.
 
@@
 
+Here are the weekly accuracy indexes for only those Kickers who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_weekly_k_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_k$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_k$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_k_subset <- round(accuracyIndices_weekly_k_subset$rsquared, digits = 2)
+me_weekly_k_subset <- round(accuracyIndices_weekly_k_subset$ME, digits = 2)
+mae_weekly_k_subset <- round(accuracyIndices_weekly_k_subset$MAE, digits = 2)
+```
+
 #### Defensive Linemen {#sec-accuracyProjectionsWeeklyDL}
+
+Here are the weekly accuracy indexes for all Defensive Linemen:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_dl <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_dl$raw_points,
+  actual = projectionsWithActuals_weekly_dl$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_dl <- round(accuracyIndices_weekly_dl$rsquared, digits = 2)
+me_weekly_dl <- round(accuracyIndices_weekly_dl$ME, digits = 2)
+mae_weekly_dl <- round(accuracyIndices_weekly_dl$MAE, digits = 2)
+```
+
+A calibration plot of the weekly projections for Defensive Linemen is in @fig-calibrationPlotWeeklyDLs.
 
@@
 
+Here are the weekly accuracy indexes for only those Defensive Linemen who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_weekly_dl_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_dl$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_dl$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_dl_subset <- round(accuracyIndices_weekly_dl_subset$rsquared, digits = 2)
+me_weekly_dl_subset <- round(accuracyIndices_weekly_dl_subset$ME, digits = 2)
+mae_weekly_dl_subset <- round(accuracyIndices_weekly_dl_subset$MAE, digits = 2)
+```
+
 #### Linebackers {#sec-accuracyProjectionsWeeklyLB}
+
+Here are the weekly accuracy indexes for all Linebackers:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_lb <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_lb$raw_points,
+  actual = projectionsWithActuals_weekly_lb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_lb <- round(accuracyIndices_weekly_lb$rsquared, digits = 2)
+me_weekly_lb <- round(accuracyIndices_weekly_lb$ME, digits = 2)
+mae_weekly_lb <- round(accuracyIndices_weekly_lb$MAE, digits = 2)
+```
+
+A calibration plot of the weekly projections for Linebackers is in @fig-calibrationPlotWeeklyLBs.
 
@@
 
+Here are the weekly accuracy indexes for only those Linebackers who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 
+```{r}
+#| include: false
+
+accuracyIndices_weekly_lb_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_lb$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_lb$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_lb_subset <- round(accuracyIndices_weekly_lb_subset$rsquared, digits = 2)
+me_weekly_lb_subset <- round(accuracyIndices_weekly_lb_subset$ME, digits = 2)
+mae_weekly_lb_subset <- round(accuracyIndices_weekly_lb_subset$MAE, digits = 2)
+```
+
 #### Defensive Backs {#sec-accuracyProjectionsWeeklyDB}
+
+Here are the weekly accuracy indexes for all Defensive Backs:
 
@@
 ```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_db <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_db$raw_points,
+  actual = projectionsWithActuals_weekly_db$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_db <- round(accuracyIndices_weekly_db$rsquared, digits = 2)
+me_weekly_db <- round(accuracyIndices_weekly_db$ME, digits = 2)
+mae_weekly_db <- round(accuracyIndices_weekly_db$MAE, digits = 2)
+```
+
+A calibration plot of the weekly projections for Linebackers is in @fig-calibrationPlotWeeklyDBs.
 
@@
 
+Here are the weekly accuracy indexes for only those Defensive Backs who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
 ```{r}
@@
 )
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_db_subset <- petersenlab::accuracyOverall(
+  predicted = projectionsWithActuals_weekly_db$raw_points[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = projectionsWithActuals_weekly_db$fantasyPoints[which(projectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_db_subset <- round(accuracyIndices_weekly_db_subset$rsquared, digits = 2)
+me_weekly_db_subset <- round(accuracyIndices_weekly_db_subset$ME, digits = 2)
+mae_weekly_db_subset <- round(accuracyIndices_weekly_db_subset$MAE, digits = 2)
+```
+
+### Summarizing the Accuracy By Position {#sec-predictionAccuracyByPosition}
+
+```{r}
+#| include: false
+
+seasonalAccuracy <- data.frame(
+  position = c(
+    "All",
+    "Quarterbacks",
+    "Running Backs",
+    "Wide Receivers",
+    "Tight Ends",
+    "Kickers",
+    "Defensive Linemen",
+    "Linebackers",
+    "Defensive Backs"
+  ), 
+  rsquared = c(
+    rsquare_seasonal_overall,
+    rsquare_seasonal_qb,
+    rsquare_seasonal_rb,
+    rsquare_seasonal_wr,
+    rsquare_seasonal_te,
+    rsquare_seasonal_k,
+    rsquare_seasonal_dl,
+    rsquare_seasonal_lb,
+    rsquare_seasonal_db
+  ),
+  me = c(
+    me_seasonal_overall,
+    me_seasonal_qb,
+    me_seasonal_rb,
+    me_seasonal_wr,
+    me_seasonal_te,
+    me_seasonal_k,
+    me_seasonal_dl,
+    me_seasonal_lb,
+    me_seasonal_db
+  ),
+  mae = c(
+    mae_seasonal_overall,
+    mae_seasonal_qb,
+    mae_seasonal_rb,
+    mae_seasonal_wr,
+    mae_seasonal_te,
+    mae_seasonal_k,
+    mae_seasonal_dl,
+    mae_seasonal_lb,
+    mae_seasonal_db
+  )
+)
+
+seasonalAccuracy_subset <- data.frame(
+  position = c(
+    "All",
+    "Quarterbacks",
+    "Running Backs",
+    "Wide Receivers",
+    "Tight Ends",
+    "Kickers",
+    "Defensive Linemen",
+    "Linebackers",
+    "Defensive Backs"
+  ), 
+  rsquared = c(
+    rsquare_seasonal_overall_subset,
+    rsquare_seasonal_qb_subset,
+    rsquare_seasonal_rb_subset,
+    rsquare_seasonal_wr_subset,
+    rsquare_seasonal_te_subset,
+    rsquare_seasonal_k_subset,
+    rsquare_seasonal_dl_subset,
+    rsquare_seasonal_lb_subset,
+    rsquare_seasonal_db_subset
+  ),
+  me = c(
+    me_seasonal_overall_subset,
+    me_seasonal_qb_subset,
+    me_seasonal_rb_subset,
+    me_seasonal_wr_subset,
+    me_seasonal_te_subset,
+    me_seasonal_k_subset,
+    me_seasonal_dl_subset,
+    me_seasonal_lb_subset,
+    me_seasonal_db_subset
+  ),
+  mae = c(
+    mae_seasonal_overall_subset,
+    mae_seasonal_qb_subset,
+    mae_seasonal_rb_subset,
+    mae_seasonal_wr_subset,
+    mae_seasonal_te_subset,
+    mae_seasonal_k_subset,
+    mae_seasonal_dl_subset,
+    mae_seasonal_lb_subset,
+    mae_seasonal_db_subset
+  )
+)
+
+weeklyAccuracy <- data.frame(
+  position = c(
+    "All",
+    "Quarterbacks",
+    "Running Backs",
+    "Wide Receivers",
+    "Tight Ends",
+    "Kickers",
+    "Defensive Linemen",
+    "Linebackers",
+    "Defensive Backs"
+  ), 
+  rsquared = c(
+    rsquare_weekly_overall,
+    rsquare_weekly_qb,
+    rsquare_weekly_rb,
+    rsquare_weekly_wr,
+    rsquare_weekly_te,
+    rsquare_weekly_k,
+    rsquare_weekly_dl,
+    rsquare_weekly_lb,
+    rsquare_weekly_db
+  ),
+  me = c(
+    me_weekly_overall,
+    me_weekly_qb,
+    me_weekly_rb,
+    me_weekly_wr,
+    me_weekly_te,
+    me_weekly_k,
+    me_weekly_dl,
+    me_weekly_lb,
+    me_weekly_db
+  ),
+  mae = c(
+    mae_weekly_overall,
+    mae_weekly_qb,
+    mae_weekly_rb,
+    mae_weekly_wr,
+    mae_weekly_te,
+    mae_weekly_k,
+    mae_weekly_dl,
+    mae_weekly_lb,
+    mae_weekly_db
+  )
+)
+
+weeklyAccuracy_subset <- data.frame(
+  position = c(
+    "All",
+    "Quarterbacks",
+    "Running Backs",
+    "Wide Receivers",
+    "Tight Ends",
+    "Kickers",
+    "Defensive Linemen",
+    "Linebackers",
+    "Defensive Backs"
+  ), 
+  rsquared = c(
+    rsquare_weekly_overall_subset,
+    rsquare_weekly_qb_subset,
+    rsquare_weekly_rb_subset,
+    rsquare_weekly_wr_subset,
+    rsquare_weekly_te_subset,
+    rsquare_weekly_k_subset,
+    rsquare_weekly_dl_subset,
+    rsquare_weekly_lb_subset,
+    rsquare_weekly_db_subset
+  ),
+  me = c(
+    me_weekly_overall_subset,
+    me_weekly_qb_subset,
+    me_weekly_rb_subset,
+    me_weekly_wr_subset,
+    me_weekly_te_subset,
+    me_weekly_k_subset,
+    me_weekly_dl_subset,
+    me_weekly_lb_subset,
+    me_weekly_db_subset
+  ),
+  mae = c(
+    mae_weekly_overall_subset,
+    mae_weekly_qb_subset,
+    mae_weekly_rb_subset,
+    mae_weekly_wr_subset,
+    mae_weekly_te_subset,
+    mae_weekly_k_subset,
+    mae_weekly_dl_subset,
+    mae_weekly_lb_subset,
+    mae_weekly_db_subset
+  )
+)
+```
+
+We summarize the accuracy of seasonal projections by position in Tables [-@tbl-seasonalAccuracyByPosition] and [-@tbl-seasonalAccuracyByPositionSubset].
+
+```{r}
+#| label: tbl-seasonalAccuracyByPosition
+#| tbl-cap: "Accuracy of Seasonal Projections By Position."
+#| echo: false
+
+knitr::kable(
+  seasonalAccuracy,
+  col.names = c("Position","$R^2$","Mean Error","Mean Absolute Error"),
+  booktabs = TRUE)
+```
+
+```{r}
+#| label: tbl-seasonalAccuracyByPositionSubset
+#| tbl-cap: "Accuracy of Seasonal Projections By Position, Among Players with Greater than 100 Projected or Actual Points."
+#| echo: false
+
+knitr::kable(
+  seasonalAccuracy_subset,
+  col.names = c("Position","$R^2$","Mean Error","Mean Absolute Error"),
+  booktabs = TRUE)
+```
+
+We summarize the accuracy of weekly projections by position in Tables [-@tbl-weeeklyAccuracyByPosition] and [-@tbl-weeklyAccuracyByPositionSubset].
+
+```{r}
+#| label: tbl-weeeklyAccuracyByPosition
+#| tbl-cap: "Accuracy of Weekly Projections By Position."
+#| echo: false
+
+knitr::kable(
+  weeklyAccuracy,
+  col.names = c("Position","$R^2$","Mean Error","Mean Absolute Error"),
+  booktabs = TRUE)
+```
+
+```{r}
+#| label: tbl-weeklyAccuracyByPositionSubset
+#| tbl-cap: "Accuracy of Weekly Projections By Position, Among Players with Greater than 100 Projected or Actual Points."
+#| echo: false
+
+knitr::kable(
+  weeklyAccuracy_subset,
+  col.names = c("Position","$R^2$","Mean Error","Mean Absolute Error"),
+  booktabs = TRUE)
 ```

### 2025-07-19 base-rates.qmd
20250719 - italics

@@
 > Two cab companies, the Green and the Blue, operate in the city.
-> You are given the following data:
+> You are given the following data:*
 > 
-> - 85% of the cabs in the city are Green and 15% are Blue.
-> - A witness identified the cab as Blue.
-> The court tested the reliability of the witness under the circumstances that existed on the night of the accident and concluded that the witness correctly identified each one of the two colors 80% of the time and failed 20% of the time.
+> - *85% of the cabs in the city are Green and 15% are Blue.*
+> - *A witness identified the cab as Blue.
+> The court tested the reliability of the witness under the circumstances that existed on the night of the accident and concluded that the witness correctly identified each one of the two colors 80% of the time and failed 20% of the time.*
 > 
-> What is the probability that the cab involved in the accident was Blue rather than Green?*
+> *What is the probability that the cab involved in the accident was Blue rather than Green?*
 > 

### 2025-07-19 evaluating-prediction-accuracy.qmd
20250719 - tweaks

@@
 
-#### Calibration for predicting a continuous outcome {#sec-calibrationContinuous}
+#### Calibration for Predicting a Continuous Outcome {#sec-calibrationContinuous}
 
@@
 The 95% confidence interval of the observed value, across all values of the predicted values, should include the diagonal reference line whose intercept is zero and whose slope is one.
-
-For instance, based on the intercept and slope of the [calibration plot](#sec-calibrationPlot) in Figure INSERT, the predictions are not well calibrated, despite having a slope near one, because the 95% confidence interval of the intercept does not include zero.
-The best-fit line is the yellow line.
-The intercept from the best-fit line is positive, as shown in the regression equation.
-This is a case of underprediction, where the predicted values are consistently less than the observed values.
-The confidence interval of the observed value (i.e., the purple band) is the interval within which we have 95% confidence that the true observed value would lie for a given predicted value, based on the model
-The 95% prediction interval of the observed value (i.e., the dashed red lines) is the interval within which we would expect that 95% of future observations would lie for a given predicted value.
-The black diagonal line indicates the reference line with an intercept of zero and a slope of one.
-The predictions would be significantly [miscalibrated](#sec-calibration) at a given level of the predicted values if the 95% confidence interval of the observed value does not include the reference line at that level of the predicted value.
-In this case, the 95% confidence interval of the observed value does not include the reference line (i.e., the actual observed value) at lower levels of the predicted values, so the predictions are [miscalibrated](#sec-calibration) lower levels of the predicted values.
-
-Gold-standard recommendations include examining the predicted values in relation to the observed values using locally estimated scatterplot smoothing (LOESS) [@Austin2014], such as in Figure INSERT.
+Gold-standard recommendations include examining the predicted values in relation to the observed values using locally estimated scatterplot smoothing (LOESS) [@Austin2014], such as in @fig-calibrationPlotSeasonalOverall.
 We can examine whether the LOESS-based 95% confidence interval of the observed value at every level of the predicted values includes the diagonal reference line (i.e., the actual observed value).
-In this case, the 95% confidence interval of the observed value does not include the reference line at lower levels of the predicted values, so the predictions are [miscalibrated](#sec-calibration) at lower levels of the predicted values.
 
@@
 
+The intercept and slope of the best-fit line can be estimated using regression, as in below:
+
+```{r}
+summary(lm(
+  fantasyPoints ~ raw_points,
+  data = projectionsWithActuals_seasonal
+))
+```
+
 A calibration plot of the seasonal projections for all players (across all positions) is in @fig-calibrationPlotSeasonalOverall.
+The best-fit LOESS line is the yellow line.
+The intercept from the best-fit line is positive; the slope is less than one.
+The predicted values are less than the observed values at lower levels of projected points, whereas the predicted values are higher than the observed values at higher levels of projected points, consistent with the overextremity form of [miscalibration](#sec-calibration).
+The confidence interval of the observed value (i.e., the purple band) is the interval within which we have 95% confidence that the true observed value would lie for a given predicted value, based on the model.
+The 95% prediction interval of the observed value (i.e., the dashed red lines) is the interval within which we would expect that 95% of future observations would lie for a given predicted value.
 We calculate the 95% prediction interval based on a LOESS model, which allows nonlinear associations, using the `loess.sd()` function of the `msir` package [@R-msir; @Scrucca2011_packages].
+The black diagonal line indicates the reference line with an intercept of zero and a slope of one.
+The predictions would be significantly [miscalibrated](#sec-calibration) at a given level of the predicted values if the 95% confidence interval of the observed value does not include the reference line at that level of the predicted value.
+In this case, based on the intercept and slope of the [calibration plot](#sec-calibrationPlot) in @fig-calibrationPlotSeasonalOverall, the 95% confidence interval of the observed value does not include the reference line (i.e., the actual observed value) at lower or upper levels of the predicted values, so the predictions are [miscalibrated](#sec-calibration) at lower and upper levels of the predicted values.
 

### 2025-07-19 evaluating-prediction-accuracy.qmd
20250719 - ROC curve

@@
 When dealing with a continuous outcome, discrimination is the strength of the association between the predictor and the outcome.
-Aspects of discrimination at a particular cutoff (e.g., sensitivity, specificity, area under the ROC curve) are described in INSERT.
+Aspects of discrimination at a particular cutoff (e.g., sensitivity, specificity, area under the ROC curve) are described in @sec-thresholdDependentAccuracy.
 
@@
 
+### Set a Cutoff/Threshold {#sec-setCutoff}
+
+For evaluating sensitivity, specificity, positive predictive value, negative predictive value, and the [receiver operating characteristic (ROC) curve](#sec-roc), we need to make a dichotomous decision/prediction.
+For this example, our goal is to predict which Running Backs or Wide Receivers will score 300 or more fantasy points.
+We will predict a player as scoring 300 or more fantasy points for a given projection source if that projection source projected the player to score 300 or more fantasy points.
+Thus, our cutoff or threshold is 300 or more projected points.
+The same player has multiple rows in the data file—one for each projection source.
+
+```{r}
+projectionsWithActuals_seasonal_subset <- projectionsWithActuals_seasonal %>% 
+  filter(position_group %in% c("RB","WR")) %>% 
+  filter(season == 2024)
+
+projectionsWithActuals_seasonal_subset$prediction <- ifelse(projectionsWithActuals_seasonal_subset$raw_points >= 300, 1, 0)
+projectionsWithActuals_seasonal_subset$truth <- ifelse(projectionsWithActuals_seasonal_subset$fantasyPoints >= 300, 1, 0)
+
+projectionsWithActuals_seasonal_subset$predictionFactor <- factor(
+  projectionsWithActuals_seasonal_subset$prediction,
+  levels = c(1,0),
+  labels = c("Prediction: Will Score 300+ Points", "Prediction: Will Not Score 300+ Points"))
+
+projectionsWithActuals_seasonal_subset$truthFactor <- factor(
+  projectionsWithActuals_seasonal_subset$truth,
+  levels = c(1,0),
+  labels = c("Truth: Scored 300+ Points", "Truth: Did Not Score 300+ Points")
+)
+
+table(projectionsWithActuals_seasonal_subset$predictionFactor)
+
+table(projectionsWithActuals_seasonal_subset$truthFactor)
+```
+
 ### Signal Detection Theory {#sec-sdt}
@@
 
-INSERT depicts an empirical ROC plot where lines are drawn to connect the hit and false alarm rates.
-
-INSERT depicts an ROC curve where a smoothed and fitted curve is drawn to connect the hit and false alarm rates.
+@fig-aucExampleEmpirical depicts an empirical ROC plot where lines are drawn to connect the hit and false alarm rates, in predicting Running Backs and Wide Receivers who score 300 or more fantasy points.
+
+```{r}
+#| label: fig-aucExampleEmpirical
+#| fig-cap: "Empirical Receiver Operating Characteristic (ROC) Curve for Predicting Running Backs and Wide Receivers that Score 300 or More Fantasy Points."
+#| fig-alt: "Empirical Receiver Operating Characteristic (ROC) Curve for Predicting Running Backs and Wide Receivers that Score 300 or More Fantasy Points."
+#| fig-width: 8
+#| fig-height: 8
+
+plot(
+  pROC::roc(
+    truth ~ raw_points,
+    projectionsWithActuals_seasonal_subset),
+  legacy.axes = TRUE,
+  print.auc = TRUE)
+```
+
+@fig-aucExample depicts an ROC curve where a smoothed and fitted curve is drawn to connect the hit and false alarm rates, in predicting Running Backs and Wide Receivers who score 300 or more fantasy points.
+
+```{r}
+#| label: fig-aucExample
+#| fig-cap: "Smoothed Receiver Operating Characteristic (ROC) Curve for Predicting Running Backs and Wide Receivers that Score 300 or More Fantasy Points."
+#| fig-alt: "Smoothed Receiver Operating Characteristic (ROC) Curve for Predicting Running Backs and Wide Receivers that Score 300 or More Fantasy Points."
+#| fig-width: 8
+#| fig-height: 8
+
+plot(
+  pROC::roc(
+    truth ~ raw_points,
+    projectionsWithActuals_seasonal_subset,
+    smooth = TRUE),
+  legacy.axes = TRUE,
+  print.auc = TRUE)
+```
 
@@
 
-To generate a 2x2 confusion matrix, we need to make a dichotomous decision/prediction.
-For this example, our goal is to predict which Running Backs or Wide Receivers will score 300 or more fantasy points.
-We will predict a player as scoring 300 or more fantasy points for a given projection source if that projection source projected the player to score 300 or more fantasy points.
-Thus, the same player has multiple rows in the data file—one for each projection source.
+Here is our 2x2 confusion matrix, predicting which Running Backs or Wide Receivers will score 300 or more fantasy points.
 

### 2025-07-19 decision-making.qmd
20250708 - track chunk execution times
20250717 - chapter overviews
20250719 - bracketing rate

@@
+{{< include _chunk-timing.qmd >}}
+
 # Decision Making in the Context of Uncertainty {#sec-decisionMaking}
+
+This chapter provides an overview of various considerations in judgment and decision making in the context of uncertainty.
 
@@
 ```
+
+### Load Data {#sec-decisionMakingLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/projectionsWithActuals_seasonal.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/projectionsWithActuals_weekly.RData", fsep = ""))
+#load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projections_seasonal.RData", fsep = ""))
+#load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projections_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/projectionsWithActuals_seasonal.RData")
+load(file = "./data/projectionsWithActuals_weekly.RData")
+#load(file = "./data/players_projections_seasonal.RData")
+#load(file = "./data/players_projections_weekly.RData")
+load(file = "./data/player_stats_seasonal.RData")
+load(file = "./data/player_stats_weekly.RData")
+```
+
+We created the `projectionsWithActuals_seasonal.RData` object in @sec-predictionAccuracyPrepareDataSeasonal; we created the `projectionsWithActuals_weekly.RData` object in @sec-predictionAccuracyPrepareDataWeekly.
 
@@
 - <https://www.actionnetwork.com/nfl/props>
+- <https://tools.32beatwriters.com>
+- <https://the-odds-api.com>
 
@@
 
+### Seasonal Projections {#sec-accuracyOfCrowdSeasonal}
+
 ```{r}
 # insert example using historical projections: wisdomOfCrowd()
+
+seasonalIndividualSources <- projectionsWithActuals_seasonal %>%
+  filter(!is.na(raw_points)) %>% 
+  filter(!is.na(fantasyPoints)) %>% 
+  group_by(player_id, season) %>%
+  filter(n() >= 2) %>% # at least 2 projections
+  summarise(
+    numProjections = n(),
+    type = "individual",
+    petersenlab::wisdomOfCrowd(
+      predicted = raw_points,
+      actual = unique(fantasyPoints),
+      dropUndefined = TRUE
+      )[1,],
+    .groups = "drop"
+  )
+
+seasonalCrowd <- projectionsWithActuals_seasonal %>%
+  filter(!is.na(raw_points)) %>% 
+  filter(!is.na(fantasyPoints)) %>% 
+  group_by(player_id, season) %>%
+  filter(n() >= 2) %>% # at least 2 projections
+  summarise(
+    numProjections = n(),
+    type = "crowd",
+    petersenlab::wisdomOfCrowd(
+      predicted = raw_points,
+      actual = unique(fantasyPoints),
+      dropUndefined = TRUE
+      )[2,],
+    .groups = "drop"
+  )
+
+seasonalIndividualSourcesAndCrowd <- bind_rows(
+  seasonalIndividualSources,
+  seasonalCrowd
+)
+
+seasonalIndividualSources <- seasonalIndividualSources %>% 
+  left_join(
+    player_stats_seasonal,
+    by = c("player_id", "season")
+  )
+
+seasonalCrowd <- seasonalCrowd %>% 
+  left_join(
+    player_stats_seasonal,
+    by = c("player_id", "season")
+  )
+
+seasonalIndividualSourcesAndCrowd <- seasonalIndividualSourcesAndCrowd %>% 
+  left_join(
+    player_stats_seasonal,
+    by = c("player_id", "season")
+  )
 ```
+
+```{r}
+#| label: fig-bracketingRateSeasonal
+#| fig-cap: "Distribution of Bracketing Rate Across Players For Seasonal Projections."
+#| fig-alt: "Distribution of Bracketing Rate Across Players For Seasonal Projections."
+
+ggplot2::ggplot(
+  data = seasonalIndividualSources %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = bracketingRate)
+) +
+  geom_histogram(
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  labs(
+    x = "Bracketing Rate",
+    y = "Count",
+    title = "Histogram of Bracketing Rate (Seasonal Projections)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
+```
+
+Proportion of players whose bracketing rate was zero:
+
+```{r}
+#length()
+```
+
+### Weekly Projections {#sec-accuracyOfCrowdWeekly}
+
+### Examples {#sec-accuracyOfCrowdExamples}
 

### 2025-07-19 evaluating-prediction-accuracy.qmd
20250719 - bracketing rate

@@
 players_projectedPoints_seasonal_combined$season <- as.integer(players_projectedPoints_seasonal_combined$season)
+players_projections_seasonal_average_merged$season <- as.integer(players_projections_seasonal_average_merged$season)
 
@@
 
+crowdAveragedProjectionsWithActuals_seasonal <- full_join(
+  player_stats_seasonal_subset_IDs,
+  players_projections_seasonal_average_merged %>% filter(avg_type == "average"),
+  by = c("mfl_id" = "id", "season"),
+  suffix = c("", "_proj"),
+)
+
 projectionsWithActuals_seasonal <- projectionsWithActuals_seasonal %>% 
@@
 
-playersWithHighProjectedOrActualPoints <- projectionsWithActuals_seasonal %>% 
-  filter(raw_points > 100 | fantasyPoints > 100) %>% 
-  select(player_id_season) %>% 
-  pull()
-
-projectionsWithActuals_seasonal_qb <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "QB")
-
-projectionsWithActuals_seasonal_rb <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "RB")
-
-projectionsWithActuals_seasonal_wr <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "WR")
-
-projectionsWithActuals_seasonal_te <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "TE")
-
-projectionsWithActuals_seasonal_k <- projectionsWithActuals_seasonal %>% 
-  filter(position == "K")
-
-projectionsWithActuals_seasonal_dl <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "DL")
-
-projectionsWithActuals_seasonal_lb <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "LB")
-
-projectionsWithActuals_seasonal_db <- projectionsWithActuals_seasonal %>% 
-  filter(position_group == "DB")
-```
-
-Players are uniquely identified by `player_id`-`season`-`data_src`:
+crowdAveragedProjectionsWithActuals_seasonal <- crowdAveragedProjectionsWithActuals_seasonal %>% 
+  unite(
+    "player_id_season",
+    player_id,
+    season,
+    remove = FALSE
+  )
+```
+
+Players in the `projectionsWithActuals_seasonal` object are (supposed to be) uniquely identified by `player_id`-`season`-`data_src`:
 
@@
 
+Players in the `crowdAveragedProjectionsWithActuals_seasonal` object are (supposed to be) uniquely identified by `player_id`-`season`-`avg_type`:
+
+```{r}
+crowdAveragedProjectionsWithActuals_seasonal %>% 
+  filter(!is.na(player_id)) %>% 
+  group_by(player_id, season, avg_type) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+We save the data object for use in other chapters:
+
+```{r}
+save(
+  projectionsWithActuals_seasonal, crowdAveragedProjectionsWithActuals_seasonal,
+  file = "./data/projectionsWithActuals_seasonal.RData"
+)
+```
+
+```{r}
+playersWithHighProjectedOrActualPoints <- projectionsWithActuals_seasonal %>% 
+  filter(raw_points > 100 | fantasyPoints > 100) %>% 
+  select(player_id_season) %>% 
+  pull()
+
+projectionsWithActuals_seasonal_qb <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "QB")
+
+projectionsWithActuals_seasonal_rb <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "RB")
+
+projectionsWithActuals_seasonal_wr <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "WR")
+
+projectionsWithActuals_seasonal_te <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "TE")
+
+projectionsWithActuals_seasonal_k <- projectionsWithActuals_seasonal %>% 
+  filter(position == "K")
+
+projectionsWithActuals_seasonal_dl <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "DL")
+
+projectionsWithActuals_seasonal_lb <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "LB")
+
+projectionsWithActuals_seasonal_db <- projectionsWithActuals_seasonal %>% 
+  filter(position_group == "DB")
+```
+
 #### Weekly Projections {#sec-predictionAccuracyPrepareDataWeekly}
@@
 players_projectedPoints_weekly_combined$season <- as.integer(players_projectedPoints_weekly_combined$season)
+players_projections_weekly_average_merged$season <- as.integer(players_projections_weekly_average_merged$season)
 
@@
   players_projectedPoints_weekly_combined,
+  by = c("mfl_id" = "id", "season", "week"),
+  suffix = c("", "_proj"),
+)
+
+crowdAveragedProjectionsWithActuals_weekly <- full_join(
+  player_stats_weekly_subset_IDs,
+  players_projections_weekly_average_merged,
   by = c("mfl_id" = "id", "season", "week"),
@@
 
-playersWithHighProjectedOrActualPoints_weekly <- projectionsWithActuals_weekly %>% 
-  filter(raw_points > 6 | fantasyPoints > 6) %>% 
-  select(player_id_season_week) %>% 
-  pull()
-
-projectionsWithActuals_weekly_qb <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "QB")
-
-projectionsWithActuals_weekly_rb <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "RB")
-
-projectionsWithActuals_weekly_wr <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "WR")
-
-projectionsWithActuals_weekly_te <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "TE")
-
-projectionsWithActuals_weekly_k <- projectionsWithActuals_weekly %>% 
-  filter(position == "K")
-
-projectionsWithActuals_weekly_dl <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "DL")
-
-projectionsWithActuals_weekly_lb <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "LB")
-
-projectionsWithActuals_weekly_db <- projectionsWithActuals_weekly %>% 
-  filter(position_group == "DB")
-```
-
-Players are uniquely identified by `player_id`-`season`-`week`-`data_src`:
+crowdAveragedProjectionsWithActuals_weekly <- crowdAveragedProjectionsWithActuals_weekly %>% 
+  unite(
+    "player_id_season_week",
+    player_id,
+    season,
+    week,
+    remove = FALSE
+  )
+```
+
+Players in the `projectionsWithActuals_weekly` object are (supposed to be) uniquely identified by `player_id`-`season`-`week`-`data_src`:
 
@@
   head()
+```
+
+Players in the `crowdAveragedProjectionsWithActuals_weekly` object are (supposed to be) uniquely identified by `player_id`-`season`-`week`-`avg_type`:
+
+```{r}
+crowdAveragedProjectionsWithActuals_weekly %>% 
+  filter(!is.na(player_id)) %>% 
+  group_by(player_id, season, week, avg_type) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+We save the data object for use in other chapters:
+
+```{r}
+save(
+  projectionsWithActuals_weekly, crowdAveragedProjectionsWithActuals_weekly,
+  file = "./data/projectionsWithActuals_weekly.RData"
+)
+```
+
+```{r}
+playersWithHighProjectedOrActualPoints_weekly <- projectionsWithActuals_weekly %>% 
+  filter(raw_points > 6 | fantasyPoints > 6) %>% 
+  select(player_id_season_week) %>% 
+  pull()
+
+projectionsWithActuals_weekly_qb <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "QB")
+
+projectionsWithActuals_weekly_rb <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "RB")
+
+projectionsWithActuals_weekly_wr <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "WR")
+
+projectionsWithActuals_weekly_te <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "TE")
+
+projectionsWithActuals_weekly_k <- projectionsWithActuals_weekly %>% 
+  filter(position == "K")
+
+projectionsWithActuals_weekly_dl <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "DL")
+
+projectionsWithActuals_weekly_lb <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "LB")
+
+projectionsWithActuals_weekly_db <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "DB")
 ```
@@
 
-Here are the weekly accuracy indexes for players across all positions, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+Here are the weekly accuracy indexes for players across all positions, but only among those who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
 
@@
 
-Here are the weekly accuracy indexes for only those Quarterbacks who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+Here are the weekly accuracy indexes for only those Quarterbacks who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
 
@@
 
-Here are the weekly accuracy indexes for only those Running Backs who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+Here are the weekly accuracy indexes for only those Running Backs who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
 
@@
 
-Here are the weekly accuracy indexes for only those Wide Receivers who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+Here are the weekly accuracy indexes for only those Wide Receivers who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
 
@@
 
-Here are the weekly accuracy indexes for only those Tight Ends who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+Here are the weekly accuracy indexes for only those Tight Ends who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
 
@@
 
-Here are the weekly accuracy indexes for only those Kickers who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+Here are the weekly accuracy indexes for only those Kickers who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
 
@@
 
-Here are the weekly accuracy indexes for only those Defensive Linemen who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+Here are the weekly accuracy indexes for only those Defensive Linemen who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
 
@@
 
-Here are the weekly accuracy indexes for only those Linebackers who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+Here are the weekly accuracy indexes for only those Linebackers who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
 
@@
 
-Here are the weekly accuracy indexes for only those Defensive Backs who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+Here are the weekly accuracy indexes for only those Defensive Backs who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
 
@@
 #| label: tbl-weeklyAccuracyByPositionSubset
-#| tbl-cap: "Accuracy of Weekly Projections By Position, Among Players with Greater than 100 Projected or Actual Points."
+#| tbl-cap: "Accuracy of Weekly Projections By Position, Among Players with Greater than 6 Projected or Actual Points."
 #| echo: false

### 2025-07-19 R-scripts/process_projections.R

@@ -17,6 +17,9 @@ filenames_week <- str_extract(filenames, "week\\d+") %>%
   str_remove("week") %>% 
   as.integer()
 
+filenames_season <- filenames_season[which(!is.na(filenames_week))]
+filenames_week <- filenames_week[which(!is.na(filenames_week))]
+
 objectNames_projectedPoints <- paste("players_projectedPoints_weekly_", filenames_season, "_", "week", filenames_week, sep = "")
 objectNames_projectedPoints_combined <- paste("players_projectedPoints_weekly_combined_", filenames_season, "_", "week", filenames_week, sep = "")
 objectNames_projections <- paste("players_projections_weekly_average_", filenames_season, "_", "week", filenames_week, sep = "")
@@ -145,12 +148,15 @@ filenamesSeasonal <- list.files(
 
 filenamesSeasonal_season <- str_extract(filenamesSeasonal, "\\d{4}")
 
-objectNames_seasonal <- paste("players_projectedPoints_seasonal_", filenamesSeasonal_season, sep = "")
+objectNames_projectedPoints_seasonal <- paste("players_projectedPoints_seasonal_", filenamesSeasonal_season, sep = "")
+objectNames_projections_seasonal <- paste("players_projections_seasonal_average_", filenamesSeasonal_season, sep = "")
 
 for(i in 1:length(filenamesSeasonal_season)){
-  objectName_seasonal_subset <- objectNames_seasonal[i]
+  objectName_projectedPoints_seasonal_subset <- objectNames_projectedPoints_seasonal[i]
+  objectName_projections_seasonal_subset <- objectNames_projections_seasonal[i]
   
   rm(players_projectedPoints_seasonal)
+  rm(players_projections_seasonal_average)
   
   fullFilepath <- filenamesSeasonal[i]
   
@@ -170,22 +176,33 @@ for(i in 1:length(filenamesSeasonal_season)){
   players_projectedPoints_seasonal_combined <- players_projectedPoints_seasonal_combined %>% 
     select(season, everything())
   
+  # Projections
+  players_projections_seasonal_average$season <- filenamesSeasonal_season[i]
+  
   # Assign to object
-  assign(objectName_seasonal_subset, players_projectedPoints_seasonal_combined)
+  assign(objectName_projectedPoints_seasonal_subset, players_projectedPoints_seasonal_combined)
+  assign(objectName_projections_seasonal_subset, players_projections_seasonal_average)
   
[…]

### 2025-07-20 decision-making.qmd
20250719 - bracketing rate

@@
 The [`petersenlab` package](https://cran.r-project.org/web/packages/petersenlab/index.html) [@R-petersenlab] has the `wisdomOfCrowd()` function that computes the overall accuracy of the crowd-averaged projections, including the bracketing rate of the individual projections.
-
-### Seasonal Projections {#sec-accuracyOfCrowdSeasonal}
-
-```{r}
-# insert example using historical projections: wisdomOfCrowd()
-
-seasonalIndividualSources <- projectionsWithActuals_seasonal %>%
-  filter(!is.na(raw_points)) %>% 
-  filter(!is.na(fantasyPoints)) %>% 
-  group_by(player_id, season) %>%
-  filter(n() >= 2) %>% # at least 2 projections
-  summarise(
-    numProjections = n(),
-    type = "individual",
-    petersenlab::wisdomOfCrowd(
-      predicted = raw_points,
-      actual = unique(fantasyPoints),
-      dropUndefined = TRUE
-      )[1,],
-    .groups = "drop"
-  )
-
-seasonalCrowd <- projectionsWithActuals_seasonal %>%
-  filter(!is.na(raw_points)) %>% 
-  filter(!is.na(fantasyPoints)) %>% 
-  group_by(player_id, season) %>%
-  filter(n() >= 2) %>% # at least 2 projections
-  summarise(
-    numProjections = n(),
-    type = "crowd",
-    petersenlab::wisdomOfCrowd(
-      predicted = raw_points,
-      actual = unique(fantasyPoints),
-      dropUndefined = TRUE
-      )[2,],
-    .groups = "drop"
-  )
-
-seasonalIndividualSourcesAndCrowd <- bind_rows(
-  seasonalIndividualSources,
-  seasonalCrowd
-)
-
-seasonalIndividualSources <- seasonalIndividualSources %>% 
-  left_join(
-    player_stats_seasonal,
-    by = c("player_id", "season")
-  )
-
-seasonalCrowd <- seasonalCrowd %>% 
-  left_join(
-    player_stats_seasonal,
-    by = c("player_id", "season")
-  )
-
-seasonalIndividualSourcesAndCrowd <- seasonalIndividualSourcesAndCrowd %>% 
-  left_join(
-    player_stats_seasonal,
-    by = c("player_id", "season")
-  )
-```
-
-```{r}
-#| label: fig-bracketingRateSeasonal
-#| fig-cap: "Distribution of Bracketing Rate Across Players For Seasonal Projections."
-#| fig-alt: "Distribution of Bracketing Rate Across Players For Seasonal Projections."
-
-ggplot2::ggplot(
-  data = seasonalIndividualSources %>%
-    filter(position_group %in% c("QB","RB","WR","TE")),
-  mapping = aes(
-    x = bracketingRate)
-) +
-  geom_histogram(
-    color = "#000000",
-    fill = "#0099F8"
-  ) +
-  labs(
-    x = "Bracketing Rate",
-    y = "Count",
-    title = "Histogram of Bracketing Rate (Seasonal Projections)"
-  ) +
-  theme_classic() +
-  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
-```
-
-Proportion of players whose bracketing rate was zero:
-
-```{r}
-#length()
-```
-
-### Weekly Projections {#sec-accuracyOfCrowdWeekly}
 
@@
 
+### Seasonal Projections {#sec-accuracyOfCrowdSeasonal}
+
+```{r}
+seasonalIndividualSources <- projectionsWithActuals_seasonal %>%
+  filter(!is.na(raw_points)) %>% 
+  filter(!is.na(fantasyPoints)) %>% 
+  group_by(player_id, season) %>%
+  filter(n() >= 2) %>% # at least 2 projections
+  summarise(
+    numProjections = n(),
+    type = "individual",
+    petersenlab::wisdomOfCrowd(
+      predicted = raw_points,
+      actual = unique(fantasyPoints),
+      dropUndefined = TRUE
+      )[1,],
+    .groups = "drop"
+  )
+
+seasonalCrowd <- projectionsWithActuals_seasonal %>%
+  filter(!is.na(raw_points)) %>% 
+  filter(!is.na(fantasyPoints)) %>% 
+  group_by(player_id, season) %>%
+  filter(n() >= 2) %>% # at least 2 projections
+  summarise(
+    numProjections = n(),
+    type = "crowd",
+    petersenlab::wisdomOfCrowd(
+      predicted = raw_points,
+      actual = unique(fantasyPoints),
+      dropUndefined = TRUE
+      )[2,],
+    .groups = "drop"
+  )
+
+seasonalIndividualSourcesAndCrowd <- bind_rows(
+  seasonalIndividualSources,
+  seasonalCrowd
+)
+
+seasonalIndividualSources <- seasonalIndividualSources %>% 
+  left_join(
+    player_stats_seasonal,
+    by = c("player_id", "season")
+  )
+
+seasonalCrowd <- seasonalCrowd %>% 
+  left_join(
+    player_stats_seasonal,
+    by = c("player_id", "season")
+  )
+
+seasonalIndividualSourcesAndCrowd <- seasonalIndividualSourcesAndCrowd %>% 
+  left_join(
+    player_stats_seasonal,
+    by = c("player_id", "season")
+  )
+```
+
+```{r}
+seasonalIndividualSources %>% 
+  select(player_display_name, position, season, numProjections, bracketingRate, ME:RMSLE)
+```
+
+```{r}
+#| label: fig-bracketingRateSeasonal
+#| fig-cap: "Distribution of Bracketing Rate Across Players For Seasonal Projections."
+#| fig-alt: "Distribution of Bracketing Rate Across Players For Seasonal Projections."
+
+ggplot2::ggplot(
+  data = seasonalIndividualSources %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = bracketingRate)
+) +
+  geom_histogram(
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  labs(
+    x = "Bracketing Rate",
+    y = "Count",
+    title = "Histogram of Bracketing Rate (Seasonal Projections)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
+```
+
+Proportion of players whose bracketing rate was zero:
+
+```{r}
+seasonalIndividualSources %>% 
+  filter(!is.na(bracketingRate)) %>% 
+  mutate(isZeroBracketingRate = ifelse(bracketingRate == 0, 1, 0)) %>% 
+  summarise(proportionZeroBracketingRate = sum(isZeroBracketingRate) / n())
+```
+
+### Weekly Projections {#sec-accuracyOfCrowdWeekly}
+
+```{r}
+weeklyIndividualSources <- projectionsWithActuals_weekly %>%
+  filter(!is.na(raw_points)) %>% 
+  filter(!is.na(fantasyPoints)) %>% 
+  group_by(player_id, season, week) %>%
+  filter(n() >= 2) %>% # at least 2 projections
+  summarise(
+    numProjections = n(),
+    type = "individual",
+    petersenlab::wisdomOfCrowd(
+      predicted = raw_points,
+      actual = unique(fantasyPoints),
+      dropUndefined = TRUE
+      )[1,],
+    .groups = "drop"
+  )
+
+weeklyCrowd <- projectionsWithActuals_weekly %>%
+  filter(!is.na(raw_points)) %>% 
+  filter(!is.na(fantasyPoints)) %>% 
+  group_by(player_id, season, week) %>%
+  filter(n() >= 2) %>% # at least 2 projections
+  summarise(
+    numProjections = n(),
+    type = "crowd",
+    petersenlab::wisdomOfCrowd(
+      predicted = raw_points,
+      actual = unique(fantasyPoints), # [1] error w/o [1] because have more than one unique value for a given player-season-week
+      dropUndefined = TRUE
+      )[2,],
+    .groups = "drop"
+  )
+
+weeklyIndividualSourcesAndCrowd <- bind_rows(
+  weeklyIndividualSources,
+  weeklyCrowd
+)
+
+weeklyIndividualSources <- weeklyIndividualSources %>% 
+  left_join(
+    player_stats_weekly,
+    by = c("player_id", "season", "week")
+  )
+
+weeklyCrowd <- weeklyCrowd %>% 
+  left_join(
+    player_stats_weekly,
+    by = c("player_id", "season", "week")
+  )
+
+weeklyIndividualSourcesAndCrowd <- weeklyIndividualSourcesAndCrowd %>% 
+  left_join(
+    player_stats_weekly,
+    by = c("player_id", "season", "week")
+  )
+```
+
+```{r}
+weeklyIndividualSources %>% 
+  select(player_display_name, position, season, week, numProjections, bracketingRate, ME:RMSLE)
+```
+
+```{r}
+#| label: fig-bracketingRateWeekly
+#| fig-cap: "Distribution of Bracketing Rate Across Players For Weekly Projections."
+#| fig-alt: "Distribution of Bracketing Rate Across Players For Weekly Projections."
+
+ggplot2::ggplot(
+  data = weeklyIndividualSources %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = bracketingRate)
+) +
+  geom_histogram(
+    color = "#000000",
+    fill = "#0099F8"
+  ) +
+  labs(
+    x = "Bracketing Rate",
+    y = "Count",
+    title = "Histogram of Bracketing Rate (Weekly Projections)"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
+```
+
+Proportion of players whose bracketing rate was zero:
+
+```{r}
+weeklyIndividualSources %>% 
+  filter(!is.na(bracketingRate)) %>% 
+  mutate(isZeroBracketingRate = ifelse(bracketingRate == 0, 1, 0)) %>% 
+  summarise(proportionZeroBracketingRate = sum(isZeroBracketingRate) / n())
+```
+
 ## How Well Do People Incorporate Advice From Others? {#sec-incoporateAdvice}

### 2025-07-20 evaluating-prediction-accuracy.qmd
20250719 - accuracy of crowd-averaged projections

@@
   filter(position_group == "DB")
+
+crowdAveragedProjectionsWithActuals_seasonal_qb <- crowdAveragedProjectionsWithActuals_seasonal %>% 
+  filter(position_group == "QB")
+
+crowdAveragedProjectionsWithActuals_seasonal_rb <- crowdAveragedProjectionsWithActuals_seasonal %>% 
+  filter(position_group == "RB")
+
+crowdAveragedProjectionsWithActuals_seasonal_wr <- crowdAveragedProjectionsWithActuals_seasonal %>% 
+  filter(position_group == "WR")
+
+crowdAveragedProjectionsWithActuals_seasonal_te <- crowdAveragedProjectionsWithActuals_seasonal %>% 
+  filter(position_group == "TE")
+
+crowdAveragedProjectionsWithActuals_seasonal_k <- crowdAveragedProjectionsWithActuals_seasonal %>% 
+  filter(position == "K")
+
+crowdAveragedProjectionsWithActuals_seasonal_dl <- crowdAveragedProjectionsWithActuals_seasonal %>% 
+  filter(position_group == "DL")
+
+crowdAveragedProjectionsWithActuals_seasonal_lb <- crowdAveragedProjectionsWithActuals_seasonal %>% 
+  filter(position_group == "LB")
+
+crowdAveragedProjectionsWithActuals_seasonal_db <- crowdAveragedProjectionsWithActuals_seasonal %>% 
+  filter(position_group == "DB")
 ```
@@
 projectionsWithActuals_weekly_db <- projectionsWithActuals_weekly %>% 
+  filter(position_group == "DB")
+
+crowdAveragedProjectionsWithActuals_weekly_qb <- crowdAveragedProjectionsWithActuals_weekly %>% 
+  filter(position_group == "QB")
+
+crowdAveragedProjectionsWithActuals_weekly_rb <- crowdAveragedProjectionsWithActuals_weekly %>% 
+  filter(position_group == "RB")
+
+crowdAveragedProjectionsWithActuals_weekly_wr <- crowdAveragedProjectionsWithActuals_weekly %>% 
+  filter(position_group == "WR")
+
+crowdAveragedProjectionsWithActuals_weekly_te <- crowdAveragedProjectionsWithActuals_weekly %>% 
+  filter(position_group == "TE")
+
+crowdAveragedProjectionsWithActuals_weekly_k <- crowdAveragedProjectionsWithActuals_weekly %>% 
+  filter(position == "K")
+
+crowdAveragedProjectionsWithActuals_weekly_dl <- crowdAveragedProjectionsWithActuals_weekly %>% 
+  filter(position_group == "DL")
+
+crowdAveragedProjectionsWithActuals_weekly_lb <- crowdAveragedProjectionsWithActuals_weekly %>% 
+  filter(position_group == "LB")
+
+crowdAveragedProjectionsWithActuals_weekly_db <- crowdAveragedProjectionsWithActuals_weekly %>% 
   filter(position_group == "DB")
@@
 
+Here are the seasonal accuracy indexes for crowd-averaged projections of all players (across all positions):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_overall_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_overall_crowd <- round(accuracyIndices_seasonal_overall_crowd$rsquared, digits = 2)
+me_seasonal_overall_crowd <- round(accuracyIndices_seasonal_overall_crowd$ME, digits = 2)
+mae_seasonal_overall_crowd <- round(accuracyIndices_seasonal_overall_crowd$MAE, digits = 2)
+```
+
+Here are the seasonal accuracy indexes for crowd-averaged projections of players across all positions, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal$points[which(crowdAveragedProjectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_overall_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal$points[which(crowdAveragedProjectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_overall_crowd_subset <- round(accuracyIndices_seasonal_overall_crowd_subset$rsquared, digits = 2)
+me_seasonal_overall_crowd_subset <- round(accuracyIndices_seasonal_overall_crowd_subset$ME, digits = 2)
+mae_seasonal_overall_crowd_subset <- round(accuracyIndices_seasonal_overall_crowd_subset$MAE, digits = 2)
+```
+
 #### Quarterbacks {#sec-accuracyProjectionsSeasonalQB}
@@
 
+Here are the seasonal accuracy indexes for crowd-averaged projections for Quarterbacks:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_qb$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_qb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_qb_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_qb$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_qb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_qb_crowd <- round(accuracyIndices_seasonal_qb_crowd$rsquared, digits = 2)
+me_seasonal_qb_crowd <- round(accuracyIndices_seasonal_qb_crowd$ME, digits = 2)
+mae_seasonal_qb_crowd <- round(accuracyIndices_seasonal_qb_crowd$MAE, digits = 2)
+```
+
+Here are the seasonal accuracy indexes for crowd-averaged projections of Quarterbacks, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_qb$points[which(crowdAveragedProjectionsWithActuals_seasonal_qb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_qb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_qb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_qb_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_qb$points[which(crowdAveragedProjectionsWithActuals_seasonal_qb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_qb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_qb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_qb_crowd_subset <- round(accuracyIndices_seasonal_qb_crowd_subset$rsquared, digits = 2)
+me_seasonal_qb_crowd_subset <- round(accuracyIndices_seasonal_qb_crowd_subset$ME, digits = 2)
+mae_seasonal_qb_crowd_subset <- round(accuracyIndices_seasonal_qb_crowd_subset$MAE, digits = 2)
+```
+
 #### Running Backs {#sec-accuracyProjectionsSeasonalRB}
@@
 
+Here are the seasonal accuracy indexes for crowd-averaged projections for Running Backs:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_rb$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_rb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_rb_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_rb$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_rb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_rb_crowd <- round(accuracyIndices_seasonal_rb_crowd$rsquared, digits = 2)
+me_seasonal_rb_crowd <- round(accuracyIndices_seasonal_rb_crowd$ME, digits = 2)
+mae_seasonal_rb_crowd <- round(accuracyIndices_seasonal_rb_crowd$MAE, digits = 2)
+```
+
+Here are the seasonal accuracy indexes for crowd-averaged projections of Running Backs, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_rb$points[which(crowdAveragedProjectionsWithActuals_seasonal_rb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_rb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_rb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_rb_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_rb$points[which(crowdAveragedProjectionsWithActuals_seasonal_rb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_rb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_rb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_rb_crowd_subset <- round(accuracyIndices_seasonal_rb_crowd_subset$rsquared, digits = 2)
+me_seasonal_rb_crowd_subset <- round(accuracyIndices_seasonal_rb_crowd_subset$ME, digits = 2)
+mae_seasonal_rb_crowd_subset <- round(accuracyIndices_seasonal_rb_crowd_subset$MAE, digits = 2)
+```
+
 #### Wide Receivers {#sec-accuracyProjectionsSeasonalWR}
@@
 
+Here are the seasonal accuracy indexes for crowd-averaged projections for Wide Receivers:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_wr$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_wr$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_wr_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_wr$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_wr$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_wr_crowd <- round(accuracyIndices_seasonal_wr_crowd$rsquared, digits = 2)
+me_seasonal_wr_crowd <- round(accuracyIndices_seasonal_wr_crowd$ME, digits = 2)
+mae_seasonal_wr_crowd <- round(accuracyIndices_seasonal_wr_crowd$MAE, digits = 2)
+```
+
+Here are the seasonal accuracy indexes for crowd-averaged projections of Wide Receivers, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_wr$points[which(crowdAveragedProjectionsWithActuals_seasonal_wr$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_wr$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_wr$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_wr_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_wr$points[which(crowdAveragedProjectionsWithActuals_seasonal_wr$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_wr$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_wr$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_wr_crowd_subset <- round(accuracyIndices_seasonal_wr_crowd_subset$rsquared, digits = 2)
+me_seasonal_wr_crowd_subset <- round(accuracyIndices_seasonal_wr_crowd_subset$ME, digits = 2)
+mae_seasonal_wr_crowd_subset <- round(accuracyIndices_seasonal_wr_crowd_subset$MAE, digits = 2)
+```
+
 #### Tight Ends {#sec-accuracyProjectionsSeasonalTE}
@@
 
+Here are the seasonal accuracy indexes for crowd-averaged projections for Tight Ends:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_te$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_te$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_te_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_te$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_te$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_te_crowd <- round(accuracyIndices_seasonal_te_crowd$rsquared, digits = 2)
+me_seasonal_te_crowd <- round(accuracyIndices_seasonal_te_crowd$ME, digits = 2)
+mae_seasonal_te_crowd <- round(accuracyIndices_seasonal_te_crowd$MAE, digits = 2)
+```
+
+Here are the seasonal accuracy indexes for crowd-averaged projections of Tight Ends, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_te$points[which(crowdAveragedProjectionsWithActuals_seasonal_te$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_te$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_te$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_te_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_te$points[which(crowdAveragedProjectionsWithActuals_seasonal_te$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_te$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_te$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_te_crowd_subset <- round(accuracyIndices_seasonal_te_crowd_subset$rsquared, digits = 2)
+me_seasonal_te_crowd_subset <- round(accuracyIndices_seasonal_te_crowd_subset$ME, digits = 2)
+mae_seasonal_te_crowd_subset <- round(accuracyIndices_seasonal_te_crowd_subset$MAE, digits = 2)
+```
+
 #### Kickers {#sec-accuracyProjectionsSeasonalK}
@@
 
+Here are the seasonal accuracy indexes for crowd-averaged projections for Kickers:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_k$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_k$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_k_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_k$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_k$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_k_crowd <- round(accuracyIndices_seasonal_k_crowd$rsquared, digits = 2)
+me_seasonal_k_crowd <- round(accuracyIndices_seasonal_k_crowd$ME, digits = 2)
+mae_seasonal_k_crowd <- round(accuracyIndices_seasonal_k_crowd$MAE, digits = 2)
+```
+
+Here are the seasonal accuracy indexes for crowd-averaged projections of Kickers, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_k$points[which(crowdAveragedProjectionsWithActuals_seasonal_k$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_k$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_k$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_k_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_k$points[which(crowdAveragedProjectionsWithActuals_seasonal_k$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_k$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_k$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_k_crowd_subset <- round(accuracyIndices_seasonal_k_crowd_subset$rsquared, digits = 2)
+me_seasonal_k_crowd_subset <- round(accuracyIndices_seasonal_k_crowd_subset$ME, digits = 2)
+mae_seasonal_k_crowd_subset <- round(accuracyIndices_seasonal_k_crowd_subset$MAE, digits = 2)
+```
+
 #### Defensive Linemen {#sec-accuracyProjectionsSeasonalDL}
@@
 
+Here are the seasonal accuracy indexes for crowd-averaged projections for Defensive Linemen:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_dl$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_dl$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_dl_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_dl$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_dl$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_dl_crowd <- round(accuracyIndices_seasonal_dl_crowd$rsquared, digits = 2)
+me_seasonal_dl_crowd <- round(accuracyIndices_seasonal_dl_crowd$ME, digits = 2)
+mae_seasonal_dl_crowd <- round(accuracyIndices_seasonal_dl_crowd$MAE, digits = 2)
+```
+
+Here are the seasonal accuracy indexes for crowd-averaged projections of Defensive Linemen, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_dl$points[which(crowdAveragedProjectionsWithActuals_seasonal_dl$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_dl$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_dl$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_dl_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_dl$points[which(crowdAveragedProjectionsWithActuals_seasonal_dl$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_dl$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_dl$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_dl_crowd_subset <- round(accuracyIndices_seasonal_dl_crowd_subset$rsquared, digits = 2)
+me_seasonal_dl_crowd_subset <- round(accuracyIndices_seasonal_dl_crowd_subset$ME, digits = 2)
+mae_seasonal_dl_crowd_subset <- round(accuracyIndices_seasonal_dl_crowd_subset$MAE, digits = 2)
+```
+
 #### Linebackers {#sec-accuracyProjectionsSeasonalLB}
@@
 
+Here are the seasonal accuracy indexes for crowd-averaged projections for Linebackers:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_lb$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_lb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_lb_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_lb$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_lb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_lb_crowd <- round(accuracyIndices_seasonal_lb_crowd$rsquared, digits = 2)
+me_seasonal_lb_crowd <- round(accuracyIndices_seasonal_lb_crowd$ME, digits = 2)
+mae_seasonal_lb_crowd <- round(accuracyIndices_seasonal_lb_crowd$MAE, digits = 2)
+```
+
+Here are the seasonal accuracy indexes for crowd-averaged projections of Linebackers, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_lb$points[which(crowdAveragedProjectionsWithActuals_seasonal_lb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_lb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_lb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_lb_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_lb$points[which(crowdAveragedProjectionsWithActuals_seasonal_lb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_lb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_lb$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_lb_crowd_subset <- round(accuracyIndices_seasonal_lb_crowd_subset$rsquared, digits = 2)
+me_seasonal_lb_crowd_subset <- round(accuracyIndices_seasonal_lb_crowd_subset$ME, digits = 2)
+mae_seasonal_lb_crowd_subset <- round(accuracyIndices_seasonal_lb_crowd_subset$MAE, digits = 2)
+```
+
 #### Defensive Backs {#sec-accuracyProjectionsSeasonalDB}
@@
 
+Here are the seasonal accuracy indexes for crowd-averaged projections for Defensive Backs:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_db$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_db$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_db_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_db$points,
+  actual = crowdAveragedProjectionsWithActuals_seasonal_db$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_db_crowd <- round(accuracyIndices_seasonal_db_crowd$rsquared, digits = 2)
+me_seasonal_db_crowd <- round(accuracyIndices_seasonal_db_crowd$ME, digits = 2)
+mae_seasonal_db_crowd <- round(accuracyIndices_seasonal_db_crowd$MAE, digits = 2)
+```
+
+Here are the seasonal accuracy indexes for crowd-averaged projections of Defensive Backs, but only among those who had high projected or actual points (i.e., players who had more than 100 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_db$points[which(crowdAveragedProjectionsWithActuals_seasonal_db$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_db$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_db$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_seasonal_db_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_seasonal_db$points[which(crowdAveragedProjectionsWithActuals_seasonal_db$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  actual = crowdAveragedProjectionsWithActuals_seasonal_db$fantasyPoints[which(crowdAveragedProjectionsWithActuals_seasonal_db$player_id_season %in% playersWithHighProjectedOrActualPoints)],
+  dropUndefined = TRUE
+)
+
+rsquare_seasonal_db_crowd_subset <- round(accuracyIndices_seasonal_db_crowd_subset$rsquared, digits = 2)
+me_seasonal_db_crowd_subset <- round(accuracyIndices_seasonal_db_crowd_subset$ME, digits = 2)
+mae_seasonal_db_crowd_subset <- round(accuracyIndices_seasonal_db_crowd_subset$MAE, digits = 2)
+```
+
 ### Weekly Projections {#sec-accuracyProjectionsWeekly}
@@
 
+Here are the weekly accuracy indexes for crowd-averaged projections of all players (across all positions):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_overall_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_overall_crowd <- round(accuracyIndices_weekly_overall_crowd$rsquared, digits = 2)
+me_weekly_overall_crowd <- round(accuracyIndices_weekly_overall_crowd$ME, digits = 2)
+mae_weekly_overall_crowd <- round(accuracyIndices_weekly_overall_crowd$MAE, digits = 2)
+```
+
+Here are the weekly accuracy indexes for crowd-averaged projections of players across all positions, but only among those who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly$points[which(crowdAveragedProjectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_overall_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly$points[which(crowdAveragedProjectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_overall_crowd_subset <- round(accuracyIndices_weekly_overall_crowd_subset$rsquared, digits = 2)
+me_weekly_overall_crowd_subset <- round(accuracyIndices_weekly_overall_crowd_subset$ME, digits = 2)
+mae_weekly_overall_crowd_subset <- round(accuracyIndices_weekly_overall_crowd_subset$MAE, digits = 2)
+```
+
 #### Quarterbacks {#sec-accuracyProjectionsWeeklyQB}
@@
 
+Here are the weekly accuracy indexes for crowd-averaged projections of Quarterbacks:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_qb$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_qb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_qb_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_qb$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_qb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_qb_crowd <- round(accuracyIndices_weekly_qb_crowd$rsquared, digits = 2)
+me_weekly_qb_crowd <- round(accuracyIndices_weekly_qb_crowd$ME, digits = 2)
+mae_weekly_qb_crowd <- round(accuracyIndices_weekly_qb_crowd$MAE, digits = 2)
+```
+
+Here are the weekly accuracy indexes for crowd-averaged projections of Quarterbacks, but only among those who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_qb$points[which(crowdAveragedProjectionsWithActuals_weekly_qb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_qb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_qb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_qb_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_qb$points[which(crowdAveragedProjectionsWithActuals_weekly_qb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_qb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_qb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_qb_crowd_subset <- round(accuracyIndices_weekly_qb_crowd_subset$rsquared, digits = 2)
+me_weekly_qb_crowd_subset <- round(accuracyIndices_weekly_qb_crowd_subset$ME, digits = 2)
+mae_weekly_qb_crowd_subset <- round(accuracyIndices_weekly_qb_crowd_subset$MAE, digits = 2)
+```
+
 #### Running Backs {#sec-accuracyProjectionsWeeklyRB}
@@
 
+Here are the weekly accuracy indexes for crowd-averaged projections of Running Backs:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_rb$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_rb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_rb_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_rb$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_rb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_rb_crowd <- round(accuracyIndices_weekly_rb_crowd$rsquared, digits = 2)
+me_weekly_rb_crowd <- round(accuracyIndices_weekly_rb_crowd$ME, digits = 2)
+mae_weekly_rb_crowd <- round(accuracyIndices_weekly_rb_crowd$MAE, digits = 2)
+```
+
+Here are the weekly accuracy indexes for crowd-averaged projections of Running Backs, but only among those who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_rb$points[which(crowdAveragedProjectionsWithActuals_weekly_rb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_rb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_rb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_rb_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_rb$points[which(crowdAveragedProjectionsWithActuals_weekly_rb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_rb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_rb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_rb_crowd_subset <- round(accuracyIndices_weekly_rb_crowd_subset$rsquared, digits = 2)
+me_weekly_rb_crowd_subset <- round(accuracyIndices_weekly_rb_crowd_subset$ME, digits = 2)
+mae_weekly_rb_crowd_subset <- round(accuracyIndices_weekly_rb_crowd_subset$MAE, digits = 2)
+```
+
 #### Wide Receivers {#sec-accuracyProjectionsWeeklyWR}
@@
 
+Here are the weekly accuracy indexes for crowd-averaged projections of Wide Receivers:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_wr$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_wr$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_wr_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_wr$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_wr$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_wr_crowd <- round(accuracyIndices_weekly_wr_crowd$rsquared, digits = 2)
+me_weekly_wr_crowd <- round(accuracyIndices_weekly_wr_crowd$ME, digits = 2)
+mae_weekly_wr_crowd <- round(accuracyIndices_weekly_wr_crowd$MAE, digits = 2)
+```
+
+Here are the weekly accuracy indexes for crowd-averaged projections of Wide Receivers, but only among those who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_wr$points[which(crowdAveragedProjectionsWithActuals_weekly_wr$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_wr$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_wr$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_wr_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_wr$points[which(crowdAveragedProjectionsWithActuals_weekly_wr$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_wr$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_wr$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_wr_crowd_subset <- round(accuracyIndices_weekly_wr_crowd_subset$rsquared, digits = 2)
+me_weekly_wr_crowd_subset <- round(accuracyIndices_weekly_wr_crowd_subset$ME, digits = 2)
+mae_weekly_wr_crowd_subset <- round(accuracyIndices_weekly_wr_crowd_subset$MAE, digits = 2)
+```
+
 #### Tight Ends {#sec-accuracyProjectionsWeeklyTE}
@@
 
+Here are the weekly accuracy indexes for crowd-averaged projections of Tight Ends:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_te$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_te$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_te_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_te$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_te$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_te_crowd <- round(accuracyIndices_weekly_te_crowd$rsquared, digits = 2)
+me_weekly_te_crowd <- round(accuracyIndices_weekly_te_crowd$ME, digits = 2)
+mae_weekly_te_crowd <- round(accuracyIndices_weekly_te_crowd$MAE, digits = 2)
+```
+
+Here are the weekly accuracy indexes for crowd-averaged projections of Tight Ends, but only among those who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_te$points[which(crowdAveragedProjectionsWithActuals_weekly_te$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_te$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_te$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_te_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_te$points[which(crowdAveragedProjectionsWithActuals_weekly_te$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_te$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_te$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_te_crowd_subset <- round(accuracyIndices_weekly_te_crowd_subset$rsquared, digits = 2)
+me_weekly_te_crowd_subset <- round(accuracyIndices_weekly_te_crowd_subset$ME, digits = 2)
+mae_weekly_te_crowd_subset <- round(accuracyIndices_weekly_te_crowd_subset$MAE, digits = 2)
+```
+
 #### Kickers {#sec-accuracyProjectionsWeeklyK}
@@
 
+Here are the weekly accuracy indexes for crowd-averaged projections of Kickers:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_k$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_k$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_k_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_k$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_k$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_k_crowd <- round(accuracyIndices_weekly_k_crowd$rsquared, digits = 2)
+me_weekly_k_crowd <- round(accuracyIndices_weekly_k_crowd$ME, digits = 2)
+mae_weekly_k_crowd <- round(accuracyIndices_weekly_k_crowd$MAE, digits = 2)
+```
+
+Here are the weekly accuracy indexes for crowd-averaged projections of Kickers, but only among those who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_k$points[which(crowdAveragedProjectionsWithActuals_weekly_k$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_k$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_k$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_k_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_k$points[which(crowdAveragedProjectionsWithActuals_weekly_k$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_k$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_k$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_k_crowd_subset <- round(accuracyIndices_weekly_k_crowd_subset$rsquared, digits = 2)
+me_weekly_k_crowd_subset <- round(accuracyIndices_weekly_k_crowd_subset$ME, digits = 2)
+mae_weekly_k_crowd_subset <- round(accuracyIndices_weekly_k_crowd_subset$MAE, digits = 2)
+```
+
 #### Defensive Linemen {#sec-accuracyProjectionsWeeklyDL}
@@
 
+Here are the weekly accuracy indexes for crowd-averaged projections of Defensive Linemen:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_dl$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_dl$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_dl_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_dl$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_dl$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_dl_crowd <- round(accuracyIndices_weekly_dl_crowd$rsquared, digits = 2)
+me_weekly_dl_crowd <- round(accuracyIndices_weekly_dl_crowd$ME, digits = 2)
+mae_weekly_dl_crowd <- round(accuracyIndices_weekly_dl_crowd$MAE, digits = 2)
+```
+
+Here are the weekly accuracy indexes for crowd-averaged projections of Defensive Linemen, but only among those who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_dl$points[which(crowdAveragedProjectionsWithActuals_weekly_dl$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_dl$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_dl$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_dl_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_dl$points[which(crowdAveragedProjectionsWithActuals_weekly_dl$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_dl$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_dl$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_dl_crowd_subset <- round(accuracyIndices_weekly_dl_crowd_subset$rsquared, digits = 2)
+me_weekly_dl_crowd_subset <- round(accuracyIndices_weekly_dl_crowd_subset$ME, digits = 2)
+mae_weekly_dl_crowd_subset <- round(accuracyIndices_weekly_dl_crowd_subset$MAE, digits = 2)
+```
+
 #### Linebackers {#sec-accuracyProjectionsWeeklyLB}
@@
 
+Here are the weekly accuracy indexes for crowd-averaged projections of Linebackers:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_lb$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_lb$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_lb_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_lb$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_lb$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_lb_crowd <- round(accuracyIndices_weekly_lb_crowd$rsquared, digits = 2)
+me_weekly_lb_crowd <- round(accuracyIndices_weekly_lb_crowd$ME, digits = 2)
+mae_weekly_lb_crowd <- round(accuracyIndices_weekly_lb_crowd$MAE, digits = 2)
+```
+
+Here are the weekly accuracy indexes for crowd-averaged projections of Linebackers, but only among those who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_lb$points[which(crowdAveragedProjectionsWithActuals_weekly_lb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_lb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_lb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_lb_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_lb$points[which(crowdAveragedProjectionsWithActuals_weekly_lb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_lb$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_lb$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_lb_crowd_subset <- round(accuracyIndices_weekly_lb_crowd_subset$rsquared, digits = 2)
+me_weekly_lb_crowd_subset <- round(accuracyIndices_weekly_lb_crowd_subset$ME, digits = 2)
+mae_weekly_lb_crowd_subset <- round(accuracyIndices_weekly_lb_crowd_subset$MAE, digits = 2)
+```
+
 #### Defensive Backs {#sec-accuracyProjectionsWeeklyDB}
@@
 
+Here are the weekly accuracy indexes for crowd-averaged projections of Defensive Backs:
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_db$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_db$fantasyPoints,
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_db_crowd <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_db$points,
+  actual = crowdAveragedProjectionsWithActuals_weekly_db$fantasyPoints,
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_db_crowd <- round(accuracyIndices_weekly_db_crowd$rsquared, digits = 2)
+me_weekly_db_crowd <- round(accuracyIndices_weekly_db_crowd$ME, digits = 2)
+mae_weekly_db_crowd <- round(accuracyIndices_weekly_db_crowd$MAE, digits = 2)
+```
+
+Here are the weekly accuracy indexes for crowd-averaged projections of Defensive Backs, but only among those who had high projected or actual points (i.e., players who had more than 6 projected or actual points; the players that are most important to differentiate between):
+
+```{r}
+petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_db$points[which(crowdAveragedProjectionsWithActuals_weekly_db$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_db$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_db$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+```
+
+```{r}
+#| include: false
+
+accuracyIndices_weekly_db_crowd_subset <- petersenlab::accuracyOverall(
+  predicted = crowdAveragedProjectionsWithActuals_weekly_db$points[which(crowdAveragedProjectionsWithActuals_weekly_db$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  actual = crowdAveragedProjectionsWithActuals_weekly_db$fantasyPoints[which(crowdAveragedProjectionsWithActuals_weekly_db$player_id_season_week %in% playersWithHighProjectedOrActualPoints_weekly)],
+  dropUndefined = TRUE
+)
+
+rsquare_weekly_db_crowd_subset <- round(accuracyIndices_weekly_db_crowd_subset$rsquared, digits = 2)
+me_weekly_db_crowd_subset <- round(accuracyIndices_weekly_db_crowd_subset$ME, digits = 2)
+mae_weekly_db_crowd_subset <- round(accuracyIndices_weekly_db_crowd_subset$MAE, digits = 2)
+```
+
 ### Summarizing the Accuracy By Position {#sec-predictionAccuracyByPosition}
@@
   position = c(
-    "All",
+    "All Positions",
     "Quarterbacks",
@@
 
+seasonalAccuracy_crowd <- data.frame(
+  position = c(
+    "All Positions",
+    "Quarterbacks",
+    "Running Backs",
+    "Wide Receivers",
+    "Tight Ends",
+    "Kickers",
+    "Defensive Linemen",
+    "Linebackers",
+    "Defensive Backs"
+  ), 
+  rsquared = c(
+    rsquare_seasonal_overall_crowd,
+    rsquare_seasonal_qb_crowd,
+    rsquare_seasonal_rb_crowd,
+    rsquare_seasonal_wr_crowd,
+    rsquare_seasonal_te_crowd,
+    rsquare_seasonal_k_crowd,
+    rsquare_seasonal_dl_crowd,
+    rsquare_seasonal_lb_crowd,
+    rsquare_seasonal_db_crowd
+  ),
+  me = c(
+    me_seasonal_overall_crowd,
+    me_seasonal_qb_crowd,
+    me_seasonal_rb_crowd,
+    me_seasonal_wr_crowd,
+    me_seasonal_te_crowd,
+    me_seasonal_k_crowd,
+    me_seasonal_dl_crowd,
+    me_seasonal_lb_crowd,
+    me_seasonal_db_crowd
+  ),
+  mae = c(
+    mae_seasonal_overall_crowd,
+    mae_seasonal_qb_crowd,
+    mae_seasonal_rb_crowd,
+    mae_seasonal_wr_crowd,
+    mae_seasonal_te_crowd,
+    mae_seasonal_k_crowd,
+    mae_seasonal_dl_crowd,
+    mae_seasonal_lb_crowd,
+    mae_seasonal_db_crowd
+  )
+)
+
+seasonalAccuracy_crowd_subset <- data.frame(
+  position = c(
+    "All Positions",
+    "Quarterbacks",
+    "Running Backs",
+    "Wide Receivers",
+    "Tight Ends",
+    "Kickers",
+    "Defensive Linemen",
+    "Linebackers",
+    "Defensive Backs"
+  ), 
+  rsquared = c(
+    rsquare_seasonal_overall_crowd_subset,
+    rsquare_seasonal_qb_crowd_subset,
+    rsquare_seasonal_rb_crowd_subset,
+    rsquare_seasonal_wr_crowd_subset,
+    rsquare_seasonal_te_crowd_subset,
+    rsquare_seasonal_k_crowd_subset,
+    rsquare_seasonal_dl_crowd_subset,
+    rsquare_seasonal_lb_crowd_subset,
+    rsquare_seasonal_db_crowd_subset
+  ),
+  me = c(
+    me_seasonal_overall_crowd_subset,
+    me_seasonal_qb_crowd_subset,
+    me_seasonal_rb_crowd_subset,
+    me_seasonal_wr_crowd_subset,
+    me_seasonal_te_crowd_subset,
+    me_seasonal_k_crowd_subset,
+    me_seasonal_dl_crowd_subset,
+    me_seasonal_lb_crowd_subset,
+    me_seasonal_db_crowd_subset
+  ),
+  mae = c(
+    mae_seasonal_overall_crowd_subset,
+    mae_seasonal_qb_crowd_subset,
+    mae_seasonal_rb_crowd_subset,
+    mae_seasonal_wr_crowd_subset,
+    mae_seasonal_te_crowd_subset,
+    mae_seasonal_k_crowd_subset,
+    mae_seasonal_dl_crowd_subset,
+    mae_seasonal_lb_crowd_subset,
+    mae_seasonal_db_crowd_subset
+  )
+)
+
 weeklyAccuracy <- data.frame(
   position = c(
-    "All",
+    "All Positions",
     "Quarterbacks",
@@
 )
+
+weeklyAccuracy_crowd <- data.frame(
+  position = c(
+    "All Positions",
+    "Quarterbacks",
+    "Running Backs",
+    "Wide Receivers",
+    "Tight Ends",
+    "Kickers",
+    "Defensive Linemen",
+    "Linebackers",
+    "Defensive Backs"
+  ), 
+  rsquared = c(
+    rsquare_weekly_overall_crowd,
+    rsquare_weekly_qb_crowd,
+    rsquare_weekly_rb_crowd,
+    rsquare_weekly_wr_crowd,
+    rsquare_weekly_te_crowd,
+    rsquare_weekly_k_crowd,
+    rsquare_weekly_dl_crowd,
+    rsquare_weekly_lb_crowd,
+    rsquare_weekly_db_crowd
+  ),
+  me = c(
+    me_weekly_overall_crowd,
+    me_weekly_qb_crowd,
+    me_weekly_rb_crowd,
+    me_weekly_wr_crowd,
+    me_weekly_te_crowd,
+    me_weekly_k_crowd,
+    me_weekly_dl_crowd,
+    me_weekly_lb_crowd,
+    me_weekly_db_crowd
+  ),
+  mae = c(
+    mae_weekly_overall_crowd,
+    mae_weekly_qb_crowd,
+    mae_weekly_rb_crowd,
+    mae_weekly_wr_crowd,
+    mae_weekly_te_crowd,
+    mae_weekly_k_crowd,
+    mae_weekly_dl_crowd,
+    mae_weekly_lb_crowd,
+    mae_weekly_db_crowd
+  )
+)
+
+weeklyAccuracy_crowd_subset <- data.frame(
+  position = c(
+    "All Positions",
+    "Quarterbacks",
+    "Running Backs",
+    "Wide Receivers",
+    "Tight Ends",
+    "Kickers",
+    "Defensive Linemen",
+    "Linebackers",
+    "Defensive Backs"
+  ), 
+  rsquared = c(
+    rsquare_weekly_overall_crowd_subset,
+    rsquare_weekly_qb_crowd_subset,
+    rsquare_weekly_rb_crowd_subset,
+    rsquare_weekly_wr_crowd_subset,
+    rsquare_weekly_te_crowd_subset,
+    rsquare_weekly_k_crowd_subset,
+    rsquare_weekly_dl_crowd_subset,
+    rsquare_weekly_lb_crowd_subset,
+    rsquare_weekly_db_crowd_subset
+  ),
+  me = c(
+    me_weekly_overall_crowd_subset,
+    me_weekly_qb_crowd_subset,
+    me_weekly_rb_crowd_subset,
+    me_weekly_wr_crowd_subset,
+    me_weekly_te_crowd_subset,
+    me_weekly_k_crowd_subset,
+    me_weekly_dl_crowd_subset,
+    me_weekly_lb_crowd_subset,
+    me_weekly_db_crowd_subset
+  ),
+  mae = c(
+    mae_weekly_overall_crowd_subset,
+    mae_weekly_qb_crowd_subset,
+    mae_weekly_rb_crowd_subset,
+    mae_weekly_wr_crowd_subset,
+    mae_weekly_te_crowd_subset,
+    mae_weekly_k_crowd_subset,
+    mae_weekly_dl_crowd_subset,
+    mae_weekly_lb_crowd_subset,
+    mae_weekly_db_crowd_subset
+  )
+)
 ```
@@
 
+```{r}
+#| label: tbl-seasonalCrowdAccuracyByPosition
+#| tbl-cap: "Accuracy of Crowd-Averaged Seasonal Projections By Position."
+#| echo: false
+
+knitr::kable(
+  seasonalAccuracy_crowd,
+  col.names = c("Position","$R^2$","Mean Error","Mean Absolute Error"),
+  booktabs = TRUE)
+```
+
+```{r}
+#| label: tbl-seasonalCrowdAccuracyByPositionSubset
+#| tbl-cap: "Accuracy of Crowd-Averaged Seasonal Projections By Position, Among Players with Greater than 100 Projected or Actual Points."
+#| echo: false
+
+knitr::kable(
+  seasonalAccuracy_crowd_subset,
+  col.names = c("Position","$R^2$","Mean Error","Mean Absolute Error"),
+  booktabs = TRUE)
+```
+
 We summarize the accuracy of weekly projections by position in Tables [-@tbl-weeeklyAccuracyByPosition] and [-@tbl-weeklyAccuracyByPositionSubset].
@@
   weeklyAccuracy_subset,
+  col.names = c("Position","$R^2$","Mean Error","Mean Absolute Error"),
+  booktabs = TRUE)
+```
+
+```{r}
+#| label: tbl-weeklyCrowdAccuracyByPosition
+#| tbl-cap: "Accuracy of Crowd-Averaged Weekly Projections By Position."
+#| echo: false
+
+knitr::kable(
+  weeklyAccuracy_crowd,
+  col.names = c("Position","$R^2$","Mean Error","Mean Absolute Error"),
+  booktabs = TRUE)
+```
+
+```{r}
+#| label: tbl-weeklyCrowdAccuracyByPositionSubset
+#| tbl-cap: "Accuracy of Crowd-Averaged Weekly Projections By Position, Among Players with Greater than 6 Projected or Actual Points."
+#| echo: false
+
+knitr::kable(
+  weeklyAccuracy_crowd_subset,
   col.names = c("Position","$R^2$","Mean Error","Mean Absolute Error"),

### 2025-07-20 decision-making.qmd
20250720 - summary of prediction accuracy

@@
 
-Even though the crowd tends to be more accurate than individual forecasters, crowd-averaged projections (at least among experts) are not necessarily highly accurate.
-In fantasy football, [DESCRIBE THEIR ACCURACY--BOTH FOR ALL PLAYERS AND WHEN SUBSETTING TO THE TOP X PLAYERS].
+Even though the crowd tends to be more accurate than individual forecasters [@Kartes2024; archived at <https://perma.cc/69F7-LLTN>; @Petersen2017; archived at <https://perma.cc/BG2W-ANUF>], crowd-averaged projections (at least among experts) are not necessarily highly accurate, as described in @sec-predictionAccuracyByPosition.
+In fantasy football, crowd-averaged seasonal projections explain ~60–75% of the variance in fantasy points among offensive players; however, the percent of variance explained drops to ~30% when considering only those players with high projected or actual points—who are the players that are the most important to distinguish, because those are the players that you are most likely trying to decide between when drafting.
 Nevertheless, individual sources tend to be even less accurate.
-[DESCRIBE THEIR ACCURACY].
+Individual projections sources tend to explain ~50–65% of the variance in fantasy points among offensive players; however, the percent of variance explained also drops considerably when considering only those players with high projected or actual points.
 

### 2025-07-20 evaluating-prediction-accuracy.qmd
20250720 - summary of prediction accuracy

@@
 Below are estimates of projection accuracy and [calibration plots](#sec-calibrationPlot) for [seasonal](#sec-accuracyProjectionsSeasonal) and [weekly](#sec-accuracyProjectionsWeekly) projections.
-In general, projections showed moderate accuracy in predicting players' fantasy performance.
-Projections showed moderate accuracy in terms of [discrimination](#sec-discrimination), based on [$R^2$](#sec-coefficientOfDetermination).
-However, projections were not highly accurate.
-Projections showed some [miscalibration](#sec-calibrationIndices) with respect to actual fantasy points.
-In general, projected showed a modest tendency to underpredict performance at lower levels of projected points and a moderate-to-strong tendency overpredict performance at higher levels of projected points, an example of the overextremity form of [miscalibration](#sec-calibration).
-As a reminder, the overextremity form of [miscalibration](#sec-calibration) is an example of the overprecision form of [overconfidence](#sec-cognitiveBiasesOverconfidence).
-That is, projection sources appear to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their projections of player performance.
-Relatedly, crowd-averaged projections tend to be more accurate than individual sources of projections [@Kartes2024; archived at <https://perma.cc/69F7-LLTN>; @Petersen2017; archived at <https://perma.cc/BG2W-ANUF>].
-
-The association between projected and actual fantasy points tended to flatten at the higher levels of projected points, for some positions, including Running Backs, Kickers, and defensive players.
-A consequence of this pattern, from the perspective of draft strategy, is that it would not make sense to draft players projected to score more than the flattening of the projected–actual curve.
-For instance, drafting a highly routed Wide Receiver may be safer than drafting a highly ranked Running Back, who may be more prone to injury.
-This finding is consistent with findings from @Petersen2017 [archived at <https://perma.cc/BG2W-ANUF>], who found that projections were more accurate for Quarterbacks and Wide Receivers than for Running Backs; projections were the least accurate for Kickers, Defensive Backs, and Defense/Special Teams.
-These finding were also extended by @Kartes2025 (archived at <https://perma.cc/R3WS-CAKD>) whose evaluation of the accuracy of projections by position found that:
-
-a.
-Quarterbacks with preseason ranks of 6–10 showed the greatest systematic underperformance relative to their projections, suggesting that you should either draft an elite Quarterback (ranked 1–5) or wait until later rounds to draft multiple lower-ranked Quarterbacks.
-a. Elite Running Backs (with preseason ranks of 1–5) showed strong underperformance (relative to projections), suggesting that you should draft Running Backs ranked 6 and below.
-a. Wide Receivers showed among the more consistent and predictable patterns in projections, suggesting that it may be safer to draft a top Wide Receiver than top-ranked players at other positions.
-a.
-Tight Ends showed relative inaccuracy of projections with limited upside concentrated among a few players, which means it is difficult, at the outset, to predict which Tight Ends will excel; based on this, @Kartes2025 suggested avoiding overpaying for mid-tier Tight Ends; to either draft a Tight End early or to wait until later rounds.
-
-This does *not* mean that you should intentionally draft a 6th ranked Running Back instead of the top-ranked Running Back.
-Instead, what it may mean is that, instead of drafting the top-ranked Running Back, you may be better off drafting the top-ranked Wide Receiver and to get a value player at Running Back in a subsequent pick.
 
@@
 mae_seasonal_overall <- round(accuracyIndices_seasonal_overall$MAE, digits = 2)
+```
+
+In the data object, each projection source has a row for each player and season.
+Thus, the $R^2$ may be higher for this aggregated object than it is for many of the individual sources.
+To calculate the accuracy by source, we could use the following code:
+
+```{r}
+accuracy_by_source <- projectionsWithActuals_seasonal %>%
+  filter(!is.na(data_src)) %>% 
+  filter(data_src != "FantasyPros") %>% # remove crowd-averaged source; we compare to crowd average later
+  group_by(data_src) %>%
+  summarise(
+    accuracy = list(
+      accuracyOverall(
+        predicted = raw_points,
+        actual = fantasyPoints,
+        dropUndefined = TRUE
+      )
+    ),
+    .groups = "drop"
+  )
+
+accuracy_by_source
+accuracy_by_source$accuracy
 ```
@@
 
+We summarize the accuracy of crowd-averaged seasonal projections by position in Tables [-@tbl-seasonalCrowdAccuracyByPosition] and [-@tbl-seasonalCrowdAccuracyByPositionSubset].
+
 ```{r}
@@
 
+We summarize the accuracy of crowd-averaged seasonal projections by position in Tables [-@tbl-weeklyCrowdAccuracyByPosition] and [-@tbl-weeklyCrowdAccuracyByPositionSubset].
+
 ```{r}
@@
 
+In general, projections showed moderate accuracy in predicting players' fantasy performance.
+Projections showed moderate accuracy in terms of [discrimination](#sec-discrimination), based on [$R^2$](#sec-coefficientOfDetermination).
+However, projections were not highly accurate.
+Projections showed some [miscalibration](#sec-calibrationIndices) with respect to actual fantasy points.
+In general, projected showed a modest tendency to underpredict performance at lower levels of projected points and a moderate-to-strong tendency overpredict performance at higher levels of projected points, an example of the overextremity form of [miscalibration](#sec-calibration).
+As a reminder, the overextremity form of [miscalibration](#sec-calibration) is an example of the overprecision form of [overconfidence](#sec-cognitiveBiasesOverconfidence).
+That is, projection sources appear to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their projections of player performance.
+
+For seasonal projections, in terms of comparison by position, projections showed greatest discrimination ($R^2$) for Quarterbacks, followed by Running Backs/Tight Ends, and Wide Receivers.
+Predictions for defensive players and kickers showed a negative $R^2$, indicating that projections would have been more accurate merely by predicting the mean of the actual values for every observation (i.e., [predicting from the base rate](#sec-predictingFromBaseRate)).
+In terms of comparison by projection source, crowd-averaged projections showed greater accuracy than individual projection sources (i.e., "experts") in terms of discrimination ($R^2$) and calibration (MAE).
+This was especially the case when examining individual sources separately (rather than in one combined object).
+
+For weekly projections, in terms of comparison by position, projections showed greatest discrimination ($R^2$) for Running Backs, followed by Wide Receivers, Quarterbacks, and Tight Ends.
+Predictions for defensive players and kickers showed a negative $R^2$, indicating that projections would have been more accurate merely by predicting the mean of the actual values for every observation (i.e., [predicting from the base rate](#sec-predictingFromBaseRate)).
+In terms of comparison by projection source, crowd-averaged projections showed greater accuracy than individual projection sources in terms of discrimination ($R^2$) and calibration (MAE).
+
+In summary, we found that, for both seasonal and weekly projections, crowd-averaged projections showed greater accuracy than individual projection sources (i.e., "experts"), consistent with prior findings [@Kartes2024; archived at <https://perma.cc/69F7-LLTN>; @Petersen2017; archived at <https://perma.cc/BG2W-ANUF>].
+Nevertheless projections—including individual sources and crowd-averaged—were not highly accurate.
+Projections tended not to explain more than about two-thirds of the variance in players' performance, and frequently explained well less than half of the variance, particularly when distinguishing between players with high (i.e, more than 100) projected or actual points—who are the players that are the most important to distinguish, because those are the players that you are most likely trying to decide between when drafting.
+Thus, although projections for offensive positions (besides Kickers) do better than chance, there remains considerable room for improvement in predictive accuracy.
+The modest accuracy of projections likely reflects that human behavior is multiply determined and hard to predict.
+
+The association between projected and actual fantasy points tended to flatten at the higher levels of projected points, for some positions, including Running Backs, Kickers, and defensive players.
+A consequence of this pattern, from the perspective of draft strategy, is that it would not make sense to draft players projected to score more than the flattening of the projected–actual curve.
+For instance, drafting a highly routed Wide Receiver may be safer than drafting a highly ranked Running Back, who may be more prone to injury.
+This finding is consistent with findings from @Petersen2017 [archived at <https://perma.cc/BG2W-ANUF>], who found that projections were more accurate for Quarterbacks and Wide Receivers than for Running Backs; projections were the least accurate for Kickers, Defensive Backs, and Defense/Special Teams.
+These finding were also extended by @Kartes2025 (archived at <https://perma.cc/R3WS-CAKD>) whose evaluation of the accuracy of projections by position found that:
+
+a.
+Quarterbacks with preseason ranks of 6–10 showed the greatest systematic underperformance relative to their projections, suggesting that you should either draft an elite Quarterback (ranked 1–5) or wait until later rounds to draft multiple lower-ranked Quarterbacks.
+a. Elite Running Backs (with preseason ranks of 1–5) showed strong underperformance (relative to projections), suggesting that you should draft Running Backs ranked 6 and below.
+a. Wide Receivers showed among the more consistent and predictable patterns in projections, suggesting that it may be safer to draft a top Wide Receiver than top-ranked players at other positions.
+a.
+Tight Ends showed relative inaccuracy of projections with limited upside concentrated among a few players, which means it is difficult, at the outset, to predict which Tight Ends will excel; based on this, @Kartes2025 suggested avoiding overpaying for mid-tier Tight Ends; to either draft a Tight End early or to wait until later rounds.
+
+This does *not* mean that you should intentionally draft a 6th ranked Running Back instead of the top-ranked Running Back.
+Instead, what it may mean is that, instead of drafting the top-ranked Running Back, you may be better off drafting the top-ranked Wide Receiver and to get a value player at Running Back in a subsequent pick.
+
 ## Ways to Improve Prediction Accuracy {#sec-waysToImprovePredictionAccuracy}
@@
 Experts' predictions from many different domains tend to be inaccurate, including political scientists [@Tetlock2017], physicians [@Koehler2002], clinical psychologists [@Oskamp1965], stock market traders and corporate financial officers [@Skala2008], seismologists' predictions of earthquakes [@Hough2016], economists' predictions about the economy [@Makridakis2009], lawyers [@Koehler2002], and business managers [@Russo1992].
-Thus, I would not put much confidence in the predictions by fantasy football fundits.
+Thus, I would not put much confidence in the predictions by fantasy football pundits.
 The most common pattern of experts' predictions is that they show overextremity, that is, their predictions have probability judgments that tend to be too extreme, as described in Section @sec-calibration.

### 2025-07-20 machine-learning.qmd
20250717 - chapter overviews
20250720 - summary of prediction accuracy

@@
 # Machine Learning {#sec-machineLearning}
+
+This chapter provides an overview of machine learning.
 
@@
 ```{r}
+#| include: false
+
+#-todo:
+#-calculate years_of_experience
+#-use common name for the same (gsis_id) ID variable
+```
+
+```{r}
 #| label: prepare-data-merging-machine-learning
@@
 # Prepare data for merging
-#-todo: calculate years_of_experience
-## Use common name for the same (gsis_id) ID variable
 
@@
 ```{r}
+#| include: false
+
+#-todo:
+#-remove redundant variables
+```
+
+```{r}
 #| label: merge-data1-machine-learning
@@
 # Create lists of objects to merge, depending on data structure: id; or id-season; or id-season-week
-#-todo: remove redundant variables
 playerListToMerge <- list(

### 2025-07-20 evaluating-prediction-accuracy.qmd
20250720 - accuracy of seasonal versus weekly projections

@@
 
+Prediction accuracy tended to be higher for seasonal projections than for weekly projections—likely because of the greater number of "trials" (i.e., plays) in a season compared to a game, so more of the random error averages out.
+The weaker accuracy weekly projections makes betting on daily fantasy sports a gamble.
+
 The association between projected and actual fantasy points tended to flatten at the higher levels of projected points, for some positions, including Running Backs, Kickers, and defensive players.

### 2025-07-20 cluster-analysis.qmd
20250708 - track chunk execution times
20250717 - chapter overviews
20250720 - conclusion sections

@@
+{{< include _chunk-timing.qmd >}}
+
 # Cluster Analysis {#sec-clusterAnalysis}
+
+This chapter provides an overview of cluster analysis.
 
@@
 Based on this analysis (and the variables included), there appear to be three types of Wide Receivers.
-Type 1 Wide Receivers includes the Elite WR1s who are strong possession receivers (note: not all players in a given cluster map on perfectly to the typology—i.e., not all Type 1 Wide Receivers are elite WR1s).
-They tend to have the lowest drop percentage, the shortest average depth of target, and the fewest receiving air yards per reception.
-They tend to have the most receptions per season and break the most tackles.
+We examined the following variables: the player's vertical jump in the NFL Combine,40-yard-dash time in the NFL Combine, height, weight, average depth of target, drop percentage, receiving air yards per reception, broken tackles, and receptions per season.
 
-Type 2 Wide Receivers includes the consistent contributor, WR2 types.
+Type 1 Wide Receivers included the Elite WR1s who are strong possession receivers (note: not all players in a given cluster map on perfectly to the typology—i.e., not all Type 1 Wide Receivers are elite WR1s).
+They tended to have the lowest drop percentage, the shortest average depth of target, and the fewest receiving air yards per reception.
+They tended to have the most receptions per season and break the most tackles.
+
+Type 2 Wide Receivers included the consistent contributor, WR2 types.
 They had fewer receptions and fewer broken tackles than Type 1 Wide Receivers.
@@
 
-Type 3 Wide Receivers includes the deep threats.
-They have the greatest average depth of target and the most receiving yards per reception.
-However, they also have the fewest receptions, the highest drop percentage, and the fewest broken tackles.
+Type 3 Wide Receivers included the deep threats.
+They had the greatest average depth of target and the most receiving yards per reception.
+However, they also had the fewest receptions, the highest drop percentage, and the fewest broken tackles.
 Thus, they may be considered the boom-or-bust Wide Receivers.
@@
 
+The goal of cluster analysis is to identify distinguishable subgroups of people.
+There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, hierarchical clustering (aka connectivity-based clustering), and others.
+The present chapter used model-based clustering to identify tiers of players based on projected points.
+Using various performance metrics of Wide Receivers, we identified three subtypes of Wide Receivers: 1) Elite WR1s who are strong possession receivers; 2) Consistent Contributor/WR2s; 3) deep threats/boom-or-bust receivers.
+The "Elite WR1s" tended to have the lowest drop percentage, the shortest average depth of target, the fewest receiving air yards per reception, the most receptions per season, and the most broken tackles.
+The "Consistent Contributor/WR2s" had fewer receptions and fewer broken tackles than the Elite WR1s; their average depth of target was longer than Elite WR1s, and they had more receiving air yards per reception than Elite WR1s.
+The "Deep Threat/Boom-or-Bust" receivers had the greatest average depth of target and the most receiving yards per reception; however, they also had the fewest receptions, the highest drop percentage, and the fewest broken tackles.
+In sum, cluster analysis can be a useful way of identifying subgroups of individuals who are more similar to one another on various characteristics.
+
 ::: {.content-visible when-format="html"}

### 2025-07-20 decision-making.qmd
20250720 - conclusion sections

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/projectionsWithActuals_weekly.RData", fsep = ""))
-#load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projections_seasonal.RData", fsep = ""))
-#load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projections_weekly.RData", fsep = ""))
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
@@
 load(file = "./data/projectionsWithActuals_weekly.RData")
-#load(file = "./data/players_projections_seasonal.RData")
-#load(file = "./data/players_projections_weekly.RData")
 load(file = "./data/player_stats_seasonal.RData")
@@
 Averaging projections from individuals tends to yield predictions that are more accurate than the accuracy of most forecasters in the group [@Mannes2014].
-Indeed, when at least some of of the projections bracket the truth, averaged predictions must be more accurate than the average individual forecaster—in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE)—and averaged predictions are often much more accurate [@Larrick2024].
+Indeed, when at least some of the projections bracket the truth, averaged predictions must be more accurate than the average individual forecaster—in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE)—and averaged predictions are often much more accurate [@Larrick2024].
 When referring to the accuracy of the "average individual forecaster", we are referring to accuracy in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE)—not to the accuracy of the forecaster at the 50th percentile.
@@
 
-The relatively modest accuracy of the projections by so-called fantasy "experts'" and of their average of their projections could occur for a number of reasons.
+The relatively modest accuracy of the projections by so-called fantasy "experts" and of their average of their projections could occur for a number of reasons.
 One possibility is that the level of expertise of the "expert" forecasters in terms of being able to provide accurate forecasts is not strong.
@@
 
-Proportion of players whose bracketing rate was zero:
+Below is the proportion of players whose bracketing rate was zero:
 
@@
 ```
+
+```{r}
+#| include: false
+
+proportionZeroBracketingRate_seasonal <- seasonalIndividualSources %>% 
+  filter(!is.na(bracketingRate)) %>% 
+  mutate(isZeroBracketingRate = ifelse(bracketingRate == 0, 1, 0)) %>% 
+  summarise(proportionZeroBracketingRate = sum(isZeroBracketingRate) / n())
+```
+
+A high proportion (`{r} petersenlab::apa(proportionZeroBracketingRate_seasonal * 100, decimals = 0)`%) of players' forecasts did not bracket the truth.
 
@@
 ```
+
+```{r}
+#| include: false
+
+proportionZeroBracketingRate_weekly <- weeklyIndividualSources %>% 
+  filter(!is.na(bracketingRate)) %>% 
+  mutate(isZeroBracketingRate = ifelse(bracketingRate == 0, 1, 0)) %>% 
+  summarise(proportionZeroBracketingRate = sum(isZeroBracketingRate) / n())
+```
+
+A high proportion (`{r} petersenlab::apa(proportionZeroBracketingRate_weekly * 100, decimals = 0)`%) of players' forecasts did not bracket the truth.
 
@@
 However, there are still unpredictable elements (e.g., injuries), and there is considerable luck involved.
-As a result, sports betting and gambling on fantasy football still involve components of [variable ratio reinforcement](#sec-reinforcementIntermittent) that lend them to being potentially highly addictive.
+As a result, sports betting and gambling on fantasy football still involve components of [variable ratio reinforcement](#sec-reinforcementIntermittent) that lend them to being potentially highly addictive despite people's tendency to lose money.
 
@@
 
+The wisdom of the crowd is the idea that the average of the predictions of many people is often more accurate than the prediction of individual experts.
+When at least some of the projections bracket—i.e., fall on opposite sides of—the truth, averaged predictions are more accurate than the average individual forecaster—in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE).
+Crowd-averaged projections tend to be most accurate when the crowd consists of individuals who hold expertise in the domain such that they will make predictions that fall close to the truth, there is relatively low variability in the expertise of the individual forecasters in terms of their ability to make accurate forecasts, there is cognitive diversity among the forecasters, the projections are made independently—i.e., the forecasters are not aware of others’ forecasts and do not discuss or interact with the other forecasters, the bracketing rate is high, and there are at least 5–10 sources of projections.
+In the case of fantasy football, however, a high proportion of players' forecasts do not bracket the truth, suggesting that the so-called experts are not accurate in predicting players' future fantasy performance.
+Nevertheless, going with the crowd-averaged projections tends to be more accurate than relying on an individual projection source.
+Performance in fantasy football is around half and half luck versus skill.
+Sports betting and gambling on fantasy football involve components of [variable ratio reinforcement](#sec-reinforcementIntermittent) that lend them to being potentially highly addictive.
+
 ::: {.content-visible when-format="html"}

### 2025-07-20 factor-analysis.qmd
20250717 - chapter overviews
20250720 - conclusion sections

@@
 > --- George Box [-@Box1979, p. 202]
+
+This chapter provides an overview of factor analysis, which encompasses a range of latent variable modeling approaches, including [exploratory](#sec-efa) and [confirmatory](#sec-cfa) factor analysis.
 
@@
 
+Factor analysis is a class of latent variable models that aims to identify the optimal, parsimonious latent structure for a group of variables.
+Latent variables are ways of studying and operationalizing theoretical constructs that cannot be directly observed or quantified.
+Factor analysis encompasses two general types: [confirmatory factor analysis](#sec-cfa) and [exploratory factor analysis](#sec-efa).
+[*Exploratory factor analysis*](#sec-efa) (EFA) is used when the researcher has no a priori hypotheses about how a set of variables is structured.
+[*Confirmatory factor analysis*](#sec-cfa) (CFA) is used when a researcher wants to evaluate how well a hypothesized model fits.
+A goal of factor analysis is to balance accuracy (i.e., variance accounted for) and parsimony (i.e., simplicity).
+Factor analysis estimates the latent factors as the common variance among the variables that load onto that factor and discards the remaining variance as "error".
+There are many [decisions to make in factor analysis](#sec-factorAnalysisDecisions).
+These decisions can have important impacts on the resulting solution.
+Thus, it can be helpful for theory and interpretability to help guide decision-making when conducting factor analysis.
+Using both [exploratory](#sec-efa) and [confirmatory](#sec-cfa) factor analysis, we were able to identify three latent factors that accounted for considerable variance in the variables we examined, pertaining to Quarterbacks: 1) usage; 2) aggressiveness; 3) performance.
+We were then able to determine which players were highest and lowest on each of these factors.
+
 ::: {.content-visible when-format="html"}

### 2025-07-20 machine-learning.qmd
20250720 - conclusion sections

@@
 
-The most accurate model, in terms of $R^2$, had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using [regression with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
+The most accurate model, in terms of [discrimination](#sec-discrimination) ($R^2$), had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using [regression with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
 The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
@@
 
-The most accurate model in terms of [mean absolute error](#sec-meanAbsoluteError) was the [tree-boosting model](#sec-treeBoosting), which had a [mean absolute error](#sec-meanAbsoluteError) of `{r} mae_minimum`.
+The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [tree-boosting model](#sec-treeBoosting), which had a [mean absolute error](#sec-meanAbsoluteError) of `{r} mae_minimum`.
 However, even in the most accurate model, the predicted values were over `{r} petersenlab::apa(mae_minimum, decimals = 0)` points away from the actual values, suggesting that the model-predicted values were not particularly accurate.
@@
 
+Machine learning attempts to maximize prediction accuracy.
+There are various types of machine learning models.
+[Supervised learning](#sec-machineLearningTypesSupervised) involves learning from data where the correct classification or outcome is known.
+[Unsupervised learning](#sec-machineLearningTypesUnsupervised) involves learning from data without known classifications.
+Machine learning can incorporate many predictor variables, so it is important to perform [cross-validation](#sec-crossValidation) to avoid [overfitting](#sec-overfitting).
+We examined multiple machine learning models to identify the model that generates the most accurate prediction of Quarterbacks' future fantasy points.
+All models explained less than half of the variance in players' fantasy points.
+The most accurate model, in terms of [discrimination](#sec-discrimination), was achieved using regression with one predictor: the player's fantasy points in the previous season.
+The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [tree-boosting model](#sec-treeBoosting), which accounted for the longitudinal nature of the data.
+The considerable unexplained variance suggests that we are missing key variables that are important predictors of future fantasy performance.
+
 ::: {.content-visible when-format="html"}

### 2025-07-20 modern-portfolio-theory.qmd
20250719 - citations
20250720 - conclusion sections

@@
 
-The tangency portfolio is the portfolio with the highest Sharpe ratio (i.e., the highest ratio of return to risk).
+The tangency portfolio is the portfolio with the highest Sharpe ratio—i.e., the highest ratio of return to risk.
 In other words, it is the portfolio with the greatest risk-adjusted returns.
@@
 You are most likely to pick the best players if you go with the [wisdom of the crowd](#sec-wisdomOfTheCrowd) (e.g., average projections across projection sources) and diversify.
+Players—like stocks—have expected returns and risk associated with them.
+You can think of constructing your team like constructing a stock portfolio, where your goal is to gain the maximum expected returns for a given level of risk.
+The greater risk you are willing to take, the greater the potential returns, but alse the lower the potential downside if the players underperform.
+Thus, you may want your team to be composed of a diverse combination of some higher risk players and some lower risk players.
+We demonstrate the [efficient frontier](#sec-efficientFrontier) as a valuable tool for identifying the combination of players with the maximum expected returns relative to their risk.
 Most projections are public information, so you might wonder whether using crowd projections gains you anything because everybody else has access to public information.

### 2025-07-20 pca.qmd
20250712 - tweaks
20250720 - conclusion sections

@@
 # Data Reduction: Principal Component Analysis {#sec-pca}
+
+This chapter provides an overview of principal component analysis as a useful technique for data reduction.
 
@@
 We generated the very simple structure (VSS) plots in Figures [-@fig-pcaVSSPlot1] and [-@fig-pcaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
-The optimal number of components based on VSS criterion is three components.
+The optimal number of components based on the VSS criterion is three components.
 
@@
 
+Principal component analysis (PCA) is a technique used for data reduction—reducing a large set of a variables down to a smaller set of components that capture most of the variance in the larger set.
+There are many [decisions to make in factor analysis](#sec-pcaDecisions).
+These decisions can have important impacts on the resulting solution.
+Thus, it can be helpful for theory and interpretability to help guide decision-making when conducting factor analysis.
+There are several differences between [factor analysis](#sec-factorAnalysis) and PCA.
+Unlike [factor analysis](#sec-factorAnalysis), which estimates the latent factors as the common variance among the variables that load onto that factor and discards the remaining variance as "error", PCA uses all variance of variables and assumes variables have no error.
+Thus, PCA does not account for measurement error.
+Using PCA, we were able to identify three PCA components that accounted for considerable variance in the variables we examined, pertaining to Quarterbacks: 1) usage; 2) aggressiveness; 3) performance.
+We were then able to determine which players were highest and lowest on each of these components.
+
 ::: {.content-visible when-format="html"}

### 2025-07-20 simulation.qmd
20250708 - track chunk execution times
20250717 - chapter overviews
20250720 - conclusion sections

@@
+{{< include _chunk-timing.qmd >}}
+
 # Simulation: Bootstrapping and the Monte Carlo Method {#sec-simulation}
+
+This chapter provides an overview of various approaches to simulation, including [bootstrapping](#sec-simulationProjectedPointsBootstrapping) and the [Monte Carlo](#sec-simulationOverviewMonteCarlo) method.
 
@@
 
+A simulation is an "imitative representation" of a phenomenon that could exist the real world.
+In statistics, simulations are computer-driven investigations to better understand a phenomenon by studying its behavior under different conditions.
+Statistical simulations can be conducted in various ways.
+Two common types of simulations are [bootstrapping](#sec-simulationOverviewBootstrapping) and [Monte Carlo](#sec-simulationOverviewMonteCarlo) simulation.
+[Bootstrapping](#sec-simulationOverviewBootstrapping) involves repeated resampling (with replacement) from observed data.
+[Monte Carlo](#sec-simulationOverviewMonteCarlo) simulation involves repeated random sampling from a known distribution.
+We demonstrated [bootstrapping](#sec-simulationOverviewBootstrapping) and [Monte Carlo](#sec-simulationOverviewMonteCarlo) approaches to simulating the most likely range of outcomes for a player in terms of fantasy points.
+
 ::: {.content-visible when-format="html"}

### 2025-07-20 sports-cognitive-psychology.qmd
20250717 - chapter overviews
20250720 - conclusion sections

@@
 # Sports and Cognitive Psychology {#sec-sportsCognitivePsychology}
+
+This chapter provides an overview of some of the roles of cognitive psychology and analytics in sports.
 
@@
 
+Football has recently seen greater use of analytics; however, its uptake has been slower compared to other sports like baseball and basketball.
+Examples of changes in football due to the use of analytics include [more often going for it on fourth down](#sec-coachingGoingForIt), a greater emphasis on the passing game, drafting Running Backs later in the draft (and, more generally, valuing Running Backs less), and trading down in the draft to obtain more low picks (rather than having fewer high picks) because top picks are frequently overvalued.
+Historically, despite the decision hurting their chances of winning, teams rarely went for it on fourth down and elected to punt or kick a field goal instead, likely due to [loss aversion bias](#sec-cognitiveBiasesLossAversion).
+[Loss aversion bias](#sec-cognitiveBiasesLossAversion) may lead coaches to make more conservative play-calling, in general, such as running the ball when passing would tend to be more advantageous.
+Rates of going for it on fourth down in the NFL were low until 2017.
+After the 2017 season, rates of going for it on fourth down increased dramatically.
+In addition to poor decisions relating to going for it on fourth down, there is not strong evidence that [icing the kicker](#sec-icingTheKicker) works, suggesting that it may better suit your team to use timeouts in other situations.
+
+There are many [psychological factors that influence sports performance](#sec-psychologicalFactorsPerformance).
+According to the Yerkes–Dodson law that describes the association between arousal and performance.
+For simple tasks, performance tends to increase with arousal.
+There is an optimal range of arousal for complex and challenging tasks, such as those in sports, which often require divided attention and multitasking.
+Too little arousal is bad because it leads to insufficient attention on the task at hand; however, too much arousal is also bad because anxiety can impair performance through its negative effects on attention (e.g., "tunnel vision"), memory, decision making, problem-solving, and multitasking.
+When a task is highly practiced or automatic, excessive attention can hurt performance.
+By contrast, for tasks that require conscious control, decision-making, and real-time adjustments, more focused attention tends to lead to better performance.
+
 ::: {.content-visible when-format="html"}

### 2025-07-20 time-series-analysis.qmd
20250708 - track chunk execution times
20250720 - conclusion sections

@@
+{{< include _chunk-timing.qmd >}}
+
 # Time Series Analysis {#sec-timeSeries}
+
+This chapter provides an overview of time series analysis.
 
@@
 
+That is, time series analysis seeks to evaluate change over time to predict future values.
+There many different types of time series analyses.
+We demonstrated use of [autoregressive integrated moving average](#sec-timeSeriesARIMA) (ARIMA) models to predict future fantasy points.
+ARIMA models aim to describe how how earlier levels of a variable are correlated with later levels of the same variable.
+ARIMA models perform best when there is a clear pattern where later values are influenced by earlier values.
+ARIMA models assume that the data are stationary (i.e., there are no long-term trends), are non-seasonal (i.e., there is no consistency of the timing of the peaks or troughs in the line), and that earlier values influence later values.
+This may not strongly be the case in fantasy football, so ARIMA models may not be particularly useful in forecasting fantasy football performance.
+We also used [Bayesian mixed models](#sec-timeSeriesBayesian) to generate forecasts of future performance and plots of individuals model-implied performance by age and position.
+
 ::: {.content-visible when-format="html"}

### 2025-07-21 decision-making.qmd
20250720 - fix

@@
   mutate(isZeroBracketingRate = ifelse(bracketingRate == 0, 1, 0)) %>% 
-  summarise(proportionZeroBracketingRate = sum(isZeroBracketingRate) / n())
+  summarise(proportionZeroBracketingRate = sum(isZeroBracketingRate) / n()) %>% 
+  as.numeric()
 ```
@@
 A key goal is to identify decisions that provide a net advantage in expected value.
-Although there is considerable variability and the outcome is not guaranteed, small advantages can yield considerable gains in the long run @Tanney2021; however, one must be able to survive the costs of losses to obtain enough long-term trials.
+Although there is considerable variability and the outcome is not guaranteed, small advantages can yield considerable gains in the long run [@Tanney2021]; however, one must be able to survive the costs of losses to obtain enough long-term trials.
 So, do not bet more than you are willing to lose.

### 2025-07-21 decision-making.qmd
20250720 - fix

@@
   mutate(isZeroBracketingRate = ifelse(bracketingRate == 0, 1, 0)) %>% 
-  summarise(proportionZeroBracketingRate = sum(isZeroBracketingRate) / n()) %>% 
-  as.numeric()
+  summarise(proportionZeroBracketingRate = sum(isZeroBracketingRate) / n())
 ```

### 2025-07-21 decision-making.qmd
20250721 - edits

@@
 When referring to the accuracy of the "average individual forecaster", we are referring to accuracy in terms of [mean absolute error](#sec-meanAbsoluteError) (MAE)—not to the accuracy of the forecaster at the 50th percentile.
-If none of the projections bracket the truth—e.g., all projections overestimate the truth—averaged predictions will be as accurate than the average individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError) [@Larrick2024].
+If none of the projections bracket the truth—e.g., all projections overestimate the truth—averaged predictions will be as accurate as the average individual forecaster in terms of [mean absolute error](#sec-meanAbsoluteError) [@Larrick2024].
 In sum, "Averaging the answers of a crowd, therefore, ensures a level of accuracy no worse than the average member of the crowd and, in some cases, a level better than nearly all members" [@Larrick2024, p.
@@
 
-There is substantial [overconfidence](#sec-cognitiveBiasesOverconfidence) (in particular overestimation) in sports betting/gambling.
+There is substantial [overconfidence](#sec-cognitiveBiasesOverconfidence) (in particular overestimation of one's actual performance) in sports betting/gambling.
 [A study](https://mattbrownecon.github.io/assets/papers/jmp/sportsbetting.pdf) (archived at <https://perma.cc/X2AW-SUBZ>) of frequent sports bettors found that they tended to predict that they would gain 0.3 cents for every dollar wagered, but in fact lost 7.5 cents for every dollar wagered [@Brown2025].

### 2025-07-21 getting-started.qmd
20250721 - if...else conditions

@@
 We can use the construction, `if()...else if()...else()` if we want to perform conditional operations.
+The typical construction of `if()...else if()...else()` operates such that it first checks if the first `if()` condition is true.
+If the first `if()` condition is true, it performs the operation specified and terminates the process.
+If the first `if()` condition is not true, it checks the `else if()` conditions in order until one of them is true.
+For the first true `else if()` condition, it performs the operation specified and terminates the process.
+If none of the `else if()` conditions is true, it performs the operation specified under `else()` and then terminates the process.
 The construction, `if()...else if()...else()` can only be used on one value at a time.

### 2025-07-21 download-football-data.qmd
20250721 - tweaks

@@
 
-@HoCombine provide a Data Dictionary for data from the combine at the following link: <https://nflreadr.nflverse.com/articles/dictionary_combine.html>
+@HoCombine provide a Data Dictionary for data from the NFL Combine at the following link: <https://nflreadr.nflverse.com/articles/dictionary_combine.html>
 

### 2025-07-21 getting-started.qmd
20250721 - tweaks

@@
 ```{r}
-convert_temperature <- function(temperature, unit = "F"){
+convert_temperature <- function(temperature, unit = "F"){ # specify the arguments and any defaults
   if(unit == "F"){ # if the input temperature(s) in Fahrenheit
@@
 
-  return(newtemp)
+  return(newtemp) # this is what is returned by the function
 }

### 2025-07-21 decision-making.qmd
20250721 - Vegas lines and the betting market; delay discounting curves

@@
 Aggregation of predictions from multiple people leverages several important features, including cognitive diversity and error cancellation.
-
 Cognitive diversity refers to the representation of individuals with different perspectives because of their "differences in knowledge, training, experience, or thinking styles" [@Rader2017, p. 8].
@@
 
-[overconfidence; confirmation bias; variable ratio reinforcement; addiction/delay discounting; wisdom of the crowd, and the challenge of beating "the market"]
-
 The National Football League (NFL) used to be adamantly against sports gambling.
@@
 
+People may make a sports bet because they believe strongly in their predicted outcome and that they "cannot be wrong".
+However, the initial Vegas lines represent the aggregation of a massive amount of information (by professional oddsmakers), including team power rankings, injuries, weather, home/away, coaching, matchups, historical betting behavior, advanced statistics, and proprietary algorithms.
+Leading up to the game, the Vegas lines are adjusted to balance the total money on each side of the bet (so that the sportsbook can be ensured it makes money).
+In this way, the Vegas lines are based on the bets of many, many people and many dollars and reflect market wisdom.
+That is, by making a sports bet, you are implicitly claiming that you have information, insight, or a model that makes you better than the market consensus at predicting the outcome—even though half of the dollars bet will lose.
+Betting markets, similar to the stock market (as described in @sec-wisdomOfTheCrowd), are often highly efficient and beating them consistently is extremely challenging.
+Moreover, people tend to show [confirmation bias](#sec-cognitiveBiasesConfirmation), such that they tend to remember their predictive successes and to forget their failures.
+So, when they they get a bet correct, they may be more likely to bet again in the future, especially because sports betting involves [intermittent reinforcement](#sec-reinforcementIntermittent) (in particular, variable ratio reinforcement), which can make it highly addictive.
+
+However, not all people are equally prone to gambling addiction.
+One of the most robust risk factors of gambling addiction is a steep delay discounting curve [@Amlung2017; @Weinsztok2021].
+The delay discounting curve can be generated from asking respondents a series of hypothetical choices, such as:
+1) Would you prefer $10 now or $15 in 1 hour?
+2) Would you prefer $10 now or $20 tomorrow?
+3) Would you prefer $10 now or $50 in one month?
+Based on these questions, one can estimate how much a respondent's valuation of a reward (in this case, money) decreases with the passage of time.
+An example of delay discounting curves of two people is in @fig-delayDiscountingCurves.
+Some individuals have a shallow delay discounting curve and will prefer more money even if they have to wait, a form of delayed gratification.
+Other individuals have a steep delay discounting curve and will prefer obtaining the money now, even if that means gaining less money in the long run, a form of impulsivity.
+In particular, individuals who have a steep delay discounting curve are most likely to develop gambling addictions, possibly because they are more sensitive to immediate rewards and are less driven by long-term consequences.
+
+::: {#fig-delayDiscountingCurves}
+![](images/delayDiscountingCurves.jpg){width=100% fig-alt="Prototypic hyperbolic delayed reward discounting curves reflecting the discounted subjective value of $100 delayed from 1 day to 1 year The curves reflect the points at which the smaller immediate reward is equal in value to the $100 delayed reward.
+For example, at a delay of 100 days, $100 has lost ∼50% of its nominal value for the low impulsivity profile and ∼90% of its nominal value for the high impulsivity profile.
+(Figure and caption adapted from Gray, J.
+C., & MacKillop, J.
+(2015).
+Impulsive delayed reward discounting as a genetically-influenced target for drug abuse prevention: A critical evaluation.
+*Frontiers in Psychology*, 6.
+https://doi.org/10.3389/fpsyg.2015.01104)"}
+
+Prototypic hyperbolic delayed reward discounting curves reflecting the discounted subjective value of $100 delayed from 1 day to 1 year The curves reflect the points at which the smaller immediate reward is equal in value to the $100 delayed reward.
+For example, at a delay of 100 days, $100 has lost ∼50% of its nominal value for the low impulsivity profile and ∼90% of its nominal value for the high impulsivity profile.
+(Figure and caption adapted from @Gray2015; Gray, J.
+C., & MacKillop, J.
+(2015).
+Impulsive delayed reward discounting as a genetically-influenced target for drug abuse prevention: A critical evaluation.
+*Frontiers in Psychology*, 6.
+[https://doi.org/10.3389/fpsyg.2015.01104](https://doi.org/10.3389/fpsyg.2015.01104)).
+:::
+
 ### Reinforcement {#sec-reinforcement}
@@
 There are various approaches to intermittent reinforcement, called schedules of reinforcement.
-The four primary reinforcement schedules are: fixed interval, fixed ratio, variable interval, adn variable ratio.
+The four primary reinforcement schedules are: fixed interval, fixed ratio, variable interval, and variable ratio.
 The reinforcement schedules are depicted in @fig-reinforcementSchedules, as adapted from @Spielman2020.

### 2025-07-21 evaluating-prediction-accuracy.qmd
20250721 - REFs

@@
 ::: {#fig-confusionMatrix}
-![](images/confusionMatrix.png){fig-alt="A Two-by-Two Confusion Matrix."}
-
-A Two-by-Two Confusion Matrix.
+![](images/confusionMatrix.png){fig-alt="A Two-by-Two Confusion Matrix. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+A Two-by-Two Confusion Matrix. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-confusionMatrix1}
-![](images/confusionMatrix1.png){fig-alt="Example of a Two-by-Two Confusion Matrix. TP = true positives; TN = true negatives; FP = false positives; FN = false negatives."}
-
-Example of a Two-by-Two Confusion Matrix. TP = true positives; TN = true negatives; FP = false positives; FN = false negatives.
+![](images/confusionMatrix1.png){fig-alt="Example of a Two-by-Two Confusion Matrix.
+TP = true positives; TN = true negatives; FP = false positives; FN = false negatives.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Two-by-Two Confusion Matrix. TP = true positives; TN = true negatives; FP = false positives; FN = false negatives. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-confusionMatrix2}
-![](images/confusionMatrix2.png){fig-alt="Example of a Two-by-Two Confusion Matrix With Marginal Sums. TP = true positives; TN = true negatives; FP = false positives; FN = false negatives."}
-
-Example of a Two-by-Two Confusion Matrix With Marginal Sums. TP = true positives; TN = true negatives; FP = false positives; FN = false negatives.
+![](images/confusionMatrix2.png){fig-alt="Example of a Two-by-Two Confusion Matrix With Marginal Sums.
+TP = true positives; TN = true negatives; FP = false positives; FN = false negatives.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Two-by-Two Confusion Matrix With Marginal Sums.
+TP = true positives; TN = true negatives; FP = false positives; FN = false negatives.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ![](images/confusionMatrix3.png){fig-alt="Example of a Two-by-Two Confusion Matrix With Marginal Sums and Marginal Probabilities.
-TP = true positives; TN = true negatives; FP = false positives; FN = false negatives; BR = base rate; SR = selection ratio."}
+TP = true positives; TN = true negatives; FP = false positives; FN = false negatives; BR = base rate; SR = selection ratio.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
@@
 TP = true positives; TN = true negatives; FP = false positives; FN = false negatives; BR = base rate; SR = selection ratio.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-confusionMatrix4}
-![](images/confusionMatrix4.png){fig-alt="Chance Expectancies in Two-by-Two Confusion Matrix. BR = base rate; SR = selection ratio."}
-
-Chance Expectancies in Two-by-Two Confusion Matrix. BR = base rate; SR = selection ratio.
+![](images/confusionMatrix4.png){fig-alt="Chance Expectancies in Two-by-Two Confusion Matrix. BR = base rate; SR = selection ratio. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Chance Expectancies in Two-by-Two Confusion Matrix. BR = base rate; SR = selection ratio. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::

### 2025-07-21 factor-analysis.qmd
20250721 - REFs

@@
 ::: {#fig-simpleStructure}
-![](images/simpleStructure.png){fig-alt="Example of a Factor Matrix That Follows Simple Structure."}
-
-Example of a Factor Matrix That Follows Simple Structure.
+![](images/simpleStructure.png){fig-alt="Example of a Factor Matrix That Follows Simple Structure. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Factor Matrix That Follows Simple Structure. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ![](images/factorSolutionSimpleStructure.png){fig-alt="Example of a Factor Analysis Model That Follows Simple Structure.
-'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems."}
-
-Example of a Factor Analysis Model That Follows Simple Structure. 'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
+'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Factor Analysis Model That Follows Simple Structure.
+'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ![](images/factorSolutionUnrotatedExample.png){fig-alt="Example of a Factor Analysis Model That Does Not Follow Simple Structure.
-'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems."}
-
-Example of a Factor Analysis Model That Does Not Follow Simple Structure. 'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
+'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Factor Analysis Model That Does Not Follow Simple Structure.
+'INT' = internalizing problems; 'EXT' = externalizing problems; 'TD' = thought-disordered problems.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-factorMatrix}
-![](images/factorMatrix.png){fig-alt="Example of a Factor Matrix."}
-
-Example of a Factor Matrix.
+![](images/factorMatrix.png){fig-alt="Example of a Factor Matrix. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Factor Matrix. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-factorSolutionUnrotated}
-![](images/factorSolutionUnrotated.png){fig-alt="Example of an Unrotated Factor Solution."}
-
-Example of an Unrotated Factor Solution.
+![](images/factorSolutionUnrotated.png){fig-alt="Example of an Unrotated Factor Solution. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of an Unrotated Factor Solution. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-factorMatrixRotated}
-![](images/factorMatrixRotated.png){fig-alt="Example of a Rotated Factor Matrix."}
-
-Example of a Rotated Factor Matrix.
+![](images/factorMatrixRotated.png){fig-alt="Example of a Rotated Factor Matrix. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Rotated Factor Matrix. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-factorSolutionRotated}
-![](images/factorSolutionRotated.png){fig-alt="Example of a Rotated Factor Solution."}
-
-Example of a Rotated Factor Solution.
+![](images/factorSolutionRotated.png){fig-alt="Example of a Rotated Factor Solution. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Rotated Factor Solution. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-orthogonalRotation}
-![](images/FactorAnalysis-08.png){fig-alt="Example of a Factor Structure From an Orthogonal Rotation."}
-
-Example of a Factor Structure From an Orthogonal Rotation.
+![](images/FactorAnalysis-08.png){fig-alt="Example of a Factor Structure From an Orthogonal Rotation. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Factor Structure From an Orthogonal Rotation. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-obliqueRotation}
-![](images/FactorAnalysis-09.png){fig-alt="Example of a Factor Structure From an Oblique Rotation."}
-
-Example of a Factor Structure From an Oblique Rotation.
+![](images/FactorAnalysis-09.png){fig-alt="Example of a Factor Structure From an Oblique Rotation. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Factor Structure From an Oblique Rotation. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-factorRotation}
-![](images/factorRotation.png){fig-alt="Example of a Factor Rotation of Neuroticism and Extraversion."}
-
-Example of a Factor Rotation of Neuroticism and Extraversion.
+![](images/factorRotation.png){fig-alt="Example of a Factor Rotation of Neuroticism and Extraversion. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
+
+Example of a Factor Rotation of Neuroticism and Extraversion. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::

### 2025-07-21 pca.qmd
20250721 - REFs

@@
 ::: {#fig-correlationMatrix2}
-![](images/correlationMatrix2.png){fig-alt="Example Correlation Matrix 2."}
+![](images/correlationMatrix2.png){fig-alt="Example Correlation Matrix 2. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
-Example Correlation Matrix 2.
+Example Correlation Matrix 2. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-pcaVsFactorAnalysis}
-![](images/pcaVsFactorAnalysis.png){fig-alt="Distinction Between Factor Analysis and Principal Component Analysis."}
+![](images/pcaVsFactorAnalysis.png){fig-alt="Distinction Between Factor Analysis and Principal Component Analysis. From @PetersenPrinciplesPsychAssessment."}
 
-Distinction Between Factor Analysis and Principal Component Analysis.
+Distinction Between Factor Analysis and Principal Component Analysis. From @PetersenPrinciplesPsychAssessment.
 :::

### 2025-07-21 research-methods.qmd
20250708 - track chunk execution times
20250717 - Conclusion sections
20250721 - REFs

@@
+{{< include _chunk-timing.qmd >}}
+
 # Research Methods {#sec-researchMethods}
+
+This chapter provides an overview of key concepts in research methods.
+These considerations are crucial for designing studies and for interpreting findings from research studies.
 
@@
 ::: {#fig-correlationAndCausation1}
-![](images/correlationAndCausation_1.png){fig-alt="Hypothesized Causal Effect Based on an Observed Association Between `X` and `Y`, Such That `X` Causes `Y`."}
+![](images/correlationAndCausation_1.png){fig-alt="Hypothesized Causal Effect Based on an Observed Association Between `X` and `Y`, Such That `X` Causes `Y`.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
-Hypothesized Causal Effect Based on an Observed Association Between `X` and `Y`, Such That `X` Causes `Y`.
+Hypothesized Causal Effect Based on an Observed Association Between `X` and `Y`, Such That `X` Causes `Y`. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-correlationAndCausation2}
-![](images/correlationAndCausation_2.png){fig-alt="Reverse (Opposite) Direction of Effect From the Hypothesized Effect, Where `Y` Causes `X`."}
+![](images/correlationAndCausation_2.png){fig-alt="Reverse (Opposite) Direction of Effect From the Hypothesized Effect, Where `Y` Causes `X`.
+From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
-Reverse (Opposite) Direction of Effect From the Hypothesized Effect, Where `Y` Causes `X`.
+Reverse (Opposite) Direction of Effect From the Hypothesized Effect, Where `Y` Causes `X`. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-correlationAndCausation3}
-![](images/correlationAndCausation_3.png){fig-alt="Bidirectional Effect Between `X` and `Y`, such that `X` causes `Y` and `Y` causes `X`."}
+![](images/correlationAndCausation_3.png){fig-alt="Bidirectional Effect Between `X` and `Y`, such that `X` causes `Y` and `Y` causes `X`. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
-Bidirectional Effect Between `X` and `Y`, such that `X` causes `Y` and `Y` causes `X`.
+Bidirectional Effect Between `X` and `Y`, such that `X` causes `Y` and `Y` causes `X`. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-correlationAndCausation4}
-![](images/correlationAndCausation_4.png){fig-alt="Confounded Association Between `X` and `Y` due to a Common Cause, `Z`."}
+![](images/correlationAndCausation_4.png){fig-alt="Confounded Association Between `X` and `Y` due to a Common Cause, `Z`. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
-Confounded Association Between `X` and `Y` due to a Common Cause, `Z`.
+Confounded Association Between `X` and `Y` due to a Common Cause, `Z`. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::
@@
 ::: {#fig-reliabilityVsValidity}
-![](images/reliabilityVsValidity.png){fig-alt="Reliability Versus Validity."}
+![](images/reliabilityVsValidity.png){fig-alt="Reliability Versus Validity. From @Petersen2024a and @PetersenPrinciplesPsychAssessment."}
 
-Reliability Versus Validity.
+Reliability Versus Validity. From @Petersen2024a and @PetersenPrinciplesPsychAssessment.
 :::

### 2025-07-21 getting-started.qmd
20250721 - tweaks

@@
 You can specify a function and its arguments either by providing values for each argument in the order indicated by the function, or by naming its arguments.
+Naming arguments explicitly (rather than merely relying on order) is considered best practice because it is safer—doing so prevents you from accidentally assigning an input to the wrong argument.
 
@@
 A vector is a series of elements that can be numeric or character.
-It has one dimension (length).
-To create a vector, use the `c()` function to combine elements into a vector.
+Character elements should be specified in quotes.
+A vector has one dimension (length).
+To create a vector, use the `c()` function to combine elements into a vector ("c" stands for combine).
 And, we use the assignment operator (`<-`) to assign the vector to an object named `exampleVector`, so we can access it later.
@@
 exampleVector <- c(40, 30, 24, 20, 18, 23, 27, 32, 26, 23, NA, 37)
+
+exampleVector2 <- c(
+  "A","B1","B2","This is a sentence.","This is another sentence.")
 ```
@@
 
+The direction of the slashes matters—you should use forward slashes (not backslashes)!
 To determine where you working directory is, you can type:
@@
 
+The direction of the slashes matters—you should use forward slashes (not backslashes)!
 The following code saves a file to an *absolute path*:
@@
 
+Logical Operators evaluate one or more elements for a condition, and return `TRUE`, `FALSE`, or if the element was missing, `NA`.
+
 ### Is Equal To: `==` {#sec-logicalOperatorEqual}
@@
 
+The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] has a convenience function, `%ni%`, for the logical operator of determining whether an element is *not* in a value of another vector:
+
+```{r}
+players$position %ni% c("RB","WR")
+```
+
 ### Is Missing: `is.na()` {#sec-logicalOperatorMissing}
@@
 
+The "and" operator (`&`) is used to string together multiple logical operators that all must evaluate as `TRUE` in order for the combined evaluation to evaluate as `TRUE`; otherwise, the combined evaluation evaluates as `FALSE`.
+
 ```{r}
@@
 
+The "or" operator (`|`) is used to string together multiple logical operators, any of which must evaluate as `TRUE` in order for the combined evaluation to evaluate as `TRUE`; otherwise, the combined evaluation evaluates as `FALSE`.
+
 ```{r}
 players$position == "WR" | players$age > 23
+```
+
+### which() {#sec-logicalOperatorWhich}
+
+The operator, `which()`, can be used to obtain the TRUE indices of an object, which is useful for [subsetting](#sec-subset).
+
+```{r}
+which(players$position == "RB")
 ```
@@
 If the first `if()` condition is not true, it checks the `else if()` conditions in order until one of them is true.
+There can be multiple `else if()` conditions.
 For the first true `else if()` condition, it performs the operation specified and terminates the process.
@@
 Now, let's put together what we have learned to create a useful function.
-Functions are useful if you want to perform an operation multiple times.
+Functions are useful if you want to perform an operation multiple times, especially if you want to be able to manipulate various settings to make slight modifications each time.
 Any operation that you want to perform multiple times, you can create a function to accomplish.
@@
 In this case, our function has two arguments: `temperature` (in degrees) and `unit` of the original temperature (`F` for Fahrenheit or `C` for Celsius, where the default unit is Fahrenheit).
+That allows us to make the slight modification of which unit is the input temperature and to change the calculation accordingly.
 

### 2025-07-22 modern-portfolio-theory.qmd
20250721 - links

@@
 Here, *risk* refers to the [variability](#sec-descriptiveStatisticsSpread) (e.g., standard deviation or variance) of returns across time.
-Given two portfolios with the same expected returns over time, people will prefer the "safer" portfolio—that is, the portfolio with less [variability](#sec-descriptiveStatisticsSpread)/volatility across time.
+Given two portfolios with the same expected returns over time, people tend to prefer the "safer" portfolio—that is, the portfolio with less [variability](#sec-descriptiveStatisticsSpread)/volatility across time.
 One of the powerful notions of modern portfolio theory is that, through diversification, one can achieve lower risk with the same expected returns.
@@
 By owning different asset classes, you can limit your downside risk without sacrificing much in terms of expected return.
+In sum, the goal of diversification is to reduce risk (not to increase returns) and thus to increase risk-adjusted returns, by reducing the amount of risk needed for a given level of return.
 
@@
 As with stocks, each player has an expected return (i.e., projection) and a degree of risk.
-In fantasy football, a player's risk might be quantified in terms of the [variability](#sec-descriptiveStatisticsSpread) of projected scores for a player across projection sources (e.g., Projection Source A, Source B, Souce C, etc.), or as historical game-to-game [variability](#sec-descriptiveStatisticsSpread).
+In fantasy football, a player's risk might be quantified in terms of the [variability](#sec-descriptiveStatisticsSpread) of projected scores for a player across projection sources (e.g., Projection Source A, Source B, Source C, etc.), or as historical game-to-game [variability](#sec-descriptiveStatisticsSpread).
 Variability of projected scores for a player across projection sources could reflect the *uncertainty* of projections for a player.
@@
 For instance, the performance of the Tight End and Running Back on the same team tends to be slightly negatively correlated, so it might not be a bad idea to start the Tight End and Running Back from the same team.
-For a correlation matrix of all positions on the team, see: <https://assets-global.website-files.com/5f1af76ed86d6771ad48324b/607a4434a565aa7763bd1312_AndyAsh-Sharpstack-RPpaper.pdf> [@Sherman2021; archived at <https://perma.cc/JQ6G-KSRT>].
+For a correlation matrix of all positions on the team, see: <https://assets-global.website-files.com/5f1af76ed86d6771ad48324b/607a4434a565aa7763bd1312_AndyAsh-Sharpstack-RPpaper.pdf> [@Sherman2021; archived at <https://perma.cc/JQ6G-KSRT>] and <https://www.4for4.com/2018/preseason/definitive-guide-stacking-draftkings> [@4for4Staff2018; archived at <https://perma.cc/JZH3-ZM5V>].
 

### 2025-07-22 getting-started.qmd
20250721 - stacking correlation matrix

@@
      "lavaangui","mice","miceadds","interactions","robustbase","ordinal","MASS",
-     "data.table","future","future.apply","SimDesign","domir"))
+     "data.table","future","future.apply","SimDesign","domir","GGally"))
    ```

### 2025-07-22 index.qmd
20250721 - stacking correlation matrix

@@
   "robustbase", "ordinal", "MASS", "data.table", "future", "future.apply", 
-  "SimDesign", "domir",
+  "SimDesign", "domir", "GGally",
   "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem

### 2025-07-22 modern-portfolio-theory.qmd
20250721 - stacking correlation matrix

@@
 library("NMOF")
+library("GGally")
 library("tidyverse")
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projectedPoints_seasonal.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_actualFantasyPoints_weekly.RData", fsep = ""))
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_playerIDs.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_schedules.RData", fsep = ""))
 ```
@@
 load(file = "./data/players_projectedPoints_seasonal.RData")
+load(file = "./data/player_stats_seasonal.RData")
 load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
 load(file = "./data/nfl_playerIDs.RData")
+load(file = "./data/nfl_schedules.RData")
 ```
@@
 For a correlation matrix of all positions on the team, see: <https://assets-global.website-files.com/5f1af76ed86d6771ad48324b/607a4434a565aa7763bd1312_AndyAsh-Sharpstack-RPpaper.pdf> [@Sherman2021; archived at <https://perma.cc/JQ6G-KSRT>] and <https://www.4for4.com/2018/preseason/definitive-guide-stacking-draftkings> [@4for4Staff2018; archived at <https://perma.cc/JZH3-ZM5V>].
+We generate a correlation matrix of positions in @sec-stacking.
 
@@
 That is, receiving higher returns generally requires taking on additional risk—at least as long as we stay along the [efficient frontier](#sec-efficientFrontier), described next.
+
+### Stacking {#sec-stacking}
+
+Stacking involves selecting players on the same team (e.g., multiple players on the Dallas Cowboys).
+To help inform the process of stacking, we evaluate the inter-correlation of performance in a given week across players on the same team and on opposing teams.
+We determine QB1, RB1, RB2, WR1, WR2, WR3, TE1, and K1 for each combination of team and season based on how many fantasy points a player scored over the entire season.
+
+```{r}
+# Identify the QB1, RB1, RB2, WR1, WR2, WR3, TE1, and K1 for each team-season combination
+qb1ByTeamSeasonPosition <- player_stats_seasonal %>% 
+  filter(!is.na(player_id) & !is.na(team) & !is.na(season) & !is.na(position_group) & !is.na(fantasyPoints)) %>% 
+  filter(position_group %in% c("QB")) %>% 
+  group_by(team, season, position_group) %>% 
+  arrange(-fantasyPoints) %>% 
+  slice_max(
+    fantasyPoints,
+    with_ties = FALSE) %>%
+  ungroup() %>% 
+  arrange(season, team, position_group) %>% 
+  select(team, season, position_group, player_id) %>% 
+  mutate(position = "QB1")
+
+rbsByTeamSeasonPosition <- player_stats_seasonal %>% 
+  filter(!is.na(player_id) & !is.na(team) & !is.na(season) & !is.na(position_group) & !is.na(fantasyPoints)) %>% 
+  filter(position_group %in% c("RB")) %>% 
+  group_by(team, season, position_group) %>% 
+  arrange(-fantasyPoints) %>% 
+  slice_max(
+    fantasyPoints,
+    n = 2,
+    with_ties = FALSE) %>% 
+  mutate(position = paste0("RB", row_number())) %>% 
+  ungroup() %>% 
+  arrange(season, team, position_group) %>% 
+  select(team, season, position_group, player_id, position)
+
+wrsByTeamSeasonPosition <- player_stats_seasonal %>% 
+  filter(!is.na(player_id) & !is.na(team) & !is.na(season) & !is.na(position_group) & !is.na(fantasyPoints)) %>% 
+  filter(position_group %in% c("WR")) %>% 
+  group_by(team, season, position_group) %>% 
+  arrange(-fantasyPoints) %>% 
+  slice_max(
+    fantasyPoints,
+    n = 3,
+    with_ties = FALSE) %>% 
+  mutate(position = paste0("WR", row_number())) %>% 
+  ungroup() %>% 
+  arrange(season, team, position_group) %>% 
+  select(team, season, position_group, player_id, position)
+
+te1ByTeamSeasonPosition <- player_stats_seasonal %>% 
+  filter(!is.na(player_id) & !is.na(team) & !is.na(season) & !is.na(position_group) & !is.na(fantasyPoints)) %>% 
+  filter(position_group %in% c("TE")) %>% 
+  group_by(team, season, position_group) %>% 
+  arrange(-fantasyPoints) %>% 
+  slice_max(
+    fantasyPoints,
+    with_ties = FALSE) %>%
+  ungroup() %>% 
+  arrange(season, team, position_group) %>% 
+  select(team, season, position_group, player_id) %>% 
+  mutate(position = "TE1")
+
+k1ByTeamSeasonPosition <- player_stats_seasonal %>% 
+  filter(!is.na(player_id) & !is.na(team) & !is.na(season) & !is.na(position_group) & !is.na(fantasyPoints)) %>% 
+  filter(position_group %in% c("K")) %>% 
+  group_by(team, season, position_group) %>% 
+  arrange(-fantasyPoints) %>% 
+  slice_max(
+    fantasyPoints,
+    with_ties = FALSE) %>%
+  ungroup() %>% 
+  arrange(season, team, position_group) %>% 
+  select(team, season, position_group, player_id) %>% 
+  mutate(position = "K1")
+
+topRankedPlayersByTeamSeasonPosition <- dplyr::bind_rows(
+  qb1ByTeamSeasonPosition,
+  rbsByTeamSeasonPosition,
+  wrsByTeamSeasonPosition,
+  te1ByTeamSeasonPosition,
+  k1ByTeamSeasonPosition
+)
+
+fantasyPointsForTopRankedPlayers_weekly <- dplyr::left_join(
+  topRankedPlayersByTeamSeasonPosition,
+  player_stats_weekly %>% select(team, season, week, player_id, fantasyPoints, game_id),
+  by = c("team","season","player_id")
+)
+
+# add game_id to dst data
+
+# first attempt: match on home_team
+dst_with_game_home <- nfl_actualFantasyPoints_dst_weekly %>%
+  left_join(
+    nfl_schedules %>% select(game_id, season, week, home_team),
+    by = c("season", "week", "team" = "home_team")
+  )
+
+# second attempt: match on away_team
+dst_with_game_away <- nfl_actualFantasyPoints_dst_weekly %>%
+  left_join(
+    nfl_schedules %>% select(game_id, season, week, away_team),
+    by = c("season", "week", "team" = "away_team")
+  )
+
+# Combine the two (one will have NA for game_id, the other will have the match)
+dst_with_game_id <- dplyr::bind_rows(
+  dst_with_game_home,
+  dst_with_game_away) %>%
+  filter(!is.na(game_id)) # keep only matched rows
+
+dst_with_game_id_subset <- dst_with_game_id %>% 
+  select(team, season, week, fantasyPoints, game_id) %>%
+  mutate(
+    position = "DST",
+    position_group = "DST")
+
+fantasyPointsSameTeam_weekly <- dplyr::bind_rows(
+  fantasyPointsForTopRankedPlayers_weekly,
+  dst_with_game_id_subset
+) %>% 
+  select(-player_id) %>% 
+  filter(!is.na(game_id))
+
+fantasyPointsOpponent_weekly <- fantasyPointsSameTeam_weekly %>%
+  rename(
+    opp_team = team,
+    opp_fantasyPoints = fantasyPoints
+  )
+
+fantasyPointsCombined_weekly <- fantasyPointsSameTeam_weekly %>%
+  inner_join(
+    fantasyPointsOpponent_weekly %>% select(game_id, opp_team, position, opp_fantasyPoints),
+    by = c("game_id", "position")
+  ) %>%
+  filter(team != opp_team) # exclude matching to own team
+
+# Pivot to wide format for team positions
+team_points_wide <- fantasyPointsCombined_weekly %>%
+  select(team, season, week, position, fantasyPoints) %>%
+  tidyr::pivot_wider(
+    names_from = position,
+    values_from = fantasyPoints)
+
+# Pivot to wide format for opponent positions
+opp_points_wide <- fantasyPointsCombined_weekly %>%
+  select(team, season, week, position, opp_fantasyPoints) %>%
+  tidyr::pivot_wider(
+    names_from = position,
+    values_from = opp_fantasyPoints,
+    names_prefix = "opp_")
+
+# Join both so each row has team and opponent positions side by side
+combined_points_wide <- dplyr::left_join(
+  team_points_wide,
+  opp_points_wide,
+  by = c("team", "season", "week")
+)
+```
+
+Here is a correlation matrix of players' weekly performance across positions within a given team and their opponent:
+
+```{r}
+cor(
+  combined_points_wide %>% select(QB1:opp_DST),
+  use = "pairwise.complete.obs") %>% 
+  round(., 2)
+```
+
+In @fig-correlationMatrixPlot, we depict a correlation matrix plot using the `ggcorr()` function of the `GGally` [@R-GGally] package.
+
+```{r}
+#| label: fig-correlationMatrixPlot
+#| fig-cap: "Correlation Matrix Plot."
+#| fig-alt: "Correlation Matrix Plot."
+
+GGally::ggcorr(combined_points_wide %>% select(QB1:opp_DST))
+```
 

### 2025-07-22 modern-portfolio-theory.qmd
20250721 - diversification

@@
 That is, receiving higher returns generally requires taking on additional risk—at least as long as we stay along the [efficient frontier](#sec-efficientFrontier), described next.
+Diversification, by contrast, can limit your upside (by also limiting your downside).
+In other words, if your goal is to win your fantasy league, you may need to be willing to carry additional risk (e.g., drafting a Quarterback and Wide Receiver from the same team), knowing full well the possibility that it will not work out as predicted.
+However, winning a fantasy football league requires multiple predictions to work out as expected, which can benefit from situations where the bets were correlated.
+If you are going to lean into variability, you might take several steps: a) draft multiple players from the same team (aka "[stacking](#sec-stacking)"; e.g., Quarterback and Wide Receiver) or go all in on a high-powered offense and b) target high-risk, high-reward players (i.e., players with a high "ceiling"), such as rookies, sleepers, and players coming off injuries, trades, and suspensions.
 

### 2025-07-22 machine-learning.qmd
20250722 - fix error after new gpboost package version

@@
 
-A learning rate of 1 is very high for boosting.
-Even if a learning rate of 1 did well in tuning, we use a lower learning rate (0.1) to avoid overfitting.
-We also added some light regularization (`lambda_l2`) for better generalization.
-We also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^5 (32).
-We set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 10.
+We use the optimal parameters identified during tuning.
+We use a low learning rate (0.1) to avoid overfitting.
+We set some light regularization (`lambda_l2`) for better generalization.
+We also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^6 (64)—though the number will not reach greater than $2^\text{max_depth} = 2^5 = 32$.
+We set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 100.
 
@@
   max_depth = 5,
-  min_data_in_leaf = 10,
-  lambda_l2 = 1,
-  num_leaves = 2^5,
+  min_data_in_leaf = 100,
+  lambda_l2 = 5,
+  num_leaves = 2^6,
   num_threads = num_cores)
 
-nrounds_qb <- 123 # identify optimal number of trees through iteration and cross-validation
+nrounds_qb <- 84 # identify optimal number of trees through iteration and cross-validation
 

### 2025-07-23 machine-learning.qmd
20250722 - edits

@@
 
-The most accurate model, in terms of [discrimination](#sec-discrimination) ($R^2$), had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using [regression with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
-The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
+The most accurate models, in terms of [discrimination](#sec-discrimination) ($R^2$), had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using the [tree-boosting model](#sec-treeBoosting) and with [regression with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
+The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and the other [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
 This suggests that the other predictors examined did not have [incremental validity](#sec-incrementalValidity)—they did not show utility above and beyond a player's fantasy points in predicting their future fantasy points.
 If anything, considering the additional predictor variables showed *decremental* validity—the inclusion of the additional variables worsened prediction, possibly by weakening the predictive association of the previous season's fantasy points and misattributing its predictive effect to other, related predictors in the model.
-
+This is consistent with the common finding that simple models (and parsimony) are often preferable.
+As an example, @Youngstrom2018 found that simple models did just as well and in some cases better than [LASSO models](#sec-lasso) in classifying bipolar disorder.
+
+<!--
 Among the machine learning models, the [random forest model](#sec-randomForest) was most accurate according to $R^2$.
 However, it was still less accurate than a [simple regression model with one predictor](#sec-regressionOnePredictor).
-This is consistent with the common finding that simple models (and parsimony) are often preferable.
-As an example, @Youngstrom2018 found that simple models did just as well and in some cases better than [LASSO models](#sec-lasso) in classifying bipolar disorder.
-
-The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [tree-boosting model](#sec-treeBoosting), which had a [mean absolute error](#sec-meanAbsoluteError) of `{r} mae_minimum`.
+-->
+
+The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [tree-boosting model](#sec-treeBoosting), which had a [mean absolute error](#sec-meanAbsoluteError) of `{r} mae_minimum`, and was just slightly more accurate than the [regression model with one predictor](#sec-regressionOnePredictor).
 However, even in the most accurate model, the predicted values were over `{r} petersenlab::apa(mae_minimum, decimals = 0)` points away from the actual values, suggesting that the model-predicted values were not particularly accurate.

### 2025-07-23 machine-learning.qmd
20250722 - tweaks

@@
 We set some light regularization (`lambda_l2`) for better generalization.
-We also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at 2^6 (64)—though the number will not reach greater than $2^\text{max_depth} = 2^5 = 32$.
+We also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at $2^6$ (64)—though the number will not reach greater than $2^{\text{max_depth}} = 2^5 = 32$.
 We set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 100.
@@
 mae_treeBoosting <- round(accuracyIndices_treeBoosting$MAE, digits = 2)
+```
+
+@fig-treeBoostingModel depicts the predicted versus actual fantasy points for the model on the test data.
+
+```{r}
+#| label: fig-treeBoostingModel
+#| fig-cap: "Predicted Versus Actual Fantasy Points for Tree-Boosting Model."
+#| fig-alt: "Predicted Versus Actual Fantasy Points for Tree-Boosting Model."
+
+# Calculate combined range for axes
+axis_limits <- range(c(predictedVsActual$predictedPoints, predictedVsActual$actualPoints), na.rm = TRUE)
+
+ggplot(
+  predictedVsActual,
+  aes(
+    x = predictedPoints,
+    y = actualPoints)) +
+  geom_point(
+    size = 2,
+    alpha = 0.6) +
+  geom_abline(
+    slope = 1,
+    intercept = 0,
+    color = "blue",
+    linetype = "dashed") +
+  coord_equal(
+    xlim = axis_limits,
+    ylim = axis_limits) +
+  labs(
+    title = "Predicted vs Actual Fantasy Points (Test Data)",
+    x = "Predicted Fantasy Points",
+    y = "Actual Fantasy Points"
+  ) +
+  theme_classic() +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) # horizontal y-axis title
 ```

### 2025-07-23 machine-learning.qmd
20250723 - tweaks

@@
 Machine learning tends to be a data-driven, atheoretical technique.
-This can result in overfitting.
+This can result in [overfitting](#sec-overfitting).
 Thus, when estimating machine learning models, it is common to keep a hold-out sample for use in [cross-validation](#sec-crossValidation) to evaluate the extent of shrinkage of model coefficients.
@@
 We use the optimal parameters identified during tuning.
-We use a low learning rate (0.1) to avoid overfitting.
+We use a low learning rate (0.1) to avoid [overfitting](#sec-overfitting).
 We set some light regularization (`lambda_l2`) for better generalization.
@@
 
-We examined multiple machine learning models to identify the model that generates the most accurate prediction of Quarterbacks' future fantasy points.
+We examined multiple machine learning models that included many predictor variables to identify the model that generates the most accurate prediction of Quarterbacks' future fantasy points.
+We also compared them to simpler models, including a [regression model with one predictor](#sec-regressionOnePredictor) and a [regression model with multiple predictors](#sec-regressionMultiplePredictors) to deteremine the [incremental validity](#sec-incrementalValidity) of the machine learning models above and beyond the simpler models.
 A summary of the accuracy of the machine learning models is in @tbl-machineLearningModelSummary.
@@
 
-The most accurate models, in terms of [discrimination](#sec-discrimination) ($R^2$), had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using the [tree-boosting model](#sec-treeBoosting) and with [regression with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
+The most accurate models, in terms of [discrimination](#sec-discrimination) ($R^2$), had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using one of two approaches—a [tree-boosting model](#sec-treeBoosting) with many predictors or a [regression model with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
 The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and the other [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
@@
 
-The most accurate of models all explained less than half of the variance in players' future fantasy points.
+All models explained less than half of the variance in players' future fantasy points.
 Thus, there remains considerable variance to be explained.
@@
 All models explained less than half of the variance in players' fantasy points.
-The most accurate model, in terms of [discrimination](#sec-discrimination), was achieved using regression with one predictor: the player's fantasy points in the previous season.
+The most accurate model, in terms of [discrimination](#sec-discrimination) ($R^2$), was achieved using regression with one predictor: the player's fantasy points in the previous season.
 The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [tree-boosting model](#sec-treeBoosting), which accounted for the longitudinal nature of the data.

### 2025-07-23 machine-learning.qmd
20250723 - fix -- don't want to have to escape underscore b/c may not render in HTML/ePub

@@
 We set some light regularization (`lambda_l2`) for better generalization.
-We also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at $2^6$ (64)—though the number will not reach greater than $2^{\text{max_depth}} = 2^5 = 32$.
+We also set the maximum tree depth (`max_depth`) at 5 to capture complex (up to 5-way) interactions, and set the maximum number of terminal nodes (`num_leaves`) per tree at $2^6$ (64)—though the number will not reach greater than $2^{\text{max depth}} = 2^5 = 32$.
 We set the minimum number of samples in any leaf (`min_data_in_leaf`) to be 100.

### 2025-07-23 basic-statistics.qmd
20250708 - track chunk execution times
20250717 - Conclusion sections
20250723 - interpretation

@@
+{{< include _chunk-timing.qmd >}}
+
 # Basic Statistics {#sec-basicStats}
+
+This chapter provides an overview of foundational statistical concepts.
 
@@
 
+Cohen's $d$ is interpreted as the mean of group 1 was _ standard deviation units [larger/smaller] than the mean of group 2.
+
 #### Standardized Regression Coefficient (Beta; $\beta$) {#sec-beta}
@@
 where $B_x$ is the unstandardized regression coefficient of the [predictor variable](#sec-correlationalStudy) $x$ in predicting the [outcome variable](#sec-correlationalStudy) $y$, $s_x$ is the standard deviation of $x$, and $s_y$ is the standard deviation of $y$.
+The standardized regression coefficient is interpreted such that, for every 1 standard deviation increase in $x$, there was a _ standard deviation [increase/decrease] in $y$.
 

### 2025-07-23 getting-started.qmd
20250723 - additions

@@
 
+To subset a vector, use brackets to specify the elements to keep:
+
+```{r}
+#| eval: false
+
+vector[elementsToKeep]
+```
+
 To subset a data frame, use brackets to specify the subset of rows and columns to keep, where the value/vector before the comma specifies the rows to keep, and the value/vector after the comma specifies the columns to keep:
@@
 
-- numeric indices of the rows/columns to keep (or drop)
+- numeric indices of the elements/rows/columns to keep (or drop)
 - names of the rows/columns to keep (or drop)
-- values of `TRUE` and `FALSE` corresponding to which rows/columns to keep
+- values of `TRUE` and `FALSE` corresponding to which elements/rows/columns to keep
 
@@
 
+Missing elements are stored in R as `NA`.
+You can determine how many missing elements there are in a data frame using the following syntax:
+
+```{r}
+length(which(is.na(nfl_players)))
+```
+
+You can determine how many missing elements there are in a variable using the following syntax:
+
 ```{r}
@@
 ### Number of Non-Missing Elements {#sec-dataCharacteristicsNonMissingElements}
+
+You can determine how many non-missing elements there are in a data frame using the following syntax:
+
+```{r}
+length(which(!is.na(nfl_players)))
+```
+
+You can determine how many rows in the data frame have no missing data using the following syntax:
+
+```{r}
+nrow(na.omit(nfl_players))
+```
+
+You can determine how many non-missing elements there are in a vector using the following syntax:
 
@@
   by = c("ID")
-)
+) %>% 
+  arrange(ID, season, week)
 
@@
 
+## Commenting Code {#sec-comment}
+
+To comment your code, using the `#` sign.
+Anything on the line that appears after the `#` sign will be interpreted as a comment and will not be evaluated by R.
+It is important to comment your code frequently with what you are doing doing and why so that you and others know how to read it.
+
+```{r}
+x <- 1 # this line will run
+       # this line will NOT run: x <- 2
+
+x
+```
+
 # Conclusion {#sec-gettingStartedConclusion}

### 2025-07-29 player-evaluation.qmd
20250728 - team quality

@@
 - team quality
+    - quality of the team's offense
+    - quality of the team's defense
 - role on team
@@
 For a Wide Receiver, it is important to consider how strong the pass blocking is, and how strong the Quarterback is.
+In general, offensive players benefit from being on stronger offenses, so it can make sense to draft players (onto your fantasy team) that are on strong offenses.
+In addition, offensive players can also benefit from being on teams with weaker defenses because that may force the offense to score more points.
 

### 2025-07-29 getting-started.qmd
20250724 - sort data
20250729 - fix

@@
 
-For instance, you can use the following code to load the `tidyverse` package [@R-tidyverse]:
-
-```{r}
+For instance, you can use the following code to load the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) [@R-petersenlab] and `tidyverse` [@R-tidyverse] packages:
+
+```{r}
+library("petersenlab")
 library("tidyverse")
@@
 
+## Sort Data {#sec-sortData}
+
+```{r}
+players %>% 
+  arrange(position, age) #sort by position (ascending) then by age (ascending)
+
+players %>% 
+  arrange(position, -age) #sort by position (ascending) then by age (descending)
+```
+
 ## Data Characteristics {#sec-dataCharacteristics}

### 2025-07-29 base-rates.qmd
20250729 - update URLs

@@
 
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test.
 
@@
 A probability nomogram calculator is available from @Schwartz2006 at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from @Chekroud2017:
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from @Chekroud2017:
 

### 2025-07-29 basic-statistics.qmd
20250729 - update URLs

@@
 The mode of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_mode`.
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `Mode()` function for computing the mode of a set of data.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `Mode()` function for computing the mode of a set of data.
 
 If you want to give some values more weight to others, you can calculate a weighted mean and a weighted median (or other quantile), while assigning a weight to each value.
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains various functions for computing the weighted median (i.e., a weighted quantile at the 0.5 quantile, which is equivalent to the 50th percentile) based on @Akinshin2023.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains various functions for computing the weighted median (i.e., a weighted quantile at the 0.5 quantile, which is equivalent to the 50th percentile) based on @Akinshin2023.
 Because some projections are outliers, we use a trimmed version of the weighted Harrell-Davis quantile estimator for greater robustness.
@@
 If you want to give some values more weight to others, you can calculate weighted quantiles, while assigning a weight to each value.
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains various functions for computing weighted quantiles based on @Akinshin2023.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains various functions for computing weighted quantiles based on @Akinshin2023.
 Because some projections are outliers, we use a trimmed version of the weighted Harrell-Davis quantile estimator for greater robustness.

### 2025-07-29 evaluating-prediction-accuracy.qmd
20250722 - link
20250722 - add URL
20250729 - update URLs

@@
 
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `accuracyAtCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `accuracyAtCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables.
 
@@
 
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `accuracyAtEachCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables at each possible cutoff.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `accuracyAtEachCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables at each possible cutoff.
 
@@
 That is, each index of accuracy described in this section provides a single numerical index of accuracy that aggregates the accuracy across all possible cutoffs.
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] contains the `accuracyOverall()` function that computes many threshold-independent accuracy indices.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `accuracyOverall()` function that computes many threshold-independent accuracy indices.
 
@@
 Note: MPE is undefined when one or more of the observed values equals zero, due to division by zero.
-The `accuracyOverall()` function of the [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
 
@@
 Note: MAPE is undefined when one or more of the observed values equals zero, due to division by zero.
-The `accuracyOverall()` function of the [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
 
@@
 Note: sMAPE is undefined when one or more of the individuals has a prediction–observed combination such that the sum of the absolute value of the predicted value and the absolute value of the observed value equals zero ($|\text{predicted}_i| + |\text{observed}_i|$), due to division by zero.
-The `accuracyOverall()` function of the [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
 
@@
 When predicted or actual values are -1, this leads to $log(0)$, which is undefined.
-The `accuracyOverall()` function of the [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
 
@@
 Below are estimates of projection accuracy and [calibration plots](#sec-calibrationPlot) for [seasonal](#sec-accuracyProjectionsSeasonal) and [weekly](#sec-accuracyProjectionsWeekly) projections.
+FantasyFootballAnalytics.net also provides tools to evaluate the accuracy of historical projections [@FFA-accuracy]: <https://fantasyfootballanalytics.net/ffanalytics-web-app-accuracy-tab>.
 

### 2025-07-29 getting-started.qmd
20250729 - update URLs

@@
 
-The [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab] has a convenience function, `%ni%`, for the logical operator of determining whether an element is *not* in a value of another vector:
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] has a convenience function, `%ni%`, for the logical operator of determining whether an element is *not* in a value of another vector:
 

### 2025-07-29 machine-learning.qmd
20250729 - update URLs

@@
 We fit the final model using the `fit()` function of the `workflows` package [@R-workflows].
-We evaluate the accuracy of the model's predictions on the test data using the `accuracyOverall()` of the [`petersenlab`](https://github.com/DevPsyLab/petersenlab) package [@R-petersenlab].
+We evaluate the accuracy of the model's predictions on the test data using the `accuracyOverall()` of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab].
 

### 2025-07-29 player-evaluation.qmd
20250729 - tweaks

@@
 We demonstrate how to import data from the NFL Combine in @sec-combine.
-There are also calculators that integrate information about body shape and information from the NFL Combine to determine a player's relative athletic score (RAS) for their position: <https://ras.football/ras-calculator/>
+There are also calculators that integrate information about body shape and information from the NFL Combine to determine a player's relative athletic score (RAS) for their position: <https://ras.football/ras-calculator/> and <https://www.nfl.com/combine/iq/>
 
@@
 - Will there be more competition for targets?
+- Is there a greater level of competition by opposing teams?
 
@@
 
-In terms of consistency, how consistent was the player they from game to game and from season to season?
+In terms of consistency, how consistent was the player from game to game and from season to season?
 For instance, we could examine the standard deviations of players' fantasy points across games in a given season.
@@
 A player's knowledge, intelligence, and reaction time can help them gain an upper-hand even when they may not be the fastest or strongest.
+The Wonderlic test was formerly used to assess a player's cognitive abilities [@OConnell2022; archived at <https://web.archive.org/web/20220303015018/https://www.nytimes.com/2022/03/02/sports/football/nfl-wonderlic-test.html>].
 A player's work ethic and mental toughness may help them be resilient and persevere in the face of challenges.

### 2025-07-30 getting-started.qmd
20250730 - updates

@@
 
+## General Workflow {#sec-generalWorkflowAnalysis}
+
+Here is a general workflow for statistical analysis:
+
+1. Load the packages you need
+1. Load the data files
+1. Do any data processing needed to prepare the data files to be merged.
+  - For example, make sure each data file has unique values on the keys.
+1. Merge the data files
+1. Do any data processing needed to prepare the data files for analysis.
+For example:
+  - calculate/recode variables
+  - transform wide to long or long to wide
+  - filter to the relevant rows
+  - examine descriptive statistics to ensure data are plausible
+- Do the analysis!
+  - evaluate analysis assumptions
+  - conduct sensitivity analyses to see extent to which findings differ using different subsets of the data, different modeling approaches/assumptions, etc.
+
 ## Install Packages {#sec-installingPackages}

### 2025-07-30 research-methods.qmd
20250730 - updates

@@
 ```
+
+## General Workflow {#sec-generalWorkflowStudy}
+
+Here is a general workflow for a study:
+
+1. Formulate your [research questions, hypotheses, resulting predictions, and exploratory questions](#sec-hypothesisVsPrediction)
+1. Obtain approval from an institutional review board (IRB) to conduct the research
+    - In the United States, analyzing publicly available data from `nflverse` is not considered human subjects research because the work is not considered to involve human subjects—the data involve publicly available information that is not sensitive (i.e., it does not pose risks to players' rights, privacy, or welfare), and the work does not involve intervention, interaction with the players, private information, or biospecimens.
+1. Design the study to test your research questions
+1. Collect the data you need to test your research questions
+1. Analyze the data to test your research questions
+1. Communicate the findings to others (e.g., papers, posters, talks, news articles, interviews, blog posts)
 
@@
 > If the hypothesis is true, I <u>*predict*</u> that players who consume sports drink during a game will score more fantasy points than players who do not consume sports drink during the game.
+
+Research questions that you do not have hypotheses and predictions for are considered exploratory questions.
+Both hypothesis-driven work and exploration are important parts of science.
+Just be honest about which one you are pursuing for each question.
+Secretly hypothesizing after the results are known in the Introduction section of a paper (known as SHARKing) is problematic [@Hollenbeck2017].
+Conducting many exploratory analyses is potentially more prone to false-positive results than conducting a single confirmatory (i.e., hypothesis-driven) test.
+Thus, presenting exploratory work as confirmatory is misrepresenting the likelihood that your results are true and can contribute to the replication crisis—the failure of many findings from many studies to replicate upon subsequent investigation.
+By contrast, it is fine to transparently hypothesize after the results are known in the Disucssion section of a paper (known as THARKing), because this can provide potential explanations for the findings that can be evaluated in subsequent analyses or studies [@Hollenbeck2017].
 

### 2025-07-31 research-methods.qmd
20250731 - addition

@@
 For example, a research question might be: "Does consumption of sports drink improve player performance?"
-A hypothesis is a proposed explanation.
+A hypothesis is a proposed explanation; it specifies the causal relation between processes.
 A prediction is "the expected result of a test that is derived, by deduction, from a hypothesis or theory" [@Eastwell2014, p. 17; archived at <https://perma.cc/8EX4-8JYN>].

### 2025-07-31 mythbusters.qmd
20250720 - conclusion sections
20250731 - research question, hypothesis, and prediction

@@
 Let's examine the question empirically.
-In order to do that, we have to make some assumptions/constraints.
+Our <u>research questions</u> is: Do players perform better in their "contract year" (i.e., the last year of their contract)?
+Our <u>hypothesis</u> is that players are motivated to get larger contracts (more money), leading players in their contract year to try harder and perform better.
+If the hypothesis is true, we <u>predict</u> that players who are in their contract year will tend to score more fantasy points than players who are not in their contract year.
+
+In order to test this question empirically, we have to make some assumptions/constraints.
 In this example, we will make the following constraints:
@@
 In sum, we did not find evidence in support of the contract year hypothesis and consider this myth debunked.
-However, we are open to this possibility being reexamined in new ways or with additional performance
+However, we are open to this possibility being reexamined in new ways or with additional performance metrics.
 

### 2025-07-31 correlation.qmd
20250731 - correlogram

@@
 library("psych")
-library("corrplot")
 library("correlation")
+library("GGally")
 library("ggridges")
@@
 
-We can depict a correlogram using the `corrplot()` function of the `corrplot` package [@R-corrplot], as shown in @fig-correlogram.
+We can depict a correlation matrix plot (correlogram) using the `ggcorr()` function of the `GGally` [@R-GGally] package, as shown in @fig-correlogram.
 
@@
 #| label: fig-correlogram
-#| fig-cap: "Correlogram."
-#| fig-alt: "Correlogram."
-
-corrplot(cor(
-  player_stats_seasonal[,c("height","weight","fantasyPoints")],
-  use = "pairwise.complete.obs"))
+#| fig-cap: "Correlation Matrix Plot (Correlogram)."
+#| fig-alt: "Correlation Matrix Plot (Correlogram)."
+
+GGally::ggcorr(player_stats_seasonal[,c("height","weight","fantasyPoints")])
 ```

### 2025-07-31 getting-started.qmd
20250731 - indentation
20250731 - correlogram

@@
      "parallelly","doParallel","missRanger","ggridges","powerjoin","caret",
-     "LongituRF","gpboost","corrplot","mgcv","rms","car","lavaan","lavaanPlot",
-     "lavaangui","mice","miceadds","interactions","robustbase","ordinal","MASS",
+     "LongituRF","gpboost","mgcv","rms","car","lavaan","lavaanPlot","lavaangui",
+     "mice","miceadds","interactions","robustbase","ordinal","MASS",
      "data.table","future","future.apply","SimDesign","domir","GGally"))

### 2025-07-31 index.qmd
20250731 - correlogram

--- 2025-07-18 index.qmd
@@
   "stringi", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
-  "caret", "LongituRF", "gpboost", "cmdstanr", "corrplot", "mgcv", "rms", 
-  "car", "lavaan", "lavaanPlot", "mice", "miceadds", "interactions", 
-  "robustbase", "ordinal", "MASS", "data.table", "future", "future.apply", 
-  "SimDesign", "domir",
+  "caret", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", "lavaan", 
+  "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", "ordinal", 
+  "MASS", "data.table", "future", "future.apply", "SimDesign", "domir", 
+  "GGally",
   "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem

### 2025-08-01 basic-statistics.qmd
20250731 - tweaks

@@
 
-Indices to describe the *spread* (variability) of a variable's data include:
+Indices to describe the *spread* (variability or dispersion) of a variable's data include:
 
@@
 
+The skewness represents the degree of asymmetry of a distribution around its mean.
 Positive skewness (right-skewed) reflects a longer or heavier right-tailed distribution, whereas negative skewness (left-skewed) reflects a longer or heavier left-tailed distribution.
@@
 
-In science, statistical significance is evaluated with the *p*-value.
+Statistical significance is the reliability (i.e., consistency) of an effect, and it is evaluated with the *p*-value.
 The *p*-value does not represent the probability that you observed the result by chance.
@@
 
+Cohen's $d$ is commonly used for evaluating the magnitude of group differences.
 Cohen's $d$ is calculated as in @eq-cohensD:
@@
 Like two-samples *t*-tests, ANOVA allows examining whether groups differ as a function of an [independent variable](#sec-experiment).
-However, unlike a *t*-test, ANOVA allows examining multiple multiple [independent variables](#sec-experiment) and more than two groups.
+However, unlike a *t*-test, ANOVA allows examining multiple [independent variables](#sec-experiment) and more than two groups.
 The null hypothesis is that the the groups' mean value does not differ systematically.

### 2025-08-01 research-methods.qmd
20250731 - tweaks

@@
 In mediation, a predictor (`X`) leads to a mediator (`M`), which leads to an outcome (`Y`).
-Mediation answers the question of, "<u>**Why (or how)** does `X` influence `Y`?</u> A <u>mediator</u> (`M`) is a variable that helps explain the assocation between two other variables, and it answers the question of why/how `X` influences `Y`.
+Mediation answers the question of, "<u>**Why (or how)** does `X` influence `Y`?</u>"
+A <u>mediator</u> (`M`) is a variable that helps explain the association between two other variables, and it answers the question of why/how `X` influences `Y`.
 That is, the mediator is the variable that helps explain how/why `X` is related to `Y`.
@@
 
-**Moderation** (sometimes called an "interaction"), on the other hand, occurs when there is a variable or condition (`M`; called a "moderator") that <u>changes the assocation</u> between `X` and `Y`.
+**Moderation** (sometimes called an "interaction"), on the other hand, occurs when there is a variable or condition (`M`; called a "moderator") that <u>changes the association</u> between `X` and `Y`.
 That is, the effect of the predictor variable on the outcome variable differs at different levels of the moderator variable.

### 2025-08-01 data-visualization.qmd
20250801 - ggplot2 syntax

@@
 Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
+Good colorblind-friendly color schemes for binned (`_b`), continuous (`_c`), and discrete (i.e., categorical; `_d`) scales are available using the `scale_color_viridis_X()` and `scale_fill_viridis_X()` functions of the `ggplot2` package [@R-ggplot2] (where `X` is `b`, `c`, or `d`, depending on whether you want a binned, continuous, discrete scale).
 There are a variety of resources for color schemes that are accessible to color-blind viewers:
@@
 - @Nunez2018
+
+## `ggplot2` Syntax {#sec-ggplot2Syntax}
+
+### General Syntax {#sec-ggplot2SyntaxGeneral}
+
+The `ggplot2` package [@R-ggplot2] is based on the grammar of graphics [@ggplot3rdEd], which is the principle that you can build any graph from three components: 1) the data set, 2) the coordinate system, and 3) the geoms—visual marks that represent data points.
+The general syntax for plots created using the `ggplot2` package [@R-ggplot2] is:
+
+```{r}
+#| eval: false
+
+ggplot2::ggplot(
+  data = INSERT, # 1) data set
+  mapping = aes(
+    x = variableName1, # name of variable on x-axis
+    y = variableName2) # name of variable on y-axis
+) +
+  coord_cartesian( # 2) coordinate system
+    xlim = c(NA, NA), # set limits (min and max) of x-axis
+    ylim = c(NA, NA)) + # set limits (min and max) of y-axis
+  INSERT_GEOM_FUNCTIONS # 3) geom(s)
+```
+
+The `coord_cartesian()` function can be used to specify the axis limits; however, this function is not required and the plot will apply sensible limits by default.
+In addition, there are optional settings you can specify or modify, such as [axis titles](#sec-scatterplotPlotLabels), [axis labels](#sec-scatterplotAxes), [faceting](#sec-faceting), and the [theme](#sec-scatterplotTheme).
+
+Here is a template including these potential modifications:
+
+```{r}
+#| eval: false
+
+ggplot2::ggplot(
+  data = INSERT, # 1) data set
+  mapping = aes(
+    x = variableName1, # name of variable on x-axis
+    y = variableName2) # name of variable on y-axis
+) +
+  coord_cartesian( # 2) coordinate system
+    xlim = c(NA, NA), # set limits (min and max) of x-axis
+    ylim = c(NA, NA)) + # set limits (min and max) of y-axis
+  INSERT_GEOM_FUNCTIONS + # 3) geom(s)
+  scale_x_continuous(
+    breaks = INSERT # specify x-axis labels
+  ) +
+  scale_y_continuous(
+    breaks = INSERT # specify y-axis labels
+  ) +
+  labs( # add plot labels
+    x = "INSERT", # x-axis title
+    y = "INSERT", # y-axis title
+    title = "INSERT", # plot title
+    subtitle = "INSERT", # plot subtitle
+    caption = "Add a caption below plot", # plot caption
+    alt = "Add alt text to the plot" # plot alt-text
+  ) +
+  INSERT_THEME_NAME + # apply theme
+  facet_wrap(vars(INSERT_VARIABLES_TO_FACET)) # facet by one or more variables
+```
+
+### Geoms {#sec-ggplot2SyntaxGeoms}
+
+Here is a list of common geoms available in `ggplot2` (there are many others):
+
+- [histogram](#sec-histogram): `geom_histogram()`
+- [density plot](#sec-densityPlot): `geom_density()`
+- [box-and-whisker plot](#sec-boxAndWhiskerPlot): `geom_boxplot()`
+- [violin plot](#sec-violinPlot): `geom_violin()`
+- [bar plot](#sec-barPlot): `geom_bar()`
+- [scatterplot](#sec-scatterplot): `geom_point()`
+- [rug plot](#sec-scatterplotRugPlot): `geom_rug()`
+- [line chart](#sec-lineChart): `geom_line()`
+- statistical summary (e.g., [best-fit line](#sec-scatterplotBestFitLine)): `geom_smooth()`
+
+### Color {#sec-ggplot2SyntaxColor}
+
+In `ggplot2`, you can specify the color of a geom using one of two parameters: 1) `fill` and 2) `color`.
+The `fill` parameter specifies the color inside the geom.
+The `color` parameter specifies the outline/border color of the geom.
+
+The colors to use in the `fill` and `color` parameters can be specified by name (as in @fig-colorByName), hexadecimal (HEX) codes (as in @fig-colorByHEX), RGB values (as in @fig-colorByRGB), or by palettes (e.g., ColorBrewer or viridis; as in Figures [-@fig-multipleColorsByPaletteColorBrewer] and [-@fig-multipleColorsByPaletteViridis]).
+
+```{r}
+#| label: fig-colorByName
+#| fig-cap: "Specifying the Color/Fill by Name."
+#| fig-alt: "Specifying the Color/Fill by Name."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasyPoints,
+    fill = position_group)
+) +
+  geom_histogram(
+    color = "black",
+    fill = "forestgreen"
+  )
+```
+
+You can see a full list of color names available using the `colors()` function:
+
+```{r}
+colors()
+```
+
+However, you can use any color by specifying its HEX code, as in @fig-colorByHEX.
+
+```{r}
+#| label: fig-colorByHEX
+#| fig-cap: "Specifying the Color/Fill by Hexadecimal Code."
+#| fig-alt: "Specifying the Color/Fill by Hexadecimal Code."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasyPoints,
+    fill = position_group)
+) +
+  geom_histogram(
+    color = "#000000",
+    fill = "#228B22"
+  )
+```
+
+you can use any color by specifying its values in red-green-blue (RGB) color space using the `rgb()` function, as in @fig-colorByRGB.
+
+```{r}
+#| label: fig-colorByRGB
+#| fig-cap: "Specifying the Color/Fill by RGB Value."
+#| fig-alt: "Specifying the Color/Fill by RGB Value."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasyPoints)
+) +
+  geom_histogram(
+    color = rgb(
+      red = 0,
+      green = 0,
+      blue = 0,
+      maxColorValue = 255
+    ),
+    fill = rgb(
+      red = 34,
+      green = 139,
+      blue = 34,
+      maxColorValue = 255
+    )
+  )
+```
+
+We can allow the colors to differ by group by specifying the `fill` parameter in the mapping aesthetics, as in @fig-colorByGroup.
+
+```{r}
+#| label: fig-colorByGroup
+#| fig-cap: "Specifying Color By Group."
+#| fig-alt: "Specifying Color By Group."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasyPoints,
+    fill = position_group)
+) +
+  geom_density(
+    color = "black",
+    alpha = 0.7
+  )
+```
+
+We can specify multiple colors manually using the `scale_fill_manual()` function, as in @fig-multipleColorsManually.
+
+```{r}
+#| label: fig-multipleColorsManually
+#| fig-cap: "Specifying Multiple Colors Manually."
+#| fig-alt: "Specifying Multiple Colors Manually."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasyPoints,
+    fill = position_group)
+) +
+  geom_density(
+    color = "black",
+    alpha = 0.7
+  ) +
+  scale_fill_manual(
+    values = c(
+      "QB" = "red", # or could use hex, rgb, etc.
+      "RB" = "blue", # or could use hex, rgb, etc.
+      "WR" = "green", # or could use hex, rgb, etc.
+      "TE" = "yellow") # or could use hex, rgb, etc.
+    )
+```
+
+We can also specify the colors using a palette, such as by the ColorBrewer palette using the `scale_fill_brewer()` function (as in @fig-multipleColorsByPaletteColorBrewer) or the viridis palette using the `scale_fill_viridis_d()` function (as in @fig-multipleColorsByPaletteViridis).
+
+```{r}
+#| label: fig-multipleColorsByPaletteColorBrewer
+#| fig-cap: "Specifying Multiple Colors Using the Color Brewer Palette."
+#| fig-alt: "Specifying Multiple Colors Using the Color Brewer Palette."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasyPoints,
+    fill = position_group)
+) +
+  geom_density(
+    color = "black",
+    alpha = 0.7
+  ) +
+  scale_fill_brewer(palette = "Set1") # see ?scale_fill_brewer for list of available palettes
+```
+
+```{r}
+#| label: fig-multipleColorsByPaletteViridis
+#| fig-cap: "Specifying Multiple Colors Using the Viridis Palette."
+#| fig-alt: "Specifying Multiple Colors Using the Viridis Palette."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(position_group %in% c("QB","RB","WR","TE")),
+  mapping = aes(
+    x = fantasyPoints,
+    fill = position_group)
+) +
+  geom_density(
+    color = "black",
+    alpha = 0.7
+  ) +
+  scale_fill_viridis_d(option = "viridis") # see ?scale_fill_viridis_d for list of available palettes
+```
+
+### Line Type {#sec-ggplot2SyntaxLineType}
+
+Common line types include "solid" lines, "dashed" lines, and "dotted" lines.
+A full list of line types available is provided here: <https://ggplot2.tidyverse.org/articles/ggplot2-specs.html> (archived at <https://perma.cc/N9ED-6LUS>).
+To specify the line type, you can use the `linetype` parameter, as in @fig-lineType.
+
+```{r}
+#| label: fig-lineType
+#| fig-cap: "Specifying Line Type."
+#| fig-alt: "Specifying Line Type."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(player_display_name %in% c("Tom Brady")),
+  mapping = aes(
+    x = season,
+    y = fantasyPoints,
+    group = player_display_name)
+) +
+  geom_line(
+    linetype = "dashed"
+  )
+```
+
+We can allow the line type to differ by group by specifying the `linetype` parameter in the mapping aesthetics, as in @fig-lineTypeByGroup.
+
+```{r}
+#| label: fig-lineTypeByGroup
+#| fig-cap: "Specifying Line Type By Group."
+#| fig-alt: "Specifying Line Type By Group."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(player_display_name %in% c("Tom Brady","Peyton Manning")),
+  mapping = aes(
+    x = season,
+    y = fantasyPoints,
+    linetype = player_display_name)
+) +
+  geom_line()
+```
+
+We can specify the line types manually using the `scale_linetype_manual()` function as in @fig-lineTypeByGroupManually.
+
+```{r}
+#| label: fig-lineTypeByGroupManually
+#| fig-cap: "Specifying Line Type By Group Manually."
+#| fig-alt: "Specifying Line Type By Group Manually."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(player_display_name %in% c("Tom Brady","Peyton Manning")),
+  mapping = aes(
+    x = season,
+    y = fantasyPoints,
+    group = player_display_name,
+    linetype = player_display_name)
+) +
+  geom_line() +
+  scale_linetype_manual(
+    values = c(
+      "Tom Brady" = "solid",
+      "Peyton Manning" = "dashed")
+    )
+```
+
+### Line Width {#sec-ggplot2SyntaxLineType}
+
+To specify the line width, you can use the `linewidth` parameter, as in @fig-linewidth.
+
+```{r}
+#| label: fig-linewidth
+#| fig-cap: "Specifying Line Width."
+#| fig-alt: "Specifying Line Width."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(player_display_name %in% c("Tom Brady")),
+  mapping = aes(
+    x = season,
+    y = fantasyPoints,
+    group = player_display_name)
+) +
+  geom_line(
+    linewidth = 4
+  )
+```
+
+### Point Shape {#sec-ggplot2SyntaxPointShape}
+
+Common point shapes include circles, triangles, squares, etc.
+A full list of point shapes available is provided here: <https://ggplot2.tidyverse.org/articles/ggplot2-specs.html> (archived at <https://perma.cc/N9ED-6LUS>).
+To specify the point shape, you can use the `shape` parameter, as in @fig-pointShape.
+
+```{r}
+#| label: fig-pointShape
+#| fig-cap: "Specifying Point Shape."
+#| fig-alt: "Specifying Point Shape."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(player_display_name %in% c("Tom Brady")),
+  mapping = aes(
+    x = season,
+    y = fantasyPoints,
+    group = player_display_name)
+) +
+  geom_point(
+    shape = 2
+  )
+```
+
+We can allow the point shape to differ by group by specifying the `shape` parameter in the mapping aesthetics, as in @fig-pointShapeByGroup.
+
+```{r}
+#| label: fig-pointShapeByGroup
+#| fig-cap: "Specifying Point Shape By Group."
+#| fig-alt: "Specifying Point Shape By Group."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(player_display_name %in% c("Tom Brady","Peyton Manning")),
+  mapping = aes(
+    x = season,
+    y = fantasyPoints,
+    shape = player_display_name)
+) +
+  geom_point()
+```
+
+We can specify the point shape manually using the `scale_shape_manual()` function as in @fig-pointShapeByGroupManually.
+
+```{r}
+#| label: fig-pointShapeByGroupManually
+#| fig-cap: "Specifying Point Shape By Group Manually."
+#| fig-alt: "Specifying Point Shape By Group Manually."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(player_display_name %in% c("Tom Brady","Peyton Manning")),
+  mapping = aes(
+    x = season,
+    y = fantasyPoints,
+    shape = player_display_name)
+) +
+  geom_point() +
+  scale_shape_manual(
+    values = c(
+      "Tom Brady" = 1,
+      "Peyton Manning" = 3)
+    )
+```
+
+### Size of Points and Text {#sec-ggplot2SyntaxSize}
+
+To specify the size of points and text, you can use the `size` and `base_size` parameters, respectively, as in @fig-sizePointsText.
+
+```{r}
+#| label: fig-sizePointsText
+#| fig-cap: "Specifying Size of Points and Text."
+#| fig-alt: "Specifying Size of Points and Text."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(player_display_name %in% c("Tom Brady")),
+  mapping = aes(
+    x = season,
+    y = fantasyPoints,
+    group = player_display_name)
+) +
+  geom_point(
+    size = 2 # size of points
+  ) +
+  theme_classic(
+    base_size = 20 # font size
+  )
+```
+
+You can also specify the font size separately for particular elements, as in @fig-sizeTextManually.
+
+```{r}
+#| label: fig-sizeTextManually
+#| fig-cap: "Specifying Size of Text Manually for Particular Elements."
+#| fig-alt: "Specifying Size of Text Manually for Particular Elements."
+
+ggplot2::ggplot(
+  data = player_stats_seasonal %>%
+    filter(player_display_name %in% c("Tom Brady")),
+  mapping = aes(
+    x = season,
+    y = fantasyPoints,
+    group = player_display_name)
+) +
+  geom_point() +
+  theme(
+    axis.title.x = element_text(size = 30), # font size for x-axis title
+    axis.title.y = element_text(size = 30), # font size for y-axis title
+    axis.text.x  = element_text(size = 12), # font size for x-axis labels
+    axis.text.y  = element_text(size = 12)  # font size for y-axis labels
+  )
+```
 
@@
   labs( # add plot labels
-    x = "Running Back's Age (years)",
-    y = "Rushing Yards (Season)",
-    title = "NFL Rushing Yards (Season) by Player Age",
-    subtitle = "(Among Running Backs)"
+    x = "Running Back's Age (years)", # x-axis title
+    y = "Rushing Yards (Season)", # y-axis title
+    title = "NFL Rushing Yards (Season) by Player Age", # plot title
+    subtitle = "(Among Running Backs)" # plot subtitle
   )

### 2025-08-01 data-visualization.qmd
20250801 - tweaks

@@
 
-"Showing the data" involves showing the data that matters the most.
+"Showing the data" involves showing the data that matter the most.
 "Reducing the clutter" involves removing non-data things that obscure the data—for example, extraneous gridlines, tick marks, data markers (e.g., symbols to distinguish between series), and complex shadings (e.g., textured or filled gradients).
@@
 
-The Data Visualization Catalogue provides examples of various types of plots depending on one's goal: <https://datavizcatalogue.com/search.html>.
+The Data Visualization Catalogue, from Data to Viz, and Data Viz Project provide examples of various types of plots depending on one's goal: <https://datavizcatalogue.com/search.html>, <https://www.data-to-viz.com>, and <https://datavizproject.com>.
 The R Graph Gallery provides examples of various types of plots and how to create them in `R`: <https://r-graph-gallery.com>.
@@
 
-The colors to use in the `fill` and `color` parameters can be specified by name (as in @fig-colorByName), hexadecimal (HEX) codes (as in @fig-colorByHEX), RGB values (as in @fig-colorByRGB), or by palettes (e.g., ColorBrewer or viridis; as in Figures [-@fig-multipleColorsByPaletteColorBrewer] and [-@fig-multipleColorsByPaletteViridis]).
+The colors to use in the `fill` and `color` parameters can be specified by name (as in @fig-colorByName), hexadecimal (HEX) codes (as in @fig-colorByHEX), values in red-green-blue (RGB) space (as in @fig-colorByRGB), or palettes (e.g., ColorBrewer or viridis; as in Figures [-@fig-multipleColorsByPaletteColorBrewer] and [-@fig-multipleColorsByPaletteViridis]).
 
@@
 
-you can use any color by specifying its values in red-green-blue (RGB) color space using the `rgb()` function, as in @fig-colorByRGB.
+You can use any color by specifying its values in red-green-blue (RGB) color space using the `rgb()` function, as in @fig-colorByRGB.
 
@@
 #| label: fig-colorByRGB
-#| fig-cap: "Specifying the Color/Fill by RGB Value."
-#| fig-alt: "Specifying the Color/Fill by RGB Value."
+#| fig-cap: "Specifying the Color/Fill by Value in Red-Green-Blue (RGB) Space."
+#| fig-alt: "Specifying the Color/Fill by Value in Red-Green-Blue (RGB) Space."
 
@@
 #| label: fig-multipleColorsByPaletteColorBrewer
-#| fig-cap: "Specifying Multiple Colors Using the Color Brewer Palette."
-#| fig-alt: "Specifying Multiple Colors Using the Color Brewer Palette."
+#| fig-cap: "Specifying Multiple Colors Using the ColorBrewer Palette."
+#| fig-alt: "Specifying Multiple Colors Using the ColorBrewer Palette."
 

### 2025-08-01 getting-started.qmd
20250801 - tweaks

@@
     - examine descriptive statistics to ensure data are plausible
-- Do the analysis!
+1. Do the analysis!
     - evaluate analysis assumptions

### 2025-08-01 data-visualization.qmd
20250801 - tweaks

@@
 
-The `ggplot2` package [@R-ggplot2] is based on the grammar of graphics [@ggplot3rdEd], which is the principle that you can build any graph from three components: 1) the data set, 2) the coordinate system, and 3) the geoms—visual marks that represent data points.
+The `ggplot2` package [@R-ggplot2] is based on the grammar of graphics [@ggplot3rdEd], which is the principle that you can build any graph from three components: 1) the data, 2) the coordinate system and aesthetics, and 3) the geometries ("geoms")—visual marks that represent data points.
 The general syntax for plots created using the `ggplot2` package [@R-ggplot2] is:
@@
 ggplot2::ggplot(
-  data = INSERT, # 1) data set
+  data = INSERT, # 1) data
   mapping = aes(
@@
 
-### Geoms {#sec-ggplot2SyntaxGeoms}
-
-Here is a list of common geoms available in `ggplot2` (there are many others):
+### Geometries {#sec-ggplot2SyntaxGeoms}
+
+Here is a list of common geometrics ("geoms") available in `ggplot2` (there are many others):
 
@@
   geom_rug(
-    col = "#800000",
+    color = "#800000",
     alpha = 0.2) +

### 2025-08-01 data-visualization.qmd
20250801 - shapes and tweaks

@@
 
-The `ggplot2` package [@R-ggplot2] is based on the grammar of graphics [@ggplot3rdEd], which is the principle that you can build any graph from three components: 1) the data, 2) the coordinate system and aesthetics, and 3) the geometries ("geoms")—visual marks that represent data points.
+The `ggplot2` package [@R-ggplot2] is based on the grammar of graphics [@ggplot3rdEd], which is the principle that you can build any graph from three components: 1) the data, 2) the coordinate system (and aesthetic attributes), and 3) the geometric objects ("geoms")—visual marks that represent data points (e.g., points, lines, and bars).
 The general syntax for plots created using the `ggplot2` package [@R-ggplot2] is:
@@
 
-### Geometries {#sec-ggplot2SyntaxGeoms}
-
-Here is a list of common geometrics ("geoms") available in `ggplot2` (there are many others):
+You can layer multiple geoms in the same plot, layer by layer.
+You can have a different data set and aesthetics associated with each geom, by specifying the `data` and `mapping` arguments for the function of that particular geom.
+For an example of a plot that uses different data sets for different geoms, see @fig-bootstrapRegressionCoefficients (@sec-multipleRegressionRobustRegression).
+
+### Geometric Objects {#sec-ggplot2SyntaxGeoms}
+
+Here is a list of common geometric objects ("geoms") available in `ggplot2` (there are many others):
 
@@
 Common point shapes include circles, triangles, squares, etc.
-A full list of point shapes available is provided here: <https://ggplot2.tidyverse.org/articles/ggplot2-specs.html> (archived at <https://perma.cc/N9ED-6LUS>).
+A full list of point shapes available is provided here: <https://ggplot2.tidyverse.org/articles/ggplot2-specs.html> (archived at <https://perma.cc/N9ED-6LUS>) and is in @fig-pointShapeOptions.
+
+::: {#fig-pointShapeOptions}
+![](images/pointShapes.png){width=100% fig-alt="Available Point Shapes in ggplot2."}
+
+Available Point Shapes in ggplot2.
+:::
+
 To specify the point shape, you can use the `shape` parameter, as in @fig-pointShape.

### 2025-08-01 multiple-regression.qmd
20250801 - shapes and tweaks

@@
 
-ggplot(
-  quadraticTermsRegressionModel_tidy,
+ggplot2::ggplot(
+  data = quadraticTermsRegressionModel_tidy,
   aes(
@@
 
-ggplot(
-  quadraticTermsRegressionModelStandardized_tidy,
+ggplot2::ggplot(
+  data = quadraticTermsRegressionModelStandardized_tidy,
   aes(
@@
 ggplot(
-  boot_coefs,
-  aes(estimate)) +
+  data = boot_coefs,
+  aes(x = estimate)) +
   geom_histogram(
@@
   geom_vline(
+    data = percentile_intervals,
     aes(
       xintercept = .lower),
-    data = percentile_intervals,
     col = "blue",
@@
   geom_vline(
+    data = percentile_intervals,
     aes(
       xintercept = .upper),
-    data = percentile_intervals,
     col = "blue",

### 2025-08-02 correlation.qmd
20250801 - correlation matrix plots
20250802 - degrees of freedom

@@
 The [effect size](#sec-practicalSignificance) of the association is small—for what is considered a small, medium, and large effect size for $r$ values, see @sec-effectSizeThresholds.
+The number of degrees of freedom ($df$) for a statistical test of a correlation coefficient is in @eq-correlationDF.
 The $p$-value is less than .05, so it is a statistically significant correlation.
 We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r(`r pearson_df`) = `r petersenlab::apa(pearson_r, decimals = 2, leading = FALSE)`$, $p `r petersenlab::pValue(pearson_p)`$); the taller a player was, the fewer fantasy points they tended to score.
+
+$$
+df = n - 2
+$$ {#eq-correlationDF}
 
@@
 
-GGally::ggcorr(player_stats_seasonal[,c("height","weight","fantasyPoints")])
+GGally::ggcorr(
+  player_stats_seasonal[,c("height","weight","fantasyPoints")],
+  label = TRUE,
+  label_round = 2)
 ```

### 2025-08-02 multiple-regression.qmd
20250802 - fix implemented for calculation of target share
https://github.com/nflverse/nflfastR/pull/505/commits/2308d30cd96d10ffdd6966a1eccb1f7c4750be47

@@
 We want to consider several [predictors](#sec-correlationalStudy), including the player's age, height, weight, and target share.
+Target share is computed as the number of targets a player receives divided by the team's total number of targets.
 We have only a few [predictors](#sec-correlationalStudy) and our sample size is large enough such that [overfitting](#sec-overfitting) is not likely a concern.
-
-On a weekly basis, target share is computed as the number of targets a player receives divided by the team's total number of targets.
-It would be nice to compute it this way for seasonal data, as well; however, for calculating seasonal target share, the `nflfastR` [@R-nflfastR] package currently sums up the player's weekly target share.
-Thus, seasonal target share (unlike weekly target share)—as it is currently calculated—does not range from 0–1.
-For example, if a player has a weekly target share of 15%, that would be a seasonal (cross-week-summed) target share of 2.55 ($0.15 \times 17 \; \text{games} = 2.55$).
 
@@
 We could also generate the model-implied prediction of fantasy points for any value of the [predictor variables](#sec-correlationalStudy).
-For instance, here are the number of fantasy points expected for a Wide Receiver who is 23 years old, is 6'2" tall (72 inches), weighs 200 pounds, and has a (cross-week-summed) target share of 3 (which reflects an approximate target share of 17.6% percent per game; i.e., $0.176 = \frac{3}{17 \; \text{games}}$):
+For instance, here are the number of fantasy points expected for a Wide Receiver who is 23 years old, is 6'2" tall (72 inches), weighs 200 pounds, and has a target share of 50% (i.e., 0.5):
 
@@
     weight = 200,
-    target_share = 3
+    target_share = .5
   )
@@
 weight <- 200
-target_share <- 3
+target_share <- .5
 

### 2025-08-03 actuarial.qmd
20250803 - shrinkage

@@
 In general, there is often shrinkage of estimates from training data set to a test data set.
-*Shrinkage* is when variables with stronger predictive power in the original data set tend to show somewhat smaller predictive power (smaller regression coefficients) when applied to new groups.
+*Shrinkage* is when variables with stronger predictive power in the original data set tend to show somewhat smaller predictive power (smaller regression coefficients and resulting accuracy metrics, such as $R^2$) when applied to new groups.
 Shrinkage reflects a model [overfitting](#sec-overfitting)—i.e., when the model explains error variance by capitalizing on chance.

### 2025-08-03 machine-learning.qmd
20250803 - shrinkage

@@
 
+accuracyIndices_onePredictor_train <- tune::collect_metrics(cv_results)
+
+rsquare_onePredictor_train <- round(as.numeric(accuracyIndices_onePredictor_train[accuracyIndices_onePredictor_train$.metric == "rsq", c("mean")]), digits = 2)
+
 accuracyIndices_onePredictor <- petersenlab::accuracyOverall(
@@
 ```
+
+There was modest [shrinkage](#sec-shrinkage) from the training model to the test model: the $R^2$ for the model on the training data was `{r} rsquare_onePredictor_train`; the $R^2$ for the same model applied to the test data was `{r} rsquare_onePredictor`.
 
@@
 
+accuracyIndices_multiplePredictors_train <- tune::collect_metrics(cv_results)
+
+rsquare_multiplePredictors_train <- round(as.numeric(accuracyIndices_multiplePredictors_train[accuracyIndices_multiplePredictors_train$.metric == "rsq", c("mean")]), digits = 2)
+
 accuracyIndices_multiplePredictors <- petersenlab::accuracyOverall(
@@
 ```
+
+There was modest [shrinkage](#sec-shrinkage) from the training model to the test model: the $R^2$ for the model on the training data was `{r} rsquare_multiplePredictors_train`; the $R^2$ for the same model applied to the test data was `{r} rsquare_multiplePredictors`.
 
@@
 
+accuracyIndices_lasso_train <- tune::collect_metrics(cv_results)
+
+rsquare_lasso_train <- accuracyIndices_lasso_train %>% 
+  filter(.metric == "rsq" & penalty == best_penalty$penalty) %>% 
+  select(mean) %>% 
+  pull() %>% 
+  as.numeric() %>% 
+  round(., 2)
+
 accuracyIndices_lasso <- petersenlab::accuracyOverall(
@@
 ```
+
+There was modest [shrinkage](#sec-shrinkage) from the training model to the test model: the $R^2$ for the model on the training data was `{r} rsquare_lasso_train`; the $R^2$ for the same model applied to the test data was `{r} rsquare_lasso`.
 
@@
 
+accuracyIndices_ridge_train <- tune::collect_metrics(cv_results)
+
+rsquare_ridge_train <- accuracyIndices_ridge_train %>% 
+  filter(.metric == "rsq" & penalty == best_penalty$penalty) %>% 
+  select(mean) %>% 
+  pull() %>% 
+  as.numeric() %>% 
+  round(., 2)
+
 accuracyIndices_ridge <- petersenlab::accuracyOverall(
@@
 ```
+
+There was modest [shrinkage](#sec-shrinkage) from the training model to the test model: the $R^2$ for the model on the training data was `{r} rsquare_ridge_train`; the $R^2$ for the same model applied to the test data was `{r} rsquare_ridge`.
 
@@
 
+accuracyIndices_elasticNet_train <- tune::collect_metrics(cv_results)
+
+rsquare_elasticNet_train <- accuracyIndices_elasticNet_train %>% 
+  filter(.metric == "rsq" & penalty == best_penalty$penalty & mixture == best_penalty$mixture) %>% 
+  select(mean) %>% 
+  pull() %>% 
+  as.numeric() %>% 
+  round(., 2)
+
 accuracyIndices_elasticNet <- petersenlab::accuracyOverall(
@@
 ```
+
+There was modest [shrinkage](#sec-shrinkage) from the training model to the test model: the $R^2$ for the model on the training data was `{r} rsquare_elasticNet_train`; the $R^2$ for the same model applied to the test data was `{r} rsquare_elasticNet`.
 
@@
 
+accuracyIndices_randomForest_train <- tune::collect_metrics(cv_results)
+
+rsquare_randomForest_train <- accuracyIndices_randomForest_train %>% 
+  filter(.metric == "rsq" & mtry == best_penalty$mtry & min_n == best_penalty$min_n) %>% 
+  select(mean) %>% 
+  pull() %>% 
+  as.numeric() %>% 
+  round(., 2)
+
 accuracyIndices_randomForest <- petersenlab::accuracyOverall(
@@
 ```
+
+There was modest [shrinkage](#sec-shrinkage) from the training model to the test model: the $R^2$ for the model on the training data was `{r} rsquare_randomForest_train`; the $R^2$ for the same model applied to the test data was `{r} rsquare_randomForest`.
 

### 2025-08-03 multiple-regression.qmd
20250803 - shrinkage

@@
 It allows obtaining a more accurate estimate of the unique contribution of a given [predictor variable](#sec-correlationalStudy), by controlling for other variables ([covariates](#sec-covariates)).
+By including multiple [predictor variables](#sec-correlationalStudy) in prediction of the [outcome variable](#sec-correlationalStudy), it also allows improved prediction accuracy.
 
@@
 Our parameter estimate for the regression coefficient, plus or minus 1 standard error, reflects the 68% confidence interval for the coefficient.
-The 95% confidence interval is computed as the parameter estimated plus or minus 1.96 standard errors.
+The 95% confidence interval is computed as the parameter estimated plus or minus 1.96 standard errors (because in a standard normal distribution, the middle 95% of the distribution lies between −1.96 and +1.96).
 For instance, if the parameter estimate for the regression coefficient is 0.50, and the standard error is 0.10, the 95% confidence interval is [0.30, 0.70]: $0.5 - (1.96 \times 0.10) = 0.3$; $0.5 + (1.96 \times 0.10) = 0.7$.
@@
 To examine whether residuals are normally distributed, we can examine quantile–quantile (QQ) plots and probability–probability (PP) plots.
+QQ plots depict quantiles of a sample distribution (y-axis) compared to the quantiles of a theoretical (in this case, normal) distribution (x-axis).
+PP plots depict cumulative probabilities of a sample distribution (y-axis) compared to those of a theoretical (in this case, normal) distribution (x-axis).
 QQ plots are particularly useful for identifying deviations from normality in the tails of the distribution; PP plots are particularly useful for identifying deviations from normality in the center of the distribution.
@@
 So, even though an overfitting statistical model appears to be accurate because it is accounting for more variance, it is not actually that accurate—it will predict new data less accurately than how accurately it accounts for the data with which the model was built.
+Overfitting is most likely to occur when the model is too complex relative to the sample size (e.g., many [predictor variables](#sec-correlationalStudy) or parameters)
 In the case of fantasy football analytics, this is especially relevant because there are hundreds if not thousands of variables we could consider for inclusion and many, many players when considering historical data.
@@
 If we were to apply the same algorithm based on the `r as.integer(format(Sys.Date(), "%Y")) - 1` prediction model to `r as.integer(format(Sys.Date(), "%Y"))` data, the prediction model would likely predict less accurately than with `r as.integer(format(Sys.Date(), "%Y")) - 1` data.
-The regression coefficients tend to become weaker when applied to new data, a phenomenon called [shrinkage](#sec-shrinkage), which is described in @sec-shrinkage.
+The regression coefficients (and resulting accuracy) tend to become weaker when applied to new data, a phenomenon called [shrinkage](#sec-shrinkage), which is described in @sec-shrinkage.
+For instance, shrinking is observed when applying multiple regression models to new data in @sec-fittingModels-regression and when applying [machine learning](#sec-machineLearning) models to new data in @sec-fittingModels-machineLearning.
 The regression coefficients in the [FILL IN]

### 2025-08-03 machine-learning.qmd
20250803 - shrinkage

@@
 
+#### Evaluate Accuracy of Model on Training Data {#sec-treeBoostingModelAccuracyTraining}
+
+```{r}
+# Model-implied predictions
+pred_train_qb <- predict(
+  gp_model_fit_qb,
+  data = data_train_qb_matrix[, pred_vars_qb],
+  group_data_pred = data_train_qb_matrix[, "gsis_id"],
+  group_rand_coef_data_pred = cbind(
+    data_train_qb_matrix[, "ageCentered20"],
+    data_train_qb_matrix[, "ageCentered20Quadratic"]),
+  predict_var = FALSE,
+  pred_latent = FALSE)
+
+petersenlab::accuracyOverall(
+  predicted = pred_train_qb[["response_mean"]],
+  actual = data_train_qb_matrix[, "fantasyPoints_lag"],
+  dropUndefined = TRUE
+)
+```
+
 #### Evaluate Accuracy of Model on Test Data {#sec-treeBoostingModelAccuracy}
@@
 
+accuracyIndices_treeBoosting_train <- petersenlab::accuracyOverall(
+  predicted = pred_train_qb[["response_mean"]],
+  actual = data_train_qb_matrix[, "fantasyPoints_lag"],
+  dropUndefined = TRUE
+)
+
+rsquare_treeBoosting_train <- round(accuracyIndices_treeBoosting_train$rsquared, digits = 2)
+
 accuracyIndices_treeBoosting <- petersenlab::accuracyOverall(
@@
 ```
+
+There was moderate [shrinkage](#sec-shrinkage) from the training model to the test model: the $R^2$ for the model on the training data was `{r} rsquare_treeBoosting_train`; the $R^2$ for the same model applied to the test data was `{r} rsquare_treeBoosting`.
 

### 2025-08-03 multiple-regression.qmd
20250803 - remove placeholder
20250803 - types of missingness

@@
 For instance, shrinking is observed when applying multiple regression models to new data in @sec-fittingModels-regression and when applying [machine learning](#sec-machineLearning) models to new data in @sec-fittingModels-machineLearning.
-The regression coefficients in the [FILL IN]
 
@@
 Multiple regression in `R` using the `lm()` function applies listwise deletion.
-*Listwise deletion* removes any row (in the data file) from analysis that has a missing value on the [outcome variable](#sec-correlationalStudy) or any of the [predictor variables](#sec-correlationalStudy).
+*Listwise deletion* (also called complete case analysis) removes any row (in the data file) from analysis that has a missing value on the [outcome variable](#sec-correlationalStudy) or any of the [predictor variables](#sec-correlationalStudy).
 Removing all rows from analysis that have any missingness in the model variables can be a problem because missingness is often not completely at random—missingness often occurs systematically (i.e., for a reason).
@@
 
+Pairwise deletion (also called available case analysis) uses all available data for each pair of variables when estimating the covariances, and the resulting covariance matrix can be used to estimate a multiple regression model.
 We can estimate a regression model that uses pairwise deletion using the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
@@
 
+Multiple imputation takes a data set with missingness and it uses available information on other variables to estimate what likely values could have been for the missing value.
+It repeats the imputation multiple times, so there are multiple imputed data sets, which can give a sense of the degree of uncertainty of each imputed value.
+The model is then fit to each of the multiply imputed data sets separately, and the results are combined.
 We can multiply impute data using the `mice` package [@vanBuuren2011_packages; @R-mice].
@@
 
+Full information maximum likelihood (FIML) estimates model parameters using all available data for each case, without imputing missing values.
+FIML is commonly used for handling missingness in approaches to latent variable modeling, such as [factor analysis](#sec-factorAnalysis) and structural equation modeling.
 We can estimate a regression model that uses full information maximum likelihood using the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
@@
 
-## Impact of Outliers {#sec-multipleRegressionOutliers}
-
-[As with correlation](#sec-correlationOutliers), multiple regression can be strongly impacted by outliers.
+## Impact of Outliers and Restricted Range {#sec-multipleRegressionOutliers}
+
+[As with correlation](#sec-correlation), multiple regression can be strongly impacted by [outliers](#sec-correlationOutliers) and [restricted range](#sec-correlationRestrictedRange).
 

### 2025-08-04 multiple-regression.qmd
20250803 - edits

@@
 
-Only the quadratic term for target share (not height or weight) shows a significant association.
+Only target share (not height or weight) shows a significant association, including its linear and quadratic terms.
 
@@
   = &`r petersenlab::apa(beta0, decimals = 2)` + `r petersenlab::apa(beta1, decimals = 2)` \cdot `r petersenlab::apa(age, decimals = 0)` + `r petersenlab::apa(beta2, decimals = 2)` \cdot `r petersenlab::apa(age, decimals = 0)`^2 + `r petersenlab::apa(beta3, decimals = 2)` \cdot `r petersenlab::apa(height, decimals = 0)` + `r petersenlab::apa(beta4, decimals = 2)` \cdot `r petersenlab::apa(height, decimals = 0)`^2 + \\
-  &`r petersenlab::apa(beta5, decimals = 2)` \cdot `r petersenlab::apa(weight, decimals = 0)` + `r petersenlab::apa(beta6, decimals = 2)` \cdot `r petersenlab::apa(weight, decimals = 0)`^2 + `r petersenlab::apa(beta7, decimals = 2)` \cdot `r petersenlab::apa(target_share, decimals = 0)` + `r petersenlab::apa(beta8, decimals = 2)` \cdot `r petersenlab::apa(target_share, decimals = 0)`^2 + \epsilon \\
+  &`r petersenlab::apa(beta5, decimals = 2)` \cdot `r petersenlab::apa(weight, decimals = 0)` + `r petersenlab::apa(beta6, decimals = 2)` \cdot `r petersenlab::apa(weight, decimals = 0)`^2 + `r petersenlab::apa(beta7, decimals = 2)` \cdot `r petersenlab::apa(target_share, decimals = 1)` + `r petersenlab::apa(beta8, decimals = 2)` \cdot `r petersenlab::apa(target_share, decimals = 1)`^2 + \epsilon \\
   = &`r petersenlab::apa(predictedFantasyPoints, decimals = 2)`

### 2025-08-04 mixed-models.qmd
20250717 - Conclusion sections
20250804 - data structures

@@
 # Mixed Models {#sec-mixedModels}
+
+This chapter provides an overview of mixed models.
 
@@
 As we will discuss, it is important to account for levels of nesting when the observations are nonindependent.
+Mixed models provide a framework for accounting for levels of nesting, in order to account for why data from the same upper-level unit are likely to be intercorrelated.
 
@@
 Examining the association between age as a fixed effect in relation to fantasy performance would examine the association between a player's age and their fantasy performance while holding the association between age and fantasy performance constant across all players.
-That is, it would assume that all players show the same trajectory such as increase for 4 years then decrease.
+That is, it would assume that all players show the same trajectory such as increase from ages 20 to 24 and then decrease.
 Examining the association between age as a random effect in relation to fantasy performance would examine the association between a player's age and their fantasy performance while allowing the association between age and fantasy performance to vary across players.
@@
 As described in @sec-fallaciesEcological, the ecological fallacy is the error of drawing inferences about an individual from group-level data.
-An type of ecological fallacy is [Simpson's paradox](#sec-simpsonsParadox).
+A type of ecological fallacy is [Simpson's paradox](#sec-simpsonsParadox).
 
@@
 It is important to evaluate how much variance is attributable to the upper-level unit.
+A way of evaluating how much variance is attributed to an upper-level unit is with the intraclass correlation coefficient (ICC).
 If substantial variance is attributable to the upper-level unit, it is important to account for the nested structure of the data.
 Mixed models are a useful approach to account for data with a nested structure so you can avoid committing the ecological fallacy.
+
+## Data Structure {#sec-mixedModelDataStructure}
+
+When estimating mixed models, it is important to know the structure of the data.
+Examples of data from nonindependent units include nested data, multiple membership data, and cross-classified data, as depicted in @fig-nestedDataStructures.
+
+::: {#fig-nestedDataStructures}
+![](images/nestedData.png){fig-alt="Various Data Structures Involving Data from Nonindependent Units, Including Nested data, Multiple Membership Data, and Cross-Classified Data."}
+
+Various Data Structures Involving Data from Nonindependent Units, Including Nested data, Multiple Membership Data, and Cross-Classified Data.
+:::
+
+One example of nonindependent data is if your data are *nested*.
+When data are collected from multiple lower-level units (e.g., people) in an upper-level unit (e.g., married couple, family, classroom, school, neighborhood), and each lower-level unit belongs to one and only one higher-level unit, the data from the lower-level units are considered nested within the upper-level unit.
+For example, multiple participants in your sample may come from the same classroom.
+Data from multiple people can be nested within the same family, classroom, school, etc.
+Longitudinal data can also be considered nested data, in which time points are nested within the participant (i.e., the same participant provides an observation across multiple time points).
+And if you have multiple informants of a child's behavior (e.g., parent-, teacher-, friend-, and self-report), the ratings could also be nested within the participant (i.e., there are ratings from multiple informants of the same child).
+
+Another form of nonindependent data is when data involve *multiple membership*.
+Data involve multiple membership when the lower-level units belong to more than one upper-level unit.
+As an example of multiple membership, children may have more than one teacher, and therefore, in a modeling sense, children "belong" to more than one teacher cluster.
+
+Another form of nonindependent data is when data are *cross-classified* (also called crossed).
+Data are cross-classified when the lower-level units are classified by two or more upper-level units, but the upper-level units are not hierarchical or nested within one another.
+For instance, children may be nested within the crossing of schools and neighborhoods.
+That is, children are nested within schools, and children are also nested within neighborhoods; however, children attending the same school may not necessarily be from the same neighborhood.
+The data would still have a two-level hierarchy, but the data would be nested in specific school-neighborhood combinations.
+That is, children are nested within the cross-classification of schools and neighborhoods.
+
+Applied to longitudinal fantasy football data, Most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players.
+However, you could get more complicated.
+For instance, you could estimate a three-level model of timepoints nested within players nested within teams.
+This would help account for players's productivity benefitting or taking a hit from being on a given team, and it would account for players changing teams.
+Alternatively, consistent with @Usami2018, you could estimate a cross-classified model of players' individual observations/measurement occasions belonging to two upper-level units: players and time points (e.g., week 1, week 2, 2025 season, etc.).
+This would help account for systematic effects that are specific to a given timepoint (e.g., week or season).
 

### 2025-08-04 mixed-models.qmd
20250804 - comparing model fit

@@
 
-## Data Structure {#sec-mixedModelDataStructure}
+### Data Structure {#sec-mixedModelDataStructure}
 
@@
 This would help account for systematic effects that are specific to a given timepoint (e.g., week or season).
+
+### Comparing Model Fit {#sec-mixedModelCompareModels}
+
+When trying to determine what predictors to include and how complex of a model to fit, it can be helpful to compare a variety a models to determine how well they fit the data (i.e., how much variance they account for).
+To compare non-nested models, you can use fit criteria such as Akaike information criterion (AIC), which penalizes model complexity.
+If two models are nested, you can compare their model fit using a likelihood ratio test.
+Two models are considered nested if one model includes all of the terms of the original model along with additional terms.
+The simpler model is considered "nested within" the more complex model.
+In the likelihood ratio test, a significant difference indicates that the more complex model is significantly better fitting than the simpler model.
+If the likelihood ratio test is non-significant, it indicates that the more complex model is not significantly better fitting, and thus you should prefer the more parsimonious model.
 
@@
 
-We compare models using the `bbmle` [@R-bbmle] and `MuMIn` [@R-MuMIn] packages.
+We compare nested models using an $F$-test with the `anova()` function.
+We compare non-nested models using the `bbmle` [@R-bbmle] and `MuMIn` [@R-MuMIn] packages.
 
@@
 anova(
+  pointsPerSeason_nullModel,
   pointsPerSeason_linearRegression,
@@
   pointsPerSeason_randomIntercepts,
-  pointsPerSeason_position
-)
-
-anova(
+  pointsPerSeason_position,
   pointsPerSeason_positionAgeFixedLinearSlopes,
@@
 
-##### Generalized Additive Model {#sec-fantasyPointsByAgeExperienceModelsAgeGAM}
+#### Generalized Additive Model {#sec-fantasyPointsByAgeExperienceModelsAgeGAM}
 
@@
 
-###### Compare Models {#sec-fantasyPointsByAgeExperienceModelsAgeGAMcompare}
+##### Compare Models {#sec-fantasyPointsByAgeExperienceModelsAgeGAMcompare}
 

### 2025-08-04 mixed-models.qmd
20250804 - Bayesian mixed model plots

@@
 )
-```
-
-Plots of model-implied fantasy points by position and age are in Figures [-@fig-fantasyPointsByAgeQuadratic]–[-@fig-fantasyPointsByAgeGAMdepth].
+
+# From Bayesian Model: All Players
+model_levels <- levels(bayesianMixedModelFit$data$player_idFactor)
+
+pointsPerSeason_positionAge_newData$player_idFactor <- factor(
+  rep(levels(player_stats_seasonal_offense_subset$player_idFactor)[1], nrow(pointsPerSeason_positionAge_newData)), 
+  levels = model_levels
+)
+
+modelPredictions_bayesian <- predict(
+  object = bayesianMixedModelFit,
+  newdata = pointsPerSeason_positionAge_newData,
+  re_formula = NA,
+  summary = TRUE
+)
+
+pointsPerSeason_positionAge_newData$fantasyPoints_bayesian <- modelPredictions_bayesian[,"Estimate"]
+```
+
+Plots of model-implied fantasy points by position and age are in Figures [-@fig-fantasyPointsByAgeQuadratic]–[-@fig-fantasyPointsByAgeBayesian].
 
@@
 
+#### Bayesian Mixed Model {#sec-fantasyPointsByAgePlotsBayesian}
+
+```{r}
+#| label: fig-fantasyPointsByAgeBayesian
+#| fig-cap: "Plot of Implied Trajectories of Fantasy Points by Age from a Bayesian Mixed Model."
+#| fig-alt: "Plot of Implied Trajectories of Fantasy Points by Age from a Bayesian Mixed Model."
+#| code-fold: true
+
+ggplot2::ggplot(
+  data = pointsPerSeason_positionAge_newData,
+  mapping = aes(
+    x = age,
+    y = fantasyPoints_bayesian,
+    color = positionFactor
+  )
+) + 
+  geom_smooth() +
+  labs(
+    x = "Player Age (years)",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by Player Age and Position",
+    subtitle = "Bayesian Mixed Model with All Players",
+    color = "Position"
+  ) +
+  theme_classic() +
+  guides(color = guide_legend(override.aes = list(fill = NA))) # transparent legend background
+```
+
 ### Plots of Individuals' Model-Implied Fantasy Points by Age {#sec-fantasyPointsIndividualByAgePlots}
+
+We provide plots of individuals' model-implied fantasy points by age from the Bayesian mixed model in @sec-timeSeriesBayesianModelImpliedPredictions.
 

### 2025-08-04 mixed-models.qmd
20250804 - tweaks

@@
 
-We will discuss a modeling framework that goes by many terms, including mixed models, mixed-effects models, multilevel models, hierarchical linear models.
+Mixed models are simply regression models that account for the nested or hierarchical structure of the data.
+Mixed models can be be used to address the [assumption in multiple regression](#sec-assumptionsRegression) that the residuals are independent (i.e., uncorrelated).
+When data come from the same group, that can lead to residuals being correlated, so accounting for the grouping-level structure can address the assumption.
+
+The modeling framework goes by many terms, including mixed models, mixed-effects models, multilevel models, hierarchical linear models.
 They are sometimes called multilevel models and hierarchical linear models, whose name emphasizes the hierarchical structure of the data because the data are nonindependent.
@@
 ```{r}
+#| label: generate-model-implied-predictions-fantasy-points-by-position-age
+
 # From Quadratic Model: All Players
@@
 pointsPerSeason_positionAge_newData$player_idFactor <- factor(
-  rep(levels(player_stats_seasonal_offense_subset$player_idFactor)[1], nrow(pointsPerSeason_positionAge_newData)), 
+  rep(model_levels[1], nrow(pointsPerSeason_positionAge_newData)), 
   levels = model_levels

### 2025-08-05 mixed-models.qmd
20250804 - tweaks
20250804 - leverage multiple cores

@@
   re_formula = NA,
-  summary = TRUE
+  cores = num_cores
 )

### 2025-08-05 mixed-models.qmd
20250804 - clear memory

--- 2025-08-04 mixed-models.qmd
@@
 )
+```
+
+```{r}
+#| label: generate-model-implied-predictions-fantasy-points-by-position-age-bayesian1
 
@@
 )
+```
+
+```{r}
+#| label: clear-memory-bayesian1
+
+gc()
+```
+
+```{r}
+#| label: generate-model-implied-predictions-fantasy-points-by-position-age-bayesian2
 
@@
   re_formula = NA,
-  summary = TRUE
-)
+  cores = num_cores
+)
+```
+
+```{r}
+#| label: clear-memory-bayesian2
+
+gc()
+```
+
+```{r}
+#| label: generate-model-implied-predictions-fantasy-points-by-position-age-bayesian3
 

### 2025-08-05 mixed-models.qmd
20250804 - hide chunks
20250805 - fix

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/bayesianMixedModelFit.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/modelPredictions_bayesian.RData", fsep = ""))
 ```
@@
 ```{r}
-#| label: clear-memory-bayesian1
-
-gc()
-```
-
-```{r}
 #| label: generate-model-implied-predictions-fantasy-points-by-position-age-bayesian2
+#| eval: false
 
@@
 ```{r}
-#| label: clear-memory-bayesian2
-
-gc()
+#| eval: false
+#| include: false
+
+save(
+  modelPredictions_bayesian,
+  file = "./data/modelPredictions_bayesian.RData")
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/modelPredictions_bayesian.RData", fsep = ""))
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/modelPredictions_bayesian.RData")
 ```

### 2025-08-05 causal-inference.qmd
20250708 - track chunk execution times
20250722 - chapter intro
20250805 - updates

@@
+{{< include _chunk-timing.qmd >}}
+
 # Causal Inference {#sec-causalInference}
+
+This chapter provides an overview of principles of causal inference and [causal diagrams](#sec-causalDiagrams).
 
@@
 
-A causal effect is an effect in which the level of one input variable results in different levels on the outcome variable.
+A causal effect is the difference in an outcome that is directly attributable to different levels or values of an input variable.
+It reflects the difference in the outcome that would be observed for the same unit under two or more different levels of the input variable, holding all other relevant factors constant.
 For instance, for a given player, the extent to which consumption of sports drink influences their performance reflects the difference between how well the player would perform if they consume sports drink compared to how well they would perform if they do not consume sports drink.
@@
 For instance, if sports drink consumption occurs only once, and it occurs only before and not after performance, then we have ruled out the possibility of reverse causation (i.e., that better performance causes players to consume sports drink).
+To help establish the first criterion, [experiments](#sec-causalInferenceExperiment) and [longitudinal designs](#sec-causalInferenceLongitudinal) are useful.
 
@@
 The third criterion involves ruling out alternative reasons why the purported cause and effect may be related.
-As noted in @sec-correlationCausality, there are four reasons why `X` may be correlated with `Y`.
+As noted in @sec-correlationCausality, there are five reasons why `X` may be correlated with `Y`.
 If we meet the first criterion of causality, we have removed the possibility that `Y` causes `X` (i.e., reverse causality).
@@
 The control group is the group players that drank only water.
-Now, consider that the players in the treatment group outperform the players in the the control group in their football game.
+Now, consider that the players in the treatment group outperform the players in the control group in their football game.
 In such a study, we observe what *did happen* when players received a treatment.
@@
 Second, through randomly assigning players to consume or not consume sports drink, this holds everything else constant (so long as the groups are evenly distributed according to other factors, such as their age, weight, etc.) and thus removes the possibility that third variable [confounds](#sec-causalDiagramConfounding) explain the effect of the [independent variable](#sec-experiment) on the [dependent variable](#sec-experiment).
+However, not everything is able to, or ethical to, be manipulated.
+In addition, as described in @sec-internalExternalValidityTradeoffs, experimental designs tend to have lower capacity for [external validity](#sec-externalValidity) than [correlational designs](#sec-correlationalStudy)—that is, compared to findings from [correlational designs](#sec-correlationalStudy), the findings from experimental designs tend to have lower capacity to be consistent with how things play out in the real world.
 
@@
 Another design feature of [longitudinal designs](#sec-causalInferenceLongitudinal) that can lead to greater [internal validity](#sec-internalValidity) is the use of within-subject analyses.
-Between-subject analyses, might examine, for instanc, whether players who consume more sports drink perform better on average compared to players who consume less sports drink.
+Between-subject analyses, might examine, for instance, whether players who consume more sports drink perform better on average compared to players who consume less sports drink.
 However, there are other between-person differences that could explain any observed between-subject associations between sports drink consumption and players performance.
@@
 For instance, you could examine whether, within the same individual, if a player consumes a sports drink, do they perform better compared to games in which they did not consume a sports drink.
-When we control for prior levels of the outcome in the prediction, we are evaluating whether the predictor is associated with witin-person *change* in the outcome.
+When we control for prior levels of the outcome in the prediction, we are evaluating whether the predictor is associated with within-person *change* in the outcome.
 Predicting within-person change provides stronger evidence consistent with causality because it uses the individual as their own control and controls for many time-invariant [confounds](#sec-causalDiagramConfounding) (i.e., [confounds](#sec-causalDiagramConfounding) that do not change across time).
@@
 
-However, in observational studies, the groups may differ on the [confounding factors](#sec-causalDiagramConfounding); people may choose the level of the they receive.
+However, in observational studies, the groups may differ on the [confounding factors](#sec-causalDiagramConfounding); for instance, due to other factors (e.g., personality, beliefs, and attitudes), people may choose (or self-select into) the level of the variable (e.g., sport drink consumption) they receive.
 Thus, we need to attempt to control for the [confounds](#sec-causalDiagramConfounding); however, in an observational study, we can only potentially adjust for the observed [confounds](#sec-causalDiagramConfounding), unlike in an [experiment](#sec-causalInferenceExperiment) in which we can control for both observed and unobserved [confounds](#sec-causalDiagramConfounding).
@@
 
+Genetically informed designs evaluate the extent to which individuals who share greater genetic relatedness are more similar in a characteristic compared to individuals who share less genetic relatedness.
 For instance, researchers may compare monozygotic twins versus dizygotic twins in some outcome—a so-called "twin study".
@@
 In general, one should not control for descendants of the outcome variable, because doing so could eliminate the apparent effect of a legitimate cause on the outcome variable [@Digitale2022].
-Moreover, as described above, when trying to understand the total causal effect of a predictor variable on an outcome variable, one should not control for descendants of the predictor variable that are also antecedents of the outcome variable (i.e., mediators of the effect of the predictor variable on the outcome variable) [@Digitale2022].
+Moreover, as described above, when trying to understand the total causal effect of a predictor variable on an outcome variable, one should not control for descendants of the predictor variable that are also ancestors of the outcome variable (i.e., mediators of the effect of the predictor variable on the outcome variable) [@Digitale2022].
 
@@
 Causal diagrams can be a useful tool for identifying the proper variables to control for (and those not to control for).
+In general, the only variables you should control for are [confound(s)](#sec-causalDiagramConfounding).
 When [confounding](#sec-causalDiagramConfounding) exists, it is important to control for the [confound(s)](#sec-causalDiagramConfounding).

### 2025-08-05 base-rates.qmd
20250805 - heuristics, biases, and fallacies

@@
 That is, if the base rate of an event or condition—such as schizophrenia—is low (e.g., ~0.5%), people overestimate the likelihood that a person has schizophrenia when given specific information about the person such as their symptoms and history.
+Moreover, @Fox1998a asked National Basketball Association (NBA) fans to estimate the probability that each of 8 teams in the playoff would win the playoffs.
+The median sum of probability judgments for the eight teams was 240% [@Fox1998a; @Kahneman2011] even though, in reality, the true probabilities must sum to 100%.
+The finding suggests that people were evaluating each team individually, considering the reasons why that particular team could win, without properly accounting for the total probability.
 
@@
 In sum, the [marginal probability](#sec-baseRate), including the [prior probability](#sec-baseRate) or [base rate](#sec-baseRate), should be weighed heavily in predictions unless there are sufficient data to indicate otherwise, i.e., to update the posterior probability based on new evidence.
+People tend to ignore base rates (i.e., [base rate neglect](#sec-fallaciesBaseRate)).
+In general, people tend to a) overestimate the likelihood of low base-rate events and b) overweight low base-rate events in their decisions [@Kahneman2011].
 Bayes' theorem specifies how prior beliefs (i.e., [base rate](#sec-baseRate) informations) should be integrated with the [predictive accuracy](#sec-predictiveValidity) of the evidence to make predictions.

### 2025-08-05 cognitive-bias.qmd
20250805 - heuristics, biases, and fallacies

@@
 [*Cognitive biases*](#sec-cognitiveBiases) are systematic errors in thinking.
-[*Fallacies*](#sec-fallacies) are forms of flawed reasoning.
+[*Fallacies*](#sec-fallacies) are forms of flawed reasoning (invoked as an argument).
+Fallacies are a characteristic of an argument, whereas cognitive biases are characteristics of the reasoning agent.
+Thus, [fallacies](#sec-fallacies) can result from [heuristics](#sec-heuristics) and [cognitive biases](#sec-cognitiveBiases).
 
-Below, we provide examples of [heuristics](#sec-heuristics), [*cognitive biases*](#sec-cognitiveBiases), and [*fallacies*](#sec-fallacies).
-These are not exhaustive lists of all the [heuristics](#sec-heuristics), [*cognitive biases*](#sec-cognitiveBiases), and [*fallacies*](#sec-fallacies).
+Below, we provide examples of [heuristics](#sec-heuristics), [cognitive biases](#sec-cognitiveBiases), and [fallacies](#sec-fallacies).
+These are not exhaustive lists of all the [heuristics](#sec-heuristics), [cognitive biases](#sec-cognitiveBiases), and [fallacies](#sec-fallacies).
 They are some examples of key ones that are relevant fantasy football.

### 2025-08-06 getting-started.qmd
20250806 - long and wide form

@@
 
-## Transform Data from Long to Wide {#sec-longToWide}
+## Long and Wide Form {#sec-longWideForm}
 
@@
 
-Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable):
+### Transform Data from Long to Wide {#sec-longToWide}
+
+Below, we transform the data from long form to wide form.
+The long data are in "`player`-`season`-`week` form".
+We widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable):
 
@@
 
-## Transform Data from Wide to Long {#sec-wideToLong}
+### Transform Data from Wide to Long {#sec-wideToLong}
 

### 2025-08-06 mixed-models.qmd
20250806 - long and wide form

@@
 
-When estimating mixed models, it is important to know the structure of the data.
+For fitting mixed models, the data should be in [long form](#sec-longWideForm)—each player (or unit of interest) should have multiple rows based on the grouping structure.
+When estimating mixed models, it is important to know the grouping structure of the data because the data structure influences the model syntax.
 Examples of data from nonindependent units include nested data, multiple membership data, and cross-classified data, as depicted in @fig-nestedDataStructures.

### 2025-08-06 cognitive-bias.qmd
20250806 - tweaks and regression to the mean

@@
 [*Cognitive biases*](#sec-cognitiveBiases) are systematic errors in thinking.
+[Cognitive biases](#sec-cognitiveBiases) can result from [heuristics](#sec-heuristics).
 [*Fallacies*](#sec-fallacies) are forms of flawed reasoning (invoked as an argument).
-Fallacies are a characteristic of an argument, whereas cognitive biases are characteristics of the reasoning agent.
+[Fallacies](#sec-fallacies) are a characteristic of an argument, whereas [cognitive biases](#sec-cognitiveBiases) are characteristics of the reasoning agent.
 Thus, [fallacies](#sec-fallacies) can result from [heuristics](#sec-heuristics) and [cognitive biases](#sec-cognitiveBiases).
+
+One might wonder why [fallacies](#sec-fallacies) and [cognitive biases](#sec-cognitiveBiases) exist, given that errors in thinking and reasoning would seem to be disadvantageous.
+However, [fallacies](#sec-fallacies) and [cognitive biases](#sec-cognitiveBiases) reflect the tradeoffs between speed and accuracy in thinking and, in some cases, may also serve to help people feel better about themselves.
 
@@
 [Regression effects](#sec-fallaciesRegression) are discussed further in @sec-fallaciesRegression.
+As another example, the first number or players named in a negotiation can serve as an anchor.
 
@@
 1. *overplacement* of one's performance relative to others
-1. *overprecision* in one's beliefs/jugments/predictions
+1. *overprecision* in one's beliefs/judgments/predictions
 
@@
 Everything seems to make sense in hindsight [@Kahneman2011].
-People tend to remember the succeess of their predictions and forget the failures of their predictions.
+People tend to remember the successes of their predictions and forget the failures of their predictions.
 For instance, if a third-string Quarterback has a breakout game, a fantasy manager may claim that they "knew it all along" that the player was going to breakout, despite not having picked up the player.
@@
 One way that self-serving bias can commonly manifest is the tendency people to attribute successes to internal factors—such as one's character, ability, effort, or actions—but to attribute failures to external factors beyond one's control—such as others, unfairness, bad luck, etc.
-That is, self-serving bias involves claiming more reponsibility for one's success than for one's failure [@Miller2013c].
+That is, self-serving bias involves claiming more responsibility for one's success than for one's failure [@Miller2013c].
 For instance, a manager who wins the league may attribute their success to their fantasy football skill and smart decisions (rather than luck).
@@
 Loss aversion is a key aspect of prospect theory and suggests that individuals tend to be biased in favor of their reference situation and toward making smaller rather than larger changes [@Kahneman2011].
-Loss aversion explains why people tend to favor the favoring the status quo; the negative consequences of a change often loom larger than the positive consequences of a change [@Kahneman2011].
+Loss aversion explains why people tend to favor the status quo; the negative consequences of a change often loom larger than the positive consequences of a change [@Kahneman2011].
 Aversion to losses and potential threats may be related to humans' survival instinct [@Kahneman2011].
@@
 The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
-The, a "low base-rate" event is an event with a low probability of occurring (i.e., an unlikely event).
+Thus, a "low base-rate" event is an event with a low probability of occurring (i.e., an unlikely event).
 Focusing on the specific information about the event (and thus underweighting the base rate) involves the [representativeness heuristic](#sec-heuristicsRepresentativeness) [@Kahneman2011]—making judgments based on how similar the event is to a mental prototype regardless of how frequent the event the event actually is.
@@
 The regression fallacy is the failure to account for the fact that things tend to naturally fluctuate around their mean and that, after an extreme fluctuation, subsequent scores tend to regress (or reverse) to the mean.
+Regression to the mean is depicted in @fig-regressionToMean.
 Regression to the mean occurs because performance—both success and failure—is influenced by talent and by luck [@Kahneman2011].
@@
 Despite peoples' tendency to ascribe such changes to causal phenomena, regression-related changes do not need a causal explanation—they are a mathematically inevitable consequence that arises from the fact that an outcome is influenced, to some degree, both by one's level on the construct and by error [@Kahneman2011].
+
+```{r}
+#| label: fig-regressionToMean
+#| fig-cap: "Regression to the Mean."
+#| fig-alt: "Regression to the Mean."
+#| code-fold: true
+
+mydata <- data.frame(
+  week = 1:7,
+  fantasyPoints = c(25,7,12,5,11,15,3)
+)
+
+ggplot(
+  data = mydata,
+  mapping = aes(
+    x = week,
+    y = fantasyPoints
+  )
+) + 
+  geom_point(
+    size = 4
+  ) +
+  geom_line(
+    linewidth = 1
+  ) +
+  geom_abline(
+    intercept = 10,
+    slope = 0,
+    linetype = "dashed"
+  ) +
+  scale_x_continuous(
+    breaks = 1:7
+  ) +
+  labs(
+    y = "Fantasy Points",
+    title = "Regression to the Mean"
+  ) +
+  annotate(
+    "segment",
+    x = 6.5,
+    xend = 7,
+    y = 17,
+    yend = 10,
+    color = "blue",
+    linewidth = 1.5,
+    alpha = 0.6,
+    arrow = arrow()) +
+  annotate(
+    "text",
+    x = 5.5,
+    y = 18,
+    label = "Average ability",
+    hjust = 0, # left-justify
+    size = 5) +
+  theme_classic(base_size = 30) +
+  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
+```
 

### 2025-08-06 cognitive-bias.qmd
20250806 - tweaks

@@
 Applied to fantasy football, consider that a Quarterback has had several lousy games in a row.
-The gambler's fallacy might lead a manager to start the player under the belief that the player "is due" for a big game, expecting a strong performance from the player merely because they player has not had a good game in a while.
+The gambler's fallacy might lead a manager to start the player under the belief that the player "is due" for a big game, expecting a strong performance from the player merely because the player has not had a good game in a while.
 
@@
 Although people tend to find stories and anecdotes compelling, they are among the weakest forms of evidence, as described in @sec-internalExternalValidityTradeoffs.
-Consider that a fantasy football commentator writes an article about how they won their league by drafting a Quarterback and Wide Receiver by drafting with their first two picks, respectively.
+Consider that a fantasy football commentator writes an article about how they won their league by drafting a Quarterback and Tight End with their first two picks, respectively.
 The anecdotal fallacy might lead a manager who reads the article to believe that they are more likely to win their league if they follow a similar strategy, despite stronger evidence otherwise.
+It is important to be skeptical of anecdotes and personal experience.
 
@@
 The *narrative fallacy* is the error of creating a story with cause-and-effect explanations from random events.
-According to @Kahneman2011, people find stories more compelling if they a) are simple (as opposed to complex), b) are concrete (as opposed to abstract), c) provide a greater attribution to talent, stupidity, and intentions (compared to luck), and d) focus on a few salient events that occured (rather than the many events that failed to occur).
+According to @Kahneman2011, people find stories more compelling if they a) are simple (as opposed to complex), b) are concrete (as opposed to abstract), c) provide a greater attribution to talent, stupidity, and intentions (compared to luck), and d) focus on a few salient events that occurred (rather than the many events that failed to occur).
 For instance, consider that a player undergoes a challenging start to the season; however, they have higher-performing games in the middle of the season.

### 2025-08-06 actuarial.qmd
20250806 - additions

@@
 Hundreds of studies have examined clinical versus actuarial prediction methods across many disciplines.
-Findings consistently show that actuarial methods are as [accurate](#sec-validity) or more [accurate](#sec-validity) than human judgment/prediction methods.
+Findings consistently show that actuarial methods are as [accurate](#sec-validity) or more [accurate](#sec-validity) than human judgment/prediction methods [@Dawes1989; @Aegisdottir2006; @Baird2000; @Grove1996; @Grove2000].
 "There is no controversy in social science that shows such a large body of qualitatively diverse studies coming out so uniformly...as this one" [@Meehl1986, pp. 373–374].
+In general, actuarial methods tend to be about 10% more accurate than human judgment [@Grove2000].
+Thus, actuarial methods have [incremental validity](#sec-incrementalValidity) over human judgment.
+As noted earlier, the superiority of actuarial approaches to clinical judgment tends to hold even when the clinician is given more information than the actuarial approach [@Dawes1989].
+In addition, actuarial predictions outperform human judgment even when the human is given the result of the actuarial prediction [@Kahneman2011].
+Allowing clinicians to override actuarial predictions consistently leads to lower predictive accuracy [@Garb2019].
 

### 2025-08-06 actuarial.qmd
20250806 - addition and misspelling

@@
 
-As an example of using the actuarial approach, Billy Beane, who was the general manager of the Oakland Athletics at the time, wanted to find ways for his team—which had less finanical resources than its competitors—to compete with teams that hard more money to sign players.
+As an example of using the actuarial approach, Billy Beane, who was the general manager of the Oakland Athletics at the time, wanted to find ways for his team—which had less financial resources than its competitors—to compete with teams that hard more money to sign players.
 Because the team did not have the resources to sign the best players, they had to find to find other ways to find the optimal players that they could afford.
@@
 A statistical formula cannot rescue inaccurate assessment data.
+Likewise, if the input data are based on measurement strategies that are biased against particular individuals or groups, the actuarial formula is not immune from bias.
 

### 2025-08-07 actuarial.qmd
20250807 - additions and tweaks

--- 2025-08-03 actuarial.qmd
@@
 
-As an example of using the actuarial approach, Billy Beane, who was the general manager of the Oakland Athletics at the time, wanted to find ways for his team—which had less finanical resources than its competitors—to compete with teams that hard more money to sign players.
+As an example of using the actuarial approach, Billy Beane, who was the general manager of the Oakland Athletics at the time, wanted to find ways for his team—which had less financial resources than its competitors—to compete with teams that had more money to sign players.
 Because the team did not have the resources to sign the best players, they had to find to find other ways to find the optimal players that they could afford.
@@
 Or, the clinical psychologist might use both human judgment and an actuarial approach independently to see whether they agree.
-That is, the clinician may make a prediction based on their judgment and might also generate a prediction from an actuarial approach.
+That is, the clinician or expert may make a prediction based on their judgment and might also generate a prediction from an actuarial approach.
 
@@
 One situation where human judgment would be important is when no actuarial method exists for the judgment or prediction.
-For instance, when no actuarial method exists for the diagnosis or disorder (e.g., suicide), it is up to the clinician.
-However, we could collect data on the outcomes or on clinicians' judgments to develop an actuarial method that will be more reliable than the clinicians' judgments.
-That is, an actuarial method developed based on clinicians' judgments will be more accurate than clinicians' judgments.
+For instance, when no actuarial method exists for the diagnosis or disorder (e.g., suicide), it is up to the clinician/expert.
+However, we could collect data on the outcomes or on experts' judgments to develop an actuarial method that will be more reliable than the experts' judgments.
+That is, an actuarial method developed based on experts' judgments will be more accurate than experts' judgments.
 In other words, we do not necessarily need outcome data to develop an actuarial method.
-We could use the client's data as predictors of the clinicians' judgments to develop a structured approach to prediction that weighs factors similarly to clinicians, but with more [reliable](#sec-reliability) predictions.
+We could use the player's data as predictors of the experts' judgments to develop a structured approach to prediction that weighs factors similarly to experts, but with more [reliable](#sec-reliability) predictions.
 
@@
 Another situation in which human judgment could be preferable is if advanced, complex theories exist.
-Computers have a difficult time adhering to complex theories, so clinicians may be better suited.
+Computers have a difficult time adhering to complex theories, so clinicians/experts may be better suited.
 However, we do not have any of these complex theories in psychology that are accurate.
@@
 
-If the prediction requires complex configural relations that a computer will have a difficult time replicating, a clinician's judgment may be preferred.
+If the prediction requires complex configural relations that a computer will have a difficult time replicating, an expert's judgment may be preferred.
 Although the likelihood that a person can accurately work through these complex relations is theoretically possible, it is highly unlikely.
@@
 Hundreds of studies have examined clinical versus actuarial prediction methods across many disciplines.
-Findings consistently show that actuarial methods are as [accurate](#sec-validity) or more [accurate](#sec-validity) than human judgment/prediction methods.
+Despite the romantic notion that professional scouts and experts being able to make accurate predictions (and identify strong players), findings consistently show that actuarial methods are as [accurate](#sec-validity) or more [accurate](#sec-validity) than human judgment/prediction methods [@Dawes1989; @Aegisdottir2006; @Baird2000; @Grove1996; @Grove2000].
 "There is no controversy in social science that shows such a large body of qualitatively diverse studies coming out so uniformly...as this one" [@Meehl1986, pp. 373–374].
+In general, actuarial methods tend to be about 10% more accurate than human judgment [@Grove2000].
+Thus, actuarial methods have [incremental validity](#sec-incrementalValidity) over human judgment.
+As noted earlier, the superiority of actuarial approaches to clinical judgment tends to hold even when the expert is given more information than the actuarial approach [@Dawes1989].
+In addition, actuarial predictions outperform human judgment even when the human is given the result of the actuarial prediction [@Kahneman2011].
+Allowing clinicians/experts to override actuarial predictions consistently leads to lower predictive accuracy [@Garb2019].
 
@@
 
-Despite actuarial methods being generally more accurate than human judgment, judgment is much more widely used by clinicians.
+Despite actuarial methods being generally more accurate than human judgment, judgment is much more widely used by clinicians/experts.
 There are several reasons why actuarial methods have not caught on; one reason is professional traditions.
@@
 
-Some clinicians do not think that group averages (e.g., in terms of which treatment is most effective) apply to an individual client.
+Some experts do not think that group averages (e.g., in terms of which treatment is most effective) apply to an individual client or player.
 This invokes the distinction between nomothetic (group-level) inferences and idiographic (individual-level) inferences.
@@
 However, one can always create an algorithm of the experts' judgments, even if one does not have access to the outcome information.
-A model of clinicians' responses tends to be more [accurate](#sec-validity) than clinicians' judgments themselves because the model gives the same outcome with the same input data—i.e., it is perfectly [reliable](#sec-reliability).
+A model of experts' responses tends to be more [accurate](#sec-validity) than experts' judgments themselves because the model gives the same outcome with the same input data—i.e., it is perfectly [reliable](#sec-reliability).
 
@@
 Knowing how people would perform in different conditions is the same thing as good understanding.
+
+Another potential reason clinicians may not use actuarial judgment is time.
+This can be a legitimate challenge to using actuarial models in some contexts.
+Actuarial methods can take time to apply—they require assessing the releving criteria, putting the assessment scores into an actuarial formula, and interpreting the output score.
+In some clinical situations, such as risk assessments in terms of whether someone is imminently suicidal or homicidal and how to proceed, clinicians must make immediate in-session decisions @Jacinto2018.
+Thus, it can be challenging or impractical in some situations to conduct all of the relevant assessments using actuarial approaches.
+Moreover, in some workplace settings, such as health care settings, clinicians may be pushed beyond their limits in terms of productivity expectations.
+In these situations, one possible approach is to exercise their best possible judgment based on the available evidence [i.e., intuitive decision-making processes; @Jacinto2018], but to check it against an actuarial model after the fact.
+It is valuable to evaluate to receive feedback on the accuracy of one's judgments and, when time permits, to evaluate one's intuitive decision-making against analytical decision-making.
 

### 2025-08-07 actuarial.qmd
20250807 - time as constraint

@@
 This can be a legitimate challenge to using actuarial models in some contexts.
-Actuarial methods can take time to apply—they require assessing the releving criteria, putting the assessment scores into an actuarial formula, and interpreting the output score.
-In some clinical situations, such as risk assessments in terms of whether someone is imminently suicidal or homicidal and how to proceed, clinicians must make immediate in-session decisions @Jacinto2018.
-Thus, it can be challenging or impractical in some situations to conduct all of the relevant assessments using actuarial approaches.
+Actuarial methods can take time to apply—they require assessing the relevant criteria, putting the assessment scores into an actuarial formula, and interpreting the output score.
+In some clinical situations, such as risk assessments in terms of whether someone is imminently suicidal or homicidal and how to proceed, clinicians must make immediate in-session decisions [@Jacinto2018].
+Thus, it can be challenging or impractical in some situations to conduct all of the relevant assessments needed to apply actuarial approaches.
 Moreover, in some workplace settings, such as health care settings, clinicians may be pushed beyond their limits in terms of productivity expectations.
-In these situations, one possible approach is to exercise their best possible judgment based on the available evidence [i.e., intuitive decision-making processes; @Jacinto2018], but to check it against an actuarial model after the fact.
-It is valuable to evaluate to receive feedback on the accuracy of one's judgments and, when time permits, to evaluate one's intuitive decision-making against analytical decision-making.
+In these situations, one possible approach is to exercise one's best possible judgment in the moment based on the available evidence [i.e., intuitive decision-making processes; @Jacinto2018], but to check it against an actuarial model after the fact.
+This is like comparing your decisions in a chess match, after the game ends, to how a computer would have moved in each of the situations.
+It is valuable to receive feedback on the accuracy of one's judgments and, when time permits, to evaluate one's intuitive decision-making against analytical decision-making.
+However, there are structured approaches to assessing suicide risk that are quick to administer [e.g., @Posner2011].
+
+Nevertheless, the lack of time is perhaps less of a concern in the sports domain.
+Teams have lots of data readily available on many players.
+And, for player selection, teams have considerable time to implement statistical models of how players will perform to inform their selections.
+There is less time during games to make in-game decisions and adjustments from complicated models; that said, teams are gaining greater adeptness with making in-game adjustments and decisions based on analytics.
+Teams use actuarial charts that inform, for instance, [when to go for it on fourth down](#sec-coachingGoingForIt).
+Now that many teams are motivated to use analytics to gain a leg up on their competitors, they are finding ways to make the time for it.
+Clinicians and other medical experts, however, still have considerable room for improvement in terms of their adoption of actuarial approaches to judgment and prediction.
 

### 2025-08-07 actuarial.qmd
20250807 - tweaks

@@
 Moreover, in some workplace settings, such as health care settings, clinicians may be pushed beyond their limits in terms of productivity expectations.
-In these situations, one possible approach is to exercise one's best possible judgment in the moment based on the available evidence [i.e., intuitive decision-making processes; @Jacinto2018], but to check it against an actuarial model after the fact.
+In these situations, one possible approach is to exercise one's best possible judgment in the moment based on the available evidence [i.e., intuitive decision-making processes; @Jacinto2018], and to check it against an actuarial model after the fact.
 This is like comparing your decisions in a chess match, after the game ends, to how a computer would have moved in each of the situations.
@@
 
-There are various challenges of data-driven approaches.
+There are various challenges of data-driven approaches (e.g., machine learning).
 First, they are sometimes not interpretable or consistent with theory.
-Second, they tend to [overfit](#sec-overfitting) the data.
+Sometimes that may because are theories are wrong; other times, it may mean we need to refine our data-driven approach.
+Second, data-driven approaches tend to [overfit](#sec-overfitting) the data.
 [Overfitting](#sec-overfitting) is described in @sec-overfitting.
-Third, as a result of [overfitting](#sec-overfitting) the data, they tend to show [shrinkage](#sec-shrinkage).
+Third, as a result of [overfitting](#sec-overfitting) the data, data-driven approaches tend to show [shrinkage](#sec-shrinkage).
 

### 2025-08-08 cognitive-bias.qmd
20250808 - regression to the mean

@@
 #| fig-alt: "Regression to the Mean."
+#| fig-width: 10
+#| fig-height: 6
+#| fig-align: center
 #| code-fold: true
@@
   week = 1:7,
-  fantasyPoints = c(25,7,12,5,11,15,3)
+  fantasyPoints = c(33,7,12,5,11,15,3)
 )
@@
   geom_abline(
+    aes(
       intercept = 10,
       slope = 0,
-    linetype = "dashed"
+      linetype = "Average ability"
+    ),
   ) +
+  scale_linetype_manual(values = "dashed") +
+  coord_cartesian(
+    ylim = c(0, NA)) +
   scale_x_continuous(
@@
     y = "Fantasy Points",
-    title = "Regression to the Mean"
+    #title = "Regression to the Mean",
+    linetype = NULL
   ) +
@@
     "segment",
-    x = 6.5,
-    xend = 7,
-    y = 17,
-    yend = 10,
+    x = 2,
+    xend = 1.75,
+    y = 19,
+    yend = 15,
     color = "blue",
@@
     "text",
-    x = 5.5,
-    y = 18,
-    label = "Average ability",
+    x = 2,
+    y = 20,
+    label = "Regression toward the mean",
     hjust = 0, # left-justify
@@
   theme_classic(base_size = 30) +
-  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
+  theme(
+    axis.title.y = element_text(angle = 0, vjust = 0.5),
+    legend.position = "inside",
+    legend.position.inside = c(0.8, 1))
 ```

### 2025-08-08 draft.qmd
20250808 - fantasy points by end-of-season position rank and position

@@
 ```{r}
+library("tidyverse")
+```
 
+### Load Data {#sec-draftLoadData}
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+```
+
+```{r}
+load(file = "./data/player_stats_seasonal.RData")
+load(file = "./data/player_stats_weekly.RData")
 ```
@@
 
+## Draft Analysis {#sec-draftAnalysis}
+
+### Fantasy Points by (End-of-Season) Position Rank and Position {#sec-expectedPointsByPositionRank}
+
+```{r}
+fantasyPointsByPositionRank <- player_stats_seasonal %>% 
+  filter(position_group %in% c("QB","RB","WR","TE","K","DL","LB","DB")) %>% 
+  group_by(position_group, season) %>% 
+  arrange(position_group, season, -fantasyPoints) %>% 
+  mutate(positionRank = row_number())
+```
+
+A plot of fantasy points by end-of-season position rank and position is depicted in @fig-expectedPointsByPositionRank1.
+
+```{r}
+#| label: fig-expectedPointsByPositionRank1
+#| fig-cap: "Fantasy Points by End-of-Season Position Rank and Position."
+#| fig-alt: "Fantasy Points by End-of-Season Position Rank and Position."
+#| code-fold: true
+
+ggplot(
+  data = fantasyPointsByPositionRank %>% filter(positionRank <= 200),
+  mapping = aes(
+    x = positionRank,
+    y = fantasyPoints
+  )
+) + 
+  geom_point(
+    alpha = 0.3
+  ) +
+  geom_smooth() +
+  coord_cartesian(
+    ylim = c(0,NA),
+    expand = FALSE) + # don't add space between axes and data
+  labs(
+    x = "End-of-Season Position Rank",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by End-of-Season Position Rank"
+  ) +
+  theme_bw() +
+  facet_wrap(vars(position_group)) # facet by position_group
+```
+
+In @fig-expectedPointsByPositionRank2, we overlay the smoothed best-fit line for each position onto the same graphing area for direct comparison.
+
+```{r}
+#| label: fig-expectedPointsByPositionRank2
+#| fig-cap: "Fantasy Points by End-of-Season Position Rank and Position."
+#| fig-alt: "Fantasy Points by End-of-Season Position Rank and Position."
+#| code-fold: true
+
+palette_OkabeIto_black <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")
+
+ggplot(
+  data = fantasyPointsByPositionRank %>% filter(positionRank <= 200),
+  mapping = aes(
+    x = positionRank,
+    y = fantasyPoints,
+    color = position_group
+  )
+) + 
+  geom_smooth(
+    linewidth = 1.5,
+    se = FALSE) +
+  coord_cartesian(
+    ylim = c(0,NA), # set limits of y-axis
+    expand = FALSE) + # don't add space between axes and data
+  scale_color_manual(
+    values = palette_OkabeIto_black
+  ) +
+  labs(
+    x = "End-of-Season Position Rank",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by End-of-Season Position Rank",
+    color = "Position Group"
+  ) +
+  theme_classic()
+```
+
+In @fig-expectedPointsByPositionRank3, we limit the position rank to ranks less than or equal to 30, for easier comparison of positions' comparative dropoff at the top position ranks.
+
+```{r}
+#| label: fig-expectedPointsByPositionRank3
+#| fig-cap: "Fantasy Points by End-of-Season Position Rank and Position."
+#| fig-alt: "Fantasy Points by End-of-Season Position Rank and Position."
+#| code-fold: true
+
+ggplot(
+  data = fantasyPointsByPositionRank %>% filter(positionRank <= 30),
+  mapping = aes(
+    x = positionRank,
+    y = fantasyPoints,
+    color = position_group
+  )
+) + 
+  geom_smooth(
+    linewidth = 1.5,
+    se = FALSE) +
+  coord_cartesian(
+    ylim = c(0,425), # set limits of y-axis
+    expand = FALSE) + # don't add space between axes and data
+  scale_color_manual(
+    values = palette_OkabeIto_black
+  ) +
+  labs(
+    x = "End-of-Season Position Rank",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by End-of-Season Position Rank",
+    subtitle = "Position Rank \u2264 30", # \u2264: <= symbol
+    color = "Position Group"
+  ) +
+  theme_classic()
+```
+
+It appears that Running Backs show the fastest dropoff (i.e., steepest negative slope), followed by Tight Ends, Quarterbacks, Wide Receivers, and then defensive positions.
+So, identifying the top-scoring Running Back can be of great benefit.
+
+However, the style of gameplay has changed.
+@fig-expectedPointsByPositionRank4 depicts the same data since 2020.
+
+```{r}
+#| label: fig-expectedPointsByPositionRank4
+#| fig-cap: "Fantasy Points by End-of-Season Position Rank and Position."
+#| fig-alt: "Fantasy Points by End-of-Season Position Rank and Position."
+#| code-fold: true
+
+ggplot(
+  data = fantasyPointsByPositionRank %>% filter(positionRank <= 30 & season >= 2020),
+  mapping = aes(
+    x = positionRank,
+    y = fantasyPoints,
+    color = position_group
+  )
+) + 
+  geom_smooth(
+    linewidth = 1.5,
+    se = FALSE) +
+  coord_cartesian(
+    ylim = c(0,425), # set limits of y-axis
+    expand = FALSE) + # don't add space between axes and data
+  scale_color_manual(
+    values = palette_OkabeIto_black
+  ) +
+  labs(
+    x = "End-of-Season Position Rank",
+    y = "Fantasy Points (Season)",
+    title = "Fantasy Points (Season) by End-of-Season Position Rank",
+    subtitle = "Position Rank \u2264 30, Since 2020", # \u2264: <= symbol
+    color = "Position Group"
+  ) +
+  theme_classic()
+```
+
+Although Running Backs appear to show the fastest dropoff, Quarterbacks and Wide Receivers have appeared to overtake them in terms of fantasy points.
+
+@fig-pointsBySeasonByPosition depicts the mean fantasy points of the top 10 players at each position by season.
+
+```{r}
+#| label: fig-pointsBySeasonByPosition
+#| fig-cap: "Mean Fantasy Points of Top 10 Players at Each Position by Season."
+#| fig-alt: "Mean Fantasy Points of Top 10 Players at Each Position by Season."
+#| code-fold: true
+
+meanFantasyPointsTop10ByPosition <- fantasyPointsByPositionRank %>% 
+  filter(positionRank <= 10) %>% 
+  summarise(
+    fantasyPoints = mean(fantasyPoints, na.rm = TRUE),
+    .groups = "drop_last"
+  )
+
+ggplot(
+  data = meanFantasyPointsTop10ByPosition,
+  mapping = aes(
+    x = season,
+    y = fantasyPoints,
+    color = position_group
+  )
+) + 
+  geom_line(
+    linewidth = 1.5) +
+  coord_cartesian(
+    ylim = c(0,425), # set limits of y-axis
+    expand = FALSE) + # don't add space between axes and data
+  scale_color_manual(
+    values = palette_OkabeIto_black
+  ) +
+  labs(
+    x = "Season",
+    y = "Mean Fantasy Points",
+    title = "Mean Fantasy Points of Top 10 Players at Each Position by Season",
+    color = "Position Group"
+  ) +
+  theme_classic()
+```
+
+The trend suggests that, since 2000, Quarterbacks have increased in fantasy points scored, and Running Backs have decreased in fantasy points scored.
+
 ## Conclusion {#sec-draftConclusion}

### 2025-08-08 draft.qmd
20250808 - tweaks

@@
 ```{r}
+#| code-fold: true
+
 fantasyPointsByPositionRank <- player_stats_seasonal %>% 
@@
 #| fig-alt: "Fantasy Points by End-of-Season Position Rank and Position."
+#| fig-height: 8
 #| code-fold: true
@@
   geom_point(
-    alpha = 0.3
+    alpha = 0.03
   ) +

### 2025-08-09 draft.qmd
20250809 - lineup optimization

@@
 library("tidyverse")
+library("Rglpk")
 ```
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
+load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/players_projections_seasonal.RData", fsep = ""))
 ```
@@
 load(file = "./data/player_stats_weekly.RData")
+load(file = "./data/players_projections_seasonal.RData")
 ```
@@
 
+FantasyFootballAnalytics.net provides tools to help you win your [snake draft](#sec-draftSnake): <https://fantasyfootballanalytics.net/win-your-snake-draft>.
+For draft tools for [snake drafts](#sec-draftSnake), see the Fantasy Football Analytics web application: <https://apps.fantasyfootballanalytics.net>.
+The web application allows users to specify their custom league settings to calculate projected points that are customized to their league and helps users select which players to draft that maximize the players' [value over replacement players](#sec-fantasyValueVORP) (VORP).
+
 ### Auction Draft {#sec-draftStrategyAuction}
@@
 Having some depth can help you offset the risk of injuries and the possibility that some of your starters may underperform their expectation, which is quite likely given how [challenging it is to predict fantasy performance](#sec-accuracyOfCrowd).
+
+FantasyFootballAnalytics.net provides tools to help you win your [auction draft](#sec-draftSnake) (<https://fantasyfootballanalytics.net/win-your-snake-draft>) and daily fantasy sports (DFS) leagues (<https://fantasyfootballanalytics.net/how-to/dfsoptimizerhelp>).
+For draft and player selection tools for [auction drafts](#sec-draftSnake) and DFS leagues, see the Fantasy Football Analytics web application: <https://apps.fantasyfootballanalytics.net>.
+The web application allows users to specify their custom league settings to calculate projected points that are customized to their league, and to identify the optimal lineup given their league settings and salary cap.
 
@@
 
+### Lineup Optimization {#sec-lineupOptimization}
+
+For [auction drafts](#sec-draftAuction) and daily fantasy sports (DFS) leagues, a key goal is to select the best lineup or team within the given cap.
+For instance, in an [auction drafts](#sec-draftAuction), you may have a \$200 cap and your goal is to draft the best team that you can with the \$200.
+Using a mixed-integer programming solver, we can identify the lineup with the most projected points given a particular salary cap.
+For projected points for a player, we use their average projected points across all available projection sources.
+For a player's estimated cost, we use their [Average Auction Value](#sec-fantasyValueADPAAV) (AAV), rounded up to the nearest integer.
+
+We first subset and prepare the data so we can perform the lineup optimization.
+
+```{r}
+mostRecentProjections <- players_projections_seasonal_average_merged %>% 
+  filter(season == max(season, na.rm = TRUE)) %>% 
+  filter(pos %in% c("QB","RB","WR","TE","K","DST")) %>% 
+  filter(avg_type == "average")
+
+# Convert NAs for fantasy points to zero
+mostRecentProjections$points[which(is.na(mostRecentProjections$points))] <- 0
+
+# Convert NA for AAV to $1
+mostRecentProjections$aav[which(is.na(mostRecentProjections$aav))] <- 1
+
+# Round up AAV to nearest integer
+mostRecentProjections$cost <- ceiling(mostRecentProjections$aav)
+
+mostRecentProjections$name <- paste(mostRecentProjections$first_name, mostRecentProjections$last_name)
+```
+
+Next, we set up the constraints.
+Our example is adapted from <https://stackoverflow.com/questions/15147398/optimize-value-with-linear-or-non-linear-constraints-in-r>.
+Our constraints were:
+
+1) 9 total players of the following number at each of following positions:
+- 1 Quarterback
+- 2 Running Backs
+- 2 Wide Receivers
+- 1 Tight End
+- 1 Flex: RB/WR/TE
+- 1 K
+- 1 DST
+
+and 2) salary cap of \$200.
+
+```{r}
+# number of players
+num.players <- nrow(mostRecentProjections)
+
+# 1 for every player
+total_players <- rep(1, num.players)
+
+# objective:
+#f <- pts
+
+# the variable are booleans
+var.types <- rep("B", num.players)
+
+# the constraints
+
+A <- rbind(
+  as.numeric(mostRecentProjections$pos == "QB"),                # num QB
+  as.numeric(mostRecentProjections$pos == "RB"),                # num RB
+  as.numeric(mostRecentProjections$pos == "WR"),                # num WR
+  as.numeric(mostRecentProjections$pos == "TE"),                # num TE
+  as.numeric(mostRecentProjections$pos == "K"),                 # num K
+  as.numeric(mostRecentProjections$pos == "DST"),               # num DST
+  as.numeric(mostRecentProjections$pos %in% c("RB","WR","TE")), # num flex
+  total_players,
+  mostRecentProjections$cost)                                   # player cost
+
+dir <- c(
+  "==",
+  ">=",
+  ">=",
+  ">=",
+  "==",
+  "==",
+  "==",
+  "==",
+  "<=")
+
+b <- c(
+  1,   # 1 QB
+  2,   # at least 2 RBs
+  2,   # at least 2 WRs
+  1,   # at least 1 TE
+  1,   # 1 K
+  1,   # 1 DST
+  6,   # 1 RB/WR/TE flex (i.e., 6 total RB/WR/TEs)
+  9,   # 9 total players
+  200) # salary cap of $200
+```
+
+The optimizer thus seeks to identify the lineup with the most projected points that meets the above constraints (within a salary cap of $200).
+We perform the optimization using the `Rglpk::Rglpk_solve_LP()` function of the `Rglpk` package [@R-Rglpk].
+
+```{r}
+sol <- Rglpk::Rglpk_solve_LP(
+  obj = mostRecentProjections$points,
+  mat = A,
+  dir = dir,
+  rhs = b,
+  types = var.types,
+  max = TRUE)
+```
+
+Here is the solution obtained by the solver:
+
+```{r}
+sol
+```
+
+A zero for `status` indicates that it found a solution:
+
+```{r}
+sol$status
+```
+
+Here is the maximum value obtained in terms of the sum of projected points for the optimal lineup:
+
+```{r}
+sol$optimum
+```
+
+Here is a vector indicating whether or not a given player was selected for the optimal lineup (0 = no; 1 = yes):
+
+```{r}
+sol$solution
+```
+
+Here is the optimal lineup that maximized the sum of projected points within the salary cap:
+
+```{r}
+mostRecentProjections[which(sol$solution == 1), c("name","pos","points")]
+```
+
+For a web application that includes a lineup optimizer for selecting the optimal lineup for [auction drafts](#sec-draftAuction) and for DFS leagues, see the Fantasy Football Analytics web application: <https://apps.fantasyfootballanalytics.net>.
+The web application allows users to specify their custom league settings, select players to draft or exclude, and exclude risky players, among other features.
+
 ## Conclusion {#sec-draftConclusion}

### 2025-08-09 index.qmd
20250805 - acknowledgment
20250805 - fix
20250806 - robust load
20250809 - lineup optimization

@@
 #| include: false
+
+# Objects loaded but not generated during GitHub Action run: betaRegression.RData, bayesianMixedModelFit.RData, modelPredictions_bayesian.RData, monteCarloSim_results.RData
 
@@
   "MASS", "data.table", "future", "future.apply", "SimDesign", "domir", 
-  "GGally",
+  "GGally","Rglpk",
   "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
@@
 I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.
+Finally, I thank my wife who, in addition to designing the book cover, is a constant source of love and support.
 

### 2025-08-09 getting-started.qmd
20250809 - Rglpk

@@
      "mice","miceadds","interactions","robustbase","ordinal","MASS",
-     "data.table","future","future.apply","SimDesign","domir","GGally"))
+     "data.table","future","future.apply","SimDesign","domir","GGally","Rglpk"))
    ```

### 2025-08-09 draft.qmd
20250809 - free tool

@@
 For draft tools for [snake drafts](#sec-draftSnake), see the Fantasy Football Analytics web application: <https://apps.fantasyfootballanalytics.net>.
-The web application allows users to specify their custom league settings to calculate projected points that are customized to their league and helps users select which players to draft that maximize the players' [value over replacement players](#sec-fantasyValueVORP) (VORP).
+The web application includes a free tool that allows users to specify their custom league settings to calculate projected points that are customized to their league and helps users select which players to draft that maximize the players' [value over replacement players](#sec-fantasyValueVORP) (VORP).
 
@@
 For draft and player selection tools for [auction drafts](#sec-draftSnake) and DFS leagues, see the Fantasy Football Analytics web application: <https://apps.fantasyfootballanalytics.net>.
-The web application allows users to specify their custom league settings to calculate projected points that are customized to their league, and to identify the optimal lineup given their league settings and salary cap.
+The web application includes a free tool that allows users to specify their custom league settings to calculate projected points that are customized to their league, and to identify the optimal lineup given their league settings and salary cap.
 

### 2025-08-09 draft.qmd
20250809 - clean up code

@@
 
-1) 9 total players of the following number at each of following positions:
+1. 9 total players of the following number at each of following positions:
     - 1 Quarterback
@@
     - 1 DST
-
-and 2) salary cap of \$200.
+1. salary cap of \$200.
 
@@
 
-# objective:
-#f <- pts
-
-# the variable are booleans
+# the variables are booleans
 var.types <- rep("B", num.players)
@@
 # the constraints
-
 A <- rbind(

### 2025-08-09 basic-statistics.qmd
20250809 - SimDesign simulation

@@
 library("psych")
+library("SimDesign")
 library("tidyverse")
@@
 
+##### Simulation {#sec-statisticalPowerOneWayANOVAsimulation}
+
 The power analysis code above assumes the groups are of equal size (i.e., a balanced design).
 If the design is unbalanced (i.e., there are different numbers of participants in each group), it may be necessary to conduct a power analysis via a simulation.
-Below is an example of evaluating the statistical power for detecting an effect unbalanced designs via simulation:
-
-```{r}
-nSim <- 1000 # number of simulations
-
-# Function to generate data and perform ANOVA
-simulate_anova <- function(nGroup1, nGroup2, f, alpha) {
+Below is an example of evaluating the statistical power for detecting an effect unbalanced designs via simulation using the `SimDesign` [@R-SimDesign; @Chalmers2020_packages] package.
+We discuss simulation more in @sec-simulation.
+
+###### 1. Design Matrix {#sec-statisticalPowerOneWayANOVAsimulationDesign}
+
+First, we set up the design matrix using the `SimDesign::createDesign()` function:
+
+```{r}
+Design <- SimDesign::createDesign(
+  nGroup1 = c(10),
+  nGroup2 = c(25, 50, 100),
+  f       = c(.10, .25, .40), # small, medium, large effect sizes
+  alpha   = 0.05
+)
+```
+
+###### 2. Generate {#sec-statisticalPowerOneWayANOVAsimulationGenerate}
+
+Second, we set up the `Generate()` function:
+
+```{r}
+Generate <- function(condition, fixed_objects = NULL) {
+  with(condition, {
     # Means for each group
@@
 
-  # Combine data
-  data <- data.frame(
+    data.frame(
       value = c(group1, group2),
-    group = factor(rep(c("Group1", "Group2"), c(nGroup1, nGroup2)))
+      group = factor(rep(c("Group1", "Group2"), 
+                         c(nGroup1, nGroup2)))
     )
-
-  # Perform ANOVA
-  aov_result <- aov(value ~ group, data = data)
+  })
+}
+```
+
+###### 3. Analyse {#sec-statisticalPowerOneWayANOVAsimulationAnalyse}
+
+Third, we set up the `Analyse()` function:
+
+```{r}
+Analyse <- function(condition, dat, fixed_objects = NULL) {
+  aov_result <- aov(value ~ group, data = dat)
   p_value <- summary(aov_result)[[1]][["Pr(>F)"]][1]
 
-  # Check if p-value is less than alpha
-  return(p_value < alpha)
+  data.frame(
+    p_value  = p_value,
+    rejectH0 = as.numeric(p_value < condition$alpha)
+  )
 }
-
-# Run simulations
-set.seed(52242) # for reproducibility
-powerSimulationOneWayAnova <- replicate(
-  nSim,
-  simulate_anova(
-    nGroup1 = 10,
-    nGroup2 = 25,
-    f = effectSize_f,
-    alpha = alpha))
-
-# Estimate power
-mean(powerSimulationOneWayAnova)
+```
+
+###### 4. Summarise {#sec-statisticalPowerOneWayANOVAsimulationSummarize}
+
+Fourth, we set up the `Summarise()` function:
+
+```{r}
+Summarise <- function(condition, results, fixed_objects = NULL) {
+  # Estimated power = proportion of rejections
+  Power <- mean(results$rejectH0)
+
+  data.frame(Power = Power)
+}
+```
+
+###### 5. Run the Simulation {#sec-statisticalPowerOneWayANOVAsimulationRun}
+
+Fifth, we run the simulation using the `SimDesign::runSimulation()` function:
+
+```{r}
+monteCarloSim_results <- SimDesign::runSimulation(
+  design       = Design,
+  replications = 1000,
+  generate     = Generate,
+  analyse      = Analyse,
+  summarise    = Summarise,
+  seed = genSeeds(Design, iseed = 52242), # for reproducibility
+  parallel = TRUE # for faster (parallel) processing
+)
+```
+
+###### Simulation Results {#sec-statisticalPowerOneWayANOVAsimulationResults}
+
+Here are the simulation results:
+
+```{r}
+monteCarloSim_results
 ```

### 2025-08-10 base-rates.qmd
20250809 - label functions

@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::pAgivenB()` function that estimates the probability of one event, $A$, given another event, $B$.
 
@@
 We can then estimate the marginal probability of the event, substititing in $P(R | \text{not } C)$, using the law of total probability.
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pA()` function that estimates the marginal probability of one event, $A$.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::pA()` function that estimates the marginal probability of one event, $A$.
 
@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pBgivenNotA()` function that estimates the probability of one event, $B$, given that another event, $A$, did not occur.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::pBgivenNotA()` function that estimates the probability of one event, $B$, given that another event, $A$, did not occur.
 
@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `pB()` function that estimates the marginal probability of one event, $B$.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::pB()` function that estimates the marginal probability of one event, $B$.
 
@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::posttestProbability()` function that estimates the posttest probability of an event, given the [pretest probability](#sec-baseRate) and the [likelihood ratio](#sec-diagnosticLikelihoodRatio), or given the [pretest probability](#sec-baseRate) and the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test.
 
@@
 A probability nomogram calculator is available from @Schwartz2006 at the following link: [http://araw.mede.uic.edu/cgi-bin/testcalc.pl](http://araw.mede.uic.edu/cgi-bin/testcalc.pl) (archived at <https://perma.cc/X8TF-7YBX>).
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from @Chekroud2017:
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::nomogrammer()` function that creates a nomogram plot using the [positive](#sec-positiveLikelihoodRatio) and [negative](#sec-negativeLikelihoodRatio) [likelihood ratio](#sec-diagnosticLikelihoodRatio) or using the [sensitivity](#sec-sensitivity) (SN) and [specificity](#sec-specificity) (SP) of the test, as adapted from @Chekroud2017:
 

### 2025-08-10 basic-statistics.qmd
20250809 - label functions

@@
 The Hodges-Lehmann statistic (aka pseudomedian) is computed as the median of all pairwise means, and it is also robust to outliers.
-The pseudomedian can be computed using the `DescTools` package [@R-DescTools].
+The pseudomedian can be computed using the `DescTools::HodgesLehmann()` function of the `DescTools` package [@R-DescTools].
 The pseudomedian of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_pseudomedian`.
@@
 The mode of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} exampleValues_mode`.
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `Mode()` function for computing the mode of a set of data.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::Mode()` function for computing the mode of a set of data.
 
@@
 
-- mean:
+- mean (`base::mean()`):
 
@@
 
-- median:
+- median (`stats::median()`):
 
@@
 
-- Hodges-Lehmann statistic (aka pseudomedian):
+- Hodges-Lehmann statistic (aka pseudomedian; `DescTools::HodgesLehmann()`):
 
@@
 
-- mode:
+- mode (`petersenlab::Mode()`):
 
@@
 
-- weighted mean:
+- weighted mean (`stats::weighted.mean()`):
 
@@
 
-- weighted median:
+- weighted median (`petersenlab::wthdquantile()`):
 
@@
 
-- standard deviation:
+- standard deviation (`stats::sd()`):
 
@@
 
-- variance:
+- variance (`stats::var()`):
 
@@
 
-- range:
+- range (`base::range()`):
 
@@
 
-- minimum and maximum:
+- minimum (`base::min()`) and maximum (`base::max()`):
 
@@
 
-- quantiles (in this case, the 10th and 90th quantiles):
+- quantiles (in this case, the 10th and 90th quantiles; `stats::quantile()`):
 
@@
 
-- weighted quantiles (in this case, the 10th and 90th quantiles):
+- weighted quantiles (in this case, the 10th and 90th quantiles; `petersenlab::wthdquantile()`):
 
@@
 
-- interquartile range (IQR):
+- interquartile range (IQR; `stats::IQR()`):
 
@@
 
-- median absolute deviation:
+- median absolute deviation (`stats::mad()`):
 
@@
 
-Below is R code to estimate each:
-
-- skewness:
+Below is R code to estimate each using the `psych::skew()` and `psych::kurtosis()` functions of the `psych` package [@R-psych]:
+
+- skewness (`psych::skew()`):
 
@@
 
-- kurtosis:
+- kurtosis (`psych::kurtosi()`):
 
@@
 
-You can easily convert a variable to a *z* score using the `scale()` function:
+You can easily convert a variable to a *z* score using the `base::scale()` function:
 
@@
 
-The area under a normal curve within one standard deviation of the mean is calculated below using the `pnorm()` function, which calculates the cumulative density function for a normal curve.
+The area under a normal curve within one standard deviation of the mean is calculated below using the `stats::pnorm()` function, which calculates the cumulative density function for a normal curve.
 
@@
 
-If you want to determine the *z* score associated with a particular percentile in a normal distribution, you can use the `qnorm()` function.
+If you want to determine the *z* score associated with a particular percentile in a normal distribution, you can use the `stats::qnorm()` function.
 For instance, the *z* score associated with the 37th percentile is:
@@
 
-One-tailed test:
+One-tailed test (`stats::qt()`):
 
@@
 
-Two-tailed test:
+Two-tailed test (`stats::qt()`):
 
@@
 
-One-tailed test:
+One-tailed test (`stats::qf()`):
 
@@
 
-Two-tailed test:
+Two-tailed test (`stats::qf()`):
 
@@
 
-Factor A (one-tailed test):
+Factor A (one-tailed test; `stats::qf()`):
 
@@
 
-Factor B (one-tailed test):
+Factor B (one-tailed test; `stats::qf()`):
 
@@
 
-Interaction (one-tailed test):
+Interaction (one-tailed test; `stats::qf()`):
 
@@
 
-Factor A (two-tailed test):
+Factor A (two-tailed test; `stats::qf()`):
 
@@
 
-Factor B (two-tailed test):
+Factor B (two-tailed test; `stats::qf()`):
 
@@
 
-Interaction (two-tailed test):
+Interaction (two-tailed test; `stats::qf()`):
 
@@
 
-One-tailed test:
+One-tailed test (`stats::qchisq()`):
 
@@
 
-Two-tailed test:
+Two-tailed test (`stats::qchisq()`):
 
@@
 
-Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance)):
+Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.t.test()`):
 
@@
 
-Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):
+Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.t.test()`):
 
@@
 
-Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):
+Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.t2n.test()`):
 
@@
 
-Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):
+Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.anova.test()`):
 
@@
 
-Solving for sample size per group needed (given [effect size](#sec-practicalSignificance), power, and [alpha level](#sec-statisticalSignificance)):
+Solving for sample size per group needed (given [effect size](#sec-practicalSignificance), power, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.anova.test()`):
 
@@
 
-Solving for the minimum detectable [effect size](#sec-practicalSignificance) (given sample size per group, power, and [alpha level](#sec-statisticalSignificance)):
+Solving for the minimum detectable [effect size](#sec-practicalSignificance) (given sample size per group, power, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.anova.test()`):
 
@@
 ```{r}
+#| output: false
+
 monteCarloSim_results <- SimDesign::runSimulation(
@@
 
-Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):
+Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance); `WebPower::wp.rmanova()`):
 
@@
 
-Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance)):
+Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size per group, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.r.test()`):
 
@@
 
-Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance)):
+Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.f2.test()`):
 
@@
 
-Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance)):
+Solving for statistical power achieved (given [effect size](#sec-practicalSignificance), sample size, and [alpha level](#sec-statisticalSignificance); `pwr::pwr.chisq.test()`):
 

### 2025-08-10 correlation.qmd
20250802 - documentation
20250809 - label functions

@@
 A partial correlation examines the association between variables after controlling for one or more variables.
-Below, we examine a partial [Pearson correlation](#sec-correlationExamplesPearson) between height and fantasy points after controlling for age using the the `partial.r()` function from the `psych` package [@R-psych].
+Below, we examine a partial [Pearson correlation](#sec-correlationExamplesPearson) between height and fantasy points after controlling for age using the the `psych::partial.r()` function from the `psych` package [@R-psych].
 
@@
   x = c("height","fantasyPoints"),
-  y = "age"
+  y = "age" # control variable(s)
 )
@@
 These estimates of nonlinear correlation only range from 0–1 (where 0 represents no association and 1 represents a strong association), so they are interpreted differently than a traditional correlation coefficient (which can be negative).
-We can estimate a distance correlation using the `correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].
+We can estimate a distance correlation using the `correlation::correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].
 
@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `cor.table()` function that generates a correlation matrix of variables, including the $r$ value, number of observations ($n$), asterisks denoting statistical significance, and $p$-value for each association.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::cor.table()` function that generates a correlation matrix of variables, including the $r$ value, number of observations ($n$), asterisks denoting statistical significance, and $p$-value for each association.
 The asterisks follow the following traditional convention for statistical significance: $^\dagger p < .1$; $^* p < .05$; $^{**} p < .01$; $^{***} p < .001$.
@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `partialcor.table()` function that generates a [partial correlation](#sec-correlationExamplesPartial) matrix.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::partialcor.table()` function that generates a [partial correlation](#sec-correlationExamplesPartial) matrix.
 Here is a [partial correlation](#sec-correlationExamplesPartial) matrix, controlling for the player's age:
@@
 
-We can depict a correlation matrix plot (correlogram) using the `ggcorr()` function of the `GGally` [@R-GGally] package, as shown in @fig-correlogram.
+We can depict a correlation matrix plot (correlogram) using the `GGally::ggcorr()` function of the `GGally` [@R-GGally] package, as shown in @fig-correlogram.
 
@@
 Another approach is the percentage bend correlation, which gives less weight to observations that are farther away from the median.
-We can estimate the each using the `correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].
+We can estimate the each using the `correlation::correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].
 

### 2025-08-10 data-visualization.qmd
20250806 - duplicate identifiers
20250809 - label functions

@@
 It is also important to use color schemes with distinguishable colors.
-Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `scale_color_brewer()` and `scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
-Good colorblind-friendly color schemes for binned (`_b`), continuous (`_c`), and discrete (i.e., categorical; `_d`) scales are available using the `scale_color_viridis_X()` and `scale_fill_viridis_X()` functions of the `ggplot2` package [@R-ggplot2] (where `X` is `b`, `c`, or `d`, depending on whether you want a binned, continuous, discrete scale).
+Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `ggplot2::scale_color_brewer()` and `ggplot2::scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
+Good colorblind-friendly color schemes for binned (`_b`), continuous (`_c`), and discrete (i.e., categorical; `_d`) scales are available using the `ggplot2::scale_color_viridis_X()` and `ggplot2::scale_fill_viridis_X()` functions of the `ggplot2` package [@R-ggplot2] (where `X` is `b`, `c`, or `d`, depending on whether you want a binned, continuous, discrete scale).
 There are a variety of resources for color schemes that are accessible to color-blind viewers:
@@
 
-The `coord_cartesian()` function can be used to specify the axis limits; however, this function is not required and the plot will apply sensible limits by default.
+The `ggplot2::coord_cartesian()` function can be used to specify the axis limits; however, this function is not required and the plot will apply sensible limits by default.
 In addition, there are optional settings you can specify or modify, such as [axis titles](#sec-scatterplotPlotLabels), [axis labels](#sec-scatterplotAxes), [faceting](#sec-faceting), and the [theme](#sec-scatterplotTheme).
@@
 
-You can see a full list of color names available using the `colors()` function:
+You can see a full list of color names available using the `grDevices::colors()` function:
 
@@
 
-You can use any color by specifying its values in red-green-blue (RGB) color space using the `rgb()` function, as in @fig-colorByRGB.
+You can use any color by specifying its values in red-green-blue (RGB) color space using the `grDevices::rgb()` function, as in @fig-colorByRGB.
 
@@
 
-We can specify multiple colors manually using the `scale_fill_manual()` function, as in @fig-multipleColorsManually.
+We can specify multiple colors manually using the `ggplot2::scale_fill_manual()` function, as in @fig-multipleColorsManually.
 
@@
 
-We can also specify the colors using a palette, such as by the ColorBrewer palette using the `scale_fill_brewer()` function (as in @fig-multipleColorsByPaletteColorBrewer) or the viridis palette using the `scale_fill_viridis_d()` function (as in @fig-multipleColorsByPaletteViridis).
+We can also specify the colors using a palette, such as by the ColorBrewer palette using the `ggplot2::scale_fill_brewer()` function (as in @fig-multipleColorsByPaletteColorBrewer) or the viridis palette using the `ggplot2::scale_fill_viridis_d()` function (as in @fig-multipleColorsByPaletteViridis).
 
@@
 
-We can specify the line types manually using the `scale_linetype_manual()` function as in @fig-lineTypeByGroupManually.
+We can specify the line types manually using the `ggplot2::scale_linetype_manual()` function as in @fig-lineTypeByGroupManually.
 
@@
 
-### Line Width {#sec-ggplot2SyntaxLineType}
+### Line Width {#sec-ggplot2SyntaxLineWidth}
 
@@
 
-We can specify the point shape manually using the `scale_shape_manual()` function as in @fig-pointShapeByGroupManually.
+We can specify the point shape manually using the `ggplot2::scale_shape_manual()` function as in @fig-pointShapeByGroupManually.
 
@@
 
-Second, we create the base layer of the plot using the `ggplot()` function of the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot1.
+Second, we create the base layer of the plot using the `ggplot2::ggplot()` function of the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot1.
 We specify the data object and the variables in the data object that are associated with the x- and y-axes:
@@
 
-Third, we create a scatterplot using the `geom_point()` function from the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot2:
+Third, we create a scatterplot using the `ggplot2::geom_point()` function from the `ggplot2` package [@R-ggplot2], as in @fig-scatterplot2:
 
@@
 
-After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages], as in @fig-scatterplot11.
+After creating our plot, we can make the plot interactive using the `plotly::ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages], as in @fig-scatterplot11.
 
@@
 
-After creating our plot, we can make the plot interactive using the `ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages].
+After creating our plot, we can make the plot interactive using the `plotly::ggplotly()` function from the `plotly` package [@R-plotly; @Sievert2020_packages].
 

### 2025-08-10 decision-making.qmd
20250809 - label functions

@@
 
-The [`petersenlab` package](https://cran.r-project.org/web/packages/petersenlab/index.html) [@R-petersenlab] has the `wisdomOfCrowd()` function that computes the overall accuracy of the crowd-averaged projections, including the bracketing rate of the individual projections.
+The [`petersenlab` package](https://cran.r-project.org/web/packages/petersenlab/index.html) [@R-petersenlab] has the `petersenlab::wisdomOfCrowd()` function that computes the overall accuracy of the crowd-averaged projections, including the bracketing rate of the individual projections.
 

### 2025-08-10 download-football-data.qmd
20250730 - updates
20250809 - label functions

@@
 
-To download play-by-play data from prior weeks and seasons, we can use the `load_pbp()` function of the `nflreadr` package [@R-nflreadr].
-We add a progress bar using the `with_progress()` function from the `progressr` package [@R-progressr] because it takes a while to run.
+To download play-by-play data from prior weeks and seasons, we can use the `nflreadr::load_pbp()` function of the `nflreadr` package [@R-nflreadr].
+We add a progress bar using the `progressr::with_progress()` function from the `progressr` package [@R-progressr] because it takes a while to run.
 @HoPBP provide a Data Dictionary for the play-by-play data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>
@@
 
-We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
+We can download historical week-by-week actual player statistics using the `nflreadr::load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
 @HoPlayerStats provide a Data Dictionary for statistics for offensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
@@
 
-A Data Dictionary for the variables is available in the `nfl_stats_variables` object that is returned when running the `calculate_stats()` function:
+A Data Dictionary for the variables is available in the `nfl_stats_variables` object that is returned when running the `nflfastR::calculate_stats()` function:
 
@@
 
-Second, we can compute the players' career statistics using the `calculate_stats()` function from the `nflfastR` package [@R-nflfastR].
+Second, we can compute the players' career statistics using the `nflfastR::calculate_stats()` function from the `nflfastR` package [@R-nflfastR].
 
@@
 
-We already load players' week-by-week statistics [above](#sec-downloadActualStats) using the `load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
-Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:
+We already load players' week-by-week statistics [above](#sec-downloadActualStats) using the `nflreadr::load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
+Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `nflfastR::calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:
 

### 2025-08-10 evaluating-prediction-accuracy.qmd
20250809 - lineup optimization
20250809 - label functions

@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `accuracyAtCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::accuracyAtCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables.
 
@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `accuracyAtEachCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables at each possible cutoff.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::accuracyAtEachCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables at each possible cutoff.
 
@@
 That is, each index of accuracy described in this section provides a single numerical index of accuracy that aggregates the accuracy across all possible cutoffs.
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `accuracyOverall()` function that computes many threshold-independent accuracy indices.
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::accuracyOverall()` function that computes many threshold-independent accuracy indices.
 
@@
 Note: MPE is undefined when one or more of the observed values equals zero, due to division by zero.
-The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+The `petersenlab::accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
 
@@
 Note: MAPE is undefined when one or more of the observed values equals zero, due to division by zero.
-The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+The `petersenlab::accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
 
@@
 Note: sMAPE is undefined when one or more of the individuals has a prediction–observed combination such that the sum of the absolute value of the predicted value and the absolute value of the observed value equals zero ($|\text{predicted}_i| + |\text{observed}_i|$), due to division by zero.
-The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+The `petersenlab::accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
 
@@
 When predicted or actual values are -1, this leads to $log(0)$, which is undefined.
-The `accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
+The `petersenlab::accuracyOverall()` function of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] provides the option in the function to drop undefined values so you can still generate an estimate of accuracy despite undefined values.
 
@@
 Larger values reflect greater accuracy.
-We can obtain standardized regression coefficients by standardizing the predictors and outcome using the `scale()` function in `R`.
+We can obtain standardized regression coefficients by standardizing the predictors and outcome using the `base::scale()` function in `R`.
 
@@
 FantasyFootballAnalytics.net also provides tools to evaluate the accuracy of historical projections [@FFA-accuracy]: <https://fantasyfootballanalytics.net/ffanalytics-web-app-accuracy-tab>.
+For instance, the web application allows evaluating the accuracy of historical projections:
+<https://apps.fantasyfootballanalytics.net>.
 
@@
 The 95% prediction interval of the observed value (i.e., the dashed red lines) is the interval within which we would expect that 95% of future observations would lie for a given predicted value.
-We calculate the 95% prediction interval based on a LOESS model, which allows nonlinear associations, using the `loess.sd()` function of the `msir` package [@R-msir; @Scrucca2011_packages].
+We calculate the 95% prediction interval based on a LOESS model, which allows nonlinear associations, using the `msir::loess.sd()` function of the `msir` package [@R-msir; @Scrucca2011_packages].
 The black diagonal line indicates the reference line with an intercept of zero and a slope of one.

### 2025-08-10 factor-analysis.qmd
20250806 - duplicate identifiers
20250809 - label functions

@@
 
-#### Specify Variables {#sec-pcaPrepareDataSpecifyVars}
+#### Specify Variables {#sec-factorAnalysisPrepareDataSpecifyVars}
 
@@
 
-#### Standardize Variables {#sec-pcaPrepareDataStandardize}
+#### Standardize Variables {#sec-factorAnalysisPrepareDataStandardize}
 
@@
 
-We generated the scree plot in @fig-efaScreePlot1 using the `fa.parallel()` function of the `psych` package [@R-psych].
+We generated the scree plot in @fig-efaScreePlot1 using the `psych::fa.parallel()` function of the `psych` package [@R-psych].
 The optimal coordinates and the acceleration factor attempt to operationalize the Cattell scree test: i.e., the "elbow" of the scree plot [@Ruscio2012].
@@
 
-We generated the scree plot in @fig-efaScreePlot2 using the `nScree()` and `plotnScree()` functions of the `nFactors` package [@R-nFactors].
+We generated the scree plot in @fig-efaScreePlot2 using the `nFactors::nScree()` and `nFactors::plotnScree()` functions of the `nFactors` package [@R-nFactors].
 
@@
 
-We generated the very simple structure (VSS) plots in Figures [-@fig-efaVSSPlot1] and [-@fig-efaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
+We generated the very simple structure (VSS) plots in Figures [-@fig-efaVSSPlot1] and [-@fig-efaVSSPlot2] using the `psych::vss()` and `psych::nfactors()` functions of the `psych` package [@R-psych].
 In addition to VSS plots, the output also provides additional criteria by which to determine the optimal number of factors, each for which lower values are better, including the Velicer minimum average partial (MAP) test, the Bayesian information criterion (BIC), the sample size-adjusted BIC (SABIC), and the root mean square error of approximation (RMSEA).
@@
 
-We fit EFA models using the `efa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
+We fit EFA models using the `lavaan::efa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
 
@@
 
-A path diagram of the three-factor EFA model in @fig-efaPathDiagram-3factor was created using the `lavaanPlot()` function of the `lavaanPlot` package [@R-lavaanPlot].
+A path diagram of the three-factor EFA model in @fig-efaPathDiagram-3factor was created using the `lavaanPlot::lavaanPlot()` function of the `lavaanPlot` package [@R-lavaanPlot].
 
@@
 
-To fit the ESEM model, we use the `sem()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
+To fit the ESEM model, we use the `lavaan::sem()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
 
@@
 
-To make the plot interactive for editing, you can use the `plot_lavaan()` function of the `lavaangui` package [@R-lavaangui; @Karch2025_packages]:
+To make the plot interactive for editing, you can use the `lavaangui::plot_lavaan()` function of the `lavaangui` package [@R-lavaangui; @Karch2025_packages]:
 
@@
 
-We fit CFA models using the `cfa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
+We fit CFA models using the `lavaan::cfa()` function of the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
 We compare one-, two-, and three-factor CFA models.

### 2025-08-10 getting-started.qmd
20250809 - label functions

--- 2025-08-06 getting-started.qmd
@@
      "mice","miceadds","interactions","robustbase","ordinal","MASS",
-     "data.table","future","future.apply","SimDesign","domir","GGally"))
+     "data.table","future","future.apply","SimDesign","domir","GGally","Rglpk",
+     "TTR"))
    ```
@@
 
-You can install `R` packages using the following syntax:
+You can install `R` packages using the `utils::install.packages()` function with the following syntax:
 
@@
 
-You can load a given library using the following syntax:
+You can load a given library using the `base::library()` function with the following syntax:
 
@@
 
-Below, we provide examples for how to learn about and use functions and arguments, by using the `seq()` function as an example.
-The `seq()` function creates a sequence of numbers.
-To learn about the `seq()` function, which creates a sequence of numbers, you can execute the following command:
+Below, we provide examples for how to learn about and use functions and arguments, by using the `base::seq()` function as an example.
+The `base::seq()` function creates a sequence of numbers.
+To learn about the `base::seq()` function, which creates a sequence of numbers, you can execute the following command:
 
@@
 
-This is what the documentation shows for the `seq()` function in the `Usage` section:
+This is what the documentation shows for the `base::seq()` function in the `Usage` section:
 
@@
 
-Based on this information, we know that the `seq()` function takes the following arguments:
+Based on this information, we know that the `base::seq()` function takes the following arguments:
 
@@
 A vector has one dimension (length).
-To create a vector, use the `c()` function to combine elements into a vector ("c" stands for combine).
+To create a vector, use the `base::c()` function to combine elements into a vector ("c" stands for combine).
 And, we use the assignment operator (`<-`) to assign the vector to an object named `exampleVector`, so we can access it later.
@@
 A data frame has two dimensions: rows and columns.
-Here is an example of creating a data frame, while using the assignment operator (`<-`) to assign the data frame to an object so we can access it later:
+Here is an example of creating a data frame using the `base::data.frame()` function, while using the assignment operator (`<-`) to assign the data frame to an object so we can access it later:
 
@@
 
+```
+
+We can also create a data frame from all combinations of variables, using the `base::expand.grid()` function:
+
+```{r}
 fantasyPoints_weekly <- expand.grid(
@@
 )
-
+```
+
+We can create a vector that is randomly sampled from values using the `base::sample()` function:
+
+```{r}
 set.seed(52242)
@@
 
-A list can store multiple data frames in one object:
+A list can store multiple data frames in one object using the `base::list()` function:
 
@@
 
-Here is how you load a `.RData` file using a *relative path* (i.e., a path relative to the working directory, where the working directory is represented by a period):
+Here is how you load a `.RData` file using a *relative path* (i.e., a path relative to the working directory, where the working directory is represented by a period) using the `base::load()` function:
 
@@
 The direction of the slashes matters—you should use forward slashes (not backslashes)!
-To determine where you working directory is, you can type:
+To determine where you working directory is, you can use the `base::getwd()` function:
 
@@
 
-To change your working directory, you can type:
+To change your working directory, you can use the `base::setwd()` function:
 
@@
 
-nfl_players <- read.csv("C:/Users/myusername/nfl_players.RData")
-```
-
-Here is how you load a `.csv` file:
+load("C:/Users/myusername/nfl_players.RData")
+```
+
+Here is how you load a `.csv` file using the `utils::read.csv()` function:
 
@@
 
-Here is how you save a `.RData` file using a *relative path* (i.e., relative to your working directory):
+Here is how you save a `.RData` file using a *relative path* (i.e., relative to your working directory) using the `base::save()` function:
 
@@
 
-Here is how you save a `.csv` file:
+Here is how you save a `.csv` file using the `utils::write.csv()` function:
 
@@
 
-To see the names of variables in a data frame, use the following syntax:
+To see the names of variables in a data frame, use the `base::names()` function:
 
@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] has a convenience function, `%ni%`, for the logical operator of determining whether an element is *not* in a value of another vector:
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] has a convenience function, `petersenlab::\`%ni%\``, for the logical operator of determining whether an element is *not* in a value of another vector:
 
@@
 
+`base::is.na()`
+
 ```{r}
@@
 
-The operator, `which()`, can be used to obtain the TRUE indices of an object, which is useful for [subsetting](#sec-subset).
+The operator, `base::which()`, can be used to obtain the TRUE indices of an object, which is useful for [subsetting](#sec-subset).
 
@@
 
-To apply conditional operations to a vector, we can use the `ifelse()` function.
+To apply conditional operations to a vector, we can use the `base::ifelse()` function.
 
@@
 
-In base `R`, you can perform piping with the `|>` expression.
-In `tidyverse` you can perform piping with the `%>%` expression.
+In base `R`, you can perform piping with the `base::\`|>\`` expression.
+In `tidyverse`, you can perform piping with the `magrittr::\`%>%\`` expression.
 
@@
 
+In `tidyverse`, you cans select particular columns using the `dplyr::select()` function.
+
 ```{r}
@@
 
+In `tidyverse`, you cans select particular columns using the `dplyr::filter()` function.
+
 ```{r}
@@
 
+In `tidyverse`, you cans select particular rows and columns using the `dplyr::filter()` and `dplyr::select()` functions, respectively.
+
 ```{r}
@@
 
-To view data, use the following syntax:
+To view data, use the `utils::View()` function:
 
@@
 
-To view only the first six rows (if a data frame) or elements (if a vector), use the following syntax:
+To view only the first six rows (if a data frame) or elements (if a vector), use the `utils::head()` function:
 
@@
 
+`utils::str()`
+
 ```{r}
@@
 
-Number of rows and columns:
+Number of rows and columns (`base::dim()`):
 
@@
 
-Number of rows:
+Number of rows (`base::nrow()`):
 
@@
 
-Number of columns:
+Number of columns (`base::ncol()`):
 
@@
 ### Number of Elements {#sec-dataCharacteristicsLength}
+
+`base::length()`
 
@@
 
-You can determine how many rows in the data frame have no missing data using the following syntax:
+You can determine how many rows in the data frame have no missing data using the following syntax and the `stats::na.omit()` function:
 
@@
 
-Recode multiple variables:
+Recode multiple variables using the `dplyr::mutate()` and `dplyr::case_match()` functions:
 
@@
 
+`dplyr::rename()`
+
 ```{r}
@@
 
-Using a vector of variable names:
+Using a vector of variable names with the `dplyr::rename_with()` function:
 
@@
 
-One variable:
+One variable (`base::factor()`; `base::as.numeric()`; `base::as.integer()`; `base::as.character()`):
 
@@
 Thus, when we merge them, the resulting object has many more rows than each respective object (even though neither object has players that the other object does not).
+Below, we merge data using the `dplyr::full_join()` function:
 
@@
 However, as described above, you should include it as a key ***only*** if you want to use its values to align the rows from each object.
-Below is an example of merging two objects with the same variable name (i.e., points) that is not used as a key.
+Below is an example of merging two objects with the same variable name (i.e., points) that is not used as a key, using the `dplyr::full_join()` function.
 
@@
 When you merge them, the resulting data object will be in `player`-`season`-`week` form.
+Below, we merge data using the `dplyr::full_join()` function:
 
@@
 It returns columns from `x` and `y`.
-Here is how to merge two data frames using a full outer join (i.e., "full join"):
+Here is how to merge two data frames using a full outer join (i.e., "full join"; `dplyr::full_join()`):
 
@@
 It returns columns from `x` and `y`.
-Here is how to merge two data frames using a left outer join ("left join"):
+Here is how to merge two data frames using a left outer join ("left join"; `dplyr::left_join()`):
 
@@
 It returns columns from `x` and `y`.
-Here is how to merge two data frames using a right outer join ("right join"):
+Here is how to merge two data frames using a right outer join ("right join"; `dplyr::right_join()`):
 
@@
 It returns columns from `x` and `y`.
-Here is how to merge two data frames using an inner join:
+Here is how to merge two data frames using an inner join (`dplyr::inner_join()`):
 
@@
 Unlike an [inner join](#innerJoin), a left semi join will never duplicate rows of `x`, and it includes columns from only `x` (not from `y`).
-Here is how to merge two data frames using a left semi join:
+Here is how to merge two data frames using a left semi join (`dplyr::semi_join()`):
 
@@
 It returns columns from only `x` (not from `y`).
-Here is how to merge two data frames using a left anti join:
+Here is how to merge two data frames using a left anti join (`dplyr::anti_join()`):
 
@@
 
-A cross join combines each row in `x` with each row in `y`.
+A cross join combines each row in `x` with each row in `y` (`dplyr::cross_join()`).
 
@@
 The long data are in "`player`-`season`-`week` form".
-We widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable):
+We widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable) using the `tidyr::pivot_wider()` function:
 
@@
 
-Conversely, we can also restructure data from wide to long.
+Conversely, we can also restructure data from wide to long using the `tidyr::pivot_longer()` function.
 Here are the data in long form, after they have been transformed from wide form using `tidyverse`:

### 2025-08-10 index.qmd
20250809 - label functions

@@
   "MASS", "data.table", "future", "future.apply", "SimDesign", "domir", 
-  "GGally","Rglpk",
+  "GGally","Rglpk","TTR",
   "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem

### 2025-08-10 machine-learning.qmd
20250809 - label functions

@@
 
-We set up the *k* folds using the `group_vfold_cv()` function of the `rsample` package [@R-rsample].
+We set up the *k* folds using the `rsample::group_vfold_cv()` function of the `rsample` package [@R-rsample].
 
@@
 
-We set up the LOO folds using the `loo_cv()` function of the `rsample` package [@R-rsample].
+We set up the LOO folds using the `rsample::loo_cv()` function of the `rsample` package [@R-rsample].
 
@@
 For each of the models, we fit and evaluate the models using the `tidymodels` ecosystem of packages [@R-tidymodels; @Kuhn2020_packages].
-we specify our model formula using the `recipe()` function of the `recipes` package [@R-recipes].
-We define the model using the `linear_reg()`, `set_engine()`, and `set_mode()` functions of the `parsnip` package [@R-parsnip].
-We specify the workflow using the `workflow()`, `add_recipe()`, and `add_model()` functions of the `workflows` package [@R-workflows].
-We fit the cross-validation model using the `fit_resamples()` function of the `tune` package [@R-tune].
-We specify the accuracy metrics to evaluate using the `metric_set()` function of the `yardstick` package [@R-yardstick].
-We fit the final model using the `fit()` function of the `workflows` package [@R-workflows].
-We evaluate the accuracy of the model's predictions on the test data using the `accuracyOverall()` of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab].
+we specify our model formula using the `recipes::recipe()` function of the `recipes` package [@R-recipes].
+We define the model using the `parsnip::linear_reg()`, `parsnip::set_engine()`, and `parsnip::set_mode()` functions of the `parsnip` package [@R-parsnip].
+We specify the workflow using the `workflows::workflow()`, `workflows::add_recipe()`, and `workflows::add_model()` functions of the `workflows` package [@R-workflows].
+We fit the cross-validation model using the `tune::fit_resamples()` function of the `tune` package [@R-tune].
+We specify the accuracy metrics to evaluate using the `yardstick::metric_set()` function of the `yardstick` package [@R-yardstick].
+We fit the final model using the `workflows::fit()` function of the `workflows` package [@R-workflows].
+We evaluate the accuracy of the model's predictions on the test data using the `petersenlab::accuracyOverall()` of the [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab].
 
@@
 
-For the machine learning models, we perform the parameter tuning using the `tune()`, `tune_grid()`, `select_best()`, and `finalize_workflow()` functions of the `tune` package [@R-tune].
-We specify the grid of possible parameter values using the `grid_regular()` function of the `dials` package [@R-dials].
+For the machine learning models, we perform the parameter tuning using the `tune::tune()`, `tune::tune_grid()`, `tune::select_best()`, and `tune::finalize_workflow()` functions of the `tune` package [@R-tune].
+We specify the grid of possible parameter values using the `dials::grid_regular()` function of the `dials` package [@R-dials].
 
@@
 The random forest models were fit using the `ranger` package [@R-ranger; @Wright2017_packages].
-We specify the grid of possible parameter values using the `grid_random()`, `update.parameters()`, `mtry()`, and `min_n()` functions of the `dials` package [@R-dials] and the `extract_parameter_set_dials()` function of the `hardhat` package [@R-hardhat].
+We specify the grid of possible parameter values using the `dials::grid_random()`, `dials::update.parameters()`, `dials::mtry()`, and `dials::min_n()` functions of the `dials` package [@R-dials] and the `hardhat::extract_parameter_set_dials()` function of the `hardhat` package [@R-hardhat].
 
@@
 Approaches to estimating random forest models with longitudinal data are described by @Hu2023.
-Below, we fit longitudinal random forest models using the `MERF()` function of the `LongituRF` package [@R-LongituRF].
+Below, we fit longitudinal random forest models using the `LongituRF::MERF()` function of the `LongituRF` package [@R-LongituRF].
 
@@
 We then use the [mean absolute error](#sec-meanAbsoluteError) as our index of prediction accuracy on the held-out validation data.
+We use the `gpboost::gpb.Dataset()` function to specify the data, the `gpboost::GPModel()` function to specify the model, and the `gpboost::gpb.grid.search.tune.parameters()` function to identify the optimal tuning parameters for that model.
 
@@
 
+We use the `gpboost::gpb.train()` function to fit the specified model with the specified tuning parameters and dataset.
+
 ```{r}
@@
 summary(gp_model_qb) # estimated random effects model
-
+```
+
+We evaluate the importance of features using the `gpboost::gpb.importance()` and `gpboost::gpb.plot.importance()` functions:
+
+```{r}
 gp_model_qb_importance <- gpboost::gpb.importance(gp_model_fit_qb)

### 2025-08-10 mixed-models.qmd
20250806 - robust load
20250809 - label functions

@@
 
-We compare nested models using an $F$-test with the `anova()` function.
-We compare non-nested models using the `bbmle` [@R-bbmle] and `MuMIn` [@R-MuMIn] packages.
+We compare nested models using an $F$-test with the `stats::anova()` function.
+We compare non-nested models using the `bbmle` [@R-bbmle] (`bbmle::AICtab()`) and `MuMIn` [@R-MuMIn] (`MuMIn::AICc()`) packages.
 
@@
 
-We estimate the multilevel models using the `lmerTest` package [@R-lmerTest], which is an extension of the `lme4` package [@R-lme4; @Bates2015_packages]:
+We estimate the multilevel models using the `lmerTest::lmer()` function of the `lmerTest` package [@R-lmerTest], which is an extension of the `lme4` package [@R-lme4; @Bates2015_packages]:
 
@@
 
-To obtain standardized regression coefficients, we use the `standardize_parameters()` function of the `effectsize` package [@R-effectsize; @BenShachar2020_packages].
-We use the `emmeans` package [@R-emmeans] to determine the model-implied number of points per season by position.
+To obtain standardized regression coefficients, we use the `effectsize::standardize_parameters()` function of the `effectsize` package [@R-effectsize; @BenShachar2020_packages].
+We use the `emmeans::emmeans()` function of the `emmeans` package [@R-emmeans] to determine the model-implied number of points per season by position.
 
@@
 
-To visualize the regression coefficients, we can use the `tidy()` function of the `broom.mixed` package [@R-broom.mixed], as in @fig-individualFantasyPointsRandomLinearFixedQuadraticSlopesInteractionCoefficients.
+To visualize the regression coefficients, we can use the `broom.mixed::tidy()` function of the `broom.mixed` package [@R-broom.mixed], as in @fig-individualFantasyPointsRandomLinearFixedQuadraticSlopesInteractionCoefficients.
 
@@
 
-load(url("https://osf.io/download/q6rjf/"))
+petersenlab::robust_load("https://osf.io/download/q6rjf/") #load(url("https://osf.io/download/q6rjf/"))
 ```
@@
 
-To visualize the regression coefficients, we can use the `tidy()` function of the `broom.mixed` package [@R-broom.mixed], as in @fig-bayesianRegressionCoefficients.
+To visualize the regression coefficients, we can use the `broom.mixed::tidy()` function of the `broom.mixed` package [@R-broom.mixed], as in @fig-bayesianRegressionCoefficients.
 

### 2025-08-10 modern-portfolio-theory.qmd
20250801 - correlation matrix plots
20250809 - label functions

@@
 library("quantmod")
+library("TTR")
 library("fPortfolio")
@@
 
-In @fig-correlationMatrixPlot, we depict a correlation matrix plot using the `ggcorr()` function of the `GGally` [@R-GGally] package.
+In @fig-correlationMatrixPlot, we depict a correlation matrix plot using the `GGally::ggcorr()` function of the `GGally` [@R-GGally] package.
 
@@
 
-GGally::ggcorr(combined_points_wide %>% select(QB1:opp_DST))
+GGally::ggcorr(
+  combined_points_wide %>% select(QB1:opp_DST),
+  label = TRUE,
+  label_size = 3)
 ```
@@
 
+We download closing prices using the `quantmod::Cl()` function of the `quantmod` package [@R-quantmod].
+We calculate returns from the closing prices using the `TTR::ROC()` function of the `TTR` package [@R-TTR].
+
 ```{r}
@@
 We use the `fPortfolio` package [@R-fPortfolio] to determine the optimal portfolio.
+We specify the portfolio specifications using the `fPortfolio::portfolioSpec()` function.
 
@@
 ### Determine the Efficient Frontier {#sec-determineEfficientFrontier}
+
+We determine the efficient frontier using the `fPortfolio::portfolioFrontier()` function.
 
@@
 In other words, it is the portfolio with the greatest risk-adjusted returns.
+We identify the tangency portfolio using the `fPortfolio::tangencyPortfolio()` function.
 
@@
   spec = portfolioSpec)
-
+```
+
+We extract the optimal weights for each asset at the tangency portfolio using the `fPortfolio::getWeights()` function.
+
+```{r}
 # Extract optimal weights
@@
 
-If the `maxreturnPortfolio()` function worked as expected, the following could should work for determining the portfolio with the maximum return at each of various risk levels.
+If the `fPortfolio::maxreturnPortfolio()` function worked as expected, the following could should work for determining the portfolio with the maximum return at each of various risk levels.
 However, there is a known bug in the `fportfolio` package [@R-fPortfolio] that the optimal portfolio (based on maximum returns) does not change when changing the target risk level, suggesting that it is not taking into account the target risk level specified by the user (see here: <https://stackoverflow.com/q/78784306/2029527>).
@@
 
-Below, we use the `mvFrontier()` function from the `NMOF` package [@R-NMOF; @Gilli2019_packages; @Schumann20112024_packages] to compute the efficient frontier.
+Below, we use the `NMOF::mvFrontier()` function from the `NMOF` package [@R-NMOF; @Gilli2019_packages; @Schumann20112024_packages] to compute the efficient frontier.
 
@@
 You can think of constructing your team like constructing a stock portfolio, where your goal is to gain the maximum expected returns for a given level of risk.
-The greater risk you are willing to take, the greater the potential returns, but alse the lower the potential downside if the players underperform.
+The greater risk you are willing to take, the greater the potential returns, but also the lower the potential downside if the players underperform.
 Thus, you may want your team to be composed of a diverse combination of some higher risk players and some lower risk players.

### 2025-08-10 multiple-regression.qmd
20250806 - robust load
20250809 - label functions

@@
 
-We fit a linear regression model using the `lm()` function.
+We fit a linear regression model using the `stats::lm()` function.
 
@@
 
-We fit a logistic regression model using the `glm()` function and specifying `family = binomial()`.
-To calculate the model $R^2$, we use the `r2()` function of the `performance` package [@R-performance; @Ludecke2021_packages].
+We fit a logistic regression model using the `stats::glm()` function and specifying `family = binomial()`.
+To calculate the model $R^2$, we use the `performance::r2()` function of the `performance` package [@R-performance; @Ludecke2021_packages].
 
@@
 
-We fit an ordinal regression model using the `clm()` function of the `ordinal` package [@R-ordinal].
+We fit an ordinal regression model using the `ordinal::clm()` function of the `ordinal` package [@R-ordinal].
 
@@
 
-We fit a Poisson regression model using the `glm()` function and specifying `family = poisson()`.
+We fit a Poisson regression model using the `stats::glm()` function and specifying `family = poisson()`.
 
@@
 
-We fit a negative binomial regression model using the `glm.nb()` function of the `MASS` package [@R-MASS].
+We fit a negative binomial regression model using the `MASS::glm.nb()` function of the `MASS` package [@R-MASS].
 
@@
 
-We fit a (Bayesian) zero-one-inflated beta regression model (to allow for zeros and ones) using the `brm()` function of the `brms` package [@R-brms] and specifying `family = zero_one_inflated_beta()`.
+We fit a (Bayesian) zero-one-inflated beta regression model (to allow for zeros and ones) using the `brms::brm()` function of the `brms` package [@R-brms] and specifying `family = zero_one_inflated_beta()`.
 
@@
 
-load(url("https://osf.io/download/fe37j/"))
+petersenlab::robust_load("https://osf.io/download/fe37j/") #load(url("https://osf.io/download/fe37j/"))
 ```
@@
 
-If we want to obtain standardized regression coefficients, we can use the `standardize_parameters()` function of the `effectsize` package [@R-effectsize; @BenShachar2020_packages].
+If we want to obtain standardized regression coefficients, we can use the `effectsize::standardize_parameters()` function of the `effectsize` package [@R-effectsize; @BenShachar2020_packages].
 
@@
 
-Or, we can standardize the [outcome variable](#sec-correlationalStudy) and each [predictor variable](#sec-correlationalStudy) using the `scale()` function:
+Or, we can standardize the [outcome variable](#sec-correlationalStudy) and each [predictor variable](#sec-correlationalStudy) using the `base::scale()` function:
 
@@
 We can perform a dominance analysis to evaluate the relative importance of the predictors in the regression model.
-To perform the dominance analysis, we use the `dominance_analysis()` function of the `parameters` package [@R-parameters; @Luedecke2020_packages], which leverages the `domir` package [@R-domir].
+To perform the dominance analysis, we use the `parameters::dominance_analysis()` function of the `parameters` package [@R-parameters; @Luedecke2020_packages], which leverages the `domir` package [@R-domir].
 
@@
 
-To visualize the regression coefficients, we can use the `tidy()` function of the `broom` package [@R-broom], as in Figures [-@fig-regressionCoefficients] and [-@fig-regressionCoefficientsStandardized].
+To visualize the regression coefficients, we can use the `broom::tidy()` function of the `broom` package [@R-broom], as in Figures [-@fig-regressionCoefficients] and [-@fig-regressionCoefficientsStandardized].
 
@@
 An important consideration in multiple regression is how missing data are handled.
-Multiple regression in `R` using the `lm()` function applies listwise deletion.
+Multiple regression in `R` using the `stats::lm()` function applies listwise deletion.
 *Listwise deletion* (also called complete case analysis) removes any row (in the data file) from analysis that has a missing value on the [outcome variable](#sec-correlationalStudy) or any of the [predictor variables](#sec-correlationalStudy).
@@
 To address outliers, there are various approaches to robust regression.
-One approach is to use an MM-type estimator, such as is used in the `lmrob()` and `glmrob()` functions of the `robustbase` package [@R-robustbase; @Todorov2009_packages].
+One approach is to use an MM-type estimator, such as is used in the `robustbase::lmrob()` and `robustbase::glmrob()` functions of the `robustbase` package [@R-robustbase; @Todorov2009_packages].
 
@@
 
-We set up the bootstrap folds using the `bootstraps()` function of the `rsample` package [@R-rsample].
+We set up the bootstrap folds using the `rsample::bootstraps()` function of the `rsample` package [@R-rsample].
 

### 2025-08-10 pca.qmd
20250809 - label functions

@@
 
-We generated the scree plot in @fig-pcaScreePlot1 using the `fa.parallel()` function of the `psych` package [@R-psych].
+We generated the scree plot in @fig-pcaScreePlot1 using the `psych::fa.parallel()` function of the `psych` package [@R-psych].
 
@@
 
-We generated the scree plot in @fig-pcaScreePlot2 using the `nScree()` and `plotnScree()` functions of the `nFactors` package [@R-nFactors].
+We generated the scree plot in @fig-pcaScreePlot2 using the `nFactors::nScree()` and `nFactors::plotnScree()` functions of the `nFactors` package [@R-nFactors].
 
@@
 
-We generated the very simple structure (VSS) plots in Figures [-@fig-pcaVSSPlot1] and [-@fig-pcaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
+We generated the very simple structure (VSS) plots in Figures [-@fig-pcaVSSPlot1] and [-@fig-pcaVSSPlot2] using the `psych::vss()` and `psych::nfactors()` functions of the `psych` package [@R-psych].
 The optimal number of components based on the VSS criterion is three components.

### 2025-08-10 simulation.qmd
20250809 - label functions

@@
 
-For performing the bootstrapping, we leverage the `data.table` [@R-data.table], `future` [@R-future], and `future.apply` [@R-future.apply] packages for speed (by using parallel processing) and memory efficiency.
-We use the `progressr` [@R-progressr] package to create a progress bar.
+For performing the bootstrapping, we leverage the `data.table` [@R-data.table] (`data.table::as.data.table()`; `data.table::data.table()`; `data.table::rbindlist()`), `future` [@R-future] (`future::plan()`; `future::multisession()`), and `future.apply` [@R-future.apply] (`future.apply::future_lapply()`) packages for speed (by using parallel processing) and memory efficiency.
+We use the `progressr` [@R-progressr] package (`progressr::handlers()`; `progressr::with_progress()`; `progressr::progressor()`) to create a progress bar.
 
@@
 
+`data.table::setnames()`
+
 ```{r}
@@
 ```
+
+`base::lapply()`
 
@@
 For each player, we identify the optimal distribution as either a normal distribution , or as a skew-normal distribution.
-The normal distribution was fit using the `fitdist()` function of the `fitdistrplus` package [@R-fitdistrplus; @DelignetteMuller2015_packages].
-The skew-normal distribution was fit using the `selm()` function of the `sn` package [@R-sn; @Azzalini2023_packages].
+The normal distribution was fit using the `fitdistrplus::fitdist()` function of the `fitdistrplus` package [@R-fitdistrplus; @DelignetteMuller2015_packages].
+The skew-normal distribution was fit using the `sn::selm()` function of the `sn` package [@R-sn; @Azzalini2023_packages].
 
@@
 
-Now, we can run the model.
+Now, we can run the model using the `SimDesign::runSimulation()` function.
 

### 2025-08-10 time-series-analysis.qmd
20250806 - robust load
20250809 - label functions

@@
 ```{r}
-load(url("https://osf.io/download/q6rjf/")) # Bayesian model object
+petersenlab::robust_load("https://osf.io/download/q6rjf/") #load(url("https://osf.io/download/q6rjf/")) # Bayesian model object
 ```
@@
 The autocorrelation function (ACF) plot depicts the autocorrelation of scores as a function of the length of the lag.
+We can generate an ACF plot using the `stats::acf()` function.
 Significant autocorrelation is detected when the autocorrelation exceeds the dashed blue lines, as is depicted in @fig-timeSeriesACFtomBrady.
@@
 
-Using the `forecast` package [@R-forecast; @Hyndman2008_packages]:
+We can fit an ARIMA model using the `forecast::auto.arima()` function of the `forecast` package [@R-forecast; @Hyndman2008_packages]:
 
@@
 ```
+
+We can generate a plot of an ARIMA model using the `stats::arima()` function.
 
@@
 
+We can generate model forecasts from the ARIMA model using the `forecast::forecast()` function.
+
 ```{r}
@@
 ### Plot the Model Forecasts {#sec-plotARIMAforecast}
+
+We can plot the model forecasts using the `forecast::autoplot()` function.
 

### 2025-08-10 draft.qmd
20250809 - fix URL

@@
 
-FantasyFootballAnalytics.net provides tools to help you win your [auction draft](#sec-draftSnake) (<https://fantasyfootballanalytics.net/win-your-snake-draft>) and daily fantasy sports (DFS) leagues (<https://fantasyfootballanalytics.net/how-to/dfsoptimizerhelp>).
+FantasyFootballAnalytics.net provides tools to help you win your [auction draft](#sec-draftSnake) (<https://fantasyfootballanalytics.net/how-to-win-your-auction-draft>) and daily fantasy sports (DFS) leagues (<https://fantasyfootballanalytics.net/how-to/dfsoptimizerhelp>).
 For draft and player selection tools for [auction drafts](#sec-draftSnake) and DFS leagues, see the Fantasy Football Analytics web application: <https://apps.fantasyfootballanalytics.net>.

### 2025-08-10 base-rates.qmd
20250810 - edits

@@
 Because they do not pay enough attention to the base rate (in this example, the base rate of HIV is .3%).
-
-In general, people tend to overestimate the likelihood of low base-rate events.
+Base rates are important to account for when making predictions and people tend to ignore them, a phenomenon called [base rate neglect](#sec-fallaciesBaseRate), which is described in @sec-fallaciesBaseRate.
+
+In general, people tend to overestimate the likelihood of low base-rate events [@Kahneman2011].
 That is, if the base rate of an event or condition—such as schizophrenia—is low (e.g., ~0.5%), people overestimate the likelihood that a person has schizophrenia when given specific information about the person such as their symptoms and history.
-Moreover, @Fox1998a asked National Basketball Association (NBA) fans to estimate the probability that each of 8 teams in the playoff would win the playoffs.
+As an example of people overestimating the likelihood of low base-rate events, @Fox1998a asked National Basketball Association (NBA) fans to estimate the probability that each of 8 teams in the playoff would win the playoffs.
 The median sum of probability judgments for the eight teams was 240% [@Fox1998a; @Kahneman2011] even though, in reality, the true probabilities must sum to 100%.
-The finding suggests that people were evaluating each team individually, considering the reasons why that particular team could win, without properly accounting for the total probability.
+The base rate for a given team to win is 12.5% (i.e., 1/8), whereas people were giving each team (on average) a 30% chance (i.e., 240/8) to win.
+The finding suggests that people were evaluating each team individually, considering the reasons why that particular team could win, without properly accounting for the total probability (and the base rate).
+In addition, people tend to overweight unlikely events in their decisions [@Kahneman2011].
+
+Another important phenomenon is that low base-rate events (i.e., unlikely events) are difficult to predict accurately.
+Predictions of lower base rate phenomena tend to be lower than predictions of more common phenomena.
+For instance, predictions of touchdowns (less common) tend to be less accurate than predictions of yardage (more common).
+
+In this chapter, we describe [ways to account for base rates](#sec-accountForBaseRates) in judgments and predictions.
 
@@
 
+To calculate the probability of the inverse conditional, we can leverage Bayesian statistics to calculate the probability of the inverse conditional given a conditional probability (i.e., likelihood) and the marginal probabilities (base rates) of both events.
+Bayesian statistics is another branch of statistics and is different from frequentist statistics and null hypothesis significance testing.
+Bayesian statistics is based on Bayes’ theorem, which allows updating our probability estimates based on prior information (e.g., base rate) and new data.
+
 ### Bayes' Theorem {#sec-bayesTheorem}
@@
 In general, people tend to a) overestimate the likelihood of low base-rate events and b) overweight low base-rate events in their decisions [@Kahneman2011].
-Bayes' theorem specifies how prior beliefs (i.e., [base rate](#sec-baseRate) informations) should be integrated with the [predictive accuracy](#sec-predictiveValidity) of the evidence to make predictions.
+Bayes' theorem specifies how prior beliefs (i.e., [base rate](#sec-baseRate) information) should be integrated with the [predictive accuracy](#sec-predictiveValidity) of the evidence to make predictions.
 It thus provides a powerful tool to [anchor](#sec-heuristicsAnchoringAdjustment) predictions to the [base rate](#sec-baseRate) unless sufficient evidence changes the posterior probability (by updating the evidence and [prior probability](#sec-baseRate)).
@@
 
-The positive likelihood ratio (LR+) compares the [true positive rate](#sec-sensitivity) to the [false positive rate](#sec-falsePositiveRate).
+The positive likelihood ratio (LR+) is the probability that a person with the disease tested positive for the disease ([true positive rate](#sec-sensitivity)) divided by the probability that a person without the disease tested positive for the disease ([false positive rate](#sec-falsePositiveRate)).
+That is, the positive likelihood ratio (LR+) compares (i.e., is a ratio of) the [true positive rate](#sec-sensitivity) to the [false positive rate](#sec-falsePositiveRate).
 Positive likelihood ratio values range from 1 to infinity.\index{positive likelihood ratio}
 Higher values reflect greater accuracy, because it indicates the degree to which a [true positive](#sec-truePositive) is more likely than a [false positive](#sec-falsePositive).
+Testing positive on a test with a high LR+ increases the probability of disease.
 The formula for calculating the positive likelihood ratio is in @eq-positiveLikelihoodRatio.
@@
 
-The negative likelihood ratio (LR−) compares the [false negative rate](#sec-falseNegativeRate) to the [true negative rate](#sec-specificity).
+The negative likelihood ratio (LR−) is the probability that a person with the disease tested negative for the disease ([false negative rate](#sec-falseNegativeRate)) divided by the probability that a person without the disease tested negative for the disease ([true negative rate](#sec-specificity)).
+That is, the negative likelihood ratio (LR−) compares (i.e., is a ratio of) the [false negative rate](#sec-falseNegativeRate) to the [true negative rate](#sec-specificity).
 Negative likelihood ratio values range from 0 to 1.
 Smaller values reflect greater accuracy, because it indicates that a [false negative](#sec-falseNegative) is less likely than a [true negative](#sec-trueNegative).
+Testing negative on a test with a low LR– decreases the probability of disease.
 The formula for calculating the negative likelihood ratio is in @eq-negativeLikelihoodRatio.

### 2025-08-10 evaluating-prediction-accuracy.qmd
20250810 - accuracy at cutoff

@@
 The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::accuracyAtCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables.
-
-```{r}
-#petersenlab::accuracyAtCutoff()
+In the example below, we compute the accuracy of predictions for which Running Backs or Wide Receivers will score 300 or more fantasy points, using a cutoff for average projected fantasy points of 300.
+
+```{r}
+petersenlab::accuracyAtCutoff(
+  predicted = projectionsWithActuals_seasonal_subset$raw_points,
+  actual = projectionsWithActuals_seasonal_subset$truth,
+  cutoff = 300
+)
 ```
@@
 The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::accuracyAtEachCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables at each possible cutoff.
-
-```{r}
-#petersenlab::accuracyAtEachCutoff()
+In the example below, we compute the accuracy of predictions for which Running Backs or Wide Receivers will score 300 or more fantasy points, at each possible cutoff for average projected fantasy points.
+
+```{r}
+petersenlab::accuracyAtEachCutoff(
+  predicted = projectionsWithActuals_seasonal_subset$raw_points,
+  actual = projectionsWithActuals_seasonal_subset$truth
+)
 ```

### 2025-08-11 evaluating-prediction-accuracy.qmd
20250810 - optimal cutoff

@@
 The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::accuracyAtCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables.
-In the example below, we compute the accuracy of predictions for which Running Backs or Wide Receivers will score 300 or more fantasy points, using a cutoff for average projected fantasy points of 300.
+In the example below, we compute the accuracy of predictions for which Running Backs or Wide Receivers will score 300 or more fantasy points, using a cutoff for projected fantasy points of 300.
 
@@
 The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::accuracyAtEachCutoff()` function that computes many accuracy indices for the prediction of categorical outcome variables at each possible cutoff.
-In the example below, we compute the accuracy of predictions for which Running Backs or Wide Receivers will score 300 or more fantasy points, at each possible cutoff for average projected fantasy points.
+In the example below, we compute the accuracy of predictions for which Running Backs or Wide Receivers will score 300 or more fantasy points, at each possible cutoff for projected fantasy points.
 
@@
 Values below zero indicate that the observed values are, on average, greater than the predicted values.
+
+### Optimal Cutoff {#sec-optimalCutoff}
+
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] contains the `petersenlab::optimalCutoff()` function that identifies the cutoff that optimizes each of various accuracy estimates.
+
+```{r}
+#| eval: false
+
+petersenlab::optimalCutoff(
+  predicted = projectionsWithActuals_seasonal_subset$raw_points,
+  actual = projectionsWithActuals_seasonal_subset$truth)
+```
+
+```{r}
+#| include: false
+
+optimalCutoff <- petersenlab::optimalCutoff(
+  predicted = projectionsWithActuals_seasonal_subset$raw_points,
+  actual = projectionsWithActuals_seasonal_subset$truth)
+```
+
+```{r}
+#| echo: false
+
+for(i in 1:length(optimalCutoff)){
+  print(head(optimalCutoff[[i]]))
+}
+```
 
@@
   print.auc = TRUE)
+```
+
+We can identify the cutoff that optimizes each of various accuracy estimates.
+
+```{r}
+petersenlab::optimalCutoff(
+  predicted = player_stats_weekly_example$rushing_yards_priorGame,
+  actual = player_stats_weekly_example$rush_200_yds)
 ```

### 2025-08-11 decision-making.qmd
20250811

@@
 - <https://the-odds-api.com>
+- <https://www.evsharps.com/ranks?format=std>
 

### 2025-08-11 player-evaluation.qmd
20250811

@@
 [Fantasy Football Analytics](https://fantasyfootballanalytics.net) creates fantasy rankings from projections that are aggregated across sources (see the webapp here: <https://apps.fantasyfootballanalytics.net>).
+There are also sites that aggregate rankings from both fantasy football analysts and from sportsbooks, which are based on Vegas lines and how people bet: <https://www.evsharps.com/ranks?format=std>.
 
@@
 Projections that are aggregated across sources are provided by [Fantasy Football Analytics](https://fantasyfootballanalytics.net) (see the webapp here: <https://apps.fantasyfootballanalytics.net>) and by [FantasyPros](https://www.fantasypros.com/nfl/auction-values/calculator.php).
-Moreover, sites also provide season-long player futures for the over/under betting lines, which are based on Vegas lines and how people bet: <https://www.rotowire.com/betting/nfl/player-futures.php>.
+There are also sites that aggregate projections from both fantasy football analysts and from sportsbooks, which are based on Vegas lines and how people bet: <https://www.evsharps.com/ranks?format=std>.
+Moreover, sites also provide season-long player futures for the over/under betting lines: <https://www.rotowire.com/betting/nfl/player-futures.php>; <https://www.evsharps.com/futures>.
 In addition, sites also provide weekly projections based on Vegas lines: <https://vegasprojections.com>; <https://www.actionnetwork.com/nfl/props>; <https://tools.32beatwriters.com>; <https://the-odds-api.com>.

### 2025-08-12 data-visualization.qmd
20250811 - function prefixes

@@
 
-- [histogram](#sec-histogram): `geom_histogram()`
-- [density plot](#sec-densityPlot): `geom_density()`
-- [box-and-whisker plot](#sec-boxAndWhiskerPlot): `geom_boxplot()`
-- [violin plot](#sec-violinPlot): `geom_violin()`
-- [bar plot](#sec-barPlot): `geom_bar()`
-- [scatterplot](#sec-scatterplot): `geom_point()`
-- [rug plot](#sec-scatterplotRugPlot): `geom_rug()`
-- [line chart](#sec-lineChart): `geom_line()`
-- statistical summary (e.g., [best-fit line](#sec-scatterplotBestFitLine)): `geom_smooth()`
+- [histogram](#sec-histogram): `ggplot2::geom_histogram()`
+- [density plot](#sec-densityPlot): `ggplot2::geom_density()`
+- [box-and-whisker plot](#sec-boxAndWhiskerPlot): `ggplot2::geom_boxplot()`
+- [violin plot](#sec-violinPlot): `ggplot2::geom_violin()`
+- [bar plot](#sec-barPlot): `ggplot2::geom_bar()`
+- [scatterplot](#sec-scatterplot): `ggplot2::geom_point()`
+- [rug plot](#sec-scatterplotRugPlot): `ggplot2::geom_rug()`
+- [line chart](#sec-lineChart): `ggplot2::geom_line()`
+- statistical summary (e.g., [best-fit line](#sec-scatterplotBestFitLine)): `ggplot2::geom_smooth()`
 

### 2025-08-12 draft.qmd
20250809 - add cost
20250812 - draft tools

@@
 FantasyFootballAnalytics.net provides tools to help you win your [snake draft](#sec-draftSnake): <https://fantasyfootballanalytics.net/win-your-snake-draft>.
-For draft tools for [snake drafts](#sec-draftSnake), see the Fantasy Football Analytics web application: <https://apps.fantasyfootballanalytics.net>.
+For draft tools for [snake drafts](#sec-draftSnake), see the Fantasy Football Analytics web application (depicted in @fig-snakeDraftToolFFA): <https://apps.fantasyfootballanalytics.net>.
 The web application includes a free tool that allows users to specify their custom league settings to calculate projected points that are customized to their league and helps users select which players to draft that maximize the players' [value over replacement players](#sec-fantasyValueVORP) (VORP).
+
+::: {#fig-snakeDraftToolFFA}
+![](images/snakeDraftTool.png){fig-alt="Free Snake Draft Tool From Fantasy Football Analytics. From <https://apps.fantasyfootballanalytics.net>."}
+
+Free Snake Draft Tool From Fantasy Football Analytics. From <https://apps.fantasyfootballanalytics.net>.
+:::
 
@@
 FantasyFootballAnalytics.net provides tools to help you win your [auction draft](#sec-draftSnake) (<https://fantasyfootballanalytics.net/how-to-win-your-auction-draft>) and daily fantasy sports (DFS) leagues (<https://fantasyfootballanalytics.net/how-to/dfsoptimizerhelp>).
-For draft and player selection tools for [auction drafts](#sec-draftSnake) and DFS leagues, see the Fantasy Football Analytics web application: <https://apps.fantasyfootballanalytics.net>.
+For draft and player selection tools for [auction drafts](#sec-draftSnake) and DFS leagues, see the Fantasy Football Analytics web application (depicted in @fig-auctionDraftToolFFA): <https://apps.fantasyfootballanalytics.net>.
 The web application includes a free tool that allows users to specify their custom league settings to calculate projected points that are customized to their league, and to identify the optimal lineup given their league settings and salary cap.
+
+::: {#fig-auctionDraftToolFFA}
+![](images/lineupOptimizer.png){fig-alt="Free Auction Draft Lineup Optimizer From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>."}
+
+Free Auction Draft Lineup Optimizer From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>.
+:::
 
@@
 ```{r}
-mostRecentProjections[which(sol$solution == 1), c("name","pos","points")]
+mostRecentProjections[which(sol$solution == 1), c("name","pos","cost","points")]
 ```

### 2025-08-12 download-football-data.qmd
20250812 - linked figures

@@
 
+You can download fantasy football projections in `R` using the `ffanalytics::scrape_data()` function of the `ffanalytics` package [@R-ffanalytics].
+Alternatively, if you prefer a graphical user interface, you can also download fantasy football projections using the Fantasy Football Analytics web application (depicted in @fig-projectionsTabFFA): <https://apps.fantasyfootballanalytics.net>.
+
+::: {#fig-projectionsTabFFA}
+[![](images/projectionsTab.png){fig-alt="Fantasy Football Projections Available From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>."}](https://apps.fantasyfootballanalytics.net)
+
+Fantasy Football Projections Available From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>.
+:::
+
 ::: {#nte-downloadProjections .callout-note title="Downloading players' seasonal projections"}

### 2025-08-12 draft.qmd
20250812 - linked figures

@@
 ::: {#fig-snakeDraftToolFFA}
-![](images/snakeDraftTool.png){fig-alt="Free Snake Draft Tool From Fantasy Football Analytics. From <https://apps.fantasyfootballanalytics.net>."}
+[![](images/snakeDraftTool.png){fig-alt="Free Snake Draft Tool From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>."}](https://apps.fantasyfootballanalytics.net)
 
-Free Snake Draft Tool From Fantasy Football Analytics. From <https://apps.fantasyfootballanalytics.net>.
+Free Snake Draft Tool From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>.
 :::
@@
 ::: {#fig-auctionDraftToolFFA}
-![](images/lineupOptimizer.png){fig-alt="Free Auction Draft Lineup Optimizer From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>."}
+[![](images/lineupOptimizer.png){fig-alt="Free Auction Draft Lineup Optimizer From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>."}](https://apps.fantasyfootballanalytics.net)
 

### 2025-08-12 download-football-data.qmd
20250812 - unlink figures

@@
 ::: {#fig-projectionsTabFFA}
-[![](images/projectionsTab.png){fig-alt="Fantasy Football Projections Available From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>."}](https://apps.fantasyfootballanalytics.net)
+![](images/projectionsTab.png){fig-alt="Fantasy Football Projections Available From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>."}
 

### 2025-08-12 draft.qmd
20250812 - unlink figures

--- 2025-08-12 draft.qmd
@@
 ::: {#fig-snakeDraftToolFFA}
-![](images/snakeDraftTool.png){fig-alt="Free Snake Draft Tool From Fantasy Football Analytics. From <https://apps.fantasyfootballanalytics.net>."}
+![](images/snakeDraftTool.png){fig-alt="Free Snake Draft Tool From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>."}
 
-Free Snake Draft Tool From Fantasy Football Analytics. From <https://apps.fantasyfootballanalytics.net>.
+Free Snake Draft Tool From Fantasy Football Analytics: <https://apps.fantasyfootballanalytics.net>.
 :::

### 2025-08-12 draft.qmd
20250812 - additions

@@
 
-- your [evaluation of players](player-evaluation.qmd)
+- your [evaluation of players](player-evaluation.qmd) and of their situation
+    - It may make sense to prioritize players, in particular Running Backs, Wide Receivers, Quarterbacks, and Tight Ends, who are are on strong offenses.
 - [expert or aggregated rankings](#sec-fantasyValueRankings)
@@
 However, if your goal is to select the best players, leave your allegiances at the door.
-Selecting players based on their playing for your favorite team—rather than based on performance—is a form of [cognitive bias](#sec-cognitiveBias).
+Selecting players based on their playing for your favorite team—rather than based on performance—is a form of [cognitive bias](#sec-cognitiveBias) called [in-group bias](#sec-cognitiveBiasesInGroup).
 

### 2025-08-13 actuarial.qmd
20250813 - tweaks

@@
 Professional scouts and coaches use judgment when making predictions or selecting players based on their impressions of the players [@DenHartigh2018].
+Experts such as scouts and coaches are often former athletes who are presumed to have unique knowledge of their sport.
+Because of their expertise, it is possible that they could make observations that a non-expert could not and that could not be captured by statistics [@DenHartigh2018].
 As an example in popular media, in the movie, "Trouble with the Curve", a professional scout makes a judgment about how well a baseball hitter will do in the major leagues from his impressions of the hitter's ability based on the sound of the ball off the player's bat.
@@
 Common [heuristics](#sec-heuristics), [cognitive biases](#sec-cognitiveBiases), and [fallacies](sec-fallacies) are described in @sec-cognitiveBias.
+Judgments of sports athletes have been shown to be biased as a function of many characteristics of the athlete, including whether they are known to have a positive reputation, and their clothing, body language, and skin color [@DenHartigh2018].
 Below, I describe a few errors to which human judgment seems particularly prone.

### 2025-08-13 draft.qmd
20250813 - tweaks

@@
 Thanks, in part, to this draft strategy, the team showed strong extended success for nearly three decades from 1989 through the late-2010s.
+
+Another approach is to embrace variability and increase the variability of a team's outcomes by [stacking](#sec-stacking) players from the same team [@Lee2022].
 
@@
 
-It appears that Running Backs show the fastest dropoff (i.e., steepest negative slope), followed by Tight Ends, Quarterbacks, Wide Receivers, and then defensive positions.
+Running Backs show the fastest intial dropoff (i.e., steepest negative slope), followed by Tight Ends, Quarterbacks, Wide Receivers, and then defensive positions.
 So, identifying the top-scoring Running Back can be of great benefit.
+After the top 10 or so players at each position, Quarterbacks show a steeper dropoff than Running Backs, Tight Ends, and Wide Receivers.
 
@@
 
-Although Running Backs appear to show the fastest dropoff, Quarterbacks and Wide Receivers have appeared to overtake them in terms of fantasy points.
+Since 2020, although Running Backs show the fastest initial dropoff, Quarterbacks and Wide Receivers have appeared to overtake them in terms of fantasy points.
+Quarterbacks continue to show the steepest dropoff after the top 10 or so players at each position.
 
@@
 The trend suggests that, since 2000, Quarterbacks have increased in fantasy points scored, and Running Backs have decreased in fantasy points scored.
+
+Key takeaways include that, based on their steep initial dropoff, it can be of great benefit (in terms of a comparative advantage against your opponents) to have a top Running Back or Tight End.
+The challenge, of course, is identifying—in advance—which of the players is going to be the top-ranked player at that position.
+It also seems important to have a top-10 Quarterback, because Quarterbacks show the steepest dropoff after position rank of ~10.
+In general, given their tendency to score lots of points and to show a relatively steep dropoff, it seems crucial to have strong Quarterbacks, Wide Receivers, and Running Backs, with the highest priority likely going to Running Backs and Wide Receivers.
 

### 2025-08-13 mixed-models.qmd
20250813 - tweaks

@@
 For instance, you could estimate a three-level model of timepoints nested within players nested within teams.
-This would help account for players's productivity benefitting or taking a hit from being on a given team, and it would account for players changing teams.
+This would help account for players' productivity benefitting or taking a hit from being on a given team, and it would account for players changing teams.
 Alternatively, consistent with @Usami2018, you could estimate a cross-classified model of players' individual observations/measurement occasions belonging to two upper-level units: players and time points (e.g., week 1, week 2, 2025 season, etc.).

### 2025-08-13 modern-portfolio-theory.qmd
20250813 - tweaks

@@
 Stacking involves selecting players on the same team (e.g., multiple players on the Dallas Cowboys).
+Having two players from the same team increases the variance of the fantasy points scored, for better or for worse, because their performance is linked [i.e., can depend on one another; @Lee2022].
+Thus, when the players' respective team performs well, the players may score many points, greatly increasing the likelihood that a fantasy team wins; however, if the players' respective team does not perform well, the players may score few points, thus increasing the likelihood that the fantasy team loses.
+It is a higher-risk strategy—but one that can pay off if the players are on a high-scoring offense.
 To help inform the process of stacking, we evaluate the inter-correlation of performance in a given week across players on the same team and on opposing teams.

### 2025-08-13 player-evaluation.qmd
20250812 - linked figures
20250812 - unlink figures
20250813 - tweaks

@@
 
+We demonstrate how to download fantasy football projections in @sec-fantasyFootballProjections.
+
 #### Benefits of Using Projections Rather than Rankings {#sec-fantasyValueProjectionsVsRankings}
@@
 The signs approach to assessment involves assessing processes that may predict performance, where the emphasis is on what the sign indicates about the underlying attribute, rather than on the specific behavior itself.
-In football, the signs approach might involve measuring skills using separate tests where players' skills are tested in isolation.
+In football, the signs approach might involve measuring skills using separate tests where players' skills are tested separately (i.e., in isolation from other skills) in controlled settings.
 For instance, the NFL Combine takes a signs approach to assessment, in which players have a separate test to assess for speed (e.g., 40-yard dash) and jumping ability (e.g., vertical jump).
+The signs approach to assessment assumes that the various skills assessed (e.g., being fast, being able to jump high, having strong endurance, being motivated) have high [predictive validity](#sec-predictiveValidity) for the target behavior [i.e., future performance in games; @DenHartigh2018].
 

### 2025-08-13 getting-started.qmd
20250813 - clarification

@@
 ::: {#nte-gettingStarted .callout-note title="If you are in Professor Petersen's class"}
-If you are in Professor Petersen's class, also perform the following steps:
+If you are in Professor Petersen's class, also perform the following steps, to set up version control and your blog:
 

### 2025-08-15 data-visualization.qmd
20250815 - Okabe-Ito color palette

@@
 It is also important to use color schemes with distinguishable colors.
-Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by ColorBrewer (<https://colorbrewer2.org>) and are available using the `ggplot2::scale_color_brewer()` and `ggplot2::scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
-Good colorblind-friendly color schemes for binned (`_b`), continuous (`_c`), and discrete (i.e., categorical; `_d`) scales are available using the `ggplot2::scale_color_viridis_X()` and `ggplot2::scale_fill_viridis_X()` functions of the `ggplot2` package [@R-ggplot2] (where `X` is `b`, `c`, or `d`, depending on whether you want a binned, continuous, discrete scale).
+Good color schemes for sequential, diverging, and qualitative (i.e., categorical) data are provided by [ColorBrewer](https://colorbrewer2.org) (<https://colorbrewer2.org>) and are available using the `ggplot2::scale_color_brewer()` and `ggplot2::scale_fill_brewer()` functions of the `ggplot2` package [@R-ggplot2], as demonstrated in @fig-barPlotColorScheme (@sec-barPlotModifiedColorScheme).
+[ColorBrewer](https://colorbrewer2.org) allows users to select colorblind safe, print friendly, and photocopy safe palettes.
+Good colorblind-friendly color schemes for binned (`_b`), continuous (`_c`), and discrete (i.e., categorical; `_d`) scales are available from the [`viridis` palette](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html) using the `ggplot2::scale_color_viridis_X()` and `ggplot2::scale_fill_viridis_X()` functions of the `ggplot2` package [@R-ggplot2] (where `X` is `b`, `c`, or `d`, depending on whether you want a binned, continuous, discrete scale).
 There are a variety of resources for color schemes that are accessible to color-blind viewers:
@@
   - <https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html> (archived at <https://perma.cc/NK9K-HL7L>)
+- [Okabe-Ito color palette](https://jfly.uni-koeln.de/color/) [@Okabe2008; archived at <https://perma.cc/VD2E-XZZU>]
 - @Nunez2018
+
+The Okabe-Ito color palette [@Okabe2008] can be specified with the following colors:
+
+```{r}
+palette_OkabeIto <- c(
+  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
+  "#0072B2", "#D55E00", "#CC79A7", "#999999")
+
+palette_OkabeIto_black <- c(
+  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
+  "#0072B2", "#D55E00", "#CC79A7", "#000000")
+```
+
+The Okabe-Ito color palette [@Okabe2008] is depicted in Figures [-@fig-okabeItoPaletteGray] and [-@fig-okabeItoPaletteBlack].
+
+```{r}
+#| label: fig-okabeItoPaletteGray
+#| fig-cap: "Okabe-Ito Color Palette With Gray."
+#| fig-alt: "Okabe-Ito Color Palette With Gray."
+
+graphics::pie(rep(1, 8), col = palette_OkabeIto)
+```
+
+```{r}
+#| label: fig-okabeItoPaletteBlack
+#| fig-cap: "Okabe-Ito Color Palette With Black."
+#| fig-alt: "Okabe-Ito Color Palette With Black."
+
+graphics::pie(rep(1, 8), col = palette_OkabeIto_black)
+```
 

### 2025-08-15 basic-statistics.qmd
20250815 - correlation updates

@@
     &= \frac{\text{Cov}(x, y)}{s_x \cdot s_y} 
-  && \quad \text{replacing numerator with covariance notation}
+  && \quad \text{replacing numerator with covariance notation} \\
+    &= \frac{\sum z_x z_y}{n - 1}
 \end{aligned}
@@
 
-where $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), $\text{Cov}(x, y)$ is the covariance between $x$ and $y$, $s_x$ is the standard deviation of $x$, and $s_y$ is the standard deviation of $y$.
+where $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), $\text{Cov}(x, y)$ is the covariance between $x$ and $y$, $s_x$ is the standard deviation of $x$, $s_y$ is the standard deviation of $y$, $n$ is the sample size, $z_x$ is the *z*-score for $x$, and $z_y$ is the *z*-score for $y$.
 

### 2025-08-15 correlation.qmd
20250815 - correlation updates

@@
 
-Also note that, although the ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength of the (linear) association between two variables, it does not reflect the slope of that association, as depicted in @fig-correlationExamples.
+Also note that, although the ([Pearson](#sec-correlationExamplesPearson)) correlation reflects the strength of the (linear) association between two variables, it does not reflect the (unstandardized) slope of that association, as depicted in @fig-correlationExamples.
 
@@
 :::
+
+The [Pearson](#sec-correlationExamplesPearson) correlation coefficient can be greatly impacted by the extent of variability in the data, differences in the shapes of the two distributions, [nonlinearity](#sec-correlationExamplesNonlinear), [outliers](#sec-correlationOutliers), characteristics of the sample, and measurement error [@Goodwin2006].
+These factors tend to have greater impact when the sample size is smaller (compared to when the sample size is larger).
+The [Pearson](#sec-correlationExamplesPearson) correlation coefficient tends to be smaller when:
+
+- there is low variability in one or both variables (i.e., [restricted range](#sec-correlationRestrictedRange))
+- the two distributions show dissimilar shapes (in terms of [skewness and kurtosis](#sec-descriptiveStatisticsShape))
+- the underlying association is nonlinear, such that [Pearson](#sec-correlationExamplesPearson) correlation does not capture the association
+- [outliers](#sec-correlationOutliers) weaken the apparent linear trend
+- the association differs across subgroups of the sample
+- the data contain random measurement error
+
+By contrast, the association between two variables can be artificially inflated by factors such as:
+
+- [outliers](#sec-correlationOutliers) that happen to fall along a linear trend
+- systematic measurement error (e.g., common method bias)
+- third variable [confounding](#sec-causalDiagramConfounding)
 
@@
 
+You can compute a correlation manually using the *z*-scores, as in @eq-pearsonCorrelation:
+
+```{r}
+sum(scale(complete_cases$height) * scale(complete_cases$fantasyPoints)) / nrow(complete_cases)
+```
+
 If you just want the Pearson correlation between two variables, you can use the following syntax.
@@
 The correlation coefficient is strongly impacted by outliers—i.e., extreme (i.e., very large or very small) values relative to the other observations.
+Outliers can arise from errors that arise from data collection, measurement, or data entry, or they could represent valid (albeit atypical) values [@Goodwin2006].
+Outliers can either artificially weaken the estimate of association—if they weaken the apparent linear trend—or artificially strengthen it—if they happen to fall along a linear trend.
 Consider the following example:

### 2025-08-16 basic-statistics.qmd
20250816 - tweaks

@@
 There are many different types of scores and scales.
-This book focuses on [raw scores](#sec-rawScores) and [*z*-scores](#sec-zScores).
-For information on other scores and scales, including percentile ranks, *T*-scores, standard scores, scaled scores, and stanine scores, see here: <https://isaactpetersen.github.io/Principles-Psychological-Assessment/scoresScales.html#scoreTransformation> [@PetersenPrinciplesPsychAssessment].
+This book focuses on [raw scores](#sec-rawScores) and [*z* scores](#sec-zScores).
+For information on other scores and scales, including percentile ranks, *T* scores, standard scores, scaled scores, and stanine scores, see here: <https://isaactpetersen.github.io/Principles-Psychological-Assessment/scoresScales.html#scoreTransformation> [@PetersenPrinciplesPsychAssessment].
 
@@
 
-where $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), $\text{Cov}(x, y)$ is the covariance between $x$ and $y$, $s_x$ is the standard deviation of $x$, $s_y$ is the standard deviation of $y$, $n$ is the sample size, $z_x$ is the *z*-score for $x$, and $z_y$ is the *z*-score for $y$.
+where $x$ is the [predictor variable](#sec-correlationalStudy), $y$ is the [outcome variable](#sec-correlationalStudy), $\text{Cov}(x, y)$ is the covariance between $x$ and $y$, $s_x$ is the standard deviation of $x$, $s_y$ is the standard deviation of $y$, $n$ is the sample size, $z_x$ is the [*z* scores](#sec-zScores) for $x$, and $z_y$ is the [*z* scores](#sec-zScores) for $y$.
 

### 2025-08-16 factor-analysis.qmd
20250816 - tweaks

@@
 PCA seeks to identify components that explain variance in the data, so if the variables are not on the same scale, some variables may contribute considerably more variance than others.
-A common way of ensuring that variables are on the same scale is to standardize them using, for example,  [*z*-scores](#zScores) that have a mean of zero and standard deviation of one.
+A common way of ensuring that variables are on the same scale is to standardize them using, for example, [*z* scores](#sec-zScores) that have a mean of zero and standard deviation of one.
 By contrast, factor analysis can better accommodate variables that are on different scales.

### 2025-08-16 multiple-regression.qmd
20250816 - tweaks

@@
 
-Standardized regression coefficients can be obtained by standardizing the variables to [*z*-scores](#sec-zScores) so they all have a mean of zero and standard deviation of one.
+Standardized regression coefficients can be obtained by standardizing the variables to [*z* scores](#sec-zScores) so they all have a mean of zero and standard deviation of one.
 The standardized regression coefficient ($\beta$) is interpreted such that, for every standard deviation change in the [predictor variable](#sec-correlationalStudy), there is a __ standard deviation change in the [outcome variable](#sec-correlationalStudy).
@@
 Thus, standardized regression coefficients tend to be more comparable across variables and models compared to unstandardized regression coefficients.
-In this way, standardized regression coefficients provide a meaningful index of [effect size](#sec-practicalSignificance).
+In this way, standardized regression coefficients provide a meaningful index of [effect size](#sec-practicalSignificance) and can be used to identify the [predictors](#sec-correlationalStudy) with the strongest [predictive validity](#sec-predictiveValidity).
 
@@
 As noted in @sec-statisticalTestsPartitionedVariance, 
-The coefficient of determination ($R^2$) reflects the proportion of variance in the [outcome (dependent) variable](#sec-correlationalStudy) that is explained by the model predictions, as in @eq-coefficientOfDeterminationPartitionedVariance: $R^2 = \frac{\text{variance explained in }Y}{\text{total variance in }Y}$.
+The coefficient of determination ($R^2$) reflects the proportion of variance in the [outcome (dependent) variable](#sec-correlationalStudy) that is explained by the model predictions (i.e., by the [predictor variable(s)](#sec-correlationalStudy)), as in @eq-coefficientOfDeterminationPartitionedVariance: $R^2 = \frac{\text{variance explained in }Y}{\text{total variance in }Y}$.
 Various formulas for $R^2$ are in @eq-rSquared.
@@
   total := abs(direct) + abs(indirect)
+
+  # proportion mediated
+  Pm := abs(indirect) / total
 '
@@
   total := abs(direct) + abs(indirectTotal)
+
+  # proportion mediated
+  Pm1 := abs(indirect1) / total
+  Pm2 := abs(indirect2) / total
+  PmTotal := abs(indirectTotal) / total
 '

### 2025-08-16 basic-statistics.qmd
20250816 - comparing mixed models to ANOVA/multiple regression

@@
 Thus, in most cases, the regression framework can be used to examine most questions regarding associations between variables or differences between groups.
+Indeed, [*t*-test](#sec-tTest), [ANOVA](#sec-anova), and [regression](#sec-regression) are all special cases of the general linear model, which in turn, is a special case of [mixed models](#sec-mixedModels) without [random effects](#sec-mixedModelsOverview) [@Brauer2018].
 
@@
 For instance, you could use MANOVA to evaluate whether one's contract length and pay differ as a function of one's position.
+
+ANOVA has key limitations.
+ANOVA cannot handle continuous [independent](#sec-experiment) or [predictor](#sec-correlationalStudy) variables that vary within units [@Brauer2018].
+In addition, ANOVA poorly handles missing data—it uses listwise deletion and thus excludes participants with missing values on any of the [independent](#sec-experiment) or [dependent](#sec-experiment) variables [@Brauer2018].
 

### 2025-08-16 mixed-models.qmd
20250816 - comparing mixed models to ANOVA/multiple regression

@@
 Fixed effects allow the mean to differ from zero.
-Thus, inclusion of random effects without the corresponding fixed effect can lead to bias in estimation of the association between the predictor variables and the outcome variable.
+Thus, inclusion of random effects without the corresponding fixed effect can lead to bias in estimation of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy).
+
+Mixed modeling is a more flexible modeling framework than [analysis of variance](#sec-anova) (ANOVA).
+Unlike [ANOVA](#sec-anova), mixed modeling can accommodate both continuous and categorical [predictor variables](#sec-correlationalStudy) that vary within units [@Brauer2018].
+[ANOVA](#sec-anova), cannot handle continuous [predictor variables](#sec-correlationalStudy) that vary within units [@Brauer2018].
+In addition, [ANOVA](#sec-anova) yields biased inferences when the same participants are exposed to the same items or stimuli [@Brauer2018].
+Additionally, mixed models better handle missing data compared to [ANOVA](#sec-anova) [@Brauer2018].
 
@@
 
-Examples of Simpson's Paradox are depicted in Figures [-@fig-simpsonParadox] and [-@fig-simpsonsParadox].
-In the example below, there is a positive association between the predictor variable and outcome variable for each group.
-However, when the groups are combined, there is a *negative* association between predictor variable and outcome variable.
+Simpson's paradox occurs when the association between between the [predictor variable](#sec-correlationalStudy) and [outcome variable](#sec-correlationalStudy) for the subgroups differ from the association when the subgroups are combined.
+Examples of Simpson's paradox are depicted in Figures [-@fig-simpsonParadox] and [-@fig-simpsonsParadox].
+In the example below, there is a positive association between the [predictor variable](#sec-correlationalStudy) and [outcome variable](#sec-correlationalStudy) for each group.
+However, when the groups are combined, there is a *negative* association between [predictor variable](#sec-correlationalStudy) and [outcome variable](#sec-correlationalStudy).
 That is, the sign of an association can differ at different levels of analysis (e.g., group level versus person level).
@@
 
-Consider that we observe a between-person association between a predictor, for example, how much sports drink a player drinks and their performance, such that players who drink more sports drink before games tend to perform better during games.
+Consider that we observe a between-person association between a [predictor](#sec-correlationalStudy), for example, how much sports drink a player drinks and their performance, such that players who drink more sports drink before games tend to perform better during games.
 However, based on this association, if we draw the inference that sports drink consumption leads to better performance, this could be a faulty inference.
@@
 It is important to evaluate how much variance is attributable to the upper-level unit.
-A way of evaluating how much variance is attributed to an upper-level unit is with the intraclass correlation coefficient (ICC).
+A way of evaluating how much variance is attributable to an upper-level unit is with the intraclass correlation coefficient (ICC).
 If substantial variance is attributable to the upper-level unit, it is important to account for the nested structure of the data.
@@
 For fitting mixed models, the data should be in [long form](#sec-longWideForm)—each player (or unit of interest) should have multiple rows based on the grouping structure.
+The cluster variable is the variable or variables that represent the upper-level unit(s), within which the data are nested.
 When estimating mixed models, it is important to know the grouping structure of the data because the data structure influences the model syntax.
@@
 For instance, children may be nested within the crossing of schools and neighborhoods.
-That is, children are nested within schools, and children are also nested within neighborhoods; however, children attending the same school may not necessarily be from the same neighborhood.
+That is, children are nested within schools, and children are also nested within neighborhoods; however, children attending the same school may not necessarily be from the same neighborhood, and children from the same neighborhood may not necessarily attend the same school.
 The data would still have a two-level hierarchy, but the data would be nested in specific school-neighborhood combinations.
@@
 
-Applied to longitudinal fantasy football data, Most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players.
+Applied to longitudinal fantasy football data, Most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players, with the player's ID as the cluster variable.
 However, you could get more complicated.
@@
 
+## Handling of Missing Data {#sec-mixedModelMissingness}
+
+Mixed models handle missing data better than [ANOVA](#sec-anova) and [multiple regression](#sec-multipleRegression) [@Brauer2018].
+[ANOVA](#sec-anova) and [multiple regression](#sec-multipleRegression) use listwise deletion by default and, thus, exclude participants who have missing values on any of the variables in the model (i.e., [independent/dependent](#sec-experiment) or [predictor/outcome](#sec-correlationalStudy)).
+By contrast, with the data in [long form](#sec-longWideForm), mixed models make use of all rows that have values on all of the [predictor variables](#sec-correlationalStudy).
+That is, mixed models exclude only those rows that have a missing value on any of the [predictor variables](#sec-correlationalStudy).
+Unlike [ANOVA](#sec-anova) and [multiple regression](#sec-multipleRegression), mixed models retain rows with missing values on the [outcome variables](#sec-correlationalStudy).
+
 ### Comparing Model Fit {#sec-mixedModelCompareModels}
 
-When trying to determine what predictors to include and how complex of a model to fit, it can be helpful to compare a variety a models to determine how well they fit the data (i.e., how much variance they account for).
+When trying to determine what [predictors](#sec-correlationalStudy) to include and how complex of a model to fit, it can be helpful to compare a variety a models to determine how well they fit the data (i.e., how much variance they account for).
 To compare non-nested models, you can use fit criteria such as Akaike information criterion (AIC), which penalizes model complexity.
@@
 In this chapter, we use mixed models to demonstrate how to model trajectories (growth curves) of players' fantasy points as a function of age.
+When estimating growth curve models, time points are nested within individuals (i.e., players).
+Mixed models provide a flexible framework for estimating growth curve models because they allow unequal time intervals between measurement occasions, time intervals that differ between individuals, and different numbers of observations for each individual [@Brauer2018].
+
 One of the challenges when modeling players' fantasy points as a function of age is that fantasy points are a function of both ability and opportunity.
@@
 
-Based on the [scatterplot](#sec-scatterplot) (and the bivariate association below), Quarterbacks' fantasy points appear to increase with age.
+Based on the [scatterplot](#sec-scatterplot) (and the bivariate association below), Quarterbacks' fantasy points appear to (slightly) increase with age.
 
@@
 Mixed models allow accounting for multiple levels or units of analysis and to include both fixed and random effects.
-Inclusion of random effects allows the association between the predictor variables (the intercept and age) and the outcome variable (fantasy points) to differ for each individual in the grouping level (in this case, each player).
+Inclusion of random effects allows the association between the [predictor variables](#sec-correlationalStudy) (the intercept and age) and the [outcome variable](#sec-correlationalStudy) (fantasy points) to differ for each individual in the grouping level (in this case, each player).
 This allows for more accurately predicting phenomena.

### 2025-08-16 multiple-regression.qmd
20250816 - comparing mixed models to ANOVA/multiple regression

@@
 
+All statistical analyses follow the same basic structure, as in @eq-analysisBasicStructure [@Brauer2018]:
+
+$$
+\text{DATA} = \text{MODEL} + \text{ERROR}
+$$ {#eq-analysisBasicStructure}
+
 Regression with one [predictor variable](#sec-correlationalStudy) takes the form of @eq-regression:
@@
 Under the hood, multiple regression seeks to identify the best weight for each [predictor](#sec-correlationalStudy).
-The intercept is the expected value of the outcome variable when all of the predictor variables have a value of zero.
+The intercept is the expected value of the [outcome variable](#sec-correlationalStudy) when all of the [predictor variables](#sec-correlationalStudy) have a value of zero.
 
@@
 
-When the outcome variable is continuous, [linear regression](#sec-multipleRegressionLinear) is common.
+When the [outcome variable](#sec-correlationalStudy) is continuous, [linear regression](#sec-multipleRegressionLinear) is common.
 However, there are other types of regression depending on type and distribution of the [outcome variable](#sec-correlationalStudy).
@@
 Added-variable plots depict the *unique* association of each [predictor variables](#sec-correlationalStudy) with the [outcome variable](#sec-correlationalStudy) when controlling for all the other [predictor variables](#sec-correlationalStudy) in the model.
-Component-plus-residual plots depict partial residuals on the y-axis as a function of each [predictor variable](#sec-correlationalStudy) on the x-axis, where a partial residual for a given [predictor](#sec-correlationalStudy) is the effect of a given predictor (thus controlling for all the other [predictor variables](#sec-correlationalStudy) in the model) plus the residual from the full model.
+Component-plus-residual plots depict partial residuals on the y-axis as a function of each [predictor variable](#sec-correlationalStudy) on the x-axis, where a partial residual for a given [predictor](#sec-correlationalStudy) is the effect of a given [predictor](#sec-correlationalStudy) (thus controlling for all the other [predictor variables](#sec-correlationalStudy) in the model) plus the residual from the full model.
 
@@
 
-Let's first examine [descriptive statistics](#sec-descriptiveStatistics) of the predictor and outcome variables.
+Let's first examine [descriptive statistics](#sec-descriptiveStatistics) of the [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variables.
 
@@
 
-We can perform a dominance analysis to evaluate the relative importance of the predictors in the regression model.
+We can perform a dominance analysis to evaluate the relative importance of the [predictors](#sec-correlationalStudy) in the regression model.
 To perform the dominance analysis, we use the `parameters::dominance_analysis()` function of the `parameters` package [@R-parameters; @Luedecke2020_packages], which leverages the `domir` package [@R-domir].
@@
 
-We can calculate this by substituting in the regression coefficients and predictor values.
+We can calculate this by substituting in the regression coefficients and [predictor](#sec-correlationalStudy) values.
 Here is the computation:
@@
 
-To address multicollinearity, you can drop a redundant predictor or you can also use [principal component analysis](#sec-pca) or [factor analysis](#sec-factorAnalysis) of the [predictors](#sec-correlationalStudy) to reduce the [predictors](#sec-correlationalStudy) down to a smaller number of meaningful [predictors](#sec-correlationalStudy).
+To address multicollinearity, you can drop a redundant [predictor](#sec-correlationalStudy) or you can also use [principal component analysis](#sec-pca) or [factor analysis](#sec-factorAnalysis) of the [predictors](#sec-correlationalStudy) to reduce the [predictors](#sec-correlationalStudy) down to a smaller number of meaningful [predictors](#sec-correlationalStudy).
 For a meaningful answer regarding [predictors](#sec-correlationalStudy) in a regression framework that is precise and confident, you need a low level of intercorrelation among [predictors](#sec-correlationalStudy), unless you have a very large sample size.
@@
 
-- NOT a predictor of the outcome: `0`
-- cluster variable: `-2`
-- fixed effect of predictor: `1`
-- fixed effect and random effect of predictor: `2`
-- include cluster mean of predictor in addition to fixed effect of predictor: `3`
-- include cluster mean of predictor in addition to fixed effect and random effect of predictor: `4`
+- NOT a [predictor](#sec-correlationalStudy) of the outcome: `0`
+- [cluster variable](#sec-mixedModelDataStructure): `-2`
+- [fixed effect](#sec-mixedModelsOverview) of [predictor](#sec-correlationalStudy): `1`
+- [fixed effect](#sec-mixedModelsOverview) and [random effect](#sec-mixedModelsOverview) of [predictor](#sec-correlationalStudy): `2`
+- include cluster mean of [predictor](#sec-correlationalStudy) in addition to [fixed effect](#sec-mixedModelsOverview) of [predictor](#sec-correlationalStudy): `3`
+- include cluster mean of [predictor](#sec-correlationalStudy) in addition to [fixed effect](#sec-mixedModelsOverview) and [random effect](#sec-mixedModelsOverview) of [predictor](#sec-correlationalStudy): `4`
 

### 2025-08-16 mixed-models.qmd
20250816 - tweak

@@
 [ANOVA](#sec-anova), cannot handle continuous [predictor variables](#sec-correlationalStudy) that vary within units [@Brauer2018].
-In addition, [ANOVA](#sec-anova) yields biased inferences when the same participants are exposed to the same items or stimuli [@Brauer2018].
+In addition, [ANOVA](#sec-anova) yields biased inferences (i.e., a higher Type I error rate) when the same participants are exposed to multiple items or stimuli [i.e., responses by the same respondent tend to be correlated across items, and responses to the same item tend to be correlated across respondents; @Brauer2018].
 Additionally, mixed models better handle missing data compared to [ANOVA](#sec-anova) [@Brauer2018].

### 2025-08-17 causal-inference.qmd
20250809 - label functions
20250816 - additions

@@
 - a third variable (i.e., [confound](#sec-causalDiagramConfounding)), `Z`, influences both `X` and `Y`
-- the association between `X` and `Y` is spurious
+- the association between `X` and `Y` is spurious (noncausal and due to random chance)
 
@@
 We have a control group, but the control group does not have the same players as the intervention group, and it is impossible for a person to simultaneously receive and not receive treatment.
+Thus, an individual causal effect cannot typically be identified [@DOnofrio2020].
 
@@
 ## Approaches for Causal Inference {#sec-approachesCausalInference}
+
+Broadly, approaches for causal inference encompass 1) analysis approaches that account for measured factors and 2) designs that account for unmeasured factors [@DOnofrio2020].
 
@@
 We cannot directly test or establish causality from a non-experimental research design.
-Nevertheless, we can leverage various design features that, in combination with other studies using different research methods, collectively strengthen our ability to make causal inferences.
+Nevertheless, we can leverage various analysis approaches and design features that, in combination with other studies using different research methods, collectively strengthen our ability to make causal inferences.
 For instance, there are are no experiments in humans showing that smoking causes cancer—randomly assigning people to smoke or not smoke would not be ethical.
@@
 
-Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax):
+Here is an alternative way of specifying the same diagram (more similar to `lavaan` syntax) using the `ggdag::dagify()` and `ggdag::ggdag()` functions:
 
@@
 The common cause—i.e., the variable that influences the two variables—is known as a confound (or confounder).
+Nevertheless, causal effects and confounding—though competing hypotheses—are not mutually exclusive [@DOnofrio2020].
+A predictor can have an association with an outcome that is partly driven by confounding and partly driven by a causal effect.
+A key goal, then is determining the extent to which the observed association reflects confounding versus a causal effect.
 An example of confounding is depicted in @fig-counfounding:

### 2025-08-17 causal-inference.qmd
20250817 - additions

@@
 
-Although [experimental designs](#sec-causalInferenceExperiment) provide the strongest tests of causality, manytimes they are impossible, unethical, or impractical to conduct.
+Although [experimental designs](#sec-causalInferenceExperiment) provide the strongest tests of causality, many times they are impossible, unethical, or impractical to conduct.
 For instance, it would likely not be practical to randomly assign National Football League (NFL) players to either consume or not consume sports drink before their games.
 Players have their pregame rituals and routines and many would likely not agree to participate in such a study.
-Thus, we often rely on quasi-experimental designs such as natural experiments and [observational/correlational designs](#sec-correlationalStudy).
+Moreover, as described in @sec-internalExternalValidityTradeoffs, [experimental designs](#sec-causalInferenceExperiment) tend to be lower in [external validity](#sec-externalValidity) and thus their findings may not always generalize to how things occur in the real world.
+In addition, [experimental designs](#sec-causalInferenceExperiment) commonly have sample sizes that are too small to study rare-but-serious outcomes [@DOnofrio2020].
+Thus, we often rely on (and need!) quasi-experimental designs such as natural experiments and [observational/correlational designs](#sec-correlationalStudy).
 
@@
 The causal inference that smoking causes cancer was derived from a combination of experimental studies in rodents and observational studies in humans.
+To the extent that findings across designs that have different threats to [research design validity](#sec-researchDesignValidity) show similar results in terms of the association between the [predictor variable](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy), the case for causality is strengthened [@DOnofrio2020].
 
@@
 
-Another design feature of [longitudinal designs](#sec-causalInferenceLongitudinal) that can lead to greater [internal validity](#sec-internalValidity) is the use of within-subject analyses.
+Another design feature of [longitudinal designs](#sec-causalInferenceLongitudinal) that can lead to greater [internal validity](#sec-internalValidity) is the use of within-subject analyses, including interrupted time series designs [@DOnofrio2020].
 Between-subject analyses, might examine, for instance, whether players who consume more sports drink perform better on average compared to players who consume less sports drink.
@@
 For example, if you want to control for a player's education level in examining the association between sports drink consumption and performance, you could statistically control for player education by including the player's education level as a [covariate](#sec-covariates) in predicting performance.
-Such a model would examine whether, when accounting for education level, there is an association between sports drink consumption and player performance.
+Such a model would examine whether, when accounting for education level (by holding education level constant—i.e., partialing out the association between education level and player performance), there is an association between sports drink consumption and player performance.
 
-Another way to control for a variable is to stratify based on the variable—e.g., you can examine the association between the variables of interest within subgroups of the control variable.
+Another way to control for a variable is to conduct a stratified analysis based on the variable—e.g., you can examine the association between the variables of interest within subgroups of the control variable [@DOnofrio2020].
 For instance, taking the example of controlling for a player's education in examining the association between sports drink consumption and performance, you could examine the association between sports drink consumption and performance separately for various education levels: a) high school or lower, b) some college, c) college degree, d) graduate degree.
@@
 Genetically informed designs are not as relevant for fantasy football analytics, but they are useful to present as one of various design features that researchers can draw upon to strengthen their ability to make causal inferences.
+One of the limitations of genetically informed designs is that they cannot, without inclusion of measured factors, rule out the influence of factors that are not shared by family members [@DOnofrio2020].
 
@@
 For instance, from a DAG, it is possible to determine what variables it is important to control for (and how)—and just as importantly, what not to control for—in order to get unbiased estimates of the effect of one variable on another variable or of the association between two variables of interest.
+Thus, when designing and conducting studies, it is good to get in the habit of drawing causal diagrams to make your hypotheses, knowledge, and assumptions explicit [@DOnofrio2020].
+Using a causal diagram (DAG), it is also valuable to specify any potential alternative hypotheses, so that the researcher can design studies to evaluate and rule out potential alternative hypotheses.
+
 To create DAGs, you can use the `R` package `dagitty` [@R-dagitty; @Textor2016_packages] or the associated browser-based extension, DAGitty: <https://dagitty.net> (archived at <https://perma.cc/U9BY-VZE2>).
@@
 Nevertheless, causal effects and confounding—though competing hypotheses—are not mutually exclusive [@DOnofrio2020].
-A predictor can have an association with an outcome that is partly driven by confounding and partly driven by a causal effect.
+A [predictor](#sec-correlationalStudy) can have an association with an [outcome](#sec-correlationalStudy) that is partly driven by confounding and partly driven by a causal effect.
 A key goal, then is determining the extent to which the observed association reflects confounding versus a causal effect.
+
+Confounding is among the most common sources of bias in [observational studies](#sec-correlationalStudy) in terms of accurately estimating the causal effect of one variable on another, because among all the processes that influence one's exposure on a risk factor (i.e., [predictor variable](#sec-correlationalStudy)), some of the processes also likely influence the [outcome variable](#sec-correlationalStudy) [i.e., and are thus confounds; @DOnofrio2020].
+Two common sources of confounding that are often neglected are previous behavior and genetics [@DOnofrio2020].
+Part of the challenge in observational studies is that there can be known and unknown sources of confounding [@DOnofrio2020].
+In addition, the sources of counfounding can be either measured or unmeasured.
+Because some sources of confounding are likely to be unknown and/or unmeasured, it is unlikely that any single observational study can account for all sources of confounding [@DOnofrio2020].
+The causal diagram is thus useful for determining what sources of confounding a given study can and cannot address [@DOnofrio2020].
+
 An example of confounding is depicted in @fig-counfounding:
@@
 That is, to obtain an unbiased estimate of the causal association between player endurance (`X`) and field goals made (`Y`), it is necessary to control for stadium altitude (`Z`).
+
+Another challenge with confounding is that all constructs are measured with error.
+Thus, including a measure of a confound as a covariate in a model will only partially account for the confound.
+The remaining effect of the confound on the outcome (after controlling for the measured confound via a covariate) is called residual confounding [@DOnofrio2020].
+
+#### Systematic Measurement Error {#sec-systematicMeasurementError}
+
+Another common source of bias, in terms of accurately estimating the causal effect of one variable on another, that functions similarly to [confounding](#sec-causalDiagramConfounding) is systematic measurement error (as opposed to random measurement error).
+Systematic measurement error is measurement error that influences scores consistently for a person or across the sample.
+Systematic measurement error is described in [@PetersenPrinciplesPsychAssessment](https://isaactpetersen.github.io/Principles-Psychological-Assessment/reliability.html#systematicError).
+Let's consider that a researcher wants to study the effects of player stress on their happiness.
+The researcher selects a questionnaire that assesses player stress and a questionnaire that assesses happiness and assesses players on both.
+To some degree, the measurements will both be influenced by nonconstruct factors, because both measurements come from the same measurement method (questionnaire; i.e., common method bias) and from the same rater (i.e., common informant bias).
+People have response styles that lead them to have a tendency to rate themselves—regardless of the construct of interest—as better or worse than they actually are.
+For instance, Person A may rate themselves as better off in terms of stress and happiness (i.e., lower stress, higher happiness) than they actually are, whereas Person B may rate themselves as worse off in terms of stress and happiness (i.e., higher stress, lower happiness) than they actually are.
+Because these response styles influence both measurements and are not the construct of interest (i.e., stress or happiness), they function like [confounding](#sec-causalDiagramConfounding) in that the influence both variables and can explain, to some degree, why the variables are correlated for reasons other than a causal effect of player stress on their happiness.
+Unlike [confounding](#sec-causalDiagramConfounding), however, systematic measurement error does not just influence observational studies—it influences [experimental designs](#sec-experiment) as well.
 
@@
 That is, we should not control for injury status (`M`) when examining the association between diet (`X`) and coaching strategy.
-Controlling for the collider leads to confounding and can artificially induce an association between the two causes of the collider despite no causal association between them [@Lederer2019].
+Controlling for the collider leads to [confounding](#sec-causalDiagramConfounding) and can artificially induce an association between the two causes of the collider despite no causal association between them [@Lederer2019].
 Obtaining a distorted or artificial association between two variables due to inappropriately controlling for a collider is known as *collider bias*.

### 2025-08-17 correlation.qmd
20250816 - tweaks
20250817 - correlation ratio

@@
 
-You can compute a correlation manually using the *z*-scores, as in @eq-pearsonCorrelation:
+You can compute a correlation manually using the [*z* scores](#sec-zScores), as in @eq-pearsonCorrelation:
 
@@
 
-An example of a nonlinear correlation is a distance correlation and the xi ($\xi$) coefficient.
+Examples of a nonlinear correlation include the correlation ratio [$\eta$; @Goodwin2006], distance correlation, and the xi ($\xi$) coefficient [@Chatterjee2021].
 These estimates of nonlinear correlation only range from 0–1 (where 0 represents no association and 1 represents a strong association), so they are interpreted differently than a traditional correlation coefficient (which can be negative).
+
+Here is how to compute the correlation ratio (eta; $\eta$) using the `stats::aov()` and `effectsize::eta_squared()` functions:
+
+```{r}
+mod <- stats::aov(
+  fantasyPoints ~ height,
+  data = player_stats_seasonal)
+
+correlationRatio <- effectsize::eta_squared(mod)
+
+cols_to_sqrt <- intersect(
+  colnames(correlationRatio),
+  c("Eta2", "CI_low", "CI_high"))
+
+correlationRatio[cols_to_sqrt] <- sqrt(correlationRatio[cols_to_sqrt])
+
+correlationRatio <- correlationRatio %>%
+  rename(Eta = Eta2)
+
+correlationRatio
+```
+
 We can estimate a distance correlation using the `correlation::correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].
@@
 ```{r}
-correlation(
-  data = player_stats_seasonal %>% select(draftround, fantasyPoints),
+correlation::correlation(
+  data = player_stats_seasonal %>% select(height, fantasyPoints),
   method = "distance"
@@
 ```{r}
-correlation(
+correlation::correlation(
   data = data.frame(v1, v2),
@@
 
-correlation(
+correlation::correlation(
   data = data.frame(v1, v2),

### 2025-08-17 correlation.qmd
20250817 - chunk labels for troubleshooting

@@
 ```{r}
+#| label: correlation-ratio
+
 mod <- stats::aov(
@@
 ```{r}
+#| label: distance-correlation
+
 correlation::correlation(
-  data = player_stats_seasonal %>% select(height, fantasyPoints),
+  data = player_stats_seasonal %>% select(draftround, fantasyPoints),
   method = "distance"
@@
 ```{r}
+#| label: xi-coefficient
+
 set.seed(52242) # for reproducibility

### 2025-08-17 correlation.qmd
20250817 - correlation ratio

@@
 mod <- stats::aov(
-  fantasyPoints ~ height,
+  fantasyPoints ~ factor(height), # convert the predictor variable to a categorical variable (i.e., factor); if there are too many unique values, can also consider binning the variable into ranges
   data = player_stats_seasonal)
 
-correlationRatio <- effectsize::eta_squared(mod)
+etaSquared <- effectsize::eta_squared(mod)
+print(etaSquared, digits = 9)
+
+correlationRatio <- etaSquared
 
@@
 
-correlationRatio[cols_to_sqrt] <- sqrt(correlationRatio[cols_to_sqrt])
+correlationRatio[cols_to_sqrt] <- sqrt(correlationRatio[cols_to_sqrt]) # take the square root of eta-squared to get eta
 
@@
 
-correlationRatio
+print(correlationRatio, digits = 9)
+
+# the correlation ratio (eta) is equivalent to square root of R-squared (when the predictor is converted to a factor)
+Rsquared <- summary(stats::lm(
+  fantasyPoints ~ factor(height),
+  data = player_stats_seasonal))$r.squared
+
+Rsquared
+
+Rvalue <- sqrt(Rsquared)
+Rvalue
 ```

### 2025-08-17 correlation.qmd
20250817 - library statement

@@
 library("correlation")
+library("effectsize")
 library("GGally")

### 2025-08-18 cognitive-bias.qmd
20250809 - label functions
20250817 - illusion of validity

@@
 That is, one's confidence is not necessarily related to the likelihood that the judgment is correct—it often reflects the "illusion of validity" [@Kahneman2011].
+The illusion of validity is the unwarranted confidence that arises from a perceived good fit between the predicted outcome and the input information [@Tversky1974].
 As one gains more knowledge in a domain, the person also tends to become unrealistically overconfident in their ability to make accurate judgments and predictions [@Kahneman2011].
@@
 
-The `dbinom()` function in R provides the density of a binomial distribution.
+The `stats::dbinom()` function in R provides the density of a binomial distribution.
 A binomial distribution is the probability of a particular number of successes (e.g., coins flipping heads-up) given a certain number of independent trials.

### 2025-08-19 actuarial.qmd
20250818 - additions

@@
 In general, many so-called experts are "pseudo-experts" who do not know the boundaries of their competence—that is, they do not know what they do not know; they have the illusion of validity of their predictions and are [overconfident](#sec-cognitiveBiasesOverconfidence) about their predictions [@Kahneman2011].
+A person's confidence is not a good indicator of their accuracy [@Kahneman2011].
 Yet, many people arrogantly proclaim to have predictive powers, including in low-validity environments such as fantasy football.
@@
 Experts may attempt to be clever and to consider complex combinations of predictors, but doing so often reduces validity [@Kahneman2011].
-Simple combinations of predictions often outperform more complex combinations [@Kahneman2011].
+Simple combinations of predictions (e.g., with few predictors and/or predictors of equal weight) often outperform more complex combinations [@Dawes1979; @Kahneman2011].
 
@@
 Actuarial methods are especially important to use in low-validity environments (like fantasy football) in which there is considerable uncertainty and unpredictability [@Kahneman2011].
+Actuarial methods may outperform human judgment in low-validity environments because formulas are better able to detect weakly valid cues and to maintain a modest degree of accuracy using such cues consistently [@Kahneman2011].
 

### 2025-08-19 base-rates.qmd
20250819 - updates

@@
 
+People tend to make judgments based on the [representativeness](#sec-heuristicsRepresentativeness) (i.e., similarity to a prototype, or stereotyping) and [availability](#sec-heuristicsAvailability) heuristics rather than the base rate [@Kahneman2011].
+For instance, professional scouts often make judgments about players based in part on their build and look—i.e., whethey they look the part—rather than just based on their performance [@Kahneman2011].
+
 Another important phenomenon is that low base-rate events (i.e., unlikely events) are difficult to predict accurately.
@@
 In general, you should [anchor](#sec-heuristicsAnchoringAdjustment) your predictions to the [base rate](#sec-baseRate) and [adjust](#sec-heuristicsAnchoringAdjustment) from there.
+It is also important to question the validity of the evidence [@Kahneman2011].
 As noted by @Kahneman2011, if you have doubts about the quality of the evidence for a particular prediction question, keep your predictions close to the [base rate](#sec-baseRate), and modify them only modifying them based on the new information.

### 2025-08-19 cognitive-bias.qmd
20250819 - updates

@@
 In general, people tend to a) overestimate the likelihood of low base-rate events and b) overweight low base-rate events in their decisions [@Kahneman2011].
-This is especially true if the unlikely event attracts attention, because of the [confirmatory bias](#sec-cognitiveBiasesConfirmation) of memory [@Kahneman2011].
+This is especially true if the unlikely event attracts attention, because of the [confirmatory bias](#sec-cognitiveBiasesConfirmation) of memory—that is, when thinking about the event, you try to make it true in your mind [@Kahneman2011].
 
@@
 
-The [framing](#sec-cognitiveBiasesFraming) of an event can also alter its weighting.
-For instance, concrete representations tend to be weighted more heavily than abstract representations [@Kahneman2011].
+The [framing](#sec-cognitiveBiasesFraming) of an event can also alter its [availability](#sec-heuristicsAvailability) in mind, judgments of probability, and weighting in decisions [@Kahneman2011].
+A more vivid description of an outcome can lead people to be more influenced by its ease of recall (i.e., [availability](#sec-heuristicsAvailability)) than its objective probability [@Kahneman2011].
+In addition, concrete representations tend to be weighted more heavily than abstract representations [@Kahneman2011].
 Abstract representations are framed in terms of "how likely" the event will occur in in terms of chance, risk, probability, or likelihood.

### 2025-08-20 multiple-regression.qmd
20250820 - tweaks

@@
 
-Pairwise deletion (also called available case analysis) uses all available data for each pair of variables when estimating the covariances, and the resulting covariance matrix can be used to estimate a multiple regression model.
+Pairwise deletion (also called available case analysis) uses all available data for each pair of variables when estimating the correlations/covariances, and the resulting correlation/covariance matrix can be used to estimate a multiple regression model.
 We can estimate a regression model that uses pairwise deletion using the `lavaan` package [@Rosseel2012_packages; @R-lavaan].
@@
 
+- rows with non-zero values are the target variables to be imputed
 - columns with non-zero values are [predictors](#sec-correlationalStudy) of the variable specified in the given row

### 2025-08-21 base-rates.qmd
20250821 - Bayesian updating of fantasy points

@@
 
+Bayesian updating can also be applied to continuous outcomes like fantasy points.
+For an example of applying Bayesian updating to fantasy points, see @Braun2012 (archived at <https://web.archive.org/web/20161028142225/http://www.bayesff.com/bayesian101/>).
+Applying Bayes' theorem to continuous outcomes, the posterior distribution is equal to the prior distribution times the likelihood (data).
+For instance, we can start with our prior belief (distribution) for a player's performance based on, for example, average draft position.
+Then, we observe the Week 1 performance.
+For instance, if Tom Brady scores 35 points in Week 1, our likelihood represents the likelihood that Tom Brady would score 35 points (the data).
+Using Bayesian updating, we can then calculate a posterior distribution that represents our new best prediction moving forward for how many points Tom Brady will score in Week 2.
+We then observe Week 2, generate a new likelihood and posterior distribution, and use that as our new prior distribution for Week 3, and so on.
+To keep things simple, we use a binary outcome below to demonstrate Bayesian updating.
+
 To perform Bayesian updating involves comparing the relative probability of two outcomes, $P(C | R)$ versus $P(\text{not } C | R)$.
@@
 The calculation of posttest (posterior) probability is described in @sec-posttestProbability.
+In its calculation, the probability nomogram automatically converts the pretest probability (i.e., base rate) to prior (pretest) odds and the posterior (posttest) odds to posterior probability, so you do not have to.
 A *probability nomogram* (aka Fagan nomogram) is a way of visually applying [Bayes' theorem](#sec-bayesTheorem) to determine the posttest probability of having a condition based on the [pretest (or prior) probability](#sec-baseRate) and [likelihood ratio](#sec-diagnosticLikelihoodRatio), as depicted in @fig-probabilityNomogram.

### 2025-08-22 draft.qmd
20250822 - draft ADP fallers

@@
 For instance, if a particular Quarterback has a slightly higher [VORP](#sec-fantasyValueVORP) than a particular Running Back, but the Quarterback is likely to be available at the manager's next pick but the Running Back is likely to be unavailable at their next pick, it might make more sense to draft the Running Back.
+Another way to draft players who are a good value at a given pick is to identify players who have fallen past their [ADP](#sec-fantasyValueADPAAV) [@Harstad2023; archived at <https://perma.cc/9MAC-PUGL>].
+For instance, if you are on pick 30, drafting a player who has fallen past their ADP might mean drafting a player whose ADP is 22.
 

### 2025-08-22 basic-statistics.qmd
20250822 - clean up code

@@
 poly1 <- poly1[poly1$x >= z_crit, ]
-poly1<-rbind(poly1, c(z_crit, 0))  # add lower-left corner
+poly1 <- rbind(poly1, c(z_crit, 0))  # add lower-left corner
 
@@
 poly2 <- poly2[poly2$x <= z_crit,]
-poly2<-rbind(poly2, c(z_crit, 0))  # add lower-left corner
+poly2 <- rbind(poly2, c(z_crit, 0))  # add lower-left corner
 
@@
 poly3 <- poly3[poly3$x >= z_crit,]
-poly3 <-rbind(poly3, c(z_crit, 0))  # add lower-left corner
+poly3 <- rbind(poly3, c(z_crit, 0))  # add lower-left corner
 
@@
   # beta arrow
-  annotate("segment", x = 0.1, y = 0.045, xend = 1.3, yend = 0.01, arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
+  annotate("segment", x = 0.1, y = 0.045, xend = 1.3, yend = 0.01, arrow = grid::arrow(length = unit(0.3, "cm")), linewidth = 1) +
   annotate("text", label="beta", x=0, y=0.05, parse=T, size=8) +
   # alpha arrow
-  annotate("segment", x = 4, y = 0.043, xend = 3.4, yend = 0.01, arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
+  annotate("segment", x = 4, y = 0.043, xend = 3.4, yend = 0.01, arrow = grid::arrow(length = unit(0.3, "cm")), linewidth = 1) +
   annotate("text", label="frac(alpha,2)", x=4.2, y=0.05, parse=T, size=8) +
   # power arrow
-  annotate("segment", x = 6, y = 0.2, xend = 4.5, yend = 0.15, arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
+  annotate("segment", x = 6, y = 0.2, xend = 4.5, yend = 0.15, arrow = grid::arrow(length = unit(0.3, "cm")), linewidth = 1) +
   annotate("text", label = expression(paste(1-beta, "  (\"power\")")), x = 6.1, y = 0.21, parse = TRUE, size = 8) +

### 2025-08-22 getting-started.qmd
20250822 - clean up code

@@
      "gsisdecoder","progressr","DescTools","pwr","pwrss","WebPower","XICOR",
-     "dagitty","ggdag","ggtext","gghighlight","ggExtra","grid","patchwork",
-     "pROC", "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats",
-     "AICcmodavg","rstan","brms","tidybayes","bbmle","fitdistrplus","sn",
-     "mclust","magrittr","viridis","viridisLite","msir","plotly","webshot2",
-     "quantmod","fPortfolio","NMOF","nFactors","xts","zoo","forecast","stringi",
+     "dagitty","ggdag","ggtext","gghighlight","ggExtra","patchwork","pROC",
+     "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats","AICcmodavg",
+     "rstan","brms","tidybayes","bbmle","fitdistrplus","sn","mclust",
+     "magrittr","viridis","viridisLite","msir","plotly","webshot2","quantmod",
+     "fPortfolio","NMOF","nFactors","xts","zoo","forecast","stringi",
      "parallelly","doParallel","missRanger","ggridges","powerjoin","caret",

### 2025-08-22 index.qmd
20250818 - acknowledgments
20250818 - escape character
20250822 - clean up code

--- 2025-08-09 index.qmd
@@
   "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
-  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
+  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "patchwork", 
   "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
@@
   "MASS", "data.table", "future", "future.apply", "SimDesign", "domir", 
-  "GGally","Rglpk",
+  "GGally","Rglpk","TTR",
+  "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))
   "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
@@
 I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.
+In addition, a special thanks to those who contributed to the book via comments, suggestions, and pull requests, including George Ellis ([\@MargareeMan](https://github.com/MargareeMan)).
 Finally, I thank my wife who, in addition to designing the book cover, is a constant source of love and support.

### 2025-08-22 getting-started.qmd
20250822 - drop stringi

@@
      "magrittr","viridis","viridisLite","msir","plotly","webshot2","quantmod",
-     "fPortfolio","NMOF","nFactors","xts","zoo","forecast","stringi",
-     "parallelly","doParallel","missRanger","ggridges","powerjoin","caret",
-     "LongituRF","gpboost","mgcv","rms","car","lavaan","lavaanPlot","lavaangui",
+     "fPortfolio","NMOF","nFactors","xts","zoo","forecast","parallelly",
+     "doParallel","missRanger","ggridges","powerjoin","caret","LongituRF",
+     "gpboost","mgcv","rms","car","lavaan","lavaanPlot","lavaangui",
      "mice","miceadds","interactions","robustbase","ordinal","MASS",

### 2025-08-22 index.qmd
20250822 - drop stringi

@@
 
-# Packages to drop: stringi (replace with stringr from tidyverse); caret (replace with tidymodels)
+# Packages to drop: caret (replace with tidymodels)
 
@@
   "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
-  "stringi", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
+  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
   "caret", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", "lavaan", 

### 2025-08-23 evaluating-prediction-accuracy.qmd
20250822 - convert caret to tidymodels
To address https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues/10

@@
 library("msir")
-library("caret")
+library("tidymodels")
 library("tidyverse")
@@
 
-For example, here is an example of estimating predictive $R^2$ using 10-fold cross-validation using the `caret` package [@R-caret].
+For example, here is an example of estimating predictive $R^2$ using [10-fold cross-validation](#sec-kFoldCrossValidation) using the `tidymodels` ecosystem of packages [@R-tidymodels; @Kuhn2020_packages].
+
+```{r}
+# Set seed for reproducibility
+set.seed(52242)
+
+# Set up Cross-Validation
+folds_kFold <- rsample::group_vfold_cv(
+  projectionsWithActuals_seasonal,
+  group = player_id, # ensures all rows for a player are in the training set or all in the validation set for each fold
+  v = 10) # 10-fold cross-validation
+
+folds <- folds_kFold
+
+# Define Recipe (Formula)
+rec <- recipes::recipe(
+  fantasyPoints ~ raw_points, # use all predictors
+  data = projectionsWithActuals_seasonal)
+
+# Define Model
+lm_spec <- parsnip::linear_reg() %>%
+  parsnip::set_engine("lm") %>%
+  parsnip::set_mode("regression")
+
+# Workflow
+lm_wf <- workflows::workflow() %>%
+  workflows::add_recipe(rec) %>%
+  workflows::add_model(lm_spec)
+
+# Fit Model with Cross-Validation
+cv_results <- tune::fit_resamples(
+  lm_wf,
+  resamples = folds,
+  metrics = yardstick::metric_set(rmse, mae, rsq),
+  control = tune::control_resamples(save_pred = TRUE)
+)
+
+# View Cross-Validation metrics
+tune::collect_metrics(cv_results)
+
+# Fold-level R-squared
+tune::collect_metrics(cv_results, summarize = FALSE) %>%
+  filter(.metric == "rsq") %>%
+  select(id, .estimate)
+```
+
+Here is an example of estimating predictive $R^2$ with a training and hold-out (testing) data set:
+
+```{r}
+# Set seed for reproducibility
+set.seed(52242)
+
+# Create 75/25 split
+split <- rsample::group_initial_split(
+  projectionsWithActuals_seasonal,
+  group = player_id,
+  prop = 0.75)
+
+# Fit model on training, evaluate on testing
+last_fit_results <- tune::last_fit(
+  lm_wf,
+  split = split,
+  metrics = yardstick::metric_set(rmse, mae, rsq)
+)
+
+# View predictive R-squared and other metrics
+tune::collect_metrics(last_fit_results)
+```
+
+Or, alternatively:
 
@@
 
+# Set seed for reproducibility
 set.seed(52242)
 
-# Set up 10-fold cross-validation
-train_control <- caret::trainControl(
-  method = "cv",
-  number = 10)
-
-# Fit the model with cross-validation
-model_cv <- caret::train(
-  outcomeVariable ~ predictorVariable1 + predictorVariable2,
-  data = mydata,
-  method = "lm",
-  na.action = na.exclude,
-  trControl = train_control,
-  metric = "Rsquared")
-
-# View R-squared across the folds
-model_cv$results$Rsquared
-```
-
-Here is an example of estimating predictive $R^2$ with a training and testing data set using the `caret` package [@R-caret].
-
-```{r}
-#| eval: false
-
-set.seed(52242)
-
-inTrain <- caret::createDataPartition(
-  y = as.factor(mydata$outcomeVariable),
-  p = .75, # the proportion of data for the training data set
-  list = FALSE
-)
-
-training <- mydata[ inTrain,]
-testing  <- mydata[-inTrain,]
-
-training_model <- lm(
-  outcomeVariable ~ predictorVariable1 + predictorVariable2,
+# Create 75/25 split
+split <- rsample::group_initial_split(
+  projectionsWithActuals_seasonal,
+  group = player_id,
+  prop = 0.75)
+
+training <- rsample::training(split)
+testing  <- rsample::testing(split)
+
+# Fit model on training set
+training_model <- parsnip::fit(
+  lm_wf,
   data = training)
 
-testing_predictions <- predict(
+# Predict on test set
+predictions <- predict(
   training_model,
-  newdata = testing)
-
-petersenlab::accuracyOverall(
-  predicted = testing_predictions,
-  actual = testing$outcomeVariable)$rsquared
+  new_data = testing) %>%
+  bind_cols(testing)
+
+# Predictive R-squared
+yardstick::rsq(
+  predictions,
+  truth = fantasyPoints,
+  estimate = .pred)
 ```

### 2025-08-23 getting-started.qmd
20250822 - replace caret package with bestNormalize()
To drop caret package and finish https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues/10

@@
      "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats","AICcmodavg",
-     "rstan","brms","tidybayes","bbmle","fitdistrplus","sn","mclust",
-     "magrittr","viridis","viridisLite","msir","plotly","webshot2","quantmod",
-     "fPortfolio","NMOF","nFactors","xts","zoo","forecast","parallelly",
-     "doParallel","missRanger","ggridges","powerjoin","caret","LongituRF",
-     "gpboost","mgcv","rms","car","lavaan","lavaanPlot","lavaangui",
-     "mice","miceadds","interactions","robustbase","ordinal","MASS",
-     "data.table","future","future.apply","SimDesign","domir","GGally","Rglpk",
-     "TTR"))
+     "rstan","brms","tidybayes","bbmle","fitdistrplus","sn","mclust","magrittr",
+     "viridis","viridisLite","msir","plotly","webshot2","quantmod","fPortfolio",
+     "NMOF","nFactors","xts","zoo","forecast","parallelly","doParallel",
+     "missRanger","ggridges","powerjoin","bestNormalize","LongituRF","gpboost",
+     "mgcv","rms","car","lavaan","lavaanPlot","lavaangui","mice","miceadds",
+     "interactions","robustbase","ordinal","MASS","data.table","future",
+     "future.apply","SimDesign","domir","GGally","Rglpk","TTR"))
    ```

### 2025-08-23 index.qmd
20250822 - replace caret package with bestNormalize()
To drop caret package and finish https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues/10

@@
 
-# Packages to drop: caret (replace with tidymodels)
+# Packages to drop:
 
@@
   "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
-  "caret", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", "lavaan", 
-  "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", "ordinal", 
-  "MASS", "data.table", "future", "future.apply", "SimDesign", "domir", 
-  "GGally","Rglpk","TTR",
+  "bestNormalize", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", 
+  "lavaan", "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", 
+  "ordinal", "MASS", "data.table", "future", "future.apply", "SimDesign", 
+  "domir", "GGally","Rglpk","TTR",
   "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))

### 2025-08-23 multiple-regression.qmd
20250822 - replace caret package with bestNormalize()
To drop caret package and finish https://github.com/isaactpetersen/Fantasy-Football-Analytics-Textbook/issues/10

@@
 library("car")
-library("caret")
+library("bestNormalize")
 library("lme4")
@@
 
-We can apply a Yeo-Johnson transformation to the [outcome variable](#sec-correlationalStudy) to generate a more normally distributed variable.
-A Yeo-Johnson transformation estimates the optimal transformation to make the variable more normally distributed.
-Let's use a Yeo-Johnson transformation of fantasy points:
-
-```{r}
-yjTransformed <- caret::preProcess(
-  player_stats_seasonal["fantasyPoints"],
-  method = c("YeoJohnson"))
-
-yjTransformed
-
-player_stats_seasonal$fantasyPoints_transformed <- predict(
-  yjTransformed,
-  newdata = player_stats_seasonal["fantasyPoints"])$fantasyPoints
+We can apply a transformation to the [outcome variable](#sec-correlationalStudy) to generate a more normally distributed variable using the `bestNormalize::bestNormalize()` function of the `bestNormalize` package [@R-bestNormalize; @Peterson2021_packages; @Peterson2020_packages].
+
+```{r}
+obj <- bestNormalize::bestNormalize(player_stats_seasonal$fantasyPoints)
+player_stats_seasonal$fantasyPoints_transformed <- predict(obj)
+
+# Can back-transform transformed values back to the original metric later if needed:
+original <- predict(
+  obj,
+  newdata = player_stats_seasonal$fantasyPoints_transformed,
+  inverse = TRUE)
 ```

### 2025-08-23 multiple-regression.qmd
20250823 - flesh out transformation

@@
 ```{r}
-obj <- bestNormalize::bestNormalize(player_stats_seasonal$fantasyPoints)
-player_stats_seasonal$fantasyPoints_transformed <- predict(obj)
+bnTransformed <- bestNormalize::bestNormalize(player_stats_seasonal$fantasyPoints)
+```
+
+Here are the transformation results—in this case, the best-fitting transformation is an ordered quantile (ORQ) normalization transformation:
+
+```{r}
+bnTransformed
+```
+
+Then, we apply the best-fitting transformation to the variable, to create a transformed variable that is more normally distributed.
+
+```{r}
+player_stats_seasonal$fantasyPoints_transformed <- predict(bnTransformed)
 
@@
 original <- predict(
-  obj,
+  bnTransformed,
   newdata = player_stats_seasonal$fantasyPoints_transformed,

### 2025-08-23 sports-cognitive-psychology.qmd
20250823 - copyediting

@@
 "Choking" refers to underperforming relative to one's ability [@Beilock2001].
+
 Distraction theories posit that the pressure situation creates a distracting environment that shifts one's attention to cues that are irrelevant to the task, such as one's worries about the situation and its consequences [@Beilock2001].
+
 An alternative possibility, based on explicit monitoring theory, is that pressure situations raise one's self-awareness and anxiety about performing well, which increases one's attention to the step-by-step procedures involved in skill processes, which may disrupt well-learned or proceduralized performance [@Beilock2001].
@@
 putting in golf among highly trained individuals), focusing too much on the task can actually hinder performance, consistent with explicit monitoring theory.
+For highly practiced experts, automatic skill tasks might include tasks like hitting/putting a golf ball, pitching a baseball, throwing/kicking a football, shooting free throws in basketball, shooting arrows in archery, serving in tennis, etc.
+Sports players often refer to highly practiced tasks as involving "muscle memory."
 When a task is highly practiced or automatic, excessive attention can hurt performance.
@@
 This is called "choking under pressure" and is based on explicit monitoring theory.
-
-Explicit monitoring theories posit that pressure situations raises one's self-awareness and anxiety about performing well, which increases one's attention to the stop-by-step procedures involved in skill processes [@Beilock2001].
-For highly practiced experts, automatic skill tasks might include tasks like hitting/putting a golf ball, pitching a baseball, throwing/kicking a football, shooting free throws in basketball, shooting arrows in archery, etc.
-Sports players often refer to highly practiced tasks as involving "muscle memory."
 

### 2025-08-26 getting-started.qmd
20250826 - updated instructions

@@
 
-To get started, follow the following steps:
-
-1. Install `R`: <https://cran.r-project.org>
-1. Install `RStudio Desktop`: <https://posit.co/download/rstudio-desktop>
-1. After installing `RStudio`, open `RStudio` and run the following code in the console to install the `R` packages used in this book (note: this will take a while):
+To get started, follow the following steps (for each step that involves installing or running software, make sure to right click and select "Run as Administrator"):
+
+1. Download and install `R`: <https://cran.r-project.org>
+1. After installing `R`, download and install `RStudio Desktop`—this is a graphical user interface for `R`: <https://posit.co/download/rstudio-desktop>
+1. Install tools to allow you to compile `R` packages so you can install packages from source, if necessary (i.e., if package binaries are not available):
+    - If on Windows, download and install the version of `RTools` (<https://cran.r-project.org/bin/windows/Rtools/>) that corresponds to your version of `R`
+    - If on Mac, download and install `R Compiler Tools for Rcpp on MacOS`: <https://thecoatlessprofessor.com/programming/cpp/r-compiler-tools-for-rcpp-on-macos/>
+1.
+After installing `R` and `RStudio`, open `RStudio` (by right clicking and selecting "Run as Administrator") and run the following code in the console to install the `R` packages used in this book (note: this will take a while):
    ```{r}

### 2025-08-26 getting-started.qmd
20250826 - formatting

@@
 
-In base `R`, you can perform piping with the `base::\`|>\`` expression.
-In `tidyverse`, you can perform piping with the `magrittr::\`%>%\`` expression.
+In base `R`, you can perform piping with the ``base::`|>` `` expression.
+In `tidyverse`, you can perform piping with the ``magrittr::`%>%` `` expression.
 

### 2025-08-26 getting-started.qmd
20250826 - formatting

@@
 
+``base::`%in%` ``:
+
 ```{r}
@@
 
-The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] has a convenience function, `petersenlab::\`%ni%\``, for the logical operator of determining whether an element is *not* in a value of another vector:
+The [`petersenlab`](https://cran.r-project.org/web/packages/petersenlab/index.html) package [@R-petersenlab] has a convenience function, ``petersenlab::`%ni%` ``, for the logical operator of determining whether an element is *not* in a value of another vector:
 

### 2025-08-27 download-football-data.qmd
20250827 - references

@@
 
-Below, we provide examples for how to download various types of NFL data using the `nflreadr` [@R-nflreadr] and `nflfastR` [@R-nflfastR] packages.
+Below, we provide examples for how to download various types of NFL data using the `nflreadr` [@R-nflreadr] and `nflfastR` [@R-nflfastR] packages that are part of the `nflverse` ecosystem of packages [@R-nflverse].
 For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.

### 2025-08-27 getting-started.qmd
20250827 - references

@@
 - Video training courses in `R` skills: <https://www.pluralsight.com/search?q=R>
-- `R` for Data Science: <https://r4ds.hadley.nz>
+- `R for Data Science` book: <https://r4ds.hadley.nz>
+    - Tutorials from the book are available via the `r4ds.tutorials` package [@R-r4ds.tutorials]: <https://ppbds.github.io/r4ds.tutorials/>
 - Browse the `Cookbook for R` to find solutions to common tasks and problems: <http://www.cookbook-r.com>
@@
 
-For instance, you can use the following code to install the `tidyverse` package [@R-tidyverse]:
+For instance, you can use the following code to install the `tidyverse` package [@R-tidyverse; @Wickham2019_packages]:
 
@@
 In base `R`, you can perform piping with the ``base::`|>` `` expression.
-In `tidyverse`, you can perform piping with the ``magrittr::`%>%` `` expression.
+In the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages], you can perform piping with the ``magrittr::`%>%` `` expression.
 
@@
 
-In `tidyverse`, you cans select particular columns using the `dplyr::select()` function.
+In the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages], you can select particular columns using the `dplyr::select()` function.
 
@@
 
-In `tidyverse`, you cans select particular columns using the `dplyr::filter()` function.
+In the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages], you can select particular columns using the `dplyr::filter()` function.
 
@@
 
-In `tidyverse`, you cans select particular rows and columns using the `dplyr::filter()` and `dplyr::select()` functions, respectively.
+In the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages], you can select particular rows and columns using the `dplyr::filter()` and `dplyr::select()` functions, respectively.
 
@@
 The long data are in "`player`-`season`-`week` form".
-We widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable) using the `tidyr::pivot_wider()` function:
+We widen the data by two variables (`season` and `week`), using the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages], so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable) using the `tidyr::pivot_wider()` function:
 
@@
 Conversely, we can also restructure data from wide to long using the `tidyr::pivot_longer()` function.
-Here are the data in long form, after they have been transformed from wide form using `tidyverse`:
+Here are the data in long form, after they have been transformed from wide form using the `tidyverse` ecosystem of packages [@R-tidyverse; @Wickham2019_packages]:
 

### 2025-08-27 index.qmd
20250827 - references

@@
   "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
-  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
+  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic", "r4ds.tutorials" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
 )
@@
 This work would not be possible without the important contributions of many people.
-This book was made possible by the hard work of the package developers of the `nflverse` ecosystem of R packages, who made football data available, including Sebastian Carl, Ben Baldwin, Lee Sharpe, Tan Ho, and John Edwards.
+This book was made possible by the hard work of the package developers of the `nflverse` ecosystem of packages [@R-nflverse], who made football data available, including Sebastian Carl, Ben Baldwin, Lee Sharpe, Tan Ho, and John Edwards.
 I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.

### 2025-08-27 evaluating-prediction-accuracy.qmd
20250827 - tweaks

@@
 - [discrimination](#sec-discrimination):
-  - [adjusted $R^2$](#sec-adjustedRsquared)
+  - [standardized regression coefficient (beta)](#sec-standardizedRegressionCoefficient)
 - [calibration](#sec-calibration):
@@
 - [general accuracy](#sec-calibration):
+  - [calibration plot](#sec-calibrationPlot)
+  - [adjusted $R^2$](#sec-adjustedRsquared)
   - [MAE](#sec-meanAbsoluteError)

### 2025-08-28 actuarial.qmd
20250827 - redundancy

@@
 Another reason why judgment methods are more widely used than actuarial methods is that so-called "experts" (and people in general) show [overconfidence](#sec-cognitiveBiasesOverconfidence) in their predictions—clinicians, experts, and humans in general think they are more accurate than they actually are.
-We see this when examining their calibration; their predictions tend to be miscalibrated (this is also the case with fantasy football projections, as described in Section @sec-accuracyProjections).
+We see this when examining their calibration; their predictions tend to be miscalibrated (this is also the case with fantasy football projections, as described in @sec-accuracyProjections).
 For example, things they report with 80% confidence occur less than 80% of the time, an example of [overprecision](#sec-cognitiveBiasesOverconfidence) in their predictions.

### 2025-08-28 correlation.qmd
20250817 - function
20250827 - redundancy

@@
 
-A nonlinear correlation can be estimated using the xi ($\xi$) coefficient from the `XICOR` package [@R-XICOR; @Chatterjee2021]:
+A nonlinear correlation can be estimated using the xi ($\xi$) coefficient from the `XICOR::xicor()` function of the `XICOR` package [@R-XICOR; @Chatterjee2021]:
 
@@
 In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @sec-assumptionsRegression.
-In section @sec-multipleRegressionNonIndependence, we demonstrate how to account for non-independence of observations using cluster-robust standard errors.
+In @sec-multipleRegressionNonIndependence, we demonstrate how to account for non-independence of observations using cluster-robust standard errors.
 

### 2025-08-28 decision-making.qmd
20250827 - redundancy

@@
 Second, there are many players who are systematically *over*predicted (e.g., due to their injury) or *under*predicted (e.g., due to their becoming the starter after a teammate becomes injured, is traded, etc.)—an example of [overextremity miscalibration](#sec-calibration).
-In general, it is likely for players who are projected to score more points to be overpredicted and for players who are projected to score fewer points to be underpredicted, as described in Section @sec-accuracyProjections.
+In general, it is likely for players who are projected to score more points to be overpredicted and for players who are projected to score fewer points to be underpredicted, as described in @sec-accuracyProjections.
 Third, the experts may interact and discuss with one another.

### 2025-08-28 evaluating-prediction-accuracy.qmd
20250827 - redundancy

@@
 The formula for calculating sensitivity is in @eq-sensitivity.
-As described in Section @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), sensitivity decreases.
+As described in @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), sensitivity decreases.
 As the cutoff decreases, sensitivity increases.
@@
 The formula for calculating specificity is in @eq-specificity.
-As described in Section @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), specificity increases.
+As described in @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), specificity increases.
 As the cutoff decreases, specificity decreases.
@@
 The same measure will have a different PPV in different contexts with different [base rates](#sec-baseRate) [@Treat2023].
-As described in Section @sec-accuracyCutoff, as the [base rate](#sec-baseRate) increases, PPV increases.
+As described in @sec-accuracyCutoff, as the [base rate](#sec-baseRate) increases, PPV increases.
 As the [base rate](#sec-baseRate) decreases, PPV decreases.
 PPV also differs as a function of the cutoff.
-As described in Section @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), PPV increases.
+As described in @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), PPV increases.
 As the cutoff decreases (becomes more liberal), PPV decreases.
@@
 The same measure will have a different NPV in different contexts with different [base rates](#sec-baseRate) [@Treat2023].
-As described in Section @sec-accuracyCutoff, as the [base rate](#sec-baseRate) increases, NPV decreases.
+As described in @sec-accuracyCutoff, as the [base rate](#sec-baseRate) increases, NPV decreases.
 As the [base rate](#sec-baseRate) decreases, NPV increases.
 NPV also differs as a function of the cutoff.
-As described in Section @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), NPV decreases.
+As described in @sec-accuracyCutoff, as the cutoff increases (becomes more conservative), NPV decreases.
 As the cutoff decreases (becomes more liberal), NPV decreases.
@@
 Thus, I would not put much confidence in the predictions by fantasy football pundits.
-The most common pattern of experts' predictions is that they show overextremity, that is, their predictions have probability judgments that tend to be too extreme, as described in Section @sec-calibration.
+The most common pattern of experts' predictions is that they show overextremity, that is, their predictions have probability judgments that tend to be too extreme, as described in @sec-calibration.
 Overextremity of experts' predictions reflects the [overprecision](#sec-cognitiveBiasesOverconfidence) type of [overconfidence bias](#sec-cognitiveBiasesOverconfidence).

### 2025-08-29 getting-started.qmd
20250827 - typo
20250829 - info about functions

@@
     1. Select "Create repository"
-1. After creating the new repository, make sure you are on the page of of your new repository and complete the following steps:
+1. After creating the new repository, make sure you are on the page of your new repository and complete the following steps:
     1. Click "Settings" (in the top of the screen)
@@
 
+Most of the time for data analysis is typically spent getting the data ready for data analysis, including merging, cleaning, and processing.
+Thus, it is important to know how to perform common steps in this workflow.
+Below, we describe commonly used techniques for data processing and analysis.
+
 ## Install Packages {#sec-installingPackages}
 
+In order for you to be able to load a package (to make its functions available for use), you first must install it.
 You can install `R` packages using the `utils::install.packages()` function with the following syntax:
@@
 
-You can load a given library using the `base::library()` function with the following syntax:
+Loading a package makes its functions available for use.
+After you have installed a package, you can load the package using the `base::library()` function with the following syntax:
 
@@
 ## Using Functions and Arguments {#sec-functionsArguments}
+
+There are many packages available for `R`.
+A package is a collection functions.
+In `R`, functions are used to perform operations given various inputs.
+For example, there are functions available for performing arithmetic, manipulating data, working with text, and creating plots.
+The function that you would use depends on what operation you intend to perform.
+If there is an operation that you would like to perform, there is a good chance that someone has already developed an `R` function to perform it.
+You can often find packages that include the necessary function(s) for performing a given operation by performing a Google search.
+And, if a function does not already exist, you can create one, as described in @sec-createFunction.
 

### 2025-08-29 getting-started.qmd
20250829 - clarify inner join

@@
 
-An inner join includes all rows that are in **both** `x` **and** `y`.
-An inner join will return one row of `x` for each matching row of `y`, and can duplicate values of records on either side (left or right) if `x` and `y` have more than one matching record.
+An inner join includes only those rows in `x` that have a matching record in `y` (i.e., the record must be in both `x` and `y`).
+An inner join will return that rows of `x` and `y` that have matching records, and can duplicate values of records on either side (left or right) if `x` and `y` have more than one matching record.
 It returns columns from `x` and `y`.

### 2025-08-30 getting-started.qmd
20250830 - merge explanation and visual

@@
 
+If we have two data frames that we want to combine into one we would perform what is called a merge.
+Let's say we want to examine how players' fantasy performance differs by age.
+Assume we have players' age in one data frame and players' fantasy points in a separate data frame.
+To examine how age and fantasy performance is related, we will need to have both variables in the same data frame.
+Thus, to examine age in relation to fantasy performance, we would need to merge both data frames so that the variables age and fantasy points are in the same data frame.
+However, to merge successfully, we need to indicate how the rows in the first data frame can be connected to the rows in the second data frame.
+For instance, if the player's ID is a variable in both data frames, we could use the player's ID as a "key" to indicate how to merge the rows from both data frames.
+The merged object, then, would retain the keys (i.e., player ID) and the variables from each object (i.e., age and fantasy points).
+An example of merge the two data frames in this example is depicted in @fig-merge.
+
+::: {#fig-merge}
+![](images/merge.png){width=100% fig-alt="Example of Merging Two Data Frames."}
+
+"Example of Merging Two Data Frames."
+:::
+
+Here is a code example:
+
+```{r}
+dataframeA <- data.frame(
+  id = c(1, 2, 3),
+  age = c(24, 26, 30)
+)
+
+dataframeB <- data.frame(
+  id = c(1, 2, 3),
+  fantasyPoints = c(200, 300, 250)
+)
+
+mergedDataframe <- dplyr::full_join(
+  dataframeA,
+  dataframeB,
+  by = "id"
+)
+
+dataframeA
+
+dataframeB
+
+mergedDataframe
+```
+
 Merging (also called joining) merges two data objects using a shared set of variables called "keys."

### 2025-09-01 machine-learning.qmd
20250901 - tidymodeling

@@
 For each of the models, we fit and evaluate the models using the `tidymodels` ecosystem of packages [@R-tidymodels; @Kuhn2020_packages].
-we specify our model formula using the `recipes::recipe()` function of the `recipes` package [@R-recipes].
+Modeling using `tidymodels` is described in @Kuhn2023: <https://www.tmwr.org>.
+We specify our model formula using the `recipes::recipe()` function of the [`recipes`](recipes::recipes) package [@R-recipes].
 We define the model using the `parsnip::linear_reg()`, `parsnip::set_engine()`, and `parsnip::set_mode()` functions of the `parsnip` package [@R-parsnip].

### 2025-09-02 player-evaluation.qmd
20250902 - additions

@@
 - work ethic and mental toughness
+- competitiveness and "will to win"
 - incentives
@@
 A player's work ethic and mental toughness may help them be resilient and persevere in the face of challenges.
+The extent to which the player has a competitiveness, drive, or "will to win" may also influence their performance.
 Contact-related incentives may lead a player to put forth greater effort.

### 2025-09-03 download-football-data.qmd
20250902 - data dictionaries

@@
 
+You can access a data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_players
+```
+
+You can download the data using the following syntax:
+
 ```{r}
@@
 
-@HoFFPlayerIDs provide a Data Dictionary for fantasy player ID data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_playerids.html>
+@HoFFPlayerIDs provide a Data Dictionary for fantasy player ID data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_playerids.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_ff_playerids
+```
+
+You can download the data using the following syntax:
 
@@
 - <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#rosters> [@SharpeRosters]
+
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_roster_status
+nflreadr::dictionary_rosters
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoSchedules provide a Data Dictionary for game schedules data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_schedules.html>
+@HoSchedules provide a Data Dictionary for game schedules data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_schedules.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_schedules
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoCombine provide a Data Dictionary for data from the NFL Combine at the following link: <https://nflreadr.nflverse.com/articles/dictionary_combine.html>
+@HoCombine provide a Data Dictionary for data from the NFL Combine at the following link: <https://nflreadr.nflverse.com/articles/dictionary_combine.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_combine
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoDraftPicks provide a Data Dictionary for draft picks data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_draft_picks.html>
+@HoDraftPicks provide a Data Dictionary for draft picks data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_draft_picks.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_draft_picks
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoDepthCharts provide a Data Dictionary for data from weekly depth charts at the following link: <https://nflreadr.nflverse.com/articles/dictionary_depth_charts.html>
+@HoDepthCharts provide a Data Dictionary for data from weekly depth charts at the following link: <https://nflreadr.nflverse.com/articles/dictionary_depth_charts.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_depth_charts
+```
+
+You can download the data using the following syntax:
 
@@
 
-@SharpeTrades provides a Data Dictionary for NFL trade data at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#trades>
+@SharpeTrades provides a Data Dictionary for NFL trade data at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md#trades>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_trades
+```
+
+You can download the data using the following syntax:
 
@@
 We add a progress bar using the `progressr::with_progress()` function from the `progressr` package [@R-progressr] because it takes a while to run.
-@HoPBP provide a Data Dictionary for the play-by-play data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>
+@HoPBP provide a Data Dictionary for the play-by-play data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_pbp
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoParticipation provide a Data Dictionary for the participation data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_participation.html>
+@HoParticipation provide a Data Dictionary for the participation data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_participation.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_participation
+```
+
+You can download the data using the following syntax:
 
@@
 @HoPlayerStatsDefense provide a Data Dictionary for statistics for defensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_player_stats
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoInjuries provide a Data Dictionary for injury data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_injuries.html>
+@HoInjuries provide a Data Dictionary for injury data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_injuries.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_injuries
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoSnapCounts provide a Data Dictionary for snap counts data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_snap_counts.html>
+@HoSnapCounts provide a Data Dictionary for snap counts data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_snap_counts.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_snap_counts
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoQBR provide a Data Dictionary for ESPN QBR data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_espn_qbr.html>
+@HoQBR provide a Data Dictionary for ESPN QBR data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_espn_qbr.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_espn_qbr
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoNextGenStats provide a Data Dictionary for NFL Next Gen Stats data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_nextgen_stats.html>
+@HoNextGenStats provide a Data Dictionary for NFL Next Gen Stats data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_nextgen_stats.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_nextgen_stats
+```
+
+You can download the data using the following syntax:
 
@@
 
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_pfr_passing
+```
+
 @PFR2024 provides advanced stats from the 2024 season at the following link: <https://www.pro-football-reference.com/years/2024/advanced.htm>
+
+You can download the data using the following syntax:
 
@@
 
-@HoContracts provide a Data Dictionary for player contracts data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_contracts.html>
+@HoContracts provide a Data Dictionary for player contracts data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_contracts.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_contracts
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoFTNCharting provide a Data Dictionary for FTN Charting data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ftn_charting.html>
+@HoFTNCharting provide a Data Dictionary for FTN Charting data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ftn_charting.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_ftn_charting
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoFantasyProsRankings provide a Data Dictionary for FantasyPros ranking data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_rankings.html>
+@HoFantasyProsRankings provide a Data Dictionary for FantasyPros ranking data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_rankings.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_ff_rankings
+```
+
+You can download the data using the following syntax:
 
@@
 
-@HoOpportunity provide a Data Dictionary for expected fantasy points data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_opportunity.html>
+@HoOpportunity provide a Data Dictionary for expected fantasy points data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_ff_opportunity.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_ff_opportunity
+```
+
+You can download the data using the following syntax:
 

### 2025-09-03 download-football-data.qmd
20250902 - historical actual team stats

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
+```
+
+### Historical Actual Team Statistics {#sec-downloadActualTeamStats}
+
+@HoTeamStats provide a Data Dictionary for historical actual team statistics data at the following link: <https://nflreadr.nflverse.com/articles/dictionary_team_stats.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_team_stats
+```
+
+You can download the data using the following syntax:
+
+```{r}
+#| eval: false
+
+nfl_actualStats_team_weekly_raw <- progressr::with_progress(
+  nflreadr::load_team_stats(
+    seasons = TRUE,
+    summary_level = "week"))
+
+nfl_actualStats_team_seasonal_reg_raw <- progressr::with_progress(
+  nflreadr::load_team_stats(
+    seasons = TRUE,
+    summary_level = "reg"))
+
+nfl_actualStats_team_seasonal_regPost_raw <- progressr::with_progress(
+  nflreadr::load_team_stats(
+    seasons = TRUE,
+    summary_level = "reg+post"))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_team_weekly_raw,
+  nfl_actualStats_team_seasonal_reg_raw,
+  nfl_actualStats_team_seasonal_regPost_raw,
+  file = "./data/nfl_actualStats_team_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_team_raw.RData")
+```
+
+The `nfl_actualStats_team_weekly_raw` object is in `season`-`week`-`team` form.
+That is, each row should be uniquely identified by the combination of `season`, `week`, and `team`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_team_weekly <- nfl_actualStats_team_weekly_raw %>% 
+  select(season, week, team, everything()) %>% 
+  arrange(season, week, team)
+```
+
+Let's check for duplicate `season`-`week`-`team` instances:
+
+```{r}
+nfl_actualStats_team_weekly %>% 
+  group_by(season, week, team) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_actualStats_team_weekly) %>% as.data.frame()
+```
+
+The `nfl_actualStats_team_seasonal_reg_raw` and `nfl_actualStats_team_seasonal_regPost_raw` objects are in `season`-`team` form.
+That is, each row should be uniquely identified by the combination of `season` and `team`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_team_seasonal_reg <- nfl_actualStats_team_seasonal_reg_raw %>% 
+  select(season, team, everything()) %>% 
+  arrange(season, team)
+
+nfl_actualStats_team_seasonal_regPost <- nfl_actualStats_team_seasonal_regPost_raw %>% 
+  select(season, team, everything()) %>% 
+  arrange(season, team)
+```
+
+Let's check for duplicate `season`-`week`-`team` instances:
+
+```{r}
+nfl_actualStats_team_seasonal_reg %>% 
+  group_by(season, team) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_team_seasonal_regPost %>% 
+  group_by(season, team) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_team_weekly,
+  nfl_actualStats_team_seasonal_reg,
+  nfl_actualStats_team_seasonal_regPost,
+  file = "./data/nfl_actualStats_team.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_team.RData")
 ```

### 2025-09-03 cluster-analysis.qmd
20250809 - label functions
20250903 - updates to address updated data files

@@
 load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
-load(file = "./data/nfl_actualStats_career.RData")
+load(file = "./data/nfl_actualStats_player_career.RData")
 ```
@@
 load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
-load(file = "./data/nfl_actualStats_career.RData")
+load(file = "./data/nfl_actualStats_player_career.RData")
 ```
+
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
 
@@
 ##### Quarterbacks {#sec-clusterAnalysisExampleNumTiersQBs}
+
+We can perform a cluster analysis using the `mclust::mclustBIC()`, `mclust::mclustICL()`, and `mclust::mclustBootstrapLRT()` functions of the `mclust` package [@R-mclust; @Scrucca2023_packages].
 
@@
 In our data, all of the following models are equivalent—i.e., they result in the same unequal variance model with a 4-cluster solution—but they arrive there in different ways.
+We can fit the cluster model to the optimal number of tiers using the `mclust::Mclust()` function.
 

### 2025-09-03 download-football-data.qmd
20250902 - ESPN QBR
20250903 - updates to address updated data files

@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation_raw.RData", fsep = ""))
-load(file = "./data/nfl_actualStats_position_weekly_raw.RData")
+load(file = "./data/nfl_actualStats_player_seasonal_raw.RData")
+load(file = "./data/nfl_actualStats_player_weekly_raw.RData")
 load(file = "./data/nfl_actualStats_weekly_raw.RData")
@@
 # Calculated Data - Raw
-load(file = "./data/nfl_actualStats_career_raw.RData")
 load(file = "./data/nfl_actualStats_seasonal_raw.RData")
+#load(file = "./data/nfl_actualStats_player_career_raw.RData")
+load(file = "./data/nfl_actualStats_weekly_raw.RData")
 load(file = "./data/nfl_actualFantasyPoints_career_raw.RData")
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
-load(file = "./data/nfl_actualStats_position_weekly.RData")
+load(file = "./data/nfl_actualStats_player_seasonal.RData")
+load(file = "./data/nfl_actualStats_player_weekly.RData")
 load(file = "./data/nfl_actualStats_weekly.RData")
@@
 # Calculated Data - Processed
-load(file = "./data/nfl_actualStats_career.RData")
 load(file = "./data/nfl_actualStats_seasonal.RData")
+load(file = "./data/nfl_actualStats_player_career.RData")
+load(file = "./data/nfl_actualStats_weekly.RData")
 load(file = "./data/nfl_actualFantasyPoints_career.RData")
@@
 nfl_players <- nfl_players_raw %>% 
-  select(gsis_id, everything()) %>% 
+  select(gsis_id, everything()) %>% # NOTE: the data structure changed in 2025: https://github.com/nflverse/nflverse-players?tab=readme-ov-file#whats-new-in-players-v2
   arrange(display_name)
@@
 nfl_depthCharts <- nfl_depthCharts_raw %>% 
-  select(gsis_id, season, week, depth_position, everything()) %>% 
+  select(gsis_id, season, week, depth_position, everything()) %>% #NOTE: the data structure will change beginning in the 2025 season: https://nflreadr.nflverse.com/articles/dictionary_depth_charts.html
   arrange(full_name, gsis_id, season, week, depth_position)
@@
 
+#nfl_trades_raw <- progressr::with_progress(
+#  nflreadr::load_trades(seasons = TRUE))
+
 nfl_trades_raw <- read_csv("https://raw.githubusercontent.com/leesharpe/nfldata/master/data/trades.csv")
@@
   nflreadr::load_participation(
-    seasons = 2016:2023, # participation data are no longer available after 2023
+    seasons = 2016:nflreadr::most_recent_season(), # participation data available after each season has ended per changelog for nflreadr v1.5.0 (https://cran.r-project.org/web/packages/nflreadr/news/news.html)
     include_pbp = TRUE))
@@
 
-We calculate historical actual season-by-season statistics in @sec-calculateActualSeasonalStats.
+We can download historical season-by-season actual player statistics using the `nflreadr::load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
+Nevertheless, we could also calculate historical actual season-by-season statistics, as in @sec-calculateActualSeasonalStats.
+@HoPlayerStats provide a Data Dictionary for statistics for offensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
+@HoPlayerStatsDefense provide a Data Dictionary for statistics for defensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_player_stats
+```
+
+You can download the data using the following syntax:
+
+```{r}
+#| eval: false
+
+nfl_actualStats_player_seasonal_reg_raw <- progressr::with_progress(
+  nflreadr::load_player_stats(
+    seasons = TRUE,
+    summary_level = "reg"))
+
+nfl_actualStats_player_seasonal_regPost_raw <- progressr::with_progress(
+  nflreadr::load_player_stats(
+    seasons = TRUE,
+    summary_level = "reg+post"))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_player_seasonal_reg_raw, nfl_actualStats_player_seasonal_regPost_raw,
+  file = "./data/nfl_actualStats_player_seasonal_raw.Rdata"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_player_seasonal_raw.Rdata")
+```
+
+The `nfl_actualStats_player_seasonal` objects are in `player`-`season` form.
+That is, each row should be uniquely identified by the combination of `player_id` and `season`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_player_seasonal_reg <- nfl_actualStats_player_seasonal_reg_raw %>% 
+  rename(team = recent_team) %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+
+nfl_actualStats_player_seasonal_regPost <- nfl_actualStats_player_seasonal_regPost_raw %>% 
+  select(player_id, season, everything()) %>% 
+  arrange(player_display_name, player_id, season)
+```
+
+Let's check for duplicate `player`-`season` instances:
+
+```{r}
+nfl_actualStats_player_seasonal_reg %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_player_seasonal_regPost %>% 
+  group_by(player_id, season) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_actualStats_player_seasonal_reg) %>% as.data.frame()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_player_seasonal_reg, nfl_actualStats_player_seasonal_regPost,
+  file = "./data/nfl_actualStats_player_seasonal.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_player_seasonal.RData")
+```
 
@@
 
-nfl_actualStats_offense_weekly_raw <- progressr::with_progress(
+nfl_actualStats_player_weekly_raw <- progressr::with_progress(
   nflreadr::load_player_stats(
     seasons = TRUE,
-    stat_type = "offense"))
-
-nfl_actualStats_defense_weekly_raw <- progressr::with_progress(
-  nflreadr::load_player_stats(
-    seasons = TRUE,
-    stat_type = "defense"))
-
-nfl_actualStats_kicking_weekly_raw <- progressr::with_progress(
-  nflreadr::load_player_stats(
-    seasons = TRUE,
-    stat_type = "kicking"))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_offense_weekly_raw, nfl_actualStats_defense_weekly_raw, nfl_actualStats_kicking_weekly_raw,
-  file = "./data/nfl_actualStats_position_weekly_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_position_weekly_raw.RData")
-```
-
-The `nfl_actualStats_weekly` objects are in `player`-`season`-`week` form.
+    summary_level = "week"))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_player_weekly_raw,
+  file = "./data/nfl_actualStats_player_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_player_weekly_raw.RData")
+```
+
+The `nfl_actualStats_weekly` object is in `player`-`season`-`week` form.
 That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
@@
 ```{r}
-nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly_raw %>% 
-  rename(team = recent_team) %>% 
+nfl_actualStats_player_weekly <- nfl_actualStats_player_weekly_raw %>% 
   select(player_id, season, week, everything()) %>% 
   arrange(player_display_name, player_id, season, week)
-
-nfl_actualStats_defense_weekly <- nfl_actualStats_defense_weekly_raw %>% 
-  select(player_id, season, week, everything()) %>% 
-  arrange(player_display_name, player_id, season, week)
-
-nfl_actualStats_kicking_weekly <- nfl_actualStats_kicking_weekly_raw %>% 
-  select(player_id, season, week, everything()) %>% 
-  arrange(player_display_name, player_id, season, week)
 ```
@@
 ```{r}
-nfl_actualStats_offense_weekly %>% 
+nfl_actualStats_player_weekly %>% 
   group_by(player_id, season, week) %>% 
@@
   head()
-
-nfl_actualStats_defense_weekly %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_kicking_weekly %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
 ```
@@
 ```{r}
-names(nfl_actualStats_offense_weekly) %>% as.data.frame()
-
-names(nfl_actualStats_defense_weekly) %>% as.data.frame()
-
-names(nfl_actualStats_kicking_weekly) %>% as.data.frame()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_offense_weekly, nfl_actualStats_defense_weekly, nfl_actualStats_kicking_weekly,
-  file = "./data/nfl_actualStats_position_weekly.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_position_weekly.RData")
+names(nfl_actualStats_player_weekly) %>% as.data.frame()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_player_weekly,
+  file = "./data/nfl_actualStats_player_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_player_weekly.RData")
 ```
@@
 ### ESPN QBR {#sec-espnQBR}
+
+ESPN provides Quarterbacks Ratings (QBR).
+QBR is described by @Katz2016 (archived at <https://perma.cc/4TP6-E57U>).
 
@@
 
-First, we can compute the players' season-by-season statistics using the `nflfastR` [@R-nflfastR] package.
+We already load players' season-by-season statistics [above](#sec-downloadActualSeasonalStats) using the `nflreadr::load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
+Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `nflfastR::calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:
 
@@
 ```{r}
-nfl_stats_variables
+nflfastR::nfl_stats_variables
 ```
@@
   nfl_actualStats_career_player, nfl_actualStats_career_player_inclPost,
-  file = "./data/nfl_actualStats_career.RData"
-)
-```
-
-```{r}
-#| eval: false
-#| include: false
-
-load(file = "./data/nfl_actualStats_career.RData")
+  file = "./data/nfl_actualStats_player_career.RData"
+)
+```
+
+```{r}
+#| eval: false
+#| include: false
+
+load(file = "./data/nfl_actualStats_player_career.RData")
 ```
@@
 
-We already load players' week-by-week statistics [above](#sec-downloadActualStats) using the `nflreadr::load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
+We already load players' week-by-week statistics [above](#sec-downloadActualWeeklyStats) using the `nflreadr::load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
 Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `nflfastR::calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:
@@
   player_game_birthdate_gameday,
-  nfl_players %>% select(-birth_date, -team_abbr, - team_seq),
+  nfl_players %>% select(-birth_date, -latest_team),
   by = c("player_id" = "gsis_id"))
@@
   nfl_actualFantasyPoints_player_seasonal,
-  nfl_players %>% select(-position, -position_group, -team_abbr, - team_seq),
+  nfl_players %>% select(-position, -position_group, -latest_team),
   by = c("player_id" = "gsis_id")

### 2025-09-03 draft.qmd
20250903 - copyediting
20250903 - updates to address updated data files

@@
 ```
+
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
+The `players_projections_seasonal` object was derived from projected points objects that were created in @sec-fantasyFootballProjections.
 
@@
 For [auction drafts](#sec-draftAuction) and daily fantasy sports (DFS) leagues, a key goal is to select the best lineup or team within the given cap.
-For instance, in an [auction drafts](#sec-draftAuction), you may have a \$200 cap and your goal is to draft the best team that you can with the \$200.
+For instance, in an [auction draft](#sec-draftAuction), you may have a \$200 cap and your goal is to draft the best team that you can with the \$200.
 Using a mixed-integer programming solver, we can identify the lineup with the most projected points given a particular salary cap.

### 2025-09-03 machine-learning.qmd
20250901 - fix link
20250903 - updates to address updated data files

@@
 # Calculated Data - Processed
-load(file = "./data/nfl_actualStats_career.RData")
+load(file = "./data/nfl_actualStats_player_career.RData")
 load(file = "./data/nfl_actualStats_seasonal.RData")
@@
 ```
+
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
 
@@
 
-nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
-  rename(gsis_id = player_id)
-
 ## Rename other variables to ensure common names
@@
     jersey_number = as.character(jersey_number),
-    gsis_it_id = as.character(gsis_it_id),
+    nfl_id = as.character(nfl_id),
     years_of_experience = as.integer(years_of_experience))
@@
     birth_date = as.Date(birth_date),
-    jersey_number = as.character(jersey_number),
-    gsis_it_id = as.character(gsis_it_id))
+    jersey_number = as.character(jersey_number))
 
@@
     by = c("gsis_id"),
-    conflict = coalesce_xy) # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
+    conflict = powerjoin::coalesce_xy) # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
 
@@
     by = c("gsis_id","season"),
-    conflict = coalesce_xy) # where the objects have the same variable name (e.g., team), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
+    conflict = powerjoin::coalesce_xy) # where the objects have the same variable name (e.g., team), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
 
@@
     by = c("gsis_id","season","week"),
-    conflict = coalesce_xy) # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
+    conflict = powerjoin::coalesce_xy) # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
 ```
@@
   playerSeasonMerged,
-  playerMerged %>% select(-age, -years_of_experience, -team, -team_abbr, -team_seq, -current_team_id), # drop variables from id objects that change from year to year (and thus are not necessarily accurate for a given season)
+  playerMerged %>% select(-age, -years_of_experience, -team, -latest_team, -last_season, -pff_status), # drop variables from id objects that change from year to year (and thus are not necessarily accurate for a given season)
   by = "gsis_id",
-  conflict = coalesce_xy # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
+  conflict = powerjoin::coalesce_xy # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
 ) %>% 
@@
   by = c("gsis_id","season"),
-  conflict = coalesce_xy # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
+  conflict = powerjoin::coalesce_xy # where the objects have the same variable name (e.g., position), keep the values from object 1, unless it's NA, in which case use the relevant value from object 2
 ) %>% 
@@
 dropVars <- c(
-  "birth_date", "loaded", "full_name", "player_name", "player_display_name", "display_name", "suffix", "headshot_url", "player", "pos",
-  "espn_id", "sportradar_id", "yahoo_id", "rotowire_id", "pff_id", "fantasy_data_id", "sleeper_id", "pfr_id",
-  "pfr_player_id", "cfb_player_id", "pfr_player_name", "esb_id", "gsis_it_id", "smart_id",
-  "college", "college_name", "team_abbr", "current_team_id", "college_conference", "draft_club", "status_description_abbr",
-  "status_short_description", "short_name", "headshot", "uniform_number", "jersey_number", "first_name", "last_name",
-  "football_name", "team")
+  "birth_date", "player_display_name", "team", "player_name", "headshot_url", "season_type", "fg_made_list", "fg_missed_list", "fg_blocked_list", "gwfg_distance_list", "pff_status", "startdate", "pos", "merge_name", "pfr_player_id", "cfb_player_id", "hof", "category", "side", "college", "car_av", "display_name", "common_first_name", "first_name", "last_name", "short_name", "football_name", "suffix", "esb_id", "nfl_id", "pff_id", "otc_id", "espn_id", "smart_id", "ngs_position_group", "ngs_position", "headshot", "college_name", "college_conference", "jersey_number", "status", "ngs_status", "ngs_status_short_description", "pff_position", "draft_team", "pfr_player_name", "pfr_id")
 
@@
     gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
-    height, weight, rookie_year, draft_number,
+    height, weight, rookie_season, draft_pick,
     fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
@@
 Modeling using `tidymodels` is described in @Kuhn2023: <https://www.tmwr.org>.
-We specify our model formula using the `recipes::recipe()` function of the [`recipes`](recipes::recipes) package [@R-recipes].
+We specify our model formula using the `recipes::recipe()` function of the `recipes` package [@R-recipes].
 We define the model using the `parsnip::linear_reg()`, `parsnip::set_engine()`, and `parsnip::set_mode()` functions of the `parsnip` package [@R-parsnip].

### 2025-09-03 pca.qmd
20250816 - tweaks
20250903 - updates to address updated data files

@@
 ```
+
+We created the `player_stats_weekly.RData` object in @sec-calculatePlayerAge.
 
@@
 PCA seeks to identify components that explain variance in the data, so if the variables are not on the same scale, some variables may contribute considerably more variance than others.
-A common way of ensuring that variables are on the same scale is to standardize them using, for example,  [*z*-scores](#sec-zScores).
+A common way of ensuring that variables are on the same scale is to standardize them using, for example,  [*z* scores](#sec-zScores).
 

### 2025-09-03 base-rates.qmd
20250903 - fix variable names

@@
   filter(years_of_experience == 0) %>% 
-  filter(draftround %in% c(1,2)) %>% 
+  filter(draft_round %in% c(1,2)) %>% 
   arrange(player_id, season) %>% 
@@
   filter(years_of_experience > 0) %>% 
-  filter(draftround %in% c(1,2)) %>% 
+  filter(draft_round %in% c(1,2)) %>% 
   mutate(

### 2025-09-03 correlation.qmd
20250903 - fix variable names

@@
 cor(
-  player_stats_seasonal[,c("draftround","fantasyPoints")],
+  player_stats_seasonal[,c("draft_round","fantasyPoints")],
   use = "pairwise.complete.obs",
@@
 cor.test(
-  ~ draftround + fantasyPoints,
+  ~ draft_round + fantasyPoints,
   data = player_stats_seasonal,
@@
 spearman_test <- cor.test(
-  ~ draftround + fantasyPoints,
+  ~ draft_round + fantasyPoints,
   data = player_stats_seasonal,
@@
 spearman_df <- cor.test(
-  ~ draftround + fantasyPoints,
+  ~ draft_round + fantasyPoints,
   data = player_stats_seasonal)$parameter
@@
   data = player_stats_seasonal,
-  x = c("draftround","fantasyPoints"),
+  x = c("draft_round","fantasyPoints"),
   y = "age",
@@
 correlation::correlation(
-  data = player_stats_seasonal %>% select(draftround, fantasyPoints),
+  data = player_stats_seasonal %>% select(draft_round, fantasyPoints),
   method = "distance"

### 2025-09-03 multiple-regression.qmd
20250823 - fix error
20250903 - fix variable names

@@
     rush_40_yds, rec_40_yds, fumbles, two_pts, return_yds,
-    rush_100_yds:draftround, height:target_share_log) %>% 
+    rush_100_yds:draft_round, height:target_share_log) %>% 
   select(-c(fantasy_points_ppr, ageCentered20, target_share_log)) # drop collinear variables
@@
 imp_long$fantasyPoints_transformed <- predict(
-  yjTransformed,
-  newdata = imp_long["fantasyPoints"])$fantasyPoints
+  bnTransformed,
+  newdata = imp_long[["fantasyPoints"]])
 
@@
   )
+
+mice::pool(imp_regression)
 

### 2025-09-04 download-football-data.qmd
20250904 - week 1 projected points

@@
 
+We can download historical actual player statistics as a function of different timeframes, including [week-by-week](#sec-downloadActualWeeklyStats) and [season-by-season](#sec-downloadActualSeasonalStats).
+
+#### Week-by-Week Statistics {#sec-downloadActualWeeklyStats}
+
+We can download historical week-by-week actual player statistics using the `nflreadr::load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
+Nevertheless, we could also calculate historical actual week-by-week statistics, as in @sec-calculateActualWeeklyStats
+@HoPlayerStats provide a Data Dictionary for statistics for offensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
+@HoPlayerStatsDefense provide a Data Dictionary for statistics for defensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.
+You can also access the data dictionary via `R` using the following syntax:
+
+```{r}
+#| eval: false
+
+nflreadr::dictionary_player_stats
+```
+
+You can download the data using the following syntax:
+
+```{r}
+#| eval: false
+
+nfl_actualStats_player_weekly_raw <- progressr::with_progress(
+  nflreadr::load_player_stats(
+    seasons = TRUE,
+    summary_level = "week"))
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_player_weekly_raw,
+  file = "./data/nfl_actualStats_player_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_player_weekly_raw.RData")
+```
+
+The `nfl_actualStats_weekly` object is in `player`-`season`-`week` form.
+That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_player_weekly <- nfl_actualStats_player_weekly_raw %>% 
+  select(player_id, season, week, everything()) %>% 
+  arrange(player_display_name, player_id, season, week)
+```
+
+Let's check for duplicate `player`-`season`-`week` instances:
+
+```{r}
+nfl_actualStats_player_weekly %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_actualStats_player_weekly) %>% as.data.frame()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_player_weekly,
+  file = "./data/nfl_actualStats_player_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_player_weekly.RData")
+```
+
 #### Season-by-Season Statistics {#sec-downloadActualSeasonalStats}
@@
 We calculate historical actual career statistics in @sec-calculateActualCareerStats.
-
-#### Week-by-Week Statistics {#sec-downloadActualWeeklyStats}
-
-We can download historical week-by-week actual player statistics using the `nflreadr::load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
-@HoPlayerStats provide a Data Dictionary for statistics for offensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
-@HoPlayerStatsDefense provide a Data Dictionary for statistics for defensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.
-You can also access the data dictionary via `R` using the following syntax:
-
-```{r}
-#| eval: false
-
-nflreadr::dictionary_player_stats
-```
-
-You can download the data using the following syntax:
-
-```{r}
-#| eval: false
-
-nfl_actualStats_player_weekly_raw <- progressr::with_progress(
-  nflreadr::load_player_stats(
-    seasons = TRUE,
-    summary_level = "week"))
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_player_weekly_raw,
-  file = "./data/nfl_actualStats_player_weekly_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_player_weekly_raw.RData")
-```
-
-The `nfl_actualStats_weekly` object is in `player`-`season`-`week` form.
-That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_actualStats_player_weekly <- nfl_actualStats_player_weekly_raw %>% 
-  select(player_id, season, week, everything()) %>% 
-  arrange(player_display_name, player_id, season, week)
-```
-
-Let's check for duplicate `player`-`season`-`week` instances:
-
-```{r}
-nfl_actualStats_player_weekly %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Here are the variable names:
-
-```{r}
-names(nfl_actualStats_player_weekly) %>% as.data.frame()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_player_weekly,
-  file = "./data/nfl_actualStats_player_weekly.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_player_weekly.RData")
-```
 
@@
   add_ecr() %>% 
-  #add_uncertainty() %>% # currently throws an error
+  add_uncertainty() %>% # comment out if throws an error
   add_player_info()
@@
 
-In addition to week-by-week actual player statistics, we can also compute historical actual player statistics as a function of different timeframes, including season-by-season and career statistics.
+We can compute historical actual player statistics as a function of different timeframes, including [week-by-week](#sec-calculateActualWeeklyStats), [season-by-season](#sec-calculateActualSeasonalStats), and the [player's career](#sec-calculateActualCareerStats).
+
+#### Week-by-Week Statistics {#sec-calculateActualWeeklyStats}
+
+We already load players' week-by-week statistics [above](#sec-downloadActualWeeklyStats) using the `nflreadr::load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
+Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `nflfastR::calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:
+
+::: {#nte-downloadWeeklyActualStats .callout-note title="Downloading week-by-week statistics"}
+Note: the following code takes a while to run.
+:::
+
+```{r}
+#| eval: false
+nfl_actualStats_weekly_player_raw <- nflfastR::calculate_stats(
+  seasons = TRUE,
+  summary_level = "week")
+
+nfl_actualStats_weekly_team_raw <- nflfastR::calculate_stats(
+  seasons = TRUE,
+  summary_level = "week",
+  stat_type = "team")
+```
+
+```{r}
+#| include: false
+
+# garbage collection
+rm(nfl_pbp)
+gc()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_weekly_player_raw, nfl_actualStats_weekly_team_raw,
+  file = "./data/nfl_actualStats_weekly_raw.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_weekly_raw.RData")
+```
+
+The `nfl_actualStats_weekly` objects are in `player`-`season`-`week` form.
+That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
+Let's rearrange the data accordingly:
+
+```{r}
+nfl_actualStats_weekly_player <- nfl_actualStats_weekly_player_raw %>% 
+  select(player_id, season, week, everything()) %>% 
+  arrange(player_display_name, player_id, season, week)
+
+nfl_actualStats_weekly_team <- nfl_actualStats_weekly_team_raw %>% 
+  select(season, week, team, everything()) %>% 
+  arrange(season, week, team)
+```
+
+Let's check for duplicate `player`-`season`-`week` instances:
+
+```{r}
+nfl_actualStats_weekly_player %>% 
+  group_by(player_id, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+
+nfl_actualStats_weekly_team %>% 
+  group_by(team, season, week) %>% 
+  filter(n() > 1) %>% 
+  head()
+```
+
+Here are the variable names:
+
+```{r}
+names(nfl_actualStats_weekly_player) %>% as.data.frame()
+
+names(nfl_actualStats_weekly_team) %>% as.data.frame()
+```
+
+```{r}
+#| eval: false
+
+save(
+  nfl_actualStats_weekly_player, nfl_actualStats_weekly_team,
+  file = "./data/nfl_actualStats_weekly.RData"
+)
+```
+
+```{r}
+#| include: false
+
+load(file = "./data/nfl_actualStats_weekly.RData")
+```
 
@@
 
-#### Week-by-Week Statistics {#sec-calculateActualWeeklyStats}
-
-We already load players' week-by-week statistics [above](#sec-downloadActualWeeklyStats) using the `nflreadr::load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
-Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `nflfastR::calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:
-
-::: {#nte-downloadWeeklyActualStats .callout-note title="Downloading week-by-week statistics"}
-Note: the following code takes a while to run.
-:::
-
-```{r}
-#| eval: false
-nfl_actualStats_weekly_player_raw <- nflfastR::calculate_stats(
-  seasons = TRUE,
-  summary_level = "week")
-
-nfl_actualStats_weekly_team_raw <- nflfastR::calculate_stats(
-  seasons = TRUE,
-  summary_level = "week",
-  stat_type = "team")
-```
-
-```{r}
-#| include: false
-
-# garbage collection
-rm(nfl_pbp)
-gc()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_weekly_player_raw, nfl_actualStats_weekly_team_raw,
-  file = "./data/nfl_actualStats_weekly_raw.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_weekly_raw.RData")
-```
-
-The `nfl_actualStats_weekly` objects are in `player`-`season`-`week` form.
-That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
-Let's rearrange the data accordingly:
-
-```{r}
-nfl_actualStats_weekly_player <- nfl_actualStats_weekly_player_raw %>% 
-  select(player_id, season, week, everything()) %>% 
-  arrange(player_display_name, player_id, season, week)
-
-nfl_actualStats_weekly_team <- nfl_actualStats_weekly_team_raw %>% 
-  select(season, week, team, everything()) %>% 
-  arrange(season, week, team)
-```
-
-Let's check for duplicate `player`-`season`-`week` instances:
-
-```{r}
-nfl_actualStats_weekly_player %>% 
-  group_by(player_id, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-
-nfl_actualStats_weekly_team %>% 
-  group_by(team, season, week) %>% 
-  filter(n() > 1) %>% 
-  head()
-```
-
-Here are the variable names:
-
-```{r}
-names(nfl_actualStats_weekly_player) %>% as.data.frame()
-
-names(nfl_actualStats_weekly_team) %>% as.data.frame()
-```
-
-```{r}
-#| eval: false
-
-save(
-  nfl_actualStats_weekly_player, nfl_actualStats_weekly_team,
-  file = "./data/nfl_actualStats_weekly.RData"
-)
-```
-
-```{r}
-#| include: false
-
-load(file = "./data/nfl_actualStats_weekly.RData")
-```
-
 ### Historical Actual Fantasy Points {#sec-calculateActualPoints}
 
+We can then calculate players' historical actual fantasy points for [week-by-week](#sec-calculateActualPointsWeekly), [season-by-season](#sec-calculateActualPointsSeasonal), and the [player's career](#sec-calculateActualPointsCareer).
 Specify scoring settings:

### 2025-09-04 R-scripts/download_weekly_projections.R

@@ -76,7 +76,7 @@ players_projections_weekly_average <- full_join(
 # Add Additional Player Information ----
 players_projections_weekly_average <- players_projections_weekly_average %>% 
   add_ecr() %>% 
-  #add_uncertainty() %>% # currently throws an error
+  add_uncertainty() %>% # comment out if throws an error
   add_player_info()
 
 # Save Projected Points ----

### 2025-09-04 draft.qmd
20250904 - herding behavior

@@
 The same was not the case for Running Backs, Wide Receivers, and Tight Ends.
+
+Engaging in herding behavior—by following a run, i.e., drafting a player at the same position as previous drafters—can be harmful.
+For instance, if the run on Quarterbacks is premature, it can lead the "herd" (i.e., the managers who draft a Quarterback after the previous drafters) to spend important draft capital that could have yielded more impactful players at other positions.
+However, engaging in herding behavior is not always harmful.
+Herding behavior can be rational if a) there is positional scarcity, or if b) not drafting a player at that position would leave with you without a serviceable option at that position.
+Position scarcity refers to the [dropoff](#sec-fantasyValueDropoff) of fantasy points after the top tier of a position—i.e., how steeply player value declines as you move down the rankings, as depicted in @sec-expectedPointsByPositionRank.
+For instance, Quarterbacks and Tight Ends tend to show greater positional scarcity than Running Backs and Wide Receivers.
+That is, there tend to be fewer elite options available at Quarterback and Tight End than at Running Back and Wide Receiver.
+So, in leagues that have two Quarterbacks on the starting roster or a superflex position (i.e., QB/RB/WR/TE), herding due to positional scarcity can be strategically sound.
+In general, though, it may be optimal to either a) be at the beginning of a run (i.e., to be the manager who drafts a player, which leads to a run on that position) to secure a scarce, elite option, or b) to avoid the run altogether (by waiting to draft that position later), rather than joining the run in the middle or at the end.
 

### 2025-09-06 base-rates.qmd
20250906 - overconfidence bias

@@
 For instance, predictions of touchdowns (less common) tend to be less accurate than predictions of yardage (more common).
+
+Teams in the NFL also tend to neglect base rates.
+For instance, NFL teams give too much weight to scouting evidence in deciding which players to draft, and fail to integrate such evidence with the prior probabilities of a player's successful future performance [i.e., the base rate of success; @Massey2013].
+"...consider how suspiciously often we hear a college prospect described as a 'once-in-a-lifetime player'." [@Massey2013, p. 1482].
 
@@
 We examine the base rates of rookie performance compared to performance of non-rookies.
+It can be challenging to predict rookie performance because they have not yet played in the NFL.
+Thus, it is important to consider the prior probabilities (base rates) of success among rookies in the NFL.
 Rookies who play tend to be better players who were drafted in early rounds of the National Football League (NFL) draft.

### 2025-09-06 cognitive-bias.qmd
20250906 - overconfidence bias

@@
 *Overestimation* involves believing that one will perform better than one actually performs.
+It thus reflects a greater mean confidence than overall accuracy [@Kahneman1996].
+In general, people tend to show excessive confidence in their beliefs; for instance, when people express 90% confidence about their belief, they may be correct only about 75% of the time [@Kahneman1996].
+As an example of overconfidence in football, early picks in the NFL draft tend to be considerably overvalued because teams believe their predictions will be more accurate than they actually are [@Massey2013].
+The probability that a player is better than the next player chosen in the draft at the same position is ~52%, or basically a coin flip [@Massey2013].
+As noted by @Massey2013 (p.
+) "...decision makers often know less than they think they know...The problem is not that future performance is difficult to predict, but that decision makers do not appreciate how difficult it is."
+This suggests that a) the presence of high stakes for the prediction does not eliminate overconfidence, and that b) teams should "trade down" to acquire more, later picks if they possess very early picks [@Massey2013].
 Overestimation can be identified with a [calibration plot](#sec-calibrationPlot) of the predicted performance versus actual performance, where the person's predicted performance is systematically higher (in at least some cases) than their actual performance.
@@
 In general, people tend to focus on what they know and to neglect what they do not know leading to overconfidence in one's beliefs [@Kahneman2011].
+Moreover, overconfidence tends to be exacerbated by information—the more information people have, the more they tend to be overconfident [@Massey2013].
+Thus, lots of scouting evidence may lead teams and fantasy managers to be overconfident in the draft.
 

### 2025-09-06 sports-cognitive-psychology.qmd
20250906 - overconfidence bias

@@
 Examples of changes in football due to the use of analytics include more often going for it on fourth down, a greater emphasis on the passing game, drafting Running Backs later in the draft (and, more generally, valuing Running Backs less), and trading down in the draft to obtain more low picks (rather than having fewer high picks) because top picks are frequently overvalued [@Massey2013].
+Although both performance and compensation decline with draft order in the NFL, compensation declines more steeply than performance [@Massey2013].
+The highest value picks—in terms of the greatest expected performance relative to compensation (i.e., performance per dollar)—are near the beginning of the second round of the NFL draft [@Massey2013].
+According to their analysis, @Massey2013 found that the least valuable pick in the first round of the NFL draft (based on expected performance relative to compensation) is the 1st overall pick!
+Thus, they refer to the first pick as the "loser's curse" [@Massey2013].
+In general, the overvaluation of picks tends to be greatest at the top of the draft, even accounting for the possibility of drafting superstars [@Massey2013].
 

### 2025-09-06 cognitive-bias.qmd
20250906 - page numbers

@@
 The probability that a player is better than the next player chosen in the draft at the same position is ~52%, or basically a coin flip [@Massey2013].
-As noted by @Massey2013 (p.
-) "...decision makers often know less than they think they know...The problem is not that future performance is difficult to predict, but that decision makers do not appreciate how difficult it is."
+As noted by @Massey2013 (pp.
+1493–1494), "The task of picking players...is an extremely difficult one...decision makers often know less than they think they know...The problem is not that future performance is difficult to predict, but that decision makers do not appreciate how difficult it is."
 This suggests that a) the presence of high stakes for the prediction does not eliminate overconfidence, and that b) teams should "trade down" to acquire more, later picks if they possess very early picks [@Massey2013].

### 2025-09-06 machine-learning.qmd
20250906 - fix error due to renamed variables

@@
 
-nfl_actualStats_offense_weekly$player_id
+#nfl_actualStats_offense_weekly$player_id
 nfl_expectedFantasyPoints_weekly$player_id
 
-nfl_rankings$id
+nfl_rankings_draft$id
+nfl_rankings_weekly$fantasypros_id
 
@@
 nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
+  rename(gsis_id = player_id) %>% 
   mutate(

### 2025-09-09 getting-started.qmd
20250909 - case sensitive
20250909 - run as administrator and don't save workspace

@@
 This chapter provides an overview of how to install and learn the software `R`, how to troubleshoot code, and how to perform various data management operations.
+Please note that R code is case sensitive—it treats lower case as different from upper case, so you must be very careful to use the correct case when referring to a given object.
 
@@
 1.
-After installing `R` and `RStudio`, open `RStudio` (by right clicking and selecting "Run as Administrator") and run the following code in the console to install the `R` packages used in this book (note: this will take a while):
+After installing `R` and `RStudio`, set the executables for `R` and `RStudio` to always run with administrator permissions (so they have write access to the main installation directory for installing packages).
+    - If on Windows, open File Explorer and find the main executable of `R` (`C:/Program Files/R/[R-VERSION]/bin/R.exe`) and `RStudio` (`C:\Program Files\RStudio\bin\RStudio.exe`).
+    Right-click it to open the contextual menu.
+    Then, click or tap on "Properties".
+    In the Properties window, go to the Compatibility tab.
+    At the bottom of the window, check the box next to the "Run this program as an administrator" option, and then click or tap on OK.
+1. Open `RStudio` (by right clicking and selecting "Run as Administrator").
+1. For [reproducibility purposes](https://devpsylab.github.io/DataAnalysis/R.html#dontSaveWorkspace), prevent `RStudio` from saving your workspaces automatically:
+    - With `RStudio` running, choose `Tools → Global Options` from the menus.
+    - In the Options dialog, change the value for `Save workspace to .RData on exit` to `Never`.
+    - Click `OK`.
+1. In `RStudio`, run the following code in the console to install the `R` packages used in this book (note: this will take a while):
    ```{r}

### 2025-09-09 getting-started.qmd
20250909 - tweaks

@@
 This chapter provides an overview of how to install and learn the software `R`, how to troubleshoot code, and how to perform various data management operations.
-Please note that R code is case sensitive—it treats lower case as different from upper case, so you must be very careful to use the correct case when referring to a given object.
+Please note that R code is case sensitive—it treats lower case as different from upper case, so you want to be careful to use the correct case when referring to a given object.
 
@@
 After installing `R` and `RStudio`, set the executables for `R` and `RStudio` to always run with administrator permissions (so they have write access to the main installation directory for installing packages).
-    - If on Windows, open File Explorer and find the main executable of `R` (`C:/Program Files/R/[R-VERSION]/bin/R.exe`) and `RStudio` (`C:\Program Files\RStudio\bin\RStudio.exe`).
+    - If on Windows, open File Explorer and find the main executable of `R` (`C:/Program Files/R/[R-VERSION]/bin/R.exe`) and `RStudio` (`C:/Program Files/RStudio/bin/RStudio.exe`).
     Right-click it to open the contextual menu.
-    Then, click or tap on "Properties".
+    Then, click on "Properties".
     In the Properties window, go to the Compatibility tab.
-    At the bottom of the window, check the box next to the "Run this program as an administrator" option, and then click or tap on OK.
+    At the bottom of the window, check the box next to the "Run this program as an administrator" option, and then click `OK`.
 1. Open `RStudio` (by right clicking and selecting "Run as Administrator").

### 2025-09-10 R-scripts/download_weekly_projections.R
20250910 - week 2 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 1
+weekNumber <- 2
 
 # Processing ----
 season <- as.character(season)
@@ -15,6 +15,7 @@ library("tidyverse")
 
 # Download Weekly Projections ----
 players_projections_weekly_raw <- ffanalytics::scrape_data(
+  #pos = c("QB", "RB", "WR", "TE", "K", "DST"), # include this line (i.e., uncomment it out) if get an error in downstream steps due to insufficient sources of IDP projections
   season = NULL, # NULL grabs the current season
   week = NULL) # NULL grabs the current week
 

### 2025-09-10 machine-learning.qmd
20250910 - machine learning additions/edits

@@
 
+Machine learning is a class of algorithmic approaches that are used to identify patterns in data.
 Machine learning takes us away from focusing on [causal inference](#sec-causalInference).
@@
 
-Machine learning can be useful for leveraging big data and lots of predictor variable to develop predictive models with greater accuracy.
+Machine learning can be useful for leveraging big data and many predictor variables to develop predictive models with greater accuracy.
 However, many machine learning techniques are black boxes—it is often unclear how or why certain predictions are made, which can make it difficult to interpret the model's decisions and understand the underlying relationships between variables.
@@
 
+Although machine learning tends to be data-driven in its execution, theory should still inform which variables are included in the model.
+
 Most machine learning methods were developed with cross-sectional data in mind.
@@
       - [elastic net regression](#sec-elasticNet)
+      - tree-based methods
+        - decision tree
         - [random forest](#sec-randomForest)
+        - [gradient tree boosting](#sec-treeBoosting)
     - categorical outcome (i.e., classification)
@@
       - [random forest](#sec-randomForest)
-      - boosting
-        - [tree boosting](#sec-treeBoosting)
+      - tree-based methods
 - unsupervised learning
@@
 - reinforcement learning
-    - deep learning
+- deep learning / neural network
 - ensemble
 
-*Ensemble* machine learning methods combine multiple machine learning approaches with the goal that combining multiple approaches might lead to more accurate predictions than any one method might be able to achieve on its own.
-
 ### Supervised Learning {#sec-machineLearningTypesSupervised}
 
-Supervised learning involves learning from data where the correct classification or outcome is known.
+Supervised learning involves learning from data where the correct classification or outcome is known (and the classification is thus part of the data).
 For instance, predicting how many points a player will score is a supervised learning task, because there is a ground truth—the actual number of points scored—that can be used to train and evaluate the model.
-
-Unlike linear and logistic regression, various machine learning techniques can handle [multicollinearity](#sec-multipleRegressionMulticollinearity), including [LASSO regression](#sec-lasso), [ridge regression](#sec-ridgeRegression), and [elastic net regression](#sec-elasticNet).
+If the outcome variable is categorical, the approach involves classification.
+If the outcome vairable is continuous, the approach involves regression.
+
+Unlike linear and logistic regression, various machine learning techniques can handle [multicollinearity](#sec-multipleRegressionMulticollinearity), including [LASSO regression](#sec-lasso), [ridge regression](#sec-ridgeRegression), and [elastic net regression](#sec-elasticNet) via regularization.
+Regularization involves penalizing model complexity to avoid [overfitting](#sec-overfitting) [@Ramasubramanian2016].
 [Least absolute shrinkage and selection option (LASSO) regression](#sec-lasso) performs selection of which predictor variables to keep in the model by shrinking some coefficients to zero, effectively removing them from the model.
@@
 We describe [cluster analysis](#sec-clusterAnalysis) in @sec-clusterAnalysis.
+We describe [factor analysis](#sec-factorAnalysis) in @sec-factorAnalysis.
 We describe [principal component analysis](#sec-pca) in @sec-pca.
@@
 Through trial and error, the agent receives feedback in the form of rewards or penalties and learns a strategy that maximizes the cumulative reward over time.
+
+### Ensemble Learing {#sec-machineLearningTypesEnsemble}
+
+*Ensemble* machine learning methods combine multiple machine learning approaches with the goal that combining multiple approaches might lead to more accurate predictions than any one method might be able to achieve on its own.
 
@@
 It is usually the largest portion of the data.
-We fit various models to the training set based on which parameters we want to evaluate (e.g., how many trees to use in a tree-boosting model).
+We fit various models to the training set based on which parameters we want to evaluate (e.g., how many trees to use in a [gradient tree boosting model](#sec-treeBoosting)).
 
@@
 However, we also provide the code to apply [leave-one-out (LOO) cross-validation](#sec-looCrossValidation).
+[*k*-fold](#sec-kFoldCrossValidation) and [LOO](#sec-looCrossValidation) cross-validation are both forms of *internal* cross-validation.
 
@@
 For example, in a LOO cross-validation with 100 players, the model is trained 100 times, each time leaving out a different player for validation.
+LOO cross-validation is a special case of [*k*-fold cross-validation](#sec-kFoldCrossValidation) where $k = n$.
 LOO cross-validation is especially useful when the dataset is small—too small to form reliable training sets in [*k*-fold cross-validation](#sec-kFoldCrossValidation) (e.g., with $k = 5$ or $k = 10$, which divide the sample into 5 or 10 folds, respectively).
@@
 
+A random forest model combines the results of many decision trees.
+A decision tree splits the data into groups based on the values of the predictors, and then predicts the outcome within each group.
+A given decision tree can be noisy and prone to [overfitting](#sec-overfitting), so random forests build many trees, where each tree is trained on a random subset of the data and considers a random subset of predictors at each split.
+The random forest then averages the predictions of all the decision trees, to improve overall predictive accuracy.
+
 #### Assuming Independent Observations {#sec-randomForestIndependent}
@@
 However, in our case, this assumption does not hold because the data are longitudinal—each player has multiple seasons, and each row corresponds to a unique player–season combination.
-In the approaches below, using a random forest model (this section) and a [tree-boosting model](#sec-treeBoosting) (in @sec-treeBoosting), we address this by explicitly accounting for the nesting of longitudinal observations within players.
+In the approaches below, using a random forest model (this section) and a [gradient tree boosting model](#sec-treeBoosting) (in @sec-treeBoosting), we address this by explicitly accounting for the nesting of longitudinal observations within players.
 
@@
 
-### Combining Tree-Boosting with Mixed Models {#sec-treeBoosting}
-
-To combine tree-boosting with mixed models, we use the `gpboost` package [@R-gpboost].
+### Combining Gradient Tree Boosting with Mixed Models {#sec-treeBoosting}
+
+To combine gradient tree boosting with mixed models, we use the `gpboost` package [@R-gpboost].
+Gradient tree boosting is similar to a [random forest model](#sec-randomForest).
+Gradient tree boosting builds trees one by one, each trying to correct the errors of the previous trees.
+By contrast, in a [random forest model](#sec-randomForest), all trees are built independently and then averaged.
 
@@
 #| label: fig-treeBoostingModel
-#| fig-cap: "Predicted Versus Actual Fantasy Points for Tree-Boosting Model."
-#| fig-alt: "Predicted Versus Actual Fantasy Points for Tree-Boosting Model."
+#| fig-cap: "Predicted Versus Actual Fantasy Points for Gradient Tree Boosting Model."
+#| fig-alt: "Predicted Versus Actual Fantasy Points for Gradient Tree Boosting Model."
 
@@
     "Random Forest",
-    "Tree-Boosting"
+    "Gradient Tree Boosting"
   ), 
@@
 
-The most accurate models, in terms of [discrimination](#sec-discrimination) ($R^2$), had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using one of two approaches—a [tree-boosting model](#sec-treeBoosting) with many predictors or a [regression model with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
-The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and the other [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
-This suggests that the other predictors examined did not have [incremental validity](#sec-incrementalValidity)—they did not show utility above and beyond a player's fantasy points in predicting their future fantasy points.
+The most accurate models, in terms of [discrimination](#sec-discrimination) ($R^2$), had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using several approaches—including a [gradient tree boosting model](#sec-treeBoosting) with many predictors and a [regression model with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
+The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and were just as accurate, if not more accurate, than other [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
+This suggests that the other predictors examined did not have [incremental validity](#sec-incrementalValidity)—they did not show utility above and beyond a player's previous fantasy points in predicting their future fantasy points.
 If anything, considering the additional predictor variables showed *decremental* validity—the inclusion of the additional variables worsened prediction, possibly by weakening the predictive association of the previous season's fantasy points and misattributing its predictive effect to other, related predictors in the model.
@@
 
-The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [tree-boosting model](#sec-treeBoosting), which had a [mean absolute error](#sec-meanAbsoluteError) of `{r} mae_minimum`, and was just slightly more accurate than the [regression model with one predictor](#sec-regressionOnePredictor).
+The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [gradient tree boosting model](#sec-treeBoosting), which had a [mean absolute error](#sec-meanAbsoluteError) of `{r} mae_minimum`, and was just slightly more accurate than the [regression model with one predictor](#sec-regressionOnePredictor).
 However, even in the most accurate model, the predicted values were over `{r} petersenlab::apa(mae_minimum, decimals = 0)` points away from the actual values, suggesting that the model-predicted values were not particularly accurate.
-In sum, the [regression model with one predictor](#sec-regressionOnePredictor) was the most accurate in terms of differentiating between players ($R^2$), whereas the [tree-boosting model](#sec-treeBoosting) was the most accurate in terms of how close the predicted values were to the actual values.
+In sum, the [regression model with one predictor](#sec-regressionOnePredictor) was the most accurate in terms of differentiating between players ($R^2$), whereas the [gradient tree boosting model](#sec-treeBoosting) was the most accurate in terms of how close the predicted values were to the actual values.
 
@@
 The most accurate model, in terms of [discrimination](#sec-discrimination) ($R^2$), was achieved using regression with one predictor: the player's fantasy points in the previous season.
-The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [tree-boosting model](#sec-treeBoosting), which accounted for the longitudinal nature of the data.
+The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [gradient tree boosting model](#sec-treeBoosting), which accounted for the longitudinal nature of the data.
 The considerable unexplained variance suggests that we are missing key variables that are important predictors of future fantasy performance.

### 2025-09-11 data-visualization.qmd
20250822 - clean up code
20250911 - geom_violin update to accommodate ggplot 4.0

@@
 The Okabe-Ito color palette [@Okabe2008] is depicted in Figures [-@fig-okabeItoPaletteGray] and [-@fig-okabeItoPaletteBlack].
+You can see a pie chart of colors using the `graphics::pie()` function:
 
@@
 
-graphics::pie(rep(1, 8), col = palette_OkabeIto)
+pie(
+  rep(1, 8),
+  col = palette_OkabeIto)
 ```
@@
 
-graphics::pie(rep(1, 8), col = palette_OkabeIto_black)
+pie(
+  rep(1, 8),
+  col = palette_OkabeIto_black)
 ```
@@
 ) +
-  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
+  geom_violin(
+    quantiles = c(0.25, 0.5, 0.75),
+    quantile.linetype = "solid"
+  ) +
   labs(

### 2025-09-12 basic-statistics.qmd
20250911 - fix error

@@
   # add line for treatment group
-  geom_line(data=df1, aes(x, y, color = "H0", group = NULL, fill = NULL), linewidth = 1.5, show_guide = FALSE) + 
+  geom_line(data=df1, aes(x, y, color = "H0", group = NULL, fill = NULL), linewidth = 1.5, show.legend = FALSE) + 
   # add line for treatment group. These lines could be combined into one dataframe.
-  geom_line(data=df2, aes(color = "HA", group = NULL, fill = NULL), linewidth = 1.5, show_guide = FALSE) +
+  geom_line(data=df2, aes(color = "HA", group = NULL, fill = NULL), linewidth = 1.5, show.legend = FALSE) +
   # add vlines for z_crit
@@
   annotate("segment", x = 6, y = 0.2, xend = 4.5, yend = 0.15, arrow = grid::arrow(length = unit(0.3, "cm")), linewidth = 1) +
-  annotate("text", label = expression(paste(1-beta, "  (\"power\")")), x = 6.1, y = 0.21, parse = TRUE, size = 8) +
+  annotate("text", label = "1 - beta ~ '(power)'", x = 6.1, y = 0.21, parse = TRUE, size = 8) + 
   # H_0 title

### 2025-09-12 basic-statistics.qmd
20250911 - clean up code

@@
 # plot with ggplot2
-ggplot(poly, aes(x,y, fill=id, group=id)) +
-  geom_polygon(show.legend=F, alpha=I(8/10)) +
+ggplot(
+  poly,
+  aes(
+    x,
+    y,
+    fill = id,
+    group = id)) +
+  geom_polygon(
+    show.legend = FALSE,
+    alpha = I(8/10)) +
   # add line for treatment group
-  geom_line(data=df1, aes(x, y, color = "H0", group = NULL, fill = NULL), linewidth = 1.5, show.legend = FALSE) + 
+  geom_line(
+    data = df1,
+    aes(
+      x,
+      y,
+      color = "H0",
+      group = NULL,
+      fill = NULL),
+    linewidth = 1.5,
+    show.legend = FALSE) + 
   # add line for treatment group. These lines could be combined into one dataframe.
-  geom_line(data=df2, aes(color = "HA", group = NULL, fill = NULL), linewidth = 1.5, show.legend = FALSE) +
+  geom_line(
+    data = df2,
+    aes(
+      color = "HA",
+      group = NULL,
+      fill = NULL),
+    linewidth = 1.5,
+    show.legend = FALSE) +
   # add vlines for z_crit
-  geom_vline(xintercept = z_crit, linewidth=1, linetype="dashed") +
+  geom_vline(
+    xintercept = z_crit,
+    linewidth = 1,
+    linetype = "dashed") +
   # change colors
@@
     values = c("HA" = "#981e0b", "H0" = "black")) +
-  scale_fill_manual("test", values= c("alpha" = "#0d6374", "beta" = "#be805e", "power"="#7cecee")) +
+  scale_fill_manual(
+    "test",
+    values = c("alpha" = "#0d6374", "beta" = "#be805e", "power"="#7cecee")) +
   # beta arrow
-  annotate("segment", x = 0.1, y = 0.045, xend = 1.3, yend = 0.01, arrow = grid::arrow(length = unit(0.3, "cm")), linewidth = 1) +
-  annotate("text", label="beta", x=0, y=0.05, parse=T, size=8) +
+  annotate(
+    "segment",
+    x = 0.1,
+    y = 0.045,
+    xend = 1.3,
+    yend = 0.01,
+    arrow = grid::arrow(length = unit(0.3, "cm")),
+    linewidth = 1) +
+  annotate(
+    "text",
+    label = "beta",
+    x = 0,
+    y = 0.05,
+    parse = TRUE,
+    size = 8) +
   # alpha arrow
-  annotate("segment", x = 4, y = 0.043, xend = 3.4, yend = 0.01, arrow = grid::arrow(length = unit(0.3, "cm")), linewidth = 1) +
-  annotate("text", label="frac(alpha,2)", x=4.2, y=0.05, parse=T, size=8) +
+  annotate(
+    "segment",
+    x = 4,
+    y = 0.043,
+    xend = 3.4,
+    yend = 0.01,
+    arrow = grid::arrow(length = unit(0.3, "cm")),
+    linewidth = 1) +
+  annotate(
+    "text",
+    label = "frac(alpha,2)",
+    x = 4.2,
+    y = 0.05,
+    parse = TRUE,
+    size = 8) +
   # power arrow
-  annotate("segment", x = 6, y = 0.2, xend = 4.5, yend = 0.15, arrow = grid::arrow(length = unit(0.3, "cm")), linewidth = 1) +
-  annotate("text", label = "1 - beta ~ '(power)'", x = 6.1, y = 0.21, parse = TRUE, size = 8) + 
+  annotate(
+    "segment",
+    x = 6,
+    y = 0.2,
+    xend = 4.5,
+    yend = 0.15,
+    arrow = grid::arrow(length = unit(0.3, "cm")),
+    linewidth = 1) +
+  annotate(
+    "text",
+    label = "1 - beta ~ '(power)'",
+    x = 6.1,
+    y = 0.21,
+    parse = TRUE,
+    size = 8) + 
   # H_0 title
-  annotate("text", label = "H[0]", x = m1, y = 0.28, parse = TRUE, size = 8) +
+  annotate(
+    "text",
+    label = "H[0]",
+    x = m1,
+    y = 0.28,
+    parse = TRUE,
+    size = 8) +
   # H_a title
-  annotate("text", label = "H[1]", x = m2, y = 0.28, parse = TRUE, size = 8) +
+  annotate(
+    "text",
+    label = "H[1]",
+    x = m2,
+    y = 0.28,
+    parse = TRUE,
+    size = 8) +
   ggtitle("Statistical Power") +
@@
     panel.background = element_blank(),
-    plot.background = element_rect(fill="white"),
+    plot.background = element_rect(fill = "white"),
     panel.border = element_blank(),
@@
     axis.title.y = element_blank(),
-    plot.title = element_text(size=22))
+    plot.title = element_text(size = 22))
 ```

### 2025-09-12 mixed-models.qmd
20250820 - visualization of mixed models
20250820 - copyediting
20250911 - prevent R chunks in notes from running

@@
 
+A visualization of mixed models, including the distinction between fixed and random intercepts and slopes, is provided by @Freeman2017: <http://mfviz.com/hierarchical-models> (archived at <https://perma.cc/H2GZ-P5RW>).
+
 Mixed modeling is a more flexible modeling framework than [analysis of variance](#sec-anova) (ANOVA).
@@
 ```{r}
+#| eval: false
+
 load(url("https://osf.io/download/q6rjf/"))

### 2025-09-12 cluster-analysis.qmd
20250910 - machine learning additions/edits
20250911 - no longer errors out

@@
 There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, hierarchical clustering (aka connectivity-based clustering), etc.
+In general, cluster analysis intends to maximize intracluster similarities and to minimize intercluster similarities [@Ramasubramanian2016].
 An overview of approaches to cluster analysis in `R` is provided by @Kassambara2017a.
@@
 
+<!--
 The model-based bootstrap clustering of Running Backs' fantasy points is unable to run due to an error:
+-->
 
@@
 
+<!--
 Thus, we cannot use the following code, which would otherwise summarize the model results, specify the number of tiers, and plot model comparisons:
+-->
 
 ```{r}
-#| eval: false
-
 numTiersRB <- as.numeric(summary(tiersRB_boostrap)[,"Length"][1]) # or could specify the number of teams manually

### 2025-09-14 download-football-data.qmd
20250914 - NFL trades

@@
 
-#nfl_trades_raw <- progressr::with_progress(
-#  nflreadr::load_trades(seasons = TRUE))
-
-nfl_trades_raw <- read_csv("https://raw.githubusercontent.com/leesharpe/nfldata/master/data/trades.csv")
+nfl_trades_raw <- progressr::with_progress(
+  nflreadr::load_trades(seasons = TRUE))
+
+#nfl_trades_raw <- read_csv("https://raw.githubusercontent.com/leesharpe/nfldata/master/data/trades.csv")
 ```

### 2025-09-16 data-visualization.qmd
20250916 - update visualization to 2025

@@
   dplyr::filter(
-    season == 2024,
+    season == 2025,
     season_type == "REG") %>%
@@
 
-Defensive EPA per play during the 2024 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA using the `nflplotR` package [@R-nflplotR].
+Defensive EPA per play during the 2025 NFL season is depicted as a function of offensive EPA per play in @fig-plotDefOffEPA using the `nflplotR` package [@R-nflplotR].
 
@@
 #| label: fig-plotDefOffEPA
-#| fig-cap: "2024 NFL Offensive and Defensive EPA Per Play."
-#| fig-alt: "2024 NFL Offensive and Defensive EPA Per Play."
+#| fig-cap: "2025 NFL Offensive and Defensive EPA Per Play."
+#| fig-alt: "2025 NFL Offensive and Defensive EPA Per Play."
 
@@
     y = "Defense EPA/play",
-    title = "2024 NFL Offensive and Defensive EPA per Play"
+    title = "2025 NFL Offensive and Defensive EPA per Play"
   ) +

### 2025-09-17 R-scripts/download_weekly_projections.R
20250917 - week 3 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 2
+weekNumber <- 3
 
 # Processing ----
 season <- as.character(season)
@@ -15,7 +15,7 @@ library("tidyverse")
 
 # Download Weekly Projections ----
 players_projections_weekly_raw <- ffanalytics::scrape_data(
-  #pos = c("QB", "RB", "WR", "TE", "K", "DST"), # include this line (i.e., uncomment it out) if get an error in downstream steps due to insufficient sources of IDP projections
+  pos = c("QB", "RB", "WR", "TE", "K", "DST"), # include this line (i.e., uncomment it out) if get an error in downstream steps due to insufficient sources of IDP projections
   season = NULL, # NULL grabs the current season
   week = NULL) # NULL grabs the current week
 

### 2025-09-18 correlation.qmd
20250903 - fix computational challenges
20250918 - edit

@@
 - [outliers](#sec-correlationOutliers) weaken the apparent linear trend
-- the association differs across subgroups of the sample
+- the association differs across subgroups of the sample (i.e., [moderation](#sec-moderation))
 - the data contain random measurement error
@@
 correlation::correlation(
-  data = player_stats_seasonal %>% select(draft_round, fantasyPoints),
+  data = player_stats_seasonal %>% 
+    filter(season == max(season, na.rm = TRUE)) %>% 
+    select(draft_round, fantasyPoints),
   method = "distance"

### 2025-09-18 basic-statistics.qmd
20250918 - ANOVA formula

@@
 
+The *F* statistic (which is used to determine statistical significance) in ANOVA is computed by first partitioning the sum of squares as in @eq-partitionSumSquares:
+
+$$
+SS_{\text{total}} = SS_{\text{between}} + SS_{\text{within}}
+$$ {#eq-partitionSumSquares}
+
+where $SS_{\text{total}}$ is the total sum of squares, which is equal to @eq-sumSquaresTotal:
+
+$$
+SS_{\text{total}} = \sum_{i = 1}^{k} \sum_{j = 1}^{n_i} \left( X_{ij} - \bar{X}_{..} \right)^2
+$$ {#eq-sumSquaresTotal}
+
+where $X_{ij}$ is the raw data point (i.e., observation $j$ within group $i$), $n_i$ is the sample size of group $i$, $\bar{X}_{i.}$ is the mean of group $i$, and $\bar{X}_{..}$ is the grand mean (the mean across all observations from all groups).
+
+$SS_{\text{between}}$ is the between-groups sum of squares, which is equal to @eq-sumSquaresBetween:
+
+$$
+SS_{\text{between}} = \sum_{i = 1}^{k} n_i \left( \bar{X}_{i.} - \bar{X}_{..} \right)^2
+$$ {#eq-sumSquaresBetween}
+
+and $SS_{\text{within}}$ is the within-groups sum of squares, which is equal to @eq-sumSquaresWithin:
+
+$$
+SS_{\text{within}} = \sum_{i = 1}^{k} \sum_{j = 1}^{n_i} \left( X_{ij} - \bar{X}_{i.} \right)^2
+$$ {#eq-sumSquaresWithin}
+
+Mean squares are then computed from the sum of squares as in @eq-meanSquares:
+
+$$
+\begin{aligned}
+  MS_{\text{between}} &= \frac{SS_{\text{between}}}{k - 1} \\
+  MS_{\text{within}} &= \frac{SS_{\text{within}}}{N - k}
+\end{aligned}
+$$ {#eq-meanSquares}
+
+where $k$ is the number of groups and $N$ is the sample size.
+
+Then, the *F* statistic is computed as the ratio of between-groups mean squares to within-groups mean squares as in @eq-fStatistic:
+
+$$
+F = \frac{MS_{\text{between}}}{MS_{\text{within}}}
+$$ {#eq-fStatistic}
+
+In sum, ANOVA examines the ratio of between-groups sum of squares to within-groups sum of squares (after adjusting for degrees of freedom).
+
 Like two-samples *t*-tests, ANOVA allows examining whether groups differ as a function of an [independent variable](#sec-experiment).

### 2025-09-21 simulation.qmd
20250911 - prevent R chunks in notes from running
20250921 - add package name to function

@@
 ```{r}
+#| eval: false
+
 load(url("https://osf.io/download/ues7n/"))
@@
   fixed_objects = list(proj_dists = proj_dists),
-  seed = genSeeds(Design, iseed = 52242), # for reproducibility
+  seed = SimDesign::genSeeds(Design, iseed = 52242), # for reproducibility
   parallel = TRUE # for faster (parallel) processing
@@
 load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/monteCarloSim_results.RData", fsep = ""))
+```
+
+```{r}
+#| include: false
+
+petersenlab::robust_load("https://osf.io/download/ues7n/") #load(url("https://osf.io/download/ues7n/"))
 ```

### 2025-09-22 basic-statistics.qmd
20250922 - add df to ANOVA formula

@@
 \begin{aligned}
-  MS_{\text{between}} &= \frac{SS_{\text{between}}}{k - 1} \\
-  MS_{\text{within}} &= \frac{SS_{\text{within}}}{N - k}
+  MS_{\text{between}} &= \frac{SS_{\text{between}}}{df_{\text{between}}} &=& \frac{SS_{\text{between}}}{k - 1} \\
+  MS_{\text{within}}  &= \frac{SS_{\text{within}}}{df_{\text{within}}}   &=& \frac{SS_{\text{within}}}{N - k}
 \end{aligned}
@@
 
-where $k$ is the number of groups and $N$ is the sample size.
+where $df_{\text{between}}$ is the between-groups degrees of freedom, $df_{\text{within}}$ is the within-groups degrees of freedom, $k$ is the number of groups, and $N$ is the sample size.
 

### 2025-09-23 basic-statistics.qmd
20250923 - clarifications

@@
 - [sample size](#sec-sampleVsPopulation)
-- [alpha level](#sec-statisticalSignificance)
-
-Knowing any three of the following, you can calculate the fourth: statistical power, [effect size](#sec-practicalSignificance), [sample size](#sec-sampleVsPopulation), and [alpha level](#sec-statisticalSignificance).
+- [alpha level](#sec-statisticalSignificance) (Type I error rate)
+
+Knowing any three of the following, you can calculate the fourth (assuming the assumptions of the statistical test are met): statistical power, [effect size](#sec-practicalSignificance), [sample size](#sec-sampleVsPopulation), and [alpha level](#sec-statisticalSignificance) (Type I error rate).
 Below is `R` code for calculating power for each of various statistical tests (i.e., a *power analysis*).

### 2025-09-23 multiple-regression.qmd
20250911 - prevent R chunks in notes from running
20250923 - update

@@
 
-When the [outcome variable](#sec-correlationalStudy) is continuous, [linear regression](#sec-multipleRegressionLinear) is common.
+When the [outcome variable](#sec-correlationalStudy) is continuous or semicontinuous, [linear regression](#sec-multipleRegressionLinear) is common.
 However, there are other types of regression depending on type and distribution of the [outcome variable](#sec-correlationalStudy).
@@
 ```{r}
+#| eval: false
+
 load(url("https://osf.io/download/fe37j/"))

### 2025-09-24 R-scripts/download_weekly_projections.R
20250924 - week 4 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 3
+weekNumber <- 4
 
 # Processing ----
 season <- as.character(season)
@@ -15,7 +15,7 @@ library("tidyverse")
 
 # Download Weekly Projections ----
 players_projections_weekly_raw <- ffanalytics::scrape_data(
-  pos = c("QB", "RB", "WR", "TE", "K", "DST"), # include this line (i.e., uncomment it out) if get an error in downstream steps due to insufficient sources of IDP projections
+  #pos = c("QB", "RB", "WR", "TE", "K", "DST"), # include this line (i.e., uncomment it out) if get an error in downstream steps due to insufficient sources of IDP projections
   season = NULL, # NULL grabs the current season
   week = NULL) # NULL grabs the current week
 

### 2025-09-24 time-series-analysis.qmd
20250903 - updates to address updated data files
20250924 - syntax to load data object

@@
 
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
 The following code loads the Bayesian model object that was fit in @sec-fantasyPointsByAgeExperienceModelsBayesian.
@@
 ```{r}
-petersenlab::robust_load("https://osf.io/download/q6rjf/") #load(url("https://osf.io/download/q6rjf/")) # Bayesian model object
+#| eval: false
+
+load(url("https://osf.io/download/q6rjf/")) # Bayesian model object
+```
+
+```{r}
+#| include: false
+
+petersenlab::robust_load("https://osf.io/download/q6rjf/") #load(url("https://osf.io/download/q6rjf/"))
 ```

### 2025-09-25 mixed-models.qmd
20250925 - cross-classification

@@
 That is, children are nested within the cross-classification of schools and neighborhoods.
-
-Applied to longitudinal fantasy football data, Most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players, with the player's ID as the cluster variable.
+An example of cross-classified data in football is that players are nested within teams and within agents.
+Agents are not nested within teams, and teams are not nested within agents.
+That is, teams are nested within the cross-classification of teams and agents.
+
+Applied to longitudinal fantasy football data, most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players, with the player's ID as the cluster variable.
 However, you could get more complicated.

### 2025-09-30 cognitive-bias.qmd
20250930 - bye weeks

@@
 If you are making optimal decisions that maximize the expected value of returns (e.g., projected fantasy points), you are following a strong process.
-That said, one thing you can count on is that players get injured.
-So, it is to your benefit to draft and compose your team to have some bench depth so you can handle injuries when they inevitably occur.
+That said, one thing you can count on is that players get injured and have bye weeks.
+So, it is to your benefit to draft and compose your team to have some bench depth so you can handle injuries and bye weeks when they inevitably occur.
 

### 2025-09-30 draft.qmd
20250930 - bye weeks

@@
 In any case, having some small amount of cap left over toward the end of the draft can help you draft good value players to fill out your bench spots for cheap (e.g., $2).
-Having some depth can help you offset the risk of injuries and the possibility that some of your starters may underperform their expectation, which is quite likely given how [challenging it is to predict fantasy performance](#sec-accuracyOfCrowd).
+Having some depth can help you offset the impacts of bye weeks, the risk of injuries, and the possibility that some of your starters may underperform their expectation, which is quite likely given how [challenging it is to predict fantasy performance](#sec-accuracyOfCrowd).
 

### 2025-09-30 causal-inference.qmd
20250930 - butterfly bias
20250930 - variables to control/not to control for

@@
 In this case, controlling for either a) coaching quality and mental health or b) diet and mental health—but not both sets—will yield an unbiased estimate of the association between off-field behavior and fantasy points.
+If you control for the both sets (`M`, `A`, and `B`), you can induce an artificial association between `X` and `Y`, which is called butterfly (or bow-tie) bias.
 
@@
 
+## Summary of Variables to Control For (and Not to Control For) {#sec-causalInferenceControlVariablesSummary}
+
+In general, you should control for all [confounds](#sec-causalDiagramConfounding).
+In addition, it is okay to control for [ancestors](#sec-ancestorsDescendants) of the [outcome variable](#sec-correlationalStudy) that are not [confounds](#sec-causalDiagramConfounding) or mediators.
+That is, it is okay to control for variables that influence `Y` that do not influence `X` and that are not influenced by `X`.
+When including these variable as control variables in a model, they are called precision variables.
+You do not need to include precision variables in the model because the estimate of the association is already unbiased if you have controlled for all [confounds](#sec-causalDiagramConfounding).
+However, including precision variables in the model reduces residual variance in the [outcome variable](#sec-correlationalStudy) and can yield more precise estimates (i.e., smaller standard errors) of the association between the [predictor variable](#sec-correlationalStudy) and [outcome variable](#sec-correlationalStudy).
+
+In addition, there are some variables that are important not to control for.
+It is important not to control for mediators of two variables for which you want to determine the estimate of the causal effect—unless you are interested in the *direct* causal effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy) above and beyond the mediator.
+In addition, it is is important not to control for a) ancestors of the [predictor variable](#sec-correlationalStudy) that are not [confounds](#sec-causalDiagramConfounding), b) [descendants](#sec-ancestorsDescendants) of the [outcome variable](#sec-correlationalStudy), and c) [colliders](#sec-colliderBias) (unless the [collider](#sec-colliderBias) is also a [confound](#sec-causalDiagramConfounding)).
+
 ## Conclusion {#sec-causalInferenceConclusion}
@@
 Causal diagrams can be a useful tool for identifying the proper variables to control for (and those not to control for).
-In general, the only variables you should control for are [confound(s)](#sec-causalDiagramConfounding).
+In general, the only variables you should control for are [confounds](#sec-causalDiagramConfounding)—however, it is okay to control for [ancestors](#sec-ancestorsDescendants) of the [outcome variable](#sec-correlationalStudy) that are not [confounds](#sec-causalDiagramConfounding) or mediators.
 When [confounding](#sec-causalDiagramConfounding) exists, it is important to control for the [confound(s)](#sec-causalDiagramConfounding).
 It is important not to control for mediators when interested in the total effect of the [predictor variable](#sec-correlationalStudy) on the [outcome variable](#sec-correlationalStudy).
-In addition, it is important not to control for [descendants](#sec-ancestorsDescendants) of the [outcome variable](#sec-correlationalStudy).
+In addition, it is important not to control for [ancestors](#sec-ancestorsDescendants) of the [predictor variable](#sec-correlationalStudy) or [descendants](#sec-ancestorsDescendants) of the [outcome variable](#sec-correlationalStudy).
 When there is a [collision](#sec-colliderBias), it is important not to control for the [collider](#sec-colliderBias) (unless the [collider](#sec-colliderBias) is also a [confound](#sec-causalDiagramConfounding)).

### 2025-10-01 R-scripts/download_weekly_projections.R
20251001 - week 5 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 4
+weekNumber <- 5
 
 # Processing ----
 season <- as.character(season)

### 2025-10-02 cognitive-bias.qmd
20251001 - copyediting

@@
 Has the player's long-term situation improved, such as moving up the depth chart, or receiving more carries/targets that is not tied to a specific opponent or game script?
-Or, alternatively, do the improvements appear to be driven by transient, game-specific factors, such as a the health of a teammate, the opponents they played, or the game script that ensued?
+Or, alternatively, do the improvements appear to be driven by transient, game-specific factors, such as the health of a teammate, the opponents they played, or the game script that ensued?
 If long-term outlook of the player has appreciably changed due to changes in the [fundamentals of a player's value](#sec-playerEvaluation), such as their [health](#sec-evalHealth), [skill](#sec-skill), or [situation](#sec-evalSituation), it is less likely that such performance improvements will [regress](#sec-fallaciesRegression) over the long run.
@@
 Pay attention to [fundamentals of a player's value](#sec-playerEvaluation), such as their [health](#sec-evalHealth), [skill](#sec-skill), or [situation](#sec-evalSituation) when considering whether a player's performance may [regress to the mean](#sec-fallaciesRegression).
-If the strong performances appear to be driven by transient, game-specific factors, such as a the health of a teammate, the opponents they played, or the game script that ensued, future performances may be more likely to [regress to the mean](#sec-fallaciesRegression).
+If the strong performances appear to be driven by transient, game-specific factors, such as the health of a teammate, the opponents they played, or the game script that ensued, future performances may be more likely to [regress to the mean](#sec-fallaciesRegression).
 In general, people tend to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their predictions.

### 2025-10-08 R-scripts/download_weekly_projections.R
20251008 - week 6

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 5
+weekNumber <- 6
 
 # Processing ----
 season <- as.character(season)

### 2025-10-09 base-rates.qmd
20251009 - update

@@
 It is also important to question the validity of the evidence [@Kahneman2011].
-As noted by @Kahneman2011, if you have doubts about the quality of the evidence for a particular prediction question, keep your predictions close to the [base rate](#sec-baseRate), and modify them only modifying them based on the new information.
+As noted by @Kahneman2011, if you have doubts about the quality of the evidence for a particular prediction question, keep your predictions close to the [base rate](#sec-baseRate), and modify them only slightly (if at all) based on the new information.
 

### 2025-10-11 machine-learning.qmd
20251011 - gradient tree boosting model

@@
 
-To combine gradient tree boosting with mixed models, we use the `gpboost` package [@R-gpboost].
+To combine gradient tree boosting with [mixed models](#sec-mixedModels), we use the `gpboost` package [@R-gpboost].
 Gradient tree boosting is similar to a [random forest model](#sec-randomForest).
@@
 By contrast, in a [random forest model](#sec-randomForest), all trees are built independently and then averaged.
+This gradient tree boosting approach includes a [mixed model](#sec-mixedModels) to account for the longitudinal data nested within players.
 
@@
 We examined multiple machine learning models that included many predictor variables to identify the model that generates the most accurate prediction of Quarterbacks' future fantasy points.
-We also compared them to simpler models, including a [regression model with one predictor](#sec-regressionOnePredictor) and a [regression model with multiple predictors](#sec-regressionMultiplePredictors) to deteremine the [incremental validity](#sec-incrementalValidity) of the machine learning models above and beyond the simpler models.
+We also compared them to simpler models, including a [regression model with one predictor](#sec-regressionOnePredictor) and a [regression model with multiple predictors](#sec-regressionMultiplePredictors) to determine the [incremental validity](#sec-incrementalValidity) of the machine learning models above and beyond the simpler models.
 A summary of the accuracy of the machine learning models is in @tbl-machineLearningModelSummary.
@@
 
-The most accurate models, in terms of [discrimination](#sec-discrimination) ($R^2$), had an $R^2$ of `{r} rsquare_maximum`, and it was achieved using several approaches—including a [gradient tree boosting model](#sec-treeBoosting) with many predictors and a [regression model with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
-The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and were just as accurate, if not more accurate, than other [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
+The most accurate models, in terms of [discrimination](#sec-discrimination) ($R^2$), had an $R^2$ near `{r} rsquare_maximum`, and it was achieved using several approaches—including a [gradient tree boosting model](#sec-treeBoosting) with many predictors and a [regression model with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season.
+The [regression model with one predictor](#sec-regressionOnePredictor) was more accurate (in terms of $R^2$) than a [regression model with multiple predictors](#sec-regressionMultiplePredictors) and was just as accurate, if not more accurate, than other [machine learning models](#sec-fittingModels-machineLearning) that leveraged multiple predictors.
 This suggests that the other predictors examined did not have [incremental validity](#sec-incrementalValidity)—they did not show utility above and beyond a player's previous fantasy points in predicting their future fantasy points.
-If anything, considering the additional predictor variables showed *decremental* validity—the inclusion of the additional variables worsened prediction, possibly by weakening the predictive association of the previous season's fantasy points and misattributing its predictive effect to other, related predictors in the model.
+If anything, considering the additional predictor variables showed *decremental* validity—the inclusion of the additional variables worsened prediction in several models, possibly by weakening the predictive association of the previous season's fantasy points and misattributing its predictive effect to other, related predictors in the model.
 This is consistent with the common finding that simple models (and parsimony) are often preferable.
@@
 The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [gradient tree boosting model](#sec-treeBoosting), which had a [mean absolute error](#sec-meanAbsoluteError) of `{r} mae_minimum`, and was just slightly more accurate than the [regression model with one predictor](#sec-regressionOnePredictor).
+The [gradient tree boosting model](#sec-treeBoosting) being the most accurate suggests that it is important to account for the longitudinal data nested within players.
 However, even in the most accurate model, the predicted values were over `{r} petersenlab::apa(mae_minimum, decimals = 0)` points away from the actual values, suggesting that the model-predicted values were not particularly accurate.
-In sum, the [regression model with one predictor](#sec-regressionOnePredictor) was the most accurate in terms of differentiating between players ($R^2$), whereas the [gradient tree boosting model](#sec-treeBoosting) was the most accurate in terms of how close the predicted values were to the actual values.
+In sum, the [regression model with one predictor](#sec-regressionOnePredictor) and [gradient tree boosting model](#sec-treeBoosting) were the most accurate in terms of differentiating between players ($R^2$), whereas the [gradient tree boosting model](#sec-treeBoosting) was the most accurate in terms of how close the predicted values were to the actual values.
 
@@
 All models explained less than half of the variance in players' fantasy points.
-The most accurate model, in terms of [discrimination](#sec-discrimination) ($R^2$), was achieved using regression with one predictor: the player's fantasy points in the previous season.
+The most accurate model, in terms of [discrimination](#sec-discrimination) ($R^2$), was achieved using a [regression model with one predictor](#sec-regressionOnePredictor): the player's fantasy points in the previous season—and a [gradient tree boosting model](#sec-treeBoosting).
 The most accurate model in terms of [calibration](#sec-calibration) ([mean absolute error](#sec-meanAbsoluteError)) was the [gradient tree boosting model](#sec-treeBoosting), which accounted for the longitudinal nature of the data.

### 2025-10-26 modern-portfolio-theory.qmd
20250903 - updates to address updated data files
20251026 - edits/additions

@@
 If they are on bye or have a down week, your team is more likely to suffer.
+Another way to diversify is to draft players with different bye weeks so you do not have multiple starting players on bye in a given week.
 Also, there are risks in picking multiple offensive players from the same team.
@@
 
-We can also examine the efficient frontier of fantasy performance based on the mean and variability of players' projected week-to-week fantasy points [@Hitchings2012; archived at <https://perma.cc/JQ6G-KSRT>].
+We can also examine the efficient frontier of fantasy performance based on the mean and variability of players' historical (actual) week-to-week fantasy points [@Hitchings2012; archived at <https://perma.cc/JQ6G-KSRT>].
 

### 2025-10-27 cluster-analysis.qmd
20251026 - cluster analysis

@@
 In general, cluster analysis intends to maximize intracluster similarities and to minimize intercluster similarities [@Ramasubramanian2016].
+Cluster analysis is a form of an [unsupervised approach to machine learning](#sec-machineLearningTypesUnsupervised) in which the groups are not known a priori.
+There is a risk of reifying clusters identified in cluster analysis when, in some cases, the clusters may reflect artifacts of the analytic method rather than true distinct groups [@Everitt2011].
+
 An overview of approaches to cluster analysis in `R` is provided by @Kassambara2017a.

### 2025-10-28 cluster-analysis.qmd
20251027 - additional detail

@@
 There is a risk of reifying clusters identified in cluster analysis when, in some cases, the clusters may reflect artifacts of the analytic method rather than true distinct groups [@Everitt2011].
+In general, the classification should be judged on the extent to which it is *useful* rather than on the extent to which it is true or false [@Everitt2011].
 
@@
 In this chapter, we focus on examples using model-based clustering with the `R` package `mclust` [@R-mclust; @Scrucca2023_packages], which uses Gaussian finite mixture modeling.
+Model-based clustering assumes the data are generated by an underlying statistical model, and the goal is to recover the structure of that model.
+Model-based clustering represents the data as coming from a mixture of probability distributions.
 The various types of `mclust` models are provided here:

### 2025-10-28 factor-analysis.qmd
20250819 - updates
20250903 - updates to address updated data files
20251028 - additional info

@@
 The goal of factor analysis is to identify simple, parsimonious factors that underlie the "junk" (i.e., scores filled with measurement error) that we observe.\index{parsimony}\index{factor analysis!confirmatory}
+That is, similar to [principal component analysis](#sec-pca), factor analysis can help us perform data reduction—i.e., it can help us reduce down a larger set of variables into a smaller set of factors.
 
@@
 An additional consideration is the scaling of the variables: whether to use raw variables, standardized variables, or dividing some variables by a constant to make the variables' variances more similar.
-Before performing a [principal component analysis](#sec-pca) (PCA), it is generally important to ensure that the variables included in the PCA are on the same scale.
-PCA seeks to identify components that explain variance in the data, so if the variables are not on the same scale, some variables may contribute considerably more variance than others.
+Before performing a [principal component analysis](#sec-pca) (PCA), it is generally important to ensure that the variables included in the [PCA](#sec-pca) are on the same scale.
+[PCA](#sec-pca) seeks to identify components that explain variance in the data, so if the variables are not on the same scale, some variables may contribute considerably more variance than others.
 A common way of ensuring that variables are on the same scale is to standardize them using, for example, [*z* scores](#sec-zScores) that have a mean of zero and standard deviation of one.
@@
 
+To make the plot interactive for editing, you can use the `lavaangui::plot_lavaan()` function of the `lavaangui` package [@R-lavaangui; @Karch2025_packages]:
+
+```{r}
+#| eval: false
+
+lavaangui::plot_lavaan(efa_fit$nf3)
+```
+
 Here is the syntax for estimating a three-factor EFA using exploratory structural equation modeling (ESEM).

### 2025-10-29 R-scripts/download_weekly_projections.R
20251029 - week 9

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 8
+weekNumber <- 9
 
 # Processing ----
 season <- as.character(season)
@@ -15,7 +15,7 @@ library("tidyverse")
 
 # Download Weekly Projections ----
 players_projections_weekly_raw <- ffanalytics::scrape_data(
-  pos = c("QB", "RB", "WR", "TE", "K", "DST"), # include this line (i.e., uncomment it out) if get an error in downstream steps due to insufficient sources of IDP projections
+  #pos = c("QB", "RB", "WR", "TE", "K", "DST"), # include this line (i.e., uncomment it out) if get an error in downstream steps due to insufficient sources of IDP projections
   season = NULL, # NULL grabs the current season
   week = NULL) # NULL grabs the current week
 

### 2025-11-01 factor-analysis.qmd
20251101 - additions and clarifications

@@
 
-## Factor Analysis and Structural Equation Modeling
+Before pursuing factor analysis, it is helpful to evaluate the extent to which the variables are intercorrelated.
+If the variables are not correlated (or are correlated only weakly), there is no reason to believe that a common factor influences them, and thus factor analysis would not be useful.
+So, before conducting a factor analysis, it is important to examine the correlation matrix of the variables to determine whether the variables intercorrelate strongly enough to justify a factor analysis.
+
+## Factor Analysis and Structural Equation Modeling {#sec-factorAnalysisSEM}
 
@@
 When the two factors are correlated in reality, if we make them uncorrelated, this would result in an inaccurate model.\index{factor analysis}\index{rotation}\index{orthogonal rotation}\index{oblique rotation}
-Oblique rotation allows for factors to be correlated and yields axes with less an angle of less than 90 degrees.\index{factor analysis}\index{rotation}\index{oblique rotation}
+Oblique rotation allows for factors to be correlated and yields axes with an angle of less than 90 degrees.\index{factor analysis}\index{rotation}\index{oblique rotation}
 However, if the factors have low correlation (e.g., .2 or less), you can likely continue with orthogonal rotation.\index{factor analysis}\index{rotation}\index{oblique rotation}
@@
 It does not make sense to replace 18 variables with 18 latent factors because that would not result in any simplification.
+The fewer the number of factors retained, the greater the parsimony, but the greater the amount of information that is discarded (i.e., the less the variance accounted for in the variables).
+The more factors we retain, the less the amount of information that is discarded (i.e., more variance is accounted for in the variables), but the less the parsimony.
 But how do you decide on the number of factors?\index{factor analysis}
@@
 
-- Kaiser-Guttman criterion: factors with eigenvalues greater than zero\index{factor analysis}\index{eigenvalue}
+- Kaiser-Guttman criterion [@Kaiser1960]: factors with eigenvalues greater than zero\index{factor analysis}\index{eigenvalue}
     - or, for [principal component analysis](#sec-pca), components with eigenvalues greater than 1\index{principal component analysis}\index{eigenvalue}
-- Cattell's scree test: the "elbow" in a scree plot minus one; sometimes operationalized with optimal coordinates (OC) or the acceleration factor (AF)\index{factor analysis}\index{scree plot}
+- Cattell's scree test [@Cattell1966]: the "elbow" (inflexion point) in a scree plot minus one; sometimes operationalized with optimal coordinates (OC) or the acceleration factor (AF)\index{factor analysis}\index{scree plot}
 - Parallel analysis: factors that explain more variance than randomly simulated data\index{factor analysis}\index{parallel analysis}
@@
 One criterion for how many factors to keep is the Kaiser-Guttman criterion.
-According to the Kaiser-Guttman criterion, you should keep any factors whose eigenvalue is greater than 1.\index{factor analysis}
+According to the Kaiser-Guttman criterion [@Kaiser1960], you should keep any principal components (from [PCA](#sec-pca)) whose eigenvalue is greater than 1 and factors (from factor analysis) whose eigenvalue is greater than 1.\index{factor analysis}\index{principal component analysis}
 That is, for the sake of simplicity, parsimony, and data reduction, you should take any factors that explain more than a single variable would explain.\index{parsimony}\index{data!reduction}\index{factor analysis}
-According to the Kaiser-Guttman criterion, we would keep three factors from @fig-screePlot that have eigenvalues greater than 1.\index{factor analysis}\index{eigenvalue}
+According to the Kaiser-Guttman criterion, we would keep three components from @fig-screePlot that have eigenvalues greater than 1.\index{factor analysis}\index{eigenvalue}
 The default in SPSS is to retain factors with eigenvalues greater than 1.\index{factor analysis}\index{eigenvalue}
@@
 
-Another criterion is Cattell's scree test, which involves selecting the number of factors from looking at the scree plot.\index{factor analysis}\index{scree plot}
+Another criterion is Cattell's scree test [@Cattell1966], which involves selecting the number of factors from looking at the scree plot.\index{factor analysis}\index{scree plot}
 "Scree" refers to the rubble of stones at the bottom of a mountain.\index{factor analysis}\index{scree plot}
@@
 
+## Benefits of Factor Analysis {#sec-factorAnalysisBenefits}
+
+Factor analysis has several benefits.
+First, factor analysis can be used for data reduction, similar to [PCA](#sec-pca).
+This is useful for reducing a large set of variables into a smaller set of factors that can be more easily analyzed.
+In addition, this data reduction can help to address [multicollinearity](#sec-multipleRegressionMulticollinearity) issues in regression analyses.
+In addition to data reduction, factor analysis can help identify the structure of measures and constructs.
+Thus, factor analysis is particularly useful in the development, validation, and improvement of measures.
+
 ## Conclusion {#sec-factorAnalysisConclusion}

### 2025-11-01 factor-analysis.qmd
20251101 - additions

@@
 Rotation leverages the idea that there are infinite solutions to the factor analysis model that fit equally well.\index{factor analysis}\index{rotation}
-*Rotation* involves changing the orientation of the factors by changing the axes so that variables end up with very high (close to one or negative one) or very low (close to zero) loadings, so that it is clear which factors include which variables.\index{factor analysis}\index{rotation}
-That is, rotation rescales the factors and tries to identify the ideal solution (factor) for each variable.\index{factor analysis}\index{rotation}
-It searches for simple structure and keeps searching until it finds a minimum.\index{factor analysis}\index{rotation}\index{simple structure}
+*Rotation* involves changing the orientation of the factors by changing the axes so that variables end up with very high (close to one or negative one) or very low (close to zero) loadings, so that it is clear which factors include which variables (and which factors each variable most strongly loads onto).\index{factor analysis!decisions}\index{factor analysis}\index{principal component analysis}\index{rotation}
+That is, rotation rescales the factors and tries to identify the ideal solution (factor) for each variable.\index{factor analysis!decisions}\index{factor analysis}\index{principal component analysis}\index{rotation}
+Rotation occurs by changing the variables' loadings while keeping the structure of correlations among the variables intact [@Field2012].
+Moreover, rotation does not change the variance explained by factors.
+Rotation searches for simple structure and keeps searching until it finds a minimum (i.e., the closest as possible to simple structure).\index{factor analysis!decisions}\index{factor analysis}\index{principal component analysis}\index{rotation}\index{simple structure}
 After rotation, if the rotation was successful for imposing simple structure, each factor will have loadings close to one (or negative one) for some variables and close to zero for other variables.\index{factor analysis}\index{rotation}\index{simple structure}

### 2025-11-03 factor-analysis.qmd
20251102 - typos

@@
 A variable's intercept is the expected value of the variable when the factor(s) (onto which it loads) is equal to zero.\index{factor analysis}\index{factor analysis!intercept}
-A [*covariance*](#sec-correlationExamplesCovariance) is the unstandardized index of the strength of association between between two variables (or factors), and it is depicted with a double-headed arrow.\index{factor analysis}\index{factor analysis!covariance}
+A [*covariance*](#sec-correlationExamplesCovariance) is the unstandardized index of the strength of association between two variables (or factors), and it is depicted with a double-headed arrow.\index{factor analysis}\index{factor analysis!covariance}
 
@@
     - or, for [principal component analysis](#sec-pca), components with eigenvalues greater than 1\index{principal component analysis}\index{eigenvalue}
-- Cattell's scree test [@Cattell1966]: the "elbow" (inflexion point) in a scree plot minus one; sometimes operationalized with optimal coordinates (OC) or the acceleration factor (AF)\index{factor analysis}\index{scree plot}
+- Cattell's scree test [@Cattell1966]: the "elbow" (inflection point) in a scree plot minus one; sometimes operationalized with optimal coordinates (OC) or the acceleration factor (AF)\index{factor analysis}\index{scree plot}
 - Parallel analysis: factors that explain more variance than randomly simulated data\index{factor analysis}\index{parallel analysis}

### 2025-11-03 mixed-models.qmd
20251102 - typos

@@
 
-Simpson's paradox occurs when the association between between the [predictor variable](#sec-correlationalStudy) and [outcome variable](#sec-correlationalStudy) for the subgroups differ from the association when the subgroups are combined.
+Simpson's paradox occurs when the association between the [predictor variable](#sec-correlationalStudy) and [outcome variable](#sec-correlationalStudy) for the subgroups differ from the association when the subgroups are combined.
 Examples of Simpson's paradox are depicted in Figures [-@fig-simpsonParadox] and [-@fig-simpsonsParadox].

### 2025-11-04 modern-portfolio-theory.qmd
20251104 - fixes

@@
 
-In @fig-teamPortfolio2, we depict the combined mean of projected fantasy points for a Quarterback/Running Back/Wide Receiver combination as a function of the standard deviation of their projected fantasy points across sources.
+In @fig-teamPortfolio2, we depict the combined mean of historical fantasy points for a Quarterback/Running Back/Wide Receiver combination as a function of the historical standard deviation of their fantasy points across games.
 

### 2025-11-05 pca.qmd
20251101 - additions and clarifications
20251104 - typos and clarifications

@@
 
-As in [factor analysis](#sec-factorAnalysis).
-The first decision when conducting a factor analysis is which variables to include and the scaling of those variables.
+As in [factor analysis](#sec-factorAnalysis), the first decision when conducting a factor analysis is which variables to include and the scaling of those variables.
 Guidance on which variables to include is in @sec-variablesToInclude-factorAnalysis.
@@
 [Factor analysis](#sec-factorAnalysis) better handles error than PCA—[factor analysis](#sec-factorAnalysis) assumes that what is in the variable is the combination of common construct variance and error.\index{principal component analysis}\index{factor analysis}
-By contrast, PCA assumes that the measures have no measurement error.\index{principal component analysis}\index{measurement error}
+By contrast, PCA uses the total variance (not just the common variance) and assumes that the measures have no measurement error.\index{principal component analysis}\index{measurement error}
 [Factor analysis](#sec-factorAnalysis) is useful for the identification of latent constructs—i.e., underlying dimensions or factors that explain (cause) observed scores.\index{factor analysis}
@@
 Based on the rule to keep factors whose eigenvalues are greater than one and based on the parallel test, we would keep nine components.
-However, based on the Cattell scree test (the "elbow" of the screen plot minus one), we would keep three components.
+However, based on the Cattell scree test (the "elbow" of the screen plot minus one) [@Cattell1966], we would keep three components.
 If using the optimal coordinates, we would keep four components; if using the acceleration factor, we would keep one component.

### 2025-11-05 R-scripts/download_weekly_projections.R
20251105 - week 10 projections

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 9
+weekNumber <- 10
 
 # Processing ----
 season <- as.character(season)

### 2025-11-05 simulation.qmd
20251105 - clarification of Monte Carlo simulation using SimDesign

@@
 
+In general, a Monte Carlo simulation study involves several steps:
+
+1. generate the data for the simulation from the relevant probabilty distribution(s)
+1. analyze the simulated data
+1. summarize the results across the simulation replications
+
+Below we conduct a Monte Carlo simulation to determine the range of each player's most likely points scored based on their projected points.
+To do this, we use the `SimDesign` [@R-SimDesign; @Chalmers2020_packages] package.
+
 #### `SimDesign` Package {#sec-simulationProjectedPointsMonteCarloSimDesign}
+
+The `SimDesign` [@R-SimDesign; @Chalmers2020_packages] package uses a five-step process to perform Monte Carlo simulations:
+
+1. Step 1—Design Grid: specify the fixed conditions to investigate in the simulation (e.g., sample size, generating distribution)
+1. Step 2—Generate: specify how to generate the data for the simulation from the relevant probabilty distribution(s)
+1. Step 3—Analyze: specify how to analyze the simulation data
+1. Step 4—Summarize: specify how to summarize the results across the simulation replications
+1. Step 5—Run the Simulation: run the simulation study (i.e., generate the data, analyze the simulated data, and summarize the results across the simulation replications)
 
@@
 
-For each player, we identify the optimal distribution as either a normal distribution , or as a skew-normal distribution.
+For each player, we identify the optimal distribution of their projected points across sources as either a normal distribution, or as a skew-normal distribution.
 The normal distribution was fit using the `fitdistrplus::fitdist()` function of the `fitdistrplus` package [@R-fitdistrplus; @DelignetteMuller2015_packages].
@@
 
-Now we build the `SimDesign` design grid based on the number of projections that each player had.
+First, we define what combinations of conditions to study, which is called our design grid.
+In our simulation study, each player retains the same number of projections as they had in the original data, ensuring the simulation reflects real-world variation in data availability.
+This number of projections defines our design grid.
 
@@
 #### `SimDesign` Step 2: Generate {#sec-simulationProjectedPointsMonteCarloGenerate}
+
+Second, we generate the data for the simulation from the relevant probabilty distribution function(s).
+In this case, we use a normal distribution or skew-normal distribution as our probability distribution function for a given player, depending on which distribution fits better to that player's distribution of projected points across sources (stored in `dist_info$type`).
 
@@
 
+Third, we analyze the simulated data.
+
 ```{r}
@@
 
+Fourth, we summarize the results across the simulation replications.
+
 ```{r}
@@
 
-Now, we can run the model using the `SimDesign::runSimulation()` function.
+Fifth, we run the simulation study.
+We run the simulation study using the `SimDesign::runSimulation()` function.
 

### 2025-11-06 simulation.qmd
20251105 - more detail

@@
 1. Step 1—Design Grid: specify the fixed conditions to investigate in the simulation (e.g., sample size, generating distribution)
-1. Step 2—Generate: specify how to generate the data for the simulation from the relevant probabilty distribution(s)
+1. Step 2—Generate: specify how to generate the data for the simulation from the relevant probability distribution(s)
 1. Step 3—Analyze: specify how to analyze the simulation data
@@
 ```
+
+When designing a Monte Carlo simulation, @Sigal2016 (p. 151) recommend the following:
+
+- beginning with a small number of replications and conditions (i.e., supplying a design object with fewer rows) to ensure that the code runs correctly, and increasing replications/conditions only when the majority of the bugs have been worked out
+- documenting strange aspects of your simulation or code that you are likely to forget
+- saving the state of your simulation in case of power failure or other mechanical failures (e.g., using the `save = TRUE` and/or `save_results = TRUE`, and/or `debug` arguments in `SimDesign::runSimulation()`), and, as necessary, performing debugging within a particular function (`debug` argument in `SimDesign::runSimulation()`)
 
@@
 
-First, we define what combinations of conditions to study, which is called our design grid.
+First, we define what combinations of conditions to study (with one row per each unique combination), which is called our design grid.
+For the design grid, the columns are the design factors (e.g., sample size, generating distribution), and each row indicates a unique combination of the factors (e.g., *N* = 200, normal distribution; *N* = 500, normal distribution; *N* = 200, Poisson distribution; *N* = 500, Poisson distribution).
+When we execute the simulation, each row of the design grid object will be run sequentially.
+
 In our simulation study, each player retains the same number of projections as they had in the original data, ensuring the simulation reflects real-world variation in data availability.
 This number of projections defines our design grid.
+The design factors (columns) are the player (`id`) and the number of projections available for that player (`n_sources`).
+The rows, then are the combinations of player and number of projections available for that player.
 
@@
     id,
-    n_sources = n_proj,
+    n_sources = n_proj, # number of projection sources available for player
     .keep = "none"
   )
-```
-
-```{r}
+
+Design
+```
+
+```{r}
+#| include: false
 #| eval: false
@@
 
-Second, we generate the data for the simulation from the relevant probabilty distribution function(s).
-In this case, we use a normal distribution or skew-normal distribution as our probability distribution function for a given player, depending on which distribution fits better to that player's distribution of projected points across sources (stored in `dist_info$type`).
+Second, we generate the data for the simulation from the relevant probability distribution function(s).
+In this case, we use a normal distribution or skew-normal distribution as our probability distribution function for a given player, depending on which distribution fits better to that player's distribution of projected points across sources (stored in `proj_dists[["INSERT_ID"]]$type`).
+The `Generate()` function has a `condition` argument, which takes a single row from the design grid object (`Design`), and which controls the data generation for subsequent analysis.
+We use the (optional) `fixed_objects` argument so we can pass an object that should be held constant across all simulation replications—i.e., a "fixed" object.
+In this case, we pass—via the `proj_dists` object—the best-fitting distribution for each player (normal or skew-normal) to be used for all simulation replications.
 
@@
 Third, we analyze the simulated data.
+For each iteration of the simulation, the `Analyse()` function will return a vector with a requested variables: player (`id`), their simulated mean (`mean_pts`) and standard deviation (`sd_pts`) of projected fantasy points across sources, their the 10th (`q10`) and 90th (`q90`) quartiles of their distributions of projected points, and the simulated probability of the player scoring at least 100 (`p100`), 150 (`p150`), 200 (`p200`), 250 (`p250`), 300 (`p300`), and 350 (`p350`) fantasy points.
+We then combine all the rows into a tibble dataframe, with each player as a different row.
+A separate dataset is created for each of the simulation replications.
 
@@
 Fourth, we summarize the results across the simulation replications.
+In this case, we calculate—for each variable—each player's mean and standard deviation across the datasets from the different simulation replications.
+
+Some potential metastatistics that can be computed include:
+
+- mean: `base::mean()`
+- standard deviation: `base::sd()`
+- bias: `SimDesign::bias()`
+- standardized bias: `SimDesign::bias(, type = "standardized")`
+- relative bias: `SimDesign::bias(, type = "relative")`
+- relative absolute bias: `SimDesign::RAB()` or `SimDesign::bias(, type = "abs_relative")`
+- mean absolute error (MAE): `SimDesign::MAE()`
+- normalized mean absolute error: `SimDesign::MAE(, type = "NMAE")`
+- standardized mean absolute error: `SimDesign::MAE(, type = "SMAE")`
+- root mean squared error (RMSE): `SimDesign::RMSE()`
+- normalized root mean squared error: `SimDesign::RMSE(, type = "NRMSE")`
+- standardized root mean squared error: `SimDesign::RMSE(, type = "SRMSE")`
+- root mean squared log error (RMSLE): `SimDesign::RMSE(, type = "RMSLE")`
+- integrated root mean squared error (IRMSE): `SimDesign::IRMSE()`
+- mean-square relative standard error (MSRSE): `SimDesign::MSRSE()`
+- coefficient of variation: `SimDesign::RMSE(, type = "CV")`
+- relative efficiency: `SimDesign::RE()`
+- empirical detection/rejection rate: `SimDesign::EDR()`
+- empirical coverage rate: `SimDesign::ECR()`
 
@@
   summarise = Summarise,
-  fixed_objects = list(proj_dists = proj_dists),
+  fixed_objects = list(proj_dists = proj_dists), # passing a list of objects that should be held constant across all simulation replications — i.e., "fixed" objects
   seed = SimDesign::genSeeds(Design, iseed = 52242), # for reproducibility

### 2025-11-06 simulation.qmd
20251106 - Monte Carlo simulation

@@
 
+Monte Carlo simulation is named after the Casino de Monte-Carlo in Monaco because Monte Carlo simulations involve random samples (random chance), as might be found in casino gambling.
 Monte Carlo simulation involves repeated random sampling from a known distribution [@Sigal2016].
@@
 
-1. generate the data for the simulation from the relevant probabilty distribution(s)
+1. generate the data for the simulation from the relevant probability distribution(s)
 1. analyze the simulated data

### 2025-11-06 modern-portfolio-theory.qmd
20251106 - typo

@@
     y = "Mean of Players' Fantasy Points Across Weeks",
-    title = "Mean vs Standard Deviation of Players' Projected Fantasy Points"
+    title = "Mean vs Standard Deviation of Players' Actual Fantasy Points"
   ) +

### 2025-11-07 multiple-regression.qmd
20251107 - suppression

@@
 
+## Suppression {#sec-multipleRegressionSuppression}
+
+Suppression in a regression model occurs when including a [predictor variable](#sec-correlationalStudy) increases the magnitude of the regression coefficient (i.e., beta) of another [predictor variable](#sec-correlationalStudy) on an [outcome variable](#sec-correlationalStudy) [@MacKinnon2000].
+The suppressor variable suppresses irrelevant variance in another [predictor variable](#sec-correlationalStudy), allowing the [predictor variable's](#sec-correlationalStudy) association with the [outcome variable](#sec-correlationalStudy) to emerge more clearly.
+Three types of suppression are depicted in @GaylordHarden2010.
+
 ## Handling of Missing Data {#sec-multipleRegressionMissingness}

### 2025-11-10 simulation.qmd
20251110 - additions and clarification

@@
 ### Bootstrapping {#sec-simulationProjectedPointsBootstrapping}
+
+In this example, we simulate players' likely distribution of (projected) fantasy points based on bootstrapping players' projected statisics (e.g., passing yards, passing touchdowns).
+That is, for each iteration of the bootstrap, we randomly draw a projected value from a projection source (with replacement), for each of various statistical categories.
+For instance, assume we have 10 projection sources for Tom Brady.
+In iteration 1, we may draw Brady's projected passing yards from ESPN and projected passing touchdowns from numberFire (i.e., randomly selected projection sources).
+In iteration 2, we may draw Brady's projected passing yards from Yahoo and projected passing touchdowns from CBS (i.e., randomly selected projection sources).
+We do this for many iterations.
+We can compute projected fantasy points at each iteration and then summarize the distribution of fantasy points for Tom Brady (and all players) across all iterations.
+This bootstrap simulation is valuable because it would give us a fuller distribution of the likely range of players' performance than just the mean (or range) of projected points across sources.
 
@@
 Third, we analyze the simulated data.
-For each iteration of the simulation, the `Analyse()` function will return a vector with a requested variables: player (`id`), their simulated mean (`mean_pts`) and standard deviation (`sd_pts`) of projected fantasy points across sources, their the 10th (`q10`) and 90th (`q90`) quartiles of their distributions of projected points, and the simulated probability of the player scoring at least 100 (`p100`), 150 (`p150`), 200 (`p200`), 250 (`p250`), 300 (`p300`), and 350 (`p350`) fantasy points.
+For each iteration of the simulation, the `Analyse()` function will return a vector with a requested variables: player (`id`), their simulated mean (`mean_pts`) and standard deviation (`sd_pts`) of projected fantasy points across sources, the 10th (`q10`) and 90th (`q90`) quartiles of their distributions of projected points, and the simulated probability of the player scoring at least 100 (`p100`), 150 (`p150`), 200 (`p200`), 250 (`p250`), 300 (`p300`), and 350 (`p350`) fantasy points.
 We then combine all the rows into a tibble dataframe, with each player as a different row.

### 2025-11-11 cluster-analysis.qmd
20251111 - cluster analysis models and fit criteria

@@
 
+There are several model fit criteria we can use to identify the optimal number of tiers in terms of the tradeoff between fit and parsimony.
+For the Bayesian Information Criterion (BIC), closer to zero is better.
+For the Integrated Complete-data Likelihood (ICL), higher is better.
+For the bootstrapped likelihood ratio tests, a significant likelihood ratio test indicates that the model with more clusters fits significantly better than the model with fewer clusters.
+
+For univariate cluster analyses in the `mclust` package [@R-mclust; @Scrucca2023_packages], there are two types of models: ones with equal variance across clusters (`modelName = "E"`) and ones with different variances across clusters (`modelName = "V"`).
+Allowing different variances across clusters is a more complex model than forcing the clusters to have equal variances.
+
+For multivariate cluster analyses in the `mclust` package [@R-mclust; @Scrucca2023_packages], models are specified by a combination of volume, shape, and orientation.
+The model name is specified such that the first letter corresponds to the volume, the second letter corresponds to shape, and the third letter corresponds to orientation.
+In terms of volume, models can be of equal volume (`E`) or variable volume (`V`) across clusters.
+In terms of shape, models can be of equal (`E`), variable (`V`), or identity (`I`) shape.
+In terms of orientation, models can be of equal (`E`), variable (`V`), or identity (`I`) orientation.
+The varying types of volume, shape, and orientation combine to produce three general types of covariance structures: spherical (low complexity), diagonal (moderate complexity), and ellipsoidal (high complexity).
+Models that allow greater differences across clusters (i.e., more `V`'s) are more flexible and typically more complex.
+
 ### Tiers of Prior Season Fantasy Points {#sec-clusterAnalysisExample}
@@
 
+We can evaluate how many clusters to keep, based on the tradeoff between fit and parsimony, using the `mclust::mclustBIC()` (BIC), `mclust::mclustICL()` (ICL), and `mclust::mclustBootstrapLRT()` (bootstrapped likelihood ratio tests) functions of the `mclust` package [@R-mclust; @Scrucca2023_packages].
+
 ##### Quarterbacks {#sec-clusterAnalysisExampleNumTiersQBs}
-
-We can perform a cluster analysis using the `mclust::mclustBIC()`, `mclust::mclustICL()`, and `mclust::mclustBootstrapLRT()` functions of the `mclust` package [@R-mclust; @Scrucca2023_packages].
 

### 2025-11-11 cluster-analysis.qmd
20251111 - model fit criteria

@@
 
-There are several model fit criteria we can use to identify the optimal number of tiers in terms of the tradeoff between fit and parsimony.
+There are several model fit criteria we can use to identify the optimal number of clusters in terms of the tradeoff between fit and parsimony.
 For the Bayesian Information Criterion (BIC), closer to zero is better.
@@
 For the bootstrapped likelihood ratio tests, a significant likelihood ratio test indicates that the model with more clusters fits significantly better than the model with fewer clusters.
+Although the model fit criteria can be helpful for identifying the optimal number of clusters, they should only be used as a guide.
+Ultimately, cluster analysis is only useful to the extent that the clusters and useful and interpretable, which may mean selecting more (or fewer) clusters than are suggested by the model fit criteria.
 

### 2025-11-12 time-series-analysis.qmd
20251111 - components of time series

@@
 We also leverage [Bayesian mixed models](#sec-fantasyPointsByAgeExperienceModelsBayesian) to generate forecasts of future performance and plots of individuals model-implied performance by age and position.
+
+There are four (potential) key components of time series data:
+
+- seasonality
+- trend
+- cycle
+- irregularity
+
+*Seasonality* corresponds to fluctuations that occur at fixed and known periods (e.g., time of day, day of week, or month of year) [@Hyndman2021].
+A *trend* represents a long-term change (i.e., increase or decrease) in the series.
+A *cycle* reflects one or more rises and falls that are not of a fixed frequency, with durations typically lasting at least two years [@Hyndman2021].
+*Irregularity* is random variation or noise.
 

### 2025-11-12 mixed-models.qmd
20251112 - fix plot

@@
 #| label: fig-bayesianRegressionCoefficientsDistributions
-#| fig-cap: "Distributions of Regression Coefficients with 66% (Thin Line) and 95% (Thin Line) Credible Intervals from Bayesian Mixed Model."
-#| fig-alt: "Distributions of Regression Coefficients with 66% (Thin Line) and 95% (Thin Line) Credible Intervals from Bayesian Mixed Model."
+#| fig-cap: "Distributions of Regression Coefficients with 66% (Thick Line) and 95% (Thin Line) Credible Intervals from Bayesian Mixed Model."
+#| fig-alt: "Distributions of Regression Coefficients with 66% (Thick Line) and 95% (Thin Line) Credible Intervals from Bayesian Mixed Model."
 
@@
   labs(
-    title = "Posterior Distributions of Regression Coefficients with 66% and 95% Credible Intervals",
+    title = "Posterior Distributions of Regression Coefficients\nwith 66% and 95% Credible Intervals",
     x = "Predictor",

### 2025-11-12 time-series-analysis.qmd
20251112 - exponential smoothing

@@
 There many different types of time series analyses.
-For simplicity, in this chapter, we use [autoregressive integrated moving average](#sec-timeSeriesARIMA) (ARIMA) models to demonstrate one approach to time series analysis.
+For simplicity, in this chapter, we use [autoregressive integrated moving average](#sec-timeSeriesARIMA) (ARIMA) and [exponential smoothing](#sec-timeSeriesExponentialSmoothing) models to demonstrate one approach to time series analysis.
 We also leverage [Bayesian mixed models](#sec-fantasyPointsByAgeExperienceModelsBayesian) to generate forecasts of future performance and plots of individuals model-implied performance by age and position.
@@
 This may not strongly be the case in fantasy football, so ARIMA models may not be particularly useful in forecasting fantasy football performance.
-Other approaches, such as exponential smoothing, may be useful for data that show longer-term trends and seasonality [@Hyndman2021].
+Other approaches, such as [exponential smoothing](#sec-timeSeriesExponentialSmoothing), may be useful for data that show longer-term trends and seasonality [@Hyndman2021].
 Nevertheless, ARIMA models are widely used in forecasting financial markets and economic indicators.
@@
 #| label: fig-timeSeriesProjectionTomBrady
-#| fig-cap: "Tom Brady's Historical and Projected Fantasy Points by Game."
-#| fig-alt: "Tom Brady's Historical and Projected Fantasy Points by Game."
+#| fig-cap: "Tom Brady's Historical and Projected Fantasy Points by Game Based on Autoregressive Integrated Moving Average (ARIMA) Model."
+#| fig-alt: "Tom Brady's Historical and Projected Fantasy Points by Game Based on Autoregressive Integrated Moving Average (ARIMA) Model."
 
@@
 #| label: fig-timeSeriesProjectionPeytonManning
-#| fig-cap: "Peyton Manning's Historical and Projected Fantasy Points by Game."
-#| fig-alt: "Peyton Manning's Historical and Projected Fantasy Points by Game."
+#| fig-cap: "Peyton Manning's Historical and Projected Fantasy Points by Game Based on Autoregressive Integrated Moving Average (ARIMA) Model."
+#| fig-alt: "Peyton Manning's Historical and Projected Fantasy Points by Game Based on Autoregressive Integrated Moving Average (ARIMA) Model."
 
 forecast::autoplot(forecast_peytonManning) + 
+  labs(
+    x = "Game Number",
+    y = "Fantasy Points",
+    title = "Peyton Manning's Historical and Projected Fantasy Points by Game",
+    subtitle = "(if he were to have continued playing additional seasons)"
+  ) +
+  theme_classic()
+```
+
+## Exponential Smoothing {#sec-timeSeriesExponentialSmoothing}
+
+### Fit the Exponential Smoothing Model {#sec-exponentialSmoothingModel}
+
+We fit exponential smoothing models using the `forecast::ets()` function of the `forecast` package [@R-forecast; @Hyndman2008_packages]:
+
+```{r}
+ets_tomBrady <- forecast::ets(ts_tomBrady)
+ets_peytonManning <- forecast::ets(ts_peytonManning)
+```
+
+We can generate a plot of an ARIMA model using the `stats::arima()` function.
+
+```{r}
+#| label: fig-etsTomBrady
+#| fig-cap: "Model Summary of Exponential Smoothing Model fit to Tom Brady's Historical Performance by Game."
+#| fig-alt: "Model Summary of Exponential Smoothing Model fit to Tom Brady's Historical Performance by Game."
+
+summary(ets_tomBrady)
+forecast::checkresiduals(ets_tomBrady)
+```
+
+```{r}
+#| label: fig-etsPeytonManning
+#| fig-cap: "Model Summary of Exponential Smoothing Model fit to Peyton Manning's Historical Performance by Game."
+#| fig-alt: "Model Summary of Exponential Smoothing Model fit to Peyton Manning's Historical Performance by Game."
+
+summary(ets_peytonManning)
+forecast::checkresiduals(ets_peytonManning)
+```
+
+### Generate the Model Forecasts {#sec-exponentialSmoothingForecasts}
+
+We can generate model forecasts from the ARIMA model using the `forecast::forecast()` function.
+
+```{r}
+forecastETS_tomBrady <- forecast::forecast(
+  ets_tomBrady,
+  level = c(80, 95)) # 80% and 95% confidence intervals
+
+forecastETS_peytonManning <- forecast::forecast(
+  ets_peytonManning,
+  level = c(80, 95)) # 80% and 95% confidence intervals
+
+forecastETS_tomBrady
+forecastETS_peytonManning
+```
+
+### Plot the Model Forecasts {#sec-plotExponentialSmoothingForecast}
+
+We can plot the model forecasts using the `forecast::autoplot()` function.
+
+```{r}
+#| label: fig-etsProjectionTomBrady
+#| fig-cap: "Tom Brady's Historical and Projected Fantasy Points by Game Based on Exponential Smoothing."
+#| fig-alt: "Tom Brady's Historical and Projected Fantasy Points by Game Based on Exponential Smoothing."
+
+forecast::autoplot(forecastETS_tomBrady) + 
+  labs(
+    x = "Game Number",
+    y = "Fantasy Points",
+    title = "Tom Brady's Historical and Projected Fantasy Points by Game",
+    subtitle = "(if he were to have continued playing additional seasons)"
+  ) +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-etsProjectionPeytonManning
+#| fig-cap: "Peyton Manning's Historical and Projected Fantasy Points by Game Based on Exponential Smoothing."
+#| fig-alt: "Peyton Manning's Historical and Projected Fantasy Points by Game Based on Exponential Smoothing."
+
+forecast::autoplot(forecastETS_peytonManning) + 
   labs(

### 2025-11-12 R-scripts/download_weekly_projections.R
20251112 - week 11

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 10
+weekNumber <- 11
 
 # Processing ----
 season <- as.character(season)

### 2025-11-12 decision-making.qmd
20250903 - updates to address updated data files
20251112 - typos

@@
 We created the `projectionsWithActuals_seasonal.RData` object in @sec-predictionAccuracyPrepareDataSeasonal; we created the `projectionsWithActuals_weekly.RData` object in @sec-predictionAccuracyPrepareDataWeekly.
+We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
 
@@
 
-The relatively modest accuracy of the projections by so-called fantasy "experts" and of their average of their projections could occur for a number of reasons.
+The relatively modest accuracy of the projections by so-called fantasy "experts" and of the average of their projections could occur for a number of reasons.
 One possibility is that the level of expertise of the "expert" forecasters in terms of being able to provide accurate forecasts is not strong.
@@
 Interaction and discussion among experts may lead them to follow the herd and conform their projections to what each other predict.
-This has been terms informational influence and may reflect the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment) [@Larrick2024].
+This has been termed "informational influence" and may reflect the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment) [@Larrick2024].
 In any case, they are able to see each other's projections and make change their projections, accordingly.
@@
 
-Sports analyst @Tanney2021 encourages people to consider framing your decisions as bets, to use a probabilistic mindset (rather than a determininistic mindset).
+Sports analyst @Tanney2021 encourages people to consider framing your decisions as bets, to use a probabilistic mindset (rather than a deterministic mindset).
 Even if something is perceived as likely to occur with 100% certainty, it is not guaranteed that such an event will occur.
@@
 Moreover, people tend to show [confirmation bias](#sec-cognitiveBiasesConfirmation), such that they tend to remember their predictive successes and to forget their failures.
-So, when they they get a bet correct, they may be more likely to bet again in the future, especially because sports betting involves [intermittent reinforcement](#sec-reinforcementIntermittent) (in particular, variable ratio reinforcement), which can make it highly addictive.
+So, when they get a bet correct, they may be more likely to bet again in the future, especially because sports betting involves [intermittent reinforcement](#sec-reinforcementIntermittent) (in particular, variable ratio reinforcement), which can make it highly addictive.
 
@@
 Extinction means cessation of the target behavior (putting the toys away, in this example).
-That is, for a previously continusouly reinforced behavior, if the behavior stops being rewarded, the person is less likely to continue the behavior.
+That is, for a previously continuously reinforced behavior, if the behavior stops being rewarded, the person is less likely to continue the behavior.
 Thus, to make a behavior less susceptible to extinction, [intermittent reinforcement](#sec-reinforcementIntermittent) may be used.

### 2025-11-14 sports-cognitive-psychology.qmd
20250910 - artificial intelligence
20251114 - typo

@@
 Head Coach Doug Peterson heavily relied on analytics with the Philadelphia Eagles to help them win the 2017 Super Bowl; for more information, see [here](https://www.forbes.com/sites/liamfox/2021/08/12/how-the-nfl-uses-analytics-according-to-the-lead-analyst-of-a-super-bowl-champion/) [@Fox2021; archived at <https://perma.cc/49KQ-R785>] and [here](https://www.nfl.com/news/super-bowl-lii-how-the-2017-philadelphia-eagles-were-built-0ap3000000912753) [@Rosenthal2018; archived at <https://perma.cc/2GRF-8Z5K>].
+Moreover, teams are now using artificial intelligence (AI) to help them prepare for games, evaluate players, make draft selections, and perform game management [@Levi2025; archived at <https://perma.cc/XF2Z-ZZZT>].
 
@@
 
-As depicted in the @fig-goForItPercentageByWinProbChange [which was adapted from a [figure by Ben Baldwin](https://www.nfl4th.com/articles/4th-down-research.html); @R-nfl4th; archived at <https://perma.cc/S5D8-3NCU>], teams went for it on fourth down nearly twice as often in 2021 compared to 2017, when when their win probability was at least 20% (given the current situation at the start of the given play) and they would increase their win probably by going for it.
+As depicted in the @fig-goForItPercentageByWinProbChange [which was adapted from a [figure by Ben Baldwin](https://www.nfl4th.com/articles/4th-down-research.html); @R-nfl4th; archived at <https://perma.cc/S5D8-3NCU>], teams went for it on fourth down nearly twice as often in 2021 compared to 2017, when their win probability was at least 20% (given the current situation at the start of the given play) and they would increase their win probability by going for it.
 We use the `ggtext` package [@R-ggtext] to annotate text:

### 2025-11-17 cluster-analysis.qmd
20251111 - reorder
20251117 - fixes

@@
 Model-based clustering represents the data as coming from a mixture of probability distributions.
-The various types of `mclust` models are provided here:
-<https://mclust-org.github.io/mclust/reference/mclustModelNames.html>.
 
@@
 Models that allow greater differences across clusters (i.e., more `V`'s) are more flexible and typically more complex.
+
+The various types of `mclust` models are provided here:
+<https://mclust-org.github.io/mclust/reference/mclustModelNames.html>.
 
@@
 
-tiersQB_boostrap <- mclust::mclustBootstrapLRT(
+tiersQB_bootstrap <- mclust::mclustBootstrapLRT(
   data = player_stats_seasonal_offense_recentQB$fantasyPoints,
@@
 
-numTiersQB <- as.numeric(summary(tiersQB_boostrap)[,"Length"][1]) # or could specify the number of teams manually
+numTiersQB <- as.numeric(summary(tiersQB_bootstrap)[,"Length"][1]) # or could specify the number of teams manually
 
-tiersQB_boostrap
+tiersQB_bootstrap
 plot(
-  tiersQB_boostrap,
+  tiersQB_bootstrap,
   G = numTiersQB - 1)
@@
 The model-based bootstrap clustering of Running Backs' fantasy points is unable to run due to an error:
+UPDATE: THE ERROR NO LONGER OCCURS
 -->
@@
 ```{r}
-tiersRB_boostrap <- mclust::mclustBootstrapLRT(
+tiersRB_bootstrap <- mclust::mclustBootstrapLRT(
   data = player_stats_seasonal_offense_recentRB$fantasyPoints,
@@
 Thus, we cannot use the following code, which would otherwise summarize the model results, specify the number of tiers, and plot model comparisons:
+UPDATE: THE ERROR NO LONGER OCCURS
 -->
@@
 ```{r}
-numTiersRB <- as.numeric(summary(tiersRB_boostrap)[,"Length"][1]) # or could specify the number of teams manually
+numTiersRB <- as.numeric(summary(tiersRB_bootstrap)[,"Length"][1]) # or could specify the number of teams manually
 
-tiersRB_boostrap
+tiersRB_bootstrap
 plot(
-  tiersRB_boostrap,
+  tiersRB_bootstrap,
   G = numTiersRB - 1)
@@
 
-tiersWR_boostrap <- mclust::mclustBootstrapLRT(
+tiersWR_bootstrap <- mclust::mclustBootstrapLRT(
   data = player_stats_seasonal_offense_recentWR$fantasyPoints,
@@
 
-numTiersWR <- as.numeric(summary(tiersWR_boostrap)[,"Length"][1]) # or could specify the number of teams manually
+numTiersWR <- as.numeric(summary(tiersWR_bootstrap)[,"Length"][1]) # or could specify the number of teams manually
 
-tiersWR_boostrap
+tiersWR_bootstrap
 plot(
-  tiersWR_boostrap,
+  tiersWR_bootstrap,
   G = numTiersWR - 1)
@@
 ##### Tight Ends {#sec-clusterAnalysisExampleNumTiersTEs}
+
+###### Model-Based Clustering {#sec-clusterAnalysisExampleNumTiersTEsModel}
 
@@
 
-tiersTE_boostrap <- mclust::mclustBootstrapLRT(
+tiersTE_bootstrap <- mclust::mclustBootstrapLRT(
   data = player_stats_seasonal_offense_recentTE$fantasyPoints,
@@
 
-numTiersTE <- as.numeric(summary(tiersTE_boostrap)[,"Length"][1]) # or could specify the number of teams manually
+numTiersTE <- as.numeric(summary(tiersTE_bootstrap)[,"Length"][1]) # or could specify the number of teams manually
 
-tiersTE_boostrap
+tiersTE_bootstrap
 plot(
-  tiersTE_boostrap,
+  tiersTE_bootstrap,
   G = numTiersTE - 1)
@@
 
-wrTypes_boostrap <- mclust::mclustBootstrapLRT(
+wrTypes_bootstrap <- mclust::mclustBootstrapLRT(
   data = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
@@
 
-wrTypes_boostrap
+wrTypes_bootstrap
 plot(
-  wrTypes_boostrap,
+  wrTypes_bootstrap,
   G = numTypesWR - 1)

### 2025-11-17 cluster-analysis.qmd
20251117 - k-means clustering

@@
 
-There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, hierarchical clustering (aka connectivity-based clustering), etc.
+There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering (e.g., *k*-means clustering), hierarchical clustering (aka connectivity-based clustering), etc.
 In general, cluster analysis intends to maximize intracluster similarities and to minimize intercluster similarities [@Ramasubramanian2016].
@@
 Model-based clustering represents the data as coming from a mixture of probability distributions.
+We also use *k*-means clustering.
 
@@
 We can evaluate how many clusters to keep, based on the tradeoff between fit and parsimony, using the `mclust::mclustBIC()` (BIC), `mclust::mclustICL()` (ICL), and `mclust::mclustBootstrapLRT()` (bootstrapped likelihood ratio tests) functions of the `mclust` package [@R-mclust; @Scrucca2023_packages].
+We can also use *k*-means clustering using the `stats::kmeans()` function.
 
 ##### Quarterbacks {#sec-clusterAnalysisExampleNumTiersQBs}
+
+###### Model-Based Clustering {#sec-clusterAnalysisExampleNumTiersQBsModel}
 
@@
 
+###### *k*-Means Clustering {#sec-clusterAnalysisExampleNumTiersQBsKmeans}
+
+```{r}
+tiersQB_kMeans <- kmeans(
+  x = player_stats_seasonal_offense_recentQB$fantasyPoints,
+  centers = numTiersQB)
+
+tiersQB_kMeans
+```
+
 ##### Running Backs {#sec-clusterAnalysisExampleNumTiersRBs}
+
+###### Model-Based Clustering {#sec-clusterAnalysisExampleNumTiersRBsModel}
 
@@
 
+###### *k*-Means Clustering {#sec-clusterAnalysisExampleNumTiersRBsKmeans}
+
+```{r}
+tiersRB_kMeans <- kmeans(
+  x = player_stats_seasonal_offense_recentRB$fantasyPoints,
+  centers = numTiersRB)
+
+tiersRB_kMeans
+```
+
 ##### Wide Receivers {#sec-clusterAnalysisExampleNumTiersWRs}
+
+###### Model-Based Clustering {#sec-clusterAnalysisExampleNumTiersWRsModel}
 
@@
   G = numTiersWR - 1)
+```
+
+###### *k*-Means Clustering {#sec-clusterAnalysisExampleNumTiersWRsKmeans}
+
+```{r}
+tiersWR_kMeans <- kmeans(
+  x = player_stats_seasonal_offense_recentWR$fantasyPoints,
+  centers = numTiersWR)
+
+tiersWR_kMeans
 ```
@@
   G = numTiersTE - 1)
+```
+
+###### *k*-Means Clustering {#sec-clusterAnalysisExampleNumTiersTEsKmeans}
+
+```{r}
+tiersTE_kMeans <- kmeans(
+  x = player_stats_seasonal_offense_recentTE$fantasyPoints,
+  centers = numTiersTE)
+
+tiersTE_kMeans
 ```
@@
 
+##### Model-Based Clustering  {#sec-clusterAnalysisNumWRtypesModel}
+
 ```{r}
@@
   G = numTypesWR - 1)
+```
+
+##### *k*-Means Clustering  {#sec-clusterAnalysisNumWRtypesKmeans}
+
+```{r}
+wrTypes_kMeans <- kmeans(
+  x = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
+  centers = numTypesWR)
+
+wrTypes_kMeans
 ```

### 2025-11-18 basic-statistics.qmd
20251111 - the end of average
20251118 - squared semi-partial correlation

@@
 For instance, the mean of the values of 0, 0, 10, 15, 20, 30, and 1000 is `{r} petersenlab::apa(exampleValues_mean, decimals = 2)`.
+Another issue with the mean is that it may not be the same as any individual's value [@Rose2016]—that is, it may be unrepresentative of all of the individuals whose values comprise the average.
 
@@
 
+#### Squared Semi-Partial Correlation ($sr^2$) {#sec-squaredSemiPartialCorrelation}
+
+Squared semi-partial correlations are described in @sec-multipleRegressionExplainedVariancePredictors.
+
 #### Eta Squared ($\eta^2$) and Partial Eta Squared ($\eta_p^2$) {#sec-etaSquared}
@@
 | Coefficient of determination ($R^2$)                | $\ge .01$   | $\ge .06$   | $\ge .14$   |
+| Squared semi-partial correlation ($sr^2$)           | $\ge .01$   | $\ge .06$   | $\ge .14$   |
 | Eta squared ($\eta^2$)                              | $\ge .01$   | $\ge .06$   | $\ge .14$   |

### 2025-11-18 getting-started.qmd
20251118 - squared semi-partial correlation

@@
      "interactions","robustbase","ordinal","MASS","data.table","future",
-     "future.apply","SimDesign","domir","GGally","Rglpk","TTR"))
+     "future.apply","SimDesign","domir","GGally","Rglpk","TTR","rockchalk"))
    ```

### 2025-11-18 index.qmd
20251118 - squared semi-partial correlation

@@
   "ordinal", "MASS", "data.table", "future", "future.apply", "SimDesign", 
-  "domir", "GGally","Rglpk","TTR",
+  "domir", "GGally","Rglpk","TTR","rockchalk",
   "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))

### 2025-11-18 multiple-regression.qmd
20251110 - model as simplification of reality
20251114 - absolute value of total effect
20251118 - squared semi-partial correlation

@@
 library("broom")
+library("rockchalk")
 library("effectsize")
@@
 
+In multiple regression, we estimate models.
+A model is a simplification of reality.
 All statistical analyses follow the same basic structure, as in @eq-analysisBasicStructure [@Brauer2018]:
@@
 
+### Unstandardized Regression Coefficient {#sec-multipleRegressionComponentsUnstandardizedCoefficient}
+
 The unstandardized regression coefficient ($B$) is interpreted such that, for every unit change in the [predictor variable](#sec-correlationalStudy), there is a __ unit change in the [outcome variable](#sec-correlationalStudy).
@@
 
+### Standardized Regression Coefficient {#sec-multipleRegressionComponentsStandardizedCoefficient}
+
 Standardized regression coefficients can be obtained by standardizing the variables to [*z* scores](#sec-zScores) so they all have a mean of zero and standard deviation of one.
@@
 
+### Standard Error {#sec-multipleRegressionComponentsStandardError}
+
 The *standard error* of a regression coefficient represents the imprecision or uncertainty of the parameter.
@@
 Otherwise said, having more data gives more precise estimates and thus increases [statistical power](#sec-statisticalPower).
+
+### Confidence Interval {#sec-multipleRegressionComponentsConfidenceInterval}
 
@@
 where $p$ is the number of [predictor variables](#sec-correlationalStudy) in the model, and $n$ is the sample size.
+
+## How Much Variance Each Predictor Explains {#sec-multipleRegressionExplainedVariancePredictors}
+
+Another estimate of [effect size](#sec-practicalSignificance) in addition to [standardized regression coefficients](#sec-multipleRegressionComponentsStandardizedCoefficient) is the squared semi-partial correlation ($sr^2$).
+The squared semi-partial correlation reflects the proportion of variance in the [outcome variable](#sec-correlationalStudy) that is uniquely explained by each [predictor variable](#sec-correlationalStudy).
+The total model $R^2$ minus the sum of all the squared semi-partial correlations reflects the proportion of variance in the [outcome variable](#sec-correlationalStudy) that is explained by the common variance among the [predictor variables](#sec-correlationalStudy).
 
@@
 
+### Squared Semi-Partial Correlation ($sr^2$) {#sec-multipleRegressionSquaredSemiPartialCorrelation}
+
+To estimate the [squared semi-partial correlation](#sec-multipleRegressionExplainedVariancePredictors) ($sr^2$), we use the `rockchalk::getDeltaRsquare()` function from the `rockchalk` package [@R-rockchalk].
+
+```{r}
+rockchalk::getDeltaRsquare(linearRegressionModel)
+```
+
 ### Dominance Analysis {#sec-multipleRegressionDominanceAnalysis}
@@
   # total effect (c)
-  total := abs(direct) + abs(indirect)
+  total := direct + indirect
+  totalAbs := abs(direct) + abs(indirect)
 
   # proportion mediated
-  Pm := abs(indirect) / total
+  Pm := abs(indirect) / totalAbs
 '

### 2025-11-19 getting-started.qmd
20251118 - commonality analysis

--- 2025-09-09 getting-started.qmd
@@
      "interactions","robustbase","ordinal","MASS","data.table","future",
-     "future.apply","SimDesign","domir","GGally","Rglpk","TTR"))
+     "future.apply","SimDesign","domir","GGally","Rglpk","TTR","rockchalk",
+     "yhat"))
    ```

### 2025-11-19 index.qmd
20251118 - commonality analysis

--- 2025-08-27 index.qmd
@@
   "ordinal", "MASS", "data.table", "future", "future.apply", "SimDesign", 
-  "domir", "GGally","Rglpk","TTR",
+  "domir", "GGally","Rglpk","TTR","rockchalk","yhat",
   "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))

### 2025-11-19 multiple-regression.qmd
20251118 - commonality analysis

@@
 library("effectsize")
+library("yhat")
 library("tidymodels")
@@
 
+### Commonality Analysis {#sec-multipleRegressionCommonalityAnalysis}
+
+Commonality analysis decomposes the model's $R^2$ into coefficients that represent the proportion of variance accounted for uniquely by each [predictor variable](#sec-correlationalStudy) and the proportion of variance accounted for that is shared in each possible combination of the [predictor variables](#sec-correlationalStudy).
+Below we conduct a commonality analysis using the `yhat::calc.yhat()`, `yhat::commonality()`, and `yhat::commonalityCoefficients()` functions of the `yhat` package [@R-yhat].
+
+```{r}
+yhat::calc.yhat(linearRegressionModel)
+
+apsOut <- yhat::aps(
+  dataMatrix = player_stats_seasonal %>% 
+    filter(position %in% c("WR")) %>% 
+    select(fantasyPoints, age, height, weight, target_share),
+  dv = "fantasyPoints",
+  ivlist = list("age", "height", "weight", "target_share"))
+
+yhat::commonality(apsOut)
+
+commonalityAnalysis <- yhat::commonalityCoefficients(
+  dataMatrix = player_stats_seasonal %>% 
+    filter(position %in% c("WR")) %>% 
+    select(fantasyPoints, age, height, weight, target_share),
+  dv = "fantasyPoints",
+  ivlist = list("age", "height", "weight", "target_share"))
+
+commonalityAnalysis
+```
+
 ### Dominance Analysis {#sec-multipleRegressionDominanceAnalysis}
@@
 parameters::dominance_analysis(linearRegressionModel)
+```
+
+We can also conduct a dominance analysis using the `yhat::dominance()` function of the `yhat` package [@R-yhat].
+
+```{r}
+yhat::dominance(apsOut)
 ```

### 2025-11-19 cluster-analysis.qmd
20251119 - tidyLPA for latent profile analysis

@@
 library("mclust")
+library("tidyLPA")
 library("plotly")
@@
 
+We can also use the `tidyLPA` package [@R-tidyLPA; @Rosenberg2018_packages], which provides an interface to the `mclust` package [@R-mclust; @Scrucca2023_packages].
+
+```{r}
+# model 1 (EEI): Equal variances and covariances fixed to 0
+# model 2 (VVI): Varying variances and covariances fixed to 0
+# model 3 (EEE): Equal variances and equal covariances
+# model 4      : Varying variances and equal covariances (not able to be fit w/ mclust but can be fit using package = "mplus", if installed)
+# model 5      : Equal variances and varying covariances (not able to be fit w/ mclust but can be fit using package = "mplus", if installed)
+# model 6 (VVV): Varying variances and varying covariances
+
+wrTypes_lpa_classes <- tidyLPA::estimate_profiles(
+  df = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
+  n_profiles = 1:6,
+  models = c(1, 2, 3) #c(1, 2, 3, 6) takes too long to run because of the model with varying variances and varying covariances (model 6)
+)
+
+wrTypes_lpa_classes
+
+tidyLPA::compare_solutions(
+  wrTypes_lpa_classes,
+  statistics = c("AIC", "BIC", "Entropy", "BLRT_p"))
+
+wrTypes_lpa <- tidyLPA::estimate_profiles(
+  df = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
+  n_profiles = numTypesWR,
+  model = 3 # equal variances and equal covariances
+)
+
+wrTypes_lpa
+tidyLPA::get_fit(wrTypes_lpa)
+```
+
 ##### *k*-Means Clustering  {#sec-clusterAnalysisNumWRtypesKmeans}
@@
   what = "density")
+```
+
+```{r}
+tidyLPA::plot_profiles(wrTypes_lpa)
 ```
@@
 
+And here are the clusters based on the `tidyLPA` [@R-tidyLPA; @Rosenberg2018_packages] solution:
+
+```{r}
+merged_dataWRs_cluster <- cbind(
+  merged_dataWRs_cluster,
+  tidyLPA::get_data(wrTypes_lpa) %>% 
+    select(model_number, classes_number, CPROB1:Class)
+)
+
+merged_dataWRs_cluster %>% 
+  select(display_name, CPROB1:Class)
+
+merged_dataWRs_cluster %>% 
+  group_by(Class) %>% 
+  summarise(across(
+    where(is.numeric),
+    ~ mean(., na.rm = TRUE)
+    )) %>% 
+  t() %>% 
+  round(., 2)
+```
+
 ## Conclusion {#sec-clusterAnalysisConclusion}

### 2025-11-19 getting-started.qmd
20251119 - tidyLPA for latent profile analysis

--- 2025-09-09 getting-started.qmd
@@
      "interactions","robustbase","ordinal","MASS","data.table","future",
-     "future.apply","SimDesign","domir","GGally","Rglpk","TTR"))
+     "future.apply","SimDesign","domir","GGally","Rglpk","TTR","rockchalk",
+     "yhat","tidyLPA"))
    ```

### 2025-11-19 index.qmd
20251119 - tidyLPA for latent profile analysis

--- 2025-08-27 index.qmd
@@
   "ordinal", "MASS", "data.table", "future", "future.apply", "SimDesign", 
-  "domir", "GGally","Rglpk","TTR",
+  "domir", "GGally","Rglpk","TTR","rockchalk","yhat","tidyLPA",
   "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))

### 2025-11-20 R-scripts/download_weekly_projections.R
20251119 - week 12

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 11
+weekNumber <- 12
 
 # Processing ----
 season <- as.character(season)

### 2025-11-24 multiple-regression.qmd
20251124 - covariates

@@
 For instance, you might include the player's age as a covariate in a model that examines whether a player's 40-yard dash time at the NFL Combine predicts their fantasy points in their rookie year, but it may not be a [confound](#sec-causalDiagramConfounding).
+
+However, it is worth noting that, for several reasons, including a variable as a covariate in a model does not necessarily fully control for that construct.
+First, your measure of the covariate has measurement error and may not fully capture the construct, which can lead to [residual confounding](#sec-causalDiagramConfounding).
+Second, including the covariate only accounts for the linear effect of that variable.
+The variable (e.g., age) could have nonlinear effects that are not captured by the linear term (and may require nonlinear terms, e.g., quadratic).
+Third, the construct could interact with other constructs, which would also not be captured by the linear term (and may require interaction terms).
+Thus, it is important to consider the reliability of the covariate, and the ways in which the covariate is related to the outcome including any nonlinearity and interactions.
 

### 2025-11-29 multiple-regression.qmd
20251129 - covariate selection

@@
 Thus, it is important to consider the reliability of the covariate, and the ways in which the covariate is related to the outcome including any nonlinearity and interactions.
+
+In general, you should select which covariates to include based on theory and your [causal diagram](#sec-causalDiagrams) of the causal process (as described in @sec-causalDiagrams), not based on whether the covariate has a [statistically significant](#sec-statisticalSignificance) association with the [outcome variable](#sec-correlationalStudy).
+Just because a [predictor variable](#sec-correlationalStudy) is not associated with an [outcome variable](#sec-correlationalStudy) does not mean that the [predictor variable](#sec-correlationalStudy) will not be useful in the model.
+For instance, the [predictor variable](#sec-correlationalStudy) could have a [suppression effect](#sec-multipleRegressionSuppression), or a [moderating effect](#sec-moderation), etc., that may not be detectable based on a bivariate association.
 

### 2025-12-01 multiple-regression.qmd
20251201 - typo

@@
 The regression coefficients (and resulting accuracy) tend to become weaker when applied to new data, a phenomenon called [shrinkage](#sec-shrinkage), which is described in @sec-shrinkage.
-For instance, shrinking is observed when applying multiple regression models to new data in @sec-fittingModels-regression and when applying [machine learning](#sec-machineLearning) models to new data in @sec-fittingModels-machineLearning.
+For instance, shrinkage is observed when applying multiple regression models to new data in @sec-fittingModels-regression and when applying [machine learning](#sec-machineLearning) models to new data in @sec-fittingModels-machineLearning.
 

### 2025-12-03 R-scripts/download_weekly_projections.R
20251203 - week 14

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 12
+weekNumber <- 14
 
 # Processing ----
 season <- as.character(season)

### 2025-12-03 download-football-data.qmd
20251203 - use nfl4th data on OSF if errors out upon load

@@
 ```{r}
-#| output: false
+#| eval: false
 
@@
   seasons = 2014:nflreadr::most_recent_season())
+```
+
+```{r}
+#| include: false
+
+tryCatch(
+  {
+    nfl_4thdown_raw <- nfl4th::load_4th_pbp(
+      seasons = 2014:nflreadr::most_recent_season()
+    )
+  },
+  error = function(e) { # if errors out when trying to load the data using nfl4th::load_4th_pbp()
+    message("Primary load failed; falling back to OSF version...")
+    petersenlab::robust_load("https://osf.io/download/nhcrj/")
+  }
+)
 ```

### 2025-12-14 multiple-regression.qmd
20251213 - typo

@@
 Homoscedasticity of the residuals means that the variance of the residuals does not differ as a function of the [outcome variable](#sec-correlationalStudy) or as a function of the [predictor variable](#sec-correlationalStudy) (i.e., the residuals show constant variance as a function of [outcome](#sec-correlationalStudy)/[predictors](#sec-correlationalStudy)).
-If the residuals differ as a function of the [outcome](#sec-correlationalStudy) or [predictor](#sec-correlationalStudy) variable, this is called heterotoscedasticity.
+If the residuals differ as a function of the [outcome](#sec-correlationalStudy) or [predictor](#sec-correlationalStudy) variable, this is called heteroscedasticity.
 

### 2025-12-31 R-scripts/download_weekly_projections.R
20251231 - week 18

@@ -3,7 +3,7 @@
 # Specify Season and Week Number ----
 
 season <- 2025
-weekNumber <- 17
+weekNumber <- 18
 
 # Processing ----
 season <- as.character(season)
